### 2017-10-03 doc/icaps-18/strips-learning.tex
icaps18 draft paper added

\def\year{2017}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai17}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{}
\setcounter{secnumdepth}{0}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{comment}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}
\usepackage{multicol}
\usepackage{arydshln}

\newcommand{\tup}[1]{{\langle #1 \rangle}}

\newcommand{\pre}{\mathsf{pre}}     % precondition
\newcommand{\del}{\mathsf{del}}     % effect
\newcommand{\add}{\mathsf{add}}     % effect
\newcommand{\eff}{\mathsf{eff}}     % effect
\newcommand{\cond}{\mathsf{cond}}   % conditional effect
\newcommand{\true}{\mathsf{true}}   % true
\newcommand{\false}{\mathsf{false}} % false
\newcommand{\PE}{\mathrm{PE}}     % precondition
\newcommand{\strips}{\textsc{Strips}}     % precondition

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

\begin{document}

\title{Learning STRIPS action models with classical planning}

\author{Diego Garc\'ia\and Sergio Jim\'enez\and Eva Onaindia\\
{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
{\small Universitat Polit\'ecnica de Val\'encia.}\\
{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
{\small \{,serjice,onaindia\}@dsic.upv.es}}

\maketitle
\begin{abstract}
This paper presents a novel approach for learning classical planning action models from minimal input knowledge and exclusively using existing classical planners. The paper defines a classical planning compilation to learn action models from examples. The compilation is flexible to different degrees of available input knowledge; it accepts partially specified action models and learning examples can range from a set of plans with their corresponding initial and final states to only a set of initial and final states.  
\end{abstract}

\section{Introduction}
Off-the-shelf planners reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}. In addition planning action models allow to effectively address further tasks than plan synthesis, like plan/goal recognition~\cite{ramirez2010probabilistic}. Unfortunately building planning action models is complex, even for planning experts, limiting the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.

Despite many Machine Learning (ML) techniques are able to compute models from examples~\cite{michalski2013machine} its application to learning planning action models is not straightforward. First, the inputs to ML algorithms usually are finite numeric vectors encoding objects features while, in planning, the input includes actions so is closer to the representation of procedures and behaviours rather than to object representation. Second, the traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks). In the case of learning planning action models the output is not a scalar but a model of the possible state transitions. Last but not least, collecting {\em informative} examples for learning planning action models is challenging. Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.

Learning classical planning action models is a well-studied problem with sophisticated algorithms, like ARMS~\cite{yang2007learning}, SLAF~\cite{amir:alearning:JAIR08} or LOCM~\cite{cresswell2013acquiring} that do not require full knowledge of all the states traversed by the example plans. Motivated by recent advances on learning generative models with classical planning~\cite{segovia2017generating} and on the effective exploration of planning state spaces~\cite{geffner:novelty:IJCAI17}, this paper introduces an innovative approach for learning classical planning action models that offers the following contribution: An inductive learning algorithm that minimizes the required input knowledge, that is flexible to different degrees of available input knowledge, and that can be defined as a classical planning compilation.

\section{Background}
This section defines the planning models used on this work.

\subsection{Classical Planning}
We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent). We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents. A {\em state} $s$ is then a total assignment of values to fluents, i.e. $|s|=|F|$, so the size of the state space $2^{|F|}$.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of positive effects $\add(a)\in\mathcal{L}(F)$, and a set of negative effects $\del(a)\in\mathcal{L}(F)$. An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\del(a))\cup\add(a)$.

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition. A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The plan $\pi$ {\em solves} $P$ if and only if $G\subseteq s_n$, i.e.~if the goal condition is satisfied following the application of $\pi$ in $I$.

In this work we assume that fluents $F$ are instantiated from predicates, as in PDDL~\cite{fox2003pddl2}. There exists a set of predicates $\Psi$, each $p\in\Psi$ with an argument list of arity $ar(p)$. Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t. $\Omega^n$ is the $n$-th Cartesian power of $\Omega$.

Likewise we assume that actions in $A$ are instantiated from operator schema. Let $\Omega_v=\{v_1,\ldots,v_v\}$, $\Omega_v\cap\Omega=\emptyset$ be a new set of objects representing variable names defined by the action with the maximum arity in a planning frame. For instance, in the blocksworld $\Omega_v=\{v_1,v_2\}$ since operators $stack$ (see Figure~\ref{fig:stack}) and $unstack$ have two parameters. Let us define a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the objects in $\Omega_v$, e.g. for the blocksworld example $F_v=\{handempty, holding(v_1), holding(v_2), clear(v_1), \\clear(v_2), ontable(v_1), ontable(v_2), on(v_1,v_1),on(v_1,v_2),\\on(v_2,v_1),on(v_2,v_2)\}$.

\begin{figure}[hbt]
\begin{footnotesize}
\begin{verbatim}
  (:action stack
   :parameters (?x1 ?x2)
   :precondition (and (holding ?x1) 
                      (clear ?x2))
   :effect (and (not (holding ?x1))
                (not (clear ?x2))
                (clear ?x1)
                (handempty)
                (on ?x1 ?x2)))
\end{verbatim}
\end{footnotesize}
 \caption{\small Example of the {\em stack} planning operator schema from the blocksworld as represented in PDDL.}
\label{fig:stack}
\end{figure}

We are ready to define an operator schema $\xi\in \Xi$ as a tuple $\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
\begin{itemize}
\item $head(\xi)$, represents the operator {\em header} defined by a pair $head(\xi)=(name(\xi),pars(\xi))$ built using an action name and a list of variables, $pars(\xi)\in\Omega_v^{ar(\xi)}$. The blocksworld only requires these four operator headers $\{pickup(v_1), putdown(v_1), stack(v_1,v_2), unstack(v_1,v_2)\}$.
\item The preconditions $pre(\xi)\subseteq F_v$, the positive effects $add(\xi)\subseteq F_v$, and the negative effects $del(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
\end{itemize}

\subsection{Classical Planning with Conditional Effects}
Conditional effects make it possible to repeatedly refer to the same action while their precise effects depend on the current state. Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$ called the {\em precondition} and a set of conditional effects $\cond(a)$. Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$ (the condition) and $E\in\mathcal{L}(F)$ (the effect).

An action $a\in A$ is applicable in state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} is
\[
\eff(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
\]
i.e.~effects whose conditions hold in $s$. The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(s,a))\cup\eff(s,a)$.

\section{Learning classical planning action models}
Learning classical planning action models from fully available input knowledge, i.e. a set of plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward: The operators schema are derived lifting the literals that change between the pre and post-state of the corresponding action executions. Preconditions are derived lifting the minimal set of literals appearing in all the pre-states that correspond to the same operator.

This section formalizes more challenging tasks, for learning classical planning action model, where less input knowledge is available. Formalization is done according to the available input knowledge.

\subsection{Learning from labeled plans}
This learning task is formalized as $\Lambda=\tup{\Psi,\Pi,\Sigma}$: 
\begin{itemize}
\item $\Psi$ the set of predicates that define the abstract state space of a given planning domain. This set includes the predicates for defining the headers of the operators schema.
\item $\Pi=\{\pi_1,\ldots,\pi_t\}$ is the given set of example plans,
\item $\Sigma=\{\sigma_1,,\ldots,\sigma_t\}$ is a set of labels s.t. each plan $\pi_i$, {\small $1\leq i\leq t$}, has an associated label $\sigma_i=(s_i,s_i')$ where $s_i'$ is the state resulting from executing $\pi_i$ starting from the state $s_i$. 
\end{itemize}

A solution to the learning task $\Lambda$ is a set of operator schema $\Xi$, with one schema for each operator header, compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.

\subsection{Learning from states}
Here we reduce the amount of input knowledge provided to the previous learning task. Now each $\Pi=\{\pi_1,\ldots,\pi_t\}$ does not contain a set of plans but the number of actions of each plan, $\Pi'=\{|\pi_1|,\ldots,|\pi_t|\}$ and redefine the learning task as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$. While the previous task can correspond to watching an agent acting in the world, this learnig tasks can be understood as watching only the results of its actions executions knowing the number of diferent actions performed by the agent.

We can go a step further and redefine the learning task as $\Lambda''=\tup{\Psi,\Sigma}$. This learnig tasks corresponds to watching only the results of the plan executions. Now a solution to the $\Lambda'$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states.

\section{Learning action models from example states using a classical planner}
Our approach for addressing $\Lambda''=\tup{\Psi,\Sigma}$ is compiling this learning task into a classical planning task $P_{\Lambda}$. The intuition behind the compilation is that a solution to $P_{\Lambda}$ is a sequence of actions that first, programs the action model (i.e. the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\Xi$) and then, validates the programmed action model in the given set of labels $\Sigma$, one after the other.

To formalize the compilation we define $t$ classical planning instances $P_1=\tup{F,A,I_1,G_1},\ldots,P_t=\tup{F,A,I_t,G_t}$, that belong to the same planning frame (share the same fluents and actions and differ only in the initial state and goals). The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the objects in $\Omega=\{o|o\in s_i\cup s_i', {\small 1\leq i\leq t}\}$, the set of objects that appear in the fluents $f\in F$ used to define the states in $\Sigma$. The set of actions is empty $A=\emptyset$. Finally the initial state $I_i$, {\small $1\leq i\leq t$}, is given by the state $s_i\in \sigma_i$ while goals $G_i$ are defined by the state $s_i'\in \sigma_i$.

Now we are ready to define the compilation for learning action models using an off-the-shelf classical planner. Given a learning task $\Lambda'=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ where:
\begin{itemize}
\item $F_{\Lambda}$ extends $F$ with:
\begin{itemize}
\item Fluents representing the programmed action model: $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ for every $f\in F_v$ and $\xi \in \Xi$.
\item Fluents $\{test_i\}_{1\leq i\leq t}$, indicating the example where the programmed model is currently being validated.
\item Fluents $prog1$, $prog2$ and $exec$ indicating whether the solution is programming the preconditions of the action schema, the effects, or it started validating the programmed action models.
\end{itemize}
\item $I_{\Lambda}$, contains the fluents from $F$ that encode the initial state $s_1\in P_1$, every fluent $pre_f(\xi)\in F_{\Lambda}$ and $prog1$. 
\item $G_{\Lambda}=\{test_i\}$,{\small $1\leq i\leq t$}, indicates that the programmed action model is validated in all the examples in $\Sigma$.
\item $A_{\Lambda}$ replaces the actions in $A$ with actions of three types:
\begin{enumerate}
\item Actions for programming:

\begin{itemize}
\item {\em Precondition} $f\in F_v$ in the action schema $\xi\in\Xi$:
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{pre_{f}(\xi),\neg prog2,\neg exec\},\\    
\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
\end{align*}
\end{small}

\item {\em Negative effect} $f\in F_v$ in the action schema $\xi\in\Xi$:
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{programDel_{f,\xi}})=&\{pre_{f}(\xi),\neg del_{f}(\xi),\neg add_{f}(\xi),\neg exec\},\\                                                   
\cond(\mathsf{programDel_{f,\xi}})=&\{\emptyset\}\rhd\{del_{f}(\xi),\\
&\{\emptyset\}\rhd\{prog2\}.
\end{align*}
\end{small}

\item {\em Positive effect} $f\in F_v$ in the action schema $\xi\in\Xi$:
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{programAdd_{f,\xi}})=&\{\neg pre_{f}(\xi)),\neg del_{f}(\xi)),\neg add_{f}(\xi),\neg exec\},\\                                                   
\cond(\mathsf{programAdd_{f,\xi}})=&\{\emptyset\}\rhd\{add_{f}(\xi),\\
&\{\emptyset\}\rhd\{prog2\}.
\end{align*}
\end{small}
\end{itemize}

\item Actions for applying an already programmed operator schema $\xi\in\Xi$ bounding it with objects $\omega\subseteq\Omega^{ar(\xi)}$
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
\cond(\mathsf{apply_{\xi,\upsilon,\upsilon'}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{\emptyset\}\rhd\{exec\}.
\end{align*}
\end{small}

For instance, these actions define that if an operator is programmed with precondition $holding(v_1)\in F_v$ it implies that $holding(block_1)\in F$ has to be true in the current state if the operator binds variable object $v_1\in\Omega_v$ with object $block_1\in\Omega_v$. The operator binding is done implicitly, i.e. variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing in the same position.

\item Actions for changing the active example where the action model is currently being validated.
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{validate_{i}})=&G_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{exec\},\\
\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i\}.
\end{align*}
\end{small}
\end{enumerate}
\end{itemize}

\begin{lemma}
Any classical plan $\pi$ that solves $P_{\Lambda}$ induces a valid action model that solves the learning task $\Lambda$.
\end{lemma}

\begin{proof}[Proof sketch]
Once an operator schema is programmed it cannot be modified and can only be applied. The only way of achieving a {\em test} fluent is by applying a sequence of programmed operator schema until achieving the goal state defined by its associated label starting from the initial state of the corresponding label. If this is done for all the labels (all the input examples) it means that the programmed action model is compliant with the learning input knowledge and hence, it is a solution to the action model learning task.
\end{proof}

The compilation is valid for partially specified action models since known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda}$. The approach allows also transfer learning where the action model generated for a given sub-task is encoded as {\em already programmed} for learning new action models in more challenging tasks.

\section{Learning action models from example plans using a classical planner}
The compilation can be extended to the learning scenario defined by $\Lambda$ and $\Lambda'$ in which a set of plans $\Pi$ (or only its lengths in the case of $\Lambda'$) is available. Each plan $\pi_i\in \Pi$, {\small $1\leq i\leq t$}, is a solution to the corresponding classical planning instance $P_i=\tup{F,A,I_i,G_i}$ defined above. The compilation extensions are: 
\begin{itemize}
\item $F_{\Lambda}$ includes the new set of fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ for encoding the $j$ steps of the {\small $1\leq i\leq t$} plans in $\Pi$ with $F_{\Pi_i}\subseteq F_{\Pi}$ the fluents encoding the plan corresponding to the $i^{th}$ example (only for the $\Lambda$ case). In addition fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, represent the plan step where the programmed action model is validated ($n$ is the max length of a plan in $\Pi$).
\item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi_1}$ that encode the plan $\pi_1\in \Pi$ for solving $P_1$, and the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the plan step where to start validating the programmed action model. Goals $G_{\Lambda}$ are like in the original compilation.
\item With respect to the actions in $A_{\Lambda}$,
\begin{enumerate}
\item The actions for programming the preconditions/effects of a given operator schema are the same.
\item The actions for applying an operator schema have an extra precondition $f\in F_{\Pi_i}$ that encodes the current plan step and extra conditional effect $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing the plan step.
\item The actions for changing the active test have an extra precondition, $at_{|\Pi_i|}$, to indicate that we simulated the full current plan $\Pi_i$ and extra conditional effects to load the next plan $\Pi_{i+1}$ where to validate the programmed action model:
\begin{small}
\begin{align*}
&\{f\}\rhd\{\neg f\}_{f\in F_{\Pi_i}},\\
&\{\emptyset\}\rhd\{f\}_{f\in F_{\Pi_i+1}},\\
&\{\emptyset\}\rhd\{\neg at_{|\pi_i|},at_1\}.
\end{align*}
\end{small}
\end{enumerate}
\end{itemize}

\section{Evaluation}

\subsection{Learning action models from example plans}
\begin{table*}[!hbt]
\begin{footnotesize}
\begin{tabular}{l||c|c|c||c|c|c||c|c|c}
handpicked/blocks 0.00 0.00 0.00 0.00 0.00 0.00
handpicked/gripper 17.78 13.70 0.00 0.00 0.00 0.00
handpicked/miconic 54.17 27.32 0.00 0.00 0.00 0.00
handpicked/visitall 0.00 0.00 0.00 0.00 0.00 0.00

\end{tabular}
\end{footnotesize}
\caption{\small Mean error and standard deviation of the learned models when using hand-picked examples and examples collected using the classical planner Fast-Downward.}
  \label{tab:eplans}
\end{table*}

The performance of our learning approach is evaluated for different degrees of available input knowledge and using different sources for collecting this input knowledge. In all the cases we assess the performance of our learning approach using the cardinality of the {\em symmetric difference} sets that are computed between the set of preconditions, del and add effects (1), in the learned model and (2), in the actual models. In all the experiments the compilation is solved using the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar}.

Table~\ref{tab:eplans} shows the mean error and standard deviation of the learned models with respect to the actual action models when (1) using {\em hand-picked} examples, (2) examples collected using the classical {\em planner} {\sc Fast-Downward}~\cite{helmert2006fast} and (3) examples collected {\em randomly}. The standard deviation provides a measure of how this error is distributed among the different operators in the domain. If this deviation is 0 it means that is equally distributed in all the domain operators.

\subsection{Learning action models from example states}
TBD.

\section{Related work}

%The  LIVE  system  (Shen  and  Simon,  1989)  was  an  extension  of  the  General  Problem  Solver  (GPS) framework (Ernst and Newell, 1969) with a learning component. LIVE alternated problem solving with model  learning  to  automatically  define  operators.  The  decision  about  when  to  alternate  depended  onsurprises,  that  is  situations  where  an  action  effects  violated  its  predicted  model.  EXPO  (Gil,  1992) generated plans with the  PRODIGY  system (Minton, 1988), monitored the plans execution, detected differences in the predicted and the observed states and constructed a set of specific hypotheses to fix those differences. Then the  EXPO  filtered the hypotheses heuristically.  OBSERVER  (Wang, 1994) learned operators by monitoring expert agents and applying the version spaces algorithm (Mitchell, 1997) to the observations. When the system already had an operator representation, the preconditions were updated by removing facts that were not present in the new observation’s pre-state; the effects were augmented by adding facts that were in the observation’s delta-state. All of these early works were based on direct liftings of the observed states. They also benefit from experience beyond simple interaction with the environment such as exploratory plans or external teachers, but  none  provided  a  theoretical  justification  for  this  second  source  of  knowledge.  The  work  recently reported in (Walsh and Littman, 2008) succeeds in bounding the number of interactions the learner must complete to learn the preconditions and effects of a STRIPS action model. This work shows that learning STRIPS  operators  from  pure  interaction  with  the  environment,  can  require  an  exponential  number  of samples, but that limiting the size of the precondition lists enable sample-efficient learning (polynomial in the number of actions and predicates of the domain). The work also proves that efficient learning is also possible without this limit if an agent has access to an external teacher that can provide solution traces on demand.

%Others systems have tried to learn more expressive action models for deterministic planning in fully observable environments. Examples would include the learning of conditional costs for AP actions (Jess Lanchas and Borrajo, 2007) or the learning of conditional effects with quantifiers (Zhuo et al., 2008).

%In addition action model learning has been studied in domains where there is partial state observability. ARMS uses the same kind od learning examples but assumes the exampls are given and proceeds in two phases. In the first phase, ARMS extracts frequent action sets from plans that share a common set of parameters. ARMS also finds some frequent literal-action pairs with the help of the initial state and the goal state that provide an initial guess on the actions preconditions, and effects. In the second phase, ARMS uses the frequent action sets and literal-action pairs to define a set of weighted constraints that must hold in order to make the plans correct. Then, ARMS solves the resulting weighted MAX-SAT problem and produces action models from the solution of the SAT problem. This process iterates until all actions are modelled. For a complex planning domain that involves hundreds of literals and actions, the corresponding weighted MAX-SAT representation is likely to be too large to be solved efficiently as the number of clauses can reach up to tens of thousands. For that reason ARMS implements a hill-climbing method that models the actions approximately. Consequently, the ARMS output is a model which may be inconsistent with the examples.

%(Amir and Chang, 2008) introduced an algorithm that tractably generates all the STRIPS-like models that  could  have  lead  to  a  set  of  observations.  Given  a  formula  representing  the  initial  belief  state,  a sequence of executed actions and the corresponding observed states(where partial observations of states are given), it builds a complete explanation of observations by models of actions through a Conjunctive Normal Form (CNF) formula. By linking the possible states of fluents to the effect propositions in the action models, the complexity of the CNF encoding can be controlled to find  exact  solutions  efficiently  in  some  circumstances.  The  learning  algorithm  updates  the  formula  of the belief state with every action and observation in the sequence. This update makes sure that the new formula represents all the transition relations consistent with the actions and observations. The formula returned at the end includes all consistent models, which can then be retrieved with additional processing. Unlike the previous approaches, the one described in (Mour ao et al., 2008) deals with both missing and noisy predicates in the observations. For each action in a given domain, they use kernel perceptrons to learn predictions of the domain properties that change because of the action execution. LOCM (Cresswellet al., 2009) induces action schemas without being provided with any information about initial, goal or intermediate state descriptions for the example action sequences. LOCM receives descriptions of plans or plan fragments, uses them to create states machines for the different domain objects and extracts the action schemas from these state machines.

\section{Conclusions}
This paper presents a novel approach for learning classical planning action models from minimal input knowledge and using exclusively existing classical planners. Learning action models from examples allows the reformulation of a domain theory. An interesting research direction is the study of domain reformulation using features that allow more compact solutions like the {\em reachable} or {\em movable} features in the Sokoban domain.

\bibliographystyle{aaai}
\bibliography{strips-learning}
\end{document}

### 2017-10-04 doc/icaps-18/strips-learning.tex
updating pages 1 and 2 for ICAPS18

@@
 \begin{abstract}
-This paper presents a novel approach for learning classical planning action models from minimal input knowledge and exclusively using existing classical planners.
-The paper defines a classical planning compilation to learn action models from examples.
-The compilation is flexible to different degrees of available input knowledge; it accepts partially specified action models and learning examples can range from a set of plans with their corresponding initial and final states to only a set of initial and final states.
+This paper presents a novel approach for learning STRIPS action models from examples that compiles the inductive learning task into a classical planning task.
+Our compilation for this learning task is flexible to different amounts of available input knowledge; it accepts partially specified action models and what is more, the input learning examples can range from a set of plans (with their corresponding initial and final states) to only a set of initial and final states where no action is observed.
 \end{abstract}
@@
 \section{Introduction}
-Off-the-shelf planners reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
-In addition planning action models allow to effectively address further tasks than plan synthesis, like plan/goal recognition~\cite{ramirez2010probabilistic}.
-Unfortunately building planning action models is complex, even for planning experts, limiting the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.
+Besides plan synthesis, planning action models are also useful for plan/goal recognition~\cite{ramirez2010probabilistic}.
+In both tasks, off-the-shelf planners require reasoning about action models that correctly and completely capture the possible world transitions~\cite{ghallab2004automated,geffner:book:2013}.
+Unfortunately, building such planning action models is complex, even for planning experts, so this knowledge acquisition bottleneck limits the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.
 
-Despite many Machine Learning (ML) techniques are able to compute models from examples~\cite{michalski2013machine} its application to learning planning action models is not straightforward.
-First, the inputs to ML algorithms usually are finite numeric vectors encoding objects features while, in planning, the input includes actions so is closer to the representation of procedures and behaviours rather than to object representation.
+On the other hand, Machine Learning (ML) techniques are able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
+However, the application of inductive ML techniques for learning planning action models is not straightforward.
+First, the inputs to ML algorithms usually are sets of finite numeric vectors encoding the value of sets of objects features.
+The input for learning planning action models traditionally are sets of observations of plan executions (each with possibly different length).
 Second, the traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
-In the case of learning planning action models the output is not a scalar but a model of the possible state transitions.
-Last but not least, collecting {\em informative} examples for learning planning action models is challenging.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+In the case of learning STRIPS action models, the output is not a scalar but a model of the preconditions and the effects of each action that defines the possible state transitions in the given planning domain.
 
-Learning classical planning action models is a well-studied problem with sophisticated algorithms, like ARMS~\cite{yang2007learning}, SLAF~\cite{amir:alearning:JAIR08} or LOCM~\cite{cresswell2013acquiring} that do not require full knowledge of all the states traversed by the example plans.
-Motivated by recent advances on learning generative models with classical planning~\cite{segovia2017generating} and on the effective exploration of planning state spaces~\cite{geffner:novelty:IJCAI17}, this paper introduces an innovative approach for learning classical planning action models that offers the following contribution: An inductive learning algorithm that minimizes the required input knowledge, that is flexible to different degrees of available input knowledge, and that can be defined as a classical planning compilation.
+Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of all the states traversed by the example plans.
+Motivated by recent advances on learning generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating} this paper introduces an innovative approach for learning classical planning action models that (1) it can be defined as a classical planning compilation and (2), it is flexible to different amounts of available input knowledge.
 
@@
 $|s|=|F|$, so the size of the state space $2^{|F|}$.
+Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions, but we often abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
 
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of positive effects $\add(a)\in\mathcal{L}(F)$, and a set of negative effects $\del(a)\in\mathcal{L}(F)$.
-An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\del(a))\cup\add(a)$.
+Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of positive effects $\eff^+(a)\in\mathcal{L}(F)$, and a set of negative effects $\eff^-(a)\in\mathcal{L}(F)$.
+An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \eff^-(a))\cup\eff^+(a)$.
 
@@
 
-In this work we assume that fluents $F$ are instantiated from predicates, as in PDDL~\cite{fox2003pddl2}.
+In this work we assume that fluents $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
 There exists a set of predicates $\Psi$, each $p\in\Psi$ with an argument list of arity $ar(p)$.
@@
 
-Likewise we assume that actions in $A$ are instantiated from operator schema.
-Let $\Omega_v=\{v_1,\ldots,v_v\}$, $\Omega_v\cap\Omega=\emptyset$ be a new set of objects representing variable names defined by the action with the maximum arity in a planning frame.
-For instance, in the blocksworld $\Omega_v=\{v_1,v_2\}$ since operators $stack$ (see Figure~\ref{fig:stack}) and $unstack$ have two parameters.
+Likewise, we assume that actions in $A$ are instantiated from an STRIPS operator schema.
+Let $\Omega_v=\{v_1,\ldots,v_v\}$, $\Omega_v\cap\Omega=\emptyset$ be a new set of objects representing {\em variable names} and that is defined by the action with the maximum arity in a planning frame.
+For instance, in a three-block blocksworld then $\Omega=\{block_1, block_2, block_3\}$ and $\Omega_v=\{v_1, v_2\}$ since the actions instantiated from the operators {\small\tt stack} (see Figure~\ref{fig:stack}) and {\small\tt unstack} have two parameters.
 Let us define a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the objects in $\Omega_v$, e.g.
-for the blocksworld example $F_v=\{handempty, holding(v_1), holding(v_2), clear(v_1), \\clear(v_2), ontable(v_1), ontable(v_2), on(v_1,v_1),on(v_1,v_2),\\on(v_2,v_1),on(v_2,v_2)\}$.
+for the blocksworld example $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.
 
@@
 \begin{footnotesize}
+\begin{scriptsize}
 \begin{verbatim}
@@
 \end{verbatim}
+\end{scriptsize}
 \end{footnotesize}
@@
 
-We are ready to define an operator schema $\xi\in \Xi$ as a tuple $\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+We are now ready to define a STRIPS operator schema as a tuple $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
 \begin{itemize}
-\item $head(\xi)$, represents the operator {\em header} defined by a pair $head(\xi)=(name(\xi),pars(\xi))$ built using an action name and a list of variables, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
-The blocksworld only requires these four operator headers $\{pickup(v_1), putdown(v_1), stack(v_1,v_2), unstack(v_1,v_2)\}$.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, represents an operator {\em header} defined by its corresponding action name and a list of variables, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
+Learning the action model for the blocksworld requires these four operator headers, $\Xi_{blocks}=${\small\tt\{pickup($v_1$), putdown($v_1$), stack($v_1,v_2$), unstack($v_1,v_2$)\}}.
 \item The preconditions $pre(\xi)\subseteq F_v$, the positive effects $add(\xi)\subseteq F_v$, and the negative effects $del(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
@@
 \subsection{Classical Planning with Conditional Effects}
-Conditional effects make it possible to repeatedly refer to the same action while their precise effects depend on the current state.
+Our approach for leaning STRIPS operator schema is compiling this leaning task into a classical planning task with conditional effects.
+Conditional effects make it possible to refer to an action while their precise effects depend on the current state.
+
 Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$ called the {\em precondition} and a set of conditional effects $\cond(a)$.
@@
 \]
-i.e.~effects whose conditions hold in $s$. The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(s,a))\cup\eff(s,a)$.
+i.e.~effects whose conditions hold in $s$.
+The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \eff^-(s,a))\cup\eff^+(s,a)$, where $\eff^-(s,a)$ and $\eff^+(s,a)$ are the negative and positive effects in $\eff(s,a)$.
 
 \section{Learning classical planning action models}
-Learning classical planning action models from fully available input knowledge, i.e.
-a set of plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward: The operators schema are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+Learning STRIPS action models from fully available input knowledge, i.e.
+a set of plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
+In this case, the operators schema are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
 Preconditions are derived lifting the minimal set of literals appearing in all the pre-states that correspond to the same operator.
 
-This section formalizes more challenging tasks, for learning classical planning action model, where less input knowledge is available. Formalization is done according to the available input knowledge.
+This section formalizes more challenging tasks, for learning STRIPS action models, where less input knowledge is available.
+Next we formalize these learning tasks according to the available input knowledge.
 
@@
 \item $\Psi$ the set of predicates that define the abstract state space of a given planning domain. This set includes the predicates for defining the headers of the operators schema.
-\item $\Pi=\{\pi_1,\ldots,\pi_t\}$ is the given set of example plans,
-\item $\Sigma=\{\sigma_1,,\ldots,\sigma_t\}$ is a set of labels s.t.
-each plan $\pi_i$, {\small $1\leq i\leq t$}, has an associated label $\sigma_i=(s_i,s_i')$ where $s_i'$ is the state resulting from executing $\pi_i$ starting from the state $s_i$.
+\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given set of example plans,
+\item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$ is a set of labels s.t.
+each plan $\pi_t$, {\small $1\leq t\leq \tau$}, has an associated label $\sigma_t=(s_0^t,s_{n+1}^t)$ where $s_{n+1}^t$ is the state resulting from executing $\pi_t$ starting from the state $s_0^t$.
 \end{itemize}
 
-A solution to the learning task $\Lambda$ is a set of operator schema $\Xi$, with one schema for each operator header, compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
+A solution to the learning task $\Lambda$ is a set of operator schema $\Xi$ (one schema for each operator header) compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
 
@@
 Here we reduce the amount of input knowledge provided to the previous learning task.
-Now each $\Pi=\{\pi_1,\ldots,\pi_t\}$ does not contain a set of plans but the number of actions of each plan, $\Pi'=\{|\pi_1|,\ldots,|\pi_t|\}$ and redefine the learning task as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$.
+Now $\Pi=\{\pi_1,\ldots,\pi_t\}$ is replaced by  $\Pi'=\{|\pi_1|,\ldots,|\pi_t|\}$ that does not contain a set of plans but the number of actions of each plan, so the learning task is redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$.
 While the previous task can correspond to watching an agent acting in the world, this learnig tasks can be understood as watching only the results of its actions executions knowing the number of diferent actions performed by the agent.
 
-We can go a step further and redefine the learning task as $\Lambda''=\tup{\Psi,\Sigma}$.
-This learnig tasks corresponds to watching only the results of the plan executions.
-Now a solution to the $\Lambda'$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states.
+Finally, we can go a step further and redefine a third learning task $\Lambda''=\tup{\Psi,\Sigma}$ that corresponds to watching only the results of the plan executions.
+Now a solution to the $\Lambda''$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states.
 
@@
 \begin{tabular}{l||c|c|c||c|c|c||c|c|c}
-handpicked/blocks 0.00 0.00 0.00 0.00 0.00 0.00
-handpicked/gripper 17.78 13.70 0.00 0.00 0.00 0.00
-handpicked/miconic 54.17 27.32 0.00 0.00 0.00 0.00
-handpicked/visitall 0.00 0.00 0.00 0.00 0.00 0.00
 
@@
 
+Last but not least, collecting {\em informative} examples for learning planning action models is challenging.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+In addition, motivated by the success of recent algorithms for exploring planning tasks~\cite{geffner:pwithsimulators:IJCAI17}, we do not assume that a learning set of plans is given apriori but instead, we autonomously collect the learning examples.
+
 \bibliographystyle{aaai}

### 2017-10-05 doc/icaps-18/strips-learning.tex
updates to page 2 icaps18 paper

@@
 
-\subsection{Classical Planning}
+\subsection{Classical planning}
 We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
@@
 
-In this work we assume that fluents $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
+In this work we assume that the fluents in $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
 There exists a set of predicates $\Psi$, each $p\in\Psi$ with an argument list of arity $ar(p)$.
 Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^n$ is the $n$-th Cartesian power of $\Omega$.
+$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
 
-Likewise, we assume that actions in $A$ are instantiated from an STRIPS operator schema.
-Let $\Omega_v=\{v_1,\ldots,v_v\}$, $\Omega_v\cap\Omega=\emptyset$ be a new set of objects representing {\em variable names} and that is defined by the action with the maximum arity in a planning frame.
-For instance, in a three-block blocksworld then $\Omega=\{block_1, block_2, block_3\}$ and $\Omega_v=\{v_1, v_2\}$ since the actions instantiated from the operators {\small\tt stack} (see Figure~\ref{fig:stack}) and {\small\tt unstack} have two parameters.
-Let us define a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the objects in $\Omega_v$, e.g.
-for the blocksworld example $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.
-
+Likewise, we assume that each action in $A$ is instantiated from an STRIPS operator schema. Figure~\ref{fig:stack} shows the {\em stack} operator schema for a STRIPS blocksworld represented in PDDL.
 \begin{figure}[hbt]
@@
   :parameters (?x1 ?x2)
-   :precondition (and (holding ?x1) 
-                      (clear ?x2))
-   :effect (and (not (holding ?x1))
-                (not (clear ?x2))
-                (clear ?x1)
-                (handempty)
-                (on ?x1 ?x2)))
+  :precondition (and (holding ?x1) (clear ?x2))
+  :effect (and (not (holding ?x1)) (not (clear ?x2))
+               (clear ?x1) (handempty) (on ?x1 ?x2)))
 \end{verbatim}
@@
 \end{footnotesize}
- \caption{\small Example of the {\em stack} planning operator schema from the blocksworld as represented in PDDL.}
+ \caption{\small Example of a {\em stack} operator schema for a STRIPS blocksworld represented in PDDL.}
 \label{fig:stack}
@@
 
-We are now ready to define a STRIPS operator schema as a tuple $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represent {\em variable names}.
+$|\Omega_v|$ is given by the action with the maximum arity in a planning frame.
+For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ and $\Omega_v=\{v_1, v_2\}$ because the operators {\small\tt stack} and {\small\tt unstack} are the ones with the maximum arity (two parameters each).
+
+Let us define a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} $\Omega_v$.
+In the blocksworld $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.
+
+We are now ready to define a STRIPS operator schema as a tuple $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$:
 \begin{itemize}
 \item $head(\xi)=\tup{name(\xi),pars(\xi)}$, represents an operator {\em header} defined by its corresponding action name and a list of variables, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
-Learning the action model for the blocksworld requires these four operator headers, $\Xi_{blocks}=${\small\tt\{pickup($v_1$), putdown($v_1$), stack($v_1,v_2$), unstack($v_1,v_2$)\}}.
+The headers for the blocksworld operators are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
 \item The preconditions $pre(\xi)\subseteq F_v$, the positive effects $add(\xi)\subseteq F_v$, and the negative effects $del(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
@@
 
-\subsection{Classical Planning with Conditional Effects}
-Our approach for leaning STRIPS operator schema is compiling this leaning task into a classical planning task with conditional effects.
-Conditional effects make it possible to refer to an action while their precise effects depend on the current state.
+\subsection{Classical planning with conditional effects}
+Our approach for leaning STRIPS action models is compiling this leaning task into a classical planning task with conditional effects.
+We use conditional effects because they allow us to compactly define actions whose effects depend on the current state.
+Many classical planners cope with conditional effects without compiling them away.
+In fact, the support of PDDL conditional effects was a requirement for participating at IPC-2014~\cite{vallati:IPC:AIM2015}.
 
@@
 
-\section{Learning classical planning action models}
+\section{Learning STRIPS action models}
 Learning STRIPS action models from fully available input knowledge, i.e.
@@
 In this case, the operators schema are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Preconditions are derived lifting the minimal set of literals appearing in all the pre-states that correspond to the same operator.
+Likewise, preconditions are derived lifting the minimal set of literals appearing in all the pre-states that correspond to the same operator.
 
@@
 \begin{itemize}
-\item $\Psi$ the set of predicates that define the abstract state space of a given planning domain. This set includes the predicates for defining the headers of the operators schema.
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given set of example plans,
-\item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$ is a set of labels s.t.
-each plan $\pi_t$, {\small $1\leq t\leq \tau$}, has an associated label $\sigma_t=(s_0^t,s_{n+1}^t)$ where $s_{n+1}^t$ is the state resulting from executing $\pi_t$ starting from the state $s_0^t$.
+\item $\Psi$, the set of predicates that define the abstract state space of a given planning domain. This set includes the predicates for defining the headers of the operators schema.
+\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans.
+\item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$, a set of labels s.t.
+each plan $\pi_t$, {\small $1\leq t\leq \tau$}, has a label $\sigma_t=(s_0^t,s_{n}^t)$ where $s_{n}^t$ is the state resulting from executing $\pi_t$ starting from the state $s_0^t$.
 \end{itemize}
@@
 
-\subsection{Learning from states}
-Here we reduce the amount of input knowledge provided to the previous learning task.
-Now $\Pi=\{\pi_1,\ldots,\pi_t\}$ is replaced by  $\Pi'=\{|\pi_1|,\ldots,|\pi_t|\}$ that does not contain a set of plans but the number of actions of each plan, so the learning task is redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$.
-While the previous task can correspond to watching an agent acting in the world, this learnig tasks can be understood as watching only the results of its actions executions knowing the number of diferent actions performed by the agent.
+\subsection{Learning from initial/final states}
+Here we reduce the amount of input knowledge provided to the learning task.
+Now $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is replaced by  $\Pi'=\{|\pi_1|,\ldots,|\pi_{\tau}|\}$ i.e.
+$\Pi'$ that does not contain a set of plans but the number of actions of each plan, so the learning task is redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$.
+While the previous learning task, $\Lambda$, corresponds to watching an agent acting in the world, this new learning task $\Lambda'$ can be understood as watching only the results of its actions executions knowing the number of diferent actions performed by the agent.
 
 Finally, we can go a step further and redefine a third learning task $\Lambda''=\tup{\Psi,\Sigma}$ that corresponds to watching only the results of the plan executions.
-Now a solution to the $\Lambda''$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states.
+In this case a solution to the $\Lambda''$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
 
-\section{Learning action models from example states using a classical planner}
-Our approach for addressing $\Lambda''=\tup{\Psi,\Sigma}$ is compiling this learning task into a classical planning task $P_{\Lambda}$.
-The intuition behind the compilation is that a solution to $P_{\Lambda}$ is a sequence of actions that first, programs the action model (i.e.
-the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\Xi$) and then, validates the programmed action model in the given set of labels $\Sigma$, one after the other.
+In these two cases, a solution must not only synthesize the action models but also the actions that produced the given labels (this information is no longer given in the learning examples).
 
-To formalize the compilation we define $t$ classical planning instances $P_1=\tup{F,A,I_1,G_1},\ldots,P_t=\tup{F,A,I_t,G_t}$, that belong to the same planning frame (share the same fluents and actions and differ only in the initial state and goals).
-The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the objects in $\Omega=\{o|o\in s_i\cup s_i', {\small 1\leq i\leq t}\}$, the set of objects that appear in the fluents $f\in F$ used to define the states in $\Sigma$.
-The set of actions is empty $A=\emptyset$.
-Finally the initial state $I_i$, {\small $1\leq i\leq t$}, is given by the state $s_i\in \sigma_i$ while goals $G_i$ are defined by the state $s_i'\in \sigma_i$.
+\section{Learning STRIPS action models with classical planning}
+Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is to compile it into a classical planning task $P_{\Lambda}$, $P_{\Lambda'}$ or $P_{\Lambda''}$.
+The intuition behind these compilations is that a solution to the resulting classical planning task is a sequence of actions that:
+\begin{enumerate}
+\item Programs the action model i.e. determines the literals in the sets $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\Xi$.
+\item Validates the programmed action model in the given set of labels $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ i.e., for every {\small $1\leq t\leq \tau$}, uses the programmed action model $\Xi$ to produce a final state $s_{n}^t$ starting from its corresponding initial state $s_0^t$.
+\end{enumerate}
 
-Now we are ready to define the compilation for learning action models using an off-the-shelf classical planner.
-Given a learning task $\Lambda'=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ where:
+To formalize these compilations we define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,A,I_t,G_t}$, that belong to the same planning frame (share the same fluents and actions and differ only in the initial state and goals).
+The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the objects appearing in the labels.
+Formally $\Omega=\{o|o\in s_0^t\cup s_n^t\cup\pi_t, {\small 1\leq t\leq \tau}\}$.
+The set of actions is empty $A=\emptyset$, is the aim of the learning tasks adressed in the paper.
+Finally the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$ are defined by the state $s_n^t\in \sigma_t$.
+
+Now we are ready to define the compilation for learning STRIPS action models using classical planning with conditional effects.
+Given a learning task $\Lambda''=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda''}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ where:
 \begin{itemize}
@@
 \item Fluents representing the programmed action model: $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ for every $f\in F_v$ and $\xi \in \Xi$.
-\item Fluents $\{test_i\}_{1\leq i\leq t}$, indicating the example where the programmed model is currently being validated.
-\item Fluents $prog1$, $prog2$ and $exec$ indicating whether the solution is programming the preconditions of the action schema, the effects, or it started validating the programmed action models.
+\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the programmed model is currently being validated.
+\item Fluents $mode_{pre}$, $mode_{eff}$ and $mode_{val}$ indicating whether currently the preconditions of the action schema are being programmed, the effects of the action schema are being programmed or the programmed action models are being validated.
 \end{itemize}
-\item $I_{\Lambda}$, contains the fluents from $F$ that encode the initial state $s_1\in P_1$, every fluent $pre_f(\xi)\in F_{\Lambda}$ and $prog1$. 
-\item $G_{\Lambda}=\{test_i\}$,{\small $1\leq i\leq t$}, indicates that the programmed action model is validated in all the examples in $\Sigma$.
-\item $A_{\Lambda}$ replaces the actions in $A$ with actions of three types:
+\item $I_{\Lambda}$, contains the fluents from $F$ that encode $s_0^1$, every fluent $pre_f(\xi)\in F_{\Lambda}$ and $mode_{pre}$. 
+\item $G_{\Lambda}=\{test_t\}$,{\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
+\item $A_{\Lambda}$ contains actions of three types:
 \begin{enumerate}

### 2017-10-05 src/compiler.py
optimization added to the compiler

@@ -110,7 +110,7 @@ predicates = get_predicates_schema_from_plans(fd_task)
 # Compilation Problem
 init_aux=copy.deepcopy(fd_task.init)
 fd_task.init=[]
-fd_task.init.append(pddl.conditions.Atom("programming1",[]))
+fd_task.init.append(pddl.conditions.Atom("modePre",[]))
 allpres=[]
 for a in actions: # All possible preconditions are initially programmed
    var_ids=[]
@@ -145,9 +145,9 @@ if input_level <= config.INPUT_LENPLAN:
    for i in range(1,MAX_STEPS+2):
       fd_task.objects.append(pddl.pddl_types.TypedObject("i"+str(i),"step"))   
    
-fd_task.predicates.append(pddl.predicates.Predicate("programming1",[]))
-fd_task.predicates.append(pddl.predicates.Predicate("programming2",[]))
-fd_task.predicates.append(pddl.predicates.Predicate("executing",[]))
+fd_task.predicates.append(pddl.predicates.Predicate("modePre",[]))
+fd_task.predicates.append(pddl.predicates.Predicate("modeEff",[]))
+fd_task.predicates.append(pddl.predicates.Predicate("modeVal",[]))
 for i in range(0,len(plans)+1):
    fd_task.predicates.append(pddl.predicates.Predicate("test"+str(i),[]))
 if input_level <= config.INPUT_LENPLAN:                  
@@ -174,7 +174,7 @@ for a in old_actions:
       params=params+[pddl.pddl_types.TypedObject("?i1","step")]
       params=params+[pddl.pddl_types.TypedObject("?i2","step")]
 
-   pre = [pddl.conditions.Atom("executing",[])]
+   pre = [pddl.conditions.Atom("modeVal",[])]
    if input_level <= config.INPUT_PLANS and input_level <config.INPUT_MINIMUM:
       pre = pre + [pddl.conditions.Atom("plan-"+a[0],["?i1"]+["?o"+str(i) for i in range(1,len(a))])]
       
@@ -225,34 +225,27 @@ for a in old_actions:
             vars = ["var"+str(t) for t in tup]
             params = []
             pre = []
-            pre = pre + [pddl.conditions.Atom("programming1",[])]
-            pre = pre + [pddl.conditions.NegatedAtom("programming2",[])]
-            pre = pre + [pddl.conditions.NegatedAtom("executing",[])]                        
[…]

### 2017-10-05 src/compiler.py
actions rename to correspond better the ICAPS18 paper

@@ -263,7 +263,7 @@ if input_level <= config.INPUT_STEPS:
       eff = eff + [pddl.effects.Effect([],pddl.conditions.Truth(),pddl.conditions.Atom("plan-"+name,["i"+str(i+1)]+params))]
       
 eff = eff + [pddl.effects.Effect([],pddl.conditions.Truth(),pddl.conditions.Atom("modeVal",[]))]
-fd_task.actions.append(pddl.actions.Action("test_0",[],0,pddl.conditions.Conjunction(pre),eff,0))
+fd_task.actions.append(pddl.actions.Action("validate_0",[],0,pddl.conditions.Conjunction(pre),eff,0))
 
 for i in range(0,len(plans)):   
    pre = []
@@ -296,7 +296,7 @@ for i in range(0,len(plans)):
             pars = ["i"+str(j+1)]+plans[i+1][j].replace("(","").replace(")","").split(" ")[1:]
             eff = eff + [pddl.effects.Effect([],pddl.conditions.Truth(),pddl.conditions.Atom(name,pars))]      
       
-   fd_task.actions.append(pddl.actions.Action("test_"+str(i+1),[],0,pddl.conditions.Conjunction(pre),eff,0))
+   fd_task.actions.append(pddl.actions.Action("validate_"+str(i+1),[],0,pddl.conditions.Conjunction(pre),eff,0))
 
    
 # Writing the compilation output domain and problem

### 2017-10-05 doc/icaps-18/strips-learning.tex
updated page 3 ICAPS18 paper

@@
 \begin{enumerate}
-\item Programs the action model i.e. determines the literals in the sets $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\Xi$.
-\item Validates the programmed action model in the given set of labels $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ i.e., for every {\small $1\leq t\leq \tau$}, uses the programmed action model $\Xi$ to produce a final state $s_{n}^t$ starting from its corresponding initial state $s_0^t$.
+\item Programs the STRIPS action model. For each $\xi\in\Xi$, determines the literals that belong to the sets $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
+\item Validates the programmed action model with the given set of labels $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$.
+For every {\small $1\leq t\leq \tau$}, the programmed action model $\Xi$ is used to produce the final states $s_{n}^t$ starting from their corresponding initial state $s_0^t$.
 \end{enumerate}
 
-To formalize these compilations we define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,A,I_t,G_t}$, that belong to the same planning frame (share the same fluents and actions and differ only in the initial state and goals).
+To formalize these compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, that belong to the same planning frame (i.e.
+share the same fluents and actions and differ only in the initial state and goals).
 The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the objects appearing in the labels.
 Formally $\Omega=\{o|o\in s_0^t\cup s_n^t\cup\pi_t, {\small 1\leq t\leq \tau}\}$.
-The set of actions is empty $A=\emptyset$, is the aim of the learning tasks adressed in the paper.
-Finally the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$ are defined by the state $s_n^t\in \sigma_t$.
+The set of actions is empty $A=\emptyset$, this is the aim of the learning tasks adressed in the paper.
+Finally the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.
 
-Now we are ready to define the compilation for learning STRIPS action models using classical planning with conditional effects.
-Given a learning task $\Lambda''=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda''}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ where:
+Now we are ready to formalize our compilations for learning STRIPS action models using classical planning with conditional effects.
+We start with $\Lambda''$ that is the learning task with the minimum amount of input knowledge and incrementally extend the formmilized compilation until addressing $\Lambda$.
+Given a learning task $\Lambda''=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda''}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
 \begin{itemize}
-\item $F_{\Lambda}$ extends $F$ with:
+\item $F_{\Lambda}$ is an extension of $F$ with:
 \begin{itemize}
 \item Fluents representing the programmed action model: $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ for every $f\in F_v$ and $\xi \in \Xi$.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect of the operator $\xi$.
 \item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the programmed model is currently being validated.
-\item Fluents $mode_{pre}$, $mode_{eff}$ and $mode_{val}$ indicating whether currently the preconditions of the action schema are being programmed, the effects of the action schema are being programmed or the programmed action models are being validated.
+\item Fluents $mode_{pre}$, $mode_{eff}$ and $mode_{val}$ indicating whether the preconditions of the action schema are being programmed, the effects of the action schema are being programmed or the programmed action models are being validated.
 \end{itemize}
-\item $I_{\Lambda}$, contains the fluents from $F$ that encode $s_0^1$, every fluent $pre_f(\xi)\in F_{\Lambda}$ and $mode_{pre}$. 
+\item $I_{\Lambda}$, contains the fluents from $F$ that encode $s_0^1$, every fluent $pre_f(\xi)\in F_{\Lambda}$ (initially all operators have all the possible preconditions) and $mode_{pre}$. 
 \item $G_{\Lambda}=\{test_t\}$,{\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
@@
 \begin{enumerate}
-\item Actions for programming:
+\item The actions for programming the operator schema. This includes the actions for removing a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
 
-\begin{itemize}
-\item {\em Precondition} $f\in F_v$ in the action schema $\xi\in\Xi$:
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{pre_{f}(\xi),\neg prog2,\neg exec\},\\    
+\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{pre_{f}(\xi),mode_{pre}, \\
+&\neg mode_{eff}, \neg mode_{val}\},\\    
 \cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
@@
 
-\item {\em Negative effect} $f\in F_v$ in the action schema $\xi\in\Xi$:
+Actions for adding a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
+
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programDel_{f,\xi}})=&\{pre_{f}(\xi),\neg del_{f}(\xi),\neg add_{f}(\xi),\neg exec\},\\                                                   
-\cond(\mathsf{programDel_{f,\xi}})=&\{\emptyset\}\rhd\{del_{f}(\xi),\\
-&\{\emptyset\}\rhd\{prog2\}.
+\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\                                                   
+&\neg mode_{val}\},\\
+\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
+&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\},\\
+&\{mode_{pre}\}\rhd\{\neg mode_{pre},mode_{eff}\}.
 \end{align*}
@@
 
-\item {\em Positive effect} $f\in F_v$ in the action schema $\xi\in\Xi$:
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programAdd_{f,\xi}})=&\{\neg pre_{f}(\xi)),\neg del_{f}(\xi)),\neg add_{f}(\xi),\neg exec\},\\                                                   
-\cond(\mathsf{programAdd_{f,\xi}})=&\{\emptyset\}\rhd\{add_{f}(\xi),\\
-&\{\emptyset\}\rhd\{prog2\}.
-\end{align*}
-\end{small}
-\end{itemize}
-
-\item Actions for applying an already programmed operator schema $\xi\in\Xi$ bounding it with objects $\omega\subseteq\Omega^{ar(\xi)}$
+\item The actions for applying an already programmed operator schema $\xi\in\Xi$ bounding it with objects $\omega\subseteq\Omega^{ar(\xi)}$
 \begin{small}
@@
 \hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-\cond(\mathsf{apply_{\xi,\upsilon,\upsilon'}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
 &\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{\emptyset\}\rhd\{exec\}.
+&\{mode_{pre}\}\rhd\{\neg mode_{pre},mode_{val}\},\\
+&\{mode_{eff}\}\rhd\{\neg mode_{eff},mode_{val}\}.
 \end{align*}
@@
 
-For instance, these actions define that if an operator is programmed with precondition $holding(v_1)\in F_v$ it implies that $holding(block_1)\in F$ has to be true in the current state if the operator binds variable object $v_1\in\Omega_v$ with object $block_1\in\Omega_v$.
+For instance, these actions define that if an operator is programmed with the precondition $holding(v_1)\in F_v$ it {\em implies} ($\implies$) that $holding(block_1)\in F$ has to be true in the current state if the operator binds the variable object $v_1\in\Omega_v$ with the regular object $block_1\in\Omega$.
 The operator binding is done implicitly, i.e.
@@
 
-\item Actions for changing the active example where the action model is currently being validated.
+\item The actions for changing the learning example {\small $1\leq t\leq \tau$} where the programmed action model is validated. 
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&G_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{exec\},\\
-\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i\}.
+\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\cup\{mode_{val}\},\\
+\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
 \end{align*}
@@
 \begin{proof}[Proof sketch]
-Once an operator schema is programmed it cannot be modified and can only be applied.
-The only way of achieving a {\em test} fluent is by applying a sequence of programmed operator schema until achieving the goal state defined by its associated label starting from the initial state of the corresponding label.
-If this is done for all the labels (all the input examples) it means that the programmed action model is compliant with the learning input knowledge and hence, it is a solution to the action model learning task.
+Once the preconditions of an operator schema $\Xi$ are programmed they cannot be modified.
+The same happens with the positive and negative effects (that can only be programmed after all the preconditions are programmed).
+Furthermore, once an operator schema is programmed it can only be applied.
+The only way of achieving a fluent $\{test_t\}$,{\small $1\leq t\leq \tau$} is by executing an applicable sequence of programmed operator schema that achieves the goal state defined by its associated label $\sigma_t$ starting from the initial state of the corresponding label.
+If this is done for all the labels (all the input examples of the learning task) it means that the programmed action model $\Xi$ is compliant with the learning input knowledge and hence, it is a solution to the action model learning task.
 \end{proof}
 
-The compilation is valid for partially specified action models since known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda}$.
-The approach allows also transfer learning where the action model generated for a given sub-task is encoded as {\em already programmed} for learning new action models in more challenging tasks.
+Interestingly, the compilation is valid for partially specified action models since known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda}$ and the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) be removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier.
 
-\section{Learning action models from example plans using a classical planner}
+\section{Constraining the hypothesis space with example plans}
 The compilation can be extended to the learning scenario defined by $\Lambda$ and $\Lambda'$ in which a set of plans $\Pi$ (or only its lengths in the case of $\Lambda'$) is available.
@@
 An interesting research direction is the study of domain reformulation using features that allow more compact solutions like the {\em reachable} or {\em movable} features in the Sokoban domain.
+The size of the compilation output depends also on the number of examples. Empirical results show that our approach is able to generate non-trivial CFGs from very small data sets.
 

### 2017-10-06 doc/icaps-18/strips-learning.tex
updates to page 4 icaps18 paper

@@
 \item Validates the programmed action model with the given set of labels $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$.
-For every {\small $1\leq t\leq \tau$}, the programmed action model $\Xi$ is used to produce the final states $s_{n}^t$ starting from their corresponding initial state $s_0^t$.
+For every {\small $1\leq t\leq \tau$}, the programmed action model $\Xi$ is used to produce a final state $s_{n}^t$ starting from their corresponding initial state $s_0^t$.
 \end{enumerate}
 
-To formalize these compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, that belong to the same planning frame (i.e.
+To formalize our compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, that belong to the same planning frame (i.e.
 share the same fluents and actions and differ only in the initial state and goals).
-The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the objects appearing in the labels.
-Formally $\Omega=\{o|o\in s_0^t\cup s_n^t\cup\pi_t, {\small 1\leq t\leq \tau}\}$.
-The set of actions is empty $A=\emptyset$, this is the aim of the learning tasks adressed in the paper.
+The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of objects appearing in the example plans or their corresponding labels.
+Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(\pi_t)\cup obj(s_0^t)\cup obj(s_n^t)\}$.
+The set of actions is empty, $A=\emptyset$, i.e.
+the action model is unkown.
 Finally the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.
@@
 Now we are ready to formalize our compilations for learning STRIPS action models using classical planning with conditional effects.
-We start with $\Lambda''$ that is the learning task with the minimum amount of input knowledge and incrementally extend the formmilized compilation until addressing $\Lambda$.
-Given a learning task $\Lambda''=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda''}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+We start with $\Lambda''$ because it is the learning task that requires the least amount of input knowledge and incrementally extend the formalized compilation until addressing $\Lambda$, the learning task with the largest amount of input knowledge.
+
+Given a learning task $\Lambda''=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda''}=\tup{F_{\Lambda''},A_{\Lambda''},I_{\Lambda''},G_{\Lambda''}}$ such that:
 \begin{itemize}
-\item $F_{\Lambda}$ is an extension of $F$ with:
+\item $F_{\Lambda''}$ extends $F$ with:
 \begin{itemize}
-\item Fluents representing the programmed action model: $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ for every $f\in F_v$ and $\xi \in \Xi$.
+\item Fluents representing the programmed action model ($pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ for every $f\in F_v$ and $\xi \in \Xi$).
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect of the operator $\xi$.
-\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the programmed model is currently being validated.
+\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the programmed model is being validated.
 \item Fluents $mode_{pre}$, $mode_{eff}$ and $mode_{val}$ indicating whether the preconditions of the action schema are being programmed, the effects of the action schema are being programmed or the programmed action models are being validated.
 \end{itemize}
-\item $I_{\Lambda}$, contains the fluents from $F$ that encode $s_0^1$, every fluent $pre_f(\xi)\in F_{\Lambda}$ (initially all operators have all the possible preconditions) and $mode_{pre}$. 
-\item $G_{\Lambda}=\{test_t\}$,{\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
-\item $A_{\Lambda}$ contains actions of three types:
+\item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (i.e.
+the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (initially all operators have all the possible preconditions) and $mode_{pre}$.
+\item $G_{\Lambda''}=\{test_t\}$,{\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
+\item $A_{\Lambda''}$ contains actions of three different types:
 \begin{enumerate}
@@
 
-Actions for adding a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
+The actions for adding a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\cup\{mode_{val}\},\\
+\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\cup\\
+&\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{mode_{val}\},\\
 \cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
@@
 \begin{lemma}
-Any classical plan $\pi$ that solves $P_{\Lambda}$ induces a valid action model that solves the learning task $\Lambda$.
+Any classical plan $\pi$ that solves $P_{\Lambda''}$ induces a valid action model that solves the learning task $\Lambda''$.
 \end{lemma}
@@
 \begin{proof}[Proof sketch]
-Once the preconditions of an operator schema $\Xi$ are programmed they cannot be modified.
-The same happens with the positive and negative effects (that can only be programmed after all the preconditions are programmed).
+\begin{small}
+Once the preconditions of an operator schema $\Xi$ are programmed, they cannot be altered.
+The same happens with the positive and negative effects that define an operator schema (besides they can only be programmed after the operator preconditions are programmed).
 Furthermore, once an operator schema is programmed it can only be applied.
-The only way of achieving a fluent $\{test_t\}$,{\small $1\leq t\leq \tau$} is by executing an applicable sequence of programmed operator schema that achieves the goal state defined by its associated label $\sigma_t$ starting from the initial state of the corresponding label.
-If this is done for all the labels (all the input examples of the learning task) it means that the programmed action model $\Xi$ is compliant with the learning input knowledge and hence, it is a solution to the action model learning task.
+The only way of achieving a goal $\{test_t\}$,{\small $1\leq t\leq \tau$} is by executing an applicable sequence of programmed operator schema that achieves the final state defined by the associated label $\sigma_t$, starting from the initial defined in that label.
+If this is done for all the input examples of the learning task (for all the labels), it means that the programmed action model $\Xi$ is compliant with the learning input knowledge and hence, it is a solution to the action model learning task.
+\end{small}
 \end{proof}
 
-Interestingly, the compilation is valid for partially specified action models since known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda}$ and the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) be removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier.
+Interestingly, the compilation is valid for partially specified action models since known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda''}$ and the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) be removed from $A_{\Lambda''}$ making the classical planning task $P_{\Lambda''}$ easier.
 
 \section{Constraining the hypothesis space with example plans}
-The compilation can be extended to the learning scenario defined by $\Lambda$ and $\Lambda'$ in which a set of plans $\Pi$ (or only its lengths in the case of $\Lambda'$) is available.
-Each plan $\pi_i\in \Pi$, {\small $1\leq i\leq t$}, is a solution to the corresponding classical planning instance $P_i=\tup{F,A,I_i,G_i}$ defined above.
-The compilation extensions are:
+Here  we extend our compilation to address the learning scenario defined by $\Lambda$ and $\Lambda'$ where a set of plans $\Pi$ (or at least its lengths in the case of $\Lambda'$) is available.
+We denote $\pi_t\in \Pi$ as a solution plan to the classical planning instances $P_t$ introduced above.
+
+Given a learning task $\Lambda=\tup{\Psi,\Pi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}''$ as follows:
 \begin{itemize}
-\item $F_{\Lambda}$ includes the new set of fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ for encoding the $j$ steps of the {\small $1\leq i\leq t$} plans in $\Pi$ with $F_{\Pi_i}\subseteq F_{\Pi}$ the fluents encoding the plan corresponding to the $i^{th}$ example (only for the $\Lambda$ case).
+\item $F_{\Lambda}$ includes the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of the plans in $\Pi$ with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding the $t^{th}$ plan.
 In addition fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, represent the plan step where the programmed action model is validated ($n$ is the max length of a plan in $\Pi$).
-\item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi_1}$ that encode the plan $\pi_1\in \Pi$ for solving $P_1$, and the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the plan step where to start validating the programmed action model.
-Goals $G_{\Lambda}$ are like in the original compilation.
-\item With respect to the actions in $A_{\Lambda}$,
+\item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi_1}$ that encode the plan $\pi_1\in \Pi$ for solving $P_1$, and the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
+$G_{\Lambda}=G_{\Lambda}''$.
+\item With respect to $A_{\Lambda}$, the actions for programming the preconditions/effects of a given operator are the same.
 \begin{enumerate}
-\item The actions for programming the preconditions/effects of a given operator schema are the same.
-\item The actions for applying an operator schema have an extra precondition $f\in F_{\Pi_i}$ that encodes the current plan step and extra conditional effect $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing the plan step.
-\item The actions for changing the active test have an extra precondition, $at_{|\Pi_i|}$, to indicate that we simulated the full current plan $\Pi_i$ and extra conditional effects to load the next plan $\Pi_{i+1}$ where to validate the programmed action model:
+\item The actions for applying an operator have an extra precondition $f\in F_{\Pi_t}$ that encodes the current plan step and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing the plan step.
+\item The actions for changing the active test have an extra precondition, $at_{|\Pi_t|}$, to indicate that the current plan $\Pi_t$ was fully executed and extra conditional effects to load the next plan $\Pi_{t+1}$ where the programmed operators are validated:
 \begin{small}
 \begin{align*}
-&\{f\}\rhd\{\neg f\}_{f\in F_{\Pi_i}},\\
-&\{\emptyset\}\rhd\{f\}_{f\in F_{\Pi_i+1}},\\
-&\{\emptyset\}\rhd\{\neg at_{|\pi_i|},at_1\}.
+&\{f\}\rhd\{\neg f\}_{f\in F_{\Pi_t}},\\
+&\{\emptyset\}\rhd\{f\}_{f\in F_{\Pi_t+1}},\\
+&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
 \end{align*}
@@
 \section{Evaluation}
+The performance of our learning approach is evaluated for different degrees of available input knowledge and using different sources for collecting this input knowledge.
 
 \subsection{Learning action models from example plans}
-\begin{table*}[!hbt]
+Here we assess the performance of our learning approach for addressing the $\Lambda$ learning task.
+The evaluation is done using the cardinality of the {\em symmetric difference} sets that are computed between the set of preconditions, del and add effects (1), in the learned model and (2), in the actual models.
+In all the experiments the compilation is solved using the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar}.
+
+\begin{table}[]
 \begin{footnotesize}
@@
 \end{footnotesize}
-\caption{\small Mean error and standard deviation of the learned models when using hand-picked examples and examples collected using the classical planner Fast-Downward.}
+\caption{\small Mean error and standard deviation of the learned models.}
   \label{tab:eplans}
-\end{table*}
+\end{table}
 
-The performance of our learning approach is evaluated for different degrees of available input knowledge and using different sources for collecting this input knowledge.
-In all the cases we assess the performance of our learning approach using the cardinality of the {\em symmetric difference} sets that are computed between the set of preconditions, del and add effects (1), in the learned model and (2), in the actual models.
-In all the experiments the compilation is solved using the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar}.
-
-Table~\ref{tab:eplans} shows the mean error and standard deviation of the learned models with respect to the actual action models when (1) using {\em hand-picked} examples, (2) examples collected using the classical {\em planner} {\sc Fast-Downward}~\cite{helmert2006fast} and (3) examples collected {\em randomly}.
+Table~\ref{tab:eplans} shows the mean error and standard deviation of the learned models with respect to the actual action models.
 The standard deviation provides a measure of how this error is distributed among the different operators in the domain.
@@
 \subsection{Learning action models from example states}
-TBD.
+When the set of input planning task is not available the {\em symmetric difference} is not a good measure for evaluating the quality of the learned models.
+In this situation the planner determines the actions for satistying the input labels so actions can be completely interpreted in totally different way.
 
-\section{Related work}
+%\section{Related work}
 
@@
 \section{Conclusions}
-This paper presents a novel approach for learning classical planning action models from minimal input knowledge and using exclusively existing classical planners.
+The paper presented a novel approach for learning STRIPS action models from examples exclusively using classical planning.
+The evaluation show that when example plans are available the approach allows to learn action models that are very close to the models used as reference.
+Interestingly when action plans are not available the results of the learning task are action models that satisfy the input knowledge but that is not constrained to the reference model.
+An interesting research direction related to this problem is then study of domain reformulation using actions in a different way, reducing the set of actions or learning features that allow more compact solutions like the {\em reachable} or {\em movable} features in the Soybean domain.
 Learning action models from examples allows the reformulation of a domain theory.
-An interesting research direction is the study of domain reformulation using features that allow more compact solutions like the {\em reachable} or {\em movable} features in the Sokoban domain.
-The size of the compilation output depends also on the number of examples. Empirical results show that our approach is able to generate non-trivial CFGs from very small data sets.
 
-Last but not least, collecting {\em informative} examples for learning planning action models is challenging.
+With regard to the efficiency of the compilation the size of the compilation output depends also on the number of examples.
+Empirical results show that our approach is able to generate non-trivial models from very small data sets.
+
+Last but not least, collecting {\em informative} examples for learning planning action models is challenging open issue.
 Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
 In addition, motivated by the success of recent algorithms for exploring planning tasks~\cite{geffner:pwithsimulators:IJCAI17}, we do not assume that a learning set of plans is given apriori but instead, we autonomously collect the learning examples.
+
+\newpage
 

### 2017-10-06 doc/icaps-18/strips-learning.tex
addded optimization to the compiler

@@
 \item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the programmed model is being validated.
-\item Fluents $mode_{pre}$, $mode_{eff}$ and $mode_{val}$ indicating whether the preconditions of the action schema are being programmed, the effects of the action schema are being programmed or the programmed action models are being validated.
+\item Fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schema are being programmed or the programmed operators are being validated.
 \end{itemize}
 \item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (i.e.
-the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (initially all operators have all the possible preconditions) and $mode_{pre}$.
+the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (initially all operators have all the possible preconditions) and $mode_{prog}$.
 \item $G_{\Lambda''}=\{test_t\}$,{\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{pre_{f}(\xi),mode_{pre}, \\
-&\neg mode_{eff}, \neg mode_{val}\},\\    
+\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{pre_{f}(\xi),\neg del_{f}(\xi),\neg add_{f}(\xi),\\
+& mode_{prog}\},\\    
 \cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
@@
 \hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\                                                   
-&\neg mode_{val}\},\\
+& mode_{prog}\},\\ 
 \cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
-&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\},\\
-&\{mode_{pre}\}\rhd\{\neg mode_{pre},mode_{eff}\}.
+&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
 \end{align*}
@@
 &\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{pre}\}\rhd\{\neg mode_{pre},mode_{val}\},\\
-&\{mode_{eff}\}\rhd\{\neg mode_{eff},mode_{val}\}.
+&\{mode_{prog}\}\rhd\{\neg mode_{prog},mode_{val}\}.
 \end{align*}

### 2017-10-06 src/compiler.py

@@ -110,7 +110,7 @@ predicates = get_predicates_schema_from_plans(fd_task)
 # Compilation Problem
 init_aux=copy.deepcopy(fd_task.init)
 fd_task.init=[]
-fd_task.init.append(pddl.conditions.Atom("modePre",[]))
+fd_task.init.append(pddl.conditions.Atom("modeProg",[]))
 allpres=[]
 for a in actions: # All possible preconditions are initially programmed
    var_ids=[]
@@ -145,8 +145,7 @@ if input_level <= config.INPUT_LENPLAN:
    for i in range(1,MAX_STEPS+2):
       fd_task.objects.append(pddl.pddl_types.TypedObject("i"+str(i),"step"))   
    
-fd_task.predicates.append(pddl.predicates.Predicate("modePre",[]))
-fd_task.predicates.append(pddl.predicates.Predicate("modeEff",[]))
+fd_task.predicates.append(pddl.predicates.Predicate("modeProg",[]))
 fd_task.predicates.append(pddl.predicates.Predicate("modeVal",[]))
 for i in range(0,len(plans)+1):
    fd_task.predicates.append(pddl.predicates.Predicate("test"+str(i),[]))
@@ -175,6 +174,7 @@ for a in old_actions:
       params=params+[pddl.pddl_types.TypedObject("?i2","step")]
 
    pre = [pddl.conditions.Atom("modeVal",[])]
+   pre = pre + [pddl.conditions.NegatedAtom("modeProg",[])]
    if input_level <= config.INPUT_PLANS and input_level <config.INPUT_MINIMUM:
       pre = pre + [pddl.conditions.Atom("plan-"+a[0],["?i1"]+["?o"+str(i) for i in range(1,len(a))])]
       
@@ -225,20 +225,20 @@ for a in old_actions:
             vars = ["var"+str(t) for t in tup]
             params = []
             pre = []
-            pre = pre + [pddl.conditions.Atom("modePre",[])]
-            pre = pre + [pddl.conditions.NegatedAtom("modeEff",[])]
+            pre = pre + [pddl.conditions.Atom("modeProg",[])]
             pre = pre + [pddl.conditions.NegatedAtom("modeVal",[])]                        
-            pre = pre + [pddl.conditions.Atom("pre_"+p[0]+"_"+a[0],vars)]   
[…]

### 2017-10-06 doc/icaps-18/strips-learning.tex
short paper format for ICAPS18

@@
 
-\subsection{Learning from labeled plans}
+\subsubsection{Learning from labeled plans.}
 This learning task is formalized as $\Lambda=\tup{\Psi,\Pi,\Sigma}$: 
@@
 
-\subsection{Learning from initial/final states}
+\subsubsection{Learning from initial/final states.}
 Here we reduce the amount of input knowledge provided to the learning task.
@@
 $\Pi'$ that does not contain a set of plans but the number of actions of each plan, so the learning task is redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$.
-While the previous learning task, $\Lambda$, corresponds to watching an agent acting in the world, this new learning task $\Lambda'$ can be understood as watching only the results of its actions executions knowing the number of diferent actions performed by the agent.
+While the previous learning task, $\Lambda$, corresponds to watching an agent acting in the world, this new learning task $\Lambda'$ can be understood as watching only the results of its actions executions knowing the number of different actions performed by the agent.
 
@@
 
-\section{Learning STRIPS action models with classical planning}
-Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is to compile it into a classical planning task $P_{\Lambda}$, $P_{\Lambda'}$ or $P_{\Lambda''}$.
+\section{Learning STRIPS action models with planning}
+Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is compiling it into a classical planning task with conditional effects $P_{\Lambda}$, $P_{\Lambda'}$ or $P_{\Lambda''}$.
 The intuition behind these compilations is that a solution to the resulting classical planning task is a sequence of actions that:
 \begin{enumerate}
-\item Programs the STRIPS action model. For each $\xi\in\Xi$, determines the literals that belong to the sets $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
-\item Validates the programmed action model with the given set of labels $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$.
-For every {\small $1\leq t\leq \tau$}, the programmed action model $\Xi$ is used to produce a final state $s_{n}^t$ starting from their corresponding initial state $s_0^t$.
+\item Programs the STRIPS action model. For each $\xi\in\Xi$, determines the literals that belong to the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Validates the programmed operator schema $\Xi$ with the given set of labels $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$.
+$\Xi$ is used, at every {\small $1\leq t\leq \tau$}, to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
 \end{enumerate}
@@
 To formalize our compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, that belong to the same planning frame (i.e.
-share the same fluents and actions and differ only in the initial state and goals).
+same fluents and actions and differ only in the initial state and goals).
 The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of objects appearing in the example plans or their corresponding labels.
 Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(\pi_t)\cup obj(s_0^t)\cup obj(s_n^t)\}$.
-The set of actions is empty, $A=\emptyset$, i.e.
-the action model is unkown.
+The set of actions is empty, $A=\emptyset$, because the action model is unknown.
 Finally the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.
 
-Now we are ready to formalize our compilations for learning STRIPS action models using classical planning with conditional effects.
-We start with $\Lambda''$ because it is the learning task that requires the least amount of input knowledge and incrementally extend the formalized compilation until addressing $\Lambda$, the learning task with the largest amount of input knowledge.
+Now we are ready to formalize our compilations for learning STRIPS action models with classical planning.
+We start with $\Lambda''$ because it is the learning task that requires the smallest amount of input knowledge and incrementally extend the formalized compilation until addressing $\Lambda$, the learning task with the largest amount of input knowledge.
 
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect of the operator $\xi$.
-\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the programmed model is being validated.
+\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed model is being validated.
 \item Fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schema are being programmed or the programmed operators are being validated.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{pre_{f}(\xi),\neg del_{f}(\xi),\neg add_{f}(\xi),\\
-& mode_{prog}\},\\    
+\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
+& mode_{prog},pre_{f}(\xi)\},\\    
 \cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
@@
 \section{Constraining the hypothesis space with example plans}
-Here  we extend our compilation to address the learning scenario defined by $\Lambda$ and $\Lambda'$ where a set of plans $\Pi$ (or at least its lengths in the case of $\Lambda'$) is available.
-We denote $\pi_t\in \Pi$ as a solution plan to the classical planning instances $P_t$ introduced above.
-
+Here  we extend our compilation to address the learning scenario defined by $\Lambda$ where a set of plans $\Pi$ is available.
 Given a learning task $\Lambda=\tup{\Psi,\Pi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}''$ as follows:
 \begin{itemize}
-\item $F_{\Lambda}$ includes the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of the plans in $\Pi$ with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding the $t^{th}$ plan.
-In addition fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, represent the plan step where the programmed action model is validated ($n$ is the max length of a plan in $\Pi$).
+\item $F_{\Lambda}$ includes the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of the plans in $\Pi$ with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding $\pi_t\in \Pi$.
+We denote $\pi_t$ as a solution plan to the classical planning instances $P_t$ introduced above.
+In addition fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, represent the plan step where the programmed action model is validated.
 \item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi_1}$ that encode the plan $\pi_1\in \Pi$ for solving $P_1$, and the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
@@
 \begin{align*}
-&\{f\}\rhd\{\neg f\}_{f\in F_{\Pi_t}},\\
-&\{\emptyset\}\rhd\{f\}_{f\in F_{\Pi_t+1}},\\
-&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
+&\{f\}\rhd\{\neg f\}_{f\in F_{\Pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\Pi_t+1}},\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
 \end{align*}
@@
 \section{Evaluation}
-The performance of our learning approach is evaluated for different degrees of available input knowledge and using different sources for collecting this input knowledge.
+We evaluated our learning approach for different amounts of available input knowledge.
+In all the experiments the compilation is solved using the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar} given its ability to deal with planning instance highly populated with dead-ends.
 
-\subsection{Learning action models from example plans}
-Here we assess the performance of our learning approach for addressing the $\Lambda$ learning task.
-The evaluation is done using the cardinality of the {\em symmetric difference} sets that are computed between the set of preconditions, del and add effects (1), in the learned model and (2), in the actual models.
-In all the experiments the compilation is solved using the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar}.
+\subsubsection{Learning action models from example plans.}
+Here we assess the performance of our learning approach when addressing the $\Lambda$ learning task.
+The quality of the learned models is quantified using the cardinality of the {\em symmetric difference} between the set of preconditions, negative and positive effects (1), in the learned model and (2), in the actual models taken as reference.
 
-\begin{table}[]
+Last but not least, collecting {\em informative} examples for learning planning action models is challenging open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+In addition, motivated by the success of recent algorithms for exploring planning tasks~\cite{geffner:pwithsimulators:IJCAI17}, we do not assume that a learning set of plans is given apriori but instead, we autonomously collect the learning examples.
+
+\begin{table}[hbt!]
 \begin{footnotesize}
-\begin{tabular}{l||c|c|c||c|c|c||c|c|c}
+\begin{tabular}{cc}
 
@@
 
-\subsection{Learning action models from example states}
-When the set of input planning task is not available the {\em symmetric difference} is not a good measure for evaluating the quality of the learned models.
-In this situation the planner determines the actions for satistying the input labels so actions can be completely interpreted in totally different way.
+\subsubsection{Learning action models from example states.}
+When the set of input planning task is not available the {\em symmetric difference} is not an acceptable measure for evaluating the quality of the learned models.
+The reason is that, in this situation, the planner determines also the actions that must satistying the input labels so actions can reformulated and still be compliant with the learning examples.
+For instance a blocksworld can be learned where the operator {\small\tt stack} is defined with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
+
+With this regard, the Table~\ref{tab:noplans} shows the planning time and number of steps required to build a set of action models compliant with the input examples.
+
+\begin{table}[hbt!]
+\begin{footnotesize}
+\begin{tabular}{cc}
+
+\end{tabular}
+\end{footnotesize}
+\caption{\small Planning time and number of steps required to build a set of action models compliant with the input examples.}
+  \label{tab:noplans}
+\end{table}
 
@@
 The paper presented a novel approach for learning STRIPS action models from examples exclusively using classical planning.
-The evaluation show that when example plans are available the approach allows to learn action models that are very close to the models used as reference.
-Interestingly when action plans are not available the results of the learning task are action models that satisfy the input knowledge but that is not constrained to the reference model.
-An interesting research direction related to this problem is then study of domain reformulation using actions in a different way, reducing the set of actions or learning features that allow more compact solutions like the {\em reachable} or {\em movable} features in the Soybean domain.
-Learning action models from examples allows the reformulation of a domain theory.
+Interestingly the approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
 
-With regard to the efficiency of the compilation the size of the compilation output depends also on the number of examples.
-Empirical results show that our approach is able to generate non-trivial models from very small data sets.
+The empirical evaluation shows that, when example plans are available, our approach can compute accurate action models.
+When action plans are not available our approach is still able to produce action models compliant with the input information.
+In this case, since learning is not constrained by actions it can change the semantics of the operators.
+An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploitin features that allow more compact solutions like the {\em reachable} or {\em movable} features in the Sokoban domain.
 
-Last but not least, collecting {\em informative} examples for learning planning action models is challenging open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-In addition, motivated by the success of recent algorithms for exploring planning tasks~\cite{geffner:pwithsimulators:IJCAI17}, we do not assume that a learning set of plans is given apriori but instead, we autonomously collect the learning examples.
+With regard to efficiency, the size of the compiled classical planning instances depend on the number of input examples.
+On the other hand the empirical results show that our approach, since is based on inferences instead of statistics, is able to generate non-trivial models from very small data sets.
+
+\subsection*{Acknowledgment}
+\begin{small}
+Diego Garc\'ia is partially supported by the {\it FPU} program funded by the Spanish government.
+Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
+\end{small}
 

### 2017-10-10 doc/icaps-18/strips-learning.tex
starting to add first EVA's comments to the ICAPS18 paper

@@
 \usepackage{arydshln}
+\usetikzlibrary{calc,backgrounds,positioning,fit}
 
@@
 
-\author{Diego Garc\'ia\and Sergio Jim\'enez\and Eva Onaindia\\
+\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
 {\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
@@
 \begin{abstract}
-This paper presents a novel approach for learning STRIPS action models from examples that compiles the inductive learning task into a classical planning task.
-Our compilation for this learning task is flexible to different amounts of available input knowledge; it accepts partially specified action models and what is more, the input learning examples can range from a set of plans (with their corresponding initial and final states) to only a set of initial and final states where no action is observed.
+This paper presents a novel approach for learning STRIPS action models from examples that compiles this inductive learning task into a classical planning task.
+Interestingly the proposed compilation is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed.
+What is more, the compilation can reuse previous knowledge (it is able to start from partially specified action models) and can be used to validate whether an observed plan follows a given action model.
 \end{abstract}
@@
 However, the application of inductive ML techniques for learning planning action models is not straightforward.
-First, the inputs to ML algorithms usually are sets of finite numeric vectors encoding the value of sets of objects features.
-The input for learning planning action models traditionally are sets of observations of plan executions (each with possibly different length).
+First, the inputs to ML algorithms usually are finite vectors encoding the value of a given set of features in a set of objects.
+The input for learning planning action models traditionally are sets of observations of plan executions (each with a possibly different length).
 Second, the traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
-In the case of learning STRIPS action models, the output is not a scalar but a model of the preconditions and the effects of each action that defines the possible state transitions in the given planning domain.
+When learning STRIPS action models, the output is not a scalar but a model of the preconditions and the effects of each action that defines the possible state transitions of a given planning domain.
 
-Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of all the states traversed by the example plans.
-Motivated by recent advances on learning generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating} this paper introduces an innovative approach for learning classical planning action models that (1) it can be defined as a classical planning compilation and (2), it is flexible to different amounts of available input knowledge.
+Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of all the intermediate states traversed by the example plans.
+Motivated by recent advances on learning generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating} this paper introduces an innovative approach for learning classical planning action models that:
+\begin{itemize}
+\item Can be defined as a classical planning compilation.
+\item It is flexible to different amounts of available input knowledge.
+The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed.
+\item Can exploit previous knowledge about the action model in the form of partially specified action models.
+\end{itemize}
+
+The paper is organized as follows.
+The first section presents the classical planning model and its extension to define conditional effects.
+The second section formalizes the learning task addressed in the paper.
+The third section describes our approach for addressing the learning of STRIPS action models by compiling this task into a classical planning task.
+The fourth section evaluates the performance of our approach.
+Finally the last section discusses  the strengths and weaknesses of our approach and proposing some future work.
 
@@
 
+\begin{figure}[hbt]
+\begin{scriptsize}
+\begin{verbatim}
+;;; Predicates as defined in PDDL
+(on ?o1 - object ?o2 - object) (ontable ?o - object) 
+(clear ?o - object) (holding ?o  - object)
+(handempty)
+\end{verbatim}
+
+\vspace{0.1cm}
+
+\begin{subfigure}{.3\textwidth}
+\begin{verbatim}
+;;; Plan
+0: (unstack A B)
+1: (put-down A)
+2: (unstack B C)
+3: (stack B A)
+4: (unstack C D)
+5: (stack C B)
+6: (pick-up D)
+\end{verbatim}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+\begin{verbatim}
+;;; Initial and final states
+\end{verbatim}
+\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
+\node (initA) []{A};
+\node (initB) [above=of initA.north]{B};
+\node (initC) [above=of initB.north]{C};
+\node (initD) [above=of initC.north]{D};
+
+\node (goalD) [right=of initA.south east]{D};
+\node (goalC) [right=of initB.south east]{C};
+\node (goalB) [right=of initC.south east]{B};
+\node (goalA) [right=of initD.south east]{A};
+\end{tikzpicture}
+
+\end{subfigure}%
+\end{scriptsize}
+ \caption{\small Example of a learning task $\Lambda$ with a single learning example that corresponds to invert a tower of four blocks.}
+\label{fig:lexample}
+\end{figure}
+
+To illustrate this, Figure~\ref{fig:lexample}, shows $\Psi$ and $\Pi=\{\pi_1\}$ for a learning task $\Lambda$ with a single learning example that corresponds to invert a tower of four blocks.
+
 A solution to the learning task $\Lambda$ is a set of operator schema $\Xi$ (one schema for each operator header) compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
@@
 \end{enumerate}
+
+\begin{figure}[hbt]
+\begin{footnotesize}
+\begin{scriptsize}
+\begin{verbatim}
+0 : (program_pre_clear_stack_var1)
+1 : (program_pre_handempty_stack_)
+2 : (program_pre_holding_stack_var2)
+3 : (program_pre_on_stack_var1_var1)
+4 : (program_pre_on_stack_var1_var2)
+5 : (program_pre_on_stack_var2_var1)
+6 : (program_pre_on_stack_var2_var2)
+7 : (program_pre_ontable_stack_var1)
+8 : (program_pre_ontable_stack_var2)
+9 : (program_eff_clear_stack_var1)
+10 : (program_eff_clear_stack_var2)
+11 : (program_eff_handempty_stack_)
+12 : (program_eff_holding_stack_var1)
+13 : (program_eff_on_stack_var1_var2)
+14 : (validate_0)
+15 : (unstack a b i1 i2)
+16 : (put-down a i2 i3)
+17 : (unstack b c i3 i4)
+18 : (stack b a i4 i5)
+19 : (unstack c d i5 i6)
+20 : (stack c b i6 i7)
+21 : (pick-up d i7 i8)
+22 : (validate_1)
+\end{verbatim}
+\end{scriptsize}
+\end{footnotesize}
+ \caption{\small Example of a plan.}
+\label{fig:plan}
+\end{figure}
 
@@
 \begin{small}
-Diego Garc\'ia is partially supported by the {\it FPU} program funded by the Spanish government.
+Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
 Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.

### 2017-10-11 doc/icaps-18/strips-learning.tex
extendind the iCAPS18 paper to the long paper format

@@
 \documentclass[letterpaper]{article} %DO NOT CHANGE THIS
-\usepackage{aaai17}  %Required
+\usepackage{aaai18}  %Required
 \usepackage{times}  %Required
@@
 Interestingly the proposed compilation is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed.
-What is more, the compilation can reuse previous knowledge (it is able to start from partially specified action models) and can be used to validate whether an observed plan follows a given action model.
+What is more, the compilation can reuse previous knowledge (it is able to start from partially specified action models) and can also be used to validate whether an observed plan follows a given STRIPS action model.
 \end{abstract}
@@
 In both tasks, off-the-shelf planners require reasoning about action models that correctly and completely capture the possible world transitions~\cite{ghallab2004automated,geffner:book:2013}.
-Unfortunately, building such planning action models is complex, even for planning experts, so this knowledge acquisition bottleneck limits the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.
+Unfortunately, building such planning action models is complex, even for planning experts, and this knowledge acquisition bottleneck limits the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.
 
 On the other hand, Machine Learning (ML) techniques are able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-However, the application of inductive ML techniques for learning planning action models is not straightforward.
-First, the inputs to ML algorithms usually are finite vectors encoding the value of a given set of features in a set of objects.
+The application of inductive ML techniques for learning planning action models is not straightforward though:
+\begin{itemize}
+\item The inputs to ML algorithms usually are finite vectors encoding the value of fixed features for a given set of objects.
 The input for learning planning action models traditionally are sets of observations of plan executions (each with a possibly different length).
-Second, the traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
+\item The traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
 When learning STRIPS action models, the output is not a scalar but a model of the preconditions and the effects of each action that defines the possible state transitions of a given planning domain.
+\end{itemize}
 
 Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of all the intermediate states traversed by the example plans.
-Motivated by recent advances on learning generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating} this paper introduces an innovative approach for learning classical planning action models that:
-\begin{itemize}
-\item Can be defined as a classical planning compilation.
+Motivated by recent advances on learning generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating} this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation so an off-the-shelf planner can be used to address the inductive learning task.
+In addition the compilation approach presents the following benefits:
+\begin{enumerate}
 \item It is flexible to different amounts of available input knowledge.
@@
 \item Can exploit previous knowledge about the action model in the form of partially specified action models.
-\end{itemize}
+\item Can be used to validate whether an observed plan follows a given STRIPS action model.
+\end{enumerate}
 
 The paper is organized as follows.
-The first section presents the classical planning model and its extension to define conditional effects.
-The second section formalizes the learning task addressed in the paper.
-The third section describes our approach for addressing the learning of STRIPS action models by compiling this task into a classical planning task.
+The first section presents the classical planning model and its extension to conditional effects (a requirement of the proposed compilation).
+The second section formalizes the task of learning STRIPS action models from observations of plan executions with regard to different amounts of input knowledge.
+The third section describes our approach for addressing these inductive learning tasks by compiling them into classical planning.
 The fourth section evaluates the performance of our approach.
-Finally the last section discusses  the strengths and weaknesses of our approach and proposing some future work.
+Finally, the last section discusses the strengths and weaknesses of our approach and proposes some future work.
 
@@
 A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent).
+A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent) and let $\neg L=\{\neg l:l\in L\}$ be the complement of $L$.
 We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
+
 A {\em state} $s$ is then a total assignment of values to fluents, i.e.
@@
 Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of positive effects $\eff^+(a)\in\mathcal{L}(F)$, and a set of negative effects $\eff^-(a)\in\mathcal{L}(F)$.
-An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \eff^-(a))\cup\eff^+(a)$.
+An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff^-(a))\cup\eff^+(a)$.
 
@@
 
-Likewise, we assume that each action in $A$ is instantiated from an STRIPS operator schema. Figure~\ref{fig:stack} shows the {\em stack} operator schema for a STRIPS blocksworld represented in PDDL.
+Likewise, we assume that each action in $A$ is instantiated from an STRIPS operator schema.
+Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld represented in PDDL.
 \begin{figure}[hbt]
 \begin{footnotesize}
-\begin{scriptsize}
 \begin{verbatim}
@@
   :parameters (?x1 ?x2)
-  :precondition (and (holding ?x1) (clear ?x2))
-  :effect (and (not (holding ?x1)) (not (clear ?x2))
-               (clear ?x1) (handempty) (on ?x1 ?x2)))
+  :precondition (and (holding ?x1) 
+                     (clear ?x2))
+  :effect (and (not (holding ?x1)) 
+               (not (clear ?x2))
+               (clear ?x1) 
+               (handempty) 
+               (on ?x1 ?x2)))
 \end{verbatim}
-\end{scriptsize}
 \end{footnotesize}
- \caption{\small Example of a {\em stack} operator schema for a STRIPS blocksworld represented in PDDL.}
+ \caption{\small Example of a STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld represented in PDDL.}
 \label{fig:stack}
@@
 
-Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represent {\em variable names}.
-$|\Omega_v|$ is given by the action with the maximum arity in a planning frame.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ and $\Omega_v=\{v_1, v_2\}$ because the operators {\small\tt stack} and {\small\tt unstack} are the ones with the maximum arity (two parameters each).
+Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names}.
+$|\Omega_v|$ is given by the maximum arity of an action in the given planning frame.
+For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators {\small\tt stack} and {\small\tt unstack} are the ones with the maximum arity (two parameters each).
 
@@
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, represents an operator {\em header} defined by its corresponding action name and a list of variables, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
-The headers for the blocksworld operators are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, represents an operator {\em header} defined by its corresponding action name and a enumeration of the variable names, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
+The headers for a 4-operator blocksworld are then {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
 \item The preconditions $pre(\xi)\subseteq F_v$, the positive effects $add(\xi)\subseteq F_v$, and the negative effects $del(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
@@
 Our approach for leaning STRIPS action models is compiling this leaning task into a classical planning task with conditional effects.
-We use conditional effects because they allow us to compactly define actions whose effects depend on the current state.
+We use conditional effects because they allow us to compactly define actions whose particular effects depend on the current state.
 Many classical planners cope with conditional effects without compiling them away.
-In fact, the support of PDDL conditional effects was a requirement for participating at IPC-2014~\cite{vallati:IPC:AIM2015}.
+In fact, the support of PDDL conditional effects was a requirement for participating at the IPC-2014~\cite{vallati:IPC:AIM2015}.
 
@@
 i.e.~effects whose conditions hold in $s$.
-The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \eff^-(s,a))\cup\eff^+(s,a)$, where $\eff^-(s,a)$ and $\eff^+(s,a)$ are the negative and positive effects in $\eff(s,a)$.
+The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg \eff^-(s,a))\cup\eff^+(s,a)$, where $\eff^-(s,a)$ and $\eff^+(s,a)$ are the negative and positive effects in $\eff(s,a)$.
 
@@
 In this case, the operators schema are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Likewise, preconditions are derived lifting the minimal set of literals appearing in all the pre-states that correspond to the same operator.
+Likewise, preconditions are derived lifting the minimal set of literals that appears in all the pre-states that correspond to the same operator.
 
 This section formalizes more challenging tasks, for learning STRIPS action models, where less input knowledge is available.
-Next we formalize these learning tasks according to the available input knowledge.
+Next, these learning tasks are formalized according to the available amount of input knowledge.
 
-\subsubsection{Learning from labeled plans.}
+\subsection{Learning from labeled plans}
 This learning task is formalized as $\Lambda=\tup{\Psi,\Pi,\Sigma}$: 
 \begin{itemize}
-\item $\Psi$, the set of predicates that define the abstract state space of a given planning domain. This set includes the predicates for defining the headers of the operators schema.
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans.
+\item $\Psi$, the set of predicates that define the abstract state space of a given planning domain. 
+\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans s.t.
+each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces a state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
 \item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$, a set of labels s.t.
-each plan $\pi_t$, {\small $1\leq t\leq \tau$}, has a label $\sigma_t=(s_0^t,s_{n}^t)$ where $s_{n}^t$ is the state resulting from executing $\pi_t$ starting from the state $s_0^t$.
+each plan $\pi_t$, {\small $1\leq t\leq \tau$}, has an associated label $\sigma_t=(s_0^t,s_{n}^t)$ where $s_{n}^t$ is the state resulting from executing the plan $\pi_t$ starting from the state $s_0^t$.
 \end{itemize}
 
+To illustrate this, Figure~\ref{fig:lexample}, shows the contain of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$ for a learning task $\Lambda$ with a single learning example from the blocksworld for inverting a tower of four blocks.
+
 \begin{figure}[hbt]
-\begin{scriptsize}
+\begin{small}
+{\tt ;;; Predicates in $\Psi$}
 \begin{verbatim}
-;;; Predicates as defined in PDDL
-(on ?o1 - object ?o2 - object) (ontable ?o - object) 
-(clear ?o - object) (holding ?o  - object)
-(handempty)
+(handempty) (holding ?o  - object)
+(clear ?o - object) (ontable ?o - object) 
+(on ?o1 - object ?o2 - object)  
 \end{verbatim}
+\end{small}
 
-\vspace{0.1cm}
+\vspace{0.3cm}
 
 \begin{subfigure}{.3\textwidth}
+\begin{small}
+{\tt ;;; Plan $\pi_1$}
 \begin{verbatim}
-;;; Plan
 0: (unstack A B)
-1: (put-down A)
+1: (putdown A)
 2: (unstack B C)
@@
 5: (stack C B)
-6: (pick-up D)
+6: (pickup D)
+7: (stack D C)
 \end{verbatim}
+\end{small}
 \end{subfigure}%
-\begin{subfigure}{.6\textwidth}
-\begin{verbatim}
-;;; Initial and final states
-\end{verbatim}
+\begin{subfigure}{.5\textwidth}
+{\small\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
+\begin{lstlisting}[mathescape]
+\end{lstlisting}
 \begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
-\node (initA) []{A};
-\node (initB) [above=of initA.north]{B};
-\node (initC) [above=of initB.north]{C};
-\node (initD) [above=of initC.north]{D};
+\node (initD) [draw]{D};
+\node (initC) [draw, above=of initD.north]{C};
+\node (initB) [draw, above=of initC.north]{B};
+\node (initA) [draw, above=of initB.north]{A};
+\draw (-1,-0.25) -- (2.5,-0.25);
 
-\node (goalD) [right=of initA.south east]{D};
-\node (goalC) [right=of initB.south east]{C};
-\node (goalB) [right=of initC.south east]{B};
-\node (goalA) [right=of initD.south east]{A};
+\node (goalA) [draw, right=10mm of initD]{A};
+\node (goalB) [draw, right=10mm of initC]{B};
+\node (goalC) [draw, right=10mm of initB]{C};
+\node (goalD) [draw, right=10mm of initA]{D};
 \end{tikzpicture}
+\vspace{0.6cm}
+\end{subfigure}%
 
-\end{subfigure}%
-\end{scriptsize}
- \caption{\small Example of a learning task $\Lambda$ with a single learning example that corresponds to invert a tower of four blocks.}
+ \caption{\small Example of a learning task with a single learning example that corresponds to invert a tower of four blocks.}
 \label{fig:lexample}
@@
 
-To illustrate this, Figure~\ref{fig:lexample}, shows $\Psi$ and $\Pi=\{\pi_1\}$ for a learning task $\Lambda$ with a single learning example that corresponds to invert a tower of four blocks.
+A solution to the learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with different name in the example plans) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
+With this regard, the operator schema shown in Figure~\label{fig:stack} is compliant with the learning task illustrated by the Figure~\ref{fig:lexample}.
 
-A solution to the learning task $\Lambda$ is a set of operator schema $\Xi$ (one schema for each operator header) compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
-
-\subsubsection{Learning from initial/final states.}
+\subsection{Learning from initial/final states}
 Here we reduce the amount of input knowledge provided to the learning task.
 Now $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is replaced by $\Pi'=\{|\pi_1|,\ldots,|\pi_{\tau}|\}$ i.e.
-$\Pi'$ that does not contain a set of plans but the number of actions of each plan, so the learning task is redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$.
-While the previous learning task, $\Lambda$, corresponds to watching an agent acting in the world, this new learning task $\Lambda'$ can be understood as watching only the results of its actions executions knowing the number of different actions performed by the agent.
+$\Pi'$ that does not contain a set of plans but the number of actions of each plan and the headers of the operators schema.
+The learning task is hence redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$.
+While the previous learning task, $\Lambda$, corresponds to watching an agent acting in the world, this new learning task $\Lambda'$ can be understood as watching only the result of its plan execution but knowing the number of actions performed by the agent.
+In the case of the learning example shown in Figure~\ref{fig:lexample}, $|\pi_1|=8$.
 
 Finally, we can go a step further and redefine a third learning task $\Lambda''=\tup{\Psi,\Sigma}$ that corresponds to watching only the results of the plan executions.
-In this case a solution to the $\Lambda''$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
+In this case no information about the executed plans is given.
+A solution to the $\Lambda''$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
 
-In these two cases, a solution must not only synthesize the action models but also the actions that produced the given labels (this information is no longer given in the learning examples).
+In these two cases $\Lambda'$ and $\Lambda''$, a solution must not only synthesize a possible action models but also the actions could have produced the given labels (this information about the actions is no longer given in the learning examples).
 
@@
 $\Xi$ is used, at every {\small $1\leq t\leq \tau$}, to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
+If information about the plans is given, then it is used to constrain the validation of the programmed opertator schema $\Xi$.
 \end{enumerate}
+
+Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ class.
+In particular the classical plan programs and validates the operator schema {\tt\small stack} using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+The details of the compilation are given in the remaining of the section.
 
@@
 \begin{footnotesize}
-\begin{scriptsize}
 \begin{verbatim}
@@
     13 : (program_eff_on_stack_var1_var2)
-14 : (validate_0)
-15 : (unstack a b i1 i2)
-16 : (put-down a i2 i3)
-17 : (unstack b c i3 i4)
-18 : (stack b a i4 i5)
-19 : (unstack c d i5 i6)
-20 : (stack c b i6 i7)
-21 : (pick-up d i7 i8)
+    14 : (unstack a b i1 i2)
+    15 : (putdown a i2 i3)
+    16 : (unstack b c i3 i4)
+    17 : (stack b a i4 i5)
+    18 : (unstack c d i5 i6)
+    19 : (stack c b i6 i7)
+    20 : (pickup d i7 i8)
+    21 : (stack d c i8 i9)
     22 : (validate_1)
 \end{verbatim}
-\end{scriptsize}
 \end{footnotesize}
- \caption{\small Example of a plan.}
+ \caption{\small Example of a plan for programing and validating the operator schema unstack using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.}
 \label{fig:plan}
@@
 
-To formalize our compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, that belong to the same planning frame (i.e.
+To formalize our compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, one for each leaning example, that belong to the same planning frame (i.e.
 same fluents and actions and differ only in the initial state and goals).
-The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of objects appearing in the example plans or their corresponding labels.
-Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(\pi_t)\cup obj(s_0^t)\cup obj(s_n^t)\}$.
-The set of actions is empty, $A=\emptyset$, because the action model is unknown.
-Finally the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.
+The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of objects appearing in the examples.
+Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
+The set of actions, $A=\emptyset$, is empty because the action model is unknown.
+Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.
 
@@
 \begin{itemize}
-\item Fluents representing the programmed action model ($pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ for every $f\in F_v$ and $\xi \in \Xi$).
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect of the operator $\xi$.
-\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed model is being validated.
+\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ for every $f\in F_v$ and $\xi \in \Xi$).
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the STRIPS operator schema $\xi\in \Xi$.
+\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
 \item Fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schema are being programmed or the programmed operators are being validated.
+They are state invariants meaning that only a mode can be true at any state.
 \end{itemize}
 \item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (i.e.
-the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (initially all operators have all the possible preconditions) and $mode_{prog}$.
-\item $G_{\Lambda''}=\{test_t\}$,{\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
+the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially all operators schema have all the possible preconditions, no positive effect and no negative effect) and $mode_{prog}$.
+\item $G_{\Lambda''}=\{test_t\}$, {\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
 \item $A_{\Lambda''}$ contains actions of three different types:
 \begin{enumerate}
-\item The actions for programming the operator schema. This includes the actions for removing a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
+\item The actions for programming the operator schema. This set includes:
+\begin{itemize}
+\item The actions for removing a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
 
@@
 
-The actions for adding a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
+\item The actions for adding a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
 
@@
 \end{small}
+\end{itemize}
 
@@
 For instance, these actions define that if an operator is programmed with the precondition $holding(v_1)\in F_v$ it {\em implies} ($\implies$) that $holding(block_1)\in F$ has to be true in the current state if the operator binds the variable object $v_1\in\Omega_v$ with the regular object $block_1\in\Omega$.
-The operator binding is done implicitly, i.e.
+The operator binding is done implicitly by order of appearance, i.e.
 variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing in the same position.
@@
 \begin{proof}[Proof sketch]
-\begin{small}
 Once the preconditions of an operator schema $\Xi$ are programmed, they cannot be altered.
@@
 If this is done for all the input examples of the learning task (for all the labels), it means that the programmed action model $\Xi$ is compliant with the learning input knowledge and hence, it is a solution to the action model learning task.
-\end{small}
 \end{proof}
 
-Interestingly, the compilation is valid for partially specified action models since known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda''}$ and the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) be removed from $A_{\Lambda''}$ making the classical planning task $P_{\Lambda''}$ easier.
+Interestingly, the compilation accepts partially specified action models since known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda''}$ and the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) be removed from $A_{\Lambda''}$ making the classical planning task $P_{\Lambda''}$ easier.
+In the extreme the compilation can also be used to validate whether an observed plan follows a given STRIPS action model.
+In this case the model to validate is coded in the initial state $I_{\Lambda''}$, any programmed action is removed from $P_{\Lambda''}$ and the only possible mode is the validation mode.
+If a solution plan is found to $P_{\Lambda''}$ it means that the given STRIPS action model is {\em valid} for the given examples.
+If $P_{\Lambda''}$ is unsolvable it means that the given STRIPS action model is invalid since it cannot satisfy the given examples.
+Tools for plan validation like VAL~\cite{howey2004val} can also be used at this point.
 
@@
 
-\subsubsection{Learning action models from example plans.}
+\subsection{Validating action models with classical planning}
+
+\subsection{Learning action models from example plans}
 Here we assess the performance of our learning approach when addressing the $\Lambda$ learning task.
@@
 
-\subsubsection{Learning action models from example states.}
+\subsection{Learning action models from example states}
 When the set of input planning task is not available the {\em symmetric difference} is not an acceptable measure for evaluating the quality of the learned models.
@@
 
-%\section{Related work}
+\section{Related work}
 
-%The  LIVE  system  (Shen  and  Simon,  1989)  was  an  extension  of  the  General  Problem  Solver  (GPS) framework (Ernst and Newell, 1969) with a learning component.
+The  LIVE  system  (Shen  and  Simon,  1989)  was  an  extension  of  the  General  Problem  Solver  (GPS) framework (Ernst and Newell, 1969) with a learning component.
 LIVE alternated problem solving with model  learning  to  automatically  define  operators.
@@
 
-%Others systems have tried to learn more expressive action models for deterministic planning in fully observable environments.
+Others systems have tried to learn more expressive action models for deterministic planning in fully observable environments.
 Examples would include the learning of conditional costs for AP actions (Jess Lanchas and Borrajo, 2007) or the learning of conditional effects with quantifiers (Zhuo et al., 2008).
 
-%In addition action model learning has been studied in domains where there is partial state observability.
+In addition action model learning has been studied in domains where there is partial state observability.
 ARMS uses the same kind od learning examples but assumes the exampls are given and proceeds in two phases.
@@
 
-%(Amir and Chang, 2008) introduced an algorithm that tractably generates all the STRIPS-like models that  could  have  lead  to  a  set  of  observations.
+(Amir and Chang, 2008) introduced an algorithm that tractably generates all the STRIPS-like models that  could  have  lead  to  a  set  of  observations.
 Given  a  formula  representing  the  initial  belief  state,  a sequence of executed actions and the corresponding observed states(where partial observations of states are given), it builds a complete explanation of observations by models of actions through a Conjunctive Normal Form (CNF) formula.
@@
 LOCM receives descriptions of plans or plan fragments, uses them to create states machines for the different domain objects and extracts the action schemas from these state machines.
+
+~\cite{stern2017efficient}.
 

### 2017-10-11 src/evaluator2.py
Added evaluator2.py: an evaluation script to compute precision and recall of the learned domains

@@ -1,0 +1,86 @@
+#! /usr/bin/env python
+# from __future__ import division
+import  sys
+import pddl, pddl_parser
+import numpy as np
+
+
+# **************************************#
+# MAIN
+# **************************************#
+try:
+   reference_domain_filename  = sys.argv[1]
+   evaluation_domain_filename  = sys.argv[2]
+   aux_problem_filename  = sys.argv[3]
+except:
+   print "Usage:"
+   print sys.argv[0] + " <reference domain> <evaluation domain>  <aux problem>"
+   sys.exit(-1)
+
+
+# reference_domain_filename = "../benchmarks/reference/blocks/domain.pddl"
+# evaluation_domain_filename = "learned_domain.pddl"
+# aux_problem_filename = "../benchmarks/handpicked/blocks/test-1.pddl"
+
+# Creating a FD task with the ref domain and the aux problem file
+fd_ref_domain = pddl_parser.pddl_file.parse_pddl_file("domain", reference_domain_filename)
+fd_problem = pddl_parser.pddl_file.parse_pddl_file("task", aux_problem_filename)
+fd_ref_task = pddl_parser.pddl_file.parsing_functions.parse_task(fd_ref_domain, fd_problem)
+
+# Creating a FD task with the domain to evaluate and the aux problem file
+fd_eva_domain = pddl_parser.pddl_file.parse_pddl_file("domain", evaluation_domain_filename)
+fd_eva_task = pddl_parser.pddl_file.parsing_functions.parse_task(fd_eva_domain, fd_problem)
+
+
+ref_pres = set()
+eva_pres = set()
+ref_adds = set()
+eva_adds = set()
+ref_dels = set()
+eva_dels = set()
+
+
+# Build the pre/add/del sets
+# Each element of the set is a tuple (action name, literal)
+for action in fd_ref_task.actions:
+    # Preconditions
+    if isinstance(action.precondition, pddl.conditions.Atom):
+        ref_pres.add((action.name,action.precondition))
+    else:
+        ref_pres.update([(action.name, x) for x in action.precondition.parts])
+    # Effects
+    for effect in action.effects:
+        if effect.literal.negated:
+            ref_dels.add((action.name, effect.literal))
+        else:
+            ref_adds.add((action.name, effect.literal))
+
+
+for action in fd_eva_task.actions:
+    # Preconditions
+    if isinstance(action.precondition, pddl.conditions.Atom):
[…]

### 2017-10-12 src/compiler.py
small optimization added to compiler.py

@@ -146,7 +146,6 @@ if input_level <= config.INPUT_LENPLAN:
       fd_task.objects.append(pddl.pddl_types.TypedObject("i"+str(i),"step"))   
    
 fd_task.predicates.append(pddl.predicates.Predicate("modeProg",[]))
-fd_task.predicates.append(pddl.predicates.Predicate("modeVal",[]))
 for i in range(0,len(plans)+1):
    fd_task.predicates.append(pddl.predicates.Predicate("test"+str(i),[]))
 if input_level <= config.INPUT_LENPLAN:                  
@@ -172,8 +171,7 @@ for a in old_actions:
    if input_level <=config.INPUT_LENPLAN and input_level <config.INPUT_MINIMUM:
       params=params+[pddl.pddl_types.TypedObject("?i1","step")]
       params=params+[pddl.pddl_types.TypedObject("?i2","step")]
-
-   pre = [pddl.conditions.Atom("modeVal",[])]
+   pre = [] 
    pre = pre + [pddl.conditions.NegatedAtom("modeProg",[])]
    if input_level <= config.INPUT_PLANS and input_level <config.INPUT_MINIMUM:
       pre = pre + [pddl.conditions.Atom("plan-"+a[0],["?i1"]+["?o"+str(i) for i in range(1,len(a))])]
@@ -226,7 +224,6 @@ for a in old_actions:
             params = []
             pre = []
             pre = pre + [pddl.conditions.Atom("modeProg",[])]
-            pre = pre + [pddl.conditions.NegatedAtom("modeVal",[])]                        
             pre = pre + [pddl.conditions.Atom("pre_"+p[0]+"_"+a[0],vars)]
             pre = pre + [pddl.conditions.NegatedAtom("del_"+p[0]+"_"+a[0],vars)]            
             pre = pre + [pddl.conditions.NegatedAtom("add_"+p[0]+"_"+a[0],vars)]            
@@ -235,7 +232,6 @@ for a in old_actions:
 
             pre = []
             pre = pre + [pddl.conditions.Atom("modeProg",[])]
-            pre = pre + [pddl.conditions.NegatedAtom("modeVal",[])]                        
             pre = pre + [pddl.conditions.NegatedAtom("del_"+p[0]+"_"+a[0],vars)]            
             pre = pre + [pddl.conditions.NegatedAtom("add_"+p[0]+"_"+a[0],vars)]
[…]

### 2017-10-12 doc/icaps-18/strips-learning.tex
added more content to the ICAPS18 paper

@@
 This paper presents a novel approach for learning STRIPS action models from examples that compiles this inductive learning task into a classical planning task.
-Interestingly the proposed compilation is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed.
-What is more, the compilation can reuse previous knowledge (it is able to start from partially specified action models) and can also be used to validate whether an observed plan follows a given STRIPS action model.
+Interestingly the proposed compilation is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where neither actions nor intermediate states are observed.
+What is more, the compilation can reuse previous knowledge (it is able to start from partially specified action models) and can also be used to validate whether observation of plan executions follow a given STRIPS action model.
 \end{abstract}
@@
 \section{Introduction}
-Besides plan synthesis, planning action models are also useful for plan/goal recognition~\cite{ramirez2010probabilistic}.
-In both tasks, off-the-shelf planners require reasoning about action models that correctly and completely capture the possible world transitions~\cite{ghallab2004automated,geffner:book:2013}.
+Besides plan synthesis~\cite{ghallab2004automated,geffner:book:2013}, planning action models are also useful for plan/goal recognition~\cite{ramirez2012plan}.
+In both tasks, off-the-shelf planners require reasoning about action models that correctly and completely capture the possible world transitions.
 Unfortunately, building such planning action models is complex, even for planning experts, and this knowledge acquisition bottleneck limits the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.
 
-On the other hand, Machine Learning (ML) techniques are able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML techniques for learning planning action models is not straightforward though:
+On the other hand, Machine Learning (ML) techniques have shown able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
+The application of inductive ML techniques for learning planning STRIPS action models from examples is not straightforward though:
 \begin{itemize}
-\item The inputs to ML algorithms usually are finite vectors encoding the value of fixed features for a given set of objects.
-The input for learning planning action models traditionally are sets of observations of plan executions (each with a possibly different length).
+\item The inputs to ML algorithms usually are finite vectors that encode the value of fixed features for a given set of objects.
+The input for learning planning action models traditionally are sets of observations of plan executions (each with possibly a different length).
 \item The traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
-When learning STRIPS action models, the output is not a scalar but a model of the preconditions and the effects of each action that defines the possible state transitions of a given planning domain.
+When learning STRIPS action models, the output is not a scalar but the sets of preconditions and effects of each action that define the possible state transitions of a given planning domain.
 \end{itemize}
 
-Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of all the intermediate states traversed by the example plans.
-Motivated by recent advances on learning generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating} this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation so an off-the-shelf planner can be used to address the inductive learning task.
-In addition the compilation approach presents the following benefits:
+Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
+Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation (so an off-the-shelf planner can be used to address the inductive learning task).
+In addition, the compilation approach presents the following benefits:
 \begin{enumerate}
@@
 \item Can exploit previous knowledge about the action model in the form of partially specified action models.
-\item Can be used to validate whether an observed plan follows a given STRIPS action model.
+\item Can be used to validate whether observed plan executions are valid for a given STRIPS action model.
 \end{enumerate}
@@
 The paper is organized as follows.
-The first section presents the classical planning model and its extension to conditional effects (a requirement of the proposed compilation).
+The first section presents the classical planning model, its extension to conditional effects (a requirement for the proposed compilation) and the STRIPS action model.
 The second section formalizes the task of learning STRIPS action models from observations of plan executions with regard to different amounts of input knowledge.
 The third section describes our approach for addressing these inductive learning tasks by compiling them into classical planning.
-The fourth section evaluates the performance of our approach.
+The fourth section reports the collected data during the evaluation of the performance of our approach.
 Finally, the last section discusses the strengths and weaknesses of our approach and proposes some future work.
@@
 We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
-A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.~$l=f$ or $l=\neg f$.
+A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
+either~$l=f$ or $l=\neg f$.
 A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent) and let $\neg L=\{\neg l:l\in L\}$ be the complement of $L$.
@@
 A {\em state} $s$ is then a total assignment of values to fluents, i.e.
-$|s|=|F|$, so the size of the state space $2^{|F|}$.
-Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions, but we often abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
+$|s|=|F|$, so the size of the state space is $2^{|F|}$.
+Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
 
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of positive effects $\eff^+(a)\in\mathcal{L}(F)$, and a set of negative effects $\eff^-(a)\in\mathcal{L}(F)$.
+Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of negative effects $\eff^-(a)\in\mathcal{L}(F)$, and a set of positive effects $\eff^+(a)\in\mathcal{L}(F)$.
 An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff^-(a))\cup\eff^+(a)$.
@@
 A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The plan $\pi$ {\em solves} $P$ if and only if $G\subseteq s_n$, i.e.~if the goal condition is satisfied following the application of $\pi$ in $I$.
+The plan $\pi$ {\em solves} $P$ if and only if $G\subseteq s_n$, i.e.~if the goal condition is satisfied in the state reached after following the application of $\pi$ in $I$.
 
-In this work we assume that the fluents in $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
+\subsection{Classical planning with conditional effects}
+Our approach for leaning STRIPS action models is compiling this leaning task into a classical planning task with conditional effects.
+We use conditional effects because they allow us to compactly define actions whose particular effects depend on the current state.
+Many classical planners cope with conditional effects without compiling them away.
+In fact, the support of conditional effects was a requirement for participating at the IPC-2014~\cite{vallati:IPC:AIM2015}.
+
+Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$ called the {\em precondition} and a set of conditional effects $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$ (the condition) and $E\in\mathcal{L}(F)$ (the effect).
+
+An action $a\in A$ is applicable in state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} is
+\[
+\eff(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
+\]
+i.e.~effects whose conditions hold in $s$.
+The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg \eff^-(s,a))\cup\eff^+(s,a)$, where $\eff^-(s,a)$ and $\eff^+(s,a)$ are the negative and positive effects in $\eff(s,a)$.
+
+\subsection{The STRIPS action model}
+In this work we address the task of learning  action  schema  in  the  STRIPS language~\cite{fikes1971strips}.
+With this regard, we assume that the fluents in $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
 There exists a set of predicates $\Psi$, each $p\in\Psi$ with an argument list of arity $ar(p)$.
@@
 
-Likewise, we assume that each action in $A$ is instantiated from an STRIPS operator schema.
-Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld represented in PDDL.
+Likewise, we assume that each action in $A$ is instantiated from a STRIPS operator schema.
+Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld~\cite{slaney2001blocks} represented in PDDL.
 \begin{figure}[hbt]
@@
 (:action stack
-  :parameters (?x1 ?x2)
+  :parameters (?x1 ?x2 - object)
   :precondition (and (holding ?x1) 
@@
 Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names}.
-$|\Omega_v|$ is given by the maximum arity of an action in the given planning frame.
+The cardinality of this set, $|\Omega_v|$, is given by the maximum arity of an action in the given planning frame.
 For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators {\small\tt stack} and {\small\tt unstack} are the ones with the maximum arity (two parameters each).
 
-Let us define a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} $\Omega_v$.
-In the blocksworld $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.
+Let us define a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} in $\Omega_v$.
+In the blocksworld this set is defined as $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.
 
-We are now ready to define a STRIPS operator schema as a tuple $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$:
+We are now ready to define a STRIPS action model $\Xi$ as a set of operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$:
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, represents an operator {\em header} defined by its corresponding action name and a enumeration of the variable names, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, represents the operator {\em header} defined by its corresponding action name and a enumeration of the variable names given by the arity of the operator, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
 The headers for a 4-operator blocksworld are then {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
-\item The preconditions $pre(\xi)\subseteq F_v$, the positive effects $add(\xi)\subseteq F_v$, and the negative effects $del(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
 \end{itemize}
-
-\subsection{Classical planning with conditional effects}
-Our approach for leaning STRIPS action models is compiling this leaning task into a classical planning task with conditional effects.
-We use conditional effects because they allow us to compactly define actions whose particular effects depend on the current state.
-Many classical planners cope with conditional effects without compiling them away.
-In fact, the support of PDDL conditional effects was a requirement for participating at the IPC-2014~\cite{vallati:IPC:AIM2015}.
-
-Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$ called the {\em precondition} and a set of conditional effects $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$ (the condition) and $E\in\mathcal{L}(F)$ (the effect).
-
-An action $a\in A$ is applicable in state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} is
-\[
-\eff(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
-\]
-i.e.~effects whose conditions hold in $s$.
-The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg \eff^-(s,a))\cup\eff^+(s,a)$, where $\eff^-(s,a)$ and $\eff^+(s,a)$ are the negative and positive effects in $\eff(s,a)$.
 
@@
 \item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans s.t.
-each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces a state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-\item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$, a set of labels s.t.
+each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces a state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+\item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$, is a set of labels s.t.
 each plan $\pi_t$, {\small $1\leq t\leq \tau$}, has an associated label $\sigma_t=(s_0^t,s_{n}^t)$ where $s_{n}^t$ is the state resulting from executing the plan $\pi_t$ starting from the state $s_0^t$.
@@
 
-To illustrate this, Figure~\ref{fig:lexample}, shows the contain of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$ for a learning task $\Lambda$ with a single learning example from the blocksworld for inverting a tower of four blocks.
+To illustrate this, Figure~\ref{fig:lexample}, shows the contain of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$ for a $\Lambda$ task that aims learning the STRIPS action model of the blocksworld.
+The shown learning task has a single learning example that corresponds to observing a plan execution for inverting a tower of four blocks.
 
@@
 \end{subfigure}%
-\begin{subfigure}{.5\textwidth}
+\begin{subfigure}{.3\textwidth}
 {\small\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
@@
 \node (initA) [draw, above=of initB.north]{A};
-\draw (-1,-0.25) -- (2.5,-0.25);
+\draw[thick] (-1,-0.25) -- (2.5,-0.25);
 
@@
 
- \caption{\small Example of a learning task with a single learning example that corresponds to invert a tower of four blocks.}
+ \caption{\small Learning task in the blocksworld with a single learning example that corresponds to inverting a tower of four blocks.}
 \label{fig:lexample}
@@
 
-A solution to the learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with different name in the example plans) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
-With this regard, the operator schema shown in Figure~\label{fig:stack} is compliant with the learning task illustrated by the Figure~\ref{fig:lexample}.
+A solution to the learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with different name in the example plans) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
+With this regard, the operator schema shown in Figure~\ref{fig:stack} is compliant with the learning task illustrated by the Figure~\ref{fig:lexample}.
 
@@
 Now $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is replaced by $\Pi'=\{|\pi_1|,\ldots,|\pi_{\tau}|\}$ i.e.
-$\Pi'$ that does not contain a set of plans but the number of actions of each plan and the headers of the operators schema.
+$\Pi'$ that does not contain a set of plans but the number of actions of each plan.
 The learning task is hence redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$.
@@
 In these two cases $\Lambda'$ and $\Lambda''$, a solution must not only synthesize a possible action models but also the actions could have produced the given labels (this information about the actions is no longer given in the learning examples).
+However, we assume that at the headers of the operators schema are known.
 
@@
 \begin{enumerate}
-\item Programs the STRIPS action model. For each $\xi\in\Xi$, determines the literals that belong to the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed operator schema $\Xi$ with the given set of labels $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$.
-$\Xi$ is used, at every {\small $1\leq t\leq \tau$}, to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
-If information about the plans is given, then it is used to constrain the validation of the programmed opertator schema $\Xi$.
+\item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan determines the literals that belong to the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Validates the programmed the STRIPS action model $\Xi$ using the available input knowledge ($\Pi$, if available, and the given set of labels $\Sigma$).
+At every learning example {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
+If information about the corresponding plan $\pi_t$ is available, then it is used to constrain the validation of the programmed STRIPS action model $\Xi$.
 \end{enumerate}
@@
 Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ class.
-In particular the classical plan programs and validates the operator schema {\tt\small stack} using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+The shown classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+In particular the steps $[0,8]$ are actions for programming the preconditions of the operator, steps $[9,13]$ are actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operator.
 The details of the compilation are given in the remaining of the section.
@@
     13 : (program_eff_on_stack_var1_var2)
-    14 : (unstack a b i1 i2)
-    15 : (putdown a i2 i3)
-    16 : (unstack b c i3 i4)
-    17 : (stack b a i4 i5)
-    18 : (unstack c d i5 i6)
-    19 : (stack c b i6 i7)
-    20 : (pickup d i7 i8)
-    21 : (stack d c i8 i9)
+    14 : (apply_unstack a b i1 i2)
+    15 : (apply_putdown a i2 i3)
+    16 : (apply_unstack b c i3 i4)
+    17 : (apply_stack b a i4 i5)
+    18 : (apply_unstack c d i5 i6)
+    19 : (apply_stack c b i6 i7)
+    20 : (apply_pickup d i7 i8)
+    21 : (apply_stack d c i8 i9)
     22 : (validate_1)
@@
 same fluents and actions and differ only in the initial state and goals).
-The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of objects appearing in the examples.
+The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of different objects appearing in the examples.
 Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
@@
 Now we are ready to formalize our compilations for learning STRIPS action models with classical planning.
-We start with $\Lambda''$ because it is the learning task that requires the smallest amount of input knowledge and incrementally extend the formalized compilation until addressing $\Lambda$, the learning task with the largest amount of input knowledge.
+We start with $\Lambda''$ because it is the learning task that requires the smallest amount of input knowledge, and incrementally extend the formalized compilation until addressing $\Lambda$, the learning task with the largest amount of input knowledge.
 
@@
 \begin{itemize}
-\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ for every $f\in F_v$ and $\xi \in \Xi$).
+\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$).
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the STRIPS operator schema $\xi\in \Xi$.
+\item A fluent $mode_{prog}$ indicating whether the operator schema are being programmed or they are already programmed operators so are being validated in the learning examples. 
 \item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
-\item Fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schema are being programmed or the programmed operators are being validated.
-They are state invariants meaning that only a mode can be true at any state.
 \end{itemize}
 \item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (i.e.
-the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially all operators schema have all the possible preconditions, no positive effect and no negative effect) and $mode_{prog}$.
+the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially, any operator schema, contains all the possible preconditions, no negative effect and no positive effect) and $mode_{prog}$.
 \item $G_{\Lambda''}=\{test_t\}$, {\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
-\item $A_{\Lambda''}$ contains actions of three different types:
+\item $A_{\Lambda''}$ contains actions of three different kinds:
 \begin{enumerate}
@@
 
-\item The actions for applying an already programmed operator schema $\xi\in\Xi$ bounding it with objects $\omega\subseteq\Omega^{ar(\xi)}$
+\item The actions for applying an already programmed operator schema $\xi\in\Xi$ bounded with the objects $\omega\subseteq\Omega^{ar(\xi)}$
 \begin{small}
@@
 &\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{prog}\}\rhd\{\neg mode_{prog},mode_{val}\}.
+&\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
 \end{align*}
@@
 The operator binding is done implicitly by order of appearance, i.e.
-variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing in the same position.
+the variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing in the same position.
 
-\item The actions for changing the learning example {\small $1\leq t\leq \tau$} where the programmed action model is validated. 
+\item The actions for updating the learning example {\tt\small $1\leq t\leq \tau$} where the programmed action model is validated. 
 \begin{small}
@@
 \hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\cup\\
-&\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{mode_{val}\},\\
+&\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
 \cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
@@
 \begin{lemma}
-Any classical plan $\pi$ that solves $P_{\Lambda''}$ induces a valid action model that solves the learning task $\Lambda''$.
+Any classical plan $\pi$ that solves $P_{\Lambda''}$ induces an action model $\Xi$ that solves the learning task $\Lambda''$.
 \end{lemma}
@@
 Once the preconditions of an operator schema $\Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema (besides they can only be programmed after the operator preconditions are programmed).
-Furthermore, once an operator schema is programmed it can only be applied.
-The only way of achieving a goal $\{test_t\}$,{\small $1\leq t\leq \tau$} is by executing an applicable sequence of programmed operator schema that achieves the final state defined by the associated label $\sigma_t$, starting from the initial defined in that label.
-If this is done for all the input examples of the learning task (for all the labels), it means that the programmed action model $\Xi$ is compliant with the learning input knowledge and hence, it is a solution to the action model learning task.
+The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after the operator preconditions are programmed, because of the preconditions of the $programPre_{f,\xi}$ actions).
+Furthermore, once the operator schema are programmed they can only be applied because of the value of the $mode_{prog}$ fluent.
+The only way of achieving a goal $\{test_t\}$,{\small $1\leq t\leq \tau$} is by executing an applicable sequence of programmed operator schema that achieves the final state defined by the associated label $\sigma_t$, starting from the initial defined in this same label.
+If this is done for all the input examples of the learning task (for all the labels $\sigma_t$, {\small $1\leq t\leq \tau$}), it means that the programmed action model $\Xi$ is compliant with all the learning input knowledge and hence, it is a solution to the action model learning task.
 \end{proof}
 
-Interestingly, the compilation accepts partially specified action models since known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda''}$ and the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) be removed from $A_{\Lambda''}$ making the classical planning task $P_{\Lambda''}$ easier.
-In the extreme the compilation can also be used to validate whether an observed plan follows a given STRIPS action model.
-In this case the model to validate is coded in the initial state $I_{\Lambda''}$, any programmed action is removed from $P_{\Lambda''}$ and the only possible mode is the validation mode.
+\subsection{Constraining the hypothesis space with previous action models}
+Interestingly, the compilation does not require to start the learning of STRIPS action models from scratch and accepts partially specified operator schema where some preconditions are effects are a priori known.
+In more detail, the known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda''}$ and the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) be removed from $A_{\Lambda''}$ reducing the number of applicable actions  and making the classical planning task $P_{\Lambda''}$ easier.
+For instance the classical plan in Figure~\ref{fig:plan} is a solution to a learning task for getting the blocksworld action model where the operator schema for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already known.
+
+In the extreme, the compilation can also be used to validate whether an observed plan follows a given STRIPS action model.
+In this case the action model to validate is coded in the initial state $I_{\Lambda''}$, any programmed action is removed from $A_{\Lambda''}$ and the $mode_{prog}$ fluent is set to false.
 If a solution plan is found to $P_{\Lambda''}$ it means that the given STRIPS action model is {\em valid} for the given examples.
 If $P_{\Lambda''}$ is unsolvable it means that the given STRIPS action model is invalid since it cannot satisfy the given examples.
-Tools for plan validation like VAL~\cite{howey2004val} can also be used at this point.
+Tools for plan validation like VAL~\cite{howey2004val} can also be used at this point but note that the compilation approach does not require a plan so the validation of an STRIPS action model can be done with just an initial and a final state.
 
-\section{Constraining the hypothesis space with example plans}
-Here  we extend our compilation to address the learning scenario defined by $\Lambda$ where a set of plans $\Pi$ is available.
+\subsection{Constraining the hypothesis space with example plans}
+Here  we extend our compilation to address the learning scenario defined by $\Lambda$ where, a set of plans $\Pi$ is available so it is used to reduce the space of possible action models.
 Given a learning task $\Lambda=\tup{\Psi,\Pi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}''$ as follows:
 \begin{itemize}
-\item $F_{\Lambda}$ includes the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of the plans in $\Pi$ with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding $\pi_t\in \Pi$.
+\item $F_{\Lambda}$ extends $F_{\Lambda''}$ by including also the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of the plans in $\Pi$, with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding $\pi_t\in \Pi$.
 We denote $\pi_t$ as a solution plan to the classical planning instances $P_t$ introduced above.
-In addition fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, represent the plan step where the programmed action model is validated.
+In addition fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the plan step where the programmed action model is validated.
 \item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi_1}$ that encode the plan $\pi_1\in \Pi$ for solving $P_1$, and the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
-$G_{\Lambda}=G_{\Lambda}''$.
-\item With respect to $A_{\Lambda}$, the actions for programming the preconditions/effects of a given operator are the same.
+The goals set is the same as in the previous compilation $G_{\Lambda}=G_{\Lambda}''$ that again requires that the programmed action model is validated in all  the learning examples.
+\item With respect to $A_{\Lambda}$.
 \begin{enumerate}
+\item The actions for programming the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
 \item The actions for applying an operator have an extra precondition $f\in F_{\Pi_t}$ that encodes the current plan step and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing the plan step.
-\item The actions for changing the active test have an extra precondition, $at_{|\Pi_t|}$, to indicate that the current plan $\Pi_t$ was fully executed and extra conditional effects to load the next plan $\Pi_{t+1}$ where the programmed operators are validated:
+\item The actions for updating the active test have an extra precondition, $at_{|\Pi_t|}$, to indicate that the current plan $\Pi_t$ was fully executed and extra conditional effects to load the next plan $\Pi_{t+1}$ where the programmed operators are validated:
 \begin{small}
@@
 Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-In addition, motivated by the success of recent algorithms for exploring planning tasks~\cite{geffner:pwithsimulators:IJCAI17}, we do not assume that a learning set of plans is given apriori but instead, we autonomously collect the learning examples.
+In addition, motivated by the success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17}, we do not assume that a learning set of plans is given apriori but instead, we autonomously collect the learning examples.
 

### 2017-10-13 doc/icaps-18/strips-learning.tex
first evaluation results added

@@
 
-Likewise, we assume that each action in $A$ is instantiated from a STRIPS operator schema.
+Likewise, we assume that actions are instantiated from a STRIPS operator schema.
 Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld~\cite{slaney2001blocks} represented in PDDL.
@@
 
-\vspace{0.3cm}
+\vspace{0.5cm}
 
-\begin{subfigure}{.3\textwidth}
+\begin{subfigure}{.25\textwidth}
 \begin{small}
@@
 \end{subfigure}%
-\begin{subfigure}{.3\textwidth}
+\begin{subfigure}{.6\textwidth}
 {\small\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
@@
 \end{lstlisting}
+\vspace{0.1cm}
 \begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
@@
 \item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan determines the literals that belong to the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed the STRIPS action model $\Xi$ using the available input knowledge ($\Pi$, if available, and the given set of labels $\Sigma$).
+\item Validates the programmed the STRIPS action model $\Xi$ using the given input knowledge ($\Pi$ if available, and the given set of labels $\Sigma$).
 At every learning example {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
@@
 The shown classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-In particular the steps $[0,8]$ are actions for programming the preconditions of the operator, steps $[9,13]$ are actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operator.
+In particular the steps $[0,8]$ are the actions for programming the preconditions of the operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are the actions for validating the programmed operator.
 The details of the compilation are given in the remaining of the section.
@@
 \begin{figure}[hbt]
-\begin{footnotesize}
+\begin{scriptsize}
 \begin{verbatim}
-     0 : (program_pre_clear_stack_var1)
-     1 : (program_pre_handempty_stack_)
-     2 : (program_pre_holding_stack_var2)
-     3 : (program_pre_on_stack_var1_var1)
-     4 : (program_pre_on_stack_var1_var2)
-     5 : (program_pre_on_stack_var2_var1)
-     6 : (program_pre_on_stack_var2_var2)
-     7 : (program_pre_ontable_stack_var1)
-     8 : (program_pre_ontable_stack_var2)
-     9 : (program_eff_clear_stack_var1)
-    10 : (program_eff_clear_stack_var2)
-    11 : (program_eff_handempty_stack_)
-    12 : (program_eff_holding_stack_var1)
-    13 : (program_eff_on_stack_var1_var2)
+     0 : (program_pre_clear_stack_v1)
+     1 : (program_pre_handempty_stack)
+     2 : (program_pre_holding_stack_v2)
+     3 : (program_pre_on_stack_v1_v1)
+     4 : (program_pre_on_stack_v1_v2)
+     5 : (program_pre_on_stack_v2_v1)
+     6 : (program_pre_on_stack_v2_v2)
+     7 : (program_pre_ontable_stack_v1)
+     8 : (program_pre_ontable_stack_v2)
+     9 : (program_eff_clear_stack_v1)
+    10 : (program_eff_clear_stack_v2)
+    11 : (program_eff_handempty_stack)
+    12 : (program_eff_holding_stack_v1)
+    13 : (program_eff_on_stack_v1_v2)
     14 : (apply_unstack a b i1 i2)
@@
 \end{verbatim}
-\end{footnotesize}
+\end{scriptsize}
  \caption{\small Example of a plan for programing and validating the operator schema unstack using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.}
@@
 same fluents and actions and differ only in the initial state and goals).
-The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of different objects appearing in the examples.
+The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of different objects appearing in the input learning examples.
 Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the STRIPS operator schema $\xi\in \Xi$.
-\item A fluent $mode_{prog}$ indicating whether the operator schema are being programmed or they are already programmed operators so are being validated in the learning examples. 
+\item A fluent $mode_{prog}$ indicating whether the operator schema are being programmed or they are already programmed operators, so are being validated in the learning examples. 
 \item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
@@
 \item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (i.e.
-the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially, any operator schema, contains all the possible preconditions, no negative effect and no positive effect) and $mode_{prog}$.
+the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (note that our compilation assumes that initially, any operator schema, contains all the possible preconditions, no negative effect and no positive effect) and the $mode_{prog}$ set to true.
 \item $G_{\Lambda''}=\{test_t\}$, {\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
@@
 \hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
-& mode_{prog},pre_{f}(\xi)\},\\    
+& pre_{f}(\xi), mode_{prog}\},\\    
 \cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
@@
 
-\item The actions for adding a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
+\item The actions for adding a {\em negative} or a {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
 
@@
 
-For instance, these actions define that if an operator is programmed with the precondition $holding(v_1)\in F_v$ it {\em implies} ($\implies$) that $holding(block_1)\in F$ has to be true in the current state if the operator binds the variable object $v_1\in\Omega_v$ with the regular object $block_1\in\Omega$.
-The operator binding is done implicitly by order of appearance, i.e.
+The binding of the operator schema is done implicitly by order of appearance, i.e.
 the variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing in the same position.
+To illustrate this, Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying an already programmed operator schema $stack$.
 
@@
 Once the preconditions of an operator schema $\Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after the operator preconditions are programmed, because of the preconditions of the $programPre_{f,\xi}$ actions).
+The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after the operator preconditions are programmed, because of the preconditions in the $programPre_{f,\xi}$ actions).
 Furthermore, once the operator schema are programmed they can only be applied because of the value of the $mode_{prog}$ fluent.
-The only way of achieving a goal $\{test_t\}$,{\small $1\leq t\leq \tau$} is by executing an applicable sequence of programmed operator schema that achieves the final state defined by the associated label $\sigma_t$, starting from the initial defined in this same label.
-If this is done for all the input examples of the learning task (for all the labels $\sigma_t$, {\small $1\leq t\leq \tau$}), it means that the programmed action model $\Xi$ is compliant with all the learning input knowledge and hence, it is a solution to the action model learning task.
+To solve the classical planning task $P_{\Lambda''}$, the only way of achieving a goal $\{test_t\}$,{\small $1\leq t\leq \tau$} is by executing an applicable sequence of programmed operator schema that achieves the final state $s_n^t$ defined by the associated label $\sigma_t$, starting from the initial $s_0^t$ defined in this same label.
+If this is done for all the input examples of the learning task (for all the labels $\sigma_t$, {\small $1\leq t\leq \tau$}), it means that the programmed action model $\Xi$ is compliant with all the input knowledge and hence, it is a solution to the action model learning task.
 \end{proof}
+
+\begin{figure}[hbt]
+\begin{tiny}
+\begin{verbatim}
+(:action apply_stack
+  :parameters (?o1 - object ?o2 - object)
+  :precondition
+   (and (or (not(pre_on_stack_v1_v1)) (on ?o1 ?o1))
+        (or (not(pre_on_stack_v1_v2)) (on ?o1 ?o2))
+        (or (not(pre_on_stack_v2_v1)) (on ?o2 ?o1))
+        (or (not(pre_on_stack_v2_v2)) (on ?o2 ?o2))
+        (or (not(pre_ontable_stack_v1)) (ontable ?o1))
+        (or (not(pre_ontable_stack_v2)) (ontable ?o2))
+        (or (not(pre_clear_stack_v1)) (clear ?o1))
+        (or (not(pre_clear_stack_v2)) (clear ?o2))
+        (or (not(pre_holding_stack_v1)) (holding ?o1))
+        (or (not(pre_holding_stack_v2)) (holding ?o2))
+        (or (not(pre_handempty_stack)) (handempty)))
+  :effect
+   (and (when (del_on_stack_v1_v1) (not(on ?o1 ?o1)))
+        (when (del_on_stack_v1_v2) (not(on ?o1 ?o2)))
+        (when (del_on_stack_v2_v1) (not(on ?o2 ?o1)))
+        (when (del_on_stack_v2_v2) (not(on ?o2 ?o2)))
+        (when (del_ontable_stack_v1) (not(ontable ?o1)))
+        (when (del_ontable_stack_v2) (not (ontable ?o2)))
+        (when (del_clear_stack_v1) (not (clear ?o1)))
+        (when (del_clear_stack_v2) (not (clear ?o2)))
+        (when (del_holding_stack_v1) (not (holding ?o1)))
+        (when (del_holding_stack_v2) (not (holding ?o2)))
+        (when (del_handempty_stack) (not (handempty)))
+        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
+        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
+        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
+        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
+        (when (add_ontable_stack_v1) (ontable ?o1))
+        (when (add_ontable_stack_v2) (ontable ?o2))
+        (when (add_clear_stack_v1) (clear ?o1))
+        (when (add_clear_stack_v2) (clear ?o2))
+        (when (add_holding_stack_v1) (holding ?o1))
+        (when (add_holding_stack_v2) (holding ?o2))
+        (when (add_handempty_stack) (handempty))
+        (when (modeProg) (not(modeProg)))))
+\end{verbatim}
+\end{tiny}
+ \caption{\small Action for applying an already programmed operator schema $stack$ as encoded in PDDL.}
+\label{fig:compilation}
+\end{figure}
 
@@
 We evaluated our learning approach for different amounts of available input knowledge.
-In all the experiments the compilation is solved using the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar} given its ability to deal with planning instance highly populated with dead-ends.
+In all the experiments the classical planning instances resulting from our compilations are adressed by the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar} given its ability to deal with planning instances populated with dead-ends.
 
@@
 Here we assess the performance of our learning approach when addressing the $\Lambda$ learning task.
-The quality of the learned models is quantified using the cardinality of the {\em symmetric difference} between the set of preconditions, negative and positive effects (1), in the learned model and (2), in the actual models taken as reference.
+The quality of the learned models is quantified using the {\em precision} and {\em recall} metrics frequently used in {\em pattern recognition}~\cite{davis2006relationship}.
+This metrics are computed in the learned models with respect to the actual models taken as reference.
 
-Last but not least, collecting {\em informative} examples for learning planning action models is challenging open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-In addition, motivated by the success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17}, we do not assume that a learning set of plans is given apriori but instead, we autonomously collect the learning examples.
+Table~\ref{tab:results_plans} summarizes the obtained results.
 
 \begin{table}[hbt!]
-\begin{footnotesize}
-\begin{tabular}{cc}
-
+	\begin{center}
+		\begin{tabular}{l|l|l|l|l|l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			 
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+			\hline
+			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Gripper & 1.0 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Miconic & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Floortile & 0.2 & 0.2 & 0.8 & 0.5 & 1.0 & 0.9 \\
+			Pegsol & - & - & - & - & - & - \\
 		\end{tabular}
-\end{footnotesize}
-\caption{\small Mean error and standard deviation of the learned models.}
-  \label{tab:eplans}
+	\end{center}
+\label{tab:results_plans}	        
+\caption{The {\em precision} and {\em recall} values achieved by the learned models in the different domains.}
 \end{table}
 
-Table~\ref{tab:eplans} shows the mean error and standard deviation of the learned models with respect to the actual action models.
-The standard deviation provides a measure of how this error is distributed among the different operators in the domain.
-If this deviation is 0 it means that is equally distributed in all the domain operators.
-
 \subsection{Learning action models from example states}
-When the set of input planning task is not available the {\em symmetric difference} is not an acceptable measure for evaluating the quality of the learned models.
-The reason is that, in this situation, the planner determines also the actions that must satistying the input labels so actions can reformulated and still be compliant with the learning examples.
+When the set of input planning task is not available, the planner determines also the actions that must satistying the input labels so actions can reformulated and still be compliant with the learning examples.
 For instance a blocksworld can be learned where the operator {\small\tt stack} is defined with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
-
-With this regard, the Table~\ref{tab:noplans} shows the planning time and number of steps required to build a set of action models compliant with the input examples.
-
-\begin{table}[hbt!]
-\begin{footnotesize}
-\begin{tabular}{cc}
-
-\end{tabular}
-\end{footnotesize}
-\caption{\small Planning time and number of steps required to build a set of action models compliant with the input examples.}
-  \label{tab:noplans}
-\end{table}
 
@@
 
+This research opens the door to the bootstrapping of action model.
+That is that a planner collects by its own the data that it is necessary to define an action model for a given planning task.
+With this regard, collecting {\em informative} examples for learning planning action models is challenging open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+The recent success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel approaches that do not assume that a learning set of plans is given apriori but instead, autonomously collect the learning examples.
+
 \subsection*{Acknowledgment}

### 2017-10-13 src/compiler.py
small optimization added

@@ -120,7 +120,7 @@ for a in actions: # All possible preconditions are initially programmed
       for tup in itertools.product(var_ids, repeat=(len(p)-1)):
          if possible_pred_for_action(p,a,tup):       
             vars = ["var"+str(t) for t in tup]         
-            fd_task.init.append(pddl.conditions.Atom("pre_"+p[0]+"_"+a[0],vars))
+            fd_task.init.append(pddl.conditions.Atom("pre_"+p[0]+"_"+a[0]+"_"+"_".join(map(str,vars)),[]))
             allpres = allpres + [str("pre_"+p[0]+"_"+a[0]+"_"+"_".join(map(str,vars)))]
 
 if input_level <= config.INPUT_LENPLAN:               
@@ -134,17 +134,14 @@ for i in range(0,len(plans)+1):
 fd_task.goal=pddl.conditions.Conjunction(goals)
                
 # Compilation Domain
-fd_task.types.append(pddl.pddl_types.Type("var","None"))
 if input_level <= config.INPUT_LENPLAN:               
    fd_task.types.append(pddl.pddl_types.Type("step","None"))
-
-for i in range(1,MAX_VARS+1):
-   fd_task.objects.append(pddl.pddl_types.TypedObject("var"+str(i),"var"))
    
 if input_level <= config.INPUT_LENPLAN:                  
    for i in range(1,MAX_STEPS+2):
       fd_task.objects.append(pddl.pddl_types.TypedObject("i"+str(i),"step"))   
-   
+
+      
 fd_task.predicates.append(pddl.predicates.Predicate("modeProg",[]))
 for i in range(0,len(plans)+1):
    fd_task.predicates.append(pddl.predicates.Predicate("test"+str(i),[]))
@@ -152,12 +149,17 @@ if input_level <= config.INPUT_LENPLAN:
    fd_task.predicates.append(pddl.predicates.Predicate("current",[pddl.pddl_types.TypedObject("?i","step")]))
    fd_task.predicates.append(pddl.predicates.Predicate("next",[pddl.pddl_types.TypedObject("?i1","step"),pddl.pddl_types.TypedObject("?i2","step")]))
 
-for a in actions:   
-   for p in predicates:   
-      if (len(p)<= len(a)):
[…]

### 2017-10-16 doc/icaps-18/strips-learning.tex
pages 1-5 updated + new results by Diego

@@
 This paper presents a novel approach for learning STRIPS action models from examples that compiles this inductive learning task into a classical planning task.
-Interestingly the proposed compilation is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where neither actions nor intermediate states are observed.
-What is more, the compilation can reuse previous knowledge (it is able to start from partially specified action models) and can also be used to validate whether observation of plan executions follow a given STRIPS action model.
+Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where neither actions nor intermediate states are given.
+What is more, partially specified action models can be used to feed the compilation and the compilation can also be used to validate whether certain observations of plan executions follow a given STRIPS action model.
 \end{abstract}
@@
 \section{Introduction}
-Besides plan synthesis~\cite{ghallab2004automated,geffner:book:2013}, planning action models are also useful for plan/goal recognition~\cite{ramirez2012plan}.
-In both tasks, off-the-shelf planners require reasoning about action models that correctly and completely capture the possible world transitions.
-Unfortunately, building such planning action models is complex, even for planning experts, and this knowledge acquisition bottleneck limits the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.
+Besides {\em plan synthesis}~\cite{ghallab2004automated,geffner:book:2013}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
+At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions.
+Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.
 
-On the other hand, Machine Learning (ML) techniques have shown able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML techniques for learning planning STRIPS action models from examples is not straightforward though:
+On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
+The application of inductive ML to the learning of STRIPS action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
 \begin{itemize}
-\item The inputs to ML algorithms usually are finite vectors that encode the value of fixed features for a given set of objects.
+\item The inputs to ML algorithms (the learning data) usually are finite vectors that encode the value of fixed features for a given set of objects.
 The input for learning planning action models traditionally are sets of observations of plan executions (each with possibly a different length).
 \item The traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
-When learning STRIPS action models, the output is not a scalar but the sets of preconditions and effects of each action that define the possible state transitions of a given planning domain.
+When learning STRIPS action models the output are, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions of a given planning domain.
 \end{itemize}
 
-Learning STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation (so an off-the-shelf planner can be used to address the inductive learning task).
-In addition, the compilation approach presents the following benefits:
+The learning of STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
+Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation.
+A compilation approach is appealing because an off-the-shelf planner can be used to address the inductive learning task opening the door to the bootstrapping of action models in planning.
+In addition, a compilation approach presents the following contributions:
 \begin{enumerate}
@@
 \item Can exploit previous knowledge about the action model in the form of partially specified action models.
-\item Can be used to validate whether observed plan executions are valid for a given STRIPS action model.
+In the extreme, it can be used to validate whether observed plan executions are valid for a given STRIPS action model.
 \end{enumerate}
@@
 The paper is organized as follows.
-The first section presents the classical planning model, its extension to conditional effects (a requirement for the proposed compilation) and the STRIPS action model.
-The second section formalizes the task of learning STRIPS action models from observations of plan executions with regard to different amounts of input knowledge.
-The third section describes our approach for addressing these inductive learning tasks by compiling them into classical planning.
-The fourth section reports the collected data during the evaluation of the performance of our approach.
-Finally, the last section discusses the strengths and weaknesses of our approach and proposes some future work.
+The first section presents the classical planning model, its extension to conditional effects (which is a requirement for the proposed compilation) and formalizes the STRIPS action model, the output of the addressed learning task.
+The second section formalizes the task of learning STRIPS action models with regard to different amounts of input knowledge (observations of plan executions).
+The third section describes our approach for addressing the learning of STRIPS action models by compiling it into classical planning and how previous knowledge can be introduced in the compilation.
+The fourth section reports the collected data during the empirical evaluation of the compilation approach.
+Finally, the last section discusses the strengths and weaknesses of the compilation approach for learning action models and proposes some opportunities of future research.
 
 \section{Background}
-This section defines the planning models used on this work.
+This section defines the planning models used on this work and the STRIPS action model, output of the addressed learning task.
 
@@
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of negative effects $\eff^-(a)\in\mathcal{L}(F)$, and a set of positive effects $\eff^+(a)\in\mathcal{L}(F)$.
-An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff^-(a))\cup\eff^+(a)$.
+Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of effects $\eff(a)\in\mathcal{L}(F)$.
+An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
 
@@
 Our approach for leaning STRIPS action models is compiling this leaning task into a classical planning task with conditional effects.
-We use conditional effects because they allow us to compactly define actions whose particular effects depend on the current state.
+Conditional effects allow us to compactly define actions whose particular effects depend on the current state.
 Many classical planners cope with conditional effects without compiling them away.
@@
 \]
-i.e.~effects whose conditions hold in $s$.
-The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg \eff^-(s,a))\cup\eff^+(s,a)$, where $\eff^-(s,a)$ and $\eff^+(s,a)$ are the negative and positive effects in $\eff(s,a)$.
+i.e.~effects whose conditions hold in $s$. The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg \eff(s,a))\cup\eff(s,a)$.
 
 \subsection{The STRIPS action model}
-In this work we address the task of learning  action  schema  in  the  STRIPS language~\cite{fikes1971strips}.
-With this regard, we assume that the fluents in $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
+This work addresses the learning  action schema in  the  STRIPS language~\cite{fikes1971strips}.
+To define the output of this learning task, we assume that the fluents in $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
 There exists a set of predicates $\Psi$, each $p\in\Psi$ with an argument list of arity $ar(p)$.
@@
 Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names}.
-The cardinality of this set, $|\Omega_v|$, is given by the maximum arity of an action in the given planning frame.
+The cardinality of this set, $|\Omega_v|$, is given by the maximum arity of an action in a given planning frame $\Phi$.
 For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators {\small\tt stack} and {\small\tt unstack} are the ones with the maximum arity (two parameters each).
+Let us define also a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} in $\Omega_v$.
+In the blocksworld $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.
 
-Let us define a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} in $\Omega_v$.
-In the blocksworld this set is defined as $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.
-
-We are now ready to define a STRIPS action model $\Xi$ as a set of operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$:
+We are now ready to define a STRIPS action model $\Xi$ as a set of operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ such that:
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, represents the operator {\em header} defined by its corresponding action name and a enumeration of the variable names given by the arity of the operator, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
-The headers for a 4-operator blocksworld are then {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its corresponding action name and a enumeration of the variable names bound by the arity of the operator, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
+The headers for a 4-operator blocksworld are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
 \item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
@@
 a set of plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
-In this case, the operators schema are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Likewise, preconditions are derived lifting the minimal set of literals that appears in all the pre-states that correspond to the same operator.
+In this case, a STRIPS operator schema is derived by lifting the literals that change between the pre and post-state of the corresponding action executions.
+Likewise, the preconditions are derived lifting the minimal set of literals that appears in all the pre-states that correspond to the same operator schema.
 
@@
 \item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans s.t.
-each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces a state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
 \item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$, is a set of labels s.t.
@@
 
-To illustrate this, Figure~\ref{fig:lexample}, shows the contain of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$ for a $\Lambda$ task that aims learning the STRIPS action model of the blocksworld.
-The shown learning task has a single learning example that corresponds to observing a plan execution for inverting a tower of four blocks.
+To illustrate this, Figure~\ref{fig:lexample} shows an example of a task for learning the STRIPS action model of the blocksworld.
+Figure~\ref{fig:lexample} shows the content of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, this learning task has a single learning example, that corresponds to observing a plan execution for inverting a tower of four blocks.
 
@@
 
- \caption{\small Learning task in the blocksworld with a single learning example that corresponds to inverting a tower of four blocks.}
+ \caption{\small Example of a task for learning the STRIPS action model of the blocksworld from a single learning example.}
 \label{fig:lexample}
@@
 
-A solution to the learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with different name in the example plans) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
-With this regard, the operator schema shown in Figure~\ref{fig:stack} is compliant with the learning task illustrated by the Figure~\ref{fig:lexample}.
+A solution to a learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
 
@@
 Here we reduce the amount of input knowledge provided to the learning task.
-Now $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is replaced by $\Pi'=\{|\pi_1|,\ldots,|\pi_{\tau}|\}$ i.e.
+First, $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is replaced by $\Pi'=\{|\pi_1|,\ldots,|\pi_{\tau}|\}$ i.e.
 $\Pi'$ that does not contain a set of plans but the number of actions of each plan.
@@
 
-Finally, we can go a step further and redefine a third learning task $\Lambda''=\tup{\Psi,\Sigma}$ that corresponds to watching only the results of the plan executions.
+Finally, we go one step further and redefine a third learning task $\Lambda''=\tup{\Psi,\Sigma}$ that corresponds to watching only the results of the plan executions.
 In this case no information about the executed plans is given.
@@
 
-In these two cases $\Lambda'$ and $\Lambda''$, a solution must not only synthesize a possible action models but also the actions could have produced the given labels (this information about the actions is no longer given in the learning examples).
-However, we assume that at the headers of the operators schema are known.
+In these two cases $\Lambda'$ and $\Lambda''$, a solution must not only determine a possible STRIPS action model but also the plans that explain the given labels using the learned model (this information about the actions is no longer given in the learning examples).
+However, we assume that the headers of the operators schema are known.
 
@@
 \item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan determines the literals that belong to the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed the STRIPS action model $\Xi$ using the given input knowledge ($\Pi$ if available, and the given set of labels $\Sigma$).
-At every learning example {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
-If information about the corresponding plan $\pi_t$ is available, then it is used to constrain the validation of the programmed STRIPS action model $\Xi$.
+\item Validates the programmed the STRIPS action model $\Xi$ using the given input knowledge (the given set of labels $\Sigma$, and $\Pi$ if available).
+For every label {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
+We call this process the validation of the programmed STRIPS $\Xi$ at the learning example {\small $1\leq t\leq \tau$}.
+If information about the corresponding plan $\pi_t$ is available, it is used to constrain the validation of $\Xi$.
 \end{enumerate}
@@
 Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ class.
-The shown classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+This classical plan programs and validates the operator schema {\tt\small stack}, from the blocksworld, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
 In particular the steps $[0,8]$ are the actions for programming the preconditions of the operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are the actions for validating the programmed operator.
@@
 \end{scriptsize}
- \caption{\small Example of a plan for programing and validating the operator schema unstack using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.}
+ \caption{\small Example of a plan for programing and validating the operator schema $unstack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schema for $pickup$, $putdown$ and $unstack$.}
 \label{fig:plan}
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the STRIPS operator schema $\xi\in \Xi$.
-\item A fluent $mode_{prog}$ indicating whether the operator schema are being programmed or they are already programmed operators, so are being validated in the learning examples. 
+\item A fluent $mode_{prog}$ indicating whether the operator schema are being programmed or they are already programmed operators that are being validated in the learning examples. 
 \item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
@@
 \item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (i.e.
-the initial state of the first learning example), every fluent $pre_f(\xi)\in F_{\Lambda}$ (note that our compilation assumes that initially, any operator schema, contains all the possible preconditions, no negative effect and no positive effect) and the $mode_{prog}$ set to true.
+the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (note that our compilation assumes that initially, any operator schema, contains all the possible preconditions, no negative effect and no positive effect) and the $mode_{prog}$ set to true.
 \item $G_{\Lambda''}=\{test_t\}$, {\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
@@
 \begin{enumerate}
-\item The actions for programming the operator schema. This set includes:
+\item The actions for programming the operator schema which includes:
 \begin{itemize}
@@
 
-The binding of the operator schema is done implicitly by order of appearance, i.e.
+The binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
 the variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing in the same position.
@@
 Once the preconditions of an operator schema $\Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after the operator preconditions are programmed, because of the preconditions in the $programPre_{f,\xi}$ actions).
-Furthermore, once the operator schema are programmed they can only be applied because of the value of the $mode_{prog}$ fluent.
-To solve the classical planning task $P_{\Lambda''}$, the only way of achieving a goal $\{test_t\}$,{\small $1\leq t\leq \tau$} is by executing an applicable sequence of programmed operator schema that achieves the final state $s_n^t$ defined by the associated label $\sigma_t$, starting from the initial $s_0^t$ defined in this same label.
-If this is done for all the input examples of the learning task (for all the labels $\sigma_t$, {\small $1\leq t\leq \tau$}), it means that the programmed action model $\Xi$ is compliant with all the input knowledge and hence, it is a solution to the action model learning task.
+The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after operator preconditions are programmed, because of the preconditions in the $programPre_{f,\xi}$ actions).
+Furthermore, once the operator schema are programmed they can only be applied because of the $mode_{prog}$ fluent.
+To solve the classical planning task $P_{\Lambda''}$, there is only one way of achieving a goal $\{test_t\}$,{\small $1\leq t\leq \tau$} that is by executing an applicable sequence of programmed operator schema that achieves the final state $s_n^t$ defined by the associated label $\sigma_t$, starting from the initial $s_0^t$ defined in this same label.
+If this is achieved for all the input examples of the learning task $\Lambda''$ (i.e.
+for all the labels $\sigma_t$, {\small $1\leq t\leq \tau$}), it means that the programmed action model $\Xi$ is compliant with all the input knowledge and hence, it is a solution to $\Lambda''$.
 \end{proof}
@@
 \begin{figure}[hbt]
-\begin{tiny}
+\begin{scriptsize}
 \begin{verbatim}
@@
   :precondition
-   (and (or (not(pre_on_stack_v1_v1)) (on ?o1 ?o1))
-        (or (not(pre_on_stack_v1_v2)) (on ?o1 ?o2))
-        (or (not(pre_on_stack_v2_v1)) (on ?o2 ?o1))
-        (or (not(pre_on_stack_v2_v2)) (on ?o2 ?o2))
-        (or (not(pre_ontable_stack_v1)) (ontable ?o1))
-        (or (not(pre_ontable_stack_v2)) (ontable ?o2))
-        (or (not(pre_clear_stack_v1)) (clear ?o1))
-        (or (not(pre_clear_stack_v2)) (clear ?o2))
-        (or (not(pre_holding_stack_v1)) (holding ?o1))
-        (or (not(pre_holding_stack_v2)) (holding ?o2))
-        (or (not(pre_handempty_stack)) (handempty)))
+   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
+        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
+        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
+        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
+        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
+        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
+        (or (not (pre_clear_stack_v1)) (clear ?o1))
+        (or (not (pre_clear_stack_v2)) (clear ?o2))
+        (or (not (pre_holding_stack_v1)) (holding ?o1))
+        (or (not (pre_holding_stack_v2)) (holding ?o2))
+        (or (not (pre_handempty_stack)) (handempty)))
   :effect
-   (and (when (del_on_stack_v1_v1) (not(on ?o1 ?o1)))
-        (when (del_on_stack_v1_v2) (not(on ?o1 ?o2)))
-        (when (del_on_stack_v2_v1) (not(on ?o2 ?o1)))
-        (when (del_on_stack_v2_v2) (not(on ?o2 ?o2)))
-        (when (del_ontable_stack_v1) (not(ontable ?o1)))
+   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
+        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
+        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
+        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
+        (when (del_ontable_stack_v1) (not (ontable ?o1)))
         (when (del_ontable_stack_v2) (not (ontable ?o2)))
@@
 \end{verbatim}
-\end{tiny}
+\end{scriptsize}
  \caption{\small Action for applying an already programmed operator schema $stack$ as encoded in PDDL.}
@@
 Interestingly, the compilation does not require to start the learning of STRIPS action models from scratch and accepts partially specified operator schema where some preconditions are effects are a priori known.
-In more detail, the known preconditions and effects (fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$) can be part of the initial state $I_{\Lambda''}$ and the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) be removed from $A_{\Lambda''}$ reducing the number of applicable actions  and making the classical planning task $P_{\Lambda''}$ easier.
-For instance the classical plan in Figure~\ref{fig:plan} is a solution to a learning task for getting the blocksworld action model where the operator schema for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already known.
+
+The known preconditions and effects can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true that are part of the initial state $I_{\Lambda''}$.
+Likewise, the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) can be removed from $A_{\Lambda''}$ reducing the number of applicable actions  and making the classical planning task $P_{\Lambda''}$ easier to be solved.
+With this regard, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task for getting the blocksworld action model where the operator schema for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are known in advance.
 
@@
 \begin{itemize}
-\item $F_{\Lambda}$ extends $F_{\Lambda''}$ by including also the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of the plans in $\Pi$, with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding $\pi_t\in \Pi$.
-We denote $\pi_t$ as a solution plan to the classical planning instances $P_t$ introduced above.
-In addition fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the plan step where the programmed action model is validated.
+\item $F_{\Lambda}$ extends $F_{\Lambda''}$ by including also the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of a plan in $\Pi$, with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding $\pi_t\in \Pi$.
+We denote $\pi_t$ a solution plan to a classical planning instance $P_t$.
+In addition fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the plan steps where the programmed action model is validated.
 \item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi_1}$ that encode the plan $\pi_1\in \Pi$ for solving $P_1$, and the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
-The goals set is the same as in the previous compilation $G_{\Lambda}=G_{\Lambda}''$ that again requires that the programmed action model is validated in all  the learning examples.
+Goals are the same as in the previous compilation $G_{\Lambda}=G_{\Lambda}''$ so again require that the programmed action model is validated in all  the learning examples.
 \item With respect to $A_{\Lambda}$.
@@
 \item The actions for applying an operator have an extra precondition $f\in F_{\Pi_t}$ that encodes the current plan step and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing the plan step.
-\item The actions for updating the active test have an extra precondition, $at_{|\Pi_t|}$, to indicate that the current plan $\Pi_t$ was fully executed and extra conditional effects to load the next plan $\Pi_{t+1}$ where the programmed operators are validated:
+\item The actions for updating the active test have an extra precondition, $at_{|\Pi_t|}$, to indicate that the current plan $\Pi_t$ was fully executed and extra conditional effects to load the next plan $\Pi_{t+1}$:
 \begin{small}
@@
 We evaluated our learning approach for different amounts of available input knowledge.
-In all the experiments the classical planning instances resulting from our compilations are adressed by the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar} given its ability to deal with planning instances populated with dead-ends.
+In all the experiments the classical planning instances resulting from our compilations are addressed by the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar} given its ability to deal with planning instances populated with dead-ends.
+
+The domains used in the evaluations are domains taken from different IPCs that satisfy the STRIPS requirement.
 
@@
 This metrics are computed in the learned models with respect to the actual models taken as reference.
-
 Table~\ref{tab:results_plans} summarizes the obtained results.
@@
 	\end{center}	        
+\caption{The {\em precision} and {\em recall} values achieved by the learned models in the different domains.}
 \label{tab:results_plans}
-\caption{The {\em precision} and {\em recall} values achieved by the learned models in the different domains.}
+\end{table}
+
+Table~\ref{tab:time_plans} reports the planning time, and prepossessing time used to solve the classical planning task that results from our compilation as well as the plan length of the found solutions.
+
+\begin{table}[hbt!]
+	\begin{center}
+		\begin{tabular}{l|l|l|l|}			 
+			Domain & Total time & Preprocess & Plan length  \\
+			\hline
+			Blocks & 0.06 & 0.01 & 72  \\
+			Visitall & 0.43 & 0.40 & 37 \\
+			Gripper & 0.03 & 0.00 & 43  \\
+			Miconic & 0.12 & 0.09 & 59  \\
+			Floortile & 0.70 & 0.57 & 141 \\
+			Pegsol & - & - & - \\
+		\end{tabular}
+	\end{center}
+	\caption{Planning time, preprocessing time and Plan length.}
+	\label{tab:time_plans}	
 \end{table}

### 2017-10-23 src/compiler2.py
fixed zenotravel domain

@@ -76,26 +76,26 @@ def possible_pred_for_action(task, p, a, tup):
 # **************************************#
 # MAIN
 # **************************************#
-# try:
-#     domain_folder_name  = sys.argv[1]
-#     domain_file = sys.argv[2]
-#     problems_prefix_filename = sys.argv[3]
-#     plans_prefix_filename = sys.argv[4]
-#     input_level = int(sys.argv[5])
-#
-#
-# except:
-#    print "Usage:"
-#    print sys.argv[0] + " <domain> <domain filename> <problems prefix>  <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
-#    sys.exit(-1)
+try:
+    domain_folder_name  = sys.argv[1]
+    domain_file = sys.argv[2]
+    problems_prefix_filename = sys.argv[3]
+    plans_prefix_filename = sys.argv[4]
+    input_level = int(sys.argv[5])
+
+
+except:
+    print "Usage:"
+    print sys.argv[0] + " <domain> <domain filename> <problems prefix>  <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
+    sys.exit(-1)
 
 
 # ../benchmarks/handpicked/blocks/ test plan 0
-domain_folder_name = "../benchmarks/handpicked/blocks/"
-domain_file = "domain"
-problems_prefix_filename = "test"
-plans_prefix_filename = "plan"
-input_level = 0
+#domain_folder_name = "../benchmarks/handpicked/blocks/"
+#domain_file = "domain"
+#problems_prefix_filename = "test"
+#plans_prefix_filename = "plan"
+#input_level = 0
 
 # Reading the example plans
 plans = []

### 2017-10-23 doc/icaps-18/strips-learning.tex
Diego's mail added

@@
 {\small Camino de Vera s/n. 46022 Valencia, Spain}\\
-{\small \{,serjice,onaindia\}@dsic.upv.es}}
+{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
 
@@
 \begin{abstract}
-This paper presents a novel approach for learning STRIPS action models from examples that compiles this inductive learning task into a classical planning task.
+This paper presents a novel approach for learning STRIPS action models from examples that compiles this particular inductive learning task into a classical planning task.
 Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where neither actions nor intermediate states are given.
@@
 \item The inputs to ML algorithms (the learning data) usually are finite vectors that encode the value of fixed features for a given set of objects.
-The input for learning planning action models traditionally are sets of observations of plan executions (each with possibly a different length).
-\item The traditional output of off-the-shelf ML techniques is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
-When learning STRIPS action models the output are, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions of a given planning domain.
+The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
+\item The traditional output of ML algorithms is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
+When learning STRIPS action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions of the corresponding planning domain.
 \end{itemize}
@@
 Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation.
-A compilation approach is appealing because an off-the-shelf planner can be used to address the inductive learning task opening the door to the bootstrapping of action models in planning.
-In addition, a compilation approach presents the following contributions:
+The compilation approach is appealing because an off-the-shelf planner can be used to address the inductive learning task opening the door to the bootstrapping of action models in planning but also because:
 \begin{enumerate}
-\item It is flexible to different amounts of available input knowledge.
-The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no actions or intermediate states are observed.
-\item Can exploit previous knowledge about the action model in the form of partially specified action models.
-In the extreme, it can be used to validate whether observed plan executions are valid for a given STRIPS action model.
+\item Is flexible to different amounts of available input knowledge.
+The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate actions or states are observed.
+\item Can exploit previous knowledge about the structure of the actions in the form of partially specified action models.
+In the extreme, the compilation can be used to validate whether observed plan executions are valid for a given STRIPS action model.
 \end{enumerate}
@@
 The first section presents the classical planning model, its extension to conditional effects (which is a requirement for the proposed compilation) and formalizes the STRIPS action model, the output of the addressed learning task.
-The second section formalizes the task of learning STRIPS action models with regard to different amounts of input knowledge (observations of plan executions).
-The third section describes our approach for addressing the learning of STRIPS action models by compiling it into classical planning and how previous knowledge can be introduced in the compilation.
-The fourth section reports the collected data during the empirical evaluation of the compilation approach.
-Finally, the last section discusses the strengths and weaknesses of the compilation approach for learning action models and proposes some opportunities of future research.
+The second section formalizes the learning of STRIPS action models with regard to different amounts of input knowledge.
+The third section describes our approach for addressing this particular learning task by compiling it into classical planning and how previous knowledge can be introduced in the compilation.
+The fourth section reports the data collected during the empirical evaluation of our approach.
+Finally, the last section discusses the strengths and weaknesses of our approach and proposes several opportunities for future research.
 
 \section{Background}
-This section defines the planning models used on this work and the STRIPS action model, output of the addressed learning task.
+This section defines the planning models used on this work as well as the STRIPS action model, the output of the learning task addressed in this paper.
 
@@
 Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of effects $\eff(a)\in\mathcal{L}(F)$.
-An action $a\in A$ is applicable in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
+An action $a\in A$ is {\em applicable} in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is the {\em succesor} state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
 
@@
 A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The plan $\pi$ {\em solves} $P$ if and only if $G\subseteq s_n$, i.e.~if the goal condition is satisfied in the state reached after following the application of $\pi$ in $I$.
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of $\pi$ in $I$.
 
@@
 Conditional effects allow us to compactly define actions whose particular effects depend on the current state.
-Many classical planners cope with conditional effects without compiling them away.
-In fact, the support of conditional effects was a requirement for participating at the IPC-2014~\cite{vallati:IPC:AIM2015}.
+Nowadays many classical planners cope with conditional effects without compiling them away.
+In fact, the support of conditional effects was a requirement for participating at the IPC-2014~\cite{vallati:IPC:AIM2015} and IPC-2018.
 
@@
 
-An action $a\in A$ is applicable in state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} is
+An action $a\in A$ is {\em applicable} in state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} is
 \[
@@
 \]
-i.e.~effects whose conditions hold in $s$. The result of applying $a$ in $s$ is a new state $\theta(s,a)=(s\setminus \neg \eff(s,a))\cup\eff(s,a)$.
+i.e.~effects whose conditions hold in $s$. The result of applying $a$ in $s$ is the {\em succesor} state $\theta(s,a)=(s\setminus \neg \eff(s,a))\cup\eff(s,a)$.
 
 \subsection{The STRIPS action model}
-This work addresses the learning  action schema in  the  STRIPS language~\cite{fikes1971strips}.
-To define the output of this learning task, we assume that the fluents in $F$ are instantiated from predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
-There exists a set of predicates $\Psi$, each $p\in\Psi$ with an argument list of arity $ar(p)$.
+This work addresses the learning of action schemes defined in the STRIPS language~\cite{fikes1971strips}.
+To formalized the output of this learning task, we assume that there exists a set of {\em predicates} $\Psi$ and that the fluents in $F$ are instantiated from these predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
+
+Formally, each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
 Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
@@
 
-Likewise, we assume that actions are instantiated from a STRIPS operator schema.
+Likewise, we assume that actions are instantiated from STRIPS operator schemes.
 Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld~\cite{slaney2001blocks} represented in PDDL.
@@
 
-Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names}.
-The cardinality of this set, $|\Omega_v|$, is given by the maximum arity of an action in a given planning frame $\Phi$.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators {\small\tt stack} and {\small\tt unstack} are the ones with the maximum arity (two parameters each).
+Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names} and whose cardinality, $|\Omega_v|$, is given by the maximum arity of an action in a given planning frame $\Phi$.
+For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because operators {\small\tt stack} and {\small\tt unstack}, the ones with the maximum arity, have two parameters.
+
 Let us define also a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} in $\Omega_v$.
@@
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its corresponding action name and a enumeration of the variable names bound by the arity of the operator, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its corresponding action name and a enumeration of the variable names whose cardinality is the operator arity, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
 The headers for a 4-operator blocksworld are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
@@
 Learning STRIPS action models from fully available input knowledge, i.e.
-a set of plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
+from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
 In this case, a STRIPS operator schema is derived by lifting the literals that change between the pre and post-state of the corresponding action executions.
@@
 
-This section formalizes more challenging tasks, for learning STRIPS action models, where less input knowledge is available.
-Next, these learning tasks are formalized according to the available amount of input knowledge.
+This section formalizes more challenging learning tasks, where less input knowledge is available.
 
@@
 
-To illustrate this, Figure~\ref{fig:lexample} shows an example of a task for learning the STRIPS action model of the blocksworld.
-Figure~\ref{fig:lexample} shows the content of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, this learning task has a single learning example, that corresponds to observing a plan execution for inverting a tower of four blocks.
+Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda$ for computing the STRIPS action model of the blocksworld.
+The figure shows the content of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$.
+This learning task has a single learning example, that corresponds to observing the execution of a plan for inverting a tower of four blocks.
 
@@
 
-A solution to a learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their labels $\Sigma$.
+A solution to a learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
 
 \subsection{Learning from initial/final states}
-Here we reduce the amount of input knowledge provided to the learning task.
+Here we reduce the amount of provided input knowledge.
 First, $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is replaced by $\Pi'=\{|\pi_1|,\ldots,|\pi_{\tau}|\}$ i.e.
@@
 While the previous learning task, $\Lambda$, corresponds to watching an agent acting in the world, this new learning task $\Lambda'$ can be understood as watching only the result of its plan execution but knowing the number of actions performed by the agent.
-In the case of the learning example shown in Figure~\ref{fig:lexample}, $|\pi_1|=8$.
+In the case of the learning example shown in Figure~\ref{fig:lexample}, $\Pi'=\{|\pi_1|=8\}$.
 
@@
 In these two cases $\Lambda'$ and $\Lambda''$, a solution must not only determine a possible STRIPS action model but also the plans that explain the given labels using the learned model (this information about the actions is no longer given in the learning examples).
-However, we assume that the headers of the operators schema are known.
+We assume however, that the headers of the operators schema are known.
 
@@
 \begin{enumerate}
-\item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan determines the literals that belong to the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed the STRIPS action model $\Xi$ using the given input knowledge (the given set of labels $\Sigma$, and $\Pi$ if available).
+\item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Validates the programmed STRIPS action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ of $\Pi'$ if available).
 For every label {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
-We call this process the validation of the programmed STRIPS $\Xi$ at the learning example {\small $1\leq t\leq \tau$}.
-If information about the corresponding plan $\pi_t$ is available, it is used to constrain the validation of $\Xi$.
+We call this process the validation of the programmed STRIPS action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
+If information about the corresponding plan $\pi_t$ is available, it is also used in the validation.
 \end{enumerate}
 
-Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ class.
-This classical plan programs and validates the operator schema {\tt\small stack}, from the blocksworld, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-In particular the steps $[0,8]$ are the actions for programming the preconditions of the operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are the actions for validating the programmed operator.
-The details of the compilation are given in the remaining of the section.
+Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ kind.
+This classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} (previously specified operator schemes for $pickup$, $putdown$ and $unstack$ are given).
+In particular the steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are the actions for validating the programmed operator.
+The details of the compilation are explained in the remaining of the section.
 
 \begin{figure}[hbt]
-\begin{scriptsize}
+\begin{footnotesize}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{scriptsize}
- \caption{\small Example of a plan for programing and validating the operator schema $unstack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schema for $pickup$, $putdown$ and $unstack$.}
+\end{footnotesize}
+ \caption{\small Example of a plan for programing and validating the operator schema $unstack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
 \label{fig:plan}
@@
 Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
-The set of actions, $A=\emptyset$, is empty because the action model is unknown.
+The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
 Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the STRIPS operator schema $\xi\in \Xi$.
-\item A fluent $mode_{prog}$ indicating whether the operator schema are being programmed or they are already programmed operators that are being validated in the learning examples. 
+\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or they are already programmed operators that are being validated. 
 \item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
 \end{itemize}
-\item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (i.e.
-the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (note that our compilation assumes that initially, any operator schema, contains all the possible preconditions, no negative effect and no positive effect) and the $mode_{prog}$ set to true.
+\item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema contains every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
 \item $G_{\Lambda''}=\{test_t\}$, {\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
@@
 \begin{enumerate}
-\item The actions for programming the operator schema which includes:
+\item The actions for programming the an operator schema $\xi\in\Xi$ which includes:
 \begin{itemize}
@@
 
-\item The actions for applying an already programmed operator schema $\xi\in\Xi$ bounded with the objects $\omega\subseteq\Omega^{ar(\xi)}$
+\item The actions for applying an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
+The binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
+the variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
+Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
 \begin{small}
@@
 \end{small}
-
-The binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
-the variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing in the same position.
-To illustrate this, Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying an already programmed operator schema $stack$.
 
@@
 \end{lemma}
-
+\begin{small}
 \begin{proof}[Proof sketch]
-Once the preconditions of an operator schema $\Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after operator preconditions are programmed, because of the preconditions in the $programPre_{f,\xi}$ actions).
-Furthermore, once the operator schema are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve the classical planning task $P_{\Lambda''}$, there is only one way of achieving a goal $\{test_t\}$,{\small $1\leq t\leq \tau$} that is by executing an applicable sequence of programmed operator schema that achieves the final state $s_n^t$ defined by the associated label $\sigma_t$, starting from the initial $s_0^t$ defined in this same label.
+The compilation forces that once the preconditions of an operator schema $\Xi$ are programmed, they cannot be altered.
+The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after the operator preconditions are programmed).
+Furthermore, once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
+To solve the classical planning task $P_{\Lambda''}$, there is only one way of achieving goals $\{test_t\}$,{\small $1\leq t\leq \tau$}: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$ defined by the associated label $\sigma_t$, starting from $s_0^t$, defined in this same label.
 If this is achieved for all the input examples of the learning task $\Lambda''$ (i.e.
-for all the labels $\sigma_t$, {\small $1\leq t\leq \tau$}), it means that the programmed action model $\Xi$ is compliant with all the input knowledge and hence, it is a solution to $\Lambda''$.
+for all the labels $\sigma_t$), it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda''$.
  \end{proof}
+\end{small}
 
@@
         (when (add_handempty_stack) (handempty))
-        (when (modeProg) (not(modeProg)))))
+        (when (modeProg) (not (modeProg)))))
 \end{verbatim}
@@
 
-\subsection{Constraining the hypothesis space with previous action models}
+\section{Constraining the learning hypothesis space with additional input knowledge}
+Here we show that further input knowledge can be used to reduce the space of the possible action models and make the learning task more practicable.
+
+\subsection{Static predicates}
+A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action.
+Given a static predicate $p$, the actions $\mathsf{programEff_{f,\xi}}$ (s.t.
+$f\in F_v$ is an instantiation of the static predicate $p$ in $\Omega_V$) are unnecessary for every $\xi\in\Xi$ and hence, they can be removed from the compilation.
+
+Furthermore, actions $\mathsf{programPre_{f,\xi}}$ can also be discarded by checking the initial and final states given the set of labels $\Sigma$.
+If a precondition $f\in F_v$ instantiated from the static predicate $p$ in $\Omega_V$ is not possible for the schema $\xi\in\Xi$ according to a given example {\small $1\leq t\leq \tau$} it means that the corresponding action can be removed.
+
+In our evaluation section we shown how to compute candidates of static predicates from the inputs to our learning tasks and how using this knowledge affects to the performance of our approach.
+
+\subsection{Previously specified action models}
 Interestingly, the compilation does not require to start the learning of STRIPS action models from scratch and accepts partially specified operator schema where some preconditions are effects are a priori known.
@@
 
-\subsection{Constraining the hypothesis space with example plans}
+\subsection{Example plans}
 Here  we extend our compilation to address the learning scenario defined by $\Lambda$ where, a set of plans $\Pi$ is available so it is used to reduce the space of possible action models.
@@
 \section{Evaluation}
-We evaluated our learning approach for different amounts of available input knowledge.
-In all the experiments the classical planning instances resulting from our compilations are addressed by the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar} given its ability to deal with planning instances populated with dead-ends.
+We evaluated our approach for learning STRIPS action models from different amounts of available input knowledge.
+All the experiments run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
+The domains used in the evaluation are domains taken from different IPCs at the {\sc planning.domains} repository~\cite{muise2016planning} and satisfy the STRIPS requirement~\cite{fox2003pddl2}.
 
-The domains used in the evaluations are domains taken from different IPCs that satisfy the STRIPS requirement.
+\subsection{The classical planner}
+The classical planner used to solve the instances that result from our compilations is the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar}.
+We decided to use a SAT-based planner because its ability to deal with planning instances populated with dead-ends.
+In addition, the benefit of SAT-based planners is that all the actions for programming the actions preconditions can be done in parallel, in a single planning step, because they do not interact.
+Likewise, all the actions for programming an action effect can also be applied in a single planning step reducing significantly the planning horizon.
 
-\subsection{Validating action models with classical planning}
+\subsection{Precision and recall}
+The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}.
+
+Formally $Precission=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
+On the other hand recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
+Intuitively precision gives a notion of {\em soundness} while and recall gives a notion of the {\em completeness} of the learned models.
 
 \subsection{Learning action models from example plans}
-Here we assess the performance of our learning approach when addressing the $\Lambda$ learning task.
-The quality of the learned models is quantified using the {\em precision} and {\em recall} metrics frequently used in {\em pattern recognition}~\cite{davis2006relationship}.
-This metrics are computed in the learned models with respect to the actual models taken as reference.
-Table~\ref{tab:results_plans} summarizes the obtained results.
+Here we assess the performance of our learning approach when addressing $\Lambda$ learning tasks.
+Here the {\em precision} and {\em recall} metrics are computed in the learned models with respect to the actual models taken as reference.
+Table~\ref{tab:results_plans} summarizes the obtained results, precision ({\bf P}) and recall ({\bf R}) is computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}).
 
 \begin{table}[hbt!]
+	\begin{small}
 	\begin{center}
@@
 			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 0.8 & 0.7 \\
+			Ferry & 0.8 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Floortile & 0.5 & 0.6 & 0.9 & 0.8 & 1.0 & 0.9 \\
 			Gripper & 1.0 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Miconic & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Floortile & 0.2 & 0.2 & 0.8 & 0.5 & 1.0 & 0.9 \\
-			Pegsol & - & - & - & - & - & - \\
+			Miconic & 0.7 & 0.3 & 1.0 & 0.7 & 0.7 & 1.0 \\
+			Satellite & 0.7 & 0.5 & 1.0 & 0.8 & 1.0 & 0.5 \\
+			Transport & 1.0 & 0.3 & 0.5 & 0.8 & 1.0 & 0.6 \\
+			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 
 		\end{tabular}
 	\end{center}
-\caption{The {\em precision} and {\em recall} values achieved by the learned models in the different domains.}
+	\end{small}
+\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks.}
 \label{tab:results_plans}
@@
 
-Table~\ref{tab:time_plans} reports the planning time, and prepossessing time used to solve the classical planning task that results from our compilation as well as the plan length of the found solutions.
+Table~\ref{tab:time_plans} reports the planning time, and preprocessing time invested by the classical planner {\sc Madagascar} to solve the classical planning tasks that result from our compilation as well as the plan length of the found solutions.
 
 \begin{table}[hbt!]
+\begin{small}
 	\begin{center}
-		\begin{tabular}{l|l|l|l|}			 
+		\begin{tabular}{l|c|c|c|}			 
 			Domain & Total time & Preprocess & Plan length  \\
 			\hline
-			Blocks & 0.06 & 0.01 & 72  \\
-			Visitall & 0.43 & 0.40 & 37 \\
-			Gripper & 0.03 & 0.00 & 43  \\
-			Miconic & 0.12 & 0.09 & 59  \\
-			Floortile & 0.70 & 0.57 & 141 \\
-			Pegsol & - & - & - \\
+			Blocks & 0.04 & 0.00 & 72  \\
+			Driverlog & 0.10 & 0.06 & 127 \\
+			Ferry & 0.20 & 0.10 & 58 \\
+			Floortile & 2.15 & 1.21 & 170 \\
+			Gripper & 0.01 & 0.00 & 45 \\
+			Miconic & 0.10 & 0.09 & 54  \\
+			Satellite & 0.21 & 0.12 & 55 \\
+			Transport & 0.20 & 0.18 & 60 \\
+			Visitall & 0.43 & 0.39 & 37 			
 		\end{tabular}
 	\end{center}
-	\caption{Planning time, preprocessing time and Plan length.}
+        \end{small}
+	\caption{\small Planning time, preprocessing time and Plan length.}
 	\label{tab:time_plans}	
@@
 
+\subsection{Learning exploiting static predicates}
+Here we repeat the previous evaluation but compute now a candidates of static predicates from the inputs to the learning task and exploit this knowledge (as explained in Section~\ref{}) to reduce the space of possible action models.
+The set of considered candidate {\em static predicates} is the set of predicates whose instantiation appear the same in the initial and final states for all the learning examples {\small $1\leq t\leq \tau$}.
+Tables~\ref{tab:results_plans_static} and~\ref{tab:time_plans_static} show the obtained results.
+
+\begin{table}[hbt!]
+\begin{small}
+	\begin{center}
+		\begin{tabular}{l|l|l|l|l|l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			 
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+			\hline 
+		\end{tabular}
+	\end{center}
+	\end{small}
+\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks exploiting the computed static predicates.}
+\label{tab:results_plans_static}
+\end{table}
+
+\begin{table}[hbt!]
+\begin{small}
+	\begin{center}
+		\begin{tabular}{l|c|c|c|}			 
+			Domain & Total time & Preprocess & Plan length  \\
+			\hline 			
+		\end{tabular}
+	\end{center}
+        \end{small}
+	\caption{\small Planning time, preprocessing time and Plan length exploiting the computed static predicates.}
+	\label{tab:time_plans_static}	
+\end{table}
+
+\subsection{Learning from partially specified action models}
+
+Here we evaluate the ability of the compilation to support partially specified action models.
+In this case instead of learning the action models from scratch the model of half of the actions is given as input of the learning task.
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} show the obtained results.
+
+\begin{table}[hbt!]
+\begin{small}
+	\begin{center}
+		\begin{tabular}{l|l|l|l|l|l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			 
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+			\hline
+                        Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Driverlog & 1.0 & 0.4 & 1.0 & 1.0 & 1.0 & 0.7 \\
+			Ferry & 1.0 & 0.6 & 0.5 & 1.0 & 1.0 & 1.0 \\
+			Floortile & 0.5 & 0.6 & 1.0 & 0.8 & 1.0 & 0.8 \\
+			Gripper & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Miconic & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Satellite & 1.0 & 0.1 & 1.0 & 0.5 & - & 0.0 \\
+			Transport & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 0.5			
+		\end{tabular}
+	\end{center}
+	\end{small}
+\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the different domains.}
+\label{tab:results_plans_partial}
+\end{table}
+
+\begin{table}[hbt!]
+\begin{small}
+	\begin{center}
+		\begin{tabular}{l|c|c|c|}			 
+			Domain & Total time & Preprocess & Plan length  \\
+			\hline
+			Blocks & 0.07 & 0.01 & 54  \\
+			Driverlog & 0.06 & 0.03 & 71 \\
+			Ferry & 0.18 & 0.14 & 50 \\
+			Floortile & 0.81 & 0.68 & 92 \\
+			Gripper & 0.01 & 0.00 & 37 \\
+			Miconic & 0.06 & 0.04 & 40  \\
+			Satellite & 0.18 & 0.12 & 43 \\
+			Transport & 0.15 & 0.14 & 33 \\
+		\end{tabular}
+	\end{center}
+        \end{small}
+	\caption{\small Planning time, preprocessing time and Plan length.}
+	\label{tab:time_plans_partial}	
+\end{table}
+
+\subsection{Validating action models with classical planning}
+Here we evaluate the performance of the compilation to validate whether a given action mode is compliant with a given set of observations.
+Table~\ref{tab:time_plans_val} reports the planning time, and prepossessing time used to solve the classical planning task that results from our compilation as well as the plan length of the found solutions.
+\begin{table}[hbt!]
+\begin{small}
+	\begin{center}
+		\begin{tabular}{l|c|c|r|}			 
+			Domain & Total time & Preprocess & Plan length  \\
+			\hline
+	                Blocks & 0.00 & 0.00 \\
+			Driverlog & 0.04 & 0.04 \\
+			Ferry & 0.01 & 0.01 \\
+			Floortile & 0.03 & 0.03 \\
+			Gripper & 0.00 & 0.00 \\
+			Miconic & 0.01 & 0.01 \\
+			Satellite & 0.30 & 0.29 \\
+			Transport & 0.02 & 0.02 \\
+			Visitall & 0.07 & 0.07  \\ 			
+		\end{tabular}
+	\end{center}
+        \end{small}
+	\caption{\small Planning time, preprocessing time and Plan length for validating a given action model.}
+	\label{tab:time_plans_val}	
+\end{table}
+
 \subsection{Learning action models from example states}
-When the set of input planning task is not available, the planner determines also the actions that must satistying the input labels so actions can reformulated and still be compliant with the learning examples.
+When the set of input planning task is not available, the planner determines also the actions that must satisfying the input labels so actions can reformulated and still be compliant with the learning examples.
 For instance a blocksworld can be learned where the operator {\small\tt stack} is defined with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
+
+\subsubsection{Learning exploiting static predicates}
+\subsubsection{Learning from partially specified action models}
 
@@
 
-This research opens the door to the bootstrapping of action model.
-That is that a planner collects by its own the data that it is necessary to define an action model for a given planning task.
+As far as we know, this is the first work on learning action models using a classical planner, which opens the door to the bootstrapping of action model.
+That is that a planner collects by its own the data necessary learn an action model for a given planning task.
 With this regard, collecting {\em informative} examples for learning planning action models is challenging open issue.

### 2017-10-24 doc/icaps-18/strips-learning.tex
paper ICAPS18 updated with the last results by Diego

@@
 Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where neither actions nor intermediate states are given.
-What is more, partially specified action models can be used to feed the compilation and the compilation can also be used to validate whether certain observations of plan executions follow a given STRIPS action model.
+What is more, the compilation accepts partially specified action models can also be used to validate whether certain observations of plan executions follow a given STRIPS action model (even if the action model is not fully specified).
 \end{abstract}
@@
 The learning of STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation.
-The compilation approach is appealing because an off-the-shelf planner can be used to address the inductive learning task opening the door to the bootstrapping of action models in planning but also because:
+Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation so an off-the-shelf planner can be used to address the learning task.
+The compilation approach is appealing because it opens the door to the bootstrapping of planning action models but also because:
 \begin{enumerate}
@@
 The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate actions or states are observed.
-\item Can exploit previous knowledge about the structure of the actions in the form of partially specified action models.
+\item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
 In the extreme, the compilation can be used to validate whether observed plan executions are valid for a given STRIPS action model.
@@
 The paper is organized as follows.
-The first section presents the classical planning model, its extension to conditional effects (which is a requirement for the proposed compilation) and formalizes the STRIPS action model, the output of the addressed learning task.
+The first section presents the classical planning model, its extension to conditional effects (which is a requirement of the proposed compilation) and formalizes the STRIPS action model (the output of the addressed learning task).
 The second section formalizes the learning of STRIPS action models with regard to different amounts of input knowledge.
-The third section describes our approach for addressing this particular learning task by compiling it into classical planning and how previous knowledge can be introduced in the compilation.
-The fourth section reports the data collected during the empirical evaluation of our approach.
-Finally, the last section discusses the strengths and weaknesses of our approach and proposes several opportunities for future research.
+The third and foruth sections describe our approach for addressing this particular learning task by compiling it into classical planning and how previous knowledge can be introduced in the compilation to reduce the space of possible action models and making learning more practicable.
+Finally the last sections report the data collected during the empirical evaluation of our approach, discusses the strengths and weaknesses of our approach and proposes several opportunities for future research.
 
@@
 \subsection{Learning from initial/final states}
-Here we reduce the amount of provided input knowledge.
-First, $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is replaced by $\Pi'=\{|\pi_1|,\ldots,|\pi_{\tau}|\}$ i.e.
-$\Pi'$ that does not contain a set of plans but the number of actions of each plan.
-The learning task is hence redefined as $\Lambda'=\tup{\Psi,\Pi',\Sigma}$.
-While the previous learning task, $\Lambda$, corresponds to watching an agent acting in the world, this new learning task $\Lambda'$ can be understood as watching only the result of its plan execution but knowing the number of actions performed by the agent.
-In the case of the learning example shown in Figure~\ref{fig:lexample}, $\Pi'=\{|\pi_1|=8\}$.
+Here we reduce the amount of provided input knowledge and define the learning task $\Lambda'=\tup{\Psi,\Sigma}$ that corresponds to watching only the results of the plan executions.
+In this case no information about the executed plans is given.
+A solution to the $\Lambda'$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
 
-Finally, we go one step further and redefine a third learning task $\Lambda''=\tup{\Psi,\Sigma}$ that corresponds to watching only the results of the plan executions.
-In this case no information about the executed plans is given.
-A solution to the $\Lambda''$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
-
-In these two cases $\Lambda'$ and $\Lambda''$, a solution must not only determine a possible STRIPS action model but also the plans that explain the given labels using the learned model (this information about the actions is no longer given in the learning examples).
+In this case, a solution must not only determine a possible STRIPS action model but also the plans that explain the given labels using the learned model (this information about the actions is no longer given in the learning examples).
 We assume however, that the headers of the operators schema are known.
@@
 \section{Learning STRIPS action models with planning}
-Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is compiling it into a classical planning task with conditional effects $P_{\Lambda}$, $P_{\Lambda'}$ or $P_{\Lambda''}$.
+Our approach for addressing a learning task $\Lambda$ or $\Lambda'$, is compiling it into a classical planning task with conditional effects $P_{\Lambda}$ or $P_{\Lambda'}$.
 The intuition behind these compilations is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed STRIPS action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ of $\Pi'$ if available).
+\item Validates the programmed STRIPS action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ if available).
 For every label {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
 We call this process the validation of the programmed STRIPS action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
-If information about the corresponding plan $\pi_t$ is available, it is also used in the validation.
+If information about the corresponding plan $\pi_t\in \Pi$ is available, it is also used in the validation.
 \end{enumerate}
@@
 \begin{figure}[hbt]
-\begin{footnotesize}
+\begin{small}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{footnotesize}
+\end{small}
  \caption{\small Example of a plan for programing and validating the operator schema $unstack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
@@
 Now we are ready to formalize our compilations for learning STRIPS action models with classical planning.
-We start with $\Lambda''$ because it is the learning task that requires the smallest amount of input knowledge, and incrementally extend the formalized compilation until addressing $\Lambda$, the learning task with the largest amount of input knowledge.
-
-Given a learning task $\Lambda''=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda''}=\tup{F_{\Lambda''},A_{\Lambda''},I_{\Lambda''},G_{\Lambda''}}$ such that:
+We start with $\Lambda'$ because it is the learning task that requires the smallest amount of input knowledge.
+Given a learning task $\Lambda'=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ such that:
 \begin{itemize}
-\item $F_{\Lambda''}$ extends $F$ with:
+\item $F_{\Lambda'}$ extends $F$ with:
 \begin{itemize}
@@
 \end{itemize}
-\item $I_{\Lambda''}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema contains every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
-\item $G_{\Lambda''}=\{test_t\}$, {\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
-\item $A_{\Lambda''}$ contains actions of three different kinds:
+\item $I_{\Lambda'}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema contains every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
+\item $G_{\Lambda'}=\{test_t\}$, {\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
+\item $A_{\Lambda'}$ contains actions of three different kinds:
 \begin{enumerate}
@@
 \begin{lemma}
-Any classical plan $\pi$ that solves $P_{\Lambda''}$ induces an action model $\Xi$ that solves the learning task $\Lambda''$.
+Any classical plan $\pi$ that solves $P_{\Lambda'}$ induces an action model $\Xi$ that solves the learning task $\Lambda'$.
 \end{lemma}
-\begin{small}
+
 \begin{proof}[Proof sketch]
@@
 Furthermore, once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve the classical planning task $P_{\Lambda''}$, there is only one way of achieving goals $\{test_t\}$,{\small $1\leq t\leq \tau$}: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$ defined by the associated label $\sigma_t$, starting from $s_0^t$, defined in this same label.
-If this is achieved for all the input examples of the learning task $\Lambda''$ (i.e.
-for all the labels $\sigma_t$), it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda''$.
+To solve the classical planning task $P_{\Lambda'}$, there is only one way of achieving goals $\{test_t\}$,{\small $1\leq t\leq \tau$}: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$ defined by the associated label $\sigma_t$, starting from $s_0^t$, defined in this same label.
+If this is achieved for all the input examples of the learning task $\Lambda'$ (i.e.
+for all the labels $\sigma_t$), it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda'$.
  \end{proof}
-\end{small}
 
@@
 \section{Constraining the learning hypothesis space with additional input knowledge}
-Here we show that further input knowledge can be used to reduce the space of the possible action models and make the learning task more practicable.
+Here we show that further input knowledge can be used to reduce the space of the possible action models and make the learning of STRIPS action models more practicable.
 
 \subsection{Static predicates}
-A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action.
-Given a static predicate $p$, the actions $\mathsf{programEff_{f,\xi}}$ (s.t.
-$f\in F_v$ is an instantiation of the static predicate $p$ in $\Omega_V$) are unnecessary for every $\xi\in\Xi$ and hence, they can be removed from the compilation.
+A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action schema $\xi\in\Xi$.
+Therefore, one can get rid of the mechanism for programming static predicates as action effects.
+Formally, given a static predicate $p$:
+\begin{itemize}
+\item The fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in $\Omega_V$, can be removed from the compilation for every $\xi\in\Xi$.
+\item The actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of the static predicate $p$ in $\Omega_V$) can also be removed from the compilation.
+\end{itemize}
 
-Furthermore, actions $\mathsf{programPre_{f,\xi}}$ can also be discarded by checking the initial and final states given the set of labels $\Sigma$.
-If a precondition $f\in F_v$ instantiated from the static predicate $p$ in $\Omega_V$ is not possible for the schema $\xi\in\Xi$ according to a given example {\small $1\leq t\leq \tau$} it means that the corresponding action can be removed.
+Furthermore, certain fluents and actions can also be removed if a precondition $f\in F_v$ is not possible for the schema $\xi\in\Xi$ according to a given example {\small $1\leq t\leq \tau$}.
+For instance in the {\em zenotravel} domain fluents $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ that indicates that a precondition {\tt\small(next ?v1 ?v2 - flevel)} is programmed can be discarded (and their corresponding programmming actions) because there is no state in the labels $\Sigma$ that can satisfy them.
 
@@
 \subsection{Previously specified action models}
-Interestingly, the compilation does not require to start the learning of STRIPS action models from scratch and accepts partially specified operator schema where some preconditions are effects are a priori known.
+Our compilation does not require to start the learning of STRIPS action models from scratch and accepts partially specified operator schema where some preconditions are effects are a priori known.
 
-The known preconditions and effects can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true that are part of the initial state $I_{\Lambda''}$.
-Likewise, the corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) can be removed from $A_{\Lambda''}$ reducing the number of applicable actions  and making the classical planning task $P_{\Lambda''}$ easier to be solved.
+The known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true that are part of the initial state $I_{\Lambda'}$.
+The corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) are removed from $A_{\Lambda'}$ reducing the number of applicable actions  and making the classical planning task $P_{\Lambda'}$ easier to be solved.
 With this regard, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task for getting the blocksworld action model where the operator schema for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are known in advance.
@@
 In the extreme, the compilation can also be used to validate whether an observed plan follows a given STRIPS action model.
-In this case the action model to validate is coded in the initial state $I_{\Lambda''}$, any programmed action is removed from $A_{\Lambda''}$ and the $mode_{prog}$ fluent is set to false.
-If a solution plan is found to $P_{\Lambda''}$ it means that the given STRIPS action model is {\em valid} for the given examples.
-If $P_{\Lambda''}$ is unsolvable it means that the given STRIPS action model is invalid since it cannot satisfy the given examples.
+In this case the action model to validate is coded in the initial state $I_{\Lambda'}$, any programmed action is removed from $A_{\Lambda'}$ and the $mode_{prog}$ fluent is set to false.
+If a solution plan is found to $P_{\Lambda'}$ it means that the given STRIPS action model is {\em valid} for the given examples.
+If $P_{\Lambda'}$ is unsolvable it means that the given STRIPS action model is invalid since it cannot satisfy the given examples.
 Tools for plan validation like VAL~\cite{howey2004val} can also be used at this point but note that the compilation approach does not require a plan so the validation of an STRIPS action model can be done with just an initial and a final state.
@@
 \subsection{Example plans}
-Here  we extend our compilation to address the learning scenario defined by $\Lambda$ where, a set of plans $\Pi$ is available so it is used to reduce the space of possible action models.
-Given a learning task $\Lambda=\tup{\Psi,\Pi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}''$ as follows:
+Here  we extend our compilation to address the learning scenario defined by the $\Lambda$ task where, a set of plans $\Pi$ is available so it is used to reduce the space of possible action models.
+Given a learning task $\Lambda=\tup{\Psi,\Pi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}'$ as follows:
 \begin{itemize}
-\item $F_{\Lambda}$ extends $F_{\Lambda''}$ by including also the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of a plan in $\Pi$, with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding $\pi_t\in \Pi$.
+\item $F_{\Lambda}$ extends $F_{\Lambda'}$ by including also the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of a plan in $\Pi$, with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding $\pi_t\in \Pi$.
 We denote $\pi_t$ a solution plan to a classical planning instance $P_t$.
@@
 \item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi_1}$ that encode the plan $\pi_1\in \Pi$ for solving $P_1$, and the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
-Goals are the same as in the previous compilation $G_{\Lambda}=G_{\Lambda}''$ so again require that the programmed action model is validated in all  the learning examples.
+Goals are the same as in the previous compilation $G_{\Lambda}=G_{\Lambda}'$ so again require that the programmed action model is validated in all  the learning examples.
 \item With respect to $A_{\Lambda}$.
@@
 We evaluated our approach for learning STRIPS action models from different amounts of available input knowledge.
+
+\subsection{Setup}
 All the experiments run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
-The domains used in the evaluation are domains taken from different IPCs at the {\sc planning.domains} repository~\cite{muise2016planning} and satisfy the STRIPS requirement~\cite{fox2003pddl2}.
+The domains used in the evaluation are IPC domains that satisfy the STRIPS requirement~\cite{fox2003pddl2}, and are taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
 
-\subsection{The classical planner}
+\subsubsection{The classical planner.}
 The classical planner used to solve the instances that result from our compilations is the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar}.
 We decided to use a SAT-based planner because its ability to deal with planning instances populated with dead-ends.
-In addition, the benefit of SAT-based planners is that all the actions for programming the actions preconditions can be done in parallel, in a single planning step, because they do not interact.
-Likewise, all the actions for programming an action effect can also be applied in a single planning step reducing significantly the planning horizon.
+In addition, in a SAT-based classical planner all the actions for programming the STRIPS actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon.
+Likewise, all the actions for programming an action effect can also be applied in a single planning step.
 
-\subsection{Precision and recall}
-The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}.
+\subsubsection{The evaluation metrics.}
+The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
+These two metrics are frequently used in pattern recognition, information retrieval and binary classification and are more informative that simply counting the number of model errors or computing the {\em symmetric difference} between the learned model and the baseline ~\cite{davis2006relationship}.
 
-Formally $Precission=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
+Intuitively precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
+Formally $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
 On the other hand recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
-Intuitively precision gives a notion of {\em soundness} while and recall gives a notion of the {\em completeness} of the learned models.
+
+\subsubsection{The learning examples.}
 
@@
 			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 0.8 & 0.7 \\
+			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 1.0 & 0.8 \\
 			Ferry & 0.8 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
@@
 			Miconic & 0.7 & 0.3 & 1.0 & 0.7 & 0.7 & 1.0 \\
-			Satellite & 0.7 & 0.5 & 1.0 & 0.8 & 1.0 & 0.5 \\
+			Satellite & 0.6 & 0.2 & 1.0 & 1.0 & 1.0 & 0.75 \\
 			Transport & 1.0 & 0.3 & 0.5 & 0.8 & 1.0 & 0.6 \\
-			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 
+			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Zenotravel & 1.0 & 0.3 & 0.7 & 0.8 & 1.0 & 0.7
 		\end{tabular}
@@
 
-Table~\ref{tab:time_plans} reports the planning time, and preprocessing time invested by the classical planner {\sc Madagascar} to solve the classical planning tasks that result from our compilation as well as the plan length of the found solutions.
+Table~\ref{tab:time_plans} reports the total planning time, and the preprocessing time (in seconds) invested by the classical planner {\sc Madagascar} to solve the classical planning tasks that result from our compilation as well as the plan length of the found solutions.
 
@@
 			Miconic & 0.10 & 0.09 & 54 \\
-			Satellite & 0.21 & 0.12 & 55 \\
+			Satellite & 0.20 & 0.14 & 67 \\
 			Transport & 0.20 & 0.18 & 60 \\
-			Visitall & 0.43 & 0.39 & 37 			
+			Visitall & 0.43 & 0.39 & 37 \\
+			Zenotravel & 2.78 & 2.76 & 70			
 		\end{tabular}
@@
 
-\subsection{Learning exploiting static predicates}
+\subsection{Exploiting static predicates}
 Here we repeat the previous evaluation but compute now a candidates of static predicates from the inputs to the learning task and exploit this knowledge (as explained in Section~\ref{}) to reduce the space of possible action models.
@@
 In this case instead of learning the action models from scratch the model of half of the actions is given as input of the learning task.
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} show the obtained results.
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarizes the obtained results.
 
@@
 			Miconic & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Satellite & 1.0 & 0.1 & 1.0 & 0.5 & - & 0.0 \\
-			Transport & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 0.5			
+			Satellite & 1.0 & 0.5 & 0.5 & 1.0 & 1.0 & 1.0 \\
+			Transport & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 0.5 \\
+			Zenotravel & 1.0 & 0.5 & 0.6 & 0.6 & 1.0 & 1.0                         
 		\end{tabular}
@@
 			Transport & 0.15 & 0.14 & 33 \\
+			Zenotravel & 2.54 & 2.54 & 42                         
 		\end{tabular}
@@
 			Visitall & 0.07 & 0.07  \\
+			Zenotravel & 11.65 & 11.39 			
 		\end{tabular}

### 2017-10-27 doc/icaps-18/strips-learning.tex
small updates and starting with Related work

@@
 Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where neither actions nor intermediate states are given.
-What is more, the compilation accepts partially specified action models can also be used to validate whether certain observations of plan executions follow a given STRIPS action model (even if the action model is not fully specified).
+What is more, the compilation accepts partially specified action models and can also be used to validate whether certain observations of plan executions follow a given STRIPS action model (even if the action model is not fully specified).
 \end{abstract}
@@
 The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
-\item The traditional output of ML algorithms is a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
+\item The traditional output of ML algorithms is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
 When learning STRIPS action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions of the corresponding planning domain.
@@
 The learning of STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation so an off-the-shelf planner can be used to address the learning task.
+Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation, so an off-the-shelf planner can address the learning task.
 The compilation approach is appealing because it opens the door to the bootstrapping of planning action models but also because:
@@
 \item Is flexible to different amounts of available input knowledge.
-The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate actions or states are observed.
+The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate state  or action is observed.
 \item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
-In the extreme, the compilation can be used to validate whether observed plan executions are valid for a given STRIPS action model.
+In the extreme, the compilation could be used to validate whether observed plan executions are valid for a given STRIPS action model.
 \end{enumerate}
@@
 The first section presents the classical planning model, its extension to conditional effects (which is a requirement of the proposed compilation) and formalizes the STRIPS action model (the output of the addressed learning task).
-The second section formalizes the learning of STRIPS action models with regard to different amounts of input knowledge.
-The third and foruth sections describe our approach for addressing this particular learning task by compiling it into classical planning and how previous knowledge can be introduced in the compilation to reduce the space of possible action models and making learning more practicable.
-Finally the last sections report the data collected during the empirical evaluation of our approach, discusses the strengths and weaknesses of our approach and proposes several opportunities for future research.
+The second section formalizes the learning of STRIPS action models with regard to different amounts of available input knowledge.
+The third and fourth sections describe our approach for addressing this particular learning task by compiling it into classical planning and how previous knowledge can be introduced in the compilation to constrain the space of possible action models and make learning more practicable.
+Finally, the last sections report the data collected during the empirical evaluation of our approach, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
 
@@
 
-A {\em state} $s$ is then a total assignment of values to fluents, i.e.
+A {\em state} $s$ is a total assignment of values to fluents, i.e.
 $|s|=|F|$, so the size of the state space is $2^{|F|}$.
@@
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of effects $\eff(a)\in\mathcal{L}(F)$.
-An action $a\in A$ is {\em applicable} in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is the {\em succesor} state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
+Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of {\em effects} $\eff(a)\in\mathcal{L}(F)$.
+An action $a\in A$ is {\em applicable} in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is the {\em successor} state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
 
@@
 
-Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$ called the {\em precondition} and a set of conditional effects $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$ (the condition) and $E\in\mathcal{L}(F)$ (the effect).
+Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called the {\em precondition}, and a set of conditional effects $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the condition, and $E\in\mathcal{L}(F)$, the effect.
 
@@
 \]
-i.e.~effects whose conditions hold in $s$. The result of applying $a$ in $s$ is the {\em succesor} state $\theta(s,a)=(s\setminus \neg \eff(s,a))\cup\eff(s,a)$.
+i.e.~effects whose conditions hold in $s$. The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=(s\setminus \neg \eff(s,a))\cup\eff(s,a)$.
 
-\subsection{The STRIPS action model}
+\subsection{The STRIPS action model and the {\em variable} objects}
 This work addresses the learning of action schemes defined in the STRIPS language~\cite{fikes1971strips}.
@@
 
-Likewise, we assume that actions are instantiated from STRIPS operator schemes.
+Likewise, we assume that STRIPS actions are instantiated from STRIPS operator schemes.
 Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld~\cite{slaney2001blocks} represented in PDDL.
-\begin{figure}[hbt]
+\begin{figure}
 \begin{footnotesize}
@@
                (not (clear ?x2))
+               (handempty)
                (clear ?x1)  
-               (handempty) 
                (on ?x1 ?x2)))
@@
 Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names} and whose cardinality, $|\Omega_v|$, is given by the maximum arity of an action in a given planning frame $\Phi$.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because operators {\small\tt stack} and {\small\tt unstack}, the ones with the maximum arity, have two parameters.
+For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because operators {\small\tt stack} and {\small\tt unstack}, with the maximum arity, have two parameters each.
 
@@
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its corresponding action name and a enumeration of the variable names whose cardinality is the operator arity, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and a enumeration of the variable objects bound by the operator arity, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
 The headers for a 4-operator blocksworld are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
@@
 \item $\Psi$, the set of predicates that define the abstract state space of a given planning domain. 
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans s.t.
-each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
 \item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$, is a set of labels s.t.
@@
 
-Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda$ for computing the STRIPS action model of the blocksworld.
-The figure shows the content of $\Psi$, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$.
-This learning task has a single learning example, that corresponds to observing the execution of a plan for inverting a tower of four blocks.
+Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda$ in the blocksworld (the figure shows the content of $\Psi$, $\Pi$ and $\Sigma$).
+This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of a plan for inverting a four-block tower.
 
-\begin{figure}[hbt]
+\begin{figure}
 \begin{small}
@@
 
- \caption{\small Example of a task for learning the STRIPS action model of the blocksworld from a single learning example.}
+ \caption{\small Example of a task for learning a STRIPS action model in the blocksworld.}
 \label{fig:lexample}
@@
 \section{Learning STRIPS action models with planning}
-Our approach for addressing a learning task $\Lambda$ or $\Lambda'$, is compiling it into a classical planning task with conditional effects $P_{\Lambda}$ or $P_{\Lambda'}$.
+Our approach for addressing a learning task $\Lambda$ or $\Lambda'$, is compiling it into a classical planning task with conditional effects.
 The intuition behind these compilations is that a solution to the resulting classical planning task is a sequence of actions that:
 \begin{enumerate}
-\item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan has a prefix that determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
 \item Validates the programmed STRIPS action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ if available).
@@
 We call this process the validation of the programmed STRIPS action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
-If information about the corresponding plan $\pi_t\in \Pi$ is available, it is also used in the validation.
+If information about the corresponding plan $\pi_t\in \Pi$ is available, is used in the validation.
 \end{enumerate}
@@
 Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ kind.
-This classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} (previously specified operator schemes for $pickup$, $putdown$ and $unstack$ are given).
-In particular the steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are the actions for validating the programmed operator.
+This classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld (previously specified operator schemes for $pickup$, $putdown$ and $unstack$ are given), using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+The plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operators.
 The details of the compilation are explained in the remaining of the section.
@@
 \end{small}
- \caption{\small Example of a plan for programing and validating the operator schema $unstack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+ \caption{\small Example of a plan for programming and validating the operator schema $unstack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
 \label{fig:plan}
@@
 same fluents and actions and differ only in the initial state and goals).
-The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the set of different objects appearing in the input learning examples.
+The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the objects appearing in the input learning examples.
 Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
@@
 Now we are ready to formalize our compilations for learning STRIPS action models with classical planning.
-We start with $\Lambda'$ because it is the learning task that requires the smallest amount of input knowledge.
+We start with $\Lambda'$ because this learning task requires smaller input knowledge.
 Given a learning task $\Lambda'=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ such that:
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the STRIPS operator schema $\xi\in \Xi$.
-\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or they are already programmed operators that are being validated. 
+\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or they are being validated (already programmed). 
 \item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
 \end{itemize}
-\item $I_{\Lambda'}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema contains every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
-\item $G_{\Lambda'}=\{test_t\}$, {\small $1\leq t\leq \tau$}, indicates that the programmed action model is validated in all the learning examples.
+\item $I_{\Lambda'}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
+\item $G_{\Lambda'}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
 \item $A_{\Lambda'}$ contains actions of three different kinds:
 \begin{enumerate}
-\item The actions for programming the an operator schema $\xi\in\Xi$ which includes:
+\item The actions for programming an operator schema $\xi\in\Xi$, which includes:
 \begin{itemize}
@@
 \begin{proof}[Proof sketch]
+\begin{small}
 The compilation forces that once the preconditions of an operator schema $\Xi$ are programmed, they cannot be altered.
@@
 for all the labels $\sigma_t$), it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda'$.
+\end{small}
 \end{proof}
+
+The compilation is {\em complete} in the sense that it does not discard any possible STRIPS action model.
 
@@
 \section{Constraining the learning hypothesis space with additional input knowledge}
-Here we show that further input knowledge can be used to reduce the space of the possible action models and make the learning of STRIPS action models more practicable.
-
-\subsection{Static predicates}
-A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action schema $\xi\in\Xi$.
-Therefore, one can get rid of the mechanism for programming static predicates as action effects.
-Formally, given a static predicate $p$:
-\begin{itemize}
-\item The fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in $\Omega_V$, can be removed from the compilation for every $\xi\in\Xi$.
-\item The actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of the static predicate $p$ in $\Omega_V$) can also be removed from the compilation.
-\end{itemize}
-
-Furthermore, certain fluents and actions can also be removed if a precondition $f\in F_v$ is not possible for the schema $\xi\in\Xi$ according to a given example {\small $1\leq t\leq \tau$}.
-For instance in the {\em zenotravel} domain fluents $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ that indicates that a precondition {\tt\small(next ?v1 ?v2 - flevel)} is programmed can be discarded (and their corresponding programmming actions) because there is no state in the labels $\Sigma$ that can satisfy them.
-
-In our evaluation section we shown how to compute candidates of static predicates from the inputs to our learning tasks and how using this knowledge affects to the performance of our approach.
-
-\subsection{Previously specified action models}
-Our compilation does not require to start the learning of STRIPS action models from scratch and accepts partially specified operator schema where some preconditions are effects are a priori known.
-
-The known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true that are part of the initial state $I_{\Lambda'}$.
-The corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) are removed from $A_{\Lambda'}$ reducing the number of applicable actions  and making the classical planning task $P_{\Lambda'}$ easier to be solved.
-With this regard, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task for getting the blocksworld action model where the operator schema for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are known in advance.
-
-In the extreme, the compilation can also be used to validate whether an observed plan follows a given STRIPS action model.
-In this case the action model to validate is coded in the initial state $I_{\Lambda'}$, any programmed action is removed from $A_{\Lambda'}$ and the $mode_{prog}$ fluent is set to false.
-If a solution plan is found to $P_{\Lambda'}$ it means that the given STRIPS action model is {\em valid} for the given examples.
-If $P_{\Lambda'}$ is unsolvable it means that the given STRIPS action model is invalid since it cannot satisfy the given examples.
-Tools for plan validation like VAL~\cite{howey2004val} can also be used at this point but note that the compilation approach does not require a plan so the validation of an STRIPS action model can be done with just an initial and a final state.
+Here we show that further input knowledge can be used to constrain the space of the possible action models and make the learning of STRIPS action models more practicable.
 
 \subsection{Example plans}
-Here  we extend our compilation to address the learning scenario defined by the $\Lambda$ task where, a set of plans $\Pi$ is available so it is used to reduce the space of possible action models.
+Here  we extend our compilation to address the learning scenario where, a set of plans is available.
 Given a learning task $\Lambda=\tup{\Psi,\Pi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}'$ as follows:
 \begin{itemize}
-\item $F_{\Lambda}$ extends $F_{\Lambda'}$ by including also the fluents $F_{\Pi}=\{plan(name(\xi),j,\Omega^{ar(\xi)})\}$ to code the $j$ steps of a plan in $\Pi$, with $F_{\Pi_t}\subseteq F_{\Pi}$ encoding $\pi_t\in \Pi$.
-We denote $\pi_t$ a solution plan to a classical planning instance $P_t$.
-In addition fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the plan steps where the programmed action model is validated.
-\item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi_1}$ that encode the plan $\pi_1\in \Pi$ for solving $P_1$, and the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
-Goals are the same as in the previous compilation $G_{\Lambda}=G_{\Lambda}'$ so again require that the programmed action model is validated in all  the learning examples.
+\item $F_{\Lambda}$ extends $F_{\Lambda'}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the $j$ steps of plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
+Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current and next plan steps.
+\item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi}$ that encode the plan $\pi_1\in \Pi$, the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
+Goals are the same as in the previous compilation $G_{\Lambda}=G_{\Lambda}'=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
 \item With respect to $A_{\Lambda}$.
@@
 \item The actions for programming the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for applying an operator have an extra precondition $f\in F_{\Pi_t}$ that encodes the current plan step and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing the plan step.
-\item The actions for updating the active test have an extra precondition, $at_{|\Pi_t|}$, to indicate that the current plan $\Pi_t$ was fully executed and extra conditional effects to load the next plan $\Pi_{t+1}$:
+\item The actions for applying an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+This mechanism forces that these actions are only applied as in the example plans.
+\item The actions for updating the active test have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to load the next plan $\pi_{t+1}$:
 \begin{small}
 \begin{align*}
-&\{f\}\rhd\{\neg f\}_{f\in F_{\Pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\Pi_t+1}},\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
+&\{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}},\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
 \end{align*}
@@
 
+\subsection{Previously specified action models}
+Our compilation does not require to start learning STRIPS action models from scratch and accepts partially specified operator schema where some preconditions and effects are a priori known.
+
+The known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that are part of the initial state $I_{\Lambda'}$.
+The corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
+To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task for getting the blocksworld action model where the schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are known in advance.
+
+In the extreme, when a fully specified STRIPS action model is given, the compilation validates whether an observed plan follows the given model.
+In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given STRIPS action model is {\em valid} for the given set of examples.
+If $P_{\Lambda'}$ is unsolvable it means that the given STRIPS action model is invalid since it is not compliant with all the given examples.
+Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
+
+\subsection{Static predicates}
+A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action schema~\cite{fox:TIM:JAIR1998}.
+Therefore, one can get rid of the mechanism for programming static predicates as action effects keeping the compilation complete.
+Formally, given a static predicate $p$:
+\begin{itemize}
+\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable} objects $\Omega_v$, can be discarded for every $\xi\in\Xi$.
+\item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
+\end{itemize}
+
+Static predicates can also be used to constrain the space of possible preconditions by looking at the given labels.
+If a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Sigma$, then the fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
+For instance in the {\em zenotravel} domain fluents $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (as well as their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} cannot be compliant with any state in $\Sigma$.
+
+Likewise, static predicates can constrain the space of possible preconditions looking at the given example plans.
+In this case, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discarded for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
+Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static fluent {\tt\small (next fl2 fl3)} but it does not include {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} it means that the only possible precondition including the static predicate is $pre\_next\_fly\_v1\_v2$.
+
+In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of candidates of static predicates from the input examples to the learning task.
+
 \section{Evaluation}
@@
 
-\subsection{Setup}
+\subsection{Experimental setup}
 All the experiments run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
@@
 \subsection{Exploiting static predicates}
-Here we repeat the previous evaluation but compute now a candidates of static predicates from the inputs to the learning task and exploit this knowledge (as explained in Section~\ref{}) to reduce the space of possible action models.
+Here we repeat the previous evaluation but compute now a candidates of static predicates from the inputs to the learning task and exploit this knowledge to reduce the space of possible action models.
 The set of considered candidate {\em static predicates} is the set of predicates whose instantiation appear the same in the initial and final states for all the learning examples {\small $1\leq t\leq \tau$}.
@@
 For each action in a given domain, they use kernel perceptrons to learn predictions of the domain properties that change because of the action execution.
-LOCM (Cresswellet al., 2009) induces action schemas without being provided with any information about initial, goal or intermediate state descriptions for the example action sequences.
-LOCM receives descriptions of plans or plan fragments, uses them to create states machines for the different domain objects and extracts the action schemas from these state machines.
+
+In the {\sf LOCM} approach for learning planning models \cite{cresswell2013acquiring}, the only required input are the example training plans, without the need for providing the initial and goal state information.
+{\sf LOCM} exploits assumptions about the kind of domain model it has to generate.
+Particularly, it assumes a domain consists of a collection of objects (called \emph{sorts}), where each object has a defined set of states that can be captured by a parameterized FSM.
+The induction of FSM draws upon assumptions like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence.
+The intuitive assumptions of {\sf LOCM} yield a learning model heavily reliant on the kind of domain structure and not capable of properly deriving domain theories in which the state of a sort is subject to different state machines.
+This limitation is later overcome by {\sf LOCM2} by forming separate state machines for a sort, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
 
@@
 The paper presented a novel approach for learning STRIPS action models from examples exclusively using classical planning.
-Interestingly the approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
+The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
 
-The empirical evaluation shows that, when example plans are available, our approach can compute accurate action models.
+The empirical evaluation shows that, when example plans are available, our approach can compute accurate action models from small sets of learning examples and investing very small learning times (less than a second time in most of the domains).
 When action plans are not available our approach is still able to produce action models compliant with the input information.
 In this case, since learning is not constrained by actions it can change the semantics of the operators.
-An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploitin features that allow more compact solutions like the {\em reachable} or {\em movable} features in the Sokoban domain.
+An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
 
-With regard to efficiency, the size of the compiled classical planning instances depend on the number of input examples.
+With regard to scalability of the compilation approach, the size of the compiled classical planning instances depends on the number of input examples.
 On the other hand the empirical results show that our approach, since is based on inferences instead of statistics, is able to generate non-trivial models from very small data sets.
+In addition the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because all the actions for programming the STRIPS actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon.
+The same happens with all the actions for programming an action effect can also be applied in a single planning step.
 
-As far as we know, this is the first work on learning action models using a classical planner, which opens the door to the bootstrapping of action model.
-That is that a planner collects by its own the data necessary learn an action model for a given planning task.
-With this regard, collecting {\em informative} examples for learning planning action models is challenging open issue.
+As far as we know, this is the first work on learning action models using a classical planner.
+This novel approach opens the door to the bootstrapping of action model that is, a planner collecting by its own the necessary data for learning an action model for a given planning task.
+Generating {\em informative} examples for learning planning action models is still a challenging open issue.
 Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-The recent success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel approaches that do not assume that a learning set of plans is given apriori but instead, autonomously collect the learning examples.
+The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques that autonomously collect the learning examples.
 
+\begin{small}
 \subsection*{Acknowledgment}
-\begin{small}
 Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
@@
 
-\newpage
-
 \bibliographystyle{aaai}

### 2017-10-27 doc/icaps-18/strips-learning.tex
validation removed from evaluation of ICAPS18 paper

@@
 \begin{table}[hbt!]
-	\begin{small}
+	\begin{footnotesize}
 	\begin{center}
@@
 	\end{center}
-	\end{small}
+	\end{footnotesize}
 \caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks.}
@@
 \begin{table}[hbt!]
-\begin{small}
+\begin{footnotesize}
 	\begin{center}
@@
 	\end{center}
-        \end{small}
+        \end{footnotesize}
 	\caption{\small Planning time, preprocessing time and Plan length.}
@@
 			\hline
+			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 1.0 & 0.8 \\
+			Ferry & 0.8 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Floortile & 0.5 & 0.6 & 0.9 & 0.8 & 1.0 & 0.9 \\
+			Gripper & 1.0 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Miconic & 0.7 & 0.3 & 1.0 & 0.7 & 0.7 & 1.0 \\
+			Satellite & 0.6 & 0.2 & 1.0 & 1.0 & 1.0 & 0.75 \\
+			Transport & 1.0 & 0.3 & 0.5 & 0.8 & 1.0 & 0.6 \\
+			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Zenotravel & 1.0 & 0.3 & 0.7 & 0.8 & 1.0 & 0.7                        
 		\end{tabular}
 	\end{center}
-	\end{small}
+	\end{footnotesize}
 \caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks exploiting the computed static predicates.}
@@
 			\hline
+			Blocks & 0.04 & 0.00 & 72 \\
+			Driverlog & 0.10 & 0.06 & 127 \\
+			Ferry & 0.20 & 0.10 & 58 \\
+			Floortile & 2.15 & 1.21 & 170 \\
+			Gripper & 0.01 & 0.00 & 45 \\
+			Miconic & 0.10 & 0.09 & 54 \\
+			Satellite & 0.20 & 0.14 & 67 \\
+			Transport & 0.20 & 0.18 & 60 \\
+			Visitall & 0.43 & 0.39 & 37 \\
+			Zenotravel & 2.78 & 2.76 & 70                        
 		\end{tabular}
 	\end{center}
-        \end{small}
+        \end{footnotesize}
 	\caption{\small Planning time, preprocessing time and Plan length exploiting the computed static predicates.}
@@
 	\end{center}
-	\end{small}
+	\end{footnotesize}
 \caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the different domains.}
@@
 	\end{center}
-        \end{small}
+        \end{footnotesize}
 	\caption{\small Planning time, preprocessing time and Plan length.}
 	\label{tab:time_plans_partial}	
-\end{table}
-
-\subsection{Validating action models with classical planning}
-Here we evaluate the performance of the compilation to validate whether a given action mode is compliant with a given set of observations.
-Table~\ref{tab:time_plans_val} reports the planning time, and prepossessing time used to solve the classical planning task that results from our compilation as well as the plan length of the found solutions.
-\begin{table}[hbt!]
-\begin{small}
-	\begin{center}
-		\begin{tabular}{l|c|c|r|}			 
-			Domain & Total time & Preprocess & Plan length  \\
-			\hline
-			Blocks & 0.00 & 0.00 \\
-			Driverlog & 0.04 & 0.04 \\
-			Ferry & 0.01 & 0.01 \\
-			Floortile & 0.03 & 0.03 \\
-			Gripper & 0.00 & 0.00 \\
-			Miconic & 0.01 & 0.01 \\
-			Satellite & 0.30 & 0.29 \\
-			Transport & 0.02 & 0.02 \\
-			Visitall & 0.07 & 0.07  \\
-			Zenotravel & 11.65 & 11.39 			
-		\end{tabular}
-	\end{center}
-        \end{small}
-	\caption{\small Planning time, preprocessing time and Plan length for validating a given action model.}
-	\label{tab:time_plans_val}	
 \end{table}
@@
 For instance a blocksworld can be learned where the operator {\small\tt stack} is defined with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
-
-\subsubsection{Learning exploiting static predicates}
-\subsubsection{Learning from partially specified action models}
 
@@
 
-With regard to scalability of the compilation approach, the size of the compiled classical planning instances depends on the number of input examples.
-On the other hand the empirical results show that our approach, since is based on inferences instead of statistics, is able to generate non-trivial models from very small data sets.
+The size of the compiled classical planning instances depends on the number of input examples.
+The empirical results show that our approach can generate non-trivial models from very small data sets.
 In addition the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because all the actions for programming the STRIPS actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon.

### 2017-10-28 doc/icaps-18/strips-learning.tex
Eva's changes

@@
 {\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-{\small Universitat Polit\'ecnica de Val\'encia.}\\
+{\small Universitat Polit\`ecnica de Val\`encia.}\\
 {\small Camino de Vera s/n. 46022 Valencia, Spain}\\
@@
 
-\section{Related work}
+\section{Past and future of learning action models}
 
-The  LIVE  system  (Shen  and  Simon,  1989)  was  an  extension  of  the  General  Problem  Solver  (GPS) framework (Ernst and Newell, 1969) with a learning component.
-LIVE alternated problem solving with model  learning  to  automatically  define  operators.
-The  decision  about  when  to  alternate  depended  onsurprises,  that  is  situations  where  an  action  effects  violated  its  predicted  model.
-EXPO  (Gil,  1992) generated plans with the  PRODIGY  system (Minton, 1988), monitored the plans execution, detected differences in the predicted and the observed states and constructed a set of specific hypotheses to fix those differences.
-Then the  EXPO  filtered the hypotheses heuristically.
-OBSERVER  (Wang, 1994) learned operators by monitoring expert agents and applying the version spaces algorithm (Mitchell, 1997) to the observations.
-When the system already had an operator representation, the preconditions were updated by removing facts that were not present in the new observation’s pre-state; the effects were augmented by adding facts that were in the observation’s delta-state.
-All of these early works were based on direct liftings of the observed states.
-They also benefit from experience beyond simple interaction with the environment such as exploratory plans or external teachers, but  none  provided  a  theoretical  justification  for  this  second  source  of  knowledge.
-The  work  recently reported in (Walsh and Littman, 2008) succeeds in bounding the number of interactions the learner must complete to learn the preconditions and effects of a STRIPS action model.
-This work shows that learning STRIPS  operators  from  pure  interaction  with  the  environment,  can  require  an  exponential  number  of samples, but that limiting the size of the precondition lists enable sample-efficient learning (polynomial in the number of actions and predicates of the domain).
-The work also proves that efficient learning is also possible without this limit if an agent has access to an external teacher that can provide solution traces on demand.
+Back in the 90's various systems were aimed at learning operators mostly via interaction with the environment.
+The proposal of the LIVE system was to capture and formulate observable features of objects and use them to acquire and refine operators \cite{ShenS89}.
+OBSERVER, on the other hand, updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
+These early works were based on direct lifting of the observed states and supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge.
+The recent work reported in \cite{WalshL08} shows that although learning STRIPS operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
 
-Others systems have tried to learn more expressive action models for deterministic planning in fully observable environments.
-Examples would include the learning of conditional costs for AP actions (Jess Lanchas and Borrajo, 2007) or the learning of conditional effects with quantifiers (Zhuo et al., 2008).
-
-In addition action model learning has been studied in domains where there is partial state observability.
-ARMS uses the same kind od learning examples but assumes the exampls are given and proceeds in two phases.
-In the first phase, ARMS extracts frequent action sets from plans that share a common set of parameters.
-ARMS also finds some frequent literal-action pairs with the help of the initial state and the goal state that provide an initial guess on the actions preconditions, and effects.
-In the second phase, ARMS uses the frequent action sets and literal-action pairs to define a set of weighted constraints that must hold in order to make the plans correct.
-Then, ARMS solves the resulting weighted MAX-SAT problem and produces action models from the solution of the SAT problem.
-This process iterates until all actions are modelled.
-For a complex planning domain that involves hundreds of literals and actions, the corresponding weighted MAX-SAT representation is likely to be too large to be solved efficiently as the number of clauses can reach up to tens of thousands.
-For that reason ARMS implements a hill-climbing method that models the actions approximately.
-Consequently, the ARMS output is a model which may be inconsistent with the examples.
-
-(Amir and Chang, 2008) introduced an algorithm that tractably generates all the STRIPS-like models that  could  have  lead  to  a  set  of  observations.
-Given  a  formula  representing  the  initial  belief  state,  a sequence of executed actions and the corresponding observed states(where partial observations of states are given), it builds a complete explanation of observations by models of actions through a Conjunctive Normal Form (CNF) formula.
-By linking the possible states of fluents to the effect propositions in the action models, the complexity of the CNF encoding can be controlled to find  exact  solutions  efficiently  in  some  circumstances.
+Whilst the aforementioned works deal with full state observability, action model learning has also been studied in domains
+where there is partial or missing state observability~\cite{jimenez2012review}.
+The {\sf ARMS} approach~\cite{yang2007learning}, which works when no or partial intermediate states are given, extracts frequent action sets from plans that share a common set of parameters and frequent literal-action pairs with the help of the initial and goal state of the plans.
+The extracted information is used then to define a set of weighted constraints that must hold for all the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver.
+Since complex planning problems may result in a too large MAX-SAT representation to be solved efficiently, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
+Consequently, the output model of {\sf ARMS} may be inconsistent with the examples.
+The approach presented in \cite{amir:alearning:JAIR08} deals with partial observability.
+Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation
+of observations by models of actions through a CNF formula.
 The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
@@
 The formula returned at the end includes all consistent models, which can then be retrieved with additional processing.
-Unlike the previous approaches, the one described in (Mour ao et al., 2008) deals with both missing and noisy predicates in the observations.
-For each action in a given domain, they use kernel perceptrons to learn predictions of the domain properties that change because of the action execution.
+Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
+An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then STRIPS rules are derived from the classifiers' parameters.
 
-In the {\sf LOCM} approach for learning planning models \cite{cresswell2013acquiring}, the only required input are the example training plans, without the need for providing the initial and goal state information.
+The approach for learning planing models that most likely works with the least information possible is {\sf LOCM} ~\cite{cresswell2013acquiring}, where the only required input are the example training plans without the need for providing the initial and goal state information.
 {\sf LOCM} exploits assumptions about the kind of domain model it has to generate.
-Particularly, it assumes a domain consists of a collection of objects (called \emph{sorts}), where each object has a defined set of states that can be captured by a parameterized FSM.
+Particularly, it assumes a domain consists of a collection of objects (called \emph{sorts}) where each object has a defined set of states that can be captured by a parameterized Finite State Machine (FSM).
 The induction of FSM draws upon assumptions like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence.
@@
 This limitation is later overcome by {\sf LOCM2} by forming separate state machines for a sort, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
+The last contribution of the {\sf LOCM} family, {\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{GregoryC16}), addresses the problem of inducing static predicates in action models by identifying the set of minimal static relations associated to an operator.
+Because {\sf LOCM} models induce similar models for domains with similar structure (e.g.
+Blocksworld and Freecell), they face problems at identifying models that do not have static relations (Blocksworld) and those that do have static predicates (Freecell).
+In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
+
+==============================================
 

### 2017-10-28 src/compiler2.py
compiler2.py: support for static predicates with different number of parameters

@@ -76,31 +76,34 @@ def get_static_predicates(tasks, predicates):
     return [p for p in predicates if p[0] in candidates]
 
 def get_static_precondition(predicate, action, plans, tasks):
-    static_preconditions = list()
+    static_preconditions = set()
     params = [pddl.pddl_types.TypedObject("?o" + str(i), action[i]) for i in range(1, len(action))]
     params = [x for x in params if x.type_name in predicate[1:]]
+    num_predicate_params = len(predicate[1:])
 
-    if len([x for x in action[1:] if x in predicate[1:]]) == len(predicate[1:]):
+    if len([x for x in action[1:] if x in predicate[1:]]) >= num_predicate_params:
         all_instances = set()
         for task in tasks:
             all_instances.update([p.args for p in task.init if p.predicate == predicate[0]])
 
         all_variables = set(sum(all_instances, ()))
+
         for a in [item for sublist in plans for item in sublist]:
             a = a.replace('(','').replace(')','').split(" ")
             if a[0] == action[0]:
                 variables = [x for x in a[1:] if x in all_variables]
-                possible_tuples = [tuple(variables)]
+                possible_tuples = list(itertools.combinations(variables, num_predicate_params))
+                possible_param_tuples = list(itertools.combinations(params, num_predicate_params))
 
-                for possible_tuple in possible_tuples:
-                    if possible_tuple in all_instances:
-                        static_preconditions.append(pddl.conditions.Atom(predicate[0], [params[0].name, params[1].name]))
-                    elif tuple(reversed(possible_tuple)) in all_instances:
-                        static_preconditions.append(pddl.conditions.Atom(predicate[0], [params[1].name, params[0].name]))
+                for i in range(len(possible_tuples)):
+                    if possible_tuples[i] in all_instances:
+                        static_preconditions.add(pddl.conditions.Atom(predicate[0], [x.name for x in possible_param_tuples[i]]))
+                    elif tuple(reversed(possible_tuples[i])) in all_instances:
+                        static_preconditions.add(pddl.conditions.Atom(predicate[0], [x.name for x in reversed(possible_param_tuples[i])]))
                 break
 
 
-    return static_preconditions
+    return list(static_preconditions)
[…]

### 2017-10-28 doc/icaps-18/strips-learning.tex
Added static predicates results to the paper

@@
 			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 1.0 & 0.8 \\
-			Ferry & 0.8 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Floortile & 0.5 & 0.6 & 0.9 & 0.8 & 1.0 & 0.9 \\
-			Gripper & 1.0 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Miconic & 0.7 & 0.3 & 1.0 & 0.7 & 0.7 & 1.0 \\
-			Satellite & 0.6 & 0.2 & 1.0 & 1.0 & 1.0 & 0.75 \\
-			Transport & 1.0 & 0.3 & 0.5 & 0.8 & 1.0 & 0.6 \\
-			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Zenotravel & 1.0 & 0.3 & 0.7 & 0.8 & 1.0 & 0.7
+			Driverlog & 1.0 & 0.5 & 0.8 & 0.9 & 1.0 & 0.7 \\
+			Ferry & 0.8 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Floortile & 0.7 & 0.8 & 0.9 & 0.8 & 1.0 & 0.9 \\
+			Gripper & 1.0 & 0.7 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Miconic & 0.9 & 0.8 & 1.0 & 0.8 & 0.8 & 1.0 \\
+			Satellite & 0.8 & 0.8 & 0.8 & 1.0 & 1.0 & 0.8 \\
+			Transport & 1.0 & 0.6 & 0.8 & 0.8 & 1.0 & 0.6 \\
+			Visitall & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Zenotravel & 1.0 & 0.6 & 0.8 & 0.9 & 1.0 & 0.7
 		\end{tabular}
@@
 			\hline
-			Blocks & 0.04 & 0.00 & 72 \\
-			Driverlog & 0.10 & 0.06 & 127 \\
-			Ferry & 0.20 & 0.10 & 58 \\
-			Floortile & 2.15 & 1.21 & 170 \\
+			Blocks & 0.03 & 0.00 & 72 \\
+			Driverlog & 0.12 & 0.07 & 102 \\
+			Ferry & 0.25 & 0.15 & 58 \\
+			Floortile & 0.73 & 0.43 & 80 \\
 			Gripper & 0.01 & 0.00 & 45 \\
-			Miconic & 0.10 & 0.09 & 54 \\
-			Satellite & 0.20 & 0.14 & 67 \\
-			Transport & 0.20 & 0.18 & 60 \\
-			Visitall & 0.43 & 0.39 & 37 \\
-			Zenotravel & 2.78 & 2.76 & 70
+			Miconic & 0.06 & 0.04 & 40 \\
+			Satellite & 0.18 & 0.12 & 60 \\
+			Transport & 0.18 & 0.18 & 44 \\
+			Visitall & 0.43 & 0.42 & 33 \\
+			Zenotravel & 0.26 & 0.25 & 53
 		\end{tabular}

### 2017-11-01 src/evaluator2.py
evaluator2.py now displays total recall and accuracy

@@ -10,7 +10,6 @@ import numpy as np
 # **************************************#
 try:
    cmdargs = sys.argv[1:]
-   print(cmdargs)
 
    if cmdargs[0] == "-p":
        partial_domain_filename = cmdargs[1]
@@ -103,5 +102,6 @@ recall_dels = np.float64(len(ref_dels.intersection(eva_dels)))/len(ref_dels)
 print("Pres: precision={}, recall={}".format(precision_pres, recall_pres))
 print("Adds: precision={}, recall={}".format(precision_adds, recall_adds))
 print("Dels: precision={}, recall={}".format(precision_dels, recall_dels))
+print("Total: precision={}, recall={}".format((precision_pres + precision_adds + precision_dels) / 3, (recall_pres + recall_adds + recall_dels) / 3))
 
 sys.exit(0)

### 2017-11-01 doc/icaps-18/strips-learning.tex
Updated tables 1 to 4

--- 2017-10-28 doc/icaps-18/strips-learning.tex
@@
 \begin{table}[hbt!]
-	\begin{footnotesize}
 	\begin{center}
-		\begin{tabular}{l|l|l|l|l|l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+		\resizebox{\columnwidth}{!}{%
+		\begin{tabular}{l|l|l|l|l|l|l|l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf All}\\ \cline{2-9}			
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
 			\hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 1.0 & 0.8 \\
-			Ferry & 0.8 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Floortile & 0.5 & 0.6 & 0.9 & 0.8 & 1.0 & 0.9 \\
-			Gripper & 1.0 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Miconic & 0.7 & 0.3 & 1.0 & 0.7 & 0.7 & 1.0 \\
-			Satellite & 0.6 & 0.2 & 1.0 & 1.0 & 1.0 & 0.75 \\
-			Transport & 1.0 & 0.3 & 0.5 & 0.8 & 1.0 & 0.6 \\
-			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Zenotravel & 1.0 & 0.3 & 0.7 & 0.8 & 1.0 & 0.7
+			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Driverlog & 1.0 & 0.43 & 0.67 & 0.86 & 1.0 & 0.86 & 0.89 & 0.71 \\
+			Ferry & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86 \\
+			Floortile & 0.54 & 0.64 & 0.90 & 0.82 & 1.0 & 0.91 & 0.81 & 0.79 \\
+			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
+			Miconic & 0.75 & 0.33 & 1.0 & 0.75 & 0.75 & 1.0 & 0.83 & 0.69 \\
+			Satellite & 0.67 & 0.29 & 1.0 & 1.0 & 1.0 & 0.75 & 0.89 & 0.68 \\
+			Transport & 1.0 & 0.30 & 0.57 & 0.80 & 1.0 & 0.60 & 0.86 & 0.57 \\
+			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+			Zenotravel & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 \\
+			\hline
+			\bf Total & 0.88 & 0.51 & 0.89 & 0.91 & 0.98 & 0.89 & 0.91 & 0.77 \\
 		\end{tabular}
+	}
 	\end{center}
-	\end{footnotesize}
 \caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks.}
@@
 			Visitall & 0.43 & 0.39 & 37 \\
-			Zenotravel & 2.78 & 2.76 & 70			
+			Zenotravel & 0.37 & 0.35 & 70			
 		\end{tabular}
@@
 \begin{table}[hbt!]
-\begin{footnotesize}
 	\begin{center}
-		\begin{tabular}{l|l|l|l|l|l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+		\resizebox{\columnwidth}{!}{%
+	\begin{tabular}{l|l|l|l|l|l|l|l|l|}
+		& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf All}\\ \cline{2-9}			
+		& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
 		\hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Driverlog & 1.0 & 0.4 & 0.6 & 0.8 & 1.0 & 0.8 \\
-			Ferry & 0.8 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Floortile & 0.5 & 0.6 & 0.9 & 0.8 & 1.0 & 0.9 \\
-			Gripper & 1.0 & 0.6 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Miconic & 0.7 & 0.3 & 1.0 & 0.7 & 0.7 & 1.0 \\
-			Satellite & 0.6 & 0.2 & 1.0 & 1.0 & 1.0 & 0.75 \\
-			Transport & 1.0 & 0.3 & 0.5 & 0.8 & 1.0 & 0.6 \\
-			Visitall & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Zenotravel & 1.0 & 0.3 & 0.7 & 0.8 & 1.0 & 0.7
+		Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+		Driverlog & 1.0 & 0.5 & 0.75 & 0.86 & 1.0 & 0.71 & 0.91 & 0.69 \\
+		Ferry & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86 \\
+		Floortile & 0.71 & 0.78 & 0.90 & 0.82 & 1.0 & 0.91 & 0.87 & 0.83 \\
+		Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
+		Miconic & 0.88 & 0.78 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.84 \\
+		Satellite & 0.85 & 0.79 & 0.83 & 1.0 & 1.0 & 0.75 & 0.89 & 0.85 \\
+		Transport & 1.0 & 0.60 & 0.80 & 0.80 & 1.0 & 0.60 & 0.93 & 0.67 \\
+		Visitall & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+		Zenotravel & 1.0 & 0.64 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.74 \\
+		\hline
+		\bf Total & 0.92 & 0.73 & 0.90 & 0.91 & 0.98 & 0.87 & 0.93 & 0.84 \\
 	\end{tabular}
+}
 	\end{center}
-	\end{footnotesize}
 \caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks exploiting the computed static predicates.}

### 2017-11-01 doc/icaps-18/strips-learning.tex
updated formalization pages 1 and 2

@@
 
-\title{Learning STRIPS action models with classical planning}
+\title{Learning \strips\ action models with classical planning}
 
@@
 \begin{abstract}
-This paper presents a novel approach for learning STRIPS action models from examples that compiles this particular inductive learning task into a classical planning task.
-Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where neither actions nor intermediate states are given.
-What is more, the compilation accepts partially specified action models and can also be used to validate whether certain observations of plan executions follow a given STRIPS action model (even if the action model is not fully specified).
+This paper presents a novel approach for learning \strips action models from examples that compiles this particular inductive learning task into classical planning.
+Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states (no intermediate action or state is given).
+What is more, the compilation accepts partially specified action models and can be used to validate whether certain observations of plan executions follow a given \strips\ action model, even if this model is not fully specified.
 \end{abstract}
@@
 On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML to the learning of STRIPS action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
+The application of inductive ML to the learning of \strips\ action models~\cite{fikes1971strips}, the vanilla action model for planning, is not straightforward though:
 \begin{itemize}
-\item The inputs to ML algorithms (the learning data) usually are finite vectors that encode the value of fixed features for a given set of objects.
-The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
+\item The inputs to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects.
+The input for learning planning action models are observations of plan executions (where each plan possibly has different length).
 \item The traditional output of ML algorithms is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning STRIPS action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions of the corresponding planning domain.
+When learning \strips\ action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions.
 \end{itemize}
 
-The learning of STRIPS action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on learning different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning STRIPS action models that can be defined as a classical planning compilation, so an off-the-shelf planner can address the learning task.
+The learning of \strips\ action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
 The compilation approach is appealing because it opens the door to the bootstrapping of planning action models but also because:
@@
 \item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
-In the extreme, the compilation could be used to validate whether observed plan executions are valid for a given STRIPS action model.
+In the extreme, the compilation could be used to validate whether observed plan executions are valid for a given \strips\ action model.
 \end{enumerate}
@@
 The paper is organized as follows.
-The first section presents the classical planning model, its extension to conditional effects (which is a requirement of the proposed compilation) and formalizes the STRIPS action model (the output of the addressed learning task).
-The second section formalizes the learning of STRIPS action models with regard to different amounts of available input knowledge.
+The first section presents the classical planning model, its extension to conditional effects (which is a requirement of the proposed compilation) and formalizes the \strips\ action model (the output of the addressed learning task).
+The second section formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
 The third and fourth sections describe our approach for addressing this particular learning task by compiling it into classical planning and how previous knowledge can be introduced in the compilation to constrain the space of possible action models and make learning more practicable.
-Finally, the last sections report the data collected during the empirical evaluation of our approach, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
+Finally, last sections report the data collected during the empirical evaluation of our approach, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
 
 \section{Background}
-This section defines the planning models used on this work as well as the STRIPS action model, the output of the learning task addressed in this paper.
+This section defines the planning models used on this work and the output of the learning task addressed in the paper.
 
@@
 either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent) and let $\neg L=\{\neg l:l\in L\}$ be the complement of $L$.
+A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent).
 We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
@@
 
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-Each action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called {\em preconditions}, a set of {\em effects} $\eff(a)\in\mathcal{L}(F)$.
-An action $a\in A$ is {\em applicable} in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is the {\em successor} state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
+A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. Each action $a\in A$ comprises three sets of literals:
+\begin{itemize}
+\item $\pre(a)$, called {\em preconditions}, that defines the literals that must hold for the action to be applicable.
+\item $\eff^+(a)$, called {\em positive effects}, that defines the fluents set to true by the action application.
+\item $\eff^-(a)$, called {\em negative effects}, that defines the fluents set to false by the application of the action. 
+\end{itemize}
+We say that an action $a\in A$ is {\em applicable} in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is the {\em successor state} $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
 
@@
 \subsection{Classical planning with conditional effects}
-Our approach for leaning STRIPS action models is compiling this leaning task into a classical planning task with conditional effects.
-Conditional effects allow us to compactly define actions whose particular effects depend on the current state.
-Nowadays many classical planners cope with conditional effects without compiling them away.
-In fact, the support of conditional effects was a requirement for participating at the IPC-2014~\cite{vallati:IPC:AIM2015} and IPC-2018.
+Our approach for leaning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
+Conditional effects allow us to compactly define actions whose effects depend on the current state.
+Many classical planners cope with conditional effects without compiling them away.
+In fact, supporting conditional effects is a requirement of IPC-2014~\cite{vallati:IPC:AIM2015} and IPC-2018.
 
-Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called the {\em precondition}, and a set of conditional effects $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the condition, and $E\in\mathcal{L}(F)$, the effect.
+Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called the {\em precondition}, and a set of {\em conditional effects} $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
 
@@
 \[
-\eff(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
+triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
 \]
-i.e.~effects whose conditions hold in $s$. The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=(s\setminus \neg \eff(s,a))\cup\eff(s,a)$.
+i.e.~effects whose conditions hold in $s$.
 
-\subsection{The STRIPS action model and the {\em variable} objects}
-This work addresses the learning of action schemes defined in the STRIPS language~\cite{fikes1971strips}.
-To formalized the output of this learning task, we assume that there exists a set of {\em predicates} $\Psi$ and that the fluents in $F$ are instantiated from these predicates, as in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}.
+The result of applying action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\del(s,a))\cup\add(s,a)\}$ where $del(s,a)\subseteq triggered(s,a)$ and $add(s,a)\subseteq triggered(s,a)$ are the triggered negative and positive effects .
 
-Formally, each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+\subsection{The \strips\ action schema and the {\em variable} objects}
+This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+Figure~\ref{fig:stack} shows the schema that corresponds to the {\em stack} action from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
 
-Likewise, we assume that STRIPS actions are instantiated from STRIPS operator schemes.
-Figure~\ref{fig:stack} shows the STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld~\cite{slaney2001blocks} represented in PDDL.
 \begin{figure}
-\begin{footnotesize}
+\begin{scriptsize}
 \begin{verbatim}
@@
   :parameters (?x1 ?x2 - object)
-  :precondition (and (holding ?x1)
-                     (clear ?x2))
-  :effect (and (not (holding ?x1))
-               (not (clear ?x2))
-               (handempty)
-               (clear ?x1)
-               (on ?x1 ?x2)))
+  :precondition (and (holding ?x1) (clear ?x2))
+  :effect (and (not (holding ?x1)) (not (clear ?x2))
+               (handempty) (clear ?x1) (on ?x1 ?x2)))
 \end{verbatim}
-\end{footnotesize}
- \caption{\small Example of a STRIPS operator schema that corresponds to the {\em stack} action from the blocksworld represented in PDDL.}
+\end{scriptsize}
+ \caption{\small Example of a \strips\ operator schema that corresponds to the {\em stack} action from the {\em blocksworld} represented in PDDL.}
 \label{fig:stack}
@@
 
-Let $\Omega_v$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names} and whose cardinality, $|\Omega_v|$, is given by the maximum arity of an action in a given planning frame $\Phi$.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because operators {\small\tt stack} and {\small\tt unstack}, with the maximum arity, have two parameters each.
+To formalize the output of the learning task, we assume that there exists a set of {\em predicates} $\Psi$ and that fluents $F$ are instantiated from these predicates, as in PDDL.
+Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
+Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
+$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+Likewise, we assume that actions $a\in A$ are instantiated from \strips\ operator schemes.
 
-Let us define also a new set of fluents $F_{v}$ that results instantiating $\Psi$ but using only the {\em variable objects} in $\Omega_v$.
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names} and that is bound by the maximum arity of an action in a given planning frame.
+For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters each.
+
+Let us define also a new set of fluents $F_{v}$, s.t.
+$F\cap F_v=\emptyset$, that results instantiating $\Psi$ using only {\em variable objects} in $\Omega_v$.
+This set defines the elements that can appear in an action schema.
 In the blocksworld $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.
 
-We are now ready to define a STRIPS action model $\Xi$ as a set of operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ such that:
+We are now ready to define $\Xi$, a set of operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ such that:
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and a enumeration of the variable objects bound by the operator arity, $pars(\xi)\in\Omega_v^{ar(\xi)}$.
-The headers for a 4-operator blocksworld are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$, an enumeration of the variable objects bound by the operator arity.
+The headers for a 4-operator blocksworld are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
 \item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
@@
 
-\section{Learning STRIPS action models}
-Learning STRIPS action models from fully available input knowledge, i.e.
+\section{Learning \strips action models}
+Learning \strips action models from fully available input knowledge, i.e.
 from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
-In this case, a STRIPS operator schema is derived by lifting the literals that change between the pre and post-state of the corresponding action executions.
+In this case, a \strips operator schema is derived by lifting the literals that change between the pre and post-state of the corresponding action executions.
 Likewise, the preconditions are derived lifting the minimal set of literals that appears in all the pre-states that correspond to the same operator schema.
@@
 
- \caption{\small Example of a task for learning a STRIPS action model in the blocksworld.}
+ \caption{\small Example of a task for learning a \strips action model in the blocksworld.}
 \label{fig:lexample}
@@
 
-A solution to a learning task $\Lambda$ is a set of STRIPS operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
+A solution to a learning task $\Lambda$ is a set of \strips operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
 
@@
 
-In this case, a solution must not only determine a possible STRIPS action model but also the plans that explain the given labels using the learned model (this information about the actions is no longer given in the learning examples).
+In this case, a solution must not only determine a possible \strips action model but also the plans that explain the given labels using the learned model (this information about the actions is no longer given in the learning examples).
 We assume however, that the headers of the operators schema are known.
 
-\section{Learning STRIPS action models with planning}
+\section{Learning \strips action models with planning}
 Our approach for addressing a learning task $\Lambda$ or $\Lambda'$, is compiling it into a classical planning task with conditional effects.
@@
 \begin{enumerate}
-\item Programs the STRIPS action model $\Xi$. For each $\xi\in\Xi$, the solution plan has a prefix that determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed STRIPS action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ if available).
+\item Programs the \strips action model $\Xi$. For each $\xi\in\Xi$, the solution plan has a prefix that determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Validates the programmed \strips action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ if available).
 For every label {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
-We call this process the validation of the programmed STRIPS action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
+We call this process the validation of the programmed \strips action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
 If information about the corresponding plan $\pi_t\in \Pi$ is available, is used in the validation.
@@
 
-Now we are ready to formalize our compilations for learning STRIPS action models with classical planning.
+Now we are ready to formalize our compilations for learning \strips action models with classical planning.
 We start with $\Lambda'$ because this learning task requires smaller input knowledge.
@@
 \item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$).
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the STRIPS operator schema $\xi\in \Xi$.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips operator schema $\xi\in \Xi$.
 \item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or they are being validated (already programmed).
@@
 
-The compilation is {\em complete} in the sense that it does not discard any possible STRIPS action model.
+The compilation is {\em complete} in the sense that it does not discard any possible \strips action model.
 
@@
 \section{Constraining the learning hypothesis space with additional input knowledge}
-Here we show that further input knowledge can be used to constrain the space of the possible action models and make the learning of STRIPS action models more practicable.
+Here we show that further input knowledge can be used to constrain the space of the possible action models and make the learning of \strips action models more practicable.
 
@@
 \subsection{Previously specified action models}
-Our compilation does not require to start learning STRIPS action models from scratch and accepts partially specified operator schema where some preconditions and effects are a priori known.
+Our compilation does not require to start learning \strips action models from scratch and accepts partially specified operator schema where some preconditions and effects are a priori known.
 
@@
 
-In the extreme, when a fully specified STRIPS action model is given, the compilation validates whether an observed plan follows the given model.
-In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given STRIPS action model is {\em valid} for the given set of examples.
-If $P_{\Lambda'}$ is unsolvable it means that the given STRIPS action model is invalid since it is not compliant with all the given examples.
+In the extreme, when a fully specified \strips action model is given, the compilation validates whether an observed plan follows the given model.
+In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips action model is {\em valid} for the given set of examples.
+If $P_{\Lambda'}$ is unsolvable it means that the given \strips action model is invalid since it is not compliant with all the given examples.
 Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
@@
 These early works were based on direct lifting of the observed states and supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge.
-The recent work reported in \cite{WalshL08} shows that although learning STRIPS operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
+The recent work reported in \cite{WalshL08} shows that although learning \strips operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
 
@@
 Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
-An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then STRIPS rules are derived from the classifiers' parameters.
+An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
 
@@
 \section{Conclusions}
-The paper presented a novel approach for learning STRIPS action models from examples exclusively using classical planning.
+The paper presented a novel approach for learning \strips action models from examples exclusively using classical planning.
 The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
@@
 The empirical results show that our approach can generate non-trivial models from very small data sets.
-In addition the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because all the actions for programming the STRIPS actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon.
+In addition the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because all the actions for programming the \strips actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon.
 The same happens with all the actions for programming an action effect can also be applied in a single planning step.

### 2017-11-01 doc/icaps-18/strips-learning.tex
pages 3 and 4 updated

@@
 
-\section{Learning \strips action models}
-Learning \strips action models from fully available input knowledge, i.e.
+\section{Learning \strips\ action models}
+Learning \strips\ action models from fully available input knowledge, i.e.
 from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
-In this case, a \strips operator schema is derived by lifting the literals that change between the pre and post-state of the corresponding action executions.
-Likewise, the preconditions are derived lifting the minimal set of literals that appears in all the pre-states that correspond to the same operator schema.
+In this case, because intermediate states are available, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+Preconditions are derived lifting the minimal set of literals that appears in all the pre-states that correspond to the same operator schema.
 
@@
 
-\subsection{Learning from labeled plans}
-This learning task is formalized as $\Lambda=\tup{\Psi,\Pi,\Sigma}$:
+\subsection{Learning from (initial, final) state pairs}
+This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
+No information about the actions in the plans is given however, we assume that the headers of the operators schema are known.
+
+This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma}$:
 \begin{itemize}
 \item $\Psi$, the set of predicates that define the abstract state space of a given planning domain.
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, the given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-\item $\Sigma=\{\sigma_1,,\ldots,\sigma_{\tau}\}$, is a set of labels s.t.
-each plan $\pi_t$, {\small $1\leq t\leq \tau$}, has an associated label $\sigma_t=(s_0^t,s_{n}^t)$ where $s_{n}^t$ is the state resulting from executing the plan $\pi_t$ starting from the state $s_0^t$.
+\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$, a set of (initial, final) state pairs s.t., for every {\tt\small $1\leq t\leq \tau$}, a pair $\sigma_t=(s_0^t,s_{n}^t)$ comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from a given {\em initial} state $s_0^t$.
 \end{itemize}
 
-Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda$ in the blocksworld (the figure shows the content of $\Psi$, $\Pi$ and $\Sigma$).
+A solution to the learning task $\Lambda$  is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
+
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$ that explain the given states $\Sigma$ using the learned \strips\ model.
+
+\subsection{Learning from labeled plans}
+Here we augment the amount of provided input knowledge and define the learning task as $\Lambda'=\tup{\Psi,\Sigma,\Pi}$.
+
+Now $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+
+A solution to a learning task $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
+
+Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda'$ in the blocksworld (the figure shows the content of $\Psi$, $\Pi$ and $\Sigma$).
 This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of a plan for inverting a four-block tower.
@@
 \begin{figure}
-\begin{small}
 {\tt ;;; Predicates in $\Psi$}
+\begin{footnotesize}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{small}
+\end{footnotesize}
 
-\vspace{0.5cm}
+\vspace{0.2cm}
 
 \begin{subfigure}{.25\textwidth}
-\begin{small}
 {\tt ;;; Plan $\pi_1$}
+\begin{footnotesize}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{small}
+\end{footnotesize}
 \end{subfigure}%
 \begin{subfigure}{.6\textwidth}
-{\small\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
+{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
 \begin{lstlisting}[mathescape]
@@
 
- \caption{\small Example of a task for learning a \strips action model in the blocksworld.}
+ \caption{\small Example of a task for learning a \strips\ action model in the blocksworld.}
 \label{fig:lexample}
@@
 
-A solution to a learning task $\Lambda$ is a set of \strips operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
-
-\subsection{Learning from initial/final states}
-Here we reduce the amount of provided input knowledge and define the learning task $\Lambda'=\tup{\Psi,\Sigma}$ that corresponds to watching only the results of the plan executions.
-In this case no information about the executed plans is given.
-A solution to the $\Lambda'$ learning task is a set of operator schema $\Xi$ that is compliant only with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
-
-In this case, a solution must not only determine a possible \strips action model but also the plans that explain the given labels using the learned model (this information about the actions is no longer given in the learning examples).
-We assume however, that the headers of the operators schema are known.
-
-\section{Learning \strips action models with planning}
+\section{Learning \strips\ action models with planning}
 Our approach for addressing a learning task $\Lambda$ or $\Lambda'$, is compiling it into a classical planning task with conditional effects.
@@
 \begin{enumerate}
-\item Programs the \strips action model $\Xi$. For each $\xi\in\Xi$, the solution plan has a prefix that determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ if available).
-For every label {\small $1\leq t\leq \tau$}, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
-We call this process the validation of the programmed \strips action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
-If information about the corresponding plan $\pi_t\in \Pi$ is available, is used in the validation.
+\item Programs the \strips\ action model $\Xi$. For each $\xi\in\Xi$, the solution plan has a prefix that determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Validates the programmed \strips\ action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ if available).
+For every label $\sigma_t\in \Sigma$, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
+We call this process the validation of the programmed \strips\ action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
+If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
 \end{enumerate}
 
-Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ kind.
-This classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld (previously specified operator schemes for $pickup$, $putdown$ and $unstack$ are given), using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-The plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operators.
-The details of the compilation are explained in the remaining of the section.
-
-\begin{figure}[hbt]
-\begin{small}
-\begin{verbatim}
-     0 : (program_pre_clear_stack_v1)
-     1 : (program_pre_handempty_stack)
-     2 : (program_pre_holding_stack_v2)
-     3 : (program_pre_on_stack_v1_v1)
-     4 : (program_pre_on_stack_v1_v2)
-     5 : (program_pre_on_stack_v2_v1)
-     6 : (program_pre_on_stack_v2_v2)
-     7 : (program_pre_ontable_stack_v1)
-     8 : (program_pre_ontable_stack_v2)
-     9 : (program_eff_clear_stack_v1)
-    10 : (program_eff_clear_stack_v2)
-    11 : (program_eff_handempty_stack)
-    12 : (program_eff_holding_stack_v1)
-    13 : (program_eff_on_stack_v1_v2)
-    14 : (apply_unstack a b i1 i2)
-    15 : (apply_putdown a i2 i3)
-    16 : (apply_unstack b c i3 i4)
-    17 : (apply_stack b a i4 i5)
-    18 : (apply_unstack c d i5 i6)
-    19 : (apply_stack c b i6 i7)
-    20 : (apply_pickup d i7 i8)
-    21 : (apply_stack d c i8 i9)
-    22 : (validate_1)
-\end{verbatim}
-\end{small}
- \caption{\small Example of a plan for programming and validating the operator schema $unstack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
-\label{fig:plan}
-\end{figure}
-
-To formalize our compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, one for each leaning example, that belong to the same planning frame (i.e.
+To formalize our compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
 same fluents and actions and differ only in the initial state and goals).
-The set of fluents $F$ is built instantiating the predicates in $\Psi$ with the objects appearing in the input learning examples.
+The fluents $F$ are built instantiating the predicates in $\Psi$ with the objects appearing in the inputs.
 Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
@@
 
-Now we are ready to formalize our compilations for learning \strips action models with classical planning.
-We start with $\Lambda'$ because this learning task requires smaller input knowledge.
-Given a learning task $\Lambda'=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ such that:
+Now we are ready to formalize the compilations.
+We start with $\Lambda$, because this learning task requires fewer input knowledge.
+Given a learning task $\Lambda=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
 \begin{itemize}
-\item $F_{\Lambda'}$ extends $F$ with:
+\item $F_{\Lambda}$ extends $F$ with:
 \begin{itemize}
-\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$).
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips operator schema $\xi\in \Xi$.
+\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$.
+For instance in the working example the preconditions of the $stack$ schema are reprenseted by the fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to true.
 \item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or they are being validated (already programmed).
@@
 \end{itemize}
-\item $I_{\Lambda'}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
-\item $G_{\Lambda'}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
-\item $A_{\Lambda'}$ contains actions of three different kinds:
+\item $I_{\Lambda}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
+\item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
+\item $A_{\Lambda}$ contains actions of three different kinds:
 \begin{enumerate}
@@
 \begin{itemize}
-\item The actions for removing a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
+\item The actions for {\bf removing} a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
 
@@
 
-\item The actions for adding a {\em negative} or a {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
+\item The actions for {\bf adding} a {\em negative} or a {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
 
@@
 \end{itemize}
-
-\begin{lemma}
-Any classical plan $\pi$ that solves $P_{\Lambda'}$ induces an action model $\Xi$ that solves the learning task $\Lambda'$.
-\end{lemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-The compilation forces that once the preconditions of an operator schema $\Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after the operator preconditions are programmed).
-Furthermore, once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve the classical planning task $P_{\Lambda'}$, there is only one way of achieving goals $\{test_t\}$,{\small $1\leq t\leq \tau$}: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$ defined by the associated label $\sigma_t$, starting from $s_0^t$, defined in this same label.
-If this is achieved for all the input examples of the learning task $\Lambda'$ (i.e.
-for all the labels $\sigma_t$), it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda'$.
-\end{small}
-\end{proof}
-
-The compilation is {\em complete} in the sense that it does not discard any possible \strips action model.
 
@@
 \label{fig:compilation}
+\end{figure}
+
+\begin{lemma}
+Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
+The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after the preconditions are programmed).
+Furthermore, once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
+To solve $P_{\Lambda}$, there is only one way of achieving goals $\{test_t\}$, {\small $1\leq t\leq \tau$}: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$ defined in $\sigma_t$, starting from $s_0^t$ (defined in this same label).
+If this is achieved for all the input examples {\small $1\leq t\leq \tau$} of the learning task, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
+\end{small}
+\end{proof}
+
+The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model.
+
+Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ kind.
+This classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld (previously specified operator schemes for $pickup$, $putdown$ and $unstack$ are given), using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+The plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operators.
+The details of the compilation are explained in the remaining of the section.
+
+\begin{figure}
+\begin{small}
+\begin{verbatim}
+     0 : (program_pre_clear_stack_v1)
+     1 : (program_pre_handempty_stack)
+     2 : (program_pre_holding_stack_v2)
+     3 : (program_pre_on_stack_v1_v1)
+     4 : (program_pre_on_stack_v1_v2)
+     5 : (program_pre_on_stack_v2_v1)
+     6 : (program_pre_on_stack_v2_v2)
+     7 : (program_pre_ontable_stack_v1)
+     8 : (program_pre_ontable_stack_v2)
+     9 : (program_eff_clear_stack_v1)
+    10 : (program_eff_clear_stack_v2)
+    11 : (program_eff_handempty_stack)
+    12 : (program_eff_holding_stack_v1)
+    13 : (program_eff_on_stack_v1_v2)
+    14 : (apply_unstack a b i1 i2)
+    15 : (apply_putdown a i2 i3)
+    16 : (apply_unstack b c i3 i4)
+    17 : (apply_stack b a i4 i5)
+    18 : (apply_unstack c d i5 i6)
+    19 : (apply_stack c b i6 i7)
+    20 : (apply_pickup d i7 i8)
+    21 : (apply_stack d c i8 i9)
+    22 : (validate_1)
+\end{verbatim}
+\end{small}
+ \caption{\small Example of a plan for programming and validating the operator schema $stack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+\label{fig:plan}
 \end{figure}

### 2017-11-01 doc/icaps-18/strips-learning.tex
updated tables 5 and 6

--- 2017-11-01 doc/icaps-18/strips-learning.tex
@@
 	\begin{center}
-		\begin{tabular}{l|l|l|l|l|l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del}  \\ \cline{2-7}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+		\resizebox{\columnwidth}{!}{%
+			\begin{tabular}{l|l|l|l|l|l|l|l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf All}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
 				\hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Driverlog & 1.0 & 0.4 & 1.0 & 1.0 & 1.0 & 0.7 \\
-			Ferry & 1.0 & 0.6 & 0.5 & 1.0 & 1.0 & 1.0 \\
-			Floortile & 0.5 & 0.6 & 1.0 & 0.8 & 1.0 & 0.8 \\
-			Gripper & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Miconic & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Satellite & 1.0 & 0.5 & 0.5 & 1.0 & 1.0 & 1.0 \\
-			Transport & 1.0 & 0.2 & 1.0 & 1.0 & 1.0 & 0.5 \\
-			Zenotravel & 1.0 & 0.5 & 0.6 & 0.6 & 1.0 & 1.0
+				Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.91 \\
+				Ferry & 1.0 & 0.67 & 0.5 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
+				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
+				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+				Miconic & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
+				Satellite & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
+				Transport & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 0.50 & 1.0 & 0.67 \\
+%				Visitall & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				Zenotravel & 1.0 & 0.83 & 0.67 & 0.67 & 1.0 & 1.0 & 0.89 & 0.83 \\
+				\hline
+				\bf Total & 0.97 & 0.68 & 0.91 & 0.94 & 1.0 & 0.92 & 0.96 & 0.85 \\
 			\end{tabular}
+		}
 	\end{center}
@@
 			Blocks & 0.07 & 0.01 & 54  \\
-			Driverlog & 0.06 & 0.03 & 71 \\
-			Ferry & 0.18 & 0.14 & 50 \\
-			Floortile & 0.81 & 0.68 & 92 \\
+			Driverlog & 0.06 & 0.03 & 59 \\
+			Ferry & 0.20 & 0.14 & 50 \\
+			Floortile & 0.54 & 0.45 & 58 \\
 			Gripper & 0.01 & 0.00 & 37 \\
-			Miconic & 0.06 & 0.04 & 40  \\
-			Satellite & 0.18 & 0.12 & 43 \\
-			Transport & 0.15 & 0.14 & 33 \\
-			Zenotravel & 2.54 & 2.54 & 42
+			Miconic & 0.06 & 0.04 & 33  \\
+			Satellite & 0.17 & 0.15 & 48 \\
+			Transport & 0.12 & 0.10 & 33 \\
+			Zenotravel & 0.26 & 0.26 & 33
 		\end{tabular}

### 2017-11-01 doc/icaps-18/strips-learning.tex
updated page 5 and merged with last results by Diego

@@
 
-Figure~\ref{fig:plan} shows a classical plan for solving a learning task of the $\Lambda$ kind.
+\section{Constraining the learning hypothesis space with additional input knowledge}
+Here we show that further input knowledge can be used to constrain the space of the possible action models and make the learning of \strips\ action models more practicable.
+
+\subsection{Example plans}
+Here  we extend our compilation to address the learning scenario where, a set of plans is available.
+Given a learning task $\Lambda'=\tup{\Psi,\Pi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
+\begin{itemize}
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the $j$ steps of plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
+Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current and next plan steps.
+\item $I_{\Lambda'}$ is extended with the fluents from $F_{\Pi}$ that encode the plan $\pi_1\in \Pi$, the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
+Goals are the same as in the previous compilation $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
+\item With respect to $A_{\Lambda'}$.
+\begin{enumerate}
+\item The actions for programming the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
+\item The actions for applying an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+This mechanism forces that these actions are only applied as in the example plans.
+\item The actions for updating the active test have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
+\begin{small}
+\begin{align*}
+&\{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}},\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
+\end{align*}
+\end{small}
+\end{enumerate}
+\end{itemize}
+
+\subsection{Previously specified action models}
+Our compilation does not require to start learning \strips\ action models from scratch and accepts partially specified operator schema where some preconditions and effects are a priori known.
+
+The known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda'}$.
+The corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
+
+To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task for getting the blocksworld action model where the schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are known in advance.
 This classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld (previously specified operator schemes for $pickup$, $putdown$ and $unstack$ are given), using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
 The plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operators.
-The details of the compilation are explained in the remaining of the section.
+
+In the extreme, when a fully specified \strips\ action model is given, the compilation validates whether an observed plan follows the given model.
+In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips\ action model is {\em valid} for the given set of examples.
+If $P_{\Lambda'}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
+Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
 
@@
 
-\section{Constraining the learning hypothesis space with additional input knowledge}
-Here we show that further input knowledge can be used to constrain the space of the possible action models and make the learning of \strips action models more practicable.
-
-\subsection{Example plans}
-Here  we extend our compilation to address the learning scenario where, a set of plans is available.
-Given a learning task $\Lambda=\tup{\Psi,\Pi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}'$ as follows:
-\begin{itemize}
-\item $F_{\Lambda}$ extends $F_{\Lambda'}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the $j$ steps of plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current and next plan steps.
-\item $I_{\Lambda}$ is extended with the fluents from $F_{\Pi}$ that encode the plan $\pi_1\in \Pi$, the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
-Goals are the same as in the previous compilation $G_{\Lambda}=G_{\Lambda}'=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
-\item With respect to $A_{\Lambda}$.
-\begin{enumerate}
-\item The actions for programming the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for applying an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism forces that these actions are only applied as in the example plans.
-\item The actions for updating the active test have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to load the next plan $\pi_{t+1}$:
-\begin{small}
-\begin{align*}
-&\{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}},\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
-\end{align*}
-\end{small}
-\end{enumerate}
-\end{itemize}
-
-\subsection{Previously specified action models}
-Our compilation does not require to start learning \strips action models from scratch and accepts partially specified operator schema where some preconditions and effects are a priori known.
-
-The known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that are part of the initial state $I_{\Lambda'}$.
-The corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task for getting the blocksworld action model where the schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are known in advance.
-
-In the extreme, when a fully specified \strips action model is given, the compilation validates whether an observed plan follows the given model.
-In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips action model is {\em valid} for the given set of examples.
-If $P_{\Lambda'}$ is unsolvable it means that the given \strips action model is invalid since it is not compliant with all the given examples.
-Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
-
 \subsection{Static predicates}
 A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action schema~\cite{fox:TIM:JAIR1998}.
-Therefore, one can get rid of the mechanism for programming static predicates as action effects keeping the compilation complete.
+Therefore, one can get rid of the mechanism for programming these predicates as the effect of any action schema while keeping the compilation complete.
 Formally, given a static predicate $p$:
@@
 
-Static predicates can also be used to constrain the space of possible preconditions by looking at the given labels.
-If a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Sigma$, then the fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
+Static predicates can be used to constrain the space of possible preconditions by looking at the given set of labels $\Sigma$.
+In particular one can assume that if a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ it means that is not possible and then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
 For instance in the {\em zenotravel} domain fluents $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (as well as their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} cannot be compliant with any state in $\Sigma$.
 
-Likewise, static predicates can constrain the space of possible preconditions looking at the given example plans.
+Likewise, static predicates can constrain the space of possible preconditions looking at the given set of example plans.
 In this case, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discarded for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
 $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
-Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static fluent {\tt\small (next fl2 fl3)} but it does not include {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} it means that the only possible precondition including the static predicate is $pre\_next\_fly\_v1\_v2$.
+Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but it does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} it means that the only possible precondition including the static predicate is $pre\_next\_fly\_v1\_v2$.
 
-In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of candidates of static predicates from the input examples to the learning task.
+In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of {\em potential} static predicates from the input examples to the learning task.
 
@@
 
+\newpage
+
 \bibliographystyle{aaai}

### 2017-11-03 src/compiler2.py
compiler2.py: support for static predicates in modes 2 and 3

@@ -73,7 +73,17 @@ def get_static_predicates(tasks, predicates):
 
         candidates = candidates.intersection(task_candidates)
 
-    return [p for p in predicates if p[0] in candidates]
+    reflexive_static_predicates = dict()
+    for candidate in candidates:
+        reflexive_static_predicates[candidate] = True
+        for task in tasks:
+            init_predicates = set([p for p in task.init if p.predicate == candidate])
+            for predicate in init_predicates:
+                if len(predicate.args) == 1 or len(set(predicate.args)) != 1:
+                    reflexive_static_predicates[candidate] = False
+                    break
+
+    return [p for p in predicates if p[0] in candidates], reflexive_static_predicates
 
 def get_static_precondition(predicate, action, plans, tasks):
     static_preconditions = set()
@@ -122,32 +132,32 @@ def possible_pred_for_action(task, p, a, tup):
 # **************************************#
 # MAIN
 # **************************************#
-try:
-    if "-s" in sys.argv:
-        check_static_predicates = True
-        sys.argv.remove("-s")
-    else:
-        check_static_predicates = False
-
-    domain_folder_name  = sys.argv[1]
-    domain_file = sys.argv[2]
-    problems_prefix_filename = sys.argv[3]
-    plans_prefix_filename = sys.argv[4]
-    input_level = int(sys.argv[5])
-
-except:
-    print "Usage:"
-    print sys.argv[0] + "[-s] <domain> <domain filename> <problems prefix>  <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
-    sys.exit(-1)
+# try:
+#     if "-s" in sys.argv:
+#         check_static_predicates = True
+#         sys.argv.remove("-s")
+#     else:
+#         check_static_predicates = False
+#
+#     domain_folder_name  = sys.argv[1]
+#     domain_file = sys.argv[2]
+#     problems_prefix_filename = sys.argv[3]
+#     plans_prefix_filename = sys.argv[4]
+#     input_level = int(sys.argv[5])
+#
+# except:
+#     print "Usage:"
+#     print sys.argv[0] + "[-s] <domain> <domain filename> <problems prefix>  <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
+#     sys.exit(-1)
 
 
 # ../benchmarks/handpicked/blocks/ test plan 0
[…]

### 2017-11-03 src/compiler2.py
compiler2.py: support for static predicates in modes 2 and 3 (input fix)

@@ -132,32 +132,32 @@ def possible_pred_for_action(task, p, a, tup):
 # **************************************#
 # MAIN
 # **************************************#
-# try:
-#     if "-s" in sys.argv:
-#         check_static_predicates = True
-#         sys.argv.remove("-s")
-#     else:
-#         check_static_predicates = False
-#
-#     domain_folder_name  = sys.argv[1]
-#     domain_file = sys.argv[2]
-#     problems_prefix_filename = sys.argv[3]
-#     plans_prefix_filename = sys.argv[4]
-#     input_level = int(sys.argv[5])
-#
-# except:
-#     print "Usage:"
-#     print sys.argv[0] + "[-s] <domain> <domain filename> <problems prefix>  <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
-#     sys.exit(-1)
+try:
+    if "-s" in sys.argv:
+        check_static_predicates = True
+        sys.argv.remove("-s")
+    else:
+        check_static_predicates = False
+
+    domain_folder_name  = sys.argv[1]
+    domain_file = sys.argv[2]
+    problems_prefix_filename = sys.argv[3]
+    plans_prefix_filename = sys.argv[4]
+    input_level = int(sys.argv[5])
+
+except:
+    print "Usage:"
+    print sys.argv[0] + "[-s] <domain> <domain filename> <problems prefix>  <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
+    sys.exit(-1)
 
 
 # ../benchmarks/handpicked/blocks/ test plan 0
-domain_folder_name = "../benchmarks/handpicked/zenotravel/"
-domain_file = "domain"
-problems_prefix_filename = "test"
-plans_prefix_filename = "plan"
-input_level = 3
-check_static_predicates = True
+# domain_folder_name = "../benchmarks/handpicked/zenotravel/"
+# domain_file = "domain"
+# problems_prefix_filename = "test"
+# plans_prefix_filename = "plan"
+# input_level = 3
+# check_static_predicates = True
 
 # Reading the example plans
 plans = []

### 2017-11-03 src/compiler2.py
compiler2.py: support for static predicates in modes 2 and 3 (more fixes)

@@ -325,10 +325,11 @@ for a in actions:
         for p in predicates:
             for tup in itertools.product(var_ids, repeat=(len(p) - 1)):
                 if possible_pred_for_action(fd_task, p, a, tup):
-                    if input_level <= config.INPUT_STEPS:
-                        continue
-                    elif not reflexive_static_predicates.get(p[0]) and len(set(tup)) == 1:
-                        continue
+                    if p in static_predicates and check_static_predicates:
+                        if input_level <= config.INPUT_STEPS:
+                            continue
+                        elif not reflexive_static_predicates.get(p[0]) and len(set(tup)) == 1:
+                            continue
                     vars = ["var" + str(t) for t in tup]
                     disjunction = pddl.conditions.Disjunction(
                         [pddl.conditions.NegatedAtom("pre_" + p[0] + "_" + a[0] + "_" + "_".join(map(str, vars)), [])] + [

### 2017-11-03 doc/icaps-18/strips-learning.tex

@@
 \begin{figure}
-\begin{scriptsize}
+\begin{footnotesize}
 \begin{verbatim}
@@
   :precondition (and (holding ?x1) (clear ?x2))
-  :effect (and (not (holding ?x1)) (not (clear ?x2))
-               (handempty) (clear ?x1) (on ?x1 ?x2)))
+  :effect (and (not (holding ?x1)) 
+               (not (clear ?x2))
+               (handempty) (clear ?x1) 
+               (on ?x1 ?x2)))
 \end{verbatim}
-\end{scriptsize}
+\end{footnotesize}
  \caption{\small Example of a \strips\ operator schema that corresponds to the {\em stack} action from the {\em blocksworld} represented in PDDL.}
@@
 \section{Constraining the learning hypothesis space with additional input knowledge}
+\label{sec:Constraining}
 Here we show that further input knowledge can be used to constrain the space of the possible action models and make the learning of \strips\ action models more practicable.
@@
 \section{Evaluation}
-We evaluated our approach for learning STRIPS action models from different amounts of available input knowledge.
+This section evaluates our approach for learning \strips\ action models starting from different amounts of available input knowledge.
 
 \subsection{Experimental setup}
-All the experiments run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
-The domains used in the evaluation are IPC domains that satisfy the STRIPS requirement~\cite{fox2003pddl2}, and are taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+All the experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
+The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, and that are taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
 
-\subsubsection{The classical planner.}
-The classical planner used to solve the instances that result from our compilations is the SAT-based classical planner {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We decided to use a SAT-based planner because its ability to deal with planning instances populated with dead-ends.
-In addition, in a SAT-based classical planner all the actions for programming the STRIPS actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon.
-Likewise, all the actions for programming an action effect can also be applied in a single planning step.
+\subsubsection{The planner.}
+The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
+We use this SAT-based planner because its ability to deal with planning instances populated with dead-ends.
+In addition, in this planner the actions for programming preconditions can be done in parallel, in a single planning step, given that these actions do not interact.
+Likewise, the actions for programming action effects are also applicable in a single planning step reducing significantly the planning horizon.
 
@@
 The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
-These two metrics are frequently used in pattern recognition, information retrieval and binary classification and are more informative that simply counting the number of model errors or computing the {\em symmetric difference} between the learned model and the baseline ~\cite{davis2006relationship}.
+These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned model and the reference model~\cite{davis2006relationship}.
 
@@
 \subsubsection{The learning examples.}
+The set of learning examples is fixed for all the experiments so we can evaluate better the effect of the different amount of input knowledge in the quality of the learned models.
+%All learning examples are generated solving a small problem from the mentioned repository.
 
 \subsection{Learning action models from example plans}
-Here we assess the performance of our learning approach when addressing $\Lambda$ learning tasks.
-Here the {\em precision} and {\em recall} metrics are computed in the learned models with respect to the actual models taken as reference.
-Table~\ref{tab:results_plans} summarizes the obtained results, precision ({\bf P}) and recall ({\bf R}) is computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}).
+Here we assess the performance of our learning approach when addressing learning tasks of the $\Lambda'$ kind.
+Table~\ref{tab:results_plans} summarizes the obtained results, precision ({\bf P}) and recall ({\bf R}) is computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report the averages.
+The last row is also an average report.
 
@@
 		\begin{tabular}{l|l|l|l|l|l|l|l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf All}\\ \cline{2-9}			
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
 			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
@@
 			\hline
-			\bf Total & 0.88 & 0.51 & 0.89 & 0.91 & 0.98 & 0.89 & 0.91 & 0.77 \\
+			\bf  & 0.88 & 0.51 & 0.89 & 0.91 & 0.98 & 0.89 & 0.91 & 0.77 \\
 		\end{tabular}
@@
 	\end{center}
-\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks.}
+\caption{\small {\em Precision} and {\em recall} values of the learned models.}
 \label{tab:results_plans}
@@
 
-Table~\ref{tab:time_plans} reports the total planning time, and the preprocessing time (in seconds) invested by the classical planner {\sc Madagascar} to solve the classical planning tasks that result from our compilation as well as the plan length of the found solutions.
+Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning tasks that result from our compilation as well as the number of actions in the solutions found.
+Remarkably all the learning tasks are solved in a few seconds time.
 
@@
 		\begin{tabular}{l|c|c|c|}			
-			Domain & Total time & Preprocess & Plan length  \\
+			 & Total time & Preprocess & Plan length  \\
 			\hline
@@
         \end{footnotesize}
-	\caption{\small Planning time, preprocessing time and Plan length.}
+	\caption{\small Planning time, preprocessing time and plan length.}
 	\label{tab:time_plans}	
@@
 \subsection{Exploiting static predicates}
-Here we repeat the previous evaluation but compute now a candidates of static predicates from the inputs to the learning task and exploit this knowledge to reduce the space of possible action models.
-The set of considered candidate {\em static predicates} is the set of predicates whose instantiation appear the same in the initial and final states for all the learning examples {\small $1\leq t\leq \tau$}.
+Here we repeat the previous evaluation but now, candidates of static predicates are computed from the learning examples.
+The considered static predicates are used to constrain the space of possible action models as explained in the previous section.
+The set of candidate {\em static predicates} considered  here is the set of predicates s.t.
+every predicate instantiation appears the same in the initial and final states, for all the learning examples {\small $1\leq t\leq \tau$}.
 Tables~\ref{tab:results_plans_static} and~\ref{tab:time_plans_static} show the obtained results.
@@
 	\begin{tabular}{l|l|l|l|l|l|l|l|l|}
-		& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf All}\\ \cline{2-9}			
+		& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c}{\bf }\\ \cline{2-9}			
 		& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
@@
 		\hline
-		\bf Total & 0.92 & 0.73 & 0.90 & 0.91 & 0.98 & 0.87 & 0.93 & 0.84 \\
+		\bf  & 0.92 & 0.73 & 0.90 & 0.91 & 0.98 & 0.87 & 0.93 & 0.84 \\
 	\end{tabular}
@@
 	\end{center}
-\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the $\Lambda$ learning tasks exploiting the computed static predicates.}
+\caption{\small The {\em precision} and {\em recall} values achieved by the learned models exploiting static predicates.}
 \label{tab:results_plans_static}
@@
         \end{footnotesize}
-	\caption{\small Planning time, preprocessing time and Plan length exploiting the computed static predicates.}
+	\caption{\small Planning time, preprocessing time and plan length exploiting static predicates.}
 	\label{tab:time_plans_static}	
 \end{table}
+
+We can observed that identifying static predicates drives to better models, in particular with larger recall values.
+This fact evidences that many of the missing preconditions in the previous models corresponded to static predicates (there were no inncetive learned them since they always hold).
+When static predicates are identified, the resulting compilation is much compact and produces smaller planning/instantiation times.
+Interestingly, one can identify the domains with static predicates by just looking at the plan length.
+In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less actions for programming preconditions are required.
 
@@
 Here we evaluate the ability of the compilation to support partially specified action models.
-In this case instead of learning the action models from scratch the model of half of the actions is given as input of the learning task.
+In this case, instead of learning the action models from scratch, the model of half of the actions is given as input of the learning task.
 Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarizes the obtained results.
@@
 			\begin{tabular}{l|l|l|l|l|l|l|l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf All}\\ \cline{2-9}			
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c}{\bf }\\ \cline{2-9}			
 				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
@@
 				\hline
-				\bf Total & 0.97 & 0.68 & 0.91 & 0.94 & 1.0 & 0.92 & 0.96 & 0.85 \\
+				\bf  & 0.97 & 0.68 & 0.91 & 0.94 & 1.0 & 0.92 & 0.96 & 0.85 \\
 			\end{tabular}
@@
 	\end{footnotesize}
-\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the different domains.}
+\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the different domains starting from partially specified action models.}
 \label{tab:results_plans_partial}
@@
         \end{footnotesize}
-	\caption{\small Planning time, preprocessing time and Plan length.}
+	\caption{\small Planning time, preprocessing time and Plan length when starting from partially specified action models.}
 	\label{tab:time_plans_partial}	
 \end{table}
+
+The trend of the previous experiment is also observed here indicating that more input knowledge produces better quality models and requires less planning times because the learning space is more constrained.
+Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is given as input).
 
@@
 
-Back in the 90's various systems were aimed at learning operators mostly via interaction with the environment.
-The proposal of the LIVE system was to capture and formulate observable features of objects and use them to acquire and refine operators \cite{ShenS89}.
-OBSERVER, on the other hand, updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
-These early works were based on direct lifting of the observed states and supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge.
-The recent work reported in \cite{WalshL08} shows that although learning \strips operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
+Back in the 90's various systems aimed learning operators mostly via interaction with the environment.
+{\sc LIVE} captured and formulated observable features of objects and used them to acquire and refine operators \cite{ShenS89}.
+{\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
+These early works were based on direct lifting of the observed states and supported by exploratory plans or external teachers.%, but none provided a theoretical justification for this second source of knowledge.
 
-Whilst the aforementioned works deal with full state observability, action model learning has also been studied in domains
-where there is partial or missing state observability~\cite{jimenez2012review}.
-The {\sf ARMS} approach~\cite{yang2007learning}, which works when no or partial intermediate states are given, extracts frequent action sets from plans that share a common set of parameters and frequent literal-action pairs with the help of the initial and goal state of the plans.
-The extracted information is used then to define a set of weighted constraints that must hold for all the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver.
-Since complex planning problems may result in a too large MAX-SAT representation to be solved efficiently, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
-Consequently, the output model of {\sf ARMS} may be inconsistent with the examples.
-The approach presented in \cite{amir:alearning:JAIR08} deals with partial observability.
-Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation
-of observations by models of actions through a CNF formula.
+%More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
+
+%Whilst the aforementioned works deal with full state observability,
+Action model learning has also been studied in domains where there is partial or missing state observability.
+{\sf ARMS} works when no or partial intermediate states are given.
+It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
+Consequently, the action models output by {\sf ARMS} may be inconsistent with some of the examples.
+{\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
+Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
 The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
@@
 The formula returned at the end includes all consistent models, which can then be retrieved with additional processing.
-Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
+
+%Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
 An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
@@
 
-==============================================
+\section{Conclusions}
+The paper presented a novel approach for learning \strips\ action models from examples using classical planning.
+The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
+As far as we know, this is the first work on learning action models using an {\em off-the-shelf} classical planner.
+~\citeauthor{stern2017efficient} recently proposed another classical planning compilation for learning action models but it followed a {\em finite domain} representation for the state variables and did not report experimental results since the compilation was formalized but not implemented.
 
-~\cite{stern2017efficient}.
-
-\section{Conclusions}
-The paper presented a novel approach for learning \strips action models from examples exclusively using classical planning.
-The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
-
-The empirical evaluation shows that, when example plans are available, our approach can compute accurate action models from small sets of learning examples and investing very small learning times (less than a second time in most of the domains).
+Our empirical evaluation shows that, when example plans are available, our approach can compute accurate action models from small sets of learning examples and investing very small learning times (less than a second time in most of the domains).
 When action plans are not available our approach is still able to produce action models compliant with the input information.
-In this case, since learning is not constrained by actions it can change the semantics of the operators.
-An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
+In this case, since learning is not constrained by actions it can reformulate operators changing their semantics.
+%An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
 
 The size of the compiled classical planning instances depends on the number of input examples.
-The empirical results show that our approach can generate non-trivial models from very small data sets.
-In addition the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because all the actions for programming the \strips actions preconditions can be done in parallel, in a single planning step, because they do not interact reducing significantly the planning horizon.
-The same happens with all the actions for programming an action effect can also be applied in a single planning step.
+The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
+In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
 
-As far as we know, this is the first work on learning action models using a classical planner.
-This novel approach opens the door to the bootstrapping of action model that is, a planner collecting by its own the necessary data for learning an action model for a given planning task.
 Generating {\em informative} examples for learning planning action models is still a challenging open issue.

### 2017-11-06 doc/icaps-18/strips-learning.tex
addressed comments by eva in the ICAPS18 paper

@@
 \begin{abstract}
-This paper presents a novel approach for learning \strips action models from examples that compiles this particular inductive learning task into classical planning.
+This paper presents a novel approach for learning \strips\ action models from examples that compiles this inductive learning task into classical planning.
 Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states (no intermediate action or state is given).
@@
 At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions.
-Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of automated planning~\cite{kambhampati:modellite:AAAI2007}.
+Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of planning~\cite{kambhampati:modellite:AAAI2007}.
 
@@
 
-The learning of \strips\ action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
+Learning \strips\ action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-The compilation approach is appealing because it opens the door to the bootstrapping of planning action models but also because:
+The compilation approach is appealing itself because it opens the door to the bootstrapping of planning action models but also because:
 \begin{enumerate}
@@
 
-The paper is organized as follows.
-The first section presents the classical planning model, its extension to conditional effects (which is a requirement of the proposed compilation) and formalizes the \strips\ action model (the output of the addressed learning task).
+The first section of the paper presents the classical planning model, its extension to conditional effects (which is a requirement of the proposed compilation) and formalizes the \strips\ action model (the output of the addressed learning task).
 The second section formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
-The third and fourth sections describe our approach for addressing this particular learning task by compiling it into classical planning and how previous knowledge can be introduced in the compilation to constrain the space of possible action models and make learning more practicable.
-Finally, last sections report the data collected during the empirical evaluation of our approach, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
+The third and fourth sections describe our approach for addressing the formalized learning tasks.
+Finally, the last sections report the data collected during the empirical evaluation of our approach, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
 
 \section{Background}
-This section defines the planning models used on this work and the output of the learning task addressed in the paper.
+This section defines the planning models used on this work and the output of the learning tasks addressed in the paper.
 
@@
 $|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
+Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
 
@@
 \begin{itemize}
-\item $\pre(a)$, called {\em preconditions}, that defines the literals that must hold for the action to be applicable.
-\item $\eff^+(a)$, called {\em positive effects}, that defines the fluents set to true by the action application.
-\item $\eff^-(a)$, called {\em negative effects}, that defines the fluents set to false by the application of the action. 
+\item $\pre(a)\subseteq\mathcal{L}(F)$, called {\em preconditions}, that defines the literals that must hold for the action to be applicable.
+\item $\eff^+(a)\subseteq\mathcal{L}(F)$, called {\em positive effects}, that defines the fluents set to true by the action application.
+\item $\eff^-(a)\subseteq\mathcal{L}(F)$, called {\em negative effects}, that defines the fluents set to false by the application of the action. 
 \end{itemize}
@@
 
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
 A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of $\pi$ in $I$.
+We denote with $|\pi|$ the {\em plan length}.
 
@@
 
-The result of applying action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\del(s,a))\cup\add(s,a)\}$ where $del(s,a)\subseteq triggered(s,a)$ and $add(s,a)\subseteq triggered(s,a)$ are the triggered negative and positive effects .
+The result of applying action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered negative and positive effects .
 
@@
 
-Let us define also a new set of fluents $F_{v}$, s.t.
-$F\cap F_v=\emptyset$, that results instantiating $\Psi$ using only {\em variable objects} in $\Omega_v$.
+Let us define also a new set of fluents $F_{v}$, $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only {\em variable objects} in $\Omega_v$.
 This set defines the elements that can appear in an action schema.
@@
 
-This section formalizes more challenging learning tasks, where less input knowledge is available.
+This section formalizes more challenging learning tasks, where less input knowledge is available:
 
-\subsection{Learning from (initial, final) state pairs}
-This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
-No information about the actions in the plans is given however, we assume that the headers of the operators schema are known.
-
+\subsubsection{Learning from (initial, final) state pairs.} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
+No information about the actions in the plans is given.
 This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma}$:
 \begin{itemize}
-\item $\Psi$, the set of predicates that define the abstract state space of a given planning domain.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$, a set of (initial, final) state pairs s.t., for every {\tt\small $1\leq t\leq \tau$}, a pair $\sigma_t=(s_0^t,s_{n}^t)$ comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from a given {\em initial} state $s_0^t$.
+\item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
+\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of labels s.t.
+each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, is an $(initial, final)$ state pair that comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from a given {\em initial} state $s_0^t$.
 \end{itemize}
 
-A solution to the learning task $\Lambda$  is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
+\subsubsection{Learning from labeled plans.}
+Here we augment the amount of provided input knowledge with the actions executed by the observed agent and define the learning task as $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
 
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$ that explain the given states $\Sigma$ using the learned \strips\ model.
+\begin{itemize}
+\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+\end{itemize}
 
-\subsection{Learning from labeled plans}
-Here we augment the amount of provided input knowledge and define the learning task as $\Lambda'=\tup{\Psi,\Sigma,\Pi}$.
-
-Now $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-
-A solution to a learning task $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) that is compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
-
-Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda'$ in the blocksworld (the figure shows the content of $\Psi$, $\Pi$ and $\Sigma$).
-This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of a plan for inverting a four-block tower.
+Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda'$ in the blocksworld.
+This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
 
@@
 \end{subfigure}%
-
  \caption{\small Example of a task for learning a \strips\ action model in the blocksworld.}
@@
 
+\subsubsection{Learning from partially specified action models.}
+We may not require to start learning the \strips\ action models from scratch so we augment here the inputs of the learning task with partially specified operator schemes.
+In this case the leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ where:
+\begin{itemize}
+\item $\Xi_0$ is a partially specified action model in which some preconditions and effects are a priori known.
+\end{itemize}
+
+A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant just with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given labels $\Sigma$ using the learned \strips\ model.
+A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
+Finally a solution to $\Lambda''$ is a set of \strips\ operator schema $\Xi$ compliant as well with the provided partially specified action model $\Xi_0$.
+
 \section{Learning \strips\ action models with planning}
-Our approach for addressing a learning task $\Lambda$ or $\Lambda'$, is compiling it into a classical planning task with conditional effects.
-The intuition behind these compilations is that a solution to the resulting classical planning task is a sequence of actions that:
+Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is compiling it into a classical planning task with conditional effects.
+The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
 \begin{enumerate}
-\item Programs the \strips\ action model $\Xi$. For each $\xi\in\Xi$, the solution plan has a prefix that determines the literals that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ if available).
-For every label $\sigma_t\in \Sigma$, then $\Xi$ is used to produce a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$.
+\item Programs the \strips\ action model $\Xi$.
+The solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Validates the programmed \strips\ action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ and/or $\Xi_0$ if available).
+For every label $\sigma_t\in \Sigma$, the solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
 We call this process the validation of the programmed \strips\ action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
-If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
+%If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
 \end{enumerate}
@@
 same fluents and actions and differ only in the initial state and goals).
-The fluents $F$ are built instantiating the predicates in $\Psi$ with the objects appearing in the inputs.
+The fluents $F$ are built instantiating the predicates in $\Psi$ with the objects appearing in the input labels $\Sigma$.
 Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
@@
 Now we are ready to formalize the compilations.
-We start with $\Lambda$, because this learning task requires fewer input knowledge.
-Given a learning task $\Lambda=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
+We start with $\Lambda$, because this learning task requires less input knowledge.
+Given a learning task $\Lambda=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
 \begin{itemize}
@@
 \end{itemize}
-\item $I_{\Lambda}$, contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
+\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
 \item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
@@
 \begin{enumerate}
-\item The actions for programming an operator schema $\xi\in\Xi$, which includes:
+\item The actions for {\em programming} an operator schema $\xi\in\Xi$, which includes:
 \begin{itemize}
-\item The actions for {\bf removing} a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
+\item Actions for {\bf removing} a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
 
@@
 \hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
-& pre_{f}(\xi), mode_{prog}\},\\
+& mode_{prog}, pre_{f}(\xi)\},\\
 \cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
@@
 
-\item The actions for {\bf adding} a {\em negative} or a {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
+\item Actions for {\bf adding} a {\em negative} or a {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
 
@@
 
-\item The actions for applying an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
-The binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
+\item The actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
+We assume that the operators headers are known so the binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
 the variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
@@
 
-\item The actions for updating the learning example {\tt\small $1\leq t\leq \tau$} where the programmed action model is validated.
+\item The actions for {\em validating} the current learning example {\tt\small $1\leq t\leq \tau$}.
 \begin{small}
@@
 
-\begin{figure}[hbt]
+\begin{figure}
 \begin{scriptsize}
@@
 The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after the preconditions are programmed).
-Furthermore, once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, there is only one way of achieving goals $\{test_t\}$, {\small $1\leq t\leq \tau$}: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$ defined in $\sigma_t$, starting from $s_0^t$ (defined in this same label).
-If this is achieved for all the input examples {\small $1\leq t\leq \tau$} of the learning task, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
+Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
+To solve $P_{\Lambda}$, there is only one way of achieving goals $\{test_t\}$, {\small $1\leq t\leq \tau$}: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
+If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
 \end{small}
@@
 
-\subsection{Example plans}
+\subsection{Labeled plans}
 Here  we extend our compilation to address the learning scenario where, a set of plans is available.
@@
 \begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the $j$ steps of plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current and next plan steps.
-\item $I_{\Lambda'}$ is extended with the fluents from $F_{\Pi}$ that encode the plan $\pi_1\in \Pi$, the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
+Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current plan step and to iterate in the steps of a plan.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with the fluents from $F_{\Pi}$ that encode the plan $\pi_1\in \Pi$, the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
 Goals are the same as in the previous compilation $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
@@
 \begin{enumerate}
-\item The actions for programming the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for applying an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
+\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
 This mechanism forces that these actions are only applied as in the example plans.
-\item The actions for updating the active test have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
+\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
 \begin{small}
 \begin{align*}
-&\{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}},\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
+&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}, \{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
 \end{align*}
@@
 
-\subsection{Previously specified action models}
-Our compilation does not require to start learning \strips\ action models from scratch and accepts partially specified operator schema where some preconditions and effects are a priori known.
-
-The known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda'}$.
+\subsection{Partially specified action models}
+Known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda'}$.
 The corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task for getting the blocksworld action model where the schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are known in advance.
+To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''$ for getting the blocksworld action model where the schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are known in advance.
 This classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld (previously specified operator schemes for $pickup$, $putdown$ and $unstack$ are given), using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-The plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operators.
+The plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operators applying the actions in the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
 
@@
 \begin{figure}
-\begin{small}
-\begin{verbatim}
-     0 : (program_pre_clear_stack_v1)
-     1 : (program_pre_handempty_stack)
-     2 : (program_pre_holding_stack_v2)
-     3 : (program_pre_on_stack_v1_v1)
-     4 : (program_pre_on_stack_v1_v2)
-     5 : (program_pre_on_stack_v2_v1)
-     6 : (program_pre_on_stack_v2_v2)
-     7 : (program_pre_ontable_stack_v1)
-     8 : (program_pre_ontable_stack_v2)
-     9 : (program_eff_clear_stack_v1)
-    10 : (program_eff_clear_stack_v2)
-    11 : (program_eff_handempty_stack)
-    12 : (program_eff_holding_stack_v1)
-    13 : (program_eff_on_stack_v1_v2)
-    14 : (apply_unstack a b i1 i2)
-    15 : (apply_putdown a i2 i3)
-    16 : (apply_unstack b c i3 i4)
-    17 : (apply_stack b a i4 i5)
-    18 : (apply_unstack c d i5 i6)
-    19 : (apply_stack c b i6 i7)
-    20 : (apply_pickup d i7 i8)
-    21 : (apply_stack d c i8 i9)
-    22 : (validate_1)
-\end{verbatim}
-\end{small}
+{\tt\small
+     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
+     01 : (program\_pre\_handempty\_stack)\\
+     02 : (program\_pre\_holding\_stack\_v2)\\
+     03 : (program\_pre\_on\_stack\_v1\_v1)\\
+     04 : (program\_pre\_on\_stack\_v1\_v2)\\
+     05 : (program\_pre\_on\_stack\_v2\_v1)\\
+     06 : (program\_pre\_on\_stack\_v2\_v2)\\
+     07 : (program\_pre\_ontable\_stack\_v1)\\
+     08 : (program\_pre\_ontable\_stack\_v2)\\
+     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
+    10 : (program\_eff\_clear\_stack\_v2)\\
+    11 : (program\_eff\_handempty\_stack)\\
+    12 : (program\_eff\_holding\_stack\_v1)\\
+    13 : (program\_eff\_on\_stack\_v1\_v2)\\
+    {\bf 14} : (apply\_unstack a b i1 i2)\\
+    15 : (apply\_putdown a i2 i3)\\
+    16 : (apply\_unstack b c i3 i4)\\
+    17 : (apply\_stack b a i4 i5)\\
+    18 : (apply\_unstack c d i5 i6)\\
+    19 : (apply\_stack c b i6 i7)\\
+    20 : (apply\_pickup d i7 i8)\\
+    21 : (apply\_stack d c i8 i9)\\
+    {\bf 22} : (validate\_1)
+}
  \caption{\small Example of a plan for programming and validating the operator schema $stack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
@@
 
-Static predicates can be used to constrain the space of possible preconditions by looking at the given set of labels $\Sigma$.
+Static predicates can also constrain the space of possible preconditions by looking at the given set of labels $\Sigma$.
 In particular one can assume that if a precondition $f\in F_v$ (s.t.
 $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ it means that is not possible and then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
-For instance in the {\em zenotravel} domain fluents $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (as well as their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} cannot be compliant with any state in $\Sigma$.
+For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (as well as their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} cannot be compliant with any state in $\Sigma$.
 
-Likewise, static predicates can constrain the space of possible preconditions looking at the given set of example plans.
-In this case, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discarded for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
+Likewise, static predicates can constrain the space of possible preconditions looking at the given set of example plans $\Pi$.
+Fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
 $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
-Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but it does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} it means that the only possible precondition including the static predicate is $pre\_next\_fly\_v1\_v2$.
+Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v1\_v2$.
 
-In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of {\em potential} static predicates from the input examples to the learning task.
+In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of {\em potential} static predicates from the input to the learning task.
 
@@
 \subsection{Experimental setup}
-All the experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
-The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, and that are taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+Experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
+The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
 
@@
 The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
-These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned model and the reference model~\cite{davis2006relationship}.
+These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
 
@@
 \subsubsection{The learning examples.}
-The set of learning examples is fixed for all the experiments so we can evaluate better the effect of the different amount of input knowledge in the quality of the learned models.
+The set of learning examples is fixed for all the experiments so we can evaluate the effect of the amount of input knowledge in the quality of the learned models.
 %All learning examples are generated solving a small problem from the mentioned repository.
 
-\subsection{Learning action models from example plans}
+\subsection{Learning action models from labeled plans}
 Here we assess the performance of our learning approach when addressing learning tasks of the $\Lambda'$ kind.
-Table~\ref{tab:results_plans} summarizes the obtained results, precision ({\bf P}) and recall ({\bf R}) is computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report the averages.
-The last row is also an average report.
+Table~\ref{tab:results_plans} summarizes the obtained results, precision ({\bf P}) and recall ({\bf R}) is computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while last two columns (and last row) report averages.
 
 \begin{table}[hbt!]
+\begin{footnotesize}
 	\begin{center}
 		\resizebox{\columnwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l|l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
 			\hline
@@
 	\end{center}
-\caption{\small {\em Precision} and {\em recall} values of the learned models.}
+\end{footnotesize}        
+\caption{\small {\em Precision} and {\em recall} of the learned models.}
 \label{tab:results_plans}
@@
 
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning tasks that result from our compilation as well as the number of actions in the solutions found.
+Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions found.
 Remarkably all the learning tasks are solved in a few seconds time.
@@
 \subsection{Exploiting static predicates}
-Here we repeat the previous evaluation but now, candidates of static predicates are computed from the learning examples.
+Here we repeat the previous evaluation but now, potential static predicates are computed from the set of labels $\Sigma$.
+The potential {\em static predicates} considered is the set of predicates s.t.
+every predicate instantiation appears untouched in the initial and final states, for every $\sigma_t\in\Sigma$.
+
 The considered static predicates are used to constrain the space of possible action models as explained in the previous section.
-The set of candidate {\em static predicates} considered  here is the set of predicates s.t.
-every predicate instantiation appears the same in the initial and final states, for all the learning examples {\small $1\leq t\leq \tau$}.
-Tables~\ref{tab:results_plans_static} and~\ref{tab:time_plans_static} show the obtained results.
+Tables~\ref{tab:results_plans_static} and~\ref{tab:time_plans_static} summarizes the obtained results.
 
 \begin{table}[hbt!]
+\begin{footnotesize}
 	\begin{center}
 		\resizebox{\columnwidth}{!}{%
-	\begin{tabular}{l|l|l|l|l|l|l|l|l|}
-		& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c}{\bf }\\ \cline{2-9}			
-		& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
 		\hline
@@
 	\end{center}
-\caption{\small The {\em precision} and {\em recall} values achieved by the learned models exploiting static predicates.}
+\end{footnotesize}        
+\caption{\small The {\em precision} and {\em recall} exploiting static predicates.}
 \label{tab:results_plans_static}
@@
         \end{footnotesize}
-	\caption{\small Planning time, preprocessing time and plan length exploiting static predicates.}
+	\caption{\small Planning time and plan length exploiting static predicates.}
 	\label{tab:time_plans_static}	
@@
 
-We can observed that identifying static predicates drives to better models, in particular with larger recall values.
-This fact evidences that many of the missing preconditions in the previous models corresponded to static predicates (there were no inncetive learned them since they always hold).
+We can observe that, identifying static predicates, drives to models with larger {\em recall}.
+This fact evidences that many of the missing preconditions in the previous models corresponded to static predicates (there were no incentive to learn these preconditions as they always hold).
 When static predicates are identified, the resulting compilation is much compact and produces smaller planning/instantiation times.
 Interestingly, one can identify the domains with static predicates by just looking at the plan length.
-In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less actions for programming preconditions are required.
+In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
 
@@
 In this case, instead of learning the action models from scratch, the model of half of the actions is given as input of the learning task.
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarizes the obtained results.
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarizes the obtained for the unknown actions (considering the known action models would mean higher scores since {\em precision} and {\em recall} in the given models is 1.0).
+The {\em Visitall} domain is excluded from this evaluation because it only contains one action schema.
 
@@
 		\resizebox{\columnwidth}{!}{%
-			\begin{tabular}{l|l|l|l|l|l|l|l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c}{\bf }\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
 			\hline
@@
 \end{footnotesize}        
-\caption{\small The {\em precision} and {\em recall} values achieved by the learned models in the different domains starting from partially specified action models.}
+\caption{\small {\em Precision} and {\em recall} in the learned models starting from partially specified actions.}
 \label{tab:results_plans_partial}
@@
         \end{footnotesize}
-	\caption{\small Planning time, preprocessing time and Plan length when starting from partially specified action models.}
+	\caption{\small Planning time and plan length when starting from partially specified action models.}
 	\label{tab:time_plans_partial}	
@@
 
-The trend of the previous experiment is also observed here indicating that more input knowledge produces better quality models and requires less planning times because the learning space is more constrained.
-Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is given as input).
+The trend of the previous experiment is also observed here: more input knowledge produces better models and requires less planning times because the learning space is more constrained.
+Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
 
 \subsection{Learning action models from example states}
-When the set of input planning task is not available, the planner determines also the actions that must satisfying the input labels so actions can reformulated and still be compliant with the learning examples.
+When the set of input plans is not available, the planner determines also the actions that must satisfy the input labels so actions can be reformulated and still be compliant with the learning examples.
 For instance a blocksworld can be learned where the operator {\small\tt stack} is defined with the preconditions and effects of the {\small\tt unstack} operator and vice versa.

### 2017-11-08 doc/icaps-18/strips-learning.tex
first version ICAPS18 paper ready for submission

@@
 
-\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
-{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-{\small Universitat Polit\`ecnica de Val\`encia.}\\
-{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
-{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
+% Commented for blind submission
+%\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
+%{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
+%{\small Universitat Polit\`ecnica de Val\`encia.}\\
+%{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
+%{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
 
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$.
-For instance in the working example the preconditions of the $stack$ schema are reprenseted by the fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to true.
+For instance in the working example the preconditions of the $stack$ schema are represented by the fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to true.
 \item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or they are being validated (already programmed).
@@
 \section{Evaluation}
-This section evaluates our approach for learning \strips\ action models starting from different amounts of available input knowledge.
+This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts of available input knowledge.
+The set of learning examples is fixed for all the experiments so we evaluate the effect of the amount of input knowledge in the quality of the learned models.
+%All learning examples are generated solving a small problem from the mentioned repository.
 
-\subsection{Experimental setup}
+\subsubsection{Setup.}
 Experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
@@
 The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use this SAT-based planner because its ability to deal with planning instances populated with dead-ends.
-In addition, in this planner the actions for programming preconditions can be done in parallel, in a single planning step, given that these actions do not interact.
-Likewise, the actions for programming action effects are also applicable in a single planning step reducing significantly the planning horizon.
+We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
+In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
+Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
 
@@
 Intuitively precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
-Formally $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
+Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
 On the other hand recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
 
-\subsubsection{The learning examples.}
-The set of learning examples is fixed for all the experiments so we can evaluate the effect of the amount of input knowledge in the quality of the learned models.
-%All learning examples are generated solving a small problem from the mentioned repository.
+\subsection{Learning from (initial,final) state pairs}
+Here we assess the performance of our learning approach when addressing learning tasks of the $\Lambda$ kind.
+This is the configuration with the minimum amount of available input knowledge, where input plans are not available so the planner must determine as well the actions that satisfy the input labels.
+Table~\ref{tab:results_states} summarizes the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while last two columns (and the last row) report averages.
 
-\subsection{Learning action models from labeled plans}
-Here we assess the performance of our learning approach when addressing learning tasks of the $\Lambda'$ kind.
-Table~\ref{tab:results_plans} summarizes the obtained results, precision ({\bf P}) and recall ({\bf R}) is computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while last two columns (and last row) report averages.
+\begin{table}
+\begin{footnotesize}
+	\begin{center}
+		\resizebox{\columnwidth}{!}{%
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+			\hline
+                        Blocks & 0.33 & 0.33 & 0.60 & 0.50 & 0.33 & 0.33 & 0.42 & 0.39 \\
+			Driverlog & 1.0 & 0.43 & 0.30 & 1.0 & 1.0 & 0.75 & 0.77 & 0.73 \\
+			Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
+			Floortile & 0.50 & 0.30 & 0.75 & 0.60 & 0.50 & 0.20 & 0.59 & 0.37 \\
+			Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+			Miconic & 1.0 & 0.25 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.75 \\
+			Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
+			Transport & nan & 0.0 & 0.29 & 1.0 & nan & 0.0 & nan & 0.33 \\
+			Visitall & - & - & - & - & - & - & - & - \\
+			Zenotravel & 1.0 & 0.17 & 0.33 & 0.33 & 1.0 & 0.33 & 0.78 & 0.28 \\
+			\hline
+			 & 0.85 & 0.31 & 0.57 & 0.83 & 0.86 & 0.62 & 0.77 & 0.59 \\				
+			\end{tabular}
+		}
+	\end{center}
+\end{footnotesize}        
+\caption{\small {\em Precision} and {\em recall} learning from state pairs.}
+\label{tab:results_states}
+\end{table}
 
-\begin{table}[hbt!]
+\begin{table}
+\begin{footnotesize}
+	\begin{center}
+		\begin{tabular}{l|c|c|c|}			
+			 & Total time & Preprocess & Plan length  \\
+			\hline
+                        Blocks & 0.73 & 0.00 & 69  \\
+			Driverlog & 0.98 & 0.04 & 104 \\
+			Ferry & 0.29 & 0.07 & 110 \\
+			Floortile & 5.68 & 0.21 & 215 \\
+			Gripper & 0.00 & 0.00 & 41 \\
+			Miconic & 0.04 & 0.00 & 47 \\
+			Satellite & 6.62 & 0.12 & 102 \\
+			Transport & 3.12 & 0.14 & 41 \\			
+			Visitall & - & - & - \\
+			Zenotravel & 15.54 & 0.10 & 82 \\
+		\end{tabular}
+	\end{center}
+        \end{footnotesize}
+	\caption{\small Planning times (secs) and plan length.}
+	\label{tab:time_states}	
+\end{table}
+
+In this scenario precision and recall have low values.
+This is not surprising because the learning hypothesis space is so low constrained that actions can be reformulated and still be compliant with the inputs (e.g.
+the blocksworld operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
+Table~\ref{tab:time_states} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
+Values for the {\em Visitall} domain are not reported since the compilation could not be solved within a time bound of 1000 seconds.
+
+\subsection{Learning from labeled plans}
+Here we assess the performance of our learning approach when addressing learning tasks of the $\Lambda'$ kind where labeled plans are available.
+Tables~\ref{tab:results_plans} and ~\ref{tab:time_plans} summarizes the obtained results.
+Precision and recall values are now acceptable, except the preconditions recall that is still low indicating there are many missing preconditions in the learned models.
+All the learning tasks are now solved in a few seconds time.
+
+\begin{table}
 \begin{footnotesize}
@@
 \end{footnotesize}        
-\caption{\small {\em Precision} and {\em recall} of the learned models.}
+\caption{\small {\em Precision} and {\em recall} learning from labeled plans.}
 \label{tab:results_plans}
@@
 
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions found.
-Remarkably all the learning tasks are solved in a few seconds time.
-
-\begin{table}[hbt!]
+\begin{table}
 \begin{footnotesize}
@@
         \end{footnotesize}
-	\caption{\small Planning time, preprocessing time and plan length.}
+	\caption{\small Planning time, preprocessing time and plan length learning from labeled plans.}
 	\label{tab:time_plans}	
@@
 
-\subsection{Exploiting static predicates}
-Here we repeat the previous evaluation but now, potential static predicates are computed from the set of labels $\Sigma$.
+We repeat now the evaluation but exploiting potential static predicates that are computed from the set of labels $\Sigma$.
 The potential {\em static predicates} considered is the set of predicates s.t.
 every predicate instantiation appears untouched in the initial and final states, for every $\sigma_t\in\Sigma$.
-
-The considered static predicates are used to constrain the space of possible action models as explained in the previous section.
+These static predicates are used to constrain the space of possible action models as explained in the previous section.
 Tables~\ref{tab:results_plans_static} and~\ref{tab:time_plans_static} summarizes the obtained results.
 
-\begin{table}[hbt!]
+\begin{table}
 \begin{footnotesize}
@@
 
-\begin{table}[hbt!]
+\begin{table}
 \begin{footnotesize}
@@
 
-We can observe that, identifying static predicates, drives to models with larger {\em recall}.
-This fact evidences that many of the missing preconditions in the previous models corresponded to static predicates (there were no incentive to learn these preconditions as they always hold).
+We can observe that, identifying static predicates, drives to models with larger precondition {\em recall}.
+This fact evidences that many of the missing preconditions corresponded to static predicates, there were no incentive to learn these preconditions as they always hold~\cite{gregory2015domain}.
 When static predicates are identified, the resulting compilation is much compact and produces smaller planning/instantiation times.
-Interestingly, one can identify the domains with static predicates by just looking at the plan length.
+Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
 In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
@@
 In this case, instead of learning the action models from scratch, the model of half of the actions is given as input of the learning task.
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarizes the obtained for the unknown actions (considering the known action models would mean higher scores since {\em precision} and {\em recall} in the given models is 1.0).
-The {\em Visitall} domain is excluded from this evaluation because it only contains one action schema.
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarizes the obtained results for only the {\em unknown} actions (considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in the {\em known} models is 1.0).
+The tables confirm the trend: more input knowledge drives to better models and requires less planning times.
+Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
+{\em Visitall} is excluded from this evaluation because it only contains one action schema.
 
-\begin{table}[hbt!]
+\begin{table}
 \begin{footnotesize}
@@
 \end{table}
-
-The trend of the previous experiment is also observed here: more input knowledge produces better models and requires less planning times because the learning space is more constrained.
-Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
-
-\subsection{Learning action models from example states}
-When the set of input plans is not available, the planner determines also the actions that must satisfy the input labels so actions can be reformulated and still be compliant with the learning examples.
-For instance a blocksworld can be learned where the operator {\small\tt stack} is defined with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
 
@@
 The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
-As far as we know, this is the first work on learning action models using an {\em off-the-shelf} classical planner.
-~\citeauthor{stern2017efficient} recently proposed another classical planning compilation for learning action models but it followed a {\em finite domain} representation for the state variables and did not report experimental results since the compilation was formalized but not implemented.
+As far as we know, this is the first work on learning action models exclusively using an {\em off-the-shelf} classical planner.
+Recently, ~\citeauthor{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
 
-Our empirical evaluation shows that, when example plans are available, our approach can compute accurate action models from small sets of learning examples and investing very small learning times (less than a second time in most of the domains).
-When action plans are not available our approach is still able to produce action models compliant with the input information.
+Our evaluation shows that, when example plans are available, we can compute accurate action models from small sets of learning examples and investing small learning times (less than a second time in most of the domains).
+When action plans are not available, our approach is still able to produce action models compliant with the input information.
 In this case, since learning is not constrained by actions it can reformulate operators changing their semantics.
@@
 
-The size of the compiled classical planning instances depends on the number of input examples.
-The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
+% The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
 In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
 
+The size of the compiled classical planning instances depends on the number of input examples.
 Generating {\em informative} examples for learning planning action models is still a challenging open issue.
@@
 
-\begin{small}
-\subsection*{Acknowledgment}
-Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
+% Commented for blind submission
+%\begin{small}
+%\subsection*{Acknowledgment}
+%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
 Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
-\end{small}
+%\end{small}
 

### 2017-11-08 experiments/scripts/generate_tables_5_6.sh
Added experiments folder with scripts and learned domains

@@ -1,0 +1,12 @@
+#!/bin/bash
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "visitall" "zenotravel")
+
+for i in "${arr[@]}"
+do
+   ../../src/compiler2.py -s ../../benchmarks/icaps18/$i/ empty_domain test plan 0
+
+   echo $i 
+   ../../src/evaluator2.py ../../benchmarks/reference/$i/domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   cat planner_out.log | grep -e 'actions\sin'
+   cat planner_out.log | grep -e 'total\stime'
+done

### 2017-11-08 experiments/scripts/run_empty_domains.sh

@@ -1,0 +1,28 @@
+#!/bin/bash
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "visitall" "zenotravel")
+
+if [ $1 == 0 ]
+then
+   mode="plans"
+elif [ $1 == 2 ]
+then
+   mode="lenplan"
+elif [ $1 == 3 ]
+then
+   mode="minimum"
+else
+   echo wrong input
+   exit
+fi
+
+
+for i in "${arr[@]}"
+do
+   ../../src/compiler2.py -s ../../benchmarks/icaps18/$i/ empty_domain test plan $1
+
+   echo $i 
+   ../../src/evaluator2.py ../../benchmarks/reference/$i/domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   cat planner_out.log | grep -e 'actions\sin'
+   cat planner_out.log | grep -e 'total\stime'
+   cp learned_domain.pddl ../domains/$i/domain.empty.$mode.pddl
+done

### 2017-11-12 doc/icaps-18/strips-learning.tex
updates evaluation results

@@
 			Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
-			Floortile & 0.50 & 0.30 & 0.75 & 0.60 & 0.50 & 0.20 & 0.59 & 0.37 \\
+			Floortile & 0.67 & 0.40 & 0.50 & 0.40 & 1.0 & 0.40 & 0.72 & 0.40 \\
 			Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-			Miconic & 1.0 & 0.25 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.75 \\
+			Miconic & 0.0 & 0.0 & 0.33 & 0.50 & 0.0 & 0.0 & 0.11 & 0.17 \\
 			Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
-			Transport & nan & 0.0 & 0.29 & 1.0 & nan & 0.0 & nan & 0.33 \\
-			Visitall & - & - & - & - & - & - & - & - \\
-			Zenotravel & 1.0 & 0.17 & 0.33 & 0.33 & 1.0 & 0.33 & 0.78 & 0.28 \\
+			Transport & 1.0 & 0.25 & 0.33 & 1.0 & 1.0 & 0.50 & 0.78 & 0.58 \\
+			Zenotravel & - & - & - & - & - & - & - & - \\
+			Grid & - & - & - & - & - & - & - & - \\
 			\hline
-			 & 0.85 & 0.31 & 0.57 & 0.83 & 0.86 & 0.62 & 0.77 & 0.59 \\				
+			 & 0.75 & 0.34 & 0.53 & 0.80 & 0.79 & 0.63 & 0.69 & 0.59 \\				
 			\end{tabular}
@@
 			Driverlog & 0.98 & 0.04 & 104 \\
-			Ferry & 0.29 & 0.07 & 110 \\
-			Floortile & 5.68 & 0.21 & 215 \\
-			Gripper & 0.00 & 0.00 & 41 \\
-			Miconic & 0.04 & 0.00 & 47 \\
-			Satellite & 6.62 & 0.12 & 102 \\
-			Transport & 3.12 & 0.14 & 41 \\			
-			Visitall & - & - & - \\
-			Zenotravel & 15.54 & 0.10 & 82 \\
+			Ferry & 0.09 & 0.00 & 107 \\
+			Floortile & 6.42 & 0.15 & 126 \\
+			Gripper & 0.03 & 0.00 & 47 \\
+			Miconic & 0.04 & 0.00 & 68 \\
+			Satellite & 4.34 & 0.10 & 126 \\
+			Transport & 12.21 & 0.53 & 54 \\			
+			Zenotravel & - & - & - \\
+			Grid & - & - & - \\
 		\end{tabular}
@@
 Table~\ref{tab:time_states} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-Values for the {\em Visitall} domain are not reported since the compilation could not be solved within a time bound of 1000 seconds.
+Values for the {\em Zenotravel} and {\em Grid} domains are not reported since the compilation could not be solved within a time bound of 1000 seconds.
 
@@
 			Ferry & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86 \\
-			Floortile & 0.54 & 0.64 & 0.90 & 0.82 & 1.0 & 0.91 & 0.81 & 0.79 \\
+			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 \\
 			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-			Miconic & 0.75 & 0.33 & 1.0 & 0.75 & 0.75 & 1.0 & 0.83 & 0.69 \\
-			Satellite & 0.67 & 0.29 & 1.0 & 1.0 & 1.0 & 0.75 & 0.89 & 0.68 \\
-			Transport & 1.0 & 0.30 & 0.57 & 0.80 & 1.0 & 0.60 & 0.86 & 0.57 \\
+			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 \\
+			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 \\
+			Transport & 1.0 & 0.40 & 0.71 & 1.0 & 1.0 & 0.80 & 0.90 & 0.73 \\
 			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-			Zenotravel & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 \\
+			Zenotravel & 1.0 & 0.36 & 0.70 & 1.0 & 1.0 & 0.71 & 0.90 & 0.69 \\
+			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+			Grid & 0.53 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.69 & 0.78 \\
 			\hline
-			\bf  & 0.88 & 0.51 & 0.89 & 0.91 & 0.98 & 0.89 & 0.91 & 0.77 \\
+			\bf  & 0.85 & 0.51 & 0.83 & 0.92 & 0.95 & 0.92 & 0.88 & 0.78 \\
 		\end{tabular}
@@
 			Driverlog & 0.10 & 0.06 & 127 \\
-			Ferry & 0.20 & 0.10 & 58 \\
-			Floortile & 2.15 & 1.21 & 170 \\
-			Gripper & 0.01 & 0.00 & 45 \\
-			Miconic & 0.10 & 0.09 & 54 \\
+			Ferry & 0.06 & 0.03 & 54 \\
+			Floortile & 2.42 & 1.64 & 168 \\
+			Gripper & 0.03 & 0.01 & 43 \\
+			Miconic & 0.06 & 0.03 & 57 \\
 			Satellite & 0.20 & 0.14 & 67 \\
-			Transport & 0.20 & 0.18 & 60 \\
-			Visitall & 0.43 & 0.39 & 37 \\
-			Zenotravel & 0.37 & 0.35 & 70			
+			Transport & 0.70 & 0.64 & 74 \\
+			Visitall & 0.21 & 0.15 & 40 \\
+			Zenotravel & 2.21 & 2.17 & 83 \\
+			Hanoi & 0.12 & 0.06 & 48 \\
+			Grid & 4.54 & 4.51 & 86 \\	
 		\end{tabular}
@@
 		Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-		Driverlog & 1.0 & 0.5 & 0.75 & 0.86 & 1.0 & 0.71 & 0.91 & 0.69 \\
+		Driverlog & 1.0 & 0.50 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.69 \\
 		Ferry & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86 \\
-		Floortile & 0.71 & 0.78 & 0.90 & 0.82 & 1.0 & 0.91 & 0.87 & 0.83 \\
+		Floortile & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74 \\
 		Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-		Miconic & 0.88 & 0.78 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.84 \\
-		Satellite & 0.85 & 0.79 & 0.83 & 1.0 & 1.0 & 0.75 & 0.89 & 0.85 \\
-		Transport & 1.0 & 0.60 & 0.80 & 0.80 & 1.0 & 0.60 & 0.93 & 0.67 \\
+		Miconic & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88 \\
+		Satellite & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80 \\
+		Transport & 1.0 & 0.70 & 0.71 & 1.0 & 1.0 & 0.80 & 0.90 & 0.83 \\
 		Visitall & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-		Zenotravel & 1.0 & 0.64 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.74 \\
+		Zenotravel & 1.0 & 0.71 & 0.70 & 1.0 & 1.0 & 0.71 & 0.90 & 0.81 \\
+		Hanoi & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92 \\
+		Grid & 0.73 & 0.65 & 1.0 & 0.86 & 0.75 & 0.86 & 0.83 & 0.79 \\
 		\hline
-		\bf  & 0.92 & 0.73 & 0.90 & 0.91 & 0.98 & 0.87 & 0.93 & 0.84 \\
+		\bf  & 0.89 & 0.73 & 0.92 & 0.93 & 0.96 & 0.89 & 0.92 & 0.85 \\
 	\end{tabular}
@@
 			Driverlog & 0.12 & 0.07 & 102 \\
-			Ferry & 0.25 & 0.15 & 58 \\
-			Floortile & 0.73 & 0.43 & 80 \\
-			Gripper & 0.01 & 0.00 & 45 \\
-			Miconic & 0.06 & 0.04 & 40 \\
+			Ferry & 0.04 & 0.03 & 54 \\
+			Floortile & 0.67 & 0.57 & 77 \\
+			Gripper & 0.01 & 0.00 & 43 \\
+			Miconic & 0.04 & 0.00 & 41 \\
 			Satellite & 0.18 & 0.12 & 60 \\
-			Transport & 0.18 & 0.18 & 44 \\
-			Visitall & 0.43 & 0.42 & 33 \\
-			Zenotravel & 0.26 & 0.25 & 53
+			Transport & 0.56 & 0.51 & 60 \\
+			Visitall & 0.17 & 0.15 & 36 \\
+			Zenotravel & 1.03 & 1.00 & 65 \\
+			Hanoi & 0.09 & 0.06 & 39 \\
+			Grid & 3.39 & 3.35 & 72 \\
 		\end{tabular}
@@
 				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.91 \\
-				Ferry & 1.0 & 0.67 & 0.5 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
+				Ferry & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
 				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
 				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-				Miconic & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
+				Miconic & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
 				Satellite & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
-				Transport & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 0.50 & 1.0 & 0.67 \\
-%				Visitall & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Zenotravel & 1.0 & 0.83 & 0.67 & 0.67 & 1.0 & 1.0 & 0.89 & 0.83 \\
+				Transport & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
+				Zenotravel & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.78 \\
+				Grid & 0.78 & 0.58 & 1.0 & 1.0 & 0.75 & 0.75 & 0.84 & 0.78 \\
 				\hline
-				\bf  & 0.97 & 0.68 & 0.91 & 0.94 & 1.0 & 0.92 & 0.96 & 0.85 \\
+				\bf  & 0.95 & 0.71 & 1.0 & 0.98 & 0.98 & 0.92 & 0.98 & 0.87 \\
 			\end{tabular}
@@
 			Driverlog & 0.06 & 0.03 & 59 \\
-			Ferry & 0.20 & 0.14 & 50 \\
-			Floortile & 0.54 & 0.45 & 58 \\
-			Gripper & 0.01 & 0.00 & 37 \\
-			Miconic & 0.06 & 0.04 & 33  \\
-			Satellite & 0.17 & 0.15 & 48 \\
-			Transport & 0.12 & 0.10 & 33 \\
-			Zenotravel & 0.26 & 0.26 & 33
+			Ferry & 0.06 & 0.03 & 45 \\
+			Floortile & 0.43 & 0.42 & 55 \\
+			Gripper & 0.03 & 0.01 & 35 \\
+			Miconic & 0.03 & 0.01 & 34  \\
+			Satellite & 0.14 & 0.14 & 47 \\
+			Transport & 0.32 & 0.32 & 42 \\
+			Zenotravel & 1.10 & 1.09 & 44 \\
+			Grid & 3.18 & 3.15 & 52 \\
 		\end{tabular}

### 2017-11-12 doc/icaps-18/strips-learning.tex
English proofreading by Eva

@@
 
-The result of applying action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered negative and positive effects .
+The result of applying action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered negative and positive effects, respectively.
 
@@
 
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects, $\Omega\cap\Omega_v=\emptyset$, that represents {\em variable names} and that is bound by the maximum arity of an action in a given planning frame.
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects denoted as {\em variable names}, $\Omega\cap\Omega_v=\emptyset$, and that is bound by the maximum arity of an action in a given planning frame.
 For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters each.
@@
 Now we are ready to formalize the compilations.
-We start with $\Lambda$, because this learning task requires less input knowledge.
+We start with $\Lambda$, because this learning task requires the less input knowledge.
 Given a learning task $\Lambda=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\cup\\
+\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\\
 &\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
@@
 \subsection{Labeled plans}
-Here  we extend our compilation to address the learning scenario where, a set of plans is available.
-Given a learning task $\Lambda'=\tup{\Psi,\Pi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
+Here we extend our compilation to address the learning scenario where a set of plans is available.
+Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
 \begin{itemize}
 \item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current plan step and to iterate in the steps of a plan.
+Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
 \item $I_{\Lambda'}$ extends $I_{\Lambda}$ with the fluents from $F_{\Pi}$ that encode the plan $\pi_1\in \Pi$, the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
@@
 \item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and an extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
 This mechanism forces that these actions are only applied as in the example plans.
@@
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''$ for getting the blocksworld action model where the schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are known in advance.
+To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the blocksworld action model where the schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} belong to $\Xi_0$ and are fully known in advance.
 This classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld (previously specified operator schemes for $pickup$, $putdown$ and $unstack$ are given), using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
@@
 
-In the extreme, when a fully specified \strips\ action model is given, the compilation validates whether an observed plan follows the given model.
+In the extreme, when a fully specified \strips\ action model $\Xi$ is given, the compilation validates whether an observed plan follows the given model.
 In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips\ action model is {\em valid} for the given set of examples.
@@
 \begin{itemize}
-\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable} objects $\Omega_v$, can be discarded for every $\xi\in\Xi$.
+\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable objects} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
 \item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
@@
 Here we assess the performance of our learning approach when addressing learning tasks of the $\Lambda'$ kind where labeled plans are available.
-Tables~\ref{tab:results_plans} and ~\ref{tab:time_plans} summarizes the obtained results.
-Precision and recall values are now acceptable, except the preconditions recall that is still low indicating there are many missing preconditions in the learned models.
+Tables~\ref{tab:results_plans} and ~\ref{tab:time_plans} summarize the obtained results.
+Precision and recall values are now acceptable, except for the still low value of the preconditions recall, indicating there are many missing preconditions in the learned models.
 All the learning tasks are now solved in a few seconds time.
@@
 The potential {\em static predicates} considered is the set of predicates s.t.
-every predicate instantiation appears untouched in the initial and final states, for every $\sigma_t\in\Sigma$.
+every predicate instantiation appears unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
 These static predicates are used to constrain the space of possible action models as explained in the previous section.
-Tables~\ref{tab:results_plans_static} and~\ref{tab:time_plans_static} summarizes the obtained results.
+Tables~\ref{tab:results_plans_static} and~\ref{tab:time_plans_static} summarize the obtained results.
 
@@
 
-We can observe that, identifying static predicates, drives to models with larger precondition {\em recall}.
-This fact evidences that many of the missing preconditions corresponded to static predicates, there were no incentive to learn these preconditions as they always hold~\cite{gregory2015domain}.
+We can observe that identifying static predicates drives to models with larger precondition {\em recall}.
+This fact evidences that since many of the missing preconditions corresponded to static predicates, there were no incentive to learn these preconditions as they always hold~\cite{gregory2015domain}.
 When static predicates are identified, the resulting compilation is much compact and produces smaller planning/instantiation times.
@@
 
-Here we evaluate the ability of the compilation to support partially specified action models.
+Here we evaluate the ability of the compilation to support partially specified action models (including identification of static predicates).
 In this case, instead of learning the action models from scratch, the model of half of the actions is given as input of the learning task.
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarizes the obtained results for only the {\em unknown} actions (considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in the {\em known} models is 1.0).
-The tables confirm the trend: more input knowledge drives to better models and requires less planning times.
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results for only the {\em unknown} actions (considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in the {\em known} models is 1.0).
+The tables confirm the trend: more input knowledge drives to better models and requires less planning time.
 Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).

### 2017-11-13 doc/icaps-18/strips-learning.tex
Modifications in the last section of the paper

@@
 
-The approach for learning planing models that most likely works with the least information possible is {\sf LOCM} ~\cite{cresswell2013acquiring}, where the only required input are the example training plans without the need for providing the initial and goal state information.
-{\sf LOCM} exploits assumptions about the kind of domain model it has to generate.
-Particularly, it assumes a domain consists of a collection of objects (called \emph{sorts}) where each object has a defined set of states that can be captured by a parameterized Finite State Machine (FSM).
-The induction of FSM draws upon assumptions like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence.
-The intuitive assumptions of {\sf LOCM} yield a learning model heavily reliant on the kind of domain structure and not capable of properly deriving domain theories in which the state of a sort is subject to different state machines.
-This limitation is later overcome by {\sf LOCM2} by forming separate state machines for a sort, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
-The last contribution of the {\sf LOCM} family, {\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{GregoryC16}), addresses the problem of inducing static predicates in action models by identifying the set of minimal static relations associated to an operator.
-Because {\sf LOCM} models induce similar models for domains with similar structure (e.g.
-Blocksworld and Freecell), they face problems at identifying models that do not have static relations (Blocksworld) and those that do have static predicates (Freecell).
+{\sf LOCM} only requires the example training plas as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
+This makes {\sf LOCM} be most likely the learning approach that works with the least information possible.
+The lack of available information is addressed by LOCM by exploting assumptions about the kind of domain model it has to generate.
+Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
+The intuitive assumptions of {\sf LOCM}
+%, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence, 
+yield a learning model heavily reliant on the kind of domain structure.
+The inability of {\sf LOCM} to properly derive domain theories where the state of a sort is subject to different FSMs is later overcome by {\sf LOCM2} by forming separate FSMs, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
+{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain}), the last contribution of the {\sf LOCM} family, addresses the problem of inducing static predicates.
+Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
+{\em blocksworld} and {\em freecell}).
 In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
+
+Compiling the learning task into a a classical planning task is a general and flexible solution as it allows to accommodate different amounts of input knowledge.
+This scheme opens up a path for solving further tasks other than learning action models or plan validation -- when a fully specified action model $\Xi$ is given.
+Thus, by replacing the learning examples of $\Pi$ by sequences of states (observations), noisy or missing fluents could be introduced in the states, and so the learning task $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi}$, where $\mathcal{O}$ is the noisy or partial observations, would amount to a plan recognition task where the domain theory is given~\cite{SohrabiRU16}.
+Furthermore, the compilation scheme can be extended to learn other type of generative models that comprise partially-specified solution plans like HTN models or full-specified solutions like FSM or behaviour trees.
 

### 2017-11-13 experiments/scripts/generate_tables_1_2.sh
updated scripts

@@ -1,12 +1,14 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "visitall" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "zenotravel" "grid")
+
+ulimit -t 1000
 
 for i in "${arr[@]}"
 do
    ../../src/compiler2.py -s ../../benchmarks/icaps18/$i/ partial_domain test plan 3 
 
    echo $i 
-   ../../src/evaluator2.py -p ../../benchmarks/icaps18/$i/partial_domain.pddl ../../benchmarks/reference/$i/domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   ../../src/evaluator2.py -p ../../benchmarks/icaps18/$i/partial_domain.pddl ../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
    cat planner_out.log | grep -e 'actions\sin'
    cat planner_out.log | grep -e 'total\stime'
 done

### 2017-11-13 experiments/scripts/generate_tables_3_4.sh

@@ -1,12 +1,14 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "visitall" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "visitall" "zenotravel" "hanoi" "grid")
+
+ulimit -t 1000
 
 for i in "${arr[@]}"
 do
    ../../src/compiler2.py ../../benchmarks/icaps18/$i/ empty_domain test plan 0
 
    echo $i 
-   ../../src/evaluator2.py ../../benchmarks/reference/$i/domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   ../../src/evaluator2.py ../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
    cat planner_out.log | grep -e 'actions\sin'
    cat planner_out.log | grep -e 'total\stime'
 done

### 2017-11-13 experiments/scripts/generate_tables_5_6.sh

@@ -1,12 +1,14 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "visitall" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "visitall" "zenotravel" "hanoi" "grid")
+
+ulimit -t 1000
 
 for i in "${arr[@]}"
 do
    ../../src/compiler2.py -s ../../benchmarks/icaps18/$i/ empty_domain test plan 0
 
    echo $i 
-   ../../src/evaluator2.py ../../benchmarks/reference/$i/domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   ../../src/evaluator2.py ../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
    cat planner_out.log | grep -e 'actions\sin'
    cat planner_out.log | grep -e 'total\stime'
 done

### 2017-11-13 experiments/scripts/generate_tables_7_8.sh

@@ -1,12 +1,14 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "visitall" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "gripper" "miconic" "satellite" "transport" "zenotravel" "hanoi" "grid")
+
+ulimit -t 1000
 
 for i in "${arr[@]}"
 do
    ../../src/compiler2.py -s ../../benchmarks/icaps18/$i/ partial_domain test plan 0 
 
    echo $i 
-   ../../src/evaluator2.py -p ../../benchmarks/icaps18/$i/partial_domain.pddl ../../benchmarks/reference/$i/domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   ../../src/evaluator2.py -p ../../benchmarks/icaps18/$i/partial_domain.pddl ../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
    cat planner_out.log | grep -e 'actions\sin'
    cat planner_out.log | grep -e 'total\stime'
 done

### 2017-11-13 doc/icaps-18/strips-learning.tex
full version paper icaps18 with reorganized results

@@
 Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states (no intermediate action or state is given).
-What is more, the compilation accepts partially specified action models and can be used to validate whether certain observations of plan executions follow a given \strips\ action model, even if this model is not fully specified.
+What is more, the compilation accepts partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
 \end{abstract}
@@
 \section{Introduction}
-Besides {\em plan synthesis}~\cite{ghallab2004automated,geffner:book:2013}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions.
-Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of planning~\cite{kambhampati:modellite:AAAI2007}.
+Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
+At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
+Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
 
 On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML to the learning of \strips\ action models~\cite{fikes1971strips}, the vanilla action model for planning, is not straightforward though:
+The application of inductive ML to the learning of \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
 \begin{itemize}
-\item The inputs to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects.
-The input for learning planning action models are observations of plan executions (where each plan possibly has different length).
-\item The traditional output of ML algorithms is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
+\item The {\em input} to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects.
+The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
+\item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
 When learning \strips\ action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions.
@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-The compilation approach is appealing itself because it opens the door to the bootstrapping of planning action models but also because:
+The compilation approach is appealing by itself because opens the door to the bootstrapping of planning action models but also because:
 \begin{enumerate}
@@
 \item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
-In the extreme, the compilation could be used to validate whether observed plan executions are valid for a given \strips\ action model.
+In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
 \end{enumerate}
 
-The first section of the paper presents the classical planning model, its extension to conditional effects (which is a requirement of the proposed compilation) and formalizes the \strips\ action model (the output of the addressed learning task).
-The second section formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
-The third and fourth sections describe our approach for addressing the formalized learning tasks.
-Finally, the last sections report the data collected during the empirical evaluation of our approach, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
+The second section of the paper formalizes the classical planning model, its extension to {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the output of the addressed learning task).
+The third section formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
+The fourth and fifth sections describe our approach for addressing the formalized learning tasks.
+Finally, the last sections report the data collected in a empirical evaluation, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
 
@@
 
-A {\em state} $s$ is a total assignment of values to fluents, i.e.
+A {\em state} $s$ is a full assignment of values to fluents, i.e.
 $|s|=|F|$, so the size of the state space is $2^{|F|}$.
@@
 \begin{itemize}
-\item $\pre(a)\subseteq\mathcal{L}(F)$, called {\em preconditions}, that defines the literals that must hold for the action to be applicable.
-\item $\eff^+(a)\subseteq\mathcal{L}(F)$, called {\em positive effects}, that defines the fluents set to true by the action application.
-\item $\eff^-(a)\subseteq\mathcal{L}(F)$, called {\em negative effects}, that defines the fluents set to false by the application of the action.
+\item $\pre(a)\subseteq\mathcal{L}(F)$, called {\em preconditions}, the literals that must hold for the action $a\in A$ to be applicable.
+\item $\eff^+(a)\subseteq\mathcal{L}(F)$, called {\em positive effects}, that defines the fluents set to true by the application of the action $a\in A$.
+\item $\eff^-(a)\subseteq\mathcal{L}(F)$, called {\em negative effects}, that defines the fluents set to false by the action application.
 \end{itemize}
-We say that an action $a\in A$ is {\em applicable} in state $s$ iff $\pre(a)\subseteq s$, and the result of applying $a$ in $s$ is the {\em successor state} $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
+The result of applying $a$ in $s$ is the {\em successor state} $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
 
@@
 A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+We denote with $|\pi|$ the {\em plan length}.
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of $\pi$ in $I$.
-We denote with $|\pi|$ the {\em plan length}.
 
 \subsection{Classical planning with conditional effects}
-Our approach for leaning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
+Our approach for learning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
 Conditional effects allow us to compactly define actions whose effects depend on the current state.
-Many classical planners cope with conditional effects without compiling them away.
-In fact, supporting conditional effects is a requirement of IPC-2014~\cite{vallati:IPC:AIM2015} and IPC-2018.
+Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
 
-Now an action $a\in A$ has a set of literals $\pre(a)\in\mathcal{L}(F)$, called the {\em precondition}, and a set of {\em conditional effects} $\cond(a)$.
+An action $a\in A$ has now a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
 Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
 
-An action $a\in A$ is {\em applicable} in state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} is
+An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} are the effects whose conditions hold in $s$:
 \[
@@
 \]
-i.e.~effects whose conditions hold in $s$.
 
-The result of applying action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered negative and positive effects, respectively.
+The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered {\em negative} and {\em positive} effects, respectively.
 
-\subsection{The \strips\ action schema and the {\em variable} objects}
+\subsection{\strips\ action schemes and {\em variable name} objects}
 This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-Figure~\ref{fig:stack} shows the schema that corresponds to the {\em stack} action from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+Figure~\ref{fig:stack} shows the schema, coded in PDDL, for the {\em stack} action from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
 
@@
 (:action stack
-  :parameters (?x1 ?x2 - object)
-  :precondition (and (holding ?x1) (clear ?x2))
-  :effect (and (not (holding ?x1))
-               (not (clear ?x2))
-               (handempty) (clear ?x1)
-               (on ?x1 ?x2)))
+  :parameters (?v1 ?v2 - object)
+  :precondition (and (holding ?v1) (clear ?v2))
+  :effect (and (not (holding ?v1))
+               (not (clear ?v2))
+               (handempty) (clear ?v1)
+               (on ?v1 ?v2)))
 \end{verbatim}
 \end{footnotesize}
- \caption{\small Example of a \strips\ operator schema that corresponds to the {\em stack} action from the {\em blocksworld} represented in PDDL.}
+ \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
 \label{fig:stack}
@@
 
-To formalize the output of the learning task, we assume that there exists a set of {\em predicates} $\Psi$ and that fluents $F$ are instantiated from these predicates, as in PDDL.
+To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
 Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of objects $\Omega$, the set of fluents $F$ is then induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
+Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
 $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-Likewise, we assume that actions $a\in A$ are instantiated from \strips\ operator schemes.
 
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects denoted as {\em variable names}, $\Omega\cap\Omega_v=\emptyset$, and that is bound by the maximum arity of an action in a given planning frame.
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects $\Omega\cap\Omega_v=\emptyset$, denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
 For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters each.
 
-Let us define also a new set of fluents $F_{v}$, $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only {\em variable objects} in $\Omega_v$.
-This set defines the elements that can appear in an action schema.
-In the blocksworld $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$),on($v_2,v_2$)\}}.
+Let us also define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and that defines the elements that can appear in an action schema.
+For instance, in the blocksworld, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
-We are now ready to define $\Xi$, a set of operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ such that:
+Finally, we assume that actions $a\in A$ are instantiated from \strips\ operator schemes $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$, an enumeration of the variable objects bound by the operator arity.
-The headers for a 4-operator blocksworld are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
+For instance, the headers for a four-operator blocksworld are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
 \item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
@@
 from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
-In this case, because intermediate states are available, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Preconditions are derived lifting the minimal set of literals that appears in all the pre-states that correspond to the same operator schema.
+When any intermediate state is available, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions.
 
@@
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of labels s.t.
-each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, is an $(initial, final)$ state pair that comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from a given {\em initial} state $s_0^t$.
+\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
+Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
 \end{itemize}
@@
 \subsubsection{Learning from labeled plans.}
-Here we augment the amount of provided input knowledge with the actions executed by the observed agent and define the learning task as $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
+Here we augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
 
 \begin{itemize}
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in the state $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
 \end{itemize}
@@
 This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
+
+\subsubsection{Learning from partially specified action models.}
+We may not require to start learning from scratch so we augment the learning input with partially specified operator schemes. This leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
+\begin{itemize}
+\item $\Xi_0$ is a partially specified action model in which some preconditions and effects are a priori known.
+\end{itemize}
+
+A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant just with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given labels $\Sigma$ using the learned \strips\ model.
+A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
+Finally a solution to $\Lambda''$ is a set of \strips\ operator schema $\Xi$ compliant as well with the provided partially specified action model $\Xi_0$.
 
@@
 \end{subfigure}%
- \caption{\small Example of a task for learning a \strips\ action model in the blocksworld.}
+ \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
 \label{fig:lexample}
 \end{figure}
-
-\subsubsection{Learning from partially specified action models.}
-We may not require to start learning the \strips\ action models from scratch so we augment here the inputs of the learning task with partially specified operator schemes.
-In this case the leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ where:
-\begin{itemize}
-\item $\Xi_0$ is a partially specified action model in which some preconditions and effects are a priori known.
-\end{itemize}
-
-A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant just with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given labels $\Sigma$ using the learned \strips\ model.
-A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (with one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
-Finally a solution to $\Lambda''$ is a set of \strips\ operator schema $\Xi$ compliant as well with the provided partially specified action model $\Xi_0$.
 
@@
 \item Programs the \strips\ action model $\Xi$.
-The solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ using the given input knowledge (the set of labels $\Sigma$, and $\Pi$ and/or $\Xi_0$ if available).
-For every label $\sigma_t\in \Sigma$, the solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
+A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (the labels $\Sigma$, and $\Pi$ and/or $\Xi_0$ if available).
+For every label $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
 We call this process the validation of the programmed \strips\ action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
@@
 
-To formalize our compilations we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
-same fluents and actions and differ only in the initial state and goals).
-The fluents $F$ are built instantiating the predicates in $\Psi$ with the objects appearing in the input labels $\Sigma$.
+To formalize our compilation we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
+same fluents and actions but differ in the initial state and goals).
+Fluents $F$ are built instantiating the predicates in $\Psi$ with the objects appearing in the input labels $\Sigma$.
 Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
@@
 Now we are ready to formalize the compilations.
-We start with $\Lambda$, because this learning task requires the less input knowledge.
+We start with $\Lambda$, because it requires less input knowledge.
 Given a learning task $\Lambda=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$.
-For instance in the working example the preconditions of the $stack$ schema are represented by the fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to true.
-\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or they are being validated (already programmed).
-\item Fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the learning example where the programmed action model is being validated.
+For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$}.
+\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is being validated.
 \end{itemize}
-\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first learning example), every $pre_f(\xi)\in F_{\Lambda}$ (our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect) and $mode_{prog}$ set to true.
+\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
+Our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect.
 \item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
-\item $A_{\Lambda}$ contains actions of three different kinds:
+\item $A_{\Lambda}$ contains actions of three kinds:
 \begin{enumerate}
-\item The actions for {\em programming} an operator schema $\xi\in\Xi$, which includes:
+\item Actions for {\em programming} an operator schema $\xi\in\Xi$:
 \begin{itemize}
@@
 
-\item Actions for {\bf adding} a {\em negative} or a {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
+\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
 
@@
 
-\item The actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
-We assume that the operators headers are known so the binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
-the variables in $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
+\item Actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
+We assume operators headers are known so the binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
+variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
 Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
@@
 
-\item The actions for {\em validating} the current learning example {\tt\small $1\leq t\leq \tau$}.
+\item Actions for {\em validating} the learning example {\tt\small $1\leq t\leq \tau$}.
 \begin{small}
@@
 \end{itemize}
+
+\begin{lemma}
+Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
+The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after preconditions are programmed).
+Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
+To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
+If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
+\end{small}
+\end{proof}
+
+The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model.
 
@@
 \end{scriptsize}
- \caption{\small Action for applying an already programmed operator schema $stack$ as encoded in PDDL.}
+ \caption{\small Action for applying an already programmed schema $stack$ as encoded in PDDL (implications coded as disjunctions).}
 \label{fig:compilation}
@@
 
-\begin{lemma}
-Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
-\end{lemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after the preconditions are programmed).
-Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, there is only one way of achieving goals $\{test_t\}$, {\small $1\leq t\leq \tau$}: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
-If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
-\end{small}
-\end{proof}
-
-The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model.
-
 \section{Constraining the learning hypothesis space with additional input knowledge}
 \label{sec:Constraining}
-Here we show that further input knowledge can be used to constrain the space of the possible action models and make the learning of \strips\ action models more practicable.
+Here we show that further input knowledge can be used to constrain the space of possible action models and make the learning of \strips\ action models more practicable.
 
 \subsection{Labeled plans}
-Here we extend our compilation to address the learning scenario where a set of plans is available.
+We extend the compilation to consider labeled plans.
 Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
@@
 Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with the fluents from $F_{\Pi}$ that encode the plan $\pi_1\in \Pi$, the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating where to start validating the programmed action model.
-Goals are the same as in the previous compilation $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the plan step where the action model is validated.
+Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
 \item With respect to $A_{\Lambda'}$.
@@
 \item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and an extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
 This mechanism forces that these actions are only applied as in the example plans.
@@
 Known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda'}$.
-The corresponding programming actions ($\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$) become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
+The corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the blocksworld action model where the schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} belong to $\Xi_0$ and are fully known in advance.
-This classical plan programs and validates the operator schema {\tt\small stack} from the blocksworld (previously specified operator schemes for $pickup$, $putdown$ and $unstack$ are given), using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-The plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are actions for validating the programmed operators applying the actions in the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
+To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the blocksworld action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
+This plan programs and validates the operator schema {\tt\small stack} from the blocksworld, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+Plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are the actions for validating the programmed operators following the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
 
-In the extreme, when a fully specified \strips\ action model $\Xi$ is given, the compilation validates whether an observed plan follows the given model.
-In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips\ action model is {\em valid} for the given set of examples.
+In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
+In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips\ action model is {\em valid} for the given examples.
 If $P_{\Lambda'}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
@@
 }
- \caption{\small Example of a plan for programming and validating the operator schema $stack$ using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+ \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi_1$ and label $\sigma_1$ (shown in Figure~\ref{fig:lexample}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
 \label{fig:plan}
@@
 A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action schema~\cite{fox:TIM:JAIR1998}.
-Therefore, one can get rid of the mechanism for programming these predicates as the effect of any action schema while keeping the compilation complete.
+Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
 Formally, given a static predicate $p$:
@@
 Static predicates can also constrain the space of possible preconditions by looking at the given set of labels $\Sigma$.
-In particular one can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ it means that is not possible and then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
-For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (as well as their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} cannot be compliant with any state in $\Sigma$.
+One can assume that if a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
+For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\Sigma$.
 
-Likewise, static predicates can constrain the space of possible preconditions looking at the given set of example plans $\Pi$.
+Likewise, static predicates can constrain the space of possible preconditions looking at the given example plans.
 Fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
 $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
-Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v1\_v2$.
+Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
 
@@
 
+\begin{table*}
+		\resizebox{\textwidth}{!}{%
+		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
+& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+			\hline
+			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Driverlog & 1.0 & 0.43 & 0.67 & 0.86 & 1.0 & 0.86 & 0.89 & 0.71 & 1.0 & 0.50 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.69\\
+			Ferry & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86 & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86\\
+			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
+			Grid & 0.53 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.69 & 0.78 & 0.73 & 0.65 & 1.0 & 0.86 & 0.75 & 0.86 & 0.83 & 0.79 \\                        
+			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
+			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\                        
+			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
+			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
+			Transport & 1.0 & 0.40 & 0.71 & 1.0 & 1.0 & 0.80 & 0.90 & 0.73 & 1.0 & 0.70 & 0.71 & 1.0 & 1.0 & 0.80 & 0.90 & 0.83\\
+			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Zenotravel & 1.0 & 0.36 & 0.70 & 1.0 & 1.0 & 0.71 & 0.90 & 0.69 &1.0 & 0.71 & 0.70 & 1.0 & 1.0 & 0.71 & 0.90 & 0.81\\
+			\hline
+			\bf  & 0.85 & 0.51 & 0.83 & 0.92 & 0.95 & 0.92 & 0.88 & 0.78 & 0.89 & 0.73 & 0.92 & 0.93 & 0.96 & 0.89 & 0.92 & 0.85\\
+		\end{tabular}
+	}
+\caption{\small {\em Precision} and {\em recall} obtained learning from labeled plans without (left) and with (right) static predicates.}
+\label{tab:results_plans}
+\end{table*}
+
+\begin{table}
+\begin{scriptsize}
+	\begin{center}
+		\begin{tabular}{l|c|c|c||c|c|c|}
+                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
+			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
+			\hline
+			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
+			Driverlog & 0.10 & 0.06 & 127 & 0.12 & 0.07 & 102 \\
+			Ferry & 0.06 & 0.03 & 54 & 0.04 & 0.03 & 54 \\
+			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
+			Grid & 4.54 & 4.51 & 86 & 3.39 & 3.35 & 72 \\ 
+			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
+                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
+                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
+			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
+			Transport & 0.70 & 0.64 & 74 & 0.56 & 0.51 & 60 \\
+			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
+			Zenotravel & 2.21 & 2.17 & 83 & 1.03 & 1.00 & 65 \\			
+		\end{tabular}
+	\end{center}
+        \end{scriptsize}
+	\caption{\small Total planning time, preprocessing time and plan length learning from labeled plans without/with static predicates.}
+	\label{tab:time_plans}	
+\end{table}
+
 \section{Evaluation}
 This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts of available input knowledge.
-The set of learning examples is fixed for all the experiments so we evaluate the effect of the amount of input knowledge in the quality of the learned models.
-%All learning examples are generated solving a small problem from the mentioned repository.
 
 \subsubsection{Setup.}
-Experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
 The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
+All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
 
-\subsubsection{The planner.}
+\subsubsection{Planner.}
 The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
@@
 
-\subsubsection{The evaluation metrics.}
+\subsubsection{Metrics.}
 The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
@@
 
+\subsection{Learning from labeled plans}
+We start assessing the performance of our learning approach when addressing learning tasks of the kind $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, where labeled plans are available, and repeat the evaluation but exploiting potential static predicates that are computed from $\Sigma$.
+The potential {\em static predicates} considered is the set of predicates s.t.
+every predicate instantiation appears unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
+These static predicates are used to constrain the space of possible action models as explained in the previous section.
+
+Table~\ref{tab:results_plans} summarizes the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while last two columns of each setting (and the last row) report averages values.
+
+We can observe that identifying static predicates drives to models with better precondition {\em recall}.
+This fact evidences that many of the missing preconditions corresponded to static predicates because there were no incentive to learn them as they always hold~\cite{gregory2015domain}.
+When static predicates are identified, the resulting compilation is much compact and produces smaller planning/instantiation times.
+
+Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
+All the learning tasks are solved in a few seconds time and interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
+In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
+
+\subsection{Learning from partially specified action models}
+
+We evaluate now the ability of the compilation to support partially specified action models, in other words, the performance of our approach when addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
+In this evaluation, instead of learning the action models from scratch, the model of half of the actions is given in $\Xi_0$ as input of the learning task .
+
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, including the identification of static predicates, but for only the {\em unknown} actions.
+Considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in the {\em known} models is 1.0.
+
+The reported results confirm the previous trend: more input knowledge drives to better models and requires less planning time.
+Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
+{\em Visitall} and {\em Hanoi} are excluded from this evaluation because it only contains one action schema.
+
+Remarkably the overall precision is now $0.98$ which means that the content of the learned models is very reliable.
+The value for the precision is a bit lower $0.87$ indicating that still the learned models are missing some information.
+In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
+
 \subsection{Learning from (initial,final) state pairs}
-Here we assess the performance of our learning approach when addressing learning tasks of the $\Lambda$ kind.
-This is the configuration with the minimum amount of available input knowledge, where input plans are not available so the planner must determine as well the actions that satisfy the input labels.
-Table~\ref{tab:results_states} summarizes the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while last two columns (and the last row) report averages.
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\resizebox{\columnwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-            Blocks & 0.33 & 0.33 & 0.60 & 0.50 & 0.33 & 0.33 & 0.42 & 0.39 \\
-            Driverlog & 1.0 & 0.43 & 0.30 & 1.0 & 1.0 & 0.75 & 0.77 & 0.73 \\
-            Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
-            Floortile & 0.67 & 0.40 & 0.50 & 0.40 & 1.0 & 0.40 & 0.72 & 0.40 \\
-            Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-            Miconic & 0.0 & 0.0 & 0.33 & 0.50 & 0.0 & 0.0 & 0.11 & 0.17 \\
-            Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
-            Transport & 1.0 & 0.25 & 0.33 & 1.0 & 1.0 & 0.50 & 0.78 & 0.58 \\
-            Zenotravel & - & - & - & - & - & - & - & - \\
-            Grid & - & - & - & - & - & - & - & - \\
-            \hline
-            & 0.75 & 0.34 & 0.53 & 0.80 & 0.79 & 0.63 & 0.69 & 0.59 \\			
-			\end{tabular}
-		}
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} learning from state pairs.}
-\label{tab:results_states}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-			\hline
-            Blocks & 0.73 & 0.00 & 69  \\
-            Driverlog & 0.98 & 0.04 & 104 \\
-            Ferry & 0.09 & 0.00 & 107 \\
-            Floortile & 6.42 & 0.15 & 126 \\
-            Gripper & 0.03 & 0.00 & 47 \\
-            Miconic & 0.04 & 0.00 & 68 \\
-            Satellite & 4.34 & 0.10 & 126 \\
-            Transport & 12.21 & 0.53 & 54 \\			
-            Zenotravel & - & - & - \\
-            Grid & - & - & - \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Planning times (secs) and plan length.}
-	\label{tab:time_states}	
-\end{table}
+Here we assess the performance of our learning approach when addressing learning tasks of the  $\Lambda=\tup{\Psi,\Sigma}$ kind.
+This is the learning configuration with the minimum amount of available input knowledge, where input plans are not available so the planner must determine as well the actions that satisfy the input labels.
+Table~\ref{tab:results_states} summarizes the obtained results when static predicates and partially specified action models are available.
 
@@
 the blocksworld operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
+We tried to minimize this effect with additional input knowledge (static predicates and partially specified action models) but still the achieved results are below the results obtained when learning from labeled plans where no other input knowledge was provided.
+
 Table~\ref{tab:time_states} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
 Values for the {\em Zenotravel} and {\em Grid} domains are not reported since the compilation could not be solved within a time bound of 1000 seconds.
-
-\subsection{Learning from labeled plans}
-Here we assess the performance of our learning approach when addressing learning tasks of the $\Lambda'$ kind where labeled plans are available.
-Tables~\ref{tab:results_plans} and ~\ref{tab:time_plans} summarize the obtained results.
-Precision and recall values are now acceptable, except for the still low value of the preconditions recall, indicating there are many missing preconditions in the learned models.
-All the learning tasks are now solved in a few seconds time.
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\resizebox{\columnwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-			Driverlog & 1.0 & 0.43 & 0.67 & 0.86 & 1.0 & 0.86 & 0.89 & 0.71 \\
-			Ferry & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86 \\
-			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 \\
-			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 \\
-			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 \\
-			Transport & 1.0 & 0.40 & 0.71 & 1.0 & 1.0 & 0.80 & 0.90 & 0.73 \\
-			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-			Zenotravel & 1.0 & 0.36 & 0.70 & 1.0 & 1.0 & 0.71 & 0.90 & 0.69 \\
-			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-			Grid & 0.53 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.69 & 0.78 \\
-			\hline
-			\bf  & 0.85 & 0.51 & 0.83 & 0.92 & 0.95 & 0.92 & 0.88 & 0.78 \\
-		\end{tabular}
-	}
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} learning from labeled plans.}
-\label{tab:results_plans}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-			\hline
-			Blocks & 0.04 & 0.00 & 72 \\
-			Driverlog & 0.10 & 0.06 & 127 \\
-			Ferry & 0.06 & 0.03 & 54 \\
-			Floortile & 2.42 & 1.64 & 168 \\
-			Gripper & 0.03 & 0.01 & 43 \\
-			Miconic & 0.06 & 0.03 & 57 \\
-			Satellite & 0.20 & 0.14 & 67 \\
-			Transport & 0.70 & 0.64 & 74 \\
-			Visitall & 0.21 & 0.15 & 40 \\
-			Zenotravel & 2.21 & 2.17 & 83 \\
-			Hanoi & 0.12 & 0.06 & 48 \\
-			Grid & 4.54 & 4.51 & 86 \\			
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Planning time, preprocessing time and plan length learning from labeled plans.}
-	\label{tab:time_plans}	
-\end{table}
-
-We repeat now the evaluation but exploiting potential static predicates that are computed from the set of labels $\Sigma$.
-The potential {\em static predicates} considered is the set of predicates s.t.
-every predicate instantiation appears unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
-These static predicates are used to constrain the space of possible action models as explained in the previous section.
-Tables~\ref{tab:results_plans_static} and~\ref{tab:time_plans_static} summarize the obtained results.
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\resizebox{\columnwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-		\hline
-		Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-		Driverlog & 1.0 & 0.50 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.69 \\
-		Ferry & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86 \\
-		Floortile & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74 \\
-		Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-		Miconic & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88 \\
-		Satellite & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80 \\
-		Transport & 1.0 & 0.70 & 0.71 & 1.0 & 1.0 & 0.80 & 0.90 & 0.83 \\
-		Visitall & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-		Zenotravel & 1.0 & 0.71 & 0.70 & 1.0 & 1.0 & 0.71 & 0.90 & 0.81 \\
-		Hanoi & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92 \\
-		Grid & 0.73 & 0.65 & 1.0 & 0.86 & 0.75 & 0.86 & 0.83 & 0.79 \\
-		\hline
-		\bf  & 0.89 & 0.73 & 0.92 & 0.93 & 0.96 & 0.89 & 0.92 & 0.85 \\
-	\end{tabular}
-}
-	\end{center}
-\end{footnotesize}
-\caption{\small The {\em precision} and {\em recall} exploiting static predicates.}
-\label{tab:results_plans_static}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-			\hline
-			Blocks & 0.03 & 0.00 & 72 \\
-			Driverlog & 0.12 & 0.07 & 102 \\
-			Ferry & 0.04 & 0.03 & 54 \\
-			Floortile & 0.67 & 0.57 & 77 \\
-			Gripper & 0.01 & 0.00 & 43 \\
-			Miconic & 0.04 & 0.00 & 41 \\
-			Satellite & 0.18 & 0.12 & 60 \\
-			Transport & 0.56 & 0.51 & 60 \\
-			Visitall & 0.17 & 0.15 & 36 \\
-			Zenotravel & 1.03 & 1.00 & 65 \\
-			Hanoi & 0.09 & 0.06 & 39 \\
-			Grid & 3.39 & 3.35 & 72 \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Planning time and plan length exploiting static predicates.}
-	\label{tab:time_plans_static}	
-\end{table}
-
-We can observe that identifying static predicates drives to models with larger precondition {\em recall}.
-This fact evidences that since many of the missing preconditions corresponded to static predicates, there were no incentive to learn these preconditions as they always hold~\cite{gregory2015domain}.
-When static predicates are identified, the resulting compilation is much compact and produces smaller planning/instantiation times.
-Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
-
-\subsection{Learning from partially specified action models}
-
-Here we evaluate the ability of the compilation to support partially specified action models (including identification of static predicates).
-In this case, instead of learning the action models from scratch, the model of half of the actions is given as input of the learning task.
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results for only the {\em unknown} actions (considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in the {\em known} models is 1.0).
-The tables confirm the trend: more input knowledge drives to better models and requires less planning time.
-Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
-{\em Visitall} is excluded from this evaluation because it only contains one action schema.
 
@@
 
+\begin{table}
+\begin{footnotesize}
+	\begin{center}
+		\resizebox{\columnwidth}{!}{%
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+			\hline
+            Blocks & 0.33 & 0.33 & 0.60 & 0.50 & 0.33 & 0.33 & 0.42 & 0.39 \\
+            Driverlog & 1.0 & 0.43 & 0.30 & 1.0 & 1.0 & 0.75 & 0.77 & 0.73 \\
+            Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
+            Floortile & 0.67 & 0.40 & 0.50 & 0.40 & 1.0 & 0.40 & 0.72 & 0.40 \\
+            Grid & - & - & - & - & - & - & - & - \\            
+            Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+            Miconic & 0.0 & 0.0 & 0.33 & 0.50 & 0.0 & 0.0 & 0.11 & 0.17 \\
+            Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
+            Transport & 1.0 & 0.25 & 0.33 & 1.0 & 1.0 & 0.50 & 0.78 & 0.58 \\
+            Zenotravel & - & - & - & - & - & - & - & - \\
+            \hline
+            & 0.75 & 0.34 & 0.53 & 0.80 & 0.79 & 0.63 & 0.69 & 0.59 \\			
+			\end{tabular}
+		}
+	\end{center}
+\end{footnotesize}
+\caption{\small {\em Precision} and {\em recall} learning from state pairs.}
+\label{tab:results_states}
+\end{table}
+
+\begin{table}
+\begin{footnotesize}
+	\begin{center}
+		\begin{tabular}{l|c|c|c|}			
+			 & Total time & Preprocess & Plan length  \\
+			\hline
+            Blocks & 0.73 & 0.00 & 69  \\
+            Driverlog & 0.98 & 0.04 & 104 \\
+            Ferry & 0.09 & 0.00 & 107 \\
+            Floortile & 6.42 & 0.15 & 126 \\
+            Grid & - & - & - \\            
+            Gripper & 0.03 & 0.00 & 47 \\
+            Miconic & 0.04 & 0.00 & 68 \\
+            Satellite & 4.34 & 0.10 & 126 \\
+            Transport & 12.21 & 0.53 & 54 \\			
+            Zenotravel & - & - & - \\
+		\end{tabular}
+	\end{center}
+        \end{footnotesize}
+	\caption{\small Planning times (secs) and plan length.}
+	\label{tab:time_states}	
+\end{table}
+
 \section{Conclusions}
@@
 The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
-As far as we know, this is the first work on learning action models exclusively using an {\em off-the-shelf} classical planner.
-Recently, ~\citeauthor{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
+In addition we introduce the {\em precision} and {\em recall} metrics, widely used in ML, to the evaluation of planning action models.
+These metrics allow to separately assess the correctness and completeness of the learned models.
 
-Our evaluation shows that, when example plans are available, we can compute accurate action models from small sets of learning examples and investing small learning times (less than a second time in most of the domains).
+As far as we know, this is the first work on learning action models exclusively using an {\em off-the-shelf} classical planner and evaluated over a wide range of different domains.
+Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
+
+Our evaluation shows that, when example plans are available, we can compute accurate action models from small sets of learning examples (only five learning examples per domain) and investing small learning times (less than a second time in most of the domains).
 When action plans are not available, our approach is still able to produce action models compliant with the input information.
-In this case, since learning is not constrained by actions it can reformulate operators changing their semantics.
+In this case, since learning is not constrained by actions it can reformulate operators changing their semantics and making hard their comparison with a reference model.
 %An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.

### 2017-11-15 experiments/icaps18/scripts/generate_tables_1_2_right.sh
compiler2.py: fixed an issue with predicate fitness in typed domains

@@ -5,10 +5,10 @@ ulimit -t 1000
 
 for i in "${arr[@]}"
 do
-   ../../src/compiler2.py -s ../../benchmarks/icaps18/$i/ empty_domain test plan 0
+   ../../../src/compiler2.py -s ../../../benchmarks/icaps18/$i/ empty_domain test plan 0
 
    echo $i 
-   ../../src/evaluator2.py ../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   ../../../src/evaluator2.py ../../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../../benchmarks/icaps18/$i/test-1.pddl
    cat planner_out.log | grep -e 'actions\sin'
    cat planner_out.log | grep -e 'total\stime'
 done

### 2017-11-16 doc/icaps-18/strips-learning.tex
Some more changes by Eva

@@
 Table~\ref{tab:results_plans} summarizes the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while last two columns of each setting (and the last row) report averages values.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
 
@@
 
-We evaluate now the ability of the compilation to support partially specified action models, in other words, the performance of our approach when addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
+We evaluate now the ability of the compilation to support partially specified action models; that is, when addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
 In this evaluation, instead of learning the action models from scratch, the model of half of the actions is given in $\Xi_0$ as input of the learning task .
 
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, including the identification of static predicates, but for only the {\em unknown} actions.
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, including the identification of static predicates, but only for the {\em unknown} actions.
 Considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in the {\em known} models is 1.0.
@@
 Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
-{\em Visitall} and {\em Hanoi} are excluded from this evaluation because it only contains one action schema.
+{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
 
-Remarkably the overall precision is now $0.98$ which means that the content of the learned models is very reliable.
-The value for the precision is a bit lower $0.87$ indicating that still the learned models are missing some information.
-In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
+Remarkably, the overall precision is now $0.98$ which means that the content of the learned models is very reliable.
+The recall value, 0.87, is a bit lower indicating that the learned models are still missing some information.
+Particularly, preconditions are again the component more difficult to be fully learned.
+We must also note that an error in this task has a greater impact than in task $\Lambda'$ because the evaluation is done over half of the actions of the domain (e.g.
+the recall of preconditions in domains like {\em Floortile} or {\em Gripper}).
+
+%In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
 
 \subsection{Learning from (initial,final) state pairs}
-Here we assess the performance of our learning approach when addressing learning tasks of the  $\Lambda=\tup{\Psi,\Sigma}$ kind.
-This is the learning configuration with the minimum amount of available input knowledge, where input plans are not available so the planner must determine as well the actions that satisfy the input labels.
-Table~\ref{tab:results_states} summarizes the obtained results when static predicates and partially specified action models are available.
+Here we assess the performance of our learning approach when addressing learning tasks of the kind $\Lambda=\tup{\Psi,\Sigma}$.
+This is the learning configuration with the minimum amount of available input knowledge, where input plans are not available, so the planner must determine as well the actions that satisfy the input labels.
+Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the obtained results.
+Values for the {\em Zenotravel} and {\em Grid} domains are not reported since the compilation could not be solved within a time bound of 1000 seconds.
 
-In this scenario precision and recall have low values.
+In this scenario, we used static predicates and partially specified action models so the particular task that is being solved is $\Lambda=\tup{\Psi,\Sigma,\emptyset,\Xi_0}$.
+As we can observe in Table~\ref{tab:results_states}, precision and recall have low values.
 This is not surprising because the learning hypothesis space is so low constrained that actions can be reformulated and still be compliant with the inputs (e.g.
@@
 
-Table~\ref{tab:time_states} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
+%Table~\ref{tab:time_states} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
 Values for the {\em Zenotravel} and {\em Grid} domains are not reported since the compilation could not be solved within a time bound of 1000 seconds.
@@
 				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
+                Grid & 0.78 & 0.58 & 1.0 & 1.0 & 0.75 & 0.75 & 0.84 & 0.78 \\
 				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
@@
 				Zenotravel & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.78 \\
-				Grid & 0.78 & 0.58 & 1.0 & 1.0 & 0.75 & 0.75 & 0.84 & 0.78 \\
 				\hline
@@
 			Floortile & 0.43 & 0.42 & 55 \\
+            Grid & 3.18 & 3.15 & 52 \\
 			Gripper & 0.03 & 0.01 & 35 \\
@@
 			Zenotravel & 1.10 & 1.09 & 44 \\
-			Grid & 3.18 & 3.15 & 52 \\
 		\end{tabular}
@@
 It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
-Consequently, the action models output by {\sf ARMS} may be inconsistent with some of the examples.
+In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately and so it may output an inconsistent model.
 {\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
@@
 Thus, by replacing the learning examples of $\Pi$ by sequences of states (observations), noisy or missing fluents could be introduced in the states, and so the learning task $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi}$, where $\mathcal{O}$ is the noisy or partial observations, would amount to a plan recognition task where the domain theory is given~\cite{SohrabiRU16}.
-Furthermore, the compilation scheme can be extended to learn other type of generative models that comprise partially-specified solution plans like HTN models or full-specified solutions like FSM or behaviour trees.
+Furthermore, we can extend this scheme to learn other type of generative models via compilation into a planning task whose solution requires less search effort than a classical planning task or no search at all (e.g.
+learning hierarchical models like HTN or behaviour trees).
+
+%Furthermore, the compilation scheme can be extended to learn other type of generative models that comprise partially-specified solution plans like HTN models or full-specified solutions like FSM or behaviour trees.
 

### 2017-11-16 doc/icaps-18/strips-learning.tex
Updated results after last fix

@@
 			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Driverlog & 1.0 & 0.43 & 0.67 & 0.86 & 1.0 & 0.86 & 0.89 & 0.71 & 1.0 & 0.50 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.69\\
-			Ferry & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86 & 0.80 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93 & 0.86\\
+			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
+			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
 			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
-			Grid & 0.53 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.69 & 0.78 & 0.73 & 0.65 & 1.0 & 0.86 & 0.75 & 0.86 & 0.83 & 0.79 \\
+			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
 			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
@@
 			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
-			Transport & 1.0 & 0.40 & 0.71 & 1.0 & 1.0 & 0.80 & 0.90 & 0.73 & 1.0 & 0.70 & 0.71 & 1.0 & 1.0 & 0.80 & 0.90 & 0.83\\
+			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
 			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Zenotravel & 1.0 & 0.36 & 0.70 & 1.0 & 1.0 & 0.71 & 0.90 & 0.69 &1.0 & 0.71 & 0.70 & 1.0 & 1.0 & 0.71 & 0.90 & 0.81\\
+			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
 			\hline
-			\bf  & 0.85 & 0.51 & 0.83 & 0.92 & 0.95 & 0.92 & 0.88 & 0.78 & 0.89 & 0.73 & 0.92 & 0.93 & 0.96 & 0.89 & 0.92 & 0.85\\
+			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
 		\end{tabular}
@@
 			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
-			Driverlog & 0.10 & 0.06 & 127 & 0.12 & 0.07 & 102 \\
-			Ferry & 0.06 & 0.03 & 54 & 0.04 & 0.03 & 54 \\
+			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
+			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
 			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
-			Grid & 4.54 & 4.51 & 86 & 3.39 & 3.35 & 72 \\
+			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
 			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
@@
 			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
-			Transport & 0.70 & 0.64 & 74 & 0.56 & 0.51 & 60 \\
+			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
 			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
-			Zenotravel & 2.21 & 2.17 & 83 & 1.03 & 1.00 & 65 \\			
+			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
 		\end{tabular}
@@
 				Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.91 \\
+				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.90 \\
 				Ferry & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
 				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
-                Grid & 0.78 & 0.58 & 1.0 & 1.0 & 0.75 & 0.75 & 0.84 & 0.78 \\
+                Grid & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.84 & 0.78 \\
 				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
@@
 				\hline
-				\bf  & 0.95 & 0.71 & 1.0 & 0.98 & 0.98 & 0.92 & 0.98 & 0.87 \\
+				\bf  & 0.98 & 0.71 & 1.0 & 0.98 & 1.0 & 0.95 & 0.98 & 0.87 \\
 			\end{tabular}
@@
 \end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} in the learned models starting from partially specified actions.}
+\caption{\small {\em Precision} and {\em recall} in the learned models starting from partially specified action models.}
 \label{tab:results_plans_partial}
@@
 			Blocks & 0.07 & 0.01 & 54  \\
-			Driverlog & 0.06 & 0.03 & 59 \\
+			Driverlog & 0.03 & 0.01 & 40 \\
 			Ferry & 0.06 & 0.03 & 45 \\
 			Floortile & 0.43 & 0.42 & 55 \\
-            Grid & 3.18 & 3.15 & 52 \\
+            Grid & 3.12 & 3.07 & 53 \\
 			Gripper & 0.03 & 0.01 & 35 \\
@@
 			Satellite & 0.14 & 0.14 & 47 \\
-			Transport & 0.32 & 0.32 & 42 \\
-			Zenotravel & 1.10 & 1.09 & 44 \\
+			Transport & 0.23 & 0.21 & 37 \\
+			Zenotravel & 0.90 & 0.89 & 40 \\
 		\end{tabular}
@@
 			\hline
-            Blocks & 0.33 & 0.33 & 0.60 & 0.50 & 0.33 & 0.33 & 0.42 & 0.39 \\
-            Driverlog & 1.0 & 0.43 & 0.30 & 1.0 & 1.0 & 0.75 & 0.77 & 0.73 \\
+            Blocks & 0.33 & 0.33 & 0.75 & 0.50 & 0.33 & 0.33 & 0.47 & 0.39 \\
+            Driverlog & 1.0 & 0.29 & 0.33 & 0.67 & 1.0 & 0.50 & 0.78 & 0.48 \\
             Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
@@
             Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
-            Transport & 1.0 & 0.25 & 0.33 & 1.0 & 1.0 & 0.50 & 0.78 & 0.58 \\
+            Transport & 0.0 & 0.0 & 0.25 & 0.5 & 0.0 & 0.0 & 0.08 & 0.17 \\
             Zenotravel & - & - & - & - & - & - & - & - \\
             \hline
-            & 0.75 & 0.34 & 0.53 & 0.80 & 0.79 & 0.63 & 0.69 & 0.59 \\			
+            & 0.63 & 0.29 & 0.54 & 0.70 & 0.67 & 0.53 & 0.61 & 0.51 \\			
 			\end{tabular}
@@
 			\hline
-            Blocks & 0.73 & 0.00 & 69  \\
-            Driverlog & 0.98 & 0.04 & 104 \\
-            Ferry & 0.09 & 0.00 & 107 \\
+            Blocks & 2.14 & 0.00 & 58  \\
+            Driverlog & 0.09 & 0.00 & 88 \\
+            Ferry & 0.17 & 0.01 & 65 \\
             Floortile & 6.42 & 0.15 & 126 \\
@@
             Satellite & 4.34 & 0.10 & 126 \\
-            Transport & 12.21 & 0.53 & 54 \\			
+            Transport & 2.57 & 0.21 & 47 \\			
             Zenotravel & - & - & - \\

### 2017-11-16 doc/icaps-18/strips-learning.tex
paper icaps18 ready for submission

@@
 \title{Learning \strips\ action models with classical planning}
+\author{\#39}
 
@@
 		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
-& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
+& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
                           \hline
@@
 
+\subsubsection{Reproducibility.}
+We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
+
 \subsubsection{Planner.}
@@
 
-Intuitively precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
+Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
 Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
-On the other hand recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
+Recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
 
 \subsection{Learning from labeled plans}
-We start assessing the performance of our learning approach when addressing learning tasks of the kind $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, where labeled plans are available, and repeat the evaluation but exploiting potential static predicates that are computed from $\Sigma$.
-The potential {\em static predicates} considered is the set of predicates s.t.
+We start evaluating our approach at $\Lambda'=\tup{\Psi,\Sigma,\Pi}$ learning tasks, where {\em labeled plans} are available, and repeat the evaluation but exploiting potential static predicates that are computed from $\Sigma$.
+The potential {\em static predicates} considered here is the set of predicates s.t.
 every predicate instantiation appears unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
-These static predicates are used to constrain the space of possible action models as explained in the previous section.
+Static predicates are used to constrain the space of possible action models as explained in the previous section.
 
@@
 Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
-
 We can observe that identifying static predicates drives to models with better precondition {\em recall}.
 This fact evidences that many of the missing preconditions corresponded to static predicates because there were no incentive to learn them as they always hold~\cite{gregory2015domain}.
-When static predicates are identified, the resulting compilation is much compact and produces smaller planning/instantiation times.
 
@@
 In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
+When static predicates are identified, the resulting compilation is also much compact and produces smaller planning/instantiation times.
 
@@
 
-We evaluate now the ability of the compilation to support partially specified action models; that is, when addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
-In this evaluation, instead of learning the action models from scratch, the model of half of the actions is given in $\Xi_0$ as input of the learning task .
+We evaluate now the ability of our approach to support partially specified action models; that is, when addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
+In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task .
 
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, including the identification of static predicates, but only for the {\em unknown} actions.
-Considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in the {\em known} models is 1.0.
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, including the identification of static predicates.
+We report only the precision and recall for the {\em unknown} actions (considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in these models is 1.0).
+We must also note that errors in $\Lambda''$ tasks have greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is just done over half of the actions (e.g.
+the preconditions recall for the {\em Floortile} or {\em Gripper} domains).
 
-The reported results confirm the previous trend: more input knowledge drives to better models and requires less planning time.
+Remarkably, the overall precision is now $0.98$ which means that the content of the learned models is highly reliable.
+Recall is a lower, 0.87, indicating that the learned models still miss some information.
+Particularly, preconditions are again the component more difficult to be fully learned.
+Overall the reported results confirm the previous trend: more input knowledge drives to better models and requires less planning time.
 Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
 {\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
-
-Remarkably, the overall precision is now $0.98$ which means that the content of the learned models is very reliable.
-The recall value, 0.87, is a bit lower indicating that the learned models are still missing some information.
-Particularly, preconditions are again the component more difficult to be fully learned.
-We must also note that an error in this task has a greater impact than in task $\Lambda'$ because the evaluation is done over half of the actions of the domain (e.g.
-the recall of preconditions in domains like {\em Floortile} or {\em Gripper}).
 
@@
 \subsection{Learning from (initial,final) state pairs}
-Here we assess the performance of our learning approach when addressing learning tasks of the kind $\Lambda=\tup{\Psi,\Sigma}$.
-This is the learning configuration with the minimum amount of available input knowledge, where input plans are not available, so the planner must determine as well the actions that satisfy the input labels.
-Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the obtained results.
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported since the compilation could not be solved within a time bound of 1000 seconds.
+Finally we evaluate our approach at learning tasks where input plans are not available, so the planner must not only compute the action models but also the plans that satisfy the input labels.
+Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the obtained results using static predicates and partially specified action models so the inputs to this learning task are $\Psi$ ,$\Sigma$ and $\Xi_0$.
 
-In this scenario, we used static predicates and partially specified action models so the particular task that is being solved is $\Lambda=\tup{\Psi,\Sigma,\emptyset,\Xi_0}$.
-As we can observe in Table~\ref{tab:results_states}, precision and recall have low values.
-This is not surprising because the learning hypothesis space is so low constrained that actions can be reformulated and still be compliant with the inputs (e.g.
-the blocksworld operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
-We tried to minimize this effect with additional input knowledge (static predicates and partially specified action models) but still the achieved results are below the results obtained when learning from labeled plans where no other input knowledge was provided.
+Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} could not solved the corresponding planning tasks within a 1000 secs time bound.
+Precision and recall have now low values.
+The learning hypothesis space is now so low constrained that actions can be reformulated and still be compliant with the inputs (e.g.
+the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
+We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) but still the achieved scores are below the scores obtained when learning from labeled plans (despite no other input knowledge was provided).
 
@@
 \end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} in the learned models starting from partially specified action models.}
+\caption{\small {\em Precision} and {\em recall} learning from partially specified action models.}
 \label{tab:results_plans_partial}
@@
         \end{footnotesize}
-	\caption{\small Planning time and plan length when starting from partially specified action models.}
+	\caption{\small Time and plan length learning from partially specified action models.}
 	\label{tab:time_plans_partial}	
@@
 {\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
-These early works were based on direct lifting of the observed states and supported by exploratory plans or external teachers.%, but none provided a theoretical justification for this second source of knowledge.
+These early works were based on lifting the observed states supported by exploratory plans or external teachers.%, but none provided a theoretical justification for this second source of knowledge.
 
@@
 Action model learning has also been studied in domains where there is partial or missing state observability.
-{\sf ARMS} works when no or partial intermediate states are given.
+{\sf ARMS} works when no partial intermediate state is given.
 It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
-In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately and so it may output an inconsistent model.
+In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
+%and so it may output an inconsistent model
 {\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
@@
 
-{\sf LOCM} only requires the example training plas as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
+{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
 This makes {\sf LOCM} be most likely the learning approach that works with the least information possible.
-The lack of available information is addressed by LOCM by exploting assumptions about the kind of domain model it has to generate.
+The lack of available information is addressed by LOCM by exploiting assumptions about the kind of domain model it has to generate.
 Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
@@
 
-Compiling the learning task into a a classical planning task is a general and flexible solution as it allows to accommodate different amounts of input knowledge.
-This scheme opens up a path for solving further tasks other than learning action models or plan validation -- when a fully specified action model $\Xi$ is given.
-Thus, by replacing the learning examples of $\Pi$ by sequences of states (observations), noisy or missing fluents could be introduced in the states, and so the learning task $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi}$, where $\mathcal{O}$ is the noisy or partial observations, would amount to a plan recognition task where the domain theory is given~\cite{SohrabiRU16}.
-Furthermore, we can extend this scheme to learn other type of generative models via compilation into a planning task whose solution requires less search effort than a classical planning task or no search at all (e.g.
-learning hierarchical models like HTN or behaviour trees).
+Compiling the learning of action models into classical planning is a general and flexible approach that allows to accommodate different amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
+For instance, the example plans in $\Pi$ could be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), so learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ could also be addressed.
+Furthermore, our approach seems extensible to learning other types of generative models (e.g.
+hierarchical models like HTN or behaviour trees), that can be more appealing than \strips\ models, since using them to compute planning solutions requires less search effort.
 
@@
         \end{footnotesize}
-	\caption{\small Planning times (secs) and plan length.}
+	\caption{\small Time and plan length when learning from state pairs.}
 	\label{tab:time_states}	
@@
 \section{Conclusions}
-The paper presented a novel approach for learning \strips\ action models from examples using classical planning.
+We presented a novel approach for learning \strips\ action models from examples using classical planning.
 The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
-In addition we introduce the {\em precision} and {\em recall} metrics, widely used in ML, to the evaluation of planning action models.
-These metrics allow to separately assess the correctness and completeness of the learned models.
+We also introduced {\em precision} and {\em recall} metrics, widely used in ML, to the evaluation of planning action models.
+These metrics separately assess the correctness and completeness of the learned models making easier the identification of flaws.
 
@@
 When action plans are not available, our approach is still able to produce action models compliant with the input information.
-In this case, since learning is not constrained by actions it can reformulate operators changing their semantics and making hard their comparison with a reference model.
+In this case, since learning is not constrained by actions it can reformulate operators changing their semantics and making tricky their comparison with a reference model.
 %An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
@@
 
-The size of the compiled classical planning instances depends on the number of input examples.
-Generating {\em informative} examples for learning planning action models is still a challenging open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, and often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques that autonomously collect the learning examples.
+The size of the compiled classical planning instances depend on the number of input examples.
+Generating {\em informative} examples for learning planning action models is an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
+The combination of such exploration techniques with our learning approach is an intriguing research direction that opens the door to the bootstrapping of planning action models.
 

### 2017-11-27 doc/ijcai-18/constraints-learning.tex
first commit ijcai 2018 paper

--- 2017-11-16 doc/icaps-18/strips-learning.tex
@@
-\def\year{2017}\relax
-%File: formatting-instruction.tex
-\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
-\usepackage{aaai18}  %Required
-\usepackage{times}  %Required
-\usepackage{helvet}  %Required
-\usepackage{courier}  %Required
-\usepackage{url}  %Required
-\usepackage{graphicx}  %Required
-\frenchspacing  %Required
-\setlength{\pdfpagewidth}{8.5in}  %Required
-\setlength{\pdfpageheight}{11in}  %Required
-%PDF Info Is Required:
-  \pdfinfo{}
-\setcounter{secnumdepth}{0}
+%%%% ijcai18.tex
+
+\typeout{IJCAI-18 Instructions for Authors}
+
+% These are the instructions for authors for IJCAI-18.
+% They are the same as the ones for IJCAI-11 with superficical wording
+%   changes only.
+
+\documentclass{article}
+\pdfpagewidth=8.5in
+\pdfpageheight=11in
+% The file ijcai18.sty is the style file for IJCAI-18 (same as ijcai08.sty).
+\usepackage{ijcai18}
+
+% Use the postscript times font!
+\usepackage{times}
+\usepackage{xcolor}
+\usepackage{soul}
+\usepackage[utf8]{inputenc}
+\usepackage[small]{caption}
 
@@
 \usepackage{comment}
-
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{listings}
-\usepackage{multicol}
-\usepackage{arydshln}
-\usetikzlibrary{calc,backgrounds,positioning,fit}
 
@@
 \newcommand{\PE}{\mathrm{PE}}     % precondition
-\newcommand{\strips}{\textsc{Strips}}     % precondition
+\newcommand{\strips}{\textsc{Strips}}     % strips
 
@@
 
+\usepackage{graphicx}
+\usepackage{caption}
+\usepackage{subcaption}
+\usepackage{listings}
+\usepackage{multicol}
+\usepackage{arydshln}
+\usetikzlibrary{calc,backgrounds,positioning,fit}
+
+% the following package is optional:
+%\usepackage{latexsym}
+
+% Following comment is from ijcai97-submit.tex:
+% The preparation of these files was supported by Schlumberger Palo Alto
+% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
+% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
+% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
+% preparation.
+
+% These instructions can be modified and used in other conferences as long
+% as credit to the authors and supporting agencies is retained, this notice
+% is not changed, and further modification or reuse is not restricted.
+% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
+% contacts for providing assistance without their prior permission.
+
+% To use for other conferences, change references to files and the
+% conference appropriate and use other authors, contacts, publishers, and
+% organizations.
+% Also change the deadline and address for returning papers and the length and
+% page charge instructions.
+% Put where the files are available in the appropriate places.
+
+\title{Learning Strips Action Models from State-Constraints}
+
+\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
+{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
+{\small Universitat Polit\`ecnica de Val\`encia.}\\
+{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
+{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
+
 \begin{document}
 
-\title{Learning \strips\ action models with classical planning}
-\author{\#39}
+\maketitle
 
-% Commented for blind submission
-%\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
-%{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-%{\small Universitat Polit\`ecnica de Val\`encia.}\\
-%{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
-%{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
-
-\maketitle
 \begin{abstract}
-This paper presents a novel approach for learning \strips\ action models from examples that compiles this inductive learning task into classical planning.
-Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states (no intermediate action or state is given).
-What is more, the compilation accepts partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
+This paper extends the classical planning compilation for learning \strips\ action models from examples with the aim of learning from sets of states (no intermediate action is given).
+The compilation accepts state-constraints that are used to reduce the space of possible action models.
 \end{abstract}
@@
 Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
-
-On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML to the learning of \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
-\begin{itemize}
-\item The {\em input} to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects.
-The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
-\item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning \strips\ action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions.
-\end{itemize}
 
@@
 \end{enumerate}
-
-The second section of the paper formalizes the classical planning model, its extension to {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the output of the addressed learning task).
-The third section formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
-The fourth and fifth sections describe our approach for addressing the formalized learning tasks.
-Finally, the last sections report the data collected in a empirical evaluation, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
 
@@
 
-\subsubsection{Learning from labeled plans.}
-Here we augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
-
-\begin{itemize}
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-\end{itemize}
-
-Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda'$ in the blocksworld.
-This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
-
-\subsubsection{Learning from partially specified action models.}
-We may not require to start learning from scratch so we augment the learning input with partially specified operator schemes. This leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
-\begin{itemize}
-\item $\Xi_0$ is a partially specified action model in which some preconditions and effects are a priori known.
-\end{itemize}
+\subsubsection{Learning from state-constraints.}
 
@@
 variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
-Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
 \begin{small}
@@
 
-\begin{figure}
-\begin{scriptsize}
-\begin{verbatim}
-(:action apply_stack
-  :parameters (?o1 - object ?o2 - object)
-  :precondition
-   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
-        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
-        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
-        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
-        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
-        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
-        (or (not (pre_clear_stack_v1)) (clear ?o1))
-        (or (not (pre_clear_stack_v2)) (clear ?o2))
-        (or (not (pre_holding_stack_v1)) (holding ?o1))
-        (or (not (pre_holding_stack_v2)) (holding ?o2))
-        (or (not (pre_handempty_stack)) (handempty)))
-  :effect
-   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
-        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
-        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
-        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
-        (when (del_ontable_stack_v1) (not (ontable ?o1)))
-        (when (del_ontable_stack_v2) (not (ontable ?o2)))
-        (when (del_clear_stack_v1) (not (clear ?o1)))
-        (when (del_clear_stack_v2) (not (clear ?o2)))
-        (when (del_holding_stack_v1) (not (holding ?o1)))
-        (when (del_holding_stack_v2) (not (holding ?o2)))
-        (when (del_handempty_stack) (not (handempty)))
-        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
-        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
-        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
-        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
-        (when (add_ontable_stack_v1) (ontable ?o1))
-        (when (add_ontable_stack_v2) (ontable ?o2))
-        (when (add_clear_stack_v1) (clear ?o1))
-        (when (add_clear_stack_v2) (clear ?o2))
-        (when (add_holding_stack_v1) (holding ?o1))
-        (when (add_holding_stack_v2) (holding ?o2))
-        (when (add_handempty_stack) (handempty))
-        (when (modeProg) (not (modeProg)))))
-\end{verbatim}
-\end{scriptsize}
- \caption{\small Action for applying an already programmed schema $stack$ as encoded in PDDL (implications coded as disjunctions).}
-\label{fig:compilation}
-\end{figure}
-
 \section{Constraining the learning hypothesis space with additional input knowledge}
@@
 Here we show that further input knowledge can be used to constrain the space of possible action models and make the learning of \strips\ action models more practicable.
-
-\subsection{Labeled plans}
-We extend the compilation to consider labeled plans.
-Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
-\begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the plan step where the action model is validated.
-Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
-\item With respect to $A_{\Lambda'}$.
-\begin{enumerate}
-\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism forces that these actions are only applied as in the example plans.
-\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
-\begin{small}
-\begin{align*}
-&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}, \{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
-\end{align*}
-\end{small}
-\end{enumerate}
-\end{itemize}
-
-\subsection{Partially specified action models}
-Known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda'}$.
-The corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
-
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the blocksworld action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
-This plan programs and validates the operator schema {\tt\small stack} from the blocksworld, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-Plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are the actions for validating the programmed operators following the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
-
-In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
-In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips\ action model is {\em valid} for the given examples.
-If $P_{\Lambda'}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
-Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
-
-\begin{figure}
-{\tt\small
-     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
-     01 : (program\_pre\_handempty\_stack)\\
-     02 : (program\_pre\_holding\_stack\_v2)\\
-     03 : (program\_pre\_on\_stack\_v1\_v1)\\
-     04 : (program\_pre\_on\_stack\_v1\_v2)\\
-     05 : (program\_pre\_on\_stack\_v2\_v1)\\
-     06 : (program\_pre\_on\_stack\_v2\_v2)\\
-     07 : (program\_pre\_ontable\_stack\_v1)\\
-     08 : (program\_pre\_ontable\_stack\_v2)\\
-     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
-    10 : (program\_eff\_clear\_stack\_v2)\\
-    11 : (program\_eff\_handempty\_stack)\\
-    12 : (program\_eff\_holding\_stack\_v1)\\
-    13 : (program\_eff\_on\_stack\_v1\_v2)\\
-    {\bf 14} : (apply\_unstack a b i1 i2)\\
-    15 : (apply\_putdown a i2 i3)\\
-    16 : (apply\_unstack b c i3 i4)\\
-    17 : (apply\_stack b a i4 i5)\\
-    18 : (apply\_unstack c d i5 i6)\\
-    19 : (apply\_stack c b i6 i7)\\
-    20 : (apply\_pickup d i7 i8)\\
-    21 : (apply\_stack d c i8 i9)\\
-    {\bf 22} : (validate\_1)
-}
- \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi_1$ and label $\sigma_1$ (shown in Figure~\ref{fig:lexample}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
-\label{fig:plan}
-\end{figure}
-
-\subsection{Static predicates}
-A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action schema~\cite{fox:TIM:JAIR1998}.
-Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
-Formally, given a static predicate $p$:
-\begin{itemize}
-\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable objects} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
-\item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
-\end{itemize}
-
-Static predicates can also constrain the space of possible preconditions by looking at the given set of labels $\Sigma$.
-One can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
-For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\Sigma$.
-
-Likewise, static predicates can constrain the space of possible preconditions looking at the given example plans.
-Fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
-Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
-
-In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of {\em potential} static predicates from the input to the learning task.
-
-\begin{table*}
-		\resizebox{\textwidth}{!}{%                
-		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
-& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
-& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-                          \hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
-			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
-			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
-			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
-			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
-			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
-			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
-			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
-			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
-			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
-			\hline
-			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
-		\end{tabular}
-	}
-\caption{\small {\em Precision} and {\em recall} obtained learning from labeled plans without (left) and with (right) static predicates.}
-\label{tab:results_plans}
-\end{table*}
-
-\begin{table}
-\begin{scriptsize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c||c|c|c|}
-                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
-			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
-                         \hline
-			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
-			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
-			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
-			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
-			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
-			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
-                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
-                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
-			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
-			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
-			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
-			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
-		\end{tabular}
-	\end{center}
-        \end{scriptsize}
-	\caption{\small Total planning time, preprocessing time and plan length learning from labeled plans without/with static predicates.}
-	\label{tab:time_plans}	
-\end{table}
 
@@
 The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
-These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
-
 Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
@@
 
-\subsection{Learning from labeled plans}
-We start evaluating our approach at $\Lambda'=\tup{\Psi,\Sigma,\Pi}$ learning tasks, where {\em labeled plans} are available, and repeat the evaluation but exploiting potential static predicates that are computed from $\Sigma$.
-The potential {\em static predicates} considered here is the set of predicates s.t.
-every predicate instantiation appears unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
-Static predicates are used to constrain the space of possible action models as explained in the previous section.
-
-Table~\ref{tab:results_plans} summarizes the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
-We can observe that identifying static predicates drives to models with better precondition {\em recall}.
-This fact evidences that many of the missing preconditions corresponded to static predicates because there were no incentive to learn them as they always hold~\cite{gregory2015domain}.
-
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-All the learning tasks are solved in a few seconds time and interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
-When static predicates are identified, the resulting compilation is also much compact and produces smaller planning/instantiation times.
-
-\subsection{Learning from partially specified action models}
-
-We evaluate now the ability of our approach to support partially specified action models; that is, when addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
-In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task .
-
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, including the identification of static predicates.
-We report only the precision and recall for the {\em unknown} actions (considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in these models is 1.0).
-We must also note that errors in $\Lambda''$ tasks have greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is just done over half of the actions (e.g.
-the preconditions recall for the {\em Floortile} or {\em Gripper} domains).
-
-Remarkably, the overall precision is now $0.98$ which means that the content of the learned models is highly reliable.
-Recall is a lower, 0.87, indicating that the learned models still miss some information.
-Particularly, preconditions are again the component more difficult to be fully learned.
-Overall the reported results confirm the previous trend: more input knowledge drives to better models and requires less planning time.
-Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
-{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
-
-%In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
-
-\subsection{Learning from (initial,final) state pairs}
-Finally we evaluate our approach at learning tasks where input plans are not available, so the planner must not only compute the action models but also the plans that satisfy the input labels.
-Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the obtained results using static predicates and partially specified action models so the inputs to this learning task are $\Psi$ ,$\Sigma$ and $\Xi_0$.
-
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} could not solved the corresponding planning tasks within a 1000 secs time bound.
-Precision and recall have now low values.
-The learning hypothesis space is now so low constrained that actions can be reformulated and still be compliant with the inputs (e.g.
-the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
-We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) but still the achieved scores are below the scores obtained when learning from labeled plans (despite no other input knowledge was provided).
-
-%Table~\ref{tab:time_states} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported since the compilation could not be solved within a time bound of 1000 seconds.
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\resizebox{\columnwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-				Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.90 \\
-				Ferry & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
-                Grid & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.84 & 0.78 \\
-				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-				Miconic & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Satellite & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
-				Transport & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
-				Zenotravel & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.78 \\
-				\hline
-				\bf  & 0.98 & 0.71 & 1.0 & 0.98 & 1.0 & 0.95 & 0.98 & 0.87 \\
-			\end{tabular}
-		}
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} learning from partially specified action models.}
-\label{tab:results_plans_partial}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-                         \hline
-			Blocks & 0.07 & 0.01 & 54  \\
-			Driverlog & 0.03 & 0.01 & 40 \\
-			Ferry & 0.06 & 0.03 & 45 \\
-			Floortile & 0.43 & 0.42 & 55 \\
-                        Grid & 3.12 & 3.07 & 53 \\
-			Gripper & 0.03 & 0.01 & 35 \\
-			Miconic & 0.03 & 0.01 & 34  \\
-			Satellite & 0.14 & 0.14 & 47 \\
-			Transport & 0.23 & 0.21 & 37 \\
-			Zenotravel & 0.90 & 0.89 & 40 \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Time and plan length learning from partially specified action models.}
-	\label{tab:time_plans_partial}	
-\end{table}
-
-\section{Past and future of learning action models}
-
-Back in the 90's various systems aimed learning operators mostly via interaction with the environment.
-{\sc LIVE} captured and formulated observable features of objects and used them to acquire and refine operators \cite{ShenS89}.
-{\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
-These early works were based on lifting the observed states supported by exploratory plans or external teachers.%, but none provided a theoretical justification for this second source of knowledge.
-
-%More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
-
-%Whilst the aforementioned works deal with full state observability,
-Action model learning has also been studied in domains where there is partial or missing state observability.
-{\sf ARMS} works when no partial intermediate state is given.
-It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
-In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
-%and so it may output an inconsistent model
-{\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
-Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
-The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
-This update makes sure that the new formula represents all the transition relations consistent with the actions and observations.
-The formula returned at the end includes all consistent models, which can then be retrieved with additional processing.
-
-%Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
-An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
-
-{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
-This makes {\sf LOCM} be most likely the learning approach that works with the least information possible.
-The lack of available information is addressed by LOCM by exploiting assumptions about the kind of domain model it has to generate.
-Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
-The intuitive assumptions of {\sf LOCM}
-%, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence,
-yield a learning model heavily reliant on the kind of domain structure.
-The inability of {\sf LOCM} to properly derive domain theories where the state of a sort is subject to different FSMs is later overcome by {\sf LOCM2} by forming separate FSMs, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
-{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain}), the last contribution of the {\sf LOCM} family, addresses the problem of inducing static predicates.
-Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
-{\em blocksworld} and {\em freecell}).
-In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
-
-Compiling the learning of action models into classical planning is a general and flexible approach that allows to accommodate different amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
-For instance, the example plans in $\Pi$ could be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), so learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ could also be addressed.
-Furthermore, our approach seems extensible to learning other types of generative models (e.g.
-hierarchical models like HTN or behaviour trees), that can be more appealing than \strips\ models, since using them to compute planning solutions requires less search effort.
-
-%Furthermore, the compilation scheme can be extended to learn other type of generative models that comprise partially-specified solution plans like HTN models or full-specified solutions like FSM or behaviour trees.
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\resizebox{\columnwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-            Blocks & 0.33 & 0.33 & 0.75 & 0.50 & 0.33 & 0.33 & 0.47 & 0.39 \\
-            Driverlog & 1.0 & 0.29 & 0.33 & 0.67 & 1.0 & 0.50 & 0.78 & 0.48 \\
-            Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
-            Floortile & 0.67 & 0.40 & 0.50 & 0.40 & 1.0 & 0.40 & 0.72 & 0.40 \\
-            Grid & - & - & - & - & - & - & - & - \\
-            Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-            Miconic & 0.0 & 0.0 & 0.33 & 0.50 & 0.0 & 0.0 & 0.11 & 0.17 \\
-            Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
-            Transport & 0.0 & 0.0 & 0.25 & 0.5 & 0.0 & 0.0 & 0.08 & 0.17 \\
-            Zenotravel & - & - & - & - & - & - & - & - \\
-            \hline
-            & 0.63 & 0.29 & 0.54 & 0.70 & 0.67 & 0.53 & 0.61 & 0.51 \\			
-			\end{tabular}
-		}
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} learning from state pairs.}
-\label{tab:results_states}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-			\hline
-            Blocks & 2.14 & 0.00 & 58  \\
-            Driverlog & 0.09 & 0.00 & 88 \\
-            Ferry & 0.17 & 0.01 & 65 \\
-            Floortile & 6.42 & 0.15 & 126 \\
-            Grid & - & - & - \\
-            Gripper & 0.03 & 0.00 & 47 \\
-            Miconic & 0.04 & 0.00 & 68 \\
-            Satellite & 4.34 & 0.10 & 126 \\
-            Transport & 2.57 & 0.21 & 47 \\			
-            Zenotravel & - & - & - \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Time and plan length when learning from state pairs.}
-	\label{tab:time_states}	
-\end{table}
-
 \section{Conclusions}
-We presented a novel approach for learning \strips\ action models from examples using classical planning.
-The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
-We also introduced {\em precision} and {\em recall} metrics, widely used in ML, to the evaluation of planning action models.
-These metrics separately assess the correctness and completeness of the learned models making easier the identification of flaws.
-
-As far as we know, this is the first work on learning action models exclusively using an {\em off-the-shelf} classical planner and evaluated over a wide range of different domains.
-Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
-
-Our evaluation shows that, when example plans are available, we can compute accurate action models from small sets of learning examples (only five learning examples per domain) and investing small learning times (less than a second time in most of the domains).
-When action plans are not available, our approach is still able to produce action models compliant with the input information.
-In this case, since learning is not constrained by actions it can reformulate operators changing their semantics and making tricky their comparison with a reference model.
-%An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
-
-% The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
-In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
-
-The size of the compiled classical planning instances depend on the number of input examples.
-Generating {\em informative} examples for learning planning action models is an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
-The combination of such exploration techniques with our learning approach is an intriguing research direction that opens the door to the bootstrapping of planning action models.
-
-% Commented for blind submission
-%\begin{small}
-%\subsection*{Acknowledgment}
-%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
-Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
-%\end{small}
 
@@
 
-\bibliographystyle{aaai}
-\bibliography{strips-learning}
+%% The file named.bst is a bibliography style file for BibTeX 0.99c
+\bibliographystyle{named}
+\bibliography{ijcai18}
+
 \end{document}

### 2017-12-04 doc/ijcai-18/constraints-learning.tex
.pdf added to ijcai18 paper
updated sketch of ijcai18 paper

@@
 \begin{abstract}
-This paper extends the classical planning compilation for learning \strips\ action models from examples with the aim of learning from sets of states (no intermediate action is given).
-The compilation accepts state-constraints that are used to reduce the space of possible action models.
+This paper presents a classical planning compilation for learning \strips\ action models from observations of plan executions.
+Interestingly the input to the compilation is a set of {\em state-constraints} that bounds the space of possible action models and that does not require the precise executed actions.
+The output is a possible action \strips\ action model, that satisfies the given state-constraints.
+Last but not least, the paper shows that our compilation is also extensible to include further information about the observed plan executions to increase the accuracy of the learned models.
 \end{abstract}
@@
 
-Learning \strips\ action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-The compilation approach is appealing by itself because opens the door to the bootstrapping of planning action models but also because:
+Learning \strips\ action models from observations of plan executions is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models from observations of plan executions that:
 \begin{enumerate}
-\item Is flexible to different amounts of available input knowledge.
-The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate state  or action is observed.
-\item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
-In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
+\item Does not require information about the the precise applied actions.
+\item Can be defined as a classical planning compilation which opens the door to the bootstrapping of planning action models.
+\item Is extensible to include further information about the observed plan executions to increase the accuracy of the learned models.
 \end{enumerate}
@@
 
+\subsection{State-constraints}
+A {\em state invariant} is a formula $\phi$ that holds in the initial state, formally $I\models \phi$, and at any state reachable from that state~\cite{rintanen:schematicInvariants:AAAI2017}, for all states $s$ reachable from $I$ $s\models \phi$.
+
+A {\em mutex} is a particualr case of state invariant that take the form of a binary clause and that represents that a pair of different properties cannot be simultaneously true and are hence mutually exclusive.
+For instance, in a three-blocks blocksworld $on(block1,block2)$ and $on(block1,block3)$ are mutex.
+
+A {\em domain invariant} is a particular state invariant that does not depend on a particular initial state, since they are instance-independent~\cite{kautz:mutex:IJCAI1999}.
+A {\em lifted invariant} is a state invariant defined usig a first order formula over the predicates of a given planning domain~\cite{rintanen:schematicInvariants:AAAI2017}.
+For instance in the blocksworld, $\forall x,y,z\ (on(x,y)\wedge on(x,z)) \implies y=z$, is a domain and lifted invariant because a block can only be on top of a single block no matter the initial state  and no matter the block.
+
+Now we are ready to define a particular case of state-constraints called {\em lifted domain mutex} as a lifted domain invariant that takes the form of a binary clause.
+For instance, in the blocksworld $\forall x,y$ then $clear(x)$ and $on(x,y)$ is a {\em lifted domain mutex}.
+
+{\em Linear Temporal Logic} (LTL) includes model operator that also allows to represent state constraints~\cite{haslum:LTL:ecai10}.
+For instance the {\em always} operator, denoted by $\square$, defines constraints that, like {\em lifted domain invariants}, must be true at any reachable state.
+
 \section{Learning \strips\ action models}
@@
 
-\subsubsection{Learning from (initial, final) state pairs.} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
+\subsubsection{Learning from state-constraints.} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
 No information about the actions in the plans is given.
-This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma}$:
+This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma,\Phi}$:
 \begin{itemize}
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
-Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
+\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of state sequences $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$ such that,  {\small $1\leq i\leq n$} and {\tt\small $1\leq t\leq \tau$}, comprises the states resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
+\item $\Phi$ is a set of {\em lifted domain mutex}.
 \end{itemize}
 
-\subsubsection{Learning from state-constraints.}
-
-A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant just with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given labels $\Sigma$ using the learned \strips\ model.
-A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
-Finally a solution to $\Lambda''$ is a set of \strips\ operator schema $\Xi$ compliant as well with the provided partially specified action model $\Xi_0$.
+A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, the given set of states $\Sigma$ and the set of state constraints $\Phi$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given set of states $\Sigma$ using the learned \strips\ model.
 
@@
 
-\begin{subfigure}{.25\textwidth}
-{\tt ;;; Plan $\pi_1$}
-\begin{footnotesize}
-\begin{verbatim}
-0: (unstack A B)
-1: (putdown A)
-2: (unstack B C)
-3: (stack B A)
-4: (unstack C D)
-5: (stack C B)
-6: (pickup D)
-7: (stack D C)
-\end{verbatim}
-\end{footnotesize}
-\end{subfigure}%
 \begin{subfigure}{.6\textwidth}
-{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
+{\tt ;;; Label $\sigma_1=(s_0^1,s_1^1,s_2^1,s_3^1)$}
 \begin{lstlisting}[mathescape]
@@
 \begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
-\node (initD) [draw]{D};
-\node (initC) [draw, above=of initD.north]{C};
-\node (initB) [draw, above=of initC.north]{B};
+\node (initB) [draw, ]{B};
 \node (initA) [draw, above=of initB.north]{A};
+\node (twoB) [draw, right=of initB]{A};
+\node (twoA) [draw, right=of twoB]{B};
+\node (three) [draw, ]{B};
+\node (three) [draw, above=of initB.north]{A};
 \draw[thick] (-1,-0.25) -- (2.5,-0.25);
-
-\node (goalA) [draw, right=10mm of initD]{A};
-\node (goalB) [draw, right=10mm of initC]{B};
-\node (goalC) [draw, right=10mm of initB]{C};
-\node (goalD) [draw, right=10mm of initA]{D};
+\node (goalB) [draw, right=10mm of initB]{A};
+\node (goalA) [draw, right=10mm of initA]{B};
 \end{tikzpicture}
@@
 \end{subfigure}%
+
+\begin{subfigure}{.6\textwidth}
+{\tt ;;; Label $\Phi$}
+\begin{lstlisting}[mathescape]
+(:derived (invariant-1-1)
+  (forall (?o1 - object)
+      (not (and (handempty) (holding ?o1)))))
+
+(:derived (invariant-1-2)
+  (forall (?o1 - object)
+      (not (and (holding ?o1) (clear ?o1)))))
+
+(:derived (invariant-1-3)
+  (forall (?o1 - object)
+      (not (and (holding ?o1) (ontable ?o1)))))
+
+(:derived (invariant-1-4)
+  (forall (?o1 - object)
+      (not (and (on ?o1 ?o1)))))
+
+;;;
+;;; Invariants with two quantified variable and size 2
+;;;
+(:derived (invariant-2-1)
+  (forall (?o1 ?o2 - object)
+      (not (and (on ?o1 ?o2) (holding ?o1)))))
+
+(:derived (invariant-2-2)
+  (forall (?o1 ?o2 - object)
+      (not (and (on ?o1 ?o2) (holding ?o2)))))
+
+(:derived (invariant-2-3)
+  (forall (?o1 ?o2 - object)
+      (not (and (on ?o1 ?o2) (clear ?o2)))))
+
+(:derived (invariant-2-4)
+  (forall (?o1 ?o2 - object)
+      (not (and (on ?o1 ?o2) (ontable ?o1)))))
+
+(:derived (invariant-2-5)
+  (forall (?o1 ?o2 - object)
+      (not (and (on ?o1 ?o2) (on ?o2 ?o1)))))
+\end{lstlisting}
+\end{subfigure}%
+
  \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
@@
 \section{Learning \strips\ action models with planning}
-Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is compiling it into a classical planning task with conditional effects.
+Our approach for addressing the learning task $\Lambda$, is compiling it into a classical planning task with conditional effects.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (the labels $\Sigma$, and $\Pi$ and/or $\Xi_0$ if available).
-For every label $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
+\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (the sequences of states $\Sigma$, and $\Phi$ if available).
+For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
 We call this process the validation of the programmed \strips\ action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
@@
 
+\subsection{State-constraints}
+
+\subsection{Plan constraints}
+\subsubsection{Learning from labeled plans.}
+Here we augment the input knowledge with the actions executed by the observed agent.
+
+\begin{itemize}
+\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+\end{itemize}
+
+We extend the compilation to consider labeled plans.
+Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
+\begin{itemize}
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
+Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the plan step where the action model is validated.
+Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
+\item With respect to $A_{\Lambda'}$.
+\begin{enumerate}
+\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
+\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+This mechanism forces that these actions are only applied as in the example plans.
+\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
+\begin{small}
+\begin{align*}
+&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}, \{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
+\end{align*}
+\end{small}
+\end{enumerate}
+\end{itemize}
+
 \section{Evaluation}
@@
 \bibliographystyle{named}
-\bibliography{ijcai18}
+\bibliography{constraints-learning}
 

### 2017-12-04 doc/ijcai-18/constraints-learning.tex
updating the sketch for the ijcai18 paper

@@
 
-\title{Learning Strips Action Models from State-Constraints}
+\title{Learning \strips\ Action Models from State-Constraints}
 
@@
 This paper presents a classical planning compilation for learning \strips\ action models from observations of plan executions.
-Interestingly the input to the compilation is a set of {\em state-constraints} that bounds the space of possible action models and that does not require the precise executed actions.
+Interestingly the compilation is fed with a set of {\em state-constraints}, that bounds the space of possible action models, and does not require the precise executed actions in the observations of the plan executions.
 The output is a possible action \strips\ action model, that satisfies the given state-constraints.
-Last but not least, the paper shows that our compilation is also extensible to include further information about the observed plan executions to increase the accuracy of the learned models.
+Last but not least, the paper shows that this compilation is also flexible to include further information about the observed plan executions and increase the accuracy of the learned models.
 \end{abstract}
@@
 
-Learning \strips\ action models from observations of plan executions is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
+An alternative approach to the hand-coding of action models is learning them from observations of plan executions.
+This is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models from observations of plan executions that:
 \begin{enumerate}
-\item Does not require information about the the precise applied actions.
+\item Does not require information about the the precise actions applied in the observed plan executions.
 \item Can be defined as a classical planning compilation which opens the door to the bootstrapping of planning action models.
-\item Is extensible to include further information about the observed plan executions to increase the accuracy of the learned models.
+\item Is extensible to further information about the observed plan executions to improve the learned models.
 \end{enumerate}
@@
 \subsection{State-constraints}
-A {\em state invariant} is a formula $\phi$ that holds in the initial state, formally $I\models \phi$, and at any state reachable from that state~\cite{rintanen:schematicInvariants:AAAI2017}, for all states $s$ reachable from $I$ $s\models \phi$.
+A {\em state invariant} is a formula $\phi$ that holds in the initial state of a given classical planning problem, formally $I\models \phi$, and at any state reachable from that state~\cite{rintanen:schematicInvariants:AAAI2017}, for all states $s$ reachable from $I$ $s\models \phi$.
+State invariants are traditionally useful for making satisfiability planning and backward search more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
 
-A {\em mutex} is a particualr case of state invariant that take the form of a binary clause and that represents that a pair of different properties cannot be simultaneously true and are hence mutually exclusive.
-For instance, in a three-blocks blocksworld $on(block1,block2)$ and $on(block1,block3)$ are mutex.
+A {\em mutex} (meaning mutually exclusive) is a particular case of state invariant that take the form of a binary clause and that represents that a pair of different properties cannot be simultaneously true.
+For instance in a three-blocks blocksworld, $on(block1,block2)$ and $on(block1,block3)$ are mutex because $block1$ can only be on top of a single block.
 
-A {\em domain invariant} is a particular state invariant that does not depend on a particular initial state, since they are instance-independent~\cite{kautz:mutex:IJCAI1999}.
-A {\em lifted invariant} is a state invariant defined usig a first order formula over the predicates of a given planning domain~\cite{rintanen:schematicInvariants:AAAI2017}.
-For instance in the blocksworld, $\forall x,y,z\ (on(x,y)\wedge on(x,z)) \implies y=z$, is a domain and lifted invariant because a block can only be on top of a single block no matter the initial state  and no matter the block.
-
-Now we are ready to define a particular case of state-constraints called {\em lifted domain mutex} as a lifted domain invariant that takes the form of a binary clause.
-For instance, in the blocksworld $\forall x,y$ then $clear(x)$ and $on(x,y)$ is a {\em lifted domain mutex}.
-
-{\em Linear Temporal Logic} (LTL) includes model operator that also allows to represent state constraints~\cite{haslum:LTL:ecai10}.
-For instance the {\em always} operator, denoted by $\square$, defines constraints that, like {\em lifted domain invariants}, must be true at any reachable state.
+A {\em lifted invariant} is a state invariant defined using a first order formula over the predicates of a given planning domain~\cite{rintanen:schematicInvariants:AAAI2017}.
+A {\em Domain invariant} is a state invariant that hold for any possible initial state~\cite{kautz:mutex:IJCAI1999}, they are instance-independent and often they are lifted invariants.
+For instance in the blocksworld, $\forall x,y,z\ (on(x,y)\wedge on(x,z)) \implies y=z$, is a domain and lifted invariant because a block can only be on top of a single block no matter the initial state of the classical planning problem and no matter the particular identity of the blocks.
 
@@
 Learning \strips\ action models from fully available input knowledge, i.e.
-from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
-When any intermediate state is available, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+from plans where every action in the plan is available as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
+In this case, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
 Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions.
 
-This section formalizes more challenging learning tasks, where less input knowledge is available:
-
-\subsubsection{Learning from state-constraints.} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
-No information about the actions in the plans is given.
+Here we formalize a more challenging learning task, where less input knowledge is available.
+This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
+No information about the particular actions is given.
 This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma,\Phi}$:
@@
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of state sequences $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$ such that,  {\small $1\leq i\leq n$} and {\tt\small $1\leq t\leq \tau$}, comprises the states resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
+\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of state sequences $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, that comprises the states resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
 \item $\Phi$ is a set of {\em lifted domain mutex}.
@@
 
-A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, the given set of states $\Sigma$ and the set of state constraints $\Phi$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given set of states $\Sigma$ using the learned \strips\ model.
+A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, the given states sequences in $\Sigma$ and the set of state constraints $\Phi$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the content of the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain $\Sigma$ and $\Phi$ with the learned \strips\ model.
 
@@
 
+\begin{footnotesize}
 \begin{subfigure}{.6\textwidth}
-{\tt ;;; Label $\Phi$}
-\begin{lstlisting}[mathescape]
+{\tt ;;; Lifted domain mutexes in $\Phi$}
+\begin{verbatim}
 (:derived (invariant-1-1)
@@
 
-;;;
-;;; Invariants with two quantified variable and size 2
-;;;
 (:derived (invariant-2-1)
@@
       (not (and (on ?o1 ?o2) (on ?o2 ?o1)))))
-\end{lstlisting}
+\end{verbatim}
 \end{subfigure}%
+\end{footnotesize}
 
@@
 A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (the sequences of states $\Sigma$, and $\Phi$ if available).
+\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (the sequences of states $\Sigma$ and $\Phi$ formulae).
 For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
@@
 
-Now we are ready to formalize the compilations.
-We start with $\Lambda$, because it requires less input knowledge.
+Now we are ready to formalize the compilation.
 Given a learning task $\Lambda=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
@@
 \item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is being validated.
+\item Fluents, $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, to code the obsevation steps and iterate through them.
 \end{itemize}
-\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
+\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), fluents $at_1$ and and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the step where the action model is validated and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
 Our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect.
@@
 &\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]},\\
 &\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
@@
 &\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
+\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\},\\
+&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
 \end{align*}
@@
 \subsection{Plan constraints}
-\subsubsection{Learning from labeled plans.}
-Here we augment the input knowledge with the actions executed by the observed agent.
+
+Here we augment the input knowledge with the actions executed by the observed agent, when available.
 
@@
 
-We extend the compilation to consider labeled plans.
-Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
+We extend the compilation to consider the actions in the executed plans.
+Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Phi,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
 \begin{itemize}
 \item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the plan step where the action model is validated.
-Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$. Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
 \item With respect to $A_{\Lambda'}$.
@@
 \item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step.
 This mechanism forces that these actions are only applied as in the example plans.
-\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
+\item The actions for {\em validating} the current learning example have an extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
 \begin{small}
 \begin{align*}
-&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}, \{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
+&\{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
 \end{align*}
@@
 
+\subsection{Learning with state-constraints}
+For each domain we provide a set of {\em lifted domain mutex},  lifted domain invariants that takes the form of a binary clauses.
+For instance, in the blocksworld $\forall x,y$ then $clear(x)$ and $on(x,y)$ is a {\em lifted domain mutex}.
+
+\subsection{Learning with plans}
+
 \section{Conclusions}
+{\em Linear Temporal Logic} (LTL) also includes model operator that also allows to represent state constraints~\cite{haslum:LTL:ecai10}.
+For instance the {\em always} operator, denoted by $\square$, defines constraints that, like {\em lifted domain invariants}, must be true at any reachable state.
 

### 2017-12-05 doc/ijcai-18/constraints-learning.tex
added updates to ijcai18 paper

@@
 This paper presents a classical planning compilation for learning \strips\ action models from observations of plan executions.
-Interestingly the compilation is fed with a set of {\em state-constraints}, that bounds the space of possible action models, and does not require the precise executed actions in the observations of the plan executions.
-The output is a possible action \strips\ action model, that satisfies the given state-constraints.
-Last but not least, the paper shows that this compilation is also flexible to include further information about the observed plan executions and increase the accuracy of the learned models.
+Remarkably the compilation is fed with {\em state-constraints}, that bound the space of possible action models, and does not require the precise actions in the plan executions.
+A plan that solves the classical planning task resulting from this compilation induces a \strips\ action model that is compliant with the given state-constraints.
+When the amount of input knowledge is too scarce, e.g.
+the state-constraints are too loose, evaluating learned \strips\ models with respect to reference action model becomes hard.
+The paper introduces a novel evaluation methodology that allows to asses the quality of learned \strips\ models even if actions are reformulated.
+Last but not least, the paper shows the flexibility of our compilation to include further information about the observed plan executions and increase the accuracy of the learned models.
 \end{abstract}
@@
 
-An alternative approach to the hand-coding of action models is learning them from observations of plan executions.
+Learning action models from observations of plan executions is a promising alternative to the hand-coding of planning action models.
 This is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
@@
 \begin{enumerate}
-\item Does not require information about the the precise actions applied in the observed plan executions.
 \item Can be defined as a classical planning compilation which opens the door to the bootstrapping of planning action models.
-\item Is extensible to further information about the observed plan executions to improve the learned models.
+\item Does not require the precise actions applied in the observed plan executions but is extensible to further information about the observed plan executions to improve the learned models.
+\item Assesses the quality of learned \strips\ models with a reference model even when the learning hypothesis space is so low constrained that the learned actions models can be reformulated and still be compliant with the inputs.
+%For instance in the {\em blocksworld}, given a low amount of input knowledge, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa) making non trivial evaluating the learned model with respect to a reference model.
 \end{enumerate}
@@
 \subsection{State-constraints}
-A {\em state invariant} is a formula $\phi$ that holds in the initial state of a given classical planning problem, formally $I\models \phi$, and at any state reachable from that state~\cite{rintanen:schematicInvariants:AAAI2017}, for all states $s$ reachable from $I$ $s\models \phi$.
+A {\em state invariant} is a formula $\phi$ that holds in the initial state of a given classical planning problem, $I\models \phi$, and at any state reachable from that state, i.e.
+for all states $s$ reachable from $I$ it holds that $s\models \phi$.
 State invariants are traditionally useful for making satisfiability planning and backward search more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
 
-A {\em mutex} (meaning mutually exclusive) is a particular case of state invariant that take the form of a binary clause and that represents that a pair of different properties cannot be simultaneously true.
+A {\em mutex} (meaning mutually exclusive) is a particular case of state invariant that takes the form of a binary clause and that represents a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
 For instance in a three-blocks blocksworld, $on(block1,block2)$ and $on(block1,block3)$ are mutex because $block1$ can only be on top of a single block.
 
-A {\em lifted invariant} is a state invariant defined using a first order formula over the predicates of a given planning domain~\cite{rintanen:schematicInvariants:AAAI2017}.
-A {\em Domain invariant} is a state invariant that hold for any possible initial state~\cite{kautz:mutex:IJCAI1999}, they are instance-independent and often they are lifted invariants.
-For instance in the blocksworld, $\forall x,y,z\ (on(x,y)\wedge on(x,z)) \implies y=z$, is a domain and lifted invariant because a block can only be on top of a single block no matter the initial state of the classical planning problem and no matter the particular identity of the blocks.
+A {\em lifted invariant} (also called schematic invariants in the literature) is a state invariant defined using a first order formula over the predicates of a given planning problem~\cite{rintanen:schematicInvariants:AAAI2017}.
+A {\em domain invariant} is a state invariant that holds for any possible initial state, they are instance-independent and often, they are lifted invariants~\cite{fox:TIM:JAIR1998}.
+For instance in the blocksworld, $\forall x,y,z:\ (on(x,y)\wedge on(x,z)) \implies y=z$, is a domain and lifted invariant because a block can only be on top of a single block no matter the initial state of the classical planning problem and no matter the particular identity of the blocks.
 
@@
 
-Here we formalize a more challenging learning task, where less input knowledge is available.
-This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
-No information about the particular actions is given.
-This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma,\Phi}$:
+We formalize a more challenging learning task, where less input knowledge is available.
+This learning task, denoted by $\Lambda=\tup{\Psi,\Sigma,\Phi}$, corresponds to observing an agent acting in the world but watching only the results of its plan executions:
 \begin{itemize}
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of state sequences $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, that comprises the states resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
+\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
+Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
 \item $\Phi$ is a set of {\em lifted domain mutex}.
@@
 A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, the given states sequences in $\Sigma$ and the set of state constraints $\Phi$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the content of the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain $\Sigma$ and $\Phi$ with the learned \strips\ model.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the content of the {\em unobserved} plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain $\Sigma$ and $\Phi$ with the learned \strips\ model $\Xi$.
 
@@
 \begin{subfigure}{.6\textwidth}
-{\tt ;;; Label $\sigma_1=(s_0^1,s_1^1,s_2^1,s_3^1)$}
+{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
 \begin{lstlisting}[mathescape]
@@
 \begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
-\node (initB) [draw, ]{B};
+\node (initD) [draw]{D};
+\node (initC) [draw, above=of initD.north]{C};
+\node (initB) [draw, above=of initC.north]{B};
 \node (initA) [draw, above=of initB.north]{A};
-\node (twoB) [draw, right=of initB]{A};
-\node (twoA) [draw, right=of twoB]{B};
-\node (three) [draw, ]{B};
-\node (three) [draw, above=of initB.north]{A};
 \draw[thick] (-1,-0.25) -- (2.5,-0.25);
-\node (goalB) [draw, right=10mm of initB]{A};
-\node (goalA) [draw, right=10mm of initA]{B};
+
+\node (goalA) [draw, right=10mm of initD]{A};
+\node (goalB) [draw, right=10mm of initC]{B};
+\node (goalC) [draw, right=10mm of initB]{C};
+\node (goalD) [draw, right=10mm of initA]{D};
 \end{tikzpicture}
@@
 \end{subfigure}%
+
+\vspace{0.2cm}
 
@@
 
-\section{Learning \strips\ action models with planning}
+\subsection{Compiling the learning of \strips\ action models into classical planning}
+
 Our approach for addressing the learning task $\Lambda$, is compiling it into a classical planning task with conditional effects.
@@
 A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (the sequences of states $\Sigma$ and $\Phi$ formulae).
-For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
+\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (states $\Sigma$ and formulae $\Phi$).
+For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying every $\phi\in\Phi$ at every reached state.
 We call this process the validation of the programmed \strips\ action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
@@
 Now we are ready to formalize the compilation.
-Given a learning task $\Lambda=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+Given a learning task $\Lambda=\tup{\Psi,\Sigma,\Phi}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
 \begin{itemize}
@@
 \item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is being validated.
-\item Fluents, $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, to code the obsevation steps and iterate through them.
 \end{itemize}
-\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), fluents $at_1$ and and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the step where the action model is validated and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
+\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label) and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
 Our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect.
@@
 variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
+Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
 \begin{small}
@@
 &\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]},\\
 &\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
@@
 
-\item Actions for {\em validating} the learning example {\tt\small $1\leq t\leq \tau$}. 
+\item Actions for {\em validating} learning example {\tt\small $1\leq t\leq \tau$}. 
 \begin{small}
@@
 &\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\},\\
-&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}.
+\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
 \end{align*}
@@
 
+\begin{figure}
+\begin{scriptsize}
+\begin{verbatim}
+(:action apply_stack
+  :parameters (?o1 - object ?o2 - object)
+  :precondition
+   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
+        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
+        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
+        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
+        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
+        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
+        (or (not (pre_clear_stack_v1)) (clear ?o1))
+        (or (not (pre_clear_stack_v2)) (clear ?o2))
+        (or (not (pre_holding_stack_v1)) (holding ?o1))
+        (or (not (pre_holding_stack_v2)) (holding ?o2))
+        (or (not (pre_handempty_stack)) (handempty)))
+  :effect
+   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
+        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
+        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
+        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
+        (when (del_ontable_stack_v1) (not (ontable ?o1)))
+        (when (del_ontable_stack_v2) (not (ontable ?o2)))
+        (when (del_clear_stack_v1) (not (clear ?o1)))
+        (when (del_clear_stack_v2) (not (clear ?o2)))
+        (when (del_holding_stack_v1) (not (holding ?o1)))
+        (when (del_holding_stack_v2) (not (holding ?o2)))
+        (when (del_handempty_stack) (not (handempty)))
+        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
+        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
+        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
+        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
+        (when (add_ontable_stack_v1) (ontable ?o1))
+        (when (add_ontable_stack_v2) (ontable ?o2))
+        (when (add_clear_stack_v1) (clear ?o1))
+        (when (add_clear_stack_v2) (clear ?o2))
+        (when (add_holding_stack_v1) (holding ?o1))
+        (when (add_holding_stack_v2) (holding ?o2))
+        (when (add_handempty_stack) (handempty))
+        (when (modeProg) (not (modeProg)))))
+\end{verbatim}
+\end{scriptsize}
+ \caption{\small Action for applying an already programmed schema $stack$ as encoded in PDDL (implications coded as disjunctions).}
+\label{fig:compilation}
+\end{figure}
+
 \section{Constraining the learning hypothesis space with additional input knowledge}
@@
 
-\subsection{State-constraints}
+\subsection{State constraints}
+Every state invariant $\phi\in\Phi$ is added as an extra precondition of actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema.
+Likewise every state invariant $\phi\in\Phi$ must hold at the last reached state so they are included aas extra goals in $G_{\Lambda}$.
+
+With regard to state trajectories.
+Now $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is not a set of $(initial, final)$ state pairs but a set of state trajectories $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, that comprises the sequence of states resulting from executing the {\em unobserved} plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
+
+\begin{itemize}
+\item Fluents, $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are added to $F_{\Lambda}$ with the aim of iterating through state trajectories.
+Likewise $I_{\Lambda}$ includes also now the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}
+\item Actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema have an extra conditional effect $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$.
+Likewise actions $\mathsf{validate_{t}}$ for {\em validating} a learning example {\tt\small $1\leq t\leq \tau$} add $at_1$ and delete $at_{|\pi_t|}$.
+\end{itemize}
 
@@
 
-Here we augment the input knowledge with the actions executed by the observed agent, when available.
+Here we augment the input knowledge with the actions executed by the observed agent, when available. So the leaerning task is now defined as $\Lambda=\tup{\Psi,\Sigma,\Phi,\Pi}$
 

### 2017-12-07 doc/ijcai-18/constraints-learning.tex
building first version ijcai18 paper

@@
 \begin{abstract}
-This paper presents a classical planning compilation for learning \strips\ action models from observations of plan executions.
-Remarkably the compilation is fed with {\em state-constraints}, that bound the space of possible action models, and does not require the precise actions in the plan executions.
-A plan that solves the classical planning task resulting from this compilation induces a \strips\ action model that is compliant with the given state-constraints.
-When the amount of input knowledge is too scarce, e.g.
-the state-constraints are too loose, evaluating learned \strips\ models with respect to reference action model becomes hard.
-The paper introduces a novel evaluation methodology that allows to asses the quality of learned \strips\ models even if actions are reformulated.
-Last but not least, the paper shows the flexibility of our compilation to include further information about the observed plan executions and increase the accuracy of the learned models.
+We present a classical planning compilation for learning \strips\ action models from state-constraints.
+Remarkably the compilation does not require the precise actions in the plan executions, despite they can be included to constraint further the space of possible action models.
+A plan that solves the classical planning task resulting from our compilation induces a \strips\ action model that is compliant with all the constraints given as input.
+The paper also shows that, when input constraints are too loose, evaluating learned \strips\ models with respect to reference model becomes non trivial (actions can be reformulated, i.e.
+their semantic can be swapped) and introduces a novel evaluation method that asses learned \strips\ models even if they are reformulated.
 \end{abstract}
@@
 
-Learning action models from observations of plan executions is a promising alternative to the hand-coding of planning action models.
-This is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models from observations of plan executions that:
+The Machine Learning of planning action models is a promising alternative to hand-coding them with sophisticated algorithms like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that:
 \begin{enumerate}
-\item Can be defined as a classical planning compilation which opens the door to the bootstrapping of planning action models.
-\item Does not require the precise actions applied in the observed plan executions but is extensible to further information about the observed plan executions to improve the learned models.
-\item Assesses the quality of learned \strips\ models with a reference model even when the learning hypothesis space is so low constrained that the learned actions models can be reformulated and still be compliant with the inputs.
+\item Is defined as a classical planning compilation which opens the door to the bootstrapping of planning action models.
+\item Does not require the actions applied in the observed plan executions but accepts this information to improve the quality of the learned models.
+\item Assesses learned \strips\ models with respect to a reference model even when the learning hypothesis space is so low constrained that actions can be reformulated and still be compliant with the inputs.
 %For instance in the {\em blocksworld}, given a low amount of input knowledge, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa) making non trivial evaluating the learned model with respect to a reference model.
@@
 \begin{figure}
-\begin{footnotesize}
+\begin{scriptsize}
 \begin{verbatim}
@@
   :precondition (and (holding ?v1) (clear ?v2))
-  :effect (and (not (holding ?v1))
-               (not (clear ?v2))
-               (handempty) (clear ?v1)
-               (on ?v1 ?v2)))
+  :effect (and (not (holding ?v1)) (not (clear ?v2))
+               (handempty) (clear ?v1) (on ?v1 ?v2)))
 \end{verbatim}
-\end{footnotesize}
+\end{scriptsize}
  \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
@@
 \subsection{State-constraints}
-A {\em state invariant} is a formula $\phi$ that holds in the initial state of a given classical planning problem, $I\models \phi$, and at any state reachable from that state, i.e.
-for all states $s$ reachable from $I$ it holds that $s\models \phi$.
-State invariants are traditionally useful for making satisfiability planning and backward search more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+A {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ reachable from $I$.
+State invariants are traditionally useful for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
 
-A {\em mutex} (meaning mutually exclusive) is a particular case of state invariant that takes the form of a binary clause and that represents a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+A {\em mutex} (mutually exclusive) is a particular state invariant that takes the form of a binary clause and that represents a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
 For instance in a three-blocks blocksworld, $on(block1,block2)$ and $on(block1,block3)$ are mutex because $block1$ can only be on top of a single block.
 
-A {\em lifted invariant} (also called schematic invariants in the literature) is a state invariant defined using a first order formula over the predicates of a given planning problem~\cite{rintanen:schematicInvariants:AAAI2017}.
-A {\em domain invariant} is a state invariant that holds for any possible initial state, they are instance-independent and often, they are lifted invariants~\cite{fox:TIM:JAIR1998}.
-For instance in the blocksworld, $\forall x,y,z:\ (on(x,y)\wedge on(x,z)) \implies y=z$, is a domain and lifted invariant because a block can only be on top of a single block no matter the initial state of the classical planning problem and no matter the particular identity of the blocks.
+A {\em lifted invariant} (also called schematic invariants in the literature) is a state invariant defined using a first order formula~\cite{rintanen:schematicInvariants:AAAI2017}.
+A {\em domain invariant} is a state invariant that is instance-independent (it holds for any possible initial state) and often, they are lifted invariants~\cite{fox:TIM:JAIR1998}.
+For instance in the blocksworld, $\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em lifted domain mutex} because the robot hand can never be empty and holding a block at the same time.
 
@@
 Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
-\item $\Phi$ is a set of {\em lifted domain mutex}.
+\item $\Phi$ is a set of {\em lifted domain invariants}.
 \end{itemize}
 
-A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, the given states sequences in $\Sigma$ and the set of state constraints $\Phi$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the content of the {\em unobserved} plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain $\Sigma$ and $\Phi$ with the learned \strips\ model $\Xi$.
+A solution to $\Lambda$ is a set of operator schema $\Xi$ compliant with the predicates in $\Psi$, the labels $\Sigma$, and the set of state-constraints $\Phi$.
+A planning compilation is a suitable approach for addressing the $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that explain $\Sigma$ and $\Phi$.
 
@@
 \vspace{0.2cm}
-
 \begin{subfigure}{.6\textwidth}
@@
 \end{lstlisting}
-\vspace{0.1cm}
+\vspace{-0.2cm}
+
 \begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
@@
 \end{tikzpicture}
-\vspace{0.6cm}
 \end{subfigure}%
 
-\vspace{0.2cm}
+\vspace{0.6cm}
 
-\begin{footnotesize}
 \begin{subfigure}{.6\textwidth}
-{\tt ;;; Lifted domain mutexes in $\Phi$}
+{\tt ;;; Lifted domain invariants in $\Phi$}
+\begin{scriptsize}
 \begin{verbatim}
-(:derived (invariant-1-1)
  (forall (?o1 - object)
@@
 
-(:derived (invariant-1-2)
  (forall (?o1 - object)
@@
 
-(:derived (invariant-1-3)
  (forall (?o1 - object)
-      (not (and (holding ?o1) (ontable ?o1)))))
+   (not (and (holding ?o1) (ontable ?o1))))
 
-(:derived (invariant-1-4)
  (forall (?o1 - object)
-      (not (and (on ?o1 ?o1)))))
+   (not (and (on ?o1 ?o1))))
 
-(:derived (invariant-2-1)
  (forall (?o1 ?o2 - object)
-      (not (and (on ?o1 ?o2) (holding ?o1)))))
+   (not (and (on ?o1 ?o2) (holding ?o1))))
 
-(:derived (invariant-2-2)
  (forall (?o1 ?o2 - object)
-      (not (and (on ?o1 ?o2) (holding ?o2)))))
+   (not (and (on ?o1 ?o2) (holding ?o2))))
 
-(:derived (invariant-2-3)
  (forall (?o1 ?o2 - object)
-      (not (and (on ?o1 ?o2) (clear ?o2)))))
+    (not (and (on ?o1 ?o2) (clear ?o2))))
 
-(:derived (invariant-2-4)
  (forall (?o1 ?o2 - object)
-      (not (and (on ?o1 ?o2) (ontable ?o1)))))
+    (not (and (on ?o1 ?o2) (ontable ?o1))))
 
-(:derived (invariant-2-5)
  (forall (?o1 ?o2 - object)
@@
 \end{verbatim}
+\end{scriptsize}
 \end{subfigure}%
-\end{footnotesize}
-
  \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
@@
 
-\subsection{Compiling the learning of \strips\ action models into classical planning}
+\subsection{Learning \strips\ action models with classical planning}
 
@@
 \item Programs the \strips\ action model $\Xi$.
-A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (states $\Sigma$ and formulae $\Phi$).
+A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Validates the programmed \strips\ action model $\Xi$ in the input constraints (labels $\Sigma$ and formulae $\Phi$).
 For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying every $\phi\in\Phi$ at every reached state.
-We call this process the validation of the programmed \strips\ action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
+We call this process the validation of the programmed \strips\ action model $\Xi$, at the $t^{th}$ learning example {\small $1\leq t\leq \tau$}.
 %If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
@@
 \item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label) and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
-Our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect.
+Our compilation assumes that initially operator schemas are programmed with every possible precondition, no negative effect and no positive effect.
 \item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
@@
 \begin{enumerate}
-\item Actions for {\em programming} an operator schema $\xi\in\Xi$:
+\item Actions for {\em programming} operator schema $\xi\in\Xi$:
 \begin{itemize}
@@
 \end{itemize}
-
-\begin{lemma}
-Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
-\end{lemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after preconditions are programmed).
-Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
-If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
-\end{small}
-\end{proof}
-
-The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model.
 
@@
 
-\section{Constraining the learning hypothesis space with additional input knowledge}
+\begin{lemma}
+Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
+The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after preconditions are programmed).
+Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
+To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
+If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
+\end{small}
+\end{proof}
+
+The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model.
+
+\section{Constraining the learning hypothesis space}
 \label{sec:Constraining}
-Here we show that further input knowledge can be used to constrain the space of possible action models and make the learning of \strips\ action models more practicable.
+Here we show that constraints can be used to reduce the space of possible action models and make the learning of \strips\ action models more practicable.
 
@@
 Every state invariant $\phi\in\Phi$ is added as an extra precondition of actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema.
-Likewise every state invariant $\phi\in\Phi$ must hold at the last reached state so they are included aas extra goals in $G_{\Lambda}$.
+Likewise, every state invariant $\phi\in\Phi$ is also added as an extra goal in $G_{\Lambda}$ because formulae $\phi\in\Phi$ must hold at every reached state, including the last reached state.
 
-With regard to state trajectories.
-Now $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is not a set of $(initial, final)$ state pairs but a set of state trajectories $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, that comprises the sequence of states resulting from executing the {\em unobserved} plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
+If the state trajectory corresponding to a plan execution is available it can also be included in the compilation.
+In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state trajectories $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, that comprises the sequence of states resulting from executing the {\em unobserved} plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
 
 \begin{itemize}
-\item Fluents, $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are added to $F_{\Lambda}$ with the aim of iterating through state trajectories.
+\item Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are added to $F_{\Lambda}$ to iterate through the state trajectories.
 Likewise $I_{\Lambda}$ includes also now the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}
-\item Actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema have an extra conditional effect $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$.
-Likewise actions $\mathsf{validate_{t}}$ for {\em validating} a learning example {\tt\small $1\leq t\leq \tau$} add $at_1$ and delete $at_{|\pi_t|}$.
+\item Actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema have extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$.
+Actions $\mathsf{validate_{t}}$ for {\em validating} the $t^{th}$ learning example add $at_1$ and delete $at_{|\pi_t|}$.
 \end{itemize}
+
+{\em Linear Temporal Logic} (LTL) allows to represent more expressive state-constraints~\cite{haslum:LTL:ecai10}.
+For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define state-constraints that, unlike {\em state invariants}, must be true at least at one of the reached states.
+LTL constraints can be included in our compilation following the ideas for compiling first-order temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) adds as new goals classical planning goals the aceptor states of the equivalent automata.
 
@@
 
-Here we augment the input knowledge with the actions executed by the observed agent, when available. So the leaerning task is now defined as $\Lambda=\tup{\Psi,\Sigma,\Phi,\Pi}$
+Here we show that our classical planning compilation is also flexible to include as input constraints, if available, the actions in the plans executed by the observed agent.
+Now the learning task is defined as $\Lambda=\tup{\Psi,\Sigma,\Phi,\Pi}$, where:
 
@@
 We extend the compilation to consider the actions in the executed plans.
-Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Phi,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
+Given a learning task $\Lambda=\tup{\Psi,\Sigma,\Phi,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}$ as follows:
 \begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$. Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
-\item With respect to $A_{\Lambda'}$.
+\item $F_{\Lambda}$ is extended with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$ and {\small $1\leq j\leq |\pi_t|$}.
+\item $I_{\Lambda}$ is extended with fluents $F_{\pi_1}$. Goals are $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
+\item With respect to $A_{\Lambda}$.
 \begin{enumerate}
-\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
+\item The actions for {\em programming} the preconditions/effects of a given schema $\xi\in\Xi$ are the same.
 \item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step.
@@
 
+When the learning hypothesis space is low constrained, the learned actions can be reformulated and still be compliant with the inputs.
+For instance in the {\em blocksworld}, given a low amount of input knowledge, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa) making non trivial to compute {\em precision} and {\em recall} with respect to a reference model.
+To address this issue we define the following evaluation methodology that deals with action reformulation.
+
+Given a reference \strips\ action model $\Xi^*$ and the learned \strips\ action model $\Xi$ we define these two bijective functions $f_p:\Xi \mapsto \Xi^*$ and $f_r:\Xi \mapsto \Xi^*$ such that $f_p$ and $f_r$ respectively maximize the the accumulated {\em precision} and {\em recall}.
+With this defined we compute the {\em precision} of an \strips\ action $\xi$ with respect to the action $f_p(\xi)$.
+Likewise, the {\em recall} of an \strips\ action $\xi$ is computed with respect to the action $f_r(\xi)$.
+
 \subsection{Learning with state-constraints}
-For each domain we provide a set of {\em lifted domain mutex},  lifted domain invariants that takes the form of a binary clauses.
-For instance, in the blocksworld $\forall x,y$ then $clear(x)$ and $on(x,y)$ is a {\em lifted domain mutex}.
+For each domain we provide a set of {\em lifted domain invariants}.
 
@@
 \section{Conclusions}
-{\em Linear Temporal Logic} (LTL) also includes model operator that also allows to represent state constraints~\cite{haslum:LTL:ecai10}.
-For instance the {\em always} operator, denoted by $\square$, defines constraints that, like {\em lifted domain invariants}, must be true at any reachable state.
+
+% Commented for blind submission
+%\begin{small}
+%\subsection*{Acknowledgment}
+%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
+Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
+%\end{small}
 

### 2017-12-18 src/tim.py
tim.py: subspace analysis and other improvements

@@ -105,6 +105,9 @@ def find_types_for_property(property, patterns):
             types.append("T{}".format(j))
     return types
 
+def is_superset_state(s, other):
+    return ([k & l for k,l in zip(s, other)] == other) and any([k - l for k,l in zip(s, other)])
+
 def build_predicate_arity_map(fd_task):
     predicate_arity_map = dict()
     for pred in fd_task.predicates:
@@ -118,7 +121,7 @@ def construct_identity_invariant(p, predicate_arity_map, patterns):
     position = int(p[-1])
 
     str = "FORALL ?x:{}. ".format(" U ".join(find_types_for_property(p, patterns)))
-    str2 = "(forall (?x "
+    str2 = "(:derived (invariant-#)\n\t(forall (?x "
 
     predicate_arity = predicate_arity_map[predicate_name]
     positions = [i for i in range(1, predicate_arity+1)]
@@ -130,7 +133,7 @@ def construct_identity_invariant(p, predicate_arity_map, patterns):
             str += "".join(["FORALL ?y{}:{}.".format(i + 1, " U ".join(
                 find_types_for_property("{}_{}".format(predicate_name, positions[j]), patterns))) for j in
                             range(len(positions))])
-        str2 += " ".join("?y{}".format(i+1) for i in range(extra_variables)) + " - object)\n(not (and "
+        str2 += " ".join("?y{}".format(i+1) for i in range(extra_variables)) + " - object)\n\t\t(not (and "
 
 
         params = ["" for x in range(predicate_arity)]
@@ -143,7 +146,7 @@ def construct_identity_invariant(p, predicate_arity_map, patterns):
             if i < 1:
                 str += " AND "
         str += " -> y1 = y2"
-        str2 += " (not (= ?y1 ?y2)) )))"
+        str2 += " (not (= ?y1 ?y2)) ))))"
 
         print(str)
         print(str2)
@@ -192,12 +195,20 @@ def construct_binary_mutexes(space, predicate_arity_map, patterns, inv_propertie
             if o in patterns[patterns.keys()[i]]['objects']:
                 associated_types.add("T{}".format(i))
 
-    if len(space['states']) == 1:
-        state = space['states'][0]
+    exclusive_states = list()
+    for s1 in space['states']:
+        is_subset = False
+        for s2 in space['states']:
+            is_subset = is_subset | is_superset_state(s2, s1)
+        if not is_subset:
+            exclusive_states.append(s1)
+
[…]

### 2017-12-18 experiments/ijcai18/evaluate_empty_domain.sh
Added ijcai18 experiments scripts

@@ -1,0 +1,14 @@
+#!/bin/bash
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "hiking" "parking" "satellite" "sokoban" "transport" "zenotravel")
+
+ulimit -t 1000
+
+for i in "${arr[@]}"
+do
+   ../../src/compiler.py -s ../../benchmarks/icaps18/$i/ empty_domain test plan 3
+
+   echo $i 
+   ../../src/evaluator2.py -r ../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   cat planner_out.log | grep -e 'actions\sin'
+   cat planner_out.log | grep -e 'total\stime'
+done

### 2017-12-18 experiments/ijcai18/evaluate_empty_invariants_domain.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "hiking" "parking" "satellite" "sokoban" "transport" "zenotravel")
+
+ulimit -t 1000
+
+for i in "${arr[@]}"
+do
+   ../../src/compiler.py -s ../../benchmarks/icaps18/$i/ tim_invariants test plan 3
+
+   echo $i 
+   ../../src/evaluator2.py -r ../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   cat planner_out.log | grep -e 'actions\sin'
+   cat planner_out.log | grep -e 'total\stime'
+done

### 2017-12-18 src/evaluator2.py

@@ -149,23 +149,36 @@ else:
 
     action_combinations = combinations_by_arity[0][1]
     for i in range(1, len(combinations_by_arity)):
-        action_combinations = [zip(x, combinations_by_arity[i][1]) for x in itertools.permutations(action_combinations, len(combinations_by_arity[i][1]))]
+        # action_combinations = [zip(x, combinations_by_arity[i][1]) for x in itertools.permutations(action_combinations, len(combinations_by_arity[i][1]))]
+        aux = list()
+        for c in itertools.product(action_combinations, combinations_by_arity[i][1]):
+            aux2 = [x for x in c[0]]
+            aux2.extend(c[1])
+            aux.append(aux2)
+        action_combinations = aux
 
-    action_combinations = [item for sublist in action_combinations for item in sublist]
 
-    matching_list = list()
-    for combination in action_combinations:
-        matches = [x for x in combination[0]]
-        for i in range(1,len(combination)):
-            matches.extend(combination[i])
-        matching_list.append(matches)
+    #
+    # if len(arities) > 1:
+    #     action_combinations = [item for sublist in action_combinations for item in sublist]
+    #
+    # matching_list = list()
+    # if len(arities) > 1:
+    #     for combination in action_combinations:
+    #         matches = [x for x in combination[0]]
+    #         for i in range(1,len(combination)):
+    #             matches.extend(combination[i])
+    #         matching_list.append(matches)
+    # else:
+    #     matching_list = action_combinations
+
+    matching_list = action_combinations
 
 best_score = 0
 best_evaluation = None
 for matches in matching_list:
     evaluation = evaluate_matching(matches, fd_eva_task, fd_ref_task)
     f1_score = 2 * (evaluation[6] * evaluation[7]) / (evaluation[6] + evaluation[7])
-    print(f1_score, matches)
     if f1_score > best_score:
         best_score = f1_score
         best_evaluation = evaluation

### 2017-12-18 doc/ijcai-18/constraints-learning.tex
updates to the ijcai18 paper

@@
 \begin{abstract}
-We present a classical planning compilation for learning \strips\ action models from state-constraints.
-Remarkably the compilation does not require the precise actions in the plan executions, despite they can be included to constraint further the space of possible action models.
-A plan that solves the classical planning task resulting from our compilation induces a \strips\ action model that is compliant with all the constraints given as input.
-The paper also shows that, when input constraints are too loose, evaluating learned \strips\ models with respect to reference model becomes non trivial (actions can be reformulated, i.e.
-their semantic can be swapped) and introduces a novel evaluation method that asses learned \strips\ models even if they are reformulated.
+This paper presents a classical planning compilation for learning \strips\ action models from state-constraints.
+A plan that solves the classical planning task resulting from the compilation induces a \strips\ action model compliant with the input constraints.
+Remarkably, the compilation approach does not require observations of the particular executed actions since they are guessed by an off-the-shelf classical planner leveraging on initial, final states, and the given set of state-constraints.
+The paper also shows that evaluating learned \strips\ models with respect to a {\em reference model} is non-trivial: when input constraints are too loose, actions can be {\em reformulated} interchanging and/or redefining their semantic.
+Last but not least, the paper introduces a novel evaluation method able to asses the learning of \strips\ models even when actions are {\em reformulated}.
 \end{abstract}
@@
 Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
+At both planning tasks, automated planners are required to reason about an action model that correctly and completely captures the possible world transitions~\cite{geffner:book:2013}.
 Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
 
-The Machine Learning of planning action models is a promising alternative to hand-coding them with sophisticated algorithms like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that:
+The Machine Learning of planning action models is a promising alternative to hand-coding them and nowadays, there exist sophisticated algorithms like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for the automatic learning of \strips\ action models that:
 \begin{enumerate}
-\item Is defined as a classical planning compilation which opens the door to the bootstrapping of planning action models.
-\item Does not require the actions applied in the observed plan executions but accepts this information to improve the quality of the learned models.
-\item Assesses learned \strips\ models with respect to a reference model even when the learning hypothesis space is so low constrained that actions can be reformulated and still be compliant with the inputs.
-%For instance in the {\em blocksworld}, given a low amount of input knowledge, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa) making non trivial evaluating the learned model with respect to a reference model.
+\item Is defined as a classical planning compilation which opens the door to the {\em bootstrapping} of planning action models.
+\item Does not require observations of the particular executed actions since they are guessed by an off-the-shelf classical planner leveraging on initial, final states, and the given set of state-constraints.
+\item Assesses learned \strips\ models with respect to a {\em reference model}, even when learning is so low constrained that actions can be reformulated and still be compliant with the learning inputs.
 \end{enumerate}
@@
 \section{Background}
-This section defines the planning models used on this work and the output of the learning tasks addressed in the paper.
+This section defines the planning models used in this work, the state constraints (input of the addressed learning task) and \strips\ action model (ouput of the learning task).
 
@@
 We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
-The result of applying $a$ in $s$ is the {\em successor state} $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
 
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 We denote with $|\pi|$ the {\em plan length}.
@@
 \subsection{Classical planning with conditional effects}
-Our approach for learning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
+Our approach for learning \strips\ action models is compiling the leaning task into a classical planning task with conditional effects.
 Conditional effects allow us to compactly define actions whose effects depend on the current state.
@@
 
-An action $a\in A$ has now a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
+An action $a\in A$ is now defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
 Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
 
-An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} are the effects whose conditions hold in $s$:
+An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
 \[
@@
 
-The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered {\em negative} and {\em positive} effects, respectively.
+The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ are the triggered {\em negative} effects and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered {\em positive} effects.
 
-\subsection{\strips\ action schemes and {\em variable name} objects}
+\subsection{State-constraints}
+The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
+If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
+In this work we focus on invariants that constrain the set of possible states of a given planning task and help us to prune the space of possible \strips\ action models for that task.
+State invariants are traditionally useful in classical planning for computing more compact state representations~\cite{helmert2009concise} and for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+
+A {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ that is reachable from $I$.
+Given a planning problem $P=\tup{F,A,I,G}$, the {\em strongest invariant} is a formula $\phi_{I,A}^*$ that exactly characterizes the set of all states reachable from $I$ and using the actions $A$.
+A {\em mutex} (mutually exclusive) is a particular state invariant that takes the form of a binary clause and that represents a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+For instance in a three-blocks {\em blocksworld}, $\phi_1=\neg on(block1,block2)\vee \neg on(block1,block3)$ are mutex because $block1$ can only be on top of a single block.
+
+A {\em lifted invariant} (also called schematic invariant) is a state invariant defined using a first order formula~\cite{rintanen:schematicInvariants:AAAI2017}.
+A {\em domain invariant} is an instance-independent state invariant, i.e.
+holds for any possible initial state, so often they are lifted invariants.
+For instance in the {\em blocksworld}, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em lifted domain mutex} because the robot hand is never empty and holding a block at the same time.
+
+\subsection{\strips\ action schemes}
 This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-Figure~\ref{fig:stack} shows the schema, coded in PDDL, for the {\em stack} action from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
 
 \begin{figure}
-\begin{scriptsize}
+\begin{footnotesize}
 \begin{verbatim}
@@
   :parameters (?v1 ?v2 - object)
-  :precondition (and (holding ?v1) (clear ?v2))
-  :effect (and (not (holding ?v1)) (not (clear ?v2))
-               (handempty) (clear ?v1) (on ?v1 ?v2)))
+  :precondition (and (holding ?v1) 
+                     (clear ?v2))
+  :effect (and (not (holding ?v1)) 
+               (not (clear ?v2))
+               (handempty) (clear ?v1) 
+               (on ?v1 ?v2)))
 \end{verbatim}
-\end{scriptsize}
+\end{footnotesize}
  \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
@@
 
-\subsection{State-constraints}
-A {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ reachable from $I$.
-State invariants are traditionally useful for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-
-A {\em mutex} (mutually exclusive) is a particular state invariant that takes the form of a binary clause and that represents a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-blocks blocksworld, $on(block1,block2)$ and $on(block1,block3)$ are mutex because $block1$ can only be on top of a single block.
-
-A {\em lifted invariant} (also called schematic invariants in the literature) is a state invariant defined using a first order formula~\cite{rintanen:schematicInvariants:AAAI2017}.
-A {\em domain invariant} is a state invariant that is instance-independent (it holds for any possible initial state) and often, they are lifted invariants~\cite{fox:TIM:JAIR1998}.
-For instance in the blocksworld, $\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em lifted domain mutex} because the robot hand can never be empty and holding a block at the same time.
-
 \section{Learning \strips\ action models}
@@
 Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
-\item $\Phi$ is a set of {\em lifted domain invariants}.
+\item $\Phi$ is a set of {\em lifted domain invariants} that do not neccesary contain the {\em strongest invarriant}.
 \end{itemize}
 
-A solution to $\Lambda$ is a set of operator schema $\Xi$ compliant with the predicates in $\Psi$, the labels $\Sigma$, and the set of state-constraints $\Phi$.
-A planning compilation is a suitable approach for addressing the $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that explain $\Sigma$ and $\Phi$.
+A solution to $\Lambda$ is a set of operator schema $\Xi$ compliant with the predicates in $\Psi$, the labels $\Sigma$, and the state-constraints $\Phi$.
+A planning compilation is a suitable approach for addressing $\Lambda$ learning tasks because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that can explain $\Sigma$ and $\Phi$.
 
 \begin{figure}
-{\tt ;;; Predicates in $\Psi$}
-\begin{footnotesize}
+{\footnotesize\tt ;;; Predicates in $\Psi$}
+\begin{scriptsize}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{footnotesize}
+\end{scriptsize}
 
@@
 \begin{subfigure}{.6\textwidth}
-{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
+{\footnotesize\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
 \begin{lstlisting}[mathescape]
@@
 \begin{subfigure}{.6\textwidth}
-{\tt ;;; Lifted domain invariants in $\Phi$}
+{\footnotesize\tt ;;; Lifted domain invariants in $\Phi$}
 \begin{scriptsize}
 \begin{verbatim}
+ (forall (?o1 - object)
+   (not (and (on ?o1 ?o1))))
+
  (forall (?o1 - object)
@@
    (not (and (holding ?o1) (ontable ?o1))))
-
- (forall (?o1 - object)
-   (not (and (on ?o1 ?o1))))
 
@@
 
-\subsection{Learning \strips\ action models with classical planning}
+\subsection{Learning with classical planning}
 
@@
 A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ in the input constraints (labels $\Sigma$ and formulae $\Phi$).
-For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying every $\phi\in\Phi$ at every reached state.
+\item Validates the programmed \strips\ action model $\Xi$ in $\Sigma$ and $\Phi$.
+For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying any $\phi\in\Phi$ at every reached state.
 We call this process the validation of the programmed \strips\ action model $\Xi$, at the $t^{th}$ learning example {\small $1\leq t\leq \tau$}.
@@
 
-To formalize our compilation we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
-same fluents and actions but differ in the initial state and goals).
+To formalize our compilation we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (same fluents and actions but different initial state and/or goals).
 Fluents $F$ are built instantiating the predicates in $\Psi$ with the objects appearing in the input labels $\Sigma$.
@@
 \item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
-\item $A_{\Lambda}$ contains actions of three kinds:
+\item $A_{\Lambda}$ comprises three kinds of actions:
 \begin{enumerate}
@@
 
+\subsection{Constraining the learning hypothesis space}
+Here we introduce the {\em state constraints} in the compilation to reduce the space of the possible \strips\ action models and make learning more practicable.
+
+With regard to the {\em state invariants} $\Phi$:
+\begin{itemize}
+\item Every invariant $\phi\in\Phi$ is added as an extra precondition of the $\mathsf{apply_{\xi,\omega}}$ actions for {\em applying} an already programmed operator schema.
+\item Every invariant $\phi\in\Phi$ is added as an extra goal to the $G_{t}$, {\tt\small $1\leq t\leq \tau$}, goal sets because $\phi$ must hold at every reached state, including the last state.
+\end{itemize}
+
+Additionally, if {\em state trajectories} $\mathcal{O_{\pi}}=(s_0,s_1,\ldots,s_{n})$ resulting from observing the execution of an {\em unobserved} plan $\pi$ are available, they can be included in the compilation to constrain further the learning hypothesis space.
+In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state pairs $\sigma_t=(s_{i},s_{i+1})$, {\tt\small $1\leq i \textless n$} so only one $\mathsf{apply_{\xi,\omega}}$ action can be used to produce a state $s_{i+1}$ from a state $s_{i}$, for every {\tt\small $1\leq i \textless n$}.
+The introduction of {\em state trajectories} as new constraints to the compilation is done straigthforward by extending $F_{\Lambda}$ with a new fluent $applied$ whose negation appears as a precondition of every $\mathsf{apply_{\xi,\omega}}$ action and that is added by the effects of these same actions.
+
+Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can also be implicitely defined using {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
+For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true at least at one of the reached states.
+LTL constraints could be included in our compilation following the ideas for compiling temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) force the solution plans to always leave the LTL automata at an aceptor state by adding new goals to the classical planning task.
+
+\subsection{Compilation properties}
+
 \begin{lemma}
-Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
 \end{lemma}
@@
 
+\begin{lemma}
+Completeness. Any \strips\ action model $\Xi$ is computable by solving the corresponding classical planning task $P_{\Lambda}$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
 The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model.
-
-\section{Constraining the learning hypothesis space}
-\label{sec:Constraining}
-Here we show that constraints can be used to reduce the space of possible action models and make the learning of \strips\ action models more practicable.
-
-\subsection{State constraints}
-Every state invariant $\phi\in\Phi$ is added as an extra precondition of actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema.
-Likewise, every state invariant $\phi\in\Phi$ is also added as an extra goal in $G_{\Lambda}$ because formulae $\phi\in\Phi$ must hold at every reached state, including the last reached state.
-
-If the state trajectory corresponding to a plan execution is available it can also be included in the compilation.
-In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state trajectories $\sigma_t=(s_0^t,s_1^t,\ldots,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, that comprises the sequence of states resulting from executing the {\em unobserved} plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
-
-\begin{itemize}
-\item Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are added to $F_{\Lambda}$ to iterate through the state trajectories.
-Likewise $I_{\Lambda}$ includes also now the fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}
-\item Actions $\mathsf{apply_{\xi,\omega}}$ for {\em applying} an already programmed operator schema have extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$.
-Actions $\mathsf{validate_{t}}$ for {\em validating} the $t^{th}$ learning example add $at_1$ and delete $at_{|\pi_t|}$.
-\end{itemize}
-
-{\em Linear Temporal Logic} (LTL) allows to represent more expressive state-constraints~\cite{haslum:LTL:ecai10}.
-For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define state-constraints that, unlike {\em state invariants}, must be true at least at one of the reached states.
-LTL constraints can be included in our compilation following the ideas for compiling first-order temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) adds as new goals classical planning goals the aceptor states of the equivalent automata.
-
-\subsection{Plan constraints}
-
-Here we show that our classical planning compilation is also flexible to include as input constraints, if available, the actions in the plans executed by the observed agent.
-Now the learning task is defined as $\Lambda=\tup{\Psi,\Sigma,\Phi,\Pi}$, where:
-
-\begin{itemize}
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-\end{itemize}
-
-We extend the compilation to consider the actions in the executed plans.
-Given a learning task $\Lambda=\tup{\Psi,\Sigma,\Phi,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ that extends $P_{\Lambda}$ as follows:
-\begin{itemize}
-\item $F_{\Lambda}$ is extended with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$ and {\small $1\leq j\leq |\pi_t|$}.
-\item $I_{\Lambda}$ is extended with fluents $F_{\pi_1}$. Goals are $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
-\item With respect to $A_{\Lambda}$.
-\begin{enumerate}
-\item The actions for {\em programming} the preconditions/effects of a given schema $\xi\in\Xi$ are the same.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step.
-This mechanism forces that these actions are only applied as in the example plans.
-\item The actions for {\em validating} the current learning example have an extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
-\begin{small}
-\begin{align*}
-&\{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
-\end{align*}
+First, $F_v$ completely captures the set of all the predicates that can appear in a \strips\ action schema $\xi\in\Xi$.
+Second, any conjunction of any length of the fluents $F_v$ in the sets $pre(\xi)$,$del(\xi)$ and $add(\xi)$ is computable with a classical plan $\pi$ that solves $P_{\Lambda}$ given the proper $\Sigma$ and $\Phi$ sets.
 \end{small}
-\end{enumerate}
-\end{itemize}
+\end{proof}
 
@@
 
-\subsubsection{Setup.}
+\subsubsection{Setup and Reproducibility}
 The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
@@
 
-\subsubsection{Reproducibility.}
-We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
-
-\subsubsection{Planner.}
 The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
@@
 
-\subsubsection{Metrics.}
+We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
+
+\subsubsection{Metrics}
 The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
@@
 
-\subsection{Learning with state-constraints}
+\subsection{Learning from state-invariants}
 For each domain we provide a set of {\em lifted domain invariants}.
 
-\subsection{Learning with plans}
+\subsection{Learning from state-trajectory constraints}
 

### 2017-12-19 experiments/ijcai18/evaluate_observations.sh
Added more ijcai18 experiments

@@ -1,0 +1,14 @@
+#!/bin/bash
+declare -a arr=("blocks" "ferry" "gripper" "hanoi")
+
+ulimit -t 1000
+
+for i in "${arr[@]}"
+do
+   ../../src/compiler.py -s ../../benchmarks/icaps18/$i/ empty_domain test plan 2
+
+   echo $i 
+   ../../src/evaluator2.py -r ../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   cat planner_out.log | grep -e 'actions\sin'
+   cat planner_out.log | grep -e 'total\stime'
+done

### 2017-12-19 experiments/ijcai18/evaluate_observations_invariants.sh

@@ -1,0 +1,14 @@
+#!/bin/bash
+declare -a arr=("blocks" "ferry" "gripper" "hanoi")
+
+ulimit -t 1000
+
+for i in "${arr[@]}"
+do
+   ../../src/compiler.py -s ../../benchmarks/icaps18/$i/ tim_invariants test plan 2
+
+   echo $i 
+   ../../src/evaluator2.py -r ../../benchmarks/icaps18/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/icaps18/$i/test-1.pddl
+   cat planner_out.log | grep -e 'actions\sin'
+   cat planner_out.log | grep -e 'total\stime'
+done

### 2017-12-19 doc/ijcai-18/constraints-learning.tex
added last results by Diego to the ijcai18 paper

@@
 This paper presents a classical planning compilation for learning \strips\ action models from state-constraints.
-A plan that solves the classical planning task resulting from the compilation induces a \strips\ action model compliant with the input constraints.
-Remarkably, the compilation approach does not require observations of the particular executed actions since they are guessed by an off-the-shelf classical planner leveraging on initial, final states, and the given set of state-constraints.
-The paper also shows that evaluating learned \strips\ models with respect to a {\em reference model} is non-trivial: when input constraints are too loose, actions can be {\em reformulated} interchanging and/or redefining their semantic.
-Last but not least, the paper introduces a novel evaluation method able to asses the learning of \strips\ models even when actions are {\em reformulated}.
+A plan that solves the classical planning task resulting from the compilation induces a \strips\ action model that is compliant with the input constraints.
+The compilation approach does not require observations of the precise executed actions since they are guessed by an off-the-shelf classical planner leveraging on the given state-constraints.
+The paper shows that evaluating the learning of \strips\ action models is non-trivial.
+When the learning task is low-constrained actions can be {\em reformulated} interchanging and/or redefining their semantic and still be compliant with the learning inputs.
+The paper ends introducing a novel evaluation able to asses the learning of \strips\ models even when actions are {\em reformulated}.
 \end{abstract}
@@
 Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-At both planning tasks, automated planners are required to reason about an action model that correctly and completely captures the possible world transitions~\cite{geffner:book:2013}.
-Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
+At these planning tasks, automated planners are required to reason about an action model that correctly and completely captures the possible world transitions~\cite{geffner:book:2013}.
+Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of {\em AI planning}~\cite{kambhampati:modellite:AAAI2007}.
 
@@
 \begin{enumerate}
-\item Is defined as a classical planning compilation which opens the door to the {\em bootstrapping} of planning action models.
-\item Does not require observations of the particular executed actions since they are guessed by an off-the-shelf classical planner leveraging on initial, final states, and the given set of state-constraints.
-\item Assesses learned \strips\ models with respect to a {\em reference model}, even when learning is so low constrained that actions can be reformulated and still be compliant with the learning inputs.
+\item Is defined as a classical planning compilation, which opens the door to the {\em bootstrapping} of planning action models.
+\item Does not require observations of the particular executed actions because they are guessed by an off-the-shelf classical planner leveraging on the given state-constraints.
+\item Can assess the quality of the learned \strips\ models with respect to a {\em reference model}, even when learning is so low constrained that actions can be reformulated and still be compliant with the learning inputs.
 \end{enumerate}
@@
 \section{Background}
-This section defines the planning models used in this work, the state constraints (input of the addressed learning task) and \strips\ action model (ouput of the learning task).
+This section defines the planning models used in this work, the input of the addressed learning task (the state constraints) and the output of the learning task, an \strips\ action model.
 
@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-We denote with $|\pi|$ the {\em plan length}.
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of $\pi$ in $I$.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+The {\em plan length} is denoted with $|\pi|=n$ .
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
 
@@
 Conditional effects allow us to compactly define actions whose effects depend on the current state.
-Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
+Supporting conditional effects is now a requirement of the {\em International Planning Competition}~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
 
@@
 \subsection{State-constraints}
-The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
+The notion of state-constraint is very general and has been used in different areas of AI and for different purposes.
 If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
-In this work we focus on invariants that constrain the set of possible states of a given planning task and help us to prune the space of possible \strips\ action models for that task.
-State invariants are traditionally useful in classical planning for computing more compact state representations~\cite{helmert2009concise} and for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+For instance to specify the states where a given action is applicable or the subset of states that are considered goal states.
 
+A useful kind of state-constraints are state invariants, traditionally used for computing more compact state representations~\cite{helmert2009concise} and/or for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
 A {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ that is reachable from $I$.
-Given a planning problem $P=\tup{F,A,I,G}$, the {\em strongest invariant} is a formula $\phi_{I,A}^*$ that exactly characterizes the set of all states reachable from $I$ and using the actions $A$.
-A {\em mutex} (mutually exclusive) is a particular state invariant that takes the form of a binary clause and that represents a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-blocks {\em blocksworld}, $\phi_1=\neg on(block1,block2)\vee \neg on(block1,block3)$ are mutex because $block1$ can only be on top of a single block.
+Given a planning problem $P=\tup{F,A,I,G}$, the formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ using the actions $A$.
+A {\em mutex} (mutually exclusive) is a particular state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+For instance in a three-blocks {\em blocksworld}, $\phi_1=\neg on(block_1,block_2)\vee \neg on(block_1,block_3)$ are mutex because $block_1$ can only be on top of a single block.
 
-A {\em lifted invariant} (also called schematic invariant) is a state invariant defined using a first order formula~\cite{rintanen:schematicInvariants:AAAI2017}.
+A {\em lifted invariant} (also called schematic invariant) is a state invariant defined as a first order formula~\cite{rintanen:schematicInvariants:AAAI2017}.
 A {\em domain invariant} is an instance-independent state invariant, i.e.
@@
 For instance in the {\em blocksworld}, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em lifted domain mutex} because the robot hand is never empty and holding a block at the same time.
+This work exploits {\em domain invariants} to constrain the space of possible \strips\ action models.
 
@@
 We formalize a more challenging learning task, where less input knowledge is available.
-This learning task, denoted by $\Lambda=\tup{\Psi,\Sigma,\Phi}$, corresponds to observing an agent acting in the world but watching only the results of its plan executions:
+This learning task, denoted by $\Lambda=\tup{\Psi,\Sigma,\Phi}$, corresponds to observing an agent acting in the world but watching only the results of its plan executions.
+The actual executed actions are unobserved:
 \begin{itemize}
@@
 \item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
-Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
-\item $\Phi$ is a set of {\em lifted domain invariants} that do not neccesary contain the {\em strongest invarriant}.
+Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unobserved plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
+\item $\Phi$ is a set of {\em lifted domain invariants} that do not necessary contain the {\em strongest invariant}.
 \end{itemize}
@@
 A planning compilation is a suitable approach for addressing $\Lambda$ learning tasks because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that can explain $\Sigma$ and $\Phi$.
+Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning a \strips\ action model in the blocksworld from a single label $\sigma_1$ and a set of nine state invariants.
 
@@
 \end{subfigure}%
- \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
+ \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single label and nine state-invariants.}
 \label{fig:lexample}
@@
 \item Programs the \strips\ action model $\Xi$.
-A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+A solution plan starts with a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
 \item Validates the programmed \strips\ action model $\Xi$ in $\Sigma$ and $\Phi$.
-For every $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying any $\phi\in\Phi$ at every reached state.
+A solution plan continues with, for every $\sigma_t\in \Sigma$, a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying any constraint $\phi\in\Phi$ at every new reached state.
 We call this process the validation of the programmed \strips\ action model $\Xi$, at the $t^{th}$ learning example {\small $1\leq t\leq \tau$}.
@@
 \subsection{Constraining the learning hypothesis space}
-Here we introduce the {\em state constraints} in the compilation to reduce the space of the possible \strips\ action models and make learning more practicable.
+Here we show how {\em state constraints} are introduced into the compilation to reduce the space of the possible \strips\ action models and make learning more practicable.
+The intuition is using the input constraints as new preconditions and goals of the classical planning task that results from the compilation.
 
@@
 
-Additionally, if {\em state trajectories} $\mathcal{O_{\pi}}=(s_0,s_1,\ldots,s_{n})$ resulting from observing the execution of an {\em unobserved} plan $\pi$ are available, they can be included in the compilation to constrain further the learning hypothesis space.
-In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state pairs $\sigma_t=(s_{i},s_{i+1})$, {\tt\small $1\leq i \textless n$} so only one $\mathsf{apply_{\xi,\omega}}$ action can be used to produce a state $s_{i+1}$ from a state $s_{i}$, for every {\tt\small $1\leq i \textless n$}.
-The introduction of {\em state trajectories} as new constraints to the compilation is done straigthforward by extending $F_{\Lambda}$ with a new fluent $applied$ whose negation appears as a precondition of every $\mathsf{apply_{\xi,\omega}}$ action and that is added by the effects of these same actions.
+With regard to {\em state trajectory constraints}.
+Additionally, if {\em state trajectories} $\mathcal{O_{\pi}}=(s_0,s_1,\ldots,s_{n})$ obtained observing the execution of an {\em unobserved} plan $\pi$ are available, they can be included in the compilation to constrain further the learning hypothesis space.
+In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state pairs $\sigma_t=(s_{i},s_{i+1})$, {\tt\small $1\leq i \textless n$} s.t.
+only one $\mathsf{apply_{\xi,\omega}}$ action can be executed to produce state $s_{i+1}$ from state $s_{i}$.
+Introducing {\em state trajectories} as new constraints to the compilation is done straightforward by:
+\begin{itemize}
+\item Extending $F_{\Lambda}$ with a new fluent $applied$.
+\item For every $\mathsf{apply_{\xi,\omega}}$ action, extending its precondition set with the negation of the $applied$ fluent while its set of positive effects is extended with the $applied$ fluent.
+\item The $applied$ fluent is set to false at the initial state $I_{\Lambda}$.
+Likewise $\mathsf{validate_{t}}$ actions set the $applied$ fluent to true while this fluent is added as an extra goal to the $G_{t}$, {\tt\small $1\leq t\leq \tau$}, goal sets.
+\end{itemize}
 
@@
 \begin{lemma}
-Completeness. Any \strips\ action model $\Xi$ is computable by solving the corresponding classical planning task $P_{\Lambda}$.
+Completeness. Any \strips\ action model $\Xi$ computable from $\Lambda=\tup{\Psi,\Sigma,\Phi}$ can be obtained by solving the corresponding classical planning task $P_{\Lambda}$.
 \end{lemma}
@@
 \begin{small}
-The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model.
-First, $F_v$ completely captures the set of all the predicates that can appear in a \strips\ action schema $\xi\in\Xi$.
-Second, any conjunction of any length of the fluents $F_v$ in the sets $pre(\xi)$,$del(\xi)$ and $add(\xi)$ is computable with a classical plan $\pi$ that solves $P_{\Lambda}$ given the proper $\Sigma$ and $\Phi$ sets.
+First, $F_v$ completely captures the set of all the predicates that can appear in a \strips\ action schema $\xi\in\Xi$ given $\Psi$, $\Sigma$ and $\Phi$.
+Second, any conjunction of any length built with the fluents $F_v$, and that hence can  appear in the sets $pre(\xi)$,$del(\xi)$ and $add(\xi)$, is computable with a classical plan $\pi$ that solves $P_{\Lambda}$ since the classical planning compilation does not discard any possible \strips\ action model compliant with $\Psi$, $\Sigma$ and $\Phi$.
 \end{small}
@@
 
-\subsubsection{Setup and Reproducibility}
+\subsubsection{Setup and reproducibility}
 The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
@@
 
+\subsection{Learning from labels}
+Figure~\ref{fig:labels} shows the precision and recall values obtained when learning from state-invariants.
+Figure~\ref{fig:planlabels} shows the planning results obtained when learning from state-invariants.
+
+\begin{table}
+		\begin{center}
+\begin{scriptsize}                
+				\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+					& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+					& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+					\hline
+					Blocks & 0.86 & 0.67 & 0.71 & 0.56 & 0.86 & 0.67 & 0.81 & 0.63 \\
+					Driverlog & 0.40 & 0.14 & 0.36 & 0.57 & 0.50 & 0.29 & 0.42 & 0.33 \\
+					Ferry & 0.50 & 0.29 & 0.33 & 0.50 & 0.50 & 0.50 & 0.44 & 0.43 \\
+					Floortile & 0.67 & 0.36 & 0.67 & 0.36 & 1.0 & 0.36 & 0.78 & 0.36 \\
+					Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
+					Gripper & 1.0 & 0.67 & 0.75 & 0.75 & 0.75 & 0.75 & 0.83 & 0.72 \\
+					Hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+					Hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
+					Parking & - & - & - & - & - & - & - & - \\ % Time out 1000s
+					Satellite & 0.40 & 0.14 & 0.43 & 0.60 & 0.67 & 0.50 & 0.50 & 0.41 \\
+					Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
+					Transport & 0.75 & 0.30 & 0.67 & 0.80 & 1.0 & 0.60 & 0.81 & 0.57 \\
+					Zenotravel & 1.0 & 0.29 & 0.50 & 0.43 & 1.0 & 0.43 & 0.83 & 0.38 \\
+					\hline
+					\bf  & 0.73 & 0.36 & 0.60 & 0.62 & 0.81 & 0.57 & 0.71 & 0.52 \\
+				\end{tabular}
+	\end{scriptsize}
+		\end{center}
+ \caption{\small Precision and recall values obtained when learning from labels.}
+\label{fig:labels}                
+\end{table}
+
+\begin{table}
+	\begin{footnotesize}
+		\begin{center}
+			\begin{tabular}{l|c|c|c|}			
+				& Total time & Preprocess & Plan length  \\
+				\hline
+				Blocks & 1.40 & 0.00 & 70  \\
+				Driverlog & 1.50 & 0.00 & 89 \\
+				Ferry & 1.49 & 0.00 & 64 \\
+				Floortile & 351.38 & 0.11 & 156 \\
+				Grid & - & - & - \\ % Segmentation fault
+				Gripper & 0.04 & 0.00 & 59 \\
+				Hanoi & 2.33 & 0.01 & 49  \\
+				Hiking & - & - & - \\ % Time out 1000s
+				Parking & - & - & - \\ % Time out 1000s
+				Satellite & 78.36 & 0.03 & 98 \\
+				Sokoban & - & - & - \\ % Time out 1000s
+				Transport & 285.61 & 0.10 & 106 \\
+				Zenotravel & 6.20 & 0.46 & 71 \\
+			\end{tabular}
+		\end{center}
+	\end{footnotesize}
+ \caption{\small Planning results obtained when learning from labels.}
+\label{fig:planlabels}                        
+\end{table}
+
 \subsection{Learning from state-invariants}
-For each domain we provide a set of {\em lifted domain invariants}.
+For each domain we provide a set of {\em lifted domain invariants} that are computed using the {\sc TIM} algorithm~\cite{fox:TIM:JAIR1998}.
+Figure~\ref{fig:invariants} shows the precision and recall values obtained when learning from state-invariants.
+Figure~\ref{fig:planinvariants} shows the planning results obtained when learning from state-invariants.
+
+\begin{table}
+	\begin{scriptsize}
+		\begin{center}
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+				\hline
+				Blocks & 0.75 & 0.67 & 0.71 & 0.56 & 0.71 & 0.56 & 0.73 & 0.59 \\
+				Driverlog & 0.67 & 0.29 & 0.38 & 0.71 & 0.60 & 0.43 & 0.55 & 0.48 \\
+				Ferry & 0.75 & 0.43 & 0.75 & 0.75 & 0.75 & 0.75 & 0.75 & 0.64 \\
+				Floortile & - & - & - & - & - & - & - & - \\ % Ctrl C
+				Grid & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
+				Hanoi & 0.50 & 0.25 & 0.0 & 0.0 & 0.50 & 0.50 & 0.33 & 0.25 \\
+				Hiking & - & - & - & - & - & - & - & - \\ % Ctrl C
+				Parking & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Satellite & 0.40 & 0.14 & 0.43 & 0.60 & 0.67 & 0.50 & 0.50 & 0.41 \\
+				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Transport & - & - & - & - & - & - & - & - \\ % Ctrl C
+				Zenotravel & 1.0 & 0.29 & 0.50 & 0.43 & 1.0 & 0.43 & 0.83 & 0.38 \\
+				\hline
+				\bf  & 0.72 & 0.39 & 0.54 & 0.58 & 0.72 & 0.60 & 0.67 & 0.52 \\
+			\end{tabular}
+		\end{center}
+	\end{scriptsize}
+ \caption{\small Precision and recall values obtained when learning from labels + invariants.}
+\label{fig:invariants}
+\end{table}
+
+\begin{table}
+	\begin{footnotesize}
+		\begin{center}
+			\begin{tabular}{l|c|c|c|}			
+				& Total time & Preprocess & Plan length  \\
+				\hline
+				Blocks & 652.70 & 0.04 & 76  \\
+				Driverlog & 14.98 & 0.10 & 65 \\
+				Ferry & 1.70 & 0.03 & 58 \\
+				Floortile & - & - & - \\ % Ctrl C
+				Grid & - & - & - \\ % Time out 1000s
+				Gripper & 0.14 & 0.00 & 47 \\
+				Hanoi & 55.30 & 0.14 & 43  \\
+				Hiking & - & - & - \\ % Ctrl C
+				Parking & - & - & - \\ % Time out 1000s
+				Satellite & 84.57 & 0.22 & 98 \\
+				Sokoban & - & - & - \\ % Time out 1000s
+				Transport & - & - & - \\ % Ctrl C
+				Zenotravel & - & - & - \\ % Ctrl C
+			\end{tabular}
+		\end{center}
+	\end{footnotesize}	
+ \caption{\small Planning results obtained when learning from labels + invariants.}
+\label{fig:planinvariants}
+\end{table}
 
 \subsection{Learning from state-trajectory constraints}
+\begin{table}
+	\begin{footnotesize}
+		\begin{center}
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+				\hline
+				Blocks & 0.71 & 0.56 & 0.60 & 0.67 & 0.71 & 0.56 & 0.68 & 0.59 \\
+				Ferry & 0.75 & 0.43 & 0.50 & 0.75 & 0.75 & 0.75 & 0.67 & 0.64 \\
+				Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
+				Hanoi & 0.50 & 0.25 & 0.0 & 0.0 & 0.50 & 0.50 & 0.33 & 0.25 \\
+				\hline
+				\bf  & 0.74 & 0.48 & 0.53 & 0.61 & 0.74 & 0.70 & 0.67 & 0.57 \\
+			\end{tabular}
+		\end{center}
+	\end{footnotesize}
+  \caption{\small Precision and recall values obtained when learning from observations.}
+\label{fig:observations}       
+\end{table}
+
+\begin{table}
+	\begin{footnotesize}
+		\begin{center}
+			\begin{tabular}{l|c|c|c|}			
+				& Total time & Preprocess & Plan length  \\
+				\hline
+				Blocks & 13.36 & 0.00 & 73  \\
+				Ferry & 89.04 & 0.03 & 63 \\
+				Gripper & 0.66 & 0.00 & 43 \\
+				Hanoi & 98.66 & 0.11 & 45  \\
+			\end{tabular}
+		\end{center}
+	\end{footnotesize}
+ \caption{\small Planning results obtained when learning from observations.}
+\label{fig:planobservations}	
+\end{table}
+
+\subsection{Empty domains with invariants}
+
+\begin{table}
+	\begin{footnotesize}
+		\begin{center}
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+				\hline
+				Blocks & - & - & - & - & - & - & - & - \\ % Ctrl C
+				Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
+				Gripper & 0.75 & 0.50 & 0.75 & 0.75 & 0.75 & 0.75 & 0.75 & 0.67 \\
+				Hanoi & 0.50 & 0.25 & 0.0 & 0.0 & 0.50 & 0.50 & 0.33 & 0.25 \\
+				\hline
+				\bf & - & - & - & - & - & - & - & - \\
+			\end{tabular}
+		\end{center}
+	\end{footnotesize}
+
+  \caption{\small Precision and recall values obtained when learning from observations + invariants.}
+\label{fig:observations2}
+\end{table}
+
+\begin{table}
+	\begin{footnotesize}
+		\begin{center}
+			\begin{tabular}{l|c|c|c|}			
+				& Total time & Preprocess & Plan length  \\
+				\hline
+				Blocks & - & - & -  \\ % Ctrl C
+				Ferry & 253.04 & 0.21 & 61 \\
+				Gripper & 1.60 & 0.04 & 43 \\
+				Hanoi & 81.76 & 0.62 & 45  \\
+			\end{tabular}
+		\end{center}
+	\end{footnotesize}
+ \caption{\small Planning results obtained when learning from observations + invariants.}
+\label{fig:planobservations2}		
+\end{table}
 
@@
 
-\newpage
-
 %% The file named.bst is a bibliography style file for BibTeX 0.99c

### 2018-01-01 doc/ijcai-18/constraints-learning.tex
format updates to the ijcai18 paper

@@
 This paper presents a classical planning compilation for learning \strips\ action models from state-constraints.
-A plan that solves the classical planning task resulting from the compilation induces a \strips\ action model that is compliant with the input constraints.
-The compilation approach does not require observations of the precise executed actions since they are guessed by an off-the-shelf classical planner leveraging on the given state-constraints.
-The paper shows that evaluating the learning of \strips\ action models is non-trivial.
-When the learning task is low-constrained actions can be {\em reformulated} interchanging and/or redefining their semantic and still be compliant with the learning inputs.
-The paper ends introducing a novel evaluation able to asses the learning of \strips\ models even when actions are {\em reformulated}.
+The compilation approach does not require observations of the precise executed actions because an off-the-shelf classical planner leverages on the given constraints to determine the executed actions.
+In addition the paper introduces a novel evaluation method to asses the learning of \strips\ models even when actions are {\em reformulated} because the learning task is too low-constrained.
 \end{abstract}
@@
 The Machine Learning of planning action models is a promising alternative to hand-coding them and nowadays, there exist sophisticated algorithms like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for the automatic learning of \strips\ action models that:
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that:
 \begin{enumerate}
 \item Is defined as a classical planning compilation, which opens the door to the {\em bootstrapping} of planning action models.
-\item Does not require observations of the particular executed actions because they are guessed by an off-the-shelf classical planner leveraging on the given state-constraints.
-\item Can assess the quality of the learned \strips\ models with respect to a {\em reference model}, even when learning is so low constrained that actions can be reformulated and still be compliant with the learning inputs.
+\item Does not require observations of the particular executed actions. An off-the-shelf classical planner leverages on the given constraints to determine the executed actions.
+\item Can assess the learned \strips\ models with respect to a {\em reference model}, even when learning is so low constrained that actions are reformulated and still compliant with the learning inputs.
 \end{enumerate}
@@
 \section{Background}
-This section defines the planning models used in this work, the input of the addressed learning task (the state constraints) and the output of the learning task, an \strips\ action model.
+This section defines the planning models used in this work as well as the input (the state constraints) and output (an \strips\ action model) of the addressed learning task.
 
@@
 Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-
 An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
@@
 \]
-
 The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ are the triggered {\em negative} effects and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered {\em positive} effects.
@@
 If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
-For instance to specify the states where a given action is applicable or the subset of states that are considered goal states.
+For instance to specify the states where a given action is applicable or that are considered goal states.
+A particular kind of state-constraints useful for planning is {\em invariants}.
+{\em State invariant} are traditionally used for computing more compact state representations~\cite{helmert2009concise} and/or for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
 
-A useful kind of state-constraints are state invariants, traditionally used for computing more compact state representations~\cite{helmert2009concise} and/or for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-A {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ that is reachable from $I$.
-Given a planning problem $P=\tup{F,A,I,G}$, the formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ using the actions $A$.
-A {\em mutex} (mutually exclusive) is a particular state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ that is reachable from $I$.
+The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ using the actions in $A$.
+A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
 For instance in a three-blocks {\em blocksworld}, $\phi_1=\neg on(block_1,block_2)\vee \neg on(block_1,block_3)$ are mutex because $block_1$ can only be on top of a single block.
 
-A {\em lifted invariant} (also called schematic invariant) is a state invariant defined as a first order formula~\cite{rintanen:schematicInvariants:AAAI2017}.
 A {\em domain invariant} is an instance-independent state invariant, i.e.
-holds for any possible initial state, so often they are lifted invariants.
-For instance in the {\em blocksworld}, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em lifted domain mutex} because the robot hand is never empty and holding a block at the same time.
-This work exploits {\em domain invariants} to constrain the space of possible \strips\ action models.
+holds for any possible initial state.
+Domain invariants are often compactly defined as {\em lifted invariant} (also called schematic invariant) that is state invariants defined as a first order formula~\cite{rintanen:schematicInvariants:AAAI2017}.
+For instance in the {\em blocksworld}, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} because the robot hand is never empty and holding a block at the same time.
+In this work we exploit {\em domain invariants} to constrain the possible \strips\ action models and reduce then the learning hypothesis space.
+
+\begin{figure}
+\begin{scriptsize}
+\begin{verbatim}
+(:action stack
+ :parameters (?v1 ?v2 - object)
+ :precondition (and (holding ?v1) (clear ?v2))
+  :effect (and (not (holding ?v1)) (not (clear ?v2))
+               (handempty) (clear ?v1) (on ?v1 ?v2)))
+\end{verbatim}
+\end{scriptsize}
+ \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
+\label{fig:stack}
+\end{figure}
 
@@
 Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{verbatim}
-(:action stack
-  :parameters (?v1 ?v2 - object)
-  :precondition (and (holding ?v1) 
-                     (clear ?v2))
-  :effect (and (not (holding ?v1)) 
-               (not (clear ?v2))
-               (handempty) (clear ?v1) 
-               (on ?v1 ?v2)))
-\end{verbatim}
-\end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
-\label{fig:stack}
-\end{figure}
 
@@
 
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects $\Omega\cap\Omega_v=\emptyset$, denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters each.
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
+For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters.
 
@@
 from plans where every action in the plan is available as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
-In this case, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+%In this case, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
 Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions.
@@
 We formalize a more challenging learning task, where less input knowledge is available.
-This learning task, denoted by $\Lambda=\tup{\Psi,\Sigma,\Phi}$, corresponds to observing an agent acting in the world but watching only the results of its plan executions.
-The actual executed actions are unobserved:
+This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
+Formally the addressed learning task is defined as $\Lambda=\tup{\Psi,\Sigma,\Phi}$:
 \begin{itemize}
-\item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
+\item $\Psi$ is the set of predicates that define the abstract state space of a given classical planning frame.
 \item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
 Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unobserved plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
-\item $\Phi$ is a set of {\em lifted domain invariants} that do not necessary contain the {\em strongest invariant}.
+\item $\Phi$ is a set of {\em domain invariants}, note that they do not necessary contain the {\em strongest invariant}.
 \end{itemize}
@@
 A solution to $\Lambda$ is a set of operator schema $\Xi$ compliant with the predicates in $\Psi$, the labels $\Sigma$, and the state-constraints $\Phi$.
-A planning compilation is a suitable approach for addressing $\Lambda$ learning tasks because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that can explain $\Sigma$ and $\Phi$.
-Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning a \strips\ action model in the blocksworld from a single label $\sigma_1$ and a set of nine state invariants.
+A planning compilation is a suitable approach for addressing $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that can explain $\Sigma$ and $\Phi$.
+Figure~\ref{fig:lexample} shows an example of a $\Lambda$ task for learning a \strips\ action model in the blocksworld from a single label $\sigma_1$ and a set of nine state invariants.
 
@@
 \end{scriptsize}
-
 \vspace{0.2cm}
@@
 \vspace{-0.2cm}
-
 \begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
@@
 \end{subfigure}%
-
 \vspace{0.6cm}
-
 \begin{subfigure}{.6\textwidth}
@@
 \item Validates the programmed \strips\ action model $\Xi$ in $\Sigma$ and $\Phi$.
-A solution plan continues with, for every $\sigma_t\in \Sigma$, a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying any constraint $\phi\in\Phi$ at every new reached state.
-We call this process the validation of the programmed \strips\ action model $\Xi$, at the $t^{th}$ learning example {\small $1\leq t\leq \tau$}.
+A solution plan continues with, for every $\sigma_t\in \Sigma$, a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying the constraints $\phi\in\Phi$ at every reached state.
+We call this process the validation of the programmed \strips\ action model $\Xi$, at the $t^{th}$ learning example, {\small $1\leq t\leq \tau$}.
 %If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$}.
-\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is being validated.
+\item Fluent $mode_{prog}$ indicating whether the operator schemes are programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is validated.
 \end{itemize}
@@
 
-\begin{figure}
+\begin{figure}[hbt!]
 \begin{scriptsize}
@@
 \subsection{Constraining the learning hypothesis space}
-Here we show how {\em state constraints} are introduced into the compilation to reduce the space of the possible \strips\ action models and make learning more practicable.
-The intuition is using the input constraints as new preconditions and goals of the classical planning task that results from the compilation.
+The compilation allows to reduce the space of the possible \strips\ action models and make learning more practicable using {\em state constraints}.
+The input constraints are introduced as new preconditions and goals of the classical planning task that results from the compilation.
 
@@
 \begin{itemize}
-\item Every invariant $\phi\in\Phi$ is added as an extra precondition of the $\mathsf{apply_{\xi,\omega}}$ actions for {\em applying} an already programmed operator schema.
+\item Every invariant $\phi\in\Phi$ is added as an extra precondition of the $\mathsf{apply_{\xi,\omega}}$ actions for {\em applying} an already programmed operator schema $\xi\in\Xi$.
 \item Every invariant $\phi\in\Phi$ is added as an extra goal to the $G_{t}$, {\tt\small $1\leq t\leq \tau$}, goal sets because $\phi$ must hold at every reached state, including the last state.
@@
 
-With regard to {\em state trajectory constraints}.
 Additionally, if {\em state trajectories} $\mathcal{O_{\pi}}=(s_0,s_1,\ldots,s_{n})$ obtained observing the execution of an {\em unobserved} plan $\pi$ are available, they can be included in the compilation to constrain further the learning hypothesis space.
-In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state pairs $\sigma_t=(s_{i},s_{i+1})$, {\tt\small $1\leq i \textless n$} s.t.
-only one $\mathsf{apply_{\xi,\omega}}$ action can be executed to produce state $s_{i+1}$ from state $s_{i}$.
-Introducing {\em state trajectories} as new constraints to the compilation is done straightforward by:
+In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state pairs $\sigma_t=(s_{i},s_{i+1})$, {\tt\small $0\leq i \textless n$} s.t.
+only one $\mathsf{apply_{\xi,\omega}}$ action can be executed to produce the state $s_{i+1}$ from state $s_{i}$.
+Introducing {\em state trajectories} as additional constraints to the compilation is done straightforward by:
 \begin{itemize}
 \item Extending $F_{\Lambda}$ with a new fluent $applied$.
-\item For every $\mathsf{apply_{\xi,\omega}}$ action, extending its precondition set with the negation of the $applied$ fluent while its set of positive effects is extended with the $applied$ fluent.
+\item For every $\mathsf{apply_{\xi,\omega}}$ action, extending its precondition set with $\neg applied$ while its set of positive effects is extended with the $applied$ fluent.   
 \item The $applied$ fluent is set to false at the initial state $I_{\Lambda}$.
@@
 \end{itemize}
-
-Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can also be implicitely defined using {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
-For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true at least at one of the reached states.
-LTL constraints could be included in our compilation following the ideas for compiling temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) force the solution plans to always leave the LTL automata at an aceptor state by adding new goals to the classical planning task.
 
@@
 The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after preconditions are programmed).
+The same happens with the positive and negative effects (furthermore, effects can only be programmed after preconditions are programmed).
 Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
@@
 \begin{small}
-First, $F_v$ completely captures the set of all the predicates that can appear in a \strips\ action schema $\xi\in\Xi$ given $\Psi$, $\Sigma$ and $\Phi$.
-Second, any conjunction of any length built with the fluents $F_v$, and that hence can  appear in the sets $pre(\xi)$,$del(\xi)$ and $add(\xi)$, is computable with a classical plan $\pi$ that solves $P_{\Lambda}$ since the classical planning compilation does not discard any possible \strips\ action model compliant with $\Psi$, $\Sigma$ and $\Phi$.
+By definition, given the set of predicates $\Psi$, then $F_v\subseteq F_\Lambda$ fully captures the set of elements that can appear in a \strips\ action schema $\xi\in\Xi$.
+If $\Sigma=\emptyset$ and $\Phi=\emptyset$ are the empty set, any possible \strips\ action schema with the fluents in $F_v$ can be computed because the compilation only constrains the application of $\mathsf{apply_{\xi,\omega}}$ actions iff they are not compliant with the $\Sigma$ and $\Phi$ sets.
+If $\Sigma\neq\emptyset$ or $\Phi\neq\emptyset$ then a classical plan $\pi$ that solves $P_{\Lambda}$ cannot discard a possible \strips\ action schema $\xi\in\Xi$ that is compliant with $\Sigma$ and $\Phi$.
 \end{small}
@@
 
-\subsubsection{Setup and reproducibility}
+\subsubsection{Reproducibility}
 The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
@@
 
-Given a reference \strips\ action model $\Xi^*$ and the learned \strips\ action model $\Xi$ we define these two bijective functions $f_p:\Xi \mapsto \Xi^*$ and $f_r:\Xi \mapsto \Xi^*$ such that $f_p$ and $f_r$ respectively maximize the the accumulated {\em precision} and {\em recall}.
-With this defined we compute the {\em precision} of an \strips\ action $\xi$ with respect to the action $f_p(\xi)$.
-Likewise, the {\em recall} of an \strips\ action $\xi$ is computed with respect to the action $f_r(\xi)$.
-
-\subsection{Learning from labels}
-Figure~\ref{fig:labels} shows the precision and recall values obtained when learning from state-invariants.
-Figure~\ref{fig:planlabels} shows the planning results obtained when learning from state-invariants.
+Precision and recall are often combined using the {\em harmonic mean}.
+This expression is called the {\em F-measure} (or the balanced {\em F-score}) and is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
+Given a reference \strips\ action model $\Xi^*$ and the learned \strips\ action model $\Xi$ we define the bijective function $f_{P\&R}:\Xi \mapsto \Xi^*$ such that $f_{P\&R}$ maximizes the accumulated {\em F-measure}.
+With this mapping defined we can compute the {\em precision} and {\em recall} of a learned \strips\ action $\xi\in\Xi$ with respect to the action $f_{P\&R}(\xi)\in \Xi^*$.
+This metric allow us to asses the quality of the learned models even if actions are reformulated in the learning process.
 
@@
 		\end{center}
- \caption{\small Precision and recall values obtained when learning from labels.}
+ \caption{\small Precision and recall values obtained when learning from labels without computing the $f_{P\&R}$ mapping.}
+\label{fig:labelsnomap}
+
+\begin{center}
+\begin{scriptsize}                
+				\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+					& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+					& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+					\hline
+					Blocks & 0.86 & 0.67 & 0.71 & 0.56 & 0.86 & 0.67 & 0.81 & 0.63 \\
+					Driverlog & 0.40 & 0.14 & 0.36 & 0.57 & 0.50 & 0.29 & 0.42 & 0.33 \\
+					Ferry & 0.50 & 0.29 & 0.33 & 0.50 & 0.50 & 0.50 & 0.44 & 0.43 \\
+					Floortile & 0.67 & 0.36 & 0.67 & 0.36 & 1.0 & 0.36 & 0.78 & 0.36 \\
+					Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
+					Gripper & 1.0 & 0.67 & 0.75 & 0.75 & 0.75 & 0.75 & 0.83 & 0.72 \\
+					Hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+					Hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
+					Parking & - & - & - & - & - & - & - & - \\ % Time out 1000s
+					Satellite & 0.40 & 0.14 & 0.43 & 0.60 & 0.67 & 0.50 & 0.50 & 0.41 \\
+					Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
+					Transport & 0.75 & 0.30 & 0.67 & 0.80 & 1.0 & 0.60 & 0.81 & 0.57 \\
+					Zenotravel & 1.0 & 0.29 & 0.50 & 0.43 & 1.0 & 0.43 & 0.83 & 0.38 \\
+					\hline
+					\bf  & 0.73 & 0.36 & 0.60 & 0.62 & 0.81 & 0.57 & 0.71 & 0.52 \\
+				\end{tabular}
+	\end{scriptsize}
+		\end{center}
+ \caption{\small Precision and recall values obtained when learning from labels but computing the $f_{P\&R}$ mapping.}
 \label{fig:labels}                
 \end{table}
+
+\subsection{Learning from labels}
+Tables~\ref{fig:labelsnomap} and~\ref{fig:labels} show the precision and recall values obtained when learning from a set of $(initial, final)$ state pairs.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
+The learning examples and learned models are the same for the both tables, the only difference here is the evaluation procedure for computing the {\em precision} and {\em recall} values.
+Table~\ref{fig:labelsnomap} does not use the $f_{P\&R}$ mapping which implies assuming that the learned actions are never reformulated.
+As the results show this is a too strong assumption.
+
+Figure~\ref{fig:planlabels} shows the planning results obtained when learning from state-invariants.
+The table reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
 
@@
 
-\subsection{Learning from state-invariants}
+\subsection{Learning from labels and state-invariants}
 For each domain we provide a set of {\em lifted domain invariants} that are computed using the {\sc TIM} algorithm~\cite{fox:TIM:JAIR1998}.
 Figure~\ref{fig:invariants} shows the precision and recall values obtained when learning from state-invariants.
-Figure~\ref{fig:planinvariants} shows the planning results obtained when learning from state-invariants.
 
-\begin{table}
+\begin{table}[hbt!]
 	\begin{scriptsize}
@@
 \subsection{Learning from state-trajectory constraints}
+Figure~\ref{fig:observations2} shows the precision and recall values obtained when learning from state trajectories and using the same state invariants from the previous evaluation.
+
 \begin{table}
@@
 \end{table}
-
-\subsection{Empty domains with invariants}
 
@@
 \section{Conclusions}
+We presented a novel approach for learning \strips\ action models from state constraints using classical planning.
+As far as we know, this is the first work on learning action models exclusively using an {\em off-the-shelf} classical planner and evaluated over a wide range of different domains.
+Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
+
+The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
+In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
+
+%The size of the compiled classical planning instances depends on the number of input examples.
+Generating {\em informative} examples for learning planning action models is an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
+The combination of such exploration techniques with our learning approach is an intriguing research direction that opens the door to the bootstrapping of planning action models.
+
+Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
+For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true {\em at least at one} of the reached states.
+Despite this is beyond the scope of this paper, LTL constraints could be included in our compilation following the ideas for compiling temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) force the solution plans to always leave the LTL automata at an acceptor state by adding new goals to the classical planning task.
 

### 2018-01-04 src/compiler.py
compiler.py: provisional support for binary mutexes

@@ -145,6 +145,75 @@ def possible_pred_for_action(task, p, a, tup):
     return all(fits)
 
 
+def is_binary_mutex(axiom):
+    return isinstance(axiom.condition, pddl.UniversalCondition) and isinstance(axiom.condition.parts[0],
+                                                                           pddl.Disjunction) and len(axiom.condition.parts[0].parts) == 2 and isinstance(
+        axiom.condition.parts[0].parts[0], pddl.NegatedAtom) and isinstance(axiom.condition.parts[0].parts[1],
+                                                                            pddl.NegatedAtom)
+
+
+def get_binary_mutexes(fd_task):
+    binary_mutexes = dict()
+    for axiom in fd_task.axioms:
+        if is_binary_mutex(axiom):
+            part1 = axiom.condition.parts[0].parts[0]
+            part2 = axiom.condition.parts[0].parts[1]
+
+            args1 = part1.args
+            args2 = part2.args
+
+            arity1 = len(args1)
+            arity2 = len(args2)
+
+            matchings = list()
+            if arity1 == 0:
+                matchings.extend([(-1,i) for i in range(arity2)])
+            elif arity2 == 0:
+                matchings.extend([(i, -1) for i in range(arity2)])
+            else:
+                for i in range(arity1):
+                    for j in range(arity2):
+                        if args1[i] == args2[j]:
+                            matchings.append((i,j))
+
+
+            print(part1, part2)
+            # print(matchings)
+            for tup in itertools.product(range(1, MAX_VARS+1), repeat=max(arity1, arity2)):
+                vars = ["var" + str(t) for t in tup]
+                # print(vars)
+                m1 = [vars[i] for i in range(arity1)]
+                for tup2 in itertools.product(vars, repeat=arity2):
+                    m2 = [t for t in tup2]
+                    # print(m1, m2)
+
+                    match_all = True
+                    for matching in matchings:
+                        if matching[0] == -1 or matching[1] == -1:
+                            continue
+                        else:
+                            match_all = match_all & (m1[matching[0]] == m2[matching[1]])
+                    if match_all:
+                        key = tuple([part1.predicate] + m1)
+                        mutex = tuple([part2.predicate] + m2)
+                        if key != mutex:
[…]

### 2018-01-04 src/evaluator2.py

@@ -191,5 +191,6 @@ print("Pres: precision={}, recall={}".format(best_evaluation[0], best_evaluation
 print("Adds: precision={}, recall={}".format(best_evaluation[2], best_evaluation[3]))
 print("Dels: precision={}, recall={}".format(best_evaluation[4], best_evaluation[5]))
 print("Total: precision={}, recall={}".format(best_evaluation[6], best_evaluation[7]))
+print(" & ".join([fd_ref_task.domain_name] + [str(round(e, 2)) for e in best_evaluation]) + " \\\\")
 
 sys.exit(0)

### 2018-01-04 doc/ijcai-18/constraints-learning.tex
small  updates  to the ijcai18 paper

@@
 This paper presents a classical planning compilation for learning \strips\ action models from state-constraints.
-The compilation approach does not require observations of the precise executed actions because an off-the-shelf classical planner leverages on the given constraints to determine the executed actions.
+The compilation approach does not require observations of the precise executed actions because an off-the-shelf classical planner leverages the given constraints to determine the executed actions.
 In addition the paper introduces a novel evaluation method to asses the learning of \strips\ models even when actions are {\em reformulated} because the learning task is too low-constrained.
@@
 Our approach for learning \strips\ action models is compiling the leaning task into a classical planning task with conditional effects.
-Conditional effects allow us to compactly define actions whose effects depend on the current state.
-Supporting conditional effects is now a requirement of the {\em International Planning Competition}~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
-
 An action $a\in A$ is now defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
@@
 A planning compilation is a suitable approach for addressing $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that can explain $\Sigma$ and $\Phi$.
-Figure~\ref{fig:lexample} shows an example of a $\Lambda$ task for learning a \strips\ action model in the blocksworld from a single label $\sigma_1$ and a set of nine state invariants.
+Figure~\ref{fig:lexample} shows an example of a $\Lambda$ task for learning a \strips\ action model in the blocksworld from a single label $\sigma_1$ and a set of nine invariants.
 
@@
 \begin{subfigure}{.6\textwidth}
-{\footnotesize\tt ;;; Lifted domain invariants in $\Phi$}
+{\footnotesize\tt ;;; Domain invariants in $\Phi$}
 \begin{scriptsize}
@@
 \end{subfigure}%
- \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single label and nine state-invariants.}
+ \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single label and nine invariants.}
 \label{fig:lexample}
@@
 
-With regard to the {\em state invariants} $\Phi$:
+With regard to the {\em invariants} $\Phi$:
 \begin{itemize}

### 2018-01-04 doc/ijcai-18/constraints-learning.tex
AMAN reference added to ijcai18 paper

@@
 
-The Machine Learning of planning action models is a promising alternative to hand-coding them and nowadays, there exist sophisticated algorithms like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
+The Machine Learning of planning action models is a promising alternative to hand-coding them and nowadays, there exist sophisticated algorithms like {\sc AMAN}~\cite{zhuo2013action}, {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that:

### 2018-01-04 src/compiler.py
compiler2.py: now compiles a reversible version of the aux_domain

@@ -177,7 +177,7 @@ def get_binary_mutexes(fd_task):
                             matchings.append((i,j))
 
 
-            print(part1, part2)
+            # print(part1, part2)
             # print(matchings)
             for tup in itertools.product(range(1, MAX_VARS+1), repeat=max(arity1, arity2)):
                 vars = ["var" + str(t) for t in tup]

### 2018-01-04 src/compiler2.py

@@ -145,6 +145,75 @@ def possible_pred_for_action(task, p, a, tup):
     return all(fits)
 
 
+def is_binary_mutex(axiom):
+    return isinstance(axiom.condition, pddl.UniversalCondition) and isinstance(axiom.condition.parts[0],
+                                                                           pddl.Disjunction) and len(axiom.condition.parts[0].parts) == 2 and isinstance(
+        axiom.condition.parts[0].parts[0], pddl.NegatedAtom) and isinstance(axiom.condition.parts[0].parts[1],
+                                                                            pddl.NegatedAtom)
+
+
+def get_binary_mutexes(fd_task):
+    binary_mutexes = dict()
+    for axiom in fd_task.axioms:
+        if is_binary_mutex(axiom):
+            part1 = axiom.condition.parts[0].parts[0]
+            part2 = axiom.condition.parts[0].parts[1]
+
+            args1 = part1.args
+            args2 = part2.args
+
+            arity1 = len(args1)
+            arity2 = len(args2)
+
+            matchings = list()
+            if arity1 == 0:
+                matchings.extend([(-1,i) for i in range(arity2)])
+            elif arity2 == 0:
+                matchings.extend([(i, -1) for i in range(arity2)])
+            else:
+                for i in range(arity1):
+                    for j in range(arity2):
+                        if args1[i] == args2[j]:
+                            matchings.append((i,j))
+
+
+            # print(part1, part2)
+            # print(matchings)
+            for tup in itertools.product(range(1, MAX_VARS+1), repeat=max(arity1, arity2)):
+                vars = ["var" + str(t) for t in tup]
+                # print(vars)
+                m1 = [vars[i] for i in range(arity1)]
+                for tup2 in itertools.product(vars, repeat=arity2):
+                    m2 = [t for t in tup2]
+                    # print(m1, m2)
+
+                    match_all = True
+                    for matching in matchings:
+                        if matching[0] == -1 or matching[1] == -1:
+                            continue
+                        else:
+                            match_all = match_all & (m1[matching[0]] == m2[matching[1]])
+                    if match_all:
+                        key = tuple([part1.predicate] + m1)
+                        mutex = tuple([part2.predicate] + m2)
+                        if key != mutex:
[…]

### 2018-01-09 doc/ijcai-18/observations-learning.tex
paper updates with regard to last comments by Eva

--- 2018-01-04 doc/ijcai-18/constraints-learning.tex
@@
 
-\title{Learning \strips\ Action Models from State-Constraints}
+\title{Learning \strips\ Action Models from State Observations}
 
@@
 \begin{abstract}
-This paper presents a classical planning compilation for learning \strips\ action models from state-constraints.
-The compilation approach does not require observations of the precise executed actions because an off-the-shelf classical planner leverages the given constraints to determine the executed actions.
-In addition the paper introduces a novel evaluation method to asses the learning of \strips\ models even when actions are {\em reformulated} because the learning task is too low-constrained.
+This paper presents a classical planning compilation for learning \strips\ action models from state observations.
+The compilation approach does not require observing the precise executed actions that produced the state observations.
+These actions are determined by an off-the-shelf classical planner given the state observations (and state invariants if available).
+The paper introduces also a supervised evaluation to assess the quality of the learned \strips\ models with respect to a reference model.
 \end{abstract}
@@
 
-The Machine Learning of planning action models is a promising alternative to hand-coding them and nowadays, there exist sophisticated algorithms like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that:
+The Machine Learning of planning action models is a promising alternative to hand-coding them and nowadays, there exist sophisticated algorithms like {\sc AMAN}~\cite{zhuo2013action}, {\sc ARMS}~\cite{yang2007learning}, {\sc LOCM}~\cite{cresswell2013acquiring} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper presents a novel approach for learning \strips\ action models that introduces the following contributions:
 \begin{enumerate}
-\item Is defined as a classical planning compilation, which opens the door to the {\em bootstrapping} of planning action models.
-\item Does not require observations of the particular executed actions. An off-the-shelf classical planner leverages on the given constraints to determine the executed actions.
-\item Can assess the learned \strips\ models with respect to a {\em reference model}, even when learning is so low constrained that actions are reformulated and still compliant with the learning inputs.
+\item Is defined as a planning compilation, which opens the door to the {\em bootstrapping} of planning action models.
+\item Does not require observing the particular executed actions. An off-the-shelf classical planner determines these actions given the state observations (and state invariants if available).
+\item Assess the learned \strips\ models with respect to a {\em reference model}, even when learning is so low constrained that the learned actions are {\em reformulated} and still compliant with the learning inputs.
 \end{enumerate}
@@
 \section{Background}
-This section defines the planning models used in this work as well as the input (the state constraints) and output (an \strips\ action model) of the addressed learning task.
+This section defines the planning models used in this work as well as the input (states and state invariants) and the output (an \strips\ action model) of the addressed learning task.
 
@@
 \]
-The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ are the triggered {\em negative} effects and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered {\em positive} effects.
+The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
 
-\subsection{State-constraints}
-The notion of state-constraint is very general and has been used in different areas of AI and for different purposes.
-If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
-For instance to specify the states where a given action is applicable or that are considered goal states.
-A particular kind of state-constraints useful for planning is {\em invariants}.
-{\em State invariant} are traditionally used for computing more compact state representations~\cite{helmert2009concise} and/or for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+\subsection{State invariants}
+{\em State invariants} is a kind of state-constraints usefull for computing more compact state representations~\cite{helmert2009concise} and making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
 
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ that is reachable from $I$.
-The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ using the actions in $A$.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ reachable from $I$.
+The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
 A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-blocks {\em blocksworld}, $\phi_1=\neg on(block_1,block_2)\vee \neg on(block_1,block_3)$ are mutex because $block_1$ can only be on top of a single block.
+For instance in a three-blocks {\em blocksworld}, $\phi_1=\neg on(block_1,block_2)\vee \neg on(block_1,block_3)$ is a mutex because $block_1$ can only be on top of a single block.
 
-A {\em domain invariant} is an instance-independent state invariant, i.e.
+A {\em domain invariant} is an instance-independent invariant, i.e.
 holds for any possible initial state.
-Domain invariants are often compactly defined as {\em lifted invariant} (also called schematic invariant) that is state invariants defined as a first order formula~\cite{rintanen:schematicInvariants:AAAI2017}.
-For instance in the {\em blocksworld}, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} because the robot hand is never empty and holding a block at the same time.
-In this work we exploit {\em domain invariants} to constrain the possible \strips\ action models and reduce then the learning hypothesis space.
+Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
+For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
+%In this work we exploit {\em domain invariants} to constrain the possible \strips\ action models and reduce then the learning hypothesis space.
+
+\subsection{\strips\ action schemes}
+This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+
+To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
+Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
+Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
+$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
+For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters.
+
+Let us also define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and that defines the elements that can appear in an action schema.
+For the blocksworld, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+
+We assume also that actions $a\in A$ are instantiated from \strips\ operator schemes $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+\begin{itemize}
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
+The headers in a four-operator blocksworld are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
+\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\end{itemize}
 
@@
 
-\subsection{\strips\ action schemes}
-This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
-
-To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters.
-
-Let us also define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and that defines the elements that can appear in an action schema.
-For instance, in the blocksworld, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-
-Finally, we assume that actions $a\in A$ are instantiated from \strips\ operator schemes $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
-\begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
-For instance, the headers for a four-operator blocksworld are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
-\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-\end{itemize}
-
 \section{Learning \strips\ action models}
@@
 Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions.
-
 We formalize a more challenging learning task, where less input knowledge is available.
 This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
-Formally the addressed learning task is defined as $\Lambda=\tup{\Psi,\Sigma,\Phi}$:
+
+The addressed learning task is defined as $\Lambda=\tup{\Psi,\Sigma,\Phi}$:
 \begin{itemize}
@@
 Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unobserved plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
-\item $\Phi$ is a set of {\em domain invariants}, note that they do not necessary contain the {\em strongest invariant}.
+\item $\Phi$ is a set of {\em domain invariants}, that do not necessary contain the {\em strongest invariant}.
 \end{itemize}
@@
 
-Our approach for addressing the learning task $\Lambda$, is compiling it into a classical planning task with conditional effects.
+Our approach for addressing the learning task, is compiling it into a classical planning task with conditional effects.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 A solution plan starts with a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ in $\Sigma$ and $\Phi$.
-A solution plan continues with, for every $\sigma_t\in \Sigma$, a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$ and satisfying the constraints $\phi\in\Phi$ at every reached state.
+\item Validates the \strips\ action model $\Xi$ in $\Sigma$ and $\Phi$.
+For for every $\sigma_t\in \Sigma$, the solution plan continues with a postfix that produces a final state $s_{n}^t$ starting from $s_0^t$, using the programmed action model $\Xi$ and satisfying the constraints $\phi\in\Phi$ at every reached state.
 We call this process the validation of the programmed \strips\ action model $\Xi$, at the $t^{th}$ learning example, {\small $1\leq t\leq \tau$}.
@@
 \bibliographystyle{named}
-\bibliography{constraints-learning}
+\bibliography{observations-learning}
 

### 2018-01-09 doc/ijcai-18/observations-learning.tex
ijcai18 paper updates with regard to last comments by Eva

@@
 This paper presents a classical planning compilation for learning \strips\ action models from state observations.
-The compilation approach does not require observing the precise executed actions that produced the state observations.
+The compilation approach does not require observing the precise actions that produced the state observations.
 These actions are determined by an off-the-shelf classical planner given the state observations (and state invariants if available).
@@
 \subsection{State invariants}
-{\em State invariants} is a kind of state-constraints usefull for computing more compact state representations~\cite{helmert2009concise} and making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+{\em State invariants} are a kind of state-constraints useful in planning for computing more compact state representations~\cite{helmert2009concise} and making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
 
 Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ reachable from $I$.
-The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
-A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+The {\em strongest invariant}, that we denote with $\phi_{I,A}^*$, exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
+A {\em mutex} (mutually exclusive) is a particular state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
 For instance in a three-blocks {\em blocksworld}, $\phi_1=\neg on(block_1,block_2)\vee \neg on(block_1,block_3)$ is a mutex because $block_1$ can only be on top of a single block.
@@
 holds for any possible initial state.
-Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
+Domain invariants are often compactly defined as {\em lifted invariants}, also called schematic invariants~\cite{rintanen:schematicInvariants:AAAI2017}.
 For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
@@
 Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters.
+For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
 
@@
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
-The headers in a four-operator blocksworld are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
+The headers of a four-operator blocksworld are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
 \item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
@@
 Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unobserved plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
-\item $\Phi$ is a set of {\em domain invariants}, that do not necessary contain the {\em strongest invariant}.
+\item $\Phi$ is a set of {\em domain invariants}, that do not necessary represents the {\em strongest invariant}.
 \end{itemize}
@@
 A solution to $\Lambda$ is a set of operator schema $\Xi$ compliant with the predicates in $\Psi$, the labels $\Sigma$, and the state-constraints $\Phi$.
-A planning compilation is a suitable approach for addressing $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that can explain $\Sigma$ and $\Phi$.
+A planning compilation is a suitable approach for addressing a $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that can explain $\Sigma$ and $\Phi$.
 Figure~\ref{fig:lexample} shows an example of a $\Lambda$ task for learning a \strips\ action model in the blocksworld from a single label $\sigma_1$ and a set of nine invariants.
@@
 The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.
+Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are given by $s_n^t\in \sigma_t$.
 
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$.
-For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$}.
+For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
 \item Fluent $mode_{prog}$ indicating whether the operator schemes are programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is validated.
@@
 \item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label) and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
-Our compilation assumes that initially operator schemas are programmed with every possible precondition, no negative effect and no positive effect.
+Our compilation assumes that initially operator schemes are programmed with every possible precondition, no negative effect and no positive effect.
 \item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
@@
 
-\begin{figure}[hbt!]
+\subsection{Constraining the learning hypothesis space}
+The compilation allows to reduce the space of the possible \strips\ action models and make learning more practicable using {\em state constraints}.
+The idea is that input constraints are introduced as new preconditions/goals of the classical planning task that results from the compilation.
+
+With regard to the {\em state invariants} in $\Phi$:
+\begin{itemize}
+\item Every $\phi\in\Phi$ is added as an extra precondition of the $\mathsf{apply_{\xi,\omega}}$ actions for {\em applying} an already programmed operator schema $\xi\in\Xi$.
+\item Every $\phi\in\Phi$ is added as an extra goal to the $G_{t}$, {\tt\small $1\leq t\leq \tau$}, goal sets because $\phi$ must hold at every reached state, including the last state.
+\end{itemize}
+
+If sequences of {\em state observations} $\mathcal{O_{\pi}}=(s_0,s_1,\ldots,s_{n})$ obtained observing the execution of an {\em unobserved} plan $\pi$ are available, they can also be included in the compilation to constrain further the learning hypothesis space.
+In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state pairs $\sigma_t=(s_{i},s_{i+1})$, {\tt\small $0\leq i \textless n$} s.t.
+only one $\mathsf{apply_{\xi,\omega}}$ action can be executed to produce the state $s_{i+1}$ from state $s_{i}$.
+Introducing {\em state observations} as additional constraints to the compilation is done straightforward by:
+\begin{itemize}
+\item Extending $F_{\Lambda}$ with a new fluent $applied$.
+\item For every $\mathsf{apply_{\xi,\omega}}$ action, extending its precondition set with $\neg applied$ while its set of positive effects is extended with the $applied$ fluent.   
+\item At the initial state $I_{\Lambda}$, the $applied$ fluent is set to false.
+Likewise $\mathsf{validate_{t}}$ actions delete the $applied$ fluent while this fluent is added as an extra goal to the $G_{t}$, {\tt\small $1\leq t\leq \tau$}, goal sets.
+\end{itemize}
+
+\subsection{Compilation properties}
+
+\begin{lemma}
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
+The same happens with the positive and negative effects (furthermore, effects can only be programmed after preconditions are programmed).
+Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
+To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
+If this is achieved for every {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, solves $\Lambda$.
+\end{small}
+\end{proof}
+
+\begin{lemma}
+Completeness. Any \strips\ action model $\Xi$ computable from $\Lambda=\tup{\Psi,\Sigma,\Phi}$ can be obtained by solving the corresponding classical planning task $P_{\Lambda}$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+By definition, given the set of predicates $\Psi$, then $F_v\subseteq F_\Lambda$ fully captures the set of elements that can appear in a \strips\ action schema $\xi\in\Xi$.
+If $\Sigma=\emptyset$ and $\Phi=\emptyset$ are the empty set, any possible \strips\ action schema with the fluents in $F_v$ can be computed because the compilation only constrains the application of $\mathsf{apply_{\xi,\omega}}$ actions iff they are not compliant with the $\Sigma$ and $\Phi$ sets.
+If $\Sigma\neq\emptyset$ or $\Phi\neq\emptyset$ then a classical plan $\pi$ that solves $P_{\Lambda}$ cannot discard a possible \strips\ action schema $\xi\in\Xi$ that is compliant with $\Sigma$ and $\Phi$.
+\end{small}
+\end{proof}
+
+\begin{figure}
 \begin{scriptsize}
@@
 
-\subsection{Constraining the learning hypothesis space}
-The compilation allows to reduce the space of the possible \strips\ action models and make learning more practicable using {\em state constraints}.
-The input constraints are introduced as new preconditions and goals of the classical planning task that results from the compilation.
-
-With regard to the {\em invariants} $\Phi$:
-\begin{itemize}
-\item Every invariant $\phi\in\Phi$ is added as an extra precondition of the $\mathsf{apply_{\xi,\omega}}$ actions for {\em applying} an already programmed operator schema $\xi\in\Xi$.
-\item Every invariant $\phi\in\Phi$ is added as an extra goal to the $G_{t}$, {\tt\small $1\leq t\leq \tau$}, goal sets because $\phi$ must hold at every reached state, including the last state.
-\end{itemize}
-
-Additionally, if {\em state trajectories} $\mathcal{O_{\pi}}=(s_0,s_1,\ldots,s_{n})$ obtained observing the execution of an {\em unobserved} plan $\pi$ are available, they can be included in the compilation to constrain further the learning hypothesis space.
-In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state pairs $\sigma_t=(s_{i},s_{i+1})$, {\tt\small $0\leq i \textless n$} s.t.
-only one $\mathsf{apply_{\xi,\omega}}$ action can be executed to produce the state $s_{i+1}$ from state $s_{i}$.
-Introducing {\em state trajectories} as additional constraints to the compilation is done straightforward by:
-\begin{itemize}
-\item Extending $F_{\Lambda}$ with a new fluent $applied$.
-\item For every $\mathsf{apply_{\xi,\omega}}$ action, extending its precondition set with $\neg applied$ while its set of positive effects is extended with the $applied$ fluent.   
-\item The $applied$ fluent is set to false at the initial state $I_{\Lambda}$.
-Likewise $\mathsf{validate_{t}}$ actions set the $applied$ fluent to true while this fluent is added as an extra goal to the $G_{t}$, {\tt\small $1\leq t\leq \tau$}, goal sets.
-\end{itemize}
-
-\subsection{Compilation properties}
-
-\begin{lemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
-\end{lemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects (furthermore, effects can only be programmed after preconditions are programmed).
-Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
-If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
-\end{small}
-\end{proof}
-
-\begin{lemma}
-Completeness. Any \strips\ action model $\Xi$ computable from $\Lambda=\tup{\Psi,\Sigma,\Phi}$ can be obtained by solving the corresponding classical planning task $P_{\Lambda}$.
-\end{lemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-By definition, given the set of predicates $\Psi$, then $F_v\subseteq F_\Lambda$ fully captures the set of elements that can appear in a \strips\ action schema $\xi\in\Xi$.
-If $\Sigma=\emptyset$ and $\Phi=\emptyset$ are the empty set, any possible \strips\ action schema with the fluents in $F_v$ can be computed because the compilation only constrains the application of $\mathsf{apply_{\xi,\omega}}$ actions iff they are not compliant with the $\Sigma$ and $\Phi$ sets.
-If $\Sigma\neq\emptyset$ or $\Phi\neq\emptyset$ then a classical plan $\pi$ that solves $P_{\Lambda}$ cannot discard a possible \strips\ action schema $\xi\in\Xi$ that is compliant with $\Sigma$ and $\Phi$.
-\end{small}
-\end{proof}
-
 \section{Evaluation}
-This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts of available input knowledge.
+This section evaluates our approach for learning \strips\ models starting from different amounts of input knowledge.
 
 \subsubsection{Reproducibility}
-The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
+For the evaluation we used IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+We only use 5 labels for each domain (25 observed states per domain when using state observations, 5 intermediate states per label) and they are fixed for all the experiments so we can evaluate the impact of the available input knowledge in the quality of the learned models.
 All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
 
-The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
-In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
+{\sc Madagascar} is the classical planner we use to solve the instances that result from our compilations because its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
+In addition, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
 Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
 
-We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
+The compilation source code, the evaluation scripts and the benchmarks are fully available at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper can be reproduced.
 
-\subsubsection{Metrics}
-The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
-Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
-Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
-Recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
+\subsubsection{Supervised evaluation of the learned models}
+For each domain the learned model is compared with the actual model and its quality is quantified with the {\em precision} and {\em recall} metrics.
+Precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
+$Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
+$Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
 
 When the learning hypothesis space is low constrained, the learned actions can be reformulated and still be compliant with the inputs.
-For instance in the {\em blocksworld}, given a low amount of input knowledge, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa) making non trivial to compute {\em precision} and {\em recall} with respect to a reference model.
-To address this issue we define the following evaluation methodology that deals with action reformulation.
+For instance in the {\em blocksworld}, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa).
+Furthermore, in a given action the role of the parameters that share the same type can be interchanged making non trivial to compute {\em precision} and {\em recall} with respect to a reference model.
 
+To address these issues we defined an evaluation method robust to action reformulation.
 Precision and recall are often combined using the {\em harmonic mean}.
@@
 Given a reference \strips\ action model $\Xi^*$ and the learned \strips\ action model $\Xi$ we define the bijective function $f_{P\&R}:\Xi \mapsto \Xi^*$ such that $f_{P\&R}$ maximizes the accumulated {\em F-measure}.
-With this mapping defined we can compute the {\em precision} and {\em recall} of a learned \strips\ action $\xi\in\Xi$ with respect to the action $f_{P\&R}(\xi)\in \Xi^*$.
-This metric allow us to asses the quality of the learned models even if actions are reformulated in the learning process.
+With this mapping defined we can compute the {\em precision} and {\em recall} of a learned \strips\ action $\xi\in\Xi$ with respect to the action $f_{P\&R}(\xi)\in \Xi^*$ even if actions are reformulated in the learning process.
 
@@
 
-\subsection{Learning from labels}
-Tables~\ref{fig:labelsnomap} and~\ref{fig:labels} show the precision and recall values obtained when learning from a set of $(initial, final)$ state pairs.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
-The learning examples and learned models are the same for the both tables, the only difference here is the evaluation procedure for computing the {\em precision} and {\em recall} values.
-Table~\ref{fig:labelsnomap} does not use the $f_{P\&R}$ mapping which implies assuming that the learned actions are never reformulated.
-As the results show this is a too strong assumption.
-
-Figure~\ref{fig:planlabels} shows the planning results obtained when learning from state-invariants.
-The table reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-
 \begin{table}
@@
 
-\subsection{Learning from labels and state-invariants}
-For each domain we provide a set of {\em lifted domain invariants} that are computed using the {\sc TIM} algorithm~\cite{fox:TIM:JAIR1998}.
-Figure~\ref{fig:invariants} shows the precision and recall values obtained when learning from state-invariants.
+\subsection{Learning from labels}
+Tables~\ref{fig:labelsnomap} and~\ref{fig:labels} show the precision and recall values obtained when learning from a set of $(initial, final)$ state pairs.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
+The learning examples and learned models are the same for both tables, the only difference is the evaluation procedure for computing the {\em precision} and {\em recall} values.
+Table~\ref{fig:labelsnomap} does not use the $f_{P\&R}$ mapping which implies assuming that the learned actions are never reformulated.
+Results show this is a too strong assumption.
 
-\begin{table}[hbt!]
+The Table~\ref{fig:planlabels} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
+
+\subsubsection{Learning from labels and state invariants}
+For each domain we provide also a set of {\em domain invariants} that are computed using the {\sc TIM} algorithm~\cite{fox:TIM:JAIR1998}.
+Table~\ref{fig:invariants} shows the precision and recall values obtained when learning from state invariants and Table~\ref{fig:planinvariants} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
+
+\begin{table}
 	\begin{scriptsize}
@@
 
-\subsection{Learning from state-trajectory constraints}
-Figure~\ref{fig:observations2} shows the precision and recall values obtained when learning from state trajectories and using the same state invariants from the previous evaluation.
+\subsection{Learning from state observations}
+This is the setting with the maximum amount of available input knowledge.
+Table~\ref{fig:observations} shows the precision and recall values obtained when learning from state observations, using the same initial and  final states and the same state invariants as in the previous evaluation.
 
@@
 
-\begin{table}
-	\begin{footnotesize}
-		\begin{center}
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-				\hline
-				Blocks & - & - & - & - & - & - & - & - \\ % Ctrl C
-				Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
-				Gripper & 0.75 & 0.50 & 0.75 & 0.75 & 0.75 & 0.75 & 0.75 & 0.67 \\
-				Hanoi & 0.50 & 0.25 & 0.0 & 0.0 & 0.50 & 0.50 & 0.33 & 0.25 \\
-				\hline
-				\bf & - & - & - & - & - & - & - & - \\
-			\end{tabular}
-		\end{center}
-	\end{footnotesize}
-
-  \caption{\small Precision and recall values obtained when learning from observations + invariants.}
-\label{fig:observations2}
-\end{table}
-
-\begin{table}
-	\begin{footnotesize}
-		\begin{center}
-			\begin{tabular}{l|c|c|c|}			
-				& Total time & Preprocess & Plan length  \\
-				\hline
-				Blocks & - & - & -  \\ % Ctrl C
-				Ferry & 253.04 & 0.21 & 61 \\
-				Gripper & 1.60 & 0.04 & 43 \\
-				Hanoi & 81.76 & 0.62 & 45  \\
-			\end{tabular}
-		\end{center}
-	\end{footnotesize}
- \caption{\small Planning results obtained when learning from observations + invariants.}
-\label{fig:planobservations2}		
-\end{table}
-
 \section{Conclusions}
-We presented a novel approach for learning \strips\ action models from state constraints using classical planning.
-As far as we know, this is the first work on learning action models exclusively using an {\em off-the-shelf} classical planner and evaluated over a wide range of different domains.
-Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
+As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains.
+%Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
 
@@
 
-Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
+%Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
 For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true {\em at least at one} of the reached states.

### 2018-01-09 doc/ijcai-18/observations-learning.tex
ijcai18 paper updates with regard to last comments by Eva

@@
 
-\subsection{Classical planning with conditional effects}
+\subsubsection{Classical planning with conditional effects}
 Our approach for learning \strips\ action models is compiling the leaning task into a classical planning task with conditional effects.

### 2018-01-10 doc/ijcai-18/observations-learning.tex

@@
 				\hline
-					Blocks & 0.86 & 0.67 & 0.71 & 0.56 & 0.86 & 0.67 & 0.81 & 0.63 \\
-					Driverlog & 0.40 & 0.14 & 0.36 & 0.57 & 0.50 & 0.29 & 0.42 & 0.33 \\
-					Ferry & 0.50 & 0.29 & 0.33 & 0.50 & 0.50 & 0.50 & 0.44 & 0.43 \\
-					Floortile & 0.67 & 0.36 & 0.67 & 0.36 & 1.0 & 0.36 & 0.78 & 0.36 \\
+				blocks & 0.57 & 0.44 & 0.63 & 0.56 & 0.57 & 0.44 & 0.59 & 0.48 \\
+				driverlog & 0.2 & 0.07 & 0.18 & 0.29 & 0.2 & 0.14 & 0.19 & 0.17 \\
+				ferry & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
+				floor-tile & 0.5 & 0.14 & 0.75 & 0.55 & 0.6 & 0.27 & 0.62 & 0.32 \\
 				Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
-					Gripper & 1.0 & 0.67 & 0.75 & 0.75 & 0.75 & 0.75 & 0.83 & 0.72 \\
-					Hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-					Hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-					Parking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-					Satellite & 0.40 & 0.14 & 0.43 & 0.60 & 0.67 & 0.50 & 0.50 & 0.41 \\
+				gripper-strips & 0.25 & 0.17 & 0.25 & 0.25 & 0.25 & 0.25 & 0.25 & 0.22 \\
+				hanoi & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
+				hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				n-puzzle & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				parking & 0.6 & 0.21 & 0.14 & 0.11 & 0.4 & 0.22 & 0.38 & 0.18 \\
+				satellite & 0.25 & 0.07 & 0.33 & 0.4 & 0.67 & 0.5 & 0.42 & 0.32 \\
 				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
-					Transport & 0.75 & 0.30 & 0.67 & 0.80 & 1.0 & 0.60 & 0.81 & 0.57 \\
-					Zenotravel & 1.0 & 0.29 & 0.50 & 0.43 & 1.0 & 0.43 & 0.83 & 0.38 \\
+				transport & 0.33 & 0.1 & 0.33 & 0.4 & 0.0 & 0.0 & 0.22 & 0.17 \\
+				zeno-travel & 0.25 & 0.07 & 0.17 & 0.14 & 0.0 & 0.0 & 0.14 & 0.07 \\
 				\hline
-					\bf  & 0.73 & 0.36 & 0.60 & 0.62 & 0.81 & 0.57 & 0.71 & 0.52 \\
+				\bf  & - & - & - & - & - & - & - & - \\
 			\end{tabular}
@@
 				\hline
-					Blocks & 0.86 & 0.67 & 0.71 & 0.56 & 0.86 & 0.67 & 0.81 & 0.63 \\
-					Driverlog & 0.40 & 0.14 & 0.36 & 0.57 & 0.50 & 0.29 & 0.42 & 0.33 \\
-					Ferry & 0.50 & 0.29 & 0.33 & 0.50 & 0.50 & 0.50 & 0.44 & 0.43 \\
-					Floortile & 0.67 & 0.36 & 0.67 & 0.36 & 1.0 & 0.36 & 0.78 & 0.36 \\
+				blocks & 0.71 & 0.56 & 0.75 & 0.67 & 0.71 & 0.56 & 0.73 & 0.59 \\
+				driverlog & 0.6 & 0.21 & 0.27 & 0.43 & 0.6 & 0.43 & 0.49 & 0.36 \\
+				ferry & 0.5 & 0.29 & 0.75 & 0.75 & 0.5 & 0.5 & 0.58 & 0.51 \\
+				floor-tile & 0.5 & 0.14 & 0.75 & 0.55 & 0.6 & 0.27 & 0.62 & 0.32 \\
 				Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
-					Gripper & 1.0 & 0.67 & 0.75 & 0.75 & 0.75 & 0.75 & 0.83 & 0.72 \\
-					Hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-					Hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-					Parking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-					Satellite & 0.40 & 0.14 & 0.43 & 0.60 & 0.67 & 0.50 & 0.50 & 0.41 \\
+				gripper-strips & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
+				hanoi & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
+				hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				n-puzzle & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				parking & 0.6 & 0.21 & 0.14 & 0.11 & 0.4 & 0.22 & 0.38 & 0.18 \\
+				satellite & 0.25 & 0.07 & 0.33 & 0.4 & 0.67 & 0.5 & 0.42 & 0.32 \\
 				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
-					Transport & 0.75 & 0.30 & 0.67 & 0.80 & 1.0 & 0.60 & 0.81 & 0.57 \\
-					Zenotravel & 1.0 & 0.29 & 0.50 & 0.43 & 1.0 & 0.43 & 0.83 & 0.38 \\
+				transport & 1.0 & 0.3 & 0.67 & 0.8 & 0.67 & 0.4 & 0.78 & 0.5 \\
+				zeno-travel & 0.5 & 0.14 & 0.5 & 0.43 & 0.33 & 0.14 & 0.44 & 0.24 \\
 				\hline
-					\bf  & 0.73 & 0.36 & 0.60 & 0.62 & 0.81 & 0.57 & 0.71 & 0.52 \\
+				\bf  & - & - & - & - & - & - & - & - \\
 			\end{tabular}
@@
 
-\begin{table}
+\begin{table}[hbt!]
 	\begin{scriptsize}
@@
 				\hline
-				Blocks & 0.75 & 0.67 & 0.71 & 0.56 & 0.71 & 0.56 & 0.73 & 0.59 \\
-				Driverlog & 0.67 & 0.29 & 0.38 & 0.71 & 0.60 & 0.43 & 0.55 & 0.48 \\
-				Ferry & 0.75 & 0.43 & 0.75 & 0.75 & 0.75 & 0.75 & 0.75 & 0.64 \\
-				Floortile & - & - & - & - & - & - & - & - \\ % Ctrl C
-				Grid & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-				Hanoi & 0.50 & 0.25 & 0.0 & 0.0 & 0.50 & 0.50 & 0.33 & 0.25 \\
-				Hiking & - & - & - & - & - & - & - & - \\ % Ctrl C
+				blocks & 0.63 & 0.56 & 0.71 & 0.56 & 0.71 & 0.56 & 0.68 & 0.56 \\
+				Driverlog & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				ferry & 0.75 & 0.43 & 0.75 & 0.75 & 0.75 & 0.75 & 0.75 & 0.64 \\
+				Floortile & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
+				gripper-strips & 0.8 & 0.67 & 0.75 & 0.75 & 1.0 & 1.0 & 0.85 & 0.81 \\
+				Hanoi & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Npuzzle & - & - & - & - & - & - & - & - \\ % Time out 1000s
 				Parking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Satellite & 0.40 & 0.14 & 0.43 & 0.60 & 0.67 & 0.50 & 0.50 & 0.41 \\
+				Satellite & - & - & - & - & - & - & - & - \\ % Time out 1000s
 				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Transport & - & - & - & - & - & - & - & - \\ % Ctrl C
-				Zenotravel & 1.0 & 0.29 & 0.50 & 0.43 & 1.0 & 0.43 & 0.83 & 0.38 \\
+				Transport & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Zenotravel & - & - & - & - & - & - & - & - \\ % Time out 1000s
 				\hline
-				\bf  & 0.72 & 0.39 & 0.54 & 0.58 & 0.72 & 0.60 & 0.67 & 0.52 \\
+				\bf  & - & - & - & - & - & - & - & - \\
 			\end{tabular}
@@
 				\hline
-				Blocks & 0.71 & 0.56 & 0.60 & 0.67 & 0.71 & 0.56 & 0.68 & 0.59 \\
-				Ferry & 0.75 & 0.43 & 0.50 & 0.75 & 0.75 & 0.75 & 0.67 & 0.64 \\
-				Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-				Hanoi & 0.50 & 0.25 & 0.0 & 0.0 & 0.50 & 0.50 & 0.33 & 0.25 \\
+				blocks & 0.78 & 0.78 & 0.78 & 0.78 & 0.78 & 0.78 & 0.78 & 0.78 \\
+				driverlog & 0.75 & 0.64 & 0.86 & 0.86 & 0.86 & 0.86 & 0.82 & 0.79 \\
+				ferry & 0.86 & 0.86 & 0.75 & 0.75 & 0.75 & 0.75 & 0.79 & 0.79 \\
+				Floortile & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
+				gripper-strips & 1.0 & 0.67 & 0.8 & 1.0 & 1.0 & 1.0 & 0.93 & 0.89 \\
+				hanoi & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
+				Hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Npuzzle & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				parking & 0.8 & 0.57 & 1.0 & 0.78 & 1.0 & 0.78 & 0.93 & 0.71 \\
+				Satellite & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				Transport & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				zeno-travel & 1.0 & 0.43 & 1.0 & 0.71 & 1.0 & 0.71 & 1.0 & 0.62 \\
 				\hline
-				\bf  & 0.74 & 0.48 & 0.53 & 0.61 & 0.74 & 0.70 & 0.67 & 0.57 \\
+				\bf  & - & - & - & - & - & - & - & - \\
 			\end{tabular}

### 2018-01-11 src/compiler2.py
compiler2.py: improvements to learning with invariants

@@ -230,6 +230,12 @@ try:
     else:
         program_with_invariants = False
 
+    if "-r" in sys.argv:
+        reversible_domain = True
+        sys.argv.remove("-r")
+    else:
+        reversible_domain = False
+
     domain_folder_name  = sys.argv[1]
     domain_file = sys.argv[2]
     problems_prefix_filename = sys.argv[3]
@@ -238,7 +244,7 @@ try:
 
 except:
     print "Usage:"
-    print sys.argv[0] + "[-s] [-i] <domain> <domain filename> <problems prefix>  <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
+    print sys.argv[0] + "[-s] [-i] [-r] <domain> <domain filename> <problems prefix>  <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
     sys.exit(-1)
 
 
@@ -368,7 +374,7 @@ for a in actions:
     is_known_action = False
 
     # Add derived predicates
-    pre.extend([invariant.condition for invariant in fd_task.axioms])
+    # pre.extend([invariant.condition for invariant in fd_task.axioms])
 
     if a in known_actions:
         is_known_action = True
@@ -474,6 +480,7 @@ for a in new_actions:
     var_ids = []
     for i in range(1, len(a)):
         var_ids = var_ids + ["" + str(i)]
+    a_vars = ["var" + str(v) for v in var_ids]
     for p in predicates:
         for tup in itertools.product(var_ids, repeat=(len(p) - 1)):
             if possible_pred_for_action(fd_task, p, a, tup):
@@ -495,8 +502,9 @@ for a in new_actions:
                 if program_with_invariants:
                     key = tuple([p[0]] + vars)
                     for mutex in binary_mutexes.get(key, set()):
-                        pre = pre + [
-                            pddl.conditions.NegatedAtom("add_" + "_".join([mutex[0]] + [a[0]] + [e for e in mutex[1:]]),
+                        if set(mutex[1:]).issubset(set(a_vars)):
+                            pre = pre + [
+                                pddl.conditions.NegatedAtom("pre_" + "_".join([mutex[0]] + [a[0]] + [e for e in mutex[1:]]),
                                                     [])]
                 eff = [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom(
                     "pre_" + "_".join([p[0]] + [a[0]] + vars), []))]
@@ -505,20 +513,21 @@ for a in new_actions:
[…]

### 2018-01-11 doc/icaps-18/rebuttal/rebuttal.txt
added first version rebuttal for icaps18 paper

Reviewer1 suggested a comparison with the ARMS system. We thoughtful considered this in the composition of the paper but decided not to include it because: while the learning task addressed by ARMS is an optimization task (maximizing the number of covered examples) our learning task is a satisfying task, aims covering all the examples. The satisfying approach is relevant when learning examples are not noisy, e.g. when learning from video-games or simulators with full observability. As a matter of fact, the metric used in the ARMS AIJ paper is not informative in our case: the STRIPS models learned by our approach always cover all the learning examples (they would always report 0 errors). One can use a testing set but the particular results would depend on the content of this set. This is the reason why we evaluate our approach comparing the learned models with respect to a reference model and why we introduced the "Precision" and "Recall" metrics. Finally, we reported the results obtained with a fixed set of 5 problems because we wanted to asses how far a classical planner can get with such a limited amount of input knowledge.

With respect to the comments of reviewer2. The classical planning instances that represent the learning examples belong to the same frame, (same fluents and actions, the actual examples are available at the anonymous repository https://github.com/anonsub/strips-learning). In the final version of the paper we will introduce an example describing how we serialize plans and step through each plan. Madagascar has shown to perform well in IPC domains populated with dead-ends such as the "floortile". We also did experiments with the FD planner but the obtained results where poorer often because of the preprocessing stages of this planner.

We appreciate the comments on the learning bias suggested by the reviewer3 and we agree with him that our compilation favors extra preconditions but fewer effects since classical planners tend to minimize plan length. We plan to introduce this discussion in the final version of the paper. Our main motivation for using a classical planner for learning strips models was opening the door to the bootstrapping of planning action models, that is a planner generating its own experience and learn from them on it own the action model for a given domain. In addition, the practicality of the compilation approach allow us to easily report results over a wide range of domains (the ARMS AIJ paper reported results on 6 domains, Depots, Driverlog, Zenotravel, Satellite, Rover, and Freecell.

### 2018-01-12 doc/icaps-18/rebuttal/rebuttal.txt
rebuttal updated

@@
-Reviewer1 suggested a comparison with the ARMS system.
-We thoughtful considered this in the composition of the paper but decided not to include it because: while the learning task addressed by ARMS is an optimization task (maximizing the number of covered examples) our learning task is a satisfying task, aims covering all the examples.
-The satisfying approach is relevant when learning examples are not noisy, e.g.
-when learning from video-games or simulators with full observability.
-As a matter of fact, the metric used in the ARMS AIJ paper is not informative in our case: the STRIPS models learned by our approach always cover all the learning examples (they would always report 0 errors).
-One can use a testing set but the particular results would depend on the content of this set.
-This is the reason why we evaluate our approach comparing the learned models with respect to a reference model and why we introduced the "Precision" and "Recall" metrics.
-Finally, we reported the results obtained with a fixed set of 5 problems because we wanted to asses how far a classical planner can get with such a limited amount of input knowledge.
+General remarks
+----------------
 
-With respect to the comments of reviewer2.
-The classical planning instances that represent the learning examples belong to the same frame, (same fluents and actions, the actual examples are available at the anonymous repository https://github.com/anonsub/strips-learning).
-In the final version of the paper we will introduce an example describing how we serialize plans and step through each plan.
-Madagascar has shown to perform well in IPC domains populated with dead-ends such as the "floortile".
-We also did experiments with the FD planner but the obtained results where poorer often because of the preprocessing stages of this planner.
+Unlike extensive-data ML, this work explores an alternative research direction to learn sound models from small amounts of correct input data (only 5 plans per domain with plan lengths <10).
+The practicality of a planning compilation approach allow us to easily report results over a wide range of planning domains (the ARMS AIJ paper reported results only on 6 domains, Depots, Driverlog, Zenotravel, Satellite, Rover, and Freecell).
+In addition, using a classical planner to learn strips models opens up a way towards the bootstrapping of planning action models.
+That is, given an abstract state space (e.g.
+a set of domain predicates), a planner could explore it and use the obtained exploration data to learn and update its domain theory, learning from its own experience and gradually avoiding over-fitting.
 
-We appreciate the comments on the learning bias suggested by the reviewer3 and we agree with him that our compilation favors extra preconditions but fewer effects since classical planners tend to minimize plan length.
-We plan to introduce this discussion in the final version of the paper.
-Our main motivation for using a classical planner for learning strips models was opening the door to the bootstrapping of planning action models, that is a planner generating its own experience and learn from them on it own the action model for a given domain.
-In addition, the practicality of the compilation approach allow us to easily report results over a wide range of domains (the ARMS AIJ paper reported results on 6 domains, Depots, Driverlog, Zenotravel, Satellite, Rover, and Freecell.
+Reviewer #1
+------------
+As reviewer1 suggests, we thoughtfully considered a comparison with the ARMS system.
+We finally decided not to include it in the paper because: (1) ARMS addresses an optimization learning task (i.e., maximizing the number of covered examples) while ours is a satisfying task which aims at covering all the training samples; (2) our satisfying approach is relevant when learning samples (plans) are not noisy (for instance, learning tasks with full observability like in video-games or simulators).
+
+As a matter of fact, the metrics used in the ARMS AIJ paper would report 0 errors in the STRIPS models learned by our approach because we cover all of the learning samples.
+We believe that the ARMS metrics are relevant though but when using testing sets.
+In this case the obtained results are subject to the contents of the particular testing set (the example plans and how they were generated).
+Since we are evaluating our approach at IPC domains we determined to compare the learned models with respect to the actual models and provide an objective evidence of the performance of our learning approach.
+
+Another interesting observation pointed out by reviewer#1, is to analyze the impact and effectiveness of our method when varying the number of learning examples per domain.
+As a rule of thumb the larger is the number of examples the larger is the compilation, (it would contain a larger number of "test" fluents and "validate" actions) and the longer is the solution plan required to derive the action model.
+These facts imply larger planning times as well.
+With regard to precision and recall, these results depend on the diversity rather than on the size of the input data.
+In this first approach, and considering that our formulation covers all the input samples, we opted for using a limited amount of input knowledge, substantially smaller that the one used by ARMS in the AIJ07 paper (they used training sets with 160 plans).
+
+Reviewer #2
+------------
+
+Regarding the comments of reviewer#2: The classical planning instances that represent the learning examples belong to the same "planning frame", (same fluents and actions, the actual examples are available at the anonymous repository https://github.com/anonsub/strips-learning).
+The encoding of plans (in our compilation with \Pi) is similar to the one for "Plan recognition as planning" by Ramirez and Geffner.
+For instance, an example of a plan for a 3-blocks blocksworld: (plan-unstack i1 A B) (plan-put-down i2 A)(plan-unstack i3 B C)(plan-stack i4 B A) (inext i1 i2) (inext i2 i3) (inext i3 i4)
+
+Madagascar has shown to perform well in IPC domains populated with dead-ends such as the "floortile" ("The deterministic part of the seventh international planning competition", AIJ2015).
+We also did experiments with others classical planners but the obtained results where not as good (in terms of planning time) as the ones we achieved with Madagascar.
+The development of specific search algorithms and heuristic for this particular planning tasks is then an interesting research question.
+We agree with the Reviewer #2 that our metrics Precision and Recall, since their nature is not semantic but syntactic, they can report lower scores for action models that are actually good but correspond to "reformulations" of the reference model, possibly because the learning task is too low constrained.
+We thank reviewer#2 for the pointer to the recent work by Asai and Fukunaga paper, that we were not aware of.
+
+Reviewer #4
+------------
+
+We appreciate the comments on the learning bias suggested by reviewer#4 and we agree with him\her that our compilation favors extra preconditions but fewer effects when classical planners tend to minimize plan length.
+This is, however, not the case of Madagascar that tends to minimize the plan makes-span and, as reported, applies all the actions for programming the preconditions in a single planning step.
+We plan to introduce this discussion in the final version of the paper, if accepted.
+We also thank reviewer #4 to advertise the typo in the effects of the validate_t action that indeed reset the corresponding initial state.

### 2018-01-13 doc/icaps-18/rebuttal/rebuttal.txt
final rebuttal

--- 2018-01-11 doc/icaps-18/rebuttal/rebuttal.txt
@@
-Reviewer1 suggested a comparison with the ARMS system.
-We thoughtful considered this in the composition of the paper but decided not to include it because: while the learning task addressed by ARMS is an optimization task (maximizing the number of covered examples) our learning task is a satisfying task, aims covering all the examples.
-The satisfying approach is relevant when learning examples are not noisy, e.g.
-when learning from video-games or simulators with full observability.
-As a matter of fact, the metric used in the ARMS AIJ paper is not informative in our case: the STRIPS models learned by our approach always cover all the learning examples (they would always report 0 errors).
-One can use a testing set but the particular results would depend on the content of this set.
-This is the reason why we evaluate our approach comparing the learned models with respect to a reference model and why we introduced the "Precision" and "Recall" metrics.
-Finally, we reported the results obtained with a fixed set of 5 problems because we wanted to asses how far a classical planner can get with such a limited amount of input knowledge.
+General remarks
+==============
 
-With respect to the comments of reviewer2.
-The classical planning instances that represent the learning examples belong to the same frame, (same fluents and actions, the actual examples are available at the anonymous repository https://github.com/anonsub/strips-learning).
-In the final version of the paper we will introduce an example describing how we serialize plans and step through each plan.
-Madagascar has shown to perform well in IPC domains populated with dead-ends such as the "floortile".
-We also did experiments with the FD planner but the obtained results where poorer often because of the preprocessing stages of this planner.
+We thank the three reviewers for their insightful comments that will definitely help to improve the contents of the paper.
 
-We appreciate the comments on the learning bias suggested by the reviewer3 and we agree with him that our compilation favors extra preconditions but fewer effects since classical planners tend to minimize plan length.
-We plan to introduce this discussion in the final version of the paper.
-Our main motivation for using a classical planner for learning strips models was opening the door to the bootstrapping of planning action models, that is a planner generating its own experience and learn from them on it own the action model for a given domain.
-In addition, the practicality of the compilation approach allow us to easily report results over a wide range of domains (the ARMS AIJ paper reported results on 6 domains, Depots, Driverlog, Zenotravel, Satellite, Rover, and Freecell.
+Unlike extensive-data ML, this work explores an alternative research direction to learn sound models from small amounts of correct input data (only 5 plans per domain with plan lengths < 10 actions).
+In addition, the practicality of the planning compilation approach allow us to easily report results over a wide range of planning domains (the ARMS AIJ paper reported results only on 6 domains, Depots, Driverlog, Zenotravel, Satellite, Rover, and Freecell).
+Furthermore, using a classical planner to learn strips models opens up a way towards the bootstrapping of planning action models.
+That is, given an abstract state space (e.g.
+the set of domain predicates), a planner could explore it and use the obtained exploration data to learn and update its domain theory, learning from its own experience and gradually avoiding over-fitting.
+
+Reviewer #1
+==========
+
+As reviewer#1 suggested, we thoughtfully considered a comparison with the ARMS system.
+We finally decided not to include it in the paper because: (1) ARMS addresses an optimization task (i.e., maximizing the number of covered examples) while ours is a satisfying task which aims at covering all the training samples; (2) our satisfying approach is relevant when learning samples (plans) are not noisy (for instance, learning tasks with full observability like in video-games or simulators).
+
+We believe that ARMS metrics are relevant when using a "training-testing" approach.
+On the other hand, we followed a "training-model comparison" approach, evaluating the learned domain with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning classical planning models from the IPC.
+As a matter of fact, reporting statistically meaningful results following an evaluation approach like the one in ARMS AIJ would require a testing set far bigger than the training set we are currently using.
+We must take into account that, on average, a single plan in ARMS (out of the 160 plans in its training sets) contains around the same number of actions as our whole set of training samples.
+
+Reviewer#1 pointed out another interesting observation, the impact and effectiveness of our method when varying the number of learning examples per domain.
+As a rule of thumb, the larger the number of examples is, the larger the compilation (a larger number of "test" fluents and "validate" actions), and also the more actions the solution plan requires to derive the action model.
+All this implies larger planning times.
+On the other hand, the precision and recall values depend on the diversity of the training samples rather than on the size of the input data.
+In this first approach, and considering that our formulation covers all the input training samples, we opted for using a limited amount of input knowledge, substantially smaller than the one used by ARMS in the AIJ07 paper (they used training sets with 160 plans).
+
+Reviewer #2 
+==========
+
+We assume that all of our learning examples belong to the same "planning frame" (same sets of fluents and actions).
+The encoding of plans (in our compilation with \Pi) is similar to the one presented in the paper "Plan recognition as planning" by Ramirez and Geffner.
+For instance, an example of a plan for a 3-block blocksworld problem is: (plan-unstack i1 A B) (plan-put-down i2 A)(plan-unstack i3 B C)(plan-stack i4 B A) (inext i1 i2) (inext i2 i3) (inext i3 i4).
+
+Madagascar has shown to perform well in the IPC domains populated with dead-ends such as the "floortile" domain ("The deterministic part of the seventh international planning competition", AIJ2015).
+We also did experiments with other classical planners but the obtained results were not as good (in terms of planning time) as the ones we achieved with Madagascar.
+The development of specific search algorithms and heuristics for this particular planning task is then an interesting research question.
+
+We agree with the reviewer#2 on the relevant observation about the precision and recall metrics.
+Given the syntax-based nature of these metrics, it may happen that these metrics report low scores for learned models that are actually good but correspond to "reformulations" of the actual model; i.e.
+a learned model semantically equivalent but syntactically different to the reference model.
+This most likely occurs when the learning task is under-constrained.
+
+We thank the reviewer for the pointer to the recent work in the paper by Asai and Fukunaga, which we were not aware of.
+
+Reviewer #4 
+==========
+
+We appreciate the comments on the learning bias suggested by the reviewer#4.
+We agree that our compilation favors extra preconditions but fewer effects when classical planners tend to minimize plan length.
+This, however, is not the case of Madagascar that tends to minimize the plan makespan and, as reported, applies all the actions for programming the preconditions of the compilation scheme in a single planning step.
+We plan to introduce this discussion in the final version of the paper, if accepted.
+
+We also thank the reviewer to notice the typo in the effects of the validate_t action, which indeed resets the corresponding initial state.

### 2018-01-18 src/validator.py
validator.py: test set evaluation

@@ -1,0 +1,501 @@
+#! /usr/bin/env python
+import glob, os, sys, copy, itertools, math
+import pddl, pddl_parser
+import config, fdtask_to_pddl
+
+def get_all_types(task, itype):
+   output=[itype]
+   for t in task.types:
+       if t.basetype_name == itype:
+           output.append(str(t.name))
+   return output
+
+
+def get_max_steps_from_plans(ps):
+    iout = 0
+    for plan in ps:
+        iout = max(iout, len(plan))
+    return iout
+
+
+def get_max_vars_from_plans(ps):
+    iout = 0
+    for plan in ps:
+        for a in plan:
+            iout = max(iout, len(a.split(" ")) - 1)
+    return iout
+
+
+def get_action_schema_from_plans(ps, task):
+    schemas = []
+    for plan in ps:
+        for a in plan:
+            counter = 0
+            name = a.replace("(", "").replace(")", "").split(" ")[0]
+            item = [name]
+            for p in a.replace("(", "").replace(")", "").split(" ")[1:]:
+                for o in task.objects:
+                    if p.upper() == o.name.upper():
+                        item.append(str(o.type_name))
+                        counter = counter + 1
+                        break
+            if item not in schemas:
+                schemas.insert(0, item)
+    return schemas
+
+
+def get_predicates_schema_from_plans(task):
+    preds = []
+    for p in task.predicates:
+        item = []
+        if p.name == "=":
+            continue
+        item.append(p.name)
+        for a in p.arguments:
+            item.append(a.type_name)
+        preds = preds + [item]
+    return preds
+
+
+def possible_pred_for_action(task, p, a, tup):
+    if (len(p) > len(a)):
+        return False
+
+    action_types = [set([a[int(tup[i])]]) for i in range(len(tup))]
+    predicate_types = [set(get_all_types(task, x)) for x in p[1:]]
+    fits = [len(action_types[i].intersection(predicate_types[i])) >= 1 for i in range(len(action_types))]
+    return all(fits)
+
+
+def get_fluents_from_model(task):
+    fluents = list()
+    for a in task.actions:
+        params = [x.name for x in a.parameters]
+        for pre in a.precondition.parts:
+            vars = list()
+            for arg in pre.args:
[…]

### 2018-01-18 doc/ijcai-18/observations-learning.tex

@@
 
-\title{Learning \strips\ Action Models from State Observations}
+\title{Learning and Recognition of \strips\ Action Models from State Observations}
 
@@
 This paper presents a classical planning compilation for learning \strips\ action models from state observations.
-The compilation approach does not require observing the precise actions that produced the state observations.
-These actions are determined by an off-the-shelf classical planner given the state observations (and state invariants if available).
-The paper introduces also a supervised evaluation to assess the quality of the learned \strips\ models with respect to a reference model.
+The compilation approach does not require observing the precise actions that produced the state observations because such actions are determined by an off-the-shelf classical planner.
+In addition, the compilation is extensible to estimate the probability distribution of the possible \strips\ models given a sequence of state observations.
+This extension allows us to address model recognition tasks as well as to semantically evaluate the quality of learned \strips\ models.
 \end{abstract}
@@
 \begin{enumerate}
-\item Is defined as a planning compilation, which opens the door to the {\em bootstrapping} of planning action models.
-\item Does not require observing the particular executed actions. An off-the-shelf classical planner determines these actions given the state observations (and state invariants if available).
-\item Assess the learned \strips\ models with respect to a {\em reference model}, even when learning is so low constrained that the learned actions are {\em reformulated} and still compliant with the learning inputs.
+\item Is defined as a planning compilation. This fact opens the door to the {\em bootstrapping} of planning action models and allows us to report results over a wide range of planning domains. 
+\item Does not require observing the particular executed actions. An off-the-shelf classical planner determines these actions given the state observations.
+\item Can estimate the probability distribution of the possible \strips\ models, given an observation sequence.
+This extension allow us to address model recognition tasks as well as to evaluate the quality of the learned models semantically, without comparing with a reference model.
 \end{enumerate}
@@
 Our approach for learning \strips\ action models is compiling the leaning task into a classical planning task with conditional effects.
+
 An action $a\in A$ is now defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
@@
 \]
+
 The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
-
-\subsection{State invariants}
-{\em State invariants} are a kind of state-constraints useful in planning for computing more compact state representations~\cite{helmert2009concise} and making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ reachable from $I$.
-The {\em strongest invariant}, that we denote with $\phi_{I,A}^*$, exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
-A {\em mutex} (mutually exclusive) is a particular state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-blocks {\em blocksworld}, $\phi_1=\neg on(block_1,block_2)\vee \neg on(block_1,block_3)$ is a mutex because $block_1$ can only be on top of a single block.
-
-A {\em domain invariant} is an instance-independent invariant, i.e.
-holds for any possible initial state.
-Domain invariants are often compactly defined as {\em lifted invariants}, also called schematic invariants~\cite{rintanen:schematicInvariants:AAAI2017}.
-For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
-%In this work we exploit {\em domain invariants} to constrain the possible \strips\ action models and reduce then the learning hypothesis space.
 
@@
 
+Finally we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in a given action schema $\xi$.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} only contains the fluents from $F_v$ that do not involve $v_2$.
+
 \begin{figure}
@@
 from plans where every action in the plan is available as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
-%In this case, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+In this case, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
 Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions.
+
 We formalize a more challenging learning task, where less input knowledge is available.
-This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
-
-The addressed learning task is defined as $\Lambda=\tup{\Psi,\Sigma,\Phi}$:
+The addressed learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
+This learning task is defined as $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$:
 \begin{itemize}
 \item $\Psi$ is the set of predicates that define the abstract state space of a given classical planning frame.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
-Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unobserved plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
-\item $\Phi$ is a set of {\em domain invariants}, that do not necessary represents the {\em strongest invariant}.
+\item $\Xi$ is the set of empty operator schemes containing only the headers with the name and parameter of each operator schema $\xi\in\Xi$.
+\item $\mathcal{O}=\{s_0,s_1,\ldots,s_{n}\}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$.
 \end{itemize}
 
-A solution to $\Lambda$ is a set of operator schema $\Xi$ compliant with the predicates in $\Psi$, the labels $\Sigma$, and the state-constraints $\Phi$.
-A planning compilation is a suitable approach for addressing a $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi$ but also, the {\em unobserved} plans $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} that can explain $\Sigma$ and $\Phi$.
-Figure~\ref{fig:lexample} shows an example of a $\Lambda$ task for learning a \strips\ action model in the blocksworld from a single label $\sigma_1$ and a set of nine invariants.
+A solution to $\Lambda$ is a set of operator schema $\Xi'$ compliant with the predicates in $\Psi$, the headers in $\Xi$ and the sequence of state observations $\mathcal{O}$.
+A planning compilation is a suitable approach for addressing a $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that can explain $\mathcal{O}$.
+Figure~\ref{fig:lexample} shows an example of a $\Lambda$ task for learning a \strips\ action model in the blocksworld from the sequence of five state observations that corresponds to inverting a 2-blocks tower.
 
@@
 \vspace{0.2cm}
-\begin{subfigure}{.6\textwidth}
-{\footnotesize\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
-\begin{lstlisting}[mathescape]
-\end{lstlisting}
-\vspace{-0.2cm}
-\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
-\node (initD) [draw]{D};
-\node (initC) [draw, above=of initD.north]{C};
-\node (initB) [draw, above=of initC.north]{B};
-\node (initA) [draw, above=of initB.north]{A};
-\draw[thick] (-1,-0.25) -- (2.5,-0.25);
-
-\node (goalA) [draw, right=10mm of initD]{A};
-\node (goalB) [draw, right=10mm of initC]{B};
-\node (goalC) [draw, right=10mm of initB]{C};
-\node (goalD) [draw, right=10mm of initA]{D};
-\end{tikzpicture}
-\end{subfigure}%
-\vspace{0.6cm}
-\begin{subfigure}{.6\textwidth}
-{\footnotesize\tt ;;; Domain invariants in $\Phi$}
+{\footnotesize\tt ;;; Headers in $\Xi$}
 \begin{scriptsize}
 \begin{verbatim}
- (forall (?o1 - object)
-   (not (and (on ?o1 ?o1))))
-
- (forall (?o1 - object)
-   (not (and (handempty) (holding ?o1)))))
-
- (forall (?o1 - object)
-   (not (and (holding ?o1) (clear ?o1)))))
-
- (forall (?o1 - object)
-   (not (and (holding ?o1) (ontable ?o1))))
-
- (forall (?o1 ?o2 - object)
-   (not (and (on ?o1 ?o2) (holding ?o1))))
-
- (forall (?o1 ?o2 - object)
-   (not (and (on ?o1 ?o2) (holding ?o2))))
-
- (forall (?o1 ?o2 - object)
-    (not (and (on ?o1 ?o2) (clear ?o2))))
-
- (forall (?o1 ?o2 - object)
-    (not (and (on ?o1 ?o2) (ontable ?o1))))
-
- (forall (?o1 ?o2 - object)
-    (not (and (on ?o1 ?o2) (on ?o2 ?o1)))))
+(pickup v1) (putdown v1) (stack v1 v2} (unstack v1 v2)
 \end{verbatim}
 \end{scriptsize}
-\end{subfigure}%
- \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single label and nine invariants.}
+\vspace{0.2cm}
+{\footnotesize\tt ;;; Observations in $\mathcal{O}$}
+\begin{scriptsize}
+\begin{verbatim}
+;;; observation #0
+(clear block2) (on block2 block1) (ontable block1) 
+(handempty)
+
+;;; observation #1
+(holding block2) (clear block1) (ontable block1)
+
+;;; observation #2
+(clear block1) (ontable block1) (clear block2) 
+(ontable block2) (handempty)
+
+;;; observation #3
+(holding block1) (clear block2) (ontable block2)
+
+;;; observation #4
+(clear block1) (on block1 block2) (ontable block2) 
+(handempty) 
+\end{verbatim}
+\end{scriptsize}
+ \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a sequence of five state observations.}
 \label{fig:lexample}
@@
 \begin{enumerate}
-\item Programs the \strips\ action model $\Xi$.
+\item Programs the \strips\ action model $\Xi'$.
 A solution plan starts with a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the \strips\ action model $\Xi$ in $\Sigma$ and $\Phi$.
-For for every $\sigma_t\in \Sigma$, the solution plan continues with a postfix that produces a final state $s_{n}^t$ starting from $s_0^t$, using the programmed action model $\Xi$ and satisfying the constraints $\phi\in\Phi$ at every reached state.
-We call this process the validation of the programmed \strips\ action model $\Xi$, at the $t^{th}$ learning example, {\small $1\leq t\leq \tau$}.
-%If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
+\item Validates the \strips\ action model $\Xi'$ in $\mathcal{O}$.
+The solution plan continues with a postfix that produces the sequence of states in $\mathcal{O}$ starting from $s_0$, and using the programmed action model $\Xi'$.
 \end{enumerate}
 
-To formalize our compilation we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (same fluents and actions but different initial state and/or goals).
-Fluents $F$ are built instantiating the predicates in $\Psi$ with the objects appearing in the input labels $\Sigma$.
-Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
-The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are given by $s_n^t\in \sigma_t$.
-
-Now we are ready to formalize the compilation.
-Given a learning task $\Lambda=\tup{\Psi,\Sigma,\Phi}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+Given a learning task $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
 \begin{itemize}
-\item $F_{\Lambda}$ extends $F$ with:
+\item $F_{\Lambda}$ contains:
 \begin{itemize}
-\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$.
+\item Fluents $F$ built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$. 
+\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$.
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$.
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
-\item Fluent $mode_{prog}$ indicating whether the operator schemes are programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is validated.
+\item Fluent $mode_{prog}$ indicating whether the operator schemes are programmed or validated (already programmed) and fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
 \end{itemize}
-\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label) and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
+\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0$ (the first observation) and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
 Our compilation assumes that initially operator schemes are programmed with every possible precondition, no negative effect and no positive effect.
-\item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
+\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, indicates that the programmed action model is validated in all the input observations.
 \item $A_{\Lambda}$ comprises three kinds of actions:
@@
 \begin{itemize}
-\item Actions for {\bf removing} a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
+\item Actions for {\bf removing} a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\Xi$.
 
@@
 
-\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
+\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\Xi$.
 
@@
 
-\item Actions for {\em validating} learning example {\tt\small $1\leq t\leq \tau$}. 
+\item Actions for {\em validating} an observation {\tt\small $1\leq i\leq n$}. 
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\\
-&\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\\
+&\cup\{\neg test_j\}_{j\in t\leq j\leq n}\cup \{\neg mode_{prog}\},\\
+\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i\}.
 \end{align*}
@@
 \end{enumerate}
-\end{itemize}
-
-\subsection{Constraining the learning hypothesis space}
-The compilation allows to reduce the space of the possible \strips\ action models and make learning more practicable using {\em state constraints}.
-The idea is that input constraints are introduced as new preconditions/goals of the classical planning task that results from the compilation.
-
-With regard to the {\em state invariants} in $\Phi$:
-\begin{itemize}
-\item Every $\phi\in\Phi$ is added as an extra precondition of the $\mathsf{apply_{\xi,\omega}}$ actions for {\em applying} an already programmed operator schema $\xi\in\Xi$.
-\item Every $\phi\in\Phi$ is added as an extra goal to the $G_{t}$, {\tt\small $1\leq t\leq \tau$}, goal sets because $\phi$ must hold at every reached state, including the last state.
-\end{itemize}
-
-If sequences of {\em state observations} $\mathcal{O_{\pi}}=(s_0,s_1,\ldots,s_{n})$ obtained observing the execution of an {\em unobserved} plan $\pi$ are available, they can also be included in the compilation to constrain further the learning hypothesis space.
-In this case $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is no longer a set of $(initial, final)$ state pairs but a set of state pairs $\sigma_t=(s_{i},s_{i+1})$, {\tt\small $0\leq i \textless n$} s.t.
-only one $\mathsf{apply_{\xi,\omega}}$ action can be executed to produce the state $s_{i+1}$ from state $s_{i}$.
-Introducing {\em state observations} as additional constraints to the compilation is done straightforward by:
-\begin{itemize}
-\item Extending $F_{\Lambda}$ with a new fluent $applied$.
-\item For every $\mathsf{apply_{\xi,\omega}}$ action, extending its precondition set with $\neg applied$ while its set of positive effects is extended with the $applied$ fluent.   
-\item At the initial state $I_{\Lambda}$, the $applied$ fluent is set to false.
-Likewise $\mathsf{validate_{t}}$ actions delete the $applied$ fluent while this fluent is added as an extra goal to the $G_{t}$, {\tt\small $1\leq t\leq \tau$}, goal sets.
 \end{itemize}
@@
 Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
-If this is achieved for every {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, solves $\Lambda$.
+To solve $P_{\Lambda}$, goals $\{test_i\}$, {\small $1\leq i\leq n$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the state $s_i$, starting from $s_0$.
+If this is achieved for every {\small $1\leq i\leq n$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, solves $\Lambda$.
 \end{small}
@@
 \begin{lemma}
-Completeness. Any \strips\ action model $\Xi$ computable from $\Lambda=\tup{\Psi,\Sigma,\Phi}$ can be obtained by solving the corresponding classical planning task $P_{\Lambda}$.
+Completeness. Any \strips\ action model $\Xi$ computable from $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ can be obtained by solving the corresponding classical planning task $P_{\Lambda}$.
 \end{lemma}
@@
 \begin{small}
-By definition, given the set of predicates $\Psi$, then $F_v\subseteq F_\Lambda$ fully captures the set of elements that can appear in a \strips\ action schema $\xi\in\Xi$.
-If $\Sigma=\emptyset$ and $\Phi=\emptyset$ are the empty set, any possible \strips\ action schema with the fluents in $F_v$ can be computed because the compilation only constrains the application of $\mathsf{apply_{\xi,\omega}}$ actions iff they are not compliant with the $\Sigma$ and $\Phi$ sets.
-If $\Sigma\neq\emptyset$ or $\Phi\neq\emptyset$ then a classical plan $\pi$ that solves $P_{\Lambda}$ cannot discard a possible \strips\ action schema $\xi\in\Xi$ that is compliant with $\Sigma$ and $\Phi$.
+By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the set of elements that can appear in a \strips\ action schema $\xi\in\Xi$.
+Any possible \strips\ action schema built with the fluents in $F_v$ can be computed because the compilation only constrains the application of $\mathsf{apply_{\xi,\omega}}$ actions iff they are not compliant with the $\mathcal{O}$ sequence.
 \end{small}
@@
 
+\section{Recognizing \strips\ action models}
+Inspired by the {\em Plan recognition as planning} approach~\cite{ramirez2009plan} we show that the previous compilation can be extended to estimate the probability distribution of the possible \strips\ models given a sequence of state observations.
+This extension allows us to address model recognition tasks as well as to semantically evaluate the quality of learned \strips\ models
+
+\subsection{The \strips\ edit distance}
+We assume that how well a given \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$, depends on the amount of edits that one has to introduce to $\Xi$ to produce the sequence of obsevations $\mathcal{O}$.
+If the given model $\Xi$ perfectly explains the observations $\mathcal{O}$, no edit have to be introduced and the probability of explanning the observations is $P(\Xi|\mathcal{O})=1$.
+On the other hand, with a probability of explanning the observations of $P(\Xi|\mathcal{O})=0$, is the model that requires the maximum number of edits to explain the sequence of observations $\mathcal{O}$.
+
+To compute the $P(\Xi|\mathcal{O})$ probability distribution, we need a metric that quantifies this amount of edits.
+To define such metric, we need to define first which are the possible edit operations that can be performed on a \strips\ action model.
+With the aim of keeping tractable the branching factor of the classical planning task resulting from our compilation, we only define two \strips\ edit operations:
+\begin{itemize}
+\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi$
+\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi$
+\end{itemize}
+
+Now we are ready to define a metric that asseses then how dissimilar two given \strips\ action models are
+\begin{definition}
+The {\bf \strips\ edit distance}, denoted as $\delta(\Xi,\Xi')$, is the minimum number of edit operations that allow to transform a given \strips\ action model $\Xi$ into the given \strips\ action model $\Xi'$.
+This edit distance satisfies the {\em metric axioms} provided that the two operations has the same positive cost.
+\end{definition}
+
+Since the size of the $F_v$ set is bound, we can define the maximum number of edits that can be introduced to a given \strips\ action model.
+\begin{definition}
+The maximum {\bf \strips\ edit distance} is an upperbound, denoted as $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$, that indicates the maximum distance required to transform a given \strips\ action model $\Xi$ into any \strips\ action model $\Xi'$.
+\end{definition}
+
+In more detail, for a given operator schema $\xi$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$.
+With regard to action effects, the maximum number of edits that can be introduce to their precondition set is $2|F_v(\xi)|$.
+
+Likewise we can define the edit distance with regard to a sequence of observations $\mathcal{O}$>
+\begin{definition}
+
+\end{definition}
+
+\subsection{Recognition and evaluation of \strips\ models}
+The probability distribution of the possible \strips\ models given a sequence of state observations allows to evaluate the quality of a learned \strips\ action models with regard to a sequence of observations that act as a test set.
+
+\subsection{Recognition with classical planning}
+
 \section{Evaluation}
@@
 \subsubsection{Reproducibility}
-For the evaluation we used IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only use 5 labels for each domain (25 observed states per domain when using state observations, 5 intermediate states per label) and they are fixed for all the experiments so we can evaluate the impact of the available input knowledge in the quality of the learned models.
+We used IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+For the learning of the \strips\ action models we used observations sequences of 25 states per domain.
 All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
@@
 
-\subsection{Learning from labels}
-Tables~\ref{fig:labelsnomap} and~\ref{fig:labels} show the precision and recall values obtained when learning from a set of $(initial, final)$ state pairs.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
-The learning examples and learned models are the same for both tables, the only difference is the evaluation procedure for computing the {\em precision} and {\em recall} values.
-Table~\ref{fig:labelsnomap} does not use the $f_{P\&R}$ mapping which implies assuming that the learned actions are never reformulated.
-Results show this is a too strong assumption.
-
-The Table~\ref{fig:planlabels} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-
-\subsubsection{Learning from labels and state invariants}
-For each domain we provide also a set of {\em domain invariants} that are computed using the {\sc TIM} algorithm~\cite{fox:TIM:JAIR1998}.
-Table~\ref{fig:invariants} shows the precision and recall values obtained when learning from state invariants and Table~\ref{fig:planinvariants} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-
-\begin{table}[hbt!]
-	\begin{scriptsize}
-		\begin{center}
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-				\hline
-				blocks & 0.63 & 0.56 & 0.71 & 0.56 & 0.71 & 0.56 & 0.68 & 0.56 \\
-				Driverlog & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				ferry & 0.75 & 0.43 & 0.75 & 0.75 & 0.75 & 0.75 & 0.75 & 0.64 \\
-				Floortile & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
-				gripper-strips & 0.8 & 0.67 & 0.75 & 0.75 & 1.0 & 1.0 & 0.85 & 0.81 \\
-				Hanoi & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Npuzzle & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Parking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Satellite & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Transport & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Zenotravel & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				\hline
-				\bf  & - & - & - & - & - & - & - & - \\
-			\end{tabular}
-		\end{center}
-	\end{scriptsize}
-	\caption{\small Precision and recall values obtained when learning from labels + invariants.}
-	\label{fig:invariants}
-\end{table}
-
-\begin{table}
-	\begin{footnotesize}
-		\begin{center}
-			\begin{tabular}{l|c|c|c|}			
-				& Total time & Preprocess & Plan length  \\
-				\hline
-				Blocks & 652.70 & 0.04 & 76  \\
-				Driverlog & 14.98 & 0.10 & 65 \\
-				Ferry & 1.70 & 0.03 & 58 \\
-				Floortile & - & - & - \\ % Ctrl C
-				Grid & - & - & - \\ % Time out 1000s
-				Gripper & 0.14 & 0.00 & 47 \\
-				Hanoi & 55.30 & 0.14 & 43  \\
-				Hiking & - & - & - \\ % Ctrl C
-				Parking & - & - & - \\ % Time out 1000s
-				Satellite & 84.57 & 0.22 & 98 \\
-				Sokoban & - & - & - \\ % Time out 1000s
-				Transport & - & - & - \\ % Ctrl C
-				Zenotravel & - & - & - \\ % Ctrl C
-			\end{tabular}
-		\end{center}
-	\end{footnotesize}	
- \caption{\small Planning results obtained when learning from labels + invariants.}
-\label{fig:planinvariants}
-\end{table}
-
-\subsection{Learning from state observations}
-This is the setting with the maximum amount of available input knowledge.
-Table~\ref{fig:observations} shows the precision and recall values obtained when learning from state observations, using the same initial and  final states and the same state invariants as in the previous evaluation.
-
-\begin{table}
-	\begin{footnotesize}
-		\begin{center}
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-				\hline
-				blocks & 0.78 & 0.78 & 0.78 & 0.78 & 0.78 & 0.78 & 0.78 & 0.78 \\
-				driverlog & 0.75 & 0.64 & 0.86 & 0.86 & 0.86 & 0.86 & 0.82 & 0.79 \\
-				ferry & 0.86 & 0.86 & 0.75 & 0.75 & 0.75 & 0.75 & 0.79 & 0.79 \\
-				Floortile & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
-				gripper-strips & 1.0 & 0.67 & 0.8 & 1.0 & 1.0 & 1.0 & 0.93 & 0.89 \\
-				hanoi & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
-				Hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Npuzzle & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				parking & 0.8 & 0.57 & 1.0 & 0.78 & 1.0 & 0.78 & 0.93 & 0.71 \\
-				Satellite & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				Transport & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				zeno-travel & 1.0 & 0.43 & 1.0 & 0.71 & 1.0 & 0.71 & 1.0 & 0.62 \\
-				\hline
-				\bf  & - & - & - & - & - & - & - & - \\
-			\end{tabular}
-		\end{center}
-	\end{footnotesize}
-	\caption{\small Precision and recall values obtained when learning from observations.}
-	\label{fig:observations}       
-\end{table}
-
-\begin{table}
-	\begin{footnotesize}
-		\begin{center}
-			\begin{tabular}{l|c|c|c|}			
-				& Total time & Preprocess & Plan length  \\
-				\hline
-				Blocks & 13.36 & 0.00 & 73  \\
-				Ferry & 89.04 & 0.03 & 63 \\
-				Gripper & 0.66 & 0.00 & 43 \\
-				Hanoi & 98.66 & 0.11 & 45  \\
-			\end{tabular}
-		\end{center}
-	\end{footnotesize}
- \caption{\small Planning results obtained when learning from observations.}
-\label{fig:planobservations}	
-\end{table}
-
 \section{Conclusions}

### 2018-01-18 src/validator.py
validator.py: removed hardcoded values

@@ -105,7 +105,7 @@ try:
 
 except:
     print "Usage:"
-    print sys.argv[0] + "<domain> <test folder> <problems prefix>  <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
+    print sys.argv[0] + "<domain> <test folder> <problems prefix> <plans prefix> <input level (0 plans, 1 steps, 2 len(plan), 3 minimum)>"
     sys.exit(-1)
 
 
@@ -489,7 +489,7 @@ for line in file:
 
 file.close()
 
-num_modifications = 5
+num_modifications = len(modifications)
 worst_case = len(allpres)*3
 mean = worst_case/2
 # p = 1/(1+math.exp(-float(mean-num_modifications)/float(mean/6)))

### 2018-01-19 doc/ijcai-18/observations-learning.tex
updates to the IJCAI18 paper after evaluation as edition distance proposal

@@
 This paper presents a classical planning compilation for learning \strips\ action models from state observations.
-The compilation approach does not require observing the precise actions that produced the state observations because such actions are determined by an off-the-shelf classical planner.
-In addition, the compilation is extensible to estimate the probability distribution of the possible \strips\ models given a sequence of state observations.
-This extension allows us to address model recognition tasks as well as to semantically evaluate the quality of learned \strips\ models.
+The compilation approach does not require observing the precise actions that produced the state observations because such actions are determined by a classical planner.
+In addition, the compilation is extensible to (1) semantically evaluate the quality of learned \strips\ models and (2), estimate the probability distribution of the possible \strips\ models given a sequence of state observations.
 \end{abstract}
@@
 \begin{enumerate}
-\item Is defined as a planning compilation. This fact opens the door to the {\em bootstrapping} of planning action models and allows us to report results over a wide range of planning domains. 
+\item Is defined as a classical planning compilation.
+This fact opens the door to the {\em bootstrapping} of planning action models and allows us to report results over a wide range of planning domains.
 \item Does not require observing the particular executed actions. An off-the-shelf classical planner determines these actions given the state observations.
-\item Can estimate the probability distribution of the possible \strips\ models, given an observation sequence.
-This extension allow us to address model recognition tasks as well as to evaluate the quality of the learned models semantically, without comparing with a reference model.
+\item Is extensible to semantically evaluate the quality of learned \strips\ models as well as to estimate the probability distribution of the possible \strips\ models given a sequence of state observations.
 \end{enumerate}
@@
 \section{Background}
-This section defines the planning models used in this work as well as the input (states and state invariants) and the output (an \strips\ action model) of the addressed learning task.
+This section defines the planning models used in this work as well as the output (an \strips\ action model) of the addressed learning task.
 
@@
 
-An action $a\in A$ is now defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
+An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
 Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
@@
 
-Finally we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in a given action schema $\xi$.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} only contains the fluents from $F_v$ that do not involve $v_2$.
+Finally we also define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in a given action schema $\xi$.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} only contains the fluents from $F_v$ that do not involve $v_2$ because the action header contained the single parameter $v_1$.
 
 \begin{figure}
-\begin{scriptsize}
+\begin{footnotesize}
 \begin{verbatim}
@@
  :precondition (and (holding ?v1) (clear ?v2))
-  :effect (and (not (holding ?v1)) (not (clear ?v2))
-               (handempty) (clear ?v1) (on ?v1 ?v2)))
+  :effect (and (not (holding ?v1)) 
+               (not (clear ?v2))
+               (handempty) (clear ?v1) 
+               (on ?v1 ?v2)))
 \end{verbatim}
-\end{scriptsize}
+\end{footnotesize}
  \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
@@
 
-We formalize a more challenging learning task, where less input knowledge is available.
+In this paper we address a more challenging learning task, where less input knowledge is available.
 The addressed learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
-This learning task is defined as $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$:
+This learning task is defined as $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$, where:
 \begin{itemize}
@@
 \item $\Xi$ is the set of empty operator schemes containing only the headers with the name and parameter of each operator schema $\xi\in\Xi$.
-\item $\mathcal{O}=\{s_0,s_1,\ldots,s_{n}\}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$.
+\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$.
 \end{itemize}
 
-A solution to $\Lambda$ is a set of operator schema $\Xi'$ compliant with the predicates in $\Psi$, the headers in $\Xi$ and the sequence of state observations $\mathcal{O}$.
-A planning compilation is a suitable approach for addressing a $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that can explain $\mathcal{O}$.
+A solution to $\Lambda$ is a set of operator schema $\Xi'$ compliant with the predicates in $\Psi$, the headers in $\Xi$ and the given sequence of state observations $\mathcal{O}$.
+A planning compilation is a suitable approach for addressing a $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains $\mathcal{O}$.
 Figure~\ref{fig:lexample} shows an example of a $\Lambda$ task for learning a \strips\ action model in the blocksworld from the sequence of five state observations that corresponds to inverting a 2-blocks tower.
 
-\begin{figure}
+\begin{figure}[hbt!]
 {\footnotesize\tt ;;; Predicates in $\Psi$}
-\begin{scriptsize}
+\begin{footnotesize}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{scriptsize}
+\end{footnotesize}
 \vspace{0.2cm}
 {\footnotesize\tt ;;; Headers in $\Xi$}
-\begin{scriptsize}
+\begin{footnotesize}
 \begin{verbatim}
-(pickup v1) (putdown v1) (stack v1 v2} (unstack v1 v2)
+(pickup v1) (putdown v1) 
+(stack v1 v2} (unstack v1 v2)
 \end{verbatim}
-\end{scriptsize}
+\end{footnotesize}
 \vspace{0.2cm}
 {\footnotesize\tt ;;; Observations in $\mathcal{O}$}
-\begin{scriptsize}
+\begin{footnotesize}
 \begin{verbatim}
 ;;; observation #0
-(clear block2) (on block2 block1) (ontable block1) 
+(clear block2) (on block2 block1) 
+(ontable block1) (handempty)
+
+;;; observation #1
+(holding block2) (clear block1) 
+(ontable block1)
+
+;;; observation #2
+(clear block1) (ontable block1) 
+(clear block2) (ontable block2) 
 (handempty)
 
-;;; observation #1
-(holding block2) (clear block1) (ontable block1)
-
-;;; observation #2
-(clear block1) (ontable block1) (clear block2) 
-(ontable block2) (handempty)
-
 ;;; observation #3
-(holding block1) (clear block2) (ontable block2)
+(holding block1) (clear block2) 
+(ontable block2)
 
 ;;; observation #4
-(clear block1) (on block1 block2) (ontable block2) 
-(handempty) 
+(clear block1) (on block1 block2) 
+(ontable block2) (handempty) 
 \end{verbatim}
-\end{scriptsize}
- \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a sequence of five state observations.}
+\end{footnotesize}
+ \caption{\small Example of a $\Lambda$ task for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
 \label{fig:lexample}
@@
 \item Programs the \strips\ action model $\Xi'$.
-A solution plan starts with a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+A solution plan starts with a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
 \item Validates the \strips\ action model $\Xi'$ in $\mathcal{O}$.
-The solution plan continues with a postfix that produces the sequence of states in $\mathcal{O}$ starting from $s_0$, and using the programmed action model $\Xi'$.
+The solution plan continues with a postfix that produces the given sequence of states $\tup{s_0,s_1,\ldots,s_{n}}$ using the programmed action model $\Xi'$.
 \end{enumerate}
@@
 \begin{itemize}
-\item Fluents $F$ built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$. 
+\item The set of fluents $F$ that is built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$. 
 \item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$.
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$.
-For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
+For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
 \item Fluent $mode_{prog}$ indicating whether the operator schemes are programmed or validated (already programmed) and fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
 \end{itemize}
-\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0$ (the first observation) and every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
-Our compilation assumes that initially operator schemes are programmed with every possible precondition, no negative effect and no positive effect.
+\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
+In addition, our compilation assumes that initially operator schemes are programmed with every possible precondition, no negative effect and no positive effect.
+With this regard, the fluents $pre_f(\xi)$ hold at the initial state for every $f\in F_v(\xi)$.
 \item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, indicates that the programmed action model is validated in all the input observations.
@@
 \item Actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
-We assume operators headers are known so the binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
+We assume that the operators headers are known so the binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
 variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
@@
 \hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\\
-&\cup\{\neg test_j\}_{j\in t\leq j\leq n}\cup \{\neg mode_{prog}\},\\
+&\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{\neg mode_{prog}\},\\
 \cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i\}.
@@
 
-\subsection{Compilation properties}
-
-\begin{lemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
-\end{lemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects (furthermore, effects can only be programmed after preconditions are programmed).
-Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, goals $\{test_i\}$, {\small $1\leq i\leq n$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the state $s_i$, starting from $s_0$.
-If this is achieved for every {\small $1\leq i\leq n$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, solves $\Lambda$.
-\end{small}
-\end{proof}
-
-\begin{lemma}
-Completeness. Any \strips\ action model $\Xi$ computable from $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ can be obtained by solving the corresponding classical planning task $P_{\Lambda}$.
-\end{lemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the set of elements that can appear in a \strips\ action schema $\xi\in\Xi$.
-Any possible \strips\ action schema built with the fluents in $F_v$ can be computed because the compilation only constrains the application of $\mathsf{apply_{\xi,\omega}}$ actions iff they are not compliant with the $\mathcal{O}$ sequence.
-\end{small}
-\end{proof}
-
-\begin{figure}
+\begin{figure}[hbt!]
 \begin{scriptsize}
@@
 
+\subsection{Compilation properties}
+
+\begin{lemma}
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi'$ that solves the learning task $\Lambda$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+The compilation forces that once the preconditions of an operator schema $\xi \in \Xi'$ are programmed, they cannot be altered.
+The same happens with the positive and negative effects.
+Furthermore because of the preconditions of the $\mathsf{programPre_{f,\xi}}$ actions, effects are only programmable after preconditions are programmed.
+Once the operator schemes $\Xi'$ are programmed, they can only be applied because of the $mode_{prog}$ fluent.
+To solve $P_{\Lambda}$, the goals $\{test_i\}$, {\small $1\leq i\leq n$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following  the sequence {\small $1\leq i\leq n$}.
+Therefore if $test_n$ is achieved, it means that the programmed action model $\Xi'$ is compliant with the provided input knowledge and hence, solves $\Lambda$.
+\end{small}
+\end{proof}
+
+\begin{lemma}
+Completeness. Any \strips\ action model $\Xi'$ that solves a $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the set of elements that can appear in a \strips\ action schema $\xi\in\Xi$.
+In addition, any possible \strips\ action schema $\Xi'$ that can be built with the fluents in $F_v$ can be computed  with the $P_{\Lambda}$ compilation.
+The only \strips\ action models that the compilation cannot compute are the ones that violate the state constraints defined by the $\mathcal{O}$ sequence, which constrain as well the solutions to the $\Lambda$ learning task.
+\end{small}
+\end{proof}
+
 \section{Recognizing \strips\ action models}
-Inspired by the {\em Plan recognition as planning} approach~\cite{ramirez2009plan} we show that the previous compilation can be extended to estimate the probability distribution of the possible \strips\ models given a sequence of state observations.
-This extension allows us to address model recognition tasks as well as to semantically evaluate the quality of learned \strips\ models
+Inspired by {\em plan recognition as planning}~\cite{ramirez2009plan}, we show that the previous compilation can be extended to (1) semantically evaluate the quality of learned \strips\ models, without comparing with a reference model and (2), address \strips\ action model recognition tasks.
 
 \subsection{The \strips\ edit distance}
-We assume that how well a given \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$, depends on the amount of edits that one has to introduce to $\Xi$ to produce the sequence of obsevations $\mathcal{O}$.
-If the given model $\Xi$ perfectly explains the observations $\mathcal{O}$, no edit have to be introduced and the probability of explanning the observations is $P(\Xi|\mathcal{O})=1$.
-On the other hand, with a probability of explanning the observations of $P(\Xi|\mathcal{O})=0$, is the model that requires the maximum number of edits to explain the sequence of observations $\mathcal{O}$.
+We assume that how well a given \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$, depends on the amount of {\em edition} that one has to introduce to $\Xi$ to produce the sequence of obsevations $\mathcal{O}$.
+In the extreme, if the given model $\Xi$ perfectly explains the observations $\mathcal{O}$, no {\em edit} to the model must be introduced.
 
-To compute the $P(\Xi|\mathcal{O})$ probability distribution, we need a metric that quantifies this amount of edits.
-To define such metric, we need to define first which are the possible edit operations that can be performed on a \strips\ action model.
-With the aim of keeping tractable the branching factor of the classical planning task resulting from our compilation, we only define two \strips\ edit operations:
+Let us define the operations allowed to edit a given \strips\ action model.
+We call these operations the {\em \strips\ edit operations}.
+With the aim of keeping tractable the branching factor of the classical planning task that results from our compilation, we only define two {\em \strips\ edit operations}:
 \begin{itemize}
-\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi$
-\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi$
+\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, such that $f\in F_v(\xi)$, is removed from the operator schema $\xi\in\Xi$.
+\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, such that $f\in F_v(\xi)$, is added to the operator schema $\xi\in\Xi$.
 \end{itemize}
 
-Now we are ready to define a metric that asseses then how dissimilar two given \strips\ action models are
+With this defined, we can now formalize an edit distance metric that assesses how dissimilar two given \strips\ action models are.
+This distance satisfies the {\em metric axioms} provided that the two {\em \strips\ edit operations} have the same positive cost.
 \begin{definition}
-The {\bf \strips\ edit distance}, denoted as $\delta(\Xi,\Xi')$, is the minimum number of edit operations that allow to transform a given \strips\ action model $\Xi$ into the given \strips\ action model $\Xi'$.
-This edit distance satisfies the {\em metric axioms} provided that the two operations has the same positive cost.
+Given two \strips\ action model $\Xi$ and $\Xi'$, such that both are built from the same set of possible elements $F_v$.
+The {\em \bf \strips\ edit distance}, denoted as $\delta(\Xi,\Xi')$, is the minimum number of {\em \strips\ edit operations} to transform $\Xi$ into $\Xi'$.
 \end{definition}
 
-Since the size of the $F_v$ set is bound, we can define the maximum number of edits that can be introduced to a given \strips\ action model.
+Since the size of the $F_v$ set is bound, the maximum number of edits that can be introduced to a given \strips\ action model is bound as well.
+In more detail, for a given operator schema $\xi\in\Xi$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$.
+With regard to effects of $\xi$, the maximum number of edits that can be introduced are two times $|F_v(\xi)|$.
+
 \begin{definition}
-The maximum {\bf \strips\ edit distance} is an upperbound, denoted as $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$, that indicates the maximum distance required to transform a given \strips\ action model $\Xi$ into any \strips\ action model $\Xi'$.
+Given a \strips\ action model $\Xi$ built from the set of possible elements $F_v$.
+The {\em \bf maximum  \strips\ edit distance} is the upperbound $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$, that indicates the maximum value of the {\em \strips\ edit distance} from $\Xi$ to any \strips\ action model definable within $F_v$.
 \end{definition}
 
-In more detail, for a given operator schema $\xi$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$.
-With regard to action effects, the maximum number of edits that can be introduce to their precondition set is $2|F_v(\xi)|$.
-
-Likewise we can define the edit distance with regard to a sequence of observations $\mathcal{O}$>
+Likewise the {\em \strips\ edit distance} can be defined with regard to a sequence of observations $\mathcal{O}$.
 \begin{definition}
-
+ Given a \strips\ action model $\Xi$ built from the set of possible elements $F_v$ and a sequence of observations $\mathcal{O}$.
+The {\em \bf observations  \strips\ edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimum number of {\em \strips\ edit operations} to transform $\Xi$ into an action model $\Xi'$ defined within $F_v$ that can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ such that $\pi$ induces the state observations $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$.
 \end{definition}
 
-\subsection{Recognition and evaluation of \strips\ models}
-The probability distribution of the possible \strips\ models given a sequence of state observations allows to evaluate the quality of a learned \strips\ action models with regard to a sequence of observations that act as a test set.
+The {\em observations  \strips\ edit distance} allow us to semantically evaluate the quality of a learned strips action model with respect to a sequence of observations that acts as a test set.
+The lower this distance is, the better the quality of the learned model.
+The semantic nature of this evaluation is robust to learning episodes where actions are reformulated and still compliant with the inputs (e.g.
+the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa or the roles of actions parameters with the same type could be interchanged).
 
-\subsection{Recognition with classical planning}
+\subsection{Recognition of \strips\ models}
+Given a set of possible \strips\ models and a sequence of state observations the {\em recognition of \strips\ models} is the task of computing the probability distribution of these models according to the given observations.
+Here we show that the {\em observations  \strips\ edit distance} is helpful to compute the probability distribution of the possible \strips\ models given a sequence of state observations.
+The main idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map these distances into likelihoods using the Bayes rule.
+
+According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
+In our scenario, the hypotheses are about the possible \strips\ action models that can be build given a set of operator headers, or in other words, given the $F_v(\xi)$ sets.
+Moreover, we assume that all the possible \strips\ action models are equiprobable so according to the Bayes rule $P(\mathcal{H}|\mathcal{O})$ is proportional to $P(\mathcal{O}|\mathcal{H})$.
+Finally, we assume that $P(\mathcal{O}|\mathcal{H})$ is given by the {\em observations  \strips\ edit distance}, mapping distances into probabilities (the larger the distance the lower the likelihood), according to the following sigmoid function $P(\mathcal{O}|\Xi)=\frac{1}{1+e^{-\delta(\Xi,\mathcal{O})}}$.
+
+With this regard, the probability distribution of the possible \strips\ models (within the set of operator headers, that is within the $F_v(\xi)$ sets) given a sequence of state observations,  $P(\Xi|\mathcal{O})$, can be computed by:
+\begin{enumerate}
+\item The {\em observations  \strips\ edit distance} is computed for every possible model.
+\item The resulting distances are plugged into the previous$P(\Xi|\mathcal{O})$ equation to map distances into likelihoods.
+\item Likelihoods are plugged into the Bayes’ rule $P(\Xi|\mathcal{O})$ to obtain the normalized posterior probabilities.
+Since we are computing a probability distribution these probabilities must sum up $1.0$.
+\end{enumerate}
+
+\subsection{\strips\ model evaluation and recognition with classical planning}
+The previous compilation can be extended to estimate the {\em observations  \strips\ edit distance}.
+In this case the tuple $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ does not represent a learning task but the edition of a given \strips\ action model $\Xi$ to cover the sequence of observations $\mathcal{O}$.
+The extended compilation outputs a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$ where:
+\begin{itemize}
+\item $F_{\Lambda}$ is defined as in the previous compilation.
+\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
+In addition, the given \strips\ action model $\Xi$ is now encoded in the initial state.
+This means that fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ with $f\in F_v(\xi)$ hold in the initial state if they appear in the given \strips\ action model $\Xi$.
+\item $G_{\Lambda}$ is defined as in the previous compilation.
+\item $A_{\Lambda}'$, comprises the three kinds of actions of $A_{\Lambda}$.
+The Actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation {\tt\small $1\leq i\leq n$} are defined exactly as in the previous compilation.
+The only difference are the actions for {\em programming} operator schema $\xi\in\Xi$ that must include now actions for adding a precondition and for removing a {\em negative} or {\em positive} effect.
+\end{itemize}
+
+The probability distribution $P(\Xi|\mathcal{O})$ would be exact if the classical planning problems resulting from our compilation are solved optimally (according to the number of edit actions), and will be approximate if they are solved with a satisfying classical planner or furthermore if what is solved is not $P_{\Lambda}'$ but a relaxation of this task such as the {\em delete relaxation}.
 
@@
 With this mapping defined we can compute the {\em precision} and {\em recall} of a learned \strips\ action $\xi\in\Xi$ with respect to the action $f_{P\&R}(\xi)\in \Xi^*$ even if actions are reformulated in the learning process.
+
+\section{Conclusions}
+As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains.
+Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
+
+%The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
+In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
+
+The size of the compiled classical planning instances depends on the number of input examples.
+Generating {\em informative} examples for learning planning action models is an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
+The combination of such exploration techniques with our learning approach is an intriguing research direction that opens the door to the bootstrapping of planning action models.
+
+%Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
+For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true {\em at least at one} of the reached states.
+Despite this is beyond the scope of this paper, LTL constraints could be included in our compilation following the ideas for compiling temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) force the solution plans to always leave the LTL automata at an acceptor state by adding new goals to the classical planning task.
+
+% Commented for blind submission
+%\begin{small}
+%\subsection*{Acknowledgment}
+%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
+Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
+%\end{small}
 
@@
 
-\section{Conclusions}
-As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains.
-%Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
-
-The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
-In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
-
-%The size of the compiled classical planning instances depends on the number of input examples.
-Generating {\em informative} examples for learning planning action models is an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
-The combination of such exploration techniques with our learning approach is an intriguing research direction that opens the door to the bootstrapping of planning action models.
-
-%Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
-For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true {\em at least at one} of the reached states.
-Despite this is beyond the scope of this paper, LTL constraints could be included in our compilation following the ideas for compiling temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) force the solution plans to always leave the LTL automata at an acceptor state by adding new goals to the classical planning task.
-
-% Commented for blind submission
-%\begin{small}
-%\subsection*{Acknowledgment}
-%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
-Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
-%\end{small}
-
 %% The file named.bst is a bibliography style file for BibTeX 0.99c

### 2018-01-20 doc/ijcai-18/observations-learning.tex
update to the likelihood function ijcai18 paper

@@
 Moreover, we assume that all the possible \strips\ action models are equiprobable so according to the Bayes rule $P(\mathcal{H}|\mathcal{O})$ is proportional to $P(\mathcal{O}|\mathcal{H})$.
-Finally, we assume that $P(\mathcal{O}|\mathcal{H})$ is given by the {\em observations  \strips\ edit distance}, mapping distances into probabilities (the larger the distance the lower the likelihood), according to the following sigmoid function $P(\mathcal{O}|\Xi)=\frac{1}{1+e^{-\delta(\Xi,\mathcal{O})}}$.
+Finally, we assume that $P(\mathcal{O}|\mathcal{H})$ is given by the {\em observations  \strips\ edit distance}, mapping distances into probabilities (the larger the distance the lower the likelihood), according to the following expression $P(\mathcal{O}|\Xi)=1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$.
 

### 2018-01-22 experiments/ijcai18/generate_table_1_2.sh
Updated experiments/ijcai18 scripts

@@ -1,16 +1,16 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "hiking" "npuzzle" "parking" "satellite" "sokoban" "transport" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "hiking" "miconic" "npuzzle" "parking" "pegsol" "satellite" "sokoban" "transport" "visitall" "zenotravel")
 
 ulimit -t 1000
 
 for i in "${arr[@]}"
 do
-   ../../src/compiler.py -s ../../benchmarks/ijcai18/labels/$i/ empty_domain test plan 3
+   ../../src/compiler.py -s ../../benchmarks/ijcai18/observations/$i/ empty_domain test plan 2
 
    echo $i "w/o reformulation"
-   ../../src/evaluator2.py ../../benchmarks/ijcai18/labels/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/ijcai18/labels/$i/test-01.pddl
+   ../../src/evaluator2.py ../../benchmarks/ijcai18/observations/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/ijcai18/observations/$i/test-01.pddl
    echo $i "with reformulation"
-   ../../src/evaluator2.py -r ../../benchmarks/ijcai18/labels/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/ijcai18/labels/$i/test-01.pddl
+   ../../src/evaluator3.py -r ../../benchmarks/ijcai18/observations/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/ijcai18/observations/$i/test-01.pddl
    cat planner_out.log | grep -e 'actions\sin'
    cat planner_out.log | grep -e 'total\stime'
 done

### 2018-01-22 experiments/ijcai18/generate_table_3.sh

@@ -1,14 +1,14 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "hiking" "npuzzle" "parking" "satellite" "sokoban" "transport" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "hiking" "miconic" "npuzzle" "parking" "satellite" "transport" "visitall" "zenotravel")
 
 ulimit -t 1000
 
 for i in "${arr[@]}"
 do
-   ../../src/compiler.py -s ../../benchmarks/ijcai18/labels/$i/ empty_invariants_domain test plan 3
+   echo $i
 
-   echo $i 
-   ../../src/evaluator2.py -r ../../benchmarks/ijcai18/labels/$i/full_domain.pddl learned_domain.pddl ../../benchmarks/ijcai18/labels/$i/test-01.pddl
-   cat planner_out.log | grep -e 'actions\sin'
-   cat planner_out.log | grep -e 'total\stime'
+   ../../src/compiler.py -s ../../benchmarks/ijcai18/observations/$i/ empty_domain test plan 2
+
+   ../../src/validator.py learned_domain.pddl ../../benchmarks/ijcai18/test/$i/ test plan 2
+
 done

### 2018-01-22 doc/ijcai-18/observations-learning.tex
Add files via upload

--- 2018-01-19 doc/ijcai-18/observations-learning.tex
@@
 \newcommand{\tup}[1]{{\langle #1 \rangle}}
+\newcommand{\argmax}{\mathop{\mbox{argmax}}}
+\newcommand{\argmin}{\mathop{\mbox{argmin}}}
 
@@
 
-\title{Learning and Recognition of \strips\ Action Models from State Observations}
+\title{Learning and Recognition of \strips\ Action Models from State Observations \\ Action Model Recognition from State Observations via Planning}
 
-\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
-{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-{\small Universitat Polit\`ecnica de Val\`encia.}\\
-{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
-{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
+\author{
+%Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
+%{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
+%{\small Universitat Polit\`ecnica de Val\`encia.}\\
+%{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
+%{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}
+}
 
@@
 The compilation approach does not require observing the precise actions that produced the state observations because such actions are determined by a classical planner.
-In addition, the compilation is extensible to (1) semantically evaluate the quality of learned \strips\ models and (2), estimate the probability distribution of the possible \strips\ models given a sequence of state observations.
+In addition, the compilation is extensible to (1) semantically evaluate the quality of learned action models and (2) estimate the probability distribution of the possible action models given a state observation sequence.
+
+\textbf{EVA}
+
+This paper presents a first novel approach to \strips\ action model recognition from state observations.
+Unlike other related works, our classical planning compilation approach not only allows learning a \strips\ action model from an observation sequence but it is also extendible to action model recognition.
+The learning task does not require observing the precise actions that produce the state observations because such actions are determined by a classical planner.
+We present a two-fold evaluation of the learned models: a syntactic and semantic assessment against a reference model and via a test dataset that calculates the probability of a model given the observation.
+Moreover, we show how these probabilistic values are used to estimate the probability of the action model recognition.
+
 \end{abstract}
@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper presents a novel approach for learning \strips\ action models that introduces the following contributions:
+
 \begin{enumerate}
@@
 \item $\Psi$ is the set of predicates that define the abstract state space of a given classical planning frame.
-\item $\Xi$ is the set of empty operator schemes containing only the headers with the name and parameter of each operator schema $\xi\in\Xi$.
+\item $\Xi$ is the set of empty operator schemes, wherein each $\xi\in\Xi$ is only composed of $head(\xi)$.
 \item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$.
@@
 
-Our approach for addressing the learning task, is compiling it into a classical planning task with conditional effects.
+Our approach for addressing the learning task is compiling $\Lambda$ into a classical planning task with conditional effects.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
+
 \begin{enumerate}
@@
 
-\section{Recognizing \strips\ action models}
+\section{\strips\ action models recognition}
 Inspired by {\em plan recognition as planning}~\cite{ramirez2009plan}, we show that the previous compilation can be extended to (1) semantically evaluate the quality of learned \strips\ models, without comparing with a reference model and (2), address \strips\ action model recognition tasks.
 
-\subsection{The \strips\ edit distance}
-We assume that how well a given \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$, depends on the amount of {\em edition} that one has to introduce to $\Xi$ to produce the sequence of obsevations $\mathcal{O}$.
+\subsection{The edit distance}
+We assume that how well a given \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$, depends on the amount of {\em edition} that one has to introduce to $\Xi$ to produce the sequence of observations $\mathcal{O}$.
 In the extreme, if the given model $\Xi$ perfectly explains the observations $\mathcal{O}$, no {\em edit} to the model must be introduced.
 
-Let us define the operations allowed to edit a given \strips\ action model.
-We call these operations the {\em \strips\ edit operations}.
+Let us define the \emph{edit operations} allowed to edit a given \strips\ action model.
 With the aim of keeping tractable the branching factor of the classical planning task that results from our compilation, we only define two {\em \strips\ edit operations}:
@@
 This distance satisfies the {\em metric axioms} provided that the two {\em \strips\ edit operations} have the same positive cost.
+
 \begin{definition}
-Given two \strips\ action model $\Xi$ and $\Xi'$, such that both are built from the same set of possible elements $F_v$.
-The {\em \bf \strips\ edit distance}, denoted as $\delta(\Xi,\Xi')$, is the minimum number of {\em \strips\ edit operations} to transform $\Xi$ into $\Xi'$.
+Let $\Xi$ and $\Xi'$ be two action models, both built from the same set of possible elements $F_v$.
+The {\bf edit distance}, denoted as $\delta(\Xi,\Xi')$, is the minimum number of {\em edit operations} to transform $\Xi$ into $\Xi'$.
 \end{definition}
 
-Since the size of the $F_v$ set is bound, the maximum number of edits that can be introduced to a given \strips\ action model is bound as well.
+Since the size of the $F_v$ set is bound, the maximum number of edits that can be introduced to a given action model is bound as well.
 In more detail, for a given operator schema $\xi\in\Xi$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$.
@@
 \begin{definition}
-Given a \strips\ action model $\Xi$ built from the set of possible elements $F_v$.
-The {\em \bf maximum  \strips\ edit distance} is the upperbound $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$, that indicates the maximum value of the {\em \strips\ edit distance} from $\Xi$ to any \strips\ action model definable within $F_v$.
+Let $\Xi$ be an action model built from the set of possible elements $F_v$.
+The \textbf{maximum edit distance}, defined as $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$, is an upper bound on the distance from $\Xi$ to any \strips\ action model definable within $F_v$.
 \end{definition}
 
-Likewise the {\em \strips\ edit distance} can be defined with regard to a sequence of observations $\mathcal{O}$.
+Likewise, we can define a distance of a model $\Xi$ to an observation sequence $\mathcal{O}$ on the basis of the \emph{edit distance} between two models.
+
 \begin{definition}
- Given a \strips\ action model $\Xi$ built from the set of possible elements $F_v$ and a sequence of observations $\mathcal{O}$.
-The {\em \bf observations  \strips\ edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimum number of {\em \strips\ edit operations} to transform $\Xi$ into an action model $\Xi'$ defined within $F_v$ that can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ such that $\pi$ induces the state observations $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$.
+Given an action model $\Xi$ built from the set of possible elements $F_v$ and a sequence of observations $\mathcal{O}$.
+The {\bf observation edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimal edit distance to transform $\Xi$ into a model $\Xi'$ which produces a plan $\pi=\tup{a_1, \ldots, a_n}$ such that $\pi$ induces the observation sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$; i.e., \[\delta(\Xi,\mathcal{O})=\argmin_{\Xi' \rightarrow \mathcal{O}} \delta(\Xi,\Xi')\]
 \end{definition}
 
-The {\em observations  \strips\ edit distance} allow us to semantically evaluate the quality of a learned strips action model with respect to a sequence of observations that acts as a test set.
+The {\em observation edit distance} allow us to semantically evaluate the quality of a learned strips action model with respect to a sequence of observations that acts as a test set.
 The lower this distance is, the better the quality of the learned model.
@@
 
-\subsection{Recognition of \strips\ models}
+\subsection{Model recognition}
 Given a set of possible \strips\ models and a sequence of state observations the {\em recognition of \strips\ models} is the task of computing the probability distribution of these models according to the given observations.
@@
 According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the possible \strips\ action models that can be build given a set of operator headers, or in other words, given the $F_v(\xi)$ sets.
+In our scenario, the hypotheses are about the possible \strips\ action models that can be built given a set of operator headers, or in other words, given the $F_v(\xi)$ sets.
 Moreover, we assume that all the possible \strips\ action models are equiprobable so according to the Bayes rule $P(\mathcal{H}|\mathcal{O})$ is proportional to $P(\mathcal{O}|\mathcal{H})$.
-Finally, we assume that $P(\mathcal{O}|\mathcal{H})$ is given by the {\em observations  \strips\ edit distance}, mapping distances into probabilities (the larger the distance the lower the likelihood), according to the following sigmoid function $P(\mathcal{O}|\Xi)=\frac{1}{1+e^{-\delta(\Xi,\mathcal{O})}}$.
+Finally, we assume that $P(\mathcal{O}|\mathcal{H})$ is given by the {\em observation edit distance}, mapping distances into probabilities (the larger the distance the lower the likelihood), according to the following expression $P(\mathcal{O}|\Xi)=1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$.
 
-With this regard, the probability distribution of the possible \strips\ models (within the set of operator headers, that is within the $F_v(\xi)$ sets) given a sequence of state observations,  $P(\Xi|\mathcal{O})$, can be computed by:
+Thus, $P(\Xi|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the set of operator headers, that is within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$, is computed by:
+
 \begin{enumerate}
-\item The {\em observations  \strips\ edit distance} is computed for every possible model.
-\item The resulting distances are plugged into the previous$P(\Xi|\mathcal{O})$ equation to map distances into likelihoods.
-\item Likelihoods are plugged into the Bayes’ rule $P(\Xi|\mathcal{O})$ to obtain the normalized posterior probabilities.
-Since we are computing a probability distribution these probabilities must sum up $1.0$.
+\item Calculating the {\em observation edit distance} $\delta(\Xi,\mathcal{O})$ for every possible model $\Xi$.
+\item Applying the resulting distances to the above formula $P(\mathcal{O}|\Xi)$ to map distances into likelihoods
+\item Applying Bayes rule to obtain the normalized posterior probabilities $P(\Xi|\mathcal{O})$. The sum of the probabilities of all the models will be 1.
 \end{enumerate}
@@
 
-When the learning hypothesis space is low constrained, the learned actions can be reformulated and still be compliant with the inputs.
+When the learning hypothesis space is under constrained, the learned actions can be reformulated and still be compliant with the inputs.
 For instance in the {\em blocksworld}, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa).

### 2018-01-22 doc/ijcai-18/observations-learning.tex
Add files via upload

@@
 This paper presents a first novel approach to \strips\ action model recognition from state observations.
-Unlike other related works, our classical planning compilation approach not only allows learning a \strips\ action model from an observation sequence but it is also extendible to action model recognition.
-The learning task does not require observing the precise actions that produce the state observations because such actions are determined by a classical planner.
-We present a two-fold evaluation of the learned models: a syntactic and semantic assessment against a reference model and via a test dataset that calculates the probability of a model given the observation.
+Unlike other related works, our classical planning compilation approach learns a \strips\ action model from state observations instead of action observations and it is also extendible to action model recognition.
+We present a two-fold evaluation of the learned models: a syntactic and semantic assessment against a reference model and calculating the probability of the model given the observations with a test dataset.
 Moreover, we show how these probabilistic values are used to estimate the probability of the action model recognition.

### 2018-01-22 doc/ijcai-18/observations-learning.tex
added last results generated by Diego

--- 2018-01-22 doc/ijcai-18/observations-learning.tex
@@
 
+\subsubsection{Unsupervised evaluation of the learned models}
+
 \section{Conclusions}
@@
 				\hline
-				blocks & 0.57 & 0.44 & 0.63 & 0.56 & 0.57 & 0.44 & 0.59 & 0.48 \\
-				driverlog & 0.2 & 0.07 & 0.18 & 0.29 & 0.2 & 0.14 & 0.19 & 0.17 \\
-				ferry & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
-				floor-tile & 0.5 & 0.14 & 0.75 & 0.55 & 0.6 & 0.27 & 0.62 & 0.32 \\
-				Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
-				gripper-strips & 0.25 & 0.17 & 0.25 & 0.25 & 0.25 & 0.25 & 0.25 & 0.22 \\
-				hanoi & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
-				hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				n-puzzle & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				parking & 0.6 & 0.21 & 0.14 & 0.11 & 0.4 & 0.22 & 0.38 & 0.18 \\
-				satellite & 0.25 & 0.07 & 0.33 & 0.4 & 0.67 & 0.5 & 0.42 & 0.32 \\
+
+				blocks & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 \\
+				driverlog & 0.0 & 0.0 & 0.25 & 0.43 & 0.0 & 0.0 & 0.08 & 0.14 \\
+				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
+				floor-tile & 0.38 & 0.55 & 0.4 & 0.18 & 0.56 & 0.45 & 0.44 & 0.39 \\
+				grid & 0.5 & 0.47 & 0.33 & 0.29 & 0.25 & 0.29 & 0.36 & 0.35 \\
+				gripper-strips & 0.83 & 0.83 & 0.75 & 0.75 & 0.75 & 0.75 & 0.78 & 0.78 \\
+				hanoi & 0.5 & 0.25 & 0.5 & 0.5 & 0.0 & 0.0 & 0.33 & 0.25 \\
+				hiking & 0.43 & 0.43 & 0.5 & 0.35 & 0.44 & 0.47 & 0.46 & 0.42 \\
+				miconic & 0.6 & 0.33 & 0.33 & 0.25 & 0.33 & 0.33 & 0.42 & 0.31 \\
+				n-puzzle & 0.33 & 0.33 & 0.0 & 0.0 & 0.0 & 0.0 & 0.11 & 0.11 \\
+				parking & 0.25 & 0.21 & 0.0 & 0.0 & 0.0 & 0.0 & 0.08 & 0.07 \\
+				pegsol & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
 				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				transport & 0.33 & 0.1 & 0.33 & 0.4 & 0.0 & 0.0 & 0.22 & 0.17 \\
-				zeno-travel & 0.25 & 0.07 & 0.17 & 0.14 & 0.0 & 0.0 & 0.14 & 0.07 \\
+				transport & 1.0 & 0.3 & 0.8 & 0.8 & 1.0 & 0.6 & 0.93 & 0.57 \\
+				grid-visit-all & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
+				zeno-travel & 0.67 & 0.29 & 0.33 & 0.29 & 0.33 & 0.14 & 0.44 & 0.24 \\
 				\hline
@@
 	\end{center}
-	\caption{\small Precision and recall values obtained when learning from labels without computing the $f_{P\&R}$ mapping.}
-	\label{fig:labelsnomap}
+	\caption{\small Precision and recall values obtained without computing the $f_{P\&R}$ mapping.}
+	\label{fig:observationsnomap}
 
@@
 				\hline
-				blocks & 0.71 & 0.56 & 0.75 & 0.67 & 0.71 & 0.56 & 0.73 & 0.59 \\
-				driverlog & 0.6 & 0.21 & 0.27 & 0.43 & 0.6 & 0.43 & 0.49 & 0.36 \\
-				ferry & 0.5 & 0.29 & 0.75 & 0.75 & 0.5 & 0.5 & 0.58 & 0.51 \\
-				floor-tile & 0.5 & 0.14 & 0.75 & 0.55 & 0.6 & 0.27 & 0.62 & 0.32 \\
-				Grid & - & - & - & - & - & - & - & - \\ % Segmentation fault
-				gripper-strips & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-				hanoi & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
-				hiking & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				n-puzzle & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				parking & 0.6 & 0.21 & 0.14 & 0.11 & 0.4 & 0.22 & 0.38 & 0.18 \\
-				satellite & 0.25 & 0.07 & 0.33 & 0.4 & 0.67 & 0.5 & 0.42 & 0.32 \\
+
+				blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				driverlog & 0.67 & 0.14 & 0.33 & 0.57 & 0.67 & 0.29 & 0.56 & 0.33 \\
+				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
+				floor-tile & 0.44 & 0.64 & 1.0 & 0.45 & 0.89 & 0.73 & 0.78 & 0.61 \\
+				grid & 0.63 & 0.59 & 0.67 & 0.57 & 0.63 & 0.71 & 0.64 & 0.62 \\
+				gripper-strips & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+				hiking & 0.78 & 0.6 & 0.93 & 0.82 & 0.88 & 0.88 & 0.87 & 0.77 \\
+				miconic & 0.8 & 0.44 & 1.0 & 0.75 & 1.0 & 1.0 & 0.93 & 0.73 \\
+				n-puzzle & 0.67 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 0.89 \\
+				parking & 0.56 & 0.36 & 0.5 & 0.33 & 0.5 & 0.33 & 0.52 & 0.34 \\
+				pegsol & - & - & - & - & - & - & - & - \\ % Time out 1000s
+				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
 				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
-				transport & 1.0 & 0.3 & 0.67 & 0.8 & 0.67 & 0.4 & 0.78 & 0.5 \\
-				zeno-travel & 0.5 & 0.14 & 0.5 & 0.43 & 0.33 & 0.14 & 0.44 & 0.24 \\
+				transport & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0 & 0.63 \\ 
+				grid-visit-all & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 \\
+				zeno-travel & 1.0 & 0.43 & 0.67 & 0.57 & 1.0 & 0.43 & 0.89 & 0.48 \\                                
 				\hline
@@
 	\end{center}
-	\caption{\small Precision and recall values obtained when learning from labels but computing the $f_{P\&R}$ mapping.}
-	\label{fig:labels}
+	\caption{\small Precision and recall values obtained when computing the $f_{P\&R}$ mapping.}
+	\label{fig:observationsmap}                
 \end{table}
@@
 \begin{table}
-	\begin{footnotesize}
 		\begin{center}
-			\begin{tabular}{l|c|c|c|}			
-				& Total time & Preprocess & Plan length  \\
+			\begin{tabular}{l|l|l|l|}
+				& $\delta(\Xi,\mathcal{O})$ & $\delta(\Xi,*)$ & $P(\mathcal{O}|\Xi)$ \\
 				\hline
-				Blocks & 1.40 & 0.00 & 70  \\
-				Driverlog & 1.50 & 0.00 & 89 \\
-				Ferry & 1.49 & 0.00 & 64 \\
-				Floortile & 351.38 & 0.11 & 156 \\
-				Grid & - & - & - \\ % Segmentation fault
-				Gripper & 0.04 & 0.00 & 59 \\
-				Hanoi & 2.33 & 0.01 & 49  \\
-				Hiking & - & - & - \\ % Time out 1000s
-				Parking & - & - & - \\ % Time out 1000s
-				Satellite & 78.36 & 0.03 & 98 \\
-				Sokoban & - & - & - \\ % Time out 1000s
-				Transport & 285.61 & 0.10 & 106 \\
-				Zenotravel & 6.20 & 0.46 & 71 \\
+				block & 0 & 90 & 1.0 \\
+				ferry & 2 & 69 & 0.97 \\
+				grid & 42 & 153 & 0.73 \\
+				gripper & 2 & 30 & 0.93 \\
+				hanoi & 1 & 63 & 0.98 \\
+				miconic & 3 & 72 & 0.96 \\
+				npuzzle & 2 & 24 & 0.92 \\
+				transport & 4 & 78 & 0.95 \\
+				visitall & 2 & 24 & 0.92 \\
+				zenotravel & 3 & 63 & 0.95 \\
+				\hline
+				\bf  & - & - & - \\
 			\end{tabular}
 		\end{center}
-	\end{footnotesize}
- \caption{\small Planning results obtained when learning from labels.}
-\label{fig:planlabels}
+	\caption{\small $P(\mathcal{O}|\Xi)=1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$}
+	\label{fig:observationstest}                
 \end{table}

### 2018-01-22 doc/ijcai-18/observations-learning.tex
added results for the 5 last domains

@@
 				block & 0 & 90 & 1.0 \\
+				driverlog & 5 & 144 & 0.97 \\
 				ferry & 2 & 69 & 0.97 \\
+				floortile & 34 & 342 & 0.90 \\
 				grid & 42 & 153 & 0.73 \\
@@
 				hanoi & 1 & 63 & 0.98 \\
+				hiking & 69 & 174 & 0.60 \\
 				miconic & 3 & 72 & 0.96 \\
 				npuzzle & 2 & 24 & 0.92 \\
+				parking & 5 & 111 & 0.95 \\
+				satellite & 24 & 75 & 0.68 \\
 				transport & 4 & 78 & 0.95 \\

### 2018-01-23 doc/ijcai-18/observations-learning.tex
ijcai18 paper updates

@@
 
-\title{Learning and Recognition of \strips\ Action Models from State Observations \\ Action Model Recognition from State Observations via Planning}
+\title{Learning Action Models from State Observations}
 
-\author{
+\author{\#1186
 %Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
@@
 This paper presents a classical planning compilation for learning \strips\ action models from state observations.
-The compilation approach does not require observing the precise actions that produced the state observations because such actions are determined by a classical planner.
-In addition, the compilation is extensible to (1) semantically evaluate the quality of learned action models and (2) estimate the probability distribution of the possible action models given a state observation sequence.
-
-\textbf{EVA}
-
-This paper presents a first novel approach to \strips\ action model recognition from state observations.
-Unlike other related works, our classical planning compilation approach learns a \strips\ action model from state observations instead of action observations and it is also extendible to action model recognition.
-We present a two-fold evaluation of the learned models: a syntactic and semantic assessment against a reference model and calculating the probability of the model given the observations with a test dataset.
-Moreover, we show how these probabilistic values are used to estimate the probability of the action model recognition.
-
+The compilation approach does not require observing the precise actions that produced the observations because such actions are determined by a planner.
+Furthermore, the presented compilation is extensible to assess how well a \strips\ action model matches a given set of observations, which enables {\em model recognition}.
+Last but not least, the paper evaluates the performance of the proposed approach by learning action models for a wide range of classical planning domains from the International Planning Competition and assessing the learned models with respect to (1) the corresponding reference models and (2), given observations test sets.
 \end{abstract}
@@
 \begin{enumerate}
-\item Is defined as a classical planning compilation.
-This fact opens the door to the {\em bootstrapping} of planning action models and allows us to report results over a wide range of planning domains.
-\item Does not require observing the particular executed actions. An off-the-shelf classical planner determines these actions given the state observations.
-\item Is extensible to semantically evaluate the quality of learned \strips\ models as well as to estimate the probability distribution of the possible \strips\ models given a sequence of state observations.
+\item {\em Learning action models as planning}.
+The practicality of the compilation approach allow us to report results over a wide range of planning domains.
+Furthermore, opens up a way towards the {\em bootstrapping} of planning action models (a planner exploring its state space and using the obtained data to learn/update its action model).
+\item {\em Learning from state observations}.  Our approach does not require observing the precise actions that produced the observations because such actions are determined by a classical planner.
+\item {\em Model evaluation}. The compilation is extensible to assess how well a \strips\ action model matches a given set of observations, which enables {\em model recognition}. 
 \end{enumerate}
@@
 \section{Background}
-This section defines the planning models used in this work as well as the output (an \strips\ action model) of the addressed learning task.
+Our approach for learning \strips\ action models is compiling the leaning task into a classical planning task with conditional effects.
 
-\subsection{Classical planning}
+\subsubsection{Classical planning with conditional effects}
 We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
@@
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
-
-\subsubsection{Classical planning with conditional effects}
-Our approach for learning \strips\ action models is compiling the leaning task into a classical planning task with conditional effects.
 
@@
 
-Finally we also define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in a given action schema $\xi$.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} only contains the fluents from $F_v$ that do not involve $v_2$ because the action header contained the single parameter $v_1$.
+Finally we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in a given action schema $\xi$.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} only contains the fluents from $F_v$ that do not involve $v_2$ because the action header contains the single parameter $v_1$.
 
@@
 
-\begin{figure}[hbt!]
+\begin{figure}
 {\footnotesize\tt ;;; Predicates in $\Psi$}
@@
 ;;; observation #1
-(holding block2) (clear block1)
-(ontable block1)
+(holding block2) (clear block1) (ontable block1)
 
@@
 (clear block1) (ontable block1)
-(clear block2) (ontable block2)
-(handempty)
+(clear block2) (ontable block2) (handempty)
 
 ;;; observation #3
-(holding block1) (clear block2)
-(ontable block2)
+(holding block1) (clear block2) (ontable block2)
 
@@
 
-\section{\strips\ action models recognition}
-Inspired by {\em plan recognition as planning}~\cite{ramirez2009plan}, we show that the previous compilation can be extended to (1) semantically evaluate the quality of learned \strips\ models, without comparing with a reference model and (2), address \strips\ action model recognition tasks.
+\section{Evaluating \strips\ action models}
+The compilation can be extended to evaluate how well a \strips\ action model matches a given set of observations.
+This allows to asses the quality of learned models without a reference model and enables the recognition of \strips\ action models.
+
+The main idea behind the compilation extension is that how well a \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$, depends on the amount of {\em edition} that one has to introduce into $\Xi$ to produce $\mathcal{O}$ (in the extreme if $\Xi$ perfectly explains $\mathcal{O}$, no {\em edit} must be done).
 
 \subsection{The edit distance}
-We assume that how well a given \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$, depends on the amount of {\em edition} that one has to introduce to $\Xi$ to produce the sequence of observations $\mathcal{O}$.
-In the extreme, if the given model $\Xi$ perfectly explains the observations $\mathcal{O}$, no {\em edit} to the model must be introduced.
-
-Let us define the \emph{edit operations} allowed to edit a given \strips\ action model.
-With the aim of keeping tractable the branching factor of the classical planning task that results from our compilation, we only define two {\em \strips\ edit operations}:
+Let us define first the \emph{operations} allowed to edit a given \strips\ action model.
+With the aim of keeping tractable the branching factor of the planning instance that results from our compilation, we only define two {\em edit operations}:
 \begin{itemize}
@@
 
-With this defined, we can now formalize an edit distance metric that assesses how dissimilar two given \strips\ action models are.
-This distance satisfies the {\em metric axioms} provided that the two {\em \strips\ edit operations} have the same positive cost.
+We can now formalize an edit distance that quantifies how dissimilar two given \strips\ action models are.
+This distance is symmetric and satisfies the {\em metric axioms} provided that the two {\em \strips\ edit operations} have the same positive cost.
 
 \begin{definition}
-Let $\Xi$ and $\Xi'$ be two action models, both built from the same set of possible elements $F_v$.
-The {\bf edit distance}, denoted as $\delta(\Xi,\Xi')$, is the minimum number of {\em edit operations} to transform $\Xi$ into $\Xi'$.
+Let $\Xi$ and $\Xi'$ be two \strips\ action models, both built from the same set of possible elements $F_v$.
+The {\bf edit distance}, denoted as $\delta(\Xi,\Xi')$, is the minimum number of {\em edit operations} required to transform $\Xi$ into $\Xi'$.
 \end{definition}
 
-Since the size of the $F_v$ set is bound, the maximum number of edits that can be introduced to a given action model is bound as well.
-In more detail, for a given operator schema $\xi\in\Xi$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$.
-With regard to effects of $\xi$, the maximum number of edits that can be introduced are two times $|F_v(\xi)|$.
+Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well.
+In more detail, for an operator schema $\xi\in\Xi$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$.
+With regard to the effects, the maximum number of edits that can be introduced are two times $|F_v(\xi)|$.
 
@@
 
-Likewise, we can define a distance of a model $\Xi$ to an observation sequence $\mathcal{O}$ on the basis of the \emph{edit distance} between two models.
+Finally, we are ready to define the distance of a model $\Xi$ to an observation sequence $\mathcal{O}$ on the basis of the \emph{edit distance} between two models.
 
@@
 Given an action model $\Xi$ built from the set of possible elements $F_v$ and a sequence of observations $\mathcal{O}$.
-The {\bf observation edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimal edit distance to transform $\Xi$ into a model $\Xi'$ which produces a plan $\pi=\tup{a_1, \ldots, a_n}$ such that $\pi$ induces the observation sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$; i.e., \[\delta(\Xi,\mathcal{O})=\argmin_{\Xi' \rightarrow \mathcal{O}} \delta(\Xi,\Xi')\]
+The {\bf observation edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimal edit distance to transform $\Xi$ into a model $\Xi'$ which: (1) is definable within $F_v$ and (2), can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ such that $\pi$ induces the observation sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$; i.e., \[\delta(\Xi,\mathcal{O})=\min_{\exists\ \Xi' s.t.
+\Xi' \rightarrow \mathcal{O}} \delta(\Xi,\Xi')\]
 \end{definition}
 
-The {\em observation edit distance} allow us to semantically evaluate the quality of a learned strips action model with respect to a sequence of observations that acts as a test set.
-The lower this distance is, the better the quality of the learned model.
+The {\em observation edit distance} asses the quality of a learned strips action model with respect to a sequence of observations (that acts as a test set).
+The lower this distance, the better the learned model.
 The semantic nature of this evaluation is robust to learning episodes where actions are reformulated and still compliant with the inputs (e.g.
@@
 
-\subsection{Model recognition}
-Given a set of possible \strips\ models and a sequence of state observations the {\em recognition of \strips\ models} is the task of computing the probability distribution of these models according to the given observations.
-Here we show that the {\em observations  \strips\ edit distance} is helpful to compute the probability distribution of the possible \strips\ models given a sequence of state observations.
-The main idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map these distances into likelihoods using the Bayes rule.
+Given a set of possible \strips\ models and a sequence of state observations, the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
+The {\em observations edit distance} is also helpful to compute the probability distribution of the possible \strips\ models given a sequence of state observations.
+The main idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods and use the Bayes rule.
 
 According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the possible \strips\ action models that can be built given a set of operator headers, or in other words, given the $F_v(\xi)$ sets.
-Moreover, we assume that all the possible \strips\ action models are equiprobable so according to the Bayes rule $P(\mathcal{H}|\mathcal{O})$ is proportional to $P(\mathcal{O}|\mathcal{H})$.
-Finally, we assume that $P(\mathcal{O}|\mathcal{H})$ is given by the {\em observation edit distance}, mapping distances into probabilities (the larger the distance the lower the likelihood), according to the following expression $P(\mathcal{O}|\Xi)=1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$.
+In our scenario, the hypotheses are about the possible \strips\ action models that can be built within a given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets).
+Moreover, we assume that $P(\mathcal{O}|\mathcal{H})$ is given by the {\em observation edit distance}, mapping distances into probabilities, according to the following expression $P(\mathcal{O}|\Xi)=1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$.
+The larger the distance the lower the likelihood.
 
-Thus, $P(\Xi|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the set of operator headers, that is within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$, is computed by:
-
+With this regard, $P(\Xi|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
 \begin{enumerate}
-\item Calculating the {\em observation edit distance} $\delta(\Xi,\mathcal{O})$ for every possible model $\Xi$.
-\item Applying the resulting distances to the above formula $P(\mathcal{O}|\Xi)$ to map distances into likelihoods
-\item Applying Bayes rule to obtain the normalized posterior probabilities $P(\Xi|\mathcal{O})$. The sum of the probabilities of all the models will be 1.
+\item Computing the {\em observation edit distance} $\delta(\Xi,\mathcal{O})$ for every possible model $\Xi$.
+\item Applying the resulting distances to the above $P(\mathcal{O}|\Xi)$ formula to map these distances into likelihoods
+\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
 \end{enumerate}
 
-\subsection{\strips\ model evaluation and recognition with classical planning}
+\subsection{Evaluating \strips\ models with planning}
 The previous compilation can be extended to estimate the {\em observations  \strips\ edit distance}.
-In this case the tuple $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ does not represent a learning task but the edition of a given \strips\ action model $\Xi$ to cover the sequence of observations $\mathcal{O}$.
+In this case the tuple $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ does not represent a learning task but the edition of a \strips\ action model $\Xi$ (the model to evaluate) to cover the sequence of observations $\mathcal{O}$.
+
 The extended compilation outputs a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$ where:
@@
 In addition, the given \strips\ action model $\Xi$ is now encoded in the initial state.
-This means that fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ with $f\in F_v(\xi)$ hold in the initial state if they appear in the given \strips\ action model $\Xi$.
+This means that fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ with $f\in F_v(\xi)$ hold in the initial state if they appear in the given $\Xi$ model.
 \item $G_{\Lambda}$ is defined as in the previous compilation.
-\item $A_{\Lambda}'$, comprises the three kinds of actions of $A_{\Lambda}$.
+\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
 The Actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation {\tt\small $1\leq i\leq n$} are defined exactly as in the previous compilation.
-The only difference are the actions for {\em programming} operator schema $\xi\in\Xi$ that must include now actions for adding a precondition and for removing a {\em negative} or {\em positive} effect.
+The only difference are the actions for {\em programming} operator schema $\xi\in\Xi$ that include now actions for adding a precondition and for removing a {\em negative} or {\em positive} effect (that is implement the {\em edit operations}).
 \end{itemize}
 
-The probability distribution $P(\Xi|\mathcal{O})$ would be exact if the classical planning problems resulting from our compilation are solved optimally (according to the number of edit actions), and will be approximate if they are solved with a satisfying classical planner or furthermore if what is solved is not $P_{\Lambda}'$ but a relaxation of this task such as the {\em delete relaxation}.
+A solution to the resulting classical planning task from the extended compilation is a sequence of actions that edits the \strips\ action model $\Xi$ and validates the edited model on the given observations $\mathcal{O}$.
+In model evaluation we are not interested in the edited model but in the number of {\em edit operations} neeeded to produce a model validable in the given observations, i.e.
+the {\em observation edit distance}.
+In more detail, the {\em observation edit distance} is exact if $P_{\Lambda}'$, the classical planning problem resulting from the extended compilation, is optimally solved (according to the number of edit actions), and is approximate if $P_{\Lambda}'$ is solved with a satisfying planner or furthermore, if what is solved is not $P_{\Lambda}'$ but its relaxation, e.g.
+the {\em delete relaxation}~\cite{bonet2001planning}.
 
 \section{Evaluation}
-This section evaluates our approach for learning \strips\ models starting from different amounts of input knowledge.
+This section presents a two-fold evaluation of the learned models: with respect to a reference model and with respect to a given set of observations.
 
-\subsubsection{Reproducibility}
-We used IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+We used 15 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
 For the learning of the \strips\ action models we used observations sequences of 25 states per domain.
@@
 
-The compilation source code, the evaluation scripts and the benchmarks are fully available at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper can be reproduced.
+The compilation source code, the evaluation scripts and the benchmarks (including learning and testing sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
 
-\subsubsection{Supervised evaluation of the learned models}
+\subsubsection{Reference model}
+Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning classical planning models from the IPC.
 For each domain the learned model is compared with the actual model and its quality is quantified with the {\em precision} and {\em recall} metrics.
@@
 
+Table~\ref{fig:observationsnomap} summarizes the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
+
 When the learning hypothesis space is under constrained, the learned actions can be reformulated and still be compliant with the inputs.
-For instance in the {\em blocksworld}, operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator (and vice versa).
-Furthermore, in a given action the role of the parameters that share the same type can be interchanged making non trivial to compute {\em precision} and {\em recall} with respect to a reference model.
-
+Given the syntax-based nature of these metrics, it may happen that these metrics report low scores for learned models that are actually good but correspond to "reformulations" of the actual model; i.e.
+a learned model semantically equivalent but syntactically different to the reference model.
 To address these issues we defined an evaluation method robust to action reformulation.
@@
 With this mapping defined we can compute the {\em precision} and {\em recall} of a learned \strips\ action $\xi\in\Xi$ with respect to the action $f_{P\&R}(\xi)\in \Xi^*$ even if actions are reformulated in the learning process.
+Table~\ref{fig:observationsmap} reports results significantly higher which shows that in many domains the learned actions or their parameters interchanged their roles with respect to the reference model.
 
-\subsubsection{Unsupervised evaluation of the learned models}
-
-\section{Conclusions}
-As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains.
-Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
-
-%The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
-In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
-
-The size of the compiled classical planning instances depends on the number of input examples.
-Generating {\em informative} examples for learning planning action models is an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
-The combination of such exploration techniques with our learning approach is an intriguing research direction that opens the door to the bootstrapping of planning action models.
-
-%Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
-For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true {\em at least at one} of the reached states.
-Despite this is beyond the scope of this paper, LTL constraints could be included in our compilation following the ideas for compiling temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) force the solution plans to always leave the LTL automata at an acceptor state by adding new goals to the classical planning task.
-
-% Commented for blind submission
-%\begin{small}
-%\subsection*{Acknowledgment}
-%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
-Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
-%\end{small}
+\subsubsection{Test set}
+When learning action models is reasonable to assume that the reference model is not available.
+In this case we compute the observations distance to asses the quality of the learned models withs respect to a set of obserations.
+Table~\ref{fig:observationstest} summarizes the obtained results.
+Results evidence that, like when using the $f_{P\&R}$ mapping, this evaluation metric is also robust to reformulation in the learned models.
 
@@
 				parking & 0.25 & 0.21 & 0.0 & 0.0 & 0.0 & 0.0 & 0.08 & 0.07 \\
-				pegsol & - & - & - & - & - & - & - & - \\ % Time out 1000s
 				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
-				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
 				transport & 1.0 & 0.3 & 0.8 & 0.8 & 1.0 & 0.6 & 0.93 & 0.57 \\
@@
 				parking & 0.56 & 0.36 & 0.5 & 0.33 & 0.5 & 0.33 & 0.52 & 0.34 \\
-				pegsol & - & - & - & - & - & - & - & - \\ % Time out 1000s
 				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
-				Sokoban & - & - & - & - & - & - & - & - \\ % Time out 1000s
 				transport & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0 & 0.63 \\ 
@@
 		\begin{center}
+                \begin{scriptsize}
 			\begin{tabular}{l|l|l|l|}
@@
 			\end{tabular}
+                        	\end{scriptsize}
 		\end{center}
-	\caption{\small $P(\mathcal{O}|\Xi)=1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$}
+	\caption{\small Evaluation of the quality of the learned models with respect to an observations test set.}
 	\label{fig:observationstest}                
 \end{table}
+
+\section{Conclusions}
+As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains.
+~\cite{asai2017classical} reported the learning of PDDL action models on a few domains from large images datasets.
+Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
+
+%The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
+In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
+
+The size of the compiled classical planning instances depends on the number of input examples.
+The larger the number of examples is, the larger the compilation (a larger number of "test" fluents and "validate" actions), and also the more actions the solution plan requires to derive the action model.
+All this implies larger planning times.
+On the other hand, the precision and recall values depend on the diversity of the training samples rather than on the size of the input data.
+
+Generating {\em informative} examples for learning planning action models is an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
+The combination of such exploration techniques with our learning approach is an intriguing research direction towards the bootstrapping of planning action models.
+
+%Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
+For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true {\em at least at one} of the reached states.
+Despite this is beyond the scope of this paper, LTL constraints could be included in our compilation following the ideas for compiling temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) force the solution plans to always leave the LTL automata at an acceptor state by adding new goals to the classical planning task.
+
+% Commented for blind submission
+%\begin{small}
+%\subsection*{Acknowledgment}
+%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
+Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
+%\end{small}
 

### 2018-01-24 doc/ijcai-18/observations-learning.tex
new updates to the ijcai18 paper

@@
 The compilation approach does not require observing the precise actions that produced the observations because such actions are determined by a planner.
-Furthermore, the presented compilation is extensible to assess how well a \strips\ action model matches a given set of observations, which enables {\em model recognition}.
+Furthermore, the presented compilation is extensible to assess how well a \strips\ action model matches a given set of observations.
 Last but not least, the paper evaluates the performance of the proposed approach by learning action models for a wide range of classical planning domains from the International Planning Competition and assessing the learned models with respect to (1) the corresponding reference models and (2), given observations test sets.
@@
 Furthermore, opens up a way towards the {\em bootstrapping} of planning action models (a planner exploring its state space and using the obtained data to learn/update its action model).
-\item {\em Learning from state observations}.  Our approach does not require observing the precise actions that produced the observations because such actions are determined by a classical planner.
+\item {\em Learning from state observations}.
+Our approach does not require the learning samples to be observations of actions but simply state observations, broadening the range of application to external observers.
 \item {\em Model evaluation}. The compilation is extensible to assess how well a \strips\ action model matches a given set of observations, which enables {\em model recognition}. 
@@
 \section{Background}
-Our approach for learning \strips\ action models is compiling the leaning task into a classical planning task with conditional effects.
+Our approach for learning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
 
@@
 
-Let us also define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and that defines the elements that can appear in an action schema.
+Let us define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and that defines the elements that can appear in an action schema.
 For the blocksworld, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
@@
 
-Finally we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in a given action schema $\xi$.
+Finally, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in a given action schema $\xi$ and that confine the space of possible action models.
 For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} only contains the fluents from $F_v$ that do not involve $v_2$ because the action header contains the single parameter $v_1$.
@@
 \end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
+ \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
 \label{fig:stack}
@@
 \section{Learning \strips\ action models}
-Learning \strips\ action models from fully available input knowledge, i.e.
-from plans where every action in the plan is available as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
+Learning \strips\ action models from plans where every action in the plan is available as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
 In this case, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions.
+Preconditions are derived lifting the minimal set of literals that appears in all the corresponding pre-states.
 
-In this paper we address a more challenging learning task, where less input knowledge is available.
+This paper addresses a more challenging learning task, where less input knowledge is available.
 The addressed learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
-This learning task is defined as $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$, where:
+This learning task is defined as $\Lambda=\tup{\Xi,\Psi,\mathcal{O}}$, where:
 \begin{itemize}
-\item $\Psi$ is the set of predicates that define the abstract state space of a given classical planning frame.
 \item $\Xi$ is the set of empty operator schemes, wherein each $\xi\in\Xi$ is only composed of $head(\xi)$.
+\item $\Psi$ is the set of predicates, that define the abstract state space of a given classical planning frame.
 \item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$.
@@
 
-A solution to $\Lambda$ is a set of operator schema $\Xi'$ compliant with the predicates in $\Psi$, the headers in $\Xi$ and the given sequence of state observations $\mathcal{O}$.
+A solution to $\Lambda$ is a set of operator schema $\Xi'$ compliant with the headers $\Xi$, the predicates $\Psi$, and the given sequence of state observations $\mathcal{O}$.
 A planning compilation is a suitable approach for addressing a $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains $\mathcal{O}$.
-Figure~\ref{fig:lexample} shows an example of a $\Lambda$ task for learning a \strips\ action model in the blocksworld from the sequence of five state observations that corresponds to inverting a 2-blocks tower.
+Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning a \strips\ action model in the blocksworld from the sequence of five state observations that corresponds to inverting a 2-blocks tower.
 
 \begin{figure}
-{\footnotesize\tt ;;; Predicates in $\Psi$}
+{\footnotesize\tt ;;;;;; Headers in $\Xi$}
+\begin{footnotesize}
+\begin{verbatim}
+(pickup v1) (putdown v1)
+(stack v1 v2} (unstack v1 v2)
+\end{verbatim}
+\end{footnotesize}
+\vspace{0.2cm}
+{\footnotesize\tt ;;;;;; Predicates $\Psi$}
 \begin{footnotesize}
@@
 \vspace{0.2cm}
-{\footnotesize\tt ;;; Headers in $\Xi$}
-\begin{footnotesize}
-\begin{verbatim}
-(pickup v1) (putdown v1)
-(stack v1 v2} (unstack v1 v2)
-\end{verbatim}
-\end{footnotesize}
-\vspace{0.2cm}
-{\footnotesize\tt ;;; Observations in $\mathcal{O}$}
+{\footnotesize\tt ;;;;;; Observations $\mathcal{O}$}
 \begin{footnotesize}
@@
 
-Our approach for addressing the learning task is compiling $\Lambda$ into a classical planning task with conditional effects.
+Our approach for addressing the learning task is compiling $\Lambda$ into a classical planning task $P_{\Lambda}$ with conditional effects.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \begin{enumerate}
-\item Programs the \strips\ action model $\Xi'$.
+\item {\bf Programs the \strips\ action model $\Xi'$}.
 A solution plan starts with a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the \strips\ action model $\Xi'$ in $\mathcal{O}$.
+\item {\bf Validates the \strips\ action model $\Xi'$ in $\mathcal{O}$}.
 The solution plan continues with a postfix that produces the given sequence of states $\tup{s_0,s_1,\ldots,s_{n}}$ using the programmed action model $\Xi'$.
@@
 
-Given a learning task $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+Given a learning task $\Lambda=\tup{\Xi,\Psi,\mathcal{O}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
 \begin{itemize}
@@
 \begin{itemize}
-\item The set of fluents $F$ that is built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$.
-\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$.
-For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$.
+\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi'$.
+For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) is represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
 \item Fluent $mode_{prog}$ indicating whether the operator schemes are programmed or validated (already programmed) and fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
@@
 \item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
-In addition, our compilation assumes that initially operator schemes are programmed with every possible precondition, no negative effect and no positive effect.
-With this regard, the fluents $pre_f(\xi)$ hold at the initial state for every $f\in F_v(\xi)$.
+Our compilation assumes that initially operator schemes are programmed with every possible precondition, no negative effect and no positive effect.
+Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
 \item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, indicates that the programmed action model is validated in all the input observations.
@@
 \item Actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
-We assume that the operators headers are known so the binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
-variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
+Given that the operators headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
 Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
@@
 \end{scriptsize}
- \caption{\small Action for applying an already programmed schema $stack$ as encoded in PDDL (implications coded as disjunctions).}
+ \caption{\small Action for applying an already programmed schema $stack$ as encoded in PDDL (implications are coded as disjunctions).}
 \label{fig:compilation}
@@
 \begin{lemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi'$ that solves the learning task $\Lambda$.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi'$ that solves the $\Lambda$ learning task.
 \end{lemma}
@@
 The same happens with the positive and negative effects.
-Furthermore because of the preconditions of the $\mathsf{programPre_{f,\xi}}$ actions, effects are only programmable after preconditions are programmed.
-Once the operator schemes $\Xi'$ are programmed, they can only be applied because of the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, the goals $\{test_i\}$, {\small $1\leq i\leq n$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following  the sequence {\small $1\leq i\leq n$}.
-Therefore if $test_n$ is achieved, it means that the programmed action model $\Xi'$ is compliant with the provided input knowledge and hence, solves $\Lambda$.
+Furthermore, once operator schemes $\Xi'$ are programmed, they can only be applied because of the $mode_{prog}$ fluent.
+Finally, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
+These goals can only be achieved executing an applicable sequence of programmed operator schemes that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence {\small $1\leq i\leq n$} which means that the programmed action model $\Xi'$ complies with the provided observations and hence, solves $\Lambda$.
 \end{small}
@@
 \begin{small}
-By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the set of elements that can appear in a \strips\ action schema $\xi\in\Xi$.
-In addition, any possible \strips\ action schema $\Xi'$ that can be built with the fluents in $F_v$ can be computed  with the $P_{\Lambda}$ compilation.
-The only \strips\ action models that the compilation cannot compute are the ones that violate the state constraints defined by the $\mathcal{O}$ sequence, which constrain as well the solutions to the $\Lambda$ learning task.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfy the state trajectory constraints given by the $\mathcal{O}$ sequence.
+By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\Xi$ given its header and the set of predicates $\Psi$.
 \end{small}
@@
 
-\section{Evaluating \strips\ action models}
-The compilation can be extended to evaluate how well a \strips\ action model matches a given set of observations.
-This allows to asses the quality of learned models without a reference model and enables the recognition of \strips\ action models.
+The size of the compiled classical planning instances depends on the number of input examples.
+The larger the number of examples, the larger the compilation (a larger number of "test" fluents and "validate" actions), and also the more actions the solution plan requires to derive the action model.
 
-The main idea behind the compilation extension is that how well a \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$, depends on the amount of {\em edition} that one has to introduce into $\Xi$ to produce $\mathcal{O}$ (in the extreme if $\Xi$ perfectly explains $\mathcal{O}$, no {\em edit} must be done).
+\section{Evaluation of \strips\ action models}
+How well a \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$ can be assessed according to the amount of {\em edition} required by $\Xi$ to produce $\mathcal{O}$ (in the extreme, if $\Xi$ perfectly explains $\mathcal{O}$, then no {\em edition} of the model is necessary).
 
-\subsection{The edit distance}
+\subsection{Edition of \strips\ action models}
 Let us define first the \emph{operations} allowed to edit a given \strips\ action model.
@@
 Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well.
-In more detail, for an operator schema $\xi\in\Xi$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$.
-With regard to the effects, the maximum number of edits that can be introduced are two times $|F_v(\xi)|$.
+In more detail, for an operator schema $\xi\in\Xi$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects are two times $|F_v(\xi)|$.
 
@@
 Let $\Xi$ be an action model built from the set of possible elements $F_v$.
-The \textbf{maximum edit distance}, defined as $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$, is an upper bound on the distance from $\Xi$ to any \strips\ action model definable within $F_v$.
+The \textbf{maximum edit distance} is $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$ and sets an upperbound on the edit distance from $\Xi$ to any other \strips\ action model definable within $F_v$.
 \end{definition}
 
-Finally, we are ready to define the distance of a model $\Xi$ to an observation sequence $\mathcal{O}$ on the basis of the \emph{edit distance} between two models.
+\subsection{The observation edit distance}
+We define now an edit distance to asses the quality of learned models with respect to a sequence of observations (that acts as a test set).
 
@@
 Given an action model $\Xi$ built from the set of possible elements $F_v$ and a sequence of observations $\mathcal{O}$.
-The {\bf observation edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimal edit distance to transform $\Xi$ into a model $\Xi'$ which: (1) is definable within $F_v$ and (2), can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ such that $\pi$ induces the observation sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$; i.e., \[\delta(\Xi,\mathcal{O})=\min_{\exists\ \Xi' s.t.
-\Xi' \rightarrow \mathcal{O}} \delta(\Xi,\Xi')\]
+The {\bf observation edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimal edit distance from $\Xi$ to a model $\Xi'$ such that: (1) $\Xi'$ is also definable within $F_v$ and (2), $\Xi'$ can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$, the given observations sequence; i.e., \[\delta(\Xi,\mathcal{O})=\min_{\forall \Xi' \rightarrow \mathcal{O}} \delta(\Xi,\Xi')\]
 \end{definition}
 
-The {\em observation edit distance} asses the quality of a learned strips action model with respect to a sequence of observations (that acts as a test set).
-The lower this distance, the better the learned model.
-The semantic nature of this evaluation is robust to learning episodes where actions are reformulated and still compliant with the inputs (e.g.
-the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa or the roles of actions parameters with the same type could be interchanged).
+Because its semantic nature, the {\em observation edit distance} is robust to learning episodes where actions are reformulated and still compliant with the inputs (e.g.
+the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa or the roles of the action parameters could be interchanged).
 
-Given a set of possible \strips\ models and a sequence of state observations, the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
-The {\em observations edit distance} is also helpful to compute the probability distribution of the possible \strips\ models given a sequence of state observations.
-The main idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods and use the Bayes rule.
+Furthermore the {\em observation edit distance} enables the recognition of \strips\ action models.
+The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
+We can map the {\em observation edit distance} into a likelihood with the following expression $P(\mathcal{O}|\Xi)=1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$.
+The larger the {\em observation edit distance} the lower this likelihood.
 
+%Given a set of possible \strips\ models and a sequence of state observations, the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
 According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
 In our scenario, the hypotheses are about the possible \strips\ action models that can be built within a given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets).
-Moreover, we assume that $P(\mathcal{O}|\mathcal{H})$ is given by the {\em observation edit distance}, mapping distances into probabilities, according to the following expression $P(\mathcal{O}|\Xi)=1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$.
-The larger the distance the lower the likelihood.
+With this regard, $P(\Xi|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
+%\begin{enumerate}
+%\item Computing the {\em observation edit distance} $\delta(\Xi,\mathcal{O})$ for every possible model $\Xi$.
+%\item Applying the resulting distances to the above $P(\mathcal{O}|\Xi)$ formula to map these distances into likelihoods
+%\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
+%\end{enumerate}
 
-With this regard, $P(\Xi|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
-\begin{enumerate}
-\item Computing the {\em observation edit distance} $\delta(\Xi,\mathcal{O})$ for every possible model $\Xi$.
-\item Applying the resulting distances to the above $P(\mathcal{O}|\Xi)$ formula to map these distances into likelihoods
-\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
-\end{enumerate}
-
-\subsection{Evaluating \strips\ models with planning}
-The previous compilation can be extended to estimate the {\em observations  \strips\ edit distance}.
-In this case the tuple $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ does not represent a learning task but the edition of a \strips\ action model $\Xi$ (the model to evaluate) to cover the sequence of observations $\mathcal{O}$.
-
+\subsection{Computing the observation edit distance}
+Our compilation can be extended to compute the {\em observations edit distance}.
+In this case the tuple $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ is not representing a learning task but the task of editing the \strips\ action model $\Xi$ to produce the observations $\mathcal{O}$.
 The extended compilation outputs a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$ where:
 \begin{itemize}
-\item $F_{\Lambda}$ is defined as in the previous compilation.
+\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
 \item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
 In addition, the given \strips\ action model $\Xi$ is now encoded in the initial state.
-This means that fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ with $f\in F_v(\xi)$ hold in the initial state if they appear in the given $\Xi$ model.
-\item $G_{\Lambda}$ is defined as in the previous compilation.
+This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, with $f\in F_v(\xi)$, hold in the initial state iff they appear in the given $\Xi$ model.
 \item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
 The Actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation {\tt\small $1\leq i\leq n$} are defined exactly as in the previous compilation.
-The only difference are the actions for {\em programming} operator schema $\xi\in\Xi$ that include now actions for adding a precondition and for removing a {\em negative} or {\em positive} effect (that is implement the {\em edit operations}).
+The only difference is the actions for {\em programming} operator schema $\xi\in\Xi$ that include now actions for adding a precondition and removing a {\em negative} or {\em positive} effect.
+They now implement the two {\em edit operations}.
 \end{itemize}
 
-A solution to the resulting classical planning task from the extended compilation is a sequence of actions that edits the \strips\ action model $\Xi$ and validates the edited model on the given observations $\mathcal{O}$.
-In model evaluation we are not interested in the edited model but in the number of {\em edit operations} neeeded to produce a model validable in the given observations, i.e.
-the {\em observation edit distance}.
-In more detail, the {\em observation edit distance} is exact if $P_{\Lambda}'$, the classical planning problem resulting from the extended compilation, is optimally solved (according to the number of edit actions), and is approximate if $P_{\Lambda}'$ is solved with a satisfying planner or furthermore, if what is solved is not $P_{\Lambda}'$ but its relaxation, e.g.
+A solution to the classical planning task resulting from the extended compilation is a sequence of actions that edits the \strips\ action model $\Xi$ and validates the edited model on the given observations $\mathcal{O}$.
+Since we are computing the {\em observations edit distance}, we are not interested in the edited model but in the number of {\em edit operations} required to produce a model validable in the given observations.
+The {\em observation edit distance} is exact if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions), and is approximate if $P_{\Lambda}'$ is solved with a satisfying planner or furthermore, if what is solved is not $P_{\Lambda}'$ but its relaxation, e.g.
 the {\em delete relaxation}~\cite{bonet2001planning}.
@@
 \section{Evaluation}
-This section presents a two-fold evaluation of the learned models: with respect to a reference model and with respect to a given set of observations.
+This section presents a two-fold evaluation of the learned models: with respect to a reference model and with respect to given sets of observations.
 
+\subsubsection{Reproducibility}
 We used 15 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-For the learning of the \strips\ action models we used observations sequences of 25 states per domain.
+For the learning of the \strips\ action models we only used observations sequences $\tup{s_0, s_1, \ldots, s_{24}}$ of twenty five states per domain.
+These sets of learning examples are fixed for all the experiments so the results reported by the different evaluation approaches is comparable.
 All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
@@
 {\sc Madagascar} is the classical planner we use to solve the instances that result from our compilations because its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
-In addition, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
+In addition, because its SAT-based nature, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
 Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
@@
 The compilation source code, the evaluation scripts and the benchmarks (including learning and testing sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
-
-\subsubsection{Reference model}
-Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning classical planning models from the IPC.
-For each domain the learned model is compared with the actual model and its quality is quantified with the {\em precision} and {\em recall} metrics.
-Precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
-$Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
-$Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
-
-Table~\ref{fig:observationsnomap} summarizes the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
-
-When the learning hypothesis space is under constrained, the learned actions can be reformulated and still be compliant with the inputs.
-Given the syntax-based nature of these metrics, it may happen that these metrics report low scores for learned models that are actually good but correspond to "reformulations" of the actual model; i.e.
-a learned model semantically equivalent but syntactically different to the reference model.
-To address these issues we defined an evaluation method robust to action reformulation.
-Precision and recall are often combined using the {\em harmonic mean}.
-This expression is called the {\em F-measure} (or the balanced {\em F-score}) and is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
-Given a reference \strips\ action model $\Xi^*$ and the learned \strips\ action model $\Xi$ we define the bijective function $f_{P\&R}:\Xi \mapsto \Xi^*$ such that $f_{P\&R}$ maximizes the accumulated {\em F-measure}.
-With this mapping defined we can compute the {\em precision} and {\em recall} of a learned \strips\ action $\xi\in\Xi$ with respect to the action $f_{P\&R}(\xi)\in \Xi^*$ even if actions are reformulated in the learning process.
-Table~\ref{fig:observationsmap} reports results significantly higher which shows that in many domains the learned actions or their parameters interchanged their roles with respect to the reference model.
-
-\subsubsection{Test set}
-When learning action models is reasonable to assume that the reference model is not available.
-In this case we compute the observations distance to asses the quality of the learned models withs respect to a set of obserations.
-Table~\ref{fig:observationstest} summarizes the obtained results.
-Results evidence that, like when using the $f_{P\&R}$ mapping, this evaluation metric is also robust to reformulation in the learned models.
 
@@
 				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
-				floor-tile & 0.38 & 0.55 & 0.4 & 0.18 & 0.56 & 0.45 & 0.44 & 0.39 \\
+				floortile & 0.38 & 0.55 & 0.4 & 0.18 & 0.56 & 0.45 & 0.44 & 0.39 \\
 				grid & 0.5 & 0.47 & 0.33 & 0.29 & 0.25 & 0.29 & 0.36 & 0.35 \\
-				gripper-strips & 0.83 & 0.83 & 0.75 & 0.75 & 0.75 & 0.75 & 0.78 & 0.78 \\
+				gripper & 0.83 & 0.83 & 0.75 & 0.75 & 0.75 & 0.75 & 0.78 & 0.78 \\
 				hanoi & 0.5 & 0.25 & 0.5 & 0.5 & 0.0 & 0.0 & 0.33 & 0.25 \\
@@
 				miconic & 0.6 & 0.33 & 0.33 & 0.25 & 0.33 & 0.33 & 0.42 & 0.31 \\
-				n-puzzle & 0.33 & 0.33 & 0.0 & 0.0 & 0.0 & 0.0 & 0.11 & 0.11 \\
+				npuzzle & 0.33 & 0.33 & 0.0 & 0.0 & 0.0 & 0.0 & 0.11 & 0.11 \\
 				parking & 0.25 & 0.21 & 0.0 & 0.0 & 0.0 & 0.0 & 0.08 & 0.07 \\
@@
 				transport & 1.0 & 0.3 & 0.8 & 0.8 & 1.0 & 0.6 & 0.93 & 0.57 \\
-				grid-visit-all & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
-				zeno-travel & 0.67 & 0.29 & 0.33 & 0.29 & 0.33 & 0.14 & 0.44 & 0.24 \\
-				\hline
-				\bf  & - & - & - & - & - & - & - & - \\
+				visitall & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
+				zenotravel & 0.67 & 0.29 & 0.33 & 0.29 & 0.33 & 0.14 & 0.44 & 0.24 
 			\end{tabular}
@@
 	\end{center}
-	\caption{\small Precision and recall values obtained without computing the $f_{P\&R}$ mapping.}
+	\caption{\small Precision and recall values obtained without computing the $f_{P\&R}$ mapping with the reference model.}
 	\label{fig:observationsnomap}
-
 \begin{center}
@@
 				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
-				floor-tile & 0.44 & 0.64 & 1.0 & 0.45 & 0.89 & 0.73 & 0.78 & 0.61 \\
+				floortile & 0.44 & 0.64 & 1.0 & 0.45 & 0.89 & 0.73 & 0.78 & 0.61 \\
 				grid & 0.63 & 0.59 & 0.67 & 0.57 & 0.63 & 0.71 & 0.64 & 0.62 \\
-				gripper-strips & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
 				hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
@@
 				miconic & 0.8 & 0.44 & 1.0 & 0.75 & 1.0 & 1.0 & 0.93 & 0.73 \\
-				n-puzzle & 0.67 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 0.89 \\
+				npuzzle & 0.67 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 0.89 \\
 				parking & 0.56 & 0.36 & 0.5 & 0.33 & 0.5 & 0.33 & 0.52 & 0.34 \\
@@
 				transport & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0 & 0.63 \\ 
-				grid-visit-all & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 \\
-				zeno-travel & 1.0 & 0.43 & 0.67 & 0.57 & 1.0 & 0.43 & 0.89 & 0.48 \\                                
-				\hline
-				\bf  & - & - & - & - & - & - & - & - \\
+				visitall & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 \\
+				zenotravel & 1.0 & 0.43 & 0.67 & 0.57 & 1.0 & 0.43 & 0.89 & 0.48 
 			\end{tabular}
@@
 	\end{center}
-	\caption{\small Precision and recall values obtained when computing the $f_{P\&R}$ mapping.}
+	\caption{\small Precision and recall values obtained when computing the $f_{P\&R}$ mapping with the reference model.}
 	\label{fig:observationsmap}                        
@@
 
+\subsection{Evaluating with a reference model}
+Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning classical planning models from the IPC.
+For each domain, the learned model is compared with the actual model and its quality is quantified with the {\em precision} and {\em recall} metrics:
+\begin{itemize}
+\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
+Gives a notion of the {\em soundness} of the learned models.
+\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
+Gives a notion of the {\em completeness} of the learned models.
+\end{itemize}
+
+Table~\ref{fig:observationsnomap} summarizes the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
+
+When the learning hypothesis space is under constrained, the learned actions can be reformulated and still be compliant with the inputs.
+Given the syntax-based nature of the {\em precision} and {\em recall} metrics, they may report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
+a learned model semantically equivalent but syntactically different to the reference model.
+To address this issue and give insight to the quality of the learned models we define a more robust method for computing {\em Precision} and {\em Recall} with regard to a reference model.
+
+Precision and recall are often combined using the {\em harmonic mean}.
+This expression is called the {\em F-measure} (or the balanced {\em F-score}) and is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
+Given a reference \strips\ action model $\Xi^*$ and the learned \strips\ action model $\Xi$ we define the bijective function $f_{P\&R}:\Xi \mapsto \Xi^*$ such that $f_{P\&R}$ maximizes the accumulated {\em F-measure}.
+Thus each learned action scheme $\xi\in\Xi$ is compared with the scheme $f_{P\&R}(\xi)\in \Xi^*$ indicated by the $f_{P\&R}$ mapping.
+Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing with the reference scheme indicated by the $f_{P\&R}$ mapping.
+This shows that in many domains, (almost all of them except the {\em ferry} and {\em satellite} domains) the learned actions models, or their parameters, were learned properly but interchanged their roles with respect to the reference model.
+
+\subsection{Evaluating with a test set}
+When a reference action model is not available, then the evaluation of the quality of the learning process can be done with respect to a test set.
+The difficulty here comes from the fact that how the observations of the testing set are generated strongly affect to the reported results.
+This is not trivial when learning planing action models since the generation of relevant and random observations of planning domains is an open issue.
+Planning domains are highly structured which makes that some states are likely to have low probability of being chosen by chance~\cite{fern2004learning}.
+
+Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an observations test set.
+The table reports for each domain, the {\em observation edit distance}computed with our extended compilation, the {\em maximum edit distance} and their ratio.
+
+Interestingly the Table~\ref{fig:observationstest} shows that using a test set results in reporting higher learning scores.
+The explanation for this is that how the observations are generated affect to the reported results.
+For instance in the floortile domain the learned action missed learning the action scheme for the {\tt paint-down} action because this action was not considered in the state observations of the learning test.
+If this same action is not used either to generate the observations of the test set it means that the learned action model will not be penalized for not learning this action scheme.
+
 \begin{table}
 		\begin{center}
-                \begin{scriptsize}
-			\begin{tabular}{l|l|l|l|}
-				& $\delta(\Xi,\mathcal{O})$ & $\delta(\Xi,*)$ & $P(\mathcal{O}|\Xi)$ \\
+                \begin{footnotesize}
+			\begin{tabular}{l|r|r|c|}
+				& $\delta(\Xi,\mathcal{O})$ & $\delta(\Xi,*)$ & $1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$ \\
 				\hline
-				block & 0 & 90 & 1.0 \\
+				blocks & 0 & 90 & 1.0 \\
 				driverlog & 5 & 144 & 0.97 \\
@@
 				visitall & 2 & 24 & 0.92 \\
-				zenotravel & 3 & 63 & 0.95 \\
-				\hline
-				\bf  & - & - & - \\
+				zenotravel & 3 & 63 & 0.95
 			\end{tabular}
-                        	\end{scriptsize}
+                        	\end{footnotesize}
 		\end{center}
@@
 As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains.
-~\cite{asai2017classical} reported the learning of PDDL action models on a few domains from large images datasets.
 Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
+~\citeauthor{asai2017classical}~\citeyear{asai2017classical} proposed an approach for learning PDDL action models on a few domains from large images datasets (thousands of observations per domain).
 
-%The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
-In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
-
-The size of the compiled classical planning instances depends on the number of input examples.
-The larger the number of examples is, the larger the compilation (a larger number of "test" fluents and "validate" actions), and also the more actions the solution plan requires to derive the action model.
-All this implies larger planning times.
-On the other hand, the precision and recall values depend on the diversity of the training samples rather than on the size of the input data.
-
-Generating {\em informative} examples for learning planning action models is an open issue.
+The empirical results show that since our approach is strongly based on inference can generate good quaility models from very small data sets (ten out of fifteen domains achieved both {\em Precision} and {\em Recall} values over 0.75 learning only from 25 observations).
+Generating {\em informative} examples for learning planning action models is however an open issue.
 Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
@@
 The combination of such exploration techniques with our learning approach is an intriguing research direction towards the bootstrapping of planning action models.
-
-%Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10}.
-For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true {\em at least at one} of the reached states.
-Despite this is beyond the scope of this paper, LTL constraints could be included in our compilation following the ideas for compiling temporally extended goals into classical planning~\cite{baier2006planning} that (1) transform the given LTL formula into an equivalent automata, (2) compute the cross product of this automata with the given classical planning task and (3) force the solution plans to always leave the LTL automata at an acceptor state by adding new goals to the classical planning task.
 
@@
 %% The file named.bst is a bibliography style file for BibTeX 0.99c
+\newpage
 \bibliographystyle{named}

### 2018-01-25 doc/ijcai-18/observations-learning.tex
section 4 of ijcai18 paper updated

@@
 \section{Evaluation of \strips\ action models}
-How well a \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$ can be assessed according to the amount of {\em edition} required by $\Xi$ to produce $\mathcal{O}$ (in the extreme, if $\Xi$ perfectly explains $\mathcal{O}$, then no {\em edition} of the model is necessary).
+We assess how well a \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} that is required by $\Xi$ to produce $\mathcal{O}$.
+In the extreme, if $\Xi$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
 
@@
 We can now formalize an edit distance that quantifies how dissimilar two given \strips\ action models are.
-This distance is symmetric and satisfies the {\em metric axioms} provided that the two {\em \strips\ edit operations} have the same positive cost.
+The distance is symmetric and meets the {\em metric axioms} provided that the two {\em \strips\ edit operations} have the same positive cost.
 
@@
 \begin{definition}
-Let $\Xi$ be an action model built from the set of possible elements $F_v$.
-The \textbf{maximum edit distance} is $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$ and sets an upperbound on the edit distance from $\Xi$ to any other \strips\ action model definable within $F_v$.
+The \textbf{maximum edit distance} of an \strips\ action model $\Xi$ built from the set of possible elements $F_v$ is $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$ and sets an upperbound on the distance from $\Xi$ to any other \strips\ action model definable within $F_v$.
 \end{definition}
@@
 Given an action model $\Xi$ built from the set of possible elements $F_v$ and a sequence of observations $\mathcal{O}$.
-The {\bf observation edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimal edit distance from $\Xi$ to a model $\Xi'$ such that: (1) $\Xi'$ is also definable within $F_v$ and (2), $\Xi'$ can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$, the given observations sequence; i.e., \[\delta(\Xi,\mathcal{O})=\min_{\forall \Xi' \rightarrow \mathcal{O}} \delta(\Xi,\Xi')\]
+The {\bf observation edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimal edit distance from $\Xi$ to any model $\Xi'$ such that: (1) $\Xi'$ is also definable within $F_v$ and (2), $\Xi'$ can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$, the given observations sequence; i.e., \[\delta(\Xi,\mathcal{O})=\min_{\forall \Xi' \rightarrow \mathcal{O}} \delta(\Xi,\Xi')\]
 \end{definition}
 
-Because its semantic nature, the {\em observation edit distance} is robust to learning episodes where actions are reformulated and still compliant with the inputs (e.g.
-the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa or the roles of the action parameters could be interchanged).
+Because its semantic nature, the {\em observation edit distance} is robust to learning episodes where actions are {\em reformulated} and still compliant with the learning inputs.
+For instance a learning episode can interchange the roles of operators with matching headers, or the roles of action parameters with matching types (e.g.
+the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
 
-Furthermore the {\em observation edit distance} enables the recognition of \strips\ action models.
+Furthermore, the {\em observation edit distance} enables the recognition of \strips\ action models.
 The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
@@
 \subsection{Computing the observation edit distance}
-Our compilation can be extended to compute the {\em observations edit distance}.
-In this case the tuple $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ is not representing a learning task but the task of editing the \strips\ action model $\Xi$ to produce the observations $\mathcal{O}$.
-The extended compilation outputs a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$ where:
+Our compilation is extensible to compute the {\em observations edit distance}. A solution to the planning task resulting from the extended compilation is a sequence of actions that:
+\begin{enumerate}
+\item {\bf Edits the \strips\ action model $\Xi$}. A solution plan starts with a {\em prefix} that edits $\Xi$.
+\item {\bf Validates the edited model $\Xi$ in $\mathcal{O}$}. The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$.
+\end{enumerate}
+
+In this case the tuple $\Lambda=\tup{\Xi,\Psi,\mathcal{O}}$ does not represent a learning task but the task of editing the \strips\ action model $\Xi$ to produce the observations $\mathcal{O}$.
+The extended compilation outputs a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
 \begin{itemize}
@@
 \item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
-The Actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation {\tt\small $1\leq i\leq n$} are defined exactly as in the previous compilation.
-The only difference is the actions for {\em programming} operator schema $\xi\in\Xi$ that include now actions for adding a precondition and removing a {\em negative} or {\em positive} effect.
-They now implement the two {\em edit operations}.
+The actions for {\em applying} an already programmed operator scheme and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
+The only difference here is the actions for {\em programming} operator schemes $\xi\in\Xi$ that now implement the two {\em edit operations} (i.e.
+include actions for adding a precondition and for removing a negative/positive effect).
 \end{itemize}
 
-A solution to the classical planning task resulting from the extended compilation is a sequence of actions that edits the \strips\ action model $\Xi$ and validates the edited model on the given observations $\mathcal{O}$.
-Since we are computing the {\em observations edit distance}, we are not interested in the edited model but in the number of {\em edit operations} required to produce a model validable in the given observations.
-The {\em observation edit distance} is exact if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions), and is approximate if $P_{\Lambda}'$ is solved with a satisfying planner or furthermore, if what is solved is not $P_{\Lambda}'$ but its relaxation, e.g.
+Since we are computing the {\em observations edit distance}, we are not interested in the resulting edited model but in the number of {\em edit operations} required to produce a model validable in the given observations.
+The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions), and is approximate if $P_{\Lambda}'$ is solved with a satisfying planner or furthermore, if what is solved is not $P_{\Lambda}'$ but its relaxation, e.g.
 the {\em delete relaxation}~\cite{bonet2001planning}.

### 2018-01-25 doc/ijcai-18/observations-learning.tex
updated section 5 ijcai18 paper

@@
 
+\begin{table}
+		\begin{center}
+                \begin{footnotesize}
+			\begin{tabular}{l|r|r|c|}
+				& $\delta(\Xi,\mathcal{O})$ & $\delta(\Xi,*)$ & $1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$ \\
+				\hline
+				blocks & 0 & 90 & 1.0 \\
+				driverlog & 5 & 144 & 0.97 \\
+				ferry & 2 & 69 & 0.97 \\
+				floortile & 34 & 342 & 0.90 \\
+				grid & 42 & 153 & 0.73 \\
+				gripper & 2 & 30 & 0.93 \\
+				hanoi & 1 & 63 & 0.98 \\
+				hiking & 69 & 174 & 0.60 \\
+				miconic & 3 & 72 & 0.96 \\
+				npuzzle & 2 & 24 & 0.92 \\
+				parking & 5 & 111 & 0.95 \\
+				satellite & 24 & 75 & 0.68 \\
+				transport & 4 & 78 & 0.95 \\
+				visitall & 2 & 24 & 0.92 \\
+				zenotravel & 3 & 63 & 0.95
+			\end{tabular}
+                        	\end{footnotesize}
+		\end{center}
+	\caption{\small Evaluation of the quality of the learned models with respect to an observations test set.}
+	\label{fig:observationstest}                
+\end{table}
+
 \section{Evaluation}
-This section presents a two-fold evaluation of the learned models: with respect to a reference model and with respect to given sets of observations.
+This section reports a two-fold evaluation of the learned models: (1) with respect to a given observation test set and (2), with respect to a reference model.
 
@@
 For the learning of the \strips\ action models we only used observations sequences $\tup{s_0, s_1, \ldots, s_{24}}$ of twenty five states per domain.
-These sets of learning examples are fixed for all the experiments so the results reported by the different evaluation approaches is comparable.
+The set of learning examples is fixed for all the experiments so the results reported by the different evaluation approaches is comparable.
 All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
@@
 
-The compilation source code, the evaluation scripts and the benchmarks (including learning and testing sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
+The compilation source code, the evaluation scripts and the benchmarks (including the learning and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
+
+\subsubsection{Evaluating with a test set}
+When a reference model is not available, the evaluation of the learning process can be done with respect to a test set.
+Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an observation test set.
+The table reports, for each domain, the {\em observation edit distance} (which is computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
+
+The reported results showed that, despite we are learning only from 25 state observations, twelve out of fifteen learned domains achieved ratios of $90\%$ or above.
+This fact evidences that the learned models required small edition amounts to match the observations given in the test set.
+
+\subsubsection{Evaluating with a reference model}
+Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning classical planning models from the IPC.
+For each domain, the learned model is compared with the actual model and its quality is quantified with the {\em precision} and {\em recall} metrics:
+\begin{itemize}
+\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates in the learned model that should not appear).
+Gives a notion of {\em soundness}.
+\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing). Gives a notion of {\em completeness}.
+\end{itemize}
+
+Table~\ref{fig:observationsnomap} summarizes the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
+
+The learning scores reported in Table~\ref{fig:observationsnomap} are poorer than in Table~\ref{fig:observationstest}.
+{\em Precision} and {\em recall},  given its syntax nature, report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model (changes of the roles of actions or their parameters).
 
@@
 
-\subsection{Evaluating with a reference model}
-Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning classical planning models from the IPC.
-For each domain, the learned model is compared with the actual model and its quality is quantified with the {\em precision} and {\em recall} metrics:
-\begin{itemize}
-\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
-Gives a notion of the {\em soundness} of the learned models.
-\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
-Gives a notion of the {\em completeness} of the learned models.
-\end{itemize}
-
-Table~\ref{fig:observationsnomap} summarizes the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
-
-When the learning hypothesis space is under constrained, the learned actions can be reformulated and still be compliant with the inputs.
-Given the syntax-based nature of the {\em precision} and {\em recall} metrics, they may report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
-a learned model semantically equivalent but syntactically different to the reference model.
-To address this issue and give insight to the quality of the learned models we define a more robust method for computing {\em Precision} and {\em Recall} with regard to a reference model.
-
+\subsubsection{Precision and recall robust to model reformulations}
+To give insight to the actual quality of the learned models we define a method for computing {\em Precision} and {\em Recall} that is robust to such model reformulations.
 Precision and recall are often combined using the {\em harmonic mean}.
 This expression is called the {\em F-measure} (or the balanced {\em F-score}) and is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
-Given a reference \strips\ action model $\Xi^*$ and the learned \strips\ action model $\Xi$ we define the bijective function $f_{P\&R}:\Xi \mapsto \Xi^*$ such that $f_{P\&R}$ maximizes the accumulated {\em F-measure}.
-Thus each learned action scheme $\xi\in\Xi$ is compared with the scheme $f_{P\&R}(\xi)\in \Xi^*$ indicated by the $f_{P\&R}$ mapping.
-Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing with the reference scheme indicated by the $f_{P\&R}$ mapping.
-This shows that in many domains, (almost all of them except the {\em ferry} and {\em satellite} domains) the learned actions models, or their parameters, were learned properly but interchanged their roles with respect to the reference model.
+Given the learned \strips\ action model $\Xi$ and the reference \strips\ action model $\Xi^*$, the bijective function $f_{P\&R}:\Xi \mapsto \Xi^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure}.
 
-\subsection{Evaluating with a test set}
-When a reference action model is not available, then the evaluation of the quality of the learning process can be done with respect to a test set.
-The difficulty here comes from the fact that how the observations of the testing set are generated strongly affect to the reported results.
-This is not trivial when learning planing action models since the generation of relevant and random observations of planning domains is an open issue.
-Planning domains are highly structured which makes that some states are likely to have low probability of being chosen by chance~\cite{fern2004learning}.
+Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing each learned action scheme $\xi\in\Xi$ with the corresponding reference scheme $f_{P\&R}(\xi)\in \Xi^*$ given by the $f_{P\&R}$ mapping.
+These results show that in all of the evaluated domains, except {\em ferry} and {\em satellite}, learning interchanged the roles of actions (or their parameters) with respect to their role in the reference model.
 
-Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an observations test set.
-The table reports for each domain, the {\em observation edit distance}computed with our extended compilation, the {\em maximum edit distance} and their ratio.
-
-Interestingly the Table~\ref{fig:observationstest} shows that using a test set results in reporting higher learning scores.
-The explanation for this is that how the observations are generated affect to the reported results.
-For instance in the floortile domain the learned action missed learning the action scheme for the {\tt paint-down} action because this action was not considered in the state observations of the learning test.
-If this same action is not used either to generate the observations of the test set it means that the learned action model will not be penalized for not learning this action scheme.
-
-\begin{table}
-		\begin{center}
-                \begin{footnotesize}
-			\begin{tabular}{l|r|r|c|}
-				& $\delta(\Xi,\mathcal{O})$ & $\delta(\Xi,*)$ & $1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$ \\
-				\hline
-				blocks & 0 & 90 & 1.0 \\
-				driverlog & 5 & 144 & 0.97 \\
-				ferry & 2 & 69 & 0.97 \\
-				floortile & 34 & 342 & 0.90 \\
-				grid & 42 & 153 & 0.73 \\
-				gripper & 2 & 30 & 0.93 \\
-				hanoi & 1 & 63 & 0.98 \\
-				hiking & 69 & 174 & 0.60 \\
-				miconic & 3 & 72 & 0.96 \\
-				npuzzle & 2 & 24 & 0.92 \\
-				parking & 5 & 111 & 0.95 \\
-				satellite & 24 & 75 & 0.68 \\
-				transport & 4 & 78 & 0.95 \\
-				visitall & 2 & 24 & 0.92 \\
-				zenotravel & 3 & 63 & 0.95
-			\end{tabular}
-                        	\end{footnotesize}
-		\end{center}
-	\caption{\small Evaluation of the quality of the learned models with respect to an observations test set.}
-	\label{fig:observationstest}                
-\end{table}
+However in few domains, Table~\ref{fig:observationsmap} still reports lower scores than Table~\ref{fig:observationstest}.
+The explanation is that, how the observations of test sets are generated, affect to the results reported on Table~\ref{fig:observationstest}.
+For instance, in the {\em driverlog} domain the learned model missed learning the action scheme for the {\tt disembark-truck} action because this action was not considered in the learning set, the same happened in the {\em floortile} with the {\tt paint-down} action or in the {\em parking} domain with the {\tt move-curb-to-curb} action.
+We realized that these actions do not appear either in the corresponding test sets so the learned action models were not be penalized for not learning these action schemes.
+The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
 

### 2018-01-25 doc/ijcai-18/observations-learning.tex
updated sections 5 and 6 ijcai18 paper

@@
 Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an observation test set.
-The table reports, for each domain, the {\em observation edit distance} (which is computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
+The table reports, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
 
 The reported results showed that, despite we are learning only from 25 state observations, twelve out of fifteen learned domains achieved ratios of $90\%$ or above.
-This fact evidences that the learned models required small edition amounts to match the observations given in the test set.
+This fact evidences that the learned models required small edition amounts to match the state observations given in the test set.
 
@@
 The learning scores reported in Table~\ref{fig:observationsnomap} are poorer than in Table~\ref{fig:observationstest}.
-{\em Precision} and {\em recall},  given its syntax nature, report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model (changes of the roles of actions or their parameters).
+{\em Precision} and {\em recall},  given its syntax nature, report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model (changes in the roles of the actions or their parameters).
 
@@
 Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing each learned action scheme $\xi\in\Xi$ with the corresponding reference scheme $f_{P\&R}(\xi)\in \Xi^*$ given by the $f_{P\&R}$ mapping.
-These results show that in all of the evaluated domains, except {\em ferry} and {\em satellite}, learning interchanged the roles of actions (or their parameters) with respect to their role in the reference model.
+These results evidenced that in all of the evaluated domains, except {\em ferry} and {\em satellite}, learning interchanged the roles of actions or parameters with respect to their role in the reference model.
 
+We can claim now that the action model for the {\em blocksworld} and the {\em gripper} domains was perfectly learned from only 25 state observations.
 However in few domains, Table~\ref{fig:observationsmap} still reports lower scores than Table~\ref{fig:observationstest}.
 The explanation is that, how the observations of test sets are generated, affect to the results reported on Table~\ref{fig:observationstest}.
-For instance, in the {\em driverlog} domain the learned model missed learning the action scheme for the {\tt disembark-truck} action because this action was not considered in the learning set, the same happened in the {\em floortile} with the {\tt paint-down} action or in the {\em parking} domain with the {\tt move-curb-to-curb} action.
-We realized that these actions do not appear either in the corresponding test sets so the learned action models were not be penalized for not learning these action schemes.
-The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
+For instance, in the {\em driverlog} domain the learned model missed learning the action scheme for the {\tt disembark-truck} action because this action was not considered in the learning set.
+The same happened in the {\em floortile} with the {\tt paint-down} action and in {\em parking} with the {\tt move-curb-to-curb} action.
+We realized that these actions do not appear either in the corresponding test sets so the learned action models were not be penalized in the scores of Table~\ref{fig:observationstest} for not learning these action schemes.
+%The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
 
 \section{Conclusions}
+Unlike extensive-data ML approaches (like~\citeauthor{asai2017classical}~\citeyear{asai2017classical}), our work explores an alternative research direction to learn sound models from small amounts of state observations.
 As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains.
 Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
-~\citeauthor{asai2017classical}~\citeyear{asai2017classical} proposed an approach for learning PDDL action models on a few domains from large images datasets (thousands of observations per domain).
 
-The empirical results show that since our approach is strongly based on inference can generate good quaility models from very small data sets (ten out of fifteen domains achieved both {\em Precision} and {\em Recall} values over 0.75 learning only from 25 observations).
-Generating {\em informative} examples for learning planning action models is however an open issue.
+The empirical results show that since our approach is strongly based on inference can generate good quality models from very small data sets.
+The action models of the {\em blocksworld} and the {\em gripper} were perfectly learned from only 25 observations besides the learned models, in ten out of fifteen domains, achieved both {\em Precision} and {\em Recall} values over 0.75.
+
+Generating {\em informative} observations for learning planning action models is however an open issue.
 Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.

### 2018-01-26 doc/ijcai-18/observations-learning.tex
Add files via upload
Section 1 included

@@
 \section{Introduction}
-Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-At these planning tasks, automated planners are required to reason about an action model that correctly and completely captures the possible world transitions~\cite{geffner:book:2013}.
-Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of {\em AI planning}~\cite{kambhampati:modellite:AAAI2007}.
 
-The Machine Learning of planning action models is a promising alternative to hand-coding them and nowadays, there exist sophisticated algorithms like {\sc AMAN}~\cite{zhuo2013action}, {\sc ARMS}~\cite{yang2007learning}, {\sc LOCM}~\cite{cresswell2013acquiring} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper presents a novel approach for learning \strips\ action models that introduces the following contributions:
+The automated learning of action models in planning is a promising approach, as an alternative to hand-coding by a human or when the action model is not available.
+Action-model acquisition has been addressed with various algorithms.
+{\sc ARMS}~\cite{yang2007learning} receives  plan traces as training samples and defines the learning task as a set of weighted constraints that must hold for the plans to be correct, which is then solved by a MAX-SAT solver.
+{\sc ARMS} validates the learned model with a test set of example plans of six domains from the IPC %checking missing preconditions and redundant effects.
+The learning algorithm of {\sc SLAF}~\cite{amir:alearning:JAIR08} generates a CNF formula which is consistent with the input plan traces and partially observed states, and the quality of the learned models is assessed with respect to the true (reference) model.
+In the case of {\sc LOCM}~\cite{cresswell2013acquiring}, the learning samples are only the examples plans, without providing information about predicates or states.
+{\sc LOCM} relies on assumptions on the kind of domain structure of the model and it was able to correctly discover the dynamics of the domain in 12 out of 20 tested domains by comparison with a reference model \cite{GregoryC16}.
+Finally, {\sc AMAN}~\cite{zhuo2013action} is able to work with incorrectly observed plan traces and the accuracy of the learned models is validated in three IPC domains with respect to the reference model.
 
-\begin{enumerate}
-\item {\em Learning action models as planning}.
-The practicality of the compilation approach allow us to report results over a wide range of planning domains.
-Furthermore, opens up a way towards the {\em bootstrapping} of planning action models (a planner exploring its state space and using the obtained data to learn/update its action model).
-\item {\em Learning from state observations}.
-Our approach does not require the learning samples to be observations of actions but simply state observations, broadening the range of application to external observers.
-\item {\em Model evaluation}. The compilation is extensible to assess how well a \strips\ action model matches a given set of observations, which enables {\em model recognition}. 
-\end{enumerate}
+The learning samples required by all the aforementioned approaches are observations of plan executions in the form of plan traces.
+A more recent work proposes acquiring symbolic action models from unstructured inputs, particularly from images that represent transitions in the environment \cite{AsaiF18}.
+Additionally, all the above approaches but {\sc ARMS} validate the accuracy of the learned model with respect to a reference model.
+
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation approach for learning \strips\ action models.
+The output of the compilation is a classical planning task for which a solution plan is a sequence of actions that determine the precondition/effects of the model operator schemas (the learned model).
+\textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards the \emph{bootstrapping} of planning action models, enabling a planner to gradually learn and update its action model.
+The practicality of the compilation approach allow us to report results over a wide range of 15 IPC planning domains.
+
+As for the type of training samples, we adopt a middle way between unstructured inputs and plan traces, wherein only samples that are state observations are required, without further information about state transitions.
+\textbf{Learning from state observations} is a relevant advancement as in many applications (e.g.
+computer games) the actual actions executed by the agent are not directly observable but instead we can observe the resulting states.
+Handling state observations broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition tasks \cite{SohrabiRU16}.
+
+We propose a two-fold model validation.
+The first method compares the learned model to a reference model, evaluating the syntactic and semantic similarity between both.
+In the second method, the learned model is tested with a set of state observation sequences.
+Since plan traces are not available, we extend the compilation to accept a learned model as input besides the observations.
+This new task consists in learning a new model that induces the observations whilst measuring the \emph{edit distance} to the input model.
+This way, the compilation scheme is also extensible to assess how well a \strips\ action model matches a given set of observations and it can thus be regarded as a first step to apply \textbf{action-model recognition}.
+
+This paper is organized as follows ...
 

### 2018-01-26 doc/ijcai-18/observations-learning.tex
ijcai18 paper updated to replace figures
ijcai18 paper updated with friday meeting comments

@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation approach for learning \strips\ action models.
-The output of the compilation is a classical planning task for which a solution plan is a sequence of actions that determine the precondition/effects of the model operator schemas (the learned model).
+The output of the compilation is a classical planning task for which a solution plan is a sequence of actions that determine the precondition/effects of the model operator schemes (the learned model).
 \textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards the \emph{bootstrapping} of planning action models, enabling a planner to gradually learn and update its action model.
@@
 
-This paper is organized as follows ...
-
 \section{Background}
@@
 
-\subsubsection{Classical planning with conditional effects}
+\subsection{Classical planning with conditional effects}
 We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
@@
 
+\begin{figure}
+\begin{footnotesize}
+\begin{verbatim}
+(:action stack
+ :parameters (?v1 ?v2 - object)
+ :precondition (and (holding ?v1) (clear ?v2))
+ :effect (and (not (holding ?v1))
+              (not (clear ?v2))
+              (handempty) (clear ?v1)
+              (on ?v1 ?v2)))
+\end{verbatim}
+\end{footnotesize}
+ \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
+\label{fig:stack}
+\end{figure}
+
 \subsection{\strips\ action schemes}
@@
 
-\begin{figure}
-\begin{footnotesize}
-\begin{verbatim}
-(:action stack
- :parameters (?v1 ?v2 - object)
- :precondition (and (holding ?v1) (clear ?v2))
- :effect (and (not (holding ?v1))
-              (not (clear ?v2))
-              (handempty) (clear ?v1)
-              (on ?v1 ?v2)))
-\end{verbatim}
-\end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
-\label{fig:stack}
-\end{figure}
-
 \section{Learning \strips\ action models}
-Learning \strips\ action models from plans where every action in the plan is available as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
+%Learning \strips\ action models from plans where every action in the plan is available, as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
 In this case, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
@@
 
-This paper addresses a more challenging learning task, where less input knowledge is available.
-The addressed learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
-This learning task is defined as $\Lambda=\tup{\Xi,\Psi,\mathcal{O}}$, where:
+This paper addresses the learning task that corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
+This learning task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$:
 \begin{itemize}
-\item $\Xi$ is the set of empty operator schemes, wherein each $\xi\in\Xi$ is only composed of $head(\xi)$.
+\item $\mathcal{M}$ is the set of empty operator schemes, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
 \item $\Psi$ is the set of predicates, that define the abstract state space of a given classical planning frame.
@@
 
-A solution to $\Lambda$ is a set of operator schema $\Xi'$ compliant with the headers $\Xi$, the predicates $\Psi$, and the given sequence of state observations $\mathcal{O}$.
-A planning compilation is a suitable approach for addressing a $\Lambda$ learning task because a solution must not only determine the \strips\ action model $\Xi'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains $\mathcal{O}$.
-Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning a \strips\ action model in the blocksworld from the sequence of five state observations that corresponds to inverting a 2-blocks tower.
+A solution to $\Lambda$ is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the given sequence of state observations $\mathcal{O}$.
+A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains $\mathcal{O}$.
+Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning a \strips\ action model in the blocksworld from the five-state observations sequence that corresponds to inverting a 2-block tower.
+
+\subsection{Learning with classical planning}
+
+Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
+The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
+
+\begin{enumerate}
+\item {\bf Programs the \strips\ action model $\mathcal{M}'$}.
+A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item {\bf Validates the \strips\ action model $\mathcal{M}'$ in $\mathcal{O}$}.
+The solution plan continues with a postfix that produces the given sequence of states $\tup{s_0,s_1,\ldots,s_{n}}$ using the programmed action model $\mathcal{M}'$.
+\end{enumerate}
 
 \begin{figure}
-{\footnotesize\tt ;;;;;; Headers in $\Xi$}
+{\footnotesize\tt ;;;;;; Headers in $\mathcal{M}$}
 \begin{footnotesize}
@@
 
-\subsection{Learning with classical planning}
-
-Our approach for addressing the learning task is compiling $\Lambda$ into a classical planning task $P_{\Lambda}$ with conditional effects.
-The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
-
-\begin{enumerate}
-\item {\bf Programs the \strips\ action model $\Xi'$}.
-A solution plan starts with a {\em prefix} that, for each $\xi\in\Xi$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item {\bf Validates the \strips\ action model $\Xi'$ in $\mathcal{O}$}.
-The solution plan continues with a postfix that produces the given sequence of states $\tup{s_0,s_1,\ldots,s_{n}}$ using the programmed action model $\Xi'$.
-\end{enumerate}
-
-Given a learning task $\Lambda=\tup{\Xi,\Psi,\mathcal{O}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
 \begin{itemize}
@@
 \item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi'$.
-For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) is represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
-\item Fluent $mode_{prog}$ indicating whether the operator schemes are programmed or validated (already programmed) and fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
+For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
+\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemes are programmed or validated and fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
 \end{itemize}
@@
 \begin{enumerate}
-\item Actions for {\em programming} operator schema $\xi\in\Xi$:
+\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$:
 \begin{itemize}
-\item Actions for {\bf removing} a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\Xi$.
+\item Actions for {\bf removing} a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.
 
@@
 
-\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\Xi$.
+\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
 
@@
 
-\item Actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
+\item Actions for {\em applying} an already programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
 Given that the operators headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
-Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
+&\cup \{\neg mode_{val}\},\\
 \cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
 &\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
+&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
+&\{\emptyset\}\rhd\{mode_{val}\}.
 \end{align*}
@@
 \hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\\
-&\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i\}.
+&\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
+\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
 \end{align*}
@@
 
-\begin{figure}[hbt!]
-\begin{scriptsize}
-\begin{verbatim}
-(:action apply_stack
-  :parameters (?o1 - object ?o2 - object)
-  :precondition
-   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
-        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
-        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
-        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
-        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
-        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
-        (or (not (pre_clear_stack_v1)) (clear ?o1))
-        (or (not (pre_clear_stack_v2)) (clear ?o2))
-        (or (not (pre_holding_stack_v1)) (holding ?o1))
-        (or (not (pre_holding_stack_v2)) (holding ?o2))
-        (or (not (pre_handempty_stack)) (handempty)))
-  :effect
-   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
-        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
-        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
-        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
-        (when (del_ontable_stack_v1) (not (ontable ?o1)))
-        (when (del_ontable_stack_v2) (not (ontable ?o2)))
-        (when (del_clear_stack_v1) (not (clear ?o1)))
-        (when (del_clear_stack_v2) (not (clear ?o2)))
-        (when (del_holding_stack_v1) (not (holding ?o1)))
-        (when (del_holding_stack_v2) (not (holding ?o2)))
-        (when (del_handempty_stack) (not (handempty)))
-        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
-        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
-        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
-        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
-        (when (add_ontable_stack_v1) (ontable ?o1))
-        (when (add_ontable_stack_v2) (ontable ?o2))
-        (when (add_clear_stack_v1) (clear ?o1))
-        (when (add_clear_stack_v2) (clear ?o2))
-        (when (add_holding_stack_v1) (holding ?o1))
-        (when (add_holding_stack_v2) (holding ?o2))
-        (when (add_handempty_stack) (handempty))
-        (when (modeProg) (not (modeProg)))))
-\end{verbatim}
-\end{scriptsize}
- \caption{\small Action for applying an already programmed schema $stack$ as encoded in PDDL (implications are coded as disjunctions).}
-\label{fig:compilation}
-\end{figure}
-
 \subsection{Compilation properties}
@@
 \begin{lemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi'$ that solves the $\Lambda$ learning task.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves the $\Lambda$ learning task.
 \end{lemma}
@@
 \begin{small}
-The compilation forces that once the preconditions of an operator schema $\xi \in \Xi'$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects.
-Furthermore, once operator schemes $\Xi'$ are programmed, they can only be applied because of the $mode_{prog}$ fluent.
-Finally, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemes that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence {\small $1\leq i\leq n$} which means that the programmed action model $\Xi'$ complies with the provided observations and hence, solves $\Lambda$.
+Once operator schemes $\mathcal{M}'$ are programmed, they can only be applied because of the $mode_{prog}$ fluent.
+In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
+These goals can only be achieved executing an applicable sequence of programmed operator schemes that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence {\small $1\leq i\leq n$}.
+This means that the programmed action model $\mathcal{M}'$ complies with the provided observations $\mathcal{O}$ and hence, solves $\Lambda$.
 \end{small}
@@
 \begin{lemma}
-Completeness. Any \strips\ action model $\Xi'$ that solves a $\Lambda=\tup{\Psi,\Xi,\mathcal{O}}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+Completeness.
+Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
 \end{lemma}
@@
 The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfy the state trajectory constraints given by the $\mathcal{O}$ sequence.
-By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\Xi$ given its header and the set of predicates $\Psi$.
+By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
 \end{small}
@@
 
-The size of the compiled classical planning instances depends on the number of input examples.
-The larger the number of examples, the larger the compilation (a larger number of "test" fluents and "validate" actions), and also the more actions the solution plan requires to derive the action model.
+The size of $P_{\Lambda}$ depends on the number of given state observations, that is $|\mathcal{O}|$. The larger this number, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions.
 
 \section{Evaluation of \strips\ action models}
-We assess how well a \strips\ action model $\Xi$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} that is required by $\Xi$ to produce $\mathcal{O}$.
-In the extreme, if $\Xi$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
+We assess how well a \strips\ action model $\mathcal{M}$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} that is required by $\mathcal{M}$ to produce $\mathcal{O}$.
+%In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
 
@@
 \begin{itemize}
-\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, such that $f\in F_v(\xi)$, is removed from the operator schema $\xi\in\Xi$.
-\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, such that $f\in F_v(\xi)$, is added to the operator schema $\xi\in\Xi$.
+\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, $f\in F_v(\xi)$.
+\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, $f\in F_v(\xi)$.
 \end{itemize}
 
-We can now formalize an edit distance that quantifies how dissimilar two given \strips\ action models are.
+Now, we can formalize an edit distance that quantifies how dissimilar two given \strips\ action models are.
 The distance is symmetric and meets the {\em metric axioms} provided that the two {\em \strips\ edit operations} have the same positive cost.
@@
 \begin{definition}
-Let $\Xi$ and $\Xi'$ be two \strips\ action models, both built from the same set of possible elements $F_v$.
-The {\bf edit distance}, denoted as $\delta(\Xi,\Xi')$, is the minimum number of {\em edit operations} required to transform $\Xi$ into $\Xi'$.
+Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, both built from the same set of possible elements $F_v$.
+The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} required to transform $\mathcal{M}$ into $\mathcal{M}'$.
 \end{definition}
@@
 Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well.
-In more detail, for an operator schema $\xi\in\Xi$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects are two times $|F_v(\xi)|$.
+In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects are two times $|F_v(\xi)|$.
 
 \begin{definition}
-The \textbf{maximum edit distance} of an \strips\ action model $\Xi$ built from the set of possible elements $F_v$ is $\delta(\Xi,*)=\sum_{\xi\in\Xi} 3|F_v(\xi)|$ and sets an upperbound on the distance from $\Xi$ to any other \strips\ action model definable within $F_v$.
+The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3|F_v(\xi)|$ and sets an upperbound on the distance from $\mathcal{M}$ to any \strips\ model definable within $F_v$.
 \end{definition}
@@
 \begin{definition}
-Given an action model $\Xi$ built from the set of possible elements $F_v$ and a sequence of observations $\mathcal{O}$.
-The {\bf observation edit distance}, denoted by  $\delta(\Xi,\mathcal{O})$, is the minimal edit distance from $\Xi$ to any model $\Xi'$ such that: (1) $\Xi'$ is also definable within $F_v$ and (2), $\Xi'$ can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$, the given observations sequence; i.e., \[\delta(\Xi,\mathcal{O})=\min_{\forall \Xi' \rightarrow \mathcal{O}} \delta(\Xi,\Xi')\]
+Given an action model $\mathcal{M}$ built from the set of possible elements $F_v$ and a sequence of observations $\mathcal{O}$.
+The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$ such that: (1) $\mathcal{M}'$ is also definable within $F_v$ and (2), $\mathcal{M}'$ can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
 
-Because its semantic nature, the {\em observation edit distance} is robust to learning episodes where actions are {\em reformulated} and still compliant with the learning inputs.
+Because its semantic nature, the {\em observation edit distance} is robust to evaluate learning episodes where actions are {\em reformulated} and still compliant with the learning inputs.
 For instance a learning episode can interchange the roles of operators with matching headers, or the roles of action parameters with matching types (e.g.
@@
 The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
-We can map the {\em observation edit distance} into a likelihood with the following expression $P(\mathcal{O}|\Xi)=1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$.
+The {\em observation edit distance} can be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
 The larger the {\em observation edit distance} the lower this likelihood.
@@
 In our scenario, the hypotheses are about the possible \strips\ action models that can be built within a given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets).
-With this regard, $P(\Xi|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
+With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
 %\begin{enumerate}
-%\item Computing the {\em observation edit distance} $\delta(\Xi,\mathcal{O})$ for every possible model $\Xi$.
-%\item Applying the resulting distances to the above $P(\mathcal{O}|\Xi)$ formula to map these distances into likelihoods
+%\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$.
+%\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
 %\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
 %\end{enumerate}
+
+\begin{figure}
+{\tt\small
+00 : (program\_add\_handempty\_stack)\\
+01 : (program\_add\_clear\_stack\_var1)\\
+02 : (apply\_unstack block2 block1)\\
+03 : (validate\_1)\\
+04 : (apply\_putdown block2)\\
+05 : (validate\_2)\\
+06 : (apply\_pickup block1)\\
+07 : (validate\_3)\\
+08 : (apply\_stack block1 block2)\\
+09 : (validate\_4)\\     
+}
+ \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:lexample}.}
+\label{fig:plan}
+\end{figure}
 
@@
 \begin{enumerate}
-\item {\bf Edits the \strips\ action model $\Xi$}. A solution plan starts with a {\em prefix} that edits $\Xi$.
-\item {\bf Validates the edited model $\Xi$ in $\mathcal{O}$}. The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$.
+\item {\bf Edits the \strips\ action model $\mathcal{M}$}. A solution plan starts with a {\em prefix} that edits the preconditions and effects of the action schemes in $\mathcal{M}$.
+\item {\bf Validates the edited model $\mathcal{M}$ in $\mathcal{O}$}. The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$.
 \end{enumerate}
 
-In this case the tuple $\Lambda=\tup{\Xi,\Psi,\mathcal{O}}$ does not represent a learning task but the task of editing the \strips\ action model $\Xi$ to produce the observations $\mathcal{O}$.
+In this case the tuple $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ does not represent a learning task but the task of editing the \strips\ action model $\mathcal{M}$ to produce the observations $\mathcal{O}$.
 The extended compilation outputs a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
@@
 \item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
-\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
-In addition, the given \strips\ action model $\Xi$ is now encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, with $f\in F_v(\xi)$, hold in the initial state iff they appear in the given $\Xi$ model.
+\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
+In addition, the given \strips\ action model $\mathcal{M}$ is now encoded in the initial state.
+This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, with $f\in F_v(\xi)$, hold in the initial state iff they appear in the given $\mathcal{M}$ model.
 \item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
 The actions for {\em applying} an already programmed operator scheme and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
-The only difference here is the actions for {\em programming} operator schemes $\xi\in\Xi$ that now implement the two {\em edit operations} (i.e.
+The only difference here is the actions for {\em programming} operator schemes $\xi\in\mathcal{M}$ now implement the two {\em edit operations} (i.e.
 include actions for adding a precondition and for removing a negative/positive effect).
@@
 
+To illustrate this, the classical plan of Figure~\ref{fig:plan} solves the task of editing a blocksworld action model where the {\tt\small stack} scheme misses the positive effects {\tt(handempty)} and {\tt (clear ?v1)}.
+This plan edits the {\tt\small stack} scheme, adding these two positive effects, and then validates the edited model at the five state observations shown in Figure~\ref{fig:lexample}.
+
 Since we are computing the {\em observations edit distance}, we are not interested in the resulting edited model but in the number of {\em edit operations} required to produce a model validable in the given observations.
-The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions), and is approximate if $P_{\Lambda}'$ is solved with a satisfying planner or furthermore, if what is solved is not $P_{\Lambda}'$ but its relaxation, e.g.
-the {\em delete relaxation}~\cite{bonet2001planning}.
+The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions), and is approximate if $P_{\Lambda}'$ is solved with a satisfying planner or furthermore, if what is solved is not $P_{\Lambda}'$ but its relaxation~\cite{bonet2001planning}.
 
@@
 			\begin{tabular}{l|r|r|c|}
-				& $\delta(\Xi,\mathcal{O})$ & $\delta(\Xi,*)$ & $1-\frac{\delta(\Xi,\mathcal{O})}{\delta(\Xi,*)}$ \\
+				& $\delta(\mathcal{M},\mathcal{O})$ & $\delta(\mathcal{M},*)$ & $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$ \\
 				\hline
@@
 \section{Evaluation}
-This section reports a two-fold evaluation of the learned models: (1) with respect to a given observation test set and (2), with respect to a reference model.
+This section evaluates the performance of our approach for learning \strips\ action models from state observations.
 
@@
 {\sc Madagascar} is the classical planner we use to solve the instances that result from our compilations because its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
-In addition, because its SAT-based nature, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
+Because its SAT-based nature, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
 Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
 
-The compilation source code, the evaluation scripts and the benchmarks (including the learning and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
+The compilation source code, evaluation scripts and benchmarks (including the learning and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
 
@@
 When a reference model is not available, the evaluation of the learning process can be done with respect to a test set.
-Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an observation test set.
+Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an state observations test set.
 The table reports, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
 
-The reported results showed that, despite we are learning only from 25 state observations, twelve out of fifteen learned domains achieved ratios of $90\%$ or above.
-This fact evidences that the learned models required small edition amounts to match the state observations given in the test set.
+The reported results showed that, despite learning only from 25 state observations, twelve out of the fifteen learned domains achieved ratios of $90\%$ or above.
+This fact evidences that the learned models required small edition amounts to match the observations given in the test set (each test set comprises between 20 and 50 observations).
 
 \subsubsection{Evaluating with a reference model}
-Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning classical planning models from the IPC.
-For each domain, the learned model is compared with the actual model and its quality is quantified with the {\em precision} and {\em recall} metrics:
+Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning models from the IPC.
+For each domain, the learned model is compared with the actual model and its quality quantified with the {\em precision} and {\em recall} metrics:
 \begin{itemize}
@@
 Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
+The scores are lower than in Table~\ref{fig:observationstest}.
+The explanation is that {\em precision} and {\em recall}, given its syntax nature, report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model (changes in the roles of the actions or their parameters).
 
-The learning scores reported in Table~\ref{fig:observationsnomap} are poorer than in Table~\ref{fig:observationstest}.
-{\em Precision} and {\em recall},  given its syntax nature, report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model (changes in the roles of the actions or their parameters).
+\subsubsection{Precision and recall robust to model reformulations}
+To give insight to the actual quality of the learned models we define a method for computing {\em Precision} and {\em Recall} that is robust to such model reformulations.
+Precision and recall are often combined using the {\em harmonic mean}.
+This expression, called the {\em F-measure} or the balanced {\em F-score}, is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
+Given the learned \strips\ action model $\mathcal{M}$ and the reference \strips\ action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure}.
 
-\begin{table}
+Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing each learned action scheme $\xi\in\mathcal{M}$ with the corresponding reference scheme $f_{P\&R}(\xi)\in \mathcal{M}^*$ given by the $f_{P\&R}$ mapping.
+These results evidenced that in all of the evaluated domains, except {\em ferry} and {\em satellite}, learning interchanged the roles of actions or parameters with respect to their role in the reference model.
+
+We can claim now that the action model for the {\em blocksworld} and the {\em gripper} domains were perfectly learned from only 25 state observations.
+In few domains, Table~\ref{fig:observationsmap} still reports lower scores than Table~\ref{fig:observationstest}.
+The explanation is that, how the observations of test sets are generated, affect to the results reported on Table~\ref{fig:observationstest}.
+For instance, the model learned for the {\em driverlog} missed the action scheme for the {\tt disembark-truck} action because this action was not considered in the learning set.
+The same happened in {\em floortile} with the {\tt paint-down} action and in {\em parking} with the {\tt move-curb-to-curb} action.
+We realized that these actions do not appear either in the corresponding test sets so the learned action models were not be penalized in the scores of Table~\ref{fig:observationstest} for not learning these action schemes.
+%The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
+
+\begin{table}[hbt!]
 	\begin{center}
@@
 \end{table}
-
-\subsubsection{Precision and recall robust to model reformulations}
-To give insight to the actual quality of the learned models we define a method for computing {\em Precision} and {\em Recall} that is robust to such model reformulations.
-Precision and recall are often combined using the {\em harmonic mean}.
-This expression is called the {\em F-measure} (or the balanced {\em F-score}) and is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
-Given the learned \strips\ action model $\Xi$ and the reference \strips\ action model $\Xi^*$, the bijective function $f_{P\&R}:\Xi \mapsto \Xi^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure}.
-
-Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing each learned action scheme $\xi\in\Xi$ with the corresponding reference scheme $f_{P\&R}(\xi)\in \Xi^*$ given by the $f_{P\&R}$ mapping.
-These results evidenced that in all of the evaluated domains, except {\em ferry} and {\em satellite}, learning interchanged the roles of actions or parameters with respect to their role in the reference model.
-
-We can claim now that the action model for the {\em blocksworld} and the {\em gripper} domains was perfectly learned from only 25 state observations.
-However in few domains, Table~\ref{fig:observationsmap} still reports lower scores than Table~\ref{fig:observationstest}.
-The explanation is that, how the observations of test sets are generated, affect to the results reported on Table~\ref{fig:observationstest}.
-For instance, in the {\em driverlog} domain the learned model missed learning the action scheme for the {\tt disembark-truck} action because this action was not considered in the learning set.
-The same happened in the {\em floortile} with the {\tt paint-down} action and in {\em parking} with the {\tt move-curb-to-curb} action.
-We realized that these actions do not appear either in the corresponding test sets so the learned action models were not be penalized in the scores of Table~\ref{fig:observationstest} for not learning these action schemes.
-%The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
-
 \section{Conclusions}
-Unlike extensive-data ML approaches (like~\citeauthor{asai2017classical}~\citeyear{asai2017classical}), our work explores an alternative research direction to learn sound models from small amounts of state observations.
+Unlike extensive-data ML approaches, our work explores an alternative research direction to learn sound models from small amounts of state observations.
 As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains.
@@
 
-The empirical results show that since our approach is strongly based on inference can generate good quality models from very small data sets.
-The action models of the {\em blocksworld} and the {\em gripper} were perfectly learned from only 25 observations besides the learned models, in ten out of fifteen domains, achieved both {\em Precision} and {\em Recall} values over 0.75.
+The empirical results show that since our approach is not statistical by inference-based, it can generate good quality models from very small data sets.
+The action models of the {\em blocksworld} or the {\em gripper} were perfectly learned from only 25 state observations besides, in ten out of fifteen domains, the learned models achieved both {\em Precision} and {\em Recall} values over $0.75$.
 
@@
 %% The file named.bst is a bibliography style file for BibTeX 0.99c
-\newpage
 \bibliographystyle{named}
-\bibliography{observations-learning}
+\bibliography{planlearnbibliography}
 

### 2018-01-27 doc/ijcai-18/observations-learning.tex
Revised paper and bibliography

@@
 {\sc ARMS}~\cite{yang2007learning} receives  plan traces as training samples and defines the learning task as a set of weighted constraints that must hold for the plans to be correct, which is then solved by a MAX-SAT solver.
-{\sc ARMS} validates the learned model with a test set of example plans of six domains from the IPC %checking missing preconditions and redundant effects.
+{\sc ARMS} validates the learned model with a test set of example plans of six domains from the IPC.
+%checking missing preconditions and redundant effects.
 The learning algorithm of {\sc SLAF}~\cite{amir:alearning:JAIR08} generates a CNF formula which is consistent with the input plan traces and partially observed states, and the quality of the learned models is assessed with respect to the true (reference) model.
@@
 
-The learning samples required by all the aforementioned approaches are observations of plan executions in the form of plan traces.
-A more recent work proposes acquiring symbolic action models from unstructured inputs, particularly from images that represent transitions in the environment \cite{AsaiF18}.
-Additionally, all the above approaches but {\sc ARMS} validate the accuracy of the learned model with respect to a reference model.
+The learning samples required by all the aforementioned approaches are observations of plan executions in the form of plan traces, and all of them but {\sc ARMS} validate the accuracy of the learned model with respect to a reference model.
+A more recent work defines the acquisition of a symbolic action model as learning the successor function from unstructured inputs, particularly from images that represent transitions in the environment \cite{AsaiF18}.
 
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation approach for learning \strips\ action models.
-The output of the compilation is a classical planning task for which a solution plan is a sequence of actions that determine the precondition/effects of the model operator schemes (the learned model).
+%A more recent work proposes acquiring symbolic action models from unstructured inputs, particularly from images that represent transitions in the environment \cite{AsaiF18}.
+
+Motivated by recent advances on the synthesis of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation approach for learning \strips\ action models.
+The output of the compilation is a classical planning task for which a solution plan is a sequence of actions that determine the precondition/effects of the model operator schemas (the learned model).
 \textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards the \emph{bootstrapping} of planning action models, enabling a planner to gradually learn and update its action model.
@@
 
-As for the type of training samples, we adopt a middle way between unstructured inputs and plan traces, wherein only samples that are state observations are required, without further information about state transitions.
+For the training samples, we adopt a middle way between unstructured inputs and plan traces, wherein only samples that are state observations are required, without further information about state transitions.
 \textbf{Learning from state observations} is a relevant advancement as in many applications (e.g.
@@
 This new task consists in learning a new model that induces the observations whilst measuring the \emph{edit distance} to the input model.
-This way, the compilation scheme is also extensible to assess how well a \strips\ action model matches a given set of observations and it can thus be regarded as a first step to apply \textbf{action-model recognition}.
+This way, the compilation approach is also extensible to assess how well a \strips\ action model matches a given set of observations and it can thus be regarded as a first step to apply \textbf{action-model recognition}.
+
+The paper is structured as follows.
+After introducing some initial planning concepts in section \ref{background}, we explain the learning task in section \ref{learning_task}.
+Section \ref{model_evaluation} presents the elements for evaluating the learned \strips\ action models and Section \ref{experimental_evaluation} presents the experimental evaluation.
+Finally, section \ref{conclusions} draws some conclusions.
 
 \section{Background}
+\label{background}
+
 Our approach for learning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
@@
 either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent).
+A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent).
 We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
@@
 
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. Each action $a\in A$ comprises three sets of literals:
-\begin{itemize}
-\item $\pre(a)\subseteq\mathcal{L}(F)$, called {\em preconditions}, the literals that must hold for the action $a\in A$ to be applicable.
-\item $\eff^+(a)\subseteq\mathcal{L}(F)$, called {\em positive effects}, that defines the fluents set to true by the application of the action $a\in A$.
-\item $\eff^-(a)\subseteq\mathcal{L}(F)$, called {\em negative effects}, that defines the fluents set to false by the action application.
-\end{itemize}
+A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
+An action $a\in A$ is defined with preconditions, $\pre(a)\subseteq\mathcal{L}(F)$, positive effects, $\eff^+(a)\subseteq\mathcal{L}(F)$, and negative effects $\eff^-(a)\subseteq\mathcal{L}(F)$.
 We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
 The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+
+%\begin{itemize}
+%\item $\pre(a)\subseteq\mathcal{L}(F)$, called {\em preconditions}, the literals that must hold for the action $a\in A$ to be applicable.
+%\item $\eff^+(a)\subseteq\mathcal{L}(F)$, called {\em positive effects}, that defines the fluents set to true by the application of the action $a\in A$.
+%\item $\eff^-(a)\subseteq\mathcal{L}(F)$, called {\em negative effects}, that defines the fluents set to false by the action application.
+%\end{itemize}
 
@@
 
-\subsection{\strips\ action schemes}
-This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+\subsection{\strips\ action schemas}
+This work addresses the learning of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
 Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
@@
 Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
+For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
 
-Let us define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and that defines the elements that can appear in an action schema.
-For the blocksworld, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+We define $F_v$, a new set of fluents s.t.
+$F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and defines the elements that can appear in an action schema.
+For the {\em blocksworld}, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
-We assume also that actions $a\in A$ are instantiated from \strips\ operator schemes $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+We assume also that actions $a\in A$ are instantiated from \strips\ operator schemas $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
 \begin{itemize}
 \item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
-The headers of a four-operator blocksworld are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
+The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
 \item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
@@
 \section{Learning \strips\ action models}
+\label{learning_task}
+
 %Learning \strips\ action models from plans where every action in the plan is available, as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
@@
 \begin{itemize}
-\item $\mathcal{M}$ is the set of empty operator schemes, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
+\item $\mathcal{M}$ is the set of empty operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
 \item $\Psi$ is the set of predicates, that define the abstract state space of a given classical planning frame.
@@
 
-A solution to $\Lambda$ is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the given sequence of state observations $\mathcal{O}$.
+A solution to $\Lambda$ is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the state observation sequence $\mathcal{O}$.
 A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains $\mathcal{O}$.
-Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning a \strips\ action model in the blocksworld from the five-state observations sequence that corresponds to inverting a 2-block tower.
+Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning a \strips\ action model in the {\em blocksworld} from the five-state observations sequence that corresponds to inverting a 2-block tower.
 
@@
 \begin{enumerate}
-\item {\bf Programs the \strips\ action model $\mathcal{M}'$}.
+\item {\bf Programs the action model $\mathcal{M}'$}.
 A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item {\bf Validates the \strips\ action model $\mathcal{M}'$ in $\mathcal{O}$}.
+\item {\bf Validates the action model $\mathcal{M}'$ in $\mathcal{O}$}.
 The solution plan continues with a postfix that produces the given sequence of states $\tup{s_0,s_1,\ldots,s_{n}}$ using the programmed action model $\mathcal{M}'$.
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
-\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemes are programmed or validated and fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
+\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemas are programmed or validated and fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
 \end{itemize}
 \item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
-Our compilation assumes that initially operator schemes are programmed with every possible precondition, no negative effect and no positive effect.
+Our compilation assumes that initially operator schemas are programmed with every possible precondition, no negative effect and no positive effect.
 Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
@@
 \begin{small}
-Once operator schemes $\mathcal{M}'$ are programmed, they can only be applied because of the $mode_{prog}$ fluent.
+Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied because of the $mode_{prog}$ fluent.
 In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemes that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence {\small $1\leq i\leq n$}.
+These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence {\small $1\leq i\leq n$}.
 This means that the programmed action model $\mathcal{M}'$ complies with the provided observations $\mathcal{O}$ and hence, solves $\Lambda$.
@@
 \section{Evaluation of \strips\ action models}
-We assess how well a \strips\ action model $\mathcal{M}$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} that is required by $\mathcal{M}$ to produce $\mathcal{O}$.
+\label{model_evaluation}
+
+We assess how well a \strips\ action model $\mathcal{M}$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} that is required to apply in $\mathcal{M}$ to induce $\mathcal{O}$.
 %In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
@@
 \subsection{Edition of \strips\ action models}
-Let us define first the \emph{operations} allowed to edit a given \strips\ action model.
-With the aim of keeping tractable the branching factor of the planning instance that results from our compilation, we only define two {\em edit operations}:
+
+We first define the allowed \emph{operations} to edit a given action model.
+With the aim of keeping a tractable branching factor of the planning instance that results from our compilation, we only define two {\em edit operations}:
 \begin{itemize}
@@
 
-Now, we can formalize an edit distance that quantifies how dissimilar two given \strips\ action models are.
-The distance is symmetric and meets the {\em metric axioms} provided that the two {\em \strips\ edit operations} have the same positive cost.
+We can now formalize an edit distance that quantifies how dissimilar two given \strips\ action models are.
+The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations} have the same positive cost.
 
@@
 Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well.
-In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects are two times $|F_v(\xi)|$.
+In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
 
 \begin{definition}
-The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3|F_v(\xi)|$ and sets an upperbound on the distance from $\mathcal{M}$ to any \strips\ model definable within $F_v$.
+The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3|F_v(\xi)|$, which sets an upper bound on the distance from $\mathcal{M}$ to any model definable within $F_v$.
 \end{definition}
@@
 \subsection{The observation edit distance}
-We define now an edit distance to asses the quality of learned models with respect to a sequence of observations (that acts as a test set).
+
+We define now an edit distance to asses the quality of a learned model with respect to a sequence of state observations; i.e., a sample of the test set.
 
@@
 
-Because its semantic nature, the {\em observation edit distance} is robust to evaluate learning episodes where actions are {\em reformulated} and still compliant with the learning inputs.
-For instance a learning episode can interchange the roles of operators with matching headers, or the roles of action parameters with matching types (e.g.
+Due to its semantic nature, the {\em observation edit distance} is a robust measure to evaluate learning episodes where actions are {\em reformulated} and are still compliant with the learning inputs.
+For instance a learning episode can interchange the roles of two operators whose headers match or the roles of action parameters that belong to the same type (e.g.
 the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
@@
 The {\em observation edit distance} can be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
-The larger the {\em observation edit distance} the lower this likelihood.
+The larger the {\em observation edit distance} the lower the likelihood.
 
@@
 \subsection{Computing the observation edit distance}
-Our compilation is extensible to compute the {\em observations edit distance}. A solution to the planning task resulting from the extended compilation is a sequence of actions that:
+
+Our compilation is extensible to compute the {\em observation edit distance} with the inclusion of a learned model $\mathcal{M}$ in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$.
+A solution to the planning task resulting from $\Lambda$ is a sequence of actions that:
+
 \begin{enumerate}
-\item {\bf Edits the \strips\ action model $\mathcal{M}$}. A solution plan starts with a {\em prefix} that edits the preconditions and effects of the action schemes in $\mathcal{M}$.
+\item {\bf Edits the action model $\mathcal{M}$}. A solution plan starts with a {\em prefix} that edits the preconditions and effects of the action schemas in $\mathcal{M}$.
 \item {\bf Validates the edited model $\mathcal{M}$ in $\mathcal{O}$}. The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$.
@@
 
-In this case the tuple $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ does not represent a learning task but the task of editing the \strips\ action model $\mathcal{M}$ to produce the observations $\mathcal{O}$.
-The extended compilation outputs a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
+In this case $\Lambda$ is not a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in a edited model $\mathcal{M}'$.
+The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
+
 \begin{itemize}
@@
 \item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
-In addition, the given \strips\ action model $\mathcal{M}$ is now encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, with $f\in F_v(\xi)$, hold in the initial state iff they appear in the given $\mathcal{M}$ model.
+In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
+This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
 \item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
-The actions for {\em applying} an already programmed operator scheme and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
-The only difference here is the actions for {\em programming} operator schemes $\xi\in\mathcal{M}$ now implement the two {\em edit operations} (i.e.
-include actions for adding a precondition and for removing a negative/positive effect).
+The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
+The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
+actions for adding a precondition and for removing a negative/positive effect).
 \end{itemize}
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan} solves the task of editing a blocksworld action model where the {\tt\small stack} scheme misses the positive effects {\tt(handempty)} and {\tt (clear ?v1)}.
-This plan edits the {\tt\small stack} scheme, adding these two positive effects, and then validates the edited model at the five state observations shown in Figure~\ref{fig:lexample}.
+To illustrate this, the plan of Figure~\ref{fig:plan} solves the task of editing a \emph{blocksworld} action model $\mathcal{M}$ where the positive effects {\tt(handempty)} and {\tt (clear ?v1)} of the {\tt\small stack} schema are missing.
+This plan edits the {\tt\small stack} schema, adding these two positive effects, and then validates the edited model $\mathcal{M}'$ with the five-observation sequence of Figure~\ref{fig:lexample}.
 
-Since we are computing the {\em observations edit distance}, we are not interested in the resulting edited model but in the number of {\em edit operations} required to produce a model validable in the given observations.
-The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions), and is approximate if $P_{\Lambda}'$ is solved with a satisfying planner or furthermore, if what is solved is not $P_{\Lambda}'$ but its relaxation~\cite{bonet2001planning}.
+When we compute the {\em observation edit distance}, our interest is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations.
+The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); and it will be an approximate value if $P_{\Lambda}'$ is solved with a satisfying planner.
+Furthermore, it will be a less accurate estimate if the task to solve is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
 
@@
 
-\section{Evaluation}
+\section{Experimental Evaluation}
+\label{experimental_evaluation}
+
 This section evaluates the performance of our approach for learning \strips\ action models from state observations.
@@
 \subsubsection{Reproducibility}
+
 We used 15 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-For the learning of the \strips\ action models we only used observations sequences $\tup{s_0, s_1, \ldots, s_{24}}$ of twenty five states per domain.
-The set of learning examples is fixed for all the experiments so the results reported by the different evaluation approaches is comparable.
+We only used observation sequences $\tup{s_0, s_1, \ldots, s_{24}}$ composed of twenty five states per domain.
+The set of training samples is fixed for all the experiments so the results reported by the different evaluation approaches is comparable.
 All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
 
-{\sc Madagascar} is the classical planner we use to solve the instances that result from our compilations because its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
-Because its SAT-based nature, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
-Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
+{\sc Madagascar} is the classical planner we used to solve the instances that result from our compilations for its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
+Due to its SAT-based nature, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because there is no interaction between these actions.
+Actions for programming effects can also be applied in a single planning step, thus significantly reducing the planning horizon.
 
-The compilation source code, evaluation scripts and benchmarks (including the learning and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
+The compilation source code, evaluation scripts and benchmarks (including the training and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
 
 \subsubsection{Evaluating with a test set}
-When a reference model is not available, the evaluation of the learning process can be done with respect to a test set.
-Table~\ref{fig:observationstest} summarizes the obtained results when evaluating the quality of the learned models with respect to an state observations test set.
-The table reports, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
 
-The reported results showed that, despite learning only from 25 state observations, twelve out of the fifteen learned domains achieved ratios of $90\%$ or above.
-This fact evidences that the learned models required small edition amounts to match the observations given in the test set (each test set comprises between 20 and 50 observations).
+When a reference model is not available, the learned models are tested with an observation set.
+Table~\ref{fig:observationstest} summarizes the results obtained when evaluating the quality of the learned models with respect to a test set of state observations.
+Each test set comprises between 20 and 50 observations and was generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
+
+The table shows, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
+The reported results show that, despite learning only from 25 state observations, 12 out of 15 learned domains yield ratios of $90\%$ or above.
+This fact evidences that the learned models require very small amounts of edition to match the observations of the given test set.
 
 \subsubsection{Evaluating with a reference model}
-Here we evaluate the learned models with respect to the actual generative model since, opposite to what usually happens in ML, this model is available when learning models from the IPC.
-For each domain, the learned model is compared with the actual model and its quality quantified with the {\em precision} and {\em recall} metrics:
+
+Here we evaluate the learned models with respect to the actual generative model since.
+Opposite to what usually happens in ML, this model is available when learning is applied to IPC domains.
+The model learned for each domain is compared with its reference model, and the quality is measured with the {\em precision} and {\em recall} metrics:
+
 \begin{itemize}
-\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates in the learned model that should not appear).
-Gives a notion of {\em soundness}.
-\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing). Gives a notion of {\em completeness}.
+\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
+Precision gives a notion of {\em soundness}.
+\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing). Recall gives a notion of {\em completeness}.
 \end{itemize}
 
-Table~\ref{fig:observationsnomap} summarizes the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
-The scores are lower than in Table~\ref{fig:observationstest}.
-The explanation is that {\em precision} and {\em recall}, given its syntax nature, report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model (changes in the roles of the actions or their parameters).
+Table~\ref{fig:observationsnomap} shows the Precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
+The reason why the scores in Table ~\ref{fig:observationsnomap} are lower than in Table~\ref{fig:observationstest} is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but actually correspond to {\em reformulations} of the actual model (changes in the roles of the actions or their parameters).
 
 \subsubsection{Precision and recall robust to model reformulations}
-To give insight to the actual quality of the learned models we define a method for computing {\em Precision} and {\em Recall} that is robust to such model reformulations.
+
+To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to model reformulations.
 Precision and recall are often combined using the {\em harmonic mean}.
 This expression, called the {\em F-measure} or the balanced {\em F-score}, is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
-Given the learned \strips\ action model $\mathcal{M}$ and the reference \strips\ action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure}.
+Given the learned action model $\mathcal{M}$ and the reference action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure}.
 
-Table~\ref{fig:observationsmap} shows that significantly higher {\em precision} and {\em recall} scores are reported when comparing each learned action scheme $\xi\in\mathcal{M}$ with the corresponding reference scheme $f_{P\&R}(\xi)\in \mathcal{M}^*$ given by the $f_{P\&R}$ mapping.
-These results evidenced that in all of the evaluated domains, except {\em ferry} and {\em satellite}, learning interchanged the roles of actions or parameters with respect to their role in the reference model.
+Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
+These results evidence that in all of the evaluated domains, except for {\em ferry} and {\em satellite}, the learning task interchanges the roles of some actions or parameters with respect to their role in the reference model.
 
-We can claim now that the action model for the {\em blocksworld} and the {\em gripper} domains were perfectly learned from only 25 state observations.
-In few domains, Table~\ref{fig:observationsmap} still reports lower scores than Table~\ref{fig:observationstest}.
-The explanation is that, how the observations of test sets are generated, affect to the results reported on Table~\ref{fig:observationstest}.
-For instance, the model learned for the {\em driverlog} missed the action scheme for the {\tt disembark-truck} action because this action was not considered in the learning set.
-The same happened in {\em floortile} with the {\tt paint-down} action and in {\em parking} with the {\tt move-curb-to-curb} action.
-We realized that these actions do not appear either in the corresponding test sets so the learned action models were not be penalized in the scores of Table~\ref{fig:observationstest} for not learning these action schemes.
+As we can see in Table~\ref{fig:observationsmap}, the {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
+On the other hand, the learning scores of half of the domains in Table~\ref{fig:observationsmap} are still lower than in Table~\ref{fig:observationstest}.
+The reason lies in the particular observations comprised by the test sets.
+%This leads us to the question on how the test set was generated and how the observations in this set affect the results in Table~\ref{fig:observationstest}. 
+As an example, in the {\em driverlog} domain, the action schema {\small \tt disembark-truck} is missing from the learned model because this action is never induced from the observations in the training set; that is, such action never appears in the \emph{unobserved} plan.
+The same happens with the {\small \tt paint-down} action of the {\em floortile} domain or {\small \tt move-curb-to-curb} in the {\em parking} domain.
+Interestingly, these actions do not appear either in the test sets and so the learned action models are not penalized in Table~\ref{fig:observationstest}.
 %The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
@@
 \end{table}
+
 \section{Conclusions}
+\label{conclusions}
+
 Unlike extensive-data ML approaches, our work explores an alternative research direction to learn sound models from small amounts of state observations.
-As far as we know, this is the first work on learning \strips\ action models from state observations, exclusively using classical planning and evaluated over a wide range of different domains.
-Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
+To the best of our knowledge, this is the first work on learning \strips\ action models from state observations, using exclusively planning technology, and evaluated over a wide range of different domains.
+Recently, the work in~\cite{SternJ17} proposes a planning compilation for learning action models from plan traces following the {\em finite domain} representation for the state variables.
+This is a theoretical study on the boundaries of the learned models and no experimental results are reported.
 
-The empirical results show that since our approach is not statistical by inference-based, it can generate good quality models from very small data sets.
-The action models of the {\em blocksworld} or the {\em gripper} were perfectly learned from only 25 state observations besides, in ten out of fifteen domains, the learned models achieved both {\em Precision} and {\em Recall} values over $0.75$.
+Unlike statistical learning, our inference-based approach is able to produce good-quality models from very small data sets.
+The action models of the {\em blocksworld} or {\em gripper} domains were perfectly learned from only 25 state observations.
+Moreover, in 10 out of the 15 domains, the learned models yield {\em Precision} and {\em Recall} values over $0.75$.
 
-Generating {\em informative} observations for learning planning action models is however an open issue.
+Generating {\em informative} observations for learning action models is still an open issue.
 Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
-The combination of such exploration techniques with our learning approach is an intriguing research direction towards the bootstrapping of planning action models.
+The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques to autonomously collect informative learning examples.
+The combination of such exploration techniques with our learning approach is an appealing research direction towards the bootstrapping of planning action models.
 

### 2018-01-29 doc/icaps-18/camera-ready/strips-learning.tex
added folder for icaps18 camera ready

--- 2017-11-16 doc/icaps-18/strips-learning.tex
@@
 % Commented for blind submission
-%\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
-%{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-%{\small Universitat Polit\`ecnica de Val\`encia.}\\
-%{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
-%{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
+\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
+{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
+{\small Universitat Polit\`ecnica de Val\`encia.}\\
+{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
+{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
 
@@
 % Commented for blind submission
-%\begin{small}
-%\subsection*{Acknowledgment}
-%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
+\begin{small}
+\subsection*{Acknowledgment}
+Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
 Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
-%\end{small}
-
-\newpage
+\end{small}
 

### 2018-01-29 doc/ijcai-18/observations-learning.tex
updates to the ijcai18 paper after 29JAN meeting

@@
 This paper presents a classical planning compilation for learning \strips\ action models from state observations.
-The compilation approach does not require observing the precise actions that produced the observations because such actions are determined by a planner.
-Furthermore, the presented compilation is extensible to assess how well a \strips\ action model matches a given set of observations.
-Last but not least, the paper evaluates the performance of the proposed approach by learning action models for a wide range of classical planning domains from the International Planning Competition and assessing the learned models with respect to (1) the corresponding reference models and (2), given observations test sets.
+The compilation approach does not require observing the precise actions that produced the observed states because such actions, are determined by a planner.
+Furthermore, the compilation is also extensible to assess how well a \strips\ action model matches a given set of observations.
+Last but not least, the paper evaluates the performance of the compilation approach by learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of state observations and (2), the true (reference) models.
 \end{abstract}
@@
 
-The automated learning of action models in planning is a promising approach, as an alternative to hand-coding by a human or when the action model is not available.
-Action-model acquisition has been addressed with various algorithms.
-{\sc ARMS}~\cite{yang2007learning} receives  plan traces as training samples and defines the learning task as a set of weighted constraints that must hold for the plans to be correct, which is then solved by a MAX-SAT solver.
-{\sc ARMS} validates the learned model with a test set of example plans of six domains from the IPC.
+Learning action models is a promising approach to relief the {\em knowledge acquisition bottleneck} of automated planning~\cite{kambhampati:modellite:AAAI2007}.
+The learning of planning action models has been addressed with various algorithms.
+{\sc ARMS}~\cite{yang2007learning} receives plan traces, as training samples, and defines the learning task as a set of weighted constraints that the input plans must hold, which is then solved by a MAX-SAT solver.
+{\sc ARMS} validates the learned models with test sets of example plans from six different IPC domains.
 %checking missing preconditions and redundant effects.
-The learning algorithm of {\sc SLAF}~\cite{amir:alearning:JAIR08} generates a CNF formula which is consistent with the input plan traces and partially observed states, and the quality of the learned models is assessed with respect to the true (reference) model.
-In the case of {\sc LOCM}~\cite{cresswell2013acquiring}, the learning samples are only the examples plans, without providing information about predicates or states.
-{\sc LOCM} relies on assumptions on the kind of domain structure of the model and it was able to correctly discover the dynamics of the domain in 12 out of 20 tested domains by comparison with a reference model \cite{GregoryC16}.
-Finally, {\sc AMAN}~\cite{zhuo2013action} is able to work with incorrectly observed plan traces and the accuracy of the learned models is validated in three IPC domains with respect to the reference model.
+The learning algorithm of {\sc SLAF}~\cite{amir:alearning:JAIR08} computes a CNF formula which is consistent with the input plan traces and partially observed states, and the quality of the learned models is assessed with respect to the corresponding reference model.
+In the case of {\sc LOCM}~\cite{cresswell2013acquiring}, the learning samples are only examples plans, without requiring any information about predicates or states.
+{\sc LOCM} relies on assumptions on the kind of domain structure of the model and compares the learned models with the corresponding reference model \cite{GregoryC16}.
+Finally, {\sc AMAN}~\cite{zhuo2013action} can work with incorrectly observed plan traces and validates the accuracy of the learned models in three IPC domains with respect to the reference model.
 
-The learning samples required by all the aforementioned approaches are observations of plan executions in the form of plan traces, and all of them but {\sc ARMS} validate the accuracy of the learned model with respect to a reference model.
-A more recent work defines the acquisition of a symbolic action model as learning the successor function from unstructured inputs, particularly from images that represent transitions in the environment \cite{AsaiF18}.
+The learning samples required by all the aforementioned approaches are observations of plan executions, in the form of plan traces, and all of them, but {\sc ARMS}, evaluate learning performance by syntactically comparing the learned models with respect to the corresponding reference model.
 
@@
 
-Motivated by recent advances on the synthesis of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation approach for learning \strips\ action models.
-The output of the compilation is a classical planning task for which a solution plan is a sequence of actions that determine the precondition/effects of the model operator schemas (the learned model).
+Motivated by recent advances on the synthesis of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation for learning \strips\ action models from state observations.
+A solution to the classical planning task output by our compilation is a sequence of actions that determines the preconditions/effects of the \strips\ operator schemas (the learned action model).
 \textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards the \emph{bootstrapping} of planning action models, enabling a planner to gradually learn and update its action model.
-The practicality of the compilation approach allow us to report results over a wide range of 15 IPC planning domains.
+The practicality of the compilation approach allow us to report learning results over fifteen IPC planning domains.
 
-For the training samples, we adopt a middle way between unstructured inputs and plan traces, wherein only samples that are state observations are required, without further information about state transitions.
-\textbf{Learning from state observations} is a relevant advancement as in many applications (e.g.
-computer games) the actual actions executed by the agent are not directly observable but instead we can observe the resulting states.
-Handling state observations broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition tasks \cite{SohrabiRU16}.
+For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required, without further information about the state transitions.
+\textbf{Learning from state observations} is a relevant advancement as in many applications the actual actions executed by the observed agent are not available but instead, we can observe the resulting states.
+Learning action models from state observations broadens then the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as the learning from unstructured data (like state images \cite{AsaiF18}).
 
-We propose a two-fold model validation.
-The first method compares the learned model to a reference model, evaluating the syntactic and semantic similarity between both.
-In the second method, the learned model is tested with a set of state observation sequences.
-Since plan traces are not available, we extend the compilation to accept a learned model as input besides the observations.
-This new task consists in learning a new model that induces the observations whilst measuring the \emph{edit distance} to the input model.
-This way, the compilation approach is also extensible to assess how well a \strips\ action model matches a given set of observations and it can thus be regarded as a first step to apply \textbf{action-model recognition}.
+Our compilation is extensible to accept a learned model as input besides the state observations.
+This extension allow us to compute a new model that induces the observations whilst assessing the amount of edition required by the input model to satisfy the given observations.
+Thus, our approach allows to {\bf assess how well a \strips\ action model matches a given set of observations}, which enables model recognition.
+Finally, we propose a two-fold validation of the learned \strips\ action models.
+First the learned models are tested with a set of state observation sequences and second, the learned models are compared to the corresponding reference model.
 
-The paper is structured as follows.
-After introducing some initial planning concepts in section \ref{background}, we explain the learning task in section \ref{learning_task}.
+%The paper is structured as follows.
+After Section \ref{background} introduces some planning concepts, Section \ref{learning_task} explains the addressed learning task.
 Section \ref{model_evaluation} presents the elements for evaluating the learned \strips\ action models and Section \ref{experimental_evaluation} presents the experimental evaluation.
@@
 
-Our approach for learning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
+Our approach for learning \strips\ action schemas from state observations is compiling this learning task into a classical planning task with conditional effects.
 
@@
 
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The {\em plan length} is denoted with $|\pi|=n$ .
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
-
 An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
@@
 
-\begin{figure}
-\begin{footnotesize}
-\begin{verbatim}
-(:action stack
- :parameters (?v1 ?v2 - object)
- :precondition (and (holding ?v1) (clear ?v2))
- :effect (and (not (holding ?v1))
-              (not (clear ?v2))
-              (handempty) (clear ?v1)
-              (on ?v1 ?v2)))
-\end{verbatim}
-\end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
-\label{fig:stack}
-\end{figure}
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+The {\em plan length} is denoted with $|\pi|=n$ .
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
 
@@
 Finally, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in a given action schema $\xi$ and that confine the space of possible action models.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} only contains the fluents from $F_v$ that do not involve $v_2$ because the action header contains the single parameter $v_1$.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header contains the single parameter $v_1$.
+
+\begin{figure}
+\begin{footnotesize}
+\begin{verbatim}
+(:action stack
+ :parameters (?v1 ?v2 - object)
+ :precondition (and (holding ?v1) (clear ?v2))
+ :effect (and (not (holding ?v1))
+              (not (clear ?v2))
+              (handempty) (clear ?v1)
+              (on ?v1 ?v2)))
+\end{verbatim}
+\end{footnotesize}
+ \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
+\label{fig:stack}
+\end{figure}
 
@@
 \item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
-Our compilation assumes that initially operator schemas are programmed with every possible precondition, no negative effect and no positive effect.
+Our compilation assumes that initially operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
 Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
@@
 \begin{small}
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfy the state trajectory constraints given by the $\mathcal{O}$ sequence.
 By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraints given by the $\mathcal{O}$ sequence.
 \end{small}
@@
 
-The size of $P_{\Lambda}$ depends on the number of given state observations, that is $|\mathcal{O}|$. The larger this number, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions.
+The size of the classcial planning task $P_{\Lambda}$ output by the compilation depends on:
+\begin{itemize}
+\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ given as input of the $\Lambda$ learning task.
+The larger these numbers, the larget the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding actions for {\em programming} operator schema $\xi\in\mathcal{M}$.
+\item The number of given state observations, that is $|\mathcal{O}|$. The larger this number, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+\end{itemize}
 
@@
 
-We define now an edit distance to asses the quality of a learned model with respect to a sequence of state observations; i.e., a sample of the test set.
+We define now an edit distance to asses the quality of a learned model with respect to a sequence of state observations.
 
@@
 
+\subsection{Computing the observation edit distance}
+
+Our compilation is extensible to compute the {\em observation edit distance} with the inclusion of a given non-empty model $\mathcal{M}$ in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$.
+In other words, now $\mathcal{M}$ is the set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$, $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+A solution to the planning task resulting from $\Lambda$ is now a sequence of actions that:
+
+\begin{enumerate}
+\item {\bf Edits the action model $\mathcal{M}$}.
+A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemas in $\mathcal{M}$ using to the two {\em edit operations}.
+\item {\bf Validates the edited model $\mathcal{M}'$ in $\mathcal{O}$}. The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$.
+\end{enumerate}
+
+In this case $\Lambda$ is not a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
+The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
+
+\begin{itemize}
+\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
+\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
+In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
+This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
+\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
+The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
+The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
+actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
+\end{itemize}
+
+To illustrate this, the plan of Figure~\ref{fig:plan} solves the task of editing a \emph{blocksworld} action model $\mathcal{M}$ where the positive effects {\tt(handempty)} and {\tt (clear ?v1)} of the {\tt\small stack} schema are missing.
+This plan edits the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited model $\mathcal{M}'$ in the five-observation sequence of Figure~\ref{fig:lexample}.
+
+When we compute the {\em observation edit distance}, our interest is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
+$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan}.
+The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); and it will be an approximate value if $P_{\Lambda}'$ is solved with a satisfying planner.
+Furthermore, it will be a less accurate estimate if the task to solve is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
 \begin{figure}
 {\tt\small
-00 : (program\_add\_handempty\_stack)\\
-01 : (program\_add\_clear\_stack\_var1)\\
+00 : (insert\_add\_handempty\_stack)\\
+01 : (insert\_add\_clear\_stack\_var1)\\
 02 : (apply\_unstack block2 block1)\\
@@
 \end{figure}
-
-\subsection{Computing the observation edit distance}
-
-Our compilation is extensible to compute the {\em observation edit distance} with the inclusion of a learned model $\mathcal{M}$ in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$.
-A solution to the planning task resulting from $\Lambda$ is a sequence of actions that:
-
-\begin{enumerate}
-\item {\bf Edits the action model $\mathcal{M}$}. A solution plan starts with a {\em prefix} that edits the preconditions and effects of the action schemas in $\mathcal{M}$.
-\item {\bf Validates the edited model $\mathcal{M}$ in $\mathcal{O}$}. The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$.
-\end{enumerate}
-
-In this case $\Lambda$ is not a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in a edited model $\mathcal{M}'$.
-The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
-
-\begin{itemize}
-\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
-\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
-In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
-\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
-The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
-The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
-actions for adding a precondition and for removing a negative/positive effect).
-\end{itemize}
-
-To illustrate this, the plan of Figure~\ref{fig:plan} solves the task of editing a \emph{blocksworld} action model $\mathcal{M}$ where the positive effects {\tt(handempty)} and {\tt (clear ?v1)} of the {\tt\small stack} schema are missing.
-This plan edits the {\tt\small stack} schema, adding these two positive effects, and then validates the edited model $\mathcal{M}'$ with the five-observation sequence of Figure~\ref{fig:lexample}.
-
-When we compute the {\em observation edit distance}, our interest is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations.
-The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); and it will be an approximate value if $P_{\Lambda}'$ is solved with a satisfying planner.
-Furthermore, it will be a less accurate estimate if the task to solve is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
 
@@
 Opposite to what usually happens in ML, this model is available when learning is applied to IPC domains.
-The model learned for each domain is compared with its reference model, and the quality is measured with the {\em precision} and {\em recall} metrics:
+The model learned for each domain is compared with its reference model, and the quality is measured with:
 
@@
 Table~\ref{fig:observationsnomap} shows the Precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
-The reason why the scores in Table ~\ref{fig:observationsnomap} are lower than in Table~\ref{fig:observationstest} is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but actually correspond to {\em reformulations} of the actual model (changes in the roles of the actions or their parameters).
+The reason why the scores in Table ~\ref{fig:observationsnomap} are lower than in Table~\ref{fig:observationstest} is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but actually correspond to {\em reformulations} of the actual model (changes in the roles of actions with matching headers or parameters with matching types).
 
@@
 
-To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to model reformulations.
+To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model reformulations.
 Precision and recall are often combined using the {\em harmonic mean}.
@@
 Interestingly, these actions do not appear either in the test sets and so the learned action models are not penalized in Table~\ref{fig:observationstest}.
+Generating {\em informative} observations for learning planning action models is still an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+
 %The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
@@
 Moreover, in 10 out of the 15 domains, the learned models yield {\em Precision} and {\em Recall} values over $0.75$.
-
-Generating {\em informative} observations for learning action models is still an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
 The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques to autonomously collect informative learning examples.

### 2018-01-30 doc/ijcai-18/observations-learning.tex
Minor changes

@@
 Motivated by recent advances on the synthesis of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation for learning \strips\ action models from state observations.
-A solution to the classical planning task output by our compilation is a sequence of actions that determines the preconditions/effects of the \strips\ operator schemas (the learned action model).
+A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions/effects of the \strips\ operator schemas (the learned action model).
 \textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards the \emph{bootstrapping} of planning action models, enabling a planner to gradually learn and update its action model.
@@
 \textbf{Learning from state observations} is a relevant advancement as in many applications the actual actions executed by the observed agent are not available but instead, we can observe the resulting states.
-Learning action models from state observations broadens then the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as the learning from unstructured data (like state images \cite{AsaiF18}).
+Learning action models from state observations broadens then the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as learning from unstructured data (like state images \cite{AsaiF18}).
 
@@
 \begin{lemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves the $\Lambda$ learning task.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$.
 \end{lemma}
@@
 
-The size of the classcial planning task $P_{\Lambda}$ output by the compilation depends on:
+The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
 \begin{itemize}
-\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ given as input of the $\Lambda$ learning task.
-The larger these numbers, the larget the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding actions for {\em programming} operator schema $\xi\in\mathcal{M}$.
+\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ given as input to the $\Lambda$ learning task.
+The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding actions for {\em programming} operator schema $\xi\in\mathcal{M}$.
 \item The number of given state observations, that is $|\mathcal{O}|$. The larger this number, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
@@
 
-To illustrate this, the plan of Figure~\ref{fig:plan} solves the task of editing a \emph{blocksworld} action model $\mathcal{M}$ where the positive effects {\tt(handempty)} and {\tt (clear ?v1)} of the {\tt\small stack} schema are missing.
+To illustrate this, the plan of Figure~\ref{fig:plan} solves the task of editing a \emph{blocksworld} action model $\mathcal{M}$ where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
 This plan edits the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited model $\mathcal{M}'$ in the five-observation sequence of Figure~\ref{fig:lexample}.

### 2018-01-30 doc/ijcai-18/observations-learning.tex
small updates to ijcai18 paper

@@
 The compilation approach does not require observing the precise actions that produced the observed states because such actions, are determined by a planner.
-Furthermore, the compilation is also extensible to assess how well a \strips\ action model matches a given set of observations.
-Last but not least, the paper evaluates the performance of the compilation approach by learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of state observations and (2), the true (reference) models.
+Furthermore, the compilation is extensible to assess how well a \strips\ action model matches a given set of observations.
+Last but not least, the paper evaluates the performance of the compilation approach by learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of state observations and (2), the true models.
 \end{abstract}
@@
 Learning action models is a promising approach to relief the {\em knowledge acquisition bottleneck} of automated planning~\cite{kambhampati:modellite:AAAI2007}.
-The learning of planning action models has been addressed with various algorithms.
+Over the last decade, the learning of planning action models has been addressed with various approaches.
 {\sc ARMS}~\cite{yang2007learning} receives plan traces, as training samples, and defines the learning task as a set of weighted constraints that the input plans must hold, which is then solved by a MAX-SAT solver.
@@
 %checking missing preconditions and redundant effects.
-The learning algorithm of {\sc SLAF}~\cite{amir:alearning:JAIR08} computes a CNF formula which is consistent with the input plan traces and partially observed states, and the quality of the learned models is assessed with respect to the corresponding reference model.
+The learning algorithm of {\sc SLAF}~\cite{amir:alearning:JAIR08} computes a CNF formula which is consistent with the input plan traces and partially observed states.
+{\sc SLAF} assesses the quality of the learned models with respect to the corresponding reference model.
 In the case of {\sc LOCM}~\cite{cresswell2013acquiring}, the learning samples are only examples plans, without requiring any information about predicates or states.
 {\sc LOCM} relies on assumptions on the kind of domain structure of the model and compares the learned models with the corresponding reference model \cite{GregoryC16}.
-Finally, {\sc AMAN}~\cite{zhuo2013action} can work with incorrectly observed plan traces and validates the accuracy of the learned models in three IPC domains with respect to the reference model.
+Finally, {\sc AMAN}~\cite{zhuo2013action} can work with incorrectly observed plan traces and evaluates the learned models in three IPC domains, regarding the corresponding reference model.
 
@@
 Motivated by recent advances on the synthesis of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, we present a novel planning compilation for learning \strips\ action models from state observations.
-A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions/effects of the \strips\ operator schemas (the learned action model).
-\textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards the \emph{bootstrapping} of planning action models, enabling a planner to gradually learn and update its action model.
+A solution to the classical planning task that results from our compilation is a sequence of actions that determines the learned action model, i.e.
+the preconditions and effects of the target \strips\ operator schemas.
+\textbf{Learning action models as planning} leverages off-the-shelf planners and opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
 The practicality of the compilation approach allow us to report learning results over fifteen IPC planning domains.
 
-For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required, without further information about the state transitions.
-\textbf{Learning from state observations} is a relevant advancement as in many applications the actual actions executed by the observed agent are not available but instead, we can observe the resulting states.
-Learning action models from state observations broadens then the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as learning from unstructured data (like state images \cite{AsaiF18}).
+For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required.
+\textbf{Learning from state observations} is a relevant advancement as in many applications, the actual actions executed by the observed agent are not available but instead, the resulting states can be observed.
+Learning action models from state observations broadens the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as learning from unstructured data (like state images \cite{AsaiF18}).
 
 Our compilation is extensible to accept a learned model as input besides the state observations.
-This extension allow us to compute a new model that induces the observations whilst assessing the amount of edition required by the input model to satisfy the given observations.
+This extension allow us to compute a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
 Thus, our approach allows to {\bf assess how well a \strips\ action model matches a given set of observations}, which enables model recognition.
-Finally, we propose a two-fold validation of the learned \strips\ action models.
-First the learned models are tested with a set of state observation sequences and second, the learned models are compared to the corresponding reference model.
+Finally, we carry out a two-fold validation of our learning approach.
+First the learned \strips\ action models are tested with a set of state observation sequences and second, the learned models are compared to the corresponding reference model.
 
@@
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-An action $a\in A$ is defined with preconditions, $\pre(a)\subseteq\mathcal{L}(F)$, positive effects, $\eff^+(a)\subseteq\mathcal{L}(F)$, and negative effects $\eff^-(a)\subseteq\mathcal{L}(F)$.
+An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$.
 We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
@@
 For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
-
 We define $F_v$, a new set of fluents s.t.
@@
 
-Finally, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in a given action schema $\xi$ and that confine the space of possible action models.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header contains the single parameter $v_1$.
+Finally, given the set of predicates $\Psi$ and the header of a \strips\ operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in the action schema $\xi$ and that confine the space of possible action models.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.
 
@@
 \begin{itemize}
-\item $\mathcal{M}$ is the set of empty operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
+\item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
 \item $\Psi$ is the set of predicates, that define the abstract state space of a given classical planning frame.
@@
 A solution to $\Lambda$ is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the state observation sequence $\mathcal{O}$.
-A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains $\mathcal{O}$.
-Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning a \strips\ action model in the {\em blocksworld} from the five-state observations sequence that corresponds to inverting a 2-block tower.
+Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
 
@@
 Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
+A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains $\mathcal{O}$.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
-\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemas are programmed or validated and fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
+\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemas are programmed or validated and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
 \end{itemize}
 \item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
-Our compilation assumes that initially operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
+Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
 Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
@@
 
-\item Actions for {\em applying} an already programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
-Given that the operators headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
+\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
+Given that the operators headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
 \begin{small}
@@
 \begin{small}
-Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied because of the $mode_{prog}$ fluent.
+Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
 In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
@@
 By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraints given by the $\mathcal{O}$ sequence.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraints given by $\mathcal{O}$.
 \end{small}
@@
 \begin{itemize}
-\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ given as input to the $\Lambda$ learning task.
-The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding actions for {\em programming} operator schema $\xi\in\mathcal{M}$.
-\item The number of given state observations, that is $|\mathcal{O}|$. The larger this number, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
+The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding set of {\em programming} actions.
+\item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
 \end{itemize}
 
-\section{Evaluation of \strips\ action models}
+\section{Evaluating \strips\ action models}
 \label{model_evaluation}
@@
 
-We first define the allowed \emph{operations} to edit a given action model.
-With the aim of keeping a tractable branching factor of the planning instance that results from our compilation, we only define two {\em edit operations}:
+We first define the allowed \emph{operations} to edit a given \strips\ action model.
+With the aim of keeping a tractable branching factor of the planning instances that results from our compilation, we only define two {\em edit operations}:
 \begin{itemize}
@@
 \begin{definition}
-The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3|F_v(\xi)|$, which sets an upper bound on the distance from $\mathcal{M}$ to any model definable within $F_v$.
+The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3|F_v(\xi)|$.
 \end{definition}
@@
 
-We define now an edit distance to asses the quality of a learned model with respect to a sequence of state observations.
+We define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
 
 \begin{definition}
-Given an action model $\mathcal{M}$ built from the set of possible elements $F_v$ and a sequence of observations $\mathcal{O}$.
-The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$ such that: (1) $\mathcal{M}'$ is also definable within $F_v$ and (2), $\mathcal{M}'$ can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
+Given an action model $\mathcal{M}$, built from $F_v$, and the observations sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$.
+The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
 
-Due to its semantic nature, the {\em observation edit distance} is a robust measure to evaluate learning episodes where actions are {\em reformulated} and are still compliant with the learning inputs.
-For instance a learning episode can interchange the roles of two operators whose headers match or the roles of action parameters that belong to the same type (e.g.
-the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
+The $\Lambda$ learning tasks can swap the roles of two operators whose headers match or two action parameters that belong to the same type (e.g.
+the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped).
+The {\em observation edit distance} is a semantic measure that is robust to role changes of this kind.
 
-Furthermore, the {\em observation edit distance} enables the recognition of \strips\ action models.
+Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning} the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
+This fact enables the recognition of \strips\ action models.
 The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
-The {\em observation edit distance} can be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
-The larger the {\em observation edit distance} the lower the likelihood.
+The {\em observation edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
 
@@
 
-Our compilation is extensible to compute the {\em observation edit distance} with the inclusion of a given non-empty model $\mathcal{M}$ in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$.
-In other words, now $\mathcal{M}$ is the set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$, $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-A solution to the planning task resulting from $\Lambda$ is now a sequence of actions that:
+Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$, is {\em non-empty}.
+In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+A solution to the planning task resulting from the extended compilation is a sequence of actions that:
 
@@
 
-In this case $\Lambda$ is not a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
+Now $\Lambda$ is not a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
 The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
@@
 
-To illustrate this, the plan of Figure~\ref{fig:plan} solves the task of editing a \emph{blocksworld} action model $\mathcal{M}$ where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-This plan edits the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited model $\mathcal{M}'$ in the five-observation sequence of Figure~\ref{fig:lexample}.
+To illustrate this, the plan of Figure~\ref{fig:plan} solves the classical planning task $P_{\Lambda}'$ that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:lexample}.
 
-When we compute the {\em observation edit distance}, our interest is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
+To compute the {\em observation edit distance}, our interest is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
 $\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan}.
-The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); and it will be an approximate value if $P_{\Lambda}'$ is solved with a satisfying planner.
-Furthermore, it will be a less accurate estimate if the task to solve is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
+The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
 \begin{figure}
@@
 				npuzzle & 2 & 24 & 0.92 \\
-				parking & 5 & 111 & 0.95 \\
+                                parking & 4 & 111 & 0.96 \\                                
 				satellite & 24 & 75 & 0.68 \\
@@
 We used 15 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only used observation sequences $\tup{s_0, s_1, \ldots, s_{24}}$ composed of twenty five states per domain.
-The set of training samples is fixed for all the experiments so the results reported by the different evaluation approaches is comparable.
+The training set comprises twenty five states per domain, i.e.
+$\tup{s_0, s_1, \ldots, s_{24}}$, and is fixed for all the experiments so the results reported by different evaluation approaches are comparable.
 All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
@@
 Table~\ref{fig:observationstest} summarizes the results obtained when evaluating the quality of the learned models with respect to a test set of state observations.
-Each test set comprises between 20 and 50 observations and was generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
+Each test set comprises between 20 and 50 observations per domain and is generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
 
@@
 
-Here we evaluate the learned models with respect to the actual generative model since.
+Here we evaluate the learned models with respect to the actual generative model.
 Opposite to what usually happens in ML, this model is available when learning is applied to IPC domains.
-The model learned for each domain is compared with its reference model, and the quality is measured with:
+The model learned for each domain is compared with its reference model using:
 
@@
 
-Table~\ref{fig:observationsnomap} shows the Precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
-The reason why the scores in Table ~\ref{fig:observationsnomap} are lower than in Table~\ref{fig:observationstest} is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but actually correspond to {\em reformulations} of the actual model (changes in the roles of actions with matching headers or parameters with matching types).
+Table~\ref{fig:observationsnomap} shows the precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
+The reason why the scores in Table ~\ref{fig:observationsnomap} are lower than in Table~\ref{fig:observationstest} is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but correspond to {\em reformulations} of the actual model (changes in the roles of actions with matching headers or parameters with matching types).
 
@@
 To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model reformulations.
+
 Precision and recall are often combined using the {\em harmonic mean}.
-This expression, called the {\em F-measure} or the balanced {\em F-score}, is formally defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
-Given the learned action model $\mathcal{M}$ and the reference action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure}.
-
+This expression, called the {\em F-measure} or the balanced {\em F-score}, is defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
+Given the learned action model $\mathcal{M}$ and the reference action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure} (considering swaps in the actions with matching headers or parameters with matching types).
 Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
-These results evidence that in all of the evaluated domains, except for {\em ferry} and {\em satellite}, the learning task interchanges the roles of some actions or parameters with respect to their role in the reference model.
+These results evidence that in all of the evaluated domains, except for {\em ferry} and {\em satellite}, the learning task swaps the roles of some actions (or parameters) with respect to their role in the reference model.
 
 As we can see in Table~\ref{fig:observationsmap}, the {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
-On the other hand, the learning scores of half of the domains in Table~\ref{fig:observationsmap} are still lower than in Table~\ref{fig:observationstest}.
+On the other hand, the learning scores of many domains in Table~\ref{fig:observationsmap} are still lower than in Table~\ref{fig:observationstest}.
 The reason lies in the particular observations comprised by the test sets.
 %This leads us to the question on how the test set was generated and how the observations in this set affect the results in Table~\ref{fig:observationstest}.
-As an example, in the {\em driverlog} domain, the action schema {\small \tt disembark-truck} is missing from the learned model because this action is never induced from the observations in the training set; that is, such action never appears in the \emph{unobserved} plan.
+As an example, in the {\em driverlog} domain, the action schema {\small \tt disembark-truck} is missing from the learned model because this action is never induced from the observations in the training set; that is, such action never appears in the corresponding \emph{unobserved} plan.
 The same happens with the {\small \tt paint-down} action of the {\em floortile} domain or {\small \tt move-curb-to-curb} in the {\em parking} domain.
@@
 The action models of the {\em blocksworld} or {\em gripper} domains were perfectly learned from only 25 state observations.
-Moreover, in 10 out of the 15 domains, the learned models yield {\em Precision} and {\em Recall} values over $0.75$.
+Moreover, in 12 out of the 15 domains, the learned models yield {\em Precision} values over $0.75$.
 The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques to autonomously collect informative learning examples.

### 2018-01-30 doc/ijcai-18/observations-learning.tex
Add files via upload

@@
 For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required.
-\textbf{Learning from state observations} is a relevant advancement as in many applications, the actual actions executed by the observed agent are not available but instead, the resulting states can be observed.
+\textbf{Learning from state observations} is a relevant advancement as, in many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
 Learning action models from state observations broadens the range of application to external observers and facilitates the representation of imperfect observability (as shown in plan recognition \cite{SohrabiRU16}) as well as learning from unstructured data (like state images \cite{AsaiF18}).
@@
 
-Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning} the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
+Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning}, the {\em observation edit distance} assesses, with a single expression, the flaws in the preconditions and effects of a given learned model.
 This fact enables the recognition of \strips\ action models.
@@
 
-To compute the {\em observation edit distance}, our interest is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
+Our interest when computing the {\em observation edit distance} is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
 $\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan}.
 The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
+
 \begin{figure}

### 2018-01-31 doc/ijcai-18/observations-learning.tex
ijcai18 paper ready for submission

@@
 
+Finally, our approach is able to {\bf assess how well a \strips\ action model matches a given set of observations}.
 Our compilation is extensible to accept a learned model as input besides the state observations.
-This extension allow us to compute a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
-Thus, our approach allows to {\bf assess how well a \strips\ action model matches a given set of observations}, which enables model recognition.
-Finally, we carry out a two-fold validation of our learning approach.
-First the learned \strips\ action models are tested with a set of state observation sequences and second, the learned models are compared to the corresponding reference model.
+This extension allows us to transform the input model into a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
+The empirical evaluation of our learning approach is two-fold: First the learned \strips\ action models are tested with a set of state observation sequences and second, the learned models are compared to the corresponding reference model.
 
@@
 
-Finally, given the set of predicates $\Psi$ and the header of a \strips\ operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in the action schema $\xi$ and that confine the space of possible action models.
+Finally, given the set of predicates $\Psi$ and the header of a \strips\ operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in the action schema $\xi$ and that confine its space of possible action models.
 For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.
@@
 \begin{itemize}
-\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$.
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$, i.e.
+{\tt\small block1} and {\tt\small block2} in Figure~\ref{fig:lexample}.
 \item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
@@
 Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
-\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, indicates that the programmed action model is validated in all the input observations.
+\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
 \item $A_{\Lambda}$ comprises three kinds of actions:
@@
 By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraints given by $\mathcal{O}$.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O}$.
 \end{small}
@@
 \item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
-The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding set of {\em programming} actions.
+The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
 \item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
@@
 
-We assess how well a \strips\ action model $\mathcal{M}$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} that is required to apply in $\mathcal{M}$ to induce $\mathcal{O}$.
+We assess how well a \strips\ action model $\mathcal{M}$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} required by $\mathcal{M}$ to induce $\mathcal{O}$.
 %In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
@@
 A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemas in $\mathcal{M}$ using to the two {\em edit operations}.
-\item {\bf Validates the edited model $\mathcal{M}'$ in $\mathcal{O}$}. The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$.
+\item {\bf Validates the edited model $\mathcal{M}'$ in $\mathcal{O}$}.
+The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$, as in Section~\ref{learning_task}.
 \end{enumerate}
@@
 The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
-actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
+include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
 \end{itemize}
@@
 $\tup{s_0, s_1, \ldots, s_{24}}$, and is fixed for all the experiments so the results reported by different evaluation approaches are comparable.
-All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
+All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM and a 600 seconds time bound.
 
@@
 
-The compilation source code, evaluation scripts and benchmarks (including the training and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
+The compilation source code, evaluation scripts and benchmarks (including the used training and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
 
@@
 
-To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model reformulations.
+To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model {\em reformulations}.
 
@@
 As we can see in Table~\ref{fig:observationsmap}, the {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
-On the other hand, the learning scores of many domains in Table~\ref{fig:observationsmap} are still lower than in Table~\ref{fig:observationstest}.
+On the other hand, the learning scores of several domains in Table~\ref{fig:observationsmap} are still lower than in Table~\ref{fig:observationstest}.
 The reason lies in the particular observations comprised by the test sets.
@@
 Interestingly, these actions do not appear either in the test sets and so the learned action models are not penalized in Table~\ref{fig:observationstest}.
-Generating {\em informative} observations for learning planning action models is still an open issue.
+Generating {\em informative} and {\em representative} observations for learning planning action models is an open issue.
 Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.

### 2018-01-31 doc/icaps-18/camera-ready/strips-learning.tex
Acknowledgements, short background, bibliography

@@
 
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. Each action $a\in A$ comprises three sets of literals:
-\begin{itemize}
-\item $\pre(a)\subseteq\mathcal{L}(F)$, called {\em preconditions}, the literals that must hold for the action $a\in A$ to be applicable.
-\item $\eff^+(a)\subseteq\mathcal{L}(F)$, called {\em positive effects}, that defines the fluents set to true by the application of the action $a\in A$.
-\item $\eff^-(a)\subseteq\mathcal{L}(F)$, called {\em negative effects}, that defines the fluents set to false by the action application.
-\end{itemize}
+A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
+An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$.
+We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
+The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+
 We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
@@
 \subsection*{Acknowledgment}
-Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
-Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
+This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.
+Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it Ramon y Cajal}, both programs funded by the Spanish government.
 \end{small}
@@
 \bibliographystyle{aaai}
-\bibliography{strips-learning}
+\bibliography{planlearnbibliography}
 \end{document}

### 2018-02-20 doc/icaps-18/camera-ready/strips-learning.tex
added reviewers comments to icaps18 camera ready

@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-The compilation approach is appealing by itself because opens the door to the bootstrapping of planning action models but also because:
+
+The compilation approach is appealing by itself, because opens the door to the bootstrapping of planning action models, but also because:
 \begin{enumerate}
-\item Is flexible to different amounts of available input knowledge.
-The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate state  or action is observed.
+\item Is flexible to various amounts of available input knowledge.
+Learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate state  or action is observed.
 \item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
@@
 The third section formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
-The fourth and fifth sections describe our approach for addressing the formalized learning tasks.
-Finally, the last sections report the data collected in a empirical evaluation, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
+The fourth and fifth sections describe our compilation approach for addressing the formalized learning tasks.
+Finally, the last sections report the data collected in a empirical evaluation, discuss the strengths and weaknesses of the compilation approach and propose several opportunities for future research.
 
 \section{Background}
-This section defines the planning models used on this work and the output of the learning tasks addressed in the paper.
+This section defines the planning model used on this work and the output of the learning tasks addressed in the paper.
 
-\subsection{Classical planning}
+\subsection{Classical planning with conditional effects}
+Our approach for learning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
+Conditional effects allow us to compactly define actions whose effects depend on the current state.
+Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
+
 We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
@@
 
-We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
-The result of applying $a$ in $s$ is the {\em successor state} $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
-
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-We denote with $|\pi|$ the {\em plan length}.
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of $\pi$ in $I$.
-
-\subsection{Classical planning with conditional effects}
-Our approach for learning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
-Conditional effects allow us to compactly define actions whose effects depend on the current state.
-Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
-
-An action $a\in A$ has now a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
+An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
 Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-
-An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} are the effects whose conditions hold in $s$:
+An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
 \[
@@
 
-The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered {\em negative} and {\em positive} effects, respectively.
+The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
+
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+The {\em plan length} is denoted with $|\pi|=n$ .
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
 
@@
 This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-Figure~\ref{fig:stack} shows the schema, coded in PDDL, for the {\em stack} action from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+Figure~\ref{fig:stack} shows the {\em stack} schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
 
@@
 \end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
+ \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from {\em blocksworld}.}
 \label{fig:stack}
@@
 
-Let us also define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and that defines the elements that can appear in an action schema.
+Let us also define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ exclusively using objects in $\Omega_v$ and that defines the elements that can appear in the preconditions and effects of an action schema.
 For instance, in the blocksworld, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
@@
 When any intermediate state is available, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions.
+Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions~\cite{jimenez2012review}.
 
@@
 
-\subsubsection{Learning from (initial, final) state pairs.} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
-No information about the actions in the plans is given.
+\subsubsection{Learning from (initial, final) state pairs.} This learning task corresponds to observing an agent acting in the world but watching only the result of its plan executions.
+No intermediate information about the actions in the plans is given.
 This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma}$:
@@
 \item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
-Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
+Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
 \end{itemize}
@@
 
-Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda'$ in the blocksworld.
+Figure~\ref{fig:lexample} shows an example of a {\em blocksworld} learning task $\Lambda'$.
 This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
@@
 \begin{itemize}
-\item $\Xi_0$ is a partially specified action model in which some preconditions and effects are a priori known.
+\item $\Xi_0$ is a partially specified action model, i.e. some preconditions and effects are a priori known.
 \end{itemize}
 
-A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant just with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
+A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
 In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given labels $\Sigma$ using the learned \strips\ model.
-A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
+A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
 Finally a solution to $\Lambda''$ is a set of \strips\ operator schema $\Xi$ compliant as well with the provided partially specified action model $\Xi_0$.
@@
 \begin{enumerate}
-\item Programs the \strips\ action model $\Xi$.
+\item {\em Programs the \strips\ action model $\Xi$}.
 A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (the labels $\Sigma$, and $\Pi$ and/or $\Xi_0$ if available).
+\item {\em Validates the programmed \strips\ action model $\Xi$ in the given input knowledge} (the labels $\Sigma$, and $\Pi$ and/or $\Xi_0$ if available).
 For every label $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
@@
 \item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
-Our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect.
+Our compilation assumes that initially any operator schema is programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
 \item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
@@
 \begin{small}
-The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after preconditions are programmed).
-Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
+Once operator schemas $\Xi$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
 To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
@@
 
-The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model.
+The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model. The size of the classical planning task $P_{\Lambda}$ depends on:
+\begin{itemize}
+\item The arity of the actions headers in $\Xi$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
+The larger these numbers, the larger the $F_v$ set, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
+\item The number of learning examples. The larger this number, the more $test_t$ fluents and $\mathsf{validate_{t}}$ actions in $P_{\Lambda}$.
+\end{itemize}
 
@@
 We extend the compilation to consider labeled plans.
-Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
+Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$:
 \begin{itemize}
 \item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the plan step where the action model is validated.
+Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$}, for indicating the plan step where the action model is validated.
 Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
@@
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the blocksworld action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
-This plan programs and validates the operator schema {\tt\small stack} from the blocksworld, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-Plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are the actions for validating the programmed operators following the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
+To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,22]$ validate the programmed operators following the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
 
@@
 \begin{figure}
-{\tt\small
+{\footnotesize\tt
      {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
@@
 \subsection{Static predicates}
-A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action schema~\cite{fox:TIM:JAIR1998}.
+A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
 Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
-Formally, given a static predicate $p$:
+Given a static predicate $p$:
 \begin{itemize}
@@
 
-Likewise, static predicates can constrain the space of possible preconditions looking at the given example plans.
-Fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
+Looking at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
 $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
-Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
+Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
 
-In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of {\em potential} static predicates from the input to the learning task.
+%In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of {\em potential} static predicates from the input to the learning task.
+
+\section{Evaluation}
+This section evaluates the performance of our approach for learning \strips\ action models starting from various amounts of available input knowledge.
+
+\subsubsection{Setup.}
+The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
+All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM.
+
+The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
+We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
+In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
+Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
+
+We make fully available the compilation source code, evaluation scripts and used benchmarks at this repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
+
+\subsubsection{Metrics.}
+The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
+These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
+
+Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
+Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
+Recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
+
+Given the syntax-based nature of these metrics, it may happen that they report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
+a learned model semantically equivalent but syntactically different to the reference model.
+This most likely occurs when the learning task is under-constrained.
+
+\subsection{Learning from labeled plans}
+We start evaluating our approach at $\Lambda'=\tup{\Psi,\Sigma,\Pi}$ learning tasks, where {\em labeled plans} are available, and repeat the evaluation but exploiting potential static predicates computed from $\Sigma$.
+The potential {\em static predicates} considered here is the set of predicates s.t.
+every predicate instantiation appears unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
+Static predicates are used to constrain the space of possible action models as explained in the previous section.
+
+Table~\ref{tab:results_plans} summarizes the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
+We can observe that identifying static predicates drives to models with better precondition {\em recall}.
+This fact evidences that many of the missing preconditions corresponded to static predicates because there were no incentive to learn them as they always hold~\cite{gregory2015domain}.
+
+Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
+All the learning tasks are solved in a few seconds time and interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
+In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
+When static predicates are identified, the resulting compilation is also much compact and produces smaller planning/instantiation times.
 
@@
 
-\section{Evaluation}
-This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts of available input knowledge.
-
-\subsubsection{Setup.}
-The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
-All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
-
-\subsubsection{Reproducibility.}
-We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
-
-\subsubsection{Planner.}
-The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
-In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
-Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
-
-\subsubsection{Metrics.}
-The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
-These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
-
-Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
-Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
-Recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
-
-\subsection{Learning from labeled plans}
-We start evaluating our approach at $\Lambda'=\tup{\Psi,\Sigma,\Pi}$ learning tasks, where {\em labeled plans} are available, and repeat the evaluation but exploiting potential static predicates that are computed from $\Sigma$.
-The potential {\em static predicates} considered here is the set of predicates s.t.
-every predicate instantiation appears unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
-Static predicates are used to constrain the space of possible action models as explained in the previous section.
-
-Table~\ref{tab:results_plans} summarizes the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
-We can observe that identifying static predicates drives to models with better precondition {\em recall}.
-This fact evidences that many of the missing preconditions corresponded to static predicates because there were no incentive to learn them as they always hold~\cite{gregory2015domain}.
-
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-All the learning tasks are solved in a few seconds time and interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
-When static predicates are identified, the resulting compilation is also much compact and produces smaller planning/instantiation times.
-
 \subsection{Learning from partially specified action models}
@@
 We report only the precision and recall for the {\em unknown} actions (considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in these models is 1.0).
-We must also note that errors in $\Lambda''$ tasks have greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is just done over half of the actions (e.g.
+Likewise, errors in $\Lambda''$ tasks have greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is just done over half of the actions (e.g.
 the preconditions recall for the {\em Floortile} or {\em Gripper} domains).
@@
 Remarkably, the overall precision is now $0.98$ which means that the content of the learned models is highly reliable.
-Recall is a lower, 0.87, indicating that the learned models still miss some information.
-Particularly, preconditions are again the component more difficult to be fully learned.
+Recall is a lower, 0.87, indicating that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
 Overall the reported results confirm the previous trend: more input knowledge drives to better models and requires less planning time.
@@
 
-%Table~\ref{tab:time_states} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported since the compilation could not be solved within a time bound of 1000 seconds.
+\section{Related work}
+
+%Back in the 90's various systems aimed learning operators mostly via interaction with the environment.
+{\sc LIVE} captured and formulated observable features of objects and used them to acquire and refine operators \cite{ShenS89}.
+{\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
+These early works were based on lifting the observed states supported by exploratory plans or external teachers.%, but none provided a theoretical justification for this second source of knowledge.
+
+%More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
+
+%Whilst the aforementioned works deal with full state observability,
+Action model learning has also been studied in domains where there is partial or missing state observability.
+{\sf ARMS} works when no partial intermediate state is given.
+It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
+In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
+%and so it may output an inconsistent model
+{\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
+Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
+The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
+This update makes sure that the new formula represents all the transition relations consistent with the actions and observations.
+The formula returned at the end includes all consistent models, which can then be retrieved with additional processing.
+
+%Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
+An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
+
+{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
+This makes {\sf LOCM} be most likely the learning approach that works with the least information possible.
+The lack of available information is addressed by LOCM by exploiting assumptions about the kind of domain model it has to generate.
+Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
+The intuitive assumptions of {\sf LOCM}
+%, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence,
+yield a learning model heavily reliant on the kind of domain structure.
+The inability of {\sf LOCM} to properly derive domain theories where the state of a sort is subject to different FSMs is later overcome by {\sf LOCM2} by forming separate FSMs, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
+{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain}), the last contribution of the {\sf LOCM} family, addresses the problem of inducing static predicates.
+Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
+{\em blocksworld} and {\em freecell}).
+In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
+
+Compiling the learning of action models into classical planning is a general and flexible approach that allows to accommodate various amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
+For instance, the example plans in $\Pi$ could be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), so learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ could also be addressed.
+Furthermore, our approach seems extensible to learning other types of generative models (e.g.
+hierarchical models like HTN or behaviour trees), that can be more appealing than \strips\ models, since using them to compute planning solutions requires less search effort.
+
+%Furthermore, the compilation scheme can be extended to learn other type of generative models that comprise partially-specified solution plans like HTN models or full-specified solutions like FSM or behaviour trees.
 
@@
 \end{table}
-
-\section{Past and future of learning action models}
-
-Back in the 90's various systems aimed learning operators mostly via interaction with the environment.
-{\sc LIVE} captured and formulated observable features of objects and used them to acquire and refine operators \cite{ShenS89}.
-{\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
-These early works were based on lifting the observed states supported by exploratory plans or external teachers.%, but none provided a theoretical justification for this second source of knowledge.
-
-%More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
-
-%Whilst the aforementioned works deal with full state observability,
-Action model learning has also been studied in domains where there is partial or missing state observability.
-{\sf ARMS} works when no partial intermediate state is given.
-It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
-In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
-%and so it may output an inconsistent model
-{\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
-Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
-The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
-This update makes sure that the new formula represents all the transition relations consistent with the actions and observations.
-The formula returned at the end includes all consistent models, which can then be retrieved with additional processing.
-
-%Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
-An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
-
-{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
-This makes {\sf LOCM} be most likely the learning approach that works with the least information possible.
-The lack of available information is addressed by LOCM by exploiting assumptions about the kind of domain model it has to generate.
-Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
-The intuitive assumptions of {\sf LOCM}
-%, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence,
-yield a learning model heavily reliant on the kind of domain structure.
-The inability of {\sf LOCM} to properly derive domain theories where the state of a sort is subject to different FSMs is later overcome by {\sf LOCM2} by forming separate FSMs, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
-{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain}), the last contribution of the {\sf LOCM} family, addresses the problem of inducing static predicates.
-Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
-{\em blocksworld} and {\em freecell}).
-In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
-
-Compiling the learning of action models into classical planning is a general and flexible approach that allows to accommodate different amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
-For instance, the example plans in $\Pi$ could be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), so learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ could also be addressed.
-Furthermore, our approach seems extensible to learning other types of generative models (e.g.
-hierarchical models like HTN or behaviour trees), that can be more appealing than \strips\ models, since using them to compute planning solutions requires less search effort.
-
-%Furthermore, the compilation scheme can be extended to learn other type of generative models that comprise partially-specified solution plans like HTN models or full-specified solutions like FSM or behaviour trees.
 
@@
 We presented a novel approach for learning \strips\ action models from examples using classical planning.
-The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
+The approach is flexible to various amounts of available input knowledge and accepts partially specified action models.
 We also introduced {\em precision} and {\em recall} metrics, widely used in ML, to the evaluation of planning action models.
@@
 As far as we know, this is the first work on learning action models exclusively using an {\em off-the-shelf} classical planner and evaluated over a wide range of different domains.
-Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
+The work in~\cite{SternJ17} proposes a planning compilation for learning action models from plan traces following the {\em finite domain} representation for the state variables.
+This is a theoretical study on the boundaries of the learned models and no experimental results are reported.
 
-Our evaluation shows that, when example plans are available, we can compute accurate action models from small sets of learning examples (only five learning examples per domain) and investing small learning times (less than a second time in most of the domains).
-When action plans are not available, our approach is still able to produce action models compliant with the input information.
-In this case, since learning is not constrained by actions it can reformulate operators changing their semantics and making tricky their comparison with a reference model.
+When example plans are available, we can compute accurate action models from small sets of learning examples (five examples per domain) and investing small learning times (less than a second).
+When action plans are not available, our approach still produces action models compliant with the input information.
+In this case, since learning is not constrained by actions, it can reformulate operators changing their semantics and making tricky their comparison with a reference model.
 %An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
@@
 
-The size of the compiled classical planning instances depend on the number of input examples.
 Generating {\em informative} examples for learning planning action models is an open issue.
 Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
+The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques able to autonomously collect informative learning examples.
 The combination of such exploration techniques with our learning approach is an intriguing research direction that opens the door to the bootstrapping of planning action models.
+
+In many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
+We plan to extend our approach for learning from state observations as it broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}.
 
@@
 This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.
-Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it Ramon y Cajal}, both programs funded by the Spanish government.
+Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it RYC15/18009}, both programs funded by the Spanish government.
 \end{small}

### 2018-02-21 doc/aij18/aij18.tex
first commit aij18 paper

% Template for Elsevier CRC journal article
% version 1.1 dated 16 March 2010

% This file (c) 2010 Elsevier Ltd.  Modifications may be freely made,
% provided the edited file is saved under a different name

% This file contains modifications for Procedia Computer Science
% but may easily be adapted to other journals

% Changes since version 1.0
% - elsarticle class option changed from 1p to 3p (to better reflect CRC layout)

%-----------------------------------------------------------------------------------

%% This template uses the elsarticle.cls document class and the extension package ecrc.sty
%% For full documentation on usage of elsarticle.cls, consult the documentation "elsdoc.pdf"
%% Further resources available at http://www.elsevier.com/latex

%-----------------------------------------------------------------------------------

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Important note on usage                  %%
%% -----------------------                  %%
%% This file must be compiled with PDFLaTeX %%
%% Using standard LaTeX will not work!      %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The '3p' and 'times' class options of elsarticle are used for Elsevier CRC
\documentclass[3p,times]{elsarticle}

%% The `ecrc' package must be called to make the CRC functionality available
\usepackage{ecrc}
%% The ecrc package defines commands needed for running heads and logos.
%% For running heads, you can set the journal name, the volume, the starting page and the authors

%% set the volume if you know. Otherwise `00'
\volume{00}

%% set the starting page if not 1
\firstpage{1}

%% Give the name of the journal
\journalname{Artificial Intelligence}

%% Give the author list to appear in the running head
%% Example \runauth{C.V. Radhakrishnan et al.}
\runauth{}

%% The choice of journal logo is determined by the \jid and \jnltitlelogo commands.
%% A user-supplied logo with the name <\jid>logo.pdf will be inserted if present.
%% e.g. if \jid{yspmi} the system will look for a file yspmilogo.pdf
%% Otherwise the content of \jnltitlelogo will be set between horizontal lines as a default logo

%% Give the abbreviation of the Journal.
\jid{procs}

%% Give a short journal name for the dummy logo (if needed)
\jnltitlelogo{Artificial Intelligence}

%% Hereafter the template follows `elsarticle'.
%% For more details see the existing template files elsarticle-template-harv.tex and elsarticle-template-num.tex.

%% Elsevier CRC generally uses a numbered reference style
%% For this, the conventions of elsarticle-template-num.tex should be followed (included below)
%% If using BibTeX, use the style file elsarticle-num.bst

%% End of ecrc-specific commands
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% The amssymb package provides various useful mathematical symbols
\usepackage{amssymb}
%% The amsthm package provides extended theorem environments
%%\usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers after \end{frontmatter}.
%% \usepackage{lineno}

%% natbib.sty is loaded by default. However, natbib options can be
%% provided with \biboptions{...} command. Following options are
%% valid:

%%   round  -  round parentheses are used (default)
%%   square -  square brackets are used   [option]
%%   curly  -  curly braces are used      {option}
%%   angle  -  angle brackets are used    <option>
%%   semicolon  -  multiple citations separated by semi-colon
%%   colon  - same as semicolon, an earlier confusion
%%   comma  -  separated by comma
%%   numbers-  selects numerical citations
%%   super  -  numerical citations as superscripts
%%   sort   -  sorts multiple citations according to order in ref. list
%%   sort&compress   -  like sort, but also compresses numerical citations
%%   compress - compresses without sorting
%%
%% \biboptions{comma,round}

% \biboptions{}

% if you have landscape tables
\usepackage[figuresright]{rotating}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{pstricks}
\usepackage{pst-node}
\usepackage{multirow}
\usepackage{listings}

\usepackage{pgf}
\usepackage{tikz}
\usepackage{subcaption}
\usetikzlibrary{arrows,automata}
\usepackage{arydshln}

% put your own definitions here:
%   \newcommand{\cZ}{\cal{Z}}
%   \newtheorem{def}{Definition}[section]
%   ...

\newtheorem{mytheorem}{Theorem}
\newtheorem{mylemma}[mytheorem]{Lemma}
\newtheorem{mydefinition}[mytheorem]{Definition}
\newtheorem{myconstruction}{Construction}

\mathchardef\mh="2D
\newcommand{\pre}{\mathsf{pre}}  % precondition
\newcommand{\eff}{\mathsf{eff}}  % effect
\newcommand{\cond}{\mathsf{cond}}   % conditional effect
\newcommand{\add}{\mathsf{add}}  % add effect
\newcommand{\del}{\mathsf{del}}  % delete effect
\newcommand{\PE}{\mathrm{PE}}     % precondition
\newcommand{\PSPACE}{\mathrm{PSPACE}}     % precondition
\newcommand{\NPSPACE}{\mathrm{NPSPACE}}     % precondition
\newcommand{\strips}{\textsc{Strips}}     % precondition

\newcommand{\pbox}[1]{\makebox[2em][l]{#1}}

\newcommand{\tup}[1]{{\langle #1 \rangle}}

\lstset{
  basicstyle=\ttfamily,
  mathescape
}

% add words to TeX's hyphenation exception list
%\hyphenation{author another created financial paper re-commend-ed Post-Script}

% declarations for front matter

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for the associated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for the associated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for the associated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%%
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \address{Address\fnref{label3}}
%% \fntext[label3]{}

\dochead{}
%% Use \dochead if there is an article header, e.g. \dochead{Short communication}

\title{Computing programs for generalized planning using a classical planner}
\author[label1]{Diego Aineto}
\author[label1]{Sergio Jim\'{e}nez Celorrio}
\author[label1]{Eva Onaindia}
\address[label1]{Department of Computer Systems and Computation, Universitat Politécnica de València. Spain}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{<author name>}
%% \address[label1]{<address>}
%% \address[label2]{<address>}

\begin{abstract}

\end{abstract}

\begin{keyword}
Classical planning\sep Planning and learning\sep Learning action models\sep Generalized planning
%% keywords here, in the form: keyword \sep keyword

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%%
%% Start line numbering here if you want
%%
% \linenumbers

%% main text

%% The Appendices part is started with the command \appendix;
%% appendix sections are then done as normal sections
%% \appendix

%% \section{}
%% \label{}

% HLP: Expressiveness is pushed when pure compilations are used. Otherwise we just use them.

\section{Introduction}
\label{sec:Section1}

\begin{small}
\subsection*{Acknowledgment}
This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R. Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it RYC15/18009}, both programs funded by the Spanish government.
\end{small}

%% References
%%
%% Following citation commands can be used in the body text:
%% Usage of \cite is as follows:
%%   \cite{key}         ==>>  [#]
%%   \cite[chap. 2]{key} ==>> [#, chap. 2]
%%

%% References with BibTeX database:

\bibliographystyle{elsarticle-num}
\bibliography{planlearnbibliography.bib}

%% Authors are advised to use a BibTeX database file for their reference list.
%% The provided style file elsarticle-num.bst formats references in the required Procedia style

%% For references without a BibTeX database:

% \begin{thebibliography}{00}

%% \bibitem must have the following form:
%%   \bibitem{key}...
%%

% \bibitem{}

% \end{thebibliography}

\end{document}

%%
%% End of file `ecrc-template.tex'.

### 2018-02-21 doc/aij18/aij18.tex
added sections 1,2 and 3 to AIJ18 paper

@@
 \usepackage{tikz}
+\usetikzlibrary{calc,backgrounds,positioning,fit}
 \usepackage{subcaption}
@@
 
-\title{Computing programs for generalized planning using a classical planner}
+\title{Learning \strips\ action models with classical planning}
 \author[label1]{Diego Aineto}
@@
 \begin{abstract}
-
+  This paper presents a novel approach for learning \strips\ action models from examples of plan executions that compiles this learning task into classical planning.
+The compilation approach is flexible to various amount and forms of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) or sequences of state observations to just a set of initial and final states (where no intermediate action or state is given).
+The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
+What is more, the compilation is extensible to assess how well a \strips\ action model matches a given set of observations.
+Last but not least, the paper evaluates the performance of the compilation approach by learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of observations and (2), the true models.
 \end{abstract}
@@
 %% or \MSC[2008] code \sep code (2000 is the default)
-
 \end{keyword}
@@
 \label{sec:Section1}
+Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
+At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
+Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
+
+On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
+The application of inductive ML to the learning of \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
+\begin{itemize}
+\item The {\em input} to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects.
+The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
+\item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
+When learning \strips\ action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions.
+\end{itemize}
+
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
+A solution to the classical planning task that results from our compilation is a sequence of actions that determines the learned action model, i.e.
+the preconditions and effects of the target \strips\ operator schemas.
+
+The compilation approach is appealing by itself, because leverages off-the-shelf planners and opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
+The practicality of the compilation approach allow us to report learning results over fifteen IPC planning domains.
+In addition,the compilation approach presents the following features:
+\begin{enumerate}
+\item Is flexible to various amounts and forms of available input knowledge.
+Learning examples can range from a set of plans (with their corresponding initial and final states) or state observations to just a set of initial and final states where no intermediate state  or action is observed.
+Learning from state observations is a relevant advancement as, in many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
+Learning action models from state observations broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}).
+\item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
+In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
+For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required.
+\item Assess how well a \strips\ action model matches a given set of observations.
+Our compilation is extensible to accept a learned model as input besides the state observations.
+This extension allows us to transform the input model into a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
+The empirical evaluation of our learning approach is two-fold: First the learned \strips\ action models are tested with a set of state observation sequences and second, the learned models are compared to the corresponding reference model.
+\end{enumerate}
+
+Section~\ref{sec:Section2} reviews related work on learning planning action models.
+Section~\ref{sec:Section3} formalizes the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the output of the addressed learning task).
+Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
+Sections describe our compilation approach for addressing the formalized learning tasks.
+Finally, the last sections report the data collected in a empirical evaluation, discuss the strengths and weaknesses of the compilation approach and propose several opportunities for future research.
+
+\section{Related work}
+\label{sec:Section2}
+Back in the 90's various systems aimed learning operators mostly via interaction with the environment.
+{\sc LIVE} captured and formulated observable features of objects and used them to acquire and refine operators \cite{ShenS89}.
+{\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
+These early works were based on lifting the observed states supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge.
+
+More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
+
+Whilst the aforementioned works deal with full state observability,action model learning has also been studied in domains where there is partial or missing state observability.
+{\sf ARMS} works when no partial intermediate state is given.
+It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
+In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
+%and so it may output an inconsistent model
+{\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
+Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
+The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
+This update makes sure that the new formula represents all the transition relations consistent with the actions and observations.
+The formula returned at the end includes all consistent models, which can then be retrieved with additional processing.
+
+Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
+An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
+
+{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
+This makes {\sf LOCM} be most likely the learning approach that works with the least information possible.
+The lack of available information is addressed by LOCM by exploiting assumptions about the kind of domain model it has to generate.
+Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
+The intuitive assumptions of {\sf LOCM}
+, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence,
+yield a learning model heavily reliant on the kind of domain structure.
+The inability of {\sf LOCM} to properly derive domain theories where the state of a sort is subject to different FSMs is later overcome by {\sf LOCM2} by forming separate FSMs, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
+{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain}), the last contribution of the {\sf LOCM} family, addresses the problem of inducing static predicates.
+Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
+{\em blocksworld} and {\em freecell}).
+In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
+
+Compiling the learning of action models into classical planning is a general and flexible approach that allows to accommodate various amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
+For instance, the example plans in $\Pi$ could be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), so learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ could also be addressed.
+Furthermore, our approach seems extensible to learning other types of generative models (e.g.
+hierarchical models like HTN or behaviour trees), that can be more appealing than \strips\ models, since using them to compute planning solutions requires less search effort.
+
+\section{Background}
+\label{sec:Section3}
+\subsection{Classical planning with conditional effects}
+We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
+A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
+either~$l=f$ or $l=\neg f$.
+A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent).
+We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
+
+A {\em state} $s$ is a full assignment of values to fluents, i.e.
+$|s|=|F|$, so the size of the state space is $2^{|F|}$.
+Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
+
+A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
+An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$.
+We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
+The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+
+%\begin{itemize}
+%\item $\pre(a)\subseteq\mathcal{L}(F)$, called {\em preconditions}, the literals that must hold for the action $a\in A$ to be applicable.
+%\item $\eff^+(a)\subseteq\mathcal{L}(F)$, called {\em positive effects}, that defines the fluents set to true by the application of the action $a\in A$.
+%\item $\eff^-(a)\subseteq\mathcal{L}(F)$, called {\em negative effects}, that defines the fluents set to false by the action application.
+%\end{itemize}
+
+An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+\[
+triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
+\]
+
+The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
+
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+The {\em plan length} is denoted with $|\pi|=n$ .
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
+
+\subsection{\strips\ action schemas}
+This work addresses the learning of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+
+To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
+Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
+Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
+$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
+For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
+We define $F_v$, a new set of fluents s.t.
+$F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and defines the elements that can appear in an action schema.
+For the {\em blocksworld}, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+
+We assume also that actions $a\in A$ are instantiated from \strips\ operator schemas $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+\begin{itemize}
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
+The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
+\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\end{itemize}
+
+Finally, given the set of predicates $\Psi$ and the header of a \strips\ operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in the action schema $\xi$ and that confine its space of possible action models.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.
+
+\begin{figure}
+\begin{footnotesize}
+\begin{verbatim}
+(:action stack
+ :parameters (?v1 ?v2 - object)
+ :precondition (and (holding ?v1) (clear ?v2))
+ :effect (and (not (holding ?v1))
+              (not (clear ?v2))
+              (handempty) (clear ?v1)
+              (on ?v1 ?v2)))
+\end{verbatim}
+\end{footnotesize}
+ \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
+\label{fig:stack}
+\end{figure}
+
+\section{Learning \strips\ action models}
+\label{sec:Section4}
+Learning \strips\ action models from fully available input knowledge, i.e.
+from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
+When any intermediate state is available, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions~\cite{jimenez2012review}.
+
+This section formalizes more challenging learning tasks, where less input knowledge is available:
+
+\subsubsection{Learning from (initial, final) state pairs.} This learning task corresponds to observing an agent acting in the world but watching only the result of its plan executions.
+No intermediate information about the actions in the plans is given.
+This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma}$:
+\begin{itemize}
+\item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
+\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
+Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
+\end{itemize}
+
+\subsubsection{Learning from labeled plans.}
+Here we augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
+
+\begin{itemize}
+\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+\end{itemize}
+
+Figure~\ref{fig:lexample} shows an example of a {\em blocksworld} learning task $\Lambda'$.
+This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
+
+\subsubsection{Learning from partially specified action models.}
+We may not require to start learning from scratch so we augment the learning input with partially specified operator schemes. This leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
+\begin{itemize}
+\item $\Xi_0$ is a partially specified action model, i.e. some preconditions and effects are a priori known.
+\end{itemize}
+
+A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given labels $\Sigma$ using the learned \strips\ model.
+A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
+Finally a solution to $\Lambda''$ is a set of \strips\ operator schema $\Xi$ compliant as well with the provided partially specified action model $\Xi_0$.
+
+\begin{figure}
+{\tt ;;; Predicates in $\Psi$}
+\begin{footnotesize}
+\begin{verbatim}
+(handempty) (holding ?o  - object)
+(clear ?o - object) (ontable ?o - object)
+(on ?o1 - object ?o2 - object)
+\end{verbatim}
+\end{footnotesize}
+
+\vspace{0.2cm}
+
+\begin{subfigure}{.25\textwidth}
+{\tt ;;; Plan $\pi_1$}
+\begin{footnotesize}
+\begin{verbatim}
+0: (unstack A B)
+1: (putdown A)
+2: (unstack B C)
+3: (stack B A)
+4: (unstack C D)
+5: (stack C B)
+6: (pickup D)
+7: (stack D C)
+\end{verbatim}
+\end{footnotesize}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
+\begin{lstlisting}[mathescape]
+\end{lstlisting}
+\vspace{0.1cm}
+\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
+\node (initD) [draw]{D};
+\node (initC) [draw, above=of initD.north]{C};
+\node (initB) [draw, above=of initC.north]{B};
+\node (initA) [draw, above=of initB.north]{A};
+\draw[thick] (-1,-0.25) -- (2.5,-0.25);
+
+\node (goalA) [draw, right=10mm of initD]{A};
+\node (goalB) [draw, right=10mm of initC]{B};
+\node (goalC) [draw, right=10mm of initB]{C};
+\node (goalD) [draw, right=10mm of initA]{D};
+\end{tikzpicture}
+\vspace{0.6cm}
+\end{subfigure}%
+ \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
+\label{fig:lexample}
+\end{figure}
+
+%Learning \strips\ action models from plans where every action in the plan is available, as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
+In this case, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+Preconditions are derived lifting the minimal set of literals that appears in all the corresponding pre-states.
+
+This paper addresses the learning task that corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
+This learning task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$:
+\begin{itemize}
+\item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
+\item $\Psi$ is the set of predicates, that define the abstract state space of a given classical planning frame.
+\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$.
+\end{itemize}
+
+A solution to $\Lambda$ is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the state observation sequence $\mathcal{O}$.
+Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
+
+\begin{figure}
+{\footnotesize\tt ;;;;;; Headers in $\mathcal{M}$}
+\begin{footnotesize}
+\begin{verbatim}
+(pickup v1) (putdown v1)
+(stack v1 v2} (unstack v1 v2)
+\end{verbatim}
+\end{footnotesize}
+\vspace{0.2cm}
+{\footnotesize\tt ;;;;;; Predicates $\Psi$}
+\begin{footnotesize}
+\begin{verbatim}
+(handempty) (holding ?o  - object)
+(clear ?o - object) (ontable ?o - object)
+(on ?o1 - object ?o2 - object)
+\end{verbatim}
+\end{footnotesize}
+\vspace{0.2cm}
+{\footnotesize\tt ;;;;;; Observations $\mathcal{O}$}
+\begin{footnotesize}
+\begin{verbatim}
+;;; observation #0
+(clear block2) (on block2 block1)
+(ontable block1) (handempty)
+
+;;; observation #1
+(holding block2) (clear block1) (ontable block1)
+
+;;; observation #2
+(clear block1) (ontable block1)
+(clear block2) (ontable block2) (handempty)
+
+;;; observation #3
+(holding block1) (clear block2) (ontable block2)
+
+;;; observation #4
+(clear block1) (on block1 block2)
+(ontable block2) (handempty)
+\end{verbatim}
+\end{footnotesize}
+ \caption{\small Example of a $\Lambda$ task for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
+\label{fig:lexample}
+\end{figure}
 

### 2018-02-21 doc/aij18/aij18.tex
serctio 4 added to AIJ18 paper

@@
 \label{sec:Section3}
+This section defines the planning model used on this work and the output of the learning tasks addressed in the paper.
+
 \subsection{Classical planning with conditional effects}
@@
 Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions~\cite{jimenez2012review}.
+This section formalizes a set of more challenging action model learning tasks, where less input knowledge is available.
 
-This section formalizes more challenging learning tasks, where less input knowledge is available:
-
-\subsubsection{Learning from (initial, final) state pairs.} This learning task corresponds to observing an agent acting in the world but watching only the result of its plan executions.
-No intermediate information about the actions in the plans is given.
-This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma}$:
+\subsection{Learning from state observations} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
+This learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\sigma}$, when no intermediate information is given:
 \begin{itemize}
+\item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
+In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but partially specified operator schemes where some preconditions and effects are a priori known.
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
-Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
+\item  $\sigma=(s_0,s_{n})$ is a  $(initial, final)$ state pair, that we call {\em label}.
+the {\em final} state $s_{n}$ resulting from executing an unknown plan $\pi_t=\tup{a_1, \ldots, a_n}$ starting from the {\em initial} state $s_0$.
+When the intermediate states are available, the learning task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$, where $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$.
 \end{itemize}
 
-\subsubsection{Learning from labeled plans.}
-Here we augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
+A solution is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the state observation sequence $\mathcal{O}$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plan $\pi$, that explain the given observations using the learned \strips\ model.
+Figure~\ref{fig:lexample} shows a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
 
-\begin{itemize}
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-\end{itemize}
-
-Figure~\ref{fig:lexample} shows an example of a {\em blocksworld} learning task $\Lambda'$.
-This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
-
-\subsubsection{Learning from partially specified action models.}
-We may not require to start learning from scratch so we augment the learning input with partially specified operator schemes. This leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
-\begin{itemize}
-\item $\Xi_0$ is a partially specified action model, i.e. some preconditions and effects are a priori known.
-\end{itemize}
-
-A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given labels $\Sigma$ using the learned \strips\ model.
-A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
-Finally a solution to $\Lambda''$ is a set of \strips\ operator schema $\Xi$ compliant as well with the provided partially specified action model $\Xi_0$.
-
-\begin{figure}
-{\tt ;;; Predicates in $\Psi$}
-\begin{footnotesize}
-\begin{verbatim}
-(handempty) (holding ?o  - object)
-(clear ?o - object) (ontable ?o - object)
-(on ?o1 - object ?o2 - object)
-\end{verbatim}
-\end{footnotesize}
-
-\vspace{0.2cm}
-
-\begin{subfigure}{.25\textwidth}
-{\tt ;;; Plan $\pi_1$}
-\begin{footnotesize}
-\begin{verbatim}
-0: (unstack A B)
-1: (putdown A)
-2: (unstack B C)
-3: (stack B A)
-4: (unstack C D)
-5: (stack C B)
-6: (pickup D)
-7: (stack D C)
-\end{verbatim}
-\end{footnotesize}
-\end{subfigure}%
-\begin{subfigure}{.6\textwidth}
-{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
-\begin{lstlisting}[mathescape]
-\end{lstlisting}
-\vspace{0.1cm}
-\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
-\node (initD) [draw]{D};
-\node (initC) [draw, above=of initD.north]{C};
-\node (initB) [draw, above=of initC.north]{B};
-\node (initA) [draw, above=of initB.north]{A};
-\draw[thick] (-1,-0.25) -- (2.5,-0.25);
-
-\node (goalA) [draw, right=10mm of initD]{A};
-\node (goalB) [draw, right=10mm of initC]{B};
-\node (goalC) [draw, right=10mm of initB]{C};
-\node (goalD) [draw, right=10mm of initA]{D};
-\end{tikzpicture}
-\vspace{0.6cm}
-\end{subfigure}%
- \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
-\label{fig:lexample}
-\end{figure}
-
-%Learning \strips\ action models from plans where every action in the plan is available, as well as its corresponding {\em pre-} and {\em post-states}, is straightforward.
-In this case, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Preconditions are derived lifting the minimal set of literals that appears in all the corresponding pre-states.
-
-This paper addresses the learning task that corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
-This learning task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$:
-\begin{itemize}
-\item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
-\item $\Psi$ is the set of predicates, that define the abstract state space of a given classical planning frame.
-\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$.
-\end{itemize}
-
-A solution to $\Lambda$ is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the state observation sequence $\mathcal{O}$.
-Figure~\ref{fig:lexample} shows a $\Lambda$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
-
-\begin{figure}
+\begin{figure}[hbt!]
 {\footnotesize\tt ;;;;;; Headers in $\mathcal{M}$}
@@
 ;;; observation #0
-(clear block2) (on block2 block1)
-(ontable block1) (handempty)
+(clear B) (on B A) (ontable A) (handempty)
 
 ;;; observation #1
-(holding block2) (clear block1) (ontable block1)
+(holding B) (clear A) (ontable A)
 
 ;;; observation #2
-(clear block1) (ontable block1)
-(clear block2) (ontable block2) (handempty)
+(clear A) (ontable A) (clear B) (ontable B) (handempty)
 
 ;;; observation #3
-(holding block1) (clear block2) (ontable block2)
+(holding A) (clear B) (ontable B)
 
 ;;; observation #4
-(clear block1) (on block1 block2)
-(ontable block2) (handempty)
+(clear A) (on blockA B) (ontable B) (handempty)
 \end{verbatim}
 \end{footnotesize}
- \caption{\small Example of a $\Lambda$ task for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
+ \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ task for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
 \label{fig:lexample}
 \end{figure}
+
+\subsection{Learning from a labeled plan}
+Here we augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$:
+
+\begin{itemize}
+\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence that induces the corresponding state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates $s_i=\theta(s_{i-1},a_i)$.
+\end{itemize}
+
+Figure~\ref{fig:lexample} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, that corresponds to observing the execution of an eight-action plan for inverting a four-block tower.
+
+\begin{figure}
+{\tt ;;; Predicates in $\Psi$}
+\begin{footnotesize}
+\begin{verbatim}
+(handempty) (holding ?o  - object)
+(clear ?o - object) (ontable ?o - object)
+(on ?o1 - object ?o2 - object)
+\end{verbatim}
+\end{footnotesize}
+
+\vspace{0.2cm}
+
+\begin{subfigure}{.25\textwidth}
+{\tt ;;; Plan $\pi$}
+\begin{footnotesize}
+\begin{verbatim}
+0: (unstack A B)
+1: (putdown A)
+2: (unstack B C)
+3: (stack B A)
+4: (unstack C D)
+5: (stack C B)
+6: (pickup D)
+7: (stack D C)
+\end{verbatim}
+\end{footnotesize}
+\end{subfigure}%
+\begin{subfigure}{.6\textwidth}
+{\tt ;;; Label $\sigma=(s_0^1,s_{n}^1)$}
+\begin{lstlisting}[mathescape]
+\end{lstlisting}
+\vspace{0.1cm}
+\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
+\node (initD) [draw]{D};
+\node (initC) [draw, above=of initD.north]{C};
+\node (initB) [draw, above=of initC.north]{B};
+\node (initA) [draw, above=of initB.north]{A};
+\draw[thick] (-1,-0.25) -- (2.5,-0.25);
+
+\node (goalA) [draw, right=10mm of initD]{A};
+\node (goalB) [draw, right=10mm of initC]{B};
+\node (goalC) [draw, right=10mm of initB]{C};
+\node (goalD) [draw, right=10mm of initA]{D};
+\end{tikzpicture}
+\vspace{0.6cm}
+\end{subfigure}%
+ \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a labeled plan.}
+\label{fig:lexample}
+\end{figure}
+
+\subsection{Learning from multiple plans}
+The previous task define the learning of planning action models from a single plan execution. These definitions can be extended to the more general case where learning from multiple plans:
+\begin{itemize}
+\item $\Lambda=\tup{\mathcal{M},\Psi,\Sigma}$ where $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
+Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
+\item $\Lambda=\tup{\mathcal{M},\Psi,\Sigma,\Pi}$ where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+\end{itemize}
+
+\section{Evaluation}
+\label{sec:Section6}
+
+\section{Conclussions}
+\label{sec:Section7}
 

### 2018-02-21 doc/aij18/aij18.tex
section 6 added to AIJ18 paper

@@
 
+\section{Learning \strips\ action models with classical planning}
+Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
+A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains $\mathcal{O}$.
+The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
+
+\begin{enumerate}
+\item {\bf Programs the action model $\mathcal{M}'$}.
+A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item {\bf Validates the action model $\mathcal{M}'$ in $\mathcal{O}$}.
+The solution plan continues with a postfix that produces the given sequence of states $\tup{s_0,s_1,\ldots,s_{n}}$ using the programmed action model $\mathcal{M}'$.
+\end{enumerate}
+
+\subsection{Learning from state observations}
+Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+\begin{itemize}
+\item $F_{\Lambda}$ contains:
+\begin{itemize}
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$, i.e.
+{\tt\small block1} and {\tt\small block2} in Figure~\ref{fig:lexample}.
+\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
+For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
+\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemas are programmed or validated and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
+\end{itemize}
+\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
+Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
+Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
+\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
+\item $A_{\Lambda}$ comprises three kinds of actions:
+\begin{enumerate}
+\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$:
+\begin{itemize}
+\item Actions for {\bf removing} a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
+& mode_{prog}, pre_{f}(\xi)\},\\
+\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
+\end{align*}
+\end{small}
+
+\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
+& mode_{prog}\},\\
+\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
+&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
+\end{align*}
+\end{small}
+\end{itemize}
+
+\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
+Given that the operators headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
+&\cup \{\neg mode_{val}\},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
+&\{\emptyset\}\rhd\{mode_{val}\}.
+\end{align*}
+\end{small}
+
+\item Actions for {\em validating} an observation {\tt\small $1\leq i\leq n$}.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\\
+&\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
+\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
+\end{align*}
+\end{small}
+\end{enumerate}
+\end{itemize}
+
+\subsubsection{Compilation properties}
+
+\begin{mylemma}
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$.
+\end{mylemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
+In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
+These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence {\small $1\leq i\leq n$}.
+This means that the programmed action model $\mathcal{M}'$ complies with the provided observations $\mathcal{O}$ and hence, solves $\Lambda$.
+\end{small}
+\end{proof}
+
+\begin{mylemma}
+Completeness.
+Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+\end{mylemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O}$.
+\end{small}
+\end{proof}
+
+The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
+\begin{itemize}
+\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
+The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
+\item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+\end{itemize}
+
+\subsection{Learning from a labeled plan}
+
+\subsection{Learning from multiple plans}
+
+\section{Evaluating \strips\ action models}
+\label{model_evaluation}
+
+We assess how well a \strips\ action model $\mathcal{M}$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} required by $\mathcal{M}$ to induce $\mathcal{O}$.
+%In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
+
+\subsection{Edition of \strips\ action models}
+
+We first define the allowed \emph{operations} to edit a given \strips\ action model.
+With the aim of keeping a tractable branching factor of the planning instances that results from our compilation, we only define two {\em edit operations}:
+\begin{itemize}
+\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, $f\in F_v(\xi)$.
+\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, $f\in F_v(\xi)$.
+\end{itemize}
+
+We can now formalize an edit distance that quantifies how dissimilar two given \strips\ action models are.
+The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations} have the same positive cost.
+
+\begin{mydefinition}
+Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, both built from the same set of possible elements $F_v$.
+The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} required to transform $\mathcal{M}$ into $\mathcal{M}'$.
+\end{mydefinition}
+
+Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well.
+In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
+
+\begin{mydefinition}
+The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3|F_v(\xi)|$.
+\end{mydefinition}
+
+\subsection{The observation edit distance}
+
+We define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
+
+\begin{mydefinition}
+Given an action model $\mathcal{M}$, built from $F_v$, and the observations sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$.
+The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
+\end{mydefinition}
+
+The $\Lambda$ learning tasks can swap the roles of two operators whose headers match or two action parameters that belong to the same type (e.g.
+the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped).
+The {\em observation edit distance} is a semantic measure that is robust to role changes of this kind.
+
+Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning}, the {\em observation edit distance} assesses, with a single expression, the flaws in the preconditions and effects of a given learned model.
+This fact enables the recognition of \strips\ action models.
+The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
+The {\em observation edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
+
+%Given a set of possible \strips\ models and a sequence of state observations, the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
+According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
+In our scenario, the hypotheses are about the possible \strips\ action models that can be built within a given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets).
+With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
+%\begin{enumerate}
+%\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$.
+%\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
+%\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
+%\end{enumerate}
+
+\subsection{Computing the observation edit distance}
+
+Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$, is {\em non-empty}.
+In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+A solution to the planning task resulting from the extended compilation is a sequence of actions that:
+
+\begin{enumerate}
+\item {\bf Edits the action model $\mathcal{M}$}.
+A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemas in $\mathcal{M}$ using to the two {\em edit operations}.
+\item {\bf Validates the edited model $\mathcal{M}'$ in $\mathcal{O}$}.
+The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$, as in Section~\ref{learning_task}.
+\end{enumerate}
+
+Now $\Lambda$ is not a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
+The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
+
+\begin{itemize}
+\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
+\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
+In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
+This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
+\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
+The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
+The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
+include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
+\end{itemize}
+
+To illustrate this, the plan of Figure~\ref{fig:plan} solves the classical planning task $P_{\Lambda}'$ that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:lexample}.
+
+Our interest when computing the {\em observation edit distance} is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
+$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan}.
+The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
+
+\begin{figure}
+{\tt\small
+00 : (insert\_add\_handempty\_stack)\\
+01 : (insert\_add\_clear\_stack\_var1)\\
+02 : (apply\_unstack block2 block1)\\
+03 : (validate\_1)\\
+04 : (apply\_putdown block2)\\
+05 : (validate\_2)\\
+06 : (apply\_pickup block1)\\
+07 : (validate\_3)\\
+08 : (apply\_stack block1 block2)\\
+09 : (validate\_4)\\
+}
+ \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:lexample}.}
+\label{fig:plan}
+\end{figure}
+
 \section{Evaluation}
-\label{sec:Section6}
+\label{sec:Section7}
+This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts of available input knowledge.
+
+\subsection{Setup}
+The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
+All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
+\begin{itemize}
+\item {\bf Reproducibility}.
+We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
+\item {\bf Planner}.
+The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
+We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
+In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
+Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
+\item {\bf Metrics}.The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
+These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
+Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
+Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
+Recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
+\end{itemize}
 
 \section{Conclussions}
-\label{sec:Section7}
+\label{sec:Section8}
 

### 2018-02-21 doc/aij18/aij18.tex
updates to sections 4 and 5

@@
 Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
-A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains $\mathcal{O}$.
+A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains the inputs to the lerning task.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \item {\bf Validates the action model $\mathcal{M}'$ in $\mathcal{O}$}.
-The solution plan continues with a postfix that produces the given sequence of states $\tup{s_0,s_1,\ldots,s_{n}}$ using the programmed action model $\mathcal{M}'$.
+The solution plan continues with a postfix that reproduces the given input knowledge using the programmed action model $\mathcal{M}'$.
 \end{enumerate}
@@
 \subsection{Learning from a labeled plan}
+We extend the compilation to consider labeled plans.
+Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$:
+\begin{itemize}
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
+Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$}, for indicating the plan step where the action model is validated.
+Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
+\item With respect to $A_{\Lambda'}$.
+\begin{enumerate}
+\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
+\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+This mechanism forces that these actions are only applied as in the example plans.
+\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
+\begin{small}
+\begin{align*}
+&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}, \{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
+\end{align*}
+\end{small}
+\end{enumerate}
+\end{itemize}
+
+\subsection{Partially specified action models}
+Known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda'}$.
+The corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
+
+To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,22]$ validate the programmed operators following the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
+
+In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
+In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips\ action model is {\em valid} for the given examples.
+If $P_{\Lambda'}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
+Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
+
+\begin{figure}[hbt!]
+{\footnotesize\tt
+     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
+     01 : (program\_pre\_handempty\_stack)\\
+     02 : (program\_pre\_holding\_stack\_v2)\\
+     03 : (program\_pre\_on\_stack\_v1\_v1)\\
+     04 : (program\_pre\_on\_stack\_v1\_v2)\\
+     05 : (program\_pre\_on\_stack\_v2\_v1)\\
+     06 : (program\_pre\_on\_stack\_v2\_v2)\\
+     07 : (program\_pre\_ontable\_stack\_v1)\\
+     08 : (program\_pre\_ontable\_stack\_v2)\\
+     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
+    10 : (program\_eff\_clear\_stack\_v2)\\
+    11 : (program\_eff\_handempty\_stack)\\
+    12 : (program\_eff\_holding\_stack\_v1)\\
+    13 : (program\_eff\_on\_stack\_v1\_v2)\\
+    {\bf 14} : (apply\_unstack a b i1 i2)\\
+    15 : (apply\_putdown a i2 i3)\\
+    16 : (apply\_unstack b c i3 i4)\\
+    17 : (apply\_stack b a i4 i5)\\
+    18 : (apply\_unstack c d i5 i6)\\
+    19 : (apply\_stack c b i6 i7)\\
+    20 : (apply\_pickup d i7 i8)\\
+    21 : (apply\_stack d c i8 i9)\\
+    {\bf 22} : (validate\_1)
+}
+ \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi_1$ and label $\sigma_1$ (shown in Figure~\ref{fig:lexample}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+\label{fig:plan}
+\end{figure}
+
+\subsection{Static predicates}
+A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
+Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
+Given a static predicate $p$:
+\begin{itemize}
+\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable objects} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
+\item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
+\end{itemize}
+
+Static predicates can also constrain the space of possible preconditions by looking at the given set of labels $\Sigma$.
+One can assume that if a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
+For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\Sigma$.
+
+Looking at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
+Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
 
 \subsection{Learning from multiple plans}
+The $\Lambda=\tup{\mathcal{M},\Psi,\Sigma}$ can be solved with the previous compilation $P_{\Lambda}$.
+On the other hand, the learning task $\Lambda=\tup{\mathcal{M},\Psi,\Sigma,\Pi}$ requires modifying the previous compilation.
+In particular, the actions for {\em validating} the labeled plan $\pi_t\in\Pi$ with the corresponding label $\sigma_t\in\Sigma$, {\tt\small $1\leq t\leq \tau$}.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\\
+&\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
+\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
+\end{align*}
+\end{small}
 
@@
 
-\subsection{The observation edit distance}
+\subsection{The \strips\ edit distance}
 
@@
 
-%Given a set of possible \strips\ models and a sequence of state observations, the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
-According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the possible \strips\ action models that can be built within a given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets).
-With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
-%\begin{enumerate}
-%\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$.
-%\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
-%\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
-%\end{enumerate}
+The edit distance can also be defined with respect to a labeled plan.
 
-\subsection{Computing the observation edit distance}
+\begin{mydefinition}
+  Given an action model $\mathcal{M}$, built from $F_v$, and the plan $\pi=\tup{a_1, \ldots, a_n}$  with the corresponding label $\sigma=(s_0,s_{n})$.
+The {\bf plan edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{\pi},\mathcal{\sigma})$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce a plan $\pi$ that induces $\sigma$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{\sigma}} \delta(\mathcal{M},\mathcal{M}')\]
+\end{mydefinition}
+
+Last but not least, the edit distance can also be defined with respect to multiple plans.
+
+\subsection{Computing the edit distance}
 
@@
 \item {\bf Validates the edited model $\mathcal{M}'$ in $\mathcal{O}$}.
-The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$, as in Section~\ref{learning_task}.
+The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$, as in Section~\ref{sec:Section4}.
 \end{enumerate}

### 2018-02-22 doc/icaps-18/camera-ready/strips-learning.tex
versión actualizada del camera-ready

@@
 \begin{abstract}
-This paper presents a novel approach for learning \strips\ action models from examples that compiles this inductive learning task into classical planning.
+This paper presents a novel approach for learning \strips\ action models from examples that compiles this inductive learning task into a classical planning task.
 Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states (no intermediate action or state is given).
-What is more, the compilation accepts partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
+Moreover, the compilation accepts partially specified action models and it can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
 \end{abstract}
@@
 The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
-\item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning \strips\ action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions.
+\item The {\em output} of ML algorithms is usually a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
+When learning action models the output is, for each action, the preconditions, negative and positive effects that define the possible state transitions.
 \end{itemize}
 
-Learning \strips\ action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
+Learning \strips\ action models is a well-studied problem with sophisticated algorithms such as {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}, which do not require full knowledge of the intermediate states traversed by the example plans.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative planning compilation approach for learning \strips\ action models.
+The compilation approach is appealing by itself because it opens the door to the bootstrapping of planning action models, but also because:
 
-The compilation approach is appealing by itself, because opens the door to the bootstrapping of planning action models, but also because:
 \begin{enumerate}
-\item Is flexible to various amounts of available input knowledge.
-Learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate state  or action is observed.
-\item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
+\item It is flexible to various amounts of available input knowledge.
+Learning examples range from a set of plans (with their corresponding initial and final states) to just a pair of initial and final states where no intermediate state or action is observed.
+\item It accepts previous knowledge about the structure of the actions in the form of partially specified action models.
 In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
@@
 The second section of the paper formalizes the classical planning model, its extension to {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the output of the addressed learning task).
-The third section formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
-The fourth and fifth sections describe our compilation approach for addressing the formalized learning tasks.
-Finally, the last sections report the data collected in a empirical evaluation, discuss the strengths and weaknesses of the compilation approach and propose several opportunities for future research.
+The third section formalizes the task of learning action models with different amounts of available input knowledge.
+The fourth and fifth sections describe our compilation approach to tackle the formalized learning tasks.
+Finally, the last sections show the experimental evaluation, discuss the strengths and weaknesses of the compilation approach and propose several opportunities for future research.
 
 \section{Background}
-This section defines the planning model used on this work and the output of the learning tasks addressed in the paper.
+This section defines the planning model and the output of the learning tasks addressed in the paper.
 
@@
 We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
-A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
+A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e.
 either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent).
-We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
+A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values).
+We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.
 
-A {\em state} $s$ is a full assignment of values to fluents, i.e.
-$|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
+A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$, so the size of the state space is $2^{|F|}$.
+Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often we will abuse of notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.
 
@@
 
-An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)$ and a set of {\em conditional effects} $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C \subseteq \mathcal{L}(F)$, the {\em condition}, and $E \subseteq \mathcal{L}(F)$, the {\em effect}.
+An action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
 \[
@@
 The {\em plan length} is denoted with $|\pi|=n$ .
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the application of the plan $\pi$ in the initial state $I$.
 
-\subsection{\strips\ action schemes and {\em variable name} objects}
-This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-Figure~\ref{fig:stack} shows the {\em stack} schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+\subsection{\strips\ action schemas and {\em variable name} objects}
+We address learning of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+Figure~\ref{fig:stack} shows the {\em stack} schema of a four-operator {\em blocksworld} domain~\cite{slaney2001blocks} encoded in PDDL.
 
@@
 \end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from {\em blocksworld}.}
+ \caption{\small The {\em stack} operator schema of the {\em blocksworld} domain specified in PDDL.}
 \label{fig:stack}
@@
 Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$, where $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
 
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects $\Omega\cap\Omega_v=\emptyset$, denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters each.
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects denoted as {\em variable names} ($\Omega\cap\Omega_v=\emptyset$).
+$\Omega_v$ is bound to the maximum arity of an action in a given planning frame.
+For instance, in a three-block \emph{blocksworld}, $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters each.
 
-Let us also define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ exclusively using objects in $\Omega_v$ and that defines the elements that can appear in the preconditions and effects of an action schema.
-For instance, in the blocksworld, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+Let $F_v$ be a new set of fluents, $F\cap F_v=\emptyset$, that results from instantiating the predicates in $\Psi$ using exclusively  objects of $\Omega_v$.
+$F_v$ defines the elements of the preconditions and effects of an action schema.
+For instance, in the \emph{blocksworld} domain, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
-Finally, we assume that actions $a\in A$ are instantiated from \strips\ operator schemes $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+Finally, we assume that an action $a\in A$ is instantiated from a \strips\ an operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
-For instance, the headers for a four-operator blocksworld are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
-\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$ is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
+For instance, the headers of a four-operator \emph{blocksworld} domain are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
+\item $pre(\xi)\subseteq F_v$ is the set of preconditions,  $del(\xi)\subseteq F_v$ the negative effects and  $add(\xi)\subseteq F_v$ the positive effects such tha $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
 \end{itemize}
@@
 \section{Learning \strips\ action models}
+
 Learning \strips\ action models from fully available input knowledge, i.e.
-from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
-When any intermediate state is available, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions~\cite{jimenez2012review}.
+from plans where the {\em pre-} and {\em post-states} of every action in the plans are available, is straightforward.
+When intermediate states are available, operator schemas are derived lifting the literals that change between the pre and post-state of each action execution.
+Preconditions of an action are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action~\cite{jimenez2012review}.
 
@@
 
-\subsubsection{Learning from (initial, final) state pairs.} This learning task corresponds to observing an agent acting in the world but watching only the result of its plan executions.
+\subsubsection{Learning from (initial, final) state pairs.} This learning task amounts to observing agents acting in the world but watching only the result of their plans execution.
 No intermediate information about the actions in the plans is given.
@@
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
-Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
+\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs called {\em labels}.
+Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ in the {\em initial} state $s_0^t$.
 \end{itemize}
@@
 \subsubsection{Learning from labeled plans.}
-Here we augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
+We augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
 
 \begin{itemize}
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that $a_i^t$, {\small $1\leq i\leq n$}, is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
 \end{itemize}
 
-Figure~\ref{fig:lexample} shows an example of a {\em blocksworld} learning task $\Lambda'$.
-This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
+Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda'$ of the {\em blocksworld} domain.
+This task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
 
 \subsubsection{Learning from partially specified action models.}
-We may not require to start learning from scratch so we augment the learning input with partially specified operator schemes. This leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
+In case that partially specified operator schemas are available, we can incorporate this information within the learning task. This leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
+
 \begin{itemize}
-\item $\Xi_0$ is a partially specified action model, i.e. some preconditions and effects are a priori known.
+\item $\Xi_0$ is a partially specified action model in whcih some preconditions and effects are known a priori.
 \end{itemize}
 
-A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given labels $\Sigma$ using the learned \strips\ model.
-A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
-Finally a solution to $\Lambda''$ is a set of \strips\ operator schema $\Xi$ compliant as well with the provided partially specified action model $\Xi_0$.
+A solution to $\Lambda$ is a set of operator schemas $\Xi$ that is compliant with the predicates in $\Psi$ and the set of initial and final states $\Sigma$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$}, that explain the given labels $\Sigma$ using the learned model.
+A solution to $\Lambda'$ is a set of \strips\ operator schemas $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action that has a different name in the example plans $\Pi$) that is compliant with the predicates $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
+Finally, a solution to $\Lambda''$ is a set of \strips\ operator schemas $\Xi$ that is also compliant with the provided partially specified action model $\Xi_0$.
 
@@
 \end{subfigure}%
- \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
+ \caption{\small Learning task of the \emph{blocksworld} domain from a single labeled plan.}
 \label{fig:lexample}
@@
 \section{Learning \strips\ action models with planning}
-Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is compiling it into a classical planning task with conditional effects.
+In our approach, a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$ is solved by compiling it into a classical planning task with conditional effects.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
+
 \begin{enumerate}
 \item {\em Programs the \strips\ action model $\Xi$}.
-A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item {\em Validates the programmed \strips\ action model $\Xi$ in the given input knowledge} (the labels $\Sigma$, and $\Pi$ and/or $\Xi_0$ if available).
-For every label $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
-We call this process the validation of the programmed \strips\ action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
+A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
+\item {\em Validates the programmed \strips\ action model $\Xi$ in the given input knowledge} (the labels $\Sigma$ and $\Pi$, and/or $\Xi_0$ if available).
+For every label $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ using the programmed action model $\Xi$ in the corresponding initial state $s_0^t$.
+This process is the validation of the programmed action model $\Xi$ with the set of learning examples {\small $1\leq t\leq \tau$}.
 %If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
@@
 
-To formalize our compilation we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
-same fluents and actions but differ in the initial state and goals).
-Fluents $F$ are built instantiating the predicates in $\Psi$ with the objects appearing in the input labels $\Sigma$.
-Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
+To formalize our compilation we first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
+same fluents and actions but different initial states and goals).
+Fluents $F$ are built instantiating the predicates in $\Psi$ with the objects of the input labels $\Sigma$.
+Formally, $\Omega=\bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
 The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.
+Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ and goals $G_t$ are defined by the state $s_n^t\in \sigma_t$.
 
-Now we are ready to formalize the compilations.
-We start with $\Lambda$, because it requires less input knowledge.
-Given a learning task $\Lambda=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+We can now formalize the compilation approach.
+We start with $\Lambda$ as it requires the least input knowledge.
+Given a learning task $\Lambda=\tup{\Psi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+
 \begin{itemize}
@@
 \item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$.
+If a fluent of $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect of the operator schema $\xi\in \Xi$.
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$}.
-\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is being validated.
+\item A fluent $mode_{prog}$ to indicate whether the operator schemas are being programmed or validated (already programmed), and fluents $\{test_t\}_{1\leq t\leq \tau}$ which represent the example where the action model is being validated.
 \end{itemize}
 \item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
-Our compilation assumes that initially any operator schema is programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
-\item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
+Our compilation assumes that any operator schema is initially programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
+\item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$ indicates that the programmed action model is validated in all the learning examples.
 \item $A_{\Lambda}$ contains actions of three kinds:
@@
 We assume operators headers are known so the binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
-variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
+variables $pars(\xi)$ are bound to the objects in $\omega$ that appear in the same position.
 Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\\
-&\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
+\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{1\leq j<t}\\
+&\cup\{\neg test_j\}_{t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
+\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\} \cup \{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\\
+&\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t}.
 \end{align*}
@@
 Once operator schemas $\Xi$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
+To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved by executing an applicable sequence of programmed operator schemas that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
 If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
@@
 \begin{itemize}
-\item The arity of the actions headers in $\Xi$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
-The larger these numbers, the larger the $F_v$ set, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
+\item The arity of the actions headers in $\Xi$ and the predicates $\Psi$ of the learning task.
+The larger the arity, the larger the $F_v$ set, which in turn defines the size of the fluent sets $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ and the corresponding set of {\em programming} actions.
 \item The number of learning examples. The larger this number, the more $test_t$ fluents and $\mathsf{validate_{t}}$ actions in $P_{\Lambda}$.
@@
 \end{scriptsize}
- \caption{\small Action for applying an already programmed schema $stack$ as encoded in PDDL (implications coded as disjunctions).}
+ \caption{\small PDDL action for applying an already programmed schema $stack$ (implications coded as disjunctions).}
 \label{fig:compilation}
@@
 \label{sec:Constraining}
-Here we show that further input knowledge can be used to constrain the space of possible action models and make the learning of \strips\ action models more practicable.
+In this section we show that further input knowledge can be used to constrain the space of possible action models and to make the learning task more practicable.
 
@@
 Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$}, for indicating the plan step where the action model is validated.
-Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$} to indicate the plan step where the action model is validated.
+As in the original compilation, $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
 \item With respect to $A_{\Lambda'}$.
@@
 \item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$ that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
 This mechanism forces that these actions are only applied as in the example plans.
@@
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
+To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the {\em blocksworld} action model where operator schemas for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
 This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
@@
 }
- \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi_1$ and label $\sigma_1$ (shown in Figure~\ref{fig:lexample}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+ \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi_1$ and label $\sigma_1$ (shown in Figure~\ref{fig:lexample}) as well as previously specified operator schemas for $pickup$, $putdown$ and $unstack$.}
 \label{fig:plan}
@@
 \subsection{Learning from labeled plans}
-We start evaluating our approach at $\Lambda'=\tup{\Psi,\Sigma,\Pi}$ learning tasks, where {\em labeled plans} are available, and repeat the evaluation but exploiting potential static predicates computed from $\Sigma$.
-The potential {\em static predicates} considered here is the set of predicates s.t.
-every predicate instantiation appears unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
+We start evaluating our approach with tasks $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, where {\em labeled plans} are available.
+We then repeat the evaluation but exploiting potential \emph{static predicates} computed from $\Sigma$, which are the predicates that appear  unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
 Static predicates are used to constrain the space of possible action models as explained in the previous section.
 
-Table~\ref{tab:results_plans} summarizes the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
-We can observe that identifying static predicates drives to models with better precondition {\em recall}.
-This fact evidences that many of the missing preconditions corresponded to static predicates because there were no incentive to learn them as they always hold~\cite{gregory2015domain}.
+Table~\ref{tab:results_plans} shows the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), while the last two columns of each setting and the last row report averages values.
+We can observe that identifying static predicates leads to models with better precondition {\em recall}.
+This fact evidences that many of the missing preconditions corresponded to static predicates because there is no incentive to learn them as they always hold~\cite{gregory2015domain}.
 
@@
 
-We evaluate now the ability of our approach to support partially specified action models; that is, when addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
+We evaluate now the ability of our approach to support partially specified action models; that is, addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
 In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task .
 
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, including the identification of static predicates.
-We report only the precision and recall for the {\em unknown} actions (considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in these models is 1.0).
-Likewise, errors in $\Lambda''$ tasks have greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is just done over half of the actions (e.g.
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, which include the identification of static predicates.
+We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
+Likewise, errors in $\Lambda''$ tasks have greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is done only over half of the actions (e.g.
 the preconditions recall for the {\em Floortile} or {\em Gripper} domains).
 
-Remarkably, the overall precision is now $0.98$ which means that the content of the learned models is highly reliable.
-Recall is a lower, 0.87, indicating that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
-Overall the reported results confirm the previous trend: more input knowledge drives to better models and requires less planning time.
-Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
+Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
+The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
+Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
+Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\Xi_0$).
 {\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
@@
 \subsection{Learning from (initial,final) state pairs}
-Finally we evaluate our approach at learning tasks where input plans are not available, so the planner must not only compute the action models but also the plans that satisfy the input labels.
-Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the obtained results using static predicates and partially specified action models so the inputs to this learning task are $\Psi$ ,$\Sigma$ and $\Xi_0$.
-
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} could not solved the corresponding planning tasks within a 1000 secs time bound.
-Precision and recall have now low values.
-The learning hypothesis space is now so low constrained that actions can be reformulated and still be compliant with the inputs (e.g.
-the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
-We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) but still the achieved scores are below the scores obtained when learning from labeled plans (despite no other input knowledge was provided).
+Finally, we evaluate our approach when input plans are not available and thereby the planner must not only compute the action models but also the plans that satisfy the input labels.
+Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the results obtained for the task $\Lambda=\tup{\Psi,\Sigma,\Xi_0}$ using static predicates.
+Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} was not able to solve the corresponding planning tasks within a 1000 sec.
+time bound.
+The values of \emph{precision} and \emph{recall} are significantly lower than in Table ~\ref{tab:results_plans}.
+Given that the learning hypothesis space is now fairly under-constrained, actions can be reformulated and still be compliant with the inputs (e.g.
+the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
+We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) and yet the results are below the scores obtained when learning from labeled plans.
 
@@
 In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
-%and so it may output an inconsistent model
+In contrast to our model comparison validation which aims at covering all the training examples, ARMS maximizes the number of covered examples from a testing set.
+
 {\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
 Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
-The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
+The learning algorithm updates the formula of the belief state with every action and observation in the sequence and thereby the final returned formula includes all consistent models.
+SLAF assesses the quality of the learned models with respect to the actual generative model.
+
+%The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
 This update makes sure that the new formula represents all the transition relations consistent with the actions and observations.
@@
 
-{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
-This makes {\sf LOCM} be most likely the learning approach that works with the least information possible.
-The lack of available information is addressed by LOCM by exploiting assumptions about the kind of domain model it has to generate.
+{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring,cresswell2011generalised}.
+The lack of available information is overcome by exploiting assumptions about the kind of domain model it has to generate.
 Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
-The intuitive assumptions of {\sf LOCM}
-%, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence,
-yield a learning model heavily reliant on the kind of domain structure.
-The inability of {\sf LOCM} to properly derive domain theories where the state of a sort is subject to different FSMs is later overcome by {\sf LOCM2} by forming separate FSMs, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
-{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain}), the last contribution of the {\sf LOCM} family, addresses the problem of inducing static predicates.
-Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
-{\em blocksworld} and {\em freecell}).
-In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
+{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain})  incorporates static predicates and applies a post-processing step after the {\sf LOCM} analysis that requires a set of optimal plans to be used in the learning phase.
+This is done to mitigate the limitation of {\sf LOCM} of inducing similar modes for domains with similar structures.
+{\sf LOP} compares the learned models with the corresponding reference model.
 
 Compiling the learning of action models into classical planning is a general and flexible approach that allows to accommodate various amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
-For instance, the example plans in $\Pi$ could be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), so learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ could also be addressed.
+For instance, the example plans in $\Pi$ can be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), and thus learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ would also be attainable.
 Furthermore, our approach seems extensible to learning other types of generative models (e.g.
@@
 \end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} learning from partially specified action models.}
+\caption{\small {\em Precision} and {\em recall} scores for learning tasks with partially specified action models.}
 \label{tab:results_plans_partial}
@@
         \end{footnotesize}
-	\caption{\small Time and plan length learning from partially specified action models.}
+	\caption{\small Time and plan length learning for learning tasks with partially specified action models.}
 	\label{tab:time_plans_partial}	
@@
 \end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} learning from state pairs.}
+\caption{\small {\em Precision} and {\em recall} scores for learning from (initial,final) state pairs.}
 \label{tab:results_states}
@@
         \end{footnotesize}
-	\caption{\small Time and plan length when learning from state pairs.}
+	\caption{\small Time and plan length when learning from (initial,final) state pairs.}
 	\label{tab:time_states}	
@@
 \begin{small}
-\subsection*{Acknowledgment}
+\subsection*{Acknowledgments}
 This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.

### 2018-02-22 doc/aij18/aij18.tex
updates added to aij18 paper

@@
 
+Our compilation is extensible to compute the {\em edit distance} between two models.
+A solution to the planning task resulting from the extended compilation is a sequence of actions that edits the action model $\mathcal{M}$ to  produce $\mathcal{M}'$ using to the two {\em edit operations}.
+
+The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
+
+\begin{itemize}
+\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
+\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
+In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
+This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
+\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
+The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
+The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
+include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
+\end{itemize}
+
+To illustrate this, the plan of Figure~\ref{fig:plan} solves the classical planning task $P_{\Lambda}'$ that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:lexample}.
+
+Our interest when computing the {\em observation edit distance} is not in $\mathcal{M}'$ but in the number of required {\em edit operations}, e.g.
+$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-edistance}.
+The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
+
+\begin{figure}[hbt!]
+{\tt\small
+00 : (insert\_add\_handempty\_stack)\\
+01 : (insert\_add\_clear\_stack\_var1)
+}
+ \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:lexample}.}
+\label{fig:plan-edistance}
+\end{figure}
+
+\subsection{Computing the observations distance}
+
 Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$, is {\em non-empty}.
@@
 Our interest when computing the {\em observation edit distance} is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
-$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan}.
+$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-odistance}.
 The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
 
-\begin{figure}
+\begin{figure}[hbt!]
 {\tt\small
@@
  \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:lexample}.}
-\label{fig:plan}
+\label{fig:plan-odistance}
 \end{figure}
 
+\section{Recognition of \strips\ action models}
+\label{sec:Section7}
+
 \section{Evaluation}
-\label{sec:Section7}
+\label{sec:Section8}
 This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts of available input knowledge.
@@
 \section{Conclussions}
-\label{sec:Section8}
+\label{sec:Section9}
 

### 2018-02-22 doc/aij18/aij18.tex
added updates to the aij18 paper

@@
 \begin{abstract}
-  This paper presents a novel approach for learning \strips\ action models from examples of plan executions that compiles this learning task into classical planning.
-The compilation approach is flexible to various amount and forms of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) or sequences of state observations to just a set of initial and final states (where no intermediate action or state is given).
+  This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning.
+The compilation approach is flexible to various amount and forms of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) or sequences of state observations, to just a set of initial and final states (where no intermediate action or state is given).
 The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
-What is more, the compilation is extensible to assess how well a \strips\ action model matches a given set of observations.
-Last but not least, the paper evaluates the performance of the compilation approach by learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of observations and (2), the true models.
+What is more, the compilation is extensible to assess how well a \strips\ action model matches observations of plan executions.
+Last but not least, the paper evaluates the performance of the compilation approach by learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of observations of plan executions and (2), the true models.
 \end{abstract}
@@
 
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
 A solution to the classical planning task that results from our compilation is a sequence of actions that determines the learned action model, i.e.
@@
 
-The compilation approach is appealing by itself, because leverages off-the-shelf planners and opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
-The practicality of the compilation approach allow us to report learning results over fifteen IPC planning domains.
-In addition,the compilation approach presents the following features:
+The compilation approach is appealing by itself, because it leverages off-the-shelf planners and because it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
+Moreover, the practicality of the compilation allow us to report learning results over a wide range of IPC planning domains.
+Apart from these, the compilation approach presents the following contributions:
 \begin{enumerate}
-\item Is flexible to various amounts and forms of available input knowledge.
-Learning examples can range from a set of plans (with their corresponding initial and final states) or state observations to just a set of initial and final states where no intermediate state  or action is observed.
-Learning from state observations is a relevant advancement as, in many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
-Learning action models from state observations broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}).
-\item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
+\item {\em Flexibility in the input knowledge}.
+The compilation is flexible to various amounts and forms of available input knowledge.
+Learning examples can range from a set of plans (with their corresponding initial and final states) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
+The compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models.
 In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
-For the training samples, we adopt a middle ground between unstructured inputs and plan traces, wherein only state observations are required.
-\item Assess how well a \strips\ action model matches a given set of observations.
-Our compilation is extensible to accept a learned model as input besides the state observations.
+
+\item {\em Model Evaluation}.
+The compilation can assess how well a \strips\ action model matches a given set of observations of plan executions.
+Our compilation is extensible to accept a learned model as input besides the observations of plan executions.
 This extension allows us to transform the input model into a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
-The empirical evaluation of our learning approach is two-fold: First the learned \strips\ action models are tested with a set of state observation sequences and second, the learned models are compared to the corresponding reference model.
 \end{enumerate}
+
+A first description of the compilation and its extensions previously appeared in several conference papers~\cite{aineto2018learning}.
+Compared to the conference papers, the present paper includes the following novel material:
+\begin{itemize}
+\item .
+\end{itemize}
 
@@
 Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
-Sections describe our compilation approach for addressing the formalized learning tasks.
-Finally, the last sections report the data collected in a empirical evaluation, discuss the strengths and weaknesses of the compilation approach and propose several opportunities for future research.
+Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks and how the compilation is extensible to assess learned action models.
+Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First the learned \strips\ action models are tested with a set of state observation sequences and second, the learned models are compared to the corresponding reference model.
+Finally, Section ~\ref{sec:Section8} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.
 
@@
 \label{sec:Section3}
-This section defines the planning model used on this work and the output of the learning tasks addressed in the paper.
+This section defines the planning model used on this work, Classical planning with conditional effects, and the output of the learning tasks addressed in the paper, a \strips\ action model.
 
@@
 
+\begin{figure}[hbt!]
+\begin{footnotesize}
+\begin{verbatim}
+(:action stack
+ :parameters (?v1 ?v2 - object)
+ :precondition (and (holding ?v1) (clear ?v2))
+ :effect (and (not (holding ?v1)) (not (clear ?v2)) (handempty) (clear ?v1) (on ?v1 ?v2))
+)
+\end{verbatim}
+\end{footnotesize}
+ \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
+\label{fig:stack}
+\end{figure}
+
 To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
@@
 For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{verbatim}
-(:action stack
- :parameters (?v1 ?v2 - object)
- :precondition (and (holding ?v1) (clear ?v2))
- :effect (and (not (holding ?v1))
-              (not (clear ?v2))
-              (handempty) (clear ?v1)
-              (on ?v1 ?v2)))
-\end{verbatim}
-\end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
-\label{fig:stack}
-\end{figure}
+Given the set of predicates $\Psi$ and the header of the operator schema $\xi$, the upper-bound of the number of possible \strips\ action model for $\xi$, is $2^{3|F_v(\xi)|}$.
 
@@
 \subsection{Learning from state observations} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
+Learning from state observations is a relevant advancement as, in many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
+Learning action models from state observations broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}).
+
 This learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\sigma}$, when no intermediate information is given:
@@
 \section{Learning \strips\ action models with classical planning}
+\label{sec:Section5}
 Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
@@
 \section{Evaluating \strips\ action models}
-\label{model_evaluation}
-
+\label{sec:Section6}
 We assess how well a \strips\ action model $\mathcal{M}$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} required by $\mathcal{M}$ to induce $\mathcal{O}$.

### 2018-02-23 doc/icaps-18/camera-ready/strips-learning.tex
Final version camera-ready

@@
 
-\title{Learning \strips\ action models with classical planning}
+\title{Learning \strips\ Action Models with Classical Planning}
 \author{\#39}
@@
 This paper presents a novel approach for learning \strips\ action models from examples that compiles this inductive learning task into a classical planning task.
-Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states (no intermediate action or state is given).
+Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a pair of initial and final states (no intermediate action or state is given).
 Moreover, the compilation accepts partially specified action models and it can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
@@
 Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
+In both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
 Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
@@
 On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML to the learning of \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
+The application of inductive ML to learning \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
 \begin{itemize}
-\item The {\em input} to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects.
-The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
+\item The {\em input} to ML algorithms (the learning/training data) is usually a finite set of vectors that represent the value of some fixed object features.
+The input for learning planning action models is, however, observations of plan executions (where each plan possibly has a different length).
 \item The {\em output} of ML algorithms is usually a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative planning compilation approach for learning \strips\ action models.
-The compilation approach is appealing by itself because it opens the door to the bootstrapping of planning action models, but also because:
+The compilation approach is appealing by itself because it opens up the door to the bootstrapping of planning action models, but also because:
 
 \begin{enumerate}
-\item It is flexible to various amounts of available input knowledge.
+\item It is flexible to various amounts of input knowledge.
 Learning examples range from a set of plans (with their corresponding initial and final states) to just a pair of initial and final states where no intermediate state or action is observed.
@@
 \subsection{Classical planning with conditional effects}
-Our approach for learning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
+Our approach to learning \strips\ action models is compiling this learning task into a classical planning task with conditional effects.
 Conditional effects allow us to compactly define actions whose effects depend on the current state.
@@
 An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)$ and a set of {\em conditional effects} $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C \subseteq \mathcal{L}(F)$, the {\em condition}, and $E \subseteq \mathcal{L}(F)$, the {\em effect}.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals: $C \subseteq \mathcal{L}(F)$, the {\em condition}, and $E \subseteq \mathcal{L}(F)$, the {\em effect}.
 An action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 The {\em plan length} is denoted with $|\pi|=n$ .
@@
 \subsection{\strips\ action schemas and {\em variable name} objects}
-We address learning of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+Our approach is aimed at learning PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
 Figure~\ref{fig:stack} shows the {\em stack} schema of a four-operator {\em blocksworld} domain~\cite{slaney2001blocks} encoded in PDDL.
@@
 
-Finally, we assume that an action $a\in A$ is instantiated from a \strips\ an operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+Finally, we assume that an action $a\in A$ is instantiated from a \strips\ operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
 
@@
 For instance, the headers of a four-operator \emph{blocksworld} domain are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
-\item $pre(\xi)\subseteq F_v$ is the set of preconditions,  $del(\xi)\subseteq F_v$ the negative effects and  $add(\xi)\subseteq F_v$ the positive effects such tha $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\item $pre(\xi)\subseteq F_v$ is the set of preconditions,  $del(\xi)\subseteq F_v$ the negative effects and  $add(\xi)\subseteq F_v$ the positive effects such that $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
 \end{itemize}
@@
 Learning \strips\ action models from fully available input knowledge, i.e.
-from plans where the {\em pre-} and {\em post-states} of every action in the plans are available, is straightforward.
+from plans where the {\em pre-} and {\em post-states} of every action in the plans are known, is straightforward.
 When intermediate states are available, operator schemas are derived lifting the literals that change between the pre and post-state of each action execution.
@@
 \subsubsection{Learning from partially specified action models.}
-In case that partially specified operator schemas are available, we can incorporate this information within the learning task. This leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
+In case that partially specified operator schemas are available, we can incorporate this information within the learning task.
+The new leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
 
 \begin{itemize}
-\item $\Xi_0$ is a partially specified action model in whcih some preconditions and effects are known a priori.
+\item $\Xi_0$ is a partially specified action model in which some preconditions and effects are known a priori.
 \end{itemize}
 
+\vspace{0.1cm}
+
 A solution to $\Lambda$ is a set of operator schemas $\Xi$ that is compliant with the predicates in $\Psi$ and the set of initial and final states $\Sigma$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$}, that explain the given labels $\Sigma$ using the learned model.
+In a $\Lambda$ learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$}, that explain the given labels $\Sigma$ using the learned model.
 A solution to $\Lambda'$ is a set of \strips\ operator schemas $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action that has a different name in the example plans $\Pi$) that is compliant with the predicates $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
@@
 Fluents $F$ are built instantiating the predicates in $\Psi$ with the objects of the input labels $\Sigma$.
-Formally, $\Omega=\bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
+Formally, $\Omega=\bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)$, where $obj$ is a function that returns the objects that appear in a fully specified state.
 The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ and goals $G_t$ are defined by the state $s_n^t\in \sigma_t$.
+Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$, and the goals $G_t$ are defined by the state $s_n^t\in \sigma_t$.
 
@@
 \begin{itemize}
-\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$.
+\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$ that represent the programmed action model.
 If a fluent of $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect of the operator schema $\xi\in \Xi$.
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$}.
-\item A fluent $mode_{prog}$ to indicate whether the operator schemas are being programmed or validated (already programmed), and fluents $\{test_t\}_{1\leq t\leq \tau}$ which represent the example where the action model is being validated.
+\item A fluent $mode_{prog}$ to indicate whether the operator schemas are being programmed or validated (when already programmed)
+\item Fluents $\{test_t\}_{1\leq t\leq \tau}$ which represent the examples where the action model will be validated.
 \end{itemize}
-\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
+
+\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), the fluents in every $pre_f(\xi)\in F_{\Lambda}$ and the fluent $mode_{prog}$ set to true.
 Our compilation assumes that any operator schema is initially programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
@@
 \begin{itemize}
-\item Actions for {\bf removing} a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
+\item Actions for {\bf removing} a {\em precondition} $f\in F_v$ from $\xi$.
 
@@
 
-\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
+\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v$ to $\xi$.
 
@@
 
-\item Actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
-We assume operators headers are known so the binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
+\item Actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound to the objects $\omega\subseteq\Omega^{ar(\xi)}$.
+We assume operators headers are known so the binding of $\xi$ is done implicitly by order of appearance of the action parameters, i.e.
 variables $pars(\xi)$ are bound to the objects in $\omega$ that appear in the same position.
@@
 \begin{lemma}
-Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
+A classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
 \end{lemma}
@@
 \begin{small}
-Once operator schemas $\Xi$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
+Once operator schemas $\Xi$ are programmed, they can only be applied and validated according to the $mode_{prog}$ fluent.
 To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved by executing an applicable sequence of programmed operator schemas that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
-If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
+If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence it is a solution to $\Lambda$.
 \end{small}
@@
 \label{sec:Constraining}
-In this section we show that further input knowledge can be used to constrain the space of possible action models and to make the learning task more practicable.
+In this section, we show that further input knowledge can be used to constrain the space of possible action models and to make the learning task more practicable.
 
@@
 Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$} to indicate the plan step where the action model is validated.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$}, to indicate the plan step where the action model is validated.
 As in the original compilation, $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
@@
 \item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$ that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism forces that these actions are only applied as in the example plans.
-\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
+With this mechanism we ensure that these actions are applied in the same order as in the example plans.
+\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ is fully executed and extra conditional effects to remove plan $\pi_{t}$ and initiate the next plan $\pi_{t+1}$:
 \begin{small}
 \begin{align*}
-&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}, \{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
+&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\} \cup \{\neg f\}_{f\in F_{\pi_t}} \cup \{f\}_{f\in F_{\pi_t+1}}.
 \end{align*}
@@
 \subsection{Partially specified action models}
-Known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda'}$.
-The corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
+The known preconditions and effects of a partially specified action model are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true in the initial state $I_{\Lambda'}$.
+The programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become now unnecessary and they are removed from $A_{\Lambda'}$, thus making the planning task $P_{\Lambda'}$ be easier to solve.
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the {\em blocksworld} action model where operator schemas for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,22]$ validate the programmed operators following the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
+To illustrate this, the plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for acquiring the {\em blocksworld} action model where operator schemas for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld} using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
+Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,22]$ validate the programmed operators following the plan $\pi_1$ shown in Figure~\ref{fig:lexample}.
 
-In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
-In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips\ action model is {\em valid} for the given examples.
-If $P_{\Lambda'}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
+In the extreme, when a fully specified \strips\ action model is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
+In this case, if a solution plan is found for $P_{\Lambda'}$, it means that the given action model is {\em valid} for the provided examples.
+If $P_{\Lambda'}$ is unsolvable then it means that the action model is invalid because it is not compliant with all the given examples.
 Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
@@
 \begin{itemize}
-\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable objects} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
+\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable names} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
 \item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
@@
 One can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
-For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\Sigma$.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ then fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
+For instance, in the {\em zenotravel} domain, $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold in any state of $\Sigma$.
 
-Looking at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
+On the other hand, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
 $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
-Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
+Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)}, the only possible precondition that would include the static predicate is $pre\_next\_fly\_v5\_v4$.
 
@@
 \section{Evaluation}
-This section evaluates the performance of our approach for learning \strips\ action models starting from various amounts of available input knowledge.
+This section evaluates the performance of our approach for learning \strips\ action models with different amounts of available input knowledge.
 
@@
 The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
+We only used 5 learning examples for each domain and we fixed the examples for all the experiments so that we can evaluate the impact of the input knowledge in the quality of the learned models.
 All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM.
 
-The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
+The classical planner we used to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
+We used {\sc Madagascar} due to its ability to deal with planning instances populated with dead-ends.
 In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
@@
 
-We make fully available the compilation source code, evaluation scripts and used benchmarks at this repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
+For the sake of reproducibility, the compilation source code, evaluation scripts and used benchmarks are fully available at the repository {\em https://github.com/anonsub/strips-learning}.
 
 \subsubsection{Metrics.}
-The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
+The quality of the learned models is measured with the {\em precision} and {\em recall} metrics.
 These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
@@
 a learned model semantically equivalent but syntactically different to the reference model.
-This most likely occurs when the learning task is under-constrained.
+This mainly occurs when the learning task is under-constrained.
 
@@
 We start evaluating our approach with tasks $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, where {\em labeled plans} are available.
-We then repeat the evaluation but exploiting potential \emph{static predicates} computed from $\Sigma$, which are the predicates that appear  unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
+We then repeat the evaluation but exploiting potential \emph{static predicates} computed from $\Sigma$, which are the predicates that appear  unaltered in the initial and final states in every $\sigma_t\in\Sigma$.
 Static predicates are used to constrain the space of possible action models as explained in the previous section.
@@
 
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-All the learning tasks are solved in a few seconds time and interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
-When static predicates are identified, the resulting compilation is also much compact and produces smaller planning/instantiation times.
+Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
+All the learning tasks are solved in a few seconds.
+Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
+In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
+When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
 
@@
 	}
-\caption{\small {\em Precision} and {\em recall} obtained learning from labeled plans without (left) and with (right) static predicates.}
+\caption{\small {\em Precision} and {\em recall} scores for learning tasks from labeled plans without (left) and with (right) static predicates.}
 \label{tab:results_plans}
@@
         \end{scriptsize}
-	\caption{\small Total planning time, preprocessing time and plan length learning from labeled plans without/with static predicates.}
+	\caption{\small Total planning time, preprocessing time and plan length for learning tasks from labeled plans without/with static predicates.}
 	\label{tab:time_plans}	
@@
 We evaluate now the ability of our approach to support partially specified action models; that is, addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
-In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task .
+In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task.
 
@@
 We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
-Likewise, errors in $\Lambda''$ tasks have greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is done only over half of the actions (e.g.
-the preconditions recall for the {\em Floortile} or {\em Gripper} domains).
+In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is done only over half of the actions.
+This occurs, for instance, in the precondition \emph{recall} of domains such as {\em Floortile}, {\em Gripper} or {\em Satellite}.
 
@@
 Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
-The learning algorithm updates the formula of the belief state with every action and observation in the sequence and thereby the final returned formula includes all consistent models.
+The learning algorithm updates the formula of the belief state with every action and observation in the sequence and thus the final returned formula includes all consistent models.
 SLAF assesses the quality of the learned models with respect to the actual generative model.
@@
 The lack of available information is overcome by exploiting assumptions about the kind of domain model it has to generate.
-Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
+Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine.
 {\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain})  incorporates static predicates and applies a post-processing step after the {\sf LOCM} analysis that requires a set of optimal plans to be used in the learning phase.
-This is done to mitigate the limitation of {\sf LOCM} of inducing similar modes for domains with similar structures.
+This is done to mitigate the limitation of {\sf LOCM} of inducing similar models for domains with similar structures.
 {\sf LOP} compares the learned models with the corresponding reference model.
 
-Compiling the learning of action models into classical planning is a general and flexible approach that allows to accommodate various amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
-For instance, the example plans in $\Pi$ can be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), and thus learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ would also be attainable.
+Compiling an action model learning task into classical planning is a general and flexible approach that allows to accommodate various amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
+For instance, the example plans in $\Pi$ can be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), and learning tasks of the kind $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ would also be attainable.
 Furthermore, our approach seems extensible to learning other types of generative models (e.g.
-hierarchical models like HTN or behaviour trees), that can be more appealing than \strips\ models, since using them to compute planning solutions requires less search effort.
+hierarchical models like HTN or behaviour trees) that can be more appealing than \strips\ models since they require less search effort to compute a a planning solution.
 
@@
 We presented a novel approach for learning \strips\ action models from examples using classical planning.
-The approach is flexible to various amounts of available input knowledge and accepts partially specified action models.
-We also introduced {\em precision} and {\em recall} metrics, widely used in ML, to the evaluation of planning action models.
-These metrics separately assess the correctness and completeness of the learned models making easier the identification of flaws.
+The approach is flexible to various amounts of input knowledge and accepts partially specified action models.
+We also introduced the {\em precision} and {\em recall} metrics, widely used in ML, for evaluating the learned action models.
+These two metrics measure the soundness and completeness of the learned models and facilitate the identification of model flaws.
 
-As far as we know, this is the first work on learning action models exclusively using an {\em off-the-shelf} classical planner and evaluated over a wide range of different domains.
+To the best of our knowledge, this is the first work on learning action models that is exhaustively evaluated over a wide range of domains and uses exclusively an {\em off-the-shelf} classical planner.
 The work in~\cite{SternJ17} proposes a planning compilation for learning action models from plan traces following the {\em finite domain} representation for the state variables.
@@
 
-When example plans are available, we can compute accurate action models from small sets of learning examples (five examples per domain) and investing small learning times (less than a second).
-When action plans are not available, our approach still produces action models compliant with the input information.
-In this case, since learning is not constrained by actions, it can reformulate operators changing their semantics and making tricky their comparison with a reference model.
+When example plans are available, we can compute accurate action models from small sets of learning examples (five examples per domain) in little computation time (less than a second).
+When action plans are not available, our approach still produces action models that are compliant with the input information.
+In this case, since learning is not constrained by actions, operators can be reformulated changing their semantics, in which case the comparison with a reference model turns out to be tricky.
+
 %An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
@@
 
-Generating {\em informative} examples for learning planning action models is an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques able to autonomously collect informative learning examples.
-The combination of such exploration techniques with our learning approach is an intriguing research direction that opens the door to the bootstrapping of planning action models.
+Generating {\em informative} examples for learning planning action models is still an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions which have low probability of being chosen by chance~\cite{fern2004learning}.
+The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques that enable to autonomously collect informative learning examples.
+The combination of such exploration techniques with our learning approach is an appealing research direction that opens up the door to the bootstrapping of planning action models.
 

### 2018-02-23 doc/aij18/aij18.tex
updated sections 1-5 AIJ18 paper

@@
 The compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models.
-In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
 
@@
 \begin{itemize}
-\item .
+\item Unifies the formulation of our approach for learning \strips\ action models from state observations and from labeled plans.
+\item Shows that our compilation is also valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
+\item Proposes a distance metric to assess how well a \strips\ action model matches a given set labeled plans.
+Defines the compilation to compute this distance as well as the edit distance between two given \strips\ action models.
 \end{itemize}
@@
 \subsection{Learning from state observations} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
-Learning from state observations is a relevant advancement as, in many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
-Learning action models from state observations broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}).
-
-This learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\sigma}$, when no intermediate information is given:
+This learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$:
 \begin{itemize}
@@
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item  $\sigma=(s_0,s_{n})$ is a  $(initial, final)$ state pair, that we call {\em label}.
-the {\em final} state $s_{n}$ resulting from executing an unknown plan $\pi_t=\tup{a_1, \ldots, a_n}$ starting from the {\em initial} state $s_0$.
-When the intermediate states are available, the learning task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$, where $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$.
+\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+Let us also define the learning task  $\Lambda=\tup{\mathcal{M},\Psi,\sigma}$ where the observed states are just the state pair $\sigma=(s_0,s_{n})$, that we call {\em label}, and that comprises the {\em final} state $s_{n}$ resulting from executing the {\em unobserved} plan $\pi$ starting from the {\em initial} state $s_0$.
 \end{itemize}
 
-A solution is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the state observation sequence $\mathcal{O}$.
+A solution to the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ learning task is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the state observation sequence $\mathcal{O}$.
 In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plan $\pi$, that explain the given observations using the learned \strips\ model.
-Figure~\ref{fig:lexample} shows a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
+Figure~\ref{fig:example-observations} shows a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
 
@@
  \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ task for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
-\label{fig:lexample}
+\label{fig:example-observations}
 \end{figure}
@@
 \subsection{Learning from a labeled plan}
-Here we augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$:
-
+Here we redefine the task of learning \strips\ action models from observations of plan executions to cover the case where the actions executed by the observed agent are known.
+In this case, the learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, where:
 \begin{itemize}
-\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence that induces the corresponding state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates $s_i=\theta(s_{i-1},a_i)$.
+\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence applicable in $s_0\in\sigma$ and that generates the state $s_n\in\sigma$.
 \end{itemize}
 
-Figure~\ref{fig:lexample} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, that corresponds to observing the execution of an eight-action plan for inverting a four-block tower.
+Figure~\ref{fig:example-lplan} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, that corresponds to observing the execution of an eight-action plan for inverting a four-block tower.
 
-\begin{figure}
+\begin{figure}[hbt!]
 {\tt ;;; Predicates in $\Psi$}
@@
 \end{subfigure}%
- \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a labeled plan.}
-\label{fig:lexample}
+ \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$ task for learning a \strips\ action model in the blocksworld from a labeled plan.}
+\label{fig:example-lplan}
 \end{figure}
@@
 \subsection{Learning from multiple plans}
-The previous task define the learning of planning action models from a single plan execution. These definitions can be extended to the more general case where learning from multiple plans:
+The previous definitions formalize the learning \strips\ action models from observations of a single plan execution.
+These definitions are extensible to the more general case where learning from the execution of multiple plans:
 \begin{itemize}
-\item $\Lambda=\tup{\mathcal{M},\Psi,\Sigma}$ where $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
-Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ starting from the {\em initial} state $s_0^t$.
-\item $\Lambda=\tup{\mathcal{M},\Psi,\Sigma,\Pi}$ where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+  \item When learning from states observations, $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ is still a valid formalization but considering that now that $\mathcal{O}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ are sequences of {\em state observations} obtained observing the execution of the corresponding {\em unobserved} plan $\pi^t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+  \item When learning from a set of labeled plans the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\Sigma,\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans and $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is the corresponding set of labels such that each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, is an action sequence applicable in $s_0^t\in\sigma_t$ and that generates the state $s_n^t\in\sigma_t$.
 \end{itemize}
@@
 Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
-A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$, that explains the inputs to the lerning task.
+A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plans that explain the inputs to the learning task.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item {\bf Validates the action model $\mathcal{M}'$ in $\mathcal{O}$}.
-The solution plan continues with a postfix that reproduces the given input knowledge using the programmed action model $\mathcal{M}'$.
+\item {\bf Validates the action model $\mathcal{M}'$}.
+The solution plan continues with a postfix that reproduces the given input knowledge (the observations of the plan executions) using the programmed action model $\mathcal{M}'$.
 \end{enumerate}
@@
 \item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$, i.e.
-{\tt\small block1} and {\tt\small block2} in Figure~\ref{fig:lexample}.
+{\tt\small block1} and {\tt\small block2} in Figure~\ref{fig:example-observations}.
 \item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
@@
 Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
+
 \item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
@@
 
-\subsubsection{Compilation properties}
+Known preconditions and effects (that is a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
+The corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
+In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
+In this case, if a solution plan is found to $P_{\Lambda}$, it means that the given \strips\ action model is {\em valid} for the given examples.
+If $P_{\Lambda}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
+%Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
 
-\begin{mylemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$.
-\end{mylemma}
+To illustrate how our compilation works, Figure~\ref{fig:plan-observations} shows a plan that solves a classical planning task resulting from the compilation.
+The plan programs and validates the $stack$ schema using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.
+Note that this compilation is valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
+Even more, the compilation is valid for learning when there are missing state observation in $\mathcal{O}$ by removing any reference to the $mode_{val}$ fluent.
+In that case the classical planner will determine how many {\em apply} actions are necessary between two {\em validate} actions.
 
-\begin{proof}[Proof sketch]
-\begin{small}
-Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
-In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence {\small $1\leq i\leq n$}.
-This means that the programmed action model $\mathcal{M}'$ complies with the provided observations $\mathcal{O}$ and hence, solves $\Lambda$.
-\end{small}
-\end{proof}
-
-\begin{mylemma}
-Completeness.
-Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
-\end{mylemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O}$.
-\end{small}
-\end{proof}
-
-The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
-\begin{itemize}
-\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
-The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
-\item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
-\end{itemize}
+\begin{figure}[hbt!]
+{\footnotesize\tt
+     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
+     01 : (program\_pre\_handempty\_stack)\\
+     02 : (program\_pre\_holding\_stack\_v2)\\
+     03 : (program\_pre\_on\_stack\_v1\_v1)\\
+     04 : (program\_pre\_on\_stack\_v1\_v2)\\
+     05 : (program\_pre\_on\_stack\_v2\_v1)\\
+     06 : (program\_pre\_on\_stack\_v2\_v2)\\
+     07 : (program\_pre\_ontable\_stack\_v1)\\
+     08 : (program\_pre\_ontable\_stack\_v2)\\
+     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
+    10 : (program\_eff\_clear\_stack\_v2)\\
+    11 : (program\_eff\_handempty\_stack)\\
+    12 : (program\_eff\_holding\_stack\_v1)\\
+    13 : (program\_eff\_on\_stack\_v1\_v2)\\
+    {\bf 14} : (apply\_unstack block2 block1)\\
+    15 : (validate\_1)\\
+    16 : (apply\_putdown block2)\\
+    17 : (validate\_2)\\
+    18 : (apply\_pickup block1)\\
+    19 : (validate\_3)\\
+    20 : (apply\_stack block1 block2)\\
+    21 : (validate\_4)
+}
+ \caption{\small Plan for programming and validating the $stack$ schema using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+\label{fig:plan-observations}
+\end{figure}
 
@@
 We extend the compilation to consider labeled plans.
-Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$:
+Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ such that:
 \begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$}, for indicating the plan step where the action model is validated.
-Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of plan $\pi=\tup{a_1, \ldots, a_n}$.
+Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< n$}, are also added to represent the current plan step and to iterate through the steps of the plan.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$}, for indicating the plan step where the action model is validated.
+Goals are $G_{\Lambda'}=\{test_1\}$, as now the programmed action model is only validated in the single state observation $s_n$, the last state reached by the plan.
+If more intermediate observed states are known they can however be used for validation adding a $validate_{i}$ action for each observation $s_{i}\in \mathcal{O}$.
 \item With respect to $A_{\Lambda'}$.
 \begin{enumerate}
-\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism forces that these actions are only applied as in the example plans.
-\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
-\begin{small}
-\begin{align*}
-&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}, \{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
-\end{align*}
-\end{small}
+\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ and the actions for {\em validating} the programmed action model in the given state observations are the same.
+\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\pi}$ that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+With this mechanism we ensure that these actions are applied in the same order as in the example plan $\pi$.
 \end{enumerate}
@@
 
-\subsection{Partially specified action models}
-Known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda'}$.
-The corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
-
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,22]$ validate the programmed operators following the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
-
-In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
-In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips\ action model is {\em valid} for the given examples.
-If $P_{\Lambda'}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
-Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
+To illustrate this, the classical plan of Figure~\ref{fig:plan-lplan} is a solution to a learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and label $\sigma$ shown in Figure~\ref{fig:example-lplan}.
+Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,22]$ validate the programmed operators following the plan $\pi$ shown in the Figure~\ref{fig:example-lplan}.
 
@@
 }
- \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi_1$ and label $\sigma_1$ (shown in Figure~\ref{fig:lexample}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
-\label{fig:plan}
+ \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi$ and label $\sigma$ (shown in Figure~\ref{fig:example-lplan}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+\label{fig:plan-lplan}
 \end{figure}
 
-\subsection{Static predicates}
+Addressing the learning task $\Lambda=\tup{\mathcal{M},\Psi,\Sigma,\Pi}$ with classical planning requires introducing a small modification to our compilation.
+In particular, the actions for {\em validating} the labeled plan $\pi_t\in\Pi$ with the corresponding label $\sigma_t\in\Sigma$, {\tt\small $1\leq t\leq \tau$} are now defined as:
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{1\leq j<t}\cup\{\neg test_j\}_{t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
+\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\} \cup \{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t}.
+\end{align*}
+\end{small}
+
+\subsection{Compilation properties}
+
+\begin{mylemma}
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}/\Sigma,\Pi}$.
+\end{mylemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
+In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
+These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence {\small $1\leq i\leq n$}.
+This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
+\end{small}
+\end{proof}
+
+\begin{mylemma}
+Completeness.
+Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}/\Sigma,\Pi}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+\end{mylemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O}/\Sigma,\Pi$.
+\end{small}
+\end{proof}
+
+The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
+\begin{itemize}
+\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
+The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
+\item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+\end{itemize}
+
+\subsection{Exploiting static predicates to optimize the compilation}
 A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
@@
 Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
-
-\subsection{Learning from multiple plans}
-The $\Lambda=\tup{\mathcal{M},\Psi,\Sigma}$ can be solved with the previous compilation $P_{\Lambda}$.
-On the other hand, the learning task $\Lambda=\tup{\mathcal{M},\Psi,\Sigma,\Pi}$ requires modifying the previous compilation.
-In particular, the actions for {\em validating} the labeled plan $\pi_t\in\Pi$ with the corresponding label $\sigma_t\in\Sigma$, {\tt\small $1\leq t\leq \tau$}.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\\
-&\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
-\end{align*}
-\end{small}
 

### 2018-02-23 doc/aij18/aij18.tex
added updates to the AIJ18 paper

@@
 \begin{itemize}
-\item Unifies the formulation of our approach for learning \strips\ action models from state observations and from labeled plans.
+\item Unifies the formulation of our approach for learning \strips\ action models from state observations and from observed plans.
 \item Shows that our compilation is also valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
-\item Proposes a distance metric to assess how well a \strips\ action model matches a given set labeled plans.
+\item Proposes a distance metric to assess how well a \strips\ action model matches a given set of plans.
 Defines the compilation to compute this distance as well as the edit distance between two given \strips\ action models.
@@
 In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
-
-Compiling the learning of action models into classical planning is a general and flexible approach that allows to accommodate various amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
-For instance, the example plans in $\Pi$ could be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), so learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ could also be addressed.
-Furthermore, our approach seems extensible to learning other types of generative models (e.g.
-hierarchical models like HTN or behaviour trees), that can be more appealing than \strips\ models, since using them to compute planning solutions requires less search effort.
 
@@
 
-\subsection{Learning from state observations} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
+\subsection{Learning from observations of plan executions}
+
+The first learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
 This learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$:
@@
 \item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-Let us also define the learning task  $\Lambda=\tup{\mathcal{M},\Psi,\sigma}$ where the observed states are just the state pair $\sigma=(s_0,s_{n})$, that we call {\em label}, and that comprises the {\em final} state $s_{n}$ resulting from executing the {\em unobserved} plan $\pi$ starting from the {\em initial} state $s_0$.
+The observed states are not required to be full states, they can be {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
 \end{itemize}
@@
 
-\subsection{Learning from a labeled plan}
 Here we redefine the task of learning \strips\ action models from observations of plan executions to cover the case where the actions executed by the observed agent are known.
-In this case, the learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, where:
-\begin{itemize}
-\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence applicable in $s_0\in\sigma$ and that generates the state $s_n\in\sigma$.
-\end{itemize}
-
-Figure~\ref{fig:example-lplan} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, that corresponds to observing the execution of an eight-action plan for inverting a four-block tower.
+In this case there may be missing state observation in $\mathcal{O}$, otherwise the learning task is trivial as previously explained.
+In this case, the learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$, where:
+\begin{itemize}
+\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence that produces the sequence of state observations given in $\mathcal{O}$.
+\end{itemize}
+
+Figure~\ref{fig:example-plans} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$, that corresponds to observing the execution of an eight-action plan for inverting a four-block tower.
+Note that $\mathcal{M},\Psi$ are skipped since they are the same as for Figure~\ref{fig:example-observations}.
 
 \begin{figure}[hbt!]
-{\tt ;;; Predicates in $\Psi$}
+{\footnotesize\tt ;;;;;; Observations $\mathcal{O}$}
 \begin{footnotesize}
 \begin{verbatim}
-(handempty) (holding ?o  - object)
-(clear ?o - object) (ontable ?o - object)
-(on ?o1 - object ?o2 - object)
+;;; observation #0
+(clear B) (on B A) (ontable A) (handempty)
+
+;;; observation #4
+(clear A) (on blockA B) (ontable B) (handempty)
 \end{verbatim}
@@
 
-\vspace{0.2cm}
-
-\begin{subfigure}{.25\textwidth}
-{\tt ;;; Plan $\pi$}
+{\footnotesize\tt ;;; Plan $\pi$}
 \begin{footnotesize}
 \begin{verbatim}
-0: (unstack A B)
-1: (putdown A)
-2: (unstack B C)
-3: (stack B A)
-4: (unstack C D)
-5: (stack C B)
-6: (pickup D)
-7: (stack D C)
+0: (unstack B A)
+1: (putdown B)
+2: (pickup A)
+3: (stack A B)
 \end{verbatim}
 \end{footnotesize}
-\end{subfigure}%
-\begin{subfigure}{.6\textwidth}
-{\tt ;;; Label $\sigma=(s_0^1,s_{n}^1)$}
-\begin{lstlisting}[mathescape]
-\end{lstlisting}
-\vspace{0.1cm}
-\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
-\node (initD) [draw]{D};
-\node (initC) [draw, above=of initD.north]{C};
-\node (initB) [draw, above=of initC.north]{B};
-\node (initA) [draw, above=of initB.north]{A};
-\draw[thick] (-1,-0.25) -- (2.5,-0.25);
-
-\node (goalA) [draw, right=10mm of initD]{A};
-\node (goalB) [draw, right=10mm of initC]{B};
-\node (goalC) [draw, right=10mm of initB]{C};
-\node (goalD) [draw, right=10mm of initA]{D};
-\end{tikzpicture}
-\vspace{0.6cm}
-\end{subfigure}%
- \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$ task for learning a \strips\ action model in the blocksworld from a labeled plan.}
-\label{fig:example-lplan}
+
+ \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$ task for learning a \strips\ action model in the blocksworld from a four-action plan and two state observations.}
+\label{fig:example-plans}
 \end{figure}
 
-\subsection{Learning from multiple plans}
 The previous definitions formalize the learning \strips\ action models from observations of a single plan execution.
@@
   \item When learning from states observations, $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ is still a valid formalization but considering that now that $\mathcal{O}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ are sequences of {\em state observations} obtained observing the execution of the corresponding {\em unobserved} plan $\pi^t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-  \item When learning from a set of labeled plans the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\Sigma,\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans and $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is the corresponding set of labels such that each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, is an action sequence applicable in $s_0^t\in\sigma_t$ and that generates the state $s_n^t\in\sigma_t$.
+  \item When learning from a set of labeled plans the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans producing the given set of state obsevations.
 \end{itemize}
@@
 Given that the operators headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
+Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
 \begin{small}
@@
 
+\begin{figure}
+\begin{scriptsize}
+\begin{verbatim}
+(:action apply_stack
+  :parameters (?o1 - object ?o2 - object)
+  :precondition
+   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
+        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
+        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
+        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
+        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
+        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
+        (or (not (pre_clear_stack_v1)) (clear ?o1))
+        (or (not (pre_clear_stack_v2)) (clear ?o2))
+        (or (not (pre_holding_stack_v1)) (holding ?o1))
+        (or (not (pre_holding_stack_v2)) (holding ?o2))
+        (or (not (pre_handempty_stack)) (handempty)))
+  :effect
+   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
+        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
+        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
+        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
+        (when (del_ontable_stack_v1) (not (ontable ?o1)))
+        (when (del_ontable_stack_v2) (not (ontable ?o2)))
+        (when (del_clear_stack_v1) (not (clear ?o1)))
+        (when (del_clear_stack_v2) (not (clear ?o2)))
+        (when (del_holding_stack_v1) (not (holding ?o1)))
+        (when (del_holding_stack_v2) (not (holding ?o2)))
+        (when (del_handempty_stack) (not (handempty)))
+        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
+        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
+        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
+        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
+        (when (add_ontable_stack_v1) (ontable ?o1))
+        (when (add_ontable_stack_v2) (ontable ?o2))
+        (when (add_clear_stack_v1) (clear ?o1))
+        (when (add_clear_stack_v2) (clear ?o2))
+        (when (add_holding_stack_v1) (holding ?o1))
+        (when (add_holding_stack_v2) (holding ?o2))
+        (when (add_handempty_stack) (handempty))
+        (when (modeProg) (not (modeProg)))))
+\end{verbatim}
+\end{scriptsize}
+ \caption{\small PDDL action for applying an already programmed schema $stack$ (implications coded as disjunctions).}
+\label{fig:compilation}
+\end{figure}
+
 Known preconditions and effects (that is a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
@@
 Note that this compilation is valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
-Even more, the compilation is valid for learning when there are missing state observation in $\mathcal{O}$ by removing any reference to the $mode_{val}$ fluent.
+Even more, the compilation is valid for learning when there are missing state observations in $\mathcal{O}$ by removing any reference to the $mode_{val}$ fluent.
 In that case the classical planner will determine how many {\em apply} actions are necessary between two {\em validate} actions.
@@
 
-\subsection{Learning from a labeled plan}
-We extend the compilation to consider labeled plans.
-Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ such that:
+\subsection{Learning from a plan}
+We extend the compilation to consider observed actions.
+Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ such that:
 \begin{itemize}
@@
 \item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$}, for indicating the plan step where the action model is validated.
-Goals are $G_{\Lambda'}=\{test_1\}$, as now the programmed action model is only validated in the single state observation $s_n$, the last state reached by the plan.
-If more intermediate observed states are known they can however be used for validation adding a $validate_{i}$ action for each observation $s_{i}\in \mathcal{O}$.
+Goals are as in the original compilation, $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
+In other words, intermediate observed states are used for validation adding a $test_t$ fluent for each observation in $\mathcal{O}$.
 \item With respect to $A_{\Lambda'}$.
@@
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan-lplan} is a solution to a learning task $\Lambda=\tup{\mathcal{M},\Psi,\sigma,\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and label $\sigma$ shown in Figure~\ref{fig:example-lplan}.
-Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,22]$ validate the programmed operators following the plan $\pi$ shown in the Figure~\ref{fig:example-lplan}.
+To illustrate this, the classical plan of Figure~\ref{fig:plan-lplan} is a solution to a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and state observations $\mathcal{O}$ shown in Figure~\ref{fig:example-plans}.
+Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the plan $\pi$ shown in the Figure~\ref{fig:example-plans}.
 
@@
     13 : (program\_eff\_on\_stack\_v1\_v2)\\
-    {\bf 14} : (apply\_unstack a b i1 i2)\\
-    15 : (apply\_putdown a i2 i3)\\
-    16 : (apply\_unstack b c i3 i4)\\
-    17 : (apply\_stack b a i4 i5)\\
-    18 : (apply\_unstack c d i5 i6)\\
-    19 : (apply\_stack c b i6 i7)\\
-    20 : (apply\_pickup d i7 i8)\\
-    21 : (apply\_stack d c i8 i9)\\
-    {\bf 22} : (validate\_1)
+    {\bf 14} : (apply\_unstack b a i1 i2)\\
+    15 : (apply\_putdown b i2 i3)\\
+    16 : (apply\_pickup a i3 i4)\\
+    17 : (apply\_stack a b i4 i5)\\
+    {\bf 18} : (validate\_1)
 }
- \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi$ and label $\sigma$ (shown in Figure~\ref{fig:example-lplan}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+ \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi$ and state observations $\mathcal{O}$ (shown in Figure~\ref{fig:example-plans}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
 \label{fig:plan-lplan}
@@
 
-Addressing the learning task $\Lambda=\tup{\mathcal{M},\Psi,\Sigma,\Pi}$ with classical planning requires introducing a small modification to our compilation.
-In particular, the actions for {\em validating} the labeled plan $\pi_t\in\Pi$ with the corresponding label $\sigma_t\in\Sigma$, {\tt\small $1\leq t\leq \tau$} are now defined as:
+Addressing the learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$ with classical planning requires introducing a small modification to our compilation.
+In particular, the actions for {\em validating} the plan $\pi_t\in\Pi$ in the corresponding sequence of state observations $\tup{s_0^t, s_1^t, \ldots, s_n^t}$, {\tt\small $1\leq t\leq \tau$}, are now defined as:
 \begin{small}
@@
 
+The compilation could be extended for learning when there are missing actions in $\pi$ by not coding the corresponding $plan(name(\xi),\Omega^{ar(\xi)},j$ fluent and not adding the extra precondition $f\in F_{\pi}$ to the corresponding corresponding {\em apply} action.
+In that case the classical planner will determine the {\em apply} action necessary for the validating the learned action model.
+
 \subsection{Compilation properties}
@@
 \begin{mylemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}/\Sigma,\Pi}$.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$.
 \end{mylemma}
@@
 Completeness.
-Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}/\Sigma,\Pi}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
 \end{mylemma}
@@
 By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O}/\Sigma,\Pi$.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O},\Pi$.
 \end{small}
@@
 
-Static predicates can also constrain the space of possible preconditions by looking at the given set of labels $\Sigma$.
+Static predicates can also constrain the space of possible preconditions by looking at the given set of state observation $\mathcal{O}$.
 One can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
-For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\Sigma$.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\mathcal{O}$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
+For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{O}$.
 
@@
 \label{sec:Section6}
-We assess how well a \strips\ action model $\mathcal{M}$ explains a given sequence of observations $\mathcal{O}$ according to the amount of {\em edition} required by $\mathcal{M}$ to induce $\mathcal{O}$.
+We assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
 %In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
 
-\subsection{Edition of \strips\ action models}
+\subsection{The \strips\ edit distance}
 
 We first define the allowed \emph{operations} to edit a given \strips\ action model.
-With the aim of keeping a tractable branching factor of the planning instances that results from our compilation, we only define two {\em edit operations}:
+With the aim of keeping a tractable branching factor of the planning instances that will results from our compilation, we only define two {\em edit operations}:
 \begin{itemize}
@@
 
-\subsection{The \strips\ edit distance}
-
 We define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
@@
 
+This edit distance can also be defined with respect to a set of plans.
+
+\begin{mydefinition}
+  Given an action model $\mathcal{M}$, built from $F_v$ and a set of plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$.
+The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\Pi)$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$}; \[\delta(\mathcal{M},\Pi)=\min_{\forall \mathcal{M}' \rightarrow \Pi} \delta(\mathcal{M},\mathcal{M}')\]
+\end{mydefinition}
+
 The $\Lambda$ learning tasks can swap the roles of two operators whose headers match or two action parameters that belong to the same type (e.g.
 the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped).
-The {\em observation edit distance} is a semantic measure that is robust to role changes of this kind.
-
-Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning}, the {\em observation edit distance} assesses, with a single expression, the flaws in the preconditions and effects of a given learned model.
+Both the {\em observation edit distance} and the {\em labeled plans edit distance} are semantic measures robust to role changes of this kind.
+
+Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning}, the {\em observation edit distance} and the and the {\em labeled plans edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
 This fact enables the recognition of \strips\ action models.
 The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
-The {\em observation edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
-
-The edit distance can also be defined with respect to a labeled plan.
-
-\begin{mydefinition}
-  Given an action model $\mathcal{M}$, built from $F_v$, and the plan $\pi=\tup{a_1, \ldots, a_n}$  with the corresponding label $\sigma=(s_0,s_{n})$.
-The {\bf plan edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{\pi},\mathcal{\sigma})$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce a plan $\pi$ that induces $\sigma$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{\sigma}} \delta(\mathcal{M},\mathcal{M}')\]
-\end{mydefinition}
-
-Last but not least, the edit distance can also be defined with respect to multiple plans.
-
-\subsection{Computing the edit distance}
-
-Our compilation is extensible to compute the {\em edit distance} between two models.
-A solution to the planning task resulting from the extended compilation is a sequence of actions that edits the action model $\mathcal{M}$ to  produce $\mathcal{M}'$ using to the two {\em edit operations}.
-
+These {\em edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
+
+\subsection{Computing the observations and plans edit distance}
+
+Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$, is {\em non-empty}.
+In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+A solution to the planning task resulting from the extended compilation is a sequence of actions that:
+
+\begin{enumerate}
+\item {\bf Edits the action model $\mathcal{M}$}.
+A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemas in $\mathcal{M}$ using to the two {\em edit operations}.
+\item {\bf Validates the edited model $\mathcal{M}'$ in observations of the plan executions}.
+The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$ (or in $\Pi$), as explained in Section~\ref{sec:Section5}.
+\end{enumerate}
+
+Now $\Lambda$ is not a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
 The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
@@
 
-To illustrate this, the plan of Figure~\ref{fig:plan} solves the classical planning task $P_{\Lambda}'$ that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:lexample}.
-
-Our interest when computing the {\em observation edit distance} is not in $\mathcal{M}'$ but in the number of required {\em edit operations}, e.g.
-$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-edistance}.
-The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
-
-\begin{figure}[hbt!]
-{\tt\small
-00 : (insert\_add\_handempty\_stack)\\
-01 : (insert\_add\_clear\_stack\_var1)
-}
- \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:lexample}.}
-\label{fig:plan-edistance}
-\end{figure}
-
-\subsection{Computing the observations distance}
-
-Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$, is {\em non-empty}.
-In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-A solution to the planning task resulting from the extended compilation is a sequence of actions that:
-
-\begin{enumerate}
-\item {\bf Edits the action model $\mathcal{M}$}.
-A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemas in $\mathcal{M}$ using to the two {\em edit operations}.
-\item {\bf Validates the edited model $\mathcal{M}'$ in $\mathcal{O}$}.
-The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$, as in Section~\ref{sec:Section4}.
-\end{enumerate}
-
-Now $\Lambda$ is not a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
-The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
-
-\begin{itemize}
-\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
-\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
-In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
-\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
-The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
-The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
-\end{itemize}
-
-To illustrate this, the plan of Figure~\ref{fig:plan} solves the classical planning task $P_{\Lambda}'$ that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:lexample}.
+To illustrate this, the plan of Figure~\ref{fig:plan-odistance} solves the classical planning task $P_{\Lambda}'$ that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:example-observations}.
 
@@
 }
+ \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:example-observations}.}
+\label{fig:plan-odistance}
+\end{figure}
+
+The modeifications to the compilation explained in  Section~\ref{sec:Section5} are useful here to redefined $\Lambda$ as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$.
+Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} schema and validating it at the plan shown in Figure~\ref{fig:example-plans}.
+
+\begin{figure}[hbt!]
+{\tt\small
+00 : (insert\_add\_handempty\_stack)\\
+01 : (insert\_add\_clear\_stack\_var1)\\
+02 : (apply\_unstack b a i1 i2)\\
+03 : (apply\_putdown b i2 i3)\\
+04 : (apply\_pickup a i3 i4)\\
+05 : (apply\_stack a b i4 i5)\\
+06 : (validate\_1)
+}
+ \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the plan shown in Figure~\ref{fig:example-plans}.}
+\label{fig:plan-pdistance}
+\end{figure}
+
+\subsection{Computing the edit distance between models}
+
+Our compilation is also extensible to compute the {\em edit distance} between two models.
+A solution to the planning task resulting from the extended compilation is a sequence of actions that edits the action model $\mathcal{M}$ to  produce $\mathcal{M}'$ using to the two {\em edit operations}.
+
+The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
+
+\begin{itemize}
+\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
+\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
+In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
+This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
+\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
+The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
+The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
+include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
+\end{itemize}
+
+To illustrate this, the plan of Figure~\ref{fig:plan} solves the classical planning task $P_{\Lambda}'$ that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:lexample}.
+
+Againn our interest is in the number of required {\em edit operations}, e.g.
+$\delta(\mathcal{M},\mathcal{M'})=2$ for the example in Figure~\ref{fig:plan-edistance}.
+Again, the {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
+
+\begin{figure}[hbt!]
+{\tt\small
+00 : (insert\_add\_handempty\_stack)\\
+01 : (insert\_add\_clear\_stack\_var1)
+}
  \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:lexample}.}
-\label{fig:plan-odistance}
+\label{fig:plan-edistance}
 \end{figure}
@@
 
+\subsection{Learning from labeled plans}
+We start evaluating our approach with tasks $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, where {\em labeled plans} are available.
+We then repeat the evaluation but exploiting potential \emph{static predicates} computed from $\Sigma$, which are the predicates that appear  unaltered in the initial and final states in every $\sigma_t\in\Sigma$.
+Static predicates are used to constrain the space of possible action models as explained in the previous section.
+
+Table~\ref{tab:results_plans} shows the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), while the last two columns of each setting and the last row report averages values.
+We can observe that identifying static predicates leads to models with better precondition {\em recall}.
+This fact evidences that many of the missing preconditions corresponded to static predicates because there is no incentive to learn them as they always hold~\cite{gregory2015domain}.
+
+Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
+All the learning tasks are solved in a few seconds.
+Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
+In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
+When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
+
+\begin{table*}
+		\resizebox{\textwidth}{!}{%
+		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
+& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
+& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+                          \hline
+			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
+			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
+			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
+			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
+			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
+			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
+			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
+			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
+			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
+			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
+			\hline
+			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
+		\end{tabular}
+	}
+\caption{\small {\em Precision} and {\em recall} scores for learning tasks from labeled plans without (left) and with (right) static predicates.}
+\label{tab:results_plans}
+\end{table*}
+
+\begin{table}
+\begin{scriptsize}
+	\begin{center}
+		\begin{tabular}{l|c|c|c||c|c|c|}
+                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
+			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
+                         \hline
+			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
+			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
+			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
+			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
+			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
+			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
+                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
+                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
+			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
+			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
+			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
+			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
+		\end{tabular}
+	\end{center}
+        \end{scriptsize}
+	\caption{\small Total planning time, preprocessing time and plan length for learning tasks from labeled plans without/with static predicates.}
+	\label{tab:time_plans}	
+\end{table}
+
+\subsection{Learning from partially specified action models}
+
+We evaluate now the ability of our approach to support partially specified action models; that is, addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
+In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task.
+
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, which include the identification of static predicates.
+We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
+In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is done only over half of the actions.
+This occurs, for instance, in the precondition \emph{recall} of domains such as {\em Floortile}, {\em Gripper} or {\em Satellite}.
+
+Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
+The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
+Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
+Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\Xi_0$).
+{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
+
+%In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
+
+\subsection{Learning from (initial,final) state pairs}
+Finally, we evaluate our approach when input plans are not available and thereby the planner must not only compute the action models but also the plans that satisfy the input labels.
+Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the results obtained for the task $\Lambda=\tup{\Psi,\Sigma,\Xi_0}$ using static predicates.
+Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} was not able to solve the corresponding planning tasks within a 1000 sec.
+time bound.
+The values of \emph{precision} and \emph{recall} are significantly lower than in Table ~\ref{tab:results_plans}.
+Given that the learning hypothesis space is now fairly under-constrained, actions can be reformulated and still be compliant with the inputs (e.g.
+the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
+We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) and yet the results are below the scores obtained when learning from labeled plans.
+
+\begin{table}
+		\begin{center}
+                \begin{footnotesize}
+			\begin{tabular}{l|r|r|c|}
+				& $\delta(\mathcal{M},\mathcal{O})$ & $\delta(\mathcal{M},*)$ & $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$ \\
+				\hline
+				blocks & 0 & 90 & 1.0 \\
+				driverlog & 5 & 144 & 0.97 \\
+				ferry & 2 & 69 & 0.97 \\
+				floortile & 34 & 342 & 0.90 \\
+				grid & 42 & 153 & 0.73 \\
+				gripper & 2 & 30 & 0.93 \\
+				hanoi & 1 & 63 & 0.98 \\
+				hiking & 69 & 174 & 0.60 \\
+				miconic & 3 & 72 & 0.96 \\
+				npuzzle & 2 & 24 & 0.92 \\
+                                parking & 4 & 111 & 0.96 \\
+				satellite & 24 & 75 & 0.68 \\
+				transport & 4 & 78 & 0.95 \\
+				visitall & 2 & 24 & 0.92 \\
+				zenotravel & 3 & 63 & 0.95
+			\end{tabular}
+                        	\end{footnotesize}
+		\end{center}
+	\caption{\small Evaluation of the quality of the learned models with respect to an observations test set.}
+	\label{fig:observationstest}
+\end{table}
+
+\subsubsection{Evaluating with a test set}
+
+When a reference model is not available, the learned models are tested with an observation set.
+Table~\ref{fig:observationstest} summarizes the results obtained when evaluating the quality of the learned models with respect to a test set of state observations.
+Each test set comprises between 20 and 50 observations per domain and is generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
+
+The table shows, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
+The reported results show that, despite learning only from 25 state observations, 12 out of 15 learned domains yield ratios of $90\%$ or above.
+This fact evidences that the learned models require very small amounts of edition to match the observations of the given test set.
+
+\subsubsection{Evaluating with a reference model}
+
+Here we evaluate the learned models with respect to the actual generative model.
+Opposite to what usually happens in ML, this model is available when learning is applied to IPC domains.
+The model learned for each domain is compared with its reference model using:
+
+\begin{itemize}
+\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
+Precision gives a notion of {\em soundness}.
+\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing). Recall gives a notion of {\em completeness}.
+\end{itemize}
+
+Table~\ref{fig:observationsnomap} shows the precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
+The reason why the scores in Table ~\ref{fig:observationsnomap} are lower than in Table~\ref{fig:observationstest} is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but correspond to {\em reformulations} of the actual model (changes in the roles of actions with matching headers or parameters with matching types).
+
+\subsubsection{Precision and recall robust to model reformulations}
+
+To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model {\em reformulations}.
+
+Precision and recall are often combined using the {\em harmonic mean}.
+This expression, called the {\em F-measure} or the balanced {\em F-score}, is defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
+Given the learned action model $\mathcal{M}$ and the reference action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure} (considering swaps in the actions with matching headers or parameters with matching types).
+Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
+These results evidence that in all of the evaluated domains, except for {\em ferry} and {\em satellite}, the learning task swaps the roles of some actions (or parameters) with respect to their role in the reference model.
+
+As we can see in Table~\ref{fig:observationsmap}, the {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
+On the other hand, the learning scores of several domains in Table~\ref{fig:observationsmap} are still lower than in Table~\ref{fig:observationstest}.
+The reason lies in the particular observations comprised by the test sets.
+%This leads us to the question on how the test set was generated and how the observations in this set affect the results in Table~\ref{fig:observationstest}.
+As an example, in the {\em driverlog} domain, the action schema {\small \tt disembark-truck} is missing from the learned model because this action is never induced from the observations in the training set; that is, such action never appears in the corresponding \emph{unobserved} plan.
+The same happens with the {\small \tt paint-down} action of the {\em floortile} domain or {\small \tt move-curb-to-curb} in the {\em parking} domain.
+Interestingly, these actions do not appear either in the test sets and so the learned action models are not penalized in Table~\ref{fig:observationstest}.
+Generating {\em informative} and {\em representative} observations for learning planning action models is an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+
+%The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
+
+\begin{table}[hbt!]
+	\begin{center}
+		\begin{scriptsize}
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+				\hline
+
+				blocks & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 \\
+				driverlog & 0.0 & 0.0 & 0.25 & 0.43 & 0.0 & 0.0 & 0.08 & 0.14 \\
+				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
+				floortile & 0.38 & 0.55 & 0.4 & 0.18 & 0.56 & 0.45 & 0.44 & 0.39 \\
+				grid & 0.5 & 0.47 & 0.33 & 0.29 & 0.25 & 0.29 & 0.36 & 0.35 \\
+				gripper & 0.83 & 0.83 & 0.75 & 0.75 & 0.75 & 0.75 & 0.78 & 0.78 \\
+				hanoi & 0.5 & 0.25 & 0.5 & 0.5 & 0.0 & 0.0 & 0.33 & 0.25 \\
+				hiking & 0.43 & 0.43 & 0.5 & 0.35 & 0.44 & 0.47 & 0.46 & 0.42 \\
+				miconic & 0.6 & 0.33 & 0.33 & 0.25 & 0.33 & 0.33 & 0.42 & 0.31 \\
+				npuzzle & 0.33 & 0.33 & 0.0 & 0.0 & 0.0 & 0.0 & 0.11 & 0.11 \\
+				parking & 0.25 & 0.21 & 0.0 & 0.0 & 0.0 & 0.0 & 0.08 & 0.07 \\
+				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
+				transport & 1.0 & 0.3 & 0.8 & 0.8 & 1.0 & 0.6 & 0.93 & 0.57 \\
+				visitall & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
+				zenotravel & 0.67 & 0.29 & 0.33 & 0.29 & 0.33 & 0.14 & 0.44 & 0.24
+			\end{tabular}
+		\end{scriptsize}
+	\end{center}
+	\caption{\small Precision and recall values obtained without computing the $f_{P\&R}$ mapping with the reference model.}
+	\label{fig:observationsnomap}
+\begin{center}
+		\begin{scriptsize}
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+				\hline
+
+				blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				driverlog & 0.67 & 0.14 & 0.33 & 0.57 & 0.67 & 0.29 & 0.56 & 0.33 \\
+				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
+				floortile & 0.44 & 0.64 & 1.0 & 0.45 & 0.89 & 0.73 & 0.78 & 0.61 \\
+				grid & 0.63 & 0.59 & 0.67 & 0.57 & 0.63 & 0.71 & 0.64 & 0.62 \\
+				gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+				hiking & 0.78 & 0.6 & 0.93 & 0.82 & 0.88 & 0.88 & 0.87 & 0.77 \\
+				miconic & 0.8 & 0.44 & 1.0 & 0.75 & 1.0 & 1.0 & 0.93 & 0.73 \\
+				npuzzle & 0.67 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 0.89 \\
+				parking & 0.56 & 0.36 & 0.5 & 0.33 & 0.5 & 0.33 & 0.52 & 0.34 \\
+				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
+				transport & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0 & 0.63 \\
+				visitall & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 \\
+				zenotravel & 1.0 & 0.43 & 0.67 & 0.57 & 1.0 & 0.43 & 0.89 & 0.48
+			\end{tabular}
+		\end{scriptsize}
+	\end{center}
+	\caption{\small Precision and recall values obtained when computing the $f_{P\&R}$ mapping with the reference model.}
+	\label{fig:observationsmap}
+\end{table}
+
 \section{Conclussions}

### 2018-03-01 doc/aij18/aij18.tex
updates added to the AIJ18 paper

@@
 
-\title{Learning \strips\ action models with classical planning}
+\title{Learning and Evaluation of \strips\ Action Models with Classical Planning}
 \author[label1]{Diego Aineto}
@@
 \author[label1]{Eva Onaindia}
-\address[label1]{Department of Computer Systems and Computation, Universitat Politécnica de València. Spain}
+\address[label1]{Department of Computer Systems and Computation, Universitat Polit\`ecnica de Val\`encia. Spain}
 
@@
   This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning.
-The compilation approach is flexible to various amount and forms of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) or sequences of state observations, to just a set of initial and final states (where no intermediate action or state is given).
+The compilation approach is flexible to various amount and kind of available input knowledge; the learning examples can range from plans (with their corresponding initial and final states), sequences of state observations or just a set of initial and final states (where no intermediate action or state is given).
 The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
 What is more, the compilation is extensible to assess how well a \strips\ action model matches observations of plan executions.
-Last but not least, the paper evaluates the performance of the compilation approach by learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of observations of plan executions and (2), the true models.
+The performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of observations of plan executions and (2), the true models.
 \end{abstract}
@@
 
-On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
+Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
 The application of inductive ML to the learning of \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
@@
 \item The {\em input} to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects.
-The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
+The input for learning planning action models are observations of plan executions (where each plan possibly has a different length and involve a different number of objects).
 \item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
@@
 \item {\em Flexibility in the input knowledge}.
-The compilation is flexible to various amounts and forms of available input knowledge.
+The compilation is flexible to various amounts and kind of available input knowledge.
 Learning examples can range from a set of plans (with their corresponding initial and final states) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
-The compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models.
+The compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known.
 
@@
 
-A first description of the compilation and its extensions previously appeared in several conference papers~\cite{aineto2018learning}.
+A first description of the compilation previously appeared in several conference papers~\cite{aineto2018learning}.
 Compared to the conference papers, the present paper includes the following novel material:
@@
 Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
-Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks and how the compilation is extensible to assess learned action models.
-Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First the learned \strips\ action models are tested with a set of state observation sequences and second, the learned models are compared to the corresponding reference model.
+Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks and how the compilation is also extensible to assess learned action models.
+Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First, the learned \strips\ action models are tested with observations of plan executions and second, the learned models are compared to the actual models.
 Finally, Section ~\ref{sec:Section8} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.
@@
 We define $F_v$, a new set of fluents s.t.
-$F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and defines the elements that can appear in an action schema.
+$F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$, i.e.
+the variable names, and that defines the elements that can appear in an action schema.
 For the {\em blocksworld}, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
@@
 For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.
-Given the set of predicates $\Psi$ and the header of the operator schema $\xi$, the upper-bound of the number of possible \strips\ action model for $\xi$, is $2^{3|F_v(\xi)|}$.
+
+Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, then $2^{3|F_v(\xi)|}$ is an upper-bound on the number of possible \strips\ action model for $\xi$.
+In practice this number is lower because the previous \strips\ constraints require that a negative effect must appear as a precondition and that cannot be a positive effect and also, that a positive effect cannot appear as a precondition.
 
@@
 
-The first learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions, the actual executed actions are unobserved.
-This learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$:
+The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its plan executions, the actual executed actions are unobserved.
+This learning task is formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,}$:
 \begin{itemize}
 \item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
-In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but partially specified operator schemes where some preconditions and effects are a priori known.
+In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but partially specified operator schemes where some preconditions and some positive or negative effects are a priori known.
+\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+The observed states are not required to be full states, they can be {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing (the are not known to be either positive or negative).
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The observed states are not required to be full states, they can be {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
-\end{itemize}
-
-A solution to the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ learning task is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the predicates $\Psi$, and the state observation sequence $\mathcal{O}$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plan $\pi$, that explain the given observations using the learned \strips\ model.
-Figure~\ref{fig:example-observations} shows a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
+This set can be given as input or inferred from the state observations given in $\mathcal{O}$ when these observations are full states.
+\end{itemize}
+
+A solution to the $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the state observation sequence $\mathcal{O}$ and the predicates $\Psi$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plan $\pi$, that explains the given observations using the learned \strips\ model.
+Figure~\ref{fig:example-observations} shows a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
 
@@
 \end{footnotesize}
- \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ task for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
+ \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
 \label{fig:example-observations}
@@
 
-Here we redefine the task of learning \strips\ action models from observations of plan executions to cover the case where the actions executed by the observed agent are known.
-In this case there may be missing state observation in $\mathcal{O}$, otherwise the learning task is trivial as previously explained.
-In this case, the learning task is formalized as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$, where:
+Here we redefine the task of learning \strips\ action models from observations of plan executions to cover the learning scenario where the actions executed by the observed agent are known.
+There may be missing state observations in $\mathcal{O}$, because otherwise the learning task is trivial as previously explained.
+In this case, the learning task is formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
 \begin{itemize}
 \item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence that produces the sequence of state observations given in $\mathcal{O}$.
-\end{itemize}
-
-Figure~\ref{fig:example-plans} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$, that corresponds to observing the execution of an eight-action plan for inverting a four-block tower.
+Note that if the plan is diverse enough it can be used to infer the set of {\em empty} operator schemas $\mathcal{M}$.
+\end{itemize}
+
+Figure~\ref{fig:example-plans} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of an eight-action plan for inverting a four-block tower.
 Note that $\mathcal{M},\Psi$ are skipped since they are the same as for Figure~\ref{fig:example-observations}.
@@
 
- \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$ task for learning a \strips\ action model in the blocksworld from a four-action plan and two state observations.}
+ \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ task for learning a \strips\ action model in the blocksworld from a four-action plan and two state observations.}
 \label{fig:example-plans}
@@
 \begin{itemize}
-  \item When learning from states observations, $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ is still a valid formalization but considering that now that $\mathcal{O}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ are sequences of {\em state observations} obtained observing the execution of the corresponding {\em unobserved} plan $\pi^t=\tup{a_1^t, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-  \item When learning from a set of labeled plans the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans producing the given set of state obsevations.
+  \item When learning from states observations, $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ is still a valid formalization by simply considering that now $\mathcal{O}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ are sequences of {\em state observations} obtained observing the execution of multiple {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} such that, for each {\small $1\leq i\leq n^t$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+  \item When learning from a set of observed plans, the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given set of example plans producing the corresponding given set of state observations.
 \end{itemize}
@@
 \item {\bf Validates the action model $\mathcal{M}'$}.
-The solution plan continues with a postfix that reproduces the given input knowledge (the observations of the plan executions) using the programmed action model $\mathcal{M}'$.
+The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) using the programmed action model $\mathcal{M}'$.
 \end{enumerate}
@@
 \subsection{Learning from state observations}
-Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
 \begin{itemize}
@@
 \begin{itemize}
-\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects appearing in the input observations $\mathcal{O}$, i.e.
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ appearing in the input observations $\mathcal{O}$, i.e.
 {\tt\small block1} and {\tt\small block2} in Figure~\ref{fig:example-observations}.
+Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
 \item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
-\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemas are programmed or validated and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the observation where the action model is validated.
+\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemas are programmed or validated and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation where the action model is validated.
 \end{itemize}
@@
 
-\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
+\item $G_{\Lambda}=\bigcup_{1\leq i\leq |\mathcal{O}|}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
 \item $A_{\Lambda}$ comprises three kinds of actions:
@@
 
-\item Actions for {\em validating} an observation {\tt\small $1\leq i\leq n$}.
+\item Actions for {\em validating} an observation {\tt\small $1\leq i\leq |\mathcal{O}|$}.
 \begin{small}
@@
 \hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\\
-&\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
+&\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\cup \{mode_{val}\},\\
 \cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
@@
 
-Known preconditions and effects (that is a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
-The corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
-In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
-In this case, if a solution plan is found to $P_{\Lambda}$, it means that the given \strips\ action model is {\em valid} for the given examples.
-If $P_{\Lambda}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
+Known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
+In this case, the corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
+In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model:
+\begin{itemize}
+\item If a solution plan is found to $P_{\Lambda}$, it means that the given \strips\ action model is {\em valid} for the given examples.
+\item If $P_{\Lambda}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
 %Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
+\end{itemize}
 
@@
 Note that this compilation is valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
-Even more, the compilation is valid for learning when there are missing state observations in $\mathcal{O}$ by removing any reference to the $mode_{val}$ fluent.
-In that case the classical planner will determine how many {\em apply} actions are necessary between two {\em validate} actions.
+
+Even more, if any reference to the $mode_{val}$ fluent is removed from the compilation then, it is is valid for learning when there are missing state observations in $\mathcal{O}$.
+In that case, the compilation is more flexible but the resulting classical planning task $P_{\Lambda}$ becomes harder to be solved since the classical planner must determine how many {\em apply} actions are necessary between two {\em validate} actions.
 
@@
 
-\subsection{Learning from a plan}
+\subsection{Learning from plans}
 We extend the compilation to consider observed actions.
-Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ such that:
-\begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of plan $\pi=\tup{a_1, \ldots, a_n}$.
-Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< n$}, are also added to represent the current plan step and to iterate through the steps of the plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$}, for indicating the plan step where the action model is validated.
-Goals are as in the original compilation, $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
-In other words, intermediate observed states are used for validation adding a $test_t$ fluent for each observation in $\mathcal{O}$.
+Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda}}$ such that:
+\begin{itemize}
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}_{\small 1\leq j\leq |\pi|}$, the fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that contains the action $name(\xi),\Omega^{ar(\xi)}$.
+Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< |\pi|$}, are also added to represent the current plan step and iterate through the steps of the plan.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<|\pi|$}, for indicating the plan step where the action model is validated.
+Goals are as in the original compilation.
+In other words, observed states are used for validation adding a $test_t$, ${\small 1\leq t\leq |\mathcal{O}|}$ fluent for each observation in $\mathcal{O}$.
 \item With respect to $A_{\Lambda'}$.
@@
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan-lplan} is a solution to a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and state observations $\mathcal{O}$ shown in Figure~\ref{fig:example-plans}.
+To illustrate this, the classical plan of Figure~\ref{fig:plan-lplan} is a solution to a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $\mathcal{O}$ shown in Figure~\ref{fig:example-plans}.
 Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the plan $\pi$ shown in the Figure~\ref{fig:example-plans}.
@@
 
-Addressing the learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$ with classical planning requires introducing a small modification to our compilation.
-In particular, the actions for {\em validating} the plan $\pi_t\in\Pi$ in the corresponding sequence of state observations $\tup{s_0^t, s_1^t, \ldots, s_n^t}$, {\tt\small $1\leq t\leq \tau$}, are now defined as:
+The compilation is extensible for the learning scenario where there are missing actions in $\pi$.
+The extension does not code the corresponding $plan(name(\xi),\Omega^{ar(\xi)},j$ fluent and does not add the extra precondition $f\in F_{\pi}$ to the corresponding corresponding {\em apply} action.
+In that case the resulting classical planning task $P_{\Lambda'}$ becomes harder to be solved since the classical planner must determine the {\em apply} action necessary for the validating the learned action model.
+
+Now we explain how to address the learning from multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
+Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
+same fluents and actions but different initial states and goals).
+The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
+Finally, the initial state $I_t$ is given by the state $s_0^t$, and the goals $G_t$ are defined by the state $s_n^t$.
+Addressing the learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ with classical planning requires introducing a small modification to our compilation.
+In particular, the actions in $P_{\Lambda'}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq \tau$}, are now defined as:
 \begin{small}
@@
 
-The compilation could be extended for learning when there are missing actions in $\pi$ by not coding the corresponding $plan(name(\xi),\Omega^{ar(\xi)},j$ fluent and not adding the extra precondition $f\in F_{\pi}$ to the corresponding corresponding {\em apply} action.
-In that case the classical planner will determine the {\em apply} action necessary for the validating the learned action model.
-
 \subsection{Compilation properties}
@@
 \begin{mylemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$.
 \end{mylemma}
@@
 
-Looking at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
-Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
+Looking at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$ and $\mathcal{O}$.
+Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
 

### 2018-03-01 doc/aij18/aij18.tex
added updates to the AIJ18 paper

@@
 We first define the allowed \emph{operations} to edit a given \strips\ action model.
-With the aim of keeping a tractable branching factor of the planning instances that will results from our compilation, we only define two {\em edit operations}:
-\begin{itemize}
-\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, $f\in F_v(\xi)$.
-\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, $f\in F_v(\xi)$.
-\end{itemize}
-
-We can now formalize an edit distance that quantifies how dissimilar two given \strips\ action models are.
-The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations} have the same positive cost.
+With the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only define two {\em edit operations}:
+\begin{itemize}
+\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_v(\xi)$.
+\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_v(\xi)$.
+\end{itemize}
+
+We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ action models are.
+The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations}, deletion and insertion, have the same positive cost.
 
 \begin{mydefinition}
-Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, both built from the same set of possible elements $F_v$.
-The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} required to transform $\mathcal{M}$ into $\mathcal{M}'$.
+Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that both models are built from the same set of possible elements $F_v$.
+The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
 \end{mydefinition}
@@
 \begin{mydefinition}
-Given an action model $\mathcal{M}$, built from $F_v$, and the observations sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$.
-The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce a plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
+  Given $\mathcal{M}$, a \strips\ action models built from $F_v$, and the observations sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$ such that each observation in $\mathcal{O}$ is built with fluents in $F$.
+The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce a valid plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{mydefinition}
 
-This edit distance can also be defined with respect to a set of plans.
-
-\begin{mydefinition}
-  Given an action model $\mathcal{M}$, built from $F_v$ and a set of plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$.
-The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\Pi)$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$}; \[\delta(\mathcal{M},\Pi)=\min_{\forall \mathcal{M}' \rightarrow \Pi} \delta(\mathcal{M},\mathcal{M}')\]
-\end{mydefinition}
-
-The $\Lambda$ learning tasks can swap the roles of two operators whose headers match or two action parameters that belong to the same type (e.g.
-the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped).
-Both the {\em observation edit distance} and the {\em labeled plans edit distance} are semantic measures robust to role changes of this kind.
-
-Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning}, the {\em observation edit distance} and the and the {\em labeled plans edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
+Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning}, the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
 This fact enables the recognition of \strips\ action models.
@@
 
+This edit distance could also be defined quantifying the amount of edition necessary to modify the observations of the plan executions to match the given \strips\ action model.
+This would imply {\em edit operations} that modify the observations instead of modifying the action schemes.
+Our definition of the edit distance is however more practical since usually $F_v$ is much smaller than $F$.
+Finally, the edit distance can also be defined with respect to a set of plans.
+
+\begin{mydefinition}
+  Given an action model $\mathcal{M}$, built from $F_v$ and a set of valid plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ that only contain actions built grounding the schemes in $\mathcal{M}$.
+The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\Pi)$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$}; \[\delta(\mathcal{M},\Pi)=\min_{\forall \mathcal{M}' \rightarrow \Pi} \delta(\mathcal{M},\mathcal{M}')\]
+\end{mydefinition}
+
+An $\Lambda$ learning task can swap the roles of two operators whose headers match or two action parameters that belong to the same type (e.g.
+the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped).
+Both the {\em observation edit distance} and the {\em plans edit distance} are semantic measures robust to role changes of this particular kind.
+
 \subsection{Computing the observations and plans edit distance}
-
-Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O}}$, is {\em non-empty}.
+Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, is {\em non-empty}.
 In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
@@
 \item {\bf Edits the action model $\mathcal{M}$}.
-A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemas in $\mathcal{M}$ using to the two {\em edit operations}.
+A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemas in $\mathcal{M}$ using to the two {\em edit operations}, deletion and insertion.
 \item {\bf Validates the edited model $\mathcal{M}'$ in observations of the plan executions}.
@@
 The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:example-observations}.
-
-Our interest when computing the {\em observation edit distance} is not in $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
-$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-odistance}.
-The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
 
@@
 
-The modeifications to the compilation explained in  Section~\ref{sec:Section5} are useful here to redefined $\Lambda$ as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$.
-Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} schema and validating it at the plan shown in Figure~\ref{fig:example-plans}.
+Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
+$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-odistance}.
+In this case $\delta(\mathcal{M},*)=4\times 3\times 11$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and te size of $F_v$ is 11 (as shown in Section~\ref{sec:Section3}).
+The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
+
+If a set of plans $\Pi$ is available, the compilation can be adapted to compute the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
+The modifications to the compilation explained in  Section~\ref{sec:Section5} are useful here to redefine the learning task as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$.
+Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+In this case the edited action model is however validated it at the plan shown in Figure~\ref{fig:example-plans}.
 
@@
 
-\subsection{Computing the edit distance between models}
-
-Our compilation is also extensible to compute the {\em edit distance} between two models.
-A solution to the planning task resulting from the extended compilation is a sequence of actions that edits the action model $\mathcal{M}$ to  produce $\mathcal{M}'$ using to the two {\em edit operations}.
-
-The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
-
-\begin{itemize}
-\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
-\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
-In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
-\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
-The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
-The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
-\end{itemize}
-
-To illustrate this, the plan of Figure~\ref{fig:plan} solves the classical planning task $P_{\Lambda}'$ that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:lexample}.
-
-Againn our interest is in the number of required {\em edit operations}, e.g.
-$\delta(\mathcal{M},\mathcal{M'})=2$ for the example in Figure~\ref{fig:plan-edistance}.
-Again, the {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
+This compilation is flexible to compute the {\em edit distance} between two \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$, built from the same set of possible elements $F_v$.
+A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion.
+$F_{\Lambda}$, $I_{\Lambda}$ are defined like in the previous compilation.
+With respect to the actions, $A_{\Lambda}$ again implement the two {\em edit operations} (i.e.
+include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect) but does not contain apply actions because the \strips\ action model are not validated in any observation of plan executions.
+Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_v(\xi)$.
+
+To illustrate this, the plan of Figure~\ref{fig:plan-mdistance} solves the classical planning task that corresponds to computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.
+The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects.
+Again our interest is in the number of required {\em edit operations}, e.g.
+$\delta(\mathcal{M},\mathcal{M'})=2$.
 
@@
 }
- \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:lexample}.}
-\label{fig:plan-edistance}
+ \caption{\small Plan for computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.}
+\label{fig:plan-mdistance}
 \end{figure}
@@
 \label{sec:Section7}
+Given a set of possible \strips\ models and set of observations of plan executions , the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
+According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
+In our scenario, the hypotheses are about the possible \strips\ action models that can be built within a given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets).
+With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
+\begin{enumerate}
+\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$.
+\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
+\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
+\end{enumerate}
+
+If a set of plans $\Pi$ is available, this same strategy can be followed using the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
 
@@
 
-\subsection{Learning from labeled plans}
-We start evaluating our approach with tasks $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, where {\em labeled plans} are available.
-We then repeat the evaluation but exploiting potential \emph{static predicates} computed from $\Sigma$, which are the predicates that appear  unaltered in the initial and final states in every $\sigma_t\in\Sigma$.
-Static predicates are used to constrain the space of possible action models as explained in the previous section.
-
-Table~\ref{tab:results_plans} shows the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), while the last two columns of each setting and the last row report averages values.
-We can observe that identifying static predicates leads to models with better precondition {\em recall}.
-This fact evidences that many of the missing preconditions corresponded to static predicates because there is no incentive to learn them as they always hold~\cite{gregory2015domain}.
-
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
-All the learning tasks are solved in a few seconds.
-Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
-When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
-
-\begin{table*}
-		\resizebox{\textwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
-& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
-& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-                          \hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
-			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
-			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
-			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
-			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
-			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
-			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
-			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
-			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
-			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
-			\hline
-			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
-		\end{tabular}
-	}
-\caption{\small {\em Precision} and {\em recall} scores for learning tasks from labeled plans without (left) and with (right) static predicates.}
-\label{tab:results_plans}
-\end{table*}
-
-\begin{table}
-\begin{scriptsize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c||c|c|c|}
-                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
-			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
-                         \hline
-			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
-			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
-			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
-			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
-			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
-			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
-                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
-                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
-			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
-			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
-			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
-			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
-		\end{tabular}
-	\end{center}
-        \end{scriptsize}
-	\caption{\small Total planning time, preprocessing time and plan length for learning tasks from labeled plans without/with static predicates.}
-	\label{tab:time_plans}	
-\end{table}
-
-\subsection{Learning from partially specified action models}
-
-We evaluate now the ability of our approach to support partially specified action models; that is, addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
-In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task.
-
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, which include the identification of static predicates.
-We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
-In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is done only over half of the actions.
-This occurs, for instance, in the precondition \emph{recall} of domains such as {\em Floortile}, {\em Gripper} or {\em Satellite}.
-
-Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
-The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
-Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
-Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\Xi_0$).
-{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
-
-%In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
-
-\subsection{Learning from (initial,final) state pairs}
-Finally, we evaluate our approach when input plans are not available and thereby the planner must not only compute the action models but also the plans that satisfy the input labels.
-Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the results obtained for the task $\Lambda=\tup{\Psi,\Sigma,\Xi_0}$ using static predicates.
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} was not able to solve the corresponding planning tasks within a 1000 sec.
-time bound.
-The values of \emph{precision} and \emph{recall} are significantly lower than in Table ~\ref{tab:results_plans}.
-Given that the learning hypothesis space is now fairly under-constrained, actions can be reformulated and still be compliant with the inputs (e.g.
-the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
-We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) and yet the results are below the scores obtained when learning from labeled plans.
-
-\begin{table}
+\subsubsection{Evaluating with a test set}
+
+When a reference model is not available, the learned models are tested with an observation set.
+Table~\ref{fig:observationstest} summarizes the results obtained when evaluating the quality of the learned models with respect to a test set of state observations.
+Each test set comprises between 20 and 50 observations per domain and is generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
+
+The table shows, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
+The reported results show that, despite learning only from 25 state observations, 12 out of 15 learned domains yield ratios of $90\%$ or above.
+This fact evidences that the learned models require very small amounts of edition to match the observations of the given test set.
+
+\begin{table}[hbt!]
 		\begin{center}
@@
 
-\subsubsection{Evaluating with a test set}
-
-When a reference model is not available, the learned models are tested with an observation set.
-Table~\ref{fig:observationstest} summarizes the results obtained when evaluating the quality of the learned models with respect to a test set of state observations.
-Each test set comprises between 20 and 50 observations per domain and is generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
-
-The table shows, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
-The reported results show that, despite learning only from 25 state observations, 12 out of 15 learned domains yield ratios of $90\%$ or above.
-This fact evidences that the learned models require very small amounts of edition to match the observations of the given test set.
-
 \subsubsection{Evaluating with a reference model}
@@
 
+\subsection{Learning from plans}
+We start evaluating our approach with tasks $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, where {\em labeled plans} are available.
+We then repeat the evaluation but exploiting potential \emph{static predicates} computed from $\Sigma$, which are the predicates that appear  unaltered in the initial and final states in every $\sigma_t\in\Sigma$.
+Static predicates are used to constrain the space of possible action models as explained in the previous section.
+
+Table~\ref{tab:results_plans} shows the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), while the last two columns of each setting and the last row report averages values.
+We can observe that identifying static predicates leads to models with better precondition {\em recall}.
+This fact evidences that many of the missing preconditions corresponded to static predicates because there is no incentive to learn them as they always hold~\cite{gregory2015domain}.
+
+Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
+All the learning tasks are solved in a few seconds.
+Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
+In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
+When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
+
+\begin{table*}[hbt!]
+		\resizebox{\textwidth}{!}{%
+		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
+& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
+& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+                          \hline
+			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
+			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
+			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
+			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
+			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
+			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
+			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
+			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
+			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
+			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
+			\hline
+			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
+		\end{tabular}
+	}
+\caption{\small {\em Precision} and {\em recall} scores for learning tasks from labeled plans without (left) and with (right) static predicates.}
+\label{tab:results_plans}
+\end{table*}
+
+\begin{table}
+\begin{scriptsize}
+	\begin{center}
+		\begin{tabular}{l|c|c|c||c|c|c|}
+                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
+			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
+                         \hline
+			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
+			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
+			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
+			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
+			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
+			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
+                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
+                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
+			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
+			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
+			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
+			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
+		\end{tabular}
+	\end{center}
+        \end{scriptsize}
+	\caption{\small Total planning time, preprocessing time and plan length for learning tasks from labeled plans without/with static predicates.}
+	\label{tab:time_plans}	
+\end{table}
+
+\subsection{Learning from partially specified action models}
+
+We evaluate now the ability of our approach to support partially specified action models; that is, addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
+In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task.
+
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, which include the identification of static predicates.
+We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
+In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is done only over half of the actions.
+This occurs, for instance, in the precondition \emph{recall} of domains such as {\em Floortile}, {\em Gripper} or {\em Satellite}.
+
+Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
+The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
+Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
+Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\Xi_0$).
+{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
+
+%In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
+
+\begin{table}[hbt!]
+\begin{footnotesize}
+	\begin{center}
+
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+			\hline
+				Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.90 \\
+				Ferry & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
+				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
+                Grid & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.84 & 0.78 \\
+				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+				Miconic & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				Satellite & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
+				Transport & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
+				Zenotravel & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.78 \\
+				\hline
+				\bf  & 0.98 & 0.71 & 1.0 & 0.98 & 1.0 & 0.95 & 0.98 & 0.87 \\
+			\end{tabular}
+
+	\end{center}
+\end{footnotesize}
+\caption{\small {\em Precision} and {\em recall} scores for learning tasks with partially specified action models.}
+\label{tab:results_plans_partial}
+\end{table}
+
+\begin{table}
+\begin{footnotesize}
+	\begin{center}
+		\begin{tabular}{l|c|c|c|}			
+			 & Total time & Preprocess & Plan length  \\
+                         \hline
+			Blocks & 0.07 & 0.01 & 54  \\
+			Driverlog & 0.03 & 0.01 & 40 \\
+			Ferry & 0.06 & 0.03 & 45 \\
+			Floortile & 0.43 & 0.42 & 55 \\
+                        Grid & 3.12 & 3.07 & 53 \\
+			Gripper & 0.03 & 0.01 & 35 \\
+			Miconic & 0.03 & 0.01 & 34  \\
+			Satellite & 0.14 & 0.14 & 47 \\
+			Transport & 0.23 & 0.21 & 37 \\
+			Zenotravel & 0.90 & 0.89 & 40 \\
+		\end{tabular}
+	\end{center}
+        \end{footnotesize}
+	\caption{\small Time and plan length learning for learning tasks with partially specified action models.}
+	\label{tab:time_plans_partial}	
+\end{table}
+
 \section{Conclussions}

### 2018-03-05 doc/aij18/aij18.tex
updates to AIJ18 paper

@@
 
+For a given operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of fluents that represent the elements that can appear in that action schema.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.
+
 We assume also that actions $a\in A$ are instantiated from \strips\ operator schemas $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
@@
 \end{itemize}
-
-Finally, given the set of predicates $\Psi$ and the header of a \strips\ operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in the action schema $\xi$ and that confine its space of possible action models.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.
-
 Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, then $2^{3|F_v(\xi)|}$ is an upper-bound on the number of possible \strips\ action model for $\xi$.
-In practice this number is lower because the previous \strips\ constraints require that a negative effect must appear as a precondition and that cannot be a positive effect and also, that a positive effect cannot appear as a precondition.
+In practice this number is often lower because the previous \strips\ constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
 
@@
 The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its plan executions, the actual executed actions are unobserved.
-This learning task is formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,}$:
+This learning task is formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$:
 \begin{itemize}
 \item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
-In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but partially specified operator schemes where some preconditions and some positive or negative effects are a priori known.
+In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes where some preconditions and some positive or negative effects are a priori known.
 \item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The observed states are not required to be full states, they can be {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing (the are not known to be either positive or negative).
-\item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-This set can be given as input or inferred from the state observations given in $\mathcal{O}$ when these observations are full states.
+The states in $\mathcal{O}$ are not required to be full states, they can be {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing (it is unknown whether their value is either positive or negative).
+We assume that observations are noiseless, meaning that if the value of a fluent is observed it is correct.
+\item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame.
+This set of predicates can be given as input or inferred from the state observations given in $\mathcal{O}$ provided that they are full states.
 \end{itemize}
@@
 \begin{verbatim}
-(pickup v1) (putdown v1)
-(stack v1 v2} (unstack v1 v2)
+(pickup v1) (putdown v1) (stack v1 v2} (unstack v1 v2)
 \end{verbatim}
@@
 \begin{verbatim}
-(handempty) (holding ?o  - object)
-(clear ?o - object) (ontable ?o - object)
+(handempty) (holding ?o  - object) (clear ?o - object) (ontable ?o - object)
 (on ?o1 - object ?o2 - object)
@@
 ;;; observation #4
-(clear A) (on blockA B) (ontable B) (handempty)
+(clear A) (on A B) (ontable B) (handempty)
 \end{verbatim}
@@
 
-Here we redefine the task of learning \strips\ action models from observations of plan executions to cover the learning scenario where the actions executed by the observed agent are known.
-There may be missing state observations in $\mathcal{O}$, because otherwise the learning task is trivial as previously explained.
+Here we redefine the $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task to cover the scenario where the actions executed by the observed agent are known.
+Here $\mathcal{O}$ may contain partial states because otherwise, the learning task becomes trivial as explained.
 In this case, the learning task is formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
@@
 \item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence that produces the sequence of state observations given in $\mathcal{O}$.
-Note that if the plan is diverse enough it can be used to infer the set of {\em empty} operator schemas $\mathcal{M}$.
-\end{itemize}
-
-Figure~\ref{fig:example-plans} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of an eight-action plan for inverting a four-block tower.
+Note that if the plan is {\em diverse} enough, i.e.
+it contains at least one ground action for each of the aimed action schemes, it can be used to infer the set of {\em empty} operator schemas $\mathcal{M}$.
+\end{itemize}
+
+Figure~\ref{fig:example-plans} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of an four-action plan for inverting a two-block tower.
 Note that $\mathcal{M},\Psi$ are skipped since they are the same as for Figure~\ref{fig:example-observations}.
@@
   \item When learning from states observations, $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ is still a valid formalization by simply considering that now $\mathcal{O}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ are sequences of {\em state observations} obtained observing the execution of multiple {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} such that, for each {\small $1\leq i\leq n^t$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-  \item When learning from a set of observed plans, the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given set of example plans producing the corresponding given set of state observations.
+  \item When learning from a set of observed plans, the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given set of example plans producing the corresponding set of state observations $\mathcal{O}$.
 \end{itemize}
@@
 Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
-A planning compilation is a suitable approach for addressing $\Lambda$ because a solution must not only determine the \strips\ action model $\mathcal{M}'$ but also, the {\em unobserved} plans that explain the inputs to the learning task.
+A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plans that explain the inputs to the learning task.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \subsection{Learning from state observations}
+Here we formalize the compilation for learning \strips\ action models with classical planning.
 Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
@@
 \item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ appearing in the input observations $\mathcal{O}$, i.e.
-{\tt\small block1} and {\tt\small block2} in Figure~\ref{fig:example-observations}.
+{\tt\small block A} and {\tt\small block B} in Figure~\ref{fig:example-observations}.
 Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
-\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemas are programmed or validated and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation where the action model is validated.
+\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemas are programmed or validated and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation in $\mathcal{O}$ where the action model is validated.
 \end{itemize}
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
-& mode_{prog}, pre_{f}(\xi)\},\\
+\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}, pre_{f}(\xi)\},\\
 \cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
-& mode_{prog}\},\\
-\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
-&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
+\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
 \end{align*}
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
-&\cup \{\neg mode_{val}\},\\
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\cup \{\neg mode_{val}\},\\
 \cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\\
-&\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\cup \{mode_{val}\},\\
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\cup \{mode_{val}\},\\
 \cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
@@
 In this case, the corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
-In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model:
-\begin{itemize}
-\item If a solution plan is found to $P_{\Lambda}$, it means that the given \strips\ action model is {\em valid} for the given examples.
-\item If $P_{\Lambda}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
-%Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
-\end{itemize}
-
-To illustrate how our compilation works, Figure~\ref{fig:plan-observations} shows a plan that solves a classical planning task resulting from the compilation.
+When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the compilation validates whether the given observations of the plan execution follows the given model.
+This feature is beyond the functionality of VAL~\cite{howey2004val} because the plan validation tool requires a plan and a full model:
+\begin{itemize}
+\item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given examples if a solution plan is found for $P_{\Lambda}$.
+\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given examples if $P_{\Lambda}$ is unsolvable since it is not compliant with all the given examples. 
+\end{itemize}
+
+Figure~\ref{fig:plan-observations} illustrate how our compilation works and shows a plan that solves a classical planning task resulting from the compilation.
 The plan programs and validates the $stack$ schema using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.
 Note that this compilation is valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
-
-Even more, if any reference to the $mode_{val}$ fluent is removed from the compilation then, it is is valid for learning when there are missing state observations in $\mathcal{O}$.
-In that case, the compilation is more flexible but the resulting classical planning task $P_{\Lambda}$ becomes harder to be solved since the classical planner must determine how many {\em apply} actions are necessary between two {\em validate} actions.
 
@@
 
+If any reference to the $mode_{val}$ fluent is removed, the compilation becomes more flexible and can learn \strips\ action models even if there are missing state observations in $\mathcal{O}$.
+On the other hand, the resulting classical planning task $P_{\Lambda}$ becomes harder to be solved since the classical planner must now determine how many {\em apply} actions are necessary between any two {\em validate} actions.
+
 \subsection{Learning from plans}
-We extend the compilation to consider observed actions.
+Now we extend the compilation to consider observed actions.
 Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda}}$ such that:
 \begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}_{\small 1\leq j\leq |\pi|}$, the fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that contains the action $name(\xi),\Omega^{ar(\xi)}$.
-Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< |\pi|$}, are also added to represent the current plan step and iterate through the steps of the plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<|\pi|$}, for indicating the plan step where the action model is validated.
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with the $F_{\pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}_{\small 1\leq j\leq |\pi|}$ fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that contains the action $name(\xi),\Omega^{ar(\xi)}$.
+The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to represent the current plan step and iterate through the steps of the plan.
+\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<|\pi|$}, for tracking the plan step where the action model is validated.
 Goals are as in the original compilation.
 In other words, observed states are used for validation adding a $test_t$, ${\small 1\leq t\leq |\mathcal{O}|}$ fluent for each observation in $\mathcal{O}$.
-\item With respect to $A_{\Lambda'}$.
+\item With respect to the set of actions $A_{\Lambda'}$.
 \begin{enumerate}
-\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ and the actions for {\em validating} the programmed action model in the given state observations are the same.
+\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ and the actions for {\em validating} the programmed action model in a given state observation are the same as in the previous compilation.
 \item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\pi}$ that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-With this mechanism we ensure that these actions are applied in the same order as in the example plan $\pi$.
+With this mechanism we ensure that these actions are applied, exclusively, in the same order as in the example plan $\pi$.
 \end{enumerate}
@@
 This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $\mathcal{O}$ shown in Figure~\ref{fig:example-plans}.
-Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the plan $\pi$ shown in the Figure~\ref{fig:example-plans}.
+Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the four-action plan $\pi$ shown in the Figure~\ref{fig:example-plans}.
 
@@
 
-The compilation is extensible for the learning scenario where there are missing actions in $\pi$.
-The extension does not code the corresponding $plan(name(\xi),\Omega^{ar(\xi)},j$ fluent and does not add the extra precondition $f\in F_{\pi}$ to the corresponding corresponding {\em apply} action.
-In that case the resulting classical planning task $P_{\Lambda'}$ becomes harder to be solved since the classical planner must determine the {\em apply} action necessary for the validating the learned action model.
-
-Now we explain how to address the learning from multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
+The compilation is flexible to the learning scenario where some of the actions in $\pi$ are missing.
+In such scenario the corresponding $plan(name(\xi),\Omega^{ar(\xi)},j$ fluents are not coded in the initial state as well as they are not added as extra preconditions of the corresponding corresponding {\em apply} actions.
+Again, the resulting classical planning task that results from this modification of the compilation becomes harder to be solved since the planner must determine which are the {\em apply} action necessary for the validating the learned action model with the given input knowledge.
+
+Now we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
 Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
@@
 The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t$, and the goals $G_t$ are defined by the state $s_n^t$.
+Finally, the initial state $I_t$ is given by the state $s_0^t$ and the plan $\pi_t$, and the goals $G_t$ are defined by the state $s_n^t$.
 Addressing the learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ with classical planning requires introducing a small modification to our compilation.
-In particular, the actions in $P_{\Lambda'}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq \tau$}, are now defined as:
+In particular, the actions in $P_{\Lambda'}$ for {\em validating} the plan $\pi_t\in\Pi$ reset the current plan, {\tt\small $1\leq t\leq \tau$}, and are now defined as:
 \begin{small}
@@
 In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence {\small $1\leq i\leq n$}.
+These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence of actions defined by the plans in $\Pi$.
 This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
@@
 The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
+This is the term that dominates the compilation size.
 \item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
@@
 $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\mathcal{O}$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
-For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{O}$.
-
-Looking at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
+For instance, in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{O}$.
+
+Furthermore looking as well at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
 $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$ and $\mathcal{O}$.
@@
 
-We first define the allowed \emph{operations} to edit a given \strips\ action model.
-With the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only define two {\em edit operations}:
+We first define the two allowed \emph{operations} to edit a given \strips\ action model:
 \begin{itemize}
@@
 \end{itemize}
+In theory, we could also define a third edit operator for {\em substituting} a fluent from a given operator schema.
+However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only define two {\em edit operations}, deletion and insertion.
 
@@
 \begin{mydefinition}
-Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that both models are built from the same set of possible elements $F_v$.
+  Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that (1) both models have operator schemes with the same headers and (2), both are built from the same set of possible elements $F_v$.
 The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
@@
 
+As the ARMS system shows, the error of a learned action model can also be computed with respect to a set of observations of plan executions~\cite{yang2007learning}.
+This kind of evaluation provides semantic information (despite it is biased by the selection of the test examples).
 We define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
@@
 
-Unlike the error function defined by {\sc ARMS}~\cite{yang2007learning}, the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
+Unlike the error function defined by {\sc ARMS}, the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
 This fact enables the recognition of \strips\ action models.
@@
 
-This edit distance could also be defined quantifying the amount of edition necessary to modify the observations of the plan executions to match the given \strips\ action model.
-This would imply {\em edit operations} that modify the observations instead of modifying the action schemes.
-Our definition of the edit distance is however more practical since usually $F_v$ is much smaller than $F$.
-Finally, the edit distance can also be defined with respect to a set of plans.
+Note that the error of a learned action model could also be defined quantifying the amount of edition necessary to modify the observations of the plan executions to match the given model.
+This would imply defining {\em edit operations} that modify the fluents in the observations instead of modifying the fluents in the action schemes.
+With this regard, our definition of the edit distance is however more practical since normally, $F_v$ is much smaller than $F$ because the number of variable objects is smaller than the number of objects in the observations.
+Finally, the edit distance can also be defined with respect to a set of plans, if they are available.
 
@@
 the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped).
+Pure syntax-based metrics can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
+a learned model semantically equivalent but syntactically different to the reference model.
 Both the {\em observation edit distance} and the {\em plans edit distance} are semantic measures robust to role changes of this particular kind.
@@
 \begin{enumerate}
-\item {\bf Edits the action model $\mathcal{M}$}.
+\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
 A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemas in $\mathcal{M}$ using to the two {\em edit operations}, deletion and insertion.
 \item {\bf Validates the edited model $\mathcal{M}'$ in observations of the plan executions}.
-The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$ (or in $\Pi$), as explained in Section~\ref{sec:Section5}.
+The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$ (an in $\Pi$ if available), as explained in Section~\ref{sec:Section5}.
 \end{enumerate}
@@
 
-To illustrate this, the plan of Figure~\ref{fig:plan-odistance} solves the classical planning task $P_{\Lambda}'$ that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+To illustrate this, the plan of Figure~\ref{fig:plan-odistance} solves the classical planning task that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
 The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:example-observations}.
@@
 $\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-odistance}.
-In this case $\delta(\mathcal{M},*)=4\times 3\times 11$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and te size of $F_v$ is 11 (as shown in Section~\ref{sec:Section3}).
-The {\em observation edit distance} is exactly computed if $P_{\Lambda}'$ is optimally solved (according to the number of edit actions); is approximated if $P_{\Lambda}'$ is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}'$~\cite{bonet2001planning}.
-
-If a set of plans $\Pi$ is available, the compilation can be adapted to compute the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
-The modifications to the compilation explained in  Section~\ref{sec:Section5} are useful here to redefine the learning task as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$.
+In this case $\delta(\mathcal{M},*)=2\times 3\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(pickup)|=|F_v(putdown)|=11$ and the size of $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
+The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}$~\cite{bonet2001planning}.
+
+When the executed plans $\Pi$ are alos available, the compilation can be adapted to compute the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
+The modifications to the compilation explained in Section~\ref{sec:Section5} are also useful here to redefine the learning task as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$.
 Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
@@
 
-This compilation is flexible to compute the {\em edit distance} between two \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$, built from the same set of possible elements $F_v$.
+Last but not least, this compilation is flexible to compute the {\em edit distance} between two \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$, built from the same set of possible elements $F_v$.
 A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion.
@@
 Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_v(\xi)$.
-
 To illustrate this, the plan of Figure~\ref{fig:plan-mdistance} solves the classical planning task that corresponds to computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.
@@
 \label{sec:Section7}
-Given a set of possible \strips\ models and set of observations of plan executions , the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
+Given a set of possible \strips\ models and set of observations of plan executions, the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
 According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.

### 2018-03-06 doc/aij18/aij18.tex
updates added to AIJ18 paper

@@
 Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, then $2^{3|F_v(\xi)|}$ is an upper-bound on the number of possible \strips\ action model for $\xi$.
-In practice this number is often lower because the previous \strips\ constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
+In practice this number is lower and is given by $2^{2|F_v(\xi)|}$ because of the previous \strips\ constraints that require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
+
+Last but not least, we say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff (1) the operator schemes of both models have the same headers and (2), both are built from the same set of possible elements $F_v$.
+In other words, for every $\xi\in\mathcal{M}$ and $\xi'\in\mathcal{M'}$ such that $head(\xi)=head(\xi')$, it holds that $F_v(\xi)=F_v(\xi')$.
 
@@
 \end{itemize}
-In theory, we could also define a third edit operator for {\em substituting} a fluent from a given operator schema.
-However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only define two {\em edit operations}, deletion and insertion.
 
@@
 \begin{mydefinition}
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that (1) both models have operator schemes with the same headers and (2), both are built from the same set of possible elements $F_v$.
+  Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that they are {\em comparable}.
 The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
@@
 In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
+In practice this number is $2\times|F_v(\xi)|$ because, as explained in Section \ref{sec:Section3}, the \strips\ constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
 
 \begin{mydefinition}
-The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3|F_v(\xi)|$.
+The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2|F_v(\xi)|$.
 \end{mydefinition}
 
-As the ARMS system shows, the error of a learned action model can also be computed with respect to a set of observations of plan executions~\cite{yang2007learning}.
-This kind of evaluation provides semantic information (despite it is biased by the selection of the test examples).
-We define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
+Normally evaluating a given learned domain with respect to the actual generative normally is not possible because this model is not available.
+As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions~\cite{yang2007learning}.
+With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
 
 \begin{mydefinition}
-  Given $\mathcal{M}$, a \strips\ action models built from $F_v$, and the observations sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$ such that each observation in $\mathcal{O}$ is built with fluents in $F$.
-The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce a valid plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
+  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and the observations sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$ such that each state observation in $\mathcal{O}$ is built with fluents in $F$.
+The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{mydefinition}
@@
 
-Note that the error of a learned action model could also be defined quantifying the amount of edition necessary to modify the observations of the plan executions to match the given model.
-This would imply defining {\em edit operations} that modify the fluents in the observations instead of modifying the fluents in the action schemes.
-With this regard, our definition of the edit distance is however more practical since normally, $F_v$ is much smaller than $F$ because the number of variable objects is smaller than the number of objects in the observations.
-Finally, the edit distance can also be defined with respect to a set of plans, if they are available.
+Note that the error of a learned action model could also be defined quantifying the amount of edition necessary to modify the observations of the plan execution to match the given model.
+This would imply defining {\em edit operations} that modify the fluents in the observations instead of {\em edit operations} that modify the action schemes.
+Our definition of the edit distance is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
+Finally, the {\em edit distance} can also be defined with respect to a set of plans, if they are available.
 
 \begin{mydefinition}
-  Given an action model $\mathcal{M}$, built from $F_v$ and a set of valid plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ that only contain actions built grounding the schemes in $\mathcal{M}$.
-The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\Pi)$, is the minimal edit distance from $\mathcal{M}$ to any model $\mathcal{M}'$, defined also within $F_v$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$}; \[\delta(\mathcal{M},\Pi)=\min_{\forall \mathcal{M}' \rightarrow \Pi} \delta(\mathcal{M},\mathcal{M}')\]
+  Given an action model $\mathcal{M}$ and a set of valid plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ that only contain actions built grounding the schemes in $\mathcal{M}$.
+The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\Pi)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$}; \[\delta(\mathcal{M},\Pi)=\min_{\forall \mathcal{M}' \rightarrow \Pi} \delta(\mathcal{M},\mathcal{M}')\]
 \end{mydefinition}
 
-An $\Lambda$ learning task can swap the roles of two operators whose headers match or two action parameters that belong to the same type (e.g.
+An $\Lambda$ learning task can swap the roles of two action schemes whose headers match or two action parameters that belong to the same type (e.g.
 the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped).
@@
 \subsection{Computing the observations and plans edit distance}
-Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, is {\em non-empty}.
+Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, is {\em non-empty}.
 In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
@@
 \item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemas in $\mathcal{M}$ using to the two {\em edit operations}, deletion and insertion.
+A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
+In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema.
+However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}.
 \item {\bf Validates the edited model $\mathcal{M}'$ in observations of the plan executions}.
-The solution plan continues with a postfix that validates the edited model on the given observations $\mathcal{O}$ (an in $\Pi$ if available), as explained in Section~\ref{sec:Section5}.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{O}$ (and on $\Pi$ if available), as explained in Section~\ref{sec:Section5}.
 \end{enumerate}
 
-Now $\Lambda$ is not a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
+Now $\Lambda$ does not formalize a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
 The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
@@
 To illustrate this, the plan of Figure~\ref{fig:plan-odistance} solves the classical planning task that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:example-observations}.
+First, the plan edits the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:example-observations}.
 
@@
 }
- \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the state observations shown in Figure~\ref{fig:example-observations}.}
+\caption{\small Plan for editing a given action model and validating it at the state observations shown in Figure~\ref{fig:example-observations}.
+The given action model is a four operator {\em blockswold} where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.}
 \label{fig:plan-odistance}
@@
 $\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-odistance}.
-In this case $\delta(\mathcal{M},*)=2\times 3\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(pickup)|=|F_v(putdown)|=11$ and the size of $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
-The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of $P_{\Lambda}$~\cite{bonet2001planning}.
-
-When the executed plans $\Pi$ are alos available, the compilation can be adapted to compute the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
+In this case $\delta(\mathcal{M},*)=2\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ and the size of $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
+The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
+
+When the executed plans $\Pi$ are also available, the compilation can be adapted to compute the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
 The modifications to the compilation explained in Section~\ref{sec:Section5} are also useful here to redefine the learning task as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$.
 Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-In this case the edited action model is however validated it at the plan shown in Figure~\ref{fig:example-plans}.
+In this case the edited action model is however validated at the plan shown in Figure~\ref{fig:example-plans}.
 
@@
 
-Last but not least, this compilation is flexible to compute the {\em edit distance} between two \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$, built from the same set of possible elements $F_v$.
+Last but not least, this compilation is flexible to compute the {\em edit distance} between two {\em comparable} \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$.
 A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion.
@@
 Given a set of possible \strips\ models and set of observations of plan executions, the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
+
 According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the possible \strips\ action models that can be built within a given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets).
+In our scenario, the hypotheses are about the set of possible \strips\ action models.
+Given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section3}.
+With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the size of the possible state observations of length $n$, that is $\mathcal{O}=s_0,\ldots,s_n$ is given by $\prod_n 2^{|F|}$.
+
  With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:

### 2018-03-09 doc/aij18/aij18.tex
added section on constraints to the AIJ18 paper

@@
 
+\section{Learning \strips\ action models with background knowledge}
+A distinctive feature of Inductive Logic Programming (ILP)~\cite{muggleton1994inductive} is that ILP can leverage {\em background knowledge} to learn explanations from data.
+Inspired by ILP, we show that learning \strips\ action models can also leverage {\em background knowledge} in this case, given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.
+
+\subsection{State constraints}
+The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
+If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
+For instance to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
+
+{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} and making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ reachable from $I$.
+
+The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
+For instance Figure~\ref{fig:strongest-invariant} show the clauses that define the {\em strongest invariant} for {\em blocksworld}.
+\begin{figure}[hbt!]
+  \begin{footnotesize}
+    \begin{center}
+$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
+$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
+$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$ 
+    \end{center}
+\end{footnotesize}
+ \caption{\small Strongest invariant for {\em blocksworld}.}
+\label{fig:strongest-invariant}
+\end{figure}
+In general there are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.
+
+A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_1,block_2)\vee \neg on(block_1,block_3)$ is a mutex because $block_1$ can only be on top of a single block.
+
+A {\em domain invariant} is an instance-independent invariant, i.e.
+holds for any possible initial state.
+Therefore if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
+Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
+For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
+
+\subsection{Trajectory constraints}
+Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10,bacchus:LTL:1998}.
+For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true {\em at least at one} of the reached states.
+
+LTL is a modal temporal logic interpreted over sequences of states.
+LTL interpreted over finite sequences of states (also called traces) has received attention from the planning community and provided a formal description called $LTL_f$~\cite{Giacomo:LTLf:AAAI2014}.
+$LTL_f$ is built up from a finite set of propositional variables $P$, the logical operators $\neg$ and $\vee$ (from which it is possible to define operators $\wedge$, $\rightarrow$ and $\leftrightarrow)$, and the temporal modal operators {\em next}($\bigcirc$), and {\em until}(${\mathcal U}$).
+An $LTL_f$ formula is then inductively defined over a set of propositions $P$:
+\begin{itemize}
+\item A proposition $p\in P$ is an $LTL_f$ formula,
+\item if $\psi$ and $\chi$ are $LTL_f$ formulae, then so they are $\neg\psi$, ($\psi\vee\chi$), $\bigcirc\psi$, ($\psi\ {\mathcal U}\ \chi$).
+\end{itemize}
+
+Given a sequence of states $\tau=(s_0,\ldots, s_n)$, such that each state $s_i$, {\small $0\leq i\leq n$} is an element in $2^P$, we say that a given $LTL_f$ formula $\phi$ holds at instant $i$ (denoted by $\tau, i\models\phi$) iff:
+\begin{itemize}
+\item $\tau, i\models p$ for a propositional variable $p\in P$, iff $p\in s_i$,
+\item $\tau, i\models \neg\psi$ iff it is not the case that $\tau, i\models\psi$,
+\item $\tau, i\models \psi\wedge\chi$ iff $\tau, i\models\psi$ and $\tau, i\models\chi$,
+\item $\tau, i\models \bigcirc\phi$ if $i<n$ and $\tau, i+1\models\phi$,
+\item $\tau, i\models (\phi_1\ {\mathcal U}\ \phi_2)$ if exists some $j\in\{1,\ldots,n\}$ such that, $\tau, j\models \phi_2$ and for all $k\in\{i,\ldots,j-1\}$ it holds that $\tau, k\models \phi_1$.
+\end{itemize}
+We say that $\tau$ is {\em valid} for a given $LTL_f$ formula $\phi$ (denoted by $\tau\models\phi$) iff for every {\small $0\leq i\leq n$} holds that $\tau, i\models\phi$.
+From the previous basic operators it is also possible to define the abbreviation $last$ (that denotes the last instant of a sequence) and the temporal modal operators {\em eventually} ($\lozenge$), {\em always} ($\square$), and {\em release} (${\mathcal R}$):
+\begin{itemize}
+\item $last$ is shorthand for $\neg\bigcirc true$ and holds only at the last state of the sequence.
+The achievement of classical planning goals $G$ can then be expressed as the $LTL_f$ formula $\phi=last\wedge G$.
+\item $\lozenge$ $\psi$ stands for $(true\ {\mathcal U}\ \psi)$, and says that $\psi$ will eventually hold before the last state (last state included).
+\item $\square\psi$ stands for $\neg \lozenge \neg\psi$ and says that from the current state till the last one the formula will always hold.
+\item $\psi\ {\mathcal R}\ \chi$ stands for $\neg(\neg\psi\ {\mathcal U}\ \neg\chi)$.
+This says that $\chi$ has to be true until and including the point where $\psi$ first becomes true; if $\psi$ never becomes true, $\chi$ must remain true forever.
+\end{itemize}
+
+$LTL$ interpreted over finite traces has the expressive power of First Order Logic over finite ordered traces.
+Satisfiability, validity and logical implication (as well as model checking) for $LTL_f$ are PSPACE-complete.
+
+\subsection{Leveraging background knowledge}
+Our approach is flexible to include constraints in the validation of the learned models.
+Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check that the learned model satisfy every constraint $\phi_i\in\Phi$:
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
+\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
+\end{align*}
+\end{small}
+
+This redefinition of validate actions applies also to trajectory constraints because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.
+
+Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
+Constraints can be used to confine the set of possible \strips\ action models and reduce then the learning hypothesis space.
+With regard to our compilation, {\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema.
+For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
+\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
+\end{align*}
+\end{small}
+
 \section{Evaluation}

### 2018-03-13 doc/aij18/aij18.tex
updates added to the aij18 paper

@@
 
-\title{Learning and Evaluation of \strips\ Action Models with Classical Planning}
+\title{Learning and Evaluating \strips\ Action Models with Classical Planning}
 \author[label1]{Diego Aineto}
@@
   This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning.
-The compilation approach is flexible to various amount and kind of available input knowledge; the learning examples can range from plans (with their corresponding initial and final states), sequences of state observations or just a set of initial and final states (where no intermediate action or state is given).
+The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state), sequences of state observations or just a set of initial and final states (where no intermediate action or state is a priori known).
 The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
-What is more, the compilation is extensible to assess how well a \strips\ action model matches observations of plan executions.
+What is more, the compilation is extensible to assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions.
 The performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of observations of plan executions and (2), the true models.
@@
 \begin{itemize}
-\item The {\em input} to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects.
-The input for learning planning action models are observations of plan executions (where each plan possibly has a different length and involve a different number of objects).
+\item The {\em input} to ML algorithms (the {\em learning/training} data) usually are finite vectors encoding the value of fixed features in a given set of objects.
+The input for learning planning action models are observations of plan executions (where each plan possibly has a different length and may involve a different number of objects).
 \item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning \strips\ action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions.
+When learning \strips\ action models the output is, for each action, the sets of {\em preconditions}, {\em negative} and {\em positive effects} that define the possible state transitions.
 \end{itemize}
@@
 A solution to the classical planning task that results from our compilation is a sequence of actions that determines the learned action model, i.e.
-the preconditions and effects of the target \strips\ operator schemas.
-
-The compilation approach is appealing by itself, because it leverages off-the-shelf planners and because it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
-Moreover, the practicality of the compilation allow us to report learning results over a wide range of IPC planning domains.
+the preconditions and effects of the target \strips\ actions.
+
+The compilation approach is appealing by itself because it leverages off-the-shelf planners for learning and because its practicality allow us to report model learning results over a wide range of IPC planning domains.
+Moreover, it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
 Apart from these, the compilation approach presents the following contributions:
 \begin{enumerate}
-\item {\em Flexibility in the input knowledge}.
-The compilation is flexible to various amounts and kind of available input knowledge.
-Learning examples can range from a set of plans (with their corresponding initial and final states) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
-The compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known.
+\item {\em Flexibility}.
+The compilation is flexible to various amount and kind of available input knowledge.
+Learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
+Last but not least, the compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known.
+% Also accepts background knowledge
 
 \item {\em Model Evaluation}.
-The compilation can assess how well a \strips\ action model matches a given set of observations of plan executions.
+The compilation can assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions.
 Our compilation is extensible to accept a learned model as input besides the observations of plan executions.
@@
 
-A first description of the compilation previously appeared in several conference papers~\cite{aineto2018learning}.
-Compared to the conference papers, the present paper includes the following novel material:
-\begin{itemize}
-\item Unifies the formulation of our approach for learning \strips\ action models from state observations and from observed plans.
-\item Shows that our compilation is also valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
+A first description of the compilation previously appeared in the conference paper~\cite{aineto2018learning}. Compared to the conference paper, this work includes the following novel material:
+\begin{itemize}
+\item Unifies the formulation of our approach for learning \strips\ action models from observed plans but also from state observations.
+\item Evaluates our compilation in the particular scenario where the observed states are not full states but {\em partial states}.
+This mean that some of the observed fluents (either with {\em positive} or {\em negative} value) are missing because they cannot be observed.
 \item Proposes a distance metric to assess how well a \strips\ action model matches a given set of plans.
-Defines the compilation to compute this distance as well as the edit distance between two given \strips\ action models.
+Extends the compilation to compute this distance as well as the edit distance between two given \strips\ action models.
+In addition defines the task of {\em model recognition} as the task of selecting from a given set, the action model that best matches a {\em test set} with observations of plan executions.
+\item Shows that learning \strips\ action models can also leverage {\em background knowledge} in this case, given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.
 \end{itemize}
@@
 Section~\ref{sec:Section2} reviews related work on learning planning action models.
-Section~\ref{sec:Section3} formalizes the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the output of the addressed learning task).
-Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
+Section~\ref{sec:Section3} introduces the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the output of the addressed learning task).
+Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.
 Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks and how the compilation is also extensible to assess learned action models.
@@
 
+Recenctly classical planning compilations have been defined to learn different kinds of generative models from examples.
+The existing compilations for computing FSCs for generalized planning follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it and hence, they tightly integrate planning and generalization.
+To keep the computation of FSCs tractable, they limit the space of possible solutions bounding the maximum size of the FSC.
+In addition, they impose that the instances to solve share, not only the domain theory (actions and predicates schemes) but the set of fluents~\cite{javi-Gplanning-ICAPS16} or a subset of {\it observable} fluents~\cite{Geffner:FSM:AAAI10}.
+Programs increase the readability of FSCs separating the control-flow structures from the primitive actions.
+Like FSCs, programs can also be computed following a {\it top-down} approach, e.g.~exploiting compilations that program and validate the program on instances with the same state and action space~\cite{javi-Gplanning-ICAPS16}.
+Since these {\it top-down} approaches search in the space of solutions, it is helpful to limit the set of different control-flow instructions.
+For instance using only {\it conditional gotos} that can both implement branching and loops~\cite{Jimenez15}.
+
 \section{Background}
@@
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$.
+An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$ and {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$.
 We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
@@
 \end{itemize}
-Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, then $2^{3|F_v(\xi)|}$ is an upper-bound on the number of possible \strips\ action model for $\xi$.
-In practice this number is lower and is given by $2^{2|F_v(\xi)|}$ because of the previous \strips\ constraints that require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
-
-Last but not least, we say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff (1) the operator schemes of both models have the same headers and (2), both are built from the same set of possible elements $F_v$.
-In other words, for every $\xi\in\mathcal{M}$ and $\xi'\in\mathcal{M'}$ such that $head(\xi)=head(\xi')$, it holds that $F_v(\xi)=F_v(\xi')$.
+Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, then $2^{3|F_v(\xi)|}$ is an upper-bound on the number of possible \strips\ action models for $\xi$.
+In practice this upper-bound is lower, $2^{2|F_v(\xi)|}$, given that the previous \strips\ constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
+
+Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same headers and are built from the same set of possible elements.
+Formally, iff $head(\xi)=head(\xi')$ and it holds that $F_v(\xi)=F_v(\xi')$.
+Likewise we can say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff, every $\xi\in\mathcal{M}$ is comparable with an action schema $\xi'\in\mathcal{M'}$ and viceversa, every $\xi'\in\mathcal{M'}$ is comparable with an action schema $\xi\in\mathcal{M}$.
 
@@
 \subsection{Learning from observations of plan executions}
-
 The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its plan executions, the actual executed actions are unobserved.
@@
 \item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
-In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes where some preconditions and some positive or negative effects are a priori known.
+In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes.
+This means that some preconditions and some positive or negative effects are a priori known.
 \item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 The states in $\mathcal{O}$ are not required to be full states, they can be {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing (it is unknown whether their value is either positive or negative).
-We assume that observations are noiseless, meaning that if the value of a fluent is observed it is correct.
+Anyway we assume that observations are noiseless, meaning that if the value of a fluent is observed it is correct.
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame.
-This set of predicates can be given as input or inferred from the state observations given in $\mathcal{O}$ provided that they are full states.
-\end{itemize}
-
-A solution to the $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ compliant with the headers in $\mathcal{M}$, the state observation sequence $\mathcal{O}$ and the predicates $\Psi$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plan $\pi$, that explains the given observations using the learned \strips\ model.
+This set of predicates can be given as input but it can also be inferred from the state observations given in $\mathcal{O}$ provided that they are full states.
+\end{itemize}
+
+A solution to the $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ compliant with the input model $\mathcal{M}$, the state observation sequence $\mathcal{O}$ and the predicates $\Psi$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the unobserved plan $\pi$, that explains the given state observations using the learned \strips\ model.
 Figure~\ref{fig:example-observations} shows a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
+In this case we are assuming full state observability so, if a fluent is not present in the observations, it means that it is set to {\em false}.
 
@@
 
-Here we redefine the $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task to cover the scenario where the actions executed by the observed agent are known.
-Here $\mathcal{O}$ may contain partial states because otherwise, the learning task becomes trivial as explained.
-In this case, the learning task is formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
+Here we redefine the learning task to cover the scenario where the actions executed by the observed agent are known.
+In this case $\mathcal{O}$ should contain partial states because otherwise, the learning task becomes trivial as explained.
+The learning task is now formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
 \begin{itemize}
 \item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence that produces the sequence of state observations given in $\mathcal{O}$.
-Note that if the plan is {\em diverse} enough, i.e.
-it contains at least one ground action for each of the aimed action schemes, it can be used to infer the set of {\em empty} operator schemas $\mathcal{M}$.
-\end{itemize}
-
-Figure~\ref{fig:example-plans} shows an example of a {\em blocksworld} learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of an four-action plan for inverting a two-block tower.
+When the plan is {\em diverse} enough, i.e.
+it contains at least one ground action for each of the aimed action schemes, the set of {\em empty} operator schemas $\mathcal{M}$ can be inferred from the input plan $\pi$.
+\end{itemize}
+
+Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of an four-action plan for inverting a two-block tower.
 Note that $\mathcal{M},\Psi$ are skipped since they are the same as for Figure~\ref{fig:example-observations}.
+Again, we are assuming full state observability and the plan is full observable as well.
+However, only the first and last states were observed.
 
@@
 
-The previous definitions formalize the learning \strips\ action models from observations of a single plan execution.
-These definitions are extensible to the more general case where learning from the execution of multiple plans:
-\begin{itemize}
-  \item When learning from states observations, $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ is still a valid formalization by simply considering that now $\mathcal{O}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ are sequences of {\em state observations} obtained observing the execution of multiple {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$} such that, for each {\small $1\leq i\leq n^t$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-  \item When learning from a set of observed plans, the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given set of example plans producing the corresponding set of state observations $\mathcal{O}$.
+The previous definitions formalize the learning of \strips\ action models from the observation of a single plan execution.
+These definitions are extensible to the more general case where the available input data is obtained from the execution of multiple plans:
+\begin{itemize}
+\item When learning from states observations, $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ is still a valid formalization by simply considering that now $\mathcal{O}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ is a sequences of {\em state observations} that are obtained observing the execution of multiple {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, one after the other.
+This means that, for each {\small $1\leq i\leq n^t$}, $a_i^t\in \pi^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+  \item When learning from a set of observed plans, the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given sequence of example plans producing the corresponding sequence of state observations $\mathcal{O}$.
 \end{itemize}
@@
 Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
-A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plans that explain the inputs to the learning task.
+A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plans that explain the given inputs to the learning task.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \item {\bf Validates the action model $\mathcal{M}'$}.
-The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) using the programmed action model $\mathcal{M}'$.
+The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) but using the programmed action model $\mathcal{M}'$.
 \end{enumerate}
@@
 \begin{itemize}
-\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ appearing in the input observations $\mathcal{O}$, i.e.
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the input observations, i.e.
 {\tt\small block A} and {\tt\small block B} in Figure~\ref{fig:example-observations}.
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
-\item The fluents $mode_{prog}$ and $mode_{val}$ indicating whether the operator schemas are programmed or validated and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation in $\mathcal{O}$ where the action model is validated.
+\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation in $\mathcal{O}$ where the action model is validated.
 \end{itemize}
@@
 \item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
-Given that the operators headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
-Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
+Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
+Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$ from {\em blocksworld}.
 \begin{small}
@@
 
-\item Actions for {\em validating} an observation {\tt\small $1\leq i\leq |\mathcal{O}|$}.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\cup \{mode_{val}\},\\
-\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
-\end{align*}
-\end{small}
-\end{enumerate}
-\end{itemize}
-
-\begin{figure}
+\begin{figure}[hbt!]
 \begin{scriptsize}
@@
 
+\item Actions for {\em validating} an observation {\tt\small $1\leq i\leq |\mathcal{O}|$}.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\cup \{mode_{val}\},\\
+\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
+\end{align*}
+\end{small}
+\end{enumerate}
+\end{itemize}
+
+As introduced, the compilation approach is flexible to different amount of input data.
 Known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
@@
 When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the compilation validates whether the given observations of the plan execution follows the given model.
-This feature is beyond the functionality of VAL~\cite{howey2004val} because the plan validation tool requires a plan and a full model:
+This feature is beyond the functionality of VAL~\cite{howey2004val} because the plan validation tool requires (1) a full action model and (2), a plan:
 \begin{itemize}
 \item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given examples if a solution plan is found for $P_{\Lambda}$.
-\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given examples if $P_{\Lambda}$ is unsolvable since it is not compliant with all the given examples. 
+\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given examples if $P_{\Lambda}$ is unsolvable since it means that $\mathcal{M}$ cannot be compliant with all the given observations.
 \end{itemize}
@@
 Figure~\ref{fig:plan-observations} illustrate how our compilation works and shows a plan that solves a classical planning task resulting from the compilation.
-The plan programs and validates the $stack$ schema using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.
-Note that this compilation is valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
+The plan programs and validates the $stack$ schema (from {\em blocksworld}) using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.
+Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,21]$ validate the programmed operators using the sequence of five state observations shown in the Figure~\ref{fig:example-observations}.
 
@@
     13 : (program\_eff\_on\_stack\_v1\_v2)\\
-    {\bf 14} : (apply\_unstack block2 block1)\\
+    {\bf 14} : (apply\_unstack blockB blockA)\\
     15 : (validate\_1)\\
-    16 : (apply\_putdown block2)\\
+    16 : (apply\_putdown blockB)\\
     17 : (validate\_2)\\
-    18 : (apply\_pickup block1)\\
+    18 : (apply\_pickup blockA)\\
     19 : (validate\_3)\\
-    20 : (apply\_stack block1 block2)\\
+    20 : (apply\_stack blockA blockB)\\
     21 : (validate\_4)
@@
 
-If any reference to the $mode_{val}$ fluent is removed, the compilation becomes more flexible and can learn \strips\ action models even if there are missing state observations in $\mathcal{O}$.
-On the other hand, the resulting classical planning task $P_{\Lambda}$ becomes harder to be solved since the classical planner must now determine how many {\em apply} actions are necessary between any two {\em validate} actions.
+Note that this compilation is valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
+If any reference to the $mode_{val}$ fluent is removed, the compilation becomes even more flexible and can learn \strips\ action models even if there are missing state observations in $\mathcal{O}$.
+On the other hand, if the $mode_{val}$ fluent is ignored, the resulting classical planning task $P_{\Lambda}$ becomes harder to be solved since the classical planner must now determine how many {\em apply} actions are necessary between any two observations, i.e.
+between the application of two {\em validate} actions.
 
@@
 \begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with the $F_{\pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}_{\small 1\leq j\leq |\pi|}$ fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that contains the action $name(\xi),\Omega^{ar(\xi)}$.
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with the $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that contains the action $a_j$.
 The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to represent the current plan step and iterate through the steps of the plan.
@@
 Goals are as in the original compilation.
-In other words, observed states are used for validation adding a $test_t$, ${\small 1\leq t\leq |\mathcal{O}|}$ fluent for each observation in $\mathcal{O}$.
+In other words, the observed states are used for validation adding a $test_t$, ${\small 1\leq t\leq |\mathcal{O}|}$ fluent for each observation in $\mathcal{O}$.
 \item With respect to the set of actions $A_{\Lambda'}$.
@@
 \item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\pi}$ that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-With this mechanism we ensure that these actions are applied, exclusively, in the same order as in the example plan $\pi$.
+This mechanism ensures that these actions are applied, exclusively, in the same order as in the example plan $\pi$.
 \end{enumerate}
@@
     13 : (program\_eff\_on\_stack\_v1\_v2)\\
-    {\bf 14} : (apply\_unstack b a i1 i2)\\
-    15 : (apply\_putdown b i2 i3)\\
-    16 : (apply\_pickup a i3 i4)\\
-    17 : (apply\_stack a b i4 i5)\\
+    {\bf 14} : (apply\_unstack blockB blockA i1 i2)\\
+    15 : (apply\_putdown blockB i2 i3)\\
+    16 : (apply\_pickup blockA i3 i4)\\
+    17 : (apply\_stack blockA blockB i4 i5)\\
     {\bf 18} : (validate\_1)
@@
 
-The compilation is flexible to the learning scenario where some of the actions in $\pi$ are missing.
-In such scenario the corresponding $plan(name(\xi),\Omega^{ar(\xi)},j$ fluents are not coded in the initial state as well as they are not added as extra preconditions of the corresponding corresponding {\em apply} actions.
-Again, the resulting classical planning task that results from this modification of the compilation becomes harder to be solved since the planner must determine which are the {\em apply} action necessary for the validating the learned action model with the given input knowledge.
+The compilation is flexible to the learning scenario where some $a_j$ actions in $\pi$ are missing.
+In such scenario the corresponding $plan(name(a_j),\Omega^{ar(a_j)},j)$ fluents are not coded in the initial state as well as they are not added as extra preconditions of the corresponding corresponding {\em apply} actions.
+Again, the resulting classical planning task that results from this modification of the compilation becomes harder to be solved since the planner must determine which are the {\em apply} action necessary for validating the learned action model with the given input knowledge.
 
@@
 Addressing the learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ with classical planning requires introducing a small modification to our compilation.
-In particular, the actions in $P_{\Lambda'}$ for {\em validating} the plan $\pi_t\in\Pi$ reset the current plan, {\tt\small $1\leq t\leq \tau$}, and are now defined as:
+In particular, the actions in $P_{\Lambda'}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq \tau$} reset the current state, and the current plan, and are now defined as:
 \begin{small}
@@
 In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from $s_0$ and following the sequence of actions defined by the plans in $\Pi$.
+These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from the corresponding initial state and following the sequence of actions defined by the plans in $\Pi$.
 This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
@@
 \item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
-The larger these numbers, the larger the $F_v(\xi)$ sets, that define the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
-This is the term that dominates the compilation size.
+The larger these numbers, the larger the size of the $F_v(\xi)$ sets.
+This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
 \item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
@@
 $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\mathcal{O}$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
-For instance, in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{O}$.
+For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{O}$.
 
@@
 \label{sec:Section6}
-We assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
+The roles of two action schemes whose headers match or the roles of two action parameters that belong to the same type can be swapped by a $\Lambda$ learning task.
+For instance, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped.
+Pure syntax-based metrics can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
+a learned model semantically equivalent but syntactically different to the reference model.
+
+Here we introduce an evaluation approach that is robust to role changes of this particular kind.
+The intuition of the approach is to assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
 %In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
@@
 
-We first define the two allowed \emph{operations} to edit a given \strips\ action model:
+We first define the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
 \begin{itemize}
@@
 In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
-In practice this number is $2\times|F_v(\xi)|$ because, as explained in Section \ref{sec:Section3}, the \strips\ constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
-
 \begin{mydefinition}
-The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2|F_v(\xi)|$.
+The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
 \end{mydefinition}
 
-Normally evaluating a given learned domain with respect to the actual generative normally is not possible because this model is not available.
+Normally evaluating a given learned domain with respect to the actual generative model is not possible because the actual model is not available.
 As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions~\cite{yang2007learning}.
@@
 The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
-These {\em edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
-
-Note that the error of a learned action model could also be defined quantifying the amount of edition necessary to modify the observations of the plan execution to match the given model.
-This would imply defining {\em edit operations} that modify the fluents in the observations instead of {\em edit operations} that modify the action schemes.
+This {\em edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
+
+The error of a learned action model could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
+This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes.
 Our definition of the edit distance is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
@@
 
-An $\Lambda$ learning task can swap the roles of two action schemes whose headers match or two action parameters that belong to the same type (e.g.
-the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped).
-Pure syntax-based metrics can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
-a learned model semantically equivalent but syntactically different to the reference model.
-Both the {\em observation edit distance} and the {\em plans edit distance} are semantic measures robust to role changes of this particular kind.
-
 \subsection{Computing the observations and plans edit distance}
-Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, is {\em non-empty}.
+Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, is {\em non-empty}.
 In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
@@
 \item {\bf Validates the edited model $\mathcal{M}'$ in observations of the plan executions}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{O}$ (and on $\Pi$ if available), as explained in Section~\ref{sec:Section5}.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{O}$ (and on $\Pi$ if available), as explained in Section~\ref{sec:Section5} for the models that are programmed from scratch.
 \end{enumerate}
@@
 01 : (insert\_add\_clear\_stack\_var1)\\
-02 : (apply\_unstack block2 block1)\\
+02 : (apply\_unstack blockB blockA)\\
 03 : (validate\_1)\\
-04 : (apply\_putdown block2)\\
+04 : (apply\_putdown blockB)\\
 05 : (validate\_2)\\
-06 : (apply\_pickup block1)\\
+06 : (apply\_pickup blockA)\\
 07 : (validate\_3)\\
-08 : (apply\_stack block1 block2)\\
+08 : (apply\_stack blockA blockB)\\
 09 : (validate\_4)\\
@@
 $\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-odistance}.
-In this case $\delta(\mathcal{M},*)=2\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ and the size of $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
-The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner (our case); and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
+In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
+The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
 
@@
 01 : (insert\_add\_clear\_stack\_var1)\\
-02 : (apply\_unstack b a i1 i2)\\
-03 : (apply\_putdown b i2 i3)\\
-04 : (apply\_pickup a i3 i4)\\
-05 : (apply\_stack a b i4 i5)\\
+02 : (apply\_unstack blockB blockA i1 i2)\\
+03 : (apply\_putdown blockB i2 i3)\\
+04 : (apply\_pickup blockA i3 i4)\\
+05 : (apply\_stack blockA blockB i4 i5)\\
 06 : (validate\_1)
@@
 A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion.
-$F_{\Lambda}$, $I_{\Lambda}$ are defined like in the previous compilation.
+In this case the edited model is not validated on a sequence of observations or plans but on the given action model $\mathcal{M}'$ that acts as a reference.
+The sets of fluents $F_{\Lambda}$ and $I_{\Lambda}$ are defined like in the previous compilation.
 With respect to the actions, $A_{\Lambda}$ again implement the two {\em edit operations} (i.e.
@@
 \label{sec:Section7}
-Given a set of possible \strips\ models and set of observations of plan executions, the {\em recognition of \strips\ models} is the task of computing the model with the highest probability according to the cited observations.
+Given a set of possible \strips\ models and set of observations of plan executions, the {\em recognition of \strips\ models} is the task of identifying which model has the highest probability of producing the given observations.
 
@@
 Given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section3}.
-With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the size of the possible state observations of length $n$, that is $\mathcal{O}=s_0,\ldots,s_n$ is given by $\prod_n 2^{|F|}$.
+With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the size of the possible state observations of length $n$, that is $\mathcal{O}=s_0,\ldots,s_n$ is given by $2^{n\times|F|}$.
 
@@
 \item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$.
+If a set of plans $\Pi$ is available, this same strategy can be followed using the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
 \item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
@@
 
-If a set of plans $\Pi$ is available, this same strategy can be followed using the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
-
 \section{Learning \strips\ action models with background knowledge}
-A distinctive feature of Inductive Logic Programming (ILP)~\cite{muggleton1994inductive} is that ILP can leverage {\em background knowledge} to learn explanations from data.
-Inspired by ILP, we show that learning \strips\ action models can also leverage {\em background knowledge} in this case, given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.
+A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn explanations from data~\cite{muggleton1994inductive}.
+Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case, given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.
 
@@
 If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
-For instance to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
-
-{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} and making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$ reachable from $I$.
+For instance, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
+
+{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$.
 
 The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
-For instance Figure~\ref{fig:strongest-invariant} show the clauses that define the {\em strongest invariant} for {\em blocksworld}.
+For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}.
+There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.
+
 \begin{figure}[hbt!]
@@
 \end{footnotesize}
- \caption{\small Strongest invariant for {\em blocksworld}.}
+ \caption{\small An example of the strongest invariant for the {\em blocksworld} domain.}
 \label{fig:strongest-invariant}
 \end{figure}
-In general there are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.
 
 A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_1,block_2)\vee \neg on(block_1,block_3)$ is a mutex because $block_1$ can only be on top of a single block.
+For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a mutex because $block_A$ can only be on top of a single block.
 
 A {\em domain invariant} is an instance-independent invariant, i.e.
-holds for any possible initial state.
-Therefore if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
+holds for any possible initial state and set of objects.
+Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
 Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
@@
 Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10,bacchus:LTL:1998}.
-For instance the LTL {\em eventually} operator, denoted by $\lozenge$, can define constraints that, unlike {\em state invariants}, must be true {\em at least at one} of the reached states.
-
 LTL is a modal temporal logic interpreted over sequences of states.
 LTL interpreted over finite sequences of states (also called traces) has received attention from the planning community and provided a formal description called $LTL_f$~\cite{Giacomo:LTLf:AAAI2014}.
+
 $LTL_f$ is built up from a finite set of propositional variables $P$, the logical operators $\neg$ and $\vee$ (from which it is possible to define operators $\wedge$, $\rightarrow$ and $\leftrightarrow)$, and the temporal modal operators {\em next}($\bigcirc$), and {\em until}(${\mathcal U}$).
@@
 
-Given a sequence of states $\tau=(s_0,\ldots, s_n)$, such that each state $s_i$, {\small $0\leq i\leq n$} is an element in $2^P$, we say that a given $LTL_f$ formula $\phi$ holds at instant $i$ (denoted by $\tau, i\models\phi$) iff:
-\begin{itemize}
-\item $\tau, i\models p$ for a propositional variable $p\in P$, iff $p\in s_i$,
-\item $\tau, i\models \neg\psi$ iff it is not the case that $\tau, i\models\psi$,
-\item $\tau, i\models \psi\wedge\chi$ iff $\tau, i\models\psi$ and $\tau, i\models\chi$,
-\item $\tau, i\models \bigcirc\phi$ if $i<n$ and $\tau, i+1\models\phi$,
-\item $\tau, i\models (\phi_1\ {\mathcal U}\ \phi_2)$ if exists some $j\in\{1,\ldots,n\}$ such that, $\tau, j\models \phi_2$ and for all $k\in\{i,\ldots,j-1\}$ it holds that $\tau, k\models \phi_1$.
-\end{itemize}
-We say that $\tau$ is {\em valid} for a given $LTL_f$ formula $\phi$ (denoted by $\tau\models\phi$) iff for every {\small $0\leq i\leq n$} holds that $\tau, i\models\phi$.
+Given a sequence of states $\mathcal{O}=(s_0,\ldots, s_n)$, we say that a given $LTL_f$ formula $\phi$ holds at instant $i$ (denoted by $\mathcal{O}, i\models\phi$) iff:
+\begin{itemize}
+\item $\mathcal{O}, i\models f$ for a propositional variable $f\in F$, iff $f\in s_i$,
+\item $\mathcal{O}, i\models \neg\psi$ iff it is not the case that $\mathcal{O}, i\models\psi$,
+\item $\mathcal{O}, i\models \psi\wedge\chi$ iff $\mathcal{O}, i\models\psi$ and $\mathcal{O}, i\models\chi$,
+\item $\mathcal{O}, i\models \bigcirc\phi$ if $i<n$ and $\mathcal{O}, i+1\models\phi$,
+\item $\mathcal{O}, i\models (\phi_1\ {\mathcal U}\ \phi_2)$ if exists some $j\in\{1,\ldots,n\}$ such that, $\tau, j\models \phi_2$ and for all $k\in\{i,\ldots,j-1\}$ it holds that $\tau, k\models \phi_1$.
+\end{itemize}
+We say that $\mathcal{O}$ is {\em valid} for a given $LTL_f$ formula $\phi$ (denoted by $\mathcal{O}\models\phi$) iff for every {\small $0\leq i\leq n$} holds that $\mathcal{O}, i\models\phi$.
 From the previous basic operators it is also possible to define the abbreviation $last$ (that denotes the last instant of a sequence) and the temporal modal operators {\em eventually} ($\lozenge$), {\em always} ($\square$), and {\em release} (${\mathcal R}$):
@@
 \subsection{Leveraging background knowledge}
-Our approach is flexible to include constraints in the validation of the learned models.
+We have shown how our approach is flexible to validate the learned action models with respect to a set of observations of executed plans and with respect to a model that acts as a reference.
+Here we show that the validation of an action model can also be done with {\em state constraints} or {\em trajectory constraints}.
 Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check that the learned model satisfy every constraint $\phi_i\in\Phi$:
@@
 Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
-Constraints can be used to confine the set of possible \strips\ action models and reduce then the learning hypothesis space.
+Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space.
 With regard to our compilation, {\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema.

### 2018-03-14 doc/aij18/aij18.tex
updates added to the aij18 paper. First full version ready

@@
 \label{sec:Section8}
-This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts of available input knowledge.
+This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts and kinds of available input knowledge.
 
@@
 We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
-All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
-\begin{itemize}
-\item {\bf Reproducibility}.
-We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
+All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM.
+\begin{itemize}
 \item {\bf Planner}.
@@
 Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
-\item {\bf Metrics}.The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
+\item {\bf Reproducibility}.
+We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this repository {\em https://github.com/sjimenezgithub/strips-learning} so any experimental data reported in the paper is fully reproducible.
+\end{itemize}
+
+\subsection{Evaluating with a reference model}
+Here we evaluate the learned models with respect to the actual generative model.
+Opposite to what usually happens in ML, this model is available when learning is applied to IPC domains.
+The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
 These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
-Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
-Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
-Recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
-\end{itemize}
-
-\subsubsection{Evaluating with a test set}
+Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
+\begin{itemize}
+\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
+\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing). 
+\end{itemize}
+
+\subsubsection{Learning from plans}
+We start evaluating our approach with $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ learning tasks, where the action of the executed plans are available and the observation sequence contains only the corresponding initial and goal states, $s_o^t$ and $s_n^t$, for every plan $\pi_t\in\Pi$.
+We then repeat the evaluation but exploiting potential \emph{static predicates} computed from the observed states in $\mathcal{O}$, which are the predicates that appear unaltered in the states that belong to the same plan.
+Static predicates are used to constrain the space of possible action models as explained in Section~\ref{sec:Section5}.
+
+Table~\ref{tab:results_plans} shows the obtained results.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), while the last two columns of each setting and the last row report averages values.
+We can observe that identifying static predicates leads to models with better precondition {\em recall}.
+This fact evidences that many of the missing preconditions corresponded to static predicates because there is no incentive to learn them as they always hold~\cite{gregory2015domain}.
+
+\begin{table*}[hbt!]
+		\resizebox{\textwidth}{!}{%
+		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
+& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
+& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+                          \hline
+			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
+			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
+			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
+			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
+			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
+			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
+			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
+			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
+			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
+			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
+			\hline
+			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
+		\end{tabular}
+	}
+\caption{\small {\em Precision} and {\em recall} scores for learning tasks from labeled plans without (left) and with (right) static predicates.}
+\label{tab:results_plans}
+\end{table*}
+
+Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
+All the learning tasks are solved in a few seconds.
+Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
+In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
+When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
+
+\begin{table}
+\begin{scriptsize}
+	\begin{center}
+		\begin{tabular}{l|c|c|c||c|c|c|}
+                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
+			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
+                         \hline
+			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
+			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
+			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
+			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
+			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
+			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
+                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
+                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
+			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
+			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
+			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
+			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
+		\end{tabular}
+	\end{center}
+        \end{scriptsize}
+	\caption{\small Total planning time, preprocessing time and plan length for learning tasks from labeled plans without/with static predicates.}
+	\label{tab:time_plans}	
+\end{table}
+
+We evaluate now the ability of our approach to support partially specified action models; that is, when the input model $\mathcal{M}$ is not empty because some preconditions and effects of the actions are initially known.
+In this particular experiment, the model of half of the actions is given in $\mathcal{M}$ as an extra input of the learning task.
+Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, which include the identification of static predicates.
+We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
+In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the previous learning tasks because the evaluation is done only over half of the actions.
+This occurs, for instance, in the precondition \emph{recall} of domains such as {\em Floortile}, {\em Gripper} or {\em Satellite}.
+
+Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
+The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
+Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
+Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\mathcal{M}$).
+{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
+
+\begin{table}[hbt!]
+\begin{footnotesize}
+	\begin{center}
+
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+			\hline
+				Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.90 \\
+				Ferry & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
+				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
+                Grid & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.84 & 0.78 \\
+				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+				Miconic & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				Satellite & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
+				Transport & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
+				Zenotravel & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.78 \\
+				\hline
+				\bf  & 0.98 & 0.71 & 1.0 & 0.98 & 1.0 & 0.95 & 0.98 & 0.87 \\
+			\end{tabular}
+
+	\end{center}
+\end{footnotesize}
+\caption{\small {\em Precision} and {\em recall} scores for learning tasks with partially specified action models.}
+\label{tab:results_plans_partial}
+\end{table}
+
+\begin{table}
+\begin{footnotesize}
+	\begin{center}
+		\begin{tabular}{l|c|c|c|}			
+			 & Total time & Preprocess & Plan length  \\
+                         \hline
+			Blocks & 0.07 & 0.01 & 54  \\
+			Driverlog & 0.03 & 0.01 & 40 \\
+			Ferry & 0.06 & 0.03 & 45 \\
+			Floortile & 0.43 & 0.42 & 55 \\
+                        Grid & 3.12 & 3.07 & 53 \\
+			Gripper & 0.03 & 0.01 & 35 \\
+			Miconic & 0.03 & 0.01 & 34  \\
+			Satellite & 0.14 & 0.14 & 47 \\
+			Transport & 0.23 & 0.21 & 37 \\
+			Zenotravel & 0.90 & 0.89 & 40 \\
+		\end{tabular}
+	\end{center}
+        \end{footnotesize}
+	\caption{\small Time and plan length learning for learning tasks with partially specified action models.}
+	\label{tab:time_plans_partial}	
+\end{table}
+
+\subsubsection{Learning from state observations}
+Here we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, where the action of the executed plans are not available but the initial and goal states are known.
+When input plans are not available, the planner must not only compute the action models but also the plans that satisfy the input observations.
+Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the results obtained for this using static predicates and partially specified models.
+Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} was not able to solve the corresponding planning tasks within a 1000 sec.
+time bound.
+The values of \emph{precision} and \emph{recall} are significantly lower than in Table ~\ref{tab:results_plans}.
+Given that the learning hypothesis space is now fairly under-constrained, actions can be reformulated and still be compliant with the inputs (e.g.
+the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
+We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) and yet the results are below the scores obtained when learning from labeled plans.
+
+\begin{table}
+\begin{footnotesize}
+	\begin{center}
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+			\hline
+            Blocks & 0.33 & 0.33 & 0.75 & 0.50 & 0.33 & 0.33 & 0.47 & 0.39 \\
+            Driverlog & 1.0 & 0.29 & 0.33 & 0.67 & 1.0 & 0.50 & 0.78 & 0.48 \\
+            Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
+            Floortile & 0.67 & 0.40 & 0.50 & 0.40 & 1.0 & 0.40 & 0.72 & 0.40 \\
+            Grid & - & - & - & - & - & - & - & - \\
+            Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+            Miconic & 0.0 & 0.0 & 0.33 & 0.50 & 0.0 & 0.0 & 0.11 & 0.17 \\
+            Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
+            Transport & 0.0 & 0.0 & 0.25 & 0.5 & 0.0 & 0.0 & 0.08 & 0.17 \\
+            Zenotravel & - & - & - & - & - & - & - & - \\
+            \hline
+            & 0.63 & 0.29 & 0.54 & 0.70 & 0.67 & 0.53 & 0.61 & 0.51 \\			
+			\end{tabular}
+	\end{center}
+\end{footnotesize}
+\caption{\small {\em Precision} and {\em recall} scores for learning from (initial,final) state pairs.}
+\label{tab:results_states}
+\end{table}
+
+\begin{table}
+\begin{footnotesize}
+	\begin{center}
+		\begin{tabular}{l|c|c|c|}			
+			 & Total time & Preprocess & Plan length  \\
+			\hline
+            Blocks & 2.14 & 0.00 & 58  \\
+            Driverlog & 0.09 & 0.00 & 88 \\
+            Ferry & 0.17 & 0.01 & 65 \\
+            Floortile & 6.42 & 0.15 & 126 \\
+            Grid & - & - & - \\
+            Gripper & 0.03 & 0.00 & 47 \\
+            Miconic & 0.04 & 0.00 & 68 \\
+            Satellite & 4.34 & 0.10 & 126 \\
+            Transport & 2.57 & 0.21 & 47 \\			
+            Zenotravel & - & - & - \\
+		\end{tabular}
+	\end{center}
+        \end{footnotesize}
+	\caption{\small Time and plan length when learning from (initial,final) state pairs.}
+	\label{tab:time_states}	
+\end{table}
+
+Now we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, where the action of the executed plans are not available but where the observation sequence $\mathcal{O}$ contains all the intermediate states, not just the initial and final states.
+Table~\ref{fig:observationsnomap} shows the precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
+The reason why the scores in Table ~\ref{fig:observationsnomap} are still low, despite more state observations are available, is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but correspond to {\em reformulations} of the actual model (changes in the roles of actions with matching headers or parameters with matching types).
+
+\begin{table}[hbt!]
+	\begin{center}
+		\begin{scriptsize}
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+				\hline
+
+				blocks & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 \\
+				driverlog & 0.0 & 0.0 & 0.25 & 0.43 & 0.0 & 0.0 & 0.08 & 0.14 \\
+				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
+				floortile & 0.38 & 0.55 & 0.4 & 0.18 & 0.56 & 0.45 & 0.44 & 0.39 \\
+				grid & 0.5 & 0.47 & 0.33 & 0.29 & 0.25 & 0.29 & 0.36 & 0.35 \\
+				gripper & 0.83 & 0.83 & 0.75 & 0.75 & 0.75 & 0.75 & 0.78 & 0.78 \\
+				hanoi & 0.5 & 0.25 & 0.5 & 0.5 & 0.0 & 0.0 & 0.33 & 0.25 \\
+				hiking & 0.43 & 0.43 & 0.5 & 0.35 & 0.44 & 0.47 & 0.46 & 0.42 \\
+				miconic & 0.6 & 0.33 & 0.33 & 0.25 & 0.33 & 0.33 & 0.42 & 0.31 \\
+				npuzzle & 0.33 & 0.33 & 0.0 & 0.0 & 0.0 & 0.0 & 0.11 & 0.11 \\
+				parking & 0.25 & 0.21 & 0.0 & 0.0 & 0.0 & 0.0 & 0.08 & 0.07 \\
+				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
+				transport & 1.0 & 0.3 & 0.8 & 0.8 & 1.0 & 0.6 & 0.93 & 0.57 \\
+				visitall & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
+				zenotravel & 0.67 & 0.29 & 0.33 & 0.29 & 0.33 & 0.14 & 0.44 & 0.24
+			\end{tabular}
+		\end{scriptsize}
+	\end{center}
+	\caption{\small Precision and recall values obtained without computing the $f_{P\&R}$ mapping with the reference model.}
+	\label{fig:observationsnomap}
+\end{table}
+
+To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model {\em reformulations}.
+Precision and recall are often combined using the {\em harmonic mean}.
+This expression, called the {\em F-measure} or the balanced {\em F-score}, is defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
+Given the learned action model $\mathcal{M}$ and the reference action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure} (considering swaps in the actions with matching headers or parameters with matching types).
+
+Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
+The {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
+These results evidence that in all of the evaluated domains, except for {\em ferry} and {\em satellite}, the learning task swaps the roles of some actions (or parameters) with respect to their role in the reference model.
+
+\begin{table}
+        \begin{center}
+
+		\begin{scriptsize}
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+				\hline
+
+				blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				driverlog & 0.67 & 0.14 & 0.33 & 0.57 & 0.67 & 0.29 & 0.56 & 0.33 \\
+				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
+				floortile & 0.44 & 0.64 & 1.0 & 0.45 & 0.89 & 0.73 & 0.78 & 0.61 \\
+				grid & 0.63 & 0.59 & 0.67 & 0.57 & 0.63 & 0.71 & 0.64 & 0.62 \\
+				gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+				hiking & 0.78 & 0.6 & 0.93 & 0.82 & 0.88 & 0.88 & 0.87 & 0.77 \\
+				miconic & 0.8 & 0.44 & 1.0 & 0.75 & 1.0 & 1.0 & 0.93 & 0.73 \\
+				npuzzle & 0.67 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 0.89 \\
+				parking & 0.56 & 0.36 & 0.5 & 0.33 & 0.5 & 0.33 & 0.52 & 0.34 \\
+				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
+				transport & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0 & 0.63 \\
+				visitall & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 \\
+				zenotravel & 1.0 & 0.43 & 0.67 & 0.57 & 1.0 & 0.43 & 0.89 & 0.48
+			\end{tabular}
+		\end{scriptsize}
+	\end{center}
+	\caption{\small Precision and recall values obtained when computing the $f_{P\&R}$ mapping with the reference model.}
+	\label{fig:observationsmap}
+\end{table}
+
+\subsection{Evaluating with a test set}
 
@@
 Each test set comprises between 20 and 50 observations per domain and is generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
-
 The table shows, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
@@
 
-\subsubsection{Evaluating with a reference model}
-
-Here we evaluate the learned models with respect to the actual generative model.
-Opposite to what usually happens in ML, this model is available when learning is applied to IPC domains.
-The model learned for each domain is compared with its reference model using:
-
-\begin{itemize}
-\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
-Precision gives a notion of {\em soundness}.
-\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing). Recall gives a notion of {\em completeness}.
-\end{itemize}
-
-Table~\ref{fig:observationsnomap} shows the precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
-The reason why the scores in Table ~\ref{fig:observationsnomap} are lower than in Table~\ref{fig:observationstest} is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but correspond to {\em reformulations} of the actual model (changes in the roles of actions with matching headers or parameters with matching types).
-
-\subsubsection{Precision and recall robust to model reformulations}
-
-To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model {\em reformulations}.
-
-Precision and recall are often combined using the {\em harmonic mean}.
-This expression, called the {\em F-measure} or the balanced {\em F-score}, is defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
-Given the learned action model $\mathcal{M}$ and the reference action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure} (considering swaps in the actions with matching headers or parameters with matching types).
-Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
-These results evidence that in all of the evaluated domains, except for {\em ferry} and {\em satellite}, the learning task swaps the roles of some actions (or parameters) with respect to their role in the reference model.
-
-As we can see in Table~\ref{fig:observationsmap}, the {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
-On the other hand, the learning scores of several domains in Table~\ref{fig:observationsmap} are still lower than in Table~\ref{fig:observationstest}.
+The learning scores of several domains in Table~\ref{fig:observationsmap} are above the ones reported in Table~\ref{fig:observationsnomap}.
 The reason lies in the particular observations comprised by the test sets.
-%This leads us to the question on how the test set was generated and how the observations in this set affect the results in Table~\ref{fig:observationstest}.
 As an example, in the {\em driverlog} domain, the action schema {\small \tt disembark-truck} is missing from the learned model because this action is never induced from the observations in the training set; that is, such action never appears in the corresponding \emph{unobserved} plan.
@@
 
-%The generation of relevant observations of planning domains is still an open research question, planning domains are highly structured which makes that some states likely to have low probability of being chosen by chance~\cite{fern2004learning}.
-
-\begin{table}[hbt!]
-	\begin{center}
-		\begin{scriptsize}
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-				\hline
-
-				blocks & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 \\
-				driverlog & 0.0 & 0.0 & 0.25 & 0.43 & 0.0 & 0.0 & 0.08 & 0.14 \\
-				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
-				floortile & 0.38 & 0.55 & 0.4 & 0.18 & 0.56 & 0.45 & 0.44 & 0.39 \\
-				grid & 0.5 & 0.47 & 0.33 & 0.29 & 0.25 & 0.29 & 0.36 & 0.35 \\
-				gripper & 0.83 & 0.83 & 0.75 & 0.75 & 0.75 & 0.75 & 0.78 & 0.78 \\
-				hanoi & 0.5 & 0.25 & 0.5 & 0.5 & 0.0 & 0.0 & 0.33 & 0.25 \\
-				hiking & 0.43 & 0.43 & 0.5 & 0.35 & 0.44 & 0.47 & 0.46 & 0.42 \\
-				miconic & 0.6 & 0.33 & 0.33 & 0.25 & 0.33 & 0.33 & 0.42 & 0.31 \\
-				npuzzle & 0.33 & 0.33 & 0.0 & 0.0 & 0.0 & 0.0 & 0.11 & 0.11 \\
-				parking & 0.25 & 0.21 & 0.0 & 0.0 & 0.0 & 0.0 & 0.08 & 0.07 \\
-				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
-				transport & 1.0 & 0.3 & 0.8 & 0.8 & 1.0 & 0.6 & 0.93 & 0.57 \\
-				visitall & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
-				zenotravel & 0.67 & 0.29 & 0.33 & 0.29 & 0.33 & 0.14 & 0.44 & 0.24
-			\end{tabular}
-		\end{scriptsize}
-	\end{center}
-	\caption{\small Precision and recall values obtained without computing the $f_{P\&R}$ mapping with the reference model.}
-	\label{fig:observationsnomap}
-\begin{center}
-		\begin{scriptsize}
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-				\hline
-
-				blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				driverlog & 0.67 & 0.14 & 0.33 & 0.57 & 0.67 & 0.29 & 0.56 & 0.33 \\
-				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
-				floortile & 0.44 & 0.64 & 1.0 & 0.45 & 0.89 & 0.73 & 0.78 & 0.61 \\
-				grid & 0.63 & 0.59 & 0.67 & 0.57 & 0.63 & 0.71 & 0.64 & 0.62 \\
-				gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-				hiking & 0.78 & 0.6 & 0.93 & 0.82 & 0.88 & 0.88 & 0.87 & 0.77 \\
-				miconic & 0.8 & 0.44 & 1.0 & 0.75 & 1.0 & 1.0 & 0.93 & 0.73 \\
-				npuzzle & 0.67 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 0.89 \\
-				parking & 0.56 & 0.36 & 0.5 & 0.33 & 0.5 & 0.33 & 0.52 & 0.34 \\
-				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
-				transport & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0 & 0.63 \\
-				visitall & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 \\
-				zenotravel & 1.0 & 0.43 & 0.67 & 0.57 & 1.0 & 0.43 & 0.89 & 0.48
-			\end{tabular}
-		\end{scriptsize}
-	\end{center}
-	\caption{\small Precision and recall values obtained when computing the $f_{P\&R}$ mapping with the reference model.}
-	\label{fig:observationsmap}
-\end{table}
-
-\subsection{Learning from plans}
-We start evaluating our approach with tasks $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, where {\em labeled plans} are available.
-We then repeat the evaluation but exploiting potential \emph{static predicates} computed from $\Sigma$, which are the predicates that appear  unaltered in the initial and final states in every $\sigma_t\in\Sigma$.
-Static predicates are used to constrain the space of possible action models as explained in the previous section.
-
-Table~\ref{tab:results_plans} shows the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), while the last two columns of each setting and the last row report averages values.
-We can observe that identifying static predicates leads to models with better precondition {\em recall}.
-This fact evidences that many of the missing preconditions corresponded to static predicates because there is no incentive to learn them as they always hold~\cite{gregory2015domain}.
-
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
-All the learning tasks are solved in a few seconds.
-Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
-When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
-
-\begin{table*}[hbt!]
-		\resizebox{\textwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
-& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
-& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-                          \hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
-			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
-			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
-			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
-			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
-			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
-			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
-			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
-			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
-			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
-			\hline
-			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
-		\end{tabular}
-	}
-\caption{\small {\em Precision} and {\em recall} scores for learning tasks from labeled plans without (left) and with (right) static predicates.}
-\label{tab:results_plans}
-\end{table*}
-
-\begin{table}
-\begin{scriptsize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c||c|c|c|}
-                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
-			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
-                         \hline
-			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
-			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
-			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
-			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
-			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
-			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
-                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
-                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
-			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
-			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
-			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
-			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
-		\end{tabular}
-	\end{center}
-        \end{scriptsize}
-	\caption{\small Total planning time, preprocessing time and plan length for learning tasks from labeled plans without/with static predicates.}
-	\label{tab:time_plans}	
-\end{table}
-
-\subsection{Learning from partially specified action models}
-
-We evaluate now the ability of our approach to support partially specified action models; that is, addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
-In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task.
-
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, which include the identification of static predicates.
-We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
-In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is done only over half of the actions.
-This occurs, for instance, in the precondition \emph{recall} of domains such as {\em Floortile}, {\em Gripper} or {\em Satellite}.
-
-Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
-The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
-Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
-Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\Xi_0$).
-{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
-
-%In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
-
-\begin{table}[hbt!]
-\begin{footnotesize}
-	\begin{center}
-
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-				Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.90 \\
-				Ferry & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
-                Grid & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.84 & 0.78 \\
-				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-				Miconic & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Satellite & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
-				Transport & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
-				Zenotravel & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.78 \\
-				\hline
-				\bf  & 0.98 & 0.71 & 1.0 & 0.98 & 1.0 & 0.95 & 0.98 & 0.87 \\
-			\end{tabular}
-
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} scores for learning tasks with partially specified action models.}
-\label{tab:results_plans_partial}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-                         \hline
-			Blocks & 0.07 & 0.01 & 54  \\
-			Driverlog & 0.03 & 0.01 & 40 \\
-			Ferry & 0.06 & 0.03 & 45 \\
-			Floortile & 0.43 & 0.42 & 55 \\
-                        Grid & 3.12 & 3.07 & 53 \\
-			Gripper & 0.03 & 0.01 & 35 \\
-			Miconic & 0.03 & 0.01 & 34  \\
-			Satellite & 0.14 & 0.14 & 47 \\
-			Transport & 0.23 & 0.21 & 37 \\
-			Zenotravel & 0.90 & 0.89 & 40 \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Time and plan length learning for learning tasks with partially specified action models.}
-	\label{tab:time_plans_partial}	
-\end{table}
-
-\section{Conclussions}
+\section{Conclusions}
 \label{sec:Section9}
-
-\begin{small}
+We presented a novel approach for learning \strips\ action models from examples using classical planning.
+The approach is flexible to various amount and kind of input knowledge and accepts partially specified action models.
+Unlike extensive-data ML approaches, our work explores an alternative research direction to learn sound models from very small data sets.
+The action models of the {\em blocksworld} or {\em gripper} domains were perfectly learned from only 25 state observations.
+Moreover, in 12 out of the 15 domains, the learned models yield {\em Precision} values over $0.75$.
+
+To the best of our knowledge, this is the first work on learning \strips\ action models from state observations, using exclusively an {\em off-the-shelf} classical planner, and evaluated over a wide range of different domains.
+Recently, the work in~\cite{SternJ17} proposes a planning compilation for learning action models from plan traces following the {\em finite domain} representation for the state variables.
+This is a theoretical study on the boundaries of the learned models and no experimental results are reported.
+
+We also introduced the {\em precision} and {\em recall} metrics, widely used in ML, for evaluating the learned action models with respect to a given reference model.
+These two metrics measure the soundness and completeness of the learned models and facilitate the identification of model flaws.
+
+When example plans are available, we can compute accurate action models from small sets of learning examples in little computation time, less than a second.
+In many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
+With this regard, we extended our approach for learning also from state observations as it broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}.
+When action plans are not available, our approach still produces action models that are compliant with the input information.
+In this case, since learning is not constrained by actions, operators can be reformulated changing their semantics, in which case the comparison with a reference model turns out to be tricky.
+
+We also introduced a semantic method for evaluating the learned \strips\ action models with respect to observations of plan executions.
+Generating {\em informative} examples for learning planning action models is still an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions which have low probability of being chosen by chance~\cite{fern2004learning}.
+The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques that enable to autonomously collect informative learning examples.
+The combination of such exploration techniques with our learning approach is an appealing research direction that opens up the door to the bootstrapping of planning action models.
+
 \subsection*{Acknowledgment}
@@
 Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it RYC15/18009}, both programs funded by the Spanish government.
-\end{small}
 

### 2018-03-20 doc/ijcai-18/rebuttal/rebuttal.txt
added first version IJCAI18 rebuttal

Rebuttal #1:
-------------
Unlike extensive-data ML, this work explores an alternative research direction to learn sound models from small amounts of correct input data. Further, this is the first work on learning and evaluating STRIPS action models from state observations that report reproducible results over a wide range of domains exclusively using an off-the-shelf classical planner. We belief this approach is relevant by itself because opens up a way towards the bootstrapping of planning action models where a planner gradually learns/updates its action model.

Rebuttal #2:
-------------
We agree with the reviewer that a SAT-based compilation could be used for the addressed learning task. With this regard, the SAT-based system ARMS is the closest system to compare with (also requires the operator heads). ARMS  addresses however a different learning task: (1) learns from plan traces (the actions observations are assumed to be correct) while we learn from state observations and (2), ARMS addresses an optimization task (i.e. maximizing the number of covered examples) while ours is a satisfying task, we compute a conservative model that is guaranteed to be applicable at ALL the given observations.

Like happens with satisfying planners, our approach is extensible to address the corresponding optimization task, e.g. in our case minimizing the number of model edits required to cover a set of partially-observable/noisy state observations. This work aimed at understanding the grounds of the compilation for learning and evaluating STRIPS models in fully-observable environments (e.g. classical planning tasks, video-games, simulators) but extending this approach to partially-observable/noisy environments is in our research agenda. With this said, we bet on classical planning because we believe that existing relaxations of the classical planning task, such as the "Relaxed plan" have an important role to play in the assessment of generative models with regard to large amounts of partially-observable/noisy examples. Further, when there is an unknown number of missing observations, planning seems more natural than SAT because the planning horizon is no longer known.

Answering the reviewer questions:
1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) because we obtained better experimental results (we attach the results we got when starting from the most general hypothesis (no preconditions). 
2. We agree with the reviewer that the "one test_i fluent" is a more natural encoding. We will evaluate it to see if we achieve some performance gain.

Rebuttal #3:
-------------
The contribution of our approach is leveraging classical planning not only for (1) LEARNING, that allows us to report quantitative results over a wide range of different domains but also for (2), EVALUATING the quality of a STRIPS model with regard to a test set of observations.

Evaluating STRIPS models with a test set of observations enables the model recognition task. This is relevant because of the high expressiveness of STRIPS (as expressive as a Turing Machine with finite tape). Different generative models like policies, programs or grammars are STRIPS compilable (see the series of work by Jorge Baier, Sheila McIlraith et al). Therefore, this work poses a general method to assess how a given generative model matches a test set of observations: for a wide range of different generative models (the STRIPS compilable ones), and exclusively using classical planning. As far as we know, this is the first work on doing so.

Rebuttal #4:
-------------
The empirical evaluation focused on understanding which kind of STRIPS action models could be learned from just 25 state observations. The results revealed that 25 state observations were enough to learn action models with few action schemes,  (e.g "blocks" or "gripper"). In some of these domains ("hanoi", "visitall", "npuzzle") the reported Precision&Recall did not reach 1.0 for the preconditions. The reason is that learning discovered preconditions that were not specified in the actual/reference domain. For instance, our approach learned for the "move" action in the "visitall" the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1), while only the (connected ?place1 ?place2) precondition is present in the actual/reference model, despite both preconditions are actually correct.

Learning accurate models from few examples becomes harder in domains with larger number of action schemes, in these cases the different actions have lower chances of being chosen. Examples are the "disembark-truck" action missing from the "driverlog" domain, the "paint-down" action missing from the "floortile" domain or the "move-curb-to-curb" from the "parking" domain. Including more examples, specially examples of these missing actions, definitely leads to better strips action models.

Because our approach is a classical planning compilation, the limit on the number of input observations is given by the limitation of the planner performance. In the particular case of current SAT-based planners, they are known to have performance issues with planning horizons beyond the 150-200 steps (this would currently limit our approach to about a hundred of input examples).

With this said, we agree with the reviewer that the set of input problems definitely affects to the quality of the learned models. Currently were are selecting problems for each domain with a small number of objects because they are easier to be solved and the result of the compilation is more tractable. An interesting research question is identifying which is the smallest number of objects per type required to learn a complete STRIPS action model for a given domain. This question has recently been addressed for Schematic Invariants (Schematic Invariants by Reduction to Ground Invariants, J. Rintanen IJCAI 2017). We believe that this research direction is promising to obtain similar results for learning STRIPS action models.

### 2018-03-21 doc/ijcai-18/rebuttal/rebuttal.txt
added todays comments to IJCAI18 rebuttal

@@
 -------------
-Unlike extensive-data ML, this work explores an alternative research direction to learn sound models from small amounts of correct input data.
-Further, this is the first work on learning and evaluating STRIPS action models from state observations that report reproducible results over a wide range of domains exclusively using an off-the-shelf classical planner.
-We belief this approach is relevant by itself because opens up a way towards the bootstrapping of planning action models where a planner gradually learns/updates its action model.
+We sincerely thank reviewers for their constructive comments and criticisms which help to clarify details of our approach.
+Unlike extensive-data ML, our approach explores an alternative research direction to learn sound action models from a small amount of correct input data.
+Further, this is the first work that learns and evaluates action models exclusively using an off-the-shelf classical planner.
+We belief this result is relevant by itself because it opens up a way towards the bootstrapping of planning action models where a planner gradually learns/updates its action model.
+
+The SAT-based system ARMS is the closest system to compare with (also requires the operator heads).
+ARMS addresses however a different learning task: (1) learns from plan traces (the actions observations are assumed to be correct) while we learn from state observations (2), ARMS addresses an optimization task (i.e.
+maximizing the # of covered examples) while ours is a satisfying task, we compute a conservative model that is guaranteed to be applicable at ALL the given observations and (3), ARMS evaluates the quality of the learned models using two different functions "error rate" that evaluates the quality of the learned preconditions and "redundancy" that evaluates the quality of the positive effects.
 
@@
 -------------
-We agree with the reviewer that a SAT-based compilation could be used for the addressed learning task.
-With this regard, the SAT-based system ARMS is the closest system to compare with (also requires the operator heads).
-ARMS  addresses however a different learning task: (1) learns from plan traces (the actions observations are assumed to be correct) while we learn from state observations and (2), ARMS addresses an optimization task (i.e.
-maximizing the number of covered examples) while ours is a satisfying task, we compute a conservative model that is guaranteed to be applicable at ALL the given observations.
+In effect a SAT-based compilation could also be used for the learning and evaluation tasks addressed in this paper.
+We bet on classical planning because when there is an unknown # of missing observations, planning seems more natural than SAT because the planning horizon is no longer known and addressing these tasks is in our research agenda.
+Further, the planning commmunity has shown a tremendous experince in the develpment of informative while inexpensiver heuristic functions able to evaluate millions of symbolic states in seconds time.
+We believe that existing relaxations of the classical planning task, such as the "Relaxed plan" have an important role to play in the assessment of generative models with regard to large amounts of partially-observable/noisy examples.
 
 Like happens with satisfying planners, our approach is extensible to address the corresponding optimization task, e.g.
-in our case minimizing the number of model edits required to cover a set of partially-observable/noisy state observations.
-This work aimed at understanding the grounds of the compilation for learning and evaluating STRIPS models in fully-observable environments (e.g.
-classical planning tasks, video-games, simulators) but extending this approach to partially-observable/noisy environments is in our research agenda.
-With this said, we bet on classical planning because we believe that existing relaxations of the classical planning task, such as the "Relaxed plan" have an important role to play in the assessment of generative models with regard to large amounts of partially-observable/noisy examples.
-Further, when there is an unknown number of missing observations, planning seems more natural than SAT because the planning horizon is no longer known.
+in our case minimizing the # of model edits required to cover a set of partially-observable/noisy state observations.
+As a first step, this work aimed at understanding the grounds of the compilation for learning and evaluating STRIPS models in fully-observable environments (e.g.
+classical planning tasks, video-games, simulators) but extending this approach to partially-observable/noisy environments is definitely in agenda.
 
 Answering the reviewer questions:
-1.
-We learn actions preconditions starting from the most specific hypothesis (all preconditions) because we obtained better experimental results (we attach the results we got when starting from the most general hypothesis (no preconditions).
-2. We agree with the reviewer that the "one test_i fluent" is a more natural encoding. We will evaluate it to see if we achieve some performance gain.
+1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained better experimental results. 
+2. We agree with the reviewer that the "one test_i fluent" is a more natural encoding. We plan to evaluate and chekc whether we achieve some performance gain.
 
@@
 -------------
-The contribution of our approach is leveraging classical planning not only for (1) LEARNING, that allows us to report quantitative results over a wide range of different domains but also for (2), EVALUATING the quality of a STRIPS model with regard to a test set of observations.
+The contribution of our approach is leveraging classical planning not only for (1) LEARNING, that allows us to report quantitative results over a wide range of different domains (15 by the time being despite we are working to cover any IPC domain with the STRIPS requirement) but also for (2), EVALUATING how well a given STRIPS model matches a test set of observations.
 
-Evaluating STRIPS models with a test set of observations enables the model recognition task.
+Evaluating STRIPS models according to a test set of observations enables the "STRIPS model recognition" task.
 This is relevant because of the high expressiveness of STRIPS (as expressive as a Turing Machine with finite tape).
-Different generative models like policies, programs or grammars are STRIPS compilable (see the series of work by Jorge Baier, Sheila McIlraith et al).
-Therefore, this work poses a general method to assess how a given generative model matches a test set of observations: for a wide range of different generative models (the STRIPS compilable ones), and exclusively using classical planning.
-As far as we know, this is the first work on doing so.
+Different generative models like policies, programs, grammars and different forms of domain-specific control knowledge are STRIPS compilable (see the series of work by Jorge Baier, Sheila McIlraith et al).
+With this regard, our work poses a general framework to assess how a given generative model (provided that it is STRIPS compilable) matches a test set of observations.
 
@@
 -------------
-The empirical evaluation focused on understanding which kind of STRIPS action models could be learned from just 25 state observations.
-The results revealed that 25 state observations were enough to learn action models with few action schemes,  (e.g "blocks" or "gripper").
-In some of these domains ("hanoi", "visitall", "npuzzle") the reported Precision&Recall did not reach 1.0 for the preconditions.
-The reason is that learning discovered preconditions that were not specified in the actual/reference domain.
-For instance, our approach learned for the "move" action in the "visitall" the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1), while only the (connected ?place1 ?place2) precondition is present in the actual/reference model, despite both preconditions are actually correct.
+The evaluation focused on understanding which models can be learned from just 25 observations.
+Results revealed that 25 was not enough to learn action models in the domains with the largest # of schemas (this # ranges [1,7] in our evaluation).
+The largest this # the lowest chance an operator of being chosen (e.g.
+"disembark-truck" action is missing from the "driverlog" domain or "paint-down" is missing from "floortile").
+Including more observations, specially of the missing actions, leads to better action models.
+Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance.
+Current SAT-based planners have performance issues with planning horizons beyond 150-200 steps (since actions for programming preconditions/effects are often applied in parallel, in a single step, this means we can handle 50-80 input observations with a SAT-based planner.
 
-Learning accurate models from few examples becomes harder in domains with larger number of action schemes, in these cases the different actions have lower chances of being chosen.
-Examples are the "disembark-truck" action missing from the "driverlog" domain, the "paint-down" action missing from the "floortile" domain or the "move-curb-to-curb" from the "parking" domain.
-Including more examples, specially examples of these missing actions, definitely leads to better strips action models.
+Precision&Recall did not reach 1.0 in some domains with few schemes because learning "discovered" preconditions that were not specified in the IPC/reference domain.
+For instance, we learned the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in "visitall", while only (connected ?place1 ?place2) was specified in the IPC/reference model.
 
-Because our approach is a classical planning compilation, the limit on the number of input observations is given by the limitation of the planner performance.
-In the particular case of current SAT-based planners, they are known to have performance issues with planning horizons beyond the 150-200 steps (this would currently limit our approach to about a hundred of input examples).
-
-With this said, we agree with the reviewer that the set of input problems definitely affects to the quality of the learned models.
-Currently were are selecting problems for each domain with a small number of objects because they are easier to be solved and the result of the compilation is more tractable.
-An interesting research question is identifying which is the smallest number of objects per type required to learn a complete STRIPS action model for a given domain.
+The input observations affects to the quality of the learned models.
+Were are using one problem for each domain with a small # of objects because the compiled task is more tractable.
+If there is IPC instances small enough, we use an IPC instance, otherwise we created a small problem.
+An interesting research question is identifying which is the smallest # of objects per type required to learn a complete STRIPS action model for a given domain.
 This question has recently been addressed for Schematic Invariants (Schematic Invariants by Reduction to Ground Invariants, J.

### 2018-03-22 doc/ijcai-18/rebuttal/rebuttal.txt
updates added to IJCAI18 rebuttal

@@
-Rebuttal #1:
+Rebuttal #12637:
 -------------
 We sincerely thank reviewers for their constructive comments and criticisms which help to clarify details of our approach.
-Unlike extensive-data ML, our approach explores an alternative research direction to learn sound action models from a small amount of correct input data.
-Further, this is the first work that learns and evaluates action models exclusively using an off-the-shelf classical planner.
+This is the first work that LEARNS and EVALUATES action models exclusively using an off-the-shelf classical planner.
 We belief this result is relevant by itself because it opens up a way towards the bootstrapping of planning action models where a planner gradually learns/updates its action model.
 
-The SAT-based system ARMS is the closest system to compare with (also requires the operator heads).
-ARMS addresses however a different learning task: (1) learns from plan traces (the actions observations are assumed to be correct) while we learn from state observations (2), ARMS addresses an optimization task (i.e.
-maximizing the # of covered examples) while ours is a satisfying task, we compute a conservative model that is guaranteed to be applicable at ALL the given observations and (3), ARMS evaluates the quality of the learned models using two different functions "error rate" that evaluates the quality of the learned preconditions and "redundancy" that evaluates the quality of the positive effects.
+The SAT-based system ARMS (AIJ07) is the closest system to compare with, since ARMS also requires the operator heads and observations of correct plans.
+ARMS addresses however a different learning task:
+(1) ARMS learns from plan traces (where the ACTIONS observations are assumed to be correct) while we learn from STATE observations.
+(2) ARMS addresses an OPTIMIZATION task (i.e.
+maximizing the # of covered examples) while ours is a SATISFYING task.
+We compute a conservative model that is guaranteed to be applicable at all the given observations.
 
-Rebuttal #2:
+Rebuttal #23413:
 -------------
-In effect a SAT-based compilation could also be used for the learning and evaluation tasks addressed in this paper.
-We bet on classical planning because when there is an unknown # of missing observations, planning seems more natural than SAT because the planning horizon is no longer known and addressing these tasks is in our research agenda.
-Further, the planning commmunity has shown a tremendous experince in the develpment of informative while inexpensiver heuristic functions able to evaluate millions of symbolic states in seconds time.
-We believe that existing relaxations of the classical planning task, such as the "Relaxed plan" have an important role to play in the assessment of generative models with regard to large amounts of partially-observable/noisy examples.
+In effect, we definitely agree with the reviewer that a SAT-based compilation could also be used for the learning and evaluation tasks addressed in this paper.
+The only reason why we bet on planning is because, when there is an unknown # of missing observations, the planning horizon is no longer known and addressing this learning task is in our research agenda.
+Further, the planning community has shown terrific expertise in the development of informative while inexpensive heuristics (current heuristics can evaluate millions of symbolic states in seconds time).
+Our believe is that this technology has a key role to play in the assessment of generative models with regard to large amounts of incomplete/noisy examples.
 
-Like happens with satisfying planners, our approach is extensible to address the corresponding optimization task, e.g.
-in our case minimizing the # of model edits required to cover a set of partially-observable/noisy state observations.
-As a first step, this work aimed at understanding the grounds of the compilation for learning and evaluating STRIPS models in fully-observable environments (e.g.
-classical planning tasks, video-games, simulators) but extending this approach to partially-observable/noisy environments is definitely in agenda.
+Like happens with satisfying planners, our approach is extensible to the corresponding optimization task, e.g.
+in our case minimizing the # of model edits required to cover a set of observations.
+As a first step, this paper aimed at understanding the grounds of our compilation for learning and evaluating STRIPS models in fully-observable environments (e.g.
+classical planning tasks) but, extending this approach to partially-observable/noisy environments, is certainly in our agenda.
 
 Answering the reviewer questions:
-1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained better experimental results. 
-2. We agree with the reviewer that the "one test_i fluent" is a more natural encoding. We plan to evaluate and chekc whether we achieve some performance gain.
+1.
+We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained slightly better experimental results.
+There is nothing that prevent us from starting with the most general hypothesis (zero preconditions).
+In fact the evaluation task starts with a given action model that may have for instance, no preconditions.
+2. We agree with the reviewer that the "one test_i fluent" encoding is more natural. We plan to evaluate it and check whether it allow us to achieve some performance gain.
+To comments on the training set we refer the reviewer to the Rebuttal #26170.
 
-Rebuttal #3:
+Rebuttal #25749:
 -------------
-The contribution of our approach is leveraging classical planning not only for (1) LEARNING, that allows us to report quantitative results over a wide range of different domains (15 by the time being despite we are working to cover any IPC domain with the STRIPS requirement) but also for (2), EVALUATING how well a given STRIPS model matches a test set of observations.
+The contribution of our approach is leveraging classical planning not only for LEARNING, that allows us to report quantitative results over a wide range of different domains (15 by the time being despite we are working to cover any IPC domain with the STRIPS requirement) but also for EVALUATING how well a given STRIPS model matches a test set of observations.
 
-Evaluating STRIPS models according to a test set of observations enables the "STRIPS model recognition" task.
+Evaluating a given STRIPS model according to a test set of observations enables the "STRIPS model recognition" task.
 This is relevant because of the high expressiveness of STRIPS (as expressive as a Turing Machine with finite tape).
 Different generative models like policies, programs, grammars and different forms of domain-specific control knowledge are STRIPS compilable (see the series of work by Jorge Baier, Sheila McIlraith et al).
-With this regard, our work poses a general framework to assess how a given generative model (provided that it is STRIPS compilable) matches a test set of observations.
+With this regard, our work poses a general framework to assess how a given generative model (provided that it is STRIPS compilable) matches a given set of observations.
 
-Rebuttal #4:
+Rebuttal #26170:
 -------------
-The evaluation focused on understanding which models can be learned from just 25 observations.
-Results revealed that 25 was not enough to learn action models in the domains with the largest # of schemas (this # ranges [1,7] in our evaluation).
-The largest this # the lowest chance an operator of being chosen (e.g.
-"disembark-truck" action is missing from the "driverlog" domain or "paint-down" is missing from "floortile").
-Including more observations, specially of the missing actions, leads to better action models.
+Our evaluation focused on understanding which models can be learned from just 25 observations.
+Results revealed that 25 was not enough to learn action models in the domains with the largest # of schemas.
+In our evaluation this # ranges [1,7].
+The largest this # the lowest chance of choosing an operator, e.g.
+"disembark-truck" is missing from the driverlog or "paint-down" is missing from floortile.
+Including more observations, specially of the missing schemes, leads to better models.
 Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance.
-Current SAT-based planners have performance issues with planning horizons beyond 150-200 steps (since actions for programming preconditions/effects are often applied in parallel, in a single step, this means we can handle 50-80 input observations with a SAT-based planner.
+Current SATplanners have performance issues with horizons beyond 150-200 steps.
+Given that actions for programming preconds/effects are often applied in parallel, in a single step, this means we could handle no more than a hundred observations with current SATplanners (we require 2 actions to validate 1 observation, Fig 3).
 
-Precision&Recall did not reach 1.0 in some domains with few schemes because learning "discovered" preconditions that were not specified in the IPC/reference domain.
-For instance, we learned the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in "visitall", while only (connected ?place1 ?place2) was specified in the IPC/reference model.
+Precision&Recall did not reach 1.0 in some domains with few schemes because learning discovered preconditions that were not specified in the IPC/reference domain.
+For instance, we learned the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in visitall, while only (connected ?place1 ?place2) was specified in the IPC model.
+Effects were perfectly learned in these domains.
 
-The input observations affects to the quality of the learned models.
-Were are using one problem for each domain with a small # of objects because the compiled task is more tractable.
-If there is IPC instances small enough, we use an IPC instance, otherwise we created a small problem.
-An interesting research question is identifying which is the smallest # of objects per type required to learn a complete STRIPS action model for a given domain.
-This question has recently been addressed for Schematic Invariants (Schematic Invariants by Reduction to Ground Invariants, J.
-Rintanen IJCAI 2017).
-We believe that this research direction is promising to obtain similar results for learning STRIPS action models.
+The input observations affect to the quality of the learned models.
+Were use 1 problem for each domain with a small # of objects because this way the compiled task is more tractable.
+If IPC instances are small enough we use them otherwise, we created a smaller problem.
+An open question is identifying the smallest # of objects per type required to learn a complete STRIPS model.
+This has recently been addressed for Schematic Invariants (Schematic Invariants by Reduction to Ground Invariants, Rintanen IJCAI17).
+We believe this research direction is promising to obtain similar results for learning STRIPS models.

### 2018-03-22 doc/ijcai-18/rebuttal/rebuttal_March22_Sergio_Diego_Eva.txt
Add files via upload

Rebuttal #12637:
-------------

We would like to thank all the reviewers for their thoughtful comments and insights as well as expressing our general agreement with the provided comments.

Lack of comparison
----------------------

The SAT-based system ARMS (AIJ07) is the closest system to compare with since ARMS also requires the operator heads and observations of correct plans. ARMS addresses, however, a different learning task:

(1) ARMS learns from plan traces (where the observed ACTIONS are assumed to be correct) while we learn from STATE observations (unobserved actions).
(2) ARMS addresses an OPTIMIZATION task (i.e. maximizing the # of covered examples) while ours is a SATISFYING task. We compute a conservative model that is guaranteed to be applicable at all the given observations.

As for LOCM, we think a better comparison would be with LOP since LOCM only learns the dynamic aspects of the domain (i.e. state changes that occur due to action application but not the static relations). However, LOP needs a set of optimal plans as input to the learning system.

Rebuttal #23413:
-------------
We definitely agree with the reviewer that a SAT-based compilation could also be used for learning and evaluating the tasks addressed in this paper. The only reason why we opt for planning is because when there is an unknown # of actions between state observations, the planning horizon is no longer known and this learning task is in our research agenda. Further, the planning community has shown terrific expertise in the development of informative and inexpensive heuristics (current heuristics are able to evaluate millions of symbolic states in seconds). Our belief is that this technology will play a key role in the assessment of generative models with inputs of large amounts of incomplete/noisy examples.

Like happens with satisfying planners, our approach is extensible to the corresponding optimization task (in our case this amounts to minimizing the # of model edits required to cover a set of observations). As a first step, this paper aims at understanding the grounds of the compilation scheme for learning and evaluating models in fully-observable environments (e.g. classical planning tasks); but, extending this approach to partially-observable/noisy environments, is certainly in our agenda.

Answers to the reviewer's questions:

1. We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained slightly better experimental results. Nothing prevents us from starting with the most general hypothesis (zero preconditions), a random model, or even a previously learnt model (as it is the case of the evaluation task). In fact, the evaluation task starts with a given action model that may contain, for instance, no preconditions at all.

2. We thank the reviewer for the suggestion that the "one test_i fluent" encoding is more natural. We will evaluate it and check if this brings some performance gain.

Regarding the remarks on how the training set was generated, we refer the reviewer to the Rebuttal #26170.

Rebuttal #25749:
-------------
The contribution of our approach is leveraging classical planning not only for LEARNING, that allows us to report quantitative results over a wide range of different domains but also for EVALUATING how well a given STRIPS model matches a test set of observations.

Evaluating a given model according to a test set of observations represents a step towards solving the "model recognition" task. Given the high expressivenes of the STRIPS models (as expressive as a Turing Machine with finite tape), the task of model recognition is certainly relevant. Different generative models like policies, programs, grammars or different forms of domain-specific control knowledge are STRIPS compilable (as stated in the series of work by Jorge Baier, Sheila McIlraith et al). In this sense, our work poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) in a given set of observations.

Regarding the remarks on the lack of comparison with other approaches, we refer the reviewer to the Rebuttal #12637.

Rebuttal #26170:
---------------

Generation of the training set
------------------------------

The 25 observations per domain were generated by solving one instance of the IPC; in some cases we modified the problem and limited the number of objects to keep the computation time low. Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance. Current SATplanners have performance issues with horizons beyond 150-200 steps. Given that actions for programming preconds/effects are often applied in parallel, in a single step, this means we could handle no more than a hundred observations with current SATplanners (note that we require 2 actions to validate one observation, Fig 3). Ultimately, our goal was to evaluate how well the approach performs with few observations and so it is more important to have representative samples than a large number of observations, which would undoubtedly lead to better models but also longer learning times.

Given the small size of the training set, there is no guarantee that the unobserved plan contains actions for all the domain schemas, which clearly has an impact in some results (e.g, the missing actions “disembark-truck” and “paint-down” of the driverlog and floortile domains, respectively). Another reason that explains that domains are not perfectly learnt (wrt the reference domain) is the appearance in the output models of preconditions that although correct they are usually skipped in the domain definition because of redundancy. For instance, we learned the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in visitall, but only (connected ?place1 ?place2) is specified in the IPC model.

An open question is the identification of the smallest # of objects per type required to learn a complete STRIPS model. This has recently been addressed for Schematic Invariants (Rintanen IJCAI17). We believe this is a promising research direction to obtain similar results in learning STRIPS models.

Regarding the remarks on the lack of comparison with other approaches, we refer the reviewer to the Rebuttal #12637.

### 2018-03-23 doc/ijcai-18/rebuttal/rebuttal.txt
rebuttal is ready for submission

@@
 -------------
-We sincerely thank reviewers for their constructive comments and criticisms which help to clarify details of our approach.
-This is the first work that LEARNS and EVALUATES action models exclusively using an off-the-shelf classical planner.
-We belief this result is relevant by itself because it opens up a way towards the bootstrapping of planning action models where a planner gradually learns/updates its action model.
 
-The SAT-based system ARMS (AIJ07) is the closest system to compare with, since ARMS also requires the operator heads and observations of correct plans.
-ARMS addresses however a different learning task:
-(1) ARMS learns from plan traces (where the ACTIONS observations are assumed to be correct) while we learn from STATE observations.
+We would like to thank all the reviewers for their thoughtful comments and insights as well as expressing our general agreement with the provided comments.
+
+The SAT-based system ARMS (AIJ07) is the closest system to compare with since ARMS also requires the operator heads and observations of correct plans.
+ARMS addresses, however, a different learning task:
+
+(1) ARMS learns from plan traces (where the observed ACTIONS are assumed to be correct) while we learn from STATE observations (the executed actions are unobserved).
 (2) ARMS addresses an OPTIMIZATION task (i.e.
@@
 
+As for LOCM, we think a better comparison would be with LOP since LOCM only learns the dynamic aspects of the domain (i.e.
+state changes that occur due to action application but not the static relations).
+However, LOP needs a set of optimal plans as input to the learning system.
+
 Rebuttal #23413:
 -------------
-In effect, we definitely agree with the reviewer that a SAT-based compilation could also be used for the learning and evaluation tasks addressed in this paper.
-The only reason why we bet on planning is because, when there is an unknown # of missing observations, the planning horizon is no longer known and addressing this learning task is in our research agenda.
-Further, the planning community has shown terrific expertise in the development of informative while inexpensive heuristics (current heuristics can evaluate millions of symbolic states in seconds time).
-Our believe is that this technology has a key role to play in the assessment of generative models with regard to large amounts of incomplete/noisy examples.
+We definitely agree with the reviewer that a SAT compilation could also be used for learning and evaluating the tasks addressed in this paper.
+The only reason why we opt for planning is because when there is an unknown # of actions between state observations, the planning horizon is no longer known and this learning task is in our research agenda.
+Further, the planning community has shown terrific expertise in the development of informative/inexpensive heuristics (current heuristics can evaluate millions of symbolic states in seconds).
+Our belief is that this technology can play a key role in the assessment of generative models with large amounts of incomplete/noisy examples.
 
-Like happens with satisfying planners, our approach is extensible to the corresponding optimization task, e.g.
-in our case minimizing the # of model edits required to cover a set of observations.
-As a first step, this paper aimed at understanding the grounds of our compilation for learning and evaluating STRIPS models in fully-observable environments (e.g.
-classical planning tasks) but, extending this approach to partially-observable/noisy environments, is certainly in our agenda.
+Like happens with satisfying planners, our approach is extensible to the corresponding optimization task (in our case this amounts to minimizing the # of model edits required to cover a set of observations).
+As a first step, this paper aims at understanding the grounds of the compilation scheme for learning and evaluating models in fully-observable environments (e.g.
+classical planning tasks); but, extending this approach to partially-observable/noisy environments, is certainly in our agenda.
 
-Answering the reviewer questions:
+Answers to the reviewer's questions:
 1.
 We learn actions preconditions starting from the most specific hypothesis (all preconditions) just because we obtained slightly better experimental results.
-There is nothing that prevent us from starting with the most general hypothesis (zero preconditions).
-In fact the evaluation task starts with a given action model that may have for instance, no preconditions.
-2. We agree with the reviewer that the "one test_i fluent" encoding is more natural. We plan to evaluate it and check whether it allow us to achieve some performance gain.
-To comments on the training set we refer the reviewer to the Rebuttal #26170.
+Nothing prevents us from starting with the most general hypothesis (zero preconditions), a random model, or even a previously learnt model (as it is the case of the evaluation task).
+In fact, the evaluation task starts with a given action model that may contain, for instance, no preconditions at all.
+
+2. We thank the reviewer for the suggestion. We agree that the "one test_i fluent" encoding is more natural. We will evaluate it and check if this brings some performance gain.
+
+Regarding the remarks on how the training set was generated, we refer the reviewer to our rebuttal to Review  #26170.
 
@@
 -------------
-The contribution of our approach is leveraging classical planning not only for LEARNING, that allows us to report quantitative results over a wide range of different domains (15 by the time being despite we are working to cover any IPC domain with the STRIPS requirement) but also for EVALUATING how well a given STRIPS model matches a test set of observations.
+The contribution of our approach is leveraging classical planning not only for LEARNING, that allows us to report quantitative results over a wide range of different domains but also for EVALUATING how well a given STRIPS model matches a test set of observations.
 
-Evaluating a given STRIPS model according to a test set of observations enables the "STRIPS model recognition" task.
-This is relevant because of the high expressiveness of STRIPS (as expressive as a Turing Machine with finite tape).
-Different generative models like policies, programs, grammars and different forms of domain-specific control knowledge are STRIPS compilable (see the series of work by Jorge Baier, Sheila McIlraith et al).
-With this regard, our work poses a general framework to assess how a given generative model (provided that it is STRIPS compilable) matches a given set of observations.
+Evaluating a given model according to a test set of observations represents a big step towards the "model recognition" task.
+Given the high expressiveness of the STRIPS models (as expressive as a Turing Machine with finite tape), the task of model recognition is certainly relevant.
+Different generative models like policies, programs, grammars or different forms of domain-specific control knowledge are STRIPS compilable (as stated in the series of work by Jorge Baier, Sheila McIlraith et al).
+In this sense, our work poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations.
+
+Regarding the remarks on the lack of comparison with other approaches, we refer the reviewer to our rebuttal to Review #12637.
 
 Rebuttal #26170:
--------------
-Our evaluation focused on understanding which models can be learned from just 25 observations.
-Results revealed that 25 was not enough to learn action models in the domains with the largest # of schemas.
-In our evaluation this # ranges [1,7].
-The largest this # the lowest chance of choosing an operator, e.g.
-"disembark-truck" is missing from the driverlog or "paint-down" is missing from floortile.
-Including more observations, specially of the missing schemes, leads to better models.
+---------------
+
+We generated 25 training observations per domain by solving 1 IPC instance per domain; in some cases we modified that instance, limiting the number of objects, to keep computation time low.
 Because our approach is a planning compilation, the limit on the # of observations is given by the planner performance.
 Current SATplanners have performance issues with horizons beyond 150-200 steps.
-Given that actions for programming preconds/effects are often applied in parallel, in a single step, this means we could handle no more than a hundred observations with current SATplanners (we require 2 actions to validate 1 observation, Fig 3).
+Given that actions for programming preconds/effects are often applied in parallel (in 1 step) this means we could handle no more than a hundred observations with current SATplanners (recall we require 2 actions to validate 1 observation, Fig 3).
 
-Precision&Recall did not reach 1.0 in some domains with few schemes because learning discovered preconditions that were not specified in the IPC/reference domain.
-For instance, we learned the preconditions (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in visitall, while only (connected ?place1 ?place2) was specified in the IPC model.
-Effects were perfectly learned in these domains.
+Given the small size of the training set, there is no guarantee that the unobserved plan contains actions for all the domain schemas, which clearly affects some results (e.g, actions 'disembark-truck' and 'paint-down' are missing from the driverlog and floortile domains, respectively).
+We aimed evaluating how well our approach performs with just 25 observations.
+A larger # of observations, specially of the missing actions, would lead to better models but also longer learning times.
 
-The input observations affect to the quality of the learned models.
-Were use 1 problem for each domain with a small # of objects because this way the compiled task is more tractable.
-If IPC instances are small enough we use them otherwise, we created a smaller problem.
+Another reason that explains that models are not perfectly learnt (wrt the IPC domains) is the appearance in the learnt models of preconds that although correct, are skipped in the IPC models because of redundancy.
+For instance, we learned preconds (connected ?place1 ?place2) and (connected ?place2 ?place1) for the "move" action in visitall, but only (connected ?place1 ?place2) is specified in the IPC model.
+
 An open question is identifying the smallest # of objects per type required to learn a complete STRIPS model.
-This has recently been addressed for Schematic Invariants (Schematic Invariants by Reduction to Ground Invariants, Rintanen IJCAI17).
-We believe this research direction is promising to obtain similar results for learning STRIPS models.
+This has recently been addressed for Schematic Invariants (Rintanen IJCAI17).
+We believe this is a promising research direction to obtain similar results for learning STRIPS models.
+Regarding the remarks on the lack of comparison with other approaches, we refer the reviewer to our rebuttal to Review #12637.

### 2018-03-23 doc/ijcai-18/rebuttal/rebuttal_March22_Sergio_Diego_Eva.txt
rebuttal is ready for submission

@@
 
-Like happens with satisfying planners, our approach is extensible to the corresponding optimization task (in our case this amounts to minimizing the # of model edits required to cover a set of observations).
+As it happens with satisfying planners, our approach is extensible to the corresponding optimization task (in our case this amounts to minimizing the # of model edits required to cover a set of observations).
 As a first step, this paper aims at understanding the grounds of the compilation scheme for learning and evaluating models in fully-observable environments (e.g.
@@
 Nothing prevents us from starting with the most general hypothesis (zero preconditions), a random model, or even a previously learnt model (as it is the case of the evaluation task).
-In fact, the evaluation task starts with a given action model that may contain, for instance, no preconditions at all.
 
@@
 Given that actions for programming preconds/effects are often applied in parallel, in a single step, this means we could handle no more than a hundred observations with current SATplanners (note that we require 2 actions to validate one observation, Fig 3).
-Ultimately, our goal was to evaluate how well the approach performs with few observations and so it is more important to have representative samples than a large number of observations, which would undoubtedly lead to better models but also longer learning times.
+Ultimately, our interest was to evaluate the scope of the compilation scheme when learning with few observations.
+Thereby, having representative samples is more relevant to us than having a large number of observations, which would undoubtedly lead to better models but also longer learning times.
 

### 2018-03-23 doc/aij18/aij18.tex
policy/goal learning added to aij18 paper

@@
 \item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$. 
-In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes.
-This means that some preconditions and some positive or negative effects are a priori known.
 \item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
@@
   \item When learning from a set of observed plans, the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given sequence of example plans producing the corresponding sequence of state observations $\mathcal{O}$.
+\end{itemize}
+
+\subsection{Learning from a partially specified model}
+In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes.
+This means that some preconditions and some positive or negative effects are a priori known.
+
+Such scenario is specially relevant to addressing two well-known learning tasks:
+\begin{itemize}
+\item {\em Policy learning}.
+A policy is function that maps states into actions and represent the conditions under which an action should be applied to achieve certain goals.
+Given an action model $\mathcal{M}$ the task of learning a policy that is compliant with a given set of observations of plan executions can be defined as the task of learning extra preconditions for each of the action schemes $\xi\in\mathcal{M}$.
+These extra preconditions capture when actions can be applied according to the policy.
+If for each possible state there is only one applicable action we say that the policy is fully specified otherwise, we say that the policy is partially specified.
+In the general case, policy learning for arbitrary planning tasks is a complex task because it may require computing new {\em high-level state features}~\cite{lotinac2016automatic}.
+
+\item {\em Goal learning}.
+The goal condition is the set of constraints that defines the subset of states that are considered goal states.
+Given an action model $\mathcal{M}$ the task of learning the goal conditions for a a given set of observations of plan executions can be defined as learning the preconditions of a extra action that is always applied in the last place and whose only effect is marking that the goals of the planning task are reached.
 \end{itemize}

### 2018-04-12 doc/aij18/aij18.tex
small updates to sections 1,2,3 aij paper

@@
 
-\title{Learning and Evaluating \strips\ Action Models with Classical Planning}
+\title{Learning, Evaluation and Recognition of \strips\ Action Models with Classical Planning}
 \author[label1]{Diego Aineto}
@@
 What is more, the compilation is extensible to assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions.
-The performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and assessing the learned models with respect to (1), test sets of observations of plan executions and (2), the true models.
+This is relevant since it allows us to assessing the learned models with respect to the true models but with respect to test sets of observations of plan executions, which is a necessary step for the task of model recognition.
+The empirical performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC).
 \end{abstract}
@@
 \begin{keyword}
-Classical planning\sep Planning and learning\sep Learning action models\sep Generalized planning
+Classical planning\sep Learning action models\sep Generalized planning\sep Model recognition
 %% keywords here, in the form: keyword \sep keyword
@@
 Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML to the learning of \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
+The application of inductive ML to the learning of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, is not straightforward though:
 \begin{itemize}
@@
 \item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning \strips\ action models the output is, for each action, the sets of {\em preconditions}, {\em negative} and {\em positive effects} that define the possible state transitions.
+When learning \strips\ action models the output is, for each action, the sets of {\em preconditions}, {\em negative} and {\em positive effects} that define which state transitions are possible.
 \end{itemize}
@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-A solution to the classical planning task that results from our compilation is a sequence of actions that determines the learned action model, i.e.
-the preconditions and effects of the target \strips\ actions.
+A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model can satisfy all the observations of plan executions that are given as input.
 
@@
 Learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
+Further, the compilation is robust to partially obervability of the intermediate and final states.
 Last but not least, the compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known.
@@
 The compilation can assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions.
-Our compilation is extensible to accept a learned model as input besides the observations of plan executions.
-This extension allows us to transform the input model into a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
+This allows to assess learning performance without having the actual model and, like in the leaning task, the evaluation of strips model with our compilation is flexible to various amount and kind of available input knowledge.
+Further, our compilation is extensible to accept a learned model as input besides the observations of plan executions to transform the input model into a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
+\item {\em Model Recognition}.
+Our mechanism for model evaluation allows us to define the model recognition task.
+This task is relevant since different generative models like policies, programs, grammars or different forms of domain-specific control knowledge are STRIPS compilable.
+In this sense, our work poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations.
 \end{enumerate}
@@
 \begin{itemize}
-\item Unifies the formulation of our approach for learning \strips\ action models from observed plans but also from state observations.
-\item Evaluates our compilation in the particular scenario where the observed states are not full states but {\em partial states}.
-This mean that some of the observed fluents (either with {\em positive} or {\em negative} value) are missing because they cannot be observed.
-\item Proposes a distance metric to assess how well a \strips\ action model matches a given set of plans.
-Extends the compilation to compute this distance as well as the edit distance between two given \strips\ action models.
-In addition defines the task of {\em model recognition} as the task of selecting from a given set, the action model that best matches a {\em test set} with observations of plan executions.
-\item Shows that learning \strips\ action models can also leverage {\em background knowledge} in this case, given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.
+\item A unified formulation for learning and evaluating \strips\ action models from observed plans but also from state observations.
+\item The formulation of {\em model recognition} as the task of selecting, from a given set, the action model that best matches a {\em test set} with observations of plan executions.
+\item Leveraging {\em background knowledge} (given as planning constraints either in the form of {\em state constraints} or {\em trajectory constraints}) for learning/evaluating/recognizing \strips\ action models.
+\item A more complete empirical evaluation of the compilation approach.
+Our evaluation analyses how the amount of input knowledge and how {\em partial state observability} (some of the fluents either with {\em positive} or {\em negative} value of the intermediate states are missing because they cannot be observed) affect to the performance of the compilation approach.
 \end{itemize}
@@
 Section~\ref{sec:Section2} reviews related work on learning planning action models.
-Section~\ref{sec:Section3} introduces the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the output of the addressed learning task).
+Section~\ref{sec:Section3} introduces the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the target of our learning/evaluation/recognition tasks).
 Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.
-Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks and how the compilation is also extensible to assess learned action models.
+Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks its extension to evaluate and recognize \strips\ action models.
 Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First, the learned \strips\ action models are tested with observations of plan executions and second, the learned models are compared to the actual models.
@@
 \label{sec:Section3}
-This section defines the planning model used on this work, Classical planning with conditional effects, and the output of the learning tasks addressed in the paper, a \strips\ action model.
+This section defines the planning model used on this work, {\em classical planning with conditional effects}, and the target of the learning/evaluation/recognition tasks addressed in the paper, a \strips\ action model.
 
@@
  :precondition (and (holding ?v1) (clear ?v2))
- :effect (and (not (holding ?v1)) (not (clear ?v2)) (handempty) (clear ?v1) (on ?v1 ?v2))
-)
+ :effect (and (not (holding ?v1)) (not (clear ?v2)) (handempty) (clear ?v1) (on ?v1 ?v2)))
 \end{verbatim}
@@
 \end{itemize}
-Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, then $2^{3|F_v(\xi)|}$ is an upper-bound on the number of possible \strips\ action models for $\xi$.
-In practice this upper-bound is lower, $2^{2|F_v(\xi)|}$, given that the previous \strips\ constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
-
-Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same headers and are built from the same set of possible elements.
+Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, $2^{2|F_v(\xi)|}$ defines the size of space of the possible \strips\ models for that operator, given that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
+For instance, this number is 4194304 for the blocksworld {\tt stack} operator while is only 1024 for the {\tt pickup} operator.
+
+Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same headers so they can be built from the same set of possible elements.
 Formally, iff $head(\xi)=head(\xi')$ and it holds that $F_v(\xi)=F_v(\xi')$.
-Likewise we can say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff, every $\xi\in\mathcal{M}$ is comparable with an action schema $\xi'\in\mathcal{M'}$ and viceversa, every $\xi'\in\mathcal{M'}$ is comparable with an action schema $\xi\in\mathcal{M}$.
+For instance we can claim that the {\tt stack} and {\tt unstack} blocksworld operators are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
+Likewise we say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there is a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.
 

### 2018-04-12 doc/aij18/aij18.tex
updates added to section 4 aij18 paper

@@
 Learning \strips\ action models from fully available input knowledge, i.e.
-from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
-When any intermediate state is available, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions~\cite{jimenez2012review}.
-This section formalizes a set of more challenging action model learning tasks, where less input knowledge is available.
+from plans where the {\em pre-} and {\em post-states} of every action are available, is straightforward~\cite{jimenez2012review}:
+\begin{itemize}
+  \item {\em Preconditions} are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions. 
+  \item {\em Effects} are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+\end{itemize}
+This section formalizes more challenging learning tasks, where less input knowledge is available.
 
 \subsection{Learning from observations of plan executions}
-The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its plan executions, the actual executed actions are unobserved.
-This learning task is formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$:
-\begin{itemize}
-\item $\mathcal{M}$ is the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$. 
-\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is a sequence of {\em state observations} obtained observing the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The states in $\mathcal{O}$ are not required to be full states, they can be {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing (it is unknown whether their value is either positive or negative).
-Anyway we assume that observations are noiseless, meaning that if the value of a fluent is observed it is correct.
+The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its actions, the actual executed actions are unobserved.
+This learning task is formalized as a tuple $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$:
+\begin{itemize}
+\item $\mathcal{M}$, the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$. 
+\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is the sequence of {\em state observations} obtained watching the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+We assume that state observations are noiseless, meaning that if the value of a fluent is observed it is correct.
+However, states in $\mathcal{O}$ are not assumed to be a fully observed states.
+States in $\mathcal{O}$ can be {\em partial states}, in the sense that some of the observed fluents are missing, because it is unknown whether their value is either positive or negative.
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame.
-This set of predicates can be given as input but it can also be inferred from the state observations given in $\mathcal{O}$ provided that they are full states.
-\end{itemize}
-
-A solution to the $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ compliant with the input model $\mathcal{M}$, the state observation sequence $\mathcal{O}$ and the predicates $\Psi$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the unobserved plan $\pi$, that explains the given state observations using the learned \strips\ model.
-Figure~\ref{fig:example-observations} shows a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
-In this case we are assuming full state observability so, if a fluent is not present in the observations, it means that it is set to {\em false}.
+This could be inferred from the state observations provided that at least a state in $\mathcal{O}$ is a full state.
+\end{itemize}
+
+A solution to a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ compliant with the input model $\mathcal{M}$, the state observation sequence $\mathcal{O}$ and the predicates $\Psi$.
+In this learning scenario, a solution must not only determine a possible \strips\ action model but also the unobserved plan $\pi$, that explains the given observations with the learned \strips\ model.
+Figure~\ref{fig:example-observations} shows an example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
 
@@
 \end{footnotesize}
- \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
+ \caption{\small Task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
 \label{fig:example-observations}
@@
 
-Here we redefine the learning task to cover the scenario where the actions executed by the observed agent are known.
-In this case $\mathcal{O}$ should contain partial states because otherwise, the learning task becomes trivial as explained.
+Here we redefine the learning task to cover the scenario where some of the actions executed by the observed agent are also available.
+In the case all actions in the plan are known, $\mathcal{O}$ should contain partial states because otherwise, the learning task becomes trivial as explained.
 The learning task is now formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
@@
 \item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence that produces the sequence of state observations given in $\mathcal{O}$.
+Again we assume that action observations are noiseless, meaning that if the value of an action is observed it is correct.
+However, some actions maybe unoserved.
 When the plan is {\em diverse} enough, i.e.
@@
 
-Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of an four-action plan for inverting a two-block tower.
-Note that $\mathcal{M},\Psi$ are skipped since they are the same as for Figure~\ref{fig:example-observations}.
-Again, we are assuming full state observability and the plan is full observable as well.
-However, only the first and last states were observed.
+Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of a four-action plan for inverting a two-block tower.
+The exaple skips $\mathcal{M},\Psi$ since they are the same as in Figure~\ref{fig:example-observations}.
+In this example only the first and last states were observed and intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
 
@@
 
- \caption{\small Example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ task for learning a \strips\ action model in the blocksworld from a four-action plan and two state observations.}
+ \caption{\small Task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for learning a {\em blocksworld} \strips\ action model from a four-action plan and two state observations.}
 \label{fig:example-plans}
@@
 
+In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes (preconditions and some positive or negative effects are a priori known).
+Such scenario is relevant for {\em Policy learning}.
+A policy is function that maps states into actions and represent the conditions under which actions should be applied to achieve certain goals.
+Given an action model $\mathcal{M}$ the task of learning a policy that is compliant with a given set of observations of plan executions can be defined as the task of learning extra preconditions for each of the action schemes $\xi\in\mathcal{M}$.
+These extra preconditions capture when actions can be applied according to the policy.
+If for each reachable state the policy defines only a single applicable action the policy is a {\em full policy} otherwise, we say it is a {\em partial policy}.
+In the general case, learning a {\em full policy} for an arbitrary planning task is complex because it may require computing {\em high-level state features}~\cite{lotinac2016automatic}.
+
+%\begin{itemize}
+%\item {\em Goal learning}.
+The goal condition is the set of constraints that defines the subset of states that are considered goal states.
+Given an action model $\mathcal{M}$ the task of learning the goal conditions for a a given set of observations of plan executions can be defined as learning the preconditions of a extra action that is always applied in the last place and whose only effect is marking that the goals of the planning task are reached.
+%\end{itemize}
+
 The previous definitions formalize the learning of \strips\ action models from the observation of a single plan execution.
@@
 \begin{itemize}
-\item When learning from states observations, $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ is still a valid formalization by simply considering that now $\mathcal{O}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ is a sequences of {\em state observations} that are obtained observing the execution of multiple {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, one after the other.
+\item When learning from states observations, $\Lambda=\tup{\mathcal{M},\mathcal{O'},\Psi}$ is still a valid formalization by simply considering that now $\mathcal{O'}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ is a sequence of {\em state observations} obtained watching the execution of a serie of {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, one after the other.
 This means that, for each {\small $1\leq i\leq n^t$}, $a_i^t\in \pi^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-  \item When learning from a set of observed plans, the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given sequence of example plans producing the corresponding sequence of state observations $\mathcal{O}$.
-\end{itemize}
-
-\subsection{Learning from a partially specified model}
-In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes.
-This means that some preconditions and some positive or negative effects are a priori known.
-
-Such scenario is specially relevant to addressing two well-known learning tasks:
-\begin{itemize}
-\item {\em Policy learning}.
-A policy is function that maps states into actions and represent the conditions under which an action should be applied to achieve certain goals.
-Given an action model $\mathcal{M}$ the task of learning a policy that is compliant with a given set of observations of plan executions can be defined as the task of learning extra preconditions for each of the action schemes $\xi\in\mathcal{M}$.
-These extra preconditions capture when actions can be applied according to the policy.
-If for each possible state there is only one applicable action we say that the policy is fully specified otherwise, we say that the policy is partially specified.
-In the general case, policy learning for arbitrary planning tasks is a complex task because it may require computing new {\em high-level state features}~\cite{lotinac2016automatic}.
-
-\item {\em Goal learning}.
-The goal condition is the set of constraints that defines the subset of states that are considered goal states.
-Given an action model $\mathcal{M}$ the task of learning the goal conditions for a a given set of observations of plan executions can be defined as learning the preconditions of a extra action that is always applied in the last place and whose only effect is marking that the goals of the planning task are reached.
+  \item When learning from a set of observed plans, the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O'},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given sequence of example plans producing the corresponding sequence of state observations $\mathcal{O'}$.
 \end{itemize}

### 2018-04-13 doc/aij18/aij18.tex
updates added to sections 4,5 aij18 paper

@@
 Further, our compilation is extensible to accept a learned model as input besides the observations of plan executions to transform the input model into a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
+
 \item {\em Model Recognition}.
@@
 In this sense, our work poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations.
+This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since we do not require a full plan or a full action model.
 \end{enumerate}
@@
 \item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is the sequence of {\em state observations} obtained watching the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-We assume that state observations are noiseless, meaning that if the value of a fluent is observed it is correct.
-However, states in $\mathcal{O}$ are not assumed to be a fully observed states.
-States in $\mathcal{O}$ can be {\em partial states}, in the sense that some of the observed fluents are missing, because it is unknown whether their value is either positive or negative.
+We assume that the initial state $s_0$ is {\em fully observable} while the rest of states in $\mathcal{O}$ can be {\em partially observable}, meaning that some of the observed fluents are missing because it is unknown whether their value is either positive or negative.
+In the extreme, full state observations could be missing but we assume that state observations are {\em noiseless}, meaning that if the value of a fluent is observed it is correct.
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame.
-This could be inferred from the state observations provided that at least a state in $\mathcal{O}$ is a full state.
-\end{itemize}
-
-A solution to a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ compliant with the input model $\mathcal{M}$, the state observation sequence $\mathcal{O}$ and the predicates $\Psi$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the unobserved plan $\pi$, that explains the given observations with the learned \strips\ model.
+Note that $\Psi$ can be inferred from the state observations provided that at least a state $s\in \mathcal{O}$ is a full state, that is $|s|=|F|$.
+\end{itemize}
+
+A solution to the $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the given state observations $\mathcal{O}$ and the predicates $\Psi$.
+Solving this learning task implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model.
 Figure~\ref{fig:example-observations} shows an example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
+Learning the action model from this example implies inferring that the unobserved plan is $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$.
 
@@
 
-Here we redefine the learning task to cover the scenario where some of the actions executed by the observed agent are also available.
-In the case all actions in the plan are known, $\mathcal{O}$ should contain partial states because otherwise, the learning task becomes trivial as explained.
+In some cases we may not require to start learning from scratch because some preconditions and/or some positive or negative effects may be a priori known.
+The operator schemas in $\mathcal{M}$, that are given as input, may be not {\em empty} but {\em partially specified}.
+Such scenario is relevant for {\em Policy learning}.
+A policy is function that maps states into actions and represents the conditions under which actions should be applied to achieve certain goals.
+Given an action model $\mathcal{M}$ the task of learning a policy that is compliant with a given set of observations of plan executions can be defined as the task of learning extra preconditions for each action scheme $\xi\in\mathcal{M}$.
+These extra preconditions capture when actions can be applied according to the policy.
+A policy that defines a single applicable action for each reachable state is a {\em full policy} otherwise, we say it is a {\em partial policy}.
+In the general case, learning a {\em full policy} for an arbitrary planning task is complex because it may require computing {\em high-level state features}~\cite{lotinac2016automatic}.
+
+Here we redefine the learning task to cover the scenario where some of the actions executed by the observed agent are available.
+If all executed actions are known, $\mathcal{O}$ should contain partial states because otherwise, the learning task becomes trivial as explained.
 The learning task is now formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
 \begin{itemize}
-\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is an action sequence that produces the sequence of state observations given in $\mathcal{O}$.
+\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is the action sequence that produces the sequence of state observations given in $\mathcal{O}$.
 Again we assume that action observations are noiseless, meaning that if the value of an action is observed it is correct.
-However, some actions maybe unoserved.
-When the plan is {\em diverse} enough, i.e.
-it contains at least one ground action for each of the aimed action schemes, the set of {\em empty} operator schemas $\mathcal{M}$ can be inferred from the input plan $\pi$.
+When the input plan is {\em diverse} enough, i.e.
+$\pi$ contains at least one ground action for each of the aimed action schemes, the set of {\em empty} operator schemas $\mathcal{M}$ can be inferred from $\pi$.
 \end{itemize}
@@
 The exaple skips $\mathcal{M},\Psi$ since they are the same as in Figure~\ref{fig:example-observations}.
-In this example only the first and last states were observed and intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
+In this example $\mathcal{O}=\tup{s_0,s_4}$ which means that only the first and last states were observed and the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
 
@@
 
-In some cases, we may not require to start learning from scratch that is, the operator schemas in $\mathcal{M}$ may be not {\em empty} but {\em partially specified} operator schemes (preconditions and some positive or negative effects are a priori known).
-Such scenario is relevant for {\em Policy learning}.
-A policy is function that maps states into actions and represent the conditions under which actions should be applied to achieve certain goals.
-Given an action model $\mathcal{M}$ the task of learning a policy that is compliant with a given set of observations of plan executions can be defined as the task of learning extra preconditions for each of the action schemes $\xi\in\mathcal{M}$.
-These extra preconditions capture when actions can be applied according to the policy.
-If for each reachable state the policy defines only a single applicable action the policy is a {\em full policy} otherwise, we say it is a {\em partial policy}.
-In the general case, learning a {\em full policy} for an arbitrary planning task is complex because it may require computing {\em high-level state features}~\cite{lotinac2016automatic}.
-
-%\begin{itemize}
-%\item {\em Goal learning}.
-The goal condition is the set of constraints that defines the subset of states that are considered goal states.
-Given an action model $\mathcal{M}$ the task of learning the goal conditions for a a given set of observations of plan executions can be defined as learning the preconditions of a extra action that is always applied in the last place and whose only effect is marking that the goals of the planning task are reached.
-%\end{itemize}
-
 The previous definitions formalize the learning of \strips\ action models from the observation of a single plan execution.
@@
 Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
-A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plans that explain the given inputs to the learning task.
+A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plans that explains the given inputs to the learning task.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \item {\bf Validates the action model $\mathcal{M}'$}.
-The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) but using the programmed action model $\mathcal{M}'$.
+The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) with the programmed action model $\mathcal{M}'$.
 \end{enumerate}
@@
 \item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the input observations, i.e.
-{\tt\small block A} and {\tt\small block B} in Figure~\ref{fig:example-observations}.
+the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-observations}.
 Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
-For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to {\em True}.
+For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
 \item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation in $\mathcal{O}$ where the action model is validated.
 \end{itemize}
-\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0$ (the first observation) and $mode_{prog}$ set to true.
+\item $I_{\Lambda}$ encodes the first observation, $s_0\subseteq F$, and sets $mode_{prog}$ to true.
 Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
@@
 \end{scriptsize}
- \caption{\small PDDL action for applying an already programmed schema $stack$ (implications coded as disjunctions).}
+ \caption{\small PDDL action for applying an already programmed schema $stack$ (implications are coded as disjunctions).}
 \label{fig:compilation}
@@
 
-\item Actions for {\em validating} an observation {\tt\small $1\leq i\leq |\mathcal{O}|$}.
+\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{O}$, {\tt\small $1\leq i< |\mathcal{O}|$}.
 \begin{small}
@@
 
-As introduced, the compilation approach is flexible to different amount of input data.
-Known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
+The compilation approach is flexible to various amount and kind of input data.
+If any reference to the $mode_{val}$ fluent is removed the compilation can learn \strips\ action models when there are missing state observations in $\mathcal{O}$.
+On the other hand, if the $mode_{val}$ fluent is ignored, $P_{\Lambda}$ becomes harder since the classical planner must determine how many {\em apply} actions are necessary between any two observations, i.e.
+between the application of two {\em validate} actions.
+Further, known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
 In this case, the corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
-When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the compilation validates whether the given observations of the plan execution follows the given model.
-This feature is beyond the functionality of VAL~\cite{howey2004val} because the plan validation tool requires (1) a full action model and (2), a plan:
-\begin{itemize}
-\item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given examples if a solution plan is found for $P_{\Lambda}$.
-\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given examples if $P_{\Lambda}$ is unsolvable since it means that $\mathcal{M}$ cannot be compliant with all the given observations.
-\end{itemize}
-
-Figure~\ref{fig:plan-observations} illustrate how our compilation works and shows a plan that solves a classical planning task resulting from the compilation.
-The plan programs and validates the $stack$ schema (from {\em blocksworld}) using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.
+When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the compilation validates whether the observation of the plan execution follows the given model:
+\begin{itemize}
+\item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given input data if a solution plan for $P_{\Lambda}$ can be found.
+\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given input data if $P_{\Lambda}$ is unsolvable.
+This means that $\mathcal{M}$ cannot be compliant with the given observation of the plan execution.
+\end{itemize}
+This feature of our compilation is beyond the functionality of VAL, the plan validation tool~\cite{howey2004val}, because VAL requires (1) a full plan and (2), a full action model for plan validation.
+
+Figure~\ref{fig:plan-observations} illustrate how our compilation works and shows a plan that solves the classical planning task that results from our compilation.
+This plan programs and validates the $stack$ schema (from {\em blocksworld}) using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.
 Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,21]$ validate the programmed operators using the sequence of five state observations shown in the Figure~\ref{fig:example-observations}.
@@
 
-Note that this compilation is valid for the particular scenario where the observed states are not full states but {\em partial states}, in the sense that some of the observed fluents (either with positive or negative value) are missing.
-If any reference to the $mode_{val}$ fluent is removed, the compilation becomes even more flexible and can learn \strips\ action models even if there are missing state observations in $\mathcal{O}$.
-On the other hand, if the $mode_{val}$ fluent is ignored, the resulting classical planning task $P_{\Lambda}$ becomes harder to be solved since the classical planner must now determine how many {\em apply} actions are necessary between any two observations, i.e.
-between the application of two {\em validate} actions.
-
 \subsection{Learning from plans}
@@
 \begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with the $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that contains the action $a_j$.
-The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to represent the current plan step and iterate through the steps of the plan.
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with the $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that corresponds to action $a_j$.
+The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to iterate through the steps of plan $\pi$.
 \item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<|\pi|$}, for tracking the plan step where the action model is validated.
@@
 \item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ and the actions for {\em validating} the programmed action model in a given state observation are the same as in the previous compilation.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\pi}$ that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism ensures that these actions are applied, exclusively, in the same order as in the example plan $\pi$.
+\item The actions for {\em applying} an already programmed operator includes the following extra conditional effects $\{at_{j},plan(name(a_j),\Omega^{ar(a_j)},j)\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+This mechanism ensures that $\mathsf{apply_{\xi,\omega}}$ actions are applied, exclusively, in the same order as in the example plan $\pi$ while it is flexible to the learning scenario where some $a_j$ actions in $\pi$ are missing.
+In such scenario $P_{\Lambda'}$ becomes harder because the planner must determine which and how many $\mathsf{apply_{\xi,\omega}}$ actions are necessary for validating the learned action model with the given input knowledge.
 \end{enumerate}
@@
 
-To illustrate this, the classical plan of Figure~\ref{fig:plan-lplan} is a solution to a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $\mathcal{O}$ shown in Figure~\ref{fig:example-plans}.
+The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $\mathcal{O}=\tup{s_0,s_4}$ shown in Figure~\ref{fig:example-plans}.
 Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the four-action plan $\pi$ shown in the Figure~\ref{fig:example-plans}.
@@
 \end{figure}
-
-The compilation is flexible to the learning scenario where some $a_j$ actions in $\pi$ are missing.
-In such scenario the corresponding $plan(name(a_j),\Omega^{ar(a_j)},j)$ fluents are not coded in the initial state as well as they are not added as extra preconditions of the corresponding corresponding {\em apply} actions.
-Again, the resulting classical planning task that results from this modification of the compilation becomes harder to be solved since the planner must determine which are the {\em apply} action necessary for validating the learned action model with the given input knowledge.
 

### 2018-04-16 doc/aij18/aij18.tex
updates added to section4 AIJ18 paper

@@
 Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same headers so they can be built from the same set of possible elements.
-Formally, iff $head(\xi)=head(\xi')$ and it holds that $F_v(\xi)=F_v(\xi')$.
+Formally, iff $head(\xi)=head(\xi')$ so it also holds that $F_v(\xi)=F_v(\xi')$.
 For instance we can claim that the {\tt stack} and {\tt unstack} blocksworld operators are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
-Likewise we say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there is a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.
+Likewise we say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.
 
@@
 \begin{itemize}
-  \item {\em Preconditions} are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions. 
+  \item {\em Preconditions} are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action, that is any action that belongs to the same operator scheme.
   \item {\em Effects} are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
 \end{itemize}
-This section formalizes more challenging learning tasks, where less input knowledge is available.
+This section formalizes more challenging learning tasks where less input knowledge is available for instance, because it cannot be observed.
 
@@
 \item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is the sequence of {\em state observations} obtained watching the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-We assume that the initial state $s_0$ is {\em fully observable} while the rest of states in $\mathcal{O}$ can be {\em partially observable}, meaning that some of the observed fluents are missing because it is unknown whether their value is either positive or negative.
-In the extreme, full state observations could be missing but we assume that state observations are {\em noiseless}, meaning that if the value of a fluent is observed it is correct.
+We assume that the initial state $s_0$ is {\em fully observable} while states $s_i$ s.t.
+{\small $1\leq i\leq |\mathcal{O}|$} can be {\em partially observable}, meaning that some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative.
+In the extreme, states $s_i$ {\small $1\leq i\leq |\mathcal{O}|$}, can be missing but we assume that any state observation is {\em noiseless}, meaning that if the value of a fluent is observed it is correct.
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame.
@@
 
-A solution to the $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the given state observations $\mathcal{O}$ and the predicates $\Psi$.
-Solving this learning task implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model.
-Figure~\ref{fig:example-observations} shows an example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence that corresponds to inverting a 2-block tower.
-Learning the action model from this example implies inferring that the unobserved plan is $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$.
+A solution to a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the given state observations $\mathcal{O}$ and the predicates $\Psi$.
+Solving $\Lambda$ implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model.
+Figure~\ref{fig:example-observations} shows an example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence obtained inverting a 2-block tower.
+Learning the action model from this example implies inferring the unobserved plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$.
 
@@
 The operator schemas in $\mathcal{M}$, that are given as input, may be not {\em empty} but {\em partially specified}.
-Such scenario is relevant for {\em Policy learning}.
+Such scenario is relevant for {\em policy learning}.
 A policy is function that maps states into actions and represents the conditions under which actions should be applied to achieve certain goals.
-Given an action model $\mathcal{M}$ the task of learning a policy that is compliant with a given set of observations of plan executions can be defined as the task of learning extra preconditions for each action scheme $\xi\in\mathcal{M}$.
+Given an action model, the task of learning a policy that is compliant with a set of observations can be defined as learning extra preconditions for each action scheme in the model.
 These extra preconditions capture when actions can be applied according to the policy.
-A policy that defines a single applicable action for each reachable state is a {\em full policy} otherwise, we say it is a {\em partial policy}.
-In the general case, learning a {\em full policy} for an arbitrary planning task is complex because it may require computing {\em high-level state features}~\cite{lotinac2016automatic}.
-
-Here we redefine the learning task to cover the scenario where some of the actions executed by the observed agent are available.
-If all executed actions are known, $\mathcal{O}$ should contain partial states because otherwise, the learning task becomes trivial as explained.
+A policy that exactly defines a single applicable action for each reachable state is a {\em full policy} otherwise, we say it is a {\em partial policy}.
+In the general case, learning a {\em full policy} for an arbitrary planning task is complex because a given action scheme may be applicable in different situations and also because the compact representation of this set of different situations may require the computation of {\em high-level state features}~\cite{lotinac2016automatic}.
+
+The $\Lambda$ learning task can also be redefined to cover the scenario where some of the actions executed by the observed agent are available.
+If all the executed actions are known then states observations should be partial otherwise, the learning task is trivial.
 The learning task is now formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
@@
 Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of a four-action plan for inverting a two-block tower.
-The exaple skips $\mathcal{M},\Psi$ since they are the same as in Figure~\ref{fig:example-observations}.
-In this example $\mathcal{O}=\tup{s_0,s_4}$ which means that only the first and last states were observed and the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
+In this example $\mathcal{O}=\tup{s_0,s_4}$ which means that only the first and last states are observed and the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
+$\mathcal{M},\Psi$ are skipped, since they are the same as in Figure~\ref{fig:example-observations}.
 
@@
 The previous definitions formalize the learning of \strips\ action models from the observation of a single plan execution.
-These definitions are extensible to the more general case where the available input data is obtained from the execution of multiple plans:
-\begin{itemize}
-\item When learning from states observations, $\Lambda=\tup{\mathcal{M},\mathcal{O'},\Psi}$ is still a valid formalization by simply considering that now $\mathcal{O'}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ is a sequence of {\em state observations} obtained watching the execution of a serie of {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, one after the other.
+These definitions are extensible to the more general case where the execution of multiple plans is observed.
+The task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O'},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given sequence of example plans producing the corresponding sequence of state observations $\mathcal{O'}$.
+Now $\mathcal{O'}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ is a sequence of {\em state observations} obtained watching the execution of a serie of {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, one after the other.
 This means that, for each {\small $1\leq i\leq n^t$}, $a_i^t\in \pi^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-  \item When learning from a set of observed plans, the task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O'},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given sequence of example plans producing the corresponding sequence of state observations $\mathcal{O'}$.
-\end{itemize}
 

### 2018-04-16 doc/aij18/aij18.tex
separate files sections 1, 2 updates 2

@@
 
-\section{Introduction}
-\label{sec:Section1}
-Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
-Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
-
-Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML to the learning of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, is not straightforward though:
-\begin{itemize}
-\item The {\em input} to ML algorithms (the {\em learning/training} data) usually are finite vectors encoding the value of fixed features in a given set of objects.
-The input for learning planning action models are observations of plan executions (where each plan possibly has a different length and may involve a different number of objects).
-\item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning \strips\ action models the output is, for each action, the sets of {\em preconditions}, {\em negative} and {\em positive effects} that define which state transitions are possible.
-\end{itemize}
-
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model can satisfy all the observations of plan executions that are given as input.
-
-The compilation approach is appealing by itself because it leverages off-the-shelf planners for learning and because its practicality allow us to report model learning results over a wide range of IPC planning domains.
-Moreover, it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
-Apart from these, the compilation approach presents the following contributions:
-\begin{enumerate}
-\item {\em Flexibility}.
-The compilation is flexible to various amount and kind of available input knowledge.
-Learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
-Further, the compilation is robust to partially obervability of the intermediate and final states.
-Last but not least, the compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known.
-% Also accepts background knowledge
-
-\item {\em Model Evaluation}.
-The compilation can assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions.
-This allows to assess learning performance without having the actual model and, like in the leaning task, the evaluation of strips model with our compilation is flexible to various amount and kind of available input knowledge.
-Further, our compilation is extensible to accept a learned model as input besides the observations of plan executions to transform the input model into a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
-
-\item {\em Model Recognition}.
-Our mechanism for model evaluation allows us to define the model recognition task.
-This task is relevant since different generative models like policies, programs, grammars or different forms of domain-specific control knowledge are STRIPS compilable.
-In this sense, our work poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations.
-This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since we do not require a full plan or a full action model.
-\end{enumerate}
-
-A first description of the compilation previously appeared in the conference paper~\cite{aineto2018learning}. Compared to the conference paper, this work includes the following novel material:
-\begin{itemize}
-\item A unified formulation for learning and evaluating \strips\ action models from observed plans but also from state observations.
-\item The formulation of {\em model recognition} as the task of selecting, from a given set, the action model that best matches a {\em test set} with observations of plan executions.
-\item Leveraging {\em background knowledge} (given as planning constraints either in the form of {\em state constraints} or {\em trajectory constraints}) for learning/evaluating/recognizing \strips\ action models.
-\item A more complete empirical evaluation of the compilation approach.
-Our evaluation analyses how the amount of input knowledge and how {\em partial state observability} (some of the fluents either with {\em positive} or {\em negative} value of the intermediate states are missing because they cannot be observed) affect to the performance of the compilation approach.
-\end{itemize}
-
-Section~\ref{sec:Section2} reviews related work on learning planning action models.
-Section~\ref{sec:Section3} introduces the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the target of our learning/evaluation/recognition tasks).
-Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.
-Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks its extension to evaluate and recognize \strips\ action models.
-Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First, the learned \strips\ action models are tested with observations of plan executions and second, the learned models are compared to the actual models.
-Finally, Section ~\ref{sec:Section8} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.
-
-\section{Related work}
-\label{sec:Section2}
-Back in the 90's various systems aimed learning operators mostly via interaction with the environment.
-{\sc LIVE} captured and formulated observable features of objects and used them to acquire and refine operators \cite{ShenS89}.
-{\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
-These early works were based on lifting the observed states supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge.
-
-More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
-
-Whilst the aforementioned works deal with full state observability,action model learning has also been studied in domains where there is partial or missing state observability.
-{\sf ARMS} works when no partial intermediate state is given.
-It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
-In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
-%and so it may output an inconsistent model
-{\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
-Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
-The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
-This update makes sure that the new formula represents all the transition relations consistent with the actions and observations.
-The formula returned at the end includes all consistent models, which can then be retrieved with additional processing.
-
-Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
-An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
-
-{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
-This makes {\sf LOCM} be most likely the learning approach that works with the least information possible.
-The lack of available information is addressed by LOCM by exploiting assumptions about the kind of domain model it has to generate.
-Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
-The intuitive assumptions of {\sf LOCM}
-, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence,
-yield a learning model heavily reliant on the kind of domain structure.
-The inability of {\sf LOCM} to properly derive domain theories where the state of a sort is subject to different FSMs is later overcome by {\sf LOCM2} by forming separate FSMs, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
-{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain}), the last contribution of the {\sf LOCM} family, addresses the problem of inducing static predicates.
-Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
-{\em blocksworld} and {\em freecell}).
-In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
-
-Recenctly classical planning compilations have been defined to learn different kinds of generative models from examples.
-The existing compilations for computing FSCs for generalized planning follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it and hence, they tightly integrate planning and generalization.
-To keep the computation of FSCs tractable, they limit the space of possible solutions bounding the maximum size of the FSC.
-In addition, they impose that the instances to solve share, not only the domain theory (actions and predicates schemes) but the set of fluents~\cite{javi-Gplanning-ICAPS16} or a subset of {\it observable} fluents~\cite{Geffner:FSM:AAAI10}.
-Programs increase the readability of FSCs separating the control-flow structures from the primitive actions.
-Like FSCs, programs can also be computed following a {\it top-down} approach, e.g.~exploiting compilations that program and validate the program on instances with the same state and action space~\cite{javi-Gplanning-ICAPS16}.
-Since these {\it top-down} approaches search in the space of solutions, it is helpful to limit the set of different control-flow instructions.
-For instance using only {\it conditional gotos} that can both implement branching and loops~\cite{Jimenez15}.
+\input{introduction}
+\input{related_work}
 
@@
 \bibliographystyle{elsarticle-num}
-\bibliography{planlearnbibliography.bib}
+\bibliography{planlearnbibliography}
 

### 2018-04-16 doc/aij18/introduction.tex
separate files sections 1, 2 updates 2

\section{Introduction}
\label{sec:Section1}

Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}. At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}. Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.

Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}. The application of inductive ML to the learning of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, is not straightforward though:
\begin{itemize}
\item The {\em input} to ML algorithms (the {\em learning/training} data) usually are finite vectors encoding the value of fixed features in a given set of objects. The input for learning planning action models are observations of plan executions (where each plan possibly has a different length and may involve a different number of objects).
\item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks). When learning \strips\ action models the output is, for each action, the sets of {\em preconditions}, {\em negative} and {\em positive effects} that define which state transitions are possible.
\end{itemize}

Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation. A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model can satisfy all the observations of plan executions that are given as input.

The compilation approach is appealing by itself because it leverages off-the-shelf planners for learning and because its practicality allow us to report model learning results over a wide range of IPC planning domains. Moreover, it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model. Apart from these, the compilation approach presents the following contributions:
\begin{enumerate}
\item {\em Flexibility}. The compilation is flexible to various amount and kind of available input knowledge. Learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed. Further, the compilation is robust to partially obervability of the intermediate and final states. Last but not least, the compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known. % Also accepts background knowledge

\item {\em Model Evaluation}. The compilation can assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions. This allows to assess learning performance without having the actual model and, like in the leaning task, the evaluation of strips model with our compilation is flexible to various amount and kind of available input knowledge. Further, our compilation is extensible to accept a learned model as input besides the observations of plan executions to transform the input model into a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.

\item {\em Model Recognition}. Our mechanism for model evaluation allows us to define the model recognition task. This task is relevant since different generative models like policies, programs, grammars or different forms of domain-specific control knowledge are STRIPS compilable. In this sense, our work poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations. This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since we do not require a full plan or a full action model.
\end{enumerate}

A first description of the compilation previously appeared in the conference paper~\cite{aineto2018learning}. Compared to the conference paper, this work includes the following novel material:
\begin{itemize}
\item A unified formulation for learning and evaluating \strips\ action models from observed plans but also from state observations.
\item The formulation of {\em model recognition} as the task of selecting, from a given set, the action model that best matches a {\em test set} with observations of plan executions.
\item Leveraging {\em background knowledge} (given as planning constraints either in the form of {\em state constraints} or {\em trajectory constraints}) for learning/evaluating/recognizing \strips\ action models.
\item A more complete empirical evaluation of the compilation approach. Our evaluation analyses how the amount of input knowledge and how {\em partial state observability} (some of the fluents either with {\em positive} or {\em negative} value of the intermediate states are missing because they cannot be observed) affect to the performance of the compilation approach.
\end{itemize}

Section~\ref{sec:Section2} reviews related work on learning planning action models. Section~\ref{sec:Section3} introduces the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the target of our learning/evaluation/recognition tasks). Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge. Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks its extension to evaluate and recognize \strips\ action models. Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First, the learned \strips\ action models are tested with observations of plan executions and second, the learned models are compared to the actual models. Finally, Section ~\ref{sec:Section8} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.

### 2018-04-16 doc/aij18/related_work.tex
separate files sections 1, 2 updates 2

--- 2017-11-16 doc/icaps-18/strips-learning.tex
@@
-\def\year{2017}\relax
-%File: formatting-instruction.tex
-\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
-\usepackage{aaai18}  %Required
-\usepackage{times}  %Required
-\usepackage{helvet}  %Required
-\usepackage{courier}  %Required
-\usepackage{url}  %Required
-\usepackage{graphicx}  %Required
-\frenchspacing  %Required
-\setlength{\pdfpagewidth}{8.5in}  %Required
-\setlength{\pdfpageheight}{11in}  %Required
-%PDF Info Is Required:
-  \pdfinfo{}
-\setcounter{secnumdepth}{0}
+\section{Related work}
+\label{sec:Section2}
 
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{multirow}
-\usepackage{tikz}
-\usepackage{comment}
+One can find many different approaches to learning action models in the literature.
+A common but distinguishing feature to all approaches is the type of input knowledge required by the each model.
+Due to the large variety of types of input knowledge and the lack of a common naming convention (e.g., plan examples, plan traces, training data, plan samples), we firstly present a glossary of terms that will ease the classification of the surveyed models accordingly to their input knowledge.
 
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{listings}
-\usepackage{multicol}
-\usepackage{arydshln}
-\usetikzlibrary{calc,backgrounds,positioning,fit}
+We will refer to a state sequence as $\Sigma = \langle s_0, s_1, \ldots, s_n \rangle$ and an action sequence as $\Omega = \langle a_1, a_2, \ldots, a_n \rangle$.
+In a full observable (FO) state sequence $\Sigma$, every state $s_i \in \Sigma$ is a complete state and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$  is given by $\Omega = \langle a_{i+1} \rangle$.
+On the other hand, $\Omega$ is a FO action sequence if there are no missing actions in $\Omega$.
 
-\newcommand{\tup}[1]{{\langle #1 \rangle}}
+A plan trace is a combination of a state and an action sequence $\tau = \langle s_0, a_1, a_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
+Additionally, $\tau$ is a plan trace with partial observable (PO) states if there is at least one missing literal in any state $s_i$.
+And $\tau$ is a plan trace with PO actions of there is at least one missing action from the action sequence.
 
-\newcommand{\pre}{\mathsf{pre}}     % precondition
-\newcommand{\del}{\mathsf{del}}     % effect
-\newcommand{\add}{\mathsf{add}}     % effect
-\newcommand{\eff}{\mathsf{eff}}     % effect
-\newcommand{\cond}{\mathsf{cond}}   % conditional effect
-\newcommand{\true}{\mathsf{true}}   % true
-\newcommand{\false}{\mathsf{false}} % false
-\newcommand{\PE}{\mathrm{PE}}     % precondition
-\newcommand{\strips}{\textsc{Strips}}     % precondition
+TABLA
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{definition}[theorem]{Definition}
-
-\begin{document}
-
-\title{Learning \strips\ action models with classical planning}
-\author{\#39}
-
-% Commented for blind submission
-%\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
-%{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-%{\small Universitat Polit\`ecnica de Val\`encia.}\\
-%{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
-%{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
-
-\maketitle
-\begin{abstract}
-This paper presents a novel approach for learning \strips\ action models from examples that compiles this inductive learning task into classical planning.
-Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states (no intermediate action or state is given).
-What is more, the compilation accepts partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
-\end{abstract}
-
-\section{Introduction}
-Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
-Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
-
-On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML to the learning of \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
-\begin{itemize}
-\item The {\em input} to ML algorithms (the learning/training data) usually are finite vectors encoding the value of fixed features in a given set of objects.
-The input for learning planning action models are observations of plan executions (where each plan possibly has a different length).
-\item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning \strips\ action models the output is, for each action, the sets of preconditions, negative and positive effects, that define the possible state transitions.
-\end{itemize}
-
-Learning \strips\ action models is a well-studied problem with sophisticated algorithms, like {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring} that do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-The compilation approach is appealing by itself because opens the door to the bootstrapping of planning action models but also because:
-\begin{enumerate}
-\item Is flexible to different amounts of available input knowledge.
-The learning examples can range from a set of plans (with their corresponding initial and final states) to just a set of initial and final states where no intermediate state  or action is observed.
-\item Accepts previous knowledge about the structure of the actions in the form of partially specified action models.
-In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
-\end{enumerate}
-
-The second section of the paper formalizes the classical planning model, its extension to {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the output of the addressed learning task).
-The third section formalizes the learning of \strips\ action models with regard to different amounts of available input knowledge.
-The fourth and fifth sections describe our approach for addressing the formalized learning tasks.
-Finally, the last sections report the data collected in a empirical evaluation, discuss the strengths and weaknesses of our approach and propose several opportunities for future research.
-
-\section{Background}
-This section defines the planning models used on this work and the output of the learning tasks addressed in the paper.
-
-\subsection{Classical planning}
-We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
-A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
-either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (WLOG we assume that $L$ does not assign conflicting values to any fluent).
-We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
-
-A {\em state} $s$ is a full assignment of values to fluents, i.e.
-$|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
-
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. Each action $a\in A$ comprises three sets of literals:
-\begin{itemize}
-\item $\pre(a)\subseteq\mathcal{L}(F)$, called {\em preconditions}, the literals that must hold for the action $a\in A$ to be applicable.
-\item $\eff^+(a)\subseteq\mathcal{L}(F)$, called {\em positive effects}, that defines the fluents set to true by the application of the action $a\in A$.
-\item $\eff^-(a)\subseteq\mathcal{L}(F)$, called {\em negative effects}, that defines the fluents set to false by the action application.
-\end{itemize}
-We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
-The result of applying $a$ in $s$ is the {\em successor state} $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
-
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces a state sequence $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-We denote with $|\pi|$ the {\em plan length}.
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of $\pi$ in $I$.
-
-\subsection{Classical planning with conditional effects}
-Our approach for learning \strips\ action models is compiling this leaning task into a classical planning task with conditional effects.
-Conditional effects allow us to compactly define actions whose effects depend on the current state.
-Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
-
-An action $a\in A$ has now a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-
-An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the resulting set of {\em triggered effects} are the effects whose conditions hold in $s$:
-\[
-triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
-\]
-
-The result of applying an action $a$ in a state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are the triggered {\em negative} and {\em positive} effects, respectively.
-
-\subsection{\strips\ action schemes and {\em variable name} objects}
-This work addresses the learning of PDDL action schemes that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-Figure~\ref{fig:stack} shows the schema, coded in PDDL, for the {\em stack} action from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{verbatim}
-(:action stack
-  :parameters (?v1 ?v2 - object)
-  :precondition (and (holding ?v1) (clear ?v2))
-  :effect (and (not (holding ?v1))
-               (not (clear ?v2))
-               (handempty) (clear ?v1)
-               (on ?v1 ?v2)))
-\end{verbatim}
-\end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from the {\em blocksworld}.}
-\label{fig:stack}
-\end{figure}
-
-To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects $\Omega\cap\Omega_v=\emptyset$, denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block blocksworld $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters each.
-
-Let us also define $F_v$, a new set of fluents $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$ and that defines the elements that can appear in an action schema.
-For instance, in the blocksworld, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-
-Finally, we assume that actions $a\in A$ are instantiated from \strips\ operator schemes $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
-\begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
-For instance, the headers for a four-operator blocksworld are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
-\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-\end{itemize}
-
-\section{Learning \strips\ action models}
-Learning \strips\ action models from fully available input knowledge, i.e.
-from plans where the {\em pre-} and {\em post-states} of every action in a plan are available, is straightforward.
-When any intermediate state is available, \strips\ operator schemes are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding actions.
-
-This section formalizes more challenging learning tasks, where less input knowledge is available:
-
-\subsubsection{Learning from (initial, final) state pairs.} This learning task corresponds to observing an agent acting in the world but watching only the results of its plan executions.
-No information about the actions in the plans is given.
-This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma}$:
-\begin{itemize}
-\item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs, that we call {\em labels}.
-Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t$ starting from the {\em initial} state $s_0^t$.
-\end{itemize}
-
-\subsubsection{Learning from labeled plans.}
-Here we augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
-
-\begin{itemize}
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where each plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that, for each {\small $1\leq i\leq n$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-\end{itemize}
-
-Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda'$ in the blocksworld.
-This learning task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
-
-\subsubsection{Learning from partially specified action models.}
-We may not require to start learning from scratch so we augment the learning input with partially specified operator schemes. This leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
-\begin{itemize}
-\item $\Xi_0$ is a partially specified action model in which some preconditions and effects are a priori known.
-\end{itemize}
-
-A solution to $\Lambda$ is a set of operator schema $\Xi$ that is compliant just with the predicates in $\Psi$, and the given set of initial and final states $\Sigma$.
-In this learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$} that explain the given labels $\Sigma$ using the learned \strips\ model.
-A solution to $\Lambda'$ is a set of \strips\ operator schema $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action with a different name in the example plans $\Pi$) compliant with the predicates in $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
-Finally a solution to $\Lambda''$ is a set of \strips\ operator schema $\Xi$ compliant as well with the provided partially specified action model $\Xi_0$.
-
-\begin{figure}
-{\tt ;;; Predicates in $\Psi$}
-\begin{footnotesize}
-\begin{verbatim}
-(handempty) (holding ?o  - object)
-(clear ?o - object) (ontable ?o - object)
-(on ?o1 - object ?o2 - object)
-\end{verbatim}
-\end{footnotesize}
-
-\vspace{0.2cm}
-
-\begin{subfigure}{.25\textwidth}
-{\tt ;;; Plan $\pi_1$}
-\begin{footnotesize}
-\begin{verbatim}
-0: (unstack A B)
-1: (putdown A)
-2: (unstack B C)
-3: (stack B A)
-4: (unstack C D)
-5: (stack C B)
-6: (pickup D)
-7: (stack D C)
-\end{verbatim}
-\end{footnotesize}
-\end{subfigure}%
-\begin{subfigure}{.6\textwidth}
-{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
-\begin{lstlisting}[mathescape]
-\end{lstlisting}
-\vspace{0.1cm}
-\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
-\node (initD) [draw]{D};
-\node (initC) [draw, above=of initD.north]{C};
-\node (initB) [draw, above=of initC.north]{B};
-\node (initA) [draw, above=of initB.north]{A};
-\draw[thick] (-1,-0.25) -- (2.5,-0.25);
-
-\node (goalA) [draw, right=10mm of initD]{A};
-\node (goalB) [draw, right=10mm of initC]{B};
-\node (goalC) [draw, right=10mm of initB]{C};
-\node (goalD) [draw, right=10mm of initA]{D};
-\end{tikzpicture}
-\vspace{0.6cm}
-\end{subfigure}%
- \caption{\small Example of a task for learning a \strips\ action model in the blocksworld from a single labeled plan.}
-\label{fig:lexample}
-\end{figure}
-
-\section{Learning \strips\ action models with planning}
-Our approach for addressing a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$, is compiling it into a classical planning task with conditional effects.
-The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
-\begin{enumerate}
-\item Programs the \strips\ action model $\Xi$.
-A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item Validates the programmed \strips\ action model $\Xi$ in the given input knowledge (the labels $\Sigma$, and $\Pi$ and/or $\Xi_0$ if available).
-For every label $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ starting from the corresponding initial state $s_0^t$ using the programmed action model $\Xi$.
-We call this process the validation of the programmed \strips\ action model $\Xi$, at the learning example {\small $1\leq t\leq \tau$}.
-%If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
-\end{enumerate}
-
-To formalize our compilation we first define {\small $1\leq t\leq \tau$} classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
-same fluents and actions but differ in the initial state and goals).
-Fluents $F$ are built instantiating the predicates in $\Psi$ with the objects appearing in the input labels $\Sigma$.
-Formally $\Omega=\{o|o\in \bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)\}$, where $obj$ is a function that returns the set of objects that appear in a fully specified state.
-The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$ while goals $G_t$, are defined by the state $s_n^t\in \sigma_t$.
-
-Now we are ready to formalize the compilations.
-We start with $\Lambda$, because it requires less input knowledge.
-Given a learning task $\Lambda=\tup{\Psi,\Sigma}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
-\begin{itemize}
-\item $F_{\Lambda}$ extends $F$ with:
-\begin{itemize}
-\item Fluents representing the programmed action model $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect in the \strips\ operator schema $\xi\in \Xi$.
-For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$}.
-\item A fluent $mode_{prog}$ indicating whether the operator schemes are being programmed or validated (already programmed) and fluents $\{test_t\}_{1\leq t\leq \tau}$, indicating the example where the action model is being validated.
-\end{itemize}
-\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), every $pre_f(\xi)\in F_{\Lambda}$ and $mode_{prog}$ set to true.
-Our compilation assumes that initially any operator schema is programmed with every possible precondition, no negative effect and no positive effect.
-\item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, indicates that the programmed action model is validated in all the learning examples.
-\item $A_{\Lambda}$ contains actions of three kinds:
-\begin{enumerate}
-\item Actions for {\em programming} an operator schema $\xi\in\Xi$:
-\begin{itemize}
-\item Actions for {\bf removing} a {\em precondition} $f\in F_v$ from the action schema $\xi\in\Xi$.
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
-& mode_{prog}, pre_{f}(\xi)\},\\
-\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
-\end{align*}
-\end{small}
-
-\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v$ to the action schema $\xi\in\Xi$.
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
-& mode_{prog}\},\\
-\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
-&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
-\end{align*}
-\end{small}
-\end{itemize}
-
-\item Actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound with the objects $\omega\subseteq\Omega^{ar(\xi)}$.
-We assume operators headers are known so the binding of the operator schema is done implicitly by order of appearance of the action parameters, i.e.
-variables $pars(\xi)$ are bound to the objects in $\omega$ appearing at the same position.
-Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
-\end{align*}
-\end{small}
-
-\item Actions for {\em validating} the learning example {\tt\small $1\leq t\leq \tau$}.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{j\in 1\leq j<t}\\
-&\cup\{\neg test_j\}_{j\in t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\}.
-\end{align*}
-\end{small}
-\end{enumerate}
-\end{itemize}
-
-\begin{lemma}
-Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
-\end{lemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-The compilation forces that once the preconditions of an operator schema $\xi \in \Xi$ are programmed, they cannot be altered.
-The same happens with the positive and negative effects that define an operator schema $\xi \in \Xi$ (besides they can only be programmed after preconditions are programmed).
-Once operator schemes are programmed they can only be applied because of the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved: executing an applicable sequence of programmed operator schemes that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
-If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence, it is a solution to $\Lambda$.
-\end{small}
-\end{proof}
-
-The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model.
-
-\begin{figure}
-\begin{scriptsize}
-\begin{verbatim}
-(:action apply_stack
-  :parameters (?o1 - object ?o2 - object)
-  :precondition
-   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
-        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
-        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
-        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
-        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
-        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
-        (or (not (pre_clear_stack_v1)) (clear ?o1))
-        (or (not (pre_clear_stack_v2)) (clear ?o2))
-        (or (not (pre_holding_stack_v1)) (holding ?o1))
-        (or (not (pre_holding_stack_v2)) (holding ?o2))
-        (or (not (pre_handempty_stack)) (handempty)))
-  :effect
-   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
-        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
-        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
-        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
-        (when (del_ontable_stack_v1) (not (ontable ?o1)))
-        (when (del_ontable_stack_v2) (not (ontable ?o2)))
-        (when (del_clear_stack_v1) (not (clear ?o1)))
-        (when (del_clear_stack_v2) (not (clear ?o2)))
-        (when (del_holding_stack_v1) (not (holding ?o1)))
-        (when (del_holding_stack_v2) (not (holding ?o2)))
-        (when (del_handempty_stack) (not (handempty)))
-        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
-        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
-        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
-        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
-        (when (add_ontable_stack_v1) (ontable ?o1))
-        (when (add_ontable_stack_v2) (ontable ?o2))
-        (when (add_clear_stack_v1) (clear ?o1))
-        (when (add_clear_stack_v2) (clear ?o2))
-        (when (add_holding_stack_v1) (holding ?o1))
-        (when (add_holding_stack_v2) (holding ?o2))
-        (when (add_handempty_stack) (handempty))
-        (when (modeProg) (not (modeProg)))))
-\end{verbatim}
-\end{scriptsize}
- \caption{\small Action for applying an already programmed schema $stack$ as encoded in PDDL (implications coded as disjunctions).}
-\label{fig:compilation}
-\end{figure}
-
-\section{Constraining the learning hypothesis space with additional input knowledge}
-\label{sec:Constraining}
-Here we show that further input knowledge can be used to constrain the space of possible action models and make the learning of \strips\ action models more practicable.
-
-\subsection{Labeled plans}
-We extend the compilation to consider labeled plans.
-Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$ that extends $P_{\Lambda}$ as follows:
-\begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-Fluents $at_j$ and $next_{j,j_2}$, {\small $1\leq j<j2\leq n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j_2}\}$, {\small $1\leq j<j2\leq n$}, for indicating the plan step where the action model is validated.
-Goals are $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$, as in the original compilation.
-\item With respect to $A_{\Lambda'}$.
-\begin{enumerate}
-\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$, that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism forces that these actions are only applied as in the example plans.
-\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ was fully executed and extra conditional effects to unload plan $\pi_{t}$ and load the next plan $\pi_{t+1}$:
-\begin{small}
-\begin{align*}
-&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\}, \{f\}\rhd\{\neg f\}_{f\in F_{\pi_t}}, \{\emptyset\}\rhd\{f\}_{f\in F_{\pi_t+1}}.
-\end{align*}
-\end{small}
-\end{enumerate}
-\end{itemize}
-
-\subsection{Partially specified action models}
-Known preconditions and effects are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda'}$.
-The corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda'}$ making the classical planning task $P_{\Lambda'}$ easier to be solved.
-
-To illustrate this, the classical plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for getting the blocksworld action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
-This plan programs and validates the operator schema {\tt\small stack} from the blocksworld, using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-Plan steps $[0,8]$ are the actions for programming the preconditions of the {\tt\small stack} operator, steps $[9,13]$ are the actions for programming the operator effects and steps $[14,22]$ are the actions for validating the programmed operators following the plan $\pi_1$ shown in the Figure~\ref{fig:lexample}.
-
-In the extreme, when a fully specified \strips\ action model $\Xi$ is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
-In this case, if a solution plan is found to $P_{\Lambda'}$, it means that the given \strips\ action model is {\em valid} for the given examples.
-If $P_{\Lambda'}$ is unsolvable it means that the given \strips\ action model is invalid since it is not compliant with all the given examples.
-Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
-
-\begin{figure}
-{\tt\small
-     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
-     01 : (program\_pre\_handempty\_stack)\\
-     02 : (program\_pre\_holding\_stack\_v2)\\
-     03 : (program\_pre\_on\_stack\_v1\_v1)\\
-     04 : (program\_pre\_on\_stack\_v1\_v2)\\
-     05 : (program\_pre\_on\_stack\_v2\_v1)\\
-     06 : (program\_pre\_on\_stack\_v2\_v2)\\
-     07 : (program\_pre\_ontable\_stack\_v1)\\
-     08 : (program\_pre\_ontable\_stack\_v2)\\
-     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
-    10 : (program\_eff\_clear\_stack\_v2)\\
-    11 : (program\_eff\_handempty\_stack)\\
-    12 : (program\_eff\_holding\_stack\_v1)\\
-    13 : (program\_eff\_on\_stack\_v1\_v2)\\
-    {\bf 14} : (apply\_unstack a b i1 i2)\\
-    15 : (apply\_putdown a i2 i3)\\
-    16 : (apply\_unstack b c i3 i4)\\
-    17 : (apply\_stack b a i4 i5)\\
-    18 : (apply\_unstack c d i5 i6)\\
-    19 : (apply\_stack c b i6 i7)\\
-    20 : (apply\_pickup d i7 i8)\\
-    21 : (apply\_stack d c i8 i9)\\
-    {\bf 22} : (validate\_1)
-}
- \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi_1$ and label $\sigma_1$ (shown in Figure~\ref{fig:lexample}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
-\label{fig:plan}
-\end{figure}
-
-\subsection{Static predicates}
-A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action schema~\cite{fox:TIM:JAIR1998}.
-Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
-Formally, given a static predicate $p$:
-\begin{itemize}
-\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable objects} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
-\item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
-\end{itemize}
-
-Static predicates can also constrain the space of possible preconditions by looking at the given set of labels $\Sigma$.
-One can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
-For instance in the {\em zenotravel} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\Sigma$.
-
-Likewise, static predicates can constrain the space of possible preconditions looking at the given example plans.
-Fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
-Back to the {\em zenotravel} domain, if a example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
-
-In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of {\em potential} static predicates from the input to the learning task.
-
-\begin{table*}
-		\resizebox{\textwidth}{!}{%                
-		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
-& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
-& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-                          \hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
-			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
-			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
-			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
-			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
-			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
-			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
-			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
-			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
-			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
-			\hline
-			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
-		\end{tabular}
-	}
-\caption{\small {\em Precision} and {\em recall} obtained learning from labeled plans without (left) and with (right) static predicates.}
-\label{tab:results_plans}
-\end{table*}
-
-\begin{table}
-\begin{scriptsize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c||c|c|c|}
-                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
-			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
-                         \hline
-			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
-			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
-			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
-			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
-			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
-			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
-                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
-                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
-			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
-			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
-			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
-			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
-		\end{tabular}
-	\end{center}
-        \end{scriptsize}
-	\caption{\small Total planning time, preprocessing time and plan length learning from labeled plans without/with static predicates.}
-	\label{tab:time_plans}	
-\end{table}
-
-\section{Evaluation}
-This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts of available input knowledge.
-
-\subsubsection{Setup.}
-The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
-All experiments are run on an Intel Core i5 3.10 GHz x 4 with 4 GB of RAM.
-
-\subsubsection{Reproducibility.}
-We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this anonymous repository {\em https://github.com/anonsub/strips-learning} so any experimental data reported in the paper is fully reproducible.
-
-\subsubsection{Planner.}
-The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
-In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
-Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
-
-\subsubsection{Metrics.}
-The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
-These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
-
-Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
-Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
-Recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
-
-\subsection{Learning from labeled plans}
-We start evaluating our approach at $\Lambda'=\tup{\Psi,\Sigma,\Pi}$ learning tasks, where {\em labeled plans} are available, and repeat the evaluation but exploiting potential static predicates that are computed from $\Sigma$.
-The potential {\em static predicates} considered here is the set of predicates s.t.
-every predicate instantiation appears unaltered in the initial and final states, for every $\sigma_t\in\Sigma$.
-Static predicates are used to constrain the space of possible action models as explained in the previous section.
-
-Table~\ref{tab:results_plans} summarizes the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns of each setting (and the last row) report averages values.
-We can observe that identifying static predicates drives to models with better precondition {\em recall}.
-This fact evidences that many of the missing preconditions corresponded to static predicates because there were no incentive to learn them as they always hold~\cite{gregory2015domain}.
-
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-All the learning tasks are solved in a few seconds time and interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some preconditions corresponding to static predicates are directly derived from the learning examples so less programming actions are required.
-When static predicates are identified, the resulting compilation is also much compact and produces smaller planning/instantiation times.
-
-\subsection{Learning from partially specified action models}
-
-We evaluate now the ability of our approach to support partially specified action models; that is, when addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
-In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task .
-
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, including the identification of static predicates.
-We report only the precision and recall for the {\em unknown} actions (considering the {\em known} action models would mean reporting higher scores since {\em precision} and {\em recall} in these models is 1.0).
-We must also note that errors in $\Lambda''$ tasks have greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is just done over half of the actions (e.g.
-the preconditions recall for the {\em Floortile} or {\em Gripper} domains).
-
-Remarkably, the overall precision is now $0.98$ which means that the content of the learned models is highly reliable.
-Recall is a lower, 0.87, indicating that the learned models still miss some information.
-Particularly, preconditions are again the component more difficult to be fully learned.
-Overall the reported results confirm the previous trend: more input knowledge drives to better models and requires less planning time.
-Likewise smaller solution plans are required since it is only necessary to program half of the actions (the other half is input knowledge).
-{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
-
-%In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
-
-\subsection{Learning from (initial,final) state pairs}
-Finally we evaluate our approach at learning tasks where input plans are not available, so the planner must not only compute the action models but also the plans that satisfy the input labels.
-Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the obtained results using static predicates and partially specified action models so the inputs to this learning task are $\Psi$ ,$\Sigma$ and $\Xi_0$.
-
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} could not solved the corresponding planning tasks within a 1000 secs time bound.
-Precision and recall have now low values.
-The learning hypothesis space is now so low constrained that actions can be reformulated and still be compliant with the inputs (e.g.
-the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
-We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) but still the achieved scores are below the scores obtained when learning from labeled plans (despite no other input knowledge was provided).
-
-%Table~\ref{tab:time_states} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the classical planning instances that result from our compilation as well as the number of actions in the solutions.
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported since the compilation could not be solved within a time bound of 1000 seconds.
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\resizebox{\columnwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-				Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.90 \\
-				Ferry & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
-                Grid & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.84 & 0.78 \\
-				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-				Miconic & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Satellite & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
-				Transport & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
-				Zenotravel & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.78 \\
-				\hline
-				\bf  & 0.98 & 0.71 & 1.0 & 0.98 & 1.0 & 0.95 & 0.98 & 0.87 \\
-			\end{tabular}
-		}
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} learning from partially specified action models.}
-\label{tab:results_plans_partial}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-                         \hline
-			Blocks & 0.07 & 0.01 & 54  \\
-			Driverlog & 0.03 & 0.01 & 40 \\
-			Ferry & 0.06 & 0.03 & 45 \\
-			Floortile & 0.43 & 0.42 & 55 \\
-                        Grid & 3.12 & 3.07 & 53 \\
-			Gripper & 0.03 & 0.01 & 35 \\
-			Miconic & 0.03 & 0.01 & 34  \\
-			Satellite & 0.14 & 0.14 & 47 \\
-			Transport & 0.23 & 0.21 & 37 \\
-			Zenotravel & 0.90 & 0.89 & 40 \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Time and plan length learning from partially specified action models.}
-	\label{tab:time_plans_partial}	
-\end{table}
-
-\section{Past and future of learning action models}
+===============================================
 
@@
 {\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
-These early works were based on lifting the observed states supported by exploratory plans or external teachers.%, but none provided a theoretical justification for this second source of knowledge.
+These early works were based on lifting the observed states supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge.
 
-%More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
+More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
 
-%Whilst the aforementioned works deal with full state observability,
-Action model learning has also been studied in domains where there is partial or missing state observability.
+Whilst the aforementioned works deal with full state observability,action model learning has also been studied in domains where there is partial or missing state observability.
 {\sf ARMS} works when no partial intermediate state is given.
@@
 
-%Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
+Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
 An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
@@
 The intuitive assumptions of {\sf LOCM}
-%, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence,
+, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence,
 yield a learning model heavily reliant on the kind of domain structure.
@@
 
-Compiling the learning of action models into classical planning is a general and flexible approach that allows to accommodate different amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
-For instance, the example plans in $\Pi$ could be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), so learning tasks $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ could also be addressed.
-Furthermore, our approach seems extensible to learning other types of generative models (e.g.
-hierarchical models like HTN or behaviour trees), that can be more appealing than \strips\ models, since using them to compute planning solutions requires less search effort.
-
-%Furthermore, the compilation scheme can be extended to learn other type of generative models that comprise partially-specified solution plans like HTN models or full-specified solutions like FSM or behaviour trees.
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\resizebox{\columnwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-            Blocks & 0.33 & 0.33 & 0.75 & 0.50 & 0.33 & 0.33 & 0.47 & 0.39 \\
-            Driverlog & 1.0 & 0.29 & 0.33 & 0.67 & 1.0 & 0.50 & 0.78 & 0.48 \\
-            Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
-            Floortile & 0.67 & 0.40 & 0.50 & 0.40 & 1.0 & 0.40 & 0.72 & 0.40 \\
-            Grid & - & - & - & - & - & - & - & - \\
-            Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-            Miconic & 0.0 & 0.0 & 0.33 & 0.50 & 0.0 & 0.0 & 0.11 & 0.17 \\
-            Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
-            Transport & 0.0 & 0.0 & 0.25 & 0.5 & 0.0 & 0.0 & 0.08 & 0.17 \\
-            Zenotravel & - & - & - & - & - & - & - & - \\
-            \hline
-            & 0.63 & 0.29 & 0.54 & 0.70 & 0.67 & 0.53 & 0.61 & 0.51 \\			
-			\end{tabular}
-		}
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} learning from state pairs.}
-\label{tab:results_states}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-			\hline
-            Blocks & 2.14 & 0.00 & 58  \\
-            Driverlog & 0.09 & 0.00 & 88 \\
-            Ferry & 0.17 & 0.01 & 65 \\
-            Floortile & 6.42 & 0.15 & 126 \\
-            Grid & - & - & - \\
-            Gripper & 0.03 & 0.00 & 47 \\
-            Miconic & 0.04 & 0.00 & 68 \\
-            Satellite & 4.34 & 0.10 & 126 \\
-            Transport & 2.57 & 0.21 & 47 \\			
-            Zenotravel & - & - & - \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Time and plan length when learning from state pairs.}
-	\label{tab:time_states}	
-\end{table}
-
-\section{Conclusions}
-We presented a novel approach for learning \strips\ action models from examples using classical planning.
-The approach is flexible to different amounts of available input knowledge and accepts partially specified action models.
-We also introduced {\em precision} and {\em recall} metrics, widely used in ML, to the evaluation of planning action models.
-These metrics separately assess the correctness and completeness of the learned models making easier the identification of flaws.
-
-As far as we know, this is the first work on learning action models exclusively using an {\em off-the-shelf} classical planner and evaluated over a wide range of different domains.
-Recently, ~\citeauthor{stern2017efficient}~\citeyear{stern2017efficient} proposed a classical planning compilation for learning action models but following the {\em finite domain} representation for the state variables and did not report experimental results since the compilation was not implemented.
-
-Our evaluation shows that, when example plans are available, we can compute accurate action models from small sets of learning examples (only five learning examples per domain) and investing small learning times (less than a second time in most of the domains).
-When action plans are not available, our approach is still able to produce action models compliant with the input information.
-In this case, since learning is not constrained by actions it can reformulate operators changing their semantics and making tricky their comparison with a reference model.
-%An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
-
-% The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
-In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
-
-The size of the compiled classical planning instances depend on the number of input examples.
-Generating {\em informative} examples for learning planning action models is an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{geffner:novelty:IJCAI17} motivates the development of novel techniques able to autonomously collect informative learning examples.
-The combination of such exploration techniques with our learning approach is an intriguing research direction that opens the door to the bootstrapping of planning action models.
-
-% Commented for blind submission
-%\begin{small}
-%\subsection*{Acknowledgment}
-%Diego Aineto is partially supported by the {\it FPU} program funded by the Spanish government.
-Sergio Jim\'enez is partially supported by the {\it Ramon y Cajal} program funded by the Spanish government.
-%\end{small}
-
-\newpage
-
-\bibliographystyle{aaai}
-\bibliography{strips-learning}
-\end{document}
+Recenctly classical planning compilations have been defined to learn different kinds of generative models from examples.
+The existing compilations for computing FSCs for generalized planning follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it and hence, they tightly integrate planning and generalization.
+To keep the computation of FSCs tractable, they limit the space of possible solutions bounding the maximum size of the FSC.
+In addition, they impose that the instances to solve share, not only the domain theory (actions and predicates schemes) but the set of fluents~\cite{javi-Gplanning-ICAPS16} or a subset of {\it observable} fluents~\cite{Geffner:FSM:AAAI10}.
+Programs increase the readability of FSCs separating the control-flow structures from the primitive actions.
+Like FSCs, programs can also be computed following a {\it top-down} approach, e.g.~exploiting compilations that program and validate the program on instances with the same state and action space~\cite{javi-Gplanning-ICAPS16}.
+Since these {\it top-down} approaches search in the space of solutions, it is helpful to limit the set of different control-flow instructions.
+For instance using only {\it conditional gotos} that can both implement branching and loops~\cite{Jimenez15}.

### 2018-04-16 doc/aij18/related_work.tex
Added input comparison table

@@
 
-TABLA
+\begin{table}
+	\small
+	\centering
+	\begin{tabular}{ l | c | c | c | c | c |}
+		& \multicolumn{1}{|p{2cm}|}{FO States sequences} 
+		& \multicolumn{1}{|p{2cm}|}{FO Actions sequences} 
+		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO states)}
+		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO actions)}
+		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO both)}   \\
+		\hline			
+		ARMS &  & \checkmark & \checkmark & & \\
+		AMAN & & & \checkmark & &\\
+		LAMP & & & \checkmark & &\\
+		SLAF & & & \checkmark & &\\
+		LOCM & & \checkmark & & &\\
+		FAMA & \checkmark & \checkmark &  \checkmark & \checkmark & \checkmark \\
+	\end{tabular}
+	\caption{Input comparison of main learning approaches. PO=Partial Observability, FO=Full Observability}
+	\label{table:input_comparison}
+\end{table}
 

### 2018-04-16 doc/aij18/aij18.tex
updates added to the remaining sections AIJ18 paper

@@
 Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
-A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plans that explains the given inputs to the learning task.
+A planning compilation is a suitable approach because solving $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plan that explains the given inputs to the learning task.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \item {\bf Validates the action model $\mathcal{M}'$}.
-The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) with the programmed action model $\mathcal{M}'$.
+The solution plan continues with a postfix that reproduces the input observations of the plan executions with the programmed action model $\mathcal{M}'$.
 \end{enumerate}
@@
 
-\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{O}$, {\tt\small $1\leq i< |\mathcal{O}|$}.
+\item Actions for {\em validating} partially observed states $s_i\in\mathcal{O}$, {\tt\small $1\leq i< |\mathcal{O}|$}.
 \begin{small}
@@
 If any reference to the $mode_{val}$ fluent is removed the compilation can learn \strips\ action models when there are missing state observations in $\mathcal{O}$.
-On the other hand, if the $mode_{val}$ fluent is ignored, $P_{\Lambda}$ becomes harder since the classical planner must determine how many {\em apply} actions are necessary between any two observations, i.e.
+In this case $P_{\Lambda}$ becomes harder because the planner must determine how many {\em apply} actions are necessary between two observations, i.e.
 between the application of two {\em validate} actions.
@@
 In this case, the corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
-When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the compilation validates whether the observation of the plan execution follows the given model:
+When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, a solution to $P_{\Lambda}$ validates whether the observation of the plan execution follows the given model:
 \begin{itemize}
@@
 \item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given input data if $P_{\Lambda}$ is unsolvable. 
-This means that $\mathcal{M}$ cannot be compliant with the given observation of the plan execution.
 \end{itemize}
@@
 \begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with the $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that corresponds to action $a_j$.
+\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with fluents $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ to code the $j^{th}$ action $a_j\in\pi$.
 The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to iterate through the steps of plan $\pi$.
@@
 \item The actions for {\em applying} an already programmed operator includes the following extra conditional effects $\{at_{j},plan(name(a_j),\Omega^{ar(a_j)},j)\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism ensures that $\mathsf{apply_{\xi,\omega}}$ actions are applied, exclusively, in the same order as in the example plan $\pi$ while it is flexible to the learning scenario where some $a_j$ actions in $\pi$ are missing.
-In such scenario $P_{\Lambda'}$ becomes harder because the planner must determine which and how many $\mathsf{apply_{\xi,\omega}}$ actions are necessary for validating the learned action model with the given input knowledge.
+This mechanism ensures that $\mathsf{apply_{\xi,\omega}}$ actions are applied in the order specified by the observed plan $\pi$ while it is flexible when some $a_j$ actions in $\pi$ are missing.
+In such scenario $P_{\Lambda'}$ becomes harder because the planner must determine the $\mathsf{apply_{\xi,\omega}}$ actions that are necessary for validating the learned action model in the given observations.
 \end{enumerate}
@@
 The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $\mathcal{O}=\tup{s_0,s_4}$ shown in Figure~\ref{fig:example-plans}.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ and the two state observations $\mathcal{O}=\tup{s_0,s_4}$ shown in Figure~\ref{fig:example-plans}.
 Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the four-action plan $\pi$ shown in the Figure~\ref{fig:example-plans}.
@@
 
-Now we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
+Finally we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
 Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
@@
 \begin{mylemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{O}',\Psi,\Pi}$.
 \end{mylemma}
@@
   \begin{small}
-Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
-In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from the corresponding initial state and following the sequence of actions defined by the plans in $\Pi$.
+    An operator schema $\xi\in\mathcal{M}'$ can only be programmed once by a plan $\pi$.
+This is forced with the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ in the preconditions of the corresponding programming actions $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$.
+
+    After an operator schema $\xi\in\mathcal{M}'$ is programmed, it can only be applied by a plan $\pi$.
+The $mode_{prog}$ fluent is a precondition of both $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$ actions that is removed after any application of an $\mathsf{apply_{\xi,\omega}}$ action.
+
+    Finally the goals of the $P_{\Lambda}$ specify that $P_{\Lambda}$ is solved iff the fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last state reached by a plan $\pi$.
+The $\{test_i\}$ fluents can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from the corresponding initial state and following the sequence of actions defined by the plans in $\Pi$.
 This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
@@
 Completeness.
-Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$ learning task, is computable with a plan $\pi$ that solves the corresponding classical planning task $P_{\Lambda}$.
 \end{mylemma}
@@
 \begin{small}
-By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O},\Pi$.
+By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its $head(\xi)$ and the set of predicates $\Psi$.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ provided that they satisfy the state trajectory constraint given by $\mathcal{O},\Pi$.
 \end{small}
@@
 \begin{itemize}
+\item The number of given state observations. There is a $test_i$ fluent and a $\mathsf{validate_{i}}$ action in $P_{\Lambda}$ for every $|\mathcal{O}|$.
 \item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
-The larger these numbers, the larger the size of the $F_v(\xi)$ sets.
-This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
-\item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+The larger these numbers, the larger the size of the $F_v(\xi)$ sets which confines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluent sets and the corresponding set of {\em programming} actions.
+The size of the $F_v(\xi)$ sets is the term that dominates the compilation size because it grows exponentially with the expresion $|pars(\xi)|^{ar(p)}$ for $p\in\Psi$ s.t.
+$ar(p)<|pars(\xi)|$.
 \end{itemize}
@@
 \label{sec:Section6}
-The roles of two action schemes whose headers match or the roles of two action parameters that belong to the same type can be swapped by a $\Lambda$ learning task.
+When the $\Lambda$ learning task is underconstrained different \strips\ action model can be compliant with the observations.
+For instance the roles of two action schemes, whose headers match, or the roles of two action parameters that belong to the same type can be swapped and still be compliant with the given input data.
 For instance, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped.
-Pure syntax-based metrics can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
+In this learning scenario pure syntax-based metrics can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
 a learned model semantically equivalent but syntactically different to the reference model.
@@
 Here we introduce an evaluation approach that is robust to role changes of this particular kind.
-The intuition of the approach is to assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
+The intuition of the approach is to assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions $\mathcal{O}$ and $\Pi$ according to the amount of {\em edition} required by $\mathcal{M}$ to reproduce that observations.
 %In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
@@
 \begin{mydefinition}
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that they are {\em comparable}.
+  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} \strips\ action models.
 The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
@@
 As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions~\cite{yang2007learning}.
-With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
+With this regard, we define an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
 
@@
 The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
-This {\em edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
-
-The error of a learned action model could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
-This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes.
-Our definition of the edit distance is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
-Finally, the {\em edit distance} can also be defined with respect to a set of plans, if they are available.
-
+This {\em edit distance} could be mapped into a $P(\mathcal{O}|\mathcal{M})$ likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
+
+The error of a learned action model could also be defined the other way around, quantifying the amount of edition required by the observations of the plan execution to match the given model.
+This would imply defining {\em edit operations} that modify the fluents in the state observations $\mathcal{O}$ instead of the {\em edit operations} that modify the action schemes $\mathcal{M}$.
+We claim that our definition of the edit distance is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
+
+Finally, the {\em observation edit distance} can also be defined with respect to a set of plans, if they are available. 
 \begin{mydefinition}
-  Given an action model $\mathcal{M}$ and a set of valid plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ that only contain actions built grounding the schemes in $\mathcal{M}$.
-The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\Pi)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$}; \[\delta(\mathcal{M},\Pi)=\min_{\forall \mathcal{M}' \rightarrow \Pi} \delta(\mathcal{M},\mathcal{M}')\]
+  Given an action model $\mathcal{M}$ and a set of valid plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ that only contain actions built grounding the schemes in $\mathcal{M}$ and that induces the corresponding state observations $\mathcal{O'}$ built with fluents in $F$.
+The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\tup{\Pi,\mathcal{O'}})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$} and the corresponding state observations $\mathcal{O'}$; \[\delta(\mathcal{M},\tup{\Pi,\mathcal{O'}})=\min_{\forall \mathcal{M}' \rightarrow \tup{\Pi,\mathcal{O'}}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{mydefinition}
@@
 In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema.
-However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}.
+With the aim of keeping tractable the planning instances that result from our compilation, we only implement {\em deletion} and {\em insertion}.
 \item {\bf Validates the edited model $\mathcal{M}'$ in observations of the plan executions}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{O}$ (and on $\Pi$ if available), as explained in Section~\ref{sec:Section5} for the models that are programmed from scratch.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{O}$ (and $\Pi$ if available), as explained in Section~\ref{sec:Section5}.
 \end{enumerate}
@@
 \item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
-\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
+\item $I_{\Lambda}'$ encodes $s_0\subseteq F$ and sets $mode_{prog}$ to true.
 In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
+This means that fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold iff they appear in $\mathcal{M}$.
 \item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
-The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
+The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are exactly defined as in the previous compilation.
 The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
@@
 
-When the executed plans $\Pi$ are also available, the compilation can be adapted to compute the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
-The modifications to the compilation explained in Section~\ref{sec:Section5} are also useful here to redefine the learning task as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$.
+When the executed plans $\Pi$ are also available, the compilation can be adapted to compute $\delta(\mathcal{M},\tup{\Pi,\mathcal{O'}})$.
+The modifications to the compilation explained in Section~\ref{sec:Section5} are also useful here to redefine the learning task as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$ as well as the correspoding observations $\mathcal{O'}$, which results in the edited model $\mathcal{M}'$.
 Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
@@
 The sets of fluents $F_{\Lambda}$ and $I_{\Lambda}$ are defined like in the previous compilation.
-With respect to the actions, $A_{\Lambda}$ again implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect) but does not contain apply actions because the \strips\ action model are not validated in any observation of plan executions.
+Again the programming actions in $A_{\Lambda}$ implement the two {\em edit operations} (i.e.
+include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect) but now, $A_{\Lambda}$ does not contain $\mathsf{apply_{\xi,\omega}}$ or $\mathsf{validate_{i}}$ actions because the \strips\ action model are not validated in any observation of a plan execution.
 Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_v(\xi)$.
+
 To illustrate this, the plan of Figure~\ref{fig:plan-mdistance} solves the classical planning task that corresponds to computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.
@@
 
-\section{Recognition of \strips\ action models}
-\label{sec:Section7}
-Given a set of possible \strips\ models and set of observations of plan executions, the {\em recognition of \strips\ models} is the task of identifying which model has the highest probability of producing the given observations.
+\subsection{Recognition of \strips\ action models}
+Given a set of possible \strips\ models and set of observations of plan executions, the {\em recognition of \strips\ models} is the task of identifying which model from the set has the highest probability of producing the given observations.
 
@@
 \item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$.
-If a set of plans $\Pi$ is available, this same strategy can be followed using the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
-\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
+If a set of plans $\Pi$ is available, this same strategy can be followed using $\delta(\mathcal{M},\tup{\Pi,\mathcal{O'}})$.
+\item  Maping this distance into a $P(\mathcal{O}|\mathcal{M})$ likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$ normalizing this ratio with regard to the size of the $F_v(\xi)$ and $\mathcal{O}$ sets.
 \item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
@@
 
+An example.
+
 \section{Learning \strips\ action models with background knowledge}
-A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn explanations from data~\cite{muggleton1994inductive}.
-Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case, given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.
+\label{sec:Section7}
+A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn explanations of the input data~\cite{muggleton1994inductive}.
+Inspired by ILP, we show that our approach for learning \strips\ action models can also leverage {\em background knowledge} that is given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.
 
@@
 
-{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$.
-
-The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
-For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}.
+{\em State invariants} is a kind of state-constraints useful for more compact state representations~\cite{helmert2009concise} or for making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s\models \phi$, built from $F$, that is reachable with a sequence of applicable actions in $A$ and starting from $I$.
+
+The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states that are reachable with a sequence of applicable actions in $A$ and starting from $I$.
+For instance, Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}.
 There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.
@@
 \subsection{Trajectory constraints}
-Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10,bacchus:LTL:1998}.
-LTL is a modal temporal logic interpreted over sequences of states.
+Instead of enumerating the full sequence of states included in certain trajectories, they can be implicitly defined with {\em state trajectory constraints}.
+Constraints of this kind can be formulated with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10,bacchus:LTL:1998}, a modal temporal logic interpreted over sequences of states.
 LTL interpreted over finite sequences of states (also called traces) has received attention from the planning community and provided a formal description called $LTL_f$~\cite{Giacomo:LTLf:AAAI2014}.
 
-$LTL_f$ is built up from a finite set of propositional variables $P$, the logical operators $\neg$ and $\vee$ (from which it is possible to define operators $\wedge$, $\rightarrow$ and $\leftrightarrow)$, and the temporal modal operators {\em next}($\bigcirc$), and {\em until}(${\mathcal U}$).
-An $LTL_f$ formula is then inductively defined over a set of propositions $P$:
-\begin{itemize}
-\item A proposition $p\in P$ is an $LTL_f$ formula,
+$LTL_f$ is built up from a finite set of fluents $F$, the logical operators $\neg$ and $\vee$ (from which it is possible to define operators $\wedge$, $\rightarrow$ and $\leftrightarrow)$, and the temporal modal operators {\em next}($\bigcirc$), and {\em until}(${\mathcal U}$).
+An $LTL_f$ formula is then inductively defined over the set of fluents $F$:
+\begin{itemize}
+\item A proposition $f\in F$ is an $LTL_f$ formula,
 \item if $\psi$ and $\chi$ are $LTL_f$ formulae, then so they are $\neg\psi$, ($\psi\vee\chi$), $\bigcirc\psi$, ($\psi\ {\mathcal U}\ \chi$).
@@
 \begin{itemize}
-\item $\mathcal{O}, i\models f$ for a propositional variable $f\in F$, iff $f\in s_i$,
+\item $\mathcal{O}, i\models f$ for a fluent $f\in F$, iff $f\in s_i$,
 \item $\mathcal{O}, i\models \neg\psi$ iff it is not the case that $\mathcal{O}, i\models\psi$,
@@
 
-\subsection{Leveraging background knowledge}
-We have shown how our approach is flexible to validate the learned action models with respect to a set of observations of executed plans and with respect to a model that acts as a reference.
-Here we show that the validation of an action model can also be done with {\em state constraints} or {\em trajectory constraints}.
+\subsection{Learning \strips\ action models with state and trajectory constraints}
+We have shown how our approach is flexible to validate the learned action models with respect to various amount and kind of input knowledge including, a set of observations of executed plans and, a model that acts as a reference.
+Here we show that the validation of a learned action model can also be done with {\em state constraints} or {\em trajectory constraints}.
 Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check that the learned model satisfy every constraint $\phi_i\in\Phi$:
@@
 
-Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
+Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned, the more efficient the computation of a solution plan.
 Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space.
@@
 \end{itemize}
+
+Dicusion on the relations to the $\delta(\mathcal{M},\mathcal{M}')$ distance.
 

### 2018-04-16 doc/aij18/related_work.tex
Added other comparison tables

@@
 	\label{table:input_comparison}
+\end{table}
+
+\begin{table}
+	\small
+	\centering
+	\begin{tabular}{ l | c | c | c |}
+		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO states)}
+		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO actions)}
+		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO both)}   \\
+		\hline			
+		ARMS & \checkmark & & \\
+		AMAN & \checkmark & & \\
+		CAMA & \checkmark & & \\
+		LAMP & \checkmark & & \\
+		SLAF & \checkmark & & \\
+		LOCM & \checkmark & & \\
+		FAMA & \checkmark & \checkmark & \checkmark \\
+	\end{tabular}
+	\caption{PO=Partial Observability, FO=Full Observability}
+	\label{table:input_comparison2}
+\end{table}
+
+\begin{table}
+	\small
+	\centering
+	\begin{tabular}{ l | c | c |}
+		& States 
+		& Actions  \\
+		\hline
+		LOCM & zero & FO\\		
+		ARMS & zero or PO & FO \\
+		AMAN & PO & FO \\
+		CAMA & PO & FO \\
+		LAMP & PO & FO \\
+		SLAF & PO & FO \\		
+		FAMA & zero or PO & zero or PO \\
+	\end{tabular}
+	\caption{PO=Partial Observability, FO=Full Observability}
+	\label{table:input_comparison3}
 \end{table}

### 2018-04-18 doc/aij18/aij18.tex
updates related work
new paper organization

--- 2018-04-16 doc/aij18/aij18.tex
@@
 \usepackage{listings}
+\usepackage{xspace}
 
@@
 \newcommand{\strips}{\textsc{Strips}}     % precondition
+
+\newcommand{\ARMS}{{\small {\sffamily ARMS}}\xspace}
+\newcommand{\CAMA}{{\small {\sffamily CAMA}}\xspace}
 
@@
 
-\title{Learning, Evaluation and Recognition of \strips\ Action Models with Classical Planning}
+\title{Learning \strips\ Action Models with Classical Planning}
 \author[label1]{Diego Aineto}
@@
 \input{introduction}
-\input{related_work}
-
-\section{Background}
-\label{sec:Section3}
-This section defines the planning model used on this work, {\em classical planning with conditional effects}, and the target of the learning/evaluation/recognition tasks addressed in the paper, a \strips\ action model.
-
-\subsection{Classical planning with conditional effects}
-We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
-A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
-either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent).
-We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
-
-A {\em state} $s$ is a full assignment of values to fluents, i.e.
-$|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
-
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$ and {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$.
-We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
-The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
-
-%\begin{itemize}
-%\item $\pre(a)\subseteq\mathcal{L}(F)$, called {\em preconditions}, the literals that must hold for the action $a\in A$ to be applicable.
-%\item $\eff^+(a)\subseteq\mathcal{L}(F)$, called {\em positive effects}, that defines the fluents set to true by the application of the action $a\in A$.
-%\item $\eff^-(a)\subseteq\mathcal{L}(F)$, called {\em negative effects}, that defines the fluents set to false by the action application.
-%\end{itemize}
-
-An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
-\[
-triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
-\]
-
-The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
-
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The {\em plan length} is denoted with $|\pi|=n$ .
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
-
-\subsection{\strips\ action schemas}
-This work addresses the learning of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
-
-\begin{figure}[hbt!]
-\begin{footnotesize}
-\begin{verbatim}
-(:action stack
- :parameters (?v1 ?v2 - object)
- :precondition (and (holding ?v1) (clear ?v2))
- :effect (and (not (holding ?v1)) (not (clear ?v2)) (handempty) (clear ?v1) (on ?v1 ?v2)))
-\end{verbatim}
-\end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
-\label{fig:stack}
-\end{figure}
-
-To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
-We define $F_v$, a new set of fluents s.t.
-$F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$, i.e.
-the variable names, and that defines the elements that can appear in an action schema.
-For the {\em blocksworld}, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-
-For a given operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of fluents that represent the elements that can appear in that action schema.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.
-
-We assume also that actions $a\in A$ are instantiated from \strips\ operator schemas $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
-\begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
-The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
-\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-\end{itemize}
-Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, $2^{2|F_v(\xi)|}$ defines the size of space of the possible \strips\ models for that operator, given that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
-For instance, this number is 4194304 for the blocksworld {\tt stack} operator while is only 1024 for the {\tt pickup} operator.
-
-Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same headers so they can be built from the same set of possible elements.
-Formally, iff $head(\xi)=head(\xi')$ so it also holds that $F_v(\xi)=F_v(\xi')$.
-For instance we can claim that the {\tt stack} and {\tt unstack} blocksworld operators are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
-Likewise we say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.
-
-\section{Learning \strips\ action models}
-\label{sec:Section4}
-Learning \strips\ action models from fully available input knowledge, i.e.
-from plans where the {\em pre-} and {\em post-states} of every action are available, is straightforward~\cite{jimenez2012review}:
-\begin{itemize}
-  \item {\em Preconditions} are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action, that is any action that belongs to the same operator scheme.
-  \item {\em Effects} are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-\end{itemize}
-This section formalizes more challenging learning tasks where less input knowledge is available for instance, because it cannot be observed.
-
-\subsection{Learning from observations of plan executions}
-The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its actions, the actual executed actions are unobserved.
-This learning task is formalized as a tuple $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$:
-\begin{itemize}
-\item $\mathcal{M}$, the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
-\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is the sequence of {\em state observations} obtained watching the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-We assume that the initial state $s_0$ is {\em fully observable} while states $s_i$ s.t.
-{\small $1\leq i\leq |\mathcal{O}|$} can be {\em partially observable}, meaning that some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative.
-In the extreme, states $s_i$ {\small $1\leq i\leq |\mathcal{O}|$}, can be missing but we assume that any state observation is {\em noiseless}, meaning that if the value of a fluent is observed it is correct.
-\item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame.
-Note that $\Psi$ can be inferred from the state observations provided that at least a state $s\in \mathcal{O}$ is a full state, that is $|s|=|F|$.
-\end{itemize}
-
-A solution to a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the given state observations $\mathcal{O}$ and the predicates $\Psi$.
-Solving $\Lambda$ implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model.
-Figure~\ref{fig:example-observations} shows an example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence obtained inverting a 2-block tower.
-Learning the action model from this example implies inferring the unobserved plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$.
-
-\begin{figure}[hbt!]
-{\footnotesize\tt ;;;;;; Headers in $\mathcal{M}$}
-\begin{footnotesize}
-\begin{verbatim}
-(pickup v1) (putdown v1) (stack v1 v2} (unstack v1 v2)
-\end{verbatim}
-\end{footnotesize}
-\vspace{0.2cm}
-{\footnotesize\tt ;;;;;; Predicates $\Psi$}
-\begin{footnotesize}
-\begin{verbatim}
-(handempty) (holding ?o  - object) (clear ?o - object) (ontable ?o - object)
-(on ?o1 - object ?o2 - object)
-\end{verbatim}
-\end{footnotesize}
-\vspace{0.2cm}
-{\footnotesize\tt ;;;;;; Observations $\mathcal{O}$}
-\begin{footnotesize}
-\begin{verbatim}
-;;; observation #0
-(clear B) (on B A) (ontable A) (handempty)
-
-;;; observation #1
-(holding B) (clear A) (ontable A)
-
-;;; observation #2
-(clear A) (ontable A) (clear B) (ontable B) (handempty)
-
-;;; observation #3
-(holding A) (clear B) (ontable B)
-
-;;; observation #4
-(clear A) (on A B) (ontable B) (handempty)
-\end{verbatim}
-\end{footnotesize}
- \caption{\small Task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
-\label{fig:example-observations}
-\end{figure}
-
-In some cases we may not require to start learning from scratch because some preconditions and/or some positive or negative effects may be a priori known.
-The operator schemas in $\mathcal{M}$, that are given as input, may be not {\em empty} but {\em partially specified}.
-Such scenario is relevant for {\em policy learning}.
-A policy is function that maps states into actions and represents the conditions under which actions should be applied to achieve certain goals.
-Given an action model, the task of learning a policy that is compliant with a set of observations can be defined as learning extra preconditions for each action scheme in the model.
-These extra preconditions capture when actions can be applied according to the policy.
-A policy that exactly defines a single applicable action for each reachable state is a {\em full policy} otherwise, we say it is a {\em partial policy}.
-In the general case, learning a {\em full policy} for an arbitrary planning task is complex because a given action scheme may be applicable in different situations and also because the compact representation of this set of different situations may require the computation of {\em high-level state features}~\cite{lotinac2016automatic}.
-
-The $\Lambda$ learning task can also be redefined to cover the scenario where some of the actions executed by the observed agent are available.
-If all the executed actions are known then states observations should be partial otherwise, the learning task is trivial.
-The learning task is now formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
-\begin{itemize}
-\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is the action sequence that produces the sequence of state observations given in $\mathcal{O}$.
-Again we assume that action observations are noiseless, meaning that if the value of an action is observed it is correct.
-When the input plan is {\em diverse} enough, i.e.
-$\pi$ contains at least one ground action for each of the aimed action schemes, the set of {\em empty} operator schemas $\mathcal{M}$ can be inferred from $\pi$.
-\end{itemize}
-
-Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of a four-action plan for inverting a two-block tower.
-In this example $\mathcal{O}=\tup{s_0,s_4}$ which means that only the first and last states are observed and the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
-$\mathcal{M},\Psi$ are skipped, since they are the same as in Figure~\ref{fig:example-observations}.
-
-\begin{figure}[hbt!]
-{\footnotesize\tt ;;;;;; Observations $\mathcal{O}$}
-\begin{footnotesize}
-\begin{verbatim}
-;;; observation #0
-(clear B) (on B A) (ontable A) (handempty)
-
-;;; observation #4
-(clear A) (on A B) (ontable B) (handempty)
-\end{verbatim}
-\end{footnotesize}
-
-{\footnotesize\tt ;;; Plan $\pi$}
-\begin{footnotesize}
-\begin{verbatim}
-0: (unstack B A)
-1: (putdown B)
-2: (pickup A)
-3: (stack A B)
-\end{verbatim}
-\end{footnotesize}
-
- \caption{\small Task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for learning a {\em blocksworld} \strips\ action model from a four-action plan and two state observations.}
-\label{fig:example-plans}
-\end{figure}
-
-The previous definitions formalize the learning of \strips\ action models from the observation of a single plan execution.
-These definitions are extensible to the more general case where the execution of multiple plans is observed.
-The task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O'},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given sequence of example plans producing the corresponding sequence of state observations $\mathcal{O'}$.
-Now $\mathcal{O'}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ is a sequence of {\em state observations} obtained watching the execution of a serie of {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, one after the other.
-This means that, for each {\small $1\leq i\leq n^t$}, $a_i^t\in \pi^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-
-\section{Learning \strips\ action models with classical planning}
-\label{sec:Section5}
-Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
-A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plans that explains the given inputs to the learning task.
-The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
-
-\begin{enumerate}
-\item {\bf Programs the action model $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item {\bf Validates the action model $\mathcal{M}'$}.
-The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) with the programmed action model $\mathcal{M}'$.
-\end{enumerate}
-
-\subsection{Learning from state observations}
-Here we formalize the compilation for learning \strips\ action models with classical planning.
-Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
-\begin{itemize}
-\item $F_{\Lambda}$ contains:
-\begin{itemize}
-\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the input observations, i.e.
-the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-observations}.
-Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
-\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
-For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
-\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation in $\mathcal{O}$ where the action model is validated.
-\end{itemize}
-\item $I_{\Lambda}$ encodes the first observation, $s_0\subseteq F$, and sets $mode_{prog}$ to true.
-Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
-Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
-
-\item $G_{\Lambda}=\bigcup_{1\leq i\leq |\mathcal{O}|}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
-\item $A_{\Lambda}$ comprises three kinds of actions:
-\begin{enumerate}
-\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$:
-\begin{itemize}
-\item Actions for {\bf removing} a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}, pre_{f}(\xi)\},\\
-\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
-\end{align*}
-\end{small}
-
-\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
-\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
-\end{align*}
-\end{small}
-\end{itemize}
-
-\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
-Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
-Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$ from {\em blocksworld}.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\cup \{\neg mode_{val}\},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
-&\{\emptyset\}\rhd\{mode_{val}\}.
-\end{align*}
-\end{small}
-
-\begin{figure}[hbt!]
-\begin{scriptsize}
-\begin{verbatim}
-(:action apply_stack
-  :parameters (?o1 - object ?o2 - object)
-  :precondition
-   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
-        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
-        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
-        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
-        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
-        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
-        (or (not (pre_clear_stack_v1)) (clear ?o1))
-        (or (not (pre_clear_stack_v2)) (clear ?o2))
-        (or (not (pre_holding_stack_v1)) (holding ?o1))
-        (or (not (pre_holding_stack_v2)) (holding ?o2))
-        (or (not (pre_handempty_stack)) (handempty)))
-  :effect
-   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
-        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
-        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
-        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
-        (when (del_ontable_stack_v1) (not (ontable ?o1)))
-        (when (del_ontable_stack_v2) (not (ontable ?o2)))
-        (when (del_clear_stack_v1) (not (clear ?o1)))
-        (when (del_clear_stack_v2) (not (clear ?o2)))
-        (when (del_holding_stack_v1) (not (holding ?o1)))
-        (when (del_holding_stack_v2) (not (holding ?o2)))
-        (when (del_handempty_stack) (not (handempty)))
-        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
-        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
-        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
-        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
-        (when (add_ontable_stack_v1) (ontable ?o1))
-        (when (add_ontable_stack_v2) (ontable ?o2))
-        (when (add_clear_stack_v1) (clear ?o1))
-        (when (add_clear_stack_v2) (clear ?o2))
-        (when (add_holding_stack_v1) (holding ?o1))
-        (when (add_holding_stack_v2) (holding ?o2))
-        (when (add_handempty_stack) (handempty))
-        (when (modeProg) (not (modeProg)))))
-\end{verbatim}
-\end{scriptsize}
- \caption{\small PDDL action for applying an already programmed schema $stack$ (implications are coded as disjunctions).}
-\label{fig:compilation}
-\end{figure}
-
-\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{O}$, {\tt\small $1\leq i< |\mathcal{O}|$}.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\cup \{mode_{val}\},\\
-\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
-\end{align*}
-\end{small}
-\end{enumerate}
-\end{itemize}
-
-The compilation approach is flexible to various amount and kind of input data.
-If any reference to the $mode_{val}$ fluent is removed the compilation can learn \strips\ action models when there are missing state observations in $\mathcal{O}$.
-On the other hand, if the $mode_{val}$ fluent is ignored, $P_{\Lambda}$ becomes harder since the classical planner must determine how many {\em apply} actions are necessary between any two observations, i.e.
-between the application of two {\em validate} actions.
-Further, known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
-In this case, the corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
-When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the compilation validates whether the observation of the plan execution follows the given model:
-\begin{itemize}
-\item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given input data if a solution plan for $P_{\Lambda}$ can be found.
-\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given input data if $P_{\Lambda}$ is unsolvable.
-This means that $\mathcal{M}$ cannot be compliant with the given observation of the plan execution.
-\end{itemize}
-This feature of our compilation is beyond the functionality of VAL, the plan validation tool~\cite{howey2004val}, because VAL requires (1) a full plan and (2), a full action model for plan validation.
-
-Figure~\ref{fig:plan-observations} illustrate how our compilation works and shows a plan that solves the classical planning task that results from our compilation.
-This plan programs and validates the $stack$ schema (from {\em blocksworld}) using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.
-Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,21]$ validate the programmed operators using the sequence of five state observations shown in the Figure~\ref{fig:example-observations}.
-
-\begin{figure}[hbt!]
-{\footnotesize\tt
-     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
-     01 : (program\_pre\_handempty\_stack)\\
-     02 : (program\_pre\_holding\_stack\_v2)\\
-     03 : (program\_pre\_on\_stack\_v1\_v1)\\
-     04 : (program\_pre\_on\_stack\_v1\_v2)\\
-     05 : (program\_pre\_on\_stack\_v2\_v1)\\
-     06 : (program\_pre\_on\_stack\_v2\_v2)\\
-     07 : (program\_pre\_ontable\_stack\_v1)\\
-     08 : (program\_pre\_ontable\_stack\_v2)\\
-     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
-    10 : (program\_eff\_clear\_stack\_v2)\\
-    11 : (program\_eff\_handempty\_stack)\\
-    12 : (program\_eff\_holding\_stack\_v1)\\
-    13 : (program\_eff\_on\_stack\_v1\_v2)\\
-    {\bf 14} : (apply\_unstack blockB blockA)\\
-    15 : (validate\_1)\\
-    16 : (apply\_putdown blockB)\\
-    17 : (validate\_2)\\
-    18 : (apply\_pickup blockA)\\
-    19 : (validate\_3)\\
-    20 : (apply\_stack blockA blockB)\\
-    21 : (validate\_4)
-}
- \caption{\small Plan for programming and validating the $stack$ schema using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
-\label{fig:plan-observations}
-\end{figure}
-
-\subsection{Learning from plans}
-Now we extend the compilation to consider observed actions.
-Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda}}$ such that:
-\begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with the $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that corresponds to action $a_j$.
-The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to iterate through the steps of plan $\pi$.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<|\pi|$}, for tracking the plan step where the action model is validated.
-Goals are as in the original compilation.
-In other words, the observed states are used for validation adding a $test_t$, ${\small 1\leq t\leq |\mathcal{O}|}$ fluent for each observation in $\mathcal{O}$.
-\item With respect to the set of actions $A_{\Lambda'}$.
-\begin{enumerate}
-\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ and the actions for {\em validating} the programmed action model in a given state observation are the same as in the previous compilation.
-\item The actions for {\em applying} an already programmed operator includes the following extra conditional effects $\{at_{j},plan(name(a_j),\Omega^{ar(a_j)},j)\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism ensures that $\mathsf{apply_{\xi,\omega}}$ actions are applied, exclusively, in the same order as in the example plan $\pi$ while it is flexible to the learning scenario where some $a_j$ actions in $\pi$ are missing.
-In such scenario $P_{\Lambda'}$ becomes harder because the planner must determine which and how many $\mathsf{apply_{\xi,\omega}}$ actions are necessary for validating the learned action model with the given input knowledge.
-\end{enumerate}
-\end{itemize}
-
-The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $\mathcal{O}=\tup{s_0,s_4}$ shown in Figure~\ref{fig:example-plans}.
-Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the four-action plan $\pi$ shown in the Figure~\ref{fig:example-plans}.
-
-\begin{figure}[hbt!]
-{\footnotesize\tt
-     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
-     01 : (program\_pre\_handempty\_stack)\\
-     02 : (program\_pre\_holding\_stack\_v2)\\
-     03 : (program\_pre\_on\_stack\_v1\_v1)\\
-     04 : (program\_pre\_on\_stack\_v1\_v2)\\
-     05 : (program\_pre\_on\_stack\_v2\_v1)\\
-     06 : (program\_pre\_on\_stack\_v2\_v2)\\
-     07 : (program\_pre\_ontable\_stack\_v1)\\
-     08 : (program\_pre\_ontable\_stack\_v2)\\
-     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
-    10 : (program\_eff\_clear\_stack\_v2)\\
-    11 : (program\_eff\_handempty\_stack)\\
-    12 : (program\_eff\_holding\_stack\_v1)\\
-    13 : (program\_eff\_on\_stack\_v1\_v2)\\
-    {\bf 14} : (apply\_unstack blockB blockA i1 i2)\\
-    15 : (apply\_putdown blockB i2 i3)\\
-    16 : (apply\_pickup blockA i3 i4)\\
-    17 : (apply\_stack blockA blockB i4 i5)\\
-    {\bf 18} : (validate\_1)
-}
- \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi$ and state observations $\mathcal{O}$ (shown in Figure~\ref{fig:example-plans}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
-\label{fig:plan-lplan}
-\end{figure}
-
-Now we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
-Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
-same fluents and actions but different initial states and goals).
-The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t$ and the plan $\pi_t$, and the goals $G_t$ are defined by the state $s_n^t$.
-Addressing the learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ with classical planning requires introducing a small modification to our compilation.
-In particular, the actions in $P_{\Lambda'}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq \tau$} reset the current state, and the current plan, and are now defined as:
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{1\leq j<t}\cup\{\neg test_j\}_{t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\} \cup \{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t}.
-\end{align*}
-\end{small}
-
-\subsection{Compilation properties}
-
-\begin{mylemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$.
-\end{mylemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
-In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from the corresponding initial state and following the sequence of actions defined by the plans in $\Pi$.
-This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
-\end{small}
-\end{proof}
-
-\begin{mylemma}
-Completeness.
-Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
-\end{mylemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O},\Pi$.
-\end{small}
-\end{proof}
-
-The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
-\begin{itemize}
-\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
-The larger these numbers, the larger the size of the $F_v(\xi)$ sets.
-This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
-\item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
-\end{itemize}
-
-\subsection{Exploiting static predicates to optimize the compilation}
-A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
-Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
-Given a static predicate $p$:
-\begin{itemize}
-\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable objects} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
-\item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
-\end{itemize}
-
-Static predicates can also constrain the space of possible preconditions by looking at the given set of state observation $\mathcal{O}$.
-One can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\mathcal{O}$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
-For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{O}$.
-
-Furthermore looking as well at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$ and $\mathcal{O}$.
-Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
-
-\section{Evaluating \strips\ action models}
-\label{sec:Section6}
-The roles of two action schemes whose headers match or the roles of two action parameters that belong to the same type can be swapped by a $\Lambda$ learning task.
-For instance, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped.
-Pure syntax-based metrics can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
-a learned model semantically equivalent but syntactically different to the reference model.
-
-Here we introduce an evaluation approach that is robust to role changes of this particular kind.
-The intuition of the approach is to assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
-%In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
-
-\subsection{The \strips\ edit distance}
-
-We first define the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
-\begin{itemize}
-\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_v(\xi)$.
-\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_v(\xi)$.
-\end{itemize}
-
-We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ action models are.
-The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations}, deletion and insertion, have the same positive cost.
-
-\begin{mydefinition}
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that they are {\em comparable}.
-The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
-\end{mydefinition}
-
-Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well.
-In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
-\begin{mydefinition}
-The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
-\end{mydefinition}
-
-Normally evaluating a given learned domain with respect to the actual generative model is not possible because the actual model is not available.
-As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions~\cite{yang2007learning}.
-With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
-
-\begin{mydefinition}
-  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and the observations sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$ such that each state observation in $\mathcal{O}$ is built with fluents in $F$.
-The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
-\end{mydefinition}
-
-Unlike the error function defined by {\sc ARMS}, the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
-This fact enables the recognition of \strips\ action models.
-The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
-This {\em edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
-
-The error of a learned action model could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
-This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes.
-Our definition of the edit distance is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
-Finally, the {\em edit distance} can also be defined with respect to a set of plans, if they are available.
-
-\begin{mydefinition}
-  Given an action model $\mathcal{M}$ and a set of valid plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ that only contain actions built grounding the schemes in $\mathcal{M}$.
-The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\Pi)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$}; \[\delta(\mathcal{M},\Pi)=\min_{\forall \mathcal{M}' \rightarrow \Pi} \delta(\mathcal{M},\mathcal{M}')\]
-\end{mydefinition}
-
-\subsection{Computing the observations and plans edit distance}
-Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, is {\em non-empty}.
-In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-A solution to the planning task resulting from the extended compilation is a sequence of actions that:
-
-\begin{enumerate}
-\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema.
-However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$ in observations of the plan executions}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{O}$ (and on $\Pi$ if available), as explained in Section~\ref{sec:Section5} for the models that are programmed from scratch.
-\end{enumerate}
-
-Now $\Lambda$ does not formalize a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
-The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
-
-\begin{itemize}
-\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
-\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
-In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
-\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
-The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
-The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
-\end{itemize}
-
-To illustrate this, the plan of Figure~\ref{fig:plan-odistance} solves the classical planning task that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-First, the plan edits the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:example-observations}.
-
-\begin{figure}[hbt!]
-{\tt\small
-00 : (insert\_add\_handempty\_stack)\\
-01 : (insert\_add\_clear\_stack\_var1)\\
-02 : (apply\_unstack blockB blockA)\\
-03 : (validate\_1)\\
-04 : (apply\_putdown blockB)\\
-05 : (validate\_2)\\
-06 : (apply\_pickup blockA)\\
-07 : (validate\_3)\\
-08 : (apply\_stack blockA blockB)\\
-09 : (validate\_4)\\
-}
-\caption{\small Plan for editing a given action model and validating it at the state observations shown in Figure~\ref{fig:example-observations}.
-The given action model is a four operator {\em blockswold} where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.}
-\label{fig:plan-odistance}
-\end{figure}
-
-Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
-$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-odistance}.
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
-The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
-
-When the executed plans $\Pi$ are also available, the compilation can be adapted to compute the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
-The modifications to the compilation explained in Section~\ref{sec:Section5} are also useful here to redefine the learning task as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$.
-Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-In this case the edited action model is however validated at the plan shown in Figure~\ref{fig:example-plans}.
-
-\begin{figure}[hbt!]
-{\tt\small
-00 : (insert\_add\_handempty\_stack)\\
-01 : (insert\_add\_clear\_stack\_var1)\\
-02 : (apply\_unstack blockB blockA i1 i2)\\
-03 : (apply\_putdown blockB i2 i3)\\
-04 : (apply\_pickup blockA i3 i4)\\
-05 : (apply\_stack blockA blockB i4 i5)\\
-06 : (validate\_1)
-}
- \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the plan shown in Figure~\ref{fig:example-plans}.}
-\label{fig:plan-pdistance}
-\end{figure}
-
-Last but not least, this compilation is flexible to compute the {\em edit distance} between two {\em comparable} \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$.
-A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion.
-In this case the edited model is not validated on a sequence of observations or plans but on the given action model $\mathcal{M}'$ that acts as a reference.
-The sets of fluents $F_{\Lambda}$ and $I_{\Lambda}$ are defined like in the previous compilation.
-With respect to the actions, $A_{\Lambda}$ again implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect) but does not contain apply actions because the \strips\ action model are not validated in any observation of plan executions.
-Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_v(\xi)$.
-To illustrate this, the plan of Figure~\ref{fig:plan-mdistance} solves the classical planning task that corresponds to computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.
-The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects.
-Again our interest is in the number of required {\em edit operations}, e.g.
-$\delta(\mathcal{M},\mathcal{M'})=2$.
-
-\begin{figure}[hbt!]
-{\tt\small
-00 : (insert\_add\_handempty\_stack)\\
-01 : (insert\_add\_clear\_stack\_var1)
-}
- \caption{\small Plan for computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.}
-\label{fig:plan-mdistance}
-\end{figure}
-
-\section{Recognition of \strips\ action models}
-\label{sec:Section7}
-Given a set of possible \strips\ models and set of observations of plan executions, the {\em recognition of \strips\ models} is the task of identifying which model has the highest probability of producing the given observations.
-
-According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the set of possible \strips\ action models.
-Given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section3}.
-With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the size of the possible state observations of length $n$, that is $\mathcal{O}=s_0,\ldots,s_n$ is given by $2^{n\times|F|}$.
-
-With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
-\begin{enumerate}
-\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$.
-If a set of plans $\Pi$ is available, this same strategy can be followed using the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
-\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
-\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
-\end{enumerate}
-
-\section{Learning \strips\ action models with background knowledge}
-A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn explanations from data~\cite{muggleton1994inductive}.
-Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case, given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.
-
-\subsection{State constraints}
-The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
-If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
-For instance, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
-
-{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$.
-
-The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
-For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}.
-There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.
-
-\begin{figure}[hbt!]
-  \begin{footnotesize}
-    \begin{center}
-$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
-$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
-$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
-    \end{center}
-\end{footnotesize}
- \caption{\small An example of the strongest invariant for the {\em blocksworld} domain.}
-\label{fig:strongest-invariant}
-\end{figure}
-
-A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a mutex because $block_A$ can only be on top of a single block.
-
-A {\em domain invariant} is an instance-independent invariant, i.e.
-holds for any possible initial state and set of objects.
-Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
-Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
-For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
-
-\subsection{Trajectory constraints}
-Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10,bacchus:LTL:1998}.
-LTL is a modal temporal logic interpreted over sequences of states.
-LTL interpreted over finite sequences of states (also called traces) has received attention from the planning community and provided a formal description called $LTL_f$~\cite{Giacomo:LTLf:AAAI2014}.
-
-$LTL_f$ is built up from a finite set of propositional variables $P$, the logical operators $\neg$ and $\vee$ (from which it is possible to define operators $\wedge$, $\rightarrow$ and $\leftrightarrow)$, and the temporal modal operators {\em next}($\bigcirc$), and {\em until}(${\mathcal U}$).
-An $LTL_f$ formula is then inductively defined over a set of propositions $P$:
-\begin{itemize}
-\item A proposition $p\in P$ is an $LTL_f$ formula,
-\item if $\psi$ and $\chi$ are $LTL_f$ formulae, then so they are $\neg\psi$, ($\psi\vee\chi$), $\bigcirc\psi$, ($\psi\ {\mathcal U}\ \chi$).
-\end{itemize}
-
-Given a sequence of states $\mathcal{O}=(s_0,\ldots, s_n)$, we say that a given $LTL_f$ formula $\phi$ holds at instant $i$ (denoted by $\mathcal{O}, i\models\phi$) iff:
-\begin{itemize}
-\item $\mathcal{O}, i\models f$ for a propositional variable $f\in F$, iff $f\in s_i$,
-\item $\mathcal{O}, i\models \neg\psi$ iff it is not the case that $\mathcal{O}, i\models\psi$,
-\item $\mathcal{O}, i\models \psi\wedge\chi$ iff $\mathcal{O}, i\models\psi$ and $\mathcal{O}, i\models\chi$,
-\item $\mathcal{O}, i\models \bigcirc\phi$ if $i<n$ and $\mathcal{O}, i+1\models\phi$,
-\item $\mathcal{O}, i\models (\phi_1\ {\mathcal U}\ \phi_2)$ if exists some $j\in\{1,\ldots,n\}$ such that, $\tau, j\models \phi_2$ and for all $k\in\{i,\ldots,j-1\}$ it holds that $\tau, k\models \phi_1$.
-\end{itemize}
-We say that $\mathcal{O}$ is {\em valid} for a given $LTL_f$ formula $\phi$ (denoted by $\mathcal{O}\models\phi$) iff for every {\small $0\leq i\leq n$} holds that $\mathcal{O}, i\models\phi$.
-From the previous basic operators it is also possible to define the abbreviation $last$ (that denotes the last instant of a sequence) and the temporal modal operators {\em eventually} ($\lozenge$), {\em always} ($\square$), and {\em release} (${\mathcal R}$):
-\begin{itemize}
-\item $last$ is shorthand for $\neg\bigcirc true$ and holds only at the last state of the sequence.
-The achievement of classical planning goals $G$ can then be expressed as the $LTL_f$ formula $\phi=last\wedge G$.
-\item $\lozenge$ $\psi$ stands for $(true\ {\mathcal U}\ \psi)$, and says that $\psi$ will eventually hold before the last state (last state included).
-\item $\square\psi$ stands for $\neg \lozenge \neg\psi$ and says that from the current state till the last one the formula will always hold.
-\item $\psi\ {\mathcal R}\ \chi$ stands for $\neg(\neg\psi\ {\mathcal U}\ \neg\chi)$.
-This says that $\chi$ has to be true until and including the point where $\psi$ first becomes true; if $\psi$ never becomes true, $\chi$ must remain true forever.
-\end{itemize}
-
-$LTL$ interpreted over finite traces has the expressive power of First Order Logic over finite ordered traces.
-Satisfiability, validity and logical implication (as well as model checking) for $LTL_f$ are PSPACE-complete.
-
-\subsection{Leveraging background knowledge}
-We have shown how our approach is flexible to validate the learned action models with respect to a set of observations of executed plans and with respect to a model that acts as a reference.
-Here we show that the validation of an action model can also be done with {\em state constraints} or {\em trajectory constraints}.
-Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check that the learned model satisfy every constraint $\phi_i\in\Phi$:
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
-\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
-\end{align*}
-\end{small}
-
-This redefinition of validate actions applies also to trajectory constraints because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.
-
-Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
-Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space.
-With regard to our compilation, {\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema.
-For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
-\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
-\end{align*}
-\end{small}
-
-\section{Evaluation}
-\label{sec:Section8}
-This section evaluates the performance of our approach for learning \strips\ action models starting from different amounts and kinds of available input knowledge.
-
-\subsection{Setup}
-The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
-All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM.
-\begin{itemize}
-\item {\bf Planner}.
-The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
-In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
-Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
-\item {\bf Reproducibility}.
-We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this repository {\em https://github.com/sjimenezgithub/strips-learning} so any experimental data reported in the paper is fully reproducible.
-\end{itemize}
-
-\subsection{Evaluating with a reference model}
-Here we evaluate the learned models with respect to the actual generative model.
-Opposite to what usually happens in ML, this model is available when learning is applied to IPC domains.
-The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
-These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
-Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
-\begin{itemize}
-\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
-\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing).
-\end{itemize}
-
-\subsubsection{Learning from plans}
-We start evaluating our approach with $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ learning tasks, where the action of the executed plans are available and the observation sequence contains only the corresponding initial and goal states, $s_o^t$ and $s_n^t$, for every plan $\pi_t\in\Pi$.
-We then repeat the evaluation but exploiting potential \emph{static predicates} computed from the observed states in $\mathcal{O}$, which are the predicates that appear unaltered in the states that belong to the same plan.
-Static predicates are used to constrain the space of possible action models as explained in Section~\ref{sec:Section5}.
-
-Table~\ref{tab:results_plans} shows the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), while the last two columns of each setting and the last row report averages values.
-We can observe that identifying static predicates leads to models with better precondition {\em recall}.
-This fact evidences that many of the missing preconditions corresponded to static predicates because there is no incentive to learn them as they always hold~\cite{gregory2015domain}.
-
-\begin{table*}[hbt!]
-		\resizebox{\textwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
-& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
-& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-                          \hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
-			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
-			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
-			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
-			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
-			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
-			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
-			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
-			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
-			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
-			\hline
-			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
-		\end{tabular}
-	}
-\caption{\small {\em Precision} and {\em recall} scores for learning tasks from labeled plans without (left) and with (right) static predicates.}
-\label{tab:results_plans}
-\end{table*}
-
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
-All the learning tasks are solved in a few seconds.
-Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
-When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
-
-\begin{table}
-\begin{scriptsize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c||c|c|c|}
-                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
-			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
-                         \hline
-			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
-			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
-			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
-			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
-			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
-			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
-                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
-                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
-			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
-			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
-			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
-			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
-		\end{tabular}
-	\end{center}
-        \end{scriptsize}
-	\caption{\small Total planning time, preprocessing time and plan length for learning tasks from labeled plans without/with static predicates.}
-	\label{tab:time_plans}	
-\end{table}
-
-We evaluate now the ability of our approach to support partially specified action models; that is, when the input model $\mathcal{M}$ is not empty because some preconditions and effects of the actions are initially known.
-In this particular experiment, the model of half of the actions is given in $\mathcal{M}$ as an extra input of the learning task.
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, which include the identification of static predicates.
-We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
-In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the previous learning tasks because the evaluation is done only over half of the actions.
-This occurs, for instance, in the precondition \emph{recall} of domains such as {\em Floortile}, {\em Gripper} or {\em Satellite}.
-
-Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
-The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
-Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
-Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\mathcal{M}$).
-{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
-
-\begin{table}[hbt!]
-\begin{footnotesize}
-	\begin{center}
-
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-				Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.90 \\
-				Ferry & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
-                Grid & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.84 & 0.78 \\
-				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-				Miconic & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Satellite & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
-				Transport & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
-				Zenotravel & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.78 \\
-				\hline
-				\bf  & 0.98 & 0.71 & 1.0 & 0.98 & 1.0 & 0.95 & 0.98 & 0.87 \\
-			\end{tabular}
-
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} scores for learning tasks with partially specified action models.}
-\label{tab:results_plans_partial}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-                         \hline
-			Blocks & 0.07 & 0.01 & 54  \\
-			Driverlog & 0.03 & 0.01 & 40 \\
-			Ferry & 0.06 & 0.03 & 45 \\
-			Floortile & 0.43 & 0.42 & 55 \\
-                        Grid & 3.12 & 3.07 & 53 \\
-			Gripper & 0.03 & 0.01 & 35 \\
-			Miconic & 0.03 & 0.01 & 34  \\
-			Satellite & 0.14 & 0.14 & 47 \\
-			Transport & 0.23 & 0.21 & 37 \\
-			Zenotravel & 0.90 & 0.89 & 40 \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Time and plan length learning for learning tasks with partially specified action models.}
-	\label{tab:time_plans_partial}	
-\end{table}
-
-\subsubsection{Learning from state observations}
-Here we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, where the action of the executed plans are not available but the initial and goal states are known.
-When input plans are not available, the planner must not only compute the action models but also the plans that satisfy the input observations.
-Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the results obtained for this using static predicates and partially specified models.
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} was not able to solve the corresponding planning tasks within a 1000 sec.
-time bound.
-The values of \emph{precision} and \emph{recall} are significantly lower than in Table ~\ref{tab:results_plans}.
-Given that the learning hypothesis space is now fairly under-constrained, actions can be reformulated and still be compliant with the inputs (e.g.
-the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
-We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) and yet the results are below the scores obtained when learning from labeled plans.
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-            Blocks & 0.33 & 0.33 & 0.75 & 0.50 & 0.33 & 0.33 & 0.47 & 0.39 \\
-            Driverlog & 1.0 & 0.29 & 0.33 & 0.67 & 1.0 & 0.50 & 0.78 & 0.48 \\
-            Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
-            Floortile & 0.67 & 0.40 & 0.50 & 0.40 & 1.0 & 0.40 & 0.72 & 0.40 \\
-            Grid & - & - & - & - & - & - & - & - \\
-            Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-            Miconic & 0.0 & 0.0 & 0.33 & 0.50 & 0.0 & 0.0 & 0.11 & 0.17 \\
-            Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
-            Transport & 0.0 & 0.0 & 0.25 & 0.5 & 0.0 & 0.0 & 0.08 & 0.17 \\
-            Zenotravel & - & - & - & - & - & - & - & - \\
-            \hline
-            & 0.63 & 0.29 & 0.54 & 0.70 & 0.67 & 0.53 & 0.61 & 0.51 \\			
-			\end{tabular}
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} scores for learning from (initial,final) state pairs.}
-\label{tab:results_states}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-			\hline
-            Blocks & 2.14 & 0.00 & 58  \\
-            Driverlog & 0.09 & 0.00 & 88 \\
-            Ferry & 0.17 & 0.01 & 65 \\
-            Floortile & 6.42 & 0.15 & 126 \\
-            Grid & - & - & - \\
-            Gripper & 0.03 & 0.00 & 47 \\
-            Miconic & 0.04 & 0.00 & 68 \\
-            Satellite & 4.34 & 0.10 & 126 \\
-            Transport & 2.57 & 0.21 & 47 \\			
-            Zenotravel & - & - & - \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Time and plan length when learning from (initial,final) state pairs.}
-	\label{tab:time_states}	
-\end{table}
-
-Now we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, where the action of the executed plans are not available but where the observation sequence $\mathcal{O}$ contains all the intermediate states, not just the initial and final states.
-Table~\ref{fig:observationsnomap} shows the precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
-The reason why the scores in Table ~\ref{fig:observationsnomap} are still low, despite more state observations are available, is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but correspond to {\em reformulations} of the actual model (changes in the roles of actions with matching headers or parameters with matching types).
-
-\begin{table}[hbt!]
-	\begin{center}
-		\begin{scriptsize}
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-				\hline
-
-				blocks & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 \\
-				driverlog & 0.0 & 0.0 & 0.25 & 0.43 & 0.0 & 0.0 & 0.08 & 0.14 \\
-				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
-				floortile & 0.38 & 0.55 & 0.4 & 0.18 & 0.56 & 0.45 & 0.44 & 0.39 \\
-				grid & 0.5 & 0.47 & 0.33 & 0.29 & 0.25 & 0.29 & 0.36 & 0.35 \\
-				gripper & 0.83 & 0.83 & 0.75 & 0.75 & 0.75 & 0.75 & 0.78 & 0.78 \\
-				hanoi & 0.5 & 0.25 & 0.5 & 0.5 & 0.0 & 0.0 & 0.33 & 0.25 \\
-				hiking & 0.43 & 0.43 & 0.5 & 0.35 & 0.44 & 0.47 & 0.46 & 0.42 \\
-				miconic & 0.6 & 0.33 & 0.33 & 0.25 & 0.33 & 0.33 & 0.42 & 0.31 \\
-				npuzzle & 0.33 & 0.33 & 0.0 & 0.0 & 0.0 & 0.0 & 0.11 & 0.11 \\
-				parking & 0.25 & 0.21 & 0.0 & 0.0 & 0.0 & 0.0 & 0.08 & 0.07 \\
-				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
-				transport & 1.0 & 0.3 & 0.8 & 0.8 & 1.0 & 0.6 & 0.93 & 0.57 \\
-				visitall & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
-				zenotravel & 0.67 & 0.29 & 0.33 & 0.29 & 0.33 & 0.14 & 0.44 & 0.24
-			\end{tabular}
-		\end{scriptsize}
-	\end{center}
-	\caption{\small Precision and recall values obtained without computing the $f_{P\&R}$ mapping with the reference model.}
-	\label{fig:observationsnomap}
-\end{table}
-
-To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model {\em reformulations}.
-Precision and recall are often combined using the {\em harmonic mean}.
-This expression, called the {\em F-measure} or the balanced {\em F-score}, is defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
-Given the learned action model $\mathcal{M}$ and the reference action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure} (considering swaps in the actions with matching headers or parameters with matching types).
-
-Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
-The {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
-These results evidence that in all of the evaluated domains, except for {\em ferry} and {\em satellite}, the learning task swaps the roles of some actions (or parameters) with respect to their role in the reference model.
-
-\begin{table}
-        \begin{center}
-
-		\begin{scriptsize}
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-				\hline
-
-				blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				driverlog & 0.67 & 0.14 & 0.33 & 0.57 & 0.67 & 0.29 & 0.56 & 0.33 \\
-				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
-				floortile & 0.44 & 0.64 & 1.0 & 0.45 & 0.89 & 0.73 & 0.78 & 0.61 \\
-				grid & 0.63 & 0.59 & 0.67 & 0.57 & 0.63 & 0.71 & 0.64 & 0.62 \\
-				gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-				hiking & 0.78 & 0.6 & 0.93 & 0.82 & 0.88 & 0.88 & 0.87 & 0.77 \\
-				miconic & 0.8 & 0.44 & 1.0 & 0.75 & 1.0 & 1.0 & 0.93 & 0.73 \\
-				npuzzle & 0.67 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 0.89 \\
-				parking & 0.56 & 0.36 & 0.5 & 0.33 & 0.5 & 0.33 & 0.52 & 0.34 \\
-				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
-				transport & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0 & 0.63 \\
-				visitall & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 \\
-				zenotravel & 1.0 & 0.43 & 0.67 & 0.57 & 1.0 & 0.43 & 0.89 & 0.48
-			\end{tabular}
-		\end{scriptsize}
-	\end{center}
-	\caption{\small Precision and recall values obtained when computing the $f_{P\&R}$ mapping with the reference model.}
-	\label{fig:observationsmap}
-\end{table}
-
-\subsection{Evaluating with a test set}
-
-When a reference model is not available, the learned models are tested with an observation set.
-Table~\ref{fig:observationstest} summarizes the results obtained when evaluating the quality of the learned models with respect to a test set of state observations.
-Each test set comprises between 20 and 50 observations per domain and is generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
-The table shows, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
-The reported results show that, despite learning only from 25 state observations, 12 out of 15 learned domains yield ratios of $90\%$ or above.
-This fact evidences that the learned models require very small amounts of edition to match the observations of the given test set.
-
-\begin{table}[hbt!]
-		\begin{center}
-                \begin{footnotesize}
-			\begin{tabular}{l|r|r|c|}
-				& $\delta(\mathcal{M},\mathcal{O})$ & $\delta(\mathcal{M},*)$ & $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$ \\
-				\hline
-				blocks & 0 & 90 & 1.0 \\
-				driverlog & 5 & 144 & 0.97 \\
-				ferry & 2 & 69 & 0.97 \\
-				floortile & 34 & 342 & 0.90 \\
-				grid & 42 & 153 & 0.73 \\
-				gripper & 2 & 30 & 0.93 \\
-				hanoi & 1 & 63 & 0.98 \\
-				hiking & 69 & 174 & 0.60 \\
-				miconic & 3 & 72 & 0.96 \\
-				npuzzle & 2 & 24 & 0.92 \\
-                                parking & 4 & 111 & 0.96 \\
-				satellite & 24 & 75 & 0.68 \\
-				transport & 4 & 78 & 0.95 \\
-				visitall & 2 & 24 & 0.92 \\
-				zenotravel & 3 & 63 & 0.95
-			\end{tabular}
-                        	\end{footnotesize}
-		\end{center}
-	\caption{\small Evaluation of the quality of the learned models with respect to an observations test set.}
-	\label{fig:observationstest}
-\end{table}
-
-The learning scores of several domains in Table~\ref{fig:observationsmap} are above the ones reported in Table~\ref{fig:observationsnomap}.
-The reason lies in the particular observations comprised by the test sets.
-As an example, in the {\em driverlog} domain, the action schema {\small \tt disembark-truck} is missing from the learned model because this action is never induced from the observations in the training set; that is, such action never appears in the corresponding \emph{unobserved} plan.
-The same happens with the {\small \tt paint-down} action of the {\em floortile} domain or {\small \tt move-curb-to-curb} in the {\em parking} domain.
-Interestingly, these actions do not appear either in the test sets and so the learned action models are not penalized in Table~\ref{fig:observationstest}.
-Generating {\em informative} and {\em representative} observations for learning planning action models is an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.
+\input{background}
+\input{learning_task}
+\input{compilation}
+\input{evaluation}
+\input{experiments}
 

### 2018-04-18 doc/aij18/background.tex
new paper organization

--- 2018-04-16 doc/aij18/related_work.tex
@@
-\section{Related work}
-\label{sec:Section2}
+\section{Background}
+\label{sec:background}
+
+This section serves two purposes.
+In subsection \ref{basic_planning} we introduce the basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper.
+Section \ref{related_work} summarizes existing approaches to learning action models in classical planning and highlights the novelty of our contribution via the related work.
+
+%This section defines the planning model used on this work, {\em classical planning with conditional effects}, and the target of the learning/evaluation/recognition tasks addressed in the paper, a \strips\ action model.
+
+\subsection{Basic planning concepts}
+\label{basic_planning}
+
+We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
+A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
+either~$l=f$ or $l=\neg f$.
+A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent).
+We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
+
+A {\em state} $s$ is a full assignment of values to fluents, i.e.
+$|s|=|F|$, so the size of the state space is $2^{|F|}$.
+Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
+
+A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. An action $a\in A$ is defined with:
+
+%{\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$ and {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$.
+
+\begin{itemize}
+\item $\pre(a)\in\mathcal{L}(F)$, the {\em preconditions} of $a$, is the set of literals that must hold for the action $a\in A$ to be applicable.
+\item $\eff^+(a)\in\mathcal{L}(F)$, the {\em positive effects} of $a$, is the set of literals that true after the application of the action $a\in A$.
+\item $\eff^-(a)\in\mathcal{L}(F)$, the {\em negative effects} of $a$, is the set of literals that are false after the application of the action.
+\end{itemize}
+
+We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
+The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+The {\em plan length} is denoted with $|\pi|=n$ .
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
+
+In this work, we will use the term \emph{observation} to refer to the input knowledge of a learning task.
+Hence, $\tup{s_0, s_1, \ldots, s_n}$ is a full observable (FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,a)=s_{i+1}$.
+Otherwise, $\tup{s_0, s_1, \ldots, s_n}$ is a partial observable (PO) state trajectory, meaning that at least one $s_i$ is a partial assignment of values to fluents in which one or more literals are missing.
+Likewise, $\tup{a_1, \ldots, a_n}$ is a FO action sequence if it contains all the necessary actions to achieve a goal from an initial state, and a PO action sequence if at least one necessary action is missing in the sequence.
+
+A \emph{plan trace} is a combination of a state and an action sequence $\tau = \langle s_0, a_1, a_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
+Additionally, $\tau$ is a plan trace with PO states if $\tup{s_0, s_1, \ldots, s_n}$ is a PO state sequence and a plan trace with PO actions if $\tup{a_1, \ldots, a_n}$ is a PO action sequence.
+
+\subsection{Related work}
+\label{related_work}
 
@@
 Due to the large variety of types of input knowledge and the lack of a common naming convention (e.g., plan examples, plan traces, training data, plan samples), we firstly present a glossary of terms that will ease the classification of the surveyed models accordingly to their input knowledge.
-
-We will refer to a state sequence as $\Sigma = \langle s_0, s_1, \ldots, s_n \rangle$ and an action sequence as $\Omega = \langle a_1, a_2, \ldots, a_n \rangle$.
-In a full observable (FO) state sequence $\Sigma$, every state $s_i \in \Sigma$ is a complete state and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$  is given by $\Omega = \langle a_{i+1} \rangle$.
-On the other hand, $\Omega$ is a FO action sequence if there are no missing actions in $\Omega$.
-
-A plan trace is a combination of a state and an action sequence $\tau = \langle s_0, a_1, a_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
-Additionally, $\tau$ is a plan trace with partial observable (PO) states if there is at least one missing literal in any state $s_i$.
-And $\tau$ is a plan trace with PO actions of there is at least one missing action from the action sequence.
 
@@
 
+The Crowdsourced Action-Model Acquisition (\CAMA) for planning explores knowledge from both crowdsourcing (human annotators) and plan traces to learn action models for planning.
+\CAMA relies on the assumption that obtaining enough training samples is often difficult and costly because there is usually a limited number of plan traces available.
+In order to overcome this limitation, \CAMA builds on a set of soft constraints based on labels \texttt{true} or \texttt{false} given by the crowd and a set of soft constraints based on the input plan traces, solves then using a MAX-SAT solver and converts the solution to action models.
+
+\CAMA uses plan traces with PO states.
+To capture the knowledge from the crowd, \CAMA enumerates all possible preconditions and effects for each action, considering all predicates whose parameters are included by the parameters of the action.
+Hence, if the parameters of a predicate $p$ are included by the parameters of $a$, \CAMA queries the crowd for three questions of the form "Is the fact $p$ a precondition/positive effect/negative effect of action $a$?", and human annotators reply "yes", "no" or "cannot tell".
+The goal is to find an optimal estimator of the \texttt{true} labels given the observation, minimizing the average bit-wise error rate.
+
+================================================
+
 Back in the 90's various systems aimed learning operators mostly via interaction with the environment.

### 2018-04-18 doc/aij18/compilation.tex
new paper organization

\section{Compilation scheme}
\label{compilation}

AQUI ES DONDE TENEMOS QUE HABLAR DE LOS CONDITIONAL EFFECTS QUE SOLO AFECTAN A LA COMPILACION

An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$. Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}. An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
\[
triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
\]

The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.

====================

Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects. A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plans that explains the given inputs to the learning task. The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:

\begin{enumerate}
\item {\bf Programs the action model $\mathcal{M}'$}. A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
\item {\bf Validates the action model $\mathcal{M}'$}. The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) with the programmed action model $\mathcal{M}'$.
\end{enumerate}

\subsection{Learning from state observations}
Here we formalize the compilation for learning \strips\ action models with classical planning. Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
\begin{itemize}
\item $F_{\Lambda}$ contains:
\begin{itemize}
\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the input observations, i.e. the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-observations}. Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model. If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$. For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation in $\mathcal{O}$ where the action model is validated.
\end{itemize}
\item $I_{\Lambda}$ encodes the first observation, $s_0\subseteq F$, and sets $mode_{prog}$ to true. Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect. Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.

\item $G_{\Lambda}=\bigcup_{1\leq i\leq |\mathcal{O}|}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
\item $A_{\Lambda}$ comprises three kinds of actions:
\begin{enumerate}
\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$:
\begin{itemize}
\item Actions for {\bf removing} a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.

\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}, pre_{f}(\xi)\},\\
\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
\end{align*}
\end{small}

\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.

\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
\end{align*}
\end{small}
\end{itemize}

\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$. Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position. Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$ from {\em blocksworld}.
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\cup \{\neg mode_{val}\},\\
\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
&\{\emptyset\}\rhd\{mode_{val}\}.
\end{align*}
\end{small}

\begin{figure}[hbt!]
\begin{scriptsize}
\begin{verbatim}
(:action apply_stack
  :parameters (?o1 - object ?o2 - object)
  :precondition
   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
        (or (not (pre_clear_stack_v1)) (clear ?o1))
        (or (not (pre_clear_stack_v2)) (clear ?o2))
        (or (not (pre_holding_stack_v1)) (holding ?o1))
        (or (not (pre_holding_stack_v2)) (holding ?o2))
        (or (not (pre_handempty_stack)) (handempty)))
  :effect
   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
        (when (del_ontable_stack_v1) (not (ontable ?o1)))
        (when (del_ontable_stack_v2) (not (ontable ?o2)))
        (when (del_clear_stack_v1) (not (clear ?o1)))
        (when (del_clear_stack_v2) (not (clear ?o2)))
        (when (del_holding_stack_v1) (not (holding ?o1)))
        (when (del_holding_stack_v2) (not (holding ?o2)))
        (when (del_handempty_stack) (not (handempty)))
        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
        (when (add_ontable_stack_v1) (ontable ?o1))
        (when (add_ontable_stack_v2) (ontable ?o2))
        (when (add_clear_stack_v1) (clear ?o1))
        (when (add_clear_stack_v2) (clear ?o2))
        (when (add_holding_stack_v1) (holding ?o1))
        (when (add_holding_stack_v2) (holding ?o2))
        (when (add_handempty_stack) (handempty))
        (when (modeProg) (not (modeProg)))))
\end{verbatim}
\end{scriptsize}
 \caption{\small PDDL action for applying an already programmed schema $stack$ (implications are coded as disjunctions).}
\label{fig:compilation}
\end{figure}

\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{O}$, {\tt\small $1\leq i< |\mathcal{O}|$}.
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\cup \{mode_{val}\},\\
\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
\end{align*}
\end{small}
\end{enumerate}
\end{itemize}

The compilation approach is flexible to various amount and kind of input data. If any reference to the $mode_{val}$ fluent is removed the compilation can learn \strips\ action models when there are missing state observations in $\mathcal{O}$. On the other hand, if the $mode_{val}$ fluent is ignored, $P_{\Lambda}$ becomes harder since the classical planner must determine how many {\em apply} actions are necessary between any two observations, i.e. between the application of two {\em validate} actions.  Further, known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$. In this case, the corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved. When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the compilation validates whether the observation of the plan execution follows the given model:
\begin{itemize}
\item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given input data if a solution plan for $P_{\Lambda}$ can be found.
\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given input data if $P_{\Lambda}$ is unsolvable. This means that $\mathcal{M}$ cannot be compliant with the given observation of the plan execution.
\end{itemize}
This feature of our compilation is beyond the functionality of VAL, the plan validation tool~\cite{howey2004val}, because VAL requires (1) a full plan and (2), a full action model for plan validation.

Figure~\ref{fig:plan-observations} illustrate how our compilation works and shows a plan that solves the classical planning task that results from our compilation. This plan programs and validates the $stack$ schema (from {\em blocksworld}) using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$. Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,21]$ validate the programmed operators using the sequence of five state observations shown in the Figure~\ref{fig:example-observations}.

\begin{figure}[hbt!]
{\footnotesize\tt
     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
     01 : (program\_pre\_handempty\_stack)\\
     02 : (program\_pre\_holding\_stack\_v2)\\
     03 : (program\_pre\_on\_stack\_v1\_v1)\\
     04 : (program\_pre\_on\_stack\_v1\_v2)\\
     05 : (program\_pre\_on\_stack\_v2\_v1)\\
     06 : (program\_pre\_on\_stack\_v2\_v2)\\
     07 : (program\_pre\_ontable\_stack\_v1)\\
     08 : (program\_pre\_ontable\_stack\_v2)\\
     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
    10 : (program\_eff\_clear\_stack\_v2)\\
    11 : (program\_eff\_handempty\_stack)\\
    12 : (program\_eff\_holding\_stack\_v1)\\
    13 : (program\_eff\_on\_stack\_v1\_v2)\\
    {\bf 14} : (apply\_unstack blockB blockA)\\
    15 : (validate\_1)\\
    16 : (apply\_putdown blockB)\\
    17 : (validate\_2)\\
    18 : (apply\_pickup blockA)\\
    19 : (validate\_3)\\
    20 : (apply\_stack blockA blockB)\\
    21 : (validate\_4)
}
 \caption{\small Plan for programming and validating the $stack$ schema using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
\label{fig:plan-observations}
\end{figure}

\subsection{Learning from plans}
Now we extend the compilation to consider observed actions. Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda}}$ such that:
\begin{itemize}
\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with the $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that corresponds to action $a_j$. The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to iterate through the steps of plan $\pi$.
\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<|\pi|$}, for tracking the plan step where the action model is validated. Goals are as in the original compilation. In other words, the observed states are used for validation adding a $test_t$, ${\small 1\leq t\leq |\mathcal{O}|}$ fluent for each observation in $\mathcal{O}$.
\item With respect to the set of actions $A_{\Lambda'}$.
\begin{enumerate}
\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ and the actions for {\em validating} the programmed action model in a given state observation are the same as in the previous compilation.
\item The actions for {\em applying} an already programmed operator includes the following extra conditional effects $\{at_{j},plan(name(a_j),\Omega^{ar(a_j)},j)\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step. This mechanism ensures that $\mathsf{apply_{\xi,\omega}}$ actions are applied, exclusively, in the same order as in the example plan $\pi$ while it is flexible to the learning scenario where some $a_j$ actions in $\pi$ are missing. In such scenario $P_{\Lambda'}$ becomes harder because the planner must determine which and how many $\mathsf{apply_{\xi,\omega}}$ actions are necessary for validating the learned action model with the given input knowledge.
\end{enumerate}
\end{itemize}

The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$. This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $\mathcal{O}=\tup{s_0,s_4}$ shown in Figure~\ref{fig:example-plans}. Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the four-action plan $\pi$ shown in the Figure~\ref{fig:example-plans}.

\begin{figure}[hbt!]
{\footnotesize\tt
     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
     01 : (program\_pre\_handempty\_stack)\\
     02 : (program\_pre\_holding\_stack\_v2)\\
     03 : (program\_pre\_on\_stack\_v1\_v1)\\
     04 : (program\_pre\_on\_stack\_v1\_v2)\\
     05 : (program\_pre\_on\_stack\_v2\_v1)\\
     06 : (program\_pre\_on\_stack\_v2\_v2)\\
     07 : (program\_pre\_ontable\_stack\_v1)\\
     08 : (program\_pre\_ontable\_stack\_v2)\\
     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
    10 : (program\_eff\_clear\_stack\_v2)\\
    11 : (program\_eff\_handempty\_stack)\\
    12 : (program\_eff\_holding\_stack\_v1)\\
    13 : (program\_eff\_on\_stack\_v1\_v2)\\
    {\bf 14} : (apply\_unstack blockB blockA i1 i2)\\
    15 : (apply\_putdown blockB i2 i3)\\
    16 : (apply\_pickup blockA i3 i4)\\
    17 : (apply\_stack blockA blockB i4 i5)\\
    {\bf 18} : (validate\_1)
}
 \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi$ and state observations $\mathcal{O}$ (shown in Figure~\ref{fig:example-plans}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
\label{fig:plan-lplan}
\end{figure}

Now we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}. Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e. same fluents and actions but different initial states and goals). The set of actions, $A=\emptyset$, is empty because the action model is initially unknown. Finally, the initial state $I_t$ is given by the state $s_0^t$ and the plan $\pi_t$, and the goals $G_t$ are defined by the state $s_n^t$. Addressing the learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ with classical planning requires introducing a small modification to our compilation. In particular, the actions in $P_{\Lambda'}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq \tau$} reset the current state, and the current plan, and are now defined as:
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{1\leq j<t}\cup\{\neg test_j\}_{t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\} \cup \{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t}.
\end{align*}
\end{small}

\subsection{Compilation properties}

\begin{mylemma}
Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$.
\end{mylemma}

\begin{proof}[Proof sketch]
\begin{small}
Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent. In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state. These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from the corresponding initial state and following the sequence of actions defined by the plans in $\Pi$. This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
\end{small}
\end{proof}

\begin{mylemma}
Completeness. Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
\end{mylemma}

\begin{proof}[Proof sketch]
\begin{small}
By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$. The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O},\Pi$.
\end{small}
\end{proof}

The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
\begin{itemize}
\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task. The larger these numbers, the larger the size of the $F_v(\xi)$ sets. This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
\item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
\end{itemize}

\subsection{Exploiting static predicates to optimize the compilation}
A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}. Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete. Given a static predicate $p$:
\begin{itemize}
\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable objects} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
\item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
\end{itemize}

Static predicates can also constrain the space of possible preconditions by looking at the given set of state observation $\mathcal{O}$. One can assume that if a precondition $f\in F_v$ (s.t. $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\mathcal{O}$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$. For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{O}$.

Furthermore looking as well at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t. $f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$ and $\mathcal{O}$. Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.

\subsection{Learning \strips\ action models with background knowledge}
A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn explanations from data~\cite{muggleton1994inductive}. Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case, given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.

\subsubsection{State constraints}
The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.  If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states. For instance, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.

{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}. Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$.

The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$. For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}. There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.

\begin{figure}[hbt!]
  \begin{footnotesize}
    \begin{center}
$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
    \end{center}
\end{footnotesize}
 \caption{\small An example of the strongest invariant for the {\em blocksworld} domain.}
\label{fig:strongest-invariant}
\end{figure}

A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}. For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a mutex because $block_A$ can only be on top of a single block.

A {\em domain invariant} is an instance-independent invariant, i.e. holds for any possible initial state and set of objects. Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state. Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}. For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.

\subsubsection{Trajectory constraints}
Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10,bacchus:LTL:1998}. LTL is a modal temporal logic interpreted over sequences of states. LTL interpreted over finite sequences of states (also called traces) has received attention from the planning community and provided a formal description called $LTL_f$~\cite{Giacomo:LTLf:AAAI2014}.

$LTL_f$ is built up from a finite set of propositional variables $P$, the logical operators $\neg$ and $\vee$ (from which it is possible to define operators $\wedge$, $\rightarrow$ and $\leftrightarrow)$, and the temporal modal operators {\em next}($\bigcirc$), and {\em until}(${\mathcal U}$). An $LTL_f$ formula is then inductively defined over a set of propositions $P$:
\begin{itemize}
\item A proposition $p\in P$ is an $LTL_f$ formula,
\item if $\psi$ and $\chi$ are $LTL_f$ formulae, then so they are $\neg\psi$, ($\psi\vee\chi$), $\bigcirc\psi$, ($\psi\ {\mathcal U}\ \chi$).
\end{itemize}

Given a sequence of states $\mathcal{O}=(s_0,\ldots, s_n)$, we say that a given $LTL_f$ formula $\phi$ holds at instant $i$ (denoted by $\mathcal{O}, i\models\phi$) iff:
\begin{itemize}
\item $\mathcal{O}, i\models f$ for a propositional variable $f\in F$, iff $f\in s_i$,
\item $\mathcal{O}, i\models \neg\psi$ iff it is not the case that $\mathcal{O}, i\models\psi$,
\item $\mathcal{O}, i\models \psi\wedge\chi$ iff $\mathcal{O}, i\models\psi$ and $\mathcal{O}, i\models\chi$,
\item $\mathcal{O}, i\models \bigcirc\phi$ if $i<n$ and $\mathcal{O}, i+1\models\phi$,
\item $\mathcal{O}, i\models (\phi_1\ {\mathcal U}\ \phi_2)$ if exists some $j\in\{1,\ldots,n\}$ such that, $\tau, j\models \phi_2$ and for all $k\in\{i,\ldots,j-1\}$ it holds that $\tau, k\models \phi_1$.
\end{itemize}
We say that $\mathcal{O}$ is {\em valid} for a given $LTL_f$ formula $\phi$ (denoted by $\mathcal{O}\models\phi$) iff for every {\small $0\leq i\leq n$} holds that $\mathcal{O}, i\models\phi$. From the previous basic operators it is also possible to define the abbreviation $last$ (that denotes the last instant of a sequence) and the temporal modal operators {\em eventually} ($\lozenge$), {\em always} ($\square$), and {\em release} (${\mathcal R}$):
\begin{itemize}
\item $last$ is shorthand for $\neg\bigcirc true$ and holds only at the last state of the sequence. The achievement of classical planning goals $G$ can then be expressed as the $LTL_f$ formula $\phi=last\wedge G$.
\item $\lozenge$ $\psi$ stands for $(true\ {\mathcal U}\ \psi)$, and says that $\psi$ will eventually hold before the last state (last state included).
\item $\square\psi$ stands for $\neg \lozenge \neg\psi$ and says that from the current state till the last one the formula will always hold.
\item $\psi\ {\mathcal R}\ \chi$ stands for $\neg(\neg\psi\ {\mathcal U}\ \neg\chi)$. This says that $\chi$ has to be true until and including the point where $\psi$ first becomes true; if $\psi$ never becomes true, $\chi$ must remain true forever.
\end{itemize}

$LTL$ interpreted over finite traces has the expressive power of First Order Logic over finite ordered traces. Satisfiability, validity and logical implication (as well as model checking) for $LTL_f$ are PSPACE-complete.

\subsubsection{Leveraging background knowledge}
We have shown how our approach is flexible to validate the learned action models with respect to a set of observations of executed plans and with respect to a model that acts as a reference. Here we show that the validation of an action model can also be done with {\em state constraints} or {\em trajectory constraints}. Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check that the learned model satisfy every constraint $\phi_i\in\Phi$:
\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
\end{align*}
\end{small}

This redefinition of validate actions applies also to trajectory constraints because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.

Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan. Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space. With regard to our compilation, {\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema. For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:

\begin{small}
\begin{align*}
\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
\end{align*}
\end{small}

### 2018-04-18 doc/aij18/evaluation.tex
new paper organization

\section{Evaluation model}
\label{evaluation}

The roles of two action schemes whose headers match or the roles of two action parameters that belong to the same type can be swapped by a $\Lambda$ learning task. For instance, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped. Pure syntax-based metrics can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e. a learned model semantically equivalent but syntactically different to the reference model.

Here we introduce an evaluation approach that is robust to role changes of this particular kind. The intuition of the approach is to assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations. %In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.

\subsection{The \strips\ edit distance}

We first define the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
\begin{itemize}
\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_v(\xi)$.
\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_v(\xi)$.
\end{itemize}

We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ action models are. The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations}, deletion and insertion, have the same positive cost.

\begin{mydefinition}
  Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that they are {\em comparable}. The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
\end{mydefinition}

Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well. In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
\begin{mydefinition}
The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
\end{mydefinition}

Normally evaluating a given learned domain with respect to the actual generative model is not possible because the actual model is not available. As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions~\cite{yang2007learning}. With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.

\begin{mydefinition}
  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and the observations sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$ such that each state observation in $\mathcal{O}$ is built with fluents in $F$. The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
\end{mydefinition}

Unlike the error function defined by {\sc ARMS}, the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model. This fact enables the recognition of \strips\ action models. The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods. This {\em edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.

The error of a learned action model could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model. This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes. Our definition of the edit distance is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations. Finally, the {\em edit distance} can also be defined with respect to a set of plans, if they are available.

\begin{mydefinition}
  Given an action model $\mathcal{M}$ and a set of valid plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ that only contain actions built grounding the schemes in $\mathcal{M}$. The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\Pi)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$}; \[\delta(\mathcal{M},\Pi)=\min_{\forall \mathcal{M}' \rightarrow \Pi} \delta(\mathcal{M},\mathcal{M}')\]
\end{mydefinition}

\subsection{Computing the observations and plans edit distance}
Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, is {\em non-empty}. In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets. A solution to the planning task resulting from the extended compilation is a sequence of actions that:

\begin{enumerate}
\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}. A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}. In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema. However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}.
\item {\bf Validates the edited model $\mathcal{M}'$ in observations of the plan executions}. The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{O}$ (and on $\Pi$ if available), as explained in Section~\ref{sec:Section5} for the models that are programmed from scratch.
\end{enumerate}

Now $\Lambda$ does not formalize a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$. The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:

\begin{itemize}
\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true. In addition, the input action model $\mathcal{M}$ is now encoded in the initial state. This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$. The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation. The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e. include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
\end{itemize}

To illustrate this, the plan of Figure~\ref{fig:plan-odistance} solves the classical planning task that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing. First, the plan edits the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:example-observations}.

\begin{figure}[hbt!]
{\tt\small
00 : (insert\_add\_handempty\_stack)\\
01 : (insert\_add\_clear\_stack\_var1)\\
02 : (apply\_unstack blockB blockA)\\
03 : (validate\_1)\\
04 : (apply\_putdown blockB)\\
05 : (validate\_2)\\
06 : (apply\_pickup blockA)\\
07 : (validate\_3)\\
08 : (apply\_stack blockA blockB)\\
09 : (validate\_4)\\
}
\caption{\small Plan for editing a given action model and validating it at the state observations shown in Figure~\ref{fig:example-observations}. The given action model is a four operator {\em blockswold} where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.}
\label{fig:plan-odistance}
\end{figure}

Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g. $\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-odistance}. In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}). The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.

When the executed plans $\Pi$ are also available, the compilation can be adapted to compute the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$. The modifications to the compilation explained in Section~\ref{sec:Section5} are also useful here to redefine the learning task as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$. Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing. In this case the edited action model is however validated at the plan shown in Figure~\ref{fig:example-plans}.

\begin{figure}[hbt!]
{\tt\small
00 : (insert\_add\_handempty\_stack)\\
01 : (insert\_add\_clear\_stack\_var1)\\
02 : (apply\_unstack blockB blockA i1 i2)\\
03 : (apply\_putdown blockB i2 i3)\\
04 : (apply\_pickup blockA i3 i4)\\
05 : (apply\_stack blockA blockB i4 i5)\\
06 : (validate\_1)
}
 \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the plan shown in Figure~\ref{fig:example-plans}.}
\label{fig:plan-pdistance}
\end{figure}

Last but not least, this compilation is flexible to compute the {\em edit distance} between two {\em comparable} \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$. A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion. In this case the edited model is not validated on a sequence of observations or plans but on the given action model $\mathcal{M}'$ that acts as a reference. The sets of fluents $F_{\Lambda}$ and $I_{\Lambda}$ are defined like in the previous compilation. With respect to the actions, $A_{\Lambda}$ again implement the two {\em edit operations} (i.e. include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect) but does not contain apply actions because the \strips\ action model are not validated in any observation of plan executions. Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_v(\xi)$. To illustrate this, the plan of Figure~\ref{fig:plan-mdistance} solves the classical planning task that corresponds to computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model. The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects. Again our interest is in the number of required {\em edit operations}, e.g. $\delta(\mathcal{M},\mathcal{M'})=2$.

\begin{figure}[hbt!]
{\tt\small
00 : (insert\_add\_handempty\_stack)\\
01 : (insert\_add\_clear\_stack\_var1)
}
 \caption{\small Plan for computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.}
\label{fig:plan-mdistance}
\end{figure}

\subsection{Recognition of \strips\ action models}
\label{sec:Section7}
Given a set of possible \strips\ models and set of observations of plan executions, the {\em recognition of \strips\ models} is the task of identifying which model has the highest probability of producing the given observations.

According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$. In our scenario, the hypotheses are about the set of possible \strips\ action models. Given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section3}. With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the size of the possible state observations of length $n$, that is $\mathcal{O}=s_0,\ldots,s_n$ is given by $2^{n\times|F|}$.

With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
\begin{enumerate}
\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$. If a set of plans $\Pi$ is available, this same strategy can be followed using the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
\end{enumerate}

### 2018-04-18 doc/aij18/experiments.tex
new paper organization

--- 2018-02-23 doc/icaps-18/camera-ready/strips-learning.tex
@@
-\def\year{2017}\relax
-%File: formatting-instruction.tex
-\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
-\usepackage{aaai18}  %Required
-\usepackage{times}  %Required
-\usepackage{helvet}  %Required
-\usepackage{courier}  %Required
-\usepackage{url}  %Required
-\usepackage{graphicx}  %Required
-\frenchspacing  %Required
-\setlength{\pdfpagewidth}{8.5in}  %Required
-\setlength{\pdfpageheight}{11in}  %Required
-%PDF Info Is Required:
-  \pdfinfo{}
-\setcounter{secnumdepth}{0}
+\section{Experimental results}
+\label{experiments}
 
-\usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{multirow}
-\usepackage{tikz}
-\usepackage{comment}
-
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{listings}
-\usepackage{multicol}
-\usepackage{arydshln}
-\usetikzlibrary{calc,backgrounds,positioning,fit}
-
-\newcommand{\tup}[1]{{\langle #1 \rangle}}
-
-\newcommand{\pre}{\mathsf{pre}}     % precondition
-\newcommand{\del}{\mathsf{del}}     % effect
-\newcommand{\add}{\mathsf{add}}     % effect
-\newcommand{\eff}{\mathsf{eff}}     % effect
-\newcommand{\cond}{\mathsf{cond}}   % conditional effect
-\newcommand{\true}{\mathsf{true}}   % true
-\newcommand{\false}{\mathsf{false}} % false
-\newcommand{\PE}{\mathrm{PE}}     % precondition
-\newcommand{\strips}{\textsc{Strips}}     % precondition
-
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{definition}[theorem]{Definition}
-
-\begin{document}
-
-\title{Learning \strips\ Action Models with Classical Planning}
-\author{\#39}
-
-% Commented for blind submission
-\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
-{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-{\small Universitat Polit\`ecnica de Val\`encia.}\\
-{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
-{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
-
-\maketitle
-\begin{abstract}
-This paper presents a novel approach for learning \strips\ action models from examples that compiles this inductive learning task into a classical planning task.
-Interestingly, the compilation approach is flexible to different amounts of available input knowledge; the learning examples can range from a set of plans (with their corresponding initial and final states) to just a pair of initial and final states (no intermediate action or state is given).
-Moreover, the compilation accepts partially specified action models and it can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
-\end{abstract}
-
-\section{Introduction}
-Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-In both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
-Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
-
-On the other hand, Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML to learning \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
+\subsection{Setup}
+The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
+All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM.
 \begin{itemize}
-\item The {\em input} to ML algorithms (the learning/training data) is usually a finite set of vectors that represent the value of some fixed object features.
-The input for learning planning action models is, however, observations of plan executions (where each plan possibly has a different length).
-\item The {\em output} of ML algorithms is usually a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning action models the output is, for each action, the preconditions, negative and positive effects that define the possible state transitions.
+\item {\bf Planner}.
+The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
+We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
+In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
+Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
+\item {\bf Reproducibility}.
+We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this repository {\em https://github.com/sjimenezgithub/strips-learning} so any experimental data reported in the paper is fully reproducible.
 \end{itemize}
 
-Learning \strips\ action models is a well-studied problem with sophisticated algorithms such as {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}, which do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative planning compilation approach for learning \strips\ action models.
-The compilation approach is appealing by itself because it opens up the door to the bootstrapping of planning action models, but also because:
-
-\begin{enumerate}
-\item It is flexible to various amounts of input knowledge.
-Learning examples range from a set of plans (with their corresponding initial and final states) to just a pair of initial and final states where no intermediate state or action is observed.
-\item It accepts previous knowledge about the structure of the actions in the form of partially specified action models.
-In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
-\end{enumerate}
-
-The second section of the paper formalizes the classical planning model, its extension to {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the output of the addressed learning task).
-The third section formalizes the task of learning action models with different amounts of available input knowledge.
-The fourth and fifth sections describe our compilation approach to tackle the formalized learning tasks.
-Finally, the last sections show the experimental evaluation, discuss the strengths and weaknesses of the compilation approach and propose several opportunities for future research.
-
-\section{Background}
-This section defines the planning model and the output of the learning tasks addressed in the paper.
-
-\subsection{Classical planning with conditional effects}
-Our approach to learning \strips\ action models is compiling this learning task into a classical planning task with conditional effects.
-Conditional effects allow us to compactly define actions whose effects depend on the current state.
-Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
-
-We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
-A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e.
-either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values).
-We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.
-
-A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often we will abuse of notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.
-
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$.
-We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
-The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
-
-An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)$ and a set of {\em conditional effects} $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals: $C \subseteq \mathcal{L}(F)$, the {\em condition}, and $E \subseteq \mathcal{L}(F)$, the {\em effect}.
-An action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
-\[
-triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
-\]
-
-The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
-
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The {\em plan length} is denoted with $|\pi|=n$ .
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the application of the plan $\pi$ in the initial state $I$.
-
-\subsection{\strips\ action schemas and {\em variable name} objects}
-Our approach is aimed at learning PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-Figure~\ref{fig:stack} shows the {\em stack} schema of a four-operator {\em blocksworld} domain~\cite{slaney2001blocks} encoded in PDDL.
-
-\begin{figure}
-\begin{footnotesize}
-\begin{verbatim}
-(:action stack
-  :parameters (?v1 ?v2 - object)
-  :precondition (and (holding ?v1) (clear ?v2))
-  :effect (and (not (holding ?v1))
-               (not (clear ?v2))
-               (handempty) (clear ?v1)
-               (on ?v1 ?v2)))
-\end{verbatim}
-\end{footnotesize}
- \caption{\small The {\em stack} operator schema of the {\em blocksworld} domain specified in PDDL.}
-\label{fig:stack}
-\end{figure}
-
-To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$, where $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects denoted as {\em variable names} ($\Omega\cap\Omega_v=\emptyset$).
-$\Omega_v$ is bound to the maximum arity of an action in a given planning frame.
-For instance, in a three-block \emph{blocksworld}, $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have two parameters each.
-
-Let $F_v$ be a new set of fluents, $F\cap F_v=\emptyset$, that results from instantiating the predicates in $\Psi$ using exclusively  objects of $\Omega_v$.
-$F_v$ defines the elements of the preconditions and effects of an action schema.
-For instance, in the \emph{blocksworld} domain, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-
-Finally, we assume that an action $a\in A$ is instantiated from a \strips\ operator schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
-
+\subsection{Evaluating with a reference model}
+Here we evaluate the learned models with respect to the actual generative model.
+Opposite to what usually happens in ML, this model is available when learning is applied to IPC domains.
+The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
+These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
+Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$ is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
-For instance, the headers of a four-operator \emph{blocksworld} domain are: {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
-\item $pre(\xi)\subseteq F_v$ is the set of preconditions,  $del(\xi)\subseteq F_v$ the negative effects and  $add(\xi)\subseteq F_v$ the positive effects such that $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
+\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing).
 \end{itemize}
 
-\section{Learning \strips\ action models}
-
-Learning \strips\ action models from fully available input knowledge, i.e.
-from plans where the {\em pre-} and {\em post-states} of every action in the plans are known, is straightforward.
-When intermediate states are available, operator schemas are derived lifting the literals that change between the pre and post-state of each action execution.
-Preconditions of an action are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action~\cite{jimenez2012review}.
-
-This section formalizes more challenging learning tasks, where less input knowledge is available:
-
-\subsubsection{Learning from (initial, final) state pairs.} This learning task amounts to observing agents acting in the world but watching only the result of their plans execution.
-No intermediate information about the actions in the plans is given.
-This learning task is formalized as $\Lambda=\tup{\Psi,\Sigma}$:
-\begin{itemize}
-\item $\Psi$ is the set of predicates that define the abstract state space of a given planning domain.
-\item $\Sigma=\{\sigma_1,\ldots,\sigma_{\tau}\}$ is a set of $(initial, final)$ state pairs called {\em labels}.
-Each label $\sigma_t=(s_0^t,s_{n}^t)$, {\tt\small $1\leq t\leq \tau$}, comprises the {\em final} state $s_{n}^t$ resulting from executing an unknown plan $\pi_t=\tup{a_1^t, \ldots, a_n^t}$ in the {\em initial} state $s_0^t$.
-\end{itemize}
-
-\subsubsection{Learning from labeled plans.}
-We augment the input knowledge with the actions executed by the observed agent and define the learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$:
-
-\begin{itemize}
-\item $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is a given set of example plans where $\pi_t=\tup{a_1^t, \ldots, a_n^t}$, {\small $1\leq t\leq \tau$}, is an action sequence that induces the corresponding state sequence $\tup{s_0^t, s_1^t, \ldots, s_n^t}$ such that $a_i^t$, {\small $1\leq i\leq n$}, is applicable in $s_{i-1}^t$ and generates $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
-\end{itemize}
-
-Figure~\ref{fig:lexample} shows an example of a learning task $\Lambda'$ of the {\em blocksworld} domain.
-This task has a single learning example, $\Pi=\{\pi_1\}$ and $\Sigma=\{\sigma_1\}$, that corresponds to observing the execution of an eight-action plan $(|\pi_1|=8)$ for inverting a four-block tower.
-
-\subsubsection{Learning from partially specified action models.}
-In case that partially specified operator schemas are available, we can incorporate this information within the learning task.
-The new leaning task is defined as $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$:
-
-\begin{itemize}
-\item $\Xi_0$ is a partially specified action model in which some preconditions and effects are known a priori.
-\end{itemize}
-
-\vspace{0.1cm}
-
-A solution to $\Lambda$ is a set of operator schemas $\Xi$ that is compliant with the predicates in $\Psi$ and the set of initial and final states $\Sigma$.
-In a $\Lambda$ learning scenario, a solution must not only determine a possible \strips\ action model but also the plans $\pi_t$, {\tt\small $1\leq t\leq \tau$}, that explain the given labels $\Sigma$ using the learned model.
-A solution to $\Lambda'$ is a set of \strips\ operator schemas $\Xi$ (one schema $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ for each action that has a different name in the example plans $\Pi$) that is compliant with the predicates $\Psi$, the example plans $\Pi$, and their corresponding labels $\Sigma$.
-Finally, a solution to $\Lambda''$ is a set of \strips\ operator schemas $\Xi$ that is also compliant with the provided partially specified action model $\Xi_0$.
-
-\begin{figure}
-{\tt ;;; Predicates in $\Psi$}
-\begin{footnotesize}
-\begin{verbatim}
-(handempty) (holding ?o  - object)
-(clear ?o - object) (ontable ?o - object)
-(on ?o1 - object ?o2 - object)
-\end{verbatim}
-\end{footnotesize}
-
-\vspace{0.2cm}
-
-\begin{subfigure}{.25\textwidth}
-{\tt ;;; Plan $\pi_1$}
-\begin{footnotesize}
-\begin{verbatim}
-0: (unstack A B)
-1: (putdown A)
-2: (unstack B C)
-3: (stack B A)
-4: (unstack C D)
-5: (stack C B)
-6: (pickup D)
-7: (stack D C)
-\end{verbatim}
-\end{footnotesize}
-\end{subfigure}%
-\begin{subfigure}{.6\textwidth}
-{\tt ;;; Label $\sigma_1=(s_0^1,s_{n}^1)$}
-\begin{lstlisting}[mathescape]
-\end{lstlisting}
-\vspace{0.1cm}
-\begin{tikzpicture}[node distance = 0mm, block/.style args = {#1,#2}{fill=#1,text width=#2,shape=square}]
-\node (initD) [draw]{D};
-\node (initC) [draw, above=of initD.north]{C};
-\node (initB) [draw, above=of initC.north]{B};
-\node (initA) [draw, above=of initB.north]{A};
-\draw[thick] (-1,-0.25) -- (2.5,-0.25);
-
-\node (goalA) [draw, right=10mm of initD]{A};
-\node (goalB) [draw, right=10mm of initC]{B};
-\node (goalC) [draw, right=10mm of initB]{C};
-\node (goalD) [draw, right=10mm of initA]{D};
-\end{tikzpicture}
-\vspace{0.6cm}
-\end{subfigure}%
- \caption{\small Learning task of the \emph{blocksworld} domain from a single labeled plan.}
-\label{fig:lexample}
-\end{figure}
-
-\section{Learning \strips\ action models with planning}
-In our approach, a learning task $\Lambda$, $\Lambda'$ or $\Lambda''$ is solved by compiling it into a classical planning task with conditional effects.
-The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
-
-\begin{enumerate}
-\item {\em Programs the \strips\ action model $\Xi$}.
-A solution plan has a {\em prefix} that, for each $\xi\in\Xi$, determines the fluents from $F_v$ that belong to $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
-\item {\em Validates the programmed \strips\ action model $\Xi$ in the given input knowledge} (the labels $\Sigma$ and $\Pi$, and/or $\Xi_0$ if available).
-For every label $\sigma_t\in \Sigma$, a solution plan has a postfix that produces a final state $s_{n}^t$ using the programmed action model $\Xi$ in the corresponding initial state $s_0^t$.
-This process is the validation of the programmed action model $\Xi$ with the set of learning examples {\small $1\leq t\leq \tau$}.
-%If information about the corresponding plan $\pi_t\in \Pi$ is available, then it is also used in the validation.
-\end{enumerate}
-
-To formalize our compilation we first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
-same fluents and actions but different initial states and goals).
-Fluents $F$ are built instantiating the predicates in $\Psi$ with the objects of the input labels $\Sigma$.
-Formally, $\Omega=\bigcup_{\small 1\leq t\leq \tau} obj(s_0^t)$, where $obj$ is a function that returns the objects that appear in a fully specified state.
-The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t\in \sigma_t$, and the goals $G_t$ are defined by the state $s_n^t\in \sigma_t$.
-
-We can now formalize the compilation approach.
-We start with $\Lambda$ as it requires the least input knowledge.
-Given a learning task $\Lambda=\tup{\Psi,\Sigma}$, the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
-
-\begin{itemize}
-\item $F_{\Lambda}$ extends $F$ with:
-\begin{itemize}
-\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v$ and $\xi \in \Xi$ that represent the programmed action model.
-If a fluent of $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative effect/positive effect of the operator schema $\xi\in \Xi$.
-For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$}.
-\item A fluent $mode_{prog}$ to indicate whether the operator schemas are being programmed or validated (when already programmed)
-\item Fluents $\{test_t\}_{1\leq t\leq \tau}$ which represent the examples where the action model will be validated.
-\end{itemize}
-
-\item $I_{\Lambda}$ contains the fluents from $F$ that encode $s_0^1$ (the initial state of the first label), the fluents in every $pre_f(\xi)\in F_{\Lambda}$ and the fluent $mode_{prog}$ set to true.
-Our compilation assumes that any operator schema is initially programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
-\item $G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$ indicates that the programmed action model is validated in all the learning examples.
-\item $A_{\Lambda}$ contains actions of three kinds:
-\begin{enumerate}
-\item Actions for {\em programming} an operator schema $\xi\in\Xi$:
-\begin{itemize}
-\item Actions for {\bf removing} a {\em precondition} $f\in F_v$ from $\xi$.
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
-& mode_{prog}, pre_{f}(\xi)\},\\
-\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
-\end{align*}
-\end{small}
-
-\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v$ to $\xi$.
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi),\\
-& mode_{prog}\},\\
-\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
-&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
-\end{align*}
-\end{small}
-\end{itemize}
-
-\item Actions for {\em applying} an already programmed operator schema $\xi\in\Xi$ bound to the objects $\omega\subseteq\Omega^{ar(\xi)}$.
-We assume operators headers are known so the binding of $\xi$ is done implicitly by order of appearance of the action parameters, i.e.
-variables $pars(\xi)$ are bound to the objects in $\omega$ that appear in the same position.
-Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
-\end{align*}
-\end{small}
-
-\item Actions for {\em validating} the learning example {\tt\small $1\leq t\leq \tau$}.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{1\leq j<t}\\
-&\cup\{\neg test_j\}_{t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\} \cup \{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\\
-&\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t}.
-\end{align*}
-\end{small}
-\end{enumerate}
-\end{itemize}
-
-\begin{lemma}
-A classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\Xi$ that solves the learning task $\Lambda$.
-\end{lemma}
-
-\begin{proof}[Proof sketch]
-\begin{small}
-Once operator schemas $\Xi$ are programmed, they can only be applied and validated according to the $mode_{prog}$ fluent.
-To solve $P_{\Lambda}$, goals $\{test_t\}$, {\small $1\leq t\leq \tau$} can only be achieved by executing an applicable sequence of programmed operator schemas that reaches the final state $s_n^t$, defined in $\sigma_t$, starting from $s_0^t$.
-If this is achieved for all the input examples {\small $1\leq t\leq \tau$}, it means that the programmed action model $\Xi$ is compliant with the provided input knowledge and hence it is a solution to $\Lambda$.
-\end{small}
-\end{proof}
-
-The compilation is {\em complete} in the sense that it does not discard any possible \strips\ action model. The size of the classical planning task $P_{\Lambda}$ depends on:
-\begin{itemize}
-\item The arity of the actions headers in $\Xi$ and the predicates $\Psi$ of the learning task.
-The larger the arity, the larger the $F_v$ set, which in turn defines the size of the fluent sets $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ and the corresponding set of {\em programming} actions.
-\item The number of learning examples. The larger this number, the more $test_t$ fluents and $\mathsf{validate_{t}}$ actions in $P_{\Lambda}$.
-\end{itemize}
-
-\begin{figure}
-\begin{scriptsize}
-\begin{verbatim}
-(:action apply_stack
-  :parameters (?o1 - object ?o2 - object)
-  :precondition
-   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
-        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
-        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
-        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
-        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
-        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
-        (or (not (pre_clear_stack_v1)) (clear ?o1))
-        (or (not (pre_clear_stack_v2)) (clear ?o2))
-        (or (not (pre_holding_stack_v1)) (holding ?o1))
-        (or (not (pre_holding_stack_v2)) (holding ?o2))
-        (or (not (pre_handempty_stack)) (handempty)))
-  :effect
-   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
-        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
-        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
-        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
-        (when (del_ontable_stack_v1) (not (ontable ?o1)))
-        (when (del_ontable_stack_v2) (not (ontable ?o2)))
-        (when (del_clear_stack_v1) (not (clear ?o1)))
-        (when (del_clear_stack_v2) (not (clear ?o2)))
-        (when (del_holding_stack_v1) (not (holding ?o1)))
-        (when (del_holding_stack_v2) (not (holding ?o2)))
-        (when (del_handempty_stack) (not (handempty)))
-        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
-        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
-        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
-        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
-        (when (add_ontable_stack_v1) (ontable ?o1))
-        (when (add_ontable_stack_v2) (ontable ?o2))
-        (when (add_clear_stack_v1) (clear ?o1))
-        (when (add_clear_stack_v2) (clear ?o2))
-        (when (add_holding_stack_v1) (holding ?o1))
-        (when (add_holding_stack_v2) (holding ?o2))
-        (when (add_handempty_stack) (handempty))
-        (when (modeProg) (not (modeProg)))))
-\end{verbatim}
-\end{scriptsize}
- \caption{\small PDDL action for applying an already programmed schema $stack$ (implications coded as disjunctions).}
-\label{fig:compilation}
-\end{figure}
-
-\section{Constraining the learning hypothesis space with additional input knowledge}
-\label{sec:Constraining}
-In this section, we show that further input knowledge can be used to constrain the space of possible action models and to make the learning task more practicable.
-
-\subsection{Labeled plans}
-We extend the compilation to consider labeled plans.
-Given a learning task $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda'}}$:
-\begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with $F_{\Pi}=\{plan(name(\xi),\Omega^{ar(\xi)},j)\}$, the fluents to code the steps of the plans in $\Pi$, where $F_{\pi_t}\subseteq F_{\Pi}$ encodes $\pi_t\in \Pi$.
-Fluents $at_j$ and $next_{j,j+1}$, {\small $1\leq j< n$}, are also added to represent the current plan step and to iterate through the steps of a plan.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi_1}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<n$}, to indicate the plan step where the action model is validated.
-As in the original compilation, $G_{\Lambda'}=G_{\Lambda}=\bigcup_{1\leq t\leq \tau}\{test_t\}$.
-\item With respect to $A_{\Lambda'}$.
-\begin{enumerate}
-\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ are the same.
-\item The actions for {\em applying} an already programmed operator have an extra precondition $f\in F_{\Pi}$ that encodes the current plan step, and extra conditional effects $\{at_{j}\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-With this mechanism we ensure that these actions are applied in the same order as in the example plans.
-\item The actions for {\em validating} the current learning example have an extra precondition, $at_{|\pi_t|}$, to indicate that the current plan $\pi_t$ is fully executed and extra conditional effects to remove plan $\pi_{t}$ and initiate the next plan $\pi_{t+1}$:
-\begin{small}
-\begin{align*}
-&\{\emptyset\}\rhd\{\neg at_{|\pi_t|},at_1\} \cup \{\neg f\}_{f\in F_{\pi_t}} \cup \{f\}_{f\in F_{\pi_t+1}}.
-\end{align*}
-\end{small}
-\end{enumerate}
-\end{itemize}
-
-\subsection{Partially specified action models}
-The known preconditions and effects of a partially specified action model are encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true in the initial state $I_{\Lambda'}$.
-The programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become now unnecessary and they are removed from $A_{\Lambda'}$, thus making the planning task $P_{\Lambda'}$ be easier to solve.
-
-To illustrate this, the plan of Figure~\ref{fig:plan} is a solution to a learning task $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$ for acquiring the {\em blocksworld} action model where operator schemas for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\Xi_0$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld} using the plan $\pi_1$ and label $\sigma_1$ shown in Figure~\ref{fig:lexample}.
-Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,22]$ validate the programmed operators following the plan $\pi_1$ shown in Figure~\ref{fig:lexample}.
-
-In the extreme, when a fully specified \strips\ action model is given in $\Xi_0$, the compilation validates whether an observed plan follows the given model.
-In this case, if a solution plan is found for $P_{\Lambda'}$, it means that the given action model is {\em valid} for the provided examples.
-If $P_{\Lambda'}$ is unsolvable then it means that the action model is invalid because it is not compliant with all the given examples.
-Tools for plan validation like VAL~\cite{howey2004val} could also be used at this point.
-
-\begin{figure}
-{\footnotesize\tt
-     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
-     01 : (program\_pre\_handempty\_stack)\\
-     02 : (program\_pre\_holding\_stack\_v2)\\
-     03 : (program\_pre\_on\_stack\_v1\_v1)\\
-     04 : (program\_pre\_on\_stack\_v1\_v2)\\
-     05 : (program\_pre\_on\_stack\_v2\_v1)\\
-     06 : (program\_pre\_on\_stack\_v2\_v2)\\
-     07 : (program\_pre\_ontable\_stack\_v1)\\
-     08 : (program\_pre\_ontable\_stack\_v2)\\
-     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
-    10 : (program\_eff\_clear\_stack\_v2)\\
-    11 : (program\_eff\_handempty\_stack)\\
-    12 : (program\_eff\_holding\_stack\_v1)\\
-    13 : (program\_eff\_on\_stack\_v1\_v2)\\
-    {\bf 14} : (apply\_unstack a b i1 i2)\\
-    15 : (apply\_putdown a i2 i3)\\
-    16 : (apply\_unstack b c i3 i4)\\
-    17 : (apply\_stack b a i4 i5)\\
-    18 : (apply\_unstack c d i5 i6)\\
-    19 : (apply\_stack c b i6 i7)\\
-    20 : (apply\_pickup d i7 i8)\\
-    21 : (apply\_stack d c i8 i9)\\
-    {\bf 22} : (validate\_1)
-}
- \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi_1$ and label $\sigma_1$ (shown in Figure~\ref{fig:lexample}) as well as previously specified operator schemas for $pickup$, $putdown$ and $unstack$.}
-\label{fig:plan}
-\end{figure}
-
-\subsection{Static predicates}
-A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
-Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
-Given a static predicate $p$:
-\begin{itemize}
-\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable names} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
-\item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
-\end{itemize}
-
-Static predicates can also constrain the space of possible preconditions by looking at the given set of labels $\Sigma$.
-One can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the labels in $\Sigma$ then fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\Xi$.
-For instance, in the {\em zenotravel} domain, $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold in any state of $\Sigma$.
-
-On the other hand, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$.
-Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding label $\sigma_t\in\Sigma$ contains the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)}, the only possible precondition that would include the static predicate is $pre\_next\_fly\_v5\_v4$.
-
-%In the evaluation of our approach we do not assume that the set of static predicates is given but compute a set of {\em potential} static predicates from the input to the learning task.
-
-\section{Evaluation}
-This section evaluates the performance of our approach for learning \strips\ action models with different amounts of available input knowledge.
-
-\subsubsection{Setup.}
-The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only used 5 learning examples for each domain and we fixed the examples for all the experiments so that we can evaluate the impact of the input knowledge in the quality of the learned models.
-All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM.
-
-The classical planner we used to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We used {\sc Madagascar} due to its ability to deal with planning instances populated with dead-ends.
-In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
-Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
-
-For the sake of reproducibility, the compilation source code, evaluation scripts and used benchmarks are fully available at the repository {\em https://github.com/anonsub/strips-learning}.
-
-\subsubsection{Metrics.}
-The quality of the learned models is measured with the {\em precision} and {\em recall} metrics.
-These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
-
-Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
-Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
-Recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
-
-Given the syntax-based nature of these metrics, it may happen that they report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
-a learned model semantically equivalent but syntactically different to the reference model.
-This mainly occurs when the learning task is under-constrained.
-
-\subsection{Learning from labeled plans}
-We start evaluating our approach with tasks $\Lambda'=\tup{\Psi,\Sigma,\Pi}$, where {\em labeled plans} are available.
-We then repeat the evaluation but exploiting potential \emph{static predicates} computed from $\Sigma$, which are the predicates that appear  unaltered in the initial and final states in every $\sigma_t\in\Sigma$.
-Static predicates are used to constrain the space of possible action models as explained in the previous section.
+\subsubsection{Learning from plans}
+We start evaluating our approach with $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ learning tasks, where the action of the executed plans are available and the observation sequence contains only the corresponding initial and goal states, $s_o^t$ and $s_n^t$, for every plan $\pi_t\in\Pi$.
+We then repeat the evaluation but exploiting potential \emph{static predicates} computed from the observed states in $\mathcal{O}$, which are the predicates that appear unaltered in the states that belong to the same plan.
+Static predicates are used to constrain the space of possible action models as explained in Section~\ref{sec:Section5}.
 
@@
 
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
-All the learning tasks are solved in a few seconds.
-Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
-When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
-
-\begin{table*}
+\begin{table*}[hbt!]
 		\resizebox{\textwidth}{!}{%
@@
 \end{table*}
+
+Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
+All the learning tasks are solved in a few seconds.
+Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
+In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
+When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
 
@@
 
-\subsection{Learning from partially specified action models}
-
-We evaluate now the ability of our approach to support partially specified action models; that is, addressing learning tasks of the kind $\Lambda''=\tup{\Psi,\Sigma,\Pi,\Xi_0}$.
-In this experiment, the model of half of the actions is given in $\Xi_0$ as an extra input of the learning task.
-
+We evaluate now the ability of our approach to support partially specified action models; that is, when the input model $\mathcal{M}$ is not empty because some preconditions and effects of the actions are initially known.
+In this particular experiment, the model of half of the actions is given in $\mathcal{M}$ as an extra input of the learning task.
 Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, which include the identification of static predicates.
 We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
-In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the corresponding $\Lambda'$ tasks because the evaluation is done only over half of the actions.
+In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the previous learning tasks because the evaluation is done only over half of the actions.
 This occurs, for instance, in the precondition \emph{recall} of domains such as {\em Floortile}, {\em Gripper} or {\em Satellite}.
@@
 Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
-Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\Xi_0$).
+Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\mathcal{M}$).
 {\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
 
-%In particular, preconditions are again the component more difficult to be fully learned, with an overall recall of $0.71$.
-
-\subsection{Learning from (initial,final) state pairs}
-Finally, we evaluate our approach when input plans are not available and thereby the planner must not only compute the action models but also the plans that satisfy the input labels.
-Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the results obtained for the task $\Lambda=\tup{\Psi,\Sigma,\Xi_0}$ using static predicates.
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} was not able to solve the corresponding planning tasks within a 1000 sec.
-time bound.
-The values of \emph{precision} and \emph{recall} are significantly lower than in Table ~\ref{tab:results_plans}.
-Given that the learning hypothesis space is now fairly under-constrained, actions can be reformulated and still be compliant with the inputs (e.g.
-the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
-We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) and yet the results are below the scores obtained when learning from labeled plans.
-
-\section{Related work}
-
-%Back in the 90's various systems aimed learning operators mostly via interaction with the environment.
-{\sc LIVE} captured and formulated observable features of objects and used them to acquire and refine operators \cite{ShenS89}.
-{\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
-These early works were based on lifting the observed states supported by exploratory plans or external teachers.%, but none provided a theoretical justification for this second source of knowledge.
-
-%More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
-
-%Whilst the aforementioned works deal with full state observability,
-Action model learning has also been studied in domains where there is partial or missing state observability.
-{\sf ARMS} works when no partial intermediate state is given.
-It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
-In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
-In contrast to our model comparison validation which aims at covering all the training examples, ARMS maximizes the number of covered examples from a testing set.
-
-{\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
-Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
-The learning algorithm updates the formula of the belief state with every action and observation in the sequence and thus the final returned formula includes all consistent models.
-SLAF assesses the quality of the learned models with respect to the actual generative model.
-
-%The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
-This update makes sure that the new formula represents all the transition relations consistent with the actions and observations.
-The formula returned at the end includes all consistent models, which can then be retrieved with additional processing.
-
-%Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
-An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
-
-{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring,cresswell2011generalised}.
-The lack of available information is overcome by exploiting assumptions about the kind of domain model it has to generate.
-Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine.
-{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain})  incorporates static predicates and applies a post-processing step after the {\sf LOCM} analysis that requires a set of optimal plans to be used in the learning phase.
-This is done to mitigate the limitation of {\sf LOCM} of inducing similar models for domains with similar structures.
-{\sf LOP} compares the learned models with the corresponding reference model.
-
-Compiling an action model learning task into classical planning is a general and flexible approach that allows to accommodate various amounts and kinds of input knowledge and opens up a path for addressing further learning and validation tasks.
-For instance, the example plans in $\Pi$ can be replaced or complemented by a set $\mathcal{O}$ of sequences of observations (i.e., fully or partial state observations with noisy or missing fluents~\cite{SohrabiRU16}), and learning tasks of the kind $\Lambda=\tup{\Psi,\Sigma,\mathcal{O},\Xi_0}$ would also be attainable.
-Furthermore, our approach seems extensible to learning other types of generative models (e.g.
-hierarchical models like HTN or behaviour trees) that can be more appealing than \strips\ models since they require less search effort to compute a a planning solution.
-
-%Furthermore, the compilation scheme can be extended to learn other type of generative models that comprise partially-specified solution plans like HTN models or full-specified solutions like FSM or behaviour trees.
-
-\begin{table}
+\begin{table}[hbt!]
 \begin{footnotesize}
 	\begin{center}
-		\resizebox{\columnwidth}{!}{%
+
 		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
@@
 			\end{tabular}
-		}
+
 	\end{center}
@@
 
+\subsubsection{Learning from state observations}
+Here we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, where the action of the executed plans are not available but the initial and goal states are known.
+When input plans are not available, the planner must not only compute the action models but also the plans that satisfy the input observations.
+Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the results obtained for this using static predicates and partially specified models.
+Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} was not able to solve the corresponding planning tasks within a 1000 sec.
+time bound.
+The values of \emph{precision} and \emph{recall} are significantly lower than in Table ~\ref{tab:results_plans}.
+Given that the learning hypothesis space is now fairly under-constrained, actions can be reformulated and still be compliant with the inputs (e.g.
+the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
+We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) and yet the results are below the scores obtained when learning from labeled plans.
+
 \begin{table}
@@
 	\begin{center}
-		\resizebox{\columnwidth}{!}{%
 		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
@@
 			\end{tabular}
-		}
 	\end{center}
@@
 
-\section{Conclusions}
-We presented a novel approach for learning \strips\ action models from examples using classical planning.
-The approach is flexible to various amounts of input knowledge and accepts partially specified action models.
-We also introduced the {\em precision} and {\em recall} metrics, widely used in ML, for evaluating the learned action models.
-These two metrics measure the soundness and completeness of the learned models and facilitate the identification of model flaws.
+Now we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, where the action of the executed plans are not available but where the observation sequence $\mathcal{O}$ contains all the intermediate states, not just the initial and final states.
+Table~\ref{fig:observationsnomap} shows the precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
+The reason why the scores in Table ~\ref{fig:observationsnomap} are still low, despite more state observations are available, is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but correspond to {\em reformulations} of the actual model (changes in the roles of actions with matching headers or parameters with matching types).
 
-To the best of our knowledge, this is the first work on learning action models that is exhaustively evaluated over a wide range of domains and uses exclusively an {\em off-the-shelf} classical planner.
-The work in~\cite{SternJ17} proposes a planning compilation for learning action models from plan traces following the {\em finite domain} representation for the state variables.
-This is a theoretical study on the boundaries of the learned models and no experimental results are reported.
+\begin{table}[hbt!]
+	\begin{center}
+		\begin{scriptsize}
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+				\hline
 
-When example plans are available, we can compute accurate action models from small sets of learning examples (five examples per domain) in little computation time (less than a second).
-When action plans are not available, our approach still produces action models that are compliant with the input information.
-In this case, since learning is not constrained by actions, operators can be reformulated changing their semantics, in which case the comparison with a reference model turns out to be tricky.
+				blocks & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 \\
+				driverlog & 0.0 & 0.0 & 0.25 & 0.43 & 0.0 & 0.0 & 0.08 & 0.14 \\
+				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
+				floortile & 0.38 & 0.55 & 0.4 & 0.18 & 0.56 & 0.45 & 0.44 & 0.39 \\
+				grid & 0.5 & 0.47 & 0.33 & 0.29 & 0.25 & 0.29 & 0.36 & 0.35 \\
+				gripper & 0.83 & 0.83 & 0.75 & 0.75 & 0.75 & 0.75 & 0.78 & 0.78 \\
+				hanoi & 0.5 & 0.25 & 0.5 & 0.5 & 0.0 & 0.0 & 0.33 & 0.25 \\
+				hiking & 0.43 & 0.43 & 0.5 & 0.35 & 0.44 & 0.47 & 0.46 & 0.42 \\
+				miconic & 0.6 & 0.33 & 0.33 & 0.25 & 0.33 & 0.33 & 0.42 & 0.31 \\
+				npuzzle & 0.33 & 0.33 & 0.0 & 0.0 & 0.0 & 0.0 & 0.11 & 0.11 \\
+				parking & 0.25 & 0.21 & 0.0 & 0.0 & 0.0 & 0.0 & 0.08 & 0.07 \\
+				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
+				transport & 1.0 & 0.3 & 0.8 & 0.8 & 1.0 & 0.6 & 0.93 & 0.57 \\
+				visitall & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
+				zenotravel & 0.67 & 0.29 & 0.33 & 0.29 & 0.33 & 0.14 & 0.44 & 0.24
+			\end{tabular}
+		\end{scriptsize}
+	\end{center}
+	\caption{\small Precision and recall values obtained without computing the $f_{P\&R}$ mapping with the reference model.}
+	\label{fig:observationsnomap}
+\end{table}
 
-%An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
+To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model {\em reformulations}.
+Precision and recall are often combined using the {\em harmonic mean}.
+This expression, called the {\em F-measure} or the balanced {\em F-score}, is defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
+Given the learned action model $\mathcal{M}$ and the reference action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure} (considering swaps in the actions with matching headers or parameters with matching types).
 
-% The empirical results show that since our approach is strongly based on inference can generate non-trivial models from very small data sets.
-In addition, the SAT-based planner {\sc Madagascar} is particularly suitable for the approach because its ability to deal with planning instances populated with dead-ends and because many actions for programming the \strips\ model can be done in parallel since they do not interact reducing significantly the planning horizon.
+Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
+The {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
+These results evidence that in all of the evaluated domains, except for {\em ferry} and {\em satellite}, the learning task swaps the roles of some actions (or parameters) with respect to their role in the reference model.
 
-Generating {\em informative} examples for learning planning action models is still an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions which have low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques that enable to autonomously collect informative learning examples.
-The combination of such exploration techniques with our learning approach is an appealing research direction that opens up the door to the bootstrapping of planning action models.
+\begin{table}
+        \begin{center}
 
-In many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
-We plan to extend our approach for learning from state observations as it broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}.
+		\begin{scriptsize}
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+				\hline
 
-% Commented for blind submission
-\begin{small}
-\subsection*{Acknowledgments}
-This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.
-Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it RYC15/18009}, both programs funded by the Spanish government.
-\end{small}
+				blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				driverlog & 0.67 & 0.14 & 0.33 & 0.57 & 0.67 & 0.29 & 0.56 & 0.33 \\
+				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
+				floortile & 0.44 & 0.64 & 1.0 & 0.45 & 0.89 & 0.73 & 0.78 & 0.61 \\
+				grid & 0.63 & 0.59 & 0.67 & 0.57 & 0.63 & 0.71 & 0.64 & 0.62 \\
+				gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
+				hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
+				hiking & 0.78 & 0.6 & 0.93 & 0.82 & 0.88 & 0.88 & 0.87 & 0.77 \\
+				miconic & 0.8 & 0.44 & 1.0 & 0.75 & 1.0 & 1.0 & 0.93 & 0.73 \\
+				npuzzle & 0.67 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 0.89 \\
+				parking & 0.56 & 0.36 & 0.5 & 0.33 & 0.5 & 0.33 & 0.52 & 0.34 \\
+				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
+				transport & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0 & 0.63 \\
+				visitall & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 \\
+				zenotravel & 1.0 & 0.43 & 0.67 & 0.57 & 1.0 & 0.43 & 0.89 & 0.48
+			\end{tabular}
+		\end{scriptsize}
+	\end{center}
+	\caption{\small Precision and recall values obtained when computing the $f_{P\&R}$ mapping with the reference model.}
+	\label{fig:observationsmap}
+\end{table}
 
-\bibliographystyle{aaai}
-\bibliography{planlearnbibliography}
-\end{document}
+\subsection{Evaluating with a test set}
+
+When a reference model is not available, the learned models are tested with an observation set.
+Table~\ref{fig:observationstest} summarizes the results obtained when evaluating the quality of the learned models with respect to a test set of state observations.
+Each test set comprises between 20 and 50 observations per domain and is generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
+The table shows, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
+The reported results show that, despite learning only from 25 state observations, 12 out of 15 learned domains yield ratios of $90\%$ or above.
+This fact evidences that the learned models require very small amounts of edition to match the observations of the given test set.
+
+\begin{table}[hbt!]
+		\begin{center}
+                \begin{footnotesize}
+			\begin{tabular}{l|r|r|c|}
+				& $\delta(\mathcal{M},\mathcal{O})$ & $\delta(\mathcal{M},*)$ & $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$ \\
+				\hline
+				blocks & 0 & 90 & 1.0 \\
+				driverlog & 5 & 144 & 0.97 \\
+				ferry & 2 & 69 & 0.97 \\
+				floortile & 34 & 342 & 0.90 \\
+				grid & 42 & 153 & 0.73 \\
+				gripper & 2 & 30 & 0.93 \\
+				hanoi & 1 & 63 & 0.98 \\
+				hiking & 69 & 174 & 0.60 \\
+				miconic & 3 & 72 & 0.96 \\
+				npuzzle & 2 & 24 & 0.92 \\
+                                parking & 4 & 111 & 0.96 \\
+				satellite & 24 & 75 & 0.68 \\
+				transport & 4 & 78 & 0.95 \\
+				visitall & 2 & 24 & 0.92 \\
+				zenotravel & 3 & 63 & 0.95
+			\end{tabular}
+                        	\end{footnotesize}
+		\end{center}
+	\caption{\small Evaluation of the quality of the learned models with respect to an observations test set.}
+	\label{fig:observationstest}
+\end{table}
+
+The learning scores of several domains in Table~\ref{fig:observationsmap} are above the ones reported in Table~\ref{fig:observationsnomap}.
+The reason lies in the particular observations comprised by the test sets.
+As an example, in the {\em driverlog} domain, the action schema {\small \tt disembark-truck} is missing from the learned model because this action is never induced from the observations in the training set; that is, such action never appears in the corresponding \emph{unobserved} plan.
+The same happens with the {\small \tt paint-down} action of the {\em floortile} domain or {\small \tt move-curb-to-curb} in the {\em parking} domain.
+Interestingly, these actions do not appear either in the test sets and so the learned action models are not penalized in Table~\ref{fig:observationstest}.
+Generating {\em informative} and {\em representative} observations for learning planning action models is an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.

### 2018-04-18 doc/aij18/learning_task.tex
new paper organization

\section{Learning task}
\label{learning_task}

\subsection{\strips\ action schemas}
This work addresses the learning of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}. Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.

\begin{figure}[hbt!]
\begin{footnotesize}
\begin{verbatim}
(:action stack
 :parameters (?v1 ?v2 - object)
 :precondition (and (holding ?v1) (clear ?v2))
 :effect (and (not (holding ?v1)) (not (clear ?v2)) (handempty) (clear ?v1) (on ?v1 ?v2)))
\end{verbatim}
\end{footnotesize}
 \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
\label{fig:stack}
\end{figure}

To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL. Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$. Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t. $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.

Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame. For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two. We define $F_v$, a new set of fluents s.t. $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$, i.e. the variable names, and that defines the elements that can appear in an action schema. For the {\em blocksworld}, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.

For a given operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of fluents that represent the elements that can appear in that action schema. For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.

We assume also that actions $a\in A$ are instantiated from \strips\ operator schemas $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
\begin{itemize}
\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$. The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
\end{itemize}
Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, $2^{2|F_v(\xi)|}$ defines the size of space of the possible \strips\ models for that operator, given that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition. For instance, this number is 4194304 for the blocksworld {\tt stack} operator while is only 1024 for the {\tt pickup} operator.

Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same headers so they can be built from the same set of possible elements. Formally, iff $head(\xi)=head(\xi')$ so it also holds that $F_v(\xi)=F_v(\xi')$. For instance we can claim that the {\tt stack} and {\tt unstack} blocksworld operators are {\em comparable} while  {\tt stack} and {\tt pickup} are not. Likewise we say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.

Learning \strips\ action models from fully available input knowledge, i.e. from plans where the {\em pre-} and {\em post-states} of every action are available, is straightforward~\cite{jimenez2012review}:
\begin{itemize}
  \item {\em Preconditions} are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action, that is any action that belongs to the same operator scheme.
  \item {\em Effects} are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
\end{itemize}
This section formalizes more challenging learning tasks where less input knowledge is available for instance, because it cannot be observed.

\subsection{Learning from observations of plan executions}
The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its actions, the actual executed actions are unobserved. This learning task is formalized as a tuple $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$:
\begin{itemize}
\item $\mathcal{M}$, the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is the sequence of {\em state observations} obtained watching the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. We assume that the initial state $s_0$ is {\em fully observable} while states $s_i$ s.t. {\small $1\leq i\leq |\mathcal{O}|$} can be {\em partially observable}, meaning that some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative. In the extreme, states $s_i$ {\small $1\leq i\leq |\mathcal{O}|$}, can be missing but we assume that any state observation is {\em noiseless}, meaning that if the value of a fluent is observed it is correct.
\item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame. Note that $\Psi$ can be inferred from the state observations provided that at least a state $s\in \mathcal{O}$ is a full state, that is $|s|=|F|$.
\end{itemize}

A solution to a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the given state observations $\mathcal{O}$ and the predicates $\Psi$. Solving $\Lambda$ implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model. Figure~\ref{fig:example-observations} shows an example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence obtained inverting a 2-block tower. Learning the action model from this example implies inferring the unobserved plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$.

\begin{figure}[hbt!]
{\footnotesize\tt ;;;;;; Headers in $\mathcal{M}$}
\begin{footnotesize}
\begin{verbatim}
(pickup v1) (putdown v1) (stack v1 v2} (unstack v1 v2)
\end{verbatim}
\end{footnotesize}
\vspace{0.2cm}
{\footnotesize\tt ;;;;;; Predicates $\Psi$}
\begin{footnotesize}
\begin{verbatim}
(handempty) (holding ?o  - object) (clear ?o - object) (ontable ?o - object)
(on ?o1 - object ?o2 - object)
\end{verbatim}
\end{footnotesize}
\vspace{0.2cm}
{\footnotesize\tt ;;;;;; Observations $\mathcal{O}$}
\begin{footnotesize}
\begin{verbatim}
;;; observation #0
(clear B) (on B A) (ontable A) (handempty)

;;; observation #1
(holding B) (clear A) (ontable A)

;;; observation #2
(clear A) (ontable A) (clear B) (ontable B) (handempty)

;;; observation #3
(holding A) (clear B) (ontable B)

;;; observation #4
(clear A) (on A B) (ontable B) (handempty)
\end{verbatim}
\end{footnotesize}
 \caption{\small Task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
\label{fig:example-observations}
\end{figure}

In some cases we may not require to start learning from scratch because some preconditions and/or some positive or negative effects may be a priori known. The operator schemas in $\mathcal{M}$, that are given as input, may be not {\em empty} but {\em partially specified}. Such scenario is relevant for {\em policy learning}. A policy is function that maps states into actions and represents the conditions under which actions should be applied to achieve certain goals. Given an action model, the task of learning a policy that is compliant with a set of observations can be defined as learning extra preconditions for each action scheme in the model. These extra preconditions capture when actions can be applied according to the policy. A policy that exactly defines a single applicable action for each reachable state is a {\em full policy} otherwise, we say it is a {\em partial policy}. In the general case, learning a {\em full policy} for an arbitrary planning task is complex because a given action scheme may be applicable in different situations and also because the compact representation of this set of different situations may require the computation of {\em high-level state features}~\cite{lotinac2016automatic}.

The $\Lambda$ learning task can also be redefined to cover the scenario where some of the actions executed by the observed agent are available. If all the executed actions are known then states observations should be partial otherwise, the learning task is trivial. The learning task is now formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
\begin{itemize}
\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is the action sequence that produces the sequence of state observations given in $\mathcal{O}$. Again we assume that action observations are noiseless, meaning that if the value of an action is observed it is correct. When the input plan is {\em diverse} enough, i.e. $\pi$ contains at least one ground action for each of the aimed action schemes, the set of {\em empty} operator schemas $\mathcal{M}$ can be inferred from $\pi$.
\end{itemize}

Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of a four-action plan for inverting a two-block tower. In this example $\mathcal{O}=\tup{s_0,s_4}$ which means that only the first and last states are observed and the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown. $\mathcal{M},\Psi$ are skipped, since they are the same as in Figure~\ref{fig:example-observations}.

\begin{figure}[hbt!]
{\footnotesize\tt ;;;;;; Observations $\mathcal{O}$}
\begin{footnotesize}
\begin{verbatim}
;;; observation #0
(clear B) (on B A) (ontable A) (handempty)

;;; observation #4
(clear A) (on A B) (ontable B) (handempty)
\end{verbatim}
\end{footnotesize}

{\footnotesize\tt ;;; Plan $\pi$}
\begin{footnotesize}
\begin{verbatim}
0: (unstack B A)
1: (putdown B)
2: (pickup A)
3: (stack A B)
\end{verbatim}
\end{footnotesize}

 \caption{\small Task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for learning a {\em blocksworld} \strips\ action model from a four-action plan and two state observations.}
\label{fig:example-plans}
\end{figure}

The previous definitions formalize the learning of \strips\ action models from the observation of a single plan execution. These definitions are extensible to the more general case where the execution of multiple plans is observed. The task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O'},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given sequence of example plans producing the corresponding sequence of state observations $\mathcal{O'}$. Now $\mathcal{O'}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ is a sequence of {\em state observations} obtained watching the execution of a serie of {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, one after the other. This means that, for each {\small $1\leq i\leq n^t$}, $a_i^t\in \pi^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.

### 2018-04-25 doc/aaai-19/tm-grammars.tex
added first notes on the learning of CSGs

\def\year{2017}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai18}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{}
\setcounter{secnumdepth}{0}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{comment}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}

\lstset{
  basicstyle=\ttfamily,
  mathescape
}

\usepackage{multicol}
\usepackage{arydshln}
\usetikzlibrary{calc,backgrounds,positioning,fit}

\newcommand{\tup}[1]{{\langle #1 \rangle}}

\newcommand{\pre}{\mathsf{pre}}     % precondition
\newcommand{\del}{\mathsf{del}}     % effect
\newcommand{\add}{\mathsf{add}}     % effect
\newcommand{\eff}{\mathsf{eff}}     % effect
\newcommand{\cond}{\mathsf{cond}}   % conditional effect
\newcommand{\true}{\mathsf{true}}   % true
\newcommand{\false}{\mathsf{false}} % false
\newcommand{\PE}{\mathrm{PE}}     % precondition
\newcommand{\strips}{\textsc{Strips}}     % precondition

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

\begin{document}

\title{One-shot Learning of Context-Sensitive Grammars with Classical Planning}
\author{\#39}

% Commented for blind submission
\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
{\small Universitat Polit\`ecnica de Val\`encia.}\\
{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}

\maketitle
\begin{abstract}
This paper presents a novel approach for learning Context-Sensitive Grammars (CSGs) from a single input string. Our approach is to compile this learning task into a classical planning problem whose solutions are sequences of actions whose prefixes build the CSG and whose post-fixes validate that the built grammar is compliant with the input string. The compilation is flexible to implement the three canonical tasks for CSGs, {\it grammar generation}, {\it string production} and {\it string recognition} within the same classical planning model. Our experimental evaluation shows that the learned CSGs achieve generalization provided that the input strings are long and diverse enough to cover all the grammar rules.  
\end{abstract}

\section{Introduction}
\label{sec:section1}
A {\em formal grammar} is a set of symbols and production rules that describe how to form the possible strings of certain formal language. Usually three canonical tasks are defined over formal grammars:
\begin{itemize}
\item {\it Learning}: Given a set of strings, compute a grammar that is compliant with the input strings.
\item {\it Production}: Given a formal grammar, generate strings that belong to the language represented by the grammar.
\item {\it Recognition} (also known as {\em parsing}): Given a formal grammar and a string, determine whether the string belongs to the language represented by the grammar.
\end{itemize}

Chomsky defined four types of formal grammars that differ in the form and generative capacity of their rules~\cite{chomsky2002syntactic}. Each grammar type generates a different class of formal language that is recognizable with a different kind of automaton: {\bf Type-0} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Turing machine} automaton. {\bf Type-1} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Turing machine} automaton. {\bf Type-2} corresponds to {\em context-free} languages that are recognizable with a {\em non-deterministic push-down automaton}. Finally, {\bf type-3} corresponds to {\em regular} languages that can be recognized with a {\em Finite state automaton}.

Figure~\ref{fig:csg}(a) shows an example of a CSG grammar that generates the $\{a^nb^nc^n : n \geq 1 \}$ language ({\em Type-1}). The grammar defines ten production rules, contains three terminal symbols ($a$, $b$ and $c$) and five non-terminal symbol ($S$,$B$,$C$,$W$ and $Z$). This CSG can generate, for instance, the string $aaabbbccc$ by applying the sequence of rules $\tup{1,1,0,2,3,4,5,2,3,4,5,2,3,4,5,6,7,7,8,9,9}$. The {\it generation chain} in Figure~\ref{fig:csg}(b) exemplifies this rule application and proves that the string $aaabbbccc$ belongs to the $\{a^nb^nc^n : n \geq 1 \}$ language defined by the CSG grammar.

\begin{figure}
    \begin{subfigure}[]{0.3\textwidth}
      \begin{lstlisting}
0. $S\rightarrow aBC$
1. $S\rightarrow aSBC$
2. $CB\rightarrow CZ$
3. $CZ\rightarrow WZ$
4. $WZ\rightarrow WC$
5. $WC\rightarrow BC$
6. $aB\rightarrow ab$
7. $bB\rightarrow bb$	
8. $bC\rightarrow bc$
9. $cC\rightarrow cc$     
      \end{lstlisting}

	\hspace*{1.8cm}(a)
  \end{subfigure}
  \begin{tiny}  
  \begin{subfigure}[]{0.15\textwidth}
    \begin{lstlisting}
S
aSBC
aaSBCBC
aaaBCBCBC
aaaBCZCBC
aaaBWZCBC
aaaBWCCBC
aaaBBCCBC
aaaBBCCZC
aaaBBCWZC
aaaBBCWCC
aaaBBCBCC
aaaBBCZCC
aaaBBWZCC
aaaBBWCCC
aaaBBBCCC
aaabBBCCC
aaabbBCCC
aaabbbCCC
aaabbbcCC
aaabbbccC
aaabbbccc 
\end{lstlisting}
	\hspace*{.3cm}(b)
  \end{subfigure}
  \end{tiny}    
  \caption{\small (a) Example of a context-sensitive grammar; (b) the corresponding {\it generation chain} for the string $aaabbbccc$.}
  \label{fig:csg}
\end{figure}

Previous work showed that a classical planning compilation can implement the three canonical tasks (namely {\it grammar generation}, {\it string production} and {\it string recognition}) for {\em Type-3} and {\em Type-2} grammars~\cite{segovia2017generating}. In this work we show that a novel classical planning compilation implements these three canonical tasks for grammars of the three last three types ({\em Type-3}, {\em Type-2} and {\em Type-1}). In addition, this novel compilation has fewer input parameters since it does not require to specify any bound on the number of non-terminal symbols or in the size of the grammar rules. The reported empirical results show that the learned CSGs achive generalization provided that the input strings {\em complete}, meaning that they are long and diverse enough to cover all the grammar rules.

\section{Background}
This section defines the formalization of CSGs and the classical planning model that we follow in this work.

\subsection{Classical planning}
Our approach for the one-shot learning of CSGs is compiling this inductive learning task into a classical planning task.

We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e. either~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values). We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.

A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$, so the size of the state space is $2^{|F|}$. Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often we will abuse of notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$. We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$. The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.

The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition. A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The {\em plan length} is denoted with $|\pi|=n$ . A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the application of the plan $\pi$ in the initial state $I$.

\subsection{Context-Sensitive Grammars}
A {\em context-sensitive grammar} (CSGs) is a formal grammar in which the left-hand sides and right-hand sides of any production rules may be surrounded by a context of terminal and nonterminal symbols. CSGs are more general than {\em context-free grammars}, in the sense that there are languages that can be described by CSG but not by context-free grammars. On the other hand CSGs are less general than unrestricted grammars.

We define a CSGs as a tuple $\mathcal{G}=\tup{V,\Sigma,R}$ where,
\begin{itemize}
\item $V$ is the finite set of non-terminal symbols, also called variables. With $v_0\in V$ is the start non-terminal symbol that represents the whole grammar.
\item $\Sigma$ is the finite set of terminal symbols, which are disjoint from the set of non-terminal symbols, i.e.~$V\cap \Sigma\neq\emptyset$. The set of terminal symbols is the alphabet of the language defined by $\mathcal{G}$
\item $R:V\rightarrow (V\cup \Sigma)^*$ is the finite set of production rules in the grammar. By definition rules $r\in R$ always contain a single non-terminal symbol on the left-hand side.
\end{itemize}

\section{Learning Context-Sensitive Grammars}

\begin{figure}
\begin{footnotesize}
\begin{verbatim}
(:action update-rule1 ;;; 0,A/1,R,B
  :parameters (?xl ?x ?xr)
  :precondition (and (head ?x) 
                     (next ?xl ?x) (next ?x ?xr) 
                     (symbol0 ?x) (stateA))
  :effect (and (not (head ?x)) 
               (not (symbol0 ?x)) 
               (not (stateA))
               (symbol1 ?x) (head ?xr) (stateB)))
\end{verbatim}
\end{footnotesize}
 \caption{\small .}
\label{fig:stack}
\end{figure}

\section{Parsing and Production of CSGs with classical planning}

\section{Evaluation}

\section{Related work}
%Our approach is related to recent work on structured prediction. Unlike existing approaches that leverage domain-specific knowledge to reduce the program hypotheses and enable fast inference, e.g., the unsupervised learning of probabilistic programs~\cite{Tenenbaum:programs:2015}, our approach is fully domain independent. On the other hand we cannot cope with noisy observations or probabilities over the space of outputs and inputs.

The learning of CFGs can also be understood in terms of activity recognition, such that the library of activities is formalized as a CFG, the library is initially unknown, and the input strings encode observations of the activities to recognize. {\it Activity recognition} is traditionally considered independent of the research done on automated planning, using handcrafted libraries of activities and specific algorithms~\cite{ravi2005activity}. An exception is the work by \citeauthor{ramirez2009plan}~[\citeyear{ramirez2009plan,ramirez2010probabilistic}] where goal recognition is formulated and solved with planning. As far as we know our work is the first that tightly integrates the tasks of (1) grammar learning, (2) recognition and (3) production using a common planning model and an off-the-shelf classical planner.

Hierarchical Task Networks (HTNs) is a powerful formalism for representing libraries of plans~\cite{nau2003shop2}. HTNs are also defined at several levels such that the tasks at one level are decomposed into other tasks at lower levels with HTN decomposition methods sharing similarities with production rules in CFGs. There is previous work in generating HTNs~\cite{hogg2008htn,conf:ecai:Lotinac16} and an interesting research direction is to extend our approach for computing HTNs from flat sequences of actions. This aim is related to Inductive Logic Programming (ILP)~\cite{muggleton1999inductive} that learns logic programs from examples. Unlike logic programs (or HTNs) the CFGs that we generate are propositional and do not include variables. Techniques for learning high level state features that include variables are promising for learning lifted grammars~\cite{damir-derived-ijcai16}.

\section{Conclusions}
There is exhaustive previous work on learning CFGs given a corpus of correctly parsed input strings~\cite{sakakibara1992efficient,langley2000learning} or using positive and negative examples~\cite{de2010grammatical,muggleton2014meta}. This work addresses generating CFGs using only a small set of positive examples (in some cases even one single string that belongs to the language). Furthermore we follow a compilation approach that benefits straightforwardly from research advances in classical planning and that is also suitable for {\it production} and {\it recognition} tasks with arbitrary CFGs.

Our compilation bounds the number of rules $m$, the length of these rules $n$, the size of the stack $\ell$ and the length of the input strings $z$. If these bounds are too small, the classical planner used to solve the output planning task will not be able to find a solution. Larger values for these bounds do not formally affect to our approach, but in practice, the performance of classical planners is sensitive to the size of the input. Interestingly our approach can also follow an incremental strategy where we generate the CFG for a given sub-language and then encode this {\it sub-grammar} as an auxiliary procedure for generating more challenging CFGs~\cite{segovia2016hierarchical}.

The size of the compilation output also depends on the number of examples. Empirical results show that our approach is able to generate non-trivial CFGs from very small data sets. Another interesting extension would be to add negative input strings, which would require a mechanism for validating that a given CFG does {\em not} generate a given string, or to accept incomplete input strings that would require combining the generation and production mechanisms.

\bibliographystyle{aaai}
\bibliography{tm-grammars}
\end{document}

### 2018-04-25 doc/aaai-19/tm-grammars.tex
further notes on the learning of CSGs

@@
 Each grammar type generates a different class of formal language that is recognizable with a different kind of automaton: {\bf Type-0} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Turing machine} automaton.
-{\bf Type-1} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Turing machine} automaton.
+{\bf Type-1} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Linear-bounded Turing machine} automaton.
 {\bf Type-2} corresponds to {\em context-free} languages that are recognizable with a {\em non-deterministic push-down automaton}.
@@
 This CSG can generate, for instance, the string $aaabbbccc$ by applying the sequence of rules $\tup{1,1,0,2,3,4,5,2,3,4,5,2,3,4,5,6,7,7,8,9,9}$.
-The {\it generation chain} in Figure~\ref{fig:csg}(b) exemplifies this rule application and proves that the string $aaabbbccc$ belongs to the $\{a^nb^nc^n : n \geq 1 \}$ language defined by the CSG grammar.
+The {\it parse tree} in Figure~\ref{fig:csg}(b) exemplifies this rule application and proves that the string $aaabbbccc$ belongs to the $\{a^nb^nc^n : n \geq 1 \}$ language defined by the CSG grammar.
 
@@
   \end{subfigure}
+  \begin{subfigure}[]{0.15\textwidth}
   \begin{tiny}  
-  \begin{subfigure}[]{0.15\textwidth}
 \begin{lstlisting}
@@
 \end{lstlisting}
+  \end{tiny}    
 \hspace*{.3cm}(b)
   \end{subfigure}
-  \end{tiny}    
-  \caption{\small (a) Example of a context-sensitive grammar; (b) the corresponding {\it generation chain} for the string $aaabbbccc$.}
+
+  \caption{\small (a) Example of a context-sensitive grammar; (b) the corresponding {\it parse tree} for the string $aaabbbccc$.}
   \label{fig:csg}
@@
 \subsection{Context-Sensitive Grammars}
-A {\em context-sensitive grammar} (CSGs) is a formal grammar in which the left-hand sides and right-hand sides of any production rules may be surrounded by a context of terminal and nonterminal symbols.
-CSGs are more general than {\em context-free grammars}, in the sense that there are languages that can be described by CSG but not by context-free grammars.
-On the other hand CSGs are less general than unrestricted grammars.
-
-We define a CSGs as a tuple $\mathcal{G}=\tup{V,\Sigma,R}$ where,
+We define a CSGs as a tuple $\mathcal{G}=\tup{V,\Sigma,R}$, where:
 \begin{itemize}
@@
 The set of terminal symbols is the alphabet of the language defined by $\mathcal{G}$
-\item $R:V\rightarrow (V\cup \Sigma)^*$ is the finite set of production rules in the grammar. By definition rules $r\in R$ always contain a single non-terminal symbol on the left-hand side.
+\item $R:\alpha v\beta\rightarrow \alpha\gamma\beta$ where $v\in V$, $\alpha,\beta\in(V\cup\Sigma)^*$.
+That is, the left-hand and right-hand sides of any production rules $r\in R $may be surrounded by a context of terminal and nonterminal symbols.
 \end{itemize}
+
+For any two strings  we say that $e_1$ {\it directly yields} $e_2$, denoted by $e_1\Rightarrow e_2$, iff $e_1$ can be written as $e_3\alpha A\beta e_4$ and $e2$ can be written as $e_3\alpha \gamma\beta e_4$, for some production rule $(\alpha A\beta\rightarrow\alpha\gamma\beta)\in R$, and some context strings $e_3,e_4\in (V\cup \Sigma)^*$.
+Furthermore we say $e_1$ {\it yields} $e_2$, denoted by $e_1\Rightarrow^* e_2$, iff $\exists k\geq 0$ and $\exists u_1, \ldots, u_k$ such that $e_1=u_1\Rightarrow u_2\Rightarrow \ldots \Rightarrow u_k=e_2$ for som $k\geq 0$ and some strings $u_1,\ldots,u_{k-1}\in(V\cup\Sigma)^*$.
+That is, the relation $\Rightarrow^*$ is the reflexive transitive closure of the $\rightarrow$ relation.
+
+For instance, Figure~\ref{fig:csg}(b) shows how the string $S$ yields the string $aaabbbccc$.
+The language of a CSG, $L(\mathcal{G})=\{e\in \Sigma^*: v_0\Rightarrow^* e\}$, is the set of strings that contain only terminal symbols and that can be yielded from the string that contains only the initial non-terminal symbol $v_0$ (which is denoted by $S$ in the example of Figure~\ref{fig:csg}(b)).
+
+Given a CSG $\mathcal{G}$ and a string $e\in L(\mathcal{G})$ that belongs to its language, we define a {\it parse tree} $t_{\mathcal{G},e}$ as an ordered, rooted tree that determines a concrete syntactic structure of $e$ according to the rules in $\mathcal{G}$:
+\begin{itemize}
+\item Each node in a parse tree $t_{\mathcal{G},e}$ is either:
+\begin{itemize}
+\item An {\it internal node} that corresponds to the application of a rule $r\in R$.
+\item A {\it leaf node} that corresponds to a terminal symbol $\sigma\in \Sigma$ and has no outgoing branches.
+\end{itemize}
+\item Edges in a parse tree $t_{\mathcal{G},e}$ connect non-terminal symbols to terminal or non-terminal symbols following the rules $R$ in $\mathcal{G}$.
+\end{itemize}
+
+\subsection{Linear-bounded Turing Machines}
+We define a {\em Turing machine} as a tuple $M=\tup{Q,q_o,Q_{\bot},\mathcal{T},\square,\Sigma,\delta}$, where:
+\begin{itemize}
+\item $Q$, is a finite and non-empty set of machine states such that $q_0\in Q$ it the initial state of the machine and $Q_{\bot}\subseteq Q$ is the subset of terminal states.  
+\item $\mathcal{T}$ is the {\em tape alphabet}, that is a finite an non-empty set of symbols that includes the blank symbol $\square\in$(the only symbol allowed to occur on the tape infinitely often at any step during the computation).
+\item $\Sigma$ is the {\em input alphabet}, the set of symbols allowed to initially appear in the tape.
+\item $\delta: (Q\setminus Q_{\bot})\times \mathcal{T}\rightarrow Q\times \mathcal{T}\times\{left,right\}$ is the {\em transition function}.
+If $\delta$ is not defined for the current state of the machine and the current tape symbol, then the machine halts.
+\end{itemize}
+
+The most common convention is to represent the transitions defined by the $\delta$ function is with a table where table rows, are indexed with the current tape symbol, while table columns, are indexed by the current machine state.
+For each possible tape symbol and state of the machine there is a table entry that defines: (1) the tape symbol to print at the current position of the header in the tape (2) whether the header is shifted {\em left} or {\em right} and (3), the new state of the machine.
+For instance, the Figure~\ref{tab:tm-anbncn} shows the table that represents a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
+In this example the tape alphabet is $\Sigma=\{a,b,c,x,y,z,\square\}$ while the set of possible machine states is $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ and \underline{$q_5$} is the acceptor state.
+
+\begin{figure}
+\begin{center}
+    \begin{tabular}{| c | c | c | c | c | c | c |}
+    \hline
+      & $q_0$ & $q_1$ & $q_2$ & $q_3$ & $q_4$ & \underline{$q_5$} \\ \hline
+    a & x,r,$q_1$ & a,r,$q_1$ & - &  a,l,$q_3$ & - & - \\ \hline
+    b & - & y,r,$q_2$ & b,r,$q_2$ & b,l,$q_3$ & - & - \\ \hline
+    c & - & - & z,l,$q_3$ & - & - & - \\ \hline
+    x & - & - & - & x,r,$q_0$ & - & - \\ \hline
+    y & y,r,$q_4$ & y,r,$q_1$ & - & y,l,$q_3$ & y,r,$q_4$ & - \\ \hline
+    z & - & - & z,r,$q_2$ & z,l,$q_3$ & z,r,$q_4$ & - \\\hline
+    $\square$ & - & - & - & - & $\square$,r,$q_5$  & - \\                
+    \hline
+    \end{tabular}
+\end{center}
+  \caption{\small Example of a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the acceptor state).}
+  \label{tab:tm-anbncn}
+\end{figure}
+
+A {\em Turing machine} is an abstract model of computation that operates on an infinite memory tape.
+An approach to implement in practice a given Turing machine, is to bound the size of its tape.
+{\em Linear bounded Turing machines} are acceptors for the class of context-sensitive languages~\cite{hopcroft:automatatheory:2001} and present the following three features:
+\begin{enumerate}
+\item Its input alphabet includes two special symbols, serving as left and right endmarkers.
+\item Its transitions may not print other symbols over the endmarkers.
+\item Its transitions may neither move to the left of the left endmarker nor to the right of the right endmarker.
+\end{enumerate}
+This means that, instead of having potentially infinite tape on which to compute, computation is restricted to the portion of the tape containing the input plus the two tape squares holding the endmarkers.
 
@@
 
-\begin{figure}
+Our approach is to leverage on classical planning to learn an \strips\ action model where each action in the model represents an entry in the table that defines the transitions of a {\em Linear-bounded Turing Machine} that recognices the string given as input.
+First we show how an entry in the table that defines the transitions of a {\em Linear-bounded Turing Machine} can be modeled as \strips\ action schema.
+Next we show how classical planning can be effectively used to learn action models of this kind.
+
+\subsection{Modeling a Linear-bounded Turing Machine with Classical Planning}
+Given $\delta$ the {\em transition function} of a {\em Linear-bounded Turing Machines} $M$, as defined in the previous section.
+Here we show that this transition can be encoded as a classical planning frame $P=\tup{F,A}$.
+
+We assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
+Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
+Given a set of {\em objects} $\Omega$ that represent each of the possible cells in the tape of the given Turing Machine $M$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$, where $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+In more detail the predicates in $\Psi$ are:
+\begin{itemize}
+\item {\tt (head ?x)} that encodes the current position of the header in the tape.
+\item {\tt (next ?x1 ?x2)} that encodes that the cell ?x2 follows cell ?x1 in the tape.
+\item {\tt (symbol-$\sigma$ ?x)} that encodes that the cell ?x in the tape contains the symbol $\sigma\in\Sigma$.
+\item {\tt (state-$q$)} that encodes that $q\in Q$ is the current machine state.
+\end{itemize}
+
+Likewise we assume that actions $a\in A$ are instantiated from \strips\ operator schema. In particualar for each transition defined in $\delta$ a \strips\ action model is defined such that:
+\begin{itemize}
+\item The name of the schema is rule-id where id uniquely identifies the transition in $\delta$. The parameters of the schema are three tape cells $(?xl\ ?x\ ?xr)$
+\item The precoditions of the schema iclude always {\tt(head ?x)} and {\tt (next ?xl ?x) (next ?x ?xr)} to force that the parameters of the schema are the current cell in the tape pointed by the header and its left and right neighbours.
+Additionally the schema includes the preconditions {\tt(symbol-$\sigma$ ?x)} and {\tt (state-$q$)} on the current symbol pointed by the header and the currrent machine state.
+\item Finally the delete effects delete always the current the current symbol pointed by the header and the currrent machine state and set the new ones.
+\end{itemize}
+
+To illustrate this, Figure~\ref{fig:update-rule} shows the rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.
+The full encoding of the Turing Machine defined in Figure~\ref{tab:tm-anbncn} will produce a total of sixteen \strips\ action schema similar to the one of Figure~\ref{fig:update-rule}.
+\begin{figure}[hbt!]
 \begin{footnotesize}
 \begin{verbatim}
-(:action update-rule1 ;;; 0,A/1,R,B
+(:action rule1 ;; a,q_0 -> x,r,q_1
   :parameters (?xl ?x ?xr)
@@
                      (next ?xl ?x) (next ?x ?xr) 
-                     (symbol0 ?x) (stateA))
+                     (symbol-a ?x) (state-q_0))
   :effect (and (not (head ?x)) 
-               (not (symbol0 ?x)) 
-               (not (stateA))
-               (symbol1 ?x) (head ?xr) (stateB)))
+               (not (symbol-a ?x)) 
+               (not (state-q_0))
+               (symbol-x ?x) (head ?xr) (state-q_1)))
 \end{verbatim}
 \end{footnotesize}
- \caption{\small .}
-\label{fig:stack}
+ \caption{\small \strips\ action schema that models the rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
+\label{fig:update-rule}
 \end{figure}
 
+\subsection{Learning CSGs with Classical Planning}
+Here we show that the existing classical planning compilation for learning \strips\ action schemes~\cite{aineto:learningSTRIPS:ICAPS2018} from plans can be adapted to address the learning task defined in this paper.
+
+Note that in this case the set of preconditions and delete effects is already known for each action $a\in A$. So the learning task reduces to learning the positive effects of these actions.
+
 \section{Parsing and Production of CSGs with classical planning}
+The task of parsing can be implemented with the planning model introduced in the previous section by defining a classical planning instace whose fluents and actions encode the transition of the corresponding Turing machine, the initial state encodes the tape with the string to recognize and the goal conditions require that the turing machine ends in on its acceptor states.
+
+The task of production requires to include actions that only write once a single symbol from the {\em tape alphabet} at the tape location indicated by the header.
 

### 2018-04-26 doc/aaai-19/tm-grammars.tex
new notes on the learning of CSGs

@@
 
-\title{One-shot Learning of Context-Sensitive Grammars with Classical Planning}
+\title{Learning of Context-Sensitive Grammars with Classical Planning}
 \author{\#39}
@@
 \begin{abstract}
-This paper presents a novel approach for learning Context-Sensitive Grammars (CSGs) from a single input string.
-Our approach is to compile this learning task into a classical planning problem whose solutions are sequences of actions whose prefixes build the CSG and whose post-fixes validate that the built grammar is compliant with the input string.
-The compilation is flexible to implement the three canonical tasks for CSGs, {\it grammar generation}, {\it string production} and {\it string recognition} within the same classical planning model.
-Our experimental evaluation shows that the learned CSGs achieve generalization provided that the input strings are long and diverse enough to cover all the grammar rules.
+This paper presents a novel approach for learning {\em Context-Sensitive Grammars} (CSGs) from small sets of input strings.
+The approach guarantees that the computed grammar generalizes provided that the input strings are long and diverse enough to cover all the grammar rules.
+Our approach is to compile this learning task into a classical planning problem whose solutions are all sequences of actions with a prefix that build the CSG and and a postfix that validates the built grammar in the input strings.
+The compilation is flexible to implement the three canonical tasks of CSGs, {\it grammar generation}, {\it string production} and {\it string recognition} within the same classical planning model.
+The experimental validation of our compilation approach reports the empirical data collected when learning the CSGs for the languages $\{a^nb^nc^n : n \geq 1 \}$ and $\{a^{2^n} : n \geq 1 \}$ and shows that overcomes the performance  of a previous compilation for learning {\em Context-Free Grammars}.
 \end{abstract}
@@
 \label{sec:section1}
-A {\em formal grammar} is a set of symbols and production rules that describe how to form the possible strings of certain formal language.
+A {\em formal grammar} is a set of symbols and production rules that describe how to form the possible strings of certain formal language~\cite{hopcroft:automatatheory:2001}.
 Usually three canonical tasks are defined over formal grammars:
@@
 Chomsky defined four types of formal grammars that differ in the form and generative capacity of their rules~\cite{chomsky2002syntactic}.
-Each grammar type generates a different class of formal language that is recognizable with a different kind of automaton: {\bf Type-0} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Turing machine} automaton.
-{\bf Type-1} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Linear-bounded Turing machine} automaton.
+Each grammar type generates a different class of formal language that is recognizable with a different kind of automaton: {\bf Type-0} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Turing machine}.
+{\bf Type-1} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Linear-bounded Turing machine}.
 {\bf Type-2} corresponds to {\em context-free} languages that are recognizable with a {\em non-deterministic push-down automaton}.
-Finally, {\bf type-3} corresponds to {\em regular} languages that can be recognized with a {\em Finite state automaton}.
+Finally, {\bf type-3} corresponds to {\em regular} languages that can be recognized with a {\em Finite state machine}.
 
-Figure~\ref{fig:csg}(a) shows an example of a CSG grammar that generates the $\{a^nb^nc^n : n \geq 1 \}$ language ({\em Type-1}).
-The grammar defines ten production rules, contains three terminal symbols ($a$, $b$ and $c$) and five non-terminal symbol ($S$,$B$,$C$,$W$ and $Z$).
+Figure~\ref{fig:csg}(a) shows an example of a CSG grammar ({\em Type-1}) that generates the $\{a^nb^nc^n : n \geq 1 \}$ language.
+The grammar defines ten production rules, contains three {\em terminal symbols} ($a$, $b$ and $c$) and five {\em non-terminal symbols} ($S$,$B$,$C$,$W$ and $Z$).
 This CSG can generate, for instance, the string $aaabbbccc$ by applying the sequence of rules $\tup{1,1,0,2,3,4,5,2,3,4,5,2,3,4,5,6,7,7,8,9,9}$.
@@
 \begin{lstlisting}
-S
-aSBC
-aaSBCBC
-aaaBCBCBC
-aaaBCZCBC
-aaaBWZCBC
-aaaBWCCBC
-aaaBBCCBC
-aaaBBCCZC
-aaaBBCWZC
-aaaBBCWCC
-aaaBBCBCC
-aaaBBCZCC
-aaaBBWZCC
-aaaBBWCCC
-aaaBBBCCC
-aaabBBCCC
-aaabbBCCC
-aaabbbCCC
-aaabbbcCC
-aaabbbccC
-aaabbbccc 
+1. S
+1. aSBC
+0. aaSBCBC
+2. aaaBCBCBC
+3. aaaBCZCBC
+4. aaaBWZCBC
+5. aaaBWCCBC
+2. aaaBBCCBC
+3. aaaBBCCZC
+4. aaaBBCWZC
+5. aaaBBCWCC
+2. aaaBBCBCC
+3. aaaBBCZCC
+4. aaaBBWZCC
+5. aaaBBWCCC
+6. aaaBBBCCC
+7. aaabBBCCC
+7. aaabbBCCC
+8. aaabbbCCC
+9. aaabbbcCC
+9. aaabbbccC
+$\bot$. aaabbbccc 
 \end{lstlisting}
@@
 In this work we show that a novel classical planning compilation implements these three canonical tasks for grammars of the three last three types ({\em Type-3}, {\em Type-2} and {\em Type-1}).
-In addition, this novel compilation has fewer input parameters since it does not require to specify any bound on the number of non-terminal symbols or in the size of the grammar rules.
-The reported empirical results show that the learned CSGs achive generalization provided that the input strings {\em complete}, meaning that they are long and diverse enough to cover all the grammar rules.
+In addition, this novel compilation has fewer input parameters since it does not require to specify any bound on the size of the grammar rules.
+The experimental validation of our compilation approach reports the empirical data collected when learning the CSGs for the languages $\{a^nb^nc^n : n \geq 1 \}$ and $\{a^{2^n} : n \geq 1 \}$ and shows that overcomes the performance  of a previous compilation for learning {\em Context-Free Grammars}.
 
@@
 \subsection{Classical planning}
-Our approach for the one-shot learning of CSGs is compiling this inductive learning task into a classical planning task.
+Our approach for learning CSGs is compiling this inductive learning task into a classical planning task.
 
@@
 The set of terminal symbols is the alphabet of the language defined by $\mathcal{G}$
-\item $R:\alpha v\beta\rightarrow \alpha\gamma\beta$ where $v\in V$, $\alpha,\beta\in(V\cup\Sigma)^*$.
-That is, the left-hand and right-hand sides of any production rules $r\in R $may be surrounded by a context of terminal and nonterminal symbols.
+\item Rules in $R$ are defined as $\alpha v\beta\rightarrow \alpha\gamma\beta$ where $v\in V$, $\alpha,\beta\in(V\cup\Sigma)^*$ and $\gamma\in(V\cup\Sigma)^+$.
+That is, the left-hand and right-hand sides of any production rules $r\in R$ may be surrounded by a context of terminal and nonterminal symbols.
 \end{itemize}
 
-For any two strings  we say that $e_1$ {\it directly yields} $e_2$, denoted by $e_1\Rightarrow e_2$, iff $e_1$ can be written as $e_3\alpha A\beta e_4$ and $e2$ can be written as $e_3\alpha \gamma\beta e_4$, for some production rule $(\alpha A\beta\rightarrow\alpha\gamma\beta)\in R$, and some context strings $e_3,e_4\in (V\cup \Sigma)^*$.
-Furthermore we say $e_1$ {\it yields} $e_2$, denoted by $e_1\Rightarrow^* e_2$, iff $\exists k\geq 0$ and $\exists u_1, \ldots, u_k$ such that $e_1=u_1\Rightarrow u_2\Rightarrow \ldots \Rightarrow u_k=e_2$ for som $k\geq 0$ and some strings $u_1,\ldots,u_{k-1}\in(V\cup\Sigma)^*$.
-That is, the relation $\Rightarrow^*$ is the reflexive transitive closure of the $\rightarrow$ relation.
+For any two strings $(e_1,e_2)$ we say that $e_1$ {\it directly yields} $e_2$, denoted by $e_1\Rightarrow e_2$, iff $e_1$ can be written as $e_3\alpha A\beta e_4$ and $e2$ can be written as $e_3\alpha \gamma\beta e_4$, for some production rule $(\alpha A\beta\rightarrow\alpha\gamma\beta)\in R$, and some context strings $e_3,e_4\in (V\cup \Sigma)^*$.
+Furthermore we say $e_1$ {\it yields} $e_2$, denoted by $e_1\Rightarrow^* e_2$, iff $\exists k\geq 0$ and $\exists u_1, \ldots, u_k$ such that $e_1=u_1\Rightarrow u_2\Rightarrow \ldots \Rightarrow u_k=e_2$ for some $k\geq 1$ and some strings $u_1,\ldots,u_{k-1}\in(V\cup\Sigma)^*$.
+In other words, the relation $\Rightarrow^*$ is the reflexive transitive closure of the $\Rightarrow$ relation.
 
@@
 \item $Q$, is a finite and non-empty set of machine states such that $q_0\in Q$ it the initial state of the machine and $Q_{\bot}\subseteq Q$ is the subset of terminal states.  
-\item $\mathcal{T}$ is the {\em tape alphabet}, that is a finite an non-empty set of symbols that includes the blank symbol $\square\in$(the only symbol allowed to occur on the tape infinitely often at any step during the computation).
+\item $\mathcal{T}$ is the {\em tape alphabet}, that is a finite an non-empty set of symbols that includes the {\em blank symbol} $\square\in\mathcal{T}$ (the only symbol allowed to occur on the tape infinitely often at any step during the computation).
 \item $\Sigma$ is the {\em input alphabet}, the set of symbols allowed to initially appear in the tape.
@@
 
-The most common convention is to represent the transitions defined by the $\delta$ function is with a table where table rows, are indexed with the current tape symbol, while table columns, are indexed by the current machine state.
-For each possible tape symbol and state of the machine there is a table entry that defines: (1) the tape symbol to print at the current position of the header in the tape (2) whether the header is shifted {\em left} or {\em right} and (3), the new state of the machine.
-For instance, the Figure~\ref{tab:tm-anbncn} shows the table that represents a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
-In this example the tape alphabet is $\Sigma=\{a,b,c,x,y,z,\square\}$ while the set of possible machine states is $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ and \underline{$q_5$} is the acceptor state.
+A table is the most common convention to represent the transitions defined by $\delta$, where the table rows are indexed with the current tape symbol, while the table columns are indexed by the current machine state.
+For each possible tape symbol and state of the machine there is a table entry that defines: (1) the tape symbol to print at the current position of the header (2) whether the header is shifted {\em left} or {\em right} after the print operation and (3), the new state of the machine after the print operation.
+For instance, Figure~\ref{tab:tm-anbncn} shows the table that represents the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
+In this example the tape alphabet is $\Sigma=\{a,b,c,x,y,z,\square\}$ while the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the acceptor state.
 
@@
 A {\em Turing machine} is an abstract model of computation that operates on an infinite memory tape.
-An approach to implement in practice a given Turing machine, is to bound the size of its tape.
-{\em Linear bounded Turing machines} are acceptors for the class of context-sensitive languages~\cite{hopcroft:automatatheory:2001} and present the following three features:
+An approach to implement in practice a Turing machine, is to bound the size of its tape.
+This can be done by:
 \begin{enumerate}
-\item Its input alphabet includes two special symbols, serving as left and right endmarkers.
-\item Its transitions may not print other symbols over the endmarkers.
-\item Its transitions may neither move to the left of the left endmarker nor to the right of the right endmarker.
+\item Extending the input alphabet with two special symbols, serving as {\em left} and {\em right endmarkers}.
+\item Limiting computation to the portion of the tape containing the input plus the two tape cells holding the endmarkers.
+This means that transitions may not print other symbols over the endmarkers and that may neither move to the left of the left endmarker nor to the right of the right endmarker.
 \end{enumerate}
-This means that, instead of having potentially infinite tape on which to compute, computation is restricted to the portion of the tape containing the input plus the two tape squares holding the endmarkers.
 
@@
 
-Our approach is to leverage on classical planning to learn an \strips\ action model where each action in the model represents an entry in the table that defines the transitions of a {\em Linear-bounded Turing Machine} that recognices the string given as input.
+Our approach to learning CSGs is to leverage on classical planning to learn an \strips\ action model s.t.
+each action schema in the model encodes a transition of a {\em Linear-bounded Turing Machine} that recognices the strings given as input.
+
 First we show how an entry in the table that defines the transitions of a {\em Linear-bounded Turing Machine} can be modeled as \strips\ action schema.
@@
 \subsection{Modeling a Linear-bounded Turing Machine with Classical Planning}
-Given $\delta$ the {\em transition function} of a {\em Linear-bounded Turing Machines} $M$, as defined in the previous section.
-Here we show that this transition can be encoded as a classical planning frame $P=\tup{F,A}$.
+Given $\delta$, the {\em transition function} of a {\em Linear-bounded Turing Machines} $M$, it can be encoded as a classical planning frame $P=\tup{F,A}$ as follows.
 
-We assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
+We assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL~\cite{fox2003pddl2}.
 Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$ that represent each of the possible cells in the tape of the given Turing Machine $M$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$, where $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine $M$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$, where $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
 In more detail the predicates in $\Psi$ are:
@@
 \item {\tt (head ?x)} that encodes the current position of the header in the tape.
-\item {\tt (next ?x1 ?x2)} that encodes that the cell ?x2 follows cell ?x1 in the tape.
-\item {\tt (symbol-$\sigma$ ?x)} that encodes that the cell ?x in the tape contains the symbol $\sigma\in\Sigma$.
+\item {\tt (next ?x1 ?x2)} that encodes that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
+\item {\tt (symbol-$\sigma$ ?x)} that encodes that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
 \item {\tt (state-$q$)} that encodes that $q\in Q$ is the current machine state.
@@
 
-Likewise we assume that actions $a\in A$ are instantiated from \strips\ operator schema. In particualar for each transition defined in $\delta$ a \strips\ action model is defined such that:
+Likewise we assume that actions $a\in A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action model is defined such that:
 \begin{itemize}
-\item The name of the schema is rule-id where id uniquely identifies the transition in $\delta$. The parameters of the schema are three tape cells $(?xl\ ?x\ ?xr)$
-\item The precoditions of the schema iclude always {\tt(head ?x)} and {\tt (next ?xl ?x) (next ?x ?xr)} to force that the parameters of the schema are the current cell in the tape pointed by the header and its left and right neighbours.
-Additionally the schema includes the preconditions {\tt(symbol-$\sigma$ ?x)} and {\tt (state-$q$)} on the current symbol pointed by the header and the currrent machine state.
-\item Finally the delete effects delete always the current the current symbol pointed by the header and the currrent machine state and set the new ones.
+\item The {\bf header} of the schema is {\tt rule-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
+\item The {\bf precoditions} of the schema includes {\tt(head ?x)} and {\tt (next ?xl ?x) (next ?x ?xr)} to force that $?x$ is the current tape cell pointed by the header and that $?xl$ and $?xr$ respectively are its left and right neighbours.
+Additionally the schema includes preconditions {\tt(symbol-$\sigma$ ?x)} and {\tt (state-$q$)} to capture the current symbol pointed by the header and the currrent machine state.
+\item The {\bf delete effects} remove the current symbol pointed by the header and the currrent machine state.
+Finally, the {\bf positive effects} set the new symbol pointed by the header and the new machine state
 \end{itemize}
@@
 To illustrate this, Figure~\ref{fig:update-rule} shows the rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.
-The full encoding of the Turing Machine defined in Figure~\ref{tab:tm-anbncn} will produce a total of sixteen \strips\ action schema similar to the one of Figure~\ref{fig:update-rule}.
+The full encoding of the Turing Machine defined in Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema with the same structure as the one of Figure~\ref{fig:update-rule}.
 \begin{figure}[hbt!]
-\begin{footnotesize}
+\begin{scriptsize}
 \begin{verbatim}
@@
 \end{verbatim}
-\end{footnotesize}
- \caption{\small \strips\ action schema that models the rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
+\end{scriptsize}
+ \caption{\small \strips\ action schema that models the first rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
 \label{fig:update-rule}
@@
 \subsection{Learning CSGs with Classical Planning}
-Here we show that the existing classical planning compilation for learning \strips\ action schemes~\cite{aineto:learningSTRIPS:ICAPS2018} from plans can be adapted to address the learning task defined in this paper.
+Here we show that the existing classical planning compilation for learning \strips\ action schemes from plan traces~\cite{aineto:learningSTRIPS:ICAPS2018} can be adapted to address the learning of CSGs.
+Note that in this case the set of preconditions and delete effects of the \strips\ action schemes is already known so the learning task reduces to learning positive effects.
 
-Note that in this case the set of preconditions and delete effects is already known for each action $a\in A$. So the learning task reduces to learning the positive effects of these actions.
+We formalize the learning of \strips\ action schemes from plan traces as a tuple $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ where:
+\begin{itemize}
+\item $\mathcal{M}$, the set of {\em partially specified} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
+\item $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_{n}}$ is the {\em plan trace} obtained watching the execution of the classical plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+\item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame. 
+\end{itemize}
+
+A solution to a $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the given state observations $\mathcal{T}$ and the predicates $\Psi$.
+
+Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ the compilation defined by~\citeauthor{aineto:learningSTRIPS:ICAPS2018} outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+\begin{itemize}
+\item $F_{\Lambda}$ contains:
+\begin{itemize}
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the input observations.
+Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
+\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
+\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation in $\mathcal{O}$ where the action model is validated.
+\end{itemize}
+\item $I_{\Lambda}$ encodes the first observation, $s_0\subseteq F$, and sets $mode_{prog}$ to true.
+Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
+Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
+
+\item $G_{\Lambda}=\bigcup_{1\leq i\leq |\mathcal{O}|}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
+\item $A_{\Lambda}$ comprises three kinds of actions:
+\begin{enumerate}
+\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$ that add a {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
+\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
+& \{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
+\end{align*}
+\end{small}
+
+\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
+Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
+&\cup \{\neg mode_{val}\},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
+&\{\emptyset\}\rhd\{mode_{val}\}.
+\end{align*}
+\end{small}
+
+\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{O}$, {\tt\small $1\leq i< |\mathcal{O}|$}.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\\
+& \cup \{mode_{val}\},\\
+\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
+\end{align*}
+\end{small}
+\end{enumerate}
+\end{itemize}
+
+\begin{lemma}
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
+In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
+These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{T}$, starting from the corresponding initial state and following the sequence of actions defined by the plans in $\Pi$.
+This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
+\end{small}
+\end{proof}
+
+\begin{lemma}
+Completeness.
+Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T},\Pi}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+\end{lemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{T},\Pi$.
+\end{small}
+\end{proof}
+
+The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
+\begin{itemize}
+\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
+The larger these numbers, the larger the size of the $F_v(\xi)$ sets.
+This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
+\item The number of given state observations. The larger $|\mathcal{T}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+\end{itemize}
 

### 2018-04-26 doc/aaai-19/tm-grammars.tex
further notes on the learning of CSGs

@@
 
-\title{Learning of Context-Sensitive Grammars with Classical Planning}
+\title{Learning Context-Sensitive Grammars with Classical Planning}
 \author{\#39}
@@
 \subsection{Modeling a Linear-bounded Turing Machine with Classical Planning}
-Given $\delta$, the {\em transition function} of a {\em Linear-bounded Turing Machines} $M$, it can be encoded as a classical planning frame $P=\tup{F,A}$ as follows.
+Given $\delta$, the {\em transition function} of a {\em Linear-bounded Turing Machines} $M$, it can be encoded as a classical planning frame $\Phi=\tup{F,A}$ as follows.
 
@@
 Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine $M$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$, where $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-In more detail the predicates in $\Psi$ are:
+In more detail there are four predicates in $\Psi$:
 \begin{itemize}
@@
 
-\subsection{Learning CSGs with Classical Planning}
+With this regard, the execution of a {\em Linear-bounded Turing Machine} can then be defined as a {\em plan trace} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$ such that $s_0$ encodes the initial state of the tape as well as the initial machine state and for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+
+\subsection{Learning \strips\ Models with Classical Planning}
 Here we show that the existing classical planning compilation for learning \strips\ action schemes from plan traces~\cite{aineto:learningSTRIPS:ICAPS2018} can be adapted to address the learning of CSGs.
-Note that in this case the set of preconditions and delete effects of the \strips\ action schemes is already known so the learning task reduces to learning positive effects.
+Note that in this case both the preconditions and negative effects of the \strips\ action schemes are already known so the learning task reduces to learning positive effects for these same action schemes.
 
@@
 \begin{itemize}
-\item $\mathcal{M}$, the set of {\em partially specified} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
-\item $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_{n}}$ is the {\em plan trace} obtained watching the execution of the classical plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+\item $\mathcal{M}$, the set of {\em partially specified} operator schemas.
+\item $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_{n}}$ is a {\em plan trace} obtained watching the execution of the classical plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 \item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame. 
@@
 
-A solution to a $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the given state observations $\mathcal{T}$ and the predicates $\Psi$.
+A solution to a $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the plan trace $\mathcal{T}$ and the predicates $\Psi$.
 
@@
 \begin{itemize}
-\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the input observations.
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the plan trace.
 Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
@@
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
-\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation in $\mathcal{O}$ where the action model is validated.
+\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state where the learned action model is validated.
 \end{itemize}
-\item $I_{\Lambda}$ encodes the first observation, $s_0\subseteq F$, and sets $mode_{prog}$ to true.
-Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
-Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
-
-\item $G_{\Lambda}=\bigcup_{1\leq i\leq |\mathcal{O}|}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
+\item $I_{\Lambda}$ encodes the first observation, $s_0\subseteq F$, the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that encode the {\em partially specified} operator schemas and sets $mode_{prog}$ to true.
+\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in all de states of the input trace.
 \item $A_{\Lambda}$ comprises three kinds of actions:
 \begin{enumerate}
-\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$ that add a {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
+\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$.
+For our purpose we only need to define the actions that add a {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
 \begin{small}
@@
 
-\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{O}$, {\tt\small $1\leq i< |\mathcal{O}|$}.
+\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{T}$, {\tt\small $1\leq i< n$}.
 \begin{small}
@@
 
+\subsection{Learning CSGs with Classical Planning}
+Now we show that the learning of CSGs is a particular case of the $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ such that:
+\begin{itemize}
+\item $\mathcal{M}$ is the set of {\em partially specified} operator schemas that encodes the rules of a {\em Linear-bounded Turing Machine} $M$.
+\item $\mathcal{T}$ represents the execution of the {\em Linear-bounded Turing Machine} $M$ on a given string.
+\item $\Psi=\{${\tt(head ?x),(next ?x1 ?x2),(symbol-$\sigma$ ?x),(state-q)}$\}$ . 
+\end{itemize}
+
+When learning a CSGs the {\em plan trace} $\mathcal{T}$ is partially observable:
+\begin{enumerate}
+\item The states $s_i$ with $1\leq i\leq n$ do not contain the fluents $(state-q)$.
+\item The exact executed actions $a_i$ are unknown while it is known that shoud be an action built instantiaing the schema corresponding to the transition rule $\sigma,q\rightarrow \sigma',{r,l},q'$ provided that $\sigma$ is the symbol at the cell of the tape that is currently pointed by the header at state $s_i$.
+\end{enumerate}
+To handle the first source of partial observability the compilation does not require any modification because $s_0$ is fully observable since initially the machine state is always $q_0$ (likewise the machine state is always $q_n$ at the last state that corresponds to a string accepted by the learned automaton).
+The second source of partial observability can be addressed by adding extra preconditions to the apply actions so only are appliable actions built instantiaing the schema corresponding to the transition rule $\sigma,q\rightarrow \sigma',{r,l},q'$ provided that $\sigma$ is the symbol at the cell of the tape that is currently pointed by the header at state $s_i$.
+
 \begin{lemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$.
 \end{lemma}
@@
 Completeness.
-Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T},\Pi}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
 \end{lemma}
@@
 By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{T},\Pi$.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{T}$.
 \end{small}
@@
 This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
-\item The number of given state observations. The larger $|\mathcal{T}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+\item The number of given plan trace. The larger $|\mathcal{T}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
 \end{itemize}

### 2018-04-30 doc/aij18/aij18.tex
paper aij18 abstract updates

@@
 
-\title{Learning \strips\ Action Models with Classical Planning}
+\title{Learning and Evaluation of \strips\ Action Models with Classical Planning}
 \author[label1]{Diego Aineto}
@@
   This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning.
-The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state), sequences of state observations or just a set of initial and final states (where no intermediate action or state is a priori known).
+The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state) to sequences of state observations or even just a set of initial and final states (where no intermediate action/state is known).
 The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
-What is more, the compilation is extensible to assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions.
-This is relevant since it allows us to assessing the learned models with respect to the true models but with respect to test sets of observations of plan executions, which is a necessary step for the task of model recognition.
-The empirical performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC).
+What is more, the compilation is extensible to assess how well a given \strips\ action model matches an observation of a plan execution.
+This extension is relevant since it allows us to evaluate the quality of the learned models with respect to the true models but also with respect to test sets of observations of plan executions, which is useful when the actual models are not available.
+The empirical performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and using these two evaluation approaches.
 \end{abstract}
@@
 \begin{keyword}
-Classical planning\sep Learning action models\sep Generalized planning\sep Model recognition
+Classical planning\sep Learning action models\sep Generalized planning
 %% keywords here, in the form: keyword \sep keyword

### 2018-04-30 doc/aij18/introduction.tex
intro updates aij18 paper

@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model can satisfy all the observations of plan executions that are given as input.
+A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model satisfies the observed plan executions that are given as input.
 
@@
 Moreover, it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
-Apart from these, the compilation approach presents the following contributions:
+Apart from these, our classical planning compilation presents the following contributions:
 \begin{enumerate}
-\item {\em Flexibility}.
+\item {\em Input Flexibility}.
 The compilation is flexible to various amount and kind of available input knowledge.
 Learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
-Further, the compilation is robust to partially obervability of the intermediate and final states.
-Last but not least, the compilation also accepts previous knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known.
+This fact makes the compilation robust to partially obervability of the intermediate input data.
+Last but not least, the compilation also accepts input knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known.
 % Also accepts background knowledge
+
+\item {\em Model Validation}.
+The compilation poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations.
+This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since we do not require a full plan or a full action model.
 
@@
 The compilation can assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions.
-This allows to assess learning performance without having the actual model and, like in the leaning task, the evaluation of strips model with our compilation is flexible to various amount and kind of available input knowledge.
-Further, our compilation is extensible to accept a learned model as input besides the observations of plan executions to transform the input model into a new model that induces the observations whilst assessing the amount of edition required by the input model to induce the given observations.
-
-\item {\em Model Recognition}.
-Our mechanism for model evaluation allows us to define the model recognition task.
-This task is relevant since different generative models like policies, programs, grammars or different forms of domain-specific control knowledge are STRIPS compilable.
-In this sense, our work poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations.
-This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since we do not require a full plan or a full action model.
+This allows us to assess learning performance without having the actual model and again, evaluation with our compilation is flexible to various amount and kind of available data in the given {\em test set}.
 \end{enumerate}
@@
 \begin{itemize}
-\item A unified formulation for learning and evaluating \strips\ action models from observed plans but also from state observations.
-\item The formulation of {\em model recognition} as the task of selecting, from a given set, the action model that best matches a {\em test set} with observations of plan executions.
+\item The formulation of {\em model evaluation} as the task of assessing the amount of edition required by the input action model to induce given observations of plan execution that serve as a test set.
+\item A unified formulation for learning and evaluating \strips\ action models from observed executions of plans.
 \item Leveraging {\em background knowledge} (given as planning constraints either in the form of {\em state constraints} or {\em trajectory constraints}) for learning/evaluating/recognizing \strips\ action models.
-\item A more complete empirical evaluation of the compilation approach.
-Our evaluation analyses how the amount of input knowledge and how {\em partial state observability} (some of the fluents either with {\em positive} or {\em negative} value of the intermediate states are missing because they cannot be observed) affect to the performance of the compilation approach.
+\item A more complete empirical evaluation of the compilation approach. Our evaluation analyses how the amount of input knowledge affects to the performance of the compilation approach.
 \end{itemize}
@@
 Section~\ref{sec:Section2} reviews related work on learning planning action models.
-Section~\ref{sec:Section3} introduces the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the target of our learning/evaluation/recognition tasks).
+Section~\ref{sec:Section3} introduces the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the target of our learning and evaluation tasks).
 Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.

### 2018-05-01 doc/aij18/learning_task.tex
updates added to section 3 aij

@@
 \subsection{\strips\ action schemas}
-This work addresses the learning of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+This work addresses the learning and evaluation of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
 Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
@@
 
-To formalize the output of the learning task, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
+To formalize the target of the learning and evaluation tasks, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
 Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
@@
 the variable names, and that defines the elements that can appear in an action schema.
-For the {\em blocksworld}, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+In {\em blocksworld} this set contains 11 elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
@@
 \end{itemize}
-Therefore, given the set of predicates $\Psi$ and the header of the operator schema $\xi$, $2^{2|F_v(\xi)|}$ defines the size of space of the possible \strips\ models for that operator, given that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
-For instance, this number is 4194304 for the blocksworld {\tt stack} operator while is only 1024 for the {\tt pickup} operator.
+Given the set of predicates $\Psi$ and the header of the operator schema $\xi$, $2^{2|F_v(\xi)|}$ defines the size of the space of possible \strips\ models for that operator.
+Note that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
+For instance, $2^{2|F_v(\xi)|}=4194304$ for the blocksworld {\tt stack} operator while for {\tt pickup} is only 1024.
 
-Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same headers so they can be built from the same set of possible elements.
-Formally, iff $head(\xi)=head(\xi')$ so it also holds that $F_v(\xi)=F_v(\xi')$.
-For instance we can claim that the {\tt stack} and {\tt unstack} blocksworld operators are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
+Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models.
+Formally, iff $pars(\xi)=pars(\xi')$ so it also holds that $F_v(\xi)=F_v(\xi')$.
+For instance we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
 Likewise we say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.
 
-Learning \strips\ action models from fully available input knowledge, i.e.
-from plans where the {\em pre-} and {\em post-states} of every action are available, is straightforward~\cite{jimenez2012review}:
+\subsection{Learning \strips\ action schemas from observations of plan executions}
+The learning tasks addressed in this paper correspond to observing an agent acting in the world. This learning task is formalized as a tuple $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$:
+\begin{itemize}
+\item $\mathcal{M}$ is the set of {\em empty} operator schemas.
+\item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame. 
+\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace that contains the sequence of {\em state observations} obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+Note that $\mathcal{M}$ can be inferred from $\mathcal{T}$ provided that actions in $\mathcal{T}$ are {\em diverse} enough.
+Meaning that $\mathcal{T}$ contains at least one instantiation of every action scheme in $\mathcal{M}$.
+Likewise $\Psi$ can be inferred from $\mathcal{T}$ when at least the observation of a state $s\in \mathcal{T}$ is a full state, that is $|s|=|F|$.
+\end{itemize}
+A solution to a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the predicates $\Psi$, and the observed plan trace $\mathcal{T}$.
+
+When all the elements in $\mathcal{T}$ are fully observed, i.e.
+learning from plans where the {\em pre-} and {\em post-states} of every action are available, solving the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is straightforward~\cite{jimenez2012review}:
 \begin{itemize}
@@
 \end{itemize}
-This section formalizes more challenging learning tasks where less input knowledge is available for instance, because it cannot be observed.
 
-\subsection{Learning from observations of plan executions}
-The first learning task corresponds to observing an agent acting in the world but watching only the states that result of its actions, the actual executed actions are unobserved.
-This learning task is formalized as a tuple $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$:
-\begin{itemize}
-\item $\mathcal{M}$, the set of {\em empty} operator schemas, wherein each $\xi\in\mathcal{M}$ is only composed of $head(\xi)$.
-\item $\mathcal{O}=\tup{s_0,s_1,\ldots,s_{n}}$ is the sequence of {\em state observations} obtained watching the execution of an {\em unobserved} plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-We assume that the initial state $s_0$ is {\em fully observable} while states $s_i$ s.t.
-{\small $1\leq i\leq |\mathcal{O}|$} can be {\em partially observable}, meaning that some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative.
-In the extreme, states $s_i$ {\small $1\leq i\leq |\mathcal{O}|$}, can be missing but we assume that any state observation is {\em noiseless}, meaning that if the value of a fluent is observed it is correct.
-\item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame.
-Note that $\Psi$ can be inferred from the state observations provided that at least a state $s\in \mathcal{O}$ is a full state, that is $|s|=|F|$.
-\end{itemize}
+In this paper we make the following assummptions over a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task:
+\begin{enumerate}
+\item Each input action scheme $\xi\in\mathcal{M}$ is only composed of $head(\xi)$,
+\item The initial state $s_0\in\mathcal{T}$ is {\em fully observable}.
+The remaining states $s_i$ s.t.
+{\small $1\leq i\leq n$} can be {\em partially observable} (some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative).
+In the extreme, states $s_i$ and actions $a_i$ {\small $1\leq i\leq n$}, can be missing.
+In such scenario solving $\Lambda$ implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model.
+\item Any state observation $s\in\mathcal{T}$ is {\em noiseless}, meaning that if the value of a fluent is observed it is correct. 
+\end{enumerate}
 
-A solution to a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the given state observations $\mathcal{O}$ and the predicates $\Psi$.
-Solving $\Lambda$ implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model.
-Figure~\ref{fig:example-observations} shows an example of a $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ task for learning the {\em blocksworld} \strips\ action model from the five-state observations sequence obtained inverting a 2-block tower.
-Learning the action model from this example implies inferring the unobserved plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$.
+Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, that corresponds to observing the execution of a four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
+In this example $\mathcal{T}=\langle s_0,${\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$,s_4\rangle$ which means that only the first and last states are observed and the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
 
 \begin{figure}[hbt!]
-{\footnotesize\tt ;;;;;; Headers in $\mathcal{M}$}
+{\footnotesize\tt ;;;;;; Action headers in $\mathcal{M}$}  
 \begin{footnotesize}  
@@
 \vspace{0.2cm}
-{\footnotesize\tt ;;;;;; Observations $\mathcal{O}$}
+{\footnotesize\tt ;;;;;; Plan trace $\mathcal{T}$}
 \begin{footnotesize}
 \begin{verbatim}
-;;; observation #0
+;;; state observation #0
 (clear B) (on B A) (ontable A) (handempty)
 
-;;; observation #1
-(holding B) (clear A) (ontable A)
-
-;;; observation #2
-(clear A) (ontable A) (clear B) (ontable B) (handempty)
-
-;;; observation #3
-(holding A) (clear B) (ontable B)
-
-;;; observation #4
-(clear A) (on A B) (ontable B) (handempty)
-\end{verbatim}
-\end{footnotesize}
- \caption{\small Task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ for learning a \strips\ action model in the {\em blocksworld} from a sequence of five state observations.}
-\label{fig:example-observations}
-\end{figure}
-
-In some cases we may not require to start learning from scratch because some preconditions and/or some positive or negative effects may be a priori known.
-The operator schemas in $\mathcal{M}$, that are given as input, may be not {\em empty} but {\em partially specified}.
-Such scenario is relevant for {\em policy learning}.
-A policy is function that maps states into actions and represents the conditions under which actions should be applied to achieve certain goals.
-Given an action model, the task of learning a policy that is compliant with a set of observations can be defined as learning extra preconditions for each action scheme in the model.
-These extra preconditions capture when actions can be applied according to the policy.
-A policy that exactly defines a single applicable action for each reachable state is a {\em full policy} otherwise, we say it is a {\em partial policy}.
-In the general case, learning a {\em full policy} for an arbitrary planning task is complex because a given action scheme may be applicable in different situations and also because the compact representation of this set of different situations may require the computation of {\em high-level state features}~\cite{lotinac2016automatic}.
-
-The $\Lambda$ learning task can also be redefined to cover the scenario where some of the actions executed by the observed agent are available.
-If all the executed actions are known then states observations should be partial otherwise, the learning task is trivial.
-The learning task is now formalized as $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, where:
-\begin{itemize}
-\item The plan $\pi=\tup{a_1, \ldots, a_n}$, is the action sequence that produces the sequence of state observations given in $\mathcal{O}$.
-Again we assume that action observations are noiseless, meaning that if the value of an action is observed it is correct.
-When the input plan is {\em diverse} enough, i.e.
-$\pi$ contains at least one ground action for each of the aimed action schemes, the set of {\em empty} operator schemas $\mathcal{M}$ can be inferred from $\pi$.
-\end{itemize}
-
-Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, that corresponds to observing the execution of a four-action plan for inverting a two-block tower.
-In this example $\mathcal{O}=\tup{s_0,s_4}$ which means that only the first and last states are observed and the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
-$\mathcal{M},\Psi$ are skipped, since they are the same as in Figure~\ref{fig:example-observations}.
-
-\begin{figure}[hbt!]
-{\footnotesize\tt ;;;;;; Observations $\mathcal{O}$}
-\begin{footnotesize}
-\begin{verbatim}
-;;; observation #0
-(clear B) (on B A) (ontable A) (handempty)
-
-;;; observation #4
+;;; state observation #4
 (clear A) (on A B) (ontable B) (handempty)
@@
 \end{footnotesize}
-
- \caption{\small Task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for learning a {\em blocksworld} \strips\ action model from a four-action plan and two state observations.}
+ \caption{\small Task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ for learning a {\em blocksworld} \strips\ action model from a four-action plan and two state observations.}
 \label{fig:example-plans}
@@
 
-The previous definitions formalize the learning of \strips\ action models from the observation of a single plan execution.
-These definitions are extensible to the more general case where the execution of multiple plans is observed.
-The task is defined as $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O'},\Pi}$, where $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ is the given sequence of example plans producing the corresponding sequence of state observations $\mathcal{O'}$.
-Now $\mathcal{O'}=\tup{s_0^1,s_1^1,\ldots,s_{n}^1,\ldots,s_0^t,s_1^t,\ldots,s_{n}^t\ldots,s_0^{\tau},s_1^{\tau},\ldots,s_{n}^{\tau}}$ is a sequence of {\em state observations} obtained watching the execution of a serie of {\em unobserved} plans $\pi^t=\tup{a_1, \ldots, a_n^t}$, {\tt\small $1\leq t\leq \tau$}, one after the other.
-This means that, for each {\small $1\leq i\leq n^t$}, $a_i^t\in \pi^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+The previous definition formalized the learning of \strips\ action models from the observation of a single plan execution.
+This definition is however extensible to the more general case where the execution of multiple plans is observed.
+In this case, $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ where each $t\in \mathcal{T}$ is a plan trace $t=\tup{s_0^t,a,_1^t,s_1^t,\ldots,a_n^t,s_{n}^t}$ such that, for each {\small $1\leq i\leq n^t$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.

### 2018-05-01 doc/aij18/compilation.tex
updates to section 4 aij18 paper

@@
 
-AQUI ES DONDE TENEMOS QUE HABLAR DE LOS CONDITIONAL EFFECTS QUE SOLO AFECTAN A LA COMPILACION
+Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
+A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plan that explains the given inputs to the learning task.
+The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
+
+\begin{enumerate}
+\item {\bf Programs the action model $\mathcal{M}'$}.
+A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item {\bf Validates the action model $\mathcal{M}'$}.
+The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) with the programmed action model $\mathcal{M}'$.
+\end{enumerate}
+
+\subsection{Conditional effects}
+Conditional effects allow us to compactly define actions whose effects depend on the current state.
+Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
 
@@
 
-====================
-
-Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
-A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plans that explains the given inputs to the learning task.
-The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
-
-\begin{enumerate}
-\item {\bf Programs the action model $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item {\bf Validates the action model $\mathcal{M}'$}.
-The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) with the programmed action model $\mathcal{M}'$.
-\end{enumerate}
-
-\subsection{Learning from state observations}
+\subsection{Learning from from observations of plan executions}
 Here we formalize the compilation for learning \strips\ action models with classical planning.
-Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
 \begin{itemize}
@@
 \item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the input observations, i.e.
-the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-observations}.
+the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-plans}.
 Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
-\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq |\mathcal{O}|}$, indicating the observation in $\mathcal{O}$ where the action model is validated.
+\item Fluents $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that corresponds to action $a_j$.
+The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to iterate through the steps of plan $\pi$
+  \item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state observation $s_i\in\mathcal{T}$ where the action model is validated.
 \end{itemize}
-\item $I_{\Lambda}$ encodes the first observation, $s_0\subseteq F$, and sets $mode_{prog}$ to true.
+\item $I_{\Lambda}$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<|\pi|$}, for tracking the plan step where the action model is validated.
 Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
@@
 
-\item $G_{\Lambda}=\bigcup_{1\leq i\leq |\mathcal{O}|}\{test_i\}$, requires that the programmed action model is validated in all the input observations.
+\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in all the input state observations $s_i\in\mathcal{T}$.
 \item $A_{\Lambda}$ comprises three kinds of actions:
@@
 \end{figure}
+The actions for {\em applying} an already programmed operator includes the following extra conditional effects $\{at_{j},plan(name(a_j),\Omega^{ar(a_j)},j)\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
+This mechanism ensures that $\mathsf{apply_{\xi,\omega}}$ actions are applied, exclusively, in the same order as in $\mathcal{T}$.
 
-\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{O}$, {\tt\small $1\leq i< |\mathcal{O}|$}.
+\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{T}$, {\tt\small $1\leq i< n$}.
 \begin{small}
@@
 
-The compilation approach is flexible to various amount and kind of input data.
-If any reference to the $mode_{val}$ fluent is removed the compilation can learn \strips\ action models when there are missing state observations in $\mathcal{O}$.
-On the other hand, if the $mode_{val}$ fluent is ignored, $P_{\Lambda}$ becomes harder since the classical planner must determine how many {\em apply} actions are necessary between any two observations, i.e.
-between the application of two {\em validate} actions.
 Further, known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
@@
 
-Figure~\ref{fig:plan-observations} illustrate how our compilation works and shows a plan that solves the classical planning task that results from our compilation.
-This plan programs and validates the $stack$ schema (from {\em blocksworld}) using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.
-Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,21]$ validate the programmed operators using the sequence of five state observations shown in the Figure~\ref{fig:example-observations}.
-
-\begin{figure}[hbt!]
-{\footnotesize\tt
-     {\bf 00} : (program\_pre\_clear\_stack\_v1)\\
-     01 : (program\_pre\_handempty\_stack)\\
-     02 : (program\_pre\_holding\_stack\_v2)\\
-     03 : (program\_pre\_on\_stack\_v1\_v1)\\
-     04 : (program\_pre\_on\_stack\_v1\_v2)\\
-     05 : (program\_pre\_on\_stack\_v2\_v1)\\
-     06 : (program\_pre\_on\_stack\_v2\_v2)\\
-     07 : (program\_pre\_ontable\_stack\_v1)\\
-     08 : (program\_pre\_ontable\_stack\_v2)\\
-     {\bf 09} : (program\_eff\_clear\_stack\_v1)\\
-    10 : (program\_eff\_clear\_stack\_v2)\\
-    11 : (program\_eff\_handempty\_stack)\\
-    12 : (program\_eff\_holding\_stack\_v1)\\
-    13 : (program\_eff\_on\_stack\_v1\_v2)\\
-    {\bf 14} : (apply\_unstack blockB blockA)\\
-    15 : (validate\_1)\\
-    16 : (apply\_putdown blockB)\\
-    17 : (validate\_2)\\
-    18 : (apply\_pickup blockA)\\
-    19 : (validate\_3)\\
-    20 : (apply\_stack blockA blockB)\\
-    21 : (validate\_4)
-}
- \caption{\small Plan for programming and validating the $stack$ schema using the five state observations shown in Figure~\ref{fig:example-observations} as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
-\label{fig:plan-observations}
-\end{figure}
-
-\subsection{Learning from plans}
-Now we extend the compilation to consider observed actions.
-Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$, the compilation outputs a classical planning task $P_{\Lambda'}=\tup{F_{\Lambda'},A_{\Lambda'},I_{\Lambda'},G_{\Lambda}}$ such that:
-\begin{itemize}
-\item $F_{\Lambda'}$ extends $F_{\Lambda}$ with the $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ fluents to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that corresponds to action $a_j$.
-The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to iterate through the steps of plan $\pi$.
-\item $I_{\Lambda'}$ extends $I_{\Lambda}$ with fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<|\pi|$}, for tracking the plan step where the action model is validated.
-Goals are as in the original compilation.
-In other words, the observed states are used for validation adding a $test_t$, ${\small 1\leq t\leq |\mathcal{O}|}$ fluent for each observation in $\mathcal{O}$.
-\item With respect to the set of actions $A_{\Lambda'}$.
-\begin{enumerate}
-\item The actions for {\em programming} the preconditions/effects of a given operator schema $\xi\in\Xi$ and the actions for {\em validating} the programmed action model in a given state observation are the same as in the previous compilation.
-\item The actions for {\em applying} an already programmed operator includes the following extra conditional effects $\{at_{j},plan(name(a_j),\Omega^{ar(a_j)},j)\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism ensures that $\mathsf{apply_{\xi,\omega}}$ actions are applied, exclusively, in the same order as in the example plan $\pi$ while it is flexible to the learning scenario where some $a_j$ actions in $\pi$ are missing.
-In such scenario $P_{\Lambda'}$ becomes harder because the planner must determine which and how many $\mathsf{apply_{\xi,\omega}}$ actions are necessary for validating the learned action model with the given input knowledge.
-\end{enumerate}
-\end{itemize}
-
-The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\pi}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $\mathcal{O}=\tup{s_0,s_4}$ shown in Figure~\ref{fig:example-plans}.
-Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the four-action plan $\pi$ shown in the Figure~\ref{fig:example-plans}.
+The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $s_0$ and $s_4$ shown in Figure~\ref{fig:example-plans}.
+Plan steps $[0,8]$ program the preconditions of the {\tt\small stack} operator, steps $[9,13]$ program the operator effects and steps $[14,18]$ validate the programmed operators following the four-action plan shown in Figure~\ref{fig:example-plans}.
 
@@
 
-Now we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
+Last but not least we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
 Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
@@
 Finally, the initial state $I_t$ is given by the state $s_0^t$ and the plan $\pi_t$, and the goals $G_t$ are defined by the state $s_n^t$.
-Addressing the learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ with classical planning requires introducing a small modification to our compilation.
-In particular, the actions in $P_{\Lambda'}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq \tau$} reset the current state, and the current plan, and are now defined as:
+Addressing the learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ where $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ requires introducing a small modification to our compilation.
+In particular, the actions in $P_{\Lambda}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq \tau$} resets the current state, and the current plan, and are now defined as:
 \begin{small}
@@
 \begin{mylemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$.
 \end{mylemma}
@@
 In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{O}$, starting from the corresponding initial state and following the sequence of actions defined by the plans in $\Pi$.
+These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{T}$, starting from the corresponding initial state and following the sequence of actions defined by the plans in $\Pi$.
 This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
@@
 Completeness.
-Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{O},\Pi}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
 \end{mylemma}
@@
 By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{O},\Pi$.
+The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{T}$.
 \end{small}
@@
 This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
-\item The number of given state observations. The larger $|\mathcal{O}|$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+\item The number of given state observations. The larger $n$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
 \end{itemize}
 
-\subsection{Exploiting static predicates to optimize the compilation}
+\subsection{Exploiting background knowledge to optimize the compilation}
+A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn explanations from data~\cite{muggleton1994inductive}.
+Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case to optimize the performance of the $P_{\Lambda}$ compilation.
+
+\subsection{Static predicates}
 A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
@@
 
-Static predicates can also constrain the space of possible preconditions by looking at the given set of state observation $\mathcal{O}$.
+Static predicates can also constrain the space of possible preconditions by looking at the given set of state observations in $\mathcal{T}$.
 One can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\mathcal{O}$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
-For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{O}$.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\mathcal{T}$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
+For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{T}$.
 
 Furthermore looking as well at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\Pi$ and $\mathcal{O}$.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\mathcal{T}$.
 Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
 
-\subsection{Learning \strips\ action models with background knowledge}
-A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn explanations from data~\cite{muggleton1994inductive}.
-Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case, given as planning constraints, either in the form of {\em state constraints} or {\em trajectory constraints}.
+Note that the $P_{\Lambda}$ compilation is flexible to an unbound number of {\em missing states} or {\em missing actions} in the input plan trace $\mathcal{T}$.
+If any reference to the $mode_{val}$ fluent is removed the compilation can learn \strips\ action models when there are missing state observations in $\mathcal{T}$.
+Likewise some of the actions in $\mathcal{T}$ may be unknown so they are not specified in $I_{\Lambda}$.
+In both cases the output planning becomes harder since the plan horizon is no longer bound (classical planning is PSPACE complete while SAT is NP) and the classical planner must determine how many {\em apply} actions are necessary between any two state observations, i.e.
+between the application of two {\em validate} actions.
 
@@
 
-\subsubsection{Trajectory constraints}
-Instead of enumerating the full sequence of states included in a trajectory, {\em state trajectory constraints} can be implicitly defined with {\em Linear Temporal Logic} (LTL)~\cite{haslum:LTL:ecai10,bacchus:LTL:1998}.
-LTL is a modal temporal logic interpreted over sequences of states.
-LTL interpreted over finite sequences of states (also called traces) has received attention from the planning community and provided a formal description called $LTL_f$~\cite{Giacomo:LTLf:AAAI2014}.
-
-$LTL_f$ is built up from a finite set of propositional variables $P$, the logical operators $\neg$ and $\vee$ (from which it is possible to define operators $\wedge$, $\rightarrow$ and $\leftrightarrow)$, and the temporal modal operators {\em next}($\bigcirc$), and {\em until}(${\mathcal U}$).
-An $LTL_f$ formula is then inductively defined over a set of propositions $P$:
-\begin{itemize}
-\item A proposition $p\in P$ is an $LTL_f$ formula,
-\item if $\psi$ and $\chi$ are $LTL_f$ formulae, then so they are $\neg\psi$, ($\psi\vee\chi$), $\bigcirc\psi$, ($\psi\ {\mathcal U}\ \chi$).
-\end{itemize}
-
-Given a sequence of states $\mathcal{O}=(s_0,\ldots, s_n)$, we say that a given $LTL_f$ formula $\phi$ holds at instant $i$ (denoted by $\mathcal{O}, i\models\phi$) iff:
-\begin{itemize}
-\item $\mathcal{O}, i\models f$ for a propositional variable $f\in F$, iff $f\in s_i$,
-\item $\mathcal{O}, i\models \neg\psi$ iff it is not the case that $\mathcal{O}, i\models\psi$,
-\item $\mathcal{O}, i\models \psi\wedge\chi$ iff $\mathcal{O}, i\models\psi$ and $\mathcal{O}, i\models\chi$,
-\item $\mathcal{O}, i\models \bigcirc\phi$ if $i<n$ and $\mathcal{O}, i+1\models\phi$,
-\item $\mathcal{O}, i\models (\phi_1\ {\mathcal U}\ \phi_2)$ if exists some $j\in\{1,\ldots,n\}$ such that, $\tau, j\models \phi_2$ and for all $k\in\{i,\ldots,j-1\}$ it holds that $\tau, k\models \phi_1$.
-\end{itemize}
-We say that $\mathcal{O}$ is {\em valid} for a given $LTL_f$ formula $\phi$ (denoted by $\mathcal{O}\models\phi$) iff for every {\small $0\leq i\leq n$} holds that $\mathcal{O}, i\models\phi$.
-From the previous basic operators it is also possible to define the abbreviation $last$ (that denotes the last instant of a sequence) and the temporal modal operators {\em eventually} ($\lozenge$), {\em always} ($\square$), and {\em release} (${\mathcal R}$):
-\begin{itemize}
-\item $last$ is shorthand for $\neg\bigcirc true$ and holds only at the last state of the sequence.
-The achievement of classical planning goals $G$ can then be expressed as the $LTL_f$ formula $\phi=last\wedge G$.
-\item $\lozenge$ $\psi$ stands for $(true\ {\mathcal U}\ \psi)$, and says that $\psi$ will eventually hold before the last state (last state included).
-\item $\square\psi$ stands for $\neg \lozenge \neg\psi$ and says that from the current state till the last one the formula will always hold.
-\item $\psi\ {\mathcal R}\ \chi$ stands for $\neg(\neg\psi\ {\mathcal U}\ \neg\chi)$.
-This says that $\chi$ has to be true until and including the point where $\psi$ first becomes true; if $\psi$ never becomes true, $\chi$ must remain true forever.
-\end{itemize}
-
-$LTL$ interpreted over finite traces has the expressive power of First Order Logic over finite ordered traces.
-Satisfiability, validity and logical implication (as well as model checking) for $LTL_f$ are PSPACE-complete.
-
-\subsubsection{Leveraging background knowledge}
-We have shown how our approach is flexible to validate the learned action models with respect to a set of observations of executed plans and with respect to a model that acts as a reference.
-Here we show that the validation of an action model can also be done with {\em state constraints} or {\em trajectory constraints}.
+The validation of an action model can also be done with {\em state constraints} or {\em trajectory constraints}.
 Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check that the learned model satisfy every constraint $\phi_i\in\Phi$:

### 2018-05-01 doc/aij18/compilation.tex
updates added to sections 5 and 6 AIJ18 paper

@@
 
+Note that the $P_{\Lambda}$ compilation is flexible to an unbound number of {\em missing states} or {\em missing actions} in the input plan trace $\mathcal{T}$.
+If any reference to the $mode_{val}$ fluent is removed the compilation can learn \strips\ action models when there are missing state observations in $\mathcal{T}$.
+Likewise some of the actions in $\mathcal{T}$ may be unknown so they are not specified in $I_{\Lambda}$.
+In both cases the output planning becomes harder since the plan horizon is no longer bound (classical planning is PSPACE complete while SAT is NP) and the classical planner must determine how many {\em apply} actions are necessary between any two state observations, i.e.
+between the application of two {\em validate} actions.
+
 Last but not least we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
@@
 
-\subsection{Exploiting background knowledge to optimize the compilation}
-A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn explanations from data~\cite{muggleton1994inductive}.
+\subsection{Optimizing the compilation with background knowledge}
+A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn logic programs from data~\cite{muggleton1994inductive}.
 Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case to optimize the performance of the $P_{\Lambda}$ compilation.
 
-\subsection{Static predicates}
+\subsubsection{Static predicates}
 A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
@@
 Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
-
-Note that the $P_{\Lambda}$ compilation is flexible to an unbound number of {\em missing states} or {\em missing actions} in the input plan trace $\mathcal{T}$.
-If any reference to the $mode_{val}$ fluent is removed the compilation can learn \strips\ action models when there are missing state observations in $\mathcal{T}$.
-Likewise some of the actions in $\mathcal{T}$ may be unknown so they are not specified in $I_{\Lambda}$.
-In both cases the output planning becomes harder since the plan horizon is no longer bound (classical planning is PSPACE complete while SAT is NP) and the classical planner must determine how many {\em apply} actions are necessary between any two state observations, i.e.
-between the application of two {\em validate} actions.
 
@@
 
-The validation of an action model can also be done with {\em state constraints} or {\em trajectory constraints}.
-Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check that the learned model satisfy every constraint $\phi_i\in\Phi$:
+An interesting contribution of our compilation is that the validation of an action model can also be done with {\em state constraints}.
+Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check that the learned model satisfy every constraint $\phi\in\Phi$:
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
 \cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.

### 2018-05-01 doc/aij18/evaluation.tex
updates added to sections 5 and 6 AIJ18 paper

@@
 
+If the actual model is available, the quality of a learned model is quantifiable with the {\em precision} and {\em recall} metrics.
+These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
+Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
+\begin{itemize}
+\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
+\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing).
+\end{itemize}
+
 The roles of two action schemes whose headers match or the roles of two action parameters that belong to the same type can be swapped by a $\Lambda$ learning task.
 For instance, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped.
-Pure syntax-based metrics can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
+Pure syntax-based evaluation metrics (like {\em precision} and {\em recall}) can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
 a learned model semantically equivalent but syntactically different to the reference model.
@@
 The intuition of the approach is to assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
-%In the extreme, if $\mathcal{M}$ perfectly explains $\mathcal{O}$, no model {\em edition} is necessary.
+Like our learning approach, our evaluation approach is also flexible to various amount and kind of available input knowledge.
 
@@
 As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions~\cite{yang2007learning}.
-With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a sequence of state observations.
+With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 
 \begin{mydefinition}
-  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and the observations sequence $\mathcal{O}=\tup{s_0, s_1, \ldots, s_n}$ such that each state observation in $\mathcal{O}$ is built with fluents in $F$.
-The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan $\pi=\tup{a_1, \ldots, a_n}$ that induces $\mathcal{O}$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
+  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and a plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ whose state observation are built with fluents in $F$.
+The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan trace $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{mydefinition}
@@
 Unlike the error function defined by {\sc ARMS}, the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
-This fact enables the recognition of \strips\ action models.
-The idea, taken from {\em plan recognition as planning}~\cite{ramirez2009plan}, is to map distances into likelihoods.
-This {\em edit distance} could be mapped into a likelihood with the following expression $P(\mathcal{O}|\mathcal{M})=1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
-
-The error of a learned action model could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
+The {\em edit distance} could be mapped into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
+Note that the error of a learned action model could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
 This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes.
 Our definition of the edit distance is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
-Finally, the {\em edit distance} can also be defined with respect to a set of plans, if they are available.
-
-\begin{mydefinition}
-  Given an action model $\mathcal{M}$ and a set of valid plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$ that only contain actions built grounding the schemes in $\mathcal{M}$.
-The {\bf plans edit distance}, denoted by  $\delta(\mathcal{M},\Pi)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce all the plans $\pi_t\in \Pi$, {\tt\small $1\leq t\leq \tau$}; \[\delta(\mathcal{M},\Pi)=\min_{\forall \mathcal{M}' \rightarrow \Pi} \delta(\mathcal{M},\mathcal{M}')\]
-\end{mydefinition}
 
 \subsection{Computing the observations and plans edit distance}
-Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, is {\em non-empty}.
+Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, is {\em non-empty}.
 In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
@@
 However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$ in observations of the plan executions}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{O}$ (and on $\Pi$ if available), as explained in Section~\ref{sec:Section5} for the models that are programmed from scratch.
+\item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:Section5} for the models that are programmed from scratch.
 \end{enumerate}
 
-Now $\Lambda$ does not formalize a learning task but the task of editing $\mathcal{M}$ to produce the observations $\mathcal{O}$, which results in the edited model $\mathcal{M}'$.
+Now $\Lambda$ does not formalize a learning task but the task of editing $\mathcal{M}$ to produce the plan trace $\mathcal{T}$, which results in the edited model $\mathcal{M}'$.
 The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
@@
 
-To illustrate this, the plan of Figure~\ref{fig:plan-odistance} solves the classical planning task that corresponds to editing a \emph{blocksworld} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-First, the plan edits the {\tt\small stack} schema, {\em inserting} these two positive effects, and then validates the edited action model in the five-observation sequence of Figure~\ref{fig:example-observations}.
-
-\begin{figure}[hbt!]
-{\tt\small
-00 : (insert\_add\_handempty\_stack)\\
-01 : (insert\_add\_clear\_stack\_var1)\\
-02 : (apply\_unstack blockB blockA)\\
-03 : (validate\_1)\\
-04 : (apply\_putdown blockB)\\
-05 : (validate\_2)\\
-06 : (apply\_pickup blockA)\\
-07 : (validate\_3)\\
-08 : (apply\_stack blockA blockB)\\
-09 : (validate\_4)\\
-}
-\caption{\small Plan for editing a given action model and validating it at the state observations shown in Figure~\ref{fig:example-observations}.
-The given action model is a four operator {\em blockswold} where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.}
-\label{fig:plan-odistance}
-\end{figure}
-
-Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
-$\delta(\mathcal{M},\mathcal{O})=2$ for the example in Figure~\ref{fig:plan-odistance}.
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
-The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
-
-When the executed plans $\Pi$ are also available, the compilation can be adapted to compute the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
-The modifications to the compilation explained in Section~\ref{sec:Section5} are also useful here to redefine the learning task as the task of editing $\mathcal{M}$ to produce the set of plans $\Pi$, which results in the edited model $\mathcal{M}'$.
-Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+To illustrate this, the plan of Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
 In this case the edited action model is however validated at the plan shown in Figure~\ref{fig:example-plans}.
@@
 \end{figure}
+
+Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
+$\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}.
+In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
+The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
 
@@
 \end{figure}
-
-\subsection{Recognition of \strips\ action models}
-\label{sec:Section7}
-Given a set of possible \strips\ models and set of observations of plan executions, the {\em recognition of \strips\ models} is the task of identifying which model has the highest probability of producing the given observations.
-
-According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the set of possible \strips\ action models.
-Given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section3}.
-With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the size of the possible state observations of length $n$, that is $\mathcal{O}=s_0,\ldots,s_n$ is given by $2^{n\times|F|}$.
-
-With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
-\begin{enumerate}
-\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$.
-If a set of plans $\Pi$ is available, this same strategy can be followed using the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
-\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
-\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
-\end{enumerate}

### 2018-05-01 doc/aij18/experiments.tex
updates added to sections 5 and 6 AIJ18 paper

@@
 Here we evaluate the learned models with respect to the actual generative model.
-Opposite to what usually happens in ML, this model is available when learning is applied to IPC domains.
-The quality of the learned models is quantified with the {\em precision} and {\em recall} metrics.
-These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
-Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
-\begin{itemize}
-\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
-\item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing).
-\end{itemize}
 
 \subsubsection{Learning from plans}
-We start evaluating our approach with $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi,\Pi}$ learning tasks, where the action of the executed plans are available and the observation sequence contains only the corresponding initial and goal states, $s_o^t$ and $s_n^t$, for every plan $\pi_t\in\Pi$.
-We then repeat the evaluation but exploiting potential \emph{static predicates} computed from the observed states in $\mathcal{O}$, which are the predicates that appear unaltered in the states that belong to the same plan.
+We start evaluating our approach with $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning tasks, where the action of the executed plans are available and the state observation sequence contains only the corresponding initial and goal states, $s_o^t$ and $s_n^t$, for every trace plan $t\in\mathcal{T}$.
+We then repeat the evaluation but exploiting potential \emph{static predicates} computed from the observed states in $\mathcal{T}$, which are the predicates that appear unaltered in the states that belong to the same plan.
 Static predicates are used to constrain the space of possible action models as explained in Section~\ref{sec:Section5}.
@@
 \subsubsection{Learning from state observations}
-Here we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, where the action of the executed plans are not available but the initial and goal states are known.
+Here we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, where the action of the executed plans are not available but the initial and goal states are known.
 When input plans are not available, the planner must not only compute the action models but also the plans that satisfy the input observations.
@@
 
-Now we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\mathcal{O},\Psi}$, where the action of the executed plans are not available but where the observation sequence $\mathcal{O}$ contains all the intermediate states, not just the initial and final states.
+Now we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, where the action of the executed plans are not available but where the plan trace $\mathcal{T}$ contains all the intermediate states, not just the initial and final states.
 Table~\ref{fig:observationsnomap} shows the precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.

### 2018-05-02 doc/aij18/aij18.tex
appendix added to aij18 paper
updates added to sections 5 and 6 AIJ18 paper
updates to the aij18 paper

@@
   This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning.
-The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state) to sequences of state observations or even just a set of initial and final states (where no intermediate action/state is known).
+The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state) to sequences of state observations or even just a set of initial and final states (where no intermediate action or state is known).
 The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
-What is more, the compilation is extensible to assess how well a given \strips\ action model matches an observation of a plan execution.
-This extension is relevant since it allows us to evaluate the quality of the learned models with respect to the true models but also with respect to test sets of observations of plan executions, which is useful when the actual models are not available.
-The empirical performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and using these two evaluation approaches.
+What is more, the compilation is extensible to assess how well a given \strips\ action model matches the observation of a plan execution.
+This extension allows us to evaluate the quality of the learned models with respect to the actual models but also, with respect to observations of plan executions that act as a {\em test set}, which is useful when the actual models are not available.
+The performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and using these two evaluation approaches.
 \end{abstract}
@@
 
+\input{appendix}
+
 %% Authors are advised to use a BibTeX database file for their reference list.

### 2018-05-02 doc/aij18/introduction.tex
updates to the aij18 paper

@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model satisfies the observed plan executions that are given as input.
+A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model satisfies a set of plan executions that is given as input.
 
@@
 \begin{enumerate}
-\item {\em Input Flexibility}.
+\item {\em Input flexibility}.
 The compilation is flexible to various amount and kind of available input knowledge.
 Learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
-This fact makes the compilation robust to partially obervability of the intermediate input data.
-Last but not least, the compilation also accepts input knowledge about the structure of the actions in the form of partially specified action models, where some preconditions and effects are a priori known.
+This fact makes the compilation robust to the partial obervability of the intermediate input data.
+Last but not least, the compilation also accepts input knowledge about the structure of the actions in the form of partially specified action models, that is learning action models when some preconditions and effects are a priori known.
 % Also accepts background knowledge
 
-\item {\em Model Validation}.
-The compilation poses a general framework to assess the validation of a generative model (provided that it is STRIPS compilable) with a given set of observations.
-This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since we do not require a full plan or a full action model.
+\item {\em Model validation}.
+The compilation poses a novel framework to assess the validation of a STRIPS model with respect to a given set of observations.
+This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since it does not require a full plan or a full action model to determine validation.
+The relevance of this validation framework overcomes classical planning because of the wide range of different generative models that are STRIPS compilable (STRIPS is {\em Turing complete}).
 
-\item {\em Model Evaluation}.
+\item {\em Model evaluation}.
 The compilation can assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions.
+The idea is to assess the amount of edition required by the input action model to induce given observations of plan execution that serve as a test set.
 This allows us to assess learning performance without having the actual model and again, evaluation with our compilation is flexible to various amount and kind of available data in the given {\em test set}.
@@
 
-A first description of the compilation previously appeared in the conference paper~\cite{aineto2018learning}. Compared to the conference paper, this work includes the following novel material:
+A first description of the compilation previously appeared in the conference paper~\cite{aineto2018learning}. Compared to that conference paper, this work includes the following novel material:
 \begin{itemize}
-\item The formulation of {\em model evaluation} as the task of assessing the amount of edition required by the input action model to induce given observations of plan execution that serve as a test set.
 \item A unified formulation for learning and evaluating \strips\ action models from observed executions of plans. 
-\item Leveraging {\em background knowledge} (given as planning constraints either in the form of {\em state constraints} or {\em trajectory constraints}) for learning/evaluating/recognizing \strips\ action models.
-\item A more complete empirical evaluation of the compilation approach. Our evaluation analyses how the amount of input knowledge affects to the performance of the compilation approach.
+\item Leveraging diverse forms of {\em background knowledge} for learning and evaluating \strips\ action models.
+\item A complete empirical evaluation of the compilation approach.
+Our evaluation analyses how the amount of input knowledge affects to the performance of the compilation approach when learning and evaluating \strips\ action models.
 \end{itemize}
@@
 Section~\ref{sec:Section2} reviews related work on learning planning action models.
-Section~\ref{sec:Section3} introduces the classical planning model with {\em conditional effects} (a requirement of the proposed compilation) and the \strips\ action model (the target of our learning and evaluation tasks).
+Section~\ref{sec:Section3} introduces the classical planning model and the \strips\ action model (the target of our learning and evaluation tasks).
 Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.

### 2018-05-02 doc/aij18/compilation.tex
new updates to the AIJ18 paper

@@
 Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
-A planning compilation is a suitable approach because computing a solution for $\Lambda$ involves, not only determining the \strips\ action model $\mathcal{M}'$, but also the {\em unobserved} plan that explains the given inputs to the learning task.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \item {\bf Validates the action model $\mathcal{M}'$}.
-The solution plan continues with a postfix that reproduces the given input knowledge (the available observations of the plan executions) with the programmed action model $\mathcal{M}'$.
+The solution plan continues with a {\em postfix} that reproduces the given input knowledge (the available observations of the plan executions) with the programmed action model $\mathcal{M}'$.
 \end{enumerate}
@@
 \begin{itemize}
-\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the input observations, i.e.
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the plan trace given as input, i.e.
 the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-plans}.
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
-\item Fluents $F_{\pi}=\{plan(name(a_j),\Omega^{ar(a_j)},j)\}_{\small 1\leq j\leq |\pi|}$ to code the $j^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that corresponds to action $a_j$.
-The static facts $next_{j,j+1}$ and the fluents $at_j$, {\small $1\leq j< |\pi|$}, are also added to iterate through the steps of plan $\pi$
+\item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that corresponds to action $a_i$.
+The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of plan $\pi$
   \item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state observation $s_i\in\mathcal{T}$ where the action model is validated.
 \end{itemize}
-\item $I_{\Lambda}$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{j,j+1}\}$, {\small $1\leq j<|\pi|$}, for tracking the plan step where the action model is validated.
+\item $I_{\Lambda}$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated.
 Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
@@
 \end{figure}
-The actions for {\em applying} an already programmed operator includes the following extra conditional effects $\{at_{j},plan(name(a_j),\Omega^{ar(a_j)},j)\}\rhd\{\neg at_{j},at_{j+1}\}_{\forall j\in [1,n]}$ for advancing to the next plan step.
-This mechanism ensures that $\mathsf{apply_{\xi,\omega}}$ actions are applied, exclusively, in the same order as in $\mathcal{T}$.
+These $\mathsf{apply_{\xi,\omega}}$ actions includes the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\cup \{mode_{val}\},\\
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
 \cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
@@
 
-Further, known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
+Known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
 In this case, the corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
-When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the compilation validates whether the observation of the plan execution follows the given model:
+When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the $P_{\Lambda}$ compilation validates whether the observation of the plan execution follows the given model:
 \begin{itemize}
-\item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given input data if a solution plan for $P_{\Lambda}$ can be found.
-\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given input data if $P_{\Lambda}$ is unsolvable.
+\item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given input data in $\mathcal{T}$ iff a solution plan for $P_{\Lambda}$ can be found.
+\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given input data $\mathcal{T}$ iff $P_{\Lambda}$ is unsolvable.
 This means that $\mathcal{M}$ cannot be compliant with the given observation of the plan execution.
@@
 This feature of our compilation is beyond the functionality of VAL, the plan validation tool~\cite{howey2004val}, because VAL requires (1) a full plan and (2), a full action model for plan validation.
+On the other hand our $P_{\Lambda}$ compilation can test {\em model validation} with a partial (or even an empty) action model and a partially observed plan trace.
 
-The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
+The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to the classical planning task $P_{\Lambda}$ that encodes a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
 This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $s_0$ and $s_4$ shown in Figure~\ref{fig:example-plans}.
@@
 }
- \caption{\small Plan for programming and validating the $stack$ schema using plan $\pi$ and state observations $\mathcal{O}$ (shown in Figure~\ref{fig:example-plans}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+ \caption{\small Plan for programming and validating the $stack$ schema (using plan $\pi$ and state observations shown in Figure~\ref{fig:example-plans}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
 \label{fig:plan-lplan}
@@
 Likewise some of the actions in $\mathcal{T}$ may be unknown so they are not specified in $I_{\Lambda}$.
-In both cases the output planning becomes harder since the plan horizon is no longer bound (classical planning is PSPACE complete while SAT is NP) and the classical planner must determine how many {\em apply} actions are necessary between any two state observations, i.e.
-between the application of two {\em validate} actions.
+In both cases the output planning problem becomes harder since the plan horizon is no longer bound and the classical planner must determine how many {\em apply} actions are necessary between any two state observations, i.e.
+between the application of two {\em validate} actions (classical planning is {\em PSPACE complete} while classical planning with a fixed plan horizon is {\em NP}~\cite{rintanen:SATplannning:AI2012}).
 
@@
 Addressing the learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ where $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ requires introducing a small modification to our compilation.
-In particular, the actions in $P_{\Lambda}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq \tau$} resets the current state, and the current plan, and are now defined as:
+In particular, the actions in $P_{\Lambda}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq \tau$} resets the current state, and the current plan, and are now redefined as:
 \begin{small}
@@
 In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{T}$, starting from the corresponding initial state and following the sequence of actions defined by the plans in $\Pi$.
+These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{T}$, starting from the corresponding initial state and following the sequence of actions defined by $\mathcal{T}$.
 This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
@@
 This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
-\item The number of given state observations. The larger $n$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+\item The number of given observations in $\mathcal{T}$. The larger $n$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
 \end{itemize}
 
-\subsection{Optimizing the compilation with background knowledge}
+\subsection{Tuning the compilation with background knowledge}
 A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn logic programs from data~\cite{muggleton1994inductive}.
@@
 {\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
 
@@
 An interesting contribution of our compilation is that the validation of an action model can also be done with {\em state constraints}.
-Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check that the learned model satisfy every constraint $\phi\in\Phi$:
+Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check also that the learned model satisfy a constraint $\phi\in\Phi$:
 \begin{small}
@@
 
-This redefinition of validate actions applies also to trajectory constraints because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.
+This redefinition of validate actions applies also to {\em trajectory constraints} because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.
 

### 2018-05-02 doc/aij18/evaluation.tex
new updates to the AIJ18 paper

@@
-\section{Evaluation model}
+\section{Evaluation of \strips\ Action Models with Classical Planning}
 \label{evaluation}
 
-If the actual model is available, the quality of a learned model is quantifiable with the {\em precision} and {\em recall} metrics.
+If the actual model is available, the quality of a learned action model is quantifiable using the {\em precision} and {\em recall} metrics.
 These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
@@
 
-The roles of two action schemes whose headers match or the roles of two action parameters that belong to the same type can be swapped by a $\Lambda$ learning task.
+The roles of two {\em comparable} action schemes or the roles of two action parameters that belong to the same type can be swapped by a $\Lambda$ learning task when the input data given in $\mathcal{T}$ is insufficient.
 For instance, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped.
@@
 
-Here we introduce an evaluation approach that is robust to role changes of this particular kind.
+Here we introduce a novel evaluation approach that is robust to role changes of this particular kind.
 The intuition of the approach is to assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
@@
 \end{mydefinition}
+Note that the number of deletions indicate the number of {\em false positives} while the number of insertions represents the number of {\em false negatives}.
+As explained, these numbers can be normalized with the {\em true positives} and {\em true negatives} to compute the {\em precision} and {\em recall} of a learned model.
 
@@
 
-Unlike the error function defined by {\sc ARMS}, the {\em observation edit distance} assess, with a single expression, the flaws in the preconditions and effects of a given learned model.
-The {\em edit distance} could be mapped into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
+The {\em observation edit distance} is mappable into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
+Further, following the previous correspondances ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can redefine the {\em precision} and {\em recall} of a learned model with respect to a partially observed plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
+In this case, for each action schema $\xi\in\mathcal{M}$, the corresponding number of {\em true positives} is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$ while the {\em true negatives} are given by $tn(\xi)=3\times|F_v(\xi)|-tp(\xi)$
+
 Note that the error of a learned action model could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
-This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes.
-Our definition of the edit distance is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
+This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes~\cite{sohrabi:precognition:IJCAI2016}.
+Our definition of the {\em observation edit distance} is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
 
-\subsection{Computing the observations and plans edit distance}
+\subsection{Computing the edit distance with classical planning}
 Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, is {\em non-empty}.
@@
 
-Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} for that $\mathcal{M}'$ is validated in the given observations, e.g.
+Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deleitions) for that $\mathcal{M}'$ is validated in the given observations, e.g.
 $\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}.
@@
 With respect to the actions, $A_{\Lambda}$ again implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect) but does not contain apply actions because the \strips\ action model are not validated in any observation of plan executions.
+include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect) but does not contain $\mathsf{apply_{\xi,\omega}}$ or $\mathsf{validate_{i}}$ actions because the \strips\ action model are not validated in any observation of plan executions.
 Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_v(\xi)$.

### 2018-05-02 doc/aij18/learning_task.tex
new updates to the AIJ18 paper

@@
 Note that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
-For instance, $2^{2|F_v(\xi)|}=4194304$ for the blocksworld {\tt stack} operator while for {\tt pickup} is only 1024.
+For instance, $2^{2|F_v(stack)|}=4194304$ for the blocksworld {\tt stack} operator while for {\tt pickup} is only 1024.
 
 Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models.
-Formally, iff $pars(\xi)=pars(\xi')$ so it also holds that $F_v(\xi)=F_v(\xi')$.
-For instance we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
-Likewise we say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.
+Formally, if $pars(\xi)=pars(\xi')$ it also holds that $F_v(\xi)=F_v(\xi')$.
+For instance, we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
+Two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.
 
 \subsection{Learning \strips\ action schemas from observations of plan executions}
-The learning tasks addressed in this paper correspond to observing an agent acting in the world. This learning task is formalized as a tuple $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$:
+The learning task addressed in this paper corresponds to learning an action model by observing an agent(s) acting in the world.
+This learning task is formalized as a tuple $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$:
 \begin{itemize}
-\item $\mathcal{M}$ is the set of {\em empty} operator schemas.
-\item $\Psi$ is the set of predicates that define the abstract state space of a given planning frame. 
-\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace that contains the sequence of {\em state observations} obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-Note that $\mathcal{M}$ can be inferred from $\mathcal{T}$ provided that actions in $\mathcal{T}$ are {\em diverse} enough.
+\item $\mathcal{M}$ is the set of {\em empty} operator schemas that define the actions of a given planning frame
+\item $\Psi$ is the set of predicates that define the abstract state space of that planning frame. 
+\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace that contains the sequence of {\em state observations} obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in state $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+\end{itemize}
+Note that $\mathcal{M}$ can be inferred from $\mathcal{T}$ provided that the actions in $\mathcal{T}$ are {\em diverse} enough.
 Meaning that $\mathcal{T}$ contains at least one instantiation of every action scheme in $\mathcal{M}$.
 Likewise $\Psi$ can be inferred from $\mathcal{T}$ when at least the observation of a state $s\in \mathcal{T}$ is a full state, that is $|s|=|F|$.
-\end{itemize}
+
 A solution to a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the predicates $\Psi$, and the observed plan trace $\mathcal{T}$.
@@
 
-In this paper we make the following assummptions over a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task:
+In this paper we address the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task with the following assummptions:
 \begin{enumerate}
-\item Each input action scheme $\xi\in\mathcal{M}$ is only composed of $head(\xi)$,
+\item Each input action scheme $\xi\in\mathcal{M}$ is at least composed of its $head(\xi)$,
 \item The initial state $s_0\in\mathcal{T}$ is {\em fully observable}.
-The remaining states $s_i$ s.t.
+The remaining states $s_i\in\mathcal{T}$ s.t.
 {\small $1\leq i\leq n$} can be {\em partially observable} (some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative).
-In the extreme, states $s_i$ and actions $a_i$ {\small $1\leq i\leq n$}, can be missing.
+In the extreme, entire states $s_i$ and actions $a_i$ {\small $1\leq i\leq n$}, can be missing.
 In such scenario solving $\Lambda$ implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model.
-\item Any state observation $s\in\mathcal{T}$ is {\em noiseless}, meaning that if the value of a fluent is observed it is correct. 
+\item Any state observation $s\in\mathcal{T}$ is {\em noiseless}, meaning that if the value of a fluent is observed it is the correct value. 
 \end{enumerate}

### 2018-05-03 doc/aaai-19/CSL.tex
first notes on model recognition

--- 2018-04-26 doc/aaai-19/tm-grammars.tex
@@
 
-\title{Learning Context-Sensitive Grammars with Classical Planning}
+\title{A Classical Planning Model for Learning, Parsing and Production in Context-Sensitive Languages}
 \author{\#39}
@@
 \begin{abstract} 
-This paper presents a novel approach for learning {\em Context-Sensitive Grammars} (CSGs) from small sets of input strings.
-The approach guarantees that the computed grammar generalizes provided that the input strings are long and diverse enough to cover all the grammar rules.
-Our approach is to compile this learning task into a classical planning problem whose solutions are all sequences of actions with a prefix that build the CSG and and a postfix that validates the built grammar in the input strings.
-The compilation is flexible to implement the three canonical tasks of CSGs, {\it grammar generation}, {\it string production} and {\it string recognition} within the same classical planning model.
-The experimental validation of our compilation approach reports the empirical data collected when learning the CSGs for the languages $\{a^nb^nc^n : n \geq 1 \}$ and $\{a^{2^n} : n \geq 1 \}$ and shows that overcomes the performance  of a previous compilation for learning {\em Context-Free Grammars}.
+This paper presents a novel approach for learning the acceptor automaton of a given {\em Context-Sensitive Language} (CSLs) from small sets of input strings.
+Our approach is to compile this learning task into a classical planning problem whose solutions are sequences of actions s.t.
+their prefix builds the acceptor automaton and their postfix validates that the built automaton accepts the input strings.
+The compilation is flexible to implement the three canonical tasks of Context-Sensitive Grammars (CSGs), {\it learning}, {\it parsing} and {\it production} within the same classical planning model.
+The paper reports the empirical data collected when learning the acceptor automata for the CSLs $\{a^nb^nc^n : n \geq 1 \}$ and $\{a^{2^n} : n \geq 1 \}$ and shows that our compilation outperforms a previous classical planning compilation for  learning {\em Context-Free Grammars}.
 \end{abstract}
@@
 \item {\it Learning}: Given a set of strings, compute a grammar that is compliant with the input strings.
-\item {\it Production}: Given a formal grammar, generate strings that belong to the language represented by the grammar.
 \item {\it Recognition} (also known as {\em parsing}): Given a formal grammar and a string, determine whether the string belongs to the language represented by the grammar.
+\item {\it Production}: Given a formal grammar, generate a string that belongs to the language represented by the grammar.
 \end{itemize}
@@
 Chomsky defined four types of formal grammars that differ in the form and generative capacity of their rules~\cite{chomsky2002syntactic}.
-Each grammar type generates a different class of formal language that is recognizable with a different kind of automaton: {\bf Type-0} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Turing machine}.
-{\bf Type-1} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Linear-bounded Turing machine}.
+Each grammar type generates a different class of formal language that is recognizable with a different kind of acceptor automaton: {\bf Type-0} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em Turing machine}.
+{\bf Type-1} corresponds to the {\em recursively enumerable} languages that can be recognized with a {\em linear-bounded Turing machine}.
 {\bf Type-2} corresponds to {\em context-free} languages that are recognizable with a {\em non-deterministic push-down automaton}.
-Finally, {\bf type-3} corresponds to {\em regular} languages that can be recognized with a {\em Finite state machine}.
+Finally, {\bf type-3} corresponds to {\em regular} languages that can be recognized with a {\em finite state machine}.
 
 Figure~\ref{fig:csg}(a) shows an example of a CSG grammar ({\em Type-1}) that generates the $\{a^nb^nc^n : n \geq 1 \}$ language.
-The grammar defines ten production rules, contains three {\em terminal symbols} ($a$, $b$ and $c$) and five {\em non-terminal symbols} ($S$,$B$,$C$,$W$ and $Z$).
-This CSG can generate, for instance, the string $aaabbbccc$ by applying the sequence of rules $\tup{1,1,0,2,3,4,5,2,3,4,5,2,3,4,5,6,7,7,8,9,9}$.
-The {\it parse tree} in Figure~\ref{fig:csg}(b) exemplifies this rule application and proves that the string $aaabbbccc$ belongs to the $\{a^nb^nc^n : n \geq 1 \}$ language defined by the CSG grammar.
+The grammar defines ten production rules over three {\em terminal symbols} ($a$, $b$ and $c$) and five {\em non-terminal symbols} ($S$,$B$,$C$,$W$ and $Z$).
+This CSG produces the string $aaabbbccc$ applying the sequence of rules $\tup{1,1,0,2,3,4,5,2,3,4,5,2,3,4,5,6,7,7,8,9,9}$.
+The {\it parse tree} in Figure~\ref{fig:csg}(b) exemplifies this rule application and proves that the string $aaabbbccc$ belongs to the $\{a^nb^nc^n : n \geq 1 \}$ language.
 
@@
 
-  \caption{\small (a) Example of a context-sensitive grammar; (b) the corresponding {\it parse tree} for the string $aaabbbccc$.}
+  \caption{\small (a) Example of a {\em context-sensitive grammar}; (b) the corresponding {\it parse tree} for that grammar and the string $aaabbbccc$.}
   \label{fig:csg}
@@
 
-Previous work showed that a classical planning compilation can implement the three canonical tasks (namely {\it grammar generation}, {\it string production} and {\it string recognition}) for {\em Type-3} and {\em Type-2} grammars~\cite{segovia2017generating}.
-In this work we show that a novel classical planning compilation implements these three canonical tasks for grammars of the three last three types ({\em Type-3}, {\em Type-2} and {\em Type-1}).
-In addition, this novel compilation has fewer input parameters since it does not require to specify any bound on the size of the grammar rules.
-The experimental validation of our compilation approach reports the empirical data collected when learning the CSGs for the languages $\{a^nb^nc^n : n \geq 1 \}$ and $\{a^{2^n} : n \geq 1 \}$ and shows that overcomes the performance  of a previous compilation for learning {\em Context-Free Grammars}.
+Previous work showed that a classical planning compilation can implement {\it learning}, {\it parsing} and {\it production} for {\em Type-3} and {\em Type-2} languages~\cite{segovia2017generating}.
+In this work we show that a novel classical planning compilation implements these three canonical tasks for {\em Type-3}, {\em Type-2} and {\em Type-1} languages.
+In addition, this novel compilation has fewer input parameters since it does not require to bound the size of the grammar rules.
+The evaluation of our compilation approach reports the empirical data collected when learning the acceptor automata for the CSLs $\{a^nb^nc^n : n \geq 1 \}$ and $\{a^{2^n} : n \geq 1 \}$ and shows that it outperforms of a previous compilation for {\it learning}, {\it parsing} and {\it production} in {\em Context-Free Languages}.
 
@@
 \begin{itemize}
-\item $V$ is the finite set of non-terminal symbols, also called variables. With $v_0\in V$ is the start non-terminal symbol that represents the whole grammar.
-\item $\Sigma$ is the finite set of terminal symbols, which are disjoint from the set of non-terminal symbols, i.e.~$V\cap \Sigma\neq\emptyset$.
-The set of terminal symbols is the alphabet of the language defined by $\mathcal{G}$
-\item Rules in $R$ are defined as $\alpha v\beta\rightarrow \alpha\gamma\beta$ where $v\in V$, $\alpha,\beta\in(V\cup\Sigma)^*$ and $\gamma\in(V\cup\Sigma)^+$.
-That is, the left-hand and right-hand sides of any production rules $r\in R$ may be surrounded by a context of terminal and nonterminal symbols.
+\item $V$ is the finite set of {\em non-terminal symbols}, also called variables. With $v_0\in V$ is the start non-terminal symbol that represents the whole grammar.
+\item $\Sigma$ is the finite set of {\em terminal symbols}, which are disjoint from the set of non-terminal symbols, i.e.~$V\cap \Sigma\neq\emptyset$.
+The set of terminal symbols is the {\em alphabet} of the language defined by $\mathcal{G}$
+\item $R$ is a finite set of $\alpha v\beta\rightarrow \alpha\gamma\beta$ rules where $v\in V$, $\alpha,\beta\in(V\cup\Sigma)^*$ and $\gamma\in(V\cup\Sigma)^+$.
+That is, the left-hand and right-hand sides of any production rule $r\in R$ may be surrounded by a context of terminal and non-terminal symbols.
 \end{itemize}
@@
 
-For instance, Figure~\ref{fig:csg}(b) shows how the string $S$ yields the string $aaabbbccc$.
-The language of a CSG, $L(\mathcal{G})=\{e\in \Sigma^*: v_0\Rightarrow^* e\}$, is the set of strings that contain only terminal symbols and that can be yielded from the string that contains only the initial non-terminal symbol $v_0$ (which is denoted by $S$ in the example of Figure~\ref{fig:csg}(b)).
+Figure~\ref{fig:csg}(b) shows how the string $S$ yields the string $aaabbbccc$.
+The language of a CSG, $L(\mathcal{G})=\{e\in \Sigma^*: v_0\Rightarrow^* e\}$, is the set of strings that contain only terminal symbols and that can be yielded from the string that contains only the initial non-terminal symbol $v_0$ (denoted by $S$ in the example of Figure~\ref{fig:csg}(b)).
 
@@
 \begin{itemize}
-\item Each node in a parse tree $t_{\mathcal{G},e}$ is either:
-\begin{itemize}
-\item An {\it internal node} that corresponds to the application of a rule $r\in R$.
-\item A {\it leaf node} that corresponds to a terminal symbol $\sigma\in \Sigma$ and has no outgoing branches.
-\end{itemize}
-\item Edges in a parse tree $t_{\mathcal{G},e}$ connect non-terminal symbols to terminal or non-terminal symbols following the rules $R$ in $\mathcal{G}$.
+\item Each node in a parse tree $t_{\mathcal{G},e}$ is either: An {\it internal node} that corresponds to the application of a rule $r\in R$ or a {\it leaf node} that corresponds to a terminal symbol $\sigma\in \Sigma$.
+\item Edges in a parse tree $t_{\mathcal{G},e}$ connect non-terminal symbols to terminal or non-terminal symbols following a rule in $\mathcal{G}$.
 \end{itemize}
@@
 \subsection{Linear-bounded Turing Machines}
-We define a {\em Turing machine} as a tuple $M=\tup{Q,q_o,Q_{\bot},\mathcal{T},\square,\Sigma,\delta}$, where:
+A {\em Turing machine} is a tuple $M=\tup{Q,q_o,Q_{\bot},\mathcal{T},\square,\Sigma,\delta}$:
 \begin{itemize}
-\item $Q$, is a finite and non-empty set of machine states such that $q_0\in Q$ it the initial state of the machine and $Q_{\bot}\subseteq Q$ is the subset of terminal states.  
-\item $\mathcal{T}$ is the {\em tape alphabet}, that is a finite an non-empty set of symbols that includes the {\em blank symbol} $\square\in\mathcal{T}$ (the only symbol allowed to occur on the tape infinitely often at any step during the computation).
-\item $\Sigma$ is the {\em input alphabet}, the set of symbols allowed to initially appear in the tape.
-\item $\delta: (Q\setminus Q_{\bot})\times \mathcal{T}\rightarrow Q\times \mathcal{T}\times\{left,right\}$ is the {\em transition function}.
-If $\delta$ is not defined for the current state of the machine and the current tape symbol, then the machine halts.
+\item $Q$, is a finite and non-empty set of machine states such that $q_0\in Q$ is the initial state of the machine and $Q_{\bot}\subseteq Q$ is the subset of acceptor states.  
+\item $\mathcal{T}$ is the {\em tape alphabet}, that is a finite non-empty set of symbols that includes the {\em blank symbol} $\square\in\mathcal{T}$ (the only symbol allowed to occur on the tape infinitely often) and that contains $\Sigma\subseteq\mathcal{T}$, the set of symbols allowed to initially appear in the tape (also called the {\em input alphabet}).
+\item $\delta: (Q\setminus Q_{\bot})\times \mathcal{T}\rightarrow Q\times\{left,right\}\times \mathcal{T}$ is the {\em transition function}.
+If $\delta$ is not defined for the current pair of machine state and tape symbol, then the machine halts.
 \end{itemize}
 
-A table is the most common convention to represent the transitions defined by $\delta$, where the table rows are indexed with the current tape symbol, while the table columns are indexed by the current machine state.
-For each possible tape symbol and state of the machine there is a table entry that defines: (1) the tape symbol to print at the current position of the header (2) whether the header is shifted {\em left} or {\em right} after the print operation and (3), the new state of the machine after the print operation.
+A table is the most common convention to represent the transitions defined by $\delta$, where the table rows are indexed by the current tape symbol, while the table columns are indexed by the current machine state.
+For each possible pair of tape symbol and machine state, there is a table entry that defines: (1) the tape symbol to print at the current position of the header (2) whether the header is shifted {\em left} or {\em right} after the print operation and (3), the new state of the machine after the print operation.
 For instance, Figure~\ref{tab:tm-anbncn} shows the table that represents the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
-In this example the tape alphabet is $\Sigma=\{a,b,c,x,y,z,\square\}$ while the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the acceptor state.
+In this example the tape alphabet is $\Sigma=\{a,b,c,x,y,z,\square\}$ while the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the only acceptor state.
 
@@
 \end{center}
-  \caption{\small Example of a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the acceptor state).}
+  \caption{\small Example of a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the only acceptor state).}
   \label{tab:tm-anbncn}
@@
 \begin{enumerate}
-\item Extending the input alphabet with two special symbols, serving as {\em left} and {\em right endmarkers}.
-\item Limiting computation to the portion of the tape containing the input plus the two tape cells holding the endmarkers.
-This means that transitions may not print other symbols over the endmarkers and that may neither move to the left of the left endmarker nor to the right of the right endmarker.
+\item Extending the input alphabet with two special symbols that serve as the {\em left} and {\em right endmarkers}.
+\item Limiting computation to the portion of the tape that contains the input plus the tape cells holding between the two endmarkers.
+This means that transitions do not print other symbols over the endmarkers and that never move neither to the left of the left endmarker nor to the right of the right endmarker.
 \end{enumerate}
@@
 
-Our approach to learning CSGs is to leverage on classical planning to learn an \strips\ action model s.t.
-each action schema in the model encodes a transition of a {\em Linear-bounded Turing Machine} that recognices the strings given as input.
-
-First we show how an entry in the table that defines the transitions of a {\em Linear-bounded Turing Machine} can be modeled as \strips\ action schema.
+Our approach is to leverage classical planning to learn an \strips\ action model s.t.
+each action schema in the model encodes a transition of the aimed {\em Linear-bounded Turing Machine}.
+First we show how an entry in the table that defines the transitions of a {\em Linear-bounded Turing Machine} is modeled as a \strips\ action schema.
 Next we show how classical planning can be effectively used to learn action models of this kind.
 
-\subsection{Modeling a Linear-bounded Turing Machine with Classical Planning}
-Given $\delta$, the {\em transition function} of a {\em Linear-bounded Turing Machines} $M$, it can be encoded as a classical planning frame $\Phi=\tup{F,A}$ as follows.
+\subsection{\strips\ Modeling of a Linear-bounded Turing Machine}
+A classical planning frame $\Phi=\tup{F,A}$ encodes the {\em transition function} $\delta$ of a {\em Linear-bounded Turing Machine} $M$ as follows.
 
@@
 Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine $M$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$, where $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-In more detail there are four predicates in $\Psi$:
+
+The predicates in $\Psi$ are:
 \begin{itemize}
 \item {\tt (head ?x)} that encodes the current position of the header in the tape.
-\item {\tt (next ?x1 ?x2)} that encodes that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
-\item {\tt (symbol-$\sigma$ ?x)} that encodes that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
-\item {\tt (state-$q$)} that encodes that $q\in Q$ is the current machine state.
+\item {\tt (next ?x1 ?x2)} encoding that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
+\item {\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
+\item {\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
 \end{itemize}
 
-Likewise we assume that actions $a\in A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action model is defined such that:
+Likewise we assume that actions $a\in A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action schema is defined such that:
 \begin{itemize}
-\item The {\bf header} of the schema is {\tt rule-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
-\item The {\bf precoditions} of the schema includes {\tt(head ?x)} and {\tt (next ?xl ?x) (next ?x ?xr)} to force that $?x$ is the current tape cell pointed by the header and that $?xl$ and $?xr$ respectively are its left and right neighbours.
-Additionally the schema includes preconditions {\tt(symbol-$\sigma$ ?x)} and {\tt (state-$q$)} to capture the current symbol pointed by the header and the currrent machine state.
-\item The {\bf delete effects} remove the current symbol pointed by the header and the currrent machine state.
-Finally, the {\bf positive effects} set the new symbol pointed by the header and the new machine state
+\item The {\bf header} of the schema is {\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
+\item The {\bf precoditions} of the schema includes {\tt(head ?x)} and {\tt (next ?xl ?x) (next ?x ?xr)} to force that $?x$ is the tape cell currently pointed by the header and that $?xl$ and $?xr$ respectively are its left and right neighbours.
+Additionally the schema includes preconditions {\tt(symbol-$\sigma$ ?x)} and {\tt (state-$q$)} to capture the symbol pointed by the header and the currrent machine state.
+\item The {\bf delete effects} remove the symbol pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state.
 \end{itemize}
 
-To illustrate this, Figure~\ref{fig:update-rule} shows the rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.
+The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.
 The full encoding of the Turing Machine defined in Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema with the same structure as the one of Figure~\ref{fig:update-rule}.
@@
 \begin{scriptsize}
-\begin{verbatim}
-(:action rule1 ;; a,q_0 -> x,r,q_1
+\begin{lstlisting}
+(:action transition-1 ;; a,$q_0\rightarrow$ x,r,$q_1$
   :parameters (?xl ?x ?xr)
   :precondition (and (head ?x)                       
-                     (next ?xl ?x) (next ?x ?xr) 
-                     (symbol-a ?x) (state-q_0))
+                     (symbol-a ?x) (state-$q_0$)
+                     (next ?xl ?x) (next ?x ?xr))
   :effect (and (not (head ?x)) 
-               (not (symbol-a ?x)) 
-               (not (state-q_0))
-               (symbol-x ?x) (head ?xr) (state-q_1)))
-\end{verbatim}
+               (not (symbol-a ?x)) (not (state-$q_0$))
+               (head ?xr) (symbol-x ?x) (state-$q_1$)))
+\end{lstlisting}
 \end{scriptsize}
- \caption{\small \strips\ action schema that models the first rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
+ \caption{\small \strips\ action schema that models the transition $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
 \label{fig:update-rule}
@@
 
-With this regard, the execution of a {\em Linear-bounded Turing Machine} can then be defined as a {\em plan trace} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$ such that $s_0$ encodes the initial state of the tape as well as the initial machine state and for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+The execution of a {\em Linear-bounded Turing Machine} can then be defined as a {\em plan trace} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$ such that $s_0$ encodes the initial state of the tape plus the initial machine state and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 
 \subsection{Learning \strips\ Models with Classical Planning}
-Here we show that the existing classical planning compilation for learning \strips\ action schemes from plan traces~\cite{aineto:learningSTRIPS:ICAPS2018} can be adapted to address the learning of CSGs.
-Note that in this case both the preconditions and negative effects of the \strips\ action schemes are already known so the learning task reduces to learning positive effects for these same action schemes.
-
-We formalize the learning of \strips\ action schemes from plan traces as a tuple $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ where:
+Learning \strips\ action schemes from plan traces is formalied as a $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ tuple, where:
 \begin{itemize}
-\item $\mathcal{M}$, the set of {\em partially specified} operator schemas.
+\item $\mathcal{M}$ is the set of {\em partially specified} operator schemas.
 \item $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_{n}}$ is a {\em plan trace} obtained watching the execution of the classical plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
@@
 
-Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ the compilation defined by~\citeauthor{aineto:learningSTRIPS:ICAPS2018} outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+Given a learning task $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ the compilation defined by~\citeauthor{aineto:learningSTRIPS:ICAPS2018} outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$.
+A solution plan to the classical planning task $P_{\Lambda}$ resulting from this compilation is a sequence of actions that:
+\begin{enumerate}
+\item Programs the action model $\mathcal{M}'$. A solution plan starts with a {\em prefix} that, for each action schema $\xi\in\mathcal{M}$, determines its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item Validates the action model $\mathcal{M}'$. The solution plan continues with a postfix that reproduces the given plan trace with the programmed action model $\mathcal{M}'$.
+\end{enumerate}
+
+\subsection{Learning CSGs with Classical Planning}
+Learning the acceptor automaton of a CSL is a particular case of the $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ learning task such that:
 \begin{itemize}
-\item $F_{\Lambda}$ contains:
-\begin{itemize}
-\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the plan trace.
-Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
-\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
-\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state where the learned action model is validated.
-\end{itemize}
-\item $I_{\Lambda}$ encodes the first observation, $s_0\subseteq F$, the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that encode the {\em partially specified} operator schemas and sets $mode_{prog}$ to true.
-\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in all de states of the input trace.
-\item $A_{\Lambda}$ comprises three kinds of actions:
-\begin{enumerate}
-\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$.
-For our purpose we only need to define the actions that add a {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
-\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
-& \{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
-\end{align*}
-\end{small}
-
-\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
-Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
-&\cup \{\neg mode_{val}\},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
-&\{\emptyset\}\rhd\{mode_{val}\}.
-\end{align*}
-\end{small}
-
-\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{T}$, {\tt\small $1\leq i< n$}.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\mathcal{O}|}\\
-& \cup \{mode_{val}\},\\
-\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
-\end{align*}
-\end{small}
-\end{enumerate}
+\item $\mathcal{M}$ is the set of {\em partially specified} operator schemas that encodes the preconditions and negative effects of the \strips\ schemas that encode the {\em transition function} $\delta$ of the {\em Linear-bounded Turing Machine} $M$.
+\item $\mathcal{T}$ represents the execution of the {\em Linear-bounded Turing Machine} $M$ for the parsing of a given string.
+\item $\Psi=\{${\small\tt(head ?x),(next ?x1 ?x2),(symbol-$\sigma$ ?x),(state-q)}$\}$ . 
 \end{itemize}
 
-\subsection{Learning CSGs with Classical Planning}
-Now we show that the learning of CSGs is a particular case of the $\Lambda=\tup{\mathcal{M},\mathcal{T},\Psi}$ such that:
-\begin{itemize}
-\item $\mathcal{M}$ is the set of {\em partially specified} operator schemas that encodes the rules of a {\em Linear-bounded Turing Machine} $M$.
-\item $\mathcal{T}$ represents the execution of the {\em Linear-bounded Turing Machine} $M$ on a given string.
-\item $\Psi=\{${\tt(head ?x),(next ?x1 ?x2),(symbol-$\sigma$ ?x),(state-q)}$\}$ . 
-\end{itemize}
+Here we show that the existing classical planning compilation $P_{\Lambda}$ for learning \strips\ action schemes from plan traces~\cite{aineto:learningSTRIPS:ICAPS2018} can be adapted to learning the acceptor automaton of a given CSL.
+When learning a CSGs the input {\em plan trace} $\mathcal{T}$ is partially observed:
+\begin{enumerate}
+\item States $s_i$, with $1\leq i\leq n$, do not contain the fluents {\tt (state-q)} s.t. $q\in Q$.
+\item At each state, $s_i$ with $1\leq i\leq n$, the exact executed action $a_i$ is unobserved while it is known that $a_i$ is built instantiaing the schema corresponding to a transition rule $\sigma,q\rightarrow \sigma',\{r,l\},q'$ provided that $\sigma$ is the symbol at the tape cell pointed by the header at state $s_i$.
+\end{enumerate}
+To handle the first source of partial observability the compilation does not require any modification because $s_0$ is fully observable since initially the machine state is always $q_0$ (likewise the machine state is always $q_n$ at the last state that corresponds to a string accepted by the learned automaton) and the compilation is robust to missing fluents in the intermediate states of the input plan trace $\mathcal{T}$.
 
-When learning a CSGs the {\em plan trace} $\mathcal{T}$ is partially observable:
-\begin{enumerate}
-\item The states $s_i$ with $1\leq i\leq n$ do not contain the fluents $(state-q)$.
-\item The exact executed actions $a_i$ are unknown while it is known that shoud be an action built instantiaing the schema corresponding to the transition rule $\sigma,q\rightarrow \sigma',{r,l},q'$ provided that $\sigma$ is the symbol at the cell of the tape that is currently pointed by the header at state $s_i$.
-\end{enumerate}
-To handle the first source of partial observability the compilation does not require any modification because $s_0$ is fully observable since initially the machine state is always $q_0$ (likewise the machine state is always $q_n$ at the last state that corresponds to a string accepted by the learned automaton).
-The second source of partial observability can be addressed by adding extra preconditions to the apply actions so only are appliable actions built instantiaing the schema corresponding to the transition rule $\sigma,q\rightarrow \sigma',{r,l},q'$ provided that $\sigma$ is the symbol at the cell of the tape that is currently pointed by the header at state $s_i$.
+Addressing the second source of partial observability requires a modification of the $P_{\Lambda}$ compilation.
+Adding extra preconditions to the apply actions so they are  only appliable actions iff they are built instantiaing the schema corresponding to the transition $\sigma,q\rightarrow \sigma',{r,l},q'$ provided that $\sigma$ is the symbol at the cell tape that is currently pointed by the header at state $s_i$.
 
@@
 \section{Parsing and Production of CSGs with classical planning}
-The task of parsing can be implemented with the planning model introduced in the previous section by defining a classical planning instace whose fluents and actions encode the transition of the corresponding Turing machine, the initial state encodes the tape with the string to recognize and the goal conditions require that the turing machine ends in on its acceptor states.
+The task of parsing can be implemented with the planning model $P_{\Lambda}$ introduced in the previous section by defining a classical planning instace whose fluents and actions encode the transition of the corresponding Turing machine, the initial state encodes the tape with the string to recognize and the goal conditions require that the turing machine ends at its acceptor states.
 
-The task of production requires to include actions that only write once a single symbol from the {\em tape alphabet} at the tape location indicated by the header.
+The task of production requires\ldots
 

### 2018-05-03 doc/aaai-19/model-recognition.tex
first notes on model recognition

\def\year{2017}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai18}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
%PDF Info Is Required:
  \pdfinfo{}
\setcounter{secnumdepth}{0}

\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{tikz}
\usepackage{comment}

\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{listings}

\lstset{
  basicstyle=\ttfamily,
  mathescape
}

\usepackage{multicol}
\usepackage{arydshln}
\usetikzlibrary{calc,backgrounds,positioning,fit}

\newcommand{\tup}[1]{{\langle #1 \rangle}}

\newcommand{\pre}{\mathsf{pre}}     % precondition
\newcommand{\del}{\mathsf{del}}     % effect
\newcommand{\add}{\mathsf{add}}     % effect
\newcommand{\eff}{\mathsf{eff}}     % effect
\newcommand{\cond}{\mathsf{cond}}   % conditional effect
\newcommand{\true}{\mathsf{true}}   % true
\newcommand{\false}{\mathsf{false}} % false
\newcommand{\PE}{\mathrm{PE}}     % precondition
\newcommand{\strips}{\textsc{Strips}}     % precondition

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

\begin{document}

\title{Planning Model Recognition}
\author{\#39}

% Commented for blind submission
\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
{\small Universitat Polit\`ecnica de Val\`encia.}\\
{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}

\maketitle
\begin{abstract} 
This paper defines the {\em Planning Model Recognition} task. Given a set of possible planning models and an observation of a plan execution, {\em Planning Model Recognition} is the task of identifying which model has the highest probability of producing the given observation. The paper proposes a metric to assess the likelihood of a given \strips\ model to produce the observation of a plan execution and a method for computing this likelihood using classical planning. This metric is robust to missing intermediate states and actions in the input observation of a plan execution. The effectiveness of our method is empirically evaluated for solving the {\em planning model recognition} task when the set of possible \strips\ models represents different finite tape {\em Turing Machines} that share the same tape alphabet and the same set of machine states.
\end{abstract}

\section{Introduction}
\label{sec:section1}
{\em Plan recognition} is the task of predicting the future actions of an agent by observing its current behaviour. {\em Goal recognition} is a subtask of plan recognition with the particular objective of discovering the terminal goal of an agent given its behaviour. A large number of methods have been proposed for plan recognition: methods that include rule-based  systems,  parsing  (both  conventional  and  stochastic), graph-covering, Bayesian nets, \ldots.

This paper defines {\em planning model recognition}, the task of identifying which model has the highest probability of producing the current behaviour of an agent. Despite related to both {\em Plan/Goal recognition}, this task is relevant because recognizing the followed model allows to apply the machinery of the crisp framework for {\em Plan recognition as planning} that leverages on a known planning model~\cite{ramirez2012plan,ramirez2009plan} besides, planning model recognition is also useful to determine whether the curent agent behaviour is optimal with respect to the most plausible model.

The paper proposes a metric to assess the likelihood of a given \strips\ model to produce the observation of a plan execution and a method for computing this likelihood using classical planning. This metric is robust to missing intermediate states and actions in the input observation of a plan execution. The effectiveness of our method is empirically evaluated for solving {\em planning model recognition} tasks when the set of possible \strips\ models represents a set of different finite tape {\em Turing Machines} that share the same tape alphabet and the same set of machine states.

\section{Background}
\label{sec:section2}
This section defines the classical planning model that we follow in this work.

\subsection{Classical planning}
Our approach for learning CSGs is compiling this inductive learning task into a classical planning task.

We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state. A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e. either~$l=f$ or $l=\neg f$. A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values). We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.

A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$, so the size of the state space is $2^{|F|}$. Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often we will abuse of notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.

A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$. We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$. The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.

The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.

A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition. A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. The {\em plan length} is denoted with $|\pi|=n$ . A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the application of the plan $\pi$ in the initial state $I$.

\subsection{\strips\ action schemas}
This work addresses the recognition of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}. Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.

\begin{figure}[hbt!]
\begin{scriptsize}
\begin{verbatim}
(:action stack
 :parameters (?v1 ?v2 - object)
 :precondition (and (holding ?v1) (clear ?v2))
 :effect (and (not (holding ?v1)) (not (clear ?v2)) 
              (handempty) (clear ?v1) (on ?v1 ?v2)))
\end{verbatim}
\end{scriptsize}
 \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
\label{fig:stack}
\end{figure}

We assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL. Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$. Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t. $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.

Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame. For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two. We define $F_v$, a new set of fluents s.t. $F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$, i.e. the variable names, and that defines the elements that can appear in an action schema. In {\em blocksworld} this set contains 11 elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.

For a given operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of fluents that represent the elements that can appear in that action schema. For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.

We assume also that actions $a\in A$ are instantiated from \strips\ operator schemas $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
\begin{itemize}
\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$. The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
\end{itemize}
Given the set of predicates $\Psi$ and the header of the operator schema $\xi$, $2^{2|F_v(\xi)|}$ defines the size of the space of possible \strips\ models for that operator. Note that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition. For instance, $2^{2|F_v(stack)|}=4194304$ for the blocksworld {\tt stack} operator while for {\tt pickup} is only 1024.

Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models. Formally, if $pars(\xi)=pars(\xi')$ it also holds that $F_v(\xi)=F_v(\xi')$. For instance, we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not. Two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.

\section{Planning Model Recognition}
\label{sec:section4}
Given a set of possible action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ and an observation of a plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$. {\em Planning Model Recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing the given observations $\mathcal{T}$.

\subsection{The \strips\ edit distance}
The intuition of our method for {\em Planning Model Recognition} is to assess how well an action model explains a given observations of plan executions according to the amount of {\em edition} required by the model to induce that observations. We first define the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
\begin{itemize}
\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_v(\xi)$.
\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_v(\xi)$.
\end{itemize}

We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ action models are. The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations}, deletion and insertion, have the same positive cost.

\begin{definition}
  Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that they are {\em comparable}. The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
\end{definition}

Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well. In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
\begin{definition}
The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
\end{definition}

We define now an edit distance to asses the mathing of a learned action model with respect to a plan trace $\mathcal{T}$.

\begin{definition}
  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and a plan trace $\mathcal{T}$ whose state observation are built with fluents in $F$. The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan trace $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
\end{definition}

Note that the distance of an action model $\mathcal{M}$ with respect to a plan trace $\mathcal{T}$ could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model. This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes~\cite{sohrabi:precognition:IJCAI2016}. Our definition of the {\em observation edit distance} is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.

\subsection{From edit distances into posterior probabilities}
The {\em observation edit distance} is mappable into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$. According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$. In our scenario, the hypotheses are about the set of possible \strips\ action models. Given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section3}. With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the size of the possible state observations of length $n$, that is $\mathcal{O}=s_0,\ldots,s_n$ is given by $2^{n\times|F|}$.

With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
\begin{enumerate}
\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$. If a set of plans $\Pi$ is available, this same strategy can be followed using the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
\end{enumerate}

\section{Computing the edit distance with classical planning}
Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, is {\em non-empty}. In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets. A solution to the planning task resulting from the extended compilation is a sequence of actions that:

\begin{enumerate}
\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}. A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}. In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema. However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}.
\item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}. The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:Section5} for the models that are programmed from scratch.
\end{enumerate}

Now $\Lambda$ does not formalize a learning task but the task of editing $\mathcal{M}$ to produce the plan trace $\mathcal{T}$, which results in the edited model $\mathcal{M}'$. The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:

\begin{itemize}
\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true. In addition, the input action model $\mathcal{M}$ is now encoded in the initial state. This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$. The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation. The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e. include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
\end{itemize}

To illustrate this, the plan of Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing. In this case the edited action model is however validated at the plan shown in Figure~\ref{fig:example-plans}.

\begin{figure}[hbt!]
{\tt\small
00 : (insert\_add\_handempty\_stack)\\
01 : (insert\_add\_clear\_stack\_var1)\\
02 : (apply\_unstack blockB blockA i1 i2)\\
03 : (apply\_putdown blockB i2 i3)\\
04 : (apply\_pickup blockA i3 i4)\\
05 : (apply\_stack blockA blockB i4 i5)\\
06 : (validate\_1)
}
 \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the plan shown in Figure~\ref{fig:example-plans}.}
\label{fig:plan-pdistance}
\end{figure}

Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deleitions) for that $\mathcal{M}'$ is validated in the given observations, e.g. $\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}. In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}). The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.

\subsection{Experiments}
\label{sec:section5}
To evaluate the empirical performance of our method for {\em planning model recognition} we defined a set of possible \strips\ models that represents a set of different finite tape {\em Turing Machines} that share the same tape alphabet and the same set of machine states.

\subsection{Turing Machines as \strips\ models}
A {\em Turing machine} is a tuple $M=\tup{Q,q_o,Q_{\bot},\mathcal{T},\square,\Sigma,\delta}$:
\begin{itemize}
\item $Q$, is a finite and non-empty set of machine states such that $q_0\in Q$ is the initial state of the machine and $Q_{\bot}\subseteq Q$ is the subset of acceptor states.  
\item $\mathcal{T}$ is the {\em tape alphabet}, that is a finite non-empty set of symbols that includes the {\em blank symbol} $\square\in\mathcal{T}$ (the only symbol allowed to occur on the tape infinitely often) and that contains $\Sigma\subseteq\mathcal{T}$, the set of symbols allowed to initially appear in the tape (also called the {\em input alphabet}).
\item $\delta: (Q\setminus Q_{\bot})\times \mathcal{T}\rightarrow Q\times\{left,right\}\times \mathcal{T}$ is the {\em transition function}. If $\delta$ is not defined for the current pair of machine state and tape symbol, then the machine halts.
\end{itemize}

A table is the most common convention to represent the transitions defined by $\delta$, where the table rows are indexed by the current tape symbol, while the table columns are indexed by the current machine state. For each possible pair of tape symbol and machine state, there is a table entry that defines: (1) the tape symbol to print at the current position of the header (2) whether the header is shifted {\em left} or {\em right} after the print operation and (3), the new state of the machine after the print operation. For instance, Figure~\ref{tab:tm-anbncn} shows the table that represents the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language. In this example the tape alphabet is $\Sigma=\{a,b,c,x,y,z,\square\}$ while the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the only acceptor state.

\begin{figure}
\begin{center}
    \begin{tabular}{| c | c | c | c | c | c | c |}
    \hline
      & $q_0$ & $q_1$ & $q_2$ & $q_3$ & $q_4$ & \underline{$q_5$} \\ \hline
    a & x,r,$q_1$ & a,r,$q_1$ & - &  a,l,$q_3$ & - & - \\ \hline
    b & - & y,r,$q_2$ & b,r,$q_2$ & b,l,$q_3$ & - & - \\ \hline
    c & - & - & z,l,$q_3$ & - & - & - \\ \hline
    x & - & - & - & x,r,$q_0$ & - & - \\ \hline
    y & y,r,$q_4$ & y,r,$q_1$ & - & y,l,$q_3$ & y,r,$q_4$ & - \\ \hline
    z & - & - & z,r,$q_2$ & z,l,$q_3$ & z,r,$q_4$ & - \\\hline
    $\square$ & - & - & - & - & $\square$,r,$q_5$  & - \\                
    \hline
    \end{tabular}
\end{center}
  \caption{\small Example of a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the only acceptor state).}
  \label{tab:tm-anbncn}
\end{figure}

A classical planning frame $\Phi=\tup{F,A}$ encodes the {\em transition function} $\delta$ of a {\em Turing Machine} $M$ as follows. We assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL~\cite{fox2003pddl2}. Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$. Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine $M$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$, where $\Omega^k$ is the $k$-th Cartesian power of $\Omega$. The predicates in $\Psi$ are:
\begin{itemize}
\item {\tt (head ?x)} that encodes the current position of the header in the tape.
\item {\tt (next ?x1 ?x2)} encoding that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
\item {\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
\item {\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
\end{itemize}

Likewise we assume that actions $a\in A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action schema is defined such that:
\begin{itemize}
\item The {\bf header} of the schema is {\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
\item The {\bf precoditions} of the schema includes {\tt(head ?x)} and {\tt (next ?xl ?x) (next ?x ?xr)} to force that $?x$ is the tape cell currently pointed by the header and that $?xl$ and $?xr$ respectively are its left and right neighbours. Additionally the schema includes preconditions {\tt(symbol-$\sigma$ ?x)} and {\tt (state-$q$)} to capture the symbol pointed by the header and the currrent machine state.
\item The {\bf delete effects} remove the symbol pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state.
\end{itemize}

The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}. The full encoding of the Turing Machine defined in Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema with the same structure as the one of Figure~\ref{fig:update-rule}. 
\begin{figure}[hbt!]
\begin{scriptsize}
\begin{lstlisting}
(:action transition-1 ;; a,$q_0\rightarrow$ x,r,$q_1$
  :parameters (?xl ?x ?xr)
  :precondition (and (head ?x)                       
                     (symbol-a ?x) (state-$q_0$)
                     (next ?xl ?x) (next ?x ?xr))
  :effect (and (not (head ?x)) 
               (not (symbol-a ?x)) (not (state-$q_0$))
               (head ?xr) (symbol-x ?x) (state-$q_1$)))
\end{lstlisting}
\end{scriptsize}
 \caption{\small \strips\ action schema that models the transition $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
\label{fig:update-rule}
\end{figure}

The execution of a {\em Turing Machine} can then be defined as a {\em plan trace} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$ such that $s_0$ encodes the initial state of the tape plus the initial machine state and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.

\subsection{Related work}
\label{sec:section6}

\subsection{Conclussions}
\label{sec:section7}

\bibliographystyle{aaai}
\bibliography{planlearnbibliography}

\end{document}

### 2018-05-03 doc/aaai-19/model-recognition.tex
more updates on model recognition

@@
 \begin{abstract} 
-This paper defines the {\em Planning Model Recognition} task.
-Given a set of possible planning models and an observation of a plan execution, {\em Planning Model Recognition} is the task of identifying which model has the highest probability of producing the given observation.
-The paper proposes a metric to assess the likelihood of a given \strips\ model to produce the observation of a plan execution and a method for computing this likelihood using classical planning.
-This metric is robust to missing intermediate states and actions in the input observation of a plan execution.
-The effectiveness of our method is empirically evaluated for solving the {\em planning model recognition} task when the set of possible \strips\ models represents different finite tape {\em Turing Machines} that share the same tape alphabet and the same set of machine states.
+Given a set of possible planning models and an observation of a plan execution, {\em Planning Model Recognition} is the task of identifying which of these models has the highest probability of producing the given observation.
+The paper formalizes the {\em Planning Model Recognition} task and proposes a metric to assess the likelihood of a given \strips\ model to produce the observation of a plan execution.
+This likelyhood metric is robust to missing intermediate states and actions in the input observation of a plan execution besides it is computable with classical planning.
+The effectiveness of our method for {\em Planning Model Recognition} is evaluated in a set of 100 possible \strips\ models each, representing a different {\em Turing Machine}, all of them sharing the same tape alphabet and same set of machine states.
 \end{abstract}
@@
 \label{sec:section1}
-{\em Plan recognition} is the task of predicting the future actions of an agent by observing its current behaviour.
-{\em Goal recognition} is a subtask of plan recognition with the particular objective of discovering the terminal goal of an agent given its behaviour.
-A large number of methods have been proposed for plan recognition: methods that include rule-based  systems,  parsing  (both  conventional  and  stochastic), graph-covering, Bayesian nets, \ldots.
+{\em Plan recognition} is the task of predicting the future actions of an agent provided the observation of its current behaviour.
+{\em Goal recognition} is a subtask of plan recognition with the particular objective of discovering the terminal goal of the observed agent.
+A large number of methods has been proposed for plan/goal recognition: rule-based systems,  parsing  (both  conventional and stochastic), graph-covering, Bayesian nets, \ldots
 
-This paper defines {\em planning model recognition}, the task of identifying which model has the highest probability of producing the current behaviour of an agent.
-Despite related to both {\em Plan/Goal recognition}, this task is relevant because recognizing the followed model allows to apply the machinery of the crisp framework for {\em Plan recognition as planning} that leverages on a known planning model~\cite{ramirez2012plan,ramirez2009plan} besides, planning model recognition is also useful to determine whether the curent agent behaviour is optimal with respect to the most plausible model.
+This paper defines {\em planning model recognition}, the task of identifying the model with the highest probability of producing the observed behaviour of an agent.
+Related to both {\em Plan/Goal recognition}, the recognition of planning models is meaningful because:
+\begin{itemize}
+\item It allows to apply the model-based machinery of the crisp framework for {\em Plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan},
+\item it is useful to determine whether the curent agent behaviour is suboptimal with respect to the identified model. For instance using an optimal planner~\cite{malte:optimalp:AAAI17}.  
+\end{itemize}
 
-The paper proposes a metric to assess the likelihood of a given \strips\ model to produce the observation of a plan execution and a method for computing this likelihood using classical planning.
-This metric is robust to missing intermediate states and actions in the input observation of a plan execution.
-The effectiveness of our method is empirically evaluated for solving {\em planning model recognition} tasks when the set of possible \strips\ models represents a set of different finite tape {\em Turing Machines} that share the same tape alphabet and the same set of machine states.
+The paper introduces a novel metric to assess the likelihood of a given \strips\ model to produce the observation of a plan execution and a method for computing this likelihood using an off-the-shelf classical planner.
+Remarkabley, this metric is robust to missing intermediate states and actions in the input observation of a plan execution.
+The effectiveness of our method for {\em Planning Model Recognition} is evaluated in a set of 100 possible \strips\ models each, representing a different {\em Turing Machine}, all of them sharing the same tape alphabet and same set of machine states.
 
@@
 \label{sec:section2}
-This section defines the classical planning model that we follow in this work.
+This section defines the classical planning model that we follow in this work and \strips\, the action model we use in the {\em Planning Model Recognition}.
 
 \subsection{Classical planning}
-Our approach for learning CSGs is compiling this inductive learning task into a classical planning task.
-
 We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
@@
 \label{sec:section4}
-Given a set of possible action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ and an observation of a plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-{\em Planning Model Recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing the given observations $\mathcal{T}$.
+Given a set of possible action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ and an observation of a plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+{\em Planning Model Recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.
 
 \subsection{The \strips\ edit distance}
-The intuition of our method for {\em Planning Model Recognition} is to assess how well an action model explains a given observations of plan executions according to the amount of {\em edition} required by the model to induce that observations.
+The intuition of our method for {\em Planning Model Recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by the model $\mathcal{M}$ to induce the observations $\mathcal{T}$.
 We first define the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
@@
 
-\subsection{From edit distances into posterior probabilities}
-The {\em observation edit distance} is mappable into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
+\subsection{The posterior probability distribution}
 According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
 In our scenario, the hypotheses are about the set of possible \strips\ action models.
-Given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section3}.
+
+Given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section2}.
+If we assume that a priori all models are equiprobable this means that $P(\mathcal{M})=\frac{1}{\prod_\xi 2^{2|F_v(\xi)|}}$.
+
 With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the size of the possible state observations of length $n$, that is $\mathcal{O}=s_0,\ldots,s_n$ is given by $2^{n\times|F|}$.
@@
 \begin{enumerate}
-\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$.
-If a set of plans $\Pi$ is available, this same strategy can be followed using the {\em plan edit distance} $\delta(\mathcal{M},\Pi)$.
-\item Applying the resulting distances to the above $P(\mathcal{O}|\mathcal{M})$ formula to map these distances into likelihoods
+\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$ and mapping the {\em observation edit distance} into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
 \item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
@@
 \label{sec:section5}
-To evaluate the empirical performance of our method for {\em planning model recognition} we defined a set of possible \strips\ models that represents a set of different finite tape {\em Turing Machines} that share the same tape alphabet and the same set of machine states.
+To evaluate the empirical performance of our method for {\em planning model recognition} we defined a set of possible \strips\ models, each representing a different {\em Turing Machines}, but all sharing the same tape alphabet and same set of machine states.
 
-\subsection{Turing Machines as \strips\ models}
+\subsection{Modeling {\em Turing Machines} with \strips\ }
 A {\em Turing machine} is a tuple $M=\tup{Q,q_o,Q_{\bot},\mathcal{T},\square,\Sigma,\delta}$:
@@
 \item The {\bf header} of the schema is {\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
-\item The {\bf precoditions} of the schema includes {\tt(head ?x)} and {\tt (next ?xl ?x) (next ?x ?xr)} to force that $?x$ is the tape cell currently pointed by the header and that $?xl$ and $?xr$ respectively are its left and right neighbours.
+\item The {\bf preconditions} of the schema includes {\tt(head ?x)} and {\tt (next ?xl ?x) (next ?x ?xr)} to force that $?x$ is the tape cell currently pointed by the header and that $?xl$ and $?xr$ respectively are its left and right neighbours.
 Additionally the schema includes preconditions {\tt(symbol-$\sigma$ ?x)} and {\tt (state-$q$)} to capture the symbol pointed by the header and the currrent machine state.

### 2018-05-04 doc/aaai-19/model-recognition.tex
udpates to model recognition

@@
 
-\title{Planning Model Recognition}
+\title{Model Recognition as Planning}
 \author{\#39}
@@
 \begin{abstract} 
-Given a set of possible planning models and an observation of a plan execution, {\em Planning Model Recognition} is the task of identifying which of these models has the highest probability of producing the given observation.
-The paper formalizes the {\em Planning Model Recognition} task and proposes a metric to assess the likelihood of a given \strips\ model to produce the observation of a plan execution.
-This likelyhood metric is robust to missing intermediate states and actions in the input observation of a plan execution besides it is computable with classical planning.
-The effectiveness of our method for {\em Planning Model Recognition} is evaluated in a set of 100 possible \strips\ models each, representing a different {\em Turing Machine}, all of them sharing the same tape alphabet and same set of machine states.
+Given a set of possible action models and an observation of a plan execution, {\em model recognition} is the task of identifying which of these action models has the highest probability of producing the observed plan execution.
+The paper formalizes the {\em model recognition} task and proposes a method to assess the probability of a given \strips\ action model to produce an observed plan execution.
+This method, that we called {\em model recognition as planning}, is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
+The effectiveness of {\em model recognition as planning} is shown in a set of 100 possible \strips\ models each, representing a different {\em Turing Machine}, all of them sharing the same tape alphabet and same set of machine states.
 \end{abstract}
@@
 
-This paper defines {\em planning model recognition}, the task of identifying the model with the highest probability of producing the observed behaviour of an agent.
-Related to both {\em Plan/Goal recognition}, the recognition of planning models is meaningful because:
+This paper defines {\em model recognition}, the task of identifying the action model with the highest probability of producing an observed plan execution.
+While related to {\em Plan/Goal recognition}, the contribution of {\em model recognition} is:
 \begin{itemize}
-\item It allows to apply the model-based machinery of the crisp framework for {\em Plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan},
-\item it is useful to determine whether the curent agent behaviour is suboptimal with respect to the identified model. For instance using an optimal planner~\cite{malte:optimalp:AAAI17}.  
+\item It allows to apply the model-based machinery of the {\em Plan recognition as planning} framework~\cite{ramirez2012plan,ramirez2009plan} with respect to the recognized.
+\item it allows to determine whether the observed plan execution is suboptimal with respect to the recognized model. For instance using an optimal planner~\cite{malte:optimalp:AAAI17}.  
 \end{itemize}
 
-The paper introduces a novel metric to assess the likelihood of a given \strips\ model to produce the observation of a plan execution and a method for computing this likelihood using an off-the-shelf classical planner.
-Remarkabley, this metric is robust to missing intermediate states and actions in the input observation of a plan execution.
-The effectiveness of our method for {\em Planning Model Recognition} is evaluated in a set of 100 possible \strips\ models each, representing a different {\em Turing Machine}, all of them sharing the same tape alphabet and same set of machine states.
+The paper introduces {\em model recognition as planning}, a novel method to assess the probability of a given \strips\ action model to produce an observed plan execution.
+This method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
+The effectiveness of {\em model recognition as planning} is shown in a set of 100 possible \strips\ models each, representing a different {\em Turing Machine}, all of them sharing the same tape alphabet and same set of machine states.
 
@@
 \label{sec:section2}
-This section defines the classical planning model that we follow in this work and \strips\, the action model we use in the {\em Planning Model Recognition}.
+This section defines the classical planning model and the \strips\ action model that we use for addressing the {\em model recognition} task.
 
@@
 
-\section{Planning Model Recognition}
+\section{Model Recognition}
 \label{sec:section4}
 Given a set of possible action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ and an observation of a plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-{\em Planning Model Recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.
+{\em Model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.
 
 \subsection{The \strips\ edit distance}
-The intuition of our method for {\em Planning Model Recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by the model $\mathcal{M}$ to induce the observations $\mathcal{T}$.
-We first define the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
+The intuition of our method for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by the model $\mathcal{M}$ to induce the observations $\mathcal{T}$.
+In this work we assume that the action models in the {\em model recognition} task are represented using the \strips\ formalism.
+We define here the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
 \begin{itemize}
@@
 \subsection{The posterior probability distribution}
-According to the Bayes rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the set of possible \strips\ action models.
+According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
+In our scenario, the hypotheses are about the set of possible \strips\ action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ while the observation is an input plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
 
-Given set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section2}.
+Given a set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section2}.
 If we assume that a priori all models are equiprobable this means that $P(\mathcal{M})=\frac{1}{\prod_\xi 2^{2|F_v(\xi)|}}$.
 
-With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the size of the possible state observations of length $n$, that is $\mathcal{O}=s_0,\ldots,s_n$ is given by $2^{n\times|F|}$.
+With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the number of possible state observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|F|}$ while the number of possible action observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|A|}$.
+If we assume that a priori all the traces are equiprobable, this means that $P(\mathcal{T})=\frac{1}{2^{n\times|F|}\times 2^{n\times|A|}}$.
 
-With this regard, $P(\mathcal{M}|\mathcal{O})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observation sequence $\mathcal{O}$ could be computed by:
+With this regard, $P(\mathcal{M}|\mathcal{T})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observed plan trace $\mathcal{T}$ can be estimated by:
 \begin{enumerate}
-\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{O})$ for every possible model $\mathcal{M}$ and mapping the {\em observation edit distance} into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
+\item Estimating the a priori probabilities $P(\mathcal{T})$ and $P(\mathcal{H})$ as explained.
+\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}$ and mapping the {\em observation edit distance} into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
 \item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
@@
 
-\section{Computing the edit distance with classical planning}
+\section{Model Recognition as Planning}
 Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, is {\em non-empty}.

### 2018-05-11 doc/aij18/aij18.tex
updates to the aij18 paper

@@
   This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning.
-The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state) to sequences of state observations or even just a set of initial and final states (where no intermediate action or state is known).
-The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
-What is more, the compilation is extensible to assess how well a given \strips\ action model matches the observation of a plan execution.
-This extension allows us to evaluate the quality of the learned models with respect to the actual models but also, with respect to observations of plan executions that act as a {\em test set}, which is useful when the actual models are not available.
-The performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and using these two evaluation approaches.
+The compilation approach is flexible to various amount and kind of available input knowledge; the action model to learn can be partially specified and the learning examples can range from plans (with their corresponding initial state) to sequences of state observations or even just a set of initial and final states (where no intermediate action or state is known).
+The compilation can also be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified or of the plan execution is not fully observed.
+What is more, the compilation is extensible to quantify how well a given \strips\ action model matches the observation of a plan execution.
+This extension allows us to evaluate the quality of the learned models with respect to the actual models ({\em syntactic evaluation}) but also, with respect to observations of plan executions ({\em semantic evaluation}).
+The performance of our compilation approach is syntactically and semantically evaluated when learning action models for a wide range of classical planning domains from the {\em International Planning Competition}.
 \end{abstract}
@@
 We presented a novel approach for learning \strips\ action models from examples using classical planning.
+The relevance of this approach goes beyond classical planning because of the wide range of different generative models that are STRIPS compilable.
 The approach is flexible to various amount and kind of input knowledge and accepts partially specified action models.
@@
 We also introduced a semantic method for evaluating the learned \strips\ action models with respect to observations of plan executions.
+Model evaluation with our compilation is also flexible to various amount and kind of available data in the given {\em test set}.
+The error of a learned action model could also be defined quantifying the amount of edition required by the observations of the plan execution $\mathcal{T}$ to match the given model $\mathcal{M}$.
+This would imply defining {\em edit operations} that modify the fluents in the observations instead of the {\em edit operations} that modify the action schemes~\cite{sohrabi:precognition:IJCAI2016}.
+Our definition of the {\em observation edit distance} is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
+
 Generating {\em informative} examples for learning planning action models is still an open issue.

### 2018-05-11 doc/aij18/compilation.tex
updates to the aij18 paper

@@
 
-Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task $P_{\Lambda}$ with conditional effects.
+Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task with conditional effects.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 
-\subsection{Conditional effects}
+\subsection{Classical planning with conditional effects}
 Conditional effects allow us to compactly define actions whose effects depend on the current state.
@@
 the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-plans}.
-Formally, $\Omega=\bigcup_{\small s\in\mathcal{O}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
+Formally, $\Omega=\bigcup_{\small s\in\mathcal{T}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
 \item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
-\item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ step of the observed plan $\pi=\tup{a_1, \ldots, a_n}$ that corresponds to action $a_i$.
-The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of plan $\pi$
+\item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\mathcal{T}$.
+The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of $\mathcal{T}$.
 \item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state observation $s_i\in\mathcal{T}$ where the action model is validated.
@@
 
-\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in all the input state observations $s_i\in\mathcal{T}$.
+\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in the state observations $s_i\in\mathcal{T}$.
 \item $A_{\Lambda}$ comprises three kinds of actions:
@@
 \end{figure}
-These $\mathsf{apply_{\xi,\omega}}$ actions includes the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.
+The extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\
 
@@
 In both cases the output planning problem becomes harder since the plan horizon is no longer bound and the classical planner must determine how many {\em apply} actions are necessary between any two state observations, i.e.
-between the application of two {\em validate} actions (classical planning is {\em PSPACE complete} while classical planning with a fixed plan horizon is {\em NP}~\cite{rintanen:SATplannning:AI2012}).
+between the application of two {\em validate} actions (classical planning is {\em PSPACE complete} while classical planning with a bound plan horizon is {\em NP}~\cite{rintanen:SATplannning:AI2012}).
 
@@
   By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{T}$.
+The compilation does not discard any possible \strips\ action schema $\mathcal{M}'$ definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{T}$.
+This means that a solution plan can be built selecting the corresponding  $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\mathcal{T}$.
 \end{small}
@@
 The larger these numbers, the larger the size of the $F_v(\xi)$ sets.
-This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents set and the corresponding set of {\em programming} actions.
-\item The number of given observations in $\mathcal{T}$. The larger $n$, the more $test_i$ fluents and $\mathsf{validate_{i}}$ actions in $P_{\Lambda}$.
+This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding set of {\em programming} actions.
+\item The length of $\mathcal{T}$, the observed plan execution. The larger $1\leq i\leq n$, the more $\{test_i\}$ fluents and $\{\mathsf{validate_{i}}\}$ actions in $P_{\Lambda}$.
 \end{itemize}
@@
 \hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
-\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
+\cond(\mathsf{programPre_{f_1,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
 \end{align*}

### 2018-05-11 doc/aij18/evaluation.tex
updates to the aij18 paper

@@
-\section{Evaluation of \strips\ Action Models with Classical Planning}
+\section{Evaluating \strips\ Action Models with Classical Planning}
 \label{evaluation}
@@
 If the actual model is available, the quality of a learned action model is quantifiable using the {\em precision} and {\em recall} metrics.
-These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
+{\em Precision} and {\em recall} are two syntactic metrics frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
 Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
@@
 
-The roles of two {\em comparable} action schemes or the roles of two action parameters that belong to the same type can be swapped by a $\Lambda$ learning task when the input data given in $\mathcal{T}$ is insufficient.
-For instance, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped.
+The roles of two {\em comparable} action schemes, or the roles of two action parameters that belong to the same type, can be swapped by a $\Lambda$ learning task when the observed input data, given in $\mathcal{T}$, is insufficient.
+For instance, when $\mathcal{T}$ does not contain any intermediate action the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped.
 Pure syntax-based evaluation metrics (like {\em precision} and {\em recall}) can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
@@
 Here we introduce a novel evaluation approach that is robust to role changes of this particular kind.
-The intuition of the approach is to assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
-Like our learning approach, our evaluation approach is also flexible to various amount and kind of available input knowledge.
+The intuition of the approach is to {\em semantically} assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
+This semantic evaluation approach is again flexible to various amount and kind of available input knowledge.
 
@@
 \end{mydefinition}
-Note that the number of deletions indicate the number of {\em false positives} while the number of insertions represents the number of {\em false negatives}.
-As explained, these numbers can be normalized with the {\em true positives} and {\em true negatives} to compute the {\em precision} and {\em recall} of a learned model.
 
@@
 
+Note that the number of deletions indicate the number of {\em false positives} while the number of insertions represents the number of {\em false negatives}.
+As explained, these numbers can be normalized with the amount of {\em true positives} to compute the {\em precision} and {\em recall} of a learned model with respect to a reference model.
 Normally evaluating a given learned domain with respect to the actual generative model is not possible because the actual model is not available.
 As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions~\cite{yang2007learning}.
-With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applied in $s_{i-1}$ and generates the successor state $s_i$.
 
@@
 
-The {\em observation edit distance} is mappable into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$.
-Further, following the previous correspondances ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can redefine the {\em precision} and {\em recall} of a learned model with respect to a partially observed plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
-In this case, for each action schema $\xi\in\mathcal{M}$, the corresponding number of {\em true positives} is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$ while the {\em true negatives} are given by $tn(\xi)=3\times|F_v(\xi)|-tp(\xi)$
-
-Note that the error of a learned action model could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
-This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes~\cite{sohrabi:precognition:IJCAI2016}.
-Our definition of the {\em observation edit distance} is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
+The {\em observation edit distance} is mappable into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
+Further, following the previous correspondances ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can redefine a {\em semantic} version of the  {\em precision} and {\em recall} metrics of a learned model with respect to a partially observed plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
+In this case, for each action schema $\xi\in\mathcal{M}$, the corresponding number of {\em true positives} is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$.
 
@@
 \begin{itemize}
-\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
-\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
-In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
+\item $F_{\Lambda}$, $I_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
+Note that, the input action model $\mathcal{M}$ is encoded in the initial state.
+This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in $I_{\Lambda}$ iff they appear in $\mathcal{M}$.
 \item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
 The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
-The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
+The only difference here is that each action for {\em programming} an operator schema now implement the two {\em edit operations} (i.e.
 include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).

### 2018-05-11 doc/aij18/introduction.tex
updates to the aij18 paper

@@
 
-The compilation approach is appealing by itself because it leverages off-the-shelf planners for learning and because its practicality allow us to report model learning results over a wide range of IPC planning domains.
+The compilation approach is appealing by itself because it leverages off-the-shelf planners for learning and because its practicality allow us to report learning results over a wide range of domains from the {\em International Planning Competition} (IPC).
 Moreover, it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
@@
 \item {\em Input flexibility}.
-The compilation is flexible to various amount and kind of available input knowledge.
-Learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
-This fact makes the compilation robust to the partial obervability of the intermediate input data.
-Last but not least, the compilation also accepts input knowledge about the structure of the actions in the form of partially specified action models, that is learning action models when some preconditions and effects are a priori known.
+Our classical planning compilation is flexible to various amount and kind of available input knowledge.
+The action model to learn can be partially specified and the learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
 % Also accepts background knowledge
@@
 \item {\em Model validation}.
-The compilation poses a novel framework to assess the validation of a STRIPS model with respect to a given set of observations.
-This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since it does not require a full plan or a full action model to determine validation.
-The relevance of this validation framework overcomes classical planning because of the wide range of different generative models that are STRIPS compilable (STRIPS is {\em Turing complete}).
+The compilation poses a novel framework to assess the validation of a STRIPS model with respect a plan execution.
+This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since it does require neither a full action model nor a fully observed plan to determine validation.
 
 \item {\em Model evaluation}.
-The compilation can assess how well a given \strips\ action model matches a {\em test set} with observations of plan executions.
-The idea is to assess the amount of edition required by the input action model to induce given observations of plan execution that serve as a test set.
-This allows us to assess learning performance without having the actual model and again, evaluation with our compilation is flexible to various amount and kind of available data in the given {\em test set}.
+The compilation can assess how well a given \strips\ action model matches a plan execution, which allows us to assess learning performance without having the actual action model.
+The idea is to assess the amount of edition required by the input action model to induce the given observation of the plan execution.
 \end{enumerate}
@@
 \item A unified formulation for learning and evaluating \strips\ action models from observed executions of plans. 
-\item Leveraging diverse forms of {\em background knowledge} for learning and evaluating \strips\ action models.
+\item Redefining the {\em precision} and {\em recall} metrics to evaluate learned action models with respect to a plan execution.
+%\item Leveraging diverse forms of {\em background knowledge} for learning and evaluating \strips\ action models.
 \item A complete empirical evaluation of the compilation approach.

### 2018-05-11 doc/aij18/learning_task.tex
updates to the aij18 paper

@@
 For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
+
 We define $F_v$, a new set of fluents s.t.
-$F\cap F_v=\emptyset$, that results from instantiating $\Psi$ using only the objects in $\Omega_v$, i.e.
-the variable names, and that defines the elements that can appear in an action schema.
+$F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}, and that defines the elements that can appear in the action schemes.
 In {\em blocksworld} this set contains 11 elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
-For a given operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of fluents that represent the elements that can appear in that action schema.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the fluents from $F_v$ that involve $v_2$ because the action header {\small\tt pickup($v_1$)} contains the single parameter $v_1$.
+In more detail, for a given operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in that action schema.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the elements from $F_v$ that involve $v_2$ because {\small\tt pickup($v_1$)} has arity one.
 
@@
 Note that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
-For instance, $2^{2|F_v(stack)|}=4194304$ for the blocksworld {\tt stack} operator while for {\tt pickup} is only 1024.
+For the {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while for the {\tt pickup} operator this number is only 1024.
 
-Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models.
-Formally, if $pars(\xi)=pars(\xi')$ it also holds that $F_v(\xi)=F_v(\xi')$.
+We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi'$).
 For instance, we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
-Two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.
+Last but not least, two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
 
-\subsection{Learning \strips\ action schemas from observations of plan executions}
-The learning task addressed in this paper corresponds to learning an action model by observing an agent(s) acting in the world.
+\subsection{Learning \strips\ action schemas from plan executions}
+The {\em learning task} addressed in this paper corresponds to learning an action model by observing an agent(s) acting in the world.
 This learning task is formalized as a tuple $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$:
 \begin{itemize}
-\item $\mathcal{M}$ is the set of {\em empty} operator schemas that define the actions of a given planning frame
+\item $\mathcal{M}$ is the set of {\em initial} operator schemas ({\em empty} or {\em partially specified} if some preconditions or effects are known) that define the actions of a given planning frame.
 \item $\Psi$ is the set of predicates that define the abstract state space of that planning frame. 
-\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace that contains the sequence of {\em state observations} obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in state $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace that contains the sequence of {\em state observations} obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applied in state $s_{i-1}$ and generates the successor state $s_i$.
 \end{itemize}
+
+A {\em solution} to a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the predicates $\Psi$, and the observed plan trace $\mathcal{T}$.
 Note that $\mathcal{M}$ can be inferred from $\mathcal{T}$ provided that the actions in $\mathcal{T}$ are {\em diverse} enough.
@@
 Likewise $\Psi$ can be inferred from $\mathcal{T}$ when at least the observation of a state $s\in \mathcal{T}$ is a full state, that is $|s|=|F|$.
-
-A solution to a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the predicates $\Psi$, and the observed plan trace $\mathcal{T}$.
 
@@
 
-In this paper we address the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task with the following assummptions:
+In this paper we address the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task with the following assumptions:
 \begin{enumerate}
@@
 \item The initial state $s_0\in\mathcal{T}$ is {\em fully observable}.
-The remaining states $s_i\in\mathcal{T}$ s.t.
-{\small $1\leq i\leq n$} can be {\em partially observable} (some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative).
-In the extreme, entire states $s_i$ and actions $a_i$ {\small $1\leq i\leq n$}, can be missing.
+\item Intermediate actions $a_i\in\mathcal{T}$ and states $s_i\in\mathcal{T}$ s.t.
+{\small $1\leq i\leq n$}, can be {\em partially observable} (some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative).
+%In the extreme, entire states $s_i$ and actions $a_i$ {\small $1\leq i\leq n$}, can be missing.
 In such scenario solving $\Lambda$ implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model.
-\item Any state observation $s\in\mathcal{T}$ is {\em noiseless}, meaning that if the value of a fluent is observed it is the correct value. 
+\item Observations in $\mathcal{T}$ are {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
 \end{enumerate}
-
 Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, that corresponds to observing the execution of a four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
@@
 \begin{verbatim}
-(pickup v1) (putdown v1) (stack v1 v2} (unstack v1 v2)
+(pickup ?v1) (putdown ?v1) (stack ?v1 ?v2} (unstack ?v1 ?v2)
 \end{verbatim}
@@
 \begin{verbatim}
-(handempty) (holding ?o  - object) (clear ?o - object) (ontable ?o - object)
-(on ?o1 - object ?o2 - object)
+(handempty) (holding ?v1) (clear ?v1) (ontable ?v1) (on ?v1 ?v2)
 \end{verbatim}
@@
 (clear B) (on B A) (ontable A) (handempty)
-
-;;; state observation #4
-(clear A) (on A B) (ontable B) (handempty)
 \end{verbatim}
@@
 
+\begin{footnotesize}
 {\footnotesize\tt ;;; Plan $\pi$}
-\begin{footnotesize}
 \begin{verbatim}
@@
 \end{footnotesize}
+
+\begin{footnotesize}
+\begin{verbatim}
+;;; state observation #4
+(clear A) (on A B) (ontable B) (handempty)
+\end{verbatim}
+\end{footnotesize}
+
  \caption{\small Task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ for learning a {\em blocksworld} \strips\ action model from a four-action plan and two state observations.}
@@
 
-The previous definition formalized the learning of \strips\ action models from the observation of a single plan execution.
-This definition is however extensible to the more general case where the execution of multiple plans is observed.
-In this case, $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ where each $t\in \mathcal{T}$ is a plan trace $t=\tup{s_0^t,a,_1^t,s_1^t,\ldots,a_n^t,s_{n}^t}$ such that, for each {\small $1\leq i\leq n^t$}, $a_i^t$ is applicable in $s_{i-1}^t$ and generates the successor state $s_i^t=\theta(s_{i-1}^t,a_i^t)$.
+The definition of the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is also extensible to the more general case where the execution of $\tau$ plans is observed.
+In this case, $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ where each $t\in \mathcal{T}$ is a plan trace $t=\tup{s_0^t,a,_1^t,s_1^t,\ldots,a_n^t,s_{n}^t}$ such that, for each {\small $1\leq i\leq n^t$}, $a_i^t$ is applied in $s_{i-1}^t$ and generates the successor state $s_i^t$.

### 2018-05-15 benchmarks/aaai18/automata/cleanup.py
added code for generating domain and probs for FSMs and TMs

@@ -1,0 +1,9 @@
+#! /usr/bin/env python
+import sys,os
+
+cmd=  "rm -rf *.pyc *.py~ "
+print cmd
+os.system(cmd)
+
+sys.exit(0)
+

### 2018-05-15 doc/aaai-19/model-recognition.tex
updates to the model recognition formalization

@@
 \begin{abstract} 
-Given a set of possible action models and an observation of a plan execution, {\em model recognition} is the task of identifying which of these action models has the highest probability of producing the observed plan execution.
-The paper formalizes the {\em model recognition} task and proposes a method to assess the probability of a given \strips\ action model to produce an observed plan execution.
+Given a set of possible action models and a partially observed plan execution, {\em model recognition} is the task of identifying which of these action models has the highest probability of producing the given plan execution.
+The {\em model recognition} task is relevant because it enables identifying algorithms by their execution and because, once an action model is recognized, the planning model-based machinery becomes applicable.
+This paper formalizes the {\em model recognition} task and proposes a method to assess the probability of a given \strips\ action model to produce a partially observed plan execution.
 This method, that we called {\em model recognition as planning}, is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
-The effectiveness of {\em model recognition as planning} is shown in a set of 100 possible \strips\ models each, representing a different {\em Turing Machine}, all of them sharing the same tape alphabet and same set of machine states.
+The effectiveness of {\em model recognition as planning} is shown with sets of \strips\ models that represent different {\em Turing Machines}, all of them defined within the same tape alphabet and same machine states.
+We show that {\em model recognition as planning} effectively identifies the executed {\em Turing Machine} despite intermediate machine state, applied transitions or tape values, are unobserved.
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:section1}
+\label{sec:introduction}
 {\em Plan recognition} is the task of predicting the future actions of an agent provided the observation of its current behaviour.
-{\em Goal recognition} is a subtask of plan recognition with the particular objective of discovering the terminal goal of the observed agent.
-A large number of methods has been proposed for plan/goal recognition: rule-based systems,  parsing  (both  conventional and stochastic), graph-covering, Bayesian nets, \ldots
+{\em Goal recognition} is a subtask of plan recognition with the particular aim of discovering the final objectives of the observed agent.
+Diverse methods has been proposed for plan and goal recognition~\cite{carberry2001techniques} such as {\em rule-based systems},  {\em parsing}  (both  conventional and stochastic), {\em graph-covering}, {\em Bayesian nets}, \ldots
 
-This paper defines {\em model recognition}, the task of identifying the action model with the highest probability of producing an observed plan execution.
-While related to {\em Plan/Goal recognition}, the contribution of {\em model recognition} is:
+{\em Plan recognition as planning} recently introduced a model-based approach for plan/goal recognition~\cite{ramirez2012plan,ramirez2009plan}.
+This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal according to the observed plan execution.
+
+In this paper we introduce the {\em model recognition} task as the task of identifying the action model with the highest probability of producing an observed plan execution.
+This task is of interest because :
 \begin{itemize}
-\item It allows to apply the model-based machinery of the {\em Plan recognition as planning} framework~\cite{ramirez2012plan,ramirez2009plan} with respect to the recognized.
-\item it allows to determine whether the observed plan execution is suboptimal with respect to the recognized model. For instance using an optimal planner~\cite{malte:optimalp:AAAI17}.  
+\item Once the action model is recognized the planning model-based machinery~\cite{ghallab2004automated,geffner:book:2013} becomes applicable.
+\item It enables identifying algorithms by their execution. Many computational models are compilable into classical planning~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2016hierarchical}.    
 \end{itemize}
@@
 This method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
-The effectiveness of {\em model recognition as planning} is shown in a set of 100 possible \strips\ models each, representing a different {\em Turing Machine}, all of them sharing the same tape alphabet and same set of machine states.
+
+Last but not least, the paper evaluates the effectiveness of {\em model recognition as planning} with sets of \strips\ models that represent different {\em Turing Machines}, all of them defined within the same tape alphabet and same machine states.
+We show that {\em model recognition as planning} effectively identifies the executed {\em Turing Machine} despite intermediate machine state, applied transitions or tape values, are unobserved.
 
 \section{Background}
-\label{sec:section2}
-This section defines the classical planning model and the \strips\ action model that we use for addressing the {\em model recognition} task.
+\label{sec:background}
+This section defines the classical planning model and the \strips\ action model that we use for formalizing the {\em model recognition as planning} method.
 
@@
 \subsection{\strips\ action schemas}
-This work addresses the recognition of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+This work addresses the learning and evaluation of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
 Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
@@
 
-We assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
+To formalize the target of the learning and evaluation tasks, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
 Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
@@
 Note that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
-For instance, $2^{2|F_v(stack)|}=4194304$ for the blocksworld {\tt stack} operator while for {\tt pickup} is only 1024.
+For the {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while for the {\small\tt pickup} operator this number is only 1024.
 
-Last but not least, we say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models.
-Formally, if $pars(\xi)=pars(\xi')$ it also holds that $F_v(\xi)=F_v(\xi')$.
-For instance, we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
-Two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and viceversa.
+We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi'$).
+For instance, we can claim that blocksworld operators {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not.
+Last but not least, two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
 
 \section{Model Recognition}
-\label{sec:section4}
-Given a set of possible action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ and an observation of a plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+\label{sec:recognition}
+Given a set of {\em comparable} action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ and a partially observed plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applied in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 {\em Model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.
@@
 The intuition of our method for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by the model $\mathcal{M}$ to induce the observations $\mathcal{T}$.
-In this work we assume that the action models in the {\em model recognition} task are represented using the \strips\ formalism.
+In this work we assume that all the action models in $M$ are represented using the \strips\ formalism.
 We define here the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
@@
 \begin{definition}
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that they are {\em comparable}.
+  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} \strips\ action models.
 The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
@@
 
-We define now an edit distance to asses the mathing of a learned action model with respect to a plan trace $\mathcal{T}$.
+We define now an edit distance to asses the matching of a learned action model with respect to a plan trace $\mathcal{T}$.
 
@@
 
-\subsection{The posterior probability distribution}
-According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$ given the observations $\mathcal{O}$ can be computed with $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the set of possible \strips\ action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ while the observation is an input plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
+\subsection{The \strips\ probability distribution}
+According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$ provided the observation $\mathcal{O}$ is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
 
-Given a set of predicates $\Psi$ and a given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$, as explained in Section~\ref{sec:Section2}.
+In our scenario, the hypotheses are about the set of possible \strips\ action models in $M$ while the observation is a partially observable input plan trace $\mathcal{T}$.
+Given the set of predicates $\Psi$ and the given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$.
 If we assume that a priori all models are equiprobable this means that $P(\mathcal{M})=\frac{1}{\prod_\xi 2^{2|F_v(\xi)|}}$.
@@
 With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the number of possible state observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|F|}$ while the number of possible action observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|A|}$.
-If we assume that a priori all the traces are equiprobable, this means that $P(\mathcal{T})=\frac{1}{2^{n\times|F|}\times 2^{n\times|A|}}$.
+If we assume that a priori all traces are equiprobable, this means $P(\mathcal{T})=\frac{1}{2^{n\times|F|}\times 2^{n\times|A|}}$.
 
-With this regard, $P(\mathcal{M}|\mathcal{T})$, the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) given an observed plan trace $\mathcal{T}$ can be estimated by:
+With this regard, we call $P(\mathcal{M}|\mathcal{T})$ the {\em \strips\ probability distribution} to the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) provided that the plan trace $\mathcal{T}$ is observed.
+The {\em \strips\ probability distribution} can be estimated by:
 \begin{enumerate}
-\item Estimating the a priori probabilities $P(\mathcal{T})$ and $P(\mathcal{H})$ as explained.
-\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}$ and mapping the {\em observation edit distance} into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
+\item Estimating the a priori probabilities $P(\mathcal{T})$ and $P(\mathcal{H})$, as explained.
+\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}\in M$ and mapping the {\em observation edit distance} into a $P(\mathcal{O}|\mathcal{H})$ likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
 \item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
@@
 \section{Model Recognition as Planning}
-Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, is {\em non-empty}.
-In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-A solution to the planning task resulting from the extended compilation is a sequence of actions that:
+\label{sec:asPlanning}
+Our method for addressing the {\em model recognition} task is compiling it into a classical planning task with conditional effects. This approach follows the following assumptions:
+\begin{enumerate}
+\item The initial state $s_0\in\mathcal{T}$ is {\em fully observable}.
+\item Intermediate actions $a_i\in\mathcal{T}$ and states $s_i\in\mathcal{T}$ s.t.
+{\small $1\leq i\leq n$}, can be {\em partially observed}.
+This means that some fluents in $s_i$ may be missing because it is unknown whether their value is either positive or negative.
+In the extreme, entire states $s_i$ can be missing.
+\item Observations in $\mathcal{T}$ are {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
+\end{enumerate}
+
+\subsection{Conditional effects}
+Conditional effects allow us to compactly define actions whose effects depend on the current state.
+An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+\[
+triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
+\]
+
+The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
+
+\subsection{The compilation}
+The compilation allow us to estimate the \strips\ edit distance of a given model $\mathcal{M}\in M$ with regard to the partially observed plan execution $\mathcal{T}$.
+The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
 
@@
 A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema.
-However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}.
 \item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:Section5} for the models that are programmed from scratch.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$.
 \end{enumerate}
-
-Now $\Lambda$ does not formalize a learning task but the task of editing $\mathcal{M}$ to produce the plan trace $\mathcal{T}$, which results in the edited model $\mathcal{M}'$.
-The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
-
-\begin{itemize}
-\item $F_{\Lambda}$ and $G_{\Lambda}$ are defined as in the previous compilation.
-\item $I_{\Lambda}'$ contains the fluents from $F$ that encode $s_0$ and $mode_{prog}$ set to true.
-In addition, the input action model $\mathcal{M}$ is now encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in the initial state iff they appear in $\mathcal{M}$.
-\item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
-The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
-The only difference here is that the actions for {\em programming} the operator schema now implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
-\end{itemize}
 
@@
 \begin{figure}[hbt!]
-{\tt\small
+{\tt\scriptsize
 00 : (insert\_add\_handempty\_stack)\\
@@
 
-\subsection{Experiments}
-\label{sec:section5}
+Given $\tup{\mathcal{M},\Psi,\mathcal{T}}$ the compilation outputs a classical planning task $P=\tup{F,A,I,G}$:
+\begin{itemize}
+\item $F$ contains:
+\begin{itemize}
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the plan trace given as input, i.e.
+the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-plans}.
+Formally, $\Omega=\bigcup_{\small s\in\mathcal{T}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
+\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
+For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
+\item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\mathcal{T}$.
+The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of $\mathcal{T}$.
+\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state observation $s_i\in\mathcal{T}$ where the action model is validated.
+\end{itemize}
+\item $I$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated.
+Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
+Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
+
+\item $G=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in the state observations $s_i\in\mathcal{T}$.
+\item $A$ comprises three kinds of actions:
+\begin{enumerate}
+\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$:
+\begin{itemize}
+\item Actions for {\bf removing} a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}, pre_{f}(\xi)\},\\
+\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
+\end{align*}
+\end{small}
+
+\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
+\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
+\end{align*}
+\end{small}
+\end{itemize}
+Besides these actions $A$ also contains the actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect.
+
+\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
+Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
+Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$ from {\em blocksworld}.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}
+&\cup \{\neg mode_{val}\},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
+&\{\emptyset\}\rhd\{mode_{val}\}.
+\end{align*}
+\end{small}
+
+\begin{figure}[hbt!]
+\begin{scriptsize}
+\begin{verbatim}
+(:action apply_stack
+  :parameters (?o1 - object ?o2 - object)
+  :precondition
+   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
+        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
+        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
+        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
+        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
+        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
+        (or (not (pre_clear_stack_v1)) (clear ?o1))
+        (or (not (pre_clear_stack_v2)) (clear ?o2))
+        (or (not (pre_holding_stack_v1)) (holding ?o1))
+        (or (not (pre_holding_stack_v2)) (holding ?o2))
+        (or (not (pre_handempty_stack)) (handempty)))
+  :effect
+   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
+        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
+        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
+        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
+        (when (del_ontable_stack_v1) (not (ontable ?o1)))
+        (when (del_ontable_stack_v2) (not (ontable ?o2)))
+        (when (del_clear_stack_v1) (not (clear ?o1)))
+        (when (del_clear_stack_v2) (not (clear ?o2)))
+        (when (del_holding_stack_v1) (not (holding ?o1)))
+        (when (del_holding_stack_v2) (not (holding ?o2)))
+        (when (del_handempty_stack) (not (handempty)))
+        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
+        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
+        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
+        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
+        (when (add_ontable_stack_v1) (ontable ?o1))
+        (when (add_ontable_stack_v2) (ontable ?o2))
+        (when (add_clear_stack_v1) (clear ?o1))
+        (when (add_clear_stack_v2) (clear ?o2))
+        (when (add_holding_stack_v1) (holding ?o1))
+        (when (add_holding_stack_v2) (holding ?o2))
+        (when (add_handempty_stack) (handempty))
+        (when (modeProg) (not (modeProg)))))
+\end{verbatim}
+\end{scriptsize}
+ \caption{\small PDDL action for applying an already programmed schema $stack$ (implications are coded as disjunctions).}
+\label{fig:compilation}
+\end{figure}
+The extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\
+
+\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{T}$, {\tt\small $1\leq i< n$}.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
+\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
+\end{align*}
+\end{small}
+\end{enumerate}
+\end{itemize}
+
+\subsection{Evaluation}
+\label{sec:evaluation}
 To evaluate the empirical performance of our method for {\em planning model recognition} we defined a set of possible \strips\ models, each representing a different {\em Turing Machines}, but all sharing the same tape alphabet and same set of machine states.
@@
 
-A classical planning frame $\Phi=\tup{F,A}$ encodes the {\em transition function} $\delta$ of a {\em Turing Machine} $M$ as follows.
-We assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL~\cite{fox2003pddl2}.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine $M$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$, where $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-The predicates in $\Psi$ are:
+A classical planning frame $\Phi=\tup{F,A}$ encodes the {\em transition function} $\delta$ of a {\em Turing Machine} $M$ as follows:
 \begin{itemize}
-\item {\tt (head ?x)} that encodes the current position of the header in the tape.
-\item {\tt (next ?x1 ?x2)} encoding that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
-\item {\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
-\item {\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
+\item The fluents $F$ are instantiated from the set of {\em predicates} $\Psi$ that contains: {\small\tt (head ?x)} that encodes the current position of the header in the tape.
+{\small\tt (next ?x1 ?x2)} encoding that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
+{\small\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
+{\small\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
+Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine $M$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
+\item The actions $a\in A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action schema is defined such that:
+\begin{itemize}
+\item The {\bf header} of the schema is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
+\item The {\bf preconditions} of the schema includes {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to force that $?x$ is the tape cell currently pointed by the header and that $?xl$ and $?xr$ respectively are its left and right neighbours.
+Additionally the schema includes preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} to capture the symbol pointed by the header and the currrent machine state.
+\item The {\bf delete effects} remove the symbol pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state.
 \end{itemize}
-
-Likewise we assume that actions $a\in A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action schema is defined such that:
-\begin{itemize}
-\item The {\bf header} of the schema is {\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
-\item The {\bf preconditions} of the schema includes {\tt(head ?x)} and {\tt (next ?xl ?x) (next ?x ?xr)} to force that $?x$ is the tape cell currently pointed by the header and that $?xl$ and $?xr$ respectively are its left and right neighbours.
-Additionally the schema includes preconditions {\tt(symbol-$\sigma$ ?x)} and {\tt (state-$q$)} to capture the symbol pointed by the header and the currrent machine state.
-\item The {\bf delete effects} remove the symbol pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state.
 \end{itemize}
@@
 \begin{lstlisting}
-(:action transition-1 ;; a,$q_0\rightarrow$ x,r,$q_1$
+(:action transition-1    ;;; a,$q_0\rightarrow$ x,r,$q_1$
   :parameters (?xl ?x ?xr)
-  :precondition (and (head ?x)                       
-                     (symbol-a ?x) (state-$q_0$)
+  :precondition (and (head ?x) (symbol-a ?x) (state-$q_0$)
                      (next ?xl ?x) (next ?x ?xr))
@@
 
-\subsection{Related work}
-\label{sec:section6}
+\subsection{Experimental setup}
+We randomly generated $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a two-symbol three-state {\em Turing Machines}.
+We randomly choose a machine $\mathcal{M}\in M$ and produce an fifty-step execution plan trace $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_{50},s_{50}}$.
 
 \subsection{Conclussions}
-\label{sec:section7}
+\label{sec:conclussions}
 

### 2018-05-16 doc/aaai-19/model-recognition.tex
updates added to the model recognition definition

@@
 \label{sec:recognition}
-Given a set of {\em comparable} action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ and a partially observed plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applied in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+Given a finite and non-empty set of {\em comparable} action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, and a partially observed plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applied in $s_{i-1}$ and generates state $s_i$.
 {\em Model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.
@@
 \subsection{The \strips\ edit distance}
-The intuition of our method for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by the model $\mathcal{M}$ to induce the observations $\mathcal{T}$.
-In this work we assume that all the action models in $M$ are represented using the \strips\ formalism.
-We define here the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
+Our approach for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by the model $\mathcal{M}$ to induce the observations $\mathcal{T}$.
+
+In this work we assume that action models $\mathcal{M}\in M$ are represented with the \strips\ formalism. We define here the two allowed \emph{operations} to edit an \strips\ action model:
 \begin{itemize}
@@
 \begin{definition}
-  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and a plan trace $\mathcal{T}$ whose state observation are built with fluents in $F$.
-The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan trace $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
+  Given $\mathcal{M}$, a \strips\ action model built from $F_v$ and $\mathcal{T}$, a plan trace built with fluents in $F$ and actions in $A$.
+The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
+$\mathcal{M}'$ produces a valid plan trace $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
 
-Note that the distance of an action model $\mathcal{M}$ with respect to a plan trace $\mathcal{T}$ could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
-This would imply defining {\em edit operations} that modify the fluents in the state observations instead of the {\em edit operations} that modify the action schemes~\cite{sohrabi:precognition:IJCAI2016}.
+The distance of an action model $\mathcal{M}$ with respect to a plan trace $\mathcal{T}$ could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
+This implies defining {\em edit operations} that modify the state observations instead of modifying the action schemes~\cite{sohrabi:precognition:IJCAI2016}.
 Our definition of the {\em observation edit distance} is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
@@
 According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$ provided the observation $\mathcal{O}$ is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
+In our scenario, the hypotheses are about the set of possible \strips\ action models $\mathcal{M}\in M$ while the observation is a partially observed plan trace $\mathcal{T}$.
 
-In our scenario, the hypotheses are about the set of possible \strips\ action models in $M$ while the observation is a partially observable input plan trace $\mathcal{T}$.
-Given the set of predicates $\Psi$ and the given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$.
+We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\mathcal{T})$, to the probability distribution of the comparable \strips\ models (within the $F_v(\xi)$ sets) provided that the plan trace $\mathcal{T}$ is observed.
+The {\em \strips\ probability distribution} can be estimated by:
+\begin{enumerate}
+\item Estimating the a priori probabilities $P(\mathcal{T})$ and $P(\mathcal{H})$.
+For instance, given the set of predicates $\Psi$ and the given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$.
 If we assume that a priori all models are equiprobable this means that $P(\mathcal{M})=\frac{1}{\prod_\xi 2^{2|F_v(\xi)|}}$.
-
 With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the number of possible state observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|F|}$ while the number of possible action observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|A|}$.
 If we assume that a priori all traces are equiprobable, this means $P(\mathcal{T})=\frac{1}{2^{n\times|F|}\times 2^{n\times|A|}}$.
-
-With this regard, we call $P(\mathcal{M}|\mathcal{T})$ the {\em \strips\ probability distribution} to the probability distribution of the possible \strips\ models (within the $F_v(\xi)$ sets) provided that the plan trace $\mathcal{T}$ is observed.
-The {\em \strips\ probability distribution} can be estimated by:
-\begin{enumerate}
-\item Estimating the a priori probabilities $P(\mathcal{T})$ and $P(\mathcal{H})$, as explained.
-\item Computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}\in M$ and mapping the {\em observation edit distance} into a $P(\mathcal{O}|\mathcal{H})$ likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
+\item Estimating the conditional probability $P(\mathcal{T}|\mathcal{M})$.
+For instance, we can compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}\in M$ and map  this distance into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
 \item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
@@
 \label{sec:asPlanning}
-Our method for addressing the {\em model recognition} task is compiling it into a classical planning task with conditional effects. This approach follows the following assumptions:
+Here we show a classical planning compilation for computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for each $\mathcal{M}\in M$. The compilation follows these assumptions
 \begin{enumerate}
@@
 {\small $1\leq i\leq n$}, can be {\em partially observed}.
-This means that some fluents in $s_i$ may be missing because it is unknown whether their value is either positive or negative.
+This means that some fluents in $s_i$ may be missing because its value is unknown.
 In the extreme, entire states $s_i$ can be missing.
@@
 \subsection{Conditional effects}
-Conditional effects allow us to compactly define actions whose effects depend on the current state.
+Conditional effects allow us to compactly define the actions output by our compilation.
 An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
@@
 \hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
-\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
+\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
+&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
 \end{align*}
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
 &\cup \{\neg mode_{val}\},\\
@@
 \label{sec:evaluation}
-To evaluate the empirical performance of our method for {\em planning model recognition} we defined a set of possible \strips\ models, each representing a different {\em Turing Machines}, but all sharing the same tape alphabet and same set of machine states.
+To evaluate the empirical performance of our method for {\em model recognition as planning} we defined a set of possible \strips\ models, each representing a different {\em Turing Machine}, but all sharing the same set of machine states and same tape alphabet.
 
 \subsection{Modeling {\em Turing Machines} with \strips\ }
-A {\em Turing machine} is a tuple $M=\tup{Q,q_o,Q_{\bot},\mathcal{T},\square,\Sigma,\delta}$:
+A {\em Turing machine} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\mathcal{T},\square,\Sigma,\delta}$:
 \begin{itemize}
 \item $Q$, is a finite and non-empty set of machine states such that $q_0\in Q$ is the initial state of the machine and $Q_{\bot}\subseteq Q$ is the subset of acceptor states.  
-\item $\mathcal{T}$ is the {\em tape alphabet}, that is a finite non-empty set of symbols that includes the {\em blank symbol} $\square\in\mathcal{T}$ (the only symbol allowed to occur on the tape infinitely often) and that contains $\Sigma\subseteq\mathcal{T}$, the set of symbols allowed to initially appear in the tape (also called the {\em input alphabet}).
+\item $\mathcal{T}$ is the {\em tape alphabet}, that is a finite non-empty set of symbols that includes the {\em blank symbol} $\square\in\mathcal{T}$ (the only symbol allowed to occur on the tape infinitely often) and the {\em input alphabet} $\Sigma\subseteq\mathcal{T}$, the subset of symbols allowed to initially appear in the tape.
 \item $\delta: (Q\setminus Q_{\bot})\times \mathcal{T}\rightarrow Q\times\{left,right\}\times \mathcal{T}$ is the {\em transition function}.
@@
 
-A table is the most common convention to represent the transitions defined by $\delta$, where the table rows are indexed by the current tape symbol, while the table columns are indexed by the current machine state.
-For each possible pair of tape symbol and machine state, there is a table entry that defines: (1) the tape symbol to print at the current position of the header (2) whether the header is shifted {\em left} or {\em right} after the print operation and (3), the new state of the machine after the print operation.
-For instance, Figure~\ref{tab:tm-anbncn} shows the table that represents the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
-In this example the tape alphabet is $\Sigma=\{a,b,c,x,y,z,\square\}$ while the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the only acceptor state.
+Figure~\ref{tab:tm-anbncn} shows the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
+For each possible pair of tape symbol and machine state, $\delta$ defines: (1) the tape symbol to print at the current position of the header (2) whether the header is shifted {\em left} or {\em right} after the print operation and (3), the new state of the machine after the print operation.
+In this example the {\em tape alphabet} is $\mathcal{T}=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} is $\Sigma=\{a,b,c,\square\}$ and the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the only acceptor state.
 
@@
 \begin{center}
-    \begin{tabular}{| c | c | c | c | c | c | c |}
+    \begin{tabular}{| c || c | c | c | c | c | c |}
     \hline
-      & $q_0$ & $q_1$ & $q_2$ & $q_3$ & $q_4$ & \underline{$q_5$} \\ \hline
+      & $q_0$ & $q_1$ & $q_2$ & $q_3$ & $q_4$ & \underline{$q_5$} \\ \hline\hline
     a & x,r,$q_1$ & a,r,$q_1$ & - &  a,l,$q_3$ & - & - \\ \hline
@@
 \end{center}
-  \caption{\small Example of a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the only acceptor state).}
+  \caption{\small Seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the only acceptor state).}
   \label{tab:tm-anbncn}
@@
 
-A classical planning frame $\Phi=\tup{F,A}$ encodes the {\em transition function} $\delta$ of a {\em Turing Machine} $M$ as follows:
+A classical planning frame $\Phi=\tup{F,A}$ encodes the {\em transition function} $\delta$ of a {\em Turing Machine} $\mathcal{M}$ as follows:
 \begin{itemize}
-\item The fluents $F$ are instantiated from the set of {\em predicates} $\Psi$ that contains: {\small\tt (head ?x)} that encodes the current position of the header in the tape.
+\item The fluents $F$ are instantiated from a set of four {\em predicates} $\Psi$: {\small\tt (head ?x)} that encodes the current position of the header in the tape.
 {\small\tt (next ?x1 ?x2)} encoding that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
-{\small\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
+{\small\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\mathcal{T}$.
 {\small\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
-Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine $M$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
-\item The actions $a\in A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action schema is defined such that:
+Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine $\mathcal{M}$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
+\item Actions $A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action schema is defined such that:
 \begin{itemize}
-\item The {\bf header} of the schema is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
-\item The {\bf preconditions} of the schema includes {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to force that $?x$ is the tape cell currently pointed by the header and that $?xl$ and $?xr$ respectively are its left and right neighbours.
-Additionally the schema includes preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} to capture the symbol pointed by the header and the currrent machine state.
+\item The {\bf header} is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
+\item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to make $?x$ the tape cell pointed by the header and $?xl$ and $?xr$ its left and right neighbours.
+Preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} are also included to capture the symbol pointed by the header and the currrent machine state.
 \item The {\bf delete effects} remove the symbol pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state.
@@
 
-The execution of a {\em Turing Machine} can then be defined as a {\em plan trace} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$ such that $s_0$ encodes the initial state of the tape plus the initial machine state and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+Since the {\em transition function} of a {\em Turing Machine} can be encoded as classical planning frame the execution of that machine can be defined as a {\em plan trace} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
+In this case $s_0$ encodes the initial state of the tape plus the initial machine state and, for each {\small $1\leq i\leq n$}, $a_i\in A$ is applied in $s_{i-1}$ to generate state $s_i$.
 

### 2018-05-16 doc/aaai-19/model-recognition.tex
updates added to the model recognition formulation

@@
 Given a set of possible action models and a partially observed plan execution, {\em model recognition} is the task of identifying which of these action models has the highest probability of producing the given plan execution.
-The {\em model recognition} task is relevant because it enables identifying algorithms by their execution and because, once an action model is recognized, the planning model-based machinery becomes applicable.
+{\em Model recognition} is relevant because it enables identifying algorithms by their execution and because, once an action model is recognized, the planning model-based machinery becomes applicable.
 This paper formalizes the {\em model recognition} task and proposes a method to assess the probability of a given \strips\ action model to produce a partially observed plan execution.
 This method, that we called {\em model recognition as planning}, is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
-The effectiveness of {\em model recognition as planning} is shown with sets of \strips\ models that represent different {\em Turing Machines}, all of them defined within the same tape alphabet and same machine states.
+The effectiveness of {\em model recognition as planning} is shown with sets of \strips\ models that represent different {\em Turing Machines}.
 We show that {\em model recognition as planning} effectively identifies the executed {\em Turing Machine} despite intermediate machine state, applied transitions or tape values, are unobserved.
@@
 
-{\em Plan recognition as planning} recently introduced a model-based approach for plan/goal recognition~\cite{ramirez2012plan,ramirez2009plan}.
+{\em Plan recognition as planning} recently defined the model-based approach for plan/goal recognition~\cite{ramirez2012plan,ramirez2009plan}.
 This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal according to the observed plan execution.
@@
 \item Once the action model is recognized the planning model-based machinery~\cite{ghallab2004automated,geffner:book:2013} becomes applicable.
-\item It enables identifying algorithms by their execution. Many computational models are compilable into classical planning~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2016hierarchical}.    
+\item It enables identifying algorithms by observing their execution.
+Many computational models are compilable into classical planning~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2016hierarchical}.
 \end{itemize}
@@
 \label{sec:recognition}
-Given a finite and non-empty set of {\em comparable} action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, and a partially observed plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applied in $s_{i-1}$ and generates state $s_i$.
+Given a finite and non-empty set of {\em comparable} action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, and a partially observed plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applied in $s_{i-1}$ and generates $s_i$.
 {\em Model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.
@@
 \subsection{The \strips\ edit distance}
-Our approach for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by the model $\mathcal{M}$ to induce the observations $\mathcal{T}$.
+Our approach for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by the model $\mathcal{M}$ to produce $\mathcal{T}$.
 
@@
 
-The distance of an action model $\mathcal{M}$ with respect to a plan trace $\mathcal{T}$ could also be defined quantifying the amount of edition required by the observations of the plan execution to match the given model.
-This implies defining {\em edit operations} that modify the state observations instead of modifying the action schemes~\cite{sohrabi:precognition:IJCAI2016}.
-Our definition of the {\em observation edit distance} is more practical since normally, $F_v$ is smaller than $F$ because the number of {\em variable objects} is smaller than the number of objects in the state observations.
+The $\delta(\mathcal{M},\mathcal{T})$ distance could also be defined assessing the edition required by the observed plan execution to match the given model.
+This implies defining {\em edit operations} that modify $\mathcal{T}$ instead of modifying $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
+Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$ (the number of {\em variable objects} is normally smaller than the number of objects in the observed states).
 
@@
 According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$ provided the observation $\mathcal{O}$ is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the set of possible \strips\ action models $\mathcal{M}\in M$ while the observation is a partially observed plan trace $\mathcal{T}$.
+In our scenario, the hypotheses are about the set of possible \strips\ action models $\mathcal{M}\in M$ while the given observation is a partially observed plan execution $\mathcal{T}$.
 
-We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\mathcal{T})$, to the probability distribution of the comparable \strips\ models (within the $F_v(\xi)$ sets) provided that the plan trace $\mathcal{T}$ is observed.
-The {\em \strips\ probability distribution} can be estimated by:
+We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\mathcal{T})$, to the probability distribution of the comparable \strips\ models (within the $F_v(\xi)$ sets) provided that the plan execution $\mathcal{T}$ is observed.
+This probability distribution can be estimated in three-steps by:
 \begin{enumerate}
@@
 For instance, we can compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}\in M$ and map  this distance into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
-\item Applying the Bayes rule to obtain the normalized posterior probabilities, these probabilities must sum 1.
+\item Applying the Bayes rule to obtain the normalized posterior probabilities since these probabilities must sum 1.
 \end{enumerate}
@@
 \label{sec:asPlanning}
-Here we show a classical planning compilation for computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for each $\mathcal{M}\in M$. The compilation follows these assumptions
+Now we show a classical planning compilation for computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for each $\mathcal{M}\in M$. The compilation follows these assumptions
 \begin{enumerate}
@@
 In the extreme, entire states $s_i$ can be missing.
-\item Observations in $\mathcal{T}$ are {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
+\item $\mathcal{T}$ is {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
 \end{enumerate}
@@
 \subsection{The compilation}
-The compilation allow us to estimate the \strips\ edit distance of a given model $\mathcal{M}\in M$ with regard to the partially observed plan execution $\mathcal{T}$.
 The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
-
 \begin{enumerate}
@@
 
-To illustrate this, the plan of Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-In this case the edited action model is however validated at the plan shown in Figure~\ref{fig:example-plans}.
+To illustrate this, Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+The edited action model is validated at a four action plan for inverting a two-block tower.
+Our interest is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deleitions) for that $\mathcal{M}'$ is validated in the given observations, e.g.
+$\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}.
+In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$.
+The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
 
@@
 }
- \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the plan shown in Figure~\ref{fig:example-plans}.}
+ \caption{\small Plan for editing and validating a given {\em blockswold} model.}
 \label{fig:plan-pdistance}
 \end{figure}
-
-Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deleitions) for that $\mathcal{M}'$ is validated in the given observations, e.g.
-$\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}.
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
-The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
 
@@
 \item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the plan trace given as input, i.e.
-the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-plans}.
+the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:plan-pdistance}.
 Formally, $\Omega=\bigcup_{\small s\in\mathcal{T}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
+\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\\
+&\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
 \cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
@@
 \label{sec:evaluation}
-To evaluate the empirical performance of our method for {\em model recognition as planning} we defined a set of possible \strips\ models, each representing a different {\em Turing Machine}, but all sharing the same set of machine states and same tape alphabet.
+To evaluate the empirical performance of {\em model recognition as planning} we defined a set of possible \strips\ models, each representing a different {\em Turing Machine}, but all sharing the same set of machine states and same tape alphabet.
 
 \subsection{Modeling {\em Turing Machines} with \strips\ }
-A {\em Turing machine} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\mathcal{T},\square,\Sigma,\delta}$:
+A {\em Turing machine} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma,\Upsilon,\square,\delta}$:
 \begin{itemize}
 \item $Q$, is a finite and non-empty set of machine states such that $q_0\in Q$ is the initial state of the machine and $Q_{\bot}\subseteq Q$ is the subset of acceptor states.  
-\item $\mathcal{T}$ is the {\em tape alphabet}, that is a finite non-empty set of symbols that includes the {\em blank symbol} $\square\in\mathcal{T}$ (the only symbol allowed to occur on the tape infinitely often) and the {\em input alphabet} $\Sigma\subseteq\mathcal{T}$, the subset of symbols allowed to initially appear in the tape.
-\item $\delta: (Q\setminus Q_{\bot})\times \mathcal{T}\rightarrow Q\times\{left,right\}\times \mathcal{T}$ is the {\em transition function}.
-If $\delta$ is not defined for the current pair of machine state and tape symbol, then the machine halts.
+\item $\Sigma$ is the {\em tape alphabet}, that is a finite non-empty set of symbols that contains the {\em input alphabet} $\Upsilon\subseteq\Sigma$ (the subset of symbols allowed to initially appear in the tape) and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
+\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}.
+For each possible pair of tape symbol and machine state, $\delta$ defines (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
+If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
 \end{itemize}
@@
 Figure~\ref{tab:tm-anbncn} shows the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
-For each possible pair of tape symbol and machine state, $\delta$ defines: (1) the tape symbol to print at the current position of the header (2) whether the header is shifted {\em left} or {\em right} after the print operation and (3), the new state of the machine after the print operation.
-In this example the {\em tape alphabet} is $\mathcal{T}=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} is $\Sigma=\{a,b,c,\square\}$ and the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the only acceptor state.
+The {\em tape alphabet} is $\Sigma=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} $\Upsilon=\{a,b,c,\square\}$ and the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the only acceptor state.
 
@@
 
-A classical planning frame $\Phi=\tup{F,A}$ encodes the {\em transition function} $\delta$ of a {\em Turing Machine} $\mathcal{M}$ as follows:
+A classical planning frame $\Phi=\tup{F,A}$ can encode the {\em transition function} $\delta$ of a {\em Turing Machine} $\mathcal{M}$ as follows:
 \begin{itemize}
-\item The fluents $F$ are instantiated from a set of four {\em predicates} $\Psi$: {\small\tt (head ?x)} that encodes the current position of the header in the tape.
+\item Fluents $F$ are instantiated from a set of four {\em predicates} $\Psi$: {\small\tt (head ?x)} that encodes the current position of the header in the tape.
 {\small\tt (next ?x1 ?x2)} encoding that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
@@
 {\small\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
-Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine $\mathcal{M}$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
-\item Actions $A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action schema is defined such that:
+Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
+\item Actions $A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action schema is defined:
 \begin{itemize}
-\item The {\bf header} is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$ and the parameters $?xl$, $?x$ and $?xr$ are tape cells.
+\item The {\bf header} is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$. Parameters $?xl$, $?x$ and $?xr$ are tape cells.
 \item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to make $?x$ the tape cell pointed by the header and $?xl$ and $?xr$ its left and right neighbours.
@@
 
-The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.
-The full encoding of the Turing Machine defined in Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema with the same structure as the one of Figure~\ref{fig:update-rule}.
+The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} (the full encoding of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema).
 \begin{figure}[hbt!]
@@
 
-Since the {\em transition function} of a {\em Turing Machine} can be encoded as classical planning frame the execution of that machine can be defined as a {\em plan trace} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
-In this case $s_0$ encodes the initial state of the tape plus the initial machine state and, for each {\small $1\leq i\leq n$}, $a_i\in A$ is applied in $s_{i-1}$ to generate state $s_i$.
+Since the {\em transition function} of a {\em Turing Machine} can be encoded as \strips\ action schemas, executions of that machine are definable as {\em plan traces} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
+In this case $s_0$ encodes the initial state of the tape plus, the position of the header and the initial machine state.
 
 \subsection{Experimental setup}
-We randomly generated $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a two-symbol three-state {\em Turing Machines}.
+We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a two-symbol three-state {\em Turing Machine}.
 We randomly choose a machine $\mathcal{M}\in M$ and produce an fifty-step execution plan trace $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_{50},s_{50}}$.
+The experiment is following our {\em model recognition as planning} method to identify the {\em Turing Machine} that produced $\mathcal{T}$.
+The exeperiment is repeated for different amounts of missing information in the input trace $\mathcal{T}$.
+
+\subsubsection{Reproducibility}
+{\sc Madagascar} is the classical planner we used to solve the instances that result from our compilations for its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
+Due to its SAT-based nature, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because there is no interaction between these actions.
+Actions for programming effects can also be applied in a single planning step, thus significantly reducing the planning horizon.
+
+The compilation source code, evaluation scripts and benchmarks (including the used training and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
+
+\subsection{Results}
 

### 2018-05-22 doc/programming-project/AI-programming.tex
added research project proposal

\documentclass[10pt,a4paper]{paper}
\usepackage[utf8]{inputenc}
\usepackage{graphicx}
\usepackage{subcaption}
\usepackage{enumitem}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage{pgfgantt}
\definecolor{TableOrange}{RGB}{255,151,46}
\definecolor{TableBlue}{RGB}{38,125,184}

\newcommand{\tup}[1]{{\langle #1 \rangle}}

\pdfoptionpdfminorversion=5

\title{Artificial Intelligence for the Automated Synthesis and Validation of Programs}
\author{Sergio Jim\'enez Celorrio\\
\footnotesize Universitat Politécnica de València, Spain\\
\small Area: \underline{Ciencias de la Computación y Tecnología Informática}}

\begin{document}
\maketitle

\begin{abstract}
Programming errors, commonly known as {\em bugs}, cause undesired software behavior making a program crash or enabling malicious users to access private data. Just for 2017, the cumulative cost of software bugs is worldwide estimated in more than one trillion US dollars.

This research project aims to investigate a {\bf novel approach for software development, that leverages Artificial Intelligence, to reduce the chances of introducing programming errors}. The project proposes to addresses automatic program synthesis and validation starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine. Our work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI}, the main international conference on Artificial Intelligence.
\end{abstract}

\section{Historial científico-técnico del Equipo de Investigadores (Apartado 5.1,bases de la convocatoria}

\subsection{Principal Investigator: Sergio Jiménez Celorrio}
Sergio is a {\em Ramón y Cajal} fellow at the Universitat Politècnica de València (2017-2022). Previously, Sergio was a post-doc at the University of Melbourne and a {\em Juan de la Cierva} fellow at the Universitat Pompeu Fabra de Barcelona. From 2004 to 2013 Sergio was assistant lecturer at the Universidad Carlos III de Madrid where he obtained the {\em Distinguished Thesis Award 2011}. Sergio is co-organizer of the {\em $7^{th}$ International Planning Competition} and program committee at top international conferences on Artificial Intelligence.

His research addresses complex {\em decision-making} problems with innovative integrations of two {\em Artificial Intelligence} paradigms: {\em Machine Learning} and {\em AI Planning}. His research trajectory on this topic has produced a novel approach to the synthesis of programs, grammars and controllers for autonomous behavior and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence (IJCAI)\footnote{IJCAI is ranked as {\it A+} by the CORE rank and with {\it h5-index} 43 at Google Scholar}, the main international conference on {\em Artificial intelligence}.

His top five scientific contributions are:
\begin{enumerate}
\item Inductive programming with classical planning. {\footnotesize~\cite{javi-ijcai17},~\cite{segovia2016generalized},~\cite{jimenez2015computing}}.
\item Generalization in AI planning. {\footnotesize~\cite{javi-icaps17},~\cite{damir-derived-ijcai16},~\cite{javi-fsc-ijcai16}}.
\item Learning to relieve the {\it knowledge acquisition bottleneck} in AI planning. {\footnotesize{~\cite{diego-icaps18},~\cite{jimenez2013integrating},~\cite{jimenez2008architecture},~\cite{jimenez2006planning},~\cite{lanchas2007learning}}}.
\item Machine learning for addressing the {\it curse of dimensionality} in AI planning. {\footnotesize~\cite{jimenez2012review},~\cite{de2011scaling},~\cite{de2008learning}}.  
\item Evaluation in AI planning. {\footnotesize~\cite{lopez2015deterministic},~\cite{lopez2013automating},~\cite{coles2012survey}}. 
\end{enumerate}

General quality indicators of scientific research: {\bf 373 citations} (287 since 2011) and {\bf h-index 11}. {\small\em Source google scholar}.

\subsection{Senior Researcher: Eva Onaindia}

\subsection{Researcher in trainnig period: Diego Aineto}

\section{Estado del arte y objetivos específicos ( Apartado 5.2.i, bases de la convocatoria)}

This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} for the automatic synthesis and validation of programs}. Remarkably, the programs synthesized with our approach are guaranteed to be bug-free over given sets of {\em input-output} tests cases.

Here we briefly introduce the technology we will rely on along the project:
\begin{itemize}
\item {\bf AI Planning (AIP)} is the Artificial Intelligence component that studies the synthesis of sets of actions to achieve some given objectives~\cite{ghallab2004automated}. AIP arose in the late ’50s from converging studies into {\em combinatorial search}, {\em theorem proving} and {\em control theory} and now, is a well formalized paradigm for problem solving with algorithms that scale-up reasonably well. State-of-the-art planners are able to synthesize plans with hundreds of actions in seconds time~\cite{geffner2013concise}.  The mainstream approach for AIP is {\em heuristic search} with heuristics derived automatically from the problem representation~\cite{mcdermott1996heuristic,bonet2001planning}.  Current planners add other ideas to this like {\it novelty exploration}~\cite{geffner:psimulators:IJCAI17}, {\it helpful actions}~\cite{hoffmann2001ff}, {\it landmarks}~\cite{helmert2006fast}, and {\it multiqueue best-first search}~\cite{richter2010lama} for combining different heuristics.

\item {\bf Test driven development (TDD)}~\cite{beck:TDD:2003} is a popular paradigm for software development that is frequently used in {\it agile methodologies}~\cite{cohen2003agile}. In TDD, test cases are created before the program code is written and they are run against the code during the development, e.g. after a code change via an automated process. Tests alert programmers of bugs before handing the code off to clients (the cost of finding a bug when the code is first written is considerably lower than the cost of detecting and fixing it later). When all tests pass, the program code is considered {\em complete} while when a test fails, it pinpoints a {\em bug} that must be fixed from the program code. Tests cases are a natural form of program specification, programmers often claim {\em 'code that is difficult to test is poorly written'}. Further, writing a thorough set of tests cases forces programmers to think through inputs, outputs, and error conditions of programs. 
\end{itemize}

Now we review the two most successful approaches for program synthesis:
\begin{itemize}
\item{\bf Programming by Example (PbE)}. PbE techniques have already been deployed in the real world and are part of the {\sc Flash Fill} feature of Excel in Office 2013 that generates programs for string transformation~\cite{gulwani2011automating}. In this case the set of synthesized programs are represented succinctly in a restricted Domain-Specific Language (DSL) using a data-structure called version space algebras~\cite{mitchell1982generalization}. The programs are computed with a domain-specific search that implements a divide and conquer approach. The approach posed in this research project falls into this category.

\item In {\bf Programming by Sketching (PbS)} programmers provide a partially specified program, i.e.~a program that expresses the high-level structure of an implementation but that leaves low level details undefined to be determined by the synthesizer~\cite{solar2006combinatorial}. This form of program synthesis relies on a programming language called {\sc SKETCH}, for sketching partial programs. PbS implements a counterexample-driven iteration over a synthesize-validate loop built from two communicating SAT solvers, the inductive synthesizer and the validator, to automatically generate test inputs and ensure that the program satisfy them. Despite, in the worst case, program synthesis is harder than NP-complete, this counterexample-driven search terminates on many real problems after solving only a few SAT instances~\cite{lake2015human}.
\end{itemize}

Closely related to the aims of the project is the work on the synthesis and validation of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}. The existing algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}. To keep the computation of FSCs tractable, they limit the space of possible solutions bounding the maximum size of the FSC. In addition, they impose that the instances to solve share, not only the domain theory (actions and predicates schemes) but the set of fluents~\cite{sergio:aprograming:icaps16} or a subset of {\it observable} fluents~\cite{Geffner:FSM:AAAI10}.

The computation of FSCs for generalized planning includes works that compile the generalized planning task into another forms of problem solving so they benefit from the last advances on off-the-shelf solvers (e.g. {\em classical planning}~\cite{sergio:aprograming:icaps16}, {\em conformant planning}~\cite{Geffner:FSM:AAAI10}, {\em CSP}~\cite{Infantes:FSC:ECAI2010} or a {\em Prolog program}~\cite{Giacomo:FSM:ICAPS13}). This last case requires a behavior specification of the FSC consisting on classified execution histories that (1) accept all legal execution histories leading to a goal-satisfying state, and (2) reject those that contain repeated configurations (indicating an infinite loop) and that cannot be extended (indicating a dead end)~\cite{Giacomo:FSM:ICAPS13}.

Finally the generation of programs from examples is a research question also addressed in inductive Logic Programming (ILP). ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn a given target concept, expressed as a  logic program,  from  examples  and  background  knowledge, that are expressed as logic facts~\cite{muggleton1991inductive}.

\section{Metodología y Plan de Trabajo ( Apartado 5.2.ii, bases de la convocatoria)}
\label{subsec:metodologia}
Our research shows that current AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16}. To illustrate this, Table~\ref{tab:programs} reports the invested time to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list or traversing a binary tree.

\begin{table*}[hbt!]
  \centering
\begin{tt}  
\begin{tabular}{l@{\hspace*{30pt}}c@{\hspace*{5pt}}}
 \textbf{Programming Task} & \textbf{Time (seconds)} \\\hline
Summatory		&	1\\
Fibonacci		&	5\\
Reverse			&	22\\
Find                    &       336 \\
Minimum                 &       284 \\
Sorting			&	30\\
Tree  		&	165
\end{tabular}
\end{tt}  
\caption{\small Time to synthesize the programs with the {\sc FD} AI planner~\cite{helmert2006fast} on a processor {\em Intel Core i5 3.10GHz x 4} and with a 4GB memory bound.}
\label{tab:programs}
\end{table*}

Our approach is modeling and solving a given TDD programming task as an AI planning task. Briefly, the state variables of these AIP tasks are {\em fluents} of the kind {\tt (var:=value)} while the initial/goal states of the AIP tasks encode the programming input/output tests. Finally, program instructions are encoded using two kinds of AIP actions:
\begin{enumerate}
\item {\it Programming actions}, that set an instruction on a given program line.
\item {\it Execution actions}, execute the instruction set on a given program line.
\end{enumerate}
We implement this encoding using standard planning languages, such as PDDL~\cite{fox2003pddl2}, so the AIP tasks resulting from our encoding can be solved with off-the-shelf planners, like {\sc FD}~\cite{helmert2006fast}. Interestingly, the encoding allows also program validation by (1), specifying the program to validate in the initial state of the AIP task and (2), disabling the mentioned {\it programming actions}.

%The space of the possible programs encoded in such a way grows exponentially in the number of possible program lines. To address challenging tasks our approach must then be combined with {\em problem decomposition}. With this regard, our encoding supports callable procedures to decompose a given programming task in simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.

\subsection{Two-year workplan}
We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that (1), takes as input a set of test cases that specify the TDD programming task to solve and (2), outputs a program code that passes these tests with a bug-free guarantee. Figure~\ref{fig:gantt} details the proposed 24-month timeline for the project.

\begin{figure}[hbt!]
\begin{ganttchart}[
  hgrid,
  group progress label node/.append style={below=3pt},
  canvas/.append style={label=below:} ]{1}{12} 
\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Design (6 months)}{1}{3} \\
\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Development (6 months)}{4}{6}\\
\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Experiments (12 months)}{7}{12}
\end{ganttchart}
\caption{\small Twentyfour-month workplan for the development of a user-interactive program synthesizer.}
\label{fig:gantt}
\end{figure}

\begin{enumerate}
\item {\bf T1. System design (months 1-6})
  \begin{small}
    \begin{enumerate}
    \item Design of the user-interaction for the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases and the available instruction set. 
    \item Experimental design. The experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve a given set of TDD programming tasks:
      \item Evaluation of different AI planners available, with special attention to the planners that get the best results at the IPC-2018. 
      \end{enumerate}
  \end{small}

{\small {\bf Deliverable:} Technical report with the specifications of the system design.}

  \item {\bf T2. Development of the system architecture (months 7-12})
    \begin{small}
      \begin{enumerate}
      \item The programming-into-planning compiler ({\em Compiler 1}). This component parses the {\em TDD programming task} and produces an {\em AIP task} encoded in the standard planning language PDDL.
      \item The plan-into-program compiler ({\em Compiler 2}). This component extracts the program code from the solution plan produced by an off-the-shelf AI planner.
      \end{enumerate}
\end{small}

\begin{figure}[hbt!]
\tikzstyle{block} = [draw, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804, rectangle, minimum height=3em, minimum width=6em]
\tikzstyle{input} = [coordinate]
\tikzstyle{output} = [coordinate]
\begin{center}
\begin{tikzpicture}[auto, node distance=2cm,>=latex']
    % We start by placing the blocks
    \node [input, name=input] {};
    \node [block, right of=input, node distance=4cm] (compiler1) {Compiler 1};
    \node [block, below of=compiler1, node distance=2cm] (planner) {AI Planner};    
    \node [block, below of=planner, node distance=2cm] (compiler2) {Compiler 2};
    \node [output, right of=compiler2, node distance=3cm] (output) {Program};

    % Once the nodes are placed, connecting them is easy. 
    \draw [->] (input) -- node {Programming task} (compiler1);
    \draw [->] (compiler1) -- node[] {AIP task} (planner);
    \draw [->] (planner) -- node[] {Solution plan} (compiler2);        
    \draw [->] (compiler2) -- node[] {Program}(output);
\end{tikzpicture}
\end{center}  
\caption{\small The system architecture.}
\label{fig:architecture}
\end{figure}

{\small {\bf Deliverable:} Open repository with the source code of the system architecture.}

\item {\bf T3. Experiments and dissemination of results (months 13-24}). 
{\small {\bf Deliverable:} Final report with the obtained conclusions and produced publications.}
\end{enumerate}

\section{Objetivos específicos del Proyecto ( Apartado 5.2.iii, bases de la convocatoria)}
\label{subsec:objectivos}
To precisely characterize the specific kind and size of the programs that we can synthesize and validate, we model programs as {\em finite state machines}~\cite{sergio:aprograming:ijcai16}. Figure~\ref{fig:list} shows a program for traversing linked lists pictured as a {\em finite state machine}. The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition under which program instructions are taken.

\begin{figure}[hbt!]
\begin{center}
\begin{tikzpicture}
\node [thick,draw,circle] at (1,1.1) (x) {$q_0$};

\node [thick,draw,circle] at (6,1.1) (y) {$q_1$};

\node [thick,draw,circle] at (11,1.1) (z) {$q_2$};

\draw (11,1.1) circle (.3cm);

\draw [thick,->] (x) -- (y);

\draw [thick,out=345,in=195,->] (x) to (z);

\draw [thick,out=165,in=15,->] (y) to (x);

\node at (4.1,0.85) {$!\mathsf{isEnd}(n)/\mathsf{visit}(n)$};
\node at (3.5,1.8) {$-/\mathsf{step}(n)$};
\node at (6,0) {$\mathsf{isEnd}(n)/a_\bot$};
\end{tikzpicture}
\end{center}
\caption{\small {\em Finite state machine} that depicts a program to solve the programming task of traversing a linked list ($n$ is a variable that points to a node of the linked list).}
\label{fig:list}
\end{figure}

Formally, we define a program as a tuple $\tup{\mathcal{Q},q_0,Q_\bot,\mathcal{V},,\mathcal{I},\Omega,T,\Gamma}$:
\begin{itemize}
\item $\mathcal{Q}$ is a finite set of program lines where $q_0\in \mathcal{Q}$ is the {\em initial program line} and $Q_\bot\subseteq \mathcal{Q}$ is the subset of {\em terminal program lines}.
\item $\mathcal{V}$ is the finite set of program variables, where the value of each variable $v\in\mathcal{V}$ is defined in the domain $D_v$.
\item $\mathcal{I}$ is a finite instruction set where each instruction $a\in \mathcal{I}$ defines an {\em update} $\theta_a$ of a subset of the program variables.
\item $\Omega:Q\times D_{v_1}\times \ldots \times D_{v_{|\mathcal{V}|}}\mapsto\{0,1\}$ is an {\em observation function} that maps a program line and a condition over the program variables into a Boolean (1 if the condition holds 0 if  it does not). 
\item $T:Q\times\{0,1\}\rightarrow Q$ is a {\em transition function} mapping pairs of a program line and the result of a {\em condition} check into the next program line.
\item $\Gamma:Q\times\{0,1\}\rightarrow \mathcal{I}$ is a {\em transition function} mapping pairs of a program line and the {\em condition} result into the corresponding instruction.
\end{itemize}

In the example of Figure~\ref{fig:list}, the edge $(q_0,q_2)$ with label $\mathsf{isEnd}(n)/a_\bot$ encodes that the condition is $\mathsf{isEnd}(n)?$, while $T(q_0,1)=q_2$, $\Gamma(q_0,1)=a_\bot$ encode the transition and associated instruction when $\mathsf{isEnd}(n)$ holds. The edge $(q_0,q_1)$ with label $!\mathsf{isEnd}(n)/\mathsf{visit}(n)$ encodes the transition and instruction when $\mathsf{isEnd}(n)$ does not hold, i.e.~$T(q_0,0)=q_1$ and $\Gamma(q_0,0)=\mathsf{visit}(n)$. The edge $(q_1,q_0)$ with label $-/\mathsf{step}(n)$ denotes that, when in $q_1$, the transition and instruction are always the same no matter the value of the program variables.

The {\em state of a program} is a pair $(q,s)$ that consists of the program line $q\in \mathcal{Q}$ and $s\in D_{v_1}\times \ldots \times D_{v_{|\mathcal{V}|}}$, an assignment to the variables compatible with their domains. Programs transition to a program state $(q',s')$ that is computed as follows:
\begin{enumerate}
\item First we observe whether the condition associated to the current progam line holds (we denote the result of this observation as $O_q(s)\in \{0,1\}$). 
\item Then we compute the new program line $q'=T(q,O_q(s))$ and the instruction $a=\Gamma(q,O_q(s))$ to apply next. Each transition $T(q,O_q(s))$ is associated with a single observation of the variable values however, programs can represent expressive conditions including the no-op action ($a_{\bot}\in\mathcal{I}$) in the instruction set. The no-op action does not modify the program variables and affects only to the next program line which allows to concatenate conjunctions of conditions like in {\em decision trees}.
\item Finally the value of the program variables $s'=\theta_a(s)$ is updated executing the corresponding instrucntion $a=\Gamma(q,O(s,q))\in \mathcal{I}$.
\end{enumerate}

The specific objective of this project is the study of the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions:
\begin{enumerate}
\item $|\mathcal{Q}|$, the maximum number of program lines.
\item $|\mathcal{V}|$, the maximum number of program variables. 
\item $|\mathcal{I}|$, the maximum size of the available instruction set.  
\end{enumerate}
Note that despite setting bounds for $|\mathcal{Q}|$, $|\mathcal{V}|$ and $|\mathcal{I}|$, challenging programming tasks can still be addressed using {\em problem decomposition}. With this regard, our AIP encoding supports callable procedures to decompose a given programming task in simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.

In this project we aim to {\bf study the performance of an Artificial Intelligence approach for the automated synthesize and validation of programs up to}: \underline{5 program lines}, \underline{50 variables with binary domain} and, an \\\underline{instruction set that comprises 10 instructions}. The performance of our approach will be evaluated with regard to the {\em computation time} and {\em memory} invested in the synthesize and validation of the characterized programs.

\section{Beneficios del proyecto. Difusión y Explotación en su caso de los Resultados.}
\label{subsec:beneficios}
This project will {\bf provide new insights into the current understanding of how Artificial Intelligence can assist programmers in the software development}.

Research in AI algorithms is too often tested with laboratory problems and AIP is not an exception. Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition (IPC)~\cite{vallati:IPC:AI15}. This project will also help to {\bf meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks}.

Promising research opportunities come from the application of AIP to {\em program synthesis} given that, {\em program synthesis} with a tests base, can also be seen as a dual to {\em program testing}. AIP has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}.

Our current work on {\em program synthesis} with AIP already produced {\bf publications at top international conferences on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}. The scientific results obtained during the development of the project will be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).

\begin{footnotesize}
\bibliography{AI-programming}
\end{footnotesize}
\bibliographystyle{ieeetr}

\end{document}

### 2018-05-23 src/compiler_new.py
compiler_new.py: added modeProg to learning_problem init

@@ -359,6 +359,8 @@ for i in range(2, MAX_ISTEPS+1):
 for i in range(1, MAX_ISTEPS + 1):
     learning_task.objects.append(pddl.pddl_types.TypedObject("i" + str(i), "step"))
 
+learning_task.init.append(pddl.conditions.Atom("modeProg", []))
+
 
 ### Write the learning task domain and problem to pddl
 fdomain = open("learning_domain.pddl", "w")

### 2018-05-23 doc/aij18/learning_task.tex

@@
 {\small $1\leq i\leq n$}, can be {\em partially observable} (some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative).
-%In the extreme, entire states $s_i$ and actions $a_i$ {\small $1\leq i\leq n$}, can be missing.
-In such scenario solving $\Lambda$ implies determining, not only the \strips\ action model $\mathcal{M}'$, but also the unobserved plan $\pi$, that explains the input observations with the learned \strips\ model.
+In the extreme, all actions $a_i$, {\small $1\leq i\leq n$}, can be missing (unobserved) provided that the final state $s_n$ is at least, partially observed.
+Likewise, all states $s_i$, {\small $1\leq i\leq n$}, can be missing (unobserved) provided that at least actions $a_i$, {\small $1\leq i\leq n$}, are partially observed.
  \item Observations in $\mathcal{T}$ are {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 

### 2018-05-23 src/compiler_new.py
compiler_new.py: added negative literals to the state

@@ -68,7 +68,7 @@ domain_name, domain_requirements, types, type_dict, constants, predicates, predi
 traces = list()
 for filename in sorted(glob.glob(domain_folder_name + "/trace" + "*")):
     trace_pddl = pddl_parser.pddl_file.parse_pddl_file("trace", filename)
-    traces.append(pddl_parser.parsing_functions.parse_trace_pddl(trace_pddl, action_observability, state_observability))
+    traces.append(pddl_parser.parsing_functions.parse_trace_pddl(trace_pddl, predicates, action_observability, state_observability))
 
 
 MAX_VARS = get_max_vars(actions)
@@ -420,8 +420,7 @@ counter = 0
 new_fd_task = copy.deepcopy(original_task)
 new_fd_task.actions = []
 for action in actions:
-    params = ["?o" + str(i + 1) for i in range(0, len(action[1:]))]
-    ps = [pddl.pddl_types.TypedObject(params[i], action[i + 1]) for i in range(0, len(params))]
+    ps = [pddl.pddl_types.TypedObject("?o"+str(i+1), action.parameters[i].type_name) for i in range(action.num_external_parameters)]
     pre = []
 
     for p in pres[counter]:
@@ -449,7 +448,7 @@ for action in actions:
                 ball = False
         if ball:
             eff = eff + [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom(p[0], args))]
-    new_fd_task.actions.append(pddl.actions.Action(action[0], ps, len(ps), pddl.conditions.Conjunction(pre), eff, 0))
+    new_fd_task.actions.append(pddl.actions.Action(action.name, ps, len(ps), pddl.conditions.Conjunction(pre), eff, 0))
     counter = counter + 1
 
 # new_fd_task.actions.extend(known_action_models)

### 2018-05-23 src/pddl_parser/parsing_functions.py

@@ -7,6 +7,7 @@ import sys
 import graph
 import pddl
 from random import random, seed
+import itertools
 
 def parse_typed_list(alist, only_variables=False,
                      constructor=pddl.TypedObject,
@@ -490,7 +491,7 @@ def parse_task_pddl(task_pddl, type_dict, predicate_dict):
         assert False, entry
 
 
-def parse_trace_pddl(trace_pddl, action_observability=1, state_observability=1):
+def parse_trace_pddl(trace_pddl, predicates, action_observability=1, state_observability=1):
     seed(123)
 
     iterator = iter(trace_pddl)
@@ -502,19 +503,34 @@ def parse_trace_pddl(trace_pddl, action_observability=1, state_observability=1):
     assert objects_opt[0] == ":objects"
     object_list = parse_typed_list(objects_opt[1:])
 
+    all_literals = set()
+    for predicate in predicates:
+        args = list()
+        for i in range(len(predicate.arguments)):
+            iargs = list()
+            for object in object_list:
+                if object.type_name == predicate.arguments[i].type_name:
+                    iargs.append(object.name)
+            args.append(iargs)
+
+        for tup in itertools.product(*args):
+            all_literals.add(pddl.Atom(predicate.name, tup))
+
+
+
     init = next(iterator)
     assert init[0] == ":init"
-    initial = parse_state(init[1:])
+    initial = parse_state(init[1:], all_literals)
 
     actions = list()
     states = list()
 
     for token in iterator:
         if token[0] == ':observations':
-            new_state = [literal for literal in parse_state(token[1:]) if random() <= state_observability]
+            new_state = [literal for literal in parse_state(token[1:], all_literals) if random() <= state_observability]
             states.append(new_state)
         elif token[0] == ':goal':
-            goal = parse_state(token[1:])
+            goal = parse_state(token[1:], all_literals)
             break
         else:
             if random() <= action_observability:
@@ -527,7 +543,7 @@ def parse_trace_pddl(trace_pddl, action_observability=1, state_observability=1):
     return pddl.Trace(object_list, initial, goal, actions, states)
 
 
-def parse_state(new_state):
+def parse_state(new_state, all_literals):
     state = []
     state_true = set()
     state_false = set()
@@ -564,6 +580,8 @@ def parse_state(new_state):
             check_atom_consistency(atom, state_true, state_false)
             state_true.add(atom)
     state.extend(state_true)
[…]

### 2018-05-23 doc/aij18/background.tex

@@
 In subsection \ref{basic_planning} we introduce the basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper.
-Section \ref{related_work} summarizes existing approaches to learning action models in classical planning and highlights the novelty of our contribution via the related work.
+Section \ref{related_work} summarizes existing approaches to learn action models in classical planning and highlights the novelty of our contribution via the related work.
 
@@
 
-A \emph{plan trace} is a combination of a state and an action sequence $\tau = \langle s_0, a_1, a_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
-Additionally, $\tau$ is a plan trace with PO states if $\tup{s_0, s_1, \ldots, s_n}$ is a PO state sequence and a plan trace with PO actions if $\tup{a_1, \ldots, a_n}$ is a PO action sequence.
+A \emph{plan trace} is a combination of a state and an action sequence $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
+A plan trace $\tau$ for a planning problem $P=\tup{F,A,I,G}$ satisfies $s_0=I$, $G \in s_n$ and every element in the action sequence of $\tau$ is an action of $A$.
+Additionally, $\tau$ is a plan trace with PO states if $\tup{s_0, s_1, \ldots, s_n}$ is a PO state sequence; and a plan trace with PO actions if $\tup{a_1, \ldots, a_n}$ is a PO action sequence.
 
@@
 
-One can find many different approaches to learning action models in the literature.
-A common but distinguishing feature to all approaches is the type of input knowledge required by the each model.
-Due to the large variety of types of input knowledge and the lack of a common naming convention (e.g., plan examples, plan traces, training data, plan samples), we firstly present a glossary of terms that will ease the classification of the surveyed models accordingly to their input knowledge.
+In this section we summarize the most recent and relevant approaches to learning action models found in the literature.
+Approaches will be examined according to the following parameters: the input knowledge (plan traces) accepted by the system, the expressiveness of the learned action model and the principal technique used for learning the action model (Table \ref{table:models_comparison1}), as well as the characteristics of the assessment method to validate the learned models (Table \ref{table:models_comparison2}).
+
+Regarding the input traces, all the approaches analyzed in this work accept partial observations of intermediate states, and some approaches also allow the sequence of intermediate states to be empty.
+With the exception of \LOCM, the minimum state sequence of a plan trace required by the approaches that appear labeled as PO states* in the first column of Table \ref{table:models_comparison1} is $\tup{s_0, G}$.
+Additionally, most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces.
+Others do not explicitly say so but the fact is that the predicates and the actions headers are easily extractable from the state sequence and action sequence of the plan traces, respectively.
+A requirement, however, for these two sets to be representative is that set of input plan traces comprises a grounded sample of all predicates and operator schemas of the domain model.
+
+The expressiveness of the learned action models varies across approaches (second column of Table \ref{table:models_comparison1}).
+All the presented systems are able to learn action models in a \textsc{Strips}  representation \cite{fikes1971strips} and some propose algorithms to learn more expressive action models that include quantifiers, logical implications or the type hierarchy of the domain.
+
+Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- most approaches rely on a comparison between the learned model and a ground-truth model), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
+
+In the following, we present a comprehensive insight of the particularities of the seven systems presented in Table \ref{table:models_comparison1} and Table \ref{table:models_comparison2}.
+This exposition will help us to highlight at the end of the section the value of our contribution \FAMA.
+
+\vspace{0.3cm}
+
+The Action-Relation Modelling System (\textbf{\ARMS}) \cite{yang2007learning} is one of the first learning algorithms able to learn from plan traces with partial or null observations of intermediate states.
+\ARMS uncovers a number of constraints from the plan traces in the training data that must hold for the plans to be correct.
+These constraints are then used to build and solve a weighted propositional satisfiability problem with a MAX-SAT solver.
+Three types of constraints are considered: 1) constraints imposed by general axioms of correct \textsc{Strips} actions, 2) constraints extracted from the distribution of actions in the plan traces and 3) constraints obtained from the PO states, if available.
+Frequent subsets of actions in which to apply the two latter types of constraints are found by means of frequent set mining.
+
+\ARMS defines an error metric and a redundancy metric to measure the correctness and conciseness of an action model over the test set of input plan traces using a cross-validation evaluation.
+The model evaluation is posed as an optimization task that returns a learned action model that minimizes the error and redundancy functions.
+This yields a model that is approximately correct (100\% correctness is not required so as to ensure generality and avoid overfitting) and approximately concise (low redundancy rates), and that can explain as many examples as possible.
+Hence, there is no guarantee that the learned model of \ARMS explains all observed plans, not even that it correctly explains any of the plan traces of the test set.
+
+The \ARMS system became a benchmark in action-model learning, showing empirically that is is feasible lo learn the models in a reasonably efficient way using a weighted MAX-SAT even when the intermediates states are not observed.
+Nevertheless, we argue that using the home-made error and redundancy metrics to evaluate the model correctness is questionable since a model that exhibits a wrong precondition or effect in any of its actions could still be eligible as the minimal-error output model.
 
@@
 	\centering
-	\begin{tabular}{ l | c | c | c | c | c |}
-		& \multicolumn{1}{|p{2cm}|}{FO States sequences}
-		& \multicolumn{1}{|p{2cm}|}{FO Actions sequences}
-		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO states)}
-		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO actions)}
-		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO both)}   \\
+	\begin{tabular}{ l | c | c | c }
+		& \multicolumn{1}{c}{Input plan traces}
+        & \multicolumn{1}{c}{Learned action model}
+        & \multicolumn{1}{c}{Technique}     \\
 		\hline			
-		ARMS &  & \checkmark & \checkmark & & \\
-		AMAN & & & \checkmark & &\\
-		LAMP & & & \checkmark & &\\
-		SLAF & & & \checkmark & &\\
-		LOCM & & \checkmark & & &\\
-		FAMA & \checkmark & \checkmark &  \checkmark & \checkmark & \checkmark \\
+		\multirow{2}{*}{\ARMS} & PO states* & \strips & MAX-SAT\\ & FO actions & & \\
+        \hline
+        \multirow{2}{*}{\SLAF} & PO states & universal quantifiers in $\eff$ & bayesian learning \\ & FO actions &  & logical filtering \\
+         \hline
+        %\multirow{2}{*}{\SLAF} & PO states and FO actions & existential quantifiers in $\pre$ \\ & & universal quantifiers in $\eff$ &  bayesian learning \\ & & & logical filtering \\
+         %\hline
+		\multirow{2}{*}{\LAMP} & PO states* &  quantifiers &  Markov logic networks \\  & FO actions & logical implications &  \\
+         \hline
+         \AMAN & PO states* & \strips & graphical model estimation \\ & noisy actions & & \\
+         \hline
+         \NOISTA & PO and noisy states & \strips &  classification \\ & FO actions & & \strips \texttt{} rules derivation \\
+         \hline
+         \CAMA & PO states &  \strips & crowdsourcing annotation\\ & FO actions &  & MAX-SAT \\
+         \hline
+         \LOCM &  FO actions &  predicates and types & \\
+         \hline
+		\FAMA & PO* states & \strips &   compilation to planning\\ & PO* actions & & \\
+         \hline
 	\end{tabular}
-	\caption{Input comparison of main learning approaches. PO=Partial Observability, FO=Full Observability}
-	\label{table:input_comparison}
+	\caption{Characteristics of action-model learning approaches (PO: Partial Observability: FO: full observability; *: accepts null observability)}
+	\label{table:models_comparison1}
 \end{table}
+
+A tractable and exact solution of action models in partially observable domains using a technique known as Simultaneous Learning and Filtering (\textbf{\SLAF}) is presented in~\cite{AmirC08}.
+\SLAF alongside \ARMS can be considered another of the precursors of the modern algorithms for action-model learning, able to learn from plan traces with PO and non-null intermediate states.
+Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, \SLAF builds a complete explanation of observations by models of actions through a CNF formula.
+The learning algorithm updates the formula of the belief state with every action and observation in the sequence such that the new transition belief formula represents all possible transition relations consistent with the actions and observations at every time step.
+
+\SLAF extracts all satisfying models of the learned formula with a SAT solver.
+For doing so, the training data set for each domain is composed of randomly generated action-observation sequences  (1000 randomly selected actions and 10 fluents uniformly selected at random per observation).
+Additional processing in the form of replacement procedures or extra axioms are run into the SAT solver when finding the satisfying models.
+The experimentally tested \SLAF version is an algorithm that learns only effects for actions that have no conditional effects and assumes that actions in the sequences are all executed successfully (without failures).
+This algorithm cannot effectively learn the unknown preconditions of the actions and in the resulting models `\emph{one can see that the learned preconditions are often inaccurate}` \cite{AmirC08}.
+On the other hand, it does not report any statistical evaluation of measurement error other than a manually comparison of the learned models with a ground-truth model.
+
+The Learning Action Models from Plan Traces (\textbf{\LAMP}) \cite{ZhuoYHL10} algorithm extends the expressiveness to learning models with universal and existential quantifiers as well as logical implications.
+The input to \LAMP is a set of plan traces with PO intermediate states, which are encoded by the algorithm into propositional formulas.
+\LAMP then uses the action headers and predicates to build a set of candidate formulas that are validated against the input set using a Markov Logic Network and effectively weighting each formula.
+The formulas with weights larger than a certain threshold will be chosen to represent preconditions and effects of the learned action models.
+
+\LAMP allows observations of intermediate states to be empty up to a minimum percentage of 1/5 of observed states as well as different degrees of PO in the number of propositions in each state.
+It uses an error metric based on counting the differences in the number of precondition and effects between the ground-truth model and the learned model.
+In general, the results show that the accuracy of the learned models is fairly sensitive to the threshold chosen to learn the weights of the candidate formulas, and that domains that feature more conditional effects are harder to learn.
 
@@
 	\centering
-	\begin{tabular}{ l | c | c | c |}
-		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO states)}
-		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO actions)}
-		& \multicolumn{1}{|p{2cm}|}{Plan traces (PO both)}   \\
+	\begin{tabular}{ l | c | c | c }
+		%& \multicolumn{1}{c}{Evaluation task}
+        & \textbf{Evaluation method} & \textbf{Metrics} & \textbf{\#tested domains/}   \\
+        &   &   & \textbf{training data size} \\
 		\hline			
-		ARMS & \checkmark & & \\
-		AMAN & \checkmark & & \\
-		CAMA & \checkmark & & \\
-		LAMP & \checkmark & & \\
-		SLAF & \checkmark & & \\
-		LOCM & \checkmark & & \\
-		FAMA & \checkmark & \checkmark & \checkmark \\
+		\multirow{1}{*}{\ARMS} & cross-validation with a test set & error counting of \#$\pre$ satisfaction  & 6  \\ 
+        & of plan traces & and redundancy & 160 plan traces\\
+        \hline
+         \SLAF &  manual checking wrt GTM &  ---   & 4\\ & & & 1,000 actions\\
+         \hline
+		\multirow{1}{*}{\LAMP} & checking wrt GTM  & error counting of extra & 4\\ & & and missing \#$\pre$ and \#$\eff$ & 100-200 plan traces\\
+         \hline
+         \AMAN & checking wrt GTM &  error counting of extra &  3 \\ & & and missing \#$\pre$ and \#$\eff$ & 40-200 plan traces\\
+         \hline
+         \NOISTA & checking wrt GTM  & error counting of extra & 5\\ & & and missing \#$\pre$ and \#$\eff$ & 5,000-20,000 actions\\
+         \hline
+         \CAMA & checking wrt GTM  &  error counting of extra & 3 \\ & & and missing \#$\pre$ and \#$\eff$ & 15-75 plan traces\\
+         \hline
+		\LOCM & checking wrt GTM & &   \\
+         \hline
+		\FAMA & checking wrt GTM  & precision and recall of & 15\\ & & \#$\pre$ and \#$\eff$ & 25 actions\\
+         \hline
 	\end{tabular}
-	\caption{PO=Partial Observability, FO=Full Observability}
-	\label{table:input_comparison2}
+	\caption{Evaluation of action models (GTM: ground-truth model)}
+	\label{table:models_comparison2}
 \end{table}
 
-\begin{table}
-	\small
-	\centering
-	\begin{tabular}{ l | c | c |}
-		& States
-		& Actions  \\
-		\hline
-		LOCM & zero & FO\\		
-		ARMS & zero or PO & FO \\
-		AMAN & PO & FO \\
-		CAMA & PO & FO \\
-		LAMP & PO & FO \\
-		SLAF & PO & FO \\		
-		FAMA & zero or PO & zero or PO \\
-	\end{tabular}
-	\caption{PO=Partial Observability, FO=Full Observability}
-	\label{table:input_comparison3}
-\end{table}
+The Action Model Acquisition from Noisy plan traces (\textbf{\AMAN}) \cite{zhuo2013action} introduces an algorithm able to learn action models from plan traces with null observations of intermediate states where actions have a probability of being observed incorrectly (noisy actions).
+The first step of the \AMAN algorithm is to build the set of candidate domain models that are compliant with the action headers and predicates.
+\AMAN then builds a graphical model to capture the domain physics; i.e., the relations between states, correct actions, observed actions and domain models.
+After that, the parameters of the graphical model are learned, computing at the same time the probability distribution of each candidate domain model.
+Finally \AMAN returns the model that maximizes a reward function defined in terms of the percentage of actions successfully executed and the percentage of goal propositions achieved after the last successfully executed action.
 
-===============================================
+\AMAN uses the same metric as \LAMP, namely counting the number of preconditions and effects that appear in the learned model and not in the ground-truth model (extra fluents) and viceversa (missing fluents).
+In a comparison between \AMAN and \ARMS on noiseless inputs, the results show that the accuracies of both \AMAN and \ARMS are very close to each other and neither dominates the other.
+On the other hand, a test comparing plan traces with and without noisy actions reveals that the accuracy in the case \emph{with noise} becomes more and more close to \emph{without noise} as the number of iterations of \AMAN algorithm increases because the distribution of noise in the plan becomes gradually closer to the real distribution.
 
-The Crowdsourced Action-Model Acquisition (\CAMA) for planning explores knowledge from both crowdsourcing (human annotators) and plan traces to learn action models for planning.
+Another interesting approach that deals with noisy and incomplete observations of states is presented in \cite{MouraoZPS12}.
+We will refer to this approach as \textbf{\NOISTA} henceforth.
+In \NOISTA, actions are correctly observed but they can obviously be unsuccessfully executed in the possibly noisy application state.
+The basis of this approach consists of two parts: a) the application of a voted Perceptron classification method to predict the effects of the actions in vectorized state descriptions and b) the derivation of explicit \strips \texttt{} action rules to predict each fluent in isolation.
+Experimentally, the error rates in \NOISTA fall below 0.1 after 5,000 training samples for the five tested domains under a maximum of 5\% noise and a minimum of 10\% of observed fluents.
+
+The Crowdsourced Action-Model Acquisition (\textbf{\CAMA}) \cite{Zhuo15} explores knowledge from both crowdsourcing (human annotators) and plan traces to learn action models for planning.
 \CAMA relies on the assumption that obtaining enough training samples is often difficult and costly because there is usually a limited number of plan traces available.
-In order to overcome this limitation, \CAMA builds on a set of soft constraints based on labels \texttt{true} or \texttt{false} given by the crowd and a set of soft constraints based on the input plan traces, solves then using a MAX-SAT solver and converts the solution to action models.
+In order to overcome this limitation, \CAMA builds on a set of soft constraints based on labels \texttt{true} or \texttt{false} given by the crowd and a set of soft constraints based on the input plan traces.
+Then it solves the constraint-based problem using a MAX-SAT solver and converts the solution to action models.
 
-\CAMA uses plan traces with PO states.
-To capture the knowledge from the crowd, \CAMA enumerates all possible preconditions and effects for each action, considering all predicates whose parameters are included by the parameters of the action.
-Hence, if the parameters of a predicate $p$ are included by the parameters of $a$, \CAMA queries the crowd for three questions of the form "Is the fact $p$ a precondition/positive effect/negative effect of action $a$?", and human annotators reply "yes", "no" or "cannot tell".
+To capture the knowledge from the crowd, \CAMA enumerates all possible preconditions and effects for each action, considering all predicates $p$  whose parameters are included by the parameters of the action $a$.
+\CAMA queries the crowed whether $p$ is a precondition/positive effect/negative effect of $a$, and human annotators reply \emph{yes}, \emph{no} or \emph{cannot tell}.
 The goal is to find an optimal estimator of the \texttt{true} labels given the observation, minimizing the average bit-wise error rate.
+An experimental comparison of a manual crowdsourcing data of \CAMA with \ARMS reveals that \CAMA outperforms \ARMS and that as expectedly the difference becomes smaller as the number of plan traces becomes larger.
+The accuracy of \CAMA for a small number of plan traces (e.g., 30) is not less than 80\%, which means that exploiting the knowledge of the crowd can help learning action models.
 
-================================================
+\vspace{0.5cm}
 
-Back in the 90's various systems aimed learning operators mostly via interaction with the environment.
-{\sc LIVE} captured and formulated observable features of objects and used them to acquire and refine operators \cite{ShenS89}.
-{\sc OBSERVER} updated preconditions and effects by removing and adding facts, respectively, accordingly to observations \cite{Wang95learningby}.
-These early works were based on lifting the observed states supported by exploratory plans or external teachers, but none provided a theoretical justification for this second source of knowledge.
+==========================================================
 
-More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
+The Learning Object-Centred Models (\LOCM) \cite{cresswell2013acquiring} system does not limit itself to learning action models and instead is able to learn full domains from plan traces without any observation of the states.
+Furthermore, \LOCM does not require any domain background knowledge such as predicates or objects and to this day can be considered the learning algorithm requiring the least amount of input data.
+\LOCM works by making some assumptions about the structure of the actions.
+Objects found in the same position in the header of actions are grouped into sorts and state machines describing the behavior of objects are built for each sort.
+In a later work \LOCMtwo extends the applicability of the \LOCM algorithm to a wider range of domains by introducing a richer representation using multiple state machines.
 
-Whilst the aforementioned works deal with full state observability,action model learning has also been studied in domains where there is partial or missing state observability.
-{\sf ARMS} works when no partial intermediate state is given.
-It defines a set of weighted constraints that must hold for the plans to be correct, and solves the weighted propositional satisfiability problem with a MAX-SAT solver~\cite{yang2007learning}.
-In order to efficiently solve the large MAX-SAT representations, {\sf ARMS} implements a hill-climbing method that models the actions approximately.
-%and so it may output an inconsistent model
-{\sc SLAF} also deals with partial observability~\cite{amir:alearning:JAIR08}.
-Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, it builds a complete explanation of observations by models of actions through a CNF formula.
-The learning algorithm updates the formula of the belief state with every action and observation in the sequence.
-This update makes sure that the new formula represents all the transition relations consistent with the actions and observations.
-The formula returned at the end includes all consistent models, which can then be retrieved with additional processing.
+%More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
 
-Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
+%Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
 An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
+
+\LOCM accepts input traces of the form  $\tau = \langle a_1, a_1, a_2, \ldots, a_n\rangle$ ..
 

### 2018-05-23 src/compiler_new.py
small comment on last planning horizon for state observability above 0

@@ -374,8 +374,13 @@ fdomain.close()
 
 ### Solvie the learning task
 starting_horizon = str(2*TOTAL_STEPS + 2)
+if state_observability>0.1:
+    ending_horizon = " -T " + starting_horizon
+else:
+    ending_horizon = ""
 
-cmd = "rm " + config.OUTPUT_FILENAME + " planner_out.log;" + config.PLANNER_PATH + "/" + config.PLANNER_NAME + " learning_domain.pddl learning_problem.pddl -F " + starting_horizon + " " + config.PLANNER_PARAMS + " > planner_out.log"
+
+cmd = "rm " + config.OUTPUT_FILENAME + " planner_out.log;" + config.PLANNER_PATH + "/" + config.PLANNER_NAME + " learning_domain.pddl learning_problem.pddl -F " + starting_horizon + " " +ending_horizon + " " + config.PLANNER_PARAMS + " > planner_out.log"
 print("\n\nExecuting... " + cmd)
 os.system(cmd)
 

### 2018-05-23 src/compiler_new.py
small comment on last planning horizon for action observability equal 1

@@ -374,7 +374,7 @@ fdomain.close()
 
 ### Solvie the learning task
 starting_horizon = str(2*TOTAL_STEPS + 2)
-if state_observability>0.1:
+if state_observability>0.1 or action_observability==1:
     ending_horizon = " -T " + starting_horizon
 else:
     ending_horizon = ""

### 2018-05-23 src/model-evaluator.py
model-evaluator.py: compatibility with aij18 compiler. Should still be compatible with previous versions

@@ -26,7 +26,7 @@ def valid_action_combination(comb):
     return all(actions_fit)
 
 
-def evaluate_matching(matchings, fd_eva_task, fd_ref_task):
+def evaluate_matching(matchings, eva_actions, ref_actions):
     ref_pres = set()
     eva_pres = set()
     ref_adds = set()
@@ -40,7 +40,7 @@ def evaluate_matching(matchings, fd_eva_task, fd_ref_task):
         param_reform = match[2:]
         # Build the pre/add/del sets
         # Each element of the set is a tuple (action name, literal)
-        for action in fd_ref_task.actions:
+        for action in ref_actions:
             if action.name == action_evaluated:
                 # Preconditions
                 if isinstance(action.precondition, pddl.conditions.Atom):
@@ -55,7 +55,7 @@ def evaluate_matching(matchings, fd_eva_task, fd_ref_task):
                         ref_adds.add((action_evaluated, effect.literal.key))
                 break
 
-        for action in fd_eva_task.actions:
+        for action in eva_actions:
             if action.name == matched_action:
                 action_args = [arg.name for arg in action.parameters]
                 # Preconditions
@@ -105,10 +105,9 @@ try:
 
    reference_domain_filename  = cmdargs[0]
    evaluation_domain_filename  = cmdargs[1]
-   aux_problem_filename  = cmdargs[2]
 except:
    print "Usage:"
-   print sys.argv[0] + " [-r] [-p <partial domain>] <reference domain> <evaluation domain>  <aux problem>"
+   print sys.argv[0] + " [-r] [-p <partial domain>] <reference domain> <evaluation domain>"
    sys.exit(-1)
 
 
@@ -117,28 +116,31 @@ except:
 # aux_problem_filename = "../benchmarks/handpicked/blocks/test-1.pddl"
 
 # Creating a FD task with the ref domain and the aux problem file
-fd_ref_domain = pddl_parser.pddl_file.parse_pddl_file("domain", reference_domain_filename)
-fd_problem = pddl_parser.pddl_file.parse_pddl_file("task", aux_problem_filename)
-fd_ref_task = pddl_parser.pddl_file.parsing_functions.parse_task(fd_ref_domain, fd_problem)
+ref_domain_pddl = pddl_parser.pddl_file.parse_pddl_file("domain", reference_domain_filename)
[…]

### 2018-05-23 doc/aaai-19/model-recognition.tex
added more notes on model recognition

@@
 
+\subsection{Getting rational}
+The {\em \strips\ probability distribution} assigns the same probabaility value to any hypothesis model $\mathcal{M}\in M$ that can reproduce the given observation $\mathcal{T}$.
+If the input plan execution $\mathcal{T}$ is partially observed, then it is easier that many different models can reproduce the same observations because the input observation is less constrained.
+A particular case of interest is when $\mathcal{T}=\tup{s_0, s_n}$ that is, when the given observation is an {\em input-output} example.
+
+If we assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}, one can go one step further and define a probability distribution that considers how an hypothesis $\mathcal{M}\in M$ reproduces an observation $\mathcal{T}$.
+Defining as more probable hypothesis that can reproduce observations in less steps for instance, using the expression: $\frac{1}{1+e^{-cost(\pi^*_{\mathcal{M},\mathcal{T}})}}$ where $\pi^*_{\mathcal{M},\mathcal{T}}$ is an optimal plan (according to $\mathcal{M}$) that reproduces $\mathcal{T}$.
+
 \section{Model Recognition as Planning}
@@
 This means that some fluents in $s_i$ may be missing because its value is unknown.
-In the extreme, entire states $s_i$ can be missing.
+In the extreme, entire states $s_i$, {\small $1\leq i\leq n$}, can be missing.
 \item $\mathcal{T}$ is {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 

### 2018-05-23 doc/aaai-19/model-recognition.tex
notes added on model recognition

--- 2018-05-16 doc/aaai-19/model-recognition.tex
@@
 
+The {\em \strips\ probability distribution} assigns the same probabaility value to any hypothesis model $\mathcal{M}\in M$ that can reproduce the given observation $\mathcal{T}$.
+As a rule of thumb, the less information observed the more different models can reproduce observations.
+If we assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}, one can define probability distributions that consider as more probable hypothesis the ones that reproduce observations in less steps.
+This an interesting research direction, specially for the particular case $\mathcal{T}=\tup{s_0, s_n}$ where observations are {\em input-output} examples.
+
 \bibliographystyle{aaai}

### 2018-05-24 doc/programming-project/AI-programming.tex
small udpates to project proposal

@@
 \section{Historial científico-técnico del Equipo de Investigadores (Apartado 5.1,bases de la convocatoria}
+
+The project will be developed within the {\em Group of Reasoning on Planning and Scheduling} (GRPS), {\tt http://users.dsic.upv.es/grupos/grps/}, that is part of the {\em Department of Computer Systems and Computation} of the UPV.
+In the last 10 years, the GRPS has led 3 research projects of the Spanish national research plan on {\em AI planning}.
+Further, the GRPS has developed its own AI planner, that is participating  at the International Planning Competition (IPC-2018).
+
+The research group for this project comprises: 1 principal investigator, 1 senior researcher and 1 PhD student. All of them are GRPS members.
 
@@
 Closely related to the aims of the project is the work on the synthesis and validation of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
-The existing algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}.
+The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}.
 To keep the computation of FSCs tractable, they limit the space of possible solutions bounding the maximum size of the FSC.
-In addition, they impose that the instances to solve share, not only the domain theory (actions and predicates schemes) but the set of fluents~\cite{sergio:aprograming:icaps16} or a subset of {\it observable} fluents~\cite{Geffner:FSM:AAAI10}.
-
-The computation of FSCs for generalized planning includes works that compile the generalized planning task into another forms of problem solving so they benefit from the last advances on off-the-shelf solvers (e.g.
+The computation of FSCs includes works that compile this task into another forms of problem solving so they benefit from the last advances on off-the-shelf solvers (e.g.
 {\em classical planning}~\cite{sergio:aprograming:icaps16}, {\em conformant planning}~\cite{Geffner:FSM:AAAI10}, {\em CSP}~\cite{Infantes:FSC:ECAI2010} or a {\em Prolog program}~\cite{Giacomo:FSM:ICAPS13}).
@@
 \label{subsec:metodologia}
-Our research shows that current AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16}.
+Our current research shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16}.
 To illustrate this, Table~\ref{tab:programs} reports the invested time to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list or traversing a binary tree.
@@
     \item Design of the user-interaction for the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases and the available instruction set. 
-    \item Experimental design. The experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve a given set of TDD programming tasks:
+    \item Experimental design.
+The experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve a given set of TDD programming tasks that are characterized as {\em finite state machines}.
       \item Evaluation of different AI planners available, with special attention to the planners that get the best results at the IPC-2018. 
@@
 \item {\bf T3. Experiments and dissemination of results (months 13-24}).
+   \begin{small}
+      \begin{enumerate}
+      \item Reporting the experimental performance of our AIP approach for solving TDD programming tasks.
+      \item Disseminating the obtained theoretical and empirical results by submitting research papers to the main conferences and journals in AI.        
+      \end{enumerate}
+\end{small}        
 {\small {\bf Deliverable:} Final report with the obtained conclusions and produced publications.}
@@
 \label{subsec:objectivos}
-To precisely characterize the specific kind and size of the programs that we can synthesize and validate, we model programs as {\em finite state machines}~\cite{sergio:aprograming:ijcai16}.
+
+In this project we aim to {\bf study the performance of an Artificial Intelligence approach for the automated synthesize and validation of programs up to}: \underline{5 program lines}, \underline{50 variables with binary domain} and, an \\\underline{instruction set that comprises 10 instructions}.
+The performance of our approach will be evaluated with regard to the {\em computation time} and {\em memory} invested in the synthesize and validation of the characterized programs.
+
+To precisely characterize the specific kind and size of the programs that we can synthesize and validate, we plan to model programs as {\em finite state machines}~\cite{sergio:aprograming:ijcai16}.
 Figure~\ref{fig:list} shows a program for traversing linked lists pictured as a {\em finite state machine}.
@@
 
-In this project we aim to {\bf study the performance of an Artificial Intelligence approach for the automated synthesize and validation of programs up to}: \underline{5 program lines}, \underline{50 variables with binary domain} and, an \\\underline{instruction set that comprises 10 instructions}.
-The performance of our approach will be evaluated with regard to the {\em computation time} and {\em memory} invested in the synthesize and validation of the characterized programs.
-
 \section{Beneficios del proyecto. Difusión y Explotación en su caso de los Resultados.}
@@
 Research in AI algorithms is too often tested with laboratory problems and AIP is not an exception.
-Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition (IPC)~\cite{vallati:IPC:AI15}.
+Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition~\cite{vallati:IPC:AI15}.
 This project will also help to {\bf meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks}.

### 2018-05-24 src/compiler_new.py
experimenter_new.py: script to learn and evaluate models

@@ -381,7 +381,7 @@ else:
 
 
 cmd = "rm " + config.OUTPUT_FILENAME + " planner_out.log;" + config.PLANNER_PATH + "/" + config.PLANNER_NAME + " learning_domain.pddl learning_problem.pddl -F " + starting_horizon + " " +ending_horizon + " " + config.PLANNER_PARAMS + " > planner_out.log"
-print("\n\nExecuting... " + cmd)
+# print("\n\nExecuting... " + cmd)
 os.system(cmd)
 
 

### 2018-05-24 src/experimenter_new.py

@@ -1,0 +1,46 @@
+#! /usr/bin/env python
+import glob, os, sys, copy, itertools, math
+import pddl, pddl_parser
+import fdtask_to_pddl
+import numpy as np
+import model_evaluator
+import config
+
+# **************************************#
+# MAIN
+# **************************************#
+
+try:
+    if "-s" in sys.argv:
+        check_static_predicates = True
+        sys.argv.remove("-s")
+    else:
+        check_static_predicates = False
+
+    domain_folder_name  = sys.argv[1]
+    action_observability = sys.argv[2]
+    state_observability = sys.argv[3]
+
+except:
+    print "Usage:"
+    print sys.argv[0] + "[-s] <domain folder> <action observability (0-1)> <state observability (0-1)>"
+    sys.exit(-1)
+
+
+# outdir = "results/" + s + "-"+ d + "/"
+# cmd = "mkdir " + outdir
+# print("\n\nExecuting... " + cmd)
+# os.system(cmd)
+
+cmd = config.PROJECT_PATH + "src/compiler_new.py " + domain_folder_name + " " + action_observability + " " + state_observability
+# print("\n\nExecuting... " + cmd)
+os.system(cmd)
+
+domain_name, best_evaluation, best_matches = model_evaluator.evaluate("learned_domain.pddl", domain_folder_name+"ref_domain", True)
+print(" & ".join(
+    [domain_name] + [str(round(e, 2)) for e in best_evaluation]) + " \\\\" + " % {}".format(
+    best_matches))
+
+
+
+sys.exit(0)

### 2018-05-24 src/model-evaluator.py

@@ -1,241 +1,0 @@
-#! /usr/bin/env python
-# from __future__ import division
-import  sys
-import pddl, pddl_parser
-import numpy as np
-import itertools
-
-def reform_literal(literal, action_args, reformulation):
-    literal_name = literal.predicate
-    params = [action_args[reformulation[action_args.index(arg)]-1] for arg in literal.args]
-
-    return tuple([literal_name] + [tuple(params)])
-
-def valid_parameter_combination(param_comb, action):
-    action_params = action_params_dict[action]
-    for i in range(len(param_comb)):
-        if action_params[i] != action_params[param_comb[i]-1]:
-            return False
-    return True
-
-def valid_action_combination(comb):
-    matched_actions = [x[1][0] for x in comb]
-    if any(matched_actions.count(x) > 1 for x in matched_actions):
-        return False
-    actions_fit = [(set(action_params_dict[p[0][0]]) == set(action_params_dict[p[1][0]])) and (len(action_params_dict[p[0][0]]) == len(action_params_dict[p[1][0]])) for p in comb]
-    return all(actions_fit)
-
-
-def evaluate_matching(matchings, eva_actions, ref_actions):
-    ref_pres = set()
-    eva_pres = set()
-    ref_adds = set()
-    eva_adds = set()
-    ref_dels = set()
-    eva_dels = set()
-
-    for match in matchings:
-        action_evaluated = match[0]
-        matched_action = match[1]
-        param_reform = match[2:]
-        # Build the pre/add/del sets
-        # Each element of the set is a tuple (action name, literal)
-        for action in ref_actions:
-            if action.name == action_evaluated:
-                # Preconditions
-                if isinstance(action.precondition, pddl.conditions.Atom):
-                    ref_pres.add((action_evaluated, action.precondition.key))
-                else:
-                    ref_pres.update([(action_evaluated, x.key) for x in action.precondition.parts])
-                # Effects
-                for effect in action.effects:
-                    if effect.literal.negated:
-                        ref_dels.add((action_evaluated, effect.literal.key))
-                    else:
-                        ref_adds.add((action_evaluated, effect.literal.key))
-                break
-
-        for action in eva_actions:
-            if action.name == matched_action:
-                action_args = [arg.name for arg in action.parameters]
-                # Preconditions
[…]

### 2018-05-24 src/model_evaluator.py

@@ -1,0 +1,244 @@
+#! /usr/bin/env python
+# from __future__ import division
+import  sys
+import pddl, pddl_parser
+import numpy as np
+import itertools
+
+def reform_literal(literal, action_args, reformulation):
+    literal_name = literal.predicate
+    params = [action_args[reformulation[action_args.index(arg)]-1] for arg in literal.args]
+
+    return tuple([literal_name] + [tuple(params)])
+
+def valid_parameter_combination(param_comb, action, action_params_dict):
+    action_params = action_params_dict[action]
+    for i in range(len(param_comb)):
+        if action_params[i] != action_params[param_comb[i]-1]:
+            return False
+    return True
+
+def valid_action_combination(comb, action_params_dict):
+    matched_actions = [x[1][0] for x in comb]
+    if any(matched_actions.count(x) > 1 for x in matched_actions):
+        return False
+    actions_fit = [(set(action_params_dict[p[0][0]]) == set(action_params_dict[p[1][0]])) and (len(action_params_dict[p[0][0]]) == len(action_params_dict[p[1][0]])) for p in comb]
+    return all(actions_fit)
+
+
+def evaluate_matching(matchings, eva_actions, ref_actions):
+    ref_pres = set()
+    eva_pres = set()
+    ref_adds = set()
+    eva_adds = set()
+    ref_dels = set()
+    eva_dels = set()
+
+    for match in matchings:
+        action_evaluated = match[0]
+        matched_action = match[1]
+        param_reform = match[2:]
+        # Build the pre/add/del sets
+        # Each element of the set is a tuple (action name, literal)
+        for action in ref_actions:
+            if action.name == action_evaluated:
+                # Preconditions
+                if isinstance(action.precondition, pddl.conditions.Atom):
+                    ref_pres.add((action_evaluated, action.precondition.key))
+                else:
+                    ref_pres.update([(action_evaluated, x.key) for x in action.precondition.parts])
+                # Effects
+                for effect in action.effects:
+                    if effect.literal.negated:
+                        ref_dels.add((action_evaluated, effect.literal.key))
+                    else:
+                        ref_adds.add((action_evaluated, effect.literal.key))
+                break
+
+        for action in eva_actions:
+            if action.name == matched_action:
+                action_args = [arg.name for arg in action.parameters]
+                # Preconditions
[…]

### 2018-05-24 doc/aaai-19/model-recognition.tex

@@
 Given a set of possible action models and a partially observed plan execution, {\em model recognition} is the task of identifying which of these action models has the highest probability of producing the given plan execution.
-{\em Model recognition} is relevant because it enables identifying algorithms by their execution and because, once an action model is recognized, the planning model-based machinery becomes applicable.
+{\em Model recognition} enables identifying algorithms by observing their execution besides if an action model is successfully recognized, the planning model-based machinery becomes applicable.
 This paper formalizes the {\em model recognition} task and proposes a method to assess the probability of a given \strips\ action model to produce a partially observed plan execution.
@@
 The effectiveness of {\em model recognition as planning} is shown with sets of \strips\ models that represent different {\em Turing Machines}.
-We show that {\em model recognition as planning} effectively identifies the executed {\em Turing Machine} despite intermediate machine state, applied transitions or tape values, are unobserved.
+We show that {\em model recognition as planning} effectively identifies the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or several tape values, are unobserved.
 \end{abstract}
@@
 
-{\em Plan recognition as planning} recently defined the model-based approach for plan/goal recognition~\cite{ramirez2012plan,ramirez2009plan}.
+{\em Plan recognition as planning} defined the model-based approach for plan/goal recognition~\cite{ramirez2012plan,ramirez2009plan}.
 This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal according to the observed plan execution.
@@
 \begin{itemize}
-\item Once the action model is recognized the planning model-based machinery~\cite{ghallab2004automated,geffner:book:2013} becomes applicable.
+\item Once the action model is recognized, then the planning model-based machinery~\cite{ghallab2004automated,geffner:book:2013} becomes applicable.
 \item It enables identifying algorithms by observing their execution.
-Many computational models are compilable into classical planning~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2016hierarchical}.
+Many computational models (like {\sc GOLOG} programs or finite state controllers) are compilable into classical planning~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2016hierarchical}.
 \end{itemize}
@@
 This method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
-
 Last but not least, the paper evaluates the effectiveness of {\em model recognition as planning} with sets of \strips\ models that represent different {\em Turing Machines}, all of them defined within the same tape alphabet and same machine states.
-We show that {\em model recognition as planning} effectively identifies the executed {\em Turing Machine} despite intermediate machine state, applied transitions or tape values, are unobserved.
+We show that {\em model recognition as planning} effectively identifies the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or several tape values, are unobserved.
 
@@
 \label{sec:background}
-This section defines the classical planning model and the \strips\ action model that we use for formalizing the {\em model recognition as planning} method.
+This section introduces the formal models for classical planning and for \strips\ actions.
 
@@
 \subsection{The \strips\ edit distance}
-Our approach for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by the model $\mathcal{M}$ to produce $\mathcal{T}$.
+Our approach for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by $\mathcal{M}$ to produce $\mathcal{T}$.
 
@@
 The $\delta(\mathcal{M},\mathcal{T})$ distance could also be defined assessing the edition required by the observed plan execution to match the given model.
-This implies defining {\em edit operations} that modify $\mathcal{T}$ instead of modifying $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
-Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$ (the number of {\em variable objects} is normally smaller than the number of objects in the observed states).
+This implies defining {\em edit operations} that modify $\mathcal{T}$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
+Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$.
+The reason is that in practice, the number of {\em variable objects} is normally smaller than the number of objects in the observed states.
 
 \subsection{The \strips\ probability distribution}
-According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$ provided the observation $\mathcal{O}$ is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
+According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$, provided the observation $\mathcal{O}$, is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
 In our scenario, the hypotheses are about the set of possible \strips\ action models $\mathcal{M}\in M$ while the given observation is a partially observed plan execution $\mathcal{T}$.
 
-We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\mathcal{T})$, to the probability distribution of the comparable \strips\ models (within the $F_v(\xi)$ sets) provided that the plan execution $\mathcal{T}$ is observed.
-This probability distribution can be estimated in three-steps by:
+We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\mathcal{T})$, to the probability distribution of the comparable \strips\ models (the definable models within the $F_v(\xi)$ sets) provided that the plan execution $\mathcal{T}$ is observed.
+This probability distribution can be estimated in three-steps:
 \begin{enumerate}
-\item Estimating the a priori probabilities $P(\mathcal{T})$ and $P(\mathcal{H})$.
+\item Estimating the {\em a priori} probabilities $P(\mathcal{T})$ and $P(\mathcal{H})$.
 For instance, given the set of predicates $\Psi$ and the given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$.
 If we assume that a priori all models are equiprobable this means that $P(\mathcal{M})=\frac{1}{\prod_\xi 2^{2|F_v(\xi)|}}$.
-With respect to the observations, given $\Psi$ and a set of objects $\Omega$, the number of possible state observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|F|}$ while the number of possible action observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|A|}$.
-If we assume that a priori all traces are equiprobable, this means $P(\mathcal{T})=\frac{1}{2^{n\times|F|}\times 2^{n\times|A|}}$.
+With respect to the observations, given the sets $F$ and $A$, the number of possible state observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|F|}$ while the number of possible action observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|A|}$.
+If we assume that a priori all traces are equiprobable, this means $P(\mathcal{T})=\frac{1}{2^{2n\times|F|\times |A|}}$.
 \item Estimating the conditional probability $P(\mathcal{T}|\mathcal{M})$.
-For instance, we can compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}\in M$ and map  this distance into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
+We can compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}\in M$ and map  this distance into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
 \item Applying the Bayes rule to obtain the normalized posterior probabilities since these probabilities must sum 1.
@@
 \label{sec:asPlanning}
-Now we show a classical planning compilation for computing the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for each $\mathcal{M}\in M$. The compilation follows these assumptions
+For each $\mathcal{M}\in M$, we compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$  using a classical planning compilation.
+The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
 \begin{enumerate}
-\item The initial state $s_0\in\mathcal{T}$ is {\em fully observable}.
-\item Intermediate actions $a_i\in\mathcal{T}$ and states $s_i\in\mathcal{T}$ s.t.
-{\small $1\leq i\leq n$}, can be {\em partially observed}.
-This means that some fluents in $s_i$ may be missing because its value is unknown.
-In the extreme, entire states $s_i$, {\small $1\leq i\leq n$}, can be missing.
-\item $\mathcal{T}$ is {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
+\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
+A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
+\item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$.
 \end{enumerate}
+
+Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+The edited action model is validated at a four action plan for inverting a two-block tower where states $s_0$ and $s_4$ are observed while the intermediate states $s_1$, $s_2$ and $s_3$ are unobserved.
+\begin{figure}
+{\tt\scriptsize
+00 : (insert\_add\_handempty\_stack)\\
+01 : (insert\_add\_clear\_stack\_var1)\\
+02 : (apply\_unstack blockB blockA i1 i2)\\
+03 : (apply\_putdown blockB i2 i3)\\
+04 : (apply\_pickup blockA i3 i4)\\
+05 : (apply\_stack blockA blockB i4 i5)\\
+06 : (validate\_1)
+}
+ \caption{\small Plan for editing (steps [0-1]) and validating (steps [2-6]) a given {\em blockswold} model.}
+\label{fig:plan-pdistance}
+\end{figure}
+
+Our interest is not in the resulting action model from the compilation $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deleitions) for that $\mathcal{M}'$ is validated in the given observations, e.g.
+$\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}.
+In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$.
+%The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
 
@@
 
-\subsection{The compilation}
-The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
-\begin{enumerate}
-\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$.
-\end{enumerate}
+\subsection{The compilation formalization}
 
-To illustrate this, Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-The edited action model is validated at a four action plan for inverting a two-block tower.
-Our interest is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deleitions) for that $\mathcal{M}'$ is validated in the given observations, e.g.
-$\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}.
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$.
-The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
-
-\begin{figure}[hbt!]
-{\tt\scriptsize
-00 : (insert\_add\_handempty\_stack)\\
-01 : (insert\_add\_clear\_stack\_var1)\\
-02 : (apply\_unstack blockB blockA i1 i2)\\
-03 : (apply\_putdown blockB i2 i3)\\
-04 : (apply\_pickup blockA i3 i4)\\
-05 : (apply\_stack blockA blockB i4 i5)\\
-06 : (validate\_1)
-}
- \caption{\small Plan for editing and validating a given {\em blockswold} model.}
-\label{fig:plan-pdistance}
-\end{figure}
-
-Given $\tup{\mathcal{M},\Psi,\mathcal{T}}$ the compilation outputs a classical planning task $P=\tup{F,A,I,G}$:
+Given $\tup{\mathcal{M},\Psi,\mathcal{T}}$ s.t.,
+\begin{itemize}
+\item The initial state $s_0\in\mathcal{T}$ is {\em fully observable}.
+\item Intermediate actions $a_i\in\mathcal{T}$ and states $s_i\in\mathcal{T}$ s.t.
+{\small $1\leq i\leq n$}, can be {\em partially observed}.
+This means that some fluents in $s_i$ may be missing because its value is unknown.
+In the extreme, entire states $s_i$, {\small $1\leq i\leq n$}, can be missing.
+\item $\mathcal{T}$ is {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
+\end{itemize}
+the compilation outputs a classical planning task $P=\tup{F,A,I,G}$:
 \begin{itemize}
@@
 {\small\tt (next ?x1 ?x2)} encoding that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
-{\small\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\mathcal{T}$.
+{\small\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
 {\small\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
 Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
-\item Actions $A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ action schema is defined:
+\item Actions $A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ schema is defined:
 \begin{itemize}
 \item The {\bf header} is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$. Parameters $?xl$, $?x$ and $?xr$ are tape cells.
-\item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to make $?x$ the tape cell pointed by the header and $?xl$ and $?xr$ its left and right neighbours.
+\item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to make $?x$ the tape cell pointed by the header and $?xl/?xr$ its left/right neighbours.
 Preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} are also included to capture the symbol pointed by the header and the currrent machine state.
-\item The {\bf delete effects} remove the symbol pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state.
+\item The {\bf delete effects} remove the symbol pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state according to $\delta$.
 \end{itemize}
@@
 
-The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} (the full encoding of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema).
+The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} (the full encoding of the {\em Turing Machine} of Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema).
 \begin{figure}[hbt!]
@@
 
-Since the {\em transition function} of a {\em Turing Machine} can be encoded as \strips\ action schemas, executions of that machine are definable as {\em plan traces} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
-In this case $s_0$ encodes the initial state of the tape plus, the position of the header and the initial machine state.
+Since the {\em transition function} of a {\em Turing Machine} can be encoded as a classical planning frame $\Phi=\tup{F,A}$, executions of a {\em Turing Machine} are definable as {\em plan traces} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
 
@@
 We randomly choose a machine $\mathcal{M}\in M$ and produce an fifty-step execution plan trace $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_{50},s_{50}}$.
-The experiment is following our {\em model recognition as planning} method to identify the {\em Turing Machine} that produced $\mathcal{T}$.
-The exeperiment is repeated for different amounts of missing information in the input trace $\mathcal{T}$.
+Finally, we follow our {\em model recognition as planning} method to identify the {\em Turing Machine} that produced $\mathcal{T}$.
+This exeperiment is repeated for different amounts of missing information in the input trace $\mathcal{T}$.
 
@@
 If we assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}, one can define probability distributions that consider as more probable hypothesis the ones that reproduce observations in less steps.
-This an interesting research direction, specially for the particular case $\mathcal{T}=\tup{s_0, s_n}$ where observations are {\em input-output} examples.
+%This an interesting research direction, specially for the particular case $\mathcal{T}=\tup{s_0, s_n}$ where observations are {\em input-output} examples.
 

### 2018-05-24 doc/programming-project/AI-programming.tex

@@
 
-This research project aims to investigate a {\bf novel approach for software development, that leverages Artificial Intelligence, to reduce the chances of introducing programming errors}.
-The project proposes to addresses automatic program synthesis and validation starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine.
+This research project aims to investigate a novel approach for software development, that leverages Artificial Intelligence, to reduce the chances of introducing programming errors.
+The project proposes to address {\bf automatic program synthesis and validation starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine}.
 Our work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI}, the main international conference on Artificial Intelligence.
@@
 
-\section{Historial científico-técnico del Equipo de Investigadores (Apartado 5.1,bases de la convocatoria}
+\section{Historial científico-técnico del Equipo de Investigadores (Apartado 5.1,bases de la convocatoria)}
 
-The project will be developed within the {\em Group of Reasoning on Planning and Scheduling} (GRPS), {\tt http://users.dsic.upv.es/grupos/grps/}, that is part of the {\em Department of Computer Systems and Computation} of the UPV.
+The project will be developed within the {\em Group of Reasoning on Planning and Scheduling} (GRPS){\footnote{\tt http://users.dsic.upv.es/grupos/grps/}}, that is part of the {\em Department of Computer Systems and Computation} of the UPV.
 In the last 10 years, the GRPS has led 3 research projects of the Spanish national research plan on {\em AI planning}.
@@
 
-\subsection{Researcher in trainnig period: Diego Aineto}
+\subsection{Researcher in training period: Diego Aineto}
 
@@
 
-This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} for the automatic synthesis and validation of programs}.
+This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} paradigm for the automatic synthesis and validation of programs}.
 Remarkably, the programs synthesized with our approach are guaranteed to be bug-free over given sets of {\em input-output} tests cases.
@@
 {\em classical planning}~\cite{sergio:aprograming:icaps16}, {\em conformant planning}~\cite{Geffner:FSM:AAAI10}, {\em CSP}~\cite{Infantes:FSC:ECAI2010} or a {\em Prolog program}~\cite{Giacomo:FSM:ICAPS13}).
-This last case requires a behavior specification of the FSC consisting on classified execution histories that (1) accept all legal execution histories leading to a goal-satisfying state, and (2) reject those that contain repeated configurations (indicating an infinite loop) and that cannot be extended (indicating a dead end)~\cite{Giacomo:FSM:ICAPS13}.
 
-Finally the generation of programs from examples is a research question also addressed in inductive Logic Programming (ILP).
-ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn a given target concept, expressed as a  logic program,  from  examples  and  background  knowledge, that are expressed as logic facts~\cite{muggleton1991inductive}.
+Finally the generation of programs from examples is a research question also addressed in inductive Logic Programming (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
+ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn a given target concept, expressed as a  logic program,  from  examples  and  background  knowledge, that are expressed as logic facts.
+
+(Please find the full {\em bibliography} at the end of the document)
 
@@
 \label{subsec:metodologia}
-Our current research shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16}.
-To illustrate this, Table~\ref{tab:programs} reports the invested time to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list or traversing a binary tree.
+Our current research already shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16}.
+Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list or traversing a binary tree.
 
@@
 \end{tt}  
-\caption{\small Time to synthesize the programs with the {\sc FD} AI planner~\cite{helmert2006fast} on a processor {\em Intel Core i5 3.10GHz x 4} and with a 4GB memory bound.}
+\caption{\small Time to synthesize the programs with the AI planner {\sc FD}~\cite{helmert2006fast} on a processor {\em Intel Core i5 3.10GHz x 4} and with a 4GB memory bound.}
 \label{tab:programs}
@@
 
-Our approach is modeling and solving a given TDD programming task as an AI planning task.
-Briefly, the state variables of these AIP tasks are {\em fluents} of the kind {\tt (var:=value)} while the initial/goal states of the AIP tasks encode the programming input/output tests.
+Given a TDD programming task our approach is modeling and solving it as it were an AI planning task.
+Briefly, the state variables of this AIP task are {\em fluents} of the kind {\tt (var:=value)} while the initial/goal states of the AIP task encode the programming input/output tests of the TDD programming task.
 Finally, program instructions are encoded using two kinds of AIP actions:
@@
 \subsection{Two-year workplan}
-We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that (1), takes as input a set of test cases that specify the TDD programming task to solve and (2), outputs a program code that passes these tests with a bug-free guarantee.
+We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that (1), takes as input a set of test cases that specify the TDD programming task to solve and (2), outputs a program source code that passes these tests with a bug-free guarantee.
 Figure~\ref{fig:gantt} details the proposed 24-month timeline for the project.
@@
 \ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Development (6 months)}{4}{6}\\
-\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Experiments (12 months)}{7}{12}
+\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Dissemination (12 months)}{7}{12}
 \end{ganttchart}
-\caption{\small Twentyfour-month workplan for the development of a user-interactive program synthesizer.}
+\caption{\small Workplan for the development of a user-interactive program synthesizer.}
 \label{fig:gantt}
@@
     \begin{enumerate}
-    \item Design of the user-interaction for the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases and the available instruction set. 
+    \item Design of the user-interaction for the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases plus the available instruction set. 
     \item Experimental design.
-The experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve a given set of TDD programming tasks that are characterized as {\em finite state machines}.
+Experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve TDD programming tasks that are characterized as {\em finite state machines}.
       \item Evaluation of different AI planners available, with special attention to the planners that get the best results at the IPC-2018. 
@@
       \item Reporting the experimental performance of our AIP approach for solving TDD programming tasks.
-      \item Disseminating the obtained theoretical and empirical results by submitting research papers to the main conferences and journals in AI.        
+      \item Dissemination of the obtained theoretical and empirical results by submitting research papers to top international conferences and journals in AI.        
       \end{enumerate}
@@
 
-In this project we aim to {\bf study the performance of an Artificial Intelligence approach for the automated synthesize and validation of programs up to}: \underline{5 program lines}, \underline{50 variables with binary domain} and, an \\\underline{instruction set that comprises 10 instructions}.
-The performance of our approach will be evaluated with regard to the {\em computation time} and {\em memory} invested in the synthesize and validation of the characterized programs.
+The specific objective of this project is the study of the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions:
+\begin{enumerate}
+\item Maximum number of program lines.
+\item Maximum number of program variables. 
+\item Maximum size of the available instruction set.  
+\end{enumerate}
 
-To precisely characterize the specific kind and size of the programs that we can synthesize and validate, we plan to model programs as {\em finite state machines}~\cite{sergio:aprograming:ijcai16}.
+In more detail, this project aims to {\bf study the performance of an Artificial Intelligence approach for the automated synthesize and validation of programs up to}: \underline{5 program lines}, \underline{50 variables with binary domain} and, an \\\underline{instruction set that comprises 10 instructions}.
+The performance of our approach will be evaluated with regard to the {\em computation time} and {\em memory} invested in the synthesize and validation of the programs.
+
+Note that despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}.
+With this regard, our AIP encoding already supports callable procedures to decompose a given programming task in simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.
+
+We precisely characterize the specific kind and size of the programs that we can synthesize and validate by modeling programs as {\em finite state machines}~\cite{sergio:aprograming:ijcai16}.
 Figure~\ref{fig:list} shows a program for traversing linked lists pictured as a {\em finite state machine}.
@@
 \item $T:Q\times\{0,1\}\rightarrow Q$ is a {\em transition function} mapping pairs of a program line and the result of a {\em condition} check into the next program line.
-\item $\Gamma:Q\times\{0,1\}\rightarrow \mathcal{I}$ is a {\em transition function} mapping pairs of a program line and the {\em condition} result into the corresponding instruction.
+Likewise $\Gamma:Q\times\{0,1\}\rightarrow \mathcal{I}$ is a {\em transition function} mapping pairs of a program line and the {\em condition} result into the corresponding instruction.
 \end{itemize}
@@
 \begin{enumerate}
-\item First we observe whether the condition associated to the current progam line holds (we denote the result of this observation as $O_q(s)\in \{0,1\}$). 
+\item First we observe whether the condition associated to the current program line holds (we denote the result of this observation as $O_q(s)\in \{0,1\}$). 
 \item Then we compute the new program line $q'=T(q,O_q(s))$ and the instruction $a=\Gamma(q,O_q(s))$ to apply next.
@@
 The no-op action does not modify the program variables and affects only to the next program line which allows to concatenate conjunctions of conditions like in {\em decision trees}.
-\item Finally the value of the program variables $s'=\theta_a(s)$ is updated executing the corresponding instrucntion $a=\Gamma(q,O(s,q))\in \mathcal{I}$.
+\item Finally the value of the program variables $s'=\theta_a(s)$ is updated executing the corresponding instruction $a=\Gamma(q,O(s,q))\in \mathcal{I}$.
 \end{enumerate}
-
-The specific objective of this project is the study of the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions:
-\begin{enumerate}
-\item $|\mathcal{Q}|$, the maximum number of program lines.
-\item $|\mathcal{V}|$, the maximum number of program variables. 
-\item $|\mathcal{I}|$, the maximum size of the available instruction set.  
-\end{enumerate}
-Note that despite setting bounds for $|\mathcal{Q}|$, $|\mathcal{V}|$ and $|\mathcal{I}|$, challenging programming tasks can still be addressed using {\em problem decomposition}.
-With this regard, our AIP encoding supports callable procedures to decompose a given programming task in simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.
 
@@
 \label{subsec:beneficios}
-This project will {\bf provide new insights into the current understanding of how Artificial Intelligence can assist programmers in the software development}.
-
+This project will provide new insights into the current understanding of how Artificial Intelligence can assist programmers in the software development.
 Research in AI algorithms is too often tested with laboratory problems and AIP is not an exception.
 Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition~\cite{vallati:IPC:AI15}.
-This project will also help to {\bf meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks}.
-
+This project will also help to meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks.
 Promising research opportunities come from the application of AIP to {\em program synthesis} given that, {\em program synthesis} with a tests base, can also be seen as a dual to {\em program testing}.
 AIP has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}.
+
+The expected benefits of this project are:
+\begin{enumerate}
+\item The definition of {\bf effective metrics for assessing how well a given program covers a set of {\em input-output} test cases}.
+\item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}.
+\item The development, evaluation and dissemination of a finished {\bf open source software for the {\em Synthesis and Validation of Programs}}.
+\end{enumerate}
 
@@
 
-\begin{footnotesize}
+\begin{tiny}
 \bibliography{AI-programming}
-\end{footnotesize}
+\end{tiny}
 \bibliographystyle{ieeetr}

### 2018-05-25 doc/programming-project/AI-programming.tex
updates to research project doc

@@
 \end{abstract}
+
+\newpage
 
@@
 
+\newpage
 \section{Estado del arte y objetivos específicos ( Apartado 5.2.i, bases de la convocatoria)}
@@
 This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} paradigm for the automatic synthesis and validation of programs}.
-Remarkably, the programs synthesized with our approach are guaranteed to be bug-free over given sets of {\em input-output} tests cases.
-
 Here we briefly introduce the technology we will rely on along the project:
@@
 
-(Please find the full {\em bibliography} at the end of the document)
+(Please find the full list of {\em references} at the end of the document)
 
+\newpage
 \section{Metodología y Plan de Trabajo ( Apartado 5.2.ii, bases de la convocatoria)}
@@
   \centering
-\begin{tt}  
+\begin{small}  
 \begin{tabular}{l@{\hspace*{30pt}}c@{\hspace*{5pt}}}
@@
 \end{tabular}
-\end{tt}  
+\end{small}  
 \caption{\small Time to synthesize the programs with the AI planner {\sc FD}~\cite{helmert2006fast} on a processor {\em Intel Core i5 3.10GHz x 4} and with a 4GB memory bound.}
@@
 
-Given a TDD programming task our approach is modeling and solving it as it were an AI planning task.
-Briefly, the state variables of this AIP task are {\em fluents} of the kind {\tt (var:=value)} while the initial/goal states of the AIP task encode the programming input/output tests of the TDD programming task.
+\subsection{Synthesis and validation of TDD programs as AI planning}
+Given a {\em TDD programming} task our approach is modeling and solving it as it were an {\em AI planning} task.
+Briefly the state variables of this AIP task are {\em fluents} of the kind {\tt var:=value} (that encode the program variables) plus fluents {\tt program(line):=instruction} (that encode the program lines) and fluents {\tt pcounter:=line} (encoding the current program line).
+The initial/goal states of the AIP task encode the value of the program variables according to the input/output tests of the TDD programming task.
 Finally, program instructions are encoded using two kinds of AIP actions:
@@
 We implement this encoding using standard planning languages, such as PDDL~\cite{fox2003pddl2}, so the AIP tasks resulting from our encoding can be solved with off-the-shelf planners, like {\sc FD}~\cite{helmert2006fast}.
-Interestingly, the encoding allows also program validation by (1), specifying the program to validate in the initial state of the AIP task and (2), disabling the mentioned {\it programming actions}.
+The programs synthesized with this approach are guaranteed to be bug-free over given sets of {\em input-output} tests cases.
+
+Interestingly, our PDDL encoding allows also program validation by (1), specifying the lines of the program to validate in the initial state of the AIP task and (2), disabling the mentioned {\it programming actions}.
+
+\subsection{Evaluation}
+The performance of our approach for the synthesis and validation of TDD programs will be evaluated in two kinds of tasks:
+\begin{itemize}
+\item {\em Theoretical benchmarks}: Programming tasks for the computation of mathematical/logic series, string manipulation and for the management of classic data structures such as {\em lists}, {\em queues}, {\em stacks}, {\em trees}, {\em heaps}, \ldots
+\item {\em Application benchmarks}: Practical programming tasks coming from {\em activity recognition} and {\em activity recommendation} domains.
+The GRPS group is currently leading a four-year research project from the Spanish National plan in which {\em activity recognition} and {\em activity recommendation} are applied to different real-world domains such as {\em tourism}, {\em domotics}, {\em traffic control} and {\em robotics}.
+\end{itemize}
 
@@
     \item Design of the user-interaction for the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases plus the available instruction set. 
-    \item Experimental design.
-Experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve TDD programming tasks that are characterized as {\em finite state machines}.
+    \item Experimental design. Experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve the given TDD programming tasks.
       \item Evaluation of different AI planners available, with special attention to the planners that get the best results at the IPC-2018. 
@@
       \item Reporting the experimental performance of our AIP approach for solving TDD programming tasks.
-      \item Dissemination of the obtained theoretical and empirical results by submitting research papers to top international conferences and journals in AI.        
+      \item Dissemination of the obtained theoretical and empirical results by submitting papers to top international conferences and journals in AI.        
       \end{enumerate}
 \end{small}        
-{\small {\bf Deliverable:} Final report with the obtained conclusions and produced publications.}
+{\small {\bf Deliverable:} Final report with the obtained conclusions and the produced publications.}
 \end{enumerate}
 
+Tasks ({\bf T1-T3}) will be developed by the three mentioned members of the research group. In addition, we plan to hire a master student to assist in the completion of task {\bf T3(a)}.
+
+\newpage
 \section{Objetivos específicos del Proyecto ( Apartado 5.2.iii, bases de la convocatoria)}
@@
 
-The specific objective of this project is the study of the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions:
-\begin{enumerate}
-\item Maximum number of program lines.
-\item Maximum number of program variables. 
-\item Maximum size of the available instruction set.  
-\end{enumerate}
-
-In more detail, this project aims to {\bf study the performance of an Artificial Intelligence approach for the automated synthesize and validation of programs up to}: \underline{5 program lines}, \underline{50 variables with binary domain} and, an \\\underline{instruction set that comprises 10 instructions}.
-The performance of our approach will be evaluated with regard to the {\em computation time} and {\em memory} invested in the synthesize and validation of the programs.
-
-Note that despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}.
-With this regard, our AIP encoding already supports callable procedures to decompose a given programming task in simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.
-
-We precisely characterize the specific kind and size of the programs that we can synthesize and validate by modeling programs as {\em finite state machines}~\cite{sergio:aprograming:ijcai16}.
-Figure~\ref{fig:list} shows a program for traversing linked lists pictured as a {\em finite state machine}.
-The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition under which program instructions are taken.
+Figure~\ref{fig:list} shows a synthesized program for traversing linked lists.
+The program is pictured as a {\em finite state machine}: the machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
 
@@
 \end{center}
-\caption{\small {\em Finite state machine} that depicts a program to solve the programming task of traversing a linked list ($n$ is a variable that points to a node of the linked list).}
+\caption{\small Synthesized program to solve the programming task of traversing a linked list ($n$ is a variable that points to a node of the linked list).}
 \label{fig:list}
@@
 
-Formally, we define a program as a tuple $\tup{\mathcal{Q},q_0,Q_\bot,\mathcal{V},,\mathcal{I},\Omega,T,\Gamma}$:
+The objective of this project is to study the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions: number of program lines, number of program variables and size of the available instruction set.
+In more detail, the two specific objectives of this project are:
 \begin{itemize}
-\item $\mathcal{Q}$ is a finite set of program lines where $q_0\in \mathcal{Q}$ is the {\em initial program line} and $Q_\bot\subseteq \mathcal{Q}$ is the subset of {\em terminal program lines}.
-\item $\mathcal{V}$ is the finite set of program variables, where the value of each variable $v\in\mathcal{V}$ is defined in the domain $D_v$.
-\item $\mathcal{I}$ is a finite instruction set where each instruction $a\in \mathcal{I}$ defines an {\em update} $\theta_a$ of a subset of the program variables.
-\item $\Omega:Q\times D_{v_1}\times \ldots \times D_{v_{|\mathcal{V}|}}\mapsto\{0,1\}$ is an {\em observation function} that maps a program line and a condition over the program variables into a Boolean (1 if the condition holds 0 if  it does not).
-\item $T:Q\times\{0,1\}\rightarrow Q$ is a {\em transition function} mapping pairs of a program line and the result of a {\em condition} check into the next program line.
-Likewise $\Gamma:Q\times\{0,1\}\rightarrow \mathcal{I}$ is a {\em transition function} mapping pairs of a program line and the {\em condition} result into the corresponding instruction.
+\item To study the performance of an Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
+\item To study the performance of an Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 20 variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
 \end{itemize}
+In both cases the performance of our approach will be evaluated with regard to the {\bf computation time} and {\bf memory} invested in the synthesize and the validation of the aimed programs.
 
-In the example of Figure~\ref{fig:list}, the edge $(q_0,q_2)$ with label $\mathsf{isEnd}(n)/a_\bot$ encodes that the condition is $\mathsf{isEnd}(n)?$, while $T(q_0,1)=q_2$, $\Gamma(q_0,1)=a_\bot$ encode the transition and associated instruction when $\mathsf{isEnd}(n)$ holds.
-The edge $(q_0,q_1)$ with label $!\mathsf{isEnd}(n)/\mathsf{visit}(n)$ encodes the transition and instruction when $\mathsf{isEnd}(n)$ does not hold, i.e.~$T(q_0,0)=q_1$ and $\Gamma(q_0,0)=\mathsf{visit}(n)$.
-The edge $(q_1,q_0)$ with label $-/\mathsf{step}(n)$ denotes that, when in $q_1$, the transition and instruction are always the same no matter the value of the program variables.
+Despite setting bounds for the programs size and kind, challenging programming tasks can still be addressed using {\em problem decomposition}.
+With this regard, our AIP encoding already supports callable procedures to decompose a given programming task in simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.
 
-The {\em state of a program} is a pair $(q,s)$ that consists of the program line $q\in \mathcal{Q}$ and $s\in D_{v_1}\times \ldots \times D_{v_{|\mathcal{V}|}}$, an assignment to the variables compatible with their domains.
-Programs transition to a program state $(q',s')$ that is computed as follows:
-\begin{enumerate}
-\item First we observe whether the condition associated to the current program line holds (we denote the result of this observation as $O_q(s)\in \{0,1\}$). 
-\item Then we compute the new program line $q'=T(q,O_q(s))$ and the instruction $a=\Gamma(q,O_q(s))$ to apply next.
-Each transition $T(q,O_q(s))$ is associated with a single observation of the variable values however, programs can represent expressive conditions including the no-op action ($a_{\bot}\in\mathcal{I}$) in the instruction set.
-The no-op action does not modify the program variables and affects only to the next program line which allows to concatenate conjunctions of conditions like in {\em decision trees}.
-\item Finally the value of the program variables $s'=\theta_a(s)$ is updated executing the corresponding instruction $a=\Gamma(q,O(s,q))\in \mathcal{I}$.
-\end{enumerate}
-
+\newpage
 \section{Beneficios del proyecto. Difusión y Explotación en su caso de los Resultados.}
 \label{subsec:beneficios}
+Promising research opportunities come from the application of AIP to {\em program synthesis} given that, {\em program synthesis} with a tests base, can also be seen as a dual to {\em program testing}.
+AIP has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}.
+
 This project will provide new insights into the current understanding of how Artificial Intelligence can assist programmers in the software development.
@@
 This project will also help to meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks.
-Promising research opportunities come from the application of AIP to {\em program synthesis} given that, {\em program synthesis} with a tests base, can also be seen as a dual to {\em program testing}.
-AIP has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}.
 
@@
 \item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}.
-\item The development, evaluation and dissemination of a finished {\bf open source software for the {\em Synthesis and Validation of Programs}}.
+\item The development, evaluation and dissemination of {\bf open source software for the {\em Synthesis and Validation of Programs}}.
 \end{enumerate}

### 2018-05-28 src/compiler_new.py
compiler_new.py: More performance improvements

@@ -179,11 +179,12 @@ for a in actions:
     # Define action validation condition
     # Example (and (plan-pickup ?i1 ?o1) (current ?i1) (inext ?i1 ?i2))
     if action_observability > 0:
-        validation_condition = [pddl.conditions.Atom("plan-" + a.name, ["?i1"] + ["?o" + str(i+1) for i in range(a.num_external_parameters) ])]
-        validation_condition += [pddl.conditions.Atom("current", ["?i1"])]
+        # validation_condition = [pddl.conditions.Atom("plan-" + a.name, ["?i1"] + ["?o" + str(i+1) for i in range(a.num_external_parameters) ])]
+        validation_condition = [pddl.conditions.Atom("current", ["?i1"])]
         validation_condition += [pddl.conditions.Atom("inext", ["?i1", "?i2"])]
 
     if action_observability == 1:
+        validation_condition += [pddl.conditions.Atom("plan-" + a.name, ["?i1"] + ["?o" + str(i+1) for i in range(a.num_external_parameters) ])]
         pre += validation_condition
         eff += [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.NegatedAtom("current", ["?i1"]))]
         eff += [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom("current", ["?i2"]))]
@@ -191,8 +192,9 @@ for a in actions:
     elif action_observability > 0:
         # Define conditional effect to validate an action in the input traces
         # This effect advances the program counter when an observed action is executed
-        eff += [pddl.effects.Effect([], pddl.conditions.Conjunction(validation_condition), pddl.conditions.NegatedAtom("current", ["?i1"]))]
-        eff = eff + [pddl.effects.Effect([], pddl.conditions.Conjunction(validation_condition), pddl.conditions.Atom("current", ["?i2"]))]
+        pre += validation_condition
+        eff += [pddl.effects.Effect([], pddl.conditions.Conjunction([pddl.conditions.Atom("plan-" + a.name, ["?i1"] + ["?o" + str(i+1) for i in range(a.num_external_parameters) ])]), pddl.conditions.NegatedAtom("current", ["?i1"]))]
[…]

### 2018-05-28 doc/programming-project/AI-programming.tex
added updates to proposal

@@
 
-The project will be developed within the {\em Group of Reasoning on Planning and Scheduling} (GRPS){\footnote{\tt http://users.dsic.upv.es/grupos/grps/}}, that is part of the {\em Department of Computer Systems and Computation} of the UPV.
+The project will be developed within the {\em Group of Reasoning on Planning and Scheduling} (GRPS, {\tt http://users.dsic.upv.es/grupos/grps/}), that is part of the {\em Department of Computer Systems and Computation} of the UPV.
 In the last 10 years, the GRPS has led 3 research projects of the Spanish national research plan on {\em AI planning}.
-Further, the GRPS has developed its own AI planner, that is participating  at the International Planning Competition (IPC-2018).
+Further, the GRPS has developed its own AI planner, that is participating  at the International Planning Competition (IPC-2018, {\tt https://ipc2018.bitbucket.io}).
 
@@
 
-General quality indicators of scientific research: {\bf 373 citations} (287 since 2011) and {\bf h-index 11}. {\small\em Source google scholar}.
+General quality indicators of scientific research: {\bf 373 citations} (287 since 2013) and {\bf h-index 11}. {\small\em - Source google scholar}.
 
 \subsection{Senior Researcher: Eva Onaindia}
+
+Eva Onaindia is a {\em Full Professor} at the Universitat Politècnica of València.
+She leads the GRPS where she conducts research in Artificial Intelligence, Automated Planning, Causal Model Learning and Recommender Systems.
+She has published numerous articles in top-tier specialized conferences such as the International Joint Conference on Artificial Intelligence (IJCAI), Autonomous Agents and Multi Agent Systems (AAMAS), International Conference on Planning and Scheduling (ICAPS) or European Conference on Artificial Intelligence (ECAI), among others; and scientific journals related to topics of Artificial Intelligence, Planning, Argumentation and Recommender Systems such as IEEE Intelligent Systems, Information Sciences, Engineering Applications of Artificial Intelligence, Knowledge and Information Systems or Applied Intelligence, among others.
+
+She has led several National Research Projects as well as served on multiple ICAPS,AAAI, IJCAI and ECAI conferences as a member of the Program Committee and as a senior member.
+She is Editor-in-Chief of the journal AI Communications, a Review Board member of the journal Applied Intelligence and she will be Program co-chair of ICAPS 2019 (the main international conference on AI planning).
+
+Eva Onaindía has supervised so far a total of 9 PhD theses and she is currently supervising four doctoral students.
+She has also participated in many diverse scientific evaluation committees at national and international level.
+In the national scene, she has collaborated with National Agencies such as ANEP and ANECA and she has been member of different expert boards and scientific committees.
+At an international level, she has participated in the evaluation of research projects for Israel, Italy or Czech Republic.
+
+Eva has also participated in several R\&D and Innovation contracts: with the firm DESTINIA S.L.
+for the development of a tourist recommender system, and currently in the construction of a system for Electricity Price Forecasting aimed at energy savings (Ayuntamiento Lliria) using neural networks and data analysis techniques.
+
+General quality indicators of scientific research: {\bf 1521 citations} (977 since 2013) and {\bf h-index 22}. {\small\em - Source google scholar}.
 
@@
 Our current research already shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16}.
-Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list or traversing a binary tree.
+Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list and traversing a binary tree.
 
@@
 Briefly the state variables of this AIP task are {\em fluents} of the kind {\tt var:=value} (that encode the program variables) plus fluents {\tt program(line):=instruction} (that encode the program lines) and fluents {\tt pcounter:=line} (encoding the current program line).
-The initial/goal states of the AIP task encode the value of the program variables according to the input/output tests of the TDD programming task.
+The initial/goal states of the AIP task encode the initial/final values of the program variables according to the input/output tests of the TDD programming task.
 Finally, program instructions are encoded using two kinds of AIP actions:
@@
 \item {\em Theoretical benchmarks}: Programming tasks for the computation of mathematical/logic series, string manipulation and for the management of classic data structures such as {\em lists}, {\em queues}, {\em stacks}, {\em trees}, {\em heaps}, \ldots
-\item {\em Application benchmarks}: Practical programming tasks coming from {\em activity recognition} and {\em activity recommendation} domains.
+\item {\em Application benchmarks}: Practical programming tasks coming from {\em activity recognition} and {\em recommendation} domains.
 The GRPS group is currently leading a four-year research project from the Spanish National plan in which {\em activity recognition} and {\em activity recommendation} are applied to different real-world domains such as {\em tourism}, {\em domotics}, {\em traffic control} and {\em robotics}.
@@
 \end{small}        
-{\small {\bf Deliverable:} Final report with the obtained conclusions and the produced publications.}
+{\small {\bf Deliverable:} Final report with the obtained conclusions and produced publications.}
 \end{enumerate}
 
-Tasks ({\bf T1-T3}) will be developed by the three mentioned members of the research group. In addition, we plan to hire a master student to assist in the completion of task {\bf T3(a)}.
+Tasks ({\bf T1-T3}) will be developed by the three mentioned members of the research group: Sergio Jiménez, Eva Onaindia and Diego Aineto.
+In addition, we plan to hire a master student to assist in the completion of the task {\bf T3(a)}.
 
@@
 
-The objective of this project is to study the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions: number of program lines, number of program variables and size of the available instruction set.
+The objective of this project is to study the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions: number of {\em program lines}, number of {\em program variables} and size of the available {\em instruction set}.
 In more detail, the two specific objectives of this project are:
@@
 Our current work on {\em program synthesis} with AIP already produced {\bf publications at top international conferences on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
+
 The scientific results obtained during the development of the project will be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).

### 2018-05-28 src/example-generator.py
added updates to the example generator to handle planning horizon

@@ -30,9 +30,12 @@ try:
    problem_filename = sys.argv[2]
    nsteps = int(sys.argv[3])
    planner = sys.argv[4]
+   nhorizon = int(sys.argv[5])      
+
+   
 except:
    print "Usage:"
-   print sys.argv[0] + " <domain> <problem> <steps> <planner>"
+   print sys.argv[0] + " <domain> <problem> <steps> <planner> <horizon>"
    sys.exit(-1)
 
 
@@ -43,16 +46,63 @@ FD_PARAMS=""
 M_PATH=config.PLANNER_PATH
 M_CALL="/M"
 M_PARAMS=config.PLANNER_PARAMS
-
 PLANNER_OUT="planner.log"
+
+
+# Creating a FD task with the domain and the problem file
+fd_domain = pddl_parser.pddl_file.parse_pddl_file("domain", domain_filename)
+fd_problem = pddl_parser.pddl_file.parse_pddl_file("task", problem_filename)
+fd_task = pddl_parser.pddl_file.parsing_functions.parse_task(fd_domain, fd_problem)
+
+# Modifying domain and problem when planning for horizon
+fd_task.types.append(pddl.pddl_types.Type("step", "None"))
+fd_task.predicates.append(pddl.predicates.Predicate("current", [pddl.pddl_types.TypedObject("?i", "step")]))
+fd_task.predicates.append(pddl.predicates.Predicate("inext", [pddl.pddl_types.TypedObject("?i1", "step"), pddl.pddl_types.TypedObject("?i2", "step")]))
+
+
+for a in fd_task.actions:
+   params = []
+   params += [pddl.pddl_types.TypedObject("?i1", "step")]
+   params += [pddl.pddl_types.TypedObject("?i2", "step")]
+   
+   pre = []
+   pre += [pddl.conditions.Atom("current", ["?i1"])]
+   pre += [pddl.conditions.Atom("inext", ["?i1", "?i2"])]
+
+   if isinstance(a.precondition, pddl.conditions.Atom):
+      pre.append(a.precondition)
+   else:
+      pre.extend([x for x in a.precondition.parts])
+   
+   a.effects += [pddl.effects.Effect(params, pddl.conditions.Conjunction(pre), pddl.conditions.NegatedAtom("current", ["?i1"]))]
[…]

### 2018-05-29 benchmarks/aaai19/automata/gen-automata-problem.py
updates added to automata generator

@@ -25,11 +25,12 @@ try:
     nStates  = int(sys.argv[1])
     nObs = int(sys.argv[2])
     str_itape = sys.argv[3]
-    str_otape = sys.argv[4]    
+    str_otape = sys.argv[4]
+    nhead = int(sys.argv[5])
 
 except:
     print "Usage:"
-    print sys.argv[0] + " <nStates> <nObservations> <itape> <otape>"
+    print sys.argv[0] + " <nStates> <nObservations> <itape> <otape> <head position>"
     sys.exit(-1)
 
    
@@ -54,7 +55,7 @@ for n in range(0,len(str_itape)):
    str_out = str_out +  " (symbolO"+str_itape[n]+" t"+str(n)+")"   
 str_out = str_out +  " )\n"
 
-str_out = str_out +  "  (:goal (and (head t"+str(len(str_otape)-1)+") "
+str_out = str_out +  "  (:goal (and (head t"+str(nhead)+") "
 for n in range(0,len(str_otape)):
    str_out = str_out +  " (symbolO"+str_otape[n]+" t"+str(n)+")"   
 str_out = str_out +  ")))"

### 2018-05-29 src/example-generator.py

@@ -13,6 +13,7 @@ def get_types(task,onames):
             output=output + [o.type_name]
    return output
 
+
 def ppossible(p,types):
    if len(p.arguments)!=len(types):
       return False
@@ -21,21 +22,22 @@ def ppossible(p,types):
       if (p.arguments[i].type_name!=types[i]):
          return False
    return True
-                        
+
+
 #**************************************#
 # MAIN
 #**************************************#   
 try:
    domain_filename  = sys.argv[1]
    problem_filename = sys.argv[2]
-   nsteps = int(sys.argv[3])
-   planner = sys.argv[4]
+   planner = sys.argv[3]   
+   nsteps = int(sys.argv[4])
    nhorizon = int(sys.argv[5])      
 
    
 except:
    print "Usage:"
-   print sys.argv[0] + " <domain> <problem> <steps> <planner> <horizon>"
+   print sys.argv[0] + " <domain> <problem> <planner> <steps> <horizon>"
    sys.exit(-1)
 
 
@@ -54,47 +56,48 @@ fd_domain = pddl_parser.pddl_file.parse_pddl_file("domain", domain_filename)
 fd_problem = pddl_parser.pddl_file.parse_pddl_file("task", problem_filename)
 fd_task = pddl_parser.pddl_file.parsing_functions.parse_task(fd_domain, fd_problem)
 
+
 # Modifying domain and problem when planning for horizon
-fd_task.types.append(pddl.pddl_types.Type("step", "None"))
-fd_task.predicates.append(pddl.predicates.Predicate("current", [pddl.pddl_types.TypedObject("?i", "step")]))
-fd_task.predicates.append(pddl.predicates.Predicate("inext", [pddl.pddl_types.TypedObject("?i1", "step"), pddl.pddl_types.TypedObject("?i2", "step")]))
+if nhorizon > 0:
+   fd_task.types.append(pddl.pddl_types.Type("step", "None"))
+   fd_task.predicates.append(pddl.predicates.Predicate("current", [pddl.pddl_types.TypedObject("?i", "step")]))
+   fd_task.predicates.append(pddl.predicates.Predicate("inext", [pddl.pddl_types.TypedObject("?i1", "step"), pddl.pddl_types.TypedObject("?i2", "step")]))
 
-
-for a in fd_task.actions:
-   params = []
[…]

### 2018-05-29 doc/programming-project/AI-programming.tex
updates added to the AI-programming proposal

@@
 The project proposes to address {\bf automatic program synthesis and validation starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine}.
-Our work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI}, the main international conference on Artificial Intelligence.
+Our current work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI}, the main international conference on Artificial Intelligence.
 \end{abstract}
@@
 
-General quality indicators of scientific research: {\bf 373 citations} (287 since 2013) and {\bf h-index 11}. {\small\em - Source google scholar}.
+\underline{General quality indicators of scientific research}: {\bf 373 citations} (287 since 2013) and {\bf h-index 11}. {\scriptsize\em Source google scholar}.
 
@@
 She leads the GRPS where she conducts research in Artificial Intelligence, Automated Planning, Causal Model Learning and Recommender Systems.
-She has published numerous articles in top-tier specialized conferences such as the International Joint Conference on Artificial Intelligence (IJCAI), Autonomous Agents and Multi Agent Systems (AAMAS), International Conference on Planning and Scheduling (ICAPS) or European Conference on Artificial Intelligence (ECAI), among others; and scientific journals related to topics of Artificial Intelligence, Planning, Argumentation and Recommender Systems such as IEEE Intelligent Systems, Information Sciences, Engineering Applications of Artificial Intelligence, Knowledge and Information Systems or Applied Intelligence, among others.
+
+Eva has published numerous articles in top-tier specialized conferences such as the International Joint Conference on Artificial Intelligence (IJCAI), Autonomous Agents and Multi Agent Systems (AAMAS), International Conference on Planning and Scheduling (ICAPS) or European Conference on Artificial Intelligence (ECAI), among others; and scientific journals related to topics of Artificial Intelligence, Planning, Argumentation and Recommender Systems such as IEEE Intelligent Systems, Information Sciences, Engineering Applications of Artificial Intelligence, Knowledge and Information Systems or Applied Intelligence, among others.
 
 She has led several National Research Projects as well as served on multiple ICAPS,AAAI, IJCAI and ECAI conferences as a member of the Program Committee and as a senior member.
-She is Editor-in-Chief of the journal AI Communications, a Review Board member of the journal Applied Intelligence and she will be Program co-chair of ICAPS 2019 (the main international conference on AI planning).
+Eva is Editor-in-Chief of the journal AI Communications, a Review Board member of the journal Applied Intelligence and she will be Program co-chair of ICAPS 2019 (the main international conference on AI planning).
 
@@
 
-General quality indicators of scientific research: {\bf 1521 citations} (977 since 2013) and {\bf h-index 22}. {\small\em - Source google scholar}.
+\underline{General quality indicators of scientific research}: {\bf 1521 citations} (977 since 2013) and {\bf h-index 22}. {\scriptsize\em Source google scholar}.
 
@@
 after a code change via an automated process.
-Tests alert programmers of bugs before handing the code off to clients (the cost of finding a bug when the code is first written is considerably lower than the cost of detecting and fixing it later).
 When all tests pass, the program code is considered {\em complete} while when a test fails, it pinpoints a {\em bug} that must be fixed from the program code.
 Tests cases are a natural form of program specification, programmers often claim {\em 'code that is difficult to test is poorly written'}.
-Further, writing a thorough set of tests cases forces programmers to think through inputs, outputs, and error conditions of programs.
+Further, tests alert programmers of bugs before handing the code off to clients (the cost of finding a bug when the code is first written is considerably lower than the cost of detecting and fixing it later).
+Last but not least, writing a thorough set of tests cases forces programmers to think through inputs, outputs, and error conditions of programs.
 \end{itemize}
@@
 
-Closely related to the aims of the project is the work on the synthesis and validation of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
+Closely related to the aims of the project is also the work on the synthesis and validation of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
 The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}.
@@
 
-Finally the generation of programs from examples is a research question also addressed in inductive Logic Programming (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
+Finally the generation of programs from examples is a research question also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
 ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn a given target concept, expressed as a  logic program,  from  examples  and  background  knowledge, that are expressed as logic facts.
-
-(Please find the full list of {\em references} at the end of the document)
+\\
+\\
+\begin{center}
+{\small (Please find the full list of {\em references} at the end of the document)}
+\end{center}
 
@@
 Given a {\em TDD programming} task our approach is modeling and solving it as it were an {\em AI planning} task.
-Briefly the state variables of this AIP task are {\em fluents} of the kind {\tt var:=value} (that encode the program variables) plus fluents {\tt program(line):=instruction} (that encode the program lines) and fluents {\tt pcounter:=line} (encoding the current program line).
+
+Briefly the state variables of this AIP task are {\em fluents} of the kind:
+\begin{itemize}
+\item {\tt var:=value}, that encode the program variables).
+\item {\tt program(line):=instruction}, that encode the program lines.
+\item {\tt pcounter:=line}, encoding the current program line.
+\end {itemize}
+
 The initial/goal states of the AIP task encode the initial/final values of the program variables according to the input/output tests of the TDD programming task.
+
 Finally, program instructions are encoded using two kinds of AIP actions:
-\begin{enumerate}
+\begin{itemize}
 \item {\it Programming actions}, that set an instruction on a given program line.
 \item {\it Execution actions}, execute the instruction set on a given program line.
-\end{enumerate}
+\end{itemize}
 We implement this encoding using standard planning languages, such as PDDL~\cite{fox2003pddl2}, so the AIP tasks resulting from our encoding can be solved with off-the-shelf planners, like {\sc FD}~\cite{helmert2006fast}.
@@
 \subsection{Evaluation}
-The performance of our approach for the synthesis and validation of TDD programs will be evaluated in two kinds of tasks:
+The performance of our approach for the synthesis and validation of TDD programs will be evaluated in two kinds of programming tasks:
 \begin{itemize}
 \item {\em Theoretical benchmarks}: Programming tasks for the computation of mathematical/logic series, string manipulation and for the management of classic data structures such as {\em lists}, {\em queues}, {\em stacks}, {\em trees}, {\em heaps}, \ldots
-\item {\em Application benchmarks}: Practical programming tasks coming from {\em activity recognition} and {\em recommendation} domains.
+\item {\em Real-world benchmarks}: Practical programming tasks coming from {\em activity recognition} and {\em recommendation} domains.
 The GRPS group is currently leading a four-year research project from the Spanish National plan in which {\em activity recognition} and {\em activity recommendation} are applied to different real-world domains such as {\em tourism}, {\em domotics}, {\em traffic control} and {\em robotics}.
+We plan to evaluate our approach in programming and validation tasks that result from these domains.
 \end{itemize}
@@
 \end{ganttchart}
-\caption{\small Workplan for the development of a user-interactive program synthesizer.}
+\caption{\small Workplan for developing a user-interactive program synthesizer/validator.}
 \label{fig:gantt}
@@
 
-{\small {\bf Deliverable:} Technical report with the specifications of the system design.}
+{\small{\bf\em  Deliverable:} Technical report with the specifications of the system design.}
 
@@
 This component parses the {\em TDD programming task} and produces an {\em AIP task} encoded in the standard planning language PDDL.
-      \item The plan-into-program compiler ({\em Compiler 2}). This component extracts the program code from the solution plan produced by an off-the-shelf AI planner.
+      \item The plan-into-program compiler ({\em Compiler 2}).
+This component extracts the program code and the corresponding validation certificate from the solution plan produced by an off-the-shelf AI planner.
       \end{enumerate}
@@
 
-{\small {\bf Deliverable:} Open repository with the source code of the system architecture.}
+{\small{\bf\em Deliverable:} Open repository with the source code of the system architecture.}
 
@@
 \end{small}        
-{\small {\bf Deliverable:} Final report with the obtained conclusions and produced publications.}
+{\small{\bf\em  Deliverable:} Final report with the obtained conclusions and produced publications.}
 \end{enumerate}
@@
 Tasks ({\bf T1-T3}) will be developed by the three mentioned members of the research group: Sergio Jiménez, Eva Onaindia and Diego Aineto.
-In addition, we plan to hire a master student to assist in the completion of the task {\bf T3(a)}.
+In addition, we plan to hire a master student to assist in the completion of task {\bf T3(a)}.
 
@@
 Figure~\ref{fig:list} shows a synthesized program for traversing linked lists.
-The program is pictured as a {\em finite state machine}: the machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
+The program is pictured as a {\em finite state machine}: The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
 
@@
 The objective of this project is to study the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions: number of {\em program lines}, number of {\em program variables} and size of the available {\em instruction set}.
+For instance, the program of Figure~\ref{fig:list} is generated using three program lines, one observable variable $\mathsf{isEnd}(n)$ with Boolean domain, and a instruction set that comprises two instructions ($\mathsf{visit}(n)$ that marks the list node assigned to $n$ as {\em visited} and $\mathsf{step}(n)$ that advances $n$ to the next node in the linked list).
+
 In more detail, the two specific objectives of this project are:
 \begin{itemize}
-\item To study the performance of an Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
-\item To study the performance of an Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 20 variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
+\item To study the performance of our Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
+\item To study the performance of our Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 20 variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
 \end{itemize}
@@
 
-Despite setting bounds for the programs size and kind, challenging programming tasks can still be addressed using {\em problem decomposition}.
-With this regard, our AIP encoding already supports callable procedures to decompose a given programming task in simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.
+Despite setting bounds for the programs size and kind, note that challenging programming tasks can be addressed using {\em problem decomposition}.
+With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.
 
@@
 \label{subsec:beneficios}
+AIP has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}.
 Promising research opportunities come from the application of AIP to {\em program synthesis} given that, {\em program synthesis} with a tests base, can also be seen as a dual to {\em program testing}.
-AIP has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}.
 
@@
 
-The expected benefits of this project are:
+The expected benefits of this project are three-fold:
 \begin{enumerate}
@@
 
-The scientific results obtained during the development of the project will be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).
+The scientific results obtained during the development of the project will also be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).
+
+\vspace{0.5cm}
 

### 2018-05-30 src/aij18generator.py
script for generating the aij18 benchmarks

@@ -1,0 +1,74 @@
+#! /usr/bin/env python
+import sys,glob,os
+import pddl, pddl_parser
+import config, fdtask_to_pddl
+
+# **************************************#
+# MAIN
+# **************************************#
+try:
+    source_folder_name  = sys.argv[1]
+    destination_folder_name  = sys.argv[2]
+
+except:
+    print "Usage:"
+    print sys.argv[0] + " <source folder name>  <destination folder name>"
+    sys.exit(-1)
+
+
+# Reading the source examples
+for item in sorted(glob.glob(source_folder_name+"/*")):
+   domain_name = item[len(source_folder_name)+1:]
+   domain_filename = source_folder_name + "/" + domain_name + "/empty_domain.pddl"
+   fd_domain = pddl_parser.pddl_file.parse_pddl_file("domain", domain_filename)
+   
+   plan=[]
+   for plan_file_name in sorted(glob.glob(source_folder_name + "/" + domain_name + "/plan-*.txt")):
+       plan_file = open(plan_file_name, 'r')
+       step = plan_file.readline().replace('\n', '')
+       plan.append(step[3:])
+       plan_file.close()
+       
+   # HEAD
+   problem_filenames = sorted(glob.glob(source_folder_name + "/" + domain_name + "/test-*.pddl"))
+   problem_filename = problem_filenames[0]
+   fd_problem = pddl_parser.pddl_file.parse_pddl_file("task", problem_filename)
+   fd_task = pddl_parser.pddl_file.parsing_functions.parse_task(fd_domain, fd_problem)
+
+   str_out = ""
+   str_out = str_out + "(solution \n"
+   str_out = str_out + "(:objects "
+   for i in sorted(set(fd_task.objects)):
+      str_out = str_out + str(i).replace(":"," - ") + " "
+   str_out = str_out + ")\n"   
+   str_out = str_out + "(:init " + fdtask_to_pddl.format_condition([i for i in fd_task.init if i.predicate!="="])+")\n"
+
+   # body
+   counter = 0
+   for test_file_name in problem_filenames:
+       fd_problem = pddl_parser.pddl_file.parse_pddl_file("task", test_file_name)
+       fd_task = pddl_parser.pddl_file.parsing_functions.parse_task(fd_domain, fd_problem)
[…]

### 2018-05-30 doc/aij18/background.tex
background and motivation

@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $s=\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 The {\em plan length} is denoted with $|\pi|=n$ .
@@
 In this work, we will use the term \emph{observation} to refer to the input knowledge of a learning task.
-Hence, $\tup{s_0, s_1, \ldots, s_n}$ is a full observable (FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,a)=s_{i+1}$.
-Otherwise, $\tup{s_0, s_1, \ldots, s_n}$ is a partial observable (PO) state trajectory, meaning that at least one $s_i$ is a partial assignment of values to fluents in which one or more literals are missing.
-Likewise, $\tup{a_1, \ldots, a_n}$ is a FO action sequence if it contains all the necessary actions to achieve a goal from an initial state, and a PO action sequence if at least one necessary action is missing in the sequence.
+Hence, $s=\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$.
+Otherwise, $s$ is a partially-observable (\PO) state trajectory, meaning that at least one $s_i$ is a partial assignment of values to fluents in which one or more literals are missing.
+This general definition gives rise to two particular cases:
+
+\begin{itemize}
+\item when all fluents of a state $s_i$ are missing, $s_i$ is a \emph{missing} or \emph{empty} state
+\item when the $n-1$ intermediate states of a trajectory $s$ are missing, $s$ is a \emph{non-observable} (\NO) state trajectory
+\end{itemize}
+
+Hence, a partially-observed state trajectory can be interpreted as a sequence that contains $n+1$ possibly incomplete states induced by an $n$-action sequence, or as a sequence with some missing intermediate states, which can be regarded as a particular case of partial observability.
+We will refer to the former as a \PO trajectory and to the latter as a \POstar state sequence.
+
+%In the rest of the paper, we will use the following notation to refer to the different levels of state visibility: $s$ denotes a \FO state trajectory; $s^{+}$ refers to a \PO state trajectory with $n+1$ possibly incomplete states; $s^{*}$ indicates a \PO state trajectory with $m$ missing intermediate states, $m < n-1$; and $s^{0}$ is a \NO state trajectory with $n-1$ missing states.
+
+Regarding an action sequence, $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to achieve the goal from an initial state, a \PO action sequence if at least one necessary action to achieve the goal is missing in the sequence and a \NO action sequence in case the sequence is empty.
 
@@
 A plan trace $\tau$ for a planning problem $P=\tup{F,A,I,G}$ satisfies $s_0=I$, $G \in s_n$ and every element in the action sequence of $\tau$ is an action of $A$.
-Additionally, $\tau$ is a plan trace with PO states if $\tup{s_0, s_1, \ldots, s_n}$ is a PO state sequence; and a plan trace with PO actions if $\tup{a_1, \ldots, a_n}$ is a PO action sequence.
+Additionally, plan traces will be classified accordingly to the type of state trajectory (\FO, \PO, \POstar or \NO) and action sequence (\FO, \PO or \NO).
+
+%Additionally, $\tau$ is a plan trace with PO states if $\tup{s_0, s_1, \ldots, s_n}$ is a PO state sequence; and a plan trace with PO actions if $\tup{a_1, \ldots, a_n}$ is a PO action sequence.
 
@@
 In this section we summarize the most recent and relevant approaches to learning action models found in the literature.
-Approaches will be examined according to the following parameters: the input knowledge (plan traces) accepted by the system, the expressiveness of the learned action model and the principal technique used for learning the action model (Table \ref{table:models_comparison1}), as well as the characteristics of the assessment method to validate the learned models (Table \ref{table:models_comparison2}).
+Approaches will be examined according to the following parameters: the input knowledge (plan traces) accepted by the system, the expressiveness of the learned action model and the principal technique used for learning the action model (Table \ref{table:models_comparison1}), as well as the characteristics of the evaluation method to validate the learned models (Table \ref{table:models_comparison2}).
 
-Regarding the input traces, all the approaches analyzed in this work accept partial observations of intermediate states, and some approaches also allow the sequence of intermediate states to be empty.
-With the exception of \LOCM, the minimum state sequence of a plan trace required by the approaches that appear labeled as PO states* in the first column of Table \ref{table:models_comparison1} is $\tup{s_0, G}$.
-Additionally, most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces.
+The first column of Table \ref{table:models_comparison1} shows that all the approaches analyzed in this work accept partial observations of intermediate states, and some approaches also allow the sequence of intermediate states to be empty.
+Although not explicitly stated in the corresponding published works, it seems reasonable to assume that handling \NO state sequences subsumes \POstar trajectories and that a \POstar plan trace can be regarded as composed of several \NO sub-traces.
+Exceptionally, a \NO state sequence in \LOCM is a fully-empty trajectory, with neither initial or final state.
+
+%With the exception of \LOCM, a \NO state trajectory is the minimum state sequence required by the approaches the minimum state sequence of a plan trace required by the approaches that appear labeled as PO states* in the first column of Table \ref{table:models_comparison1} is $\tup{s_0, s_n}$.
+
+Most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces.
 Others do not explicitly say so but the fact is that the predicates and the actions headers are easily extractable from the state sequence and action sequence of the plan traces, respectively.
@@
 
-Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- most approaches rely on a comparison between the learned model and a ground-truth model), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
+Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- almost all approaches rely on a comparison between the learned model and a ground-truth model), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
 
 In the following, we present a comprehensive insight of the particularities of the seven systems presented in Table \ref{table:models_comparison1} and Table \ref{table:models_comparison2}.
-This exposition will help us to highlight at the end of the section the value of our contribution \FAMA.
+This exposition will help us to highlight in section \ref{sec:motivation} the value of our contribution \FAMA.
 
@@
 These constraints are then used to build and solve a weighted propositional satisfiability problem with a MAX-SAT solver.
-Three types of constraints are considered: 1) constraints imposed by general axioms of correct \textsc{Strips} actions, 2) constraints extracted from the distribution of actions in the plan traces and 3) constraints obtained from the PO states, if available.
+Three types of constraints are considered: 1) constraints imposed by general axioms of correct \textsc{Strips} actions, 2) constraints extracted from the distribution of actions in the plan traces and 3) constraints obtained from the \PO states, if available.
 Frequent subsets of actions in which to apply the two latter types of constraints are found by means of frequent set mining.
@@
 \ARMS defines an error metric and a redundancy metric to measure the correctness and conciseness of an action model over the test set of input plan traces using a cross-validation evaluation.
-The model evaluation is posed as an optimization task that returns a learned action model that minimizes the error and redundancy functions.
-This yields a model that is approximately correct (100\% correctness is not required so as to ensure generality and avoid overfitting) and approximately concise (low redundancy rates), and that can explain as many examples as possible.
+The model evaluation is posed as an optimization task that returns the model that best explains the input traces by minimizing the error and redundancy functions.
+This yields a model that is approximately correct (100\% correctness is not required so as to ensure generality and avoid overfitting), approximately concise (low redundancy rates), and that can explain as many examples as possible.
 Hence, there is no guarantee that the learned model of \ARMS explains all observed plans, not even that it correctly explains any of the plan traces of the test set.
 
-The \ARMS system became a benchmark in action-model learning, showing empirically that is is feasible lo learn the models in a reasonably efficient way using a weighted MAX-SAT even when the intermediates states are not observed.
+The \ARMS system became a benchmark in action-model learning, showing empirically that is is feasible lo learn a model in a reasonably efficient way using a weighted MAX-SAT even with \NO state trajectories.
 Nevertheless, we argue that using the home-made error and redundancy metrics to evaluate the model correctness is questionable since a model that exhibits a wrong precondition or effect in any of its actions could still be eligible as the minimal-error output model.
@@
 		\hline			
-		\multirow{2}{*}{\ARMS} & PO states* & \strips & MAX-SAT\\ & FO actions & & \\
+		\multirow{2}{*}{\ARMS} & \PO and \NO states & \strips & MAX-SAT \\ & \FO actions & & \\
         \hline
-        \multirow{2}{*}{\SLAF} & PO states & universal quantifiers in $\eff$ & bayesian learning \\ & FO actions &  & logical filtering \\
+        \multirow{2}{*}{\SLAF} & \PO states  & universal quantifiers in $\eff$ & logical inference \\ & \FO actions &  & SAT solver \\
          \hline
@@
          %\hline
-		\multirow{2}{*}{\LAMP} & PO states* &  quantifiers &  Markov logic networks \\  & FO actions & logical implications &  \\
+		\multirow{2}{*}{\LAMP} & \PO and \PO* states &  quantifiers &  Markov logic networks \\  & \FO actions & logical implications &  \\
          \hline
-         \AMAN & PO states* & \strips & graphical model estimation \\ & noisy actions & & \\
+         \AMAN & \NO states & \strips & graphical model estimation \\ & noisy actions & & \\
          \hline
-         \NOISTA & PO and noisy states & \strips &  classification \\ & FO actions & & \strips \texttt{} rules derivation \\
+         \NOISTA & \PO and noisy states & \strips &  classification \\ & \FO actions & & \strips \texttt{} rules derivation \\
          \hline
-         \CAMA & PO states &  \strips & crowdsourcing annotation\\ & FO actions &  & MAX-SAT \\
+         \CAMA & \PO and \PO* states &  \strips & crowdsourcing annotation\\ & \FO actions &  & MAX-SAT \\
          \hline
-         \LOCM &  FO actions &  predicates and types & \\
+         \LOCMtwo & \NO states &  predicates and types & Finite State Machines \\ & \FO actions & & \\
          \hline
-		\FAMA & PO* states & \strips &   compilation to planning\\ & PO* actions & & \\
+		\FAMA &  \PO, \PO* and \NO states & \strips &   compilation to planning\\ & \PO and \NO actions & & \\
          \hline
 	\end{tabular}
-	\caption{Characteristics of action-model learning approaches (PO: Partial Observability: FO: full observability; *: accepts null observability)}
+	\caption{Characteristics of action-model learning approaches}
 	\label{table:models_comparison1}
@@
 \SLAF extracts all satisfying models of the learned formula with a SAT solver.
-For doing so, the training data set for each domain is composed of randomly generated action-observation sequences  (1000 randomly selected actions and 10 fluents uniformly selected at random per observation).
+For doing so, the training data set for each domain is composed of randomly generated action-observation sequences  (1,000 randomly selected actions and 10 fluents uniformly selected at random per observation).
 Additional processing in the form of replacement procedures or extra axioms are run into the SAT solver when finding the satisfying models.
@@
 The Learning Action Models from Plan Traces (\textbf{\LAMP}) \cite{ZhuoYHL10} algorithm extends the expressiveness to learning models with universal and existential quantifiers as well as logical implications.
-The input to \LAMP is a set of plan traces with PO intermediate states, which are encoded by the algorithm into propositional formulas.
+The input to \LAMP is a set of plan traces with intermediate states, which are encoded by the algorithm into propositional formulas.
 \LAMP then uses the action headers and predicates to build a set of candidate formulas that are validated against the input set using a Markov Logic Network and effectively weighting each formula.
-The formulas with weights larger than a certain threshold will be chosen to represent preconditions and effects of the learned action models.
+The formulas with weights larger than a certain threshold are chosen to represent preconditions and effects of the learned action models.
 
-\LAMP allows observations of intermediate states to be empty up to a minimum percentage of 1/5 of observed states as well as different degrees of PO in the number of propositions in each state.
+\LAMP allows \POstar state observations up to a minimum percentage of 1/5 of the observed states as well as \PO state trajectories with different degrees of observability in the number of propositions in each state.
 It uses an error metric based on counting the differences in the number of precondition and effects between the ground-truth model and the learned model.
 In general, the results show that the accuracy of the learned models is fairly sensitive to the threshold chosen to learn the weights of the candidate formulas, and that domains that feature more conditional effects are harder to learn.
+
+%\LAMP allows observations of intermediate states to be empty up to a minimum percentage of 1/5 of observed states as well as different degrees of PO in the number of propositions in each state.
 
@@
 		\multirow{1}{*}{\ARMS} & cross-validation with a test set & error counting of \#$\pre$ satisfaction  & 6  \\
-        & of plan traces & and redundancy & 160 plan traces\\
+        & of plan traces & and redundancy & 1,600-4,320 actions\\ & & & (160 plan traces) \\
         \hline
@@
          \hline
-		\multirow{1}{*}{\LAMP} & checking wrt GTM  & error counting of extra & 4\\ & & and missing \#$\pre$ and \#$\eff$ & 100-200 plan traces\\
+		\multirow{1}{*}{\LAMP} & checking wrt GTM  & error counting of extra & 4\\ & & and missing \#$\pre$ and \#$\eff$ & 1,300-6,100 actions\\
+           & & & (100-200 plan traces) \\
          \hline
@@
          \hline
-		\LOCM & checking wrt GTM & &   \\
+		\LOCMtwo & manual checking wrt GTM &  ---  &   --- \\
          \hline
@@
 
-The Action Model Acquisition from Noisy plan traces (\textbf{\AMAN}) \cite{zhuo2013action} introduces an algorithm able to learn action models from plan traces with null observations of intermediate states where actions have a probability of being observed incorrectly (noisy actions).
+The Action Model Acquisition from Noisy plan traces (\textbf{\AMAN}) \cite{zhuo2013action} introduces an algorithm able to learn action models from plan traces with \NO state sequences where actions have a probability of being observed incorrectly (noisy actions).
 The first step of the \AMAN algorithm is to build the set of candidate domain models that are compliant with the action headers and predicates.
@@
 Then it solves the constraint-based problem using a MAX-SAT solver and converts the solution to action models.
+An experimental comparison reveals that a manual crowdsourcing of \CAMA outperforms \ARMS and that as expectedly the difference becomes smaller as the number of plan traces becomes larger.
+The accuracy of \CAMA for a small number of plan traces (e.g., 30) is not less than 80\%, which means that exploiting the knowledge of the crowd can help learning action models.
 
-To capture the knowledge from the crowd, \CAMA enumerates all possible preconditions and effects for each action, considering all predicates $p$  whose parameters are included by the parameters of the action $a$.
+%To capture the knowledge from the crowd, \CAMA enumerates all possible preconditions and effects for each action, considering all predicates $p$  whose parameters are included by the parameters of the action $a$.
 \CAMA queries the crowed whether $p$ is a precondition/positive effect/negative effect of $a$, and human annotators reply \emph{yes}, \emph{no} or \emph{cannot tell}.
 The goal is to find an optimal estimator of the \texttt{true} labels given the observation, minimizing the average bit-wise error rate.
-An experimental comparison of a manual crowdsourcing data of \CAMA with \ARMS reveals that \CAMA outperforms \ARMS and that as expectedly the difference becomes smaller as the number of plan traces becomes larger.
-The accuracy of \CAMA for a small number of plan traces (e.g., 30) is not less than 80\%, which means that exploiting the knowledge of the crowd can help learning action models.
 
-\vspace{0.5cm}
+The Learning Object-Centred Models (\textbf{\LOCM}) is an approach that only requires the \FO action sequence as input knowledge, without need for providing any information about the predicates or the state trajectory, not even the initial or final state~cite{cresswell2013acquiring}.
+The lack of available information is overcome by exploiting assumptions about the structure of the actions.
+Particularly, it assumes that objects found in the same position in the header of actions are grouped as a collection of objects (sorts) whose defined set of states is captured by a parameterized Finite State Machine (FSM).
+The intuitive assumptions of \LOCM, like the continuity of object transitions or the association of parameters between consecutive actions in the training datset, yield a learning model heavily reliant on the kind of domain structure.
+A later work, \textbf{\LOCMtwo}, extends the applicability of the \LOCM algorithm to a wider range of domains by introducing a richer representation that allows using multiple FSMs to represent the state of a sort~\cite{cresswell2011generalised}.
 
-==========================================================
+\LOCMtwo is not experimentally evaluated and only the outcome of running the \LOCMtwo algorithm on several benchmark domains wrt to the reference model is reported.
+It is worth noting the last contribution of \LOCM, called \textbf{\LOP} (\LOCM with Optimized Plans), addresses the problem of inducing static predicates~\cite{GregoryC16}.
+\LOP applies a post-processing step after the \LOCM analysis and it requires additional input information, particularly a set of optimal plans besides the suboptimal \FO action sequences.
 
-The Learning Object-Centred Models (\LOCM) \cite{cresswell2013acquiring} system does not limit itself to learning action models and instead is able to learn full domains from plan traces without any observation of the states.
-Furthermore, \LOCM does not require any domain background knowledge such as predicates or objects and to this day can be considered the learning algorithm requiring the least amount of input data.
-\LOCM works by making some assumptions about the structure of the actions.
-Objects found in the same position in the header of actions are grouped into sorts and state machines describing the behavior of objects are built for each sort.
-In a later work \LOCMtwo extends the applicability of the \LOCM algorithm to a wider range of domains by introducing a richer representation using multiple state machines.
-
-%More recent work on learning planning action models \cite{WalshL08} shows that although learning \strips\ operators from pure interaction with the environment requires an exponential number of samples, access to an external teacher can provide solution traces on demand.
-
-%Unlike the previous approaches, the one described in \cite{MouraoZPS12} deals with both missing and noisy predicates in the observations.
-An action model is first learnt by constructing a set of kernel classifiers which tolerate noise and partial observability and then \strips rules are derived from the classifiers' parameters.
-
-\LOCM accepts input traces of the form  $\tau = \langle a_1, a_1, a_2, \ldots, a_n\rangle$ ..
-
-{\sf LOCM} only requires the example plans as input without need for providing information about predicates or states~\cite{cresswell2013acquiring}.
-This makes {\sf LOCM} be most likely the learning approach that works with the least information possible.
-The lack of available information is addressed by LOCM by exploiting assumptions about the kind of domain model it has to generate.
-Particularly, it assumes a domain consists of a collection of objects (sorts) whose defined set of states can be captured by a parameterized Finite State Machine (FSM).
-The intuitive assumptions of {\sf LOCM}
-, like the continuity of object transitions or the association of parameters between consecutive actions in the training sequence,
-yield a learning model heavily reliant on the kind of domain structure.
-The inability of {\sf LOCM} to properly derive domain theories where the state of a sort is subject to different FSMs is later overcome by {\sf LOCM2} by forming separate FSMs, each containing a subset of the full transition set for the sort~\cite{cresswell2011generalised}.
-{\sf LOP} ({\sf LOCM} with Optimized Plans ~\cite{gregory2015domain}), the last contribution of the {\sf LOCM} family, addresses the problem of inducing static predicates.
-Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
+ %Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
 {\em blocksworld} and {\em freecell}).
@@
 
-Recenctly classical planning compilations have been defined to learn different kinds of generative models from examples.
+%Recently classical planning compilations have been defined to learn different kinds of generative models from examples.
 The existing compilations for computing FSCs for generalized planning follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it and hence, they tightly integrate planning and generalization.

### 2018-05-30 doc/aij18/motivation.tex
background and motivation

\section{Motivation}
\label{sec:motivation}

The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which makes the learning task more challenging. Essentially, this brings two key differences: a) the transition function between two observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_n})=s_{i+1}$; and b) the horizon of the input plan traces is not longer known. These two factors justify the use of planning techniques for solving the learning task over other methods such as a SAT solver.

We set a probability threshold of observability to each fluent and action, which determines the percentage of literals in each state and in turn the percentage of observed states (when the probability of every fluent is above the threshold, the result is a missing state). This way, \FAMA must always work under the assumption that the length of the input plan traces is unknown.

The common assumption of having \FO action sequences in a learning task is rather unrealistic in many domains because it implies the existence of human observers that annotate the observed action sequences. In some real-world applications, the observed and collected data are sensory data (e.g., home automation, robotics) or images (e.g. traffic) and one cannot rely on human intervention for labeling actions. Actually, learning the executed actions can also be part of the action-model learning task. Learning from unstructured data involves some prior processing to transform the sensor or image information into a predicate-like format before applying the action-model learning approach~\cite{AsaiF18}, and it also requires the ability of identifying action symbols. In this sense, this work represents a step ahead towards learning action models without assuming observed actions.

Obviously, learning from \PO and \NO action sequences is much more complex. When the number of actions and the number of states is unknown, solving the learning task is as hard as solving a \strips\texttt{} planning problem; i.e., it is PSPACE-complete. Otherwise, the complexity of the task is NP-complete, in which case it can be translated into SAT, which is known to be a NP-complete problem (see Table \ref{tab:complex}). This is the reason that SAT solving is a common technique in the approaches presented in section \ref{sec:background}.

\begin{table}[ht]
\centering
\begin{tabular}{c|c|c|c|c|}
	& \multicolumn{4}{c|}{\emph{state observability}} \\ \cline{2-5}
	\multirow{1}{*}{\emph{action}} & \FO & \PO & \POstar & \NO\\ {\emph{observability}} & & & & \\ \hline
	\FO & NP-complete & NP-complete & NP-complete & NP-complete \\ \hline
	\PO & NP-complete & PSPACE-complete & PSPACE-complete & PSPACE-complete \\ \hline
	\NO & NP-complete & PSPACE-complete & PSPACE-complete & PSPACE-complete\\ \hline
\end{tabular}
\caption{Complexity of learning tasks according to the type of input trace}
\label{tab:complex}
\end{table}

Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists in counting the missing and extra fluents that appear in the learned model wrt the GTM and normalizing by the the total number of all the possible preconditions and effects of an action model. This optimistic metric returns error rates that do not reflect the accuracy of the learned model as it could output a rate below 100\% for a totally wrong learned model that only includes a subset of all possible preconditions and effects of an action model. To overcome this limitation, we propose to use two standard metrics, {\em precision} and {\em recall}. These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.

A striking figure of Table \ref{table:models_comparison2} that emphasizes a relevant feature of \FAMA is the small size of the training dataset it requires in comparison to other approaches. Unlike extensive-data ML approaches, our work explores an alternative research direction to learn sound models from small amounts of plan traces. This is an important advantage, particularly in domains in which it is costly or impossible to obtain a significant number of training samples. Unlike \CAMA, our approach does not require human intervention to label samples as it is able to learn from very small datasets.

Finally, we would also like to point out that, as shown in section \ref{experiments}, \FAMA is exhaustively evaluated over a wide range of domains (15 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner.

### 2018-05-31 doc/aij18/background.tex
small updates to sections 1 and 2

@@
 
-This section serves two purposes.
-In subsection \ref{basic_planning} we introduce the basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper.
-Section \ref{related_work} summarizes existing approaches to learn action models in classical planning and highlights the novelty of our contribution via the related work.
+This section serves two purposes: First we introduce the basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper.
+Second, we summarize the existing approaches to learn action models in classical planning and highlight our contributions via the related work.
 
@@
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
+A solution plan is {\em optimal} if it has minimum length.
 
-In this work, we will use the term \emph{observation} to refer to the input knowledge of a learning task.
-Hence, $s=\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$.
-Otherwise, $s$ is a partially-observable (\PO) state trajectory, meaning that at least one $s_i$ is a partial assignment of values to fluents in which one or more literals are missing.
-This general definition gives rise to two particular cases:
-
+In this work, we will use the term \emph{plan trace} to refer to the observation of a plan execution that starts on a given initial state.
+\emph{Plan traces} are the input knowledge of the learning tasks that are addressed in this paper.
+A \emph{plan trace} is a combination of a state and an action sequence $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
+This sequence is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applied in state $s_{i-1}$ and generates the successor state $s_i$.
+We name different classes of \emph{plan traces} according to the kind of knowledge that is observed:
 \begin{itemize}
-\item when all fluents of a state $s_i$ are missing, $s_i$ is a \emph{missing} or \emph{empty} state
-\item when the $n-1$ intermediate states of a trajectory $s$ are missing, $s$ is a \emph{non-observable} (\NO) state trajectory
+  \item[{\bf\FO}] We say that a plan trace $\mathcal{T}=\tup{s_0, s_1, \ldots, s_n}$ is a {\em fully-observable} (\FO) {\em state trajectory} if every state $s_i$, {\small $1\leq i\leq n$}, is a full assignment of values to fluents and the transition from states $s_i$ to states $s_{i+1}$ is caused by the execution of a single action $a_i$ that is unobserved; that is, $\theta(s_i,\tup{a_i})=s_{i+1}$.
+  \item[{\bf\PO}] A plan trace $\mathcal{T}$ is a {\em partially-observable} (\PO) state trajectory, meaning that at least one $s_i\in\mathcal{T}$ is a partial assignment of values to fluents in which one or more literals are missing, formally there exists a state $s_i$, {\small $1\textless i$} such that $|s_i|\textless|F|$.
+  \item[{\bf\POstar}] \emph{Plan traces} with some missing intermediate states is regarded as a particular case of partial observability (that we name \POstar). 
+  \item[{\bf\NO}]A state trajectory $\mathcal{T}$ where every intermediate state $s_i\in\mathcal{T}$, {\small $1\textless i$}, holds that $|s_i|=0$ is a \emph{non-observable} (\NO) state trajectory.
 \end{itemize}
 
-Hence, a partially-observed state trajectory can be interpreted as a sequence that contains $n+1$ possibly incomplete states induced by an $n$-action sequence, or as a sequence with some missing intermediate states, which can be regarded as a particular case of partial observability.
-We will refer to the former as a \PO trajectory and to the latter as a \POstar state sequence.
-
-%In the rest of the paper, we will use the following notation to refer to the different levels of state visibility: $s$ denotes a \FO state trajectory; $s^{+}$ refers to a \PO state trajectory with $n+1$ possibly incomplete states; $s^{*}$ indicates a \PO state trajectory with $m$ missing intermediate states, $m < n-1$; and $s^{0}$ is a \NO state trajectory with $n-1$ missing states.
-
-Regarding an action sequence, $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to achieve the goal from an initial state, a \PO action sequence if at least one necessary action to achieve the goal is missing in the sequence and a \NO action sequence in case the sequence is empty.
-
-A \emph{plan trace} is a combination of a state and an action sequence $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
-A plan trace $\tau$ for a planning problem $P=\tup{F,A,I,G}$ satisfies $s_0=I$, $G \in s_n$ and every element in the action sequence of $\tau$ is an action of $A$.
-Additionally, plan traces will be classified accordingly to the type of state trajectory (\FO, \PO, \POstar or \NO) and action sequence (\FO, \PO or \NO).
-
-%Additionally, $\tau$ is a plan trace with PO states if $\tup{s_0, s_1, \ldots, s_n}$ is a PO state sequence; and a plan trace with PO actions if $\tup{a_1, \ldots, a_n}$ is a PO action sequence.
+Regarding to the action sequence, $\tup{a_1, \ldots, a_n}$, it is called a \FO action sequence if it contains all the necessary actions $a_i$ to transit every state $s_i$ to the corresponding succesor state $s_{i+1}$, for each {\small $1\leq i\leq n$}.
+We say that an action sequence is \PO if at least one of these necessary $a_i$ actions is missing in the sequence and \NO, in case that the action sequence is empty.
 

### 2018-05-31 doc/aij18/introduction.tex
small updates to sections 1 and 2

@@
 \section{Introduction}
-\label{sec:Section1}
+\label{sec:introduction}
 
 Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-At both planning tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
-Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
+At both planning tasks, automated planners reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
+Unfortunately, modeling planning actions is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
 
@@
 \item The {\em input} to ML algorithms (the {\em learning/training} data) usually are finite vectors encoding the value of fixed features in a given set of objects.
-The input for learning planning action models are observations of plan executions (where each plan possibly has a different length and may involve a different number of objects).
+The input for learning planning action models are observations of plan executions (where each observed plan possibly has a different number of steps and involves a different number of objects).
 \item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning \strips\ action models the output is, for each action, the sets of {\em preconditions}, {\em negative} and {\em positive effects} that define which state transitions are possible.
+When learning \strips\ action models the output is, for each action, the sets of {\em preconditions}, {\em negative} and {\em positive effects} that define the possible state transitions.
 \end{itemize}
@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this \strips\ model satisfies a set of plan executions that is given as input.
+A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this model satisfies the plan executions given as input.
 
-The compilation approach is appealing by itself because it leverages off-the-shelf planners for learning and because its practicality allow us to report learning results over a wide range of domains from the {\em International Planning Competition} (IPC).
+The compilation approach is appealing by itself because it leverages off-the-shelf planners and because its practicality allow us to report learning results over a wide range of domains from the {\em International Planning Competition} (IPC).
 Moreover, it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
-Apart from these, our classical planning compilation presents the following contributions:
+Apart from these, our compilation exhibits the following contributions:
 \begin{enumerate}
-\item {\em Input flexibility}.
+\item {\bf\em Input flexibility}.
 Our classical planning compilation is flexible to various amount and kind of available input knowledge.
@@
 
-\item {\em Model validation}.
+\item {\bf\em Model validation}.
 The compilation poses a novel framework to assess the validation of a STRIPS model with respect a plan execution.
@@
 
-\item {\em Model evaluation}.
+\item {\bf\em Model evaluation}.
 The compilation can assess how well a given \strips\ action model matches a plan execution, which allows us to assess learning performance without having the actual action model.
-The idea is to assess the amount of edition required by the input action model to induce the given observation of the plan execution.
+The idea is to assess the amount of edition that is required by the input action model to induce the given observation of the plan execution.
 \end{enumerate}
 
-A first description of the compilation previously appeared in the conference paper~\cite{aineto2018learning}. Compared to that conference paper, this work includes the following novel material:
+A first description of the compilation previously appeared in our conference paper~\cite{aineto2018learning}. Compared to that conference paper, this work includes the following novel material:
 \begin{itemize}
-\item A unified formulation for learning and evaluating \strips\ action models from observed executions of plans. 
-\item Redefining the {\em precision} and {\em recall} metrics to evaluate learned action models with respect to a plan execution.
+\item A unified formulation for learning and evaluating \strips\ action models from observed executions of plans. Further, these executions can only comprise state observations.
+\item A redefinition of the ML metrics {\em precision} and {\em recall} to evaluate \strips\ action models with respect to observations of plan executions.
 %\item Leveraging diverse forms of {\em background knowledge} for learning and evaluating \strips\ action models.
@@
 
-Section~\ref{sec:Section2} reviews related work on learning planning action models.
+Section~\ref{sec:background} introduces classical planning and reviews related work on learning planning action models.
 Section~\ref{sec:Section3} introduces the classical planning model and the \strips\ action model (the target of our learning and evaluation tasks).

### 2018-05-31 doc/aij18/motivation.tex
small updates to sections 1,2,3

@@
 
-The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which makes the learning task more challenging.
-Essentially, this brings two key differences: a) the transition function between two observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_n})=s_{i+1}$; and b) the horizon of the input plan traces is not longer known.
-These two factors justify the use of planning techniques for solving the learning task over other methods such as a SAT solver.
+The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which defines a more challenging learning task.
+Essentially, this brings one key difference:
+\begin{itemize}
+\item The transition between two given observed states may now involve the execution of more than one action; formally this means that $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$ where $k$ is unkown; The horizon of the input plan traces is not longer known what justify the use of planning techniques for solving the learning task over other methods such as a SAT solver.
+\end{itemize}
 

### 2018-06-01 doc/aij18/motivation.tex
aij18 motivation section revised

--- 2018-05-30 doc/aij18/motivation.tex
@@
 
-The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which makes the learning task more challenging.
-Essentially, this brings two key differences: a) the transition function between two observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_n})=s_{i+1}$; and b) the horizon of the input plan traces is not longer known.
-These two factors justify the use of planning techniques for solving the learning task over other methods such as a SAT solver.
+The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling input plan traces whose length is a priori unbound. This brings one essential difference:
+\begin{itemize}
+\item The transition between two given observed states may now involve the execution of more than a single action.
+Formally $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unkown and unbound;
+\end{itemize}
 
-We set a probability threshold of observability to each fluent and action, which determines the percentage of literals in each state and in turn the percentage of observed states (when the probability of every fluent is above the threshold, the result is a missing state).
-This way, \FAMA must always work under the assumption that the length of the input plan traces is unknown.
+In this particular scenario the number of actual plan traces that can correspond with the given input observations is also unbound and grows exponentially with the actual length of the plan trace (that is now unknown).
+Otherwise, the learning task is SAT compilable, which is known to be a NP-complete task~\cite{russell2016artificial}.
+This is the reason that SAT solving is a common technique in the approaches presented in section \ref{sec:background}.
 
+Under this new assumption, the learning of \strips\ action models that satisfy a set of observations of plan executions becomes PSPACE-complete instead of NP-complete, which motivates planning as a more  suitable approach for the learning of \strips\ action models.
+The length a plan trace, that represent the observation of a plan execution, is a priori unbound because of the conjunction of the following two causes:
+\begin{enumerate}
+\item There may be an unbound number of missing intermediate states in the trace because of partial state observability.
+The assumption of having \FO state trajectory means that the sensors are able to capture every state change at every instant which typically is unrealistic.
+Normally the process of getting state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two sensor readings.
+
+\item There may be an unbound number of missing intermediate actions in the trace because of partial observability.
 The common assumption of having \FO action sequences in a learning task is rather unrealistic in many domains because it implies the existence of human observers that annotate the observed action sequences.
@@
 In this sense, this work represents a step ahead towards learning action models without assuming observed actions.
+\end{enumerate}
 
-Obviously, learning from \PO and \NO action sequences is much more complex.
-When the number of actions and the number of states is unknown, solving the learning task is as hard as solving a \strips\texttt{} planning problem; i.e., it is PSPACE-complete.
-Otherwise, the complexity of the task is NP-complete, in which case it can be translated into SAT, which is known to be a NP-complete problem (see Table \ref{tab:complex}).
-This is the reason that SAT solving is a common technique in the approaches presented in section \ref{sec:background}.
-
+The Table~\ref{tab:complex} precisely characterizes the scenarios where the length of the input plan traces cannot longer be assumed to be unbound.
 \begin{table}[ht]
@@
 \end{tabular}
-\caption{Complexity of learning tasks according to the type of input trace}
+\caption{Complexity of learning tasks according to the kinf of the input plan traces.}
 \label{tab:complex}
@@
 Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists in counting the missing and extra fluents that appear in the learned model wrt the GTM and normalizing by the the total number of all the possible preconditions and effects of an action model.
-This optimistic metric returns error rates that do not reflect the accuracy of the learned model as it could output a rate below 100\% for a totally wrong learned model that only includes a subset of all possible preconditions and effects of an action model.
-To overcome this limitation, we propose to use two standard metrics, {\em precision} and {\em recall}.
+This is an {\em optimistic} metric since it returns error rates that are not normalized with respect to the size of the actual models.
+For intance these metrics it could output a rate below 100\% for a totally wrong learned model that only includes a subset of all possible preconditions and effects of an action model.
+
+To overcome this limitation, we propose to use two standard ML metrics, {\em precision} and {\em recall}.
 These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
@@
 Finally, we would also like to point out that, as shown in section \ref{experiments}, \FAMA is exhaustively evaluated over a wide range of domains (15 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner.
+
+%We set a probability threshold of observability to each fluent and action, which determines the percentage of literals in each state and in turn the percentage of observed states (when the probability of every fluent is above the threshold, the result is a missing state).
+This way, \FAMA must always work under the assumption that the length of the input plan traces is unknown.

### 2018-06-04 doc/aaai-19/model-recognition.tex
citing model reconciliation
updates added to aaai19 proposal

@@
 \title{Model Recognition as Planning}
-\author{\#39}
 
 % Commented for blind submission
-\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia\\
+\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia \and Miquel Ram\'irez\\
 {\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
@@
 \begin{abstract} 
-Given a set of possible action models and a partially observed plan execution, {\em model recognition} is the task of identifying which of these action models has the highest probability of producing the given plan execution.
-{\em Model recognition} enables identifying algorithms by observing their execution besides if an action model is successfully recognized, the planning model-based machinery becomes applicable.
-This paper formalizes the {\em model recognition} task and proposes a method to assess the probability of a given \strips\ action model to produce a partially observed plan execution.
+Given a set of planning models and a partially observed plan execution, we define {\em model recognition} as the task of identifying which model in the set has the highest probability of producing the input observation.
+The paper formalizes the {\em model recognition} task and proposes a method to assess the probability of a given \strips\ model to produce a partially observed plan execution.
 This method, that we called {\em model recognition as planning}, is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
-The effectiveness of {\em model recognition as planning} is shown with sets of \strips\ models that represent different {\em Turing Machines}.
-We show that {\em model recognition as planning} effectively identifies the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or several tape values, are unobserved.
+The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different {\em Turing Machines}.
+We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or the values of several tape cells, are unknown.
 \end{abstract}
@@
 \label{sec:introduction}
-{\em Plan recognition} is the task of predicting the future actions of an agent provided the observation of its current behaviour.
-{\em Goal recognition} is a subtask of plan recognition with the particular aim of discovering the final objectives of the observed agent.
-Diverse methods has been proposed for plan and goal recognition~\cite{carberry2001techniques} such as {\em rule-based systems},  {\em parsing}  (both  conventional and stochastic), {\em graph-covering}, {\em Bayesian nets}, \ldots
+{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behaviour.
+Plan recognition is considered {\em automated planning} in reverse; while automated planning aims to compute a sequence of actions that accounts for a given goals, plan recognition aims to compute the goals that account for an observed sequence of actions~\cite{geffner:book:2013}.
 
-{\em Plan recognition as planning} defined the model-based approach for plan/goal recognition~\cite{ramirez2012plan,ramirez2009plan}.
+Diverse approaches has been proposed for plan recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{carberry2001techniques}.
+{\em Plan recognition as planning} is the model-based approach for plan recognition~\cite{ramirez2012plan,ramirez2009plan}.
 This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal according to the observed plan execution.
 
-In this paper we introduce the {\em model recognition} task as the task of identifying the action model with the highest probability of producing an observed plan execution.
-This task is of interest because :
+In this paper we introduce the task of {\em model recognition}.
+Given a set of planning models and a partially observed plan execution, {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation.
+{\em Model recognition} is of interest because:
 \begin{itemize}
-\item Once the action model is recognized, then the planning model-based machinery~\cite{ghallab2004automated,geffner:book:2013} becomes applicable.
-\item It enables identifying algorithms by observing their execution.
-Many computational models (like {\sc GOLOG} programs or finite state controllers) are compilable into classical planning~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2016hierarchical}.
+\item Once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
+\item It enables recognizing algorithms by observing their execution.
+Planning models can encode diverse forms of algorithm representation, like {\em {\sc GOLOG} programs}, {\em finite state controllers} or {\em push-down automata}~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}.
 \end{itemize}
 
-The paper introduces {\em model recognition as planning}, a novel method to assess the probability of a given \strips\ action model to produce an observed plan execution.
-This method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
-Last but not least, the paper evaluates the effectiveness of {\em model recognition as planning} with sets of \strips\ models that represent different {\em Turing Machines}, all of them defined within the same tape alphabet and same machine states.
-We show that {\em model recognition as planning} effectively identifies the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or several tape values, are unobserved.
+This paper introduces {\em model recognition as planning}, a novel method to assess the probability of a given \strips\ model to produce an observed plan execution.
+The method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
+The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em Turing Machines} (all of them defined within the same {\em tape alphabet} and same {\em machine states} but different {\em transition functions}).
+We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or the values of several tape cells, are unknown.
 
@@
 \label{sec:background}
-This section introduces the formal models for classical planning and for \strips\ actions.
+This section formalizes classical planning and characterizes the space of \strips\ models.
 
@@
 
-The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
-
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
@@
 
-\subsection{\strips\ action schemas}
-This work addresses the learning and evaluation of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+\subsection{The space of \strips\ models}
+Like in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
+Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
+Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
+$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+We assume also that actions $A$ are instantiated from a set of action schemas.
 Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
-
-\begin{figure}[hbt!]
+\begin{figure}
 \begin{scriptsize}
@@
 \end{scriptsize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
+ \caption{\small \strips\ action schema coded in PDDL.}
 \label{fig:stack}
@@
 
-To formalize the target of the learning and evaluation tasks, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be an additional set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
+For instance, in a three-block four-operator {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
+We define $F_v$, a new set of fluents, $F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}.
+In the mentioned {\em blocksworld} this set contains 11 elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
-
-We define $F_v$, a new set of fluents s.t.
-$F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}, and that defines the elements that can appear in the action schemes.
-In {\em blocksworld} this set contains 11 elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-
-In more detail, for a given operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in that action schema.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the elements from $F_v$ that involve $v_2$ because {\small\tt pickup($v_1$)} has arity one.
-
-We assume also that actions $a\in A$ are instantiated from \strips\ operator schemas $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+We define an \strips\ action schema as the tuple $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
 The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
@@
 \end{itemize}
-Given the set of predicates $\Psi$ and the header of the operator schema $\xi$, $2^{2|F_v(\xi)|}$ defines the size of the space of possible \strips\ models for that operator.
-Note that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
-For the {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while for the {\small\tt pickup} operator this number is only 1024.
 
-We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi'$).
-For instance, we can claim that blocksworld operators {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not.
-Last but not least, two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
+For a given $head(\xi)$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in the preconditions and effects of the corresponding action schema.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes any element from $F_v$ that involves $v_2$ because {\small\tt pickup($v_1$)} has arity one.
+The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2|F_v(\xi)|}$.
+Note that \strips\ constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
+For the mentioned {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while $2^{2|F_v(pickup)|}=1024$.
+
+We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi')$).
+For instance, we claim that blocksworld operators {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not.
+Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
 
@@
 
+Recently {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} also proposes model edition but in this case the aim is to conform the different PDDL models of two different agents.
+
 \bibliographystyle{aaai}

### 2018-06-04 doc/aaai-19/model-recognition.tex
updates added to aaai19 proposal

@@
 A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The {\em plan length} is denoted with $|\pi|=n$ .
+The {\em plan length} is denoted with $|\pi|=n$.
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the application of the plan $\pi$ in the initial state $I$.
@@
 
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be an additional set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be an additional set of objects ($\Omega\cap\Omega_v=\emptyset$), that we denote as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
 For instance, in a three-block four-operator {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
 We define $F_v$, a new set of fluents, $F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}.
-In the mentioned {\em blocksworld} this set contains 11 elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+This set contains eleven elements for the mentioned {\em blocksworld}, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
@@
 
-For a given $head(\xi)$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in the preconditions and effects of the corresponding action schema.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes any element from $F_v$ that involves $v_2$ because {\small\tt pickup($v_1$)} has arity one.
+For a given $head(\xi)$, then $F_v(\xi)\subseteq F_v$ defines the subset of elements that can appear in the preconditions and effects of the corresponding action schema.
+For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes any element in $F_v$ that involves $v_2$ because {\small\tt pickup($v_1$)} has arity one.
 The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2|F_v(\xi)|}$.
-Note that \strips\ constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
+Note that \strips\ constraints require negative effects appearing as preconditions and that they cannot be positive effects and also, that positive effects cannot appear as preconditions.
 For the mentioned {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while $2^{2|F_v(pickup)|}=1024$.
 
-We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi')$).
+We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas share the same space of possible \strips\ models, i.e.
+iff $pars(\xi)=pars(\xi')$.
 For instance, we claim that blocksworld operators {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not.
-Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
+Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
 
@@
 \label{sec:recognition}
-Given a finite and non-empty set of {\em comparable} action models $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, and a partially observed plan execution $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ that is obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ s.t., for each {\small $1\leq i\leq n$}, $a_i$ is applied in $s_{i-1}$ and generates $s_i$.
-{\em Model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.
+Given $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, a finite and non-empty set of planning models and $\mathcal{T}$, the observation of a plan execution, {\em model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.
+Our approach is to estimate this probability according to the amount of {\em edition} required by $\mathcal{M}$ to produce a plan execution that is compliant with $\mathcal{T}$.
+
+This work assumes that models $\mathcal{M}\in M$ are {\em comparable} \strips\ planning models, so {\em edition} is focused on the action schemas $\xi\in\mathcal{M}$.
+In addition, the observation $\mathcal{T}$ is a sequence of {\em partially-observed} states and actions, $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ s.t.:
+\begin{itemize}
+\item Intermediate states $s_i\in\mathcal{T}$, {\small $1\textless i$}, are partial assignment of values to fluents in which one or more literals can be missing, formally $0\leq|s_i|\leq|F|$.
+In the extreme full intermediate states can be missing, i.e, $|s_i|=0$.
+\item Intermediate actions $a_i$, {\small $1\leq i$}, that transit between an observed state, $s_i$, to the next observed state, $s_{i+1}$, can be missing. 
+\end{itemize}
+The conjunction of these two assumptions makes that transitions between two consecutive observed states may involve the execution of more than a single action.
+Formally $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound.
+This means that the actual length of the plan whose execution is observed is unknown.
 
 \subsection{The \strips\ edit distance}
-Our approach for {\em model recognition} is to assess how well an action model $\mathcal{M}\in M$ explains $\mathcal{T}$ according to the amount of {\em edition} required by $\mathcal{M}$ to produce $\mathcal{T}$.
-
-In this work we assume that action models $\mathcal{M}\in M$ are represented with the \strips\ formalism. We define here the two allowed \emph{operations} to edit an \strips\ action model:
+We define two edit \emph{operations} on a \strips\ model $\mathcal{M}\in M$:
 \begin{itemize}
@@
 \begin{definition}
-The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
+The \textbf{maximum edit distance} of an \strips\ model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
 \end{definition}
 
-We define now an edit distance to asses the matching of a learned action model with respect to a plan trace $\mathcal{T}$.
+We define now an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.
 
 \begin{definition}
-  Given $\mathcal{M}$, a \strips\ action model built from $F_v$ and $\mathcal{T}$, a plan trace built with fluents in $F$ and actions in $A$.
-The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a valid plan trace $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
+  Given $\mathcal{M}$, a \strips\ action model built from $F_v$ and $\mathcal{T}$, an observation built with fluents in $F$ and actions in $A$.
+The {\bf observation edit distance}, $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
+$\mathcal{M}'$ produces a plan execution compliant with $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
-
-The $\delta(\mathcal{M},\mathcal{T})$ distance could also be defined assessing the edition required by the observed plan execution to match the given model.
-This implies defining {\em edit operations} that modify $\mathcal{T}$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
-Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$.
-The reason is that in practice, the number of {\em variable objects} is normally smaller than the number of objects in the observed states.
 
@@
 According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$, provided the observation $\mathcal{O}$, is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, the hypotheses are about the set of possible \strips\ action models $\mathcal{M}\in M$ while the given observation is a partially observed plan execution $\mathcal{T}$.
+In our scenario, hypotheses are about the comparable \strips\ models $\mathcal{M}\in M$ while the given observation is the partially observed plan execution $\mathcal{T}$.
 
-We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\mathcal{T})$, to the probability distribution of the comparable \strips\ models (the definable models within the $F_v(\xi)$ sets) provided that the plan execution $\mathcal{T}$ is observed.
+We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\mathcal{T})$, to the probability distribution of models $\mathcal{M}\in M$ provided that the plan execution $\mathcal{T}$ is observed.
 This probability distribution can be estimated in three-steps:
 \begin{enumerate}
-\item Estimating the {\em a priori} probabilities $P(\mathcal{T})$ and $P(\mathcal{H})$.
-For instance, given the set of predicates $\Psi$ and the given a set of operator headers (in other words, given the $F_v(\xi)$ sets) the size of the set of possible \strips\ models set is $\prod_\xi 2^{2|F_v(\xi)|}$.
-If we assume that a priori all models are equiprobable this means that $P(\mathcal{M})=\frac{1}{\prod_\xi 2^{2|F_v(\xi)|}}$.
-With respect to the observations, given the sets $F$ and $A$, the number of possible state observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|F|}$ while the number of possible action observations of length $n$ in a trace $\mathcal{T}$ is $2^{n\times|A|}$.
-If we assume that a priori all traces are equiprobable, this means $P(\mathcal{T})=\frac{1}{2^{2n\times|F|\times |A|}}$.
+\item Estimating the {\em a priori} probabilities $P(\mathcal{T})$ and $P(\mathcal{M})$.
+For instance given the $F_v(\xi)$ sets, if we assume that a priori all models are equiprobable $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2|F_v(\xi)|}}$.
+With respect to the observation, if we assume that all the observations of plan executions with a maximum of $n$ steps are equiprobable, $P(\mathcal{T})=\frac{1}{2^{2n\times|F|\times |A|}}$.
 \item Estimating the conditional probability $P(\mathcal{T}|\mathcal{M})$.
-We can compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ for every possible model $\mathcal{M}\in M$ and map  this distance into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
-\item Applying the Bayes rule to obtain the normalized posterior probabilities since these probabilities must sum 1.
+For every possible model $\mathcal{M}\in M$, we can compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ and map it into a $[0,1]$ likelihood with the expression, $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
+\item Applying the Bayes rule to obtain the normalized posterior probabilities (these probabilities must sum 1).
 \end{enumerate}
@@
 A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}.
+\item {\bf Validates the edited model $\mathcal{M}'$ in the observation $\mathcal{T}$}.
 The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$.
@@
 
-Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-The edited action model is validated at a four action plan for inverting a two-block tower where states $s_0$ and $s_4$ are observed while the intermediate states $s_1$, $s_2$ and $s_3$ are unobserved.
+Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+The edited action model is validated at the observation of a four action plan for inverting a two-block tower where: states $s_0$ and $s_4$ are fully observed while the intermediate states, $s_1$, $s_2$ and $s_3$, are unobserved.
 \begin{figure}
@@
 }
- \caption{\small Plan for editing (steps [0-1]) and validating (steps [2-6]) a given {\em blockswold} model.}
+ \caption{\small Plan for editing (steps [0-1]) and validating (steps [2-6]) a given \strips\ planning model for the {\em blockswold}.}
 \label{fig:plan-pdistance}
@@
 
-Our interest is not in the resulting action model from the compilation $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deleitions) for that $\mathcal{M}'$ is validated in the given observations, e.g.
+Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deleitions) required by $\mathcal{M}'$ to be validated in the given observation, e.g.
 $\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}.
@@
 \subsection{Conditional effects}
-Conditional effects allow us to compactly define the actions output by our compilation.
+Conditional effects allow us to compactly define our compilation.
 An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
@@
 \subsection{The compilation formalization}
-
-Given $\tup{\mathcal{M},\Psi,\mathcal{T}}$ s.t.,
-\begin{itemize}
-\item The initial state $s_0\in\mathcal{T}$ is {\em fully observable}.
-\item Intermediate actions $a_i\in\mathcal{T}$ and states $s_i\in\mathcal{T}$ s.t.
-{\small $1\leq i\leq n$}, can be {\em partially observed}.
-This means that some fluents in $s_i$ may be missing because its value is unknown.
-In the extreme, entire states $s_i$, {\small $1\leq i\leq n$}, can be missing.
-\item $\mathcal{T}$ is {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
-\end{itemize}
-the compilation outputs a classical planning task $P=\tup{F,A,I,G}$:
+Given a \strips\ model $\mathcal{M}\in M$ and the observation of a plan execution $\mathcal{T}$, our compilation outputs a classical planning task $P=\tup{F,A,I,G}$ such that:
 \begin{itemize}
@@
 \begin{itemize}
-\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the plan trace given as input, i.e.
+\item The fluents built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in any state/action observed in $\mathcal{T}$, i.e.
 the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:plan-pdistance}.
-Formally, $\Omega=\bigcup_{\small s\in\mathcal{T}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
-\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
+\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the edited action model.
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
@@
 The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of $\mathcal{T}$.
-\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state observation $s_i\in\mathcal{T}$ where the action model is validated.
+\item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schemas are edited or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state observation $s_i\in\mathcal{T}$ where the action model is validated.
 \end{itemize}
-\item $I$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated.
-Our compilation assumes that initially, operator schemas are programmed with every possible precondition (the most specific learning hypothesis), no negative effect and no positive effect.
-Therefore fluents $pre_f(\xi)$, for every $f\in F_v(\xi)$, hold also at the initial state.
+\item $I$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{edit}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated.
+Our compilation assumes that initially $\mathcal{M}'$ is defined as $\mathcal{M}$.
+Therefore fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ hold as given by $\mathcal{M}$.
 
-\item $G=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in the state observations $s_i\in\mathcal{T}$.
-\item $A$ comprises three kinds of actions:
+\item $G=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the edited model is validated in the state observations $s_i\in\mathcal{T}$.
+\item $A$ comprises three kinds of actions with conditional effects:
 \begin{enumerate}
-\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$:
+\item Actions for {\em editing} operator schema $\xi\in\mathcal{M}$:
 \begin{itemize}
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}, pre_{f}(\xi)\},\\
+\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{edit}, pre_{f}(\xi)\},\\
 \cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
+\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{edit}\},\\
 \cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
@@
 \end{itemize}
-Besides these actions $A$ also contains the actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect.
+Besides these actions, $A$ also contains the actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect.
 
-\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
+\item Actions for {\em applying} an edited operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
 Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
@@
 &\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
+&\{mode_{edit}\}\rhd\{\neg mode_{edit}\},\\
 &\{\emptyset\}\rhd\{mode_{val}\}.
@@
 
-\begin{figure}[hbt!]
+\begin{figure}
 \begin{scriptsize}
@@
 \end{figure}
-The extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\
+When the observation $\mathcal{T}$ includes observed actions, then extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\
 
@@
 \label{sec:evaluation}
-To evaluate the empirical performance of {\em model recognition as planning} we defined a set of possible \strips\ models, each representing a different {\em Turing Machine}, but all sharing the same set of machine states and same tape alphabet.
+To evaluate the empirical performance of {\em model recognition as planning} we defined a set of possible \strips\ models, each representing a different {\em Turing Machine}, but all sharing the same set of {\em machine states} and same {\em tape alphabet}.
 
@@
 \item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}.
-For each possible pair of tape symbol and machine state, $\delta$ defines (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
+For each possible pair of tape symbol and non-terminal machine state $\delta$ defines (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
 If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
@@
 \item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to make $?x$ the tape cell pointed by the header and $?xl/?xr$ its left/right neighbours.
-Preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} are also included to capture the symbol pointed by the header and the currrent machine state.
-\item The {\bf delete effects} remove the symbol pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state according to $\delta$.
+Preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} are also included to capture the symbol currently pointed by the header and the current machine state.
+\item The {\bf delete effects} remove the symbol currently pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state according to $\delta$.
 \end{itemize}
@@
 The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} (the full encoding of the {\em Turing Machine} of Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema).
-\begin{figure}[hbt!]
+\begin{figure}
 \begin{scriptsize}
@@
 
-Since the {\em transition function} of a {\em Turing Machine} can be encoded as a classical planning frame $\Phi=\tup{F,A}$, executions of a {\em Turing Machine} are definable as {\em plan traces} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
+Since the {\em transition function} of a {\em Turing Machine} can be encoded as a classical planning frame $\Phi=\tup{F,A}$, executions of that {\em Turing Machine} are definable as {\em plan traces} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
 
 \subsection{Experimental setup}
-We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a two-symbol three-state {\em Turing Machine}.
+We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a seven-symbol six-state {\em Turing Machine}.
 We randomly choose a machine $\mathcal{M}\in M$ and produce an fifty-step execution plan trace $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_{50},s_{50}}$.
 Finally, we follow our {\em model recognition as planning} method to identify the {\em Turing Machine} that produced $\mathcal{T}$.
-This exeperiment is repeated for different amounts of missing information in the input trace $\mathcal{T}$.
+This exeperiment is repeated for different amounts of missing information in the input trace $\mathcal{T}$: unknown applied transitions, unknown internal machine state and unknown values of several tape cells.
 
@@
 {\sc Madagascar} is the classical planner we used to solve the instances that result from our compilations for its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
-Due to its SAT-based nature, {\sc Madagascar} can apply the actions for programming preconditions in a single planning step (in parallel) because there is no interaction between these actions.
-Actions for programming effects can also be applied in a single planning step, thus significantly reducing the planning horizon.
+Due to its SAT-based nature, {\sc Madagascar} can apply the actions for editing preconditions in a single planning step (in parallel) because there is no interaction between them.
+Actions for editing effects can also be applied in a single planning step, thus significantly reducing the planning horizon.
 
-The compilation source code, evaluation scripts and benchmarks (including the used training and test sets) are fully available at this anonymous repository {\em https://github.com/anonsub/observations-learning} so any experimental data reported in the paper can be reproduced.
+The compilation source code, evaluation scripts and benchmarks (including the used training and test sets) are fully available at this anonymous repository {\em } so any experimental data reported in the paper can be reproduced.
 
@@
 
+Our $\delta(\mathcal{M},\mathcal{T})$ distance could also be defined assessing the edition required by the observed plan execution to match the given model.
+This implies defining {\em edit operations} that modify $\mathcal{T}$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
+Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$.
+In practice, the number of {\em variable objects} is normally smaller than the number of objects in the observed states.
+
 The {\em \strips\ probability distribution} assigns the same probabaility value to any hypothesis model $\mathcal{M}\in M$ that can reproduce the given observation $\mathcal{T}$.
@@
 If we assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}, one can define probability distributions that consider as more probable hypothesis the ones that reproduce observations in less steps.
-%This an interesting research direction, specially for the particular case $\mathcal{T}=\tup{s_0, s_n}$ where observations are {\em input-output} examples.
-
-Recently {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} also proposes model edition but in this case the aim is to conform the different PDDL models of two different agents.
+A related approach is recently followed in {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the different PDDL models of two different agents.
 

### 2018-06-07 util/HTNML/experimenter_ARMS.py
Small changes to the pddl_parser to allow more flexibility in the inputs

@@ -1,0 +1,51 @@
+#! /usr/bin/env python
+import glob, os, sys, copy, itertools, math
+import src.pddl, src.pddl_parser
+import src.fdtask_to_pddl
+import numpy as np
+import src.model_evaluator
+import src.config
+
+# **************************************#
+# MAIN
+# **************************************#
+
+try:
+    domain_folder_name  = sys.argv[1]
+    state_observability = sys.argv[2]
+except:
+    print "Usage:"
+    print sys.argv[0] + " <domain folder> <state observability (0-1)>"
+    sys.exit(-1)
+
+domain_path = os.path.expanduser(src.config.PROJECT_PATH + "util/HTML/" + domain_folder_name)
+
+profile = "{}\n" \
+          "1\n" \
+          "{}\n" \
+          "0.1\n" \
+          "{}/Maxsat.in\n" \
+          "{}/Maxsat1.in\n" \
+          "{}/Result.out\n".format(domain_path, state_observability, domain_path, domain_path, domain_path)
+
+profile_path = os.path.expanduser(src.config.PROJECT_PATH + "util/HTNML/Profile")
+with open(profile_path, "w") as f:
+    f.write(profile)
+
+
+
+cmd = src.config.PROJECT_PATH + "util/HTNML/HTNML"
+# print("\n\nExecuting... " + cmd)
+os.system(cmd)
+
+try:
+    domain_name, best_evaluation, best_matches = src.model_evaluator.evaluate("learned_domain.pddl", domain_folder_name + "ref_domain", True)
+    print(" & ".join(
+    [domain_name] + [str(round(e, 2)) for e in best_evaluation]) + " \\\\" + " % {}".format(
+    best_matches))
+except:
+    print("No solution found")
+
+
+
+sys.exit(0)

### 2018-06-07 src/compiler_new.py
model_evaluator.py: returned domain name corresponds to the reference domain

@@ -36,6 +36,76 @@ def possible_pred_for_action(task, p, a, tup):
     fits = [len(action_types[i].intersection(predicate_types[i])) >= 1 for i in range(len(action_types))]
     return all(fits)
 
+def get_static_predicates(traces, predicates):
+
+    candidates = set([p.name for p in predicates])
+
+    for trace in traces:
+        trace_candidates = set()
+        for predicate in candidates:
+            static = True
+            init_literals = set([l for l in trace.init if l.predicate == predicate])
+            for state in trace.states:
+                state_literals = set([l for l in state if l.predicate == predicate])
+
+                if len(init_literals.intersection(state_literals)) != len(state_literals):
+                    static = False
+                    break
+
+            if static:
+                trace_candidates.add(predicate)
+
+        candidates = candidates.intersection(trace_candidates)
+
+    reflexive_static_predicates = dict()
+    for candidate in candidates:
+        reflexive_static_predicates[candidate] = True
+        for trace in traces:
+            init_literals = set([l for l in trace.init if l.predicate == candidate])
+            for literal in init_literals:
+                if len(literal.args) == 1 or len(set(literal.args)) != 1:
+                    reflexive_static_predicates[candidate] = False
+                    break
+
+    return [p for p in predicates if p.name in candidates], reflexive_static_predicates
+
+
+def get_static_precondition(predicate, action, plans, tasks):
+    static_preconditions = set()
+    params = [pddl.pddl_types.TypedObject("?o" + str(i), action[i]) for i in range(1, len(action))]
+    params = [x for x in params if x.type_name in predicate[1:]]
+    num_predicate_params = len(predicate[1:])
+    possible_param_tuples = list(itertools.combinations(params, num_predicate_params))
+    for t in possible_param_tuples:
+        static_preconditions.add(pddl.conditions.Atom(predicate[0], [x.name for x in t]))
+        static_preconditions.add(pddl.conditions.Atom(predicate[0], [x.name for x in reversed(t)]))
+
+    if len([x for x in action[1:] if x in predicate[1:]]) >= num_predicate_params:
+        all_instances = set()
+        for task in tasks:
+            all_instances.update([p.args for p in task.init if p.predicate == predicate[0]])
+
+        all_variables = set(sum(all_instances, ()))
+
+        for a in [item for sublist in plans for item in sublist]:
[…]

### 2018-06-07 src/model_evaluator.py

@@ -91,6 +91,7 @@ def evaluate(evaluation_domain_filename, reference_domain_filename, reformulatio
     domain_name, domain_requirements, types, type_dict, constants, predicates, predicate_dict, functions, actions, axioms \
                      = pddl_parser.parsing_functions.parse_domain_pddl(ref_domain_pddl)
     ref_actions = actions
+    name = domain_name
 
     # Creating a FD task with the domain to evaluate and the aux problem file
     eva_domain_pddl = pddl_parser.pddl_file.parse_pddl_file("domain", evaluation_domain_filename)
@@ -98,8 +99,8 @@ def evaluate(evaluation_domain_filename, reference_domain_filename, reformulatio
                      = pddl_parser.parsing_functions.parse_domain_pddl(eva_domain_pddl)
     eva_actions = actions
 
-    known_actions = list()
 
+    known_actions = list()
     if partial_domain_filename:
         # Creating a FD task with the partial domain and the aux problem file
         partial_domain_pddl = pddl_parser.pddl_file.parse_pddl_file("domain", partial_domain_filename)
@@ -150,7 +151,7 @@ def evaluate(evaluation_domain_filename, reference_domain_filename, reformulatio
             proper_actions = [tuple([x]+[i for i in range(1,arity+1)]) for x in actions_name_by_arity[i][1]]
 
             combinations = [zip(proper_actions, x) for x in itertools.permutations(actions, len(proper_actions))]
-            # combinations = [zip(x, actions) for x in itertools.product(proper_actions, actions)]
+
             if len(proper_actions) > 1:
                 for comb in combinations:
                     if not valid_action_combination(comb, action_params_dict):
@@ -198,7 +199,7 @@ def evaluate(evaluation_domain_filename, reference_domain_filename, reformulatio
             best_matches = matches
         # print(f1_score, matches)
 
-    return domain_name, best_evaluation, best_matches
+    return name, best_evaluation, best_matches
 
 
 

### 2018-06-07 doc/aij18/aij18.tex
Add files via upload
Background updated
background and motivation

--- 2018-05-02 doc/aij18/aij18.tex
@@
 \newcommand{\CAMA}{{\small {\sffamily CAMA}}\xspace}
+\newcommand{\SLAF}{{\small {\sffamily SLAF}}\xspace}
+\newcommand{\LAMP}{{\small {\sffamily LAMP}}\xspace}
+\newcommand{\NOISTA}{{\small {\sffamily NOISTA}}\xspace}
+\newcommand{\LOCM}{{\small {\sffamily LOCM}}\xspace}
+\newcommand{\LOCMtwo}{{\small {\sffamily LOCM2}}\xspace}
+\newcommand{\LOP}{{\small {\sffamily LOP}}\xspace}
+\newcommand{\AMAN}{{\small {\sffamily AMAN}}\xspace}
+\newcommand{\FAMA}{{\small {\sffamily FAMA}}\xspace}
+
+\newcommand{\FO}{{\small {\sffamily FO}}\xspace}
+\newcommand{\PO}{{\small {\sffamily PO}}\xspace}
+\newcommand{\POstar}{{\small {\sffamily PO*}}\xspace}
+\newcommand{\NO}{{\small {\sffamily NO}}\xspace}
 
@@
 The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
-What is more, the compilation is extensible to assess how well a given \strips\ action model matches the observation of a plan execution.
-This extension allows us to evaluate the quality of the learned models with respect to the actual models but also, with respect to observations of plan executions that act as a {\em test set}, which is useful when the actual models are not available.
-The performance of our compilation approach is evaluated learning action models for a wide range of classical planning domains from the International Planning Competition (IPC) and using these two evaluation approaches.
+On the other hand, the compilation is extensible to assess how well a given \strips\ action model matches the observation of a plan execution.
+This extension allows us to evaluate the quality of the learned models with respect to the actual models but also, with respect to a {\em test set} of observations of plan executions.
+The performance of our compilation approach is evaluated learning action models, for a wide range of classical planning domains from the International Planning Competition (IPC), and following these two evaluation approaches.
 \end{abstract}
@@
 \input{background}
+\input{motivation}
 \input{learning_task}

### 2018-06-07 doc/aij18/background.tex

@@
 
-In this section we summarize the most recent and relevant approaches to learning action models found in the literature.
-Approaches will be examined according to the following parameters: the input knowledge (plan traces) accepted by the system, the expressiveness of the learned action model and the principal technique used for learning the action model (Table \ref{table:models_comparison1}), as well as the characteristics of the evaluation method to validate the learned models (Table \ref{table:models_comparison2}).
+Here we summarize the most recent and relevant approaches to learning action models that is found in the AI planning literature.
+These approaches will be examined according to the following parameters: the {\em input knowledge} that is accepted by the approach (i.e.
+the information observed in the given plan traces), the {\em expressiveness} of the learned action models and, the principal {\em technique} that is used for learning the action models (Table \ref{table:models_comparison1}).
+In addition, we examine the characteristics of the {\em evaluation} methods that are used to assess the quality of the learned models (Table \ref{table:models_comparison2}).
 
-The first column of Table \ref{table:models_comparison1} shows that all the approaches analyzed in this work accept partial observations of intermediate states, and some approaches also allow the sequence of intermediate states to be empty.
+The first column of Table \ref{table:models_comparison1} shows that all the approaches analyzed in this work accept partial observations of intermediate states, and that some approaches also allow the sequence of intermediate states to be empty.
 Although not explicitly stated in the corresponding published works, it seems reasonable to assume that handling \NO state sequences subsumes \POstar trajectories and that a \POstar plan trace can be regarded as composed of several \NO sub-traces.
@@
 Most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces.
-Others do not explicitly say so but the fact is that the predicates and the actions headers are easily extractable from the state sequence and action sequence of the plan traces, respectively.
-A requirement, however, for these two sets to be representative is that set of input plan traces comprises a grounded sample of all predicates and operator schemas of the domain model.
+Others do not explicitly say so but the fact is that the predicates and the actions headers are easily extractable from the state sequence and action sequence of the given plan traces, respectively.
+A requirement, however, for these two sets to be representative is that the set of input plan traces comprises a grounded sample of all predicates and operator schemas of the domain model.
 
 The expressiveness of the learned action models varies across approaches (second column of Table \ref{table:models_comparison1}).
-All the presented systems are able to learn action models in a \textsc{Strips}  representation \cite{fikes1971strips} and some propose algorithms to learn more expressive action models that include quantifiers, logical implications or the type hierarchy of the domain.
+All the presented systems are able to learn action models in a \textsc{Strips}  representation \cite{fikes1971strips} and some propose algorithms to learn more expressive action models that include {\em quantifiers}, {\em logical implications} or the {\em type hierarchy} of the domain.
 
-Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- almost all approaches rely on a comparison between the learned model and a ground-truth model), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
+Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- almost all approaches rely on a syntactic comparison between the learned model and a ground-truth model), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
 
@@
 These constraints are then used to build and solve a weighted propositional satisfiability problem with a MAX-SAT solver.
-Three types of constraints are considered: 1) constraints imposed by general axioms of correct \textsc{Strips} actions, 2) constraints extracted from the distribution of actions in the plan traces and 3) constraints obtained from the \PO states, if available.
+Three types of constraints are considered: 1) constraints imposed by general syntactic axioms of correct \textsc{Strips} actions, 2) constraints extracted from the distribution of actions in the plan traces and 3) constraints obtained from the \PO states, if available.
 Frequent subsets of actions in which to apply the two latter types of constraints are found by means of frequent set mining.
@@
 \ARMS defines an error metric and a redundancy metric to measure the correctness and conciseness of an action model over the test set of input plan traces using a cross-validation evaluation.
-The model evaluation is posed as an optimization task that returns the model that best explains the input traces by minimizing the error and redundancy functions.
+Learning is posed as an optimization task that returns the model that best explains the input traces by minimizing the error and redundancy functions.
 This yields a model that is approximately correct (100\% correctness is not required so as to ensure generality and avoid overfitting), approximately concise (low redundancy rates), and that can explain as many examples as possible.
@@
 The \ARMS system became a benchmark in action-model learning, showing empirically that is is feasible lo learn a model in a reasonably efficient way using a weighted MAX-SAT even with \NO state trajectories.
-Nevertheless, we argue that using the home-made error and redundancy metrics to evaluate the model correctness is questionable since a model that exhibits a wrong precondition or effect in any of its actions could still be eligible as the minimal-error output model.
+Nevertheless, we argue that using the home-made error and redundancy metrics to evaluate the model correctness is questionable since models that exhibit a wrong preconditions or effects in any of its actions could still be eligible as the minimal-error output model.
 
@@
 	\begin{tabular}{ l | c | c | c }
-		& \multicolumn{1}{c}{Input plan traces}
-        & \multicolumn{1}{c}{Learned action model}
-        & \multicolumn{1}{c}{Technique}     \\
+        & \multicolumn{1}{c|}{\bf Input plan traces}
+        & \multicolumn{1}{c|}{\bf Learned action model}
+        & \multicolumn{1}{c}{\bf Technique}     \\
 		\hline			
@@
          \hline
-		\FAMA &  \PO, \PO* and \NO states & \strips &   compilation to planning\\ & \PO and \NO actions & & \\
-         \hline
+		\FAMA &  \PO, \PO* and \NO states & \strips &   compilation to planning\\ & \PO and \NO actions & & 
 	\end{tabular}
@@
 
-A tractable and exact solution of action models in partially observable domains using a technique known as Simultaneous Learning and Filtering (\textbf{\SLAF}) is presented in~\cite{AmirC08}.
+Simultaneous Learning and Filtering (\textbf{\SLAF})~\cite{AmirC08} is a tractable and exact method for learning action models in partially observable domains.
 \SLAF alongside \ARMS can be considered another of the precursors of the modern algorithms for action-model learning, able to learn from plan traces with PO and non-null intermediate states.
@@
 
-The Learning Action Models from Plan Traces (\textbf{\LAMP}) \cite{ZhuoYHL10} algorithm extends the expressiveness to learning models with universal and existential quantifiers as well as logical implications.
+The Learning Action Models from Plan Traces (\textbf{\LAMP}) \cite{ZhuoYHL10} algorithm extends the expressiveness of \ARMS to learning models with {\em universal} and {\em existential} quantifiers as well as {\em logical implications}.
 The input to \LAMP is a set of plan traces with intermediate states, which are encoded by the algorithm into propositional formulas.
@@
 	 \LOCMtwo & manual checking wrt GTM &  ---  &   --- \\
+		 &  &    &   \\         
          \hline
-		\FAMA & checking wrt GTM  & precision and recall of & 15\\ & & \#$\pre$ and \#$\eff$ & 25 actions\\
-         \hline
+		\FAMA & checking wrt GTM  & precision and recall of & 15\\ & & \#$\pre$ and \#$\eff$ & 25 actions
 	\end{tabular}

### 2018-06-07 doc/programming-project/AI-programming.tex

@@
 
-The research group for this project comprises: 1 principal investigator, 1 senior researcher and 1 PhD student. All of them are GRPS members.
+The research group for this project comprises three researchers: 1 principal investigator, 1 senior researcher and 1 PhD student. All of them are GRPS members.
 
@@
 \subsection{Senior Researcher: Eva Onaindia}
-
 Eva Onaindia is a {\em Full Professor} at the Universitat Politècnica of València.
@@
 
-She has led several National Research Projects as well as served on multiple ICAPS,AAAI, IJCAI and ECAI conferences as a member of the Program Committee and as a senior member.
+She has led several projects from the {\em National research plan} as well as served on multiple ICAPS,AAAI, IJCAI and ECAI conferences as a member of the Program Committee and as a senior member.
 Eva is Editor-in-Chief of the journal AI Communications, a Review Board member of the journal Applied Intelligence and she will be Program co-chair of ICAPS 2019 (the main international conference on AI planning).
 
-Eva Onaindía has supervised so far a total of 9 PhD theses and she is currently supervising four doctoral students.
-She has also participated in many diverse scientific evaluation committees at national and international level.
-In the national scene, she has collaborated with National Agencies such as ANEP and ANECA and she has been member of different expert boards and scientific committees.
+Eva Onaindía has supervised so far a total of 9 PhD theses and she currently supervising four doctoral students.
+She has also participated in many diverse scientific evaluation committees at both national and international level.
+In the national scene, she has collaborated with National agencies ANEP and ANECA and she has been member of different expert boards and scientific committees.
 At an international level, she has participated in the evaluation of research projects for Israel, Italy or Czech Republic.
 
-Eva has also participated in several R\&D and Innovation contracts: with the firm DESTINIA S.L.
-for the development of a tourist recommender system, and currently in the construction of a system for Electricity Price Forecasting aimed at energy savings (Ayuntamiento Lliria) using neural networks and data analysis techniques.
+Eva has also participated in several R\&D and Innovation contracts: with the firm {\sc DESTINIA S.L.} ({\tt https://destinia.com/}) for the development of a tourist recommender system, and currently in the construction of a system for {\em electricity price forecasting} aimed at energy savings (Ayuntamiento Lliria) using neural networks and data analysis techniques.
 
@@
 \subsection{Researcher in training period: Diego Aineto}
+Diego Aineto is a PhD student at the {\em Department of Computer Systems and Computation} of the UPV.
+
+Since 2017 Diego is holding a four-year FPU grant on {\em Causal discovery for predictive analysis}.
+Diego is also involved on the current research project of the GPRS from the Spanish national research plan on {\em activity recognition}.
+
+With Eva Onaindia and Sergio Jiménez, Diego is already co-author of one paper on the International Conference on Planning and Scheduling (ICAPS)~\cite{diego-icaps18}, and of one journal paper on research topics that are relevant to the development of the project~\cite{onaindia2018common}.
 
@@
 
-Now we review the two most successful approaches for program synthesis:
+Now we review the two most successful approaches for program synthesis. The research posed in this project falls into the first approach:
 \begin{itemize}
@@
 The programs are computed with a domain-specific search that implements a divide and conquer approach.
-The approach posed in this research project falls into this category.
 
@@
 
-Briefly the state variables of this AIP task are {\em fluents} of the kind:
+Briefly the {\bf state variables} of this AIP task are {\em fluents} of the kind:
 \begin{itemize}
-\item {\tt var:=value}, that encode the program variables).
+\item {\tt var:=value}, that encode the program variables.
 \item {\tt program(line):=instruction}, that encode the program lines.
@@
 
-The initial/goal states of the AIP task encode the initial/final values of the program variables according to the input/output tests of the TDD programming task.
+The {\bf initial/goal states} of the AIP task encode the initial/final values of the program variables according to the input/output tests of the TDD programming task.
 
-Finally, program instructions are encoded using two kinds of AIP actions:
+Finally, program instructions are encoded using two kinds of AIP {\bf actions}:
 \begin{itemize}
@@
 
-Interestingly, our PDDL encoding allows also program validation by (1), specifying the lines of the program to validate in the initial state of the AIP task and (2), disabling the mentioned {\it programming actions}.
+Interestingly, our PDDL encoding allows also program validation by (1), specifying the lines of the program to validate in the initial state of the AIP task and (2), disabling the mentioned {\it programming actions} so only {\it execution actions} are applicable.
 
@@
 \begin{itemize}
-\item {\em Theoretical benchmarks}: Programming tasks for the computation of mathematical/logic series, string manipulation and for the management of classic data structures such as {\em lists}, {\em queues}, {\em stacks}, {\em trees}, {\em heaps}, \ldots
-\item {\em Real-world benchmarks}: Practical programming tasks coming from {\em activity recognition} and {\em recommendation} domains.
-The GRPS group is currently leading a four-year research project from the Spanish National plan in which {\em activity recognition} and {\em activity recommendation} are applied to different real-world domains such as {\em tourism}, {\em domotics}, {\em traffic control} and {\em robotics}.
-We plan to evaluate our approach in programming and validation tasks that result from these domains.
+\item {\em Real-world benchmarks}: The GRPS group is currently leading a four-year research project from the {\em Spanish national plan} in which {\em activity recognition} is applied to different real-world domains such as {\em domotics}, {\em tourism}, {\em traffic control} and {\em robotics}.
+We plan to evaluate our approach in synthesis and validation tasks coming from these real-world domains.
+Figure~\ref{fig:activity} shows a four-line program (pictured as a finite state machine) that represents the sequence of instructions required for the {\em making a buttered toast} activity.
+
+\begin{figure}[hbt!]
+\begin{center}
+\begin{tikzpicture}
+\node [thick,draw,circle] at (1,1.1) (x) {$q_0$};
+\node [thick,draw,circle] at (4,1.1) (y) {$q_1$};
+\node [thick,draw,circle] at (7,1.1) (v) {$q_2$};
+\node [thick,draw,circle] at (10,1.1) (w) {$q_3$};
+\node [thick,draw,circle] at (13,1.1) (z) {$q_3$};
+\draw (13,1.1) circle (.3cm);
+
+\draw [thick,->] (x) to (y);
+\draw [thick,->] (y) to (v);
+\draw [thick,->] (v) to (w);
+\draw [thick,->] (w) to (z);
+
+\node at (2.4,1.5) {$\mathsf{take}(bread)$};
+\node at (5.4,1.5) {$\mathsf{use}(toaster)$};
+\node at (8.4,1.5) {$\mathsf{take}(knife)$};
+\node at (11.4,1.5) {$\mathsf{spread}(butter)$};
+
+\end{tikzpicture}
+\end{center}
+\caption{\small Four-line program representing the activity of {\em making a buttered toast}.}
+\label{fig:activity}
+\end{figure}
+
+\item {\em Theoretical benchmarks}: Programming tasks for the computation of mathematical/logic series, string manipulation and for the management of classic data structures such as {\em lists}, {\em queues}, {\em stacks} or {\em trees}.
+Figure~\ref{fig:list} shows a synthesized program for traversing linked lists.
+The program is pictured as a {\em finite state machine}: The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
 \end{itemize}
+
+\begin{figure}[hbt!]
+\begin{center}
+\begin{tikzpicture}
+\node [thick,draw,circle] at (1,1.1) (x) {$q_0$};
+
+\node [thick,draw,circle] at (6,1.1) (y) {$q_1$};
+
+\node [thick,draw,circle] at (11,1.1) (z) {$q_2$};
+
+\draw (11,1.1) circle (.3cm);
+
+\draw [thick,->] (x) -- (y);
+
+\draw [thick,out=345,in=195,->] (x) to (z);
+
+\draw [thick,out=165,in=15,->] (y) to (x);
+
+\node at (4.1,0.85) {$!\mathsf{isEnd}(n)/\mathsf{visit}(n)$};
+\node at (3.5,1.8) {$-/\mathsf{step}(n)$};
+\node at (6,0) {$\mathsf{isEnd}(n)/a_\bot$};
+\end{tikzpicture}
+\end{center}
+\caption{\small Synthesized program to solve the programming task of traversing a linked list ($n$ is a variable that points to a node of the linked list).}
+\label{fig:list}
+\end{figure}
 
@@
 We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that (1), takes as input a set of test cases that specify the TDD programming task to solve and (2), outputs a program source code that passes these tests with a bug-free guarantee.
+
+In the particular case that the {\em program synthesizer} receives an additional input specifying the program source code, it outputs a validation certificate guaranteeing that the input program passes the given test cases.
 Figure~\ref{fig:gantt} details the proposed 24-month timeline for the project.
@@
     \item Design of the user-interaction for the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases plus the available instruction set. 
-    \item Experimental design. Experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve the given TDD programming tasks.
+    \item Experimental design. Experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve the given TDD programming/validation tasks.
       \item Evaluation of different AI planners available, with special attention to the planners that get the best results at the IPC-2018. 
@@
       \begin{enumerate}
-      \item Reporting the experimental performance of our AIP approach for solving TDD programming tasks.
+      \item Reporting the experimental performance of our AIP approach for solving diverse TDD programming tasks.
       \item Dissemination of the obtained theoretical and empirical results by submitting papers to top international conferences and journals in AI.        
@@
 \end{small}        
-{\small{\bf\em  Deliverable:} Final report with the obtained conclusions and produced publications.}
+{\small{\bf\em  Deliverable:} Final report with obtained conclusions and produced publications.}
 \end{enumerate}
 
-Tasks ({\bf T1-T3}) will be developed by the three mentioned members of the research group: Sergio Jiménez, Eva Onaindia and Diego Aineto.
-In addition, we plan to hire a master student to assist in the completion of task {\bf T3(a)}.
+{\em Tasks ({\bf T1-T3})} will be developed by the three mentioned members of the research group: Sergio Jiménez, Eva Onaindia and Diego Aineto.
+In addition, we plan to hire a master student to assist in the completion of {\em task {\bf T3(a)}}.
 
@@
 
-Figure~\ref{fig:list} shows a synthesized program for traversing linked lists.
-The program is pictured as a {\em finite state machine}: The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
+The objective of this project is to study the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions:
+\begin{enumerate}
+\item Number of {\em program lines}.
+\item Number of observable {\em program variables}.
+\item Size of the available {\em instruction set}.
+\end{enumerate}
 
-\begin{figure}[hbt!]
-\begin{center}
-\begin{tikzpicture}
-\node [thick,draw,circle] at (1,1.1) (x) {$q_0$};
-
-\node [thick,draw,circle] at (6,1.1) (y) {$q_1$};
-
-\node [thick,draw,circle] at (11,1.1) (z) {$q_2$};
-
-\draw (11,1.1) circle (.3cm);
-
-\draw [thick,->] (x) -- (y);
-
-\draw [thick,out=345,in=195,->] (x) to (z);
-
-\draw [thick,out=165,in=15,->] (y) to (x);
-
-\node at (4.1,0.85) {$!\mathsf{isEnd}(n)/\mathsf{visit}(n)$};
-\node at (3.5,1.8) {$-/\mathsf{step}(n)$};
-\node at (6,0) {$\mathsf{isEnd}(n)/a_\bot$};
-\end{tikzpicture}
-\end{center}
-\caption{\small Synthesized program to solve the programming task of traversing a linked list ($n$ is a variable that points to a node of the linked list).}
-\label{fig:list}
-\end{figure}
-
-The objective of this project is to study the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions: number of {\em program lines}, number of {\em program variables} and size of the available {\em instruction set}.
-For instance, the program of Figure~\ref{fig:list} is generated using three program lines, one observable variable $\mathsf{isEnd}(n)$ with Boolean domain, and a instruction set that comprises two instructions ($\mathsf{visit}(n)$ that marks the list node assigned to $n$ as {\em visited} and $\mathsf{step}(n)$ that advances $n$ to the next node in the linked list).
+For instance, the program of Figure~\ref{fig:list} is generated using three program lines, one observable variable ($\mathsf{isEnd}(n)$ that has Boolean domain and indicates when $n$ points to the end of the linked list), and a instruction set that comprises two instructions ($\mathsf{visit}(n)$ that marks the list node assigned to $n$ as {\em visited} and $\mathsf{step}(n)$ that advances $n$ to the next node in the linked list).
 
@@
 \begin{itemize}
-\item To study the performance of our Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
-\item To study the performance of our Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 20 variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
+\item To study the performance of our Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
+\item To study the performance of our Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}.
 \end{itemize}
@@
 AIP has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}.
-Promising research opportunities come from the application of AIP to {\em program synthesis} given that, {\em program synthesis} with a tests base, can also be seen as a dual to {\em program testing}.
+Promising research opportunities come from the application of AIP to {\em program synthesis} given that, {\em program synthesis} with a tests base, can be seen as the {\em program testing} dual.
 
-This project will provide new insights into the current understanding of how Artificial Intelligence can assist programmers in the software development.
+In fact, our current work on {\em program synthesis} with AIP already produced {\bf publications at top international conferences on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
+
+In more detail, the expected benefits of this research project are four-fold:
+\begin{enumerate}
+\item The definition of {\bf effective metrics for assessing how well a given program covers a set of {\em input-output} test cases}.
+Our belief is that these metrics can provide new insights into the current understanding of how Artificial Intelligence can assist programmers in the software development.
+\item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}.
 Research in AI algorithms is too often tested with laboratory problems and AIP is not an exception.
@@
 This project will also help to meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks.
-
-The expected benefits of this project are three-fold:
-\begin{enumerate}
-\item The definition of {\bf effective metrics for assessing how well a given program covers a set of {\em input-output} test cases}.
-\item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}.
-\item The development, evaluation and dissemination of {\bf open source software for the {\em Synthesis and Validation of Programs}}.
+\item The {\bf development and evaluation of software and benchmarks for the {\em Synthesis and Validation of Programs}}.
+We strongly belief that reproducibility and open knowledge are essential to the advance of the research on computer science.
+With this regard we plan to develop a {\em github} repository where we make available the developed source code and benchmarks.
+\item The {\bf dissemination of the obtained scientific results}.
+The scientific results obtained during the development of the project will also be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).
 \end{enumerate}
 
-Our current work on {\em program synthesis} with AIP already produced {\bf publications at top international conferences on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
-
-The scientific results obtained during the development of the project will also be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).
-
-\vspace{0.5cm}
+\vspace{0.3cm}
 

### 2018-06-07 doc/programming-project/AI-programming.tex
updates added to upv research project

@@
 \author{Sergio Jim\'enez Celorrio\\
-\footnotesize Universitat Politécnica de València, Spain\\
+\footnotesize Universitat Politècnica de València, Spain\\
 \small Area: \underline{Ciencias de la Computación y Tecnología Informática}}
@@
 \begin{abstract}
-Programming errors, commonly known as {\em bugs}, cause undesired software behavior making a program crash or enabling malicious users to access private data.
+  Programming is still a manual handcraft task, even when the complexity of programs is not high enough to justify the cost of the human workforce to produce them.
+The need to automate programming tasks increases every day: Programming errors, commonly known as {\em bugs}, cause undesired software behavior making programs crash or enabling malicious users to access private data.
 Just for 2017, the cumulative cost of software bugs is worldwide estimated in more than one trillion US dollars.
 
-This research project aims to investigate a novel approach for software development, that leverages Artificial Intelligence, to reduce the chances of introducing programming errors.
+  This research project investigates a novel approach for software development, that leverages Artificial Intelligence, to increase the automation of the programming process and hence, reduce the chances of introducing software bugs.
 The project proposes to address {\bf automatic program synthesis and validation starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine}.
@@
 
-The project will be developed within the {\em Group of Reasoning on Planning and Scheduling} (GRPS, {\tt http://users.dsic.upv.es/grupos/grps/}), that is part of the {\em Department of Computer Systems and Computation} of the UPV.
-In the last 10 years, the GRPS has led 3 research projects of the Spanish national research plan on {\em AI planning}.
+The project will be developed within the {\em Group of Reasoning on Planning and Scheduling} (GRPS, {\tt http://users.dsic.upv.es/grupos/grps/}), that is part of the {\em Department of Computer Systems and Computation} of the UPV (DSIC, {\tt www.dsic.upv.es}).
+
+In the last 10 years, the GRPS has led three research projects of the Spanish national research plan on {\em AI planning}.
 Further, the GRPS has developed its own AI planner, that is participating  at the International Planning Competition (IPC-2018, {\tt https://ipc2018.bitbucket.io}).
@@
 From 2004 to 2013 Sergio was assistant lecturer at the Universidad Carlos III de Madrid where he obtained the {\em Distinguished Thesis Award 2011}.
-Sergio is co-organizer of the {\em $7^{th}$ International Planning Competition} and program committee at top international conferences on Artificial Intelligence.
+Sergio is co-organizer of the {\em $7^{th}$ International Planning Competition} and program committee at top international conferences on Artificial Intelligence (including ICAPS that is the main international conference on {\em AI planning}).
 
@@
 \begin{enumerate}
-\item Inductive programming with classical planning. {\footnotesize~\cite{javi-ijcai17},~\cite{segovia2016generalized},~\cite{jimenez2015computing}}.
-\item Generalization in AI planning. {\footnotesize~\cite{javi-icaps17},~\cite{damir-derived-ijcai16},~\cite{javi-fsc-ijcai16}}.
-\item Learning to relieve the {\it knowledge acquisition bottleneck} in AI planning.
-{\footnotesize{~\cite{diego-icaps18},~\cite{jimenez2013integrating},~\cite{jimenez2008architecture},~\cite{jimenez2006planning},~\cite{lanchas2007learning}}}.
-\item Machine learning for addressing the {\it curse of dimensionality} in AI planning. {\footnotesize~\cite{jimenez2012review},~\cite{de2011scaling},~\cite{de2008learning}}.  
-\item Evaluation in AI planning. {\footnotesize~\cite{lopez2015deterministic},~\cite{lopez2013automating},~\cite{coles2012survey}}. 
+\item Inductive programming with AI planning{\footnotesize~\cite{javi-ijcai17},~\cite{segovia2016generalized},~\cite{jimenez2015computing}}.
+\item Generalization in AI planning{\footnotesize~\cite{javi-icaps17},~\cite{damir-derived-ijcai16},~\cite{javi-fsc-ijcai16}}.
+\item Learning to relieve the {\it knowledge acquisition bottleneck} in AI planning{\footnotesize{~\cite{diego-icaps18},~\cite{jimenez2013integrating},~\cite{jimenez2008architecture},~\cite{jimenez2006planning},~\cite{lanchas2007learning}}}.
+\item Machine learning for addressing the {\it curse of dimensionality} in AI planning{\footnotesize~\cite{jimenez2012review},~\cite{de2011scaling},~\cite{de2008learning}}.  
+\item Evaluation in AI planning{\footnotesize~\cite{lopez2015deterministic},~\cite{lopez2013automating},~\cite{coles2012survey}}. 
 \end{enumerate}
 
-\underline{General quality indicators of scientific research}: {\bf 373 citations} (287 since 2013) and {\bf h-index 11}. {\scriptsize\em Source google scholar}.
+\underline{General quality indicators of scientific research}: {\bf 381 citations} (295 since 2013) and {\bf h-index 11}. {\scriptsize\em Source google scholar}.
 
@@
 
-She has led several projects from the {\em National research plan} as well as served on multiple ICAPS,AAAI, IJCAI and ECAI conferences as a member of the Program Committee and as a senior member.
+She has led several projects from the {\em National research plan} as well as served on multiple ICAPS, AAAI, IJCAI and ECAI conferences as a member of the Program Committee and as a senior member.
 Eva is Editor-in-Chief of the journal AI Communications, a Review Board member of the journal Applied Intelligence and she will be Program co-chair of ICAPS 2019 (the main international conference on AI planning).
 
-Eva Onaindía has supervised so far a total of 9 PhD theses and she currently supervising four doctoral students.
+Eva Onaindía has supervised so far a total of nine PhD theses and she currently supervising four doctoral students.
 She has also participated in many diverse scientific evaluation committees at both national and international level.
@@
 
-\underline{General quality indicators of scientific research}: {\bf 1521 citations} (977 since 2013) and {\bf h-index 22}. {\scriptsize\em Source google scholar}.
+\underline{General quality indicators of scientific research}: {\bf 1532 citations} (989 since 2013) and {\bf h-index 22}. {\scriptsize\em Source google scholar}.
 
@@
 
-Since 2017 Diego is holding a four-year FPU grant on {\em Causal discovery for predictive analysis}.
+Since 2016 Diego is holding a four-year FPU grant on {\em Causal discovery for predictive analysis} (FPU16/03184).
 Diego is also involved on the current research project of the GPRS from the Spanish national research plan on {\em activity recognition}.
 
-With Eva Onaindia and Sergio Jiménez, Diego is already co-author of one paper on the International Conference on Planning and Scheduling (ICAPS)~\cite{diego-icaps18}, and of one journal paper on research topics that are relevant to the development of the project~\cite{onaindia2018common}.
+With Eva Onaindia and Sergio Jiménez, Diego is already co-author of one paper on the International Conference on Planning and Scheduling (ICAPS)~\cite{diego-icaps18}, and one journal paper~\cite{onaindia2018common}.
+Both papers on research topics related to the development of this project.
 
@@
 Further, tests alert programmers of bugs before handing the code off to clients (the cost of finding a bug when the code is first written is considerably lower than the cost of detecting and fixing it later).
-Last but not least, writing a thorough set of tests cases forces programmers to think through inputs, outputs, and error conditions of programs.
+%Last but not least, writing a thorough set of tests cases forces programmers to think through inputs, outputs, and error conditions of programs.
 \end{itemize}
 
-Now we review the two most successful approaches for program synthesis. The research posed in this project falls into the first approach:
+{\em Program synthesis} is the task of computing a program that satisfies a given formal specification.
+On the other hand, {\em program validation} is the task of proving, or disproving, the correctness of a given program with respect to a certain formal specification or property.
+Program validation is considered a necessary step for program synthesis.
+In 2008 the PhD Thesis work by Armando Solar-Lezama, at the University of California Berkeley, showed that it is possible to encode program synthesis problems in Boolean logic and therefore, use SAT algorithms to automatically compute programs~\cite{lezama2008program}.
+Since then, there has been a surge of practical interest in the idea of program synthesis in the formal verification community and related fields~\cite{alur2013syntax}.
+To illustrate this, in 2013, a unified framework for program synthesis problems was proposed and since 2014 there is a yearly program synthesis competition, comparing the different algorithms for program synthesis in a competitive event ({\tt www.sygus.org}).
+
+Now we review the two most successful approaches for automated program synthesis. 
 \begin{itemize}
@@
 
-Closely related to the aims of the project is also the work on the synthesis and validation of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
+Closely related to the aims of the project is also the work on the synthesis of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
 The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}.
@@
 ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn a given target concept, expressed as a  logic program,  from  examples  and  background  knowledge, that are expressed as logic facts.
-\\
-\\
+
 \begin{center}
@@
 \item {\tt var:=value}, that encode the program variables.
-\item {\tt program(line):=instruction}, that encode the program lines.
+\item {\tt program(line):=instruction}, that encode the instructions at the different program lines.
 \item {\tt pcounter:=line}, encoding the current program line.
@@
 
-The {\bf initial/goal states} of the AIP task encode the initial/final values of the program variables according to the input/output tests of the TDD programming task.
+The {\bf initial/goal states} of the AIP task encode the initial/final values of the program variables that are given by the input/output tests of the TDD programming task.
 
@@
 
-Interestingly, our PDDL encoding allows also program validation by (1), specifying the lines of the program to validate in the initial state of the AIP task and (2), disabling the mentioned {\it programming actions} so only {\it execution actions} are applicable.
+Interestingly, our PDDL encoding allows also program validation by (1), specifying the lines of the program to validate (i.e.
+the {\tt program(line):=instruction} fluents) in the initial state of the AIP task and (2), disabling the mentioned {\it programming actions} so only {\it execution actions} are applicable.
 
 \subsection{Evaluation}
+\label{sec:evaluation}
 The performance of our approach for the synthesis and validation of TDD programs will be evaluated in two kinds of programming tasks:
@@
 \item {\em Real-world benchmarks}: The GRPS group is currently leading a four-year research project from the {\em Spanish national plan} in which {\em activity recognition} is applied to different real-world domains such as {\em domotics}, {\em tourism}, {\em traffic control} and {\em robotics}.
+Interestingly many activities in these domains can be understood as simple programs.
+For instance, Figure~\ref{fig:activity} shows a four-line program (pictured as a finite state machine) that represents the sequence of instructions required for the {\em making a buttered toast} activity.
 We plan to evaluate our approach in synthesis and validation tasks coming from these real-world domains.
-Figure~\ref{fig:activity} shows a four-line program (pictured as a finite state machine) that represents the sequence of instructions required for the {\em making a buttered toast} activity.
 
@@
 
-\item {\em Theoretical benchmarks}: Programming tasks for the computation of mathematical/logic series, string manipulation and for the management of classic data structures such as {\em lists}, {\em queues}, {\em stacks} or {\em trees}.
-Figure~\ref{fig:list} shows a synthesized program for traversing linked lists.
+\item {\em Theoretical benchmarks}: Classic programming tasks are a neat touchstone to assess the performance of our approach.
+For instance, programs for the computation of mathematical/logic series, string manipulation and for the management of data structures such as {\em lists}, {\em queues}, {\em stacks} or {\em trees}.
+Figure~\ref{fig:list} shows a synthesized program for traversing a linked list.
 The program is pictured as a {\em finite state machine}: The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
@@
 
-\subsection{Two-year workplan}
+\subsection{Workplan}
 We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that (1), takes as input a set of test cases that specify the TDD programming task to solve and (2), outputs a program source code that passes these tests with a bug-free guarantee.
@@
 \ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Development (6 months)}{4}{6}\\
-\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Dissemination (12 months)}{7}{12}
+\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Experiments \& dissemination (12 months)}{7}{12}
 \end{ganttchart}
-\caption{\small Workplan for developing a user-interactive program synthesizer/validator.}
+\caption{\small Work-plan for developing a user-interactive program synthesizer/validator.}
 \label{fig:gantt}
@@
 
-{\small{\bf\em  Deliverable:} Technical report with the specifications of the system design.}
+{\small{\bf\em  Deliverable T1:} Technical report with the specifications of the system design.}
 
@@
 \end{center}  
-\caption{\small The system architecture.}
+\caption{\small System architecture for the synthesis and validation of TDD programs.}
 \label{fig:architecture}
@@
 
-{\small{\bf\em Deliverable:} Open repository with the source code of the system architecture.}
+{\small{\bf\em Deliverable T2:} Open repository with the source code of the system architecture.}
 
@@
       \begin{enumerate}
-      \item Reporting the experimental performance of our AIP approach for solving diverse TDD programming tasks.
+      \item Reporting the experimental performance of our AIP approach for solving diverse TDD programming tasks. This task will follow an iterative workflow over the following subtasks:
+      \begin{enumerate}
+      \item Executing the system architecture in the {\em theoretical benchmarks} described in Section~\ref{sec:evaluation}.
+      \item Analysis and validation of the obtained results.
+      \item Tuning and repairing the system components, evaluation metrics and benchmarks according to the obtained results.                 
+      \item Executing the system architecture in the {\em real-world benchmarks} introduced in Section~\ref{sec:evaluation}.
+      \item Analysis and validation of the obtained results.         
+      \item Tuning and repairing the system components, evaluation metrics and benchmarks according to the obtained results.                 
+      \end{enumerate}
       \item Dissemination of the obtained theoretical and empirical results by submitting papers to top international conferences and journals in AI.        
@@
 \end{small}        
-{\small{\bf\em  Deliverable:} Final report with obtained conclusions and produced publications.}
+{\small{\bf\em  Deliverable T3:} Final report with the obtained conclusions and produced publications.}
 \end{enumerate}
 
+\subsection{Workload}
+
 {\em Tasks ({\bf T1-T3})} will be developed by the three mentioned members of the research group: Sergio Jiménez, Eva Onaindia and Diego Aineto.
-In addition, we plan to hire a master student to assist in the completion of {\em task {\bf T3(a)}}.
+
+In addition, we plan to hire a master student to assist in the completion of two tasks, {\bf T3(a,i)} and {\bf T3(a,iv)}. His detailed responsability will be:
+\begin{enumerate}
+\item Executing the scripts of the system architecture.
+\item Collecting the output data.
+\item Presenting the output data in a easy-readable format.
+\end{enumerate}
 
@@
 \item Number of {\em program lines}.
-\item Number of observable {\em program variables}.
+\item Number and domain of the observable {\em program variables}.
 \item Size of the available {\em instruction set}.
@@
 
-In more detail, the expected benefits of this research project are four-fold:
+In more detail, the expected benefits for this particular research project are four-fold:
 \begin{enumerate}
 \item The definition of {\bf effective metrics for assessing how well a given program covers a set of {\em input-output} test cases}.
-Our belief is that these metrics can provide new insights into the current understanding of how Artificial Intelligence can assist programmers in the software development.
+These metrics can provide new insights into the current understanding of how Artificial Intelligence can assist programmers in the software development.
 \item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}.
@@
 This project will also help to meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks.
-\item The {\bf development and evaluation of software and benchmarks for the {\em Synthesis and Validation of Programs}}.
+\item The {\bf development of open software and benchmarks for the {\em Synthesis and Validation of Programs}}.
 We strongly belief that reproducibility and open knowledge are essential to the advance of the research on computer science.
 With this regard we plan to develop a {\em github} repository where we make available the developed source code and benchmarks.
-\item The {\bf dissemination of the obtained scientific results}.
+\item International {\bf dissemination of the obtained scientific results}.
 The scientific results obtained during the development of the project will also be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).

### 2018-06-10 doc/aij18/background.tex
Background and Motivation completed
Background and Motivation completed

--- 2018-05-30 doc/aij18/background.tex
@@
 
-This section serves two purposes.
-In subsection \ref{basic_planning} we introduce the basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper.
-Section \ref{related_work} summarizes existing approaches to learn action models in classical planning and highlights the novelty of our contribution via the related work.
-
-%This section defines the planning model used on this work, {\em classical planning with conditional effects}, and the target of the learning/evaluation/recognition tasks addressed in the paper, a \strips\ action model.
+This section serves two purposes: first we introduce the basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper; and secondly, we summarize the existing approaches to learn action models in classical planning and highlight our contributions via the related work.
 
@@
 $|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as is common in \strips\ planning.
+Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.
 
@@
 \item $\pre(a)\in\mathcal{L}(F)$, the {\em preconditions} of $a$, is the set of literals that must hold for the action $a\in A$ to be applicable.
-\item $\eff^+(a)\in\mathcal{L}(F)$, the {\em positive effects} of $a$, is the set of literals that true after the application of the action $a\in A$.
+\item $\eff^+(a)\in\mathcal{L}(F)$, the {\em positive effects} of $a$, is the set of literals that are true after the application of the action $a\in A$.
 \item $\eff^-(a)\in\mathcal{L}(F)$, the {\em negative effects} of $a$, is the set of literals that are false after the application of the action.
@@
 The {\em plan length} is denoted with $|\pi|=n$ .
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.,~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
+A solution plan is {\em optimal} if it has minimum length.
 
-In this work, we will use the term \emph{observation} to refer to the input knowledge of a learning task.
-Hence, $s=\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$.
+In this work, the term \emph{plan trace} refers to the \emph{observation} of a plan execution that starts on a given initial state, which constitutes the input knowledge of a learning task.
+Specifically, $s=\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$.
 Otherwise, $s$ is a partially-observable (\PO) state trajectory, meaning that at least one $s_i$ is a partial assignment of values to fluents in which one or more literals are missing.
-This general definition gives rise to two particular cases:
+This implies that all fluents of a state $s_i$ may be missing, in which case, $s_i$ is a \emph{missing} or \emph{empty} state.
+This general definition of \PO gives rise to two particular cases:
 
 \begin{itemize}
-\item when all fluents of a state $s_i$ are missing, $s_i$ is a \emph{missing} or \emph{empty} state
-\item when the $n-1$ intermediate states of a trajectory $s$ are missing, $s$ is a \emph{non-observable} (\NO) state trajectory
+\item when \textbf{all} the $n-1$ intermediate states of a trajectory $s$ are \textbf{missing}, $s$ is a \emph{non-observable} (\NO) state trajectory
+\item when \textbf{none} of the $n-1$ intermediate states of a trajectory $s$ is \textbf{missing}, we will refer to $s$ as a \POstar state trajectory
 \end{itemize}
 
-Hence, a partially-observed state trajectory can be interpreted as a sequence that contains $n+1$ possibly incomplete states induced by an $n$-action sequence, or as a sequence with some missing intermediate states, which can be regarded as a particular case of partial observability.
-We will refer to the former as a \PO trajectory and to the latter as a \POstar state sequence.
+Table \ref{tab:state_trajectory} summarizes the four types of state trajectories according to the observed information, which ultimately affects the number of observed intermediate states and the number of literals comprised in each intermediate state.
+\PO comprises both \POstar and \NO, and it thus encompasses trajectories with some missing state.
 
-%In the rest of the paper, we will use the following notation to refer to the different levels of state visibility: $s$ denotes a \FO state trajectory; $s^{+}$ refers to a \PO state trajectory with $n+1$ possibly incomplete states; $s^{*}$ indicates a \PO state trajectory with $m$ missing intermediate states, $m < n-1$; and $s^{0}$ is a \NO state trajectory with $n-1$ missing states.
+\begin{table}[ht]
+\centering
+\begin{tabular}{c|c|c|}
+	     & \# intermediate states & state type \\ \hline
+	\FO & $n-1$  & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ is a full assignment \\ \hline
+    \multirow{1}{*}{\POstar} & $n-1$ & {\small $\exists i, 1 \leq i < n$}  \\ & & $s_i$ is a partial assignment \\ \hline
+    \multirow{1}{*}{\PO} & $\leq n-1$ & {\small $\exists i, 1 \leq i < n$}   \\  & & $s_i$ is a partial assignment \\ \hline
+    \NO & 0  & ---  \\ \hline
+\end{tabular}
+\caption{Classification of state trajectories accordingly to observation}
+\label{tab:state_trajectory}
+\end{table}
 
-Regarding an action sequence, $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to achieve the goal from an initial state, a \PO action sequence if at least one necessary action to achieve the goal is missing in the sequence and a \NO action sequence in case the sequence is empty.
+Regarding an action sequence, $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to transit every state $s_{i-1}$ to the corresponding successor state $s_{i}$, for each {\small $1\leq i\le n$}.
+We say it is a \PO action sequence if at least one of these necessary actions is missing in the sequence and a \NO action sequence in case the sequence is empty.
 
-A \emph{plan trace} is a combination of a state and an action sequence $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
+A \emph{plan trace} $\tau$ is a combination of a state and an action sequence: $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
 A plan trace $\tau$ for a planning problem $P=\tup{F,A,I,G}$ satisfies $s_0=I$, $G \in s_n$ and every element in the action sequence of $\tau$ is an action of $A$.
-Additionally, plan traces will be classified accordingly to the type of state trajectory (\FO, \PO, \POstar or \NO) and action sequence (\FO, \PO or \NO).
+Additionally, plan traces will be classified accordingly to the type of state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
 
@@
 
-The first column of Table \ref{table:models_comparison1} shows that all the approaches analyzed in this work accept partial observations of intermediate states, and some approaches also allow the sequence of intermediate states to be empty.
-Although not explicitly stated in the corresponding published works, it seems reasonable to assume that handling \NO state sequences subsumes \POstar trajectories and that a \POstar plan trace can be regarded as composed of several \NO sub-traces.
+The first column of Table \ref{table:models_comparison1} shows the constraints imposed on the input plan traces with regards to observability.
+Since all approaches except ours deal only with \FO action sequences, constraints are exclusively concerned with the type of state trajectory.
+This directly affects the complexity of the task, which can be sorted from the least to the most constrained following this order: 1) \NO, 2) \PO, 3) \POstar, and 4) \FO.
+Note that \PO is less constrained than \POstar because it also includes the possibility of having some missing state in the trajectory.
+
+The task of learning from less constrained traces subsumes learning from more constrained ones.
+Consequently, approaches to learning from, for instance, traces with \PO state trajectories will also be able to learn from traces with \POstar state trajectories.
+All the approaches analyzed in this work accept the more constrained definition of partial observations of intermediate states \POstar, and most of them also allow the sequence of intermediate states to be empty.
 Exceptionally, a \NO state sequence in \LOCM is a fully-empty trajectory, with neither initial or final state.
@@
 The expressiveness of the learned action models varies across approaches (second column of Table \ref{table:models_comparison1}).
-All the presented systems are able to learn action models in a \textsc{Strips}  representation \cite{fikes1971strips} and some propose algorithms to learn more expressive action models that include quantifiers, logical implications or the type hierarchy of the domain.
+All the presented systems are able to learn action models in a \textsc{Strips}  representation \cite{fikes1971strips} and some propose algorithms to learn more expressive action models that include quantifiers, logical implications or the type hierarchy of a PDDL domain.
 
@@
 The \ARMS system became a benchmark in action-model learning, showing empirically that is is feasible lo learn a model in a reasonably efficient way using a weighted MAX-SAT even with \NO state trajectories.
-Nevertheless, we argue that using the home-made error and redundancy metrics to evaluate the model correctness is questionable since a model that exhibits a wrong precondition or effect in any of its actions could still be eligible as the minimal-error output model.
+%Nevertheless, we argue that using the home-made error and redundancy metrics to evaluate the model correctness is questionable since a model that exhibits a wrong precondition or effect in any of its actions could still be eligible as the minimal-error output model.
 
@@
 		\hline			
-		\multirow{2}{*}{\ARMS} & \PO and \NO states & \strips & MAX-SAT \\ & \FO actions & & \\
+		\multirow{2}{*}{\ARMS} & \NO states & \strips & MAX-SAT \\ & \FO actions & & \\
         \hline
-        \multirow{2}{*}{\SLAF} & \PO states  & universal quantifiers in $\eff$ & logical inference \\ & \FO actions &  & SAT solver \\
+        \multirow{2}{*}{\SLAF} & \POstar states  & universal quantifiers in $\eff$ & logical inference \\ & \FO actions &  & SAT solver \\
          \hline
@@
          %\hline
-		\multirow{2}{*}{\LAMP} & \PO and \PO* states &  quantifiers &  Markov logic networks \\  & \FO actions & logical implications &  \\
+		\multirow{2}{*}{\LAMP} & \PO states &  quantifiers &  Markov logic networks \\  & \FO actions & logical implications &  \\
          \hline
@@
          \hline
-         \NOISTA & \PO and noisy states & \strips &  classification \\ & \FO actions & & \strips \texttt{} rules derivation \\
+         \NOISTA & \POstar and noisy states & \strips &  classification \\ & \FO actions & & \strips \texttt{} rules derivation \\
          \hline
-         \CAMA & \PO and \PO* states &  \strips & crowdsourcing annotation\\ & \FO actions &  & MAX-SAT \\
+         \CAMA & \PO states &  \strips & crowdsourcing annotation\\ & \FO actions &  & MAX-SAT \\
          \hline
@@
          \hline
-		\FAMA &  \PO, \PO* and \NO states & \strips &   compilation to planning\\ & \PO and \NO actions & & \\
+		\FAMA & \NO states & \strips &   compilation to planning\\ & \NO actions & & \\
          \hline
@@
 A tractable and exact solution of action models in partially observable domains using a technique known as Simultaneous Learning and Filtering (\textbf{\SLAF}) is presented in~\cite{AmirC08}.
-\SLAF alongside \ARMS can be considered another of the precursors of the modern algorithms for action-model learning, able to learn from plan traces with PO and non-null intermediate states.
+\SLAF alongside \ARMS can be considered another of the precursors of the modern algorithms for action-model learning, able to learn from partially observable states.
 Given a formula representing the initial belief state, a sequence of executed actions and the corresponding partially observed states, \SLAF builds a complete explanation of observations by models of actions through a CNF formula.
@@
 
-\LAMP allows \POstar state observations up to a minimum percentage of 1/5 of the observed states as well as \PO state trajectories with different degrees of observability in the number of propositions in each state.
+\LAMP allows \PO state trajectories up to a minimum percentage of 1/5 of non-empty states as well as \POstar state trajectories with different degrees of observability in the number of propositions in each state.
 It uses an error metric based on counting the differences in the number of precondition and effects between the ground-truth model and the learned model.
@@
 After that, the parameters of the graphical model are learned, computing at the same time the probability distribution of each candidate domain model.
-Finally \AMAN returns the model that maximizes a reward function defined in terms of the percentage of actions successfully executed and the percentage of goal propositions achieved after the last successfully executed action.
+\AMAN finally returns the model that maximizes a reward function defined in terms of the percentage of actions successfully executed and the percentage of goal propositions achieved after the last successfully executed action.
 
@@
 In a comparison between \AMAN and \ARMS on noiseless inputs, the results show that the accuracies of both \AMAN and \ARMS are very close to each other and neither dominates the other.
-On the other hand, a test comparing plan traces with and without noisy actions reveals that the accuracy in the case \emph{with noise} becomes more and more close to \emph{without noise} as the number of iterations of \AMAN algorithm increases because the distribution of noise in the plan becomes gradually closer to the real distribution.
+The convergence property of \AMAN guarantees that the accuracy of the learned model with noisy input traces becomes more and more close to the case  \emph{without noise} because the distribution of noise in the plan becomes gradually closer to real distribution with the number of iterations.
+
+%On the other hand, a test comparing plan traces with and without noisy actions reveals that the accuracy in the case \emph{with noise} becomes more and more close to \emph{without noise} as the number of iterations of \AMAN algorithm increases because the distribution of noise in the plan becomes gradually closer to the real distribution.
 
@@
 Then it solves the constraint-based problem using a MAX-SAT solver and converts the solution to action models.
+
+Plan traces in \CAMA are composed of 80\% of empty states and each partial state was selected by 50\% of propositions in the corresponding full state.
 An experimental comparison reveals that a manual crowdsourcing of \CAMA outperforms \ARMS and that as expectedly the difference becomes smaller as the number of plan traces becomes larger.
-The accuracy of \CAMA for a small number of plan traces (e.g., 30) is not less than 80\%, which means that exploiting the knowledge of the crowd can help learning action models.
+The accuracy of \CAMA for a small number of plan traces (e.g., 30) is not less than 80\%, thus revealing that exploiting the knowledge of the crowd can help learning action models.
 
@@
 
-The Learning Object-Centred Models (\textbf{\LOCM}) is an approach that only requires the \FO action sequence as input knowledge, without need for providing any information about the predicates or the state trajectory, not even the initial or final state~cite{cresswell2013acquiring}.
-The lack of available information is overcome by exploiting assumptions about the structure of the actions.
-Particularly, it assumes that objects found in the same position in the header of actions are grouped as a collection of objects (sorts) whose defined set of states is captured by a parameterized Finite State Machine (FSM).
-The intuitive assumptions of \LOCM, like the continuity of object transitions or the association of parameters between consecutive actions in the training datset, yield a learning model heavily reliant on the kind of domain structure.
+The Learning Object-Centred Models (\textbf{\LOCM}) is an approach that only requires the \FO action sequence as input knowledge, without need for providing any information about the predicates or the state trajectory, not even the initial or final state~\cite{CresswellMW09,cresswell2013acquiring}.
+The lack of available state information is overcome by exploiting assumptions about the structure of the actions.
+Particularly, \LOCM assumes that objects found in the same position in the header of actions are grouped as a collection of objects (sorts) whose defined set of states is captured by a parameterized Finite State Machine (FSM).
+The intuitive assumptions of \LOCM, like the continuity of object transitions or the association of parameters between consecutive actions in the training dataset, yield a learning model heavily reliant on the kind of domain structure.
 A later work, \textbf{\LOCMtwo}, extends the applicability of the \LOCM algorithm to a wider range of domains by introducing a richer representation that allows using multiple FSMs to represent the state of a sort~\cite{cresswell2011generalised}.
 
-\LOCMtwo is not experimentally evaluated and only the outcome of running the \LOCMtwo algorithm on several benchmark domains wrt to the reference model is reported.
-It is worth noting the last contribution of \LOCM, called \textbf{\LOP} (\LOCM with Optimized Plans), addresses the problem of inducing static predicates~\cite{GregoryC16}.
+\LOCMtwo is not experimentally evaluated, only the outcome of running the \LOCMtwo algorithm on several benchmark domains wrt to the reference model is reported in ~\cite{cresswell2011generalised}.
+It is worth noting the last contribution of the \LOCM family, called \textbf{\LOP} (\LOCM with Optimized Plans), addresses the problem of inducing static predicates~\cite{GregoryC16}.
 \LOP applies a post-processing step after the \LOCM analysis and it requires additional input information, particularly a set of optimal plans besides the suboptimal \FO action sequences.

### 2018-06-10 doc/programming-project/AI-programming.tex
added comments on model checking

@@
 {\em Program synthesis} is the task of computing a program that satisfies a given formal specification.
-On the other hand, {\em program validation} is the task of proving, or disproving, the correctness of a given program with respect to a certain formal specification or property.
-Program validation is considered a necessary step for program synthesis.
-In 2008 the PhD Thesis work by Armando Solar-Lezama, at the University of California Berkeley, showed that it is possible to encode program synthesis problems in Boolean logic and therefore, use SAT algorithms to automatically compute programs~\cite{lezama2008program}.
+The 2008 PhD Thesis work by Armando Solar-Lezama, at the University of California Berkeley, showed that it is possible to encode program synthesis problems in Boolean logic and therefore, use SAT algorithms to automatically compute programs~\cite{lezama2008program}.
 Since then, there has been a surge of practical interest in the idea of program synthesis in the formal verification community and related fields~\cite{alur2013syntax}.
@@
 
-Closely related to the aims of the project is also the work on the synthesis of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
+The synthesis of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15} is closely related to program synthesis.
 The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}.
@@
 {\em classical planning}~\cite{sergio:aprograming:icaps16}, {\em conformant planning}~\cite{Geffner:FSM:AAAI10}, {\em CSP}~\cite{Infantes:FSC:ECAI2010} or a {\em Prolog program}~\cite{Giacomo:FSM:ICAPS13}).
+The generation of programs from examples is a research question also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
+ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn a given target concept, expressed as a  logic program,  from  examples  and  background  knowledge, that are expressed as logic facts.
 
-Finally the generation of programs from examples is a research question also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
-ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn a given target concept, expressed as a  logic program,  from  examples  and  background  knowledge, that are expressed as logic facts.
+On the other hand, {\em program validation} is the task of proving, or disproving, the correctness of a given program with respect to a certain formal specification or property.
+Program validation is considered a necessary step for program synthesis.
+
+Closely related to {\em program validation} are the aims of {\em model checking}~\cite{clarke1999model}.
+In model checking a given model of a system, is exhaustively and automatically checked to verify  whether this model meets a given specification.
+Typically the specification contains safety requirements such as the absence of deadlocks and similar critical states that can cause the system to crash.
+Both the model of the system and the specification are formulated in a formal language.
+To this end, the problem is formulated as a task in logic, namely to check whether a given structure satisfies a given logical formula.
+
+Currrent approaches for model checking reduces to graph search.
+Instead of enumerating reachable states one at a time, the state space can sometimes be traversed more efficiently by considering large numbers of states at a single step.
+When such state space traversal is based on representations of set of states and transition relations as logical formulas, binary decision diagrams (BDD) or other related data structures, the model-checking method is symbolic~\cite{mcmillan1993symbolic}.
 

### 2018-06-10 doc/aij18/motivation.tex

--- 2018-05-30 doc/aij18/motivation.tex
@@
 
-The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which makes the learning task more challenging.
-Essentially, this brings two key differences: a) the transition function between two observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_n})=s_{i+1}$; and b) the horizon of the input plan traces is not longer known.
-These two factors justify the use of planning techniques for solving the learning task over other methods such as a SAT solver.
+The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories makes the learning task more challenging.
+This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \leq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now.
 
-We set a probability threshold of observability to each fluent and action, which determines the percentage of literals in each state and in turn the percentage of observed states (when the probability of every fluent is above the threshold, the result is a missing state).
-This way, \FAMA must always work under the assumption that the length of the input plan traces is unknown.
+In this particular scenario, the actual number of plan traces that can correspond with the given input observations is also unbound and grows exponentially with the actual length of the plan trace (that is now unknown).
+Otherwise, the learning task is SAT compilable, which is known to be a NP-complete task~\cite{russell2016artificial}.
+This is the reason that SAT solving is a common technique in the approaches presented in section \ref{sec:background}.
 
+When we assume partial observability in both the sequence of actions and the state trajectory, we must always consider the length of the input plan traces to be unknown and so the task of learning an action model becomes as hard as solving a \strips\ planning problem.
+Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning techniques, as our proposal of compiling the learning task to a planning problem.
+
+%In \FAMA we set a probability threshold of observability to each fluent and action, which determines the percentage of literals in each state and in turn the percentage of observed states (when the probability of every fluent is above the threshold, the result is a missing state).
+This way, \FAMA must always work under the assumption that the number and length of the input plan traces are unknown and so the task of learning an action model becomes as hard as solving a \strips planning problem.
+Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning techniques, as our proposal of compiling the learning task to a planning problem.
+
+%Thereby, solving this type of learning tasks justifies the use of techniques other than SAT solvers, as our proposal of compiling the learning task to a planning problem.
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{c|c|c|c|c|}
+	& \multicolumn{4}{c|}{\emph{state observability}} \\ \cline{2-5}
+	\multirow{1}{*}{\emph{action}} & \FO & \POstar & \PO & \NO\\ {\emph{observability}} & & & & \\ \hline
+	\FO & NP-complete & NP-complete & NP-complete & NP-complete \\ \hline
+	\PO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
+	\NO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
+\end{tabular}
+\caption{Complexity of learning tasks according to the type of input trace}
+\label{tab:complex}
+\end{table}
+
+Table \ref{tab:complex} displays the complexity of the learning task according to the type of input trace.
+The PSPACE-complexity happens when the length of the plan trace is a priori unbound, which results from the combination of the following two causes:
+
+\begin{enumerate}
+\item As many of the approaches in section \ref{sec:background} assume, there may be an unknown number of missing intermediate states in the trace because of partial state observability (\PO and \NO).
+The assumption of having \FO state trajectory means that the sensors are able to capture every state change at every instant which typically is unrealistic.
+Normally the process of getting state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two sensor readings.
+
+\item There may be also an unbound number of missing actions in the plan trace because of partial observability.
 The common assumption of having \FO action sequences in a learning task is rather unrealistic in many domains because it implies the existence of human observers that annotate the observed action sequences.
@@
 Learning from unstructured data involves some prior processing to transform the sensor or image information into a predicate-like format before applying the action-model learning approach~\cite{AsaiF18}, and it also requires the ability of identifying action symbols.
-In this sense, this work represents a step ahead towards learning action models without assuming observed actions.
-
-Obviously, learning from \PO and \NO action sequences is much more complex.
-When the number of actions and the number of states is unknown, solving the learning task is as hard as solving a \strips\texttt{} planning problem; i.e., it is PSPACE-complete.
-Otherwise, the complexity of the task is NP-complete, in which case it can be translated into SAT, which is known to be a NP-complete problem (see Table \ref{tab:complex}).
-This is the reason that SAT solving is a common technique in the approaches presented in section \ref{sec:background}.
-
-\begin{table}[ht]
-\centering
-\begin{tabular}{c|c|c|c|c|}
-	& \multicolumn{4}{c|}{\emph{state observability}} \\ \cline{2-5}
-	\multirow{1}{*}{\emph{action}} & \FO & \PO & \POstar & \NO\\ {\emph{observability}} & & & & \\ \hline
-	\FO & NP-complete & NP-complete & NP-complete & NP-complete \\ \hline
-	\PO & NP-complete & PSPACE-complete & PSPACE-complete & PSPACE-complete \\ \hline
-	\NO & NP-complete & PSPACE-complete & PSPACE-complete & PSPACE-complete\\ \hline
-\end{tabular}
-\caption{Complexity of learning tasks according to the type of input trace}
-\label{tab:complex}
-\end{table}
+In this sense, \FAMA represents a step ahead towards learning action models without assuming observed actions.
+\end{enumerate}
 
 Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists in counting the missing and extra fluents that appear in the learned model wrt the GTM and normalizing by the the total number of all the possible preconditions and effects of an action model.
-This optimistic metric returns error rates that do not reflect the accuracy of the learned model as it could output a rate below 100\% for a totally wrong learned model that only includes a subset of all possible preconditions and effects of an action model.
-To overcome this limitation, we propose to use two standard metrics, {\em precision} and {\em recall}.
+This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual model.
+Hence, because the set of preconditions and effects of a model is usually small compared to the set of all possible preconditions and effects, it turns out the metric may output an error rate below 100\% for a totally wrong learned model.
+
+In order to overcome this limitation, we propose to use two standard metrics, {\em precision} and {\em recall}.
 These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.

### 2018-06-10 doc/programming-project/AI-programming.tex
project reference added

@@
 Since 2016 Diego is holding a four-year FPU grant on {\em Causal discovery for predictive analysis} (FPU16/03184).
-Diego is also involved on the current research project of the GPRS from the Spanish national research plan on {\em activity recognition}.
+Diego is also involved on the current research project TIN2017-88476-C2-1-R of the GPRS from the Spanish national research plan with title {\small\em RECONOCIMIENTO DE ACTIVIDADES Y PLANIFICACION AUTOMATICA PARA EL DISEÑO DE ASISTENTES INTELIGENTES}.
 
@@
 Program validation is considered a necessary step for program synthesis.
-
-Closely related to {\em program validation} are the aims of {\em model checking}~\cite{clarke1999model}.
+{\em Model checking} is the best-known approach for program validation~\cite{clarke1999model}.
 In model checking a given model of a system, is exhaustively and automatically checked to verify  whether this model meets a given specification.

### 2018-06-11 doc/aij18/background.tex
small updates added to first sections

@@
 
-In this work, the term \emph{plan trace} refers to the \emph{observation} of a plan execution that starts on a given initial state, which constitutes the input knowledge of a learning task.
-Specifically, $s=\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$.
-Otherwise, $s$ is a partially-observable (\PO) state trajectory, meaning that at least one $s_i$ is a partial assignment of values to fluents in which one or more literals are missing.
-This implies that all fluents of a state $s_i$ may be missing, in which case, $s_i$ is a \emph{missing} or \emph{empty} state.
+In this work, the term \emph{plan trace} refers to the \emph{observation} of a plan execution that starts on a given initial state.
+\emph{Plan traces} constitute the input knowledge of the learning tasks addressed in this paper.
+With regard to the observed states in a plan trace, we say that $\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$.
+Otherwise, it is a partially-observable (\PO) state trajectory, meaning that at least one state $s_i$ is a partial assignment of values to fluents in which one or more literals are missing (formally, $|s_i|<|F|$).
+This implies that in a \PO state trajectory all fluents of a state $s_i$ may be missing, in which case, $s_i$ is a \emph{missing} or \emph{empty} state.
 This general definition of \PO gives rise to two particular cases:
@@
 \item when \textbf{all} the $n-1$ intermediate states of a trajectory $s$ are \textbf{missing}, $s$ is a \emph{non-observable} (\NO) state trajectory
-\item when \textbf{none} of the $n-1$ intermediate states of a trajectory $s$ is \textbf{missing}, we will refer to $s$ as a \POstar state trajectory
+\item when \textbf{none} of the $n-1$ intermediate states of a trajectory $s$ is \textbf{missing}, we will refer to $s$ as a \POstar state trajectory. 
 \end{itemize}
@@
 
-\begin{table}[ht]
+\begin{table}[hbt!]
 \centering
@@
     \multirow{1}{*}{\PO} & $\leq n-1$ & {\small $\exists i, 1 \leq i < n$}   \\  & & $s_i$ is a partial assignment \\ \hline
-    \NO & 0  & ---  \\ \hline
+    \NO & 0 & {\small $\forall i, 1 \leq i < n$}  \\  & & $|s_i|=0$  \\\hline
 \end{tabular}
@@
 
-Regarding an action sequence, $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to transit every state $s_{i-1}$ to the corresponding successor state $s_{i}$, for each {\small $1\leq i\le n$}.
-We say it is a \PO action sequence if at least one of these necessary actions is missing in the sequence and a \NO action sequence in case the sequence is empty.
+With regard to the observed actions in a plan trace, we say that an action sequence, $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to transit every state $s_{i-1}$ to the corresponding successor state $s_{i}$, for each {\small $1\leq i\le n$}.
+We say it is a \PO action sequence if at least one of these necessary actions is missing in the sequence and a \NO action sequence in case the sequence of observed actions is empty.
 
-A \emph{plan trace} $\tau$ is a combination of a state and an action sequence: $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
+A \emph{plan trace} $\tau$ is an interleaved combination of a state and an action sequence: $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
 A plan trace $\tau$ for a planning problem $P=\tup{F,A,I,G}$ satisfies $s_0=I$, $G \in s_n$ and every element in the action sequence of $\tau$ is an action of $A$.
-Additionally, plan traces will be classified accordingly to the type of state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
+Additionally, plan traces will be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
 
@@
 Others do not explicitly say so but the fact is that the predicates and the actions headers are easily extractable from the state sequence and action sequence of the plan traces, respectively.
-A requirement, however, for these two sets to be representative is that set of input plan traces comprises a grounded sample of all predicates and operator schemas of the domain model.
+A requirement, however, for these two sets to be representative is that set of input plan traces comprises at least a grounded sample of all predicates and operator schemas of the domain model.
 

### 2018-06-11 doc/aij18/introduction.tex
small updates added to first sections

@@
 \item {\bf\em Model validation}.
-The compilation poses a novel framework to assess the validation of a STRIPS model with respect a plan execution.
+The compilation poses a novel framework to assess the validation of a STRIPS model with respect to plan executions.
 This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since it does require neither a full action model nor a fully observed plan to determine validation.
@@
 \item {\bf\em Model evaluation}.
-The compilation can assess how well a given \strips\ action model matches a plan execution, which allows us to assess learning performance without having the actual action model.
-The idea is to assess the amount of edition that is required by the input action model to induce the given observation of the plan execution.
+The compilation can assess how well a given \strips\ action model matches a plan execution, which allows us to assess learning performance without knowing the actual action model.
+The idea is to assess the amount of edition that is required by the input action model to induce the given observations of plan executions.
 \end{enumerate}
 
-A first description of the compilation previously appeared in our conference paper~\cite{aineto2018learning}. Compared to that conference paper, this work includes the following novel material:
+A first description of the compilation previously appeared in our previous conference paper~\cite{aineto2018learning}. Compared to that paper, this work includes the following novel material:
 \begin{itemize}
@@
 %\item Leveraging diverse forms of {\em background knowledge} for learning and evaluating \strips\ action models.
-\item A complete empirical evaluation of the compilation approach.
-Our evaluation analyses how the amount of input knowledge affects to the performance of the compilation approach when learning and evaluating \strips\ action models.
+\item A complete empirical evaluation of the compilation approach for learning of \strips\ action models.
+Our evaluation analyses how input knowledge affects to the performance of the compilation approach when learning and evaluating \strips\ action models.
 \end{itemize}
@@
 Section~\ref{sec:background} introduces classical planning and reviews related work on learning planning action models.
-Section~\ref{sec:Section3} introduces the classical planning model and the \strips\ action model (the target of our learning and evaluation tasks).
+Section~\ref{sec:motivation} motivates our compilation approach for learning of \strips\ action models from observations of plan executions.
+Section~\ref{sec:section3} introduces the classical planning model and the \strips\ action model (the target of our learning and evaluation tasks).
 Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.

### 2018-06-11 doc/aij18/motivation.tex
small updates added to first sections

@@
 
-The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories makes the learning task more challenging.
+The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging.
 This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \leq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now.
@@
 
-When we assume partial observability in both the sequence of actions and the state trajectory, we must always consider the length of the input plan traces to be unknown and so the task of learning an action model becomes as hard as solving a \strips\ planning problem.
+When we assume partial observability in both the sequence of actions and the state trajectory, a complete approach must consider the length of the input plan traces to be unknown.
+This work shows that classical planning is a complete approach for this particular scenario.
 Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning techniques, as our proposal of compiling the learning task to a planning problem.
@@
 	\multirow{1}{*}{\emph{action}} & \FO & \POstar & \PO & \NO\\ {\emph{observability}} & & & & \\ \hline
-	\FO & NP-complete & NP-complete & NP-complete & NP-complete \\ \hline
+	\FO & - & NP-complete & NP-complete & NP-complete \\ \hline
 	\PO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
@@
 \item There may be also an unbound number of missing actions in the plan trace because of partial observability.
-The common assumption of having \FO action sequences in a learning task is rather unrealistic in many domains because it implies the existence of human observers that annotate the observed action sequences.
+The common assumption of having \FO action sequences in a learning task is unrealistic in many domains because it implies the existence of human observers that annotate the observed action sequences.
 In some real-world applications, the observed and collected data are sensory data (e.g., home automation, robotics) or images (e.g.
@@
 
-Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists in counting the missing and extra fluents that appear in the learned model wrt the GTM and normalizing by the the total number of all the possible preconditions and effects of an action model.
-This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual model.
-Hence, because the set of preconditions and effects of a model is usually small compared to the set of all possible preconditions and effects, it turns out the metric may output an error rate below 100\% for a totally wrong learned model.
+Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists of (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
+This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM model.
+Hence, because the set of preconditions and effects of the GTM model is usually smaller than the set of all possible preconditions and effects, it turns out the metric may output error rates just slightly below 100\% for totally wrong learned models.
 
-In order to overcome this limitation, we propose to use two standard metrics, {\em precision} and {\em recall}.
+In order to overcome this limitation, we propose to use two standard metrics from ML, {\em precision} and {\em recall}.
 These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
@@
 
-Finally, we would also like to point out that, as shown in section \ref{experiments}, \FAMA is exhaustively evaluated over a wide range of domains (15 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner.
+Finally, we would also like to point out that, as shown in section \ref{experiments}, \FAMA is exhaustively evaluated over a wide range of domains (15 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner so it can benefit straightforward from the last advances in classical planning.

### 2018-06-11 doc/aij18/introduction.tex
updates to the learning task

@@
 Section~\ref{sec:motivation} motivates our compilation approach for learning of \strips\ action models from observations of plan executions.
-Section~\ref{sec:section3} introduces the classical planning model and the \strips\ action model (the target of our learning and evaluation tasks).
-Section~\ref{sec:Section4} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.
+Section~\ref{sec:learning} introduces the \strips\ action model (the target of our learning and evaluation tasks) and formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.
 Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks its extension to evaluate and recognize \strips\ action models.

### 2018-06-11 doc/aij18/learning_task.tex
updates to the learning task

@@
 \section{Learning task}
-\label{learning_task}
+\label{sec:learning}
+Here we formalize the \strips\ action model and the task of learning \strips\ action models from observations of plan executions.
 
@@
 \item $\Psi$ is the set of predicates that define the abstract state space of that planning frame. 
-\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace that contains the sequence of {\em state observations} obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applied in state $s_{i-1}$ and generates the successor state $s_i$.
+\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace that contains the sequence of {\em state observations} obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$. 
 \end{itemize}
@@
 Likewise $\Psi$ can be inferred from $\mathcal{T}$ when at least the observation of a state $s\in \mathcal{T}$ is a full state, that is $|s|=|F|$.
-
-When all the elements in $\mathcal{T}$ are fully observed, i.e.
-learning from plans where the {\em pre-} and {\em post-states} of every action are available, solving the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is straightforward~\cite{jimenez2012review}:
-\begin{itemize}
-  \item {\em Preconditions} are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action, that is any action that belongs to the same operator scheme.
-  \item {\em Effects} are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-\end{itemize}
 
@@
 The definition of the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is also extensible to the more general case where the execution of $\tau$ plans is observed.
-In this case, $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ where each $t\in \mathcal{T}$ is a plan trace $t=\tup{s_0^t,a,_1^t,s_1^t,\ldots,a_n^t,s_{n}^t}$ such that, for each {\small $1\leq i\leq n^t$}, $a_i^t$ is applied in $s_{i-1}^t$ and generates the successor state $s_i^t$.
+In this case, $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ where each $t\in \mathcal{T}$ is a plan trace $t=\tup{s_0^t,a,_1^t,s_1^t,\ldots,a_n^t,s_{n}^t}$ that satisfies the previous [2-4] assumptions.

### 2018-06-11 doc/aij18/motivation.tex
updates to the learning task

@@
 Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning techniques, as our proposal of compiling the learning task to a planning problem.
+
+When the plan trace is fully observed, learning \strips\ action models is straightforward~\cite{jimenez2012review}.
+In this case the {\em pre-} and {\em post-states} of every action are available so action {\em effects} are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+Likewise {\em preconditions} are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action.
 

### 2018-06-12 doc/aij18/compilation.tex
updates added to aij18 paper

@@
-\section{Compilation scheme}
+\section{Compilation}
 \label{compilation}
@@
 \item {\bf Validates the action model $\mathcal{M}'$}.
-The solution plan continues with a {\em postfix} that reproduces the given input knowledge (the available observations of the plan executions) with the programmed action model $\mathcal{M}'$.
+The solution plan continues with a {\em postfix} that reproduces the given input knowledge (the available plan traces) with the programmed action model $\mathcal{M}'$.
 \end{enumerate}
@@
 
-\subsection{Learning from from observations of plan executions}
+\subsection{Learning from observations of plan executions}
 Here we formalize the compilation for learning \strips\ action models with classical planning.
-Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$:
+Given a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
 \begin{itemize}
@@
 the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-plans}.
-Formally, $\Omega=\bigcup_{\small s\in\mathcal{T}} obj(s)$, where $obj$ is a function that returns the objects that appear in a given state.
 \item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
@@
 For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
-\item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\mathcal{T}$.
-The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of $\mathcal{T}$.
-\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state observation $s_i\in\mathcal{T}$ where the action model is validated.
+\item Fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in a plan trace $\mathcal{T}$.
+The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ actions of $\mathcal{T}$ that are observed.
+In the example of Figure~\ref{fig:example-plans} four actions were observed $\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$.
+\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{T}$ where the action model is validated.
+In the example of Figure~\ref{fig:example-plans} two states were observed $\tup{s_0,s_4}$.
 \end{itemize}
@@
 
-\item $G_{\Lambda}=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the programmed action model is validated in the state observations $s_i\in\mathcal{T}$.
+\item $G_{\Lambda}=\{at_n\}\cup\{test_m\}$, requires that the programmed action model is validated in all the actions and states observed from the input plan trace $\mathcal{T}$.
 \item $A_{\Lambda}$ comprises three kinds of actions:
@@
 \end{figure}
-The extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\
+When the input plan trace contain observed actions, then the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\
 
-\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{T}$, {\tt\small $1\leq i< n$}.
+\item Actions for {\em validating} the partially observed state $s_j\in\mathcal{T}$, {\tt\small $1\leq j< m$}.
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
-\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\}\cup \{mode_{val}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\neg mode_{val}\}.
 \end{align*}
@@
 \end{itemize}
-This feature of our compilation is beyond the functionality of VAL, the plan validation tool~\cite{howey2004val}, because VAL requires (1) a full plan and (2), a full action model for plan validation.
-On the other hand our $P_{\Lambda}$ compilation can test {\em model validation} with a partial (or even an empty) action model and a partially observed plan trace.
+This validation capacity of our compilation is beyond the functionality of VAL (the plan validation tool~\cite{howey2004val}) because, our $P_{\Lambda}$ compilation can test {\em model validation} with a partial (or even an empty) action model and a partially observed plan trace.
+On the other hand, VAL requires (1) a full plan and (2), a full action model for plan validation.
 
@@
 
-Note that the $P_{\Lambda}$ compilation is flexible to an unbound number of {\em missing states} or {\em missing actions} in the input plan trace $\mathcal{T}$.
-If any reference to the $mode_{val}$ fluent is removed the compilation can learn \strips\ action models when there are missing state observations in $\mathcal{T}$.
-Likewise some of the actions in $\mathcal{T}$ may be unknown so they are not specified in $I_{\Lambda}$.
-In both cases the output planning problem becomes harder since the plan horizon is no longer bound and the classical planner must determine how many {\em apply} actions are necessary between any two state observations, i.e.
-between the application of two {\em validate} actions (classical planning is {\em PSPACE complete} while classical planning with a bound plan horizon is {\em NP}~\cite{rintanen:SATplannning:AI2012}).
+Note that the $P_{\Lambda}$ compilation is flexible to a bound number of {\em missing states} or {\em missing actions} in the input plan trace $\mathcal{T}$.
+In both cases the output planning problem becomes simpler since the plan horizon is bound and the classical planner does not require to determine how many {\em apply} actions are necessary between any two state observations, i.e.
+between the application of two {\em validate} actions.
 
-Last but not least we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$, {\tt\small $1\leq t\leq \tau$}.
-Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$ that belong to the same planning frame (i.e.
+Last but not least we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_{\tau}\}$.
+Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, {\tt\small $1\leq t\leq \tau$}, that belong to the same planning frame (i.e.
 same fluents and actions but different initial states and goals).
 The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t$ and the plan $\pi_t$, and the goals $G_t$ are defined by the state $s_n^t$.
+Finally, the initial state $I_t$ is given by the state $s_0^t$ and the observed actions from the plan $\pi_t$, and the goals $G_t$ are defined by the number of observed actions and states from the corresponding plan trace.
 Addressing the learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ where $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ requires introducing a small modification to our compilation.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_j\}_{1\leq j<t}\cup\{\neg test_j\}_{t\leq j\leq \tau}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{test_t\} \cup \{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t}.
+\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_{j-1}\}\cup \{\neg mode_{prog}\},\\
+\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_t\} \cup \{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t}.
 \end{align*}
@@
 Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
-In addition, $P_{\Lambda}$ is only solvable if fluents $\{test_i\}$, {\small $1\leq i\leq n$} hold at the last reached state.
+In addition, $P_{\Lambda}$ is only solvable if fluents {\tt\small $at_n$} and {\tt\small $test_m$} hold at the last reached state.
 These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{T}$, starting from the corresponding initial state and following the sequence of actions defined by $\mathcal{T}$.
@@
 This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding set of {\em programming} actions.
-\item The length of $\mathcal{T}$, the observed plan execution. The larger $1\leq i\leq n$, the more $\{test_i\}$ fluents and $\{\mathsf{validate_{i}}\}$ actions in $P_{\Lambda}$.
+\item The length of $\mathcal{T}$, the observed plan execution. The larger $1\leq j\leq n$, the more $\{test_j\}$ fluents and $\{\mathsf{validate_{j}}\}$ actions in $P_{\Lambda}$.
 \end{itemize}

### 2018-06-12 doc/aij18/learning_task.tex
updates added to aij18 paper

@@
 \label{sec:learning}
-Here we formalize the \strips\ action model and the task of learning \strips\ action models from observations of plan executions.
+Here we formalize the \strips\ action model and the task of learning a \strips\ action model from observations of plan executions.
 
@@
 Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
+For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the actions with the maximum arity have arity two (the instantiations of the {\small\tt stack} and {\small\tt unstack} schemas).
 
@@
 $F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}, and that defines the elements that can appear in the action schemes.
-In {\em blocksworld} this set contains 11 elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+In {\em blocksworld} this set contains eleven elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
 In more detail, for a given operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in that action schema.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the elements from $F_v$ that involve $v_2$ because {\small\tt pickup($v_1$)} has arity one.
+For instance, $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the elements from $F_v$ that involve $v_2$ because {\small\tt pickup} has arity one.
 
@@
 \begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
+\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$, representing the operator parameters.
 The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
@@
 Given the set of predicates $\Psi$ and the header of the operator schema $\xi$, $2^{2|F_v(\xi)|}$ defines the size of the space of possible \strips\ models for that operator.
-Note that the previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
+The previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
 For the {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while for the {\tt pickup} operator this number is only 1024.
@@
 We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi'$).
-For instance, we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
-Last but not least, two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
+Therefore  we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
+We say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
 
@@
 \begin{itemize}
-\item $\mathcal{M}$ is the set of {\em initial} operator schemas ({\em empty} or {\em partially specified} if some preconditions or effects are known) that define the actions of a given planning frame.
+\item $\mathcal{M}$ is the {\em initial} set of given operator schemas ({\em empty} or {\em partially specified} if some preconditions or effects are known) that induces the actions of a planning frame.
 \item $\Psi$ is the set of predicates that define the abstract state space of that planning frame. 
-\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace that contains the sequence of {\em state observations} obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$. 
+\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$. 
 \end{itemize}
@@
 A {\em solution} to a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the predicates $\Psi$, and the observed plan trace $\mathcal{T}$.
-Note that $\mathcal{M}$ can be inferred from $\mathcal{T}$ provided that the actions in $\mathcal{T}$ are {\em diverse} enough.
-Meaning that $\mathcal{T}$ contains at least one instantiation of every action scheme in $\mathcal{M}$.
-Likewise $\Psi$ can be inferred from $\mathcal{T}$ when at least the observation of a state $s\in \mathcal{T}$ is a full state, that is $|s|=|F|$.
+Note that $\mathcal{M}$ can be inferred from $\mathcal{T}$ provided that the actions in $\mathcal{T}$ are {\em diverse} enough (meaning that $\mathcal{T}$ contains at least one instantiation of every action scheme in $\mathcal{M}$).
+Likewise $\Psi$ can be inferred from $\mathcal{T}$ when at least the observation of a state $s_i\in \mathcal{T}$ is a full state, that is $|s_i|=|F|$.
 
@@
 \begin{enumerate}
-\item Each input action scheme $\xi\in\mathcal{M}$ is at least composed of its $head(\xi)$,
-\item The initial state $s_0\in\mathcal{T}$ is {\em fully observable}.
+\item Each input action scheme $\xi\in\mathcal{M}$ is at least composed of its $head(\xi)$.
+\item The initial state $s_0\in\mathcal{T}$ of the given plan trace is {\em fully observable}, $|s_0|=|F|$.
 \item Intermediate actions $a_i\in\mathcal{T}$ and states $s_i\in\mathcal{T}$ s.t.
-{\small $1\leq i\leq n$}, can be {\em partially observable} (some fluents in $s_i$ are missing because it is unknown whether their value is either positive or negative).
+{\small $1\leq i\leq n$}, are {\em partially observable}.
 In the extreme, all actions $a_i$, {\small $1\leq i\leq n$}, can be missing (unobserved) provided that the final state $s_n$ is at least, partially observed.
-Likewise, all states $s_i$, {\small $1\leq i\leq n$}, can be missing (unobserved) provided that at least actions $a_i$, {\small $1\leq i\leq n$}, are partially observed.
+Some fluents in $s_i$ may be missing because it is unknown whether their value is either positive or negative, $0\leq |s_i|\leq |F|$.
+All states $s_i$, {\small $1\leq i\leq n$}, might be missing (unobserved) provided that at least actions $a_i$, {\small $1\leq i\leq n$}, are partially observed.
  \item Observations in $\mathcal{T}$ are {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
 \end{enumerate}
-Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, that corresponds to observing the execution of a four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
+Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, that corresponds to observing the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
 In this example $\mathcal{T}=\langle s_0,${\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$,s_4\rangle$ which means that only the first and last states are observed and the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
@@
 
-The definition of the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is also extensible to the more general case where the execution of $\tau$ plans is observed.
-In this case, $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ where each $t\in \mathcal{T}$ is a plan trace $t=\tup{s_0^t,a,_1^t,s_1^t,\ldots,a_n^t,s_{n}^t}$ that satisfies the previous [2-4] assumptions.
+The definition of the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is extensible to the more general case where the execution of $\tau$ plans is observed.
+In this case, $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ where each $t\in \mathcal{T}$ is a plan trace $t=\tup{s_0^t,a,_1^t,s_1^t,\ldots,a_n^t,s_{n}^t}$ that satisfies the previous [2,4] assumptions of our learning task.

### 2018-06-12 doc/aij18/compilation.tex
updates added to aij18 paper

@@
 
-\subsection{Tuning the compilation with background knowledge}
+\subsection{Optimizing the compilation with background knowledge}
 A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn logic programs from data~\cite{muggleton1994inductive}.
@@
 
-\subsubsection{Static predicates}
+%\subsubsection{Static predicates}
 A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
@@
 
-\subsubsection{State constraints}
-The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
+%\subsubsection{State constraints}
+%The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
 If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
@@
 
-{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+%{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
 Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
 
-The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
+%The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
 For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}.
@@
 
-\begin{figure}[hbt!]
-  \begin{footnotesize}
-    \begin{center}
-$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
-$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
-$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
-    \end{center}
-\end{footnotesize}
- \caption{\small An example of the strongest invariant for the {\em blocksworld} domain.}
-\label{fig:strongest-invariant}
-\end{figure}
+%\begin{figure}[hbt!]
+%  \begin{footnotesize}
+%    \begin{center}
+%$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
+%$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
+%$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
+%$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
+%$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
+%    \end{center}
+%\end{footnotesize}
+% \caption{\small An example of the strongest invariant for the {\em blocksworld} domain.}
+%\label{fig:strongest-invariant}
+%\end{figure}
 
-A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+%A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
 For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a mutex because $block_A$ can only be on top of a single block.
 
-A {\em domain invariant} is an instance-independent invariant, i.e.
+%A {\em domain invariant} is an instance-independent invariant, i.e.
 holds for any possible initial state and set of objects.
@@
 
-An interesting contribution of our compilation is that the validation of an action model can also be done with {\em state constraints}.
+%An interesting contribution of our compilation is that the validation of an action model can also be done with {\em state constraints}.
 Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check also that the learned model satisfy a constraint $\phi\in\Phi$:
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
-\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
-\end{align*}
-\end{small}
+%\begin{small}
+%\begin{align*}
+%\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
+%\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
+%\end{align*}
+%\end{small}
 
-This redefinition of validate actions applies also to {\em trajectory constraints} because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.
+%This redefinition of validate actions applies also to {\em trajectory constraints} because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.
 
-Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
+%Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
 Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space.
@@
 
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
-\cond(\mathsf{programPre_{f_1,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
-\end{align*}
-\end{small}
+%\begin{small}
+%\begin{align*}
+%\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
+%\cond(\mathsf{programPre_{f_1,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
+%\end{align*}
+%\end{small}

### 2018-06-12 doc/aij18/evaluation.tex
updates added to aij18 paper

@@
 
-If the actual model is available, the quality of a learned action model is quantifiable using the {\em precision} and {\em recall} metrics.
-{\em Precision} and {\em recall} are two syntactic metrics frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
+If the actual GTM model is available then the quality of a learned action model is quantifiable using the ML metrics, {\em precision} and {\em recall}.
+{\em Precision} and {\em recall} are two syntactic metrics frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative than counting the number of errors between the learned and the reference model~\cite{davis2006relationship}.
 Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
 \begin{itemize}
-\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
+\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (in our case, predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
 \item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing).
@@
 
-The roles of two {\em comparable} action schemes, or the roles of two action parameters that belong to the same type, can be swapped by a $\Lambda$ learning task when the observed input data, given in $\mathcal{T}$, is insufficient.
-For instance, when $\mathcal{T}$ does not contain any intermediate action the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa, or the parameters of the {\small\tt stack} operator can be swapped.
-Pure syntax-based evaluation metrics (like {\em precision} and {\em recall}) can report low scores for learned models that are actually good but correspond to {\em reformulations} of the actual model; i.e.
+In a $\Lambda$ learning task, the roles of two {\em comparable} action schemes (or the roles of two action parameters with the same type) can be swapped.
+These role swaps tipically happen when the observed input data, given in $\mathcal{T}$, is scarce.
+For instance when $\mathcal{T}$ does not contain any intermediate action, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
+Further, the roles of parameters of the {\small\tt stack} (or the {\small\tt unstack}) operator could be swapped and these learned models would still be sematically correct with respect to the given input observations.
+
+Pure syntax-based evaluation metrics (like {\em precision} and {\em recall}) can report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the actual model; i.e.
 a learned model semantically equivalent but syntactically different to the reference model.
-
 Here we introduce a novel evaluation approach that is robust to role changes of this particular kind.
@@
 
-Note that the number of deletions indicate the number of {\em false positives} while the number of insertions represents the number of {\em false negatives}.
-As explained, these numbers can be normalized with the amount of {\em true positives} to compute the {\em precision} and {\em recall} of a learned model with respect to a reference model.
-Normally evaluating a given learned domain with respect to the actual generative model is not possible because the actual model is not available.
-As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions~\cite{yang2007learning}.
-With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$ such that, for each {\small $1\leq i\leq n$}, $a_i$ is applied in $s_{i-1}$ and generates the successor state $s_i$.
+Note that the number of deletions indicate the number of {\em false positives} while the number of insertions corresponds to the number of {\em false negatives}.
+As explained, these numbers can be normalized with the amount of {\em true positives} to compute the {\em precision} and {\em recall} of a given learned model with respect to a reference model.
+
+Normally evaluating a given learned domain with respect to the actual GTM model is not possible because the actual GTM model is unknown.
+As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions that are assumed to be generated with that model~\cite{yang2007learning}.
+With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a partially observed plan trace $\mathcal{T}$.
 
@@
 
-The {\em observation edit distance} is mappable into a likelihood with the following expression $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
-Further, following the previous correspondances ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can redefine a {\em semantic} version of the  {\em precision} and {\em recall} metrics of a learned model with respect to a partially observed plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
-In this case, for each action schema $\xi\in\mathcal{M}$, the corresponding number of {\em true positives} is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$.
+Following the previous correspondances ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can define a {\em semantic} version of the {\em precision} and {\em recall} metrics of a learned model.
+In this case {\em precision} and {\em recall} is not computed with respect to a GTM but with respect to a partially observed plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
+To define the {\em semantic} version of {\em precision} and {\em recall} the only term to reformalize is the number of {\em true positives} that, for each action schema $\xi\in\mathcal{M}$, is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$.
 

### 2018-06-12 doc/aij18/aij18.tex
Background and Motivation completed
Background and Motivation completed
updates added to aij18 paper

@@
 
-\title{Learning and Evaluation of \strips\ Action Models with Classical Planning}
+\title{A general framework for learning planning action models}
 \author[label1]{Diego Aineto}
@@
 
+In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema.
+However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}
+
 \subsection*{Acknowledgment}

### 2018-06-12 doc/aij18/background.tex
updates added to aij18 paper

@@
 
-The Action-Relation Modelling System (\textbf{\ARMS}) \cite{yang2007learning} is one of the first learning algorithms able to learn from plan traces with partial or null observations of intermediate states.
+The Action-Relation Modeling System (\textbf{\ARMS}) \cite{yang2007learning} is one of the first learning algorithms able to learn from plan traces with partial or null observations of intermediate states.
 \ARMS uncovers a number of constraints from the plan traces in the training data that must hold for the plans to be correct.
@@
 \AMAN uses the same metric as \LAMP, namely counting the number of preconditions and effects that appear in the learned model and not in the ground-truth model (extra fluents) and viceversa (missing fluents).
-In a comparison between \AMAN and \ARMS on noiseless inputs, the results show that the accuracies of both \AMAN and \ARMS are very close to each other and neither dominates the other.
+In a comparison between \AMAN and \ARMS on noiseless inputs, the results show that the accuracy performance of both \AMAN and \ARMS are very close to each other and neither dominates the other.
 The convergence property of \AMAN guarantees that the accuracy of the learned model with noisy input traces becomes more and more close to the case  \emph{without noise} because the distribution of noise in the plan becomes gradually closer to real distribution with the number of iterations.
@@
 Plan traces in \CAMA are composed of 80\% of empty states and each partial state was selected by 50\% of propositions in the corresponding full state.
-An experimental comparison reveals that a manual crowdsourcing of \CAMA outperforms \ARMS and that as expectedly the difference becomes smaller as the number of plan traces becomes larger.
+An experimental comparison reveals that a manual crowdsourcing of \CAMA outperforms \ARMS and that as expected the difference becomes smaller as the number of plan traces becomes larger.
 The accuracy of \CAMA for a small number of plan traces (e.g., 30) is not less than 80\%, thus revealing that exploiting the knowledge of the crowd can help learning action models.

### 2018-06-12 doc/aij18/evaluation.tex
updates added to aij18 paper

@@
 In a $\Lambda$ learning task, the roles of two {\em comparable} action schemes (or the roles of two action parameters with the same type) can be swapped.
-These role swaps tipically happen when the observed input data, given in $\mathcal{T}$, is scarce.
+These role swaps typically happen when the observed input data, given in $\mathcal{T}$, is scarce.
 For instance when $\mathcal{T}$ does not contain any intermediate action, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
-Further, the roles of parameters of the {\small\tt stack} (or the {\small\tt unstack}) operator could be swapped and these learned models would still be sematically correct with respect to the given input observations.
+Further, the roles of parameters of the {\small\tt stack} (or the {\small\tt unstack}) operator could be swapped and these learned models would still be semantically correct with respect to the given input observations.
 
@@
 
-Following the previous correspondances ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can define a {\em semantic} version of the {\em precision} and {\em recall} metrics of a learned model.
+Following the previous correspondences ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can define a {\em semantic} version of the {\em precision} and {\em recall} metrics of a learned model.
 In this case {\em precision} and {\em recall} is not computed with respect to a GTM but with respect to a partially observed plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
-To define the {\em semantic} version of {\em precision} and {\em recall} the only term to reformalize is the number of {\em true positives} that, for each action schema $\xi\in\mathcal{M}$, is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$.
+To define the {\em semantic} version of {\em precision} and {\em recall} the only term to re-formalize is the number of {\em true positives} that, for each action schema $\xi\in\mathcal{M}$, is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$.
 
 \subsection{Computing the edit distance with classical planning}
-Our compilation is extensible to compute the {\em observation edit distance} by simply considering that the input \strips\ model $\mathcal{M}$, given in a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, is {\em non-empty}.
+Our compilation is extensible to compute the {\em observation edit distance} and hence, the semantic versions of the {\em precision} and {\em recall} metrics.
+This extension considers that the input \strips\ model $\mathcal{M}$, is {\em non-empty} so instead of learning an action model from scratch we simply edit $\mathcal{M}$ until it satisfies the given input observations.
 In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-A solution to the planning task resulting from the extended compilation is a sequence of actions that:
-
+A solution to the classical planning task resulting from the extended compilation is a sequence of actions that:
 \begin{enumerate}
@@
 A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema.
-However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}.
 \item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:Section5} for the models that are programmed from scratch.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:compilation} for the models that are programmed from scratch.
 \end{enumerate}
 
-Now $\Lambda$ does not formalize a learning task but the task of editing $\mathcal{M}$ to produce the plan trace $\mathcal{T}$, which results in the edited model $\mathcal{M}'$.
-The output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda}',G_{\Lambda}}$:
-
+In more detail given $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, the output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda},G_{\Lambda}}$ such that:
 \begin{itemize}
@@
 
-Our interest when computing the {\em observation edit distance} is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deleitions) for that $\mathcal{M}'$ is validated in the given observations, e.g.
-$\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}.
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:Section3}).
+Our interest when solving the classical planning task that is output by our extended compilation is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deletions).
+For instance, in the example of Figure~\ref{fig:plan-pdistance}, $\mathcal{M}'$ is validated in the given observations after inserting two new positive effects to the initial action model, e.g.
+$\delta(\mathcal{M},\mathcal{T})=2$ .
+In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:learning}).
 The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
@@
 The sets of fluents $F_{\Lambda}$ and $I_{\Lambda}$ are defined like in the previous compilation.
-With respect to the actions, $A_{\Lambda}$ again implement the two {\em edit operations} (i.e.
+With respect to the actions, $A_{\Lambda}'$ again implement the two {\em edit operations} (i.e.
 include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect) but does not contain $\mathsf{apply_{\xi,\omega}}$ or $\mathsf{validate_{i}}$ actions because the \strips\ action model are not validated in any observation of plan executions.

### 2018-06-12 doc/aij18/motivation.tex
minor changes

@@
 The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging.
-This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \leq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now.
+This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now.
 
@@
 
-Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists of (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
+Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
 This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM model.
-Hence, because the set of preconditions and effects of the GTM model is usually smaller than the set of all possible preconditions and effects, it turns out the metric may output error rates just slightly below 100\% for totally wrong learned models.
+Hence, because the set of preconditions and effects of the GTM model is usually smaller than the set of all possible preconditions and effects, it turns out the metric may output error rates below 100\% for totally wrong learned models.
 
 In order to overcome this limitation, we propose to use two standard metrics from ML, {\em precision} and {\em recall}.
-These two metrics are frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
+These two metrics are frequently used in pattern recognition, information retrieval and binary classification and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
 

### 2018-06-14 doc/programming-project/AI-programming.tex
updates added to project proposal

--- 2018-06-07 doc/programming-project/AI-programming.tex
@@
 \begin{abstract}
-  Programming is still a manual handcraft task, even when the complexity of programs is not high enough to justify the cost of the human workforce to produce them.
-The need to automate programming tasks increases every day: Programming errors, commonly known as {\em bugs}, cause undesired software behavior making programs crash or enabling malicious users to access private data.
+  Programming is nowadays a manual handcraft task however, the need to automate programming tasks increases every day: Programming errors, commonly known as {\em bugs}, cause undesired software behavior making programs crash or enabling malicious users to access private data.
 Just for 2017, the cumulative cost of software bugs is worldwide estimated in more than one trillion US dollars.
 
-  This research project investigates a novel approach for software development, that leverages Artificial Intelligence, to increase the automation of the programming process and hence, reduce the chances of introducing software bugs.
+  This research project investigates a novel approach for software development, that leverages {\em Artificial Intelligence} (AI), to increase the automation of the programming process and hence, reduce the chances of introducing software bugs.
 The project proposes to address {\bf automatic program synthesis and validation starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine}.
-Our current work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI}, the main international conference on Artificial Intelligence.
+Our current work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on Artificial Intelligence).
 \end{abstract}
@@
 In the last 10 years, the GRPS has led three research projects of the Spanish national research plan on {\em AI planning}.
-Further, the GRPS has developed its own AI planner, that is participating  at the International Planning Competition (IPC-2018, {\tt https://ipc2018.bitbucket.io}).
+Further, the GRPS has developed its own AI planner, that is participating  at the International Planning Competition (IPC-2018, {\tt http://ipc2018.bitbucket.io}).
 
-The research group for this project comprises three researchers: 1 principal investigator, 1 senior researcher and 1 PhD student. All of them are GRPS members.
+The research group for developing this project would comprise three researchers: 1 principal investigator, 1 senior researcher and 1 PhD student. All of them are GRPS members.
 
@@
 
-\underline{General quality indicators of scientific research}: {\bf 381 citations} (295 since 2013) and {\bf h-index 11}. {\scriptsize\em Source google scholar}.
+General quality indicators of scientific research: {\bf 382 citations} (296 since 2013) and {\bf h-index 11}. {\scriptsize\em Source google scholar}.
 
@@
 
-Eva Onaindía has supervised so far a total of nine PhD theses and she currently supervising four doctoral students.
+Eva Onaindía has supervised so far a total of nine PhD theses and she is currently supervising four doctoral students.
 She has also participated in many diverse scientific evaluation committees at both national and international level.
@@
 
-\underline{General quality indicators of scientific research}: {\bf 1532 citations} (989 since 2013) and {\bf h-index 22}. {\scriptsize\em Source google scholar}.
+General quality indicators of scientific research: {\bf 1536 citations} (992 since 2013) and {\bf h-index 22}. {\scriptsize\em Source google scholar}.
 
 \subsection{Researcher in training period: Diego Aineto}
-Diego Aineto is a PhD student at the {\em Department of Computer Systems and Computation} of the UPV.
+Diego Aineto is a PhD student at the {\em Department of Computer Systems and Computation} of the Universitat Politècnica of València.
 
 Since 2016 Diego is holding a four-year FPU grant on {\em Causal discovery for predictive analysis} (FPU16/03184).
-Diego is also involved on the current research project of the GPRS from the Spanish national research plan on {\em activity recognition}.
+Diego is also involved on the current research project TIN2017-88476-C2-1-R, funded by the Spanish national research plan and with title, {\em Reconocimiento de actividades y planificación automática para el diseño de asistentes inteligentes}.
 
@@
 This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} paradigm for the automatic synthesis and validation of programs}.
-Here we briefly introduce the technology we will rely on along the project:
+Here we briefly introduce this technology:
 \begin{itemize}
@@
 {\em Program synthesis} is the task of computing a program that satisfies a given formal specification.
-On the other hand, {\em program validation} is the task of proving, or disproving, the correctness of a given program with respect to a certain formal specification or property.
-Program validation is considered a necessary step for program synthesis.
-In 2008 the PhD Thesis work by Armando Solar-Lezama, at the University of California Berkeley, showed that it is possible to encode program synthesis problems in Boolean logic and therefore, use SAT algorithms to automatically compute programs~\cite{lezama2008program}.
+In 2008 the PhD Thesis work by Armando Solar-Lezama, at the University of California Berkeley, showed that it is possible to encode program synthesis problems in Boolean logic and therefore, use {\em satisfiability modulo theories}~\cite{barrett:SMT:2009} to automatically compute programs~\cite{lezama2008program}.
 Since then, there has been a surge of practical interest in the idea of program synthesis in the formal verification community and related fields~\cite{alur2013syntax}.
-To illustrate this, in 2013, a unified framework for program synthesis problems was proposed and since 2014 there is a yearly program synthesis competition, comparing the different algorithms for program synthesis in a competitive event ({\tt www.sygus.org}).
+To illustrate this, in 2013, a unified framework for program synthesis problems was proposed and since 2014 there is a yearly program synthesis competition, comparing the different algorithms for program synthesis in a competitive event ({\tt http://www.sygus.org}).
+Further the US National Science Foundation is funding the ExCAPE research project\footnote{\tt https://www.nsf.gov/awardsearch/showAward?AWD\textunderscore ID=1138996}, to transform the way programmers develop software by advancing the theory and practice of software synthesis (since 2012 the ExCAPE research project is funded with a total amount of $\$3,750,000$).
 
@@
 
-\item In {\bf Programming by Sketching (PbS)} programmers provide a partially specified program, i.e.~a program that expresses the high-level structure of an implementation but that leaves low level details undefined to be determined by the synthesizer~\cite{solar2006combinatorial}.
+\item In {\bf Programming by Sketching (PbS)} programmers provide a partially specified program, i.e.~a program with the high-level structure of an implementation but that leaves low level details undefined to be determined by the synthesizer~\cite{solar2006combinatorial}.
 This form of program synthesis relies on a programming language called {\sc SKETCH}, for sketching partial programs.
@@
 
-Closely related to the aims of the project is also the work on the synthesis of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
+On the other hand, {\em program validation} is the task of proving, or disproving, the correctness of a given program with respect to a certain formal specification or property.
+Program validation is considered a necessary step for program synthesis.
+{\em Model checking} is the mainstream AI approach for the formal validation of programs and controllers~\cite{clarke1999model}.
+In model checking a given model of a system (the program or controller to validate), is exhaustively and automatically checked to verify  whether this model meets a given specification.
+Typically the specification contains safety requirements such as the absence of deadlocks and similar critical states that can cause the system to crash.
+Both the model of the system and the specification are formulated in a formal language.
+Current approaches for model checking reduces to graph search.
+Instead of enumerating reachable states one at a time, the state space can sometimes be traversed more efficiently by considering large numbers of states at a single step.
+For instance, representing set of states and transition relations as logical formulas or binary decision diagrams, like in {\em symbolic model-checking} ~\cite{mcmillan1993symbolic}.
+
+Closely related to the aims of the project is the work on the automated synthesis of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
 The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}.
-To keep the computation of FSCs tractable, they limit the space of possible solutions bounding the maximum size of the FSC.
+To keep the computation of FSCs tractable, the space of possible solutions is bound by the maximum size of the FSCs.
 The computation of FSCs includes works that compile this task into another forms of problem solving so they benefit from the last advances on off-the-shelf solvers (e.g.
 {\em classical planning}~\cite{sergio:aprograming:icaps16}, {\em conformant planning}~\cite{Geffner:FSM:AAAI10}, {\em CSP}~\cite{Infantes:FSC:ECAI2010} or a {\em Prolog program}~\cite{Giacomo:FSM:ICAPS13}).
-
-Finally the generation of programs from examples is a research question also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
-ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn a given target concept, expressed as a  logic program,  from  examples  and  background  knowledge, that are expressed as logic facts.
+The validation and synthesis of programs from examples are research questions also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
+ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn logic programs from  examples and background knowledge, that are expressed as logic facts.
 
@@
 \subsection{Synthesis and validation of TDD programs as AI planning}
-Given a {\em TDD programming} task our approach is modeling and solving it as it were an {\em AI planning} task.
+Given a {\em TDD programming} task, our approach is modeling and solving it as it were an {\em AI planning} task.
 
@@
 
-The {\bf initial/goal states} of the AIP task encode the initial/final values of the program variables that are given by the input/output tests of the TDD programming task.
+The {\bf initial/goal states} of the AIP task encode the initial/final values of the program variables and are given by the input/output tests of the TDD programming task.
 
@@
 \begin{itemize}
-\item {\it Programming actions}, that set an instruction on a given program line.
-\item {\it Execution actions}, execute the instruction set on a given program line.
+\item {\it Programming actions}, that assign an instruction to a given program line.
+\item {\it Execution actions}, execute the instruction assigned to the current program line.
 \end{itemize}
-We implement this encoding using standard planning languages, such as PDDL~\cite{fox2003pddl2}, so the AIP tasks resulting from our encoding can be solved with off-the-shelf planners, like {\sc FD}~\cite{helmert2006fast}.
+We implement this encoding using standard planning languages, such as PDDL~\cite{fox2003pddl2}, so the AIP tasks resulting from our encoding can be solved with off-the-shelf planners, like the {\sc FD} planning system~\cite{helmert2006fast}.
 The programs synthesized with this approach are guaranteed to be bug-free over given sets of {\em input-output} tests cases.
@@
 \begin{itemize}
-\item {\em Real-world benchmarks}: The GRPS group is currently leading a four-year research project from the {\em Spanish national plan} in which {\em activity recognition} is applied to different real-world domains such as {\em domotics}, {\em tourism}, {\em traffic control} and {\em robotics}.
+\item {\em Real-world benchmarks}: The GRPS group is currently leading the four-year research project TIN2017-88476-C2-1-R from the {\em Spanish national plan} in which {\em AI Planning} and {\em activity recognition} is applied to different real-world domains such as {\em domotics}, {\em tourism}, {\em traffic control} and {\em robotics}.
 Interestingly many activities in these domains can be understood as simple programs.
@@
 Figure~\ref{fig:gantt} details the proposed 24-month timeline for the project.
+A deliverable is provided at the end of each task.
 
@@
     \begin{enumerate}
-    \item Design of the user-interaction for the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases plus the available instruction set. 
+    \item Design of the test-case specification. Programming tasks are specified as a set of {\em input-output} tests cases plus the available instruction set. 
     \item Experimental design. Experiments will comprise taking time and memory measurements to evaluate the resources required by our approach to solve the given TDD programming/validation tasks.
-      \item Evaluation of different AI planners available, with special attention to the planners that get the best results at the IPC-2018. 
+    \item Evaluation of the different AI planners available, with special attention to the planners that get the best results at the IPC-2018. 
       \end{enumerate}
@@
       \item The programming-into-planning compiler ({\em Compiler 1}).
-This component parses the {\em TDD programming task} and produces an {\em AIP task} encoded in the standard planning language PDDL.
+This system component parses the {\em TDD programming task} and produces an {\em AIP task} encoded in the standard planning language PDDL.
       \item The plan-into-program compiler ({\em Compiler 2}).
-This component extracts the program code and the corresponding validation certificate from the solution plan produced by an off-the-shelf AI planner.
+This system component extracts the program code and the corresponding validation certificate from the solution plan produced by an off-the-shelf AI planner.
       \end{enumerate}
@@
 
-{\small{\bf\em Deliverable T2:} Open repository with the source code of the system architecture.}
+{\small{\bf\em Deliverable T2:} Open repository with the source code of the system architecture and the corresponding benchmarks.}
 
@@
 
-In addition, we plan to hire a master student to assist in the completion of two tasks, {\bf T3(a,i)} and {\bf T3(a,iv)}. His detailed responsability will be:
+In addition, we plan to hire a master student for 6 months to assist in the completion of two tasks, {\bf T3(a,i)} and {\bf T3(a,iv)}. The detailed responsability of the student will be:
 \begin{enumerate}
@@
 
-The objective of this project is to study the performance of our approach for the automated synthesis and validation of programs that are characterized by these tree dimensions:
+The specific objective of this project is to study the performance of our AIP approach for the automated synthesis and validation of programs that are characterized by these tree dimensions:
 \begin{enumerate}
@@
 
-For instance, the program of Figure~\ref{fig:list} is generated using three program lines, one observable variable ($\mathsf{isEnd}(n)$ that has Boolean domain and indicates when $n$ points to the end of the linked list), and a instruction set that comprises two instructions ($\mathsf{visit}(n)$ that marks the list node assigned to $n$ as {\em visited} and $\mathsf{step}(n)$ that advances $n$ to the next node in the linked list).
+For instance the program of Figure~\ref{fig:list}, for traversing linked lists, is generated using three program lines, one observable variable ($\mathsf{isEnd}(n)$ that has Boolean domain and indicates when $n$ points to the end of the given linked list), and an instruction set that comprises two instructions ($\mathsf{visit}(n)$, that marks the list node assigned to $n$ as {\em visited} and $\mathsf{step}(n)$, that advances $n$ to the next node in the linked list).
 
 In more detail, the two specific objectives of this project are:
-\begin{itemize}
+\begin{enumerate}
 \item To study the performance of our Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
 \item To study the performance of our Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}.
-\end{itemize}
+\end{enumerate}
 In both cases the performance of our approach will be evaluated with regard to the {\bf computation time} and {\bf memory} invested in the synthesize and the validation of the aimed programs.
 
-Despite setting bounds for the programs size and kind, note that challenging programming tasks can be addressed using {\em problem decomposition}.
+Despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}.
 With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.
@@
 
-In fact, our current work on {\em program synthesis} with AIP already produced {\bf publications at top international conferences on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
+In fact, our current work on {\em program synthesis} with AIP already produced several {\bf publications at top international conferences on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
 
+The main benefit of this project is to provide new insights into the current understanding of how AI can assist programmers in the software development.
 In more detail, the expected benefits for this particular research project are four-fold:
 \begin{enumerate}
-\item The definition of {\bf effective metrics for assessing how well a given program covers a set of {\em input-output} test cases}.
-These metrics can provide new insights into the current understanding of how Artificial Intelligence can assist programmers in the software development.
+\item A new {\bf evaluation methodology for the {\em Synthesis and Validation of Programs}}.
+The application of the exiting AIP technology to program synthesis can provide new evaluation metrics that assess how well a program covers a set of {\em input-output} test cases.
 \item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}.
@@
 Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition~\cite{vallati:IPC:AI15}.
-This project will also help to meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks.
-\item The {\bf development of open software and benchmarks for the {\em Synthesis and Validation of Programs}}.
+This project will help to meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks.
+\item The {\bf development of open software and open benchmarks for the {\em Synthesis and Validation of Programs}}.
 We strongly belief that reproducibility and open knowledge are essential to the advance of the research on computer science.
-With this regard we plan to develop a {\em github} repository where we make available the developed source code and benchmarks.
-\item International {\bf dissemination of the obtained scientific results}.
-The scientific results obtained during the development of the project will also be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).
+With this regard, we plan to develop a {\em github} repository where we make available the developed source code and benchmarks.
+\item {\bf International dissemination of the obtained scientific results}.
+The scientific results obtained during the development of the project will be submitted to top Artificial Intelligence conferences (such as IJCAI, AAAI, ICML and ICAPS) and to the main journals in the AI field (such as AIJ, JMLR and JAIR).
 \end{enumerate}
@@
 
-\begin{tiny}
+\begin{scriptsize}
 \bibliography{AI-programming}
-\end{tiny}
+\end{scriptsize}
 \bibliographystyle{ieeetr}

### 2018-06-14 doc/aij18/background.tex

@@
 \AMAN uses the same metric as \LAMP, namely counting the number of preconditions and effects that appear in the learned model and not in the ground-truth model (extra fluents) and viceversa (missing fluents).
-In a comparison between \AMAN and \ARMS on noiseless inputs, the results show that the accuracy performance of both \AMAN and \ARMS are very close to each other and neither dominates the other.
+In a comparison between \AMAN and \ARMS on noiseless inputs, the results show that the accuracy of the learnt models are very close to each other and neither dominates the other.
 The convergence property of \AMAN guarantees that the accuracy of the learned model with noisy input traces becomes more and more close to the case  \emph{without noise} because the distribution of noise in the plan becomes gradually closer to real distribution with the number of iterations.

### 2018-06-18 doc/aij18/background.tex
updates added to learning section

@@
 
+Like in PDDL, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
+Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
+Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
+$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+
 A {\em state} $s$ is a full assignment of values to fluents, i.e.
@@
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. An action $a\in A$ is defined with:
-
-%{\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$ and {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$.
-
 \begin{itemize}
@@
 \end{itemize}
-
+We assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$ and that actions $a\in A$ are instantiated from given action schemas.
 We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.

### 2018-06-18 doc/aij18/learning_task.tex
updates added to learning section

@@
-\section{Learning task}
+\section{Learning action models from plan executions}
 \label{sec:learning}
-Here we formalize the \strips\ action model and the task of learning a \strips\ action model from observations of plan executions.
+The {\em learning task} addressed in this paper corresponds to learning a classical planning action model by observing an agent(s) acting in the world.
+This learning task is formalized by the pair $\Lambda=\tup{\mathcal{M},\tau}$:
+\begin{itemize}
+\item $\mathcal{M}$ is the {\bf initial action model}. This model is {\em empty}, when learning from scratch, or {\em partially specified}, when some fragments of the model are a priori known.
+\item $\tau$ is the observed {\bf plan trace} such that:
+\begin{enumerate}
+ \item Observations in $\tau$ are {\em noiseless}, meaning that if the value of a fluent or an action is observed, then the observation is correct. 
+\item The initial state $s_0\in\tau$ is {\em fully observed}, i.e.~$|s_0|=|F|$.
+This means that the set of fluents $F$, and the corresponding set of predicates $\Psi$ that shapes the fluents in $F$, are inferrable from $s_0$.
+\item The set of actions $A$ is inferrable either from $\mathcal{M}$ or $\tau$.
+This means that the action name and parameters that shape the actions in $A$ are given by $\mathcal{M}$ otherwise, $\tau$ contains at least one instantiation of every action model in $\mathcal{M}$.
+\item Intermediate actions $a_i\in\tau$, and intermediate states $s_i\in\tau$, {\small $1\leq i$}, are {\em partially observable}.
+In the extreme all $a_i\in\tau$, {\small $1\leq i\leq n$} actions could be unobserved, provided that the final state $s_n$ is at least, partially observed.
+Likewise states $s_i$, {\small $1\leq i\leq n$}, might be unobserved provided that at least actions $a_i$, {\small $1\leq i\leq n$}, are partially observed.
 
-\subsection{\strips\ action schemas}
+\end{enumerate}
+\end{itemize}
+
+A {\em solution} to a $\Lambda=\tup{\mathcal{M},\tau}$ learning task is an action model $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$ and the observed plan trace $\tau$.
+
+Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\tau}$, that corresponds to observing the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
+In this example $\tau=\langle s_0,${\small\tt (putdown\ B), (stack\ A\ B)}, $s_4\rangle$ which means that only the first and last states are observed (the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown) and that actions $a_1$ and $a_4$ are also unknown.
+
+\begin{figure}[hbt!]
+{\footnotesize\tt ;;;;;; Action headers in $\mathcal{M}$}  
+\begin{footnotesize}  
+\begin{verbatim}
+(pickup ?v1) (putdown ?v1) (stack ?v1 ?v2} (unstack ?v1 ?v2)
+\end{verbatim}
+\end{footnotesize}
+\vspace{0.2cm}
+{\footnotesize\tt ;;;;;; Plan trace $\tau$}
+\begin{footnotesize}
+\begin{verbatim}
+;;; Initial state observation
+(clear B) (ontable A) (handempty) (on B A)
+(not (clear A)) (not (ontable B)) (not (holding A)) (not (holding B)) 
+(not (on A A)) (not (on A B)) (not (on B B))
+\end{verbatim}
+\end{footnotesize}
+
+\begin{footnotesize}
+\begin{verbatim}
+;;; State observation 
+(clear A) (on A B) (ontable B) (handempty)
+\end{verbatim}
+\end{footnotesize}
+
+\begin{footnotesize}
+\begin{verbatim}
+;;; Actions observation
+(putdown B) (stack A B)
+\end{verbatim}
+\end{footnotesize}
+
+ \caption{\small Task $\Lambda=\tup{\mathcal{M},\tau}$ for learning a {\em blocksworld} \strips\ action model from a four-action plan and two state observations.}
+\label{fig:example-plans}
+\end{figure}
+
+The definition of the learning task is extensible to the more general case where the execution of several plans generated from the same action model is observed.
+In this case, $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_{k}\}$ such that each $\tau\in \mathcal{T}$ is a plan trace that satisfies the previous [1,4] assumptions.
+In this case a {\em solution} action model $\mathcal{M}'$ have to be compliant with the input model $\mathcal{M}$ but also with every observed plan trace $\tau\in \mathcal{T}$.
+
+\subsection{A propositional encoding for \strips\ action schemas}
 This work addresses the learning and evaluation of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+An \strips\ action schema is a tuple $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$:
+\begin{itemize}
+\item The name, $name(\xi)$, and parameters, $pars(\xi)$, of the schema that define the {\em header} of that schema. 
+\item The preconditions $pre(\xi)\subseteq \Psi$, the negative effects $del(\xi)\subseteq \Psi$, and the positive effects $add(\xi)\subseteq \Psi$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\end{itemize}
 Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
@@
 \end{footnotesize}
- \caption{\small \strips\ operator schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
+ \caption{\small \strips\ action schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
 \label{fig:stack}
 \end{figure}
+We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi'$).
+Therefore  we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
+We say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
 
-To formalize the target of the learning and evaluation tasks, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$, as in PDDL.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+Now we formalize a propositional encoding for the target of the learning task addressed in the paper, the \strips\ action model.
+This encoding is the core of our compilation approach for learning \strips\ action models from observations of plan executions.
 
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
+First, let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
 For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the actions with the maximum arity have arity two (the instantiations of the {\small\tt stack} and {\small\tt unstack} schemas).
@@
 $F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}, and that defines the elements that can appear in the action schemes.
-In {\em blocksworld} this set contains eleven elements, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-
-In more detail, for a given operator schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in that action schema.
-For instance, $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes the elements from $F_v$ that involve $v_2$ because {\small\tt pickup} has arity one.
-
-We assume also that actions $a\in A$ are instantiated from \strips\ operator schemas $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
-\begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the operator {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$, representing the operator parameters.
-The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
-\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-\end{itemize}
-Given the set of predicates $\Psi$ and the header of the operator schema $\xi$, $2^{2|F_v(\xi)|}$ defines the size of the space of possible \strips\ models for that operator.
-The previous constraints require that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition.
+In {\em blocksworld} this set contains eleven elements, $F_v$={\small\tt\{handempty, holding\_$v_1$, holding\_$v_2$, clear\_$v_1$, clear\_$v_2$, ontable\_$v_1$, ontable\_$v_2$, on\_$v_1$\_$v_1$, on\_$v_1$\_$v_2$, on\_$v_2$\_$v_1$, on\_$v_2$\_$v_2$\}}.
+In more detail, for a given schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in that action schema.
+For instance, $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding\_$v_1$, clear\_$v_1$, ontable\_$v_1$, on\_$v_1$\_$v_1$\}} excludes the elements from $F_v$ that involve $v_2$ because {\small\tt pickup} actions have arity one.
+The size of the space of possible \strips\ models for a given schema $\xi$ is $2^{2|F_v(\xi)|}$ (recall that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition).
 For the {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while for the {\tt pickup} operator this number is only 1024.
 
-We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi'$).
-Therefore  we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
-We say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
+Now we are ready to define the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the propositional encoding for the preconditions, negative and positive effects of an action schema $\xi$.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in that schema.
+For instance, Figure~\ref{fig:encodedstack} shows the propositional encoding for the preconditions, negative and positive effects of the {\em stack} schema from a four-operator {\em blocksworld}.
 
-\subsection{Learning \strips\ action schemas from plan executions}
-The {\em learning task} addressed in this paper corresponds to learning an action model by observing an agent(s) acting in the world.
-This learning task is formalized as a tuple $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$:
+\begin{figure}[hbt!]
+\begin{footnotesize}
+\begin{verbatim}
+(pre_holding_stack_v1) (pre_clear_stack_v2)
+(del_holding_stack_v1) (del_clear_stack_v2)
+(add_handempty_stack) (add_clear_stack_v1) (add_on_stack_v1_v2)
+\end{verbatim}
+\end{footnotesize}
+ \caption{\small Propositional encoding for the {\em stack} schema from a four-operator {\em blocksworld}.}
+\label{fig:encodedstack}
+\end{figure}
+
+\subsection{Classical planning with conditional effects}
+Conditional effects allow us to compactly define actions whose effects depend on the current state.
+Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
+
+An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+\[
+triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
+\]
+
+The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
+
+\subsection{Compilation}
+%\label{sec:compilation}
+
+Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task with conditional effects.
+The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
+
+\begin{enumerate}
+\item {\bf Programs the action model $\mathcal{M}'$}.
+A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+\item {\bf Validates the action model $\mathcal{M}'$}.
+The solution plan continues with a {\em postfix} that reproduces the given input knowledge (the available plan traces) with the programmed action model $\mathcal{M}'$.
+\end{enumerate}
+
+%\subsection{Learning from observations of plan executions}
+Here we formalize the compilation for learning \strips\ action models with classical planning.
+Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
 \begin{itemize}
-\item $\mathcal{M}$ is the {\em initial} set of given operator schemas ({\em empty} or {\em partially specified} if some preconditions or effects are known) that induces the actions of a planning frame.
-\item $\Psi$ is the set of predicates that define the abstract state space of that planning frame. 
-\item $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ is a plan trace obtained watching the execution of a plan $\pi=\tup{a_1, \ldots, a_n}$. 
+\item $F_{\Lambda}$ contains:
+\begin{itemize}
+\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the plan trace given as input, i.e.
+the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-plans}.
+\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
+\item Fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in a plan trace $\tau$.
+The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ actions of $\tau$ that are observed.
+In the example of Figure~\ref{fig:example-plans} four actions were observed $\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$.
+\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\tau$ where the action model is validated.
+In the example of Figure~\ref{fig:example-plans} two states were observed $\tup{s_0,s_4}$.
+\end{itemize}
+\item $I_{\Lambda}$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated.
+Our compilation assumes that initially, operator schemas are programmed with no precondition, no negative effect and no positive effect.
+
+\item $G_{\Lambda}=\{at_n\}\cup\{test_m\}$, requires that the programmed action model is validated in all the actions and states observed from the input plan trace $\tau$.
+\item $A_{\Lambda}$ comprises three kinds of actions:
+\begin{enumerate}
+\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$:
+\begin{itemize}
+\item Actions for adding a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg pre_{f}(\xi), \neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
+\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{pre_{f}(\xi)\}.
+\end{align*}
+\end{small}
+
+\item Actions for adding a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
+\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
+\end{align*}
+\end{small}
 \end{itemize}
 
-A {\em solution} to a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is a set of operator schema $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$, the predicates $\Psi$, and the observed plan trace $\mathcal{T}$.
-Note that $\mathcal{M}$ can be inferred from $\mathcal{T}$ provided that the actions in $\mathcal{T}$ are {\em diverse} enough (meaning that $\mathcal{T}$ contains at least one instantiation of every action scheme in $\mathcal{M}$).
-Likewise $\Psi$ can be inferred from $\mathcal{T}$ when at least the observation of a state $s_i\in \mathcal{T}$ is a full state, that is $|s_i|=|F|$.
-
-In this paper we address the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task with the following assumptions:
-\begin{enumerate}
-\item Each input action scheme $\xi\in\mathcal{M}$ is at least composed of its $head(\xi)$.
-\item The initial state $s_0\in\mathcal{T}$ of the given plan trace is {\em fully observable}, $|s_0|=|F|$.
-\item Intermediate actions $a_i\in\mathcal{T}$ and states $s_i\in\mathcal{T}$ s.t.
-{\small $1\leq i\leq n$}, are {\em partially observable}.
-In the extreme, all actions $a_i$, {\small $1\leq i\leq n$}, can be missing (unobserved) provided that the final state $s_n$ is at least, partially observed.
-Some fluents in $s_i$ may be missing because it is unknown whether their value is either positive or negative, $0\leq |s_i|\leq |F|$.
-All states $s_i$, {\small $1\leq i\leq n$}, might be missing (unobserved) provided that at least actions $a_i$, {\small $1\leq i\leq n$}, are partially observed.
- \item Observations in $\mathcal{T}$ are {\em noiseless}, meaning that if a fluent or an action is observed it is the correct value. 
-\end{enumerate}
-Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, that corresponds to observing the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
-In this example $\mathcal{T}=\langle s_0,${\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$,s_4\rangle$ which means that only the first and last states are observed and the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown.
+\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
+Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
+Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$ from {\em blocksworld}.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\cup \{\neg mode_{val}\},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
+&\{\emptyset\}\rhd\{mode_{val}\}.
+\end{align*}
+\end{small}
 
 \begin{figure}[hbt!]
-{\footnotesize\tt ;;;;;; Action headers in $\mathcal{M}$}  
-\begin{footnotesize}  
+\begin{scriptsize}
 \begin{verbatim}
-(pickup ?v1) (putdown ?v1) (stack ?v1 ?v2} (unstack ?v1 ?v2)
+(:action apply_stack
+  :parameters (?o1 - object ?o2 - object)
+  :precondition
+   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
+        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
+        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
+        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
+        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
+        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
+        (or (not (pre_clear_stack_v1)) (clear ?o1))
+        (or (not (pre_clear_stack_v2)) (clear ?o2))
+        (or (not (pre_holding_stack_v1)) (holding ?o1))
+        (or (not (pre_holding_stack_v2)) (holding ?o2))
+        (or (not (pre_handempty_stack)) (handempty)))
+  :effect
+   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
+        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
+        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
+        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
+        (when (del_ontable_stack_v1) (not (ontable ?o1)))
+        (when (del_ontable_stack_v2) (not (ontable ?o2)))
+        (when (del_clear_stack_v1) (not (clear ?o1)))
+        (when (del_clear_stack_v2) (not (clear ?o2)))
+        (when (del_holding_stack_v1) (not (holding ?o1)))
+        (when (del_holding_stack_v2) (not (holding ?o2)))
+        (when (del_handempty_stack) (not (handempty)))
+        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
+        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
+        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
+        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
+        (when (add_ontable_stack_v1) (ontable ?o1))
+        (when (add_ontable_stack_v2) (ontable ?o2))
+        (when (add_clear_stack_v1) (clear ?o1))
+        (when (add_clear_stack_v2) (clear ?o2))
+        (when (add_holding_stack_v1) (holding ?o1))
+        (when (add_holding_stack_v2) (holding ?o2))
+        (when (add_handempty_stack) (handempty))
+        (when (modeProg) (not (modeProg)))))
 \end{verbatim}
-\end{footnotesize}
-\vspace{0.2cm}
-{\footnotesize\tt ;;;;;; Predicates $\Psi$}
-\begin{footnotesize}
-\begin{verbatim}
-(handempty) (holding ?v1) (clear ?v1) (ontable ?v1) (on ?v1 ?v2)
-\end{verbatim}
-\end{footnotesize}
-\vspace{0.2cm}
-{\footnotesize\tt ;;;;;; Plan trace $\mathcal{T}$}
-\begin{footnotesize}
-\begin{verbatim}
-;;; state observation #0
-(clear B) (on B A) (ontable A) (handempty)
-\end{verbatim}
-\end{footnotesize}
+\end{scriptsize}
+ \caption{\small PDDL action for applying an already programmed schema $stack$ (implications are coded as disjunctions).}
+\label{fig:compilation}
+\end{figure}
+When the input plan trace contain observed actions, then the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\
 
-\begin{footnotesize}
-{\footnotesize\tt ;;; Plan $\pi$}
-\begin{verbatim}
-0: (unstack B A)
-1: (putdown B)
-2: (pickup A)
-3: (stack A B)
-\end{verbatim}
-\end{footnotesize}
+\item Actions for {\em validating} the partially observed state $s_j\in\tau$, {\tt\small $1\leq j< m$}.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\}\cup \{mode_{val}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\neg mode_{val}\}.
+\end{align*}
+\end{small}
+\end{enumerate}
+\end{itemize}
 
-\begin{footnotesize}
-\begin{verbatim}
-;;; state observation #4
-(clear A) (on A B) (ontable B) (handempty)
-\end{verbatim}
-\end{footnotesize}
+Known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
+In this case, the corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
+When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the $P_{\Lambda}$ compilation validates whether the observation of the plan execution follows the given model:
+\begin{itemize}
+\item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given input data in $\tau$ iff a solution plan for $P_{\Lambda}$ can be found.
+\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given input data $\tau$ iff $P_{\Lambda}$ is unsolvable.
+This means that $\mathcal{M}$ cannot be compliant with the given observation of the plan execution.
+\end{itemize}
+This validation capacity of our compilation is beyond the functionality of VAL (the plan validation tool~\cite{howey2004val}) because, our $P_{\Lambda}$ compilation can test {\em model validation} with a partial (or even an empty) action model and a partially observed plan trace.
+On the other hand, VAL requires (1) a full plan and (2), a full action model for plan validation.
 
- \caption{\small Task $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ for learning a {\em blocksworld} \strips\ action model from a four-action plan and two state observations.}
-\label{fig:example-plans}
+The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to the classical planning task $P_{\Lambda}$ that encodes a $\Lambda=\tup{\mathcal{M},\tau}$ learning task for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $s_0$ and $s_4$ shown in Figure~\ref{fig:example-plans}.
+Plan steps $[0,1]$ program the preconditions of the {\tt\small stack} operator, steps $[2,6]$ program the operator effects and steps $[7,11]$ validate the programmed operators following the four-action plan shown in Figure~\ref{fig:example-plans}.
+
+\begin{figure}[hbt!]
+{\footnotesize\tt
+  {\bf 00} : (program\_pre\_holding\_stack\_v1) \\
+     01 : (program\_pre\_clear\_stack\_v2)\\
+  {\bf 02} : (program\_eff\_clear\_stack\_v1)\\
+    03 : (program\_eff\_clear\_stack\_v2)\\
+    04 : (program\_eff\_handempty\_stack)\\
+    05 : (program\_eff\_holding\_stack\_v1)\\
+    06 : (program\_eff\_on\_stack\_v1\_v2)\\
+    {\bf 07} : (apply\_unstack blockB blockA i1 i2)\\
+    08 : (apply\_putdown blockB i2 i3)\\
+    09 : (apply\_pickup blockA i3 i4)\\
+    10 : (apply\_stack blockA blockB i4 i5)\\
+    {\bf 11} : (validate\_1)
+}
+ \caption{\small Plan for programming and validating the $stack$ schema (using plan $\pi$ and state observations shown in Figure~\ref{fig:example-plans}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+\label{fig:plan-lplan}
 \end{figure}
 
-The definition of the $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task is extensible to the more general case where the execution of $\tau$ plans is observed.
-In this case, $\mathcal{T}=\{t_1,\ldots,t_{\tau}\}$ where each $t\in \mathcal{T}$ is a plan trace $t=\tup{s_0^t,a,_1^t,s_1^t,\ldots,a_n^t,s_{n}^t}$ that satisfies the previous [2,4] assumptions of our learning task.
+Note that the $P_{\Lambda}$ compilation is flexible to a bound number of {\em missing states} or {\em missing actions} in the input plan trace $\tau$.
+In both cases the output planning problem becomes simpler since the plan horizon is bound and the classical planner does not require to determine how many {\em apply} actions are necessary between any two state observations, i.e.
+between the application of two {\em validate} actions.
+
+Last but not least we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_k\}$.
+Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, {\tt\small $1\leq t\leq k$}, that belong to the same planning frame (i.e.
+same fluents and actions but different initial states and goals).
+The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
+Finally, the initial state $I_t$ is given by the state $s_0^t$ and the observed actions from the plan $\pi_t$, and the goals $G_t$ are defined by the number of observed actions and states from the corresponding plan trace.
+Addressing the learning task $\Lambda=\tup{\mathcal{M},\mathcal{T}}$ where $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ requires introducing a small modification to our compilation.
+In particular, the actions in $P_{\Lambda}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq k$} resets the current state, and the current plan, and are now redefined as:
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_{j-1}\}\cup \{\neg mode_{prog}\},\\
+\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_t\} \cup \{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t}.
+\end{align*}
+\end{small}
+
+\subsection{Compilation properties}
+
+\begin{mylemma}
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$.
+\end{mylemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
+In addition, $P_{\Lambda}$ is only solvable if fluents {\tt\small $at_n$} and {\tt\small $test_m$} hold at the last reached state.
+These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{T}$, starting from the corresponding initial state and following the sequence of actions defined by $\mathcal{T}$.
+This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
+\end{small}
+\end{proof}
+
+\begin{mylemma}
+Completeness.
+Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+\end{mylemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+  By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
+The compilation does not discard any possible \strips\ action schema $\mathcal{M}'$ definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{T}$.
+This means that a solution plan can be built selecting the corresponding  $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\mathcal{T}$.
+\end{small}
+\end{proof}
+
+The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
+\begin{itemize}
+\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
+The larger these numbers, the larger the size of the $F_v(\xi)$ sets.
+This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding set of {\em programming} actions.
+\item The length of $\mathcal{T}$, the observed plan execution. The larger $1\leq j\leq n$, the more $\{test_j\}$ fluents and $\{\mathsf{validate_{j}}\}$ actions in $P_{\Lambda}$.
+\end{itemize}
+
+\subsection{Optimizing the compilation with background knowledge}
+A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn logic programs from data~\cite{muggleton1994inductive}.
+Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case to optimize the performance of the $P_{\Lambda}$ compilation.
+
+%\subsubsection{Static predicates}
+A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
+Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
+Given a static predicate $p$:
+\begin{itemize}
+\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable objects} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
+\item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
+\end{itemize}
+
+Static predicates can also constrain the space of possible preconditions by looking at the given set of state observations in $\mathcal{T}$.
+One can assume that if a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\mathcal{T}$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
+For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{T}$.
+
+Furthermore looking as well at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\mathcal{T}$.
+Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
+
+%\subsubsection{State constraints}
+%The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
+If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
+For instance, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
+
+%{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
+
+%The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
+For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}.
+There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.
+
+%\begin{figure}[hbt!]
+%  \begin{footnotesize}
+%    \begin{center}
+%$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
+%$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
+%$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
+%$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
+%$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
+%    \end{center}
+%\end{footnotesize}
+% \caption{\small An example of the strongest invariant for the {\em blocksworld} domain.}
+%\label{fig:strongest-invariant}
+%\end{figure}
+
+%A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a mutex because $block_A$ can only be on top of a single block.
+
+%A {\em domain invariant} is an instance-independent invariant, i.e.
+holds for any possible initial state and set of objects.
+Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
+Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
+For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
+
+%An interesting contribution of our compilation is that the validation of an action model can also be done with {\em state constraints}.
+Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check also that the learned model satisfy a constraint $\phi\in\Phi$:
+%\begin{small}
+%\begin{align*}
+%\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
+%\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
+%\end{align*}
+%\end{small}
+
+%This redefinition of validate actions applies also to {\em trajectory constraints} because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.
+
+%Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
+Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space.
+With regard to our compilation, {\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema.
+For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:
+
+%\begin{small}
+%\begin{align*}
+%\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
+%\cond(\mathsf{programPre_{f_1,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
+%\end{align*}
+%\end{small}

### 2018-06-19 doc/aij18/background.tex
updates added to sections 1-4

@@
 
-Like in PDDL, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
+Like in PDDL~\cite{fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
 Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
@@
 \end{itemize}
-We assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$ and that actions $a\in A$ are instantiated from given action schemas.
+We assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$ and that actions $a\in A$ are instantiated from given action schemas, as in PDDL.
 We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
@@
 
-In this work, the term \emph{plan trace} refers to the \emph{observation} of a plan execution that starts on a given initial state.
+In this work, the term \emph{plan trace} refers to the \emph{observation} of a plan execution that starts on a given initial state and interleaves observations of the executed actions and the produced states.
 \emph{Plan traces} constitute the input knowledge of the learning tasks addressed in this paper.
+
 With regard to the observed states in a plan trace, we say that $\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$.
@@
 
-A \emph{plan trace} $\tau$ is an interleaved combination of a state and an action sequence: $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
-A plan trace $\tau$ for a planning problem $P=\tup{F,A,I,G}$ satisfies $s_0=I$, $G \in s_n$ and every element in the action sequence of $\tau$ is an action of $A$.
-Additionally, plan traces will be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
+A \emph{plan trace} $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_m \rangle$ is an interleaved combination of a state sequence $\tup{s_0, s_1, \ldots, s_m}$ and an action sequence $\tup{a_1, \ldots, a_n}$.
+A plan trace $\tau$ for a planning frame $\Phi=\tup{F,A}$ satisfies that every action $a_i\in A$ for every {\small $1\leq i\leq n$} and that every state $s_j\in\mathcal{L}(F)$ for every {\small $1\leq j\leq m$}.
+Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
 

### 2018-06-19 doc/aij18/introduction.tex
updates added to sections 1-4

@@
 Section~\ref{sec:motivation} motivates our compilation approach for learning of \strips\ action models from observations of plan executions.
-Section~\ref{sec:learning} introduces the \strips\ action model (the target of our learning and evaluation tasks) and formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.
-Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks its extension to evaluate and recognize \strips\ action models.
-Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First, the learned \strips\ action models are tested with observations of plan executions and second, the learned models are compared to the actual models.
-Finally, Section ~\ref{sec:Section8} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.
+Section~\ref{sec:learning} formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge and describe how to address this learning task with a classical planning compilation.
+Sections~\ref{sec:evaluation} describes our compilation approach for evaluating the learned \strips\ action models with respect to the actual models but also, with respect to a {\em test set} of observations of plan executions.
+Section~\ref{sec:experiments} reports the data collected in a two-fold empirical evaluation of our learning approach: First, the learned \strips\ action models are tested with observations of plan executions and second, the learned models are compared to the actual models.
+Finally, Section~\ref{sec:conclusions} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.

### 2018-06-19 doc/aij18/learning_task.tex
updates added to sections 1-4

@@
 \label{sec:learning}
-The {\em learning task} addressed in this paper corresponds to learning a classical planning action model by observing an agent(s) acting in the world.
+The {\em learning task} addressed in this paper corresponds to learning a classical planning action model by observing an agent(s) acting in a world that is defined by a {\em classical planning frame} $\Phi=\tup{F,A}$.
 This learning task is formalized by the pair $\Lambda=\tup{\mathcal{M},\tau}$:
@@
 \begin{enumerate}
- \item Observations in $\tau$ are {\em noiseless}, meaning that if the value of a fluent or an action is observed, then the observation is correct. 
+ \item Observations in $\tau$ are {\em noiseless}, meaning that if the value of a fluent or an action is observed in $\tau$, then the observation is correct. 
 \item The initial state $s_0\in\tau$ is {\em fully observed}, i.e.~$|s_0|=|F|$.
@@
 \item The set of actions $A$ is inferrable either from $\mathcal{M}$ or $\tau$.
-This means that the action name and parameters that shape the actions in $A$ are given by $\mathcal{M}$ otherwise, $\tau$ contains at least one instantiation of every action model in $\mathcal{M}$.
-\item Intermediate actions $a_i\in\tau$, and intermediate states $s_i\in\tau$, {\small $1\leq i$}, are {\em partially observable}.
-In the extreme all $a_i\in\tau$, {\small $1\leq i\leq n$} actions could be unobserved, provided that the final state $s_n$ is at least, partially observed.
-Likewise states $s_i$, {\small $1\leq i\leq n$}, might be unobserved provided that at least actions $a_i$, {\small $1\leq i\leq n$}, are partially observed.
+In other words, the action name and parameters that shape the actions in $A$ are given by $\mathcal{M}$ otherwise, $\tau$ contains at least one instantiation of every action model in $\mathcal{M}$.
+\item Intermediate actions $a_i\in\tau$, and intermediate states $s_j\in\tau$, {\small $1\leq i,j$} are {\em partially observable}.
+In the extreme all $a_i\in\tau$, {\small $1\leq i\leq n$} actions could be unobserved, provided that the final state $s_m$ is at least, partially observed.
+Likewise states $s_j$, {\small $1\leq j\leq m$}, might be unobserved provided that at least actions $a_i$, {\small $1\leq i\leq n$}, are partially observed.
 
@@
 
-A {\em solution} to a $\Lambda=\tup{\mathcal{M},\tau}$ learning task is an action model $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$ and the observed plan trace $\tau$.
+A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is an action model $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$ and the observed plan trace $\tau$.
 
 Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\tau}$, that corresponds to observing the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
-In this example $\tau=\langle s_0,${\small\tt (putdown\ B), (stack\ A\ B)}, $s_4\rangle$ which means that only the first and last states are observed (the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown) and that actions $a_1$ and $a_4$ are also unknown.
+In this example $\tau=\langle s_0,${\small\tt (putdown\ B), (stack\ A\ B)}, $s_4\rangle$ which means that only the first and last states are observed (the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown) and that actions $a_2$ and $a_3$ are observed while $a_1$ and $a_4$ are unknown.
 
@@
 \begin{verbatim}
-;;; Actions observation
-(putdown B) (stack A B)
+;;; Action observation
+(putdown B)
+
+;;; Action observation
+(stack A B)
 \end{verbatim}
@@
 \begin{itemize}
-\item The name, $name(\xi)$, and parameters, $pars(\xi)$, of the schema that define the {\em header} of that schema. 
-\item The preconditions $pre(\xi)\subseteq \Psi$, the negative effects $del(\xi)\subseteq \Psi$, and the positive effects $add(\xi)\subseteq \Psi$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\item The $name(\xi)$ and parameters, $pars(\xi)$, of the schema that define the {\em header} for that schema. 
+\item The preconditions $pre(\xi)\subseteq \Psi$, negative effects $del(\xi)\subseteq \Psi$, and positive effects $add(\xi)\subseteq \Psi$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
 \end{itemize}
-Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+As an example Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
 
@@
 Now we formalize a propositional encoding for the target of the learning task addressed in the paper, the \strips\ action model.
-This encoding is the core of our compilation approach for learning \strips\ action models from observations of plan executions.
+This encoding is the core of our compilation approach for learning \strips\ action models with classical planning.
 
 First, let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the actions with the maximum arity have arity two (the instantiations of the {\small\tt stack} and {\small\tt unstack} schemas).
+For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the actions with the maximum arity have arity two (any instantiation of the {\small\tt stack} or the {\small\tt unstack} schemas).
 
@@
 In {\em blocksworld} this set contains eleven elements, $F_v$={\small\tt\{handempty, holding\_$v_1$, holding\_$v_2$, clear\_$v_1$, clear\_$v_2$, ontable\_$v_1$, ontable\_$v_2$, on\_$v_1$\_$v_1$, on\_$v_1$\_$v_2$, on\_$v_2$\_$v_1$, on\_$v_2$\_$v_2$\}}.
-In more detail, for a given schema $\xi$, we define $F_v(\xi)\subseteq F_v$ as the subset of elements that can appear in that action schema.
-For instance, $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding\_$v_1$, clear\_$v_1$, ontable\_$v_1$, on\_$v_1$\_$v_1$\}} excludes the elements from $F_v$ that involve $v_2$ because {\small\tt pickup} actions have arity one.
-The size of the space of possible \strips\ models for a given schema $\xi$ is $2^{2|F_v(\xi)|}$ (recall that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition).
-For the {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while for the {\tt pickup} operator this number is only 1024.
+In more detail, for a given schema $\xi$, we define $F_{\xi}\subseteq F_v$ as the subset of elements that can appear in that action schema.
+For instance, $F_{\tt stack}=F_v$ while $F_{\tt pickup}$={\small\tt\{handempty, holding\_$v_1$, clear\_$v_1$, ontable\_$v_1$, on\_$v_1$\_$v_1$\}} excludes the elements from $F_v$ that involve $v_2$ because {\small\tt pickup} actions have arity one.
+The size of the space of possible \strips\ models for a given schema $\xi$ is $2^{2|F_{\xi}|}$ (recall that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition).
+For the {\em blocksworld}, $2^{2|F_{\tt stack}|}=4194304$ while for the {\tt pickup} operator this number is only 1024.
 
-Now we are ready to define the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the propositional encoding for the preconditions, negative and positive effects of an action schema $\xi$.
+Now we are ready to define the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_{\xi}$, that represent the propositional encoding for the preconditions, negative and positive effects of an action schema $\xi$.
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in that schema.
-For instance, Figure~\ref{fig:encodedstack} shows the propositional encoding for the preconditions, negative and positive effects of the {\em stack} schema from a four-operator {\em blocksworld}.
+For instance, Figure~\ref{fig:encodedstack} shows the conjunction of fluents that represents he propositional encoding for the preconditions, negative and positive effects of the {\em stack} schema shown in Figure~\ref{fig:stack}.
 
@@
 \item {\bf Programs the action model $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_v(\xi)$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_{\xi}$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
 \item {\bf Validates the action model $\mathcal{M}'$}.
@@
 the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-plans}.
-\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the programmed action model.
+\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_{\xi}$, that represent the programmed action model.
 If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
@@
 The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ actions of $\tau$ that are observed.
-In the example of Figure~\ref{fig:example-plans} four actions were observed $\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$.
-\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated, and the fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\tau$ where the action model is validated.
-In the example of Figure~\ref{fig:example-plans} two states were observed $\tup{s_0,s_4}$.
+In the example of Figure~\ref{fig:example-plans} two actions were observed $\tup{\small\tt (putdown\ B), (stack\ A\ B)}$.
+\item Fluents $\{test_j\}_{0\leq j\leq m}$, indicating the state observation $s_j\in\tau$ where the action model is validated.
+In the example of Figure~\ref{fig:example-plans} two tests are required to validate the programmed action model at $\tup{s_0,s_4}$.
+\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated. 
 \end{itemize}
-\item $I_{\Lambda}$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated.
-Our compilation assumes that initially, operator schemas are programmed with no precondition, no negative effect and no positive effect.
+\item $I_{\Lambda}$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as fluent $test_0$ and the fluents $F_{\pi}$ plus $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the action where the programmed model is validated.
+Our compilation assumes that initially, schemas are programmed with no precondition, no negative effect and no positive effect.
 
-\item $G_{\Lambda}=\{at_n\}\cup\{test_m\}$, requires that the programmed action model is validated in all the actions and states observed from the input plan trace $\tau$.
+\item $G_{\Lambda}=\{at_n, test_m\}$, requires that the programmed action model is validated in all the actions and states observed from the input plan trace $\tau$.
 \item $A_{\Lambda}$ comprises three kinds of actions:
@@
 \begin{itemize}
-\item Actions for adding a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.
+\item Actions for adding a {\em precondition} $f\in F_{\xi}$ from the action schema $\xi\in\mathcal{M}$.
 
@@
 
-\item Actions for adding a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
+\item Actions for adding a {\em negative} or {\em positive} effect $f\in F_{\xi}$ to the action schema $\xi\in\mathcal{M}$.
 
@@
 
-The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to the classical planning task $P_{\Lambda}$ that encodes a $\Lambda=\tup{\mathcal{M},\tau}$ learning task for getting the {\em blocksworld} action model where operator schemes for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan $\pi$ and the two state observations $s_0$ and $s_4$ shown in Figure~\ref{fig:example-plans}.
+The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to the classical planning task $P_{\Lambda}$ that encodes a $\Lambda=\tup{\mathcal{M},\tau}$ learning task for getting the {\em blocksworld} action model where the headers for the schemes {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
+This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan trace shown in Figure~\ref{fig:example-plans}.
 Plan steps $[0,1]$ program the preconditions of the {\tt\small stack} operator, steps $[2,6]$ program the operator effects and steps $[7,11]$ validate the programmed operators following the four-action plan shown in Figure~\ref{fig:example-plans}.
@@
 
-Note that the $P_{\Lambda}$ compilation is flexible to a bound number of {\em missing states} or {\em missing actions} in the input plan trace $\tau$.
-In both cases the output planning problem becomes simpler since the plan horizon is bound and the classical planner does not require to determine how many {\em apply} actions are necessary between any two state observations, i.e.
+The compilation is flexible to the particular scenario where the number of {\em missing states} or {\em missing actions} in the input plan trace $\tau$ is bound.
+In both cases the output planning problem $P_{\Lambda}$ becomes simpler since the plan horizon is bound so the classical planner does not require to determine how many {\em apply} actions are necessary between any two state observations, i.e.
 between the application of two {\em validate} actions.
 
-Last but not least we explain how to address learning \strips\ action models from the observation of the execution of multiple plans $\Pi=\{\pi_1,\ldots,\pi_k\}$.
-Let us first define a set of classical planning instances $P_t=\tup{F,\emptyset,I_t,G_t}$, {\tt\small $1\leq t\leq k$}, that belong to the same planning frame (i.e.
+Last but not least we explain how to address learning \strips\ action models from $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ observations of the execution of plans $\Pi=\{\pi_1,\ldots,\pi_k\}$ in a {\em classical planning frame} $\Phi=\tup{F,A}$.
+Let us first define a set of classical planning instances $P_t=\tup{F,A,I_t,G_t}$, {\tt\small $1\leq t\leq k$}, that belong to $\Phi$ (i.e.
 same fluents and actions but different initial states and goals).
-The set of actions, $A=\emptyset$, is empty because the action model is initially unknown.
-Finally, the initial state $I_t$ is given by the state $s_0^t$ and the observed actions from the plan $\pi_t$, and the goals $G_t$ are defined by the number of observed actions and states from the corresponding plan trace.
+The initial state $I_t$ is given by the state $s_0^t\in \tau_t$ and the observed actions from the plan $\pi_t$ while the goals $G_t$ are defined by the number of observed actions and states from the corresponding plan trace $\tau_t$.
 Addressing the learning task $\Lambda=\tup{\mathcal{M},\mathcal{T}}$ where $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ requires introducing a small modification to our compilation.
-In particular, the actions in $P_{\Lambda}$ for {\em validating} the plan $\pi_t\in\Pi$, {\tt\small $1\leq t\leq k$} resets the current state, and the current plan, and are now redefined as:
+In particular, the actions in $P_{\Lambda}$ for {\em validating} last state $s_m^t\in \tau_t$ of a plan trace $\tau_t$ reset the current state and the current plan.
+These actions are now redefined as:
 \begin{small}
@@
 \hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_{j-1}\}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_t\} \cup \{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t}.
+\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_t\} \cup \\
+&\{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t},\\
+&\{\neg f\}_{\forall f\in F_{\pi_t}}\cup \{f\}_{\forall f\in F_{\pi_{t+1}}}.\\
 \end{align*}

### 2018-06-19 doc/aij18/motivation.tex
updates added to sections 1-4

@@
 
-Finally, we would also like to point out that, as shown in section \ref{experiments}, \FAMA is exhaustively evaluated over a wide range of domains (15 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner so it can benefit straightforward from the last advances in classical planning.
+Finally, we would also like to point out that, as shown in section \ref{sec:experiments}, \FAMA is exhaustively evaluated over a wide range of domains (15 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner so it can benefit straightforward from the last advances in classical planning.

### 2018-06-19 doc/aij18/learning_task.tex
section 4 aij18 paper updated

@@
 \subsection{Compilation properties}
+Given a classical plan $\pi$ that solves $P_{\Lambda}$, then an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ can be computed in linear time and space.
+In this case $\pi$ has to be executed starting from the initial state $I_{\Lambda}$ and the action model $\mathcal{M}'$ will be given by the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that are set to true at the last state reached by $\pi$.
 
 \begin{mylemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$.
 \end{mylemma}
@@
   Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
-In addition, $P_{\Lambda}$ is only solvable if fluents {\tt\small $at_n$} and {\tt\small $test_m$} hold at the last reached state.
-These goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\mathcal{T}$, starting from the corresponding initial state and following the sequence of actions defined by $\mathcal{T}$.
-This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, solves $\Lambda$.
+In addition, $P_{\Lambda}$ is only solvable if fluents {\tt\small $at_n$} and {\tt\small $test_m$} hold at the last state reached by $\pi$.
+By the definition of the $\mathsf{apply_{\xi,\omega}}$ and the $\mathsf{validate_{j}}$ actions, these goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\tau$, starting from the corresponding initial state and following the sequence of $n$ actions that is defined by $\tau$.
+This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, that $\mathcal{M}'$ is a solution to $\Lambda$.
 \end{small}
@@
 \begin{mylemma}
-Completeness.
-Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+Completeness. Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\tau}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
 \end{mylemma}
@@
 \begin{small}
-  By definition, $F_v(\xi)\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$ given its header and the set of predicates $\Psi$.
-The compilation does not discard any possible \strips\ action schema $\mathcal{M}'$ definable within $F_v$ that satisfies the state trajectory constraint given by $\mathcal{T}$.
-This means that a solution plan can be built selecting the corresponding  $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\mathcal{T}$.
+  By definition, $F_{\xi}\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$.
+The compilation does not discard any possible \strips\ action schema $\mathcal{M}'$ definable within $F_v$ that satisfies the state trajectory constraint and plan constrains given by $\tau$.
+This means that a solution plan $\pi$ can be built selecting the corresponding  $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\tau$.
 \end{small}
@@
 \begin{itemize}
-\item The arity of the actions headers in $\mathcal{M}$ and the predicates $\Psi$ that are given as input to the $\Lambda$ learning task.
-The larger these numbers, the larger the size of the $F_v(\xi)$ sets.
+\item The arity of the actions and the fluents in $\tau$ that are given as input to the $\Lambda$ learning task.
+The larger these numbers, the larger the size of the $F_{\xi}$ sets.
 This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding set of {\em programming} actions.
-\item The length of $\mathcal{T}$, the observed plan execution. The larger $1\leq j\leq n$, the more $\{test_j\}$ fluents and $\{\mathsf{validate_{j}}\}$ actions in $P_{\Lambda}$.
+\item The length of the action sequence and the state trajectory in the observed plan execution $\tau$.
+The larger the number of observed actions, $a_i\in\tau$ s.t.
+$1\leq i\leq n$, the more $\{at_i\}$ fluents.
+The larger the number of observed states, $s_j\in\tau$ s.t.
+$1\leq j\leq m$, the more $\{test_j\}$ fluents and $\{\mathsf{validate_{j}}\}$ actions in $P_{\Lambda}$.
 \end{itemize}
@@
 
-Static predicates can also constrain the space of possible preconditions by looking at the given set of state observations in $\mathcal{T}$.
+Static predicates can also constrain the space of possible preconditions by looking at the given set of state observations in $\tau$.
 One can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\mathcal{T}$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
-For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\mathcal{T}$.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\tau$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
+For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\tau$.
 
-Furthermore looking as well at the given example plans, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\mathcal{T}$.
-Back to the {\em zenotravel} domain, if an example plan $\pi_t\in \Pi$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
+Furthermore looking as well at the given plan traces, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
+$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\tau$.
+Back to the {\em zenotravel} domain, if a plan trace $\tau$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
 

### 2018-06-19 doc/aij18/motivation.tex
added notes on semantic metrics

@@
 
-Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
+Regarding the evaluation of the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntactic metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
 This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM model.
 Hence, because the set of preconditions and effects of the GTM model is usually smaller than the set of all possible preconditions and effects, it turns out the metric may output error rates below 100\% for totally wrong learned models.
+To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}.
+These two metrics are frequently used in pattern recognition, information retrieval and binary classification and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
 
-In order to overcome this limitation, we propose to use two standard metrics from ML, {\em precision} and {\em recall}.
-These two metrics are frequently used in pattern recognition, information retrieval and binary classification and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
+The ARMS systems implement semantic metrics for evaluating the learned action models with respect to observations of plan executions that act as a {\em test set}.
+This semantic evaluation approach is suitable for scenarios where the GTM is not available, e.g.
+the traditional ML setting.
+For this scenario we propose semantic versions of the {\em precision} and {\em recall} that provide a notion of the soundness and completeness of the learned models with respect to input plan traces.
+Unlike the metrics defined by ARMS, our new semantic metrics do not accumulate errors caused by the same flaw in the learned model.
 

### 2018-06-19 doc/aij18/motivation.tex
small updates added to the motivation section

--- 2018-06-12 doc/aij18/motivation.tex
@@
 \label{sec:motivation}
+This section details the contributions of our approach with respect to the reviewed related  work.
 
-The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging.
+\subsection{Learning action models from plan traces of unbound horizon}
+When Learning action models from observations of plan executions there are two main sources of partial observability:
+\begin{enumerate}
+\item As many of the approaches in Section~\ref{sec:background} assume, there may be an unknown number of missing intermediate states in the trace because of partial state observability (\PO and \NO).
+The assumption of having \FO state trajectory means that the sensors are able to capture every state change at every instant which typically is unrealistic.
+Normally the process of getting state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two sensor readings.
+
+\item There may be also an unbound number of missing actions in the plan trace because of partial observability.
+The common assumption of having \FO action sequences in a learning task is unrealistic in many domains because it implies the existence of human observers that annotate the observed action sequences.
+In some real-world applications, the observed and collected data are sensory data (e.g., home automation, robotics) or images (e.g.
+traffic) and one cannot rely on human intervention for labeling actions.
+Actually, learning the executed actions can also be part of the action-model learning task.
+Learning from unstructured data involves some prior processing to transform the sensor or image information into a predicate-like format before applying the action-model learning approach~\cite{AsaiF18}, and it also requires the ability of identifying action symbols.
+In this sense, \FAMA represents a step ahead towards learning action models without assuming observed actions.
+\end{enumerate}
+
+\begin{table}[ht]
+\centering
+\begin{tabular}{c|c|c|c|c|}
+	& \multicolumn{4}{c|}{\emph{state observability}} \\ \cline{2-5}
+	\multirow{1}{*}{\emph{action}} & \FO & \POstar & \PO & \NO\\ {\emph{observability}} & & & & \\ \hline
+	\FO & - & NP-complete & NP-complete & NP-complete \\ \hline
+	\PO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
+	\NO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
+\end{tabular}
+\caption{Complexity of learning tasks according to the type of input trace}
+\label{tab:complex}
+\end{table}
+
+The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging (as Table \ref{tab:complex} shows, PSPACE-complexity happens when the length of the plan trace is a priori unbound, which results from the combination of the previous two sources partial observability).
 This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now.
@@
 
-\begin{table}[ht]
-\centering
-\begin{tabular}{c|c|c|c|c|}
-	& \multicolumn{4}{c|}{\emph{state observability}} \\ \cline{2-5}
-	\multirow{1}{*}{\emph{action}} & \FO & \POstar & \PO & \NO\\ {\emph{observability}} & & & & \\ \hline
-	\FO & - & NP-complete & NP-complete & NP-complete \\ \hline
-	\PO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
-	\NO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
-\end{tabular}
-\caption{Complexity of learning tasks according to the type of input trace}
-\label{tab:complex}
-\end{table}
-
-Table \ref{tab:complex} displays the complexity of the learning task according to the type of input trace.
-The PSPACE-complexity happens when the length of the plan trace is a priori unbound, which results from the combination of the following two causes:
-
-\begin{enumerate}
-\item As many of the approaches in section \ref{sec:background} assume, there may be an unknown number of missing intermediate states in the trace because of partial state observability (\PO and \NO).
-The assumption of having \FO state trajectory means that the sensors are able to capture every state change at every instant which typically is unrealistic.
-Normally the process of getting state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two sensor readings.
-
-\item There may be also an unbound number of missing actions in the plan trace because of partial observability.
-The common assumption of having \FO action sequences in a learning task is unrealistic in many domains because it implies the existence of human observers that annotate the observed action sequences.
-In some real-world applications, the observed and collected data are sensory data (e.g., home automation, robotics) or images (e.g.
-traffic) and one cannot rely on human intervention for labeling actions.
-Actually, learning the executed actions can also be part of the action-model learning task.
-Learning from unstructured data involves some prior processing to transform the sensor or image information into a predicate-like format before applying the action-model learning approach~\cite{AsaiF18}, and it also requires the ability of identifying action symbols.
-In this sense, \FAMA represents a step ahead towards learning action models without assuming observed actions.
-\end{enumerate}
-
-Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
+\subsection{Evaluation}
+Regarding the evaluation of the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntactic metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
 This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM model.
 Hence, because the set of preconditions and effects of the GTM model is usually smaller than the set of all possible preconditions and effects, it turns out the metric may output error rates below 100\% for totally wrong learned models.
-
-In order to overcome this limitation, we propose to use two standard metrics from ML, {\em precision} and {\em recall}.
+To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}.
 These two metrics are frequently used in pattern recognition, information retrieval and binary classification and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
 
+When learning action models from observations of plan executions, the roles of two {\em comparable} action schemes (or the roles of two action parameters with the same type) can be swapped.
+These role swaps typically happen when the observed input data is scarce.
+For instance the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
+Further, the roles of parameters of the {\small\tt stack} (or the {\small\tt unstack}) operator could be swapped and these learned models would still be semantically correct with respect to the given input observations.
+
+Pure syntax-based evaluation metrics can report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the GTM model; i.e.
+a learned model semantically equivalent but syntactically different to the reference model.
+The ARMS system implemented semantic metrics for evaluating the learned action models with respect to observations of plan executions that act as a {\em test set}.
+This semantic evaluation approach is suitable for scenarios where the GTM is not available, e.g.
+the traditional ML setting.
+For this scenario our work proposes semantic versions of {\em precision} and {\em recall} that provide a notion of the soundness and completeness of the learned models with respect to input plan traces.
+Unlike the metrics defined by ARMS, our semantic metrics do not accumulate errors that are caused by the same flaw in the learned model.
+
+As shown in section \ref{sec:experiments}, \FAMA is exhaustively evaluated, syntactically and semantically, over a wide range of domains (15 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner so it can benefit straightforward from the last advances in classical planning.
+
+\subsection{Input data}
 A striking figure of Table \ref{table:models_comparison2} that emphasizes a relevant feature of \FAMA is the small size of the training dataset it requires in comparison to other approaches.
@@
 Unlike \CAMA, our approach does not require human intervention to label samples as it is able to learn from very small datasets.
-
-Finally, we would also like to point out that, as shown in section \ref{experiments}, \FAMA is exhaustively evaluated over a wide range of domains (15 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner so it can benefit straightforward from the last advances in classical planning.

### 2018-06-19 doc/aij18/evaluation.tex
updates added to sections 1-4
updates added to section 5

@@
 \section{Evaluating \strips\ Action Models with Classical Planning}
-\label{evaluation}
+\label{sec:evaluation}
 
-If the actual GTM model is available then the quality of a learned action model is quantifiable using the ML metrics, {\em precision} and {\em recall}.
-{\em Precision} and {\em recall} are two syntactic metrics frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative than counting the number of errors between the learned and the reference model~\cite{davis2006relationship}.
+If the actual GTM model is available then the quality of learned action model is quantifiable using the well-studied ML metrics, {\em precision} and {\em recall}.
+{\em Precision} and {\em recall} are two syntactic metrics frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative than counting the number of errors between the learned and the GTM model~\cite{davis2006relationship}.
 Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
@@
 
-In a $\Lambda$ learning task, the roles of two {\em comparable} action schemes (or the roles of two action parameters with the same type) can be swapped.
-These role swaps typically happen when the observed input data, given in $\mathcal{T}$, is scarce.
-For instance when $\mathcal{T}$ does not contain any intermediate action, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
+When learning action models from observations of plan executions the semantic of the learned models can be easy altered.
+For instance, the roles of two {\em comparable} action schemes (or the roles of two action parameters with the same type) can be swapped.
+These role swaps typically happen when the observed input data, given in $\tau$, is scarce (e.g.
+if $\tau$ does not contain any intermediate action, the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
 Further, the roles of parameters of the {\small\tt stack} (or the {\small\tt unstack}) operator could be swapped and these learned models would still be semantically correct with respect to the given input observations.
 
-Pure syntax-based evaluation metrics (like {\em precision} and {\em recall}) can report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the actual model; i.e.
-a learned model semantically equivalent but syntactically different to the reference model.
+Pure syntax-based evaluation metrics (like {\em precision} and {\em recall}) can report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the GTM; i.e.
+a learned model semantically equivalent but syntactically different to the GTM.
 Here we introduce a novel evaluation approach that is robust to role changes of this particular kind.
@@
 \begin{itemize}
-\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_v(\xi)$.
-\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_v(\xi)$.
+\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_{\xi}$.
+\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_{\xi}$.
 \end{itemize}
@@
 
+Note that the number of {\em deletions} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$ matches our previous definition for the number of {\em false positives} when: (1) $\mathcal{M}$ is a learned model and (2), $\mathcal{M}'$ is the given GTM that serves as reference.
+Likewise the number of insertions required to transform $\mathcal{M}$ into $\mathcal{M}'$ corresponds to the number of {\em false negatives} in the learned model $\mathcal{M}$ with respect to the GTM $\mathcal{M}'$.
+As explained, these numbers can be normalized with the amount of {\em true positives} to compute the {\em precision} and {\em recall} of a given learned model with respect to a reference model.
+Further, we can claim that all these numbers are bounded.
 Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well.
-In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
+In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_{\xi}|$.
+This is the same max number of edits that can be introduced to the positive and to the negative effects of $\xi\in\mathcal{M}$.
 \begin{mydefinition}
-The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
+The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is given by the expression $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_{\xi}|$.
 \end{mydefinition}
 
-Note that the number of deletions indicate the number of {\em false positives} while the number of insertions corresponds to the number of {\em false negatives}.
-As explained, these numbers can be normalized with the amount of {\em true positives} to compute the {\em precision} and {\em recall} of a given learned model with respect to a reference model.
-
-Normally evaluating a given learned domain with respect to the actual GTM model is not possible because the actual GTM model is unknown.
-As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions that are assumed to be generated with that model~\cite{yang2007learning}.
-With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a partially observed plan trace $\mathcal{T}$.
+In practice evaluating a given learned domain with respect to the actual GTM model is often impossible because the actual GTM model is unknown.
+The ARMS system showed that the error of a learned action model can also be estimated with respect to a set of observations of plan executions by assuming that the observations were generated with the learned model~\cite{yang2007learning}.
+Inspired by this approach for the semantic evaluation of learned action models, we define now an edit distance to asses the quality of a learned action model with respect to a partially observed plan trace.
 
 \begin{mydefinition}
-  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and a plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ whose state observation are built with fluents in $F$.
-The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan trace $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
+  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and a plan trace $\tau=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{m}}$ whose state observation are built with fluents in $F$ and observed actions are all in $A$.
+The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan trace $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
 \end{mydefinition}
 
+Unlike the semantic metrics defined by ARMS, thes $\delta(\mathcal{M},\tau)$ distance does not accumulate errors that are caused by the same flaw in the learned model $\mathcal{M}'$.
+Further, our {\bf observation edit distance}, allows us to define a semantic version of the well-studied {\em precision} and {\em recall} metrics.
 Following the previous correspondences ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can define a {\em semantic} version of the {\em precision} and {\em recall} metrics of a learned model.
-In this case {\em precision} and {\em recall} is not computed with respect to a GTM but with respect to a partially observed plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
-To define the {\em semantic} version of {\em precision} and {\em recall} the only term to re-formalize is the number of {\em true positives} that, for each action schema $\xi\in\mathcal{M}$, is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$.
+In this case {\em precision} and {\em recall} is not computed with respect to a GTM but with respect to a partially observed plan trace $\tau=\tup{s_0,a,_1,s_1,\ldots,a_n,s_m}$.
+To define the {\em semantic} version of {\em precision} and {\em recall} the only term to re-formalize is the number of {\em true positives} that, for each action schema $\xi\in\mathcal{M}$, is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$, where $fp(\xi)$ is the number of deletions done in the schema $\xi\in\mathcal{M}$.
 
-\subsection{Computing the edit distance with classical planning}
-Our compilation is extensible to compute the {\em observation edit distance} and hence, the semantic versions of the {\em precision} and {\em recall} metrics.
+\subsection{Computing edit distances with classical planning}
+Our compilation is extensible to compute the {\em observation edit distance} and hence, our semantic versions of the {\em precision} and {\em recall} metrics.
 This extension considers that the input \strips\ model $\mathcal{M}$, is {\em non-empty} so instead of learning an action model from scratch we simply edit $\mathcal{M}$ until it satisfies the given input observations.
-In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
 A solution to the classical planning task resulting from the extended compilation is a sequence of actions that:
@@
 \item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:compilation} for the models that are programmed from scratch.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:learning} for the models that are programmed from scratch.
 \end{enumerate}
 
-In more detail given $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, the output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda},G_{\Lambda}}$ such that:
+Given $\Lambda=\tup{\mathcal{M},\tau}$, the output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda},G_{\Lambda}}$:
 \begin{itemize}
@@
 The only difference here is that each action for {\em programming} an operator schema now implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
+include also the actions for {\em deleting} a precondition or negative/positive effect from a given action schema).
 \end{itemize}
 
-To illustrate this, the plan of Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-In this case the edited action model is however validated at the plan shown in Figure~\ref{fig:example-plans}.
+Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where only the two positve effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+In this case the edited action model is again validated in the plan and state observations shown in Figure~\ref{fig:example-plans}.
 
 \begin{figure}[hbt!]
-{\tt\small
-00 : (insert\_add\_handempty\_stack)\\
+{\footnotesize\tt
+  {\bf 00} : (insert\_add\_handempty\_stack)\\
   01 : (insert\_add\_clear\_stack\_var1)\\
-02 : (apply\_unstack blockB blockA i1 i2)\\
+  {\bf 02} : (apply\_unstack blockB blockA i1 i2)\\
   03 : (apply\_putdown blockB i2 i3)\\
@@
   05 : (apply\_stack blockA blockB i4 i5)\\
-06 : (validate\_1)
+  {\bf 06} : (validate\_1)
 }
- \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the plan shown in Figure~\ref{fig:example-plans}.}
+\caption{\small Plan for editing and validating a blockswold action model where only the positve effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.}
 \label{fig:plan-pdistance}
@@
 
-Our interest when solving the classical planning task that is output by our extended compilation is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deletions).
+Our interest when solving the classical planning task that is output by our extended compilation is not in the resulting action model $\mathcal{M}'$ but in the number of {\em edit operations} (insertions and deletions) that is required to compute $\mathcal{M}'$.
 For instance, in the example of Figure~\ref{fig:plan-pdistance}, $\mathcal{M}'$ is validated in the given observations after inserting two new positive effects to the initial action model, e.g.
 $\delta(\mathcal{M},\mathcal{T})=2$ .
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:learning}).
+In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_{\tt stack}|=|F_{\tt unstack}|=11$ while $|F_{\tt pickup}|=|F_{\tt putdown}|=5$  (as shown in Section~\ref{sec:learning}).
 The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
 
-Last but not least, this compilation is flexible to compute the {\em edit distance} between two {\em comparable} \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$.
+Last but not least, the compilation approach is also flexible to compute the {\em edit distance} between two {\em comparable} \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$.
 A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion.
@@
 The sets of fluents $F_{\Lambda}$ and $I_{\Lambda}$ are defined like in the previous compilation.
-With respect to the actions, $A_{\Lambda}'$ again implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect) but does not contain $\mathsf{apply_{\xi,\omega}}$ or $\mathsf{validate_{i}}$ actions because the \strips\ action model are not validated in any observation of plan executions.
-Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_v(\xi)$.
+With respect to the actions, $A_{\Lambda}'$ again implement the two {\em edit operations} ({\em insertion} and {\em deletion} for both preconditions and effects) but does not contain $\mathsf{apply_{\xi,\omega}}$ or $\mathsf{validate_{i}}$ actions because the edited \strips\ action model are not validated in any observation of plan executions.
+Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_{\xi}$.
 To illustrate this, the plan of Figure~\ref{fig:plan-mdistance} solves the classical planning task that corresponds to computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.

### 2018-06-20 doc/aaai-19/model-recognition.tex
smal updates added to aaai19 paper

@@
 \begin{abstract} 
-Given a set of planning models and a partially observed plan execution, we define {\em model recognition} as the task of identifying which model in the set has the highest probability of producing the input observation.
-The paper formalizes the {\em model recognition} task and proposes a method to assess the probability of a given \strips\ model to produce a partially observed plan execution.
+Given a set of planning models and a partially observed plan execution, we define {\em model recognition} as the task of identifying which model in the set produced the observed execution.
+The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a given \strips\ model to produce a partially observed plan execution.
 This method, that we called {\em model recognition as planning}, is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
 The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different {\em Turing Machines}.
-We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or the values of several tape cells, are unknown.
+We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, internal machine state or the values of several tape cells, are unknown.
 \end{abstract}
@@
 
-In this paper we introduce the task of {\em model recognition}.
+This paper introduces the task of {\em model recognition}.
 Given a set of planning models and a partially observed plan execution, {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation.
@@
 \begin{itemize}
-\item Once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
-\item It enables recognizing algorithms by observing their execution.
-Planning models can encode diverse forms of algorithm representation, like {\em {\sc GOLOG} programs}, {\em finite state controllers} or {\em push-down automata}~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}.
+\item Once the planning model is recognized with certainty, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
+\item It enables the recognition of algorithms by observing their execution.
+Classical planning models are expressive enough to encode diverse forms of algorithm representation, like different {\em {\sc GOLOG} programs}, {\em finite state controllers} or {\em push-down automata}~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}.
 \end{itemize}
 
-This paper introduces {\em model recognition as planning}, a novel method to assess the probability of a given \strips\ model to produce an observed plan execution.
+The paper introduces also {\em model recognition as planning}; a novel method to assess the probability of a given \strips\ model to produce an observed plan execution.
 The method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
@@
 $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-We assume also that actions $A$ are instantiated from a set of action schemas.
-Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+We assume also that actions $A$ are instantiated from a set of action schemas that shape the preconditions and effects of the ground actions.
+Figure~\ref{fig:stack} shows the PDDL code for the {\em stack} actions, taken from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
 \begin{figure}
@@
 Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be an additional set of objects ($\Omega\cap\Omega_v=\emptyset$), that we denote as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block four-operator {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the operators with the maximum arity, {\small\tt stack} and {\small\tt unstack}, have arity two.
+For instance, in a three-block four-operator {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the {\small\tt stack} and {\small\tt unstack} actions have arity two.
 We define $F_v$, a new set of fluents, $F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}.
@@
 
-For a given $head(\xi)$, then $F_v(\xi)\subseteq F_v$ defines the subset of elements that can appear in the preconditions and effects of the corresponding action schema.
-For instance, for the {\em stack} action schema $F_v({\tt stack})=F_v$ while $F_v({\tt pickup})$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes any element in $F_v$ that involves $v_2$ because {\small\tt pickup($v_1$)} has arity one.
-The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2|F_v(\xi)|}$.
+We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$ i.e.~both share the same space of possible \strips\ models.
+For instance, we claim that blocksworld schemas {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not.
+Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
+
+For a given action schema $\xi$, then $F_{\xi}\subseteq F_v$ defines the subset of elements that can appear in the preconditions and effects of that action schema.
+For instance $F_{\tt stack}=F_v$ while $F_{\tt pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes any element in $F_v$ that involves $v_2$ because {\small\tt pickup} actions have arity one.
+The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2|F_{\xi}|}$.
 Note that \strips\ constraints require negative effects appearing as preconditions and that they cannot be positive effects and also, that positive effects cannot appear as preconditions.
-For the mentioned {\em blocksworld}, $2^{2|F_v(stack)|}=4194304$ while $2^{2|F_v(pickup)|}=1024$.
-
-We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas share the same space of possible \strips\ models, i.e.
-iff $pars(\xi)=pars(\xi')$.
-For instance, we claim that blocksworld operators {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not.
-Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
+For the mentioned {\em blocksworld}, $2^{2|F_{\tt stack}|}=4194304$ while $2^{2|F_{\tt pickup}|}=1024$.
 
@@
 \label{sec:recognition}
-Given $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, a finite and non-empty set of planning models and $\mathcal{T}$, the observation of a plan execution, {\em model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\mathcal{T}$.
-Our approach is to estimate this probability according to the amount of {\em edition} required by $\mathcal{M}$ to produce a plan execution that is compliant with $\mathcal{T}$.
+Given $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, a finite and non-empty set of planning models and $\tau$, the observation of a plan execution, {\em model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\tau$.
 
-This work assumes that models $\mathcal{M}\in M$ are {\em comparable} \strips\ planning models, so {\em edition} is focused on the action schemas $\xi\in\mathcal{M}$.
-In addition, the observation $\mathcal{T}$ is a sequence of {\em partially-observed} states and actions, $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ s.t.:
-\begin{itemize}
-\item Intermediate states $s_i\in\mathcal{T}$, {\small $1\textless i$}, are partial assignment of values to fluents in which one or more literals can be missing, formally $0\leq|s_i|\leq|F|$.
-In the extreme full intermediate states can be missing, i.e, $|s_i|=0$.
-\item Intermediate actions $a_i$, {\small $1\leq i$}, that transit between an observed state, $s_i$, to the next observed state, $s_{i+1}$, can be missing. 
-\end{itemize}
+Our approach is to estimate this probability according to the amount of {\em edition} required by the action schemas in $\mathcal{M}$ to produce a plan that is compliant with $\tau$.
+We say that a plan $\pi$ is compliant with the observation of a plan execution $\tau$ when (1) the observed actions $\pi$ but with certain actions omitted, (2) the observed states is the sequence of states traversed by $\pi$ with certain states omitted or cerntain fluents in the states omitted.
+
+To estimate the probability of a model in $M$ to produce an observation $\tau$ of a plan execution we assume that:
+\begin{enumerate}
+\item All the planning models in $M$ are {\em comparable} \strips\ models.
+\item $\tau=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{m}}$ is an interleaved sequence of {\em partially-observed} states and actions.
 The conjunction of these two assumptions makes that transitions between two consecutive observed states may involve the execution of more than a single action.
@@
 This means that the actual length of the plan whose execution is observed is unknown.
+\end{enumerate}
 
@@
   Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} \strips\ action models.
-The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
+The {\bf edit distance} $\delta(\mathcal{M},\mathcal{M}')$ is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
 \end{definition}
@@
 Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well. 
-In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
 \begin{definition}
-The \textbf{maximum edit distance} of an \strips\ model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
+The \textbf{maximum edit distance} of an \strips\ model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_{\xi}|$.
 \end{definition}
@@
 \begin{definition}
-  Given $\mathcal{M}$, a \strips\ action model built from $F_v$ and $\mathcal{T}$, an observation built with fluents in $F$ and actions in $A$.
-The {\bf observation edit distance}, $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a plan execution compliant with $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
+  Given $\mathcal{M}$, a \strips\ action model built from $F_v$ and $\tau$, an observation built with fluents in $F$ and actions in $A$.
+The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
+$\mathcal{M}'$ produces a plan execution compliant with $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
@@
 According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$, provided the observation $\mathcal{O}$, is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, hypotheses are about the comparable \strips\ models $\mathcal{M}\in M$ while the given observation is the partially observed plan execution $\mathcal{T}$.
+In our scenario, hypotheses are about the comparable \strips\ models $\mathcal{M}\in M$ while the given observation is the partially observed plan execution $\tau$.
 
-We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\mathcal{T})$, to the probability distribution of models $\mathcal{M}\in M$ provided that the plan execution $\mathcal{T}$ is observed.
+We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\tau)$, to the probability distribution of models $\mathcal{M}\in M$ provided that the plan execution $\tau$ is observed.
 This probability distribution can be estimated in three-steps:
 \begin{enumerate}
-\item Estimating the {\em a priori} probabilities $P(\mathcal{T})$ and $P(\mathcal{M})$.
-For instance given the $F_v(\xi)$ sets, if we assume that a priori all models are equiprobable $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2|F_v(\xi)|}}$.
-With respect to the observation, if we assume that all the observations of plan executions with a maximum of $n$ steps are equiprobable, $P(\mathcal{T})=\frac{1}{2^{2n\times|F|\times |A|}}$.
-\item Estimating the conditional probability $P(\mathcal{T}|\mathcal{M})$.
-For every possible model $\mathcal{M}\in M$, we can compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$ and map it into a $[0,1]$ likelihood with the expression, $1-\frac{\delta(\mathcal{M},\mathcal{T})}{\delta(\mathcal{M},*)}$.
+\item Estimating the {\em a priori} probabilities $P(\tau)$ and $P(\mathcal{M})$.
+For instance given the $F_v(\xi)$ sets, if we assume that a priori all models are equiprobable because there are no reasons to assume that
+one is a priori more likely than the others then $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2|F_v(\xi)|}}$.
+With respect to the observation $P(\tau)$ measures how surprising is the given observation.
+If we assume that all the observations of plan executions with a maximum of $n$ steps are equiprobable, $P(\tau)=\frac{1}{2^{2n\times|F|\times |A|}}$.
+\item Estimating the conditional probability $P(\tau|\mathcal{M})$.
+For every possible model $\mathcal{M}\in M$, we can compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$ and map it into a $[0,1]$ likelihood with the expression, $1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
 \item Applying the Bayes rule to obtain the normalized posterior probabilities (these probabilities must sum 1).
@@
 \label{sec:asPlanning}
-For each $\mathcal{M}\in M$, we compute the {\em observation edit distance} $\delta(\mathcal{M},\mathcal{T})$  using a classical planning compilation.
+For each $\mathcal{M}\in M$, we compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$  using a classical planning compilation.
 The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$ in the observation $\mathcal{T}$}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$.
+\item {\bf Validates the edited model $\mathcal{M}'$ in the observation $\tau$}.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\tau$.
 \end{enumerate}
@@
 Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deleitions) required by $\mathcal{M}'$ to be validated in the given observation, e.g.
-$\delta(\mathcal{M},\mathcal{T})=2$ for the example in Figure~\ref{fig:plan-pdistance}.
+$\delta(\mathcal{M},\tau)=2$ for the example in Figure~\ref{fig:plan-pdistance}.
 In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$.

### 2018-06-20 doc/aaai-19/model-recognition.tex
added goals to the definition of the model recognition task

@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that, when executed starting from the initial state $I$, induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 The {\em plan length} is denoted with $|\pi|=n$.
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the application of the plan $\pi$ in the initial state $I$.
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the execution of the plan $\pi$ in the initial state $I$.
+
+\subsection{Observations of plan executions}
+An observation $\tau$ of a plan $\pi$ executed for solving a given classical planning problem $P=\tup{F,A,I,G}$ is an interleaved combinations of a sequence of {\em partially-observed} actions $\tup{a_1,\ldots,a_n}$ and a sequence of {\em partially-observed} states $\tup{s_1,\ldots,s_m}$ such that $n\leq |\pi|$ and $m\leq |\pi|+1$.
+Note that, because of the partial observability, the transitions between two consecutive observed states in $\tau$ may involve the execution of more than a single action.
+Formally $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound.
+This means that the actual length of the plan whose execution is observed is unknown.
+
+The observation of a plan execution $\tau$ is {\em compliant} with a plan $\pi$ iff:
+\begin{itemize}
+\item The sequence of actions in $\tau$ is the same sequence of actions $\pi$ but with certain actions omitted.
+\item The sequence of states in $\tau$ is the same sequence of states traversed by $\pi$ but (1), with certain states omitted and/or (2), the value of certain fluents omitted in that states.
+\end{itemize}
+
+The observation of a plan execution $\tau$ is {\em noiseless} iff the value of any fluent or any action observed in $\tau$ is correct.
 
@@
 \label{sec:recognition}
-Given $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, a finite and non-empty set of planning models and $\tau$, the observation of a plan execution, {\em model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\tau$.
+Given a classical planning problem $P=\tup{F,A,I,G}$ where the semantic of the actions in $A$ is unknown.
+$\tau$ the observation of the execution of a plan for solving $P$ and $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, a finite and non-empty set of action models that define different semantics for the actions in $A$, then {\em model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\tau$.
 
-Our approach is to estimate this probability according to the amount of {\em edition} required by the action schemas in $\mathcal{M}$ to produce a plan that is compliant with $\tau$.
-We say that a plan $\pi$ is compliant with the observation of a plan execution $\tau$ when (1) the observed actions $\pi$ but with certain actions omitted, (2) the observed states is the sequence of states traversed by $\pi$ with certain states omitted or cerntain fluents in the states omitted.
-
-To estimate the probability of a model in $M$ to produce an observation $\tau$ of a plan execution we assume that:
-\begin{enumerate}
-\item All the planning models in $M$ are {\em comparable} \strips\ models.
-\item $\tau=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{m}}$ is an interleaved sequence of {\em partially-observed} states and actions.
-The conjunction of these two assumptions makes that transitions between two consecutive observed states may involve the execution of more than a single action.
-Formally $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound.
-This means that the actual length of the plan whose execution is observed is unknown.
-\end{enumerate}
+Our approach is to estimate this probability, for each $\mathcal{M}\in M$, according to the amount of {\em edition} required by $\mathcal{M}$ to produce a plan that is compliant with $\tau$.
 
@@
 \label{sec:asPlanning}
-For each $\mathcal{M}\in M$, we compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$  using a classical planning compilation.
+For each $\mathcal{M}\in M$, we compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$  using a classical planning compilation that follows these assumptions:
+\begin{enumerate}
+\item Observations in $\tau$ are {\em noiseless}. 
+\item The initial state $I$ is {\em fully observed}. This means that the set of fluents $F$, and the corresponding set of predicates $\Psi$ that shapes the fluents in $F$, are inferrable from $I$.
+\end{enumerate}
+
 The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:

### 2018-06-20 doc/aaai-19/model-recognition.tex
added the notion of optimality to model recognition

@@
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$.
-We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
-The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+The dynamics of the actions in $A$ can be specified with two functions: $app(s)$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)\}$ that denotes the {\em successor state} that results of applying $a$ in $s$.
 
@@
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the execution of the plan $\pi$ in the initial state $I$.
+A solution plan is {\em optimal} if it has minimum length.
 
 \subsection{Observations of plan executions}
-An observation $\tau$ of a plan $\pi$ executed for solving a given classical planning problem $P=\tup{F,A,I,G}$ is an interleaved combinations of a sequence of {\em partially-observed} actions $\tup{a_1,\ldots,a_n}$ and a sequence of {\em partially-observed} states $\tup{s_1,\ldots,s_m}$ such that $n\leq |\pi|$ and $m\leq |\pi|+1$.
+An observation $\tau$ of a plan $\pi$ executed for solving a given classical planning problem $P=\tup{F,A,I,G}$ is an interleaved combination of a sequence of {\em partially-observed} actions $\tup{a_1,\ldots,a_n}$ and a sequence of {\em partially-observed} states $\tup{s_1,\ldots,s_m}$ such that $n\leq |\pi|$ and $m\leq |\pi|+1$.
 Note that, because of the partial observability, the transitions between two consecutive observed states in $\tau$ may involve the execution of more than a single action.
 Formally $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound.
-This means that the actual length of the plan whose execution is observed is unknown.
+This means knowing $\tau$ does not mean knowing the actual length of $\pi$.
 
-The observation of a plan execution $\tau$ is {\em compliant} with a plan $\pi$ iff:
+We say that the observation of a plan execution $\tau$ is {\em compliant} with a plan $\pi$ iff:
 \begin{itemize}
@@
 
-The observation of a plan execution $\tau$ is {\em noiseless} iff the value of any fluent or any action observed in $\tau$ is correct.
+\section{Model Recognition}
+\label{sec:recognition}
+The {\em model recognition} task is a tuple $\tup{P,M,\tau}$ where:
+\begin{itemize}
+\item $P=\tup{F,A,I,G}$ is a classical planning problem such that the dynamics of the actions in $A$ is unknown (i.e. the $app(s)$ and $\theta(s,a)\}$ functions are unknown).
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty set of action models that define different semantics for the actions in $A$
+\item $\tau$ is the observation of the execution of a solution plan for $P$.
+\end{itemize}
+A solution to the {\em model recognition} task is a discrete probability distributoin $P(\mathcal{M}|\tau)$ that express for each model $\mathcal{M}\in M$ its probability of producing $\tau$.
+Our approach is to estimate the $P(\mathcal{M}|\tau)$ probability, for each $\mathcal{M}\in M$, according to the amount of {\em edition} required by $\mathcal{M}$ to produce a plan that is an optimal solution for $P$ while it is compliant with $\tau$.
 
 \subsection{The space of \strips\ models}
+In this work we assume that the dynamics of the actions in $A$ are specified using \strips\ action schemas.
 Like in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
@@
 
-\section{Model Recognition}
-\label{sec:recognition}
-Given a classical planning problem $P=\tup{F,A,I,G}$ where the semantic of the actions in $A$ is unknown.
-$\tau$ the observation of the execution of a plan for solving $P$ and $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, a finite and non-empty set of action models that define different semantics for the actions in $A$, then {\em model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\tau$.
+\subsection{The \strips\ edit distance}
 
-Our approach is to estimate this probability, for each $\mathcal{M}\in M$, according to the amount of {\em edition} required by $\mathcal{M}$ to produce a plan that is compliant with $\tau$.
-
-\subsection{The \strips\ edit distance}
 We define two edit \emph{operations} on a \strips\ model $\mathcal{M}\in M$:
@@
 The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a plan execution compliant with $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
+$\mathcal{M}'$ produces an optimal plan $\pi^*$ s.t., $\tau$ is compliant with $\pi^*$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}

### 2018-06-21 doc/aaai-19/model-recognition.tex
goals added to the model recognition formalism

--- 2018-06-20 doc/aaai-19/model-recognition.tex
@@
 \begin{abstract} 
-Given a set of planning models and a partially observed plan execution, we define {\em model recognition} as the task of identifying which model in the set produced the observed execution.
+Given a partially observed plan execution and a set of possible planning models, {\em model recognition} is the task of identifying which model in the set produced the observed execution.
 The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a given \strips\ model to produce a partially observed plan execution.
@@
 This paper introduces the task of {\em model recognition}.
-Given a set of planning models and a partially observed plan execution, {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation.
+Given a partially observed plan execution and a set of possible planning models, {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation.
 {\em Model recognition} is of interest because:
@@
 \item Once the planning model is recognized with certainty, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
-\item It enables the recognition of algorithms by observing their execution.
-Classical planning models are expressive enough to encode diverse forms of algorithm representation, like different {\em {\sc GOLOG} programs}, {\em finite state controllers} or {\em push-down automata}~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}.
+\item It enables the recognition of algorithms by observing their execution given that diverse algorithm representations, like {\em {\sc GOLOG} programs}, {\em finite state controllers} or {\em push-down automata} can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}.
 \end{itemize}
@@
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\subseteq\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\subseteq\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\subseteq\mathcal{L}(F)$.
-We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
-The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+The dynamics of the actions in $A$ are specified with two functions: $App(s)$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
 
@@
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the execution of the plan $\pi$ in the initial state $I$.
+A solution plan is {\em optimal} if it has minimum length.
 
-\subsection{Observations of plan executions}
-An observation $\tau$ of a plan $\pi$ executed for solving a given classical planning problem $P=\tup{F,A,I,G}$ is an interleaved combinations of a sequence of {\em partially-observed} actions $\tup{a_1,\ldots,a_n}$ and a sequence of {\em partially-observed} states $\tup{s_1,\ldots,s_m}$ such that $n\leq |\pi|$ and $m\leq |\pi|+1$.
-Note that, because of the partial observability, the transitions between two consecutive observed states in $\tau$ may involve the execution of more than a single action.
+\subsection{The observation model}
+The observation of a plan $\pi$ executed for solving a given classical planning problem $P=\tup{F,A,I,G}$ is an interleaved combination $\tau=\tup{s_1, a_1, \ldots , a_n, s_m}$ of a {\em partially-observed} sequence of actions and a sequence of {\em partially-observed} states.
+Because of the partial observability $0\leq n\leq |\pi|$ and $0\leq m\leq |\pi|+1$ and hence, the transitions between two consecutive observed states in $\tau$ may involve the execution of more than a single action.
 Formally $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound.
-This means that the actual length of the plan whose execution is observed is unknown.
+This means that having $\tau$ does not implies knowing the actual length of $\pi$.
 
-The observation of a plan execution $\tau$ is {\em compliant} with a plan $\pi$ iff:
+We say that a given observation $\tau$ is {\em compliant} with a given plan $\pi$ iff:
 \begin{itemize}
-\item The sequence of actions in $\tau$ is the same sequence of actions $\pi$ but with certain actions omitted.
-\item The sequence of states in $\tau$ is the same sequence of states traversed by $\pi$ but (1), with certain states omitted and/or (2), the value of certain fluents omitted in that states.
+\item The sequence of actions $\tup{a_1, \ldots, a_n}$ in $\tau$ is the same sequence of actions $\pi$ but with certain actions omitted.
+\item The sequence of states $\tup{s_1, \ldots, s_m}$ in $\tau$ is the same sequence of states traversed by $\pi$ but (1), with certain states omitted and/or (2), the value of certain fluents omitted in that states, i.e.~$|s_i|\leq |F|$ for every $0\leq i\leq m$.
 \end{itemize}
 
-The observation of a plan execution $\tau$ is {\em noiseless} iff the value of any fluent or any action observed in $\tau$ is correct.
+\section{Model Recognition}
+\label{sec:recognition}
+The {\em model recognition} task is a tuple $\tup{P,M,\tau}$ where:
+\begin{itemize}
+\item $P=\tup{F,A,I,G}$ is a classical planning problem such that the dynamics of the actions in $A$ is unknown (i.e. the $App(s)$ and $\theta(s,a)$ functions are undefined).
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty set of action models, each defining a different dynamics (different $App(s)$ and $\theta(s,a)$ functions for the actions in $A$).
+\item $\tau$ is the observation of the execution of a classical plan for solving $P$.
+\end{itemize}
+A solution to the {\em model recognition} task is a discrete probability distribution $P(\mathcal{M}|\tau)$ that expresses for each model $\mathcal{M}\in M$ its probability of producing $\tau$.
+
+According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$, provided the observation $\mathcal{O}$, is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
+In our scenario, hypotheses are about the possible action models models $\mathcal{M}\in M$ while the given observation is the partially observed plan execution $\tau$.
+The $P(\mathcal{M}|\tau)$ probability distribution can then be estimated in three steps:
+\begin{enumerate}
+\item Estimating the {\em a priori} probabilities.
+$P(\tau)$, that measures how surprising is the given observation and $P(\mathcal{M})$, that expresses if one model is a priori more likely than the others.
+\item Estimating the conditional probability $P(\tau|\mathcal{M})$.
+Our approach is to estimate this value according to the minimum amount of {\em edition} required by $\mathcal{M}$ to produce a plan $\pi^*_\tau$ such that:
+\begin{enumerate}
+\item $\pi^*_\tau$ is an optimal solution for the classical planning problem $P$. 
+\item $\pi^*_\tau$ is compliant with the given observation $\tau$.
+\end{enumerate}
+\item Applying the Bayes rule to obtain the normalized posterior probabilities. Since $P(\mathcal{M}|\tau)$ is a probability distribution these probabilities must sum 1 for all the $\mathcal{M}\in M$.
+\end{enumerate}
+
+We assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
+A related approach is recently followed in {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the different PDDL models of two different agents.
+
+\section{Recognition of \strips\ models}
+\label{sec:asPlanning}
+Here we study the particular instantiation of the model recognition task where the dynamics of the actions $A\in P$ are specified with \strips\ action schemas.
+Figure~\ref{fig:stack} shows the PDDL code for the {\em stack} actions, taken from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
 
@@
 $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-We assume also that actions $A$ are instantiated from a set of action schemas that shape the preconditions and effects of the ground actions.
-Figure~\ref{fig:stack} shows the PDDL code for the {\em stack} actions, taken from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
 \begin{figure}
@@
 
-We define an \strips\ action schema as the tuple $\xi=\tup{head(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
-\begin{itemize}
-\item $head(\xi)=\tup{name(\xi),pars(\xi)}$, is the {\em header} defined by its name and the corresponding {\em variable names}, $pars(\xi)=\{v_i\}_{i=1}^{ar(\xi)}$.
-The headers of a four-operator {\em blocksworld} are {\small\tt pickup($v_1$), putdown($v_1$), stack($v_1,v_2$)} and {\small\tt unstack($v_1,v_2$)}.
-\item The preconditions $pre(\xi)\subseteq F_v$, the negative effects $del(\xi)\subseteq F_v$, and the positive effects $add(\xi)\subseteq F_v$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-\end{itemize}
-
-We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$ i.e.~both share the same space of possible \strips\ models.
-For instance, we claim that blocksworld schemas {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not.
-Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
-
 For a given action schema $\xi$, then $F_{\xi}\subseteq F_v$ defines the subset of elements that can appear in the preconditions and effects of that action schema.
 For instance $F_{\tt stack}=F_v$ while $F_{\tt pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes any element in $F_v$ that involves $v_2$ because {\small\tt pickup} actions have arity one.
+
+Now we are ready to define the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_{\xi}$, that represent the propositional encoding for the preconditions, negative and positive effects of an action schema $\xi$.
+If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in that schema.
+For instance, Figure~\ref{fig:encodedstack} shows the conjunction of fluents that represents he propositional encoding for the preconditions, negative and positive effects of the {\em stack} schema shown in Figure~\ref{fig:stack}.
+
+\begin{figure}
+\begin{scriptsize}
+\begin{verbatim}
+(pre_holding_stack_v1) (pre_clear_stack_v2)
+(del_holding_stack_v1) (del_clear_stack_v2)
+(add_handempty_stack) (add_clear_stack_v1) 
+(add_on_stack_v1_v2)
+\end{verbatim}
+\end{scriptsize}
+ \caption{\small Propositional encoding for the {\em stack} schema from a four-operator {\em blocksworld}.}
+\label{fig:encodedstack}
+\end{figure}
+
 The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2|F_{\xi}|}$.
@@
 For the mentioned {\em blocksworld}, $2^{2|F_{\tt stack}|}=4194304$ while $2^{2|F_{\tt pickup}|}=1024$.
-
-\section{Model Recognition}
-\label{sec:recognition}
-Given a classical planning problem $P=\tup{F,A,I,G}$ where the semantic of the actions in $A$ is unknown.
-$\tau$ the observation of the execution of a plan for solving $P$ and $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$, a finite and non-empty set of action models that define different semantics for the actions in $A$, then {\em model recognition} is the task of identifying which model $\mathcal{M}\in M$ has the highest probability of producing $\tau$.
-
-Our approach is to estimate this probability, for each $\mathcal{M}\in M$, according to the amount of {\em edition} required by $\mathcal{M}$ to produce a plan that is compliant with $\tau$.
+We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same space of possible \strips\ models.
+For instance, we claim that blocksworld schemas {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not.
+Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
 
@@
 \begin{definition}
-  Given $\mathcal{M}$, a \strips\ action model built from $F_v$ and $\tau$, an observation built with fluents in $F$ and actions in $A$.
+  Given $\tau$, an observation of the execution of a classical plan for solving $P$ and $\mathcal{M}$, a \strips\ action model built from $F_v$.
 The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a plan execution compliant with $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
+$\mathcal{M}'$ produces a plan $\pi^*_\tau$ that is optimal for $P$ and is compliant with $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
 
-\subsection{The \strips\ probability distribution}
-According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$, provided the observation $\mathcal{O}$, is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, hypotheses are about the comparable \strips\ models $\mathcal{M}\in M$ while the given observation is the partially observed plan execution $\tau$.
+Our $\delta(\mathcal{M},\tau)$ distance could also be defined assessing the edition required by the observed plan execution to match the given model.
+This implies defining {\em edit operations} that modify $\tau$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
+Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$.
+In practice, the number of {\em variable objects} is normally smaller than the number of objects in the observed states.
 
-We call the {\em \strips\ probability distribution}, denoted by $P(\mathcal{M}|\tau)$, to the probability distribution of models $\mathcal{M}\in M$ provided that the plan execution $\tau$ is observed.
-This probability distribution can be estimated in three-steps:
+\subsection{The \strips\ $P(\mathcal{M}|\tau)$ probability distribution}
+Since we assume that the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, we can compute the $P(\mathcal{M}|\tau)$ probability as follows:
 \begin{enumerate}
-\item Estimating the {\em a priori} probabilities $P(\tau)$ and $P(\mathcal{M})$.
-For instance given the $F_v(\xi)$ sets, if we assume that a priori all models are equiprobable because there are no reasons to assume that
-one is a priori more likely than the others then $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2|F_v(\xi)|}}$.
-With respect to the observation $P(\tau)$ measures how surprising is the given observation.
-If we assume that all the observations of plan executions with a maximum of $n$ steps are equiprobable, $P(\tau)=\frac{1}{2^{2n\times|F|\times |A|}}$.
-\item Estimating the conditional probability $P(\tau|\mathcal{M})$.
-For every possible model $\mathcal{M}\in M$, we can compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$ and map it into a $[0,1]$ likelihood with the expression, $1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
-\item Applying the Bayes rule to obtain the normalized posterior probabilities (these probabilities must sum 1).
+\item If we assume that a priori all models are equiprobable because there are no reasons to assume that one is a priori more likely than the others then $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2|F_v(\xi)|}}$.
+\item With respect to $P(\tau)$, if we assume that all the observations of plan executions with a maximum of $n$ observed actions and $m$ observed states are equiprobable then $P(\tau)=\frac{1}{2^{n\times|A|}\times2^{m\times|F|}}$.
+\item To estimate the conditional probability $P(\tau|\mathcal{M})$ we compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$ and map it into a $[0,1]$ likelihood with the expression, $1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
 \end{enumerate}
 
-\section{Model Recognition as Planning}
-\label{sec:asPlanning}
-For each $\mathcal{M}\in M$, we compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$  using a classical planning compilation that follows these assumptions:
-\begin{enumerate}
-\item Observations in $\tau$ are {\em noiseless}. 
-\item The initial state $I$ is {\em fully observed}. This means that the set of fluents $F$, and the corresponding set of predicates $\Psi$ that shapes the fluents in $F$, are inferrable from $I$.
-\end{enumerate}
-
+\section{Model recognition as planning}
+This section shows that, when the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, then the {\em observation edit distance} can be computed with a compilation of a classical planning with conditional effects.
 The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \subsection{The compilation formalization}
-Given a \strips\ model $\mathcal{M}\in M$ and the observation of a plan execution $\mathcal{T}$, our compilation outputs a classical planning task $P=\tup{F,A,I,G}$ such that:
+Given a \strips\ model $\mathcal{M}\in M$ and the observation of a plan execution $\tau$, our compilation outputs a classical planning task $P'=\tup{F',A',I',G'}$ such that:
 \begin{itemize}
-\item $F$ contains:
+\item $F'$ contains:
 \begin{itemize}
-\item The fluents built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in any state/action observed in $\mathcal{T}$, i.e.
-the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:plan-pdistance}.
-\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_v(\xi)$, that represent the edited action model.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
-For instance, the preconditions of the $stack$ schema (Figure~\ref{fig:stack}) are represented by the pair of fluents {\small\tt pre\_holding\_stack\_$v_1$} and {\small\tt pre\_clear\_stack\_$v_2$} set to True.
-\item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\mathcal{T}$.
-The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of $\mathcal{T}$.
-\item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schemas are edited or validated, and the fluents $\{test_i\}_{1\leq i\leq n}$, indicating the state observation $s_i\in\mathcal{T}$ where the action model is validated.
+\item The original fluents $F$. 
+\item Fluents modeling the space of \strips\ models. That is, for every $f\in F_v(\xi)$, the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that represent the edited action model.
+\item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\tau$.
+The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of $\tau$.
+\item The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\tau$ where the action model is validated.
+\item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schemas are edited or validated.
 \end{itemize}
-\item $I$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{edit}$ as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated.
+\item $I'$ extends the original initial state $I$ with the fluent $mode_{edit}$ set to true as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated.
 Our compilation assumes that initially $\mathcal{M}'$ is defined as $\mathcal{M}$.
@@
 
-\item $G=\bigcup_{1\leq i\leq n}\{test_i\}$, requires that the edited model is validated in the state observations $s_i\in\mathcal{T}$.
-\item $A$ comprises three kinds of actions with conditional effects:
+\item $G'=G\bigcup\{at_n,test_m\}$.
+\item $A'$ comprises three kinds of actions with conditional effects:
 \begin{enumerate}
@@
 \end{figure}
-When the observation $\mathcal{T}$ includes observed actions, then extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\
+When the observation $\tau$ includes observed actions, then extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\tau$.\\
 
-\item Actions for {\em validating} the partially observed state $s_i\in\mathcal{T}$, {\tt\small $1\leq i< n$}.
+\item Actions for {\em validating} the partially observed state $s_j\in\tau$, {\tt\small $1\leq j< m$}.
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{i}})=&s_i\cup\{test_j\}_{j\in 1\leq j<i}\\
-&\cup\{\neg test_j\}_{j\in i\leq j\leq n}\cup \{mode_{val}\},\\
-\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_i\cup\{test_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\neg mode_{val}\}.
 \end{align*}
@@
 
-Since the {\em transition function} of a {\em Turing Machine} can be encoded as a classical planning frame $\Phi=\tup{F,A}$, executions of that {\em Turing Machine} are definable as {\em plan traces} $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
+Since the {\em transition function} of a {\em Turing Machine} can be encoded as a classical planning frame $\Phi=\tup{F,A}$, executions of that {\em Turing Machine} are definable as {\em plan traces} $\tau=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
 
@@
 We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a seven-symbol six-state {\em Turing Machine}.
-We randomly choose a machine $\mathcal{M}\in M$ and produce an fifty-step execution plan trace $\mathcal{T}=\tup{s_0,a_1,s_1,\ldots,a_{50},s_{50}}$.
-Finally, we follow our {\em model recognition as planning} method to identify the {\em Turing Machine} that produced $\mathcal{T}$.
-This exeperiment is repeated for different amounts of missing information in the input trace $\mathcal{T}$: unknown applied transitions, unknown internal machine state and unknown values of several tape cells.
+We randomly choose a machine $\mathcal{M}\in M$ and produce an fifty-step execution plan trace $\tau=\tup{s_0,a_1,s_1,\ldots,a_{50},s_{50}}$.
+Finally, we follow our {\em model recognition as planning} method to identify the {\em Turing Machine} that produced $\tau$.
+This exeperiment is repeated for different amounts of missing information in the input trace $\tau$: unknown applied transitions, unknown internal machine state and unknown values of several tape cells.
 
@@
 
-Our $\delta(\mathcal{M},\mathcal{T})$ distance could also be defined assessing the edition required by the observed plan execution to match the given model.
-This implies defining {\em edit operations} that modify $\mathcal{T}$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
-Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$.
-In practice, the number of {\em variable objects} is normally smaller than the number of objects in the observed states.
-
-The {\em \strips\ probability distribution} assigns the same probabaility value to any hypothesis model $\mathcal{M}\in M$ that can reproduce the given observation $\mathcal{T}$.
-As a rule of thumb, the less information observed the more different models can reproduce observations.
-If we assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}, one can define probability distributions that consider as more probable hypothesis the ones that reproduce observations in less steps.
-A related approach is recently followed in {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the different PDDL models of two different agents.
-
 \bibliographystyle{aaai}

### 2018-06-21 doc/aaai-19/model-recognition.tex
goals added to the models recognition formalism

@@
 \begin{abstract} 
-Given a partially observed plan execution and a set of possible planning models, {\em model recognition} is the task of identifying which model in the set produced the observed execution.
-The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a given \strips\ model to produce a partially observed plan execution.
+Given a partially observed plan execution and a set of possible planning models, {\em model recognition} is the task of identifying the model in the set that produced the observed plan execution.
+The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a \strips\ model to produce a given partially observed plan execution.
 This method, that we called {\em model recognition as planning}, is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
@@
 {\em Plan recognition as planning} is the model-based approach for plan recognition~\cite{ramirez2012plan,ramirez2009plan}.
-This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal according to the observed plan execution.
+This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal of the agent according to the observed plan execution.
 
@@
 \label{sec:background}
-This section formalizes classical planning and characterizes the space of \strips\ models.
+This section formalizes classical planning and the observation of the execution of a classical plan.
 
@@
 The {\em plan length} is denoted with $|\pi|=n$.
+
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the execution of the plan $\pi$ in the initial state $I$.
@@
 \subsection{The observation model}
-The observation of a plan $\pi$ executed for solving a given classical planning problem $P=\tup{F,A,I,G}$ is an interleaved combination $\tau=\tup{s_1, a_1, \ldots , a_n, s_m}$ of a {\em partially-observed} sequence of actions and a sequence of {\em partially-observed} states.
+Given classical planning problem $P=\tup{F,A,I,G}$, the observation of the execution of a plan $\pi$ that solves $P$ is an interleaved combination $\tau=\tup{s_1, a_1, \ldots , a_n, s_m}$ of a {\em partially-observed} sequence of actions and a sequence of {\em partially-observed} states.
 Because of the partial observability $0\leq n\leq |\pi|$ and $0\leq m\leq |\pi|+1$ and hence, the transitions between two consecutive observed states in $\tau$ may involve the execution of more than a single action.
@@
 \begin{itemize}
-\item The sequence of actions $\tup{a_1, \ldots, a_n}$ in $\tau$ is the same sequence of actions $\pi$ but with certain actions omitted.
+\item The sequence of actions $\tup{a_1, \ldots, a_n}$ in $\tau$ is a subsequence of $\pi$.
 \item The sequence of states $\tup{s_1, \ldots, s_m}$ in $\tau$ is the same sequence of states traversed by $\pi$ but (1), with certain states omitted and/or (2), the value of certain fluents omitted in that states, i.e.~$|s_i|\leq |F|$ for every $0\leq i\leq m$.
@@
 \begin{itemize}
-\item $P=\tup{F,A,I,G}$ is a classical planning problem such that the dynamics of the actions in $A$ is unknown (i.e. the $App(s)$ and $\theta(s,a)$ functions are undefined).
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty set of action models, each defining a different dynamics (different $App(s)$ and $\theta(s,a)$ functions for the actions in $A$).
-\item $\tau$ is the observation of the execution of a classical plan for solving $P$.
+\item $P=\tup{F,A,I,G}$ is a classical planning problem such that the dynamics of the actions in $A$ is unknown because the $App(s)$ and $\theta(s,a)$ functions are undefined.
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty set of action models, each defining a different dynamics (a different $App(s)$ and $\theta(s,a)$ pair) for the actions in $A$.
+\item $\tau$ is the observation of the execution of a solution plan $\pi$ for $P$.
 \end{itemize}
-A solution to the {\em model recognition} task is a discrete probability distribution $P(\mathcal{M}|\tau)$ that expresses for each model $\mathcal{M}\in M$ its probability of producing $\tau$.
+A solution to the {\em model recognition} task is the discrete probability distribution $P(\mathcal{M}|\tau)$ that expresses, for each model $\mathcal{M}\in M$, its probability of producing the observation $\tau$.
 
 According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$, provided the observation $\mathcal{O}$, is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In our scenario, hypotheses are about the possible action models models $\mathcal{M}\in M$ while the given observation is the partially observed plan execution $\tau$.
+In the {\em model recognition} task, hypotheses are about the possible action models $\mathcal{M}\in M$ while the given observation is the partially observed plan execution $\tau$.
 The $P(\mathcal{M}|\tau)$ probability distribution can then be estimated in three steps:
 \begin{enumerate}
-\item Estimating the {\em a priori} probabilities.
+\item Computing the {\em a priori} probabilities.
 $P(\tau)$, that measures how surprising is the given observation and $P(\mathcal{M})$, that expresses if one model is a priori more likely than the others.
-\item Estimating the conditional probability $P(\tau|\mathcal{M})$.
-Our approach is to estimate this value according to the minimum amount of {\em edition} required by $\mathcal{M}$ to produce a plan $\pi^*_\tau$ such that:
+\item Computing the conditional probability $P(\tau|\mathcal{M})$.
+Our approach is to estimate this value according to the minimum amount of {\em edition} that is required by $\mathcal{M}$ to produce a plan $\pi^*_\tau$ such that:
 \begin{enumerate}
@@
 \end{enumerate}
-\item Applying the Bayes rule to obtain the normalized posterior probabilities. Since $P(\mathcal{M}|\tau)$ is a probability distribution these probabilities must sum 1 for all the $\mathcal{M}\in M$.
+\item Applying the Bayes rule to obtain the normalized posterior probabilities (the $P(\mathcal{M}|\tau)$ probabilities must sum 1 for all the $\mathcal{M}\in M$).
 \end{enumerate}
 
-We assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
-A related approach is recently followed in {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the different PDDL models of two different agents.
+Note that the second step assumes that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
+A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the different PDDL models of two different agents.
 
@@
 \label{sec:asPlanning}
-Here we study the particular instantiation of the model recognition task where the dynamics of the actions $A\in P$ are specified with \strips\ action schemas.
+Here we analyze the particular instantiation of the {\em model recognition} task where the dynamics of the actions $A\in P$ (i.e.
+the $App(s)$ and $\theta(s,a)$ functions) are specified with \strips\ action schemas.
 Figure~\ref{fig:stack} shows the PDDL code for the {\em stack} actions, taken from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
@@
 
-The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2|F_{\xi}|}$.
-Note that \strips\ constraints require negative effects appearing as preconditions and that they cannot be positive effects and also, that positive effects cannot appear as preconditions.
+The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions and that they cannot be positive effects and also, that positive effects cannot appear as preconditions).
 For the mentioned {\em blocksworld}, $2^{2|F_{\tt stack}|}=4194304$ while $2^{2|F_{\tt pickup}|}=1024$.
@@
 \begin{definition}
-  Given $\tau$, an observation of the execution of a classical plan for solving $P$ and $\mathcal{M}$, a \strips\ action model built from $F_v$.
+  Given $\tau$, an observation of the execution of a plan for solving $P$ and $\mathcal{M}$, a \strips\ action model built from $F_v$.
 The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a plan $\pi^*_\tau$ that is optimal for $P$ and is compliant with $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
+$\mathcal{M}'$ produces a plan $\pi^*_\tau$ optimal for $P$ and compliant with $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
@@
 Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$.
-In practice, the number of {\em variable objects} is normally smaller than the number of objects in the observed states.
+In practice, the number of {\em variable objects} is normally smaller than the number of objects in a planning problem.
 
-\subsection{The \strips\ $P(\mathcal{M}|\tau)$ probability distribution}
-Since we assume that the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, we can compute the $P(\mathcal{M}|\tau)$ probability as follows:
+\subsection{The $P(\mathcal{M}|\tau)$ probability for \strips\ models}
+Since we assume that the dynamics of the actions $A\in P$ is specified with \strips\ action schemas, we can estimate the $P(\mathcal{M}|\tau)$ probability as follows:
 \begin{enumerate}
-\item If we assume that a priori all models are equiprobable because there are no reasons to assume that one is a priori more likely than the others then $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2|F_v(\xi)|}}$.
-\item With respect to $P(\tau)$, if we assume that all the observations of plan executions with a maximum of $n$ observed actions and $m$ observed states are equiprobable then $P(\tau)=\frac{1}{2^{n\times|A|}\times2^{m\times|F|}}$.
+\item Assuming that a priori, all action models $\mathcal{M}\in M$ are equiprobable because there are no reasons to assume that one is a priori more likely than the others.
+Therefore, $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2|F_{\xi}|}}$.
+\item Assuming that all the observations of plan executions with a maximum of $n$ observed actions and $m$ observed states are equiprobable then $P(\tau)=\frac{1}{2^{n\times|A|}\times2^{m\times|F|}}$.
 \item To estimate the conditional probability $P(\tau|\mathcal{M})$ we compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$ and map it into a $[0,1]$ likelihood with the expression, $1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
@@
 \section{Model recognition as planning}
-This section shows that, when the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, then the {\em observation edit distance} can be computed with a compilation of a classical planning with conditional effects.
+This section shows that, when the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, then $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects.
 The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-The edited action model is validated at the observation of a four action plan for inverting a two-block tower where: states $s_0$ and $s_4$ are fully observed while the intermediate states, $s_1$, $s_2$ and $s_3$, are unobserved.
+The edited action model is validated at the observation of a four action plan for inverting a two-block tower where the intermediate states, $s_1$, $s_2$ and $s_3$, are unobserved.
 \begin{figure}
@@
 
-Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deleitions) required by $\mathcal{M}'$ to be validated in the given observation, e.g.
+Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deletions) required by $\mathcal{M}'$ to be validated in the given observation, e.g.
 $\delta(\mathcal{M},\tau)=2$ for the example in Figure~\ref{fig:plan-pdistance}.
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$.
+In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_{\tt stack}|=|F_{\tt unstack}|=11$ while $|F_{\tt pickup}|=|F_{\tt putdown}|=5$.
 %The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
@@
 \subsection{The compilation formalization}
-Given a \strips\ model $\mathcal{M}\in M$ and the observation of a plan execution $\tau$, our compilation outputs a classical planning task $P'=\tup{F',A',I',G'}$ such that:
+Given a \strips\ model $\mathcal{M}\in M$ and the observation $\tau$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task $P'=\tup{F',A',I',G'}$ such that:
 \begin{itemize}
@@
 \item The original fluents $F$. 
-\item Fluents modeling the space of \strips\ models. That is, for every $f\in F_v(\xi)$, the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that represent the edited action model.
+\item Fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ modeling the space of \strips\ models. 
 \item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\tau$.
@@
 \begin{itemize}
-\item Actions for {\bf removing} a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.
+\item Actions for adding a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{edit}, pre_{f}(\xi)\},\\
-\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg pre_{f}(\xi), \neg del_{f}(\xi),\neg add_{f}(\xi), mode_{edit}\},\\
+\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{pre_{f}(\xi)\}.
 \end{align*}
@@
 
-\item Actions for {\bf adding} a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
+\item Actions for adding a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
 
@@
 \end{itemize}
-Besides these actions, $A$ also contains the actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect.
+Besides these actions, the $A'$ set also contains the actions for {\em deleting} a precondition and a negative/positive effect.
 
@@
 \end{figure}
-When the observation $\tau$ includes observed actions, then extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\tau$.\\
+When the observation $\tau$ includes observed actions, then the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as they appear in $\tau$.\\
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_i\cup\{test_{j-1}\},\\
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
 \cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\neg mode_{val}\}.
@@
 \item The {\bf header} is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$. Parameters $?xl$, $?x$ and $?xr$ are tape cells.
-\item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to make $?x$ the tape cell pointed by the header and $?xl/?xr$ its left/right neighbours.
+\item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to make $?x$ the tape cell pointed by the header and $?xl/?xr$ its left/right neighbors.
 Preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} are also included to capture the symbol currently pointed by the header and the current machine state.
-\item The {\bf delete effects} remove the symbol currently pointed by the header and the currrent machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state according to $\delta$.
+\item The {\bf delete effects} remove the symbol currently pointed by the header and the current machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state according to $\delta$.
 \end{itemize}
@@
 
-Since the {\em transition function} of a {\em Turing Machine} can be encoded as a classical planning frame $\Phi=\tup{F,A}$, executions of that {\em Turing Machine} are definable as {\em plan traces} $\tau=\tup{s_0,a_1,s_1,\ldots,a_n,s_n}$.
+Since the {\em transition function} of a {\em Turing Machine} can be encoded as a classical planning frame $\Phi=\tup{F,A}$, executions of that {\em Turing Machine} are definable as an action sequence $\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 
@@
 We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a seven-symbol six-state {\em Turing Machine}.
-We randomly choose a machine $\mathcal{M}\in M$ and produce an fifty-step execution plan trace $\tau=\tup{s_0,a_1,s_1,\ldots,a_{50},s_{50}}$.
+The classical planning frame $\Phi=\tup{F,A}$ is the same for all the {\em Turing Machines}, there is an $a\in A$ action for each pair of tape symbol and non-terminal state machine, while the $\theta(s,a)$ function is defined differently for each the machines (using a different \strips\ action model).
+
+We randomly choose a machine $\mathcal{M}\in M$ and produce the observation $\tau$ of a fifty-step execution plan.
 Finally, we follow our {\em model recognition as planning} method to identify the {\em Turing Machine} that produced $\tau$.
-This exeperiment is repeated for different amounts of missing information in the input trace $\tau$: unknown applied transitions, unknown internal machine state and unknown values of several tape cells.
+This experiment is repeated for different amounts of missing information in the input trace $\tau$: unknown applied transitions, unknown internal machine state and unknown values of several tape cells.
 
@@
 
-\subsection{Conclussions}
+\subsection{Conclusions}
 \label{sec:conclussions}

### 2018-07-08 doc/aij18/background.tex
Add files via upload
Sections 2, 3, 4 and 5 finished

@@
 Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
 
@@
 $|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.
+We explicitly include negative literals $\neg f$ in states since our approach handles with partial observability, thus adopting the \emph{open world assumption}.
+
+%simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.
 
@@
 
-In this work, the term \emph{plan trace} refers to the \emph{observation} of a plan execution that starts on a given initial state and interleaves observations of the executed actions and the produced states.
+In this work, the term \emph{plan trace} refers to the \emph{observation} of a plan execution that starts on a given initial state.
+A plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ is generally defined as an interleaved combination of a sequence of executed actions $\tup{a_1, \ldots, a_n}$ and the induced state trajectory $\tup{s_0, s_1, \ldots, s_n}$.
 \emph{Plan traces} constitute the input knowledge of the learning tasks addressed in this paper.
@@
 
-A \emph{plan trace} $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_m \rangle$ is an interleaved combination of a state sequence $\tup{s_0, s_1, \ldots, s_m}$ and an action sequence $\tup{a_1, \ldots, a_n}$.
-A plan trace $\tau$ for a planning frame $\Phi=\tup{F,A}$ satisfies that every action $a_i\in A$ for every {\small $1\leq i\leq n$} and that every state $s_j\in\mathcal{L}(F)$ for every {\small $1\leq j\leq m$}.
+Given a  plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$, it holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}.
 Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
@@
 
-The Action-Relation Modeling System (\textbf{\ARMS}) \cite{yang2007learning} is one of the first learning algorithms able to learn from plan traces with partial or null observations of intermediate states.
+The Action-Relation Modeling System (\textbf{\ARMS})~\cite{yang2007learning} is one of the first learning algorithms able to learn from plan traces with partial or null observations of intermediate states.
 \ARMS uncovers a number of constraints from the plan traces in the training data that must hold for the plans to be correct.

### 2018-07-08 doc/aij18/evaluation.tex
Add files via upload
Sections 2, 3, 4 and 5 finished

--- 2018-06-12 doc/aij18/evaluation.tex
@@
-\section{Evaluating \strips\ Action Models with Classical Planning}
-\label{evaluation}
+\section{Evaluation of action models}
+\label{sec:evaluation}
 
-If the actual GTM model is available then the quality of a learned action model is quantifiable using the ML metrics, {\em precision} and {\em recall}.
-{\em Precision} and {\em recall} are two syntactic metrics frequently used in {\em pattern recognition}, {\em information retrieval} and {\em binary classification} and are more informative than counting the number of errors between the learned and the reference model~\cite{davis2006relationship}.
+In this section we introduce the metrics used by \FAMA to evaluate the action models that result from solving a learning task $\Lambda$.
+First, we will describe two standard syntactic metrics and then a semantic evaluation measure that leverages these two syntactic metrics is defined in section \ref{semantic_precision_recall}.
+Finally, section \ref{edit_distance} explains how \FAMA computes the semantic-based metric.
+
+\vspace{0.1cm}
+
+When the planning reference model of the input observations (i.e., the GTM) is available, the quality of the learned action models is measurable using two well-studied syntax-based metrics, {\em precision} and {\em recall}, commonly used in tasks such as information retrieval and recommender systems~\cite{davis2006relationship}.
+These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM.
 Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
+
 \begin{itemize}
@@
 
-In a $\Lambda$ learning task, the roles of two {\em comparable} action schemes (or the roles of two action parameters with the same type) can be swapped.
-These role swaps typically happen when the observed input data, given in $\mathcal{T}$, is scarce.
-For instance when $\mathcal{T}$ does not contain any intermediate action, the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
-Further, the roles of parameters of the {\small\tt stack} (or the {\small\tt unstack}) operator could be swapped and these learned models would still be semantically correct with respect to the given input observations.
+Introducing semantic-based evaluation metrics can be justified on several grounds:
 
-Pure syntax-based evaluation metrics (like {\em precision} and {\em recall}) can report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the actual model; i.e.
-a learned model semantically equivalent but syntactically different to the reference model.
-Here we introduce a novel evaluation approach that is robust to role changes of this particular kind.
-The intuition of the approach is to {\em semantically} assess how well a \strips\ action model $\mathcal{M}$ explains given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce that observations.
-This semantic evaluation approach is again flexible to various amount and kind of available input knowledge.
-
-\subsection{The \strips\ edit distance}
-
-We first define the two allowed \emph{operations} to edit a given \strips\ action model $\mathcal{M}$:
 \begin{itemize}
-\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_v(\xi)$.
-\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_v(\xi)$.
+\item The GTM is unknown
+\item Though a planning reference model exists, a test-based model evaluation on a dataset is preferable or needed as complementary to a syntactic evaluation
+\item The semantics of an action model learned from observations of plan executions can be easily altered.
+That is, it is very likely to learn a semantically correct but syntactically incorrect model.
+We will refer to this problem as \emph{reformulation}.
 \end{itemize}
 
-We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ action models are.
-The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations}, deletion and insertion, have the same positive cost.
+\emph{Reformulation} has a large impact in the action-model learning task.
+For instance, the roles of two {\em comparable} action models can be swapped.
+Two action models $\xi$ and $\xi'$ are comparable if both have the same parameters (iff $pars(\xi)=pars(\xi'$)) and so they share the same space of possible models.
+Hence, the {\em blocksworld} operator {\small\tt stack} could be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator, and viceversa, because they are comparable.
+On the contrary, this reformulation will not happen between the {\tt stack} and {\tt pickup} because they are not comparable.
+In the same way, the roles of two action parameters that share the same type can also be swapped (e.g., interchanging the role of the two parameters of the operator {\small\tt stack} or the opreator {\small\tt unstack}) and yet the learned models would be semantically correct with respect to the given input observations.
+A more complex kind of reformulation occurs when two or more action models are learned in a single \emph{macro-action}.
+
+Semantic alterations typically happen when the observed input data given in $\tau$ is scarce.
+Defining a proper semantic evaluation metric is key because the application of syntax-based metrics may report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the GTM.
+In the following sections, we introduce a novel evaluation metric that is robust to different types of reformulation.
+
+\subsection{Semantic-based precision and recall}
+\label{semantic_precision_recall}
+
+We define semantic-based metrics which are conceptually grounded on the precision and recall metrics introduced above.
+The rationale behind these novel metrics lies in counting the number of \emph{editing operations} that are necessary to match two sets of action models.
+Given $\mathcal{M}$, the two allowed editing operations are:
+
+\begin{itemize}
+\item {\em Deletion}. A fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ is removable from $\xi\in\mathcal{M}$.
+\item {\em Insertion}. A fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ can be added to $\xi\in\mathcal{M}$.
+\end{itemize}
+
+%The \ARMS system showed that the error of the learned action models can also be estimated executing a set of plan traces with the learned model ~\cite{yang2007learning}.
+ARMS tackles the semantic evaluation by counting the number of preconditions not met during the execution of the plan trace with the learned models.
+
+The \ARMS system showed that a semantic evaluation can be done via validation of a set of plan traces with the learned model~\cite{yang2007learning}.
+The underlying idea is that an error indication of the learned action models is obtained by counting the number of preconditions that are not satisfied during the execution of the plan trace with the learned models.
+This approach can be understood as modifying the plan trace (by adding the necessary preconditions to the intermediate states) so as to allow the execution of the observed actions using the learned models.
+In other words, modifying the plan trace to fit the model.
+Inspired by this approach, we present an alternative method that, instead, modifies the learned models so that they can explain the given plan traces.
+
+%Inspired by this approach, we present an alternative method, in which instead of modifying the trace, we modify the learned models so they can explain the given plan traces.
+
+\vspace{0.1cm}
+
+We interpret the semantic evaluation of action models as a learning task $\Lambda = \tup{\mathcal{M}, \mathcal{T}}$, where:
+
+\begin{itemize}
+	\item $\mathcal{M}$ is a previously \textbf{learned set of action models} obtained using any learning approach such as \FAMA.
+
+	\item$\mathcal{T}$ is a set of plan traces  used for \textbf{testing}.
+\end{itemize}
+
+A solution to this task is an \textbf{edited set of action models} $\mathcal{M'}$ able to explain $\mathcal{T}$, and obtained by applying \emph{deletion} and \emph{insertion} operations to $\mathcal{M}$.
+While it is always recommended for the test set to be different from the one used during learning, this is specially important for satisfying approaches such as \FAMA; otherwise $\mathcal{M'}$ = $\mathcal{M}$, since $\mathcal{M}$ can already explain $\mathcal{T}$ without any modification.
+
+We now provide formal definitions of $INS(\mathcal{M},\mathcal{M'})$ and $DEL(\mathcal{M},\mathcal{M'})$, the insertions and deletions needed to transform the set of action models $\mathcal{M}$ into $\mathcal{M'}$.
+
+%We say that two action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M'}$ that maps every $\xi\in\mathcal{M}$ to a comparable action model $\xi'\in\mathcal{M'}$ and vice versa.
+
+%\begin{mydefinition}
+%  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} sets of action models. We compute $ins(\mathcal{M}, \mathcal{M'})$ and $del(\mathcal{M}, \mathcal{M'})$ as:
+%  \begin{small}
+%  	\begin{align*}
+%  	ins(\mathcal{M}, \mathcal{M'})=&\{pre_p(\xi) ~|~ p \notin pre(\xi), p \in pre(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')} \cup\\
+%  	&\{add_p(\xi) ~|~ p \notin add(\xi), p \in add(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')} \cup\\
+%  	&\{del_p(\xi) ~|~ p \notin del(\xi), p \in del(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')}\\
+%  	\\
+%  	del(\mathcal{M}, \mathcal{M'})=&\{pre_p(\xi) ~|~ p \in pre(\xi), p \notin pre(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')} \cup\\
+%  	&\{add_p(\xi) ~|~ p \in add(\xi), p \notin add(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')} \cup\\
+%  	&\{del_p(\xi) ~|~ p \in del(\xi), p \notin del(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')}
+%  	\end{align*}
+%  \end{small}
+%\end{mydefinition}
 
 \begin{mydefinition}
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two \strips\ action models, such that they are {\em comparable}.
-The {\bf edit distance}, denoted as $\delta(\mathcal{M},\mathcal{M}')$, is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
+	Let $PRE(\xi) = \underset{\forall p \in pre(\xi)}{\bigcup} pre_p(\xi)$, $ADD(\xi) = \underset{\forall p \in add(\xi)}{\bigcup} add_p(\xi)$, and $DEL(\xi) = \underset{\forall p \in del(\xi)}{\bigcup} del_p(\xi)$ be the set of propositional fluents that represent preconditions, positive and negative effects of a given action model $\xi$.
+We define:
+	\begin{small}
+		\begin{align*}
+		INS(\mathcal{M}, \mathcal{M'})=&PRE(\xi') \backslash PRE(\xi) \cup\\
+		&ADD(\xi') \backslash ADD(\xi) \cup \\
+		&DEL(\xi') \backslash DEL(\xi),\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t.\ name(\xi) = name(\xi')\\
+		\\
+		DEL(\mathcal{M}, \mathcal{M'})=&PRE(\xi) \backslash PRE(\xi') \cup\\
+		&ADD(\xi) \backslash ADD(\xi') \cup \\
+		&DEL(\xi) \backslash DEL(\xi'),\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t.\ name(\xi) = name(\xi')\\
+		\end{align*}
+	\end{small}
 \end{mydefinition}
 
+Note that the number of {\em deletions} ($\left|DEL(\mathcal{M},\mathcal{M'})\right|$) that is required to transform $\mathcal{M}$ into $\mathcal{M}'$ matches our previous definition for the number of {\em false positives} when (1) $\mathcal{M}$ is the set of learned action models and (2) $\mathcal{M}'$ is the given GTM that serves as reference.
+Likewise the number of insertions ($\left|INS(\mathcal{M},\mathcal{M'})\right|$) required to transform $\mathcal{M}$ into $\mathcal{M}'$ corresponds to the number of {\em false negatives} in the learned models $\mathcal{M}$ with respect to the GTM $\mathcal{M}'$.
+%As explained, these numbers can be normalized with the amount of {\em true positives} to compute the {\em precision} and {\em recall} of a given learned model with respect to a reference model.
+Further, we can claim that all these numbers are bounded.
 Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well.
-In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_v(\xi)|$ while the max number of edits that can be introduced to the effects is twice $|F_v(\xi)|$.
+In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_{\xi}|$.
+This is the same max number of edits that can be introduced to the positive and to the negative effects of $\xi\in\mathcal{M}$.
+
+%Inspired by this approach for the semantic evaluation of learned action models, we define now the semantic version of precision and recall to asses the quality of a learned action model with respect to a partially observed plan trace.
+
+Given that the evaluation task is defined in terms of a learning task $\Lambda$, there might exist potentially many edited models $\mathcal{M'}$ which are solution to this task.
+Although the actual GTM is included among the solution set, it is impossible to identify it, so we define the best solution based on its proximity to the input model.
+
 \begin{mydefinition}
-The \textbf{maximum edit distance} of an \strips\ action model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_v(\xi)|$.
+  Given a set of action models $\mathcal{M}$, and all the sets of action models $\mathcal{M'}$ able to explain the plan traces $\mathcal{T}$.
+The {\bf closest compliant set of action models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) and able to explain $\mathcal{T}$;
+  \[\mathcal{M^*}=\underset{\forall \mathcal{M}' \rightarrow \mathcal{T}}{\arg\min} \ \left| INS(\mathcal{M},\mathcal{M'}) \cup DEL(\mathcal{M},\mathcal{M'}) \right|\]
 \end{mydefinition}
 
-Note that the number of deletions indicate the number of {\em false positives} while the number of insertions corresponds to the number of {\em false negatives}.
-As explained, these numbers can be normalized with the amount of {\em true positives} to compute the {\em precision} and {\em recall} of a given learned model with respect to a reference model.
+The {\bf closest compliant set of action models} allows us to define a semantic version of {\em precision} and {\em recall}, following the previous correspondences ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}).
+Also, by definition, the number of preconditions and effects of the learned action models is equal to the sum of {\em true positives} and {\em false positives}; that is, given $size(\mathcal{M}) = \left|pre(\xi)\right| + \left|add(\xi)\right| + \left|del(\xi)\right| \forall \xi \in \mathcal{M}$, we know that $size(\mathcal{M}) = tp + fp$.
 
-Normally evaluating a given learned domain with respect to the actual GTM model is not possible because the actual GTM model is unknown.
-As the ARMS system shows, the error of a learned action model can also be estimated with respect to a set of observations of plan executions that are assumed to be generated with that model~\cite{yang2007learning}.
-With this regard, we define now an edit distance to asses the quality of a learned action model with respect to a partially observed plan trace $\mathcal{T}$.
+The semantic-based definition of precision and recall are given by:
 
-\begin{mydefinition}
-  Given $\mathcal{M}$, a \strips\ action model built from $F_v$, and a plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$ whose state observation are built with fluents in $F$.
-The {\bf observation edit distance}, denoted by  $\delta(\mathcal{M},\mathcal{T})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$, such that $\mathcal{M}'$ can produce a valid plan trace $\mathcal{T}$; \[\delta(\mathcal{M},\mathcal{T})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{T}} \delta(\mathcal{M},\mathcal{M}')\]
-\end{mydefinition}
+\begin{small}
+	\begin{align*}
+	sem\text{-}Precision=&\frac{tp}{tp + fp}=\frac{size(\mathcal{M})- \left|del(\mathcal{M},\mathcal{M^*})\right|}{size(\mathcal{M})}\\
+	sem\text{-}Recall=&\frac{tp}{tp + fn}=\frac{size(\mathcal{M})- \left|del(\mathcal{M},\mathcal{M^*})\right|}{size(\mathcal{M}) - \left|del(\mathcal{M},\mathcal{M^*})\right| + \left|ins(\mathcal{M},\mathcal{M^*})\right|}
+	\end{align*}
+\end{small}
 
-Following the previous correspondences ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}) we can define a {\em semantic} version of the {\em precision} and {\em recall} metrics of a learned model.
-In this case {\em precision} and {\em recall} is not computed with respect to a GTM but with respect to a partially observed plan trace $\mathcal{T}=\tup{s_0,a,_1,s_1,\ldots,a_n,s_{n}}$.
-To define the {\em semantic} version of {\em precision} and {\em recall} the only term to re-formalize is the number of {\em true positives} that, for each action schema $\xi\in\mathcal{M}$, is given by $tp(\xi)=|pre(\xi)|+|del(\xi)|+|add(\xi)|-fp(\xi)$.
+As we can observe in the two above formulas, precision and recall are not computed with respect to a GTM but with respect to the closest compliant set of action models that explain the test set of plan traces $\mathcal{T}$.
+Thereby, in the case that the {\bf closest compliant set of action models} $\mathcal{M^*}$ is the GTM, we conclude that the values of the syntactic measures {\em Precision} and {\em Recall} are exactly the same as {\em sem\text{-}Precision} and {\em sem\text{-}Recall}.
 
-\subsection{Computing the edit distance with classical planning}
-Our compilation is extensible to compute the {\em observation edit distance} and hence, the semantic versions of the {\em precision} and {\em recall} metrics.
-This extension considers that the input \strips\ model $\mathcal{M}$, is {\em non-empty} so instead of learning an action model from scratch we simply edit $\mathcal{M}$ until it satisfies the given input observations.
-In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains $head(\xi)$ but also the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
+The intuition behind this evaluation is to {\em semantically} assess how well the learned action models $\mathcal{M}$ explain a set of given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce the observations.
+Unlike the semantic metric defined by ARMS, our novel semantic definitions of precision and recall are not sensitive to flaws that appear more than once in the plan traces since the flaws are corrected only once in the learned models instead of at every intermediate state of the plan traces.
+
+%This semantic evaluation approach is again flexible to various amount and kind of available input knowledge.
+
+\subsection{Semantic evaluation with classical planning}
+\label{edit_distance}
+
+%Our compilation is extensible to compute the {\em closest compliant set of action models} and hence, our semantic versions of the {\em precision} and {\em recall} metrics.
+This extension considers that the input models $\mathcal{M}$, is {\em non-empty} so instead of learning an action model from scratch we simply edit $\mathcal{M}$ until it satisfies the given input observations.
+In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
 A solution to the classical planning task resulting from the extended compilation is a sequence of actions that:
+
+The compilation scheme presented in section \ref{sec:compilation} is extensible to address the evaluation task $\Lambda=\tup{\mathcal{M}, \mathcal{T}}$ defined in section \ref{semantic_precision_recall}.
+In this extended task, $\mathcal{M}$ represents a set of previously learned action models; therefore, rather than learning the action models from scratch, we simply edit $\mathcal{M}$ until it satisfies the given test set of plan traces $\mathcal{T}$.
+A solution to the classical planning task resulting from the extended compilation is a plan that:
+
 \begin{enumerate}
-\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan trace}.
+\item {\bf Edits the action models $\mathcal{M}$ to build $\mathcal{M}'$}.
+A solution plan starts with a prefix that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using the two {\em editing operations} defined above, {\em deletion} and {\em insertion}.
+\item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan traces}.
 The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:compilation} for the models that are programmed from scratch.
@@
 
-In more detail given $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, the output of the extended compilation is a classical planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda},G_{\Lambda}}$ such that:
+Given $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, the output of the extended compilation is a planning task $P_{\Lambda}'=\tup{F_{\Lambda},A_{\Lambda}',I_{\Lambda},G_{\Lambda}}$ such that:
+
 \begin{itemize}
@@
 Note that, the input action model $\mathcal{M}$ is encoded in the initial state.
-This means that the fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$, $f\in F_v(\xi)$, hold in $I_{\Lambda}$ iff they appear in $\mathcal{M}$.
+This means that the fluents $pre_p(\xi)/del_p(\xi)/add_p(\xi)$, $p\in \Psi_\xi$, hold in $I_{\Lambda}$ iff they appear in $\mathcal{M}$.
 \item $A_{\Lambda}'$, comprises the same three kinds of actions of $A_{\Lambda}$.
-The actions for {\em applying} an already programmed operator schema and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
-The only difference here is that each action for {\em programming} an operator schema now implement the two {\em edit operations} (i.e.
-include actions for {\em inserting} a precondition and for {\em deleting} a negative/positive effect).
+The actions for {\em applying} an already programmed action model and the actions for {\em validating} an observation are defined exactly as in the previous compilation.
+The only difference here is that the {\em programming actions} now implement the two editing operations (i.e., they also include the actions for {\em deleting} a precondition or negative/positive effect from an action model).
 \end{itemize}
 
-To illustrate this, the plan of Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} action model where again the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-In this case the edited action model is however validated at the plan shown in Figure~\ref{fig:example-plans}.
+Figure~\ref{fig:plan-pdistance} shows the plan for editing the action model of the operator {\tt\small stack} of the {\em blocksworld} domain where only the two positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} are missing.
+In this case the edited action model is again validated in the plan trace shown in Figure~\ref{fig:example-plans}.
 
 \begin{figure}[hbt!]
-{\tt\small
-00 : (insert\_add\_handempty\_stack)\\
-01 : (insert\_add\_clear\_stack\_var1)\\
-02 : (apply\_unstack blockB blockA i1 i2)\\
+{\footnotesize\tt
+  {\bf 00} : (insert\_add\_stack\_handempty)\\
+  01 : (insert\_add\_stack\_clear\_var1)\\
+  {\bf 02} : (apply\_unstack blockB blockA i1 i2)\\
   03 : (apply\_putdown blockB i2 i3)\\
@@
   05 : (apply\_stack blockA blockB i4 i5)\\
-06 : (validate\_1)
+  {\bf 06} : (validate\_1)
 }
- \caption{\small Plan for editing a given {\em blockswold} schema and validating it at the plan shown in Figure~\ref{fig:example-plans}.}
+\caption{\small Plan for editing and validating the action model {\tt\small{stack}} in which the positive effects {\tt\small{(handempty)}} and {\tt\small{(clear ?v1)}} are missing.}
 \label{fig:plan-pdistance}
@@
 
-Our interest when solving the classical planning task that is output by our extended compilation is not in the resulting action model $\mathcal{M}'$ but in the number of required {\em edit operations} (insertions and deletions).
+Assuming we are using an optimal planner to solve $P_{\Lambda}'$, the solution plan of this problem will induce the \emph{closest compliant set of action models} $\mathcal{M^*}$.
+Therefore, our compilation enables the computation of the semantic versions of \emph{precision} and \emph{recall}.
+An argument can be made, however, that solving optimally $P_{\Lambda}'$ may turn the evaluation process very time consuming.
+Considering this, $sem\text{-}Precision$ and $sem\text{-}Recall$ can be approximated if $P_{\Lambda}'$ is solved with a satisfying planner.
+In this case, no guarantees can be made that the edited models will be the closest compliant ones, but a planner will always try to minimize the length of the plan and hence the number of editing operations applied to the input models.
+
+%Our interest when solving the classical planning task that is output by our extended compilation is not in the resulting action model $\mathcal{M}'$ but in the number of {\em edit operations} (insertions and deletions) that is required to compute $\mathcal{M}'$.
 For instance, in the example of Figure~\ref{fig:plan-pdistance}, $\mathcal{M}'$ is validated in the given observations after inserting two new positive effects to the initial action model, e.g.
 $\delta(\mathcal{M},\mathcal{T})=2$ .
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_v(stack)|=|F_v(unstack)|=11$ while $|F_v(pickup)|=|F_v(putdown)|=5$  (as shown in Section~\ref{sec:learning}).
+In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_{\tt stack}|=|F_{\tt unstack}|=11$ while $|F_{\tt pickup}|=|F_{\tt putdown}|=5$  (as shown in Section~\ref{sec:learning}).
 The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
 
-Last but not least, this compilation is flexible to compute the {\em edit distance} between two {\em comparable} \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$.
+%Last but not least, the compilation approach is also flexible to compute the {\em edit distance} between two {\em comparable} \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$.
 A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion.
@@
 
-\begin{figure}[hbt!]
-{\tt\small
-00 : (insert\_add\_handempty\_stack)\\
-01 : (insert\_add\_clear\_stack\_var1)
-}
- \caption{\small Plan for computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.}
-\label{fig:plan-mdistance}
-\end{figure}
+%\begin{figure}[hbt!]
+%{\tt\small
+%00 : (insert\_add\_handempty\_stack)\\
+%01 : (insert\_add\_clear\_stack\_var1)
+%}
+% \caption{\small Plan for computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.}
+%\label{fig:plan-mdistance}
+%\end{figure}

### 2018-07-08 doc/aij18/learning_task.tex
Add files via upload
Sections 2, 3, 4 and 5 finished

@@
 \label{sec:learning}
-The {\em learning task} addressed in this paper corresponds to learning a classical planning action model by observing an agent(s) acting in a world that is defined by a {\em classical planning frame} $\Phi=\tup{F,A}$.
-This learning task is formalized by the pair $\Lambda=\tup{\mathcal{M},\tau}$:
+
+This section details the \FAMA approach for learning action models.
+The notion of learning task is defined in section \ref{learning_task} and the components of the compilation scheme are described in sections \ref{propositional_encoding} and \ref{conditional_effects}.
+Subsequently, the compilation approach is fully detailed in section \ref{compilation} and the last section presents some theoretical properties of the compilation scheme.
+
+\FAMA addresses the learning and evaluation of PDDL action models that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
+An \strips\ action model is a tuple $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+
 \begin{itemize}
-\item $\mathcal{M}$ is the {\bf initial action model}. This model is {\em empty}, when learning from scratch, or {\em partially specified}, when some fragments of the model are a priori known.
+	\item The name, $name(\xi)$, and parameters, $pars(\xi)$, of the action model define the {\em header} of the model.
+	\item $pre(\xi)$, $del(\xi)$ and $add(\xi)$ represent the preconditions, negative effects and positive effects of the action model, respectively, such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\end{itemize}
+
+As an example, Figure~\ref{fig:stack} shows the action model of the {\em stack} operator from the four-operator {\em blocksworld} domain~\cite{slaney2001blocks} encoded in PDDL.
+
+\begin{figure}[hbt!]
+	\begin{footnotesize}
+		\begin{verbatim}
+		(:action stack
+		:parameters (?v1 ?v2 - object)
+		:precondition (and (holding ?v1) (clear ?v2))
+		:effect (and (not (holding ?v1)) (not (clear ?v2)) (handempty) (clear ?v1) (on ?v1 ?v2)))
+		\end{verbatim}
+	\end{footnotesize}
+	\caption{PDDL encoding of the action model of the {\em stack} operator from the four-operator {\em blocksworld} domain.}
+	\label{fig:stack}
+\end{figure}
+
+\subsection{Learning Task}
+\label{learning_task}
+
+Our {\em learning task} consists in learning classical planning action models by observing one or more agents acting in a world, which is defined by a {\em classical planning frame} $\Phi=\tup{F,A}$.
+The learning task is formalized by the pair $\Lambda=\tup{\mathcal{M},\tau}$:
+
+\begin{itemize}
+\item $\mathcal{M}$ is the set of {\bf initial action models}.
+This set is {\em empty}, when learning from scratch, or {\em partially specified}, when some fragments of the action models are known a priori.
 \item $\tau$ is the observed {\bf plan trace} such that:
@@
  \item Observations in $\tau$ are {\em noiseless}, meaning that if the value of a fluent or an action is observed in $\tau$, then the observation is correct.
-\item The initial state $s_0\in\tau$ is {\em fully observed}, i.e.~$|s_0|=|F|$.
-This means that the set of fluents $F$, and the corresponding set of predicates $\Psi$ that shapes the fluents in $F$, are inferrable from $s_0$.
-\item The set of actions $A$ is inferrable either from $\mathcal{M}$ or $\tau$.
+\item The initial state $s_0\in\tau$ is a {\em fully observed} state including positive and negative fluents, i.e.~$|s_0|=|F|$.
+Consequently, the corresponding set of predicates $\Psi$ and objects $\Omega$ that shape the fluents in $F$ are inferrable from $s_0$.
+\item The header of an action model is either given by $\mathcal{M}$ or inferrable from $\tau$.
+In the latter case, $\tau$ must contain at least one instantiation of the respective action model header.
+%\item The set of actions $A$ is inferrable either from $\mathcal{M}$ or $\tau$.
 In other words, the action name and parameters that shape the actions in $A$ are given by $\mathcal{M}$ otherwise, $\tau$ contains at least one instantiation of every action model in $\mathcal{M}$.
-\item Intermediate actions $a_i\in\tau$, and intermediate states $s_j\in\tau$, {\small $1\leq i,j$} are {\em partially observable}.
-In the extreme all $a_i\in\tau$, {\small $1\leq i\leq n$} actions could be unobserved, provided that the final state $s_m$ is at least, partially observed.
-Likewise states $s_j$, {\small $1\leq j\leq m$}, might be unobserved provided that at least actions $a_i$, {\small $1\leq i\leq n$}, are partially observed.
-
+\item We allow plan traces with \NO state trajectories and action sequences.
+In the extreme, all actions and intermediate states may be missing, provided that the final state is at least partially observed.
+The least informative plan trace is thus $\tau = \tup {s_0,s_n}$.
 \end{enumerate}
@@
 
-A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is an action model $\mathcal{M}'$ that is compliant with the input model $\mathcal{M}$ and the observed plan trace $\tau$.
+Ultimately, we can always assume that $\mathcal{M}$ will contain the predicates $\Psi$ and the headers of the actions models, either explicitly provided in $\mathcal{M}$ or inferred from $\tau$.
+A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a set of action models $\mathcal{M}'$ that is compliant with the input models $\mathcal{M}$ and the observed plan trace $\tau$.
 
-Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\tau}$, that corresponds to observing the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
-In this example $\tau=\langle s_0,${\small\tt (putdown\ B), (stack\ A\ B)}, $s_4\rangle$ which means that only the first and last states are observed (the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown) and that actions $a_2$ and $a_3$ are observed while $a_1$ and $a_4$ are unknown.
+Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\tau}$ corresponding to the observation of the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
+In this example $\tau=\langle s_0,${\small\tt (putdown\ B),(stack\ A\ B)}, $s_4\rangle$, which means that only the first and the last state are observed (the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown) and that actions $a_2$ and $a_3$ are observed while $a_1$ and $a_4$ are unknown.
+The set of initial action models $\mathcal{M}$ only contains two of the four headers needed, but can be completed with the headers {\small\tt(putdown ?v1)} and {\small\tt(stack ?v1 ?v2)} inferred from $\tau$.
 
@@
 \begin{verbatim}
-(pickup ?v1) (putdown ?v1) (stack ?v1 ?v2} (unstack ?v1 ?v2)
+(pickup ?v1) (unstack ?v1 ?v2)
 \end{verbatim}
@@
 \begin{verbatim}
-;;; State observation 
-(clear A) (on A B) (ontable B) (handempty)
-\end{verbatim}
-\end{footnotesize}
-
-\begin{footnotesize}
-\begin{verbatim}
 ;;; Action observation
@@
 
- \caption{\small Task $\Lambda=\tup{\mathcal{M},\tau}$ for learning a {\em blocksworld} \strips\ action model from a four-action plan and two state observations.}
+\begin{footnotesize}
+	\begin{verbatim}
+	;;; State observation
+	(clear A) (on A B) (ontable B) (handempty)
+	\end{verbatim}
+\end{footnotesize}
+
+ \caption{\small Task $\Lambda=\tup{\mathcal{M},\tau}$ associated to the observation $\tau=\langle s_0,${\small\tt (putdown\ B),(stack\ A\ B)}, $s_4\rangle$}
 \label{fig:example-plans}
@@
 
-The definition of the learning task is extensible to the more general case where the execution of several plans generated from the same action model is observed.
-In this case, $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_{k}\}$ such that each $\tau\in \mathcal{T}$ is a plan trace that satisfies the previous [1,4] assumptions.
-In this case a {\em solution} action model $\mathcal{M}'$ have to be compliant with the input model $\mathcal{M}$ but also with every observed plan trace $\tau\in \mathcal{T}$.
+The definition of the learning task is extensible to the more general case where the execution of several plans from the same action models are observed.
+In this case, $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_{k}\}$ such that each $\tau\in \mathcal{T}$ is a plan trace that satisfies the previous 1--4 assumptions.
+In this case, the {\em learned} action models $\mathcal{M}'$ have to be compliant with the input models $\mathcal{M}$ and also with every observed plan trace $\tau\in \mathcal{T}$.
 
-\subsection{A propositional encoding for \strips\ action schemas}
-This work addresses the learning and evaluation of PDDL action schemas that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-An \strips\ action schema is a tuple $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$:
+\subsection{Propositional encoding of an action model}
+\label{propositional_encoding}
+
+%We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi'$).
+Therefore  we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
+We say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
+
+In this section we formalize the propositional encoding of an \strips\ action model.
+This encoding is at the core of the \FAMA compilation approach for addressing the learning task defined in section \ref{learning_task}.
+
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, which is bound to the maximum arity of an action in a given planning frame.
+For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1,v_2\}$ because the actions with the maximum arity have arity two; i.e., any instantiation of the {\small\tt stack} or the {\small\tt unstack} models.
+
+We define $\Psi_v$ as the set of predicates $\Psi$ parameterized with the {\em variable names} of $\Omega_v$ as arguments.
+$\Psi_v$  defines the elements that can appear in the preconditions and effects of the action models.
+In the {\em blocksworld} domain, this set contains eleven elements, $\Psi_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+In more detail, for a given action model $\xi$, we define $\Psi_{\xi}\subseteq \Psi_v$ as the subset of elements of $\Psi_v$ that appear in $\xi$.
+For instance, $\Psi_{\tt stack}=\Psi_v$ whereas $\Psi_{\tt pickup}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$),on($v_1,v_1$)\}} excludes the elements from $\Psi_v$ that involve $v_2$ because {\small\tt pickup} actions have arity one.
+The size of the space of possible \strips\ models for a given $\xi$ is $2^{2|\Psi_{\xi}|}$ (recall that negative effects appear as preconditions and that they cannot be positive effects, and also that a positive effect cannot appear as a precondition).
+For the {\em blocksworld}, $2^{2|\Psi_{\tt stack}|}=4194304$ while for the {\tt pickup} operator this number is only 1024.
+
+We are now ready to define the propositional encoding of $pre(\xi)$, $del(\xi)$ and $add(\xi)$. For every $\xi$ and $p\in \Psi_{\xi}$, we create:
+
 \begin{itemize}
-\item The $name(\xi)$ and parameters, $pars(\xi)$, of the schema that define the {\em header} for that schema. 
-\item The preconditions $pre(\xi)\subseteq \Psi$, negative effects $del(\xi)\subseteq \Psi$, and positive effects $add(\xi)\subseteq \Psi$ such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\item $pre_p(\xi)$: fluent formed by the combination of the prefixes {\small \texttt{pre}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
+{\small \texttt{pre\_stack\_on\_v1\_v2}}, for $name(\xi)={\small\texttt{stack}}$ and $p={\small\texttt{on($v_1,v_2$)}}$)
+\item $del_p(\xi)$: fluent formed by the combination of the prefixes {\small \texttt{del}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
+{\small  \texttt{del\_stack\_on\_v1\_v2}})
+\item $add_p(\xi)$: fluent formed by the combination of the prefixes {\small \texttt{add}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
+{\small \texttt{add\_stack\_on\_v1\_v2}})
 \end{itemize}
-As an example Figure~\ref{fig:stack} shows the {\em stack} action schema, coded in PDDL, from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+
+%The fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $p\in \Psi_{\xi}$, that represent the propositional encoding for the preconditions, negative and positive effects of an action model $\xi$.
+If a fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ holds, it means that $p$ is a precondition/negative/positive effect in that action model.
+
+For a given action model $\xi$, if a fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ holds in a state, it means that $p$ is a precondition/negative/positive effect of $\xi$.
+For instance, Figure~\ref{fig:encodedstack} shows the conjunction of fluents that represents the propositional encoding for the preconditions, negative effects and positive effects of the action model of the {\em stack} operator shown in Figure~\ref{fig:stack}.
 
@@
 \begin{verbatim}
-(:action stack
- :parameters (?v1 ?v2 - object)
- :precondition (and (holding ?v1) (clear ?v2))
- :effect (and (not (holding ?v1)) (not (clear ?v2)) (handempty) (clear ?v1) (on ?v1 ?v2)))
+(pre_stack_holding_v1) (pre_stack_clear_v2)
+(del_stack_holding_v1) (del_stack_clear_v2)
+(add_stack_handempty) (add_stack_clear_v1) (add_stack_on_v1_v2)
 \end{verbatim}
 \end{footnotesize}
- \caption{\small \strips\ action schema coding, in PDDL, the {\em stack} action from a four-operator {\em blocksworld}.}
-\label{fig:stack}
-\end{figure}
-We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi'$).
-Therefore  we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
-We say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
-
-Now we formalize a propositional encoding for the target of the learning task addressed in the paper, the \strips\ action model.
-This encoding is the core of our compilation approach for learning \strips\ action models with classical planning.
-
-First, let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the actions with the maximum arity have arity two (any instantiation of the {\small\tt stack} or the {\small\tt unstack} schemas).
-
-We define $F_v$, a new set of fluents s.t.
-$F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}, and that defines the elements that can appear in the action schemes.
-In {\em blocksworld} this set contains eleven elements, $F_v$={\small\tt\{handempty, holding\_$v_1$, holding\_$v_2$, clear\_$v_1$, clear\_$v_2$, ontable\_$v_1$, ontable\_$v_2$, on\_$v_1$\_$v_1$, on\_$v_1$\_$v_2$, on\_$v_2$\_$v_1$, on\_$v_2$\_$v_2$\}}.
-In more detail, for a given schema $\xi$, we define $F_{\xi}\subseteq F_v$ as the subset of elements that can appear in that action schema.
-For instance, $F_{\tt stack}=F_v$ while $F_{\tt pickup}$={\small\tt\{handempty, holding\_$v_1$, clear\_$v_1$, ontable\_$v_1$, on\_$v_1$\_$v_1$\}} excludes the elements from $F_v$ that involve $v_2$ because {\small\tt pickup} actions have arity one.
-The size of the space of possible \strips\ models for a given schema $\xi$ is $2^{2|F_{\xi}|}$ (recall that negative effects appear as preconditions and that they cannot be positive effects and also, that a positive effect cannot appear as a precondition).
-For the {\em blocksworld}, $2^{2|F_{\tt stack}|}=4194304$ while for the {\tt pickup} operator this number is only 1024.
-
-Now we are ready to define the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_{\xi}$, that represent the propositional encoding for the preconditions, negative and positive effects of an action schema $\xi$.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in that schema.
-For instance, Figure~\ref{fig:encodedstack} shows the conjunction of fluents that represents he propositional encoding for the preconditions, negative and positive effects of the {\em stack} schema shown in Figure~\ref{fig:stack}.
-
-\begin{figure}[hbt!]
-\begin{footnotesize}
-\begin{verbatim}
-(pre_holding_stack_v1) (pre_clear_stack_v2)
-(del_holding_stack_v1) (del_clear_stack_v2)
-(add_handempty_stack) (add_clear_stack_v1) (add_on_stack_v1_v2)
-\end{verbatim}
-\end{footnotesize}
- \caption{\small Propositional encoding for the {\em stack} schema from a four-operator {\em blocksworld}.}
+ \caption{\small Propositional encoding for the {\em stack} action model from a four-operator {\em blocksworld}.}
 \label{fig:encodedstack}
@@
 \subsection{Classical planning with conditional effects}
+\label{conditional_effects}
+
 Conditional effects allow us to compactly define actions whose effects depend on the current state.
@@
 \subsection{Compilation}
-%\label{sec:compilation}
+\label{sec:compilation}
 
-Our approach for addressing a $\Lambda$ learning task is compiling it into a classical planning task with conditional effects.
-The intuition behind the compilation is that a solution to the resulting classical planning task is a sequence of actions that:
+In section \ref{sec:motivation}, we exposed our reasons to solve the learning task via compilation to planning.
+Our compilation scheme  builds upon the approach presented in \cite{aineto2018learning}.
+\FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.
+
+Our approach to solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to compile $\Lambda$ into a classical planning task $P_{\Lambda}$ with conditional effects.
+The intuition behind the compilation is that a solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output set of action models $\mathcal{M}'$ that solves $\Lambda$.
+Specifically, a solution plan $\pi_\Lambda$ serves two purposes:
 
 \begin{enumerate}
-\item {\bf Programs the action model $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that, for each $\xi\in\mathcal{M}$, determines which fluents $f\in F_{\xi}$ belong to its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-\item {\bf Validates the action model $\mathcal{M}'$}.
-The solution plan continues with a {\em postfix} that reproduces the given input knowledge (the available plan traces) with the programmed action model $\mathcal{M}'$.
+\item {\bf To program the set of action models $\mathcal{M}'$}.
+$\pi_\Lambda$ comprises a plan {\em prefix} whose actions determine the predicates $p\in \Psi_{\xi}$ that belong to $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\mathcal{M}$.
+\item {\bf To validate the set of action models $\mathcal{M}'$}.
+$\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the recognition of the available plan trace $\tau$ with the programmed action models $\mathcal{M}'$.
 \end{enumerate}
@@
 Here we formalize the compilation for learning \strips\ action models with classical planning.
-Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
+Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$, with $\tau$ formed by an $n$-action sequence $\tup{a_1, \ldots, a_n}$ and a $m$-state trajectory $\tup{s_0, s_1, \ldots, s_m}$ (i.e., $\tau = \langle s_0, a_1, \ldots, a_n, s_m \rangle$), the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
+
 \begin{itemize}
+
 \item $F_{\Lambda}$ contains:
 \begin{itemize}
-\item The set of fluents $F$ built instantiating the predicates $\Psi$ with the objects $\Omega$ that appear in the plan trace given as input, i.e.
-the blocks {\tt\small A} and {\tt\small B} in the example of Figure~\ref{fig:example-plans}.
-\item Fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_{\xi}$, that represent the programmed action model.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in the schema $\xi\in \mathcal{M}'$.
-\item Fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in a plan trace $\tau$.
-The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ actions of $\tau$ that are observed.
-In the example of Figure~\ref{fig:example-plans} two actions were observed $\tup{\small\tt (putdown\ B), (stack\ A\ B)}$.
-\item Fluents $\{test_j\}_{0\leq j\leq m}$, indicating the state observation $s_j\in\tau$ where the action model is validated.
-In the example of Figure~\ref{fig:example-plans} two tests are required to validate the programmed action model at $\tup{s_0,s_4}$.
-\item The fluents $mode_{prog}$ and $mode_{val}$ to indicate whether the operator schemas are programmed or validated. 
+\item The set of fluents of $s_0$; i.e., $F$.
+\item The fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, as explained in section \ref{propositional_encoding}
+\item A set of fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to represent the $i^{th}$ observable action of $\tau$.
+In the example of Figure~\ref{fig:example-plans}, the two observed actions {\small \texttt{(putdown B)}} and {\small \texttt{(stack  A  B)}} would be encoded as fluents  {\small \texttt{(plan-putdown B i1)}} and {\small \texttt{(plan-stack A B i2)}} to indicate that {\small \texttt{(putdown B)}} is observed in the first place and {\small \texttt{(stack  A  B)}} is the second observed action.
+\item Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions of $\tau$.
+The former is used to ensure that actions are executed in the same order as they are observed in $\tau$.
+The latter is used to iterate to the next planning step when solving $P_{\Lambda}$.
+\item A set of fluents $\{test_j\}_{0\leq j\leq m}$, to point at the state observation $s_j\in\tau$ where the action model is
+validated. In the example of Figure~\ref{fig:example-plans} two tests are required to validate the programmed action model, one test at $s_0$ and another one at $s_4$.
+\item A fluent, $mode_{prog}$, to indicate whether the action models are being programmed or validated.
 \end{itemize}
-\item $I_{\Lambda}$ encodes the first state observation, $s_0\subseteq F$ and sets to true $mode_{prog}$ as well as fluent $test_0$ and the fluents $F_{\pi}$ plus $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the action where the programmed model is validated.
-Our compilation assumes that initially, schemas are programmed with no precondition, no negative effect and no positive effect.
 
-\item $G_{\Lambda}=\{at_n, test_m\}$, requires that the programmed action model is validated in all the actions and states observed from the input plan trace $\tau$.
+\item $I_{\Lambda}$ encodes $s_0$ and the following fluents set to true: $mode_{prog}$, $test_0$, $F_{\pi}$, $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i \leq n$}.
+Our compilation assumes that action models are initially programmed with no precondition, no negative effect and no positive effect.
+
+\item $G_{\Lambda}$ includes the positive fluents $at_n$ and $test_m$.
+When these two goals are achieved by the solution plan $\pi_\Lambda$, we will be certain that the programmed action models are validated in all the actions and states observed in the input plan trace $\tau$.
+
 \item $A_{\Lambda}$ comprises three kinds of actions:
 \begin{enumerate}
-\item Actions for {\em programming} operator schema $\xi\in\mathcal{M}$:
+\item Actions for {\em programming} an action model $\xi\in\mathcal{M}$.
+These actions will form the prefix of the solution plan $\pi_\Lambda$ and they are aimed at generating the appropriate state configuration to shape $\xi$.
+Among the programming actions, we find:
 \begin{itemize}
-\item Actions for adding a {\em precondition} $f\in F_{\xi}$ from the action schema $\xi\in\mathcal{M}$.
+\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg pre_{f}(\xi), \neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
-\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{pre_{f}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{programPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
+\cond(\mathsf{programPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
 \end{align*}
@@
 
-\item Actions for adding a {\em negative} or {\em positive} effect $f\in F_{\xi}$ to the action schema $\xi\in\mathcal{M}$.
+\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{prog}\},\\
-\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{programEff_{p,\xi}})=&\{\neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
+\cond(\mathsf{programEff_{p,\xi}})=&\{pre_{p}(\xi)\}\rhd\{del_{p}(\xi)\},\{\neg pre_{p}(\xi)\}\rhd\{add_{p}(\xi)\}.
 \end{align*}
@@
 
-\item Actions for {\em applying} a programmed operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
-Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
-Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$ from {\em blocksworld}.
+\item Actions for {\em applying} a programmed action model $\xi\in\mathcal{M}$ bound to objects $\omega\subseteq\Omega^{ar(\xi)}$.
+These actions will be part of the postfix of the solution plan $\pi_\Lambda$ and they execute $\xi$ according to the current state configuration, i.e., the values of $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$.
+Since action headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear in the same position.
+Figure~\ref{fig:compilation} shows the PDDL encoding for applying a programmed action model of the {\em stack} operator from the {\em blocksworld} domain.
+
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\cup \{\neg mode_{val}\},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
-&\{\emptyset\}\rhd\{mode_{val}\}.
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
+&\{add_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi},\\
+&\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
 \end{align*}
 \end{small}
+
+%\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\cup \{\neg mode_{val}\},\\
+%\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+%&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+%&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
+%&\{\emptyset\}\rhd\{mode_{val}\}.
 
@@
   :precondition
-   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
-        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
-        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
-        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
-        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
-        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
-        (or (not (pre_clear_stack_v1)) (clear ?o1))
-        (or (not (pre_clear_stack_v2)) (clear ?o2))
-        (or (not (pre_holding_stack_v1)) (holding ?o1))
-        (or (not (pre_holding_stack_v2)) (holding ?o2))
-        (or (not (pre_handempty_stack)) (handempty)))
+   (and (or (not (pre_stack_on_v1_v1)) (on ?o1 ?o1))
+        (or (not (pre_stack_on_v1_v2)) (on ?o1 ?o2))
+        (or (not (pre_stack_on_v2_v1)) (on ?o2 ?o1))
+        (or (not (pre_stack_on_v2_v2)) (on ?o2 ?o2))
+        (or (not (pre_stack_ontable_v1)) (ontable ?o1))
+        (or (not (pre_stack_ontable_v2)) (ontable ?o2))
+        (or (not (pre_stack_clear_v1)) (clear ?o1))
+        (or (not (pre_stack_clear_v2)) (clear ?o2))
+        (or (not (pre_stack_holding_v1)) (holding ?o1))
+        (or (not (pre_stack_holding_v2)) (holding ?o2))
+        (or (not (pre_stack_handempty)) (handempty)))
   :effect
-   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
-        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
-        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
-        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
-        (when (del_ontable_stack_v1) (not (ontable ?o1)))
-        (when (del_ontable_stack_v2) (not (ontable ?o2)))
-        (when (del_clear_stack_v1) (not (clear ?o1)))
-        (when (del_clear_stack_v2) (not (clear ?o2)))
-        (when (del_holding_stack_v1) (not (holding ?o1)))
-        (when (del_holding_stack_v2) (not (holding ?o2)))
-        (when (del_handempty_stack) (not (handempty)))
-        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
-        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
-        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
-        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
-        (when (add_ontable_stack_v1) (ontable ?o1))
-        (when (add_ontable_stack_v2) (ontable ?o2))
-        (when (add_clear_stack_v1) (clear ?o1))
-        (when (add_clear_stack_v2) (clear ?o2))
-        (when (add_holding_stack_v1) (holding ?o1))
-        (when (add_holding_stack_v2) (holding ?o2))
-        (when (add_handempty_stack) (handempty))
+   (and (when (del_stack_on_v1_v1) (not (on ?o1 ?o1)))
+        (when (del_stack_on_v1_v2) (not (on ?o1 ?o2)))
+        (when (del_stack_on_v2_v1) (not (on ?o2 ?o1)))
+        (when (del_stack_on_v2_v2) (not (on ?o2 ?o2)))
+        (when (del_stack_ontable_v1) (not (ontable ?o1)))
+        (when (del_stack_ontable_v2) (not (ontable ?o2)))
+        (when (del_stack_clear_v1) (not (clear ?o1)))
+        (when (del_stack_clear_v2) (not (clear ?o2)))
+        (when (del_stack_holding_v1) (not (holding ?o1)))
+        (when (del_stack_holding_v2) (not (holding ?o2)))
+        (when (del_stack_handempty) (not (handempty)))
+        (when (add_stack_on_v1_v1) (on ?o1 ?o1))
+        (when (add_stack_on_v1_v2) (on ?o1 ?o2))
+        (when (add_stack_on_v2_v1) (on ?o2 ?o1))
+        (when (add_stack_on_v2_v2) (on ?o2 ?o2))
+        (when (add_stack_ontable_v1) (ontable ?o1))
+        (when (add_stack_ontable_v2) (ontable ?o2))
+        (when (add_stack_clear_v1) (clear ?o1))
+        (when (add_stack_clear_v2) (clear ?o2))
+        (when (add_stack_holding_v1) (holding ?o1))
+        (when (add_stack_holding_v2) (holding ?o2))
+        (when (add_stack_handempty) (handempty))
         (when (modeProg) (not (modeProg)))))
@@
 \end{scriptsize}
- \caption{\small PDDL action for applying an already programmed schema $stack$ (implications are coded as disjunctions).}
+ \caption{\small PDDL action for applying an already programmed model for $stack$ (implications are coded as disjunctions).}
 \label{fig:compilation}
 \end{figure}
-When the input plan trace contain observed actions, then the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as in $\mathcal{T}$.\\
+
+When the input plan trace contains observed actions, the extra conditional effects
+
+$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as in $\tau$.\\
 
 \item Actions for {\em validating} the partially observed state $s_j\in\tau$, {\tt\small $1\leq j< m$}.
+These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the input plan trace $\tau$ follows after the apply actions.
+
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\}\cup \{mode_{val}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\neg mode_{val}\}.
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j\}.
 \end{align*}
@@
 
-Known preconditions and effects (that is, a partially specified \strips\ action model) can be encoded as fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ set to true at the initial state $I_{\Lambda}$.
-In this case, the corresponding programming actions, $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$ making the classical planning task $P_{\Lambda}$ easier to be solved.
-When a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given, the $P_{\Lambda}$ compilation validates whether the observation of the plan execution follows the given model:
-\begin{itemize}
-\item $\mathcal{M}$ is proved to be a {\em valid} \strips\ action model for the given input data in $\tau$ iff a solution plan for $P_{\Lambda}$ can be found.
-\item $\mathcal{M}$ is proved to be a {\em invalid} \strips\ action model for the given input data $\tau$ iff $P_{\Lambda}$ is unsolvable.
-This means that $\mathcal{M}$ cannot be compliant with the given observation of the plan execution.
-\end{itemize}
-This validation capacity of our compilation is beyond the functionality of VAL (the plan validation tool~\cite{howey2004val}) because, our $P_{\Lambda}$ compilation can test {\em model validation} with a partial (or even an empty) action model and a partially observed plan trace.
-On the other hand, VAL requires (1) a full plan and (2), a full action model for plan validation.
+In some contexts, it is reasonable to assume that some parts of the action model are known and so there is no need to learn the entire model from scratch \cite{ZhuoNK13}.
+In \FAMA, when an action model $\xi$ is partially specified, the known preconditions and effects are encoded as fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$ set to true in the initial state $I_{\Lambda}$.
+In this case, the corresponding programming actions, $\mathsf{programPre_{p,\xi}}$ and $\mathsf{programEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, thereby making the classical planning task $P_{\Lambda}$ easier to be solved.
 
-The classical plan of Figure~\ref{fig:plan-lplan} shows a solution to the classical planning task $P_{\Lambda}$ that encodes a $\Lambda=\tup{\mathcal{M},\tau}$ learning task for getting the {\em blocksworld} action model where the headers for the schemes {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are specified in $\mathcal{M}$.
-This plan programs and validates the operator schema {\tt\small stack} from {\em blocksworld}, using the plan trace shown in Figure~\ref{fig:example-plans}.
-Plan steps $[0,1]$ program the preconditions of the {\tt\small stack} operator, steps $[2,6]$ program the operator effects and steps $[7,11]$ validate the programmed operators following the four-action plan shown in Figure~\ref{fig:example-plans}.
+%The compilation is flexible to the particular scenario where the number of {\em missing states} or {\em missing actions} in the input plan trace $\tau$ is bound.
+In both cases the output planning problem $P_{\Lambda}$ becomes simpler since the plan horizon is bound so the classical planner does not require to determine how many {\em apply} actions are necessary between any two state observations, i.e.
+between the application of two {\em validate} actions.
+
+So far we have explained the compilation for learning from a single input trace.
+However, the compilation is extensible to the more general case $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ is a set of plan traces.
+Taking this into account, a small modification is required in our compilation approach.
+In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_m^t\in \tau_t$, {\tt\small $1\leq t\leq k$} of a plan trace $\tau_t$ have now to reset the current state and the current plan.
+These actions are now redefined as:
+
+\begin{small}
+	\begin{align*}
+	\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_m^t\cup\{test_{j-1}\}\cup \{\neg mode_{prog}\},\\
+	\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_j\} \cup \\
+	&\{\neg f\}_{\forall f\in s_m^t, f \notin s_0^{t+1}}\cup \{f\}_{\forall f\in s_0^{t+1}, f \notin s_m^t},\\
+	&\{\neg f\}_{\forall f\in F_{\pi_t}}\cup \{f\}_{\forall f\in F_{\pi_{t+1}}}.\\
+	\end{align*}
+\end{small}
+
+Finally, we will detail the composition of a solution plan $\pi_\Lambda$ to a planning task $P_\Lambda$ and the mechanism to extract the output set of action models $\mathcal{M}'$ from $\pi_\Lambda$.
+The plan of Figure~\ref{fig:plan-lplan} shows a solution to the task $P_{\Lambda}$ that encodes a learning task $\Lambda=\tup{\mathcal{M},\tau}$ for obtaining the action models of the {\em blocksworld} domain, where the models for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already specified in $\mathcal{M}$.
+Therefore, this plan programs and validates the action model for {\tt\small stack}, using the plan trace of Figure~\ref{fig:example-plans}.
+Plan steps $00-01$ program the preconditions of the {\tt\small stack} model, steps $02-06$ program the action model effects and steps $07-11$ validate the programmed model following the four-action plan shown in Figure~\ref{fig:example-plans}.
 
@@
 	{\footnotesize\tt
-  {\bf 00} : (program\_pre\_holding\_stack\_v1) \\
-     01 : (program\_pre\_clear\_stack\_v2)\\
-  {\bf 02} : (program\_eff\_clear\_stack\_v1)\\
-    03 : (program\_eff\_clear\_stack\_v2)\\
-    04 : (program\_eff\_handempty\_stack)\\
-    05 : (program\_eff\_holding\_stack\_v1)\\
-    06 : (program\_eff\_on\_stack\_v1\_v2)\\
+		{\bf 00} : (program\_pre\_stack\_holding\_v1) \\
+		01 : (program\_pre\_stack\_clear\_v2)\\
+		{\bf 02} : (program\_eff\_stack\_clear\_v1)\\
+		03 : (program\_eff\_stack\_clear\_v2)\\
+		04 : (program\_eff\_stack\_handempty)\\
+		05 : (program\_eff\_stack\_holding\_v1)\\
+		06 : (program\_eff\_stack\_on\_v1\_v2)\\
 		{\bf 07} : (apply\_unstack blockB blockA i1 i2)\\
@@
 	}
- \caption{\small Plan for programming and validating the $stack$ schema (using plan $\pi$ and state observations shown in Figure~\ref{fig:example-plans}) as well as previously specified operator schemes for $pickup$, $putdown$ and $unstack$.}
+	\caption{\small Plan for programming and validating the $stack$ action model (using the plan trace $\tau$ of Figure~\ref{fig:example-plans}) as well as previously specified action models for $pickup$, $putdown$ and $unstack$.}
 	\label{fig:plan-lplan}
@@
 
-The compilation is flexible to the particular scenario where the number of {\em missing states} or {\em missing actions} in the input plan trace $\tau$ is bound.
-In both cases the output planning problem $P_{\Lambda}$ becomes simpler since the plan horizon is bound so the classical planner does not require to determine how many {\em apply} actions are necessary between any two state observations, i.e.
-between the application of two {\em validate} actions.
+Given a solution plan $\pi_\Lambda$ that solves $P_{\Lambda}$, the set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ are computed in linear time and space.
+In order to do so, $\pi_\Lambda$ is executed in the initial state $I_{\Lambda}$ and the action model $\mathcal{M}'$ will be given by the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that are set to true in the last state reached by $\pi_\Lambda$, $s_g=\theta(I_\Lambda,\pi_\Lambda)$.
+For each $\xi \in \mathcal{M'}$, we build the sets of preconditions, positive effects and negative effects as follows:
 
-Last but not least we explain how to address learning \strips\ action models from $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ observations of the execution of plans $\Pi=\{\pi_1,\ldots,\pi_k\}$ in a {\em classical planning frame} $\Phi=\tup{F,A}$.
-Let us first define a set of classical planning instances $P_t=\tup{F,A,I_t,G_t}$, {\tt\small $1\leq t\leq k$}, that belong to $\Phi$ (i.e.
-same fluents and actions but different initial states and goals).
-The initial state $I_t$ is given by the state $s_0^t\in \tau_t$ and the observed actions from the plan $\pi_t$ while the goals $G_t$ are defined by the number of observed actions and states from the corresponding plan trace $\tau_t$.
-Addressing the learning task $\Lambda=\tup{\mathcal{M},\mathcal{T}}$ where $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ requires introducing a small modification to our compilation.
-In particular, the actions in $P_{\Lambda}$ for {\em validating} last state $s_m^t\in \tau_t$ of a plan trace $\tau_t$ reset the current state and the current plan.
-These actions are now redefined as:
+%As shown in Figure \ref{fig:plan-lplan}, the learned action models $\mathcal{M'}$ can be easily inferred from the solution plan $\pi_\Lambda$, just by looking at the programming actions.
+However, for the shake of formalizing the extraction of the learned models, we will infer them from the final state of the solution plan $s_g=\theta(I_\Lambda,\pi_\Lambda)$ by looking at the values for the fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$ which encode the preconditions and effects of the validated action models.
+For each $\xi \in \mathcal{M'}$, we build the sets of preconditions, positive effects and negative effects as follows:
+
 \begin{small}
 	\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{t}})=&G_t\cup\{test_{j-1}\}\cup \{\neg mode_{prog}\},\\
-\cond(\mathsf{validate_{t}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_t\} \cup \\
-&\{\neg f\}_{\forall f\in G_t, f \notin I_{t+1}}\cup \{f\}_{\forall f\in I_{t+1}, f \notin G_t},\\
-&\{\neg f\}_{\forall f\in F_{\pi_t}}\cup \{f\}_{\forall f\in F_{\pi_{t+1}}}.\\
+	\hspace*{7pt}pre(\xi)=& \{p ~|~ pre_p(\xi) \in s_g\}_{\forall p \in \Psi_\xi},\\
+	\hspace*{7pt}add(\xi)=& \{p ~|~ add_p(\xi) \in s_g\}_{\forall p \in \Psi_\xi},\\
+	\hspace*{7pt}del(\xi)=& \{p ~|~ del_p(\xi) \in s_g\}_{\forall p \in \Psi_\xi}.
 	\end{align*}
@@
 
-\subsection{Compilation properties}
-Given a classical plan $\pi$ that solves $P_{\Lambda}$, then an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ can be computed in linear time and space.
-In this case $\pi$ has to be executed starting from the initial state $I_{\Lambda}$ and the action model $\mathcal{M}'$ will be given by the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that are set to true at the last state reached by $\pi$.
+\subsection{Properties of the compilation}
+
+Given a plan $\pi_\Lambda$ that solves $P_\Lambda$, and the learned action models $\mathcal{M'}$ that solve $\Lambda=\tup{\mathcal{M}, \tau}$, \FAMA contributes with the following compilation properties:
 
 \begin{mylemma}
-Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ induces an action model $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$.
+Soundness. Any classical plan $\pi_\Lambda$ that solves $P_{\Lambda}$ induces a set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$.
 \end{mylemma}
@@
 \begin{small}
-  Once operator schemas $\mathcal{M}'$ are programmed, they can only be applied and validated, because of the $mode_{prog}$ fluent.
-In addition, $P_{\Lambda}$ is only solvable if fluents {\tt\small $at_n$} and {\tt\small $test_m$} hold at the last state reached by $\pi$.
-By the definition of the $\mathsf{apply_{\xi,\omega}}$ and the $\mathsf{validate_{j}}$ actions, these goals can only be achieved executing an applicable sequence of programmed operator schemas that reaches every state $s_i\in\tau$, starting from the corresponding initial state and following the sequence of $n$ actions that is defined by $\tau$.
+  Once action models $\mathcal{M}'$ are programmed, they can only be applied and validated because of the $mode_{prog}$ fluent.
+In addition, $P_{\Lambda}$ is only solvable if fluents {\tt\small $at_n$} and {\tt\small $test_m$} hold at the last state reached by $\pi_\Lambda$.
+By the definition of the $\mathsf{apply_{\xi,\omega}}$ and the $\mathsf{validate_{j}}$ actions, these goals can only be achieved executing an applicable sequence of programmed action models that reaches every state $s_j\in\tau$, starting in the corresponding initial state and following the sequence of $n$ observed actions of $\tau$.
 This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, that $\mathcal{M}'$ is a solution to $\Lambda$.
@@
 \begin{mylemma}
-Completeness. Any \strips\ action model $\mathcal{M}'$ that solves a $\Lambda=\tup{\mathcal{M},\tau}$ learning task, is computable solving the corresponding classical planning task $P_{\Lambda}$.
+Completeness. Any set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ is computable solving the corresponding classical planning task $P_{\Lambda}$.
 \end{mylemma}
@@
 \begin{small}
-  By definition, $F_{\xi}\subseteq F_\Lambda$ fully captures the full set of elements that can appear in a \strips\ action schema $\xi\in\mathcal{M}$.
-The compilation does not discard any possible \strips\ action schema $\mathcal{M}'$ definable within $F_v$ that satisfies the state trajectory constraint and plan constrains given by $\tau$.
-This means that a solution plan $\pi$ can be built selecting the corresponding  $\mathsf{programPre_{f,\xi}}$ and $\mathsf{programEff_{f,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\tau$.
+  By definition, $\Psi_{\xi}\subseteq \Psi_v$ fully captures the set of elements that can appear in an action model $\xi\in\mathcal{M}$.
+The compilation does not discard any possible set of action models $\mathcal{M}'$ definable within $\Psi_v$ that satisfies the observed state trajectory and action sequence of $\tau$.
+This means that for every $\mathcal{M}'$ that solves $\Lambda$, there exists a plan $\pi_\Lambda$ that can be built selecting the appropriate programming, apply and validate actions from the $P_{\Lambda}$ compilation.
+  %a solution plan $\pi_\Lambda$ can be built selecting the corresponding  $\mathsf{programPre_{p,\xi}}$ and $\mathsf{programEff_{p,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\tau$.
 \end{small}
@@
 
-The size of the classical planning task $P_{\Lambda}$ output by the compilation depends on:
+The size of the planning task $P_{\Lambda}$ output by the compilation approach depends on:
+
 \begin{itemize}
-\item The arity of the actions and the fluents in $\tau$ that are given as input to the $\Lambda$ learning task.
-The larger these numbers, the larger the size of the $F_{\xi}$ sets.
-This is the term that dominates the compilation size because it defines the $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ fluents and the corresponding set of {\em programming} actions.
-\item The length of the action sequence and the state trajectory in the observed plan execution $\tau$.
+\item The arity of the actions and the fluents in $\tau$ given as input in $\Lambda$.
+The larger the arity, the larger the size of the $\Psi_{\xi}$ sets.
+This is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents and the corresponding set of {\em programming} actions.
+\item The length of the observed action sequence and the state trajectory of $\tau$.
 The larger the number of observed actions, $a_i\in\tau$ s.t.
@@
 
-\subsection{Optimizing the compilation with background knowledge}
-A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn logic programs from data~\cite{muggleton1994inductive}.
+An interesting aspect of our approach is that when a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given in $\Lambda$, the $P_{\Lambda}$ compilation also serves to validate whether the observed $\tau$ follows the given model $\mathcal{M}$:
+
+\begin{itemize}
+	\item $\mathcal{M}$ is proved to be a {\em valid} action model for the given input data in $\tau$ iff a solution plan for $P_{\Lambda}$ can be found.
+	\item $\mathcal{M}$ is proved to be a {\em invalid} action model for the given input data $\tau$ iff $P_{\Lambda}$ is unsolvable.
+This means that $\mathcal{M}$ cannot be compliant with the given observation of the plan execution.
+\end{itemize}
+
+The validation capacity of our compilation is beyond the functionality of VAL (the plan validation tool~\cite{howey2004val}) because our $P_{\Lambda}$ compilation is able to address {\em model validation} of a partial (or even an empty) action model with a partially observed plan trace.
+On the other hand, VAL requires (1) a full plan and (2), a full action model for plan validation.
+
+%\subsection{Optimizing the compilation with background knowledge}
+%A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn logic programs from data~\cite{muggleton1994inductive}.
 Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case to optimize the performance of the $P_{\Lambda}$ compilation.
@@
 %\subsubsection{Static predicates}
-A {\em static predicate} $p \in \Psi$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
+%A {\em static predicate} $p \in \Psi_\Lambda$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
 Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
 Given a static predicate $p$:
-\begin{itemize}
-\item Fluents $del_f(\xi)$ and $add_f(\xi)$, such that $f\in F_v$ is an instantiation of the static predicate $p$ in the set of {\em variable objects} $\Omega_v$, can be discarded for every $\xi\in\Xi$.
-\item Actions $\mathsf{programEff_{f,\xi}}$ (s.t. $f\in F_v$ is an instantiation of $p$ in $\Omega_v$) can also be discarded for every $\xi\in\Xi$.
-\end{itemize}
+%\begin{itemize}
+%\item Fluents $del_p(\xi)$ and $add_p(\xi)$, such that $p\in \Psi_v$ is the static predicate, can be discarded for every $\xi\in\mathcal{M}$.
+%\item Actions $\mathsf{programEff_{p,\xi}}$ (s.t. $p\in \Psi_v$ is a parameterized static predicate) can also be discarded for every $\xi\in\mathcal(M)$.
+%\end{itemize}
 
-Static predicates can also constrain the space of possible preconditions by looking at the given set of state observations in $\tau$.
-One can assume that if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not compliant with the observations in $\tau$ then, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
+%Static predicates can also constrain the space of possible preconditions by looking at the given set of state observations in $\tau$.
+One can assume that if a precondition $p\in \Psi_v$ (s.t.
+$p\in \Psi_v$ is a static predicate parameterized with $\Omega_v$) is not compliant with the observations in $\tau$ then, fluents $pre_p(\xi)$ and actions $\mathsf{programPre_{p,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
 For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\tau$.
 
-Furthermore looking as well at the given plan traces, fluents $pre_f(\xi)$ and actions $\mathsf{programPre_{f,\xi}}$ are also discardable for every $\xi\in\Xi$ if a precondition $f\in F_v$ (s.t.
-$f\in F_v$ is an instantiation of a static predicate in $\Omega_v$) is not possible according to $\tau$.
+%Furthermore looking as well at the given plan traces, fluents $pre_p(\xi)$ and actions $\mathsf{programPre_{p,\xi}}$ are also discardable for every $\xi\in\mathcal(M)$ if a precondition $p\in \Psi_v$ (s.t.
+$p\in \Psi_v$ is a static predicate parameterized with $\Omega_v$) is not possible according to $\tau$.
 Back to the {\em zenotravel} domain, if a plan trace $\tau$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.

### 2018-07-08 doc/aij18/motivation.tex
Add files via upload
Sections 2, 3, 4 and 5 finished

--- 2018-06-19 doc/aij18/motivation.tex
@@
 
+In this section, we will highlight the principal distinctive features of our approach \FAMA with respect to the related work reviewed in Section~\ref{related_work}.
+
+\vspace{0.1cm}
+
+%\subsection{Learning action models from plan traces of unbound horizon}
+When learning action models from observations of plan executions there are two main sources of partial observability:
+\begin{enumerate}
+\item As many of the approaches in Section~\ref{sec:background} assume, there may be an unknown number of missing intermediate states in the trace because of the partial state observability (\PO and \NO).
+The assumption of having \FO state trajectories means that the sensors are able to capture every state change at every instant, which typically is unrealistic.
+Normally, the process for obtaining state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two sensor readings.
+
+\item There may be also an unbound number of missing actions in the plan trace because of partial observability.
+The common assumption of having \FO action sequences in a learning task is unrealistic in many domains as it implies the existence of human observers that annotate the observed action sequences.
+In some real-world applications, the observed and collected data are sensory data (e.g., home automation, robotics) or images (e.g.
+traffic) and one cannot rely on human intervention for labeling actions.
+Actually, learning the executed actions can also be part of the action-model learning task.
+Learning, for instance, from unstructured data involves transforming the sensor or image information into a predicate-like format before applying the action-model learning approach, and it also requires the ability of identifying action symbols~\cite{AsaiF18}.
+\end{enumerate}
+
+\FAMA represents a step ahead towards learning action models without assuming observed actions.
 The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging.
+As Table \ref{tab:complex} shows, PSPACE-complexity happens when the length of the plan trace is a priori unbound, which results from the combination of \PO and \NO state trajectories and action sequences.
 This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now.
 
-In this particular scenario, the actual number of plan traces that can correspond with the given input observations is also unbound and grows exponentially with the actual length of the plan trace (that is now unknown).
+In this particular scenario, the actual number of plan traces associated to a given input observation is also unbound and grows exponentially with the actual length of the plan trace (that is now unknown).
 Otherwise, the learning task is SAT compilable, which is known to be a NP-complete task~\cite{russell2016artificial}.
 This is the reason that SAT solving is a common technique in the approaches presented in section \ref{sec:background}.
-
-When we assume partial observability in both the sequence of actions and the state trajectory, a complete approach must consider the length of the input plan traces to be unknown.
-This work shows that classical planning is a complete approach for this particular scenario.
-Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning techniques, as our proposal of compiling the learning task to a planning problem.
-
-When the plan trace is fully observed, learning \strips\ action models is straightforward~\cite{jimenez2012review}.
-In this case the {\em pre-} and {\em post-states} of every action are available so action {\em effects} are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
-Likewise {\em preconditions} are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action.
-
-%In \FAMA we set a probability threshold of observability to each fluent and action, which determines the percentage of literals in each state and in turn the percentage of observed states (when the probability of every fluent is above the threshold, the result is a missing state).
-This way, \FAMA must always work under the assumption that the number and length of the input plan traces are unknown and so the task of learning an action model becomes as hard as solving a \strips planning problem.
-Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning techniques, as our proposal of compiling the learning task to a planning problem.
-
-%Thereby, solving this type of learning tasks justifies the use of techniques other than SAT solvers, as our proposal of compiling the learning task to a planning problem.
 
@@
 
-Table \ref{tab:complex} displays the complexity of the learning task according to the type of input trace.
-The PSPACE-complexity happens when the length of the plan trace is a priori unbound, which results from the combination of the following two causes:
+When we assume partial observability in both actions and states, a complete approach must consider the length of the input plan traces to be unknown.
+\FAMA shows that classical planning is a complete approach for this particular scenario.
+Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning techniques, as our proposal of compiling the learning task to a planning problem.
 
-\begin{enumerate}
-\item As many of the approaches in section \ref{sec:background} assume, there may be an unknown number of missing intermediate states in the trace because of partial state observability (\PO and \NO).
-The assumption of having \FO state trajectory means that the sensors are able to capture every state change at every instant which typically is unrealistic.
-Normally the process of getting state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two sensor readings.
+When the plan trace is fully observed, learning \strips\ action models is straightforward~\cite{jimenez2012review}.
+In this case the {\em pre-} and {\em post-states} of every action are available and so action {\em effects} are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
+Likewise {\em preconditions} are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action.
 
-\item There may be also an unbound number of missing actions in the plan trace because of partial observability.
-The common assumption of having \FO action sequences in a learning task is unrealistic in many domains because it implies the existence of human observers that annotate the observed action sequences.
-In some real-world applications, the observed and collected data are sensory data (e.g., home automation, robotics) or images (e.g.
-traffic) and one cannot rely on human intervention for labeling actions.
-Actually, learning the executed actions can also be part of the action-model learning task.
-Learning from unstructured data involves some prior processing to transform the sensor or image information into a predicate-like format before applying the action-model learning approach~\cite{AsaiF18}, and it also requires the ability of identifying action symbols.
-In this sense, \FAMA represents a step ahead towards learning action models without assuming observed actions.
-\end{enumerate}
+%In \FAMA we set a probability threshold of observability to each fluent and action, which determines the percentage of literals in each state and in turn the percentage of observed states (when the probability of every fluent is above the threshold, the result is a missing state).
+This way, \FAMA must always work under the assumption that the number and length of the input plan traces are unknown and so the task of learning an action model becomes as hard as solving a \strips planning problem.
+Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning techniques, as our proposal of compiling the learning task to a planning problem.
 
-Regarding the metric to evaluate the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
-This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM model.
-Hence, because the set of preconditions and effects of the GTM model is usually smaller than the set of all possible preconditions and effects, it turns out the metric may output error rates below 100\% for totally wrong learned models.
+%Thereby, solving this type of learning tasks justifies the use of techniques other than SAT solvers, as our proposal of compiling the learning task to a planning problem.
 
-In order to overcome this limitation, we propose to use two standard metrics from ML, {\em precision} and {\em recall}.
-These two metrics are frequently used in pattern recognition, information retrieval and binary classification and are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model~\cite{davis2006relationship}.
+\vspace{0.1cm}
+Regarding the evaluation of the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
+This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM.
+Hence, since the set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects, it turns out that the syntactic metric may output an error rate below 100\% for a totally wrong learned model.
+To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}.
+These two syntactic metrics are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model.
 
-A striking figure of Table \ref{table:models_comparison2} that emphasizes a relevant feature of \FAMA is the small size of the training dataset it requires in comparison to other approaches.
+Pure syntax-based evaluation metrics, like the ones mentioned in the above paragraph, can report low scores for learned models that are actually {\em sound} and {\em complete} but syntactically different from the GTM.
+In this sense, \FAMA contributes with a novel semantic-based error measure that builds upon the precision and recall metrics.
+Semantic evaluation metrics add a distinctive value over the syntactic ones, which is that they evaluate the learned model with a set of observations of plan executions and hence they are very appropriate for scenarios where the GTM is not available.
+Unlike the semantic metric used in \ARMS~\cite{yang2007learning}, our proposal is not sensitive to the repetition of one same flaw in the model evaluation.
+
+%When learning action models from observations of plan executions, the roles of two {\em comparable} action schemes (or the roles of two action parameters with the same type) can be swapped.
+These role swaps typically happen when the observed input data is scarce.
+For instance the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa.
+Further, the roles of parameters of the {\small\tt stack} (or the {\small\tt unstack}) operator could be swapped and these learned models would still be semantically correct with respect to the given input observations.
+
+%Pure syntax-based evaluation metrics can report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the GTM model; i.e.
+a learned model semantically equivalent but syntactically different to the reference model.
+The ARMS system implemented semantic metrics for evaluating the learned action models with respect to observations of plan executions that act as a {\em test set}.
+This semantic evaluation approach is suitable for scenarios where the GTM is not available, e.g.
+the traditional ML setting.
+For this scenario our work proposes semantic versions of {\em precision} and {\em recall} that provide a notion of the soundness and completeness of the learned models with respect to input plan traces.
+Unlike the metrics defined by ARMS, our semantic metrics do not accumulate errors that are caused by the same flaw in the learned model.
+
+%\subsection{Input data}
+
+On the other hand, a striking figure of Table \ref{table:models_comparison2} that emphasizes a relevant feature of \FAMA is the small size of the training dataset it requires in comparison to other approaches.
 Unlike extensive-data ML approaches, our work explores an alternative research direction to learn sound models from small amounts of plan traces.
@@
 
-Finally, we would also like to point out that, as shown in section \ref{sec:experiments}, \FAMA is exhaustively evaluated over a wide range of domains (15 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner so it can benefit straightforward from the last advances in classical planning.
+Finally, as it will be shown in section \ref{sec:experiments}, \FAMA is exhaustively evaluated, syntactically and semantically, over a wide range of domains (13 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner so it can benefit straightforward from the last advances in classical planning.

### 2018-07-09 doc/aij18/background.tex
small updates to background section aij18 paper

@@
 
-This section serves two purposes: first we introduce the basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper; and secondly, we summarize the existing approaches to learn action models in classical planning and highlight our contributions via the related work.
+This section serves two purposes: first we introduce basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper; and secondly, we summarize the existing approaches to learn classical planning action models and highlight our contributions comparing with related work.
 
@@
 
+A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states, i.e.
+$|s|=|F|$, so the size of the state space is $2^{|F|}$.
 Like in PDDL~\cite{fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
@@
 Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-
-A {\em state} $s$ is a full assignment of values to fluents, i.e.
-$|s|=|F|$, so the size of the state space is $2^{|F|}$.
-We explicitly include negative literals $\neg f$ in states since our approach handles with partial observability, thus adopting the \emph{open world assumption}.
 
@@
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.,~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
-A solution plan is {\em optimal} if it has minimum length.
+A solution plan for $P$ is {\em optimal} if it has minimum length.
 
@@
 
+Our approach copes with the partial observability of the plan execution adopting the \emph{open world assumption}.
 With regard to the observed states in a plan trace, we say that $\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$.
 Otherwise, it is a partially-observable (\PO) state trajectory, meaning that at least one state $s_i$ is a partial assignment of values to fluents in which one or more literals are missing (formally, $|s_i|<|F|$).
-This implies that in a \PO state trajectory all fluents of a state $s_i$ may be missing, in which case, $s_i$ is a \emph{missing} or \emph{empty} state.
+This implies that in a \PO state trajectory all fluents of a state $s_i$ might be missing, in which case, $s_i$ is a \emph{missing} or \emph{empty} state.
 This general definition of \PO gives rise to two particular cases:
@@
 \begin{itemize}
-\item when \textbf{all} the $n-1$ intermediate states of a trajectory $s$ are \textbf{missing}, $s$ is a \emph{non-observable} (\NO) state trajectory
+\item when \textbf{all} the $n-1$ intermediate states of a trajectory $s$ are \textbf{missing}, $s$ is a \emph{non-observable} (\NO) state trajectory.
 \item when \textbf{none} of the $n-1$ intermediate states of a trajectory $s$ is \textbf{missing}, we will refer to $s$ as a \POstar state trajectory.
@@
 \begin{tabular}{c|c|c|}
-	     & \# intermediate states & state type \\ \hline
+	     & {\bf \# intermediate states} & {\bf state type} \\ \hline
     \FO & $n-1$  & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ is a full assignment \\ \hline
@@
     \multirow{1}{*}{\PO} & $\leq n-1$ & {\small $\exists i, 1 \leq i < n$}   \\  & & $s_i$ is a partial assignment \\ \hline
-    \NO & 0 & {\small $\forall i, 1 \leq i < n$}  \\  & & $|s_i|=0$  \\\hline
+    \NO & 0 & {\small $\forall i, 1 \leq i < n$}  \\  & & $|s_i|=0$  
 \end{tabular}
-\caption{Classification of state trajectories accordingly to observation}
+\caption{Classification of state trajectories accordingly to the observed information.}
 \label{tab:state_trajectory}
@@
 With regard to the observed actions in a plan trace, we say that an action sequence, $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to transit every state $s_{i-1}$ to the corresponding successor state $s_{i}$, for each {\small $1\leq i\le n$}.
-We say it is a \PO action sequence if at least one of these necessary actions is missing in the sequence and a \NO action sequence in case the sequence of observed actions is empty.
+We say that the action sequence is \PO, if at least one of these necessary actions is missing in the sequence and that is \NO, when the sequence of observed actions is empty.
 
-Given a  plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$, it holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}.
+A plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$, holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}.
 Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
@@
 
-The first column of Table \ref{table:models_comparison1} shows the constraints imposed on the input plan traces with regards to observability.
+The first column of Table \ref{table:models_comparison1} shows the constraints imposed on the input plan traces with regard to observability.
 Since all approaches except ours deal only with \FO action sequences, constraints are exclusively concerned with the type of state trajectory.
 This directly affects the complexity of the task, which can be sorted from the least to the most constrained following this order: 1) \NO, 2) \PO, 3) \POstar, and 4) \FO.
-Note that \PO is less constrained than \POstar because it also includes the possibility of having some missing state in the trajectory.
+Note that \PO is less constrained than \POstar because \PO considers the possibility of having some missing state in the trajectory.
 
 The task of learning from less constrained traces subsumes learning from more constrained ones.
-Consequently, approaches to learning from, for instance, traces with \PO state trajectories will also be able to learn from traces with \POstar state trajectories.
+Consequently, approaches to learning from, for instance traces with \PO state trajectories, will also be able to learn from traces with \POstar state trajectories.
 All the approaches analyzed in this work accept the more constrained definition of partial observations of intermediate states \POstar, and most of them also allow the sequence of intermediate states to be empty.
@@
 Most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces.
-Others do not explicitly say so but the fact is that the predicates and the actions headers are easily extractable from the state sequence and action sequence of the plan traces, respectively.
-A requirement, however, for these two sets to be representative is that set of input plan traces comprises at least a grounded sample of all predicates and operator schemas of the domain model.
+Others do not explicitly say so but the fact is that predicates and actions headers are easily extractable from the state sequence and action sequence of the input plan traces, respectively.
+A requirement, however, is that the input plan traces comprise at least a grounded sample of every predicate and operator schema in the domain model.
 
@@
 
-Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- almost all approaches rely on a comparison between the learned model and a ground-truth model), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
+Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- almost all approaches rely on a comparison between the learned model and a {\em Ground-Truth Model}), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
 
@@
          \hline
-		\FAMA & checking wrt GTM  & precision and recall of & 15\\ & & \#$\pre$ and \#$\eff$ & 25 actions\\
+		\FAMA & checking wrt GTM  & precision and recall & 15\\ & &  & 25 actions\\
          \hline

### 2018-07-09 doc/aij18/learning_task.tex
small updates to learning section aij18 paper

@@
 This section details the \FAMA approach for learning action models.
-The notion of learning task is defined in section \ref{learning_task} and the components of the compilation scheme are described in sections \ref{propositional_encoding} and \ref{conditional_effects}.
+The notion of the learning task is defined in section \ref{learning_task} and the components of the compilation scheme are described in sections \ref{propositional_encoding} and \ref{conditional_effects}.
 Subsequently, the compilation approach is fully detailed in section \ref{compilation} and the last section presents some theoretical properties of the compilation scheme.
@@
 
-Our {\em learning task} consists in learning classical planning action models by observing one or more agents acting in a world, which is defined by a {\em classical planning frame} $\Phi=\tup{F,A}$.
+Our {\em learning task} consists in learning classical planning action models by observing one or more agents acting in a world definable by a {\em classical planning frame} $\Phi=\tup{F,A}$.
 The learning task is formalized by the pair $\Lambda=\tup{\mathcal{M},\tau}$:
@@
 
-Ultimately, we can always assume that $\mathcal{M}$ will contain the predicates $\Psi$ and the headers of the actions models, either explicitly provided in $\mathcal{M}$ or inferred from $\tau$.
+Ultimately, we can always assume that $\mathcal{M}$ will contain predicates in $\Psi$ as well as the headers of the actions models, either explicitly provided in $\mathcal{M}$ or inferred from $\tau$.
 A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a set of action models $\mathcal{M}'$ that is compliant with the input models $\mathcal{M}$ and the observed plan trace $\tau$.
@@
 	;;; State observation
-	(clear A) (on A B) (ontable B) (handempty)
+	(clear A) (on A B) (ontable B) 
 	\end{verbatim}
@@
 
-\subsection{Propositional encoding of an action model}
+\subsection{A propositional encoding for \strips\ action models}
 \label{propositional_encoding}
@@
 
-In this section we formalize the propositional encoding of an \strips\ action model.
+In this section we formalize a propositional encoding of an \strips\ action model.
 This encoding is at the core of the \FAMA compilation approach for addressing the learning task defined in section \ref{learning_task}.
@@
 We define $\Psi_v$ as the set of predicates $\Psi$ parameterized with the {\em variable names} of $\Omega_v$ as arguments.
-$\Psi_v$  defines the elements that can appear in the preconditions and effects of the action models.
+The set $\Psi_v$  defines the elements that can appear in the preconditions and effects of the action models.
 In the {\em blocksworld} domain, this set contains eleven elements, $\Psi_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-In more detail, for a given action model $\xi$, we define $\Psi_{\xi}\subseteq \Psi_v$ as the subset of elements of $\Psi_v$ that appear in $\xi$.
+In more detail, for a given action model $\xi$, we define $\Psi_{\xi}\subseteq \Psi_v$ as the subset of elements of $\Psi_v$ that can appear in $\xi$.
 For instance, $\Psi_{\tt stack}=\Psi_v$ whereas $\Psi_{\tt pickup}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$),on($v_1,v_1$)\}} excludes the elements from $\Psi_v$ that involve $v_2$ because {\small\tt pickup} actions have arity one.
 The size of the space of possible \strips\ models for a given $\xi$ is $2^{2|\Psi_{\xi}|}$ (recall that negative effects appear as preconditions and that they cannot be positive effects, and also that a positive effect cannot appear as a precondition).
-For the {\em blocksworld}, $2^{2|\Psi_{\tt stack}|}=4194304$ while for the {\tt pickup} operator this number is only 1024.
+For the {\em blocksworld}, $2^{2|\Psi_{\tt stack}|}=4,194,304$ while for the {\tt pickup} operator this number is only 1024.
 
@@
 
+Our approach to learning action models is to compile this learning task into a classical planning task with conditional effects.
 Conditional effects allow us to compactly define actions whose effects depend on the current state.
@@
 \subsection{Compilation}
-\label{sec:compilation}
+\label{compilation}
 
-In section \ref{sec:motivation}, we exposed our reasons to solve the learning task via compilation to planning.
-Our compilation scheme  builds upon the approach presented in \cite{aineto2018learning}.
-\FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.
+In section \ref{sec:motivation}, we exposed our reasons to solve the learning task $\Lambda=\tup{\mathcal{M},\tau}$ via compiling it into a classical planning task $P_{\Lambda}$.
+Our compilation scheme  builds upon the approach presented in \cite{aineto2018learning} but \FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.
 
-Our approach to solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to compile $\Lambda$ into a classical planning task $P_{\Lambda}$ with conditional effects.
-The intuition behind the compilation is that a solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output set of action models $\mathcal{M}'$ that solves $\Lambda$.
+The intuition behind the \FAMA compilation is that a solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output set of action models $\mathcal{M}'$ that solves $\Lambda$.
 Specifically, a solution plan $\pi_\Lambda$ serves two purposes:
@@
 \item {\bf To validate the set of action models $\mathcal{M}'$}.
-$\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the recognition of the available plan trace $\tau$ with the programmed action models $\mathcal{M}'$.
+$\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the validation of the available plan trace $\tau$ with the programmed action models $\mathcal{M}'$.
 \end{enumerate}

### 2018-07-09 doc/aij18/learning_task.tex
small updates to learning section aij18 paper

@@
 \item {\bf To validate the set of action models $\mathcal{M}'$}.
-$\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the validation of the available plan trace $\tau$ with the programmed action models $\mathcal{M}'$.
+$\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the validation of the observed plan trace $\tau$ with the programmed action models $\mathcal{M}'$.
 \end{enumerate}
@@
 Here we formalize the compilation for learning \strips\ action models with classical planning.
-Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$, with $\tau$ formed by an $n$-action sequence $\tup{a_1, \ldots, a_n}$ and a $m$-state trajectory $\tup{s_0, s_1, \ldots, s_m}$ (i.e., $\tau = \langle s_0, a_1, \ldots, a_n, s_m \rangle$), the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
+Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$, with $\tau$ formed by an $n$-action sequence $\tup{a_1, \ldots, a_n}$ and a $m$-state trajectory $\tup{s_0, s_1, \ldots, s_m}$ ($\tau = \langle s_0, a_1, \ldots, a_n, s_m \rangle$), the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
 
@@
 \begin{itemize}
-\item The set of fluents of $s_0$; i.e., $F$.
-\item The fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, as explained in section \ref{propositional_encoding}
+\item The set of fluents obtained from $s_0$; i.e., $F$.
+\item The fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section \ref{propositional_encoding}
 \item A set of fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to represent the $i^{th}$ observable action of $\tau$.
@@
 validated. In the example of Figure~\ref{fig:example-plans} two tests are required to validate the programmed action model, one test at $s_0$ and another one at $s_4$.
-\item A fluent, $mode_{prog}$, to indicate whether the action models are being programmed or validated.
+\item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated.
 \end{itemize}
@@
 \hspace*{7pt}\pre(\mathsf{programEff_{p,\xi}})=&\{\neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
-\cond(\mathsf{programEff_{p,\xi}})=&\{pre_{p}(\xi)\}\rhd\{del_{p}(\xi)\},\{\neg pre_{p}(\xi)\}\rhd\{add_{p}(\xi)\}.
+\cond(\mathsf{programEff_{p,\xi}})=&\{pre_{p}(\xi)\}\rhd\{del_{p}(\xi)\},\\
+& \{\neg pre_{p}(\xi)\}\rhd\{add_{p}(\xi)\}.
 \end{align*}
@@
 Taking this into account, a small modification is required in our compilation approach.
-In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_m^t\in \tau_t$, {\tt\small $1\leq t\leq k$} of a plan trace $\tau_t$ have now to reset the current state and the current plan.
+In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_m^t\in \tau_t$, {\tt\small $1\leq t\leq k$} of a plan trace $\tau_t$ reset the current state and the current plan.
 These actions are now redefined as:
@@
 The plan of Figure~\ref{fig:plan-lplan} shows a solution to the task $P_{\Lambda}$ that encodes a learning task $\Lambda=\tup{\mathcal{M},\tau}$ for obtaining the action models of the {\em blocksworld} domain, where the models for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already specified in $\mathcal{M}$.
-Therefore, this plan programs and validates the action model for {\tt\small stack}, using the plan trace of Figure~\ref{fig:example-plans}.
-Plan steps $00-01$ program the preconditions of the {\tt\small stack} model, steps $02-06$ program the action model effects and steps $07-11$ validate the programmed model following the four-action plan shown in Figure~\ref{fig:example-plans}.
+Therefore, this plan programs and validates the action model for {\tt\small stack}, using the input plan trace of Figure~\ref{fig:example-plans}.
+Plan steps $00-01$ program the preconditions of the {\tt\small stack} model, steps $02-06$ program the action model effects and steps $07-11$ is the plan postfix that validates the programmed model following the trace of Figure~\ref{fig:example-plans}.
 
@@
 
-Given a plan $\pi_\Lambda$ that solves $P_\Lambda$, and the learned action models $\mathcal{M'}$ that solve $\Lambda=\tup{\mathcal{M}, \tau}$, \FAMA contributes with the following compilation properties:
+%Given a plan $\pi_\Lambda$ that solves $P_\Lambda$, and the learned action models $\mathcal{M'}$ that solve $\Lambda=\tup{\mathcal{M}, \tau}$, \FAMA contributes with the following compilation properties:
 
@@
 This is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents and the corresponding set of {\em programming} actions.
-\item The length of the observed action sequence and the state trajectory of $\tau$.
+\item The length of the observed action sequence and state trajectory of $\tau$.
 The larger the number of observed actions, $a_i\in\tau$ s.t.

### 2018-07-09 doc/aij18/background.tex
small updates to background section aij18 paper

@@
 	\begin{tabular}{ l | c | c | c }
-		& \multicolumn{1}{c}{Input plan traces}
-        & \multicolumn{1}{c}{Learned action model}
-        & \multicolumn{1}{c}{Technique}     \\
+		& \multicolumn{1}{c|}{\bf Input plan traces}
+        & \multicolumn{1}{c|}{\bf Learned action model}
+        & \multicolumn{1}{c}{\bf Technique}     \\
 		\hline			

### 2018-07-09 src/compiler_new.py
model_evaluator.py: returned domain name corresponds to the reference domain

@@ -36,19 +36,19 @@ def possible_pred_for_action(task, p, a, tup):
     fits = [len(action_types[i].intersection(predicate_types[i])) >= 1 for i in range(len(action_types))]
     return all(fits)
 
-def get_static_predicates(traces, predicates):
+def get_static_predicates(state_trajectories, predicates):
 
     candidates = set([p.name for p in predicates])
 
-    for trace in traces:
+    for trajectory in state_trajectories:
         trace_candidates = set()
         for predicate in candidates:
             static = True
-            init_literals = set([l for l in trace.init if l.predicate == predicate])
-            for state in trace.states:
+            init_literals = set([l for l in trajectory[0] if l.predicate == predicate])
+            for state in trajectory[1:]:
                 state_literals = set([l for l in state if l.predicate == predicate])
 
-                if len(init_literals.intersection(state_literals)) != len(state_literals):
+                if init_literals != state_literals:
                     static = False
                     break
 
@@ -57,17 +57,18 @@ def get_static_predicates(traces, predicates):
 
         candidates = candidates.intersection(trace_candidates)
 
-    reflexive_static_predicates = dict()
-    for candidate in candidates:
-        reflexive_static_predicates[candidate] = True
-        for trace in traces:
-            init_literals = set([l for l in trace.init if l.predicate == candidate])
-            for literal in init_literals:
-                if len(literal.args) == 1 or len(set(literal.args)) != 1:
-                    reflexive_static_predicates[candidate] = False
-                    break
+    # reflexive_static_predicates = dict()
+    # for candidate in candidates:
+    #     reflexive_static_predicates[candidate] = True
+    #     for trace in traces:
+    #         init_literals = set([l for l in trace.init if l.predicate == candidate])
+    #         for literal in init_literals:
+    #             if len(literal.args) == 1 or len(set(literal.args)) != 1:
+    #                 reflexive_static_predicates[candidate] = False
+    #                 break
 
-    return [p for p in predicates if p.name in candidates], reflexive_static_predicates
+    return candidates
+    # return [p for p in predicates if p.name in candidates]
 
 
 def get_static_precondition(predicate, action, plans, tasks):
@@ -123,6 +124,22 @@ try:
[…]

### 2018-07-09 src/experimenter_new.py

@@ -1,5 +1,5 @@
 #! /usr/bin/env python
-import glob, os, sys, copy, itertools, math
+import glob, os, sys, copy, itertools, math, time
 import pddl, pddl_parser
 import fdtask_to_pddl
 import numpy as np
@@ -12,10 +12,27 @@ import config
 
 try:
     if "-s" in sys.argv:
-        check_static_predicates = True
+        check_static_predicates = "-s "
         sys.argv.remove("-s")
     else:
-        check_static_predicates = False
+        check_static_predicates = ""
+
+
+    if "-t" in sys.argv:
+        index = sys.argv.index("-t")
+        trace_prefix = sys.argv[index+1]
+        sys.argv.remove("-t")
+        sys.argv.remove(trace_prefix)
+    else:
+        trace_prefix = "trace"
+
+    if "-l" in sys.argv:
+        index = sys.argv.index("-l")
+        trace_limit = sys.argv[index+1]
+        sys.argv.remove("-l")
+        sys.argv.remove(sys.argv[index])
+    else:
+        trace_limit = None
 
     domain_folder_name  = sys.argv[1]
     action_observability = sys.argv[2]
@@ -32,15 +49,21 @@ except:
 # print("\n\nExecuting... " + cmd)
 # os.system(cmd)
 
-cmd = config.PROJECT_PATH + "src/compiler_new.py " + domain_folder_name + " " + action_observability + " " + state_observability
+cmd = config.PROJECT_PATH + "src/compiler_new.py " + check_static_predicates + domain_folder_name + " " + action_observability + " " + state_observability + " -t " + trace_prefix
+if trace_limit:
+    cmd += " -l " + trace_limit
 # print("\n\nExecuting... " + cmd)
+
+tic = time.time()
 os.system(cmd)
+toc = time.time()
+
+processing_time = toc - tic
 
 try:
     domain_name, best_evaluation, best_matches = model_evaluator.evaluate("learned_domain.pddl", domain_folder_name+"ref_domain", True)
     print(" & ".join(
-    [domain_name] + [str(round(e, 2)) for e in best_evaluation]) + " \\\\" + " % {}".format(
-    best_matches))
+    [domain_name] + [str(round(e, 2)) for e in best_evaluation]) + "& {}".format(round(processing_time,2)) + " \\\\" + " % {}".format(best_matches))
 except:
     print("No solution found")
 

### 2018-07-09 src/pddl_parser/parsing_functions.py

@@ -6,7 +6,7 @@ import sys
 
 import graph
 import pddl
-from random import random, seed
+import random
 import itertools
 
 def parse_typed_list(alist, only_variables=False,
@@ -496,7 +496,7 @@ def parse_task_pddl(task_pddl, type_dict, predicate_dict):
 
 
 def parse_trace_pddl(trace_pddl, predicates, action_observability=1, state_observability=1):
-    seed(123)
+    random.seed(123)
 
     iterator = iter(trace_pddl)
 
@@ -531,13 +531,13 @@ def parse_trace_pddl(trace_pddl, predicates, action_observability=1, state_obser
 
     for token in iterator:
         if token[0] == ':observations':
-            new_state = [literal for literal in parse_state(token[1:], all_literals) if random() <= state_observability]
+            new_state = [literal for literal in parse_state(token[1:], all_literals) if random.random() <= state_observability]
             states.append(new_state)
         elif token[0] == ':goal':
             goal = parse_state(token[1:], all_literals)
             break
         else:
-            if random() <= action_observability:
+            if random.random() <= action_observability:
                 actions.append(token)
             else:
                 actions.append([])
@@ -586,7 +586,7 @@ def parse_state(new_state, all_literals):
     state.extend(state_true)
     for atom in all_literals.difference(state_true):
         state.append(pddl.NegatedAtom(atom.predicate, atom.args))
-    return state
+    return sorted(state)
 
 
 def check_atom_consistency(atom, same_truth_value, other_truth_value, atom_is_true=True):

### 2018-07-09 doc/aij18/aij18.tex
updates added to learning section
updates added to sections 1-4

@@
 
-\title{A general framework for learning planning action models}
+\title{Learning action models with minimal observability}
 \author[label1]{Diego Aineto}
@@
 \input{learning_task}
-\input{compilation}
 \input{evaluation}
@@
 \section{Conclusions}
-\label{sec:Section9}
+\label{sec:conclusions}
 We presented a novel approach for learning \strips\ action models from examples using classical planning.

### 2018-07-09 doc/aij18/motivation.tex

@@
 The assumption of having \FO state trajectories means that the sensors are able to capture every state change at every instant, which typically is unrealistic.
-Normally, the process for obtaining state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two sensor readings.
+Normally, the process for obtaining state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two subsequent sensor readings.
 
@@
 
-\FAMA represents a step ahead towards learning action models without assuming observed actions.
-The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging.
-As Table \ref{tab:complex} shows, PSPACE-complexity happens when the length of the plan trace is a priori unbound, which results from the combination of \PO and \NO state trajectories and action sequences.
+\FAMA represents one step ahead towards learning action models without assuming observed actions.
+The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging (the actual length of the input plan trace becomes a priori unbound).
 This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now.
+Table \ref{tab:complex} shows when the worst case complexity of learning \strips\ action models becomes PSPACE-complete.
 
@@
 \FAMA shows that classical planning is a complete approach for this particular scenario.
-Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning techniques, as our proposal of compiling the learning task to a planning problem.
+Consequently, the new learning scenario features PSPACE-complete instead of NP-complete tasks, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.
 
@@
 This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM.
-Hence, since the set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects, it turns out that the syntactic metric may output an error rate below 100\% for a totally wrong learned model.
+The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects hence, it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models.
 To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}.
-These two syntactic metrics are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model.
+%{\em Precision} and {\em recall} are more informative that simply counting the number of errors in the learned model or computing the {\em symmetric difference} between the learned and the reference model.
 
 Pure syntax-based evaluation metrics, like the ones mentioned in the above paragraph, can report low scores for learned models that are actually {\em sound} and {\em complete} but syntactically different from the GTM.
-In this sense, \FAMA contributes with a novel semantic-based error measure that builds upon the precision and recall metrics.
-Semantic evaluation metrics add a distinctive value over the syntactic ones, which is that they evaluate the learned model with a set of observations of plan executions and hence they are very appropriate for scenarios where the GTM is not available.
-Unlike the semantic metric used in \ARMS~\cite{yang2007learning}, our proposal is not sensitive to the repetition of one same flaw in the model evaluation.
+Semantic evaluation metrics add a distinctive value over the syntactic ones, which is that they evaluate the learned model with a set of observations of plan executions and hence they are appropriate for scenarios where the GTM is not available.
+In this sense, \FAMA contributes also with a novel semantic-based error measure that builds upon the {\em precision} and {\em recall} metrics.
+Unlike the semantic metric used in \ARMS~\cite{yang2007learning}, our semantic version of {\em precision} and {\em recall} is not sensitive to the repetition of one same flaw in the model evaluation.
 

### 2018-07-12 src/example-generator.py
example-generator.py: replaced "inext" and "current" conditional effects for regular effects

@@ -59,13 +59,26 @@ if nhorizon > 0:
       params = []
       params += [pddl.pddl_types.TypedObject("?i1", "step")]
       params += [pddl.pddl_types.TypedObject("?i2", "step")]
+
+      a.parameters += params
    
       pre = []
       pre += [pddl.conditions.Atom("current", ["?i1"])]
       pre += [pddl.conditions.Atom("inext", ["?i1", "?i2"])]
+
+      if a.precondition.parts:
+         original_pre = list(a.precondition.parts)
+      else:
+         original_pre = [a.precondition]
+
+      a.precondition = pddl.conditions.Conjunction(original_pre + pre)
    
-      a.effects += [pddl.effects.Effect(params, pddl.conditions.Conjunction(pre), pddl.conditions.NegatedAtom("current", ["?i1"]))]
-      a.effects += [pddl.effects.Effect(params, pddl.conditions.Conjunction(pre), pddl.conditions.Atom("current", ["?i2"]))]
+      # a.effects += [pddl.effects.Effect(params, pddl.conditions.Conjunction(pre), pddl.conditions.NegatedAtom("current", ["?i1"]))]
+      # a.effects += [pddl.effects.Effect(params, pddl.conditions.Conjunction(pre), pddl.conditions.Atom("current", ["?i2"]))]
+      a.effects += [
+         pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.NegatedAtom("current", ["?i1"]))]
+      a.effects += [
+         pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom("current", ["?i2"]))]
       a.effects += [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom("applied-"+a.name, []))]      
       
    for i in range(1, nhorizon + 1):
@@ -77,10 +90,10 @@ if nhorizon > 0:
    fd_task.init.append(pddl.conditions.Atom("current", ["i1"]))
 
    new_goals = []
-   for a in fd_task.actions:
-      new_goals.append(pddl.conditions.Atom("applied-"+a.name, []))
-   while(len(new_goals)>6):
-      new_goals.pop(random.randint(0,len(new_goals)-1))
+   # for a in fd_task.actions:
+   #    new_goals.append(pddl.conditions.Atom("applied-"+a.name, []))
+   # while(len(new_goals)>6):
+   #    new_goals.pop(random.randint(0,len(new_goals)-1))
    
[…]

### 2018-07-12 doc/aij18/experiments.tex
updates added to sections 1-4

@@
 \section{Experimental results}
-\label{experiments}
+\label{sec:experiments}
 
 \subsection{Setup}
-The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only use 5 learning examples for each domain and they are fixed for all the experiments so we can evaluate the impact of the input knowledge in the quality of the learned models.
-All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM.
+In order to asses the performance of our approach, we evaluate \FAMA on a wide range of domains.
+All tested domains are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+Table \ref{tab:domain_features} compiles the features of the thirteen domains used in the experiments.
+From left to right, the columns report the name of the domain, number of actions, number of predicates, maximum arity of the actions, and maximum arity of the predicates.
+These are all features which affect the size of $P_\Lambda$ and, hence, have an impact on the complexity of the learning task.
+As can be seen in table \ref{tab:domain_features}, the selected domains range from simple ones, such as \emph{blocks} and \emph{visitall}, to complex ones, like \emph{floortile} and \emph{satellite}, with higher number of actions and predicates and with higher arity.
+
+Following, we explain the details of our experimental setup:
+
 \begin{itemize}
+
+\item {\bf Plan traces}.
+For each domain, we have generated 10 plan traces with 10 actions/intermediate states each using random walks.
+Depending on the experiment, these traces may be used for training or testing, so more details will be provided later.
+
 \item {\bf Planner}.
 The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use {\sc Madagascar} because its ability to deal with planning instances populated with dead-ends.
-In addition, SAT-based planners can apply the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
-Actions for programming action effects can also be applied in a single planning step reducing significantly the planning horizon.
+We use {\sc Madagascar} for two reasons: 1) its ability to deal with planning instances populated with dead-ends, and 2) when the length of the plan traces is known (\FO action sequences or \FO state trajectories), the horizon of the solution plan is also known and can be solved as a NP-complete problem.
+This is because, SAT-based planners can apply the actions for programming preconditions in parallel during a single planning step (and the same for actions programming effects) because these actions do not interact.
+Hence, we know that the programming prefix of the solution plan can be solved in two steps, independently of the number of programming actions applied.
+
+\item {\bf Machine}. All experiments are run on an Intel Core i5 3.50 GHz x 4 with 16 GB of RAM.
+
 \item {\bf Reproducibility}.
@@
 
-\subsection{Evaluating with a reference model}
-Here we evaluate the learned models with respect to the actual generative model.
+\begin{table}[hbt!]
+		\begin{center}
+			\begin{tabular}{l|c|c|c|c|}	
+				& \multicolumn{4}{c|}{Domain features}\\ \cline{2-5}
+				Domain & \# actions & \# predicates & max action arity & max predicate arity  \\
+				\hline
+				Blocks & 4 & 5 & 2 & 2  \\
+				Driverlog & 6 & 5 & 4 & 2  \\
+				Ferry & 3 & 5 & 2 & 2  \\
+				Floortile & 7 & 10 & 4 & 2  \\
+				Grid & 5 & 9 & 4 & 2  \\
+				Gripper & 3 & 4 & 3 & 2  \\
+				Hanoi & 1 & 3 & 3 & 2  \\
+				Miconic & 4 & 6 & 2 & 2  \\
+				Npuzzle & 1 & 3 & 3 & 2  \\
+				Parking & 4 & 5 & 3 & 2  \\
+				Satellite & 5 & 8 & 4 & 2  \\
+				Transport & 3 & 5 & 5 & 2  \\
+				Visitall & 1 & 3 & 2 & 2  \\
+			\end{tabular}
+		\end{center}
+	\caption{\small Feature description of the 13 domains used in the experiments.}
+	\label{tab:domain_features}	
+\end{table}
 
-\subsubsection{Learning from plans}
-We start evaluating our approach with $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$ learning tasks, where the action of the executed plans are available and the state observation sequence contains only the corresponding initial and goal states, $s_o^t$ and $s_n^t$, for every trace plan $t\in\mathcal{T}$.
-We then repeat the evaluation but exploiting potential \emph{static predicates} computed from the observed states in $\mathcal{T}$, which are the predicates that appear unaltered in the states that belong to the same plan.
-Static predicates are used to constrain the space of possible action models as explained in Section~\ref{sec:Section5}.
+\subsection{Impact of the size of the input knowledge}
+In our first experiment, we evaluate how the size of the input knowledge, i.e., $\left|\mathcal{T}\right|$ affects \FAMA.
+The experiment consists in increasing the size of $\mathcal{T}$ from 1 to 10, and analyze the evolution of the computation time as well as the quality of the learned models.
+Our goals with this experiment are:
+\begin{enumerate}
+	\item Identify the amount of input required by \FAMA to learn good models,
+	\item Evaluate the scalability of \FAMA wrt the input size which, as stated in section \ref{properties}, is one of the limiting factors of our approach.
+\end{enumerate}
 
-Table~\ref{tab:results_plans} shows the obtained results.
-Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), while the last two columns of each setting and the last row report averages values.
+With this in mind we have defined two case studies:
+\begin{itemize}
+	\item \textbf{\FO action sequence and \PO state trajectory}: This is the typical case solved by most approaches, which corresponds to a NP-complete scenario.
+We are assuming a degree of observability of 10\% for the state trajectory, meaning that each literal of a state has a 10\% chance of being observed.
+	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a PSPACE-complete scenario where both the action sequence and state trajectory are completely empty and only initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
+\end{itemize}
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=0.8\linewidth]{figures/input_size_100_10_precision.pdf}
+	\caption{Evaluation of the impact of the input size on the quality of the learned models when learning from plan traces with \FO action sequences and \PO state trajectories with 10\% observability}
+\end{figure}
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=0.8\linewidth]{figures/input_size_100_10_time.pdf}
+	\caption{Evaluation of the impact of the input size on the computation time when learning from plan traces with \FO action sequences and \PO state trajectories with 10\% observability}
+\end{figure}
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=0.8\linewidth]{figures/input_size_0_0_precision.pdf}
+	\caption{Evaluation of the impact of the input size on the quality of the learned models when learning from plan traces with \NO action sequences and \NO state trajectories}
+\end{figure}
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=0.8\linewidth]{figures/input_size_0_0_time.pdf}
+	\caption{Evaluation of the impact of the input size on the computation time when learning from plan traces with \NO action sequences and \NO state trajectories}
+\end{figure}
+
+\subsection{Comparison with \ARMS}
+Here we compare the performance of \FAMA to that of \ARMS, one of the most well-known approaches in the action model acquisition field.
+\ARMS works under the assumption of plan traces with \FO action sequences and \NO state trajectories, so we will also adopt this assumption.
+In more detail, we define a \emph{degree of observability} $\sigma$ for the state trajectory, ranging from 0\% to 100\%, that measures the probability of observing a literal, and evaluate both approaches as $\sigma$ increases.
+Note that when $\sigma = 0$ we have a \NO state trajectory and when $\sigma=100$ we have a \FO state trajectory, all cases in-between correspond to the \PO scenario.
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=.8\linewidth]{figures/comparison_precision.pdf}
+	\caption{Precision comparison between \FAMA and \ARMS}
+	\label{fig:comparison_precision}
+\end{figure}
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=.8\linewidth]{figures/comparison_recall.pdf}
+	\caption{Recall comparison between \FAMA and \ARMS}
+	\label{fig:comparison_recall}
+\end{figure}
+
+Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of precision and recall.
+The horizontal axis represents the degree of observability, while the vertical axis show the average precision (figure \ref{fig:comparison_precision}) or recall (figure \ref{fig:comparison_recall}) computed over the 13 tested domains.
+
+\subsection{Experiments with minimal input knowledge}
+In the previous experiments we have given a general view of the performance of \FAMA under different conditions.
+So far, experiments have shown that \FAMA is able to learn with very small amounts of input knowledge, be it due to low observability or few training samples.
+In this section we want to take a closer look at the action models learned from minimal input knowledge.
+To that end, we will limit the input to only 2 plan traces and analyze the results under different levels of observability.
+
+All tables in this section (tables \ref{tab:results_minimum_0_0}, \ref{tab:results_minimum_0_20}) follow the same structure.
+In these tables precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), and also globally ({\bf Global}).
+The last column reports the computation time in seconds needed to obtain the learned models.
 We can observe that identifying static predicates leads to models with better precondition {\em recall}.
@@
 
-\begin{table*}[hbt!]
-		\resizebox{\textwidth}{!}{%
-		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|l|l|l|l|l||l|l|}
-& \multicolumn{8}{|c||}{\bf No Static}& \multicolumn{8}{|c|}{\bf Static}\\\cline{2-17}
-& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c||}{\bf}& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c|}{\bf Del} & \multicolumn{2}{|c|}{\bf}\\
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R}& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
+\begin{table}[hbt!]
+		\begin{center}
+
+			\begin{tabular}{l|l|l|l|l|l|l||l|l|l|}
+				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c|}{\bf Global} & \\ \cline{2-9}			
+				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & {\bf Time} \\
 				\hline
-			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Driverlog & 1.0 & 0.36 & 0.75 & 0.86 & 1.0 & 0.71 & 0.92 & 0.64 & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
-			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
-			Floortile & 0.52 & 0.68 & 0.64 & 0.82 & 0.83 & 0.91 & 0.66 & 0.80 & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
-			Grid & 0.62 & 0.47 & 0.75 & 0.86 & 0.78 & 1.0 & 0.71 & 0.78 & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
-			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
-			Hanoi & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
-			Miconic & 0.75 & 0.33 & 0.50 & 0.50 & 0.75 & 1.0 & 0.67 & 0.61 & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
-			Satellite & 0.60 & 0.21 & 1.0 & 1.0 & 1.0 & 0.75 & 0.87 & 0.65 & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
-			Transport & 1.0 & 0.40 & 1.0 & 1.0 & 1.0 & 0.80 & 1.0 & 0.73 & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
-			Visitall & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
-			Zenotravel & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0 & 0.69 &1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
+				blocks & 0.6 & 0.67 & 0.29 & 0.22 & 0.75 & 0.33 & 0.55 & 0.41& 0.22 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'unstack', 2, 1), (u'unstack', u'stack', 2, 1)]
+				driverlog & 0.45 & 0.36 & 0.36 & 0.57 & 0.17 & 0.14 & 0.33 & 0.36& 1.59 \\ % [(u'load-truck', u'walk', 2, 1, 3), (u'unload-truck', u'disembark-truck', 1, 2, 3), (u'board-truck', u'unload-truck', 1, 2, 3), (u'disembark-truck', u'board-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
+				ferry & 0.6 & 0.43 & 0.4 & 0.5 & 0.5 & 0.5 & 0.5 & 0.48& 0.51 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
+				floor-tile & 0.67 & 0.45 & 0.6 & 0.55 & 0.88 & 0.64 & 0.71 & 0.55& 359.43 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'up', 1, 3, 2), (u'down', u'down', 1, 3, 2), (u'right', u'right', 1, 2, 3), (u'left', u'left', 1, 2, 3), (u'paint-up', u'paint-up', 1, 3, 2, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
+				grid & 0.37 & 0.41 & 0.5 & 0.71 & 0.33 & 0.43 & 0.4 & 0.52& 81.71 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'putdown', 1, 2), (u'putdown', u'pickup', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 3, 2), (u'unlock', u'unlock', 1, 2, 3, 4)]
+				gripper-strips & 0.8 & 0.67 & 0.6 & 0.75 & 0.8 & 1.0 & 0.73 & 0.81& 0.09 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
+				hanoi & 1.0 & 0.5 & 0.5 & 0.5 & 1.0 & 1.0 & 0.83 & 0.67& 0.87 \\ % [(u'move', u'move', 1, 3, 2)]
+				miconic & 0.67 & 0.22 & 0.67 & 0.5 & 1.0 & 0.67 & 0.78 & 0.46& 0.25 \\ % [(u'board', u'depart', 1, 2), (u'depart', u'board', 1, 2), (u'up', u'up', 2, 1), (u'down', u'down', 1, 2)]
+				n-puzzle & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 0.47 \\ % [(u'move', u'move', 1, 2, 3)]
+				parking & 0.64 & 0.5 & 0.63 & 0.56 & 0.5 & 0.44 & 0.59 & 0.5& 12.56 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 3, 2), (u'move-curb-to-car', u'move-curb-to-car', 3, 2, 1), (u'move-car-to-curb', u'move-car-to-curb', 2, 1, 3), (u'move-car-to-car', u'move-car-to-car', 1, 3, 2)]
+				satellite & 0.38 & 0.21 & 0.5 & 0.6 & 0.75 & 0.75 & 0.54 & 0.52& 1.86 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
+				transport & 0.43 & 0.3 & 0.4 & 0.4 & 0.0 & 0.0 & 0.28 & 0.23& 0.88 \\ % [(u'drive', u'drive', 1, 3, 2), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
+				grid-visit-all & 0.0 & 0.0 & 1.0 & 0.5 & 0.0 & 0.0 & 0.33 & 0.17& 1.54 \\ % [(u'move', u'move', 2, 1)]			
 				\hline
-			\bf  & 0.88 & 0.50 & 0.88 & 0.92 & 0.95 & 0.91 & 0.90 & 0.78 & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
+				\bf & 0.51& 0.36 & 0.50 & 0.49 & 0.51 & 0.45 & 0.51 & 0.44 & 35.54 \\
 			\end{tabular}
-	}
-\caption{\small {\em Precision} and {\em recall} scores for learning tasks from labeled plans without (left) and with (right) static predicates.}
-\label{tab:results_plans}
-\end{table*}
 
-Table~\ref{tab:time_plans} reports the total planning time, the preprocessing time (in seconds) invested by {\sc Madagascar} to solve the planning instances that result from our compilation as well as the number of actions of the solution plans.
-All the learning tasks are solved in a few seconds.
-Interestingly, one can identify the domains with static predicates by just looking at the reported plan length.
-In these domains some of the preconditions that correspond to static predicates are directly derived from the learning examples and therefore fewer programming actions are required.
-When static predicates are identified, the resulting compilation is also much more compact and produces smaller planning/instantiation times.
-
-\begin{table}
-\begin{scriptsize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c||c|c|c|}
-                         & \multicolumn{3}{|c||}{\bf No Static}& \multicolumn{3}{|c|}{\bf Static}\\
-			 & Total & Preprocess & Length  & Total & Preprocess &  Length\\
-                         \hline
-			Blocks & 0.04 & 0.00 & 72  & 0.03 & 0.00 & 72 \\
-			Driverlog & 0.14 & 0.09 & 83 & 0.06 & 0.03 & 59 \\
-			Ferry & 0.06 & 0.03 & 55 & 0.06 & 0.03 & 55 \\
-			Floortile & 2.42 & 1.64 & 168 & 0.67 & 0.57 & 77 \\
-			Grid & 4.82 & 4.75 & 88 & 3.39 & 3.35 & 72 \\
-			Gripper & 0.03 & 0.01 & 43 & 0.01 & 0.00 & 43 \\
-                        Hanoi & 0.12 & 0.06 & 48 & 0.09 & 0.06 & 39 \\
-                        Miconic & 0.06 & 0.03 & 57 & 0.04 & 0.00 & 41 \\
-			Satellite & 0.20 & 0.14 & 67 & 0.18 & 0.12 & 60 \\
-			Transport & 0.59 & 0.53 & 61 & 0.39 & 0.35 & 48 \\
-			Visitall & 0.21 & 0.15 & 40 & 0.17 & 0.15 & 36 \\
-			Zenotravel & 2.07 & 2.04 & 71 & 1.01 & 1.00 & 55 \\			
-		\end{tabular}
 		\end{center}
-        \end{scriptsize}
-	\caption{\small Total planning time, preprocessing time and plan length for learning tasks from labeled plans without/with static predicates.}
-	\label{tab:time_plans}	
+	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \NO action sequences and \NO state trajectories}
+	\label{tab:results_minimum_0_0}
 \end{table}
 
-We evaluate now the ability of our approach to support partially specified action models; that is, when the input model $\mathcal{M}$ is not empty because some preconditions and effects of the actions are initially known.
-In this particular experiment, the model of half of the actions is given in $\mathcal{M}$ as an extra input of the learning task.
-Tables~\ref{tab:results_plans_partial} and~\ref{tab:time_plans_partial} summarize the obtained results, which include the identification of static predicates.
-We only report the {\em precision} and {\em recall} of the {\em unknown} actions since the values of the metrics of the {\em known} action models is 1.0.
-In this experiment, a low value of {\em precision} or {\em recall} has a greater impact than in the previous learning tasks because the evaluation is done only over half of the actions.
-This occurs, for instance, in the precondition \emph{recall} of domains such as {\em Floortile}, {\em Gripper} or {\em Satellite}.
+\begin{table}[hbt!]
+	\begin{center}
 
-Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|l|}
+			& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c|}{\bf Global} & \\ \cline{2-9}			
+			& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & {\bf Time} \\
+			\hline
+			blocks & 0.78 & 0.78 & 0.6 & 0.67 & 0.8 & 0.44 & 0.73 & 0.63& 0.27 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'unstack', 1, 2), (u'unstack', u'stack', 2, 1)]
+			driverlog & 0.5 & 0.07 & 0.18 & 0.57 & 0.0 & 0.0 & 0.23 & 0.21& 1.01 \\ % [(u'load-truck', u'walk', 2, 1, 3), (u'unload-truck', u'board-truck', 1, 2, 3), (u'board-truck', u'unload-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
+			ferry & 0.83 & 0.71 & 0.75 & 0.75 & 0.8 & 1.0 & 0.79 & 0.82& 0.37 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'debark', 1, 2), (u'debark', u'board', 1, 2)]
+			floor-tile & 0.68 & 0.59 & 0.62 & 0.73 & 0.75 & 0.55 & 0.68 & 0.62& 186.92 \\ % [(u'change-color', u'change-color', 1, 3, 2), (u'up', u'right', 1, 2, 3), (u'down', u'up', 1, 3, 2), (u'right', u'down', 1, 2, 3), (u'left', u'left', 1, 3, 2), (u'paint-up', u'paint-down', 1, 2, 3, 4), (u'paint-down', u'paint-up', 1, 2, 3, 4)]
+			grid & 0.46 & 0.35 & 0.25 & 0.43 & 0.63 & 0.71 & 0.45 & 0.5& 107.9 \\ % [(u'move', u'move', 2, 1), (u'pickup', u'pickup', 1, 2), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
+			gripper-strips & 0.57 & 0.67 & 0.67 & 0.5 & 1.0 & 0.75 & 0.75 & 0.64& 0.06 \\ % [(u'move', u'move', 1, 2), (u'pick', u'drop', 1, 2, 3), (u'drop', u'pick', 1, 2, 3)]
+			hanoi & 0.6 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.87 & 0.92& 1.09 \\ % [(u'move', u'move', 1, 3, 2)]
+			miconic & 1.0 & 0.11 & 0.33 & 0.5 & 1.0 & 0.33 & 0.78 & 0.31& 0.17 \\ % [(u'board', u'board', 1, 2), (u'depart', u'up', 1, 2), (u'up', u'depart', 1, 2), (u'down', u'up', 2, 1)]
+			n-puzzle & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 1.0& 2.5 \\ % [(u'move', u'move', 1, 2, 3)]
+			parking & 0.45 & 0.36 & 0.4 & 0.44 & 0.63 & 0.56 & 0.49 & 0.45& 5.87 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 3, 2), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-car', 2, 1, 3), (u'move-car-to-car', u'move-car-to-car', 3, 1, 2)]
+			satellite & 0.67 & 0.43 & 0.25 & 0.4 & 0.29 & 0.5 & 0.4 & 0.44& 1.34 \\ % [(u'switch-on', u'switch-off', 1, 2), (u'switch-off', u'switch-on', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
+			transport & 0.33 & 0.1 & 0.4 & 0.4 & 0.0 & 0.0 & 0.24 & 0.17& 1.14 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
+			grid-visit-all & 0.5 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.83& 1.08 \\ % [(u'move', u'move', 2, 1)]
+
+			\hline
+			\bf & 0.62 & 0.49 & 0.57 & 0.65 & 0.68 & 0.60 & 0.63 & 0.58	& 23.82	 \\
+		\end{tabular}
+
+	\end{center}
+	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \NO action sequences and \PO state trajectories with 20\% observability}
+	\label{tab:results_minimum_0_20}
+\end{table}
+
+%Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
 The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
@@
 
-\begin{table}[hbt!]
-\begin{footnotesize}
-	\begin{center}
-
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-				Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Driverlog & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.90 \\
-				Ferry & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 \\
-				Floortile & 0.75 & 0.60 & 1.0 & 0.80 & 1.0 & 0.80 & 0.92 & 0.73 \\
-                Grid & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.84 & 0.78 \\
-				Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-				Miconic & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				Satellite & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 \\
-				Transport & 1.0 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 \\
-				Zenotravel & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.78 \\
-				\hline
-				\bf  & 0.98 & 0.71 & 1.0 & 0.98 & 1.0 & 0.95 & 0.98 & 0.87 \\
-			\end{tabular}
-
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} scores for learning tasks with partially specified action models.}
-\label{tab:results_plans_partial}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-                         \hline
-			Blocks & 0.07 & 0.01 & 54  \\
-			Driverlog & 0.03 & 0.01 & 40 \\
-			Ferry & 0.06 & 0.03 & 45 \\
-			Floortile & 0.43 & 0.42 & 55 \\
-                        Grid & 3.12 & 3.07 & 53 \\
-			Gripper & 0.03 & 0.01 & 35 \\
-			Miconic & 0.03 & 0.01 & 34  \\
-			Satellite & 0.14 & 0.14 & 47 \\
-			Transport & 0.23 & 0.21 & 37 \\
-			Zenotravel & 0.90 & 0.89 & 40 \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Time and plan length learning for learning tasks with partially specified action models.}
-	\label{tab:time_plans_partial}	
-\end{table}
-
-\subsubsection{Learning from state observations}
-Here we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, where the action of the executed plans are not available but the initial and goal states are known.
+%Here we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, where the action of the executed plans are not available but the initial and goal states are known.
 When input plans are not available, the planner must not only compute the action models but also the plans that satisfy the input observations.
@@
 
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-			 & \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-			  & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-			\hline
-            Blocks & 0.33 & 0.33 & 0.75 & 0.50 & 0.33 & 0.33 & 0.47 & 0.39 \\
-            Driverlog & 1.0 & 0.29 & 0.33 & 0.67 & 1.0 & 0.50 & 0.78 & 0.48 \\
-            Ferry & 1.0 & 0.67 & 0.50 & 1.0 & 1.0 & 1.0 & 0.83 & 0.89 \\
-            Floortile & 0.67 & 0.40 & 0.50 & 0.40 & 1.0 & 0.40 & 0.72 & 0.40 \\
-            Grid & - & - & - & - & - & - & - & - \\
-            Gripper & 1.0 & 0.50 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-            Miconic & 0.0 & 0.0 & 0.33 & 0.50 & 0.0 & 0.0 & 0.11 & 0.17 \\
-            Satellite & 1.0 & 0.14 & 0.67 & 1.0 & 1.0 & 1.0 & 0.89 & 0.71 \\
-            Transport & 0.0 & 0.0 & 0.25 & 0.5 & 0.0 & 0.0 & 0.08 & 0.17 \\
-            Zenotravel & - & - & - & - & - & - & - & - \\
-            \hline
-            & 0.63 & 0.29 & 0.54 & 0.70 & 0.67 & 0.53 & 0.61 & 0.51 \\			
-			\end{tabular}
-	\end{center}
-\end{footnotesize}
-\caption{\small {\em Precision} and {\em recall} scores for learning from (initial,final) state pairs.}
-\label{tab:results_states}
-\end{table}
-
-\begin{table}
-\begin{footnotesize}
-	\begin{center}
-		\begin{tabular}{l|c|c|c|}			
-			 & Total time & Preprocess & Plan length  \\
-			\hline
-            Blocks & 2.14 & 0.00 & 58  \\
-            Driverlog & 0.09 & 0.00 & 88 \\
-            Ferry & 0.17 & 0.01 & 65 \\
-            Floortile & 6.42 & 0.15 & 126 \\
-            Grid & - & - & - \\
-            Gripper & 0.03 & 0.00 & 47 \\
-            Miconic & 0.04 & 0.00 & 68 \\
-            Satellite & 4.34 & 0.10 & 126 \\
-            Transport & 2.57 & 0.21 & 47 \\			
-            Zenotravel & - & - & - \\
-		\end{tabular}
-	\end{center}
-        \end{footnotesize}
-	\caption{\small Time and plan length when learning from (initial,final) state pairs.}
-	\label{tab:time_states}	
-\end{table}
-
-Now we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, where the action of the executed plans are not available but where the plan trace $\mathcal{T}$ contains all the intermediate states, not just the initial and final states.
-Table~\ref{fig:observationsnomap} shows the precision ({\bf P}) and recall ({\bf R}) computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}) while the last two columns report averages values.
-The reason why the scores in Table ~\ref{fig:observationsnomap} are still low, despite more state observations are available, is because the syntax-based nature of {\em precision} and {\em recall} make these two metrics report low scores for learned models that are semantically correct but correspond to {\em reformulations} of the actual model (changes in the roles of actions with matching headers or parameters with matching types).
-
-\begin{table}[hbt!]
-	\begin{center}
-		\begin{scriptsize}
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-				\hline
-
-				blocks & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 & 0.44 \\
-				driverlog & 0.0 & 0.0 & 0.25 & 0.43 & 0.0 & 0.0 & 0.08 & 0.14 \\
-				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
-				floortile & 0.38 & 0.55 & 0.4 & 0.18 & 0.56 & 0.45 & 0.44 & 0.39 \\
-				grid & 0.5 & 0.47 & 0.33 & 0.29 & 0.25 & 0.29 & 0.36 & 0.35 \\
-				gripper & 0.83 & 0.83 & 0.75 & 0.75 & 0.75 & 0.75 & 0.78 & 0.78 \\
-				hanoi & 0.5 & 0.25 & 0.5 & 0.5 & 0.0 & 0.0 & 0.33 & 0.25 \\
-				hiking & 0.43 & 0.43 & 0.5 & 0.35 & 0.44 & 0.47 & 0.46 & 0.42 \\
-				miconic & 0.6 & 0.33 & 0.33 & 0.25 & 0.33 & 0.33 & 0.42 & 0.31 \\
-				npuzzle & 0.33 & 0.33 & 0.0 & 0.0 & 0.0 & 0.0 & 0.11 & 0.11 \\
-				parking & 0.25 & 0.21 & 0.0 & 0.0 & 0.0 & 0.0 & 0.08 & 0.07 \\
-				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
-				transport & 1.0 & 0.3 & 0.8 & 0.8 & 1.0 & 0.6 & 0.93 & 0.57 \\
-				visitall & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 \\
-				zenotravel & 0.67 & 0.29 & 0.33 & 0.29 & 0.33 & 0.14 & 0.44 & 0.24
-			\end{tabular}
-		\end{scriptsize}
-	\end{center}
-	\caption{\small Precision and recall values obtained without computing the $f_{P\&R}$ mapping with the reference model.}
-	\label{fig:observationsnomap}
-\end{table}
-
-To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model {\em reformulations}.
+%To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model {\em reformulations}.
 Precision and recall are often combined using the {\em harmonic mean}.
@@
 
-Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
+%Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
 The {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
@@
 
-\begin{table}
-        \begin{center}
-
-		\begin{scriptsize}
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|}
-				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c}{\bf}\\ \cline{2-9}			
-				& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} \\
-				\hline
-
-				blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				driverlog & 0.67 & 0.14 & 0.33 & 0.57 & 0.67 & 0.29 & 0.56 & 0.33 \\
-				ferry & 1.0 & 0.71 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.9 \\
-				floortile & 0.44 & 0.64 & 1.0 & 0.45 & 0.89 & 0.73 & 0.78 & 0.61 \\
-				grid & 0.63 & 0.59 & 0.67 & 0.57 & 0.63 & 0.71 & 0.64 & 0.62 \\
-				gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 \\
-				hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 \\
-				hiking & 0.78 & 0.6 & 0.93 & 0.82 & 0.88 & 0.88 & 0.87 & 0.77 \\
-				miconic & 0.8 & 0.44 & 1.0 & 0.75 & 1.0 & 1.0 & 0.93 & 0.73 \\
-				npuzzle & 0.67 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 0.89 \\
-				parking & 0.56 & 0.36 & 0.5 & 0.33 & 0.5 & 0.33 & 0.52 & 0.34 \\
-				satellite & 0.6 & 0.21 & 0.8 & 0.8 & 1.0 & 0.5 & 0.8 & 0.5 \\
-				transport & 1.0 & 0.3 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0 & 0.63 \\
-				visitall & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89 & 1.0 \\
-				zenotravel & 1.0 & 0.43 & 0.67 & 0.57 & 1.0 & 0.43 & 0.89 & 0.48
-			\end{tabular}
-		\end{scriptsize}
-	\end{center}
-	\caption{\small Precision and recall values obtained when computing the $f_{P\&R}$ mapping with the reference model.}
-	\label{fig:observationsmap}
-\end{table}
-
-\subsection{Evaluating with a test set}
-
-When a reference model is not available, the learned models are tested with an observation set.
-Table~\ref{fig:observationstest} summarizes the results obtained when evaluating the quality of the learned models with respect to a test set of state observations.
-Each test set comprises between 20 and 50 observations per domain and is generated executing the plans for various instances of the IPC domains and collecting the intermediate states.
-The table shows, for each domain, the {\em observation edit distance} (computed with our extended compilation), the {\em maximum edit distance}, and their ratio.
-The reported results show that, despite learning only from 25 state observations, 12 out of 15 learned domains yield ratios of $90\%$ or above.
-This fact evidences that the learned models require very small amounts of edition to match the observations of the given test set.
-
-\begin{table}[hbt!]
-		\begin{center}
-                \begin{footnotesize}
-			\begin{tabular}{l|r|r|c|}
-				& $\delta(\mathcal{M},\mathcal{O})$ & $\delta(\mathcal{M},*)$ & $1-\frac{\delta(\mathcal{M},\mathcal{O})}{\delta(\mathcal{M},*)}$ \\
-				\hline
-				blocks & 0 & 90 & 1.0 \\
-				driverlog & 5 & 144 & 0.97 \\
-				ferry & 2 & 69 & 0.97 \\
-				floortile & 34 & 342 & 0.90 \\
-				grid & 42 & 153 & 0.73 \\
-				gripper & 2 & 30 & 0.93 \\
-				hanoi & 1 & 63 & 0.98 \\
-				hiking & 69 & 174 & 0.60 \\
-				miconic & 3 & 72 & 0.96 \\
-				npuzzle & 2 & 24 & 0.92 \\
-                                parking & 4 & 111 & 0.96 \\
-				satellite & 24 & 75 & 0.68 \\
-				transport & 4 & 78 & 0.95 \\
-				visitall & 2 & 24 & 0.92 \\
-				zenotravel & 3 & 63 & 0.95
-			\end{tabular}
-                        	\end{footnotesize}
-		\end{center}
-	\caption{\small Evaluation of the quality of the learned models with respect to an observations test set.}
-	\label{fig:observationstest}
-\end{table}
-
-The learning scores of several domains in Table~\ref{fig:observationsmap} are above the ones reported in Table~\ref{fig:observationsnomap}.
+%The learning scores of several domains in Table~\ref{fig:observationsmap} are above the ones reported in Table~\ref{fig:observationsnomap}.
 The reason lies in the particular observations comprised by the test sets.

### 2018-07-16 src/example-generator.py
new observations computed from bad quality plans

@@ -1,5 +1,5 @@
 #! /usr/bin/env python
-import sys, os, copy, glob, itertools, random
+import sys, os, glob
 import pddl, pddl_parser
 import fdtask_to_pddl, planning
 import config
@@ -8,7 +8,7 @@ import config
 # Madagascar details
 M_PATH=config.PLANNER_PATH
 M_CALL="/"+config.PLANNER_NAME
-M_PARAMS=" -W " + config.PLANNER_PARAMS
+M_PARAMS=" -W -T 150 -F 150 -P 1 -S 20 -Q -o " + config.OUTPUT_FILENAME
 
 # FD details
 FD_PATH="/home/slimbook/software/fd/"
@@ -29,104 +29,54 @@ try:
    problem_filename = sys.argv[2]
    planner = sys.argv[3]   
    nsteps = int(sys.argv[4])
-
-   if "-h" in sys.argv:
-      sys.argv.remove("-h")      
-      nhorizon = int(sys.argv[5])
-   else:
-      nhorizon = 0                
-   
+   nhorizon = int(sys.argv[5])      
 except:
    print "Usage:"
-   print sys.argv[0] + " <domain> <problem> <planner> <steps> <-h horizon>"
+   print sys.argv[0] + " <domain> <problem> <planner> <steps> <horizon>"
    sys.exit(-1)
 
+   
 # Creating a FD task with the domain and the problem file
 fd_domain = pddl_parser.pddl_file.parse_pddl_file("domain", domain_filename)
 fd_problem = pddl_parser.pddl_file.parse_pddl_file("task", problem_filename)
 fd_task = pddl_parser.pddl_file.parsing_functions.parse_task(fd_domain, fd_problem)
 
-# Modifying domain and problem when planning for horizon
-if nhorizon > 0:
-  
-   fd_task.types.append(pddl.pddl_types.Type("step", "None"))
-   fd_task.predicates.append(pddl.predicates.Predicate("current", [pddl.pddl_types.TypedObject("?i", "step")]))
-   fd_task.predicates.append(pddl.predicates.Predicate("inext", [pddl.pddl_types.TypedObject("?i1", "step"), pddl.pddl_types.TypedObject("?i2", "step")]))
-   for a in fd_task.actions:
-      fd_task.predicates.append(pddl.predicates.Predicate("applied-"+a.name, []))  
-
-   for a in fd_task.actions:
-      params = []
-      params += [pddl.pddl_types.TypedObject("?i1", "step")]
[…]

### 2018-07-16 doc/aij18/background.tex
small updates

@@
 
-A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states, i.e.
+A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states; i.e.
 $|s|=|F|$, so the size of the state space is $2^{|F|}$.
@@
 Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
 
@@
 
-With regard to the observed actions in a plan trace, we say that an action sequence, $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to transit every state $s_{i-1}$ to the corresponding successor state $s_{i}$, for each {\small $1\leq i\le n$}.
-We say that the action sequence is \PO, if at least one of these necessary actions is missing in the sequence and that is \NO, when the sequence of observed actions is empty.
+With regard to the observed actions in a plan trace, we say that $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to transit every state $s_{i-1}$ to the corresponding successor state $s_{i}$, for each {\small $1\leq i\le n$};  we say it is a \PO action sequence if at least one of these necessary actions is missing in the sequence, and a \NO action sequence when the sequence of observed actions is empty.
 
-A plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$, holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}.
+A plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$ holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}.
 Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).

### 2018-07-21 doc/aij18/evaluation.tex
Add files via upload

@@
 In this section we introduce the metrics used by \FAMA to evaluate the action models that result from solving a learning task $\Lambda$.
-First, we will describe two standard syntactic metrics and then a semantic evaluation measure that leverages these two syntactic metrics is defined in section \ref{semantic_precision_recall}.
-Finally, section \ref{edit_distance} explains how \FAMA computes the semantic-based metric.
+First, we will describe two standard syntactic metrics ({\em precision} and {\em recall}) and then section \ref{semantic_precision_recall} will define a semantic evaluation measure that builds upon {\em precision} and {\em recall}.
+Finally, section \ref{edit_distance} explains how \FAMA computes our novel semantic-based metrics.
 
@@
 \begin{itemize}
-\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (in our case, predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
+\item $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of {\em true positives} (in our particular case, predicates that correctly appear in the action model) and $fp$ is the number of {\em false positives} (predicates of the learned model that should not appear).
 \item $Recall=\frac{tp}{tp+fn}$, where $fn$ is the number of {\em false negatives} (predicates that should appear in the learned model but are missing).
@@
 
-Introducing semantic-based evaluation metrics can be justified on several grounds:
+Introducing semantic-based evaluation metrics can be justified on two grounds:
 
-\begin{itemize}
-\item The GTM is unknown
-\item Though a planning reference model exists, a test-based model evaluation on a dataset is preferable or needed as complementary to a syntactic evaluation
-\item The semantics of an action model learned from observations of plan executions can be easily altered.
-That is, it is very likely to learn a semantically correct but syntactically incorrect model.
-We will refer to this problem as \emph{reformulation}.
-\end{itemize}
+\begin{enumerate}
+\item When the GTM is unknown. This is the most common scenario in ML, where models are both learned and evaluated with respect to datasets.
+\item When the GTM is known but a test-based evaluation on a dataset is preferable (or needed as complementary to a syntactic evaluation).
+As a rule of thumb, it is preferable to evaluate the learned models wrt a dataset because a learned model can be semantically correct though syntactically incorrect (different from the GTM).
+We refer to this phenomenon as \emph{model reformulation}.
+\end{enumerate}
 
-\emph{Reformulation} has a large impact in the action-model learning task.
-For instance, the roles of two {\em comparable} action models can be swapped.
+An example of \emph{model reformulation} is the swapping of the roles of two {\em comparable} action models.
 Two action models $\xi$ and $\xi'$ are comparable if both have the same parameters (iff $pars(\xi)=pars(\xi'$)) and so they share the same space of possible models.
@@
 
-Semantic alterations typically happen when the observed input data given in $\tau$ is scarce.
-Defining a proper semantic evaluation metric is key because the application of syntax-based metrics may report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the GTM.
+These semantic alterations typically appear in the learned models when the observed input data given in $\tau$ is scarce.
+Defining a proper semantic evaluation is key because the application of syntax-based metrics may report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the GTM.
 In the following sections, we introduce a novel evaluation metric that is robust to different types of reformulation.
@@
 
-We define semantic-based metrics which are conceptually grounded on the precision and recall metrics introduced above.
-The rationale behind these novel metrics lies in counting the number of \emph{editing operations} that are necessary to match two sets of action models.
-Given $\mathcal{M}$, the two allowed editing operations are:
+The \ARMS system showed that a semantic evaluation can be done via validation of a set of plan traces with the learned model~\cite{yang2007learning}.
+The underlying idea is that an error indication of the learned action models is obtained by counting the number of preconditions that are not satisfied during the execution of the plan trace with the learned models.
+This approach can be understood as modifying the plan trace (by adding the necessary preconditions to the intermediate states) so as to allow the execution of the observed actions using the learned models.
+In other words, modifying the plan trace to fit the model.
+Inspired by this approach, we present here an alternative evaluation that, instead, determines the modifications required by a learned model to explain the given plan traces.
+
+The rationale behind our novel metrics lies in counting the \emph{edit operations} that need to be applied in a set of action models $\mathcal{M}$ to fit the plan traces.
+Given a set of action models $\mathcal{M}$, the two allowed edit operations are:
 
@@
 
-%The \ARMS system showed that the error of the learned action models can also be estimated executing a set of plan traces with the learned model ~\cite{yang2007learning}.
-ARMS tackles the semantic evaluation by counting the number of preconditions not met during the execution of the plan trace with the learned models.
-
-The \ARMS system showed that a semantic evaluation can be done via validation of a set of plan traces with the learned model~\cite{yang2007learning}.
-The underlying idea is that an error indication of the learned action models is obtained by counting the number of preconditions that are not satisfied during the execution of the plan trace with the learned models.
-This approach can be understood as modifying the plan trace (by adding the necessary preconditions to the intermediate states) so as to allow the execution of the observed actions using the learned models.
-In other words, modifying the plan trace to fit the model.
-Inspired by this approach, we present an alternative method that, instead, modifies the learned models so that they can explain the given plan traces.
-
-%Inspired by this approach, we present an alternative method, in which instead of modifying the trace, we modify the learned models so they can explain the given plan traces.
-
-\vspace{0.1cm}
-
-We interpret the semantic evaluation of action models as a learning task $\Lambda = \tup{\mathcal{M}, \mathcal{T}}$, where:
-
-\begin{itemize}
-	\item $\mathcal{M}$ is a previously \textbf{learned set of action models} obtained using any learning approach such as \FAMA.
-
-	\item$\mathcal{T}$ is a set of plan traces  used for \textbf{testing}.
-\end{itemize}
-
-A solution to this task is an \textbf{edited set of action models} $\mathcal{M'}$ able to explain $\mathcal{T}$, and obtained by applying \emph{deletion} and \emph{insertion} operations to $\mathcal{M}$.
-While it is always recommended for the test set to be different from the one used during learning, this is specially important for satisfying approaches such as \FAMA; otherwise $\mathcal{M'}$ = $\mathcal{M}$, since $\mathcal{M}$ can already explain $\mathcal{T}$ without any modification.
-
-We now provide formal definitions of $INS(\mathcal{M},\mathcal{M'})$ and $DEL(\mathcal{M},\mathcal{M'})$, the insertions and deletions needed to transform the set of action models $\mathcal{M}$ into $\mathcal{M'}$.
-
-%We say that two action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M'}$ that maps every $\xi\in\mathcal{M}$ to a comparable action model $\xi'\in\mathcal{M'}$ and vice versa.
-
-%\begin{mydefinition}
-%  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} sets of action models. We compute $ins(\mathcal{M}, \mathcal{M'})$ and $del(\mathcal{M}, \mathcal{M'})$ as:
-%  \begin{small}
-%  	\begin{align*}
-%  	ins(\mathcal{M}, \mathcal{M'})=&\{pre_p(\xi) ~|~ p \notin pre(\xi), p \in pre(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')} \cup\\
-%  	&\{add_p(\xi) ~|~ p \notin add(\xi), p \in add(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')} \cup\\
-%  	&\{del_p(\xi) ~|~ p \notin del(\xi), p \in del(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')}\\
-%  	\\
-%  	del(\mathcal{M}, \mathcal{M'})=&\{pre_p(\xi) ~|~ p \in pre(\xi), p \notin pre(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')} \cup\\
-%  	&\{add_p(\xi) ~|~ p \in add(\xi), p \notin add(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')} \cup\\
-%  	&\{del_p(\xi) ~|~ p \in del(\xi), p \notin del(\xi')\}_{\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t. name(\xi) = name(\xi')}
-%  	\end{align*}
-%  \end{small}
-%\end{mydefinition}
+We now provide formal definitions of $INS(\mathcal{M},\mathcal{M'})$ and $DEL(\mathcal{M},\mathcal{M'})$, the sets of insertions and deletions, respectively, that are needed to transform a set of action models $\mathcal{M}$ into a new set of action models $\mathcal{M'}$.
 
@@
 
-Note that the number of {\em deletions} ($\left|DEL(\mathcal{M},\mathcal{M'})\right|$) that is required to transform $\mathcal{M}$ into $\mathcal{M}'$ matches our previous definition for the number of {\em false positives} when (1) $\mathcal{M}$ is the set of learned action models and (2) $\mathcal{M}'$ is the given GTM that serves as reference.
-Likewise the number of insertions ($\left|INS(\mathcal{M},\mathcal{M'})\right|$) required to transform $\mathcal{M}$ into $\mathcal{M}'$ corresponds to the number of {\em false negatives} in the learned models $\mathcal{M}$ with respect to the GTM $\mathcal{M}'$.
-%As explained, these numbers can be normalized with the amount of {\em true positives} to compute the {\em precision} and {\em recall} of a given learned model with respect to a reference model.
-Further, we can claim that all these numbers are bounded.
-Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well.
-In more detail, for an operator schema $\xi\in\mathcal{M}$ the maximum number of edits that can be introduced to their precondition set is $|F_{\xi}|$.
-This is the same max number of edits that can be introduced to the positive and to the negative effects of $\xi\in\mathcal{M}$.
+With these ingredients in mind, we first adapt the definitions of syntactic precision and recall to sets of action models.
+Let $\mathcal{M}$ be a set of learned action models and let $\mathcal{M'}$ be the GTM.
+We know that $size(\mathcal{M}) = \left|pre(\xi)\right| + \left|add(\xi)\right| + \left|del(\xi)\right| \forall \xi \in \mathcal{M}$ and by definition the number of preconditions and effects of the learned action models is equal to the sum of {\em true positives} and {\em false positives}; that is, $size(\mathcal{M}) = tp + fp$.
 
-%Inspired by this approach for the semantic evaluation of learned action models, we define now the semantic version of precision and recall to asses the quality of a learned action model with respect to a partially observed plan trace.
+The number of \emph{deletions} required to transform $\mathcal{M}$ into $\mathcal{M'}$ ($\left|DEL(\mathcal{M},\mathcal{M'})\right|$) matches our previous definition of the number of {\em false positives}; and $\left|INS(\mathcal{M},\mathcal{M'})\right|$, the number of \emph{insertions} required to transform $\mathcal{M}$ into $\mathcal{M'}$, corresponds to the number of {\em false negatives} of $\mathcal{M}$.
+Then we can affirm that $size(\mathcal{M'}) = size(\mathcal{M}) - \left|DEL(\mathcal{M},\mathcal{M'})\right| + \left|INS(\mathcal{M},\mathcal{M'})\right|$.
 
-Given that the evaluation task is defined in terms of a learning task $\Lambda$, there might exist potentially many edited models $\mathcal{M'}$ which are solution to this task.
+\begin{mydefinition} \label{syn-precision} The precision of $\mathcal{M}$ relative to the GTM is defined as the fraction of the common predicates and effects between $\mathcal{M}$ and the GTM among all predicates and effects of $\mathcal{M}$.
+This gives an intuitive measure of the soundness of $\mathcal{M}$.
+\begin{small}
+	\begin{align*}
+     Precision=\frac{tp}{tp+fp}=\frac{size(\mathcal{M})- \left|DEL(\mathcal{M},GTM)\right|}{size(\mathcal{M})}
+	\end{align*}
+\end{small}
+\end{mydefinition}
+
+\begin{mydefinition} \label{syn-recall} The recall of $\mathcal{M}$ relative to the GTM is defined as the fraction of the common predicates and effects between $\mathcal{M}$ and the GTM among all predicates and effects of the GTM.
+This gives an intuitive measure of the completeness of $\mathcal{M}$.
+
+\begin{small}
+	\begin{align*}
+     Recall= \frac{tp}{tp+fn}=
+\frac{size(\mathcal{M})- \left|DEL(\mathcal{M},GTM)\right|}{size(\mathcal{M}) - \left|DEL(\mathcal{M},GTM)\right| + \left|INS(\mathcal{M},GTM)\right|}
+	\end{align*}
+\end{small}
+\end{mydefinition}
+
+Definitions \ref{syn-precision} and \ref{syn-recall} are syntax-based metrics to evaluate $\mathcal{M}$ with respect to the GTM.
+A semantic metric, on the other hand, evaluates $\mathcal{M}$ with respect to a given set of plan traces.
+
+We interpret the semantic evaluation of action models as a learning task $\Lambda = \tup{\mathcal{M}, \mathcal{T}}$, where:
+
+\begin{itemize}
+	\item $\mathcal{M}$ is a \textbf{set of learned action models} obtained using any learning approach such as \FAMA; in general, $\mathcal{M}$ can be any given input set of action models even manually encoded.
+	\item$\mathcal{T}$ is a set of plan traces  used for \textbf{testing}.
+\end{itemize}
+
+A solution to this task is an \textbf{edited set of action models} $\mathcal{M'}$ such that (1) $\mathcal{M'}$ is obtained by exclusively applying a finite sequence of \emph{deletion} and \emph{insertion} operations to $\mathcal{M}$ and (2) $\mathcal{M'}$ explains $\mathcal{T}$; i.e.
+$\mathcal{M'}$ is {\em compliant} with every plan trace $\tau\in\mathcal{T}$.
+It is always recommended for the test set to be different from the one used during learning and this is specially important for satisfying approaches such as \FAMA; otherwise $\mathcal{M'}$ = $\mathcal{M}$ since $\mathcal{M}$ would be able to explain $\mathcal{T}$ without any modification.
+
+Since we are defining the semantic evaluation task in terms of a learning task $\Lambda$, there might exist potentially many edited models $\mathcal{M'}$ which are solution to this task.
 Although the actual GTM is included among the solution set, it is impossible to identify it, so we define the best solution based on its proximity to the input model.
@@
   Given a set of action models $\mathcal{M}$, and all the sets of action models $\mathcal{M'}$ able to explain the plan traces $\mathcal{T}$.
-The {\bf closest compliant set of action models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) and able to explain $\mathcal{T}$;
+The {\bf closest compliant set of action models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) that is able to explain $\mathcal{T}$;
   \[\mathcal{M^*}=\underset{\forall \mathcal{M}' \rightarrow \mathcal{T}}{\arg\min} \ \left| INS(\mathcal{M},\mathcal{M'}) \cup DEL(\mathcal{M},\mathcal{M'}) \right|\]
@@
 
-The {\bf closest compliant set of action models} allows us to define a semantic version of {\em precision} and {\em recall}, following the previous correspondences ({\em false positives} $\equiv$ {\em deletions} and {\em false negatives} $\equiv$ {\em insertions}).
-Also, by definition, the number of preconditions and effects of the learned action models is equal to the sum of {\em true positives} and {\em false positives}; that is, given $size(\mathcal{M}) = \left|pre(\xi)\right| + \left|add(\xi)\right| + \left|del(\xi)\right| \forall \xi \in \mathcal{M}$, we know that $size(\mathcal{M}) = tp + fp$.
-
-The semantic-based definition of precision and recall are given by:
+The closest compliant set of action models $\mathcal{M^*}$ allows us to define a semantic version of {\em precision} and {\em recall} following definitions \ref{syn-precision} and \ref{syn-recall}.
 
@@
 	\begin{align*}
-	sem\text{-}Precision=&\frac{tp}{tp + fp}=\frac{size(\mathcal{M})- \left|del(\mathcal{M},\mathcal{M^*})\right|}{size(\mathcal{M})}\\
-	sem\text{-}Recall=&\frac{tp}{tp + fn}=\frac{size(\mathcal{M})- \left|del(\mathcal{M},\mathcal{M^*})\right|}{size(\mathcal{M}) - \left|del(\mathcal{M},\mathcal{M^*})\right| + \left|ins(\mathcal{M},\mathcal{M^*})\right|}
+	sem\text{-}Precision=&\frac{size(\mathcal{M})- \left|DEL(\mathcal{M},\mathcal{M^*})\right|}{size(\mathcal{M})}\\
+    \vspace{0.5cm}
+	sem\text{-}Recall=&\frac{size(\mathcal{M})- \left|DEL(\mathcal{M},\mathcal{M^*})\right|}{size(\mathcal{M}) - \left|DEL(\mathcal{M},\mathcal{M^*})\right| + \left|INS(\mathcal{M},\mathcal{M^*})\right|}
 	\end{align*}
@@
 
-As we can observe in the two above formulas, precision and recall are not computed with respect to a GTM but with respect to the closest compliant set of action models that explain the test set of plan traces $\mathcal{T}$.
-Thereby, in the case that the {\bf closest compliant set of action models} $\mathcal{M^*}$ is the GTM, we conclude that the values of the syntactic measures {\em Precision} and {\em Recall} are exactly the same as {\em sem\text{-}Precision} and {\em sem\text{-}Recall}.
+\begin{myproposition}
+When the closest compliant set of action models $\mathcal{M^*}$ of an evaluation task $\Lambda = \tup{\mathcal{M}, \mathcal{T}}$ is the GTM, the syntactic and semantic evaluation of $\mathcal{M}$ return the same values; that is, $Precision=sem\text{-}Precision=$ and $Recall=sem\text{-}Recall$.
+\end{myproposition}
 
@@
 
-The compilation scheme presented in section \ref{sec:compilation} is extensible to address the evaluation task $\Lambda=\tup{\mathcal{M}, \mathcal{T}}$ defined in section \ref{semantic_precision_recall}.
+The compilation scheme presented in section \ref{compilation} is extensible to address the evaluation task $\Lambda=\tup{\mathcal{M}, \mathcal{T}}$ defined in section \ref{semantic_precision_recall}.
 In this extended task, $\mathcal{M}$ represents a set of previously learned action models; therefore, rather than learning the action models from scratch, we simply edit $\mathcal{M}$ until it satisfies the given test set of plan traces $\mathcal{T}$.
@@
 \item {\bf Edits the action models $\mathcal{M}$ to build $\mathcal{M}'$}.
-A solution plan starts with a prefix that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using the two {\em editing operations} defined above, {\em deletion} and {\em insertion}.
+A solution plan starts with a prefix that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
 \item {\bf Validates the edited model $\mathcal{M}'$ in the observed plan traces}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{sec:compilation} for the models that are programmed from scratch.
+The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\mathcal{T}$, as explained in Section~\ref{compilation} for the models that are programmed from scratch.
 \end{enumerate}
@@
 Assuming we are using an optimal planner to solve $P_{\Lambda}'$, the solution plan of this problem will induce the \emph{closest compliant set of action models} $\mathcal{M^*}$.
-Therefore, our compilation enables the computation of the semantic versions of \emph{precision} and \emph{recall}.
+Therefore, our compilation enables the straightforward computation of the semantic versions of \emph{precision} and \emph{recall}.
 An argument can be made, however, that solving optimally $P_{\Lambda}'$ may turn the evaluation process very time consuming.
 Considering this, $sem\text{-}Precision$ and $sem\text{-}Recall$ can be approximated if $P_{\Lambda}'$ is solved with a satisfying planner.
-In this case, no guarantees can be made that the edited models will be the closest compliant ones, but a planner will always try to minimize the length of the plan and hence the number of editing operations applied to the input models.
-
-%Our interest when solving the classical planning task that is output by our extended compilation is not in the resulting action model $\mathcal{M}'$ but in the number of {\em edit operations} (insertions and deletions) that is required to compute $\mathcal{M}'$.
-For instance, in the example of Figure~\ref{fig:plan-pdistance}, $\mathcal{M}'$ is validated in the given observations after inserting two new positive effects to the initial action model, e.g.
-$\delta(\mathcal{M},\mathcal{T})=2$ .
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_{\tt stack}|=|F_{\tt unstack}|=11$ while $|F_{\tt pickup}|=|F_{\tt putdown}|=5$  (as shown in Section~\ref{sec:learning}).
-The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
-
-%Last but not least, the compilation approach is also flexible to compute the {\em edit distance} between two {\em comparable} \strips\ action models, $\mathcal{M}$ and $\mathcal{M}'$.
-A solution to the planning task resulting from this compilation is a sequence of actions that edits the action model $\mathcal{M}$ to produce $\mathcal{M}'$ using to the two {\em edit operations}, deletion and insertion.
-In this case the edited model is not validated on a sequence of observations or plans but on the given action model $\mathcal{M}'$ that acts as a reference.
-The sets of fluents $F_{\Lambda}$ and $I_{\Lambda}$ are defined like in the previous compilation.
-With respect to the actions, $A_{\Lambda}'$ again implement the two {\em edit operations} ({\em insertion} and {\em deletion} for both preconditions and effects) but does not contain $\mathsf{apply_{\xi,\omega}}$ or $\mathsf{validate_{i}}$ actions because the edited \strips\ action model are not validated in any observation of plan executions.
-Finally, the goals are also different and are now defined by the set of fluents, $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ that represent all the operator schema $\xi\in\mathcal{M'}$, such that $f\in F_{\xi}$.
-To illustrate this, the plan of Figure~\ref{fig:plan-mdistance} solves the classical planning task that corresponds to computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.
-The plan edits first the {\tt\small stack} schema, {\em inserting} these two positive effects.
-Again our interest is in the number of required {\em edit operations}, e.g.
-$\delta(\mathcal{M},\mathcal{M'})=2$.
-
-%\begin{figure}[hbt!]
-%{\tt\small
-%00 : (insert\_add\_handempty\_stack)\\
-%01 : (insert\_add\_clear\_stack\_var1)
-%}
-% \caption{\small Plan for computing the distance between a \emph{blocksworld} action model, where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing, and the actual four-operator {\em blocksworld} model.}
-%\label{fig:plan-mdistance}
-%\end{figure}
+In this case, no guarantees can be made that the edited models will be the closest compliant ones, but a classical planner will always try to minimize the solution plan length and hence the number of edit operations applied to the input models.

### 2018-07-23 doc/aij18/experiments.tex

@@
 
-\item {\bf Machine}. All experiments are run on an Intel Core i5 3.50 GHz x 4 with 16 GB of RAM.
+\item {\bf Machine}. All experiments are run on an Intel Core i5 3.50 GHz x 4 with 16 GB of RAM, with a timeout of 1000 s after which the process is killed.
 
@@
 				Visitall & 1 & 3 & 2 & 2  \\
+				Zenotravel & 5 & 4 & 5 & 2
 			\end{tabular}
@@
 \begin{enumerate}
-	\item Identify the amount of input required by \FAMA to learn good models,
+	\item Identify the amount of input required by \FAMA to learn sound and complete models,
 	\item Evaluate the scalability of \FAMA wrt the input size which, as stated in section \ref{properties}, is one of the limiting factors of our approach.
@@
 	\caption{Evaluation of the impact of the input size on the quality of the learned models when learning from plan traces with \FO action sequences and \PO state trajectories with 10\% observability}
+	\label{fig:np_quality}
 \end{figure}
@@
 	\caption{Evaluation of the impact of the input size on the computation time when learning from plan traces with \FO action sequences and \PO state trajectories with 10\% observability}
+	\label{fig:np_time}
 \end{figure}
+
+Figures \ref{fig:np_quality} and \ref{fig:np_time} show the quality of the models and computation time, respectively, for the NP-complete scenario.
+In figure \ref{fig:np_quality} we see precision stabilizes at 0.86 after 5 traces while recall stabilizes at 0.95 after 3 traces.
+These results show that \FAMA does, in fact, not need big amounts of input to learn sound and complete models as opposite to other approaches in the literature where models are learned using around a hundred traces (see table \ref{table:models_comparison2}).
+
+With regards to the scalability of \FAMA, figure \ref{fig:np_time} paints a very interesting picture.
+In this figure, we can see an exponential increase in computation time for input sizes beyond 5 traces.
+Until an input size of 4 traces the computation time is below 1 s, but it reaches 153 s when the input is composed of 10 traces.
+These results match the performance of {\sc Madagascar}, since this planner is known to struggle with plan horizons beyond 160 steps (which corresponds to 8 traces in this case).
 
@@
 	\caption{Evaluation of the impact of the input size on the quality of the learned models when learning from plan traces with \NO action sequences and \NO state trajectories}
+	\label{fig:pspace_quality}
 \end{figure}
@@
 	\caption{Evaluation of the impact of the input size on the computation time when learning from plan traces with \NO action sequences and \NO state trajectories}
+	\label{fig:pspace_time}
 \end{figure}
+
+Next, we evaluate the PSPACE-complete scenario.
+As expected, the quality of the learned models, shown in figure \ref{fig:pspace_quality}, is inferior than in the NP-complete case.
+The results contradict the previous trend and we find a drop in quality as the input knowledge increases.
+This is, in fact, caused by an increasing number of timeouts in the experiments, meaning that no solution is found.
+The computation time in this scenario (figure \ref{fig:pspace_time}) is also considerably higher given the complexity of the task, which is reflected on the amount of timeouts.
+
+The conclusion we can draw is that learning from few input samples is both our strong point and our limitation.
+The NP-complete results show that the learned models are considerably sound and complete, although, to put these results into perspective, we will compare ourselves with another approach in the following section.
+The PSPACE-complete scenario, on the other hand, show that the learned models have less quality.
+Nevertheless, we argue that syntax-based metrics are not appropriate for the PSPACE-complete scenario due to the vast amount of reformulations that can appear in such under-constrained learning task and we will show proof of this in section \ref{minimal}.
 
@@
 Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of precision and recall.
-The horizontal axis represents the degree of observability, while the vertical axis show the average precision (figure \ref{fig:comparison_precision}) or recall (figure \ref{fig:comparison_recall}) computed over the 13 tested domains.
+The horizontal axis represents the degree of observability, while the vertical axis show the average precision (figure \ref{fig:comparison_precision}) or recall (figure \ref{fig:comparison_recall}) computed over the 14 tested domains.
+Remarkably, figure \ref{fig:comparison_precision} shows that \FAMA dominates in terms of precision in all cases save for \FO state trajectories.
+Furthermore, in average, the models learned by our approach are 15\% to 24\% more precise than those learned by \ARMS.
+A similar conclusion can be drawn for recall (figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our models are more complete.
+
+%Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
+The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
+Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
+Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\mathcal{M}$).
+{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
+
+The overall results indicate that \FAMA outperforms \ARMS in this particular experiment, which is one with very limited input.
+By no means do these results imply that our approach is better, just that it is able to learn better models when few plan traces are available.
+We are not able to perform experiments where far more input traces are available, since \FAMA cannot handle so much input, the same as \ARMS (or any other approach) cannot handle PSPACE-complete scenarios.
 
 \subsection{Experiments with minimal input knowledge}
+\label{minimal}
 In the previous experiments we have given a general view of the performance of \FAMA under different conditions.
@@
 				\hline
-				blocks & 0.6 & 0.67 & 0.29 & 0.22 & 0.75 & 0.33 & 0.55 & 0.41& 0.22 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'unstack', 2, 1), (u'unstack', u'stack', 2, 1)]
-				driverlog & 0.45 & 0.36 & 0.36 & 0.57 & 0.17 & 0.14 & 0.33 & 0.36& 1.59 \\ % [(u'load-truck', u'walk', 2, 1, 3), (u'unload-truck', u'disembark-truck', 1, 2, 3), (u'board-truck', u'unload-truck', 1, 2, 3), (u'disembark-truck', u'board-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
-				ferry & 0.6 & 0.43 & 0.4 & 0.5 & 0.5 & 0.5 & 0.5 & 0.48& 0.51 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
-				floor-tile & 0.67 & 0.45 & 0.6 & 0.55 & 0.88 & 0.64 & 0.71 & 0.55& 359.43 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'up', 1, 3, 2), (u'down', u'down', 1, 3, 2), (u'right', u'right', 1, 2, 3), (u'left', u'left', 1, 2, 3), (u'paint-up', u'paint-up', 1, 3, 2, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
-				grid & 0.37 & 0.41 & 0.5 & 0.71 & 0.33 & 0.43 & 0.4 & 0.52& 81.71 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'putdown', 1, 2), (u'putdown', u'pickup', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 3, 2), (u'unlock', u'unlock', 1, 2, 3, 4)]
-				gripper-strips & 0.8 & 0.67 & 0.6 & 0.75 & 0.8 & 1.0 & 0.73 & 0.81& 0.09 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
-				hanoi & 1.0 & 0.5 & 0.5 & 0.5 & 1.0 & 1.0 & 0.83 & 0.67& 0.87 \\ % [(u'move', u'move', 1, 3, 2)]
-				miconic & 0.67 & 0.22 & 0.67 & 0.5 & 1.0 & 0.67 & 0.78 & 0.46& 0.25 \\ % [(u'board', u'depart', 1, 2), (u'depart', u'board', 1, 2), (u'up', u'up', 2, 1), (u'down', u'down', 1, 2)]
-				n-puzzle & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 0.47 \\ % [(u'move', u'move', 1, 2, 3)]
-				parking & 0.64 & 0.5 & 0.63 & 0.56 & 0.5 & 0.44 & 0.59 & 0.5& 12.56 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 3, 2), (u'move-curb-to-car', u'move-curb-to-car', 3, 2, 1), (u'move-car-to-curb', u'move-car-to-curb', 2, 1, 3), (u'move-car-to-car', u'move-car-to-car', 1, 3, 2)]
-				satellite & 0.38 & 0.21 & 0.5 & 0.6 & 0.75 & 0.75 & 0.54 & 0.52& 1.86 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
-				transport & 0.43 & 0.3 & 0.4 & 0.4 & 0.0 & 0.0 & 0.28 & 0.23& 0.88 \\ % [(u'drive', u'drive', 1, 3, 2), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
-				grid-visit-all & 0.0 & 0.0 & 1.0 & 0.5 & 0.0 & 0.0 & 0.33 & 0.17& 1.54 \\ % [(u'move', u'move', 2, 1)]			
+				blocks & 0.86 & 0.67 & 0.86 & 0.67 & 0.8 & 0.44 & 0.84 & 0.59& 0.21 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
+				driverlog & 0.68 & 0.93 & 0.5 & 0.57 & 0.83 & 0.71 & 0.67 & 0.79& 0.27 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'unload-truck', 1, 2, 3), (u'board-truck', u'board-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
+				ferry & 0.78 & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0& 0.4 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
+				floor-tile & 0.69 & 1.0 & 0.45 & 0.91 & 1.0 & 0.82 & 0.65 & 0.93& 1.42 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'up', 1, 2, 3), (u'down', u'down', 1, 2, 3), (u'right', u'right', 1, 2, 3), (u'left', u'left', 1, 2, 3), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
+				grid & 0.76 & 0.94 & 0.55 & 0.86 & 1.0 & 1.0 & 0.74 & 0.94& 0.7 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'pickup', 1, 2), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
+				gripper-strips & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 0.16 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
+				hanoi & 0.67 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.73 & 1.0& 0.31 \\ % [(u'move', u'move', 1, 2, 3)]
+				miconic & 1.0 & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 0.8 & 1.0& 0.29 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 1, 2), (u'down', u'down', 1, 2)]
+				n-puzzle & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.88 & 1.0& 0.28 \\ % [(u'move', u'move', 1, 2, 3)]
+				parking & 0.73 & 0.79 & 0.7 & 0.78 & 1.0 & 0.78 & 0.78 & 0.78& 0.65 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-car-to-car', u'move-car-to-car', 1, 2, 3)]
+				satellite & 0.93 & 1.0 & 0.63 & 1.0 & 1.0 & 0.75 & 0.85 & 0.96& 0.36 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
+				transport & 1.0 & 1.0 & 0.71 & 1.0 & 1.0 & 0.6 & 0.9 & 0.9& 0.22 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
+				grid-visit-all & 0.67 & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 0.63 & 1.0& 0.87 \\ % [(u'move', u'move', 1, 2)]
+				zeno-travel & 0.92 & 0.79 & 0.6 & 0.86 & 0.8 & 0.57 & 0.78 & 0.75& 0.26 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 4, 5), (u'zoom', u'zoom', 1, 2, 3, 4, 5, 6)]
 				\hline
-				\bf & 0.51& 0.36 & 0.50 & 0.49 & 0.51 & 0.45 & 0.51 & 0.44 & 35.54 \\
+				\bf & 0.82 & 0.94 & 0.66 & 0.90 & 0.96 & 0.83 & 0.78 & 0.90 & 0.46 \\
 			\end{tabular}
@@
 		\end{center}
-	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \NO action sequences and \NO state trajectories}
-	\label{tab:results_minimum_0_0}
+	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \FO action sequences and \PO state trajectories with 10\% observability}
+	\label{tab:results_minimum_100_10}
 \end{table}
@@
 			\hline
-			blocks & 0.78 & 0.78 & 0.6 & 0.67 & 0.8 & 0.44 & 0.73 & 0.63& 0.27 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'unstack', 1, 2), (u'unstack', u'stack', 2, 1)]
-			driverlog & 0.5 & 0.07 & 0.18 & 0.57 & 0.0 & 0.0 & 0.23 & 0.21& 1.01 \\ % [(u'load-truck', u'walk', 2, 1, 3), (u'unload-truck', u'board-truck', 1, 2, 3), (u'board-truck', u'unload-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
-			ferry & 0.83 & 0.71 & 0.75 & 0.75 & 0.8 & 1.0 & 0.79 & 0.82& 0.37 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'debark', 1, 2), (u'debark', u'board', 1, 2)]
-			floor-tile & 0.68 & 0.59 & 0.62 & 0.73 & 0.75 & 0.55 & 0.68 & 0.62& 186.92 \\ % [(u'change-color', u'change-color', 1, 3, 2), (u'up', u'right', 1, 2, 3), (u'down', u'up', 1, 3, 2), (u'right', u'down', 1, 2, 3), (u'left', u'left', 1, 3, 2), (u'paint-up', u'paint-down', 1, 2, 3, 4), (u'paint-down', u'paint-up', 1, 2, 3, 4)]
-			grid & 0.46 & 0.35 & 0.25 & 0.43 & 0.63 & 0.71 & 0.45 & 0.5& 107.9 \\ % [(u'move', u'move', 2, 1), (u'pickup', u'pickup', 1, 2), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
-			gripper-strips & 0.57 & 0.67 & 0.67 & 0.5 & 1.0 & 0.75 & 0.75 & 0.64& 0.06 \\ % [(u'move', u'move', 1, 2), (u'pick', u'drop', 1, 2, 3), (u'drop', u'pick', 1, 2, 3)]
-			hanoi & 0.6 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.87 & 0.92& 1.09 \\ % [(u'move', u'move', 1, 3, 2)]
-			miconic & 1.0 & 0.11 & 0.33 & 0.5 & 1.0 & 0.33 & 0.78 & 0.31& 0.17 \\ % [(u'board', u'board', 1, 2), (u'depart', u'up', 1, 2), (u'up', u'depart', 1, 2), (u'down', u'up', 2, 1)]
-			n-puzzle & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 1.0& 2.5 \\ % [(u'move', u'move', 1, 2, 3)]
-			parking & 0.45 & 0.36 & 0.4 & 0.44 & 0.63 & 0.56 & 0.49 & 0.45& 5.87 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 3, 2), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-car', 2, 1, 3), (u'move-car-to-car', u'move-car-to-car', 3, 1, 2)]
-			satellite & 0.67 & 0.43 & 0.25 & 0.4 & 0.29 & 0.5 & 0.4 & 0.44& 1.34 \\ % [(u'switch-on', u'switch-off', 1, 2), (u'switch-off', u'switch-on', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
-			transport & 0.33 & 0.1 & 0.4 & 0.4 & 0.0 & 0.0 & 0.24 & 0.17& 1.14 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
-			grid-visit-all & 0.5 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 0.83 & 0.83& 1.08 \\ % [(u'move', u'move', 2, 1)]
-
+			blocks & 1.0 & 0.89 & 0.9 & 1.0 & 1.0 & 0.89 & 0.96 & 0.93& 3.51 \\ % [(u'pick-up', u'put-down', 1), (u'put-down', u'pick-up', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
+			driverlog & 0.3 & 0.21 & 0.31 & 0.57 & 0.29 & 0.29 & 0.3 & 0.32& 24.41 \\ % [(u'load-truck', u'unload-truck', 1, 2, 3), (u'unload-truck', u'load-truck', 1, 2, 3), (u'board-truck', u'board-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 2, 1, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
+			ferry & 0.83 & 0.71 & 0.8 & 1.0 & 1.0 & 1.0 & 0.87 & 0.87& 4.39 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
+			floor-tile & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1003.08 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'change-color', 1, 3, 2), (u'down', u'up', 1, 2, 3), (u'right', u'up', 1, 3, 2), (u'left', u'down', 1, 3, 2), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
+			grid & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1001.07 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'move', 2, 1), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
+			gripper-strips & 1.0 & 0.67 & 0.8 & 1.0 & 1.0 & 1.0 & 0.92 & 0.86& 1.31 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
+			hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.75& 1566.44 \\ % [(u'move', u'move', 1, 2, 3)]
+			miconic & 0.75 & 0.33 & 0.5 & 0.75 & 0.5 & 0.67 & 0.57 & 0.5& 0.86 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 2, 1), (u'down', u'down', 1, 2)]
+			n-puzzle & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 62.82 \\ % [(u'move', u'move', 1, 2, 3)]
+			parking & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1000.37 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-curb', 1, 3, 2), (u'move-car-to-curb', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-car', u'move-car-to-curb', 1, 2, 3)]
+			satellite & 0.78 & 0.5 & 0.57 & 0.8 & 0.33 & 0.25 & 0.63 & 0.52& 22.76 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
+			transport & 0.5 & 0.3 & 0.43 & 0.6 & 0.5 & 0.4 & 0.47 & 0.4& 31.07 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'pick-up', 1, 2, 3, 5, 4), (u'drop', u'drop', 1, 2, 3, 5, 4)]
+			grid-visit-all & 1.0 & 0.5 & 0.5 & 0.5 & 1.0 & 1.0 & 0.75 & 0.6& 3.11 \\ % [(u'move', u'move', 1, 2)]
+			zeno-travel & 0.67 & 0.29 & 0.75 & 0.43 & 0.75 & 0.43 & 0.71 & 0.36& 494.2 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 3, 2, 5, 4), (u'zoom', u'zoom', 1, 2, 3, 4, 5, 6)]
 			\hline
-			\bf & 0.62 & 0.49 & 0.57 & 0.65 & 0.68 & 0.60 & 0.63 & 0.58	& 23.82	 \\
+			\bf & 0.63 & 0.42 & 0.54 & 0.62 & 0.60 & 0.57 & 0.58 & 0.51 & 372.81
 		\end{tabular}
@@
 	\end{center}
-	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \NO action sequences and \PO state trajectories with 20\% observability}
-	\label{tab:results_minimum_0_20}
+	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \PO action sequences and \PO state trajectories with 30\% observability}
+	\label{tab:results_minimum_30_30}
 \end{table}
+
+\begin{table}[hbt!]
+	\begin{center}		
+		\begin{tabular}{l|l|l|l|l|l|l||l|l|l|}
+			& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c|}{\bf Global} & \\ \cline{2-9}			
+			& \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c||}{\bf R} &  \multicolumn{1}{|c|}{\bf P} & \multicolumn{1}{|c|}{\bf R} & {\bf Time} \\
+			\hline
+			blocks & 0.5 & 0.56 & 0.5 & 0.33 & 0.75 & 0.33 & 0.55 & 0.41& 0.27 \\ % [(u'pick-up', u'put-down', 1), (u'put-down', u'pick-up', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
+			driverlog & 0.13 & 0.07 & 0.38 & 0.71 & 0.0 & 0.0 & 0.25 & 0.21& 0.98 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'unload-truck', 1, 2, 3), (u'board-truck', u'disembark-truck', 1, 2, 3), (u'disembark-truck', u'board-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
+			ferry & 0.5 & 0.29 & 0.5 & 0.5 & 0.67 & 0.5 & 0.55 & 0.4& 0.47 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
+			floor-tile & 0.34 & 0.64 & 0.5 & 0.36 & 0.44 & 0.73 & 0.39 & 0.59& 165.92 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'up', 1, 3, 2), (u'down', u'left', 1, 2, 3), (u'right', u'right', 1, 3, 2), (u'left', u'left', 1, 3, 2), (u'paint-up', u'paint-up', 1, 3, 2, 4), (u'paint-down', u'paint-down', 1, 3, 2, 4)]
+			grid & 0.47 & 0.41 & 0.38 & 0.43 & 0.25 & 0.29 & 0.39 & 0.39& 214.87 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'pickup', 1, 2), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 3, 2), (u'unlock', u'unlock', 2, 1, 3, 4)]
+			gripper-strips & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93& 0.2 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
+			hanoi & 0.6 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.78 & 0.88& 3.23 \\ % [(u'move', u'move', 1, 3, 2)]
+			miconic & 0.4 & 0.44 & 0.6 & 0.75 & 0.25 & 0.33 & 0.42 & 0.5& 0.25 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 2, 1), (u'down', u'down', 2, 1)]
+			n-puzzle & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 4.15 \\ % [(u'move', u'move', 1, 2, 3)]
+			parking & 0.67 & 0.57 & 0.43 & 0.33 & 0.5 & 0.44 & 0.56 & 0.47& 7.61 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 3, 2), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-car-to-car', u'move-car-to-car', 2, 1, 3)]
+			satellite & 0.5 & 0.14 & 0.5 & 0.6 & 0.75 & 0.75 & 0.57 & 0.35& 2.06 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 3, 2), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
+			transport & 0.6 & 0.3 & 0.38 & 0.6 & 0.5 & 0.2 & 0.47 & 0.35& 0.83 \\ % [(u'drive', u'drive', 1, 3, 2), (u'pick-up', u'drop', 1, 2, 3, 4, 5), (u'drop', u'pick-up', 1, 2, 3, 4, 5)]
+			grid-visit-all & 0.0 & 0.0 & 1.0 & 0.5 & 0.0 & 0.0 & 1.0 & 0.2& 1.24 \\ % [(u'move', u'move', 2, 1)]
+			zeno-travel & 0.86 & 0.43 & 0.29 & 0.29 & 0.33 & 0.14 & 0.53 & 0.32& 28.4 \\ % [(u'board', u'debark', 1, 2, 3), (u'debark', u'board', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 4, 5), (u'zoom', u'zoom', 1, 2, 3, 5, 6, 4)]
+			\hline
+			\bf & 0.54 & 0.46 & 0.60 & 0.60 & 0.53 & 0.48 & 0.60 & 0.50 & 30.75
+		\end{tabular}
+	\end{center}
+	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \NO action sequences and \NO state trajectories}
+	\label{tab:results_minimum_0_0}
+\end{table}
+
+\begin{figure}[hbt!]
+	\begin{footnotesize}
+		\begin{verbatim}
+(:action stack
+ :parameters (?o1 - object ?o2 - object)
+ :precondition (and (on ?o1 ?o2)(handempty ))
+ :effect (and (not (on ?o1 ?o2))(clear ?o1)(clear ?o2)(ontable ?o1)))
+		\end{verbatim}
+	\end{footnotesize}
+	\caption{PDDL encoding of the learned action model of the {\em stack} operator from the four-operator {\em blocksworld} domain.}
+	\label{fig:macroaction}
+\end{figure}
 

### 2018-07-23 doc/aij18/learning_task.tex

@@
 
-Our approach to learning action models is to compile this learning task into a classical planning task with conditional effects.
 Conditional effects allow us to compactly define actions whose effects depend on the current state.
@@
 
-In section \ref{sec:motivation}, we exposed our reasons to solve the learning task $\Lambda=\tup{\mathcal{M},\tau}$ via compiling it into a classical planning task $P_{\Lambda}$.
+In section \ref{sec:motivation}, we exposed our reasons to solve the learning task via compiling it into a classical planning task.
 Our compilation scheme  builds upon the approach presented in \cite{aineto2018learning} but \FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.
 
+Our approach to solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to compile $\Lambda$ into a classical planning task $P_{\Lambda}$ with conditional effects.
 The intuition behind the \FAMA compilation is that a solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output set of action models $\mathcal{M}'$ that solves $\Lambda$.
@@
 Therefore, this plan programs and validates the action model for {\tt\small stack}, using the input plan trace of Figure~\ref{fig:example-plans}.
-Plan steps $00-01$ program the preconditions of the {\tt\small stack} model, steps $02-06$ program the action model effects and steps $07-11$ is the plan postfix that validates the programmed model following the trace of Figure~\ref{fig:example-plans}.
+Plan steps $00-01$ program the preconditions of the {\tt\small stack} model, steps $02-06$ program the action model effects and steps $07-11$ is the plan postfix that validates the programmed model following the plan trace of Figure~\ref{fig:example-plans}.
 

### 2018-07-23 doc/aij18/motivation.tex

@@
 \FAMA represents one step ahead towards learning action models without assuming observed actions.
-The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging (the actual length of the input plan trace becomes a priori unbound).
+The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging.
 This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now.
-Table \ref{tab:complex} shows when the worst case complexity of learning \strips\ action models becomes PSPACE-complete.
+Table \ref{tab:complex} shows that the worst case complexity of learning \strips\ action models becomes PSPACE-complete when combining \PO and \NO state trajectories and action sequences.
 
@@
 \vspace{0.1cm}
-Regarding the evaluation of the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2), normalizing this error by the the total number of all the possible preconditions and effects of an action model.
+Regarding the evaluation of the learned action models, we can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the the total number of all the possible preconditions and effects of an action model.
 This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM.
-The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects hence, it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models.
+The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects and thereby it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models.
 To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}.
@@
 Semantic evaluation metrics add a distinctive value over the syntactic ones, which is that they evaluate the learned model with a set of observations of plan executions and hence they are appropriate for scenarios where the GTM is not available.
-In this sense, \FAMA contributes also with a novel semantic-based error measure that builds upon the {\em precision} and {\em recall} metrics.
+In this sense, \FAMA also contributes with a novel semantic-based error measure that builds upon the {\em precision} and {\em recall} metrics.
 Unlike the semantic metric used in \ARMS~\cite{yang2007learning}, our semantic version of {\em precision} and {\em recall} is not sensitive to the repetition of one same flaw in the model evaluation.

### 2018-07-23 doc/aij18/experiments.tex
updates to experiments section

@@
 
-All tables in this section (tables \ref{tab:results_minimum_0_0}, \ref{tab:results_minimum_0_20}) follow the same structure.
+All tables in this section (tables \ref{tab:results_minimum_100_10}, \ref{tab:results_minimum_0_0}, \ref{tab:results_minimum_30_30}) follow the same structure.
 In these tables precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), and also globally ({\bf Global}).
 The last column reports the computation time in seconds needed to obtain the learned models.
-We can observe that identifying static predicates leads to models with better precondition {\em recall}.
-This fact evidences that many of the missing preconditions corresponded to static predicates because there is no incentive to learn them as they always hold~\cite{gregory2015domain}.
+
+In this experiment we evaluate three case studies, the first one being an NP-complete scenario and the other two being PSPACE-complete:
+\begin{itemize}
+	\item \textbf{\FO action sequence and \PO state trajectory}: We are, once again, assuming a degree of observability of 10\% for the state trajectory.
+The results of this case study are detailed in table \ref{tab:results_minimum_100_10}
+	\item  \textbf{\PO action sequence and \PO state trajectory}: In this case study we are assuming a degree of observability of 30\% for both the action sequence and state trajectory.
+The results of this case study are detailed in table \ref{tab:results_minimum_30_30}
+	\item  \textbf{\NO action sequence and \NO state trajectory}: Both the action sequence and state trajectory are completely empty and only initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
+The results of this case study are detailed in table \ref{tab:results_minimum_0_0}
+\end{itemize}
 
@@
 			ferry & 0.83 & 0.71 & 0.8 & 1.0 & 1.0 & 1.0 & 0.87 & 0.87& 4.39 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
-			floor-tile & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1003.08 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'change-color', 1, 3, 2), (u'down', u'up', 1, 2, 3), (u'right', u'up', 1, 3, 2), (u'left', u'down', 1, 3, 2), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
-			grid & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1001.07 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'move', 2, 1), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
+			floor-tile & - & - & - & - & - & - & - & - & - \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'change-color', 1, 3, 2), (u'down', u'up', 1, 2, 3), (u'right', u'up', 1, 3, 2), (u'left', u'down', 1, 3, 2), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
+			grid & - & - & - & - & - & - & - & - & - \\ % [(u'move', u'move', 1, 2), (u'pickup', u'move', 2, 1), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
 			gripper-strips & 1.0 & 0.67 & 0.8 & 1.0 & 1.0 & 1.0 & 0.92 & 0.86& 1.31 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
@@
 			n-puzzle & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 62.82 \\ % [(u'move', u'move', 1, 2, 3)]
-			parking & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1000.37 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-curb', 1, 3, 2), (u'move-car-to-curb', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-car', u'move-car-to-curb', 1, 2, 3)]
+			parking & - & - & - & - & - & - & - & - & - \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-curb', 1, 3, 2), (u'move-car-to-curb', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-car', u'move-car-to-curb', 1, 2, 3)]
 			satellite & 0.78 & 0.5 & 0.57 & 0.8 & 0.33 & 0.25 & 0.63 & 0.52& 22.76 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
@@
 			\hline
-			\bf & 0.63 & 0.42 & 0.54 & 0.62 & 0.60 & 0.57 & 0.58 & 0.51 & 372.81
+			\bf & 0.80 & 0.54 & 0.69 & 0.79 & 0.76 & 0.72 & 0.74 & 0.65 & 201.35
+
 		\end{tabular}
@@
 	\end{center}
-	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \PO action sequences and \PO state trajectories with 30\% observability}
+	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \PO action sequences and \PO state trajectories with 30\% observability in both cases}
 	\label{tab:results_minimum_30_30}
 \end{table}
+
+The first thing to notice in table \ref{tab:results_minimum_30_30}, which collects the result of the PSPACE case study with 30\% observability, is that the values for some domains are missing.
+This is the case for "floor-tile", "grid", and "parking" which not only are fairly complex domains, but are also considered \emph{puzzle-like}, a quality that is known to put a strain on the planners.
+Interestingly enough, "hanoi" also qualifies as a \emph{puzzle-like} domain, and we can see how this is reflected on the computation time.
+With regards to the quality of the models, we find that the learned models retrain a similar level of soundness, while their completeness has drastically worsened with respect to the previous case study.
+This is specially the case for the preconditions, which have gone from 0.94 recall to 0.54.
+This results are difficult to explain, given that in this case study the intermediate states contain more information, but our intuition is that ...
 
@@
 
+Next, we analyze the PSPACE case study with \NO action sequences and state trajectories (table \ref{tab:results_minimum_0_0}).
+Contrary to what might be expected by looking at the previous table, we are able to find solutions for all domains.
+This is because, although the search is less guided, there are far more possible solutions for this learning task.
+This wider range of solutions is also evidenced by the quality of the learned models, which in spite of being compliant, are more removed from the original GTM.
+In the table we find that both precision and recall have seen their scores decreased to 0.6 and 0.5, respectively.
+We argue, however, that syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the \emph{reformulations} that occur in these circumstances.
+
 \begin{figure}[hbt!]
@@
 \end{figure}
+
+In order to illustrate this, figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the stack operator learned from plan traces with \NO action sequences and state trajectories.
+This action model removes a block from on top of another block and puts it down on the table, all done in a single step.
+There are two main differences from the expected action model for stack: 1) it is unstacking a block instead of stacking it, and 2) instead of the arm holding the block by the end of the action, the block ends in the table.
+We refer to the first difference as \emph{role swapping} and it happens when actions are removed from the trace.
+If no actions are present in the input traces, the name of the actions lose their utility and they are effectively anonymous actions, meaning that they can interchange their behavior with any comparable action model.
+The second difference is commonly known as \emph{macro-action} (in this case unstack + put-down) and it can happen when there are missing states in the input traces.
+Reformulated action models are perfectly valid and can be used to solve planning tasks.
+For instance, any blocks-world problem can be solved by unstacking all the blocks to the table (unstack + put-down) and then stacking them to meet the goal condition (pick-up + stack), so reformulated action models like the one in the figure can be used to solve common planning tasks.
+
+As we have shown, the \NO/\NO case study meets all the criteria needed for reformulations to happen, and this is the reason why scenarios such as this are better evaluated using semantic-based metrics.
+
+\subsection{Syntactic versus semantic evaluation}
+\begin{table}[hbt!]
+	\begin{center}		
+		\begin{tabular}{l|c|c|c|c|}		
+			& {\bf Precision} & {\bf Recall} & {\bf sem-Precision} & {\bf sem-Recall} \\
+			\hline
+			blocks & 0.84 & 0.59 & 0.84 & 0.64 \\
+			driverlog & 0.67 & 0.79 & 0.70 & 0.92 \\
+			ferry & 0.71 & 1.0 & 1.0 & 1.0 \\
+			floor-tile & 0.65 & 0.93 & 0.95 & 0.95 \\
+			grid & 0.74 & 0.94 & 1.0 & 0.98 \\
+			gripper-strips & 1.0 & 1.0 & 1.0 & 1.0 \\
+			hanoi & 0.73 & 1.0 & 0.91 & 1.0 \\
+			miconic & 0.8 & 1.0	& 1.0 & 1.0 \\
+			n-puzzle & 0.88 & 1.0 & 1.0 & 1.0 \\
+			parking & 0.78 & 0.78 & 0.97 & 0.84 \\
+			satellite & 0.85 & 0.96 & 0.96 & 1.0 \\
+			transport & 0.9 & 0.9 & 1.0 & 1.0 \\
+			grid-visit-all & 0.63 & 1 & 1.0 & 1.0 \\
+			zeno-travel & 0.78 & 0.75 & 0.85 & 0.96 \\
+			\hline
+			& 0.78 & 0.9 & 0.94 & 0.95
+		\end{tabular}
+	\end{center}
+	\caption{\small Syntactic and semantic metric scores for learning tasks with \FO action sequences and \PO state trajectories with 10\% observability}
+	\label{tab:metric_comparison_100_10}
+\end{table}
+
+\begin{table}[hbt!]
+	\begin{center}		
+		\begin{tabular}{l|c|c|c|c|}		
+			& {\bf Precision} & {\bf Recall} & {\bf sem-Precision} & {\bf sem-Recall} \\
+			\hline
+			blocks & 0.55 & 0.41 & 0.9 & 0.86 \\
+			driverlog & 0.25 & 0.21 & 0.54 & 0.72 \\
+			ferry & 0.55 & 0.4 & 1.0 & 0.79 \\
+			floor-tile & 0.39 & 0.59 & - & - \\
+			grid & 0.39 & 0.39 & - & - \\
+			gripper-strips & 1.0 & 0.93 & 1.0 & 1.0 \\
+			hanoi & 0.78 & 0.88 & 0.89 & 1.0 \\
+			miconic & 0.42 & 0.5 & 0.89 & 0.85 \\
+			n-puzzle & 1.0 & 1.0 & 1.0 & 1.0 \\
+			parking & 0.56 & 0.47 & - & - \\
+			satellite & 0.57 & 0.35 & - & - \\
+			transport & 0.47 & 0.35 & 0.93 & 0.93 \\
+			grid-visit-all & 1.0 & 0.2 & 1.0 & 1.0 \\
+			zeno-travel & 0.53 & 0.32 & - & - \\
+			\hline
+			& 0.6 & 0.5 & 0.91 & 0.91
+		\end{tabular}
+	\end{center}
+	\caption{\small Syntactic and semantic metric scores for learning tasks with \NO action sequences and \NO state trajectories}
+	\label{tab:metric_comparison_0_0}
+\end{table}
 

### 2018-07-24 doc/aij18/experiments.tex
Completed experiments sections and added post-process description

@@
 \label{sec:experiments}
+
+In this section we present a set of experiments designed to evaluate different aspects of \FAMA, from its performance to the quality of the learned models.
+Whenever applicable, we will consider both NP-complete and PSPACE-complete scenarios, in order to draw conclusions at both levels of complexity.
+We will also devote one experiment to assess the suitability of the metrics proposed in this paper as evaluation metrics.
 
@@
 
-\item {\bf Machine}. All experiments are run on an Intel Core i5 3.50 GHz x 4 with 16 GB of RAM, with a timeout of 1000 s after which the process is killed.
+\item {\bf Machine}. All experiments are run on an Intel Core i5 3.50 GHz x 4 with 16 GB of RAM.
 
@@
 The experiment consists in increasing the size of $\mathcal{T}$ from 1 to 10, and analyze the evolution of the computation time as well as the quality of the learned models.
+To keep the experiment practicable, we introduce a 1000 s timeout after which the process is killed and the learned models are assigned a score of 0 for precision and recall.
 Our goals with this experiment are:
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{figures/input_size_100_10_precision.pdf}
+	\includegraphics[width=0.8\linewidth]{figures/input_size_100_10_precision.eps}
 	\caption{Evaluation of the impact of the input size on the quality of the learned models when learning from plan traces with \FO action sequences and \PO state trajectories with 10\% observability}
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{figures/input_size_100_10_time.pdf}
+	\includegraphics[width=0.8\linewidth]{figures/input_size_100_10_time.eps}
 	\caption{Evaluation of the impact of the input size on the computation time when learning from plan traces with \FO action sequences and \PO state trajectories with 10\% observability}
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{figures/input_size_0_0_precision.pdf}
+	\includegraphics[width=0.8\linewidth]{figures/input_size_0_0_precision.eps}
 	\caption{Evaluation of the impact of the input size on the quality of the learned models when learning from plan traces with \NO action sequences and \NO state trajectories}
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{figures/input_size_0_0_time.pdf}
+	\includegraphics[width=0.8\linewidth]{figures/input_size_0_0_time.eps}
 	\caption{Evaluation of the impact of the input size on the computation time when learning from plan traces with \NO action sequences and \NO state trajectories}
@@
 \ARMS works under the assumption of plan traces with \FO action sequences and \NO state trajectories, so we will also adopt this assumption.
-In more detail, we define a \emph{degree of observability} $\sigma$ for the state trajectory, ranging from 0\% to 100\%, that measures the probability of observing a literal, and evaluate both approaches as $\sigma$ increases.
+In more detail, we define a \emph{degree of observability} $\sigma$ for the state trajectory, ranging from 0\% to 100\%, that measures the probability of observing a literal, and evaluate both approaches as $\sigma$ increases using 5 plan traces as input.
 Note that when $\sigma = 0$ we have a \NO state trajectory and when $\sigma=100$ we have a \FO state trajectory, all cases in-between correspond to the \PO scenario.
@@
 	\centering
-	\includegraphics[width=.8\linewidth]{figures/comparison_precision.pdf}
+	\includegraphics[width=.8\linewidth]{figures/comparison_precision.eps}
 	\caption{Precision comparison between \FAMA and \ARMS}
@@
 	\centering
-	\includegraphics[width=.8\linewidth]{figures/comparison_recall.pdf}
+	\includegraphics[width=.8\linewidth]{figures/comparison_recall.eps}
 	\caption{Recall comparison between \FAMA and \ARMS}
@@
 
-All tables in this section (tables \ref{tab:results_minimum_100_10}, \ref{tab:results_minimum_0_0}, \ref{tab:results_minimum_30_30}) follow the same structure.
-In these tables precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), and also globally ({\bf Global}).
-The last column reports the computation time in seconds needed to obtain the learned models.
-
 In this experiment we evaluate three case studies, the first one being an NP-complete scenario and the other two being PSPACE-complete:
@@
 \end{itemize}
+
+All tables in this section (tables \ref{tab:results_minimum_100_10}, \ref{tab:results_minimum_0_0}, \ref{tab:results_minimum_30_30}) follow the same structure.
+In these tables precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), and also globally ({\bf Global}).
+The last column reports the computation time in seconds needed to obtain the learned models.
+Missing values in the tables correspond to domains where no solution was found given a timeout of 1800 s.
 
@@
 \end{table}
+
+Table \ref{tab:results_minimum_100_10} shows the results of the first case study.
+As was expected from the previous experiments, the recall scores are generally higher than the precision ones, and, in fact, the models learned for 6 of the tested domains were perfectly complete.
+Although precision is overall lower, it is interesting to notice that the soundness of most learned sets of positive effects is flawless.
+With regards to the computation time, we find times below the second in all cases save one, "floor-tile", the most complex among the tested domains.
 
@@
 Interestingly enough, "hanoi" also qualifies as a \emph{puzzle-like} domain, and we can see how this is reflected on the computation time.
-With regards to the quality of the models, we find that the learned models retrain a similar level of soundness, while their completeness has drastically worsened with respect to the previous case study.
+With regards to the quality of the models, we find that the learned models retain a similar level of soundness, while their completeness has drastically worsened with respect to the previous case study.
 This is specially the case for the preconditions, which have gone from 0.94 recall to 0.54.
-This results are difficult to explain, given that in this case study the intermediate states contain more information, but our intuition is that ...
+This results are difficult to explain, given that in this case study the intermediate states contain more information, but our intuition is that the sequence of actions plays a big role on the quality of the learned models or, at least, on how close they are to the GTM.
 
@@
 
-In order to illustrate this, figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the stack operator learned from plan traces with \NO action sequences and state trajectories.
+In order to illustrate this, figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the {\tt stack} operator learned from plan traces with \NO action sequences and state trajectories.
 This action model removes a block from on top of another block and puts it down on the table, all done in a single step.
-There are two main differences from the expected action model for stack: 1) it is unstacking a block instead of stacking it, and 2) instead of the arm holding the block by the end of the action, the block ends in the table.
+There are two main differences from the expected action model for {\tt stack}: 1) it is unstacking a block instead of stacking it, and 2) instead of the arm holding the block by the end of the action, the block ends in the table.
 We refer to the first difference as \emph{role swapping} and it happens when actions are removed from the trace.
@@
 \subsection{Syntactic versus semantic evaluation}
+
+In this experiment we compare the scores provided by the syntactic and semantic versions of precision and recall.
+For that purpose, we will evaluate the models learned in \ref{minimal} both syntactically, using the GTM, and semantically, computing the closest compliant set of action models with our approach.
+For the semantic evaluation we will be using 5 plan traces.
+Since we are using {\sc Madagascar}, a satisfying planner, the scores for sem-Precision and sem-Recall are approximated.
+
+Our goal with this experiment is to gauge the suitability of the semantic metrics we propose with respect to their well-known counterparts. With that in mind, we define two case studies:
+
+\begin{itemize}
+	\item \textbf{\FO action sequence and \PO state trajectory}: In this case study the full sequence of actions is known and no states are missing, which makes it practically impossible for reformulated models to appear.
+In fact, in all our experiments with \FAMA and other approaches we have never observed reformulations when the full sequence of actions is known.
+	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a case study that favors reformulations in the learned models, as we discussed previously.
+\end{itemize}
+
 \begin{table}[hbt!]
@@
 
+Table \ref{tab:metric_comparison_100_10} shows the results of the first case study.
+Looking at the high scores of the syntactic metrics, specially recall, we can conclude that the learned models are, in fact, fairly similar to the GTM.
+This supports our conclusion that no reformulation can occur in this case study, which also means that the space of possible solutions is restricted to models close to the GTM.
+The values for sem-Precision and sem-Recall are very high across the table, which is exactly the desired behavior for the metric given that the solutions are very close to the GTM.
+In comparison, both versions of recall show very similar scores, while the difference is bigger for the precision metrics.
+This means that the semantic version is more lenient towards extra preconditions or effects.
+This matches what we have observed during the experiments, where it is usual for the learned models to contain redundant or implicit preconditions which are penalized by the Precision metric.
+An example of this can be found in the {\tt move} action of the "hanoi" domain, where the learned action model specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
+
 \begin{table}[hbt!]
@@
 \end{table}
+
+Following, we take a look at table \ref{tab:metric_comparison_0_0} that details the results of the \NO/\NO case study, and we can see that using 5 traces in this PSPACE-complete scenario made it impossible to find a solution for some of the most demanding domains.
+Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger here.
+Comparing the scores of both versions, we find that models that originally achieved mediocre scores when using the GTM as reference, are considerably sound and complete, reaching and overall score of 0.91 for both sem-Precision and sem-Recall.
+
+Looking at the results of both case studies we can draw two conclusions with regards to the semantic metrics proposed in this paper.
+The first is that these metrics behave similarly to their syntactic counterparts, which means they are a good substitute when the GTM is not available.
+The second conclusion is that sem-Precision and sem-Recall are better suited to evaluate reformulated models than the original syntactic metrics, since they contemplate valid solutions outside the GTM.
 

### 2018-07-24 doc/aij18/learning_task.tex
Completed experiments sections and added post-process description

@@
 
+The logical inference process we base our approach on has trouble learning preconditions that are not, at the same time, negative effects, since no change can be observed between the pre-state and post-state of an action.
+This is specially the case for static predicates that never change and, hence, can only be preconditions.
+In order to address this shortcoming, we apply a post-process based on the one proposed in \cite{long2018building}.
+In their proposal, they obtain a full trace by applying the sequence of actions of the input trace and infer the preconditions from this fully observable trace.
+In our case, since the sequence of actions of the input trace might not be fully observable, we produce the traces by applying the actions found in the validation part of the solution plan.
+For instance, in the example of the figure \ref{fig:plan-lplan}, the sequence of actions used to produce the full trace would be \text{(unstack blockB blockA)}, \text{(put-down blockB)}, \text{(pick-up blockA)}, and \text{(stack blockA blockB)}.
+This post-process allows \FAMA to learn more complete lists of preconditions and deal with the always problematic static predicates.
+
 \subsection{Properties of the compilation}

### 2018-07-24 doc/aij18/experiments.tex
small updates added after reviewing section 6

@@
 Whenever applicable, we will consider both NP-complete and PSPACE-complete scenarios, in order to draw conclusions at both levels of complexity.
-We will also devote one experiment to assess the suitability of the metrics proposed in this paper as evaluation metrics.
+We will also devote one experiment to assess the suitability of the evaluation metrics proposed in this paper.
 
 \subsection{Setup}
-In order to asses the performance of our approach, we evaluate \FAMA on a wide range of domains.
-All tested domains are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-Table \ref{tab:domain_features} compiles the features of the thirteen domains used in the experiments.
+We evaluate \FAMA on fourteen IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2} and that are taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+Table \ref{tab:domain_features} compiles the features of these fourteen domains used in the experiments.
 From left to right, the columns report the name of the domain, number of actions, number of predicates, maximum arity of the actions, and maximum arity of the predicates.
-These are all features which affect the size of $P_\Lambda$ and, hence, have an impact on the complexity of the learning task.
-As can be seen in table \ref{tab:domain_features}, the selected domains range from simple ones, such as \emph{blocks} and \emph{visitall}, to complex ones, like \emph{floortile} and \emph{satellite}, with higher number of actions and predicates and with higher arity.
+These are all features which affect the size of $P_\Lambda$ and hence, have an impact on the complexity of the learning task.
+%As can be seen in table \ref{tab:domain_features}, the selected domains range from, such as \emph{blocks} and \emph{visitall}, to complex ones, like \emph{floortile} and \emph{satellite}, with higher number of actions and predicates and with higher arity.
 
-Following, we explain the details of our experimental setup:
+In the following, we explain the details of our experimental setup:
 
@@
 For each domain, we have generated 10 plan traces with 10 actions/intermediate states each using random walks.
-Depending on the experiment, these traces may be used for training or testing, so more details will be provided later.
+Depending on the experiment, these traces are used for training or testing, so more details will be provided later.
 
 \item {\bf Planner}.
-The classical planner we use to solve the instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use {\sc Madagascar} for two reasons: 1) its ability to deal with planning instances populated with dead-ends, and 2) when the length of the plan traces is known (\FO action sequences or \FO state trajectories), the horizon of the solution plan is also known and can be solved as a NP-complete problem.
+The classical planner we use to solve the $P_\Lambda$ instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
+We use {\sc Madagascar} for two reasons: 1) its ability to deal with planning instances populated with dead-ends, and 2) when the length of the plan traces is known (\FO action sequences or \FO state trajectories), the horizon of the solution plan is also known and can be provided to {\sc Madagascar}, addressing in this case NP-complete problems.
 This is because, SAT-based planners can apply the actions for programming preconditions in parallel during a single planning step (and the same for actions programming effects) because these actions do not interact.
@@
 				& \multicolumn{4}{c|}{Domain features}\\ \cline{2-5}
-				Domain & \# actions & \# predicates & max action arity & max predicate arity  \\
+				 & {\bf \# actions} & {\bf \# predicates} & {\bf max action arity} & {\bf max predicate arity}  \\
 				\hline
@@
 \subsection{Impact of the size of the input knowledge}
-In our first experiment, we evaluate how the size of the input knowledge, i.e., $\left|\mathcal{T}\right|$ affects \FAMA.
-The experiment consists in increasing the size of $\mathcal{T}$ from 1 to 10, and analyze the evolution of the computation time as well as the quality of the learned models.
-To keep the experiment practicable, we introduce a 1000 s timeout after which the process is killed and the learned models are assigned a score of 0 for precision and recall.
-Our goals with this experiment are:
+This experiment evaluates how the size of the input knowledge, i.e., $\left|\mathcal{T}\right|$ affects the performance of \FAMA.
+The experiment analyzes the evolution of the computation time as well as the quality of the learned models as the size of $\mathcal{T}$ increases (from 1 to 10).
+Recall that each trace in $\mathcal{T}$ comprises 10 actions/intermediate states.
+To keep the experiment practicable, we introduce a 1000s timeout after which the learning process is killed and the learned models are assigned a score of 0 for both precision and recall.
+Our goals with this experiment are to:
 \begin{enumerate}
 	\item Identify the amount of input required by \FAMA to learn sound and complete models,
-	\item Evaluate the scalability of \FAMA wrt the input size which, as stated in section \ref{properties}, is one of the limiting factors of our approach.
+	\item Evaluate the scalability of \FAMA wrt the input size which, as stated in section \ref{sec:properties}, is one of the limiting factors of our approach.
 \end{enumerate}
 
-With this in mind we have defined two case studies:
+With this in mind we defined two case studies:
 \begin{itemize}
-	\item \textbf{\FO action sequence and \PO state trajectory}: This is the typical case solved by most approaches, which corresponds to a NP-complete scenario.
-We are assuming a degree of observability of 10\% for the state trajectory, meaning that each literal of a state has a 10\% chance of being observed.
+	\item \textbf{\FO action sequence and \PO state trajectory}: This is the typical case addressed by most previous work, which corresponds to a NP-complete scenario.
+We are assuming a degree of observability of just 10\% for the state trajectory, meaning that each literal of a state has a 10\% chance of being observed.
 	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a PSPACE-complete scenario where both the action sequence and state trajectory are completely empty and only initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
@@
 Until an input size of 4 traces the computation time is below 1 s, but it reaches 153 s when the input is composed of 10 traces.
-These results match the performance of {\sc Madagascar}, since this planner is known to struggle with plan horizons beyond 160 steps (which corresponds to 8 traces in this case).
+These results match the performance of {\sc Madagascar}, since this planner is known to struggle with plan horizons beyond 150-200 steps (in our case 160 steps corresponds to 8 traces).
 
@@
 The conclusion we can draw is that learning from few input samples is both our strong point and our limitation.
-The NP-complete results show that the learned models are considerably sound and complete, although, to put these results into perspective, we will compare ourselves with another approach in the following section.
-The PSPACE-complete scenario, on the other hand, show that the learned models have less quality.
+Results on the NP-complete scenario show that the learned models are considerably sound and complete (to put these results into a richer perspective we will also compare ourselves with another approach in the following section).
+The PSPACE-complete scenario, on the other hand, shows that the learned models have less quality.
 Nevertheless, we argue that syntax-based metrics are not appropriate for the PSPACE-complete scenario due to the vast amount of reformulations that can appear in such under-constrained learning task and we will show proof of this in section \ref{minimal}.
@@
 \subsection{Comparison with \ARMS}
-Here we compare the performance of \FAMA to that of \ARMS, one of the most well-known approaches in the action model acquisition field.
-\ARMS works under the assumption of plan traces with \FO action sequences and \NO state trajectories, so we will also adopt this assumption.
-In more detail, we define a \emph{degree of observability} $\sigma$ for the state trajectory, ranging from 0\% to 100\%, that measures the probability of observing a literal, and evaluate both approaches as $\sigma$ increases using 5 plan traces as input.
-Note that when $\sigma = 0$ we have a \NO state trajectory and when $\sigma=100$ we have a \FO state trajectory, all cases in-between correspond to the \PO scenario.
+Here we compare the performance of \FAMA to that of \ARMS, one of the most well-known approaches learning planning models.
+\ARMS works under the assumption of plan traces with \FO action sequences and \NO state trajectories, so we will also adopt this assumption (\ARMS, as any other previous approaches for learning planning models cannot handle the PSPACE-complete scenarios).
+
+We define a \emph{degree of observability} $\sigma$ for the state trajectory, ranging from 0\% to 100\%, that measures the probability of observing a literal, and evaluate both approaches as $\sigma$ increases using 5 plan traces as input.
+Note that when $\sigma = 0$ we have a \NO state trajectory, when $\sigma=100$ we have a \FO state trajectory and all cases in-between correspond to the \PO scenario.
+Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of precision and recall.
+The horizontal axis represents the degree of observability, while the vertical axis show the average precision (figure \ref{fig:comparison_precision}) or recall (figure \ref{fig:comparison_recall}) computed over the fourteen tested domains.
+Remarkably, figure \ref{fig:comparison_precision} shows that \FAMA dominates in terms of precision in all cases save for \FO state trajectories.
+Furthermore, in average, the models learned by our approach are 15\% to 24\% more precise than those learned by \ARMS.
+A similar conclusion can be drawn for recall (figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our models are more complete.
 
@@
 
-Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of precision and recall.
-The horizontal axis represents the degree of observability, while the vertical axis show the average precision (figure \ref{fig:comparison_precision}) or recall (figure \ref{fig:comparison_recall}) computed over the 14 tested domains.
-Remarkably, figure \ref{fig:comparison_precision} shows that \FAMA dominates in terms of precision in all cases save for \FO state trajectories.
-Furthermore, in average, the models learned by our approach are 15\% to 24\% more precise than those learned by \ARMS.
-A similar conclusion can be drawn for recall (figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our models are more complete.
-
 %Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
@@
 
-The overall results indicate that \FAMA outperforms \ARMS in this particular experiment, which is one with very limited input.
-By no means do these results imply that our approach is better, just that it is able to learn better models when few plan traces are available.
-We are not able to perform experiments where far more input traces are available, since \FAMA cannot handle so much input, the same as \ARMS (or any other approach) cannot handle PSPACE-complete scenarios.
+\FAMA outperforms \ARMS in this particular experiment, which is one with very limited input.
+By no means do these results imply that our approach is overall better for the NP-complete scenarios, just that it is able to learn better models when few plan traces are available.
+We are not reporting experiments where far more input traces are available, since \FAMA cannot handle so much input.
 
@@
 \label{minimal}
-In the previous experiments we have given a general view of the performance of \FAMA under different conditions.
+%In the previous experiments we have given a general view of the performance of \FAMA under different conditions.
 So far, experiments have shown that \FAMA is able to learn with very small amounts of input knowledge, be it due to low observability or few training samples.
@@
 Although precision is overall lower, it is interesting to notice that the soundness of most learned sets of positive effects is flawless.
-With regards to the computation time, we find times below the second in all cases save one, "floor-tile", the most complex among the tested domains.
+With regards to the computation time, we find times below the second in all cases except for one, "floor-tile", the most complex among the tested domains.
 
@@
 The first thing to notice in table \ref{tab:results_minimum_30_30}, which collects the result of the PSPACE case study with 30\% observability, is that the values for some domains are missing.
-This is the case for "floor-tile", "grid", and "parking" which not only are fairly complex domains, but are also considered \emph{puzzle-like}, a quality that is known to put a strain on the planners.
+This is the case for "floor-tile", and "parking" which not only are fairly complex domains, but are also considered \emph{puzzle-like}, a quality that is known to put a strain on the planners.
 Interestingly enough, "hanoi" also qualifies as a \emph{puzzle-like} domain, and we can see how this is reflected on the computation time.
@@
 This is specially the case for the preconditions, which have gone from 0.94 recall to 0.54.
-This results are difficult to explain, given that in this case study the intermediate states contain more information, but our intuition is that the sequence of actions plays a big role on the quality of the learned models or, at least, on how close they are to the GTM.
+%This results are difficult to explain, given that in this case study the intermediate states contain more information, but
+Our intuition is that the input actions play a big role on how close the learned model are with respect to the GTM.
+
+Next, we analyze the PSPACE case study with \NO action sequences and state trajectories (Table \ref{tab:results_minimum_0_0}).
+Contrary to what might be expected by looking at the previous table, we are able to find solutions for all domains.
+This is because, although the search is less guided, there are far more possible solutions for this learning task.
+This wider range of solutions is also evidenced by the quality of the learned models, which in spite of being compliant with the input data, are further from the original GTM.
+In the table we find that both precision and recall have seen their scores decreased to 0.6 and 0.5, respectively.
 
@@
 
-Next, we analyze the PSPACE case study with \NO action sequences and state trajectories (table \ref{tab:results_minimum_0_0}).
-Contrary to what might be expected by looking at the previous table, we are able to find solutions for all domains.
-This is because, although the search is less guided, there are far more possible solutions for this learning task.
-This wider range of solutions is also evidenced by the quality of the learned models, which in spite of being compliant, are more removed from the original GTM.
-In the table we find that both precision and recall have seen their scores decreased to 0.6 and 0.5, respectively.
-We argue, however, that syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the \emph{reformulations} that occur in these circumstances.
+We argue, however, that syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the \emph{reformulations} that frequently occur in these circumstances.
+To illustrate this, Figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the {\tt stack} operator learned from plan traces with \NO action sequences and state trajectories.
 
@@
 
-In order to illustrate this, figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the {\tt stack} operator learned from plan traces with \NO action sequences and state trajectories.
 This action model removes a block from on top of another block and puts it down on the table, all done in a single step.
-There are two main differences from the expected action model for {\tt stack}: 1) it is unstacking a block instead of stacking it, and 2) instead of the arm holding the block by the end of the action, the block ends in the table.
+There are two main differences from the expected action model for {\tt stack}: 1) it is unstacking a block instead of stacking it, and 2) instead of the arm holding the block by the end of the action, the block ends on the table.
 We refer to the first difference as \emph{role swapping} and it happens when actions are removed from the trace.
@@
 For instance, any blocks-world problem can be solved by unstacking all the blocks to the table (unstack + put-down) and then stacking them to meet the goal condition (pick-up + stack), so reformulated action models like the one in the figure can be used to solve common planning tasks.
-
 As we have shown, the \NO/\NO case study meets all the criteria needed for reformulations to happen, and this is the reason why scenarios such as this are better evaluated using semantic-based metrics.
@@
 
+Next, we take a look at table \ref{tab:metric_comparison_0_0} that details the results of the \NO/\NO case study, and we can see that using 5 traces in this PSPACE-complete scenario made it impossible to find a solution for some of the most demanding domains.
+Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger here.
+Comparing the scores of both versions, we find that models that originally achieved mediocre scores when using the GTM as reference, are considerably sound and complete, reaching and overall score of 0.91 for both sem-Precision and sem-Recall.
+
 \begin{table}[hbt!]
@@
 \end{table}
-
-Following, we take a look at table \ref{tab:metric_comparison_0_0} that details the results of the \NO/\NO case study, and we can see that using 5 traces in this PSPACE-complete scenario made it impossible to find a solution for some of the most demanding domains.
-Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger here.
-Comparing the scores of both versions, we find that models that originally achieved mediocre scores when using the GTM as reference, are considerably sound and complete, reaching and overall score of 0.91 for both sem-Precision and sem-Recall.
 

### 2018-07-24 doc/aij18/aij18.tex
Add files via upload
Introduction done

@@
 \newtheorem{mydefinition}[mytheorem]{Definition}
+\newtheorem{myproposition}[mytheorem]{Proposition}
 \newtheorem{myconstruction}{Construction}
@@
 \begin{abstract}
-  This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning.
+  \textcolor[rgb]{1.00,0.00,0.00}{This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning.
 The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state) to sequences of state observations or even just a set of initial and final states (where no intermediate action or state is known).
@@
 This extension allows us to evaluate the quality of the learned models with respect to the actual models but also, with respect to a {\em test set} of observations of plan executions.
-The performance of our compilation approach is evaluated learning action models, for a wide range of classical planning domains from the International Planning Competition (IPC), and following these two evaluation approaches.
+The performance of our compilation approach is evaluated learning action models, for a wide range of classical planning domains from the International Planning Competition (IPC), and following these two evaluation approaches.}
 \end{abstract}
@@
 \begin{keyword}
-Classical planning\sep Learning action models\sep Generalized planning
+Action model learning\sep AI planning\sep Machine Learning
 %% keywords here, in the form: keyword \sep keyword

### 2018-07-24 doc/aij18/evaluation.tex
Introduction done

@@
 The \ARMS system showed that a semantic evaluation can be done via validation of a set of plan traces with the learned model~\cite{yang2007learning}.
-The underlying idea is that an error indication of the learned action models is obtained by counting the number of preconditions that are not satisfied during the execution of the plan trace with the learned models.
+The underlying idea is that an error indication of the learned action models is obtained by counting the number of preconditions that are not satisfied during the execution of the plan trace with the learned models, similarly to the functionality provided by the automatic validation tool VAL~\cite{howey2004val} used in the IPCs.
 This approach can be understood as modifying the plan trace (by adding the necessary preconditions to the intermediate states) so as to allow the execution of the observed actions using the learned models.

### 2018-07-24 doc/aij18/introduction.tex
Introduction done

--- 2018-06-11 doc/aij18/introduction.tex
@@
 
-Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-At both planning tasks, automated planners reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
-Unfortunately, modeling planning actions is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
+There is common agreement in the planning community that the unavailability of a complete domain model is a bottleneck in the applicability of planning technology to many real-world domains~\cite{kambhampati:modellite:AAAI2007}.
+Motivated by the difficulty and cost of crafting action models, research in action-model learning has seen huge advances since the emergence of the pioneer learning system ARMS~\cite{yang2007learning}.
+Ever since we have seen models able to learn action models with quantifiers~\cite{AmirC08,ZhuoYHL10}, from noisy actions or noisy states~\cite{zhuo2013action,MouraoZPS12}, from null state information~\cite{cresswell2013acquiring}, from incomplete domain models~\cite{ZhuoNK13,ZhuoK17} and many more.
 
-Machine Learning (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
-The application of inductive ML to the learning of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, is not straightforward though:
+The primary underlying motivation for acquiring planning action models is to be able to solve model-based planning tasks afterwards.
+A learning system receives as an input observations of the agent's plan execution and outputs an abstract version of the capability model that reflects the physics of the real-world domain.
+Recently, model-based representations are also becoming popular in explainable AI planning as they form a common basis for communicating with users and facilitate the generation of transparent and explainable decisions~\cite{FoxLM17} as well as explanations in terms of the differences with a human mental model~\cite{ChakrabortiSK18}.
+
+Two different types of data sources are generally identified in the generation of explanations~\cite{Sheh17,Roberts18}: (1) data that embody enough information to fully trace the steps of the decision-making process of the agent and so accurate and detailed explanations can be generated; (2) data that do not accurately represent the decision-making process because the observed external behaviour of the agent responds to a black-box model.
+A similar distinction can be established with the type of observations that algorithms use to learn planning action models.
+Up to date, all existing learning approaches assume that a given input plan trace contains a fully observed sequence of the executed actions.
+This seriously restricts the applicability of the learning approach to contexts in which the behaviour of the agent is fully observable and a human annotator correctly labels the observed actions.
+
+In this paper we claim that a planning model is learnable even though an accurate representation of the agent's behaviour is not available.
+Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under a minimum input knowledge.
+While current learning systems accept input plan traces to a greater or lesser degree of observability in the states traversed by the sequence of observed actions, \FAMA extends the degree of observability to the action sequence.
+This way, we can define a percentage of observability of the actions up to a null observability.
+The minimal observability case acceptable by \FAMA is when the algorithm is only fed with the input and output of a plan trace, namely the initial and final state of a planning task.
+Like many Machine Learning (ML) techniques, \FAMA is able to operate with only correct input/ouput pairs and an unknown or a partially known model of the agent.
+Unlike ML algorithms, \FAMA requires a symbolic structured representation of the input knowledge.
+In this sense, recent investigations tackle the problem of learning symbolic representations from low-level sensing information and unstructured data~\cite{KonidarisKL18,AsaiF18}.
+
+\FAMA is a solving schema, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}.
+\FAMA is thereby a model-based approach that automatically builds its own planning model by logical inference from the input observations.
+A solution to this planning task is a sequence of actions that determines the preconditions and effects of the \strips\ action models.
+The construct of the planning model starts out with the input observations of the agent's behaviour that, as commented above, may comprise none of the actions executed by the agent.
+\FAMA can thus be regarded as a solving process that generates an instance of a planning model, and whose solution must be \emph{compliant} with the input plan traces.
+This behaviour largely differs from ML techniques, which aim to minimize an error function on the training data.
+Moreover, \FAMA requires far less sample data (input observations) than ML algorithms, thus alleviating the dependency on the assumption that there are enough data for learning the action models~\cite{Zhuo15}.
+
+Like other learning approaches, \FAMA also accepts as an input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}.
+In this case, the output model of \FAMA is compliant with both the input model and with the observations.
+Following this behaviour, we can define a distance measure between the two models which yields an assessment of the accuracy with which the input model explains the observations.
+In other words, the distance measure provides a mechanism to explicitly correct or adjust a model to the observations.
+By using this distance-based concept, \FAMA contributes with two novel semantic evaluation metrics that are generally more informative than counting the number of errors between two models.
+Moreover, this new evaluation method opens up a way towards model reconciliation~\cite{KulkarniCZVZK16}.
+
+All in all, \FAMA is a planning-based solving scheme that outputs a \strips\ action model using an automatically built planning model from minimal input knowledge.
+Unlike extensive data-ML approaches, \FAMA only requires a small amount of input plan traces.
+Unlike most relevant action-model learning algorithms, \FAMA does not require the traces to contain any observed action executed by the agent.
+Ultimately, \FAMA is a model-based planning approach that adopts certain ML-like characteristics, like the ability to learn only from data that do not explicitly exhibit the agent¡s behaviour or the capacity of correcting an initially given action model.
+
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper presents \FAMA, an innovative approach for learning \strips\ action models defined as a classical planning compilation.
+A first description of the compilation scheme previously appeared in our previous conference paper~\cite{aineto2018learning}.
+\FAMA brings the following contributions over the first version of the compilation:
+
 \begin{itemize}
-\item The {\em input} to ML algorithms (the {\em learning/training} data) usually are finite vectors encoding the value of fixed features in a given set of objects.
-The input for learning planning action models are observations of plan executions (where each observed plan possibly has a different number of steps and involves a different number of objects).
-\item The {\em output} of ML algorithms usually is a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning \strips\ action models the output is, for each action, the sets of {\em preconditions}, {\em negative} and {\em positive effects} that define the possible state transitions.
+\item A unified formulation for learning and evaluating action models from observations of plan executions.
+In the case of minimal observability, these execution only comprise the initial and final state of the plan traces.
+\item A thorough elaboration of two semantic evaluation metrics that build upon two well-known syntax-based ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate the output action models with respect to observations of plan executions.
+\item An exhaustive empirical evaluation over 14 domains from the International Planning Competitions (IPCs).
+We include an analysis of the impact of the size of the input knowledge in the accuracy and performance of \FAMA, a comparison with \ARMS and a detailed experimentation when \FAMA is executed with minimal input knowledge.
 \end{itemize}
 
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper introduces an innovative approach for learning \strips\ action models that can be defined as a classical planning compilation.
-A solution to the classical planning task that results from our compilation is a sequence of actions that determines the preconditions and effects of a \strips\ model such that this model satisfies the plan executions given as input.
-
-The compilation approach is appealing by itself because it leverages off-the-shelf planners and because its practicality allow us to report learning results over a wide range of domains from the {\em International Planning Competition} (IPC).
-Moreover, it opens up a way towards \emph{bootstrapping} planning action models, enabling a planner to gradually learn/update its action model.
-Apart from these, our compilation exhibits the following contributions:
-\begin{enumerate}
-\item {\bf\em Input flexibility}.
-Our classical planning compilation is flexible to various amount and kind of available input knowledge.
-The action model to learn can be partially specified and the learning examples can range from a set of plans (with their corresponding initial state) or state observations, to just a set of initial and final states where no intermediate action or state is observed.
-% Also accepts background knowledge
-
-\item {\bf\em Model validation}.
-The compilation poses a novel framework to assess the validation of a STRIPS model with respect to plan executions.
-This validation capacity goes beyond the functionality of VAL~\cite{howey2004val} since it does require neither a full action model nor a fully observed plan to determine validation.
-
-\item {\bf\em Model evaluation}.
-The compilation can assess how well a given \strips\ action model matches a plan execution, which allows us to assess learning performance without knowing the actual action model.
-The idea is to assess the amount of edition that is required by the input action model to induce the given observations of plan executions.
-\end{enumerate}
-
-A first description of the compilation previously appeared in our previous conference paper~\cite{aineto2018learning}. Compared to that paper, this work includes the following novel material:
-\begin{itemize}
-\item A unified formulation for learning and evaluating \strips\ action models from observed executions of plans. Further, these executions can only comprise state observations.
-\item A redefinition of the ML metrics {\em precision} and {\em recall} to evaluate \strips\ action models with respect to observations of plan executions.
-%\item Leveraging diverse forms of {\em background knowledge} for learning and evaluating \strips\ action models.
-\item A complete empirical evaluation of the compilation approach for learning of \strips\ action models.
-Our evaluation analyses how input knowledge affects to the performance of the compilation approach when learning and evaluating \strips\ action models.
-\end{itemize}
-
-Section~\ref{sec:background} introduces classical planning and reviews related work on learning planning action models.
-Section~\ref{sec:motivation} motivates our compilation approach for learning of \strips\ action models from observations of plan executions.
-Section~\ref{sec:learning} introduces the \strips\ action model (the target of our learning and evaluation tasks) and formalizes the learning of \strips\ action models with regard to different amount and kind of available input knowledge.
-Sections~\ref{sec:Section5} and ~\ref{sec:Section6} describe our compilation approach for addressing the formalized learning tasks its extension to evaluate and recognize \strips\ action models.
-Section~\ref{sec:Section7} reports the data collected in a two-fold empirical evaluation of our learning approach: First, the learned \strips\ action models are tested with observations of plan executions and second, the learned models are compared to the actual models.
-Finally, Section ~\ref{sec:Section8} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.
+The paper is organized as follows.
+Section~\ref{sec:background} introduces classical planning concepts and reviews related work on learning planning action models.
+Section~\ref{sec:motivation} motivates our compilation-to-planning approach for learning action models.
+Section~\ref{sec:learning} formalizes the learning task and presents the compilation scheme, the core of \FAMA.
+Sections~\ref{sec:evaluation} explains the evaluation of a learned model with respect to a reference model (syntactic evaluation) and with respect to a set of plan traces (semantic evaluation).
+Section~\ref{sec:experiments} reports the results of the experimental evaluation and, finally, Section~\ref{sec:conclusions} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.

### 2018-07-26 doc/aij18/introduction.tex
Add files via upload
Updates Introduction

@@
 Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under a minimum input knowledge.
-While current learning systems accept input plan traces to a greater or lesser degree of observability in the states traversed by the sequence of observed actions, \FAMA extends the degree of observability to the action sequence.
+While current learning systems accept varying degrees of observability in the states traversed by the input observed actions, \FAMA extends partial observability to the action sequence.
 This way, we can define a percentage of observability of the actions up to a null observability.
@@
 
-Like other learning approaches, \FAMA also accepts as an input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}.
+A key aspect in action-model learning is the evaluation method to assess the quality and performance of the learning approach.
+The most common method is to use a syntax-based evaluation that compares the learned model with a reference model.
+\FAMA instead proposes two novel semantic evaluation metrics that build upon two well-known syntactic ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate the output action models with respect to observations of plan executions.
+Our semantic evaluation is generally more informative that counting the number of error between two models and alleviates the limitations of a purely syntax-based assessment: (a) that the learned model is syntactically different from the reference model but semantically correct and (b) that the learned model comprises correct though unnecessary preconditions in regards to the reference model.
+This latter issue is concerned with the qualification problem, the actual impossibility of listing all the preconditions required for a real world action to have its intended effects~\cite{GinsbergS88}.
+
+The semantic evaluation in \FAMA is carried out using the same compilation scheme for solving a learning task.
+In particular, \FAMA also accepts as an input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}, alongside the observations.
+In this case, \FAMA returns a model that follows the input model and is compliant with the observations.
+Following this behaviour, we design an adjustment mechanism that serves to correct the input model to the output model, which in turn defines an assessment of the accuracy with which the input model explains the observations.
+Interpreting the adjustment value as a distance-based concept between two models could be exploited in model reconciliation~\cite{KulkarniCZVZK16}.
+
+%Like other learning approaches, \FAMA also accepts as an input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}.
 In this case, the output model of \FAMA is compliant with both the input model and with the observations.
@@
 Unlike most relevant action-model learning algorithms, \FAMA does not require the traces to contain any observed action executed by the agent.
-Ultimately, \FAMA is a model-based planning approach that adopts certain ML-like characteristics, like the ability to learn only from data that do not explicitly exhibit the agent¡s behaviour or the capacity of correcting an initially given action model.
+Ultimately, \FAMA adopts certain ML-like characteristics, like the ability to learn from data that do not explicitly exhibit the agent¡s behaviour or the capacity of correcting an initially given action model.
 
@@
 \item A unified formulation for learning and evaluating action models from observations of plan executions.
-In the case of minimal observability, these execution only comprise the initial and final state of the plan traces.
-\item A thorough elaboration of two semantic evaluation metrics that build upon two well-known syntax-based ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate the output action models with respect to observations of plan executions.
+In the case of minimal observability, these executions only comprise the initial and final state of the plan traces.
+\item A thorough elaboration of two semantic evaluation metrics that build upon the notions of {\em precision} and {\em recall} to evaluate the output action models with respect to observations of plan executions.
 \item An exhaustive empirical evaluation over 14 domains from the International Planning Competitions (IPCs).
-We include an analysis of the impact of the size of the input knowledge in the accuracy and performance of \FAMA, a comparison with \ARMS and a detailed experimentation when \FAMA is executed with minimal input knowledge.
+We include an analysis of the impact that the size of the input knowledge has in the performance of \FAMA, a comparison with \ARMS and a detailed experimentation when \FAMA is executed with minimal input knowledge.
 \end{itemize}

### 2018-07-26 doc/aij18/introduction.tex
minor changes to the introduction

@@
 Motivated by the difficulty and cost of crafting action models, research in action-model learning has seen huge advances since the emergence of the pioneer learning system ARMS~\cite{yang2007learning}.
-Ever since we have seen models able to learn action models with quantifiers~\cite{AmirC08,ZhuoYHL10}, from noisy actions or noisy states~\cite{zhuo2013action,MouraoZPS12}, from null state information~\cite{cresswell2013acquiring}, from incomplete domain models~\cite{ZhuoNK13,ZhuoK17} and many more.
+Ever since, we have seen models able to learn action models with quantifiers~\cite{AmirC08,ZhuoYHL10}, from noisy actions or noisy states~\cite{zhuo2013action,MouraoZPS12}, from null state information~\cite{cresswell2013acquiring}, from incomplete domain models~\cite{ZhuoNK13,ZhuoK17} and many more.
 
@@
 In this paper we claim that a planning model is learnable even though an accurate representation of the agent's behaviour is not available.
-Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under a minimum input knowledge.
-While current learning systems accept varying degrees of observability in the states traversed by the input observed actions, \FAMA extends partial observability to the action sequence.
-This way, we can define a percentage of observability of the actions up to a null observability.
-The minimal observability case acceptable by \FAMA is when the algorithm is only fed with the input and output of a plan trace, namely the initial and final state of a planning task.
+Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimum input knowledge.
+While current learning systems accept varying degrees of observability in the states traversed by the input observed actions, none of them allow partial observability in the sequence of actions.
+\FAMA, on the other hand, allows incomplete or empty action sequences.
+This way, the minimal observability case acceptable by \FAMA is when the algorithm is only fed with the input and output of a plan trace, namely the initial and final state of a planning task.
 Like many Machine Learning (ML) techniques, \FAMA is able to operate with only correct input/ouput pairs and an unknown or a partially known model of the agent.

### 2018-07-31 doc/aij18/experiments.tex
first updates added to the experiments section aij18 paper

@@
 
-In this section we present a set of experiments designed to evaluate different aspects of \FAMA, from its performance to the quality of the learned models.
-Whenever applicable, we will consider both NP-complete and PSPACE-complete scenarios, in order to draw conclusions at both levels of complexity.
-We will also devote one experiment to assess the suitability of the evaluation metrics proposed in this paper.
+This section evaluates different aspects of \FAMA that range, from CPU-time performance to the quality of the learned models.
+Whenever applicable, we will consider both NP-complete and PSPACE-complete scenarios to draw conclusions at both levels of worst case complexity.
+We will also devote one experiment to assess the suitability of the brand new evaluation metrics introduced in this paper.
 
 \subsection{Setup}
-We evaluate \FAMA on fourteen IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2} and that are taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+We evaluate \FAMA on fourteen IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, all taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
 Table \ref{tab:domain_features} compiles the features of these fourteen domains used in the experiments.
-From left to right, the columns report the name of the domain, number of actions, number of predicates, maximum arity of the actions, and maximum arity of the predicates.
-These are all features which affect the size of $P_\Lambda$ and hence, have an impact on the complexity of the learning task.
-%As can be seen in table \ref{tab:domain_features}, the selected domains range from, such as \emph{blocks} and \emph{visitall}, to complex ones, like \emph{floortile} and \emph{satellite}, with higher number of actions and predicates and with higher arity.
+From left to right columns report the domain, the number of actions, number of predicates, maximum arity of the actions, and maximum arity of the predicates in the domain.
+All these features affect the size of our compilation.
+%and hence, have an impact on the complexity of the learning task.
+As can be seen in table \ref{tab:domain_features}, the selected domains range from, such as \emph{blocks} and \emph{visitall}, to complex ones, like \emph{floortile} and \emph{satellite}, with higher number of actions and predicates and with higher arity.
 
-In the following, we explain the details of our experimental setup:
+Now we explain the details of our experimental setup:
 
@@
 \item {\bf Plan traces}.
-For each domain, we have generated 10 plan traces with 10 actions/intermediate states each using random walks.
-Depending on the experiment, these traces are used for training or testing, so more details will be provided later.
+For each domain, we generated 10 plan traces (with 10 actions/intermediate states each) using random walks.
+Depending on the experiment, these traces are used for training or testing purposes, details on this are provided at the particular experiment.
 
@@
 The classical planner we use to solve the $P_\Lambda$ instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use {\sc Madagascar} for two reasons: 1) its ability to deal with planning instances populated with dead-ends, and 2) when the length of the plan traces is known (\FO action sequences or \FO state trajectories), the horizon of the solution plan is also known and can be provided to {\sc Madagascar}, addressing in this case NP-complete problems.
-This is because, SAT-based planners can apply the actions for programming preconditions in parallel during a single planning step (and the same for actions programming effects) because these actions do not interact.
-Hence, we know that the programming prefix of the solution plan can be solved in two steps, independently of the number of programming actions applied.
+We use {\sc Madagascar} for two reasons:
+  \begin{enumerate}
+  \item  Its ability to deal with planning instances populated with dead-ends.
+  \item  It can levarage the horizon of the solution plan, when it is known.
+Recall that, when the length of the plan traces is known (\FO action sequences or \FO state trajectories) the horizon of the solution plan is also known.
+SAT-based planners can apply the actions for programming preconditions in parallel during a single planning step (and the same for actions programming effects) because these actions do not interact.
+Hence, we know that the programming prefix of a solution plan can be solved in two steps, independently of the number of programming actions applied.
+  \end{enumerate}
 
-\item {\bf Machine}. All experiments are run on an Intel Core i5 3.50 GHz x 4 with 16 GB of RAM.
+\item {\bf Hardware}. All experiments are run on an Intel Core i5 3.50 GHz x 4 with 16 GB of RAM.
 
@@
 		\end{center}
-	\caption{\small Feature description of the 13 domains used in the experiments.}
+	\caption{\small Feature description of the domains used in the experiments.}
 	\label{tab:domain_features}	
@@
 \subsection{Impact of the size of the input knowledge}
-This experiment evaluates how the size of the input knowledge, i.e., $\left|\mathcal{T}\right|$ affects the performance of \FAMA.
-The experiment analyzes the evolution of the computation time as well as the quality of the learned models as the size of $\mathcal{T}$ increases (from 1 to 10).
-Recall that each trace in $\mathcal{T}$ comprises 10 actions/intermediate states.
-To keep the experiment practicable, we introduce a 1000s timeout after which the learning process is killed and the learned models are assigned a score of 0 for both precision and recall.
-Our goals with this experiment are to:
+This experiment evaluates how the size of the input knowledge affects the performance of \FAMA to:
 \begin{enumerate}
-	\item Identify the amount of input required by \FAMA to learn sound and complete models,
-	\item Evaluate the scalability of \FAMA wrt the input size which, as stated in section \ref{sec:properties}, is one of the limiting factors of our approach.
+	\item Identify the minimal amount of input knowledge required by \FAMA to learn {\em sound} and {\em complete} models,
+	\item Evaluate the scalability of \FAMA with respect to the size of the input knowledge. % which, as stated in section \ref{sec:properties}, is one of the limiting factors of our approach.
 \end{enumerate}
 
-With this in mind we defined two case studies:
+The experiment analyzes the evolution of the CPU-time and quality of the learned models as $|\mathcal{T}|$ increases from 1 to 10 (recall that each trace in $\mathcal{T}$ comprises 10 actions/intermediate states).
+To keep the experiment practicable, we introduce a 1000s timeout after which the learning process is killed and $0$ scores, for both {\em precision} and {\em recall}, are given to the learned model.
+We defined two case studies:
 \begin{itemize}
-	\item \textbf{\FO action sequence and \PO state trajectory}: This is the typical case addressed by most previous work, which corresponds to a NP-complete scenario.
-We are assuming a degree of observability of just 10\% for the state trajectory, meaning that each literal of a state has a 10\% chance of being observed.
-	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a PSPACE-complete scenario where both the action sequence and state trajectory are completely empty and only initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
+	\item \textbf{\FO action sequence and \PO state trajectory}: This is the typical case addressed by most of previous work, which corresponds to NP-complete scenarios.
+In this experiment we assume a degree of state observability of just 10\%, meaning that each literal of a state has a 10\% chance of being observed.
+	\item  \textbf{\NO action sequence and \NO state trajectory}: A PSPACE-complete scenario where both the input action sequence and state trajectory are {\em empty}.
+Only the initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
 \end{itemize}
+
+Figures \ref{fig:np_quality} and \ref{fig:np_time} show the quality of the models and computation time, respectively, for the NP-complete scenario.
+In Figure \ref{fig:np_quality} we see precision stabilizes at 0.86 after 5 traces while recall stabilizes at 0.95 after 3 traces.
+These results show that \FAMA does, in fact, not need big amounts of input to learn sound and complete models as opposite to other approaches in the literature where models are learned using a hundred of traces (see Table \ref{table:models_comparison2}).
+
+With regards to the scalability of \FAMA, Figure \ref{fig:np_time} paints an exponential increase in computation time.
+Until an input size of 4 traces the computation time is below 1 s, but it reaches 153 s when the input is composed of 10 traces.
+These results match the expected performance of {\sc Madagascar}, since this planner is known to struggle with plan horizons beyond 150-200 steps (in our case 160 steps corresponds to 8 traces).
 
@@
 
-Figures \ref{fig:np_quality} and \ref{fig:np_time} show the quality of the models and computation time, respectively, for the NP-complete scenario.
-In figure \ref{fig:np_quality} we see precision stabilizes at 0.86 after 5 traces while recall stabilizes at 0.95 after 3 traces.
-These results show that \FAMA does, in fact, not need big amounts of input to learn sound and complete models as opposite to other approaches in the literature where models are learned using around a hundred traces (see table \ref{table:models_comparison2}).
-
-With regards to the scalability of \FAMA, figure \ref{fig:np_time} paints a very interesting picture.
-In this figure, we can see an exponential increase in computation time for input sizes beyond 5 traces.
-Until an input size of 4 traces the computation time is below 1 s, but it reaches 153 s when the input is composed of 10 traces.
-These results match the performance of {\sc Madagascar}, since this planner is known to struggle with plan horizons beyond 150-200 steps (in our case 160 steps corresponds to 8 traces).
+\newpage
+Figure~\ref{fig:pspace_quality} reports the {\em precision} and {\em recall} of the models learned by our approach in the PSPACE-complete scenario (NO action sequence and NO state trajectory).
+The quality of the learned models is now lower than in the previous case (the NP-complete scenario) and a quality drop appears as the amount of input knowledge increases.
+This is caused by an increasing number of timeouts in the experiments, meaning that no solution is found.
+The computation time in this scenario (figure \ref{fig:pspace_time}) is also considerably higher given the complexity of the task, which is reflected on the amount of timeouts.
 
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{figures/input_size_0_0_precision.eps}
+	\includegraphics[width=0.65\linewidth]{figures/input_size_0_0_precision.eps}
 	\caption{Evaluation of the impact of the input size on the quality of the learned models when learning from plan traces with \NO action sequences and \NO state trajectories}
@@
 \end{figure}
-
 \begin{figure}[hbt!]
 	\centering
-	\includegraphics[width=0.8\linewidth]{figures/input_size_0_0_time.eps}
+	\includegraphics[width=0.65\linewidth]{figures/input_size_0_0_time.eps}
 	\caption{Evaluation of the impact of the input size on the computation time when learning from plan traces with \NO action sequences and \NO state trajectories}
@@
 \end{figure}
-
-Next, we evaluate the PSPACE-complete scenario.
-As expected, the quality of the learned models, shown in figure \ref{fig:pspace_quality}, is inferior than in the NP-complete case.
-The results contradict the previous trend and we find a drop in quality as the input knowledge increases.
-This is, in fact, caused by an increasing number of timeouts in the experiments, meaning that no solution is found.
-The computation time in this scenario (figure \ref{fig:pspace_time}) is also considerably higher given the complexity of the task, which is reflected on the amount of timeouts.
 

### 2018-07-31 doc/aij18/experiments.tex
updated section 6 AIJ18 paper

@@
   \item  Its ability to deal with planning instances populated with dead-ends.
-  \item  It can levarage the horizon of the solution plan, when it is known.
+  \item  It can leverage the horizon of the solution plan, when it is known.
 Recall that, when the length of the plan traces is known (\FO action sequences or \FO state trajectories) the horizon of the solution plan is also known.
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{figures/input_size_100_10_precision.eps}
-	\caption{Evaluation of the impact of the input size on the quality of the learned models when learning from plan traces with \FO action sequences and \PO state trajectories with 10\% observability}
+	\includegraphics[width=0.7\linewidth]{figures/input_size_100_10_precision.eps}
+	\caption{{\em Precision} and {\em recall} when learning from [1-10] plan traces with \FO action sequences and \PO state trajectories with 10\% observability.}
 	\label{fig:np_quality}
@@
 	\centering
-	\includegraphics[width=0.8\linewidth]{figures/input_size_100_10_time.eps}
-	\caption{Evaluation of the impact of the input size on the computation time when learning from plan traces with \FO action sequences and \PO state trajectories with 10\% observability}
+	\includegraphics[width=0.7\linewidth]{figures/input_size_100_10_time.eps}
+	\caption{Computation time when learning from [1-10] plan traces with \FO action sequences and \PO state trajectories with 10\% observability.}
 	\label{fig:np_time}
@@
 \newpage
-Figure~\ref{fig:pspace_quality} reports the {\em precision} and {\em recall} of the models learned by our approach in the PSPACE-complete scenario (NO action sequence and NO state trajectory).
-The quality of the learned models is now lower than in the previous case (the NP-complete scenario) and a quality drop appears as the amount of input knowledge increases.
-This is caused by an increasing number of timeouts in the experiments, meaning that no solution is found.
-The computation time in this scenario (figure \ref{fig:pspace_time}) is also considerably higher given the complexity of the task, which is reflected on the amount of timeouts.
+Figure~\ref{fig:pspace_quality} reports {\em precision} and {\em recall} for the models learned by our approach in the PSPACE-complete scenario (NO action sequence and NO state trajectory).
+The {\em precision} and {\em recall} values are now lower and they drop as the amount of input knowledge increases.
+For the PSPACE-complete scenario, {\em precision} and {\em recall} scores have to be taken with a grain of salt because {\bf pure syntax-based metrics are not informative when reformulations appear}.
+This is the case of such an under-constrained learning task and we provide experimental evidence of this in Section \ref{minimal}.
+Further, the score drop reflects an increasing number of timeouts meaning that, no solution is found for several domains within the given time-bound.
+Computation time (Figure \ref{fig:pspace_time}) is also higher explained again by this larger number of timeouts produced in a scenario with larger worst case complexity.
 
@@
 	\centering
-	\includegraphics[width=0.65\linewidth]{figures/input_size_0_0_precision.eps}
-	\caption{Evaluation of the impact of the input size on the quality of the learned models when learning from plan traces with \NO action sequences and \NO state trajectories}
+	\includegraphics[width=0.7\linewidth]{figures/input_size_0_0_precision.eps}
+	\caption{{\em Precision} and {\em recall} when learning from [1-10] plan traces with \NO action sequences and \NO state trajectories.}
 	\label{fig:pspace_quality}
@@
 	\centering
-	\includegraphics[width=0.65\linewidth]{figures/input_size_0_0_time.eps}
-	\caption{Evaluation of the impact of the input size on the computation time when learning from plan traces with \NO action sequences and \NO state trajectories}
+	\includegraphics[width=0.7\linewidth]{figures/input_size_0_0_time.eps}
+	\caption{Computation time when learning from [1-10] plan traces with \NO action sequences and \NO state trajectories.}
 	\label{fig:pspace_time}
@@
 
-The conclusion we can draw is that learning from few input samples is both our strong point and our limitation.
+%The conclusion we can draw is that learning from few input samples is both our strong point and our limitation.
 Results on the NP-complete scenario show that the learned models are considerably sound and complete (to put these results into a richer perspective we will also compare ourselves with another approach in the following section).
@@
 \subsection{Comparison with \ARMS}
-Here we compare the performance of \FAMA to that of \ARMS, one of the most well-known approaches learning planning models.
-\ARMS works under the assumption of plan traces with \FO action sequences and \NO state trajectories, so we will also adopt this assumption (\ARMS, as any other previous approaches for learning planning models cannot handle the PSPACE-complete scenarios).
+Now we compare the performance of \FAMA to that of \ARMS.
+\ARMS assumes input plan traces with \FO action sequences and \NO state trajectories so here, we will also adopt this assumption (\ARMS, as any other previous approaches for learning planning models cannot handle the PSPACE-complete scenario).
 
 We define a \emph{degree of observability} $\sigma$ for the state trajectory, ranging from 0\% to 100\%, that measures the probability of observing a literal, and evaluate both approaches as $\sigma$ increases using 5 plan traces as input.
-Note that when $\sigma = 0$ we have a \NO state trajectory, when $\sigma=100$ we have a \FO state trajectory and all cases in-between correspond to the \PO scenario.
-Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of precision and recall.
-The horizontal axis represents the degree of observability, while the vertical axis show the average precision (figure \ref{fig:comparison_precision}) or recall (figure \ref{fig:comparison_recall}) computed over the fourteen tested domains.
-Remarkably, figure \ref{fig:comparison_precision} shows that \FAMA dominates in terms of precision in all cases save for \FO state trajectories.
-Furthermore, in average, the models learned by our approach are 15\% to 24\% more precise than those learned by \ARMS.
-A similar conclusion can be drawn for recall (figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our models are more complete.
+When $\sigma = 0$ we have a \NO state trajectory, when $\sigma=100$ we have a \FO state trajectory and all cases in-between correspond to the \PO scenario.
+Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of {\em precision} and {\em recall}.
+Horizontal axes represents the degree of observability, while vertical axes show the average {\em precision} (Figure \ref{fig:comparison_precision}) or {\em recall} (Figure \ref{fig:comparison_recall}).
+\FAMA dominates in terms of {\em precision} in all cases (the models learned by our approach are 15\% to 24\% more precise than those learned by \ARMS) except for \FO state trajectories.
+A similar trend is drawn for {\em recall} (Figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our learned models are more complete.
+These results evidence the \FAMA capacity to learn better models when few plan traces are available but do not imply that our approach is overall better for NP-complete scenarios (Figure~\ref{fig:np_time} showed that \FAMA cannot handle so much input as \ARMS).
 
@@
 	\centering
-	\includegraphics[width=.8\linewidth]{figures/comparison_precision.eps}
-	\caption{Precision comparison between \FAMA and \ARMS}
+	\includegraphics[width=.65\linewidth]{figures/comparison_precision.eps}
+	\caption{{\em Precision} comparison between \FAMA and \ARMS for different \emph{degrees of observability}.}
 	\label{fig:comparison_precision}
@@
 	\centering
-	\includegraphics[width=.8\linewidth]{figures/comparison_recall.eps}
-	\caption{Recall comparison between \FAMA and \ARMS}
+	\includegraphics[width=.65\linewidth]{figures/comparison_recall.eps}
+	\caption{{\em Recall} comparison between \FAMA and \ARMS for different \emph{degrees of observability}.}
 	\label{fig:comparison_recall}
@@
 
-\FAMA outperforms \ARMS in this particular experiment, which is one with very limited input.
-By no means do these results imply that our approach is overall better for the NP-complete scenarios, just that it is able to learn better models when few plan traces are available.
-We are not reporting experiments where far more input traces are available, since \FAMA cannot handle so much input.
-
-\subsection{Experiments with minimal input knowledge}
+\subsection{Learning with minimal input knowledge}
 \label{minimal}
@@
 So far, experiments have shown that \FAMA is able to learn with very small amounts of input knowledge, be it due to low observability or few training samples.
-In this section we want to take a closer look at the action models learned from minimal input knowledge.
+Let us take a closer look at the action models learned from minimal input knowledge.
 To that end, we will limit the input to only 2 plan traces and analyze the results under different levels of observability.
-
-In this experiment we evaluate three case studies, the first one being an NP-complete scenario and the other two being PSPACE-complete:
+In this experiment we evaluate three cases of study, the first one being an NP-complete scenario and the other two being PSPACE-complete:
 \begin{itemize}
 	\item \textbf{\FO action sequence and \PO state trajectory}: We are, once again, assuming a degree of observability of 10\% for the state trajectory.
-The results of this case study are detailed in table \ref{tab:results_minimum_100_10}
+The results of this case study are detailed in Table \ref{tab:results_minimum_100_10}.
 	\item  \textbf{\PO action sequence and \PO state trajectory}: In this case study we are assuming a degree of observability of 30\% for both the action sequence and state trajectory.
-The results of this case study are detailed in table \ref{tab:results_minimum_30_30}
-	\item  \textbf{\NO action sequence and \NO state trajectory}: Both the action sequence and state trajectory are completely empty and only initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
-The results of this case study are detailed in table \ref{tab:results_minimum_0_0}
+The results of this case study are detailed in Table \ref{tab:results_minimum_30_30}.
+	\item  \textbf{\NO action sequence and \NO state trajectory}: Both the action sequence and state trajectory are completely empty so only the initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
+The results of this case study are detailed in Table \ref{tab:results_minimum_0_0}.
 \end{itemize}
 
-All tables in this section (tables \ref{tab:results_minimum_100_10}, \ref{tab:results_minimum_0_0}, \ref{tab:results_minimum_30_30}) follow the same structure.
-In these tables precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), and also globally ({\bf Global}).
-The last column reports the computation time in seconds needed to obtain the learned models.
-Missing values in the tables correspond to domains where no solution was found given a timeout of 1800 s.
+All tables in this section (Tables \ref{tab:results_minimum_100_10}, \ref{tab:results_minimum_30_30} and \ref{tab:results_minimum_0_0}) follow the same structure.
+{\em Precision} ({\bf P}) and {\em Recall} ({\bf R}) scores are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), and also globally ({\bf Global}).
+The last column reports the computation time (in seconds) needed to obtain the learned models.
+Missing values in the tables (-) correspond to domains where no solution was found given a timeout of 1800 s.
+
+Table \ref{tab:results_minimum_100_10} shows the results of the first case study.
+Recall scores are generally higher than the precision ones, and, in fact, the models learned for six out of the fourteen domains were perfectly complete.
+Although precision is overall lower, it is interesting to notice that the soundness of most learned sets of positive effects is flawless.
+With regards to the computation time, we find times below the second in all cases except for the {\em floor-tile} domain.
 
@@
 		\begin{center}
-
-			\begin{tabular}{l|l|l|l|l|l|l||l|l|l|}
+                  \begin{footnotesize}			
+			\begin{tabular}{l|l|l|l|l|l|l||l|l||l|}
 				& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c|}{\bf Global} & \\ \cline{2-9}			
@@
 			\end{tabular}
-
+                  \end{footnotesize}			
 		\end{center}
-	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \FO action sequences and \PO state trajectories with 10\% observability}
+	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \FO action sequences and \PO state trajectories with 10\% observability.}
 	\label{tab:results_minimum_100_10}
@@
 
-Table \ref{tab:results_minimum_100_10} shows the results of the first case study.
-As was expected from the previous experiments, the recall scores are generally higher than the precision ones, and, in fact, the models learned for 6 of the tested domains were perfectly complete.
-Although precision is overall lower, it is interesting to notice that the soundness of most learned sets of positive effects is flawless.
-With regards to the computation time, we find times below the second in all cases except for one, "floor-tile", the most complex among the tested domains.
+In Table \ref{tab:results_minimum_30_30}, which collects the result of the PSPACE case study with 30\% observability, the scores for some domains are missing.
+This is the case for {\em floor-tile}, and {\em parking} both considered \emph{puzzle-like} domains and that put a strain on the planners.
+Interestingly enough, {\em hanoi} also qualifies as a \emph{puzzle-like} domain, and we can see how this is reflected on the computation time.
+With regards to the learned models, we find that they retain similar level of soundness, while their completeness is lower than in the previous case study.
+This is specially the case for the preconditions, which have gone from 0.94 recall to 0.54.
+%This results are difficult to explain, given that in this case study the intermediate states contain more information, but
+The explanation for these scores is that input actions are strong constraints that play a key role on how close the learned model are with respect to the GTM.
+The more these actions are missing the more present is reformulation and pure syntax-based metrics become less relevant.
 
@@
 	\begin{center}
-
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|l|}
+		\begin{footnotesize}
+		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|}
 			& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c|}{\bf Global} & \\ \cline{2-9}			
@@
 		\end{tabular}
-
+		\end{footnotesize}
 	\end{center}
-	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \PO action sequences and \PO state trajectories with 30\% observability in both cases}
+	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \PO action sequences and \PO state trajectories with 30\% observability in both cases.}
 	\label{tab:results_minimum_30_30}
@@
 
-The first thing to notice in table \ref{tab:results_minimum_30_30}, which collects the result of the PSPACE case study with 30\% observability, is that the values for some domains are missing.
-This is the case for "floor-tile", and "parking" which not only are fairly complex domains, but are also considered \emph{puzzle-like}, a quality that is known to put a strain on the planners.
-Interestingly enough, "hanoi" also qualifies as a \emph{puzzle-like} domain, and we can see how this is reflected on the computation time.
-With regards to the quality of the models, we find that the learned models retain a similar level of soundness, while their completeness has drastically worsened with respect to the previous case study.
-This is specially the case for the preconditions, which have gone from 0.94 recall to 0.54.
-%This results are difficult to explain, given that in this case study the intermediate states contain more information, but
-Our intuition is that the input actions play a big role on how close the learned model are with respect to the GTM.
+The effect of missing input actions is more evident in our last case study with \NO action sequences and state trajectories (Table \ref{tab:results_minimum_0_0}).
+In this case we are able to find solutions for all domains.
+Search is even less constrained so there are far more possible solutions for this learning task.
+This wider range of solutions is also evidenced by the kind of the learned models, which in spite of being compliant with the input data, are further from the original GTM.
+In the table we find that both precision and recall have seen their scores decreased to 0.6 and 0.5, respectively.
+Syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the \emph{reformulations} that frequently occur in these circumstances.
 
-Next, we analyze the PSPACE case study with \NO action sequences and state trajectories (Table \ref{tab:results_minimum_0_0}).
-Contrary to what might be expected by looking at the previous table, we are able to find solutions for all domains.
-This is because, although the search is less guided, there are far more possible solutions for this learning task.
-This wider range of solutions is also evidenced by the quality of the learned models, which in spite of being compliant with the input data, are further from the original GTM.
-In the table we find that both precision and recall have seen their scores decreased to 0.6 and 0.5, respectively.
+To illustrate this, Figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the {\tt stack} operator learned from plan traces with \NO action sequences and state trajectories.
+This action model removes a block from on top of another block and puts it down on the table, all done in a single step.
+There are two main differences from the GTM for {\tt stack}: 1) the learned action is unstacking a block instead of stacking it, and 2) the block ends on the table after the action is applied, working as an {\tt unstack + put-down} \emph{macro-action}.
+We refer to the first difference as \emph{role swapping} and it happens when actions are removed from the trace.
+If no actions are present in the input traces, the name of actions becomes meanless (actions are effectively anonymous), meaning that they can interchange their behavior with any comparable action model.
+The second issue can happen if there are missing states in the input traces.
 
@@
   \begin{center}
-		\begin{tabular}{l|l|l|l|l|l|l||l|l|l|}
+    \begin{footnotesize}
+		\begin{tabular}{l|l|l|l|l|l|l||l|l||l|}
 			& \multicolumn{2}{|c|}{\bf Pre} & \multicolumn{2}{|c|}{\bf Add} & \multicolumn{2}{|c||}{\bf Del} & \multicolumn{2}{|c|}{\bf Global} & \\ \cline{2-9}			
@@
 		\end{tabular}
+            \end{footnotesize}    
 	\end{center}
-	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \NO action sequences and \NO state trajectories}
+	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \NO action sequences and \NO state trajectories.}
 	\label{tab:results_minimum_0_0}
 \end{table}
-
-We argue, however, that syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the \emph{reformulations} that frequently occur in these circumstances.
-To illustrate this, Figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the {\tt stack} operator learned from plan traces with \NO action sequences and state trajectories.
 
@@
 
-This action model removes a block from on top of another block and puts it down on the table, all done in a single step.
-There are two main differences from the expected action model for {\tt stack}: 1) it is unstacking a block instead of stacking it, and 2) instead of the arm holding the block by the end of the action, the block ends on the table.
-We refer to the first difference as \emph{role swapping} and it happens when actions are removed from the trace.
-If no actions are present in the input traces, the name of the actions lose their utility and they are effectively anonymous actions, meaning that they can interchange their behavior with any comparable action model.
-The second difference is commonly known as \emph{macro-action} (in this case unstack + put-down) and it can happen when there are missing states in the input traces.
 Reformulated action models are perfectly valid and can be used to solve planning tasks.
@@
 
-In this experiment we compare the scores provided by the syntactic and semantic versions of precision and recall.
+This experiment compares the scores provided by the syntactic and semantic versions of {\em precision} and {\em recall}.
 For that purpose, we will evaluate the models learned in \ref{minimal} both syntactically, using the GTM, and semantically, computing the closest compliant set of action models with our approach.
 For the semantic evaluation we will be using 5 plan traces.
-Since we are using {\sc Madagascar}, a satisfying planner, the scores for sem-Precision and sem-Recall are approximated.
-
-Our goal with this experiment is to gauge the suitability of the semantic metrics we propose with respect to their well-known counterparts. With that in mind, we define two case studies:
+Our goal with this experiment is to gauge the suitability of the semantic metrics we propose with respect to their well-known counterparts.
+With that in mind, we define two case studies:
 
@@
 In fact, in all our experiments with \FAMA and other approaches we have never observed reformulations when the full sequence of actions is known.
-	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a case study that favors reformulations in the learned models, as we discussed previously.
+	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a case study that favors reformulations in the learned models, as previously discussed.
 \end{itemize}
@@
 \begin{table}[hbt!]
+     \begin{footnotesize}
 	\begin{center}		
@@
 	\end{center}
-	\caption{\small Syntactic and semantic metric scores for learning tasks with \FO action sequences and \PO state trajectories with 10\% observability}
+     \end{footnotesize}
+     \caption{\small Syntactic and semantic metric scores for learning tasks with \FO action sequences and \PO state trajectories with 10\% observability.}
      \label{tab:metric_comparison_100_10}
@@
 Looking at the high scores of the syntactic metrics, specially recall, we can conclude that the learned models are, in fact, fairly similar to the GTM.
-This supports our conclusion that no reformulation can occur in this case study, which also means that the space of possible solutions is restricted to models close to the GTM.
-The values for sem-Precision and sem-Recall are very high across the table, which is exactly the desired behavior for the metric given that the solutions are very close to the GTM.
+This supports our conclusion that no reformulation occurs in this case study, which also means that the space of possible solutions is restricted to models close to the GTM.
+The values for {\em sem-Precision} and {\em sem-Recall} are very high across the table, which is exactly the desired behavior for the metric given that the solutions are close to the GTM.
 In comparison, both versions of recall show very similar scores, while the difference is bigger for the precision metrics.
 This means that the semantic version is more lenient towards extra preconditions or effects.
-This matches what we have observed during the experiments, where it is usual for the learned models to contain redundant or implicit preconditions which are penalized by the Precision metric.
-An example of this can be found in the {\tt move} action of the "hanoi" domain, where the learned action model specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
+This matches what we have observed during the experiments, where it is usual for the learned models to contain redundant or implicit preconditions which are penalized by the {\em precision} metric.
+For instance the {\tt move} action of the {\em hanoi} domain, where the learned model specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
+This learned model is semantically correct but sintactically different from the GTM.
 
-Next, we take a look at table \ref{tab:metric_comparison_0_0} that details the results of the \NO/\NO case study, and we can see that using 5 traces in this PSPACE-complete scenario made it impossible to find a solution for some of the most demanding domains.
+Finally Table \ref{tab:metric_comparison_0_0} details the results of the \NO/\NO case study.
+We can see that using 5 traces in this PSPACE-complete scenario made it impossible to find a solution for some of the most demanding domains.
 Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger here.
-Comparing the scores of both versions, we find that models that originally achieved mediocre scores when using the GTM as reference, are considerably sound and complete, reaching and overall score of 0.91 for both sem-Precision and sem-Recall.
+Comparing the scores of both versions, we find that models that originally achieved mediocre scores when using the GTM as reference, are in fact reasonable sound and complete, reaching and overall score of 0.91 for both {\em sem-Precision} and {\em sem-Recall}.
+This means that the learned models require very few editions to explain the traces given as input.
 
 \begin{table}[hbt!]
+     \begin{footnotesize}
 	\begin{center}		
@@
 	\end{center}
-	\caption{\small Syntactic and semantic metric scores for learning tasks with \NO action sequences and \NO state trajectories}
+        \end{footnotesize}
+	\caption{\small Syntactic and semantic metric scores for learning tasks with \NO action sequences and \NO state trajectories.}
 	\label{tab:metric_comparison_0_0}
@@
 Looking at the results of both case studies we can draw two conclusions with regards to the semantic metrics proposed in this paper.
-The first is that these metrics behave similarly to their syntactic counterparts, which means they are a good substitute when the GTM is not available.
-The second conclusion is that sem-Precision and sem-Recall are better suited to evaluate reformulated models than the original syntactic metrics, since they contemplate valid solutions outside the GTM.
+The first one is that, when no reformulation occurs, these metrics behave similarly to their syntactic counterparts, which means they are a good substitute when the GTM is not available.
+The second conclusion is that sem-Precision and sem-Recall are better suited to evaluate reformulated models than the original syntactic metrics, since they contemplate valid solutions outside the GTM that successfully explain the given input data.
+%Since we are using {\sc Madagascar}, a satisfying planner, the scores for sem-Precision and sem-Recall are approximated.
 

### 2018-08-01 doc/aij18/experiments.tex
section 6 aij18 paper revised

@@
 Whenever applicable, we will consider both NP-complete and PSPACE-complete scenarios to draw conclusions at both levels of worst case complexity.
-We will also devote one experiment to assess the suitability of the brand new evaluation metrics introduced in this paper.
+We will also devote one experiment to show the suitability of the brand new evaluation metrics introduced in this paper.
 
@@
 We evaluate \FAMA on fourteen IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, all taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-Table \ref{tab:domain_features} compiles the features of these fourteen domains used in the experiments.
-From left to right columns report the domain, the number of actions, number of predicates, maximum arity of the actions, and maximum arity of the predicates in the domain.
-All these features affect the size of our compilation.
+Table \ref{tab:domain_features} compiles the features of the domains, all these features affect the size of our compilation.
+From left to right columns report, for each domain, its number of actions, number of predicates, maximum arity of the actions, and maximum arity of the predicates.
 %and hence, have an impact on the complexity of the learning task.
@@
 Now we explain the details of our experimental setup:
-
 \begin{itemize}
-
 \item {\bf Plan traces}.
 For each domain, we generated 10 plan traces (with 10 actions/intermediate states each) using random walks.
-Depending on the experiment, these traces are used for training or testing purposes, details on this are provided at the particular experiment.
+Depending on the experiment, these traces are used for training or testing purposes (more details on this are provided at the particular experiment).
 
@@
   \begin{enumerate}
-  \item  Its ability to deal with planning instances populated with dead-ends.
-  \item  It can leverage the horizon of the solution plan, when it is known.
-Recall that, when the length of the plan traces is known (\FO action sequences or \FO state trajectories) the horizon of the solution plan is also known.
+  \item  Its ability to deal with planning instances populated with dead-ends~\cite{lopez2015deterministic}.
+  \item  It can leverage the horizon of the solution plan, when it is known in advance.
+When the length of plan traces is known (e.g.
+for \FO action sequences or \FO state trajectories) the horizon of the solution plan is also known.
 SAT-based planners can apply the actions for programming preconditions in parallel during a single planning step (and the same for actions programming effects) because these actions do not interact.
-Hence, we know that the programming prefix of a solution plan can be solved in two steps, independently of the number of programming actions applied.
+Hence, we also know that the programming prefix of a solution plan can be solved in two steps, independently of the number of programming actions applied.
   \end{enumerate}
@@
 \begin{table}[hbt!]
+  \begin{footnotesize}			  
 		\begin{center}
@@
 		\end{center}
+                \end{footnotesize}			                
 	\caption{\small Feature description of the domains used in the experiments.}
@@
 
-The experiment analyzes the evolution of the CPU-time and quality of the learned models as $|\mathcal{T}|$ increases from 1 to 10 (recall that each trace in $\mathcal{T}$ comprises 10 actions/intermediate states).
+The experiment analyzes the evolution of the CPU-time, as well as the {\em precision} and {\em recall} of the learned models wrt the GTM, as $|\mathcal{T}|$ increases from 1 to 10 (each trace in $\mathcal{T}$ comprises 10 actions/intermediate states).
 To keep the experiment practicable, we introduce a 1000s timeout after which the learning process is killed and $0$ scores, for both {\em precision} and {\em recall}, are given to the learned model.
@@
 With regards to the scalability of \FAMA, Figure \ref{fig:np_time} paints an exponential increase in computation time.
-Until an input size of 4 traces the computation time is below 1 s, but it reaches 153 s when the input is composed of 10 traces.
+Until an input size of 4 traces the computation time is below 1 sec, but it reaches 153 secs when the input is composed of 10 traces.
 These results match the expected performance of {\sc Madagascar}, since this planner is known to struggle with plan horizons beyond 150-200 steps (in our case 160 steps corresponds to 8 traces).
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{figures/input_size_100_10_precision.eps}
+	\includegraphics[width=0.65\linewidth]{figures/input_size_100_10_precision.eps}
 	\caption{{\em Precision} and {\em recall} when learning from [1-10] plan traces with \FO action sequences and \PO state trajectories with 10\% observability.}
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{figures/input_size_100_10_time.eps}
+	\includegraphics[width=0.65\linewidth]{figures/input_size_100_10_time.eps}
 	\caption{Computation time when learning from [1-10] plan traces with \FO action sequences and \PO state trajectories with 10\% observability.}
@@
 This is the case of such an under-constrained learning task and we provide experimental evidence of this in Section \ref{minimal}.
-Further, the score drop reflects an increasing number of timeouts meaning that, no solution is found for several domains within the given time-bound.
-Computation time (Figure \ref{fig:pspace_time}) is also higher explained again by this larger number of timeouts produced in a scenario with larger worst case complexity.
+Further, the score drop reflects an increasing number of timeouts meaning that, no solution is found for several domains within the given time-bound (the worst case complexity of this scenario is larger).
+Computation time (Figure \ref{fig:pspace_time}) is also higher explained again by this larger number of timeouts.
 
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{figures/input_size_0_0_precision.eps}
+	\includegraphics[width=0.65\linewidth]{figures/input_size_0_0_precision.eps}
 	\caption{{\em Precision} and {\em recall} when learning from [1-10] plan traces with \NO action sequences and \NO state trajectories.}
@@
 	\centering
-	\includegraphics[width=0.7\linewidth]{figures/input_size_0_0_time.eps}
+	\includegraphics[width=0.65\linewidth]{figures/input_size_0_0_time.eps}
 	\caption{Computation time when learning from [1-10] plan traces with \NO action sequences and \NO state trajectories.}
@@
 Now we compare the performance of \FAMA to that of \ARMS.
-\ARMS assumes input plan traces with \FO action sequences and \NO state trajectories so here, we will also adopt this assumption (\ARMS, as any other previous approaches for learning planning models cannot handle the PSPACE-complete scenario).
+\ARMS, as any other previous approaches for learning planning models cannot handle the PSPACE-complete scenario.
+On the other hand, \ARMS assumes input plan traces with \FO action sequences and \NO state trajectories so here, we will also adopt this assumption.
 
@@
 Let us take a closer look at the action models learned from minimal input knowledge.
-To that end, we will limit the input to only 2 plan traces and analyze the results under different levels of observability.
+To that end, we will limit the input to only 2 plan traces and analyze the results under different degrees of observability.
 In this experiment we evaluate three cases of study, the first one being an NP-complete scenario and the other two being PSPACE-complete:
@@
 	\item \textbf{\FO action sequence and \PO state trajectory}: We are, once again, assuming a degree of observability of 10\% for the state trajectory.
-The results of this case study are detailed in Table \ref{tab:results_minimum_100_10}.
-	\item  \textbf{\PO action sequence and \PO state trajectory}: In this case study we are assuming a degree of observability of 30\% for both the action sequence and state trajectory.
-The results of this case study are detailed in Table \ref{tab:results_minimum_30_30}.
+Results of this case study are detailed in Table \ref{tab:results_minimum_100_10}.
+	\item  \textbf{\PO action sequence and \PO state trajectory}: In this case study (results shown in Table \ref{tab:results_minimum_30_30}) we are assuming a degree of observability of 30\% for both the action sequence and state trajectory.
 	\item  \textbf{\NO action sequence and \NO state trajectory}: Both the action sequence and state trajectory are completely empty so only the initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
-The results of this case study are detailed in Table \ref{tab:results_minimum_0_0}.
+Results of this case study are reported in Table \ref{tab:results_minimum_0_0}.
 \end{itemize}
@@
 The last column reports the computation time (in seconds) needed to obtain the learned models.
-Missing values in the tables (-) correspond to domains where no solution was found given a timeout of 1800 s.
+Missing values in the tables (reported as -) correspond to domains where no solution was found given a 1800 seconds timeout.
 
@@
 Recall scores are generally higher than the precision ones, and, in fact, the models learned for six out of the fourteen domains were perfectly complete.
-Although precision is overall lower, it is interesting to notice that the soundness of most learned sets of positive effects is flawless.
-With regards to the computation time, we find times below the second in all cases except for the {\em floor-tile} domain.
+Although precision is overall lower, it is interesting that most learned sets of negative effects are flawless.
+With regards to the computation time, we find times below one second in all cases except for the {\em floor-tile} domain.
 
@@
 	\end{center}
-	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \PO action sequences and \PO state trajectories with 30\% observability in both cases.}
+	\caption{\small {\em Precision} and {\em recall} when learning with \PO action sequences and \PO state trajectories, 30\% observability in both cases.}
 	\label{tab:results_minimum_30_30}
@@
 The effect of missing input actions is more evident in our last case study with \NO action sequences and state trajectories (Table \ref{tab:results_minimum_0_0}).
-In this case we are able to find solutions for all domains.
+In this case we find solutions for all domains.
 Search is even less constrained so there are far more possible solutions for this learning task.
-This wider range of solutions is also evidenced by the kind of the learned models, which in spite of being compliant with the input data, are further from the original GTM.
-In the table we find that both precision and recall have seen their scores decreased to 0.6 and 0.5, respectively.
+This wider range of solutions is also evidenced by the kind of the learned models, which in spite of being compliant with the input data, are further from the GTM.
+Both {\em precision} and {\em recall} decreased to 0.6 and 0.5, respectively.
 Syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the \emph{reformulations} that frequently occur in these circumstances.
@@
 To illustrate this, Figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the {\tt stack} operator learned from plan traces with \NO action sequences and state trajectories.
-This action model removes a block from on top of another block and puts it down on the table, all done in a single step.
-There are two main differences from the GTM for {\tt stack}: 1) the learned action is unstacking a block instead of stacking it, and 2) the block ends on the table after the action is applied, working as an {\tt unstack + put-down} \emph{macro-action}.
-We refer to the first difference as \emph{role swapping} and it happens when actions are removed from the trace.
-If no actions are present in the input traces, the name of actions becomes meanless (actions are effectively anonymous), meaning that they can interchange their behavior with any comparable action model.
-The second issue can happen if there are missing states in the input traces.
+This learned action model removes a block from on top of another block and puts it down on the table.
+There are two main differences with respect to the GTM: (1), the learned action is unstacking a block instead of stacking it and (2), the block ends on the table after the action is applied, working as an {\tt unstack\&putdown} \emph{macro-action}.
+We refer to the first difference as \emph{role swapping} and it happens when there are missing actions in the plan trace.
+If no actions are present in the input traces, the name of actions becomes meanless (actions are effectively anonymous), meaning that they can interchange their behavior with any other comparable action model.
+The second issue happens when there are missing states in the input traces.
 
@@
 
-Reformulated action models are perfectly valid and can be used to solve planning tasks.
-For instance, any blocks-world problem can be solved by unstacking all the blocks to the table (unstack + put-down) and then stacking them to meet the goal condition (pick-up + stack), so reformulated action models like the one in the figure can be used to solve common planning tasks.
-As we have shown, the \NO/\NO case study meets all the criteria needed for reformulations to happen, and this is the reason why scenarios such as this are better evaluated using semantic-based metrics.
+Reformulated action models, like the one in Figure \ref{fig:macroaction}, can be used to solve planning tasks.
+For instance, any blocks-world problem can be solved unstacking all the blocks to the table (unstack + put-down) and then stacking them to meet the goal conditions (pick-up + stack).
+The \NO/\NO case study meets all the conditions where reformulations happens, and this is the reason why scenarios such as this are better evaluated using {\em semantic-based metrics}.
 
@@
 
-This experiment compares the scores provided by the syntactic and semantic versions of {\em precision} and {\em recall}.
-For that purpose, we will evaluate the models learned in \ref{minimal} both syntactically, using the GTM, and semantically, computing the closest compliant set of action models with our approach.
-For the semantic evaluation we will be using 5 plan traces.
+The last experiment compares the scores provided by the syntactic and semantic versions of {\em precision} and {\em recall}.
+For that purpose, we will evaluate the models learned in \ref{minimal} both syntactically, using the GTM, and semantically, using a testing set of 5 plan traces.
 Our goal with this experiment is to gauge the suitability of the semantic metrics we propose with respect to their well-known counterparts.
@@
 	\item \textbf{\FO action sequence and \PO state trajectory}: In this case study the full sequence of actions is known and no states are missing, which makes it practically impossible for reformulated models to appear.
-In fact, in all our experiments with \FAMA and other approaches we have never observed reformulations when the full sequence of actions is known.
+In fact, in all our experiments with \FAMA and other approaches we never observed reformulations when the full sequence of actions is known.
 	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a case study that favors reformulations in the learned models, as previously discussed.
@@
      \end{footnotesize}
-     \caption{\small Syntactic and semantic metric scores for learning tasks with \FO action sequences and \PO state trajectories with 10\% observability.}
+     \caption{\small Syntactic and semantic scores when learning with \FO action sequences and \PO state trajectories with 10\% observability.}
      \label{tab:metric_comparison_100_10}
@@
 Table \ref{tab:metric_comparison_100_10} shows the results of the first case study.
-Looking at the high scores of the syntactic metrics, specially recall, we can conclude that the learned models are, in fact, fairly similar to the GTM.
+Looking at the high scores of the syntactic metrics, specially {\em recall}, we can conclude that the learned models are, in fact, fairly similar to the GTM.
 This supports our conclusion that no reformulation occurs in this case study, which also means that the space of possible solutions is restricted to models close to the GTM.
-The values for {\em sem-Precision} and {\em sem-Recall} are very high across the table, which is exactly the desired behavior for the metric given that the solutions are close to the GTM.
-In comparison, both versions of recall show very similar scores, while the difference is bigger for the precision metrics.
+The values for {\em sem-Precision} and {\em sem-Recall} are also high across the table, which is exactly the desired behavior for these metrics given that solutions are close to the GTM.
+In comparison, both versions of {\em recall} show similar scores, while the difference is bigger for the precision metrics.
 This means that the semantic version is more lenient towards extra preconditions or effects.
-This matches what we have observed during the experiments, where it is usual for the learned models to contain redundant or implicit preconditions which are penalized by the {\em precision} metric.
-For instance the {\tt move} action of the {\em hanoi} domain, where the learned model specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
-This learned model is semantically correct but sintactically different from the GTM.
+This matches what we have observed during the experiments for instance, the learned model for the {\tt move} action of the {\em hanoi} domain, specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
+This learned model is semantically correct but sintactically different from the GTM and hence, penalized by the {\em precision} metric.
 
 Finally Table \ref{tab:metric_comparison_0_0} details the results of the \NO/\NO case study.
-We can see that using 5 traces in this PSPACE-complete scenario made it impossible to find a solution for some of the most demanding domains.
 Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger here.
-Comparing the scores of both versions, we find that models that originally achieved mediocre scores when using the GTM as reference, are in fact reasonable sound and complete, reaching and overall score of 0.91 for both {\em sem-Precision} and {\em sem-Recall}.
-This means that the learned models require very few editions to explain the traces given as input.
+Comparing the scores of both versions, we find that learned models that achieved mediocre scores when using the GTM as reference, are in fact reasonable sound and complete, reaching and overall score of 0.91 for both {\em sem-Precision} and {\em sem-Recall}.
+This means that the learned models by our approach, despite syntactically different from the GTM, require very few editions to explain the testing traces given as input.
 
@@
 The first one is that, when no reformulation occurs, these metrics behave similarly to their syntactic counterparts, which means they are a good substitute when the GTM is not available.
-The second conclusion is that sem-Precision and sem-Recall are better suited to evaluate reformulated models than the original syntactic metrics, since they contemplate valid solutions outside the GTM that successfully explain the given input data.
+The second conclusion is that {\em sem-Precision} and {\em sem-Recall} are better suited to evaluate reformulated models than the original syntactic metrics, since they contemplate valid solutions outside the GTM that successfully explain the given input data.
 %Since we are using {\sc Madagascar}, a satisfying planner, the scores for sem-Precision and sem-Recall are approximated.

### 2018-08-01 doc/aij18/aij18.tex
serction 7 of aij18 paper revised

@@
 The approach is flexible to various amount and kind of input knowledge and accepts partially specified action models.
-Unlike extensive-data ML approaches, our work explores an alternative research direction to learn sound models from very small data sets.
-The action models of the {\em blocksworld} or {\em gripper} domains were perfectly learned from only 25 state observations.
-Moreover, in 12 out of the 15 domains, the learned models yield {\em Precision} values over $0.75$.
+Unlike extensive-data ML approaches, our work explores an alternative research direction that exploits logic reasoning to learn sound models from minimal input knowledge.
+Remarkably, the action models of the {\em gripper}, {\em npuzzle} or {\em visitall} domains were perfectly learned from only observations of the initial and final states.
 
-To the best of our knowledge, this is the first work on learning \strips\ action models from state observations, using exclusively an {\em off-the-shelf} classical planner, and evaluated over a wide range of different domains.
-Recently, the work in~\cite{SternJ17} proposes a planning compilation for learning action models from plan traces following the {\em finite domain} representation for the state variables.
-This is a theoretical study on the boundaries of the learned models and no experimental results are reported.
-
-We also introduced the {\em precision} and {\em recall} metrics, widely used in ML, for evaluating the learned action models with respect to a given reference model.
-These two metrics measure the soundness and completeness of the learned models and facilitate the identification of model flaws.
-
-When example plans are available, we can compute accurate action models from small sets of learning examples in little computation time, less than a second.
+To the best of our knowledge, this is the first work on learning and evaluating \strips\ action models from state observations, using exclusively an {\em off-the-shelf} classical planner, and tested over a wide range of different domains.
 In many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
-With this regard, we extended our approach for learning also from state observations as it broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}.
+With this regard, our approach can learn from state observations as it broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}.
 When action plans are not available, our approach still produces action models that are compliant with the input information.
@@
 
-We also introduced a semantic method for evaluating the learned \strips\ action models with respect to observations of plan executions.
+When example plans are available, our approach is also able to compute accurate action models from small sets of learning examples.
+Our experimental results report that \FAMA achieves {\em precision} and {\em recall} scores that are over 0.7 for ten out of the fourteen IPC domains, using only two plan traces (with ten actions each) and in little computation time (less than a second).
+Our experimental evaluation also shows that, in the particular setting of minimal input data, \FAMA outperforms \ARMS.
+
+On the other hand, this work also introduced a semantic version of the {\em precision} and {\em recall} metrics.
+These two metrics, widely used in ML, separately measure the soundness and completeness of the learned models, which facilitate the identification of model flaws.
+Our brand new semantic version of {\em precision} and {\em recall} do not require a GTM and evaluate \strips\ action models with respect to a given set of observations.
+In addition, when no reformulation occurs, these new metrics behave similarly to their syntactic counterparts.
+
 Generating {\em informative} examples for learning planning action models is still an open issue.
@@
 
-In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema.
+%In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema.
 However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}

### 2018-08-01 doc/aij18/introduction.tex
small updates added to the introduction

@@
 There is common agreement in the planning community that the unavailability of a complete domain model is a bottleneck in the applicability of planning technology to many real-world domains~\cite{kambhampati:modellite:AAAI2007}.
-Motivated by the difficulty and cost of crafting action models, research in action-model learning has seen huge advances since the emergence of the pioneer learning system ARMS~\cite{yang2007learning}.
-Ever since, we have seen models able to learn action models with quantifiers~\cite{AmirC08,ZhuoYHL10}, from noisy actions or noisy states~\cite{zhuo2013action,MouraoZPS12}, from null state information~\cite{cresswell2013acquiring}, from incomplete domain models~\cite{ZhuoNK13,ZhuoK17} and many more.
+Motivated by the difficulty and cost of crafting action models, research in action-model learning has seen huge advances.
+Since the emergence of pioneer learning systems like ARMS~\cite{yang2007learning}, we have seen systems able to learn action models with quantifiers~\cite{AmirC08,ZhuoYHL10}, from noisy actions or noisy states~\cite{zhuo2013action,MouraoZPS12}, from null state information~\cite{cresswell2013acquiring}, from incomplete domain models~\cite{ZhuoNK13,ZhuoK17} and many more.
 
-The primary underlying motivation for acquiring planning action models is to be able to solve model-based planning tasks afterwards.
-A learning system receives as an input observations of the agent's plan execution and outputs an abstract version of the capability model that reflects the physics of the real-world domain.
+A system for learning planning action models, receives as an input observations of the agent's plan execution and outputs an abstract version of the capability model that reflects the physics of the real-world domain.
+The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards.
 Recently, model-based representations are also becoming popular in explainable AI planning as they form a common basis for communicating with users and facilitate the generation of transparent and explainable decisions~\cite{FoxLM17} as well as explanations in terms of the differences with a human mental model~\cite{ChakrabortiSK18}.
 
-Two different types of data sources are generally identified in the generation of explanations~\cite{Sheh17,Roberts18}: (1) data that embody enough information to fully trace the steps of the decision-making process of the agent and so accurate and detailed explanations can be generated; (2) data that do not accurately represent the decision-making process because the observed external behaviour of the agent responds to a black-box model.
+Two different types of data sources are generally identified in the generation of explanations~\cite{Sheh17,Roberts18}: (1) data that embody enough information to fully trace the steps of the decision-making process of the agent and so accurate and detailed explanations can be generated; (2) data that do not capture the decision-making process because the observed external behaviour of the agent responds to a black-box.
 A similar distinction can be established with the type of observations that algorithms use to learn planning action models.
 Up to date, all existing learning approaches assume that a given input plan trace contains a fully observed sequence of the executed actions.
-This seriously restricts the applicability of the learning approach to contexts in which the behaviour of the agent is fully observable and a human annotator correctly labels the observed actions.
+This seriously restricts the applicability of the learning approach to contexts in which the behaviour of the agent is fully observable and a human annotator correctly labels the executed actions.
 
-In this paper we claim that a planning model is learnable even though an accurate representation of the agent's behaviour is not available.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behaviour is not available.
 Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimum input knowledge.
-While current learning systems accept varying degrees of observability in the states traversed by the input observed actions, none of them allow partial observability in the sequence of actions.
+
+While current learning systems accept varying degrees of observability in the states traversed by the input observed actions, none of them allow partial observability in the sequence of executed actions.
 \FAMA, on the other hand, allows incomplete or empty action sequences.
 This way, the minimal observability case acceptable by \FAMA is when the algorithm is only fed with the input and output of a plan trace, namely the initial and final state of a planning task.
-Like many Machine Learning (ML) techniques, \FAMA is able to operate with only correct input/ouput pairs and an unknown or a partially known model of the agent.
+Like many Machine Learning (ML) techniques, \FAMA is able to operate with only input/output pairs and an unknown or a partially known model of the agent.
 Unlike ML algorithms, \FAMA requires a symbolic structured representation of the input knowledge.
@@
 A solution to this planning task is a sequence of actions that determines the preconditions and effects of the \strips\ action models.
-The construct of the planning model starts out with the input observations of the agent's behaviour that, as commented above, may comprise none of the actions executed by the agent.
+The construct of the planning model starts out with the input observations of the agent's behaviour that, as mentioned above, may comprise none of the actions executed by the agent.
 \FAMA can thus be regarded as a solving process that generates an instance of a planning model, and whose solution must be \emph{compliant} with the input plan traces.
 This behaviour largely differs from ML techniques, which aim to minimize an error function on the training data.
-Moreover, \FAMA requires far less sample data (input observations) than ML algorithms, thus alleviating the dependency on the assumption that there are enough data for learning the action models~\cite{Zhuo15}.
+Moreover, \FAMA requires far less sample data (input observations) than typical ML algorithms, thus alleviating the dependency on the assumption that there are enough data for learning the action models~\cite{Zhuo15}.
 
@@
 The most common method is to use a syntax-based evaluation that compares the learned model with a reference model.
-\FAMA instead proposes two novel semantic evaluation metrics that build upon two well-known syntactic ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate the output action models with respect to observations of plan executions.
+\FAMA instead proposes two novel semantic evaluation metrics that build upon two well-known ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate action models with respect to observations of plan executions.
 Our semantic evaluation is generally more informative that counting the number of error between two models and alleviates the limitations of a purely syntax-based assessment: (a) that the learned model is syntactically different from the reference model but semantically correct and (b) that the learned model comprises correct though unnecessary preconditions in regards to the reference model.
@@
 
-The semantic evaluation in \FAMA is carried out using the same compilation scheme for solving a learning task.
+Our semantic evaluation method is built on the same compilation scheme for solving a learning task.
 In particular, \FAMA also accepts as an input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}, alongside the observations.
 In this case, \FAMA returns a model that follows the input model and is compliant with the observations.
-Following this behaviour, we design an adjustment mechanism that serves to correct the input model to the output model, which in turn defines an assessment of the accuracy with which the input model explains the observations.
-Interpreting the adjustment value as a distance-based concept between two models could be exploited in model reconciliation~\cite{KulkarniCZVZK16}.
+We designed an {\em edition} mechanism that serves to correct the input model to the output model, which in turn defines an assessment of the accuracy with which the input model explains the observations.
+Interpreting {\em edition} as a distance-based concept between two models could be exploited in model reconciliation~\cite{KulkarniCZVZK16}.
 
@@
 Ultimately, \FAMA adopts certain ML-like characteristics, like the ability to learn from data that do not explicitly exhibit the agent¡s behaviour or the capacity of correcting an initially given action model.
-
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, this paper presents \FAMA, an innovative approach for learning \strips\ action models defined as a classical planning compilation.
-A first description of the compilation scheme previously appeared in our previous conference paper~\cite{aineto2018learning}.
-\FAMA brings the following contributions over the first version of the compilation:
-
+A first description of the \FAMA compilation scheme previously appeared in our previous conference paper~\cite{aineto2018learning}.
+This paper brings the following contributions over the first version of the compilation:
 \begin{itemize}

### 2018-08-02 doc/aij18/aij18.tex
Added abstract

@@
 \begin{abstract}
-  \textcolor[rgb]{1.00,0.00,0.00}{This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this learning task into classical planning.
-The compilation approach is flexible to various amount and kind of available input knowledge; learning examples can range from plans (with their corresponding initial state) to sequences of state observations or even just a set of initial and final states (where no intermediate action or state is known).
-The compilation accepts also partially specified action models and can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified.
-On the other hand, the compilation is extensible to assess how well a given \strips\ action model matches the observation of a plan execution.
-This extension allows us to evaluate the quality of the learned models with respect to the actual models but also, with respect to a {\em test set} of observations of plan executions.
-The performance of our compilation approach is evaluated learning action models, for a wide range of classical planning domains from the International Planning Competition (IPC), and following these two evaluation approaches.}
+  This paper presents \FAMA, a novel approach for learning \strips\ action models from observations of plan executions that compiles the learning task into a classical planning task.
+Unlike all existing learning systems, \FAMA is able to learn when the actions of the plan executions are partially or totally unobservable and information on intermediate states is partially provided.
+This flexibility makes \FAMA an ideal learning approach in domains where only sensoring data are accessible.
+Additionally, we leverage the compilation scheme and extend it to come up with an evaluation method that allows us to assess the quality of a learned model syntactically, that is, with respect to the actual model; and, semantically, that is, with respect to a set of observations of plan executions.
+We also show that the extended compilation scheme can be used to lay the foundations of a framework for action model comparison.
+\FAMA is exhaustively evaluated over a wide range of IPC domains and its performance is compared to \ARMS, a state-of-the-art benchmark in action model learning.
 \end{abstract}

### 2018-08-02 doc/aij18/introduction.tex
Minor changes to introduction

@@
 
-A system for learning planning action models, receives as an input observations of the agent's plan execution and outputs an abstract version of the capability model that reflects the physics of the real-world domain.
+A system for learning planning action models, receives as input observations of the agent's plan execution and outputs an abstract version of the capability model that reflects the physics of the real-world domain.
 The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards.
@@
 
-Two different types of data sources are generally identified in the generation of explanations~\cite{Sheh17,Roberts18}: (1) data that embody enough information to fully trace the steps of the decision-making process of the agent and so accurate and detailed explanations can be generated; (2) data that do not capture the decision-making process because the observed external behaviour of the agent responds to a black-box.
+Two different types of data sources are generally identified in the generation of explanations~\cite{Sheh17,Roberts18}: (1) data that embody enough information to fully trace the steps of the decision-making process of the agent and so that accurate and detailed explanations can be generated; (2) data that do not capture the decision-making process because the observed external behaviour of the agent responds to a black-box.
 A similar distinction can be established with the type of observations that algorithms use to learn planning action models.
 Up to date, all existing learning approaches assume that a given input plan trace contains a fully observed sequence of the executed actions.
-This seriously restricts the applicability of the learning approach to contexts in which the behaviour of the agent is fully observable and a human annotator correctly labels the executed actions.
+This heavily restricts the applicability of the learning approach to contexts in which the behaviour of the agent is fully observable and a human annotator correctly labels the executed actions.
 
@@
 \FAMA instead proposes two novel semantic evaluation metrics that build upon two well-known ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate action models with respect to observations of plan executions.
-Our semantic evaluation is generally more informative that counting the number of error between two models and alleviates the limitations of a purely syntax-based assessment: (a) that the learned model is syntactically different from the reference model but semantically correct and (b) that the learned model comprises correct though unnecessary preconditions in regards to the reference model.
+Our semantic evaluation is generally more informative than counting the number of errors between two models and alleviates the limitations of a purely syntax-based assessment: (a) that the learned model is syntactically different from the reference model but semantically correct and (b) that the learned model comprises correct though unnecessary preconditions in regards to the reference model.
 This latter issue is concerned with the qualification problem, the actual impossibility of listing all the preconditions required for a real world action to have its intended effects~\cite{GinsbergS88}.
@@
 Our semantic evaluation method is built on the same compilation scheme for solving a learning task.
-In particular, \FAMA also accepts as an input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}, alongside the observations.
+In particular, \FAMA also accepts as input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}, alongside the observations.
 In this case, \FAMA returns a model that follows the input model and is compliant with the observations.
@@
 Ultimately, \FAMA adopts certain ML-like characteristics, like the ability to learn from data that do not explicitly exhibit the agent¡s behaviour or the capacity of correcting an initially given action model.
-A first description of the \FAMA compilation scheme previously appeared in our previous conference paper~\cite{aineto2018learning}.
+A first description of the \FAMA compilation scheme already appeared in our previous conference paper~\cite{aineto2018learning}.
 This paper brings the following contributions over the first version of the compilation:
@@
 \item An exhaustive empirical evaluation over 14 domains from the International Planning Competitions (IPCs).
-We include an analysis of the impact that the size of the input knowledge has in the performance of \FAMA, a comparison with \ARMS and a detailed experimentation when \FAMA is executed with minimal input knowledge.
+We include an analysis of the impact that the size of the input knowledge has in the performance of \FAMA, a comparison with \ARMS, and a detailed experimentation when \FAMA is executed with minimal input knowledge.
 \end{itemize}

### 2018-08-02 doc/aij18/experiments.tex
Minor changes to experiments

@@
 \subsection{Setup}
-We evaluate \FAMA on fourteen IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, all taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-Table \ref{tab:domain_features} compiles the features of the domains, all these features affect the size of our compilation.
+We evaluate \FAMA on 14 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, all taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+Table \ref{tab:domain_features} compiles the features of the tested domains that affect the size of our compilation.
 From left to right columns report, for each domain, its number of actions, number of predicates, maximum arity of the actions, and maximum arity of the predicates.
@@
 The {\em precision} and {\em recall} values are now lower and they drop as the amount of input knowledge increases.
-For the PSPACE-complete scenario, {\em precision} and {\em recall} scores have to be taken with a grain of salt because {\bf pure syntax-based metrics are not informative when reformulations appear}.
-This is the case of such an under-constrained learning task and we provide experimental evidence of this in Section \ref{minimal}.
-Further, the score drop reflects an increasing number of timeouts meaning that, no solution is found for several domains within the given time-bound (the worst case complexity of this scenario is larger).
+For the PSPACE-complete scenario, {\em precision} and {\em recall} scores have to be taken with a grain of salt because {\bf pure syntax-based metrics are not informative when reformulations appear} (we provide experimental evidence of this in Section \ref{minimal}).
+Further, the score drop reflects an increasing number of timeouts meaning that no solution is found for several domains within the given time-bound, which hints to the higher worst case complexity of this scenario.
 Computation time (Figure \ref{fig:pspace_time}) is also higher explained again by this larger number of timeouts.
@@
 Now we compare the performance of \FAMA to that of \ARMS.
-\ARMS, as any other previous approaches for learning planning models cannot handle the PSPACE-complete scenario.
-On the other hand, \ARMS assumes input plan traces with \FO action sequences and \NO state trajectories so here, we will also adopt this assumption.
+\ARMS, as all previous approaches, cannot handle the PSPACE-complete scenario since it requires input plan traces with \FO action sequences so, in this experiment, we will also adopt this assumption.
 
@@
 A similar trend is drawn for {\em recall} (Figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our learned models are more complete.
-These results evidence the \FAMA capacity to learn better models when few plan traces are available but do not imply that our approach is overall better for NP-complete scenarios (Figure~\ref{fig:np_time} showed that \FAMA cannot handle so much input as \ARMS).
+These results evidence the \FAMA capacity to learn better models when few plan traces are available but do not imply that our approach is overall better for NP-complete scenarios (Figure~\ref{fig:np_time} showed that \FAMA cannot handle as much input as \ARMS).
 
@@
 In Table \ref{tab:results_minimum_30_30}, which collects the result of the PSPACE case study with 30\% observability, the scores for some domains are missing.
-This is the case for {\em floor-tile}, and {\em parking} both considered \emph{puzzle-like} domains and that put a strain on the planners.
+This is the case for {\em floor-tile}, and {\em parking} both considered \emph{puzzle-like} domains, a trait known for putting a strain on the planners.
 Interestingly enough, {\em hanoi} also qualifies as a \emph{puzzle-like} domain, and we can see how this is reflected on the computation time.
-With regards to the learned models, we find that they retain similar level of soundness, while their completeness is lower than in the previous case study.
+With regards to the learned models, we find that they retain similar levels of soundness, while their completeness is lower than in the previous case study.
 This is specially the case for the preconditions, which have gone from 0.94 recall to 0.54.
-%This results are difficult to explain, given that in this case study the intermediate states contain more information, but
 The explanation for these scores is that input actions are strong constraints that play a key role on how close the learned model are with respect to the GTM.
-The more these actions are missing the more present is reformulation and pure syntax-based metrics become less relevant.
+The more these actions are missing the more likely it is for reformulations to occur so pure syntax-based metrics become less relevant.
 
@@
 The effect of missing input actions is more evident in our last case study with \NO action sequences and state trajectories (Table \ref{tab:results_minimum_0_0}).
-In this case we find solutions for all domains.
-Search is even less constrained so there are far more possible solutions for this learning task.
-This wider range of solutions is also evidenced by the kind of the learned models, which in spite of being compliant with the input data, are further from the GTM.
-Both {\em precision} and {\em recall} decreased to 0.6 and 0.5, respectively.
-Syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the \emph{reformulations} that frequently occur in these circumstances.
+Contrary to what might be expected by looking at the previous table, in this case we are able to find solutions for all domains.
+The reason for this is that the search is less constrained so there are far more possible solutions for this learning task.
+This wider range of solutions is also evidenced by the learned models, which in spite of being compliant, are more removed from the GTM.
+In the table we find that both precision and recall have seen their scores decreased to 0.6 and 0.5, respectively.
+We argue, however, that syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the \emph{reformulations} that frequently occur in these circumstances.
 
@@
 We refer to the first difference as \emph{role swapping} and it happens when there are missing actions in the plan trace.
-If no actions are present in the input traces, the name of actions becomes meanless (actions are effectively anonymous), meaning that they can interchange their behavior with any other comparable action model.
-The second issue happens when there are missing states in the input traces.
+If no actions are present in the input traces, the names of actions become meaningless (actions are effectively anonymous), meaning that they can interchange their behavior with any other comparable action model.
+The second issue happens when there are missing states in the input traces, since \emph{macro-actions} apply more than one action in a single step so some intermediate states are skipped.
 
@@
 This means that the semantic version is more lenient towards extra preconditions or effects.
-This matches what we have observed during the experiments for instance, the learned model for the {\tt move} action of the {\em hanoi} domain, specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
+This matches what we have observed during the experiments.
+For instance, the learned model for the {\tt move} action of the {\em hanoi} domain, specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
 This learned model is semantically correct but sintactically different from the GTM and hence, penalized by the {\em precision} metric.
 
-Finally Table \ref{tab:metric_comparison_0_0} details the results of the \NO/\NO case study.
+Finally, Table \ref{tab:metric_comparison_0_0} details the results of the \NO/\NO case study.
 Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger here.
 Comparing the scores of both versions, we find that learned models that achieved mediocre scores when using the GTM as reference, are in fact reasonable sound and complete, reaching and overall score of 0.91 for both {\em sem-Precision} and {\em sem-Recall}.
-This means that the learned models by our approach, despite syntactically different from the GTM, require very few editions to explain the testing traces given as input.
+This means that the models learned by our approach, despite syntactically different from the GTM, require very few editions to explain the testing traces given as input.
 
@@
 The second conclusion is that {\em sem-Precision} and {\em sem-Recall} are better suited to evaluate reformulated models than the original syntactic metrics, since they contemplate valid solutions outside the GTM that successfully explain the given input data.
-%Since we are using {\sc Madagascar}, a satisfying planner, the scores for sem-Precision and sem-Recall are approximated.
-
-%Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
-The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
-Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
-Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\mathcal{M}$).
-{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
-
-%Here we evaluate our approach with learning tasks of the kind $\Lambda=\tup{\mathcal{M},\Psi,\mathcal{T}}$, where the action of the executed plans are not available but the initial and goal states are known.
-When input plans are not available, the planner must not only compute the action models but also the plans that satisfy the input observations.
-Table~\ref{tab:results_states} and ~\ref{tab:time_states} summarize the results obtained for this using static predicates and partially specified models.
-Values for the {\em Zenotravel} and {\em Grid} domains are not reported because {\sc Madagascar} was not able to solve the corresponding planning tasks within a 1000 sec.
-time bound.
-The values of \emph{precision} and \emph{recall} are significantly lower than in Table ~\ref{tab:results_plans}.
-Given that the learning hypothesis space is now fairly under-constrained, actions can be reformulated and still be compliant with the inputs (e.g.
-the {\em blocksworld} operator {\small\tt stack} can be {\em learned} with the preconditions and effects of the {\small\tt unstack} operator and vice versa).
-We tried to minimize this effect with the additional input knowledge (static predicates and partially specified action models) and yet the results are below the scores obtained when learning from labeled plans.
-
-%To give an insight of the actual quality of the learned models, we defined a method for computing {\em Precision} and {\em Recall} that is robust to the mentioned model {\em reformulations}.
-Precision and recall are often combined using the {\em harmonic mean}.
-This expression, called the {\em F-measure} or the balanced {\em F-score}, is defined as $F=2\times\frac{Precision\times Recall}{Precision+Recall}$.
-Given the learned action model $\mathcal{M}$ and the reference action model $\mathcal{M}^*$, the bijective function $f_{P\&R}:\mathcal{M} \mapsto \mathcal{M}^*$ is the mapping between the learned and the reference model that maximizes the accumulated {\em F-measure} (considering swaps in the actions with matching headers or parameters with matching types).
-
-%Table~\ref{fig:observationsmap} shows that significantly higher values of {\em precision} and {\em recall} are reported when a learned action schema, $\xi\in\mathcal{M}$, is compared to its corresponding reference schema given by the $f_{P\&R}$ mapping ($f_{P\&R}(\xi)\in \mathcal{M}^*$).
-The {\em blocksworld} and {\em gripper} domains are perfectly learned from only 25 state observations.
-These results evidence that in all of the evaluated domains, except for {\em ferry} and {\em satellite}, the learning task swaps the roles of some actions (or parameters) with respect to their role in the reference model.
-
-%The learning scores of several domains in Table~\ref{fig:observationsmap} are above the ones reported in Table~\ref{fig:observationsnomap}.
-The reason lies in the particular observations comprised by the test sets.
-As an example, in the {\em driverlog} domain, the action schema {\small \tt disembark-truck} is missing from the learned model because this action is never induced from the observations in the training set; that is, such action never appears in the corresponding \emph{unobserved} plan.
-The same happens with the {\small \tt paint-down} action of the {\em floortile} domain or {\small \tt move-curb-to-curb} in the {\em parking} domain.
-Interestingly, these actions do not appear either in the test sets and so the learned action models are not penalized in Table~\ref{fig:observationstest}.
-Generating {\em informative} and {\em representative} observations for learning planning action models is an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions, often, with a low probability of being chosen by chance~\cite{fern2004learning}.

### 2018-08-04 introduction.tex
Add files via upload
final version

--- 2018-08-02 doc/aij18/introduction.tex
@@
 
-A system for learning planning action models, receives as input observations of the agent's plan execution and outputs an abstract version of the capability model that reflects the physics of the real-world domain.
+A system for learning planning action models receives as an input observations of the agent's plan execution and outputs an abstract version of the capability model that reflects the physics of the real-world domain.
 The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards.
@@
 
-Two different types of data sources are generally identified in the generation of explanations~\cite{Sheh17,Roberts18}: (1) data that embody enough information to fully trace the steps of the decision-making process of the agent and so that accurate and detailed explanations can be generated; (2) data that do not capture the decision-making process because the observed external behaviour of the agent responds to a black-box.
+Two different types of data are generally identified in the generation of explanations~\cite{Sheh17,Roberts18}: (1) data that embody enough information to fully trace the steps of the decision-making process of the agent and therefore accurate and detailed explanations can be generated; (2) data that do not accurately represent the decision-making process because the observed external behaviour of the agent responds to a black-box model.
 A similar distinction can be established with the type of observations that algorithms use to learn planning action models.
@@
 The most common method is to use a syntax-based evaluation that compares the learned model with a reference model.
-\FAMA instead proposes two novel semantic evaluation metrics that build upon two well-known ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate action models with respect to observations of plan executions.
-Our semantic evaluation is generally more informative than counting the number of errors between two models and alleviates the limitations of a purely syntax-based assessment: (a) that the learned model is syntactically different from the reference model but semantically correct and (b) that the learned model comprises correct though unnecessary preconditions in regards to the reference model.
-This latter issue is concerned with the qualification problem, the actual impossibility of listing all the preconditions required for a real world action to have its intended effects~\cite{GinsbergS88}.
+\FAMA instead proposes two novel semantic evaluation metrics that build upon two well-known ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate the learned action models with respect to observations of plan executions.
+Our semantic evaluation is generally more informative than counting the number of errors between two models and alleviates two important limitations of a purely syntax-based assessment: (a) that the learned model is syntactically different from the reference model but semantically correct and (b) that the learned model comprises correct though unnecessary preconditions in regards to the reference model.
+This latter issue is concerned with the qualification problem, which is defined as the actual impossibility of listing all the preconditions required for a real world action to have its intended effects~\cite{GinsbergS88}.
 
 Our semantic evaluation method is built on the same compilation scheme for solving a learning task.
-In particular, \FAMA also accepts as input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}, alongside the observations.
+In particular, \FAMA also accepts as an input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}, alongside the observations.
 In this case, \FAMA returns a model that follows the input model and is compliant with the observations.
 We designed an {\em edition} mechanism that serves to correct the input model to the output model, which in turn defines an assessment of the accuracy with which the input model explains the observations.
-Interpreting {\em edition} as a distance-based concept between two models could be exploited in model reconciliation~\cite{KulkarniCZVZK16}.
+Interpreting the edition measure as a distance-based concept between two models could be exploited in model reconciliation~\cite{KulkarniCZVZK16}.
 
@@
 
-All in all, \FAMA is a planning-based solving scheme that outputs a \strips\ action model using an automatically built planning model from minimal input knowledge.
-Unlike extensive data-ML approaches, \FAMA only requires a small amount of input plan traces.
+All in all, \FAMA is a planning-based solving scheme that outputs a \strips\ action model using a planning model that is automatically built from minimal input knowledge.
+Unlike extensive-data ML approaches, \FAMA only requires a small amount of input plan traces.
 Unlike most relevant action-model learning algorithms, \FAMA does not require the traces to contain any observed action executed by the agent.
-Ultimately, \FAMA adopts certain ML-like characteristics, like the ability to learn from data that do not explicitly exhibit the agent¡s behaviour or the capacity of correcting an initially given action model.
+Ultimately, \FAMA adopts certain ML-like characteristics, like the ability to learn from data that do not explicitly exhibit the agent's behaviour or the capacity of correcting an initially given action model.
 A first description of the \FAMA compilation scheme already appeared in our previous conference paper~\cite{aineto2018learning}.
 This paper brings the following contributions over the first version of the compilation:
+
 \begin{itemize}

### 2018-08-04 doc/aij18/evaluation.tex
Add files via upload
añadida una etiqueta

@@
 
-\begin{mydefinition}
+\begin{mydefinition} \label{compliant}
   Given a set of action models $\mathcal{M}$, and all the sets of action models $\mathcal{M'}$ able to explain the plan traces $\mathcal{T}$.

### 2018-08-05 doc/aij18/motivation.tex
corregido #domains y minor comment

@@
 This is an important advantage, particularly in domains in which it is costly or impossible to obtain a significant number of training samples.
-Unlike \CAMA, our approach does not require human intervention to label samples as it is able to learn from very small datasets.
+Unlike \CAMA, our approach does not require human intervention to label samples as it is able to learn from empty sequences of observed actions.
 
-Finally, as it will be shown in section \ref{sec:experiments}, \FAMA is exhaustively evaluated, syntactically and semantically, over a wide range of domains (13 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner so it can benefit straightforward from the last advances in classical planning.
+%very small datasets.
+
+Finally, as it will be shown in section \ref{sec:experiments}, \FAMA is exhaustively evaluated, syntactically and semantically, over a wide range of domains (14 domains compared to the scarce number of tested domains of the rest of the approaches in Table \ref{table:models_comparison2}) and uses exclusively an \emph{off-the shelf} classical planner so it can benefit straightforward from the last advances in classical planning.

### 2018-08-06 doc/aij18/aij18.tex
final conclusions

@@
 \label{sec:conclusions}
-We presented a novel approach for learning \strips\ action models from examples using classical planning.
-The approach is flexible to various amount and kind of input knowledge and accepts partially specified action models.
+
+In this paper we have presented \FAMA, an approach for learning \strips\ action models based on the compilation of the learning task into a planning task.
+The distinctive characteristic of \FAMA over other state-of-the-art approaches is its ability to learn from minimal input knowledge and, more particularly, from minimal observability plan traces with no observed actions and no observed intermediate states.
+By lightening the input constraints, \FAMA opens up the way for action model learning to operate on real-world problems, as opposite to current approaches where the heavy input restrictions have limited their applicability to synthetic benchmarks.
+Our approach is thus very well suited for learning action models in data-based applications where the only observable information is a possibly incomplete sequence of partially observed states.
+
+Besides its capacity of working with minimal observability, \FAMA is also able to learn from very small amounts of input knowledge, a clear advantage in domains where obtaining enough training samples is difficult or costly.
+While the experimental evaluation shows in general an exponential increase of the computation time as the number of training traces augments, \FAMA is able to learn action models more accurate than those of \ARMS with very limited input knowledge.
 Unlike extensive-data ML approaches, our work explores an alternative research direction that exploits logic reasoning to learn sound models from minimal input knowledge.
-Remarkably, the action models of the {\em gripper}, {\em npuzzle} or {\em visitall} domains were perfectly learned from only observations of the initial and final states.
 
-To the best of our knowledge, this is the first work on learning and evaluating \strips\ action models from state observations, using exclusively an {\em off-the-shelf} classical planner, and tested over a wide range of different domains.
-In many applications, the actual actions executed by the observed agent are not available but, instead, the resulting states can be observed.
-With this regard, our approach can learn from state observations as it broadens the range of application to external observers and facilitates the representation of imperfect observability, as shown in plan recognition \cite{SohrabiRU16}, as well as learning from unstructured data, like state images \cite{AsaiF18}.
-When action plans are not available, our approach still produces action models that are compliant with the input information.
+This great flexibility of \FAMA to different amounts of minimal input knowledge impacts the complexity of the planning tasks.
+When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, the planning task to be solved becomes a PSPACE-complete scenario.
+This is one of the reasons that justifies our compilation-to-planning scheme in contrast to the most extended SAT-solving scheme used in many existing learning systems.
+Additionally, our planning-based solving scheme allows us to leverage {\em off-the-shelf} classical planners and benefit from the multiple advances in classical planning research.
+
+A key contribution of this work is the proposal of two novel metrics to semantically evaluate the learned action models.
+These two metrics build upon the well-known syntax-based metrics \emph{precision} and \emph{recall}.
+Our semantic evaluation mitigates the common issue known as \emph{reformulation} that appears when training sets of minimal observability are used.
+Due to the lack of observable information, \FAMA can learn semantically valid models that are syntactically different from the reference model.
+The application of the semantic-based precision and recall allows us to assess the validity of the learned models even in domains where a reference model is not available.
+%Remarkably, the action models of the {\em gripper-strips}, {\em n-puzzle} and {\em grid-visit-all} domains were perfectly learned from only observations of the initial and final states.
+
+We highlight that the semantic evaluation of \FAMA is done via the same compilation-to-planning scheme that we use to learn the action models.
+Given that the input of our learning task definition accepts an initial action model of the agent's behaviour, either complete or partially specified, this solving scheme is exploitable to computing a model that follows the initial model and is compliant with a test set of plan traces.
+In other words, \FAMA is applicable in model reconciliation tasks by defining a distance metric that measures how close the two models are~\cite{KulkarniCZVZK16}.
+
+More importantly, \FAMA is applicable not only in environments where the domain model is unknown but also in environments where the executable actions are unknown as well.
+This ability  broadens the range of application to goal reasoning tasks such as plan recognition under imperfect observability \cite{SohrabiRU16}, planning for transparency~\cite{MacNallyLRP18} or counterplanning~\cite{PozancoEFB18}.
+The application of \FAMA to these tasks offers a plan-based solution where the assumption of a known domain model can be removed.
+In other words, \FAMA opens up the way towards domain-free goal reasoning.
+
+Finally, we would like to add a note on the open issue of generating {\em informative} plan traces for learning planning action models.
+Planning actions include preconditions that are only satisfied by specific sequences of actions which have low probability of being chosen by chance~\cite{fern2004learning}.
+The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques that enable to autonomously collect informative learning examples.
+The combination of such exploration techniques with our learning approach is an appealing research direction towards the bootstrapping of planning action models.
+
+%When action plans are not available, our approach still produces action models that are compliant with the input information.
 In this case, since learning is not constrained by actions, operators can be reformulated changing their semantics, in which case the comparison with a reference model turns out to be tricky.
 
-When example plans are available, our approach is also able to compute accurate action models from small sets of learning examples.
+%When example plans are available, our approach is also able to compute accurate action models from small sets of learning examples.
 Our experimental results report that \FAMA achieves {\em precision} and {\em recall} scores that are over 0.7 for ten out of the fourteen IPC domains, using only two plan traces (with ten actions each) and in little computation time (less than a second).
@@
 
-On the other hand, this work also introduced a semantic version of the {\em precision} and {\em recall} metrics.
+%On the other hand, this work also introduced a semantic version of the {\em precision} and {\em recall} metrics.
 These two metrics, widely used in ML, separately measure the soundness and completeness of the learned models, which facilitate the identification of model flaws.
@@
 In addition, when no reformulation occurs, these new metrics behave similarly to their syntactic counterparts.
-
-Generating {\em informative} examples for learning planning action models is still an open issue.
-Planning actions include preconditions that are only satisfied by specific sequences of actions which have low probability of being chosen by chance~\cite{fern2004learning}.
-The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques that enable to autonomously collect informative learning examples.
-The combination of such exploration techniques with our learning approach is an appealing research direction that opens up the door to the bootstrapping of planning action models.
-
-%In theory, we could implement a third edit operation for {\em substituting} a fluent from a given operator schema.
-However, and with the aim of keeping a tractable branching factor of the planning instances that result from our compilations, we only implement {\em deletion} and {\em insertion}
 

### 2018-08-16 benchmarks/aaai19/automata/gen-automata-domain.py
first steps for aaai19 experiments

@@ -5,7 +5,7 @@ REGULAR = 0
 STACK = 1
 TURING = 2
 
-def baseb(n, b):
+def decodeID(n, b):
     e = n//b
     q = n%b
     if n == 0:
@@ -13,7 +13,7 @@ def baseb(n, b):
     elif e == 0:
         return str(q)
     else:
-        return baseb(e, b) + str(q)
+        return decodeID(e, b) + str(q)
 
 
 # **************************************#
@@ -30,20 +30,18 @@ except:
     print sys.argv[0] + " <nStates> <nObservations> <kind (0 regular, 1 stack, 2 Turing Machine)> <idAutomata [0-states^(states*observations))>"
     sys.exit(-1)
 
-    
+str_ID = ""    
 if nKind == REGULAR:
     str_automata="S"+str(nStates)+"-O"+str(nObs)+"-REGULAR"+"-DET-"+str(idAutomata)
-    MAX_ID = int(math.pow(nStates,(nStates*nObs)))-1
-    str_MAX = baseb(MAX_ID,nStates)
-    str_ID = baseb(idAutomata,nStates)
-
+    str_ID = decodeID(idAutomata,nStates)
+    
 if nKind == TURING:
     str_automata="S"+str(nStates)+"-O"+str(nObs)+"-TURING"+"-DET-"+str(idAutomata)
-    unit =  max(nStates,nObs,2)
-    MAX_ID=int(math.pow(unit*unit*unit,(nStates*nObs)))-1
-    str_MAX = baseb(MAX_ID,unit*unit*unit)
-    str_ID = baseb(idAutomata,unit*unit*unit)
+    str_ID = decodeID(idAutomata,nStates*nObs*2)
 
+str_MAX = "0"*nStates*nObs    
+for aux in range(len(str_MAX)-len(str_ID)):
+    str_ID = "0" + str_ID    
     
 str_out=""
 str_out = str_out +  "(define (domain S"+str(nStates)+"-O"+str(nObs)+")    ;;; "+str_automata+"\n"
@@ -51,16 +49,17 @@ str_out = str_out +  "  (:requirements :strips)\n"
 str_out = str_out +  "  (:predicates (head ?x)\n"
 str_out = str_out +  "               (next ?x1 ?x2)"
 str_out = str_out + "\n              "
-for i in range(0,nStates):
+for i in range(nStates):
    str_istate="S"+str(i)
    str_out = str_out +  " (state"+str_istate+")"
[…]

### 2018-08-16 benchmarks/aaai19/automata/gen-automata-problem.py

@@ -2,7 +2,6 @@
 import sys, math
 
 
-
 # **************************************#
 # MAIN
 # **************************************#
@@ -30,11 +29,11 @@ str_istate="S"+str(0)
 str_ihead="t"+str(0)
 str_out = str_out +  "  (:init (state"+str_istate+") (head "+str_ihead+")\n        "
 
-for n in range(0,len(str_itape)-1):
+for n in range(len(str_itape)-1):
    str_out = str_out +  " (next t"+str(n)+" t"+str(n+1)+")"
 str_out = str_out +  "\n        "
 
-for n in range(0,len(str_itape)):
+for n in range(len(str_itape)):
    str_out = str_out +  " (symbolO"+str_itape[n]+" t"+str(n)+")"   
 str_out = str_out +  " )\n"
 

### 2018-08-16 doc/aij18/experiments.tex

@@
 
-This section evaluates different aspects of \FAMA that range, from CPU-time performance to the quality of the learned models.
+This section presents several experiments to evaluate the performance of \FAMA and the quality of the learned models.
 Whenever applicable, we will consider both NP-complete and PSPACE-complete scenarios to draw conclusions at both levels of worst case complexity.
-We will also devote one experiment to show the suitability of the brand new evaluation metrics introduced in this paper.
+We also devote one experiment to show the suitability of the novel evaluation metrics introduced in section \ref{sec:evaluation}.
 
 \subsection{Setup}
+
 We evaluate \FAMA on 14 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, all taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-Table \ref{tab:domain_features} compiles the features of the tested domains that affect the size of our compilation.
-From left to right columns report, for each domain, its number of actions, number of predicates, maximum arity of the actions, and maximum arity of the predicates.
-%and hence, have an impact on the complexity of the learning task.
-As can be seen in table \ref{tab:domain_features}, the selected domains range from, such as \emph{blocks} and \emph{visitall}, to complex ones, like \emph{floortile} and \emph{satellite}, with higher number of actions and predicates and with higher arity.
+Table \ref{tab:domain_features} presents the features of the tested domains that affect the size of the planning task $P_\Lambda$ that results from the compilation.
+For each domain, the columns report, from left to right, the number of actions, the number of predicates, the maximum arity of the actions, and the maximum arity of the predicates.
 
-Now we explain the details of our experimental setup:
+The details of our experimental setup are the following:
+
 \begin{itemize}
 \item {\bf Plan traces}.
-For each domain, we generated 10 plan traces (with 10 actions/intermediate states each) using random walks.
-Depending on the experiment, these traces are used for training or testing purposes (more details on this are provided at the particular experiment).
+For each domain, we generated 10 plan traces (each with 10 actions and 10 intermediate states) using random walks.
+Depending on the experiment, the traces are used for training or testing purposes (more details on this issue are provided at the particular experiment).
 
 \item {\bf Planner}.
-The classical planner we use to solve the $P_\Lambda$ instances that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We use {\sc Madagascar} for two reasons:
+The classical planner we used to solve the instances of $P_\Lambda$ that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
+We used {\sc Madagascar} for two reasons:
+
   \begin{enumerate}
-  \item  Its ability to deal with planning instances populated with dead-ends~\cite{lopez2015deterministic}.
-  \item  It can leverage the horizon of the solution plan, when it is known in advance.
-When the length of plan traces is known (e.g.
-for \FO action sequences or \FO state trajectories) the horizon of the solution plan is also known.
-SAT-based planners can apply the actions for programming preconditions in parallel during a single planning step (and the same for actions programming effects) because these actions do not interact.
-Hence, we also know that the programming prefix of a solution plan can be solved in two steps, independently of the number of programming actions applied.
+  \item  its ability to deal with planning instances populated with dead-ends~\cite{lopez2015deterministic}.
+  \item  its ability to leverage the horizon of the solution plan when it is known.
+Specifically, when the length of the plan traces is known; i.e., \FO action sequences or \FO state trajectories, the horizon of the solution plan is also known, and this can be exploited by a SAT-based planner like MADAGASCAR.
+SAT-based planners are able to apply the actions for programming preconditions in parallel in a single planning step (and the same for the actions programming the effects) since these actions do not interact.
+Hence, we also know that the programming of the prefix of a solution plan is solvable in two steps, independently of the number of programming actions applied.
   \end{enumerate}
 
-\item {\bf Hardware}. All experiments are run on an Intel Core i5 3.50 GHz x 4 with 16 GB of RAM.
+\item {\bf Hardware}. All experiments were run on an Intel Core i5 3.50 GHz x 4 with 16 GB of RAM.
 
@@
 \subsection{Impact of the size of the input knowledge}
-This experiment evaluates how the size of the input knowledge affects the performance of \FAMA to:
+
+This experiment evaluates the impact of $\left|\mathcal{T}\right|$, the size of the input knowledge, in the performance of \FAMA in order to:
+
 \begin{enumerate}
-	\item Identify the minimal amount of input knowledge required by \FAMA to learn {\em sound} and {\em complete} models,
-	\item Evaluate the scalability of \FAMA with respect to the size of the input knowledge. % which, as stated in section \ref{sec:properties}, is one of the limiting factors of our approach.
+	\item Identify the minimal amount of input knowledge required by \FAMA to learn sound and complete models,
+	\item Evaluate the scalability of \FAMA with respect to the size of the input knowledge.
 \end{enumerate}
 
-The experiment analyzes the evolution of the CPU-time, as well as the {\em precision} and {\em recall} of the learned models wrt the GTM, as $|\mathcal{T}|$ increases from 1 to 10 (each trace in $\mathcal{T}$ comprises 10 actions/intermediate states).
-To keep the experiment practicable, we introduce a 1000s timeout after which the learning process is killed and $0$ scores, for both {\em precision} and {\em recall}, are given to the learned model.
+The experiment analyzes the evolution of the CPU-time and the Precision and Recall of the learned models wrt the GTM as $|\mathcal{T}|$ increases from 1 to 10 plan traces.
+To keep the experiment practicable, we introduced a 1000s timeout, after which the learning process is killed and a score of $0$ is given to both the Precision and Recall of the learned model.
 We defined two case studies:
+
 \begin{itemize}
-	\item \textbf{\FO action sequence and \PO state trajectory}: This is the typical case addressed by most of previous work, which corresponds to NP-complete scenarios.
-In this experiment we assume a degree of state observability of just 10\%, meaning that each literal of a state has a 10\% chance of being observed.
-	\item  \textbf{\NO action sequence and \NO state trajectory}: A PSPACE-complete scenario where both the input action sequence and state trajectory are {\em empty}.
-Only the initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
+	\item \textbf{\FO action sequence and \PO state trajectory}: This is the common case addressed by most of the state-of-the-art learning
+approaches, which corresponds to a NP-complete scenario.
+In this experiment we assume a degree of observability of only 10\% for the state trajectory, meaning that each literal of a state has a 10\% chance of being observed.
+	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a PSPACE-complete scenario where both the input action sequence and state trajectory are {\em empty}.
+Only the initial and final states are observed; i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
 \end{itemize}
@@
 Figures \ref{fig:np_quality} and \ref{fig:np_time} show the quality of the models and computation time, respectively, for the NP-complete scenario.
-In Figure \ref{fig:np_quality} we see precision stabilizes at 0.86 after 5 traces while recall stabilizes at 0.95 after 3 traces.
-These results show that \FAMA does, in fact, not need big amounts of input to learn sound and complete models as opposite to other approaches in the literature where models are learned using a hundred of traces (see Table \ref{table:models_comparison2}).
+In Figure \ref{fig:np_quality} we see that Precision stabilizes at 0.86 after five traces whereas Recall stabilizes at 0.95 after three traces.
+These results show that \FAMA does, in fact, not need big amounts of input knowledge to learn sound and complete models as opposite to other approaches in the literature where models are learned using around a hundred of traces (see Table \ref{table:models_comparison2}).
 
-With regards to the scalability of \FAMA, Figure \ref{fig:np_time} paints an exponential increase in computation time.
-Until an input size of 4 traces the computation time is below 1 sec, but it reaches 153 secs when the input is composed of 10 traces.
-These results match the expected performance of {\sc Madagascar}, since this planner is known to struggle with plan horizons beyond 150-200 steps (in our case 160 steps corresponds to 8 traces).
+Figure \ref{fig:np_time} displays the scalability of \FAMA.
+Interestingly, we can observe an exponential increase in computation time for input sizes beyond five traces.
+Until an input size of four traces the computation time is below 1 sec but it reaches 153 secs when the input is composed of 10 traces.
+These results match the expected performance of {\sc Madagascar} since this planner is known to struggle with plan horizons beyond 150-200 steps (in our case 160 steps corresponds to 8 traces since each trace has 10 actions and 10 intermediate states).
 
@@
 	\includegraphics[width=0.65\linewidth]{figures/input_size_100_10_precision.eps}
-	\caption{{\em Precision} and {\em recall} when learning from [1-10] plan traces with \FO action sequences and \PO state trajectories with 10\% observability.}
+	\caption{Precision and Recall when learning from [1-10] plan traces with \FO action sequences and \PO state trajectories with 10\% observability.}
 	\label{fig:np_quality}
@@
 
-\newpage
-Figure~\ref{fig:pspace_quality} reports {\em precision} and {\em recall} for the models learned by our approach in the PSPACE-complete scenario (NO action sequence and NO state trajectory).
-The {\em precision} and {\em recall} values are now lower and they drop as the amount of input knowledge increases.
-For the PSPACE-complete scenario, {\em precision} and {\em recall} scores have to be taken with a grain of salt because {\bf pure syntax-based metrics are not informative when reformulations appear} (we provide experimental evidence of this in Section \ref{minimal}).
-Further, the score drop reflects an increasing number of timeouts meaning that no solution is found for several domains within the given time-bound, which hints to the higher worst case complexity of this scenario.
-Computation time (Figure \ref{fig:pspace_time}) is also higher explained again by this larger number of timeouts.
+Figure~\ref{fig:pspace_quality} displays the quality of the learned models in the PSPACE-complete scenario.
+As expectedly, the values of Precision and Recall are lower than in the NP-complete case.
+We can also observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
+The drop in the score is caused by an increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for Precision and Recall is assigned to these experiments.
+Figure \ref{fig:pspace_time}, on the other hand, reflects that the computation time of the PSPACE-complete scenario is also higher than the NP-complete scenario, which is again explained by the large number of timeouts.
+
+The conclusions we draw from these experiments is that learning with few input samples yield action models that are fairly sound and almost totally complete in the NP-complete scenario and less accurate in the PSPACE-complete scenarios.
+In the latter, while being timeouts the main cause of the drop in the score, we must point out that pure syntax-based metrics are not adequate to evaluate such under-constrained tasks since the phenomenon of \emph{reformulation} occurs and this largely impacts the results.
+We will provide experimental evidence of this in section \ref{minimal}.
+
+%For the PSPACE-complete scenario, {\em precision} and {\em recall} scores have to be taken with a grain of salt because {\bf pure syntax-based metrics are not informative when reformulations appear} (we provide experimental evidence of this in Section \ref{minimal}).
 
@@
 	\includegraphics[width=0.65\linewidth]{figures/input_size_0_0_precision.eps}
-	\caption{{\em Precision} and {\em recall} when learning from [1-10] plan traces with \NO action sequences and \NO state trajectories.}
+	\caption{Precision and Recall when learning from [1-10] plan traces with \NO action sequences and \NO state trajectories.}
 	\label{fig:pspace_quality}
@@
 \subsection{Comparison with \ARMS}
-Now we compare the performance of \FAMA to that of \ARMS.
-\ARMS, as all previous approaches, cannot handle the PSPACE-complete scenario since it requires input plan traces with \FO action sequences so, in this experiment, we will also adopt this assumption.
 
-We define a \emph{degree of observability} $\sigma$ for the state trajectory, ranging from 0\% to 100\%, that measures the probability of observing a literal, and evaluate both approaches as $\sigma$ increases using 5 plan traces as input.
+In this section we analyze the performance of \FAMA compared to \ARMS, one of the most well-known approaches to learning planning models.
+\ARMS, as well as most of the existing current learning systems, works under the assumption of plan traces with \FO action sequences and \NO state trajectories and therefore is not able to handle the PSPACE-complete scenario.
+We will thereby restrict the experimentation to the cases manageable by \ARMS.
+
+In this experiment, we defined a \emph{degree of observability} $\sigma$ for the state trajectory, ranging from 0\% to 100\%, that measures the probability of observing a literal, and evaluated both \FAMA and \ARMS for increasing values of $\sigma$ using five traces as input knowledge.
 When $\sigma = 0$ we have a \NO state trajectory, when $\sigma=100$ we have a \FO state trajectory and all cases in-between correspond to the \PO scenario.
-Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of {\em precision} and {\em recall}.
-Horizontal axes represents the degree of observability, while vertical axes show the average {\em precision} (Figure \ref{fig:comparison_precision}) or {\em recall} (Figure \ref{fig:comparison_recall}).
-\FAMA dominates in terms of {\em precision} in all cases (the models learned by our approach are 15\% to 24\% more precise than those learned by \ARMS) except for \FO state trajectories.
-A similar trend is drawn for {\em recall} (Figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our learned models are more complete.
-These results evidence the \FAMA capacity to learn better models when few plan traces are available but do not imply that our approach is overall better for NP-complete scenarios (Figure~\ref{fig:np_time} showed that \FAMA cannot handle as much input as \ARMS).
+
+Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of Precision and Recall.
+The horizontal axes represent the degree of observability and vertical axes show the average Precision (Figure \ref{fig:comparison_precision}) and Recall (Figure \ref{fig:comparison_recall}) computed over the 14 tested domains.
+Remarkably, \FAMA dominates in terms of Precision in all cases except for the \FO state trajectories.
+Particularly, the models learned by \FAMA are between 15\% to 24\% more precise than those learned by \ARMS.
+A similar trend is observed for Recall (Figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our learned models are more complete.
+
+The results highlight that \FAMA outperforms \ARMS when very few plan traces are available.
+This by no means is conclusive that \FAMA is overall better in NP-complete scenarios but only that is able to learn better with very limited input knowledge (actually, Figure~\ref{fig:np_time} reflects the exponential behaviour of \FAMA with more than five traces).
 
@@
 	\includegraphics[width=.65\linewidth]{figures/comparison_precision.eps}
-	\caption{{\em Precision} comparison between \FAMA and \ARMS for different \emph{degrees of observability}.}
+	\caption{Precision comparison between \FAMA and \ARMS for different \emph{degrees of observability}.}
 	\label{fig:comparison_precision}
@@
 	\includegraphics[width=.65\linewidth]{figures/comparison_recall.eps}
-	\caption{{\em Recall} comparison between \FAMA and \ARMS for different \emph{degrees of observability}.}
+	\caption{Recall comparison between \FAMA and \ARMS for different \emph{degrees of observability}.}
 	\label{fig:comparison_recall}
@@
 \label{minimal}
+
 %In the previous experiments we have given a general view of the performance of \FAMA under different conditions.
 So far, experiments have shown that \FAMA is able to learn with very small amounts of input knowledge, be it due to low observability or few training samples.
-Let us take a closer look at the action models learned from minimal input knowledge.
-To that end, we will limit the input to only 2 plan traces and analyze the results under different degrees of observability.
-In this experiment we evaluate three cases of study, the first one being an NP-complete scenario and the other two being PSPACE-complete:
+
+In this section, we will take a closer look at the action models learned from minimal input knowledge.
+To that end, we will limit the input to only two plan traces and analyze the results under different degrees of observability.
+We evaluate three case studies, the first one is an NP-complete scenario and the other two are PSPACE-complete:
+
 \begin{itemize}
@@
 Results of this case study are detailed in Table \ref{tab:results_minimum_100_10}.
-	\item  \textbf{\PO action sequence and \PO state trajectory}: In this case study (results shown in Table \ref{tab:results_minimum_30_30}) we are assuming a degree of observability of 30\% for both the action sequence and state trajectory.
-	\item  \textbf{\NO action sequence and \NO state trajectory}: Both the action sequence and state trajectory are completely empty so only the initial and final states are observed, i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
+	\item  \textbf{\PO action sequence and \PO state trajectory}: In this case study we are assuming a degree of observability of 30\% for both the action sequence and state trajectory.
+Results are shown in Table \ref{tab:results_minimum_30_30}.
+	\item  \textbf{\NO action sequence and \NO state trajectory}: Both the action sequence and state trajectory are completely empty so only the initial and final states are observed; i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
 Results of this case study are reported in Table \ref{tab:results_minimum_0_0}.
@@
 All tables in this section (Tables \ref{tab:results_minimum_100_10}, \ref{tab:results_minimum_30_30} and \ref{tab:results_minimum_0_0}) follow the same structure.
-{\em Precision} ({\bf P}) and {\em Recall} ({\bf R}) scores are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), and also globally ({\bf Global}).
+Precision ({\bf P}) and Recall ({\bf R}) scores are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), and also globally ({\bf Global}).
 The last column reports the computation time (in seconds) needed to obtain the learned models.
-Missing values in the tables (reported as -) correspond to domains where no solution was found given a 1800 seconds timeout.
+Missing values in the tables (reported as -) correspond to domains where no solution was found within a 1800s timeout.
 
 Table \ref{tab:results_minimum_100_10} shows the results of the first case study.
-Recall scores are generally higher than the precision ones, and, in fact, the models learned for six out of the fourteen domains were perfectly complete.
-Although precision is overall lower, it is interesting that most learned sets of negative effects are flawless.
-With regards to the computation time, we find times below one second in all cases except for the {\em floor-tile} domain.
+Recall scores are generally higher than the precision ones, and, in fact, the models learned for six out of the 14 domains were perfectly complete.
+Although precision is overall lower, it is interesting to notice that the learned sets of negative effects are mostly flawless.
+With regards to the computation time, we can observe times are below one second in all cases except for the {\em floor-tile} domain.
 
@@
 
-In Table \ref{tab:results_minimum_30_30}, which collects the result of the PSPACE case study with 30\% observability, the scores for some domains are missing.
-This is the case for {\em floor-tile}, and {\em parking} both considered \emph{puzzle-like} domains, a trait known for putting a strain on the planners.
-Interestingly enough, {\em hanoi} also qualifies as a \emph{puzzle-like} domain, and we can see how this is reflected on the computation time.
-With regards to the learned models, we find that they retain similar levels of soundness, while their completeness is lower than in the previous case study.
-This is specially the case for the preconditions, which have gone from 0.94 recall to 0.54.
-The explanation for these scores is that input actions are strong constraints that play a key role on how close the learned model are with respect to the GTM.
-The more these actions are missing the more likely it is for reformulations to occur so pure syntax-based metrics become less relevant.
+Table \ref{tab:results_minimum_30_30} gathers the results of the PSPACE-complete case study with 30\% observability.
+We can see in the table that the scores of some domains are missing.
+This is the case of {\em floor-tile} and {\em parking}, which not only are fairly complex domains, but also categorized as \emph{puzzle-like} domains, a feature that is known for putting a strain in the planners.
+Interestingly enough, we note the huge computation time of {\em hanoi}, which also qualifies as a \emph{puzzle-like} domain.
+Regarding quality, we find that the learned models retain a level of soundness similar to Table~\ref{tab:results_minimum_100_10} but the completeness is lower than in the previous case study.
+This is specially noticeable in the preconditions, where recall values drop from from 0.94 to 0.54.
+This is because the input actions act as strong constraints playing a key role on the closeness of the learned model to the GTM.
+The more actions are missing in the input knowledge, the more likely the occurrence of reformulations.
 
@@
 
-The effect of missing input actions is more evident in our last case study with \NO action sequences and state trajectories (Table \ref{tab:results_minimum_0_0}).
-Contrary to what might be expected by looking at the previous table, in this case we are able to find solutions for all domains.
-The reason for this is that the search is less constrained so there are far more possible solutions for this learning task.
-This wider range of solutions is also evidenced by the learned models, which in spite of being compliant, are more removed from the GTM.
-In the table we find that both precision and recall have seen their scores decreased to 0.6 and 0.5, respectively.
-We argue, however, that syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the \emph{reformulations} that frequently occur in these circumstances.
+We now analyze the PSPACE case study with \NO action sequences and \NO state trajectories (Table~\ref{tab:results_minimum_0_0}).
+A first outstanding observation is that, contrary to what might be expected by looking at the previous table, we are able in this case to find solutions for all the domains.
+This happens because the search is less constrained and consequently there are far more possible solutions for this learning task.
+This broader space of solutions is also stressed in a diminished quality of the learned models.
+Thus, despite the learned models are compliant with the input data, they are further from the original GTM.
+In Table~\ref{tab:results_minimum_0_0} we can observe the global values of Precision and Recall drop to 0.6 and 0.5, respectively.
 
-To illustrate this, Figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the {\tt stack} operator learned from plan traces with \NO action sequences and state trajectories.
-This learned action model removes a block from on top of another block and puts it down on the table.
-There are two main differences with respect to the GTM: (1), the learned action is unstacking a block instead of stacking it and (2), the block ends on the table after the action is applied, working as an {\tt unstack\&putdown} \emph{macro-action}.
-We refer to the first difference as \emph{role swapping} and it happens when there are missing actions in the plan trace.
-If no actions are present in the input traces, the names of actions become meaningless (actions are effectively anonymous), meaning that they can interchange their behavior with any other comparable action model.
-The second issue happens when there are missing states in the input traces, since \emph{macro-actions} apply more than one action in a single step so some intermediate states are skipped.
+We argue, however, that syntax-based metrics are not appropriate for scenarios with minimal observability as they cannot cope with the reformulations that frequently occur in these circumstances.
+To illustrate this, Figure \ref{fig:macroaction} shows the PDDL encoding of the action model of the {\tt\small stack} operator learned from plan traces with \NO action sequences and \NO state trajectories.
+This learned action model removes a block from on top of another block and puts it down on the table in a single step.
+There are two main differences with respect to the model of the {\tt\small stack} operator of the GTM: (1) the learned action is actually \emph{unstacking} a block instead of stacking it and (2) the block on the top ends on the table, not held by the robot arm.
+We refer to the first difference as \emph{role swapping} and it happens when there are missing actions in the input plan trace.
+If no actions are present in the input traces, the names of actions become meaningless, in which case the effectively anonymous actions can interchange their behaviour with any other comparable action model.
+The second difference indeed reveals that the learned action model is working as an {\tt\small unstack+put-down} \emph{macro-action}.
+This happens when there are missing states in the input traces since a \emph{macro-action} can be seen as the application of more than one action in a single step, thus skipping some intermediate states.
+
+Reformulated action models, like the one in Figure \ref{fig:macroaction}, are indeed sound models that can be used to solve planning tasks.
+For instance, any \emph{blocks-world} problem can be solved unstacking all the blocks to the table ({\tt\small unstack+put-down}) and then stacking them to meet the goal conditions ({\tt\small pick-up+stack}).
+Hence, the \NO/\NO case study features all the conditions for reformulation to happen, and this is the reason why scenarios such as this one are better evaluated using {\em semantic-based metrics}.
 
@@
 
-Reformulated action models, like the one in Figure \ref{fig:macroaction}, can be used to solve planning tasks.
-For instance, any blocks-world problem can be solved unstacking all the blocks to the table (unstack + put-down) and then stacking them to meet the goal conditions (pick-up + stack).
-The \NO/\NO case study meets all the conditions where reformulations happens, and this is the reason why scenarios such as this are better evaluated using {\em semantic-based metrics}.
-
 \subsection{Syntactic versus semantic evaluation}
 
-The last experiment compares the scores provided by the syntactic and semantic versions of {\em precision} and {\em recall}.
-For that purpose, we will evaluate the models learned in \ref{minimal} both syntactically, using the GTM, and semantically, using a testing set of 5 plan traces.
-Our goal with this experiment is to gauge the suitability of the semantic metrics we propose with respect to their well-known counterparts.
+Our last experiment is devoted to compare the scores provided by the syntactic and semantic versions of Precision and Recall.
+For that purpose, we will evaluate the models learned in Section \ref{minimal} both syntactically, using the GTM, and semantically, computing the set of action models closest to the learned models that is compliant with a testing set of five traces (see Definition \ref{compliant}).
+We must note that since we are using {\sc Madagascar}, a satisficing planner, the solution to the model evaluation may not be the closest compliant model, reason why the scores of sem-Precision and sem-Recall are approximate values.
+Our goal with this experiment is to gauge the suitability of the semantic metrics proposed in section \ref{sec:evaluation} with respect to their well-known counterparts.
 With that in mind, we define two case studies:
@@
 	\item \textbf{\FO action sequence and \PO state trajectory}: In this case study the full sequence of actions is known and no states are missing, which makes it practically impossible for reformulated models to appear.
-In fact, in all our experiments with \FAMA and other approaches we never observed reformulations when the full sequence of actions is known.
+In fact, in all our experimentation with \FAMA and other approaches we never observed reformulations when the full sequence of actions is known.
 	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a case study that favors reformulations in the learned models, as previously discussed.
@@
 Table \ref{tab:metric_comparison_100_10} shows the results of the first case study.
-Looking at the high scores of the syntactic metrics, specially {\em recall}, we can conclude that the learned models are, in fact, fairly similar to the GTM.
+Looking at the high scores of the syntactic metrics, specially the value of Recall, we can conclude that the learned models are, in fact, fairly similar to the GTM.
 This supports our conclusion that no reformulation occurs in this case study, which also means that the space of possible solutions is restricted to models close to the GTM.
-The values for {\em sem-Precision} and {\em sem-Recall} are also high across the table, which is exactly the desired behavior for these metrics given that solutions are close to the GTM.
-In comparison, both versions of {\em recall} show similar scores, while the difference is bigger for the precision metrics.
-This means that the semantic version is more lenient towards extra preconditions or effects.
-This matches what we have observed during the experiments.
-For instance, the learned model for the {\tt move} action of the {\em hanoi} domain, specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
-This learned model is semantically correct but sintactically different from the GTM and hence, penalized by the {\em precision} metric.
+The values of sem-Precision and sem-Recall are also very high across the table, which is exactly the desired behavior for these metrics given that solutions are very close to the GTM.
+In comparison, Recall and sem-Recall show similar scores, while sem-Precision is significantly higher than Precision, thus showing that the sem-Precision is more lenient towards extra preconditions or effects.
+This is in line with the results of the previous experiments, where the common appearance of redundant or implicit preconditions in the learned models is penalized by the Precision metric.
+We can interpret this phenomenon as a manifestation of the qualification problem~\cite{GinsbergS88}.
+For instance, the model learned for the {\tt\small move} action of the {\em hanoi} domain specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
+This learned model is semantically correct but syntactically different from the GTM and hence penalized by the Precision metric.
 
-Finally, Table \ref{tab:metric_comparison_0_0} details the results of the \NO/\NO case study.
-Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger here.
-Comparing the scores of both versions, we find that learned models that achieved mediocre scores when using the GTM as reference, are in fact reasonable sound and complete, reaching and overall score of 0.91 for both {\em sem-Precision} and {\em sem-Recall}.
-This means that the models learned by our approach, despite syntactically different from the GTM, require very few editions to explain the testing traces given as input.
+Table~\ref{tab:metric_comparison_0_0} details the results of the \NO/\NO case study.
+One first observation is the impossibility of applying a semantic evaluation in some of the most complex domains with five traces.
+Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger in this PSPACE-complete scenario.
+Comparing the scores of both versions, we find that learned models that achieved mediocre scores when using the GTM as reference (syntactic metrics), are in fact reasonably sound and complete, reaching an overall score of 0.91 in both sem-Precision and sem-Recall.
+This is an indication that the models learned by our approach, despite syntactically different from the GTM, require very few editions to explain the testing set of traces.
 
@@
 The first one is that, when no reformulation occurs, these metrics behave similarly to their syntactic counterparts, which means they are a good substitute when the GTM is not available.
-The second conclusion is that {\em sem-Precision} and {\em sem-Recall} are better suited to evaluate reformulated models than the original syntactic metrics, since they contemplate valid solutions outside the GTM that successfully explain the given input data.
+The second conclusion is that sem-Precision and sem-Recall are better suited to evaluate reformulated models than the original syntactic metrics since they contemplate valid solutions outside the GTM that successfully explain the given input data.

### 2018-08-16 benchmarks/aaai19/automata/experiment.py
updating experiments script for aaai19 paper

@@ -1,5 +1,5 @@
 #! /usr/bin/env python
-import sys,os,random
+import sys,os,random,glob
 
 # **************************************#
 # MAIN
@@ -16,30 +16,63 @@ except:
     sys.exit(-1)
 
 
-# Generate the n possible automatas
-for i in range(0,nModels):
-    cmd=  "./gen-automata-domain.py " + str(nStates) + " " + str(nObs) + " 2 " + str(i)
-    print cmd
-    os.system(cmd)
-
-    cmd=  "mv domain.pddl models/domain-"+str(i)+".pddl"
-    print cmd
-    os.system(cmd)
-
-
-# Pick one automata
+# Random generation of n candidate automatas
 random.seed()
+candidates=[]
+
+while (len(candidates) < nModels):
+    automata_id = random.randint(0, (nStates*nObs)-1)
+
+    if not automata_id in candidates:
+        candidates.append(automata_id)
+
+        cmd =  "mkdir models"
+        print cmd
+        os.system(cmd)
+        
+   
+        cmd =  "./gen-automata-domain.py " + str(nStates) + " " + str(nObs) + " 0 " + str(automata_id)
+        print cmd
+        os.system(cmd)
+
+        cmd =  "mv domain.pddl models/domain-"+str(automata_id)+".pddl"
+        print cmd
+        os.system(cmd)
+
+        
+# Pick one automata
 picked_id = random.randint(0, nModels-1)
-cmd =  "cp models/domain-"+str(picked_id)+".pddl domain.pddl"
+domain_filename = "domain-"+str(candidates[picked_id])+".pddl"
+cmd =  "cp models/" + domain_filename + " ./"
 print cmd
 os.system(cmd)
 
 
 # Generate trace for picked automata
-cmd=  "./gen-automata-trace.py " + str(nStates) + " " + str(nObs) + " " + str(nlen)
+cmd=  "rm problem.pddl; ./gen-automata-problem.py "  + str(nStates) + " " + str(nObs) + " " + "0" * nlen
+print cmd
+os.system(cmd)
+    
+cmd=  "/home/slimbook/research/strips-learning/src/example-generator.py " + domain_filename + " problem.pddl M " + str(nlen) + " " + str(5)
 print cmd
 os.system(cmd)
 
 
+# Evaluate the trace for all the candidate automata
+scores=[]
+for item in sorted(glob.glob("./models/*")):
+    score_filename = item + ".score"
[…]

### 2018-08-16 benchmarks/aaai19/automata/gen-automata-domain.py

@@ -59,7 +59,6 @@ for j in range(nObs):
 str_out = str_out +  ")\n"
 str_out = str_out + "\n"
 
-print str_ID
 
 counter=0
 for i in range(0,nStates):

### 2018-08-16 benchmarks/aaai19/automata/gen-automata-trace.py

@@ -1,26 +1,0 @@
-#! /usr/bin/env python
-import sys,os
-
-# **************************************#
-# MAIN
-# **************************************#
-try:
-    nStates  = int(sys.argv[1])
-    nObs = int(sys.argv[2])
-    nlen = int(sys.argv[3])
-
-except:
-    print "Usage:"
-    print sys.argv[0] + " <nStates> <nObservations> <length> "
-    sys.exit(-1)
-
-cmd=  "./gen-automata-problem.py "  + str(nStates) + " " + str(nObs) + " " + "0" * nlen
-print cmd
-os.system(cmd)
-    
-cmd=  "../../../src/example-generator.py domain.pddl problem.pddl M " + str(nlen-1) + " " + str(nlen)
-print cmd
-os.system(cmd)
-
-sys.exit(0)
-

### 2018-08-16 benchmarks/aaai19/automata/experiment.py
updating experiments script for aaai19 paper

@@ -60,16 +60,18 @@ os.system(cmd)
 
 # Evaluate the trace for all the candidate automata
 scores=[]
-for item in sorted(glob.glob("./models/*")):
-    score_filename = item + ".score"
-    cmd=  "/home/slimbook/research/strips-learning/src/compiler_new.py -v " + item + " ./ 100 10 -t ten-observation-01 > " + score_filename
+for item in candidates:
+    domain_filename = "models/domain-"+str(item)+".pddl"
+    score_filename = "domain-"+str(item)+ ".score"
+    cmd=  "/home/slimbook/research/strips-learning/src/compiler_new.py -v " + domain_filename + " ./ 100 10 -t ten-observation-01 > " + score_filename
     print cmd
     os.system(cmd)
 
     score_file = open(score_filename, 'r')
     scores.append(float(score_file.readline().split(" & ")[1]))
     score_file.close()
-
+    
+print candidates
 print scores
     
 sys.exit(0)

### 2018-09-03 doc/icaps-19/model-recognition.tex
new folder for the ICAPS19 paper on model recognition

--- 2018-06-21 doc/aaai-19/model-recognition.tex
@@
 \begin{abstract} 
-Given a partially observed plan execution and a set of possible planning models, {\em model recognition} is the task of identifying the model in the set that produced the observed plan execution.
+Given a partially observed plan execution and a set of possible planning models, {\em model recognition} is the task of identifying which model in the set produced the observed plan execution.
 The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a \strips\ model to produce a given partially observed plan execution.
-This method, that we called {\em model recognition as planning}, is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
+This method, that we called {\em model recognition as planning}, is robust to missing data in the observed plan execution (i.e.
+missing intermediate states and actions) given as input besides, it is computable with an off-the-shelf classical planner.
 The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different {\em Turing Machines}.
@@
 {\em Plan recognition as planning} is the model-based approach for plan recognition~\cite{ramirez2012plan,ramirez2009plan}.
-This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal of the agent according to the observed plan execution.
+This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal of the agent, according to the observed plan execution.
 
@@
 \begin{itemize}
-\item Once the planning model is recognized with certainty, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
-\item It enables the recognition of algorithms by observing their execution given that diverse algorithm representations, like {\em {\sc GOLOG} programs}, {\em finite state controllers} or {\em push-down automata} can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}.
+\item Once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
+\item It enables indentifying algorithms by observing their execution.
+Diverse algorithm representations, like {\em finite state controllers}, {\em push-down automata} or {\em {\sc GOLOG} programs}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}.
 \end{itemize}
@@
 The method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
-The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em Turing Machines} (all of them defined within the same {\em tape alphabet} and same {\em machine states} but different {\em transition functions}).
+The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em Turing Machines}.
+All of these {\em Turing Machines} are defined within the same {\em tape alphabet} and same {\em machine states} but different {\em transition functions}.
 We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or the values of several tape cells, are unknown.
@@
 
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
-The dynamics of the actions in $A$ are specified with two functions: $App(s)$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state, $G\subseteq\mathcal{L}(F)$ is a goal condition and $A$ is a set of actions whose dynamics are specified with two functions: $App(s)$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
 
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\subseteq\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that, when executed starting from the initial state $I$, induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that, when executed starting from the initial state $I$, induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and forall {\small $1\leq i\leq n$}, an action $a_i$ is applicable in the corresponding state $a_i\in App(s_{i-1})$ and  generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 The {\em plan length} is denoted with $|\pi|=n$.
@@
 \subsection{The observation model}
-Given classical planning problem $P=\tup{F,A,I,G}$, the observation of the execution of a plan $\pi$ that solves $P$ is an interleaved combination $\tau=\tup{s_1, a_1, \ldots , a_n, s_m}$ of a {\em partially-observed} sequence of actions and a sequence of {\em partially-observed} states.
-Because of the partial observability $0\leq n\leq |\pi|$ and $0\leq m\leq |\pi|+1$ and hence, the transitions between two consecutive observed states in $\tau$ may involve the execution of more than a single action.
-Formally $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound.
+Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$, the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{s_1, a_1, \ldots , a_n, s_m}$, an interleaved combination of {\small $1\leq m\leq |\pi|+1$} observed states and {\small $1\leq n\leq |\pi|$} observed actions:
+\begin{itemize}
+\item Observed states may be partial states. The value of certain fluents may be omitted in that states, i.e.~$|s_i|\leq |F|$ for every $0\leq i\leq m$.
+
+\item The sequence of observed states $\tup{s_1, \ldots, s_m}$ in $\tau$ is the same sequence of states traversed by $\pi$ but certain states may also be omitted.
+Therefore the transitions between two consecutive observed states in $\tau$ may require the execution of more than a single action.
+Formally, $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound.
 This means that having $\tau$ does not implies knowing the actual length of $\pi$.
+\item The sequence of observed actions $\tup{a_1, \ldots, a_n}$ in $\tau$ is a sub-sequence of the solution plan $\pi$.
+\end{itemize}
 
-We say that a given observation $\tau$ is {\em compliant} with a given plan $\pi$ iff:
-\begin{itemize}
-\item The sequence of actions $\tup{a_1, \ldots, a_n}$ in $\tau$ is a subsequence of $\pi$.
-\item The sequence of states $\tup{s_1, \ldots, s_m}$ in $\tau$ is the same sequence of states traversed by $\pi$ but (1), with certain states omitted and/or (2), the value of certain fluents omitted in that states, i.e.~$|s_i|\leq |F|$ for every $0\leq i\leq m$.
-\end{itemize}
+\begin{definition}[$\Phi$-observation]
+Given a subset of fluents $\Phi\subseteq F$ we say that $\tau$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every $0\leq i\leq m$, each observed state $s_i$ only contains fluents in $\Phi$.
+\end{definition}
 
@@
 \begin{itemize}
-\item $P=\tup{F,A,I,G}$ is a classical planning problem such that the dynamics of the actions in $A$ is unknown because the $App(s)$ and $\theta(s,a)$ functions are undefined.
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty set of action models, each defining a different dynamics (a different $App(s)$ and $\theta(s,a)$ pair) for the actions in $A$.
-\item $\tau$ is the observation of the execution of a solution plan $\pi$ for $P$.
+\item $P=\tup{F,A,I,G}$ is a classical planning problem such that the dynamics of the actions in $A$ is unknown because functions $App(s)$ and/or $\theta(s,a)$ are undefined.
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty set of models for the actions in $A$. Each model in $\mathcal{M}\in M$ defines a different function pair ($App$,$\theta$).
+\item $\tau$ is an observation of the execution of a solution plan $\pi$ for $P$.
 \end{itemize}
@@
 \item $\pi^*_\tau$ is an optimal solution for the classical planning problem $P$. 
-\item $\pi^*_\tau$ is compliant with the given observation $\tau$.
+\item $\tau$ is an observation of the execution of $\pi^*_\tau$ on the classical planning problem $P$.
 \end{enumerate}
@@
 Note that the second step assumes that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
-A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the different PDDL models of two different agents.
+A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
 
@@
 \label{sec:asPlanning}
-Here we analyze the particular instantiation of the {\em model recognition} task where the dynamics of the actions $A\in P$ (i.e.
-the $App(s)$ and $\theta(s,a)$ functions) are specified with \strips\ action schemas.
+Here we analyze the particular instantiation of the {\em model recognition} task where the dynamics of the actions $A$ (i.e.
+the $App$ and $\theta$ functions) are specified with \strips\ action schemas.
 Figure~\ref{fig:stack} shows the PDDL code for the {\em stack} actions, taken from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
@@
 
-Now we are ready to define the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_{\xi}$, that represent the propositional encoding for the preconditions, negative and positive effects of an action schema $\xi$.
-If a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in that schema.
-For instance, Figure~\ref{fig:encodedstack} shows the conjunction of fluents that represents he propositional encoding for the preconditions, negative and positive effects of the {\em stack} schema shown in Figure~\ref{fig:stack}.
+Now we are ready to define the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_{\xi}$.
+These fluents represent a propositional encoding for the preconditions, negative and positive effects of an action schema $\xi$ and hence, they specify the $App$ and $\theta$ functions for all the actions $a\in A$ built from the corresponding \strips\ action schemas.
+In more detail, if a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in that schema.
+For instance, Figure~\ref{fig:encodedstack} shows the conjunction of fluents representing the propositional encoding for the preconditions, negative and positive effects of the {\em stack} schema (see Figure~\ref{fig:stack}).
 
@@
 
-The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions and that they cannot be positive effects and also, that positive effects cannot appear as preconditions).
-For the mentioned {\em blocksworld}, $2^{2|F_{\tt stack}|}=4194304$ while $2^{2|F_{\tt pickup}|}=1024$.
+The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2\times|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions).
+For the mentioned {\em blocksworld}, $2^{2\times|F_{\tt stack}|}=4,194,304$ while $2^{2\times|F_{\tt pickup}|}=1,024$.
 We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same space of possible \strips\ models.
@@
 
-Our $\delta(\mathcal{M},\tau)$ distance could also be defined assessing the edition required by the observed plan execution to match the given model.
+The $\delta(\mathcal{M},\tau)$ distance could also be defined assessing the edition required by the observed plan execution to match the given model.
 This implies defining {\em edit operations} that modify $\tau$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
 Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$.
-In practice, the number of {\em variable objects} is normally smaller than the number of objects in a planning problem.
+In practice, the number of {\em variable objects} is smaller than the number of objects in a planning problem.
 
@@
 \begin{enumerate}
-\item Assuming that a priori, all action models $\mathcal{M}\in M$ are equiprobable because there are no reasons to assume that one is a priori more likely than the others.
-Therefore, $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2|F_{\xi}|}}$.
-\item Assuming that all the observations of plan executions with a maximum of $n$ observed actions and $m$ observed states are equiprobable then $P(\tau)=\frac{1}{2^{n\times|A|}\times2^{m\times|F|}}$.
-\item To estimate the conditional probability $P(\tau|\mathcal{M})$ we compute the {\em observation edit distance} $\delta(\mathcal{M},\tau)$ and map it into a $[0,1]$ likelihood with the expression, $1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
+\item Assuming {\em a priori} all action models $\mathcal{M}\in M$ are equiprobable.
+There are no reasons to assume that one model is {\em a priori} more likely than the others so, $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2\times|F_{\xi}|}}$.
+\item Assuming that all the observations of plan executions with maximum $n$ observed actions and $m$ observed states are equiprobable then, $P(\tau)=\frac{1}{|A|^{n}\times2^{m\times|F|}}$.
+\item Estimating the conditional probability $P(\tau|\mathcal{M})$ by mapping the {\em observation edit distance} into a $[0,1]$ likelihood, $1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
 \end{enumerate}

### 2018-09-03 doc/icaps-19/model-recognition.tex
added TM automata to the introduction

@@
 \usepackage{tikz}
+\usetikzlibrary{arrows,automata}
 \usepackage{comment}
@@
 The method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
+
 The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em Turing Machines}.
 All of these {\em Turing Machines} are defined within the same {\em tape alphabet} and same {\em machine states} but different {\em transition functions}.
+For instance, Figure~\ref{fig:automatae} illustrate a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
 We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or the values of several tape cells, are unknown.
+
+\begin{figure}
+  \begin{scriptsize}
+  \begin{center}
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
+	  \node[state] (A)              {$Q_0$};
+	  \node[state] (B) [right of=A] {$Q_1$};
+	  \node[state] (C) [right of=B] {$Q_2$};
+	  \node[state] (D) [right of=C] {$Q_3$};
+	  \node[state] (E) [below of=A] {$Q_4$};
+	  \node[state] (F) [right of=E] {$Q_5$};
+
+	  \path
+(A) edge  node {a/x,right} (B)
+(B) edge [align=left, loop above] node {a/a,right\\y/y,right} (B)
+(B) edge  node {b/y,right} (C)
+(C) edge [align=left, loop above] node {z/z,right\\b/b,right} (C)
+(C) edge  node {c/z,left} (D)
+(D) edge [align=left, loop right] node {a/a,left\\b/b,left\\y/y,left\\z/z,left} (D)
+(D) edge [bend left=45] node {x/x,right} (A)
+
+(A) edge  node {y/y,right} (E)
+(E) edge [align=left, loop left] node {y/y,right\\z/z,right} (E)
+(E) edge  node {b/b,right} (F)
+;
+	\end{tikzpicture}
+  \end{center}
+  \end{scriptsize}  
+ \caption{\small Seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.}
+\label{fig:automatae}
+\end{figure}
 
@@
 
+\section{Recognition of {\em Turing Machines}}
+\label{sec:asPlanning}
+We analyze now {\em model recognition} when the input set of given \strips\ models represent different {\em Turing Machines}.
+
+A {\em Turing machine} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma,\Upsilon,\square,\delta}$:
+\begin{itemize}
+\item $Q$ is a finite and non-empty set of machine states with the {\em initial state} $q_0\in Q$ and the {\em terminal states} $Q_{\bot}\subseteq Q$.  
+\item $\Sigma$ is the {\em tape alphabet}, a finite non-empty set of symbols with the {\em input alphabet} $\Upsilon\subseteq\Sigma$ (symbols allowed to initially appear in the tape) and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
+\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}.
+For each pair of tape symbol and non-terminal machine state $\delta$ defines: (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
+If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
+\end{itemize}
+
+Figure~\ref{tab:tm-anbncn} shows the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
+The {\em tape alphabet} is $\Sigma=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} $\Upsilon=\{a,b,c,\square\}$ and the machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ (where \underline{$q_5$} is the only acceptor state).
+
+\begin{figure}
+\begin{center}
+    \begin{tabular}{| c || c | c | c | c | c | c |}
+    \hline
+      & $q_0$ & $q_1$ & $q_2$ & $q_3$ & $q_4$ & \underline{$q_5$} \\ \hline\hline
+    a & x,r,$q_1$ & a,r,$q_1$ & - &  a,l,$q_3$ & - & - \\ \hline
+    b & - & y,r,$q_2$ & b,r,$q_2$ & b,l,$q_3$ & - & - \\ \hline
+    c & - & - & z,l,$q_3$ & - & - & - \\ \hline
+    x & - & - & - & x,r,$q_0$ & - & - \\ \hline
+    y & y,r,$q_4$ & y,r,$q_1$ & - & y,l,$q_3$ & y,r,$q_4$ & - \\ \hline
+    z & - & - & z,r,$q_2$ & z,l,$q_3$ & z,r,$q_4$ & - \\\hline
+    $\square$ & - & - & - & - & $\square$,r,$q_5$  & - \\                
+    \hline
+    \end{tabular}
+\end{center}
+  \caption{\small Seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the only acceptor state).}
+  \label{tab:tm-anbncn}
+\end{figure}
+
+\subsection{\strips\ modeling of {\em Turing Machines}}
+Our \strips\ model of the {\em transition function} $\delta$ of a given {\em Turing Machine} $\mathcal{M}$ is built as follows:
+\begin{itemize}
+\item Fluents $F$ are instantiated from a set $\Psi$ of four {\em predicates}: {\small\tt (head ?x)} that encodes the current position of the header in the tape.
+{\small\tt (next ?x1 ?x2)} encoding that cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
+{\small\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
+{\small\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
+Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
+\item Actions $A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ schema is defined:
+\begin{itemize}
+\item The {\bf header} is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$. Parameters $?xl$, $?x$ and $?xr$ are tape cells.
+\item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} so $?x$ is the tape cell pointed by the header and $?xl/?xr$ its left/right neighbors.
+Preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} are also included to capture the symbol pointed by the header and the current machine state.
+\item The {\bf delete effects} remove the symbol currently pointed by the header and the current machine state while {\bf positive effects} set the new symbol pointed by the header and the new machine state according to $\delta$.
+\end{itemize}
+\end{itemize}
+
+The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn}.
+The full encoding of the {\em Turing Machine} of Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema.
+\begin{figure}
+\begin{scriptsize}
+\begin{lstlisting}
+(:action transition-1      ;;; a,$q_0\rightarrow$ x,r,$q_1$
+  :parameters (?xl ?x ?xr)
+  :precondition (and (head ?x) (symbol-a ?x) (state-$q_0$)
+                     (next ?xl ?x) (next ?x ?xr))
+  :effect (and (not (head ?x)) 
+               (not (symbol-a ?x)) (not (state-$q_0$))
+               (head ?xr) (symbol-x ?x) (state-$q_1$)))
+\end{lstlisting}
+\end{scriptsize}
+ \caption{\small \strips\ action schema that models the transition $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
+\label{fig:update-rule}
+\end{figure}
+
+With regard to our \strips\ model for {\em Turing Machines}, executions of a {\em Turing Machine} are definable as an action sequence $\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+For instance, the execution of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} with an initial tape $abc\square\square\square\ldots$ produces the eight-action plan {\small $(a,q_0\rightarrow x,r,q_1)$, $(b,q_1\rightarrow y,r,q_2)$, $(c,q_2\rightarrow z,l,q_3)$, $(y,q_3\rightarrow y,l,q_3)$, $(x,q_3\rightarrow x,r,q_0)$, $(y,q_0\rightarrow y,r,q_4)$, $(z,q_4\rightarrow z,r,q_4)$, $(\square,q_4\rightarrow \square,r,\underline{q_5})$}.
+
+\subsection{The $P(\mathcal{M}|\tau)$ probability for Turing Machines}
+Assuming that the actual applied transitions is unknown means that the observation $\tau$ of the execution of a Turing Machine contains no actions, it is simply a sequence of states $\tau=\tup{s_1, \ldots , s_m}$.
+Further, assuming that the internal machine state is unknown means that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt (state-$q$)} fluents, with $q\in Q$ and $q\neq q_0$.
+Finally, assuming that the values of several tape cells is unknown means that fluents of the kind {\small\tt (symbol-$\sigma$ ?x)} are missing (i.e.
+unobserved) for some state $s_i\in \tau$ s.t.
+$1\leq i\leq m$.
+These facts affects to the a priori probaility of the possible observations.
+
+On the other hand, the model edition for Turing Machines is limited to these subset of possible positive effects: {\tt (head ?xr)} or {\tt (head ?xl)}, {\tt(symbol-$\sigma$ ?x)} with $\sigma \in \Sigma$ and last but not least, {\tt (state-q)} with $q\in Q$.
+No other positive effects, preconditions, or negative effects are required to be edited.
+This fact reduces the possible edition operations and affects to the a priori probaility of the possible action models.
+
 \section{Model recognition as planning}
@@
 A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$ in the observation $\tau$}.
-The solution plan continues with a postfix that validates the edited model $\mathcal{M}'$ on the given observations $\tau$.
+\item {\bf Validates the edited model $\mathcal{M}'$}. The solution plan continues with a postfix that:
+\begin{enumerate}
+\item Induces an optimal solution plan $\pi^*_\tau$ for the original classical planning problem $P$.
+\item Validates that $\tau$ is an observation of the execution of $\pi^*_\tau$ on the classical planning problem $P$.
+\end{enumerate}
 \end{enumerate}
 
-Figure~\ref{fig:plan-pdistance} shows the plan for editing a given {\em blockswold} model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-The edited action model is validated at the observation of a four action plan for inverting a two-block tower where the intermediate states, $s_1$, $s_2$ and $s_3$, are unobserved.
+Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,1]) for editing a given {\em blockswold} model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
+The postfix of the plan (steps [2,6]) validates the edited action model at the observation of a four action plan for inverting a two-block tower where intermediate states, $s_1$, $s_2$ and $s_3$, are unobserved.
 \begin{figure}
@@
 
-Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deletions) required by $\mathcal{M}'$ to be validated in the given observation, e.g.
-$\delta(\mathcal{M},\tau)=2$ for the example in Figure~\ref{fig:plan-pdistance}.
-In this case $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) and $|F_v|=|F_{\tt stack}|=|F_{\tt unstack}|=11$ while $|F_{\tt pickup}|=|F_{\tt putdown}|=5$.
-%The {\em observation edit distance} is exactly computed if the classical planning task resulting from our compilation is optimally solved (according to the number of edit actions); is approximated if it is solved with a satisfying planner; and is a less accurate estimate (but faster to be computed) if the solved task is a relaxation of the classical planning task that results from our compilation~\cite{bonet2001planning}.
+Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deletions) required by $\mathcal{M}'$ to be validated.
+In the example of Figure~\ref{fig:plan-pdistance} $\delta(\mathcal{M},\tau)=2$ and $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) s.t.
+$|F_v|=|F_{\tt stack}|=|F_{\tt unstack}|=11$ while $|F_{\tt pickup}|=|F_{\tt putdown}|=5$.
 
@@
 Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+
+An action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
 \[
@@
 
-\subsection{Modeling {\em Turing Machines} with \strips\ }
-A {\em Turing machine} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma,\Upsilon,\square,\delta}$:
-\begin{itemize}
-\item $Q$, is a finite and non-empty set of machine states such that $q_0\in Q$ is the initial state of the machine and $Q_{\bot}\subseteq Q$ is the subset of acceptor states.  
-\item $\Sigma$ is the {\em tape alphabet}, that is a finite non-empty set of symbols that contains the {\em input alphabet} $\Upsilon\subseteq\Sigma$ (the subset of symbols allowed to initially appear in the tape) and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
-\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}.
-For each possible pair of tape symbol and non-terminal machine state $\delta$ defines (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
-If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
-\end{itemize}
-
-Figure~\ref{tab:tm-anbncn} shows the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
-The {\em tape alphabet} is $\Sigma=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} $\Upsilon=\{a,b,c,\square\}$ and the possible machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ where \underline{$q_5$} is the only acceptor state.
-
-\begin{figure}
-\begin{center}
-    \begin{tabular}{| c || c | c | c | c | c | c |}
-    \hline
-      & $q_0$ & $q_1$ & $q_2$ & $q_3$ & $q_4$ & \underline{$q_5$} \\ \hline\hline
-    a & x,r,$q_1$ & a,r,$q_1$ & - &  a,l,$q_3$ & - & - \\ \hline
-    b & - & y,r,$q_2$ & b,r,$q_2$ & b,l,$q_3$ & - & - \\ \hline
-    c & - & - & z,l,$q_3$ & - & - & - \\ \hline
-    x & - & - & - & x,r,$q_0$ & - & - \\ \hline
-    y & y,r,$q_4$ & y,r,$q_1$ & - & y,l,$q_3$ & y,r,$q_4$ & - \\ \hline
-    z & - & - & z,r,$q_2$ & z,l,$q_3$ & z,r,$q_4$ & - \\\hline
-    $\square$ & - & - & - & - & $\square$,r,$q_5$  & - \\                
-    \hline
-    \end{tabular}
-\end{center}
-  \caption{\small Seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the only acceptor state).}
-  \label{tab:tm-anbncn}
-\end{figure}
-
-A classical planning frame $\Phi=\tup{F,A}$ can encode the {\em transition function} $\delta$ of a {\em Turing Machine} $\mathcal{M}$ as follows:
-\begin{itemize}
-\item Fluents $F$ are instantiated from a set of four {\em predicates} $\Psi$: {\small\tt (head ?x)} that encodes the current position of the header in the tape.
-{\small\tt (next ?x1 ?x2)} encoding that the cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
-{\small\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
-{\small\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
-Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
-\item Actions $A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ schema is defined:
-\begin{itemize}
-\item The {\bf header} is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$. Parameters $?xl$, $?x$ and $?xr$ are tape cells.
-\item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} to make $?x$ the tape cell pointed by the header and $?xl/?xr$ its left/right neighbors.
-Preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} are also included to capture the symbol currently pointed by the header and the current machine state.
-\item The {\bf delete effects} remove the symbol currently pointed by the header and the current machine state while the {\bf positive effects} set the new symbol pointed by the header and the new machine state according to $\delta$.
-\end{itemize}
-\end{itemize}
-
-The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} (the full encoding of the {\em Turing Machine} of Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema).
-\begin{figure}
-\begin{scriptsize}
-\begin{lstlisting}
-(:action transition-1      ;;; a,$q_0\rightarrow$ x,r,$q_1$
-  :parameters (?xl ?x ?xr)
-  :precondition (and (head ?x) (symbol-a ?x) (state-$q_0$)
-                     (next ?xl ?x) (next ?x ?xr))
-  :effect (and (not (head ?x)) 
-               (not (symbol-a ?x)) (not (state-$q_0$))
-               (head ?xr) (symbol-x ?x) (state-$q_1$)))
-\end{lstlisting}
-\end{scriptsize}
- \caption{\small \strips\ action schema that models the transition $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
-\label{fig:update-rule}
-\end{figure}
-
-Since the {\em transition function} of a {\em Turing Machine} can be encoded as a classical planning frame $\Phi=\tup{F,A}$, executions of that {\em Turing Machine} are definable as an action sequence $\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-
 \subsection{Experimental setup}

### 2018-09-06 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
draft added for conselleria research project

--- 2018-06-14 doc/programming-project/AI-programming.tex
@@
 \title{Artificial Intelligence for the Automated Synthesis and Validation of Programs}
-\author{Sergio Jim\'enez Celorrio\\
-\footnotesize Universitat Politècnica de València, Spain\\
-\small Area: \underline{Ciencias de la Computación y Tecnología Informática}}
-
 \begin{document}
@@
   This research project investigates a novel approach for software development, that leverages {\em Artificial Intelligence} (AI), to increase the automation of the programming process and hence, reduce the chances of introducing software bugs.
-The project proposes to address {\bf automatic program synthesis and validation starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine}.
-Our current work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on Artificial Intelligence).
+The project proposes to address {\bf program synthesis and program validation starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine}.
+
+  Our work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and has recently been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for publishing research results in AI.
+
 \end{abstract}
+{\footnotesize {\bf Keywords:} Computer Science, Artificial Intelligence, Automatic Programming, Program Validation, AI planning.}
 
-\newpage
-
-\section{Historial científico-técnico del Equipo de Investigadores (Apartado 5.1,bases de la convocatoria)}
-
-The project will be developed within the {\em Group of Reasoning on Planning and Scheduling} (GRPS, {\tt http://users.dsic.upv.es/grupos/grps/}), that is part of the {\em Department of Computer Systems and Computation} of the UPV (DSIC, {\tt www.dsic.upv.es}).
-
-In the last 10 years, the GRPS has led three research projects of the Spanish national research plan on {\em AI planning}.
-Further, the GRPS has developed its own AI planner, that is participating  at the International Planning Competition (IPC-2018, {\tt http://ipc2018.bitbucket.io}).
-
-The research group for developing this project would comprise three researchers: 1 principal investigator, 1 senior researcher and 1 PhD student. All of them are GRPS members.
-
-\subsection{Principal Investigator: Sergio Jiménez Celorrio}
-Sergio is a {\em Ramón y Cajal} fellow at the Universitat Politècnica de València (2017-2022).
-Previously, Sergio was a post-doc at the University of Melbourne and a {\em Juan de la Cierva} fellow at the Universitat Pompeu Fabra de Barcelona.
-From 2004 to 2013 Sergio was assistant lecturer at the Universidad Carlos III de Madrid where he obtained the {\em Distinguished Thesis Award 2011}.
-Sergio is co-organizer of the {\em $7^{th}$ International Planning Competition} and program committee at top international conferences on Artificial Intelligence (including ICAPS that is the main international conference on {\em AI planning}).
-
-His research addresses complex {\em decision-making} problems with innovative integrations of two {\em Artificial Intelligence} paradigms: {\em Machine Learning} and {\em AI Planning}.
-His research trajectory on this topic has produced a novel approach to the synthesis of programs, grammars and controllers for autonomous behavior and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence (IJCAI)\footnote{IJCAI is ranked as {\it A+} by the CORE rank and with {\it h5-index} 43 at Google Scholar}, the main international conference on {\em Artificial intelligence}.
-
-His top five scientific contributions are:
-\begin{enumerate}
-\item Inductive programming with AI planning{\footnotesize~\cite{javi-ijcai17},~\cite{segovia2016generalized},~\cite{jimenez2015computing}}.
-\item Generalization in AI planning{\footnotesize~\cite{javi-icaps17},~\cite{damir-derived-ijcai16},~\cite{javi-fsc-ijcai16}}.
-\item Learning to relieve the {\it knowledge acquisition bottleneck} in AI planning{\footnotesize{~\cite{diego-icaps18},~\cite{jimenez2013integrating},~\cite{jimenez2008architecture},~\cite{jimenez2006planning},~\cite{lanchas2007learning}}}.
-\item Machine learning for addressing the {\it curse of dimensionality} in AI planning{\footnotesize~\cite{jimenez2012review},~\cite{de2011scaling},~\cite{de2008learning}}.  
-\item Evaluation in AI planning{\footnotesize~\cite{lopez2015deterministic},~\cite{lopez2013automating},~\cite{coles2012survey}}. 
-\end{enumerate}
-
-General quality indicators of scientific research: {\bf 382 citations} (296 since 2013) and {\bf h-index 11}. {\scriptsize\em Source google scholar}.
-
-\subsection{Senior Researcher: Eva Onaindia}
-Eva Onaindia is a {\em Full Professor} at the Universitat Politècnica of València.
-She leads the GRPS where she conducts research in Artificial Intelligence, Automated Planning, Causal Model Learning and Recommender Systems.
-
-Eva has published numerous articles in top-tier specialized conferences such as the International Joint Conference on Artificial Intelligence (IJCAI), Autonomous Agents and Multi Agent Systems (AAMAS), International Conference on Planning and Scheduling (ICAPS) or European Conference on Artificial Intelligence (ECAI), among others; and scientific journals related to topics of Artificial Intelligence, Planning, Argumentation and Recommender Systems such as IEEE Intelligent Systems, Information Sciences, Engineering Applications of Artificial Intelligence, Knowledge and Information Systems or Applied Intelligence, among others.
-
-She has led several projects from the {\em National research plan} as well as served on multiple ICAPS, AAAI, IJCAI and ECAI conferences as a member of the Program Committee and as a senior member.
-Eva is Editor-in-Chief of the journal AI Communications, a Review Board member of the journal Applied Intelligence and she will be Program co-chair of ICAPS 2019 (the main international conference on AI planning).
-
-Eva Onaindía has supervised so far a total of nine PhD theses and she is currently supervising four doctoral students.
-She has also participated in many diverse scientific evaluation committees at both national and international level.
-In the national scene, she has collaborated with National agencies ANEP and ANECA and she has been member of different expert boards and scientific committees.
-At an international level, she has participated in the evaluation of research projects for Israel, Italy or Czech Republic.
-
-Eva has also participated in several R\&D and Innovation contracts: with the firm {\sc DESTINIA S.L.} ({\tt https://destinia.com/}) for the development of a tourist recommender system, and currently in the construction of a system for {\em electricity price forecasting} aimed at energy savings (Ayuntamiento Lliria) using neural networks and data analysis techniques.
-
-General quality indicators of scientific research: {\bf 1536 citations} (992 since 2013) and {\bf h-index 22}. {\scriptsize\em Source google scholar}.
-
-\subsection{Researcher in training period: Diego Aineto}
-Diego Aineto is a PhD student at the {\em Department of Computer Systems and Computation} of the Universitat Politècnica of València.
-
-Since 2016 Diego is holding a four-year FPU grant on {\em Causal discovery for predictive analysis} (FPU16/03184).
-Diego is also involved on the current research project TIN2017-88476-C2-1-R, funded by the Spanish national research plan and with title, {\em Reconocimiento de actividades y planificación automática para el diseño de asistentes inteligentes}.
-
-With Eva Onaindia and Sergio Jiménez, Diego is already co-author of one paper on the International Conference on Planning and Scheduling (ICAPS)~\cite{diego-icaps18}, and one journal paper~\cite{onaindia2018common}.
-Both papers on research topics related to the development of this project.
-
-\newpage
-\section{Estado del arte y objetivos específicos ( Apartado 5.2.i, bases de la convocatoria)}
-
-This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} paradigm for the automatic synthesis and validation of programs}.
-Here we briefly introduce this technology:
-\begin{itemize}
-\item {\bf AI Planning (AIP)} is the Artificial Intelligence component that studies the synthesis of sets of actions to achieve some given objectives~\cite{ghallab2004automated}.
-AIP arose in the late ’50s from converging studies into {\em combinatorial search}, {\em theorem proving} and {\em control theory} and now, is a well formalized paradigm for problem solving with algorithms that scale-up reasonably well.
-State-of-the-art planners are able to synthesize plans with hundreds of actions in seconds time~\cite{geffner2013concise}.
-The mainstream approach for AIP is {\em heuristic search} with heuristics derived automatically from the problem representation~\cite{mcdermott1996heuristic,bonet2001planning}.
-Current planners add other ideas to this like {\it novelty exploration}~\cite{geffner:psimulators:IJCAI17}, {\it helpful actions}~\cite{hoffmann2001ff}, {\it landmarks}~\cite{helmert2006fast}, and {\it multiqueue best-first search}~\cite{richter2010lama} for combining different heuristics.
-
-\item {\bf Test driven development (TDD)}~\cite{beck:TDD:2003} is a popular paradigm for software development that is frequently used in {\it agile methodologies}~\cite{cohen2003agile}.
-In TDD, test cases are created before the program code is written and they are run against the code during the development, e.g.
-after a code change via an automated process.
-When all tests pass, the program code is considered {\em complete} while when a test fails, it pinpoints a {\em bug} that must be fixed from the program code.
-Tests cases are a natural form of program specification, programmers often claim {\em 'code that is difficult to test is poorly written'}.
-Further, tests alert programmers of bugs before handing the code off to clients (the cost of finding a bug when the code is first written is considerably lower than the cost of detecting and fixing it later).
-%Last but not least, writing a thorough set of tests cases forces programmers to think through inputs, outputs, and error conditions of programs.
-\end{itemize}
+\section{Introduction}
+\label{sec:introduction}
 
@@
 
-\begin{center}
-{\small (Please find the full list of {\em references} at the end of the document)}
-\end{center}
+\section{Methodology}
+\label{sec:methodology}
+This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} paradigm for the automatic synthesis and validation of programs}.
+Here we briefly introduce this technology:
+\begin{itemize}
+\item {\bf AI Planning (AIP)} is the Artificial Intelligence component that studies the synthesis of sets of actions to achieve some given objectives~\cite{ghallab2004automated}.
+AIP arose in the late ’50s from converging studies into {\em combinatorial search}, {\em theorem proving} and {\em control theory} and now, is a well formalized paradigm for problem solving with algorithms that scale-up reasonably well.
+State-of-the-art planners are able to synthesize plans with hundreds of actions in seconds time~\cite{geffner2013concise}.
+The mainstream approach for AIP is {\em heuristic search} with heuristics derived automatically from the problem representation~\cite{mcdermott1996heuristic,bonet2001planning}.
+Current planners add other ideas to this like {\it novelty exploration}~\cite{geffner:psimulators:IJCAI17}, {\it helpful actions}~\cite{hoffmann2001ff}, {\it landmarks}~\cite{helmert2006fast}, and {\it multiqueue best-first search}~\cite{richter2010lama} for combining different heuristics.
 
-\newpage
-\section{Metodología y Plan de Trabajo ( Apartado 5.2.ii, bases de la convocatoria)}
-\label{subsec:metodologia}
-Our current research already shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16}.
+\item {\bf Test driven development (TDD)}~\cite{beck:TDD:2003} is a popular paradigm for software development that is frequently used in {\it agile methodologies}~\cite{cohen2003agile}.
+In TDD, test cases are created before the program code is written and they are run against the code during the development, e.g.
+after a code change via an automated process.
+When all tests pass, the program code is considered {\em complete} while when a test fails, it pinpoints a {\em bug} that must be fixed from the program code.
+Tests cases are a natural form of program specification, programmers often claim {\em 'code that is difficult to test is poorly written'}.
+Further, tests alert programmers of bugs before handing the code off to clients (the cost of finding a bug when the code is first written is considerably lower than the cost of detecting and fixing it later).
+%Last but not least, writing a thorough set of tests cases forces programmers to think through inputs, outputs, and error conditions of programs.
+\end{itemize}
+
+Our current research already shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
 Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list and traversing a binary tree.
@@
 
-%The space of the possible programs encoded in such a way grows exponentially in the number of possible program lines.
-To address challenging tasks our approach must then be combined with {\em problem decomposition}.
-With this regard, our encoding supports callable procedures to decompose a given programming task in simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.
+In both cases the performance of our approach will be evaluated with regard to the {\bf computation time} and {\bf memory} invested in the synthesize and the validation of the aimed programs.
 
-\subsection{Workplan}
+\subsection{Specific Objectives}
+\label{sec:objectivos}
+
+The specific objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of programs that are characterized by these tree dimensions:
+\begin{enumerate}
+\item Number of {\em program lines}.
+\item Number and domain of the observable {\em program variables}.
+\item Size of the available {\em instruction set}.
+\end{enumerate}
+
+For instance the program of Figure~\ref{fig:list}, for traversing linked lists, is generated using three program lines, one observable variable ($\mathsf{isEnd}(n)$ that has Boolean domain and indicates when $n$ points to the end of the given linked list), and an instruction set that comprises two instructions ($\mathsf{visit}(n)$, that marks the list node assigned to $n$ as {\em visited} and $\mathsf{step}(n)$, that advances $n$ to the next node in the linked list).
+
+In more detail, the two specific objectives of this project are:
+\begin{enumerate}
+\item To study the performance of our Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
+\item To study the performance of our Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}.
+\end{enumerate}
+
+Despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}.
+With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
+
+\section{Workplan}
+\label{sec:workplan}
+
 We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that (1), takes as input a set of test cases that specify the TDD programming task to solve and (2), outputs a program source code that passes these tests with a bug-free guarantee.
@@
 
-\newpage
-\section{Objetivos específicos del Proyecto ( Apartado 5.2.iii, bases de la convocatoria)}
-\label{subsec:objectivos}
+\subsection{Budget}
+Detailed desctiption of the two-year budget for the development of the proyect.
 
-The specific objective of this project is to study the performance of our AIP approach for the automated synthesis and validation of programs that are characterized by these tree dimensions:
-\begin{enumerate}
-\item Number of {\em program lines}.
-\item Number and domain of the observable {\em program variables}.
-\item Size of the available {\em instruction set}.
-\end{enumerate}
+\begin{table}[hbt!]
+\begin{small}  
+  \begin{tabular}{clc|l}
+    {\bf Priority} & {\bf Description} & {\bf Term} & {\bf Euros} \\\hline
+1 & {\scriptsize Difusión de las actividades del grupo} & 2018, 2019 & 2,000\\    
+2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2018 & 1,500\\
+2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2019 & 1,500\\
+3 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2018 & 1,500\\
+3 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2018 & 1,500\\\hline
+\multicolumn{3}{l|}{} & 8,000 \\
+  \end{tabular}
+\end{small}
+\caption{\small Two-year budget for the development of the proyect.}
+\end{table}
 
-For instance the program of Figure~\ref{fig:list}, for traversing linked lists, is generated using three program lines, one observable variable ($\mathsf{isEnd}(n)$ that has Boolean domain and indicates when $n$ points to the end of the given linked list), and an instruction set that comprises two instructions ($\mathsf{visit}(n)$, that marks the list node assigned to $n$ as {\em visited} and $\mathsf{step}(n)$, that advances $n$ to the next node in the linked list).
-
-In more detail, the two specific objectives of this project are:
-\begin{enumerate}
-\item To study the performance of our Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
-\item To study the performance of our Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}.
-\end{enumerate}
-In both cases the performance of our approach will be evaluated with regard to the {\bf computation time} and {\bf memory} invested in the synthesize and the validation of the aimed programs.
-
-Despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}.
-With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16}.
-
-\newpage
-\section{Beneficios del proyecto. Difusión y Explotación en su caso de los Resultados.}
+\section{Disemination and exploitation of the research results.}
 \label{subsec:beneficios}
@@
 
-In fact, our current work on {\em program synthesis} with AIP already produced several {\bf publications at top international conferences on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
+In fact, our current work on {\em program synthesis} with AIP already produced several {\bf publications at top international conferences and journals on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16,segovia:FSC:JAIR2018,segovia:programs:AIJ19} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
 

### 2018-09-07 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
first udpates to project proposal

@@
 
-  Our work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and has recently been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for publishing research results in AI.
+  Our work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and has recently been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for research in AI.
 
 \end{abstract}
-{\footnotesize {\bf Keywords:} Computer Science, Artificial Intelligence, Automatic Programming, Program Validation, AI planning.}
+{\footnotesize {\bf Keywords:} Computer Science, Artificial Intelligence, Program Synthesis, Program Validation, AI planning.}
 
@@
 
-{\em Program synthesis} is the task of computing a program that satisfies a given formal specification.
-In 2008 the PhD Thesis work by Armando Solar-Lezama, at the University of California Berkeley, showed that it is possible to encode program synthesis problems in Boolean logic and therefore, use {\em satisfiability modulo theories}~\cite{barrett:SMT:2009} to automatically compute programs~\cite{lezama2008program}.
-Since then, there has been a surge of practical interest in the idea of program synthesis in the formal verification community and related fields~\cite{alur2013syntax}.
-To illustrate this, in 2013, a unified framework for program synthesis problems was proposed and since 2014 there is a yearly program synthesis competition, comparing the different algorithms for program synthesis in a competitive event ({\tt http://www.sygus.org}).
-Further the US National Science Foundation is funding the ExCAPE research project\footnote{\tt https://www.nsf.gov/awardsearch/showAward?AWD\textunderscore ID=1138996}, to transform the way programmers develop software by advancing the theory and practice of software synthesis (since 2012 the ExCAPE research project is funded with a total amount of $\$3,750,000$).
+{\underline{Program Synthesis}} is the task of computing a program that satisfies a given formal specification of semantic correctness.
 
-Now we review the two most successful approaches for automated program synthesis. 
-\begin{itemize}
-\item{\bf Programming by Example (PbE)}.
-PbE techniques have already been deployed in the real world and are part of the {\sc Flash Fill} feature of Excel in Office 2013 that generates programs for string transformation~\cite{gulwani2011automating}.
-In this case the set of synthesized programs are represented succinctly in a restricted Domain-Specific Language (DSL) using a data-structure called version space algebras~\cite{mitchell1982generalization}.
-The programs are computed with a domain-specific search that implements a divide and conquer approach.
+The 2008 PhD Thesis work by Armando Solar-Lezama, at the University of California Berkeley, showed that it is possible to encode program synthesis as a {\em Boolean logic SAT} problem and therefore, use {\em satisfiability modulo theories}~\cite{barrett:SMT:2009} to automatically compute programs~\cite{lezama2008program}.
+Since then, there has been a surge of practical interest in the idea of program synthesis in the formal verification community and related fields.
 
-\item In {\bf Programming by Sketching (PbS)} programmers provide a partially specified program, i.e.~a program with the high-level structure of an implementation but that leaves low level details undefined to be determined by the synthesizer~\cite{solar2006combinatorial}.
-This form of program synthesis relies on a programming language called {\sc SKETCH}, for sketching partial programs.
-PbS implements a counterexample-driven iteration over a synthesize-validate loop built from two communicating SAT solvers, the inductive synthesizer and the validator, to automatically generate test inputs and ensure that the program satisfy them.
-Despite, in the worst case, program synthesis is harder than NP-complete, this counterexample-driven search terminates on many real problems after solving only a few SAT instances~\cite{lake2015human}.
-\end{itemize}
+To illustrate this interest, in 2013 the unified framework {\sc SyGuS} for program synthesis problems was proposed~\cite{alur2013syntax}.
+Since then, there is a yearly competition ({\tt http://www.sygus.org}) comparing different approaches for program synthesis.
+Further since 2012, the {\sc US National Science Foundation} is funding the ExCAPE research project\footnote{\tt https://www.nsf.gov/awardsearch/showAward?AWD\textunderscore ID=1138996}, to transform the way programmers develop software by advancing the theory and practice of software synthesis (the ExCAPE research project is funded with a total amount of $\$3,750,000$).
+Last but not last, program synthesis have already been deployed in the real world and is part of the {\sc Flash Fill} feature of {\sc Microsoft Excel} that generates programs for string transformation~\cite{gulwani2011automating}.
 
-On the other hand, {\em program validation} is the task of proving, or disproving, the correctness of a given program with respect to a certain formal specification or property.
-Program validation is considered a necessary step for program synthesis.
-{\em Model checking} is the mainstream AI approach for the formal validation of programs and controllers~\cite{clarke1999model}.
-In model checking a given model of a system (the program or controller to validate), is exhaustively and automatically checked to verify  whether this model meets a given specification.
-Typically the specification contains safety requirements such as the absence of deadlocks and similar critical states that can cause the system to crash.
-Both the model of the system and the specification are formulated in a formal language.
-Current approaches for model checking reduces to graph search.
-Instead of enumerating reachable states one at a time, the state space can sometimes be traversed more efficiently by considering large numbers of states at a single step.
-For instance, representing set of states and transition relations as logical formulas or binary decision diagrams, like in {\em symbolic model-checking} ~\cite{mcmillan1993symbolic}.
+Algorithmic approaches to program synthesis range over a wide spectrum, from {\em deductive synthesis} to {\em inductive synthesis}.
+In deductive program synthesis, a program is synthesized by constructively proving a theorem, employing logical inference and constraint solving~\cite{manna1986deductive}.
+On the other hand, inductive synthesis seeks to find a program matching a set of input-output examples~\cite{summers1977methodology,shapiro1983algorithmic}.
+It is thus an instance of learning from examples, also termed as {\em inductive inference} or {\em Machine Learning} (ML).
+Many current approaches to synthesis blend induction and deduction~\cite{seshia2015combining}; syntax guidance is usually a key ingredient in these approaches.
+Inductive synthesizers generalize from examples by searching a restricted space of programs.
+In ML, this space is called the {\em concept class}, and each element of that space is often called a {\em candidate concept}.
+The {\em concept class} is usually specified syntactically.
+Inductive learning is thus a natural fit for the syntax-guided synthesis problem introduced in this paper: the concept class is simply the set of permissible expressions.
 
-Closely related to the aims of the project is the work on the automated synthesis of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
-The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16}.
+Closely related to the aims of the project is the synthesis of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
+The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16,segovia:FSC:JAIR2018}.
 To keep the computation of FSCs tractable, the space of possible solutions is bound by the maximum size of the FSCs.
@@
 {\em classical planning}~\cite{sergio:aprograming:icaps16}, {\em conformant planning}~\cite{Geffner:FSM:AAAI10}, {\em CSP}~\cite{Infantes:FSC:ECAI2010} or a {\em Prolog program}~\cite{Giacomo:FSM:ICAPS13}).
-The validation and synthesis of programs from examples are research questions also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
+The synthesis of programs from examples is also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
 ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn logic programs from  examples and background knowledge, that are expressed as logic facts.
+
+{\underline{\em Program Validation}} is the task of proving, or disproving, the correctness of a given program with respect to a certain formal specification of its aimed semantics.
+Program validation is considered a necessary step for program synthesis.
+{\em Model checking} is the mainstream AI approach for the formal validation of programs and controllers~\cite{clarke1999model}.
+Current approaches for model checking reduces to graph search but, instead of enumerating reachable states one at a time, the state space is traversed considering large numbers of states at a single step.
+For instance, representing set of states and transition relations as logical formulas or binary decision diagrams, like in {\em symbolic model-checking} ~\cite{mcmillan1993symbolic}.
 
@@
 
-Our current research already shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
-Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list and traversing a binary tree.
+\subsection{Synthesis and validation of programs as AI planning}
+Our current research already shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia2017generating,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
+Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list and traversing a binary tree, or building a {\em parser} for simple arithmetic operations.
 
@@
 \begin{small}  
-\begin{tabular}{l@{\hspace*{30pt}}c@{\hspace*{5pt}}}
+\begin{tabular}{l@{\hspace*{30pt}}r@{\hspace*{5pt}}}
  \textbf{Programming Task} & \textbf{Time (seconds)} \\\hline
@@
 Sorting			&	30\\
-Tree  		&	165
+Tree  		        &	165\\
+Parser                  &       45
 \end{tabular}
@@
 
-\subsection{Synthesis and validation of TDD programs as AI planning}
+We now give more details of our method for program synthesis and validation.
 Given a {\em TDD programming} task, our approach is modeling and solving it as it were an {\em AI planning} task.
-
 Briefly the {\bf state variables} of this AIP task are {\em fluents} of the kind:
 \begin{itemize}
-\item {\tt var:=value}, that encode the program variables.
+\item {\tt var:=value}, that encode the value of the program variables.
 \item {\tt program(line):=instruction}, that encode the instructions at the different program lines.
-\item {\tt pcounter:=line}, encoding the current program line.
+\item {\tt pcounter:=line}, encoding which is the current program line.
 \end {itemize}
@@
 \begin{itemize}
-\item {\it Programming actions}, that assign an instruction to a given program line.
+\item {\it Programming actions}, that assign an instruction in the {\em instruction set} to a given program line.
 \item {\it Execution actions}, execute the instruction assigned to the current program line.
@@
 \label{sec:evaluation}
-The performance of our approach for the synthesis and validation of TDD programs will be evaluated in two kinds of programming tasks:
+
+The performance of our approach for the synthesis and validation of TDD programming tasks will be evaluated with regard to (1 )the {\bf computation time} and (2), the {\bf memory} invested in the synthesize and the validation of the aimed programs.
+
+We will evaluate our approach in programming tasks of two different kinds:
 \begin{itemize}
@@
 
-In both cases the performance of our approach will be evaluated with regard to the {\bf computation time} and {\bf memory} invested in the synthesize and the validation of the aimed programs.
-
 \subsection{Specific Objectives}
@@
 
-The specific objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of programs that are characterized by these tree dimensions:
+The programs objects of study in this project are characterized by these tree dimensions:
 \begin{enumerate}
@@
 
+The specific objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of TDD programming tasks coming from the domains described in the previous subsection.
 In more detail, the two specific objectives of this project are:
@@
 \subsection{Budget}
-Detailed desctiption of the two-year budget for the development of the proyect.
+Here we show the detailed desctiption of the two-year budget for the development of the proyect.
 
@@
 2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2018 & 1,500\\
-2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2019 & 1,500\\
-3 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2018 & 1,500\\
-3 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2018 & 1,500\\\hline
+3 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2019 & 1,500\\
+4 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2018 & 1,500\\
+5 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2019 & 1,500\\\hline
 \multicolumn{3}{l|}{} & 8,000 \\

### 2018-09-10 experiments/aij18/comparison_FAMA.sh
Added aij18 scripts to experiments folder

@@ -1,0 +1,25 @@
+#!/bin/bash
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "satellite" "transport" "visitall" "zenotravel")
+
+ulimit -t 1000
+rm results*
+rm sem-results*
+
+LANG=en_US
+
+for j in {0..100..10}
+do
+touch results.100.$j
+touch sem-results.100.$j
+
+#echo $j
+
+for d in "${arr[@]}"
+do
+   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 100 $j -t ten-observation -l 5 >> results.100.$j
+
+   ../../src/compiler_new.py -v learned_domain.pddl ../../benchmarks/aij18/$d/ 100 $j -t ten-observation -l 5 >> sem-results.100.$j
+
+done
+
+done

### 2018-09-10 experiments/aij18/minimal_100_10.sh

@@ -1,0 +1,19 @@
+#!/bin/bash
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "satellite" "transport" "visitall" "zenotravel")
+
+ulimit -t 1000
+rm results*
+rm sem-results*
+
+LANG=en_US
+
+touch results.100.10
+touch sem-results.100.10
+#echo $j
+
+for d in "${arr[@]}"
+do
+   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 100 10 -t ten-observation -l 2 >> results.100.10
+
+   ../../src/compiler_new.py -v learned_domain.pddl ../../benchmarks/aij18/$d/ 100 10 -t ten-observation -l 5 >> sem-results.100.10 
+done

### 2018-09-10 experiments/aij18/minimal_30_30.sh

@@ -1,0 +1,16 @@
+#!/bin/bash
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "satellite" "transport" "visitall" "zenotravel")
+
+ulimit -t 1000
+rm results*
+
+LANG=en_US
+
+touch results.30.30
+#echo $j
+
+for d in "${arr[@]}"
+do
+   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 30 30 -t ten-observation -l 2  >> results.30.30
+
+done

### 2018-09-10 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updated introduction project proposal

@@
 
-  Our work on this particular research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and has recently been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for research in AI.
+  Our current work on this research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and recently, it is been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for research in AI.
 
@@
 
-{\underline{Program Synthesis}} is the task of computing a program that satisfies a given formal specification of semantic correctness.
+{\underline{\em Program Synthesis}} is the task of computing a program such that it satisfies a given formal specification of semantic correctness.
 
@@
 
-To illustrate this interest, in 2013 the unified framework {\sc SyGuS} for program synthesis problems was proposed~\cite{alur2013syntax}.
-Since then, there is a yearly competition ({\tt http://www.sygus.org}) comparing different approaches for program synthesis.
-Further since 2012, the {\sc US National Science Foundation} is funding the ExCAPE research project\footnote{\tt https://www.nsf.gov/awardsearch/showAward?AWD\textunderscore ID=1138996}, to transform the way programmers develop software by advancing the theory and practice of software synthesis (the ExCAPE research project is funded with a total amount of $\$3,750,000$).
-Last but not last, program synthesis have already been deployed in the real world and is part of the {\sc Flash Fill} feature of {\sc Microsoft Excel} that generates programs for string transformation~\cite{gulwani2011automating}.
+To illustrate this interest, since 2012 the {\sc US National Science Foundation} is funding the ExCAPE research project\footnote{\tt https://www.nsf.gov/awardsearch/showAward?AWD\textunderscore ID=1138996} (an amount of $\$3,750,000$) to advance the theory and practice of program synthesis.
+In 2013 a unified framework for program synthesis was defined~\cite{alur2013syntax}.
+Since then, there is a yearly international competition ({\tt http://www.sygus.org}) to compare the different approaches for program synthesis.
+Last but not least, program synthesis has already been deployed in the real world and it is part of the {\sc Flash Fill} feature of {\sc Microsoft Excel} that automatically generates programs for string transformation~\cite{gulwani2011automating}.
 
@@
 In deductive program synthesis, a program is synthesized by constructively proving a theorem, employing logical inference and constraint solving~\cite{manna1986deductive}.
-On the other hand, inductive synthesis seeks to find a program matching a set of input-output examples~\cite{summers1977methodology,shapiro1983algorithmic}.
+On the other hand, inductive synthesis seeks to find a program matching a set of input-output examples by searching in a restricted space of programs~\cite{summers1977methodology,shapiro1983algorithmic}.
 It is thus an instance of learning from examples, also termed as {\em inductive inference} or {\em Machine Learning} (ML).
 Many current approaches to synthesis blend induction and deduction~\cite{seshia2015combining}; syntax guidance is usually a key ingredient in these approaches.
-Inductive synthesizers generalize from examples by searching a restricted space of programs.
-In ML, this space is called the {\em concept class}, and each element of that space is often called a {\em candidate concept}.
-The {\em concept class} is usually specified syntactically.
-Inductive learning is thus a natural fit for the syntax-guided synthesis problem introduced in this paper: the concept class is simply the set of permissible expressions.
 
 Closely related to the aims of the project is the synthesis of {\it Finite State Controllers} (FSCs)~\cite{geffner:policies:IJCAI15}.
-The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it~\cite{sergio:aprograming:ijcai16,segovia:FSC:JAIR2018}.
+The state-of-the-art algorithms for computing FSCs follow a {\it top-down} approach that interleaves {\it programming} the FSC with {\em validating} it~\cite{sergio:aprograming:ijcai16,segovia:FSC:JAIR2018}.
 To keep the computation of FSCs tractable, the space of possible solutions is bound by the maximum size of the FSCs.
@@
 The synthesis of programs from examples is also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
-ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn logic programs from  examples and background knowledge, that are expressed as logic facts.
+ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn {\em logic programs} from examples and background knowledge, that are expressed as logic facts.
 
-{\underline{\em Program Validation}} is the task of proving, or disproving, the correctness of a given program with respect to a certain formal specification of its aimed semantics.
+{\underline{\em Program Validation}} is the task of proving (or disproving) the correctness of a given program with respect to a given formal specification of its aimed semantics.
 Program validation is considered a necessary step for program synthesis.
@@
 For instance, representing set of states and transition relations as logical formulas or binary decision diagrams, like in {\em symbolic model-checking} ~\cite{mcmillan1993symbolic}.
+
+Program validation is also compilable into classical planning.
+Examples are the compilations for {\em {\sc GOLOG} procedures}~\cite{baier2007exploiting}, {\em planning programs}~\cite{segovia:programs:AIJ19}, or {\em Finite State Controllers}~\cite{Geffner:FSM:AAAI10,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018}.
+Briefly, these compilations encode the {\em cross product} of a given planning instance and the automata corresponding to the program to validate.
+In this approach a {\em validation proof} is provided if a solution is found (the program is {\em correct}).
+Otherwise, if a classical planner proves that no solution exists, then the program is {\em incorrect} (its execution necessarily failed).
+When actions have non-deterministic effects, program validation becomes more complex since it requires proving that all the possible program executions reach the goals.
+In such a scenario, {\em model checking} is a more suitable approach.
+\newpage
 

### 2018-09-10 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updated methodology project proposal

@@
 This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} paradigm for the automatic synthesis and validation of programs}.
-Here we briefly introduce this technology:
+In more detail, we use {\em test cases} to specify the semantic correctness of programs, as in Test Driven Development.
+Then, we encode the program synthesis and program validation tasks as AI planning problems and finally, we use off-the-shelf AI planners to compute the solution to these problems.
+
+Our current research already shows that this method can synthesize and validate programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia2017generating,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
+Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list and traversing a binary tree, or building a {\em parser} for simple arithmetic operations.
+
+\begin{table*}[hbt!]
+  \centering
+\begin{small}  
+\begin{tabular}{l@{\hspace*{30pt}}|r@{\hspace*{5pt}}}
+ \textbf{Programming Task} & \textbf{Time (seconds)} \\\hline
+Summatory		&	1\\
+Fibonacci		&	5\\
+Reverse			&	22\\
+Find                    &       336 \\
+Minimum                 &       284 \\
+Sorting			&	30\\
+Tree  		        &	165\\
+Parser                  &       45
+\end{tabular}
+\end{small}  
+\caption{\small Time to synthesize the programs with the AI planner {\sc FD}~\cite{helmert2006fast} on a processor {\em Intel Core i5 3.10GHz x 4} and with a 4GB memory bound.}
+\label{tab:programs}
+\end{table*}
+
+\subsection{Background}
+Here we briefly introduce the technology required by our AI approach for the automated synthesis and validation of programs:
 \begin{itemize}
@@
 \subsection{Synthesis and validation of programs as AI planning}
-Our current research already shows that AI planners can synthesize programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia2017generating,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
-Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list and traversing a binary tree, or building a {\em parser} for simple arithmetic operations.
+We now give more details of our method for program synthesis and validation.
+Given a {\em TDD programming} task, our approach is modeling and solving it as it were an {\em AI planning} problem.
+Formally, an AI planning problem is defined as a tuple $\tup{X,D,A,I,G}$ where:
+\begin{itemize}
+\item $X=\tup{x_0, \ldots, x_n}$ is the set of $n$ {\em state variables} with finite domain.
+The respective domains are $D=\tup{D_{x_0}, \ldots, D_{x_n}}$ defining the set of possible values for each variable $x\in X$.
+\item $A$ is a set of actions whose dynamics are specified with two functions: $App(s)\subseteq A$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)=s'$ that denotes the {\em successor state} that results of applying $a$ in $s$.
+\item $I$ is an initial state, i.e. a full assigment of values to the state variables.
+\item $G$ is the set of goal conditions bounding the possible values of the variables in $X$ and defining the set of goal states.  
+\end{itemize}
+A solution to an AI planning problem is a sequence of applicable actions such that, starting from the initial state, its application reach a state where all the goal conditions are met.
 
-\begin{table*}[hbt!]
-  \centering
-\begin{small}  
-\begin{tabular}{l@{\hspace*{30pt}}r@{\hspace*{5pt}}}
- \textbf{Programming Task} & \textbf{Time (seconds)} \\\hline
-Summatory		&	1\\
-Fibonacci		&	5\\
-Reverse			&	22\\
-Find                    &       336 \\
-Minimum                 &       284 \\
-Sorting			&	30\\
-Tree  		        &	165\\
-Parser                  &       45
-\end{tabular}
-\end{small}  
-\caption{\small Time to synthesize the programs with the AI planner {\sc FD}~\cite{helmert2006fast} on a processor {\em Intel Core i5 3.10GHz x 4} and with a 4GB memory bound.}
-\label{tab:programs}
-\end{table*}
-
-We now give more details of our method for program synthesis and validation.
-Given a {\em TDD programming} task, our approach is modeling and solving it as it were an {\em AI planning} task.
-Briefly the {\bf state variables} of this AIP task are {\em fluents} of the kind:
+Now we detail our encoding of the program synthesis task as an AI planning problem. Briefly the {\bf state variables} of this AIP task are {\em fluents} of the kind:
 \begin{itemize}
-\item {\tt var:=value}, that encode the value of the program variables.
-\item {\tt program(line):=instruction}, that encode the instructions at the different program lines.
-\item {\tt pcounter:=line}, encoding which is the current program line.
+\item {\tt var := value}, that encode the value of the program variables.
+\item {\tt program(line) := instruction}, that encode the instructions at the different program lines.
+\item {\tt pcounter := line}, encoding which is the current program line.
 \end {itemize}
@@
 
-The performance of our approach for the synthesis and validation of TDD programming tasks will be evaluated with regard to (1 )the {\bf computation time} and (2), the {\bf memory} invested in the synthesize and the validation of the aimed programs.
+The performance of our approach for the synthesis and validation of TDD programming tasks will be evaluated with regard to (1) the {\bf computation time} and (2), the {\bf memory} invested in the synthesize and the validation of the aimed programs.
 

### 2018-09-10 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updated methodology project proposal

@@
 \subsection{Synthesis and validation of programs as AI planning}
-We now give more details of our method for program synthesis and validation.
-Given a {\em TDD programming} task, our approach is modeling and solving it as it were an {\em AI planning} problem.
-Formally, an AI planning problem is defined as a tuple $\tup{X,D,A,I,G}$ where:
+Here we explain the details of our method for program synthesis and validation. Given a {\em TDD programming} task, our approach is modeling and solving it as it were an {\em AI planning} problem.
+
+\subsubsection{The AI planning problem}
+An {\em AI planning problem} is defined as a tuple $\tup{V,D,A,I,G}$ where:
 \begin{itemize}
-\item $X=\tup{x_0, \ldots, x_n}$ is the set of $n$ {\em state variables} with finite domain.
-The respective domains are $D=\tup{D_{x_0}, \ldots, D_{x_n}}$ defining the set of possible values for each variable $x\in X$.
-\item $A$ is a set of actions whose dynamics are specified with two functions: $App(s)\subseteq A$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)=s'$ that denotes the {\em successor state} that results of applying $a$ in $s$.
-\item $I$ is an initial state, i.e. a full assigment of values to the state variables.
-\item $G$ is the set of goal conditions bounding the possible values of the variables in $X$ and defining the set of goal states.  
+\item $V=\tup{v_0, \ldots, v_n}$ is the set of $n$ {\em state variables} with finite domain.
+The respective domains are $D=\tup{D_{v_0}, \ldots, D_{v_n}}$ defining the set of possible values for each variable $v\in V$.
+\item $A$ is a set of actions whose dynamics are specified with two functions:
+\begin{itemize}
+\item $App(s)\subseteq A$, defines the subset of actions applicable in a state $s$.
+\item $\theta(s,a)=s'$, that defines the {\em successor state} $s'$ that results of applying $a$ in $s$.
+\end{itemize}  
+\item $I$ is an initial state, i.e. a full assignment of values to the state variables.
+\item $G$ is the set of goal conditions bounding the possible values of the state variables in the goal states.  
 \end{itemize}
@@
 
-Now we detail our encoding of the program synthesis task as an AI planning problem. Briefly the {\bf state variables} of this AIP task are {\em fluents} of the kind:
+\subsubsection{The TDD programming task}
+We define a {\em TDD programming task} as a programming task whose semantic correctness is specified with a set of {\em test cases}.
+Formally a {\em TDD programming} task is a tuple $\tup{{\mathcal V},{\mathcal D},{\mathcal A}, {\mathcal T}}$ where:
 \begin{itemize}
-\item {\tt var := value}, that encode the value of the program variables.
+\item ${\mathcal V}$ is the set of program variables.
+The respective domains are ${\mathcal D}=\tup{{\mathcal D}_{v_0}, \ldots, {\mathcal D}_{v_n}}$ defining the set of possible values for each variable $v\in V$.
+\item ${\mathcal A}$ is the instruction set. That is the set of instructions that can appear in a program.
+\item ${\mathcal T}$ is a set of test cases where each test $t\in {\mathcal T}$ is a pair $t=\tup{{\mathcal I}_t,{\mathcal G}_t}$ that represent the input value for the program variables and the aimed output for these variables.
+\end{itemize}
+A solution to {\em TDD programming task} is a program whose execution succeeds to solve all the test cases given as input.
+
+\subsubsection{TDD programming as AI planning}
+Our encoding of the program synthesis task as an AI planning problem defines {\bf state variables} of three kinds:
+\begin{itemize}
 \item {\tt program(line) := instruction}, that encode the instructions at the different program lines.
 \item {\tt pcounter := line}, encoding which is the current program line.
+\item {\tt var := value}, that encode the value of the program variables.
 \end {itemize}
@@
 \begin{tikzpicture}
-\node [thick,draw,circle] at (1,1.1) (x) {$q_0$};
-\node [thick,draw,circle] at (4,1.1) (y) {$q_1$};
-\node [thick,draw,circle] at (7,1.1) (v) {$q_2$};
-\node [thick,draw,circle] at (10,1.1) (w) {$q_3$};
-\node [thick,draw,circle] at (13,1.1) (z) {$q_3$};
-\draw (13,1.1) circle (.3cm);
+\node [thick,draw,circle] at (1,1.1) (a) {$q_0$};
+\node [thick,draw,circle] at (4,1.1) (b) {$q_1$};
+\node [thick,draw,circle] at (7,1.1) (c) {$q_2$};
+\node [thick,draw,circle] at (10,1.1) (d) {$q_3$};
+\node [thick,draw,circle] at (13,1.1) (e) {$q_4$};
+\node [thick,draw,circle, below of=e]  (f) {$q_5$};
+\draw (13,0.1) circle (.3cm);
 
-\draw [thick,->] (x) to (y);
-\draw [thick,->] (y) to (v);
-\draw [thick,->] (v) to (w);
-\draw [thick,->] (w) to (z);
+\draw [thick,->] (a) to (b);
+\draw [thick,->] (b) to (c);
+\draw [thick,->] (c) to (d);
+\draw [thick,->] (d) to (e);
+\draw [thick,->, bend left] (e) to (a);
+\draw [thick,->] (e) to (f);
 
-\node at (2.4,1.5) {$\mathsf{take}(bread)$};
-\node at (5.4,1.5) {$\mathsf{use}(toaster)$};
-\node at (8.4,1.5) {$\mathsf{take}(knife)$};
-\node at (11.4,1.5) {$\mathsf{spread}(butter)$};
+\node at (2.4,1.5) {$\mathsf{take}(orange)$};
+\node at (5.4,1.5) {$\mathsf{take}(knife)$};
+\node at (8.4,1.5) {$\mathsf{split}(orange)$};
+\node at (11.4,1.5) {$\mathsf{squeeze}(orange)$};
 
@@
 \node at (4.1,0.85) {$!\mathsf{isEnd}(n)/\mathsf{visit}(n)$};
-\node at (3.5,1.8) {$-/\mathsf{step}(n)$};
+\node at (3.5,1.8) {$\mathsf{step}(n)$};
 \node at (6,0) {$\mathsf{isEnd}(n)/a_\bot$};

### 2018-09-10 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updated automatae project proposal

@@
 \usepackage{tikz}
+\usetikzlibrary{arrows,automata}
 \usetikzlibrary{calc}
@@
 \begin{itemize}
+\item {\em Theoretical benchmarks}: Classic programming tasks are a neat touchstone to assess the performance of our approach.
+For instance, programs for the computation of mathematical/logic series, string manipulation and for the management of data structures such as {\em lists}, {\em queues}, {\em stacks} or {\em trees}.
+Figure~\ref{fig:list} shows a synthesized program for computing $y = \sum_0^N x$.
+The program is pictured as a {\em finite state machine}: The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
+The program assumes there are three variables and the Boolean variables resulting from the comparison of these values ${\mathcal V}=\{n,x,y,x=y,x=n,y=n,\}$.
+The instruction set is ${\mathcal A}=\{x = x + 1, y = y + 1, n = n + 1, y = y + x, y = y + n, x = x + y, x = x + n, n = n + x, n = n + y\}$.
+The input test cases are $0 = \sum_0^0 x$, $1 = \sum_0^1 1$, $3 = \sum_0^2 x$, $6 = \sum_0^3 x$, $10 = \sum_0^4 x$ so the domains of variables $x,y$ and $n$ is [0,10]:
+
+\begin{figure}[hbt!]
+  \begin{center}
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=4cm,semithick]
+	  \node[state] (A)              {$Q_0$};
+	  \node[state] (B) [right of=A] {$Q_1$};
+	  \node[state] (C) [right of=B] {$Q_2$};
+	  \node[state] (D) [right of=C] {$Q_3$};          
+	  \path
+          (A) edge [align=center, bend left] node {$x=N$/\\ - } (D)          
+          (A) edge [align=center] node {$x\neq N$ /\\y = y + x} (B)
+          (B) edge [align=center] node { - /\\ $x = x + 1$} (C)
+          (C) edge [align=center] node {$x=N$/\\ - } (D)
+          (C) edge [align=center, bend left] node {$x\neq N$/\\ - } (A)
+          ;          
+	\end{tikzpicture}
+\end{center}
+\caption{\small Three-line program to solve the programming task of computing $y = \sum_0^N x$.}
+\label{fig:list}
+\end{figure}
+
 \item {\em Real-world benchmarks}: The GRPS group is currently leading the four-year research project TIN2017-88476-C2-1-R from the {\em Spanish national plan} in which {\em AI Planning} and {\em activity recognition} is applied to different real-world domains such as {\em domotics}, {\em tourism}, {\em traffic control} and {\em robotics}.
 Interestingly many activities in these domains can be understood as simple programs.
-For instance, Figure~\ref{fig:activity} shows a four-line program (pictured as a finite state machine) that represents the sequence of instructions required for the {\em making a buttered toast} activity.
+For instance, Figure~\ref{fig:activity} shows a four-line program (pictured as a finite state machine) that represents the sequence of instructions required for the activity of {\em preparing an orange juice} .
 We plan to evaluate our approach in synthesis and validation tasks coming from these real-world domains.
@@
   \begin{center}
-\begin{tikzpicture}
-\node [thick,draw,circle] at (1,1.1) (a) {$q_0$};
-\node [thick,draw,circle] at (4,1.1) (b) {$q_1$};
-\node [thick,draw,circle] at (7,1.1) (c) {$q_2$};
-\node [thick,draw,circle] at (10,1.1) (d) {$q_3$};
-\node [thick,draw,circle] at (13,1.1) (e) {$q_4$};
-\node [thick,draw,circle, below of=e]  (f) {$q_5$};
-\draw (13,0.1) circle (.3cm);
-
-\draw [thick,->] (a) to (b);
-\draw [thick,->] (b) to (c);
-\draw [thick,->] (c) to (d);
-\draw [thick,->] (d) to (e);
-\draw [thick,->, bend left] (e) to (a);
-\draw [thick,->] (e) to (f);
-
-\node at (2.4,1.5) {$\mathsf{take}(orange)$};
-\node at (5.4,1.5) {$\mathsf{take}(knife)$};
-\node at (8.4,1.5) {$\mathsf{split}(orange)$};
-\node at (11.4,1.5) {$\mathsf{squeeze}(orange)$};
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.3cm,semithick]
+	  \node[state] (A)              {$Q_0$};
+	  \node[state] (B) [right of=A] {$Q_1$};
+	  \node[state] (C) [right of=B] {$Q_2$};
+	  \node[state] (D) [right of=C] {$Q_3$};
+	  \node[state] (E) [right of=D] {$Q_4$};
+	  \node[state] (F) [below of=E] {$Q_5$};                    
+	  \path
+          (A) edge [align=center] node { - /\\ $\mathsf{take}(orange)$} (B)          
+          (B) edge [align=center] node { - /\\$\mathsf{take}(knife)$} (C)
+          (C) edge [align=center] node { - /\\ $\mathsf{split}(orange)$} (D)
+          (D) edge [align=center] node { - /\\ $\mathsf{squeeze}(orange)$} (E)
+          (E) edge [align=center] node { ready(juice) /\\ - } (F)
+          (E) edge [align=center, bend left] node { $\neg ready(juice)$ /\\ - } (A);
 
@@
 \end{center}
-\caption{\small Four-line program representing the activity of {\em making a buttered toast}.}
+\caption{\small Five-line program representing the activity of {\em preparing an orange juice}.}
 \label{fig:activity}
@@
 
-\item {\em Theoretical benchmarks}: Classic programming tasks are a neat touchstone to assess the performance of our approach.
-For instance, programs for the computation of mathematical/logic series, string manipulation and for the management of data structures such as {\em lists}, {\em queues}, {\em stacks} or {\em trees}.
-Figure~\ref{fig:list} shows a synthesized program for traversing a linked list.
-The program is pictured as a {\em finite state machine}: The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
 \end{itemize}
-
-\begin{figure}[hbt!]
-\begin{center}
-\begin{tikzpicture}
-\node [thick,draw,circle] at (1,1.1) (x) {$q_0$};
-
-\node [thick,draw,circle] at (6,1.1) (y) {$q_1$};
-
-\node [thick,draw,circle] at (11,1.1) (z) {$q_2$};
-
-\draw (11,1.1) circle (.3cm);
-
-\draw [thick,->] (x) -- (y);
-
-\draw [thick,out=345,in=195,->] (x) to (z);
-
-\draw [thick,out=165,in=15,->] (y) to (x);
-
-\node at (4.1,0.85) {$!\mathsf{isEnd}(n)/\mathsf{visit}(n)$};
-\node at (3.5,1.8) {$\mathsf{step}(n)$};
-\node at (6,0) {$\mathsf{isEnd}(n)/a_\bot$};
-\end{tikzpicture}
-\end{center}
-\caption{\small Synthesized program to solve the programming task of traversing a linked list ($n$ is a variable that points to a node of the linked list).}
-\label{fig:list}
-\end{figure}
 

### 2018-09-11 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updated evaluation project proposal

@@
 This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} paradigm for the automatic synthesis and validation of programs}.
-In more detail, we use {\em test cases} to specify the semantic correctness of programs, as in Test Driven Development.
-Then, we encode the program synthesis and program validation tasks as AI planning problems and finally, we use off-the-shelf AI planners to compute the solution to these problems.
+In more detail, we use {\em test cases} to specify the semantic of the aimed programs, as in Test Driven Development.
+Then, we encode the {\em program synthesis} (and {\em program validation}) tasks as AI planning problems and finally, we use off-the-shelf AI planners to compute solutions to these problems.
 
 Our current research already shows that this method can synthesize and validate programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia2017generating,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
-Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list and traversing a binary tree, or building a {\em parser} for simple arithmetic operations.
+Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list, traversing a binary tree, or building a {\em parser} for simple arithmetic operations.
 
@@
 \subsection{Background}
-Here we briefly introduce the technology required by our AI approach for the automated synthesis and validation of programs:
+First we briefly introduce the technology required by our AI method for the automated synthesis and validation of programs:
 \begin{itemize}
@@
 after a code change via an automated process.
-When all tests pass, the program code is considered {\em complete} while when a test fails, it pinpoints a {\em bug} that must be fixed from the program code.
+When all tests pass, the program code is considered {\em correct} while when a test fails, it pinpoints a {\em bug} that must be fixed from the program code.
 Tests cases are a natural form of program specification, programmers often claim {\em 'code that is difficult to test is poorly written'}.
@@
 \subsection{Synthesis and validation of programs as AI planning}
-Here we explain the details of our method for program synthesis and validation. Given a {\em TDD programming} task, our approach is modeling and solving it as it were an {\em AI planning} problem.
+Here we explain the details of our AI method for program synthesis and validation. Given a {\em TDD programming} task, our approach is modeling and solving it as it were an {\em AI planning} problem.
 
@@
 \item $V=\tup{v_0, \ldots, v_n}$ is the set of $n$ {\em state variables} with finite domain.
-The respective domains are $D=\tup{D_{v_0}, \ldots, D_{v_n}}$ defining the set of possible values for each variable $v\in V$.
-\item $A$ is a set of actions whose dynamics are specified with two functions:
+The respective domains are $D=\tup{D_{v_0}, \ldots, D_{v_n}}$ defining, for each variable $v\in V$, its set of possible values.
+\item $A$ is a set of {\em actions} whose dynamics are specified with two functions:
 \begin{itemize}
 \item $App(s)\subseteq A$, defines the subset of actions applicable in a state $s$.
-\item $\theta(s,a)=s'$, that defines the {\em successor state} $s'$ that results of applying $a$ in $s$.
+\item $\theta(s,a)=s'$, that defines the {\em successor state} $s'$ that results of applying an action $a$ in a state $s$.
 \end{itemize}  
-\item $I$ is an initial state, i.e. a full assignment of values to the state variables.
-\item $G$ is the set of goal conditions bounding the possible values of the state variables in the goal states.  
+\item $I$ is an {\em initial state}, i.e. a full assignment of values to the state variables.
+\item $G$ is the set of {\em goal conditions} bounding the possible values of the state variables in the goal states.  
 \end{itemize}
-A solution to an AI planning problem is a sequence of applicable actions such that, starting from the initial state, its application reach a state where all the goal conditions are met.
+A {\em solution} to an AI planning problem is a sequence of applicable actions such that its application, starting from the initial state, reach a state where all goal conditions are met.
 
 \subsubsection{The TDD programming task}
-We define a {\em TDD programming task} as a programming task whose semantic correctness is specified with a set of {\em test cases}.
-Formally a {\em TDD programming} task is a tuple $\tup{{\mathcal V},{\mathcal D},{\mathcal A}, {\mathcal T}}$ where:
+We formalize a {\em TDD programming task} as a programming task whose semantic correctness is specified with a set of input-output {\em test cases}.
+A {\em TDD programming} task is a tuple $\tup{{\mathcal V},{\mathcal D},{\mathcal A}, {\mathcal T}}$ where:
 \begin{itemize}
-\item ${\mathcal V}$ is the set of program variables.
-The respective domains are ${\mathcal D}=\tup{{\mathcal D}_{v_0}, \ldots, {\mathcal D}_{v_n}}$ defining the set of possible values for each variable $v\in V$.
-\item ${\mathcal A}$ is the instruction set. That is the set of instructions that can appear in a program.
-\item ${\mathcal T}$ is a set of test cases where each test $t\in {\mathcal T}$ is a pair $t=\tup{{\mathcal I}_t,{\mathcal G}_t}$ that represent the input value for the program variables and the aimed output for these variables.
+\item ${\mathcal V}$ is the set of $m$ {\em program variables}.
+The respective domains are ${\mathcal D}=\tup{{\mathcal D}_{v_0}, \ldots, {\mathcal D}_{v_m}}$ defining the set of possible values for each variable $v\in {\mathcal V}$.
+\item ${\mathcal A}$ is the {\em instruction set}, i.e. the set of different instructions that can appear in a program.
+\item ${\mathcal T}$ is a set of input-output {\em test cases} where each test $t\in {\mathcal T}$ is a pair $t=\tup{{\mathcal I}_t,{\mathcal G}_t}$ that represent the input value for the program variables and the aimed output for these variables.
 \end{itemize}
-A solution to {\em TDD programming task} is a program whose execution succeeds to solve all the test cases given as input.
+A {\em solution} to TDD programming task is a program whose execution succeeds to solve every given test case.
 
-\subsubsection{TDD programming as AI planning}
-Our encoding of the program synthesis task as an AI planning problem defines {\bf state variables} of three kinds:
+\subsubsection{Program synthesis and validation as AI planning}
+Here we describe how to encode a {\em TDD programming task} task as an AI planning problem. Our encoding defines {\bf state variables} of three kinds:
 \begin{itemize}
-\item {\tt program(line) := instruction}, that encode the instructions at the different program lines.
-\item {\tt pcounter := line}, encoding which is the current program line.
-\item {\tt var := value}, that encode the value of the program variables.
+\item {\tt program(line) := instruction}, that encode the instructions (from the instruction set ${\mathcal A}$) at the different program lines. Initialy all program lines are empty.
+\item {\tt pcounter := line}, encoding which is the current program line. Initially the program counter points to the first program line.
+\item {\tt var := value}, that encode the value of the program variables (the ${\mathcal V}$ set).
 \end {itemize}
 
-The {\bf initial/goal states} of the AIP task encode the initial/final values of the program variables and are given by the input/output tests of the TDD programming task.
-
-Finally, program instructions are encoded using two kinds of AIP {\bf actions}:
+The {\bf initial/goal states} of the AIP task encode the initial/final values of the program variables.
+These values are specified by the test cases ${\mathcal T}$ of the TDD programming task.
+Finally, every program instruction ($w\in {\mathcal A}$) is encoded with two AIP {\bf actions}:
 \begin{itemize}
-\item {\it Programming actions}, that assign an instruction in the {\em instruction set} to a given program line.
+\item {\it Programming actions} that assign an instruction in the {\em instruction set} to a given program line, i.e. change the value of variable {\tt program(line)}.
 \item {\it Execution actions}, execute the instruction assigned to the current program line.
@@
 We implement this encoding using standard planning languages, such as PDDL~\cite{fox2003pddl2}, so the AIP tasks resulting from our encoding can be solved with off-the-shelf planners, like the {\sc FD} planning system~\cite{helmert2006fast}.
-The programs synthesized with this approach are guaranteed to be bug-free over given sets of {\em input-output} tests cases.
+Our encoding is {\em complete} and {\em correct}~\cite{segovia:programs:AIJ19}, which means that the programs synthesized with this approach are guaranteed to be bug-free over given sets of {\em input-output} tests cases.
 
@@
 
-The performance of our approach for the synthesis and validation of TDD programming tasks will be evaluated with regard to (1) the {\bf computation time} and (2), the {\bf memory} invested in the synthesize and the validation of the aimed programs.
+The performance of our AI method for the synthesis and validation of TDD programming tasks will be evaluated with regard to (1) {\bf computation time} and (2), {\bf memory} invested in the synthesize and the validation of the aimed programs.
 
-We will evaluate our approach in programming tasks of two different kinds:
+The programming tasks used for the evaluation are of two different kinds:
 \begin{itemize}
@@
 For instance, programs for the computation of mathematical/logic series, string manipulation and for the management of data structures such as {\em lists}, {\em queues}, {\em stacks} or {\em trees}.
-Figure~\ref{fig:list} shows a synthesized program for computing $y = \sum_0^N x$.
+Figure~\ref{fig:sum} shows a synthesized program for computing $y = \sum_0^N x$.
 The program is pictured as a {\em finite state machine}: The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
-The program assumes there are three variables and the Boolean variables resulting from the comparison of these values ${\mathcal V}=\{n,x,y,x=y,x=n,y=n,\}$.
-The instruction set is ${\mathcal A}=\{x = x + 1, y = y + 1, n = n + 1, y = y + x, y = y + n, x = x + y, x = x + n, n = n + x, n = n + y\}$.
-The input test cases are $0 = \sum_0^0 x$, $1 = \sum_0^1 1$, $3 = \sum_0^2 x$, $6 = \sum_0^3 x$, $10 = \sum_0^4 x$ so the domains of variables $x,y$ and $n$ is [0,10]:
+The program in Figure~\ref{fig:sum} assumes that the program variables are ${\mathcal V}=\{x,y,x=N,y=N\}$ and that the value of $x$ and $y$ is initially 0.
+The instruction set is ${\mathcal A}=\{x = x + 1, y = y + 1, x = x + y, y = y + x\}$.
+The input test cases are $\{0 = \sum_0^0 x, 1 = \sum_0^1 1, 3 = \sum_0^2 x, 6 = \sum_0^3 x, 10 = \sum_0^4 x\}$ so the domains of variables $x,y$ is ${\mathcal D}_x = {\mathcal D}_y = [0,10]$ ($x=N$ and $y=N$ are Boolean variables whose value depends on the current value of $x$ and $y$).
 
@@
           (A) edge [align=center, bend left] node {$x=N$/\\ - } (D)          
-          (A) edge [align=center] node {$x\neq N$ /\\y = y + x} (B)
+          (A) edge [align=right] node {$x\neq N$ /\\y = y + x} (B)
           (B) edge [align=center] node { - /\\ $x = x + 1$} (C)
-          (C) edge [align=center] node {$x=N$/\\ - } (D)
+          (C) edge [align=left] node {$x=N$/\\ - } (D)
           (C) edge [align=center, bend left] node {$x\neq N$/\\ - } (A)
@@
 \caption{\small Three-line program to solve the programming task of computing $y = \sum_0^N x$.}
-\label{fig:list}
+\label{fig:sum}
 \end{figure}
@@
 Interestingly many activities in these domains can be understood as simple programs.
-For instance, Figure~\ref{fig:activity} shows a four-line program (pictured as a finite state machine) that represents the sequence of instructions required for the activity of {\em preparing an orange juice} .
+For instance, Figure~\ref{fig:activity} shows a five-line program (pictured as a finite state machine) that represents the sequence of instructions required for the activity of {\em preparing an orange juice}.
 We plan to evaluate our approach in synthesis and validation tasks coming from these real-world domains.
@@
 \end{figure}
-
 \end{itemize}
 
-\subsection{Specific Objectives}
+\subsection{Specific objectives}
 \label{sec:objectivos}
-
-The programs objects of study in this project are characterized by these tree dimensions:
+We characterize the programs objects of study in this project by these tree dimensions:
 \begin{enumerate}
@@
 
-For instance the program of Figure~\ref{fig:list}, for traversing linked lists, is generated using three program lines, one observable variable ($\mathsf{isEnd}(n)$ that has Boolean domain and indicates when $n$ points to the end of the given linked list), and an instruction set that comprises two instructions ($\mathsf{visit}(n)$, that marks the list node assigned to $n$ as {\em visited} and $\mathsf{step}(n)$, that advances $n$ to the next node in the linked list).
+To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ is generated using three program lines, two observable Boolean variables ($x=N$ and $y=N$), and an instruction set that comprises four instructions.
+Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em preparing an orange juice}, is generated using five program lines, the observable Boolean variable $ready(juice)$ that holds when the orange juice is recognized to be ready, and an instruction set that comprises four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$).
 
-The specific objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of TDD programming tasks coming from the domains described in the previous subsection.
-In more detail, the two specific objectives of this project are:
+The specific objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of TDD programming tasks coming from the two kinds of domains described in the previous subsection.
+In further detail, the two specific objectives of this project are:
 \begin{enumerate}
@@
 
-\section{Workplan}
+\section{Workplan and Budget}
 \label{sec:workplan}

### 2018-09-11 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updated evaluation project proposal

@@
 
-  Our current work on this research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and recently, it is been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for research in AI.
+  Our current work on this research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and it is been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for research in AI.
 
@@
 \label{sec:objectivos}
-We characterize the programs objects of study in this project by these tree dimensions:
-\begin{enumerate}
+We characterize the programs objects of study in this project by these three dimensions:
+\begin{itemize}
 \item Number of {\em program lines}.
@@
 \item Size of the available {\em instruction set}.
-\end{enumerate}
+\end{itemize}
 
-To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ is generated using three program lines, two observable Boolean variables ($x=N$ and $y=N$), and an instruction set that comprises four instructions.
-Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em preparing an orange juice}, is generated using five program lines, the observable Boolean variable $ready(juice)$ that holds when the orange juice is recognized to be ready, and an instruction set that comprises four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$).
+To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ is generated using: three program lines, two observable Boolean variables ($x=N$ and $y=N$), and an instruction set that comprises four instructions.
+Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em preparing an orange juice}, is generated using: five program lines, the observable Boolean variable $ready(juice)$ that holds when the orange juice is recognized to be ready, and an instruction set that comprises four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$).
 
-The specific objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of TDD programming tasks coming from the two kinds of domains described in the previous subsection.
-In further detail, the two specific objectives of this project are:
+The objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of TDD programming tasks that come from the two kinds of domains described in the previous subsection.
+In further detail, the specific objective can be itemized into two specific objectives:
 \begin{enumerate}
@@
 
-Despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}.
+Note that despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}.
 With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
@@
 
-We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that (1), takes as input a set of test cases that specify the TDD programming task to solve and (2), outputs a program source code that passes these tests with a bug-free guarantee.
+We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that:
+\begin{enumerate}
+\item Takes as input a set of test cases that specify the TDD programming task to solve and outputs a program source code that passes these tests with a bug-free guarantee.
+\item The {\em program synthesizer} must also work in {\em validation a mode}.
+In this particular setting, the {\em program synthesizer} receives an additional input specifying the program source code and it outputs a {\em validation certificate} guaranteeing that the input program passes (or fails) at the given test cases.
+\end{enumerate}
 
-In the particular case that the {\em program synthesizer} receives an additional input specifying the program source code, it outputs a validation certificate guaranteeing that the input program passes the given test cases.
-Figure~\ref{fig:gantt} details the proposed 24-month timeline for the project.
-A deliverable is provided at the end of each task.
+Figure~\ref{fig:gantt} details the proposed 24-month timeline for the project. A deliverable is provided at the end of each task.
 
@@
 
-\subsection{Workload}
-
-{\em Tasks ({\bf T1-T3})} will be developed by the three mentioned members of the research group: Sergio Jiménez, Eva Onaindia and Diego Aineto.
-
-In addition, we plan to hire a master student for 6 months to assist in the completion of two tasks, {\bf T3(a,i)} and {\bf T3(a,iv)}. The detailed responsability of the student will be:
-\begin{enumerate}
-\item Executing the scripts of the system architecture.
-\item Collecting the output data.
-\item Presenting the output data in a easy-readable format.
-\end{enumerate}
-
 \subsection{Budget}

### 2018-09-11 src/trace_generator.py
Added trace_generator.py for generating traces from a problem instance

@@ -1,0 +1,96 @@
+#! /usr/bin/env python
+import sys, os, copy, glob, itertools, random
+import pddl, pddl_parser
+import fdtask_to_pddl, planning
+import config
+
+# Madagascar details
+M_PATH = config.PLANNER_PATH
+M_CALL = "/" + config.PLANNER_NAME
+M_PARAMS = " -W " + config.PLANNER_PARAMS
+
+# FD details
+FD_PATH = "/home/slimbook/software/fd/"
+FD_CALL = "/fast-downward.py --alias seq-sat-lama-2011 "
+FD_PARAMS = ""
+
+# LPG details
+LPG_PATH = config.PROJECT_PATH + "util/lpg/"
+LPG_CALL = "lpg-td-1.0 "
+LPG_PARAMS = " -n 1 -v off -out sas_plan"
+
+# **************************************#
+# MAIN
+# **************************************#
+try:
+    domain_filename = sys.argv[1]
+    problem_filename = sys.argv[2]
+    planner = sys.argv[3]
+
+except:
+    print "Usage:"
+    print sys.argv[0] + " <domain> <problem> <planner> "
+    sys.exit(-1)
+
+# Creating a FD task with the domain and the problem file
+fd_domain = pddl_parser.pddl_file.parse_pddl_file("domain", domain_filename)
+fd_problem = pddl_parser.pddl_file.parse_pddl_file("task", problem_filename)
+fd_task = pddl_parser.pddl_file.parsing_functions.parse_task(fd_domain, fd_problem)
+
+# Running the planner
+PLANNER_OUT = "aux_planner.log"
+if planner == "FD":
+    cmd = "rm sas_plan*; ulimit -t 200;" + FD_PATH + FD_CALL + " " + domain_filename + " " + problem_filename + " " + FD_PARAMS + " > " + PLANNER_OUT
+if planner == "M":
+    cmd = "rm sas_plan*; ulimit -t 20;" + M_PATH + M_CALL + " " + domain_filename + " " + problem_filename + " " + M_PARAMS + " > " + PLANNER_OUT
+if planner == "LPG":
+    cmd = "rm sas_plan*; ulimit -t 20;" + LPG_PATH + LPG_CALL + " -o " + domain_filename + " -f " + problem_filename + " " + LPG_PARAMS + " > " + PLANNER_OUT
+print("\n\nExecuting... " + cmd)
+os.system(cmd)
+
+try:
+    # Loading the plan
[…]

### 2018-09-11 src/FAMA_FSM.py
First version of FAMA_FSM.py

@@ -1,0 +1,656 @@
+#! /usr/bin/env python
+import glob, os, sys, copy, itertools
+import pddl, pddl_parser
+import config, fdtask_to_pddl
+import numpy as np
+
+
+def get_max_vars(actions):
+    max_vars = 0
+    for a in actions:
+        max_vars = max(max_vars, a.num_external_parameters)
+    return max_vars
+
+def get_max_steps(traces):
+    traces_steps = list()
+
+    for trace in traces:
+        not_empty_states = len([state for state in trace.states if state != []])
+        not_empty_actions = len([action for action in trace.actions if action != []])
+        traces_steps.append(max(not_empty_states, not_empty_actions))
+    return sum(traces_steps), max(traces_steps)
+
+def get_all_types(task, itype):
+   output=[itype]
+   for t in task.types:
+       if t.basetype_name == itype:
+           output.append(str(t.name))
+   return output
+
+def possible_pred_for_action(task, p, a, tup):
+    if (len(p.arguments) > len(a.parameters)):
+        return False
+
+    action_types = [set([a.parameters[int(tup[i])-1].type_name]) for i in range(len(tup))]
+    predicate_types = [set(get_all_types(task, x.type_name)) for x in p.arguments]
+
+    fits = [len(action_types[i].intersection(predicate_types[i])) >= 1 for i in range(len(action_types))]
+    return all(fits)
+
+def get_static_predicates(state_trajectories, predicates):
+
+    candidates = set([p.name for p in predicates])
+
+    for trajectory in state_trajectories:
+        trace_candidates = set()
+        for predicate in candidates:
+            static = True
+            init_literals = set([l for l in trajectory[0] if l.predicate == predicate])
+            for state in trajectory[1:]:
+                state_literals = set([l for l in state if l.predicate == predicate])
+
+                if init_literals != state_literals:
+                    static = False
+                    break
+
+            if static:
+                trace_candidates.add(predicate)
+
+        candidates = candidates.intersection(trace_candidates)
+
+    # reflexive_static_predicates = dict()
+    # for candidate in candidates:
+    #     reflexive_static_predicates[candidate] = True
+    #     for trace in traces:
+    #         init_literals = set([l for l in trace.init if l.predicate == candidate])
+    #         for literal in init_literals:
[…]

### 2018-09-13 src/FAMA_FSM.py
FAMA_FSM.py: enabled validation mode for Model Recognition

@@ -126,7 +126,8 @@ except:
     sys.exit(-1)
 
 
-trace_filter = ["symbolo1", "symbolo2", "symbolo3"]
+# trace_filter = ["symbolo1", "symbolo2", "symbolo3"]
+trace_filter = ["head"]
 pres_filter = []
 adds_filter = ['states0', 'states1', 'states2', 'states3', 'states4']
 dels_filter = []
@@ -306,48 +307,29 @@ for a in actions:
                 vars = ["var" + str(t) for t in tup]
                 params = []
 
-                if not validation_mode:
-                    # Action for programming the effects
-                    pre = []
-                    pre += [pddl.conditions.Atom("modeProg", [])]
-                    pre += [pddl.conditions.NegatedAtom("add_" + "_".join([fp.name] + [a.name] + vars), []) for fp in filtered_predicates]
+                # Action for inserting a positive effect
+                pre = []
+                pre += [pddl.conditions.Atom("modeProg", [])]
+                pre += [pddl.conditions.NegatedAtom("add_" + "_".join([fp.name] + [a.name] + vars), []) for fp in filtered_predicates]
 
-                    eff = [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom(
-                        "add_" + "_".join([p.name] + [a.name] + vars), []))]
+                eff = [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom(
+                    "add_" + "_".join([p.name] + [a.name] + vars), []))]
 
-                    learning_task.actions.append(
-                        pddl.actions.Action("program_add_" + "_".join([p.name]+[a.name]+vars), params,
-                                            len(params), pddl.conditions.Conjunction(pre), eff, 0))
-                else:
-                    # Action for inserting negative effects
-                    pre = []
-                    pre += [pddl.conditions.Atom("modeProg", [])]
-                    pre += [pddl.conditions.NegatedAtom("del_" + "_".join([p.name] + [a.name] + vars), [])]
-                    pre += [pddl.conditions.NegatedAtom("add_" + "_".join([p.name] + [a.name] + vars), [])]
-                    pre += [pddl.conditions.Atom("pre_" + "_".join([p.name] + [a.name] + vars), [])]
-
-
-                    eff = []
-                    eff = eff + [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom(
-                        "del_" + "_".join([p.name] + [a.name] + vars), []))]
-
-                    learning_task.actions.append(
[…]

### 2018-09-13 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updates to the project proposal

@@
 \usepackage{tikz}
+\usepackage[hidelinks]{hyperref}
 \usetikzlibrary{arrows,automata}
@@
   This research project investigates a novel approach for software development, that leverages {\em Artificial Intelligence} (AI), to increase the automation of the programming process and hence, reduce the chances of introducing software bugs.
-The project proposes to address {\bf program synthesis and program validation starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine}.
+The project proposes to address {\bf program synthesis and program validation, starting from {\em input-output} tests cases, and using {\em AI planning} as a problem solving engine}.
 
-  Our current work on this research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and it is been accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for research in AI.
+  Our current work on this research topic is the recipient of the {\em 2016 distinguished paper award} at {\sc IJCAI} (the main international conference on AI) and it is accepted for publication at the {\em Artificial Intelligence Journal}, the premier international journal for research in AI.
 
 \end{abstract}
-{\footnotesize {\bf Keywords:} Computer Science, Artificial Intelligence, Program Synthesis, Program Validation, AI planning.}
+{\em\scriptsize {\bf Keywords:} Computer Science, Artificial Intelligence, Program Synthesis, Program Validation, AI planning.}
 
@@
 
-{\underline{\em Program Synthesis}} is the task of computing a program such that it satisfies a given formal specification of semantic correctness.
+{\underline{\em Program Synthesis}} is the task of computing a program that satisfies a given semantic and formal specification of correctness.
 
-The 2008 PhD Thesis work by Armando Solar-Lezama, at the University of California Berkeley, showed that it is possible to encode program synthesis as a {\em Boolean logic SAT} problem and therefore, use {\em satisfiability modulo theories}~\cite{barrett:SMT:2009} to automatically compute programs~\cite{lezama2008program}.
+The 2008 PhD Thesis work by Armando Solar-Lezama, at the University of California Berkeley, showed that it is possible to encode program synthesis as a {\em Boolean logic SAT} problem and therefore, use {\em Satisfiability Modulo Theories}~\cite{barrett:SMT:2009} to automatically compute programs~\cite{lezama2008program}.
 Since then, there has been a surge of practical interest in the idea of program synthesis in the formal verification community and related fields.
 
-To illustrate this interest, since 2012 the {\sc US National Science Foundation} is funding the ExCAPE research project\footnote{\tt https://www.nsf.gov/awardsearch/showAward?AWD\textunderscore ID=1138996} (an amount of $\$3,750,000$) to advance the theory and practice of program synthesis.
-In 2013 a unified framework for program synthesis was defined~\cite{alur2013syntax}.
-Since then, there is a yearly international competition ({\tt http://www.sygus.org}) to compare the different approaches for program synthesis.
-Last but not least, program synthesis has already been deployed in the real world and it is part of the {\sc Flash Fill} feature of {\sc Microsoft Excel} that automatically generates programs for string transformation~\cite{gulwani2011automating}.
+To illustrate this interest: Since 2012 the {\sc US National Science Foundation} funds the ExCAPE research project\footnote{\tt\url{https://www.nsf.gov/awardsearch/showAward?AWD_ID=1138996}} ($\$3,750,000$) to advance the theory and practice of program synthesis.
+In 2013 a standard framework for program synthesis was defined~\cite{alur2013syntax}.
+Since then, two yearly international competitions are established ({\small\tt\url{http://www.sygus.org}}), to compare the different approaches for program synthesis, and the sister competition ({\small\tt\url{http://www.syntcomp.org/}}), to compare different approaches for the synthesize of reactive programs.
+Program synthesis has already been deployed in the real world and it is part of the {\sc Flash Fill} feature of {\sc Microsoft Excel} that automatically generates programs for string transformation~\cite{gulwani2011automating}.
 
-Algorithmic approaches to program synthesis range over a wide spectrum, from {\em deductive synthesis} to {\em inductive synthesis}.
+Computational approaches to program synthesis range over a wide spectrum, from {\em deductive synthesis} to {\em inductive synthesis}.
 In deductive program synthesis, a program is synthesized by constructively proving a theorem, employing logical inference and constraint solving~\cite{manna1986deductive}.
-On the other hand, inductive synthesis seeks to find a program matching a set of input-output examples by searching in a restricted space of programs~\cite{summers1977methodology,shapiro1983algorithmic}.
-It is thus an instance of learning from examples, also termed as {\em inductive inference} or {\em Machine Learning} (ML).
-Many current approaches to synthesis blend induction and deduction~\cite{seshia2015combining}; syntax guidance is usually a key ingredient in these approaches.
+On the other hand, inductive synthesis aims to compute a program matching a set of input-output examples by searching in a restricted space of programs~\cite{summers1977methodology,shapiro1983algorithmic}.
+It is thus an instance of {\em Machine Learning} (ML).
+Most of current systems for program synthesis blend induction and deduction~\cite{seshia2015combining} and usually syntax guidance is a key ingredient in these systems.
 
@@
 {\em classical planning}~\cite{sergio:aprograming:icaps16}, {\em conformant planning}~\cite{Geffner:FSM:AAAI10}, {\em CSP}~\cite{Infantes:FSC:ECAI2010} or a {\em Prolog program}~\cite{Giacomo:FSM:ICAPS13}).
-The synthesis of programs from examples is also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
+Last but not least, the synthesis of programs from examples is also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
 ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn {\em logic programs} from examples and background knowledge, that are expressed as logic facts.
 
-{\underline{\em Program Validation}} is the task of proving (or disproving) the correctness of a given program with respect to a given formal specification of its aimed semantics.
-Program validation is considered a necessary step for program synthesis.
-{\em Model checking} is the mainstream AI approach for the formal validation of programs and controllers~\cite{clarke1999model}.
-Current approaches for model checking reduces to graph search but, instead of enumerating reachable states one at a time, the state space is traversed considering large numbers of states at a single step.
-For instance, representing set of states and transition relations as logical formulas or binary decision diagrams, like in {\em symbolic model-checking} ~\cite{mcmillan1993symbolic}.
+{\underline{\em Program Validation}} is the task of proving (or disproving) the correctness of a given program with respect to a formal specification of the aimed program semantics.
+Program validation is considered a {\em necessary step} for program synthesis and {\em model checking} is the mainstream AI approach for the formal validation of programs and controllers~\cite{clarke1999model}.
+Current approaches for model checking reduces to graph search but, instead of enumerating reachable states one at a time, they traverse the state space considering large numbers of states at a single step.
+For instance, representing sets of states and transition relations as logical formulas or {\em binary decision diagrams}, like in {\em symbolic model-checking} ~\cite{mcmillan1993symbolic}.
 
@@
 Examples are the compilations for {\em {\sc GOLOG} procedures}~\cite{baier2007exploiting}, {\em planning programs}~\cite{segovia:programs:AIJ19}, or {\em Finite State Controllers}~\cite{Geffner:FSM:AAAI10,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018}.
-Briefly, these compilations encode the {\em cross product} of a given planning instance and the automata corresponding to the program to validate.
-In this approach a {\em validation proof} is provided if a solution is found (the program is {\em correct}).
-Otherwise, if a classical planner proves that no solution exists, then the program is {\em incorrect} (its execution necessarily failed).
+Briefly these compilations encode the {\em cross product} of a given planning instance and the automata corresponding to the program to validate.
+A {\em validation proof} is provided if a solution to the compiled instance is found (i.e.
+the program is {\em correct}).
+Otherwise, if a classical planner proves that no solution exists for the compiled instance, then the program is {\em incorrect} because its execution necessarily failed.
 When actions have non-deterministic effects, program validation becomes more complex since it requires proving that all the possible program executions reach the goals.
@@
 \label{sec:methodology}
-This research project will {\bf investigate the integration of {\em AI planning} into the {\em Test Driven Development} paradigm for the automatic synthesis and validation of programs}.
-In more detail, we use {\em test cases} to specify the semantic of the aimed programs, as in Test Driven Development.
-Then, we encode the {\em program synthesis} (and {\em program validation}) tasks as AI planning problems and finally, we use off-the-shelf AI planners to compute solutions to these problems.
+This research project will investigate a computation method that {\bf integrates {\em AI planning} into the {\em Test Driven Development} paradigm for the automatic synthesis and validation of programs}.
 
@@
 \begin{small}  
-\begin{tabular}{l@{\hspace*{30pt}}|r@{\hspace*{5pt}}}
+\begin{tabular}{c@{\hspace*{10pt}}|r@{\hspace*{5pt}}}
  \textbf{Programming Task} & \textbf{Time (seconds)} \\\hline
@@
 
-\subsection{Synthesis and validation of programs as AI planning}
-Here we explain the details of our AI method for program synthesis and validation. Given a {\em TDD programming} task, our approach is modeling and solving it as it were an {\em AI planning} problem.
+\subsection{Automated synthesis and validation of programs}
+We use input-output {\em test cases} to specify the semantics of the aimed programs, as in Test Driven Development.
+Then, we encode the {\em TDD programming} (or validation) task as an AI planning problem and finally, we use an off-the-shelf AI planner to compute solutions to these problems.
+
+\subsubsection{The TDD programming task}
+A {\em TDD programming task} is a programming task where the semantics of the aimed program is specified with a set of input-output {\em test cases}.
+Formally, a {\em TDD programming} task is a tuple $\tup{{\mathcal V},{\mathcal D},{\mathcal A}, {\mathcal T}}$ where:
+\begin{itemize}
+\item ${\mathcal V}$ is the set of $n$ {\em program variables}.
+The respective domains are ${\mathcal D}=\tup{{\mathcal D}_{v_0}, \ldots, {\mathcal D}_{v_n}}$ defining the set of possible values for each variable $v\in {\mathcal V}$.
+\item ${\mathcal A}$ is the {\em instruction set}, i.e. the set of different instructions that can appear in a program.
+\item ${\mathcal T}$ is a set of input-output {\em test cases} where each test $t\in {\mathcal T}$ is a pair $t=\tup{{\mathcal I}_t,{\mathcal G}_t}$ indicating the input value for the program variables and the aimed output (i.e.
+after the program is executed) for these same variables.
+\end{itemize}
+A {\em solution} to TDD programming task is a program whose execution succeeds to solve every given test case.
 
 \subsubsection{The AI planning problem}
-An {\em AI planning problem} is defined as a tuple $\tup{V,D,A,I,G}$ where:
+An {\em AI planning problem} is formalized as a tuple $\tup{V,D,A,I,G}$ where:
 \begin{itemize}
-\item $V=\tup{v_0, \ldots, v_n}$ is the set of $n$ {\em state variables} with finite domain.
-The respective domains are $D=\tup{D_{v_0}, \ldots, D_{v_n}}$ defining, for each variable $v\in V$, its set of possible values.
+\item $V=\tup{v_0, \ldots, v_m}$ is the set of $m$ {\em state variables} with finite domain.
+The respective domains $D=\tup{D_{v_0}, \ldots, D_{v_m}}$ define, for each variable $v\in V$, its set of possible values.
 \item $A$ is a set of {\em actions} whose dynamics are specified with two functions:
@@
 \item $App(s)\subseteq A$, defines the subset of actions applicable in a state $s$.
-\item $\theta(s,a)=s'$, that defines the {\em successor state} $s'$ that results of applying an action $a$ in a state $s$.
+\item $\theta(s,a)=s'$, defines the {\em successor state} $s'$ that results of applying an action $a$ in a state $s$.
 \end{itemize}  
 \item $I$ is an {\em initial state}, i.e. a full assignment of values to the state variables.
-\item $G$ is the set of {\em goal conditions} bounding the possible values of the state variables in the goal states.  
+\item $G$ is the set of {\em goal conditions} constraining the possible values of the state variables in the goal states.  
 \end{itemize}
@@
 
-\subsubsection{The TDD programming task}
-We formalize a {\em TDD programming task} as a programming task whose semantic correctness is specified with a set of input-output {\em test cases}.
-A {\em TDD programming} task is a tuple $\tup{{\mathcal V},{\mathcal D},{\mathcal A}, {\mathcal T}}$ where:
+\subsubsection{Program synthesis and validation as AI planning}
+Here we introduce how to encode a {\em TDD programming task} as an {\em AI planning problem}.
+For further details on the encoding we refer the reader to~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia2017generating,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
+
+Breifly, our encoding defines an {\em AI planning problem} with {\bf state variables} of three kinds:
 \begin{itemize}
-\item ${\mathcal V}$ is the set of $m$ {\em program variables}.
-The respective domains are ${\mathcal D}=\tup{{\mathcal D}_{v_0}, \ldots, {\mathcal D}_{v_m}}$ defining the set of possible values for each variable $v\in {\mathcal V}$.
-\item ${\mathcal A}$ is the {\em instruction set}, i.e. the set of different instructions that can appear in a program.
-\item ${\mathcal T}$ is a set of input-output {\em test cases} where each test $t\in {\mathcal T}$ is a pair $t=\tup{{\mathcal I}_t,{\mathcal G}_t}$ that represent the input value for the program variables and the aimed output for these variables.
-\end{itemize}
-A {\em solution} to TDD programming task is a program whose execution succeeds to solve every given test case.
-
-\subsubsection{Program synthesis and validation as AI planning}
-Here we describe how to encode a {\em TDD programming task} task as an AI planning problem. Our encoding defines {\bf state variables} of three kinds:
-\begin{itemize}
-\item {\tt program(line) := instruction}, that encode the instructions (from the instruction set ${\mathcal A}$) at the different program lines. Initialy all program lines are empty.
+\item {\tt program(line) := instruction}, that encode the instructions (from the instruction set ${\mathcal A}$) at the different program lines. Initially all program lines are empty.
 \item {\tt pcounter := line}, encoding which is the current program line. Initially the program counter points to the first program line.
@@
 
-The {\bf initial/goal states} of the AIP task encode the initial/final values of the program variables.
-These values are specified by the test cases ${\mathcal T}$ of the TDD programming task.
-Finally, every program instruction ($w\in {\mathcal A}$) is encoded with two AIP {\bf actions}:
+The {\bf initial/goal states} of the AI planning problem encode the input/output values of the program variables (as specified by the test cases ${\mathcal T}$ of the TDD programming task).
+Finally, every program instruction ($w\in {\mathcal A}$) is encoded into the {\em AI planning problem} with two {\bf actions}:
 \begin{itemize}
@@
 We implement this encoding using standard planning languages, such as PDDL~\cite{fox2003pddl2}, so the AIP tasks resulting from our encoding can be solved with off-the-shelf planners, like the {\sc FD} planning system~\cite{helmert2006fast}.
-Our encoding is {\em complete} and {\em correct}~\cite{segovia:programs:AIJ19}, which means that the programs synthesized with this approach are guaranteed to be bug-free over given sets of {\em input-output} tests cases.
+Our encoding is {\em complete} and {\em correct}~\cite{segovia:programs:AIJ19}, which means that the programs synthesized with this approach are guaranteed to be bug-free over the given set of {\em input-output} tests cases.
 
 Interestingly, our PDDL encoding allows also program validation by (1), specifying the lines of the program to validate (i.e.
-the {\tt program(line):=instruction} fluents) in the initial state of the AIP task and (2), disabling the mentioned {\it programming actions} so only {\it execution actions} are applicable.
+the {\tt program(line):=instruction} fluents) in the initial state of the AI planning problem and (2), disabling the mentioned {\it programming actions} so only {\it execution actions} are applicable.
 
@@
 
-The performance of our AI method for the synthesis and validation of TDD programming tasks will be evaluated with regard to (1) {\bf computation time} and (2), {\bf memory} invested in the synthesize and the validation of the aimed programs.
-
-The programming tasks used for the evaluation are of two different kinds:
+The performance of our AI method for the synthesis and validation of programs will be evaluated with regard to (1) {\bf computation time} and (2), {\bf memory}.
+Further, the programming and validation tasks used for the evaluation of the performance of our approach are of two different kinds:
 \begin{itemize}
@@
 Figure~\ref{fig:sum} shows a synthesized program for computing $y = \sum_0^N x$.
-The program is pictured as a {\em finite state machine}: The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
+The program is pictured as a {\em finite state machine}: Machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
 The program in Figure~\ref{fig:sum} assumes that the program variables are ${\mathcal V}=\{x,y,x=N,y=N\}$ and that the value of $x$ and $y$ is initially 0.
-The instruction set is ${\mathcal A}=\{x = x + 1, y = y + 1, x = x + y, y = y + x\}$.
-The input test cases are $\{0 = \sum_0^0 x, 1 = \sum_0^1 1, 3 = \sum_0^2 x, 6 = \sum_0^3 x, 10 = \sum_0^4 x\}$ so the domains of variables $x,y$ is ${\mathcal D}_x = {\mathcal D}_y = [0,10]$ ($x=N$ and $y=N$ are Boolean variables whose value depends on the current value of $x$ and $y$).
+The instruction set is ${\mathcal A}=\{x:=x+1, y:=y+1, x:=x+y, y:=y+x\}$.
+For the synthesis of this program we are using these five test cases $\{0 = \sum_0^0 x, 1 = \sum_0^1 x, 3 = \sum_0^2 x, 6 = \sum_0^3 x, 10 = \sum_0^4 x\}$ so the domains of variables $x$ and $y$ are ${\mathcal D}_x = {\mathcal D}_y = [0,10]$.
+Note that $x=N$ and $y=N$ are Boolean variables whose value depends on the current value of $x$ and $y$.
 
@@
   \begin{center}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=4cm,semithick]
+  \begin{scriptsize}    
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3cm,semithick]
 	  \node[state] (A)              {$Q_0$};
@@
           (B) edge [align=center] node { - /\\ $x = x + 1$} (C)
-          (C) edge [align=left] node {$x=N$/\\ - } (D)
+          (C) edge [align=center] node {$x=N$/\\ - } (D)
           (C) edge [align=center, bend left] node {$x\neq N$/\\ - } (A)
@@
 	\end{tikzpicture}
+  \end{scriptsize}            
 \end{center}
@@
 
-\item {\em Real-world benchmarks}: The GRPS group is currently leading the four-year research project TIN2017-88476-C2-1-R from the {\em Spanish national plan} in which {\em AI Planning} and {\em activity recognition} is applied to different real-world domains such as {\em domotics}, {\em tourism}, {\em traffic control} and {\em robotics}.
-Interestingly many activities in these domains can be understood as simple programs.
-For instance, Figure~\ref{fig:activity} shows a five-line program (pictured as a finite state machine) that represents the sequence of instructions required for the activity of {\em preparing an orange juice}.
+\item {\em Real-world benchmarks}: We are involved in the four-year research project TIN2017-88476-C2-1-R from the {\em Spanish national plan} in which {\em AI Planning} and {\em activity recognition} is applied to different real-world domains such as {\em domotics}, {\em tourism}, {\em traffic control} and {\em robotics}.
+Interestingly many activities from these domains can be understood as simple programs.
+For instance, Figure~\ref{fig:activity} shows a five-line program (pictured as a finite state machine) that represents the sequence of instructions required for {\em making an orange juice}.
 We plan to evaluate our approach in synthesis and validation tasks coming from these real-world domains.
@@
   \begin{center}
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=3.3cm,semithick]
+      \begin{scriptsize}   
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.8cm,semithick]
 	  \node[state] (A)              {$Q_0$};
@@
           (E) edge [align=center, bend left] node { $\neg ready(juice)$ /\\ - } (A);                    
-
     \end{tikzpicture}
+     \end{scriptsize}            
 \end{center}
-\caption{\small Five-line program representing the activity of {\em preparing an orange juice}.}
+\caption{\small Five-line program representing the {\em making an orange juice} activity from the {\em domotics} domain.}
 \label{fig:activity}
@@
 \item Number of {\em program lines}.
-\item Number and domain of the observable {\em program variables}.
 \item Size of the available {\em instruction set}.
+\item Number and domain of {\em observable} program variables. That is, the subset of program variables whose value can condition the program flow.
 \end{itemize}
 
-To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ is generated using: three program lines, two observable Boolean variables ($x=N$ and $y=N$), and an instruction set that comprises four instructions.
-Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em preparing an orange juice}, is generated using: five program lines, the observable Boolean variable $ready(juice)$ that holds when the orange juice is recognized to be ready, and an instruction set that comprises four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$).
+To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ have three program lines, an instruction set that comprises four instructions and a single observable Boolean variable ($x=N$).
+Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em preparing an orange juice} have five program lines, an instruction set with four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$) and a single observable Boolean variable, ${\tt ready(juice)}$, that holds when the orange juice is recognized to be ready.
 
-The objective of this project is to study the performance of our approach (in terms of computation time and memory) for the automated synthesis and validation of TDD programming tasks that come from the two kinds of domains described in the previous subsection.
-In further detail, the specific objective can be itemized into two specific objectives:
+The objective of this project is to study the performance of our approach (in terms of  {\bf computation time} and {\bf memory}) in the synthesis and validation of {\em TDD programming tasks} that come from the two kinds of domains described in the previous subsection.
+This objective can be itemized into two specific objectives:
 \begin{enumerate}
@@
 Note that despite setting bounds for the programs size and kind, challenging programming tasks can be addressed using {\em problem decomposition}.
-With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
+With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16,jimenez2018review,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
 
@@
 The main benefit of this project is to provide new insights into the current understanding of how AI can assist programmers in the software development.
+Automated program synnthesis can produce huge cost savings in comparison to conventional programmming and increase the quality, trustbility and modifiability of programs.
+A longer term objective is the deeper and deeper embedding of intelligence in all manner of software systems, as well as the design and management of organizational processes by automated means.
+
 In more detail, the expected benefits for this particular research project are four-fold:

### 2018-09-13 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
first full version project proposal

@@
 
-The performance of our AI method for the synthesis and validation of programs will be evaluated with regard to (1) {\bf computation time} and (2), {\bf memory}.
-Further, the programming and validation tasks used for the evaluation of the performance of our approach are of two different kinds:
+The performance of our AI method for the synthesis and validation of programs will be evaluated with regard to (1), {\bf computation time} and (2), {\bf memory}.
+Further, the evaluation will be carried out in programming and validation tasks of two different kinds:
 \begin{itemize}
-\item {\em Theoretical benchmarks}: Classic programming tasks are a neat touchstone to assess the performance of our approach.
+\item {\bf\em Theoretical benchmarks}: Classic programming tasks are a neat touchstone to assess the performance of our approach.
 For instance, programs for the computation of mathematical/logic series, string manipulation and for the management of data structures such as {\em lists}, {\em queues}, {\em stacks} or {\em trees}.
-Figure~\ref{fig:sum} shows a synthesized program for computing $y = \sum_0^N x$.
-The program is pictured as a {\em finite state machine}: Machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
+As an example Figure~\ref{fig:sum} shows a synthesized program, pictured as a {\em finite state machine}, for computing $y = \sum_0^N x$.
+The machine nodes mount to the different program lines while edges are tagged with a {\em condition/instruction} label, that denotes the condition (over the program variables) under which program instructions are taken.
 The program in Figure~\ref{fig:sum} assumes that the program variables are ${\mathcal V}=\{x,y,x=N,y=N\}$ and that the value of $x$ and $y$ is initially 0.
 The instruction set is ${\mathcal A}=\{x:=x+1, y:=y+1, x:=x+y, y:=y+x\}$.
-For the synthesis of this program we are using these five test cases $\{0 = \sum_0^0 x, 1 = \sum_0^1 x, 3 = \sum_0^2 x, 6 = \sum_0^3 x, 10 = \sum_0^4 x\}$ so the domains of variables $x$ and $y$ are ${\mathcal D}_x = {\mathcal D}_y = [0,10]$.
+For the synthesis of this program we are using the following five test cases $\{0 = \sum_0^0 x, 1 = \sum_0^1 x, 3 = \sum_0^2 x, 6 = \sum_0^3 x, 10 = \sum_0^4 x\}$ so the domains of variables $x$ and $y$ are ${\mathcal D}_x = {\mathcal D}_y = [0,10]$.
 Note that $x=N$ and $y=N$ are Boolean variables whose value depends on the current value of $x$ and $y$.
@@
 
-\item {\em Real-world benchmarks}: We are involved in the four-year research project TIN2017-88476-C2-1-R from the {\em Spanish national plan} in which {\em AI Planning} and {\em activity recognition} is applied to different real-world domains such as {\em domotics}, {\em tourism}, {\em traffic control} and {\em robotics}.
-Interestingly many activities from these domains can be understood as simple programs.
-For instance, Figure~\ref{fig:activity} shows a five-line program (pictured as a finite state machine) that represents the sequence of instructions required for {\em making an orange juice}.
+\item {\bf\em Real-world benchmarks}: We are involved in the four-year research project {\sc ARPIA} (TIN2017-88476-C2-1-R, {\tt\small\url{http://arpia.blogs.upv.es/}}) fund by the {\em Spanish national plan} in which {\em AI Planning} and {\em activity recognition} is applied to different real-world domains such as {\em domotics}, {\em tourism}, {\em traffic control} and {\em robotics}.
+Interestingly many activities that are objects of study in these domains can be modeled as simple programs.
+Figure~\ref{fig:activity} shows a five-line program (pictured as a finite state machine) that represents the sequence of instructions required for {\em making an orange juice} from the {\em domotics} domain.
 We plan to evaluate our approach in synthesis and validation tasks coming from these real-world domains.
@@
       \begin{scriptsize}   
-    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.8cm,semithick]
+    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2.7cm,semithick]
 	  \node[state] (A)              {$Q_0$};
@@
           (E) edge [align=center] node { ready(juice) /\\ - } (F)
-          (E) edge [align=center, bend left] node { $\neg ready(juice)$ /\\ - } (A);                    
+          (E) edge [align=center, bend left] node {$\neg ready(juice)$ /\\ $\mathsf{take}(orange)$} (B);                    
     \end{tikzpicture}
@@
 \label{sec:objectivos}
-We characterize the programs objects of study in this project by these three dimensions:
+We characterize the programs that are objects of study in this project by these three dimensions:
 \begin{itemize}
-\item Number of {\em program lines}.
-\item Size of the available {\em instruction set}.
-\item Number and domain of {\em observable} program variables. That is, the subset of program variables whose value can condition the program flow.
+\item Number of {\bf\em program lines}.
+\item Size of the available {\bf\em instruction set}.
+\item Number and domain of {\bf\em observable program variables}. That is, the subset of program variables whose value can condition the program flow.
 \end{itemize}
@@
 To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ have three program lines, an instruction set that comprises four instructions and a single observable Boolean variable ($x=N$).
-Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em preparing an orange juice} have five program lines, an instruction set with four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$) and a single observable Boolean variable, ${\tt ready(juice)}$, that holds when the orange juice is recognized to be ready.
+Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em making an orange juice} have five program lines, an instruction set with four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$) and the single observable Boolean variable, ${\tt ready(juice)}$, that holds when the orange juice is recognized to be ready.
 
-The objective of this project is to study the performance of our approach (in terms of  {\bf computation time} and {\bf memory}) in the synthesis and validation of {\em TDD programming tasks} that come from the two kinds of domains described in the previous subsection.
-This objective can be itemized into two specific objectives:
+The objective of this project is to study the performance of our approach (in terms of  {\em computation time} and {\em memory}) in the synthesis and validation of {\em TDD programming tasks} that come from domains of the two kinds described in the previous subsection.
+This general objective is itemized into two specific objectives:
 \begin{enumerate}
-\item To study the performance of our Artificial Intelligence approach for the {\bf synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
-\item To study the performance of our Artificial Intelligence approach for the {\bf validation of programs} up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}.
+\item Analyze the performance of our Artificial Intelligence approach for the {\bf synthesis} of programs up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
+\item Analyze the performance of our Artificial Intelligence approach for the {\bf validation} of programs up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}.
 \end{enumerate}
@@
 
-We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that:
+We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that works in two different modes:
 \begin{enumerate}
-\item Takes as input a set of test cases that specify the TDD programming task to solve and outputs a program source code that passes these tests with a bug-free guarantee.
-\item The {\em program synthesizer} must also work in {\em validation a mode}.
-In this particular setting, the {\em program synthesizer} receives an additional input specifying the program source code and it outputs a {\em validation certificate} guaranteeing that the input program passes (or fails) at the given test cases.
+\item {\em Synthesis mode}. Takes as input a set of test cases that specify the TDD programming task to solve and outputs a program source code that passes these tests with a bug-free guarantee.
+\item {\em Validation mode}.
+In this setting the {\em program synthesizer} receives an additional input, specifying the source code of the program to validate, and it outputs a {\em validation certificate} guaranteeing that the input program either succeeds (or fails) to solve each test case from the TDD programming task.
 \end{enumerate}
 
-Figure~\ref{fig:gantt} details the proposed 24-month timeline for the project. A deliverable is provided at the end of each task.
+Figure~\ref{fig:gantt} details the proposed 24-month timeline for the development of the project. A deliverable is provided at the end of each task.
 
@@
       \item The programming-into-planning compiler ({\em Compiler 1}).
-This system component parses the {\em TDD programming task} and produces an {\em AIP task} encoded in the standard planning language PDDL.
+This system component parses the {\em TDD programming task} and produces an {\em AI Planning problem} encoded in the standard planning language PDDL.
       \item The plan-into-program compiler ({\em Compiler 2}).
@@
       \begin{enumerate}
-      \item Reporting the experimental performance of our AIP approach for solving diverse TDD programming tasks. This task will follow an iterative workflow over the following subtasks:
+      \item Reporting the experimental performance of our AI approach for solving diverse TDD programming tasks. This task will follow an iterative workflow over the following subtasks:
       \begin{enumerate}
@@
 \end{small}        
-{\small{\bf\em  Deliverable T3:} Final report with the obtained conclusions and produced publications.}
+{\small{\bf\em  Deliverable T3:} Final report with obtained conclusions and produced publications.}
 \end{enumerate}
 
-\subsection{Budget}
-Here we show the detailed desctiption of the two-year budget for the development of the proyect.
+\subsection{Presupuesto} 
+Esta sección presenta la descripción detallada del presupuesto, de dos anualidades, para el desarrollo del proyecto:
 
@@
 \begin{small}  
-  \begin{tabular}{clc|l}
-    {\bf Priority} & {\bf Description} & {\bf Term} & {\bf Euros} \\\hline
-1 & {\scriptsize Difusión de las actividades del grupo} & 2018, 2019 & 2,000\\    
-2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2018 & 1,500\\
-3 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2019 & 1,500\\
-4 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2018 & 1,500\\
-5 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2019 & 1,500\\\hline
-\multicolumn{3}{l|}{} & 8,000 \\
+  \begin{tabular}{clc|r}
+    {\large\bf Prioridad} & {\large\bf Descripcion} & {\large\bf Año} & {\large\bf Euros} \\\hline
+1 & {\scriptsize Difusión de las actividades del grupo} & 1 & 4,000\\    
+2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 1 & 2,000\\
+3 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 1 & 2,000\\\hline
+1 & {\scriptsize Difusión de las actividades del grupo} & 2 & 4,000\\    
+2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2 & 2,000\\
+3 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2 & 2,000\\\hline
+\multicolumn{3}{l|}{} & 16,000 \\
   \end{tabular}
 \end{small}
-\caption{\small Two-year budget for the development of the proyect.}
+\caption{\small Presupuesto de dos anualidades para el desarrollo del proyecto.}
 \end{table}
 
-\section{Disemination and exploitation of the research results.}
+\begin{enumerate}
+\item {\em Difusión de las actividades del grupo}.
+Esta partida contempla las inscripciones a congresos internacionales en Inteligencia Artificial, previsiblemente los tres siguientes: The International Conference on Automated Planning and Scheduling (h5-index 27).
+The International Joint Conference on Artificial Intelligence (h5-index 61).
+The AAAI Conference on Artificial Intelligence (h5-index 69).
+The International Conference on Machine Learning (h5-index 113).
+Todos los h5-index calculados de acuerdo a Google Scholar.
+\item {\em Viajes, manutención y alojamiento grupo de investigación}.
+Esta partida contempla los gastos generados por la asistencia a los congresos internacionales mencionados anteriormente y la visita a grupos de investigacion relacionados con los objetivos del proyecto.
+\item {\em Viajes, manutención, alojamiento y ponencias investigadores invitados}.
+Sergio Jimenez, PI del presente proyecto, mantiene un activa colaboración con el grupo {\em A.I.
+and Autonomy Lab} de la universidad de Melbourne {\tt\small\url{https://cis.unimelb.edu.au/agentlab}} asi como con el {\em Artificial Intelligence and Machine Learning Research group} ({\tt\small\url{https://www.upf.edu/web/ai-ml}}) de la Universitat Pompeu Fabra de Barcelona.
+Esta partida contempla los gastos generados para el desarrollo de actividades conjuntas en el ámbito de la Comunitat Valenciana.
+\end{enumerate}
+
+\section{Expected benefits of the research.}
 \label{subsec:beneficios}
-AIP has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}.
-Promising research opportunities come from the application of AIP to {\em program synthesis} given that, {\em program synthesis} with a tests base, can be seen as the {\em program testing} dual.
+{\em AI Planning} has recently shown successful in {\em program testing} to generate {\em attack plans} that completed non-trivial software security tests~\cite{hoffmann2015simulated,steinmetz2016revisiting,shmaryahu2016constructing,steinmetz2016goal}.
+Promising research opportunities come from the application of {\em AI Planning} to {\em program synthesis} given that, {\em program synthesis} with test cases, can be seen as the {\em program testing} dual.
+In fact, our current work on {\em program synthesis} with AIP already produced several {\bf publications at the top international conferences and journals on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16,segovia:FSC:JAIR2018,segovia:programs:AIJ19} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
 
-In fact, our current work on {\em program synthesis} with AIP already produced several {\bf publications at top international conferences and journals on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16,segovia:FSC:JAIR2018,segovia:programs:AIJ19} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
-
-The main benefit of this project is to provide new insights into the current understanding of how AI can assist programmers in the software development.
-Automated program synnthesis can produce huge cost savings in comparison to conventional programmming and increase the quality, trustbility and modifiability of programs.
-A longer term objective is the deeper and deeper embedding of intelligence in all manner of software systems, as well as the design and management of organizational processes by automated means.
-
+The main benefit of this project is to {\bf provide new insights into the current understanding of how AI can assist programmers in the software development}.
+Automated program synthesis promises to produce huge cost savings in comparison to conventional software development and to increase the quality, reliability and modifiability of programs.
 In more detail, the expected benefits for this particular research project are four-fold:
@@
 \item A new {\bf evaluation methodology for the {\em Synthesis and Validation of Programs}}.
-The application of the exiting AIP technology to program synthesis can provide new evaluation metrics that assess how well a program covers a set of {\em input-output} test cases.
+The application of the exiting {\em AI Planning} technology to program synthesis can provide new evaluation metrics that assess how well a program covers a set of {\em input-output} test cases.
 \item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}.

### 2018-09-14 doc/icaps-19/model-recognition.tex
updates after first meeting after Miquels comments

@@
 missing intermediate states and actions) given as input besides, it is computable with an off-the-shelf classical planner.
-The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different {\em Turing Machines}.
-We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, internal machine state or the values of several tape cells, are unknown.
+The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different {\em regular automatae} and different {\em Turing machines}.
+We show that {\em model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unknown.
 \end{abstract}
+
+\begin{figure}
+  \begin{scriptsize}
+  \begin{center}
+\begin{tikzpicture}[scale=.4]
+          \begin{scope}
+            \draw (0, 0) grid (5, 5);
+            \node[anchor=center] at (3.5, 1.5) {R};
+            \node[anchor=center] at (0.5, 0.5) {I};
+            \draw[thick,style=dotted] (0.7,0.5) -- (4.5,0.5);
+            \draw[thick,style=dotted] (4.5,0.5) -- (4.5,1.5);
+            \draw[thick,style=dotted,->] (4.5,1.5) -- (3.7,1.5);                        
+          \end{scope}
+        \end{tikzpicture}
+  \end{center}
+  \end{scriptsize}  
+ \caption{\small Plan execution of a robot navigating a $5\times 5$ grid.}
+\label{fig:grid-example}
+\end{figure}
+
+\begin{figure}
+  \begin{tiny}
+  \begin{verbatim}       
+  (:action inc-x
+    :parameters (?px1 ?px2)
+    :precondition (and (x-pos ?px1) (next ?px1 ?px2) (even))
+    :effect (and (not (x-pos ?px1)) (x-pos ?px2)))
+
+  (:action dec-x
+    :parameters (?px1 ?px2)
+    :precondition (and (x-pos ?px1) (next ?px2 ?px1) (odd))
+    :effect (and (not (x-pos ?px1)) (x-pos ?px2)))
+
+  (:action inc-y-min-odd
+    :parameters (?py1 ?py2)
+    :precondition (and (y-pos ?py1) (not (next ?py2 ?py1)) (odd))
+    :effect (and (not (y-pos ?py1)) (y-pos ?py2) (not (odd)) (even)))
+
+  (:action inc-y-min-even
+    :parameters (?py1 ?py2)
+    :precondition (and (y-pos ?py1) (not (next ?py2 ?py1)) (even))
+    :effect (and (not (y-pos ?py1)) (y-pos ?py2) (not (even)) (odd)))
+
+  (:action inc-y-max-odd
+    :parameters (?py1 ?py2)
+    :precondition (and (y-pos ?py1) (not (next ?py1 ?py2)) (odd))
+    :effect (and (not (y-pos ?py1)) (y-pos ?py2) (not (odd)) (even)))
+
+  (:action inc-y-max-even
+    :parameters (?py1 ?py2)
+    :precondition (and (y-pos ?py1) (not (next ?py1 ?py2)) (even))
+    :effect (and (not (y-pos ?py1)) (y-pos ?py2) (not (even)) (odd)))
+  \end{verbatim}           
+  \end{tiny}  
+ \caption{\small Action model for a robot navigation in a $5\times 5$ grid.}
+\label{fig:model-example}
+\end{figure}
 
@@
 Given a partially observed plan execution and a set of possible planning models, {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation.
-{\em Model recognition} is of interest because:
-\begin{itemize}
-\item Once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
-\item It enables indentifying algorithms by observing their execution.
-Diverse algorithm representations, like {\em finite state controllers}, {\em push-down automata} or {\em {\sc GOLOG} programs}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}.
-\end{itemize}
+To better illustrate {\em model recognition}, imagine a robot in a grid whose navigation is determined by the \strips\ model shown in Figure~\ref{fig:model-example}.
+According to this navigation model the robot increments its {\em x-coordinate} when it is at an {\em even} row while, at {\em odd rows}, the {\em x-coordinate} is decremented.
+Further, the robot only increments the {\em y-coordinate} either at the end or at the begining of rows.
+Apart from this particular navigation model, different models can be defined within the same state variables and these models can determine different kinds of robot navigation.
+Given an observation of a plan execution (e.g, the six-step plan illustrated by Figure~\ref{fig:grid-example}) the {\em model recognition} task aims here to identify which model is more likely of producing the given observation.
+
+{\em Model recognition} is of interest because once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
+In addition, it enables indentifying different kinds of automatae by observing their execution (it is well-known that diverse automatae representations, like {\em finite state controllers}, {\em push-down automata} or {\em {\sc GOLOG} programs}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}).
 
@@
 The method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
-
-The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em Turing Machines}.
-All of these {\em Turing Machines} are defined within the same {\em tape alphabet} and same {\em machine states} but different {\em transition functions}.
-For instance, Figure~\ref{fig:automatae} illustrate a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
-We show that {\em model recognition as planning} succeeds to identify the executed {\em Turing Machine} despite the actual applied transitions, the internal machine state or the values of several tape cells, are unknown.
-
-\begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-	  \node[state] (A)              {$Q_0$};
-	  \node[state] (B) [right of=A] {$Q_1$};
-	  \node[state] (C) [right of=B] {$Q_2$};
-	  \node[state] (D) [right of=C] {$Q_3$};
-	  \node[state] (E) [below of=A] {$Q_4$};
-	  \node[state] (F) [right of=E] {$Q_5$};
-
-	  \path
-(A) edge  node {a/x,right} (B)
-(B) edge [align=left, loop above] node {a/a,right\\y/y,right} (B)
-(B) edge  node {b/y,right} (C)
-(C) edge [align=left, loop above] node {z/z,right\\b/b,right} (C)
-(C) edge  node {c/z,left} (D)
-(D) edge [align=left, loop right] node {a/a,left\\b/b,left\\y/y,left\\z/z,left} (D)
-(D) edge [bend left=45] node {x/x,right} (A)
-
-(A) edge  node {y/y,right} (E)
-(E) edge [align=left, loop left] node {y/y,right\\z/z,right} (E)
-(E) edge  node {b/b,right} (F)
-;
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}  
- \caption{\small Seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.}
-\label{fig:automatae}
-\end{figure}
+The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em regular automatae} and different {\em Turing Machines}.
+All of these {\em automatae} are defined within the same {\em alphabet} and same {\em machine states} but different {\em transition functions}.
+We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite its internal machine state or actual applied transitions are unknown.
 
@@
 We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
-A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e.
+A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
 either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values).
-We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.
+A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent).
+We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
 
-A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Explicitly including negative literals $\neg f$ in states simplifies subsequent definitions but often we will abuse of notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.
+A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states; i.e.
+$|s|=|F|$, so the size of the state space is $2^{|F|}$.
+Like in PDDL~\cite{fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
+Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
+Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
 
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state, $G\subseteq\mathcal{L}(F)$ is a goal condition and $A$ is a set of actions whose dynamics are specified with two functions: $App(s)$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
+A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose dynamics are specified with two functions: $App(s)$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
+In this work we are specifying action dynamics with the \strips\ model.
+With this regard, an action $a\in A$ is defined with:
+\begin{itemize}
+\item $\pre(a)\in\mathcal{L}(F)$, the {\em preconditions} of $a$, is the set of literals that must hold for the action $a\in A$ to be applicable.
+\item $\eff^+(a)\in\mathcal{L}(F)$, the {\em positive effects} of $a$, is the set of literals that are true after the application of the action $a\in A$.
+\item $\eff^-(a)\in\mathcal{L}(F)$, the {\em negative effects} of $a$, is the set of literals that are false after the application of the action.
+\end{itemize}
+We assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$ and that actions $a\in A$ are instantiated from given action schemas, as in PDDL.
+We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
+The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
 
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that, when executed starting from the initial state $I$, induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and forall {\small $1\leq i\leq n$}, an action $a_i$ is applicable in the corresponding state $a_i\in App(s_{i-1})$ and  generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The {\em plan length} is denoted with $|\pi|=n$.
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
+A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $s=\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+The {\em plan length} is denoted with $|\pi|=n$ .
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.,~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
+A solution plan for $P$ is {\em optimal} if it has minimum length.
 
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$; i.e.~if the goal condition is satisfied in the last state resulting from the execution of the plan $\pi$ in the initial state $I$.
-A solution plan is {\em optimal} if it has minimum length.
+\subsection{Conditional effects}
+An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+
+An action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+\[
+triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
+\]
+
+The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
 
@@
 the $App$ and $\theta$ functions) are specified with \strips\ action schemas.
-Figure~\ref{fig:stack} shows the PDDL code for the {\em stack} actions, taken from a four-operator {\em blocksworld}~\cite{slaney2001blocks}.
+Figure~\ref{fig:model-example} shows the PDDL code for the action schema of the mentioned navigation model (Figure).
 
@@
 $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-\begin{figure}
-\begin{scriptsize}
-\begin{verbatim}
-(:action stack
- :parameters (?v1 ?v2 - object)
- :precondition (and (holding ?v1) (clear ?v2))
- :effect (and (not (holding ?v1)) (not (clear ?v2))
-              (handempty) (clear ?v1) (on ?v1 ?v2)))
-\end{verbatim}
-\end{scriptsize}
- \caption{\small \strips\ action schema coded in PDDL.}
-\label{fig:stack}
-\end{figure}
 
 Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be an additional set of objects ($\Omega\cap\Omega_v=\emptyset$), that we denote as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a three-block four-operator {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1, v_2\}$ because the {\small\tt stack} and {\small\tt unstack} actions have arity two.
+For instance, in a $5\times 5$ grid $\Omega=\{p_1, p_2, p_3, p_4, p_5\}$ while $\Omega_v=\{v_1, v_2\}$ because all the actions have arity two.
 We define $F_v$, a new set of fluents, $F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}.
-This set contains eleven elements for the mentioned {\em blocksworld}, $F_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+This set contains ten elements for the mentioned navigation model, $F_v$={\small\tt\{x-pos($v_1$), x-pos($v_2$), y-pos($v_1$), y-pos($v_2$), odd(), even(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
 
 For a given action schema $\xi$, then $F_{\xi}\subseteq F_v$ defines the subset of elements that can appear in the preconditions and effects of that action schema.
-For instance $F_{\tt stack}=F_v$ while $F_{\tt pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} excludes any element in $F_v$ that involves $v_2$ because {\small\tt pickup} actions have arity one.
+For instance $F_{\tt inc-x}=F_{\tt dec-x}=F_v$ because all actions have arity two.
 
@@
 In more detail, if a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in that schema.
-For instance, Figure~\ref{fig:encodedstack} shows the conjunction of fluents representing the propositional encoding for the preconditions, negative and positive effects of the {\em stack} schema (see Figure~\ref{fig:stack}).
+For instance, Figure~\ref{fig:encoding} shows the conjunction of fluents representing the propositional encoding for the preconditions, negative and positive effects of the {\em inc-x} schema previously defined in Figure~\ref{fig:model-example}.
 
@@
 \begin{verbatim}
-(pre_holding_stack_v1) (pre_clear_stack_v2)
-(del_holding_stack_v1) (del_clear_stack_v2)
-(add_handempty_stack) (add_clear_stack_v1) 
-(add_on_stack_v1_v2)
+(pre_x-pos_inc-x_v1) (pre_next_inc-x_v1_v2) (pre_even_inc-x_)
+(del_x-pos_inc-x_v1) (add_x-pos_inc-x_v2)
 \end{verbatim}
 \end{scriptsize}
- \caption{\small Propositional encoding for the {\em stack} schema from a four-operator {\em blocksworld}.}
-\label{fig:encodedstack}
+ \caption{\small Propositional encoding for the {\tt inc-x} schema previously defined in Figure~\ref{fig:model-example}.}
+\label{fig:encoding}
 \end{figure}
@@
 The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2\times|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions).
-For the mentioned {\em blocksworld}, $2^{2\times|F_{\tt stack}|}=4,194,304$ while $2^{2\times|F_{\tt pickup}|}=1,024$.
-We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same space of possible \strips\ models.
-For instance, we claim that blocksworld schemas {\small\tt stack} and {\small\tt unstack} are {\em comparable} while  {\small\tt stack} and {\small\tt pickup} are not.
+For the mentioned navigation model, $2^{2\times|F_{\xi}|}=1,048,576$ for every action schema.
+We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same space of possible \strips\ models (same parameters and same constraints).
+For instance, we claim that all the schemas in the mentioned navigation model are {\em comparable} while {\small\tt stack} and {\small\tt pickup} from a four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
 Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
@@
 
-\section{Recognition of {\em Turing Machines}}
-\label{sec:asPlanning}
-We analyze now {\em model recognition} when the input set of given \strips\ models represent different {\em Turing Machines}.
-
-A {\em Turing machine} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma,\Upsilon,\square,\delta}$:
-\begin{itemize}
-\item $Q$ is a finite and non-empty set of machine states with the {\em initial state} $q_0\in Q$ and the {\em terminal states} $Q_{\bot}\subseteq Q$.  
-\item $\Sigma$ is the {\em tape alphabet}, a finite non-empty set of symbols with the {\em input alphabet} $\Upsilon\subseteq\Sigma$ (symbols allowed to initially appear in the tape) and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
-\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}.
-For each pair of tape symbol and non-terminal machine state $\delta$ defines: (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
-If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
-\end{itemize}
-
-Figure~\ref{tab:tm-anbncn} shows the $\delta$ function of a {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
-The {\em tape alphabet} is $\Sigma=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} $\Upsilon=\{a,b,c,\square\}$ and the machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ (where \underline{$q_5$} is the only acceptor state).
-
-\begin{figure}
-\begin{center}
-    \begin{tabular}{| c || c | c | c | c | c | c |}
-    \hline
-      & $q_0$ & $q_1$ & $q_2$ & $q_3$ & $q_4$ & \underline{$q_5$} \\ \hline\hline
-    a & x,r,$q_1$ & a,r,$q_1$ & - &  a,l,$q_3$ & - & - \\ \hline
-    b & - & y,r,$q_2$ & b,r,$q_2$ & b,l,$q_3$ & - & - \\ \hline
-    c & - & - & z,l,$q_3$ & - & - & - \\ \hline
-    x & - & - & - & x,r,$q_0$ & - & - \\ \hline
-    y & y,r,$q_4$ & y,r,$q_1$ & - & y,l,$q_3$ & y,r,$q_4$ & - \\ \hline
-    z & - & - & z,r,$q_2$ & z,l,$q_3$ & z,r,$q_4$ & - \\\hline
-    $\square$ & - & - & - & - & $\square$,r,$q_5$  & - \\                
-    \hline
-    \end{tabular}
-\end{center}
-  \caption{\small Seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language (\underline{$q_5$} is the only acceptor state).}
-  \label{tab:tm-anbncn}
-\end{figure}
-
-\subsection{\strips\ modeling of {\em Turing Machines}}
-Our \strips\ model of the {\em transition function} $\delta$ of a given {\em Turing Machine} $\mathcal{M}$ is built as follows:
-\begin{itemize}
-\item Fluents $F$ are instantiated from a set $\Psi$ of four {\em predicates}: {\small\tt (head ?x)} that encodes the current position of the header in the tape.
-{\small\tt (next ?x1 ?x2)} encoding that cell {\tt ?x2} follows cell {\tt ?x1} in the tape.
-{\small\tt (symbol-$\sigma$ ?x)} encoding that the tape cell {\tt ?x} contains the symbol $\sigma\in\Sigma$.
-{\small\tt (state-$q$)} encoding that $q\in Q$ is the current machine state.
-Given a set of {\em objects} $\Omega$ that represent the cells in the tape of the given Turing Machine, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of the predicates in $\Psi$.
-\item Actions $A$ are instantiated from \strips\ operator schema. For each transition in $\delta$, a \strips\ schema is defined:
-\begin{itemize}
-\item The {\bf header} is {\small\tt transition-id(?xl ?x ?xr)} where $id$ uniquely identifies the transition in $\delta$. Parameters $?xl$, $?x$ and $?xr$ are tape cells.
-\item The {\bf preconditions} are {\small\tt(head ?x)} and {\small\tt (next ?xl ?x) (next ?x ?xr)} so $?x$ is the tape cell pointed by the header and $?xl/?xr$ its left/right neighbors.
-Preconditions {\small\tt(symbol-$\sigma$ ?x)} and {\small\tt (state-$q$)} are also included to capture the symbol pointed by the header and the current machine state.
-\item The {\bf delete effects} remove the symbol currently pointed by the header and the current machine state while {\bf positive effects} set the new symbol pointed by the header and the new machine state according to $\delta$.
-\end{itemize}
-\end{itemize}
-
-The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn}.
-The full encoding of the {\em Turing Machine} of Figure~\ref{tab:tm-anbncn} produces a total of sixteen \strips\ action schema.
-\begin{figure}
-\begin{scriptsize}
-\begin{lstlisting}
-(:action transition-1      ;;; a,$q_0\rightarrow$ x,r,$q_1$
-  :parameters (?xl ?x ?xr)
-  :precondition (and (head ?x) (symbol-a ?x) (state-$q_0$)
-                     (next ?xl ?x) (next ?x ?xr))
-  :effect (and (not (head ?x)) 
-               (not (symbol-a ?x)) (not (state-$q_0$))
-               (head ?xr) (symbol-x ?x) (state-$q_1$)))
-\end{lstlisting}
-\end{scriptsize}
- \caption{\small \strips\ action schema that models the transition $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{tab:tm-anbncn}.}
-\label{fig:update-rule}
-\end{figure}
-
-With regard to our \strips\ model for {\em Turing Machines}, executions of a {\em Turing Machine} are definable as an action sequence $\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-For instance, the execution of the {\em Turing Machine} defined in Figure~\ref{tab:tm-anbncn} with an initial tape $abc\square\square\square\ldots$ produces the eight-action plan {\small $(a,q_0\rightarrow x,r,q_1)$, $(b,q_1\rightarrow y,r,q_2)$, $(c,q_2\rightarrow z,l,q_3)$, $(y,q_3\rightarrow y,l,q_3)$, $(x,q_3\rightarrow x,r,q_0)$, $(y,q_0\rightarrow y,r,q_4)$, $(z,q_4\rightarrow z,r,q_4)$, $(\square,q_4\rightarrow \square,r,\underline{q_5})$}.
-
-\subsection{The $P(\mathcal{M}|\tau)$ probability for Turing Machines}
-Assuming that the actual applied transitions is unknown means that the observation $\tau$ of the execution of a Turing Machine contains no actions, it is simply a sequence of states $\tau=\tup{s_1, \ldots , s_m}$.
-Further, assuming that the internal machine state is unknown means that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt (state-$q$)} fluents, with $q\in Q$ and $q\neq q_0$.
-Finally, assuming that the values of several tape cells is unknown means that fluents of the kind {\small\tt (symbol-$\sigma$ ?x)} are missing (i.e.
-unobserved) for some state $s_i\in \tau$ s.t.
-$1\leq i\leq m$.
-These facts affects to the a priori probaility of the possible observations.
-
-On the other hand, the model edition for Turing Machines is limited to these subset of possible positive effects: {\tt (head ?xr)} or {\tt (head ?xl)}, {\tt(symbol-$\sigma$ ?x)} with $\sigma \in \Sigma$ and last but not least, {\tt (state-q)} with $q\in Q$.
-No other positive effects, preconditions, or negative effects are required to be edited.
-This fact reduces the possible edition operations and affects to the a priori probaility of the possible action models.
-
 \section{Model recognition as planning}
@@
 
-\subsection{Conditional effects}
+\subsection{The compilation formalization}
 Conditional effects allow us to compactly define our compilation.
-An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-
-An action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
-\[
-triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
-\]
-
-The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
-
-\subsection{The compilation formalization}
-Given a \strips\ model $\mathcal{M}\in M$ and the observation $\tau$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task $P'=\tup{F',A',I',G'}$ such that:
+Given a \strips\ model $\mathcal{M}\in M$ and the observation $\tau$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditionl effects $P'=\tup{F',A',I',G'}$ such that:
 \begin{itemize}
@@
 
+\subsection{Recognition of {\em Regular Automatae}}
+We analyze now {\em model recognition} when the input set of given set of models represent different {\em regular automatae}.
+
+A {\em Regular automatae} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma\square,\delta}$:
+\begin{itemize}
+\item $Q$ is a finite and non-empty set of machine states with the {\em initial state} $q_0\in Q$ and the {\em terminal states} $Q_{\bot}\subseteq Q$.  
+\item $\Sigma$ is the {\em input alphabet}, a finite non-empty set of symbols and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
+\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}.
+For each pair of tape symbol and non-terminal machine state $\delta$ defines: (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
+If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
+\end{itemize}
+
+Figure~\ref{fig:regautomatae} illustrate a four-symbol four-state {\em regular automata} for recognizing the $\{(abc)^n : n \geq 1 \}$ language.
+The {\em input alphabet} is $\Sigma=\{a,b,c,\square\}$, and the machine states are $Q=\{q_0,q_1,q_2,\underline{q_3}\}$ (where \underline{$q_3$} is the only acceptor state).
+
+\begin{figure}
+  \begin{scriptsize}
+  \begin{center}
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
+	  \node[state] (A)              {$Q_0$};
+	  \node[state] (B) [right of=A] {$Q_1$};
+	  \node[state] (C) [right of=B] {$Q_2$};
+	  \node[state] (D) [right of=C] {$Q_3$};          
+	  \path
+(A) edge  node {a} (B)
+(B) edge  node {b} (C)
+(C) edge  node {c} (D)
+(D) edge [bend left]  node {a} (B)
+;
+	\end{tikzpicture}
+  \end{center}
+  \end{scriptsize}  
+ \caption{\small Four-symbol four-state {\em regular automata} for recognizing the $\{(abc)^n : n \geq 1 \}$ language.}
+\label{fig:regautomatae}
+\end{figure}
+
+The \strips\ action schema of Figure~\ref{fig:reg-update-rule} models the rule $a,q_0\rightarrow q_1$ of the {\em regular automatae} defined in Figure~\ref{fig:regautomatae}.
+The full encoding of the {\em automata} of Figure~\ref{fig:regautomatae} produces a total of sixteen \strips\ action schema.
+
+\begin{figure}
+\begin{scriptsize}
+\begin{lstlisting}
+(:action transition-1      ;;; a,$q_0\rightarrow$ $q_1$
+  :parameters (?x ?xr)
+  :precondition (and (head ?x) (symbol-a ?x) (state-$q_0$)
+                     (next ?x ?xr))
+  :effect (and (not (head ?x)) 
+               (not (symbol-a ?x)) (not (state-$q_0$))
+               (head ?xr) (state-$q_1$)))
+\end{lstlisting}
+\end{scriptsize}
+ \caption{\small \strips\ action schema that models the transition $q_0\rightarrow q_1$ of the automata defined in Figure~\ref{fig:regautomatae}.}
+\label{fig:regupdate-rule}
+\end{figure}
+
+\subsection{Recognition of {\em Turing Machines}}
+We analyze now {\em model recognition} when the input set of given set of models represent different {\em Turing machines}.
+
+A {\em Turing machine} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma,\Upsilon,\square,\delta}$:
+\begin{itemize}
+\item $Q$ is a finite and non-empty set of machine states with the {\em initial state} $q_0\in Q$ and the {\em terminal states} $Q_{\bot}\subseteq Q$.  
+\item $\Sigma$ is the {\em tape alphabet}, a finite non-empty set of symbols with the {\em input alphabet} $\Upsilon\subseteq\Sigma$ (symbols allowed to initially appear in the tape) and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
+\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}.
+For each pair of tape symbol and non-terminal machine state $\delta$ defines: (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
+If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
+\end{itemize}
+
+Figure~\ref{fig:tautomatae} illustrate a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
+The {\em tape alphabet} is $\Sigma=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} $\Upsilon=\{a,b,c,\square\}$ and the machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ (where \underline{$q_5$} is the only acceptor state).
+
+\begin{figure}
+  \begin{scriptsize}
+  \begin{center}
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.8cm,semithick]
+	  \node[state] (A)              {$Q_0$};
+	  \node[state] (B) [right of=A] {$Q_1$};
+	  \node[state] (C) [right of=B] {$Q_2$};
+	  \node[state] (D) [right of=C] {$Q_3$};
+	  \node[state] (E) [below of=A] {$Q_4$};
+	  \node[state] (F) [right of=E] {$Q_5$};
+
+	  \path
+(A) edge  node {a/x,right} (B)
+(B) edge [align=left, loop above] node {a/a,right\\y/y,right} (B)
+(B) edge  node {b/y,right} (C)
+(C) edge [align=left, loop above] node {z/z,right\\b/b,right} (C)
+(C) edge  node {c/z,left} (D)
+(D) edge [align=left, loop right] node {a/a,left\\b/b,left\\y/y,left\\z/z,left} (D)
+(D) edge [bend left=45] node {x/x,right} (A)
+
+(A) edge  node {y/y,right} (E)
+(E) edge [align=left, loop left] node {y/y,right\\z/z,right} (E)
+(E) edge  node {b/b,right} (F)
+;
+	\end{tikzpicture}
+  \end{center}
+  \end{scriptsize}  
+ \caption{\small Seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.}
+\label{fig:tautomatae}
+\end{figure}
+
+The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{fig:tautomatae}.
+The full encoding of the {\em Turing Machine} of Figure~\ref{fig:tautomatae} produces a total of sixteen \strips\ action schema.
+
+\begin{figure}
+\begin{scriptsize}
+\begin{lstlisting}
+(:action transition-1      ;;; a,$q_0\rightarrow$ x,r,$q_1$
+  :parameters (?xl ?x ?xr)
+  :precondition (and (head ?x) (symbol-a ?x) (state-$q_0$)
+                     (next ?xl ?x) (next ?x ?xr))
+  :effect (and (not (head ?x)) 
+               (not (symbol-a ?x)) (not (state-$q_0$))
+               (head ?xr) (symbol-x ?x) (state-$q_1$)))
+\end{lstlisting}
+\end{scriptsize}
+ \caption{\small \strips\ action schema that models the transition $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{fig:tautomatae}.}
+\label{fig:update-rule}
+\end{figure}
+
+With regard to our \strips\ model for {\em Turing Machines}, executions of a {\em Turing Machine} are definable as an action sequence $\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+For instance, the execution of the {\em Turing Machine} defined in Figure~\ref{fig:tautomatae} with an initial tape $abc\square\square\square\ldots$ produces the eight-action plan {\small $(a,q_0\rightarrow x,r,q_1)$, $(b,q_1\rightarrow y,r,q_2)$, $(c,q_2\rightarrow z,l,q_3)$, $(y,q_3\rightarrow y,l,q_3)$, $(x,q_3\rightarrow x,r,q_0)$, $(y,q_0\rightarrow y,r,q_4)$, $(z,q_4\rightarrow z,r,q_4)$, $(\square,q_4\rightarrow \square,r,\underline{q_5})$}.
+
+Assuming that the actual applied transitions is unknown means that the observation $\tau$ of the execution of a Turing Machine contains no actions, it is simply a sequence of states $\tau=\tup{s_1, \ldots , s_m}$.
+Further, assuming that the internal machine state is unknown means that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt (state-$q$)} fluents, with $q\in Q$ and $q\neq q_0$.
+Finally, assuming that the values of several tape cells is unknown means that fluents of the kind {\small\tt (symbol-$\sigma$ ?x)} are missing (i.e.
+unobserved) for some state $s_i\in \tau$ s.t.
+$1\leq i\leq m$.
+These facts affects to the a priori probaility of the possible observations.
+
+On the other hand, the model edition for Turing Machines is limited to these subset of possible positive effects: {\tt (head ?xr)} or {\tt (head ?xl)}, {\tt(symbol-$\sigma$ ?x)} with $\sigma \in \Sigma$ and last but not least, {\tt (state-q)} with $q\in Q$.
+No other positive effects, preconditions, or negative effects are required to be edited.
+This fact reduces the possible edition operations and affects to the a priori probaility of the possible action models.
+
 \subsection{Results}
@@
 \label{sec:conclussions}
+The task of {\em model recognition} can also be understood as a classification task where each class is represented with a different planning model and the observed plan execution is the example to classify.
+With this regard, planning model that is associated to a class is acting as a class prototype that the summarizes all the plan executions that could be synthesized with that model or in other words, all the examples that belong to that class.
 

### 2018-09-17 doc/icaps-19/model-recognition.tex
updating pages 1,2  icaps19 paper

@@
 \begin{abstract} 
-Given a partially observed plan execution and a set of possible planning models, {\em model recognition} is the task of identifying which model in the set produced the observed plan execution.
+Given a partially observed plan execution and a set of possible planning models (share same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set produced the observed plan execution.
 The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a \strips\ model to produce a given partially observed plan execution.
-This method, that we called {\em model recognition as planning}, is robust to missing data in the observed plan execution (i.e.
-missing intermediate states and actions) given as input besides, it is computable with an off-the-shelf classical planner.
+This method, that we called {\em model recognition as planning}, is robust to missing data in the given observed plan execution (i.e.
+missing intermediate states and actions) besides, it is computable with an off-the-shelf classical planner.
 The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different {\em regular automatae} and different {\em Turing machines}.
-We show that {\em model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unknown.
+We show that {\em model recognition as planning} succeeds to identify the executed automata despite the internal machine states or actual applied transitions, are unknown.
 \end{abstract}
@@
   \end{scriptsize}  
- \caption{\small Plan execution of a robot navigating a $5\times 5$ grid.}
+ \caption{\small Execution of a robot navigation plan in a $5\times 5$ grid.}
 \label{fig:grid-example}
@@
   (:action inc-x
-    :parameters (?px1 ?px2)
-    :precondition (and (x-pos ?px1) (next ?px1 ?px2) (even))
-    :effect (and (not (x-pos ?px1)) (x-pos ?px2)))
+    :parameters (?x1 ?p2)
+    :precondition (and (x-coord ?x1) (next ?x1 ?x2) (even))
+    :effect (and (not (x-coord ?x1)) (x-coord ?x2)))
 
   (:action dec-x
-    :parameters (?px1 ?px2)
-    :precondition (and (x-pos ?px1) (next ?px2 ?px1) (odd))
-    :effect (and (not (x-pos ?px1)) (x-pos ?px2)))
+    :parameters (?x1 ?px2)
+    :precondition (and (x-coord ?x1) (next ?x2 ?x1) (odd))
+    :effect (and (not (x-coord ?x1)) (x-coord ?x2)))
 
   (:action inc-y-min-odd
-    :parameters (?py1 ?py2)
-    :precondition (and (y-pos ?py1) (not (next ?py2 ?py1)) (odd))
-    :effect (and (not (y-pos ?py1)) (y-pos ?py2) (not (odd)) (even)))
+    :parameters (?y1 ?y2)
+    :precondition (and (y-coord ?y1) (not (next ?y2 ?y1)) (odd))
+    :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (odd)) (even)))
 
   (:action inc-y-min-even
-    :parameters (?py1 ?py2)
-    :precondition (and (y-pos ?py1) (not (next ?py2 ?py1)) (even))
-    :effect (and (not (y-pos ?py1)) (y-pos ?py2) (not (even)) (odd)))
+    :parameters (?y1 ?y2)
+    :precondition (and (y-coord ?y1) (not (next ?y2 ?y1)) (even))
+    :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (even)) (odd)))
 
@@
     :parameters (?py1 ?py2)
-    :precondition (and (y-pos ?py1) (not (next ?py1 ?py2)) (odd))
-    :effect (and (not (y-pos ?py1)) (y-pos ?py2) (not (odd)) (even)))
+    :precondition (and (y-coord ?y1) (not (next ?y1 ?y2)) (odd))
+    :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (odd)) (even)))
 
   (:action inc-y-max-even
-    :parameters (?py1 ?py2)
-    :precondition (and (y-pos ?py1) (not (next ?py1 ?py2)) (even))
-    :effect (and (not (y-pos ?py1)) (y-pos ?py2) (not (even)) (odd)))
+    :parameters (?y1 ?y2)
+    :precondition (and (y-coord ?y1) (not (next ?y1 ?y2)) (even))
+    :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (even)) (odd)))
   \end{verbatim}           
   \end{tiny}  
- \caption{\small Action model for a robot navigation in a $5\times 5$ grid.}
+ \caption{\small Action model for a robot navigation in a $n\times n$ grid.}
 \label{fig:model-example}
@@
 This paper introduces the task of {\em model recognition}.
-Given a partially observed plan execution and a set of possible planning models, {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation.
-To better illustrate {\em model recognition}, imagine a robot in a grid whose navigation is determined by the \strips\ model shown in Figure~\ref{fig:model-example}.
-According to this navigation model the robot increments its {\em x-coordinate} when it is at an {\em even} row while, at {\em odd rows}, the {\em x-coordinate} is decremented.
-Further, the robot only increments the {\em y-coordinate} either at the end or at the begining of rows.
+Given a partially observed plan execution and a set of possible planning models (share the same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation.
+To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model shown in Figure~\ref{fig:model-example}.
+According to this model the robot increments its {\em x-coordinate} when it is at an {\em even} row while, at {\em odd rows}, decrements the {\em x-coordinate}.
+Further, the robot exclusively increments the {\em y-coordinate} either at the {\em end} or at the {\em begining} of a row.
 Apart from this particular navigation model, different models can be defined within the same state variables and these models can determine different kinds of robot navigation.
-Given an observation of a plan execution (e.g, the six-step plan illustrated by Figure~\ref{fig:grid-example}) the {\em model recognition} task aims here to identify which model is more likely of producing the given observation.
+Given an observation of a plan execution (e.g, the six-step plan illustrated by Figure~\ref{fig:grid-example}) the {\em model recognition} task aims here to identify which navigation model produced that observation.
 
@@
 
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose dynamics are specified with two functions: $App(s)$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
-In this work we are specifying action dynamics with the \strips\ model.
+A {\em classical planning frame} is a pair $\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose dynamics are specified with two functions: $App(s)$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
+In this work we specify action dynamics with the \strips\ model.
 With this regard, an action $a\in A$ is defined with:
@@
 \item $\pre(a)\in\mathcal{L}(F)$, the {\em preconditions} of $a$, is the set of literals that must hold for the action $a\in A$ to be applicable.
-\item $\eff^+(a)\in\mathcal{L}(F)$, the {\em positive effects} of $a$, is the set of literals that are true after the application of the action $a\in A$.
+\item $\eff^+(a)\in\mathcal{L}(F)$, the {\em positive effects} of $a$, is the set of literals that are true after the application of action $a\in A$.
 \item $\eff^-(a)\in\mathcal{L}(F)$, the {\em negative effects} of $a$, is the set of literals that are false after the application of the action.
 \end{itemize}
-We assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$ and that actions $a\in A$ are instantiated from given action schemas, as in PDDL.
 We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$, the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{s_1, a_1, \ldots , a_n, s_m}$, an interleaved combination of {\small $1\leq m\leq |\pi|+1$} observed states and {\small $1\leq n\leq |\pi|$} observed actions:
+Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$, the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{s_1, a_1, \ldots , a_n, s_m}$, an interleaved combination of {\small $1\leq m\leq |\pi|+1$} observed states and {\small $1\leq n\leq |\pi|$} observed actions such that:
 \begin{itemize}
@@
 \item Computing the conditional probability $P(\tau|\mathcal{M})$.
-Our approach is to estimate this value according to the minimum amount of {\em edition} that is required by $\mathcal{M}$ to produce a plan $\pi^*_\tau$ such that:
-\begin{enumerate}
-\item $\pi^*_\tau$ is an optimal solution for the classical planning problem $P$. 
-\item $\tau$ is an observation of the execution of $\pi^*_\tau$ on the classical planning problem $P$.
-\end{enumerate}
+Our approach is to estimate this value according to the minimum {\em amount of edits} that is required by the model $\mathcal{M}$ to produce a plan $\pi_\tau$ such that $\tau$ is an observation of the execution of $\pi_\tau$ on the classical planning problem $P$.
+The {\em edit distance} is a similarity metric, traditionally computed over strings or graphs, that has been proved succesful for {\em pattern recognition} ~\cite{masek1980faster,bunke1997relation}.
+In this work this similarity metric refers to the edition of classical planning models.
 \item Applying the Bayes rule to obtain the normalized posterior probabilities (the $P(\mathcal{M}|\tau)$ probabilities must sum 1 for all the $\mathcal{M}\in M$).
 \end{enumerate}
-
-Note that the second step assumes that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
-A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
 
@@
 Here we analyze the particular instantiation of the {\em model recognition} task where the dynamics of the actions $A$ (i.e.
-the $App$ and $\theta$ functions) are specified with \strips\ action schemas.
-Figure~\ref{fig:model-example} shows the PDDL code for the action schema of the mentioned navigation model (Figure).
+the $App$ and $\theta$ functions) are specified with \strips\ action schemas like the ones shown in Figure~\ref{fig:model-example}.
 
 \subsection{The space of \strips\ models}
+We assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$ and that actions $a\in A$ are instantiated from given action schemas, as in PDDL.
+
 Like in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
@@
 
+In this work we do not assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
+A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
+
 \bibliographystyle{aaai}

### 2018-09-17 doc/icaps-19/model-recognition.tex
reformulating the space of possible models paper ICAPS19

@@
   (:action dec-x
-    :parameters (?x1 ?px2)
+    :parameters (?x1 ?x2)
     :precondition (and (x-coord ?x1) (next ?x2 ?x1) (odd))
@@
 
-  (:action inc-y-min-odd
+  (:action inc-y-odd
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (not (next ?y2 ?y1)) (odd))
+    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (odd))
     :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (odd)) (even)))
 
-  (:action inc-y-min-even
+  (:action inc-y-even
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (not (next ?y2 ?y1)) (even))
+    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (even))
     :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (even)) (odd)))
 
-  (:action inc-y-max-odd
-    :parameters (?py1 ?py2)
-    :precondition (and (y-coord ?y1) (not (next ?y1 ?y2)) (odd))
+  (:action dec-y-odd
+    :parameters (?y1 ?y2)
+    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (odd))
     :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (odd)) (even)))
 
-  (:action inc-y-max-even
+  (:action dec-y-even
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (not (next ?y1 ?y2)) (even))
+    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (even))
     :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (even)) (odd)))
@@
 To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model shown in Figure~\ref{fig:model-example}.
-According to this model the robot increments its {\em x-coordinate} when it is at an {\em even} row while, at {\em odd rows}, decrements the {\em x-coordinate}.
-Further, the robot exclusively increments the {\em y-coordinate} either at the {\em end} or at the {\em begining} of a row.
+According to this model the robot can increment its {\em x-coordinate} when it is at an {\em even} row while, at {\em odd rows}, can decrement the {\em x-coordinate}.
 Apart from this particular navigation model, different models can be defined within the same state variables and these models can determine different kinds of robot navigation.
@@
 Here we analyze the particular instantiation of the {\em model recognition} task where the dynamics of the actions $A$ (i.e.
-the $App$ and $\theta$ functions) are specified with \strips\ action schemas like the ones shown in Figure~\ref{fig:model-example}.
+the $App$ and $\theta$ functions) are specified with \strips\ action schemas.
+We start formalizing the full space of possible \strips\ models, next we formalize an {\em edit distance} for \strips\ models and eventually, we define the $P(\mathcal{M}|\tau)$ probability distribution for \strips\ models.
 
 \subsection{The space of \strips\ models}
-We assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$ and that actions $a\in A$ are instantiated from given action schemas, as in PDDL.
+An \strips\ schema $\xi$ is defined by three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the fluents that can appear in the preconditions, negative effects and positivie effects of an /strips/ action.
 
-Like in PDDL~\cite{mcdermott1998pddl,fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$, i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ s.t.
-$\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same parameters and same constraints.
+For instance, we claim that all the schemas in the mentioned navigation model are {\em comparable} while {\small\tt stack} and {\small\tt pickup} from a four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
+Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
 
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be an additional set of objects ($\Omega\cap\Omega_v=\emptyset$), that we denote as {\em variable names}, and that is bound by the maximum arity of an action in a given planning frame.
-For instance, in a $5\times 5$ grid $\Omega=\{p_1, p_2, p_3, p_4, p_5\}$ while $\Omega_v=\{v_1, v_2\}$ because all the actions have arity two.
-We define $F_v$, a new set of fluents, $F\cap F_v=\emptyset$, produced instantiating $\Psi$ using only {\em variable names}.
-This set contains ten elements for the mentioned navigation model, $F_v$={\small\tt\{x-pos($v_1$), x-pos($v_2$), y-pos($v_1$), y-pos($v_2$), odd(), even(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+The space of possible \strips\ models is bound by these three elements:
+\begin{itemize}
+\item $\Psi$, is the set of {\em state predicates}. A set of FOL predicates  that shape the propositional state varibles.
+\item $\Psi\cap\Psi_a =\emptyset$, is a set of additional predicates that shape the action headers.
+An action header of an action schema $\xi$ is defined by a predicate $p_\xi=name_\xi(pars_\xi)$ such that $name_\xi$ is the name of action schema $\xi$ and $pars$ is the list of the corresponding parameters.
+For instance, the set of predictes induced from the PDDL schema shown in Figure~\ref{fig:model-example} is $\{${\small\tt inc-x(X1,X2), dec-x(X1,X2), inc-y-odd(X1,X2), inc-y-even(X1,X2), inc-y-odd (X1,X2), inc-y-even(X1,X2)}$\}$.
+\item ${\mathcal C}$, a set of {\em syntactic constraints}.
+These constraints include the \strips\ constraints (we assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$) and domain-specific constraints.
+For instace, in a {\em robot navigation} domain like the modeled in Figure~\ref{fig:model-example}, predicates {\small\tt even()} and {\small\tt odd()} are exclusive so they cannot hold a the same time.
+\end{itemize}
 
-For a given action schema $\xi$, then $F_{\xi}\subseteq F_v$ defines the subset of elements that can appear in the preconditions and effects of that action schema.
-For instance $F_{\tt inc-x}=F_{\tt dec-x}=F_v$ because all actions have arity two.
+Predicates $\Psi_a$ also shape an additional set of objects ($\Omega\cap\Omega_v=\emptyset$) that we denote as {\em variable names} and that is bound by the maximum arity of the given action headers, i.e., $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{\xi} ar(\xi)}$.
+For instance, in a $5\times 5$ grid (like the one in Figure~\label{fig:grid-example} and modeled in Figure~\ref{fig:model-example}) $\Omega=\{o_1, o_2, o_3, o_4, o_5\}$ while $\Omega_v=\{v_1, v_2\}$ because all the actions schema have arity two.
 
+Given the set of {\em state predicates} $\Psi$ and the set of {\em variable names} $\Omega_v$, then $F_{\xi}\subseteq F_v$ defines the subset of elements that can appear in the preconditions and effects of that action schema $\xi$.
+This set is the same contains ten elements for the mentioned navigation model, $F_v$={\small\tt\{x-coord($v_1$), x-coord($v_2$), y-coord($v_1$), y-coord($v_2$), odd(), even(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+
+\begin{definition}[The space of \strips\ models]
+Given the set of {\em state predicates} $\Psi$, the set of {\em variable names} $\Omega_v$ and the set of {\em syntactic constraints} ${\mathcal C}$, the space of models for a given op predicate $p_\xi$ that represents the header of an action schema $\xi$ is given by three sets of inter
+\end{definition}
+
+The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2\times|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions).
+For the mentioned navigation model, $2^{2\times|F_{\xi}|}=1,048,576$ for every action schema.
+
+\subsection{A propositional encoding for \strips\ models}
 Now we are ready to define the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_{\xi}$.
@@
 \begin{figure}
-\begin{scriptsize}
+\begin{tiny}
 \begin{verbatim}
-(pre_x-pos_inc-x_v1) (pre_next_inc-x_v1_v2) (pre_even_inc-x_)
-(del_x-pos_inc-x_v1) (add_x-pos_inc-x_v2)
+;;; Propositional encoding for inc-x(?x1 ?p2)
+(pre_x-coord_inc-x_v1) (pre_next_inc-x_v1_v2) (pre_even_inc-x_)
+(del_x-coord_inc-x_v1) (add_x-coord_inc-x_v2)
+
+;;; Propositional encoding for dec-x(?x1 ?p2)
+(pre_x-coord_inc-x_v1) (pre_next_inc-x_v2_v1) (pre_odd_inc-x_)
+(del_x-coord_inc-x_v1) (add_x-coord_inc-x_v2)
+
+;;; Propositional encoding for inc-y-odd(?y1 ?y2)
+(pre_y-coord_inc-y-odd_v1) (pre_next_inc-y-odd_v1_v2) 
+(pre_even_inc-y-odd_)
+(del_y-coord_inc-y-odd_v1) (del_odd_inc-y-odd_)
+(add_y-coord_inc-y-odd_v2) (add_even_inc-y-odd_)
+
+;;; Propositional encoding for inc-y-even(?y1 ?y2)
+(pre_y-coord_inc-y-even_v1) (pre_next_inc-y-even_v1_v2)
+(pre_even_inc-y-even_)
+(del_y-coord_inc-y-even_v1) (del_even_inc-y-even_)
+(add_y-coord_inc-y-even_v2) (add_odd_inc-y-even_)
+
+;;; Propositional encoding for dec-y-odd(?y1 ?y2)
+(pre_y-coord_dec-y-odd_v1) (pre_next_dec-y-odd_v2_v1)
+(pre_even_dec-y-odd_)
+(del_y-coord_dec-y-odd_v1) (del_odd_dec-y-odd_)
+(add_y-coord_dec-y-odd_v2) (add_even_dec-y-odd_)
+
+;;; Propositional encoding for inc-y-even(?y1 ?y2)
+(pre_y-coord_dec-y-even_v1) (pre_next_dec-y-even_v2_v1)
+(pre_even_dec-y-even_)
+(del_y-coord_dec-y-even_v1) (del_even_dec-y-even_)
+(add_y-coord_dec-y-even_v2) (add_odd_dec-y-even_)
+
 \end{verbatim}
-\end{scriptsize}
- \caption{\small Propositional encoding for the {\tt inc-x} schema previously defined in Figure~\ref{fig:model-example}.}
+\end{tiny}
+ \caption{\small Propositional encoding for the action schema previously defined in Figure~\ref{fig:model-example}.}
 \label{fig:encoding}
 \end{figure}
-
-The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2\times|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions).
-For the mentioned navigation model, $2^{2\times|F_{\xi}|}=1,048,576$ for every action schema.
-We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same space of possible \strips\ models (same parameters and same constraints).
-For instance, we claim that all the schemas in the mentioned navigation model are {\em comparable} while {\small\tt stack} and {\small\tt pickup} from a four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
-Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
 

### 2018-09-18 doc/icaps-19/model-recognition.tex
introducing programmabble inc-x action

@@
   (:action inc-x
-    :parameters (?x1 ?p2)
-    :precondition (and (x-coord ?x1) (next ?x1 ?x2) (even))
-    :effect (and (not (x-coord ?x1)) (x-coord ?x2)))
+    :parameters (?v1 ?v2)
+    :precondition (and (x-coord ?v1) (next ?v1 ?v2) (even))
+    :effect (and (not (x-coord ?v1)) (x-coord ?v2)))
 
   (:action dec-x
-    :parameters (?x1 ?x2)
-    :precondition (and (x-coord ?x1) (next ?x2 ?x1) (odd))
-    :effect (and (not (x-coord ?x1)) (x-coord ?x2)))
+    :parameters (?v1 ?v2)
+    :precondition (and (x-coord ?v1) (next ?v2 ?v1) (odd))
+    :effect (and (not (x-coord ?v1)) (x-coord ?v2)))
 
@@
 the $App$ and $\theta$ functions) are specified with \strips\ action schemas.
-We start formalizing the full space of possible \strips\ models, next we formalize an {\em edit distance} for \strips\ models and eventually, we define the $P(\mathcal{M}|\tau)$ probability distribution for \strips\ models.
+We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} to estimate the $P(\mathcal{M}|\tau)$ probability distribution for \strips\ models.
 
-\subsection{The space of \strips\ models}
-An \strips\ schema $\xi$ is defined by three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the fluents that can appear in the preconditions, negative effects and positivie effects of an /strips/ action.
+\subsection{A propositional encoding for \strips\ action schema}
+\strips\ action schema are a compact representation for specifying planning action models.
+In more detail, a \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+Figure~\ref{fig:model-example} shows six action schema codded in PDDL for a robot navigation in a $n\times n$ grid.
 
-We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same parameters and same constraints.
-For instance, we claim that all the schemas in the mentioned navigation model are {\em comparable} while {\small\tt stack} and {\small\tt pickup} from a four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
+We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same list of parameters.
+For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while {\small\tt stack} and {\small\tt pickup} from a four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
 Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
 
-The space of possible \strips\ models is bound by these three elements:
-\begin{itemize}
-\item $\Psi$, is the set of {\em state predicates}. A set of FOL predicates  that shape the propositional state varibles.
-\item $\Psi\cap\Psi_a =\emptyset$, is a set of additional predicates that shape the action headers.
-An action header of an action schema $\xi$ is defined by a predicate $p_\xi=name_\xi(pars_\xi)$ such that $name_\xi$ is the name of action schema $\xi$ and $pars$ is the list of the corresponding parameters.
-For instance, the set of predictes induced from the PDDL schema shown in Figure~\ref{fig:model-example} is $\{${\small\tt inc-x(X1,X2), dec-x(X1,X2), inc-y-odd(X1,X2), inc-y-even(X1,X2), inc-y-odd (X1,X2), inc-y-even(X1,X2)}$\}$.
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents $[pre|del|add]\_p\_name(\xi)$.
+Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}.
+The interest of this propositional encoding for \strips\ action schema is that it allows to define {\em programmable action schema} that is, actions whose semantics is given by the value of these fluents on the current state.
+For instance, Figure~\ref{fig:programmable} shows a programmable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.
+Note that this programmable action, when the fluents of Figure~\ref{fig:encoding} holds, behaves exactly as is defined in Figure~\ref{fig:model-example}.
+
+\begin{figure}
+\begin{tiny}
+\begin{verbatim}
+;;; Propositional encoding for inc-x(?v1 ?v2)
+(pre_x-coord_v1_inc-x) (pre_next_v1_v2_inc-x) (pre_even__inc-x)
+(del_x-coord_v1_inc-x) (add_x-coord_v2_inc-x)
+
+;;; Propositional encoding for dec-x(?v1 ?v2)
+(pre_x-coord_v1_dec-x) (pre_next_v2_v1_dec-x) (pre_odd___dec-x)
+(del_x-coord_v1_dec-x) (add_x-coord_v2_dec-x)
+
+;;; Propositional encoding for inc-y-odd(?v1 ?v2)
+(pre_y-coord_v1_inc-y-odd) (pre_next_v1_v2_inc-y-odd) 
+(pre_even__inc-y-odd)
+(del_y-coord_v1_inc-y-odd) (del_odd__inc-y-odd)
+(add_y-coord_v2_inc-y-odd) (add_even__inc-y-odd)
+
+;;; Propositional encoding for inc-y-even(?v1 ?v2)
+(pre_y-coord_v1_inc-y-even) (pre_next_v1_v2_inc-y-even)
+(pre_even__inc-y-even)
+(del_y-coord_v1_inc-y-even) (del_even__inc-y-even)
+(add_y-coord_v2_inc-y-even) (add_odd__inc-y-even)
+
+;;; Propositional encoding for dec-y-odd(?v1 ?v2)
+(pre_y-coord_v1_dec-y-odd) (pre_next_v2_v1_dec-y-odd)
+(pre_odd__dec-y-odd)
+(del_y-coord_v1_dec-y-odd) (del_odd__dec-y-odd)
+(add_y-coord_v2_dec-y-odd) (add_even__dec-y-odd)
+
+;;; Propositional encoding for inc-y-even(v1 ?v2)
+(pre_y-coord_v1_dec-y-even) (pre_next_v2_v1_dec-y-even)
+(pre_even__dec-y-even)
+(del_y-coord_v1_dec-y-even) (del_even__dec-y-even)
+(add_y-coord_v2_dec-y-even) (add_odd__dec-y-even)
+\end{verbatim}
+\end{tiny}
+ \caption{\small Propositional encoding for the six schema from Figure~\ref{fig:model-example}.}
+\label{fig:encoding}
+\end{figure}
+
+\begin{figure}
+  \begin{scriptsize}  
+  \begin{verbatim}
+(:action inc-x
+  :parameters (?v1 ?v2)
+  :precondition
+    (and (or (not (pre_x-coord_v1_inc-x)) (x-coord ?v1))
+         (or (not (pre_x-coord_v2_inc-x)) (x-coord ?v2))
+         (or (not (pre_y-coord_v1_inc-x)) (x-coord ?v1))                       
+         (or (not (pre_y-coord_v2_inc-x)) (x-coord ?v2))
+         (or (not (pre_even__inc-x)) (even))
+         (or (not (pre_odd__inc-x)) (odd)))
+         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
+         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
+         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
+         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2)))         
+    )
+    :effect (and
+       (when (del_x-coord_v1_inc-x) (not (x-coord ?v1)))
+       (when (del_x-coord_v2_inc-x) (not (x-coord ?v2)))
+       (when (del_y-coord_v1_inc-x) (not (x-coord ?v1)))
+       (when (del_y-coord_v2_inc-x) (not (x-coord ?v2)))
+       (when (del_even__inc-x) (not (even)))
+       (when (del_odd__inc-x) (not (odd)))
+       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
+       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
+       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
+       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))
+
+       (when (add_x-coord_v1_inc-x) (x-coord ?v1))
+       (when (add_x-coord_v2_inc-x) (x-coord ?v2))
+       (when (add_y-coord_v1_inc-x) (x-coord ?v1))
+       (when (add_y-coord_v2_inc-x) (x-coord ?v2))
+       (when (add_even__inc-x) (even))
+       (when (add_odd__inc-x) (odd))
+       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
+       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
+       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
+       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
+  \end{verbatim}           
+  \end{scriptsize}  
+ \caption{\small Programmable action {\tt\small inc-x(?v1,?v2)} for robot navigation in a $n\times n$ grid.}
+\label{fig:programmable}
+\end{figure}
+
+\subsection{The space of \strips\ schema}
+The space of possible \strips\ schema is bound by these elements:
+\begin{enumerate}
+\item $\Psi$, is the set of {\em state predicates}. A set of FOL predicates that shape the propositional state variables.
+\item $pars(\xi)$, the list of {\em parameters} of that schema. 
+\end{enumerate}
+
+For each action schema $\xi$, given the set of {\em state predicates} $\Psi$, then $F_{\xi}$ defines the subset of elements that can appear in the preconditions and effects of that action schema.
+This set is the same contains ten elements for the mentioned navigation model, {\small\tt\{x-coord($v_1$), x-coord($v_2$), y-coord($v_1$), y-coord($v_2$), odd(), even(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+
+\begin{enumerate}
+\setcounter{enumi}{2}
 \item ${\mathcal C}$, a set of {\em syntactic constraints}.
@@
 For instace, in a {\em robot navigation} domain like the modeled in Figure~\ref{fig:model-example}, predicates {\small\tt even()} and {\small\tt odd()} are exclusive so they cannot hold a the same time.
-\end{itemize}
+\end{enumerate}
 
@@
 For instance, in a $5\times 5$ grid (like the one in Figure~\label{fig:grid-example} and modeled in Figure~\ref{fig:model-example}) $\Omega=\{o_1, o_2, o_3, o_4, o_5\}$ while $\Omega_v=\{v_1, v_2\}$ because all the actions schema have arity two.
-
-Given the set of {\em state predicates} $\Psi$ and the set of {\em variable names} $\Omega_v$, then $F_{\xi}\subseteq F_v$ defines the subset of elements that can appear in the preconditions and effects of that action schema $\xi$.
-This set is the same contains ten elements for the mentioned navigation model, $F_v$={\small\tt\{x-coord($v_1$), x-coord($v_2$), y-coord($v_1$), y-coord($v_2$), odd(), even(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
 
@@
 For the mentioned navigation model, $2^{2\times|F_{\xi}|}=1,048,576$ for every action schema.
-
-\subsection{A propositional encoding for \strips\ models}
-Now we are ready to define the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$, for every $f\in F_{\xi}$.
-These fluents represent a propositional encoding for the preconditions, negative and positive effects of an action schema $\xi$ and hence, they specify the $App$ and $\theta$ functions for all the actions $a\in A$ built from the corresponding \strips\ action schemas.
-In more detail, if a fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ holds, it means that $f$ is a precondition/negative/positive effect in that schema.
-For instance, Figure~\ref{fig:encoding} shows the conjunction of fluents representing the propositional encoding for the preconditions, negative and positive effects of the {\em inc-x} schema previously defined in Figure~\ref{fig:model-example}.
-
-\begin{figure}
-\begin{tiny}
-\begin{verbatim}
-;;; Propositional encoding for inc-x(?x1 ?p2)
-(pre_x-coord_inc-x_v1) (pre_next_inc-x_v1_v2) (pre_even_inc-x_)
-(del_x-coord_inc-x_v1) (add_x-coord_inc-x_v2)
-
-;;; Propositional encoding for dec-x(?x1 ?p2)
-(pre_x-coord_inc-x_v1) (pre_next_inc-x_v2_v1) (pre_odd_inc-x_)
-(del_x-coord_inc-x_v1) (add_x-coord_inc-x_v2)
-
-;;; Propositional encoding for inc-y-odd(?y1 ?y2)
-(pre_y-coord_inc-y-odd_v1) (pre_next_inc-y-odd_v1_v2) 
-(pre_even_inc-y-odd_)
-(del_y-coord_inc-y-odd_v1) (del_odd_inc-y-odd_)
-(add_y-coord_inc-y-odd_v2) (add_even_inc-y-odd_)
-
-;;; Propositional encoding for inc-y-even(?y1 ?y2)
-(pre_y-coord_inc-y-even_v1) (pre_next_inc-y-even_v1_v2)
-(pre_even_inc-y-even_)
-(del_y-coord_inc-y-even_v1) (del_even_inc-y-even_)
-(add_y-coord_inc-y-even_v2) (add_odd_inc-y-even_)
-
-;;; Propositional encoding for dec-y-odd(?y1 ?y2)
-(pre_y-coord_dec-y-odd_v1) (pre_next_dec-y-odd_v2_v1)
-(pre_even_dec-y-odd_)
-(del_y-coord_dec-y-odd_v1) (del_odd_dec-y-odd_)
-(add_y-coord_dec-y-odd_v2) (add_even_dec-y-odd_)
-
-;;; Propositional encoding for inc-y-even(?y1 ?y2)
-(pre_y-coord_dec-y-even_v1) (pre_next_dec-y-even_v2_v1)
-(pre_even_dec-y-even_)
-(del_y-coord_dec-y-even_v1) (del_even_dec-y-even_)
-(add_y-coord_dec-y-even_v2) (add_odd_dec-y-even_)
-
-\end{verbatim}
-\end{tiny}
- \caption{\small Propositional encoding for the action schema previously defined in Figure~\ref{fig:model-example}.}
-\label{fig:encoding}
-\end{figure}
 

### 2018-09-18 doc/icaps-19/model-recognition.tex
updating pages 1,2,3 icaps19 paper

@@
 \begin{abstract} 
-Given a partially observed plan execution and a set of possible planning models (share same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set produced the observed plan execution.
+Given a partially observed plan execution and a set of possible planning models (share same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set produced the input observation.
 The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a \strips\ model to produce a given partially observed plan execution.
@@
 The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different {\em regular automatae} and different {\em Turing machines}.
-We show that {\em model recognition as planning} succeeds to identify the executed automata despite the internal machine states or actual applied transitions, are unknown.
+We show that {\em model recognition as planning} succeeds to identify the executed automata despite internal machine states or actual applied transitions, are unobserved.
 \end{abstract}
@@
             \node[anchor=center] at (3.5, 1.5) {R};
-            \node[anchor=center] at (0.5, 0.5) {I};
-            \draw[thick,style=dotted] (0.7,0.5) -- (4.5,0.5);
+            \node[anchor=center] at (0.5, 0.5) {};
+            \draw[thick,style=dotted] (0.5,0.5) -- (4.5,0.5);
             \draw[thick,style=dotted] (4.5,0.5) -- (4.5,1.5);
@@
   \end{scriptsize}  
- \caption{\small Execution of a robot navigation plan in a $5\times 5$ grid.}
+ \caption{\small Observation of the execution of a robot navigation plan in a $5\times 5$ grid.}
 \label{fig:grid-example}
@@
 \label{sec:introduction}
-{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behaviour.
+{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behavior.
 Plan recognition is considered {\em automated planning} in reverse; while automated planning aims to compute a sequence of actions that accounts for a given goals, plan recognition aims to compute the goals that account for an observed sequence of actions~\cite{geffner:book:2013}.
@@
 Given a partially observed plan execution and a set of possible planning models (share the same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation.
-To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model shown in Figure~\ref{fig:model-example}.
+To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
 According to this model the robot can increment its {\em x-coordinate} when it is at an {\em even} row while, at {\em odd rows}, can decrement the {\em x-coordinate}.
 Apart from this particular navigation model, different models can be defined within the same state variables and these models can determine different kinds of robot navigation.
-Given an observation of a plan execution (e.g, the six-step plan illustrated by Figure~\ref{fig:grid-example}) the {\em model recognition} task aims here to identify which navigation model produced that observation.
+Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} aims here to identify which navigation model produced that observation.
 
 {\em Model recognition} is of interest because once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
-In addition, it enables indentifying different kinds of automatae by observing their execution (it is well-known that diverse automatae representations, like {\em finite state controllers}, {\em push-down automata} or {\em {\sc GOLOG} programs}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}).
+In addition, it enables identifying different kinds of automatae by observing their execution (it is well-known that diverse automatae representations, like {\em finite state controllers}, {\em push-down automata} or {\em {\sc GOLOG} programs}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}).
 
@@
 All of these {\em automatae} are defined within the same {\em alphabet} and same {\em machine states} but different {\em transition functions}.
-We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite its internal machine state or actual applied transitions are unknown.
+We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite internal machine states or actual applied transitions are unobserved.
 
@@
 
-A {\em classical planning frame} is a pair $\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose dynamics are specified with two functions: $App(s)$ that denotes the subset of actions applicable in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
-In this work we specify action dynamics with the \strips\ model.
+A {\em classical planning frame} is a pair $\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose semantics are specified with two functions: $App(s)$ that denotes the subset of actions {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
+In this work we specify the particular action semantics with the semantics of the \strips\ model.
 With this regard, an action $a\in A$ is defined with:
@@
 \subsection{Conditional effects}
+Conditional effects allow classical planning actions to have different semantics according to different values of the current state.
+This model of action effects is useful for compactly defining our {\em model recognition as planning} method.
+
 An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
@@
 \begin{itemize}
-\item $P=\tup{F,A,I,G}$ is a classical planning problem such that the dynamics of the actions in $A$ is unknown because functions $App(s)$ and/or $\theta(s,a)$ are undefined.
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty set of models for the actions in $A$. Each model in $\mathcal{M}\in M$ defines a different function pair ($App$,$\theta$).
-\item $\tau$ is an observation of the execution of a solution plan $\pi$ for $P$.
+\item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} such that the semantics of the actions in $A$ is unknown because functions $App(s)$ and/or $\theta(s,a)$ are undefined.
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty {\em set of models} for the actions in $A$.
+Each model in $\mathcal{M}\in M$ defines a different function pair ($App$,$\theta$).
+\item $\tau$ is an {\em observation} of the execution of a solution plan $\pi$ for $P$.
 \end{itemize}
@@
 \item Computing the conditional probability $P(\tau|\mathcal{M})$.
-Our approach is to estimate this value according to the minimum {\em amount of edits} that is required by the model $\mathcal{M}$ to produce a plan $\pi_\tau$ such that $\tau$ is an observation of the execution of $\pi_\tau$ on the classical planning problem $P$.
-The {\em edit distance} is a similarity metric, traditionally computed over strings or graphs, that has been proved succesful for {\em pattern recognition} ~\cite{masek1980faster,bunke1997relation}.
+Our approach is to estimate this value according to the {\em amount of edits} required by the model $\mathcal{M}$ to produce a plan $\pi_\tau$ such that $\tau$ is an observation of the execution of $\pi_\tau$ on the classical planning problem $P$.
+The {\em edit distance} is a similarity metric that is traditionally computed over strings or graphs and that has been proved succesful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
 In this work this similarity metric refers to the edition of classical planning models.
-\item Applying the Bayes rule to obtain the normalized posterior probabilities (the $P(\mathcal{M}|\tau)$ probabilities must sum 1 for all the $\mathcal{M}\in M$).
+\item Applying the Bayes rule to obtain the normalized posterior probabilities. The $P(\mathcal{M}|\tau)$ probabilities must sum 1 for all the $\mathcal{M}\in M$.
 \end{enumerate}
@@
 \label{sec:asPlanning}
-Here we analyze the particular instantiation of the {\em model recognition} task where the dynamics of the actions $A$ (i.e.
+Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of the actions $A$ (i.e.
 the $App$ and $\theta$ functions) are specified with \strips\ action schemas.
-We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} to estimate the $P(\mathcal{M}|\tau)$ probability distribution for \strips\ models.
-
-\subsection{A propositional encoding for \strips\ action schema}
-\strips\ action schema are a compact representation for specifying planning action models.
-In more detail, a \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-Figure~\ref{fig:model-example} shows six action schema codded in PDDL for a robot navigation in a $n\times n$ grid.
-
-We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same list of parameters.
-For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while {\small\tt stack} and {\small\tt pickup} from a four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
-Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
-
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents $[pre|del|add]\_p\_name(\xi)$.
-Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}.
-The interest of this propositional encoding for \strips\ action schema is that it allows to define {\em programmable action schema} that is, actions whose semantics is given by the value of these fluents on the current state.
-For instance, Figure~\ref{fig:programmable} shows a programmable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.
-Note that this programmable action, when the fluents of Figure~\ref{fig:encoding} holds, behaves exactly as is defined in Figure~\ref{fig:model-example}.
+We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} to estimate the $P(\mathcal{M}|\tau)$ probabilities for \strips\ models.
 
@@
 
+\subsection{A propositional encoding for \strips\ action schema}
+\strips\ action schema provide a compact representation for specifying classical planning action models.
+In more detail, a \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+Figure~\ref{fig:model-example} shows six action schema codded in PDDL for a robot navigation in a $n\times n$ grid.
+
+We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same list of parameters.
+For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} from a four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
+Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
+
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents $[pre|del|add]\_p\_name(\xi)$.
+Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}.
+The interest of having a propositional encoding for \strips\ action schema is that it allows to define {\em programmable actions} that is, actions whose semantics is given by the value of these particular fluents on the current state.
+For instance, Figure~\ref{fig:programmable} shows the programmable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.
+Note that this programmable schema, when the fluents of Figure~\ref{fig:encoding} holds, behaves exactly as is defined in Figure~\ref{fig:model-example}.
+Further it allows to swap the semantics of two programable schemas provided that they are {\em comparable}.
+
 \begin{figure}
-  \begin{scriptsize}  
+  \begin{tiny}  
   \begin{verbatim}
@@
   \end{verbatim}           
-  \end{scriptsize}  
- \caption{\small Programmable action {\tt\small inc-x(?v1,?v2)} for robot navigation in a $n\times n$ grid.}
+  \end{tiny}  
+ \caption{\small Programmable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
 \label{fig:programmable}
@@
 \begin{enumerate}
-\item $\Psi$, is the set of {\em state predicates}. A set of FOL predicates that shape the propositional state variables.
-\item $pars(\xi)$, the list of {\em parameters} of that schema. 
+\item $\Psi$, is the set of {\em state predicates} that shape the propositional state variables of the model.
+\item $pars(\xi)$, the list of {\em parameters} of the corresponding schema $\xi$. 
 \end{enumerate}
@@
 For each action schema $\xi$, given the set of {\em state predicates} $\Psi$, then $F_{\xi}$ defines the subset of elements that can appear in the preconditions and effects of that action schema.
-This set is the same contains ten elements for the mentioned navigation model, {\small\tt\{x-coord($v_1$), x-coord($v_2$), y-coord($v_1$), y-coord($v_2$), odd(), even(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+For the actions schema defined in~\ref{fig:model-example} this set is the same and contains the following ten elements, {\small\tt\{x-coord($v_1$), x-coord($v_2$), y-coord($v_1$), y-coord($v_2$), odd(), even(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
 

### 2018-09-19 doc/icaps-19/model-recognition.tex
updated page 1 icaps19 paper

@@
 \begin{abstract} 
-Given a partially observed plan execution and a set of possible planning models (share same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set produced the input observation.
+Given the partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set produced the given observation.
 The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a \strips\ model to produce a given partially observed plan execution.
-This method, that we called {\em model recognition as planning}, is robust to missing data in the given observed plan execution (i.e.
-missing intermediate states and actions) besides, it is computable with an off-the-shelf classical planner.
-The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different {\em regular automatae} and different {\em Turing machines}.
+This method, that we called {\em model recognition as planning}, is robust to missing data (missing intermediate states and actions) in the observed plan execution  besides, it is computable with an off-the-shelf classical planner.
+The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different kinds of {\em finite state machines}.
 We show that {\em model recognition as planning} succeeds to identify the executed automata despite internal machine states or actual applied transitions, are unobserved.
 \end{abstract}
+
+\section{Introduction}
+\label{sec:introduction}
+{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behavior.
+Plan recognition is considered {\em automated planning} in reverse; while automated planning aims to compute a sequence of actions that accounts for a given goals, plan recognition aims to compute the goals that account for an observed sequence of actions~\cite{geffner:book:2013}.
+
+Diverse approaches has been proposed for plan recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{carberry2001techniques,sukthankar2014plan}.
+{\em Plan recognition as planning} is the model-based approach for plan recognition~\cite{ramirez2012plan,ramirez2009plan}.
+This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal of the agent, according to the observed plan execution.
+
+This paper introduces the {\em model recognition} task where the target of the recognition process is not a goal but a {\em planning model}.
+Given a partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the given observation.
 
@@
 
+To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
+According to this model the robot can increment its {\em x-coordinate} when it is at an {\em even} row while, at {\em odd rows}, can decrement the {\em x-coordinate}.
+Apart from this particular navigation model, different models can be defined within the same state variables (e.g.
+altering the way {\small\tt (state-q0)} and {\small\tt (state-q1)} are required and updated) and these models can determine different kinds of robot navigation.
+Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} aims here to identify which navigation model produced that observation, despite key information is unobserved (e.g.
+the value of {\small\tt (state-q0)} and {\small\tt (state-q1)} or the particular applied actions).
+
+{\em Model recognition} is of interest because once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
+In addition, it enables identifying different kinds of automatae by observing their execution.
+It is well-known that diverse automatae representations, like {\em finite state controllers}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating,ivankovic2015optimal}.
+
+The paper introduces also {\em model recognition as planning}; a novel method to assess the probability of a given \strips\ model to produce an observed plan execution.
+The method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
+The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em finite state machines}.
+All of these {\em automatae} are defined within the same {\em alphabet} and same {\em machine states} but different {\em transition functions}.
+We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite internal machine states or actual applied transitions are unobserved.
+
 \begin{figure}
@@
     :parameters (?v1 ?v2)
-    :precondition (and (x-coord ?v1) (next ?v1 ?v2) (even))
+    :precondition (and (x-coord ?v1) (next ?v1 ?v2) (state-q0))
     :effect (and (not (x-coord ?v1)) (x-coord ?v2)))
@@
     :parameters (?v1 ?v2)
-    :precondition (and (x-coord ?v1) (next ?v2 ?v1) (odd))
+    :precondition (and (x-coord ?v1) (next ?v2 ?v1) (state-q1))
     :effect (and (not (x-coord ?v1)) (x-coord ?v2)))
+
+  (:action inc-y-even
+    :parameters (?y1 ?y2)
+    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (state-q0))
+    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
+                 (not (state-q0)) (state-q1))
 
@@
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (odd))
-    :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (odd)) (even)))
+    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (state-q1)))
+    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
+                 (not (state-q1)) (state-q0)))
 
-  (:action inc-y-even
+  (:action dec-y-even
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (even))
-    :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (even)) (odd)))
+    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (state-q0))
+    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
+                 (not (state-q0)) (state-q1)))
 
@@
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (odd))
-    :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (odd)) (even)))
-
-  (:action dec-y-even
-    :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (even))
-    :effect (and (not (y-coord ?y1)) (y-coord ?y2) (not (even)) (odd)))
+    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (state-q1))
+    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
+                 (not (state-q1)) (state-q0)))
   \end{verbatim}           
@@
 \end{figure}
-
-\section{Introduction}
-\label{sec:introduction}
-{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behavior.
-Plan recognition is considered {\em automated planning} in reverse; while automated planning aims to compute a sequence of actions that accounts for a given goals, plan recognition aims to compute the goals that account for an observed sequence of actions~\cite{geffner:book:2013}.
-
-Diverse approaches has been proposed for plan recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{carberry2001techniques}.
-{\em Plan recognition as planning} is the model-based approach for plan recognition~\cite{ramirez2012plan,ramirez2009plan}.
-This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal of the agent, according to the observed plan execution.
-
-This paper introduces the task of {\em model recognition}.
-Given a partially observed plan execution and a set of possible planning models (share the same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the input observation.
-To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model the robot can increment its {\em x-coordinate} when it is at an {\em even} row while, at {\em odd rows}, can decrement the {\em x-coordinate}.
-Apart from this particular navigation model, different models can be defined within the same state variables and these models can determine different kinds of robot navigation.
-Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} aims here to identify which navigation model produced that observation.
-
-{\em Model recognition} is of interest because once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
-In addition, it enables identifying different kinds of automatae by observing their execution (it is well-known that diverse automatae representations, like {\em finite state controllers}, {\em push-down automata} or {\em {\sc GOLOG} programs}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating}).
-
-The paper introduces also {\em model recognition as planning}; a novel method to assess the probability of a given \strips\ model to produce an observed plan execution.
-The method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
-The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em regular automatae} and different {\em Turing Machines}.
-All of these {\em automatae} are defined within the same {\em alphabet} and same {\em machine states} but different {\em transition functions}.
-We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite internal machine states or actual applied transitions are unobserved.
 

### 2018-09-19 doc/icaps-19/model-recognition.tex
updated page 2 icaps19 paper

@@
 The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a \strips\ model to produce a given partially observed plan execution.
-This method, that we called {\em model recognition as planning}, is robust to missing data (missing intermediate states and actions) in the observed plan execution  besides, it is computable with an off-the-shelf classical planner.
+This method, that we called {\em model recognition as planning}, is robust to missing intermediate states and actions in the observed plan execution besides, it is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
 The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different kinds of {\em finite state machines}.
@@
 The paper introduces also {\em model recognition as planning}; a novel method to assess the probability of a given \strips\ model to produce an observed plan execution.
-The method is robust to missing data in the intermediate states and actions of the observed plan execution besides, it is computable with an off-the-shelf classical planner.
+The method is robust to missing intermediate states and actions in the observed plan execution besides, it is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
 The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em finite state machines}.
@@
 \label{sec:background}
-This section formalizes classical planning and the observation of the execution of a classical plan.
+This section formalizes the models for classical planning and for the observation of the execution of a classical plan.
 
@@
 
-A {\em classical planning frame} is a pair $\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose semantics are specified with two functions: $App(s)$ that denotes the subset of actions {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
-In this work we specify the particular action semantics with the semantics of the \strips\ model.
+A {\em classical planning frame} is a pair $\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose semantics are specified with two functions: $\rho(s,a)$ that denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
+In this work we specify action semantics with the \strips\ model.
 With this regard, an action $a\in A$ is defined with:
@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
-A {\em plan} for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $s=\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $s=\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 The {\em plan length} is denoted with $|\pi|=n$ .
@@
 \subsection{Conditional effects}
-Conditional effects allow classical planning actions to have different semantics according to different values of the current state.
-This model of action effects is useful for compactly defining our {\em model recognition as planning} method.
+Conditional effects are useful for compactly defining our {\em model recognition as planning} method because they allow planning actions to have different semantics according to the value of the current state.
 
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$, the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{s_1, a_1, \ldots , a_n, s_m}$, an interleaved combination of {\small $1\leq m\leq |\pi|+1$} observed states and {\small $1\leq n\leq |\pi|$} observed actions such that:
+Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$, the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{a_1, \ldots , a_n, s_m}$, an interleaved combination of {\small $1\leq n\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|+1$} observed states such that:
 \begin{itemize}
-\item Observed states may be partial states. The value of certain fluents may be omitted in that states, i.e.~$|s_i|\leq |F|$ for every $0\leq i\leq m$.
+\item Observed actions are {\em consistent} with $\pi$~\cite{ramirez2009plan}.
+This means that the sequence of observed actions $\tup{a_1, \ldots, a_n}$ in $\tau$ is a sub-sequence of the solution plan $\pi$.
+\item Observed states  are {\em consistent} with $\pi$.
+On the one hand, the initial state $s_0=I$ is fully observed while the remaining observed states may be partial, the value of certain fluents in these states may be omitted, i.e.~$|s_i|\leq |F|$ for every $1\leq i\leq m$.
+On the other hand, the sequence of observed states $\tup{s_1, \ldots, s_m}$ in $\tau$ is the same sequence of states traversed by $\pi$ but certain states may also be omitted.
+Formally, $0\leq|s_i|$ for every $1\leq i\leq m$.
+This means that the transitions between two consecutive observed states in $\tau$ may require the execution of more than a single action ($\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound).
+Therefore having $\tau$ does not implies knowing the actual length of $\pi$.
 
-\item The sequence of observed states $\tup{s_1, \ldots, s_m}$ in $\tau$ is the same sequence of states traversed by $\pi$ but certain states may also be omitted.
-Therefore the transitions between two consecutive observed states in $\tau$ may require the execution of more than a single action.
-Formally, $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound.
-This means that having $\tau$ does not implies knowing the actual length of $\pi$.
-\item The sequence of observed actions $\tup{a_1, \ldots, a_n}$ in $\tau$ is a sub-sequence of the solution plan $\pi$.
 \end{itemize}
@@
 
+For instance, the observation of the plan execution shown in Figure~\ref{fig:grid-example} illustrates a $\Phi$-observation because only the fluents of the kind {\tt\small (x-coord ?v)} and {\tt\small (y-coord ?v)} are observed while the value of the fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (state-q0)} and {\tt\small (state-q1)} is unknown.
+
 \section{Model Recognition}
@@
 \begin{itemize}
-\item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} such that the semantics of the actions in $A$ is unknown because functions $App(s)$ and/or $\theta(s,a)$ are undefined.
+\item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} such that the semantics of the actions in $A$ is unknown because functions $\rho(s,a)$ and/or $\theta(s,a)$ are undefined.
 \item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty {\em set of models} for the actions in $A$.
-Each model in $\mathcal{M}\in M$ defines a different function pair ($App$,$\theta$).
+Each model in $\mathcal{M}\in M$ defines a different function pair ($\rho$,$\theta$).
 \item $\tau$ is an {\em observation} of the execution of a solution plan $\pi$ for $P$.
@@
 Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of the actions $A$ (i.e.
-the $App$ and $\theta$ functions) are specified with \strips\ action schemas.
+the $\rho$ and $\theta$ functions) are specified with \strips\ action schemas.
 We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} to estimate the $P(\mathcal{M}|\tau)$ probabilities for \strips\ models.

### 2018-09-19 doc/icaps-19/model-recognition.tex
updated page 2 icaps19 paper

@@
 \item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} such that the semantics of the actions in $A$ is unknown because functions $\rho(s,a)$ and/or $\theta(s,a)$ are undefined.
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty {\em set of models} for the actions in $A$.
-Each model in $\mathcal{M}\in M$ defines a different function pair ($\rho$,$\theta$).
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty {\em set of models} for the actions in $A$ such that, each model in $\mathcal{M}\in M$, defines a different function pair $\tup{\rho,\theta}$.
 \item $\tau$ is an {\em observation} of the execution of a solution plan $\pi$ for $P$.
 \end{itemize}
-A solution to the {\em model recognition} task is the discrete probability distribution $P(\mathcal{M}|\tau)$ that expresses, for each model $\mathcal{M}\in M$, its probability of producing the observation $\tau$.
+
+The {\em discrete probability distribution} $P(\mathcal{M}|\tau)$ expresses, for each model $\mathcal{M}\in M$, its probability of producing the observation $\tau$.
+The {\em naive Bayes classifier} assigns a model $\mathcal{M}\in M$ to the given observation $\tau$ wrt the following expression, $argmax_{\mathcal{M}\in M} P(\mathcal{M})\times P(\tau|\mathcal{M})$.
+With this regard, the solution to the {\em model recognition} task is the model $\mathcal{M}\in M$ that maximizes the previous expression.
 
@@
 \begin{enumerate}
-\item Computing the {\em a priori} probabilities.
+\item Estimating the {\em a priori} probabilities, if they are not given as input.
 $P(\tau)$, that measures how surprising is the given observation and $P(\mathcal{M})$, that expresses if one model is a priori more likely than the others.
-\item Computing the conditional probability $P(\tau|\mathcal{M})$.
-Our approach is to estimate this value according to the {\em amount of edits} required by the model $\mathcal{M}$ to produce a plan $\pi_\tau$ such that $\tau$ is an observation of the execution of $\pi_\tau$ on the classical planning problem $P$.
+\item Computing likelihoods $P(\tau|\mathcal{M})$.  Our approach is to estimate this value according to the {\em amount of edits} required by the model $\mathcal{M}$ to produce a plan $\pi$ such that:
+\begin{enumerate}
+\item $\pi$ solves $P$ and,
+\item $\tau$ is an observation of the execution of $\pi$ on the classical planning problem $P$. 
+\end{enumerate}
 The {\em edit distance} is a similarity metric that is traditionally computed over strings or graphs and that has been proved succesful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
@@
 ;;; Propositional encoding for inc-x(?v1 ?v2)
-(pre_x-coord_v1_inc-x) (pre_next_v1_v2_inc-x) (pre_even__inc-x)
+(pre_x-coord_v1_inc-x) (pre_next_v1_v2_inc-x) 
+(pre_state-q0__inc-x)
 (del_x-coord_v1_inc-x) (add_x-coord_v2_inc-x)
@@
 ;;; Propositional encoding for dec-x(?v1 ?v2)
-(pre_x-coord_v1_dec-x) (pre_next_v2_v1_dec-x) (pre_odd___dec-x)
+(pre_x-coord_v1_dec-x) (pre_next_v2_v1_dec-x) 
+(pre_state-q1___dec-x)
 (del_x-coord_v1_dec-x) (add_x-coord_v2_dec-x)
@@
 (pre_y-coord_v1_inc-y-odd) (pre_next_v1_v2_inc-y-odd) 
-(pre_even__inc-y-odd)
-(del_y-coord_v1_inc-y-odd) (del_odd__inc-y-odd)
-(add_y-coord_v2_inc-y-odd) (add_even__inc-y-odd)
+(pre_state-q0__inc-y-odd)
+(del_y-coord_v1_inc-y-odd) (del_state-q1__inc-y-odd)
+(add_y-coord_v2_inc-y-odd) (add_state-q0__inc-y-odd)
 
@@
 (pre_y-coord_v1_inc-y-even) (pre_next_v1_v2_inc-y-even)
-(pre_even__inc-y-even)
-(del_y-coord_v1_inc-y-even) (del_even__inc-y-even)
-(add_y-coord_v2_inc-y-even) (add_odd__inc-y-even)
+(pre_state-q0__inc-y-even)
+(del_y-coord_v1_inc-y-even) (del_state-q0__inc-y-even)
+(add_y-coord_v2_inc-y-even) (add_state-q1__inc-y-even)
 
@@
 (pre_y-coord_v1_dec-y-odd) (pre_next_v2_v1_dec-y-odd)
-(pre_odd__dec-y-odd)
-(del_y-coord_v1_dec-y-odd) (del_odd__dec-y-odd)
-(add_y-coord_v2_dec-y-odd) (add_even__dec-y-odd)
+(pre_state-q1__dec-y-odd)
+(del_y-coord_v1_dec-y-odd) (del_state-q1__dec-y-odd)
+(add_y-coord_v2_dec-y-odd) (add_state-q0__dec-y-odd)
 
@@
 (pre_y-coord_v1_dec-y-even) (pre_next_v2_v1_dec-y-even)
-(pre_even__dec-y-even)
-(del_y-coord_v1_dec-y-even) (del_even__dec-y-even)
-(add_y-coord_v2_dec-y-even) (add_odd__dec-y-even)
+(pre_state-q0__dec-y-even)
+(del_y-coord_v1_dec-y-even) (del_state-q0__dec-y-even)
+(add_y-coord_v2_dec-y-even) (add_state-q1__dec-y-even)
 \end{verbatim}
@@
          (or (not (pre_y-coord_v2_inc-x)) (x-coord ?v2))
-         (or (not (pre_even__inc-x)) (even))
-         (or (not (pre_odd__inc-x)) (odd)))
+         (or (not (pre_state-q0__inc-x)) (state-q0))
+         (or (not (pre_state-q1__inc-x)) (state-q1)))
          (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
@@
        (when (del_y-coord_v2_inc-x) (not (x-coord ?v2)))
-       (when (del_even__inc-x) (not (even)))
-       (when (del_odd__inc-x) (not (odd)))
+       (when (del_state-q0__inc-x) (not (state-q0)))
+       (when (del_state-q1__inc-x) (not (state-q1)))
        (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
@@
        (when (add_y-coord_v2_inc-x) (x-coord ?v2))
-       (when (add_even__inc-x) (even))
-       (when (add_odd__inc-x) (odd))
+       (when (add_state-q0__inc-x) (state-q0))
+       (when (add_state-q1__inc-x) (state-q1))
        (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
@@
 For each action schema $\xi$, given the set of {\em state predicates} $\Psi$, then $F_{\xi}$ defines the subset of elements that can appear in the preconditions and effects of that action schema.
-For the actions schema defined in~\ref{fig:model-example} this set is the same and contains the following ten elements, {\small\tt\{x-coord($v_1$), x-coord($v_2$), y-coord($v_1$), y-coord($v_2$), odd(), even(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+For the actions schema defined in~\ref{fig:model-example} this set is the same and contains the following ten elements, {\small\tt\{x-coord($v_1$), x-coord($v_2$), y-coord($v_1$), y-coord($v_2$), state-q0(), state-q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
 
@@
 These constraints include the \strips\ constraints (we assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$) and domain-specific constraints.
-For instace, in a {\em robot navigation} domain like the modeled in Figure~\ref{fig:model-example}, predicates {\small\tt even()} and {\small\tt odd()} are exclusive so they cannot hold a the same time.
+For instace, in a {\em robot navigation} domain like the modeled in Figure~\ref{fig:model-example}, predicates {\small\tt state-q0()} and {\small\tt state-q1()} are exclusive so they cannot hold a the same time.
 \end{enumerate}
@@
 
-\subsection{The $P(\mathcal{M}|\tau)$ probability for \strips\ models}
-Since we assume that the dynamics of the actions $A\in P$ is specified with \strips\ action schemas, we can estimate the $P(\mathcal{M}|\tau)$ probability as follows:
-\begin{enumerate}
-\item Assuming {\em a priori} all action models $\mathcal{M}\in M$ are equiprobable.
-There are no reasons to assume that one model is {\em a priori} more likely than the others so, $P(\mathcal{M})=\frac{1}{\prod_{\xi\in\mathcal{M}} 2^{2\times|F_{\xi}|}}$.
-\item Assuming that all the observations of plan executions with maximum $n$ observed actions and $m$ observed states are equiprobable then, $P(\tau)=\frac{1}{|A|^{n}\times2^{m\times|F|}}$.
-\item Estimating the conditional probability $P(\tau|\mathcal{M})$ by mapping the {\em observation edit distance} into a $[0,1]$ likelihood, $1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
-\end{enumerate}
+Estimating the conditional probability $P(\tau|\mathcal{M})$ by mapping the {\em observation edit distance} into a $[0,1]$ likelihood, $1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
 
@@
 Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
-Figure~\ref{fig:compilation} shows the PDDL encoding of the action for applying a programmed operator $stack$ from {\em blocksworld}.
+Figure~\ref{fig:programmable} shows the PDDL encoding of the action for applying a programmed operator.
 \begin{small}
@@
 
-\begin{figure}
-\begin{scriptsize}
-\begin{verbatim}
-(:action apply_stack
-  :parameters (?o1 - object ?o2 - object)
-  :precondition
-   (and (or (not (pre_on_stack_v1_v1)) (on ?o1 ?o1))
-        (or (not (pre_on_stack_v1_v2)) (on ?o1 ?o2))
-        (or (not (pre_on_stack_v2_v1)) (on ?o2 ?o1))
-        (or (not (pre_on_stack_v2_v2)) (on ?o2 ?o2))
-        (or (not (pre_ontable_stack_v1)) (ontable ?o1))
-        (or (not (pre_ontable_stack_v2)) (ontable ?o2))
-        (or (not (pre_clear_stack_v1)) (clear ?o1))
-        (or (not (pre_clear_stack_v2)) (clear ?o2))
-        (or (not (pre_holding_stack_v1)) (holding ?o1))
-        (or (not (pre_holding_stack_v2)) (holding ?o2))
-        (or (not (pre_handempty_stack)) (handempty)))
-  :effect
-   (and (when (del_on_stack_v1_v1) (not (on ?o1 ?o1)))
-        (when (del_on_stack_v1_v2) (not (on ?o1 ?o2)))
-        (when (del_on_stack_v2_v1) (not (on ?o2 ?o1)))
-        (when (del_on_stack_v2_v2) (not (on ?o2 ?o2)))
-        (when (del_ontable_stack_v1) (not (ontable ?o1)))
-        (when (del_ontable_stack_v2) (not (ontable ?o2)))
-        (when (del_clear_stack_v1) (not (clear ?o1)))
-        (when (del_clear_stack_v2) (not (clear ?o2)))
-        (when (del_holding_stack_v1) (not (holding ?o1)))
-        (when (del_holding_stack_v2) (not (holding ?o2)))
-        (when (del_handempty_stack) (not (handempty)))
-        (when (add_on_stack_v1_v1) (on ?o1 ?o1))
-        (when (add_on_stack_v1_v2) (on ?o1 ?o2))
-        (when (add_on_stack_v2_v1) (on ?o2 ?o1))
-        (when (add_on_stack_v2_v2) (on ?o2 ?o2))
-        (when (add_ontable_stack_v1) (ontable ?o1))
-        (when (add_ontable_stack_v2) (ontable ?o2))
-        (when (add_clear_stack_v1) (clear ?o1))
-        (when (add_clear_stack_v2) (clear ?o2))
-        (when (add_holding_stack_v1) (holding ?o1))
-        (when (add_holding_stack_v2) (holding ?o2))
-        (when (add_handempty_stack) (handempty))
-        (when (modeProg) (not (modeProg)))))
-\end{verbatim}
-\end{scriptsize}
- \caption{\small PDDL action for applying an already programmed schema $stack$ (implications are coded as disjunctions).}
-\label{fig:compilation}
-\end{figure}
 When the observation $\tau$ includes observed actions, then the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as they appear in $\tau$.\\

### 2018-09-20 doc/icaps-19/model-recognition.tex
paper icaps19: sections 1,2 and 3 stable. Ready for review

@@
 \begin{abstract} 
-Given the partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set produced the given observation.
-The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a \strips\ model to produce a given partially observed plan execution.
-This method, that we called {\em model recognition as planning}, is robust to missing intermediate states and actions in the observed plan execution besides, it is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
+Given the partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set explains (produced) the given observation.
+The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a \strips\ model to produce a partial observation of a plan execution.
+This method, that we called {\em model recognition as planning}, is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
+{\em Model recognition as planning} is robust to missing intermediate states and actions in the observed plan execution.
 The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different kinds of {\em finite state machines}.
-We show that {\em model recognition as planning} succeeds to identify the executed automata despite internal machine states or actual applied transitions, are unobserved.
+We show that {\em model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unobserved.
 \end{abstract}
@@
 To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model the robot can increment its {\em x-coordinate} when it is at an {\em even} row while, at {\em odd rows}, can decrement the {\em x-coordinate}.
-Apart from this particular navigation model, different models can be defined within the same state variables (e.g.
+According to this model the robot could increment its {\em x-coordinate} at {\em even} rows and decrement it at the {\em odd rows}.
+Apart from this particular navigation model, different models could be defined within the same state variables (e.g.
 altering the way {\small\tt (state-q0)} and {\small\tt (state-q1)} are required and updated) and these models can determine different kinds of robot navigation.
@@
 In addition, it enables identifying different kinds of automatae by observing their execution.
-It is well-known that diverse automatae representations, like {\em finite state controllers}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,segovia2017generating,ivankovic2015optimal}.
+It is well-known that diverse automatae representations, like {\em finite state controllers}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,ivankovic2015optimal,segovia2017generating}.
 
 The paper introduces also {\em model recognition as planning}; a novel method to assess the probability of a given \strips\ model to produce an observed plan execution.
-The method is robust to missing intermediate states and actions in the observed plan execution besides, it is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
-The paper evaluates the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em finite state machines}.
+Our method is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
+{\em Model recognition as planning} is robust to missing intermediate states and actions in the observed plan execution.
+We evaluate the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em finite state machines}.
 All of these {\em automatae} are defined within the same {\em alphabet} and same {\em machine states} but different {\em transition functions}.
@@
 \label{sec:background}
-This section formalizes the models for classical planning and for the observation of the execution of a classical plan.
+This section formalizes the models for {\em classical planning} and for the {\em observations} of a plan execution.
 
@@
 
-A {\em classical planning frame} is a pair $\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose semantics are specified with two functions: $\rho(s,a)$ that denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
+A {\em classical planning frame} is a pair $\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose semantics are specified with two functions: $\rho(s,a)$ that denotes whether an action $a\in A$ is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
 In this work we specify action semantics with the \strips\ model.
@@
 \subsection{Conditional effects}
-Conditional effects are useful for compactly defining our {\em model recognition as planning} method because they allow planning actions to have different semantics according to the value of the current state.
+{\em Conditional effects} allow planning actions to have different semantics according to the value of the current state.
+This feature is useful for compactly defining our method for {\em model recognition as planning}.
 
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$, the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{a_1, \ldots , a_n, s_m}$, an interleaved combination of {\small $1\leq n\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|+1$} observed states such that:
+Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$, the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{a_1^o, \ldots , a_n^o, s_m^o}$, an interleaved combination of {\small $1\leq n\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|+1$} observed states such that:
 \begin{itemize}
 \item Observed actions are {\em consistent} with $\pi$~\cite{ramirez2009plan}.
-This means that the sequence of observed actions $\tup{a_1, \ldots, a_n}$ in $\tau$ is a sub-sequence of the solution plan $\pi$.
-\item Observed states  are {\em consistent} with $\pi$.
-On the one hand, the initial state $s_0=I$ is fully observed while the remaining observed states may be partial, the value of certain fluents in these states may be omitted, i.e.~$|s_i|\leq |F|$ for every $1\leq i\leq m$.
-On the other hand, the sequence of observed states $\tup{s_1, \ldots, s_m}$ in $\tau$ is the same sequence of states traversed by $\pi$ but certain states may also be omitted.
-Formally, $0\leq|s_i|$ for every $1\leq i\leq m$.
-This means that the transitions between two consecutive observed states in $\tau$ may require the execution of more than a single action ($\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, where $k\geq 1$ is unknown and unbound).
-Therefore having $\tau$ does not implies knowing the actual length of $\pi$.
-
+This means that the sequence of observed actions $\tup{a_1^o, \ldots, a_n^o}$ in $\tau$ is a sub-sequence of the solution plan $\pi$.
+\item Observed states are a sub-sequence  of partial states that is {\em consistent} with the sequence of states traversed by $\pi$. 
 \end{itemize}
 
+On the one hand, the initial state $I$ is fully observed while the observed states in $\tau$ may be partial, i.e.
+the value of certain fluents in the intermediate states may be omitted (~$|s_i^o|\leq |F|$ for every $1\leq i\leq m$).
+On the other hand, the sequence of observed states $\tup{s_1^o, \ldots, s_m^o}$ in $\tau$ is the same sequence of states traversed by $\pi$ but certain states may also be omitted.
+Formally, $0\leq|s_i^o|$ for every $1\leq i\leq m$.
+This means that the transitions between two consecutive observed states in $\tau$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where $k\geq 1$ is unknown and unbound).
+Therefore we can conclude that having $\tau$ does not implies knowing the actual length of $\pi$.
+
 \begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\tau$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every $0\leq i\leq m$, each observed state $s_i$ only contains fluents in $\Phi$.
+Given a subset of fluents $\Phi\subseteq F$ we say that $\tau$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every $0\leq i\leq m$, each observed state $s_i^o$ only contains fluents in $\Phi$.
 \end{definition}
 
-For instance, the observation of the plan execution shown in Figure~\ref{fig:grid-example} illustrates a $\Phi$-observation because only the fluents of the kind {\tt\small (x-coord ?v)} and {\tt\small (y-coord ?v)} are observed while the value of the fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (state-q0)} and {\tt\small (state-q1)} is unknown.
+For instance, Figure~\ref{fig:grid-example} illustrates a $\Phi$-observation where $\Phi$ only contains fluents of the kind {\tt\small (x-coord ?v)} and {\tt\small (y-coord ?v)}.
+This means that, for each observed state, only the value of fluents {\tt\small (x-coord ?v)} and {\tt\small (y-coord ?v)} is known while the value of the fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (state-q0)} and {\tt\small (state-q1)} is unknown.
 
@@
 
-The {\em discrete probability distribution} $P(\mathcal{M}|\tau)$ expresses, for each model $\mathcal{M}\in M$, its probability of producing the observation $\tau$.
-The {\em naive Bayes classifier} assigns a model $\mathcal{M}\in M$ to the given observation $\tau$ wrt the following expression, $argmax_{\mathcal{M}\in M} P(\mathcal{M})\times P(\tau|\mathcal{M})$.
-With this regard, the solution to the {\em model recognition} task is the model $\mathcal{M}\in M$ that maximizes the previous expression.
-
-According to the {\em Bayes} rule, the probability of an hypothesis $\mathcal{H}$, provided the observation $\mathcal{O}$, is given by the expression $P(\mathcal{H}|\mathcal{O})=\frac{P(\mathcal{O}|\mathcal{H})P(\mathcal{H})}{P(\mathcal{O})}$.
-In the {\em model recognition} task, hypotheses are about the possible action models $\mathcal{M}\in M$ while the given observation is the partially observed plan execution $\tau$.
-The $P(\mathcal{M}|\tau)$ probability distribution can then be estimated in three steps:
+In the {\em model recognition} task, hypotheses are about the possible action models $\mathcal{M}\in M$ while $\tau$ represents the given observation.
+The {\em discrete probability distribution} $P(\mathcal{M}|\tau)$ expresses, for each model $\mathcal{M}\in M$, its probability of producing the input observation.
+According to the {\em Bayes rule} this probability distribution is given by $P(\mathcal{M}|\tau)=\frac{P(\tau|\mathcal{M})P(\mathcal{M})}{P(\tau)}$ and can be estimated in these three steps:
 \begin{enumerate}
-\item Estimating the {\em a priori} probabilities, if they are not given as input.
-$P(\tau)$, that measures how surprising is the given observation and $P(\mathcal{M})$, that expresses if one model is a priori more likely than the others.
-\item Computing likelihoods $P(\tau|\mathcal{M})$.  Our approach is to estimate this value according to the {\em amount of edits} required by the model $\mathcal{M}$ to produce a plan $\pi$ such that:
+\item Estimating the {\em a priori probabilities} (if they are not already given as input).
+$P(\tau)$, that measures how surprising is the given observation and $P(\mathcal{M})$, that expresses if one model is known to be a priori more likely than the others.
+\item Computing likelihoods $P(\tau|\mathcal{M})$.
+Our approach is to estimate these likelihoods according to the {\em amount of edits} required by the model $\mathcal{M}$ to produce a plan $\pi$ such that:
 \begin{enumerate}
 \item $\pi$ solves $P$ and,
-\item $\tau$ is an observation of the execution of $\pi$ on the classical planning problem $P$. 
+\item $\tau$ is a consistent observation of the execution of $\pi$ on the classical planning problem $P$. 
 \end{enumerate}
-The {\em edit distance} is a similarity metric that is traditionally computed over strings or graphs and that has been proved succesful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
+The {\em edit distance} is a similarity metric that is traditionally computed over {\em strings} or {\em graphs} and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
 In this work this similarity metric refers to the edition of classical planning models.
-\item Applying the Bayes rule to obtain the normalized posterior probabilities. The $P(\mathcal{M}|\tau)$ probabilities must sum 1 for all the $\mathcal{M}\in M$.
+\item Applying the {\em Bayes rule} to obtain the normalized posterior probabilities. The $P(\mathcal{M}|\tau)$ probabilities must sum 1 for all the $\mathcal{M}\in M$.
 \end{enumerate}
+
+The {\em naive Bayes classifier} assigns a model $\mathcal{M}\in M$ to the given observation $\tau$ wrt the following expression, $argmax_{\mathcal{M}\in M} P(\mathcal{M})\times P(\tau|\mathcal{M})$.
+The {\em solution} to the {\em model recognition} task is the model $\mathcal{M}\in M$ (or subset of models in $M$) that maximizes the previous expression.
 

### 2018-09-20 doc/icaps-19/model-recognition.tex
paper icaps19: pages 1,2 and 3 stable. Ready for review

@@
 \label{sec:asPlanning}
-Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of the actions $A$ (i.e.
-the $\rho$ and $\theta$ functions) are specified with \strips\ action schemas.
-We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} to estimate the $P(\mathcal{M}|\tau)$ probabilities for \strips\ models.
+Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of actions (in other words, the corresponding $\rho$ and $\theta$ functions) are specified with \strips\ action schemas.
+
+We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} which allows us to estimate the $P(\tau|\mathcal{M})$ likelihoods for observations of the execution of classical plans and planning \strips\ models.
 
@@
 \subsection{A propositional encoding for \strips\ action schema}
-\strips\ action schema provide a compact representation for specifying classical planning action models.
-In more detail, a \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-Figure~\ref{fig:model-example} shows six action schema codded in PDDL for a robot navigation in a $n\times n$ grid.
+\strips\ action schema provide a compact representation for specifying classical planning models.
+For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema codded in PDDL for a robot navigation in a $n\times n$ grid.
+A \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 
-We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff both share the same list of parameters.
-For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} from a four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
-Last but not least, two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
+We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
+For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
+Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
 
@@
 We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a seven-symbol six-state {\em Turing Machine}.
-The classical planning frame $\Phi=\tup{F,A}$ is the same for all the {\em Turing Machines}, there is an $a\in A$ action for each pair of tape symbol and non-terminal state machine, while the $\theta(s,a)$ function is defined differently for each the machines (using a different \strips\ action model).
+The classical planning frame $\tup{F,A}$ is the same for all the {\em Turing Machines}, there is an $a\in A$ action for each pair of tape symbol and non-terminal state machine, while the $\rho$ and $\theta$ function is defined differently for each the machines (using a different \strips\ action model).
 
@@
 
+Remarkably, the extension of this piece of work to the FOND planning setting~\cite{muise2012improved} is straightforward by simply considering the {\em all-outcomes} determiniztion of the actions with non-determinitic effects~\cite{yoon2007ff}.
+An interesting research direction is however to understand how to apply our approach to planning models where the planning models include actions with probabilistic effects~\cite{younes2005first}.
+
 \bibliographystyle{aaai}

### 2018-09-20 doc/icaps-19/model-recognition.tex
paper icaps19: sections 1,2 and 3 stable. Ready for review

@@
         \end{tikzpicture}
+\hspace*{1cm}        
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
+	  \node[state] (A)              {$q_0$};
+	  \node[state] (B) [right of=A] {$q_1$};          
+	  \path
+(A) edge [bend left] node {inc-y, dec-y} (B)
+(B) edge [bend left] node {inc-y, dec-y} (A);
+	\end{tikzpicture}
   \end{center}
   \end{scriptsize}  
- \caption{\small Observation of the execution of a robot navigation plan in a $5\times 5$ grid.}
+ \caption{\small Observation of the execution of a robot navigation plan in a $5\times 5$ grid and a two state automata for controlling that the robot only increments its x-coordinate at {\em even} rows (i.e.
+at state $q_0$).}
 \label{fig:grid-example}
@@
 Apart from this particular navigation model, different models could be defined within the same state variables (e.g.
-altering the way {\small\tt (state-q0)} and {\small\tt (state-q1)} are required and updated) and these models can determine different kinds of robot navigation.
+altering the way {\small\tt (q0)} and {\small\tt (q1)} are required and updated) and these models can determine different kinds of robot navigation.
 Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} aims here to identify which navigation model produced that observation, despite key information is unobserved (e.g.
-the value of {\small\tt (state-q0)} and {\small\tt (state-q1)} or the particular applied actions).
+the value of {\small\tt (q0)} and {\small\tt (q1)} or the particular applied actions).
 
@@
     :parameters (?v1 ?v2)
-    :precondition (and (x-coord ?v1) (next ?v1 ?v2) (state-q0))
-    :effect (and (not (x-coord ?v1)) (x-coord ?v2)))
+    :precondition (and (xcoord ?v1) (next ?v1 ?v2) (q0))
+    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))
 
@@
     :parameters (?v1 ?v2)
-    :precondition (and (x-coord ?v1) (next ?v2 ?v1) (state-q1))
-    :effect (and (not (x-coord ?v1)) (x-coord ?v2)))
+    :precondition (and (xcoord ?v1) (next ?v2 ?v1) (q1))
+    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))
 
@@
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (state-q0))
-    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
-                 (not (state-q0)) (state-q1))
+    :precondition (and (ycoord ?y1) (next ?y1 ?y2) (q0))
+    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
+                 (not (q0)) (q1))
 
@@
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (next ?y1 ?y2) (state-q1)))
-    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
-                 (not (state-q1)) (state-q0)))
+    :precondition (and (ycoord ?y1) (next ?y1 ?y2) (q1)))
+    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
+                 (not (q1)) (q0)))
 
@@
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (state-q0))
-    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
-                 (not (state-q0)) (state-q1)))
+    :precondition (and (ycoord ?y1) (next ?y2 ?y1) (q0))
+    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
+                 (not (q0)) (q1)))
 
@@
     :parameters (?y1 ?y2)
-    :precondition (and (y-coord ?y1) (next ?y2 ?y1) (state-q1))
-    :effect (and (not (y-coord ?y1)) (y-coord ?y2)
-                 (not (state-q1)) (state-q0)))
+    :precondition (and (ycoord ?y1) (next ?y2 ?y1) (q1))
+    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
+                 (not (q1)) (q0)))
   \end{verbatim}           
@@
 
-For instance, Figure~\ref{fig:grid-example} illustrates a $\Phi$-observation where $\Phi$ only contains fluents of the kind {\tt\small (x-coord ?v)} and {\tt\small (y-coord ?v)}.
-This means that, for each observed state, only the value of fluents {\tt\small (x-coord ?v)} and {\tt\small (y-coord ?v)} is known while the value of the fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (state-q0)} and {\tt\small (state-q1)} is unknown.
+For instance, Figure~\ref{fig:grid-example} illustrates the $\Phi$-observation \{{\tt\scriptsize<(xcoord 1)(ycoord 1)>, <(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\} where $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
+This means that, for each observed state, only the value of fluents {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} is known while the value of the fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)} is unknown.
 
@@
 
-We first formalize \strips\ schemas, then define the full space of possible \strips\ schema and eventually, we introduce an {\em edit distance} which allows us to estimate the $P(\tau|\mathcal{M})$ likelihoods for observations of the execution of classical plans and planning \strips\ models.
+We first formalize \strips\ schemas and define the full space of possible \strips\ schema.
+Eventually, we introduce an {\em edit distance} which allows us to estimate the $P(\tau|\mathcal{M})$ likelihoods for observations of the execution of classical plans and planning \strips\ models.
 
-\begin{figure}
-\begin{tiny}
-\begin{verbatim}
-;;; Propositional encoding for inc-x(?v1 ?v2)
-(pre_x-coord_v1_inc-x) (pre_next_v1_v2_inc-x) 
-(pre_state-q0__inc-x)
-(del_x-coord_v1_inc-x) (add_x-coord_v2_inc-x)
-
-;;; Propositional encoding for dec-x(?v1 ?v2)
-(pre_x-coord_v1_dec-x) (pre_next_v2_v1_dec-x) 
-(pre_state-q1___dec-x)
-(del_x-coord_v1_dec-x) (add_x-coord_v2_dec-x)
-
-;;; Propositional encoding for inc-y-odd(?v1 ?v2)
-(pre_y-coord_v1_inc-y-odd) (pre_next_v1_v2_inc-y-odd) 
-(pre_state-q0__inc-y-odd)
-(del_y-coord_v1_inc-y-odd) (del_state-q1__inc-y-odd)
-(add_y-coord_v2_inc-y-odd) (add_state-q0__inc-y-odd)
-
-;;; Propositional encoding for inc-y-even(?v1 ?v2)
-(pre_y-coord_v1_inc-y-even) (pre_next_v1_v2_inc-y-even)
-(pre_state-q0__inc-y-even)
-(del_y-coord_v1_inc-y-even) (del_state-q0__inc-y-even)
-(add_y-coord_v2_inc-y-even) (add_state-q1__inc-y-even)
-
-;;; Propositional encoding for dec-y-odd(?v1 ?v2)
-(pre_y-coord_v1_dec-y-odd) (pre_next_v2_v1_dec-y-odd)
-(pre_state-q1__dec-y-odd)
-(del_y-coord_v1_dec-y-odd) (del_state-q1__dec-y-odd)
-(add_y-coord_v2_dec-y-odd) (add_state-q0__dec-y-odd)
-
-;;; Propositional encoding for inc-y-even(v1 ?v2)
-(pre_y-coord_v1_dec-y-even) (pre_next_v2_v1_dec-y-even)
-(pre_state-q0__dec-y-even)
-(del_y-coord_v1_dec-y-even) (del_state-q0__dec-y-even)
-(add_y-coord_v2_dec-y-even) (add_state-q1__dec-y-even)
-\end{verbatim}
-\end{tiny}
- \caption{\small Propositional encoding for the six schema from Figure~\ref{fig:model-example}.}
-\label{fig:encoding}
-\end{figure}
-
-\subsection{A propositional encoding for \strips\ action schema}
+\subsection{The space of \strips\ schema}
 \strips\ action schema provide a compact representation for specifying classical planning models.
 For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema codded in PDDL for a robot navigation in a $n\times n$ grid.
+
+\begin{definition}[A \strips\ action schema $\xi$]
 A \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-
-We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
-For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
-Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
-
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents $[pre|del|add]\_p\_name(\xi)$.
-Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}.
-The interest of having a propositional encoding for \strips\ action schema is that it allows to define {\em programmable actions} that is, actions whose semantics is given by the value of these particular fluents on the current state.
-For instance, Figure~\ref{fig:programmable} shows the programmable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.
-Note that this programmable schema, when the fluents of Figure~\ref{fig:encoding} holds, behaves exactly as is defined in Figure~\ref{fig:model-example}.
-Further it allows to swap the semantics of two programable schemas provided that they are {\em comparable}.
-
-\begin{figure}
-  \begin{tiny}  
-  \begin{verbatim}
-(:action inc-x
-  :parameters (?v1 ?v2)
-  :precondition
-    (and (or (not (pre_x-coord_v1_inc-x)) (x-coord ?v1))
-         (or (not (pre_x-coord_v2_inc-x)) (x-coord ?v2))
-         (or (not (pre_y-coord_v1_inc-x)) (x-coord ?v1))                       
-         (or (not (pre_y-coord_v2_inc-x)) (x-coord ?v2))
-         (or (not (pre_state-q0__inc-x)) (state-q0))
-         (or (not (pre_state-q1__inc-x)) (state-q1)))
-         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
-         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
-         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
-         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2)))         
-    )
-    :effect (and
-       (when (del_x-coord_v1_inc-x) (not (x-coord ?v1)))
-       (when (del_x-coord_v2_inc-x) (not (x-coord ?v2)))
-       (when (del_y-coord_v1_inc-x) (not (x-coord ?v1)))
-       (when (del_y-coord_v2_inc-x) (not (x-coord ?v2)))
-       (when (del_state-q0__inc-x) (not (state-q0)))
-       (when (del_state-q1__inc-x) (not (state-q1)))
-       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
-       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
-       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
-       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))
-
-       (when (add_x-coord_v1_inc-x) (x-coord ?v1))
-       (when (add_x-coord_v2_inc-x) (x-coord ?v2))
-       (when (add_y-coord_v1_inc-x) (x-coord ?v1))
-       (when (add_y-coord_v2_inc-x) (x-coord ?v2))
-       (when (add_state-q0__inc-x) (state-q0))
-       (when (add_state-q1__inc-x) (state-q1))
-       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
-       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
-       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
-       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
-  \end{verbatim}           
-  \end{tiny}  
- \caption{\small Programmable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
-\label{fig:programmable}
-\end{figure}
-
-\subsection{The space of \strips\ schema}
-The space of possible \strips\ schema is bound by these elements:
-\begin{enumerate}
-\item $\Psi$, is the set of {\em state predicates} that shape the propositional state variables of the model.
-\item $pars(\xi)$, the list of {\em parameters} of the corresponding schema $\xi$. 
-\end{enumerate}
-
-For each action schema $\xi$, given the set of {\em state predicates} $\Psi$, then $F_{\xi}$ defines the subset of elements that can appear in the preconditions and effects of that action schema.
-For the actions schema defined in~\ref{fig:model-example} this set is the same and contains the following ten elements, {\small\tt\{x-coord($v_1$), x-coord($v_2$), y-coord($v_1$), y-coord($v_2$), state-q0(), state-q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
-
-\begin{enumerate}
-\setcounter{enumi}{2}
-\item ${\mathcal C}$, a set of {\em syntactic constraints}.
-These constraints include the \strips\ constraints (we assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$) and domain-specific constraints.
-For instace, in a {\em robot navigation} domain like the modeled in Figure~\ref{fig:model-example}, predicates {\small\tt state-q0()} and {\small\tt state-q1()} are exclusive so they cannot hold a the same time.
-\end{enumerate}
-
-Predicates $\Psi_a$ also shape an additional set of objects ($\Omega\cap\Omega_v=\emptyset$) that we denote as {\em variable names} and that is bound by the maximum arity of the given action headers, i.e., $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{\xi} ar(\xi)}$.
-For instance, in a $5\times 5$ grid (like the one in Figure~\label{fig:grid-example} and modeled in Figure~\ref{fig:model-example}) $\Omega=\{o_1, o_2, o_3, o_4, o_5\}$ while $\Omega_v=\{v_1, v_2\}$ because all the actions schema have arity two.
-
-\begin{definition}[The space of \strips\ models]
-Given the set of {\em state predicates} $\Psi$, the set of {\em variable names} $\Omega_v$ and the set of {\em syntactic constraints} ${\mathcal C}$, the space of models for a given op predicate $p_\xi$ that represents the header of an action schema $\xi$ is given by three sets of inter
 \end{definition}
 
-The size of the space of possible \strips\ models for an action schema $\xi$ is given by the expression, $2^{2\times|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions).
+Let us define additional set of objects ($\Omega\cap\Omega_\xi=\emptyset$) that we denote as {\em variable names} and that contains one variable name for each paramter in $pars(\xi)$, that is $\Omega_\xi=\{v_i\}_{i=1}^{|pars(\xi)|}$.
+For instance, for all the \strips\ schema shown in Figure~\ref{fig:model-example} $\Omega_\xi=\{v_1,v_2\}$.
+Note that, for any of the actions schema defined in~\ref{fig:model-example}, the FOL interpretations of $\Psi$ (the {\em state predicates} that shape the propositional state variables) over the corresponding set of objects $\Omega_\xi$ contains the same ten elements, {\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+
+In addition, we can define constraints that bound the space of possible \strips\ schema:
+${\mathcal C}$
+\begin{itemize}
+\item {\em Syntactic constraints}.
+These constraints include the \strips\ constraints (we assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$).
+The size of the space of possible \strips\ schema is given by the expression, $2^{2\times|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions).
 For the mentioned navigation model, $2^{2\times|F_{\xi}|}=1,048,576$ for every action schema.
+\item {\em Domain-specific constraints}.
+For instace, in a {\em robot navigation} model, like the one in Figure~\ref{fig:model-example}, predicates {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold a the same time.
+Further, next($v_1$,$v_1$) and next($v_2$,$v_2$) is not appearing neither at the preconditions nor the effects of any possible \strips\ schema.
+This domain-specific knowledge can be used to constraint further the space of possible \strips\ schema.
+\end{itemize}
+
+Now we are ready to define the space of \strips\ schema that are possible for $\xi$.
+\begin{definition}[The space of possible \strips\ schema for $\xi$]
+The space of possible \strips\ schema for $\xi$ is the set ${\mathcal I}_{\Psi,\xi,{\mathcal C}}$ of the FOL interpretations of $\Psi$ (the {\em state predicates} that shape the propositional state variables) over the corresponding set of objects $\Omega_\xi$ and that satisfy the given shcema constraints.
+\end{definition}
 
@@
 
-\section{Model recognition as planning}
+\section{Model recognition with classical planning}
 This section shows that, when the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, then $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects.
@@
 $|F_v|=|F_{\tt stack}|=|F_{\tt unstack}|=11$ while $|F_{\tt pickup}|=|F_{\tt putdown}|=5$.
+
+\begin{figure}
+\begin{tiny}
+\begin{verbatim}
+;;; Propositional encoding for inc-x(?v1 ?v2)
+(pre_xcoord_v1_inc-x) (pre_next_v1_v2_inc-x) 
+(pre_q0__inc-x)
+(del_xcoord_v1_inc-x) (add_xcoord_v2_inc-x)
+
+;;; Propositional encoding for dec-x(?v1 ?v2)
+(pre_xcoord_v1_dec-x) (pre_next_v2_v1_dec-x) 
+(pre_q1___dec-x)
+(del_xcoord_v1_dec-x) (add_xcoord_v2_dec-x)
+
+;;; Propositional encoding for inc-y-even(?v1 ?v2)
+(pre_ycoord_v1_inc-y-even) (pre_next_v1_v2_inc-y-even)
+(pre_q0__inc-y-even)
+(del_ycoord_v1_inc-y-even) (del_q0__inc-y-even)
+(add_ycoord_v2_inc-y-even) (add_q1__inc-y-even)
+
+;;; Propositional encoding for inc-y-odd(?v1 ?v2)
+(pre_ycoord_v1_inc-y-odd) (pre_next_v1_v2_inc-y-odd) 
+(pre_q0__inc-y-odd)
+(del_ycoord_v1_inc-y-odd) (del_q1__inc-y-odd)
+(add_ycoord_v2_inc-y-odd) (add_q0__inc-y-odd)
+
+;;; Propositional encoding for dec-y-even(v1 ?v2)
+(pre_ycoord_v1_dec-y-even) (pre_next_v2_v1_dec-y-even)
+(pre_q0__dec-y-even)
+(del_ycoord_v1_dec-y-even) (del_q0__dec-y-even)
+(add_ycoord_v2_dec-y-even) (add_q1__dec-y-even)
+
+;;; Propositional encoding for dec-y-odd(?v1 ?v2)
+(pre_ycoord_v1_dec-y-odd) (pre_next_v2_v1_dec-y-odd)
+(pre_q1__dec-y-odd)
+(del_ycoord_v1_dec-y-odd) (del_q1__dec-y-odd)
+(add_ycoord_v2_dec-y-odd) (add_q0__dec-y-odd)
+\end{verbatim}
+\end{tiny}
+ \caption{\small Propositional encoding for the six schema from Figure~\ref{fig:model-example}.}
+\label{fig:encoding}
+\end{figure}
+
+\subsection{A propositional encoding for \strips\ action schema}
+We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
+For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
+Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
+
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents $[pre|del|add]\_p\_name(\xi)$.
+Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}.
+The interest of having a propositional encoding for \strips\ action schema is that it allows to define {\em editable actions} that is, actions whose semantics is given by the value of these particular fluents on the current state.
+
+Formally given an operator schema $\xi\in\mathcal{M}$ we can build its {\em editable} version as follows:
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{editable_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
+\cond(\mathsf{editable_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
+&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}.
+\end{align*}
+\end{small}
+bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
+Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
+Figure~\ref{fig:editable} shows the PDDL encoding of the action for applying a programmed operator.
+For instance, Figure~\ref{fig:editable} shows the editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (see Figure~\ref{fig:model-example}).
+Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} holds, behaves exactly as is defined in Figure~\ref{fig:model-example}.
+Further it allows to swap the semantics of two programable schemas provided that they are {\em comparable}.
+
+\begin{figure}
+  \begin{tiny}  
+  \begin{verbatim}
+(:action inc-x
+  :parameters (?v1 ?v2)
+  :precondition
+    (and (or (not (pre_xcoord_v1_inc-x)) (xcoord ?v1))
+         (or (not (pre_xcoord_v2_inc-x)) (xcoord ?v2))
+         (or (not (pre_ycoord_v1_inc-x)) (xcoord ?v1))                       
+         (or (not (pre_ycoord_v2_inc-x)) (xcoord ?v2))
+         (or (not (pre_q0__inc-x)) (q0))
+         (or (not (pre_q1__inc-x)) (q1)))
+         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
+         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
+         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
+         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2))))
+    :effect (and
+       (when (del_xcoord_v1_inc-x) (not (xcoord ?v1)))
+       (when (del_xcoord_v2_inc-x) (not (xcoord ?v2)))
+       (when (del_ycoord_v1_inc-x) (not (xcoord ?v1)))
+       (when (del_ycoord_v2_inc-x) (not (xcoord ?v2)))
+       (when (del_q0__inc-x) (not (q0)))
+       (when (del_q1__inc-x) (not (q1)))
+       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
+       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
+       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
+       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))
+
+       (when (add_xcoord_v1_inc-x) (xcoord ?v1))
+       (when (add_xcoord_v2_inc-x) (xcoord ?v2))
+       (when (add_ycoord_v1_inc-x) (xcoord ?v1))
+       (when (add_ycoord_v2_inc-x) (xcoord ?v2))
+       (when (add_q0__inc-x) (q0))
+       (when (add_q1__inc-x) (q1))
+       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
+       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
+       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
+       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
+  \end{verbatim}           
+  \end{tiny}  
+ \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
+\label{fig:editable}
+\end{figure}
 
@@
 Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
-Figure~\ref{fig:programmable} shows the PDDL encoding of the action for applying a programmed operator.
+Figure~\ref{fig:editable} shows the PDDL encoding of the action for applying a programmed operator.
 \begin{small}
@@
 \subsection{Recognition of {\em Regular Automatae}}
-We analyze now {\em model recognition} when the input set of given set of models represent different {\em regular automatae}.
-
+We analyze now {\em model recognition} when the given set of models represent different {\em regular automatae}.
 A {\em Regular automatae} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma\square,\delta}$:
@@
   :parameters (?x ?xr)
-  :precondition (and (head ?x) (symbol-a ?x) (state-$q_0$)
+  :precondition (and (head ?x) (symbol-a ?x) ($q_0$)
                      (next ?x ?xr))
   :effect (and (not (head ?x)) 
-               (not (symbol-a ?x)) (not (state-$q_0$))
-               (head ?xr) (state-$q_1$)))
+               (not (symbol-a ?x)) (not ($q_0$))
+               (head ?xr) ($q_1$)))
 \end{lstlisting}
@@
   :parameters (?xl ?x ?xr)
-  :precondition (and (head ?x) (symbol-a ?x) (state-$q_0$)
+  :precondition (and (head ?x) (symbol-a ?x) ($q_0$)
                      (next ?xl ?x) (next ?x ?xr))
   :effect (and (not (head ?x)) 
-               (not (symbol-a ?x)) (not (state-$q_0$))
-               (head ?xr) (symbol-x ?x) (state-$q_1$)))
+               (not (symbol-a ?x)) (not ($q_0$))
+               (head ?xr) (symbol-x ?x) ($q_1$)))
 \end{lstlisting}

### 2018-09-21 doc/icaps-19/model-recognition.tex
paper icaps19: sections 1,2, 3 and 4 stable. Ready for review

@@
 	  \path
-(A) edge [bend left] node {inc-y, dec-y} (B)
-(B) edge [bend left] node {inc-y, dec-y} (A);
+(A) edge [bend left, align=center] node {inc-y,\\dec-y} (B)
+(B) edge [bend left, align=center] node {inc-y,\\dec-y} (A);
 	\end{tikzpicture}
@@
 
-We first formalize \strips\ schemas and define the full space of possible \strips\ schema.
-Eventually, we introduce an {\em edit distance} which allows us to estimate the $P(\tau|\mathcal{M})$ likelihoods for observations of the execution of classical plans and planning \strips\ models.
+We start formalizing the \strips\ schema and define then the full space of possible \strips\ schema.
+Eventually, we introduce an {\em edit distance} for \strips\ schema which allows us to estimate the $P(\tau|\mathcal{M})$ likelihoods for planning models.
 
-\subsection{The space of \strips\ schema}
+\subsection{Well-defined \strips\ action schema}
 \strips\ action schema provide a compact representation for specifying classical planning models.
-For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema codded in PDDL for a robot navigation in a $n\times n$ grid.
+For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema, codded in PDDL, that determine a robot navigation in a $n\times n$ grid.
 
-\begin{definition}[A \strips\ action schema $\xi$]
-A \strips\ action schema $\xi$ with name, $name(\xi)$, defines a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-\end{definition}
+A \strips\ action schema $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from the schema.
 
-Let us define additional set of objects ($\Omega\cap\Omega_\xi=\emptyset$) that we denote as {\em variable names} and that contains one variable name for each paramter in $pars(\xi)$, that is $\Omega_\xi=\{v_i\}_{i=1}^{|pars(\xi)|}$.
-For instance, for all the \strips\ schema shown in Figure~\ref{fig:model-example} $\Omega_\xi=\{v_1,v_2\}$.
-Note that, for any of the actions schema defined in~\ref{fig:model-example}, the FOL interpretations of $\Psi$ (the {\em state predicates} that shape the propositional state variables) over the corresponding set of objects $\Omega_\xi$ contains the same ten elements, {\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+Given a \strips\ action schema $\xi$, let us define an additional set of objects ($\Omega\cap\Omega_\xi=\emptyset$), that we denote as {\em variable names}, and that contains one variable name for each parameter in $pars(\xi)$, that is $\Omega_\xi=\{v_i\}_{i=1}^{|pars(\xi)|}$.
+Note that, for any of the actions schema defined in~\ref{fig:model-example}, $\Omega_\xi=\{v_1,v_2\}$.
 
-In addition, we can define constraints that bound the space of possible \strips\ schema:
-${\mathcal C}$
+The set of FOL interpretations of $\Psi$ (the {\em predicates} that shape the propositional state variables) over the corresponding $\Omega_\xi$ objects (the {\em variable names} for that schema), confines the elements that can appear in the $\xi$ \strips\ action schema, that is in each of its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
+We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
+For any of the actions schema defined in~\ref{fig:model-example} this set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+
+Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of a schema $\xi$.
+The space of possible \strips\ schema for $\xi$ is bound further by a set of constraints ${\mathcal C}$ of the following three kinds:
 \begin{itemize}
 \item {\em Syntactic constraints}.
-These constraints include the \strips\ constraints (we assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$).
-The size of the space of possible \strips\ schema is given by the expression, $2^{2\times|F_{\xi}|}$ (\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions).
-For the mentioned navigation model, $2^{2\times|F_{\xi}|}=1,048,576$ for every action schema.
+\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions.
+Formally, $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$).
+Given exclusively these syntactic constraints, the size of the space of possible \strips\ schema is given by the expression, $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
+For the navigation model of Figure~\ref{fig:model-example}, $2^{2\times 10}=1,048,576$ for every action schema.
 \item {\em Domain-specific constraints}.
-For instace, in a {\em robot navigation} model, like the one in Figure~\ref{fig:model-example}, predicates {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold a the same time.
-Further, next($v_1$,$v_1$) and next($v_2$,$v_2$) is not appearing neither at the preconditions nor the effects of any possible \strips\ schema.
-This domain-specific knowledge can be used to constraint further the space of possible \strips\ schema.
+One can also introduce domain-specific knowledge to more precisely bound the space of possible \strips\ schema for a particular domain.
+For instance, in a {\em robot navigation} model, like the one in Figure~\ref{fig:model-example}, predicates {\small\tt q0()} and {\small\tt q1()} are always exclusive so they cannot hold a the same time in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
+Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate is coding the {\em succesor} function for {\em natural numbers}.
+In this case, introducing this domain-specific constraints would reduce the size of the space of possible schema to $2^{2\times 7}=16384$, for every action schema.
 \end{itemize}
 
-Now we are ready to define the space of \strips\ schema that are possible for $\xi$.
-\begin{definition}[The space of possible \strips\ schema for $\xi$]
-The space of possible \strips\ schema for $\xi$ is the set ${\mathcal I}_{\Psi,\xi,{\mathcal C}}$ of the FOL interpretations of $\Psi$ (the {\em state predicates} that shape the propositional state variables) over the corresponding set of objects $\Omega_\xi$ and that satisfy the given shcema constraints.
+Now we are ready to define what is a well-defined \strips\ action schema.
+\begin{definition}[Well-defined \strips\ action schema]
+Given a set of {\em predicates} $\Psi$, a list of action {\em parameters} $pars(\xi)$, and set of FOL constraints ${\mathcal C}$ we say that $\xi$ is a {\bf well-defined \strips\ action schema} iff its three lists $pre(\xi)\subseteq {\mathcal I}_{\Psi,\xi}$, $del(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ and $add(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ only contain elements in ${\mathcal I}_{\Psi,\xi}$ and they satisfy all the constraints in ${\mathcal C}$.
 \end{definition}
@@
 \subsection{The \strips\ edit distance}
-We define two edit \emph{operations} on a \strips\ model $\mathcal{M}\in M$:
+We define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
 \begin{itemize}
-\item {\em Deletion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is removed from the operator schema $\xi\in\mathcal{M}$, such that $f\in F_{\xi}$.
-\item {\em Insertion}. A fluent $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ is added to the operator schema $\xi\in\mathcal{M}$, s.t. $f\in F_{\xi}$.
+\item {\em Deletion}.
+An element in ${\mathcal I}_{\Psi,\xi}$ is removed from any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ such that the resulting schema is a {\em well-defined} \strips\ action schema.
+\item {\em Insertion}.
+An element in ${\mathcal I}_{\Psi,\xi}$ is added to any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ s.t.
+the resulting schema is {\em well-defined}.
 \end{itemize}
@@
 
-\begin{definition}
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} \strips\ action models.
+\begin{definition}[Edit distance]
+  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} \strips\ action models defined within the same set of predicates $\Psi$.
 The {\bf edit distance} $\delta(\mathcal{M},\mathcal{M}')$ is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
@@
 
-Since $F_v$ is a bound set, the maximum number of edits that can be introduced to a given action model defined within $F_v$ is bound as well. 
-\begin{definition}
-The \textbf{maximum edit distance} of an \strips\ model $\mathcal{M}$ built from the set of possible elements $F_v$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|F_{\xi}|$.
+Since ${\mathcal I}_{\Psi,\xi}$ are bound sets, the maximum number of edits that can be introduced to a given action model is bound as well. 
+\begin{definition}[Maximum edit distance]
+The \textbf{maximum edit distance} of an \strips\ model $\mathcal{M}$ built within the set of predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$.
 \end{definition}
 
+The observation of plan executions that are generated with $\xi$ action schema constrain further the space of possible schema.
+In this case we talk about {\em semantic constraints} that can also be added the $\mathcal{C}$ set.
 We define now an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.
 
-\begin{definition}
-  Given $\tau$, an observation of the execution of a plan for solving $P$ and $\mathcal{M}$, a \strips\ action model built from $F_v$.
+\begin{definition}[Observation edit distance]
+  Given $\tau$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, all defined within the same set of predicates $\Psi$.
 The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
@@
 
-The $\delta(\mathcal{M},\tau)$ distance could also be defined assessing the edition required by the observed plan execution to match the given model.
+Remarkably, the {\em observation edit distance} allows us to elicit the likelihood of the observations given a candidate model.
+It can be argued that the shorter this  distance the better the given model explains the given observation and hence, the higher the likelihood $P(\tau|\mathcal{M})$.
+In particular this distance is maximum when it requires fully editing all the schemas in the model while it is minimum when the given model can produce the input observation without introducing any change.
+Intermmediate distance values reflect how far are the models from explaining the input observations.
+In this work we map the {\em observation edit distance} into a likelihood with the following expression $P(\tau|\mathcal{M}) = 1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
+
+Note that the {\em observation edit distance} could also be defined assessing the edition required by the observed plan execution to match the given model.
 This implies defining {\em edit operations} that modify $\tau$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
-Our definition of the {\em observation edit distance} is more practical since normally $F_v$ is smaller than $F$.
-In practice, the number of {\em variable objects} is smaller than the number of objects in a planning problem.
-
-Estimating the conditional probability $P(\tau|\mathcal{M})$ by mapping the {\em observation edit distance} into a $[0,1]$ likelihood, $1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
+Our definition of the {\em observation edit distance} is more practical since normally ${\mathcal I}_{\Psi,\xi}$ is much smaller than $F$.
+In practice, the number of {\em variable objects} should be smaller than the number of objects in a planning problem.
 

### 2018-09-21 doc/icaps-19/model-recognition.tex
paper icaps19: pages [1-4] stable. Ready for review

@@
 
+The task of {\em model recognition} can also be understood as a classification task where each class is represented with a different planning model and the observed plan execution is the example to classify.
+In this case, the planning model that is associated to a class is acting as a class prototype that the summarizes all the plan executions that could be synthesized with that model or in other words, all the examples that belong to that class.
+
 In the {\em model recognition} task, hypotheses are about the possible action models $\mathcal{M}\in M$ while $\tau$ represents the given observation.
@@
 
-\subsection{The \strips\ edit distance}
+\subsection{Edit distances for the \strips\ planning model}
 We define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
@@
 \item {\em Deletion}.
-An element in ${\mathcal I}_{\Psi,\xi}$ is removed from any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ such that the resulting schema is a {\em well-defined} \strips\ action schema.
+An element is removed from any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ such that the resulting schema is a {\em well-defined} \strips\ action schema.
 \item {\em Insertion}.
@@
 
-The observation of plan executions that are generated with $\xi$ action schema constrain further the space of possible schema.
-In this case we talk about {\em semantic constraints} that can also be added the $\mathcal{C}$ set.
-We define now an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.
+The observation of plan executions, generated with an action model $\mathcal{M}$, is {\em semantic knowledge} that constrain further the space of possible schema $\xi\in \mathcal{M}$.
+In this case, we talk about {\em observation constraints} that can also be added to the $\mathcal{C}$ set.
+In addition, this new kind of model constraints allow us to define an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.
 
@@
 Remarkably, the {\em observation edit distance} allows us to elicit the likelihood of the observations given a candidate model.
-It can be argued that the shorter this  distance the better the given model explains the given observation and hence, the higher the likelihood $P(\tau|\mathcal{M})$.
-In particular this distance is maximum when it requires fully editing all the schemas in the model while it is minimum when the given model can produce the input observation without introducing any change.
-Intermmediate distance values reflect how far are the models from explaining the input observations.
-In this work we map the {\em observation edit distance} into a likelihood with the following expression $P(\tau|\mathcal{M}) = 1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
+It can be argued that the shorter this  distance the better the given model explains the given observation and hence, the higher the $P(\tau|\mathcal{M})$ likelihood.
+In particular this distance is maximum when it requires fully editing all the schemas in the model while it is minimum when the given model is already able to produce the input observation without introducing any change.
+Intermmediate distance values reflect how far models are from explaining the input observations.
+In this work we map the {\em observation edit distance} into a $P(\tau|\mathcal{M})$ likelihood with the following expression, $P(\tau|\mathcal{M}) = 1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
 
@@
 \label{sec:conclussions}
-The task of {\em model recognition} can also be understood as a classification task where each class is represented with a different planning model and the observed plan execution is the example to classify.
-With this regard, planning model that is associated to a class is acting as a class prototype that the summarizes all the plan executions that could be synthesized with that model or in other words, all the examples that belong to that class.
-
 In this work we do not assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.

### 2018-09-21 doc/icaps-19/model-recognition.tex
paper icaps19: pages [1-5] stable. Ready for review

@@
 
-The task of {\em model recognition} can also be understood as a classification task where each class is represented with a different planning model and the observed plan execution is the example to classify.
+The task of {\em model recognition} can be understood as a classification task where each class is represented with a different planning model and the observed plan execution is the example to classify.
 In this case, the planning model that is associated to a class is acting as a class prototype that the summarizes all the plan executions that could be synthesized with that model or in other words, all the examples that belong to that class.
@@
 A \strips\ action schema $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from the schema.
+
+We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
+For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
+Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
 
@@
 \section{Model recognition with classical planning}
-This section shows that, when the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, then $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects.
+This section shows that, for \strips\ planning models, the $P(\tau|\mathcal{M})$ likelihood can be estimated with a classical planning compilation.
+In more detail, when the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, then $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects.
+
 The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 
-Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,1]) for editing a given {\em blockswold} model where the positive effects {\tt\small (handempty)} and {\tt\small (clear ?v1)} of the {\tt\small stack} schema are missing.
-The postfix of the plan (steps [2,6]) validates the edited action model at the observation of a four action plan for inverting a two-block tower where intermediate states, $s_1$, $s_2$ and $s_3$, are unobserved.
+Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its action {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
+The postfix of the plan (steps [7,13]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
 \begin{figure}
 {\tt\scriptsize
-00 : (insert\_add\_handempty\_stack)\\
-01 : (insert\_add\_clear\_stack\_var1)\\
-02 : (apply\_unstack blockB blockA i1 i2)\\
-03 : (apply\_putdown blockB i2 i3)\\
-04 : (apply\_pickup blockA i3 i4)\\
-05 : (apply\_stack blockA blockB i4 i5)\\
-06 : (validate\_1)
+00 : (insert\_pre\_xcoord\_v1\_inc-x)\\
+01 : (insert\_pre\_next\_v1\_v2\_inc-x)\\
+02 : (insert\_pre\_q0\_inc-x)\\
+03 : (delete\_del\_xcoord\_v2\_inc-x)\\
+04 : (delete\_add\_xcoord\_v1\_inc-x)\\
+05 : (insert\_del\_xcoord\_v1\_inc-x)\\
+06 : (insert\_add\_xcoord\_v2\_inc-x)\\
+07 : (editable\_inc-x 1 2)\\
+08 : (editable\_inc-x 2 3)\\
+09 : (editable\_inc-x 3 4)\\
+10 : (editable\_inc-x 4 5)\\
+11 : (editable\_inc-y-even 1 2)\\
+12 : (editable\_dec-x 5 4)\\
+13 : (validate\_1)
 }
- \caption{\small Plan for editing (steps [0-1]) and validating (steps [2-6]) a given \strips\ planning model for the {\em blockswold}.}
+ \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-13]) the planning model of Figure\ref{fig:model-example} when action {\tt\small inc-x} has no preconditions and positive/negative effects are swapped wrt Figure\ref{fig:model-example}.}
 \label{fig:plan-pdistance}
@@
 Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deletions) required by $\mathcal{M}'$ to be validated.
-In the example of Figure~\ref{fig:plan-pdistance} $\delta(\mathcal{M},\tau)=2$ and $\delta(\mathcal{M},*)=3\times 2\times (11+5)$ since there are 4 action schemes ({\small\tt pickup}, {\small\tt putdown}, {\small\tt stack} and {\small\tt unstack}) s.t.
-$|F_v|=|F_{\tt stack}|=|F_{\tt unstack}|=11$ while $|F_{\tt pickup}|=|F_{\tt putdown}|=5$.
+In the example of Figure~\ref{fig:plan-pdistance} $\delta(\mathcal{M},\tau)=7$ and $\delta(\mathcal{M},*)=6\times 3\times 10$ since there are 6 action schemes for which $|{\mathcal I}_{\Psi,\xi}|=10$.
 
@@
 \subsection{A propositional encoding for \strips\ action schema}
-We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
-For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
-Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_name(\xi)$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding objects $\Omega_\xi$.
+Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}.
 
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents $[pre|del|add]\_p\_name(\xi)$.
-Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}.
 The interest of having a propositional encoding for \strips\ action schema is that it allows to define {\em editable actions} that is, actions whose semantics is given by the value of these particular fluents on the current state.
-
-Formally given an operator schema $\xi\in\mathcal{M}$ we can build its {\em editable} version as follows:
+Formally given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version can be compactly specified with {\em conditional effects}:
 \begin{small}
@@
 \end{small}
-bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
-Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
-Figure~\ref{fig:editable} shows the PDDL encoding of the action for applying a programmed operator.
-For instance, Figure~\ref{fig:editable} shows the editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (see Figure~\ref{fig:model-example}).
-Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} holds, behaves exactly as is defined in Figure~\ref{fig:model-example}.
-Further it allows to swap the semantics of two programable schemas provided that they are {\em comparable}.
+Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega\subseteq\Omega^{ar(\xi)}$ that appear at the same position.
+Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable version} of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (see Figure~\ref{fig:model-example}).
+Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
+Further, it allows to swap the semantics of any two schema provided that they are {\em comparable}.
 
@@
   \begin{verbatim}
-(:action inc-x
+(:action editable_inc-x
   :parameters (?v1 ?v2)

### 2018-09-21 doc/icaps-19/model-recognition.tex
paper icaps19: pages [1-5] stable. Ready for review

@@
 
+The compilation is an extension of the classical planning compilation for the learning of \strips\ planning models~\cite{aineto2018learning}.
 The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
@@
 \begin{figure}
-{\tt\scriptsize
+{\tt\footnotesize
 00 : (insert\_pre\_xcoord\_v1\_inc-x)\\
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{editable_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
-\cond(\mathsf{editable_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}.
+\hspace*{7pt}\pre(\mathsf{editable_{\xi,\omega}})=&\{pre\_e\_name(\xi)\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
+\cond(\mathsf{editable_{\xi,\omega}})=&\{del\_e\_name(\xi)\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+&\{add\_e\_name(\xi)\}\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
 \end{align*}
 \end{small}
-Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega\subseteq\Omega^{ar(\xi)}$ that appear at the same position.
+Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects that appear at the same position.
 Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable version} of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (see Figure~\ref{fig:model-example}).
@@
 \item The original fluents $F$. 
-\item Fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ modeling the space of \strips\ models. 
-\item The fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\tau$.
+\item Fluents $[pre|del|add]\_e\_name(\xi)$ to model the possible \strips\ schema. 
+\item The fluents to code the {\em observation constraints} and that are of these kinds:
+\begin{itemize}
+\item $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\tau$.
 The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of $\tau$.
 \item The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\tau$ where the action model is validated.
+\end{itemize}
 \item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schemas are edited or validated.
@@
 Our compilation assumes that initially $\mathcal{M}'$ is defined as $\mathcal{M}$.
-Therefore fluents $pre_f(\xi)/del_f(\xi)/add_f(\xi)$ hold as given by $\mathcal{M}$.
+Therefore fluents $[pre|del|add]\_e\_name(\xi)$ hold as given by $\mathcal{M}$.
 
@@
 \begin{enumerate}
+\item The {\em editable} version of the original actions in $A$.
+This actions have now  an extra preconditions  becuase they can only be applied in the {\em edition} mode (i.e.
+when $mode_{edit}$ holds).
+When the observation $\tau$ includes observed actions, they also include the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ to validate that actions are applied, exclusively, in the same order as they appear in $\tau$.\\
+
 \item Actions for {\em editing} operator schema $\xi\in\mathcal{M}$:
 \begin{itemize}
-\item Actions for adding a {\em precondition} $f\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$.
+\item Actions for inserting a into an action schema $\xi\in\mathcal{M}$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f,\xi}})=&\{\neg pre_{f}(\xi), \neg del_{f}(\xi),\neg add_{f}(\xi), mode_{edit}\},\\
-\cond(\mathsf{programPre_{f,\xi}})=&\{\emptyset\}\rhd\{pre_{f}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{programPre_{e,\xi}})=&\{\neg pre\_e\_name(\xi), \neg del\_e\_name(\xi),\\
+& \neg add\_e\_name(\xi), mode_{edit}\},\\
+\cond(\mathsf{programPre_{e,\xi}})=&\{\emptyset\}\rhd\{pre\_e\_name(\xi)\}.
 \end{align*}
@@
 
-\item Actions for adding a {\em negative} or {\em positive} effect $f\in F_v(\xi)$ to the action schema $\xi\in\mathcal{M}$.
+\item Actions for inserting adding a {\em negative} or {\em positive} effect into the action schema $\xi\in\mathcal{M}$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{f,\xi}})=&\{\neg del_{f}(\xi),\neg add_{f}(\xi), mode_{edit}\},\\
-\cond(\mathsf{programEff_{f,\xi}})=&\{pre_{f}(\xi)\}\rhd\{del_{f}(\xi)\},\\
-&\{\neg pre_{f}(\xi)\}\rhd\{add_{f}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{programEff_{e,\xi}})=&\{\neg del\_e\_name(\xi),\\
+& \neg add\_e\_name(\xi),mode_{edit}\},\\
+\cond(\mathsf{programEff_{e,\xi}})=&\{pre\_e\_name(\xi)\}\rhd\{del\_e\_name(\xi)\},\\
+&\{\neg pre\_e\_name(\xi)\}\rhd\{add\_e\_name(\xi)\}.
 \end{align*}
@@
 Besides these actions, the $A'$ set also contains the actions for {\em deleting} a precondition and a negative/positive effect.
-
-\item Actions for {\em applying} an edited operator schema $\xi\in\mathcal{M}$ bound with objects $\omega\subseteq\Omega^{ar(\xi)}$.
-Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear at the same position.
-Figure~\ref{fig:editable} shows the PDDL encoding of the action for applying a programmed operator.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\\
-&\cup \{\neg mode_{val}\},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-&\{mode_{edit}\}\rhd\{\neg mode_{edit}\},\\
-&\{\emptyset\}\rhd\{mode_{val}\}.
-\end{align*}
-\end{small}
-
-When the observation $\tau$ includes observed actions, then the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to validate that actions are applied, exclusively, in the same order as they appear in $\tau$.\\
 

### 2018-09-24 doc/icaps-19/model-recognition.tex
icaps19 paper evaluation section updated

@@
 \begin{definition}[Edit distance]
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} \strips\ action models defined within the same set of predicates $\Psi$.
+  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} and {\em well-defined} \strips\ action models defined within the same set of predicates $\Psi$.
 The {\bf edit distance} $\delta(\mathcal{M},\mathcal{M}')$ is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
@@
 
-\subsection{Evaluation}
+\section{Evaluation}
 \label{sec:evaluation}
-To evaluate the empirical performance of {\em model recognition as planning} we defined a set of possible \strips\ models, each representing a different {\em Turing Machine}, but all sharing the same set of {\em machine states} and same {\em tape alphabet}.
-
-\subsection{Experimental setup}
-We randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{100}\}$ set of one-hundred different {\em Turing Machines} where each $\mathcal{M}\in M$ is a seven-symbol six-state {\em Turing Machine}.
-The classical planning frame $\tup{F,A}$ is the same for all the {\em Turing Machines}, there is an $a\in A$ action for each pair of tape symbol and non-terminal state machine, while the $\rho$ and $\theta$ function is defined differently for each the machines (using a different \strips\ action model).
-
-We randomly choose a machine $\mathcal{M}\in M$ and produce the observation $\tau$ of a fifty-step execution plan.
-Finally, we follow our {\em model recognition as planning} method to identify the {\em Turing Machine} that produced $\tau$.
-This experiment is repeated for different amounts of missing information in the input trace $\tau$: unknown applied transitions, unknown internal machine state and unknown values of several tape cells.
+To evaluate the empirical performance of {\em model recognition as planning} we collect a set $M$ of possible \strips\ models, that share the same state variables but define different action models.
+Then, we randomly choose one of these models $\mathcal{M}\in M$ and use it to produce an observation $\tau$ of a plan execution.
+Finally, we follow our {\em model recognition as planning} method to identify the model $\mathcal{M}\in M$ that produced $\tau$.
+The experiment is repeated for models of different kind and different observability of the given plan execution.
 
@@
 \subsection{Recognition of {\em Regular Automatae}}
-We analyze now {\em model recognition} when the given set of models represent different {\em regular automatae}.
-A {\em Regular automatae} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma\square,\delta}$:
-\begin{itemize}
-\item $Q$ is a finite and non-empty set of machine states with the {\em initial state} $q_0\in Q$ and the {\em terminal states} $Q_{\bot}\subseteq Q$.  
-\item $\Sigma$ is the {\em input alphabet}, a finite non-empty set of symbols and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
-\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}.
-For each pair of tape symbol and non-terminal machine state $\delta$ defines: (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
-If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
-\end{itemize}
-
+In this experiment the models in $M$ represent different {\em regular automatae}.
 Figure~\ref{fig:regautomatae} illustrate a four-symbol four-state {\em regular automata} for recognizing the $\{(abc)^n : n \geq 1 \}$ language.
@@
 	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-	  \node[state] (A)              {$Q_0$};
-	  \node[state] (B) [right of=A] {$Q_1$};
-	  \node[state] (C) [right of=B] {$Q_2$};
-	  \node[state] (D) [right of=C] {$Q_3$};          
+	  \node[state] (A)              {$q_0$};
+	  \node[state] (B) [right of=A] {$q_1$};
+	  \node[state] (C) [right of=B] {$q_2$};
+	  \node[state] (D) [right of=C] {$q_3$};          
 	  \path
@@
   \end{scriptsize}  
- \caption{\small Four-symbol four-state {\em regular automata} for recognizing the $\{(abc)^n : n \geq 1 \}$ language.}
+ \caption{\small Four-symbol four-state {\em regular automata} for recognizing the $\{(abc)^n : n \geq 1 \}$ language (\underline{$q_3$} is the acceptor state).}
 \label{fig:regautomatae}
@@
 
-The \strips\ action schema of Figure~\ref{fig:reg-update-rule} models the rule $a,q_0\rightarrow q_1$ of the {\em regular automatae} defined in Figure~\ref{fig:regautomatae}.
-The full encoding of the {\em automata} of Figure~\ref{fig:regautomatae} produces a total of sixteen \strips\ action schema.
+In more detail, we randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{50}\}$ set of fifty different \strips\ models that encode different {\em regular automata}.
+Each $\mathcal{M}\in M$ encodes a different five-state {\em regular automata} with a five-symbol input alphabet.
+Each automata transition is encoded with a planning action.
+For instance, the execution of the {\em regular automatae} defined in Figure~\ref{fig:tautomatae}, with the sequence of input symbols $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
 
-\begin{figure}
-\begin{scriptsize}
-\begin{lstlisting}
-(:action transition-1      ;;; a,$q_0\rightarrow$ $q_1$
-  :parameters (?x ?xr)
-  :precondition (and (head ?x) (symbol-a ?x) ($q_0$)
-                     (next ?x ?xr))
-  :effect (and (not (head ?x)) 
-               (not (symbol-a ?x)) (not ($q_0$))
-               (head ?xr) ($q_1$)))
-\end{lstlisting}
-\end{scriptsize}
- \caption{\small \strips\ action schema that models the transition $q_0\rightarrow q_1$ of the automata defined in Figure~\ref{fig:regautomatae}.}
-\label{fig:regupdate-rule}
-\end{figure}
+Here we assume that the actual applied transitions are unknown as well as the internal machine state.
+Assuming that the actual applied transitions is unknown means that the observation $\tau$ of the execution of a Turing Machine contains no actions, it is simply a sequence of states $\tau=\tup{s_1, \ldots , s_m}$.
+Assuming that the internal machine state is unknown means that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt $q$()} fluents, with $q\in Q$ and $q\neq q_0$.
 
 \subsection{Recognition of {\em Turing Machines}}
-We analyze now {\em model recognition} when the input set of given set of models represent different {\em Turing machines}.
-
-A {\em Turing machine} is a tuple $\mathcal{M}=\tup{Q,q_o,Q_{\bot},\Sigma,\Upsilon,\square,\delta}$:
-\begin{itemize}
-\item $Q$ is a finite and non-empty set of machine states with the {\em initial state} $q_0\in Q$ and the {\em terminal states} $Q_{\bot}\subseteq Q$.  
-\item $\Sigma$ is the {\em tape alphabet}, a finite non-empty set of symbols with the {\em input alphabet} $\Upsilon\subseteq\Sigma$ (symbols allowed to initially appear in the tape) and the {\em blank symbol} $\square\in\Upsilon$ (the only symbol allowed to occur on the tape infinitely often).
-\item $\delta: \Sigma\times (Q\setminus Q_{\bot}) \rightarrow \Sigma\times Q\times\{left,right\}$ is the {\em transition function}.
-For each pair of tape symbol and non-terminal machine state $\delta$ defines: (1), the tape symbol to print at the current position of the header (2), the new state of the machine and (3), whether the header is shifted {\em left} or {\em right} after the print operation.
-If $\delta$ is not defined for the current pair of tape symbol and machine state, the machine halts.
-\end{itemize}
-
+In this experiment the given models in $M$ represent different {\em Turing machines}~\cite{bylander1994computational,porco2013automatic}.
 Figure~\ref{fig:tautomatae} illustrate a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
 The {\em tape alphabet} is $\Sigma=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} $\Upsilon=\{a,b,c,\square\}$ and the machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ (where \underline{$q_5$} is the only acceptor state).
+
+Here we randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{50}\}$ set of fifty different \strips\ models that encode different five-symbol five-state {\em Turing Machines} with circular tapes.
+There is a planning action $a\in A$ for each machine transition, e.g.
+the full encoding of the {\em Turing Machine} of Figure~\ref{fig:tautomatae} produces a total of sixteen \strips\ action schema.
+With this regard, the execution of the {\em Turing Machine} defined in Figure~\ref{fig:tautomatae} with an initial tape $abc\square\square\square$ produces the eight-action plan {\small $(a,q_0\rightarrow x,r,q_1)$, $(b,q_1\rightarrow y,r,q_2)$, $(c,q_2\rightarrow z,l,q_3)$, $(y,q_3\rightarrow y,l,q_3)$, $(x,q_3\rightarrow x,r,q_0)$, $(y,q_0\rightarrow y,r,q_4)$, $(z,q_4\rightarrow z,r,q_4)$, $(\square,q_4\rightarrow \square,r,\underline{q_5})$}.
+
+Again we assume that the actual applied transitions is unknown ($\tau$ contains no actions).
+Further, we assume that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt $q$()} fluents, with $q\in Q$ and $q\neq q_0$ and that the values of several tape cells is unknown.
 
@@
 
-The \strips\ action schema of Figure~\ref{fig:update-rule} models the rule $a,q_0\rightarrow x,r,q_1$ of the {\em Turing Machine} defined in Figure~\ref{fig:tautomatae}.
-The full encoding of the {\em Turing Machine} of Figure~\ref{fig:tautomatae} produces a total of sixteen \strips\ action schema.
+\subsection{Recognition of navigation models}
+In this experiment the given models in $M$ represent different navigation models that are computed as the cross product of a regular automata with a four-operator \strips\ model for navigating a $n\times n$ grid.
 
-\begin{figure}
-\begin{scriptsize}
-\begin{lstlisting}
-(:action transition-1      ;;; a,$q_0\rightarrow$ x,r,$q_1$
-  :parameters (?xl ?x ?xr)
-  :precondition (and (head ?x) (symbol-a ?x) ($q_0$)
-                     (next ?xl ?x) (next ?x ?xr))
-  :effect (and (not (head ?x)) 
-               (not (symbol-a ?x)) (not ($q_0$))
-               (head ?xr) (symbol-x ?x) ($q_1$)))
-\end{lstlisting}
-\end{scriptsize}
- \caption{\small \strips\ action schema that models the transition $a,q_0\rightarrow x,r,q_1$ of the Turing Machine defined in Figure~\ref{fig:tautomatae}.}
-\label{fig:update-rule}
-\end{figure}
-
-With regard to our \strips\ model for {\em Turing Machines}, executions of a {\em Turing Machine} are definable as an action sequence $\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $a_i$ ({\small $1\leq i\leq n$}) is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-For instance, the execution of the {\em Turing Machine} defined in Figure~\ref{fig:tautomatae} with an initial tape $abc\square\square\square\ldots$ produces the eight-action plan {\small $(a,q_0\rightarrow x,r,q_1)$, $(b,q_1\rightarrow y,r,q_2)$, $(c,q_2\rightarrow z,l,q_3)$, $(y,q_3\rightarrow y,l,q_3)$, $(x,q_3\rightarrow x,r,q_0)$, $(y,q_0\rightarrow y,r,q_4)$, $(z,q_4\rightarrow z,r,q_4)$, $(\square,q_4\rightarrow \square,r,\underline{q_5})$}.
-
-Assuming that the actual applied transitions is unknown means that the observation $\tau$ of the execution of a Turing Machine contains no actions, it is simply a sequence of states $\tau=\tup{s_1, \ldots , s_m}$.
-Further, assuming that the internal machine state is unknown means that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt (state-$q$)} fluents, with $q\in Q$ and $q\neq q_0$.
-Finally, assuming that the values of several tape cells is unknown means that fluents of the kind {\small\tt (symbol-$\sigma$ ?x)} are missing (i.e.
-unobserved) for some state $s_i\in \tau$ s.t.
-$1\leq i\leq m$.
-These facts affects to the a priori probaility of the possible observations.
-
-On the other hand, the model edition for Turing Machines is limited to these subset of possible positive effects: {\tt (head ?xr)} or {\tt (head ?xl)}, {\tt(symbol-$\sigma$ ?x)} with $\sigma \in \Sigma$ and last but not least, {\tt (state-q)} with $q\in Q$.
-No other positive effects, preconditions, or negative effects are required to be edited.
-This fact reduces the possible edition operations and affects to the a priori probaility of the possible action models.
+In this case the regular automata constrain the applications of the navigation actions producing different navigation policies e.g.
+like the one in Figure~\ref{fig:model-example}.
+Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, here the task is to identify which navigation model produced that observation, despite the the applied actions are unobserved.
+In addition, for each observed state, only the value of fluents encoding the x and y coordinates of the agent are known while the value of the regular automata conditioning the navigation policy is unknown.
 
 \subsection{Results}
+
+\subsection{Related Work}
+Plan recognition.
+
+Instace Based Classification.
+
+Model Reconciliation.
 

### 2018-09-24 doc/icaps-19/model-recognition.tex
new stable version icaps19 paper. Ready for a second review

@@
 \begin{abstract} 
-Given the partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set explains (produced) the given observation.
-The paper formalizes the {\em model recognition} task and proposes a novel method to assess the probability of a \strips\ model to produce a partial observation of a plan execution.
-This method, that we called {\em model recognition as planning}, is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
+Given the partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models to update these variables), {\em model recognition} is the task of identifying which model in the set explains (produced) the given observation.
+The paper formalizes the {\em model recognition} task and proposes a novel method to estimate the probability of a \strips\ model to produce a partial observation of a plan execution.
+This method, that we called {\em model recognition as planning}, is built on top of off-the-shelf classical planning algorithms and elicit the likelihood of the observation of a plan execution given a candidate model.
 {\em Model recognition as planning} is robust to missing intermediate states and actions in the observed plan execution.
@@
 
-This paper introduces the {\em model recognition} task where the target of the recognition process is not a goal but a {\em planning model}.
-Given a partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models), {\em model recognition} is the task of identifying which model in the set has the highest probability of producing the given observation.
+This paper introduces the {\em model recognition} task where the object to recognize is not a goal but a {\em planning model}.
+Given a partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models to update these variables), {\em model recognition} is the task of identifying which model in the set has the highest probability of explaining (producing) the given observation.
 
@@
   \end{scriptsize}  
- \caption{\small Observation of the execution of a robot navigation plan in a $5\times 5$ grid and a two state automata for controlling that the robot only increments its x-coordinate at {\em even} rows (i.e.
-at state $q_0$).}
+ \caption{\small Observation of the execution of a robot navigation plan in a $5\times 5$ grid and a 2-state automata for controlling that the robot only increments its x-coordinate at {\em even} rows (i.e.
+when $q_0$ holds).}
 \label{fig:grid-example}
@@
 To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model the robot could increment its {\em x-coordinate} at {\em even} rows and decrement it at the {\em odd rows}.
-Apart from this particular navigation model, different models could be defined within the same state variables (e.g.
-altering the way {\small\tt (q0)} and {\small\tt (q1)} are required and updated) and these models can determine different kinds of robot navigation.
+According to this model the robot could increment its {\em x-coordinate} at {\em even} rows (when {\tt\small $q0$} holds) and decrement it at the {\em odd rows} (when {\tt\small $q1$} holds).
+Apart from this particular navigation model, different action models could be defined within the same state variables (e.g.
+altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated) and these models can determine different kinds of robot navigation.
 Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} aims here to identify which navigation model produced that observation, despite key information is unobserved (e.g.
-the value of {\small\tt (q0)} and {\small\tt (q1)} or the particular applied actions).
+the particular applied actions or the value of {\small\tt $q0$} and {\small\tt $q1$}).
 
@@
 
-The paper introduces also {\em model recognition as planning}; a novel method to assess the probability of a given \strips\ model to produce an observed plan execution.
+The paper introduces also {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
 Our method is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
 {\em Model recognition as planning} is robust to missing intermediate states and actions in the observed plan execution.
+
 We evaluate the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em finite state machines}.
-All of these {\em automatae} are defined within the same {\em alphabet} and same {\em machine states} but different {\em transition functions}.
+All of these {\em automatae} are defined within the same input {\em alphabet} and same {\em machine states} but different {\em transition functions}.
 We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite internal machine states or actual applied transitions are unobserved.
@@
   \end{tiny}  
- \caption{\small Action model for a robot navigation in a $n\times n$ grid.}
+ \caption{\small Example of a \strips\ action model (codded in PDDL) for robot navigation in a $n\times n$ grid.}
 \label{fig:model-example}
@@
 \label{sec:background}
-This section formalizes the models for {\em classical planning} and for the {\em observations} of a plan execution.
+This section formalizes the models for {\em classical planning} and for the {\em observation} of a plan execution.
 
@@
 A {\em classical planning frame} is a pair $\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose semantics are specified with two functions: $\rho(s,a)$ that denotes whether an action $a\in A$ is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
-In this work we specify action semantics with the \strips\ model.
+In this work we specify action semantics (the $\rho$ and $\theta$ functions) with the \strips\ model.
 With this regard, an action $a\in A$ is defined with:
@@
 Formally, $0\leq|s_i^o|$ for every $1\leq i\leq m$.
-This means that the transitions between two consecutive observed states in $\tau$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where $k\geq 1$ is unknown and unbound).
+This means that the transitions between two consecutive observed states in $\tau$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown and unbound).
 Therefore we can conclude that having $\tau$ does not implies knowing the actual length of $\pi$.
@@
 \begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\tau$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every $0\leq i\leq m$, each observed state $s_i^o$ only contains fluents in $\Phi$.
+Given a subset of fluents $\Phi\subseteq F$ we say that $\tau$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 0\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
 \end{definition}
 
-For instance, Figure~\ref{fig:grid-example} illustrates the $\Phi$-observation \{{\tt\scriptsize<(xcoord 1)(ycoord 1)>, <(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\} where $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
-This means that, for each observed state, only the value of fluents {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} is known while the value of the fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)} is unknown.
+For instance, Figure~\ref{fig:grid-example} illustrates the six-state $\Phi$-observation \{{\tt\footnotesize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\} where $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
+This means that, for each observed state, only the value of fluents {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} is known while the value of the remaining fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)} is unknown.
 
@@
 \begin{itemize}
-\item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} such that the semantics of the actions in $A$ is unknown because functions $\rho(s,a)$ and/or $\theta(s,a)$ are undefined.
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty {\em set of models} for the actions in $A$ such that, each model in $\mathcal{M}\in M$, defines a different function pair $\tup{\rho,\theta}$.
+\item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} s.t. the semantics of actions $a\in A$ is unknown because the corresponding functions, $\rho(s,a)$ and/or $\theta(s,a)$, are undefined.
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty {\em set of models} for the actions in $A$ s.t.
+each model in $\mathcal{M}\in M$, defines a different function pair $\tup{\rho,\theta}$.
 \item $\tau$ is an {\em observation} of the execution of a solution plan $\pi$ for $P$.
@@
 
-The task of {\em model recognition} can be understood as a classification task where each class is represented with a different planning model and the observed plan execution is the example to classify.
-In this case, the planning model that is associated to a class is acting as a class prototype that the summarizes all the plan executions that could be synthesized with that model or in other words, all the examples that belong to that class.
+{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model and the observed plan execution is the single example to classify.
+In this case, the planning model that is associated to a class is acting as the {\em class prototype} and summarizes all the plan executions that could be synthesized with that model (all the examples that belong to that class).
 
-In the {\em model recognition} task, hypotheses are about the possible action models $\mathcal{M}\in M$ while $\tau$ represents the given observation.
-The {\em discrete probability distribution} $P(\mathcal{M}|\tau)$ expresses, for each model $\mathcal{M}\in M$, its probability of producing the input observation.
-According to the {\em Bayes rule} this probability distribution is given by $P(\mathcal{M}|\tau)=\frac{P(\tau|\mathcal{M})P(\mathcal{M})}{P(\tau)}$ and can be estimated in these three steps:
-\begin{enumerate}
-\item Estimating the {\em a priori probabilities} (if they are not already given as input).
-$P(\tau)$, that measures how surprising is the given observation and $P(\mathcal{M})$, that expresses if one model is known to be a priori more likely than the others.
-\item Computing likelihoods $P(\tau|\mathcal{M})$.
-Our approach is to estimate these likelihoods according to the {\em amount of edits} required by the model $\mathcal{M}$ to produce a plan $\pi$ such that:
-\begin{enumerate}
-\item $\pi$ solves $P$ and,
-\item $\tau$ is a consistent observation of the execution of $\pi$ on the classical planning problem $P$. 
-\end{enumerate}
-The {\em edit distance} is a similarity metric that is traditionally computed over {\em strings} or {\em graphs} and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
-In this work this similarity metric refers to the edition of classical planning models.
-\item Applying the {\em Bayes rule} to obtain the normalized posterior probabilities. The $P(\mathcal{M}|\tau)$ probabilities must sum 1 for all the $\mathcal{M}\in M$.
-\end{enumerate}
-
+In {\em model recognition}, hypotheses are then about the possible action models $\mathcal{M}\in M$ while the observation $\tau$ of a plan execution represents the input observation.
 The {\em naive Bayes classifier} assigns a model $\mathcal{M}\in M$ to the given observation $\tau$ wrt the following expression, $argmax_{\mathcal{M}\in M} P(\mathcal{M})\times P(\tau|\mathcal{M})$.
@@
 
+The $P(\mathcal{M})$ probability expresses if one model is known to be a priori more likely than the others.
+If this probability is not given as input it is reasonable to assume that {\em a priori} all models are equiprobable.
+Our approach for {\em model recognition} is then to estimate the $P(\tau|\mathcal{M})$ likelihood according to the {\em amount of edits} required by the model $\mathcal{M}$ to produce a plan $\pi$ such that:
+\begin{enumerate}
+\item $\pi$ solves $P$ and,
+\item $\tau$ is a {\em consistent} observation of the execution of $\pi$ on the classical planning problem $P$. 
+\end{enumerate}
+The {\em edit distance} is a similarity metric that is traditionally computed over {\em strings} or {\em graphs} and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
+In this work we show how to compute this similarity metric referred to the edition of classical planning models.
+
 \section{Recognition of \strips\ models}
 \label{sec:asPlanning}
-Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of actions (in other words, the corresponding $\rho$ and $\theta$ functions) are specified with \strips\ action schemas.
+Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of actions (i.e. $\rho$ and $\theta$ functions) are specified with \strips\ action schemas.
 
-We start formalizing the \strips\ schema and define then the full space of possible \strips\ schema.
-Eventually, we introduce an {\em edit distance} for \strips\ schema which allows us to estimate the $P(\tau|\mathcal{M})$ likelihoods for planning models.
+We start formalizing the \strips\ schema and define the full space of possible \strips\ schema.
+Eventually, we introduce an {\em edit distance} for \strips\ schema to estimate the $P(\tau|\mathcal{M})$ likelihoods for classical planning models.
 
@@
 
-A \strips\ action schema $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and the three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from the schema.
+A \strips\ action schema $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 
@@
 Given a \strips\ action schema $\xi$, let us define an additional set of objects ($\Omega\cap\Omega_\xi=\emptyset$), that we denote as {\em variable names}, and that contains one variable name for each parameter in $pars(\xi)$, that is $\Omega_\xi=\{v_i\}_{i=1}^{|pars(\xi)|}$.
-Note that, for any of the actions schema defined in~\ref{fig:model-example}, $\Omega_\xi=\{v_1,v_2\}$.
+Note that, for any of the six schema defined in Figure~\ref{fig:model-example}, $\Omega_\xi=\{v_1,v_2\}$.
 
-The set of FOL interpretations of $\Psi$ (the {\em predicates} that shape the propositional state variables) over the corresponding $\Omega_\xi$ objects (the {\em variable names} for that schema), confines the elements that can appear in the $\xi$ \strips\ action schema, that is in each of its $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
+Given a \strips\ action schema $\xi$ and a set of {\em predicates} $\Psi$ that shape the propositional state variables.
+The set of FOL interpretations of $\Psi$ over the corresponding $\Omega_\xi$ objects (the {\em variable names} for schema $\xi$), confines the elements that can appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
 We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
-For any of the actions schema defined in~\ref{fig:model-example} this set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+For any of the six schema defined in Figure~\ref{fig:model-example} the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
 
 Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of a schema $\xi$.
-The space of possible \strips\ schema for $\xi$ is bound further by a set of constraints ${\mathcal C}$ of the following three kinds:
+The space of possible \strips\ schema is bound further by a set of constraints ${\mathcal C}$ of the following kinds:
 \begin{itemize}
@@
 \strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions.
-Formally, $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$).
-Given exclusively these syntactic constraints, the size of the space of possible \strips\ schema is given by the expression, $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
+Formally, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$).
+Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schema is given by the expression, $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
 For the navigation model of Figure~\ref{fig:model-example}, $2^{2\times 10}=1,048,576$ for every action schema.
@@
 One can also introduce domain-specific knowledge to more precisely bound the space of possible \strips\ schema for a particular domain.
-For instance, in a {\em robot navigation} model, like the one in Figure~\ref{fig:model-example}, predicates {\small\tt q0()} and {\small\tt q1()} are always exclusive so they cannot hold a the same time in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
+For instance, in a {\em robot navigation} model, like the one in Figure~\ref{fig:model-example}, predicates {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold a the same time in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
 Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate is coding the {\em succesor} function for {\em natural numbers}.
-In this case, introducing this domain-specific constraints would reduce the size of the space of possible schema to $2^{2\times 7}=16384$, for every action schema.
+In this case, introducing these domain-specific constraints would reduce the size of the space of possible schema to $2^{2\times 7}=16,384$ for every action schema.
 \end{itemize}
@@
 \end{definition}
+We say that an action model is well-defined if all its \strips\ action schema are well-defined.
 
-\subsection{Edit distances for the \strips\ planning model}
+\subsection{Edit distances for \strips\ action models}
 We define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
@@
 
-The observation of plan executions, generated with an action model $\mathcal{M}$, is {\em semantic knowledge} that constrain further the space of possible schema $\xi\in \mathcal{M}$.
+The observation of a plan execution, generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrain further the space of the possible schema $\xi\in \mathcal{M}$.
 In this case, we talk about {\em observation constraints} that can also be added to the $\mathcal{C}$ set.
@@
   Given $\tau$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, all defined within the same set of predicates $\Psi$.
-The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a plan $\pi^*_\tau$ optimal for $P$ and compliant with $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
+The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
+$\mathcal{M}'$ produces a plan $\pi_\tau$ for $P$ {\em consistent} with $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
@@
 \section{Model recognition with classical planning}
-This section shows that, for \strips\ planning models, the $P(\tau|\mathcal{M})$ likelihood can be estimated with a classical planning compilation.
-In more detail, when the dynamics of the actions $A\in P$ are specified with \strips\ action schemas, then $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects.
+This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects (and hence the $P(\tau|\mathcal{M})$ likelihood can be estimated).
 
@@
 \begin{enumerate}
-\item Induces an optimal solution plan $\pi^*_\tau$ for the original classical planning problem $P$.
-\item Validates that $\tau$ is an observation of the execution of $\pi^*_\tau$ on the classical planning problem $P$.
+\item Induces a solution plan $\pi_\tau$ for the original classical planning problem $P$.
+\item Validates that $\tau$ is an observation of the execution of $\pi_\tau$ on the classical planning problem $P$.
 \end{enumerate}
@@
 
-Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its action {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
-The postfix of the plan (steps [7,13]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
+Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
+The postfix of the plan (steps [7,14]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
 \begin{figure}
@@
 06 : (insert\_add\_xcoord\_v2\_inc-x)\\
-07 : (editable\_inc-x 1 2)\\
-08 : (editable\_inc-x 2 3)\\
-09 : (editable\_inc-x 3 4)\\
-10 : (editable\_inc-x 4 5)\\
-11 : (editable\_inc-y-even 1 2)\\
-12 : (editable\_dec-x 5 4)\\
-13 : (validate\_1)
+07 : (validate\_0)
+08 : (editable\_inc-x 1 2)\\
+09 : (editable\_inc-x 2 3)\\
+10 : (editable\_inc-x 3 4)\\
+11 : (editable\_inc-x 4 5)\\
+12 : (editable\_inc-y-even 1 2)\\
+13 : (editable\_dec-x 5 4)\\
+14 : (validate\_1)
 }
@@
 \subsection{A propositional encoding for \strips\ action schema}
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_name(\xi)$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding objects $\Omega_\xi$.
-Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema previously defined in Figure~\ref{fig:model-example}.
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding objects $\Omega_\xi$.
+Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.
 
-The interest of having a propositional encoding for \strips\ action schema is that it allows to define {\em editable actions} that is, actions whose semantics is given by the value of these particular fluents on the current state.
-Formally given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version can be compactly specified with {\em conditional effects}:
+The interest of having a propositional encoding for \strips\ action schema is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}.
+Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents on the current state.
+Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
 \begin{small}  
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{editable_{\xi,\omega}})=&\{pre\_e\_name(\xi)\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
-\cond(\mathsf{editable_{\xi,\omega}})=&\{del\_e\_name(\xi)\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
-&\{add\_e\_name(\xi)\}\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
+\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
+\cond(\mathsf{editable_{\xi}})=&\{del\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+&\{add\_e\_\xi\}\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
 \end{align*}
 \end{small}
-Since operators headers are given as input, the variables $pars(\xi)$ are bound to the objects that appear at the same position.
+
 Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable version} of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (see Figure~\ref{fig:model-example}).
 Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
-Further, it allows to swap the semantics of any two schema provided that they are {\em comparable}.
 
@@
 \subsection{The compilation formalization}
-Conditional effects allow us to compactly define our compilation.
-Given a \strips\ model $\mathcal{M}\in M$ and the observation $\tau$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditionl effects $P'=\tup{F',A',I',G'}$ such that:
+Conditional effects allow us to compactly define our compilation for computing $\delta(\mathcal{M},\tau)$ and hence, estimate the $P(\tau|\mathcal{M})$ likelihood.
+Given a \strips\ model $\mathcal{M}\in M$ and the observation $\tau$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditional effects $P'=\tup{F',A',I',G'}$ such that:
 \begin{itemize}
-\item $F'$ contains:
+\item $F'$ extends the original fluents $F$ with:
 \begin{itemize}
-\item The original fluents $F$. 
-\item Fluents $[pre|del|add]\_e\_name(\xi)$ to model the possible \strips\ schema. 
+\item Fluents $[pre|del|add]\_e\_\xi$ to model the possible \strips\ schema. 
 \item The fluents to code the {\em observation constraints} and that are of these kinds:
@@
 Our compilation assumes that initially $\mathcal{M}'$ is defined as $\mathcal{M}$.
-Therefore fluents $[pre|del|add]\_e\_name(\xi)$ hold as given by $\mathcal{M}$.
+Therefore fluents $[pre|del|add]\_e\_\xi$ hold as given by $\mathcal{M}$.
 
@@
 \item The {\em editable} version of the original actions in $A$.
-This actions have now  an extra preconditions  becuase they can only be applied in the {\em edition} mode (i.e.
-when $mode_{edit}$ holds).
+This actions have now an extra preconditions because they can only be applied in the {\em validation} mode (i.e.
+when $mode_{val}$ holds).
 When the observation $\tau$ includes observed actions, they also include the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ to validate that actions are applied, exclusively, in the same order as they appear in $\tau$.\\
@@
 \begin{itemize}
-\item Actions for inserting a into an action schema $\xi\in\mathcal{M}$.
+\item Actions for inserting a new precondition into an action schema $\xi\in\mathcal{M}$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{e,\xi}})=&\{\neg pre\_e\_name(\xi), \neg del\_e\_name(\xi),\\
-& \neg add\_e\_name(\xi), mode_{edit}\},\\
-\cond(\mathsf{programPre_{e,\xi}})=&\{\emptyset\}\rhd\{pre\_e\_name(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{programPre_{e,\xi}})=&\{\neg pre\_e\_\xi, \neg del\_e\_\xi,\\
+& \neg add\_e\_\xi, mode_{edit}\},\\
+\cond(\mathsf{programPre_{e,\xi}})=&\{\emptyset\}\rhd\{pre\_e\_\xi\}.
 \end{align*}
@@
 
-\item Actions for inserting adding a {\em negative} or {\em positive} effect into the action schema $\xi\in\mathcal{M}$.
+\item Actions for inserting a new {\em negative} or {\em positive} effect into the action schema $\xi\in\mathcal{M}$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{e,\xi}})=&\{\neg del\_e\_name(\xi),\\
-& \neg add\_e\_name(\xi),mode_{edit}\},\\
-\cond(\mathsf{programEff_{e,\xi}})=&\{pre\_e\_name(\xi)\}\rhd\{del\_e\_name(\xi)\},\\
-&\{\neg pre\_e\_name(\xi)\}\rhd\{add\_e\_name(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{programEff_{e,\xi}})=&\{\neg del\_e\_\xi,\\
+& \neg add\_e\_\xi,mode_{edit}\},\\
+\cond(\mathsf{programEff_{e,\xi}})=&\{pre\_e\_\xi\}\rhd\{del\_e\_\xi\},\\
+&\{\neg pre\_e\_\xi\}\rhd\{add\_e\_\xi\}.
 \end{align*}
@@
 
-\item Actions for {\em validating} the partially observed state $s_j\in\tau$, {\tt\small $1\leq j< m$}.
+\item Actions for {\em validating} the edited models at the $s_j$ observed states, {\tt\small $0\leq j< m$}.
 \begin{small}
@@
 \hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\neg mode_{val}\}.
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\\
+                            &\{mode_{edit}\}\rhd\{\neg mode_{edit}, mode_{val}\}.
 \end{align*}
@@
 Here we assume that the actual applied transitions are unknown as well as the internal machine state.
-Assuming that the actual applied transitions is unknown means that the observation $\tau$ of the execution of a Turing Machine contains no actions, it is simply a sequence of states $\tau=\tup{s_1, \ldots , s_m}$.
-Assuming that the internal machine state is unknown means that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt $q$()} fluents, with $q\in Q$ and $q\neq q_0$.
+Assuming that the actual applied transitions is unknown means that the observation $\tau$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\tau=\tup{s_1, \ldots , s_m}$.
+Assuming that the internal machine state is unknown means that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
 
@@
 
-\subsection{Related Work}
+\section{Related Work}
+\label{sec:related}
 Plan recognition.
+
+Model Reconciliation.
 
@@
 
-Model Reconciliation.
-
-\subsection{Conclusions}
+\section{Conclusions}
 \label{sec:conclussions}

### 2018-09-26 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updates to conselleria research project. One year budget

@@
 Since then, two yearly international competitions are established ({\small\tt\url{http://www.sygus.org}}), to compare the different approaches for program synthesis, and the sister competition ({\small\tt\url{http://www.syntcomp.org/}}), to compare different approaches for the synthesize of reactive programs.
-Program synthesis has already been deployed in the real world and it is part of the {\sc Flash Fill} feature of {\sc Microsoft Excel} that automatically generates programs for string transformation~\cite{gulwani2011automating}.
+Further, program synthesis has already been deployed in the real world and it is part of the {\sc Flash Fill} feature of {\sc Microsoft Excel} that automatically generates programs for string transformation~\cite{gulwani2011automating}.
 
@@
 Program validation is also compilable into classical planning.
-Examples are the compilations for {\em {\sc GOLOG} procedures}~\cite{baier2007exploiting}, {\em planning programs}~\cite{segovia:programs:AIJ19}, or {\em Finite State Controllers}~\cite{Geffner:FSM:AAAI10,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018}.
+Examples are the compilations for {\em GOLOG procedures}, {\em planning programs}, {\em reactive policies}, or {\em Finite State Controllers}~\cite{baier2007exploiting,Geffner:FSM:AAAI10,ivankovic2015optimal,sergio:aprograming:ijcai16,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
 Briefly these compilations encode the {\em cross product} of a given planning instance and the automata corresponding to the program to validate.
-A {\em validation proof} is provided if a solution to the compiled instance is found (i.e.
+A {\em validation proof} is provided if a solution to the compiled planning instance is found (i.e.
 the program is {\em correct}).
-Otherwise, if a classical planner proves that no solution exists for the compiled instance, then the program is {\em incorrect} because its execution necessarily failed.
+Otherwise, if an AI planner proves that no solution exists for the compiled instance, then the program is {\em incorrect} because its execution necessarily failed.
 When actions have non-deterministic effects, program validation becomes more complex since it requires proving that all the possible program executions reach the goals.
@@
 \subsection{Background}
-First we briefly introduce the technology required by our AI method for the automated synthesis and validation of programs:
+We introduce the technology required by our AI method for the automated synthesis and validation of programs:
+
+\subsubsection{AI Planning}
+{\em AI Planning (AIP)} is the Artificial Intelligence component that studies the synthesis of sets of actions to achieve some given objectives~\cite{ghallab2004automated}.
+AIP arose in the late ’50s from converging studies into {\em combinatorial search}, {\em theorem proving} and {\em control theory} and now, is a well formalized paradigm for problem solving with algorithms that scale-up reasonably well.
+
+An {\em AI planning problem} is formalized as a tuple $\tup{V,D,A,I,G}$ where:
 \begin{itemize}
-\item {\bf AI Planning (AIP)} is the Artificial Intelligence component that studies the synthesis of sets of actions to achieve some given objectives~\cite{ghallab2004automated}.
-AIP arose in the late ’50s from converging studies into {\em combinatorial search}, {\em theorem proving} and {\em control theory} and now, is a well formalized paradigm for problem solving with algorithms that scale-up reasonably well.
+\item $V=\tup{v_0, \ldots, v_m}$ is the set of $m$ {\em state variables} with finite domain.
+The respective domains $D=\tup{D_{v_0}, \ldots, D_{v_m}}$ define, for each variable $v\in V$, its set of possible values.
+\item $A$ is a set of {\em actions} whose dynamics are specified with two functions:
+\begin{itemize}
+\item $\rho(s,a)=[0|1]$, determines whether action $a\in A$ is applicable in a state $s$.
+\item $\theta(s,a)=s'$, defines the {\em successor state} $s'$ that results of applying an action $a$ in a state $s$.
+\end{itemize}  
+\item $I$ is an {\em initial state}, i.e. a full assignment of values to the state variables.
+\item $G$ is the set of {\em goal conditions} constraining the possible values of the state variables in the goal states.  
+\end{itemize}
+A {\em solution} to an AI planning problem is a sequence of applicable actions such that its application, starting from the initial state, reach a state where all goal conditions are met.
+
 State-of-the-art planners are able to synthesize plans with hundreds of actions in seconds time~\cite{geffner2013concise}.
@@
 
-\item {\bf Test driven development (TDD)}~\cite{beck:TDD:2003} is a popular paradigm for software development that is frequently used in {\it agile methodologies}~\cite{cohen2003agile}.
+\subsubsection{TDD programming}
+{\em Test driven development (TDD)}~\cite{beck:TDD:2003} is a popular paradigm for software development that is frequently used in {\it agile methodologies}~\cite{cohen2003agile}.
 In TDD, test cases are created before the program code is written and they are run against the code during the development, e.g.
@@
 Further, tests alert programmers of bugs before handing the code off to clients (the cost of finding a bug when the code is first written is considerably lower than the cost of detecting and fixing it later).
-%Last but not least, writing a thorough set of tests cases forces programmers to think through inputs, outputs, and error conditions of programs.
-\end{itemize}
 
-\subsection{Automated synthesis and validation of programs}
-We use input-output {\em test cases} to specify the semantics of the aimed programs, as in Test Driven Development.
-Then, we encode the {\em TDD programming} (or validation) task as an AI planning problem and finally, we use an off-the-shelf AI planner to compute solutions to these problems.
-
-\subsubsection{The TDD programming task}
 A {\em TDD programming task} is a programming task where the semantics of the aimed program is specified with a set of input-output {\em test cases}.
@@
 
-\subsubsection{The AI planning problem}
-An {\em AI planning problem} is formalized as a tuple $\tup{V,D,A,I,G}$ where:
-\begin{itemize}
-\item $V=\tup{v_0, \ldots, v_m}$ is the set of $m$ {\em state variables} with finite domain.
-The respective domains $D=\tup{D_{v_0}, \ldots, D_{v_m}}$ define, for each variable $v\in V$, its set of possible values.
-\item $A$ is a set of {\em actions} whose dynamics are specified with two functions:
-\begin{itemize}
-\item $App(s)\subseteq A$, defines the subset of actions applicable in a state $s$.
-\item $\theta(s,a)=s'$, defines the {\em successor state} $s'$ that results of applying an action $a$ in a state $s$.
-\end{itemize}  
-\item $I$ is an {\em initial state}, i.e. a full assignment of values to the state variables.
-\item $G$ is the set of {\em goal conditions} constraining the possible values of the state variables in the goal states.  
-\end{itemize}
-A {\em solution} to an AI planning problem is a sequence of applicable actions such that its application, starting from the initial state, reach a state where all goal conditions are met.
-
-\subsubsection{Program synthesis and validation as AI planning}
+\subsection{Program synthesis and validation as AI planning}
+We use input-output {\em test cases} to specify the semantics of the aimed programs, as in Test Driven Development.
+Then, we encode the {\em TDD programming} (or validation) task as an AI planning problem and finally, we use an off-the-shelf AI planner to compute solutions to these problems.
 Here we introduce how to encode a {\em TDD programming task} as an {\em AI planning problem}.
@@
 The program in Figure~\ref{fig:sum} assumes that the program variables are ${\mathcal V}=\{x,y,x=N,y=N\}$ and that the value of $x$ and $y$ is initially 0.
-The instruction set is ${\mathcal A}=\{x:=x+1, y:=y+1, x:=x+y, y:=y+x\}$.
+The instruction set is ${\mathcal A}=\{x:=x+1, y:=y+1, x:=x-1, y:=y-1, x:=x+y, y:=y+x, x:=x-y, y:=y-x\}$.
 For the synthesis of this program we are using the following five test cases $\{0 = \sum_0^0 x, 1 = \sum_0^1 x, 3 = \sum_0^2 x, 6 = \sum_0^3 x, 10 = \sum_0^4 x\}$ so the domains of variables $x$ and $y$ are ${\mathcal D}_x = {\mathcal D}_y = [0,10]$.
@@
 Figure~\ref{fig:activity} shows a five-line program (pictured as a finite state machine) that represents the sequence of instructions required for {\em making an orange juice} from the {\em domotics} domain.
-We plan to evaluate our approach in synthesis and validation tasks coming from these real-world domains.
+We plan to evaluate our approach in synthesis and validation tasks coming from different activities in these real-world domains.
 
@@
 
-To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ have three program lines, an instruction set that comprises four instructions and a single observable Boolean variable ($x=N$).
+To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ have three program lines, an instruction set that comprises eight instructions and a single observable Boolean variable ($x=N$).
 Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em making an orange juice} have five program lines, an instruction set with four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$) and the single observable Boolean variable, ${\tt ready(juice)}$, that holds when the orange juice is recognized to be ready.
@@
 
-We designed a 24-month workplan plan for the {\bf development of a user-interactive program synthesizer} that works in two different modes:
+We designed a 12-month workplan plan for the {\bf development of a user-interactive program synthesizer} that works in two different modes:
 \begin{enumerate}
-\item {\em Synthesis mode}. Takes as input a set of test cases that specify the TDD programming task to solve and outputs a program source code that passes these tests with a bug-free guarantee.
+\item {\em Synthesis mode}.
+Takes as input a set of test cases that specify the TDD programming task to solve and outputs a program source code that passes the input test cases with a bug-free guarantee.
 \item {\em Validation mode}.
-In this setting the {\em program synthesizer} receives an additional input, specifying the source code of the program to validate, and it outputs a {\em validation certificate} guaranteeing that the input program either succeeds (or fails) to solve each test case from the TDD programming task.
+In this setting the {\em program synthesizer} receives an additional input: the source code of the program to validate.
+The output is a {\em validation certificate} guaranteeing that the input program either succeeds (or fails) to solve each test case in the TDD programming task.
 \end{enumerate}
 
-Figure~\ref{fig:gantt} details the proposed 24-month timeline for the development of the project. A deliverable is provided at the end of each task.
+Figure~\ref{fig:gantt} details the proposed 12-month timeline for the development of the project. A deliverable is provided at the end of each task.
 
@@
   canvas/.append style={label=below:} ]{1}{12} 
-\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Design (6 months)}{1}{3} \\
-\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Development (6 months)}{4}{6}\\
-\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Experiments \& dissemination (12 months)}{7}{12}
+\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Design (3 months)}{1}{3} \\
+\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Development (3 months)}{4}{6}\\
+\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Experiments \& dissemination (6 months)}{7}{12}
 \end{ganttchart}
@@
 \begin{enumerate}
-\item {\bf T1. System design (months 1-6})
+\item {\bf Task1. System design (months 1-3})
   \begin{small}
@@
 
-  \item {\bf T2. Development of the system architecture (months 7-12})
+  \item {\bf Task2. Development of the system architecture (months 4-6})
     \begin{small}
@@
 
-\item {\bf T3. Experiments and dissemination of results (months 13-24}).
+\item {\bf Task3. Experiments and dissemination of results (months 7-12}).
    \begin{small}
@@
 \subsection{Presupuesto} 
-Esta sección presenta la descripción detallada del presupuesto, de dos anualidades, para el desarrollo del proyecto:
+Esta sección presenta la descripción del presupuesto, de una anualidad, para el desarrollo del proyecto:
 
@@
 \begin{small}  
-  \begin{tabular}{clc|r}
-    {\large\bf Prioridad} & {\large\bf Descripcion} & {\large\bf Año} & {\large\bf Euros} \\\hline
-1 & {\scriptsize Difusión de las actividades del grupo} & 1 & 4,000\\    
-2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 1 & 2,000\\
-3 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 1 & 2,000\\\hline
-1 & {\scriptsize Difusión de las actividades del grupo} & 2 & 4,000\\    
-2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2 & 2,000\\
-3 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2 & 2,000\\\hline
-\multicolumn{3}{l|}{} & 16,000 \\
+  \begin{tabular}{cl|r}
+    {\large\bf Prioridad} & {\large\bf Descripcion} & {\large\bf Euros} \\\hline
+1 & {\scriptsize Difusión de las actividades del grupo}  & 4,000\\    
+2 & {\scriptsize Viajes, manutención y alojamiento grupo de investigación} & 2,000\\
+3 & {\scriptsize Viajes, manutención, alojamiento y ponencias investigadores invitados} & 2,000\\\hline
+\multicolumn{2}{l|}{} & 8,000 \\
   \end{tabular}
@@
 
+Aquí detallamos cada una de las tres partidas que componen el presupuesto del proyecto:
 \begin{enumerate}
 \item {\em Difusión de las actividades del grupo}.
-Esta partida contempla las inscripciones a congresos internacionales en Inteligencia Artificial, previsiblemente los tres siguientes: The International Conference on Automated Planning and Scheduling (h5-index 27).
+Esta partida contempla las inscripciones a congresos internacionales en Inteligencia Artificial, previsiblemente los siguientes: The International Conference on Automated Planning and Scheduling (h5-index 27).
 The International Joint Conference on Artificial Intelligence (h5-index 61).
@@
 \begin{enumerate}
-\item A new {\bf evaluation methodology for the {\em Synthesis and Validation of Programs}}.
-The application of the exiting {\em AI Planning} technology to program synthesis can provide new evaluation metrics that assess how well a program covers a set of {\em input-output} test cases.
 \item An empirical {\bf study on the performance of the state-of-the-art AI planners for the {\em Synthesis and Validation of Programs}}.
@@
 This project will help to meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks.
+\item A new {\bf evaluation methodology for the {\em Synthesis and Validation of Programs}}.
+The application of the exiting {\em AI Planning} technology to program synthesis can provide new evaluation metrics that assess how well a program covers a set of {\em input-output} test cases.
 \item The {\bf development of open software and open benchmarks for the {\em Synthesis and Validation of Programs}}.

### 2018-10-01 doc/icaps-19/model-recognition.tex
added first justification of the likelyhoods estimation

@@
 \begin{abstract} 
-Given the partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models to update these variables), {\em model recognition} is the task of identifying which model in the set explains (produced) the given observation.
+Given the partial observation of a plan execution and a set of possible planning models (models that share the same state variables but update these variables according to different action models), {\em model recognition} is the task of identifying which model in the set explains (produced) the given observation.
 The paper formalizes the {\em model recognition} task and proposes a novel method to estimate the probability of a \strips\ model to produce a partial observation of a plan execution.
-This method, that we called {\em model recognition as planning}, is built on top of off-the-shelf classical planning algorithms and elicit the likelihood of the observation of a plan execution given a candidate model.
-{\em Model recognition as planning} is robust to missing intermediate states and actions in the observed plan execution.
-The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different kinds of {\em finite state machines}.
+This method, that we called {\em model recognition as planning}, is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.
+The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different kinds of {\em automata}.
 We show that {\em model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unobserved.
@@
 \label{sec:introduction}
-{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behavior.
-Plan recognition is considered {\em automated planning} in reverse; while automated planning aims to compute a sequence of actions that accounts for a given goals, plan recognition aims to compute the goals that account for an observed sequence of actions~\cite{geffner:book:2013}.
+{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behavior~\cite{carberry2001techniques}.
+{\em Goal recognition} is a closely related task that aims identifying the goals of the observed agent.
+Plan/goal recognition is considered {\em automated planning} in reverse; while automated planning compute sequences of actions that accounts for a given goals, plan/goal recognition compute goals that account for an observed sequence of actions~\cite{geffner:book:2013}.
 
-Diverse approaches has been proposed for plan recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{carberry2001techniques,sukthankar2014plan}.
+Diverse approaches has been proposed for plan recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{sukthankar2014plan}.
 {\em Plan recognition as planning} is the model-based approach for plan recognition~\cite{ramirez2012plan,ramirez2009plan}.
@@
 
-This paper introduces the {\em model recognition} task where the object to recognize is not a goal but a {\em planning model}.
-Given a partial observation of a plan execution and a set of possible planning models (that share the same state variables but define different action models to update these variables), {\em model recognition} is the task of identifying which model in the set has the highest probability of explaining (producing) the given observation.
+This paper formalizes the {\em model recognition} task where the object to recognize is not a goal but the {\em planning model} that determines the behavior of the observed agent.
+Given a partial observation of a plan execution and a set of possible planning models (models that share the same state variables but update these variables with different action models), {\em model recognition} is the task of identifying the model in the set with the highest probability of explaining (producing) the given observation.
 
@@
   \end{scriptsize}  
- \caption{\small Observation of the execution of a robot navigation plan in a $5\times 5$ grid and a 2-state automata for controlling that the robot only increments its x-coordinate at {\em even} rows (i.e.
+ \caption{\small ({\em Left)} Robot navigation plan in a $5\times 5$ grid.
+{\em (Right)} 2-state automata to control that the robot only increments its x-coordinate at {\em even} rows (i.e.
 when $q_0$ holds).}
@@
 To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model the robot could increment its {\em x-coordinate} at {\em even} rows (when {\tt\small $q0$} holds) and decrement it at the {\em odd rows} (when {\tt\small $q1$} holds).
+According to this model the robot could increment its {\em x-coordinate} at {\em even rows} (when {\tt\small $q0$} holds) and decrement it at the {\em odd rows} (when {\tt\small $q1$} holds).
 Apart from this particular navigation model, different action models could be defined within the same state variables (e.g.
 altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated) and these models can determine different kinds of robot navigation.
-Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} aims here to identify which navigation model produced that observation, despite key information is unobserved (e.g.
-the particular applied actions or the value of {\small\tt $q0$} and {\small\tt $q1$}).
+Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} would aim here to identify which navigation model produced that observation, despite key information is unobserved (e.g.
+the particular applied actions or the value of the state variables {\small\tt $q0$} and {\small\tt $q1$}).
 
 {\em Model recognition} is of interest because once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
-In addition, it enables identifying different kinds of automatae by observing their execution.
+In addition, it enables identifying different kinds of automata by observing their execution.
 It is well-known that diverse automatae representations, like {\em finite state controllers}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,ivankovic2015optimal,segovia2017generating}.
 
-The paper introduces also {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
+The paper introduces {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
 Our method is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
@@
 
-We evaluate the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em finite state machines}.
-All of these {\em automatae} are defined within the same input {\em alphabet} and same {\em machine states} but different {\em transition functions}.
+We evaluate the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em automata}.
+All of these {\em automata} are defined within the same state variables (same input {\em alphabet} and same {\em machine states}) but different {\em transition functions}.
 We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite internal machine states or actual applied transitions are unobserved.
@@
 
-A {\em classical planning frame} is a pair $\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions whose semantics are specified with two functions: $\rho(s,a)$ that denotes whether an action $a\in A$ is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying $a$ in $s$.
-In this work we specify action semantics (the $\rho$ and $\theta$ functions) with the \strips\ model.
-With this regard, an action $a\in A$ is defined with:
+A {\em classical planning frame} is a $\tup{F,A}$ pair, where $F$ is a set of fluents and $A$ is a set of actions.
+The semantics of actions $a\in A$ are specified with two functions: $\rho(s,a)$ that denotes whether the action is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying the action $a$ in a state $s$.
+
+In this work we specify action semantics (the $\rho$ and $\theta$ functions) with the \strips\ model. With this regard, an action $a\in A$ is defined with:
 \begin{itemize}
@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
-A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $s=\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 The {\em plan length} is denoted with $|\pi|=n$ .
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$, the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{a_1^o, \ldots , a_n^o, s_m^o}$, an interleaved combination of {\small $1\leq n\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|+1$} observed states such that:
+Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$; the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{a_1^o,s_1^o \ldots , a_n^o, s_m^o}$, an interleaved combination of {\small $1\leq n\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states such that:
 \begin{itemize}
@@
 This means that the sequence of observed actions $\tup{a_1^o, \ldots, a_n^o}$ in $\tau$ is a sub-sequence of the solution plan $\pi$.
-\item Observed states are a sub-sequence  of partial states that is {\em consistent} with the sequence of states traversed by $\pi$. 
+\item Observed states are a sub-sequence  of partial states {\em consistent} with the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ traversed by $\pi$. 
 \end{itemize}
@@
 \begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\tau$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 0\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
+Given a subset of fluents $\Phi\subseteq F$ we say that $\tau$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
 \end{definition}
 
-For instance, Figure~\ref{fig:grid-example} illustrates the six-state $\Phi$-observation \{{\tt\footnotesize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\} where $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
-This means that, for each observed state, only the value of fluents {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} is known while the value of the remaining fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)} is unknown.
+For instance, Figure~\ref{fig:grid-example} illustrates the six-state $\Phi$-observation \{{\tt\footnotesize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\}, where $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
+This means that, for each observed state, only the value of fluents {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} is known while the value of the remaining fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unknown.
 
@@
 \item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} s.t. the semantics of actions $a\in A$ is unknown because the corresponding functions, $\rho(s,a)$ and/or $\theta(s,a)$, are undefined.
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite and non-empty {\em set of models} for the actions in $A$ s.t.
-each model in $\mathcal{M}\in M$, defines a different function pair $\tup{\rho,\theta}$.
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite non-empty {\em set of models} for the actions in $A$ s.t.
+each model in $\mathcal{M}\in M$, defines different function pairs $\tup{\rho,\theta}$ over the state variables $F$.
 \item $\tau$ is an {\em observation} of the execution of a solution plan $\pi$ for $P$.
@@
 
-{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model and the observed plan execution is the single example to classify.
-In this case, the planning model that is associated to a class is acting as the {\em class prototype} and summarizes all the plan executions that could be synthesized with that model (all the examples that belong to that class).
+{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model in $M$ and the observed plan execution $\tau$ is the single example to classify.
+Further, the planning model associated to a class acts as a {\em class prototype} that summarizes all the plan executions that could be synthesized with that model (i.e.
+all the examples that belong to that class).
 
-In {\em model recognition}, hypotheses are then about the possible action models $\mathcal{M}\in M$ while the observation $\tau$ of a plan execution represents the input observation.
-The {\em naive Bayes classifier} assigns a model $\mathcal{M}\in M$ to the given observation $\tau$ wrt the following expression, $argmax_{\mathcal{M}\in M} P(\mathcal{M})\times P(\tau|\mathcal{M})$.
+We use the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\tau$ wrt the following expression, $argmax_{\mathcal{M}\in M} P(\tau|\mathcal{M})\times P(\mathcal{M})$.
+The hypotheses in {\em model recognition} are then about the possible action models $\mathcal{M}\in M$, while the observation $\tau$ of a plan execution represents the input observation.
 The {\em solution} to the {\em model recognition} task is the model $\mathcal{M}\in M$ (or subset of models in $M$) that maximizes the previous expression.
@@
 If this probability is not given as input it is reasonable to assume that {\em a priori} all models are equiprobable.
-Our approach for {\em model recognition} is then to estimate the $P(\tau|\mathcal{M})$ likelihood according to the {\em amount of edits} required by the model $\mathcal{M}$ to produce a plan $\pi$ such that:
-\begin{enumerate}
-\item $\pi$ solves $P$ and,
-\item $\tau$ is a {\em consistent} observation of the execution of $\pi$ on the classical planning problem $P$. 
-\end{enumerate}
-The {\em edit distance} is a similarity metric that is traditionally computed over {\em strings} or {\em graphs} and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
-In this work we show how to compute this similarity metric referred to the edition of classical planning models.
+The challenge in our formulation for {\em model recognition} is the definition of the $P(\tau|\mathcal{M})$ {\em likelihoods} that express the probability of observing $\tau$ when $\mathcal{M}$ is the used planning model.
 
@@
 \label{sec:asPlanning}
-Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of actions (i.e. $\rho$ and $\theta$ functions) are specified with \strips\ action schemas.
-
+Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of actions (i.e.
+$\rho$ and $\theta$ functions) are specified with \strips\ action schemas.
 We start formalizing the \strips\ schema and define the full space of possible \strips\ schema.
@@
 \strips\ action schema provide a compact representation for specifying classical planning models.
-For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema, codded in PDDL, that determine a robot navigation in a $n\times n$ grid.
+For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema, codded in PDDL, that determine robot navigation in $n\times n$ grids.
 
@@
 Given a \strips\ action schema $\xi$, let us define an additional set of objects ($\Omega\cap\Omega_\xi=\emptyset$), that we denote as {\em variable names}, and that contains one variable name for each parameter in $pars(\xi)$, that is $\Omega_\xi=\{v_i\}_{i=1}^{|pars(\xi)|}$.
-Note that, for any of the six schema defined in Figure~\ref{fig:model-example}, $\Omega_\xi=\{v_1,v_2\}$.
+For any of the six schema defined in Figure~\ref{fig:model-example}, $|pars(\xi)|=2$ so $\Omega_\xi=\{v_1,v_2\}$.
 
@@
 One can also introduce domain-specific knowledge to more precisely bound the space of possible \strips\ schema for a particular domain.
-For instance, in a {\em robot navigation} model, like the one in Figure~\ref{fig:model-example}, predicates {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold a the same time in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
+For instance, in a {\em robot navigation} model, like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold a the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
 Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate is coding the {\em succesor} function for {\em natural numbers}.
-In this case, introducing these domain-specific constraints would reduce the size of the space of possible schema to $2^{2\times 7}=16,384$ for every action schema.
+In this case, introducing these domain-specific constraints reduces the size of the space of possible schema to $2^{2\times 7}=16,384$ for every action schema.
 \end{itemize}
 
-Now we are ready to define what is a well-defined \strips\ action schema.
+Now we are ready to define what is a {\em well-defined} \strips\ action schema.
 \begin{definition}[Well-defined \strips\ action schema]
@@
 \end{definition}
-We say that an action model is well-defined if all its \strips\ action schema are well-defined.
+We say that an action model is {\em well-defined} if all its \strips\ action schema are {\em well-defined}.
 
 \subsection{Edit distances for \strips\ action models}
-We define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
+{\em Edit distances} are similarity metrics traditionally computed over {\em strings} or {\em graphs} and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
+In this work we formalize and compute edit distances that are referred to \strips\ planning models.
+
+First, we define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
 \begin{itemize}
@@
 The observation of a plan execution, generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrain further the space of the possible schema $\xi\in \mathcal{M}$.
-In this case, we talk about {\em observation constraints} that can also be added to the $\mathcal{C}$ set.
-In addition, this new kind of model constraints allow us to define an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.
+In this case, we talk about {\em observation constraints} (that can also be added to the $\mathcal{C}$ set).
+In addition, {\em observation constraints} allow us to define an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.
 
@@
 Remarkably, the {\em observation edit distance} allows us to elicit the likelihood of the observations given a candidate model.
-It can be argued that the shorter this  distance the better the given model explains the given observation and hence, the higher the $P(\tau|\mathcal{M})$ likelihood.
-In particular this distance is maximum when it requires fully editing all the schemas in the model while it is minimum when the given model is already able to produce the input observation without introducing any change.
-Intermmediate distance values reflect how far models are from explaining the input observations.
-In this work we map the {\em observation edit distance} into a $P(\tau|\mathcal{M})$ likelihood with the following expression, $P(\tau|\mathcal{M}) = 1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$.
-
+It can be argued that the shorter this  distance the better the given model explains the given observation.
 Note that the {\em observation edit distance} could also be defined assessing the edition required by the observed plan execution to match the given model.
@@
 
-\section{Model recognition with classical planning}
+\begin{definition}[{\em closest compliant set of models}] \label{compliant}
+Given a model $\mathcal{M}$.
+The {\bf closest compliant set of models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) that can produce a solution plan consistent with $\tau$;
+  \[\mathcal{M^*}=\underset{\forall \mathcal{M}' \rightarrow \tau}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
+\end{definition}
+
+\begin{figure}
+  \begin{scriptsize}
+  \begin{center}
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
+	  \node[state] (A)              {$\mathcal{M}$};
+	  \node[state] (B) [right of=A] {$\mathcal{M}'$};
+	  \node[state] (C) [right of=B] {$\pi$};
+	  \node[state] (D) [right of=C] {$\tau$};          
+	  \path
+(A) edge  node {} (B)
+(B) edge  node {} (C)
+(C) edge  node {} (D)
+;
+	\end{tikzpicture}
+  \end{center}
+  \end{scriptsize}  
+ \caption{\small {\em Bayesian network} abstracting that $\mathcal{M}'$ is a model that results from editing the input model $\mathcal{M}$ s.t.
+$\mathcal{M}'$ can produce a plan $\pi$ that that solves $P$ and is consistent with $\tau$.}
+\label{fig:net}
+\end{figure}
+
+\subsection{Approximating the $P(\tau|\mathcal{M})$ {\em likelihood}}
+Intuitively, our approach to estimate the $P(\tau|\mathcal{M})$ {\em likelihoods} is to assess how far is $\mathcal{M}$ from a model $\mathcal{M'}$ that can produce a plan $\pi$ that solves $P$ and s.t.
+$\tau$ is a {\em consistent} observation of the execution of $\pi$ on $P$.
+
+Figure~\ref{fig:net} shows the four-variable {\em Bayesian network} abstracting that $\mathcal{M}'$ is a model that results from editing the input model $\mathcal{M}$ s.t.
+$\mathcal{M}'$ can produce a plan $\pi$ that that solves $P$ and is consistent with $\tau$.
+Regarding the network of Figure~\ref{fig:net}, a natural way to formulate the aimed {\em likelihood} is $\displaystyle P(\tau|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi}\frac{P(\mathcal{M},\mathcal{M}',\pi,\tau)}{P(\mathcal{M})}$ where $\mathcal{M}'$ ranges over all the models that can be generated editing $\mathcal{M}$ and $\pi$ ranges over all the plans that can be synthesized with a model $\mathcal{M}'$.
+Since the joint probability distribution can be marginalized as $\displaystyle P(\mathcal{M},\mathcal{M}',\pi,\tau)=P(\mathcal{M})P(\mathcal{M'}|\mathcal{M})P(\mathcal{\pi}|\mathcal{M'})P(\tau|\mathcal{\pi})$ we have that the aimed {\em likelihoods} can be rewritten as $\displaystyle P(\tau|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi} P(\mathcal{M'}|\mathcal{M})P(\mathcal{\pi}|\mathcal{M'})P(\tau|\mathcal{\pi})$.
+
+We know that $P(\tau|\mathcal{\pi})=0$ for plans that are not consistent with $\tau$ likewise, $P(\mathcal{\pi}|\mathcal{M'})=0$ for plans that cannot be generated with a model $\mathcal{M'}$, i.e.
+for {\em invalid plans}~\cite{howey2004val}.
+Therefore, we can assume that $\pi$ ranges only over the valid plans consistent with $\tau$ and further, that $\mathcal{M}'$ ranges over the models that can synthesize {\em valid plans} consistent with $\tau$.
+
+The exact computation of $P(\tau|\mathcal{M})$ is still likely to be intractable since for most planning problem $P$, the set of {\em valid} plans consistent with an observation $\tau$ can easily be very large (is infinite in the case of planning tasks without dead-ends).
+Here we propose to estimate $P(\tau|\mathcal{M})$ making the following assumptions:
+\begin{enumerate}
+\item The further the compliant model the lower the likelihood so we assume that the sum in the previous equation is dominated by the closest compliant set.
+Further we assume that the {\em closest compliant set of models} is a singleton, so that probabilities corresponding to different compliant models are not added up.
+This also means that $P(\pi|\mathcal{M^*})=1$ since the closest compliant model can by definition generate a plan $\pi$ that solves $P$.
+Therefore, $\displaystyle P(\tau|\mathcal{M}) \approx \sum_{\pi} P(\mathcal{M^*}|\mathcal{M})P(\tau|\mathcal{\pi})$ where $\pi$ ranges now over the over the valid plans consistent with $\tau$ and valid for the closest compliant model $\mathcal{M^*}$.
+
+\item If we assume that all the observations of a given plan are equiprobable, then the value of $P(\tau|\mathcal{\pi})$ is given by a normalizing constant $\alpha$.
+
+\item $P(\mathcal{M^*}|\mathcal{M})$ is propotional the {\em observation edit distance}.
+In particular this distance is maximum when it requires fully editing all the schemas in the model while it is minimum when the given model is already able to produce the input observation without introducing any change.
+Intermmediate distance values reflect how far models are from explaining the input observations.
+In this work we get map the {\em observation edit distance} into a $P(\tau|\mathcal{M})$ likelihood with the following expression, $P(\tau|\mathcal{M}) \approx \alpha' 1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$, where $\alpha'$ is a normalizing constant.
+\end{enumerate}
+
+\section{Model Recognition as classical planning}
 This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects (and hence the $P(\tau|\mathcal{M})$ likelihood can be estimated).
@@
 \begin{figure}
-{\tt\footnotesize
+{\tt\scriptsize
 00 : (insert\_pre\_xcoord\_v1\_inc-x)\\
@@
 06 : (insert\_add\_xcoord\_v2\_inc-x)\\
-07 : (validate\_0)
+07 : (validate\_0)\\
 08 : (editable\_inc-x 1 2)\\
@@
 }
- \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-13]) the planning model of Figure\ref{fig:model-example} when action {\tt\small inc-x} has no preconditions and positive/negative effects are swapped wrt Figure\ref{fig:model-example}.}
+ \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-14]) the planning model of Figure\ref{fig:model-example} when action {\tt\small inc-x} has no preconditions and positive/negative effects are swapped wrt Figure\ref{fig:model-example}.}
 \label{fig:plan-pdistance}
@@
 	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.8cm,semithick]
-	  \node[state] (A)              {$Q_0$};
-	  \node[state] (B) [right of=A] {$Q_1$};
-	  \node[state] (C) [right of=B] {$Q_2$};
-	  \node[state] (D) [right of=C] {$Q_3$};
-	  \node[state] (E) [below of=A] {$Q_4$};
-	  \node[state] (F) [right of=E] {$Q_5$};
+	  \node[state] (A)              {$q_0$};
+	  \node[state] (B) [right of=A] {$q_1$};
+	  \node[state] (C) [right of=B] {$q_2$};
+	  \node[state] (D) [right of=C] {$q_3$};
+	  \node[state] (E) [below of=A] {$q_4$};
+	  \node[state] (F) [right of=E] {$q_5$};
 
@@
 \label{sec:related}
-Plan recognition.
-
-Model Reconciliation.
-
-Instace Based Classification.
+In this work we do not assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
+A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
+Also  related to this paper is the work on {\em Goal Recognition Design} but in that work the action model is given in advance and fixed~\cite{keren2014goal}.
 
@@
 \label{sec:conclussions}
-In this work we do not assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
-A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
+This paper formalized the {\em model recognition} task and proposed, {\em model recognition as planning}, a method built on top of off-the-shelf classical planning algorithms to estimate the probability of a \strips\ model to produce a partial observation of a plan execution.
+The paper shows the effectiveness of {\em model recognition as planning} in a set of \strips\ models encoding different kinds of {\em automata}.
+{\em Model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unobserved.
 

### 2018-10-02 doc/icaps-19/model-recognition.tex
new full version icaps19 paper. Ready for reviews

@@
 \begin{itemize}
-\item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} s.t. the semantics of actions $a\in A$ is unknown because the corresponding functions, $\rho(s,a)$ and/or $\theta(s,a)$, are undefined.
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite non-empty {\em set of models} for the actions in $A$ s.t.
-each model in $\mathcal{M}\in M$, defines different function pairs $\tup{\rho,\theta}$ over the state variables $F$.
+\item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} s.t. the semantics of actions $a\in A$ is unknown, i.e. the corresponding functions $\rho(s,a)$ and/or $\theta(s,a)$ are undefined.
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite non-empty {\em set of models} for the actions in $A$ s.t., each model in $\mathcal{M}\in M$, defines different function pairs $\tup{\rho,\theta}$ over the state variables $F$.
 \item $\tau$ is an {\em observation} of the execution of a solution plan $\pi$ for $P$.
@@
 
-{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model in $M$ and the observed plan execution $\tau$ is the single example to classify.
-Further, the planning model associated to a class acts as a {\em class prototype} that summarizes all the plan executions that could be synthesized with that model (i.e.
-all the examples that belong to that class).
+{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model $\mathcal{M}\in M$ and the observed plan execution $\tau$ is the single example to classify.
+Further, the planning model associated to a class acts as a {\em class prototype} that summarizes all the plan executions that could be synthesized with that model (all the examples that belong to that class).
 
-We use the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\tau$ wrt the following expression, $argmax_{\mathcal{M}\in M} P(\tau|\mathcal{M})\times P(\mathcal{M})$.
+We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\tau$ wrt the following expression, $argmax_{\mathcal{M}\in M} P(\tau|\mathcal{M})\times P(\mathcal{M})$.
 The hypotheses in {\em model recognition} are then about the possible action models $\mathcal{M}\in M$, while the observation $\tau$ of a plan execution represents the input observation.
@@
 If this probability is not given as input it is reasonable to assume that {\em a priori} all models are equiprobable.
-The challenge in our formulation for {\em model recognition} is the definition of the $P(\tau|\mathcal{M})$ {\em likelihoods} that express the probability of observing $\tau$ when $\mathcal{M}$ is the used planning model.
+The challenge in our formulation for {\em model recognition} is the definition of the $P(\tau|\mathcal{M})$ {\em likelihood} that express the probability of observing $\tau$ when $\mathcal{M}$ is the planning model.
 
@@
 
-\begin{definition}[{\em closest compliant set of models}] \label{compliant}
+\begin{definition}[{\em Closest compliant models}] \label{compliant}
 Given a model $\mathcal{M}$.
-The {\bf closest compliant set of models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) that can produce a solution plan consistent with $\tau$;
+The {\bf closest compliant models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) that can produce a solution plan consistent with $\tau$;
   \[\mathcal{M^*}=\underset{\forall \mathcal{M}' \rightarrow \tau}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
@@
   \end{scriptsize}  
- \caption{\small {\em Bayesian network} abstracting that $\mathcal{M}'$ is a model that results from editing the input model $\mathcal{M}$ s.t.
-$\mathcal{M}'$ can produce a plan $\pi$ that that solves $P$ and is consistent with $\tau$.}
+ \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}'$ results from editing $\mathcal{M}$ and that can produce a plan $\pi$ consistent with $\tau$.}
 \label{fig:net}
@@
 \subsection{Approximating the $P(\tau|\mathcal{M})$ {\em likelihood}}
-Intuitively, our approach to estimate the $P(\tau|\mathcal{M})$ {\em likelihoods} is to assess how far is $\mathcal{M}$ from a model $\mathcal{M'}$ that can produce a plan $\pi$ that solves $P$ and s.t.
+Intuitively, our approach to estimate the $P(\tau|\mathcal{M})$ {\em likelihood} is to assess how far is $\mathcal{M}$ from a model $\mathcal{M'}$ that can produce a plan $\pi$ that solves $P$ and s.t.
 $\tau$ is a {\em consistent} observation of the execution of $\pi$ on $P$.
+Figure~\ref{fig:net} shows the four-variable {\em Bayesian network} abstracting that $\mathcal{M}'$ is a model that results from editing the input model $\mathcal{M}$ s.t.
+$\mathcal{M}'$ can produce a solution plan $\pi$ consistent with $\tau$.
+Regarding this network, a natural way to formulate the aimed {\em likelihood} is
+\[
+ P(\tau|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi}\frac{P(\mathcal{M},\mathcal{M}',\pi,\tau)}{P(\mathcal{M})},
+\]
+where $\mathcal{M}'$ ranges over all the models that can be generated editing $\mathcal{M}$ and $\pi$ ranges over all the plans that can be synthesized with a model $\mathcal{M}'$.
 
-Figure~\ref{fig:net} shows the four-variable {\em Bayesian network} abstracting that $\mathcal{M}'$ is a model that results from editing the input model $\mathcal{M}$ s.t.
-$\mathcal{M}'$ can produce a plan $\pi$ that that solves $P$ and is consistent with $\tau$.
-Regarding the network of Figure~\ref{fig:net}, a natural way to formulate the aimed {\em likelihood} is $\displaystyle P(\tau|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi}\frac{P(\mathcal{M},\mathcal{M}',\pi,\tau)}{P(\mathcal{M})}$ where $\mathcal{M}'$ ranges over all the models that can be generated editing $\mathcal{M}$ and $\pi$ ranges over all the plans that can be synthesized with a model $\mathcal{M}'$.
-Since the joint probability distribution can be marginalized as $\displaystyle P(\mathcal{M},\mathcal{M}',\pi,\tau)=P(\mathcal{M})P(\mathcal{M'}|\mathcal{M})P(\mathcal{\pi}|\mathcal{M'})P(\tau|\mathcal{\pi})$ we have that the aimed {\em likelihoods} can be rewritten as $\displaystyle P(\tau|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi} P(\mathcal{M'}|\mathcal{M})P(\mathcal{\pi}|\mathcal{M'})P(\tau|\mathcal{\pi})$.
+Given that the joint probability distribution can be marginalized as $\displaystyle P(\mathcal{M},\mathcal{M}',\pi,\tau)=P(\mathcal{M})P(\mathcal{M'}|\mathcal{M})P(\mathcal{\pi}|\mathcal{M'})P(\tau|\mathcal{\pi})$, that $P(\tau|\mathcal{\pi})=1$ for plans consistent with $\tau$ and that $P(\tau|\mathcal{\pi})=0$ for $\tau$ inconsistent plans.
+Likewise, $P(\mathcal{\pi}|\mathcal{M'})=0$ for plans that cannot be generated with a model $\mathcal{M'}$, i.e.
+for {\em invalid plans}~\cite{howey2004val}.
+We have that
+\[
+P(\tau|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi} P(\mathcal{M'}|\mathcal{M})P(\mathcal{\pi}|\mathcal{M'}),
+\]
+where $\mathcal{M}'$ now ranges over the models that can synthesize {\em valid plans} and $\pi$ ranges only over the valid plans for $\mathcal{M}'$ consistent with $\tau$.
 
-We know that $P(\tau|\mathcal{\pi})=0$ for plans that are not consistent with $\tau$ likewise, $P(\mathcal{\pi}|\mathcal{M'})=0$ for plans that cannot be generated with a model $\mathcal{M'}$, i.e.
-for {\em invalid plans}~\cite{howey2004val}.
-Therefore, we can assume that $\pi$ ranges only over the valid plans consistent with $\tau$ and further, that $\mathcal{M}'$ ranges over the models that can synthesize {\em valid plans} consistent with $\tau$.
+Despite this simplification, the exact computation of $P(\tau|\mathcal{M})$ is still likely to be intractable since for most planning problems, the set of {\em valid} plans consistent with an observation can easily be very large (infinite in the case of planning problems without dead-ends).
+In this work we propose to estimate $P(\tau|\mathcal{M})$ making the following assumptions:
+\begin{enumerate}
+\item The sum in the previous equation is dominated by the {\em closest compliant set}.
+The rationale behind this assumption is that the further the compliant model, the lower the likelihood.
+Further, we assume that the {\em closest compliant set of models} is a singleton, so that probabilities corresponding to different compliant models are not added up
+\[
+P(\tau|\mathcal{M}) \approx \sum_{\pi} P(\mathcal{M^*}|\mathcal{M})P(\pi|\mathcal{M^*}),
+\]
+where $\pi$ ranges now over the valid plans consistent with $\tau$ that are valid for the closest compliant model $\mathcal{M^*}$.
 
-The exact computation of $P(\tau|\mathcal{M})$ is still likely to be intractable since for most planning problem $P$, the set of {\em valid} plans consistent with an observation $\tau$ can easily be very large (is infinite in the case of planning tasks without dead-ends).
-Here we propose to estimate $P(\tau|\mathcal{M})$ making the following assumptions:
-\begin{enumerate}
-\item The further the compliant model the lower the likelihood so we assume that the sum in the previous equation is dominated by the closest compliant set.
-Further we assume that the {\em closest compliant set of models} is a singleton, so that probabilities corresponding to different compliant models are not added up.
-This also means that $P(\pi|\mathcal{M^*})=1$ since the closest compliant model can by definition generate a plan $\pi$ that solves $P$.
-Therefore, $\displaystyle P(\tau|\mathcal{M}) \approx \sum_{\pi} P(\mathcal{M^*}|\mathcal{M})P(\tau|\mathcal{\pi})$ where $\pi$ ranges now over the over the valid plans consistent with $\tau$ and valid for the closest compliant model $\mathcal{M^*}$.
-
-\item If we assume that all the observations of a given plan are equiprobable, then the value of $P(\tau|\mathcal{\pi})$ is given by a normalizing constant $\alpha$.
-
-\item $P(\mathcal{M^*}|\mathcal{M})$ is propotional the {\em observation edit distance}.
-In particular this distance is maximum when it requires fully editing all the schemas in the model while it is minimum when the given model is already able to produce the input observation without introducing any change.
-Intermmediate distance values reflect how far models are from explaining the input observations.
-In this work we get map the {\em observation edit distance} into a $P(\tau|\mathcal{M})$ likelihood with the following expression, $P(\tau|\mathcal{M}) \approx \alpha' 1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}$, where $\alpha'$ is a normalizing constant.
+\item Solution plans, that can be synthesized with a given model, are equiprobable.  
 \end{enumerate}
 
-\section{Model Recognition as classical planning}
-This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\tau)$ can be computed with a compilation of a classical planning with conditional effects (and hence the $P(\tau|\mathcal{M})$ likelihood can be estimated).
+With this regard, our approximation of the $P(\tau|\mathcal{M})$ likelihood becomes proportional to $P(\mathcal{M^*}|\mathcal{M})$, which is given by the {\em observation edit distance}.
+This distance is maximum when it requires fully editing all the schema in the model while it is minimum when the given model is already able to produce the input observation without introducing any change.
+Intermediate distance values reflect how far models are from explaining the input observations.
+In more detail we approximate the $P(\tau|\mathcal{M})$ likelihood with the expression:
+\[
+P(\tau|\mathcal{M}) \approx \alpha\times (1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}),
+\]
+where $\alpha$ is a normalizing constant that reflects the amount of plans valid for $\mathcal{M^*}$ and consistent with $\tau$.
+Since the value of $\delta(\mathcal{M},*)$ is {\em a priori} known, the only missing ingredient to estimate the $P(\tau|\mathcal{M})$ likelihood is the $\delta(\mathcal{M},\tau)$ distance.
 
-The compilation is an extension of the classical planning compilation for the learning of \strips\ planning models~\cite{aineto2018learning}.
-The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
-\begin{enumerate}
-\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$}. The solution plan continues with a postfix that:
-\begin{enumerate}
-\item Induces a solution plan $\pi_\tau$ for the original classical planning problem $P$.
-\item Validates that $\tau$ is an observation of the execution of $\pi_\tau$ on the classical planning problem $P$.
-\end{enumerate}
-\end{enumerate}
-
-Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
-The postfix of the plan (steps [7,14]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
 \begin{figure}
-{\tt\scriptsize
+{\tt\tiny
 00 : (insert\_pre\_xcoord\_v1\_inc-x)\\
@@
 \end{figure}
+
+\section{Model Recognition as classical planning}
+This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\tau)$ can be computed (and hence an approximation of the $P(\tau|\mathcal{M})$ likelihood) with a compilation of a classical planning with conditional effects.
+
+The compilation is an extension of the classical planning compilation for the learning of \strips\ planning models~\cite{aineto2018learning}.
+The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
+\begin{enumerate}
+\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
+A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
+\item {\bf Validates the edited model $\mathcal{M}'$}. The solution plan continues with a postfix that:
+\begin{enumerate}
+\item Induces a solution plan $\pi_\tau$ for the original classical planning problem $P$.
+\item Validates that $\tau$ is an observation of the execution of $\pi_\tau$ on the classical planning problem $P$.
+\end{enumerate}
+\end{enumerate}
+
+Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
+The postfix of the plan (steps [7,14]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
 
@@
 \subsection{A propositional encoding for \strips\ action schema}
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding objects $\Omega_\xi$.
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with the fluents
+
+of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding objects $\Omega_\xi$.
 Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.
@@
 \item Fluents $[pre|del|add]\_e\_\xi$ to model the possible \strips\ schema. 
-\item The fluents to code the {\em observation constraints} and that are of these kinds:
+\item The fluents to code the {\em observation constraints}:
 \begin{itemize}
@@
 
-\item Actions for {\em editing} operator schema $\xi\in\mathcal{M}$:
-\begin{itemize}
-\item Actions for inserting a new precondition into an action schema $\xi\in\mathcal{M}$.
-
+\item Actions for {\em editing} operator schema $\xi\in\mathcal{M}$.
+This includes the actions for {\em inserting} a new {\em precondition} into an action schema $\xi\in\mathcal{M}$ and for inserting a new {\em negative} or {\em positive} effect into the action schema $\xi\in\mathcal{M}$
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{e,\xi}})=&\{\neg pre\_e\_\xi, \neg del\_e\_\xi,\\
+\hspace*{7pt}\pre(\mathsf{insertPre_{e,\xi}})=&\{\neg pre\_e\_\xi, \neg del\_e\_\xi,\\
 & \neg add\_e\_\xi, mode_{edit}\},\\
-\cond(\mathsf{programPre_{e,\xi}})=&\{\emptyset\}\rhd\{pre\_e\_\xi\}.
+\cond(\mathsf{insertPre_{e,\xi}})=&\{\emptyset\}\rhd\{pre\_e\_\xi\}.
 \end{align*}
 \end{small}
-
-\item Actions for inserting a new {\em negative} or {\em positive} effect into the action schema $\xi\in\mathcal{M}$.
-
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{e,\xi}})=&\{\neg del\_e\_\xi,\\
-& \neg add\_e\_\xi,mode_{edit}\},\\
-\cond(\mathsf{programEff_{e,\xi}})=&\{pre\_e\_\xi\}\rhd\{del\_e\_\xi\},\\
+\hspace*{7pt}\pre(\mathsf{insertEff_{e,\xi}})=&\{\neg del\_e\_\xi, \neg add\_e\_\xi,mode_{edit}\},\\
+\cond(\mathsf{insertEff_{e,\xi}})=&\{pre\_e\_\xi\}\rhd\{del\_e\_\xi\},\\
 &\{\neg pre\_e\_\xi\}\rhd\{add\_e\_\xi\}.
@@
 \end{small}
-\end{itemize}
-Besides these actions, the $A'$ set also contains the actions for {\em deleting} a precondition and a negative/positive effect.
+Besides these actions, $A'$ also contains the actions for {\em deleting} a {\em precondition} and a negative/positive {\em effect}.
 
@@
 
+Previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
+Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
+
 Remarkably, the extension of this piece of work to the FOND planning setting~\cite{muise2012improved} is straightforward by simply considering the {\em all-outcomes} determiniztion of the actions with non-determinitic effects~\cite{yoon2007ff}.

### 2018-10-03 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updates to the conselleria project proposal

@@
 
-To illustrate this interest: Since 2012 the {\sc US National Science Foundation} funds the ExCAPE research project\footnote{\tt\url{https://www.nsf.gov/awardsearch/showAward?AWD_ID=1138996}} ($\$3,750,000$) to advance the theory and practice of program synthesis.
+A few facts that illustrate this interest: Since 2012 the {\sc US National Science Foundation} funds the ExCAPE research project\footnote{\tt\url{https://www.nsf.gov/awardsearch/showAward?AWD_ID=1138996}} ($\$3,750,000$) to advance the theory and practice of program synthesis.
 In 2013 a standard framework for program synthesis was defined~\cite{alur2013syntax}.
 Since then, two yearly international competitions are established ({\small\tt\url{http://www.sygus.org}}), to compare the different approaches for program synthesis, and the sister competition ({\small\tt\url{http://www.syntcomp.org/}}), to compare different approaches for the synthesize of reactive programs.
-Further, program synthesis has already been deployed in the real world and it is part of the {\sc Flash Fill} feature of {\sc Microsoft Excel} that automatically generates programs for string transformation~\cite{gulwani2011automating}.
+Further, program synthesis has already been deployed in the real world and it is part of the {\sc Flash Fill} feature of {\sc Microsoft Excel}~\cite{gulwani2011automating}.
 
-Computational approaches to program synthesis range over a wide spectrum, from {\em deductive synthesis} to {\em inductive synthesis}.
-In deductive program synthesis, a program is synthesized by constructively proving a theorem, employing logical inference and constraint solving~\cite{manna1986deductive}.
+Computational approaches to program synthesis range over a wide spectrum, from {\em deductive} to {\em inductive synthesis}.
+In deductive synthesis, program are synthesized by constructively proving a theorem, employing logical inference and constraint solving~\cite{manna1986deductive}.
 On the other hand, inductive synthesis aims to compute a program matching a set of input-output examples by searching in a restricted space of programs~\cite{summers1977methodology,shapiro1983algorithmic}.
@@
 Last but not least, the synthesis of programs from examples is also addressed in the classic AI field of {\em Inductive Logic Programming} (ILP)~\cite{muggleton1991inductive,Raedt:relationalML:book2008}.
-ILP arises from the intersection of {\em Machine Learning} and {\em Logic Programming} and deals with the development of inductive techniques to learn {\em logic programs} from examples and background knowledge, that are expressed as logic facts.
+ILP deals with the development of inductive techniques to learn {\em logic programs} from examples and background knowledge, that are expressed as {\em logic facts}.
 
-{\underline{\em Program Validation}} is the task of proving (or disproving) the correctness of a given program with respect to a formal specification of the aimed program semantics.
-Program validation is considered a {\em necessary step} for program synthesis and {\em model checking} is the mainstream AI approach for the formal validation of programs and controllers~\cite{clarke1999model}.
-Current approaches for model checking reduces to graph search but, instead of enumerating reachable states one at a time, they traverse the state space considering large numbers of states at a single step.
-For instance, representing sets of states and transition relations as logical formulas or {\em binary decision diagrams}, like in {\em symbolic model-checking} ~\cite{mcmillan1993symbolic}.
+{\underline{\em Program Validation}} is the task of proving (or disproving) the correctness of a program with respect to a formal specification of the aimed program semantics.
+Program validation is considered a {\em necessary step} for program synthesis and {\em model checking}, is the mainstream AI approach for the formal validation of programs and controllers~\cite{clarke1999model}.
+Current approaches for model checking reduces to graph search but, instead of enumerating reachable states one at a time, they traverse the state space considering large numbers of states at a single step~\cite{mcmillan1993symbolic}.
+For instance, representing sets of states and transition functions as logical formulas or {\em Binary Decision Diagrams}~\cite{bryant1992symbolic}.
 
@@
 Our current research already shows that this method can synthesize and validate programs for non trivial tasks like sorting lists, traversing graphs or manipulating strings~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia2017generating,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
-Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list, traversing a binary tree, or building a {\em parser} for simple arithmetic operations.
+Table~\ref{tab:programs} reports the time invested by the AI planner {\sc FD}~\cite{helmert2006fast} to solve the following programming tasks: computing the $n^{th}$ term of the {\em summatory} and  {\em Fibonacci} series, {\em reversing} a list, {\em finding} an element (and the {\em minimum} element) in a list, {\em sorting} a list, visiting all the nodes of a binary {\em tree}, or building a {\em parser} for simple arithmetic operations.
 
@@
 \subsection{Background}
-We introduce the technology required by our AI method for the automated synthesis and validation of programs:
+We introduce the technology required by our method for the automated synthesis and validation of programs:
 
@@
 The respective domains $D=\tup{D_{v_0}, \ldots, D_{v_m}}$ define, for each variable $v\in V$, its set of possible values.
-\item $A$ is a set of {\em actions} whose dynamics are specified with two functions:
+\item $A$ is the set of {\em actions} that update the value of the {\em state variables}. The dynamics of an action $a\in A$ is specified with two functions:
 \begin{itemize}
@@
 \end{itemize}  
-\item $I$ is an {\em initial state}, i.e. a full assignment of values to the state variables.
+\item $I$ is the {\em initial state}, i.e. a full assignment of values to the state variables.
 \item $G$ is the set of {\em goal conditions} constraining the possible values of the state variables in the goal states.  
@@
 
-State-of-the-art planners are able to synthesize plans with hundreds of actions in seconds time~\cite{geffner2013concise}.
+State-of-the-art planners can synthesize plans with hundreds of actions in seconds time~\cite{geffner2013concise}.
 The mainstream approach for AIP is {\em heuristic search} with heuristics derived automatically from the problem representation~\cite{mcdermott1996heuristic,bonet2001planning}.
@@
 
-\subsubsection{TDD programming}
+\subsubsection{Test driven development}
 {\em Test driven development (TDD)}~\cite{beck:TDD:2003} is a popular paradigm for software development that is frequently used in {\it agile methodologies}~\cite{cohen2003agile}.
@@
 
-A {\em TDD programming task} is a programming task where the semantics of the aimed program is specified with a set of input-output {\em test cases}.
+We define a {\em TDD programming task} as a programming task where the semantics of the aimed program is specified by a set of input-output {\em test cases}.
 Formally, a {\em TDD programming} task is a tuple $\tup{{\mathcal V},{\mathcal D},{\mathcal A}, {\mathcal T}}$ where:
@@
 \item ${\mathcal V}$ is the set of $n$ {\em program variables}.
-The respective domains are ${\mathcal D}=\tup{{\mathcal D}_{v_0}, \ldots, {\mathcal D}_{v_n}}$ defining the set of possible values for each variable $v\in {\mathcal V}$.
+The respective domains are ${\mathcal D}=\tup{{\mathcal D}_{v_0}, \ldots, {\mathcal D}_{v_n}}$ and define the possible values of each variable $v\in {\mathcal V}$.
 \item ${\mathcal A}$ is the {\em instruction set}, i.e. the set of different instructions that can appear in a program.
-\item ${\mathcal T}$ is a set of input-output {\em test cases} where each test $t\in {\mathcal T}$ is a pair $t=\tup{{\mathcal I}_t,{\mathcal G}_t}$ indicating the input value for the program variables and the aimed output (i.e.
-after the program is executed) for these same variables.
+\item ${\mathcal T}$ is the set of {\em test cases} where each test $t\in {\mathcal T}$ is a pair $t=\tup{{\mathcal I}_t,{\mathcal G}_t}$ indicating the {\em input} value for the program variables and the corresponding aimed {\em output} (after the program run) for these same variables.
 \end{itemize}
@@
 \subsection{Program synthesis and validation as AI planning}
-We use input-output {\em test cases} to specify the semantics of the aimed programs, as in Test Driven Development.
-Then, we encode the {\em TDD programming} (or validation) task as an AI planning problem and finally, we use an off-the-shelf AI planner to compute solutions to these problems.
-Here we introduce how to encode a {\em TDD programming task} as an {\em AI planning problem}.
-For further details on the encoding we refer the reader to~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia2017generating,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
+Now we are ready to describe the details of our method for the automated synthesis and validation of programs. Our method follows this three-step process:
+\begin{enumerate}
+\item Input-output {\em test cases} are used to specify the semantics of the aimed program, as in Test Driven Development.
+\item The resulting {\em TDD programming} (or validation) task is compiled (automatically encoded) as an AI planning problem.
+\item An off-the-shelf AI planner is used to compute solutions to the resulting planning problem.
+\end{enumerate}
 
-Breifly, our encoding defines an {\em AI planning problem} with {\bf state variables} of three kinds:
+Here we introduce the core of our method that is, how to encode a {\em TDD programming task} as an {\em AI planning problem} (step 2).
+For further details on our encoding we refer the reader to~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia2017generating,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
+Briefly, our encoding defines an {\em AI planning problem} with {\bf state variables} of three kinds:
 \begin{itemize}
-\item {\tt program(line) := instruction}, that encode the instructions (from the instruction set ${\mathcal A}$) at the different program lines. Initially all program lines are empty.
-\item {\tt pcounter := line}, encoding which is the current program line. Initially the program counter points to the first program line.
+\item {\tt program(line) := instruction}, that encode the instructions (from the instruction set ${\mathcal A}$) at the different program lines. 
+\item {\tt pcounter := line}, encoding which is the current program line. 
 \item {\tt var := value}, that encode the value of the program variables (the ${\mathcal V}$ set). 
@@
 
-The {\bf initial/goal states} of the AI planning problem encode the input/output values of the program variables (as specified by the test cases ${\mathcal T}$ of the TDD programming task).
-Finally, every program instruction ($w\in {\mathcal A}$) is encoded into the {\em AI planning problem} with two {\bf actions}:
+The {\bf initial state} specifies that initially, all program lines are empty, the program counter points to the first line and that the value of the program variables is as given by the TDD test cases.
+The {\bf goal conditions} of the AI planning problem encode the aimed output values of the program variables as specified by the TDD test cases.
+
+Finally, every program instruction ($w\in {\mathcal A}$) is encoded into the {\em AI planning problem} with {\bf actions} of two kinds:
 \begin{itemize}
@@
 We implement this encoding using standard planning languages, such as PDDL~\cite{fox2003pddl2}, so the AIP tasks resulting from our encoding can be solved with off-the-shelf planners, like the {\sc FD} planning system~\cite{helmert2006fast}.
-Our encoding is {\em complete} and {\em correct}~\cite{segovia:programs:AIJ19}, which means that the programs synthesized with this approach are guaranteed to be bug-free over the given set of {\em input-output} tests cases.
+Our encoding is {\em complete} and {\em correct}~\cite{segovia:programs:AIJ19}, which means that the programs synthesized with out method are guaranteed to be bug-free over the given set of {\em input-output} tests cases.
 
@@
 \label{sec:evaluation}
-
-The performance of our AI method for the synthesis and validation of programs will be evaluated with regard to (1), {\bf computation time} and (2), {\bf memory}.
+The aptitude of a programming process can be assessed with regard to different metrics.
+In this project the performance of our AI method for the synthesis and validation of programs will be evaluated with regard to (1), {\bf computation time} and (2), {\bf memory}.
 Further, the evaluation will be carried out in programming and validation tasks of two different kinds:
@@
 The instruction set is ${\mathcal A}=\{x:=x+1, y:=y+1, x:=x-1, y:=y-1, x:=x+y, y:=y+x, x:=x-y, y:=y-x\}$.
-For the synthesis of this program we are using the following five test cases $\{0 = \sum_0^0 x, 1 = \sum_0^1 x, 3 = \sum_0^2 x, 6 = \sum_0^3 x, 10 = \sum_0^4 x\}$ so the domains of variables $x$ and $y$ are ${\mathcal D}_x = {\mathcal D}_y = [0,10]$.
-Note that $x=N$ and $y=N$ are Boolean variables whose value depends on the current value of $x$ and $y$.
+We synthesize this program with five test cases $\{0=\sum_0^0 x, 1=\sum_0^1 x, 3=\sum_0^2 x, 6=\sum_0^3 x, 10=\sum_0^4 x\}$ so the domains of variables $x$ and $y$ are ${\mathcal D}_x = {\mathcal D}_y = [0,10]$.
+Note that $x=N$ and $y=N$ are Boolean variables whose value depends on the current value of variables $x$ and $y$.
 
@@
 Interestingly many activities that are objects of study in these domains can be modeled as simple programs.
-Figure~\ref{fig:activity} shows a five-line program (pictured as a finite state machine) that represents the sequence of instructions required for {\em making an orange juice} from the {\em domotics} domain.
+Figure~\ref{fig:activity} shows a five-line program (pictured as a finite state machine) that represents the sequences of instructions required for {\em making an orange juice} from the {\em domotics} domain.
 We plan to evaluate our approach in synthesis and validation tasks coming from different activities in these real-world domains.
@@
 \label{sec:objectivos}
-We characterize the programs that are objects of study in this project by these three dimensions:
+We characterize the kind of the programs, that are objects of study, in this particular research project by these three dimensions:
 \begin{itemize}
-\item Number of {\bf\em program lines}.
-\item Size of the available {\bf\em instruction set}.
-\item Number and domain of {\bf\em observable program variables}. That is, the subset of program variables whose value can condition the program flow.
+\item Number of {\em program lines}.
+\item Size of the available {\em instruction set}.
+\item Number and domain of {\em observable program variables}. That is, the subset of program variables whose value can condition the program flow.
 \end{itemize}
 
-To illustrate this, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ have three program lines, an instruction set that comprises eight instructions and a single observable Boolean variable ($x=N$).
-Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em making an orange juice} have five program lines, an instruction set with four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$) and the single observable Boolean variable, ${\tt ready(juice)}$, that holds when the orange juice is recognized to be ready.
+To illustrate this characterization, the program of Figure~\ref{fig:sum} for computing $y = \sum_0^N x$ has: three program lines, an instruction set that comprises eight instructions and a single observable Boolean variable ($x=N$).
+Likewise, the program of Figure~\ref{fig:activity}, representing the activity of {\em making an orange juice} has five program lines, an instruction set with four instructions (namely $\mathsf{take}(orange)$, $\mathsf{take}(knife)$, $\mathsf{split}(orange)$ and $\mathsf{squeeze}(orange)$) and the single observable Boolean variable, ${\tt ready(juice)}$, that holds when the orange juice is recognized to be ready.
 
-The objective of this project is to study the performance of our approach (in terms of  {\em computation time} and {\em memory}) in the synthesis and validation of {\em TDD programming tasks} that come from domains of the two kinds described in the previous subsection.
+The objective of this research project is to study the performance of our approach (in terms of  {\em computation time} and {\em memory}) in the synthesis and validation of programs that come from the two kinds of domains previously described.
 This general objective is itemized into two specific objectives:
 \begin{enumerate}
-\item Analyze the performance of our Artificial Intelligence approach for the {\bf synthesis} of programs up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
-\item Analyze the performance of our Artificial Intelligence approach for the {\bf validation} of programs up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}.
+\item Analyze the performance of our Artificial Intelligence approach for the \underline{\em synthesis of programs} up to: {\bf 10 program lines}, {\bf 10 observable variables} with binary domain and, an instruction set that comprises {\bf 10 instructions}.
+\item Analyze the performance of our Artificial Intelligence approach for the \underline{\em validation of programs} up to: {\bf 15 program lines}, {\bf 15 observable variables} with binary domain and, an instruction set that comprises {\bf 15 instructions}.
 \end{enumerate}
@@
 With this regard, our AIP encoding already supports callable procedures to decompose a given programming task into simpler modules and to enable recursive solutions~\cite{sergio:aprograming:icaps16,sergio:aprograming:ijcai16,jimenez2018review,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
+\newpage
 
-\section{Workplan and Budget}
+\section{Workplan}
 \label{sec:workplan}
@@
 In this setting the {\em program synthesizer} receives an additional input: the source code of the program to validate.
-The output is a {\em validation certificate} guaranteeing that the input program either succeeds (or fails) to solve each test case in the TDD programming task.
+The output is a {\em validation certificate} guaranteeing that the input program either succeeds or fails to solve each test case of the given TDD programming task.
 \end{enumerate}
 
-Figure~\ref{fig:gantt} details the proposed 12-month timeline for the development of the project. A deliverable is provided at the end of each task.
+The proposed 12-month timeline for the development of the project (Figure~\ref{fig:gantt}) is split in three tasks. A deliverable is provided at the end of each task.
 
@@
 \ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Development (3 months)}{4}{6}\\
-\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Experiments \& dissemination (6 months)}{7}{12}
+\ganttbar[bar/.append style={line width=1pt, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804}]{Experiments \& Dissemination (6 months)}{7}{12}
 \end{ganttchart}
-\caption{\small Work-plan for developing a user-interactive program synthesizer/validator.}
+\caption{\small Work-plan for developing the user-interactive program synthesizer/validator.}
 \label{fig:gantt}
@@
 \begin{enumerate}
-\item {\bf Task1. System design (months 1-3})
+\item {\bf Task1. System design} {\tt\small (months 1-3)}
   \begin{small}
@@
 
-  \item {\bf Task2. Development of the system architecture (months 4-6})
+\item {\bf Task2. Development of the system architecture} {\tt\small (months 4-6)}
     \begin{small}
       \begin{enumerate}
-      \item The programming-into-planning compiler ({\em Compiler 1}).
+      \item Programming-into-planning compiler ({\em Compiler 1}).
 This system component parses the {\em TDD programming task} and produces an {\em AI Planning problem} encoded in the standard planning language PDDL.
-      \item The plan-into-program compiler ({\em Compiler 2}).
+      \item Plan-into-program compiler ({\em Compiler 2}).
 This system component extracts the program code and the corresponding validation certificate from the solution plan produced by an off-the-shelf AI planner.
@@
 \end{small}      
+The development of the system architecture (see Figure ~\ref{fig:architecture}) comprises the development of {\em Compiler 1} and {\em Compiler 2}.
+Our aim is reusing, as possible, the code generated by our current work in the research topic~\cite{jimenez2015computing,sergio:aprograming:icaps16,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,segovia2017generating,segovia:FSC:JAIR2018,segovia:programs:AIJ19}.
+
+{\small{\bf\em Deliverable T2:} Open repository with the source code of the system architecture ({\em Compiler 1} and {\em Compiler 2}) and the corresponding benchmarks.}
 
@@
   \tikzstyle{block} = [draw, draw=TableBlue,fill=TableBlue,fill opacity=0.6274509804, rectangle, minimum height=3em, minimum width=6em]
+\tikzstyle{block2} = [draw, draw=TableBlue,rectangle, minimum height=3em, minimum width=6em]  
 \tikzstyle{input} = [coordinate]
@@
     \node [block, right of=input, node distance=4cm] (compiler1) {Compiler 1};
-    \node [block, below of=compiler1, node distance=2cm] (planner) {AI Planner};    
+    \node [block2, below of=compiler1, node distance=2cm] (planner) {AI Planner};    
     \node [block, below of=planner, node distance=2cm] (compiler2) {Compiler 2};
@@
 
-{\small{\bf\em Deliverable T2:} Open repository with the source code of the system architecture and the corresponding benchmarks.}
-
-\item {\bf Task3. Experiments and dissemination of results (months 7-12}).
+\item {\bf Task3. Experiments and dissemination of results} {\tt\small (months 7-12)}.
    \begin{small}
@@
       \item Analysis and validation of the obtained results.
-      \item Tuning and repairing the system components, evaluation metrics and benchmarks according to the obtained results.                 
+      \item Repairing the system components, evaluation metrics and benchmarks according to the obtained results.                 
       \item Executing the system architecture in the {\em real-world benchmarks} introduced in Section~\ref{sec:evaluation}.
       \item Analysis and validation of the obtained results.         
-      \item Tuning and repairing the system components, evaluation metrics and benchmarks according to the obtained results.                 
+      \item Repairing the system components, evaluation metrics and benchmarks according to the obtained results.                 
       \end{enumerate}
@@
 \end{small}        
-{\small{\bf\em  Deliverable T3:} Final report with obtained conclusions and produced publications.}
+{\small{\bf\em  Deliverable T3:} Final report with the obtained conclusions and produced publications.}
 \end{enumerate}
 
-\subsection{Presupuesto} 
-Esta sección presenta la descripción del presupuesto, de una anualidad, para el desarrollo del proyecto:
+\subsection{Research Group}
+\label{sec:workplan}
+The group for developing this project comprises six researchers with proven research experience in AI planning.
+All these researchers are currently members of the {\em Group of Reasoning on Planning and Scheduling} (GRPS) \footnote{\tt\url http://users.dsic.upv.es/grupos/grps/}, that is part of the {\em Department of Computer Systems and Computation} of the {\em Universitat Politècnica de València}.
+In the last 10 years, the GRPS research group has led three research projects of the Spanish national research plan on {\em AI planning}.
+Further, the GRPS has developed its own AI planner, that is participating at the {\em International Planning Competition} (IPC-2018, {\tt\url http://ipc2018.bitbucket.io}).
+In addition, the Principal Investigator is a previous organizer of the {\em International Planning Competition} (in 2008), is invited speaker of the {\em summer school} of the {\em International Conference on Automated Planning and Scheduling} (2013) and is a current member of the program committee of this same conference, the main international forum for research on AI planning.
+
+\begin{table}[]
+\begin{footnotesize}  
+\begin{tabular}{llccr}
+ & {\bf Name} & {\bf Current position} & {\bf Role in the project} & {\bf Tasks}\\\hline
+1 & Dr. Sergio Jimenez  &  {\em 'Ramón y Cajal' fellow @UPV}  &  Principal Investigator & 1,2,3\\
+2 & Dr. Inmaculada Garcia  & {\em Associate professor @UPV}   &  Senior researcher & 1,3\\
+3 & Dr. Antonio Garrido & {\em Associate professor @UPV}   &  Senior researcher & 1,3\\
+4 & Dr. Oscar Sapena & {\em Associate professor @UPV}   &  Senior researcher & 1,3\\
+5 & Dr. Eliseo J. Marzal & {\em Lecturer @UPV} & Junior researcher & 2,3\\
+6 & Diego Aineto & {\em PhD candidate @UPV} & in-Training researcher  & 2,3
+\end{tabular}
+\end{footnotesize}
+\caption{\small Group for the development of the research project.}
+\label{tab:group}
+\end{table}
+
+As Table~\ref{tab:group} sumarizes, the role of the {\em senior researchers} in this project is to coollaborate in the {\em design of the system architecture} ({\bf Task1}) and in the {\em analysis and validation of the obtained results} ({\bf Task3.a.ii} and {\bf Task3.a.v}).
+The {\em Junior researcher} and the {\em in-training researcher} will participate at the {\em development of the system architecture} as well as at the development of experiments and the dissemination of the obtained results, that is tasks {\bf Task2} and {\bf Task3}.
+Last but not least, the {\em Principal Investigator} is in charge of all the project tasks ({\bf Task1}, {\bf Task2} and {\bf Task3}) and the corresponding deliverables.
+
+\subsection{Budget (Spanish only)} 
+Esta sección presenta la descripción del presupuesto, de un año de duración, para el desarrollo del proyecto:
 
@@
   \begin{tabular}{cl|r}
-    {\large\bf Prioridad} & {\large\bf Descripcion} & {\large\bf Euros} \\\hline
+    {\bf Prioridad} & {\bf Descripción} & {\bf Euros} \\\hline
 1 & {\scriptsize Difusión de las actividades del grupo}  & 4,000\\    
@@
 \end{small}
-\caption{\small Presupuesto de dos anualidades para el desarrollo del proyecto.}
+\caption{\small Resumen del presupuesto para el desarrollo del proyecto.}
 \end{table}
@@
 \begin{enumerate}
-\item {\em Difusión de las actividades del grupo}.
-Esta partida contempla las inscripciones a congresos internacionales en Inteligencia Artificial, previsiblemente los siguientes: The International Conference on Automated Planning and Scheduling (h5-index 27).
-The International Joint Conference on Artificial Intelligence (h5-index 61).
-The AAAI Conference on Artificial Intelligence (h5-index 69).
-The International Conference on Machine Learning (h5-index 113).
-Todos los h5-index calculados de acuerdo a Google Scholar.
-\item {\em Viajes, manutención y alojamiento grupo de investigación}.
+\item {\bf Difusión de las actividades del grupo}.
+Esta partida contempla las inscripciones a congresos internacionales en Inteligencia Artificial, previsiblemente los siguientes: The {\em International Conference on Automated Planning and Scheduling} (h5-index 27).
+The {\em International Joint Conference on Artificial Intelligence} (h5-index 61).
+The {\em AAAI Conference on Artificial Intelligence} (h5-index 69).
+The {\em International Conference on Machine Learning} (h5-index 113).
+Todos los h-index calculados de acuerdo a Google Scholar.
+\item {\bf Viajes, manutención y alojamiento grupo de investigación}.
 Esta partida contempla los gastos generados por la asistencia a los congresos internacionales mencionados anteriormente y la visita a grupos de investigacion relacionados con los objetivos del proyecto.
-\item {\em Viajes, manutención, alojamiento y ponencias investigadores invitados}.
-Sergio Jimenez, PI del presente proyecto, mantiene un activa colaboración con el grupo {\em A.I.
-and Autonomy Lab} de la universidad de Melbourne {\tt\small\url{https://cis.unimelb.edu.au/agentlab}} asi como con el {\em Artificial Intelligence and Machine Learning Research group} ({\tt\small\url{https://www.upf.edu/web/ai-ml}}) de la Universitat Pompeu Fabra de Barcelona.
-Esta partida contempla los gastos generados para el desarrollo de actividades conjuntas en el ámbito de la Comunitat Valenciana.
+\item {\bf Viajes, manutención, alojamiento y ponencias investigadores invitados}.
+Sergio Jiménez, PI del presente proyecto, mantiene una activa colaboración con el grupo {\em A.I.
+and Autonomy Lab} de la universidad de Melbourne ({\tt\small\url{https://cis.unimelb.edu.au/agentlab}}), con el {\em Artificial Intelligence and Machine Learning Research group} ({\tt\small\url{https://www.upf.edu/web/ai-ml}}) de la Universitat Pompeu Fabra de Barcelona y con el {\em Planning and Learning group} de la universidad Carlos III de Madrid ({\tt\small\url{http://www.plg.inf.uc3m.es/}}).
+Esta partida contempla los gastos generados para el desarrollo de actividades conjuntas relacionadas con el presente proyecto de investigación y en el ámbito de la Comunitat Valenciana.
 \end{enumerate}
 
-\section{Expected benefits of the research.}
+\section{Expected benefits of the research}
 \label{subsec:beneficios}
@@
 
+\subsubsection{Gender sensitive programming}
+
+A programming process is {\em gender sensitive} when the gender dimension is systematically integrated into everystep of the process~\cite{leduc2009guidelines}.
+From defining the problem, to identifying potential solutions, in the methodology and approach to implementing the project, in stakeholders analysis and the choice of partners, in defining the objective, outcomes, outputs, and activities, in the composition of the implementation and management team, in budgeting, in the monitoring and evaluation process, and in policy dialogue.
+
+When designing a programme, it is important to take this dimension into account.
+However it is not easy with the current lack of representation of women, minorities in the IT work forces~\cite{arnold2001global}.
+Futher, statistically the roles of women in the IT sector have  greater representation  maintaining  legacy  systems  (using languages such as) rather than engineering new systems (using languages such as Java or C++~\cite{Dattero:2004:PLG:962081.962087}.
+
+In this propoject we propose to increase the automation of the programming process (which arguably could improve the chances of developing programs less gender biased).
+However human are still required in the loop for deffining the set of input-output test cases that define the program behaviour.
+
 \vspace{0.3cm}

### 2018-10-03 doc/programming-projects/conselleria-emergente-18/AI-programming.tex
updates to the conselleria project proposal

@@
 Promising research opportunities come from the application of {\em AI Planning} to {\em program synthesis} given that, {\em program synthesis} with test cases, can be seen as the {\em program testing} dual.
-In fact, our current work on {\em program synthesis} with AIP already produced several {\bf publications at the top international conferences and journals on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16,segovia:FSC:JAIR2018,segovia:programs:AIJ19} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
+In fact, our current work on {\em program synthesis} with AIP already produced several {\bf publications at top international conferences and journals on Artificial Intelligence}~\cite{segovia2017generating,sergio:aprogramingb:ijcai16,sergio:aprograming:ijcai16,sergio:aprograming:icaps16,javi-icaps17,segovia:FSC:JAIR2018,segovia:programs:AIJ19} and is the recipient of the {\it 2016 distinguished paper award} at the International Joint Conference on Artificial Intelligence, the main international conference on {\em Artificial intelligence}.
 
-The main benefit of this project is to {\bf provide new insights into the current understanding of how AI can assist programmers in the software development}.
+The main benefit of this project is to provide {\bf new insights into the current understanding of how AI can assist programmers} in the software development.
 Automated program synthesis promises to produce huge cost savings in comparison to conventional software development and to increase the quality, reliability and modifiability of programs.
@@
 Most of the new planning algorithms are only tested within the benchmarks of the International Planing Competition~\cite{vallati:IPC:AI15}.
-This project will help to meet the computational and expressiveness limits of AI planners when addressing real-world programming tasks.
+This project will help to meet the computational and expressiveness limits of the off-the-shelf AI planners when addressing real-world programming tasks.
 \item A new {\bf evaluation methodology for the {\em Synthesis and Validation of Programs}}.
 The application of the exiting {\em AI Planning} technology to program synthesis can provide new evaluation metrics that assess how well a program covers a set of {\em input-output} test cases.
-\item The {\bf development of open software and open benchmarks for the {\em Synthesis and Validation of Programs}}.
+\item The development of {\bf open software and benchmarks for the {\em Synthesis and Validation of Programs}}.
 We strongly belief that reproducibility and open knowledge are essential to the advance of the research on computer science.
@@
 
-\subsubsection{Gender sensitive programming}
+\subsubsection*{Automated Synthesis and Validation of Programs: A gender perspective}
+Programs are {\em gender sensitive} when the gender dimension is systematically integrated into every step of the programming process~\cite{leduc2009guidelines}.
+From defining the problem, to identifying potential solutions, in the methodology and the final approach followed to eventually implement the program.
 
-A programming process is {\em gender sensitive} when the gender dimension is systematically integrated into everystep of the process~\cite{leduc2009guidelines}.
-From defining the problem, to identifying potential solutions, in the methodology and approach to implementing the project, in stakeholders analysis and the choice of partners, in defining the objective, outcomes, outputs, and activities, in the composition of the implementation and management team, in budgeting, in the monitoring and evaluation process, and in policy dialogue.
+The consideration of the gender dimension at all the steps that are comprised in a programming process is nowadays a chimera.
+Specially, given the current lack of representation of women in the IT work forces~\cite{arnold2001global}.
+In this project we propose to increase the automation of the programming process, which arguably, could improve the chances of developing programs that result less gender-biased.
 
-When designing a programme, it is important to take this dimension into account.
-However it is not easy with the current lack of representation of women, minorities in the IT work forces~\cite{arnold2001global}.
-Futher, statistically the roles of women in the IT sector have  greater representation  maintaining  legacy  systems  (using languages such as) rather than engineering new systems (using languages such as Java or C++~\cite{Dattero:2004:PLG:962081.962087}.
-
-In this propoject we propose to increase the automation of the programming process (which arguably could improve the chances of developing programs less gender biased).
-However human are still required in the loop for deffining the set of input-output test cases that define the program behaviour.
+In practice, our approach for automated program synthesis still requires humans in the loop (e.g.
+to specify the set of test cases that determine the program behavior).
+Therefore, even programs automatically synthesized can result gender-biased provided a gender-biased selection of test cases.
+This concern becomes more evident when we consider that statistically, the current roles of women in the IT sector have greater representation in maintaining {\em legacy systems} rather than in the engineering of {\em cutting-edge} systems~\cite{Dattero:2004:PLG:962081.962087}, like Artificial Intelligence systems.
 

### 2018-10-09 doc/icaps-19/model-recognition.tex
new full version icaps19 paper. Ready for reviews

@@
 \documentclass[letterpaper]{article} %DO NOT CHANGE THIS
-\usepackage{aaai18}  %Required
+\usepackage{aaai19}  %Required
 \usepackage{times}  %Required
@@
 \begin{abstract} 
-Given the partial observation of a plan execution and a set of possible planning models (models that share the same state variables but update these variables according to different action models), {\em model recognition} is the task of identifying which model in the set explains (produced) the given observation.
-The paper formalizes the {\em model recognition} task and proposes a novel method to estimate the probability of a \strips\ model to produce a partial observation of a plan execution.
+Given the partial observation of a plan execution and a set of possible planning models (models that share the same state variables but that update these variables according to different action models), {\em model recognition} is the task of identifying which model in the set explains/produced the given observation.
+The paper formalizes the {\em model recognition} task and introduces a novel method to estimate the probability of a \strips\ model to produce an observation of a plan execution.
 This method, that we called {\em model recognition as planning}, is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.
@@
 {\em Goal recognition} is a closely related task that aims identifying the goals of the observed agent.
-Plan/goal recognition is considered {\em automated planning} in reverse; while automated planning compute sequences of actions that accounts for a given goals, plan/goal recognition compute goals that account for an observed sequence of actions~\cite{geffner:book:2013}.
+Goal recognition is considered {\em automated planning} in reverse; while automated planning compute sequences of actions that accounts for a given goals, goal recognition compute goals that account for an observed sequence of actions~\cite{geffner:book:2013}.
 
-Diverse approaches has been proposed for plan recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{sukthankar2014plan}.
-{\em Plan recognition as planning} is the model-based approach for plan recognition~\cite{ramirez2012plan,ramirez2009plan}.
+Diverse approaches has been proposed for plan/goal recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{sukthankar2014plan}.
+{\em Plan recognition as planning} is the model-based approach for plan/goal recognition~\cite{ramirez2012plan,ramirez2009plan}.
 This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal of the agent, according to the observed plan execution.
@@
 This paper formalizes the {\em model recognition} task where the object to recognize is not a goal but the {\em planning model} that determines the behavior of the observed agent.
-Given a partial observation of a plan execution and a set of possible planning models (models that share the same state variables but update these variables with different action models), {\em model recognition} is the task of identifying the model in the set with the highest probability of explaining (producing) the given observation.
+Given a partial observation of a plan execution and a set of possible planning models (models that share the same state variables but that update these variables with different action models), {\em model recognition} is the task of identifying the model in the set with the highest probability of producing/explaining the given observation.
 
@@
   \end{scriptsize}  
- \caption{\small ({\em Left)} Robot navigation plan in a $5\times 5$ grid.
-{\em (Right)} 2-state automata to control that the robot only increments its x-coordinate at {\em even} rows (i.e.
-when $q_0$ holds).}
+ \caption{\small ({\em Left)} Robot navigating a $5\times 5$ grid.
+{\em (Right)} Automata controling that the robot only increments its x-coordinate at {\em even} rows (when $q_0$ holds, given that actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and the automata state).}
 \label{fig:grid-example}
@@
 altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated) and these models can determine different kinds of robot navigation.
-Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, {\em model recognition} would aim here to identify which navigation model produced that observation, despite key information is unobserved (e.g.
+Given an observation of a plan execution (like the one illustrated at Figure~\ref{fig:grid-example}) {\em model recognition} would aim here to identify which navigation model produced/explains that observation, despite key information is unobserved (e.g.
 the particular applied actions or the value of the state variables {\small\tt $q0$} and {\small\tt $q1$}).
@@
 
-The paper introduces {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
-Our method is built on top of off-the-shelf classical planning algorithms, that are used to elicit the likelihood of the observations given a candidate model.
-{\em Model recognition as planning} is robust to missing intermediate states and actions in the observed plan execution.
+The paper also introduces {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
+Our method is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.
 
-We evaluate the effectiveness of {\em model recognition as planning} with a set of \strips\ models that represent different {\em automata}.
-All of these {\em automata} are defined within the same state variables (same input {\em alphabet} and same {\em machine states}) but different {\em transition functions}.
+We evaluate the effectiveness of {\em model recognition as planning} with sets of \strips\ models that represent different {\em automata}.
+All the {\em automata} in a set are defined within the same state variables (same input {\em alphabet} and same {\em machine states}) but different {\em transition functions}.
 We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite internal machine states or actual applied transitions are unobserved.
@@
 \label{sec:background}
-This section formalizes the models for {\em classical planning} and for the {\em observation} of a plan execution.
+This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
 
@@
 
-A {\em classical planning frame} is a $\tup{F,A}$ pair, where $F$ is a set of fluents and $A$ is a set of actions.
+A {\em classical planning frame} is a $\tup{F,A}$ pair, where $F$ is a set of {\em fluents} and $A$ is a set of {\em actions}.
 The semantics of actions $a\in A$ are specified with two functions: $\rho(s,a)$ that denotes whether the action is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying the action $a$ in a state $s$.
-
-In this work we specify action semantics (the $\rho$ and $\theta$ functions) with the \strips\ model. With this regard, an action $a\in A$ is defined with:
+In this work we specify action semantics (the $\rho$ and $\theta$ functions) with the \strips\ model.
+With this regard, an action $a\in A$ is defined by:
 \begin{itemize}
@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
-A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The {\em plan length} is denoted with $|\pi|=n$ .
+A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em trajectory} $\tau(\pi,s_0)=\tup{a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+The {\em plan length} is denoted with $|\pi|=n$.
 A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.,~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$; the observation of the execution of  $\pi$ on $P$ is $\tau=\tup{a_1^o,s_1^o \ldots , a_n^o, s_m^o}$, an interleaved combination of {\small $1\leq n\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states such that:
+Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$; the observation of the execution of $\pi$ on $P$ is $\mathcal{O}(\pi,I)=\tup{a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, an interleaved combination of {\small $1\leq l\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states such that:
 \begin{itemize}
-\item Observed actions are {\em consistent} with $\pi$~\cite{ramirez2009plan}.
-This means that the sequence of observed actions $\tup{a_1^o, \ldots, a_n^o}$ in $\tau$ is a sub-sequence of the solution plan $\pi$.
-\item Observed states are a sub-sequence  of partial states {\em consistent} with the {\em state trajectory} $\tup{s_0, s_1, \ldots, s_n}$ traversed by $\pi$. 
+\item Observed actions are {\em consistent} with $\pi$~\cite{ramirez2009plan}. This means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of the solution plan $\pi$.
+\item Observed states are a sub-sequence  of partial states {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by $\pi$. 
 \end{itemize}
 
-On the one hand, the initial state $I$ is fully observed while the observed states in $\tau$ may be partial, i.e.
+On the one hand, the initial state $I$ is fully observed while the observed states in $\mathcal{O}(\pi,I)$ may be partial, i.e.
 the value of certain fluents in the intermediate states may be omitted (~$|s_i^o|\leq |F|$ for every $1\leq i\leq m$).
-On the other hand, the sequence of observed states $\tup{s_1^o, \ldots, s_m^o}$ in $\tau$ is the same sequence of states traversed by $\pi$ but certain states may also be omitted.
+On the other hand, the sequence of observed states $\tup{s_1^o, \ldots, s_m^o}$ in $\mathcal{O}(\pi,I)$ is the same sequence of states traversed by $\pi$ but certain states may also be omitted.
 Formally, $0\leq|s_i^o|$ for every $1\leq i\leq m$.
-This means that the transitions between two consecutive observed states in $\tau$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown and unbound).
-Therefore we can conclude that having $\tau$ does not implies knowing the actual length of $\pi$.
+This means that the transitions between two consecutive observed states in $\mathcal{O}(\pi,I)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown and unbound).
+Therefore we can conclude that having $\mathcal{O}(\pi,I)$ does not implies knowing the actual length of $\pi$.
 
 \begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\tau$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
+Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\pi,I)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
 \end{definition}
 
-For instance, Figure~\ref{fig:grid-example} illustrates the six-state $\Phi$-observation \{{\tt\footnotesize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\}, where $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
-This means that, for each observed state, only the value of fluents {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} is known while the value of the remaining fluents {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unknown.
+Figure~\ref{fig:grid-example} illustrates the six-state $\Phi$-observation \{{\tt\footnotesize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\}, where $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
+This means that, for each observed state, only the value of fluents {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} is known while the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unknown.
 
@@
 \label{sec:recognition}
-The {\em model recognition} task is a tuple $\tup{P,M,\tau}$ where:
+The {\em model recognition} task is a tuple $\tup{P,M,\mathcal{O}}$ where:
 \begin{itemize}
-\item $P=\tup{F,A,I,G}$ is a {\em classical planning problem} s.t. the semantics of actions $a\in A$ is unknown, i.e. the corresponding functions $\rho(s,a)$ and/or $\theta(s,a)$ are undefined.
+\item $P=\tup{F,A[\cdot],I,G}$ is a {\em classical planning problem} s.t.
+the semantics of actions $a\in A[\cdot]$ is unknown, i.e.
+the corresponding functions $\rho(s,a)$ and/or $\theta(s,a)$ are undefined.
 \item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite non-empty {\em set of models} for the actions in $A$ s.t., each model in $\mathcal{M}\in M$, defines different function pairs $\tup{\rho,\theta}$ over the state variables $F$.
-\item $\tau$ is an {\em observation} of the execution of a solution plan $\pi$ for $P$.
+\item $\mathcal{O}(\pi,I)$ is an {\em observation} of the execution of an unknown solution plan $\pi$ for the planning problem $P$.
 \end{itemize}
 
-{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model $\mathcal{M}\in M$ and the observed plan execution $\tau$ is the single example to classify.
-Further, the planning model associated to a class acts as a {\em class prototype} that summarizes all the plan executions that could be synthesized with that model (all the examples that belong to that class).
+{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model $\mathcal{M}\in M$ and the observed plan execution $\mathcal{O}(\pi,I)$ is the single example to classify.
+Further, the planning model associated to each class acts as the corresponding {\em class prototype} that summarizes all the plan executions that could be synthesized with that model (i.e.
+all the examples that belong to that class).
 
-We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\tau$ wrt the following expression, $argmax_{\mathcal{M}\in M} P(\tau|\mathcal{M})\times P(\mathcal{M})$.
-The hypotheses in {\em model recognition} are then about the possible action models $\mathcal{M}\in M$, while the observation $\tau$ of a plan execution represents the input observation.
+In this work we follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\mathcal{O}(\pi,I)$ with respect to the expression:
+\begin{align}
+argmax_{\mathcal{M}\in M} P(\mathcal{O}|\mathcal{M})\times P(\mathcal{M}).
+\end{align}
+The hypotheses in {\em model recognition} are then about the possible action models $\mathcal{M}\in M$, while the observation $\mathcal{O}(\pi,I)$ of a plan execution represents the input observation.
 The {\em solution} to the {\em model recognition} task is the model $\mathcal{M}\in M$ (or subset of models in $M$) that maximizes the previous expression.
 
-The $P(\mathcal{M})$ probability expresses if one model is known to be a priori more likely than the others.
+The $P(\mathcal{M})$ probability expresses whether one model is known to be a priori more likely than the others.
 If this probability is not given as input it is reasonable to assume that {\em a priori} all models are equiprobable.
-The challenge in our formulation for {\em model recognition} is the definition of the $P(\tau|\mathcal{M})$ {\em likelihood} that express the probability of observing $\tau$ when $\mathcal{M}$ is the planning model.
+The challenge in our formulation for {\em model recognition} is the definition of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} that expresses the probability of observing $\mathcal{O}(\pi,I)$ when $\mathcal{M}$ is the planning model.
+
+\begin{figure}
+  \begin{scriptsize}
+  \begin{center}
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
+	  \node[state] (A)              {$\mathcal{M}$};
+	  \node[state] (B) [right of=A] {$\mathcal{M}'$};
+	  \node[state] (C) [right of=B] {$\pi$};
+	  \node[state] (D) [right of=C] {$\mathcal{O}$};          
+	  \path
+(A) edge  node {} (B)
+(B) edge  node {} (C)
+(C) edge  node {} (D)
+;
+	\end{tikzpicture}
+  \end{center}
+  \end{scriptsize}  
+ \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}'$ results from editing $\mathcal{M}$ and that can produce a plan $\pi$ consistent with $\mathcal{O}(\pi,I)$.}
+\label{fig:net}
+\end{figure}
+
+Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess how far is $\mathcal{M}$ from a model $\mathcal{M'}$ that can produce a plan $\pi$ that solves $P$ and s.t.
+$\mathcal{O}(\pi,I)$ is a {\em consistent} observation of the execution of $\pi$ on $P$.
+Figure~\ref{fig:net} shows the four-variable {\em Bayesian network} abstracting this process.
+Regarding this network we have that:
+\begin{align}
+ P(\mathcal{O}|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi} P(\mathcal{M'}|\mathcal{M})P(\pi|\mathcal{M'})P(\mathcal{O}|\pi),
+\end{align}
+where $\mathcal{M}'$ ranges over all the models that can be generated editing $\mathcal{M}$ and $\pi$ ranges over all the plans that can be synthesized with a model $\mathcal{M}'$
+
+\subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
+Following the previous equation (2) the exact computation of $P(\mathcal{O}|\mathcal{M})$ is intractable because, for most planning problems, the set of {\em valid} plans consistent with an observation can easily be very large (infinite in the case of planning problems without dead-ends).
+Instead, in this work we propose to estimate $P(\mathcal{O}|\mathcal{M})$ making the following assumptions:
+\begin{enumerate}
+\item The sum in the previous equation is dominated by the {\em closest compliant set}, i.e.
+the models closest to $\mathcal{M}$ that can produce a solution plan consistent with $\mathcal{O}(\pi,I)$.
+The rationale behind this assumption is that the further the compliant model, the lower the $P(\mathcal{O}|\mathcal{M})$ likelihood.
+\end{enumerate}
+
+The {\em full observability of the executed plan} is a too strong assumption but, as we explain next, it allows us to build reasonable estimates of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the more general case where the observability of the executed plan is partial.
+
+Note that in this baseline scenario where there is {\em full observability of the executed plan}:
+\begin{itemize}
+\item There is only a single possible plan {\em consistent} with the given observation, besides $P(\mathcal{O}|\pi)=1$.
+\item If we consider assumption 1, the {\em closest compliant set} $\mathcal{M}^*$ becomes a singleton for this particular scenario, so that probabilities corresponding to different compliant models are not added up.
+Further, the $P(\pi|\mathcal{M^*})$ probability equals also to $1$ since there is only one possible plan consistent with the observation and by definition, the {\em closest compliant} model produces that plan.
+\end{itemize}
+As a consequence, when there is {\em full obsevability of the executed plan} and under the previous assumption, we have that expression (1) simplifies to
+\begin{align}
+argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}).
+\end{align}
+
+Next we show how to approximate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the more general case where there is partial observability of the executed plan.
+First we start with the simpler scenario where the length of the observed plan is bound and add these two new assumptions.
+\begin{enumerate}\addtocounter{enumi}{1}
+\item The {\em closest compliant set} is a singleton, so that probabilities corresponding to different compliant models are not added up.
+\item All solutions of a given {\em plan length} are equiprobable, with shorter plans more likely than longer plans.  
+\end{enumerate}
+
+In such an scenario $P(\mathcal{O}|\pi)$ equals to a constant $\alpha$ that indicates the number of plans of bounded length that are consistent with $\mathcal{O}$.
+Likewise, $P(\pi|\mathcal{M^*})$ is also one given the definition of the {\em closest compliant set}.
+As a consequence, when there is {\em partial obsevability of the executed plan} (but the plan length is bound), and under the previous three assumptions, we have that expression (1) simplifies to
+\begin{align}
+argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M})\times \alpha.
+\end{align}
+where $\alpha$ is a normalizing constant that has the same value for every $\mathcal{M}\in M$, so it can be ignored.
+
+Finally, we can say that the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be estimated for the general case of partial observability of the executed plan (where the length of the plan is not bound) by adding a fourth assumption:
+\begin{enumerate}\addtocounter{enumi}{3}
+\item The observed plan $\pi$ that solves $P$ is an optimal plan since we are assuming that the observed agent is acting rationally~\cite{ramirez2012plan}.
+\end{enumerate}
+In this case the length of the plan becomes bound by the optimal plan length and hence, we can again use expression (4) for the  likelihood estimation.
 
@@
 We start formalizing the \strips\ schema and define the full space of possible \strips\ schema.
-Eventually, we introduce an {\em edit distance} for \strips\ schema to estimate the $P(\tau|\mathcal{M})$ likelihoods for classical planning models.
+Eventually, we introduce an {\em edit distance} for \strips\ schema to estimate the $P(\mathcal{O}|\mathcal{M})$ likelihoods for classical planning models.
 
@@
 \strips\ action schema provide a compact representation for specifying classical planning models.
-For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema, codded in PDDL, that determine robot navigation in $n\times n$ grids.
+For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema, codded in PDDL, that determine a particular kind of robot navigation in $n\times n$ grids.
 
-A \strips\ action schema $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and three list of predicates ($pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+A \strips\ action schema $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 
 We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
-For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
+For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
 Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
@@
 Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of a schema $\xi$.
-The space of possible \strips\ schema is bound further by a set of constraints ${\mathcal C}$ of the following kinds:
+The space of possible \strips\ schema is bound further by ${\mathcal C}$, a set of constraints of the following kinds:
 \begin{itemize}
@@
 \subsection{Edit distances for \strips\ action models}
-{\em Edit distances} are similarity metrics traditionally computed over {\em strings} or {\em graphs} and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
+{\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
 In this work we formalize and compute edit distances that are referred to \strips\ planning models.
-
 First, we define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
@@
 The observation of a plan execution, generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrain further the space of the possible schema $\xi\in \mathcal{M}$.
-In this case, we talk about {\em observation constraints} (that can also be added to the $\mathcal{C}$ set).
+In this case, we talk about {\em observation constraints} (that can also be included into the $\mathcal{C}$ set).
 In addition, {\em observation constraints} allow us to define an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.
@@
 \begin{definition}[Observation edit distance]
-  Given $\tau$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, all defined within the same set of predicates $\Psi$.
-The {\bf observation edit distance}, $\delta(\mathcal{M},\tau)$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a plan $\pi_\tau$ for $P$ {\em consistent} with $\tau$; \[\delta(\mathcal{M},\tau)=\min_{\forall \mathcal{M}' \rightarrow \tau} \delta(\mathcal{M},\mathcal{M}')\]
+  Given $\mathcal{O}(\pi,I)$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, all defined within the same set of predicates $\Psi$.
+The {\bf observation edit distance}, $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
+$\mathcal{M}'$ produces a plan $\pi$ that solves $P$ and that is {\em consistent} with $\mathcal{O}(\pi,I)$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
@@
 Note that the {\em observation edit distance} could also be defined assessing the edition required by the observed plan execution to match the given model.
-This implies defining {\em edit operations} that modify $\tau$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
+This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\pi,I)$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
 Our definition of the {\em observation edit distance} is more practical since normally ${\mathcal I}_{\Psi,\xi}$ is much smaller than $F$.
@@
 Given a model $\mathcal{M}$.
-The {\bf closest compliant models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) that can produce a solution plan consistent with $\tau$;
-  \[\mathcal{M^*}=\underset{\forall \mathcal{M}' \rightarrow \tau}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
+The {\bf closest compliant models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) that can produce a solution plan consistent with $\mathcal{O}(\pi,I)$;
+  \[\mathcal{M^*}=\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
 \end{definition}
@@
 \begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-	  \node[state] (A)              {$\mathcal{M}$};
-	  \node[state] (B) [right of=A] {$\mathcal{M}'$};
-	  \node[state] (C) [right of=B] {$\pi$};
-	  \node[state] (D) [right of=C] {$\tau$};          
-	  \path
-(A) edge  node {} (B)
-(B) edge  node {} (C)
-(C) edge  node {} (D)
-;
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}  
- \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}'$ results from editing $\mathcal{M}$ and that can produce a plan $\pi$ consistent with $\tau$.}
-\label{fig:net}
-\end{figure}
-
-\subsection{Approximating the $P(\tau|\mathcal{M})$ {\em likelihood}}
-Intuitively, our approach to estimate the $P(\tau|\mathcal{M})$ {\em likelihood} is to assess how far is $\mathcal{M}$ from a model $\mathcal{M'}$ that can produce a plan $\pi$ that solves $P$ and s.t.
-$\tau$ is a {\em consistent} observation of the execution of $\pi$ on $P$.
-Figure~\ref{fig:net} shows the four-variable {\em Bayesian network} abstracting that $\mathcal{M}'$ is a model that results from editing the input model $\mathcal{M}$ s.t.
-$\mathcal{M}'$ can produce a solution plan $\pi$ consistent with $\tau$.
-Regarding this network, a natural way to formulate the aimed {\em likelihood} is
-\[
- P(\tau|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi}\frac{P(\mathcal{M},\mathcal{M}',\pi,\tau)}{P(\mathcal{M})},
-\]
-where $\mathcal{M}'$ ranges over all the models that can be generated editing $\mathcal{M}$ and $\pi$ ranges over all the plans that can be synthesized with a model $\mathcal{M}'$.
-
-Given that the joint probability distribution can be marginalized as $\displaystyle P(\mathcal{M},\mathcal{M}',\pi,\tau)=P(\mathcal{M})P(\mathcal{M'}|\mathcal{M})P(\mathcal{\pi}|\mathcal{M'})P(\tau|\mathcal{\pi})$, that $P(\tau|\mathcal{\pi})=1$ for plans consistent with $\tau$ and that $P(\tau|\mathcal{\pi})=0$ for $\tau$ inconsistent plans.
-Likewise, $P(\mathcal{\pi}|\mathcal{M'})=0$ for plans that cannot be generated with a model $\mathcal{M'}$, i.e.
-for {\em invalid plans}~\cite{howey2004val}.
-We have that
-\[
-P(\tau|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi} P(\mathcal{M'}|\mathcal{M})P(\mathcal{\pi}|\mathcal{M'}),
-\]
-where $\mathcal{M}'$ now ranges over the models that can synthesize {\em valid plans} and $\pi$ ranges only over the valid plans for $\mathcal{M}'$ consistent with $\tau$.
-
-Despite this simplification, the exact computation of $P(\tau|\mathcal{M})$ is still likely to be intractable since for most planning problems, the set of {\em valid} plans consistent with an observation can easily be very large (infinite in the case of planning problems without dead-ends).
-In this work we propose to estimate $P(\tau|\mathcal{M})$ making the following assumptions:
-\begin{enumerate}
-\item The sum in the previous equation is dominated by the {\em closest compliant set}.
-The rationale behind this assumption is that the further the compliant model, the lower the likelihood.
-Further, we assume that the {\em closest compliant set of models} is a singleton, so that probabilities corresponding to different compliant models are not added up
-\[
-P(\tau|\mathcal{M}) \approx \sum_{\pi} P(\mathcal{M^*}|\mathcal{M})P(\pi|\mathcal{M^*}),
-\]
-where $\pi$ ranges now over the valid plans consistent with $\tau$ that are valid for the closest compliant model $\mathcal{M^*}$.
-
-\item Solution plans, that can be synthesized with a given model, are equiprobable.  
-\end{enumerate}
-
-With this regard, our approximation of the $P(\tau|\mathcal{M})$ likelihood becomes proportional to $P(\mathcal{M^*}|\mathcal{M})$, which is given by the {\em observation edit distance}.
-This distance is maximum when it requires fully editing all the schema in the model while it is minimum when the given model is already able to produce the input observation without introducing any change.
-Intermediate distance values reflect how far models are from explaining the input observations.
-In more detail we approximate the $P(\tau|\mathcal{M})$ likelihood with the expression:
-\[
-P(\tau|\mathcal{M}) \approx \alpha\times (1-\frac{\delta(\mathcal{M},\tau)}{\delta(\mathcal{M},*)}),
-\]
-where $\alpha$ is a normalizing constant that reflects the amount of plans valid for $\mathcal{M^*}$ and consistent with $\tau$.
-Since the value of $\delta(\mathcal{M},*)$ is {\em a priori} known, the only missing ingredient to estimate the $P(\tau|\mathcal{M})$ likelihood is the $\delta(\mathcal{M},\tau)$ distance.
-
-\begin{figure}
 {\tt\tiny
-00 : (insert\_pre\_xcoord\_v1\_inc-x)\\
-01 : (insert\_pre\_next\_v1\_v2\_inc-x)\\
-02 : (insert\_pre\_q0\_inc-x)\\
-03 : (delete\_del\_xcoord\_v2\_inc-x)\\
-04 : (delete\_add\_xcoord\_v1\_inc-x)\\
-05 : (insert\_del\_xcoord\_v1\_inc-x)\\
-06 : (insert\_add\_xcoord\_v2\_inc-x)\\
-07 : (validate\_0)\\
-08 : (editable\_inc-x 1 2)\\
-09 : (editable\_inc-x 2 3)\\
-10 : (editable\_inc-x 3 4)\\
-11 : (editable\_inc-x 4 5)\\
-12 : (editable\_inc-y-even 1 2)\\
-13 : (editable\_dec-x 5 4)\\
-14 : (validate\_1)
+\begin{tabular}{ll}
+00 : (insert\_pre\_xcoord\_v1\_inc-x)   & 07 : (validate\_0)\\
+01 : (insert\_pre\_next\_v1\_v2\_inc-x) & 08 : (editable\_inc-x 1 2)\\
+02 : (insert\_pre\_q0\_inc-x)           & 09 : (editable\_inc-x 2 3)\\
+03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (editable\_inc-x 3 4) \\
+04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (editable\_inc-x 4 5)\\
+05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (editable\_inc-y-even 1 2)\\
+06 : (insert\_add\_xcoord\_v2\_inc-x)   & 13 : (editable\_dec-x 5 4)\\
+& 14 : (validate\_1)
+\end{tabular}
 }
@@
 \section{Model Recognition as classical planning}
-This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\tau)$ can be computed (and hence an approximation of the $P(\tau|\mathcal{M})$ likelihood) with a compilation of a classical planning with conditional effects.
+This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\mathcal{O})$ can be computed (and hence an approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood) with a compilation of a classical planning with conditional effects.
 
@@
 \begin{enumerate}
-\item Induces a solution plan $\pi_\tau$ for the original classical planning problem $P$.
-\item Validates that $\tau$ is an observation of the execution of $\pi_\tau$ on the classical planning problem $P$.
+\item Induces a solution plan $\pi$ for the original classical planning problem $P$.
+\item Validates that $\mathcal{O}(\pi,I)$ is an observation of the execution of $\pi$ on the classical planning problem $P$.
 \end{enumerate}
@@
 
-Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
-The postfix of the plan (steps [7,14]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
+Figure~\ref{fig:plan-pdistance} shows the plan with a prefix ({\em steps [0,6]}) for editing the planning model of Figure\ref{fig:model-example} when its schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
+The postfix of the plan ({\em steps [7,14]}) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
 
 Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deletions) required by $\mathcal{M}'$ to be validated.
-In the example of Figure~\ref{fig:plan-pdistance} $\delta(\mathcal{M},\tau)=7$ and $\delta(\mathcal{M},*)=6\times 3\times 10$ since there are 6 action schemes for which $|{\mathcal I}_{\Psi,\xi}|=10$.
+In the example of Figure~\ref{fig:plan-pdistance} $\delta(\mathcal{M},\mathcal{O})=7$ and $\delta(\mathcal{M},*)=6\times 3\times 10$ since there are 6 action schemes for which $|{\mathcal I}_{\Psi,\xi}|=10$.
 
@@
 \subsection{A propositional encoding for \strips\ action schema}
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with the fluents
-
-of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding objects $\Omega_\xi$.
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with the fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding objects $\Omega_\xi$.
 Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.
@@
 \subsection{The compilation formalization}
-Conditional effects allow us to compactly define our compilation for computing $\delta(\mathcal{M},\tau)$ and hence, estimate the $P(\tau|\mathcal{M})$ likelihood.
-Given a \strips\ model $\mathcal{M}\in M$ and the observation $\tau$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditional effects $P'=\tup{F',A',I',G'}$ such that:
+Conditional effects allow us to compactly define our compilation for computing $\delta(\mathcal{M},\mathcal{O})$ and hence, estimate the $P(\mathcal{O}|\mathcal{M})$ likelihood.
+Given a \strips\ model $\mathcal{M}\in M$ and the observation $\mathcal{O}(\pi,I)$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditional effects $P'=\tup{F',A',I',G'}$ such that:
 \begin{itemize}
@@
 \begin{itemize}
-\item $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to code the $i^{th}$ action in $\tau$.
-The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< n$}, are also added to iterate through the $n$ steps of $\tau$.
-\item The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\tau$ where the action model is validated.
+\item $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq l}$ to code the $i^{th}$ action in $\mathcal{O}(\pi,I)$.
+The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< l$}, are also added to iterate through the $l$ actions in $\mathcal{O}(\pi,I)$.
+\item The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\pi,I)$ where the action model is validated.
 \end{itemize}
@@
 \end{itemize}
-\item $I'$ extends the original initial state $I$ with the fluent $mode_{edit}$ set to true as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<n$}, for tracking the plan step where the action model is validated.
+\item $I'$ extends the original initial state $I$ with the fluent $mode_{edit}$ set to true as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<l$}, for tracking the plan step where the action model is validated.
 Our compilation assumes that initially $\mathcal{M}'$ is defined as $\mathcal{M}$.
@@
 when $mode_{val}$ holds).
-When the observation $\tau$ includes observed actions, they also include the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ to validate that actions are applied, exclusively, in the same order as they appear in $\tau$.\\
+When the observation $\mathcal{O}(\pi,I)$ includes observed actions, they also include the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,l]}$ to validate that actions are applied, exclusively, in the same order as they appear in $\mathcal{O}(\pi,I)$.\\
 
@@
 \end{align*}
-\end{small}
-\begin{small}
 \begin{align*}
@@
 To evaluate the empirical performance of {\em model recognition as planning} we collect a set $M$ of possible \strips\ models, that share the same state variables but define different action models.
-Then, we randomly choose one of these models $\mathcal{M}\in M$ and use it to produce an observation $\tau$ of a plan execution.
-Finally, we follow our {\em model recognition as planning} method to identify the model $\mathcal{M}\in M$ that produced $\tau$.
+Then, we randomly choose one of these models $\mathcal{M}\in M$ and use it to produce an observation $\mathcal{O}(\pi,I)$ of a plan execution.
+Finally, we follow our {\em model recognition as planning} method to identify the model $\mathcal{M}\in M$ that produced $\mathcal{O}(\pi,I)$.
 The experiment is repeated for models of different kind and different observability of the given plan execution.
 
-\subsubsection{Reproducibility}
-{\sc Madagascar} is the classical planner we used to solve the instances that result from our compilations for its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
+\subsubsection{Reproducibility.} {\sc Madagascar} is the classical planner we used to solve the instances that result from our compilations for its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
 Due to its SAT-based nature, {\sc Madagascar} can apply the actions for editing preconditions in a single planning step (in parallel) because there is no interaction between them.
@@
 
-\subsection{Recognition of {\em Regular Automatae}}
-In this experiment the models in $M$ represent different {\em regular automatae}.
+\subsubsection{Recognition of {\em Regular Automatae}.} In this experiment the models in $M$ represent different {\em regular automatae}.
 Figure~\ref{fig:regautomatae} illustrate a four-symbol four-state {\em regular automata} for recognizing the $\{(abc)^n : n \geq 1 \}$ language.
@@
 Here we assume that the actual applied transitions are unknown as well as the internal machine state.
-Assuming that the actual applied transitions is unknown means that the observation $\tau$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\tau=\tup{s_1, \ldots , s_m}$.
-Assuming that the internal machine state is unknown means that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
+Assuming that the actual applied transitions is unknown means that the observation $\mathcal{O}(\pi,I)$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\mathcal{O}(\pi,I)=\tup{s_1, \ldots , s_m}$.
+Assuming that the internal machine state is unknown means that $\mathcal{O}(\pi,I)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
 
-\subsection{Recognition of {\em Turing Machines}}
-In this experiment the given models in $M$ represent different {\em Turing machines}~\cite{bylander1994computational,porco2013automatic}.
+\subsubsection{Recognition of {\em Turing Machines}.} In this experiment the given models in $M$ represent different {\em Turing machines}~\cite{bylander1994computational,porco2013automatic}.
 Figure~\ref{fig:tautomatae} illustrate a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
@@
 
-Again we assume that the actual applied transitions is unknown ($\tau$ contains no actions).
-Further, we assume that $\tau$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt $q$()} fluents, with $q\in Q$ and $q\neq q_0$ and that the values of several tape cells is unknown.
+Again we assume that the actual applied transitions is unknown ($\mathcal{O}(\pi,I)$ contains no actions).
+Further, $\mathcal{O}(\pi,I)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt $q$()} fluents, with $q\in Q$ and $q\neq q_0$ and that the values of several tape cells is unknown.
 
@@
 
-\subsection{Recognition of navigation models}
-In this experiment the given models in $M$ represent different navigation models that are computed as the cross product of a regular automata with a four-operator \strips\ model for navigating a $n\times n$ grid.
+\subsubsection{Recognition of navigation models.} In this experiment the given models in $M$ represent different navigation models that are computed as the cross product of a regular automata with a four-operator \strips\ model for navigating a $n\times n$ grid.
 
@@
 
-\section{Related Work}
-\label{sec:related}
-In this work we do not assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
-A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
-Also  related to this paper is the work on {\em Goal Recognition Design} but in that work the action model is given in advance and fixed~\cite{keren2014goal}.
-
 \section{Conclusions}
@@
 {\em Model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unobserved.
+
+In this work we do not assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
+A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
+Also  related to this paper is the work on {\em Goal Recognition Design} but in that work the action model is given in advance and fixed~\cite{keren2014goal}.
 

### 2018-10-12 doc/icaps-19/model-recognition.tex
new full version icaps19 ready for review

@@
 \begin{abstract} 
-Given the partial observation of a plan execution and a set of possible planning models (models that share the same state variables but that update these variables according to different action models), {\em model recognition} is the task of identifying which model in the set explains/produced the given observation.
+Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but that update these variables according to different action models); {\em model recognition} is the task of identifying which model in the set produced/explains the given observation.
 The paper formalizes the {\em model recognition} task and introduces a novel method to estimate the probability of a \strips\ model to produce an observation of a plan execution.
-This method, that we called {\em model recognition as planning}, is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.
+This method, that we called {\em model recognition as planning}, is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observation of the plan execution.
 The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different kinds of {\em automata}.
-We show that {\em model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unobserved.
+We show that {\em model recognition as planning} succeeds to identify the executed automata despite some state variables (e.g.
+the internal machine state) or the actual applied transitions are unobserved.
 \end{abstract}
@@
 This paper formalizes the {\em model recognition} task where the object to recognize is not a goal but the {\em planning model} that determines the behavior of the observed agent.
-Given a partial observation of a plan execution and a set of possible planning models (models that share the same state variables but that update these variables with different action models), {\em model recognition} is the task of identifying the model in the set with the highest probability of producing/explaining the given observation.
+Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but that update these variables with different action models); {\em model recognition} is the task of identifying the model in the set with the highest probability of producing/explaining the given observation.
 
@@
  \caption{\small ({\em Left)} Robot navigating a $5\times 5$ grid.
-{\em (Right)} Automata controling that the robot only increments its x-coordinate at {\em even} rows (when $q_0$ holds, given that actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and the automata state).}
+{\em (Right)} Automata controlling that the robot only increments its x-coordinate at {\em even} rows (when $q_0$ holds, given that actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and the automata state).}
 \label{fig:grid-example}
@@
 To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model the robot could increment its {\em x-coordinate} at {\em even rows} (when {\tt\small $q0$} holds) and decrement it at the {\em odd rows} (when {\tt\small $q1$} holds).
+According to this model the robot could increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
+at {\em even rows} if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at {\em odd rows} if {\tt\small $q0$} holds initially).
 Apart from this particular navigation model, different action models could be defined within the same state variables (e.g.
@@
 Given an observation of a plan execution (like the one illustrated at Figure~\ref{fig:grid-example}) {\em model recognition} would aim here to identify which navigation model produced/explains that observation, despite key information is unobserved (e.g.
-the particular applied actions or the value of the state variables {\small\tt $q0$} and {\small\tt $q1$}).
+the particular applied actions or the value of state variables {\small\tt $q0$} and {\small\tt $q1$}).
 
@@
 In addition, it enables identifying different kinds of automata by observing their execution.
-It is well-known that diverse automatae representations, like {\em finite state controllers}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,ivankovic2015optimal,segovia2017generating}.
+It is well-known that diverse automata representations, like {\em finite state controllers}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,ivankovic2015optimal,segovia2017generating}.
 
@@
 Our method is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.
-
 We evaluate the effectiveness of {\em model recognition as planning} with sets of \strips\ models that represent different {\em automata}.
-All the {\em automata} in a set are defined within the same state variables (same input {\em alphabet} and same {\em machine states}) but different {\em transition functions}.
-We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite internal machine states or actual applied transitions are unobserved.
+All the {\em automata} in a set are defined within the same state variables but different {\em transition functions}.
+We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite some state variables (e.g.
+the internal machine state) or the actual applied transitions are unobserved.
 
@@
 A {\em classical planning frame} is a $\tup{F,A}$ pair, where $F$ is a set of {\em fluents} and $A$ is a set of {\em actions}.
-The semantics of actions $a\in A$ are specified with two functions: $\rho(s,a)$ that denotes whether the action is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying the action $a$ in a state $s$.
-In this work we specify action semantics (the $\rho$ and $\theta$ functions) with the \strips\ model.
+The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ that denotes whether the action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying the action $a$ in a state $s$.
+In this work we compactly represent the $\rho$ and $\theta$ functions following the \strips\ model.
 With this regard, an action $a\in A$ is defined by:
@@
 \subsection{Conditional effects}
-{\em Conditional effects} allow planning actions to have different semantics according to the value of the current state.
+{\em Conditional effects} allow classical planning actions to have different semantics according to the value of the current state.
 This feature is useful for compactly defining our method for {\em model recognition as planning}.
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$; the observation of the execution of $\pi$ on $P$ is $\mathcal{O}(\pi,I)=\tup{a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, an interleaved combination of {\small $1\leq l\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states such that:
+Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$; the observation of the execution of $\pi$ on $P$ is $\mathcal{O}(\pi,P)=\tup{a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, an interleaved combination of {\small $1\leq l\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states such that:
 \begin{itemize}
 \item Observed actions are {\em consistent} with $\pi$~\cite{ramirez2009plan}. This means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of the solution plan $\pi$.
-\item Observed states are a sub-sequence  of partial states {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by $\pi$. 
+\item Observed states are a sequence of partial states that is {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by the execution of $\pi$ on $P$. 
 \end{itemize}
 
-On the one hand, the initial state $I$ is fully observed while the observed states in $\mathcal{O}(\pi,I)$ may be partial, i.e.
+On the one hand, the initial state $I$ is fully observed while the observed states in $\mathcal{O}(\pi,P)$ may be partial, i.e.
 the value of certain fluents in the intermediate states may be omitted (~$|s_i^o|\leq |F|$ for every $1\leq i\leq m$).
-On the other hand, the sequence of observed states $\tup{s_1^o, \ldots, s_m^o}$ in $\mathcal{O}(\pi,I)$ is the same sequence of states traversed by $\pi$ but certain states may also be omitted.
-Formally, $0\leq|s_i^o|$ for every $1\leq i\leq m$.
-This means that the transitions between two consecutive observed states in $\mathcal{O}(\pi,I)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown and unbound).
-Therefore we can conclude that having $\mathcal{O}(\pi,I)$ does not implies knowing the actual length of $\pi$.
+On the other hand, the sequence of observed states $\tup{s_1^o, \ldots, s_m^o}$ in $\mathcal{O}(\pi,P)$ corresponds to the same sequence of states traversed by $\pi$ but certain states may also be omitted ($0\leq|s_i^o|$ for every $1\leq i\leq m$).
+This means that the transitions between two consecutive observed states in $\mathcal{O}(\pi,P)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown and unbound).
+Therefore we can conclude that having $\mathcal{O}(\pi,P)$ does not imply knowing the actual length of $\pi$.
 
 \begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\pi,I)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
+Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\pi,P)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
 \end{definition}
 
-Figure~\ref{fig:grid-example} illustrates the six-state $\Phi$-observation \{{\tt\footnotesize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\}, where $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
-This means that, for each observed state, only the value of fluents {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} is known while the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unknown.
+Figure~\ref{fig:grid-example} illustrates the six-state $\Phi$-observation \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\}.
+In this case $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} so that the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unknown.
 
@@
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$ is a {\em classical planning problem} s.t.
-the semantics of actions $a\in A[\cdot]$ is unknown, i.e.
-the corresponding functions $\rho(s,a)$ and/or $\theta(s,a)$ are undefined.
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite non-empty {\em set of models} for the actions in $A$ s.t., each model in $\mathcal{M}\in M$, defines different function pairs $\tup{\rho,\theta}$ over the state variables $F$.
-\item $\mathcal{O}(\pi,I)$ is an {\em observation} of the execution of an unknown solution plan $\pi$ for the planning problem $P$.
+\item $P=\tup{F,A[\cdot],I,G}$ is a {\em classical planning problem} s.t. the semantics of actions $a\in A[\cdot]$ is unknown (i.e. the corresponding $\rho$ and/or $\theta$ functions are undefined).
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite non-empty {\em set of models} for the actions in $A[\cdot]$ s.t., each model in $\mathcal{M}\in M$, defines different function pairs $\tup{\rho,\theta}$ over the state variables $F$.
+\item $\mathcal{O}(\pi,P)$ is an {\em observation} of the execution of an unknown solution plan $\pi$ for the planning problem $P$.
 \end{itemize}
 
-{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model $\mathcal{M}\in M$ and the observed plan execution $\mathcal{O}(\pi,I)$ is the single example to classify.
-Further, the planning model associated to each class acts as the corresponding {\em class prototype} that summarizes all the plan executions that could be synthesized with that model (i.e.
+{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model $\mathcal{M}\in M$ and the observed plan execution $\mathcal{O}(\pi,P)$, is the single example to classify.
+The planning model associated to each class acts as the corresponding {\em class prototype} and summarizes any observation of a plan execution that could be synthesized with that model (i.e.
 all the examples that belong to that class).
 
-In this work we follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\mathcal{O}(\pi,I)$ with respect to the expression:
+We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\mathcal{O}(\pi,P)$ with respect to the expression:
 \begin{align}
@@
 \end{align}
-The hypotheses in {\em model recognition} are then about the possible action models $\mathcal{M}\in M$, while the observation $\mathcal{O}(\pi,I)$ of a plan execution represents the input observation.
-The {\em solution} to the {\em model recognition} task is the model $\mathcal{M}\in M$ (or subset of models in $M$) that maximizes the previous expression.
+The hypotheses in {\em model recognition} are then about the possible action models $\mathcal{M}\in M$, while the input observation represents the observation $\mathcal{O}(\pi,P)$ of the execution of a plan $\pi$ that solves $P$.
+The {\em solution} to the {\em model recognition} task is the subset of models in $M$ that maximizes the previous expression (1).
 
+\subsection{Formulating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
 The $P(\mathcal{M})$ probability expresses whether one model is known to be a priori more likely than the others.
-If this probability is not given as input it is reasonable to assume that {\em a priori} all models are equiprobable.
-The challenge in our formulation for {\em model recognition} is the definition of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} that expresses the probability of observing $\mathcal{O}(\pi,I)$ when $\mathcal{M}$ is the planning model.
+If this probability is not given as input it is reasonable to assume that, {\em a priori}, all models are equiprobable.
+The challenge in our formulation for {\em model recognition} is the definition of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} that expresses the probability of observing $\mathcal{O}(\pi,P)$ when $\mathcal{M}$ is the planning model.
 
@@
   \end{scriptsize}  
- \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}'$ results from editing $\mathcal{M}$ and that can produce a plan $\pi$ consistent with $\mathcal{O}(\pi,I)$.}
+ \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}$ can be {\em transformed} into a model $\mathcal{M}'$ that is able to produce a plan $\pi$ consistent with observation $\mathcal{O}(\pi,P)$.}
 \label{fig:net}
@@
 
-Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess how far is $\mathcal{M}$ from a model $\mathcal{M'}$ that can produce a plan $\pi$ that solves $P$ and s.t.
-$\mathcal{O}(\pi,I)$ is a {\em consistent} observation of the execution of $\pi$ on $P$.
+Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess the cost of {\em transforming} $\mathcal{M}$ into a model $\mathcal{M'}$ that can produce a plan $\pi$ that solves $P$ and s.t.
+$\mathcal{O}(\pi,P)$ is {\em consistent} with $\pi$.
 Figure~\ref{fig:net} shows the four-variable {\em Bayesian network} abstracting this process.
-Regarding this network we have that:
+Regarding this network we have the following formulation:
 \begin{align}
@@
 \end{align}
-where $\mathcal{M}'$ ranges over all the models that can be generated editing $\mathcal{M}$ and $\pi$ ranges over all the plans that can be synthesized with a model $\mathcal{M}'$
+where $\pi$ ranges over all the plans that can be synthesized with a model $\mathcal{M}'$ and $\mathcal{M}'$ ranges over all the models that can be generated {\em transforming} $\mathcal{M}$.
 
-\subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
-Following the previous equation (2) the exact computation of $P(\mathcal{O}|\mathcal{M})$ is intractable because, for most planning problems, the set of {\em valid} plans consistent with an observation can easily be very large (infinite in the case of planning problems without dead-ends).
-Instead, in this work we propose to estimate $P(\mathcal{O}|\mathcal{M})$ making the following assumptions:
-\begin{enumerate}
-\item The sum in the previous equation is dominated by the {\em closest compliant set}, i.e.
-the models closest to $\mathcal{M}$ that can produce a solution plan consistent with $\mathcal{O}(\pi,I)$.
-The rationale behind this assumption is that the further the compliant model, the lower the $P(\mathcal{O}|\mathcal{M})$ likelihood.
-\end{enumerate}
-
-The {\em full observability of the executed plan} is a too strong assumption but, as we explain next, it allows us to build reasonable estimates of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the more general case where the observability of the executed plan is partial.
-
-Note that in this baseline scenario where there is {\em full observability of the executed plan}:
-\begin{itemize}
-\item There is only a single possible plan {\em consistent} with the given observation, besides $P(\mathcal{O}|\pi)=1$.
-\item If we consider assumption 1, the {\em closest compliant set} $\mathcal{M}^*$ becomes a singleton for this particular scenario, so that probabilities corresponding to different compliant models are not added up.
-Further, the $P(\pi|\mathcal{M^*})$ probability equals also to $1$ since there is only one possible plan consistent with the observation and by definition, the {\em closest compliant} model produces that plan.
-\end{itemize}
-As a consequence, when there is {\em full obsevability of the executed plan} and under the previous assumption, we have that expression (1) simplifies to
-\begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}).
-\end{align}
-
-Next we show how to approximate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the more general case where there is partial observability of the executed plan.
-First we start with the simpler scenario where the length of the observed plan is bound and add these two new assumptions.
-\begin{enumerate}\addtocounter{enumi}{1}
-\item The {\em closest compliant set} is a singleton, so that probabilities corresponding to different compliant models are not added up.
-\item All solutions of a given {\em plan length} are equiprobable, with shorter plans more likely than longer plans.  
-\end{enumerate}
-
-In such an scenario $P(\mathcal{O}|\pi)$ equals to a constant $\alpha$ that indicates the number of plans of bounded length that are consistent with $\mathcal{O}$.
-Likewise, $P(\pi|\mathcal{M^*})$ is also one given the definition of the {\em closest compliant set}.
-As a consequence, when there is {\em partial obsevability of the executed plan} (but the plan length is bound), and under the previous three assumptions, we have that expression (1) simplifies to
-\begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M})\times \alpha.
-\end{align}
-where $\alpha$ is a normalizing constant that has the same value for every $\mathcal{M}\in M$, so it can be ignored.
-
-Finally, we can say that the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be estimated for the general case of partial observability of the executed plan (where the length of the plan is not bound) by adding a fourth assumption:
-\begin{enumerate}\addtocounter{enumi}{3}
-\item The observed plan $\pi$ that solves $P$ is an optimal plan since we are assuming that the observed agent is acting rationally~\cite{ramirez2012plan}.
-\end{enumerate}
-In this case the length of the plan becomes bound by the optimal plan length and hence, we can again use expression (4) for the  likelihood estimation.
+Edit distances are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
+We assess the cost of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ formalizing and computing {\em edit distances} that, in this work, are referred to \strips\ planning models.
 
@@
 Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of actions (i.e.
-$\rho$ and $\theta$ functions) are specified with \strips\ action schemas.
+$\rho$ and $\theta$ functions) are specified with \strips\ action schema.
 We start formalizing the \strips\ schema and define the full space of possible \strips\ schema.
@@
 
-A \strips\ action schema $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+{\em A \strips\ action schema} $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 
 We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
-For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four opertator {\em blocksworld}~\cite{slaney2001blocks} are not.
+For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four operator {\em blocksworld}~\cite{slaney2001blocks} are not.
 Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
@@
 
-Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of a schema $\xi$.
-The space of possible \strips\ schema is bound further by ${\mathcal C}$, a set of constraints of the following kinds:
+Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a schema $\xi$, the space of possible \strips\ schema is constrained further by ${\mathcal C}$ which includes:
 \begin{itemize}
@@
 Formally, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$).
-Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schema is given by the expression, $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-For the navigation model of Figure~\ref{fig:model-example}, $2^{2\times 10}=1,048,576$ for every action schema.
+Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schema is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
+For every action schema in the navigation model of Figure~\ref{fig:model-example} then $2^{2\times 10}=1,048,576$.
 \item {\em Domain-specific constraints}.
-One can also introduce domain-specific knowledge to more precisely bound the space of possible \strips\ schema for a particular domain.
-For instance, in a {\em robot navigation} model, like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold a the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
-Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate is coding the {\em succesor} function for {\em natural numbers}.
-In this case, introducing these domain-specific constraints reduces the size of the space of possible schema to $2^{2\times 7}=16,384$ for every action schema.
+One can also introduce domain-specific knowledge to more precisely constrain the space of possible schema for a particular domain.
+For instance in a {\em robot navigation} model like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold at the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
+Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate is coding the {\em successor} function for {\em natural numbers}.
+These domain-specific constraints reduces the size of the space of possible schema to $2^{2\times 7}=16,384$ for every action schema.
 \end{itemize}
 
-Now we are ready to define what is a {\em well-defined} \strips\ action schema.
 \begin{definition}[Well-defined \strips\ action schema]
@@
 \end{definition}
-We say that an action model is {\em well-defined} if all its \strips\ action schema are {\em well-defined}.
+We say a classical planning action model is {\em well-defined} if all its corresponding  \strips\ action schema are {\em well-defined}.
 
 \subsection{Edit distances for \strips\ action models}
-{\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
-In this work we formalize and compute edit distances that are referred to \strips\ planning models.
 First, we define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
@@
 
-The observation of a plan execution, generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrain further the space of the possible schema $\xi\in \mathcal{M}$.
+An observation $\mathcal{O}(\pi,P)$ of the execution of a plan generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrains further the space of the possible schema $\xi\in \mathcal{M}$.
 In this case, we talk about {\em observation constraints} (that can also be included into the $\mathcal{C}$ set).
-In addition, {\em observation constraints} allow us to define an edit distance to asses the matching of a \strips\ model with respect to an observation of a plan execution.
+In addition, {\em observation constraints} allow us to define an edit distance to elicit the $P(\mathcal{O}|\mathcal{M})$ likelihoods.
+It can be argued that the shorter this  distance the better the given model explains the given observation.
 
 \begin{definition}[Observation edit distance]
-  Given $\mathcal{O}(\pi,I)$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, all defined within the same set of predicates $\Psi$.
+  Given $\mathcal{O}(\pi,P)$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, all defined within the same set of predicates $\Psi$.
 The {\bf observation edit distance}, $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a plan $\pi$ that solves $P$ and that is {\em consistent} with $\mathcal{O}(\pi,I)$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
+$\mathcal{M}'$ produces a plan $\pi$ that solves $P$ and that is {\em consistent} with $\mathcal{O}(\pi,P)$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
 
-Remarkably, the {\em observation edit distance} allows us to elicit the likelihood of the observations given a candidate model.
-It can be argued that the shorter this  distance the better the given model explains the given observation.
 Note that the {\em observation edit distance} could also be defined assessing the edition required by the observed plan execution to match the given model.
-This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\pi,I)$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
+This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\pi,P)$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
 Our definition of the {\em observation edit distance} is more practical since normally ${\mathcal I}_{\Psi,\xi}$ is much smaller than $F$.
@@
 
-\begin{definition}[{\em Closest compliant models}] \label{compliant}
+\begin{definition}[{\em Closest consistent models}] \label{consistent}
 Given a model $\mathcal{M}$.
-The {\bf closest compliant models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) that can produce a solution plan consistent with $\mathcal{O}(\pi,I)$;
-  \[\mathcal{M^*}=\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
+The {\bf closest consistent models} is the comparable set $M^*$ of action models $\mathcal{M}'$, closest to $\mathcal{M}$ in terms of editions, and that can produce a solution plan consistent with $\mathcal{O}(\pi,P)$;
+  \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
 \end{definition}
+
+\subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
+Following the previous equation (2) the exact computation of $P(\mathcal{O}|\mathcal{M})$ is intractable because, for most planning problems, the set of {\em valid} plans consistent with an observation can easily be huge (infinite in the case of planning problems without dead-ends).
+Moreover the number of models that can be generated {\em editing} a given classical planning model explodes combinatorially.
+
+Instead, in this work we propose to estimate $P(\mathcal{O}|\mathcal{M})$ making the following assumptions:
+\begin{enumerate}
+\item The sum in the previous equation (2) is dominated by the {\em closest consistent set} $M^*$, i.e.
+the {\em closest} models to $\mathcal{M}$ (at the lowest {\em edit distance}) that can produce a solution plan {\em consistent} with $\mathcal{O}(\pi,P)$.
+The rationale behind this assumption is that the further the consistent model, the lower the $P(\mathcal{O}|\mathcal{M})$ likelihood.
+\end{enumerate}
+
+\subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to build reasonable estimates of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the general case.
+In this baseline scenario:
+\begin{itemize}
+\item There is only a single possible plan $\pi$ {\em consistent} with the given observation, besides $P(\mathcal{O}|\pi)=1$.
+\item If we consider assumption 1, the {\em closest consistent set} becomes a singleton $\mathcal{M}^*$ for this particular scenario of full observability, so that probabilities corresponding to different consistent models are not added up.
+Further, $P(\pi|\mathcal{M^*})=1$ since, by definition, the {\em closest consistent model} produces $\pi$ consistent with $\mathcal{O}(\pi,P)$.
+\end{itemize}
+As a consequence, when there is {\em full observability of the executed plan} and under the previous assumption, expression (1) simplifies to:
+\begin{align}
+argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}).
+\end{align}
+
+\subsubsection{Partial observability of the executed plan.} A similar approximation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be built for the general case where the executed plan is partially observed.
+We start again with a simpler scenario where, there is partial observability, but the length of the observed plan is bound and we add these two new assumptions.
+\begin{enumerate}\addtocounter{enumi}{1}
+\item All solutions of a given {\em plan length} are equiprobable, with shorter plans more likely than longer plans.
+Therefore $P(\mathcal{O}|\pi)$ equals to a normalizing constant indicating the number of bounded length plans consistent with $\mathcal{O}(\pi,P)$.
+\item The {\em closest consistent set} is a singleton, so that probabilities corresponding to different consistent models are not added up.
+Again, because of the definition of the {\em closest consistent set} $P(\pi|\mathcal{M^*})=1$.
+\end{enumerate}
+
+As a consequence, when there is {\em partial observability of the executed plan} (but the plan length is bound), and under the previous three assumptions, expression (1) simplifies to
+\begin{align}
+argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M})\times \gamma.
+\end{align}
+where $\gamma$ is a normalizing constant that has the same value for every $\mathcal{M}\in M$, so it can be ignored.
+
+Finally we introduce a fourth assumption to provide an estimate of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} when there is partial observability of the executed plan and the length of the observed plan is not bound:
+\begin{enumerate}\addtocounter{enumi}{3}
+\item The observed plan that solves $P$ and is consistent with the given observation is an optimal plan $\pi^*$, i.e.
+we are assuming that the observed agent is acting rationally~\cite{ramirez2012plan}.
+\end{enumerate}
+Under this fourth assumption the length of plans becomes bound by the optimal plan length and again, expression (1) can be simplified to (4).
+
+\subsubsection{The $P(\mathcal{M^*}|\mathcal{M})$ probability distribution.} In this work $P(\mathcal{M'}|\mathcal{M})$ indicates the probability of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ by exclusively using the two {\em edit operations} previously defined, {\em deletion} and {\em insertion}.
+
+Our approach to formulate the $P(\mathcal{M'}|\mathcal{M})$ probability distribution is to map the distance $\delta(\mathcal{M},\mathcal{M'})$ according to a {\em Beta-binomial distribution}~\cite{wilcox1981review}:
+\[
+P(\mathcal{M'}|\mathcal{M})=Beta(X=\delta(\mathcal{M},\mathcal{M'}),\alpha,\beta).
+\]
+The Beta-binomial is a discrete probability distributions frequently used in {\em Bayesian} statistics with two parameters, $\alpha$ and $\beta$, that allows us to take into account the distribution of the number of models that are at a fixed distance of a given model.
+
+Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ according to a {\em Beta-binomial distribution} but in this case, mapping the observation distance $\delta(\mathcal{M},\mathcal{O})$.
+\[
+P(\mathcal{M^*}|\mathcal{M})=Beta(X=\delta(\mathcal{M},\mathcal{O}),\alpha,\beta).
+\]
 
@@
 }
- \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-14]) the planning model of Figure\ref{fig:model-example} when action {\tt\small inc-x} has no preconditions and positive/negative effects are swapped wrt Figure\ref{fig:model-example}.}
+ \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-14]) the model of Figure\ref{fig:model-example} when schema {\tt\small inc-x} has no {\em preconditions} and positive/negative {\em effects} are swapped wrt Figure\ref{fig:model-example}.}
 \label{fig:plan-pdistance}
@@
 \section{Model Recognition as classical planning}
-This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\mathcal{O})$ can be computed (and hence an approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood) with a compilation of a classical planning with conditional effects.
+This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\mathcal{O})$ can be computed (and hence an approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood) with a a classical planning compilation.
 
@@
 \item Induces a solution plan $\pi$ for the original classical planning problem $P$.
-\item Validates that $\mathcal{O}(\pi,I)$ is an observation of the execution of $\pi$ on the classical planning problem $P$.
+\item Validates that $\mathcal{O}(\pi,P)$ is an observation of the execution of $\pi$ on the classical planning problem $P$.
 \end{enumerate}
@@
 
-Figure~\ref{fig:plan-pdistance} shows the plan with a prefix ({\em steps [0,6]}) for editing the planning model of Figure\ref{fig:model-example} when its schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
-The postfix of the plan ({\em steps [7,14]}) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
-
+Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
+The postfix of the plan (steps [7,14]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
 Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deletions) required by $\mathcal{M}'$ to be validated.
-In the example of Figure~\ref{fig:plan-pdistance} $\delta(\mathcal{M},\mathcal{O})=7$ and $\delta(\mathcal{M},*)=6\times 3\times 10$ since there are 6 action schemes for which $|{\mathcal I}_{\Psi,\xi}|=10$.
+In the example of Figure~\ref{fig:plan-pdistance}, $\delta(\mathcal{M},\mathcal{O})=7$.
 
@@
 \subsection{A propositional encoding for \strips\ action schema}
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with the fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding objects $\Omega_\xi$.
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
 Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.
@@
 The interest of having a propositional encoding for \strips\ action schema is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}.
-Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents on the current state.
+Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state.
 Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
@@
 
-Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable version} of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (see Figure~\ref{fig:model-example}).
+Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}).
 Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
@@
 Conditional effects allow us to compactly define our compilation for computing $\delta(\mathcal{M},\mathcal{O})$ and hence, estimate the $P(\mathcal{O}|\mathcal{M})$ likelihood.
-Given a \strips\ model $\mathcal{M}\in M$ and the observation $\mathcal{O}(\pi,I)$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditional effects $P'=\tup{F',A',I',G'}$ such that:
+Given a \strips\ model $\mathcal{M}\in M$ and the observation $\mathcal{O}(\pi,P)$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditional effects $P'=\tup{F',A',I',G'}$:
 \begin{itemize}
@@
 \begin{itemize}
-\item $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq l}$ to code the $i^{th}$ action in $\mathcal{O}(\pi,I)$.
-The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< l$}, are also added to iterate through the $l$ actions in $\mathcal{O}(\pi,I)$.
-\item The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\pi,I)$ where the action model is validated.
+\item $F_{\pi}=\{plan(a_i,i)\}_{\small 1\leq i\leq l}$ to code the $i^{th}$ action in $\mathcal{O}(\pi,P)$.
+The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< l$}, are also added to iterate through the $l$ actions in $\mathcal{O}(\pi,P)$.
+\item The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\pi,P)$ where the action model is validated.
 \end{itemize}
-\item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schemas are edited or validated.
+\item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schema are edited or validated.
 \end{itemize}
@@
 \begin{enumerate}
-\item The {\em editable} version of the original actions in $A$.
+\item The {\em editable} version of the original actions given by $\mathcal{M}$.
 This actions have now an extra preconditions because they can only be applied in the {\em validation} mode (i.e.
 when $mode_{val}$ holds).
-When the observation $\mathcal{O}(\pi,I)$ includes observed actions, they also include the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,l]}$ to validate that actions are applied, exclusively, in the same order as they appear in $\mathcal{O}(\pi,I)$.\\
+When the observation $\mathcal{O}(\pi,P)$ includes observed actions, they also include the extra conditional effects $\{at_{i},plan(a_i,i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,l]}$ to validate that actions are applied, exclusively, in the same order as they appear in $\mathcal{O}(\pi,P)$.\\
 
@@
 To evaluate the empirical performance of {\em model recognition as planning} we collect a set $M$ of possible \strips\ models, that share the same state variables but define different action models.
-Then, we randomly choose one of these models $\mathcal{M}\in M$ and use it to produce an observation $\mathcal{O}(\pi,I)$ of a plan execution.
-Finally, we follow our {\em model recognition as planning} method to identify the model $\mathcal{M}\in M$ that produced $\mathcal{O}(\pi,I)$.
+Then, we randomly choose one of these models $\mathcal{M}\in M$ and use it to produce an observation $\mathcal{O}(\pi,P)$ of a plan execution.
+Finally, we follow our {\em model recognition as planning} method to identify the model $\mathcal{M}\in M$ that produced $\mathcal{O}(\pi,P)$.
 The experiment is repeated for models of different kind and different observability of the given plan execution.
@@
 
-\subsubsection{Recognition of {\em Regular Automatae}.} In this experiment the models in $M$ represent different {\em regular automatae}.
+\subsubsection{Recognition of {\em Regular Automata}.} In this experiment the models in $M$ represent different {\em regular automata}.
 Figure~\ref{fig:regautomatae} illustrate a four-symbol four-state {\em regular automata} for recognizing the $\{(abc)^n : n \geq 1 \}$ language.
@@
 
-In more detail, we randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{50}\}$ set of fifty different \strips\ models that encode different {\em regular automata}.
-Each $\mathcal{M}\in M$ encodes a different five-state {\em regular automata} with a five-symbol input alphabet.
+We generated a $M$ set of different \strips\ models that encode different {\em regular automata}.
+Each $\mathcal{M}\in M$ encodes a different {\em regular automata}.
 Each automata transition is encoded with a planning action.
-For instance, the execution of the {\em regular automatae} defined in Figure~\ref{fig:tautomatae}, with the sequence of input symbols $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
+For instance, the execution of the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, with the sequence of input symbols $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
 
 Here we assume that the actual applied transitions are unknown as well as the internal machine state.
-Assuming that the actual applied transitions is unknown means that the observation $\mathcal{O}(\pi,I)$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\mathcal{O}(\pi,I)=\tup{s_1, \ldots , s_m}$.
-Assuming that the internal machine state is unknown means that $\mathcal{O}(\pi,I)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
-
-\subsubsection{Recognition of {\em Turing Machines}.} In this experiment the given models in $M$ represent different {\em Turing machines}~\cite{bylander1994computational,porco2013automatic}.
-Figure~\ref{fig:tautomatae} illustrate a seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.
-The {\em tape alphabet} is $\Sigma=\{a,b,c,x,y,z,\square\}$, the {\em input alphabet} $\Upsilon=\{a,b,c,\square\}$ and the machine states are $Q=\{q_0,q_1,q_2,q_3,q_4,\underline{q_5}\}$ (where \underline{$q_5$} is the only acceptor state).
-
-Here we randomly generated a $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_{50}\}$ set of fifty different \strips\ models that encode different five-symbol five-state {\em Turing Machines} with circular tapes.
-There is a planning action $a\in A$ for each machine transition, e.g.
-the full encoding of the {\em Turing Machine} of Figure~\ref{fig:tautomatae} produces a total of sixteen \strips\ action schema.
-With this regard, the execution of the {\em Turing Machine} defined in Figure~\ref{fig:tautomatae} with an initial tape $abc\square\square\square$ produces the eight-action plan {\small $(a,q_0\rightarrow x,r,q_1)$, $(b,q_1\rightarrow y,r,q_2)$, $(c,q_2\rightarrow z,l,q_3)$, $(y,q_3\rightarrow y,l,q_3)$, $(x,q_3\rightarrow x,r,q_0)$, $(y,q_0\rightarrow y,r,q_4)$, $(z,q_4\rightarrow z,r,q_4)$, $(\square,q_4\rightarrow \square,r,\underline{q_5})$}.
-
-Again we assume that the actual applied transitions is unknown ($\mathcal{O}(\pi,I)$ contains no actions).
-Further, $\mathcal{O}(\pi,I)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt $q$()} fluents, with $q\in Q$ and $q\neq q_0$ and that the values of several tape cells is unknown.
-
-\begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.8cm,semithick]
-	  \node[state] (A)              {$q_0$};
-	  \node[state] (B) [right of=A] {$q_1$};
-	  \node[state] (C) [right of=B] {$q_2$};
-	  \node[state] (D) [right of=C] {$q_3$};
-	  \node[state] (E) [below of=A] {$q_4$};
-	  \node[state] (F) [right of=E] {$q_5$};
-
-	  \path
-(A) edge  node {a/x,right} (B)
-(B) edge [align=left, loop above] node {a/a,right\\y/y,right} (B)
-(B) edge  node {b/y,right} (C)
-(C) edge [align=left, loop above] node {z/z,right\\b/b,right} (C)
-(C) edge  node {c/z,left} (D)
-(D) edge [align=left, loop right] node {a/a,left\\b/b,left\\y/y,left\\z/z,left} (D)
-(D) edge [bend left=45] node {x/x,right} (A)
-
-(A) edge  node {y/y,right} (E)
- (E) edge [align=left, loop left] node {y/y,right\\z/z,right} (E)
-(E) edge  node {b/b,right} (F)
-;
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}  
- \caption{\small Seven-symbol six-state {\em Turing Machine} for recognizing the $\{a^nb^nc^n : n \geq 1 \}$ language.}
-\label{fig:tautomatae}
-\end{figure}
+Assuming that the actual applied transitions is unknown means that the observation $\mathcal{O}(\pi,P)$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\mathcal{O}(\pi,P)=\tup{s_1, \ldots , s_m}$.
+Assuming that the internal machine state is unknown means that $\mathcal{O}(\pi,P)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
 
@@
 
-In this work we do not assume that the observed agent is acting rationally, like in {\em plan recognition as planning}~\cite{ramirez2012plan,ramirez2009plan}.
+Previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
+Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
 A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
@@
 
-Previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
-Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
-
-Remarkably, the extension of this piece of work to the FOND planning setting~\cite{muise2012improved} is straightforward by simply considering the {\em all-outcomes} determiniztion of the actions with non-determinitic effects~\cite{yoon2007ff}.
+%Remarkably, the extension of this piece of work to the FOND planning setting~\cite{muise2012improved} is straightforward by simply considering the {\em all-outcomes} determiniztion of the actions with non-determinitic effects~\cite{yoon2007ff}.
 An interesting research direction is however to understand how to apply our approach to planning models where the planning models include actions with probabilistic effects~\cite{younes2005first}.

### 2018-10-16 doc/icaps-19/model-recognition.tex
added last discussion on likelihoods estimation

@@
 We show that {\em model recognition as planning} succeeds to identify the executed automata despite some state variables (e.g.
-the internal machine state) or the actual applied transitions are unobserved.
+the internal automata state) or the actual applied transitions are unobserved.
 \end{abstract}
@@
 Diverse approaches has been proposed for plan/goal recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{sukthankar2014plan}.
-{\em Plan recognition as planning} is the model-based approach for plan/goal recognition~\cite{ramirez2012plan,ramirez2009plan}.
-This approach assumes that the action model of the observed agent is known and leverages it to compute the most likely goal of the agent, according to the observed plan execution.
+{\em Plan recognition as planning} is the model-based approach for plan/goal recognition that leverages the action model of the observed agent to compute its most likely goal~\cite{ramirez2012plan,ramirez2009plan}.
 
-This paper formalizes the {\em model recognition} task where the object to recognize is not a goal but the {\em planning model} that determines the behavior of the observed agent.
+This paper formalizes the {\em model recognition} task where the object to recognize is not a goal but the {\em planning model} that is determining the behavior of the observed agent.
 Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but that update these variables with different action models); {\em model recognition} is the task of identifying the model in the set with the highest probability of producing/explaining the given observation.
@@
  \caption{\small ({\em Left)} Robot navigating a $5\times 5$ grid.
-{\em (Right)} Automata controlling that the robot only increments its x-coordinate at {\em even} rows (when $q_0$ holds, given that actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and the automata state).}
+{\em (Right)} Automata to control that the robot only increments its x-coordinate when $q_0$ holds (actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and switch the automata state).}
 \label{fig:grid-example}
@@
 Given an observation of a plan execution (like the one illustrated at Figure~\ref{fig:grid-example}) {\em model recognition} would aim here to identify which navigation model produced/explains that observation, despite key information is unobserved (e.g.
-the particular applied actions or the value of state variables {\small\tt $q0$} and {\small\tt $q1$}).
+the applied actions or the value of state variables {\small\tt $q0$} and {\small\tt $q1$}).
 
@@
 We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite some state variables (e.g.
-the internal machine state) or the actual applied transitions are unobserved.
+the internal automata state) or the actual applied transitions are unobserved.
 
@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
-A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em trajectory} $\tau(\pi,s_0)=\tup{a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em trajectory} $\tau(\pi,P)=\tup{a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
 The {\em plan length} is denoted with $|\pi|=n$.
@@
 \item Observed actions are {\em consistent} with $\pi$~\cite{ramirez2009plan}. This means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of the solution plan $\pi$.
-\item Observed states are a sequence of partial states that is {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by the execution of $\pi$ on $P$. 
+\item Observed states are a sequence of partial states $\tup{s_1^o, \ldots, s_m^o}$ that is {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by the execution of $\pi$ on $P$.
 \end{itemize}
@@
 {\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model $\mathcal{M}\in M$ and the observed plan execution $\mathcal{O}(\pi,P)$, is the single example to classify.
-The planning model associated to each class acts as the corresponding {\em class prototype} and summarizes any observation of a plan execution that could be synthesized with that model (i.e.
+The planning model associated to each class acts as the corresponding {\em class prototype} and it summarizes any observation of a plan execution that could be synthesized with that model (i.e.
 all the examples that belong to that class).
@@
 \end{align}
-The hypotheses in {\em model recognition} are then about the possible action models $\mathcal{M}\in M$, while the input observation represents the observation $\mathcal{O}(\pi,P)$ of the execution of a plan $\pi$ that solves $P$.
-The {\em solution} to the {\em model recognition} task is the subset of models in $M$ that maximizes the previous expression (1).
+The {\em solution} to the {\em model recognition} task is the subset of models in $M$ that maximizes expression (1).
 
@@
 	  \node[state] (B) [right of=A] {$\mathcal{M}'$};
-	  \node[state] (C) [right of=B] {$\pi$};
+	  \node[state] (C) [right of=B] {$\tau$};
 	  \node[state] (D) [right of=C] {$\mathcal{O}$};          
@@
   \end{scriptsize}  
- \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}$ can be {\em transformed} into a model $\mathcal{M}'$ that is able to produce a plan $\pi$ consistent with observation $\mathcal{O}(\pi,P)$.}
+ \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}$ can be {\em transformed} into a model $\mathcal{M}'$ that is able to produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is consistent with the observation $\mathcal{O}(\pi,P)$.}
 \label{fig:net}
@@
 
-Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess the cost of {\em transforming} $\mathcal{M}$ into a model $\mathcal{M'}$ that can produce a plan $\pi$ that solves $P$ and s.t.
-$\mathcal{O}(\pi,P)$ is {\em consistent} with $\pi$.
+Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess the cost of {\em transforming} $\mathcal{M}$ into a model $\mathcal{M'}$ that can produce a trajectory $\tau(\pi,P)$ that that reaches the goals in $P$ and is consistent with the observation $\mathcal{O}(\pi,P)$.
 Figure~\ref{fig:net} shows the four-variable {\em Bayesian network} abstracting this process.
@@
 \begin{align}
- P(\mathcal{O}|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\pi} P(\mathcal{M'}|\mathcal{M})P(\pi|\mathcal{M'})P(\mathcal{O}|\pi),
+ P(\mathcal{O}|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\tau} P(\mathcal{M'}|\mathcal{M})P(\tau|\mathcal{M'})P(\mathcal{O}|\tau),
 \end{align}
-where $\pi$ ranges over all the plans that can be synthesized with a model $\mathcal{M}'$ and $\mathcal{M}'$ ranges over all the models that can be generated {\em transforming} $\mathcal{M}$.
+where $\tau$ ranges over all the trajectories that can be synthesized with a model $\mathcal{M}'$ and $\mathcal{M}'$ ranges over all the models that can be generated {\em transforming} $\mathcal{M}$.
 
-Edit distances are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
+{\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
 We assess the cost of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ formalizing and computing {\em edit distances} that, in this work, are referred to \strips\ planning models.
@@
 \strips\ action schema provide a compact representation for specifying classical planning models.
-For instance, Figure~\ref{fig:model-example} shows six \strips\ action schema, codded in PDDL, that determine a particular kind of robot navigation in $n\times n$ grids.
+Figure~\ref{fig:model-example} shows six \strips\ action schema, codded in PDDL, that determine a particular kind of robot navigation in $n\times n$ grids.
 
@@
 Given a \strips\ action schema $\xi$ and a set of {\em predicates} $\Psi$ that shape the propositional state variables.
-The set of FOL interpretations of $\Psi$ over the corresponding $\Omega_\xi$ objects (the {\em variable names} for schema $\xi$), confines the elements that can appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
+The set of FOL interpretations of $\Psi$ over the corresponding $\Omega_\xi$ objects ({\em variable names} for schema $\xi$), confines the elements that can appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
 We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
@@
 
-Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a schema $\xi$, the space of possible \strips\ schema is constrained further by ${\mathcal C}$ which includes:
+Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a schema $\xi$, the space of possible \strips\ schema is constrained by the set ${\mathcal C}$ that includes:
 \begin{itemize}
@@
 \item {\em Domain-specific constraints}.
-One can also introduce domain-specific knowledge to more precisely constrain the space of possible schema for a particular domain.
+One can also introduce domain-specific knowledge to more precisely constrain the space of possible schema.
 For instance in a {\em robot navigation} model like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold at the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
 Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate is coding the {\em successor} function for {\em natural numbers}.
-These domain-specific constraints reduces the size of the space of possible schema to $2^{2\times 7}=16,384$ for every action schema.
+These domain-specific constraints reduces the size of the space of possible action schema to $2^{2\times 7}=16,384$ (for every schema).
 \end{itemize}
@@
 Since ${\mathcal I}_{\Psi,\xi}$ are bound sets, the maximum number of edits that can be introduced to a given action model is bound as well.
-\begin{definition}[Maximum edit distance]
-The \textbf{maximum edit distance} of an \strips\ model $\mathcal{M}$ built within the set of predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$.
-\end{definition}
+The \textbf{maximum edit distance} of a \strips\ model $\mathcal{M}$ built within predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$ (note that, if we consider the syntactic \strips\ constraints, $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2\times|{\mathcal I}_{\Psi,\xi}|$).
 
 An observation $\mathcal{O}(\pi,P)$ of the execution of a plan generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrains further the space of the possible schema $\xi\in \mathcal{M}$.
-In this case, we talk about {\em observation constraints} (that can also be included into the $\mathcal{C}$ set).
+In this case, we talk about a third kind of constraints that we call {\em observation constraints} and that can also be included into the $\mathcal{C}$ set.
 In addition, {\em observation constraints} allow us to define an edit distance to elicit the $P(\mathcal{O}|\mathcal{M})$ likelihoods.
@@
 \begin{definition}[Observation edit distance]
-  Given $\mathcal{O}(\pi,P)$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, all defined within the same set of predicates $\Psi$.
+  Given $\mathcal{O}(\pi,P)$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, defined within the same predicates $\Psi$.
 The {\bf observation edit distance}, $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a plan $\pi$ that solves $P$ and that is {\em consistent} with $\mathcal{O}(\pi,P)$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
+$\mathcal{M}'$ produces a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\pi,P)$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
@@
 Given a model $\mathcal{M}$.
-The {\bf closest consistent models} is the comparable set $M^*$ of action models $\mathcal{M}'$, closest to $\mathcal{M}$ in terms of editions, and that can produce a solution plan consistent with $\mathcal{O}(\pi,P)$;
+The set $M^*$ of {\bf closest consistent models} is the set of action models $\mathcal{M}'$, closest to $\mathcal{M}$ in terms of editions, and that can produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\pi,P)$;
   \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
@@
 \subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
-Following the previous equation (2) the exact computation of $P(\mathcal{O}|\mathcal{M})$ is intractable because, for most planning problems, the set of {\em valid} plans consistent with an observation can easily be huge (infinite in the case of planning problems without dead-ends).
-Moreover the number of models that can be generated {\em editing} a given classical planning model explodes combinatorially.
+Following equation (2), the exact computation of $P(\mathcal{O}|\mathcal{M})$ is intractable.
+For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge (infinite in the case of planning problems without dead-ends).
+Even worse, the number of models $\mathcal{M}'$ that can be generated {\em editing} a given classical planning model $\mathcal{M}$ explodes combinatorially.
 
-Instead, in this work we propose to estimate $P(\mathcal{O}|\mathcal{M})$ making the following assumptions:
-\begin{enumerate}
-\item The sum in the previous equation (2) is dominated by the {\em closest consistent set} $M^*$, i.e.
-the {\em closest} models to $\mathcal{M}$ (at the lowest {\em edit distance}) that can produce a solution plan {\em consistent} with $\mathcal{O}(\pi,P)$.
-The rationale behind this assumption is that the further the consistent model, the lower the $P(\mathcal{O}|\mathcal{M})$ likelihood.
-\end{enumerate}
+Instead, our approach is to formulate an effectively estimate of $P(\mathcal{O}|\mathcal{M})$.
 
 \subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to build reasonable estimates of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the general case.
-In this baseline scenario:
-\begin{itemize}
-\item There is only a single possible plan $\pi$ {\em consistent} with the given observation, besides $P(\mathcal{O}|\pi)=1$.
-\item If we consider assumption 1, the {\em closest consistent set} becomes a singleton $\mathcal{M}^*$ for this particular scenario of full observability, so that probabilities corresponding to different consistent models are not added up.
-Further, $P(\pi|\mathcal{M^*})=1$ since, by definition, the {\em closest consistent model} produces $\pi$ consistent with $\mathcal{O}(\pi,P)$.
-\end{itemize}
-As a consequence, when there is {\em full observability of the executed plan} and under the previous assumption, expression (1) simplifies to:
+
+In this baseline scenario there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the given observation (because of full observability), so $P(\mathcal{O}|\tau^*)=1$.
+Further, there is also a single model $M^*$ that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory).
+
+If there is a single possible trajectory and a single possible model $\mathcal{M}^*$ consistent with the input observation then, the probabilities of expression (1) are not added up so expression (1) simplifies to:
 \begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}).
+argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}) \times P(\mathcal{M}).
 \end{align}
+Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because is independent of the model $\mathcal{M}\in M$.
 
-\subsubsection{Partial observability of the executed plan.} A similar approximation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be built for the general case where the executed plan is partially observed.
-We start again with a simpler scenario where, there is partial observability, but the length of the observed plan is bound and we add these two new assumptions.
-\begin{enumerate}\addtocounter{enumi}{1}
-\item All solutions of a given {\em plan length} are equiprobable, with shorter plans more likely than longer plans.
-Therefore $P(\mathcal{O}|\pi)$ equals to a normalizing constant indicating the number of bounded length plans consistent with $\mathcal{O}(\pi,P)$.
-\item The {\em closest consistent set} is a singleton, so that probabilities corresponding to different consistent models are not added up.
-Again, because of the definition of the {\em closest consistent set} $P(\pi|\mathcal{M^*})=1$.
+\subsubsection{Partial observability of the executed plan.} In a similar way, an approximation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be built for the general case where the executed plan is partially observed.
+To deal with the new scenario we add the following two assumptions:
+\begin{enumerate}
+\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state.
+\item {\em The best plans for $P$ are unique or have different cost}.  
 \end{enumerate}
+Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}.
+They allow us to claim that $\sum_{\tau}$ in equation (2) is dominated by its largest term, so that probabilities corresponding to different trajectories that reach the goals of $P$ (and are consistent with the given observation) are not added up.
+Likewise, $P(\mathcal{O}|\tau*)=\alpha$, where $\tau^*$ is the single trajectory that corresponds to the single optimal plan $\pi^*$ for $P$ (consistent with $\mathcal{O}$ and $\alpha$ is a normalizing constant referred to $|\pi^*|$ (the number of possible observations of a bounded trajectory is bounded too).
 
-As a consequence, when there is {\em partial observability of the executed plan} (but the plan length is bound), and under the previous three assumptions, expression (1) simplifies to
+Again, there is also a single model $\mathcal{M^*}$ that can produce that single trajectory $\tau*$.
+As a consequence, when there is {\em partial observability of the executed plan}, and under the previous assumptions, expression (1) simplifies again to expression (3)
+
+\subsubsection{The $P(\mathcal{M^*}|\mathcal{M})$ probability distribution.} $P(\mathcal{M'}|\mathcal{M})$ indicates the probability of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ by exclusively using the two {\em edit operations} previously defined, {\em deletion} and {\em insertion}.
+
+We know from {\em pattern recognition} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}.
+Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
 \begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M})\times \gamma.
+CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d(1-p)^{N-d}
 \end{align}
-where $\gamma$ is a normalizing constant that has the same value for every $\mathcal{M}\in M$, so it can be ignored.
+where $N$ is the length of the string, and $p$ is $\frac{3}{4}$.
 
-Finally we introduce a fourth assumption to provide an estimate of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} when there is partial observability of the executed plan and the length of the observed plan is not bound:
-\begin{enumerate}\addtocounter{enumi}{3}
-\item The observed plan that solves $P$ and is consistent with the given observation is an optimal plan $\pi^*$, i.e.
-we are assuming that the observed agent is acting rationally~\cite{ramirez2012plan}.
-\end{enumerate}
-Under this fourth assumption the length of plans becomes bound by the optimal plan length and again, expression (1) can be simplified to (4).
-
-\subsubsection{The $P(\mathcal{M^*}|\mathcal{M})$ probability distribution.} In this work $P(\mathcal{M'}|\mathcal{M})$ indicates the probability of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ by exclusively using the two {\em edit operations} previously defined, {\em deletion} and {\em insertion}.
-
-Our approach to formulate the $P(\mathcal{M'}|\mathcal{M})$ probability distribution is to map the distance $\delta(\mathcal{M},\mathcal{M'})$ according to a {\em Beta-binomial distribution}~\cite{wilcox1981review}:
-\[
-P(\mathcal{M'}|\mathcal{M})=Beta(X=\delta(\mathcal{M},\mathcal{M'}),\alpha,\beta).
-\]
-The Beta-binomial is a discrete probability distributions frequently used in {\em Bayesian} statistics with two parameters, $\alpha$ and $\beta$, that allows us to take into account the distribution of the number of models that are at a fixed distance of a given model.
-
-Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ according to a {\em Beta-binomial distribution} but in this case, mapping the observation distance $\delta(\mathcal{M},\mathcal{O})$.
-\[
-P(\mathcal{M^*}|\mathcal{M})=Beta(X=\delta(\mathcal{M},\mathcal{O}),\alpha,\beta).
-\]
+With this regard, and considering that \strips\ models $\mathcal{M}\in M$ can be encoded with a propositional representation of fixed length $N$, we formulate the $P(\mathcal{M'}|\mathcal{M})$ probability distribution mapping the distance $\delta(\mathcal{M},\mathcal{M'})$ according to equation:
+\begin{align}
+P(\mathcal{M'}|\mathcal{M}) = 0.75^d \times 0.25^{N-d}
+\end{align}
+where $d=\delta(\mathcal{M},\mathcal{M}')$.
+Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation but now, the parameter $d$ is given by the {\em observation distance}, $d=\delta(\mathcal{M},\mathcal{O})$.
 

### 2018-10-17 doc/icaps-19/model-recognition.tex
added comments by Diego

@@
 
-\subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to build reasonable estimates of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the general case.
+\subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to understand how to build a reasonable estimate of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the general case.
 
 In this baseline scenario there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the given observation (because of full observability), so $P(\mathcal{O}|\tau^*)=1$.
-Further, there is also a single model $M^*$ that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory).
+Further, there is also a single model $M^*$ that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory so they can be considered the same model).
 
-If there is a single possible trajectory and a single possible model $\mathcal{M}^*$ consistent with the input observation then, the probabilities of expression (1) are not added up so expression (1) simplifies to:
+If there is a single possible trajectory and a single possible model $\mathcal{M}^*$ consistent with the input observation then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
 \begin{align}
@@
 \subsubsection{Partial observability of the executed plan.} In a similar way, an approximation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be built for the general case where the executed plan is partially observed.
-To deal with the new scenario we add the following two assumptions:
+To deal with this general scenario we add the following two assumptions:
 \begin{enumerate}
@@
 Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}.
-They allow us to claim that $\sum_{\tau}$ in equation (2) is dominated by its largest term, so that probabilities corresponding to different trajectories that reach the goals of $P$ (and are consistent with the given observation) are not added up.
-Likewise, $P(\mathcal{O}|\tau*)=\alpha$, where $\tau^*$ is the single trajectory that corresponds to the single optimal plan $\pi^*$ for $P$ (consistent with $\mathcal{O}$ and $\alpha$ is a normalizing constant referred to $|\pi^*|$ (the number of possible observations of a bounded trajectory is bounded too).
+They allow us to claim that the sum in equation (2) is dominated by its largest term; the term corresponding to the shortest trajectory $\tau^*$ (which is assumed to be unique) consistent with the observation, so other terms in the sum are not added up.
+Again, there is a single model $M^*$ that can produce $\tau^*$.
+With this regard, both $P(O|\tau^*)$ and $P(\tau^*|\mathcal{M}^*)$ are independent of $\mathcal{M} \in M$ so equation (1) simplifies once again to equation (3) under the previous assumptions.
 
-Again, there is also a single model $\mathcal{M^*}$ that can produce that single trajectory $\tau*$.
-As a consequence, when there is {\em partial observability of the executed plan}, and under the previous assumptions, expression (1) simplifies again to expression (3)
+Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more reasonable to assume that there is only one plan consistent with the given observation.
 
@@
 \begin{align}
-CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d(1-p)^{N-d}
+CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  \times p^d \times (1-p)^{N-d}
 \end{align}
-where $N$ is the length of the string, and $p$ is $\frac{3}{4}$.
+where $N$ is the length of the string, and $p<0.5$ since the cost of applying an {\em edit operation} is higher than do not applying it.
 
@@
 \begin{align}
-P(\mathcal{M'}|\mathcal{M}) = 0.75^d \times 0.25^{N-d}
+P(\mathcal{M'}|\mathcal{M}) = p^d \times (1-p)^{N-d}
 \end{align}
 where $d=\delta(\mathcal{M},\mathcal{M}')$.
-Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation but now, the parameter $d$ is given by the {\em observation distance}, $d=\delta(\mathcal{M},\mathcal{O})$.
+Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation (5) but instead, the parameter $d$ is given by the {\em observation distance}, $d=\delta(\mathcal{M},\mathcal{O})$.
 
-\begin{figure}
-{\tt\tiny
-\begin{tabular}{ll}
-00 : (insert\_pre\_xcoord\_v1\_inc-x)   & 07 : (validate\_0)\\
-01 : (insert\_pre\_next\_v1\_v2\_inc-x) & 08 : (editable\_inc-x 1 2)\\
-02 : (insert\_pre\_q0\_inc-x)           & 09 : (editable\_inc-x 2 3)\\
-03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (editable\_inc-x 3 4) \\
-04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (editable\_inc-x 4 5)\\
-05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (editable\_inc-y-even 1 2)\\
-06 : (insert\_add\_xcoord\_v2\_inc-x)   & 13 : (editable\_dec-x 5 4)\\
-& 14 : (validate\_1)
-\end{tabular}
-}
- \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-14]) the model of Figure\ref{fig:model-example} when schema {\tt\small inc-x} has no {\em preconditions} and positive/negative {\em effects} are swapped wrt Figure\ref{fig:model-example}.}
-\label{fig:plan-pdistance}
-\end{figure}
+In other words, we are modeling the edition of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every event $P(X=\top)=p$ and $P(X=\perp)=1-p$.
 
@@
 In the example of Figure~\ref{fig:plan-pdistance}, $\delta(\mathcal{M},\mathcal{O})=7$.
+
+\begin{figure}
+{\tt\tiny
+\begin{tabular}{ll}
+00 : (insert\_pre\_xcoord\_v1\_inc-x)   & 07 : (validate\_0)\\
+01 : (insert\_pre\_next\_v1\_v2\_inc-x) & 08 : (editable\_inc-x 1 2)\\
+02 : (insert\_pre\_q0\_inc-x)           & 09 : (editable\_inc-x 2 3)\\
+03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (editable\_inc-x 3 4) \\
+04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (editable\_inc-x 4 5)\\
+05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (editable\_inc-y-even 1 2)\\
+06 : (insert\_add\_xcoord\_v2\_inc-x)   & 13 : (editable\_dec-x 5 4)\\
+& 14 : (validate\_1)
+\end{tabular}
+}
+ \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-14]) the model of Figure\ref{fig:model-example} when schema {\tt\small inc-x} has no {\em preconditions} and positive/negative {\em effects} are swapped wrt Figure\ref{fig:model-example}.}
+\label{fig:plan-pdistance}
+\end{figure}
+
+\subsection{A propositional encoding for \strips\ action schema}
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
+Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.
+
+The interest of having a propositional encoding for \strips\ action schema is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}.
+Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state.
+Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
+\begin{small}  
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
+\cond(\mathsf{editable_{\xi}})=&\{del\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+&\{add\_e\_\xi\}\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
+\end{align*}
+\end{small}
+
+Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}).
+Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
 
@@
 \label{fig:encoding}
-\end{figure}
-
-\subsection{A propositional encoding for \strips\ action schema}
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
-Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.
-
-The interest of having a propositional encoding for \strips\ action schema is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}.
-Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state.
-Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
-\begin{small}  
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
-\cond(\mathsf{editable_{\xi}})=&\{del\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
-&\{add\_e\_\xi\}\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
-\end{align*}
-\end{small}
-
-Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}).
-Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
-
-\begin{figure}
-  \begin{tiny}  
-  \begin{verbatim}
-(:action editable_inc-x
-  :parameters (?v1 ?v2)
-  :precondition
-    (and (or (not (pre_xcoord_v1_inc-x)) (xcoord ?v1))
-         (or (not (pre_xcoord_v2_inc-x)) (xcoord ?v2))
-         (or (not (pre_ycoord_v1_inc-x)) (xcoord ?v1))                       
-         (or (not (pre_ycoord_v2_inc-x)) (xcoord ?v2))
-         (or (not (pre_q0__inc-x)) (q0))
-         (or (not (pre_q1__inc-x)) (q1)))
-         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
-         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
-         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
-         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2))))
-    :effect (and
-       (when (del_xcoord_v1_inc-x) (not (xcoord ?v1)))
-       (when (del_xcoord_v2_inc-x) (not (xcoord ?v2)))
-       (when (del_ycoord_v1_inc-x) (not (xcoord ?v1)))
-       (when (del_ycoord_v2_inc-x) (not (xcoord ?v2)))
-       (when (del_q0__inc-x) (not (q0)))
-       (when (del_q1__inc-x) (not (q1)))
-       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
-       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
-       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
-       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))
-
-       (when (add_xcoord_v1_inc-x) (xcoord ?v1))
-       (when (add_xcoord_v2_inc-x) (xcoord ?v2))
-       (when (add_ycoord_v1_inc-x) (xcoord ?v1))
-       (when (add_ycoord_v2_inc-x) (xcoord ?v2))
-       (when (add_q0__inc-x) (q0))
-       (when (add_q1__inc-x) (q1))
-       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
-       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
-       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
-       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
-  \end{verbatim}           
-  \end{tiny}  
- \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
-\label{fig:editable}
 \end{figure}
@@
 
+\begin{figure}
+  \begin{tiny}  
+  \begin{verbatim}
+(:action editable_inc-x
+  :parameters (?v1 ?v2)
+  :precondition
+    (and (or (not (pre_xcoord_v1_inc-x)) (xcoord ?v1))
+         (or (not (pre_xcoord_v2_inc-x)) (xcoord ?v2))
+         (or (not (pre_ycoord_v1_inc-x)) (xcoord ?v1))                       
+         (or (not (pre_ycoord_v2_inc-x)) (xcoord ?v2))
+         (or (not (pre_q0__inc-x)) (q0))
+         (or (not (pre_q1__inc-x)) (q1)))
+         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
+         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
+         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
+         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2))))
+    :effect (and
+       (when (del_xcoord_v1_inc-x) (not (xcoord ?v1)))
+       (when (del_xcoord_v2_inc-x) (not (xcoord ?v2)))
+       (when (del_ycoord_v1_inc-x) (not (xcoord ?v1)))
+       (when (del_ycoord_v2_inc-x) (not (xcoord ?v2)))
+       (when (del_q0__inc-x) (not (q0)))
+       (when (del_q1__inc-x) (not (q1)))
+       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
+       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
+       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
+       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))
+
+       (when (add_xcoord_v1_inc-x) (xcoord ?v1))
+       (when (add_xcoord_v2_inc-x) (xcoord ?v2))
+       (when (add_ycoord_v1_inc-x) (xcoord ?v1))
+       (when (add_ycoord_v2_inc-x) (xcoord ?v2))
+       (when (add_q0__inc-x) (q0))
+       (when (add_q1__inc-x) (q1))
+       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
+       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
+       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
+       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
+  \end{verbatim}           
+  \end{tiny}  
+ \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
+\label{fig:editable}
+\end{figure}
+
 \section{Evaluation}
@@
 The {\em input alphabet} is $\Sigma=\{a,b,c,\square\}$, and the machine states are $Q=\{q_0,q_1,q_2,\underline{q_3}\}$ (where \underline{$q_3$} is the only acceptor state).
+
+We generated a $M$ set of different \strips\ models that encode different {\em regular automata}.
+Each $\mathcal{M}\in M$ encodes a different {\em regular automata}.
+Each automata transition is encoded with a planning action.
+For instance, the execution of the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, with the sequence of input symbols $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
+
+Here we assume that the actual applied transitions are unknown as well as the internal machine state.
+Assuming that the actual applied transitions is unknown means that the observation $\mathcal{O}(\pi,P)$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\mathcal{O}(\pi,P)=\tup{s_1, \ldots , s_m}$.
+Assuming that the internal machine state is unknown means that $\mathcal{O}(\pi,P)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
 
@@
 \end{figure}
-
-We generated a $M$ set of different \strips\ models that encode different {\em regular automata}.
-Each $\mathcal{M}\in M$ encodes a different {\em regular automata}.
-Each automata transition is encoded with a planning action.
-For instance, the execution of the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, with the sequence of input symbols $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
-
-Here we assume that the actual applied transitions are unknown as well as the internal machine state.
-Assuming that the actual applied transitions is unknown means that the observation $\mathcal{O}(\pi,P)$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\mathcal{O}(\pi,P)=\tup{s_1, \ldots , s_m}$.
-Assuming that the internal machine state is unknown means that $\mathcal{O}(\pi,P)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
 

### 2018-10-24 benchmarks/navigation-models/gen-navigation-domain.py
added navigation domains and generators

@@ -10,19 +10,17 @@ try:
     pre_inc_horizontal = int(sys.argv[2])
     eff_inc_horizontal = int(sys.argv[3])
     pre_dec_horizontal = int(sys.argv[4])
-    eff_dec_horizontal = int(sys.argv[5])
-    pre_inc_vertical = int(sys.argv[6])
-    eff_inc_vertical = int(sys.argv[7])
-    pre_dec_vertical = int(sys.argv[8])
-    eff_dec_vertical = int(sys.argv[9])
+    pre_inc_vertical = int(sys.argv[5])
+    eff_inc_vertical = int(sys.argv[6])
+    pre_dec_vertical = int(sys.argv[7])
     
 except:
     print "Usage:"
-    print sys.argv[0] + " <nStates> <pre-inc-horizontal [-1, nStates]> <eff-inc-horizontal [-1,0,1]>  <pre-dec-horizontal [-1, nStates]> <eff-dec-horizontal [-1,0,1]> <pre-inc-vertical [-1, nStates]>  <eff-inc-vertical [-1,0,1]> <pre-dec-vertical [-1, nStates]>  <eff-dec-vertical [-1,0,1]> "    
+    print sys.argv[0] + " <nStates> <pre-inc-horizontal [-3, nStates]> <eff-inc-horizontal [-1,0,1]>  <pre-dec-horizontal [-3, nStates]>  <pre-inc-vertical [-3, nStates]>  <eff-inc-vertical [-1,0,1]> <pre-dec-vertical [-3, nStates]>  "    
     sys.exit(-1)
 
 
-str_id = sys.argv[1]+"-"+sys.argv[2]+"-"+sys.argv[3]+"-"+sys.argv[4]+"-"+sys.argv[5]+"-"+sys.argv[6]+"-"+sys.argv[7]+"-"+sys.argv[8]+"-"+sys.argv[9]
+str_id = sys.argv[1]+"-"+sys.argv[2]+"-"+sys.argv[3]+"-"+sys.argv[4]+"-"+sys.argv[5]+"-"+sys.argv[6]+"-"+sys.argv[7]
     
 str_out=""
 str_out = str_out +  "(define (domain grid-navigation)    ;;; Parameters: "+str_id+"\n"
@@ -30,6 +28,8 @@ str_out = str_out +  "  (:requirements :typing :strips)\n"
 str_out = str_out +  "  (:types value - object)\n"
 str_out = str_out +  "  (:predicates (xcoord ?v - value)\n"
[…]

### 2018-10-24 benchmarks/navigation-models/gen-navigation-problem.py

@@ -25,9 +25,9 @@ str_out = str_out +  " - value)\n"
 
 str_istate="S"+str(0)
 str_ihead="t"+str(0)
-str_out = str_out +  "  (:init (xcoord v0) (ycoord v0)\n        "
-
-
+str_out = str_out +  "  (:init (xcoord v0) (ycoord v0)\n"
+str_out = str_out +  "         (min v0) (max v" + str(nSize-1) + ")\n"
+str_out = str_out +  "        "
 for n in range(nSize-1):
    str_out = str_out +  " (next v"+str(n)+" v"+str(n+1)+")"
 

### 2018-10-24 src/FAMA_FSM.py
FAMA_FSM.py: fixes to allow loops in automata

@@ -126,8 +126,8 @@ except:
     sys.exit(-1)
 
 
-# trace_filter = ["symbolo1", "symbolo2", "symbolo3"]
-trace_filter = ["head"]
+trace_filter = ["symbolo0", "symbolo1", "symbolo2", "next", "head"]
+# trace_filter = ["head"]
 pres_filter = []
 adds_filter = ['states0', 'states1', 'states2', 'states3', 'states4']
 dels_filter = []
@@ -243,6 +243,8 @@ for a in actions:
     eff = list()
 
     known_effects = list(a.effects)
+    del_state_effects = [e for e in known_effects if e.literal.predicate in adds_filter and e.literal.negated]
+    known_effects = [e for e in known_effects if e.literal.predicate not in adds_filter or not e.literal.negated]
     for known_effect in known_effects:
         if not known_effect.literal.negated:
             eff += [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom(known_effect.literal.predicate, ["?o" + str(original_params.index(arg) + 1) for arg in known_effect.literal.args]))]
@@ -290,6 +292,10 @@ for a in actions:
                     eff = eff + [
                     pddl.effects.Effect([], condition, pddl.conditions.Atom(p.name, ["?o" + str(t) for t in tup]))]
 
+    for del_state_effect in del_state_effects:
+        condition = pddl.conditions.Conjunction([pddl.conditions.NegatedAtom("add_" + "_".join([del_state_effect.literal.predicate] + [a.name] ), [])])
+        eff = eff + [
+            pddl.effects.Effect([], condition, pddl.conditions.NegatedAtom(del_state_effect.literal.predicate, []))]
 
     learning_task.actions.append(pddl.actions.Action(a.name, params, len(params), pddl.conditions.Conjunction(pre), eff, 0))
 
@@ -414,6 +420,12 @@ for j in range(len(traces)):
                 for atom in new_atoms:
                     eff += [pddl.effects.Effect([], pddl.conditions.Truth(),
                                                 pddl.conditions.Atom(atom.predicate, atom.args))]
+
+                reset_state = [p for p in filtered_predicates if p.name not in [atom.predicate for atom in new_atoms]]
+                for p in reset_state:
+                    eff += [pddl.effects.Effect([], pddl.conditions.Truth(),
+                                                pddl.conditions.NegatedAtom(p.name, []))]
+
 
 
 

### 2018-10-24 benchmarks/automata-generator/gen-automata-domain.py
gen-automata-domain.py: -1 for empty domain

@@ -7,6 +7,8 @@ def decodeID(n, b):
     q = n%b
     if n == 0:
         return '0'
+    elif n == -1:
+        return '0'
     elif e == 0:
         return str(q)
     else:
@@ -17,6 +19,15 @@ def decodeID(n, b):
 # MAIN
 # **************************************#
 try:
+
+    if "-o" in sys.argv:
+        index = sys.argv.index("-o")
+        observations_file = sys.argv[index+1]
+        sys.argv.remove("-o")
+        sys.argv.remove(observations_file)
+    else:
+        observations_file = None
+
     nStates  = int(sys.argv[1])
     nObs = int(sys.argv[2])
     nKind = int(sys.argv[3])
@@ -24,8 +35,25 @@ try:
 
 except:
     print "Usage:"
-    print sys.argv[0] + " <nStates> <nObservations> <kind (0 Regular, 1 HMM, 2 Turing Machine)> <idAutomata>"
+    print sys.argv[0] + " <nStates> <nObservations> <kind (0 Regular, 1 HMM, 2 Turing Machine)> <idAutomata> [-o observations_file]"
     sys.exit(-1)
+
+
+observations = set()
+if observations_file:
+    for line in open(observations_file):
+        obs = line.strip().split(',')
+        obs = [o.replace('-', '').replace(' ', '') for o in obs]
+
+        observations.update(obs)
+
+else:
+    for i in range(nObs):
+        observations.add('symbol' + str(i))
+
+observations = list(observations)
+
+
 
 str_ID = ""    
 if nKind == config.REGULAR:
@@ -59,9 +87,11 @@ for i in range(nStates):
    str_istate="S"+str(i)
    str_out = str_out +  " (state"+str_istate+")"
 str_out = str_out + "\n              "
-for j in range(nObs):
-   str_iobs="O"+str(j)
-   str_out = str_out +  " (symbol"+str_iobs+" ?x)"
+# for j in range(nObs):
+#    str_iobs="O"+str(j)
+#    str_out = str_out +  " (symbol"+str_iobs+" ?x)"
+for i in range(len(observations)):
+    str_out += " ({} ?x)".format(observations[i])
 
 if nKind == config.HMM:
     str_out = str_out + "\n              "    
@@ -75,20 +105,24 @@ set_index_obs = set([])
 set_index_state = set([])
 
 for i in range(nStates):
-   for j in range(nObs):
+   for j in range(len(observations)):
       str_istate="S"+str(i)
[…]

### 2018-10-24 doc/icaps-19/model-recognition.tex

@@
     :effect (and (not (ycoord ?y1)) (ycoord ?y2)
-                 (not (q0)) (q1))
+                 (not (q0)) (q1)))
 
@@
     :parameters (?y1 ?y2)
-    :precondition (and (ycoord ?y1) (next ?y1 ?y2) (q1)))
+    :precondition (and (ycoord ?y1) (next ?y1 ?y2) (q1))
     :effect (and (not (ycoord ?y1)) (ycoord ?y2)

### 2018-10-30 doc/icaps-19/model-recognition.tex
first icaps19 paper review to address new comment by Miquel

@@
 
-This paper formalizes the {\em model recognition} task where the object to recognize is not a goal but the {\em planning model} that is determining the behavior of the observed agent.
+This paper formalizes the {\em model recognition} task where the object to recognize is not a goal but the {\em planning model} that shapes the behavior of the observed agent.
 Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but that update these variables with different action models); {\em model recognition} is the task of identifying the model in the set with the highest probability of producing/explaining the given observation.
@@
 To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model the robot could increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
+According to this model, the robot can increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
 at {\em even rows} if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at {\em odd rows} if {\tt\small $q0$} holds initially).
@@
 altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated) and these models can determine different kinds of robot navigation.
-Given an observation of a plan execution (like the one illustrated at Figure~\ref{fig:grid-example}) {\em model recognition} would aim here to identify which navigation model produced/explains that observation, despite key information is unobserved (e.g.
+Given an observation of a plan execution (like the one illustrated at Figure~\ref{fig:grid-example}) {\em model recognition} aims to identify the action model that produced/explains that observation, despite relevant information is unobserved (e.g.
 the applied actions or the value of state variables {\small\tt $q0$} and {\small\tt $q1$}).
@@
 In addition, it enables identifying different kinds of automata by observing their execution.
-It is well-known that diverse automata representations, like {\em finite state controllers}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,ivankovic2015optimal,segovia2017generating}.
+It is well-known that diverse automata representations, like {\em finite state controllers}, {\em B\"{u}chi automata}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,patrizi2011computing,ivankovic2015optimal,segovia2017generating}.
 
@@
   (:action inc-y-even
-    :parameters (?y1 ?y2)
-    :precondition (and (ycoord ?y1) (next ?y1 ?y2) (q0))
-    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
+    :parameters (?v1 ?v2)
+    :precondition (and (ycoord ?v1) (next ?v1 ?v2) (q0))
+    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
                  (not (q0)) (q1)))
@@
   (:action inc-y-odd
-    :parameters (?y1 ?y2)
-    :precondition (and (ycoord ?y1) (next ?y1 ?y2) (q1))
-    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
+    :parameters (?v1 ?v2)
+    :precondition (and (ycoord ?v1) (next ?v1 ?v2) (q1))
+    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
                  (not (q1)) (q0)))
@@
   (:action dec-y-even
-    :parameters (?y1 ?y2)
-    :precondition (and (ycoord ?y1) (next ?y2 ?y1) (q0))
-    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
+    :parameters (?v1 ?v2)
+    :precondition (and (ycoord ?v1) (next ?v2 ?v1) (q0))
+    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
                  (not (q0)) (q1)))
@@
   (:action dec-y-odd
-    :parameters (?y1 ?y2)
-    :precondition (and (ycoord ?y1) (next ?y2 ?y1) (q1))
-    :effect (and (not (ycoord ?y1)) (ycoord ?y2)
+    :parameters (?v1 ?v2)
+    :precondition (and (ycoord ?v1) (next ?v2 ?v1) (q1))
+    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
                  (not (q1)) (q0)))
@@
   \end{tiny}  
- \caption{\small Example of a \strips\ action model (codded in PDDL) for robot navigation in a $n\times n$ grid.}
+ \caption{\small Example of a \strips\ action model (given in PDDL) for robot navigation in a $n\times n$ grid.}
 \label{fig:model-example}
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$ that solves $P$; the observation of the execution of $\pi$ on $P$ is $\mathcal{O}(\pi,P)=\tup{a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, an interleaved combination of {\small $1\leq l\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states such that:
+Given a classical planning problem $P=\tup{F,A,I,G}$ and a trajectory $\tau(\pi,P)$ generated by the execution of a plan $\pi$ that solves $P$; the observation of the execution of $\pi$ on $P$ is $\mathcal{O}(\tau)=\tup{a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, an interleaved combination of {\small $1\leq l\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states such that:
 \begin{itemize}
-\item Observed actions are {\em consistent} with $\pi$~\cite{ramirez2009plan}. This means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of the solution plan $\pi$.
-\item Observed states are a sequence of partial states $\tup{s_1^o, \ldots, s_m^o}$ that is {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by the execution of $\pi$ on $P$.
+\item Observed actions are {\em consistent} with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$. 
+\item Observed states are a sequence of {\em partial states} $\tup{s_1^o, \ldots, s_m^o}$ that is {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by the trajectory $\tau(\pi,P)$.
+The initial state $I$ is fully observed while the observed states in $\mathcal{O}(\tau)$ may be partial states, i.e.
+the value of certain fluents in the intermediate states may not be observed (~$|s_i^o|\leq |F|$ for every $1\leq i\leq m$).
+Certain states may even be fully unobserved ($0\leq|s_i^o|$ for every $1\leq i\leq m$).
 \end{itemize}
 
-On the one hand, the initial state $I$ is fully observed while the observed states in $\mathcal{O}(\pi,P)$ may be partial, i.e.
-the value of certain fluents in the intermediate states may be omitted (~$|s_i^o|\leq |F|$ for every $1\leq i\leq m$).
-On the other hand, the sequence of observed states $\tup{s_1^o, \ldots, s_m^o}$ in $\mathcal{O}(\pi,P)$ corresponds to the same sequence of states traversed by $\pi$ but certain states may also be omitted ($0\leq|s_i^o|$ for every $1\leq i\leq m$).
-This means that the transitions between two consecutive observed states in $\mathcal{O}(\pi,P)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown and unbound).
-Therefore we can conclude that having $\mathcal{O}(\pi,P)$ does not imply knowing the actual length of $\pi$.
+With this regard, we are assuming that a bijective monotone mapping exists between plans and observations~\cite{ramirez2009plan}.
+We can then  conclude that transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite).
+In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
+
+Figure~\ref{fig:grid-example} illustrates the {\em partial observation} of a six-state trajectory \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\}.
+In more detail, this observation only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} because the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, cannot be observed.
 
 \begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\pi,P)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
+Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
 \end{definition}
-
-Figure~\ref{fig:grid-example} illustrates the six-state $\Phi$-observation \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\}.
-In this case $\Phi$ only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} so that the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unknown.
 
@@
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$ is a {\em classical planning problem} s.t. the semantics of actions $a\in A[\cdot]$ is unknown (i.e. the corresponding $\rho$ and/or $\theta$ functions are undefined).
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_m\}$ is a finite non-empty {\em set of models} for the actions in $A[\cdot]$ s.t., each model in $\mathcal{M}\in M$, defines different function pairs $\tup{\rho,\theta}$ over the state variables $F$.
-\item $\mathcal{O}(\pi,P)$ is an {\em observation} of the execution of an unknown solution plan $\pi$ for the planning problem $P$.
+\item $P=\tup{F,A[\cdot],I,G}$ is a {\em classical planning problem} s.t.
+the semantics of actions $a\in A[\cdot]$ is unknown (i.e.
+for each $a\in A[\cdot]$ its corresponding $\rho$ and/or $\theta$ functions are undefined).
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_k\}$ is a non-empty {\em set of k models} for the actions in $A[\cdot]$ s.t., each model in $\mathcal{M}\in M$, defines different function pairs $\tup{\rho,\theta}$ over the state variables $F$.
+\item $\mathcal{O}(\tau)$ is an {\em observation} of the execution of an unknown solution plan $\pi$ for the planning problem $P$.
 \end{itemize}
 
-{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model $\mathcal{M}\in M$ and the observed plan execution $\mathcal{O}(\pi,P)$, is the single example to classify.
+{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model $\mathcal{M}\in M$ and the observed plan execution $\mathcal{O}(\tau)$, is the single example to classify.
 The planning model associated to each class acts as the corresponding {\em class prototype} and it summarizes any observation of a plan execution that could be synthesized with that model (i.e.
-all the examples that belong to that class).
+the set of all the examples that belong to that class).
 
-We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\mathcal{O}(\pi,P)$ with respect to the expression:
+We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\mathcal{O}(\tau)$.
+Therefore the {\em solution} to the {\em model recognition} task is the subset of models in $M$ that maximizes this expression.
 \begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{O}|\mathcal{M})\times P(\mathcal{M}).
+argmax_{\mathcal{M}\in M} P(\mathcal{O}|\mathcal{M}) P(\mathcal{M}).
 \end{align}
-The {\em solution} to the {\em model recognition} task is the subset of models in $M$ that maximizes expression (1).
 
@@
 If this probability is not given as input it is reasonable to assume that, {\em a priori}, all models are equiprobable.
-The challenge in our formulation for {\em model recognition} is the definition of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} that expresses the probability of observing $\mathcal{O}(\pi,P)$ when $\mathcal{M}$ is the planning model.
+The challenge in our formulation for {\em model recognition} is the definition of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} that expresses the probability of observing $\mathcal{O}(\tau)$ when $\mathcal{M}$ is the planning model.
 
@@
   \end{scriptsize}  
- \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}$ can be {\em transformed} into a model $\mathcal{M}'$ that is able to produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is consistent with the observation $\mathcal{O}(\pi,P)$.}
+ \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}$ can be {\em transformed} into a model $\mathcal{M}'$ that is able to produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is consistent with the observation $\mathcal{O}(\tau)$.}
 \label{fig:net}
@@
 
-Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess the cost of {\em transforming} $\mathcal{M}$ into a model $\mathcal{M'}$ that can produce a trajectory $\tau(\pi,P)$ that that reaches the goals in $P$ and is consistent with the observation $\mathcal{O}(\pi,P)$.
+Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess the cost of {\em transforming} $\mathcal{M}$ into a model $\mathcal{M'}$ such that: $\mathcal{M'}$ that can produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is consistent with the observation $\mathcal{O}(\tau)$.
 Figure~\ref{fig:net} shows the four-variable {\em Bayesian network} abstracting this process.
@@
 
+Following this equation, the exact computation of $P(\mathcal{O}|\mathcal{M})$ is intractable.
+For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge (infinite in the case of planning problems without dead-ends)~\cite{lesh1995sound}.
+Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially.
+Instead, our approach is to formulate an effectively estimate of $P(\mathcal{O}|\mathcal{M})$ using the {\em edit distance} referred to \strips\ planning models.
 {\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
-We assess the cost of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ formalizing and computing {\em edit distances} that, in this work, are referred to \strips\ planning models.
+In this work, we assess the cost of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ formalizing and computing {\em edit distances} that are referred to \strips\ planning models.
 
@@
 Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of actions (i.e.
-$\rho$ and $\theta$ functions) are specified with \strips\ action schema.
-We start formalizing the \strips\ schema and define the full space of possible \strips\ schema.
-Eventually, we introduce an {\em edit distance} for \strips\ schema to estimate the $P(\mathcal{O}|\mathcal{M})$ likelihoods for classical planning models.
+$\rho$ and $\theta$ functions) are specified with \strips\ action schemata.
+We start formalizing the \strips\ schemata and define the full space of possible \strips\ schemata.
+Eventually, we introduce an {\em edit distance} for \strips\ schemata to estimate the $P(\mathcal{O}|\mathcal{M})$ likelihoods for classical planning models.
 
-\subsection{Well-defined \strips\ action schema}
-\strips\ action schema provide a compact representation for specifying classical planning models.
-Figure~\ref{fig:model-example} shows six \strips\ action schema, codded in PDDL, that determine a particular kind of robot navigation in $n\times n$ grids.
+\subsection{Well-defined \strips\ action schemata}
+\strips\ action schemata provide a compact representation for specifying classical planning models.
+Figure~\ref{fig:model-example} shows six \strips\ action schemata, codded in PDDL, that determine a particular kind of robot navigation in $n\times n$ grids.
 
-{\em A \strips\ action schema} $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+{\em A \strips\ action schemata} $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schemata.
 
-We say that two \strips\ schemes $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
-For instance, we claim that the six action schema of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemes from the four operator {\em blocksworld}~\cite{slaney2001blocks} are not.
-Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schema $\xi'\in\mathcal{M'}$ and vice versa.
+We say that two \strips\ schemata $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
+For instance, we claim that the six action schemata of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemata from the four operator {\em blocksworld}~\cite{slaney2001blocks} are not.
+Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schemata $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
 
-Given a \strips\ action schema $\xi$, let us define an additional set of objects ($\Omega\cap\Omega_\xi=\emptyset$), that we denote as {\em variable names}, and that contains one variable name for each parameter in $pars(\xi)$, that is $\Omega_\xi=\{v_i\}_{i=1}^{|pars(\xi)|}$.
-For any of the six schema defined in Figure~\ref{fig:model-example}, $|pars(\xi)|=2$ so $\Omega_\xi=\{v_1,v_2\}$.
+Given a \strips\ action schemata $\xi$ and a set of {\em predicates} $\Psi$ that shape the propositional state variables.
+The set of FOL interpretations of $\Psi$ over the corresponding parameters of the schemata $\xi$, confines the elements that can appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
+We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
+For any of the six action schemata defined in Figure~\ref{fig:model-example} the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
 
-Given a \strips\ action schema $\xi$ and a set of {\em predicates} $\Psi$ that shape the propositional state variables.
-The set of FOL interpretations of $\Psi$ over the corresponding $\Omega_\xi$ objects ({\em variable names} for schema $\xi$), confines the elements that can appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
-We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
-For any of the six schema defined in Figure~\ref{fig:model-example} the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
-
-Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a schema $\xi$, the space of possible \strips\ schema is constrained by the set ${\mathcal C}$ that includes:
+Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a schemata $\xi$, the space of possible \strips\ schemata is constrained by the set ${\mathcal C}$ that includes:
 \begin{itemize}
@@
 Formally, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$).
-Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schema is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-For every action schema in the navigation model of Figure~\ref{fig:model-example} then $2^{2\times 10}=1,048,576$.
+Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
+For every action schemata in the navigation model of Figure~\ref{fig:model-example} then $2^{2\times 10}=1,048,576$.
 \item {\em Domain-specific constraints}.
-One can also introduce domain-specific knowledge to more precisely constrain the space of possible schema.
+One can also introduce domain-specific knowledge to more precisely constrain the space of possible schemata.
 For instance in a {\em robot navigation} model like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold at the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
 Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate is coding the {\em successor} function for {\em natural numbers}.
-These domain-specific constraints reduces the size of the space of possible action schema to $2^{2\times 7}=16,384$ (for every schema).
+These domain-specific constraints reduce further the size of the space of possible action schemata to $2^{2\times 7}=16,384$ (for every schemata in the navigation model of Figure~\ref{fig:model-example}).
 \end{itemize}
 
-\begin{definition}[Well-defined \strips\ action schema]
-Given a set of {\em predicates} $\Psi$, a list of action {\em parameters} $pars(\xi)$, and set of FOL constraints ${\mathcal C}$ we say that $\xi$ is a {\bf well-defined \strips\ action schema} iff its three lists $pre(\xi)\subseteq {\mathcal I}_{\Psi,\xi}$, $del(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ and $add(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ only contain elements in ${\mathcal I}_{\Psi,\xi}$ and they satisfy all the constraints in ${\mathcal C}$.
+\begin{definition}[Well-defined \strips\ action schemata]
+Given a set of {\em predicates} $\Psi$, a list of action {\em parameters} $pars(\xi)$, and set of FOL constraints ${\mathcal C}$ we say that $\xi$ is a {\bf well-defined \strips\ action schemata} iff its three lists $pre(\xi)\subseteq {\mathcal I}_{\Psi,\xi}$, $del(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ and $add(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ only contain elements in ${\mathcal I}_{\Psi,\xi}$ and they satisfy all the constraints in ${\mathcal C}$.
 \end{definition}
-We say a classical planning action model is {\em well-defined} if all its corresponding  \strips\ action schema are {\em well-defined}.
+We say a classical planning action model is {\em well-defined} if all its corresponding  \strips\ action schemata are {\em well-defined}.
 
 \subsection{Edit distances for \strips\ action models}
-First, we define two edit \emph{operations} on a schema $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
+First, we define the two edit \emph{operations} on a schemata $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
 \begin{itemize}
 \item {\em Deletion}.
-An element is removed from any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ such that the resulting schema is a {\em well-defined} \strips\ action schema.
+An element is removed from any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schemata $\xi\in\mathcal{M}$ such that the resulting schemata is a {\em well-defined} \strips\ action schemata.
 \item {\em Insertion}.
-An element in ${\mathcal I}_{\Psi,\xi}$ is added to any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schema $\xi\in\mathcal{M}$ s.t.
-the resulting schema is {\em well-defined}.
+An element in ${\mathcal I}_{\Psi,\xi}$ is added to any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schemata $\xi\in\mathcal{M}$ s.t.
+the resulting schemata is {\em well-defined}.
 \end{itemize}
@@
 
-Since ${\mathcal I}_{\Psi,\xi}$ are bound sets, the maximum number of edits that can be introduced to a given action model is bound as well.
+Since ${\mathcal I}_{\Psi,\xi}$ are bounded sets, the maximum number of edits that can be introduced to a given action model is bounded as well.
 The \textbf{maximum edit distance} of a \strips\ model $\mathcal{M}$ built within predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$ (note that, if we consider the syntactic \strips\ constraints, $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2\times|{\mathcal I}_{\Psi,\xi}|$).
 
-An observation $\mathcal{O}(\pi,P)$ of the execution of a plan generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrains further the space of the possible schema $\xi\in \mathcal{M}$.
-In this case, we talk about a third kind of constraints that we call {\em observation constraints} and that can also be included into the $\mathcal{C}$ set.
+An observation $\mathcal{O}(\tau)$ of the execution of a plan generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrains further the space of the possible schemata $\xi\in \mathcal{M}$.
+We talk then about a third kind of constraints, that we call {\em observation constraints}, and that can also be included into the $\mathcal{C}$ set.
 In addition, {\em observation constraints} allow us to define an edit distance to elicit the $P(\mathcal{O}|\mathcal{M})$ likelihoods.
@@
 \begin{definition}[Observation edit distance]
-  Given $\mathcal{O}(\pi,P)$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, defined within the same predicates $\Psi$.
+  Given $\mathcal{O}(\tau)$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, defined within the same predicates $\Psi$.
 The {\bf observation edit distance}, $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\pi,P)$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
+$\mathcal{M}'$ produces a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
 
-Note that the {\em observation edit distance} could also be defined assessing the edition required by the observed plan execution to match the given model.
-This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\pi,P)$ instead of $\mathcal{M}$~\cite{sohrabi:precognition:IJCAI2016}.
-Our definition of the {\em observation edit distance} is more practical since normally ${\mathcal I}_{\Psi,\xi}$ is much smaller than $F$.
-In practice, the number of {\em variable objects} should be smaller than the number of objects in a planning problem.
+The {\em observation edit distance} could also be defined assessing the edition required by the observed plan execution to match the given model.
+This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of $\mathcal{M}$~\cite{yang2007learning,sohrabi:precognition:IJCAI2016}.
+Our definition of the {\em observation edit distance} is more practical since normally, ${\mathcal I}_{\Psi,\xi}$ is much smaller than $F$ (the number of variables in the schemata should be smaller than the number of objects in a planning problem).
+In addition, our definition of the {\em observation edit distance} does not accumulate mismatches between the observation and the model that are due to the same source (e.g.
+the same model flaw).
 
@@
 Given a model $\mathcal{M}$.
-The set $M^*$ of {\bf closest consistent models} is the set of action models $\mathcal{M}'$, closest to $\mathcal{M}$ in terms of editions, and that can produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\pi,P)$;
+The set $M^*$ of {\bf closest consistent models} is the set of action models $\mathcal{M}'$, closest to $\mathcal{M}$ in terms of editions, and that can produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$;
   \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
@@
 \subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
-Following equation (2), the exact computation of $P(\mathcal{O}|\mathcal{M})$ is intractable.
-For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge (infinite in the case of planning problems without dead-ends).
-Even worse, the number of models $\mathcal{M}'$ that can be generated {\em editing} a given classical planning model $\mathcal{M}$ explodes combinatorially.
-
-Instead, our approach is to formulate an effectively estimate of $P(\mathcal{O}|\mathcal{M})$.
+Now we we are ready to formulate an effectively estimate of $P(\mathcal{O}|\mathcal{M})$ for the particular instantiation of the {\em model recognition} task where actions are specified with \strips\ schemata.
 
@@
 
-In this baseline scenario there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the given observation (because of full observability), so $P(\mathcal{O}|\tau^*)=1$.
+Under the assumption of {\em full observability} there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the given observation (because of full observability), so $P(\mathcal{O}|\tau^*)=1$.
 Further, there is also a single model $M^*$ that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory so they can be considered the same model).
 
-If there is a single possible trajectory and a single possible model $\mathcal{M}^*$ consistent with the input observation then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
+Provided that there is a single possible trajectory and a single possible model $\mathcal{M}^*$ consistent with the input observation then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
 \begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}) \times P(\mathcal{M}).
+argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}) P(\mathcal{M}).
 \end{align}
@@
 \begin{enumerate}
-\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state.
+\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state~\cite{dennett1983intentional}.
 \item {\em The best plans for $P$ are unique or have different cost}.  
@@
 Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}.
-They allow us to claim that the sum in equation (2) is dominated by its largest term; the term corresponding to the shortest trajectory $\tau^*$ (which is assumed to be unique) consistent with the observation, so other terms in the sum are not added up.
+They allow us to claim that the sum in equation (2) is dominated by its largest term which corresponds to the shortest trajectory $\tau^*$ (that is assumed to be unique) consistent with the observation, so other terms in the sum are not added up.
 Again, there is a single model $M^*$ that can produce $\tau^*$.
@@
 
-Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more reasonable to assume that there is only one plan consistent with the given observation.
+Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more likely to assume that there is only one plan consistent with the given observation.
 
@@
 
-We know from {\em pattern recognition} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}.
+We know from {\em pattern recognition} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}~\cite{}.
 Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
 \begin{align}
-CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  \times p^d \times (1-p)^{N-d}
+CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d (1-p)^{N-d}
 \end{align}
@@
 \begin{align}
-P(\mathcal{M'}|\mathcal{M}) = p^d \times (1-p)^{N-d}
+P(\mathcal{M'}|\mathcal{M}) = p^d  (1-p)^{N-d}
 \end{align}
@@
 \item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
+A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemata in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
 \item {\bf Validates the edited model $\mathcal{M}'$}. The solution plan continues with a postfix that:
@@
 \item Induces a solution plan $\pi$ for the original classical planning problem $P$.
-\item Validates that $\mathcal{O}(\pi,P)$ is an observation of the execution of $\pi$ on the classical planning problem $P$.
+\item Validates that $\mathcal{O}(\tau)$ is an observation of the execution of $\pi$ on the classical planning problem $P$.
 \end{enumerate}
@@
 
-Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
+Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schemata {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
 The postfix of the plan (steps [7,14]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
@@
 }
- \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-14]) the model of Figure\ref{fig:model-example} when schema {\tt\small inc-x} has no {\em preconditions} and positive/negative {\em effects} are swapped wrt Figure\ref{fig:model-example}.}
+ \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-14]) the model of Figure\ref{fig:model-example} when schemata {\tt\small inc-x} has no {\em preconditions} and positive/negative {\em effects} are swapped wrt Figure\ref{fig:model-example}.}
 \label{fig:plan-pdistance}
@@
 
-\subsection{A propositional encoding for \strips\ action schema}
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
-Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.
+\subsection{A propositional encoding for \strips\ action schemata}
+Given a \strips\ action schemata $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schemata can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
+Figure~\ref{fig:encoding} shows the propositional encoding for the six action schemata defined in Figure~\ref{fig:model-example}.
 
-The interest of having a propositional encoding for \strips\ action schema is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}.
+The interest of having a propositional encoding for \strips\ action schemata is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}.
 Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state.
-Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
+Given an operator schemata $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
 \begin{small}  
@@
 
-Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}).
-Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
+Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schemata for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}).
+Note that this editable schemata, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
 
@@
 \end{tiny}
- \caption{\small Propositional encoding for the six schema from Figure~\ref{fig:model-example}.}
+ \caption{\small Propositional encoding for the six schemata from Figure~\ref{fig:model-example}.}
 \label{fig:encoding}
@@
 Conditional effects allow us to compactly define our compilation for computing $\delta(\mathcal{M},\mathcal{O})$ and hence, estimate the $P(\mathcal{O}|\mathcal{M})$ likelihood.
-Given a \strips\ model $\mathcal{M}\in M$ and the observation $\mathcal{O}(\pi,P)$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditional effects $P'=\tup{F',A',I',G'}$:
+Given a \strips\ model $\mathcal{M}\in M$ and the observation $\mathcal{O}(\tau)$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditional effects $P'=\tup{F',A',I',G'}$:
 \begin{itemize}
@@
 \begin{itemize}
-\item Fluents $[pre|del|add]\_e\_\xi$ to model the possible \strips\ schema. 
+\item Fluents $[pre|del|add]\_e\_\xi$ to model the possible \strips\ schemata. 
 \item The fluents to code the {\em observation constraints}:
 \begin{itemize}
-\item $F_{\pi}=\{plan(a_i,i)\}_{\small 1\leq i\leq l}$ to code the $i^{th}$ action in $\mathcal{O}(\pi,P)$.
-The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< l$}, are also added to iterate through the $l$ actions in $\mathcal{O}(\pi,P)$.
-\item The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\pi,P)$ where the action model is validated.
+\item $F_{\pi}=\{plan(a_i,i)\}_{\small 1\leq i\leq l}$ to code the $i^{th}$ action in $\mathcal{O}(\tau)$.
+The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< l$}, are also added to iterate through the $l$ actions in $\mathcal{O}(\tau)$.
+\item The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated.
 \end{itemize}
-\item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schema are edited or validated.
+\item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schemata are edited or validated.
 \end{itemize}
@@
 when $mode_{val}$ holds).
-When the observation $\mathcal{O}(\pi,P)$ includes observed actions, they also include the extra conditional effects $\{at_{i},plan(a_i,i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,l]}$ to validate that actions are applied, exclusively, in the same order as they appear in $\mathcal{O}(\pi,P)$.\\
+When the observation $\mathcal{O}(\tau)$ includes observed actions, they also include the extra conditional effects $\{at_{i},plan(a_i,i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,l]}$ to validate that actions are applied, exclusively, in the same order as they appear in $\mathcal{O}(\tau)$.\\
 
-\item Actions for {\em editing} operator schema $\xi\in\mathcal{M}$.
-This includes the actions for {\em inserting} a new {\em precondition} into an action schema $\xi\in\mathcal{M}$ and for inserting a new {\em negative} or {\em positive} effect into the action schema $\xi\in\mathcal{M}$
+\item Actions for {\em editing} operator schemata $\xi\in\mathcal{M}$.
+This includes the actions for {\em inserting} a new {\em precondition} into an action schemata $\xi\in\mathcal{M}$ and for inserting a new {\em negative} or {\em positive} effect into the action schemata $\xi\in\mathcal{M}$
 \begin{small}
@@
   \end{tiny}  
- \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
+ \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schemata for robot navigation in a $n\times n$ grid.}
 \label{fig:editable}
@@
 To evaluate the empirical performance of {\em model recognition as planning} we collect a set $M$ of possible \strips\ models, that share the same state variables but define different action models.
-Then, we randomly choose one of these models $\mathcal{M}\in M$ and use it to produce an observation $\mathcal{O}(\pi,P)$ of a plan execution.
-Finally, we follow our {\em model recognition as planning} method to identify the model $\mathcal{M}\in M$ that produced $\mathcal{O}(\pi,P)$.
+Then, we randomly choose one of these models $\mathcal{M}\in M$ and use it to produce an observation $\mathcal{O}(\tau)$ of a plan execution.
+Finally, we follow our {\em model recognition as planning} method to identify the model $\mathcal{M}\in M$ that produced $\mathcal{O}(\tau)$.
 The experiment is repeated for models of different kind and different observability of the given plan execution.
@@
 Here we assume that the actual applied transitions are unknown as well as the internal machine state.
-Assuming that the actual applied transitions is unknown means that the observation $\mathcal{O}(\pi,P)$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\mathcal{O}(\pi,P)=\tup{s_1, \ldots , s_m}$.
-Assuming that the internal machine state is unknown means that $\mathcal{O}(\pi,P)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
+Assuming that the actual applied transitions is unknown means that the observation $\mathcal{O}(\tau)$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\mathcal{O}(\tau)=\tup{s_1, \ldots, s_m}$.
+Assuming that the internal machine state is unknown means that $\mathcal{O}(\tau)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
 

### 2018-11-02 doc/icaps-19/model-recognition.tex
all comments by Miquel processed

@@
 % Commented for blind submission
-\author{Diego Aineto\and Sergio Jim\'enez\and Eva Onaindia \and Miquel Ram\'irez\\
-{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-{\small Universitat Polit\`ecnica de Val\`encia.}\\
-{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
-{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
+\author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
+{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
+{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
+{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
+{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
+\And Miquel Ram\'irez\\
+{\scriptsize School of Computing and Information Systems}\\
+{\scriptsize The University of Melbourne}\\
+{\scriptsize Melbourne, Victoria. Australia}\\
+{\scriptsize miquel.ramirez@nimelb.edu.au}}
 
@@
 The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different kinds of {\em automata}.
-We show that {\em model recognition as planning} succeeds to identify the executed automata despite some state variables (e.g.
-the internal automata state) or the actual applied transitions are unobserved.
+We show that {\em model recognition as planning} succeeds to identify the executed automata despite the actual applied transitions, plus some state variables e.g.
+the internal automata state, are unobserved.
 \end{abstract}
@@
             \draw (0, 0) grid (5, 5);
-            \node[anchor=center] at (3.5, 1.5) {R};
+            \node[anchor=center] at (2.5, 1.5) {R};
             \node[anchor=center] at (0.5, 0.5) {};
-            \draw[thick,style=dotted] (0.5,0.5) -- (4.5,0.5);
-            \draw[thick,style=dotted] (4.5,0.5) -- (4.5,1.5);
-            \draw[thick,style=dotted,->] (4.5,1.5) -- (3.7,1.5);                        
+            \draw[thick,style=dotted] (0.5,0.5) -- (3.5,0.5);
+            \draw[thick,style=dotted] (3.5,0.5) -- (3.5,1.5);
+            \draw[thick,style=dotted,->] (3.5,1.5) -- (2.7,1.5);                        
           \end{scope}
@@
 
-To better illustrate {\em model recognition}, imagine a robot in a $n\times n$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
+To better illustrate {\em model recognition}, imagine a robot in a $N\times N$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
 According to this model, the robot can increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
 at {\em even rows} if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at {\em odd rows} if {\tt\small $q0$} holds initially).
-Apart from this particular navigation model, different action models could be defined within the same state variables (e.g.
-altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated) and these models can determine different kinds of robot navigation.
-Given an observation of a plan execution (like the one illustrated at Figure~\ref{fig:grid-example}) {\em model recognition} aims to identify the action model that produced/explains that observation, despite relevant information is unobserved (e.g.
-the applied actions or the value of state variables {\small\tt $q0$} and {\small\tt $q1$}).
+Apart from the model of Figure~\ref{fig:model-example}, different models can be defined within the same state variables (e.g.
+altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated) and these models can shape different kinds of robot navigation.
+Given a partially observed plan execution (like the one illustrated at Figure~\ref{fig:grid-example} where state variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved) {\em model recognition} aims to identify the action model that produced/explains that observation.
 
@@
 We evaluate the effectiveness of {\em model recognition as planning} with sets of \strips\ models that represent different {\em automata}.
-All the {\em automata} in a set are defined within the same state variables but different {\em transition functions}.
-We show that {\em model recognition as planning} succeeds to identify the executed {\em automata} despite some state variables (e.g.
-the internal automata state) or the actual applied transitions are unobserved.
+All the {\em automata} in a set are defined within the same state variables but with different {\em transition functions}.
+We show that {\em model recognition as planning} identifies the executed {\em automata} despite the actual applied transitions, plus some state variables (e.g.
+the internal automata state), are unobserved.
 
@@
   \end{tiny}  
- \caption{\small Example of a \strips\ action model (given in PDDL) for robot navigation in a $n\times n$ grid.}
+ \caption{\small Example of a \strips\ action model (given in PDDL) for robot navigation in a $N\times N$ grid.}
 \label{fig:model-example}
@@
 A {\em classical planning frame} is a $\tup{F,A}$ pair, where $F$ is a set of {\em fluents} and $A$ is a set of {\em actions}.
-The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ that denotes whether the action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying the action $a$ in a state $s$.
-In this work we compactly represent the $\rho$ and $\theta$ functions following the \strips\ model.
-With this regard, an action $a\in A$ is defined by:
+The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ that denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying action $a$ in a state $s$.
+In this work we compactly represent the $\rho$ and $\theta$ functions of an action $a\in A$ with the \strips\ model:
 \begin{itemize}
@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
-A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ that induces the {\em trajectory} $\tau(\pi,P)=\tup{a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-The {\em plan length} is denoted with $|\pi|=n$.
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.,~if the goal condition is satisfied at the last state reached after following the application of the plan $\pi$ in the initial state $I$.
-A solution plan for $P$ is {\em optimal} if it has minimum length.
+A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ and $|\pi|=n$ denotes its {\em plan length}.
+The execution of a plan $\pi$ on a classical planning problem $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
+A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.,~if the goal condition is satisfied at the last state reached after following the execution of plan $\pi$ on problem $P$.
+A solution plan for $P$ is {\em optimal} iff it has minimum length.
 
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$ and a trajectory $\tau(\pi,P)$ generated by the execution of a plan $\pi$ that solves $P$; the observation of the execution of $\pi$ on $P$ is $\mathcal{O}(\tau)=\tup{a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, an interleaved combination of {\small $1\leq l\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states such that:
+Given a classical planning problem $P=\tup{F,A,I,G}$, a solution plan $\pi$ for $P$, and a trajectory $\tau(\pi,P)$ generated by the execution of $\pi$ on $P$; the observation of the execution of $\pi$ on $P$ is $\mathcal{O}(\tau)=\tup{a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, an interleaved combination of {\small $1\leq l\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states:
 \begin{itemize}
-\item Observed actions are {\em consistent} with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$. 
-\item Observed states are a sequence of {\em partial states} $\tup{s_1^o, \ldots, s_m^o}$ that is {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by the trajectory $\tau(\pi,P)$.
-The initial state $I$ is fully observed while the observed states in $\mathcal{O}(\tau)$ may be partial states, i.e.
-the value of certain fluents in the intermediate states may not be observed (~$|s_i^o|\leq |F|$ for every $1\leq i\leq m$).
-Certain states may even be fully unobserved ($0\leq|s_i^o|$ for every $1\leq i\leq m$).
+\item {\bf Observed actions} are {\em consistent} with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$. 
+\item {\bf Observed states} are a sequence of {\em partial states} $\tup{s_1^o, \ldots, s_m^o}$ that is {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by the trajectory $\tau(\pi,P)$.
+The initial state $I$ is fully observed while $\tup{s_1^o, \ldots, s_m^o}$ may be partial states, i.e.
+the value of certain fluents in the intermediate states may not be observed (~$|s_i^o|\leq |F|$ for $1\leq i\leq m$).
+Even more, certain intermediate states may be fully unobserved ($0\leq|s_i^o|$ for $1\leq i\textless m$).
 \end{itemize}
 
-With this regard, we are assuming that a bijective monotone mapping exists between plans and observations~\cite{ramirez2009plan}.
-We can then  conclude that transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite).
+We are assuming that a bijective monotone mapping exists between plans and observations~\cite{ramirez2009plan}.
+Therefore transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite).
 In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
 
-Figure~\ref{fig:grid-example} illustrates the {\em partial observation} of a six-state trajectory \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 5)(ycoord 1)>, <(xcoord 5)(ycoord 2)>, <(xcoord 4)(ycoord 2)>}\}.
-In more detail, this observation only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)} because the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, cannot be observed.
+Figure~\ref{fig:grid-example} illustrates the {\em partial observation} of a five-state trajectory \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 4)(ycoord 2)>, <(xcoord 3)(ycoord 2)>}\}.
+In more detail, this observation only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
+The value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, cannot be observed.
 
@@
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$ is a {\em classical planning problem} s.t.
+\item $P=\tup{F,A[\cdot],I,G}$ is a {\em classical planning problem} where $A[\cdot]$ is a set of actions s.t.
 the semantics of actions $a\in A[\cdot]$ is unknown (i.e.
 for each $a\in A[\cdot]$ its corresponding $\rho$ and/or $\theta$ functions are undefined).
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_k\}$ is a non-empty {\em set of k models} for the actions in $A[\cdot]$ s.t., each model in $\mathcal{M}\in M$, defines different function pairs $\tup{\rho,\theta}$ over the state variables $F$.
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_k\}$ is a {\em set of k different models} for the actions in $A[\cdot]$ s.t., each model in $\mathcal{M}\in M$, defines different semantics for the actions in $A[\cdot]$ (i.e.
+different function pairs $\tup{\rho,\theta}$) but using the same set of state variables $F$.
 \item $\mathcal{O}(\tau)$ is an {\em observation} of the execution of an unknown solution plan $\pi$ for the planning problem $P$.
@@
 the set of all the examples that belong to that class).
-
-We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to the given observation $\mathcal{O}(\tau)$.
-Therefore the {\em solution} to the {\em model recognition} task is the subset of models in $M$ that maximizes this expression.
+To assign a model $\mathcal{M}\in M$ to a given observation $\mathcal{O}(\tau)$, we follow the {\em naive Bayes classifier}.
+The {\em solution} to the {\em model recognition} task is then the subset of models in $M$ that maximizes this expression.
 \begin{align}
@@
 
-\subsection{Formulating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
+\subsection{Formulating the $P(\mathcal{O}|\mathcal{M})$ likelihood}
 The $P(\mathcal{M})$ probability expresses whether one model is known to be a priori more likely than the others.
@@
   \end{scriptsize}  
- \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}$ can be {\em transformed} into a model $\mathcal{M}'$ that is able to produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is consistent with the observation $\mathcal{O}(\tau)$.}
+ \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}$ can be {\em transformed} into a model $\mathcal{M}'$ that produces a trajectory $\tau(\pi,P)$ that (1), reaches the goals in $P$ and (2), is consistent with $\mathcal{O}(\tau)$.}
 \label{fig:net}
@@
 
-Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess the cost of {\em transforming} $\mathcal{M}$ into a model $\mathcal{M'}$ such that: $\mathcal{M'}$ that can produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is consistent with the observation $\mathcal{O}(\tau)$.
-Figure~\ref{fig:net} shows the four-variable {\em Bayesian network} abstracting this process.
-Regarding this network we have the following formulation:
+Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess the cost of {\em transforming} $\mathcal{M}$ into a model $\mathcal{M'}$ that produces a trajectory $\tau(\pi,P)$ such that: (1), $\tau(\pi,P)$ reaches the goals in $P$ and (2), $\tau(\pi,P)$ is consistent with observation $\mathcal{O}(\tau)$.
+Figure~\ref{fig:net} shows the {\em Bayesian network} that abstracts this procedure.
+Regarding this network we have the following formulation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}:
 \begin{align}
@@
 \end{align}
-where $\tau$ ranges over all the trajectories that can be synthesized with a model $\mathcal{M}'$ and $\mathcal{M}'$ ranges over all the models that can be generated {\em transforming} $\mathcal{M}$.
+where $\tau$ ranges over all the trajectories consistent with $\mathcal{O}(\tau)$ that can be synthesized with a model $\mathcal{M}'$ and $\mathcal{M}'$ ranges over all the models that can be generated {\em transforming} $\mathcal{M}$.
 
-Following this equation, the exact computation of $P(\mathcal{O}|\mathcal{M})$ is intractable.
-For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge (infinite in the case of planning problems without dead-ends)~\cite{lesh1995sound}.
-Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially.
-Instead, our approach is to formulate an effectively estimate of $P(\mathcal{O}|\mathcal{M})$ using the {\em edit distance} referred to \strips\ planning models.
+The exact computation of $P(\mathcal{O}|\mathcal{M})$ with equation (2) is intractable.
+For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge, infinite in the case of planning problems without dead-ends~\cite{lesh1995sound}.
+Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially with the number of state variables.
+Instead, our approach is to formulate an effectively estimate of $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} referred to \strips\ planning models.
 {\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
-In this work, we assess the cost of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ formalizing and computing {\em edit distances} that are referred to \strips\ planning models.
+In this work, we assess the cost of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ computing {\em edit distances} between \strips\ planning models.
 
@@
 \label{sec:asPlanning}
-Here we analyze the particular instantiation of the {\em model recognition} task where the semantics of actions (i.e.
-$\rho$ and $\theta$ functions) are specified with \strips\ action schemata.
-We start formalizing the \strips\ schemata and define the full space of possible \strips\ schemata.
+Here we study the particular instantiation of {\em model recognition} where the semantics of models $\mathcal{M}\in M$ (i.e.
+the corresponding $\rho$ and $\theta$ functions) is specified with \strips\ action schemata.
+We start formalizing \strips\ schemata and define the space of possible \strips\ schemata for a given set of state variables.
 Eventually, we introduce an {\em edit distance} for \strips\ schemata to estimate the $P(\mathcal{O}|\mathcal{M})$ likelihoods for classical planning models.
@@
 \strips\ action schemata provide a compact representation for specifying classical planning models.
-Figure~\ref{fig:model-example} shows six \strips\ action schemata, codded in PDDL, that determine a particular kind of robot navigation in $n\times n$ grids.
+Figure~\ref{fig:model-example} shows six \strips\ action schemata that shape a particular kind of robot navigation in $N\times N$ grids (no matter the grid size).
 
-{\em A \strips\ action schemata} $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schemata.
+{\em A \strips\ action schemata} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schemata.
 
-We say that two \strips\ schemata $\xi$ and $\xi'$ are {\em comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
-For instance, we claim that the six action schemata of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemata from the four operator {\em blocksworld}~\cite{slaney2001blocks} are not.
+\begin{definition}[Comparable \strips\ action schemata]
+Two \strips\ schemata $\xi$ and $\xi'$ are {\bf comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
+\end{definition}
+For instance, we claim that the six action schemata of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are not.
 Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schemata $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
 
-Given a \strips\ action schemata $\xi$ and a set of {\em predicates} $\Psi$ that shape the propositional state variables.
-The set of FOL interpretations of $\Psi$ over the corresponding parameters of the schemata $\xi$, confines the elements that can appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists.
+Given the set of {\em predicates} $\Psi$, that shape the propositional state variables, and a list of {\em parameters} $pars(\xi)$.
+The set of elements that can appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schemata $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
 We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
@@
 
-Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a schemata $\xi$, the space of possible \strips\ schemata is constrained by the set ${\mathcal C}$ that includes:
+Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schemata $\xi$, the space of possible \strips\ schemata is constrained by a set ${\mathcal C}$ that includes:
 \begin{itemize}
 \item {\em Syntactic constraints}.
-\strips\ constraints require negative effects appearing as preconditions, negative effects cannot be positive effects at the same time and also, positive effects cannot appear as preconditions.
-Formally, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$).
+\strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
 Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
@@
 \item {\em Domain-specific constraints}.
-One can also introduce domain-specific knowledge to more precisely constrain the space of possible schemata.
+One can introduce domain-specific knowledge to constrain further the space of possible schemata.
 For instance in a {\em robot navigation} model like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold at the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
-Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate is coding the {\em successor} function for {\em natural numbers}.
+Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate codes the {\em successor} function for {\em natural numbers}.
 These domain-specific constraints reduce further the size of the space of possible action schemata to $2^{2\times 7}=16,384$ (for every schemata in the navigation model of Figure~\ref{fig:model-example}).
@@
 \begin{definition}[Well-defined \strips\ action schemata]
-Given a set of {\em predicates} $\Psi$, a list of action {\em parameters} $pars(\xi)$, and set of FOL constraints ${\mathcal C}$ we say that $\xi$ is a {\bf well-defined \strips\ action schemata} iff its three lists $pre(\xi)\subseteq {\mathcal I}_{\Psi,\xi}$, $del(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ and $add(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ only contain elements in ${\mathcal I}_{\Psi,\xi}$ and they satisfy all the constraints in ${\mathcal C}$.
+Given a set of {\em predicates} $\Psi$, a list of action {\em parameters} $pars(\xi)$, and set of FOL constraints ${\mathcal C}$ then $\xi$ is a {\bf well-defined \strips\ action schemata} iff its three lists $pre(\xi)\subseteq {\mathcal I}_{\Psi,\xi}$, $del(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ and $add(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ only contain elements in ${\mathcal I}_{\Psi,\xi}$ and they satisfy all the constraints in ${\mathcal C}$.
 \end{definition}
-We say a classical planning action model is {\em well-defined} if all its corresponding  \strips\ action schemata are {\em well-defined}.
+We say a classical planning action model is {\em well-defined} if all its \strips\ action schemata are {\em well-defined}.
 
@@
 \item {\em Deletion}.
-An element is removed from any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schemata $\xi\in\mathcal{M}$ such that the resulting schemata is a {\em well-defined} \strips\ action schemata.
+An element is removed from a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list of operator schemata $\xi\in\mathcal{M}$ such that the resulting schemata is a {\em well-defined} \strips\ action schemata.
 \item {\em Insertion}.
-An element in ${\mathcal I}_{\Psi,\xi}$ is added to any of these three lists $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the operator schemata $\xi\in\mathcal{M}$ s.t.
+An element in ${\mathcal I}_{\Psi,\xi}$ is added to a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list of operator schemata $\xi\in\mathcal{M}$ s.t.
 the resulting schemata is {\em well-defined}.
@@
 We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ action models are.
-The distance is symmetric and meets the {\em metric axioms} provided that the two {\em edit operations}, deletion and insertion, have the same positive cost.
+The distance is symmetric and meets the {\em metric axioms} provided that the two edit operations, {\em deletion} and {\em insertion}, have the same positive cost.
 
@@
 
-Since ${\mathcal I}_{\Psi,\xi}$ are bounded sets, the maximum number of edits that can be introduced to a given action model is bounded as well.
+Since ${\mathcal I}_{\Psi,\xi}$ is a bounded set, the maximum number of edits that can be introduced to an action model is bounded as well.
 The \textbf{maximum edit distance} of a \strips\ model $\mathcal{M}$ built within predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$ (note that, if we consider the syntactic \strips\ constraints, $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2\times|{\mathcal I}_{\Psi,\xi}|$).
 
-An observation $\mathcal{O}(\tau)$ of the execution of a plan generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrains further the space of the possible schemata $\xi\in \mathcal{M}$.
-We talk then about a third kind of constraints, that we call {\em observation constraints}, and that can also be included into the $\mathcal{C}$ set.
+An observation $\mathcal{O}(\tau)$ of the execution of a plan generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrains further the space of possible schemata $\xi\in \mathcal{M}$.
+We talk then about a third kind of constraints, that we call {\em Observation constraints}, and that can be added to the $\mathcal{C}$ set.
 In addition, {\em observation constraints} allow us to define an edit distance to elicit the $P(\mathcal{O}|\mathcal{M})$ likelihoods.
@@
 \begin{definition}[Observation edit distance]
-  Given $\mathcal{O}(\tau)$, an observation of the execution of a plan for solving $P$ and a \strips\ action model $\mathcal{M}$, defined within the same predicates $\Psi$.
+Given a classical planning problem $P$, an observation $\mathcal{O}(\tau)$ of the execution of a plan for solving that problem and a \strips\ action model $\mathcal{M}$ (all defined within the same predicates $\Psi$).
 The {\bf observation edit distance}, $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
@@
 
-The {\em observation edit distance} could also be defined assessing the edition required by the observed plan execution to match the given model.
-This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of $\mathcal{M}$~\cite{yang2007learning,sohrabi:precognition:IJCAI2016}.
+The {\em observation edit distance} could also be defined assessing the edition required by the observed plan execution to fit the given model.
+This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of modifying the model $\mathcal{M}$~\cite{yang2007learning,sohrabi:precognition:IJCAI2016}.
 Our definition of the {\em observation edit distance} is more practical since normally, ${\mathcal I}_{\Psi,\xi}$ is much smaller than $F$ (the number of variables in the schemata should be smaller than the number of objects in a planning problem).
-In addition, our definition of the {\em observation edit distance} does not accumulate mismatches between the observation and the model that are due to the same source (e.g.
-the same model flaw).
+What is more, our definition of the {\em observation edit distance} does not accumulate mismatches between the observation and the model that are due to the same source (the same model flaw).
 
 \begin{definition}[{\em Closest consistent models}] \label{consistent}
-Given a model $\mathcal{M}$.
-The set $M^*$ of {\bf closest consistent models} is the set of action models $\mathcal{M}'$, closest to $\mathcal{M}$ in terms of editions, and that can produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$;
+Given a model $\mathcal{M}$, the set $M^*$ of the {\bf closest consistent models} is the set of models $\mathcal{M}'$ that: (1) can produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$ and (2), their {\em edit distance} to $\mathcal{M}$ is minimal;
   \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
@@
 \subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
-Now we we are ready to formulate an effectively estimate of $P(\mathcal{O}|\mathcal{M})$ for the particular instantiation of the {\em model recognition} task where actions are specified with \strips\ schemata.
+Now we we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ schemata.
 
-\subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to understand how to build a reasonable estimate of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} for the general case.
+\subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to understand how to build a reasonable estimate of $P(\mathcal{O}|\mathcal{M})$ for the general case.
 
-Under the assumption of {\em full observability} there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the given observation (because of full observability), so $P(\mathcal{O}|\tau^*)=1$.
-Further, there is also a single model $M^*$ that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory so they can be considered the same model).
+Under the assumption of {\em full observability}, there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the input observation so $P(\mathcal{O}|\tau^*)=1$.
+Further, there is also a single model that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory so they can be considered the same model).
 
-Provided that there is a single possible trajectory and a single possible model $\mathcal{M}^*$ consistent with the input observation then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
+Provided that there is a single possible trajectory and a single possible model consistent with the input observation, i.e.
+$M^*=\{\mathcal{M}^*\}$ then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
 \begin{align}
@@
 \end{align}
-Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because is independent of the model $\mathcal{M}\in M$.
+Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because is independent of the input model $\mathcal{M}\in M$.
 
-\subsubsection{Partial observability of the executed plan.} In a similar way, an approximation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be built for the general case where the executed plan is partially observed.
-To deal with this general scenario we add the following two assumptions:
+\subsubsection{Partial observability of the executed plan.} In a similar way, an approximation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be built for the general case, where the executed plan is partially observed.
+We add the following two assumptions to deal with this general case:
 \begin{enumerate}
@@
 Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}.
-They allow us to claim that the sum in equation (2) is dominated by its largest term which corresponds to the shortest trajectory $\tau^*$ (that is assumed to be unique) consistent with the observation, so other terms in the sum are not added up.
-Again, there is a single model $M^*$ that can produce $\tau^*$.
-With this regard, both $P(O|\tau^*)$ and $P(\tau^*|\mathcal{M}^*)$ are independent of $\mathcal{M} \in M$ so equation (1) simplifies once again to equation (3) under the previous assumptions.
+They allow us to assume that the sum in equation (2) is dominated by its largest term, so other terms in the sum are not added up.
+The largest term corresponds here to the shortest trajectory $\tau^*$ consistent with the observation (which now is assumed to be unique).
+Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more likely to assume that there is only one trajectory consistent with the input observation.
 
-Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more likely to assume that there is only one plan consistent with the given observation.
+Again, we have that there is a single model $\mathcal{M}^*$ that can produce a given trajectory $\tau^*$.
+With this said, both $P(O|\tau^*)$ and $P(\tau^*|\mathcal{M}^*)$ are independent of $\mathcal{M} \in M$ so equation (1) simplifies once again to equation (3) under the previous two assumptions.
 
@@
 
-We know from {\em pattern recognition} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}~\cite{}.
+We know from {\em pattern recognition}~\cite{devroye2013probabilistic} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}.
 Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
@@
 \end{align}
-where $N$ is the length of the string, and $p<0.5$ since the cost of applying an {\em edit operation} is higher than do not applying it.
+where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than do not applying it.
 
@@
 
-In other words, we are modeling the edition of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every event $P(X=\top)=p$ and $P(X=\perp)=1-p$.
+In other words, we are modeling the edition of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.
 
 \section{Model Recognition as classical planning}
-This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\mathcal{O})$ can be computed (and hence an approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood) with a a classical planning compilation.
-
-The compilation is an extension of the classical planning compilation for the learning of \strips\ planning models~\cite{aineto2018learning}.
-The intuition behind this compilation is that a solution to the resulting classical planning task is a sequence of actions that:
+This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\mathcal{O})$ can be computed (and hence an approximation $P(\mathcal{O}|\mathcal{M})$) with a a classical planning compilation.
+The compilation is inspired by the classical planning compilation for the learning of \strips\ planning models~\cite{aineto2018learning}.
+The intuition behind our compilation is that a solution to the resulting classical planning task is a sequence of actions that:
 \begin{enumerate}
@@
 A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemata in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$}. The solution plan continues with a postfix that:
-\begin{enumerate}
-\item Induces a solution plan $\pi$ for the original classical planning problem $P$.
-\item Validates that $\mathcal{O}(\tau)$ is an observation of the execution of $\pi$ on the classical planning problem $P$.
+\item {\bf Validates the edited model $\mathcal{M}'$}.
+The solution plan continues with a postfix that induces a solution plan $\pi$ for the original classical planning problem $P$ and validates that $\mathcal{O}(\tau)$ is an observation of the execution of $\pi$ on $P$.
 \end{enumerate}
-\end{enumerate}
+The value of the {\em observation distance} $\delta(\mathcal{M},\mathcal{O})$ is given by the number of {\em edit operations} (insertions and deletions) required by $\mathcal{M}$ to be validated in the input observation.
 
-Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schemata {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
-The postfix of the plan (steps [7,14]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
-Note that our interest is not in $\mathcal{M}'$, the edited model resulting from the compilation, but in the number of required {\em edit operations} (insertions and deletions) required by $\mathcal{M}'$ to be validated.
-In the example of Figure~\ref{fig:plan-pdistance}, $\delta(\mathcal{M},\mathcal{O})=7$.
+Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schemata {\tt\small inc-x} is defined without preconditions and its positive/negative effects are all swapped wrt Figure~\ref{fig:model-example}.
+The postfix of the plan (steps [7,13]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
+As explained our interest is not the edited model but the number of  {\em edit operations} introduced to validate the given model in the input observation, $\delta(\mathcal{M},\mathcal{O})=7$ in the example of Figure~\ref{fig:plan-pdistance}.
 
@@
 03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (editable\_inc-x 3 4) \\
-04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (editable\_inc-x 4 5)\\
-05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (editable\_inc-y-even 1 2)\\
-06 : (insert\_add\_xcoord\_v2\_inc-x)   & 13 : (editable\_dec-x 5 4)\\
-& 14 : (validate\_1)
+04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (editable\_inc-y-even 1 2)\\
+05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (editable\_dec-x 4 3)\\
+06 : (insert\_add\_xcoord\_v2\_inc-x)   & 13 : (validate\_1) 
 \end{tabular}
 }
- \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-14]) the model of Figure\ref{fig:model-example} when schemata {\tt\small inc-x} has no {\em preconditions} and positive/negative {\em effects} are swapped wrt Figure\ref{fig:model-example}.}
+ \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-13]) the model of Figure\ref{fig:model-example} when schemata {\tt\small inc-x} has no {\em preconditions} and positive/negative {\em effects} are swapped wrt Figure\ref{fig:model-example}.}
 \label{fig:plan-pdistance}
@@
 \subsection{A propositional encoding for \strips\ action schemata}
-Given a \strips\ action schemata $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schemata can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
-Figure~\ref{fig:encoding} shows the propositional encoding for the six action schemata defined in Figure~\ref{fig:model-example}.
+Given a \strips\ action schemata $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schemata can be represented with fluents of the kind $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
+We call these fluents the set of {\em editable fluents}.
+Figure~\ref{fig:encoding} shows the propositional encoding for the six action schemata defined in Figure~\ref{fig:model-example} using the {\em editable fluents}.
 
-The interest of having a propositional encoding for \strips\ action schemata is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}.
-Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state.
+We use {\em conditional effects} to compactly define {\em editable actions}.
+Actions whose semantics is given by the value of the {\em editable fluents} $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ at the current state.
 Given an operator schemata $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
@@
 
-Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schemata for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}).
+Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schemata for robot navigation in a $N\times N$ grid.
 Note that this editable schemata, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
@@
 \begin{itemize}
-\item Fluents $[pre|del|add]\_e\_\xi$ to model the possible \strips\ schemata. 
-\item The fluents to code the {\em observation constraints}:
-\begin{itemize}
-\item $F_{\pi}=\{plan(a_i,i)\}_{\small 1\leq i\leq l}$ to code the $i^{th}$ action in $\mathcal{O}(\tau)$.
+\item The set of {\em editable fluents} $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to model the possible \strips\ schemata. 
+\item The fluents to code the {\em observation constraints}: $F_{\pi}=\{plan(a_i,i)\}_{\small 1\leq i\leq l}$ to code the $i^{th}$ action in $\mathcal{O}(\tau)$.
 The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< l$}, are also added to iterate through the $l$ actions in $\mathcal{O}(\tau)$.
-\item The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated.
-\end{itemize}
+The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated.
 \item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schemata are edited or validated.
@@
 Our compilation assumes that initially $\mathcal{M}'$ is defined as $\mathcal{M}$.
-Therefore fluents $[pre|del|add]\_e\_\xi$ hold as given by $\mathcal{M}$.
+Therefore the {\em editable fluents} $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ hold as given by $\mathcal{M}$.
 
@@
   \end{tiny}  
- \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schemata for robot navigation in a $n\times n$ grid.}
+ \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schemata for robot navigation in a $N\times N$ grid.}
 \label{fig:editable}
@@
 
-\subsubsection{Recognition of navigation models.} In this experiment the given models in $M$ represent different navigation models that are computed as the cross product of a regular automata with a four-operator \strips\ model for navigating a $n\times n$ grid.
+\subsubsection{Recognition of navigation models.} In this experiment the given models in $M$ represent different navigation models that are computed as the cross product of a regular automata with a four-operator \strips\ model for navigating a $N\times N$ grid.
 

### 2018-11-08 src/trace_generator.py
trace_generator.py: increased timeout

@@ -42,7 +42,7 @@ PLANNER_OUT = "aux_planner.log"
 if planner == "FD":
     cmd = "rm sas_plan*; ulimit -t 200;" + FD_PATH + FD_CALL + " " + domain_filename + " " + problem_filename + " " + FD_PARAMS + " > " + PLANNER_OUT
 if planner == "M":
-    cmd = "rm sas_plan*; ulimit -t 20;" + M_PATH + M_CALL + " " + domain_filename + " " + problem_filename + " " + M_PARAMS + " > " + PLANNER_OUT
+    cmd = "rm sas_plan*; ulimit -t 200;" + M_PATH + M_CALL + " " + domain_filename + " " + problem_filename + " " + M_PARAMS + " > " + PLANNER_OUT
 if planner == "LPG":
     cmd = "rm sas_plan*; ulimit -t 20;" + LPG_PATH + LPG_CALL + " -o " + domain_filename + " -f " + problem_filename + " " + LPG_PARAMS + " > " + PLANNER_OUT
 print("\n\nExecuting... " + cmd)

### 2018-11-08 src/FAMA_FSM.py
compiler_new.py: some FF provisional code

@@ -128,13 +128,13 @@ except:
     sys.exit(-1)
 
 
-trace_filter = ["a", "b", "c", "next", "head"]
-acceptor_state = "states3"
+trace_filter = ["a", "b", "c", "d", "next", "head"]
+acceptor_state = "states4"
 # trace_filter = ["head"]
 pres_filter = []
-adds_filter = ['states0', 'states1', 'states2', 'states3']
+adds_filter = ['states0', 'states1', 'states2', 'states3', 'states4']
 # adds_filter = []
-dels_filter = ['states0', 'states1', 'states2', 'states3']
+dels_filter = ['states0', 'states1', 'states2', 'states3', 'states4']
 
 
 
@@ -229,6 +229,9 @@ if action_observability > 0:
         learning_task.predicates.append(pddl.predicates.Predicate("plan-" + a.name,
                                                         [pddl.pddl_types.TypedObject("?i", "step")] + a.parameters))
 
+learning_task.predicates.append(pddl.predicates.Predicate("action_applied", []))
+
+
 
 # Original domain actions
 for a in actions:
@@ -252,6 +255,9 @@ for a in actions:
             eff += [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom(known_effect.literal.predicate, ["?o" + str(original_params.index(arg) + 1) for arg in known_effect.literal.args]))]
         elif known_effect.literal.negated and known_effect.literal.predicate not in dels_filter:
             eff += [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.NegatedAtom(known_effect.literal.predicate, ["?o" + str(original_params.index(arg) + 1) for arg in known_effect.literal.args]))]
+
+    # action_applied predicate
+    eff += [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.Atom("action_applied", []))]
 
 
     # Add "step" parameters to the original actions
@@ -395,6 +401,11 @@ for j in range(len(traces)):
                 pddl.actions.Action("validate_" + str(states_seen), [], 0, pddl.conditions.Conjunction(pre), eff, 0))
 
             pre = [pddl.conditions.NegatedAtom("modeProg", [])]
+
+            # action_applied
+            pre += [pddl.conditions.Atom("action_applied", [])]
+            eff += [pddl.effects.Effect([], pddl.conditions.Truth(), pddl.conditions.NegatedAtom("action_applied", []))]
+
[…]

### 2018-11-08 src/compiler_new.py

@@ -695,8 +695,29 @@ if validation_mode:
     ending_horizon = ""
 
 cmd = "rm " + config.OUTPUT_FILENAME + " planner_out.log;" + config.PLANNER_PATH + "/" + config.PLANNER_NAME + " learning_domain.pddl learning_problem.pddl -F " + starting_horizon + " " +ending_horizon + " " + plan_type + " " + config.PLANNER_PARAMS + " > planner_out.log"
+# cmd = "rm " + config.OUTPUT_FILENAME + " planner_out.log;" + config.PROJECT_PATH + "util/FF-v2.3/ff" + " -o learning_domain.pddl -f learning_problem.pddl > planner_out.log"
 # print("\n\nExecuting... " + cmd)
 os.system(cmd)
+
+# file = open("planner_out.log", "r")
+# lines = file.readlines()
+# for i in range(len(lines)):
+#     if "step" in lines[i]:
+#         break
+# lines = lines[i:]
+# for i in range(len(lines)):
+#     if lines[i] == "\n":
+#         break
+# lines = lines[:i-1]
+# file.close()
+#
+# lines = [l.lower() for l in lines]
+# file = open(config.OUTPUT_FILENAME, "w")
+# file.writelines(lines)
+# file.close()
+
+
+
 
 
 ### Read the solution plan to the learning task

### 2018-11-09 doc/icaps-19/model-recognition.tex
Add files via upload
Compilation changed

@@
 \begin{abstract}
-Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but that update these variables according to different action models); {\em model recognition} is the task of identifying which model in the set produced/explains the given observation.
+Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action models), {\em model recognition} is the task of identifying the model that explains the observation.
 The paper formalizes the {\em model recognition} task and introduces a novel method to estimate the probability of a \strips\ model to produce an observation of a plan execution.
-This method, that we called {\em model recognition as planning}, is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observation of the plan execution.
-The effectiveness of {\em model recognition as planning} is shown in a set of \strips\ models encoding different kinds of {\em automata}.
-We show that {\em model recognition as planning} succeeds to identify the executed automata despite the actual applied transitions, plus some state variables e.g.
+This method builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and intermediate states in the observation.
+The effectiveness of the method is shown in a set of \strips\ models encoding different kinds of {\em automata}.
+The presented approach succeeds to identify the executed automata despite the actual applied transitions, plus some state variables e.g.
 the internal automata state, are unobserved.
@@
 \label{sec:introduction}
-{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behavior~\cite{carberry2001techniques}.
+{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behavior~\cite{Carberry01}.
 {\em Goal recognition} is a closely related task that aims identifying the goals of the observed agent.
@@
 
-Diverse approaches has been proposed for plan/goal recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{sukthankar2014plan}.
+Diverse approaches has been proposed for plan/goal recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{sukthankar14}.
 {\em Plan recognition as planning} is the model-based approach for plan/goal recognition that leverages the action model of the observed agent to compute its most likely goal~\cite{ramirez2012plan,ramirez2009plan}.
@@
 In addition, it enables identifying different kinds of automata by observing their execution.
-It is well-known that diverse automata representations, like {\em finite state controllers}, {\em B\"{u}chi automata}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{baier2007exploiting,Geffner:FSM:AAAI10,patrizi2011computing,ivankovic2015optimal,segovia2017generating}.
+It is well-known that diverse automata representations, like {\em finite state controllers}, {\em B\"{u}chi automata}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,patrizi2011computing,IvankovicH15,segovia2017generating}.
 
@@
 
-\subsection{Classical planning}
-We use $F$ to denote the set of {\em fluents} (propositional variables) describing a state.
+\subsection{Classical planning with conditional effects}
+$F$ is the set of {\em fluents} or {\em state variables} (propositional variables).
 A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
 either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent).
-We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
+$L$ is a set of literals that represents a partial assignment of values to fluents, and $\mathcal{L}(F)$ is the set of all literals sets on $F$, i.e.~all partial assignments of values to fluents.
+A {\em state} $s$ is a full assignment of values to fluents.
+We explicitly include negative literals $\neg f$ in states and so $|s|=|F|$ and the size of the state space is $2^{|F|}$.
 
-A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states; i.e.
+%A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states; i.e.
 $|s|=|F|$, so the size of the state space is $2^{|F|}$.
@@
 
-A {\em classical planning frame} is a $\tup{F,A}$ pair, where $F$ is a set of {\em fluents} and $A$ is a set of {\em actions}.
-The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ that denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ that denotes the {\em successor state} that results of applying action $a$ in a state $s$.
-In this work we compactly represent the $\rho$ and $\theta$ functions of an action $a\in A$ with the \strips\ model:
-\begin{itemize}
-\item $\pre(a)\in\mathcal{L}(F)$, the {\em preconditions} of $a$, is the set of literals that must hold for the action $a\in A$ to be applicable.
-\item $\eff^+(a)\in\mathcal{L}(F)$, the {\em positive effects} of $a$, is the set of literals that are true after the application of action $a\in A$.
-\item $\eff^-(a)\in\mathcal{L}(F)$, the {\em negative effects} of $a$, is the set of literals that are false after the application of the action.
-\end{itemize}
-We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
-The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+A {\em planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
+An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\in\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\in\mathcal{L}(F)$.
+The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
+Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$.
+And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
 
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is an initial state and $G\in\mathcal{L}(F)$ is a goal condition.
-A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$ and $|\pi|=n$ denotes its {\em plan length}.
-The execution of a plan $\pi$ on a classical planning problem $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $a_i$ is applicable in $s_{i-1}$ and generates the successor state $s_i=\theta(s_{i-1},a_i)$.
-A plan $\pi$ {\em solves} $P$ iff $G\subseteq s_n$, i.e.,~if the goal condition is satisfied at the last state reached after following the execution of plan $\pi$ on problem $P$.
-A solution plan for $P$ is {\em optimal} iff it has minimum length.
+A {\em planning problem} is defined as a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state in which all the fluents of $F$ are assigned a value true/false and $G$ is the goal set.
+A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, and $|\pi|=n$ denotes its {\em plan length}.
+The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A trajectory $\tau(\pi,P)$ that solves $P$ is one in which $G \subseteq s_n$.
 
-\subsection{Conditional effects}
-{\em Conditional effects} allow classical planning actions to have different semantics according to the value of the current state.
-This feature is useful for compactly defining our method for {\em model recognition as planning}.
-
-An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-
-An action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
+Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
 \[
-triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
+triggered(s,a_c)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
 \]
 
-The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
+The result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$, where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
 
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$, a solution plan $\pi$ for $P$, and a trajectory $\tau(\pi,P)$ generated by the execution of $\pi$ on $P$; the observation of the execution of $\pi$ on $P$ is $\mathcal{O}(\tau)=\tup{a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, an interleaved combination of {\small $1\leq l\leq |\pi|$} observed actions and {\small $1\leq m\leq |\pi|$} observed states:
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
+Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
+
+%which indicates that we observe $l$ actions, $1\leq l\leq |\pi|$, and $m$ states, $1\leq m \leq |\pi|+1$, from $\tau(\pi,P)$:
+
 \begin{itemize}
-\item {\bf Observed actions} are {\em consistent} with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$. 
-\item {\bf Observed states} are a sequence of {\em partial states} $\tup{s_1^o, \ldots, s_m^o}$ that is {\em consistent} with the sequence of states $\tup{s_0, s_1, \ldots, s_n}$ traversed by the trajectory $\tau(\pi,P)$.
-The initial state $I$ is fully observed while $\tup{s_1^o, \ldots, s_m^o}$ may be partial states, i.e.
-the value of certain fluents in the intermediate states may not be observed (~$|s_i^o|\leq |F|$ for $1\leq i\leq m$).
-Even more, certain intermediate states may be fully unobserved ($0\leq|s_i^o|$ for $1\leq i\textless m$).
+\item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
+Specifically, the number of observed actions, $l$, can range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
+\item The {\bf observed states} $\tup{s_0^0, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
+A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
+Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^0 \subseteq s_i$.
+In practive, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
+Exceptionally, $s_m^o$ cannot be fully unobservable for the purpose of our task (we will elaborate on this issue later on).
 \end{itemize}
 
-We are assuming that a bijective monotone mapping exists between plans and observations~\cite{ramirez2009plan}.
-Therefore transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite).
+We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
 In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
 
-Figure~\ref{fig:grid-example} illustrates the {\em partial observation} of a five-state trajectory \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 4)(ycoord 2)>, <(xcoord 3)(ycoord 2)>}\}.
-In more detail, this observation only contains fluents of the kind {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
-The value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, cannot be observed.
-
-\begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
-\end{definition}
+In our model recognition task, we will always assume an \emph{empty sequence of observed actions} and so we will only work with the observed states.
+Figure~\ref{fig:grid-example} illustrates a {\em partial observation} of a five-state trajectory \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 4)(ycoord 2)>, <(xcoord 3)(ycoord 2)>}\}.
+Note that this observation only contains fluents of the type {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is not observable in any of the five states.
 
@@
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$ is a {\em classical planning problem} where $A[\cdot]$ is a set of actions s.t.
-the semantics of actions $a\in A[\cdot]$ is unknown (i.e.
-for each $a\in A[\cdot]$ its corresponding $\rho$ and/or $\theta$ functions are undefined).
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_k\}$ is a {\em set of k different models} for the actions in $A[\cdot]$ s.t., each model in $\mathcal{M}\in M$, defines different semantics for the actions in $A[\cdot]$ (i.e.
-different function pairs $\tup{\rho,\theta}$) but using the same set of state variables $F$.
-\item $\mathcal{O}(\tau)$ is an {\em observation} of the execution of an unknown solution plan $\pi$ for the planning problem $P$.
+\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions.
+For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
+the functions $\rho$ and/or $\theta$ of $a$ are undefined.
+\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_k\}$ is a set of {\em k different planning models} for the actions in $A[\cdot]$.
+A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$.
+Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
+\item $\mathcal{O}(\tau)$ is an observation of the execution of an unknown solution plan $\pi$ for the planning problem $P$.
 \end{itemize}
 
-{\em Model recognition} can be understood as a {\em classification task} where each class is represented with a different planning model $\mathcal{M}\in M$ and the observed plan execution $\mathcal{O}(\tau)$, is the single example to classify.
-The planning model associated to each class acts as the corresponding {\em class prototype} and it summarizes any observation of a plan execution that could be synthesized with that model (i.e.
+Model recognition can be understood as a {\em classification task} where each class is represented by a different planning model $\mathcal{M}\in M$, and the observed plan execution $\mathcal{O}(\tau)$ is the single example to classify.
+The planning model associated to each class acts as the corresponding {\em class prototype} and it summarizes any observation of a plan execution that could be synthesized with such model (i.e.
 the set of all the examples that belong to that class).
-To assign a model $\mathcal{M}\in M$ to a given observation $\mathcal{O}(\tau)$, we follow the {\em naive Bayes classifier}.
-The {\em solution} to the {\em model recognition} task is then the subset of models in $M$ that maximizes this expression.
+We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to a given observation $\mathcal{O}(\tau)$.
+The {\em solution} to the model recognition task is then the subset of models in $M$ that maximizes this expression.
+
 \begin{align}
@@
 \subsection{Formulating the $P(\mathcal{O}|\mathcal{M})$ likelihood}
+
 The $P(\mathcal{M})$ probability expresses whether one model is known to be a priori more likely than the others.
 If this probability is not given as input it is reasonable to assume that, {\em a priori}, all models are equiprobable.
-The challenge in our formulation for {\em model recognition} is the definition of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} that expresses the probability of observing $\mathcal{O}(\tau)$ when $\mathcal{M}$ is the planning model.
+The challenge of our formulation to the model recognition task is the definition of $P(\mathcal{O}|\mathcal{M})$, the likelihood that expresses the probability of observing $\mathcal{O}(\tau)$ when $\mathcal{M}$ is the planning model.
 
@@
   \end{scriptsize}
- \caption{\small {\em Bayesian network} abstracting that model $\mathcal{M}$ can be {\em transformed} into a model $\mathcal{M}'$ that produces a trajectory $\tau(\pi,P)$ that (1), reaches the goals in $P$ and (2), is consistent with $\mathcal{O}(\tau)$.}
+ \caption{\small {\em Bayesian network} representing that model $\mathcal{M}$ is {\em transformable} into a model $\mathcal{M}'$ that produces a trajectory $\tau(\pi,P)$ that (1) reaches the goals in $P$ and (2) it is consistent with $\mathcal{O}(\tau)$.}
 \label{fig:net}
@@
 
-Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} is to assess the cost of {\em transforming} $\mathcal{M}$ into a model $\mathcal{M'}$ that produces a trajectory $\tau(\pi,P)$ such that: (1), $\tau(\pi,P)$ reaches the goals in $P$ and (2), $\tau(\pi,P)$ is consistent with observation $\mathcal{O}(\tau)$.
-Figure~\ref{fig:net} shows the {\em Bayesian network} that abstracts this procedure.
+Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ likelihood is to assess the cost of transforming $\mathcal{M}$ into a model $\mathcal{M'}$ that produces a trajectory $\tau(\pi,P)$ such that: (1) $\tau(\pi,P)$ reaches the goals in $P$ and (2) $\tau(\pi,P)$ is consistent with observation $\mathcal{O}(\tau)$.
+Figure~\ref{fig:net} shows the {\em Bayesian network} that embodies this procedure.
 Regarding this network we have the following formulation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}:
+
 \begin{align}
@@
 \end{align}
+
 where $\tau$ ranges over all the trajectories consistent with $\mathcal{O}(\tau)$ that can be synthesized with a model $\mathcal{M}'$ and $\mathcal{M}'$ ranges over all the models that can be generated {\em transforming} $\mathcal{M}$.
@@
 Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially with the number of state variables.
-Instead, our approach is to formulate an effectively estimate of $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} referred to \strips\ planning models.
-{\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, and that has been proved successful for {\em pattern recognition}~\cite{masek1980faster,bunke1997relation}.
-In this work, we assess the cost of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ computing {\em edit distances} between \strips\ planning models.
+Instead, our approach is to estimate of $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} referred to \strips\ planning models.
+{\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, which have been proved successful for {\em pattern recognition}~\cite{MasekP80,Bunke97}.
+In this work, we assess the cost of {\em transforming} $\mathcal{M}$ into $\mathcal{M'}$ computing {\em edit distances} between \strips\ planning models.
 
-\section{Recognition of \strips\ models}
+\section{Recognition of \strips\ planning models}
 \label{sec:asPlanning}
-Here we study the particular instantiation of {\em model recognition} where the semantics of models $\mathcal{M}\in M$ (i.e.
-the corresponding $\rho$ and $\theta$ functions) is specified with \strips\ action schemata.
-We start formalizing \strips\ schemata and define the space of possible \strips\ schemata for a given set of state variables.
-Eventually, we introduce an {\em edit distance} for \strips\ schemata to estimate the $P(\mathcal{O}|\mathcal{M})$ likelihoods for classical planning models.
+
+Our formalization of a model recognition task is valid for any planning model but our \emph{edit distance} measure is exclusively for \strips\ models and so we restrict our attention to \strips\texttt{-}compilable planning models\footnote{An edit distance for other planning models is also definable}.
+Thus, we focus on the recognition of $\mathcal{M}\in M$, where $M$ is a set of \strips\ planning models and $\mathcal{M}$ defines the semantics of the actions through a set of \strips\ action schemata.
 
@@
 
-{\em A \strips\ action schemata} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schemata.
+{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 
 \begin{definition}[Comparable \strips\ action schemata]
-Two \strips\ schemata $\xi$ and $\xi'$ are {\bf comparable} iff $pars(\xi)=pars(\xi')$, both share the same list of parameters.
+Two \strips\ schemata $\xi$ and $\xi'$ are {\bf comparable} iff $pars(\xi)=pars(\xi')$, i.e, both share the same list of parameters.\footnote{In \strips\ models, $pars(\xi)=pars(\xi')$ implies the number of parameters must be the same.
+For other planning models that allow object typing, the equality implies that parameters share the same type}
 \end{definition}
+
 For instance, we claim that the six action schemata of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are not.
-Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schemata $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
+Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
 
-Given the set of {\em predicates} $\Psi$, that shape the propositional state variables, and a list of {\em parameters} $pars(\xi)$.
-The set of elements that can appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schemata $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
+Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
+The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
 We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
-For any of the six action schemata defined in Figure~\ref{fig:model-example} the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+For any of the six action schemata of Figure~\ref{fig:model-example}, the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
 
-Despite any element from ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schemata $\xi$, the space of possible \strips\ schemata is constrained by a set ${\mathcal C}$ that includes:
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is constrained by a set ${\mathcal C}$ that includes:
+
 \begin{itemize}
@@
 Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-For every action schemata in the navigation model of Figure~\ref{fig:model-example} then $2^{2\times 10}=1,048,576$.
+For every action schema in the navigation model of Figure~\ref{fig:model-example} then $2^{2\times 10}=1,048,576$.
+
 \item {\em Domain-specific constraints}.
 One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance in a {\em robot navigation} model like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold at the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
-Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear at any of these lists because the {\tt\small next} predicate codes the {\em successor} function for {\em natural numbers}.
-These domain-specific constraints reduce further the size of the space of possible action schemata to $2^{2\times 7}=16,384$ (for every schemata in the navigation model of Figure~\ref{fig:model-example}).
+For instance, in a {\em robot navigation} model like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold at the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
+Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear in any of these lists because the {\tt\small next} predicate codes the {\em successor} function for {\em natural numbers}.
+These domain-specific constraints reduce further the size of the space of possible action schemata to $2^{2\times 7}=16,384$ (for every schema in the navigation model of Figure~\ref{fig:model-example}).
 \end{itemize}
@@
 \begin{definition}[Well-defined \strips\ action schemata]
-Given a set of {\em predicates} $\Psi$, a list of action {\em parameters} $pars(\xi)$, and set of FOL constraints ${\mathcal C}$ then $\xi$ is a {\bf well-defined \strips\ action schemata} iff its three lists $pre(\xi)\subseteq {\mathcal I}_{\Psi,\xi}$, $del(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ and $add(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ only contain elements in ${\mathcal I}_{\Psi,\xi}$ and they satisfy all the constraints in ${\mathcal C}$.
+Given a set of {\em predicates} $\Psi$, a list of action {\em parameters} $pars(\xi)$, and set of FOL constraints ${\mathcal C}$, $\xi$ is a {\bf well-defined \strips\ action schema} iff its three lists $pre(\xi)\subseteq {\mathcal I}_{\Psi,\xi}$, $del(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ and $add(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ only contain elements in ${\mathcal I}_{\Psi,\xi}$ and they satisfy all the constraints in ${\mathcal C}$.
 \end{definition}
-We say a classical planning action model is {\em well-defined} if all its \strips\ action schemata are {\em well-defined}.
 
-\subsection{Edit distances for \strips\ action models}
-First, we define the two edit \emph{operations} on a schemata $\xi\in\mathcal{M}$ that belongs to a \strips\ action model $\mathcal{M}\in M$:
+We say a planning model $\mathcal{M}$ is {\em well-defined} if all its \strips\ action schemata are {\em well-defined}.
+
+\subsection{Edit distances for \strips\ planning models}
+First, we define the two edit \emph{operations} on a schema $\xi$ that belongs to a \strips\ model $\mathcal{M}\in M$:
+
 \begin{itemize}
-\item {\em Deletion}.
-An element is removed from a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list of operator schemata $\xi\in\mathcal{M}$ such that the resulting schemata is a {\em well-defined} \strips\ action schemata.
+\item {\em Deletion}. Given $\xi\in\mathcal{M}$, an element from any of the lists $pre(\xi)$/$del(\xi)$/$add(\xi)$ is removed such that the result is a {\em well-defined} \strips\ action schema.
 \item {\em Insertion}.
-An element in ${\mathcal I}_{\Psi,\xi}$ is added to a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list of operator schemata $\xi\in\mathcal{M}$ s.t.
-the resulting schemata is {\em well-defined}.
+Given $\xi\in\mathcal{M}$, an element in ${\mathcal I}_{\Psi,\xi}$ is added to any of the lists $pre(\xi)$/$del(\xi)$/$add(\xi)$ such that the result is a {\em well-defined} action schema.
 \end{itemize}
 
-We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ action models are.
+We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ models are.
 The distance is symmetric and meets the {\em metric axioms} provided that the two edit operations, {\em deletion} and {\em insertion}, have the same positive cost.
@@
 \begin{definition}[Edit distance]
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} and {\em well-defined} \strips\ action models defined within the same set of predicates $\Psi$.
+  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} and {\em well-defined} \strips\ planning models within the same set of predicates $\Psi$.
 The {\bf edit distance} $\delta(\mathcal{M},\mathcal{M}')$ is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
@@
 Since ${\mathcal I}_{\Psi,\xi}$ is a bounded set, the maximum number of edits that can be introduced to an action model is bounded as well.
-The \textbf{maximum edit distance} of a \strips\ model $\mathcal{M}$ built within predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$ (note that, if we consider the syntactic \strips\ constraints, $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2\times|{\mathcal I}_{\Psi,\xi}|$).
+The \textbf{maximum edit distance} of a \strips\ model $\mathcal{M}$ built with predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$ (note that if we consider the set of syntactic constraints then $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2\times|{\mathcal I}_{\Psi,\xi}|$).
 
-An observation $\mathcal{O}(\tau)$ of the execution of a plan generated with an action model $\mathcal{M}$, reflects {\em semantic knowledge} that constrains further the space of possible schemata $\xi\in \mathcal{M}$.
-We talk then about a third kind of constraints, that we call {\em Observation constraints}, and that can be added to the $\mathcal{C}$ set.
-In addition, {\em observation constraints} allow us to define an edit distance to elicit the $P(\mathcal{O}|\mathcal{M})$ likelihoods.
+\vspace{0.02cm}
+
+An observation of the execution of a plan generated with an action model $\mathcal{M}$ further constraints the space of possible action schemata of $\mathcal{M}$.
+The \emph{semantic knowledge} included in the observations introduce a third type of constraints, that we will call {\em observation constraints}, and that can be added to the set $\mathcal{C}$.
+In addition, {\em observation constraints} allow us to define an edit distance to elicit the value of $P(\mathcal{O}|\mathcal{M})$.
 It can be argued that the shorter this distance the better the given model explains the given observation.
@@
 \begin{definition}[Observation edit distance]
-Given a classical planning problem $P$, an observation $\mathcal{O}(\tau)$ of the execution of a plan for solving that problem and a \strips\ action model $\mathcal{M}$ (all defined within the same predicates $\Psi$).
-The {\bf observation edit distance}, $\delta(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$; \[\delta(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
+Given a planning problem $P$, an observation $\mathcal{O}(\tau)$ of the execution of a plan that solves $P$ and a \strips\ planning model $\mathcal{M}$ (all defined within the same set of predicates $\Psi$).
+The {\bf observation edit distance}, $\delta^o(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
+$\mathcal{M}'$ produces a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$; \[\delta^o(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
 \end{definition}
 
-The {\em observation edit distance} could also be defined assessing the edition required by the observed plan execution to fit the given model.
-This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of modifying the model $\mathcal{M}$~\cite{yang2007learning,sohrabi:precognition:IJCAI2016}.
-Our definition of the {\em observation edit distance} is more practical since normally, ${\mathcal I}_{\Psi,\xi}$ is much smaller than $F$ (the number of variables in the schemata should be smaller than the number of objects in a planning problem).
-What is more, our definition of the {\em observation edit distance} does not accumulate mismatches between the observation and the model that are due to the same source (the same model flaw).
+$\delta^o(\mathcal{M},\mathcal{O})$ can also be defined through the edition that the observation $\mathcal{O}(\tau)$ requires to fit $\mathcal{M}$.
+This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of the model $\mathcal{M}$~\cite{yang2007learning,SohrabiRU16}.
+Our definition of {\em observation edit distance} is more practical since the size of ${\mathcal I}_{\Psi,\xi}$ is usually much smaller than $F$ (the number of variables in the action schemata should normally be lower than the number of objects in a planning problem).
 
 \begin{definition}[{\em Closest consistent models}] \label{consistent}
-Given a model $\mathcal{M}$, the set $M^*$ of the {\bf closest consistent models} is the set of models $\mathcal{M}'$ that: (1) can produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$ and (2), their {\em edit distance} to $\mathcal{M}$ is minimal;
+Given a model $\mathcal{M}$, the set $M^*$ of the {\bf closest consistent models} is the set of models $\mathcal{M}'$ that: (1) produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$ and (2) their {\em edit distance} to $\mathcal{M}$ is minimal;
   \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
@@
 \subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
-Now we we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ schemata.
+
+Now we we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ action schemata.
 
@@
 
-\subsubsection{Partial observability of the executed plan.} In a similar way, an approximation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood} can be built for the general case, where the executed plan is partially observed.
+\subsubsection{Partial observability of the executed plan.} In a similar way, an approximation to $P(\mathcal{O}|\mathcal{M})$ can be built for the general case, where the executed plan is partially observed.
 We add the following two assumptions to deal with this general case:
+
 \begin{enumerate}
-\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state~\cite{dennett1983intentional}.
+\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state~\cite{Dennett83}.
 \item {\em The best plans for $P$ are unique or have different cost}.
 \end{enumerate}
+
 Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}.
@@
 where $d=\delta(\mathcal{M},\mathcal{M}')$.
-Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation (5) but instead, the parameter $d$ is given by the {\em observation distance}, $d=\delta(\mathcal{M},\mathcal{O})$.
+Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation (5) but instead, the parameter $d$ is given by the {\em observation distance}, $d=\delta^o(\mathcal{M},\mathcal{O})$.
 
@@
 
-\section{Model Recognition as classical planning}
-This section shows that, for \strips\ planning models, $\delta(\mathcal{M},\mathcal{O})$ can be computed (and hence an approximation $P(\mathcal{O}|\mathcal{M})$) with a a classical planning compilation.
-The compilation is inspired by the classical planning compilation for the learning of \strips\ planning models~\cite{aineto2018learning}.
-The intuition behind our compilation is that a solution to the resulting classical planning task is a sequence of actions that:
-\begin{enumerate}
-\item {\bf Edits the action model $\mathcal{M}$ to build $\mathcal{M}'$}.
-A solution plan starts with a {\em prefix} that modifies the preconditions and effects of the action schemata in $\mathcal{M}$ using to the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
-\item {\bf Validates the edited model $\mathcal{M}'$}.
-The solution plan continues with a postfix that induces a solution plan $\pi$ for the original classical planning problem $P$ and validates that $\mathcal{O}(\tau)$ is an observation of the execution of $\pi$ on $P$.
-\end{enumerate}
-The value of the {\em observation distance} $\delta(\mathcal{M},\mathcal{O})$ is given by the number of {\em edit operations} (insertions and deletions) required by $\mathcal{M}$ to be validated in the input observation.
+\section{Model Recognition as planning}
 
-Figure~\ref{fig:plan-pdistance} shows the plan with a prefix (steps [0,6]) for editing the planning model of Figure\ref{fig:model-example} when its schemata {\tt\small inc-x} is defined without preconditions and its positive/negative effects are all swapped wrt Figure~\ref{fig:model-example}.
-The postfix of the plan (steps [7,13]) validates the edited action model at the observation of the plan execution illustrated at Figure~\ref{fig:grid-example}.
-As explained our interest is not the edited model but the number of  {\em edit operations} introduced to validate the given model in the input observation, $\delta(\mathcal{M},\mathcal{O})=7$ in the example of Figure~\ref{fig:plan-pdistance}.
+This section shows that $\delta^o(\mathcal{M},\mathcal{O})$, and hence an approximation to $P(\mathcal{O}|\mathcal{M})$), can be computed with a compilation-to-planning approach as the one proposed in~\cite{aineto2018learning} (AJO approach hereafter) for learning \strips\ models.
+The AJO approach receives as input an empty model $\mathcal{M}$, which only contains the headers of the action schemata formed of $\xi=\tup{name(\xi),pars(\xi)}$, and a set of observations of plan executions (i.e., a set of  $\mathcal{O}(\tau)$), and it returns a model $\mathcal{M'}$ with specification of preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$ following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1},a_i)$ for every observed action of $\mathcal{O}(\tau)$ and $s_i=\theta(s_{i-1},a_i)$ for every observed state of $\mathcal{O}(\tau)$.
+
+In essence, the task $\tup{\mathcal{M},\mathcal{O}}$ of the AJO approach -- learning $\mathcal{M}$ that satisfies $\mathcal{O}$- can be interpreted as editing the empty action schemata of $\mathcal{M}$ introducing preconditions and effects until $\mathcal{O}$ is validated with the resulting model.
+We leverage the same idea to compute $\delta^o(\mathcal{M},\mathcal{O})$ with the exception that now our $\mathcal{M}$ is not empty but $\mathcal{M}=\{\xi_1, \ldots, \xi_m\}$, where $\xi_i=\tup{name(\xi_i),pars(\xi_i),pre(\xi_i),add(\xi_i),del(\xi_i)}$, $1 \leq i \leq m$.
+
+Editing/learning the action schemata of $\mathcal{M}$ in the AJO approach is addressed converting the task into a classical planning problem, which is later solved with a planner.
+The intuition behind this compilation is that a solution plan to the problem is a sequence of: (a) \emph{edit actions} on the schemata of $\mathcal{M}$ to build $\mathcal{M'}$ and (b) \emph{validate actions} that apply $\mathcal{M'}$ in $\mathcal{O(\tau)}$.
+The adaptation of this compilation scheme for solving $\delta^o(\mathcal{M},\mathcal{O})$ results in a planning problem $P'=\tup{F',A',I',G'}$ whose objective is to determine the preconditions and effects of the action schemata of $\mathcal{M}$ that need to be added or deleted so as to satisfy $\mathcal{O}$.
+The accomplishment of this task requires therefore a propositional encoding of the components of the action schemata.
+Specifically:
+
+\begin{itemize}
+\item $F'$ contains the necessary fluents to represent the $pre$, $add$ and $del$ of the action schemata.
+Specifically, it is a set of \emph{editable fluents} of the type $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
+Additionally, $F$ also contains fluents to encode the observation constraints; that is, fluents to iterate through the $l$ observed actions and $m$ observed states of $\mathcal{O(\tau)}$.
+\item $A'$ comprises two types of actions with conditional effects:
+    \begin{itemize}
+    \item actions for editing $\xi \in \mathcal{M}$ that follow the \emph{syntactic constraints} for well-defined \strips\ action schemata.
+Hence, here are actions for inserting a new precondition $pre\_e\_\xi$ or a new positive effect $add\_e\_\xi$ in $\xi$ when neither $pre\_e\_\xi$, $del\_e\_\xi$ or $add\_e\_\xi$ exist in $\xi$, and actions for inserting a new negative effect when neither $del\_e\_\xi$ or $add\_e\_\xi$ appear in $\xi$ but $pre\_e\_\xi$ does.
+    \item actions for applying the new action schemata of the edited model $\mathcal{M'}$ and validating the observed states of $\mathcal{O(\tau)}$.
+Particularly, the \emph{apply} actions check the behavior of the propositional version of the new action schemata with the editable fluents of the current state.
+    \end{itemize}
+\item $I'$ encodes the editable fluents $pre\_e\_\xi, del\_e\_\xi$ and $add\_e\_\xi$ that hold in the action schemata of $\mathcal{M}$.
+\item $G'$ contains the necessary fluents to check that all the observed actions and states of $\mathcal{O(\tau)}$ are generated correctly following the edited model $\mathcal{M'}$.
+    %Here we include the necessary fluents to check the final reached state is consistent with the final state $s_m^o$ of $\mathcal{O(\tau)}$.
+\end{itemize}
+
+We now show an example of a solution plan to a planning problem $P'$ that results from compiling a specific task $\tup{\mathcal{M},\mathcal{O}}$.
+Consider that $\mathcal{M}$ is the model of Figure~\ref{fig:model-example} except that the schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
+And consider an observation that only contains the initial and final state of Figure~\ref{fig:grid-example}; i.e., $\mathcal{O(\tau)}=\tup{s_0^o,s_m^o}=\tup{(1,1),(3,2)}$.
+The plan found by a planner to $P'$ is shown in figure~\ref{fig:plan-pdistance}.
+The first seven steps are the edit actions to \emph{fix} the schema {\tt\small inc-x}; step 07 is a validate action that sets the robot in the state $s_0^o=(1,1)$; step 08 applies the action {\tt\small (inc-x 1 2)} and so moves the robot one cell to the right to $(2,1)$; steps 09-11 move also the robot one cell to the right; step 11 applies an action that increases coordinate $y$ to an even number, thus moving robot to row $2$; step 12 moves the robot one cell to the left and, finally, action {\tt\small (validate\_1)} checks the robot position is consistent with the final state $s_m^o=(3,2)$.
 
@@
 00 : (insert\_pre\_xcoord\_v1\_inc-x)   & 07 : (validate\_0)\\
-01 : (insert\_pre\_next\_v1\_v2\_inc-x) & 08 : (editable\_inc-x 1 2)\\
-02 : (insert\_pre\_q0\_inc-x)           & 09 : (editable\_inc-x 2 3)\\
-03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (editable\_inc-x 3 4) \\
-04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (editable\_inc-y-even 1 2)\\
-05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (editable\_dec-x 4 3)\\
+01 : (insert\_pre\_next\_v1\_v2\_inc-x) & 08 : (apply\_inc-x 1 2)\\
+02 : (insert\_pre\_q0\_inc-x)           & 09 : (apply\_inc-x 2 3)\\
+03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (apply\_inc-x 3 4) \\
+04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (apply\_inc-y-even 1 2)\\
+05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (apply\_dec-x 4 3)\\
 06 : (insert\_add\_xcoord\_v2\_inc-x)   & 13 : (validate\_1)
@@
 }
- \caption{\small Plan for editing (steps [0-6]) and validating (steps [7-13]) the model of Figure\ref{fig:model-example} when schemata {\tt\small inc-x} has no {\em preconditions} and positive/negative {\em effects} are swapped wrt Figure\ref{fig:model-example}.}
+ \caption{\small A plan for $\tup{\mathcal{M},\mathcal{O}}$ where $\mathcal{M}$ is a modification of the model of Figure \ref{fig:model-example} and $\mathcal{O}$ a partial observation from Figure~\ref{fig:grid-example}.}
 \label{fig:plan-pdistance}
@@
 
-\subsection{A propositional encoding for \strips\ action schemata}
-Given a \strips\ action schemata $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schemata can be represented with fluents of the kind $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
-We call these fluents the set of {\em editable fluents}.
-Figure~\ref{fig:encoding} shows the propositional encoding for the six action schemata defined in Figure~\ref{fig:model-example} using the {\em editable fluents}.
-
-We use {\em conditional effects} to compactly define {\em editable actions}.
-Actions whose semantics is given by the value of the {\em editable fluents} $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ at the current state.
-Given an operator schemata $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
-\begin{small}  
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
-\cond(\mathsf{editable_{\xi}})=&\{del\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
-&\{add\_e\_\xi\}\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
-\end{align*}
-\end{small}
-
-Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schemata for robot navigation in a $N\times N$ grid.
-Note that this editable schemata, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
-
-\begin{figure}
-\begin{tiny}
-\begin{verbatim}
-;;; Propositional encoding for inc-x(?v1 ?v2)
-(pre_xcoord_v1_inc-x) (pre_next_v1_v2_inc-x) 
-(pre_q0__inc-x)
-(del_xcoord_v1_inc-x) (add_xcoord_v2_inc-x)
-
-;;; Propositional encoding for dec-x(?v1 ?v2)
-(pre_xcoord_v1_dec-x) (pre_next_v2_v1_dec-x) 
-(pre_q1___dec-x)
-(del_xcoord_v1_dec-x) (add_xcoord_v2_dec-x)
-
-;;; Propositional encoding for inc-y-even(?v1 ?v2)
-(pre_ycoord_v1_inc-y-even) (pre_next_v1_v2_inc-y-even)
-(pre_q0__inc-y-even)
-(del_ycoord_v1_inc-y-even) (del_q0__inc-y-even)
-(add_ycoord_v2_inc-y-even) (add_q1__inc-y-even)
-
-;;; Propositional encoding for inc-y-odd(?v1 ?v2)
-(pre_ycoord_v1_inc-y-odd) (pre_next_v1_v2_inc-y-odd) 
-(pre_q0__inc-y-odd)
-(del_ycoord_v1_inc-y-odd) (del_q1__inc-y-odd)
-(add_ycoord_v2_inc-y-odd) (add_q0__inc-y-odd)
-
-;;; Propositional encoding for dec-y-even(v1 ?v2)
-(pre_ycoord_v1_dec-y-even) (pre_next_v2_v1_dec-y-even)
-(pre_q0__dec-y-even)
-(del_ycoord_v1_dec-y-even) (del_q0__dec-y-even)
-(add_ycoord_v2_dec-y-even) (add_q1__dec-y-even)
-
-;;; Propositional encoding for dec-y-odd(?v1 ?v2)
-(pre_ycoord_v1_dec-y-odd) (pre_next_v2_v1_dec-y-odd)
-(pre_q1__dec-y-odd)
-(del_ycoord_v1_dec-y-odd) (del_q1__dec-y-odd)
-(add_ycoord_v2_dec-y-odd) (add_q0__dec-y-odd)
-\end{verbatim}
-\end{tiny}
- \caption{\small Propositional encoding for the six schemata from Figure~\ref{fig:model-example}.}
-\label{fig:encoding}
-\end{figure}
-
-\subsection{The compilation formalization}
-Conditional effects allow us to compactly define our compilation for computing $\delta(\mathcal{M},\mathcal{O})$ and hence, estimate the $P(\mathcal{O}|\mathcal{M})$ likelihood.
-Given a \strips\ model $\mathcal{M}\in M$ and the observation $\mathcal{O}(\tau)$ of the execution of a plan for solving $P=\tup{F,A,I,G}$, our compilation outputs a classical planning task with conditional effects $P'=\tup{F',A',I',G'}$:
-\begin{itemize}
-\item $F'$ extends the original fluents $F$ with:
-\begin{itemize}
-\item The set of {\em editable fluents} $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to model the possible \strips\ schemata. 
-\item The fluents to code the {\em observation constraints}: $F_{\pi}=\{plan(a_i,i)\}_{\small 1\leq i\leq l}$ to code the $i^{th}$ action in $\mathcal{O}(\tau)$.
-The static facts $next_{i,i+1}$ and the fluents $at_i$, {\small $1\leq i< l$}, are also added to iterate through the $l$ actions in $\mathcal{O}(\tau)$.
-The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated.
-\item The fluents $mode_{edit}$ and $mode_{val}$ to indicate whether the operator schemata are edited or validated.
-\end{itemize}
-\item $I'$ extends the original initial state $I$ with the fluent $mode_{edit}$ set to true as well as the fluents $F_{\pi}$ plus fluents $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i<l$}, for tracking the plan step where the action model is validated.
-Our compilation assumes that initially $\mathcal{M}'$ is defined as $\mathcal{M}$.
-Therefore the {\em editable fluents} $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ hold as given by $\mathcal{M}$.
-
-\item $G'=G\bigcup\{at_n,test_m\}$.
-\item $A'$ comprises three kinds of actions with conditional effects:
-\begin{enumerate}
-\item The {\em editable} version of the original actions given by $\mathcal{M}$.
-This actions have now an extra preconditions because they can only be applied in the {\em validation} mode (i.e.
-when $mode_{val}$ holds).
-When the observation $\mathcal{O}(\tau)$ includes observed actions, they also include the extra conditional effects $\{at_{i},plan(a_i,i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,l]}$ to validate that actions are applied, exclusively, in the same order as they appear in $\mathcal{O}(\tau)$.\\
-
-\item Actions for {\em editing} operator schemata $\xi\in\mathcal{M}$.
-This includes the actions for {\em inserting} a new {\em precondition} into an action schemata $\xi\in\mathcal{M}$ and for inserting a new {\em negative} or {\em positive} effect into the action schemata $\xi\in\mathcal{M}$
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{e,\xi}})=&\{\neg pre\_e\_\xi, \neg del\_e\_\xi,\\
-& \neg add\_e\_\xi, mode_{edit}\},\\
-\cond(\mathsf{insertPre_{e,\xi}})=&\{\emptyset\}\rhd\{pre\_e\_\xi\}.
-\end{align*}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertEff_{e,\xi}})=&\{\neg del\_e\_\xi, \neg add\_e\_\xi,mode_{edit}\},\\
-\cond(\mathsf{insertEff_{e,\xi}})=&\{pre\_e\_\xi\}\rhd\{del\_e\_\xi\},\\
-&\{\neg pre\_e\_\xi\}\rhd\{add\_e\_\xi\}.
-\end{align*}
-\end{small}
-Besides these actions, $A'$ also contains the actions for {\em deleting} a {\em precondition} and a negative/positive {\em effect}.
-
-\item Actions for {\em validating} the edited models at the $s_j$ observed states, {\tt\small $0\leq j< m$}.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\\
-                            &\{mode_{edit}\}\rhd\{\neg mode_{edit}, mode_{val}\}.
-\end{align*}
-\end{small}
-\end{enumerate}
-\end{itemize}
-
-\begin{figure}
-  \begin{tiny}  
-  \begin{verbatim}
-(:action editable_inc-x
-  :parameters (?v1 ?v2)
-  :precondition
-    (and (or (not (pre_xcoord_v1_inc-x)) (xcoord ?v1))
-         (or (not (pre_xcoord_v2_inc-x)) (xcoord ?v2))
-         (or (not (pre_ycoord_v1_inc-x)) (xcoord ?v1))                       
-         (or (not (pre_ycoord_v2_inc-x)) (xcoord ?v2))
-         (or (not (pre_q0__inc-x)) (q0))
-         (or (not (pre_q1__inc-x)) (q1)))
-         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
-         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
-         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
-         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2))))
-    :effect (and
-       (when (del_xcoord_v1_inc-x) (not (xcoord ?v1)))
-       (when (del_xcoord_v2_inc-x) (not (xcoord ?v2)))
-       (when (del_ycoord_v1_inc-x) (not (xcoord ?v1)))
-       (when (del_ycoord_v2_inc-x) (not (xcoord ?v2)))
-       (when (del_q0__inc-x) (not (q0)))
-       (when (del_q1__inc-x) (not (q1)))
-       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
-       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
-       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
-       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))
-
-       (when (add_xcoord_v1_inc-x) (xcoord ?v1))
-       (when (add_xcoord_v2_inc-x) (xcoord ?v2))
-       (when (add_ycoord_v1_inc-x) (xcoord ?v1))
-       (when (add_ycoord_v2_inc-x) (xcoord ?v2))
-       (when (add_q0__inc-x) (q0))
-       (when (add_q1__inc-x) (q1))
-       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
-       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
-       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
-       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
-  \end{verbatim}           
-  \end{tiny}  
- \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schemata for robot navigation in a $N\times N$ grid.}
-\label{fig:editable}
-\end{figure}
+The value of the {\em observation distance} $\delta^o(\mathcal{M},\mathcal{O})$ is given by the number of {\em edit operations} (insertions and deletions) required by $\mathcal{M}$ to be validated in the input observation.
+In the case of the above example, the distance equals $7$.
 
@@
 \label{sec:evaluation}
+
 To evaluate the empirical performance of {\em model recognition as planning} we collect a set $M$ of possible \strips\ models, that share the same state variables but define different action models.
@@
 The experiment is repeated for models of different kind and different observability of the given plan execution.
+
+--------------------------
+
+INCLUIR EN ALGUN MOMENTO:
+
+\begin{definition}[$\Phi$-observation]
+Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
+\end{definition}
+
+TAMBIEN:
+
+Moreover, our definition of the {\em observation edit distance} does not accumulate mismatches between the observation and the model that are due to the same source (the same model flaw).
+
+--------------------------
 
@@
 Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
-A related approach is recently followed for {\em model reconciliation}~\cite{Kambhampati:mreconciliation:ijcai17} where model edition is used to conform the PDDL models of two different agents.
-Also  related to this paper is the work on {\em Goal Recognition Design} but in that work the action model is given in advance and fixed~\cite{keren2014goal}.
+A related approach is recently followed for {\em model reconciliation}~\cite{ChakrabortiSZK17} where model edition is used to conform the PDDL models of two different agents.
+Also  related to this paper is the work on {\em Goal Recognition Design} but in that work the action model is given in advance and fixed~\cite{KerenGK14}.
 
@@
 An interesting research direction is however to understand how to apply our approach to planning models where the planning models include actions with probabilistic effects~\cite{younes2005first}.
+
+\newpage
 

### 2018-11-09 doc/icaps-19/model-recognition.tex
Compilation scheme explained

@@
 This section shows that $\delta^o(\mathcal{M},\mathcal{O})$, and hence an approximation to $P(\mathcal{O}|\mathcal{M})$), can be computed with a compilation-to-planning approach as the one proposed in~\cite{aineto2018learning} (AJO approach hereafter) for learning \strips\ models.
-The AJO approach receives as input an empty model $\mathcal{M}$, which only contains the headers of the action schemata formed of $\xi=\tup{name(\xi),pars(\xi)}$, and a set of observations of plan executions (i.e., a set of  $\mathcal{O}(\tau)$), and it returns a model $\mathcal{M'}$ with specification of preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$ following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1},a_i)$ for every observed action of $\mathcal{O}(\tau)$ and $s_i=\theta(s_{i-1},a_i)$ for every observed state of $\mathcal{O}(\tau)$.
+The AJO approach receives as input an empty model $\mathcal{M}$, which only contains the headers of the action schemata formed of $\xi=\tup{name(\xi),pars(\xi)}$, and an observation of a plan execution $\mathcal{O}(\tau)$ (extensible to a set of observations), and it returns a model $\mathcal{M'}$ with specification of preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$ following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1},a_i)$ for every observed action of $\mathcal{O}(\tau)$ and $s_i=\theta(s_{i-1},a_i)$ for every observed state of $\mathcal{O}(\tau)$.
 
-In essence, the task $\tup{\mathcal{M},\mathcal{O}}$ of the AJO approach -- learning $\mathcal{M}$ that satisfies $\mathcal{O}$- can be interpreted as editing the empty action schemata of $\mathcal{M}$ introducing preconditions and effects until $\mathcal{O}$ is validated with the resulting model.
+In essence, the task $\tup{\mathcal{M},\mathcal{O}}$ of the AJO approach -- learning $\mathcal{M}$ that satisfies $\mathcal{O}$-- can be interpreted as editing the empty action schemata of $\mathcal{M}$ introducing preconditions and effects until $\mathcal{O}$ is validated with the resulting model.
 We leverage the same idea to compute $\delta^o(\mathcal{M},\mathcal{O})$ with the exception that now our $\mathcal{M}$ is not empty but $\mathcal{M}=\{\xi_1, \ldots, \xi_m\}$, where $\xi_i=\tup{name(\xi_i),pars(\xi_i),pre(\xi_i),add(\xi_i),del(\xi_i)}$, $1 \leq i \leq m$.
@@
 The intuition behind this compilation is that a solution plan to the problem is a sequence of: (a) \emph{edit actions} on the schemata of $\mathcal{M}$ to build $\mathcal{M'}$ and (b) \emph{validate actions} that apply $\mathcal{M'}$ in $\mathcal{O(\tau)}$.
-The adaptation of this compilation scheme for solving $\delta^o(\mathcal{M},\mathcal{O})$ results in a planning problem $P'=\tup{F',A',I',G'}$ whose objective is to determine the preconditions and effects of the action schemata of $\mathcal{M}$ that need to be added or deleted so as to satisfy $\mathcal{O}$.
+The adaptation of this compilation scheme for solving $\delta^o(\mathcal{M},\mathcal{O})$ results in a planning problem $P'=\tup{F',A',I',G'}$ whose objective is to determine the preconditions and effects that need to be added or deleted to the action schemata of $\mathcal{M}$ so as to satisfy $\mathcal{O}$.
 The accomplishment of this task requires therefore a propositional encoding of the components of the action schemata.
@@
 \item $F'$ contains the necessary fluents to represent the $pre$, $add$ and $del$ of the action schemata.
-Specifically, it is a set of \emph{editable fluents} of the type $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
-Additionally, $F$ also contains fluents to encode the observation constraints; that is, fluents to iterate through the $l$ observed actions and $m$ observed states of $\mathcal{O(\tau)}$.
+It is a set of \emph{editable fluents} of the type $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
+Additionally, $F$ also contains fluents to encode the \emph{observation constraints}; that is, fluents to iterate through the $l$ observed actions and $m$ observed states of $\mathcal{O(\tau)}$.
 \item $A'$ comprises two types of actions with conditional effects:
@@
     \item actions for editing $\xi \in \mathcal{M}$ that follow the \emph{syntactic constraints} for well-defined \strips\ action schemata.
-Hence, here are actions for inserting a new precondition $pre\_e\_\xi$ or a new positive effect $add\_e\_\xi$ in $\xi$ when neither $pre\_e\_\xi$, $del\_e\_\xi$ or $add\_e\_\xi$ exist in $\xi$, and actions for inserting a new negative effect when neither $del\_e\_\xi$ or $add\_e\_\xi$ appear in $\xi$ but $pre\_e\_\xi$ does.
+Hence, $A'$ contains actions for inserting a new precondition $pre\_e\_\xi$ or a new positive effect $add\_e\_\xi$ in $\xi$ when neither $pre\_e\_\xi$, $del\_e\_\xi$ or $add\_e\_\xi$ exist in $\xi$, and actions for inserting a new negative effect when neither $del\_e\_\xi$ or $add\_e\_\xi$ appear in $\xi$ but $pre\_e\_\xi$ does.
     \item actions for applying the new action schemata of the edited model $\mathcal{M'}$ and validating the observed states of $\mathcal{O(\tau)}$.
-Particularly, the \emph{apply} actions check the behavior of the propositional version of the new action schemata with the editable fluents of the current state.
+Particularly, the \emph{apply} actions check the preconditions and produce the effects defined by the \emph{editable fluents}.
     \end{itemize}
@@
 The plan found by a planner to $P'$ is shown in figure~\ref{fig:plan-pdistance}.
-The first seven steps are the edit actions to \emph{fix} the schema {\tt\small inc-x}; step 07 is a validate action that sets the robot in the state $s_0^o=(1,1)$; step 08 applies the action {\tt\small (inc-x 1 2)} and so moves the robot one cell to the right to $(2,1)$; steps 09-11 move also the robot one cell to the right; step 11 applies an action that increases coordinate $y$ to an even number, thus moving robot to row $2$; step 12 moves the robot one cell to the left and, finally, action {\tt\small (validate\_1)} checks the robot position is consistent with the final state $s_m^o=(3,2)$.
+The first seven steps are the edit actions to \emph{fix} the schema {\tt\small inc-x}; step 07 is a validate action that sets the robot in the initial state $s_0^o=(1,1)$; step 08 applies the action {\tt\small (inc-x 1 2)} and moves the robot one cell to the right to $(2,1)$; steps 09-11 also move the robot one cell to the right; step 11 applies an action that increases coordinate $y$ to an even number, thus moving robot to row $2$; step 12 moves the robot one cell to the left and, finally, action {\tt\small (validate\_1)} checks the robot position is consistent with the final state $s_m^o=(3,2)$.
 

### 2018-11-11 doc/icaps-19/model-recognition.tex
Evaluation section update

@@
 
-To evaluate the empirical performance of {\em model recognition as planning} we collect a set $M$ of possible \strips\ models, that share the same state variables but define different action models.
-Then, we randomly choose one of these models $\mathcal{M}\in M$ and use it to produce an observation $\mathcal{O}(\tau)$ of a plan execution.
+This section evaluates the empirical performance of {\em model recognition as planning}.
+In order to do so, we collect a set $M$ of planning models, that share the same state variables but define different action schemata, and generate observations $\mathcal{O}(\tau)$ of plan trajectories for each model.
 Finally, we follow our {\em model recognition as planning} method to identify the model $\mathcal{M}\in M$ that produced $\mathcal{O}(\tau)$.
-The experiment is repeated for models of different kind and different observability of the given plan execution.
+For all experiments, we will assume equal prior probabilities, meaning $P(\mathcal{M}) = \frac{1}{|M|}$ for all $\mathcal{M} \in M$.
 
---------------------------
-
-INCLUIR EN ALGUN MOMENTO:
-
+For the purpose of the experiments below, we introduce a new form of observation where only some types of fluents are observed.
+This new concept allows us to distinguish between observable data, which may come from, for example, sensors, and hidden or latent factors.
 \begin{definition}[$\Phi$-observation]
@@
 \end{definition}
+
+--------------------------
 
@@
 
-\subsubsection{Reproducibility.} {\sc Madagascar} is the classical planner we used to solve the instances that result from our compilations for its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
-Due to its SAT-based nature, {\sc Madagascar} can apply the actions for editing preconditions in a single planning step (in parallel) because there is no interaction between them.
-Actions for editing effects can also be applied in a single planning step, thus significantly reducing the planning horizon.
+\subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar} for its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
 
@@
 
-\subsubsection{Recognition of {\em Regular Automata}.} In this experiment the models in $M$ represent different {\em regular automata}.
-Figure~\ref{fig:regautomatae} illustrate a four-symbol four-state {\em regular automata} for recognizing the $\{(abc)^n : n \geq 1 \}$ language.
-The {\em input alphabet} is $\Sigma=\{a,b,c,\square\}$, and the machine states are $Q=\{q_0,q_1,q_2,\underline{q_3}\}$ (where \underline{$q_3$} is the only acceptor state).
+\subsubsection{Recognition of {\em Regular Automata}.} For our first experiment, which doubles as a proof of concept, we perform one of the most typical classification problems, string classification.
+In this experiment, the system receives a set of regular automata (coded as planning models) that represent the possible classes and a string (the observation), and assigns a class to the string.
 
-We generated a $M$ set of different \strips\ models that encode different {\em regular automata}.
-Each $\mathcal{M}\in M$ encodes a different {\em regular automata}.
-Each automata transition is encoded with a planning action.
-For instance, the execution of the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, with the sequence of input symbols $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
+Figure~\ref{fig:regautomatae} illustrates a three-symbol four-state {\em regular automata} for recognizing the $(abc)^+$ language.
+The {\em input alphabet} is $\Sigma=\{a,b,c\}$, and the machine states are $Q=\{q_0,q_1,q_2,\underline{q_3}\}$ (where \underline{$q_3$} is the only acceptor state).
+For instance, the execution of the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, with the string $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
 
-Here we assume that the actual applied transitions are unknown as well as the internal machine state.
+For this experiment, $M$ comprises 5 planning models representing regular automata with 5 states and an alphabet of 4 symbols. The regular languages defined by these automata are the following:
+\begin{itemize}
+	\item $\mathcal{L}$1: $a^+(b|c)d(dd)^*a$
+	\item $\mathcal{L}$2: $bd(abd)^*cd^*c^+$
+	\item $\mathcal{L}$3: $d^*c(ac)^*db^*ac^*b^+$
+	\item $\mathcal{L}$4: $(cc)^+bd(abd)^*$
+	\item $\mathcal{L}$5: $(d|a)a(ba)^*c^+d(dc^*d)^*$
+\end{itemize}
+
+We have generated 20 random strings (i.e.
+observations) of length 20 to 30 for each regular language for a total of 100 strings.
+In an attemp to bring this experiment closer to an actual string classification problem, we assume that observations only contain the string of symbols.
+This means that $\mathcal{O}(\tau)$ is a $\Phi$-observation where fluents representing the internal machine state are unknown and the applied transitions (actions) unobserved.
+
+%the actual applied transitions are unknown as well as the internal machine state.
 Assuming that the actual applied transitions is unknown means that the observation $\mathcal{O}(\tau)$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\mathcal{O}(\tau)=\tup{s_1, \ldots, s_m}$.
-Assuming that the internal machine state is unknown means that $\mathcal{O}(\tau)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$ and $q\neq q_0$.
+Assuming that the internal machine state is unknown means that $\mathcal{O}(\tau)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$.
 
@@
   \end{scriptsize}
- \caption{\small Four-symbol four-state {\em regular automata} for recognizing the $\{(abc)^n : n \geq 1 \}$ language (\underline{$q_3$} is the acceptor state).}
+ \caption{\small Four-symbol four-state {\em regular automata} for recognizing the $(abc)^+$ language (\underline{$q_3$} is the acceptor state).}
 \label{fig:regautomatae}
@@
 
-\subsubsection{Recognition of navigation models.} In this experiment the given models in $M$ represent different navigation models that are computed as the cross product of a regular automata with a four-operator \strips\ model for navigating a $N\times N$ grid.
+Table \ref{tab:conf_matrix} shows the confusion matrix resulting from classifying the 100 strings.
+In this matrix, rows represent the real class and columns the predicted class.
+Even though we are using a suboptimal planner, we can see that all strings were correctly labeled.
+These results proof that our approach is able to perform classic classification tasks.
 
-In this case the regular automata constrain the applications of the navigation actions producing different navigation policies e.g.
+\begin{table}[hbt!]
+	\centering
+	\begin{tabular}{c|c|c|c|c|c}
+		& $\mathcal{L}$1 & $\mathcal{L}$2 & $\mathcal{L}$3 & $\mathcal{L}$4 & $\mathcal{L}$5\\ \hline
+		$\mathcal{L}$1 & 20 & 0 & 0 & 0 & 0 \\
+		$\mathcal{L}$2 & 0 & 20 & 0 & 0 & 0 \\
+		$\mathcal{L}$3 & 0 & 0 & 20 & 0 & 0 \\
+		$\mathcal{L}$4 & 0 & 0 & 0 & 20 & 0 \\
+		$\mathcal{L}$5 & 0 & 0 & 0 & 0 & 20 \\
+	\end{tabular}
+	\caption{Confusion matrix of the regular automata classification task.}
+	\label{tab:conf_matrix}
+\end{table}
+
+\subsubsection{Recognition of navigation models.} In this experiment the given models in $M$ represent different navigation models, that is, different ways to navigate a $N\times N$ grid.
+In particular, we have constrained the applications of the navigation actions producing different navigation policies e.g.
 like the one in Figure~\ref{fig:model-example}.
-Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, here the task is to identify which navigation model produced that observation, despite the the applied actions are unobserved.
-In addition, for each observed state, only the value of fluents encoding the x and y coordinates of the agent are known while the value of the regular automata conditioning the navigation policy is unknown.
 
-\subsection{Results}
+Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, here the task is to identify which navigation model produced that observation, despite the the applied actions being unobserved.
+In addition, for each observed state, only the value of fluents encoding the x and y coordinates of the agent are known ($\Phi =$\{xcoord, ycoord\}) while the value of the hidden factors conditioning the navigation policy are unknown.
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=\linewidth]{figures/navigation.eps}
+	\caption{Classification accuracy for the navigation domain.}
+	\label{fig:navigation_acc}
+\end{figure}
 

### 2018-11-12 doc/icaps-19/model-recognition.tex
Updated evaluation section with navigation domain results

@@
 
---------------------------
-
-TAMBIEN:
-
-Moreover, our definition of the {\em observation edit distance} does not accumulate mismatches between the observation and the model that are due to the same source (the same model flaw).
-
---------------------------
-
 \subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar} for its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
@@
 
+\subsubsection{Recognition of failures in blocks-world domain}
+
 \subsubsection{Recognition of navigation models.} In this experiment the given models in $M$ represent different navigation models, that is, different ways to navigate a $N\times N$ grid.
-In particular, we have constrained the applications of the navigation actions producing different navigation policies e.g.
-like the one in Figure~\ref{fig:model-example}.
+In particular, we have taken a classical navigation domain with actions to move freely in each of the four directions (up, down, right, left), and constrained some of its actions to limit their applicability producing 8 different navigation policies.
+For example, the navigation policy in Figure~\ref{fig:model-example}, moves right in even rows and left in odd rows, producing a zig-zag pattern when navigating the grid.
 
-Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, here the task is to identify which navigation model produced that observation, despite the the applied actions being unobserved.
+By solving the planning problem of visiting all the tiles in a $5\times 5$ grid with these navigation models, we produced 8 different plan trajectories depicting the path followed by each model to traverse the grid.
+As in previous experiments, the observations contain no actions, and additionally, for each observed state, only the value of fluents encoding the x and y coordinates of the agent are known ($\Phi =$\{xcoord, ycoord\})
+
+% Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, here the task is to identify which navigation model produced that observation, despite the the applied actions being unobserved.
 In addition, for each observed state, only the value of fluents encoding the x and y coordinates of the agent are known ($\Phi =$\{xcoord, ycoord\}) while the value of the hidden factors conditioning the navigation policy are unknown.
+
+Similar to the previous experiment, we measure the accuracy of our approach in identifying the navigation model that generated each observation under different degrees of observability.
+The novel aspect of this experiment is that now we do not garantee that observed states contain at least one literal, meaning that some states may be missing.
+In this scenario more than a single action may be required to reach one observed state from another.
+
+Another interesting aspect of this experiment is that all navigation policies are at a maximum distance of 4 from the base navigation model which can move in any direction at any given state.
+This means that, for any given model and observation, we will need to perform a maximum of 4 editions to the input model to explain the observation.
+This aspect heavily constrains the discriminating power of the $P(\mathcal{O}|\mathcal{M})$ likelihood, and added to the low observability of the states produces a very harsh benchmark for our approach.
 
@@
 \end{figure}
+
+Figure \ref{fig:navigation_acc} shows the accuracy of our approach for degrees of observability from 0\% to 100\% with 10\% increments.
+Note that in this experiment we include the 0\% case which corresponds to observations where only the initial and final states are observed.
+The figure shows that for 0\% observability we were unable to unmistakenly identify the generating navigation models, but starting from 10\% we where able to correctly classify most of the observations.
+Accuracy stabilizes in the range 0.875 to 1 which means that, at most, only 1 observation of the 8 was not correctly classified.
 

### 2018-11-13 doc/icaps-19/model-recognition.tex
first two experiments in evaluation section revised

--- 2018-11-11 doc/icaps-19/model-recognition.tex
@@
 This section evaluates the empirical performance of {\em model recognition as planning}.
-In order to do so, we collect a set $M$ of planning models, that share the same state variables but define different action schemata, and generate observations $\mathcal{O}(\tau)$ of plan trajectories for each model.
-Finally, we follow our {\em model recognition as planning} method to identify the model $\mathcal{M}\in M$ that produced $\mathcal{O}(\tau)$.
-For all experiments, we will assume equal prior probabilities, meaning $P(\mathcal{M}) = \frac{1}{|M|}$ for all $\mathcal{M} \in M$.
+To do so, we define a set $M$ of different planning models, that share the same state variables but that update these variables according to different action schemata.
+Then for each model $\mathcal{M}\in M$, we generate {\em partial observations} $\mathcal{O}(\tau)$ of {\em plan trajectories} that are computed with that model.
+Finally, we apply our {\em model recognition as planning} method to identify which model, among on the models in $M$, was actually used for generating a given $\mathcal{O}(\tau)$ observation.
+For all the experiments, we assume equal prior probabilities, meaning $P(\mathcal{M}) = \frac{1}{|M|}$ for every $\mathcal{M} \in M$.
 
-For the purpose of the experiments below, we introduce a new form of observation where only some types of fluents are observed.
-This new concept allows us to distinguish between observable data, which may come from, for example, sensors, and hidden or latent factors.
+For the purpose of the experiments below, we introduce a particular class of $\mathcal{O}(\tau)$ observations.
+This new class allows us to distinguish between observable data, which may come from sensors, and hidden or latent factors that cannot be observed.
 \begin{definition}[$\Phi$-observation]
@@
 
---------------------------
-
-TAMBIEN:
-
-Moreover, our definition of the {\em observation edit distance} does not accumulate mismatches between the observation and the model that are due to the same source (the same model flaw).
-
---------------------------
-
-\subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar} for its ability to deal with dead-ends~\cite{rintanen2014madagascar}.
+\subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar}~\cite{rintanen2014madagascar} because of its ability to deal with classical planning problems with dead-ends~\cite{lopez2015deterministic}.
 
@@
 
-\subsubsection{Recognition of {\em Regular Automata}.} For our first experiment, which doubles as a proof of concept, we perform one of the most typical classification problems, string classification.
-In this experiment, the system receives a set of regular automata (coded as planning models) that represent the possible classes and a string (the observation), and assigns a class to the string.
+\subsection{Recognition of {\em Regular Automata}.} The first experiment, which doubles as a proof of concept, leverages {\em model recognition as planning} to perform a typical {\em string classification} problem.
+In this experiment, the system receives (1) the string to classify (the $\mathcal{O}(\tau)$ observation) and (2), a set $M$ of different planning models, where each $\mathcal{M}\in M$ codes a different class (i.e.
+a different regular automata that accepts any string that belong to the class).
 
@@
 The {\em input alphabet} is $\Sigma=\{a,b,c\}$, and the machine states are $Q=\{q_0,q_1,q_2,\underline{q_3}\}$ (where \underline{$q_3$} is the only acceptor state).
-For instance, the execution of the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, with the string $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
-
-For this experiment, $M$ comprises 5 planning models representing regular automata with 5 states and an alphabet of 4 symbols. The regular languages defined by these automata are the following:
-\begin{itemize}
-	\item $\mathcal{L}$1: $a^+(b|c)d(dd)^*a$
-	\item $\mathcal{L}$2: $bd(abd)^*cd^*c^+$
-	\item $\mathcal{L}$3: $d^*c(ac)^*db^*ac^*b^+$
-	\item $\mathcal{L}$4: $(cc)^+bd(abd)^*$
-	\item $\mathcal{L}$5: $(d|a)a(ba)^*c^+d(dc^*d)^*$
-\end{itemize}
-
-We have generated 20 random strings (i.e.
-observations) of length 20 to 30 for each regular language for a total of 100 strings.
-In an attemp to bring this experiment closer to an actual string classification problem, we assume that observations only contain the string of symbols.
-This means that $\mathcal{O}(\tau)$ is a $\Phi$-observation where fluents representing the internal machine state are unknown and the applied transitions (actions) unobserved.
-
-%the actual applied transitions are unknown as well as the internal machine state.
-Assuming that the actual applied transitions is unknown means that the observation $\mathcal{O}(\tau)$ of the execution of a regular automata contains no actions, it is simply a sequence of states $\mathcal{O}(\tau)=\tup{s_1, \ldots, s_m}$.
-Assuming that the internal machine state is unknown means that $\mathcal{O}(\tau)$ is a $\Phi$-observation and that the $\Phi$ subset does not contain {\small\tt ($q$)} fluents, with $q\in Q$.
+For instance, the execution of the planning model coding the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, and with the input string $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
 
@@
 	  \node[state] (C) [right of=B] {$q_2$};
-	  \node[state] (D) [right of=C] {$q_3$};
+	  \node[state] (D) [right of=C] {\underline{$q_3$}};
 	  \path
@@
   \end{scriptsize}
- \caption{\small Four-symbol four-state {\em regular automata} for recognizing the $(abc)^+$ language (\underline{$q_3$} is the acceptor state).}
+ \caption{\small Three-symbol four-state {\em regular automata} for recognizing the $(abc)^+$ language (\underline{$q_3$} is the acceptor state).}
 \label{fig:regautomatae}
@@
 
-Table \ref{tab:conf_matrix} shows the confusion matrix resulting from classifying the 100 strings.
-In this matrix, rows represent the real class and columns the predicted class.
-Even though we are using a suboptimal planner, we can see that all strings were correctly labeled.
-These results proof that our approach is able to perform classic classification tasks.
+Note that the the string to classify is a {\em $\Phi$-observation} since it does not contain any information about the structure of the possible regular automata.
+This means that $\mathcal{O}(\tau)$ is a $\Phi$-observation where fluents representing the internal machine state are unknown and the applied transitions (actions) are unobserved.
+
+In this experiment, $M$ comprises 5 planning models representing different regular automata defined with 5 states and a 4-symbol alphabet.
+The regular languages defined by these five automata are the following:
+\begin{itemize}
+	\item $\mathcal{L}$1: $a^+(b|c)d(dd)^*a$
+	\item $\mathcal{L}$2: $bd(abd)^*cd^*c^+$
+	\item $\mathcal{L}$3: $d^*c(ac)^*db^*ac^*b^+$
+	\item $\mathcal{L}$4: $(cc)^+bd(abd)^*$
+	\item $\mathcal{L}$5: $(d|a)a(ba)^*c^+d(dc^*d)^*$
+\end{itemize}
+
+For each regular language, we generated $20$ random strings (observations) of length 20 to 30  for a total of 100 strings (observations).
+Table \ref{tab:conf_matrix} shows the confusion matrix resulting from classifying these 100 strings with {\em model recognition as planning}.
+In this matrix, rows represent the actual class and columns the class predicted by our {\em model recognition as planning} approach.
+Even though we are using a suboptimal classical planner, we can see that the class for all the 100 input strings was correctly recognized.
+These results proof that {\em model recognition as planning} is suitable for classification tasks where classes are given as generative planning models.
 
@@
 	\end{tabular}
-	\caption{Confusion matrix of the regular automata classification task.}
+	\caption{Confusion matrix for {\em regular automata recognition}.}
 	\label{tab:conf_matrix}
@@
 
-\subsubsection{Recognition of navigation models.} In this experiment the given models in $M$ represent different navigation models, that is, different ways to navigate a $N\times N$ grid.
-In particular, we have constrained the applications of the navigation actions producing different navigation policies e.g.
-like the one in Figure~\ref{fig:model-example}.
+\subsection{Recognition of failures in non-deterministic {\em blocksworld} models}
 
-Given an observation of a plan execution, like the one illustrated at Figure~\ref{fig:grid-example}, here the task is to identify which navigation model produced that observation, despite the the applied actions being unobserved.
-In addition, for each observed state, only the value of fluents encoding the x and y coordinates of the agent are known ($\Phi =$\{xcoord, ycoord\}) while the value of the hidden factors conditioning the navigation policy are unknown.
+\subsection{Recognition of navigation models.} In this experiment the planning models given in $M$ represent different navigation models for $N\times N$ grids.
+In particular, we took a {\em classical navigation model} with actions {\tt\small up, down, right, left} (to move one cell in each of these four directions) and use it to define 8 different navigation policies with respect to these additional state variables {\tt\small \{(q0), (q1), (min ?v), (max ?v)\}}.
+An example is the navigation policy, shown in Figure~\ref{fig:model-example}.
+This policy allows to move right when $q_0$ holds while it allows to move left when $q_1$ holds, producing a zig-zag pattern when navigating for visiting all the cells of $N\times N$ grids.
+
+On the other hand we generated observations from 8 different trajectories, one for each of the 8 previous planning models.
+The trajectories depict the path followed by each navigation model to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
+As in the previous experiments, observations contain no actions and here, for each observed state, they only contain the value of the fluents encoding the {\em x and y coordinates} of the agent are known (i.e.
+$\Phi =$\{{\small\tt xcoord, ycoord}\}).
+
+Again, this experiment evaluates the accuracy of {\em model recognition as planning} to identify the model that generated each observation under different degrees of partial observability.
+The novel aspect of this experiment is that now, we do not guarantee that observed states contain at least one literal, meaning that some intermediate states may be missing.
+Therefore, in this new scenario more than a single action may be required to reach one observed state from another.
+
+Another interesting aspect of this experiment is that all navigation policies are at a maximum distance of just 4 editions from the base {\em classical navigation model} that can move in any direction at any given state.
+This means that, for any given model and observation, we will need to perform a maximum of 4 editions to the {\em classical navigation model} that can explain any observation.
+This aspect heavily constrains the discriminating power of the $P(\mathcal{O}|\mathcal{M})$ likelihood, and added to the low observability of the states produces a very benchmark for {\em model recognition as planning}.
+
+Figure \ref{fig:navigation_acc} shows the classification {\em accuracy} achieved by {\em model recognition as planning} with respect to a range of different degrees of observability, from 0\% to 100\% (full observability) with 10\% increments.
+In this experiment we included the 0\% case which corresponds to observations where only the initial and final states are observed.
+The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation models, but starting from 10\% we start to correctly classify most of the observations.
+Accuracy stabilizes in the range 0.875 to 1 which means that, at most, only 1 observation of the 8 was not correctly classified.
 
@@
 	\centering
-	\includegraphics[width=\linewidth]{figures/navigation.eps}
-	\caption{Classification accuracy for the navigation domain.}
+	\includegraphics[width=0.9\linewidth]{figures/navigation.eps}
+	\caption{Classification accuracy for the recognition of navigation models.}
 	\label{fig:navigation_acc}

### 2018-11-13 doc/icaps-19/model-recognition.tex
Updated evaluation section for blocks domain (WIP)

@@
 
-\subsection{Recognition of failures in non-deterministic {\em blocksworld} models}
+\subsection{Recognition of failures in {\em blocksworld} domain}
+For our second experiments we have generated a set of observations for the {\em blocksworld} domain where some of the actions may fail. In more detail, the failures we consider are:
+
+\begin{itemize}
+	\item Stack fails and does nothing.
+	\item Unstack fails and does nothing.
+	\item Unstack fails and drops the block on the table.
+\end{itemize}
+
+We limited the errors to one type of error per observation, but it would be possible to recognize combinations of errors.
+The input models extend a 4-schemata {\em blocksworld} domain with an additional action schema that encodes one of the above failures.
+This means that the failure can be identified by finding the model that better explains the observation.
+
+Our aim with this experiment is to validate the assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood.
 

### 2018-11-15 doc/icaps-19/model-recognition.tex
Added failure detection in blocksworld to the evaluation section

@@
 Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action models), {\em model recognition} is the task of identifying the model that explains the observation.
-The paper formalizes the {\em model recognition} task and introduces a novel method to estimate the probability of a \strips\ model to produce an observation of a plan execution.
+The paper formalizes this task and introduces a novel method to estimate the probability of a \strips\ model that produce an observation of a plan execution.
 This method builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and intermediate states in the observation.
-The effectiveness of the method is shown in a set of \strips\ models encoding different kinds of {\em automata}.
-The presented approach succeeds to identify the executed automata despite the actual applied transitions, plus some state variables e.g.
-the internal automata state, are unobserved.
+The effectiveness of the method is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a classical string classification task; (2) identification of the model that causes a failure in an observation; and (3) recognition of a robot navigation policy.
 \end{abstract}
@@
 Then for each model $\mathcal{M}\in M$, we generate {\em partial observations} $\mathcal{O}(\tau)$ of {\em plan trajectories} that are computed with that model.
-Finally, we apply our {\em model recognition as planning} method to identify which model, among on the models in $M$, was actually used for generating a given $\mathcal{O}(\tau)$ observation.
+Finally, we apply our {\em model recognition as planning} method to identify which model, among the models in $M$, was actually used for generating a given $\mathcal{O}(\tau)$ observation.
 For all the experiments, we assume equal prior probabilities, meaning $P(\mathcal{M}) = \frac{1}{|M|}$ for every $\mathcal{M} \in M$.
@@
 \subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar}~\cite{rintanen2014madagascar} because of its ability to deal with classical planning problems with dead-ends~\cite{lopez2015deterministic}.
+We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned maximum distance.
 
-The compilation source code, evaluation scripts and benchmarks (including the used training and test sets) are fully available at this anonymous repository {\em } so any experimental data reported in the paper can be reproduced.
+The compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em } so any experimental data reported in the paper can be reproduced.
 
@@
 
-Note that the the string to classify is a {\em $\Phi$-observation} since it does not contain any information about the structure of the possible regular automata.
+%Note that the the string to classify is a {\em $\Phi$-observation} since it does not contain any information about the structure of the possible regular automata.
 This means that $\mathcal{O}(\tau)$ is a $\Phi$-observation where fluents representing the internal machine state are unknown and the applied transitions (actions) are unobserved.
@@
 
-For each regular language, we generated $20$ random strings (observations) of length 20 to 30  for a total of 100 strings (observations).
+For each regular language, we generated $20$ random strings (observations) of length 20 to 30  for a total of 100 strings.
+These observations contain no information about the structure of the possible automata.
+This means that $\mathcal{O}(\tau)$ is a $\Phi$-observation where fluents representing the internal machine state are unknown and the applied transitions (actions) are unobserved.
+
 Table \ref{tab:conf_matrix} shows the confusion matrix resulting from classifying these 100 strings with {\em model recognition as planning}.
@@
 \subsection{Recognition of failures in {\em blocksworld} domain}
-For our second experiments we have generated a set of observations for the {\em blocksworld} domain where some of the actions may fail. In more detail, the failures we consider are:
+For our second experiment we have generated a set of observations for the {\em blocksworld} domain where some of the actions may fail. In more detail, the failures we consider are:
 
@@
 
-We limited the errors to one type of error per observation, but it would be possible to recognize combinations of errors.
+We limited the failures to one type per observation, but it would be possible to recognize combinations of failures.
 The input models extend a 4-schemata {\em blocksworld} domain with an additional action schema that encodes one of the above failures.
@@
 Our aim with this experiment is to validate the assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood.
+These assumption being that the agent acts rational, so the computation of the $P(\mathcal{O}|\mathcal{M})$ likelihood is dominated by the shortest trajectory that explains the observation, and that there exists a single such trajectory.
+In order to proof this, we assume that the length of the real trajectory is known, thus forcing the explanation to be optimal, and measure the accuracy of our approach across different degrees of observability.
+The accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a literal in an intermediate state.
+Note that this definition allows for missing states (no literals were observed), but for this experiment we will make it so at least one literal was observed at every state so the length of the plan trajectory can be known from the observation.
+If our assumptions hold, we expect the results of this experiment to show high accuracy and to find an increasing trend as observability increases and the number of possible explanations is reduced.
 
-\subsection{Recognition of navigation models.} In this experiment the planning models given in $M$ represent different navigation models for $N\times N$ grids.
+%In this experiment, we measure the accuracy of our approach.
+Accuracy is defined as the number of correct predictions over all predictions made so, for this experiment, it translates to the percentage of correctly identified failures in the observations.
+
+%We measure the accuracy across different degrees of observability, which indicate the probability of observing a literal in an intermediate state.
+Note that this definition allows for missing states (no literals were observed), but for this experiment we will make it so at least one literal was observed at every state so the length of the plan trajectory can be known from the observation, and in the next experiment we will forego this constraint.
+By knowing the length of the real trajectory we force the trajectory found by our approach to be rational, i.e, the shortest explanation for the observation.
+If our assumption holds, we expect the results to
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=0.9\linewidth]{figures/blocks.eps}
+	\caption{Accuracy for the identification of failures in the \em{blocksworld} domain.}
+	\label{fig:blocks_acc}
+\end{figure}
+
+Figure \ref{fig:blocks_acc} shows the accuracy of our approach in identifying failures over 30 observations.
+The figure showcases a positive trend proving that the more complete the observations are the more accurate our stimation becomes.
+The overall high accuracy also validates our assumption that the trajectory that better explains the observation is the shortest one.
+
+\subsection{Recognition of navigation policies.} In this experiment the planning models given in $M$ represent different navigation models for $N\times N$ grids.
 In particular, we took a {\em classical navigation model} with actions {\tt\small up, down, right, left} (to move one cell in each of these four directions) and use it to define 8 different navigation policies with respect to these additional state variables {\tt\small \{(q0), (q1), (min ?v), (max ?v)\}}.
 An example is the navigation policy, shown in Figure~\ref{fig:model-example}.
-This policy allows to move right when $q_0$ holds while it allows to move left when $q_1$ holds, producing a zig-zag pattern when navigating for visiting all the cells of $N\times N$ grids.
+This policy allows to move right when $q_0$ holds while it allows to move left when $q_1$ holds, producing a zig-zag pattern when visiting all the cells of $N\times N$ grids.
 
 On the other hand we generated observations from 8 different trajectories, one for each of the 8 previous planning models.
-The trajectories depict the path followed by each navigation model to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
-As in the previous experiments, observations contain no actions and here, for each observed state, they only contain the value of the fluents encoding the {\em x and y coordinates} of the agent are known (i.e.
+The trajectories depict the path followed by each navigation policy to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
+As in the previous experiments, observations contain no actions and here, for each observed state, only the values of the fluents encoding the {\em x and y coordinates} of the agent are known (i.e.
 $\Phi =$\{{\small\tt xcoord, ycoord}\}).
 
-Again, this experiment evaluates the accuracy of {\em model recognition as planning} to identify the model that generated each observation under different degrees of partial observability.
+Again, this experiment evaluates the accuracy of {\em model recognition as planning} to identify the model that generated each observation under different degrees of observability.
 The novel aspect of this experiment is that now, we do not guarantee that observed states contain at least one literal, meaning that some intermediate states may be missing.
@@
 Another interesting aspect of this experiment is that all navigation policies are at a maximum distance of just 4 editions from the base {\em classical navigation model} that can move in any direction at any given state.
-This means that, for any given model and observation, we will need to perform a maximum of 4 editions to the {\em classical navigation model} that can explain any observation.
+This means that, for any given model and observation, we will need to perform a maximum of 4 editions to the input model to explain the observation.
 This aspect heavily constrains the discriminating power of the $P(\mathcal{O}|\mathcal{M})$ likelihood, and added to the low observability of the states produces a very benchmark for {\em model recognition as planning}.
 
-Figure \ref{fig:navigation_acc} shows the classification {\em accuracy} achieved by {\em model recognition as planning} with respect to a range of different degrees of observability, from 0\% to 100\% (full observability) with 10\% increments.
+Figure \ref{fig:navigation_acc} shows the classification {\em accuracy} achieved by {\em model recognition as planning} with respect to a range of different degrees of observability, from 0\% to 100\% with 10\% increments.
 In this experiment we included the 0\% case which corresponds to observations where only the initial and final states are observed.
-The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation models, but starting from 10\% we start to correctly classify most of the observations.
+The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation policies, but starting from 10\% we start to correctly classify most of the observations.
 Accuracy stabilizes in the range 0.875 to 1 which means that, at most, only 1 observation of the 8 was not correctly classified.

### 2018-11-15 doc/icaps-19/model-recognition.tex
Updated abstract

@@
 Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action models), {\em model recognition} is the task of identifying the model that explains the observation.
-The paper formalizes this task and introduces a novel method to estimate the probability of a \strips\ model that produce an observation of a plan execution.
+The paper formalizes this task and introduces a novel method that estimates the probability of a \strips\ model to produce an observation of a plan execution.
 This method builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and intermediate states in the observation.
-The effectiveness of the method is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a classical string classification task; (2) identification of the model that causes a failure in an observation; and (3) recognition of a robot navigation policy.
+The effectiveness of the method is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a classical string classification task; (2) identification of the model that encodes a failure present in an observation; and (3) recognition of a robot navigation policy.
 \end{abstract}

### 2018-11-15 doc/icaps-19/model-recognition.tex
evaluation revised icaps19 paper

--- 2018-11-15 doc/icaps-19/model-recognition.tex
@@
 
-This section evaluates the empirical performance of {\em model recognition as planning}.
-To do so, we define a set $M$ of different planning models, that share the same state variables but that update these variables according to different action schemata.
-Then for each model $\mathcal{M}\in M$, we generate {\em partial observations} $\mathcal{O}(\tau)$ of {\em plan trajectories} that are computed with that model.
+This section presents three different applications of {\em model recognition} and evaluates the empirical performance of {\em model recognition as planning} when addressing the three of them.
+
+In all the experiments we define a set $M$ of different planning models, that share the same state variables but that update these variables according to different action schemata.
+Then for each model $\mathcal{M}\in M$, we generate {\em partial observations} $\mathcal{O}(\tau)$ of {\em plan trajectories} that were computed with that model.
 Finally, we apply our {\em model recognition as planning} method to identify which model, among the models in $M$, was actually used for generating a given $\mathcal{O}(\tau)$ observation.
@@
 
-For the purpose of the experiments below, we introduce a particular class of $\mathcal{O}(\tau)$ observations.
-This new class allows us to distinguish between observable data, which may come from sensors, and hidden or latent factors that cannot be observed.
+To provide a better explanation of the experiments below, we introduce a particular class of $\mathcal{O}(\tau)$ observations.
+This new class allows us to distinguish between {\em observable} state variables, whose value may come from sensors, and {\em hidden} or {\em latent} state variables that cannot be observed.
 \begin{definition}[$\Phi$-observation]
@@
 \end{definition}
+Therefore the value of variables in $\Phi\subseteq F$ can be {\em observed} while the value of the variables in $F\setminus\Phi$ is {\em hidden}.
 
@@
 We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned maximum distance.
+Compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em } so any experimental data reported in the paper can be reproduced.
 
-The compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em } so any experimental data reported in the paper can be reproduced.
-
-\subsection{Recognition of {\em Regular Automata}.} The first experiment, which doubles as a proof of concept, leverages {\em model recognition as planning} to perform a typical {\em string classification} problem.
+\subsection{Recognition of {\em regular automata}.} The first experiment, which doubles as a proof of concept, leverages {\em model recognition as planning} to perform a typical {\em string classification} problem.
 In this experiment, the system receives (1) the string to classify (the $\mathcal{O}(\tau)$ observation) and (2), a set $M$ of different planning models, where each $\mathcal{M}\in M$ codes a different class (i.e.
@@
 The {\em input alphabet} is $\Sigma=\{a,b,c\}$, and the machine states are $Q=\{q_0,q_1,q_2,\underline{q_3}\}$ (where \underline{$q_3$} is the only acceptor state).
-For instance, the execution of the planning model coding the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, and with the input string $abcabc$, produces the following six-action plan {\small $(a,q_0\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$, $(a,q_3\rightarrow q_1)$, $(b,q_1\rightarrow q_2)$, $(c,q_2\rightarrow q_3)$}.
+For instance, the execution of the planning model coding the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, and with the input string $abcabc$, produces the following six-action plan {\small $(\langle a,q_0\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow \underline{q_3})$, $(\langle a,\underline{q_3}\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow \underline{q_3})$}.
 
@@
 In this experiment, $M$ comprises 5 planning models representing different regular automata defined with 5 states and a 4-symbol alphabet.
-The regular languages defined by these five automata are the following:
+The five regular languages defined by these five automata are the following:
 \begin{itemize}
@@
 
-For each regular language, we generated $20$ random strings (observations) of length 20 to 30  for a total of 100 strings.
-These observations contain no information about the structure of the possible automata.
-This means that $\mathcal{O}(\tau)$ is a $\Phi$-observation where fluents representing the internal machine state are unknown and the applied transitions (actions) are unobserved.
-
-Table \ref{tab:conf_matrix} shows the confusion matrix resulting from classifying these 100 strings with {\em model recognition as planning}.
+For each regular language, we generated $20$ random strings, $\mathcal{O}(\tau)$ observations, of length 20 to 30 for a total of 100 strings.
+More precisely observations $\mathcal{O}(\tau)$ are $\Phi$-observations where: the applied automata transitions and the fluents representing the internal automata state are unobserved.
+Table \ref{tab:conf_matrix} shows the {\em confusion matrix} resulting from classifying these 100 strings with {\em model recognition as planning}.
 In this matrix, rows represent the actual class and columns the class predicted by our {\em model recognition as planning} approach.
 Even though we are using a suboptimal classical planner, we can see that the class for all the 100 input strings was correctly recognized.
-These results proof that {\em model recognition as planning} is suitable for classification tasks where classes are given as generative planning models.
+These results proof the suitability of {\em model recognition as planning} for classification tasks where classes are given as generative planning models and examples are sequence of observations.
 
-\begin{table}[hbt!]
+\begin{table}
 	\centering
@@
 
-\subsection{Recognition of failures in {\em blocksworld} domain}
-For our second experiment we have generated a set of observations for the {\em blocksworld} domain where some of the actions may fail. In more detail, the failures we consider are:
+\subsection{Recognizing failures in a non-deterministic {\em blocksworld}}
+With this second experiment we aim to validate the assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood.
+These assumptions being that the agent acts rational, so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation, and that there exists a single such trajectory.
 
+In this particular experiment we assume that the length of the trajectory is known (to force the explanation to be optimal) and hence, measure the {\em  accuracy} of our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood across different degrees of observability.
+More precisely, we assume that at least one literal is observed at every state so the length of the plan trajectory is fixed by the observation.
+
+The input observations $\mathcal{O}(\tau)$ come from a non-deterministic {\em blocksworld} in which some of the executed actions failed. We considered failures of three kinds:
 \begin{itemize}
-	\item Stack fails and does nothing.
-	\item Unstack fails and does nothing.
-	\item Unstack fails and drops the block on the table.
+	\item $\mathcal{F}$1: the execution of a {\tt\small stack} action fails causing no effect.
+	\item $\mathcal{F}$2: {\tt\small unstack} fails dropping the block to unstack on the table.        
+	\item $\mathcal{F}$3: {\tt\small unstack} fails doing nothing.
 \end{itemize}
+The input models in $M$ are then three different {\em extended} 4-operator {\em blocksworlds} such that each of these models includes an additional action schema that encodes one of the three failures above.
+This means that failures are identified by finding the model that explains the best the input observation.
 
-We limited the failures to one type per observation, but it would be possible to recognize combinations of failures.
-The input models extend a 4-schemata {\em blocksworld} domain with an additional action schema that encodes one of the above failures.
-This means that the failure can be identified by finding the model that better explains the observation.
+Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations.
+{\em Accuracy} is defined as the number of correct predictions over all predictions made while the {\em degree of observability} indicates the probability of observing a literal in an intermediate state.
+The figure showcases a positive trend proving that the more complete the observations are the more accurate our estimation becomes.
+The overall high accuracy validates our assumption that the trajectory that better explains the observation is the shortest one.
+%We limited the failures to be of one kind per observation however, it would be possible to recognize different combinations of them.
 
-Our aim with this experiment is to validate the assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood.
-These assumption being that the agent acts rational, so the computation of the $P(\mathcal{O}|\mathcal{M})$ likelihood is dominated by the shortest trajectory that explains the observation, and that there exists a single such trajectory.
-In order to proof this, we assume that the length of the real trajectory is known, thus forcing the explanation to be optimal, and measure the accuracy of our approach across different degrees of observability.
-The accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a literal in an intermediate state.
-Note that this definition allows for missing states (no literals were observed), but for this experiment we will make it so at least one literal was observed at every state so the length of the plan trajectory can be known from the observation.
-If our assumptions hold, we expect the results of this experiment to show high accuracy and to find an increasing trend as observability increases and the number of possible explanations is reduced.
-
-%In this experiment, we measure the accuracy of our approach.
-Accuracy is defined as the number of correct predictions over all predictions made so, for this experiment, it translates to the percentage of correctly identified failures in the observations.
-
-%We measure the accuracy across different degrees of observability, which indicate the probability of observing a literal in an intermediate state.
-Note that this definition allows for missing states (no literals were observed), but for this experiment we will make it so at least one literal was observed at every state so the length of the plan trajectory can be known from the observation, and in the next experiment we will forego this constraint.
-By knowing the length of the real trajectory we force the trajectory found by our approach to be rational, i.e, the shortest explanation for the observation.
-If our assumption holds, we expect the results to
-
-\begin{figure}[hbt!]
+\begin{figure}
 	\centering
 	\includegraphics[width=0.9\linewidth]{figures/blocks.eps}
-	\caption{Accuracy for the identification of failures in the \em{blocksworld} domain.}
+	\caption{Accuracy for the recognition of failures in a non-deterministic {\em blocksworld}.}
 	\label{fig:blocks_acc}
@@
 
-Figure \ref{fig:blocks_acc} shows the accuracy of our approach in identifying failures over 30 observations.
-The figure showcases a positive trend proving that the more complete the observations are the more accurate our stimation becomes.
-The overall high accuracy also validates our assumption that the trajectory that better explains the observation is the shortest one.
+\subsection{Recognition of navigation policies}
+Again, this experiment evaluates the accuracy of {\em model recognition as planning} to identify the model that generated each observation under different {\em degrees of observability}.
+The novel aspect of this experiment is that now, we do not guarantee that observed states contain at least one literal, meaning that some intermediate states may be missing.
+Therefore, in this new scenario more than a single action may be required to reach one observed state from another.
 
-\subsection{Recognition of navigation policies.} In this experiment the planning models given in $M$ represent different navigation models for $N\times N$ grids.
-In particular, we took a {\em classical navigation model} with actions {\tt\small up, down, right, left} (to move one cell in each of these four directions) and use it to define 8 different navigation policies with respect to these additional state variables {\tt\small \{(q0), (q1), (min ?v), (max ?v)\}}.
+In this experiment the planning models given in $M$ represent different navigation models for $N\times N$ grids.
+In particular, we took a {\em classical navigation model} with actions {\tt\small up, down, right, left} (to move one cell in each of these four directions) and extend it to define 8 different navigation policies with respect to these additional state variables {\tt\small \{(q0), (q1), (min ?v), (max ?v)\}}.
 An example is the navigation policy, shown in Figure~\ref{fig:model-example}.
@@
 
+Note that all navigation policies are at a maximum distance of just 4 editions from the base {\em classical navigation model} that can move in any direction at any given state.
+This aspect constrains the discriminating power of the $P(\mathcal{O}|\mathcal{M})$ likelihood, and added to the low observability of the states, produces a touchstone for {\em model recognition as planning}.
+
 On the other hand we generated observations from 8 different trajectories, one for each of the 8 previous planning models.
-The trajectories depict the path followed by each navigation policy to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
+The trajectories depict paths followed by each navigation policy to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
 As in the previous experiments, observations contain no actions and here, for each observed state, only the values of the fluents encoding the {\em x and y coordinates} of the agent are known (i.e.
 $\Phi =$\{{\small\tt xcoord, ycoord}\}).
-
-Again, this experiment evaluates the accuracy of {\em model recognition as planning} to identify the model that generated each observation under different degrees of observability.
-The novel aspect of this experiment is that now, we do not guarantee that observed states contain at least one literal, meaning that some intermediate states may be missing.
-Therefore, in this new scenario more than a single action may be required to reach one observed state from another.
-
-Another interesting aspect of this experiment is that all navigation policies are at a maximum distance of just 4 editions from the base {\em classical navigation model} that can move in any direction at any given state.
-This means that, for any given model and observation, we will need to perform a maximum of 4 editions to the input model to explain the observation.
-This aspect heavily constrains the discriminating power of the $P(\mathcal{O}|\mathcal{M})$ likelihood, and added to the low observability of the states produces a very benchmark for {\em model recognition as planning}.
 
@@
 
-\begin{figure}[hbt!]
+\begin{figure}
 	\centering
@@
 
+\section{Related work}
+\label{sec:related}
+A related approach is recently followed for {\em model reconciliation}~\cite{ChakrabortiSZK17} where model edition is used to conform the PDDL models of two different agents.
+Also  related to this paper is the work on {\em Goal Recognition Design} but in that work the action model is given in advance and fixed~\cite{KerenGK14}.
+
+Previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
+Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
+
 \section{Conclusions}
@@
 {\em Model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unobserved.
-
-Previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
-Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
-A related approach is recently followed for {\em model reconciliation}~\cite{ChakrabortiSZK17} where model edition is used to conform the PDDL models of two different agents.
-Also  related to this paper is the work on {\em Goal Recognition Design} but in that work the action model is given in advance and fixed~\cite{KerenGK14}.
 

### 2018-11-16 doc/icaps-19/model-recognition.tex
added related work section to icaps19 paper

@@
 \label{sec:related}
-A related approach is recently followed for {\em model reconciliation}~\cite{ChakrabortiSZK17} where model edition is used to conform the PDDL models of two different agents.
-Also  related to this paper is the work on {\em Goal Recognition Design} but in that work the action model is given in advance and fixed~\cite{KerenGK14}.
+An intriguing research line related with {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17} where model edition is used to conform the PDDL models of two different agents wrt an observation of a plan execution.
+In {\em model recognition} we also conform every input model $\mathcal{M}$ in the $M$ set with another model $\mathcal{M}'$ s.t.
+it is consistent with the input observation.
+Note however that in the {\em model recognition} task, the model $\mathcal{M}'$ is not given as input but instead, this model $\mathcal{M}'$ is computed for every $\mathcal{M}\in M$.
+Further, {\em model recognition as planning} shows that two input planning models can be conformed using an off-the-shelf classical planner and wrt an observation of a planning execution that contains no information of the applied actions.
 
-Previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
+Once the planning model of the observed agent is recognized, then the model-based machinery for automated planning becomes fully applicable.
+Of course, this includes the current approaches for {\em goal recognition}~\cite{ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} and {\em counter-planning}~\cite{pozanco2018counterplanning} that require to have a model, in advance and fixed, of the observed agent.
+Last but not least, our approach allows to predict the future actions of an observed agent whose behavior is not ruled by a deliberative planning process that aims to achieve certain goals.
+As an example, our experiments show that {\em model recognition} is suitable for anticipating the actions of agents that are ruled by the execution of a fixed program or a finite state machine, such as a {\em regular automata}.
+
+%Last but not least, previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
 Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.

### 2018-11-16 doc/icaps-19/model-recognition.tex
added related work section to icaps19 paper

@@
 \label{sec:related}
-An intriguing research line related with {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17} where model edition is used to conform the PDDL models of two different agents wrt an observation of a plan execution.
+An intriguing research line related with {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17} where model edition is used to conform the PDDL models of two different agents (or the model of an agent with an {\em annotated model}~\cite{sreedharan2018handling}) wrt an observation of a plan execution.
 In {\em model recognition} we also conform every input model $\mathcal{M}$ in the $M$ set with another model $\mathcal{M}'$ s.t.

### 2018-11-17 doc/icaps-19/model-recognition.tex

--- 2018-11-16 doc/icaps-19/model-recognition.tex
@@
 \label{sec:related}
-An intriguing research line related with {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17} where model edition is used to conform the PDDL models of two different agents wrt an observation of a plan execution.
-In {\em model recognition} we also conform every input model $\mathcal{M}$ in the $M$ set with another model $\mathcal{M}'$ s.t.
-it is consistent with the input observation.
-Note however that in the {\em model recognition} task, the model $\mathcal{M}'$ is not given as input but instead, this model $\mathcal{M}'$ is computed for every $\mathcal{M}\in M$.
-Further, {\em model recognition as planning} shows that two input planning models can be conformed using an off-the-shelf classical planner and wrt an observation of a planning execution that contains no information of the applied actions.
+An interesting research line related with {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}.
+In {\em model reconciliation} model edition is used to conform the PDDL models of two different agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, and with respect to a fully observed {\em optimal plan} computed using one of the models.
+It can be thought that {\em model recognition} conforms every input model $\mathcal{M}$ in the $M$ set with another model $\mathcal{M}'$ s.t.
+it is {\em consistent} with a partial observation of a plan execution.
+Note however that in the {\em model recognition} task, the model $\mathcal{M}'$ is not given as input and instead, this model $\mathcal{M}'$ is computed for every $\mathcal{M}\in M$.
+Further, {\em model recognition as planning} shows that the probability of an input planning model to explain a partial observation of a plan execution (that contains no information of the applied actions and where intermediate states may be incomplete or even missing) can be estimated with an off-the-shelf classical planner.
 
@@
 Last but not least, our approach allows to predict the future actions of an observed agent whose behavior is not ruled by a deliberative planning process that aims to achieve certain goals.
-As an example, our experiments show that {\em model recognition} is suitable for anticipating the actions of agents that are ruled by the execution of a fixed program or a finite state machine, such as a {\em regular automata}.
+As an example, our experiments show that {\em model recognition} is suitable for anticipating the actions of agents that are ruled by the execution of fixed programs or finite state machines, such as a {\em regular automata}.
 

### 2018-11-17 doc/icaps-19/model-recognition.tex
Updated navigation policies results

@@
 
-Figure \ref{fig:navigation_acc} shows the classification {\em accuracy} achieved by {\em model recognition as planning} with respect to a range of different degrees of observability, from 0\% to 100\% with 10\% increments.
+Figure \ref{fig:navigation_acc} shows the classification {\em accuracy} achieved by {\em model recognition as planning} with respect to a range of degrees of observability, from 0\% to 100\% with 10\% increments.
 In this experiment we included the 0\% case which corresponds to observations where only the initial and final states are observed.
-The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation policies, but starting from 10\% we start to correctly classify most of the observations.
-Accuracy stabilizes in the range 0.875 to 1 which means that, at most, only 1 observation of the 8 was not correctly classified.
+The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation policies, but starting from 10\% we start to correctly classify half of the observations.
+Accuracy stabilizes after 40\% observability in the range 0.875 to 1 which means that, at most, only 1 observation of the 8 was not correctly classified.
 

### 2018-11-18 doc/icaps-19/model-recognition.tex
versión casi final paper

--- 2018-11-17 doc/icaps-19/model-recognition.tex
@@
 \begin{abstract}
-Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action models), {\em model recognition} is the task of identifying the model that explains the observation.
+Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action schemata), {\em model recognition} is the task of identifying the model that explains the observation.
 The paper formalizes this task and introduces a novel method that estimates the probability of a \strips\ model to produce an observation of a plan execution.
@@
 \label{sec:introduction}
-{\em Plan recognition} is the task of predicting the future actions of an agent provided observations of its current behavior~\cite{Carberry01}.
-{\em Goal recognition} is a closely related task that aims identifying the goals of the observed agent.
-Goal recognition is considered {\em automated planning} in reverse; while automated planning compute sequences of actions that accounts for a given goals, goal recognition compute goals that account for an observed sequence of actions~\cite{geffner:book:2013}.
 
-Diverse approaches has been proposed for plan/goal recognition such as {\em rule-based systems}, {\em parsing}, {\em graph-covering}, {\em Bayesian nets}, etc~\cite{sukthankar14}.
-{\em Plan recognition as planning} is the model-based approach for plan/goal recognition that leverages the action model of the observed agent to compute its most likely goal~\cite{ramirez2012plan,ramirez2009plan}.
+A growing interest in addressing the task of recognizing the mission (plans and goals) of a planning agent from a set of given observations has recently emerged.
+Approaches based on rule-based systems, parsing, graph-covering or Bayesian nets have been proposed for the plan/goal recognition tasks~\cite{sukthankar14}.
+{\em Plan recognition as planning} is the model-based approach for this task that leverages the action model of the observed agent to compute its most likely goal and predict its future actions~\cite{ramirez2009plan,ramirez2012plan}.
+While action models are learnable from observations of the agent~\cite{yang2007learning,cresswell2013acquiring,aineto2018learning}, real-world applications developing in a particular context usually dispose of approximate behavioral models of the involved agents and there is rarely a need to learn a model from scratch.
+A more relevant task in these scenarios is thereby to predict the behavior of an agent from a set of given models.
 
-This paper formalizes the {\em model recognition} task where the object to recognize is not a goal but the {\em planning model} that shapes the behavior of the observed agent.
-Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but that update these variables with different action models); {\em model recognition} is the task of identifying the model in the set with the highest probability of producing/explaining the given observation.
+This paper formalizes the novel task of {\em model recognition}, where the object to recognize is not a goal or a plan but the model that shapes the behavior of the observed agent.
+Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action schemata), {\em model recognition} is the task of identifying the model in the set with the highest probability of producing the given observation.
+We posit this problem as a \emph{classification task} where classes are given as generative planning models and samples are sequences of observations.
+We accomplish this task with off-the-shelf classical planning algorithms that compute for each model (class) $\mathcal{M}$ in the set a model $\mathcal{M'}$ that explains the observation, and subsequently distances between the two models are measured.
+Two key contributions of our {\em model recognition as planning} approach are: (1) the ability of handling partial observations of plan executions that contain no actions and incomplete -- even empty- intermediate states; and (2) a thorough formalization that induces a probability distribution of a model $\mathcal{M}$ to explain an observed plan execution.
 
@@
 According to this model, the robot can increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
-at {\em even rows} if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at {\em odd rows} if {\tt\small $q0$} holds initially).
-Apart from the model of Figure~\ref{fig:model-example}, different models can be defined within the same state variables (e.g.
-altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated) and these models can shape different kinds of robot navigation.
-Given a partially observed plan execution (like the one illustrated at Figure~\ref{fig:grid-example} where state variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved) {\em model recognition} aims to identify the action model that produced/explains that observation.
+at even rows if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at odd rows if {\tt\small $q0$} holds initially).
+Besides this model, other models that shape different kinds of robot navigation can be defined within the same state variables (e.g.
+altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated).
+Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where state variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, {\em model recognition} aims to identify the robot action schemata that explains this observation.
 
-{\em Model recognition} is of interest because once the planning model is recognized, then the model-based machinery for automated planning becomes applicable~\cite{ghallab2004automated}.
-In addition, it enables identifying different kinds of automata by observing their execution.
-It is well-known that diverse automata representations, like {\em finite state controllers}, {\em B\"{u}chi automata}, {\em push-down automata}, {\em {\sc GOLOG} programs} or {\em reactive policies}, can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,patrizi2011computing,IvankovicH15,segovia2017generating}.
+We provide a formalization of {\em model recognition} that is applicable to any planning model but we restrict our attention to \strips\texttt{-}compilable models.
+It is well-known that diverse automata representations like finite state controllers, {\sc GOLOG} programs or reactive policies can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,segovia2017generating}.
+As for the evaluation, the effectiveness of our approach is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a proof-of-concept classification task; (2) a simulation of a non-deterministic \emph{blocksworld} under different degrees of observability; and (3) a navigation task where observations contain an unbounded number of unobservable states.
 
-The paper also introduces {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
+%: a classical string classification task; identification of the model that encodes a failure present in an observation; and recognition of a robot navigation policy.
+
+%The paper also introduces {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
 Our method is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.
-We evaluate the effectiveness of {\em model recognition as planning} with sets of \strips\ models that represent different {\em automata}.
-All the {\em automata} in a set are defined within the same state variables but with different {\em transition functions}.
-We show that {\em model recognition as planning} identifies the executed {\em automata} despite the actual applied transitions, plus some state variables (e.g.
-the internal automata state), are unobserved.
 
@@
   \end{tiny}
- \caption{\small Example of a \strips\ action model (given in PDDL) for robot navigation in a $N\times N$ grid.}
+ \caption{\small Example of a \strips\ action schemata (given in PDDL) for robot navigation in a $N\times N$ grid.}
 \label{fig:model-example}
@@
 Specifically, the number of observed actions, $l$, can range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
-\item The {\bf observed states} $\tup{s_0^0, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
+\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
 A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
 Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
-Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^0 \subseteq s_i$.
-In practive, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
-Exceptionally, $s_m^o$ cannot be fully unobservable for the purpose of our task (we will elaborate on this issue later on).
+Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
+    %Exceptionally, $s_m^o$ cannot be fully unobservable for the purpose of our task (we will elaborate on this issue later on).
 \end{itemize}
@@
 
-In our model recognition task, we will always assume an \emph{empty sequence of observed actions} and so we will only work with the observed states.
 Figure~\ref{fig:grid-example} illustrates a {\em partial observation} of a five-state trajectory \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 4)(ycoord 2)>, <(xcoord 3)(ycoord 2)>}\}.
@@
 Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
-\item $\mathcal{O}(\tau)$ is an observation of the execution of an unknown solution plan $\pi$ for the planning problem $P$.
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
 \end{itemize}
@@
 
-Since ${\mathcal I}_{\Psi,\xi}$ is a bounded set, the maximum number of edits that can be introduced to an action model is bounded as well.
+Since ${\mathcal I}_{\Psi,\xi}$ is a bounded set, the maximum number of edits that can be introduced to an action schema is bounded as well.
 The \textbf{maximum edit distance} of a \strips\ model $\mathcal{M}$ built with predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$ (note that if we consider the set of syntactic constraints then $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2\times|{\mathcal I}_{\Psi,\xi}|$).
@@
 
-An observation of the execution of a plan generated with an action model $\mathcal{M}$ further constraints the space of possible action schemata of $\mathcal{M}$.
+An observation of the execution of a plan generated with $\mathcal{M}$ further constraints the space of possible action schemata of $\mathcal{M}$.
 The \emph{semantic knowledge} included in the observations introduce a third type of constraints, that we will call {\em observation constraints}, and that can be added to the set $\mathcal{C}$.
@@
 
-In essence, the task $\tup{\mathcal{M},\mathcal{O}}$ of the AJO approach -- learning $\mathcal{M}$ that satisfies $\mathcal{O}$-- can be interpreted as editing the empty action schemata of $\mathcal{M}$ introducing preconditions and effects until $\mathcal{O}$ is validated with the resulting model.
+Essentially, the task $\tup{\mathcal{M},\mathcal{O}}$ of the AJO approach -- learning $\mathcal{M}$ that satisfies $\mathcal{O}$-- is interpreted as editing the empty action schemata of $\mathcal{M}$ introducing preconditions and effects until $\mathcal{O}$ is validated with the resulting model.
 We leverage the same idea to compute $\delta^o(\mathcal{M},\mathcal{O})$ with the exception that now our $\mathcal{M}$ is not empty but $\mathcal{M}=\{\xi_1, \ldots, \xi_m\}$, where $\xi_i=\tup{name(\xi_i),pars(\xi_i),pre(\xi_i),add(\xi_i),del(\xi_i)}$, $1 \leq i \leq m$.
@@
 We now show an example of a solution plan to a planning problem $P'$ that results from compiling a specific task $\tup{\mathcal{M},\mathcal{O}}$.
-Consider that $\mathcal{M}$ is the model of Figure~\ref{fig:model-example} except that the schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped wrt Figure~\ref{fig:model-example}.
-And consider an observation that only contains the initial and final state of Figure~\ref{fig:grid-example}; i.e., $\mathcal{O(\tau)}=\tup{s_0^o,s_m^o}=\tup{(1,1),(3,2)}$.
+Consider that $\mathcal{M}$ is the model of Figure~\ref{fig:model-example} except that the schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped with respect to Figure~\ref{fig:model-example}.
+And consider an observation that only contains the initial and final state of Figure~\ref{fig:grid-example}; i.e., $\mathcal{O(\tau)}=\tup{s_0^o,s_m^o}=\tup{\{{\texttt{\small{(xcoord 0)(ycoord 0)}}}\}, \{{\texttt{\small{(xcoord 2)(ycoord 1)}}}\}}$.
 The plan found by a planner to $P'$ is shown in figure~\ref{fig:plan-pdistance}.
-The first seven steps are the edit actions to \emph{fix} the schema {\tt\small inc-x}; step 07 is a validate action that sets the robot in the initial state $s_0^o=(1,1)$; step 08 applies the action {\tt\small (inc-x 1 2)} and moves the robot one cell to the right to $(2,1)$; steps 09-11 also move the robot one cell to the right; step 11 applies an action that increases coordinate $y$ to an even number, thus moving robot to row $2$; step 12 moves the robot one cell to the left and, finally, action {\tt\small (validate\_1)} checks the robot position is consistent with the final state $s_m^o=(3,2)$.
+The first seven steps are the edit actions to \emph{fix} the schema {\tt\small inc-x}; step 07 is a validate action that sets the robot in the initial state $s_0^o=\tup{\{{\texttt{\small{(xcoord 0)(ycoord 0)}}}\}}$; step 08 applies the action {\tt\small (inc-x 0 1)} and moves the robot one cell to the right to position $(1,0)$; steps 09-11 also move the robot one cell to the right; step 11 applies an action that increases coordinate $y$ from a position in an even row number, thus moving robot to row $1$; step 12 moves the robot one cell to the left and, finally, action {\tt\small (validate\_1)} checks the robot position is consistent with the final state $s_m^o=\tup{\{{\texttt{\small{(xcoord 2)(ycoord 1)}}}\}}$.
 
@@
 00 : (insert\_pre\_xcoord\_v1\_inc-x)   & 07 : (validate\_0)\\
-01 : (insert\_pre\_next\_v1\_v2\_inc-x) & 08 : (apply\_inc-x 1 2)\\
-02 : (insert\_pre\_q0\_inc-x)           & 09 : (apply\_inc-x 2 3)\\
-03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (apply\_inc-x 3 4) \\
-04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (apply\_inc-y-even 1 2)\\
-05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (apply\_dec-x 4 3)\\
+01 : (insert\_pre\_next\_v1\_v2\_inc-x) & 08 : (apply\_inc-x 0 1)\\
+02 : (insert\_pre\_q0\_inc-x)           & 09 : (apply\_inc-x 1 2)\\
+03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (apply\_inc-x 2 3) \\
+04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (apply\_inc-y-even 0 1)\\
+05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (apply\_dec-x 3 2)\\
 06 : (insert\_add\_xcoord\_v2\_inc-x)   & 13 : (validate\_1)
@@
 
-This section presents three different applications of {\em model recognition} and evaluates the empirical performance of {\em model recognition as planning} when addressing the three of them.
+In this section, we evaluate the empirical performance of our approach in three different applications of {\em model recognition}.
+We will assume in the three experiments that $\mathcal{O}(\tau)$ contains an \emph{empty sequence of observed actions} and so we will only work with the available observed states.
 
-In all the experiments we define a set $M$ of different planning models, that share the same state variables but that update these variables according to different action schemata.
-Then for each model $\mathcal{M}\in M$, we generate {\em partial observations} $\mathcal{O}(\tau)$ of {\em plan trajectories} that were computed with that model.
-Finally, we apply our {\em model recognition as planning} method to identify which model, among the models in $M$, was actually used for generating a given $\mathcal{O}(\tau)$ observation.
-For all the experiments, we assume equal prior probabilities, meaning $P(\mathcal{M}) = \frac{1}{|M|}$ for every $\mathcal{M} \in M$.
+For each experiment, we define a set $M$ of different planning models that share the same state variables but update the variables using different action schemata.
+For every $\mathcal{M}\in M$, we generated the same number of partial observations of plan trajectories ($\mathcal{O}(\tau)$) with such a model.
+Finally, we applied our {\em model recognition as planning} approach to identify the model, among the models in $M$, which was actually used for generating a given $\mathcal{O}(\tau)$ observation.
+We assume equiprobable prior probabilities in all the experiments; i.e., $P(\mathcal{M}) = \frac{1}{|M|}$ for every $\mathcal{M} \in M$.
 
-To provide a better explanation of the experiments below, we introduce a particular class of $\mathcal{O}(\tau)$ observations.
-This new class allows us to distinguish between {\em observable} state variables, whose value may come from sensors, and {\em hidden} or {\em latent} state variables that cannot be observed.
+In order to provide a better explanation of the experiments below, we introduce a particular class of $\mathcal{O}(\tau)$ observations.
+This new class allows us to distinguish between {\em observable} state variables, whose value may be read from sensors, and {\em hidden} or {\em latent} state variables that cannot be observed.
+
 \begin{definition}[$\Phi$-observation]
@@
 \end{definition}
-Therefore the value of variables in $\Phi\subseteq F$ can be {\em observed} while the value of the variables in $F\setminus\Phi$ is {\em hidden}.
+
+Hence the value of the variables in $\Phi\subseteq F$ is observable while the value of the variables in $F\setminus\Phi$ is unobservable.
 
 \subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar}~\cite{rintanen2014madagascar} because of its ability to deal with classical planning problems with dead-ends~\cite{lopez2015deterministic}.
-We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned maximum distance.
-Compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em } so any experimental data reported in the paper can be reproduced.
+We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned the maximum distance obtained.
+The compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em https://github.com/anonsub/model-recognition} so any experimental data reported in the paper can be reproduced.
 
-\subsection{Recognition of {\em regular automata}.} The first experiment, which doubles as a proof of concept, leverages {\em model recognition as planning} to perform a typical {\em string classification} problem.
-In this experiment, the system receives (1) the string to classify (the $\mathcal{O}(\tau)$ observation) and (2), a set $M$ of different planning models, where each $\mathcal{M}\in M$ codes a different class (i.e.
-a different regular automata that accepts any string that belong to the class).
+\subsection{Recognition of {\em regular automata}.} The first experiment, which doubles as a proof of concept, exploits {\em model recognition as planning} for a classical {\em string classification} problem.
+In this experiment, the system receives (1) the string to classify (the $\mathcal{O}(\tau)$ observation) and (2) a set $M$ of different planning models, where each $\mathcal{M}\in M$ represents a different class; i.e.
+a different regular automata that accepts strings that belong to the class.
 
-Figure~\ref{fig:regautomatae} illustrates a three-symbol four-state {\em regular automata} for recognizing the $(abc)^+$ language.
-The {\em input alphabet} is $\Sigma=\{a,b,c\}$, and the machine states are $Q=\{q_0,q_1,q_2,\underline{q_3}\}$ (where \underline{$q_3$} is the only acceptor state).
-For instance, the execution of the planning model coding the {\em regular automata} defined in Figure~\ref{fig:regautomatae}, and with the input string $abcabc$, produces the following six-action plan {\small $(\langle a,q_0\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow \underline{q_3})$, $(\langle a,\underline{q_3}\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow \underline{q_3})$}.
+Figure~\ref{fig:regautomatae} illustrates a 4-symbol and 5-state {\em regular automata} for recognizing the $(abcd)^+$ language.
+The {\em input alphabet} is $\Sigma=\{a,b,c,d\}$, and the machine states are $Q=\{q_0,q_1,q_2,q_3,\underline{q_4}\}$, where \underline{$q_4$} is the only acceptor state.
+For instance, executing the planning model that encodes the {\em regular automata} of Figure~\ref{fig:regautomatae} with the input string $abcdabcd$ produces the following 8-action plan {\small $(\langle a,q_0\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow q_3)$, $(\langle d,q_3\rangle\rightarrow \underline{q_4})$, $(\langle a,\underline{q_4}\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow \underline{q_3})$, $(\langle d,q_3\rangle\rightarrow \underline{q_4})$}.
+
+%\begin{figure}
+%  \begin{scriptsize}
+%  \begin{center}
+%	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
+%	  \node[state] (A)              {$q_0$};
+%	  \node[state] (B) [right of=A] {$q_1$};
+%	  \node[state] (C) [right of=B] {$q_2$};
+%	  \node[state] (D) [right of=C] {\underline{$q_3$}};
+%	  \path
+%(A) edge  node {a} (B)
+%(B) edge  node {b} (C)
+%(C) edge  node {c} (D)
+%(D) edge [bend left]  node {a} (B)
+%;
+%	\end{tikzpicture}
+%  \end{center}
+%  \end{scriptsize}
+% \caption{\small A 3-symbol 4-state {\em regular automata} for recognizing the $(abc)^+$ language (\underline{$q_3$} is the acceptor state).}
+%\label{fig:regautomatae}
+%\end{figure}
 
@@
   \begin{center}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.8cm,semithick]
 	  \node[state] (A)              {$q_0$};
@@
 	  \node[state] (C) [right of=B] {$q_2$};
-	  \node[state] (D) [right of=C] {\underline{$q_3$}};
+	  \node[state] (D) [right of=C] {$q_3$};
+	  \node[state] (E) [right of=D] {\underline{$q_4$}};
 	  \path
@@
 (C) edge  node {c} (D)
-(D) edge [bend left]  node {a} (B)
+(D) edge  node {d} (E)
+(E) edge [bend left]  node {a} (B)
 ;
@@
   \end{scriptsize}
- \caption{\small Three-symbol four-state {\em regular automata} for recognizing the $(abc)^+$ language (\underline{$q_3$} is the acceptor state).}
+ \caption{\small A 4-symbol and 5-state {\em regular automata} for recognizing the $(abcd)^+$ language (\underline{$q_4$} is the acceptor state).}
 \label{fig:regautomatae}
@@
 
-In this experiment, $M$ comprises 5 planning models representing different regular automata defined with 5 states and a 4-symbol alphabet.
-The five regular languages defined by these five automata are the following:
+In this experiment, $M$ comprises five planning models, each representing a different 5-state and 4-symbol regular automata. The five regular languages defined by these automata are the following:
+
 \begin{itemize}
@@
 
-For each regular language, we generated $20$ random strings, $\mathcal{O}(\tau)$ observations, of length 20 to 30 for a total of 100 strings.
-More precisely observations $\mathcal{O}(\tau)$ are $\Phi$-observations where: the applied automata transitions and the fluents representing the internal automata state are unobserved.
-Table \ref{tab:conf_matrix} shows the {\em confusion matrix} resulting from classifying these 100 strings with {\em model recognition as planning}.
-In this matrix, rows represent the actual class and columns the class predicted by our {\em model recognition as planning} approach.
-Even though we are using a suboptimal classical planner, we can see that the class for all the 100 input strings was correctly recognized.
-These results proof the suitability of {\em model recognition as planning} for classification tasks where classes are given as generative planning models and examples are sequence of observations.
+For each regular language, we generated $20$ random strings, $\mathcal{O}(\tau)$ observations, which length range from 20 to 30 symbols, thus generating a total of 100 strings.
+More precisely, $\mathcal{O}(\tau)$ are observations in which the applied transitions (actions) are unobservable and the fluents of the automata internal state are also unobservable.
+Consequently, $\mathcal{O}(\tau)$ are $\Phi$-observations where the problem states are only represented with the fluents that specify the string to classify.
+
+%More precisely observations $\mathcal{O}(\tau)$ are $\Phi$-observations where: the applied automata transitions and the fluents representing the internal automata state are unobserved.
+
+Table \ref{tab:conf_matrix} shows the {\em confusion matrix} resulting from classifying the 100 strings with our method.
+In this matrix, rows represent the actual class and columns represent the class predicted by our {\em model recognition as planning} approach.
+Despite using a suboptimal classical planner, we can observe that the class for all the 100 input strings was correctly recognized, which proves the feasibility of our method for classification tasks.
+The outstanding results reveal that our approach is very suitable for the classification of generative planning models that are more restrictive than \strips\ models, as it is the case of regular automata.
+
+%These results prove the suitability of {\em model recognition as planning} for classification tasks where classes are given as generative planning models and examples are sequence of observations.
 
@@
 \subsection{Recognizing failures in a non-deterministic {\em blocksworld}}
-With this second experiment we aim to validate the assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood.
-These assumptions being that the agent acts rational, so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation, and that there exists a single such trajectory.
+With this second experiment we aim to validate the two assumptions underlying our approximation of $P(\mathcal{O}|\mathcal{M})$: (1) that agents act rationally and so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation $\mathcal{O}(\tau)$ and (2) that there exists a single such trajectory.
+Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions.
+In order to show the assumptions validation:
 
-In this particular experiment we assume that the length of the trajectory is known (to force the explanation to be optimal) and hence, measure the {\em  accuracy} of our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood across different degrees of observability.
-More precisely, we assume that at least one literal is observed at every state so the length of the plan trajectory is fixed by the observation.
+\begin{itemize}
+\item we will measure the {\em  accuracy} of our approximation of $P(\mathcal{O}|\mathcal{M})$ across different degrees of observability; the accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a fluent in an intermediate state of $\mathcal{O}(\tau)$
+\item we will assume that the length of the trajectory is known so as to force the optimal explanation of $\mathcal{O}(\tau)$; more precisely, at least one fluent of every state of $\mathcal{O}(\tau)$ is observed and the length of the plan trajectory is thus fixed by the observation
+\end{itemize}
 
-The input observations $\mathcal{O}(\tau)$ come from a non-deterministic {\em blocksworld} in which some of the executed actions failed. We considered failures of three kinds:
+The input observations $\mathcal{O}(\tau)$ come from a non-deterministic {\em blocksworld} in which some of the executed actions may have failed. We considered failures of three kinds:
+
 \begin{itemize}
-	\item $\mathcal{F}$1: the execution of a {\tt\small stack} action fails causing no effect.
-	\item $\mathcal{F}$2: {\tt\small unstack} fails dropping the block to unstack on the table.        
-	\item $\mathcal{F}$3: {\tt\small unstack} fails doing nothing.
+	\item the execution of a {\tt\small stack} action fails and causes no effect
+    \item the execution of {\tt\small unstack} fails and causes no effect
+	\item {\tt\small unstack} fails and drops the block on the table
 \end{itemize}
-The input models in $M$ are then three different {\em extended} 4-operator {\em blocksworlds} such that each of these models includes an additional action schema that encodes one of the three failures above.
-This means that failures are identified by finding the model that explains the best the input observation.
+
+$M$ comprises three different {\em extended} 4-schema {\em blocksworld} models.
+Each model in $M$ includes an additional action schema that encodes one of the three above failures.
+This means that failures are identified by finding the model that better explains the input observation.
 
 Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations.
-{\em Accuracy} is defined as the number of correct predictions over all predictions made while the {\em degree of observability} indicates the probability of observing a literal in an intermediate state.
-The figure showcases a positive trend proving that the more complete the observations are the more accurate our estimation becomes.
-The overall high accuracy validates our assumption that the trajectory that better explains the observation is the shortest one.
+The increasing trend of the accuracy as observability increases, and so the number of possible explanation lowers, validates our two assumptions.
+
+%The overall high accuracy validates our assumption that the trajectory that better explains the observation is the shortest one.
 %We limited the failures to be of one kind per observation however, it would be possible to recognize different combinations of them.
@@
 \subsection{Recognition of navigation policies}
-Again, this experiment evaluates the accuracy of {\em model recognition as planning} to identify the model that generated each observation under different {\em degrees of observability}.
-The novel aspect of this experiment is that now, we do not guarantee that observed states contain at least one literal, meaning that some intermediate states may be missing.
-Therefore, in this new scenario more than a single action may be required to reach one observed state from another.
+%This experiment is also aimed at validating the accuracy of our approach to identify the model that generates an observation under different degrees of observability.
+The novel aspect of this experiment is that, unlike the previous experiment, we do not guarantee that the observed states contain at least one fluent.
+This means that some intermediate states of $\mathcal{O}(\tau)$ may be missing (all fluents are unobservable) and, consequently, more than a single action may be required to reach an observed state from another one.
 
-In this experiment the planning models given in $M$ represent different navigation models for $N\times N$ grids.
-In particular, we took a {\em classical navigation model} with actions {\tt\small up, down, right, left} (to move one cell in each of these four directions) and extend it to define 8 different navigation policies with respect to these additional state variables {\tt\small \{(q0), (q1), (min ?v), (max ?v)\}}.
-An example is the navigation policy, shown in Figure~\ref{fig:model-example}.
-This policy allows to move right when $q_0$ holds while it allows to move left when $q_1$ holds, producing a zig-zag pattern when visiting all the cells of $N\times N$ grids.
+In this experiment, the planning models of $M$ represent different navigation models for $N\times N$ grids.
+In particular, we adopt a {\em classical navigation model} with actions {\tt\small up, down, right, left} (to move one cell in each of these four directions) and extend it to define 8 different navigation policies with respect to these additional state variables {\tt\small \{(q0), (q1), (min ?v), (max ?v)\}}.
+An example of a navigation policy is shown in Figure~\ref{fig:model-example}.
+This policy allows to move right when $q_0$ holds while it allows to move left when $q_1$ holds, producing a zigzag pattern when visiting all the cells of $N\times N$ grids.
 
-Note that all navigation policies are at a maximum distance of just 4 editions from the base {\em classical navigation model} that can move in any direction at any given state.
-This aspect constrains the discriminating power of the $P(\mathcal{O}|\mathcal{M})$ likelihood, and added to the low observability of the states, produces a touchstone for {\em model recognition as planning}.
+Another interesting aspect of this experiment is that all navigation policies are at a maximum distance of four editions from the base {\em classical navigation model} that can move the robot in any direction at any given state.
+Thus, for any given model and observation, a maximum of four editions are needed for the input model to explain the observation.
+This aspect heavily constrains the discriminating power of $P(\mathcal{O}|\mathcal{M})$, which along with the low degree of observability of the states, gives rise to a meaningful benchmark for {\em model recognition as planning}.
 
-On the other hand we generated observations from 8 different trajectories, one for each of the 8 previous planning models.
-The trajectories depict paths followed by each navigation policy to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
-As in the previous experiments, observations contain no actions and here, for each observed state, only the values of the fluents encoding the {\em x and y coordinates} of the agent are known (i.e.
+On the other hand, we generated observations from 8 different trajectories, one for each of the 8 previous planning models.
+The trajectories depict paths followed by a navigation policy to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
+As in the previous experiments, observations contain no actions and here, for each observed state, only the values of the fluents encoding the {\em x} and {\em y coordinates} of the agent are known (i.e.
 $\Phi =$\{{\small\tt xcoord, ycoord}\}).
 
-Figure \ref{fig:navigation_acc} shows the classification {\em accuracy} achieved by {\em model recognition as planning} with respect to a range of different degrees of observability, from 0\% to 100\% with 10\% increments.
+Figure \ref{fig:navigation_acc} shows the classification {\em accuracy} achieved by our approach with respect to a range of degrees of observability, from 0\% to 100\% with 10\% increments.
 In this experiment we included the 0\% case which corresponds to observations where only the initial and final states are observed.
-The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation policies, but starting from 10\% we start to correctly classify most of the observations.
-Accuracy stabilizes in the range 0.875 to 1 which means that, at most, only 1 observation of the 8 was not correctly classified.
+The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation policies, but from 10\% of observability onwards, we start to correctly classify half of the observations.
+Accuracy stabilizes after 40\% observability in the range 0.875 to 1 which means that at most only one out of the 8 observations was not correctly classified.
 
@@
 An interesting research line related with {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}.
-In {\em model reconciliation} model edition is used to conform the PDDL models of two different agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, and with respect to a fully observed {\em optimal plan} computed using one of the models.
-It can be thought that {\em model recognition} conforms every input model $\mathcal{M}$ in the $M$ set with another model $\mathcal{M}'$ s.t.
-it is {\em consistent} with a partial observation of a plan execution.
-Note however that in the {\em model recognition} task, the model $\mathcal{M}'$ is not given as input and instead, this model $\mathcal{M}'$ is computed for every $\mathcal{M}\in M$.
+In {\em model reconciliation} model edition is used to conform the PDDL models of two different agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed {\em optimal plan} computed using one of the models.
+{\em Model recognition} conforms every input model $\mathcal{M}$ in the $M$ set with another model $\mathcal{M}'$ that is consistent with a partial observation of a plan execution.
+Note, however, that in {\em model recognition} $\mathcal{M}'$ is not given as input but instead is computed for every $\mathcal{M}\in M$.
 Further, {\em model recognition as planning} shows that the probability of an input planning model to explain a partial observation of a plan execution (that contains no information of the applied actions and where intermediate states may be incomplete or even missing) can be estimated with an off-the-shelf classical planner.
@@
 Once the planning model of the observed agent is recognized, then the model-based machinery for automated planning becomes fully applicable.
-Of course, this includes the current approaches for {\em goal recognition}~\cite{ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} and {\em counter-planning}~\cite{pozanco2018counterplanning} that require to have a model, in advance and fixed, of the observed agent.
-Last but not least, our approach allows to predict the future actions of an observed agent whose behavior is not ruled by a deliberative planning process that aims to achieve certain goals.
-As an example, our experiments show that {\em model recognition} is suitable for anticipating the actions of agents that are ruled by the execution of fixed programs or finite state machines, such as a {\em regular automata}.
+This includes current approaches for {\em goal recognition}~\cite{ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} and {\em counter-planning}~\cite{PozancoEFB18} that require to have in advance a fixed model of the observed agent.
+Last but not least, our approach allows predicting the future actions of an agent whose behavior is not governed by a deliberative planning process aimed to achieve some goals.
+Our experiments show that {\em model recognition} is suitable for anticipating the actions of agents that are ruled by the execution of fixed programs or finite state machines, such as a {\em regular automata}.
 

### 2018-11-18 doc/icaps-19/model-recognition.tex
minor changes by Diego

@@
 
-For each regular language, we generated $20$ random strings, $\mathcal{O}(\tau)$ observations, which length range from 20 to 30 symbols, thus generating a total of 100 strings.
+For each regular language, we generated $20$ random strings, $\mathcal{O}(\tau)$ observations, which lengths range from 20 to 30 symbols, thus generating a total of 100 strings.
 More precisely, $\mathcal{O}(\tau)$ are observations in which the applied transitions (actions) are unobservable and the fluents of the automata internal state are also unobservable.
@@
 \subsection{Recognizing failures in a non-deterministic {\em blocksworld}}
-With this second experiment we aim to validate the two assumptions underlying our approximation of $P(\mathcal{O}|\mathcal{M})$: (1) that agents act rationally and so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation $\mathcal{O}(\tau)$ and (2) that there exists a single such trajectory.
-Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions.
+With this second experiment we aim to validate the two assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood: (1) that agents act rationally and so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation $\mathcal{O}(\tau)$ and (2) that there exists a single such trajectory.
+Our assumptions are more correct with higher observability and become tautological when the trajectory is fully observed ($\mathcal{O(\tau)} = \tau$).
+In order to show the assumptions validation:
+
+% Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions.
 In order to show the assumptions validation:
@@
 \item we will measure the {\em  accuracy} of our approximation of $P(\mathcal{O}|\mathcal{M})$ across different degrees of observability; the accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a fluent in an intermediate state of $\mathcal{O}(\tau)$
-\item we will assume that the length of the trajectory is known so as to force the optimal explanation of $\mathcal{O}(\tau)$; more precisely, at least one fluent of every state of $\mathcal{O}(\tau)$ is observed and the length of the plan trajectory is thus fixed by the observation
+\item we will assume that the length of the trajectory is known so as to keep the possible explanations bounded and force the optimal explanation of $\mathcal{O}(\tau)$; more precisely, at least one fluent of every state of $\mathcal{O}(\tau)$ is observed and the length of the plan trajectory is thus fixed by the observation
 \end{itemize}
@@
 
-$M$ comprises three different {\em extended} 4-schema {\em blocksworld} models.
-Each model in $M$ includes an additional action schema that encodes one of the three above failures.
+$M$ comprises three different 4-schema {\em blocksworld} models, each one extended with an additional action schema that encodes one of the three above failures.
 This means that failures are identified by finding the model that better explains the input observation.
@@
 Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations.
-The increasing trend of the accuracy as observability increases, and so the number of possible explanation lowers, validates our two assumptions.
+The positive trend seen in the figure proves that the more complete the observations are the more accurate our estimation becomes.
 
 %The overall high accuracy validates our assumption that the trajectory that better explains the observation is the shortest one.
+
+%The increasing trend of the accuracy as observability increases and so the number of possible explanation lowers, validates our two assumptions.
+
 %We limited the failures to be of one kind per observation however, it would be possible to recognize different combinations of them.

### 2018-11-19 doc/icaps-19/model-recognition.tex
blocks results changes

@@
 \subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar}~\cite{rintanen2014madagascar} because of its ability to deal with classical planning problems with dead-ends~\cite{lopez2015deterministic}.
+Other planners, such as FastDownward were also tested but provided worse experimental results.
 We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned the maximum distance obtained.
@@
 Our assumptions are more correct with higher observability and become tautological when the trajectory is fully observed ($\mathcal{O(\tau)} = \tau$).
-In order to show the assumptions validation:
-
-% Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions.
-In order to show the assumptions validation:
-
-\begin{itemize}
-\item we will measure the {\em  accuracy} of our approximation of $P(\mathcal{O}|\mathcal{M})$ across different degrees of observability; the accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a fluent in an intermediate state of $\mathcal{O}(\tau)$
-\item we will assume that the length of the trajectory is known so as to keep the possible explanations bounded and force the optimal explanation of $\mathcal{O}(\tau)$; more precisely, at least one fluent of every state of $\mathcal{O}(\tau)$ is observed and the length of the plan trajectory is thus fixed by the observation
-\end{itemize}
 
@@
 This means that failures are identified by finding the model that better explains the input observation.
+
+The validation of the assumptions is done by measuring the {\em  accuracy} of our approach across different degrees of observability.
+The accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a fluent in an intermediate state of $\mathcal{O}(\tau)$.
+In this experiment, we also assume that the length of the trajectory is known.
+More precisely, at least one fluent of every state of $\mathcal{O}(\tau)$ is observed and the length of the plan trajectory is thus fixed by the observation.
+This allows us to produce a controlled experiment in which the number of possible trajectories able to explain an observation is bounded.
+
+% Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions.
+In order to show the assumptions validation:
 
@@
 The novel aspect of this experiment is that, unlike the previous experiment, we do not guarantee that the observed states contain at least one fluent.
-This means that some intermediate states of $\mathcal{O}(\tau)$ may be missing (all fluents are unobservable) and, consequently, more than a single action may be required to reach an observed state from another one.
+This means that some intermediate states of $\mathcal{O}(\tau)$ may be missing (no fluents are observed) and, consequently, more than a single action may be required to reach an observed state from another one.
 

### 2018-11-20 doc/icaps-19/model-recognition.tex
final version

@@
 % Commented for blind submission
-\author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
-{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
-{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
-{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
-\And Miquel Ram\'irez\\
-{\scriptsize School of Computing and Information Systems}\\
-{\scriptsize The University of Melbourne}\\
-{\scriptsize Melbourne, Victoria. Australia}\\
-{\scriptsize miquel.ramirez@nimelb.edu.au}}
+%\author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
+%{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
+%{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
+%{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
+%{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
+%\And Miquel Ram\'irez\\
+%{\scriptsize School of Computing and Information Systems}\\
+%{\scriptsize The University of Melbourne}\\
+%{\scriptsize Melbourne, Victoria. Australia}\\
+%{\scriptsize miquel.ramirez@nimelb.edu.au}}
 
@@
 {\em Plan recognition as planning} is the model-based approach for this task that leverages the action model of the observed agent to compute its most likely goal and predict its future actions~\cite{ramirez2009plan,ramirez2012plan}.
-While action models are learnable from observations of the agent~\cite{yang2007learning,cresswell2013acquiring,aineto2018learning}, real-world applications developing in a particular context usually dispose of approximate behavioral models of the involved agents and there is rarely a need to learn a model from scratch.
-A more relevant task in these scenarios is thereby to predict the behavior of an agent from a set of given models.
+While action models are learnable from observations of the agent~\cite{yang2007learning,cresswell2013acquiring,aineto2018learning}, in real-world applications developing in a particular context, approximate behavioral models of the involved agents (manually-introduced or automatically-learned) are already available, so there is rarely a need to learn a model from scratch.
+A more relevant and applicable task in these scenarios is thereby to predict the behavior of an agent from a set of given models.
 
@@
 
-Figure~\ref{fig:grid-example} illustrates a {\em partial observation} of a five-state trajectory \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 4)(ycoord 2)>, <(xcoord 3)(ycoord 2)>}\}.
-Note that this observation only contains fluents of the type {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is not observable in any of the five states.
+Figure~\ref{fig:grid-example} illustrates a partial observation of a six-state trajectory \{{\tt\scriptsize<(xcoord 0)(ycoord 0)>, <(xcoord 1)(ycoord 0)>, <(xcoord 2)(ycoord 0)>, <(xcoord 3)(ycoord 0)>, <(xcoord 3)(ycoord 1)>, <(xcoord 2)(ycoord 1)>}\}.
+This observation only contains fluents of the predicates {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, corresponding to predicates {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unobservable in the six states.
 
@@
 The adaptation of this compilation scheme for solving $\delta^o(\mathcal{M},\mathcal{O})$ results in a planning problem $P'=\tup{F',A',I',G'}$ whose objective is to determine the preconditions and effects that need to be added or deleted to the action schemata of $\mathcal{M}$ so as to satisfy $\mathcal{O}$.
-The accomplishment of this task requires therefore a propositional encoding of the components of the action schemata.
-Specifically:
+The accomplishment of this task requires therefore a propositional encoding of the components of the action schemata:
 
@@
 
-Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations.
-The positive trend seen in the figure proves that the more complete the observations are the more accurate our estimation becomes.
+Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations across different degrees of observability.
+The positive trend seen in the figure proves that the more complete the observations are, the more accurate our estimation becomes.
 
@@
 
-In this experiment, the planning models of $M$ represent different navigation models for $N\times N$ grids.
+The planning models of $M$ represent different navigation models for $N\times N$ grids.
 In particular, we adopt a {\em classical navigation model} with actions {\tt\small up, down, right, left} (to move one cell in each of these four directions) and extend it to define 8 different navigation policies with respect to these additional state variables {\tt\small \{(q0), (q1), (min ?v), (max ?v)\}}.
@@
 The trajectories depict paths followed by a navigation policy to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
-As in the previous experiments, observations contain no actions and here, for each observed state, only the values of the fluents encoding the {\em x} and {\em y coordinates} of the agent are known (i.e.
-$\Phi =$\{{\small\tt xcoord, ycoord}\}).
+As in the previous experiments, observations contain no actions and here, for each observed state, only the values of the fluents encoding the {\em x} and {\em y coordinates} of the agent are known; i.e.
+$\Phi$ comprises all the fluents instantiated with predicates {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}.
 
@@
 
-\section{Related work}
+\section{Conclusions and Related work}
 \label{sec:related}
-An interesting research line related with {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}.
-In {\em model reconciliation} model edition is used to conform the PDDL models of two different agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed {\em optimal plan} computed using one of the models.
-{\em Model recognition} conforms every input model $\mathcal{M}$ in the $M$ set with another model $\mathcal{M}'$ that is consistent with a partial observation of a plan execution.
-Note, however, that in {\em model recognition} $\mathcal{M}'$ is not given as input but instead is computed for every $\mathcal{M}\in M$.
-Further, {\em model recognition as planning} shows that the probability of an input planning model to explain a partial observation of a plan execution (that contains no information of the applied actions and where intermediate states may be incomplete or even missing) can be estimated with an off-the-shelf classical planner.
 
-Once the planning model of the observed agent is recognized, then the model-based machinery for automated planning becomes fully applicable.
-This includes current approaches for {\em goal recognition}~\cite{ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} and {\em counter-planning}~\cite{PozancoEFB18} that require to have in advance a fixed model of the observed agent.
-Last but not least, our approach allows predicting the future actions of an agent whose behavior is not governed by a deliberative planning process aimed to achieve some goals.
-Our experiments show that {\em model recognition} is suitable for anticipating the actions of agents that are ruled by the execution of fixed programs or finite state machines, such as a {\em regular automata}.
+This paper formalizes the task of {\em model recognition} and introduces a novel method that estimates the probability of a \strips\ model to explain a partial observation of a plan execution.
+Our {\em model recognition as planning} approach builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and incomplete or missing intermediate states in the observation.
+Once the planning model of the observed agent is recognized, the model-based machinery for automated planning becomes fully applicable for other recognition tasks like {\em goal recognition}~\cite{ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} or {\em counter-planning}~\cite{PozancoEFB18}, which require beforehand a model of the observed agent.
+
+We show the effectiveness of our approach in three experiments that encode a set of different \strips\ models. 
+The first experiment is a proof-of-concept \emph{classification task} that perfectly recognizes the regular automata of a given string.
+Additionally, the two assumptions underpinning our formal theory are experimentally proved in the second experiment by varying the degree of observability.
+Finally, the robustness of the approach is shown in the third experiment where actions and intermediates states are unobservable.
+A meaningful lesson from the experiments is that our approach allows us to predict the future actions of agents whose behavior is ruled by a  model other than a deliberative planning model aimed at achieving some goals; this includes, for instance, models of fixed programs or finite state machines (regular automata).
+
+An interesting research line related to {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}, in which model edition is used to conform the PDDL models of two agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed optimal plan computed with one of the two models.
+{\em Model recognition}, however, conforms every input model $\mathcal{M}$ with another model $\mathcal{M}'$ that is not given as input but instead computed for every $\mathcal{M}$, and which is consistent with a partial observation of a plan execution.
 
@@
 Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
-
-\section{Conclusions}
-\label{sec:conclussions}
-This paper formalized the {\em model recognition} task and proposed, {\em model recognition as planning}, a method built on top of off-the-shelf classical planning algorithms to estimate the probability of a \strips\ model to produce a partial observation of a plan execution.
-The paper shows the effectiveness of {\em model recognition as planning} in a set of \strips\ models encoding different kinds of {\em automata}.
-{\em Model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unobserved.
 

### 2018-11-23 doc/aij18/major-revision/letter.tex
added comments by reviewer1 to editor letter format

\documentclass{article}

\usepackage[framemethod=tikz]{mdframed}

\mathchardef\mh="2D
\newcommand{\pre}{\mathsf{pre}}       % precondition
\newcommand{\eff}{\mathsf{eff}}       % effect
\newcommand{\cond}{\mathsf{cond}}     % conditional effect
\newcommand{\add}{\mathsf{add}}       % add effect
\newcommand{\del}{\mathsf{del}}       % delete effect
\newcommand{\PE}{\mathrm{PE}}         % precondition
\newcommand{\PSPACE}{\mathrm{PSPACE}} % PSPACE
\newcommand{\strips}{\textsc{Strips}} % Strips

\begin{document}

\subsection*{Reviewer \#1}

\vspace*{.3cm}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
I suggest to describe the overall concept of the learning approach first (see above for my hopefully correct summary) and to give the technical details later, rather than describing the technical details of compilation only without providing the overall picture. Going top down would make reading much easier. The second reason for revision is the comparison to ARMS. I suggest to compare with LOUGA that is better than ARMS in all aspects (runtime and quality of obtained models) so ARMS is no more state-of-the-art.
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
 Page 3, Section 2.1: The definition of effects is strange. The authors define positive and negative effects as literals. What if a negated predicate (which is a literal) is among positive effects? Does it mean the same as having the positive version of the predicate among negative effects? That is very confusing.

Also, there are additional assumptions about the model, namely positive effects cannot be among preconditions while negative effects must be among preconditions. It would be useful to discuss these assumptions more as there are domain models that do not satisfy these assumptions. Do we really need them? 
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Page 4: There is a nice discussion about partial traces, but I have found it incomplete because the relation between actions and state partial observability is not fully clear. For example, if we have FO for actions, does it mean FO or PO* for states? Also PO* for states looks like a special case of PO based on the definition. I would suggest to include some discussion there.
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Page 5, Table 2: I suggest including LOUGA there too.
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Page 10, Section 4.2: on(v1,v1) looks a bit strange. Do we really need such predicate with repeated variables?
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Page 11, Section 4.3: Maybe this section fits better in the section Background. Do you use the assumption about having negative effects among preconditions and not having positive effects among preconditions like before? This is not clear from the definition of triggered.
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Page 11, Section 4.4. This section is hard to understand, at least at the beginning. One needs to read it completely before understanding the details. I suggest to describe the concept at high level first and going to details then.
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Page 11, Section 4.4. This section is hard to understand, at least at the beginning. One needs to read it completely before understanding the details. I suggest to describe the concept at high level first and going to details then.
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Page 12: As I understand, the programming actions are building the model by adding preconditions and effects to the model. This immediately brings the question about symmetries as these programming actions can be used in any order (any permutation of them is fine). It would be useful to discuss this now. There are a few words about using SAT-based planner later, where parallel actions can be used. I think that this is critical for efficiency of the approach. Note also that the proposed method looks like a generate-and-test approach - we first generate a candidate model and then we verify that the model complies with plan traces. This does not seem particularly efficient.
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Page 13: There is implication used in the precondition. It was was never mentioned before that a general formula can be used.
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  My major problem here is that it is not clear how the actions validating states interleave with actions applying the learned actions. This must be clarified as the states are generated by actions.
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Page 20: In the graphs, it is not clear how the values are obtained. Are these average values between the domains or mean values or is it just for one selected domain (which?).
  The graph with runtimes show what I was afraid about - the efficiency degrades fast with larger example plans. This raises the question if longer plans can be used if needed for some domains.
\end{mdframed}

\subsection*{Reviewer \#2}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
\end{mdframed}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
\end{mdframed}

\subsection*{Reviewer \#3}

\end{document}

### 2018-11-23 doc/aij18/major-revision/letter.tex
added comments by reviewer3 with editor letter format

@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  It took me a while to figure out that the authors focus on a "consistency" kind of problem, in which the task is to find a model that is consistent with the observations.
+This is given in section 4.1, but I would liked an organized problem definition.
+This is important because throughout the paper you talk about soundness, completeness and computational complexity of the problem, but this only makes sense w.r.t a clear problem definition.
+
+For example, the authors discuss in section 3 the complexity of the learning task with different assumptions, but this is done before the problem is defined.
+Also, all the complexity of the results mentioned in page 8 are given without a proof, and the only reference in this context is to the Russell and Norvig book.
+The authors write in page 8 that FAMA solves the learning problem by compiling it to a planning problem, and consequently the planning problem is P-SPACE complete, but this is not a proof.
+In fact, it seems like this is an incorrect reduction: if planning is P-SPACE complete and learning can be solved by planning, it does not mean that learning is P-SPACE complete, as it may be easier.
 \end{mdframed}
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  There are several cases where authors repeat themselves several times. Here are a few examples:
+
+-          (page 2) "While current learning systems …. none of them allow partial observability in the sequence of executed actions "
+
+Two paragraphs later:
+
+" … unlike most relevant action-model learning algorithms, FAMA does not require the traces to contain any observed action …"
+
+(Page 8) "FAMA represents one step ahead … without assuming observed actions."
+
+-          A STRIPS action model is defined twice: once in section 2 and another time in section 4.
+
+-          The syntax-based notions of false positives and false negatives are defined twice: once in the second paragraph of section 5, and twice in Def.
+4-5.
+In fact, most of page 17 seems redundant to me, and the text should be merged with that in the previous page.
 \end{mdframed}
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Surprisingly, adding more traces causes the precision and recall of FAMA to drop for the NO action NO states setting.
+The authors say this is because more timeouts were observed due to the complexity of the model learning planning problem.
+To support this, the authors should report the number of timeouts for the different number of traces.
+In particular, since we do not see this drop in precision and recall in the other settings, I expect a comparison of the number of timeouts as a function of the number of traces for both settings.
+If we see that the number of timeouts for this setting is larger than the simpler setting, that would support the authors conclusion.
+
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  It was a bit weird to me that the authors chose to compile to a planning and not to a SAT problem like many prior work on learning action models.
+
+One reason to compile to planning is when the length of the solution plan is unknown, but the authors say that this is known in this case as well (see my question on this later).
+
+From my understanding of this work, there's a variable for every tuple of action name, fluent, and precondition/add-effect/delete-effect option.
+Then, there are constraints to hold that all this fall together in a consistent way with the observed trajectories.
+
+So, my question is: can the authors motivate the choice of planning over SAT?
+
+Since planning problems can be solved by compilation to SAT, and SAT problems can be solved by compilation to planning, it does not really matter to which problem you compile to, but still, I wonder what motivated the authors to take this approach.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Page 1.
+In general, the introduction left me very confusing, and I only really understood it after reading the paper.
+I wish the authors could do a better job in having the introduction not rely on understanding the rest of the paper so much.
+Concretely, talking about FAMA as a "solving scheme" "whose solution must be compliant with the input plan traces" did not make sense to me in the first pass.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+    Page 1.
+"… an abstract version of the capability model that reflects … " – I am not sure that "abstract" is the right term here, since the learned model is not an abstraction of the real-world but an approximation of it.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+    Page 1.
+"Two different types of data are generally identified in the generation of explanations … (2) data that do not accurately represent the decision-making process because the observed external behavior of the agent responds to a black-box model." – First, I think you meant "corresponds" and not "responds".
+Second, if the agent is treated as a black box, then this is not really a way to generate explanation.
+Please better explain the text after (2) .
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+      Page 2. "… the synthesis of different kinds of generative models with classical planning …" – I do not understand this sentence. What do you mean by generative models in this context?
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+      Page 2. "FAMA is a solving scheme … " – FAMA is a learning scheme, not a solving scheme. I agree that it is a learning scheme that learns by solving a planning problem.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+      Page 2. "… by the input observed actions, …" – What is "input observed actions"? please rephrase.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+      Page 3. "… the open world assumption" – Either explain what this means or provide a reference.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+      Page 3.
+"…the minimal action sequence to transit from state $s_i$ to $s_{i+1}$ is composed of a single action…" – This can be stated in a simpler way: "there exists a single action that transitions $s_i$ to $s_{i+1}$"
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+      Page 4. "A plan trace …. For a planning frame …. holds that … for every action in  … and that …" – this sentence is not clear. Please rephrase.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+      Page 4. The term action header is not defined.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+      Page 4. "… of the evaluation method to validate …" – missing "used" before "to"
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Page 4."Exceptionally, a NO state sequence in LOCM is a fully-empty trajectory, with neither initial or final state.
+" – I don't understand this sentence.
+Isn't a NO state sequence a fully empty trajectory for all algorithms?
+Also, it should be "nor" and not "or".
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+   Page 4."Most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces" – What about FAMA? I think it also requires this.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Pages 5-7. When describing prior work, it is sometimes not clear when you describe the algorithm and when you describe their specific implementation and experiments. Specifically:
+
+1.
+When describing SLAF, you write that 1,000 action-observation sequences and 10 fluents are selected.
+Is this part of the algorithm or just how they specifically implemented it in their experiments?
+
+2.
+When talking about LAMP, you write that it allows trajectories up to a "minimum percentage of 1/5 …" – again, is this requirement mandatory in their algorithm, or is it just what they happened to get in their experiments?
+Also, if you write 1/5 then you should write "ratio" and not "percentage".
+
+3.       The same question applies to your description of CAMA, when talking about 80\% empty states.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+    Pages 5-7. "… the state of a sort" – what is a "sort"?
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+    Pages 5-7. "As many of the approaches in section 2 assume, there may be an …" – this text is redundant. It just repeats again what was said twice already.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Page 8."… an unbound number of missing actions …" – should be "unbounded'
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Page 8.I disagree that having FO implies having human observers annotating the traces.
+For example, if one monitors a robot performing actions by monitoring the communications between the robot and its controller.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Page 8."When the plan traces is fully observed, learning STRIPS action models is straightforward" – The reference if for work done in 2018, while learning an action model in this setting has been discussed in IJCAI 2017 by Stern and Juba (Safe, model-free learning of action models").
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Page 8.The use of syntax and semantics in this stage was confusing to me. I am not sure how to improve this, though.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Pages 9-13
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Pages 9-13
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+   Pages 9-13
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+   Pages 9-13
 \end{mdframed}
@@
 
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+First sentence of the Introduction: there is no satisfactory or generally accepted definition of “complete” domain model, given that a domain model is an abstraction of the domain being modelled.
+If the authors use such a term they should  make clear what their definition is.
+“Adequate” is a better word to use - in the sense of being adequate enough for use with a planner to solve a set of planning  tasks.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  In this part and the rest of the paper I do not fully agree with the author’s hype
+
+about FAMA with respect to the input information required. The minimal amount of
+
+information that is needed (if I have got this right) is to
+
+specify every predicate that will be needed in the application (this is done in the initial state)
+
+the number of steps that are in the training plan, the set of all action headers including their
+
+list of parameters, and a goal state which refers to the same predicates as the initial state.
+
+ Comparing this to, say, LOCM, where no state/predicate information at all
+
+needs to be specified, I would say that in fact LOCM requires a lot less ‘information’
+
+a priori. What FAMA does not require is the identity of all the actions at each step,
+
+nor the instantiation of their headers, but it does require the definition of all
+
+the predicates that may appear in any state. The truth it seems to me is that
+
+FAMA would be good for some applications which fit this set of constraints, which is
+
+a different set of constraints than for other domain model learners.
+
+A way forward might be to put the claims about FAMA into a smaller set of learning systems
+
+(certainly ARMS will be in that set) where it clearly needs less input information.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  I totally agree with the authors in their assertion that the evaluation methods for
+
+previous model learners is currently poor, and amount to how close the learned model
+
+is to a GTM. I am not convinced, however, that the use of
+
+the statistic methods proposed is a great improvement. Precision and Recall
+
+are heavily used in data mining where (generally) there are huge data sets and there
+
+is not generally a hand crafted model for comparison. In planning, however, the idea of
+
+separating out a set of traces into training and testing appears odd, especially where
+
+it is known that the learning will eventually create a model that explains all the training
+
+traces. This seems particularly odd if there are only a few traces to be split up (and FAMA appears
+
+only able to cope  with a few traces because of complexity problems).
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  - The method of learning: This is to compile the learning problem into a conditional-operater
+
+classical planning problem. This method appears to be highly innovative to this line of research
+
+(stemming from an earlier ICAPS paper by the authors).
+
+The description of the compilation method did not appear to me to be at the level
+
+of the detail to allow easy reading. It took me a good deal of time to work out
+
+what going on. This is a complex process and I recommend much more use of examples
+
+throughout - perhaps a running example to show the full run of a compilation.
+
+It might also improve the paper if an algorithmic description of compilation
+
+were included? One aspect that was not obvious from the description was how a
+
+SET of models are delivered by planning rather than just one; this needs spelling out.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  - The evaluations of the domain model using classification metrics of precision and recall:
+
+I like the distinction of the methods labelled "syntactic" (using the GTM) vs "semantic" here;
+
+however I think the use of the word - "semantic” a bit misleading. This implies
+
+taking into account the meaning of the features within the application domain which are
+
+left out or in erroneously compared to the GTM e.g. that a missing "on(x,y)” in a learned model
+
+is somewhat better/worse from a missing ”holding(x)” - because of their meaning, rather than
+
+counting each predicate as ‘1’.
+
+In the end your method reverts to counting the number of different predicates, but in the context
+
+of the recall/precision formula with respect to domain model(s) that are consistent with the test
+
+sets, not only the GTM.
+
+The definition of the metrics was not too clear. Intuitively, I think I understood 5.1 but it
+
+was not well stated -the definition of some terms seemed syntactically wrong or at least sloppy.
+
+E.g. in the definition of INS and DEL, should not the universally quantified variables be further
+
+qualified by stating that there is a union of the already unioned sets? Or again, should there not
+
+be an extra "Sigma" to represent the addition of the addition of the number of precons and
+
+effects in size(M) definition?  In proposition 7 you say the closest set "is the GTM"
+
+do you mean equal to the singleton set of the GTM?
+
+Also - the structure of 5.1 - it would be best to clearly state that
+
+the syntactic evaluation is going to be introduced, then as a further subsection the semantic
+
+evaluation.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  - I advise changing title Experimental Results => Experimental Evaluation
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  - At the beginning of the evaluation you state the obvious - that the purpose is to evaluate
+
+the performance of FAMA and the quality of the models. At this stage to clarify the experimental
+
+section I suggest that you indicate, before detailing the set up, what would be success, for example
+
+how do you intend to show that your technique is better than or comparable to others? This will
+
+no doubt mean introducing the metrics you intend to use and giving an idea of success
+
+in terms of them. Also it may mean taking text from the start of the sections 6.2 - 6.5
+
+ into the first part of the evaluation section. This will however give a good rational for the
+
+'set-up’, the choice of the particular experimental scenarios, and for the discussion of the
+
+results.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  - Why "14" models? please give some explanation (e.g. there are only 14 strips models there)
+
+so we know why/if you have left any out
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  was the max predicate arity of "2" in all of the 14 a big issue? Given the inherent computational
+
+problems in planning I guess so?
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  so for the NO / NO case study, we know the full IS , GS, the number of action steps,
+
+and the space of possible action headers for each step - is that correct?
+
+If so please make it more explicit - it not say explicitly what it means.
+
+This was not very explicit in the introduction/compilation process
+
+description either.
+\end{mdframed}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  Typos spotted
+
+“around a hundred of traces”p.20
+
+Please replace “fairly sound and almost totally complete” with something more precise. p.21
+
+“reason why the scores of” p.26
+\end{mdframed}
+
 \end{document}

### 2018-11-23 doc/aij18/major-revision/introduction.tex
added comments by reviewer3 with editor letter format

--- 2018-08-04 introduction.tex
@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behaviour is not available.
-Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimum input knowledge.
+Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimal input knowledge.
 
@@
 \FAMA, on the other hand, allows incomplete or empty action sequences.
-This way, the minimal observability case acceptable by \FAMA is when the algorithm is only fed with the input and output of a plan trace, namely the initial and final state of a planning task.
+This way, the minimum observability case acceptable by \FAMA is when the algorithm is only fed with the input and output of a plan trace, namely the initial and final state of a planning task.
 Like many Machine Learning (ML) techniques, \FAMA is able to operate with only input/output pairs and an unknown or a partially known model of the agent.
@@
 \item A unified formulation for learning and evaluating action models from observations of plan executions.
-In the case of minimal observability, these executions only comprise the initial and final state of the plan traces.
+In the case of minimum observability, these executions only comprise the initial and final state of the plan traces.
 \item A thorough elaboration of two semantic evaluation metrics that build upon the notions of {\em precision} and {\em recall} to evaluate the output action models with respect to observations of plan executions.

### 2018-11-23 doc/aij18/major-revision/related_work.tex
added comments by reviewer3 with editor letter format

--- 2018-04-16 doc/aij18/related_work.tex
@@
 A plan trace is a combination of a state and an action sequence $\tau = \langle s_0, a_1, a_1, a_2, s_2, \ldots, a_n, s_n \rangle$.
-Additionally, $\tau$ is a plan trace with partial observable (PO) states if there is at least one missing literal in any state $s_i$.
-And $\tau$ is a plan trace with PO actions of there is at least one missing action from the action sequence.
+Additionally, $\tau$ is a plan trace with partial observable (PO) states if one or more propositions are missing in any state $s_i$.
+And $\tau$ is a plan trace with PO actions if at least one action is missing in $\Omega$.
 

### 2018-11-23 doc/aij18/major-revision/letter.tex
added comments by reviewer2 with editor letter format

@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  Pages 9-13
+-          "On the other hand, a striking figure …" – I could not understand this sentence.
+
+-          "Let $\Omega_v = {v_i}_{i=1}^{max_{a\in A} ar(a)}$ be a new set of objects …" – I think "ar(a)" should be "|ar(a)|"
+
+-          "… which is bound to …" – I think should be " which is bounded by"
+
+-          "In more details, for a given …" – delete the "In more details" as you are defining something new, not providing additional details.
+
+-          Please elaborate on the computation of the size of the space of possible STRIPS models.
+
+-          The definition of conditional effects given in 4.3 should be moved to the background section, where you define a STRIPS model.
+
+-          The explanation about the plan prefix and postfix is confusing and not helpful.
+I recommend removing it.
+Only after you explain your solution it starts to make sense.
+Also, I am not sure the term "postfix" is suitable.
+Maybe "suffix"?
+
+-          The formula for the precondition of apply uses the notation $\Rightarrow$ which was not defined.
 \end{mdframed}
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  Pages 9-13
+Pages 15-19.
+
+-          I'd appreciate more details on the mentioned post-processing, e.g., a pseudo code.
+
+-          At the bottom of the page, you write that VAL cannot address the model validation of a partial model and on the other hand it requires a full plan and full action model.
+I don't see how this is "on the other hand"?
+it is the same "hand"
+
+-          When introducing the semantic evaluation metrics, you provide two justification.
+One is that there is no GTM.
+The other is that a test-based evaluation is preferable.
+The latter justification is not really a justification.
+It is like saying "we justify using a test-based evaluation by the fact that test-based evaluation is sometimes preferred".
+
+-          What does it mean that a learned action model is sound and complete?
+
+-          "… to flaws that appear more than once in the plan traces …" – I recommend to clarify: "flaws in the action model that manifest more than once in the plan traces".
+
+-          "… generated 10 plan traces (each with 10 actions …" – the parenthesis should be removed.
+
+-          "… the horizon of the solution plan is also known.
+" – how do you know the horizon?
+You know the length of the given traces, but how do you know how many actions are needed to build the action model?
+
+-          "… a solution plan is solvable in two steps" – which two steps? I thought you put it all in the planner and get both action model and validation in the same step.   
 \end{mdframed}
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-   Pages 9-13
-\end{mdframed}
+   Pages 20-end of paper:
 
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-   Pages 9-13
-\end{mdframed}
+-          "… the size of the input knowledge, in the performance …" – should be "on the performance"
 
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-\end{mdframed}
+-          "… the Precision and Recall" – why capitalized?
 
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+-          "retain a level of soundness similar to " – what is this "level of soundness"? do you talk about precision?
+
+-          "By lightening the input constraints …" – maybe "relaxing" instead of "lightening"?
 \end{mdframed}

### 2018-11-26 doc/ijcai-19-conformant/conformant-learning.tex
adding the classical planning and observation model to ijca19 paper on learning with conformant plannign

--- 2018-11-18 doc/icaps-19/model-recognition.tex
@@
-\def\year{2017}\relax
-%File: formatting-instruction.tex
-\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
-\usepackage{aaai19}  %Required
-\usepackage{times}  %Required
-\usepackage{helvet}  %Required
-\usepackage{courier}  %Required
-\usepackage{url}  %Required
-\usepackage{graphicx}  %Required
-\frenchspacing  %Required
-\setlength{\pdfpagewidth}{8.5in}  %Required
-\setlength{\pdfpageheight}{11in}  %Required
-%PDF Info Is Required:
-  \pdfinfo{}
-\setcounter{secnumdepth}{0}
+%%%% ijcai19.tex
 
+\typeout{IJCAI-19 Instructions for Authors}
+
+% These are the instructions for authors for IJCAI-19.
+
+\documentclass{article}
+\pdfpagewidth=8.5in
+\pdfpageheight=11in
+% The file ijcai19.sty is NOT the same than previous years'
+\usepackage{ijcai19}
+
+% Use the postscript times font!
+\usepackage{times}
+\usepackage{soul}
+\usepackage{url}
+\usepackage[hidelinks]{hyperref}
+\usepackage[utf8]{inputenc}
+\usepackage[small]{caption}
+\usepackage{graphicx}
 \usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{multirow}
-\usepackage{tikz}
-\usetikzlibrary{arrows,automata}
-\usepackage{comment}
+\usepackage{booktabs}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\urlstyle{same}
 
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{listings}
-
-\lstset{
-  basicstyle=\ttfamily,
-  mathescape
-}
-
-\usepackage{multicol}
-\usepackage{arydshln}
-\usetikzlibrary{calc,backgrounds,positioning,fit}
-
+%%%%%%%%%%%%%%%%%% Added for this paper
+\usepackage{ wasysym }
 \newcommand{\tup}[1]{{\langle #1 \rangle}}
-
 \newcommand{\pre}{\mathsf{pre}}     % precondition
@@
 \newcommand{\PE}{\mathrm{PE}}     % precondition
-\newcommand{\strips}{\textsc{Strips}}     % precondition
+\newcommand{\strips}{\textsc{Strips}}
 
@@
 
+%%%%%%%%%%%%%55
+
+% the following package is optional:
+%\usepackage{latexsym}
+
+% Following comment is from ijcai97-submit.tex:
+% The preparation of these files was supported by Schlumberger Palo Alto
+% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
+% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
+% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
+% preparation.
+
+% These instructions can be modified and used in other conferences as long
+% as credit to the authors and supporting agencies is retained, this notice
+% is not changed, and further modification or reuse is not restricted.
+% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
+% contacts for providing assistance without their prior permission.
+
+% To use for other conferences, change references to files and the
+% conference appropriate and use other authors, contacts, publishers, and
+% organizations.
+% Also change the deadline and address for returning papers and the length and
+% page charge instructions.
+% Put where the files are available in the appropriate places.
+
+\title{Computing the least commitment action models from observations of plan executions}
+
+% Single author syntax
+%\author{
+%    Sarit Kraus
+%    \affiliations
+%    Department of Computer Science, Bar-Ilan University, Israel \emails
+%    pcchair@ijcai19.org
+%}
+
+% Multiple author syntax (remove the single-author syntax above and the \iffalse ... \fi here)
+% Check the ijcai19-multiauthor.tex file for detailed instructions
+\author{
+Diego Aineto$^1$\and
+Sergio Jim\'enez$^1$\and
+Eva Onaindia$^1$\And
+\and
+Blai Bonet$^2$
+\affiliations
+$^1${\small Departamento de Sistemas Inform\'aticos y Computaci\'on. Universitat Polit\`ecnica de Val\`encia. Valencia, Spain}\\
+$^2${\small Departamento de Computaci\'on. Universidad Sim\'on Bolívar. Caracas, Venezuela}
+\emails
+{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es, bonet@usb.ve}}
+
 \begin{document}
 
-\title{Model Recognition as Planning}
+\maketitle
 
-% Commented for blind submission
-\author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
-{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
-{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
-{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
-\And Miquel Ram\'irez\\
-{\scriptsize School of Computing and Information Systems}\\
-{\scriptsize The University of Melbourne}\\
-{\scriptsize Melbourne, Victoria. Australia}\\
-{\scriptsize miquel.ramirez@nimelb.edu.au}}
+\begin{abstract}
 
-\maketitle
-\begin{abstract}
-Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action schemata), {\em model recognition} is the task of identifying the model that explains the observation.
-The paper formalizes this task and introduces a novel method that estimates the probability of a \strips\ model to produce an observation of a plan execution.
-This method builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and intermediate states in the observation.
-The effectiveness of the method is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a classical string classification task; (2) identification of the model that encodes a failure present in an observation; and (3) recognition of a robot navigation policy.
 \end{abstract}
@@
 \section{Introduction}
-\label{sec:introduction}
-
-A growing interest in addressing the task of recognizing the mission (plans and goals) of a planning agent from a set of given observations has recently emerged.
-Approaches based on rule-based systems, parsing, graph-covering or Bayesian nets have been proposed for the plan/goal recognition tasks~\cite{sukthankar14}.
-{\em Plan recognition as planning} is the model-based approach for this task that leverages the action model of the observed agent to compute its most likely goal and predict its future actions~\cite{ramirez2009plan,ramirez2012plan}.
-While action models are learnable from observations of the agent~\cite{yang2007learning,cresswell2013acquiring,aineto2018learning}, real-world applications developing in a particular context usually dispose of approximate behavioral models of the involved agents and there is rarely a need to learn a model from scratch.
-A more relevant task in these scenarios is thereby to predict the behavior of an agent from a set of given models.
-
-This paper formalizes the novel task of {\em model recognition}, where the object to recognize is not a goal or a plan but the model that shapes the behavior of the observed agent.
-Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action schemata), {\em model recognition} is the task of identifying the model in the set with the highest probability of producing the given observation.
-We posit this problem as a \emph{classification task} where classes are given as generative planning models and samples are sequences of observations.
-We accomplish this task with off-the-shelf classical planning algorithms that compute for each model (class) $\mathcal{M}$ in the set a model $\mathcal{M'}$ that explains the observation, and subsequently distances between the two models are measured.
-Two key contributions of our {\em model recognition as planning} approach are: (1) the ability of handling partial observations of plan executions that contain no actions and incomplete -- even empty- intermediate states; and (2) a thorough formalization that induces a probability distribution of a model $\mathcal{M}$ to explain an observed plan execution.
-
-\begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-  \begin{tikzpicture}[scale=.4]
-          \begin{scope}
-            \draw (0, 0) grid (5, 5);
-            \node[anchor=center] at (2.5, 1.5) {R};
-            \node[anchor=center] at (0.5, 0.5) {};
-            \draw[thick,style=dotted] (0.5,0.5) -- (3.5,0.5);
-            \draw[thick,style=dotted] (3.5,0.5) -- (3.5,1.5);
-            \draw[thick,style=dotted,->] (3.5,1.5) -- (2.7,1.5);
-          \end{scope}
-        \end{tikzpicture}
-\hspace*{1cm}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-	  \node[state] (A)              {$q_0$};
-	  \node[state] (B) [right of=A] {$q_1$};
-	  \path
-(A) edge [bend left, align=center] node {inc-y,\\dec-y} (B)
-(B) edge [bend left, align=center] node {inc-y,\\dec-y} (A);
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}
- \caption{\small ({\em Left)} Robot navigating a $5\times 5$ grid.
-{\em (Right)} Automata to control that the robot only increments its x-coordinate when $q_0$ holds (actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and switch the automata state).}
-\label{fig:grid-example}
-\end{figure}
-
-To better illustrate {\em model recognition}, imagine a robot in a $N\times N$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model, the robot can increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
-at even rows if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at odd rows if {\tt\small $q0$} holds initially).
-Besides this model, other models that shape different kinds of robot navigation can be defined within the same state variables (e.g.
-altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated).
-Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where state variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, {\em model recognition} aims to identify the robot action schemata that explains this observation.
-
-We provide a formalization of {\em model recognition} that is applicable to any planning model but we restrict our attention to \strips\texttt{-}compilable models.
-It is well-known that diverse automata representations like finite state controllers, {\sc GOLOG} programs or reactive policies can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,segovia2017generating}.
-As for the evaluation, the effectiveness of our approach is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a proof-of-concept classification task; (2) a simulation of a non-deterministic \emph{blocksworld} under different degrees of observability; and (3) a navigation task where observations contain an unbounded number of unobservable states.
-
-%: a classical string classification task; identification of the model that encodes a failure present in an observation; and recognition of a robot navigation policy.
-
-%The paper also introduces {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
-Our method is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.
-
-\begin{figure}
-  \begin{tiny}
-  \begin{verbatim}
-  (:action inc-x
-    :parameters (?v1 ?v2)
-    :precondition (and (xcoord ?v1) (next ?v1 ?v2) (q0))
-    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))
-
-  (:action dec-x
-    :parameters (?v1 ?v2)
-    :precondition (and (xcoord ?v1) (next ?v2 ?v1) (q1))
-    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))
-
-  (:action inc-y-even
-    :parameters (?v1 ?v2)
-    :precondition (and (ycoord ?v1) (next ?v1 ?v2) (q0))
-    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
-                 (not (q0)) (q1)))
-
-  (:action inc-y-odd
-    :parameters (?v1 ?v2)
-    :precondition (and (ycoord ?v1) (next ?v1 ?v2) (q1))
-    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
-                 (not (q1)) (q0)))
-
-  (:action dec-y-even
-    :parameters (?v1 ?v2)
-    :precondition (and (ycoord ?v1) (next ?v2 ?v1) (q0))
-    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
-                 (not (q0)) (q1)))
-
-  (:action dec-y-odd
-    :parameters (?v1 ?v2)
-    :precondition (and (ycoord ?v1) (next ?v2 ?v1) (q1))
-    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
-                 (not (q1)) (q0)))
-  \end{verbatim}
-  \end{tiny}
- \caption{\small Example of a \strips\ action schemata (given in PDDL) for robot navigation in a $N\times N$ grid.}
-\label{fig:model-example}
-\end{figure}
 
 \section{Background}
-\label{sec:background}
 This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
-
 \subsection{Classical planning with conditional effects}
@@
 
-Figure~\ref{fig:grid-example} illustrates a {\em partial observation} of a five-state trajectory \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 4)(ycoord 2)>, <(xcoord 3)(ycoord 2)>}\}.
-Note that this observation only contains fluents of the type {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is not observable in any of the five states.
+\subsection{Conformant planning}
 
-\section{Model Recognition}
-\label{sec:recognition}
-The {\em model recognition} task is a tuple $\tup{P,M,\mathcal{O}}$ where:
-\begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions.
-For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined.
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_k\}$ is a set of {\em k different planning models} for the actions in $A[\cdot]$.
-A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$.
-Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
-\end{itemize}
+\section{Learning the least commitment action models from observations}
 
-Model recognition can be understood as a {\em classification task} where each class is represented by a different planning model $\mathcal{M}\in M$, and the observed plan execution $\mathcal{O}(\tau)$ is the single example to classify.
-The planning model associated to each class acts as the corresponding {\em class prototype} and it summarizes any observation of a plan execution that could be synthesized with such model (i.e.
-the set of all the examples that belong to that class).
-We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to a given observation $\mathcal{O}(\tau)$.
-The {\em solution} to the model recognition task is then the subset of models in $M$ that maximizes this expression.
+\subsection{The learning task}
 
-\begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{O}|\mathcal{M}) P(\mathcal{M}).
-\end{align}
-
-\subsection{Formulating the $P(\mathcal{O}|\mathcal{M})$ likelihood}
-
-The $P(\mathcal{M})$ probability expresses whether one model is known to be a priori more likely than the others.
-If this probability is not given as input it is reasonable to assume that, {\em a priori}, all models are equiprobable.
-The challenge of our formulation to the model recognition task is the definition of $P(\mathcal{O}|\mathcal{M})$, the likelihood that expresses the probability of observing $\mathcal{O}(\tau)$ when $\mathcal{M}$ is the planning model.
-
-\begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-	  \node[state] (A)              {$\mathcal{M}$};
-	  \node[state] (B) [right of=A] {$\mathcal{M}'$};
-	  \node[state] (C) [right of=B] {$\tau$};
-	  \node[state] (D) [right of=C] {$\mathcal{O}$};
-	  \path
-(A) edge  node {} (B)
-(B) edge  node {} (C)
-(C) edge  node {} (D)
-;
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}
- \caption{\small {\em Bayesian network} representing that model $\mathcal{M}$ is {\em transformable} into a model $\mathcal{M}'$ that produces a trajectory $\tau(\pi,P)$ that (1) reaches the goals in $P$ and (2) it is consistent with $\mathcal{O}(\tau)$.}
-\label{fig:net}
-\end{figure}
-
-Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ likelihood is to assess the cost of transforming $\mathcal{M}$ into a model $\mathcal{M'}$ that produces a trajectory $\tau(\pi,P)$ such that: (1) $\tau(\pi,P)$ reaches the goals in $P$ and (2) $\tau(\pi,P)$ is consistent with observation $\mathcal{O}(\tau)$.
-Figure~\ref{fig:net} shows the {\em Bayesian network} that embodies this procedure.
-Regarding this network we have the following formulation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}:
-
-\begin{align}
- P(\mathcal{O}|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\tau} P(\mathcal{M'}|\mathcal{M})P(\tau|\mathcal{M'})P(\mathcal{O}|\tau),
-\end{align}
-
-where $\tau$ ranges over all the trajectories consistent with $\mathcal{O}(\tau)$ that can be synthesized with a model $\mathcal{M}'$ and $\mathcal{M}'$ ranges over all the models that can be generated {\em transforming} $\mathcal{M}$.
-
-The exact computation of $P(\mathcal{O}|\mathcal{M})$ with equation (2) is intractable.
-For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge, infinite in the case of planning problems without dead-ends~\cite{lesh1995sound}.
-Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially with the number of state variables.
-Instead, our approach is to estimate of $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} referred to \strips\ planning models.
-{\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, which have been proved successful for {\em pattern recognition}~\cite{MasekP80,Bunke97}.
-In this work, we assess the cost of {\em transforming} $\mathcal{M}$ into $\mathcal{M'}$ computing {\em edit distances} between \strips\ planning models.
-
-\section{Recognition of \strips\ planning models}
-\label{sec:asPlanning}
-
-Our formalization of a model recognition task is valid for any planning model but our \emph{edit distance} measure is exclusively for \strips\ models and so we restrict our attention to \strips\texttt{-}compilable planning models\footnote{An edit distance for other planning models is also definable}.
-Thus, we focus on the recognition of $\mathcal{M}\in M$, where $M$ is a set of \strips\ planning models and $\mathcal{M}$ defines the semantics of the actions through a set of \strips\ action schemata.
-
-\subsection{Well-defined \strips\ action schemata}
+\subsection{The hypothesis space}
 \strips\ action schemata provide a compact representation for specifying classical planning models.
-Figure~\ref{fig:model-example} shows six \strips\ action schemata that shape a particular kind of robot navigation in $N\times N$ grids (no matter the grid size).
-
 {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
@@
 
-For instance, we claim that the six action schemata of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are not.
+For instance, the {\small\tt stack(?v1,?v2)} and {\small\tt unsstack(?v1,?v2)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} are not.
 Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
@@
 We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
-For any of the six action schemata of Figure~\ref{fig:model-example}, the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
 
@@
 Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-For every action schema in the navigation model of Figure~\ref{fig:model-example} then $2^{2\times 10}=1,048,576$.
 
@@
 One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in a {\em robot navigation} model like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold at the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
-Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear in any of these lists because the {\tt\small next} predicate codes the {\em successor} function for {\em natural numbers}.
-These domain-specific constraints reduce further the size of the space of possible action schemata to $2^{2\times 7}=16,384$ (for every schema in the navigation model of Figure~\ref{fig:model-example}).
+For instance, in a {\em blocksworld} one can claim that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in any of these lists because of the semantic of the {\tt\small on} predicate that codes which block is in top of other block.
 \end{itemize}
@@
 
-\subsection{Edit distances for \strips\ planning models}
-First, we define the two edit \emph{operations} on a schema $\xi$ that belongs to a \strips\ model $\mathcal{M}\in M$:
+\subsection{A propositional encoding for \strips\ action schema}
+Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
+Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.
 
-\begin{itemize}
-\item {\em Deletion}. Given $\xi\in\mathcal{M}$, an element from any of the lists $pre(\xi)$/$del(\xi)$/$add(\xi)$ is removed such that the result is a {\em well-defined} \strips\ action schema.
-\item {\em Insertion}.
-Given $\xi\in\mathcal{M}$, an element in ${\mathcal I}_{\Psi,\xi}$ is added to any of the lists $pre(\xi)$/$del(\xi)$/$add(\xi)$ such that the result is a {\em well-defined} action schema.
-\end{itemize}
+The interest of having a propositional encoding for \strips\ action schema is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}.
+Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state.
+Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
+\begin{small}  
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
+\cond(\mathsf{editable_{\xi}})=&\{del\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+&\{add\_e\_\xi\}\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
+\end{align*}
+\end{small}
 
-We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ models are.
-The distance is symmetric and meets the {\em metric axioms} provided that the two edit operations, {\em deletion} and {\em insertion}, have the same positive cost.
-
-\begin{definition}[Edit distance]
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} and {\em well-defined} \strips\ planning models within the same set of predicates $\Psi$.
-The {\bf edit distance} $\delta(\mathcal{M},\mathcal{M}')$ is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
-\end{definition}
-
-Since ${\mathcal I}_{\Psi,\xi}$ is a bounded set, the maximum number of edits that can be introduced to an action schema is bounded as well.
-The \textbf{maximum edit distance} of a \strips\ model $\mathcal{M}$ built with predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$ (note that if we consider the set of syntactic constraints then $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2\times|{\mathcal I}_{\Psi,\xi}|$).
-
-\vspace{0.02cm}
-
-An observation of the execution of a plan generated with $\mathcal{M}$ further constraints the space of possible action schemata of $\mathcal{M}$.
-The \emph{semantic knowledge} included in the observations introduce a third type of constraints, that we will call {\em observation constraints}, and that can be added to the set $\mathcal{C}$.
-In addition, {\em observation constraints} allow us to define an edit distance to elicit the value of $P(\mathcal{O}|\mathcal{M})$.
-It can be argued that the shorter this distance the better the given model explains the given observation.
-
-\begin{definition}[Observation edit distance]
-Given a planning problem $P$, an observation $\mathcal{O}(\tau)$ of the execution of a plan that solves $P$ and a \strips\ planning model $\mathcal{M}$ (all defined within the same set of predicates $\Psi$).
-The {\bf observation edit distance}, $\delta^o(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$; \[\delta^o(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
-\end{definition}
-
-$\delta^o(\mathcal{M},\mathcal{O})$ can also be defined through the edition that the observation $\mathcal{O}(\tau)$ requires to fit $\mathcal{M}$.
-This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of the model $\mathcal{M}$~\cite{yang2007learning,SohrabiRU16}.
-Our definition of {\em observation edit distance} is more practical since the size of ${\mathcal I}_{\Psi,\xi}$ is usually much smaller than $F$ (the number of variables in the action schemata should normally be lower than the number of objects in a planning problem).
-
-\begin{definition}[{\em Closest consistent models}] \label{consistent}
-Given a model $\mathcal{M}$, the set $M^*$ of the {\bf closest consistent models} is the set of models $\mathcal{M}'$ that: (1) produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$ and (2) their {\em edit distance} to $\mathcal{M}$ is minimal;
-  \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
-\end{definition}
-
-\subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
-
-Now we we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ action schemata.
-
-\subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to understand how to build a reasonable estimate of $P(\mathcal{O}|\mathcal{M})$ for the general case.
-
-Under the assumption of {\em full observability}, there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the input observation so $P(\mathcal{O}|\tau^*)=1$.
-Further, there is also a single model that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory so they can be considered the same model).
-
-Provided that there is a single possible trajectory and a single possible model consistent with the input observation, i.e.
-$M^*=\{\mathcal{M}^*\}$ then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
-\begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}) P(\mathcal{M}).
-\end{align}
-Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because is independent of the input model $\mathcal{M}\in M$.
-
-\subsubsection{Partial observability of the executed plan.} In a similar way, an approximation to $P(\mathcal{O}|\mathcal{M})$ can be built for the general case, where the executed plan is partially observed.
-We add the following two assumptions to deal with this general case:
-
-\begin{enumerate}
-\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state~\cite{Dennett83}.
-\item {\em The best plans for $P$ are unique or have different cost}.
-\end{enumerate}
-
-Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}.
-They allow us to assume that the sum in equation (2) is dominated by its largest term, so other terms in the sum are not added up.
-The largest term corresponds here to the shortest trajectory $\tau^*$ consistent with the observation (which now is assumed to be unique).
-Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more likely to assume that there is only one trajectory consistent with the input observation.
-
-Again, we have that there is a single model $\mathcal{M}^*$ that can produce a given trajectory $\tau^*$.
-With this said, both $P(O|\tau^*)$ and $P(\tau^*|\mathcal{M}^*)$ are independent of $\mathcal{M} \in M$ so equation (1) simplifies once again to equation (3) under the previous two assumptions.
-
-\subsubsection{The $P(\mathcal{M^*}|\mathcal{M})$ probability distribution.} $P(\mathcal{M'}|\mathcal{M})$ indicates the probability of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ by exclusively using the two {\em edit operations} previously defined, {\em deletion} and {\em insertion}.
-
-We know from {\em pattern recognition}~\cite{devroye2013probabilistic} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}.
-Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
-\begin{align}
-CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d (1-p)^{N-d}
-\end{align}
-where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than do not applying it.
-
-With this regard, and considering that \strips\ models $\mathcal{M}\in M$ can be encoded with a propositional representation of fixed length $N$, we formulate the $P(\mathcal{M'}|\mathcal{M})$ probability distribution mapping the distance $\delta(\mathcal{M},\mathcal{M'})$ according to equation:
-\begin{align}
-P(\mathcal{M'}|\mathcal{M}) = p^d  (1-p)^{N-d}
-\end{align}
-where $d=\delta(\mathcal{M},\mathcal{M}')$.
-Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation (5) but instead, the parameter $d$ is given by the {\em observation distance}, $d=\delta^o(\mathcal{M},\mathcal{O})$.
-
-In other words, we are modeling the edition of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.
-
-\section{Model Recognition as planning}
-
-This section shows that $\delta^o(\mathcal{M},\mathcal{O})$, and hence an approximation to $P(\mathcal{O}|\mathcal{M})$), can be computed with a compilation-to-planning approach as the one proposed in~\cite{aineto2018learning} (AJO approach hereafter) for learning \strips\ models.
-The AJO approach receives as input an empty model $\mathcal{M}$, which only contains the headers of the action schemata formed of $\xi=\tup{name(\xi),pars(\xi)}$, and an observation of a plan execution $\mathcal{O}(\tau)$ (extensible to a set of observations), and it returns a model $\mathcal{M'}$ with specification of preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$ following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1},a_i)$ for every observed action of $\mathcal{O}(\tau)$ and $s_i=\theta(s_{i-1},a_i)$ for every observed state of $\mathcal{O}(\tau)$.
-
-Essentially, the task $\tup{\mathcal{M},\mathcal{O}}$ of the AJO approach -- learning $\mathcal{M}$ that satisfies $\mathcal{O}$-- is interpreted as editing the empty action schemata of $\mathcal{M}$ introducing preconditions and effects until $\mathcal{O}$ is validated with the resulting model.
-We leverage the same idea to compute $\delta^o(\mathcal{M},\mathcal{O})$ with the exception that now our $\mathcal{M}$ is not empty but $\mathcal{M}=\{\xi_1, \ldots, \xi_m\}$, where $\xi_i=\tup{name(\xi_i),pars(\xi_i),pre(\xi_i),add(\xi_i),del(\xi_i)}$, $1 \leq i \leq m$.
-
-Editing/learning the action schemata of $\mathcal{M}$ in the AJO approach is addressed converting the task into a classical planning problem, which is later solved with a planner.
-The intuition behind this compilation is that a solution plan to the problem is a sequence of: (a) \emph{edit actions} on the schemata of $\mathcal{M}$ to build $\mathcal{M'}$ and (b) \emph{validate actions} that apply $\mathcal{M'}$ in $\mathcal{O(\tau)}$.
-The adaptation of this compilation scheme for solving $\delta^o(\mathcal{M},\mathcal{O})$ results in a planning problem $P'=\tup{F',A',I',G'}$ whose objective is to determine the preconditions and effects that need to be added or deleted to the action schemata of $\mathcal{M}$ so as to satisfy $\mathcal{O}$.
-The accomplishment of this task requires therefore a propositional encoding of the components of the action schemata.
-Specifically:
-
-\begin{itemize}
-\item $F'$ contains the necessary fluents to represent the $pre$, $add$ and $del$ of the action schemata.
-It is a set of \emph{editable fluents} of the type $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
-Additionally, $F$ also contains fluents to encode the \emph{observation constraints}; that is, fluents to iterate through the $l$ observed actions and $m$ observed states of $\mathcal{O(\tau)}$.
-\item $A'$ comprises two types of actions with conditional effects:
-    \begin{itemize}
-    \item actions for editing $\xi \in \mathcal{M}$ that follow the \emph{syntactic constraints} for well-defined \strips\ action schemata.
-Hence, $A'$ contains actions for inserting a new precondition $pre\_e\_\xi$ or a new positive effect $add\_e\_\xi$ in $\xi$ when neither $pre\_e\_\xi$, $del\_e\_\xi$ or $add\_e\_\xi$ exist in $\xi$, and actions for inserting a new negative effect when neither $del\_e\_\xi$ or $add\_e\_\xi$ appear in $\xi$ but $pre\_e\_\xi$ does.
-    \item actions for applying the new action schemata of the edited model $\mathcal{M'}$ and validating the observed states of $\mathcal{O(\tau)}$.
-Particularly, the \emph{apply} actions check the preconditions and produce the effects defined by the \emph{editable fluents}.
-    \end{itemize}
-\item $I'$ encodes the editable fluents $pre\_e\_\xi, del\_e\_\xi$ and $add\_e\_\xi$ that hold in the action schemata of $\mathcal{M}$.
-\item $G'$ contains the necessary fluents to check that all the observed actions and states of $\mathcal{O(\tau)}$ are generated correctly following the edited model $\mathcal{M'}$.
-    %Here we include the necessary fluents to check the final reached state is consistent with the final state $s_m^o$ of $\mathcal{O(\tau)}$.
-\end{itemize}
-
-We now show an example of a solution plan to a planning problem $P'$ that results from compiling a specific task $\tup{\mathcal{M},\mathcal{O}}$.
-Consider that $\mathcal{M}$ is the model of Figure~\ref{fig:model-example} except that the schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped with respect to Figure~\ref{fig:model-example}.
-And consider an observation that only contains the initial and final state of Figure~\ref{fig:grid-example}; i.e., $\mathcal{O(\tau)}=\tup{s_0^o,s_m^o}=\tup{\{{\texttt{\small{(xcoord 0)(ycoord 0)}}}\}, \{{\texttt{\small{(xcoord 2)(ycoord 1)}}}\}}$.
-The plan found by a planner to $P'$ is shown in figure~\ref{fig:plan-pdistance}.
-The first seven steps are the edit actions to \emph{fix} the schema {\tt\small inc-x}; step 07 is a validate action that sets the robot in the initial state $s_0^o=\tup{\{{\texttt{\small{(xcoord 0)(ycoord 0)}}}\}}$; step 08 applies the action {\tt\small (inc-x 0 1)} and moves the robot one cell to the right to position $(1,0)$; steps 09-11 also move the robot one cell to the right; step 11 applies an action that increases coordinate $y$ from a position in an even row number, thus moving robot to row $1$; step 12 moves the robot one cell to the left and, finally, action {\tt\small (validate\_1)} checks the robot position is consistent with the final state $s_m^o=\tup{\{{\texttt{\small{(xcoord 2)(ycoord 1)}}}\}}$.
+Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}).
+Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
 
 \begin{figure}
-{\tt\tiny
-\begin{tabular}{ll}
-00 : (insert\_pre\_xcoord\_v1\_inc-x)   & 07 : (validate\_0)\\
-01 : (insert\_pre\_next\_v1\_v2\_inc-x) & 08 : (apply\_inc-x 0 1)\\
-02 : (insert\_pre\_q0\_inc-x)           & 09 : (apply\_inc-x 1 2)\\
-03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (apply\_inc-x 2 3) \\
-04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (apply\_inc-y-even 0 1)\\
-05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (apply\_dec-x 3 2)\\
-06 : (insert\_add\_xcoord\_v2\_inc-x)   & 13 : (validate\_1)
-\end{tabular}
-}
- \caption{\small A plan for $\tup{\mathcal{M},\mathcal{O}}$ where $\mathcal{M}$ is a modification of the model of Figure \ref{fig:model-example} and $\mathcal{O}$ a partial observation from Figure~\ref{fig:grid-example}.}
-\label{fig:plan-pdistance}
+  \begin{tiny}  
+  \begin{verbatim}
+(:action editable_inc-x
+  :parameters (?v1 ?v2)
+  :precondition
+    (and (or (not (pre_xcoord_v1_inc-x)) (xcoord ?v1))
+         (or (not (pre_xcoord_v2_inc-x)) (xcoord ?v2))
+         (or (not (pre_ycoord_v1_inc-x)) (xcoord ?v1))                       
+         (or (not (pre_ycoord_v2_inc-x)) (xcoord ?v2))
+         (or (not (pre_q0__inc-x)) (q0))
+         (or (not (pre_q1__inc-x)) (q1)))
+         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
+         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
+         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
+         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2))))
+    :effect (and
+       (when (del_xcoord_v1_inc-x) (not (xcoord ?v1)))
+       (when (del_xcoord_v2_inc-x) (not (xcoord ?v2)))
+       (when (del_ycoord_v1_inc-x) (not (xcoord ?v1)))
+       (when (del_ycoord_v2_inc-x) (not (xcoord ?v2)))
+       (when (del_q0__inc-x) (not (q0)))
+       (when (del_q1__inc-x) (not (q1)))
+       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
+       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
+       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
+       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))
+
+       (when (add_xcoord_v1_inc-x) (xcoord ?v1))
+       (when (add_xcoord_v2_inc-x) (xcoord ?v2))
+       (when (add_ycoord_v1_inc-x) (xcoord ?v1))
+       (when (add_ycoord_v2_inc-x) (xcoord ?v2))
+       (when (add_q0__inc-x) (q0))
+       (when (add_q1__inc-x) (q1))
+       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
+       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
+       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
+       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
+  \end{verbatim}           
+  \end{tiny}  
+ \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
+\label{fig:editable}
 \end{figure}
 
-The value of the {\em observation distance} $\delta^o(\mathcal{M},\mathcal{O})$ is given by the number of {\em edit operations} (insertions and deletions) required by $\mathcal{M}$ to be validated in the input observation.
-In the case of the above example, the distance equals $7$.
+\subsection{Learning the least commitment models with conformant planning}
 
 \section{Evaluation}
-\label{sec:evaluation}
+\section{Conclusions}
 
-In this section, we evaluate the empirical performance of our approach in three different applications of {\em model recognition}.
-We will assume in the three experiments that $\mathcal{O}(\tau)$ contains an \emph{empty sequence of observed actions} and so we will only work with the available observed states.
-
-For each experiment, we define a set $M$ of different planning models that share the same state variables but update the variables using different action schemata.
-For every $\mathcal{M}\in M$, we generated the same number of partial observations of plan trajectories ($\mathcal{O}(\tau)$) with such a model.
-Finally, we applied our {\em model recognition as planning} approach to identify the model, among the models in $M$, which was actually used for generating a given $\mathcal{O}(\tau)$ observation.
-We assume equiprobable prior probabilities in all the experiments; i.e., $P(\mathcal{M}) = \frac{1}{|M|}$ for every $\mathcal{M} \in M$.
-
-In order to provide a better explanation of the experiments below, we introduce a particular class of $\mathcal{O}(\tau)$ observations.
-This new class allows us to distinguish between {\em observable} state variables, whose value may be read from sensors, and {\em hidden} or {\em latent} state variables that cannot be observed.
-
-\begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
-\end{definition}
-
-Hence the value of the variables in $\Phi\subseteq F$ is observable while the value of the variables in $F\setminus\Phi$ is unobservable.
-
-\subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar}~\cite{rintanen2014madagascar} because of its ability to deal with classical planning problems with dead-ends~\cite{lopez2015deterministic}.
-We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned the maximum distance obtained.
-The compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em https://github.com/anonsub/model-recognition} so any experimental data reported in the paper can be reproduced.
-
-\subsection{Recognition of {\em regular automata}.} The first experiment, which doubles as a proof of concept, exploits {\em model recognition as planning} for a classical {\em string classification} problem.
-In this experiment, the system receives (1) the string to classify (the $\mathcal{O}(\tau)$ observation) and (2) a set $M$ of different planning models, where each $\mathcal{M}\in M$ represents a different class; i.e.
-a different regular automata that accepts strings that belong to the class.
-
-Figure~\ref{fig:regautomatae} illustrates a 4-symbol and 5-state {\em regular automata} for recognizing the $(abcd)^+$ language.
-The {\em input alphabet} is $\Sigma=\{a,b,c,d\}$, and the machine states are $Q=\{q_0,q_1,q_2,q_3,\underline{q_4}\}$, where \underline{$q_4$} is the only acceptor state.
-For instance, executing the planning model that encodes the {\em regular automata} of Figure~\ref{fig:regautomatae} with the input string $abcdabcd$ produces the following 8-action plan {\small $(\langle a,q_0\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow q_3)$, $(\langle d,q_3\rangle\rightarrow \underline{q_4})$, $(\langle a,\underline{q_4}\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow \underline{q_3})$, $(\langle d,q_3\rangle\rightarrow \underline{q_4})$}.
-
-%\begin{figure}
-%  \begin{scriptsize}
-%  \begin{center}
-%	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-%	  \node[state] (A)              {$q_0$};
-%	  \node[state] (B) [right of=A] {$q_1$};
-%	  \node[state] (C) [right of=B] {$q_2$};
-%	  \node[state] (D) [right of=C] {\underline{$q_3$}};
-%	  \path
-%(A) edge  node {a} (B)
-%(B) edge  node {b} (C)
-%(C) edge  node {c} (D)
-%(D) edge [bend left]  node {a} (B)
-%;
-%	\end{tikzpicture}
-%  \end{center}
-%  \end{scriptsize}
-% \caption{\small A 3-symbol 4-state {\em regular automata} for recognizing the $(abc)^+$ language (\underline{$q_3$} is the acceptor state).}
-%\label{fig:regautomatae}
-%\end{figure}
-
-\begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.8cm,semithick]
-	  \node[state] (A)              {$q_0$};
-	  \node[state] (B) [right of=A] {$q_1$};
-	  \node[state] (C) [right of=B] {$q_2$};
-	  \node[state] (D) [right of=C] {$q_3$};
-	  \node[state] (E) [right of=D] {\underline{$q_4$}};
-	  \path
-(A) edge  node {a} (B)
-(B) edge  node {b} (C)
-(C) edge  node {c} (D)
-(D) edge  node {d} (E)
-(E) edge [bend left]  node {a} (B)
-;
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}
- \caption{\small A 4-symbol and 5-state {\em regular automata} for recognizing the $(abcd)^+$ language (\underline{$q_4$} is the acceptor state).}
-\label{fig:regautomatae}
-\end{figure}
-
-%Note that the the string to classify is a {\em $\Phi$-observation} since it does not contain any information about the structure of the possible regular automata.
-This means that $\mathcal{O}(\tau)$ is a $\Phi$-observation where fluents representing the internal machine state are unknown and the applied transitions (actions) are unobserved.
-
-In this experiment, $M$ comprises five planning models, each representing a different 5-state and 4-symbol regular automata. The five regular languages defined by these automata are the following:
-
-\begin{itemize}
-	\item $\mathcal{L}$1: $a^+(b|c)d(dd)^*a$
-	\item $\mathcal{L}$2: $bd(abd)^*cd^*c^+$
-	\item $\mathcal{L}$3: $d^*c(ac)^*db^*ac^*b^+$
-	\item $\mathcal{L}$4: $(cc)^+bd(abd)^*$
-	\item $\mathcal{L}$5: $(d|a)a(ba)^*c^+d(dc^*d)^*$
-\end{itemize}
-
-For each regular language, we generated $20$ random strings, $\mathcal{O}(\tau)$ observations, which lengths range from 20 to 30 symbols, thus generating a total of 100 strings.
-More precisely, $\mathcal{O}(\tau)$ are observations in which the applied transitions (actions) are unobservable and the fluents of the automata internal state are also unobservable.
-Consequently, $\mathcal{O}(\tau)$ are $\Phi$-observations where the problem states are only represented with the fluents that specify the string to classify.
-
-%More precisely observations $\mathcal{O}(\tau)$ are $\Phi$-observations where: the applied automata transitions and the fluents representing the internal automata state are unobserved.
-
-Table \ref{tab:conf_matrix} shows the {\em confusion matrix} resulting from classifying the 100 strings with our method.
-In this matrix, rows represent the actual class and columns represent the class predicted by our {\em model recognition as planning} approach.
-Despite using a suboptimal classical planner, we can observe that the class for all the 100 input strings was correctly recognized, which proves the feasibility of our method for classification tasks.
-The outstanding results reveal that our approach is very suitable for the classification of generative planning models that are more restrictive than \strips\ models, as it is the case of regular automata.
-
-%These results prove the suitability of {\em model recognition as planning} for classification tasks where classes are given as generative planning models and examples are sequence of observations.
-
-\begin{table}
-	\centering
-	\begin{tabular}{c|c|c|c|c|c}
-		& $\mathcal{L}$1 & $\mathcal{L}$2 & $\mathcal{L}$3 & $\mathcal{L}$4 & $\mathcal{L}$5\\ \hline
-		$\mathcal{L}$1 & 20 & 0 & 0 & 0 & 0 \\
-		$\mathcal{L}$2 & 0 & 20 & 0 & 0 & 0 \\
-		$\mathcal{L}$3 & 0 & 0 & 20 & 0 & 0 \\
-		$\mathcal{L}$4 & 0 & 0 & 0 & 20 & 0 \\
-		$\mathcal{L}$5 & 0 & 0 & 0 & 0 & 20 \\
-	\end{tabular}
-	\caption{Confusion matrix for {\em regular automata recognition}.}
-	\label{tab:conf_matrix}
-\end{table}
-
-\subsection{Recognizing failures in a non-deterministic {\em blocksworld}}
-With this second experiment we aim to validate the two assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood: (1) that agents act rationally and so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation $\mathcal{O}(\tau)$ and (2) that there exists a single such trajectory.
-Our assumptions are more correct with higher observability and become tautological when the trajectory is fully observed ($\mathcal{O(\tau)} = \tau$).
-In order to show the assumptions validation:
-
-% Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions.
-In order to show the assumptions validation:
-
-\begin{itemize}
-\item we will measure the {\em  accuracy} of our approximation of $P(\mathcal{O}|\mathcal{M})$ across different degrees of observability; the accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a fluent in an intermediate state of $\mathcal{O}(\tau)$
-\item we will assume that the length of the trajectory is known so as to keep the possible explanations bounded and force the optimal explanation of $\mathcal{O}(\tau)$; more precisely, at least one fluent of every state of $\mathcal{O}(\tau)$ is observed and the length of the plan trajectory is thus fixed by the observation
-\end{itemize}
-
-The input observations $\mathcal{O}(\tau)$ come from a non-deterministic {\em blocksworld} in which some of the executed actions may have failed. We considered failures of three kinds:
-
-\begin{itemize}
-	\item the execution of a {\tt\small stack} action fails and causes no effect
-    \item the execution of {\tt\small unstack} fails and causes no effect
-	\item {\tt\small unstack} fails and drops the block on the table
-\end{itemize}
-
-$M$ comprises three different 4-schema {\em blocksworld} models, each one extended with an additional action schema that encodes one of the three above failures.
-This means that failures are identified by finding the model that better explains the input observation.
-
-Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations.
-The positive trend seen in the figure proves that the more complete the observations are the more accurate our estimation becomes.
-
-%The overall high accuracy validates our assumption that the trajectory that better explains the observation is the shortest one.
-
-%The increasing trend of the accuracy as observability increases and so the number of possible explanation lowers, validates our two assumptions.
-
-%We limited the failures to be of one kind per observation however, it would be possible to recognize different combinations of them.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{figures/blocks.eps}
-	\caption{Accuracy for the recognition of failures in a non-deterministic {\em blocksworld}.}
-	\label{fig:blocks_acc}
-\end{figure}
-
-\subsection{Recognition of navigation policies}
-%This experiment is also aimed at validating the accuracy of our approach to identify the model that generates an observation under different degrees of observability.
-The novel aspect of this experiment is that, unlike the previous experiment, we do not guarantee that the observed states contain at least one fluent.
-This means that some intermediate states of $\mathcal{O}(\tau)$ may be missing (all fluents are unobservable) and, consequently, more than a single action may be required to reach an observed state from another one.
-
-In this experiment, the planning models of $M$ represent different navigation models for $N\times N$ grids.
-In particular, we adopt a {\em classical navigation model} with actions {\tt\small up, down, right, left} (to move one cell in each of these four directions) and extend it to define 8 different navigation policies with respect to these additional state variables {\tt\small \{(q0), (q1), (min ?v), (max ?v)\}}.
-An example of a navigation policy is shown in Figure~\ref{fig:model-example}.
-This policy allows to move right when $q_0$ holds while it allows to move left when $q_1$ holds, producing a zigzag pattern when visiting all the cells of $N\times N$ grids.
-
-Another interesting aspect of this experiment is that all navigation policies are at a maximum distance of four editions from the base {\em classical navigation model} that can move the robot in any direction at any given state.
-Thus, for any given model and observation, a maximum of four editions are needed for the input model to explain the observation.
-This aspect heavily constrains the discriminating power of $P(\mathcal{O}|\mathcal{M})$, which along with the low degree of observability of the states, gives rise to a meaningful benchmark for {\em model recognition as planning}.
-
-On the other hand, we generated observations from 8 different trajectories, one for each of the 8 previous planning models.
-The trajectories depict paths followed by a navigation policy to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
-As in the previous experiments, observations contain no actions and here, for each observed state, only the values of the fluents encoding the {\em x} and {\em y coordinates} of the agent are known (i.e.
-$\Phi =$\{{\small\tt xcoord, ycoord}\}).
-
-Figure \ref{fig:navigation_acc} shows the classification {\em accuracy} achieved by our approach with respect to a range of degrees of observability, from 0\% to 100\% with 10\% increments.
-In this experiment we included the 0\% case which corresponds to observations where only the initial and final states are observed.
-The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation policies, but from 10\% of observability onwards, we start to correctly classify half of the observations.
-Accuracy stabilizes after 40\% observability in the range 0.875 to 1 which means that at most only one out of the 8 observations was not correctly classified.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{figures/navigation.eps}
-	\caption{Classification accuracy for the recognition of navigation models.}
-	\label{fig:navigation_acc}
-\end{figure}
-
-\section{Related work}
-\label{sec:related}
-An interesting research line related with {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}.
-In {\em model reconciliation} model edition is used to conform the PDDL models of two different agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed {\em optimal plan} computed using one of the models.
-{\em Model recognition} conforms every input model $\mathcal{M}$ in the $M$ set with another model $\mathcal{M}'$ that is consistent with a partial observation of a plan execution.
-Note, however, that in {\em model recognition} $\mathcal{M}'$ is not given as input but instead is computed for every $\mathcal{M}\in M$.
-Further, {\em model recognition as planning} shows that the probability of an input planning model to explain a partial observation of a plan execution (that contains no information of the applied actions and where intermediate states may be incomplete or even missing) can be estimated with an off-the-shelf classical planner.
-
-Once the planning model of the observed agent is recognized, then the model-based machinery for automated planning becomes fully applicable.
-This includes current approaches for {\em goal recognition}~\cite{ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} and {\em counter-planning}~\cite{PozancoEFB18} that require to have in advance a fixed model of the observed agent.
-Last but not least, our approach allows predicting the future actions of an agent whose behavior is not governed by a deliberative planning process aimed to achieve some goals.
-Our experiments show that {\em model recognition} is suitable for anticipating the actions of agents that are ruled by the execution of fixed programs or finite state machines, such as a {\em regular automata}.
-
-%Last but not least, previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
-Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
-
-\section{Conclusions}
-\label{sec:conclussions}
-This paper formalized the {\em model recognition} task and proposed, {\em model recognition as planning}, a method built on top of off-the-shelf classical planning algorithms to estimate the probability of a \strips\ model to produce a partial observation of a plan execution.
-The paper shows the effectiveness of {\em model recognition as planning} in a set of \strips\ models encoding different kinds of {\em automata}.
-{\em Model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unobserved.
-
-%Remarkably, the extension of this piece of work to the FOND planning setting~\cite{muise2012improved} is straightforward by simply considering the {\em all-outcomes} determiniztion of the actions with non-determinitic effects~\cite{yoon2007ff}.
-An interesting research direction is however to understand how to apply our approach to planning models where the planning models include actions with probabilistic effects~\cite{younes2005first}.
-
-\newpage
-
-\bibliographystyle{aaai}
-\bibliography{planlearnbibliography}
+%% The file named.bst is a bibliography style file for BibTeX 0.99c
+\bibliographystyle{named}
+\bibliography{planlearnbibliography.bib}
 

### 2018-11-26 doc/ijcai-19-conformant/conformant-learning.tex
added conformant planning model

@@
 
-\title{Computing the least commitment action models from observations of plan executions}
+\title{Computing {\em the least commitment} action model from observations of plan executions}
 
@@
 \section{Introduction}
+Given the partial observation of a plan execution, this paper formalizes the task of computing the exact set of action models that are {\em conformant} with the given observation.
+This task is of interest because it allows to scallably learn action models from arbitrary large sets of observations generated with that models.
+The approach we follow for addressing this learning task is compiling it into {\em conformant planning}.
 
 \section{Background}
-This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
+This section formalizes the {\em classical} and the {\em conformant} planning models as well as the {\em observation} model we follow to observe the execution of sequential plans.
 \subsection{Classical planning with conditional effects}
@@
 
-A {\em planning problem} is defined as a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state in which all the fluents of $F$ are assigned a value true/false and $G$ is the goal set.
+A {\em planning problem} is defined as a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state in which all the fluents of $F$ are assigned a value {\em true\textbar false} and $G$ is the set of goal conditions over the state variables in $F$.
 A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, and $|\pi|=n$ denotes its {\em plan length}.
-The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
-A trajectory $\tau(\pi,P)$ that solves $P$ is one in which $G \subseteq s_n$.
+The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$.
 
@@
 \subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
 Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
-
-%which indicates that we observe $l$ actions, $1\leq l\leq |\pi|$, and $m$ states, $1\leq m \leq |\pi|+1$, from $\tau(\pi,P)$:
-
 \begin{itemize}
@@
 Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
-Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,s_0)$, meaning that $\forall i, s_i^o \subseteq s_i$.
 In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
@@
 \subsection{Conformant planning}
+{\em Conformant planning} refers to planning with incomplete information about the initial situation, no sensing, and where goals are to be achieved with certainty in spite of the uncertainty in the initial situation~\cite{smith1998conformant,goldman1996expressive}.
 
-\section{Learning the least commitment action models from observations}
+Syntactically, conformant planning problems are expressed in compact form through a set of state variables.
+A {\em conformant planning problem} is defined as a tuple $P_c=\tup{F,A,\Upsilon,G}$, where $\Upsilon$ is a set of clauses over literals $l=f$ or $l=\neg f$ (for $f\in F$) that define the set of possible initial states.
 
-\subsection{The learning task}
+A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state.
+More precisely an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0$ is a valuation that satisfies $\Upsilon$ and it holds that the trajectory reaches $G \subseteq s_n$.
+
+\section{Learning {\em the least commitment} action model from observations}
 
@@
 
-\begin{itemize}
+\begin{enumerate}
 \item {\em Syntactic constraints}.
@@
 For instance, in a {\em blocksworld} one can claim that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in any of these lists because of the semantic of the {\tt\small on} predicate that codes which block is in top of other block.
-\end{itemize}
+\end{enumerate}
 
@@
 
-\subsection{A propositional encoding for \strips\ action schema}
+\begin{enumerate}
+\item[3.] {\em Observation constraints}.
+An observation of the execution of a plan further constraints the space of possible action schemata.
+This \emph{semantic knowledge} included in the observations introduce a third type of constraints, that we will call {\em observation constraints}, and that can be added to the set $\mathcal{C}$.
+\end{enumerate}
+
+\subsection{The learning task}
+
+\subsection{Learning the least commitment models with conformant planning}
 Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
@@
 
-\subsection{Learning the least commitment models with conformant planning}
+\section{Evaluation}
 
-\section{Evaluation}
 \section{Conclusions}

### 2018-11-26 doc/ijcai-19-conformant/conformant-learning.tex
added first definition for least commitment action model

@@
 A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state.
-More precisely an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0$ is a valuation that satisfies $\Upsilon$ and it holds that the trajectory reaches $G \subseteq s_n$.
+More precisely an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$ and it holds that the trajectory reaches $G \subseteq s_n$.
 
@@
 \subsection{The hypothesis space}
-\strips\ action schemata provide a compact representation for specifying classical planning models.
+\strips\ action schemata provide a compact representation specifying action models.
 {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
@@
 
-For instance, the {\small\tt stack(?v1,?v2)} and {\small\tt unsstack(?v1,?v2)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} are not.
-Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
+For instance, the {\small\tt stack(?v1,?v2)} and {\small\tt unstack(?v1,?v2)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} are not.
+Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
 
@@
 One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in a {\em blocksworld} one can claim that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in any of these lists because of the semantic of the {\tt\small on} predicate that codes which block is in top of other block.
+For instance, in the {\em blocksworld} one can claim that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of an action schema $\xi$ because of the semantic of the {\tt\small on} predicate (i.e.
+only one block can be on top of another block).
 \end{enumerate}
@@
 An observation of the execution of a plan further constraints the space of possible action schemata.
-This \emph{semantic knowledge} included in the observations introduce a third type of constraints, that we will call {\em observation constraints}, and that can be added to the set $\mathcal{C}$.
+This \emph{semantic knowledge} included in the observations introduce a third type of constraints and that can also be added to the set $\mathcal{C}$.
 \end{enumerate}
 
-\subsection{The learning task}
+\subsection{The least commitment action model}
+Now we are ready to formalize the learning task that we are addressing in this paper. The task of computing {\em the least commitment} action model is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:
+
+\begin{itemize}
+\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions.
+For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
+the functions $\rho$ and/or $\theta$ of $a$ are undefined.
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
+\item $M$ is the set of different planning models for the actions in $A[\cdot]$.
+A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$.
+Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
+\end{itemize}
+
+The {\em solution} to this task is {\em the least commitment} action model which is defined as the minimum subset of models $M^*\in M$ that is compliant with the input observation $\mathcal{O}(\tau)$.
 

### 2018-11-26 doc/ijcai-19-conformant/conformant-learning.tex
removing actions from the observation model

@@
 \subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
-Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
-\begin{itemize}
-\item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
-Specifically, the number of observed actions, $l$, can range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
-\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as sequence of partial states that represents the observation from the execution of $\pi$ in $P$.
+Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$ and $s_0^o=I$.
+
+More precisely $\mathcal{O}(\tau)$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
 A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
@@
 In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
-    %Exceptionally, $s_m^o$ cannot be fully unobservable for the purpose of our task (we will elaborate on this issue later on).
-\end{itemize}
 
@@
 \section{Learning {\em the least commitment} action model from observations}
+
+\subsection{The {\em least commitment} action model}
+Now we are ready to formalize the learning task that we are addressing in this paper. The task of computing {\em the least commitment} action model is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:
+
+\begin{itemize}
+\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions.
+For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
+the functions $\rho$ and/or $\theta$ of $a$ are undefined.
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
+\item $M$ is the set of different planning models for the actions in $A[\cdot]$.
+A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$.
+Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
+\end{itemize}
+
+The {\em solution} to this task is {\em the least commitment} action model which is defined as the minimum subset of models $M^*\in M$ that is compliant with the input observation $\mathcal{O}(\tau)$.
 
@@
 
-\subsection{The least commitment action model}
-Now we are ready to formalize the learning task that we are addressing in this paper. The task of computing {\em the least commitment} action model is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:
-
-\begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions.
-For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined.
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
-\item $M$ is the set of different planning models for the actions in $A[\cdot]$.
-A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$.
-Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
-\end{itemize}
-
-The {\em solution} to this task is {\em the least commitment} action model which is defined as the minimum subset of models $M^*\in M$ that is compliant with the input observation $\mathcal{O}(\tau)$.
-
-\subsection{Learning the least commitment models with conformant planning}
+\subsection{Learning {\em the least commitment} model with conformant planning}
 Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
@@
 
+The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated.
+The new goals $G'=G\bigcup\{at_n,test_m\}$ and the new actions for {\em validating} the edited models at the $s_j$ observed states, {\tt\small $0\leq j< m$}.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\\
+                            &\{mode_{edit}\}\rhd\{\neg mode_{edit}, mode_{val}\}.
+\end{align*}
+\end{small}
+
 \section{Evaluation}

### 2018-11-26 doc/ijcai-19-conformant/conformant-learning.tex
added formalization for conformant planning

@@
 
-\title{Computing {\em the least commitment} action model from observations of plan executions}
+\title{Computing {\em the least commitment} action model from state observations}
 
@@
 \section{Introduction}
-Given the partial observation of a plan execution, this paper formalizes the task of computing the exact set of action models that are {\em conformant} with the given observation.
-This task is of interest because it allows to scallably learn action models from arbitrary large sets of observations generated with that models.
-The approach we follow for addressing this learning task is compiling it into {\em conformant planning}.
+Given a sequence of partially observed states, this paper formalizes the task of computing the exact set of action models that are {\em conformant} with the given observation.
+This task is of interest because it allows to, incrementally and scalably, learn action models from arbitrary large sets of state observations.
+
+In addition, the paper presets a new method to compute {\em the least commitment} action model from state observations that compiles this task into {\em conformant planning}.
 
 \section{Background}
-This section formalizes the {\em classical} and the {\em conformant} planning models as well as the {\em observation} model we follow to observe the execution of sequential plans.
+This section formalizes the {\em classical} and {\em conformant} planning models as well as the kind of input observations for the computation of {\em the least commitment} action model.
+
 \subsection{Classical planning with conditional effects}
@@
 A {\em state} $s$ is a full assignment of values to fluents.
-We explicitly include negative literals $\neg f$ in states and so $|s|=|F|$ and the size of the state space is $2^{|F|}$.
-
-%A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states; i.e.
-$|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Like in PDDL~\cite{fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+We explicitly include negative literals $\neg f$ in states s.t.
+$|s|=|F|$ and the size of the state space is $2^{|F|}$.
 
@@
 
-A {\em planning problem} is defined as a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state in which all the fluents of $F$ are assigned a value {\em true\textbar false} and $G$ is the set of goal conditions over the state variables in $F$.
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state in which all the fluents of $F$ are assigned a {\em true\textbar false} value and $G\in\mathcal{L}(F)$ is the set of goal conditions over the state variables.
 A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, and $|\pi|=n$ denotes its {\em plan length}.
@@
 \subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as sequence of partial states that represents the observation from the execution of $\pi$ in $P$.
-Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$ and $s_0^o=I$.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as sequence of partial states that represents the observation from the execution of $\pi$ in $P$.
+Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$ where $s_0^o=I$.
 
-More precisely $\mathcal{O}(\tau)$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
 A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
@@
 \subsection{Conformant planning}
-{\em Conformant planning} refers to planning with incomplete information about the initial situation, no sensing, and where goals are to be achieved with certainty in spite of the uncertainty in the initial situation~\cite{smith1998conformant,goldman1996expressive}.
+{\em Conformant planning} refers to planning with incomplete information about the initial state, no sensing, and where goals have to be achieved with certainty (despite the uncertainty of the initial state)~\cite{smith1998conformant,goldman1996expressive}.
 
 Syntactically, conformant planning problems are expressed in compact form through a set of state variables.
-A {\em conformant planning problem} is defined as a tuple $P_c=\tup{F,A,\Upsilon,G}$, where $\Upsilon$ is a set of clauses over literals $l=f$ or $l=\neg f$ (for $f\in F$) that define the set of possible initial states.
+A {\em conformant planning problem} is then defined as a tuple $P_c=\tup{F,A,\Upsilon,G}$ where $F$, $A$ and $G$ are the set of fluents, actions and goals (as previously defined for the classical planing model).
+Now $\Upsilon$ is a set of clauses over literals $l=f$ or $l=\neg f$ (for $f\in F$) that define the set of possible initial states.
 
 A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state.
-More precisely an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$ and it holds that the trajectory reaches $G \subseteq s_n$.
+More precisely, an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$, such that $s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$, then the $\tau(\pi,s_0)$ trajectory reaches a final state $G \subseteq s_n$.
 

### 2018-11-26 doc/ijcai-19-conformant/conformant-learning.tex
first conformant compilation version ijcai19 paper

@@
 
-In addition, the paper presets a new method to compute {\em the least commitment} action model from state observations that compiles this task into {\em conformant planning}.
+In addition, the paper presets a new method to compute {\em the least commitment} action model from state observations.
+The method assumes that action models are specified as \strips\ action schema and builds on top of off-the-shelf {\em conformant planning} algorithms.
 
@@
 
-\section{Learning {\em the least commitment} action model from observations}
+\section{Learning {\em the least commitment} action model from state observations}
+Now we are ready to formalize the learning of the {\em the least commitment} action model from state observations and our approach to address it via {\em conformant planning}.
 
 \subsection{The {\em least commitment} action model}
-Now we are ready to formalize the learning task that we are addressing in this paper. The task of computing {\em the least commitment} action model is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:
+The task of computing {\em the least commitment} action model is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:
 
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions.
+\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a planning problem where $A[\cdot]$ is a set of actions.
 For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined.
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
+The set of goal conditions $G[\cdot]$ is also unknown.
 \item $M$ is the set of different planning models for the actions in $A[\cdot]$.
@@
 Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
+\item $\mathcal{O}(\tau)$ is a sequence of state observations coming from the observation of a trajectory $\tau(\pi,s_0)$ produced by the execution of certain unknown plan $\pi$.
 \end{itemize}
 
-The {\em solution} to this task is {\em the least commitment} action model which is defined as the minimum subset of models $M^*\in M$ that is compliant with the input observation $\mathcal{O}(\tau)$.
+The {\em solution} to this task is {\em the least commitment} action model which defines the smallest subset of models $M^*\subseteq M$ that is {\em conformant} with the input observation $\mathcal{O}(\tau)$.
 
-\subsection{The hypothesis space}
-\strips\ action schemata provide a compact representation specifying action models.
+\subsection{The space of possible action models}
+This work focuses on the lerning of action models that are specified as \strips\ action schema.
+\strips\ action schemata provide a compact representation for specifying action models.
 {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
@@
 One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in the {\em blocksworld} one can claim that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of an action schema $\xi$ because of the semantic of the {\tt\small on} predicate (i.e.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because of the semantic of the {\tt\small on} predicate (i.e.
 only one block can be on top of another block).
@@
 \item[3.] {\em Observation constraints}.
-An observation of the execution of a plan further constraints the space of possible action schemata.
-This \emph{semantic knowledge} included in the observations introduce a third type of constraints and that can also be added to the set $\mathcal{C}$.
+A sequence of state observations $\mathcal{O}(\tau)$ constraints further the space of possible action schemata.
+This \emph{semantic knowledge} included in the observations introduce a third type of constraints and can also be added to the set $\mathcal{C}$.
 \end{enumerate}
@@
 \subsection{Learning {\em the least commitment} model with conformant planning}
-Given a \strips\ action schema $\xi$, a propositional encoding for the {\em preconditions}, {\em negative} and {\em positive} effects of that schema can be represented with fluents of the kind $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of interpretations of predicates $\Psi$ over the corresponding variable names $\Omega_\xi$.
-Figure~\ref{fig:encoding} shows the propositional encoding for the six action schema defined in Figure~\ref{fig:model-example}.
+Given a classical planning problem $P=\tup{F,A[\cdot],I,G[\cdot]}$ and a sequence of state observations $\mathcal{O}(\tau)$ coming from the observation of a $\tau(\pi,I)$ trajectory, we show here that we can build a {\em conformant planning problem} $P_c$ in linear time and space and whose solution is a plan that induces the {\em least commitment} model that is conformant with $\mathcal{O}(\tau)$.
 
-The interest of having a propositional encoding for \strips\ action schema is that, using {\em conditional effects}, it allows to compactly define {\em editable actions}.
-Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state.
+In more detail, our compilation defines a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G}$ such that:
+\begin{itemize}
+\item The set of fluents $F_c$ extends $F$ with two new sets of fluents:
+      \begin{itemize}
+      \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
+      \item The fluents $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ for a propositional encoding of the {\em preconditions}, {\em negative} and {\em positive} effects of schema $\xi$.
+Figure~\ref{fig:propositional} shows the propositional encoding of the unstack action from the blocksworld.
+      \end {itemize}
+
+\begin{figure}
+  \begin{tiny}  
+  \begin{verbatim}
+(pre_holding_v1_stack) (pre_clear_v2_stack)
+(del_holding_v1_stack) (del_clear_v2_stack)
+(add_clear_v1_stack) (add_handempty_stack) (add_on_v1_v2_stack)
+  \end{verbatim}           
+  \end{tiny}  
+ \caption{\small Propositional encoding of the {\tt\small stack(?v1,?v2)} schema.}
+\label{fig:propositional}
+\end{figure}
+
+      \item The set of actions $A_c$ contains now actions of two different kinds:
+\begin{itemize}
+      \item Actions for {\em validating} the edited models at the $s_j$ observed states, {\tt\small $0\leq j< m$}
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\\
+                            &\{mode_{edit}\}\rhd\{\neg mode_{edit}, mode_{val}\}.
+\end{align*}
+\end{small}
+
+      \item Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state.
 Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
@@
 \end{small}
-
 Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}).
 Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
+      \end{itemize}
+
+\item The new goals are $G_c=\{test_m\}$ .
+\end{itemize}
 
@@
   \begin{verbatim}
-(:action editable_inc-x
-  :parameters (?v1 ?v2)
+(:action stack
+  :parameters (?o1 - object ?o2 - object)
   :precondition
-    (and (or (not (pre_xcoord_v1_inc-x)) (xcoord ?v1))
-         (or (not (pre_xcoord_v2_inc-x)) (xcoord ?v2))
-         (or (not (pre_ycoord_v1_inc-x)) (xcoord ?v1))                       
-         (or (not (pre_ycoord_v2_inc-x)) (xcoord ?v2))
-         (or (not (pre_q0__inc-x)) (q0))
-         (or (not (pre_q1__inc-x)) (q1)))
-         (or (not (pre_next_v1_v1_inc-x)) (next ?v1 ?v1)))
-         (or (not (pre_next_v1_v2_inc-x)) (next ?v1 ?v2)))
-         (or (not (pre_next_v2_v1_inc-x)) (next ?v2 ?v1)))
-         (or (not (pre_next_v2_v2_inc-x)) (next ?v2 ?v2))))
-    :effect (and
-       (when (del_xcoord_v1_inc-x) (not (xcoord ?v1)))
-       (when (del_xcoord_v2_inc-x) (not (xcoord ?v2)))
-       (when (del_ycoord_v1_inc-x) (not (xcoord ?v1)))
-       (when (del_ycoord_v2_inc-x) (not (xcoord ?v2)))
-       (when (del_q0__inc-x) (not (q0)))
-       (when (del_q1__inc-x) (not (q1)))
-       (when (del_next_v1_v1_inc-x) (not (next ?v1 ?v1)))
-       (when (del_next_v1_v2_inc-x) (not (next ?v1 ?v2)))
-       (when (del_next_v2_v1_inc-x) (not (next ?v2 ?v1)))
-       (when (del_next_v2_v2_inc-x) (not (next ?v2 ?v2)))
-
-       (when (add_xcoord_v1_inc-x) (xcoord ?v1))
-       (when (add_xcoord_v2_inc-x) (xcoord ?v2))
-       (when (add_ycoord_v1_inc-x) (xcoord ?v1))
-       (when (add_ycoord_v2_inc-x) (xcoord ?v2))
-       (when (add_q0__inc-x) (q0))
-       (when (add_q1__inc-x) (q1))
-       (when (add_next_v1_v1_inc-x) (next ?v1 ?v1))
-       (when (add_next_v1_v2_inc-x) (next ?v1 ?v2))
-       (when (add_next_v2_v1_inc-x) (next ?v2 ?v1))
-       (when (add_next_v2_v2_inc-x) (next ?v2 ?v2)))
+   (and (or (not (pre_on_v1_v1_stack)) (on ?o1 ?o1))
+        (or (not (pre_on_v1_v2_stack)) (on ?o1 ?o2))
+        (or (not (pre_on_v2_v1_stack)) (on ?o2 ?o1))
+        (or (not (pre_on_v2_v2_stack)) (on ?o2 ?o2))
+        (or (not (pre_ontable_v1_stack)) (ontable ?o1))
+        (or (not (pre_ontable_v2_stack)) (ontable ?o2))
+        (or (not (pre_clear_v1_stack)) (clear ?o1))
+        (or (not (pre_clear_v2_stack)) (clear ?o2))
+        (or (not (pre_holding_v1_stack)) (holding ?o1))
+        (or (not (pre_holding_v2_stack)) (holding ?o2))
+        (or (not (pre_handempty_stack)) (handempty)))
+  :effect
+   (and (when (del_on_v1_v1_stack) (not (on ?o1 ?o1)))
+        (when (del_on_v1_v2_stack) (not (on ?o1 ?o2)))
+        (when (del_on_v2_v1_stack) (not (on ?o2 ?o1)))
+        (when (del_on_v2_v2_stack) (not (on ?o2 ?o2)))
+        (when (del_ontable_v1_stack) (not (ontable ?o1)))
+        (when (del_ontable_v2_stack) (not (ontable ?o2)))
+        (when (del_clear_v1_stack) (not (clear ?o1)))
+        (when (del_clear_v2_stack) (not (clear ?o2)))
+        (when (del_holding_v1_stack) (not (holding ?o1)))
+        (when (del_holding_v2_stack) (not (holding ?o2)))
+        (when (del_handempty_stack) (not (handempty)))
+        (when (add_on_v1_v1_stack) (on ?o1 ?o1))
+        (when (add_on_v1_v2_stack) (on ?o1 ?o2))
+        (when (add_on_v2_v1_stack) (on ?o2 ?o1))
+        (when (add_on_v2_v2_stack) (on ?o2 ?o2))
+        (when (add_ontable_v1_stack) (ontable ?o1))
+        (when (add_ontable_v2_stack) (ontable ?o2))
+        (when (add_clear_v1_stack) (clear ?o1))
+        (when (add_clear_v2_stack) (clear ?o2))
+        (when (add_holding_v1_stack) (holding ?o1))
+        (when (add_holding_v2_stack) (holding ?o2))
+        (when (add_handempty_stack) (handempty))))
   \end{verbatim}           
   \end{tiny}  
- \caption{\small Editable version of the {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid.}
+ \caption{\small Editable version of the {\tt\small stack(?v1,?v2)} schema.}
 \label{fig:editable}
@@
 
-The fluents $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated.
-The new goals $G'=G\bigcup\{at_n,test_m\}$ and the new actions for {\em validating} the edited models at the $s_j$ observed states, {\tt\small $0\leq j< m$}.
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\\
-                            &\{mode_{edit}\}\rhd\{\neg mode_{edit}, mode_{val}\}.
-\end{align*}
-\end{small}
+Given a plan $\pi$ that solves the {\em conformant planning} problem $P_c$ that results from our compilation then, the {\em least commitment} model that is conformant with $\mathcal{O}(\tau)$ is extracted in linear time from the last state reached by $\pi$.
+
+\subsection{Compilation properties}
 

### 2018-11-27 doc/ijcai-19-conformant/conformant-learning.tex
first version of commit actions

@@
 only one block can be on top of another block).
+Invariants constraining the space of possible states in a given domain belong also to this group~\cite{fox1998automatic}.
 \end{enumerate}
@@
   \begin{verbatim}
+(:action stack
+   :parameters (?x ?y)
+   :precondition (and (holding ?x) (clear ?y))
+   :effect (and (not (holding ?x)) (not (clear ?y))
+                (clear ?x) (handempty) (on ?x ?y)))
+
 (pre_holding_v1_stack) (pre_clear_v2_stack)
-(del_holding_v1_stack) (del_clear_v2_stack)
-(add_clear_v1_stack) (add_handempty_stack) (add_on_v1_v2_stack)
+(eff_holding_v1_stack) (eff_clear_v2_stack)
+(eff_clear_v1_stack) (eff_handempty_stack) (eff_on_v1_v2_stack)
   \end{verbatim}           
   \end{tiny}  
- \caption{\small Propositional encoding of the {\tt\small stack(?v1,?v2)} schema.}
+ \caption{\small PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same action.}
 \label{fig:propositional}
@@
 
-      \item The set of actions $A_c$ contains now actions of two different kinds:
+      \item The set of actions $A_c$ contains now actions of three different kinds:
 \begin{itemize}
+      \item Actions for {\em committing} {\small $pre\_e\_\xi$} fluents to a positive and a negative value.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{set\top\_pre\_e\_\xi})=&\{\},\\
+\cond(\mathsf{set\top\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{pre\_e\_\xi\},\\
+                                    &\{\neg pre\_e\_\xi\}\rhd\{pre\_e\_\xi\}.\\
+\hspace*{7pt}\pre(\mathsf{set\bot\_pre\_e\_\xi})=&\{\},\\
+\cond(\mathsf{set\bot\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{\neg pre\_e\_\xi\},\\
+                                    &\{\neg pre\_e\_\xi\}\rhd\{\neg pre\_e\_\xi\}.                                    
+\end{align*}
+\end{small}
+In addition, similar actions are defined for {\em committing} {\small $eff\_e\_\xi$} fluents to a positive value and for {\em committing} {\small $pre\_e\_\xi$} fluents to a negative value.
+
       \item Actions for {\em validating} the edited models at the $s_j$ observed states, {\tt\small $0\leq j< m$}
@@
 
-      \item Actions whose semantics is given by the value of the $[pre|del|add]\_e\_\xi$ fluents at the current state.
+      \item Actions whose semantics is given by the value of the {\small $pre\_e\_\xi$}, {\small $pre\_e\_\xi$} fluents at the current state.
 Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
@@
 \hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
-\cond(\mathsf{editable_{\xi}})=&\{del\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
-&\{add\_e\_\xi\}\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
+\cond(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi, eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+&\{\neg pre\_e\_\xi, eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
 \end{align*}
 \end{small}
-Figure~\ref{fig:editable} shows the PDDL encoding of the {\em editable} {\tt\small inc-x(?v1,?v2)} schema for robot navigation in a $n\times n$ grid (Figure~\ref{fig:model-example}).
-Note that this editable schema, when the fluents of Figure~\ref{fig:encoding} hold, behaves exactly as defined in Figure~\ref{fig:model-example}.
+Figure~\ref{fig:editable} shows the PDDL encoding of the editable version of the {\tt\small stack(?v1,?v2)} schema.
+Note that this editable schema, when the fluents {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)} hold, it behaves exactly as defined in Figure~\ref{fig:propositional}.
 \end{itemize}
-
 \item The new goals are $G_c=\{test_m\}$ .
@@
         (or (not (pre_handempty_stack)) (handempty)))
-  :effect
-   (and (when (del_on_v1_v1_stack) (not (on ?o1 ?o1)))
-        (when (del_on_v1_v2_stack) (not (on ?o1 ?o2)))
-        (when (del_on_v2_v1_stack) (not (on ?o2 ?o1)))
-        (when (del_on_v2_v2_stack) (not (on ?o2 ?o2)))
-        (when (del_ontable_v1_stack) (not (ontable ?o1)))
-        (when (del_ontable_v2_stack) (not (ontable ?o2)))
-        (when (del_clear_v1_stack) (not (clear ?o1)))
-        (when (del_clear_v2_stack) (not (clear ?o2)))
-        (when (del_holding_v1_stack) (not (holding ?o1)))
-        (when (del_holding_v2_stack) (not (holding ?o2)))
-        (when (del_handempty_stack) (not (handempty)))
-        (when (add_on_v1_v1_stack) (on ?o1 ?o1))
-        (when (add_on_v1_v2_stack) (on ?o1 ?o2))
-        (when (add_on_v2_v1_stack) (on ?o2 ?o1))
-        (when (add_on_v2_v2_stack) (on ?o2 ?o2))
-        (when (add_ontable_v1_stack) (ontable ?o1))
-        (when (add_ontable_v2_stack) (ontable ?o2))
-        (when (add_clear_v1_stack) (clear ?o1))
-        (when (add_clear_v2_stack) (clear ?o2))
-        (when (add_holding_v1_stack) (holding ?o1))
-        (when (add_holding_v2_stack) (holding ?o2))
-        (when (add_handempty_stack) (handempty))))
+ :effect (and
+   (when (and (pre_on_v1_v1_stack)(eff_on_v1_v1_stack)) (not (on ?o1 ?o1)))
+   (when (and (pre_on_v1_v2_stack)(eff_on_v1_v2_stack)) (not (on ?o1 ?o2)))
+   (when (and (pre_on_v2_v1_stack)(eff_on_v2_v1_stack)) (not (on ?o2 ?o1)))
+   (when (and (pre_on_v2_v2_stack)(eff_on_v2_v2_stack)) (not (on ?o2 ?o2)))
+   (when (and (pre_ontable_v1_stack)(eff_ontable_v1_stack)) (not (ontable ?o1)))
+   (when (and (pre_ontable_v2_stack)(eff_ontable_v2_stack)) (not (ontable ?o2)))
+   (when (and (pre_clear_v1_stack)(eff_clear_v1_stack)) (not (clear ?o1)))
+   (when (and (pre_clear_v2_stack)(eff_clear_v2_stack)) (not (clear ?o2)))
+   (when (and (pre_holding_v1_stack)(eff_holding_v1_stack)) (not (holding ?o1)))
+   (when (and (pre_holding_v2_stack)(eff_holding_v2_stack)) (not (holding ?o2)))
+   (when (and (pre_handempty_stack)(eff_handempty_stack)) (not (handempty)))
+   (when (and (not(pre_on_v1_v1_stack))(eff_on_v1_v1_stack)) (on ?o1 ?o1))
+   (when (and (not(pre_on_v1_v2_stack))(eff_on_v1_v2_stack)) (on ?o1 ?o2))
+   (when (and (not(pre_on_v2_v1_stack))(eff_on_v2_v1_stack)) (on ?o2 ?o1))
+   (when (and (not(pre_on_v2_v2_stack))(eff_on_v2_v2_stack)) (on ?o2 ?o2))
+   (when (and (not(pre_ontable_v1_stack))(eff_ontable_v1_stack)) (ontable ?o1))
+   (when (and (not(pre_ontable_v2_stack))(eff_ontable_v2_stack)) (ontable ?o2))
+   (when (and (not(pre_clear_v1_stack))(eff_clear_v1_stack)) (clear ?o1))
+   (when (and (not(pre_clear_v2_stack))(eff_clear_v2_stack)) (clear ?o2))
+   (when (and (not(pre_holding_v1_stack))(eff_holding_v1_stack)) (holding ?o1))
+   (when (and (not(pre_holding_v2_stack))(eff_holding_v2_stack)) (holding ?o2))
+   (when (and (not(pre_handempty_stack))(eff_handempty_stack)) (handempty))))
   \end{verbatim}           
   \end{tiny}  
- \caption{\small Editable version of the {\tt\small stack(?v1,?v2)} schema.}
+ \caption{\small PDDL encoding of the editable version of the {\tt\small stack(?v1,?v2)} schema.}
 \label{fig:editable}

### 2018-11-27 doc/ijcai-19-conformant/conformant-learning.tex
added definition for the explanation of an observation, ijcai19 paper

@@
 \section{Introduction}
-Given a sequence of partially observed states, this paper formalizes the task of computing the exact set of action models that are {\em conformant} with the given observation.
-This task is of interest because it allows to, incrementally and scalably, learn action models from arbitrary large sets of state observations.
+Given a sequence of partially observed states, this paper formalizes the task of computing {\em the least commitment} action model that can explain the given observation.
+This task is of interest because it allows the incremental learning of action models from arbitrary large sets of state observations.
 
-In addition, the paper presets a new method to compute {\em the least commitment} action model from state observations.
-The method assumes that action models are specified as \strips\ action schema and builds on top of off-the-shelf {\em conformant planning} algorithms.
+In addition, the paper introduces a new method to compute {\em the least commitment} action model from a sequence of state observations.
+The method assumes that action models are specified as \strips\ action schema and it builds on top of off-the-shelf algorithms for {\em conformant planning}.
 
 \section{Background}
-This section formalizes the {\em classical} and {\em conformant} planning models as well as the kind of input observations for the computation of {\em the least commitment} action model.
+This section formalizes the model of {\em classical planning with conditional effects}, the {\em conformant planning} model, and the kind of input observations for the computation of {\em the least commitment} action model.
 
 \subsection{Classical planning with conditional effects}
-$F$ is the set of {\em fluents} or {\em state variables} (propositional variables).
+Let $F$ be the set of {\em fluents} or {\em state variables} (propositional variables).
 A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
@@
 
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state in which all the fluents of $F$ are assigned a {\em true\textbar false} value and $G\in\mathcal{L}(F)$ is the set of goal conditions over the state variables.
-A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, and $|\pi|=n$ denotes its {\em plan length}.
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\in\mathcal{L}(F)$ is the set of goal conditions over the state variables.
+A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length}.
 The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
 A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$.
+A solution plan is optimal iff its length is minimal.
 
@@
 Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,s_0)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
+In practice, the number of observed states, $m$, ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
 
-We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+We assume a bijective monotone mapping between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
 Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
@@
 \subsection{Conformant planning}
-{\em Conformant planning} refers to planning with incomplete information about the initial state, no sensing, and where goals have to be achieved with certainty (despite the uncertainty of the initial state)~\cite{smith1998conformant,goldman1996expressive}.
+{\em Conformant planning} is planning with incomplete information about the initial state, no sensing, and where goals have to be achieved with certainty (despite this uncertainty of the initial state)~\cite{smith1998conformant,goldman1996expressive}.
 
 Syntactically, conformant planning problems are expressed in compact form through a set of state variables.
-A {\em conformant planning problem} is then defined as a tuple $P_c=\tup{F,A,\Upsilon,G}$ where $F$, $A$ and $G$ are the set of fluents, actions and goals (as previously defined for the classical planing model).
+A {\em conformant planning problem} can be defined as a tuple $P_c=\tup{F,A,\Upsilon,G}$ where $F$, $A$ and $G$ are the set of {\em fluents}, {\em actions} and {\em goals} (as previously defined for {\em classical planning}).
 Now $\Upsilon$ is a set of clauses over literals $l=f$ or $l=\neg f$ (for $f\in F$) that define the set of possible initial states.
@@
 A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state.
-More precisely, an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$, such that $s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$, then the $\tau(\pi,s_0)$ trajectory reaches a final state $G \subseteq s_n$.
+More precisely, an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$, such that $s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$, then the $\tau(\pi,s_0)$ trajectory reaches a final state $G \subseteq s_n$ in which all the goal conditions are met.
 
 \section{Learning {\em the least commitment} action model from state observations}
-Now we are ready to formalize the learning of the {\em the least commitment} action model from state observations and our approach to address it via {\em conformant planning}.
+This section formalizes the notion of {\em the least commitment action model} that is able to {\em explain} a sequence of state observations and our approach to compute such model via {\em conformant planning}.
 
@@
 The task of computing {\em the least commitment} action model is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:
-
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a planning problem where $A[\cdot]$ is a set of actions.
+\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions.
 For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined.
-The set of goal conditions $G[\cdot]$ is also unknown.
-\item $M$ is the set of different planning models for the actions in $A[\cdot]$.
+\item $M$ is the set of possible models for the actions in $A[\cdot]$.
 A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$.
-Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
-\item $\mathcal{O}(\tau)$ is a sequence of state observations coming from the observation of a trajectory $\tau(\pi,s_0)$ produced by the execution of certain unknown plan $\pi$.
+Planning models $\mathcal{M}\in M$ differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
+\item $\mathcal{O}(\tau)$ is a sequence of partial states coming from the observation of a trajectory $\tau(\pi,s_0)$ produced by the execution of certain unknown plan $\pi$ that solves $P$.
 \end{itemize}
 
-The {\em solution} to this task is {\em the least commitment} action model which defines the smallest subset of models $M^*\subseteq M$ that is {\em conformant} with the input observation $\mathcal{O}(\tau)$.
+The {\em solution} to this task is {\em the least commitment} action model but, before formalizing it, we introduce the following necessary definition.
+
+\begin{definition}[Explanation of an observation]
+Given a model $\mathcal{M}\in M$ for the actions $A[\cdot]$ in $P$, and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that {\em the model explains the observation}, that we denote by $\mathcal{M}\mapsto\mathcal{O}(\tau)$, iff there exists a plan $\pi$ that solves $P$ and that it is consistent with $\pi$.
+We say that $\pi$ is the {\em best explanation} for $\mathcal{O}(\tau)$ iff, in additon, $\pi$ is optimal.
+\end{definition}
+
+Now we are ready to define {\em the least commitment} action model as the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ explains the input observation $\mathcal{O}(\tau)$.
 
 \subsection{The space of possible action models}
-This work focuses on the lerning of action models that are specified as \strips\ action schema.
+In spite of previous definitions are general, this work focuses on action models that are specified as \strips\ action schemata.
 \strips\ action schemata provide a compact representation for specifying action models.
 {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-
-\begin{definition}[Comparable \strips\ action schemata]
-Two \strips\ schemata $\xi$ and $\xi'$ are {\bf comparable} iff $pars(\xi)=pars(\xi')$, i.e, both share the same list of parameters.\footnote{In \strips\ models, $pars(\xi)=pars(\xi')$ implies the number of parameters must be the same.
-For other planning models that allow object typing, the equality implies that parameters share the same type}
-\end{definition}
-
-For instance, the {\small\tt stack(?v1,?v2)} and {\small\tt unstack(?v1,?v2)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are {\em comparable} while {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} are not.
-Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
 
@@
 only one block can be on top of another block).
-Invariants constraining the space of possible states in a given domain belong also to this group~\cite{fox1998automatic}.
+As a rule of thumb, {\it invariants} constraining the possible states of a given planning domain belong to this class of constraints~\cite{fox1998automatic}.
 \end{enumerate}

### 2018-11-27 doc/ijcai-19-conformant/conformant-learning.tex
first formalization completed of the conformant planning task

@@
 
-\section{Learning {\em the least commitment} action model from state observations}
+\section{Computing {\em the least commitment} action model from state observations}
 This section formalizes the notion of {\em the least commitment action model} that is able to {\em explain} a sequence of state observations and our approach to compute such model via {\em conformant planning}.
@@
 
-\subsection{Learning {\em the least commitment} model with conformant planning}
-Given a classical planning problem $P=\tup{F,A[\cdot],I,G[\cdot]}$ and a sequence of state observations $\mathcal{O}(\tau)$ coming from the observation of a $\tau(\pi,I)$ trajectory, we show here that we can build a {\em conformant planning problem} $P_c$ in linear time and space and whose solution is a plan that induces the {\em least commitment} model that is conformant with $\mathcal{O}(\tau)$.
+\subsection{Computing {\em the least commitment} model via conformant planning}
+Given a classical planning problem $P=\tup{F,A[\cdot],I,G}$ and a sequence of partial states $\mathcal{O}(\tau)$ coming from the observation of a $\tau(\pi,I)$ trajectory, we show here that we can build a {\em conformant planning problem} $P_c$ whose solution is a plan that induces the {\em least commitment} model.
 
@@
       \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
-      \item The fluents $[pre|del|add]\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ for a propositional encoding of the {\em preconditions}, {\em negative} and {\em positive} effects of schema $\xi$.
-Figure~\ref{fig:propositional} shows the propositional encoding of the unstack action from the blocksworld.
+      \item The fluents $pre\_e\_\xi$ and $eff\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ for a propositional encoding of the {\em preconditions}, {\em negative} and {\em positive} effects of schema $\xi$.
+Note that our encoding exploits the syntactic constraint of \strips.
+According to our encoding, if $pre\_e\_\xi$ and $eff\_e\_\xi$ holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ and if $pre\_e\_\xi$ does not hold but $eff\_e\_\xi$ holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same action.
       \end {itemize}
@@
 (:action stack
-   :parameters (?x ?y)
-   :precondition (and (holding ?x) (clear ?y))
-   :effect (and (not (holding ?x)) (not (clear ?y))
-                (clear ?x) (handempty) (on ?x ?y)))
+   :parameters (?v1 ?v2)
+   :precondition (and (holding ?v1) (clear ?v2))
+   :effect (and (not (holding ?v1)) (not (clear ?v2))
+                (clear ?v1) (handempty) (on ?v1 ?v2)))
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{set\top\_pre\_e\_\xi})=&\{\},\\
-\cond(\mathsf{set\top\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{pre\_e\_\xi\},\\
+\hspace*{7pt}\pre(\mathsf{commit\top\_pre\_e\_\xi})=&\{mode_{commit}\},\\
+\cond(\mathsf{commit\top\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{pre\_e\_\xi\},\\
                                     &\{\neg pre\_e\_\xi\}\rhd\{pre\_e\_\xi\}.\\
-\hspace*{7pt}\pre(\mathsf{set\bot\_pre\_e\_\xi})=&\{\},\\
-\cond(\mathsf{set\bot\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{\neg pre\_e\_\xi\},\\
+\hspace*{7pt}\pre(\mathsf{commit\bot\_pre\_e\_\xi})=&\{mode_{commit}\},\\
+\cond(\mathsf{commit\bot\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{\neg pre\_e\_\xi\},\\
                                     &\{\neg pre\_e\_\xi\}\rhd\{\neg pre\_e\_\xi\}.                                    
@@
 \end{small}
-In addition, similar actions are defined for {\em committing} {\small $eff\_e\_\xi$} fluents to a positive value and for {\em committing} {\small $pre\_e\_\xi$} fluents to a negative value.
+In addition, similar actions are defined for {\em committing} {\small $eff\_e\_\xi$} fluents to a positive value and to a negative value.
 
-      \item Actions for {\em validating} the edited models at the $s_j$ observed states, {\tt\small $0\leq j< m$}
+      \item Actions for {\em validating} the committed models at the $s_j$ observed states, {\tt\small $0\leq j< m$}
 \begin{small}
@@
 \cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\\
-                            &\{mode_{edit}\}\rhd\{\neg mode_{edit}, mode_{val}\}.
+                            &\{mode_{commit}\}\rhd\{\neg mode_{commit}, mode_{val}\}.
 \end{align*}
@@
 
-      \item Actions whose semantics is given by the value of the {\small $pre\_e\_\xi$}, {\small $pre\_e\_\xi$} fluents at the current state.
+      \item Actions whose semantics is given by the value of the {\small $pre\_e\_\xi$}, {\small $eff\_e\_\xi$} fluents at the current state.
 Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
@@
 \end{itemize}
-\item The new goals are $G_c=\{test_m\}$ .
+
+\item The clauses in $\Upsilon$ extend the set of fluents that hold in the initial state $I=s_0$ and $mode_{prog}$ set to true.
+In addition, $\Upsilon$ includes the two clauses $pre\_e\_\xi$ xor $\neg pre\_e\_\xi$ and $eff\_e\_\xi$ xor $\neg eff\_e\_\xi$ for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$.
+This means that the actual value of these fluents is actually unknown, so any model from the \strips\ space of models can initially be part of the {\em the least commitment} action model.
+
+\item The new goals are $G_c=\{test_m\}$.
 \end{itemize}
@@
 
-Given a plan $\pi$ that solves the {\em conformant planning} problem $P_c$ that results from our compilation then, the {\em least commitment} model that is conformant with $\mathcal{O}(\tau)$ is extracted in linear time from the last state reached by $\pi$.
-
 \subsection{Compilation properties}

### 2018-11-28 doc/ijcai-19-conformant/conformant-learning.tex
small updates added to ijcai19 paper

@@
 
-\title{Computing {\em the least commitment} action model from state observations}
+\title{Computing the {\em least-commitment} action model from state observations}
 
@@
 \section{Introduction}
-Given a sequence of partially observed states, this paper formalizes the task of computing {\em the least commitment} action model that can explain the given observation.
-This task is of interest because it allows the incremental learning of action models from arbitrary large sets of state observations.
+Given a sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to explain the given observation.
+This task is of interest because it allows the incremental learning of action models from arbitrary large sets of partial observations.
 
-In addition, the paper introduces a new method to compute {\em the least commitment} action model from a sequence of state observations.
+In addition, the paper introduces a new method to compute the {\em least-commitment} action model from a sequence of partially observed states.
 The method assumes that action models are specified as \strips\ action schema and it builds on top of off-the-shelf algorithms for {\em conformant planning}.
@@
 \section{Background}
-This section formalizes the model of {\em classical planning with conditional effects}, the {\em conformant planning} model, and the kind of input observations for the computation of {\em the least commitment} action model.
+This section formalizes the planning models we use in the paper as well as the kind of input observations for the computation of the {\em least-commitment} action model.
 
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as sequence of partial states that represents the observation from the execution of $\pi$ in $P$.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as a sequence of partial states coming from observing the execution of $\pi$ in $P$.
 Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$ where $s_0^o=I$.
@@
 
-We assume a bijective monotone mapping between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+In other words, we assume there is a bijective monotone mapping between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
 Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
@@
 \subsection{Conformant planning}
-{\em Conformant planning} is planning with incomplete information about the initial state, no sensing, and where goals have to be achieved with certainty (despite this uncertainty of the initial state)~\cite{smith1998conformant,goldman1996expressive}.
+{\em Conformant planning} is planning with incomplete information about the initial state, no sensing, and validating that goals are achieved with certainty (despite the uncertainty of the initial state)~\cite{smith1998conformant,goldman1996expressive}.
 
@@
 A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state.
-More precisely, an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$, such that $s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$, then the $\tau(\pi,s_0)$ trajectory reaches a final state $G \subseteq s_n$ in which all the goal conditions are met.
+More precisely, an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$, such that $s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$, then $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$ where all goal conditions are met.
 
-\section{Computing {\em the least commitment} action model from state observations}
-This section formalizes the notion of {\em the least commitment action model} that is able to {\em explain} a sequence of state observations and our approach to compute such model via {\em conformant planning}.
+\section{Computing the {\em least-commitment} action model from state observations}
+First, this section formalizes the notion of the {\em least-commitment} action model that is able to {\em explain} a sequence of partially observed states.
+Next, the section describes our approach to compute such model via {\em conformant planning}.
 
-\subsection{The {\em least commitment} action model}
-The task of computing {\em the least commitment} action model is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:
+\subsection{The {\em least-commitment} action model}
+The task of computing the {\em least-commitment} action model from state observations is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions.
-For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
+\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions s.t.
+for each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined.
@@
 
-The {\em solution} to this task is {\em the least commitment} action model but, before formalizing it, we introduce the following necessary definition.
+The {\em solution} to this task is the {\em least-commitment} action model, before formalizing it we introduce the following necessary definition.
 
 \begin{definition}[Explanation of an observation]
-Given a model $\mathcal{M}\in M$ for the actions $A[\cdot]$ in $P$, and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that {\em the model explains the observation}, that we denote by $\mathcal{M}\mapsto\mathcal{O}(\tau)$, iff there exists a plan $\pi$ that solves $P$ and that it is consistent with $\pi$.
+Given a model $\mathcal{M}\in M$ for the actions $A[\cdot]$ in $P$, and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that {\em the model explains the observation} (denoted by $\mathcal{M}\mapsto\mathcal{O}(\tau)$) iff there exists a plan $\pi$ that solves $P$ and that it is consistent with $\pi$.
 We say that $\pi$ is the {\em best explanation} for $\mathcal{O}(\tau)$ iff, in additon, $\pi$ is optimal.
@@
 
-Now we are ready to define {\em the least commitment} action model as the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ explains the input observation $\mathcal{O}(\tau)$.
+Now we are ready to define the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$ as the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ {\em explains} the input observation.
 
-\subsection{The space of possible action models}
-In spite of previous definitions are general, this work focuses on action models that are specified as \strips\ action schemata.
-\strips\ action schemata provide a compact representation for specifying action models.
+\subsection{The space of \strips\ action models}
+Despite previous definitions are general, this work focuses on the particular kind of action models that are specified as \strips\ action schemata.
+
 {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-
 Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
@@
 One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because of the semantic of the {\tt\small on} predicate (i.e.
-only one block can be on top of another block).
-As a rule of thumb, {\it invariants} constraining the possible states of a given planning domain belong to this class of constraints~\cite{fox1998automatic}.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this particular domain, only one block can be on top of another block.
+As a rule of thumb, {\it state invariants} constraining the possible states of a given planning domain belong to this second class of constraints~\cite{fox1998automatic}.
 \end{enumerate}
@@
 
-\subsection{Computing {\em the least commitment} model via conformant planning}
-Given a classical planning problem $P=\tup{F,A[\cdot],I,G}$ and a sequence of partial states $\mathcal{O}(\tau)$ coming from the observation of a $\tau(\pi,I)$ trajectory, we show here that we can build a {\em conformant planning problem} $P_c$ whose solution is a plan that induces the {\em least commitment} model.
+\subsection{Computing the {\em least-commitment} model via conformant planning}
+Given a sequence of partial states $\mathcal{O}(\tau)$, and a set of actions $A[\cdot]$ with known list of parameters, but unknown $\rho$ and $\theta$ functions, we can build a classical planning problem $P=\tup{F,A[\cdot],I,G}$ such that $\mathcal{O}(\tau)$ represents the observation of a $\tau(\pi,I)$ trajectory that solves $P$.
+In this section we show that starting from $\mathcal{O}(\tau)$ and $P$ we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for observation $\mathcal{O}(\tau)$.
 
-In more detail, our compilation defines a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G}$ such that:
+In more detail, we build a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G}$ such that:
 \begin{itemize}
@@
       \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
-      \item The fluents $pre\_e\_\xi$ and $eff\_e\_\xi$ such that $e\in{\mathcal I}_{\Psi,\xi}$ for a propositional encoding of the {\em preconditions}, {\em negative} and {\em positive} effects of schema $\xi$.
-Note that our encoding exploits the syntactic constraint of \strips.
-According to our encoding, if $pre\_e\_\xi$ and $eff\_e\_\xi$ holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ and if $pre\_e\_\xi$ does not hold but $eff\_e\_\xi$ holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same action.
+      \item The fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$) for a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$.
+Our encoding exploits the syntactic constraint of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ and if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
       \end {itemize}
@@
   \end{tiny}  
- \caption{\small PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same action.}
+ \caption{\small PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.}
 \label{fig:propositional}
@@
 \begin{itemize}
-      \item Actions for {\em committing} {\small $pre\_e\_\xi$} fluents to a positive and a negative value.
+      \item Actions for {\em committing} {\tt\small pre\_e\_$\xi$} fluents to a positive/negative value (similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} fluents to a positive/negative value).
 \begin{small}
@@
 \end{small}
-In addition, similar actions are defined for {\em committing} {\small $eff\_e\_\xi$} fluents to a positive value and to a negative value.
 
-      \item Actions for {\em validating} the committed models at the $s_j$ observed states, {\tt\small $0\leq j< m$}
+      \item Actions for {\em validating} that the committed models explain the $s_j$ observed states, {\tt\small $0\leq j< m$}
 \begin{small}
@@
 
-      \item Actions whose semantics is given by the value of the {\small $pre\_e\_\xi$}, {\small $eff\_e\_\xi$} fluents at the current state.
+      \item Actions whose semantics is given by the value of the {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} fluents at the current state.
 Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
@@
 Figure~\ref{fig:editable} shows the PDDL encoding of the editable version of the {\tt\small stack(?v1,?v2)} schema.
-Note that this editable schema, when the fluents {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)} hold, it behaves exactly as defined in Figure~\ref{fig:propositional}.
+Note that this editable schema, when the following set of fluents holds, {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)}, it behaves exactly as defined in Figure~\ref{fig:propositional}.
 \end{itemize}
@@
 \item The clauses in $\Upsilon$ extend the set of fluents that hold in the initial state $I=s_0$ and $mode_{prog}$ set to true.
-In addition, $\Upsilon$ includes the two clauses $pre\_e\_\xi$ xor $\neg pre\_e\_\xi$ and $eff\_e\_\xi$ xor $\neg eff\_e\_\xi$ for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$.
-This means that the actual value of these fluents is actually unknown, so any model from the \strips\ space of models can initially be part of the {\em the least commitment} action model.
+In addition, $\Upsilon$ includes the two clauses {\tt\small pre\_e\_$\xi$} xor {\tt\small $\neg$pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} xor {\tt\small $\neg$eff\_e\_$\xi$} for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$.
+This means that the actual value of these fluents is unknown, so any model from the \strips\ space of models can initially be part of the {\em least-commitment} action model.
 

### 2018-11-29 doc/ijcai-19-conformant/conformant-learning.tex
small updates added to ijcai19 paper
added definition for partially specified model

@@
 \subsection{The {\em least-commitment} action model}
-The task of computing the {\em least-commitment} action model from state observations is defined as a tuple $\tup{P,M,\mathcal{O}}$ where:
+The task of computing the {\em least-commitment} action model from state observations is defined as a $\tup{P,\mathcal{O}}$ pair:
 \begin{itemize}
@@
 for each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined.
-\item $M$ is the set of possible models for the actions in $A[\cdot]$.
-A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$.
-Planning models $\mathcal{M}\in M$ differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
+the corresponding $\tup{\rho,\theta}$ functions are undefined.
 \item $\mathcal{O}(\tau)$ is a sequence of partial states coming from the observation of a trajectory $\tau(\pi,s_0)$ produced by the execution of certain unknown plan $\pi$ that solves $P$.
@@
 
-The {\em solution} to this task is the {\em least-commitment} action model, before formalizing it we introduce the following necessary definition.
+Before formalizing the solution to this task, i.e.
+the {\em least-commitment} action model, we introduce several necessary definitions.
+We first start defining a {\em partially specified action model} inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}.
+\begin{definition}[Partially specified action model]
+Given a set of actions $A[\cdot]$ and a set of fluents $F$ then, a {\em partially specified action model} $M$ is a set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ defines the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$.
+(Note that if $M$ is a singleton it represents a {\em fully specified action model}).
+\end{definition}
 
 \begin{definition}[Explanation of an observation]
-Given a model $\mathcal{M}\in M$ for the actions $A[\cdot]$ in $P$, and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that {\em the model explains the observation} (denoted by $\mathcal{M}\mapsto\mathcal{O}(\tau)$) iff there exists a plan $\pi$ that solves $P$ and that it is consistent with $\pi$.
-We say that $\pi$ is the {\em best explanation} for $\mathcal{O}(\tau)$ iff, in additon, $\pi$ is optimal.
+Given a fully specified action model $\mathcal{M}$ for the actions $A[\cdot]$ in $P$, and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that {\em the model explains the observation} (denoted $\mathcal{M}\mapsto\mathcal{O}(\tau)$) iff there exists a plan $\pi$, consistent with $\mathcal{O}(\tau)$, that solves $P$.
+We say that $\pi$ is the {\em best explanation} for $\mathcal{O}(\tau)$ iff, in addition, $\pi$ is optimal.
 \end{definition}
 
-Now we are ready to define the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$ as the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ {\em explains} the input observation.
+Now we are ready to define the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$.
+\begin{definition}[The {\em least-commitment} action model]
+Given a {\em partially specified action model} $M$ that represents the space of possible action models and a sequence of partially observed states $\mathcal{O}(\tau)$, then the {\em least-commitment} action model is another {\em partially specified action model} that represents the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ {\em explains} the input observation.
+\end{definition}
 
@@
 \subsection{Computing the {\em least-commitment} model via conformant planning}
-Given a sequence of partial states $\mathcal{O}(\tau)$, and a set of actions $A[\cdot]$ with known list of parameters, but unknown $\rho$ and $\theta$ functions, we can build a classical planning problem $P=\tup{F,A[\cdot],I,G}$ such that $\mathcal{O}(\tau)$ represents the observation of a $\tau(\pi,I)$ trajectory that solves $P$.
+Given the set of actions $A[\cdot]$ with unknown $\rho$ and $\theta$ functions and a sequence of partial states $\mathcal{O}(\tau)$, we can build a classical planning problem $P=\tup{F,A[\cdot],I,G}$ such that $\mathcal{O}(\tau)$ represents the observation of a $\tau(\pi,I)$ trajectory that solves $P$.
+
 In this section we show that starting from $\mathcal{O}(\tau)$ and $P$ we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for observation $\mathcal{O}(\tau)$.
-
 In more detail, we build a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G}$ such that:
@@
 
-\item The clauses in $\Upsilon$ extend the set of fluents that hold in the initial state $I=s_0$ and $mode_{prog}$ set to true.
+\item The clauses in $\Upsilon$ extend the set of fluents that hold in the initial state $I=s_0$ and $mode_{commit}$ set to true.
 In addition, $\Upsilon$ includes the two clauses {\tt\small pre\_e\_$\xi$} xor {\tt\small $\neg$pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} xor {\tt\small $\neg$eff\_e\_$\xi$} for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$.
@@
 \section{Conclusions}
+~\cite{SternJ17}
 

### 2018-11-29 doc/ijcai-19-conformant/conformant-learning.tex
added updates to ijcai19 paper to make the new initial state clear

@@
 Given the set of actions $A[\cdot]$ with unknown $\rho$ and $\theta$ functions and a sequence of partial states $\mathcal{O}(\tau)$, we can build a classical planning problem $P=\tup{F,A[\cdot],I,G}$ such that $\mathcal{O}(\tau)$ represents the observation of a $\tau(\pi,I)$ trajectory that solves $P$.
+In this section we show that starting from $\mathcal{O}(\tau)$ and $P$ we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for the input observation $\mathcal{O}(\tau)$.
 
-In this section we show that starting from $\mathcal{O}(\tau)$ and $P$ we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for observation $\mathcal{O}(\tau)$.
 In more detail, we build a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G}$ such that:
@@
       \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
-      \item The fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$) for a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$.
+      \item The fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$) implementing a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$.
 Our encoding exploits the syntactic constraint of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ and if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
@@
 
-      \item Actions for {\em validating} that the committed models explain the $s_j$ observed states, {\tt\small $0\leq j< m$}
+      \item Actions for {\em validating} that committed models explain the $s_j$ observed states, {\tt\small $0\leq j< m$}.
 \begin{small}
@@
 
-      \item Actions whose semantics is given by the value of the {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} fluents at the current state.
-Given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is formalized as:
+      \item {\em Editable} actions whose semantics is given by the value of the {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} fluents at the current state.
+Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema.
+Note that this editable schema when {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)} holds, it behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional}.
+Formally, given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is:
 \begin{small}  
@@
 \end{small}
-Figure~\ref{fig:editable} shows the PDDL encoding of the editable version of the {\tt\small stack(?v1,?v2)} schema.
-Note that this editable schema, when the following set of fluents holds, {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)}, it behaves exactly as defined in Figure~\ref{fig:propositional}.
+
 \end{itemize}
 
-\item The clauses in $\Upsilon$ extend the set of fluents that hold in the initial state $I=s_0$ and $mode_{commit}$ set to true.
-In addition, $\Upsilon$ includes the two clauses {\tt\small pre\_e\_$\xi$} xor {\tt\small $\neg$pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} xor {\tt\small $\neg$eff\_e\_$\xi$} for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$.
-This means that the actual value of these fluents is unknown, so any model from the \strips\ space of models can initially be part of the {\em least-commitment} action model.
-
+\item The clauses in $\Upsilon$ comprises:
+      \begin{enumerate}
+      \item The {\em unit clauses} given by the fluents that hold in the initial state $I=s_0$ and $mode_{commit}$ set to true.
+      \item The clauses representing that the actual value of fluents {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} is unknown.
+In other words, that any model from the \strips\ space of models can initially be part of the {\em least-commitment} action model.
+Formally, for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$, then $\Upsilon$ includes these two clauses:
+            \begin{itemize}
+            \item {\tt\small pre\_e\_$\xi$} xor {\tt\small $\neg$pre\_e\_$\xi$}.
+            \item {\tt\small eff\_e\_$\xi$} xor {\tt\small $\neg$eff\_e\_$\xi$}.           
+            \end{itemize}
+      \end{enumerate}
 \item The new goals are $G_c=\{test_m\}$.

### 2018-11-29 doc/ijcai-19-conformant/conformant-learning.tex
updates added to address the comments by Diego

@@
 \section{Introduction}
-Given a sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to explain the given observation.
+Given a sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to {\em explain} the given observation.
 This task is of interest because it allows the incremental learning of action models from arbitrary large sets of partial observations.
@@
 In addition, the paper introduces a new method to compute the {\em least-commitment} action model from a sequence of partially observed states.
-The method assumes that action models are specified as \strips\ action schema and it builds on top of off-the-shelf algorithms for {\em conformant planning}.
+The method assumes that action models are specified as \strips\ action schemata and it is built on top of off-the-shelf algorithms for {\em conformant planning}.
 
 \section{Background}
-This section formalizes the planning models we use in the paper as well as the kind of input observations for the computation of the {\em least-commitment} action model.
+This section formalizes the planning models we use in the paper as well as the kind of observations that are given as input for the computation of the {\em least-commitment} action model.
 
@@
 
-A {\em planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
+A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
 An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\in\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\in\mathcal{L}(F)$.
@@
 A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length}.
-The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
-A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$.
+The execution of $\pi$ on the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
 A solution plan is optimal iff its length is minimal.
@@
 \subsection{The observation model}
-Given a classical planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as a sequence of partial states coming from observing the execution of $\pi$ in $P$.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as a sequence of partial states that results from observing the execution of $\pi$ on $I$.
 Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$ where $s_0^o=I$.
@@
 
+\begin{definition}[Explaning a $\mathcal{O}(\tau)$ observation]
+Given a {\em classical planning problem} $P$ and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that $P$ {\em explains the observation} (denoted $P\mapsto\mathcal{O}(\tau)$) iff there exists a solution plan $\pi$ for $P$ that is consistent with $\mathcal{O}(\tau)$.
+If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for $\mathcal{O}(\tau)$.
+\end{definition}
+
 \subsection{Conformant planning}
-{\em Conformant planning} is planning with incomplete information about the initial state, no sensing, and validating that goals are achieved with certainty (despite the uncertainty of the initial state)~\cite{smith1998conformant,goldman1996expressive}.
+{\em Conformant planning} is planning with incomplete information about the initial state, no sensing, and validating that goals are achieved with certainty (despite the uncertainty of the initial state)~\cite{goldman1996expressive,smith1998conformant,bonet2000planning}.
 
@@
 A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state.
-More precisely, an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$, such that $s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$, then $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$ where all goal conditions are met.
+More precisely, an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff, for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ s.t.
+$s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$, then $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$.
 
@@
 \subsection{The {\em least-commitment} action model}
-The task of computing the {\em least-commitment} action model from state observations is defined as a $\tup{P,\mathcal{O}}$ pair:
+The task of computing the {\em least-commitment} action model from a sequence of state observations is defined as $\tup{P,\mathcal{O}(\tau)}$:
 \begin{itemize}
-\item  $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions s.t.
-for each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
+\item $P=\tup{F,A[\cdot],I,G}$ is a classical planning problem where $A[\cdot]$ is a set of actions s.t.
+the semantics of each $a\in A[\cdot]$ is unknown; i.e.
 the corresponding $\tup{\rho,\theta}$ functions are undefined.
-\item $\mathcal{O}(\tau)$ is a sequence of partial states coming from the observation of a trajectory $\tau(\pi,s_0)$ produced by the execution of certain unknown plan $\pi$ that solves $P$.
+The set of goals $G$ can also be unknown.
+\item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the observation of a trajectory $\tau(\pi,s_0)$ produced by the execution of certain unknown plan $\pi$ that solves $P$.
 \end{itemize}
@@
 \begin{definition}[Partially specified action model]
-Given a set of actions $A[\cdot]$ and a set of fluents $F$ then, a {\em partially specified action model} $M$ is a set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ defines the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$.
+Given a set of actions $A[\cdot]$ and a set of fluents $F$ then, a {\em partially specified action model} $M$ is a set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$.
 (Note that if $M$ is a singleton it represents a {\em fully specified action model}).
-\end{definition}
-
-\begin{definition}[Explanation of an observation]
-Given a fully specified action model $\mathcal{M}$ for the actions $A[\cdot]$ in $P$, and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that {\em the model explains the observation} (denoted $\mathcal{M}\mapsto\mathcal{O}(\tau)$) iff there exists a plan $\pi$, consistent with $\mathcal{O}(\tau)$, that solves $P$.
-We say that $\pi$ is the {\em best explanation} for $\mathcal{O}(\tau)$ iff, in addition, $\pi$ is optimal.
 \end{definition}
@@
 \begin{definition}[The {\em least-commitment} action model]
-Given a {\em partially specified action model} $M$ that represents the space of possible action models and a sequence of partially observed states $\mathcal{O}(\tau)$, then the {\em least-commitment} action model is another {\em partially specified action model} that represents the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ {\em explains} the input observation.
+Given a $\tup{P,\mathcal{O}(\tau)}$ task and the {\em partially specified action model} $M$ that represents the full space of possible action models for the actions in $A[\cdot]\in P$, then the {\em least-commitment} action model is another {\em partially specified action model} that represents the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ {\em explains} the input observation.
 \end{definition}
@@
 We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
+For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contain eleven elements for the {\tt stack} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is constrained by a set ${\mathcal C}$ that includes:
-
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is constrained by constraints of three kinds:
 \begin{enumerate}
@@
 Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-
 \item {\em Domain-specific constraints}.
 One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this particular domain, only one block can be on top of another block.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this particular domain, a block cannot be on top of itself.
 As a rule of thumb, {\it state invariants} constraining the possible states of a given planning domain belong to this second class of constraints~\cite{fox1998automatic}.
-\end{enumerate}
-
-\begin{definition}[Well-defined \strips\ action schemata]
-Given a set of {\em predicates} $\Psi$, a list of action {\em parameters} $pars(\xi)$, and set of FOL constraints ${\mathcal C}$, $\xi$ is a {\bf well-defined \strips\ action schema} iff its three lists $pre(\xi)\subseteq {\mathcal I}_{\Psi,\xi}$, $del(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ and $add(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ only contain elements in ${\mathcal I}_{\Psi,\xi}$ and they satisfy all the constraints in ${\mathcal C}$.
-\end{definition}
-
-We say a planning model $\mathcal{M}$ is {\em well-defined} if all its \strips\ action schemata are {\em well-defined}.
-
-\begin{enumerate}
-\item[3.] {\em Observation constraints}.
-A sequence of state observations $\mathcal{O}(\tau)$ constraints further the space of possible action schemata.
-This \emph{semantic knowledge} included in the observations introduce a third type of constraints and can also be added to the set $\mathcal{C}$.
+\item {\em Observation constraints}. A sequence of state observations $\mathcal{O}(\tau)$ is \emph{semantic knowledge} that constraints further the space of possible action schemata. 
 \end{enumerate}
@@
 \subsection{Computing the {\em least-commitment} model via conformant planning}
-Given the set of actions $A[\cdot]$ with unknown $\rho$ and $\theta$ functions and a sequence of partial states $\mathcal{O}(\tau)$, we can build a classical planning problem $P=\tup{F,A[\cdot],I,G}$ such that $\mathcal{O}(\tau)$ represents the observation of a $\tau(\pi,I)$ trajectory that solves $P$.
-In this section we show that starting from $\mathcal{O}(\tau)$ and $P$ we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for the input observation $\mathcal{O}(\tau)$.
+Given the set of actions $A[\cdot]$ (with unknown $\tup{\rho,\theta}$ functions) and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
+In other words, a classical planning problem such that $\mathcal{O}(\tau)$ is the observation of a $\tau(\pi,I)$ trajectory that solves that problem.
 
+In this section we show that starting from a $\tup{P_\mathcal{O},\mathcal{O}(\tau)}$ task we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for the input observation $\mathcal{O}(\tau)$.
 In more detail, we build a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G}$ such that:
@@
       \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
-      \item The fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$) implementing a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$.
-Our encoding exploits the syntactic constraint of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ and if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+      \item Fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$) implementing a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$.
+Our encoding exploits the syntactic constraint of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
 Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
@@
 \cond(\mathsf{commit\top\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{pre\_e\_\xi\},\\
-                                    &\{\neg pre\_e\_\xi\}\rhd\{pre\_e\_\xi\}.\\
+                                    &\{\neg pre\_e\_\xi\}\rhd\{pre\_e\_\xi\}.\\\\
 \hspace*{7pt}\pre(\mathsf{commit\bot\_pre\_e\_\xi})=&\{mode_{commit}\},\\
@@
 
-      \item {\em Editable} actions whose semantics is given by the value of the {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} fluents at the current state.
+      \item {\em Editable} actions whose semantics is given by the value of {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} fluents at the current state.
 Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema.
-Note that this editable schema when {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)} holds, it behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional}.
+Note that this editable schema when {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)} hold, then it behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional}.
 Formally, given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is:
@@
       \item The clauses representing that the actual value of fluents {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} is unknown.
-In other words, that any model from the \strips\ space of models can initially be part of the {\em least-commitment} action model.
+In other words, that any model from the \strips\ space of models (following the previously mentioned syntactic constraints) can initially be part of the {\em least-commitment} action model.
 Formally, for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$, then $\Upsilon$ includes these two clauses:
@@
       \end{enumerate}
+One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier to be solved for a particular domain. 
 \item The new goals are $G_c=\{test_m\}$.

### 2018-11-30 doc/ijcai-19-conformant/conformant-learning.tex
adding comments by Eva and Diego from todays meeting

@@
 
-A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+A partially observable state $s_i^o$, {\small $0<i<m$}, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
 Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
 Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,s_0)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-In practice, the number of observed states, $m$, ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
+In practice, the number of observed states $m$, ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
 
-In other words, we assume there is a bijective monotone mapping between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+In other words, we assume there is a {\em bijective monotone mapping} between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
 Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
@@
 \begin{definition}[Explaning a $\mathcal{O}(\tau)$ observation]
-Given a {\em classical planning problem} $P$ and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that $P$ {\em explains the observation} (denoted $P\mapsto\mathcal{O}(\tau)$) iff there exists a solution plan $\pi$ for $P$ that is consistent with $\mathcal{O}(\tau)$.
+Given a {\em classical planning problem} $P$ and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that a plan $\pi$ {\em explains the observation} (denoted $\pi\mapsto\mathcal{O}(\tau)$) iff $\pi$ is a solution for $P$ that is consistent with $\mathcal{O}(\tau)$.
 If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for $\mathcal{O}(\tau)$.
@@
 \subsection{The {\em least-commitment} action model}
-The task of computing the {\em least-commitment} action model from a sequence of state observations is defined as $\tup{P,\mathcal{O}(\tau)}$:
+The task of computing the {\em least-commitment} action model from a sequence of state observations is defined as $\tup{\Phi,\mathcal{O}(\tau)}$:
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$ is a classical planning problem where $A[\cdot]$ is a set of actions s.t.
+\item $\Phi=\tup{F,A[\cdot]}$ is a {\em classical planning frame} where $A[\cdot]$ is a set of actions s.t.
 the semantics of each $a\in A[\cdot]$ is unknown; i.e.
 the corresponding $\tup{\rho,\theta}$ functions are undefined.
-The set of goals $G$ can also be unknown.
-\item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the observation of a trajectory $\tau(\pi,s_0)$ produced by the execution of certain unknown plan $\pi$ that solves $P$.
+\item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the partial observation of a trajectory $\tau(\pi,s_0)$ that is defined within the {\em classical planning frame} $\Phi$.
 \end{itemize}
@@
 the {\em least-commitment} action model, we introduce several necessary definitions.
-We first start defining a {\em partially specified action model} inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}.
+Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
+We say that an action model $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$.
+Further, we say that a model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff there exists a solution plan for $P_\mathcal{O}$ when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$.
+
+Now, we define a {\em partially specified action model} inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}.
 \begin{definition}[Partially specified action model]
-Given a set of actions $A[\cdot]$ and a set of fluents $F$ then, a {\em partially specified action model} $M$ is a set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$.
+Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ then, a {\em partially specified action model} $M$ is a set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$.
 (Note that if $M$ is a singleton it represents a {\em fully specified action model}).
@@
 \begin{definition}[The {\em least-commitment} action model]
-Given a $\tup{P,\mathcal{O}(\tau)}$ task and the {\em partially specified action model} $M$ that represents the full space of possible action models for the actions in $A[\cdot]\in P$, then the {\em least-commitment} action model is another {\em partially specified action model} that represents the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ {\em explains} the input observation.
+Given a $\tup{\Phi,\mathcal{O}(\tau)}$ task and the {\em partially specified action model} $M$ that represents the full space of possible action models for the actions in $A[\cdot]\in \Phi$, then the {\em least-commitment} action model is another {\em partially specified action model} that represents the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ {\em explains} the input observation.
 \end{definition}
@@
 We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
-For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contain eleven elements for the {\tt stack} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contain eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is constrained by constraints of three kinds:
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is bounded by constraints of three kinds:
 \begin{enumerate}
@@
 One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this particular domain, a block cannot be on top of itself.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
 As a rule of thumb, {\it state invariants} constraining the possible states of a given planning domain belong to this second class of constraints~\cite{fox1998automatic}.
-\item {\em Observation constraints}. A sequence of state observations $\mathcal{O}(\tau)$ is \emph{semantic knowledge} that constraints further the space of possible action schemata. 
+\item {\em Observation constraints}. A sequence of state observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata. 
 \end{enumerate}
@@
 \subsection{Computing the {\em least-commitment} model via conformant planning}
-Given the set of actions $A[\cdot]$ (with unknown $\tup{\rho,\theta}$ functions) and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
-In other words, a classical planning problem such that $\mathcal{O}(\tau)$ is the observation of a $\tau(\pi,I)$ trajectory that solves that problem.
-
-In this section we show that starting from a $\tup{P_\mathcal{O},\mathcal{O}(\tau)}$ task we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for the input observation $\mathcal{O}(\tau)$.
-In more detail, we build a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G}$ such that:
+Given a task $\tup{\Phi,\mathcal{O}(\tau)}$, this section shows that we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for the input observation $\mathcal{O}(\tau)$.
+In more detail, we build a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G_c}$ such that:
 \begin{itemize}
@@
 Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema.
-Note that this editable schema when {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)} hold, then it behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional}.
+Note that this editable schema when the set of fluents {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)} hold at the current state, then it behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional}.
 Formally, given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is:
@@
             \begin{itemize}
-            \item {\tt\small pre\_e\_$\xi$} xor {\tt\small $\neg$pre\_e\_$\xi$}.
-            \item {\tt\small eff\_e\_$\xi$} xor {\tt\small $\neg$eff\_e\_$\xi$}.           
+            \item {\tt\small pre\_e\_$\xi$} $\vee$ {\tt\small $\neg$pre\_e\_$\xi$}.
+            \item {\tt\small eff\_e\_$\xi$} $\vee$ {\tt\small $\neg$eff\_e\_$\xi$}.
             \end{itemize}
       \end{enumerate}
-One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier to be solved for a particular domain. 
-\item The new goals are $G_c=\{test_m\}$.
+One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier to be solved for a specific domain. 
+\item The new goals are $G_c=s_m\cup\{test_m\}$.
 \end{itemize}
@@
 
+\subsection{Optimization of the compilation}
+In fact, we do not need to compute the {\em least-commitment} action model (i.e.
+solve the $P_c$ conformant planning problem) from scratch for all the $\tup{s_0^o,s_1^o \ldots , s_m^o}$ observations.
+A more efficient approach is to compute the {\em least-commitment} action model from scratch only for the first observation, that is for $\tup{s_0^o,s_1^o}$.
+Then, the obtained {\em least-commitment} action model can be used as input for improving the computation of the two first observations $\tup{s_0^o,s_1^o,s_2^o}$.
+In other words, including the commits found in the previous step as new {\em unit clauses} in set of the {\em initial clauses} $\Upsilon$.
+This process is repeated until the {\em least-commitment} action model is computed  for the full sequence of state observations $\tup{s_0^o,s_1^o \ldots , s_m^o}$.
+
 \subsection{Compilation properties}
@@
 \section{Conclusions}
-~\cite{SternJ17}
+Related work~\cite{SternJ17}.
 
@@
 \bibliographystyle{named}
-\bibliography{planlearnbibliography.bib}
+\bibliography{planlearnbibliography}
 

### 2018-11-30 doc/ijcai-19-conformant/conformant-learning.tex
adding comments by Eva and Diego from todays meeting

@@
 We say that an action model $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$.
-Further, we say that a model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff there exists a solution plan for $P_\mathcal{O}$ when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$.
+Further we say that a model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that explains $\mathcal{O}(\tau)$.
 
@@
       \item Fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$) implementing a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$.
-Our encoding exploits the syntactic constraint of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+{\bf Our encoding exploits the syntactic constraint of \strips\ } so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
 Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
@@
       \end{enumerate}
-One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier to be solved for a specific domain. 
+One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier for a specific domain. 
 \item The new goals are $G_c=s_m\cup\{test_m\}$.

### 2018-12-01 doc/aij18/major-revision/letter.tex
updates in response letter

@@
 \usepackage[framemethod=tikz]{mdframed}
+\usepackage{xspace}
 
@@
 \newcommand{\strips}{\textsc{Strips}} % Strips
+\newcommand{\FAMA}{{\small {\sffamily FAMA}}\xspace}
 
+\title{\textbf{Response Letter}}
 \begin{document}
+\maketitle
+
+Firstly, we would like to thank the three reviewers for their insightful comments and recommendations.
+We truly believe the observations pointed by the three reviewers have been very helpful to make the paper more readable.
+
+This response letter is structured as follows.
+The next section summarizes the most relevant modifications addressed in the paper.
+Subsequent sections are devoted to thoroughly reply all the reviewers' concerns.
+
+\subsection*{Main modifications addressed in the paper}
+
+We identified four main points to be addressed as per the reviewers' comments and observations.
+In this section, we provide details of these four main modifications, and we will point at them when replying to each of the reviewers' comments in the remainder sections.
+
+\begin{itemize}
+\item \textbf{Introduction}.
+We performed a profound revision of the Introduction in order to make it more self-explanatory.
+First, we carefully revised all of the comments pointed by reviewer \#2.
+Additionally, as indicated by reviewer \#1, we provide at the end of the section the overall picture of our approach.
+Our modification is intended to adequately position our contribution within the concerned research field, and to provide a global picture of the \FAMA approach that explains our pursued objective.
+
+\item \textbf{Problem definition and complexity}.
+As indicated by reviewer \#2, and also supported by reviewers \#1 and \#3, the problem definition is not clearly introduced,.
+We believe this is the reason that our claims about the complexity of the learning task were not fully comprehensible.
+We acknowledge this lack of problem definition, and we introduce a subsection 3.1 to clearly define upfront the basic and essence of the learning task we are aimed to solve.
+
+\item \textbf{Compilation scheme}.
+Reviewers \#1 and \#3 show several concerns about the understanding of section 4.4., the compilation approach.
+We acknowledge their concerns and we have thus introduced an introductory explanatory subsection that explains ..........
+before going into the technical details.
+We hope this new subsection will help the reader to have first a broad view of the compilation scheme before diving into the technicalities.
+
+\end{itemize}
 
@@
 The second reason for revision is the comparison to ARMS.
+\end{mdframed}
+
+We agree with the reviewer that we should provide ....
+
+\begin{mdframed} [hidealllines=true,backgroundcolor=gray!20]
 I suggest to compare with LOUGA that is better than ARMS in all aspects (runtime and quality of obtained models) so ARMS is no more state-of-the-art.
 \end{mdframed}
+
+As per the reviewer's recommendation, we requested the code of LOUGA to the first author of the publication entitled \emph{LOUGA: learning planning operators using genetic algorithms}, which was published in the Pacific Rim Knowledge Acquisition Workshop.
+The author of LOUGA promptly provided us with the code, so we would first like to thank him for his gentle response.
+However, the issue we faced with LOUGA is that we were not able to automate the processing and output of the results.
+LOUGA throws results in a graphical interface, which makes it hard to easily retrieve the data to be subsequently processed.
+Since the input of our automated process for extracting the experimental results are plain text files, we could not adapt the graphical output of LOUGA to our required input.
+Nevertheless, we did test a couple of ....
 

### 2018-12-01 doc/aij18/major-revision/letter.tex
updates in response letter

@@
 
-We identified four main points to be addressed as per the reviewers' comments and observations.
-In this section, we provide details of these four main modifications, and we will point at them when replying to each of the reviewers' comments in the remainder sections.
+We identified three main points to be addressed as per the reviewers' comments and observations.
+In this section, we provide details of these three main modifications, and we will point at them when replying to each of the reviewers' comments in the remainder sections.
 

### 2018-12-03 doc/aij18/major-revision/aij18.tex
First round of revision

--- 2018-08-06 doc/aij18/aij18.tex
@@
 While the experimental evaluation shows in general an exponential increase of the computation time as the number of training traces augments, \FAMA is able to learn action models more accurate than those of \ARMS with very limited input knowledge.
-Unlike extensive-data ML approaches, our work explores an alternative research direction that exploits logic reasoning to learn sound models from minimal input knowledge.
+Unlike extensive-data ML approaches, our work explores an alternative research direction that exploits logic reasoning to learn sound models from minimal input observability.
 

### 2018-12-04 src/compiler_new.py
compiler_new.py: added -f option to indicate known horizon

@@ -136,8 +136,8 @@ try:
     action_observability = float(sys.argv[2])/100
     state_observability = float(sys.argv[3])/100
 
-    if action_observability == 1 or state_observability == 1:
-        finite_steps = True
+    # if action_observability == 1 or state_observability == 1:
+    #     finite_steps = True
 
 except:
     print "Usage:"
@@ -265,6 +265,9 @@ for a in actions:
         if known_precondition.predicate not in pres_filter:
             pre += [pddl.conditions.Atom(known_precondition.predicate,
                                      ["?o" + str(original_params.index(arg) + 1) for arg in known_precondition.args])]
+
+    if finite_steps:
+        pre += [pddl.conditions.NegatedAtom("action_applied", [])]
 
     eff = list()
 
@@ -684,7 +687,7 @@ validation_steps = max(states_seen-1, total_actions_seen)*2 + 1
 if action_observability == 1 and state_observability == 0:
     validation_steps = states_seen + total_actions_seen
 starting_horizon = str(validation_steps + 2)
-if finite_steps:
+if action_observability == 1 or state_observability == 1:
     ending_horizon = " -T " + starting_horizon
 else:
     ending_horizon = ""

### 2018-12-04 doc/aij18/major-revision/appendix.tex

\section*{Appendix}
\label{sec:appendix}

\begin{figure}[hbtp!]
\begin{scriptsize}  
\begin{verbatim}
(define (domain BLOCKS)
  (:requirements :strips)
  (:predicates (on ?x ?y)(ontable ?x)(clear ?x)(handempty)(holding ?x))

  (:action pick-up
	     :parameters (?x)
	     :precondition (and (clear ?x) (ontable ?x) (handempty))
	     :effect (and (not (ontable ?x)) (not (clear ?x)) (not (handempty)) (holding ?x)))

  (:action put-down
	     :parameters (?x)
	     :precondition (holding ?x)
	     :effect (and (not (holding ?x)) (clear ?x) (handempty) (ontable ?x)))

  (:action stack
	     :parameters (?x ?y)
	     :precondition (and (holding ?x) (clear ?y))
	     :effect (and (not (holding ?x)) (not (clear ?y)) (clear ?x) (handempty) (on ?x ?y)))

  (:action unstack
	     :parameters (?x ?y)
	     :precondition (and (on ?x ?y) (clear ?x) (handempty))
	     :effect (and (holding ?x) (clear ?y) (not (clear ?x)) (not (handempty)) (not (on ?x ?y)))))
  \end{verbatim}
\end{scriptsize}  
\caption{\small PDDL domain file for the blocksworld domain.}
\label{fig:bw-domain}
\end{figure}

Compiled PDDL domain file for learning the blocksworld action models from two initial and final states.

\begin{scriptsize}  
\begin{verbatim}
(define (domain blocks)
 (:requirements :strips)
 (:types object - None )
 (:constants a -  object b -  object c -  object d -  object e -  object f -  object g -  object)
 (:predicates (add_clear_pick-up_var1) (add_clear_put-down_var1) (add_clear_stack_var1) (add_clear_stack_var2) 
(add_clear_unstack_var1) (add_clear_unstack_var2) (add_handempty_pick-up) (add_handempty_put-down) 
(add_handempty_stack) (add_handempty_unstack) (add_holding_pick-up_var1) (add_holding_put-down_var1) 
(add_holding_stack_var1) (add_holding_stack_var2) (add_holding_unstack_var1) (add_holding_unstack_var2) 
(add_on_stack_var1_var1) (add_on_stack_var1_var2) (add_on_stack_var2_var1) (add_on_stack_var2_var2) 
(add_on_unstack_var1_var1) (add_on_unstack_var1_var2) (add_on_unstack_var2_var1) (add_on_unstack_var2_var2) 
(add_ontable_pick-up_var1) (add_ontable_put-down_var1) (add_ontable_stack_var1) (add_ontable_stack_var2) 
(add_ontable_unstack_var1) (add_ontable_unstack_var2) (clear ?o1 - object) (del_clear_pick-up_var1) 
(del_clear_put-down_var1) (del_clear_stack_var1) (del_clear_stack_var2) (del_clear_unstack_var1) 
(del_clear_unstack_var2) (del_handempty_pick-up) (del_handempty_put-down) (del_handempty_stack) 
(del_handempty_unstack) (del_holding_pick-up_var1) (del_holding_put-down_var1) (del_holding_stack_var1) 
(del_holding_stack_var2) (del_holding_unstack_var1) (del_holding_unstack_var2) (del_on_stack_var1_var1) 
(del_on_stack_var1_var2) (del_on_stack_var2_var1) (del_on_stack_var2_var2) (del_on_unstack_var1_var1) 
(del_on_unstack_var1_var2) (del_on_unstack_var2_var1) (del_on_unstack_var2_var2) (del_ontable_pick-up_var1) 
(del_ontable_put-down_var1) (del_ontable_stack_var1) (del_ontable_stack_var2) (del_ontable_unstack_var1) 
(del_ontable_unstack_var2) (handempty) (holding ?o1 - object) (modeProg) (on ?o1 - object ?o2 - object) 
(ontable ?o1 - object) (pre_clear_pick-up_var1) (pre_clear_put-down_var1) (pre_clear_stack_var1) 
(pre_clear_stack_var2) (pre_clear_unstack_var1) (pre_clear_unstack_var2) (pre_handempty_pick-up) 
(pre_handempty_put-down) (pre_handempty_stack) (pre_handempty_unstack) (pre_holding_pick-up_var1) 
(pre_holding_put-down_var1) (pre_holding_stack_var1) (pre_holding_stack_var2) (pre_holding_unstack_var1) 
(pre_holding_unstack_var2) (pre_on_stack_var1_var1) (pre_on_stack_var1_var2) (pre_on_stack_var2_var1) 
(pre_on_stack_var2_var2) (pre_on_unstack_var1_var1) (pre_on_unstack_var1_var2) (pre_on_unstack_var2_var1) 
(pre_on_unstack_var2_var2) (pre_ontable_pick-up_var1) (pre_ontable_put-down_var1) (pre_ontable_stack_var1) 
(pre_ontable_stack_var2) (pre_ontable_unstack_var1) (pre_ontable_unstack_var2) (test1) (test2) (test3))

 (:action pick-up
   :parameters (?o1 - object)
   :precondition (and (not (modeProg ))(or (not (pre_ontable_pick-up_var1 ))(ontable ?o1))
(or (not (pre_clear_pick-up_var1 ))(clear ?o1))(or (not (pre_handempty_pick-up ))(handempty ))
(or (not (pre_holding_pick-up_var1 ))(holding ?o1)))
   :effect (and (when (and (del_ontable_pick-up_var1 ))(not (ontable ?o1)))
(when (and (add_ontable_pick-up_var1 ))(ontable ?o1))(when (and (del_clear_pick-up_var1 ))(not (clear ?o1)))
(when (and (add_clear_pick-up_var1 ))(clear ?o1))(when (and (del_handempty_pick-up ))(not (handempty )))
(when (and (add_handempty_pick-up ))(handempty ))(when (and (del_holding_pick-up_var1 ))(not (holding ?o1)))
(when (and (add_holding_pick-up_var1 ))(holding ?o1))))

 (:action put-down
   :parameters (?o1 - object)
   :precondition (and (not (modeProg ))(or (not (pre_ontable_put-down_var1 ))(ontable ?o1))
(or (not (pre_clear_put-down_var1 ))(clear ?o1))(or (not (pre_handempty_put-down ))(handempty ))
(or (not (pre_holding_put-down_var1 ))(holding ?o1)))
   :effect (and (when (and (del_ontable_put-down_var1 ))(not (ontable ?o1)))
(when (and (add_ontable_put-down_var1 ))(ontable ?o1))(when (and (del_clear_put-down_var1 ))(not (clear ?o1)))
(when (and (add_clear_put-down_var1 ))(clear ?o1))(when (and (del_handempty_put-down ))(not (handempty )))
(when (and (add_handempty_put-down ))(handempty ))(when (and (del_holding_put-down_var1 ))(not (holding ?o1)))
(when (and (add_holding_put-down_var1 ))(holding ?o1))))

 (:action stack
   :parameters (?o1 - object ?o2 - object)
   :precondition (and (not (modeProg ))(or (not (pre_on_stack_var1_var1 ))(on ?o1 ?o1))
(or (not (pre_on_stack_var1_var2 ))(on ?o1 ?o2))(or (not (pre_on_stack_var2_var1 ))(on ?o2 ?o1))
(or (not (pre_on_stack_var2_var2 ))(on ?o2 ?o2))(or (not (pre_ontable_stack_var1 ))(ontable ?o1))
(or (not (pre_ontable_stack_var2 ))(ontable ?o2))(or (not (pre_clear_stack_var1 ))(clear ?o1))
(or (not (pre_clear_stack_var2 ))(clear ?o2))(or (not (pre_handempty_stack ))(handempty ))
(or (not (pre_holding_stack_var1 ))(holding ?o1))(or (not (pre_holding_stack_var2 ))(holding ?o2)))
   :effect (and (when (and (del_on_stack_var1_var1 ))(not (on ?o1 ?o1)))(when (and (add_on_stack_var1_var1 ))(on ?o1 ?o1))
(when (and (del_on_stack_var1_var2 ))(not (on ?o1 ?o2)))(when (and (add_on_stack_var1_var2 ))(on ?o1 ?o2))
(when (and (del_on_stack_var2_var1 ))(not (on ?o2 ?o1)))(when (and (add_on_stack_var2_var1 ))(on ?o2 ?o1))
(when (and (del_on_stack_var2_var2 ))(not (on ?o2 ?o2)))(when (and (add_on_stack_var2_var2 ))(on ?o2 ?o2))
(when (and (del_ontable_stack_var1 ))(not (ontable ?o1)))(when (and (add_ontable_stack_var1 ))(ontable ?o1))
(when (and (del_ontable_stack_var2 ))(not (ontable ?o2)))(when (and (add_ontable_stack_var2 ))(ontable ?o2))
(when (and (del_clear_stack_var1 ))(not (clear ?o1)))(when (and (add_clear_stack_var1 ))(clear ?o1))
(when (and (del_clear_stack_var2 ))(not (clear ?o2)))(when (and (add_clear_stack_var2 ))(clear ?o2))
(when (and (del_handempty_stack ))(not (handempty )))(when (and (add_handempty_stack ))(handempty ))
(when (and (del_holding_stack_var1 ))(not (holding ?o1)))(when (and (add_holding_stack_var1 ))(holding ?o1))
(when (and (del_holding_stack_var2 ))(not (holding ?o2)))(when (and (add_holding_stack_var2 ))(holding ?o2))))

 (:action unstack
   :parameters (?o1 - object ?o2 - object)
   :precondition (and (not (modeProg ))(or (not (pre_on_unstack_var1_var1 ))(on ?o1 ?o1))
(or (not (pre_on_unstack_var1_var2 ))(on ?o1 ?o2))(or (not (pre_on_unstack_var2_var1 ))(on ?o2 ?o1))
(or (not (pre_on_unstack_var2_var2 ))(on ?o2 ?o2))(or (not (pre_ontable_unstack_var1 ))(ontable ?o1))
(or (not (pre_ontable_unstack_var2 ))(ontable ?o2))(or (not (pre_clear_unstack_var1 ))(clear ?o1))
(or (not (pre_clear_unstack_var2 ))(clear ?o2))(or (not (pre_handempty_unstack ))(handempty ))
(or (not (pre_holding_unstack_var1 ))(holding ?o1))(or (not (pre_holding_unstack_var2 ))(holding ?o2)))
   :effect (and (when (and (del_on_unstack_var1_var1 ))(not (on ?o1 ?o1)))
(when (and (add_on_unstack_var1_var1 ))(on ?o1 ?o1))(when (and (del_on_unstack_var1_var2 ))(not (on ?o1 ?o2)))
(when (and (add_on_unstack_var1_var2 ))(on ?o1 ?o2))(when (and (del_on_unstack_var2_var1 ))(not (on ?o2 ?o1)))
(when (and (add_on_unstack_var2_var1 ))(on ?o2 ?o1))(when (and (del_on_unstack_var2_var2 ))(not (on ?o2 ?o2)))
(when (and (add_on_unstack_var2_var2 ))(on ?o2 ?o2))(when (and (del_ontable_unstack_var1 ))(not (ontable ?o1)))
(when (and (add_ontable_unstack_var1 ))(ontable ?o1))(when (and (del_ontable_unstack_var2 ))(not (ontable ?o2)))
(when (and (add_ontable_unstack_var2 ))(ontable ?o2))(when (and (del_clear_unstack_var1 ))(not (clear ?o1)))
(when (and (add_clear_unstack_var1 ))(clear ?o1))(when (and (del_clear_unstack_var2 ))(not (clear ?o2)))
(when (and (add_clear_unstack_var2 ))(clear ?o2))(when (and (del_handempty_unstack ))(not (handempty )))
(when (and (add_handempty_unstack ))(handempty ))(when (and (del_holding_unstack_var1 ))(not (holding ?o1)))
(when (and (add_holding_unstack_var1 ))(holding ?o1))(when (and (del_holding_unstack_var2 ))(not (holding ?o2)))
(when (and (add_holding_unstack_var2 ))(holding ?o2))))

 (:action program_pre_ontable_pick-up_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_pick-up_var1 ))(not (del_ontable_pick-up_var1 ))
  (not (add_ontable_pick-up_var1 )))
   :effect (and (pre_ontable_pick-up_var1 )))

 (:action program_eff_ontable_pick-up_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_pick-up_var1 ))(not (add_ontable_pick-up_var1 )))
  :effect (and (when (pre_ontable_pick-up_var1 )(del_ontable_pick-up_var1 ))
  (when (not (pre_ontable_pick-up_var1 ))(add_ontable_pick-up_var1 ))))

 (:action program_pre_clear_pick-up_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_pick-up_var1 ))(not (del_clear_pick-up_var1 ))
  (not (add_clear_pick-up_var1 )))
   :effect (and (pre_clear_pick-up_var1 )))

 (:action program_eff_clear_pick-up_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_pick-up_var1 ))(not (add_clear_pick-up_var1 )))
  :effect (and (when (pre_clear_pick-up_var1 )(del_clear_pick-up_var1 ))
  (when (not (pre_clear_pick-up_var1 ))(add_clear_pick-up_var1 ))))

 (:action program_pre_handempty_pick-up
   :parameters ()
  :precondition (and (modeProg )(not (pre_handempty_pick-up ))(not (del_handempty_pick-up ))
  (not (add_handempty_pick-up )))
   :effect (and (pre_handempty_pick-up )))

 (:action program_eff_handempty_pick-up
   :parameters ()
   :precondition (and (modeProg )(not (del_handempty_pick-up ))(not (add_handempty_pick-up )))
  :effect (and (when (pre_handempty_pick-up )(del_handempty_pick-up ))
  (when (not (pre_handempty_pick-up ))(add_handempty_pick-up ))))

 (:action program_pre_holding_pick-up_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_pick-up_var1 ))(not (del_holding_pick-up_var1 ))
  (not (add_holding_pick-up_var1 )))
   :effect (and (pre_holding_pick-up_var1 )))

 (:action program_eff_holding_pick-up_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_pick-up_var1 ))(not (add_holding_pick-up_var1 )))
  :effect (and (when (pre_holding_pick-up_var1 )(del_holding_pick-up_var1 ))
  (when (not (pre_holding_pick-up_var1 ))(add_holding_pick-up_var1 ))))

 (:action program_pre_ontable_put-down_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_put-down_var1 ))(not (del_ontable_put-down_var1 ))
  (not (add_ontable_put-down_var1 )))
   :effect (and (pre_ontable_put-down_var1 )))

 (:action program_eff_ontable_put-down_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_put-down_var1 ))(not (add_ontable_put-down_var1 )))
  :effect (and (when (pre_ontable_put-down_var1 )(del_ontable_put-down_var1 ))
  (when (not (pre_ontable_put-down_var1 ))(add_ontable_put-down_var1 ))))

 (:action program_pre_clear_put-down_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_put-down_var1 ))(not (del_clear_put-down_var1 ))
  (not (add_clear_put-down_var1 )))
   :effect (and (pre_clear_put-down_var1 )))

 (:action program_eff_clear_put-down_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_put-down_var1 ))(not (add_clear_put-down_var1 )))
  :effect (and (when (pre_clear_put-down_var1 )(del_clear_put-down_var1 ))
  (when (not (pre_clear_put-down_var1 ))(add_clear_put-down_var1 ))))

 (:action program_pre_handempty_put-down
   :parameters ()
  :precondition (and (modeProg )(not (pre_handempty_put-down ))(not (del_handempty_put-down ))
  (not (add_handempty_put-down )))
   :effect (and (pre_handempty_put-down )))

 (:action program_eff_handempty_put-down
   :parameters ()
   :precondition (and (modeProg )(not (del_handempty_put-down ))(not (add_handempty_put-down )))
  :effect (and (when (pre_handempty_put-down )(del_handempty_put-down ))
  (when (not (pre_handempty_put-down ))(add_handempty_put-down ))))

 (:action program_pre_holding_put-down_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_put-down_var1 ))(not (del_holding_put-down_var1 ))
  (not (add_holding_put-down_var1 )))
   :effect (and (pre_holding_put-down_var1 )))

 (:action program_eff_holding_put-down_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_put-down_var1 ))(not (add_holding_put-down_var1 )))
  :effect (and (when (pre_holding_put-down_var1 )(del_holding_put-down_var1 ))
  (when (not (pre_holding_put-down_var1 ))(add_holding_put-down_var1 ))))

 (:action program_pre_on_stack_var1_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_stack_var1_var1 ))(not (del_on_stack_var1_var1 ))
  (not (add_on_stack_var1_var1 )))
   :effect (and (pre_on_stack_var1_var1 )))

 (:action program_eff_on_stack_var1_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_on_stack_var1_var1 ))(not (add_on_stack_var1_var1 )))
  :effect (and (when (pre_on_stack_var1_var1 )(del_on_stack_var1_var1 ))
  (when (not (pre_on_stack_var1_var1 ))(add_on_stack_var1_var1 ))))

 (:action program_pre_on_stack_var1_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_stack_var1_var2 ))(not (del_on_stack_var1_var2 ))
  (not (add_on_stack_var1_var2 )))
   :effect (and (pre_on_stack_var1_var2 )))

 (:action program_eff_on_stack_var1_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_on_stack_var1_var2 ))(not (add_on_stack_var1_var2 )))
  :effect (and (when (pre_on_stack_var1_var2 )(del_on_stack_var1_var2 ))
  (when (not (pre_on_stack_var1_var2 ))(add_on_stack_var1_var2 ))))

 (:action program_pre_on_stack_var2_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_stack_var2_var1 ))(not (del_on_stack_var2_var1 ))
  (not (add_on_stack_var2_var1 )))
   :effect (and (pre_on_stack_var2_var1 )))

 (:action program_eff_on_stack_var2_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_on_stack_var2_var1 ))(not (add_on_stack_var2_var1 )))
  :effect (and (when (pre_on_stack_var2_var1 )(del_on_stack_var2_var1 ))
  (when (not (pre_on_stack_var2_var1 ))(add_on_stack_var2_var1 ))))

 (:action program_pre_on_stack_var2_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_stack_var2_var2 ))(not (del_on_stack_var2_var2 ))
  (not (add_on_stack_var2_var2 )))
   :effect (and (pre_on_stack_var2_var2 )))

 (:action program_eff_on_stack_var2_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_on_stack_var2_var2 ))(not (add_on_stack_var2_var2 )))
  :effect (and (when (pre_on_stack_var2_var2 )(del_on_stack_var2_var2 ))
  (when (not (pre_on_stack_var2_var2 ))(add_on_stack_var2_var2 ))))

 (:action program_pre_ontable_stack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_stack_var1 ))(not (del_ontable_stack_var1 ))
  (not (add_ontable_stack_var1 )))
   :effect (and (pre_ontable_stack_var1 )))

 (:action program_eff_ontable_stack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_stack_var1 ))(not (add_ontable_stack_var1 )))
  :effect (and (when (pre_ontable_stack_var1 )(del_ontable_stack_var1 ))
  (when (not (pre_ontable_stack_var1 ))(add_ontable_stack_var1 ))))

 (:action program_pre_ontable_stack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_stack_var2 ))(not (del_ontable_stack_var2 ))
  (not (add_ontable_stack_var2 )))
   :effect (and (pre_ontable_stack_var2 )))

 (:action program_eff_ontable_stack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_stack_var2 ))(not (add_ontable_stack_var2 )))
  :effect (and (when (pre_ontable_stack_var2 )(del_ontable_stack_var2 ))
  (when (not (pre_ontable_stack_var2 ))(add_ontable_stack_var2 ))))

 (:action program_pre_clear_stack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_stack_var1 ))(not (del_clear_stack_var1 ))
  (not (add_clear_stack_var1 )))
   :effect (and (pre_clear_stack_var1 )))

 (:action program_eff_clear_stack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_stack_var1 ))(not (add_clear_stack_var1 )))
  :effect (and (when (pre_clear_stack_var1 )(del_clear_stack_var1 ))
  (when (not (pre_clear_stack_var1 ))(add_clear_stack_var1 ))))

 (:action program_pre_clear_stack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_stack_var2 ))(not (del_clear_stack_var2 ))
  (not (add_clear_stack_var2 )))
   :effect (and (pre_clear_stack_var2 )))

 (:action program_eff_clear_stack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_stack_var2 ))(not (add_clear_stack_var2 )))
  :effect (and (when (pre_clear_stack_var2 )(del_clear_stack_var2 ))
  (when (not (pre_clear_stack_var2 ))(add_clear_stack_var2 ))))

 (:action program_pre_handempty_stack
   :parameters ()
  :precondition (and (modeProg )(not (pre_handempty_stack ))(not (del_handempty_stack ))
  (not (add_handempty_stack )))
   :effect (and (pre_handempty_stack )))

 (:action program_eff_handempty_stack
   :parameters ()
   :precondition (and (modeProg )(not (del_handempty_stack ))(not (add_handempty_stack )))
  :effect (and (when (pre_handempty_stack )(del_handempty_stack ))
  (when (not (pre_handempty_stack ))(add_handempty_stack ))))

 (:action program_pre_holding_stack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_stack_var1 ))(not (del_holding_stack_var1 ))
  (not (add_holding_stack_var1 )))
   :effect (and (pre_holding_stack_var1 )))

 (:action program_eff_holding_stack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_stack_var1 ))(not (add_holding_stack_var1 )))
  :effect (and (when (pre_holding_stack_var1 )(del_holding_stack_var1 ))
  (when (not (pre_holding_stack_var1 ))(add_holding_stack_var1 ))))

 (:action program_pre_holding_stack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_stack_var2 ))(not (del_holding_stack_var2 ))
  (not (add_holding_stack_var2 )))
   :effect (and (pre_holding_stack_var2 )))

 (:action program_eff_holding_stack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_stack_var2 ))(not (add_holding_stack_var2 )))
  :effect (and (when (pre_holding_stack_var2 )(del_holding_stack_var2 ))
  (when (not (pre_holding_stack_var2 ))(add_holding_stack_var2 ))))

 (:action program_pre_on_unstack_var1_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_unstack_var1_var1 ))(not (del_on_unstack_var1_var1 ))
  (not (add_on_unstack_var1_var1 )))
   :effect (and (pre_on_unstack_var1_var1 )))

 (:action program_eff_on_unstack_var1_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_on_unstack_var1_var1 ))(not (add_on_unstack_var1_var1 )))
  :effect (and (when (pre_on_unstack_var1_var1 )(del_on_unstack_var1_var1 ))
  (when (not (pre_on_unstack_var1_var1 ))(add_on_unstack_var1_var1 ))))

 (:action program_pre_on_unstack_var1_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_unstack_var1_var2 ))(not (del_on_unstack_var1_var2 ))
  (not (add_on_unstack_var1_var2 )))
   :effect (and (pre_on_unstack_var1_var2 )))

 (:action program_eff_on_unstack_var1_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_on_unstack_var1_var2 ))(not (add_on_unstack_var1_var2 )))
  :effect (and (when (pre_on_unstack_var1_var2 )(del_on_unstack_var1_var2 ))
  (when (not (pre_on_unstack_var1_var2 ))(add_on_unstack_var1_var2 ))))

 (:action program_pre_on_unstack_var2_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_unstack_var2_var1 ))(not (del_on_unstack_var2_var1 ))
  (not (add_on_unstack_var2_var1 )))
   :effect (and (pre_on_unstack_var2_var1 )))

 (:action program_eff_on_unstack_var2_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_on_unstack_var2_var1 ))(not (add_on_unstack_var2_var1 )))
  :effect (and (when (pre_on_unstack_var2_var1 )(del_on_unstack_var2_var1 ))
  (when (not (pre_on_unstack_var2_var1 ))(add_on_unstack_var2_var1 ))))

 (:action program_pre_on_unstack_var2_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_on_unstack_var2_var2 ))(not (del_on_unstack_var2_var2 ))
  (not (add_on_unstack_var2_var2 )))
   :effect (and (pre_on_unstack_var2_var2 )))

 (:action program_eff_on_unstack_var2_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_on_unstack_var2_var2 ))(not (add_on_unstack_var2_var2 )))
  :effect (and (when (pre_on_unstack_var2_var2 )(del_on_unstack_var2_var2 ))
  (when (not (pre_on_unstack_var2_var2 ))(add_on_unstack_var2_var2 ))))

 (:action program_pre_ontable_unstack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_unstack_var1 ))(not (del_ontable_unstack_var1 ))
  (not (add_ontable_unstack_var1 )))
   :effect (and (pre_ontable_unstack_var1 )))

 (:action program_eff_ontable_unstack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_unstack_var1 ))(not (add_ontable_unstack_var1 )))
  :effect (and (when (pre_ontable_unstack_var1 )(del_ontable_unstack_var1 ))
  (when (not (pre_ontable_unstack_var1 ))(add_ontable_unstack_var1 ))))

 (:action program_pre_ontable_unstack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_ontable_unstack_var2 ))(not (del_ontable_unstack_var2 ))
  (not (add_ontable_unstack_var2 )))
   :effect (and (pre_ontable_unstack_var2 )))

 (:action program_eff_ontable_unstack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_ontable_unstack_var2 ))(not (add_ontable_unstack_var2 )))
  :effect (and (when (pre_ontable_unstack_var2 )(del_ontable_unstack_var2 ))
  (when (not (pre_ontable_unstack_var2 ))(add_ontable_unstack_var2 ))))

 (:action program_pre_clear_unstack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_unstack_var1 ))(not (del_clear_unstack_var1 ))
  (not (add_clear_unstack_var1 )))
   :effect (and (pre_clear_unstack_var1 )))

 (:action program_eff_clear_unstack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_unstack_var1 ))(not (add_clear_unstack_var1 )))
  :effect (and (when (pre_clear_unstack_var1 )(del_clear_unstack_var1 ))
  (when (not (pre_clear_unstack_var1 ))(add_clear_unstack_var1 ))))

 (:action program_pre_clear_unstack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_clear_unstack_var2 ))(not (del_clear_unstack_var2 ))
  (not (add_clear_unstack_var2 )))
   :effect (and (pre_clear_unstack_var2 )))

 (:action program_eff_clear_unstack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_clear_unstack_var2 ))(not (add_clear_unstack_var2 )))
  :effect (and (when (pre_clear_unstack_var2 )(del_clear_unstack_var2 ))
  (when (not (pre_clear_unstack_var2 ))(add_clear_unstack_var2 ))))

 (:action program_pre_handempty_unstack
   :parameters ()
  :precondition (and (modeProg )(not (pre_handempty_unstack ))(not (del_handempty_unstack ))
  (not (add_handempty_unstack )))
   :effect (and (pre_handempty_unstack )))

 (:action program_eff_handempty_unstack
   :parameters ()
   :precondition (and (modeProg )(not (del_handempty_unstack ))(not (add_handempty_unstack )))
  :effect (and (when (pre_handempty_unstack )(del_handempty_unstack ))
  (when (not (pre_handempty_unstack ))(add_handempty_unstack ))))

 (:action program_pre_holding_unstack_var1
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_unstack_var1 ))(not (del_holding_unstack_var1 ))
  (not (add_holding_unstack_var1 )))
   :effect (and (pre_holding_unstack_var1 )))

 (:action program_eff_holding_unstack_var1
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_unstack_var1 ))(not (add_holding_unstack_var1 )))
  :effect (and (when (pre_holding_unstack_var1 )(del_holding_unstack_var1 ))
  (when (not (pre_holding_unstack_var1 ))(add_holding_unstack_var1 ))))

 (:action program_pre_holding_unstack_var2
   :parameters ()
  :precondition (and (modeProg )(not (pre_holding_unstack_var2 ))(not (del_holding_unstack_var2 ))
  (not (add_holding_unstack_var2 )))
   :effect (and (pre_holding_unstack_var2 )))

 (:action program_eff_holding_unstack_var2
   :parameters ()
   :precondition (and (modeProg )(not (del_holding_unstack_var2 ))(not (add_holding_unstack_var2 )))
  :effect (and (when (pre_holding_unstack_var2 )(del_holding_unstack_var2 ))
  (when (not (pre_holding_unstack_var2 ))(add_holding_unstack_var2 ))))

 (:action validate_1
   :parameters ()
   :precondition (and (modeProg ))
  :effect (and (not (modeProg ))(clear a)(not (clear b))(clear c)(not (clear d))(not (clear e))(not (clear f))
  (not (clear g))(handempty )(not (holding a))(not (holding b))(not (holding c))(not (holding d))
  (not (holding e))(not (holding f))(not (holding g))(not (on a a))(not (on a b))(not (on a c))(not (on a d))
  (not (on a e))(not (on a f))(on a g)(not (on b a))(not (on b b))(not (on b c))(not (on b d))(on b e)
  (not (on b f))(not (on b g))(not (on c a))(not (on c b))(not (on c c))(on c d)(not (on c e))(not (on c f))
  (not (on c g))(not (on d a))(on d b)(not (on d c))(not (on d d))(not (on d e))(not (on d f))(not (on d g))
  (not (on e a))(not (on e b))(not (on e c))(not (on e d))(not (on e e))(on e f)(not (on e g))(not (on f a))
  (not (on f b))(not (on f c))(not (on f d))(not (on f e))(not (on f f))(not (on f g))(not (on g a))
  (not (on g b))(not (on g c))(not (on g d))(not (on g e))(not (on g f))(not (on g g))(not (ontable a))
  (not (ontable b))(not (ontable c))(not (ontable d))(not (ontable e))(ontable f)(ontable g)(test1 )))

 (:action validate_2
   :parameters ()
  :precondition (and (not (modeProg ))(clear a)(clear b)(clear c)(clear d)(clear e)(not (clear f))(not (clear g))
  (handempty )(not (holding a))(not (holding b))(not (holding c))(not (holding d))(not (holding e))
  (not (holding f))(not (holding g))(not (on a a))(not (on a b))(not (on a c))(not (on a d))(not (on a e))
  (not (on a f))(on a g)(not (on b a))(not (on b b))(not (on b c))(not (on b d))(not (on b e))(not (on b f))
  (not (on b g))(not (on c a))(not (on c b))(not (on c c))(not (on c d))(not (on c e))(not (on c f))(not (on c g))
  (not (on d a))(not (on d b))(not (on d c))(not (on d d))(not (on d e))(not (on d f))(not (on d g))(not (on e a))
  (not (on e b))(not (on e c))(not (on e d))(not (on e e))(on e f)(not (on e g))(not (on f a))(not (on f b))
  (not (on f c))(not (on f d))(not (on f e))(not (on f f))(not (on f g))(not (on g a))(not (on g b))(not (on g c))
  (not (on g d))(not (on g e))(not (on g f))(not (on g g))(not (ontable a))(ontable b)(ontable c)(ontable d)
  (not (ontable e))(ontable f)(ontable g)(test1 ))
   :effect (and (test2 )(not (test1 ))))

 (:action validate_3
   :parameters ()
  :precondition (and (not (modeProg ))(clear a)(clear b)(clear c)(clear d)(clear e)(clear f)(clear g)(handempty )
  (not (holding a))(not (holding b))(not (holding c))(not (holding d))(not (holding e))(not (holding f))
  (not (holding g))(not (on a a))(not (on a b))(not (on a c))(not (on a d))(not (on a e))(not (on a f))
  (not (on a g))(not (on b a))(not (on b b))(not (on b c))(not (on b d))(not (on b e))(not (on b f))(not (on b g))
  (not (on c a))(not (on c b))(not (on c c))(not (on c d))(not (on c e))(not (on c f))(not (on c g))(not (on d a))
  (not (on d b))(not (on d c))(not (on d d))(not (on d e))(not (on d f))(not (on d g))(not (on e a))(not (on e b))
  (not (on e c))(not (on e d))(not (on e e))(not (on e f))(not (on e g))(not (on f a))(not (on f b))(not (on f c))
  (not (on f d))(not (on f e))(not (on f f))(not (on f g))(not (on g a))(not (on g b))(not (on g c))(not (on g d))
  (not (on g e))(not (on g f))(not (on g g))(ontable a)(ontable b)(ontable c)(ontable d)(ontable e)(ontable f)
  (ontable g)(test2 ))
   :effect (and (not (test2 ))(test3 )))

)

  \end{verbatim}
\end{scriptsize}

\begin{figure}[hbtp!]
\begin{scriptsize}  
  \begin{verbatim}
(define (problem learning_problem)
  (:domain blocks)
  (:objects g -  object c -  object d -  object e -  object a -  object b -  object f -  object )
  (:init (modeProg ) )
  (:goal (and (test3 ))))
  \end{verbatim}
\end{scriptsize}  
\caption{\small Compiled PDDL problem file for learning the blocksworld action models from two initial and final states.}
\label{fig:compiled-problem}
\end{figure}

### 2018-12-04 doc/aij18/major-revision/introduction.tex

@@
 
-There is common agreement in the planning community that the unavailability of a complete domain model is a bottleneck in the applicability of planning technology to many real-world domains~\cite{kambhampati:modellite:AAAI2007}.
+There is common agreement in the planning community that the unavailability of an \textcolor[rgb]{1.00,0.00,0.00}{adequate} domain model is a bottleneck in the applicability of planning technology to many real-world domains~\cite{kambhampati:modellite:AAAI2007}.
 Motivated by the difficulty and cost of crafting action models, research in action-model learning has seen huge advances.
@@
 
-A system for learning planning action models receives as an input observations of the agent's plan execution and outputs an abstract version of the capability model that reflects the physics of the real-world domain.
+A system for learning planning action models receives as an input observations of the agent's plan execution and \textcolor[rgb]{1.00,0.00,0.00}{outputs an approximation of the actions that embody the physics of the real-world domain being modeled}.
 The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards.
@@
 
-Two different types of data are generally identified in the generation of explanations~\cite{Sheh17,Roberts18}: (1) data that embody enough information to fully trace the steps of the decision-making process of the agent and therefore accurate and detailed explanations can be generated; (2) data that do not accurately represent the decision-making process because the observed external behaviour of the agent responds to a black-box model.
-A similar distinction can be established with the type of observations that algorithms use to learn planning action models.
+\textcolor[rgb]{1.00,0.00,0.00}{Providing an effective explanation of the decision-making process to humans and other agents depends, among other factors, on the type of data that are used to generate the explanation.
+Roughly speaking, two different types of data are identified~\cite{Sheh17,Roberts18}: data based on \emph{introspection} into the underlying model that provide enough information to fully trace the decision making process, and data that are based on observing the external behavior of the agent as a black box.
+In the latter case, the agent builds \emph{post-hoc rationalizations} that explain a decision without accurately representing the underlying decision making processes.} A similar distinction can be established with the type of observations that algorithms use to learn planning action models.
 Up to date, all existing learning approaches assume that a given input plan trace contains a fully observed sequence of the executed actions.
@@
 
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behaviour is not available.
-Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimal input knowledge.
+\textcolor[rgb]{1.00,0.00,0.00}{Motivated by recent advances on the use of classical planning for the generation of different types of planning models (regular automata, context-free grammars, finite-state machines, \strips)} \cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behavior is not available.
+Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimal observability.
 
-While current learning systems accept varying degrees of observability in the states traversed by the input observed actions, none of them allow partial observability in the sequence of executed actions.
-\FAMA, on the other hand, allows incomplete or empty action sequences.
-This way, the minimum observability case acceptable by \FAMA is when the algorithm is only fed with the input and output of a plan trace, namely the initial and final state of a planning task.
-Like many Machine Learning (ML) techniques, \FAMA is able to operate with only input/output pairs and an unknown or a partially known model of the agent.
+\textcolor[rgb]{1.00,0.00,0.00}{All current learning approaches require as input information a plan trace that includes the complete sequence of actions executed by the agent; i.e, all the actions performed by the agent must be observable.
+Additionally, the existing systems accept a varying degree of observability of the states traversed by these actions, which ranges from fully observable to fully unobservable states (see section \ref{sec:background} for details)}.
+In contrast, \FAMA allows for an incomplete or empty sequence of observable actions, and the minimum observability case acceptable by \FAMA is when the algorithm is only fed with the initial and final state of a plan trace.
+Like many Machine Learning (ML) techniques, \FAMA is able to operate with only input/output pairs of states and an unknown or a partially known model of the agent.
 Unlike ML algorithms, \FAMA requires a symbolic structured representation of the input knowledge.
@@
 
-\FAMA is a solving schema, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}.
-\FAMA is thereby a model-based approach that automatically builds its own planning model by logical inference from the input observations.
-A solution to this planning task is a sequence of actions that determines the preconditions and effects of the \strips\ action models.
-The construct of the planning model starts out with the input observations of the agent's behaviour that, as mentioned above, may comprise none of the actions executed by the agent.
-\FAMA can thus be regarded as a solving process that generates an instance of a planning model, and whose solution must be \emph{compliant} with the input plan traces.
+\textcolor[rgb]{1.00,0.00,0.00}{\FAMA is a new learning approach, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}.
+The construction of a \strips\ planning model starts out from a set of plan traces containing the observation of several plan executions.
+As mentioned above, a plan trace may comprise none of the actions executed by the agent but must include, at least, the initial state $s_0$ and final state $s_f$ of the execution.
+The compilation scheme lies in defining a planning task from the set of plan traces using a set of \emph{programming actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces.
+A solution to this planning task is thus a plan that determines the preconditions and effects of a \strips\ planning model $M$.
+The rationale of this process draws its strength in that when the problem $\tup{s_0,s_f}$ is solved with the learned model $M$, a plan $\pi$ is generated such that $\pi$ will contain the observed actions of the plan trace, if any, and the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.
+In other words, the plan $\pi$ generated when solving the problem $\tup{s_0,s_f}$ with $M$ is \emph{compliant} with the plan trace.
+}
+
+Hence, \FAMA is a model-based approach that automatically builds its own planning model by logical inference from the input plan traces that contain the observations of the agent execution.
 This behaviour largely differs from ML techniques, which aim to minimize an error function on the training data.
-Moreover, \FAMA requires far less sample data (input observations) than typical ML algorithms, thus alleviating the dependency on the assumption that there are enough data for learning the action models~\cite{Zhuo15}.
+Moreover, \FAMA requires far less sample data (example plan traces) than typical ML algorithms, thus alleviating the dependency on the assumption that there are enough data for learning the action models~\cite{Zhuo15}.
 
@@
 The most common method is to use a syntax-based evaluation that compares the learned model with a reference model.
-\FAMA instead proposes two novel semantic evaluation metrics that build upon two well-known ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate the learned action models with respect to observations of plan executions.
+\FAMA proposes instead two novel semantic evaluation metrics that build upon two well-known ML metrics, {\em precision} and {\em recall}~\cite{davis2006relationship}, to evaluate the learned action models with respect to observations of plan executions.
 Our semantic evaluation is generally more informative than counting the number of errors between two models and alleviates two important limitations of a purely syntax-based assessment: (a) that the learned model is syntactically different from the reference model but semantically correct and (b) that the learned model comprises correct though unnecessary preconditions in regards to the reference model.
@@
 Our semantic evaluation method is built on the same compilation scheme for solving a learning task.
-In particular, \FAMA also accepts as an input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}, alongside the observations.
-In this case, \FAMA returns a model that follows the input model and is compliant with the observations.
-We designed an {\em edition} mechanism that serves to correct the input model to the output model, which in turn defines an assessment of the accuracy with which the input model explains the observations.
-Interpreting the edition measure as a distance-based concept between two models could be exploited in model reconciliation~\cite{KulkarniCZVZK16}.
+In particular, \FAMA also accepts an input initial action model $M$ of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}, alongside the observation of the agent execution.
+In this case, \FAMA returns a model $M'$ that follows the input model $M$ and is compliant with the observations.
+We designed an {\em edition} mechanism that serves to correct the input model to the output model, which in turn defines an assessment of the accuracy with which $M$ explains the observations.
+Interpreting the edition measure as a distance-based concept between two models can also be exploitable in model reconciliation~\cite{KulkarniCZVZK16}.
 
@@
 
-All in all, \FAMA is a planning-based solving scheme that outputs a \strips\ action model using a planning model that is automatically built from minimal input knowledge.
+%All in all, \FAMA is a planning-based solving scheme that outputs a \strips\ action model using a planning model that is automatically built from minimal input knowledge.
 Unlike extensive-data ML approaches, \FAMA only requires a small amount of input plan traces.
 Unlike most relevant action-model learning algorithms, \FAMA does not require the traces to contain any observed action executed by the agent.
+
+\textcolor[rgb]{1.00,0.00,0.00}{In summary, \FAMA is a novel learning approach characterized by:
+\begin{itemize}
+\item it outputs a \strips\ action model building automatically a planning task from a set of input plan traces
+\item plan traces are correct (no noise is considered in the observations) but may be incomplete
+\item the task of learning the action model is compiled to a planning task which specifies actions for programming the components of the \strips\ actions and actions for validating the learned \strips\ actions
+\item a semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model
+\end{itemize} }
+
 Ultimately, \FAMA adopts certain ML-like characteristics, like the ability to learn from data that do not explicitly exhibit the agent's behaviour or the capacity of correcting an initially given action model.

### 2018-12-04 doc/icaps-19/model-recognition.tex

@@
 
+\author{\#92}
 % Commented for blind submission
@@
 {\em Plan recognition as planning} is the model-based approach for this task that leverages the action model of the observed agent to compute its most likely goal and predict its future actions~\cite{ramirez2009plan,ramirez2012plan}.
-While action models are learnable from observations of the agent~\cite{yang2007learning,cresswell2013acquiring,aineto2018learning}, in real-world applications developing in a particular context, approximate behavioral models of the involved agents (manually-introduced or automatically-learned) are already available, so there is rarely a need to learn a model from scratch.
+While action models are learnable from observations of the agent~\cite{yang2007learning,cresswell2013acquiring,aineto2018learning}, in real-world applications developing in a particular context, approximate behavioral models (manually-introduced or
+automatically-learned) of the involved agents are usually already available, so there is rarely a need to learn a model from scratch.
 A more relevant and applicable task in these scenarios is thereby to predict the behavior of an agent from a set of given models.
@@
 Finally, the robustness of the approach is shown in the third experiment where actions and intermediates states are unobservable.
-A meaningful lesson from the experiments is that our approach allows us to predict the future actions of agents whose behavior is ruled by a  model other than a deliberative planning model aimed at achieving some goals; this includes, for instance, models of fixed programs or finite state machines (regular automata).
+A meaningful lesson from the experiments is that our approach allows us to predict the future actions of agents whose behavior is ruled by a model other than a goal-driven deliberative planning model; this includes, for instance, models of fixed programs or finite state machines (regular automata).
 

### 2018-12-09 doc/aij18/major-revision/background.tex
Sections 1 and 2 updated

--- 2018-07-16 doc/aij18/background.tex
@@
 
-This section serves two purposes: first we introduce basic planning concepts as well as the classical planning model that we will use throughout the rest of the paper; and secondly, we summarize the existing approaches to learn classical planning action models and highlight our contributions comparing with related work.
+This section serves two purposes; first, we introduce basic planning concepts and define the classical planning model we aim to learn; secondly, we summarize the most relevant existing approaches to learn classical planning action models.
 
@@
 A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent).
-We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
+\textcolor[rgb]{1.00,0.00,0.00}{The complement of $L$ is defined as $\neg L=\{\neg l:l\in L\}$.} We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
 
-A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states; i.e.
+\textcolor[rgb]{1.00,0.00,0.00}{In \FAMA, we need to implicitly represent the unobserved literals of states.
+In order to do so, we will adopt the \emph{open world assumption}; i.e., what is not known to be true in a state is unknown.
+Consequently, states will explicitly include positive literals ($f$) and negative literals ($\neg f$) such that literals that are not in a state are unknown or unobserved.} Hence, a {\em state} $s$ is a full assignment of values to fluents; i.e.
 $|s|=|F|$, so the size of the state space is $2^{|F|}$.
@@
 
-%simplifies subsequent definitions but often, we will abuse notation by defining a state $s$ only in terms of the fluents that are true in $s$, as it is common in \strips\ planning.
+A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions.
+An action $a\in A$ has a set of preconditions $\pre(a)\in\mathcal{L}(F)$ and a set of effects $\eff(a)\in\mathcal{L}(F)$.
+An action $a\in A$ is applicable in a given state $s$ iff $pre(a)\subseteq s$, i.e.~if the literals $pre(a)$ hold in $s$.
+\textcolor[rgb]{1.00,0.00,0.00}{The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$.
+Note that subtracting the complement of $\eff(a)$ from $s$ ensures that $\theta(s,a)$ remains a well-defined state with positive and negative literals.
+Then:}
 
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of actions. An action $a\in A$ is defined with:
 \begin{itemize}
-\item $\pre(a)\in\mathcal{L}(F)$, the {\em preconditions} of $a$, is the set of literals that must hold for the action $a\in A$ to be applicable.
-\item $\eff^+(a)\in\mathcal{L}(F)$, the {\em positive effects} of $a$, is the set of literals that are true after the application of the action $a\in A$.
-\item $\eff^-(a)\in\mathcal{L}(F)$, the {\em negative effects} of $a$, is the set of literals that are false after the application of the action.
+\item \textcolor[rgb]{1.00,0.00,0.00}{$\eff^+(a)\in\mathcal{L}(F)$ is the {\em positive effects} of $a$, the subset of action effects that assert a positive literal in the state resulting after the application of $a$}
+\item \textcolor[rgb]{1.00,0.00,0.00}{$\eff^-(a)\in\mathcal{L}(F)$ is the {\em negative effects} of $a$, the subset of action effects that assert a negative literal in the state resulting after the application of $a$}
 \end{itemize}
-We assume that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$ and that actions $a\in A$ are instantiated from given action schemas, as in PDDL.
-We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
+
+\textcolor[rgb]{1.00,0.00,0.00}{Since we restrict our attention to \strips\ action models learning, we will assume the set of syntactic constraints imposed by \strips\ models, namely }that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$.
+Additionally, actions $a\in A$ are instantiated from given action schemas, as in PDDL.
+
+%We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
 The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
@@
 
-Our approach copes with the partial observability of the plan execution adopting the \emph{open world assumption}.
-With regard to the observed states in a plan trace, we say that $\tup{s_0, s_1, \ldots, s_n}$ is a fully-observable (\FO) state trajectory if every state $s_i$ is a full assignment of values to fluents and the minimal action sequence to transit from state $s_i$ to state $s_{i+1}$ is composed of a single action; that is, $\theta(s_i,\tup{a})=s_{i+1}$.
-Otherwise, it is a partially-observable (\PO) state trajectory, meaning that at least one state $s_i$ is a partial assignment of values to fluents in which one or more literals are missing (formally, $|s_i|<|F|$).
-This implies that in a \PO state trajectory all fluents of a state $s_i$ might be missing, in which case, $s_i$ is a \emph{missing} or \emph{empty} state.
-This general definition of \PO gives rise to two particular cases:
+Our approach copes with partial observability in the plan traces.
+Let $\pi=\tup{a_1, \ldots, a_n}$ be the plan executed by an agent that induces the state trajectory $s=\tup{s_0, s_1, \ldots, s_n}$, and let $\tau = \langle s_0, \ldots, a_i, \ldots, s_j, \ldots, s_n \rangle$ be the plan trace observed from the plan execution.
+With regards to the observed states of $\tau$, that we will refer to as $\tau_s$, we identify two general cases of observability:
 
-\begin{itemize}
-\item when \textbf{all} the $n-1$ intermediate states of a trajectory $s$ are \textbf{missing}, $s$ is a \emph{non-observable} (\NO) state trajectory.
-\item when \textbf{none} of the $n-1$ intermediate states of a trajectory $s$ is \textbf{missing}, we will refer to $s$ as a \POstar state trajectory.
-\end{itemize}
+\begin{enumerate}
+\item we say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and \textcolor[rgb]{1.00,0.00,0.00}{there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$};  that is $\theta(s_i,\tup{a})=s_{i+1}$.
+\textcolor[rgb]{1.00,0.00,0.00}{This case clearly states that $\tau_s = s$, meaning that $\forall s_i \in \tau_s$, $s_i$ comprises all the literals of the corresponding state in the trajectory $s$ of the plan $\pi$.
+Formally, $\forall i, 1 \leq i  < n, |s_i| = |F|$.}
+\item we say that $\tau_s$ is a partially-observable (\PO) state trajectory if at least one intermediate state of $\tau_s$ is a partial assignment of values to fluents.
+\textcolor[rgb]{1.00,0.00,0.00}{Formally, $\exists i, 1 \leq i  < n, |s_i| < |F|$.} This means that one or more literals are missing in the intermediate $s_i$, all of which may be missing.
+When all literals are missing, $s_i$ is a \emph{missing} or \emph{empty state} ($s_i = \emptyset$).
+\end{enumerate}
+
+The general definition of a \PO state trajectory gives rise to two special cases:
+
+\begin{enumerate}
+\item when \textbf{all} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, $\tau_s$ is a \emph{non-observable} (\NO) state trajectory.
+\textcolor[rgb]{1.00,0.00,0.00}{Formally, $\forall i, 1 \leq i < n,  s_i= \emptyset$; i.e., $|s_i| = 0$.}
+\item when \textbf{none} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, we will refer to $\tau_s$ as a \POstar state trajectory.
+\textcolor[rgb]{1.00,0.00,0.00}{Formally, $\exists i, 1 \leq i < n, s_i \neq \emptyset$; i.e., $0 < |s_i| < |F|$.}
+\end{enumerate}
 
@@
 	     & {\bf \# intermediate states} & {\bf state type} \\ \hline
-    \FO & $n-1$  & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ is a full assignment \\ \hline
-    \multirow{1}{*}{\POstar} & $n-1$ & {\small $\exists i, 1 \leq i < n$}  \\ & & $s_i$ is a partial assignment \\ \hline
-    \multirow{1}{*}{\PO} & $\leq n-1$ & {\small $\exists i, 1 \leq i < n$}   \\  & & $s_i$ is a partial assignment \\ \hline
-    \NO & 0 & {\small $\forall i, 1 \leq i < n$}  \\  & & $|s_i|=0$
+    \FO & $n-1$  & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ is a full assignment \textcolor[rgb]{1.00,0.00,0.00}{$|s_i|=|F|$} \\ \hline
+    \multirow{1}{*}{\POstar} & $n-1$ & {\small $\exists i, 1 \leq i < n$}  \\ & & $s_i$ is a partial assignment \textcolor[rgb]{1.00,0.00,0.00}{$0 < |s_i|< |F|$}\\ \hline
+    \multirow{1}{*}{\PO} & $\leq n-1$ & {\small $\exists i, 1 \leq i < n$}   \\  & & $s_i$ is a partial assignment \textcolor[rgb]{1.00,0.00,0.00}{$|s_i|< |F|$}\\ \hline
+    \NO & 0 & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ \textcolor[rgb]{1.00,0.00,0.00}{is an empty state  $|s_i|=0$}
 \end{tabular}
@@
 
-With regard to the observed actions in a plan trace, we say that $\tup{a_1, \ldots, a_n}$ is a \FO action sequence if it contains all the necessary actions to transit every state $s_{i-1}$ to the corresponding successor state $s_{i}$, for each {\small $1\leq i\le n$};  we say it is a \PO action sequence if at least one of these necessary actions is missing in the sequence, and a \NO action sequence when the sequence of observed actions is empty.
+\textcolor[rgb]{1.00,0.00,0.00}{\FAMA can also deal with partial observability in the observed actions of $\tau$, that we will refer to as $\tau_a$.
+We identify three levels of observability, from the greatest to the lowest:}
+
+\begin{enumerate}
+\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$.
+In this case, $\tau_a$ contains all the necessary actions to transit every state $s_{i-1}$ to its corresponding successor state $s_{i}$, from $s_0$ to $s_n$.
+This is the type of input trace accepted by all the existing learning approaches (see section \ref{related_work} for details)}.
+\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence.
+In this case, at least one of the necessary actions of the plan $\pi$ is missing in $\tau_a$.
+Formally, $\exists i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.}
+\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence.
+Formally, $\forall i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.
+That is, $\tau_a = \emptyset$.}
+\end{enumerate}
+
+================
 
 A plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$ holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}.
+
+NO TENGO CLARO LO QUE QUEREMOS DECIR CON ESTA FRASE (??).
+ADEMAS, NO ES CORRECTA PORQUE LOS ESTADOS DE UNA TRAZA PUEDEN SER FULL-OBSERVABLE Y $\mathcal{L}(F)$ SON PARTIAL ASSIGNMENTS OF VALUES TO FLUENTS.
+
+=================
+
 Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
+\textcolor[rgb]{1.00,0.00,0.00}{In section \ref{overview}, we expose the impact of all combinations of observed state trajectories and observed action sequences in the complexity of the learning task.}
 
@@
 In this section we summarize the most recent and relevant approaches to learning action models found in the literature.
-Approaches will be examined according to the following parameters: the input knowledge (plan traces) accepted by the system, the expressiveness of the learned action model and the principal technique used for learning the action model (Table \ref{table:models_comparison1}), as well as the characteristics of the evaluation method to validate the learned models (Table \ref{table:models_comparison2}).
+Approaches will be examined according to the following parameters: the observability of the plan traces accepted by the system, the expressiveness of the learned action model and the principal technique used for learning the action model (Table \ref{table:models_comparison1}), as well as the characteristics of the evaluation method used to validate the learned models (Table \ref{table:models_comparison2}).
 
@@
 The task of learning from less constrained traces subsumes learning from more constrained ones.
-Consequently, approaches to learning from, for instance traces with \PO state trajectories, will also be able to learn from traces with \POstar state trajectories.
-All the approaches analyzed in this work accept the more constrained definition of partial observations of intermediate states \POstar, and most of them also allow the sequence of intermediate states to be empty.
-Exceptionally, a \NO state sequence in \LOCM is a fully-empty trajectory, with neither initial or final state.
+Consequently, approaches to learning from, say traces with \PO state trajectories, will also enable learning from traces with \POstar state trajectories.
+All the approaches analyzed in this work accept the more constrained definition of partial observations of intermediate states \POstar, two of them also allow the sequence of intermediate states to be empty (\PO) and the majority accept \NO state trajectories.
+\textcolor[rgb]{1.00,0.00,0.00}{Exceptionally, \LOCM is the only approach capable of learning from a fully-empty state trajectory, with neither initial nor final state.}
 
@@
          \hline
-         \LOCMtwo & \NO states &  predicates and types & Finite State Machines \\ & \FO actions & & \\
+         \textcolor[rgb]{1.00,0.00,0.00}{\LOUGA} & \textcolor[rgb]{1.00,0.00,0.00}{\NO states} & \textcolor[rgb]{1.00,0.00,0.00}{\strips} & \textcolor[rgb]{1.00,0.00,0.00}{Genetic algorithm} \\ & \textcolor[rgb]{1.00,0.00,0.00}{\FO actions} & \textcolor[rgb]{1.00,0.00,0.00}{negative preconditions} & \\
+         \hline
+         \LOCMtwo & \textcolor[rgb]{1.00,0.00,0.00}{--- }&  predicates and types & Finite State Machines \\ & \FO actions & & \\
          \hline
@@
          \hline
+       \textcolor{red}{\LOUGA} & \textcolor[rgb]{1.00,0.00,0.00}{cross-validation with} & \textcolor[rgb]{1.00,0.00,0.00}{redundant effects} & \textcolor[rgb]{1.00,0.00,0.00}{5} \\  & \textcolor[rgb]{1.00,0.00,0.00}{a test set of plan traces} & \textcolor[rgb]{1.00,0.00,0.00}{differences wrt the test set} & \textcolor[rgb]{1.00,0.00,0.00}{800 - 3200 actions (160 traces)}\\
+         \hline
 		\LOCMtwo & manual checking wrt GTM &  ---  &   --- \\
          \hline
-		\FAMA & checking wrt GTM  & precision and recall & 15\\ & &  & 25 actions\\
+		\FAMA & checking wrt GTM  & precision and recall & 14\\  & validation with a test set &  & 20-50 actions\\
          \hline
@@
 
-The Learning Object-Centred Models (\textbf{\LOCM}) is an approach that only requires the \FO action sequence as input knowledge, without need for providing any information about the predicates or the state trajectory, not even the initial or final state~\cite{CresswellMW09,cresswell2013acquiring}.
+\textcolor[rgb]{1.00,0.00,0.00}{One of the latest incorporations to the family of action model learning algorithms is \LOUGA.
+This system uses a genetic algorithm to learn the effects of actions.
+In order to do this, each gene in the genome encodes whether a predicate is a positive effect, negative effect or none for a particular action, and the fitness of an individual is evaluated by reproducing the trace with the model encoded in the individual.
+After a solution for the effects is found, an ad-hoc algorithm is used to infer preconditions by finding those literals that are always present before the execution of an action.
+\LOUGA evaluates the learned models via cross-validation using the same metrics they use in their fitness function.
+In more detail, they measure (1) redundant positive and negative effects, (2) preconditions not met, and (3) literals observed in the input trace but not in the corresponding execution of the plan with the learned model.}
+
+\vspace{0.15cm}
+
+\textcolor[rgb]{1.00,0.00,0.00}{Finally, we present the Learning Object-Centred Models (\textbf{\LOCM}), possibly the most distinctive learning system due to its ability of learning with minimal input knowledge.} \textbf{\LOCM} only requires the \FO action sequence of the plan trace, without need for providing any information about the predicates or the state trajectory, not even the initial or final state~\cite{CresswellMW09,cresswell2013acquiring}.
 The lack of available state information is overcome by exploiting assumptions about the structure of the actions.
@@
 \LOP applies a post-processing step after the \LOCM analysis and it requires additional input information, particularly a set of optimal plans besides the suboptimal \FO action sequences.
+
+\textcolor[rgb]{1.00,0.00,0.00}{The distinctive feature of \LOCMtwo lies in the need to learn the state variables (fluents) because predicates are not either provided as input nor they are deducible from the plan trace as no state observability is allowed.
+In contrast, \FAMA and the rest of approaches either assume the set of predicates are provided alongside the input traces or assume they are extractable from the observed states of the plan trace, in which case the plan trace must comprise at least a grounded sample of every predicate.
+Similarly, the syntax of an action header (the action name and its parameters) is either extractable from the action sequence of the plan trace or it must be explicitly provided.}
 

### 2018-12-09 doc/aij18/major-revision/introduction.tex
Sections 1 and 2 updated

@@
 
-A system for learning planning action models receives as an input observations of the agent's plan execution and \textcolor[rgb]{1.00,0.00,0.00}{outputs an approximation of the actions that embody the physics of the real-world domain being modeled}.
-The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards.
-Recently, model-based representations are also becoming popular in explainable AI planning as they form a common basis for communicating with users and facilitate the generation of transparent and explainable decisions~\cite{FoxLM17} as well as explanations in terms of the differences with a human mental model~\cite{ChakrabortiSK18}.
+A system for learning planning action models receives as an input observations of the agent's plan execution and \textcolor[rgb]{1.00,0.00,0.00}{outputs an approximation of the actions that embody the physics of the real-world domain being modeled.} The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards, \textcolor[rgb]{1.00,0.00,0.00}{but there exists as well a large variety of planning-related tasks that rely upon the existence of a planning model.
+Among these tasks, we might cite: \emph{goal and plan recognition} approaches based on a domain theory~\cite{ramirez2009plan,ramirez2012plan,SohrabiRU16}; \emph{transparent planning}, in which an agent implicitly communicates its true goal by making its intentions and its action selection transparent (recognizable) to observers~\cite{MacNallyLRP18}; or \emph{deceptive path-planning}, which draws on the definition of path-planning domains and aims at finding a path such that the probability of an observer identifying the final destination is minimised~\cite{MastersS17}.
+Planning models are also used in \emph{explainable AI planning} to form a common basis for communicating with users and facilitate the generation of transparent and explainable decisions~\cite{FoxLM17} as well as explanations in terms of the differences with a human mental model~\cite{ChakrabortiSK18}.
+\emph{Counterplanning} requires a model of the opponent agent in order to recognize its goals \cite{PozancoEFB18} and \emph{model reconciliation} aims to conform the models of two agents with respect to an observation of a plan computed with one of the two models~\cite{ChakrabortiSZK17}.}
 
-\textcolor[rgb]{1.00,0.00,0.00}{Providing an effective explanation of the decision-making process to humans and other agents depends, among other factors, on the type of data that are used to generate the explanation.
-Roughly speaking, two different types of data are identified~\cite{Sheh17,Roberts18}: data based on \emph{introspection} into the underlying model that provide enough information to fully trace the decision making process, and data that are based on observing the external behavior of the agent as a black box.
-In the latter case, the agent builds \emph{post-hoc rationalizations} that explain a decision without accurately representing the underlying decision making processes.} A similar distinction can be established with the type of observations that algorithms use to learn planning action models.
-Up to date, all existing learning approaches assume that a given input plan trace contains a fully observed sequence of the executed actions.
+\textcolor[rgb]{1.00,0.00,0.00}{Motivated by the requirement for a planning model in many different tasks and the recent advances on the use of classical planning for the generation of different types of planning models (regular automata, context-free grammars, finite-state machines, \strips)} \cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behavior is not available.
+Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimal observability.
+
+%Up to date, all existing learning approaches assume that a given input plan trace contains a fully observed sequence of the executed actions.
 This heavily restricts the applicability of the learning approach to contexts in which the behaviour of the agent is fully observable and a human annotator correctly labels the executed actions.
 
-\textcolor[rgb]{1.00,0.00,0.00}{Motivated by recent advances on the use of classical planning for the generation of different types of planning models (regular automata, context-free grammars, finite-state machines, \strips)} \cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behavior is not available.
-Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimal observability.
-
-\textcolor[rgb]{1.00,0.00,0.00}{All current learning approaches require as input information a plan trace that includes the complete sequence of actions executed by the agent; i.e, all the actions performed by the agent must be observable.
-Additionally, the existing systems accept a varying degree of observability of the states traversed by these actions, which ranges from fully observable to fully unobservable states (see section \ref{sec:background} for details)}.
-In contrast, \FAMA allows for an incomplete or empty sequence of observable actions, and the minimum observability case acceptable by \FAMA is when the algorithm is only fed with the initial and final state of a plan trace.
+\textcolor[rgb]{1.00,0.00,0.00}{Current learning approaches assume that the observation of the agent's plan execution (plan trace) encompasses the fully observed sequence of the executed actions; i.e, they assume all the actions performed by the agent are observable.
+This heavily restricts the applicability of the learning approaches to contexts where the behaviour of the agent is fully observable and a human annotator correctly labels the executed actions.
+On the other hand, learning approaches accept a varying degree of observability in the states traversed in the plan trace, ranging from fully observable to fully unobservable states (see section \ref{related_work} for details).} In contrast, \FAMA allows for an incomplete or empty sequence of observable actions, and the minimum observability case acceptable by \FAMA is when the algorithm is only fed with the initial and final state of a plan trace.
 Like many Machine Learning (ML) techniques, \FAMA is able to operate with only input/output pairs of states and an unknown or a partially known model of the agent.
@@
 The compilation scheme lies in defining a planning task from the set of plan traces using a set of \emph{programming actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces.
-A solution to this planning task is thus a plan that determines the preconditions and effects of a \strips\ planning model $M$.
+Hence, a solution to this planning task is a plan that determines the preconditions and effects of a \strips\ planning model $M$.
 The rationale of this process draws its strength in that when the problem $\tup{s_0,s_f}$ is solved with the learned model $M$, a plan $\pi$ is generated such that $\pi$ will contain the observed actions of the plan trace, if any, and the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.
@@
 
-Hence, \FAMA is a model-based approach that automatically builds its own planning model by logical inference from the input plan traces that contain the observations of the agent execution.
+\FAMA is thus a model-based approach that automatically builds its own planning model by logical inference from the input plan traces that contain the observations of the agent execution.
 This behaviour largely differs from ML techniques, which aim to minimize an error function on the training data.
@@
 \begin{itemize}
-\item it outputs a \strips\ action model building automatically a planning task from a set of input plan traces
-\item plan traces are correct (no noise is considered in the observations) but may be incomplete
-\item the task of learning the action model is compiled to a planning task which specifies actions for programming the components of the \strips\ actions and actions for validating the learned \strips\ actions
+\item compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces
+\item the plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states
+\item the planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions
 \item a semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model

### 2018-12-09 doc/aij18/major-revision/motivation.tex
Sections 1 and 2 updated

--- 2018-08-05 doc/aij18/motivation.tex
@@
-\section{Motivation}
-\label{sec:motivation}
+\section{Overview of \FAMA}
+\label{overview}
+
+\subsection{Problem definition}
+\label{probkem_definition}
+
+\subsection{Complexity of the learning tasks: a justification for using planning}
+\label{complexity}
 
@@
 Otherwise, the learning task is SAT compilable, which is known to be a NP-complete task~\cite{russell2016artificial}.
-This is the reason that SAT solving is a common technique in the approaches presented in section \ref{sec:background}.
+This is the reason why SAT solving is a common technique in the approaches presented in section \ref{sec:background}.
 

### 2018-12-17 doc/aij18/major-revision/evaluation.tex
Section 5 revised

--- 2018-07-24 doc/aij18/evaluation.tex
@@
 In this section we introduce the metrics used by \FAMA to evaluate the action models that result from solving a learning task $\Lambda$.
-First, we will describe two standard syntactic metrics ({\em precision} and {\em recall}) and then section \ref{semantic_precision_recall} will define a semantic evaluation measure that builds upon {\em precision} and {\em recall}.
+\textcolor{red}{First, we will motivate the uso of two standard syntactic metrics ({\em precision} and {\em recall}).
+Then, section \ref{syntactic_precision_recall} will define these metrics for planning models, and section \ref{semantic_precision_recall} will introduce a semantic evaluation measure that builds upon {\em precision} and {\em recall}}.
 Finally, section \ref{edit_distance} explains how \FAMA computes our novel semantic-based metrics.
 
-\vspace{0.1cm}
+\textcolor{red}{We can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the the total number of all the possible preconditions and effects of an action model.
+This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM.
+The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects and thereby it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models.
+To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}.
+These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM:}
 
-When the planning reference model of the input observations (i.e., the GTM) is available, the quality of the learned action models is measurable using two well-studied syntax-based metrics, {\em precision} and {\em recall}, commonly used in tasks such as information retrieval and recommender systems~\cite{davis2006relationship}.
+%When the planning reference model of the input observations (i.e., the GTM) is available, the quality of the learned action models is measurable using two well-studied syntax-based metrics, {\em precision} and {\em recall}, commonly used in tasks such as information retrieval and recommender systems~\cite{davis2006relationship}.
 These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM.
@@
 
-Introducing semantic-based evaluation metrics can be justified on two grounds:
+\subsection{Syntactic-based precision and recall for planning models}
+\label{syntactic_precision_recall}
+
+The rationale behind the adaptation of precision and recall for planning models lies in counting the \emph{edit operations} that need to be applied in domain model $\mathcal{M}$ to transform it into the reference model $\mathcal{M'}$.
+Given a set of action models $\mathcal{M}$, the two allowed edit operations are:
+
+\begin{itemize}
+	\item {\em Deletion}. A fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ is removable from $\xi\in\mathcal{M}$.
+	\item {\em Insertion}. A fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ can be added to $\xi\in\mathcal{M}$.
+\end{itemize}
+
+We now provide formal definitions of $INS(\mathcal{M},\mathcal{M'})$ and $DEL(\mathcal{M},\mathcal{M'})$, the sets of insertions and deletions, respectively, that are needed to transform a domain model $\mathcal{M}$ into a new domain model $\mathcal{M'}$.
+
+\begin{mydefinition}
+	Let $PRE(\xi) = \underset{p \in pre(\xi)}{\bigcup} pre_p(\xi)$, $ADD(\xi) = \underset{p \in add(\xi)}{\bigcup} add_p(\xi)$, and $DEL(\xi) = \underset{p \in del(\xi)}{\bigcup} del_p(\xi)$ be the set of propositional fluents that represent preconditions, positive and negative effects of a given action model $\xi$.
+We define:
+	\begin{small}
+		\textcolor{red}{
+			\begin{align*}
+			INS(\mathcal{M}, \mathcal{M'})=&\bigcup\limits_{\xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t.\ name(\xi) = name(\xi')} 
+			PRE(\xi') \backslash PRE(\xi) \cup
+			ADD(\xi') \backslash ADD(\xi) \cup 
+			DEL(\xi') \backslash DEL(\xi)
+			\\
+			\\
+			DEL(\mathcal{M}, \mathcal{M'})=&\bigcup\limits_{\xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t.\ name(\xi) = name(\xi')}
+			PRE(\xi) \backslash PRE(\xi') \cup
+			ADD(\xi) \backslash ADD(\xi') \cup 
+			DEL(\xi) \backslash DEL(\xi')
+			\end{align*}
+		}
+	\end{small}
+\end{mydefinition}
+
+With these ingredients in mind, we adapt the definitions of syntactic precision and recall to domain models.
+Let $\mathcal{M}$ be a domain model and let $\mathcal{M'}$ be the GTM.
+We know that \textcolor{red}{$size(\mathcal{M}) = \sum_{\xi \in \mathcal{M}}\left|pre(\xi)\right| + \left|add(\xi)\right| + \left|del(\xi)\right|$} and by definition the number of preconditions and effects of the learned action models is equal to the sum of {\em true positives} and {\em false positives}; that is, $size(\mathcal{M}) = tp + fp$.
+
+The number of \emph{deletions} required to transform $\mathcal{M}$ into $\mathcal{M'}$ ($\left|DEL(\mathcal{M},\mathcal{M'})\right|$) matches our previous definition of the number of {\em false positives}; and $\left|INS(\mathcal{M},\mathcal{M'})\right|$, the number of \emph{insertions} required to transform $\mathcal{M}$ into $\mathcal{M'}$, corresponds to the number of {\em false negatives} of $\mathcal{M}$.
+Then we can affirm that $size(\mathcal{M'}) = size(\mathcal{M}) - \left|DEL(\mathcal{M},\mathcal{M'})\right| + \left|INS(\mathcal{M},\mathcal{M'})\right|$.
+
+\begin{mydefinition} \label{syn-precision} The precision of $\mathcal{M}$ relative to the GTM is defined as the fraction of the common preconditions and effects between $\mathcal{M}$ and the GTM among all prediconditions and effects of $\mathcal{M}$.
+%This gives an intuitive measure of the soundness of $\mathcal{M}$.
+	\begin{small}
+		\begin{align*}
+		Precision=\frac{tp}{tp+fp}=\frac{size(\mathcal{M})- \left|DEL(\mathcal{M},GTM)\right|}{size(\mathcal{M})}
+		\end{align*}
+	\end{small}
+\end{mydefinition}
+
+\begin{mydefinition} \label{syn-recall} The recall of $\mathcal{M}$ relative to the GTM is defined as the fraction of the common preconditions and effects between $\mathcal{M}$ and the GTM among all preconditions and effects of the GTM.
+%This gives an intuitive measure of the completeness of $\mathcal{M}$.
+
+	\begin{small}
+		\begin{align*}
+		Recall= \frac{tp}{tp+fn}=
+		\frac{size(\mathcal{M})- \left|DEL(\mathcal{M},GTM)\right|}{size(\mathcal{M}) - \left|DEL(\mathcal{M},GTM)\right| + \left|INS(\mathcal{M},GTM)\right|}
+		\end{align*}
+	\end{small}
+\end{mydefinition}
+
+\textcolor{red}{Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
+%In more detail, precision is the fraction of relevant predicates among the predicates that make up the learned model, and recall is the fraction of relevant predicates in the learned model over all relevant predicates, understanding a relevant predicate as one that appears in the GTM.
+As an example, a precision of 0.5 means that only half of the predicates that make up the learned domain model are present in the GTM, while a recall of 0.5 means that only half of the predicates that make up the GTM are present in the learned domain model.}
+
+\subsection{Semantic-based precision and recall for planning models}
+\label{semantic_precision_recall}
+
+Pure syntax-based evaluation metrics can report low scores for learned models that are actually {\em sound} and {\em complete} but syntactically different from the GTM.
+Semantic evaluation metrics add a distinctive value over the syntactic ones, which is that they evaluate the learned model with a set of observations of plan executions.
+\textcolor{red}{These metrics measure how well a model can reproduce a given plan execution, so the use of the word "semantic" here is in reference to the degree to which the learned model is able to capture the physics of the domain.} Semantic metrics are appropriate for scenarios where:
 
 \begin{enumerate}
-\item When the GTM is unknown. This is the most common scenario in ML, where models are both learned and evaluated with respect to datasets.
-\item When the GTM is known but a test-based evaluation on a dataset is preferable (or needed as complementary to a syntactic evaluation).
+	\item The GTM is unknown. This is the most common scenario in ML, where models are both learned and evaluated with respect to datasets.
+	\item The GTM is known but a test-based evaluation on a dataset is preferable (or needed as complementary to a syntactic evaluation).
 As a rule of thumb, it is preferable to evaluate the learned models wrt a dataset because a learned model can be semantically correct though syntactically incorrect (different from the GTM).
@@
 A more complex kind of reformulation occurs when two or more action models are learned in a single \emph{macro-action}.
+These semantic alterations typically appear in the learned models when the observed input data given in $\tau$ is scarce.
 
-These semantic alterations typically appear in the learned models when the observed input data given in $\tau$ is scarce.
+%These semantic alterations typically appear in the learned models when the observed input data given in $\tau$ is scarce.
 Defining a proper semantic evaluation is key because the application of syntax-based metrics may report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the GTM.
-In the following sections, we introduce a novel evaluation metric that is robust to different types of reformulation.
+In the following, we introduce a novel evaluation metric that is robust to different types of reformulation.
 
-\subsection{Semantic-based precision and recall}
-\label{semantic_precision_recall}
-
-The \ARMS system showed that a semantic evaluation can be done via validation of a set of plan traces with the learned model~\cite{yang2007learning}.
+\textcolor{red}{The \ARMS system was the first to show that a semantic evaluation can be done via validation of a set of plan traces with the learned model~\cite{yang2007learning}.
 The underlying idea is that an error indication of the learned action models is obtained by counting the number of preconditions that are not satisfied during the execution of the plan trace with the learned models, similarly to the functionality provided by the automatic validation tool VAL~\cite{howey2004val} used in the IPCs.
@@
 In other words, modifying the plan trace to fit the model.
-Inspired by this approach, we present here an alternative evaluation that, instead, determines the modifications required by a learned model to explain the given plan traces.
+Inspired by this approach, we present novel semantic-based error measure that builds upon the {\em precision} and {\em recall} metrics, and instead, determines the modifications required by a learned model to explain the given plan traces.}
 
-The rationale behind our novel metrics lies in counting the \emph{edit operations} that need to be applied in a set of action models $\mathcal{M}$ to fit the plan traces.
-Given a set of action models $\mathcal{M}$, the two allowed edit operations are:
-
-\begin{itemize}
-\item {\em Deletion}. A fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ is removable from $\xi\in\mathcal{M}$.
-\item {\em Insertion}. A fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ can be added to $\xi\in\mathcal{M}$.
-\end{itemize}
-
-We now provide formal definitions of $INS(\mathcal{M},\mathcal{M'})$ and $DEL(\mathcal{M},\mathcal{M'})$, the sets of insertions and deletions, respectively, that are needed to transform a set of action models $\mathcal{M}$ into a new set of action models $\mathcal{M'}$.
-
-\begin{mydefinition}
-	Let $PRE(\xi) = \underset{\forall p \in pre(\xi)}{\bigcup} pre_p(\xi)$, $ADD(\xi) = \underset{\forall p \in add(\xi)}{\bigcup} add_p(\xi)$, and $DEL(\xi) = \underset{\forall p \in del(\xi)}{\bigcup} del_p(\xi)$ be the set of propositional fluents that represent preconditions, positive and negative effects of a given action model $\xi$.
-We define:
-	\begin{small}
-		\begin{align*}
-		INS(\mathcal{M}, \mathcal{M'})=&PRE(\xi') \backslash PRE(\xi) \cup\\
-		&ADD(\xi') \backslash ADD(\xi) \cup \\
-		&DEL(\xi') \backslash DEL(\xi),\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t.\ name(\xi) = name(\xi')\\
-		\\
-		DEL(\mathcal{M}, \mathcal{M'})=&PRE(\xi) \backslash PRE(\xi') \cup\\
-		&ADD(\xi) \backslash ADD(\xi') \cup \\
-		&DEL(\xi) \backslash DEL(\xi'),\forall \xi\in\mathcal{M}, \xi'\in\mathcal{M'} s.t.\ name(\xi) = name(\xi')\\
-		\end{align*}
-	\end{small}
-\end{mydefinition}
-
-With these ingredients in mind, we first adapt the definitions of syntactic precision and recall to sets of action models.
-Let $\mathcal{M}$ be a set of learned action models and let $\mathcal{M'}$ be the GTM.
-We know that $size(\mathcal{M}) = \left|pre(\xi)\right| + \left|add(\xi)\right| + \left|del(\xi)\right| \forall \xi \in \mathcal{M}$ and by definition the number of preconditions and effects of the learned action models is equal to the sum of {\em true positives} and {\em false positives}; that is, $size(\mathcal{M}) = tp + fp$.
-
-The number of \emph{deletions} required to transform $\mathcal{M}$ into $\mathcal{M'}$ ($\left|DEL(\mathcal{M},\mathcal{M'})\right|$) matches our previous definition of the number of {\em false positives}; and $\left|INS(\mathcal{M},\mathcal{M'})\right|$, the number of \emph{insertions} required to transform $\mathcal{M}$ into $\mathcal{M'}$, corresponds to the number of {\em false negatives} of $\mathcal{M}$.
-Then we can affirm that $size(\mathcal{M'}) = size(\mathcal{M}) - \left|DEL(\mathcal{M},\mathcal{M'})\right| + \left|INS(\mathcal{M},\mathcal{M'})\right|$.
-
-\begin{mydefinition} \label{syn-precision} The precision of $\mathcal{M}$ relative to the GTM is defined as the fraction of the common predicates and effects between $\mathcal{M}$ and the GTM among all predicates and effects of $\mathcal{M}$.
-This gives an intuitive measure of the soundness of $\mathcal{M}$.
-\begin{small}
-	\begin{align*}
-     Precision=\frac{tp}{tp+fp}=\frac{size(\mathcal{M})- \left|DEL(\mathcal{M},GTM)\right|}{size(\mathcal{M})}
-	\end{align*}
-\end{small}
-\end{mydefinition}
-
-\begin{mydefinition} \label{syn-recall} The recall of $\mathcal{M}$ relative to the GTM is defined as the fraction of the common predicates and effects between $\mathcal{M}$ and the GTM among all predicates and effects of the GTM.
-This gives an intuitive measure of the completeness of $\mathcal{M}$.
-
-\begin{small}
-	\begin{align*}
-     Recall= \frac{tp}{tp+fn}=
-\frac{size(\mathcal{M})- \left|DEL(\mathcal{M},GTM)\right|}{size(\mathcal{M}) - \left|DEL(\mathcal{M},GTM)\right| + \left|INS(\mathcal{M},GTM)\right|}
-	\end{align*}
-\end{small}
-\end{mydefinition}
-
-Definitions \ref{syn-precision} and \ref{syn-recall} are syntax-based metrics to evaluate $\mathcal{M}$ with respect to the GTM.
+%Definitions \ref{syn-precision} and \ref{syn-recall} are syntax-based metrics to evaluate $\mathcal{M}$ with respect to the GTM.
 A semantic metric, on the other hand, evaluates $\mathcal{M}$ with respect to a given set of plan traces.
@@
 \begin{itemize}
-	\item $\mathcal{M}$ is a \textbf{set of learned action models} obtained using any learning approach such as \FAMA; in general, $\mathcal{M}$ can be any given input set of action models even manually encoded.
+	\item $\mathcal{M}$ is a \textbf{learned domain model} obtained using any learning approach such as \FAMA; in general, $\mathcal{M}$ can be any given input domain model even manually encoded.
 	\item$\mathcal{T}$ is a set of plan traces  used for \textbf{testing}.
@@
 
-A solution to this task is an \textbf{edited set of action models} $\mathcal{M'}$ such that (1) $\mathcal{M'}$ is obtained by exclusively applying a finite sequence of \emph{deletion} and \emph{insertion} operations to $\mathcal{M}$ and (2) $\mathcal{M'}$ explains $\mathcal{T}$; i.e.
+A solution to this task is an \textbf{edited domain model} $\mathcal{M'}$ such that (1) $\mathcal{M'}$ is obtained by exclusively applying a finite sequence of \emph{deletion} and \emph{insertion} operations to $\mathcal{M}$ and (2) $\mathcal{M'}$ explains $\mathcal{T}$; i.e.
 $\mathcal{M'}$ is {\em compliant} with every plan trace $\tau\in\mathcal{T}$.
@@
 
-\begin{mydefinition}
-  Given a set of action models $\mathcal{M}$, and all the sets of action models $\mathcal{M'}$ able to explain the plan traces $\mathcal{T}$.
-The {\bf closest compliant set of action models}, $\mathcal{M^*}$, is the comparable set of action models closest to $\mathcal{M}$ (in terms of editions) that is able to explain $\mathcal{T}$;
+\begin{mydefinition} \label{compliant}
+  Given a domain model $\mathcal{M}$, and all the domain models $\mathcal{M'}$ able to explain the plan traces $\mathcal{T}$.
+The {\bf closest compliant domain model}, $\mathcal{M^*}$, is the comparable domain model closest to $\mathcal{M}$ (in terms of editions) that is able to explain $\mathcal{T}$;
   \[\mathcal{M^*}=\underset{\forall \mathcal{M}' \rightarrow \mathcal{T}}{\arg\min} \ \left| INS(\mathcal{M},\mathcal{M'}) \cup DEL(\mathcal{M},\mathcal{M'}) \right|\]
@@
 
-The closest compliant set of action models $\mathcal{M^*}$ allows us to define a semantic version of {\em precision} and {\em recall} following definitions \ref{syn-precision} and \ref{syn-recall}.
+The closest compliant domain model $\mathcal{M^*}$ allows us to define a semantic version of {\em precision} and {\em recall} following definitions \ref{syn-precision} and \ref{syn-recall}.
 
@@
 \begin{myproposition}
-When the closest compliant set of action models $\mathcal{M^*}$ of an evaluation task $\Lambda = \tup{\mathcal{M}, \mathcal{T}}$ is the GTM, the syntactic and semantic evaluation of $\mathcal{M}$ return the same values; that is, $Precision=sem\text{-}Precision=$ and $Recall=sem\text{-}Recall$.
+When the closest compliant set of action models $\mathcal{M^*}$ of an evaluation task $\Lambda = \tup{\mathcal{M}, \mathcal{T}}$ is the GTM, the syntactic and semantic evaluation of $\mathcal{M}$ return the same values; that is, $Precision=sem\text{-}Precision$ and $Recall=sem\text{-}Recall$.
 \end{myproposition}
 
-The intuition behind this evaluation is to {\em semantically} assess how well the learned action models $\mathcal{M}$ explain a set of given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce the observations.
-Unlike the semantic metric defined by ARMS, our novel semantic definitions of precision and recall are not sensitive to flaws that appear more than once in the plan traces since the flaws are corrected only once in the learned models instead of at every intermediate state of the plan traces.
+The intuition behind this evaluation is to {\em semantically} assess how well the learned domain model $\mathcal{M}$ explain a set of given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce the observations.
+Unlike the semantic metric defined by ARMS, our novel semantic definitions of precision and recall are not sensitive to flaws in the action model that manifest more than once in the plan traces since the flaws are corrected only once in the learned models instead of at every intermediate state of the plan traces.
 
@@
 The compilation scheme presented in section \ref{compilation} is extensible to address the evaluation task $\Lambda=\tup{\mathcal{M}, \mathcal{T}}$ defined in section \ref{semantic_precision_recall}.
-In this extended task, $\mathcal{M}$ represents a set of previously learned action models; therefore, rather than learning the action models from scratch, we simply edit $\mathcal{M}$ until it satisfies the given test set of plan traces $\mathcal{T}$.
+In this extended task, $\mathcal{M}$ represents a previously learned domain model; therefore, rather than learning the action models from scratch, we simply edit $\mathcal{M}$ until it satisfies the given test set of plan traces $\mathcal{T}$.
 A solution to the classical planning task resulting from the extended compilation is a plan that:
@@
 \begin{enumerate}
-\item {\bf Edits the action models $\mathcal{M}$ to build $\mathcal{M}'$}.
+\item {\bf Edits the domain model $\mathcal{M}$ to build $\mathcal{M}'$}.
 A solution plan starts with a prefix that modifies the preconditions and effects of the action schemes in $\mathcal{M}$ using the two {\em edit operations} defined above, {\em deletion} and {\em insertion}.
@@
 
-Assuming we are using an optimal planner to solve $P_{\Lambda}'$, the solution plan of this problem will induce the \emph{closest compliant set of action models} $\mathcal{M^*}$.
+Assuming we are using an optimal planner to solve $P_{\Lambda}'$, the solution plan of this problem will induce the \emph{closest compliant domain model} $\mathcal{M^*}$.
 Therefore, our compilation enables the straightforward computation of the semantic versions of \emph{precision} and \emph{recall}.
@@
 Considering this, $sem\text{-}Precision$ and $sem\text{-}Recall$ can be approximated if $P_{\Lambda}'$ is solved with a satisfying planner.
-In this case, no guarantees can be made that the edited models will be the closest compliant ones, but a classical planner will always try to minimize the solution plan length and hence the number of edit operations applied to the input models.
+In this case, no guarantees can be made that the edited model will be the closest compliant one, but a classical planner will always try to minimize the solution plan length and hence the number of edit operations applied to the input model.

### 2018-12-17 doc/ijcai-19-conformant/conformant-learning.tex
conformant-learning.adding comments by Eva and Diego from todays meeting
updates added after meeting for discussing the points made by blai

@@
 \section{Introduction}
-Given a sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to {\em explain} the given observation.
+Given an input sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to {\em explain} the given observation.
 This task is of interest because it allows the incremental learning of action models from arbitrary large sets of partial observations.
 
-In addition, the paper introduces a new method to compute the {\em least-commitment} action model from a sequence of partially observed states.
+In addition, the paper introduces a new method to compute the {\em least-commitment} action model for an input sequence of partially observed states.
 The method assumes that action models are specified as \strips\ action schemata and it is built on top of off-the-shelf algorithms for {\em conformant planning}.
@@
 \section{Background}
-This section formalizes the planning models we use in the paper as well as the kind of observations that are given as input for the computation of the {\em least-commitment} action model.
+This section formalizes the {\em planning models} we use in the paper as well as the kind of state {\em observations} that are given as input for computing the {\em least-commitment} action model.
 
 \subsection{Classical planning with conditional effects}
-Let $F$ be the set of {\em fluents} or {\em state variables} (propositional variables).
-A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
+Let $F$ be the set of {\em fluents} or {\em state variables} (propositional variables) describing a state.
+A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e.
 either~$l=f$ or $l=\neg f$.
-$L$ is a set of literals that represents a partial assignment of values to fluents, and $\mathcal{L}(F)$ is the set of all literals sets on $F$, i.e.~all partial assignments of values to fluents.
-A {\em state} $s$ is a full assignment of values to fluents.
-We explicitly include negative literals $\neg f$ in states s.t.
-$|s|=|F|$ and the size of the state space is $2^{|F|}$.
+A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values).
+Given $L$, let $\neg L=\{\neg l:l\in L\}$ be its complement.
+We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.
+A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$.
 
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
-An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\in\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\in\mathcal{L}(F)$.
+Each classical planning action $a\in A$ has a precondition $\pre(a)\in\mathcal{L}(F)$ and a set of effects $\eff(a)\in\mathcal{L}(F)$.
 The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
-Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$.
-And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$, i.e.~if its precondition holds in $s$.
+The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
+Subtracting the complement of $\eff(a)$ from $s$ ensures that $\theta(s,a)$ remains a well-defined state.
+The subset of action effects that assign a positive value to a state fluent is called {\em positive effects} and denoted by $\eff^+(a)\in \eff(a)$ while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects} of an action $a\in A$.
 
@@
 Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\neg\eff_c(s,a)\cup\eff_c(s,a)\}$ where $\eff_c(s,a)$ are the {\em triggered effects} resulting from the action application (conditional effects whose conditions hold in $s$):
 \[
-triggered(s,a_c)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
+\eff_c(s,a)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
 \]
-
-The result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$, where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
 
@@
 
-A partially observable state $s_i^o$, {\small $0<i<m$}, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+A {\em partial state} $s_i^o$, {\small $0<i<m$}, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
 Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
-Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,s_0)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be {\em consistent} with the sequence of states of $\tau(\pi,s_0)$, meaning that $\forall i, s_i^o \subseteq s_i$.
 In practice, the number of observed states $m$, ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
 
-In other words, we assume there is a {\em bijective monotone mapping} between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+\newpage
+
+We are assuming then that there is a {\em bijective monotone mapping} between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
 Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
@@
 \begin{definition}[Explaning a $\mathcal{O}(\tau)$ observation]
-Given a {\em classical planning problem} $P$ and a sequence of partially observed states $\mathcal{O}(\tau)$, we say that a plan $\pi$ {\em explains the observation} (denoted $\pi\mapsto\mathcal{O}(\tau)$) iff $\pi$ is a solution for $P$ that is consistent with $\mathcal{O}(\tau)$.
-If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for $\mathcal{O}(\tau)$.
+Given a {\em classical planning problem} $P$ and a sequence of partial states $\mathcal{O}(\tau)$, a plan $\pi$ {\em explains} $\mathcal{O}(\tau)$ (denoted $\pi\mapsto\mathcal{O}(\tau)$) iff $\pi$ is a solution for $P$ {\em consistent} with $\mathcal{O}(\tau)$.
 \end{definition}
+If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}(\tau)$.
+
+Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
+We say that an action model $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$.
+Further we say that a model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that explains $\mathcal{O}(\tau)$.
 
@@
 \begin{itemize}
-\item $\Phi=\tup{F,A[\cdot]}$ is a {\em classical planning frame} where $A[\cdot]$ is a set of actions s.t.
-the semantics of each $a\in A[\cdot]$ is unknown; i.e.
-the corresponding $\tup{\rho,\theta}$ functions are undefined.
-\item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the partial observation of a trajectory $\tau(\pi,s_0)$ that is defined within the {\em classical planning frame} $\Phi$.
+\item $\Phi=\tup{F,A[\cdot]}$ is a {\em classical planning frame} where the semantics of each action $a\in A[\cdot]$ is unknown; i.e. the corresponding $\tup{\rho,\theta}$ functions are undefined. 
+\item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the partial observation of a trajectory $\tau(\pi,s_0)$ defined within the {\em classical planning frame} $\Phi$.
 \end{itemize}
 
-Before formalizing the solution to this task, i.e.
-the {\em least-commitment} action model, we introduce several necessary definitions.
-Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
-We say that an action model $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$.
-Further we say that a model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that explains $\mathcal{O}(\tau)$.
+Before formalizing the solution to this task, i.e. the {\em least-commitment} action model, we introduce several necessary definitions.
+
+\begin{definition}[Model Space]
+Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ the model space $M$ is the set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$.
+\end{definition}
 
@@
 \begin{definition}[Partially specified action model]
-Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ then, a {\em partially specified action model} $M$ is a set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$.
-(Note that if $M$ is a singleton it represents a {\em fully specified action model}).
+A {\em partially specified action model} is a subset of models in a given {\em model space $M$}.
 \end{definition}
+If the {\em partially specified action model} is a singleton, it represents a {\em fully specified action model}.
+On the other hand, if its size is $|M|$ the {\em partially specified action model} represents the full model space.
 
@@
 Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
+With this regard, we implement a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$ using only fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+This encoding exploits the syntactic constraints of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
 \item {\em Domain-specific constraints}.
@@
 For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-As a rule of thumb, {\it state invariants} constraining the possible states of a given planning domain belong to this second class of constraints~\cite{fox1998automatic}.
-\item {\em Observation constraints}. A sequence of state observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata. 
+As a rule of thumb, {\em typing constraints} and {\it state invariants} belong to this second class of constraints~\cite{fox1998automatic}.
+\item {\em Observation constraints}.
+A sequence of state observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.
+For instance, since we assume that the initial state $s_0$ is fully observed, we have that $\bigcup_{\xi}$ {\tt\small $\neg\pre\_e\_\xi$} $\wedge$ {\tt\small $\eff\_e\_\xi$} if there is an instantiation of $e$ that holds in $s_i^o$ ({\small $0<i$}) but not in $s_0$.
+
 \end{enumerate}
-
-\subsection{Computing the {\em least-commitment} model via conformant planning}
-Given a task $\tup{\Phi,\mathcal{O}(\tau)}$, this section shows that we can build a {\em conformant planning problem} $P_c$ whose solution induces the {\em least-commitment} action model for the input observation $\mathcal{O}(\tau)$.
-In more detail, we build a {\em conformant planning problem} $P_c=\tup{F_c,A_c,\Upsilon,G_c}$ such that:
-\begin{itemize}
-\item The set of fluents $F_c$ extends $F$ with two new sets of fluents:
-      \begin{itemize}
-      \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
-      \item Fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$) implementing a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$.
-{\bf Our encoding exploits the syntactic constraint of \strips\ } so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
-      \end {itemize}
 
@@
 
-      \item The set of actions $A_c$ contains now actions of three different kinds:
+\subsection{Computing the {\em least-commitment} model via classical planning}
+Inspired by the {\em classical planning compilation} $K_{s_0}$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose solution induces the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$:
 \begin{itemize}
-      \item Actions for {\em committing} {\tt\small pre\_e\_$\xi$} fluents to a positive/negative value (similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} fluents to a positive/negative value).
+\item The set of fluents $F'$ extends $F$ with two new sets of fluents:
+      \begin{itemize}
+      \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
+      \item Fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$} and {\tt\small $K\eff\_e\_\xi$} encoding the {\em knowledge level} representation of the space of possible \strips\ action models (note that fluents {\tt\small $K\neg\eff\_e\_\xi$} are not necessary in our encoding).
+      \end {itemize}
+\item The set of actions $A'$ contains now actions of three different kinds:
+\begin{itemize}
+      \item Actions for {\em committing} {\tt\small pre\_e\_$\xi$} to a positive/negative value.
+Similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} but only to a positive value (recall that fluents {\tt\small $K\neg\eff\_e\_\xi$} are not necessary in our encoding).
+The value of an effect {\tt\small eff\_e\_$\xi$} can only be committed once the value of the corresponding precondition {\tt\small pre\_e\_$\xi$} is committed (i.e.
+once either $\mathsf{K\pre\_e\_\xi}$ or $\mathsf{K\neg\pre\_e\_\xi}$ hold in the current state).
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{commit\top\_pre\_e\_\xi})=&\{mode_{commit}\},\\
-\cond(\mathsf{commit\top\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{pre\_e\_\xi\},\\
-                                    &\{\neg pre\_e\_\xi\}\rhd\{pre\_e\_\xi\}.\\\\
-\hspace*{7pt}\pre(\mathsf{commit\bot\_pre\_e\_\xi})=&\{mode_{commit}\},\\
-\cond(\mathsf{commit\bot\_pre\_e\_\xi})=&\{pre\_e\_\xi\}\rhd\{\neg pre\_e\_\xi\},\\
-                                    &\{\neg pre\_e\_\xi\}\rhd\{\neg pre\_e\_\xi\}.                                    
+\hspace*{7pt}\pre(\mathsf{commit\top\_pre\_e\_\xi})=&\{mode_{commit}, \\
+&\mathsf{\neg K\pre\_e\_\xi}, \mathsf{\neg K\neg\pre\_e\_\xi}\},\\
+\cond(\mathsf{commit\top\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\pre\_e\_\xi}\}.\\\\
+\hspace*{7pt}\pre(\mathsf{commit\bot\_pre\_e\_\xi})=&\{mode_{commit}, \\
+&\mathsf{\neg K\pre\_e\_\xi},\mathsf{\neg K\neg\pre\_e\_\xi}\},\\
+\cond(\mathsf{commit\bot\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\neg\pre\_e\_\xi}\}.\\\\
+\hspace*{7pt}\pre(\mathsf{commit\top\_eff\_e\_\xi})=&\{mode_{commit}, \\
+&\mathsf{\neg K\eff\_e\_\xi},\\
+&\mathsf{K\pre\_e\_\xi}\vee \mathsf{K\neg\pre\_e\_\xi}\},\\
+\cond(\mathsf{commit\top\_eff\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\eff\_e\_\xi}\}.\\\\
 \end{align*}
@@
 \hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j,\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j\},\\
                             &\{mode_{commit}\}\rhd\{\neg mode_{commit}, mode_{val}\}.
@@
 
-      \item {\em Editable} actions whose semantics is given by the value of {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} fluents at the current state.
+      \item {\em Editable} actions whose semantics is given by the value of fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$} and {\tt\small $K\eff\_e\_\xi$} at the current state.
 Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema.
-Note that this editable schema when the set of fluents {\tiny\tt (pre\_holding\_v1\_stack) (pre\_clear\_v2\_stack) (eff\_holding\_v1\_stack) (eff\_clear\_v2\_stack) (eff\_clear\_v1\_stack) (eff\_handempty\_stack) (eff\_on\_v1\_v2\_stack)} hold at the current state, then it behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional}.
+This editable schema behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional} when the set of fluents {\tiny\tt (Kpre\_holding\_v1\_stack) (Kpre\_clear\_v2\_stack) (Keff\_holding\_v1\_stack) (Keff\_clear\_v2\_stack) (Keff\_clear\_v1\_stack) (Keff\_handempty\_stack) (Keff\_on\_v1\_v2\_stack)} hold at the current state.
 Formally, given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is:
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
-\cond(\mathsf{editable_{\xi}})=&\{pre\_e\_\xi, eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
-&\{\neg pre\_e\_\xi, eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
+\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
+\cond(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi, K\eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+&\{K\neg\pre\_e\_\xi, K\eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
 \end{align*}
@@
 
-\item The clauses in $\Upsilon$ comprises:
-      \begin{enumerate}
-      \item The {\em unit clauses} given by the fluents that hold in the initial state $I=s_0$ and $mode_{commit}$ set to true.
-      \item The clauses representing that the actual value of fluents {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} is unknown.
-In other words, that any model from the \strips\ space of models (following the previously mentioned syntactic constraints) can initially be part of the {\em least-commitment} action model.
-Formally, for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$, then $\Upsilon$ includes these two clauses:
-            \begin{itemize}
-            \item {\tt\small pre\_e\_$\xi$} $\vee$ {\tt\small $\neg$pre\_e\_$\xi$}.
-            \item {\tt\small eff\_e\_$\xi$} $\vee$ {\tt\small $\neg$eff\_e\_$\xi$}.
-            \end{itemize}
-      \end{enumerate}
-One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier for a specific domain. 
-\item The new goals are $G_c=s_m\cup\{test_m\}$.
+\item The new initial state $I'=I \cup\{mode_{commit}\}$ while the new goals are $G'=s_m\cup\{test_m\}$.
 \end{itemize}
@@
  :precondition
-   (and (or (not (pre_on_v1_v1_stack)) (on ?o1 ?o1))
-        (or (not (pre_on_v1_v2_stack)) (on ?o1 ?o2))
-        (or (not (pre_on_v2_v1_stack)) (on ?o2 ?o1))
-        (or (not (pre_on_v2_v2_stack)) (on ?o2 ?o2))
-        (or (not (pre_ontable_v1_stack)) (ontable ?o1))
-        (or (not (pre_ontable_v2_stack)) (ontable ?o2))
-        (or (not (pre_clear_v1_stack)) (clear ?o1))
-        (or (not (pre_clear_v2_stack)) (clear ?o2))
-        (or (not (pre_holding_v1_stack)) (holding ?o1))
-        (or (not (pre_holding_v2_stack)) (holding ?o2))
-        (or (not (pre_handempty_stack)) (handempty)))
+   (and (or (not (Kpre_on_v1_v1_stack)) (on ?o1 ?o1))
+        (or (not (Kpre_on_v1_v2_stack)) (on ?o1 ?o2))
+        (or (not (Kpre_on_v2_v1_stack)) (on ?o2 ?o1))
+        (or (not (Kpre_on_v2_v2_stack)) (on ?o2 ?o2))
+        (or (not (Kpre_ontable_v1_stack)) (ontable ?o1))
+        (or (not (Kpre_ontable_v2_stack)) (ontable ?o2))
+        (or (not (Kpre_clear_v1_stack)) (clear ?o1))
+        (or (not (Kpre_clear_v2_stack)) (clear ?o2))
+        (or (not (Kpre_holding_v1_stack)) (holding ?o1))
+        (or (not (Kpre_holding_v2_stack)) (holding ?o2))
+        (or (not (Kpre_handempty_stack)) (handempty)))
  :effect (and
-   (when (and (pre_on_v1_v1_stack)(eff_on_v1_v1_stack)) (not (on ?o1 ?o1)))
-   (when (and (pre_on_v1_v2_stack)(eff_on_v1_v2_stack)) (not (on ?o1 ?o2)))
-   (when (and (pre_on_v2_v1_stack)(eff_on_v2_v1_stack)) (not (on ?o2 ?o1)))
-   (when (and (pre_on_v2_v2_stack)(eff_on_v2_v2_stack)) (not (on ?o2 ?o2)))
-   (when (and (pre_ontable_v1_stack)(eff_ontable_v1_stack)) (not (ontable ?o1)))
-   (when (and (pre_ontable_v2_stack)(eff_ontable_v2_stack)) (not (ontable ?o2)))
-   (when (and (pre_clear_v1_stack)(eff_clear_v1_stack)) (not (clear ?o1)))
-   (when (and (pre_clear_v2_stack)(eff_clear_v2_stack)) (not (clear ?o2)))
-   (when (and (pre_holding_v1_stack)(eff_holding_v1_stack)) (not (holding ?o1)))
-   (when (and (pre_holding_v2_stack)(eff_holding_v2_stack)) (not (holding ?o2)))
-   (when (and (pre_handempty_stack)(eff_handempty_stack)) (not (handempty)))
-   (when (and (not(pre_on_v1_v1_stack))(eff_on_v1_v1_stack)) (on ?o1 ?o1))
-   (when (and (not(pre_on_v1_v2_stack))(eff_on_v1_v2_stack)) (on ?o1 ?o2))
-   (when (and (not(pre_on_v2_v1_stack))(eff_on_v2_v1_stack)) (on ?o2 ?o1))
-   (when (and (not(pre_on_v2_v2_stack))(eff_on_v2_v2_stack)) (on ?o2 ?o2))
-   (when (and (not(pre_ontable_v1_stack))(eff_ontable_v1_stack)) (ontable ?o1))
-   (when (and (not(pre_ontable_v2_stack))(eff_ontable_v2_stack)) (ontable ?o2))
-   (when (and (not(pre_clear_v1_stack))(eff_clear_v1_stack)) (clear ?o1))
-   (when (and (not(pre_clear_v2_stack))(eff_clear_v2_stack)) (clear ?o2))
-   (when (and (not(pre_holding_v1_stack))(eff_holding_v1_stack)) (holding ?o1))
-   (when (and (not(pre_holding_v2_stack))(eff_holding_v2_stack)) (holding ?o2))
-   (when (and (not(pre_handempty_stack))(eff_handempty_stack)) (handempty))))
+   (when (and (Kpre_on_v1_v1_stack)(Keff_on_v1_v1_stack)) (not (on ?o1 ?o1)))
+   (when (and (Kpre_on_v1_v2_stack)(Keff_on_v1_v2_stack)) (not (on ?o1 ?o2)))
+   (when (and (Kpre_on_v2_v1_stack)(Keff_on_v2_v1_stack)) (not (on ?o2 ?o1)))
+   (when (and (Kpre_on_v2_v2_stack)(Keff_on_v2_v2_stack)) (not (on ?o2 ?o2)))
+   (when (and (Kpre_ontable_v1_stack)(Keff_ontable_v1_stack)) (not (ontable ?o1)))
+   (when (and (Kpre_ontable_v2_stack)(Keff_ontable_v2_stack)) (not (ontable ?o2)))
+   (when (and (Kpre_clear_v1_stack)(Keff_clear_v1_stack)) (not (clear ?o1)))
+   (when (and (Kpre_clear_v2_stack)(Keff_clear_v2_stack)) (not (clear ?o2)))
+   (when (and (Kpre_holding_v1_stack)(Keff_holding_v1_stack)) (not (holding ?o1)))
+   (when (and (Kpre_holding_v2_stack)(Keff_holding_v2_stack)) (not (holding ?o2)))
+   (when (and (Kpre_handempty_stack)(Keff_handempty_stack)) (not (handempty)))
+   (when (and (Knot_pre_on_v1_v1_stack)(Keff_on_v1_v1_stack)) (on ?o1 ?o1))
+   (when (and (Knot_pre_on_v1_v2_stack)(Keff_on_v1_v2_stack)) (on ?o1 ?o2))
+   (when (and (Knot_pre_on_v2_v1_stack)(Keff_on_v2_v1_stack)) (on ?o2 ?o1))
+   (when (and (Knot_pre_on_v2_v2_stack)(Keff_on_v2_v2_stack)) (on ?o2 ?o2))
+   (when (and (Knot_pre_ontable_v1_stack)(Keff_ontable_v1_stack)) (ontable ?o1))
+   (when (and (Knot_pre_ontable_v2_stack)(Keff_ontable_v2_stack)) (ontable ?o2))
+   (when (and (Knot_pre_clear_v1_stack)(Keff_clear_v1_stack)) (clear ?o1))
+   (when (and (Knot_pre_clear_v2_stack)(Keff_clear_v2_stack)) (clear ?o2))
+   (when (and (Knot_pre_holding_v1_stack)(Keff_holding_v1_stack)) (holding ?o1))
+   (when (and (Knot_pre_holding_v2_stack)(Keff_holding_v2_stack)) (holding ?o2))
+   (when (and (Knot_pre_handempty_stack)(Keff_handempty_stack)) (handempty))))
   \end{verbatim}           
@@
 \end{figure}
-
-\subsection{Optimization of the compilation}
-In fact, we do not need to compute the {\em least-commitment} action model (i.e.
-solve the $P_c$ conformant planning problem) from scratch for all the $\tup{s_0^o,s_1^o \ldots , s_m^o}$ observations.
-A more efficient approach is to compute the {\em least-commitment} action model from scratch only for the first observation, that is for $\tup{s_0^o,s_1^o}$.
-Then, the obtained {\em least-commitment} action model can be used as input for improving the computation of the two first observations $\tup{s_0^o,s_1^o,s_2^o}$.
-In other words, including the commits found in the previous step as new {\em unit clauses} in set of the {\em initial clauses} $\Upsilon$.
-This process is repeated until the {\em least-commitment} action model is computed  for the full sequence of state observations $\tup{s_0^o,s_1^o \ldots , s_m^o}$.
 

### 2018-12-18 doc/ijcai-19-conformant/conformant-learning.tex
first updates added after discussing the points made by blai

@@
 Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-With this regard, we implement a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of an action schema $\xi$ using only fluents {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
 \item {\em Domain-specific constraints}.
@@
 For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-As a rule of thumb, {\em typing constraints} and {\it state invariants} belong to this second class of constraints~\cite{fox1998automatic}.
-\item {\em Observation constraints}.
-A sequence of state observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.
-For instance, since we assume that the initial state $s_0$ is fully observed, we have that $\bigcup_{\xi}$ {\tt\small $\neg\pre\_e\_\xi$} $\wedge$ {\tt\small $\eff\_e\_\xi$} if there is an instantiation of $e$ that holds in $s_i^o$ ({\small $0<i$}) but not in $s_0$.
+{\em Typing constraints} and {\it state invariants} are also constraints of this kind~\cite{fox1998automatic}.
+\item {\em Observation constraints}. A sequence of state observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
+\end{enumerate}
 
-\end{enumerate}
+\subsection{A propositional encoding for \strips\ action models}
+With this regard, we implement a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+This encoding exploits the syntactic constraints of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
+There is a total number of $2^{2\times|11|}=4,194,304$ different models for that schema.
 
@@
       \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
-      \item Fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$} and {\tt\small $K\eff\_e\_\xi$} encoding the {\em knowledge level} representation of the space of possible \strips\ action models (note that fluents {\tt\small $K\neg\eff\_e\_\xi$} are not necessary in our encoding).
+      \item Fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} encoding the {\em knowledge level} representation of the space of possible \strips\ action models.
       \end {itemize}
@@
       \item Actions for {\em committing} {\tt\small pre\_e\_$\xi$} to a positive/negative value.
-Similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} but only to a positive value (recall that fluents {\tt\small $K\neg\eff\_e\_\xi$} are not necessary in our encoding).
-The value of an effect {\tt\small eff\_e\_$\xi$} can only be committed once the value of the corresponding precondition {\tt\small pre\_e\_$\xi$} is committed (i.e.
-once either $\mathsf{K\pre\_e\_\xi}$ or $\mathsf{K\neg\pre\_e\_\xi}$ hold in the current state).
+Similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} to a positive/negative value but the value of {\tt\small eff\_e\_$\xi$} can only be committed once the value of the corresponding {\tt\small pre\_e\_$\xi$} is committed (i.e.
+once either $\mathsf{K\pre\_e\_\xi}$ or $\mathsf{K\neg\pre\_e\_\xi}$ holds in the current state).
 \begin{small}
@@
 &\mathsf{\neg K\pre\_e\_\xi},\mathsf{\neg K\neg\pre\_e\_\xi}\},\\
-\cond(\mathsf{commit\bot\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\neg\pre\_e\_\xi}\}.\\\\
-\hspace*{7pt}\pre(\mathsf{commit\top\_eff\_e\_\xi})=&\{mode_{commit}, \\
-&\mathsf{\neg K\eff\_e\_\xi},\\
-&\mathsf{K\pre\_e\_\xi}\vee \mathsf{K\neg\pre\_e\_\xi}\},\\
-\cond(\mathsf{commit\top\_eff\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\eff\_e\_\xi}\}.\\\\
+\cond(\mathsf{commit\bot\_pre\_e\_\xi})=&\{\emptyset\}\rhd\{\mathsf{K\neg\pre\_e\_\xi}\}.
 \end{align*}
@@
 
-      \item {\em Editable} actions whose semantics is given by the value of fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$} and {\tt\small $K\eff\_e\_\xi$} at the current state.
+      \item {\em Editable} actions whose semantics is given by the value of the {\em knowledge level} fluents ({\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$}) at the current state.
 Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
-\cond(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi, K\eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
-&\{K\neg\pre\_e\_\xi, K\eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
+\hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{\neg K\neg\pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
+\cond(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi, \neg K\neg\eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+&\{K\neg\pre\_e\_\xi, \neg K\neg\eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
 \end{align*}
@@
  :precondition
-   (and (or (not (Kpre_on_v1_v1_stack)) (on ?o1 ?o1))
-        (or (not (Kpre_on_v1_v2_stack)) (on ?o1 ?o2))
-        (or (not (Kpre_on_v2_v1_stack)) (on ?o2 ?o1))
-        (or (not (Kpre_on_v2_v2_stack)) (on ?o2 ?o2))
-        (or (not (Kpre_ontable_v1_stack)) (ontable ?o1))
-        (or (not (Kpre_ontable_v2_stack)) (ontable ?o2))
-        (or (not (Kpre_clear_v1_stack)) (clear ?o1))
-        (or (not (Kpre_clear_v2_stack)) (clear ?o2))
-        (or (not (Kpre_holding_v1_stack)) (holding ?o1))
-        (or (not (Kpre_holding_v2_stack)) (holding ?o2))
-        (or (not (Kpre_handempty_stack)) (handempty)))
+   (and (or (Knotpre_on_v1_v1_stack) (on ?o1 ?o1))
+        (or (Knotpre_on_v1_v2_stack) (on ?o1 ?o2))
+        (or (Knotpre_on_v2_v1_stack) (on ?o2 ?o1))
+        (or (Knotpre_on_v2_v2_stack) (on ?o2 ?o2))
+        (or (Knotpre_ontable_v1_stack) (ontable ?o1))
+        (or (Knotpre_ontable_v2_stack) (ontable ?o2))
+        (or (Knotpre_clear_v1_stack) (clear ?o1))
+        (or (Knotpre_clear_v2_stack) (clear ?o2))
+        (or (Knotpre_holding_v1_stack) (holding ?o1))
+        (or (Knotpre_holding_v2_stack) (holding ?o2))
+        (or (Knotpre_handempty_stack) (handempty)))
  :effect (and

### 2018-12-18 doc/ijcai-19-conformant/conformant-learning.tex
first updates added after discussing the points made by blai

@@
 %%%%%%%%%%%%%%%%%% Added for this paper
+\usepackage{listings}% http://ctan.org/pkg/listings
+\lstset{
+  basicstyle=\ttfamily,
+  mathescape
+}
 \usepackage{ wasysym }
@@
 In addition, the paper introduces a new method to compute the {\em least-commitment} action model for an input sequence of partially observed states.
-The method assumes that action models are specified as \strips\ action schemata and it is built on top of off-the-shelf algorithms for {\em conformant planning}.
+The method assumes that action models are specified as \strips\ action schemata and it is built on top of off-the-shelf algorithms for {\em classical planning}.
 
@@
 
-\subsection{Conformant planning}
-{\em Conformant planning} is planning with incomplete information about the initial state, no sensing, and validating that goals are achieved with certainty (despite the uncertainty of the initial state)~\cite{goldman1996expressive,smith1998conformant,bonet2000planning}.
-
-Syntactically, conformant planning problems are expressed in compact form through a set of state variables.
-A {\em conformant planning problem} can be defined as a tuple $P_c=\tup{F,A,\Upsilon,G}$ where $F$, $A$ and $G$ are the set of {\em fluents}, {\em actions} and {\em goals} (as previously defined for {\em classical planning}).
-Now $\Upsilon$ is a set of clauses over literals $l=f$ or $l=\neg f$ (for $f\in F$) that define the set of possible initial states.
-
-A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state.
-More precisely, an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff, for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ s.t.
-$s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$, then $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$.
-
 \section{Computing the {\em least-commitment} action model from state observations}
@@
 Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
+{\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
 \item {\em Domain-specific constraints}.
@@
 
-\subsection{A propositional encoding for \strips\ action models}
-With this regard, we implement a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\ so, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+\subsection{A propositional encoding for {\em partially specified} \strips\ models}
+First we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+This encoding exploits the syntactic constraints of \strips\ so is more compact that the one previuosly proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
 
@@
 \label{fig:propositional}
+\end{figure}
+
+Now we show that this encoding can be extended to the {\em knwoledge level} to compactly represent a partially specified action model.
+The extension replaces each proposition {\tt\small $pre\_e\_\xi$} with two new propositions {\tt\small $K\pre\_e\_\xi$} and {\tt\small $K\neg\pre\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is a precondition of $\xi$.
+The same is done for each {\tt\small $eff\_e\_\xi$} that is replaced by {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is an effect of $\xi$ or that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not an effect of $\xi$.
+Figure~\ref{fig:partial} shows a {\em partial specification} of the  {\tt\small stack(?v1,?v2)} schema that follows our propositional representation.
+
+With respect to this formalization, the {\em full space} of possible \strips\ schemas for $xi$ is compactly represented as {\tt\small $\bigcap_{e\in{\mathcal I}_{\Psi,\xi}} \neg K\pre\_e\_\xi\wedge \neg K\neg\pre\_e\_\xi \wedge\neg K\eff\_e\_\xi\wedge \neg K\neg\eff\_e\_\xi$}.
+
+\begin{figure}
+  \begin{tiny}  
+ \begin{lstlisting} 
+(K$\neg\pre$_holding_v2_stack) (K$\neg\pre$_clear_v1_stack)
+(K$\neg\pre$_handempty_stack) 
+(K$\neg\pre$_ontable_v1_stack) (K$\neg\pre$_ontable_v2_stack)
+(K$\neg\pre$_on_v1_v2_stack) (K$\neg\pre$_on_v2_v1_stack)
+(K$\neg\pre$_on_v1_v1_stack) (K$\neg\pre$_on_v2_v2_stack)
+
+(K$\neg\eff$_holding_v2_stack) (K$\neg\eff$_clear_v2_stack)
+(K$\neg\eff$_clear_v2_stack) (K$\neg\eff$_on_v2_v2_stack)
+(K$\neg\eff$_on_v1_v1_stack) (K$\neg\eff$_on_v2_v1_stack)
+  \end{lstlisting}           
+  \end{tiny}  
+ \caption{\small {\em Partial specification} of the  {\tt\small stack(?v1,?v2)} schema that follows our propositional representation.}
+\label{fig:partial}
 \end{figure}
@@
  :effect (and
-   (when (and (Kpre_on_v1_v1_stack)(Keff_on_v1_v1_stack)) (not (on ?o1 ?o1)))
-   (when (and (Kpre_on_v1_v2_stack)(Keff_on_v1_v2_stack)) (not (on ?o1 ?o2)))
-   (when (and (Kpre_on_v2_v1_stack)(Keff_on_v2_v1_stack)) (not (on ?o2 ?o1)))
-   (when (and (Kpre_on_v2_v2_stack)(Keff_on_v2_v2_stack)) (not (on ?o2 ?o2)))
-   (when (and (Kpre_ontable_v1_stack)(Keff_ontable_v1_stack)) (not (ontable ?o1)))
-   (when (and (Kpre_ontable_v2_stack)(Keff_ontable_v2_stack)) (not (ontable ?o2)))
-   (when (and (Kpre_clear_v1_stack)(Keff_clear_v1_stack)) (not (clear ?o1)))
-   (when (and (Kpre_clear_v2_stack)(Keff_clear_v2_stack)) (not (clear ?o2)))
-   (when (and (Kpre_holding_v1_stack)(Keff_holding_v1_stack)) (not (holding ?o1)))
-   (when (and (Kpre_holding_v2_stack)(Keff_holding_v2_stack)) (not (holding ?o2)))
-   (when (and (Kpre_handempty_stack)(Keff_handempty_stack)) (not (handempty)))
-   (when (and (Knot_pre_on_v1_v1_stack)(Keff_on_v1_v1_stack)) (on ?o1 ?o1))
-   (when (and (Knot_pre_on_v1_v2_stack)(Keff_on_v1_v2_stack)) (on ?o1 ?o2))
-   (when (and (Knot_pre_on_v2_v1_stack)(Keff_on_v2_v1_stack)) (on ?o2 ?o1))
-   (when (and (Knot_pre_on_v2_v2_stack)(Keff_on_v2_v2_stack)) (on ?o2 ?o2))
-   (when (and (Knot_pre_ontable_v1_stack)(Keff_ontable_v1_stack)) (ontable ?o1))
-   (when (and (Knot_pre_ontable_v2_stack)(Keff_ontable_v2_stack)) (ontable ?o2))
-   (when (and (Knot_pre_clear_v1_stack)(Keff_clear_v1_stack)) (clear ?o1))
-   (when (and (Knot_pre_clear_v2_stack)(Keff_clear_v2_stack)) (clear ?o2))
-   (when (and (Knot_pre_holding_v1_stack)(Keff_holding_v1_stack)) (holding ?o1))
-   (when (and (Knot_pre_holding_v2_stack)(Keff_holding_v2_stack)) (holding ?o2))
-   (when (and (Knot_pre_handempty_stack)(Keff_handempty_stack)) (handempty))))
+   (when (and (Kpre_on_v1_v1_stack)(not(Knot_eff_on_v1_v1_stack))) (not (on ?o1 ?o1)))
+   (when (and (Kpre_on_v1_v2_stack)(not(Knot_eff_on_v1_v2_stack))) (not (on ?o1 ?o2)))
+   (when (and (Kpre_on_v2_v1_stack)(not(Knot_eff_on_v2_v1_stack))) (not (on ?o2 ?o1)))
+   (when (and (Kpre_on_v2_v2_stack)(not(Knot_eff_on_v2_v2_stack))) (not (on ?o2 ?o2)))
+   (when (and (Kpre_ontable_v1_stack)(not(Knot_eff_ontable_v1_stack))) (not (ontable ?o1)))
+   (when (and (Kpre_ontable_v2_stack)(not(Knot_eff_ontable_v2_stack))) (not (ontable ?o2)))
+   (when (and (Kpre_clear_v1_stack)(not(Knot_eff_clear_v1_stack))) (not (clear ?o1)))
+   (when (and (Kpre_clear_v2_stack)(not(Knot_eff_clear_v2_stack))) (not (clear ?o2)))
+   (when (and (Kpre_holding_v1_stack)(not(Knot_eff_holding_v1_stack))) (not (holding ?o1)))
+   (when (and (Kpre_holding_v2_stack)(not(Knot_eff_holding_v2_stack))) (not (holding ?o2)))
+   (when (and (Kpre_handempty_stack)(not(Knot_eff_handempty_stack))) (not (handempty)))
+   (when (and (Knot_pre_on_v1_v1_stack)(not(Knot_eff_on_v1_v1_stack))) (on ?o1 ?o1))
+   (when (and (Knot_pre_on_v1_v2_stack)(not(Knot_eff_on_v1_v2_stack))) (on ?o1 ?o2))
+   (when (and (Knot_pre_on_v2_v1_stack)(not(Knot_eff_on_v2_v1_stack))) (on ?o2 ?o1))
+   (when (and (Knot_pre_on_v2_v2_stack)(not(Knot_eff_on_v2_v2_stack))) (on ?o2 ?o2))
+   (when (and (Knot_pre_ontable_v1_stack)(not(Knot_eff_ontable_v1_stack))) (ontable ?o1))
+   (when (and (Knot_pre_ontable_v2_stack)(not(Knot_eff_ontable_v2_stack))) (ontable ?o2))
+   (when (and (Knot_pre_clear_v1_stack)(not(Knot_eff_clear_v1_stack))) (clear ?o1))
+   (when (and (Knot_pre_clear_v2_stack)(not(Knot_eff_clear_v2_stack))) (clear ?o2))
+   (when (and (Knot_pre_holding_v1_stack)(not(Knot_eff_holding_v1_stack))) (holding ?o1))
+   (when (and (Knot_pre_holding_v2_stack)(not(Knot_eff_holding_v2_stack))) (holding ?o2))
+   (when (and (Knot_pre_handempty_stack)(not(Knot_eff_handempty_stack))) (handempty))))
   \end{verbatim}           

### 2018-12-19 doc/ijcai-19-conformant/conformant-learning.tex
first updates added after discussing the points made by blai

@@
 \begin{definition}[Explaning a $\mathcal{O}(\tau)$ observation]
-Given a {\em classical planning problem} $P$ and a sequence of partial states $\mathcal{O}(\tau)$, a plan $\pi$ {\em explains} $\mathcal{O}(\tau)$ (denoted $\pi\mapsto\mathcal{O}(\tau)$) iff $\pi$ is a solution for $P$ {\em consistent} with $\mathcal{O}(\tau)$.
+Given a {\em classical planning problem} $P$ and an observation $\mathcal{O}(\tau)$, a plan $\pi$ {\em explains} $\mathcal{O}(\tau)$ (denoted $\pi\mapsto\mathcal{O}(\tau)$) iff $\pi$ is a solution for $P$ that is {\em consistent} with the state trajectory constraints imposed by the sequence of partial states $\mathcal{O}(\tau)$.
 \end{definition}
@@
 
-Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
-We say that an action model $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$.
-Further we say that a model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that explains $\mathcal{O}(\tau)$.
-
-\section{Computing the {\em least-commitment} action model from state observations}
-First, this section formalizes the notion of the {\em least-commitment} action model that is able to {\em explain} a sequence of partially observed states.
-Next, the section describes our approach to compute such model via {\em conformant planning}.
+\section{Computing the {\em least-commitment} action model from observations}
+First, this section formalizes the notion of the {\em least-commitment} action model that is able to {\em explain} an observation. Next, the section describes our approach to compute such model.
 
@@
 \begin{itemize}
-\item $\Phi=\tup{F,A[\cdot]}$ is a {\em classical planning frame} where the semantics of each action $a\in A[\cdot]$ is unknown; i.e. the corresponding $\tup{\rho,\theta}$ functions are undefined. 
-\item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the partial observation of a trajectory $\tau(\pi,s_0)$ defined within the {\em classical planning frame} $\Phi$.
+\item $\Phi=\tup{F,A[\cdot]}$ is a {\em classical planning frame} where the semantics of each action $a\in A[\cdot]$ is unknown; i.e.
+the corresponding $\tup{\rho,\theta}$ functions are undefined.
+We say that an {\em action model} $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$.
+\item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the partial observation of a trajectory $\tau(\pi,s_0)$ within the {\em classical planning frame} $\Phi$.
 \end{itemize}
 
-Before formalizing the solution to this task, i.e. the {\em least-commitment} action model, we introduce several necessary definitions.
-
-\begin{definition}[Model Space]
-Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ the model space $M$ is the set of possible models for the actions in $A[\cdot]$ such that: (1), any model $\mathcal{M}\in M$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$ and (2), for every $\mathcal{M}\in M$ the $\tup{\rho,\theta}$ functions are defined in the set of state variables $F$.
+Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
+A model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}(\tau)$.
+\begin{definition}[The {\em least-commitment} action model]
+Given a $\tup{\Phi,\mathcal{O}(\tau)}$ task (and let $M$ be the set of action models that represents the full space of possible action models for the actions in $A[\cdot]\in \Phi$), the {\em least-commitment} action model is the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ {\em explains} the input observation.
 \end{definition}
 
-Now, we define a {\em partially specified action model} inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}.
-\begin{definition}[Partially specified action model]
-A {\em partially specified action model} is a subset of models in a given {\em model space $M$}.
-\end{definition}
-If the {\em partially specified action model} is a singleton, it represents a {\em fully specified action model}.
-On the other hand, if its size is $|M|$ the {\em partially specified action model} represents the full model space.
-
-Now we are ready to define the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$.
-\begin{definition}[The {\em least-commitment} action model]
-Given a $\tup{\Phi,\mathcal{O}(\tau)}$ task and the {\em partially specified action model} $M$ that represents the full space of possible action models for the actions in $A[\cdot]\in \Phi$, then the {\em least-commitment} action model is another {\em partially specified action model} that represents the largest subset of models $M^*\subseteq M$ such that every model $\mathcal{M}\in M^*$ {\em explains} the input observation.
-\end{definition}
+This work focuses on the particular task of computing the {\em least-commitment} action model when action models are specified as \strips\ action schemata.
 
 \subsection{The space of \strips\ action models}
-Despite previous definitions are general, this work focuses on the particular kind of action models that are specified as \strips\ action schemata.
-
 {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
@@
 For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\em Typing constraints} and {\it state invariants} are also constraints of this kind~\cite{fox1998automatic}.
-\item {\em Observation constraints}. A sequence of state observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
+{\it State invariants} are also constraints of this kind~\cite{fox1998automatic}.
+\item {\em Observation constraints}. An observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
 \end{enumerate}
-
-\subsection{A propositional encoding for {\em partially specified} \strips\ models}
-First we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\ so is more compact that the one previuosly proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
-In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
-
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
-There is a total number of $2^{2\times|11|}=4,194,304$ different models for that schema.
 
@@
 
-Now we show that this encoding can be extended to the {\em knwoledge level} to compactly represent a partially specified action model.
-The extension replaces each proposition {\tt\small $pre\_e\_\xi$} with two new propositions {\tt\small $K\pre\_e\_\xi$} and {\tt\small $K\neg\pre\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is a precondition of $\xi$.
-The same is done for each {\tt\small $eff\_e\_\xi$} that is replaced by {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is an effect of $\xi$ or that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not an effect of $\xi$.
-Figure~\ref{fig:partial} shows a {\em partial specification} of the  {\tt\small stack(?v1,?v2)} schema that follows our propositional representation.
+In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+This encoding exploits the syntactic constraints of \strips\ so is more compact that the one previuosly proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
 
-With respect to this formalization, the {\em full space} of possible \strips\ schemas for $xi$ is compactly represented as {\tt\small $\bigcap_{e\in{\mathcal I}_{\Psi,\xi}} \neg K\pre\_e\_\xi\wedge \neg K\neg\pre\_e\_\xi \wedge\neg K\eff\_e\_\xi\wedge \neg K\neg\eff\_e\_\xi$}.
+\subsection{{\em Partially specified} \strips\ action models}
+A set of action models can be defined {\em explicitely}, by enumerating all the models that belong to the set or {\em implictely}, by enumerating the constraints that must satisfy any model that belongs to the set.
+Inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}, we introduce here {\em partially specified} \strips\ action models, a formalism for the {\em implicit} representation of a set of \strips\ schema.
+
+We show now that extending to the {\em knwoledge level} our propositional encoding of \strips\ action schemes, we can compactly represent a set of \strips\ schema.
+The extension defines, for each proposition {\tt\small $pre\_e\_\xi$}, with two propositions {\tt\small $K\pre\_e\_\xi$} and {\tt\small $K\neg\pre\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is a precondition of $\xi$ and that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not a precondition of $\xi$.
+Likewise for each {\tt\small $eff\_e\_\xi$} proposition we define these two, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is an effect of $\xi$ and that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not an effect of $\xi$.
+With respect to this formalization, the {\em full space} of possible \strips\ schemas for $\xi$ is compactly represented as {\tt\small $\bigcap_{e\in{\mathcal I}_{\Psi,\xi}} \neg K\pre\_e\_\xi\wedge \neg K\neg\pre\_e\_\xi \wedge\neg K\eff\_e\_\xi\wedge \neg K\neg\eff\_e\_\xi$}.
+
+Figure~\ref{fig:partial} shows a {\em partially specified} \strips\ model that represents of a set of action models for the {\tt\small stack(?v1,?v2)} schema.
+This {\em partially specified} \strips\ model represents a set of $2^5$ models since (1), the actual five effects of the action are {\em unkwnon}, (2) preconditions are known to appear as in the actual definition of the {\tt\small stack(?v1,?v2)} schema for the blocksworld and (3), it is {\em known} that the preconditions and effects that are not part of the actual definition of the {\tt\small stack(?v1,?v2)} schema are not part of this {\em partially specified} \strips\ model.
 
@@
  \begin{lstlisting}
+(K$\pre$_holding_v1_stack) (K$\pre$_clear_v2_stack)
 (K$\neg\pre$_holding_v2_stack) (K$\neg\pre$_clear_v1_stack)
@@
   \end{tiny}  
- \caption{\small {\em Partial specification} of the  {\tt\small stack(?v1,?v2)} schema that follows our propositional representation.}
+ \caption{\small {\em Partially specified} \strips\ model that represents of a set of action models for the {\tt\small stack(?v1,?v2)} schema.}
 \label{fig:partial}

### 2018-12-19 doc/ijcai-19-conformant/conformant-learning.tex
first updates added after discussing the points made by blai

@@
       \item $\{test_j\}_{1\leq j\leq m}$, indicating the state observation $s_j\in\mathcal{O}(\tau)$ where the action model is validated
-      \item Fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} encoding the {\em knowledge level} representation of the space of possible \strips\ action models.
+      \item The {\em knowledge level} fluents {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} encoding the space of possible {\em partially specified} action models.
       \end {itemize}
@@
 Figure~\ref{fig:editable} shows the PDDL encoding of an {\em editable} {\tt\small stack(?v1,?v2)} schema.
-This editable schema behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional} when the set of fluents {\tiny\tt (Kpre\_holding\_v1\_stack) (Kpre\_clear\_v2\_stack) (Keff\_holding\_v1\_stack) (Keff\_clear\_v2\_stack) (Keff\_clear\_v1\_stack) (Keff\_handempty\_stack) (Keff\_on\_v1\_v2\_stack)} hold at the current state.
+This editable schema behaves exactly as the original PDDL schema defined in Figure~\ref{fig:propositional} when the set of fluents {\tiny\tt (Kpre\_holding\_v1\_stack) (Kpre\_clear\_v2\_stack) (Keff\_holding\_v1\_stack) (Keff\_clear\_v2\_stack) (Keff\_clear\_v1\_stack) (Keff\_handempty\_stack) (Keff\_on\_v1\_v2\_stack)} hold at the current state as well as the remaining {\tt\small $K\neg\pre\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} for all the preconditions and effects that are not part of the actual {\tt\small stack(?v1,?v2)} schema.
 Formally, given an operator schema $\xi\in\mathcal{M}$ its {\em editable} version is:

### 2018-12-19 doc/ijcai-19-conformant/conformant-learning.tex
first updates added after discussing the points made by blai

@@
 
-\begin{definition}[Explaning a $\mathcal{O}(\tau)$ observation]
+\begin{definition}[Plan explanation]
 Given a {\em classical planning problem} $P$ and an observation $\mathcal{O}(\tau)$, a plan $\pi$ {\em explains} $\mathcal{O}(\tau)$ (denoted $\pi\mapsto\mathcal{O}(\tau)$) iff $\pi$ is a solution for $P$ that is {\em consistent} with the state trajectory constraints imposed by the sequence of partial states $\mathcal{O}(\tau)$.
@@
 
-Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a sequence of partial states $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we can build the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
-A model $\mathcal{M}$ {\em explains} a sequence of observations $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}(\tau)$.
+Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and an observation $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$ let $P_\mathcal{O}$ be the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
+\begin{definition}[Model explanation]
+A model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}(\tau)$.
+\end{definition}
+
 \begin{definition}[The {\em least-commitment} action model]
@@
 
+\begin{definition}[Partially specified model explanation]
+
+\end{definition}
+
 \subsection{Computing the {\em least-commitment} model via classical planning}
-Inspired by the {\em classical planning compilation} $K_{s_0}$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose solution induces the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$:
+Inspired by the {\em classical planning compilation} $K_0$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose solution induces the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$:
 \begin{itemize}

### 2018-12-19 doc/ijcai-19-conformant/conformant-learning.tex
first updates added after discussing the points made by blai

@@
 \begin{definition}[Model explanation]
-A model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}(\tau)$.
+A model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that is the {\em best explanation} for $\mathcal{O}(\tau)$.
 \end{definition}

### 2018-12-19 doc/ijcai-19-conformant/conformant-learning.tex
first updates added after discussing the points made by blai

@@
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
-Each classical planning action $a\in A$ has a precondition $\pre(a)\in\mathcal{L}(F)$ and a set of effects $\eff(a)\in\mathcal{L}(F)$.
+Each classical planning action $a\in A$ has a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $c(a)$.
 The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
@@
 A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
-A solution plan is optimal iff its length is minimal.
+A solution plan is {\em optimal} iff it minimizes the sum of action costs.
 
@@
 If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}(\tau)$.
+
+\subsection{Conformant planning}
+{\em Conformant planning} is planning with incomplete information about the initial state, no sensing, and validating that goals are achieved with certainty (despite the uncertainty of the initial state)~\cite{goldman1996expressive,smith1998conformant,bonet2000planning}.
+
+Syntactically, conformant planning problems are expressed in compact form through a set of state variables.
+A {\em conformant planning problem} can be defined as a tuple $P_c=\tup{F,A,\Upsilon,G}$ where $F$, $A$ and $G$ are the set of {\em fluents}, {\em actions} and {\em goals} (as previously defined for {\em classical planning}).
+Now $\Upsilon$ is a set of clauses over literals $l=f$ or $l=\neg f$ (for $f\in F$) that define the set of possible initial states.
+
+A solution to a conformant planning problem is an action sequence that maps each possible initial state into a goal state.
+More precisely, an action sequence $\pi=\tup{a_1, \ldots, a_n}$ is a {\em conformant plan} for $P_c$ iff, for each possible {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ s.t.
+$s_0$ is a valuation of the fluents in $F$ that satisfies $\Upsilon$, then $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$.
 
@@
 \begin{definition}[Partially specified model explanation]
-
+A partially specified action schema $\xi[\cdot]$ (that defines the set of action models $M_\xi$) {\em explains} an observation $\mathcal{O}(\tau)$, iff every model $\mathcal{M}\in M_{\xi}$ {\em explains} the input observation.
 \end{definition}

### 2018-12-19 doc/ijcai-19-conformant/conformant-learning.tex
first updates added after discussing the points made by blai

@@
 \begin{definition}[Model explanation]
-A model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that is the {\em best explanation} for $\mathcal{O}(\tau)$.
+A model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}(\tau)$.
 \end{definition}
@@
 In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\ so is more compact that the one previuosly proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+This encoding exploits the syntactic constraints of \strips\ so is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
 In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
@@
 \subsection{{\em Partially specified} \strips\ action models}
-A set of action models can be defined {\em explicitely}, by enumerating all the models that belong to the set or {\em implictely}, by enumerating the constraints that must satisfy any model that belongs to the set.
+A set of action models can be defined {\em explicitly}, by enumerating all the models that belong to the set or {\em impolitely}, by enumerating the constraints that must satisfy any model that belongs to the set.
 Inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}, we introduce here {\em partially specified} \strips\ action models, a formalism for the {\em implicit} representation of a set of \strips\ schema.
 
-We show now that extending to the {\em knwoledge level} our propositional encoding of \strips\ action schemes, we can compactly represent a set of \strips\ schema.
+We show now that extending to the {\em knowledge level} our propositional encoding of \strips\ action schemes, we can compactly represent a set of \strips\ schema.
 The extension defines, for each proposition {\tt\small $pre\_e\_\xi$}, with two propositions {\tt\small $K\pre\_e\_\xi$} and {\tt\small $K\neg\pre\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is a precondition of $\xi$ and that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not a precondition of $\xi$.
@@
 Figure~\ref{fig:partial} shows a {\em partially specified} \strips\ model that represents of a set of action models for the {\tt\small stack(?v1,?v2)} schema.
-This {\em partially specified} \strips\ model represents a set of $2^5$ models since (1), the actual five effects of the action are {\em unkwnon}, (2) preconditions are known to appear as in the actual definition of the {\tt\small stack(?v1,?v2)} schema for the blocksworld and (3), it is {\em known} that the preconditions and effects that are not part of the actual definition of the {\tt\small stack(?v1,?v2)} schema are not part of this {\em partially specified} \strips\ model.
+This {\em partially specified} \strips\ model compactly represents a set of $2^5$ models since (1), the actual five effects of the action are {\em unknown}, (2) preconditions are known to appear as in the actual definition of the {\tt\small stack(?v1,?v2)} schema for the {\em blocksworld} and (3), it is {\em known} that the preconditions and effects that are not part of the actual definition of the {\tt\small stack(?v1,?v2)} schema are not part of this {\em partially specified} \strips\ model.
 
@@
 
+\subsection{Computing the {\em least-commitment} model via conformant planning}
+The compilation for learning \strips\ action models defined by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} can be used to compute a {\em partially specified model} that explains a given observation by producing a conformant planning task instead of a classical planning task.
+To do so the initial state $\Upsilon$ of the conformant planning task does not code the {\em empty} action model but the full space of action models.
+In more detail, the clauses in $\Upsilon$ comprises:
+      \begin{enumerate}
+      \item The {\em unit clauses} given by the fluents that hold in the initial state $I=s_0$.
+      \item The clauses representing that the actual value of fluents {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} is unknown.
+In other words, that any model from the \strips\ space of models (following the previously mentioned syntactic constraints) can initially be part of the {\em least-commitment} action model.
+Formally, for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$, then $\Upsilon$ includes these two clauses:
+            \begin{itemize}
+            \item {\tt\small pre\_e\_$\xi$} $\vee$ {\tt\small $\neg$pre\_e\_$\xi$}.
+            \item {\tt\small eff\_e\_$\xi$} $\vee$ {\tt\small $\neg$eff\_e\_$\xi$}.
+            \end{itemize}
+            One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier for a specific domain.
+      \end{enumerate}
+A conformant plan that solves this task is a {\em partially specified model} that {\em explains} the input observation.
+To compute the {\em least-commitment} action model, we have to associate a cost of 1 to the actions that {\em program} a given precondition or effect of an action schema, a cost of 0 to the remaining actions, and compute the {\em optimal} plan for the resulting conformant planning task.
+
 \subsection{Computing the {\em least-commitment} model via classical planning}
-Inspired by the {\em classical planning compilation} $K_0$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose solution induces the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$:
+Inspired by the {\em classical planning compilation} $K_0$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose optimal solution induces the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$:
 \begin{itemize}

### 2018-12-21 doc/aij18/major-revision/learning_task_def.tex
Section 3 updated

\section{Learning task: on the use of planning for solving the task.}
\label{task_definition}

In this section, we firstly define the concept of \emph{learning task}. Subsequently, we examine the particularities of the learning task to the different types of plan traces according to the observed state trajectories and action sequences. The analysis will serve to justify the use of planning for solving the learning task as well as to highlight the principal distinctive features of our approach \FAMA with respect to the related work reviewed in Section~\ref{related_work}.

\vspace{0.1cm}

\FAMA addresses the learning and evaluation of PDDL action models that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}. An \strips\ action model is a tuple $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$ where:

\begin{itemize}
	\item The name, $name(\xi)$, and parameters, $pars(\xi)$, of the action model define the {\em header} of the model.
	\item $pre(\xi)$, $del(\xi)$ and $add(\xi)$ represent the preconditions, negative effects and positive effects of the action model, respectively, \textcolor[rgb]{1.00,0.00,0.00}{which follow the set of syntactic \strips\ constraints defined in section \ref{basic_planning}; specifically,} $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
\end{itemize}

As an example, Figure~\ref{fig:stack} shows the action model of the {\em stack} operator from the four-operator {\em blocksworld} domain~\cite{slaney2001blocks} encoded in PDDL.

\begin{figure}[hbt!]
	\begin{footnotesize}
		\begin{verbatim}
		(:action stack
		:parameters (?v1 ?v2 - object)
		:precondition (and (holding ?v1) (clear ?v2))
		:effect (and (not (holding ?v1)) (not (clear ?v2)) (handempty) (clear ?v1) (on ?v1 ?v2)))
		\end{verbatim}
	\end{footnotesize}
	\caption{PDDL encoding of the action model of the {\em stack} operator from the four-operator {\em blocksworld} domain.}
	\label{fig:stack}
\end{figure}

Our {\em learning task} consists in learning a classical domain model by observing one or more agents acting in a world definable by a {\em classical planning frame} $\Phi=\tup{F,A}$. The learning task is formalized by the pair $\Lambda=\tup{\mathcal{M},\tau}$:

\begin{itemize}
\item $\mathcal{M}$ is the {\bf initial domain model} (set of action models). This set is {\em empty}, when learning from scratch, or {\em partially specified}, when some fragments of the action models are known a priori.
\item $\tau$ is the observed {\bf plan trace} such that:
\begin{enumerate}
 \item Observations in $\tau$ are {\em noiseless}, meaning that if the value of a fluent or an action is observed in $\tau$, then the observation is correct.
\item The initial state $s_0\in\tau$ is a {\em fully observed} state including positive and negative fluents; i.e.~$|s_0|=|F|$. Consequently, the corresponding set of predicates $\Psi$ and objects $\Omega$ that shape the fluents in $F$ can be inferred from $s_0$.
\item The header of an action model is either given by $\mathcal{M}$ or inferable from $\tau$. In the latter case, $\tau$ must contain at least one instantiation of the respective action model header.
\item We allow plan traces with \NO state trajectories and \NO action sequences. In the extreme, all actions and intermediate states may be missing, provided that the final state is at least partially observed. The least informative plan trace is thus $\tau = \tup {s_0,s_n}$.
\end{enumerate}
\end{itemize}

Ultimately, we can always assume that $\Lambda$ will contain the predicates $\Psi$ as well as the headers of the actions models, either explicitly provided in $\mathcal{M}$ or deducible from $\tau$.

\vspace{0.1cm}

A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a domain model $\mathcal{M}'$ that complies with the information of $\mathcal{M}$ and with the observed plan trace $\tau$. \textcolor[rgb]{1.00,0.00,0.00}{This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan encompass the plan trace $\tau$.}

Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\tau}$ corresponding to the observation of the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower. In this example $\tau=\langle s_0,${\small\tt (putdown\ B),(stack\ A\ B)}, $s_4\rangle$. \textcolor[rgb]{1.00,0.00,0.00}{Therefore, $\tau$ contains a \NO state trajectory because only the initial and final state are observed and the three intermediate states, $s_1$, $s_2$ and $s_3$, are missing; and a \PO action sequence where actions $a_2$ and $a_3$ are observed while $a_1$ and $a_4$ are unknown.} The initial domain model $\mathcal{M}$ only contains two of the four needed headers, but can be completed with the headers {\small\tt(putdown ?v1)} and {\small\tt(stack ?v1 ?v2)} inferred from $\tau$.

\begin{figure}[hbt!]
{\footnotesize\tt ;;;;;; Action headers in $\mathcal{M}$}
\begin{footnotesize}
\begin{verbatim}
(pickup ?v1) (unstack ?v1 ?v2)
\end{verbatim}
\end{footnotesize}
\vspace{0.2cm}
{\footnotesize\tt ;;;;;; Plan trace $\tau$}
\begin{footnotesize}
\begin{verbatim}
;;; Initial state observation
(clear B) (ontable A) (handempty) (on B A)
(not (clear A)) (not (ontable B)) (not (holding A)) (not (holding B))
(not (on A A)) (not (on A B)) (not (on B B))
\end{verbatim}
\end{footnotesize}

\begin{footnotesize}
\begin{verbatim}
;;; Action observation
(putdown B)

;;; Action observation
(stack A B)
\end{verbatim}
\end{footnotesize}

\begin{footnotesize}
	\begin{verbatim}
	;;; State observation
	(clear A) (on A B) (ontable B)
	\end{verbatim}
\end{footnotesize}

 \caption{\small Task $\Lambda=\tup{\mathcal{M},\tau}$ associated to the observation $\tau=\langle s_0,${\small\tt (putdown\ B),(stack\ A\ B)}, $s_4\rangle$}
\label{fig:example-plans}
\end{figure}

The definition of the learning task is extensible to the more general case where the execution of several plans from the same action models are observed. In this case, $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_{k}\}$ such that each $\tau\in \mathcal{T}$ is a plan trace that satisfies the previous 1--4 assumptions. In this case, the {\em learned} domain model $\mathcal{M}'$ must be compliant with the input model $\mathcal{M}$ and with every observed plan trace $\tau\in \mathcal{T}$.

\vspace{0.15cm}

\textcolor[rgb]{1.00,0.00,0.00}{The key to understanding the intricacies of solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ lies in the type of plan trace $\tau = \tup {s_0, \ldots, s_n}$. Let $\pi$ be the plan that solves the planning problem $\tup {s_0,s_n}$ with a learned domain model $\mathcal{M}'$ and $\tau$ be the observation of $\pi$. Since $\tau$ contains a (partial) observation of the execution of $\pi$, actions, fluents or states traversed by $\pi$ may be missing in $\tau$ (we will use $\tau_s$ and $\tau_a$ to refer to the observed states and observed actions of $\tau$, respectively). We distinguish two well-identified cases:}

\begin{enumerate}
\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ determines the length of $\pi$}. This happens when $\tau_a$ is a \FO action sequence (we know the actions of $\pi$), $\tau_s$ is a \FO state trajectory (the states induced by the actions of $\pi$ are fully known) or $\tau$ is a \POstar state trajectory (we have at least one fluent for every state induced by $\pi$, in which case we know there is a single action that transitions from every state of $\tau_s$).}

    \begin{itemize}
    \item The common assumption of having \FO action sequences in a learning task, as is the case of all the learning approaches presented in section \ref{related_work} except for \FAMA, is unrealistic in many domains as it commonly implies the existence of human observers that annotate the observed action sequences. In some real-world applications, the observed and collected data are sensory data (e.g., home automation, robotics) or images (e.g. traffic) and one cannot rely on human intervention for labeling actions. Actually, learning the executed actions can also be part of the action-model learning task. Learning, for instance, from unstructured data involves transforming the sensor or image information into a predicate-like format before applying the action-model learning approach, and it also requires the ability of identifying action symbols~\cite{AsaiF18}.
    \item The assumption of having \FO state trajectories means that the sensors are able to capture every state change at every instant, which is also typically unrealistic. Normally, the process of obtaining state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two subsequent sensor readings.
    \item \textcolor[rgb]{1.00,0.00,0.00}{The assumption of having \POstar state trajectories seems more appropriate to reflect a real-world sensor reading but still requires that at least one fluent of every state traversed by $\pi$ is captured by the sensors.}
    \end{itemize}

When the length of $\pi$ is given by $\tau$, the learning task is \textbf{SAT compilable}, and it is known that a boolean satisfiability problem is a NP-complete task~\cite{cook1971complexity}. This is the reason why SAT solving is a common technique in the approaches presented in section \ref{related_work}. Particularly, when $\tau_a$ is a \FO action sequence and $\tau_s$ is a \FO state trajectory, learning \strips\ action models is straightforward~\cite{jimenez2012review}. In this case the {\em pre-} and {\em post-states} of every action are available and so action effects are derived lifting the literals that change between the pre and post-state of the corresponding action executions. Likewise preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action.

\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ does not indicate the length of $\pi$}. This happens when $\tau_a$ and $\tau_s$ are both partially observed (\PO) or non-observable (\NO) action/state trajectories. In this case, we are unaware of the number of actions of $\pi$ and the number of states induced by $\pi$. This gives rise to a completely different scenario and a more challenging learning task that brings one key difference: the transition between two given observed states of $\tau$ may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbounded, and so the horizon of $\pi$ is no longer known. This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$. As it is also well known, the worst case complexity of \strips\ planning is PSPACE-complete so we can conclude our learning task becomes PSPACE-complete instead of NP-complete when combining \PO and \NO state trajectories and action sequences.}

   \textcolor[rgb]{1.00,0.00,0.00}{ In this particular scenario, the actual number of plans compliant with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown). Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.}

%\FAMA shows that classical planning is a suitable approach for this particular scenario
\end{enumerate}

%\FAMA represents one step ahead towards learning action models without assuming observed actions. The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging. This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now. Table \ref{tab:complex} shows that the worst case complexity of learning \strips\ action models becomes PSPACE-complete when combining \PO and \NO state trajectories and action sequences.

%\begin{table}[ht]
%\centering
%\begin{tabular}{c|c|c|c|c|}
%	& \multicolumn{4}{c|}{\emph{state observability}} \\ \cline{2-5}
%	\multirow{1}{*}{\emph{action}} & \FO & \POstar & \PO & \NO\\ {\emph{observability}} & & & & \\ \hline
%	\FO & - & NP-complete & NP-complete & NP-complete \\ \hline
%	\PO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
%	\NO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
%\end{tabular}
%\caption{Complexity of learning tasks according to the type of input trace}
%\label{tab:complex}
%\end{table}

### 2018-12-21 doc/aij18/major-revision/experiments.tex
Updated experiment results with new rovers domain

--- 2018-08-16 doc/aij18/experiments.tex
@@
-\section{Experimental results}
+\section{Experimental evaluation}
 \label{sec:experiments}
@@
 				Parking & 4 & 5 & 3 & 2  \\
+				Rovers & 9 & 25 & 6 & 3 \\
 				Satellite & 5 & 8 & 4 & 2  \\
@@
 				\hline
-				blocks & 0.86 & 0.67 & 0.86 & 0.67 & 0.8 & 0.44 & 0.84 & 0.59& 0.21 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
-				driverlog & 0.68 & 0.93 & 0.5 & 0.57 & 0.83 & 0.71 & 0.67 & 0.79& 0.27 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'unload-truck', 1, 2, 3), (u'board-truck', u'board-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
-				ferry & 0.78 & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 0.71 & 1.0& 0.4 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
-				floor-tile & 0.69 & 1.0 & 0.45 & 0.91 & 1.0 & 0.82 & 0.65 & 0.93& 1.42 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'up', 1, 2, 3), (u'down', u'down', 1, 2, 3), (u'right', u'right', 1, 2, 3), (u'left', u'left', 1, 2, 3), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
-				grid & 0.76 & 0.94 & 0.55 & 0.86 & 1.0 & 1.0 & 0.74 & 0.94& 0.7 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'pickup', 1, 2), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
-				gripper-strips & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 0.16 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
-				hanoi & 0.67 & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 0.73 & 1.0& 0.31 \\ % [(u'move', u'move', 1, 2, 3)]
-				miconic & 1.0 & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 0.8 & 1.0& 0.29 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 1, 2), (u'down', u'down', 1, 2)]
-				n-puzzle & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.88 & 1.0& 0.28 \\ % [(u'move', u'move', 1, 2, 3)]
-				parking & 0.73 & 0.79 & 0.7 & 0.78 & 1.0 & 0.78 & 0.78 & 0.78& 0.65 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-car-to-car', u'move-car-to-car', 1, 2, 3)]
-				satellite & 0.93 & 1.0 & 0.63 & 1.0 & 1.0 & 0.75 & 0.85 & 0.96& 0.36 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
-				transport & 1.0 & 1.0 & 0.71 & 1.0 & 1.0 & 0.6 & 0.9 & 0.9& 0.22 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
-				grid-visit-all & 0.67 & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 0.63 & 1.0& 0.87 \\ % [(u'move', u'move', 1, 2)]
-				zeno-travel & 0.92 & 0.79 & 0.6 & 0.86 & 0.8 & 0.57 & 0.78 & 0.75& 0.26 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 4, 5), (u'zoom', u'zoom', 1, 2, 3, 4, 5, 6)]
+				Blocks & 0.86 & 0.67 & 1.0 & 0.67 & 0.8 & 0.44 & 0.89 & 0.59& 0.24 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
+				Driverlog & 0.6 & 0.86 & 0.36 & 0.57 & 0.67 & 0.29 & 0.53 & 0.64& 0.58 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'unload-truck', 1, 2, 3), (u'board-truck', u'board-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
+				Ferry & 0.7 & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0& 0.37 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
+				Floortile & 0.69 & 1.0 & 0.55 & 1.0 & 1.0 & 0.82 & 0.69 & 0.95& 1.38 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'up', 1, 2, 3), (u'down', u'down', 1, 2, 3), (u'right', u'right', 1, 2, 3), (u'left', u'left', 1, 2, 3), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
+				Grid & 0.68 & 0.88 & 0.5 & 0.86 & 0.88 & 1.0 & 0.67 & 0.9& 0.65 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'pickup', 1, 2), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
+				Gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 0.18 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
+				Hanoi & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.8 & 1.0& 0.36 \\ % [(u'move', u'move', 1, 2, 3)]
+				Miconic & 1.0 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 0.84 & 1.0& 0.3 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 1, 2), (u'down', u'down', 1, 2)]
+				Npuzzle & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.88 & 1.0& 0.26 \\ % [(u'move', u'move', 1, 2, 3)]
+				Parking & 0.78 & 1.0 & 0.69 & 1.0 & 1.0 & 1.0 & 0.8 & 1.0& 0.24 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-car-to-car', u'move-car-to-car', 1, 2, 3)]
+				Rovers & 0.54 & 1.0 & 0.3 & 0.76 & 1.0 & 0.46 & 0.48 & 0.85 & 2.14 \\ % [(u'drop', u'drop', 1, 2), (u'navigate', u'navigate', 1, 2, 3), (u'sample-soil', u'sample-soil', 1, 2, 3), (u'sample-rock', u'sample-rock', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3, 4), (u'take-image', u'take-image', 1, 2, 3, 4, 5), (u'communicate-soil-data', u'communicate-soil-data', 1, 2, 3, 4, 5), (u'communicate-rock-data', u'communicate-soil-data', 1, 2, 4, 3, 5), (u'communicate-image-data', u'communicate-image-data', 1, 2, 3, 4, 5, 6)]
+				Satellite & 0.93 & 1.0 & 0.56 & 1.0 & 1.0 & 0.75 & 0.81 & 0.96& 0.4 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
+				Transport & 0.83 & 1.0 & 0.5 & 1.0 & 0.6 & 0.6 & 0.67 & 0.9& 0.19 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
+				Visitall & 1.0 & 1.0 & 0.25 & 0.5 & 1.0 & 1.0 & 0.57 & 0.8& 1.31 \\ % [(u'move', u'move', 1, 2)]
+				Zenotravel & 0.9 & 0.64 & 0.5 & 0.71 & 0.83 & 0.71 & 0.73 & 0.68& 0.25 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 4, 5), (u'zoom', u'zoom', 1, 2, 3, 4, 5, 6)]
 				\hline
-				\bf & 0.82 & 0.94 & 0.66 & 0.90 & 0.96 & 0.83 & 0.78 & 0.90 & 0.46 \\
+				\bf & 0.8 & 0.94 & 0.61 & 0.87 & 0.92 & 0.8 & 0.73 & 0.88 & 0.59 \\
 			\end{tabular}
@@
 			\hline
-			blocks & 1.0 & 0.89 & 0.9 & 1.0 & 1.0 & 0.89 & 0.96 & 0.93& 3.51 \\ % [(u'pick-up', u'put-down', 1), (u'put-down', u'pick-up', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
-			driverlog & 0.3 & 0.21 & 0.31 & 0.57 & 0.29 & 0.29 & 0.3 & 0.32& 24.41 \\ % [(u'load-truck', u'unload-truck', 1, 2, 3), (u'unload-truck', u'load-truck', 1, 2, 3), (u'board-truck', u'board-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 2, 1, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
-			ferry & 0.83 & 0.71 & 0.8 & 1.0 & 1.0 & 1.0 & 0.87 & 0.87& 4.39 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
-			floor-tile & - & - & - & - & - & - & - & - & - \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'change-color', 1, 3, 2), (u'down', u'up', 1, 2, 3), (u'right', u'up', 1, 3, 2), (u'left', u'down', 1, 3, 2), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
-			grid & - & - & - & - & - & - & - & - & - \\ % [(u'move', u'move', 1, 2), (u'pickup', u'move', 2, 1), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
-			gripper-strips & 1.0 & 0.67 & 0.8 & 1.0 & 1.0 & 1.0 & 0.92 & 0.86& 1.31 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
-			hanoi & 1.0 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.75& 1566.44 \\ % [(u'move', u'move', 1, 2, 3)]
-			miconic & 0.75 & 0.33 & 0.5 & 0.75 & 0.5 & 0.67 & 0.57 & 0.5& 0.86 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 2, 1), (u'down', u'down', 1, 2)]
-			n-puzzle & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 62.82 \\ % [(u'move', u'move', 1, 2, 3)]
-			parking & - & - & - & - & - & - & - & - & - \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-curb', 1, 3, 2), (u'move-car-to-curb', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-car', u'move-car-to-curb', 1, 2, 3)]
-			satellite & 0.78 & 0.5 & 0.57 & 0.8 & 0.33 & 0.25 & 0.63 & 0.52& 22.76 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
-			transport & 0.5 & 0.3 & 0.43 & 0.6 & 0.5 & 0.4 & 0.47 & 0.4& 31.07 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'pick-up', 1, 2, 3, 5, 4), (u'drop', u'drop', 1, 2, 3, 5, 4)]
-			grid-visit-all & 1.0 & 0.5 & 0.5 & 0.5 & 1.0 & 1.0 & 0.75 & 0.6& 3.11 \\ % [(u'move', u'move', 1, 2)]
-			zeno-travel & 0.67 & 0.29 & 0.75 & 0.43 & 0.75 & 0.43 & 0.71 & 0.36& 494.2 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 3, 2, 5, 4), (u'zoom', u'zoom', 1, 2, 3, 4, 5, 6)]
+			Blocks & 0.89 & 0.89 & 0.8 & 0.89 & 0.83 & 0.56 & 0.84 & 0.78& 0.77 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
+			Driverlog & 0.57 & 0.29 & 0.31 & 0.57 & 0.4 & 0.29 & 0.4 & 0.36& 7.35 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'unload-truck', 1, 2, 3), (u'board-truck', u'walk', 1, 2, 3), (u'disembark-truck', u'board-truck', 1, 2, 3), (u'walk', u'walk', 2, 1, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
+			Ferry & 0.83 & 0.71 & 0.36 & 1.0 & 1.0 & 1.0 & 0.62 & 0.87& 3.38 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
+			Floortile & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1001.6 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'change-color', 1, 3, 2), (u'down', u'up', 1, 2, 3), (u'right', u'up', 1, 3, 2), (u'left', u'down', 1, 3, 2), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
+			Grid & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1001.26 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'move', 2, 1), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
+			Gripper & 1.0 & 1.0 & 0.8 & 1.0 & 1.0 & 1.0 & 0.93 & 1.0& 0.17 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
+			Hanoi & 0.67 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 0.75& 132.69 \\ % [(u'move', u'move', 1, 2, 3)]
+			Miconic & 1.0 & 0.33 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.56& 0.71 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 1, 2), (u'down', u'down', 1, 2)]
+			Npuzzle & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86& 13.48 \\ % [(u'move', u'move', 1, 2, 3)]
+			Parking & 0.83 & 0.36 & 1.0 & 0.89 & 0.83 & 0.56 & 0.9 & 0.56& 160.64 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-car-to-car', u'move-car-to-car', 1, 2, 3)]
+			Rover & 0.43 & 0.73 & 0.24 & 0.47 & 0.56 & 0.38 & 0.39 & 0.61& 31.13 \\ % [(u'drop', u'drop', 1, 2), (u'navigate', u'navigate', 1, 3, 2), (u'sample-soil', u'sample-soil', 1, 2, 3), (u'sample-rock', u'sample-rock', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3, 4), (u'take-image', u'take-image', 1, 2, 3, 4, 5), (u'communicate-soil-data', u'communicate-rock-data', 1, 2, 4, 3, 5), (u'communicate-rock-data', u'communicate-rock-data', 1, 2, 3, 5, 4), (u'communicate-image-data', u'communicate-image-data', 1, 2, 3, 4, 6, 5)]
+			Satellite & 0.6 & 0.21 & 0.56 & 1.0 & 0.5 & 0.5 & 0.56 & 0.43& 11.55 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
+			Transport & 1.0 & 0.2 & 0.57 & 0.8 & 1.0 & 0.4 & 0.73 & 0.4& 23.39 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'drop', 1, 2, 3, 4, 5), (u'drop', u'pick-up', 1, 2, 3, 4, 5)]
+			Visitall & 0.67 & 1.0 & 0.5 & 0.5 & 1.0 & 1.0 & 0.67 & 0.8& 3.13 \\ % [(u'move', u'move', 1, 2)]
+			Zenotravel & 1.0 & 0.36 & 0.4 & 0.29 & 1.0 & 0.43 & 0.77 & 0.36& 226.27 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 5, 4), (u'zoom', u'zoom', 1, 2, 3, 5, 6, 4)]
 			\hline
-			\bf & 0.80 & 0.54 & 0.69 & 0.79 & 0.76 & 0.72 & 0.74 & 0.65 & 201.35
+			\bf & 0.70 & 0.48 & 0.57 & 0.69 & 0.74 & 0.59 & 0.64 & 0.56 & 174.5
 
@@
 			\hline
-			blocks & 0.5 & 0.56 & 0.5 & 0.33 & 0.75 & 0.33 & 0.55 & 0.41& 0.27 \\ % [(u'pick-up', u'put-down', 1), (u'put-down', u'pick-up', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
-			driverlog & 0.13 & 0.07 & 0.38 & 0.71 & 0.0 & 0.0 & 0.25 & 0.21& 0.98 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'unload-truck', 1, 2, 3), (u'board-truck', u'disembark-truck', 1, 2, 3), (u'disembark-truck', u'board-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
-			ferry & 0.5 & 0.29 & 0.5 & 0.5 & 0.67 & 0.5 & 0.55 & 0.4& 0.47 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
-			floor-tile & 0.34 & 0.64 & 0.5 & 0.36 & 0.44 & 0.73 & 0.39 & 0.59& 165.92 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'up', 1, 3, 2), (u'down', u'left', 1, 2, 3), (u'right', u'right', 1, 3, 2), (u'left', u'left', 1, 3, 2), (u'paint-up', u'paint-up', 1, 3, 2, 4), (u'paint-down', u'paint-down', 1, 3, 2, 4)]
-			grid & 0.47 & 0.41 & 0.38 & 0.43 & 0.25 & 0.29 & 0.39 & 0.39& 214.87 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'pickup', 1, 2), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 3, 2), (u'unlock', u'unlock', 2, 1, 3, 4)]
-			gripper-strips & 1.0 & 0.83 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.93& 0.2 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
-			hanoi & 0.6 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.78 & 0.88& 3.23 \\ % [(u'move', u'move', 1, 3, 2)]
-			miconic & 0.4 & 0.44 & 0.6 & 0.75 & 0.25 & 0.33 & 0.42 & 0.5& 0.25 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 2, 1), (u'down', u'down', 2, 1)]
-			n-puzzle & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 4.15 \\ % [(u'move', u'move', 1, 2, 3)]
-			parking & 0.67 & 0.57 & 0.43 & 0.33 & 0.5 & 0.44 & 0.56 & 0.47& 7.61 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 3, 2), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-car-to-car', u'move-car-to-car', 2, 1, 3)]
-			satellite & 0.5 & 0.14 & 0.5 & 0.6 & 0.75 & 0.75 & 0.57 & 0.35& 2.06 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 3, 2), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
-			transport & 0.6 & 0.3 & 0.38 & 0.6 & 0.5 & 0.2 & 0.47 & 0.35& 0.83 \\ % [(u'drive', u'drive', 1, 3, 2), (u'pick-up', u'drop', 1, 2, 3, 4, 5), (u'drop', u'pick-up', 1, 2, 3, 4, 5)]
-			grid-visit-all & 0.0 & 0.0 & 1.0 & 0.5 & 0.0 & 0.0 & 1.0 & 0.2& 1.24 \\ % [(u'move', u'move', 2, 1)]
-			zeno-travel & 0.86 & 0.43 & 0.29 & 0.29 & 0.33 & 0.14 & 0.53 & 0.32& 28.4 \\ % [(u'board', u'debark', 1, 2, 3), (u'debark', u'board', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 4, 5), (u'zoom', u'zoom', 1, 2, 3, 5, 6, 4)]
+			Blocks & 0.6 & 0.67 & 0.33 & 0.22 & 0.67 & 0.44 & 0.55 & 0.44& 0.26 \\ % [(u'pick-up', u'put-down', 1), (u'put-down', u'pick-up', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
+			Driverlog & 0.5 & 0.29 & 0.33 & 0.57 & 0.0 & 0.0 & 0.38 & 0.29& 0.88 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'board-truck', 1, 2, 3), (u'board-truck', u'unload-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 2, 1, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
+			Ferry & 0.5 & 0.43 & 0.5 & 0.25 & 0.67 & 0.5 & 0.55 & 0.4& 0.45 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
+			Floortile & 0.48 & 0.45 & 0.27 & 0.36 & 0.46 & 0.55 & 0.41 & 0.45& 58.38 \\ % [(u'change-color', u'change-color', 1, 3, 2), (u'up', u'left', 1, 3, 2), (u'down', u'up', 1, 3, 2), (u'right', u'down', 1, 3, 2), (u'left', u'right', 1, 3, 2), (u'paint-up', u'paint-up', 1, 3, 2, 4), (u'paint-down', u'paint-down', 1, 3, 2, 4)]
+			Grid & 0.25 & 0.24 & 0.33 & 0.43 & 0.14 & 0.14 & 0.25 & 0.26& 234.63 \\ % [(u'move', u'move', 2, 1), (u'pickup', u'putdown', 1, 2), (u'putdown', u'pickup', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 2, 1, 3, 4)]
+			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86& 0.16 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
+			Hanoi & 0.6 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.78 & 0.88& 6.32 \\ % [(u'move', u'move', 3, 2, 1)]
+			Miconic & 0.63 & 0.56 & 0.6 & 0.75 & 0.25 & 0.33 & 0.53 & 0.56& 0.25 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'down', 1, 2), (u'down', u'up', 2, 1)]
+			Npuzzle & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 1.52 \\ % [(u'move', u'move', 1, 2, 3)]
+			Parking & 0.57 & 0.29 & 0.2 & 0.11 & 0.8 & 0.44 & 0.53 & 0.28& 17.43 \\ % [(u'move-curb-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-curb', 1, 3, 2), (u'move-car-to-curb', u'move-car-to-car', 3, 2, 1), (u'move-car-to-car', u'move-car-to-car', 1, 3, 2)]
+			Rovers & 0.38 & 0.64 & 0.07 & 0.24 & 0.13 & 0.54 & 0.22 & 0.53& 1.74 \\ % [(u'drop', u'drop', 1, 2), (u'navigate', u'navigate', 1, 2, 3), (u'sample-soil', u'sample-rock', 1, 2, 3), (u'sample-rock', u'sample-soil', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3, 4), (u'take-image', u'take-image', 1, 2, 3, 4, 5), (u'communicate-soil-data', u'communicate-soil-data', 1, 2, 5, 3, 4), (u'communicate-rock-data', u'communicate-soil-data', 1, 2, 3, 5, 4), (u'communicate-image-data', u'communicate-image-data', 1, 2, 3, 4, 5, 6)]
+			Satellite & 0.86 & 0.43 & 0.43 & 0.6 & 0.75 & 0.75 & 0.67 & 0.52& 3.15 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
+			Transport & 0.29 & 0.2 & 0.38 & 0.6 & 0.67 & 0.4 & 0.39 & 0.35& 1.06 \\ % [(u'drive', u'drive', 1, 3, 2), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
+			Visitall & 0.0 & 0.0 & 1.0 & 0.5 & 0.0 & 0.0 & 1.0 & 0.2& 1.21 \\ % [(u'move', u'move', 2, 1)]
+			Zenotravel & 0.4 & 0.29 & 0.25 & 0.29 & 0.2 & 0.14 & 0.3 & 0.25& 17.48 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 5, 4), (u'zoom', u'zoom', 1, 2, 3, 4, 5, 6)]
+
 			\hline
-			\bf & 0.54 & 0.46 & 0.60 & 0.60 & 0.53 & 0.48 & 0.60 & 0.50 & 30.75
+			\bf & 0.54 & 0.46 & 0.51 & 0.53 & 0.52 & 0.48 & 0.57 & 0.48 & 22.99
 		\end{tabular}
@@
 			\hline
-			blocks & 0.84 & 0.59 & 0.84 & 0.64 \\
-			driverlog & 0.67 & 0.79 & 0.70 & 0.92 \\
-			ferry & 0.71 & 1.0 & 1.0 & 1.0 \\
-			floor-tile & 0.65 & 0.93 & 0.95 & 0.95 \\
-			grid & 0.74 & 0.94 & 1.0 & 0.98 \\
-			gripper-strips & 1.0 & 1.0 & 1.0 & 1.0 \\
-			hanoi & 0.73 & 1.0 & 0.91 & 1.0 \\
-			miconic & 0.8 & 1.0	& 1.0 & 1.0 \\
-			n-puzzle & 0.88 & 1.0 & 1.0 & 1.0 \\
-			parking & 0.78 & 0.78 & 0.97 & 0.84 \\
-			satellite & 0.85 & 0.96 & 0.96 & 1.0 \\
-			transport & 0.9 & 0.9 & 1.0 & 1.0 \\
-			grid-visit-all & 0.63 & 1 & 1.0 & 1.0 \\
-			zeno-travel & 0.78 & 0.75 & 0.85 & 0.96 \\
+			Blocks & 0.84 & 0.59 & 0.89 & 0.64 \\
+			Driverlog & 0.53 & 0.64 & 0.71 & 0.83 \\
+			Ferry & 0.6 & 1.0 & 0.96 & 1.0 \\
+			Floortile & 0.69 & 0.95 & 0.97 & 0.95 \\
+			Grid & 0.67 & 0.9 & 0.95 & 0.98 \\
+			Gripper & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Hanoi & 0.8 & 1.0 & 0.9 & 1.0 \\
+			Miconic & 0.84 & 1.0 & 1.0 & 1.0 \\
+			Npuzzle & 0.88 & 1.0 & 1.0 & 1.0 \\
+			Parking & 0.8 & 1.0 & 0.98 & 1.0 \\
+			Rovers & 0.31 & 0.72 & 0.94 & 0.99 \\ 
+			Satellite & 0.81 & 0.96 & 0.93 & 1.0 \\
+			Transport & 0.67 & 0.9 & 1.0 & 1.0 \\
+			Visitall & 0.57 & 1.0 & 1.0 & 1.0 \\
+			Zenotravel & 0.73 & 0.68 & 0.85 & 0.88 \\
 			\hline
-			& 0.78 & 0.9 & 0.94 & 0.95
+			& 0.72 & 0.88 & 0.94 & 0.95
 		\end{tabular}
@@
 			\hline
-			blocks & 0.55 & 0.41 & 0.9 & 0.86 \\
-			driverlog & 0.25 & 0.21 & 0.54 & 0.72 \\
-			ferry & 0.55 & 0.4 & 1.0 & 0.79 \\
-			floor-tile & 0.39 & 0.59 & - & - \\
-			grid & 0.39 & 0.39 & - & - \\
-			gripper-strips & 1.0 & 0.93 & 1.0 & 1.0 \\
-			hanoi & 0.78 & 0.88 & 0.89 & 1.0 \\
-			miconic & 0.42 & 0.5 & 0.89 & 0.85 \\
-			n-puzzle & 1.0 & 1.0 & 1.0 & 1.0 \\
-			parking & 0.56 & 0.47 & - & - \\
-			satellite & 0.57 & 0.35 & - & - \\
-			transport & 0.47 & 0.35 & 0.93 & 0.93 \\
-			grid-visit-all & 1.0 & 0.2 & 1.0 & 1.0 \\
-			zeno-travel & 0.53 & 0.32 & - & - \\
+			Blocks & 0.55 & 0.44 & 0.77 & 0.77 \\
+			Driverlog & 0.38 & 0.29 & 0.86 & 0.86 \\
+			Ferry & 0.55 & 0.4 & 0.82 & 0.53 \\
+			Floortile & 0.41 & 0.45 & - & - \\
+			Grid & 0.25 & 0.26 & - & - \\
+			Gripper & 1.0 & 0.86 & 1.0 & 1.0 \\
+			Hanoi & 0.78 & 0.88 & 0.89 & 1.0 \\
+			Miconic & 0.53 & 0.56 & 1.0 & 0.89 \\
+			Npuzzle & 1.0 & 1.0 & 1.0 & 1.0 \\
+			Parking & 0.53 & 0.28 & - & - \\
+			Rovers & 0.22 & 0.53 & - & - \\
+			Satellite & 0.67 & 0.52 & - & - \\
+			Transport & 0.39 & 0.35 & 0.94 & 1.0 \\
+			Visitall & 1.0 & 0.2 & 1.0 & 1.0 \\
+			Zenotravel & 0.57 & 0.48 & - & - \\
 			\hline
-			& 0.6 & 0.5 & 0.91 & 0.91
+			& 0.57 & 0.48 & 0.92 & 0.89
 		\end{tabular}

### 2018-12-21 doc/aij18/major-revision/learning_action_models.tex

--- 2018-07-24 doc/aij18/learning_task.tex
@@
 Subsequently, the compilation approach is fully detailed in section \ref{compilation} and the last section presents some theoretical properties of the compilation scheme.
-
-\FAMA addresses the learning and evaluation of PDDL action models that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-An \strips\ action model is a tuple $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
-
-\begin{itemize}
-	\item The name, $name(\xi)$, and parameters, $pars(\xi)$, of the action model define the {\em header} of the model.
-	\item $pre(\xi)$, $del(\xi)$ and $add(\xi)$ represent the preconditions, negative effects and positive effects of the action model, respectively, such that, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-\end{itemize}
-
-As an example, Figure~\ref{fig:stack} shows the action model of the {\em stack} operator from the four-operator {\em blocksworld} domain~\cite{slaney2001blocks} encoded in PDDL.
-
-\begin{figure}[hbt!]
-	\begin{footnotesize}
-		\begin{verbatim}
-		(:action stack
-		:parameters (?v1 ?v2 - object)
-		:precondition (and (holding ?v1) (clear ?v2))
-		:effect (and (not (holding ?v1)) (not (clear ?v2)) (handempty) (clear ?v1) (on ?v1 ?v2)))
-		\end{verbatim}
-	\end{footnotesize}
-	\caption{PDDL encoding of the action model of the {\em stack} operator from the four-operator {\em blocksworld} domain.}
-	\label{fig:stack}
-\end{figure}
-
-\subsection{Learning Task}
-\label{learning_task}
-
-Our {\em learning task} consists in learning classical planning action models by observing one or more agents acting in a world definable by a {\em classical planning frame} $\Phi=\tup{F,A}$.
-The learning task is formalized by the pair $\Lambda=\tup{\mathcal{M},\tau}$:
-
-\begin{itemize}
-\item $\mathcal{M}$ is the set of {\bf initial action models}.
-This set is {\em empty}, when learning from scratch, or {\em partially specified}, when some fragments of the action models are known a priori.
-\item $\tau$ is the observed {\bf plan trace} such that:
-\begin{enumerate}
- \item Observations in $\tau$ are {\em noiseless}, meaning that if the value of a fluent or an action is observed in $\tau$, then the observation is correct.
-\item The initial state $s_0\in\tau$ is a {\em fully observed} state including positive and negative fluents, i.e.~$|s_0|=|F|$.
-Consequently, the corresponding set of predicates $\Psi$ and objects $\Omega$ that shape the fluents in $F$ are inferrable from $s_0$.
-\item The header of an action model is either given by $\mathcal{M}$ or inferrable from $\tau$.
-In the latter case, $\tau$ must contain at least one instantiation of the respective action model header.
-%\item The set of actions $A$ is inferrable either from $\mathcal{M}$ or $\tau$.
-In other words, the action name and parameters that shape the actions in $A$ are given by $\mathcal{M}$ otherwise, $\tau$ contains at least one instantiation of every action model in $\mathcal{M}$.
-\item We allow plan traces with \NO state trajectories and action sequences.
-In the extreme, all actions and intermediate states may be missing, provided that the final state is at least partially observed.
-The least informative plan trace is thus $\tau = \tup {s_0,s_n}$.
-\end{enumerate}
-\end{itemize}
-
-Ultimately, we can always assume that $\mathcal{M}$ will contain predicates in $\Psi$ as well as the headers of the actions models, either explicitly provided in $\mathcal{M}$ or inferred from $\tau$.
-A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a set of action models $\mathcal{M}'$ that is compliant with the input models $\mathcal{M}$ and the observed plan trace $\tau$.
-
-Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\tau}$ corresponding to the observation of the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
-In this example $\tau=\langle s_0,${\small\tt (putdown\ B),(stack\ A\ B)}, $s_4\rangle$, which means that only the first and the last state are observed (the three intermediate states $s_1$, $s_2$ and $s_3$ are fully unknown) and that actions $a_2$ and $a_3$ are observed while $a_1$ and $a_4$ are unknown.
-The set of initial action models $\mathcal{M}$ only contains two of the four headers needed, but can be completed with the headers {\small\tt(putdown ?v1)} and {\small\tt(stack ?v1 ?v2)} inferred from $\tau$.
-
-\begin{figure}[hbt!]
-{\footnotesize\tt ;;;;;; Action headers in $\mathcal{M}$}
-\begin{footnotesize}
-\begin{verbatim}
-(pickup ?v1) (unstack ?v1 ?v2)
-\end{verbatim}
-\end{footnotesize}
-\vspace{0.2cm}
-{\footnotesize\tt ;;;;;; Plan trace $\tau$}
-\begin{footnotesize}
-\begin{verbatim}
-;;; Initial state observation
-(clear B) (ontable A) (handempty) (on B A)
-(not (clear A)) (not (ontable B)) (not (holding A)) (not (holding B))
-(not (on A A)) (not (on A B)) (not (on B B))
-\end{verbatim}
-\end{footnotesize}
-
-\begin{footnotesize}
-\begin{verbatim}
-;;; Action observation
-(putdown B)
-
-;;; Action observation
-(stack A B)
-\end{verbatim}
-\end{footnotesize}
-
-\begin{footnotesize}
-	\begin{verbatim}
-	;;; State observation
-	(clear A) (on A B) (ontable B)
-	\end{verbatim}
-\end{footnotesize}
-
- \caption{\small Task $\Lambda=\tup{\mathcal{M},\tau}$ associated to the observation $\tau=\langle s_0,${\small\tt (putdown\ B),(stack\ A\ B)}, $s_4\rangle$}
-\label{fig:example-plans}
-\end{figure}
-
-The definition of the learning task is extensible to the more general case where the execution of several plans from the same action models are observed.
-In this case, $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_{k}\}$ such that each $\tau\in \mathcal{T}$ is a plan trace that satisfies the previous 1--4 assumptions.
-In this case, the {\em learned} action models $\mathcal{M}'$ have to be compliant with the input models $\mathcal{M}$ and also with every observed plan trace $\tau\in \mathcal{T}$.
 
@@
 This is specially the case for static predicates that never change and, hence, can only be preconditions.
-In order to address this shortcoming, we apply a post-process based on the one proposed in \cite{long2018building}.
+In order to address this shortcoming, we apply a post-process based on the one proposed in \cite{kuvcera2018louga}.
 In their proposal, they obtain a full trace by applying the sequence of actions of the input trace and infer the preconditions from this fully observable trace.

### 2018-12-21 doc/aij18/major-revision/experiments.tex
Sentence Madagascar changed

@@
 The classical planner we used to solve the instances of $P_\Lambda$ that result from our compilations is {\sc Madagascar}~\cite{rintanen2014madagascar}.
-We used {\sc Madagascar} for two reasons:
+We used {\sc Madagascar} for several reasons:
 
   \begin{enumerate}
-  \item  its ability to deal with planning instances populated with dead-ends~\cite{lopez2015deterministic}.
-  \item  its ability to leverage the horizon of the solution plan when it is known.
-Specifically, when the length of the plan traces is known; i.e., \FO action sequences or \FO state trajectories, the horizon of the solution plan is also known, and this can be exploited by a SAT-based planner like MADAGASCAR.
-SAT-based planners are able to apply the actions for programming preconditions in parallel in a single planning step (and the same for the actions programming the effects) since these actions do not interact.
-Hence, we also know that the programming of the prefix of a solution plan is solvable in two steps, independently of the number of programming actions applied.
+  \item \textcolor[rgb]{1.00,0.00,0.00}{Other planners such as {\sc FastDownward} were also tested but provided worse experimental results}
+  \item \textcolor[rgb]{1.00,0.00,0.00}{A SAT-based planner like {\sc Madagascar} is particularly suitable for tasks where the observed plan trace $\tau$ determines the horizon of the solution plan (case 1 presented in section \ref{task_definition})}.
+In this case, the prefix of the solution plan is solvable in two steps because the actions for programming preconditions can be applied in parallel in a single step and the same for the actions programming the effects.
+  \item \textcolor[rgb]{1.00,0.00,0.00}{When the horizon of the plan is unknown (case 2 presented in section \ref{task_definition}), the ability of {\sc Madagascar} to deal with instances populated with dead-ends is very helpful~\cite{lopez2015deterministic}.}
   \end{enumerate}

### 2018-12-22 doc/aij18/major-revision/experiments.tex
Revised section 6

@@
 This section presents several experiments to evaluate the performance of \FAMA and the quality of the learned models.
-Whenever applicable, we will consider both NP-complete and PSPACE-complete scenarios to draw conclusions at both levels of worst case complexity.
+Whenever applicable, we will consider scenarios with both known and unknown plan horizon to draw conclusions at both levels of complexity.
 We also devote one experiment to show the suitability of the novel evaluation metrics introduced in section \ref{sec:evaluation}.
@@
 
-We evaluate \FAMA on 14 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, all taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+We evaluate \FAMA on 15 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, all taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
 Table \ref{tab:domain_features} presents the features of the tested domains that affect the size of the planning task $P_\Lambda$ that results from the compilation.
@@
 \item {\bf Plan traces}.
-For each domain, we generated 10 plan traces (each with 10 actions and 10 intermediate states) using random walks.
+For each domain, we generated 10 plan traces, each with 10 actions and 10 intermediate states, using random walks.
 Depending on the experiment, the traces are used for training or testing purposes (more details on this issue are provided at the particular experiment).
@@
 
-The experiment analyzes the evolution of the CPU-time and the Precision and Recall of the learned models wrt the GTM as $|\mathcal{T}|$ increases from 1 to 10 plan traces.
-To keep the experiment practicable, we introduced a 1000s timeout, after which the learning process is killed and a score of $0$ is given to both the Precision and Recall of the learned model.
+The experiment analyzes the evolution of the CPU-time and the precision and recall of the learned models wrt the GTM as $|\mathcal{T}|$ increases from 1 to 10 plan traces.
+To keep the experiment practicable, we introduced a 1000s timeout, after which the learning process is killed and a score of $0$ is given to both the precision and recall of the learned model.
 We defined two case studies:
@@
 \begin{itemize}
-	\item \textbf{\FO action sequence and \PO state trajectory}: This is the common case addressed by most of the state-of-the-art learning
-approaches, which corresponds to a NP-complete scenario.
+	\item \textbf{\FO action sequence and \PO state trajectory}: This is the common case addressed by most of the state-of-the-art learning approaches, which corresponds to a scenario where the plan horizon is given by the action sequence.
 In this experiment we assume a degree of observability of only 10\% for the state trajectory, meaning that each literal of a state has a 10\% chance of being observed.
-	\item  \textbf{\NO action sequence and \NO state trajectory}: This is a PSPACE-complete scenario where both the input action sequence and state trajectory are {\em empty}.
+	\item  \textbf{\NO action sequence and \NO state trajectory}: In this case study, both the input action sequence and state trajectory are {\em empty} and the length of the plan is unknown.
 Only the initial and final states are observed; i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
@@
 
-Figures \ref{fig:np_quality} and \ref{fig:np_time} show the quality of the models and computation time, respectively, for the NP-complete scenario.
-In Figure \ref{fig:np_quality} we see that Precision stabilizes at 0.86 after five traces whereas Recall stabilizes at 0.95 after three traces.
-These results show that \FAMA does, in fact, not need big amounts of input knowledge to learn sound and complete models as opposite to other approaches in the literature where models are learned using around a hundred of traces (see Table \ref{table:models_comparison2}).
+Figures \ref{fig:np_quality} and \ref{fig:np_time} show the quality of the models and computation time, respectively, for the first case study.
+\textcolor{red}{The values plotted in these figures are averages over the 15 domains.} In Figure \ref{fig:np_quality} we see that, after three traces, precision stabilizes at 0.84 whereas recall stabilizes at 0.95.
+These results show that \FAMA does, in fact, not need big amounts of input knowledge to learn sound and complete models as opposite to other approaches in the literature where models are learned using around 100 traces (see Table \ref{table:models_comparison2}).
 
@@
 Interestingly, we can observe an exponential increase in computation time for input sizes beyond five traces.
-Until an input size of four traces the computation time is below 1 sec but it reaches 153 secs when the input is composed of 10 traces.
+Until an input size of four traces the computation time is below 1 sec but it reaches 166 secs when the input is composed of 10 traces.
 These results match the expected performance of {\sc Madagascar} since this planner is known to struggle with plan horizons beyond 150-200 steps (in our case 160 steps corresponds to 8 traces since each trace has 10 actions and 10 intermediate states).
@@
 	\includegraphics[width=0.65\linewidth]{figures/input_size_100_10_precision.eps}
-	\caption{Precision and Recall when learning from [1-10] plan traces with \FO action sequences and \PO state trajectories with 10\% observability.}
+	\caption{Precision and recall when learning from [1-10] plan traces with \FO action sequences and \PO state trajectories with 10\% observability.}
 	\label{fig:np_quality}
@@
 
-Figure~\ref{fig:pspace_quality} displays the quality of the learned models in the PSPACE-complete scenario.
-As expectedly, the values of Precision and Recall are lower than in the NP-complete case.
+Figure~\ref{fig:pspace_quality} displays \textcolor{red}{the average precision and recall of the 15 learned models} in the scenario with unknown plan horizon.
+As expected, the quality of the learned models is lower than when the horizon of the plan is known.
 We can also observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
-The drop in the score is caused by an increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for Precision and Recall is assigned to these experiments.
-Figure \ref{fig:pspace_time}, on the other hand, reflects that the computation time of the PSPACE-complete scenario is also higher than the NP-complete scenario, which is again explained by the large number of timeouts.
+The drop in the score is caused by an increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for precision and recall is assigned to these experiments.
+\textcolor{red}{The first timeout is found at $|\mathcal{T}| = 3$ in the \emph{grid} domain, and the number of timeouts increases until it reaches 6 with $|\mathcal{T}| = 10$.} Figure \ref{fig:pspace_time}, on the other hand, reflects that the computation time of the first case study is also higher than in the second case, which is explained by both the higher complexity and the large number of timeouts.
 
-The conclusions we draw from these experiments is that learning with few input samples yield action models that are fairly sound and almost totally complete in the NP-complete scenario and less accurate in the PSPACE-complete scenarios.
-In the latter, while being timeouts the main cause of the drop in the score, we must point out that pure syntax-based metrics are not adequate to evaluate such under-constrained tasks since the phenomenon of \emph{reformulation} occurs and this largely impacts the results.
+The conclusions we draw from these experiments is that, when the plan horizon is known, learning with few input samples yield action models that contain 95\% of the preconditions and effects of the GTM plus some extra ones as indicated by the values of precision and recall.
+With unkown plan horizons, on the other hand, the learned models are generally more different from the GTM; while timeouts are the main cause of the drop in the score, we must point out that pure syntax-based metrics are not adequate to evaluate such under-constrained tasks since the phenomenon of \emph{reformulation} occurs and this largely impacts the results.
 We will provide experimental evidence of this in section \ref{minimal}.
@@
 	\includegraphics[width=0.65\linewidth]{figures/input_size_0_0_precision.eps}
-	\caption{Precision and Recall when learning from [1-10] plan traces with \NO action sequences and \NO state trajectories.}
+	\caption{Precision and recall when learning from [1-10] plan traces with \NO action sequences and \NO state trajectories.}
 	\label{fig:pspace_quality}
@@
 In this section we analyze the performance of \FAMA compared to \ARMS, one of the most well-known approaches to learning planning models.
-\ARMS, as well as most of the existing current learning systems, works under the assumption of plan traces with \FO action sequences and \NO state trajectories and therefore is not able to handle the PSPACE-complete scenario.
+\ARMS, as well as most of the existing current learning systems, works under the assumption of plan traces with \FO action sequences and \NO state trajectories and therefore is not able to handle the scenarios where the plan horizon is unknown.
 We will thereby restrict the experimentation to the cases manageable by \ARMS.
@@
 
-Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of Precision and Recall.
-The horizontal axes represent the degree of observability and vertical axes show the average Precision (Figure \ref{fig:comparison_precision}) and Recall (Figure \ref{fig:comparison_recall}) computed over the 14 tested domains.
-Remarkably, \FAMA dominates in terms of Precision in all cases except for the \FO state trajectories.
-Particularly, the models learned by \FAMA are between 15\% to 24\% more precise than those learned by \ARMS.
-A similar trend is observed for Recall (Figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our learned models are more complete.
+Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare \FAMA and \ARMS in terms of precision and recall.
+The horizontal axes represent the degree of observability and vertical axes show the average precision (Figure \ref{fig:comparison_precision}) and recall (Figure \ref{fig:comparison_recall}) computed over the 15 tested domains.
+Remarkably, \FAMA dominates in terms of precision in all cases except for the \FO state trajectories.
+Particularly, the models learned by \FAMA are between 13\% to 34\% more precise than those learned by \ARMS.
+A similar trend is observed for recall (Figure \ref{fig:comparison_recall}), where the difference is even larger, meaning that our learned models are more complete.
 
 The results highlight that \FAMA outperforms \ARMS when very few plan traces are available.
-This by no means is conclusive that \FAMA is overall better in NP-complete scenarios but only that is able to learn better with very limited input knowledge (actually, Figure~\ref{fig:np_time} reflects the exponential behaviour of \FAMA with more than five traces).
+This by no means is conclusive that \FAMA is overall better in NP-complete scenarios but only that it is able to learn better with very limited input knowledge (actually, Figure~\ref{fig:np_time} reflects the exponential behaviour of \FAMA with more than five traces).
 
@@
 To that end, we will limit the input to only two plan traces and analyze the results under different degrees of observability.
-We evaluate three case studies, the first one is an NP-complete scenario and the other two are PSPACE-complete:
+We evaluate three case studies:
 
@@
 All tables in this section (Tables \ref{tab:results_minimum_100_10}, \ref{tab:results_minimum_30_30} and \ref{tab:results_minimum_0_0}) follow the same structure.
-Precision ({\bf P}) and Recall ({\bf R}) scores are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative Effects ({\bf Del}), and also globally ({\bf Global}).
+precision ({\bf P}) and recall ({\bf R}) scores are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative effects ({\bf Del}), and also globally ({\bf Global}).
 The last column reports the computation time (in seconds) needed to obtain the learned models.
@@
 Table \ref{tab:results_minimum_100_10} shows the results of the first case study.
-Recall scores are generally higher than the precision ones, and, in fact, the models learned for six out of the 14 domains were perfectly complete.
+Recall scores are generally higher than the precision ones, and, in fact, the models learned for six out of the 15 domains were perfectly complete.
 Although precision is overall lower, it is interesting to notice that the learned sets of negative effects are mostly flawless.
-With regards to the computation time, we can observe times are below one second in all cases except for the {\em floor-tile} domain.
+With regards to the computation time, we can observe times are below one second in most cases except for some of the more complex domains.
 
@@
 		\end{center}
-	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \FO action sequences and \PO state trajectories with 10\% observability.}
+	\caption{\small Precision and recall scores for learning tasks with \FO action sequences and \PO state trajectories with 10\% observability.}
 	\label{tab:results_minimum_100_10}
@@
 
-Table \ref{tab:results_minimum_30_30} gathers the results of the PSPACE-complete case study with 30\% observability.
+Table \ref{tab:results_minimum_30_30} gathers the results of the case study with 30\% observability.
 We can see in the table that the scores of some domains are missing.
-This is the case of {\em floor-tile} and {\em parking}, which not only are fairly complex domains, but also categorized as \emph{puzzle-like} domains, a feature that is known for putting a strain in the planners.
-Interestingly enough, we note the huge computation time of {\em hanoi}, which also qualifies as a \emph{puzzle-like} domain.
+This is the case of {\em floor-tile} and {\em grid}, which not only are fairly complex domains, but also categorized as \emph{puzzle-like} domains, a feature that is known for putting a strain in the planners.
+Interestingly enough, we note the high computation time of {\em hanoi} and {\em parking}, which also qualify as a \emph{puzzle-like} domains.
 Regarding quality, we find that the learned models retain a level of soundness similar to Table~\ref{tab:results_minimum_100_10} but the completeness is lower than in the previous case study.
-This is specially noticeable in the preconditions, where recall values drop from from 0.94 to 0.54.
+This is specially noticeable in the preconditions, where recall values drop from from 0.88 to 0.64.
 This is because the input actions act as strong constraints playing a key role on the closeness of the learned model to the GTM.
@@
 			Ferry & 0.83 & 0.71 & 0.36 & 1.0 & 1.0 & 1.0 & 0.62 & 0.87& 3.38 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
-			Floortile & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1001.6 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'change-color', 1, 3, 2), (u'down', u'up', 1, 2, 3), (u'right', u'up', 1, 3, 2), (u'left', u'down', 1, 3, 2), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
-			Grid & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0 & 0.0& 1001.26 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'move', 2, 1), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
+			Floortile & - & - & - & - & - & - & - & - & - \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'change-color', 1, 3, 2), (u'down', u'up', 1, 2, 3), (u'right', u'up', 1, 3, 2), (u'left', u'down', 1, 3, 2), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
+			Grid & - & - & - & - & - & - & - & - & - \\ % [(u'move', u'move', 1, 2), (u'pickup', u'move', 2, 1), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
 			Gripper & 1.0 & 1.0 & 0.8 & 1.0 & 1.0 & 1.0 & 0.93 & 1.0& 0.17 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
@@
 			\hline
-			\bf & 0.70 & 0.48 & 0.57 & 0.69 & 0.74 & 0.59 & 0.64 & 0.56 & 174.5
+			\bf & 0.81 & 0.56 & 0.66 & 0.8 & 0.86 & 0.68 & 0.74 & 0.64 & 47.28
 
@@
 	\end{center}
-	\caption{\small {\em Precision} and {\em recall} when learning with \PO action sequences and \PO state trajectories, 30\% observability in both cases.}
+	\caption{\small Precision and recall when learning with \PO action sequences and \PO state trajectories, 30\% observability in both cases.}
 	\label{tab:results_minimum_30_30}
@@
 
-We now analyze the PSPACE case study with \NO action sequences and \NO state trajectories (Table~\ref{tab:results_minimum_0_0}).
+We now analyze the case study with \NO action sequences and \NO state trajectories (Table~\ref{tab:results_minimum_0_0}).
 A first outstanding observation is that, contrary to what might be expected by looking at the previous table, we are able in this case to find solutions for all the domains.
@@
 This broader space of solutions is also stressed in a diminished quality of the learned models.
-Thus, despite the learned models are compliant with the input data, they are further from the original GTM.
-In Table~\ref{tab:results_minimum_0_0} we can observe the global values of Precision and Recall drop to 0.6 and 0.5, respectively.
+Thus, despite the learned models being compliant with the input data, they are further from the original GTM.
+In Table~\ref{tab:results_minimum_0_0} we can observe the global values of precision and recall drop to 0.57 and 0.48, respectively.
 
@@
 	\end{center}
-	\caption{\small {\em Precision} and {\em recall} scores for learning tasks with \NO action sequences and \NO state trajectories.}
+	\caption{\small Precision and recall scores for learning tasks with \NO action sequences and \NO state trajectories.}
 	\label{tab:results_minimum_0_0}
@@
 
-Our last experiment is devoted to compare the scores provided by the syntactic and semantic versions of Precision and Recall.
+Our last experiment is devoted to compare the scores provided by the syntactic and semantic versions of precision and recall.
 For that purpose, we will evaluate the models learned in Section \ref{minimal} both syntactically, using the GTM, and semantically, computing the set of action models closest to the learned models that is compliant with a testing set of five traces (see Definition \ref{compliant}).
-We must note that since we are using {\sc Madagascar}, a satisficing planner, the solution to the model evaluation may not be the closest compliant model, reason why the scores of sem-Precision and sem-Recall are approximate values.
+We must note that since we are using {\sc Madagascar}, a satisficing planner, the solution to the model evaluation may not be the closest compliant model, so the scores of sem-Precision and sem-Recall are approximate values.
 Our goal with this experiment is to gauge the suitability of the semantic metrics proposed in section \ref{sec:evaluation} with respect to their well-known counterparts.
@@
 Table \ref{tab:metric_comparison_100_10} shows the results of the first case study.
-Looking at the high scores of the syntactic metrics, specially the value of Recall, we can conclude that the learned models are, in fact, fairly similar to the GTM.
+Looking at the high scores of the syntactic metrics, specially the value of recall, we can conclude that the learned models are, in fact, fairly similar to the GTM.
 This supports our conclusion that no reformulation occurs in this case study, which also means that the space of possible solutions is restricted to models close to the GTM.
 The values of sem-Precision and sem-Recall are also very high across the table, which is exactly the desired behavior for these metrics given that solutions are very close to the GTM.
-In comparison, Recall and sem-Recall show similar scores, while sem-Precision is significantly higher than Precision, thus showing that the sem-Precision is more lenient towards extra preconditions or effects.
-This is in line with the results of the previous experiments, where the common appearance of redundant or implicit preconditions in the learned models is penalized by the Precision metric.
+In comparison, recall and sem-Recall show similar scores, while sem-Precision is significantly higher than precision, thus showing that the sem-Precision is more lenient towards extra preconditions or effects.
+This is in line with the results of the previous experiments, where the common appearance of redundant or implicit preconditions in the learned models is penalized by the precision metric.
 We can interpret this phenomenon as a manifestation of the qualification problem~\cite{GinsbergS88}.
 For instance, the model learned for the {\tt\small move} action of the {\em hanoi} domain specifies that both the origin and destination disks must be bigger than the one moving, but the GTM contains only one of these preconditions.
-This learned model is semantically correct but syntactically different from the GTM and hence penalized by the Precision metric.
+This learned model is semantically correct but syntactically different from the GTM and hence penalized by the precision metric.
 
@@
 One first observation is the impossibility of applying a semantic evaluation in some of the most complex domains with five traces.
-Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger in this PSPACE-complete scenario.
-Comparing the scores of both versions, we find that learned models that achieved mediocre scores when using the GTM as reference (syntactic metrics), are in fact reasonably sound and complete, reaching an overall score of 0.91 in both sem-Precision and sem-Recall.
+Contrary to the previous case study, the difference between the syntactic and semantic metrics is larger in this scenario with unknown plan horizon.
+Comparing the scores of both versions, we find that learned models that achieved mediocre scores when using the GTM as reference (syntactic metrics), are in fact reasonably sound and complete, reaching overall scores of 0.92 and 0.89 in sem-Precision and sem-Recall.
 This is an indication that the models learned by our approach, despite syntactically different from the GTM, require very few editions to explain the testing set of traces.

### 2018-12-22 doc/aij18/major-revision/aij18.tex
Sections 1 and 2 updated
Added answers related to section 6 to the letter

@@
 \newcommand{\AMAN}{{\small {\sffamily AMAN}}\xspace}
+\newcommand{\LOUGA}{{\small {\sffamily LOUGA}}\xspace}
 \newcommand{\FAMA}{{\small {\sffamily FAMA}}\xspace}
@@
 \input{background}
-\input{motivation}
-\input{learning_task}
+%\input{motivation}
+\input{learning_task_def}
+\input{learning_action_models}
 \input{evaluation}
@@
 The distinctive characteristic of \FAMA over other state-of-the-art approaches is its ability to learn from minimal input knowledge and, more particularly, from minimal observability plan traces with no observed actions and no observed intermediate states.
-By lightening the input constraints, \FAMA opens up the way for action model learning to operate on real-world problems, as opposite to current approaches where the heavy input restrictions have limited their applicability to synthetic benchmarks.
+By relaxing the input constraints, \FAMA opens up the way for action model learning to operate on real-world problems, as opposite to current approaches where the heavy input restrictions have limited their applicability to synthetic benchmarks.
 Our approach is thus very well suited for learning action models in data-based applications where the only observable information is a possibly incomplete sequence of partially observed states.

### 2018-12-22 doc/aij18/major-revision/experiments.tex
Added answers related to section 6 to the letter

@@
 
-This experiment evaluates the impact of $\left|\mathcal{T}\right|$, the size of the input knowledge, in the performance of \FAMA in order to:
+This experiment evaluates the impact of $\left|\mathcal{T}\right|$, the size of the input knowledge, on the performance of \FAMA in order to:
 
@@
 As expected, the quality of the learned models is lower than when the horizon of the plan is known.
-We can also observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
-The drop in the score is caused by an increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for precision and recall is assigned to these experiments.
-\textcolor{red}{The first timeout is found at $|\mathcal{T}| = 3$ in the \emph{grid} domain, and the number of timeouts increases until it reaches 6 with $|\mathcal{T}| = 10$.} Figure \ref{fig:pspace_time}, on the other hand, reflects that the computation time of the first case study is also higher than in the second case, which is explained by both the higher complexity and the large number of timeouts.
+\textcolor{red}{The higher complexity of this setting is also reflected in the appearance of some timeouts when solving the learning task.
+The first timeout is found at $|\mathcal{T}| = 3$ in the \emph{grid} domain, and the number of timeouts increases until it reaches 6 with $|\mathcal{T}| = 10$.} We can observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
+The drop in the score is caused by the increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for precision and recall is assigned to these experiments.
+Figure \ref{fig:pspace_time}, on the other hand, reflects that the computation time of the first case study is also higher than in the second case, which is explained by both the higher complexity and the large number of timeouts.
 

### 2018-12-22 doc/aij18/major-revision/evaluation.tex
Further revisions on sections 5 and 6

@@
 	\item The GTM is unknown. This is the most common scenario in ML, where models are both learned and evaluated with respect to datasets.
-	\item The GTM is known but a test-based evaluation on a dataset is preferable (or needed as complementary to a syntactic evaluation).
+%	\item The GTM is known but a test-based evaluation on a dataset is preferable (or needed as complementary to a syntactic evaluation).
 As a rule of thumb, it is preferable to evaluate the learned models wrt a dataset because a learned model can be semantically correct though syntactically incorrect (different from the GTM).
+We refer to this phenomenon as \emph{model reformulation}.
+	\item \textcolor{red}{We are interested in measuring the ability of a model to explain a given plan trace, which is a good indicator of how the model will perform in actual planning tasks.} As a rule of thumb, it is preferable to evaluate the learned models wrt a dataset because a learned model can be semantically correct though syntactically incorrect (different from the GTM).
 We refer to this phenomenon as \emph{model reformulation}.

### 2018-12-22 doc/aij18/major-revision/experiments.tex
Further revisions on sections 5 and 6

@@
 For that purpose, we will evaluate the models learned in Section \ref{minimal} both syntactically, using the GTM, and semantically, computing the set of action models closest to the learned models that is compliant with a testing set of five traces (see Definition \ref{compliant}).
-We must note that since we are using {\sc Madagascar}, a satisficing planner, the solution to the model evaluation may not be the closest compliant model, so the scores of sem-Precision and sem-Recall are approximate values.
+We must note that since we are using {\sc Madagascar}, a satisficing planner, the solution to the model evaluation may not be the closest compliant domain model, so the scores of sem-Precision and sem-Recall are approximate values.
 Our goal with this experiment is to gauge the suitability of the semantic metrics proposed in section \ref{sec:evaluation} with respect to their well-known counterparts.

### 2018-12-23 doc/aij18/major-revision/learning_action_models.tex
Section 4 updated

@@
 
-This section details the \FAMA approach for learning action models.
-The notion of the learning task is defined in section \ref{learning_task} and the components of the compilation scheme are described in sections \ref{propositional_encoding} and \ref{conditional_effects}.
-Subsequently, the compilation approach is fully detailed in section \ref{compilation} and the last section presents some theoretical properties of the compilation scheme.
+\textcolor[rgb]{1.00,0.00,0.00}{Our proposal to address a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to transform $\Lambda$ into a planning task $P_{\Lambda}$.
+The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that determine the composition of the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ complies with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$.
+Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition, a positive effect or a negative effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that specify the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\tau$, including that the final reached state comprises $s_n$.
+Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the data generated through the action application must be validated with the data of $\tau$.}
+
+\vspace{0.1cm}
+
+The specification of $P_{\Lambda}$ requires a propositional encoding of the components of the action models $\xi$s, which is explained in the following section.
+The compilation approach is fully detailed in section \ref{compilation} and section \ref{properties} presents some theoretical properties of the compilation scheme.
 
@@
 
-%We say that two \strips\ operator schemes $\xi$ and $\xi'$ are {\em comparable} if both schemas have the same parameters so they share the same space of possible \strips\ models (formally, iff $pars(\xi)=pars(\xi'$).
-Therefore  we can claim that blocksworld operators {\tt stack} and {\tt unstack} are {\em comparable} while  {\tt stack} and {\tt pickup} are not.
-We say that two \strips\ action models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every $\xi\in\mathcal{M}$ to a comparable action schema $\xi'\in\mathcal{M'}$ and vice versa.
-
-In this section we formalize a propositional encoding of an \strips\ action model.
-This encoding is at the core of the \FAMA compilation approach for addressing the learning task defined in section \ref{learning_task}.
+In this section we formalize a propositional encoding of an \strips\ action model $\xi$.
+This encoding is at the core of the \FAMA compilation approach for addressing the learning task defined in section \ref{task_definition}.
 
@@
 
-We are now ready to define the propositional encoding of $pre(\xi)$, $del(\xi)$ and $add(\xi)$. For every $\xi$ and $p\in \Psi_{\xi}$, we create:
+We are now ready to define the propositional encoding of the \textcolor[rgb]{1.00,0.00,0.00}{model fluents} of $pre(\xi)$, $del(\xi)$ and $add(\xi)$. For every $\xi$ and $p\in \Psi_{\xi}$, we create:
 
 \begin{itemize}
-\item $pre_p(\xi)$: fluent formed by the combination of the prefixes {\small \texttt{pre}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
+\item $pre_p(\xi)$: \textcolor[rgb]{1.00,0.00,0.00}{model fluent} formed by the combination of the prefixes {\small \texttt{pre}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
 {\small \texttt{pre\_stack\_on\_v1\_v2}}, for $name(\xi)={\small\texttt{stack}}$ and $p={\small\texttt{on($v_1,v_2$)}}$)
-\item $del_p(\xi)$: fluent formed by the combination of the prefixes {\small \texttt{del}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
+\item $del_p(\xi)$: \textcolor[rgb]{1.00,0.00,0.00}{model fluent} formed by the combination of the prefixes {\small \texttt{del}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
 {\small  \texttt{del\_stack\_on\_v1\_v2}})
-\item $add_p(\xi)$: fluent formed by the combination of the prefixes {\small \texttt{add}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
+\item $add_p(\xi)$: \textcolor[rgb]{1.00,0.00,0.00}{model fluent} formed by the combination of the prefixes {\small \texttt{add}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
 {\small \texttt{add\_stack\_on\_v1\_v2}})
@@
 
-%The fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $p\in \Psi_{\xi}$, that represent the propositional encoding for the preconditions, negative and positive effects of an action model $\xi$.
-If a fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ holds, it means that $p$ is a precondition/negative/positive effect in that action model.
-
 For a given action model $\xi$, if a fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ holds in a state, it means that $p$ is a precondition/negative/positive effect of $\xi$.
-For instance, Figure~\ref{fig:encodedstack} shows the conjunction of fluents that represents the propositional encoding for the preconditions, negative effects and positive effects of the action model of the {\em stack} operator shown in Figure~\ref{fig:stack}.
+For instance, Figure~\ref{fig:encodedstack} shows the conjunction of \textcolor[rgb]{1.00,0.00,0.00}{model fluents} that represents the propositional encoding of the preconditions, negative effects and positive effects of the action model corresponding to the {\em stack} operator shown in Figure~\ref{fig:stack}.
 
@@
 
-\subsection{Classical planning with conditional effects}
-\label{conditional_effects}
-
-Conditional effects allow us to compactly define actions whose effects depend on the current state.
-Supporting conditional effects is now a requirement of the IPC~\cite{vallati:IPC:AIM2015} and many classical planners cope with conditional effects without compiling them away.
-
-An action $a\in A$ with conditional effects is defined as a set of {\em preconditions} $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a\in A$ is {\em applicable} in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
-\[
-triggered(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E,
-\]
-
-The result of applying action $a$ in state $s$ is the {\em successor} state $\theta(s,a)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$ where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
-
 \subsection{Compilation}
@@
 
-In section \ref{sec:motivation}, we exposed our reasons to solve the learning task via compiling it into a classical planning task.
 Our compilation scheme builds upon the approach presented in \cite{aineto2018learning} but \FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.
 
-Our approach to solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to compile $\Lambda$ into a classical planning task $P_{\Lambda}$ with conditional effects.
-The intuition behind the \FAMA compilation is that a solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output set of action models $\mathcal{M}'$ that solves $\Lambda$.
-Specifically, a solution plan $\pi_\Lambda$ serves two purposes:
+\vspace{0.1cm}
+
+A learning task $\Lambda=\tup{\mathcal{M},\tau}$ is compiled into a planning task $P_{\Lambda}$ with conditional effects in the context of a planning frame $\Phi=\tup{F,A}$.
+We use conditional effects because they allow us to compactly define actions whose effects depend on the current state.
+An action $a\in A$ with conditional effects is defined as a set of preconditions $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+An action $a\in A$ is applicable in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$; that is, $triggered(s,a)=\bigcup\limits_{C\rhd E\in\cond(a),C\subseteq s} E$.
+The result of applying $a$ in state $s$ follows the same definition of successor state, $\theta(s,a)$, introduced in section \ref{basic_planning} but applied to the conditional effects in $triggered(s,a)$.
+
+\vspace{0.25cm}
+
+A solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output domain model $\mathcal{M}'$ that solves the learning task $\Lambda$. Specifically, a solution plan $\pi_\Lambda$ serves two purposes:
 
 \begin{enumerate}
-\item {\bf To program the set of action models $\mathcal{M}'$}.
-$\pi_\Lambda$ comprises a plan {\em prefix} whose actions determine the predicates $p\in \Psi_{\xi}$ that belong to $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\mathcal{M}$.
-\item {\bf To validate the set of action models $\mathcal{M}'$}.
-$\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the validation of the observed plan trace $\tau$ with the programmed action models $\mathcal{M}'$.
+\item {\bf To specify the action models of $\mathcal{M}'$}.
+$\pi_\Lambda$ comprises a plan {\em prefix} whose actions set the predicates $p\in \Psi_{\xi}$ that belong to $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\mathcal{M}$.
+\item {\bf To validate the action models of $\mathcal{M}'$}.
+$\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the validation of the observed plan trace $\tau$ with the specified action models $\mathcal{M}'$.
 \end{enumerate}
@@
 %\subsection{Learning from observations of plan executions}
-Here we formalize the compilation for learning \strips\ action models with classical planning.
 Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$, with $\tau$ formed by an $n$-action sequence $\tup{a_1, \ldots, a_n}$ and a $m$-state trajectory $\tup{s_0, s_1, \ldots, s_m}$ ($\tau = \langle s_0, a_1, \ldots, a_n, s_m \rangle$), the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
@@
 
-\item $F_{\Lambda}$ contains:
+\item $F_{\Lambda}$ \textcolor[rgb]{1.00,0.00,0.00}{extends $F$ with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of $\tau$}.
+Specifically, $F_{\Lambda}$ contains:
 \begin{itemize}
 \item The set of fluents obtained from $s_0$; i.e., $F$.
-\item The fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section \ref{propositional_encoding}
+\item The \textcolor[rgb]{1.00,0.00,0.00}{model fluents} $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section \ref{propositional_encoding}
 \item A set of fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to represent the $i^{th}$ observable action of $\tau$.
@@
 
-\item $G_{\Lambda}$ includes the positive fluents $at_n$ and $test_m$.
-When these two goals are achieved by the solution plan $\pi_\Lambda$, we will be certain that the programmed action models are validated in all the actions and states observed in the input plan trace $\tau$.
+\item $G_{\Lambda}$ includes the positive literals $at_n$ and $test_m$.
+When these two goals are achieved by the solution plan $\pi_\Lambda$, we will be certain that the action models of $\mathcal{M'}$ are validated in all the actions and states observed in the input plan trace $\tau$.
 
-\item $A_{\Lambda}$ comprises three kinds of actions:
+\item $A_{\Lambda}$ includes \textcolor[rgb]{1.00,0.00,0.00}{three types of actions that give rise to the actions of $\pi_\Lambda$}. 
 \begin{enumerate}
-\item Actions for {\em programming} an action model $\xi\in\mathcal{M}$.
-These actions will form the prefix of the solution plan $\pi_\Lambda$ and they are aimed at generating the appropriate state configuration to shape $\xi$.
-Among the programming actions, we find:
+\item Actions for \textcolor[rgb]{1.00,0.00,0.00}{{\em inserting} a component (precondition, positive effect or negative effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models}.
+These actions will form the prefix of the solution plan $\pi_\Lambda$.
+Among the \emph{inserting} actions, we find:
 \begin{itemize}
 \item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
+\textcolor[rgb]{1.00,0.00,0.00}{A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $del_p$ nor $add_p$ exist in $\xi$.}
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
-\cond(\mathsf{programPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
+\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
 \end{align*}
@@
 \item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
+\textcolor[rgb]{1.00,0.00,0.00}{A positive effect is inserted in $\xi$ under the same conditions of a precondition insertion, and a negative effect is inserted in $\xi$ when neither $del_p$ nor $add_p$ appear in $\xi$ but $pre_p$ does.}
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programEff_{p,\xi}})=&\{\neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
-\cond(\mathsf{programEff_{p,\xi}})=&\{pre_{p}(\xi)\}\rhd\{del_{p}(\xi)\},\\
+\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
+\cond(\mathsf{insertEff_{p,\xi}})=&\{pre_{p}(\xi)\}\rhd\{del_{p}(\xi)\},\\
 & \{\neg pre_{p}(\xi)\}\rhd\{add_{p}(\xi)\}.
@@
 
-\item Actions for {\em applying} a programmed action model $\xi\in\mathcal{M}$ bound to objects $\omega\subseteq\Omega^{ar(\xi)}$.
-These actions will be part of the postfix of the solution plan $\pi_\Lambda$ and they execute $\xi$ according to the current state configuration, i.e., the values of $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$.
+\textcolor[rgb]{1.00,0.00,0.00}{For instance, given $name(\xi)=${\tt{\small stack}} and $C_{pre-stack}=\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_holding\_v2), (pre\_stack\_on\_v1\_v2),(pre\_stack\_clear\_v1),(pre\_stack\_clear\_v1),}}$\ldots \}$, the insertion of each item $c \in C_{pre-stack}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$ as long as $c \notin pre(\xi)$,  $c \notin add(\xi)$ and $c \notin del(\xi)$.
+The same applies to effects with respect to sets $C_{add-stack}$ and $C_{del-stack}$ that would include all fluents starting with prefix  {\tt{\small add}} and {\tt{\small del}}, respectively.}
+
+\vspace{0.1cm}
+
+\textcolor[rgb]{1.00,0.00,0.00}{Note that executing an insert action, e.g.{\tt{\small (insert\_pre\_stack\_holding\_v1)}}, will add the corresponding model fluent {\tt{\small (pre\_stack\_holding\_v1)}} to the successor state.
+Hence, the execution of the insert actions of $\pi_\Lambda$ yield a state containing the valuation of the model fluents that shape every $\xi \in \mathcal{M}$.
+For example, executing the insert actions that shape the action model $name(\xi)=${\tt{\small putdown}} leads to a state containing the positive literals {\tt{\small (pre\_putdown\_holding\_v1),(eff\_putdown\_holding\_v1),\\ (eff\_putdown\_clear\_v1),
+(eff\_putdown\_ontable\_v1),(eff\_putdown\_handempty)}}.}
+
+\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ specified by the insert actions bound to objects $\omega\subseteq\Omega^{ar(\xi)}$.
 Since action headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear in the same position.
-Figure~\ref{fig:compilation} shows the PDDL encoding for applying a programmed action model of the {\em stack} operator from the {\em blocksworld} domain.
 
@@
 
-%\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{f}(\xi)\implies p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))}\cup \{\neg mode_{val}\},\\
-%\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{f}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-%&\{add_{f}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi,f=p(pars(\xi))},\\
-%&\{mode_{prog}\}\rhd\{\neg mode_{prog}\},\\
-%&\{\emptyset\}\rhd\{mode_{val}\}.
+    These actions will be part of the postfix of the plan $\pi_\Lambda$ and \textcolor[rgb]{1.00,0.00,0.00}{they determine the application of the learned action models according to the values of the model fluents in the current state configuration}.
+Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator.
+\textcolor[rgb]{1.00,0.00,0.00}{Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$.
+Executing this action in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$.
+For example, if  $\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}}$\} \subset s$ then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
+Otherwise, a different set of precondition literals will be checked.
+The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.}
+
+\textcolor[rgb]{1.00,0.00,0.00}{    Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if $name(\xi)=${\tt{\small stack}} has been correctly programmed by the insert actions.
+Hence, while \textbf{insert actions} add the values of the \textbf{model fluents} that shape $\xi$, the \textbf{apply actions} add the values of the \textbf{fluents of $F$} that result from the execution of $\xi$.}
 
 \begin{figure}[hbt!]
+\begin{center}
 \begin{scriptsize}
@@
 \label{fig:compilation}
+\end{center}
 \end{figure}
+
+\vspace{0.1cm}
 
@@
 
-$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as in $\tau$.\\
+$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.\\
 
 \item Actions for {\em validating} the partially observed state $s_j\in\tau$, {\tt\small $1\leq j< m$}.
-These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the input plan trace $\tau$ follows after the apply actions.
+These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observable data of the input plan trace $\tau$ follows after the execution of the apply actions.
 
@@
 In \FAMA, when an action model $\xi$ is partially specified, the known preconditions and effects are encoded as fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$ set to true in the initial state $I_{\Lambda}$.
-In this case, the corresponding programming actions, $\mathsf{programPre_{p,\xi}}$ and $\mathsf{programEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, thereby making the classical planning task $P_{\Lambda}$ easier to be solved.
-
-%The compilation is flexible to the particular scenario where the number of {\em missing states} or {\em missing actions} in the input plan trace $\tau$ is bound.
-In both cases the output planning problem $P_{\Lambda}$ becomes simpler since the plan horizon is bound so the classical planner does not require to determine how many {\em apply} actions are necessary between any two state observations, i.e.
-between the application of two {\em validate} actions.
+In this case, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, thereby making the classical planning task $P_{\Lambda}$ easier to be solved.
 
@@
 
-Finally, we will detail the composition of a solution plan $\pi_\Lambda$ to a planning task $P_\Lambda$ and the mechanism to extract the output set of action models $\mathcal{M}'$ from $\pi_\Lambda$.
+Finally, we will detail the composition of a solution plan $\pi_\Lambda$ to a planning task $P_\Lambda$ and the mechanism to extract the action models of $\mathcal{M}'$ from $\pi_\Lambda$.
 The plan of Figure~\ref{fig:plan-lplan} shows a solution to the task $P_{\Lambda}$ that encodes a learning task $\Lambda=\tup{\mathcal{M},\tau}$ for obtaining the action models of the {\em blocksworld} domain, where the models for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already specified in $\mathcal{M}$.
-Therefore, this plan programs and validates the action model for {\tt\small stack}, using the input plan trace of Figure~\ref{fig:example-plans}.
-Plan steps $00-01$ program the preconditions of the {\tt\small stack} model, steps $02-06$ program the action model effects and steps $07-11$ is the plan postfix that validates the programmed model following the plan trace of Figure~\ref{fig:example-plans}.
+Therefore, the plan shows the insert actions and validate action for the action model {\tt\small stack} using the input plan trace of Figure~\ref{fig:example-plans}.
+Plan steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects, and steps $07-11$ form the plan postfix that applies the action models (only the {\tt\small stack} model is learned) and validates the result in the plan trace of Figure~\ref{fig:example-plans}.
 
@@
 	{\footnotesize\tt
-		{\bf 00} : (program\_pre\_stack\_holding\_v1) \\
-		01 : (program\_pre\_stack\_clear\_v2)\\
+		{\bf 00} : (insert\_pre\_stack\_holding\_v1) \\
+		01 : (insert\_pre\_stack\_clear\_v2)\\
 		{\bf 02} : (program\_eff\_stack\_clear\_v1)\\
-		03 : (program\_eff\_stack\_clear\_v2)\\
-		04 : (program\_eff\_stack\_handempty)\\
-		05 : (program\_eff\_stack\_holding\_v1)\\
-		06 : (program\_eff\_stack\_on\_v1\_v2)\\
+		03 : (insert\_eff\_stack\_clear\_v2)\\
+		04 : (insert\_eff\_stack\_handempty)\\
+		05 : (insert\_eff\_stack\_holding\_v1)\\
+		06 : (insert\_eff\_stack\_on\_v1\_v2)\\
 		{\bf 07} : (apply\_unstack blockB blockA i1 i2)\\
@@
 
-The logical inference process we base our approach on has trouble learning preconditions that are not, at the same time, negative effects, since no change can be observed between the pre-state and post-state of an action.
+The logical inference process our approach is based on has trouble learning preconditions that are not, simultaneously, negative effects as well, since in this case no change can be observed between the pre-state and post-state of an action.
 This is specially the case for static predicates that never change and, hence, can only be preconditions.
@@
 In our case, since the sequence of actions of the input trace might not be fully observable, we produce the traces by applying the actions found in the validation part of the solution plan.
-For instance, in the example of the figure \ref{fig:plan-lplan}, the sequence of actions used to produce the full trace would be \text{(unstack blockB blockA)}, \text{(put-down blockB)}, \text{(pick-up blockA)}, and \text{(stack blockA blockB)}.
+For instance, in the example of the figure \ref{fig:plan-lplan}, the sequence of actions used to produce the full trace would be {\tt{\small(unstack blockB blockA)}}, {\tt{\small(put-down blockB)}}, {\tt{\small(pick-up blockA)}}, and {\tt{\small(stack blockA blockB)}}.
 This post-process allows \FAMA to learn more complete lists of preconditions and deal with the always problematic static predicates.
@@
 \subsection{Properties of the compilation}
+\label{properties}
 

### 2018-12-24 doc/aij18/major-revision/experiments.tex
Updated Dec 24 10.20

@@
   \item \textcolor[rgb]{1.00,0.00,0.00}{Other planners such as {\sc FastDownward} were also tested but provided worse experimental results}
-  \item \textcolor[rgb]{1.00,0.00,0.00}{A SAT-based planner like {\sc Madagascar} is particularly suitable for tasks where the observed plan trace $\tau$ determines the horizon of the solution plan (case 1 presented in section \ref{task_definition})}.
-In this case, the prefix of the solution plan is solvable in two steps because the actions for programming preconditions can be applied in parallel in a single step and the same for the actions programming the effects.
-  \item \textcolor[rgb]{1.00,0.00,0.00}{When the horizon of the plan is unknown (case 2 presented in section \ref{task_definition}), the ability of {\sc Madagascar} to deal with instances populated with dead-ends is very helpful~\cite{lopez2015deterministic}.}
+  \item \textcolor[rgb]{1.00,0.00,0.00}{A SAT-based planner like {\sc Madagascar} is particularly suitable for tasks where the observed plan trace $\tau$ determines the length of the solution plan (case 1 presented in section \ref{task_definition})}.
+In this case, the prefix of the solution plan is solvable in two steps (planning horizon length 2) because the actions for inserting preconditions can be applied in parallel in a single step and the same for the actions inserting the effects.
+  \item \textcolor[rgb]{1.00,0.00,0.00}{When the length of the plan is unknown (case 2 presented in section \ref{task_definition}), the ability of {\sc Madagascar} to deal with instances populated with dead-ends is very helpful~\cite{lopez2015deterministic}.}
   \end{enumerate}

### 2018-12-24 doc/aij18/major-revision/introduction.tex
Updated Dec 24 10.20

@@
 \textcolor[rgb]{1.00,0.00,0.00}{Current learning approaches assume that the observation of the agent's plan execution (plan trace) encompasses the fully observed sequence of the executed actions; i.e, they assume all the actions performed by the agent are observable.
-This heavily restricts the applicability of the learning approaches to contexts where the behaviour of the agent is fully observable and a human annotator correctly labels the executed actions.
+This heavily restricts the applicability of the learning approaches to contexts where the behaviour of the agent is fully observable, which also commonly entails a human annotator that correctly labels the executed actions.
 On the other hand, learning approaches accept a varying degree of observability in the states traversed in the plan trace, ranging from fully observable to fully unobservable states (see section \ref{related_work} for details).} In contrast, \FAMA allows for an incomplete or empty sequence of observable actions, and the minimum observability case acceptable by \FAMA is when the algorithm is only fed with the initial and final state of a plan trace.
@@
 As mentioned above, a plan trace may comprise none of the actions executed by the agent but must include, at least, the initial state $s_0$ and final state $s_f$ of the execution.
-The compilation scheme lies in defining a planning task from the set of plan traces using a set of \emph{programming actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces.
-Hence, a solution to this planning task is a plan that determines the preconditions and effects of a \strips\ planning model $M$.
+The compilation scheme lies in defining a planning task from the set of plan traces using a set of \emph{building actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces.
+Hence, a solution to this planning task is a plan that determines the preconditions and effects of the actions of a \strips\ planning model $M$.
 The rationale of this process draws its strength in that when the problem $\tup{s_0,s_f}$ is solved with the learned model $M$, a plan $\pi$ is generated such that $\pi$ will contain the observed actions of the plan trace, if any, and the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.
-In other words, the plan $\pi$ generated when solving the problem $\tup{s_0,s_f}$ with $M$ is \emph{compliant} with the plan trace.
-}
+In other words, the plan $\pi$ generated when solving the problem $\tup{s_0,s_f}$ with $M$ is \emph{compliant} with the plan trace.}
 

### 2018-12-24 doc/aij18/major-revision/learning_action_models.tex
Updated Dec 24 10.20

@@
 \textcolor[rgb]{1.00,0.00,0.00}{Our proposal to address a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to transform $\Lambda$ into a planning task $P_{\Lambda}$.
-The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that determine the composition of the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ complies with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$.
-Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition, a positive effect or a negative effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that specify the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\tau$, including that the final reached state comprises $s_n$.
+The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ complies with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$.
+Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition, a positive effect or a negative effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\tau$, including that the final reached state comprises $s_n$.
 Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the data generated through the action application must be validated with the data of $\tau$.}
@@
 
-Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, which is bound to the maximum arity of an action in a given planning frame.
+Let $\Omega_v=\{v_i\}_{i=1}^{\operatorname*{max}_{a\in A} ar(a)}$ be a new set of objects ($\Omega\cap\Omega_v=\emptyset$), denoted as {\em variable names}, which is bounded to the maximum arity of an action in a given planning frame.
 For instance, in a three-block {\em blocksworld} $\Omega=\{block_1, block_2, block_3\}$ while $\Omega_v=\{v_1,v_2\}$ because the actions with the maximum arity have arity two; i.e., any instantiation of the {\small\tt stack} or the {\small\tt unstack} models.
@@
 In the {\em blocksworld} domain, this set contains eleven elements, $\Psi_v$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-In more detail, for a given action model $\xi$, we define $\Psi_{\xi}\subseteq \Psi_v$ as the subset of elements of $\Psi_v$ that can appear in $\xi$.
+For a given action model $\xi$, we define $\Psi_{\xi}\subseteq \Psi_v$ as the subset of elements of $\Psi_v$ that can appear in $\xi$.
 For instance, $\Psi_{\tt stack}=\Psi_v$ whereas $\Psi_{\tt pickup}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$),on($v_1,v_1$)\}} excludes the elements from $\Psi_v$ that involve $v_2$ because {\small\tt pickup} actions have arity one.
@@
 \begin{enumerate}
-\item {\bf To specify the action models of $\mathcal{M}'$}.
-$\pi_\Lambda$ comprises a plan {\em prefix} whose actions set the predicates $p\in \Psi_{\xi}$ that belong to $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\mathcal{M}$.
+\item {\bf To build the action models of $\mathcal{M}'$}.
+$\pi_\Lambda$ comprises a first block of actions (plan {\em prefix}) that set the predicates $p\in \Psi_{\xi}$ of $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\mathcal{M}$.
 \item {\bf To validate the action models of $\mathcal{M}'$}.
-$\pi_\Lambda$ also comprises a plan {\em postfix} whose actions target the validation of the observed plan trace $\tau$ with the specified action models $\mathcal{M}'$.
+$\pi_\Lambda$ also comprises a second block of actions (plan {\em postfix}) which is aimed at validating of the observed plan trace $\tau$ with the built action models $\mathcal{M}'$.
 \end{enumerate}
@@
 
-\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ specified by the insert actions bound to objects $\omega\subseteq\Omega^{ar(\xi)}$.
-Since action headers are known, the variables $pars(\xi)$ are bound to the objects in $\omega$ that appear in the same position.
+\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{ar(\xi)}$.
+Since action headers are known, the variables $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position.
 
@@
 \end{small}
+
+\textcolor[rgb]{1.00,0.00,0.00}{      There will be a validate action in $\pi_\Lambda$ per observed state in $\tau$.
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.}
+
 \end{enumerate}
@@
 		01 : (insert\_pre\_stack\_clear\_v2)\\
-		{\bf 02} : (program\_eff\_stack\_clear\_v1)\\
+		{\bf 02} : (insert\_eff\_stack\_clear\_v1)\\
 		03 : (insert\_eff\_stack\_clear\_v2)\\
@@
 The validation capacity of our compilation is beyond the functionality of VAL (the plan validation tool~\cite{howey2004val}) because our $P_{\Lambda}$ compilation is able to address {\em model validation} of a partial (or even an empty) action model with a partially observed plan trace.
-On the other hand, VAL requires (1) a full plan and (2), a full action model for plan validation.
+VAL, however, requires a full plan and a full action model for plan validation.
 

### 2018-12-24 doc/aij18/major-revision/letter.tex
Updated Dec 24 10.20

--- 2018-11-23 doc/aij18/major-revision/letter.tex
@@
 \newcommand{\strips}{\textsc{Strips}} % Strips
+\newcommand{\FAMA}{{\small {\sffamily FAMA}}\xspace}
 
+\usepackage{amsthm}
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{times}
+\usepackage{helvet}
+\usepackage{courier}
+\usepackage{pstricks}
+\usepackage{pst-node}
+\usepackage{multirow}
+\usepackage{listings}
+\usepackage{xspace}
+
+\title{\textbf{Response Letter}}
 \begin{document}
+\maketitle
 
+Firstly, we would like to thank the three reviewers for their insightful comments and recommendations.
+We truly believe the observations pointed by the three reviewers have been very helpful to make the paper more readable.
+
+This response letter is structured as follows.
+The next section summarizes the most relevant modifications addressed in the paper.
+Subsequent sections are devoted to thoroughly reply all the reviewers' concerns.
+
+\subsection*{Main modifications addressed in the paper}
+
+We identified three main points to be addressed as per the reviewers' comments and observations.
+In this section, we provide details of these three main modifications, and we will point at them when replying to each of the reviewers' comments in the remainder sections.
+By main modifications we mean significantly large portions of new text in the paper that incorporate new explanations.
+
+\begin{enumerate}
+\item \textbf{Introduction}.
+We thoroughly revised the Introduction to make it more self-explanatory.
+First, we carefully revised all of the comments pointed by reviewer \#2.
+Additionally, as indicated by reviewer \#1, we provide an outline of the overall picture of our \FAMA approach.
+Our modification is intended to adequately position our contribution within the concerned research field, and to provide a global picture of \FAMA that explains our pursued objective.
+For this purpose, we have introduced a new paragraph that descriptively presents the functioning of \FAMA.
+Additionally, at the end of the section we highlight the main distinctive characteristics of \FAMA which describe the overall concept of our learning approach.
+
+\item \textbf{Problem definition and complexity}.
+As indicated by reviewer \#2, and also supported by reviewers \#1 and \#3, the problem definition was not clearly introduced.
+We believe this is the reason that our claim about using planning for solving the learning task was not fully comprehensible.
+We have reshaped the old section 3 (Motivation), which now is entitled "Learning task: on the use of planning for solving the task".
+First, we introduce the definition of the learning task (which was in section 4 in the previous version of the paper), and then we analyze the intricacies of the learning task to the different variants of observed plan traces.
+We organised the explanation around two well differentiated cases: when the observation identifies the length of the plan being executed and when it does not.
+Now, the discussion revolves around the particularities of solving the learning task in these two cases and we make some side comments on the complexity of the task.
+Hopefully, the significant remodeling of this section will clarify the need of using planning techniques for solving the learning task.
+
+\item \textbf{Compilation scheme}.
+Reviewers \#1 and \#3 show several concerns about the understanding of section 4.4, the compilation approach.
+We acknowledge their concerns and we changed section 4 significantly.
+First, a new introductory paragraph has been added at the beginning of the section so as to provide a rough overview of the compilation approach, making explicit references to the actions that will appear in a plan that solves the planning task that results from the compilation.
+Second, the explanation of the set of actions that make up the planning task has substantially improved, and we provide concrete examples throughout that will help grasp the target of the three types of actions.
+We have also changed other minor things (e.g., the first set of actions is now renamed as \textbf{insert actions} instead of program actions) with the purpose to ease a smooth reading.
+\end{enumerate}
+
+\textcolor{blue}{In the following, our responses to the reviewers are shown up in blue color.}
+
+\vspace{0.3cm}
+
+\textcolor{red}{Additionally, all the changes are highlighted in red in the document file.}
+
+%##################################################
 \subsection*{Reviewer \#1}
+%##################################################
 
@@
 Going top down would make reading much easier.
+\end{mdframed}
+
+\textcolor{blue}{We agree with the reviewer that the paper lacks a description of the overall approach.
+We refer the reviewer to the main modification \#1 exposed above.
+Following the notes of the reviewer's summary, we have also included a brief summary at the end of the section Introduction that serves to wrap up the main features of \FAMA.
+Additionally, and as requested by the other reviewers too, we introduced the main modification \#2.
+The new section 3 defines the learning task and provides insight of the nature of the problem to solve according to the type of the observed plan trace.
+We believe that this explanation also helps to grasp the overall picture of \FAMA.}
+
+\begin{mdframed} [hidealllines=true,backgroundcolor=gray!20]
 The second reason for revision is the comparison to ARMS.
@@
 \end{mdframed}
+
+\textcolor{blue}{As per the reviewer's recommendation, we requested the code of LOUGA to the first author of the publication entitled \emph{LOUGA: learning planning operators using genetic algorithms}, which was published in the Pacific Rim Knowledge Acquisition Workshop.
+The author of LOUGA promptly provided us with the code, and so we thank him for his gentle response.
+The issue we faced though with LOUGA is that this system is built as a graphical interface application, which makes it impossible to automate the model learning process and evaluation process via a script for the 154 tested cases.
+Nevertheless, we did manually test LOUGA on a few domains, namely the \emph{blocksworld}, \emph{floor-tile}, \emph{n-puzzle} and \emph{transport} domains.
+We evaluated LOUGA and \FAMA in the experiment with FO action sequences and PO state trajectories with 10\% of observability using 5 traces.
+The global precision and recall values for each domain are shown below as LOUGA/\FAMA: }
+
+\begin{table}[h!]
+\begin{center}
+\begin{tabular}{lll}
+        & Precision  &  Recall    \\
+ blocks &  0.78/1.0 &  0.78/0.93    \\
+ floor-tile & 0.83/0.75  &  1.0/1.0 \\
+ n-puzzle & 0.88/1.0 & 1.0/1.0  \\
+ transport &  0.70/0.79 &  0.8/0.95  \\
+\textcolor[rgb]{0.00,0.00,1.00}{AVERAGE}  & \textcolor{blue}{0.82/0.89} & \textcolor{blue}{0.90/0.97}
+\end{tabular}
+\end{center}
+\end{table}
 
@@
 That is very confusing.
+ \end{mdframed}
 
+\textcolor{blue}{The reviewer is right in that the definition of effects was confusing.
+It all stems from the adoption of the open world assumption, which leads to states that contain positive and negative literals, but we did not adapt the definition of effects appropriately.
+In section 2.1, we have now included a brief explanation of the open world assumption and the representation of states, and we have adapted the definition of effects accordingly.}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
  Also, there are additional assumptions about the model, namely positive effects cannot be among preconditions while negative effects must be among preconditions.
@@
 \end{mdframed}
+
+\textcolor{blue}{The additional assumptions are the syntactic constrains imposed by \strips\ planning models.
+We agree with the reviewer that other domain models do not satisfy these assumptions.
+We intend to extend the compilation scheme to planning models other than \strips\ as future work, but the focus of this paper is exclusively on \strips\ models.
+We have included a sentence before the description of the assumptions to clearly state that they derive from the syntactic constraints of \strips.}
 
@@
 
+\textcolor{blue}{We have reformulated the explanation about the observability in plan traces of section 2.1 to make it more comprehensible and added a few more details that help distinguish the different cases of observability.
+We have also highlighted the two special cases, PO* and NO, that derive from PO.
+With respect to the relation between actions and state partial observability, we refer the reviewer to the main modification \#2.
+The new section 3 presents the combinations of observed state trajectories and
+observed action sequences and the impact in the learning task solving.
+Particularly, as it is written in the text at the end of page 10, the case FO actions with FO states is straightforward because in this case every state is available and so preconditions and effects are derivable by lifting the corresponding literals.
+The case FO actions with PO* states is a common case handled by some learning approaches as shown in Table 2 of section 2.2.
+In general, all learning approaches but \FAMA assume FO action sequences, which may come along with PO*, PO or NO state trajectories.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 
+\textcolor[rgb]{0.00,0.07,1.00}{Done. LOUGA is included in Table 2 and Table 3 and an explanatory paragraph about the main characteristics of LOUGA has been incorporated in section 2.2.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 \end{mdframed}
+
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ....}}
 
@@
 \end{mdframed}
+
+\textcolor{blue}{Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects.
+This is the reason why conditional effects are introduced in the compilation scheme since they only intervene in the specification of $P_{\Lambda}$ and do not condition the characteristics of the models that \FAMA is able to learn.
+Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}
+
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...  SE SIGUEN CUMPLIENDO LAS SYNTACTIC STRIPS CONSTRAINTS CON CONDITIONAL EFFECTS????? }}
 
@@
 
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  Page 11, Section 4.4.
-This section is hard to understand, at least at the beginning.
-One needs to read it completely before understanding the details.
-I suggest to describe the concept at high level first and going to details then.
-\end{mdframed}
+\textcolor{blue}{We refer the reviewer to the main modification \#3.
+Section 4 has been polished significantly.
+An introductory paragraph provides a high level description of the approach and examples have been incorporated throughout the text.}
 
@@
 
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ... TENEMOS QUE VER ESTO ... }}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 
+\textcolor{blue}{The implications can actually be encoded as a disjunction, as it is the case of the {\tt{\small apply\_stack}} action of Figure 4.
+The caption of the figure indicates that "implications are coded as disjunctions".
+The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 \end{mdframed}
+
+\textcolor{blue}{The positions of the validate actions among the apply actions is determined by the planner, which will verify that the state that results after the execution of an apply action comprises the observed state of the plan trace.
+We have added the following sentence in the document to clarify this point: "There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$.
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$."}
 
@@
 
+\textcolor{blue}{All plotted values are averages over the 14 (now 15) tested domains.
+The reviewer is right, however, to point out that this is not said in the experiment of section 6.2, so we have added some lines making it explicit.}
+
+%##################################################
 \subsection*{Reviewer \#2}
+%##################################################
 
@@
 
+\textcolor{blue}{Following the reviewer's advice, we have completely remodeled section 3 and we have introduced major improvements in section 4.
+We refer the reviewer to our main modifications \#2 and \#3.
+We removed the complexity table of page 8, focusing now the discussion around the two main cases that determine the difficulty of the learning task and giving an indication of the task complexity according to the characteristics that define these two cases.
+We also changed the reference and added instead the original work that proves that boolean SAT problems are NP-complete.}
+
+\textcolor{blue}{The take-home lesson of section 3 is that when the observed plan trace does not provide information of the solution plan length (case 2), planning is needed to find the solution and so the learning task is PSPACE-complete.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+Repetitions.
+
   There are several cases where authors repeat themselves several times. Here are a few examples:
 
--          (page 2) "While current learning systems …. none of them allow partial observability in the sequence of executed actions "
-
-Two paragraphs later:
-
-" … unlike most relevant action-model learning algorithms, FAMA does not require the traces to contain any observed action …"
-
-(Page 8) "FAMA represents one step ahead … without assuming observed actions."
+- (page 2) "While current learning systems ... none of them allow partial observability in the sequence of executed actions "
+Two paragraphs later: " ... unlike most relevant action-model learning algorithms, FAMA does not require the traces to contain any observed action ..."
+(Page 8) "FAMA represents one step ahead ... without assuming observed actions."
 
@@
 
+\textcolor{blue}{We have removed repetitions as much as possible.
+In the first example pointed by the reviewer, the explanation about partial observability of \FAMA and the other learning systems is concentrated in a single paragraph and the repetition is eliminated.
+Nevertheless, we think that at some points it is necessary to recall a particular concept in order to enhance the exposition of a further concept.}
+
+\textcolor{blue}{With respect to the repetition of the definition of a \strips\ action model, conditional effects are introduced in section 4 because these are exclusively used for modeling the task that results from the compilation, $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$.
+Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}
+
+\textcolor{blue}{With respect to the third observation, we have introduced several changes in section 5.
+First, we provide the standard definitions of precision and recall at the beginning of the section.
+Then, we adapt these two concepts to planning problems in Def.
+4 and 5, respectively.
+Subsequently, in section 5.2, we present our novel semantic-based metrics that build upon definitions 4 and 5.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 If we see that the number of timeouts for this setting is larger than the simpler setting, that would support the authors conclusion.
+\end{mdframed}
 
-\end{mdframed}
+\textcolor{blue}{As indicated by the reviewer, timeouts are only observed in the setting with NO action sequences and NO state trajectories.
+We have added a few sentences providing more details on this issue.}
+
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{SOLO HE VISTO UNA FRASE DE TIMEOUTS SOBRE EL GRID ... Y EL RESTO DE DOMINIOS? NO SE PUEDE DECIR NADA???? }}
 
@@
 
+\textcolor{blue}{The reviewer is totally right in that planning is needed when the length of the solution plan is unknown.
+Precisely, this is the point that motivates our approach but that we did not successfully convey in the previous version of the paper.
+When the observed plan trace comprises a partially observed action sequence and a partially observed state trajectory, we ignore the length of the solution plan and so we need to resort to planning.
+The new section 3 explains this in detail (main modification \#2).
+When the length of the plan is known, SAT-solvers can be used; when it is not known, planning is required.
+In our case, we use a planner to solve all learning tasks independently of their nature.}
+
+\textcolor{blue}{On the other hand, we happen to use a SAT-based planner (Madagascar) for solving the learning tasks, which may lead to wrongly interpret that the length of the solution plan is always known.
+The reasons for using Madagascar have been highlighted in red at the beginning of section 6.
+In a nutshell, we use Madagascar because it returned the best results compared to other planners and because of its ability to handle problems with dead-ends.
+Additionally, Madagascar is obviously very suitable to solve learning tasks when the length of the plan is known.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-    Page 1.
-"… an abstract version of the capability model that reflects … " – I am not sure that "abstract" is the right term here, since the learned model is not an abstraction of the real-world but an approximation of it.
-\end{mdframed}
+\textcolor{blue}{We acknowledge the reviewer's concern and refer the reviewer to the main modification \#1 exposed at the beginning of this response letter.}
+\vspace{0.1cm}
 
@@
     Page 1.
-"Two different types of data are generally identified in the generation of explanations … (2) data that do not accurately represent the decision-making process because the observed external behavior of the agent responds to a black-box model." – First, I think you meant "corresponds" and not "responds".
+" an abstract version of the capability model that reflects ...
+" -- I am not sure that "abstract" is the right term here, since the learned model is not an abstraction of the real-world but an approximation of it.
+\end{mdframed}
+
+\textcolor{blue}{We have replaced the word 'abstract' by 'approximation' and rephrased the sentence a little bit.}
+\vspace{0.1cm}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+    Page 1.
+"Two different types of data are generally identified in the generation of explanations ‚Ä¶ (2) data that do not accurately represent the decision-making process because the observed external behavior of the agent responds to a black-box model." -- First, I think you meant "corresponds" and not "responds".
 Second, if the agent is treated as a black box, then this is not really a way to generate explanation.
@@
 
+\textcolor{blue}{We think this paragraph was rather confusing and that we did not clearly come through our intention to set an analogy between the data sources used in explainable AI and the data required by the learning approaches.
+For this reason, we have removed the paragraph and included instead a list of planning-related tasks that require a planning model for being solved.
+We think this rather complete but non-exhaustive list is a pretty good indication of the need for learning planning models.}
+
+\vspace{0.1cm}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-      Page 2. "… the synthesis of different kinds of generative models with classical planning …" – I do not understand this sentence. What do you mean by generative models in this context?
+      Page 2. " ... the synthesis of different kinds of generative models with classical planning ..." -- I do not understand this sentence. What do you mean by generative models in this context?
+
+      Page 2. "FAMA is a solving scheme ..." -- FAMA is a learning scheme, not a solving scheme. I agree that it is a learning scheme that learns by solving a planning problem.
+
+      Page 2. " ... by the input observed actions, ... " What is "input observed actions"? please rephrase.
 \end{mdframed}
 
+\textcolor{blue}{We have rephrased all the paragraphs and sentences pointed by the reviewer.
+With respect to the first comment, we attempt to make it clear that classical planning techniques can be used to generate a large variety of planning models, not only \strips\ models, and that these techniques underpin some relevant concepts that are also reusable in learning models when the behaviour of the agent is not available.}
+
+\textcolor{blue}{With respect to the other two comments, besides using different expressions to refer to the two concepts, we have included some additional side explanations that we hope help understand the meaning of the sentences.}
+\vspace{0.1cm}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-      Page 2. "FAMA is a solving scheme … " – FAMA is a learning scheme, not a solving scheme. I agree that it is a learning scheme that learns by solving a planning problem.
+      Page 3. " ... the open world assumption" -- Either explain what this means or provide a reference.
 \end{mdframed}
 
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-      Page 2. "… by the input observed actions, …" – What is "input observed actions"? please rephrase.
-\end{mdframed}
-
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-      Page 3. "… the open world assumption" – Either explain what this means or provide a reference.
-\end{mdframed}
+\textcolor{blue}{We have included an explanation of the open world assumption and how this affects the representation of states in Section 2.1.}
 
@@
       Page 3.
-"…the minimal action sequence to transit from state $s_i$ to $s_{i+1}$ is composed of a single action…" – This can be stated in a simpler way: "there exists a single action that transitions $s_i$ to $s_{i+1}$"
+" ...
+the minimal action sequence to transit from state $s_i$ to $s_{i+1}$ is composed of a single action ..." - This can be stated in a simpler way: "there exists a single action that transitions $s_i$ to $s_{i+1}$"
 \end{mdframed}
 
+\textcolor[rgb]{0.00,0.07,1.00}{Fixed.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-      Page 4. "A plan trace …. For a planning frame …. holds that … for every action in  … and that …" – this sentence is not clear. Please rephrase.
+      Page 4. "A plan trace .... For a planning frame ... holds that ... for every action in ... and that ..." - this sentence is not clear. Please rephrase.
 \end{mdframed}
+
+\textcolor[rgb]{0.00,0.07,1.00}{The sentence was, indeed, unclear.
+Anyway, we removed it in the revised version of the paper since it was an expendable phrase that did not fit very well in this section either.}
 
@@
 
+\textcolor[rgb]{0.00,0.07,1.00}{We have moved the 4th paragraph of the Related Work, where we commented on the predicates and action headers of the approaches, to the end of the section since it makes more sense to comment on this issue when comparing with LOCM.
+Also, we clarified the meaning of action header (action name and parameters) because the formal definition of an action header is actually introduced in section 4.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-      Page 4. "… of the evaluation method to validate …" – missing "used" before "to"
+      Page 4. " ... of the evaluation method to validate ..."  - missing "used" before "to"
 \end{mdframed}
+
+\textcolor[rgb]{0.00,0.07,1.00}{Fixed.}
 
@@
   Page 4."Exceptionally, a NO state sequence in LOCM is a fully-empty trajectory, with neither initial or final state.
-" – I don't understand this sentence.
+" - I don't understand this sentence.
 Isn't a NO state sequence a fully empty trajectory for all algorithms?
@@
 
+\textcolor[rgb]{0.00,0.07,1.00}{The labels FO, PO*, PO and NO concern exclusively the intermediate states of the state trajectories as indicated in Table 1.
+We have improved the explanation of the four types of state trajectories in section 2.1.
+As for LOCM, this approach does not require any state, not even the initial or final state.
+We changed the label of LOCM in the first column of Table 2 to "---" to indicate that no state information at all is needed.
+We have also rephrased the sentence about LOCM.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-   Page 4."Most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces" – What about FAMA? I think it also requires this.
+   Page 4."Most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces" - What about FAMA? I think it also requires this.
 \end{mdframed}
+
+\textcolor[rgb]{0.00,0.07,1.00}{The reviewer is right in that \FAMA also assumes that predicates and action headers are provided or otherwise are extractable from the plan traces.
+We believe it is more appropriate to move this comment after LOCM is introduced since LOCM is an exception that learns the predicates using Finite-State Machines (FSM) because no observed states are given in the plan trace.}
 
@@
 2.
-When talking about LAMP, you write that it allows trajectories up to a "minimum percentage of 1/5 …" – again, is this requirement mandatory in their algorithm, or is it just what they happened to get in their experiments?
+When talking about LAMP, you write that it allows trajectories up to a "minimum percentage of 1/5 ‚Ä¶" ‚Äì again, is this requirement mandatory in their algorithm, or is it just what they happened to get in their experiments?
 Also, if you write 1/5 then you should write "ratio" and not "percentage".
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-    Pages 5-7. "… the state of a sort" – what is a "sort"?
+    Pages 5-7. " ... the state of a sort"  -- " what is a "sort"?
 \end{mdframed}
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-    Pages 5-7. "As many of the approaches in section 2 assume, there may be an …" – this text is redundant. It just repeats again what was said twice already.
+    Pages 5-7. "As many of the approaches in section 2 assume, there may be an ..."  - this text is redundant. It just repeats again what was said twice already.
 \end{mdframed}
 
+\textcolor[rgb]{0.00,0.07,1.00}{We refer the reviewer to the main modification \#2. Section 3 has substantially changed and this sentence does no longer appear in the text.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  Page 8."… an unbound number of missing actions …" – should be "unbounded'
+  Page 8." ... an unbound number of missing actions ..."  - should be "unbounded'
 \end{mdframed}
+
+\textcolor[rgb]{0.00,0.07,1.00}{Corrected.}
 
@@
 
+\textcolor[rgb]{0.00,0.07,1.00}{It is true that a machinery for capturing and labeling the executed actions can be deployed.
+In either case, whether a human or an automated mechanism, the need of a fully observable behaviour of the agent imposes a limitation in learning systems.
+We have soften the sentence including the word "\emph{commonly}" in our exposition.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  Page 8."When the plan traces is fully observed, learning STRIPS action models is straightforward" – The reference if for work done in 2018, while learning an action model in this setting has been discussed in IJCAI 2017 by Stern and Juba (Safe, model-free learning of action models").
+  Page 8."When the plan traces is fully observed, learning STRIPS action models is straightforward" - The reference if for work done in 2018, while learning an action model in this setting has been discussed in IJCAI 2017 by Stern and Juba (Safe, model-free learning of action models").
 \end{mdframed}
@@
 
+\textcolor[rgb]{0.00,0.07,1.00}{We have removed the paragraph regarding the evaluation of the learned models from section 3 and moved to section 5.
+We also clarify the use of the word "semantic" as requested by another reviewer as well.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
--          "On the other hand, a striking figure …" – I could not understand this sentence.
+- "On the other hand, a striking figure ..."  - I could not understand this sentence.
+\end{mdframed}
 
--          "Let $\Omega_v = {v_i}_{i=1}^{max_{a\in A} ar(a)}$ be a new set of objects …" – I think "ar(a)" should be "|ar(a)|"
+\textcolor[rgb]{0.00,0.07,1.00}{In the remodeling of section 3, the paragraph about the size of the training dataset has been removed too. The ideas presented here have been moved to section 6.}
 
--          "… which is bound to …" – I think should be " which is bounded by"
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ... NO HE VISTO ESTE PARRAFO EN LA SECCION 6, NO PUSISTE ESTA INFORMACION EN 6??}}
 
--          "In more details, for a given …" – delete the "In more details" as you are defining something new, not providing additional details.
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+- "Let $\Omega_v = {v_i}_{i=1}^{max_{a\in A} ar(a)}$ be a new set of objects ‚Ä¶" ‚Äì I think "ar(a)" should be "|ar(a)|"
+\end{mdframed}
 
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...}}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+- "... which is bound to ..."  - I think should be " which is bounded by"
+\end{mdframed}
+
+\textcolor[rgb]{0.00,0.07,1.00}{Corrected.}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+- "In more details, for a given ..." - delete the "In more details" as you are defining something new, not providing additional details.
+\end{mdframed}
+
+\textcolor[rgb]{0.00,0.07,1.00}{Corrected.}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
 - Please elaborate on the computation of the size of the space of possible STRIPS models.
+\end{mdframed}
 
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...}}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
 - The definition of conditional effects given in 4.3 should be moved to the background section, where you define a STRIPS model.
+\end{mdframed}
 
+\textcolor{blue}{Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects.
+This is the reason why conditional effects are introduced in the compilation scheme since they only intervene in the specification of $P_{\Lambda}$ and do not condition the characteristics of the models that \FAMA is able to learn.
+Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
 - The explanation about the plan prefix and postfix is confusing and not helpful.
@@
 Maybe "suffix"?
+\end{mdframed}
 
+\textcolor[rgb]{0.00,0.07,1.00}{We have introduced a new paragraph at the beginning of section 4 that gives an overview of the compilation approach and explains the two blocks of actions of the solution plan to the learning task.
+We have reused the concept of "block of actions" to refer to the prefix and postfix of $\pi_{\Lambda}$ in section 4.2 "Compilation".}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
 - The formula for the precondition of apply uses the notation $\Rightarrow$ which was not defined.
@@
 
+\textcolor[rgb]{0.00,0.07,1.00}{$\Rightarrow$ is the general logic symbol read as implication.
+The implications can actually be encoded as a disjunction, as it is the case of the {\tt{\small apply\_stack}} action of Figure 4.
+The caption of the figure indicates that "implications are coded as disjunctions".
+The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$ }
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
 Pages 15-19.
+- I'd appreciate more details on the mentioned post-processing, e.g., a pseudo code.
+\end{mdframed}
 
--          I'd appreciate more details on the mentioned post-processing, e.g., a pseudo code.
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...  SE PUEDE DECIR ALGO MAS O IGNORAMOS EL COMENTARIO?}}
 
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
 - At the bottom of the page, you write that VAL cannot address the model validation of a partial model and on the other hand it requires a full plan and full action model.
@@
 it is the same "hand"
+\end{mdframed}
+
+\textcolor[rgb]{0.00,0.07,1.00}{Corrected.}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+Pages 15-19.
 
@@
 It is like saying "we justify using a test-based evaluation by the fact that test-based evaluation is sometimes preferred".
+\end{mdframed}
 
+\textcolor{blue}{We have provided better justification for the use of semantic evaluation metrics in section 5.}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+Pages 15-19.
 -  What does it mean that a learned action model is sound and complete?
+\end{mdframed}
 
--          "… to flaws that appear more than once in the plan traces …" – I recommend to clarify: "flaws in the action model that manifest more than once in the plan traces".
+\textcolor{blue}{The precision and recall metrics give a notion of the soundness and completeness of a model.
+By a sound model we understand one where all preconditions and effects are correct, either wrt to the GTM or a test set, so there is no need to remove anything.
+A complete model is one where no precondition or effect is missing.
+We have added a sentence to clarify this.}
 
--          "… generated 10 plan traces (each with 10 actions …" – the parenthesis should be removed.
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...  HE A—ADIDO ESTA FRASE EN EL TEXTO PERO SOLO CON RESPECTO AL GTM. MIRA A VER SI TE PARECE BIEN.}}
 
--          "… the horizon of the solution plan is also known.
-" – how do you know the horizon?
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+Pages 15-19.
+
+- "... to flaws that appear more than once in the plan traces ..." - I recommend to clarify: "flaws in the action model that manifest more than once in the plan traces".
+
+- "... generated 10 plan traces (each with 10 actions ..." - the parenthesis should be removed.
+\end{mdframed}
+
+\textcolor{blue}{Fixed.}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+Pages 15-19.
+-  "...
+the horizon of the solution plan is also known.
+" - how do you know the horizon?
 You know the length of the given traces, but how do you know how many actions are needed to build the action model?
+\end{mdframed}
 
--          "… a solution plan is solvable in two steps" – which two steps? I thought you put it all in the planner and get both action model and validation in the same step.   
+\textcolor{blue}{Hopefully the new section 3 (see main modification \#2) helps clarify this question.
+In a nutshell, sometimes a plan trace indicates the length of the plan (this is the case when SAT solvers can be used) and others it does not (this is the case when planning is required).}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+-  "... a solution plan is solvable in two steps" - which two steps? I thought you put it all in the planner and get both action model and validation in the same step.
 \end{mdframed}
+
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...  HE A—ADIDO ALGO PARA CLARIFICAR ESTO EN EL TEXTO PERO NO SE SI EST¡ BIEN ... COMPROBAR}}
+
+\textcolor{blue}{Hopefully the new section 3 (see main modification \#2) helps the reviewer clarify this question. }
 
@@
 
--          "… the size of the input knowledge, in the performance …" – should be "on the performance"
+- "... the size of the input knowledge, in the performance ..." - should be "on the performance"
 
--          "… the Precision and Recall" – why capitalized?
+- "... the Precision and Recall" - why capitalized?
 
--          "retain a level of soundness similar to " – what is this "level of soundness"? do you talk about precision?
+- "By lightening the input constraints ..." - maybe "relaxing" instead of "lightening"?
+\end{mdframed}
 
--          "By lightening the input constraints …" – maybe "relaxing" instead of "lightening"?
+\textcolor{blue}{Fixed.}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+-          "retain a level of soundness similar to " ... what is this "level of soundness"? do you talk about precision?
 \end{mdframed}
+
+\textcolor{blue}{Yes, precision gives a notion of soundness since it measures how many of the learned preconditions and effects are correct.
+We have improved the explanation of the metrics in section 5 to make this more clear.}
 
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-First sentence of the Introduction: there is no satisfactory or generally accepted definition of “complete” domain model, given that a domain model is an abstraction of the domain being modelled.
+First sentence of the Introduction: there is no satisfactory or generally accepted definition of "complete"ù domain model, given that a domain model is an abstraction of the domain being modelled.
 If the authors use such a term they should  make clear what their definition is.
-“Adequate” is a better word to use - in the sense of being adequate enough for use with a planner to solve a set of planning  tasks.
+"Adequate"ùis a better word to use - in the sense of being adequate enough for use with a planner to solve a set of planning tasks.
+\end{mdframed}
+
+\textcolor{blue}{We have replaced the word "complete" by "adequate" as suggested by the reviewer.}
+
+\vspace{0.1cm}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  In this part and the rest of the paper I do not fully agree with the author's hype about FAMA with respect to the input information required.
+The minimal amount of information that is needed (if I have got this right) is to specify every predicate that will be needed in the application (this is done in the initial state) the number of steps that are in the training plan, the set of all action headers including their list of parameters, and a goal state which refers to the same predicates as the initial state.
+
+  Comparing this to, say, LOCM, where no state/predicate information at all needs to be specified, I would say that in fact LOCM requires a lot less 'information' a priori.
+What FAMA does not require is the identity of all the actions at each step, nor the instantiation of their headers, but it does require the definition of all the predicates that may appear in any state.
+The truth it seems to me is that FAMA would be good for some applications which fit this set of constraints, which is a different set of constraints than for other domain model learners.
+
+  A way forward might be to put the claims about FAMA into a smaller set of learning systems (certainly ARMS will be in that set) where it clearly needs less input information.
+\end{mdframed}
+
+\textcolor{blue}{We agree with the reviewer's observation, and we believe the misunderstanding stems from us mixing up the words \emph{information} and \emph{observability}.
+In our claim of FAMA using the minimal amount of information, we should actually say that FAMA uses the \emph{minimal observability} possible, being the extreme case to have only the initial state and final state as observations.
+In this respect, LOCM is indeed the learning approach that requires the \emph{minimal information} since no state or predicate is needed.}
+
+\textcolor{blue}{We think that LOCM is a learning approach that notably differs from the rest of systems presented in Table XXX and, in this regard, we have highlighted this distinction in section XXXXX.}
+
+\textcolor[rgb]{1.00,0.00,0.00}{NO ESTA HECHO TODAVIA}
+
+\vspace{0.1cm}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  I totally agree with the authors in their assertion that the evaluation methods for previous model learners is currently poor, and amount to how close the learned model is to a GTM.
+I am not convinced, however, that the use of the statistic methods proposed is a great improvement.
+Precision and Recall are heavily used in data mining where (generally) there are huge data sets and there is not generally a hand crafted model for comparison.
+In planning, however, the idea of separating out a set of traces into training and testing appears odd, especially where it is known that the learning will eventually create a model that explains all the training traces.
+This seems particularly odd if there are only a few traces to be split up (and FAMA appears only able to cope  with a few traces because of complexity problems).
+\end{mdframed}
+
+\textcolor{blue}{The syntactic metrics commonly used by other approaches count the number of differences wrt the GTM and normalize by the total number of possible preconditions and effects.
+This produces an optimistic metric since the set of possible preconditions and effects is usually much larger than the set of learned ones, resulting in low error rates even when the model is largely flawed.
+In contrast, precision and recall normalize by the size of the model, removing this problem.
+Furthermore, precision and recall give a notion of soundness and completeness by splitting the differences into extra and missing preconditions/effects.
+We have added some paragraphs to section 5 to better explain these concepts.}
+
+\textcolor{blue}{As the reviewer points out, the learning will create a model that explains all the training traces.
+There is, however, no guarantee that the training traces are able to capture all the dynamics of the domain.
+This is specially true for scenarios with low observability like the ones explored in this paper.
+Hence, it is always interesting to measure how well the learned model is able to explain new plan traces.}
+
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  - The method of learning: This is to compile the learning problem into a conditional-operater classical planning problem.
+This method appears to be highly innovative to this line of research (stemming from an earlier ICAPS paper by the authors).
+
+  The description of the compilation method did not appear to me to be at the level of the detail to allow easy reading.
+It took me a good deal of time to work out what going on.
+This is a complex process and I recommend much more use of examples throughout - perhaps a running example to show the full run of a compilation.
+
+  It might also improve the paper if an algorithmic description of compilation were included?
+One aspect that was not obvious from the description was how a SET of models are delivered by planning rather than just one; this needs spelling out.
 \end{mdframed}
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  In this part and the rest of the paper I do not fully agree with the author’s hype
+  - The evaluations of the domain model using classification metrics of precision and recall: I like the distinction of the methods labelled "syntactic" (using the GTM) vs "semantic" here; however I think the use of the word - "semantic"ù a bit misleading.
+This implies taking into account the meaning of the features within the application domain which are left out or in erroneously compared to the GTM e.g.
+that a missing "on(x,y)"ù in a learned model is somewhat better/worse from a missing ‚Äùholding(x)‚Äù - because of their meaning, rather than counting each predicate as '1'.
 
-about FAMA with respect to the input information required. The minimal amount of
+  In the end your method reverts to counting the number of different predicates, but in the context of the recall/precision formula with respect to domain model(s) that are consistent with the test sets, not only the GTM.
 
-information that is needed (if I have got this right) is to
+  The definition of the metrics was not too clear. Intuitively, I think I understood 5.1 but it was not well stated -the definition of some terms seemed syntactically wrong or at least sloppy.
 
-specify every predicate that will be needed in the application (this is done in the initial state)
+  E.g.
+in the definition of INS and DEL, should not the universally quantified variables be further qualified by stating that there is a union of the already unioned sets?
+Or again, should there not be an extra "Sigma" to represent the addition of the addition of the number of precons and effects in size(M) definition?
+In proposition 7 you say the closest set "is the GTM" do you mean equal to the singleton set of the GTM?
 
-the number of steps that are in the training plan, the set of all action headers including their
+  Also - the structure of 5.1 - it would be best to clearly state that the syntactic evaluation is going to be introduced, then as a further subsection the semantic evaluation.
+\end{mdframed}
 
-list of parameters, and a goal state which refers to the same predicates as the initial state.
+\textcolor{blue}{The semantic evaluation is used to measure the ability of a learned model to produce a given plan trace.
+In this context, the use of the word "semantic" refers to the physics of the domain, not the actual meaning of the predicates or actions.
+We have added a comment clarifying this point.}
 
- Comparing this to, say, LOCM, where no state/predicate information at all
+\textcolor{blue}{We agree with the reviewers concerns about the definitions of INS, DEL and size(M) and have corrected them.}
 
-needs to be specified, I would say that in fact LOCM requires a lot less ‘information’
+\textcolor{blue}{Following the reviewer recommendation, we have splitted section 5.1 into two sections: one for the syntactic evaluation and another for the semantic evaluation}
 
-a priori. What FAMA does not require is the identity of all the actions at each step,
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  - I advise changing title Experimental Results $=>$ Experimental Evaluation
+\end{mdframed}
 
-nor the instantiation of their headers, but it does require the definition of all
+\textcolor{blue}{We changed the title as recommended by the reviewer.}
 
-the predicates that may appear in any state. The truth it seems to me is that
+\vspace{0.1cm}
 
-FAMA would be good for some applications which fit this set of constraints, which is
-
-a different set of constraints than for other domain model learners.
-
-A way forward might be to put the claims about FAMA into a smaller set of learning systems
-
-(certainly ARMS will be in that set) where it clearly needs less input information.
+\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
+  - At the beginning of the evaluation you state the obvious - that the purpose is to evaluate the performance of FAMA and the quality of the models.
+At this stage to clarify the experimental section I suggest that you indicate, before detailing the set up, what would be success, for example how do you intend to show that your technique is better than or comparable to others?
+This will no doubt mean introducing the metrics you intend to use and giving an idea of success in terms of them.
+Also it may mean taking text from the start of the sections 6.2 - 6.5  into the first part of the evaluation section.
+This will however give a good rational for the 'set-up‚Äô, the choice of the particular experimental scenarios, and for the discussion of the results.
 \end{mdframed}
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  I totally agree with the authors in their assertion that the evaluation methods for
-
-previous model learners is currently poor, and amount to how close the learned model
-
-is to a GTM. I am not convinced, however, that the use of
-
-the statistic methods proposed is a great improvement. Precision and Recall
-
-are heavily used in data mining where (generally) there are huge data sets and there
-
-is not generally a hand crafted model for comparison. In planning, however, the idea of
-
-separating out a set of traces into training and testing appears odd, especially where
-
-it is known that the learning will eventually create a model that explains all the training
-
-traces. This seems particularly odd if there are only a few traces to be split up (and FAMA appears
-
-only able to cope  with a few traces because of complexity problems).
+  - Why "14" models? please give some explanation (e.g. there are only 14 strips models there) so we know why/if you have left any out
 \end{mdframed}
 
+\textcolor{blue}{We keep adding new domains as we work on the approach so that number is representative of how many domains we have tested our approach on.
+We did, however, leave one domain out, {\em Sokoban}.
+As a matter of fact, this domain has some of the features more difficult to handle for our approach, those are puzzle-like domains and domains with grid-like structure.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  - The method of learning: This is to compile the learning problem into a conditional-operater
-
-classical planning problem. This method appears to be highly innovative to this line of research
-
-(stemming from an earlier ICAPS paper by the authors).
-
-The description of the compilation method did not appear to me to be at the level
-
-of the detail to allow easy reading. It took me a good deal of time to work out
-
-what going on. This is a complex process and I recommend much more use of examples
-
-throughout - perhaps a running example to show the full run of a compilation.
-
-It might also improve the paper if an algorithmic description of compilation
-
-were included? One aspect that was not obvious from the description was how a
-
-SET of models are delivered by planning rather than just one; this needs spelling out.
+  was the max predicate arity of "2" in all of the 14 a big issue? Given the inherent computational problems in planning I guess so?
 \end{mdframed}
 
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  - The evaluations of the domain model using classification metrics of precision and recall:
-
-I like the distinction of the methods labelled "syntactic" (using the GTM) vs "semantic" here;
-
-however I think the use of the word - "semantic” a bit misleading. This implies
-
-taking into account the meaning of the features within the application domain which are
-
-left out or in erroneously compared to the GTM e.g. that a missing "on(x,y)” in a learned model
-
-is somewhat better/worse from a missing ”holding(x)” - because of their meaning, rather than
-
-counting each predicate as ‘1’.
-
-In the end your method reverts to counting the number of different predicates, but in the context
-
-of the recall/precision formula with respect to domain model(s) that are consistent with the test
-
-sets, not only the GTM.
-
-The definition of the metrics was not too clear. Intuitively, I think I understood 5.1 but it
-
-was not well stated -the definition of some terms seemed syntactically wrong or at least sloppy.
-
-E.g. in the definition of INS and DEL, should not the universally quantified variables be further
-
-qualified by stating that there is a union of the already unioned sets? Or again, should there not
-
-be an extra "Sigma" to represent the addition of the addition of the number of precons and
-
-effects in size(M) definition?  In proposition 7 you say the closest set "is the GTM"
-
-do you mean equal to the singleton set of the GTM?
-
-Also - the structure of 5.1 - it would be best to clearly state that
-
-the syntactic evaluation is going to be introduced, then as a further subsection the semantic
-
-evaluation.
-\end{mdframed}
+\textcolor{blue}{Predicates with arity 3 are certainly harder but still manageable. In order to proof this, we have added the {\em rovers} domain to the list of domains tested in section 6. }
 
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  - I advise changing title Experimental Results => Experimental Evaluation
-\end{mdframed}
-
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  - At the beginning of the evaluation you state the obvious - that the purpose is to evaluate
-
-the performance of FAMA and the quality of the models. At this stage to clarify the experimental
-
-section I suggest that you indicate, before detailing the set up, what would be success, for example
-
-how do you intend to show that your technique is better than or comparable to others? This will
-
-no doubt mean introducing the metrics you intend to use and giving an idea of success
-
-in terms of them. Also it may mean taking text from the start of the sections 6.2 - 6.5
-
- into the first part of the evaluation section. This will however give a good rational for the
-
-'set-up’, the choice of the particular experimental scenarios, and for the discussion of the
-
-results.
-\end{mdframed}
-
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  - Why "14" models? please give some explanation (e.g. there are only 14 strips models there)
-
-so we know why/if you have left any out
-\end{mdframed}
-
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  was the max predicate arity of "2" in all of the 14 a big issue? Given the inherent computational
-
-problems in planning I guess so?
-\end{mdframed}
-
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-  so for the NO / NO case study, we know the full IS , GS, the number of action steps,
-
-and the space of possible action headers for each step - is that correct?
-
+  so for the NO / NO case study, we know the full IS , GS, the number of action steps, and the space of possible action headers for each step - is that correct?
 If so please make it more explicit - it not say explicitly what it means.
-
-This was not very explicit in the introduction/compilation process
-
-description either.
+This was not very explicit in the introduction/compilation process description either.
 \end{mdframed}
@@
 
-“around a hundred of traces”p.20
+"around a hundred of traces"ùp.20
 
-Please replace “fairly sound and almost totally complete” with something more precise. p.21
+Please replace "fairly sound and almost totally complete"ù with something more precise. p.21
 
-“reason why the scores of” p.26
+"reason why the scores of"ù p.26
 \end{mdframed}
 
+\textcolor{blue}{Fixed.}
+
 \end{document}

### 2018-12-24 doc/aij18/major-revision/introduction.tex
section 1 aij paper revised

@@
 
-There is common agreement in the planning community that the unavailability of an \textcolor[rgb]{1.00,0.00,0.00}{adequate} domain model is a bottleneck in the applicability of planning technology to many real-world domains~\cite{kambhampati:modellite:AAAI2007}.
+There is common agreement in the planning community that the unavailability of an adequate domain model is a bottleneck in the applicability of planning technology to many real-world domains~\cite{kambhampati:modellite:AAAI2007}.
 Motivated by the difficulty and cost of crafting action models, research in action-model learning has seen huge advances.
@@
 
-A system for learning planning action models receives as an input observations of the agent's plan execution and \textcolor[rgb]{1.00,0.00,0.00}{outputs an approximation of the actions that embody the physics of the real-world domain being modeled.} The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards, \textcolor[rgb]{1.00,0.00,0.00}{but there exists as well a large variety of planning-related tasks that rely upon the existence of a planning model.
-Among these tasks, we might cite: \emph{goal and plan recognition} approaches based on a domain theory~\cite{ramirez2009plan,ramirez2012plan,SohrabiRU16}; \emph{transparent planning}, in which an agent implicitly communicates its true goal by making its intentions and its action selection transparent (recognizable) to observers~\cite{MacNallyLRP18}; or \emph{deceptive path-planning}, which draws on the definition of path-planning domains and aims at finding a path such that the probability of an observer identifying the final destination is minimised~\cite{MastersS17}.
+A system for learning planning action models receives as input observations of the agent's plan execution and \textcolor[rgb]{1.00,0.00,0.00}{outputs an approximation of the actions that embody the physics of the modeled domain.} The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards, \textcolor[rgb]{1.00,0.00,0.00}{but there exists as well a large variety of planning-related tasks that rely upon the existence of a planning model.
+Among these tasks we might cite: \emph{goal and plan recognition} approaches based on a domain theory~\cite{ramirez2009plan,ramirez2012plan,SohrabiRU16}; \emph{transparent planning}, in which an agent implicitly communicates its true goal by making its intentions and its action selection {\em transparent} (recognizable) to observers~\cite{MacNallyLRP18}; or \emph{deceptive path-planning}, which aims at finding a path such that the probability of an observer identifying the final destination is minimised~\cite{MastersS17}.
 Planning models are also used in \emph{explainable AI planning} to form a common basis for communicating with users and facilitate the generation of transparent and explainable decisions~\cite{FoxLM17} as well as explanations in terms of the differences with a human mental model~\cite{ChakrabortiSK18}.
-\emph{Counterplanning} requires a model of the opponent agent in order to recognize its goals \cite{PozancoEFB18} and \emph{model reconciliation} aims to conform the models of two agents with respect to an observation of a plan computed with one of the two models~\cite{ChakrabortiSZK17}.}
+\emph{Counterplanning} requires also a model of the opponent agent in order to recognize its goals \cite{PozancoEFB18} and \emph{model reconciliation} aims to conform the models of two agents with respect to an observation of a plan computed with one of the two models~\cite{ChakrabortiSZK17}.}
 
-\textcolor[rgb]{1.00,0.00,0.00}{Motivated by the requirement for a planning model in many different tasks and the recent advances on the use of classical planning for the generation of different types of planning models (regular automata, context-free grammars, finite-state machines, \strips)} \cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behavior is not available.
+\textcolor[rgb]{1.00,0.00,0.00}{Motivated by recent advances on the use of classical planning for the generation of different types of planning models (regular automata, context-free grammars, finite-state machines)} \cite{bonet2009automatic,segovia2016generalized,segovia2017generating,segovia:HFSCs:JAIR18}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behavior is not available.
 Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimal observability.
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{Current learning approaches assume that the observation of the agent's plan execution (plan trace) encompasses the fully observed sequence of the executed actions; i.e, they assume all the actions performed by the agent are observable.
+\textcolor[rgb]{1.00,0.00,0.00}{Most learning approaches assume that the observation of the agent's plan execution (plan trace) encompasses the fully observed sequence of the executed actions; i.e, they assume all the actions performed by the agent are observable.
 This heavily restricts the applicability of the learning approaches to contexts where the behaviour of the agent is fully observable, which also commonly entails a human annotator that correctly labels the executed actions.
 On the other hand, learning approaches accept a varying degree of observability in the states traversed in the plan trace, ranging from fully observable to fully unobservable states (see section \ref{related_work} for details).} In contrast, \FAMA allows for an incomplete or empty sequence of observable actions, and the minimum observability case acceptable by \FAMA is when the algorithm is only fed with the initial and final state of a plan trace.
-Like many Machine Learning (ML) techniques, \FAMA is able to operate with only input/output pairs of states and an unknown or a partially known model of the agent.
+Like many Machine Learning (ML) techniques, \FAMA is able to operate with only input/output pairs of states (pairs of initial and final states) and an unknown (or partially known) model of the agent.
 Unlike ML algorithms, \FAMA requires a symbolic structured representation of the input knowledge.
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{\FAMA is a new learning approach, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}.
+\textcolor[rgb]{1.00,0.00,0.00}{\FAMA is a new learning approach, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a classical planning task which is then solved with an off-the-shelf planner~\cite{aineto2018learning}.
 The construction of a \strips\ planning model starts out from a set of plan traces containing the observation of several plan executions.
 As mentioned above, a plan trace may comprise none of the actions executed by the agent but must include, at least, the initial state $s_0$ and final state $s_f$ of the execution.
-The compilation scheme lies in defining a planning task from the set of plan traces using a set of \emph{building actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces.
+The compilation scheme defines a planning task from the set of plan traces using a set of \emph{building actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces.
 Hence, a solution to this planning task is a plan that determines the preconditions and effects of the actions of a \strips\ planning model $M$.
 The rationale of this process draws its strength in that when the problem $\tup{s_0,s_f}$ is solved with the learned model $M$, a plan $\pi$ is generated such that $\pi$ will contain the observed actions of the plan trace, if any, and the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.
-In other words, the plan $\pi$ generated when solving the problem $\tup{s_0,s_f}$ with $M$ is \emph{compliant} with the plan trace.}
+In other words, the plan $\pi$ generated when solving the problem $\tup{s_0,s_f}$ with $M$ is \emph{consistent} with the plan trace.}
 
 \FAMA is thus a model-based approach that automatically builds its own planning model by logical inference from the input plan traces that contain the observations of the agent execution.
-This behaviour largely differs from ML techniques, which aim to minimize an error function on the training data.
+This behaviour diverges from ML techniques, which aim to minimize an error function on the training data.
 Moreover, \FAMA requires far less sample data (example plan traces) than typical ML algorithms, thus alleviating the dependency on the assumption that there are enough data for learning the action models~\cite{Zhuo15}.
@@
 Our semantic evaluation is generally more informative than counting the number of errors between two models and alleviates two important limitations of a purely syntax-based assessment: (a) that the learned model is syntactically different from the reference model but semantically correct and (b) that the learned model comprises correct though unnecessary preconditions in regards to the reference model.
-This latter issue is concerned with the qualification problem, which is defined as the actual impossibility of listing all the preconditions required for a real world action to have its intended effects~\cite{GinsbergS88}.
+This latter issue is concerned with the {\em qualification problem}, which is defined as the actual impossibility of listing all the preconditions required for a real world action to have its intended effects~\cite{GinsbergS88}.
 
@@
 In particular, \FAMA also accepts an input initial action model $M$ of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}, alongside the observation of the agent execution.
-In this case, \FAMA returns a model $M'$ that follows the input model $M$ and is compliant with the observations.
+In this case, \FAMA returns a model $M'$ that follows the input model $M$ and is consistent with the observations.
 We designed an {\em edition} mechanism that serves to correct the input model to the output model, which in turn defines an assessment of the accuracy with which $M$ explains the observations.
@@
 \begin{itemize}
-\item compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces
-\item the plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states
-\item the planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions
-\item a semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model
+\item Compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces.
+\item The plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states.
+\item The planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions.
+\item A semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model.
 \end{itemize} }
 
-Ultimately, \FAMA adopts certain ML-like characteristics, like the ability to learn from data that do not explicitly exhibit the agent's behaviour or the capacity of correcting an initially given action model.
 A first description of the \FAMA compilation scheme already appeared in our previous conference paper~\cite{aineto2018learning}.
@@
 \item A thorough elaboration of two semantic evaluation metrics that build upon the notions of {\em precision} and {\em recall} to evaluate the output action models with respect to observations of plan executions.
-\item An exhaustive empirical evaluation over 14 domains from the International Planning Competitions (IPCs).
+\item An exhaustive empirical evaluation over 15 domains from the International Planning Competitions (IPCs).
 We include an analysis of the impact that the size of the input knowledge has in the performance of \FAMA, a comparison with \ARMS, and a detailed experimentation when \FAMA is executed with minimal input knowledge.
@@
 Section~\ref{sec:background} introduces classical planning concepts and reviews related work on learning planning action models.
-Section~\ref{sec:motivation} motivates our compilation-to-planning approach for learning action models.
-Section~\ref{sec:learning} formalizes the learning task and presents the compilation scheme, the core of \FAMA.
-Sections~\ref{sec:evaluation} explains the evaluation of a learned model with respect to a reference model (syntactic evaluation) and with respect to a set of plan traces (semantic evaluation).
-Section~\ref{sec:experiments} reports the results of the experimental evaluation and, finally, Section~\ref{sec:conclusions} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.
+Section~\ref{task_definition} formalizes the addressed learning task and Section~\ref{sec:learning} presents the compilation scheme, the core of \FAMA.
+Sections~\ref{sec:evaluation} explains the evaluation of a learned model with respect to a reference model ({\em syntactic} evaluation) and with respect to a set of plan traces ({\em semantic} evaluation).
+Section~\ref{sec:experiments} reports the results of the experimental evaluation and finally, Section~\ref{sec:conclusions} discusses the strengths and weaknesses of the compilation approach and proposes several opportunities for future research.

### 2018-12-24 doc/aij18/major-revision/background.tex
section 2 aij paper revised

@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{In \FAMA, we need to implicitly represent the unobserved literals of states.
-In order to do so, we will adopt the \emph{open world assumption}; i.e., what is not known to be true in a state is unknown.
+\textcolor[rgb]{1.00,0.00,0.00}{\FAMA adopts the \emph{open world assumption}; i.e., what is not known to be true in a state is unknown.
 Consequently, states will explicitly include positive literals ($f$) and negative literals ($\neg f$) such that literals that are not in a state are unknown or unobserved.} Hence, a {\em state} $s$ is a full assignment of values to fluents; i.e.
@@
 Note that subtracting the complement of $\eff(a)$ from $s$ ensures that $\theta(s,a)$ remains a well-defined state with positive and negative literals.
-Then:}
+With this regard we say that:}
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{Since we restrict our attention to \strips\ action models learning, we will assume the set of syntactic constraints imposed by \strips\ models, namely }that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$.
+\textcolor[rgb]{1.00,0.00,0.00}{Since we restrict our attention to learning \strips\ action models, we will assume the syntactic constraints imposed by \strips\ models, namely } $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$.
 Additionally, actions $a\in A$ are instantiated from given action schemas, as in PDDL.
@@
 \begin{enumerate}
-\item we say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and \textcolor[rgb]{1.00,0.00,0.00}{there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$};  that is $\theta(s_i,\tup{a})=s_{i+1}$.
+\item We say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and \textcolor[rgb]{1.00,0.00,0.00}{there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$};  that is $\theta(s_i,\tup{a})=s_{i+1}$.
 \textcolor[rgb]{1.00,0.00,0.00}{This case clearly states that $\tau_s = s$, meaning that $\forall s_i \in \tau_s$, $s_i$ comprises all the literals of the corresponding state in the trajectory $s$ of the plan $\pi$.
 Formally, $\forall i, 1 \leq i  < n, |s_i| = |F|$.}
-\item we say that $\tau_s$ is a partially-observable (\PO) state trajectory if at least one intermediate state of $\tau_s$ is a partial assignment of values to fluents.
+\item We say that $\tau_s$ is a partially-observable (\PO) state trajectory if at least one intermediate state of $\tau_s$ is a partial assignment of values to fluents.
 \textcolor[rgb]{1.00,0.00,0.00}{Formally, $\exists i, 1 \leq i  < n, |s_i| < |F|$.} This means that one or more literals are missing in the intermediate $s_i$, all of which may be missing.
@@
 \begin{enumerate}
-\item when \textbf{all} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, $\tau_s$ is a \emph{non-observable} (\NO) state trajectory.
+\item When \textbf{all} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, $\tau_s$ is a \emph{non-observable} (\NO) state trajectory.
 \textcolor[rgb]{1.00,0.00,0.00}{Formally, $\forall i, 1 \leq i < n,  s_i= \emptyset$; i.e., $|s_i| = 0$.}
-\item when \textbf{none} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, we will refer to $\tau_s$ as a \POstar state trajectory.
+\item When \textbf{none} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, we will refer to $\tau_s$ as a \POstar state trajectory.
 \textcolor[rgb]{1.00,0.00,0.00}{Formally, $\exists i, 1 \leq i < n, s_i \neq \emptyset$; i.e., $0 < |s_i| < |F|$.}
@@
 \begin{enumerate}
-\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$.
+\item \textcolor[rgb]{1.00,0.00,0.00}{When \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$.
 In this case, $\tau_a$ contains all the necessary actions to transit every state $s_{i-1}$ to its corresponding successor state $s_{i}$, from $s_0$ to $s_n$.
 This is the type of input trace accepted by all the existing learning approaches (see section \ref{related_work} for details)}.
-\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence.
+\item \textcolor[rgb]{1.00,0.00,0.00}{When \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence.
 In this case, at least one of the necessary actions of the plan $\pi$ is missing in $\tau_a$.
 Formally, $\exists i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.}
-\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence.
+\item \textcolor[rgb]{1.00,0.00,0.00}{When \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence.
 Formally, $\forall i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.
@@
 
-================
-
-A plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$ holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}.
-
-NO TENGO CLARO LO QUE QUEREMOS DECIR CON ESTA FRASE (??).
-ADEMAS, NO ES CORRECTA PORQUE LOS ESTADOS DE UNA TRAZA PUEDEN SER FULL-OBSERVABLE Y $\mathcal{L}(F)$ SON PARTIAL ASSIGNMENTS OF VALUES TO FLUENTS.
-
-=================
-
-Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
-\textcolor[rgb]{1.00,0.00,0.00}{In section \ref{overview}, we expose the impact of all combinations of observed state trajectories and observed action sequences in the complexity of the learning task.}
+Plan traces can then be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
+%\textcolor[rgb]{1.00,0.00,0.00}{Section~\ref{task_definition} formalizes the addressed learning task according to the particular combination of observed state trajectories and observed action sequences.}
 
@@
 Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- almost all approaches rely on a comparison between the learned model and a {\em Ground-Truth Model}), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
-
 In the following, we present a comprehensive insight of the particularities of the seven systems presented in Table \ref{table:models_comparison1} and Table \ref{table:models_comparison2}.
-This exposition will help us to highlight in section \ref{sec:motivation} the value of our contribution \FAMA.
+%This exposition will help us to highlight in section \ref{sec:motivation} the value of our contribution \FAMA.
 
@@
          \hline
-		\FAMA & checking wrt GTM  & precision and recall & 14\\  & validation with a test set &  & 20-50 actions\\
+		\FAMA & checking wrt GTM  & precision and recall & 15\\  & validation with a test set &  & 20-50 actions\\
          \hline
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{One of the latest incorporations to the family of action model learning algorithms is \LOUGA.
+\textcolor[rgb]{1.00,0.00,0.00}{One of the latest entries to the family of action model learning algorithms is \LOUGA.
 This system uses a genetic algorithm to learn the effects of actions.
-In order to do this, each gene in the genome encodes whether a predicate is a positive effect, negative effect or none for a particular action, and the fitness of an individual is evaluated by reproducing the trace with the model encoded in the individual.
+Each gene in the genome encodes whether a predicate is a positive effect, negative effect or none for a particular action, and the fitness of an individual is evaluated by reproducing the trace with the model encoded in the individual.
 After a solution for the effects is found, an ad-hoc algorithm is used to infer preconditions by finding those literals that are always present before the execution of an action.

### 2018-12-24 doc/aij18/major-revision/experiments.tex
More answers for the response letter

@@
 The conclusions we draw from these experiments is that, when the plan horizon is known, learning with few input samples yield action models that contain 95\% of the preconditions and effects of the GTM plus some extra ones as indicated by the values of precision and recall.
-With unkown plan horizons, on the other hand, the learned models are generally more different from the GTM; while timeouts are the main cause of the drop in the score, we must point out that pure syntax-based metrics are not adequate to evaluate such under-constrained tasks since the phenomenon of \emph{reformulation} occurs and this largely impacts the results.
-We will provide experimental evidence of this in section \ref{minimal}.
+With unknown plan horizons, on the other hand, the learned models are generally more different from the GTM; while timeouts are the main cause of the drop in the score, we must point out that pure syntax-based metrics are not adequate to evaluate such under-constrained tasks since the phenomenon of \emph{reformulation} occurs and this largely impacts the results (we will provide experimental evidence of this in section \ref{minimal}).
+\textcolor{red}{These results emphatize a relevant feature our approach, the small size of the training set required by \FAMA in comparison with other approaches (see Table \ref{table:models_comparison2}).
+Unlike extensive-data approaches, our work explores an alternative research direction to learn action models from small amounts of plan traces.
+This is an important advantage, particularly in domains in which it is costly or impossible to obtain a significant number of training samples.}
 

### 2018-12-24 doc/aij18/major-revision/letter.tex
More answers for the response letter

@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ....}}
+\textcolor[rgb]{1.00,0.00,0.50}{That depends on the domain we are trying to learn.
+Certainly, most planning domains do not need predicates with repeated variables, but in order to ensure the completeness of our learning approach we have opted to include it.}
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ... TENEMOS QUE VER ESTO ... }}
+\textcolor[rgb]{1.00,0.00,0.50}{Removing symmetries would admittedly improve performance for heuristic planners.
+This removal can be achieved by ensuring the actions are programmed following some order, which can be done via some additional fluents.
+We think, however, that the compilation scheme should remain as simple and general as possible to facilitate its understanding.
+}
+
+\textcolor[rgb]{1.00,0.00,0.50}{Nevertheless, we want to thank the reviewer for the great idea and we will look into adding this optimization to the approach as a compilation option.}
 
@@
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-- "Let $\Omega_v = {v_i}_{i=1}^{max_{a\in A} ar(a)}$ be a new set of objects ‚Ä¶" ‚Äì I think "ar(a)" should be "|ar(a)|"
+- "Let $\Omega_v = {v_i}_{i=1}^{max_{a\in A} ar(a)}$ be a new set of objects "ì I think "$ar(a)$" should be "$|ar(a)|$"
 \end{mdframed}
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...}}
+\textcolor[rgb]{1.00,0.00,0.50}{The function $ar(\cdot)$ returns the arity of an action or predicate (the length of its argument list) and is first introduced in the Background section.}
 

### 2018-12-24 doc/aij18/major-revision/learning_task_def.tex
Updates in response letter

@@
-\section{Learning task: on the use of planning for solving the task.}
+\section{Learning task}
 \label{task_definition}
 
-In this section, we firstly define the concept of \emph{learning task}.
-Subsequently, we examine the particularities of the learning task to the different types of plan traces according to the observed state trajectories and action sequences.
+In this section we first define the addressed \emph{learning task}.
+Subsequently, we examine the particularities of the learning task to the different types of input plan traces (according to the observed state trajectories and action sequences).
 The analysis will serve to justify the use of planning for solving the learning task as well as to highlight the principal distinctive features of our approach \FAMA with respect to the related work reviewed in Section~\ref{related_work}.
@@
 An \strips\ action model is a tuple $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
-
 \begin{itemize}
 	\item The name, $name(\xi)$, and parameters, $pars(\xi)$, of the action model define the {\em header} of the model.
-	\item $pre(\xi)$, $del(\xi)$ and $add(\xi)$ represent the preconditions, negative effects and positive effects of the action model, respectively, \textcolor[rgb]{1.00,0.00,0.00}{which follow the set of syntactic \strips\ constraints defined in section \ref{basic_planning}; specifically,} $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+	\item $pre(\xi)$, $del(\xi)$ and $add(\xi)$ represent the preconditions, negative effects and positive effects of the action model (i.e.
+$del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$).
 \end{itemize}
-
 As an example, Figure~\ref{fig:stack} shows the action model of the {\em stack} operator from the four-operator {\em blocksworld} domain~\cite{slaney2001blocks} encoded in PDDL.
@@
 Ultimately, we can always assume that $\Lambda$ will contain the predicates $\Psi$ as well as the headers of the actions models, either explicitly provided in $\mathcal{M}$ or deducible from $\tau$.
-
-\vspace{0.1cm}
-
-A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a domain model $\mathcal{M}'$ that complies with the information of $\mathcal{M}$ and with the observed plan trace $\tau$.
-\textcolor[rgb]{1.00,0.00,0.00}{This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan encompass the plan trace $\tau$.}
-
-Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\tau}$ corresponding to the observation of the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
+Figure~\ref{fig:example-plans} shows the learning task $\Lambda=\tup{\mathcal{M},\tau}$ corresponding to the observation of the execution of the four-action plan $\pi=\tup{\tiny\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
 In this example $\tau=\langle s_0,${\small\tt (putdown\ B),(stack\ A\ B)}, $s_4\rangle$.
@@
 
-The definition of the learning task is extensible to the more general case where the execution of several plans from the same action models are observed.
+A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a domain model $\mathcal{M}'$ that is consistent with the information of $\mathcal{M}$ and the observed plan trace $\tau$.
+\textcolor[rgb]{1.00,0.00,0.00}{This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan encompass the plan trace $\tau$.}
+
+Our definition of the learning task is extensible to the more general case where the execution of several plans from the same action models are observed.
 In this case, $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_{k}\}$ such that each $\tau\in \mathcal{T}$ is a plan trace that satisfies the previous 1--4 assumptions.
-In this case, the {\em learned} domain model $\mathcal{M}'$ must be compliant with the input model $\mathcal{M}$ and with every observed plan trace $\tau\in \mathcal{T}$.
+In this case, the {\em learned} domain model $\mathcal{M}'$ must be consistent with the input model $\mathcal{M}$ and every observed plan trace $\tau\in \mathcal{T}$.
 
-\vspace{0.15cm}
+\subsection{On the use of planning for learning action models}
 
-\textcolor[rgb]{1.00,0.00,0.00}{The key to understanding the intricacies of solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ lies in the type of plan trace $\tau = \tup {s_0, \ldots, s_n}$.
-Let $\pi$ be the plan that solves the planning problem $\tup {s_0,s_n}$ with a learned domain model $\mathcal{M}'$ and $\tau$ be the observation of $\pi$.
-Since $\tau$ contains a (partial) observation of the execution of $\pi$, actions, fluents or states traversed by $\pi$ may be missing in $\tau$ (we will use $\tau_s$ and $\tau_a$ to refer to the observed states and observed actions of $\tau$, respectively).
-We distinguish two well-identified cases:}
+\textcolor[rgb]{1.00,0.00,0.00}{Since $\tau$ is a (partial) observation of the execution of $\pi$, actions, fluents or states traversed by $\pi$ may be missing in $\tau$ (we will use $\tau_s$ and $\tau_a$ to refer to the observed states and observed actions of $\tau$, respectively).
+We distinguish two well differentiated cases:}
+
+\begin{itemize}
+\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ determines the length of $\pi$}.
+This happens in three different scenarios (1), when $\tau_a$ is a \FO action sequence (we know the actions of $\pi$) (2), when $\tau_s$ is a \FO state trajectory (the states induced by the actions of $\pi$ are fully known) or (3), when $\tau$ is a \POstar state trajectory (we have at least one fluent for every state induced by $\pi$, in which case we know there is a single action that transitions from every state of $\tau_s$).}
 
     \begin{enumerate}
-\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ determines the length of $\pi$}.
-This happens when $\tau_a$ is a \FO action sequence (we know the actions of $\pi$), $\tau_s$ is a \FO state trajectory (the states induced by the actions of $\pi$ are fully known) or $\tau$ is a \POstar state trajectory (we have at least one fluent for every state induced by $\pi$, in which case we know there is a single action that transitions from every state of $\tau_s$).}
-
-    \begin{itemize}
-    \item The common assumption of having \FO action sequences in a learning task, as is the case of all the learning approaches presented in section \ref{related_work} except for \FAMA, is unrealistic in many domains as it commonly implies the existence of human observers that annotate the observed action sequences.
+    \item[(1)] The common assumption of having \FO action sequences in a learning task, as is the case of all the learning approaches presented in section \ref{related_work} except for \FAMA, is unrealistic in many domains as it commonly implies the existence of human observers that annotate the observed action sequences.
 In some real-world applications, the observed and collected data are sensory data (e.g., home automation, robotics) or images (e.g.
@@
 Learning, for instance, from unstructured data involves transforming the sensor or image information into a predicate-like format before applying the action-model learning approach, and it also requires the ability of identifying action symbols~\cite{AsaiF18}.
-    \item The assumption of having \FO state trajectories means that the sensors are able to capture every state change at every instant, which is also typically unrealistic.
+    \item[(2)] The assumption of having \FO state trajectories means that the sensors are able to capture every state change at every instant, which is also typically unrealistic.
 Normally, the process of obtaining state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two subsequent sensor readings.
-    \item \textcolor[rgb]{1.00,0.00,0.00}{The assumption of having \POstar state trajectories seems more appropriate to reflect a real-world sensor reading but still requires that at least one fluent of every state traversed by $\pi$ is captured by the sensors.}
-    \end{itemize}
+    \item[(3)] \textcolor[rgb]{1.00,0.00,0.00}{The assumption of having \POstar state trajectories seems more appropriate to reflect sensor readings but still requires that at least one fluent of every state traversed by $\pi$ is captured by the sensors.}
+    \end{enumerate}
 
-When the length of $\pi$ is given by $\tau$, the learning task is \textbf{SAT compilable}, and it is known that a boolean satisfiability problem is a NP-complete task~\cite{cook1971complexity}.
-This is the reason why SAT solving is a common technique in the approaches presented in section \ref{related_work}.
-Particularly, when $\tau_a$ is a \FO action sequence and $\tau_s$ is a \FO state trajectory, learning \strips\ action models is straightforward~\cite{jimenez2012review}.
+In these three scenarios the length of $\pi$ is given by $\tau$, the learning task is \textbf{SAT compilable}, and it is known that a Boolean satisfiability problem is a NP-complete task~\cite{cook1971complexity}.
+This is the reason why SAT solvers are commonly used in the approaches presented in section \ref{related_work}.
+%Particularly, when $\tau_a$ is a \FO action sequence and $\tau_s$ is a \FO state trajectory, learning \strips\ action models is straightforward~\cite{jimenez2012review}.
 In this case the {\em pre-} and {\em post-states} of every action are available and so action effects are derived lifting the literals that change between the pre and post-state of the corresponding action executions.
@@
 
-\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ does not indicate the length of $\pi$}.
+\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ does not identify the length of $\pi$}.
 This happens when $\tau_a$ and $\tau_s$ are both partially observed (\PO) or non-observable (\NO) action/state trajectories.
 In this case, we are unaware of the number of actions of $\pi$ and the number of states induced by $\pi$.
-This gives rise to a completely different scenario and a more challenging learning task that brings one key difference: the transition between two given observed states of $\tau$ may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbounded, and so the horizon of $\pi$ is no longer known.
+This gives rise to a completely different scenario and a more challenging learning task that brings one key difference: the transition between two given observed states of $\tau$ may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbounded, and so {\bf the horizon of $\pi$ is no longer known}.
 This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$.
-As it is also well known, the worst case complexity of \strips\ planning is PSPACE-complete so we can conclude our learning task becomes PSPACE-complete instead of NP-complete when combining \PO and \NO state trajectories and action sequences.}
+As it is also well known, the worst case complexity of \strips\ planning is PSPACE-complete so the learning task we address in this paper is PSPACE-complete instead of NP-complete.}
 
-   \textcolor[rgb]{1.00,0.00,0.00}{ In this particular scenario, the actual number of plans compliant with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).
+   \textcolor[rgb]{1.00,0.00,0.00}{ In this particular scenario, the actual number of plans that are consistent with the input plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).
 Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.}
@@
 %\FAMA shows that classical planning is a suitable approach for this particular scenario
-\end{enumerate}
+\end{itemize}
 

### 2018-12-24 doc/aij18/major-revision/learning_task_def.tex
Add files via upload

--- 2018-12-21 doc/aij18/major-revision/learning_task_def.tex
@@
 Since $\tau$ contains a (partial) observation of the execution of $\pi$, actions, fluents or states traversed by $\pi$ may be missing in $\tau$ (we will use $\tau_s$ and $\tau_a$ to refer to the observed states and observed actions of $\tau$, respectively).
-We distinguish two well-identified cases:}
+We distinguish two well differentiated cases:}
 
@@
 When the length of $\pi$ is given by $\tau$, the learning task is \textbf{SAT compilable}, and it is known that a boolean satisfiability problem is a NP-complete task~\cite{cook1971complexity}.
-This is the reason why SAT solving is a common technique in the approaches presented in section \ref{related_work}.
+This is the reason why SAT solvers are commonly used in the approaches presented in section \ref{related_work}.
 Particularly, when $\tau_a$ is a \FO action sequence and $\tau_s$ is a \FO state trajectory, learning \strips\ action models is straightforward~\cite{jimenez2012review}.
@@
 
-\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ does not indicate the length of $\pi$}.
+\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ does not identify the length of $\pi$}.
 This happens when $\tau_a$ and $\tau_s$ are both partially observed (\PO) or non-observable (\NO) action/state trajectories.

### 2018-12-24 doc/aij18/major-revision/experiments.tex
More details on timeouts

@@
 \textcolor{red}{The higher complexity of this setting is also reflected in the appearance of some timeouts when solving the learning task.
-The first timeout is found at $|\mathcal{T}| = 3$ in the \emph{grid} domain, and the number of timeouts increases until it reaches 6 with $|\mathcal{T}| = 10$.} We can observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
+The first timeout is found at $|\mathcal{T}| = 3$ in the \emph{grid} domain, then, with $|\mathcal{T}| = 4$ \emph{floortile} also times out, and by the time $|\mathcal{T}|$ reaches 10 the number of domains were no solution is found is 6, with the addition of the \emph{npuzzle, parking, rovers and zenotravel} domains.} We can observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
 The drop in the score is caused by the increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for precision and recall is assigned to these experiments.

### 2018-12-24 doc/aij18/major-revision/background.tex

--- 2018-12-09 doc/aij18/major-revision/background.tex
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{In \FAMA, we need to implicitly represent the unobserved literals of states.
-In order to do so, we will adopt the \emph{open world assumption}; i.e., what is not known to be true in a state is unknown.
+\textcolor[rgb]{1.00,0.00,0.00}{We will adopt the \emph{open world assumption}, that is, what is not known to be true in a state is unknown, to implicitly represent the unobserved literals of states.
 Consequently, states will explicitly include positive literals ($f$) and negative literals ($\neg f$) such that literals that are not in a state are unknown or unobserved.} Hence, a {\em state} $s$ is a full assignment of values to fluents; i.e.
@@
 
-================
-
-A plan trace $\tau = \langle s_0, a_1, s_1, a_2, s_2, \ldots, a_n, s_n \rangle$ for a planning frame $\Phi=\tup{F,A}$ holds that $a_i\in A$ for every action in $\tau$ and that $s_i \in \mathcal{L}(F)$ for each {\small $1\leq i\le n$}.
-
-NO TENGO CLARO LO QUE QUEREMOS DECIR CON ESTA FRASE (??).
-ADEMAS, NO ES CORRECTA PORQUE LOS ESTADOS DE UNA TRAZA PUEDEN SER FULL-OBSERVABLE Y $\mathcal{L}(F)$ SON PARTIAL ASSIGNMENTS OF VALUES TO FLUENTS.
-
-=================
-
 Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
-\textcolor[rgb]{1.00,0.00,0.00}{In section \ref{overview}, we expose the impact of all combinations of observed state trajectories and observed action sequences in the complexity of the learning task.}
+\textcolor[rgb]{1.00,0.00,0.00}{In section \ref{task_definition}, we expose the impact of the combinations of observed state trajectories and observed action sequences when solving a learning task.}
 
@@
 
-Most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces.
+%Most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces.
 Others do not explicitly say so but the fact is that predicates and actions headers are easily extractable from the state sequence and action sequence of the input plan traces, respectively.
@@
 In the following, we present a comprehensive insight of the particularities of the seven systems presented in Table \ref{table:models_comparison1} and Table \ref{table:models_comparison2}.
-This exposition will help us to highlight in section \ref{sec:motivation} the value of our contribution \FAMA.
+This exposition will also help us to highlight in section \label{task_definition} the value of our contribution \FAMA.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{The distinctive feature of \LOCMtwo lies in the need to learn the state variables (fluents) because predicates are not either provided as input nor they are deducible from the plan trace as no state observability is allowed.
+\textcolor[rgb]{1.00,0.00,0.00}{The distinctive feature of the \LOCM family lies in the need to learn the state variables (fluents) because predicates are neither provided as input nor they are deducible from the plan trace as no state observability is allowed.
 In contrast, \FAMA and the rest of approaches either assume the set of predicates are provided alongside the input traces or assume they are extractable from the observed states of the plan trace, in which case the plan trace must comprise at least a grounded sample of every predicate.

### 2018-12-24 doc/ijcai-19-conformant/conformant-learning.tex
adding updates to the IJCAI19 paper on conformant planning

@@
 \section{Introduction}
-Given an input sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to {\em explain} the given observation.
-This task is of interest because it allows the incremental learning of action models from arbitrary large sets of partial observations.
+Given an input sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to {\em explain} that input observation.
+
+This task is of interest because, if we can keep the set of all the action models that are consistent with the observation, i.e.
+the {\em version space}~\cite{mitchell1982generalization}, then each new observation has either no effect on the {\em version space} or eliminates some models from it.
+This allows to implement an incremental prunning of the {\em version space} and hence, to learn planning action models from arbitrary large sets of partial observations (one has never to go back and re-process old observations).
 
@@
 
-\newpage
-
 We are assuming then that there is a {\em bijective monotone mapping} between trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
@@
 
-\section{Computing the {\em least-commitment} action model from observations}
-First, this section formalizes the notion of the {\em least-commitment} action model that is able to {\em explain} an observation. Next, the section describes our approach to compute such model.
-
-\subsection{The {\em least-commitment} action model}
+\section{The {\em least-commitment} action model}
 The task of computing the {\em least-commitment} action model from a sequence of state observations is defined as $\tup{\Phi,\mathcal{O}(\tau)}$:
@@
 \subsection{{\em Partially specified} \strips\ action models}
-A set of action models can be defined {\em explicitly}, by enumerating all the models that belong to the set or {\em impolitely}, by enumerating the constraints that must satisfy any model that belongs to the set.
+A set of action models can be defined {\em explicitly}, enumerating all the models that belong to the set or {\em implicitly}, enumerating the constraints that must satisfy any model that belongs to the set.
 Inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}, we introduce here {\em partially specified} \strips\ action models, a formalism for the {\em implicit} representation of a set of \strips\ schema.
@@
 We show now that extending to the {\em knowledge level} our propositional encoding of \strips\ action schemes, we can compactly represent a set of \strips\ schema.
-The extension defines, for each proposition {\tt\small $pre\_e\_\xi$}, with two propositions {\tt\small $K\pre\_e\_\xi$} and {\tt\small $K\neg\pre\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is a precondition of $\xi$ and that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not a precondition of $\xi$.
+The extension defines, for each proposition {\tt\small $pre\_e\_\xi$}, two propositions {\tt\small $K\pre\_e\_\xi$} and {\tt\small $K\neg\pre\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is a precondition of $\xi$ and that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not a precondition of $\xi$.
 Likewise for each {\tt\small $eff\_e\_\xi$} proposition we define these two, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is an effect of $\xi$ and that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not an effect of $\xi$.
-With respect to this formalization, the {\em full space} of possible \strips\ schemas for $\xi$ is compactly represented as {\tt\small $\bigcap_{e\in{\mathcal I}_{\Psi,\xi}} \neg K\pre\_e\_\xi\wedge \neg K\neg\pre\_e\_\xi \wedge\neg K\eff\_e\_\xi\wedge \neg K\neg\eff\_e\_\xi$}.
 
+\begin{definition}[Partially specified model]
+A {\em partially specified action schema} $\xi[\cdot]$ is an aritrary formula over the {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} propositions ($e\in{\mathcal I}_{\Psi,\xi}$) whose valuations express the enumeration of the set of action models $M_\xi$.
+\end{definition}
+
+With respect to this formalization, the {\em full space} of possible \strips\ schemas for $\xi$ is compactly represented by the {\em partially specified action schema} {\tt\small $\bigcap_{e\in{\mathcal I}_{\Psi,\xi}} \neg K\pre\_e\_\xi\wedge \neg K\neg\pre\_e\_\xi \wedge\neg K\eff\_e\_\xi\wedge \neg K\neg\eff\_e\_\xi$}.
 Figure~\ref{fig:partial} shows a {\em partially specified} \strips\ model that represents of a set of action models for the {\tt\small stack(?v1,?v2)} schema.
@@
 
+\section{Incremental learning of action models from observations}
+This section sows how to implement an incremental learning of action models by prunning action models that are not consistent with the input observations.
+
 \subsection{Computing the {\em least-commitment} model via conformant planning}
 The compilation for learning \strips\ action models defined by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} can be used to compute a {\em partially specified model} that explains a given observation by producing a conformant planning task instead of a classical planning task.
-To do so the initial state $\Upsilon$ of the conformant planning task does not code the {\em empty} action model but the full space of action models.
+The initial state $\Upsilon$ of the conformant planning task does not code an {\em initial} action model (typically {\em empty}) but the full space of action models.
 In more detail, the clauses in $\Upsilon$ comprises:
@@
       \item The clauses representing that the actual value of fluents {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} is unknown.
-In other words, that any model from the \strips\ space of models (following the previously mentioned syntactic constraints) can initially be part of the {\em least-commitment} action model.
+In other words, that any model from the \strips\ space of models (following the previously mentioned {\em syntactic constraints}) can initially be part of the {\em least-commitment} action model.
 Formally, for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$, then $\Upsilon$ includes these two clauses:
@@
             \end{itemize}
-            One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier for a specific domain.
+            One can also add here clauses that encode {\em domain-specific constraints} (as mentioned in the previous section) to make the conformant planning problem easier to be solved for a specific domain.
       \end{enumerate}
-A conformant plan that solves this task is a {\em partially specified model} that {\em explains} the input observation.
-To compute the {\em least-commitment} action model, we have to associate a cost of 1 to the actions that {\em program} a given precondition or effect of an action schema, a cost of 0 to the remaining actions, and compute the {\em optimal} plan for the resulting conformant planning task.
+A conformant plan that solves this task induces a {\em partially specified model} that {\em explains} the input observation.
+To get the {\em least-commitment} action model, we have to compute the {\em optimal} plan for the resulting conformant planning task when: The actions that {\em program} a given precondition (or effect) of an action schema have an associated cost of 1.
+These actions represent now a reduction of one unit in the {\em uncertainty} of the initial {\em partially specified model}.
+In other words, these actions implement the different possible {\em immediate specification} of the current {\em partially specified model}.
+The remaining actions have a cost of 0.
 
@@
 
+\subsection{Learning algorithm}
+Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ where the semantics of each action $a\in A[\cdot]$ is unknown and a set of state observations $\mathcal{O}_1,\ldots,\mathcal{O}_T$.
+Our learning algorithm for the incremental learning of action models from state observations is defined as follows:
+\begin{enumerate}
+\item Build the {\em least-commitment} task $\tup{\Phi,\mathcal{O}_t}$ to process observation $\mathcal{O}_t$, {\tt\small $1\leq t\leq T$}.
+\item Build the classical planning task $P=\tup{F',A',I',G'}$ to solve the previous task.
+\item Extract the {\em partially specified} model from the previous solution.
+\item Continue until, the {\em version space} is a singleton or no more observations left
+\end{enumerate}
+
+Assuming and a set of independent observations $\{\mathcal{O}_1,ldots,\mathcal{O}_t\}$ we need to see at least $\frac{ln{1-\delta} + ln{}}{\epsilon}$ to be sure that the prediction error is at most $\epsilon$ with probability $(1-\delta)$.
+
 \section{Evaluation}

### 2018-12-24 doc/aij18/major-revision/experiments.tex
update 24 Dec 20.00

--- 2018-12-24 doc/aij18/major-revision/experiments.tex
@@
   \item \textcolor[rgb]{1.00,0.00,0.00}{Other planners such as {\sc FastDownward} were also tested but provided worse experimental results}
-  \item \textcolor[rgb]{1.00,0.00,0.00}{A SAT-based planner like {\sc Madagascar} is particularly suitable for tasks where the observed plan trace $\tau$ determines the length of the solution plan (case 1 presented in section \ref{task_definition})}.
-In this case, the prefix of the solution plan is solvable in two steps (planning horizon length 2) because the actions for inserting preconditions can be applied in parallel in a single step and the same for the actions inserting the effects.
+  \item \textcolor[rgb]{1.00,0.00,0.00}{A SAT-based planner like {\sc Madagascar} is particularly suitable for tasks where the observed plan trace $\tau$ determines the horizon of the solution plan (case 1 presented in section \ref{task_definition})}.
+In this case, a SAT-based planner solves the prefix of the solution plan in two time steps because the actions for inserting preconditions can be applied in parallel in a single time step and the same for the actions inserting the effects.
   \item \textcolor[rgb]{1.00,0.00,0.00}{When the length of the plan is unknown (case 2 presented in section \ref{task_definition}), the ability of {\sc Madagascar} to deal with instances populated with dead-ends is very helpful~\cite{lopez2015deterministic}.}
@@
 \textcolor{red}{The higher complexity of this setting is also reflected in the appearance of some timeouts when solving the learning task.
-The first timeout is found at $|\mathcal{T}| = 3$ in the \emph{grid} domain, and the number of timeouts increases until it reaches 6 with $|\mathcal{T}| = 10$.} We can observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
+The first timeout is found in the \emph{grid} domain at $|\mathcal{T}| = 3$; the \emph{floortile} times out with $|\mathcal{T}| = 4$; and by the time $|\mathcal{T}|$ reaches 10 the number of domains were no solution is found is 6, adding the \emph{npuzzle, parking, rovers and zenotravel} domains.} We can observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
 The drop in the score is caused by the increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for precision and recall is assigned to these experiments.
@@
 With unknown plan horizons, on the other hand, the learned models are generally more different from the GTM; while timeouts are the main cause of the drop in the score, we must point out that pure syntax-based metrics are not adequate to evaluate such under-constrained tasks since the phenomenon of \emph{reformulation} occurs and this largely impacts the results (we will provide experimental evidence of this in section \ref{minimal}).
-\textcolor{red}{These results emphatize a relevant feature our approach, the small size of the training set required by \FAMA in comparison with other approaches (see Table \ref{table:models_comparison2}).
+\textcolor{red}{These results emphasize a relevant feature our approach: the small size of the training set required by \FAMA in comparison with other approaches (see Table \ref{table:models_comparison2}).
 Unlike extensive-data approaches, our work explores an alternative research direction to learn action models from small amounts of plan traces.

### 2018-12-24 doc/aij18/major-revision/learning_action_models.tex
update 24 Dec 20.00

@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{      There will be a validate action in $\pi_\Lambda$ per observed state in $\tau$.
-The position of the validate actions in $\pi_\Lambda$ will be determined by the planner checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.}
+\textcolor[rgb]{1.00,0.00,0.00}{There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$.
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.}
 
@@
 In order to address this shortcoming, we apply a post-process based on the one proposed in \cite{kuvcera2018louga}.
-In their proposal, they obtain a full trace by applying the sequence of actions of the input trace and infer the preconditions from this fully observable trace.
+In their proposal, they obtain a full trace by applying the sequence of actions of the input trace and infering the preconditions from this fully observable trace.
 In our case, since the sequence of actions of the input trace might not be fully observable, we produce the traces by applying the actions found in the validation part of the solution plan.

### 2018-12-24 doc/aij18/major-revision/letter.tex
Updated 24 Dec 20.00

@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{That depends on the domain we are trying to learn.
+\textcolor{blue}{It depends on the domain we are trying to learn.
 Certainly, most planning domains do not need predicates with repeated variables, but in order to ensure the completeness of our learning approach we have opted to include it.}
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...  SE SIGUEN CUMPLIENDO LAS SYNTACTIC STRIPS CONSTRAINTS CON CONDITIONAL EFFECTS????? }}
+\textcolor{blue}{On the other hand, the assumptions regarding the syntactic \strips\ constraints still apply when using conditional effects}
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{Removing symmetries would admittedly improve performance for heuristic planners.
+\textcolor{blue}{Removing symmetries would admittedly improve performance for heuristic planners.
 This removal can be achieved by ensuring the actions are programmed following some order, which can be done via some additional fluents.
-We think, however, that the compilation scheme should remain as simple and general as possible to facilitate its understanding.
-}
+We think, however, that the compilation scheme should remain as simple and general as possible to facilitate its understanding.}
 
-\textcolor[rgb]{1.00,0.00,0.50}{Nevertheless, we want to thank the reviewer for the great idea and we will look into adding this optimization to the approach as a compilation option.}
+\textcolor{blue}{Nevertheless, we want to thank the reviewer for the great idea and we will look into adding this optimization to the approach as a compilation option.}
 
@@
 We have added the following sentence in the document to clarify this point: "There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$.
-The position of the validate actions in $\pi_\Lambda$ will be determined by the planner checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$."}
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$."}
 
@@
 \textcolor{blue}{With respect to the repetition of the definition of a \strips\ action model, conditional effects are introduced in section 4 because these are exclusively used for modeling the task that results from the compilation, $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$.
-Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}
+We have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}
 
@@
 
-\textcolor{blue}{As indicated by the reviewer, timeouts are only observed in the setting with NO action sequences and NO state trajectories.
-We have added a few sentences providing more details on this issue.}
-
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{SOLO HE VISTO UNA FRASE DE TIMEOUTS SOBRE EL GRID ... Y EL RESTO DE DOMINIOS? NO SE PUEDE DECIR NADA???? }}
+\textcolor{blue}{Timeouts are only produced in the setting with NO action sequences and NO state trajectories.
+We have added some details on the domains where timeouts are found: in the grid domain when $|\mathcal{T}| = 3$, floortile when $|\mathcal{T}| = 4$ and four other domains.
+We refer the reviewer to the comment added in page 23 of section 6.2.}
 
@@
 
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{EVA ...}}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 \end{mdframed}
+
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{EVA ...}}
 
@@
 
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{SERGIO ...}}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{In the remodeling of section 3, the paragraph about the size of the training dataset has been removed too. The ideas presented here have been moved to section 6.}
-
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ... NO HE VISTO ESTE PARRAFO EN LA SECCION 6, NO PUSISTE ESTA INFORMACION EN 6??}}
+\textcolor[rgb]{0.00,0.07,1.00}{In the remodeling of section 3, the paragraph about the size of the training dataset has been removed as well.
+The ideas presented here have been moved to page 23, at the end of section 6.2.}
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{The function $ar(\cdot)$ returns the arity of an action or predicate (the length of its argument list) and is first introduced in the Background section.}
+\textcolor{blue}{The function $ar(\cdot)$ does not return a set, it returns the arity of an action or predicate; i.e., the length of its argument list.
+This is introduced in the Background section for the first time.}
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...}}
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{SERGIO ...}}
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...  SE PUEDE DECIR ALGO MAS O IGNORAMOS EL COMENTARIO?}}
+\textcolor{blue}{The post-processing is applied to learning preconditions that are static literals and are never deleted or added.
+We inherit the same procedure for learning preconditions presented in [40].
+The pseudo-code is presented in [40] (algorithm 2).
+}
+
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{PENDIENTE DIEGO MIRE UNA FRASE QUE LE HE MANDADO POR EMAIL}}
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...  HE A—ADIDO ESTA FRASE EN EL TEXTO PERO SOLO CON RESPECTO AL GTM. MIRA A VER SI TE PARECE BIEN.}}
-
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{DIEGO ...  HE A—ADIDO ALGO PARA CLARIFICAR ESTO EN EL TEXTO PERO NO SE SI EST¡ BIEN ... COMPROBAR}}
-
-\textcolor{blue}{Hopefully the new section 3 (see main modification \#2) helps the reviewer clarify this question. }
+\textcolor{blue}{Yes, the reviewer is right in that we put it all in the planner and get the overall plan, both building actions and validating actions.
+What we mean by "solvable in two steps" is that a SAT-based planner internally solves the construction of the model in two \emph{time steps}: in one time step the truth value of all the model fluents that represent the preconditions is set, and in a second time step the value of all the model fluents that represent the effects is set.
+We have slightly modified the sentence to make it clear we refer to the SAT solving process of Madagascar.}
 
@@
 
+%#########################################
 \subsection*{Reviewer \#3}
+%#########################################
 
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
-First sentence of the Introduction: there is no satisfactory or generally accepted definition of "complete"ù domain model, given that a domain model is an abstraction of the domain being modelled.
+First sentence of the Introduction: there is no satisfactory or generally accepted definition of "complete" domain model, given that a domain model is an abstraction of the domain being modelled.
 If the authors use such a term they should  make clear what their definition is.
-"Adequate"ùis a better word to use - in the sense of being adequate enough for use with a planner to solve a set of planning tasks.
+"Adequate" is a better word to use - in the sense of being adequate enough for use with a planner to solve a set of planning tasks.
 \end{mdframed}
@@
 
+\textcolor{blue}{We have substantially changed the explanation of the compilation scheme.
+We refer the reviewer to the main modification \#3.
+Besides an introductory paragraph that provides an overview of the compilation, we have included examples throughout the text to clarify the role of the insert actions, the apply actions and the validate actions.
+We have also renamed some concepts with more illustrative names.}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 
+\textcolor[rgb]{1.00,0.00,0.50}{\textbf{PENDIENTE  .... y hay que revisar resultados, alguna cosa no cuadra ... }}
+
 \begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
@@
 
-\textcolor{blue}{We keep adding new domains as we work on the approach so that number is representative of how many domains we have tested our approach on.
+\textcolor{blue}{We keep adding new domains as we work on the approach so this number is representative of how many domains we have tested our approach on.
 We did, however, leave one domain out, {\em Sokoban}.
-As a matter of fact, this domain has some of the features more difficult to handle for our approach, those are puzzle-like domains and domains with grid-like structure.}
+As a matter of fact, this domain presents some of the most difficult features to be able to be handled by our approach, particularly features intrinsic in puzzle-like domains and domains with grid-like structure.
+We have though included one more domain, the {\em rovers} domain.}
 
@@
 
-\textcolor{blue}{Predicates with arity 3 are certainly harder but still manageable. In order to proof this, we have added the {\em rovers} domain to the list of domains tested in section 6. }
+\textcolor{blue}{Predicates with arity 3 are certainly harder but still manageable. In order to prove this, we have added the {\em rovers} domain to the list of domains tested in section 6. }
 

### 2018-12-25 doc/aij18/major-revision/experiments.tex
Update 25 Dec 11.00

@@
 Whenever applicable, we will consider scenarios with both known and unknown plan horizon to draw conclusions at both levels of complexity.
-We also devote one experiment to show the suitability of the novel evaluation metrics introduced in section \ref{sec:evaluation}.
+
+\textcolor[rgb]{1.00,0.00,0.00}{After presenting the setup of the experiments in section \ref{setup}, we introduce three experiments that measure different aspects like the minimal number of traces necessary to obtain good quality models (section \ref{size_knowledge}), a comparison of the quality of models obtained by \FAMA compared to \ARMS (section \ref{compare_ARMS}) and an analysis of the quality of models when using a very small number of input traces in section \ref{minimal} In these three experiments, the quality of the models is measured using the syntactic-based precision and recall metrics presented in section \ref{syntactic_precision_recall}, that is, we measure the precision and recall of the learned models with respect to the GTM for each domain.
+Finally, the syntactic evaluation of section \ref{minimal} is compared against a semantic evaluation, i.e.
+how well the learned models are able to reproduce the input plan traces, in section \ref{semantic}.}
+
+%We also devote one experiment to show the suitability of the novel evaluation metrics introduced in section \ref{sec:evaluation}.
 
 \subsection{Setup}
+\label{setup}
 
@@
 \subsection{Impact of the size of the input knowledge}
+\label{size_knowledge}
 
@@
 The drop in the score is caused by the increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for precision and recall is assigned to these experiments.
-Figure \ref{fig:pspace_time}, on the other hand, reflects that the computation time of the first case study is also higher than in the second case, which is explained by both the higher complexity and the large number of timeouts.
+Figure \ref{fig:pspace_time}, on the other hand, \textcolor[rgb]{1.00,0.00,0.50}{\textbf{reflects that the computation time of the first case study is also higher than in the second case},} which is explained by both the higher complexity and the large number of timeouts.
 
@@
 \subsection{Comparison with \ARMS}
+\label{compare_ARMS}
 
@@
 \subsection{Syntactic versus semantic evaluation}
+\label{semantic}
 

### 2018-12-25 doc/aij18/major-revision/letter.tex
Update 25 Dec 11.00

@@
 Also it may mean taking text from the start of the sections 6.2 - 6.5  into the first part of the evaluation section.
-This will however give a good rational for the 'set-up‚Äô, the choice of the particular experimental scenarios, and for the discussion of the results.
+This will however give a good rational for the 'set-up', the choice of the particular experimental scenarios, and for the discussion of the results.
 \end{mdframed}
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{PENDIENTE  .... y hay que revisar resultados, alguna cosa no cuadra ... }}
+\textcolor{blue}{We have introduced a paragraph at the beginning of section 6 that gives an overview of the experiments presented in this section.
+The quality of the models (\emph{success} in the words of the reviewer) of the three first experiments is measured by the syntactic metrics introduced in the evaluation section; that is, the quality of the models is compared against the corresponding GTM.
+The last section presents a syntatic vs semantic evaluation of the learned models.
+The semantic evaluation measures the precision and recall of the models to reproduce the input plan traces.
+All this is explained in the introductory paragraph.
+With respect to moving text from the subsections of section 6 to section 5, we think the text of these subsections is particularly contextualized to the experiment in question and seems better to leave the text in section 6.
+Nevertheless, we have improved some explanations in section 6.}
 

### 2018-12-25 doc/aij18/major-revision/letter.tex
Update Dec 25 11.30

@@
 \textcolor{blue}{The syntactic metrics commonly used by other approaches count the number of differences wrt the GTM and normalize by the total number of possible preconditions and effects.
-This produces an optimistic metric since the set of possible preconditions and effects is usually much larger than the set of learned ones, resulting in low error rates even when the model is largely flawed.
-In contrast, precision and recall normalize by the size of the model, removing this problem.
-Furthermore, precision and recall give a notion of soundness and completeness by splitting the differences into extra and missing preconditions/effects.
+This gives rise to an optimistic measure of the error since the set of possible preconditions and effects is usually much larger than the size of the actual GTM, resulting in low error rates even when the model is largely flawed.
+In contrast, precision and recall normalize by the size of the GTM, thus removing this problem.
+Furthermore, precision and recall provide an intuitive notion of soundness and completeness, respectively; whereas a sound model is interpreted as a model in which all preconditions and effects are correct wrt the GTM, a complete model is one in which no precondition or effect is missing.
 We have added some paragraphs to section 5 to better explain these concepts.}
 
-\textcolor{blue}{As the reviewer points out, the learning will create a model that explains all the training traces.
-There is, however, no guarantee that the training traces are able to capture all the dynamics of the domain.
+\textcolor{blue}{As the reviewer points out, \FAMA will create a model that explains all the training traces because it is a satisfying approach (that satisfies all the training samples) and so the training error is 0.
+There is, however, no guarantee that the training traces are able to capture the whole dynamics of the domain.
 This is specially true for scenarios with low observability like the ones explored in this paper.
-Hence, it is always interesting to measure how well the learned model is able to explain new plan traces.}
+Hence, it is always interesting to measure how well the learned model is able to explain a new set of plan traces.}
 

### 2018-12-25 doc/aij18/major-revision/learning_task_def.tex
Update Dec 25 at 12.02

@@
-\section{Learning task: on the use of planning for solving the task.}
+\section{Learning task}
 \label{task_definition}
@@
 
-A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a domain model $\mathcal{M}'$ that complies with the information of $\mathcal{M}$ and with the observed plan trace $\tau$.
-\textcolor[rgb]{1.00,0.00,0.00}{This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan encompass the plan trace $\tau$.}
-
 Figure~\ref{fig:example-plans} shows an example of a learning task $\Lambda=\tup{\mathcal{M},\tau}$ corresponding to the observation of the execution of the four-action plan $\pi=\tup{\small\tt (unstack\ B\ A), (putdown\ B), (pickup\ A), (stack\ A\ B)}$ for inverting a two-block tower.
@@
 \textcolor[rgb]{1.00,0.00,0.00}{Therefore, $\tau$ contains a \NO state trajectory because only the initial and final state are observed and the three intermediate states, $s_1$, $s_2$ and $s_3$, are missing; and a \PO action sequence where actions $a_2$ and $a_3$ are observed while $a_1$ and $a_4$ are unknown.} The initial domain model $\mathcal{M}$ only contains two of the four needed headers, but can be completed with the headers {\small\tt(putdown ?v1)} and {\small\tt(stack ?v1 ?v2)} inferred from $\tau$.
+
+A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a domain model $\mathcal{M}'$ that complies with the information of $\mathcal{M}$ and with the observed plan trace $\tau$.
+\textcolor[rgb]{1.00,0.00,0.00}{This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan encompass the plan trace $\tau$.}
 
@@
 
-The definition of the learning task is extensible to the more general case where the execution of several plans from the same action models are observed.
+Our definition of the learning task is extensible to the more general case where the execution of several plans from the same action models are observed.
 In this case, $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_{k}\}$ such that each $\tau\in \mathcal{T}$ is a plan trace that satisfies the previous 1--4 assumptions.
@@
 
-\vspace{0.15cm}
+\subsection{On the use of planning for solving the task.}
 
@@
 Let $\pi$ be the plan that solves the planning problem $\tup {s_0,s_n}$ with a learned domain model $\mathcal{M}'$ and $\tau$ be the observation of $\pi$.
-Since $\tau$ contains a (partial) observation of the execution of $\pi$, actions, fluents or states traversed by $\pi$ may be missing in $\tau$ (we will use $\tau_s$ and $\tau_a$ to refer to the observed states and observed actions of $\tau$, respectively).
+Since $\tau$ is a (partial) observation of the execution of $\pi$, actions, fluents or states traversed by $\pi$ may be missing in $\tau$ (we will use $\tau_s$ and $\tau_a$ to refer to the observed states and observed actions of $\tau$, respectively).
 We distinguish two well differentiated cases:}
@@
 \item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ determines the length of $\pi$}.
-This happens when $\tau_a$ is a \FO action sequence (we know the actions of $\pi$), $\tau_s$ is a \FO state trajectory (the states induced by the actions of $\pi$ are fully known) or $\tau$ is a \POstar state trajectory (we have at least one fluent for every state induced by $\pi$, in which case we know there is a single action that transitions from every state of $\tau_s$).}
+This happens in three scenarios: (1) when $\tau_a$ is a \FO action sequence (we know the actions of $\pi$); (2) when $\tau_s$ is a \FO state trajectory (the states induced by the actions of $\pi$ are fully known); or (3) when $\tau$ is a \POstar state trajectory (we have at least one fluent for every state induced by $\pi$, in which case we know there is a single action that transitions from every state of $\tau_s$).}
 
@@
 In this case, we are unaware of the number of actions of $\pi$ and the number of states induced by $\pi$.
-This gives rise to a completely different scenario and a more challenging learning task that brings one key difference: the transition between two given observed states of $\tau$ may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbounded, and so the horizon of $\pi$ is no longer known.
+This gives rise to a completely different scenario and a more challenging learning task that brings one key difference: the transition between two given observed states of $\tau$ may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbounded, and so \textbf{the horizon of $\pi$ is no longer known}.
 This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$.

### 2018-12-25 doc/aij18/major-revision/letter.tex
Update Dec 25 at 12.02

@@
 We believe this is the reason that our claim about using planning for solving the learning task was not fully comprehensible.
-We have reshaped the old section 3 (Motivation), which now is entitled "Learning task: on the use of planning for solving the task".
-First, we introduce the definition of the learning task (which was in section 4 in the previous version of the paper), and then we analyze the intricacies of the learning task to the different variants of observed plan traces.
-We organised the explanation around two well differentiated cases: when the observation identifies the length of the plan being executed and when it does not.
+We have reshaped the old section 3 (Motivation), which now is entitled "Learning task" and contains the subsection 3.1 "On the use of planning for solving the task".
+First, we introduce the definition of the learning task (which was in section 4 in the previous version of the paper), and then we analyze the intricacies of the learning task to the different variants of observed plan traces in section 3.1.
+We organised the explanation around two well differentiated cases: when the observation identifies the length of the executed plan and when it does not.
 Now, the discussion revolves around the particularities of solving the learning task in these two cases and we make some side comments on the complexity of the task.
@@
 Additionally, and as requested by the other reviewers too, we introduced the main modification \#2.
-The new section 3 defines the learning task and provides insight of the nature of the problem to solve according to the type of the observed plan trace.
+The new section 3 defines the learning task and provides insight of the nature of the problem to solve according to the type of the observed plan trace in section 3.1.
 We believe that this explanation also helps to grasp the overall picture of \FAMA.}
@@
 With respect to the relation between actions and state partial observability, we refer the reviewer to the main modification \#2.
-The new section 3 presents the combinations of observed state trajectories and
+The new section 3.1 presents the combinations of observed state trajectories and
 observed action sequences and the impact in the learning task solving.

### 2018-12-25 doc/aij18/major-revision/letter.tex
Update Dec 25 at 12.18

@@
 \textcolor{blue}{We agree with the reviewer's observation, and we believe the misunderstanding stems from us mixing up the words \emph{information} and \emph{observability}.
-In our claim of FAMA using the minimal amount of information, we should actually say that FAMA uses the \emph{minimal observability} possible, being the extreme case to have only the initial state and final state as observations.
+In our claim of \FAMA using the minimal amount of information, we should actually say that \FAMA uses the \emph{minimal observability} possible, being the extreme case to have no actions at all and only the initial state and final state as observations.
 In this respect, LOCM is indeed the learning approach that requires the \emph{minimal information} since no state or predicate is needed.}
 
-\textcolor{blue}{We think that LOCM is a learning approach that notably differs from the rest of systems presented in Table XXX and, in this regard, we have highlighted this distinction in section XXXXX.}
+\textcolor{blue}{We think that LOCM is a learning approach that notably differs from the rest of systems presented in Table 2 and, in this regard, we have highlighted this distinction in the explanation of section 2.2.
+Besides adding a sentence that stresses the ability of LOCM to learning with minimal input knowledge, we devote the last paragraph of section 2.2.
+to explain the distinctive features of the LOCM family.}
 

### 2018-12-25 doc/aij18/major-revision/letter.tex
Update Dec 25 at 12.20
Update Dec 25 at 13.00

@@
 
-\textcolor{blue}{The post-processing is applied to learning preconditions that are static literals and are never deleted or added.
-We inherit the same procedure for learning preconditions presented in [40].
-The pseudo-code is presented in [40] (algorithm 2).
-}
-
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{PENDIENTE DIEGO MIRE UNA FRASE QUE LE HE MANDADO POR EMAIL}}
+\textcolor{blue}{We have added a new sentence that overviews the post-process.
+We have also rephrases some part of the text to make it more comprehensible.
+We believe the text provides now a clear idea of the post-process and that the pseudo-code is not necessary.
+Nevertheless, we refer the reviewer to the pseudo-code presented in [40] (algorithm 2) for a more detailed explanation.}
 
@@
 to explain the distinctive features of the LOCM family.}
-
-\textcolor[rgb]{1.00,0.00,0.00}{NO ESTA HECHO TODAVIA}
 

### 2018-12-25 doc/ijcai-19-conformant/conformant-learning.tex
adding updates to the IJCAI19 paper on conformant planning

@@
 \section{Introduction}
-Given an input sequence of partially observed states, this paper formalizes the task of computing the {\em least-commitment} action model that is able to {\em explain} that input observation.
+Given an input sequence of partially observed states, this paper formalizes the task of computing a compact representation of the set of all action models that are consistent with that input observation.
+That is the {\em least-commitment} action model.
 
-This task is of interest because, if we can keep the set of all the action models that are consistent with the observation, i.e.
-the {\em version space}~\cite{mitchell1982generalization}, then each new observation has either no effect on the {\em version space} or eliminates some models from it.
-This allows to implement an incremental prunning of the {\em version space} and hence, to learn planning action models from arbitrary large sets of partial observations (one has never to go back and re-process old observations).
+This task is of interest because, if we can compute the {\em least-commitment} action model, then a new input observation has either no effect on the {\em least-commitment} action model or eliminates some models from it.
+This property allows to implement an incremental algorithm for learning planning action models from arbitrary large sets of partial observations (one has never to go back and re-process old observations)~\cite{mitchell1982generalization}.
 
@@
 A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
-Each classical planning action $a\in A$ has a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $c(a)$.
+Each classical planning action $a\in A$ has a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $cost(a)$.
 The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
@@
 A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\in\mathcal{L}(F)$ is the set of goal conditions over the state variables.
-A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length}.
+A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length} and $cost(\pi)=\sum_{a\in\pi} cost(a)$ its {\em plan cost}.
 The execution of $\pi$ on the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
@@
 
-An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
+An {\em action with conditional effects} $a_c\in A$ is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
 Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
@@
 Given a classical planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,s_0)$, we define the \emph{observation of the trajectory} as a sequence of partial states that results from observing the execution of $\pi$ on $I$.
-Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$ where $s_0^o=I$.
+Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$ where $s_0^o=I$ and $m$ ranges from 1 (the initial state, at least) to $|\pi|+1$.
+The observed intermediate states are {\em partial states}.
 
-A {\em partial state} $s_i^o$, {\small $0<i<m$}, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+A {\em partial state} $s_i^o$, {\small $0<i<m$} resulting from observing the corresponding state $s_i$, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
 Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
 Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be {\em consistent} with the sequence of states of $\tau(\pi,s_0)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-In practice, the number of observed states $m$, ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
 
@@
 \subsection{Conformant planning}
-{\em Conformant planning} is planning with incomplete information about the initial state, no sensing, and validating that goals are achieved with certainty (despite the uncertainty of the initial state)~\cite{goldman1996expressive,smith1998conformant,bonet2000planning}.
+{\em Conformant planning} is planning with incomplete information about the initial state, no sensing, and validating that goals are achieved with certainty (despite there is uncertainty at the initial state and no sensing is available)~\cite{goldman1996expressive,smith1998conformant,bonet2000planning}.
 
@@
 \begin{itemize}
-\item $\Phi=\tup{F,A[\cdot]}$ is a {\em classical planning frame} where the semantics of each action $a\in A[\cdot]$ is unknown; i.e.
-the corresponding $\tup{\rho,\theta}$ functions are undefined.
-We say that an {\em action model} $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$.
+\item $\Phi=\tup{F,A[\cdot]}$ is a {\em classical planning frame} where the semantics of each action $a\in A[\cdot]$ is unknown; i.e. the corresponding $\tup{\rho,\theta}$ functions are undefined. 
 \item $\mathcal{O}(\tau)$ is a sequence of partial states that results from the partial observation of a trajectory $\tau(\pi,s_0)$ within the {\em classical planning frame} $\Phi$.
@@
 
+\begin{definition}[Action model]
+An {\em action model} $\mathcal{M}$ is a definition of the $\tup{\rho,\theta}$ functions of every action in $A[\cdot]$. 
+\end{definition}
+
 Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and an observation $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$ let $P_\mathcal{O}$ be the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$.
-\begin{definition}[Model explanation]
-A model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}(\tau)$.
-\end{definition}
+A model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}(\tau)$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are defined by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}(\tau)$.
 
@@
 
+\subsection{The space of \strips\ action models}
 This work focuses on the particular task of computing the {\em least-commitment} action model when action models are specified as \strips\ action schemata.
 
-\subsection{The space of \strips\ action models}
 {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
-The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
-We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
-For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contain eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
+
+For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contains five elements for a {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for a {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
@@
 In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
 
 \subsection{{\em Partially specified} \strips\ action models}
-A set of action models can be defined {\em explicitly}, enumerating all the models that belong to the set or {\em implicitly}, enumerating the constraints that must satisfy any model that belongs to the set.
-Inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}, we introduce here {\em partially specified} \strips\ action models, a formalism for the {\em implicit} representation of a set of \strips\ schema.
+A set of action models can be defined {\em explicitly}, enumerating all the models that belong to the set or {\em implicitly}, enumerating all the constraints that must satisfy any model that belongs to the set.
+Inspired by the notion of {\em incomplete (annotated) model}~\cite{sreedharan2018handling}, we define here a {\em partially specified} \strips\ action model, a formalism for the {\em implicit} representation of a set of \strips\ schemes.
 
-We show now that extending to the {\em knowledge level} our propositional encoding of \strips\ action schemes, we can compactly represent a set of \strips\ schema.
+Extending our propositional encoding of \strips\ action schemes to the {\em knowledge level}, we can compactly represent a set of \strips\ schema.
 The extension defines, for each proposition {\tt\small $pre\_e\_\xi$}, two propositions {\tt\small $K\pre\_e\_\xi$} and {\tt\small $K\neg\pre\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is a precondition of $\xi$ and that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not a precondition of $\xi$.
-Likewise for each {\tt\small $eff\_e\_\xi$} proposition we define these two, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$}, meaning that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is an effect of $\xi$ and that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not an effect of $\xi$.
+Likewise {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} represent that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is an effect of $\xi$ and that is {\em known} that $e\in{\mathcal I}_{\Psi,\xi}$ is not an effect of $\xi$.
 
-\begin{definition}[Partially specified model]
-A {\em partially specified action schema} $\xi[\cdot]$ is an aritrary formula over the {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} propositions ($e\in{\mathcal I}_{\Psi,\xi}$) whose valuations express the enumeration of the set of action models $M_\xi$.
+\begin{definition}[Partially specified \strips\ model]
+A {\em partially specified action schema} $\xi[\cdot]$ is an aritrary formula over the {\tt\small $K\pre\_e\_\xi$}, {\tt\small $K\neg\pre\_e\_\xi$}, {\tt\small $K\eff\_e\_\xi$} and {\tt\small $K\neg\eff\_e\_\xi$} propositions ($e\in{\mathcal I}_{\Psi,\xi}$) whose valuations enumerate the set of action models $M_\xi$.
 \end{definition}
@@
 \begin{definition}[Partially specified model explanation]
-A partially specified action schema $\xi[\cdot]$ (that defines the set of action models $M_\xi$) {\em explains} an observation $\mathcal{O}(\tau)$, iff every model $\mathcal{M}\in M_{\xi}$ {\em explains} the input observation.
+A partially specified action schema $\xi[\cdot]$ {\em explains} a given observation $\mathcal{O}(\tau)$, iff every model $\mathcal{M}\in M_{\xi}$ {\em explains} that observation.
 \end{definition}
 
-\section{Incremental learning of action models from observations}
-This section sows how to implement an incremental learning of action models by prunning action models that are not consistent with the input observations.
+\section{Batch learning of action models}
+Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a batch of state observations $\mathcal{O}_1,\ldots,\mathcal{O}_T$, this section shows how to implement a model learning algorithm that incrementally eliminates the action models that are not {\em consistent} with the input observations.
+The algorithm is defined as follows.
+\begin{enumerate}
+\item Initialize a {\em partially specified} model to the full space of possible action models, $\xi[\cdot]:=M$.
+\item For each observation $\mathcal{O}_t$, {\tt\small $1\leq t\leq T$}:
+\begin{enumerate}
+\item Compute $M^*_t$, the {\em least-commitment} model that solves the corresponding $\tup{\Phi,\mathcal{O}_t}$ task.
+\item Update the {\em partially specified} model intersecting with the {\em least-commitment} model, $\xi[\cdot]:=\xi[\cdot]\cap M^*_t$.
+\item Continue until, the {\em partially specified} model is a singleton or no more observations left.
+\end{enumerate}
+\item Return $\xi[\cdot]$
+\end{enumerate}
+
+%Assuming that $\mathcal{O}_1,\ldots,\mathcal{O}_T$ are independent observations, and according to the PAC learning theory~\cite{kearns:PAC:1994}, we need to see at least $\frac{ln{(1/\delta)}+ln{|H|}}{\epsilon}$ observations to be sure that the prediction error is at most $\epsilon$ with probability $(1-\delta)$.
+In our case $|H|=\Pi_\xi2^{2\times|{\mathcal I}_{\Psi,\xi}|}$
 
@@
 
-\subsection{Learning algorithm}
-Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ where the semantics of each action $a\in A[\cdot]$ is unknown and a set of state observations $\mathcal{O}_1,\ldots,\mathcal{O}_T$.
-Our learning algorithm for the incremental learning of action models from state observations is defined as follows:
-\begin{enumerate}
-\item Build the {\em least-commitment} task $\tup{\Phi,\mathcal{O}_t}$ to process observation $\mathcal{O}_t$, {\tt\small $1\leq t\leq T$}.
-\item Build the classical planning task $P=\tup{F',A',I',G'}$ to solve the previous task.
-\item Extract the {\em partially specified} model from the previous solution.
-\item Continue until, the {\em version space} is a singleton or no more observations left
-\end{enumerate}
-
-Assuming and a set of independent observations $\{\mathcal{O}_1,ldots,\mathcal{O}_t\}$ we need to see at least $\frac{ln{1-\delta} + ln{}}{\epsilon}$ to be sure that the prediction error is at most $\epsilon$ with probability $(1-\delta)$.
-
 \section{Evaluation}

### 2018-12-25 doc/aij18/major-revision/experiments.tex

@@
 The drop in the score is caused by the increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for precision and recall is assigned to these experiments.
-Figure \ref{fig:pspace_time}, on the other hand, \textcolor[rgb]{1.00,0.00,0.50}{\textbf{reflects that the computation time of the first case study is also higher than in the second case},} which is explained by both the higher complexity and the large number of timeouts.
+Figure \ref{fig:pspace_time}, on the other hand, reflects that the computation time of the second case study is also higher than in the first case, which is explained by both the higher complexity and the large number of timeouts.
 
@@
 			\hline
-			Blocks & 0.84 & 0.59 & 0.89 & 0.64 \\
+			Blocks & 0.89 & 0.59 & 0.89 & 0.64 \\
 			Driverlog & 0.53 & 0.64 & 0.71 & 0.83 \\
@@
 			Parking & 0.8 & 1.0 & 0.98 & 1.0 \\
-			Rovers & 0.31 & 0.72 & 0.94 & 0.99 \\
+			Rovers & 0.48 & 0.85 & 0.94 & 0.99 \\
 			Satellite & 0.81 & 0.96 & 0.93 & 1.0 \\
 			Transport & 0.67 & 0.9 & 1.0 & 1.0 \\
-			Visitall & 0.57 & 1.0 & 1.0 & 1.0 \\
+			Visitall & 0.57 & 0.8 & 1.0 & 1.0 \\
 			Zenotravel & 0.73 & 0.68 & 0.85 & 0.88 \\
 			\hline
-			& 0.72 & 0.88 & 0.94 & 0.95
+			& 0.73 & 0.88 & 0.94 & 0.95
 		\end{tabular}
@@
 			Visitall & 1.0 & 0.2 & 1.0 & 1.0 \\
-			Zenotravel & 0.57 & 0.48 & - & - \\
+			Zenotravel & 0.3 & 0.25 & - & - \\
 			\hline

### 2018-12-25 doc/aij18/major-revision/letter.tex
last update Dec 25 13:37

@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{EVA ...}}
+\textcolor{blue}{The figures that appear in the third column of Table 3 for each learning system were taken from the respective published papers.
+They all refer to the specific experimental setup used in the evaluation of the approaches.
+In general, since there is not a common benchmark for the evaluation of learning approaches, each system selects the configuration of training data that best adapts to the particularities of the approach.
+Hence, all the figures in the 3rd column of Table 3 refer to the specific implementation used in the experiments.
+What we tried to do in this section is to somehow compare the approaches by establishing some common patterns of evaluation, namely number of tested domains, size of the training data in terms of action and/or traces, degree of observability, etc.}
 

### 2018-12-25 doc/aij18/major-revision/background.tex
update 24 Dec 20.00
Update Dec 25 21:20

@@
 
-\LAMP allows \PO state trajectories up to a minimum percentage of 1/5 of non-empty states as well as \POstar state trajectories with different degrees of observability in the number of propositions in each state.
+\LAMP allows \PO state trajectories up to a minimum observability of 1/5 of non-empty states as well as \POstar state trajectories with different degrees of observability in the number of propositions in each state.
 It uses an error metric based on counting the differences in the number of precondition and effects between the ground-truth model and the learned model.
@@
 The lack of available state information is overcome by exploiting assumptions about the structure of the actions.
-Particularly, \LOCM assumes that objects found in the same position in the header of actions are grouped as a collection of objects (sorts) whose defined set of states is captured by a parameterized Finite State Machine (FSM).
+Particularly, \LOCM assumes that objects found in the same position in the header of actions are grouped as a collection of objects \textcolor[rgb]{1.00,0.00,0.00}{named \emph{sort}} whose defined set of states is captured by a parameterized Finite State Machine (FSM).
 The intuitive assumptions of \LOCM, like the continuity of object transitions or the association of parameters between consecutive actions in the training dataset, yield a learning model heavily reliant on the kind of domain structure.
-A later work, \textbf{\LOCMtwo}, extends the applicability of the \LOCM algorithm to a wider range of domains by introducing a richer representation that allows using multiple FSMs to represent the state of a sort~\cite{cresswell2011generalised}.
+A later work, \textbf{\LOCMtwo}, extends the applicability of the \LOCM algorithm to a wider range of domains by introducing a richer representation that allows using multiple FSMs to represent the state of a \emph{sort}~\cite{cresswell2011generalised}.
 

### 2018-12-25 doc/aij18/major-revision/letter.tex
Update Dec 25 21:20

--- 2018-12-25 doc/aij18/major-revision/letter.tex
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{EVA ...}}
+\textcolor{blue}{The figures that appear in the third column of Table 3 for each learning system were taken from the respective published papers.
+They all refer to the specific experimental setup used in the evaluation of the approaches.
+In general, since there is not a common benchmark for the evaluation of learning approaches, each system selects the configuration of training data that best suits the particularities of the approach.
+Hence, all the figures in the 3rd column of Table 3 refer to the specific implementation used in the experiments.
+What we tried to do in this section is to somehow compare the approaches by establishing some common elements of evaluation, namely number of tested domains, size of the training data in terms of action and/or traces, degree of observability, etc.}
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{EVA ...}}
+\textcolor{blue}{A sort is a group of objects found in the same position in the header of an action.
+Although this was already defined in the text, we have changed the font of the word to italics and we explicitly say now this set is "named \emph{sort"}.}
 

### 2018-12-25 doc/aij18/major-revision/learning_action_models.tex
Update Dec 25 21:20

@@
 
-The logical inference process our approach is based on has trouble learning preconditions that are not, simultaneously, negative effects as well, since in this case no change can be observed between the pre-state and post-state of an action.
-This is specially the case for static predicates that never change and, hence, can only be preconditions.
-In order to address this shortcoming, we apply a post-process based on the one proposed in \cite{kuvcera2018louga}.
-In their proposal, they obtain a full trace by applying the sequence of actions of the input trace and infering the preconditions from this fully observable trace.
+The logical inference process our approach is based on has trouble learning preconditions that do not appear as negative effects since in this case no change is observed between the pre-state and post-state of an action.
+This is specially relevant for static predicates that never change and, hence, only appear as preconditions in the actions.
+In order to address this shortcoming \textcolor[rgb]{1.00,0.00,0.00}{and complete the list of learned preconditions}, we apply a post-process based on the one proposed in~\cite{kuvcera2018louga}.
+\textcolor[rgb]{1.00,0.00,0.00}{The idea lies in going through every action and counting the number of cases where a literal is present before the action is executed and the number of cases where it is not present.
+If a literal is present in all the cases before the action, the literal is considered to be a precondition.}
+
+\textcolor[rgb]{1.00,0.00,0.00}{In order to obtain a complete trace, the proposal in ~\cite{kuvcera2018louga} applies the sequence of actions of the input trace and infers the preconditions from this \FO action sequence}.
 In our case, since the sequence of actions of the input trace might not be fully observable, we produce the traces by applying the actions found in the validation part of the solution plan.
-For instance, in the example of the figure \ref{fig:plan-lplan}, the sequence of actions used to produce the full trace would be {\tt{\small(unstack blockB blockA)}}, {\tt{\small(put-down blockB)}}, {\tt{\small(pick-up blockA)}}, and {\tt{\small(stack blockA blockB)}}.
-This post-process allows \FAMA to learn more complete lists of preconditions and deal with the always problematic static predicates.
+For instance, in the example of the figure \ref{fig:plan-lplan}, the sequence of actions used to produce the complete trace would be {\tt{\small(unstack blockB blockA)}}, {\tt{\small(put-down blockB)}}, {\tt{\small(pick-up blockA)}}, and {\tt{\small(stack blockA blockB)}}.
+
+%This post-process allows \FAMA to learn more complete lists of preconditions and deal with the always problematic static predicates.
 

### 2018-12-25 doc/aij18/major-revision/experiments.tex
Update Dec 25 21:40

@@
 \begin{itemize}
-	\item \textbf{\FO action sequence and \PO state trajectory}: This is the common case addressed by most of the state-of-the-art learning approaches, which corresponds to a scenario where the plan horizon is given by the action sequence.
+	\item \textbf{\FO action sequence and \PO state trajectory (known plan horizon)}: This is the common case addressed by most of the state-of-the-art learning approaches, which corresponds to a scenario where the plan horizon is given by the action sequence.
 In this experiment we assume a degree of observability of only 10\% for the state trajectory, meaning that each literal of a state has a 10\% chance of being observed.
-	\item  \textbf{\NO action sequence and \NO state trajectory}: In this case study, both the input action sequence and state trajectory are {\em empty} and the length of the plan is unknown.
+	\item  \textbf{\NO action sequence and \NO state trajectory (unknown plan horizon)}: In this case study, both the input action sequence and state trajectory are {\em empty} and the length of the plan is unknown.
 Only the initial and final states are observed; i.e., $\tau = \tup{s_0, s_m}, \forall \tau \in \mathcal{T}$.
@@
 
-Figures \ref{fig:np_quality} and \ref{fig:np_time} show the quality of the models and computation time, respectively, for the first case study.
+Figures \ref{fig:np_quality} and \ref{fig:np_time} show the quality of the models and computation time, respectively, for the case \FO/\PO.
 \textcolor{red}{The values plotted in these figures are averages over the 15 domains.} In Figure \ref{fig:np_quality} we see that, after three traces, precision stabilizes at 0.84 whereas recall stabilizes at 0.95.
@@
 
-Table \ref{tab:results_minimum_100_10} shows the results of the first case study.
+Table \ref{tab:results_minimum_100_10} shows the results of the case study \FO/\PO.
 Recall scores are generally higher than the precision ones, and, in fact, the models learned for six out of the 15 domains were perfectly complete.
@@
 
-Table \ref{tab:results_minimum_30_30} gathers the results of the case study with 30\% observability.
+Table \ref{tab:results_minimum_30_30} gathers the results of the case study \PO/\PO with 30\% observability.
 We can see in the table that the scores of some domains are missing.
@@
 Regarding quality, we find that the learned models retain a level of soundness similar to Table~\ref{tab:results_minimum_100_10} but the completeness is lower than in the previous case study.
-This is specially noticeable in the preconditions, where recall values drop from from 0.88 to 0.64.
+This is specially noticeable in the preconditions, where recall values drop from 0.88 to 0.64.
 This is because the input actions act as strong constraints playing a key role on the closeness of the learned model to the GTM.
@@
 
-Table \ref{tab:metric_comparison_100_10} shows the results of the first case study.
+Table \ref{tab:metric_comparison_100_10} shows the results of the case study \FO/\PO.
 Looking at the high scores of the syntactic metrics, specially the value of recall, we can conclude that the learned models are, in fact, fairly similar to the GTM.
@@
 
-Table~\ref{tab:metric_comparison_0_0} details the results of the \NO/\NO case study.
+Table~\ref{tab:metric_comparison_0_0} details the results of the case study \NO/\NO.
 One first observation is the impossibility of applying a semantic evaluation in some of the most complex domains with five traces.

### 2018-12-26 doc/aij18/major-revision/aij18.tex
AIJ final ahora si que si

@@
 We highlight that the semantic evaluation of \FAMA is done via the same compilation-to-planning scheme that we use to learn the action models.
-Given that the input of our learning task definition accepts an initial action model of the agent's behaviour, either complete or partially specified, this solving scheme is exploitable to computing a model that follows the initial model and is compliant with a test set of plan traces.
+Given that the input of our learning task definition accepts an initial action model of the agent's behaviour, either complete or partially specified, this solving scheme is exploitable to computing a model that follows the initial model and is consistent with a test set of plan traces.
 In other words, \FAMA is applicable in model reconciliation tasks by defining a distance metric that measures how close the two models are~\cite{KulkarniCZVZK16}.

### 2018-12-26 doc/aij18/major-revision/background.tex
AIJ final ahora si que si

@@
 The Action Model Acquisition from Noisy plan traces (\textbf{\AMAN}) \cite{zhuo2013action} introduces an algorithm able to learn action models from plan traces with \NO state sequences where actions have a probability of being observed incorrectly (noisy actions).
-The first step of the \AMAN algorithm is to build the set of candidate domain models that are compliant with the action headers and predicates.
+The first step of the \AMAN algorithm is to build the set of candidate domain models that are consistent with the action headers and predicates.
 \AMAN then builds a graphical model to capture the domain physics; i.e., the relations between states, correct actions, observed actions and domain models.
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{The distinctive feature of the \LOCM family lies in the need to learn the state variables (fluents) because predicates are neither provided as input nor they are deducible from the plan trace as no state observability is allowed.
+\textcolor[rgb]{1.00,0.00,0.00}{The distinctive feature of the \LOCM family lies in the capacity to learn the state variables (fluents) because predicates are neither provided as input nor they are deducible from the plan trace as no state observability is allowed.
 In contrast, \FAMA and the rest of approaches either assume the set of predicates are provided alongside the input traces or assume they are extractable from the observed states of the plan trace, in which case the plan trace must comprise at least a grounded sample of every predicate.

### 2018-12-26 doc/aij18/major-revision/introduction.tex
AIJ final ahora si que si

--- 2018-12-24 doc/aij18/major-revision/introduction.tex
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{\FAMA is a new learning approach, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}.
+\textcolor{red}{\FAMA is a new learning approach, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}.
 The construction of a \strips\ planning model starts out from a set of plan traces containing the observation of several plan executions.
@@
 The compilation scheme lies in defining a planning task from the set of plan traces using a set of \emph{building actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces.
-Hence, a solution to this planning task is a plan that determines the preconditions and effects of the actions of a \strips\ planning model $M$.
-The rationale of this process draws its strength in that when the problem $\tup{s_0,s_f}$ is solved with the learned model $M$, a plan $\pi$ is generated such that $\pi$ will contain the observed actions of the plan trace, if any, and the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.
-In other words, the plan $\pi$ generated when solving the problem $\tup{s_0,s_f}$ with $M$ is \emph{compliant} with the plan trace.}
+Hence, a solution to this planning task is a plan that determines the preconditions and effects of the actions of a \strips\ planning model $M$ while ensuring consistency with the input traces.
+We say that a model $M$ is \emph{consistent} with a plan trace when $M$ can produce a solution $\pi$ to the planning problem $\tup{s_0,s_f}$ so that: (1) $\pi$ contains the observed actions of the plan trace, if any, and (2) the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.}
 
@@
 \begin{itemize}
-\item compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces
-\item the plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states
-\item the planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions
-\item a semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model
+\item compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces.
+\item the plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states.
+\item the planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions.
+\item a semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model.
 \end{itemize} }
 
-Ultimately, \FAMA adopts certain ML-like characteristics, like the ability to learn from data that do not explicitly exhibit the agent's behaviour or the capacity of correcting an initially given action model.
 A first description of the \FAMA compilation scheme already appeared in our previous conference paper~\cite{aineto2018learning}.
@@
 Section~\ref{sec:background} introduces classical planning concepts and reviews related work on learning planning action models.
-Section~\ref{sec:motivation} motivates our compilation-to-planning approach for learning action models.
-Section~\ref{sec:learning} formalizes the learning task and presents the compilation scheme, the core of \FAMA.
+\textcolor{red}{Section~\ref{task_definition} formally defines the learning task and motivates our compilation-to-planning approach for learning action models.} Section~\ref{sec:learning} presents the compilation scheme, the core of \FAMA.
 Sections~\ref{sec:evaluation} explains the evaluation of a learned model with respect to a reference model (syntactic evaluation) and with respect to a set of plan traces (semantic evaluation).

### 2018-12-26 doc/aij18/major-revision/learning_action_models.tex
AIJ final ahora si que si

@@
 \textcolor[rgb]{1.00,0.00,0.00}{Our proposal to address a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to transform $\Lambda$ into a planning task $P_{\Lambda}$.
-The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ complies with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$.
+The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$.
 Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition, a positive effect or a negative effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\tau$, including that the final reached state comprises $s_n$.
-Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the data generated through the action application must be validated with the data of $\tau$.}
+Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states of $\tau$.}
 
@@
 By the definition of the $\mathsf{apply_{\xi,\omega}}$ and the $\mathsf{validate_{j}}$ actions, these goals can only be achieved executing an applicable sequence of programmed action models that reaches every state $s_j\in\tau$, starting in the corresponding initial state and following the sequence of $n$ observed actions of $\tau$.
-This means that the programmed action model $\mathcal{M}'$ complies with the provided input knowledge and hence, that $\mathcal{M}'$ is a solution to $\Lambda$.
+This means that the programmed action model $\mathcal{M}'$ is consistent with the provided input knowledge and hence, that $\mathcal{M}'$ is a solution to $\Lambda$.
 \end{small}
@@
 	\item $\mathcal{M}$ is proved to be a {\em invalid} action model for the given input data $\tau$ iff $P_{\Lambda}$ is unsolvable.
-This means that $\mathcal{M}$ cannot be compliant with the given observation of the plan execution.
+This means that $\mathcal{M}$ cannot be consistent with the given observation of the plan execution.
 \end{itemize}

### 2018-12-26 doc/aij18/major-revision/learning_task_def.tex
AIJ final ahora si que si

@@
 
-A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a domain model $\mathcal{M}'$ that complies with the information of $\mathcal{M}$ and with the observed plan trace $\tau$.
+A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a domain model $\mathcal{M}'$ that is consistent with the information of $\mathcal{M}$ and with the observed plan trace $\tau$.
 \textcolor[rgb]{1.00,0.00,0.00}{This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan encompass the plan trace $\tau$.}
@@
 In this case, $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_{k}\}$ such that each $\tau\in \mathcal{T}$ is a plan trace that satisfies the previous 1--4 assumptions.
-In this case, the {\em learned} domain model $\mathcal{M}'$ must be compliant with the input model $\mathcal{M}$ and with every observed plan trace $\tau\in \mathcal{T}$.
+In this case, the {\em learned} domain model $\mathcal{M}'$ must be consistent with the input model $\mathcal{M}$ and with every observed plan trace $\tau\in \mathcal{T}$.
 
@@
 
-   \textcolor[rgb]{1.00,0.00,0.00}{ In this particular scenario, the actual number of plans compliant with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).
+   \textcolor[rgb]{1.00,0.00,0.00}{ In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).
 Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.}

### 2018-12-26 doc/aij18/major-revision/letter.tex
AIJ final ahora si que si

--- 2018-12-25 doc/aij18/major-revision/letter.tex
@@
 
-\textcolor{blue}{In the following, our responses to the reviewers are shown up in blue color.}
-
-\vspace{0.3cm}
-
-\textcolor{red}{Additionally, all the changes are highlighted in red in the document file.}
+\textcolor{red}{All the changes are highlighted in red in the document file.}
 
@@
 
-\textcolor{blue}{We agree with the reviewer that the paper lacks a description of the overall approach.
+We agree with the reviewer that the paper lacks a description of the overall approach.
 We refer the reviewer to the main modification \#1 exposed above.
@@
 The new section 3 defines the learning task and provides insight of the nature of the problem to solve according to the type of the observed plan trace in section 3.1.
-We believe that this explanation also helps to grasp the overall picture of \FAMA.}
+We believe that this explanation also helps to grasp the overall picture of \FAMA.
 
@@
 
-\textcolor{blue}{As per the reviewer's recommendation, we requested the code of LOUGA to the first author of the publication entitled \emph{LOUGA: learning planning operators using genetic algorithms}, which was published in the Pacific Rim Knowledge Acquisition Workshop.
+As per the reviewer's recommendation, we requested the code of LOUGA to the first author of the publication entitled \emph{LOUGA: learning planning operators using genetic algorithms}, which was published in the Pacific Rim Knowledge Acquisition Workshop.
 The author of LOUGA promptly provided us with the code, and so we thank him for his gentle response.
@@
 We evaluated LOUGA and \FAMA in the experiment with FO action sequences and PO state trajectories with 10\% of observability using 5 traces.
-The global precision and recall values for each domain are shown below as LOUGA/\FAMA: }
+The global precision and recall values for each domain are shown below as LOUGA/\FAMA:
 
@@
  transport &  0.70/0.79 &  0.8/0.95  \\
-\textcolor[rgb]{0.00,0.00,1.00}{AVERAGE}  & \textcolor{blue}{0.82/0.89} & \textcolor{blue}{0.90/0.97}
+AVERAGE  & 0.82/0.89 & 0.90/0.97
 \end{tabular}
@@
 
-\textcolor{blue}{The reviewer is right in that the definition of effects was confusing.
+The reviewer is right in that the definition of effects was confusing.
 It all stems from the adoption of the open world assumption, which leads to states that contain positive and negative literals, but we did not adapt the definition of effects appropriately.
-In section 2.1, we have now included a brief explanation of the open world assumption and the representation of states, and we have adapted the definition of effects accordingly.}
+In section 2.1, we have now included a brief explanation of the open world assumption and the representation of states, and we have adapted the definition of effects accordingly.
 
@@
 
-\textcolor{blue}{The additional assumptions are the syntactic constrains imposed by \strips\ planning models.
+The additional assumptions are the syntactic constrains imposed by \strips\ planning models.
 We agree with the reviewer that other domain models do not satisfy these assumptions.
 We intend to extend the compilation scheme to planning models other than \strips\ as future work, but the focus of this paper is exclusively on \strips\ models.
-We have included a sentence before the description of the assumptions to clearly state that they derive from the syntactic constraints of \strips.}
+We have included a sentence before the description of the assumptions to clearly state that they derive from the syntactic constraints of \strips.
 
@@
 
-\textcolor{blue}{We have reformulated the explanation about the observability in plan traces of section 2.1 to make it more comprehensible and added a few more details that help distinguish the different cases of observability.
+We have reformulated the explanation about the observability in plan traces of section 2.1 to make it more comprehensible and added a few more details that help distinguish the different cases of observability.
 We have also highlighted the two special cases, PO* and NO, that derive from PO.
@@
 The case FO actions with PO* states is a common case handled by some learning approaches as shown in Table 2 of section 2.2.
-In general, all learning approaches but \FAMA assume FO action sequences, which may come along with PO*, PO or NO state trajectories.}
+In general, all learning approaches but \FAMA assume FO action sequences, which may come along with PO*, PO or NO state trajectories.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{Done. LOUGA is included in Table 2 and Table 3 and an explanatory paragraph about the main characteristics of LOUGA has been incorporated in section 2.2.}
+Done. LOUGA is included in Table 2 and Table 3 and an explanatory paragraph about the main characteristics of LOUGA has been incorporated in section 2.2.
 
@@
 
-\textcolor{blue}{It depends on the domain we are trying to learn.
-Certainly, most planning domains do not need predicates with repeated variables, but in order to ensure the completeness of our learning approach we have opted to include it.}
+It depends on the domain we are trying to learn.
+Certainly, most planning domains do not need predicates with repeated variables, but in order to ensure the completeness of our learning approach we have opted to include it.
 
@@
 
-\textcolor{blue}{Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects.
+Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects.
 This is the reason why conditional effects are introduced in the compilation scheme since they only intervene in the specification of $P_{\Lambda}$ and do not condition the characteristics of the models that \FAMA is able to learn.
-Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}
+Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.
 
-\textcolor{blue}{On the other hand, the assumptions regarding the syntactic \strips\ constraints still apply when using conditional effects}
+On the other hand, the assumptions regarding the syntactic \strips\ constraints still apply when using conditional effects
 
@@
 
-\textcolor{blue}{We refer the reviewer to the main modification \#3.
+We refer the reviewer to the main modification \#3.
 Section 4 has been polished significantly.
-An introductory paragraph provides a high level description of the approach and examples have been incorporated throughout the text.}
+An introductory paragraph provides a high level description of the approach and examples have been incorporated throughout the text.
 
@@
 
-\textcolor{blue}{Removing symmetries would admittedly improve performance for heuristic planners.
+Removing symmetries would admittedly improve performance for heuristic planners.
 This removal can be achieved by ensuring the actions are programmed following some order, which can be done via some additional fluents.
-We think, however, that the compilation scheme should remain as simple and general as possible to facilitate its understanding.}
+We think, however, that the compilation scheme should remain as simple and general as possible to facilitate its understanding.
 
-\textcolor{blue}{Nevertheless, we want to thank the reviewer for the great idea and we will look into adding this optimization to the approach as a compilation option.}
+Nevertheless, we want to thank the reviewer for the great idea and we will look into adding this optimization to the approach as a compilation option.
 
@@
 
-\textcolor{blue}{The implications can actually be encoded as a disjunction, as it is the case of the {\tt{\small apply\_stack}} action of Figure 4.
+The implications can actually be encoded as a disjunction, as it is the case of the {\tt{\small apply\_stack}} action of Figure 4.
 The caption of the figure indicates that "implications are coded as disjunctions".
-The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$.}
+The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$.
 
@@
 
-\textcolor{blue}{The positions of the validate actions among the apply actions is determined by the planner, which will verify that the state that results after the execution of an apply action comprises the observed state of the plan trace.
+The positions of the validate actions among the apply actions is determined by the planner, which will verify that the state that results after the execution of an apply action comprises the observed state of the plan trace.
 We have added the following sentence in the document to clarify this point: "There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$.
-The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$."}
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$."
 
@@
 
-\textcolor{blue}{All plotted values are averages over the 14 (now 15) tested domains.
-The reviewer is right, however, to point out that this is not said in the experiment of section 6.2, so we have added some lines making it explicit.}
+All plotted values are averages over the 14 (now 15) tested domains.
+The reviewer is right, however, to point out that this is not said in the experiment of section 6.2, so we have added some lines making it explicit.
 
@@
 
-\textcolor{blue}{Following the reviewer's advice, we have completely remodeled section 3 and we have introduced major improvements in section 4.
+Following the reviewer's advice, we have completely remodeled section 3 and we have introduced major improvements in section 4.
 We refer the reviewer to our main modifications \#2 and \#3.
 We removed the complexity table of page 8, focusing now the discussion around the two main cases that determine the difficulty of the learning task and giving an indication of the task complexity according to the characteristics that define these two cases.
-We also changed the reference and added instead the original work that proves that boolean SAT problems are NP-complete.}
+We also changed the reference and added instead the original work that proves that boolean SAT problems are NP-complete.
 
-\textcolor{blue}{The take-home lesson of section 3 is that when the observed plan trace does not provide information of the solution plan length (case 2), planning is needed to find the solution and so the learning task is PSPACE-complete.}
+The take-home lesson of section 3 is that when the observed plan trace does not provide information of the solution plan length (case 2), planning is needed to find the solution and so the learning task is PSPACE-complete.
 
@@
 
-\textcolor{blue}{We have removed repetitions as much as possible.
+We have removed repetitions as much as possible.
 In the first example pointed by the reviewer, the explanation about partial observability of \FAMA and the other learning systems is concentrated in a single paragraph and the repetition is eliminated.
-Nevertheless, we think that at some points it is necessary to recall a particular concept in order to enhance the exposition of a further concept.}
+Nevertheless, we think that at some points it is necessary to recall a particular concept in order to enhance the exposition of a further concept.
 
-\textcolor{blue}{With respect to the repetition of the definition of a \strips\ action model, conditional effects are introduced in section 4 because these are exclusively used for modeling the task that results from the compilation, $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$.
-We have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}
+With respect to the repetition of the definition of a \strips\ action model, conditional effects are introduced in section 4 because these are exclusively used for modeling the task that results from the compilation, $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$.
+We have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.
 
-\textcolor{blue}{With respect to the third observation, we have introduced several changes in section 5.
+With respect to the third observation, we have introduced several changes in section 5.
 First, we provide the standard definitions of precision and recall at the beginning of the section.
@@
 4 and 5, respectively.
-Subsequently, in section 5.2, we present our novel semantic-based metrics that build upon definitions 4 and 5.}
+Subsequently, in section 5.2, we present our novel semantic-based metrics that build upon definitions 4 and 5.
 
@@
 
-\textcolor{blue}{Timeouts are only produced in the setting with NO action sequences and NO state trajectories.
+Timeouts are only produced in the setting with NO action sequences and NO state trajectories.
 We have added some details on the domains where timeouts are found: in the grid domain when $|\mathcal{T}| = 3$, floortile when $|\mathcal{T}| = 4$ and four other domains.
-We refer the reviewer to the comment added in page 23 of section 6.2.}
+We refer the reviewer to the comment added in page 23 of section 6.2.
 
@@
 
-\textcolor{blue}{The reviewer is totally right in that planning is needed when the length of the solution plan is unknown.
+The reviewer is totally right in that planning is needed when the length of the solution plan is unknown.
 Precisely, this is the point that motivates our approach but that we did not successfully convey in the previous version of the paper.
@@
 When the length of the plan is known, SAT-solvers can be used; when it is not known, planning is required.
-In our case, we use a planner to solve all learning tasks independently of their nature.}
+In our case, we use a planner to solve all learning tasks independently of their nature.
 
-\textcolor{blue}{On the other hand, we happen to use a SAT-based planner (Madagascar) for solving the learning tasks, which may lead to wrongly interpret that the length of the solution plan is always known.
+On the other hand, we happen to use a SAT-based planner (Madagascar) for solving the learning tasks, which may lead to wrongly interpret that the length of the solution plan is always known.
 The reasons for using Madagascar have been highlighted in red at the beginning of section 6.
 In a nutshell, we use Madagascar because it returned the best results compared to other planners and because of its ability to handle problems with dead-ends.
-Additionally, Madagascar is obviously very suitable to solve learning tasks when the length of the plan is known.}
+Additionally, Madagascar is obviously very suitable to solve learning tasks when the length of the plan is known.
 
@@
 
-\textcolor{blue}{We acknowledge the reviewer's concern and refer the reviewer to the main modification \#1 exposed at the beginning of this response letter.}
+We acknowledge the reviewer's concern and refer the reviewer to the main modification \#1 exposed at the beginning of this response letter.
 \vspace{0.1cm}
@@
 
-\textcolor{blue}{We have replaced the word 'abstract' by 'approximation' and rephrased the sentence a little bit.}
-\vspace{0.1cm}
+We have replaced the word 'abstract' by 'approximation' and rephrased the sentence a little bit.
 
@@
 
-\textcolor{blue}{We think this paragraph was rather confusing and that we did not clearly come through our intention to set an analogy between the data sources used in explainable AI and the data required by the learning approaches.
+We think this paragraph was rather confusing and that we did not clearly come through our intention to set an analogy between the data sources used in explainable AI and the data required by the learning approaches.
 For this reason, we have removed the paragraph and included instead a list of planning-related tasks that require a planning model for being solved.
-We think this rather complete but non-exhaustive list is a pretty good indication of the need for learning planning models.}
+We think this rather complete but non-exhaustive list is a pretty good indication of the need for learning planning models.
 
@@
 
-\textcolor{blue}{We have rephrased all the paragraphs and sentences pointed by the reviewer.
-With respect to the first comment, we attempt to make it clear that classical planning techniques can be used to generate a large variety of planning models, not only \strips\ models, and that these techniques underpin some relevant concepts that are also reusable in learning models when the behaviour of the agent is not available.}
+We have rephrased all the paragraphs and sentences pointed by the reviewer.
+With respect to the first comment, we attempt to make it clear that classical planning techniques can be used to generate a large variety of planning models, not only \strips\ models, and that these techniques underpin some relevant concepts that are also reusable in learning models when the behaviour of the agent is not available.
 
-\textcolor{blue}{With respect to the other two comments, besides using different expressions to refer to the two concepts, we have included some additional side explanations that we hope help understand the meaning of the sentences.}
-\vspace{0.1cm}
+With respect to the other two comments, besides using different expressions to refer to the two concepts, we have included some additional side explanations that we hope help understand the meaning of the sentences.
 
@@
 
-\textcolor{blue}{We have included an explanation of the open world assumption and how this affects the representation of states in Section 2.1.}
+We have included an explanation of the open world assumption and how this affects the representation of states in Section 2.1.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{Fixed.}
+Fixed.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{The sentence was, indeed, unclear.
-Anyway, we removed it in the revised version of the paper since it was an expendable phrase that did not fit very well in this section either.}
+The sentence was, indeed, unclear. Anyway, we removed it in the revised version of the paper since it was an expendable phrase that did not fit very well in this section either.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{We have moved the 4th paragraph of the Related Work, where we commented on the predicates and action headers of the approaches, to the end of the section since it makes more sense to comment on this issue when comparing with LOCM.
-Also, we clarified the meaning of action header (action name and parameters) because the formal definition of an action header is actually introduced in section 4.}
+We have moved the 4th paragraph of the Related Work, where we commented on the predicates and action headers of the approaches, to the end of the section since it makes more sense to comment on this issue when comparing with LOCM.
+Also, we clarified the meaning of action header (action name and parameters) because the formal definition of an action header is actually introduced in section 4.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{The labels FO, PO*, PO and NO concern exclusively the intermediate states of the state trajectories as indicated in Table 1.
+The labels FO, PO*, PO and NO concern exclusively the intermediate states of the state trajectories as indicated in Table 1.
 We have improved the explanation of the four types of state trajectories in section 2.1.
@@
 We changed the label of LOCM in the first column of Table 2 to "---" to indicate that no state information at all is needed.
-We have also rephrased the sentence about LOCM.}
+We have also rephrased the sentence about LOCM.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{The reviewer is right in that \FAMA also assumes that predicates and action headers are provided or otherwise are extractable from the plan traces.
-We believe it is more appropriate to move this comment after LOCM is introduced since LOCM is an exception that learns the predicates using Finite-State Machines (FSM) because no observed states are given in the plan trace.}
+The reviewer is right in that \FAMA also assumes that predicates and action headers are provided or otherwise are extractable from the plan traces.
+We believe it is more appropriate to move this comment after LOCM is introduced since LOCM is an exception that learns the predicates using Finite-State Machines (FSM) because no observed states are given in the plan trace.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{EVA ...}}
+The figures that appear in the third column of Table 3 for each learning system were taken from the respective published papers.
+They all refer to the specific experimental setup used in the evaluation of the approaches.
+In general, since there is not a common benchmark for the evaluation of learning approaches, each system selects the configuration of training data that best suits the particularities of the approach.
+Hence, all the figures in the 3rd column of Table 3 refer to the specific implementation used in the experiments.
+What we tried to do in this section is to somehow compare the approaches by establishing some common elements of evaluation, namely number of tested domains, size of the training data in terms of action and/or traces, degree of observability, etc.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{EVA ...}}
+A sort is a group of objects found in the same position in the header of an action.
+Although this was already defined in the text, we have changed the font of the word to italics and we explicitly say now this set is "named \emph{sort"}.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{We refer the reviewer to the main modification \#2. Section 3 has substantially changed and this sentence does no longer appear in the text.}
+We refer the reviewer to the main modification \#2. Section 3 has substantially changed and this sentence does no longer appear in the text.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{Corrected.}
+Corrected.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{It is true that a machinery for capturing and labeling the executed actions can be deployed.
+It is true that a machinery for capturing and labeling the executed actions can be deployed.
 In either case, whether a human or an automated mechanism, the need of a fully observable behaviour of the agent imposes a limitation in learning systems.
-We have soften the sentence including the word "\emph{commonly}" in our exposition.}
+We have soften the sentence including the word "\emph{commonly}" in our exposition.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{SERGIO ...}}
+We agree with the reviewer that relevant results have recently been achieved in the setting of full observability.
+Now we acknowledge, in Section 3.1.
+of our paper, the work by Stern and Juba that allows computing the least number of examples that are necessary to learn models within a given error rate.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{We have removed the paragraph regarding the evaluation of the learned models from section 3 and moved to section 5.
-We also clarify the use of the word "semantic" as requested by another reviewer as well.}
+We have removed the paragraph regarding the evaluation of the learned models from section 3 and moved to section 5.
+We also clarify the use of the word "semantic" as requested by another reviewer as well.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{In the remodeling of section 3, the paragraph about the size of the training dataset has been removed as well.
-The ideas presented here have been moved to page 23, at the end of section 6.2.}
+In the remodeling of section 3, the paragraph about the size of the training dataset has been removed as well. The ideas presented here have been moved to page 23, at the end of section 6.2.
 
@@
 
-\textcolor{blue}{The function $ar(\cdot)$ does not return a set, it returns the arity of an action or predicate; i.e., the length of its argument list.
-This is introduced in the Background section for the first time.}
+The function $ar(\cdot)$ does not return a set, it returns the arity of an action or predicate; i.e., the length of its argument list. This is introduced in the Background section for the first time.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.50}{\textbf{SERGIO ...}}
+The size of the space of possible \strips\ models for a given $\xi$ is $2^{2|\Psi_{\xi}|}$ because the value of two bits is enough to code how a given element from $\Psi_{\xi}$ appears in a given action model $\xi$: (0,0) indicates it does not appear in $\xi$, (1,0) indicates it appears as precondition, (1,1) indicates it appears as precondition and a negative effect and (0,1) indicates it appears as a positive effect.
 
@@
 
-\textcolor{blue}{Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects.
+Our approach is aimed at learning \strips\ models for classical planning problems $P=\langle F,A,I,G \rangle$ but $P_{\Lambda}=\langle F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda} \rangle$, the task that results from the compilation, is modeled as a planning problem with conditional effects.
 This is the reason why conditional effects are introduced in the compilation scheme since they only intervene in the specification of $P_{\Lambda}$ and do not condition the characteristics of the models that \FAMA is able to learn.
-Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.}
+Nevertheless, we have now incorporated the explanation about conditional effects within the section 4.2 'Compilation' since this is a merely syntactic aspect of the task formalization.
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{We have introduced a new paragraph at the beginning of section 4 that gives an overview of the compilation approach and explains the two blocks of actions of the solution plan to the learning task.
-We have reused the concept of "block of actions" to refer to the prefix and postfix of $\pi_{\Lambda}$ in section 4.2 "Compilation".}
+We have introduced a new paragraph at the beginning of section 4 that gives an overview of the compilation approach and explains the two blocks of actions of the solution plan to the learning task.
+We have reused the concept of "block of actions" to refer to the prefix and postfix of $\pi_{\Lambda}$ in section 4.2 "Compilation".
 
@@
 
-\textcolor[rgb]{0.00,0.07,1.00}{$\Rightarrow$ is the general logic symbol read as implication.
+$\Rightarrow$ is the general logic symbol read as implication.
 The implications can actually be encoded as a disjunction, as it is the case of the {\tt{\small apply\_stack}} action of Figure 4.
 The caption of the figure indicates that "implications are coded as disjunctions".
-The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$ }
+The implication of the formula can then be encoded as $\{\neg pre_{p}(\xi) \vee p(\omega)\}_{\forall p\in\Psi_\xi}$
 
@@
 
-\textcolor{blue}{We have added a new sentence that overviews the post-process.
-We have also rephrases some part of the text to make it more comprehensible.
-We believe the text provides now a clear idea of the post-process and that the pseudo-code is not necessary.
-Nevertheless, we refer the reviewer to the pseudo-code presented in [40] (algorithm 2) for a more detailed explanation.}
+We have added a new sentence that overviews the post-process.
+We have also rephrased some part of the text to make it more comprehensible.
+We believe the text provides now a clear idea of the post-process.
+Nevertheless, we refer the reviewer to the pseudo-code presented in [40] (algorithm 2) for a more detailed explanation.
 
@@
 
-\textcolor{blue}{We have provided better justification for the use of semantic evaluation metrics in section 5.}
+We have provided better justification for the use of semantic evaluation metrics in section 5.
 
@@
 
-\textcolor{blue}{The precision and recall metrics give a notion of the soundness and completeness of a model.
+The precision and recall metrics give a notion of the soundness and completeness of a model.
 By a sound model we understand one where all preconditions and effects are correct, either wrt to the GTM or a test set, so there is no need to remove anything.
 A complete model is one where no precondition or effect is missing.
-We have added a sentence to clarify this.}
+We have added a sentence to clarify this.
 
@@
 
-\textcolor{blue}{Fixed.}
+Fixed.
 
@@
 Pages 15-19.
+
 -  "...
@@
 You know the length of the given traces, but how do you know how many actions are needed to build the action model?
-\end{mdframed}
 
-\textcolor{blue}{Hopefully the new section 3 (see main modification \#2) helps clarify this question.
-In a nutshell, sometimes a plan trace indicates the length of the plan (this is the case when SAT solvers can be used) and others it does not (this is the case when planning is required).}
-
-\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
 -  "... a solution plan is solvable in two steps" - which two steps? I thought you put it all in the planner and get both action model and validation in the same step.
@@
 
-\textcolor{blue}{Yes, the reviewer is right in that we put it all in the planner and get the overall plan, both building actions and validating actions.
+Yes, the reviewer is right in that we put it all in the planner and get the overall plan, both building actions and validating actions.
 What we mean by "solvable in two steps" is that a SAT-based planner internally solves the construction of the model in two \emph{time steps}: in one time step the truth value of all the model fluents that represent the preconditions is set, and in a second time step the value of all the model fluents that represent the effects is set.
-We have slightly modified the sentence to make it clear we refer to the SAT solving process of Madagascar.}
+
+%We have slightly modified the sentence to make it clear we refer to the SAT solving process of Madagascar.
 
@@
 
-\textcolor{blue}{Yes, precision gives a notion of soundness since it measures how many of the learned preconditions and effects are correct.
-We have improved the explanation of the metrics in section 5 to make this more clear.}
+Yes, precision gives a notion of soundness since it measures how many of the learned preconditions and effects are correct.
+We have improved the explanation of the metrics in section 5 to make this more clear.
 
@@
 
-\textcolor{blue}{We have replaced the word "complete" by "adequate" as suggested by the reviewer.}
+We have replaced the word "complete" by "adequate" as suggested by the reviewer.
 
@@
 
-\textcolor{blue}{We agree with the reviewer's observation, and we believe the misunderstanding stems from us mixing up the words \emph{information} and \emph{observability}.
+We agree with the reviewer's observation, and we believe the misunderstanding stems from us mixing up the words \emph{information} and \emph{observability}.
 In our claim of \FAMA using the minimal amount of information, we should actually say that \FAMA uses the \emph{minimal observability} possible, being the extreme case to have no actions at all and only the initial state and final state as observations.
-In this respect, LOCM is indeed the learning approach that requires the \emph{minimal information} since no state or predicate is needed.}
+In this respect, LOCM is indeed the learning approach that requires the \emph{minimal information} since no state or predicate is needed.
+We want to point out, however, that \FAMA does not require the number of steps that are in the training plan, and we hope the new section 3 (see main modification \#2) does a better job at conveying this point.
 
-\textcolor{blue}{We think that LOCM is a learning approach that notably differs from the rest of systems presented in Table 2 and, in this regard, we have highlighted this distinction in the explanation of section 2.2.
+We think that LOCM is a learning approach that notably differs from the rest of systems presented in Table 2 and, in this regard, we have highlighted this distinction in the explanation of section 2.2.
 Besides adding a sentence that stresses the ability of LOCM to learning with minimal input knowledge, we devote the last paragraph of section 2.2.
-to explain the distinctive features of the LOCM family.}
-
-\vspace{0.1cm}
+to explain the distinctive features of the LOCM family.
 
@@
 
-\textcolor{blue}{The syntactic metrics commonly used by other approaches count the number of differences wrt the GTM and normalize by the total number of possible preconditions and effects.
+The syntactic metrics commonly used by other approaches count the number of differences wrt the GTM and normalize by the total number of possible preconditions and effects.
 This gives rise to an optimistic measure of the error since the set of possible preconditions and effects is usually much larger than the size of the actual GTM, resulting in low error rates even when the model is largely flawed.
@@
 Furthermore, precision and recall provide an intuitive notion of soundness and completeness, respectively; whereas a sound model is interpreted as a model in which all preconditions and effects are correct wrt the GTM, a complete model is one in which no precondition or effect is missing.
-We have added some paragraphs to section 5 to better explain these concepts.}
+We have added some paragraphs to section 5 to better explain these concepts.
 
-\textcolor{blue}{As the reviewer points out, \FAMA will create a model that explains all the training traces because it is a satisfying approach (that satisfies all the training samples) and so the training error is 0.
+As the reviewer points out, \FAMA will create a model that explains all the training traces because it is a satisfying approach (that satisfies all the training samples) and so the training error is 0.
 There is, however, no guarantee that the training traces are able to capture the whole dynamics of the domain.
 This is specially true for scenarios with low observability like the ones explored in this paper.
-Hence, it is always interesting to measure how well the learned model is able to explain a new set of plan traces.}
+Hence, it is always interesting to measure how well the learned model is able to explain a new set of plan traces.
 
@@
 
-\textcolor{blue}{We have substantially changed the explanation of the compilation scheme.
+We have substantially changed the explanation of the compilation scheme.
 We refer the reviewer to the main modification \#3.
 Besides an introductory paragraph that provides an overview of the compilation, we have included examples throughout the text to clarify the role of the insert actions, the apply actions and the validate actions.
-We have also renamed some concepts with more illustrative names.}
+We have also renamed some concepts with more illustrative names.
+One of the concepts we have renamed is the \emph{set of action models} $M$, which is now named \emph{domain model} to make it clear that \FAMA outputs a single model.
 
@@
 
-\textcolor{blue}{The semantic evaluation is used to measure the ability of a learned model to produce a given plan trace.
+The semantic evaluation is used to measure the ability of a learned model to produce a given plan trace.
 In this context, the use of the word "semantic" refers to the physics of the domain, not the actual meaning of the predicates or actions.
-We have added a comment clarifying this point.}
+We have added a comment clarifying this point.
 
-\textcolor{blue}{We agree with the reviewers concerns about the definitions of INS, DEL and size(M) and have corrected them.}
+We agree with the reviewers concerns about the definitions of INS, DEL and size(M) and have corrected them.
 
-\textcolor{blue}{Following the reviewer recommendation, we have splitted section 5.1 into two sections: one for the syntactic evaluation and another for the semantic evaluation}
+Following the reviewer recommendation, we have splitted section 5.1 into two sections: one for the syntactic evaluation and another for the semantic evaluation.
 
@@
 
-\textcolor{blue}{We changed the title as recommended by the reviewer.}
-
-\vspace{0.1cm}
+We changed the title as recommended by the reviewer.
 
@@
 
-\textcolor{blue}{We have introduced a paragraph at the beginning of section 6 that gives an overview of the experiments presented in this section.
+We have introduced a paragraph at the beginning of section 6 that gives an overview of the experiments presented in this section.
 The quality of the models (\emph{success} in the words of the reviewer) of the three first experiments is measured by the syntactic metrics introduced in the evaluation section; that is, the quality of the models is compared against the corresponding GTM.
@@
 With respect to moving text from the subsections of section 6 to section 5, we think the text of these subsections is particularly contextualized to the experiment in question and seems better to leave the text in section 6.
-Nevertheless, we have improved some explanations in section 6.}
+Nevertheless, we have improved some explanations in section 6.
 
@@
 
-\textcolor{blue}{We keep adding new domains as we work on the approach so this number is representative of how many domains we have tested our approach on.
+We keep adding new domains as we work on the approach so this number is representative of how many domains we have tested our approach on.
 We did, however, leave one domain out, {\em Sokoban}.
 As a matter of fact, this domain presents some of the most difficult features to be able to be handled by our approach, particularly features intrinsic in puzzle-like domains and domains with grid-like structure.
-We have though included one more domain, the {\em rovers} domain.}
+We have though included one more domain, the {\em rovers} domain.
 
@@
 
-\textcolor{blue}{Predicates with arity 3 are certainly harder but still manageable. In order to prove this, we have added the {\em rovers} domain to the list of domains tested in section 6. }
+Predicates with arity 3 are certainly harder but still manageable. In order to prove this, we have added the {\em rovers} domain to the list of domains tested in section 6.
 
@@
 \end{mdframed}
+
+In the NO / NO case study, the number of action steps is not known. The changes introduced in main modification \#3 should make the distinction between known and unknown horizon more explicit.
 

### 2019-01-24 doc/ijcai-19-goalRecognition/goalRecognition.tex
adding old content to the ijcai19 paper on goal recognition

--- 2018-11-19 doc/icaps-19/model-recognition.tex
@@
-\def\year{2017}\relax
-%File: formatting-instruction.tex
-\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
-\usepackage{aaai19}  %Required
-\usepackage{times}  %Required
-\usepackage{helvet}  %Required
-\usepackage{courier}  %Required
-\usepackage{url}  %Required
-\usepackage{graphicx}  %Required
-\frenchspacing  %Required
-\setlength{\pdfpagewidth}{8.5in}  %Required
-\setlength{\pdfpageheight}{11in}  %Required
-%PDF Info Is Required:
-  \pdfinfo{}
-\setcounter{secnumdepth}{0}
+%%%% ijcai19.tex
 
+\typeout{IJCAI-19 Instructions for Authors}
+
+% These are the instructions for authors for IJCAI-19.
+
+\documentclass{article}
+\pdfpagewidth=8.5in
+\pdfpageheight=11in
+% The file ijcai19.sty is NOT the same than previous years'
+\usepackage{ijcai19}
+
+% Use the postscript times font!
+\usepackage{times}
+\usepackage{soul}
+\usepackage{url}
+\usepackage[hidelinks]{hyperref}
+\usepackage[utf8]{inputenc}
+\usepackage[small]{caption}
+\usepackage{graphicx}
 \usepackage{amsmath}
-\usepackage{amssymb}
-\usepackage{amsthm}
-\usepackage{multirow}
-\usepackage{tikz}
-\usetikzlibrary{arrows,automata}
-\usepackage{comment}
+\usepackage{booktabs}
+\usepackage{algorithm}
+\usepackage{algorithmic}
+\urlstyle{same}
 
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{listings}
-
+%%%%%%%%%%%%%%%%%% Added for this paper
+\usepackage{listings}% http://ctan.org/pkg/listings
 \lstset{
@@
 }
-
-\usepackage{multicol}
-\usepackage{arydshln}
-\usetikzlibrary{calc,backgrounds,positioning,fit}
-
+\usepackage{ wasysym }
 \newcommand{\tup}[1]{{\langle #1 \rangle}}
-
 \newcommand{\pre}{\mathsf{pre}}     % precondition
@@
 
+%%%%%%%%%%%%%
+
+% the following package is optional:
+%\usepackage{latexsym}
+
+% Following comment is from ijcai97-submit.tex:
+% The preparation of these files was supported by Schlumberger Palo Alto
+% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
+% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
+% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
+% preparation.
+
+% These instructions can be modified and used in other conferences as long
+% as credit to the authors and supporting agencies is retained, this notice
+% is not changed, and further modification or reuse is not restricted.
+% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
+% contacts for providing assistance without their prior permission.
+
+% To use for other conferences, change references to files and the
+% conference appropriate and use other authors, contacts, publishers, and
+% organizations.
+% Also change the deadline and address for returning papers and the length and
+% page charge instructions.
+% Put where the files are available in the appropriate places.
+
+\title{Model-Based Goal Recognition with Unknown Domain Models}
+
+% Single author syntax
+%\author{
+%    Sarit Kraus
+%    \affiliations
+%    Department of Computer Science, Bar-Ilan University, Israel \emails
+%    pcchair@ijcai19.org
+%}
+
+% Multiple author syntax (remove the single-author syntax above and the \iffalse ... \fi here)
+% Check the ijcai19-multiauthor.tex file for detailed instructions
+\author{
+Diego Aineto$^1$\and
+Sergio Jim\'enez$^1$\and
+Eva Onaindia$^1$\And
+\and
+Miquel Ram\'irez$^2$
+\affiliations
+$^1${\small Departamento de Sistemas Inform\'aticos y Computaci\'on. Universitat Polit\`ecnica de Val\`encia. Valencia, Spain}\\
+$^2${\small School of Computing and Information Systems. The University of Melbourne. Melbourne, Victoria. Australia}
+\emails
+{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es, miquel.ramirez@unimelb.edu.au}}
+
 \begin{document}
+\maketitle
 
-\title{Model Recognition as Planning}
-
-% Commented for blind submission
-\author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
-{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
-{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
-{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
-\And Miquel Ram\'irez\\
-{\scriptsize School of Computing and Information Systems}\\
-{\scriptsize The University of Melbourne}\\
-{\scriptsize Melbourne, Victoria. Australia}\\
-{\scriptsize miquel.ramirez@nimelb.edu.au}}
-
-\maketitle
 \begin{abstract}
-Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action schemata), {\em model recognition} is the task of identifying the model that explains the observation.
-The paper formalizes this task and introduces a novel method that estimates the probability of a \strips\ model to produce an observation of a plan execution.
-This method builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and intermediate states in the observation.
-The effectiveness of the method is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a classical string classification task; (2) identification of the model that encodes a failure present in an observation; and (3) recognition of a robot navigation policy.
+The paper shows how to relax one key assumption of the {\em plan recognition as planning} for {\em goal recognition} that is knowing the action model of the observed agent.
+The paper introduces a novel formulation that fits together the {\em learning of planning action models} with the {\em plan recognition as planning} approach.
+The empirical evaluation evidences that our novel formulation allows to solve standard goal recognition benchmarks without having knowing the action model of the observed agent.
 \end{abstract}
@@
 \label{sec:introduction}
+{\em Goal recognition} is a particular classification task in which each class represents a different goal and each example is an observation of an agent acting to achieve one of that goals.
+Despite there is a wide range of diffenrent approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most popular since it is at the core of several interesting tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
 
-A growing interest in addressing the task of recognizing the mission (plans and goals) of a planning agent from a set of given observations has recently emerged.
-Approaches based on rule-based systems, parsing, graph-covering or Bayesian nets have been proposed for the plan/goal recognition tasks~\cite{sukthankar14}.
-{\em Plan recognition as planning} is the model-based approach for this task that leverages the action model of the observed agent to compute its most likely goal and predict its future actions~\cite{ramirez2009plan,ramirez2012plan}.
-While action models are learnable from observations of the agent~\cite{yang2007learning,cresswell2013acquiring,aineto2018learning}, real-world applications developing in a particular context usually dispose of approximate behavioral models of the involved agents and there is rarely a need to learn a model from scratch.
-A more relevant task in these scenarios is thereby to predict the behavior of an agent from a set of given models.
-
-This paper formalizes the novel task of {\em model recognition}, where the object to recognize is not a goal or a plan but the model that shapes the behavior of the observed agent.
-Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action schemata), {\em model recognition} is the task of identifying the model in the set with the highest probability of producing the given observation.
-We posit this problem as a \emph{classification task} where classes are given as generative planning models and samples are sequences of observations.
-We accomplish this task with off-the-shelf classical planning algorithms that compute for each model (class) $\mathcal{M}$ in the set a model $\mathcal{M'}$ that explains the observation, and subsequently distances between the two models are measured.
-Two key contributions of our {\em model recognition as planning} approach are: (1) the ability of handling partial observations of plan executions that contain no actions and incomplete -- even empty- intermediate states; and (2) a thorough formalization that induces a probability distribution of a model $\mathcal{M}$ to explain an observed plan execution.
-
-\begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-  \begin{tikzpicture}[scale=.4]
-          \begin{scope}
-            \draw (0, 0) grid (5, 5);
-            \node[anchor=center] at (2.5, 1.5) {R};
-            \node[anchor=center] at (0.5, 0.5) {};
-            \draw[thick,style=dotted] (0.5,0.5) -- (3.5,0.5);
-            \draw[thick,style=dotted] (3.5,0.5) -- (3.5,1.5);
-            \draw[thick,style=dotted,->] (3.5,1.5) -- (2.7,1.5);
-          \end{scope}
-        \end{tikzpicture}
-\hspace*{1cm}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-	  \node[state] (A)              {$q_0$};
-	  \node[state] (B) [right of=A] {$q_1$};
-	  \path
-(A) edge [bend left, align=center] node {inc-y,\\dec-y} (B)
-(B) edge [bend left, align=center] node {inc-y,\\dec-y} (A);
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}
- \caption{\small ({\em Left)} Robot navigating a $5\times 5$ grid.
-{\em (Right)} Automata to control that the robot only increments its x-coordinate when $q_0$ holds (actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and switch the automata state).}
-\label{fig:grid-example}
-\end{figure}
-
-To better illustrate {\em model recognition}, imagine a robot in a $N\times N$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model, the robot can increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
-at even rows if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at odd rows if {\tt\small $q0$} holds initially).
-Besides this model, other models that shape different kinds of robot navigation can be defined within the same state variables (e.g.
-altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated).
-Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where state variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, {\em model recognition} aims to identify the robot action schemata that explains this observation.
-
-We provide a formalization of {\em model recognition} that is applicable to any planning model but we restrict our attention to \strips\texttt{-}compilable models.
-It is well-known that diverse automata representations like finite state controllers, {\sc GOLOG} programs or reactive policies can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,segovia2017generating}.
-As for the evaluation, the effectiveness of our approach is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a proof-of-concept classification task; (2) a simulation of a non-deterministic \emph{blocksworld} under different degrees of observability; and (3) a navigation task where observations contain an unbounded number of unobservable states.
-
-%: a classical string classification task; identification of the model that encodes a failure present in an observation; and recognition of a robot navigation policy.
-
-%The paper also introduces {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
-Our method is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.
-
-\begin{figure}
-  \begin{tiny}
-  \begin{verbatim}
-  (:action inc-x
-    :parameters (?v1 ?v2)
-    :precondition (and (xcoord ?v1) (next ?v1 ?v2) (q0))
-    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))
-
-  (:action dec-x
-    :parameters (?v1 ?v2)
-    :precondition (and (xcoord ?v1) (next ?v2 ?v1) (q1))
-    :effect (and (not (xcoord ?v1)) (xcoord ?v2)))
-
-  (:action inc-y-even
-    :parameters (?v1 ?v2)
-    :precondition (and (ycoord ?v1) (next ?v1 ?v2) (q0))
-    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
-                 (not (q0)) (q1)))
-
-  (:action inc-y-odd
-    :parameters (?v1 ?v2)
-    :precondition (and (ycoord ?v1) (next ?v1 ?v2) (q1))
-    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
-                 (not (q1)) (q0)))
-
-  (:action dec-y-even
-    :parameters (?v1 ?v2)
-    :precondition (and (ycoord ?v1) (next ?v2 ?v1) (q0))
-    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
-                 (not (q0)) (q1)))
-
-  (:action dec-y-odd
-    :parameters (?v1 ?v2)
-    :precondition (and (ycoord ?v1) (next ?v2 ?v1) (q1))
-    :effect (and (not (ycoord ?v1)) (ycoord ?v2)
-                 (not (q1)) (q0)))
-  \end{verbatim}
-  \end{tiny}
- \caption{\small Example of a \strips\ action schemata (given in PDDL) for robot navigation in a $N\times N$ grid.}
-\label{fig:model-example}
-\end{figure}
+{\em Plan recognition as planning} leverages the action model of the observed agent and an off-the-shelf classical planner to compute the most likely goal of the agent.
+In this paper we show that we can relax the key assumption of {\em plan recognition as planning} for {\em Goal recognition} that is knowing the action model of the observed agent.
+In particular, the paper introduces a novel formulation that fits together the {\em learning of planning action models} with the {\em plan recognition as planning} approach.
+The evaluation of our formulation evidences that it allows to solve goal recognition tasks, even when the action model of the observed is unknown.
 
@@
 \label{sec:background}
-This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
+This section formalizes the {\em planning model} we follow as well as the kind of {\em observations} that are given as classification examples for the {\em goal recognition} task.
 
 \subsection{Classical planning with conditional effects}
-$F$ is the set of {\em fluents} or {\em state variables} (propositional variables).
-A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
+Let $F$ be the set of {\em fluents} or {\em state variables} (propositional variables) describing a state.
+A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e.
 either~$l=f$ or $l=\neg f$.
-$L$ is a set of literals that represents a partial assignment of values to fluents, and $\mathcal{L}(F)$ is the set of all literals sets on $F$, i.e.~all partial assignments of values to fluents.
-A {\em state} $s$ is a full assignment of values to fluents.
-We explicitly include negative literals $\neg f$ in states and so $|s|=|F|$ and the size of the state space is $2^{|F|}$.
+A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values).
+Given $L$, let $\neg L=\{\neg l:l\in L\}$ be its complement.
+We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.
+A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$.
 
-%A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states; i.e.
-$|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Like in PDDL~\cite{fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
+Each classical planning action $a\in A$ has a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $cost(a)$.
+The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
+Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$, i.e.~if its precondition holds in $s$.
+The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
+Subtracting the complement of $\eff(a)$ from $s$ ensures that $\theta(s,a)$ remains a well-defined state.
+The subset of action effects that assign a positive value to a state fluent is called {\em positive effects} and denoted by $\eff^+(a)\in \eff(a)$ while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects} of an action $a\in A$.
 
-A {\em planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
-An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\in\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\in\mathcal{L}(F)$.
-The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
-Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$.
-And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\in\mathcal{L}(F)$ is the set of goal conditions over the state variables.
+A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length} and $cost(\pi)=\sum_{a\in\pi} cost(a)$ its {\em plan cost}.
+The execution of $\pi$ on the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
+A solution plan is {\em optimal} iff it minimizes the sum of action costs.
 
-A {\em planning problem} is defined as a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state in which all the fluents of $F$ are assigned a value true/false and $G$ is the goal set.
-A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, and $|\pi|=n$ denotes its {\em plan length}.
-The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
-A trajectory $\tau(\pi,P)$ that solves $P$ is one in which $G \subseteq s_n$.
-
-An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
+An {\em action with conditional effects} $a_c\in A$ is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
 Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\neg\eff_c(s,a)\cup\eff_c(s,a)\}$ where $\eff_c(s,a)$ are the {\em triggered effects} resulting from the action application (conditional effects whose conditions hold in $s$):
 \[
-triggered(s,a_c)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
+\eff_c(s,a)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
 \]
-
-The result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$, where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
 
@@
 
-Figure~\ref{fig:grid-example} illustrates a {\em partial observation} of a five-state trajectory \{{\tt\scriptsize<(xcoord 2)(ycoord 1)>, <(xcoord 3)(ycoord 1)>, <(xcoord 4)(ycoord 1)>, <(xcoord 4)(ycoord 2)>, <(xcoord 3)(ycoord 2)>}\}.
-Note that this observation only contains fluents of the type {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, namely {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is not observable in any of the five states.
+\subsection{Model-Based Goal Recognition}
+We formalize the {\em goal recognition} task following the standard  formalization of Ram\'irez and Geffner for plan recognition~\cite{ramirez2012plan}.
 
-\section{Model Recognition}
-\label{sec:recognition}
-The {\em model recognition} task is a tuple $\tup{P,M,\mathcal{O}}$ where:
+The {\em model-based goal recognition} is a classification task defined by a tuple $\tup{P,\mathcal{O}}$, where:
 \begin{itemize} 
-\item $P=\tup{F,A[\cdot],I,G}$ is a planning problem where $A[\cdot]$ is a set of actions.
-For each $a\in A[\cdot]$, the semantics of $a$ is unknown; i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined.
-\item $M=\{\mathcal{M}_1,\ldots,\mathcal{M}_k\}$ is a set of {\em k different planning models} for the actions in $A[\cdot]$.
-A model $\mathcal{M}\in M$ defines the semantics of every action in $A[\cdot]$.
-Planning models differ in the $\tup{\rho,\theta}$ functions of the actions but they all use the same set of state variables $F$.
+\item $P=\tup{F,A,I,G[\cdot]}$ is a planning problem where $G[\cdot]$ is a set of possible goals.
 \item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
@@
 
-Model recognition can be understood as a {\em classification task} where each class is represented by a different planning model $\mathcal{M}\in M$, and the observed plan execution $\mathcal{O}(\tau)$ is the single example to classify.
-The planning model associated to each class acts as the corresponding {\em class prototype} and it summarizes any observation of a plan execution that could be synthesized with such model (i.e.
-the set of all the examples that belong to that class).
-We follow the {\em naive Bayes classifier} to assign a model $\mathcal{M}\in M$ to a given observation $\mathcal{O}(\tau)$.
-The {\em solution} to the model recognition task is then the subset of models in $M$ that maximizes this expression.
+Following the {\em naive Bayes classifier}, the {\em solution} to the {\em model-based goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
 
 \begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{O}|\mathcal{M}) P(\mathcal{M}).
+argmax_{g\in G[\cdot]} P(\mathcal{O}|g) P(g).
 \end{align}
 
-\subsection{Formulating the $P(\mathcal{O}|\mathcal{M})$ likelihood}
+The {\em Plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelyhood using an off-the-shelf classical planner.
+Recent works show that faster, but less accurate estimates, of this $P(\mathcal{O}|g)$ likelyhood can also be computed using relaxations of the classical planning task.
 
-The $P(\mathcal{M})$ probability expresses whether one model is known to be a priori more likely than the others.
-If this probability is not given as input it is reasonable to assume that, {\em a priori}, all models are equiprobable.
-The challenge of our formulation to the model recognition task is the definition of $P(\mathcal{O}|\mathcal{M})$, the likelihood that expresses the probability of observing $\mathcal{O}(\tau)$ when $\mathcal{M}$ is the planning model.
-
-\begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-	  \node[state] (A)              {$\mathcal{M}$};
-	  \node[state] (B) [right of=A] {$\mathcal{M}'$};
-	  \node[state] (C) [right of=B] {$\tau$};
-	  \node[state] (D) [right of=C] {$\mathcal{O}$};
-	  \path
-(A) edge  node {} (B)
-(B) edge  node {} (C)
-(C) edge  node {} (D)
-;
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}
- \caption{\small {\em Bayesian network} representing that model $\mathcal{M}$ is {\em transformable} into a model $\mathcal{M}'$ that produces a trajectory $\tau(\pi,P)$ that (1) reaches the goals in $P$ and (2) it is consistent with $\mathcal{O}(\tau)$.}
-\label{fig:net}
-\end{figure}
-
-Our approach to formulate the $P(\mathcal{O}|\mathcal{M})$ likelihood is to assess the cost of transforming $\mathcal{M}$ into a model $\mathcal{M'}$ that produces a trajectory $\tau(\pi,P)$ such that: (1) $\tau(\pi,P)$ reaches the goals in $P$ and (2) $\tau(\pi,P)$ is consistent with observation $\mathcal{O}(\tau)$.
-Figure~\ref{fig:net} shows the {\em Bayesian network} that embodies this procedure.
-Regarding this network we have the following formulation of the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}:
-
-\begin{align}
- P(\mathcal{O}|\mathcal{M})=\sum_{\mathcal{M}'}\sum_{\tau} P(\mathcal{M'}|\mathcal{M})P(\tau|\mathcal{M'})P(\mathcal{O}|\tau),
-\end{align}
-
-where $\tau$ ranges over all the trajectories consistent with $\mathcal{O}(\tau)$ that can be synthesized with a model $\mathcal{M}'$ and $\mathcal{M}'$ ranges over all the models that can be generated {\em transforming} $\mathcal{M}$.
-
-The exact computation of $P(\mathcal{O}|\mathcal{M})$ with equation (2) is intractable.
-For most planning problems the set of trajectories consistent with an arbitrary observation can easily be huge, infinite in the case of planning problems without dead-ends~\cite{lesh1995sound}.
-Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially with the number of state variables.
-Instead, our approach is to estimate of $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} referred to \strips\ planning models.
-{\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, which have been proved successful for {\em pattern recognition}~\cite{MasekP80,Bunke97}.
-In this work, we assess the cost of {\em transforming} $\mathcal{M}$ into $\mathcal{M'}$ computing {\em edit distances} between \strips\ planning models.
-
-\section{Recognition of \strips\ planning models}
-\label{sec:asPlanning}
-
-Our formalization of a model recognition task is valid for any planning model but our \emph{edit distance} measure is exclusively for \strips\ models and so we restrict our attention to \strips\texttt{-}compilable planning models\footnote{An edit distance for other planning models is also definable}.
-Thus, we focus on the recognition of $\mathcal{M}\in M$, where $M$ is a set of \strips\ planning models and $\mathcal{M}$ defines the semantics of the actions through a set of \strips\ action schemata.
+\section{Model-Based Goal Recognition with Unknown Domain Models}
+\label{sec:recognition}
 
 \subsection{Well-defined \strips\ action schemata}
-\strips\ action schemata provide a compact representation for specifying classical planning models.
-Figure~\ref{fig:model-example} shows six \strips\ action schemata that shape a particular kind of robot navigation in $N\times N$ grids (no matter the grid size).
-
-{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-
-\begin{definition}[Comparable \strips\ action schemata]
-Two \strips\ schemata $\xi$ and $\xi'$ are {\bf comparable} iff $pars(\xi)=pars(\xi')$, i.e, both share the same list of parameters.\footnote{In \strips\ models, $pars(\xi)=pars(\xi')$ implies the number of parameters must be the same.
-For other planning models that allow object typing, the equality implies that parameters share the same type}
-\end{definition}
-
-For instance, we claim that the six action schemata of Figure~\ref{fig:model-example} are {\em comparable} while, for example, the {\small\tt stack(?v1,?v2)} and {\small\tt pickup(?v1)} schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are not.
-Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
+\strips\ action schemata provide a compact representation for specifying classical planning models.{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 
 Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
-The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
-We denote this set of FOL interpretations as ${\mathcal I}_{\Psi,\xi}$.
-For any of the six action schemata of Figure~\ref{fig:model-example}, the ${\mathcal I}_{\Psi,\xi}$ set contains the same ten elements, ${\mathcal I}_{\Psi,\xi}=${\small\tt\{xcoord($v_1$), xcoord($v_2$), ycoord($v_1$), ycoord($v_2$), q0(), q1(), next($v_1$,$v_1$), next($v_1$,$v_2$), next($v_2$,$v_1$), next($v_2$,$v_2$))\}}.
+The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
+For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contains five elements for a {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for a {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is constrained by a set ${\mathcal C}$ that includes:
-
-\begin{itemize}
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is bounded by a set of constraints ${\mathcal C}$ of three kinds:
+\begin{enumerate}
 \item {\em Syntactic constraints}.
@@
 Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-For every action schema in the navigation model of Figure~\ref{fig:model-example} then $2^{2\times 10}=1,048,576$.
-
+{\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
 \item {\em Domain-specific constraints}.
 One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in a {\em robot navigation} model like the one in Figure~\ref{fig:model-example}, {\small\tt q0()} and {\small\tt q1()} are exclusive so they cannot hold at the same time in a $pre(\xi)$/$del(\xi)$/$add(\xi)$ list.
-Further, {\small\tt next($v_1$,$v_1$)} and {\small\tt next($v_2$,$v_2$)} will not appear in any of these lists because the {\tt\small next} predicate codes the {\em successor} function for {\em natural numbers}.
-These domain-specific constraints reduce further the size of the space of possible action schemata to $2^{2\times 7}=16,384$ (for every schema in the navigation model of Figure~\ref{fig:model-example}).
-\end{itemize}
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
+{\it State invariants} are also constraints of this kind~\cite{fox1998automatic}.
+\item {\em Observation constraints}. An observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
+\end{enumerate}
 
@@
 
-We can now formalize an {\em edit distance} that quantifies how similar two given \strips\ models are.
+Second, let us define when to action models are comparable.
+For instance, we claim that the {\small\tt stack(?v1,?v2)} and {\small\tt unstack(?v1,?v2)} actions schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are comparable while, the {\small\tt stack(?v1,?v2)} and {\small\tt pick-up(?v1)} schemata are not.
+Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
+
+\begin{definition}[Comparable \strips\ action schemata]
+Two \strips\ schemata $\xi$ and $\xi'$ are {\bf comparable} iff $pars(\xi)=pars(\xi')$, i.e, both share the same list of parameters.\footnote{In \strips\ models, $pars(\xi)=pars(\xi')$ implies the number of parameters must be the same.
+For other planning models that allow object typing, the equality implies that parameters share the same type}
+\end{definition}
+
+We are now ready to formalize an {\em edit distance} that quantifies how similar two given \strips\ models are.
 The distance is symmetric and meets the {\em metric axioms} provided that the two edit operations, {\em deletion} and {\em insertion}, have the same positive cost.
@@
 
-\subsection{Approximating the $P(\mathcal{O}|\mathcal{M})$ {\em likelihood}}
-
-Now we we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ action schemata.
-
-\subsubsection{Full observability of the executed plan.} The {\em full observability of the executed plan} is a too strong assumption for {\em model recognition} but it allows us to understand how to build a reasonable estimate of $P(\mathcal{O}|\mathcal{M})$ for the general case.
-
-Under the assumption of {\em full observability}, there is only a single possible trajectory $\tau^*(\pi,P)$ {\em consistent} with the input observation so $P(\mathcal{O}|\tau^*)=1$.
-Further, there is also a single model that can exactly produce that trajectory (otherwise models are identical, at least, in the actions relevant to the observed trajectory so they can be considered the same model).
-
-Provided that there is a single possible trajectory and a single possible model consistent with the input observation, i.e.
-$M^*=\{\mathcal{M}^*\}$ then, the probabilities of expression (2) are not added up and expression (1) simplifies to:
-\begin{align}
-argmax_{\mathcal{M}\in M} P(\mathcal{M^*}|\mathcal{M}) P(\mathcal{M}).
-\end{align}
-Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because is independent of the input model $\mathcal{M}\in M$.
-
-\subsubsection{Partial observability of the executed plan.} In a similar way, an approximation to $P(\mathcal{O}|\mathcal{M})$ can be built for the general case, where the executed plan is partially observed.
-We add the following two assumptions to deal with this general case:
-
-\begin{enumerate}
-\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state~\cite{Dennett83}.
-\item {\em The best plans for $P$ are unique or have different cost}.
-\end{enumerate}
-
-Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}.
-They allow us to assume that the sum in equation (2) is dominated by its largest term, so other terms in the sum are not added up.
-The largest term corresponds here to the shortest trajectory $\tau^*$ consistent with the observation (which now is assumed to be unique).
-Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more likely to assume that there is only one trajectory consistent with the input observation.
-
-Again, we have that there is a single model $\mathcal{M}^*$ that can produce a given trajectory $\tau^*$.
-With this said, both $P(O|\tau^*)$ and $P(\tau^*|\mathcal{M}^*)$ are independent of $\mathcal{M} \in M$ so equation (1) simplifies once again to equation (3) under the previous two assumptions.
-
-\subsubsection{The $P(\mathcal{M^*}|\mathcal{M})$ probability distribution.} $P(\mathcal{M'}|\mathcal{M})$ indicates the probability of {\em transforming} a classical planning model $\mathcal{M}$ into a model $\mathcal{M'}$ by exclusively using the two {\em edit operations} previously defined, {\em deletion} and {\em insertion}.
-
-We know from {\em pattern recognition}~\cite{devroye2013probabilistic} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}.
-Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
-\begin{align}
-CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d (1-p)^{N-d}
-\end{align}
-where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than do not applying it.
-
-With this regard, and considering that \strips\ models $\mathcal{M}\in M$ can be encoded with a propositional representation of fixed length $N$, we formulate the $P(\mathcal{M'}|\mathcal{M})$ probability distribution mapping the distance $\delta(\mathcal{M},\mathcal{M'})$ according to equation:
-\begin{align}
-P(\mathcal{M'}|\mathcal{M}) = p^d  (1-p)^{N-d}
-\end{align}
-where $d=\delta(\mathcal{M},\mathcal{M}')$.
-Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation (5) but instead, the parameter $d$ is given by the {\em observation distance}, $d=\delta^o(\mathcal{M},\mathcal{O})$.
-
-In other words, we are modeling the edition of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.
-
-\section{Model Recognition as planning}
-
-This section shows that $\delta^o(\mathcal{M},\mathcal{O})$, and hence an approximation to $P(\mathcal{O}|\mathcal{M})$), can be computed with a compilation-to-planning approach as the one proposed in~\cite{aineto2018learning} (AJO approach hereafter) for learning \strips\ models.
-The AJO approach receives as input an empty model $\mathcal{M}$, which only contains the headers of the action schemata formed of $\xi=\tup{name(\xi),pars(\xi)}$, and an observation of a plan execution $\mathcal{O}(\tau)$ (extensible to a set of observations), and it returns a model $\mathcal{M'}$ with specification of preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$ following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1},a_i)$ for every observed action of $\mathcal{O}(\tau)$ and $s_i=\theta(s_{i-1},a_i)$ for every observed state of $\mathcal{O}(\tau)$.
-
-Essentially, the task $\tup{\mathcal{M},\mathcal{O}}$ of the AJO approach -- learning $\mathcal{M}$ that satisfies $\mathcal{O}$-- is interpreted as editing the empty action schemata of $\mathcal{M}$ introducing preconditions and effects until $\mathcal{O}$ is validated with the resulting model.
-We leverage the same idea to compute $\delta^o(\mathcal{M},\mathcal{O})$ with the exception that now our $\mathcal{M}$ is not empty but $\mathcal{M}=\{\xi_1, \ldots, \xi_m\}$, where $\xi_i=\tup{name(\xi_i),pars(\xi_i),pre(\xi_i),add(\xi_i),del(\xi_i)}$, $1 \leq i \leq m$.
-
-Editing/learning the action schemata of $\mathcal{M}$ in the AJO approach is addressed converting the task into a classical planning problem, which is later solved with a planner.
-The intuition behind this compilation is that a solution plan to the problem is a sequence of: (a) \emph{edit actions} on the schemata of $\mathcal{M}$ to build $\mathcal{M'}$ and (b) \emph{validate actions} that apply $\mathcal{M'}$ in $\mathcal{O(\tau)}$.
-The adaptation of this compilation scheme for solving $\delta^o(\mathcal{M},\mathcal{O})$ results in a planning problem $P'=\tup{F',A',I',G'}$ whose objective is to determine the preconditions and effects that need to be added or deleted to the action schemata of $\mathcal{M}$ so as to satisfy $\mathcal{O}$.
-The accomplishment of this task requires therefore a propositional encoding of the components of the action schemata.
-Specifically:
-
-\begin{itemize}
-\item $F'$ contains the necessary fluents to represent the $pre$, $add$ and $del$ of the action schemata.
-It is a set of \emph{editable fluents} of the type $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
-Additionally, $F$ also contains fluents to encode the \emph{observation constraints}; that is, fluents to iterate through the $l$ observed actions and $m$ observed states of $\mathcal{O(\tau)}$.
-\item $A'$ comprises two types of actions with conditional effects:
-    \begin{itemize}
-    \item actions for editing $\xi \in \mathcal{M}$ that follow the \emph{syntactic constraints} for well-defined \strips\ action schemata.
-Hence, $A'$ contains actions for inserting a new precondition $pre\_e\_\xi$ or a new positive effect $add\_e\_\xi$ in $\xi$ when neither $pre\_e\_\xi$, $del\_e\_\xi$ or $add\_e\_\xi$ exist in $\xi$, and actions for inserting a new negative effect when neither $del\_e\_\xi$ or $add\_e\_\xi$ appear in $\xi$ but $pre\_e\_\xi$ does.
-    \item actions for applying the new action schemata of the edited model $\mathcal{M'}$ and validating the observed states of $\mathcal{O(\tau)}$.
-Particularly, the \emph{apply} actions check the preconditions and produce the effects defined by the \emph{editable fluents}.
-    \end{itemize}
-\item $I'$ encodes the editable fluents $pre\_e\_\xi, del\_e\_\xi$ and $add\_e\_\xi$ that hold in the action schemata of $\mathcal{M}$.
-\item $G'$ contains the necessary fluents to check that all the observed actions and states of $\mathcal{O(\tau)}$ are generated correctly following the edited model $\mathcal{M'}$.
-    %Here we include the necessary fluents to check the final reached state is consistent with the final state $s_m^o$ of $\mathcal{O(\tau)}$.
-\end{itemize}
-
-We now show an example of a solution plan to a planning problem $P'$ that results from compiling a specific task $\tup{\mathcal{M},\mathcal{O}}$.
-Consider that $\mathcal{M}$ is the model of Figure~\ref{fig:model-example} except that the schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped with respect to Figure~\ref{fig:model-example}.
-And consider an observation that only contains the initial and final state of Figure~\ref{fig:grid-example}; i.e., $\mathcal{O(\tau)}=\tup{s_0^o,s_m^o}=\tup{\{{\texttt{\small{(xcoord 0)(ycoord 0)}}}\}, \{{\texttt{\small{(xcoord 2)(ycoord 1)}}}\}}$.
-The plan found by a planner to $P'$ is shown in figure~\ref{fig:plan-pdistance}.
-The first seven steps are the edit actions to \emph{fix} the schema {\tt\small inc-x}; step 07 is a validate action that sets the robot in the initial state $s_0^o=\tup{\{{\texttt{\small{(xcoord 0)(ycoord 0)}}}\}}$; step 08 applies the action {\tt\small (inc-x 0 1)} and moves the robot one cell to the right to position $(1,0)$; steps 09-11 also move the robot one cell to the right; step 11 applies an action that increases coordinate $y$ from a position in an even row number, thus moving robot to row $1$; step 12 moves the robot one cell to the left and, finally, action {\tt\small (validate\_1)} checks the robot position is consistent with the final state $s_m^o=\tup{\{{\texttt{\small{(xcoord 2)(ycoord 1)}}}\}}$.
-
-\begin{figure}
-{\tt\tiny
-\begin{tabular}{ll}
-00 : (insert\_pre\_xcoord\_v1\_inc-x)   & 07 : (validate\_0)\\
-01 : (insert\_pre\_next\_v1\_v2\_inc-x) & 08 : (apply\_inc-x 0 1)\\
-02 : (insert\_pre\_q0\_inc-x)           & 09 : (apply\_inc-x 1 2)\\
-03 : (delete\_del\_xcoord\_v2\_inc-x)   & 10 : (apply\_inc-x 2 3) \\
-04 : (delete\_add\_xcoord\_v1\_inc-x)   & 11 : (apply\_inc-y-even 0 1)\\
-05 : (insert\_del\_xcoord\_v1\_inc-x)   & 12 : (apply\_dec-x 3 2)\\
-06 : (insert\_add\_xcoord\_v2\_inc-x)   & 13 : (validate\_1)
-\end{tabular}
-}
- \caption{\small A plan for $\tup{\mathcal{M},\mathcal{O}}$ where $\mathcal{M}$ is a modification of the model of Figure \ref{fig:model-example} and $\mathcal{O}$ a partial observation from Figure~\ref{fig:grid-example}.}
-\label{fig:plan-pdistance}
-\end{figure}
-
-The value of the {\em observation distance} $\delta^o(\mathcal{M},\mathcal{O})$ is given by the number of {\em edit operations} (insertions and deletions) required by $\mathcal{M}$ to be validated in the input observation.
-In the case of the above example, the distance equals $7$.
-
 \section{Evaluation}
@@
 
-In this section, we evaluate the empirical performance of our approach in three different applications of {\em model recognition}.
-We will assume in the three experiments that $\mathcal{O}(\tau)$ contains an \emph{empty sequence of observed actions} and so we will only work with the available observed states.
+\section{Conclusions}
+\label{sec:conclusions}
 
-For each experiment, we define a set $M$ of different planning models that share the same state variables but update the variables using different action schemata.
-For every $\mathcal{M}\in M$, we generated the same number of partial observations of plan trajectories ($\mathcal{O}(\tau)$) with such a model.
-Finally, we applied our {\em model recognition as planning} approach to identify the model, among the models in $M$, which was actually used for generating a given $\mathcal{O}(\tau)$ observation.
-We assume equiprobable prior probabilities in all the experiments; i.e., $P(\mathcal{M}) = \frac{1}{|M|}$ for every $\mathcal{M} \in M$.
-
-In order to provide a better explanation of the experiments below, we introduce a particular class of $\mathcal{O}(\tau)$ observations.
-This new class allows us to distinguish between {\em observable} state variables, whose value may be read from sensors, and {\em hidden} or {\em latent} state variables that cannot be observed.
-
-\begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
-\end{definition}
-
-Hence the value of the variables in $\Phi\subseteq F$ is observable while the value of the variables in $F\setminus\Phi$ is unobservable.
-
-\subsubsection{Reproducibility.} All experiments were run in an Intel Core i5 3.10GHz x 4 16GB of RAM and the classical planner we used to solve the instances that result from the compilation was {\sc Madagascar}~\cite{rintanen2014madagascar} because of its ability to deal with classical planning problems with dead-ends~\cite{lopez2015deterministic}.
-Other planners, such as FastDownward were also tested but provided worse experimental results.
-We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned the maximum distance obtained.
-The compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em https://github.com/anonsub/model-recognition} so any experimental data reported in the paper can be reproduced.
-
-\subsection{Recognition of {\em regular automata}.} The first experiment, which doubles as a proof of concept, exploits {\em model recognition as planning} for a classical {\em string classification} problem.
-In this experiment, the system receives (1) the string to classify (the $\mathcal{O}(\tau)$ observation) and (2) a set $M$ of different planning models, where each $\mathcal{M}\in M$ represents a different class; i.e.
-a different regular automata that accepts strings that belong to the class.
-
-Figure~\ref{fig:regautomatae} illustrates a 4-symbol and 5-state {\em regular automata} for recognizing the $(abcd)^+$ language.
-The {\em input alphabet} is $\Sigma=\{a,b,c,d\}$, and the machine states are $Q=\{q_0,q_1,q_2,q_3,\underline{q_4}\}$, where \underline{$q_4$} is the only acceptor state.
-For instance, executing the planning model that encodes the {\em regular automata} of Figure~\ref{fig:regautomatae} with the input string $abcdabcd$ produces the following 8-action plan {\small $(\langle a,q_0\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow q_3)$, $(\langle d,q_3\rangle\rightarrow \underline{q_4})$, $(\langle a,\underline{q_4}\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow \underline{q_3})$, $(\langle d,q_3\rangle\rightarrow \underline{q_4})$}.
-
-%\begin{figure}
-%  \begin{scriptsize}
-%  \begin{center}
-%	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-%	  \node[state] (A)              {$q_0$};
-%	  \node[state] (B) [right of=A] {$q_1$};
-%	  \node[state] (C) [right of=B] {$q_2$};
-%	  \node[state] (D) [right of=C] {\underline{$q_3$}};
-%	  \path
-%(A) edge  node {a} (B)
-%(B) edge  node {b} (C)
-%(C) edge  node {c} (D)
-%(D) edge [bend left]  node {a} (B)
-%;
-%	\end{tikzpicture}
-%  \end{center}
-%  \end{scriptsize}
-% \caption{\small A 3-symbol 4-state {\em regular automata} for recognizing the $(abc)^+$ language (\underline{$q_3$} is the acceptor state).}
-%\label{fig:regautomatae}
-%\end{figure}
-
-\begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1.8cm,semithick]
-	  \node[state] (A)              {$q_0$};
-	  \node[state] (B) [right of=A] {$q_1$};
-	  \node[state] (C) [right of=B] {$q_2$};
-	  \node[state] (D) [right of=C] {$q_3$};
-	  \node[state] (E) [right of=D] {\underline{$q_4$}};
-	  \path
-(A) edge  node {a} (B)
-(B) edge  node {b} (C)
-(C) edge  node {c} (D)
-(D) edge  node {d} (E)
-(E) edge [bend left]  node {a} (B)
-;
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}
- \caption{\small A 4-symbol and 5-state {\em regular automata} for recognizing the $(abcd)^+$ language (\underline{$q_4$} is the acceptor state).}
-\label{fig:regautomatae}
-\end{figure}
-
-%Note that the the string to classify is a {\em $\Phi$-observation} since it does not contain any information about the structure of the possible regular automata.
-This means that $\mathcal{O}(\tau)$ is a $\Phi$-observation where fluents representing the internal machine state are unknown and the applied transitions (actions) are unobserved.
-
-In this experiment, $M$ comprises five planning models, each representing a different 5-state and 4-symbol regular automata. The five regular languages defined by these automata are the following:
-
-\begin{itemize}
-	\item $\mathcal{L}$1: $a^+(b|c)d(dd)^*a$
-	\item $\mathcal{L}$2: $bd(abd)^*cd^*c^+$
-	\item $\mathcal{L}$3: $d^*c(ac)^*db^*ac^*b^+$
-	\item $\mathcal{L}$4: $(cc)^+bd(abd)^*$
-	\item $\mathcal{L}$5: $(d|a)a(ba)^*c^+d(dc^*d)^*$
-\end{itemize}
-
-For each regular language, we generated $20$ random strings, $\mathcal{O}(\tau)$ observations, which lengths range from 20 to 30 symbols, thus generating a total of 100 strings.
-More precisely, $\mathcal{O}(\tau)$ are observations in which the applied transitions (actions) are unobservable and the fluents of the automata internal state are also unobservable.
-Consequently, $\mathcal{O}(\tau)$ are $\Phi$-observations where the problem states are only represented with the fluents that specify the string to classify.
-
-%More precisely observations $\mathcal{O}(\tau)$ are $\Phi$-observations where: the applied automata transitions and the fluents representing the internal automata state are unobserved.
-
-Table \ref{tab:conf_matrix} shows the {\em confusion matrix} resulting from classifying the 100 strings with our method.
-In this matrix, rows represent the actual class and columns represent the class predicted by our {\em model recognition as planning} approach.
-Despite using a suboptimal classical planner, we can observe that the class for all the 100 input strings was correctly recognized, which proves the feasibility of our method for classification tasks.
-The outstanding results reveal that our approach is very suitable for the classification of generative planning models that are more restrictive than \strips\ models, as it is the case of regular automata.
-
-%These results prove the suitability of {\em model recognition as planning} for classification tasks where classes are given as generative planning models and examples are sequence of observations.
-
-\begin{table}
-	\centering
-	\begin{tabular}{c|c|c|c|c|c}
-		& $\mathcal{L}$1 & $\mathcal{L}$2 & $\mathcal{L}$3 & $\mathcal{L}$4 & $\mathcal{L}$5\\ \hline
-		$\mathcal{L}$1 & 20 & 0 & 0 & 0 & 0 \\
-		$\mathcal{L}$2 & 0 & 20 & 0 & 0 & 0 \\
-		$\mathcal{L}$3 & 0 & 0 & 20 & 0 & 0 \\
-		$\mathcal{L}$4 & 0 & 0 & 0 & 20 & 0 \\
-		$\mathcal{L}$5 & 0 & 0 & 0 & 0 & 20 \\
-	\end{tabular}
-	\caption{Confusion matrix for {\em regular automata recognition}.}
-	\label{tab:conf_matrix}
-\end{table}
-
-\subsection{Recognizing failures in a non-deterministic {\em blocksworld}}
-With this second experiment we aim to validate the two assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood: (1) that agents act rationally and so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation $\mathcal{O}(\tau)$ and (2) that there exists a single such trajectory.
-Our assumptions are more correct with higher observability and become tautological when the trajectory is fully observed ($\mathcal{O(\tau)} = \tau$).
-
-The input observations $\mathcal{O}(\tau)$ come from a non-deterministic {\em blocksworld} in which some of the executed actions may have failed. We considered failures of three kinds:
-
-\begin{itemize}
-	\item the execution of a {\tt\small stack} action fails and causes no effect
-	\item the execution of {\tt\small unstack} fails and causes no effect
-	\item {\tt\small unstack} fails and drops the block on the table
-\end{itemize}
-
-$M$ comprises three different 4-schema {\em blocksworld} models, each one extended with an additional action schema that encodes one of the three above failures.
-This means that failures are identified by finding the model that better explains the input observation.
-
-The validation of the assumptions is done by measuring the {\em  accuracy} of our approach across different degrees of observability.
-The accuracy metric is defined as the number of correct predictions over all predictions made, and the degree of observability indicates the probability of observing a fluent in an intermediate state of $\mathcal{O}(\tau)$.
-In this experiment, we also assume that the length of the trajectory is known.
-More precisely, at least one fluent of every state of $\mathcal{O}(\tau)$ is observed and the length of the plan trajectory is thus fixed by the observation.
-This allows us to produce a controlled experiment in which the number of possible trajectories able to explain an observation is bounded.
-
-% Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions.
-In order to show the assumptions validation:
-
-Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations.
-The positive trend seen in the figure proves that the more complete the observations are the more accurate our estimation becomes.
-
-%The overall high accuracy validates our assumption that the trajectory that better explains the observation is the shortest one.
-
-%The increasing trend of the accuracy as observability increases and so the number of possible explanation lowers, validates our two assumptions.
-
-%We limited the failures to be of one kind per observation however, it would be possible to recognize different combinations of them.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{figures/blocks.eps}
-	\caption{Accuracy for the recognition of failures in a non-deterministic {\em blocksworld}.}
-	\label{fig:blocks_acc}
-\end{figure}
-
-\subsection{Recognition of navigation policies}
-%This experiment is also aimed at validating the accuracy of our approach to identify the model that generates an observation under different degrees of observability.
-The novel aspect of this experiment is that, unlike the previous experiment, we do not guarantee that the observed states contain at least one fluent.
-This means that some intermediate states of $\mathcal{O}(\tau)$ may be missing (no fluents are observed) and, consequently, more than a single action may be required to reach an observed state from another one.
-
-In this experiment, the planning models of $M$ represent different navigation models for $N\times N$ grids.
-In particular, we adopt a {\em classical navigation model} with actions {\tt\small up, down, right, left} (to move one cell in each of these four directions) and extend it to define 8 different navigation policies with respect to these additional state variables {\tt\small \{(q0), (q1), (min ?v), (max ?v)\}}.
-An example of a navigation policy is shown in Figure~\ref{fig:model-example}.
-This policy allows to move right when $q_0$ holds while it allows to move left when $q_1$ holds, producing a zigzag pattern when visiting all the cells of $N\times N$ grids.
-
-Another interesting aspect of this experiment is that all navigation policies are at a maximum distance of four editions from the base {\em classical navigation model} that can move the robot in any direction at any given state.
-Thus, for any given model and observation, a maximum of four editions are needed for the input model to explain the observation.
-This aspect heavily constrains the discriminating power of $P(\mathcal{O}|\mathcal{M})$, which along with the low degree of observability of the states, gives rise to a meaningful benchmark for {\em model recognition as planning}.
-
-On the other hand, we generated observations from 8 different trajectories, one for each of the 8 previous planning models.
-The trajectories depict paths followed by a navigation policy to solve the planning problem of {\em visiting all} cells in a $5\times 5$ grid.
-As in the previous experiments, observations contain no actions and here, for each observed state, only the values of the fluents encoding the {\em x} and {\em y coordinates} of the agent are known (i.e.
-$\Phi =$\{{\small\tt xcoord, ycoord}\}).
-
-Figure \ref{fig:navigation_acc} shows the classification {\em accuracy} achieved by our approach with respect to a range of degrees of observability, from 0\% to 100\% with 10\% increments.
-In this experiment we included the 0\% case which corresponds to observations where only the initial and final states are observed.
-The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation policies, but from 10\% of observability onwards, we start to correctly classify half of the observations.
-Accuracy stabilizes after 40\% observability in the range 0.875 to 1 which means that at most only one out of the 8 observations was not correctly classified.
-
-\begin{figure}
-	\centering
-	\includegraphics[width=0.9\linewidth]{figures/navigation.eps}
-	\caption{Classification accuracy for the recognition of navigation models.}
-	\label{fig:navigation_acc}
-\end{figure}
-
-\section{Related work}
-\label{sec:related}
-An interesting research line related with {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}.
-In {\em model reconciliation} model edition is used to conform the PDDL models of two different agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed {\em optimal plan} computed using one of the models.
-{\em Model recognition} conforms every input model $\mathcal{M}$ in the $M$ set with another model $\mathcal{M}'$ that is consistent with a partial observation of a plan execution.
-Note, however, that in {\em model recognition} $\mathcal{M}'$ is not given as input but instead is computed for every $\mathcal{M}\in M$.
-Further, {\em model recognition as planning} shows that the probability of an input planning model to explain a partial observation of a plan execution (that contains no information of the applied actions and where intermediate states may be incomplete or even missing) can be estimated with an off-the-shelf classical planner.
-
-Once the planning model of the observed agent is recognized, then the model-based machinery for automated planning becomes fully applicable.
-This includes current approaches for {\em goal recognition}~\cite{ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} and {\em counter-planning}~\cite{PozancoEFB18} that require to have in advance a fixed model of the observed agent.
-Last but not least, our approach allows predicting the future actions of an agent whose behavior is not governed by a deliberative planning process aimed to achieve some goals.
-Our experiments show that {\em model recognition} is suitable for anticipating the actions of agents that are ruled by the execution of fixed programs or finite state machines, such as a {\em regular automata}.
-
-%Last but not least, previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
-Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
-
-\section{Conclusions}
-\label{sec:conclussions}
-This paper formalized the {\em model recognition} task and proposed, {\em model recognition as planning}, a method built on top of off-the-shelf classical planning algorithms to estimate the probability of a \strips\ model to produce a partial observation of a plan execution.
-The paper shows the effectiveness of {\em model recognition as planning} in a set of \strips\ models encoding different kinds of {\em automata}.
-{\em Model recognition as planning} succeeds to identify the executed automata despite the internal machine state or actual applied transitions, are unobserved.
-
-%Remarkably, the extension of this piece of work to the FOND planning setting~\cite{muise2012improved} is straightforward by simply considering the {\em all-outcomes} determiniztion of the actions with non-determinitic effects~\cite{yoon2007ff}.
-An interesting research direction is however to understand how to apply our approach to planning models where the planning models include actions with probabilistic effects~\cite{younes2005first}.
-
-\newpage
-
-\bibliographystyle{aaai}
+%% The file named.bst is a bibliography style file for BibTeX 0.99c
+\bibliographystyle{named}
 \bibliography{planlearnbibliography}

### 2019-01-24 doc/ijcai-19-goalRecognition/goalRecognition.tex
adding old content to the ijcai19 paper on goal recognition

@@
 
-\subsection{Model-Based Goal Recognition}
-We formalize the {\em goal recognition} task following the standard  formalization of Ram\'irez and Geffner for plan recognition~\cite{ramirez2012plan}.
+\section{Model-Based Goal Recognition with Unknown Domain Models}
+\label{sec:recognition}
 
-The {\em model-based goal recognition} is a classification task defined by a tuple $\tup{P,\mathcal{O}}$, where:
-\begin{itemize} 
-\item $P=\tup{F,A,I,G[\cdot]}$ is a planning problem where $G[\cdot]$ is a set of possible goals.
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves the planning problem $P$.
-\end{itemize}
-
-Following the {\em naive Bayes classifier}, the {\em solution} to the {\em model-based goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal and each example is an observation of an agent acting to achieve one of that goals.
+Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
 
@@
 
-The {\em Plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelyhood using an off-the-shelf classical planner.
-Recent works show that faster, but less accurate estimates, of this $P(\mathcal{O}|g)$ likelyhood can also be computed using relaxations of the classical planning task.
-
-\section{Model-Based Goal Recognition with Unknown Domain Models}
-\label{sec:recognition}
+The {\em Plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelyhood leveraging the action model of the observed agent and an off-the-shelf classical planner.
+Recent works show that faster, but less accurate estimates, of this $P(\mathcal{O}|g)$ likelyhood can also be computed using relaxations of the classical planning tasks~\cite{pereira2017landmark}.
 

### 2019-01-25 doc/ijcai-19-goalRecognition/goalRecognition.tex
extended background of ijcai19 paper on goal recognition

@@
 
-\title{Model-Based Goal Recognition with Unknown Domain Models}
+\title{Goal Recognition as Planning with Unknown Domain Models}
 
@@
 \begin{abstract}
-The paper shows how to relax one key assumption of the {\em plan recognition as planning} for {\em goal recognition} that is knowing the action model of the observed agent.
-The paper introduces a novel formulation that fits together the {\em learning of planning action models} with the {\em plan recognition as planning} approach.
-The empirical evaluation evidences that our novel formulation allows to solve standard goal recognition benchmarks without having knowing the action model of the observed agent.
+The paper shows how to relax one key assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is knowing the action model of the observed agent.
+The paper introduces a novel formulation that fits together the {\em learning of planning action models} with {\em plan recognition as planning}.
+The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks without {\em a priori} knowing the action model of the observed agent.
 \end{abstract}
@@
 {\em Goal recognition} is a particular classification task in which each class represents a different goal and each example is an observation of an agent acting to achieve one of that goals.
-Despite there is a wide range of diffenrent approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most popular since it is at the core of several interesting tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
+Despite there exists a wide range of diffenrent approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most appealing since it is at the core of various activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
 
-{\em Plan recognition as planning} leverages the action model of the observed agent and an off-the-shelf classical planner to compute the most likely goal of the agent.
-In this paper we show that we can relax the key assumption of {\em plan recognition as planning} for {\em Goal recognition} that is knowing the action model of the observed agent.
+{\em Plan recognition as planning} leverages the action model of the observed agent and an off-the-shelf classical planner to compute the most likely goal of that agent.
+In this paper we show that we can relax the key assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is having an action model of the observed agent.
 In particular, the paper introduces a novel formulation that fits together the {\em learning of planning action models} with the {\em plan recognition as planning} approach.
-The evaluation of our formulation evidences that it allows to solve goal recognition tasks, even when the action model of the observed is unknown.
+The evaluation of our formulation evidences that it allows to solve goal recognition tasks, even when the action model of the observed is unknown, using an off-the-shelf classical planner.
 
@@
 \label{sec:background}
-This section formalizes the {\em planning model} we follow as well as the kind of {\em observations} that are given as classification examples for the {\em goal recognition} task.
+This section formalizes the {\em planning model} we follow as well as the kind of {\em observations} that are given as input to the {\em goal recognition} task.
 
 \subsection{Classical planning with conditional effects}
-Let $F$ be the set of {\em fluents} or {\em state variables} (propositional variables) describing a state.
+Let $F$ be the set of  propositional state variables ({\em fluents}) describing a state.
 A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e.
@@
 A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
-A solution plan is {\em optimal} iff it minimizes the sum of action costs.
+A solution plan is {\em optimal} iff its cost is minimal.
 
@@
 
-\section{Model-Based Goal Recognition with Unknown Domain Models}
-\label{sec:recognition}
-
-{\em Goal recognition} is a particular classification task in which each class represents a different goal and each example is an observation of an agent acting to achieve one of that goals.
+\subsection{Goal recognition as classical planning}
+{\em Goal recognition} is a particular classification task in which each class represents a different goal $g\in G[\cdot]$ and each example is an $\mathcal{O}(\tau)$ observation of an agent acting to achieve one of the input goals in $G[\cdot]$.
 Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
@@
 
-The {\em Plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelyhood leveraging the action model of the observed agent and an off-the-shelf classical planner.
-Recent works show that faster, but less accurate estimates, of this $P(\mathcal{O}|g)$ likelyhood can also be computed using relaxations of the classical planning tasks~\cite{pereira2017landmark}.
-
-\subsection{Well-defined \strips\ action schemata}
-\strips\ action schemata provide a compact representation for specifying classical planning models.{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-
-Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
-The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
-For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contains five elements for a {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for a {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is bounded by a set of constraints ${\mathcal C}$ of three kinds:
-\begin{enumerate}
-\item {\em Syntactic constraints}.
-\strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-{\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
-\item {\em Domain-specific constraints}.
-One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\it State invariants} are also constraints of this kind~\cite{fox1998automatic}.
-\item {\em Observation constraints}. An observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
-\end{enumerate}
-
-\begin{definition}[Well-defined \strips\ action schemata]
-Given a set of {\em predicates} $\Psi$, a list of action {\em parameters} $pars(\xi)$, and set of FOL constraints ${\mathcal C}$, $\xi$ is a {\bf well-defined \strips\ action schema} iff its three lists $pre(\xi)\subseteq {\mathcal I}_{\Psi,\xi}$, $del(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ and $add(\xi)\subseteq{\mathcal I}_{\Psi,\xi}$ only contain elements in ${\mathcal I}_{\Psi,\xi}$ and they satisfy all the constraints in ${\mathcal C}$.
-\end{definition}
-
-We say a planning model $\mathcal{M}$ is {\em well-defined} if all its \strips\ action schemata are {\em well-defined}.
-
-\subsection{Edit distances for \strips\ planning models}
-First, we define the two edit \emph{operations} on a schema $\xi$ that belongs to a \strips\ model $\mathcal{M}\in M$:
-
+The {\em plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelihood leveraging the action model of the observed agent and an off-the-shelf classical planner.
+More precisely, given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$, where $G[\cdot]$ represents the set of possible goals, then the {\em plan recognition as planning} approach estimates the $P(\mathcal{O}|g)$ by computing the cost difference between of solution plans to these two different classical planning problems:
 \begin{itemize}
-\item {\em Deletion}. Given $\xi\in\mathcal{M}$, an element from any of the lists $pre(\xi)$/$del(\xi)$/$add(\xi)$ is removed such that the result is a {\em well-defined} \strips\ action schema.
-\item {\em Insertion}.
-Given $\xi\in\mathcal{M}$, an element in ${\mathcal I}_{\Psi,\xi}$ is added to any of the lists $pre(\xi)$/$del(\xi)$/$add(\xi)$ such that the result is a {\em well-defined} action schema.
+\item $P'_{\top}$, that constrains problem $P$ to achieve $g\in G[\cdot]$ through a plan $\pi_\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
+\item $P'_{\bot}$, that constrains problem $P$ to achieve $g\in G[\cdot]$ through a plan $\pi_\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$.
 \end{itemize}
 
-Second, let us define when to action models are comparable.
-For instance, we claim that the {\small\tt stack(?v1,?v2)} and {\small\tt unstack(?v1,?v2)} actions schemata from a four operator {\em blocksworld}~\cite{slaney2001blocks} are comparable while, the {\small\tt stack(?v1,?v2)} and {\small\tt pick-up(?v1)} schemata are not.
-Last but not least, we say that two \strips\ models $\mathcal{M}$ and $\mathcal{M}'$ are {\em comparable} iff there exists a bijective function $\mathcal{M} \mapsto \mathcal{M}^*$ that maps every action schema $\xi\in\mathcal{M}$ to a comparable schemata $\xi'\in\mathcal{M'}$ and vice versa.
+The higher the value of this cost difference $\Delta(cost(\pi_\top),cost(\pi_\bot))$, the better $g\in G[\cdot]$ predicts $\mathcal{O}(\tau)$ and hence, the higher $P(\mathcal{O}|g)$ likelihood.
+The function used by the {\em plan recognition as planning} approach for mapping the previous cost difference into likelihoods is the sigmoid function:
 
-\begin{definition}[Comparable \strips\ action schemata]
-Two \strips\ schemata $\xi$ and $\xi'$ are {\bf comparable} iff $pars(\xi)=pars(\xi')$, i.e, both share the same list of parameters.\footnote{In \strips\ models, $pars(\xi)=pars(\xi')$ implies the number of parameters must be the same.
-For other planning models that allow object typing, the equality implies that parameters share the same type}
-\end{definition}
+\begin{align}
+P(\mathcal{O}|g) = \frac{1}{1+e^{-\beta\Delta(cost(\pi_\top),cost(\pi_\bot))}}
+\end{align}
 
-We are now ready to formalize an {\em edit distance} that quantifies how similar two given \strips\ models are.
-The distance is symmetric and meets the {\em metric axioms} provided that the two edit operations, {\em deletion} and {\em insertion}, have the same positive cost.
+This expression is derived from the assumption that while the observed agent is not perfectly rational, he is more likely to follow cheaper plans, according to a Boltzmann distribution.
+The larger the value of $\beta$, the more rational the agent, and the less likely that he will follow suboptimal plans.
 
-\begin{definition}[Edit distance]
-  Let $\mathcal{M}$ and $\mathcal{M}'$ be two {\em comparable} and {\em well-defined} \strips\ planning models within the same set of predicates $\Psi$.
-The {\bf edit distance} $\delta(\mathcal{M},\mathcal{M}')$ is the minimum number of {\em edit operations} that is required to transform $\mathcal{M}$ into $\mathcal{M}'$.
-\end{definition}
+Recent works show that faster estimates of this $P(\mathcal{O}|g)$ likelihood can also be computed using relaxations of the $P'_{\top}$ and $P'_{\bot}$ classical planning tasks~\cite{pereira2017landmark}.
 
-Since ${\mathcal I}_{\Psi,\xi}$ is a bounded set, the maximum number of edits that can be introduced to an action schema is bounded as well.
-The \textbf{maximum edit distance} of a \strips\ model $\mathcal{M}$ built with predicates $\Psi$ is $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 3\times|{\mathcal I}_{\Psi,\xi}|$ (note that if we consider the set of syntactic constraints then $\delta(\mathcal{M},*)=\sum_{\xi\in\mathcal{M}} 2\times|{\mathcal I}_{\Psi,\xi}|$).
-
-\vspace{0.02cm}
-
-An observation of the execution of a plan generated with $\mathcal{M}$ further constraints the space of possible action schemata of $\mathcal{M}$.
-The \emph{semantic knowledge} included in the observations introduce a third type of constraints, that we will call {\em observation constraints}, and that can be added to the set $\mathcal{C}$.
-In addition, {\em observation constraints} allow us to define an edit distance to elicit the value of $P(\mathcal{O}|\mathcal{M})$.
-It can be argued that the shorter this distance the better the given model explains the given observation.
-
-\begin{definition}[Observation edit distance]
-Given a planning problem $P$, an observation $\mathcal{O}(\tau)$ of the execution of a plan that solves $P$ and a \strips\ planning model $\mathcal{M}$ (all defined within the same set of predicates $\Psi$).
-The {\bf observation edit distance}, $\delta^o(\mathcal{M},\mathcal{O})$, is the minimal edit distance from $\mathcal{M}$ to any {\em comparable} and well-defined model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ produces a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$; \[\delta^o(\mathcal{M},\mathcal{O})=\min_{\forall \mathcal{M}' \rightarrow \mathcal{O}} \delta(\mathcal{M},\mathcal{M}')\]
-\end{definition}
-
-$\delta^o(\mathcal{M},\mathcal{O})$ can also be defined through the edition that the observation $\mathcal{O}(\tau)$ requires to fit $\mathcal{M}$.
-This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of the model $\mathcal{M}$~\cite{yang2007learning,SohrabiRU16}.
-Our definition of {\em observation edit distance} is more practical since the size of ${\mathcal I}_{\Psi,\xi}$ is usually much smaller than $F$ (the number of variables in the action schemata should normally be lower than the number of objects in a planning problem).
-
-\begin{definition}[{\em Closest consistent models}] \label{consistent}
-Given a model $\mathcal{M}$, the set $M^*$ of the {\bf closest consistent models} is the set of models $\mathcal{M}'$ that: (1) produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$ and (2) their {\em edit distance} to $\mathcal{M}$ is minimal;
-  \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
-\end{definition}
+\section{Goal Recognition as Planning with Unknown Domain Models}
+\label{sec:recognition}
+This section shows that $cost(\pi_\top)$ and $cost(\pi_\bot)$, and hence an approximation to the $P(\mathcal{O}|g)$ likelihood, can also be computed with classical planing when the action model of the observed agent is {\em unknown}.
 

### 2019-01-25 doc/ijcai-19-goalRecognition/goalRecognition.tex
section 3 added to ijcai19 paper on goal recognition

@@
 The {\em plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelihood leveraging the action model of the observed agent and an off-the-shelf classical planner.
-More precisely, given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$, where $G[\cdot]$ represents the set of possible goals, then the {\em plan recognition as planning} approach estimates the $P(\mathcal{O}|g)$ by computing the cost difference between of solution plans to these two different classical planning problems:
+More precisely, given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$, where $G[\cdot]$ represents the set of possible goals, then the {\em plan recognition as planning} approach estimates the $P(\mathcal{O}|g)$ by computing the cost difference of the solutions to these two different classical planning problems:
 \begin{itemize}
-\item $P'_{\top}$, that constrains problem $P$ to achieve $g\in G[\cdot]$ through a plan $\pi_\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-\item $P'_{\bot}$, that constrains problem $P$ to achieve $g\in G[\cdot]$ through a plan $\pi_\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$.
+\item $P^{\top}$, that constrains the classical planning problem $P=\tup{F,A,I,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
+\item $P^{\bot}$, that constrains $P=\tup{F,A,I,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$.
 \end{itemize}
 
-The higher the value of this cost difference $\Delta(cost(\pi_\top),cost(\pi_\bot))$, the better $g\in G[\cdot]$ predicts $\mathcal{O}(\tau)$ and hence, the higher $P(\mathcal{O}|g)$ likelihood.
-The function used by the {\em plan recognition as planning} approach for mapping the previous cost difference into likelihoods is the sigmoid function:
+The higher the value of this difference $\Delta(cost(\pi^\top),cost(\pi^\bot))$, the higher $P(\mathcal{O}|g)$ likelihood.
+{\em Plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:
 
 \begin{align}
-P(\mathcal{O}|g) = \frac{1}{1+e^{-\beta\Delta(cost(\pi_\top),cost(\pi_\bot))}}
+P(\mathcal{O}|g) = \frac{1}{1+e^{-\beta\Delta(cost(\pi^\top),cost(\pi^\bot))}}
 \end{align}
 
-This expression is derived from the assumption that while the observed agent is not perfectly rational, he is more likely to follow cheaper plans, according to a Boltzmann distribution.
+This expression is derived from the assumption that while the observed agent is not perfectly rational, he is more likely to follow cheaper plans, according to a {\em Boltzmann} distribution.
 The larger the value of $\beta$, the more rational the agent, and the less likely that he will follow suboptimal plans.
 
-Recent works show that faster estimates of this $P(\mathcal{O}|g)$ likelihood can also be computed using relaxations of the $P'_{\top}$ and $P'_{\bot}$ classical planning tasks~\cite{pereira2017landmark}.
+Recent works show that estimates of the $P(\mathcal{O}|g)$ likelihood can be faster computed using relaxations of the $P^{\top}$ and $P^{\bot}$ classical planning tasks~\cite{pereira2017landmark}.
 
-\section{Goal Recognition as Planning with Unknown Domain Models}
+\section{Goal recognition as planning with unknown domain models}
 \label{sec:recognition}
-This section shows that $cost(\pi_\top)$ and $cost(\pi_\bot)$, and hence an approximation to the $P(\mathcal{O}|g)$ likelihood, can also be computed with classical planing when the action model of the observed agent is {\em unknown}.
+
+We define the task of {\em goal recognition with unknown domain models} as a $\tup{P,\mathcal{O}(\tau)}$ pair, where:
+\begin{itemize}
+\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a planning problem where $G[\cdot]$ is the set of possible goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
+the functions $\rho$ and/or $\theta$ of $a$ are undefined).
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,I)$ produced by the execution of an unknown plan $\pi$ that reaches a goal $g\in G[\cdot]$ starting from the given initial state.
+\end{itemize}
+
+The {\em solution} to the {\em goal recognition with unknown domain models} task is again the subset of goals in $G[\cdot]$ that maximizes expression (1).
+
+Next we show that plans $\pi^\top$ and $\pi^\bot$, and hence an approximation to the $P(\mathcal{O}|g)$ likelihood, can also be computed with an off-the-shelf classical planner despite the action model of the observed agent (i.e., the semantics of actions $a\in A[\cdot]$) is {\em unknown}.
+Our approach to compute $\pi^\top$ and $\pi^\bot$ plans with unknown domain models is to adapt the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
+First we introduce how we encode a space of STRIPS action models as a set of propositional variables and associated constraints and then, we show how the classical planning compilation is adapted for the computation of the $\pi^\top$ and $\pi^\bot$ plans.
+
+\subsection{A propositional encoding for the space of STRIPS action models}
+{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
+The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
+
+For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contains five elements for a {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for a {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is bounded by constraints of three kinds:
+\begin{enumerate}
+\item {\em Syntactic constraints}.
+\strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
+{\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
+\item {\em Domain-specific constraints}.
+One can introduce domain-specific knowledge to constrain further the space of possible schemata.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
+{\it State invariants} are also constraints of this kind~\cite{fox1998automatic}.
+\item {\em Observation constraints}. An observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
+\end{enumerate}
+
+\begin{figure}
+  \begin{tiny}  
+  \begin{verbatim}
+(:action stack
+   :parameters (?v1 ?v2)
+   :precondition (and (holding ?v1) (clear ?v2))
+   :effect (and (not (holding ?v1)) (not (clear ?v2))
+                (clear ?v1) (handempty) (on ?v1 ?v2)))
+
+(pre_holding_v1_stack) (pre_clear_v2_stack)
+(eff_holding_v1_stack) (eff_clear_v2_stack)
+(eff_clear_v1_stack) (eff_handempty_stack) (eff_on_v1_v2_stack)
+  \end{verbatim}           
+  \end{tiny}  
+ \caption{\small PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.}
+\label{fig:propositional}
+\end{figure}
+
+In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+This encoding exploits the syntactic constraints of \strips\ so is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
+
+\subsection{Explaining observations with unknown domain models}
+The classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} uses a set of $\mathcal{O}(\tau)$ input observations to complete a given classical planning frame $P=\tup{F,A[\cdot]}$, where $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
+the functions $\rho$ and/or $\theta$ of $a$ are undefined).
+
+The output of the compilation is a new classical planning problem $P_\Lambda$ s.t a solution plan $\pi_\Lambda$ for $P_\Lambda$ is a sequence of actions that:
+\begin{enumerate}
+\item Builds the action models of the {\em learned} domain model. 
+\item Uses the {\em learned} domain model to build a plan that is consistent with the given input observations.
+\end{enumerate}
+Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the preconditions and effects of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\mathcal{O}(\tau)$.
+
+Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states of $\mathcal{O}(\tau)$.
+
+\subsection{Computing the $P(\mathcal{O}|g)$ with unknown domain models}
+Now we are ready to compute the target distribution $P(g|\mathcal{O})$ over the possible goals $g\in G[\cdot]$ given the observation $\mathcal{O}(\tau)$:
+\begin{enumerate}
+\item For each goal, we define the $P^{\top}$, that constrains the classical planning problem $P=\tup{F,A[\cdot],s_0,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
+Note that $s_0$ is the initial state in the given observation $\mathcal{O}(\tau)$.
+Likewise we define $P^{\bot}$, that constrains $P=\tup{F,A[\cdot],s_0,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$.
+\item We use the adapted compilation to compute the classcial planning tasks $P^{\top}_\lambda$ and $P^{\bot}_\lambda$ and solve them using an off-the-shelf-classical planner.
+\item We compute the cost difference $\Delta(cost(\pi_\top),cost(\pi_\bot))$ where these costs are defined as the length of the postfix of the $\pi^{\top}_\lambda$ and $\pi^{\bot}_\lambda$ plans and plug this cost difference into equation (2) to get the $P(g|\mathcal{O})$ likelihoods.
+\item Finally the previous likelihoods are plugged into the Bayes rule from which the goal posterior probabilities are obtained.
+In this case the $P\mathcal{O}(\tau)$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
+\end{enumerate}
 

### 2019-01-28 doc/ijcai-19-goalRecognition/goalRecognition.tex
added comments by Diego and Eva to the IJCAI19 paper on goal recognition

@@
 \label{sec:introduction}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal and each example is an observation of an agent acting to achieve one of that goals.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of an agent acting to achieve one of that goals.
 Despite there exists a wide range of diffenrent approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most appealing since it is at the core of various activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
@@
 In particular, the paper introduces a novel formulation that fits together the {\em learning of planning action models} with the {\em plan recognition as planning} approach.
-The evaluation of our formulation evidences that it allows to solve goal recognition tasks, even when the action model of the observed is unknown, using an off-the-shelf classical planner.
+The evaluation of our formulation evidences that it allows to solve goal recognition tasks, even when the action model of the observed is unknown, with an off-the-shelf classical planner.
 
@@
 \label{sec:background}
-This section formalizes the {\em planning model} we follow as well as the kind of {\em observations} that are given as input to the {\em goal recognition} task.
+This section formalizes the {\em planning model} we follow, the kind of {\em observations} that are given as input to the {\em goal recognition} task and the {\em plan recognition as planning} approach for {\em goal recognition}.
 
@@
 
-\subsection{Goal recognition as classical planning}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal $g\in G[\cdot]$ and each example is an $\mathcal{O}(\tau)$ observation of an agent acting to achieve one of the input goals in $G[\cdot]$.
+\subsection{Goal recognition with classical planning}
+{\em Goal recognition} is a particular classification task in which each class represents a different goal $g\in G[\cdot]$ and there is a single classification example $\mathcal{O}(\tau)$ that represents the observation of an agent acting to achieve one of the input goals in $G[\cdot]$.
 Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
@@
 The larger the value of $\beta$, the more rational the agent, and the less likely that he will follow suboptimal plans.
-
 Recent works show that estimates of the $P(\mathcal{O}|g)$ likelihood can be faster computed using relaxations of the $P^{\top}$ and $P^{\bot}$ classical planning tasks~\cite{pereira2017landmark}.
 
-\section{Goal recognition as planning with unknown domain models}
-\label{sec:recognition}
+The original work on {\em plan recognition as planning} assumes a more restricted observation model where observations are only about actions~\cite{ramirez2009plan}.
+However the same approach applies to more expressive observation models that consider also state observations, like the one define above, with a trivial three-fold extension: (1) One fluent $\{test_j\}_{0\leq j\leq m}$ to point at every $s_j\in\mathcal{O}(\tau)$ state observation, (2) adding $test_m$ to every possible goal $g\in G[\cdot]$ to constraint $\pi^\top$ to be consistent with all the state observations and (3), one $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j\in\mathcal{O}(\tau)$ input state observation.
 
-We define the task of {\em goal recognition with unknown domain models} as a $\tup{P,\mathcal{O}(\tau)}$ pair, where:
-\begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a planning problem where $G[\cdot]$ is the set of possible goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j\}.
+\end{align*}
+\end{small}
+
+\section{Classical planning with unknown domain models}
+\label{sec:planning}
+This section shows how to solve a classical planning problem $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,I)$ produced by the execution of an unknown plan $\pi$ that reaches a goal $g\in G[\cdot]$ starting from the given initial state.
-\end{itemize}
-
-The {\em solution} to the {\em goal recognition with unknown domain models} task is again the subset of goals in $G[\cdot]$ that maximizes expression (1).
-
-Next we show that plans $\pi^\top$ and $\pi^\bot$, and hence an approximation to the $P(\mathcal{O}|g)$ likelihood, can also be computed with an off-the-shelf classical planner despite the action model of the observed agent (i.e., the semantics of actions $a\in A[\cdot]$) is {\em unknown}.
-Our approach to compute $\pi^\top$ and $\pi^\bot$ plans with unknown domain models is to adapt the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
-First we introduce how we encode a space of STRIPS action models as a set of propositional variables and associated constraints and then, we show how the classical planning compilation is adapted for the computation of the $\pi^\top$ and $\pi^\bot$ plans.
+First, se show how to encode the possible models for the actions in $A[\cdot]$ with a set of propositional variables and a set of  constraints and then we show how to exploit this encoding to compute a solution to the $P=\tup{F,A[\cdot],I,G}$ problem with an off-the-shelf classical planner.
 
@@
 
-\subsection{Explaining observations with unknown domain models}
-The classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} uses a set of $\mathcal{O}(\tau)$ input observations to complete a given classical planning frame $P=\tup{F,A[\cdot]}$, where $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
+\subsection{A classical planning compilation for planning with unknown domain models}
+To solve a classical planning problem $P=\tup{F,A[\cdot],I,G}$ we create another classical planning problem $P=\tup{F',A',I,G}$ such that:
+\begin{itemize}
+\item $F'$ extends $F$ with the necessary fluents to encode the corresponding space of STRIPS action models.
+This is a set of \emph{editable fluents} of the type $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
+
+\item $A'$ replaces the actoin in $A$ with two types of actions.
+\begin{enumerate}
+\item Actions for {\em inserting} a component (precondition, positive effect or negative effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models. 
+\begin{itemize}
+\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
+A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi)\},\\
+\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
+\end{align*}
+\end{small}
+
+\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p}(\xi)\},\\
+\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff_{p}(\xi)\}.
+\end{align*}
+\end{small}
+\end{itemize}
+
+\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{ar(\xi)}$.
+Since action headers are known, the variables $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
+&\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\}.
+\end{align*}
+\end{small}
+
+These actions determine the application of the learned action models according to the values of the model fluents in the current state configuration.
+Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator.
+Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$.
+Executing this action in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$.
+For example, if  $\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}}$\} \subset s$ then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
+Otherwise, a different set of precondition literals will be checked.
+The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.
+
+Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if {\tt{\small stack}} has been correctly programmed by the insert actions.
+Hence, while \textbf{insert actions} add the values of the \textbf{model fluents} that shape $\xi$, the \textbf{apply actions} add the values of the \textbf{fluents of $F$} that result from the execution of $\xi$.
+
+\begin{figure}[hbt!]
+\begin{center}
+\begin{scriptsize}
+\begin{verbatim}
+(:action apply_stack
+  :parameters (?o1 - object ?o2 - object)
+  :precondition
+   (and (or (not (pre_stack_on_v1_v1)) (on ?o1 ?o1))
+        (or (not (pre_stack_on_v1_v2)) (on ?o1 ?o2))
+        (or (not (pre_stack_on_v2_v1)) (on ?o2 ?o1))
+        (or (not (pre_stack_on_v2_v2)) (on ?o2 ?o2))
+        (or (not (pre_stack_ontable_v1)) (ontable ?o1))
+        (or (not (pre_stack_ontable_v2)) (ontable ?o2))
+        (or (not (pre_stack_clear_v1)) (clear ?o1))
+        (or (not (pre_stack_clear_v2)) (clear ?o2))
+        (or (not (pre_stack_holding_v1)) (holding ?o1))
+        (or (not (pre_stack_holding_v2)) (holding ?o2))
+        (or (not (pre_stack_handempty)) (handempty)))
+  :effect
+   (and (when (del_stack_on_v1_v1) (not (on ?o1 ?o1)))
+        (when (del_stack_on_v1_v2) (not (on ?o1 ?o2)))
+        (when (del_stack_on_v2_v1) (not (on ?o2 ?o1)))
+        (when (del_stack_on_v2_v2) (not (on ?o2 ?o2)))
+        (when (del_stack_ontable_v1) (not (ontable ?o1)))
+        (when (del_stack_ontable_v2) (not (ontable ?o2)))
+        (when (del_stack_clear_v1) (not (clear ?o1)))
+        (when (del_stack_clear_v2) (not (clear ?o2)))
+        (when (del_stack_holding_v1) (not (holding ?o1)))
+        (when (del_stack_holding_v2) (not (holding ?o2)))
+        (when (del_stack_handempty) (not (handempty)))
+        (when (add_stack_on_v1_v1) (on ?o1 ?o1))
+        (when (add_stack_on_v1_v2) (on ?o1 ?o2))
+        (when (add_stack_on_v2_v1) (on ?o2 ?o1))
+        (when (add_stack_on_v2_v2) (on ?o2 ?o2))
+        (when (add_stack_ontable_v1) (ontable ?o1))
+        (when (add_stack_ontable_v2) (ontable ?o2))
+        (when (add_stack_clear_v1) (clear ?o1))
+        (when (add_stack_clear_v2) (clear ?o2))
+        (when (add_stack_holding_v1) (holding ?o1))
+        (when (add_stack_holding_v2) (holding ?o2))
+        (when (add_stack_handempty) (handempty))
+        (when (modeProg) (not (modeProg)))))
+\end{verbatim}
+\end{scriptsize}
+ \caption{\small PDDL action for applying an already programmed model for $stack$ (implications are coded as disjunctions).}
+\label{fig:compilation}
+\end{center}
+\end{figure}
+\end{enumerate}
+\end{itemize}
+
+\subsection{The bias of the initially empty action model}
+Our compilation introduces a bias to solve the $P=\tup{F,A[\cdot],I,G}$ classical planning task using action models that have the smallest number of preconditions and effects, because this means less actions to solve the resulting compilation.
+This bias can be eliminated using action cost if the actions for determinign a precndition or an action effect has no cost.
+In out case we use a SAT-based planner that applies the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
+Actions for programming action effects can also be applied in a single planning step reducing also the planning horizon.
+
+\section{Goal recognition as planning with unknown domain models}
+\label{sec:recognition}
+
+We define the task of {\em goal recognition with unknown domain models} as a $\tup{P,\mathcal{O}(\tau)}$ pair, where:
+\begin{itemize}
+\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a planning problem where $G[\cdot]$ is the set of possible goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,I)$ produced by the execution of an unknown plan $\pi$ that reaches a goal $g\in G[\cdot]$ starting from the given initial state.
+\end{itemize}
 
-The output of the compilation is a new classical planning problem $P_\Lambda$ s.t a solution plan $\pi_\Lambda$ for $P_\Lambda$ is a sequence of actions that:
-\begin{enumerate}
-\item Builds the action models of the {\em learned} domain model. 
-\item Uses the {\em learned} domain model to build a plan that is consistent with the given input observations.
-\end{enumerate}
-Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the preconditions and effects of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\mathcal{O}(\tau)$.
-
-Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states of $\mathcal{O}(\tau)$.
+The {\em solution} to the {\em goal recognition with unknown domain models} task is again the subset of goals in $G[\cdot]$ that maximizes expression (1).
 
@@
 Note that $s_0$ is the initial state in the given observation $\mathcal{O}(\tau)$.
-Likewise we define $P^{\bot}$, that constrains $P=\tup{F,A[\cdot],s_0,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$.
-\item We use the adapted compilation to compute the classcial planning tasks $P^{\top}_\lambda$ and $P^{\bot}_\lambda$ and solve them using an off-the-shelf-classical planner.
+We use our adapted compilation to compute the classcial planning tasks $P^{\top}_\lambda$ and solve them using an off-the-shelf-classical planner.
+\item For each goal, we define $P^{\bot}$, that constrains $P=\tup{F,A,s_0,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ and that uses the action model $A$ used by the corresponding solution $\pi^\top$.
 \item We compute the cost difference $\Delta(cost(\pi_\top),cost(\pi_\bot))$ where these costs are defined as the length of the postfix of the $\pi^{\top}_\lambda$ and $\pi^{\bot}_\lambda$ plans and plug this cost difference into equation (2) to get the $P(g|\mathcal{O})$ likelihoods.

### 2019-01-29 doc/ijcai-19-goalRecognition/goalRecognition.tex
updates added to the IJCAI19 paper on goal recognition

@@
 
-\newtheorem{theorem}{Theorem}
-\newtheorem{lemma}[theorem]{Lemma}
-\newtheorem{definition}[theorem]{Definition}
+\newtheorem{mytheorem}{Theorem}
+\newtheorem{mylemma}[mytheorem]{Lemma}
+\newtheorem{mydefinition}[mytheorem]{Definition}
 
@@
 \begin{abstract}
-The paper shows how to relax one key assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is knowing the action model of the observed agent.
-The paper introduces a novel formulation that fits together the {\em learning of planning action models} with {\em plan recognition as planning}.
-The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks without {\em a priori} knowing the action model of the observed agent.
+The paper shows how to relax one key assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is knowing the action model of the observed agents.
+The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are given) and it shows how this formulation neatly fits  with the {\em plan recognition as planning} approach.
+The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks without {\em a priori} knowing the action model of the observed agents and using an off-the-shelf classical planner.
 \end{abstract}
@@
 \label{sec:introduction}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of an agent acting to achieve one of that goals.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal and the classification examples are observations of agents acting to achieve one of that goals.
 Despite there exists a wide range of diffenrent approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most appealing since it is at the core of various activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
 
-{\em Plan recognition as planning} leverages the action model of the observed agent and an off-the-shelf classical planner to compute the most likely goal of that agent.
-In this paper we show that we can relax the key assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is having an action model of the observed agent.
-In particular, the paper introduces a novel formulation that fits together the {\em learning of planning action models} with the {\em plan recognition as planning} approach.
-The evaluation of our formulation evidences that it allows to solve goal recognition tasks, even when the action model of the observed is unknown, with an off-the-shelf classical planner.
+{\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to compute the most likely goal of that agents.
+In this paper we show that we can relax the key assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is a priori having an action model of the observed agents.
+In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are given) and it shows how this formulation neatly fits with the {\em plan recognition as planning} approach.
+The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks without {\em a priori} knowing the action model of the observed agents and using an off-the-shelf classical planner.
 
@@
 \label{sec:background}
-This section formalizes the {\em planning model} we follow, the kind of {\em observations} that are given as input to the {\em goal recognition} task and the {\em plan recognition as planning} approach for {\em goal recognition}.
+This section formalizes the {\em planning model} we follow, the kind of {\em observations} that are given as input to the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.
 
@@
 \subsection{Goal recognition with classical planning}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal $g\in G[\cdot]$ and there is a single classification example $\mathcal{O}(\tau)$ that represents the observation of an agent acting to achieve one of the input goals in $G[\cdot]$.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal $g\in G[\cdot]$ and there is a single classification example $\mathcal{O}(\tau)$ that represents the observation of an agent acting to achieve one of the input goals in $g\in G[\cdot]$.
 Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
-
 \begin{align}
@@
 The {\em plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelihood leveraging the action model of the observed agent and an off-the-shelf classical planner.
-More precisely, given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$, where $G[\cdot]$ represents the set of possible goals, then the {\em plan recognition as planning} approach estimates the $P(\mathcal{O}|g)$ by computing the cost difference of the solutions to these two different classical planning problems:
+More precisely, given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$, where $G[\cdot]$ represents the set of possible goals, then the {\em plan recognition as planning} approach estimates the $P(\mathcal{O}|g)$.
+This estimate is computed by calculating, for each goal $g\in G[\cdot]$, the cost difference of the solutions to these two different classical planning problems:
 \begin{itemize}
-\item $P^{\top}$, that constrains the classical planning problem $P=\tup{F,A,I,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-\item $P^{\bot}$, that constrains $P=\tup{F,A,I,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$.
+\item $P^{\top}_g$, that is a classical planning problem built constraining $P=\tup{F,A,I,g}$ to achieve the particular goal $g\in G[\cdot]$ through a plan $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
+\item $P^{\bot}_g$, that constrains $P=\tup{F,A,I,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$.
 \end{itemize}
 
-The higher the value of this difference $\Delta(cost(\pi^\top),cost(\pi^\bot))$, the higher $P(\mathcal{O}|g)$ likelihood.
-{\em Plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:
+The higher the value of this cost difference $cost(\pi^\top)-cost(\pi^\bot)$, the higher probability of aiming to achieve the $g\in G[\cdot]$ goal.
+With this regard, {\em plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:
 
 \begin{align}
-P(\mathcal{O}|g) = \frac{1}{1+e^{-\beta\Delta(cost(\pi^\top),cost(\pi^\bot))}}
+P(\mathcal{O}|g) = \frac{1}{1+e^{-\beta(cost(\pi^\top)-cost(\pi^\bot))}}
 \end{align}
@@
 The larger the value of $\beta$, the more rational the agent, and the less likely that he will follow suboptimal plans.
-Recent works show that estimates of the $P(\mathcal{O}|g)$ likelihood can be faster computed using relaxations of the $P^{\top}$ and $P^{\bot}$ classical planning tasks~\cite{pereira2017landmark}.
+Recent works show that estimates of the $P(\mathcal{O}|g)$ likelihood can be faster computed using relaxations of the classical planning tasks~\cite{pereira2017landmark}.
 
-The original work on {\em plan recognition as planning} assumes a more restricted observation model where observations are only about actions~\cite{ramirez2009plan}.
-However the same approach applies to more expressive observation models that consider also state observations, like the one define above, with a trivial three-fold extension: (1) One fluent $\{test_j\}_{0\leq j\leq m}$ to point at every $s_j\in\mathcal{O}(\tau)$ state observation, (2) adding $test_m$ to every possible goal $g\in G[\cdot]$ to constraint $\pi^\top$ to be consistent with all the state observations and (3), one $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j\in\mathcal{O}(\tau)$ input state observation.
-
+The original work on {\em plan recognition as planning} assumes less expressive observation model where observations are refer only about logs of executed actions~\cite{ramirez2009plan}.
+However the same approach applies to more expressive observation models that consider also state observations, like the one defined above, with a trivial three-fold extension:
+\begin{itemize}
+\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j\in\mathcal{O}(\tau)$ state observation.
+\item Adding $validated_m$ to every possible goal $g\in G[\cdot]$ to constraint the solution plans $\pi^\top$ to be consistent with all the state observations.
+\item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
+\end{itemize}
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j\}.
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{validated_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg validated_{j-1}, validated_j\}.
 \end{align*}
@@
 This section shows how to solve a classical planning problem $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-First, se show how to encode the possible models for the actions in $A[\cdot]$ with a set of propositional variables and a set of  constraints and then we show how to exploit this encoding to compute a solution to the $P=\tup{F,A[\cdot],I,G}$ problem with an off-the-shelf classical planner.
+the functions $\rho$ and/or $\theta$ of $a$ are undefined but the corresponding action headers are known).
+First, se show how to encode the possible models for the actions in $A[\cdot]$ with a set of propositional variables and a set of constraints over that variables and finally, we show how to exploit this encoding to compute a solution to the $P=\tup{F,A[\cdot],I,G}$ problem with an off-the-shelf classical planner.
 
@@
 
-For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contains five elements for a {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for a {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for a {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for a {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
@@
 \begin{itemize}
-\item $F'$ extends $F$ with the necessary fluents to encode the corresponding space of STRIPS action models.
-This is a set of \emph{editable fluents} of the type $\{pre\_e\_\xi, del\_e\_\xi, add\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
+\item $F'$ extends $F$ with the necessary fluents for the propositional encoding of the corresponding space of STRIPS action models.
+This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
 
-\item $A'$ replaces the actoin in $A$ with two types of actions.
+\item $A'$ replaces the actions in $A$ with two new types of actions.
 \begin{enumerate}
@@
 \end{small}
-
-These actions determine the application of the learned action models according to the values of the model fluents in the current state configuration.
-Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator.
-Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$.
-Executing this action in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$.
-For example, if  $\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}}$\} \subset s$ then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
-Otherwise, a different set of precondition literals will be checked.
-The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.
-
-Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if {\tt{\small stack}} has been correctly programmed by the insert actions.
-Hence, while \textbf{insert actions} add the values of the \textbf{model fluents} that shape $\xi$, the \textbf{apply actions} add the values of the \textbf{fluents of $F$} that result from the execution of $\xi$.
 
@@
 
-\subsection{The bias of the initially empty action model}
-Our compilation introduces a bias to solve the $P=\tup{F,A[\cdot],I,G}$ classical planning task using action models that have the smallest number of preconditions and effects, because this means less actions to solve the resulting compilation.
-This bias can be eliminated using action cost if the actions for determinign a precndition or an action effect has no cost.
-In out case we use a SAT-based planner that applies the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
-Actions for programming action effects can also be applied in a single planning step reducing also the planning horizon.
+The dynamics of the actions for {\em applying} an action model $\xi\in\mathcal{M}$ is determined by the values of the model the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in the current state.
+Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator.
+
+For instance, executing the action {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$.
+This means that if the current state $s$ holds {\scriptsize$\{${\tt (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}$\} \subset s$}, then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
+Otherwise, a different set of precondition literals will be checked.
+The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.
+Note that executing {\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if {\tt{\small stack}} has been correctly programmed by the insert actions.
+
+\subsection{Compilation properties}
+Now we present some theoretical properties of the compilation scheme.
+
+\subsubsection{Soundness and completeness}
+
+\begin{mylemma}
+Soundness. Any classical plan $\pi_\Lambda$ that solves $P_{\Lambda}$ induces a set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$.
+\end{mylemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+  Once action models $\mathcal{M}'$ are programmed, they can only be applied and validated because of the $mode_{prog}$ fluent.
+In addition, $P_{\Lambda}$ is only solvable if fluents {\tt\small $at_n$} and {\tt\small $test_m$} hold at the last state reached by $\pi_\Lambda$.
+By the definition of the $\mathsf{apply_{\xi,\omega}}$ and the $\mathsf{validate_{j}}$ actions, these goals can only be achieved executing an applicable sequence of programmed action models that reaches every state $s_j\in\tau$, starting in the corresponding initial state and following the sequence of $n$ observed actions of $\tau$.
+This means that the programmed action model $\mathcal{M}'$ is consistent with the provided input knowledge and hence, that $\mathcal{M}'$ is a solution to $\Lambda$.
+\end{small}
+\end{proof}
+
+\begin{mylemma}
+Completeness. Any set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ is computable solving the corresponding classical planning task $P_{\Lambda}$.
+\end{mylemma}
+
+\begin{proof}[Proof sketch]
+\begin{small}
+  By definition, $\Psi_{\xi}\subseteq \Psi_v$ fully captures the set of elements that can appear in an action model $\xi\in\mathcal{M}$.
+The compilation does not discard any possible set of action models $\mathcal{M}'$ definable within $\Psi_v$ that satisfies the observed state trajectory and action sequence of $\tau$.
+This means that for every $\mathcal{M}'$ that solves $\Lambda$, there exists a plan $\pi_\Lambda$ that can be built selecting the appropriate programming, apply and validate actions from the $P_{\Lambda}$ compilation.
+  %a solution plan $\pi_\Lambda$ can be built selecting the corresponding  $\mathsf{programPre_{p,\xi}}$ and $\mathsf{programEff_{p,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\tau$.
+\end{small}
+\end{proof}
+
+\subsubsection{Size}
+The size of the planning task $P_{\Lambda}$ output by the compilation approach depends on:
+\begin{itemize}
+\item The arity of the actions and the fluents in $\tau$ given as input in $\Lambda$.
+The larger the arity, the larger the size of the $\Psi_{\xi}$ sets.
+This is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents and the corresponding set of {\em programming} actions.
+\item The length of the observed action sequence and state trajectory of $\tau$.
+The larger the number of observed actions, $a_i\in\tau$ s.t.
+$1\leq i\leq n$, the more $\{at_i\}$ fluents.
+The larger the number of observed states, $s_j\in\tau$ s.t.
+$1\leq j\leq m$, the more $\{test_j\}$ fluents and $\{\mathsf{validate_{j}}\}$ actions in $P_{\Lambda}$.
+\end{itemize}
+
+\subsubsection{The bias of the initially empty action model}
+Since in the initial state of the classical planning compilation all the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ are false, our compilation introduces a bias to solve the $P=\tup{F,A[\cdot],I,G}$ classical planning task.
+This bias can be eliminated definin a cost landscape where any actions for determinign a precondition or an action effect has zero cost.
+Since classical planners are nor proficincy when optimizing the cost of solutions with this kind of cost landscapes we use a different aprpoach to disregard the cost of the actions that add a precondtion/effect to an action model.
+Our approach is to use a SAT-based planner that applies the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
+Actions for programming action effects can also be applied in a single planning step so the plan horizon for programming any action model is always 2 which in adddition, significantly reduces the planning horizon.
 

### 2019-01-29 doc/ijcai-19-goalRecognition/goalRecognition.tex
updates added to the ijcai19 paper on goal recognition

@@
 
-The original work on {\em plan recognition as planning} assumes less expressive observation model where observations are refer only about logs of executed actions~\cite{ramirez2009plan}.
+The original work on {\em plan recognition as planning} assumes a less expressive observation model where observations are referred only about logs of executed actions~\cite{ramirez2009plan}.
 However the same approach applies to more expressive observation models that consider also state observations, like the one defined above, with a trivial three-fold extension:
@@
 \label{sec:planning}
-This section shows how to solve a classical planning problem $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
+This section defines a classical planning problem $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined but the corresponding action headers are known).
-First, se show how to encode the possible models for the actions in $A[\cdot]$ with a set of propositional variables and a set of constraints over that variables and finally, we show how to exploit this encoding to compute a solution to the $P=\tup{F,A[\cdot],I,G}$ problem with an off-the-shelf classical planner.
+
+A solution to this task is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a {\em trajectory} $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and it holds that {\em there exists} at least one model in the space of possible action models (e.g.
+one possible definition of the $\rho$ and $\theta$ functions with the given state variables and action headers) that satisfies that $\rho(s_{i-1},a_i)$ and  $s_i=\theta(s_{i-1},a_i)$, for each {\small $1\leq i\leq n$}, as well as it reaches a final state where all goal conditions are met, $G \subseteq s_n$.
+
+First, se show how to encode the space of possible models for the actions in $A[\cdot]$ with a set of propositional variables and a set of constraints over that variables and finally, we show how to exploit this encoding to compute a solution to the $P=\tup{F,A[\cdot],I,G}$ problem with an off-the-shelf classical planner.
 

### 2019-01-29 doc/ijcai-19-goalRecognition/goalRecognition.tex
updates added to the ijcai19 paper on goal recognition

--- 2019-01-29 doc/ijcai-19-goalRecognition/goalRecognition.tex
@@
 %%%%%%%%%%%%%%%%%% Added for this paper
+\usepackage{amsthm}
 \usepackage{listings}% http://ctan.org/pkg/listings
@@
 \begin{abstract}
-The paper shows how to relax one key assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is knowing the action model of the observed agents.
-The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are given) and it shows how this formulation neatly fits  with the {\em plan recognition as planning} approach.
+This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is a priori having an action model of the observed agents.
+The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are known) and it shows that this formulation neatly fits with the popular {\em plan recognition as planning} approach for {\em goal recognition}.
 The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks without {\em a priori} knowing the action model of the observed agents and using an off-the-shelf classical planner.
@@
 {\em Goal recognition} is a particular classification task in which each class represents a different goal and the classification examples are observations of agents acting to achieve one of that goals.
-Despite there exists a wide range of diffenrent approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most appealing since it is at the core of various activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
+Despite there exists a wide range of different approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most appealing and it is currently at the core of various activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
 
 {\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to compute the most likely goal of that agents.
-In this paper we show that we can relax the key assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is a priori having an action model of the observed agents.
-In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are given) and it shows how this formulation neatly fits with the {\em plan recognition as planning} approach.
+In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which frequently becomes a too strong assumption when using {\em plan recognition as planning} at real-world applications.
+In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are given) and it shows that this formulation neatly fits with the {\em plan recognition as planning} approach.
 The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks without {\em a priori} knowing the action model of the observed agents and using an off-the-shelf classical planner.
@@
 
-A {\em classical planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
-Each classical planning action $a\in A$ has a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $cost(a)$.
+A classical planning action $a\in A$ has: a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $cost(a)$.
 The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
@@
 \subsection{Goal recognition with classical planning}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal $g\in G[\cdot]$ and there is a single classification example $\mathcal{O}(\tau)$ that represents the observation of an agent acting to achieve one of the input goals in $g\in G[\cdot]$.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal $g\in G[\cdot]$ and there is a single classification example $\mathcal{O}(\tau)$ that represents the observation of agents acting to achieve one of the input goals in $g\in G[\cdot]$.
 Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
@@
 
-The {\em plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelihood leveraging the action model of the observed agent and an off-the-shelf classical planner.
-More precisely, given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$, where $G[\cdot]$ represents the set of possible goals, then the {\em plan recognition as planning} approach estimates the $P(\mathcal{O}|g)$.
-This estimate is computed by calculating, for each goal $g\in G[\cdot]$, the cost difference of the solutions to these two different classical planning problems:
+The {\em plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelihood leveraging the action model of the observed agent and an off-the-shelf classical planner~\cite{ramirez2012plan}.
+Given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$, where $G[\cdot]$ represents the set of possible goals $P(\mathcal{O}|g)$ is estimated computing, for each goal $g\in G[\cdot]$, the cost difference of the solution plans to these two different classical planning problems:
 \begin{itemize}
-\item $P^{\top}_g$, that is a classical planning problem built constraining $P=\tup{F,A,I,g}$ to achieve the particular goal $g\in G[\cdot]$ through a plan $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-\item $P^{\bot}_g$, that constrains $P=\tup{F,A,I,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$.
+\item $P^{\top}_g$, the classical planning problem built constraining $P=\tup{F,A,I,g}$ to achieve the particular goal $g\in G[\cdot]$ through a plan $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
+\item $P^{\bot}_g$, the classical planning problem that constrains the solutions of $P=\tup{F,A,I,g}$ to plans $\pi^\bot$ that achieve $g\in G[\cdot]$ but are {\em inconsistent} with $\mathcal{O}(\tau)$.
 \end{itemize}
@@
 
-The original work on {\em plan recognition as planning} assumes less expressive observation model where observations are refer only about logs of executed actions~\cite{ramirez2009plan}.
-However the same approach applies to more expressive observation models that consider also state observations, like the one defined above, with a trivial three-fold extension:
+The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions~\cite{ramirez2009plan}.
+However, the approach applies to more expressive observation models that consider also state observations, like the one defined above, with a trivial three-fold extension:
 \begin{itemize}
@@
 \label{sec:planning}
-This section shows how to solve a classical planning problem $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined but the corresponding action headers are known).
-First, se show how to encode the possible models for the actions in $A[\cdot]$ with a set of propositional variables and a set of constraints over that variables and finally, we show how to exploit this encoding to compute a solution to the $P=\tup{F,A[\cdot],I,G}$ problem with an off-the-shelf classical planner.
+This section introduces a novel formulation for classical planning in a setting where no action model is given.
+This setting of classical planning is very related to the learning of action models for planning~\cite{SternJ17}.
+In particular it can be consired an extreme learning scenario when a model has to be learned from a single learning example that contains only two state observations, an initial state and the goals.
+
+A {\em classical planning with unknown domain models} is a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
+the functions $\rho$ and/or $\theta$ of $a$ are undefined).
+
+A solution to this task is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
+one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for each {\small $1\leq i\leq n$}, and such that the reached final state met the goal conditions, $G \subseteq s_n$.
+
+Next we show that the space of possible STRIPS action models can be encoded as a set of propositional variables and a set of constraints over those variables.
+Then, we show how to exploit this encoding to solve $P=\tup{F,A[\cdot],I,G}$ problems with an off-the-shelf classical planner as well as the properties of this approach.
 
@@
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is bounded by constraints of three kinds:
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the actual space of possible \strips\ schemata is bounded by constraints of three kinds:
 \begin{enumerate}
@@
 This encoding exploits the syntactic constraints of \strips\ so is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
-In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a negative effect in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a positive effect in $\xi$.
+In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
 Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
@@
 \subsection{A classical planning compilation for planning with unknown domain models}
-To solve a classical planning problem $P=\tup{F,A[\cdot],I,G}$ we create another classical planning problem $P=\tup{F',A',I,G}$ such that:
+Now we show how we adapt the classical planning compilation~\cite{aineto2018learning} to our propositional encoding.
+In more detail, given a classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$ we create another classical planning problem $P'=\tup{F',A',I,G}$ such that:
 \begin{itemize}
-\item $F'$ extends $F$ with the necessary fluents for the propositional encoding of the corresponding space of STRIPS action models.
+\item $F'$ extends $F$ with the fluents for the propositional encoding of the corresponding space of STRIPS action models.
 This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
@@
 
-\begin{figure}[hbt!]
-\begin{center}
-\begin{scriptsize}
-\begin{verbatim}
-(:action apply_stack
-  :parameters (?o1 - object ?o2 - object)
-  :precondition
-   (and (or (not (pre_stack_on_v1_v1)) (on ?o1 ?o1))
-        (or (not (pre_stack_on_v1_v2)) (on ?o1 ?o2))
-        (or (not (pre_stack_on_v2_v1)) (on ?o2 ?o1))
-        (or (not (pre_stack_on_v2_v2)) (on ?o2 ?o2))
-        (or (not (pre_stack_ontable_v1)) (ontable ?o1))
-        (or (not (pre_stack_ontable_v2)) (ontable ?o2))
-        (or (not (pre_stack_clear_v1)) (clear ?o1))
-        (or (not (pre_stack_clear_v2)) (clear ?o2))
-        (or (not (pre_stack_holding_v1)) (holding ?o1))
-        (or (not (pre_stack_holding_v2)) (holding ?o2))
-        (or (not (pre_stack_handempty)) (handempty)))
-  :effect
-   (and (when (del_stack_on_v1_v1) (not (on ?o1 ?o1)))
-        (when (del_stack_on_v1_v2) (not (on ?o1 ?o2)))
-        (when (del_stack_on_v2_v1) (not (on ?o2 ?o1)))
-        (when (del_stack_on_v2_v2) (not (on ?o2 ?o2)))
-        (when (del_stack_ontable_v1) (not (ontable ?o1)))
-        (when (del_stack_ontable_v2) (not (ontable ?o2)))
-        (when (del_stack_clear_v1) (not (clear ?o1)))
-        (when (del_stack_clear_v2) (not (clear ?o2)))
-        (when (del_stack_holding_v1) (not (holding ?o1)))
-        (when (del_stack_holding_v2) (not (holding ?o2)))
-        (when (del_stack_handempty) (not (handempty)))
-        (when (add_stack_on_v1_v1) (on ?o1 ?o1))
-        (when (add_stack_on_v1_v2) (on ?o1 ?o2))
-        (when (add_stack_on_v2_v1) (on ?o2 ?o1))
-        (when (add_stack_on_v2_v2) (on ?o2 ?o2))
-        (when (add_stack_ontable_v1) (ontable ?o1))
-        (when (add_stack_ontable_v2) (ontable ?o2))
-        (when (add_stack_clear_v1) (clear ?o1))
-        (when (add_stack_clear_v2) (clear ?o2))
-        (when (add_stack_holding_v1) (holding ?o1))
-        (when (add_stack_holding_v2) (holding ?o2))
-        (when (add_stack_handempty) (handempty))
-        (when (modeProg) (not (modeProg)))))
-\end{verbatim}
-\end{scriptsize}
- \caption{\small PDDL action for applying an already programmed model for $stack$ (implications are coded as disjunctions).}
-\label{fig:compilation}
-\end{center}
-\end{figure}
 \end{enumerate}
@@
 The dynamics of the actions for {\em applying} an action model $\xi\in\mathcal{M}$ is determined by the values of the model the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in the current state.
-Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator.
-
 For instance, executing the action {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$.
@@
 Note that executing {\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if {\tt{\small stack}} has been correctly programmed by the insert actions.
+
+\begin{figure}[hbt!]
+	{\scriptsize\tt
+		{\bf 00} : (insert\_pre\_stack\_holding\_v1) \\
+		01 : (insert\_pre\_stack\_clear\_v2)\\
+		{\bf 02} : (insert\_eff\_stack\_clear\_v1)\\
+		03 : (insert\_eff\_stack\_clear\_v2)\\
+		04 : (insert\_eff\_stack\_handempty)\\
+		05 : (insert\_eff\_stack\_holding\_v1)\\
+		06 : (insert\_eff\_stack\_on\_v1\_v2)\\
+		{\bf 07} : (apply\_stack blockA blockB)\\
+	}
+	\caption{\small Plan computed when solving the classical planning problem output by our compiltion for solving a classical planning with unknown domain models.}
+	\label{fig:plan-lplan}
+\end{figure}
+
+The plan of Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving the classical planning problem output by our compiltion for solving a classical planning with unknown domain models.
+In the initial state of that problem the robot is holding the blockA while the blockB is clear and the problem goal is having blockA on blockB.
+The plan shows the insert actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the action models to achive the goals.
+Note that an equivalent solution could be found by inserting the same preconditions and effects into the {\em unstack} action model and then applying action (unstack blockA blockB) instead of (stack blockA blockB).
 
@@
 \begin{mylemma}
-Soundness. Any classical plan $\pi_\Lambda$ that solves $P_{\Lambda}$ induces a set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$.
+Soundness. Any classical plan $\pi'$ that solves $P'$ produces a solution to the classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$.
 \end{mylemma}
 
-\begin{proof}[Proof sketch]
+\begin{proof}[Proof]
 \begin{small}
-  Once action models $\mathcal{M}'$ are programmed, they can only be applied and validated because of the $mode_{prog}$ fluent.
-In addition, $P_{\Lambda}$ is only solvable if fluents {\tt\small $at_n$} and {\tt\small $test_m$} hold at the last state reached by $\pi_\Lambda$.
-By the definition of the $\mathsf{apply_{\xi,\omega}}$ and the $\mathsf{validate_{j}}$ actions, these goals can only be achieved executing an applicable sequence of programmed action models that reaches every state $s_j\in\tau$, starting in the corresponding initial state and following the sequence of $n$ observed actions of $\tau$.
-This means that the programmed action model $\mathcal{M}'$ is consistent with the provided input knowledge and hence, that $\mathcal{M}'$ is a solution to $\Lambda$.
+Once a given precondition or effect is inserted into an action model it can never be removed back.
+In addition, $P'$ is only solvable if all the goals $G$ hold at the last state reached by $\pi'$.
+In the compiled problem the value of $F$, the set of fluents of the original problem, can exclusively be modified using $\mathsf{apply_{\xi,\omega}}$ actions.
+By the definition of these $\mathsf{apply_{\xi,\omega}}$ actions, the set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$.
+This means that the action model used by the $\mathsf{apply_{\xi,\omega}}$ actions has to be consistent with all the produced intermediate states and hence, that the subsequence of $\mathsf{apply_{\xi,\omega}}$ in $\pi'$ is a solution to $P=\tup{F,A[\cdot],I,G}$.
 \end{small}
@@
 \begin{mylemma}
-Completeness. Any set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ is computable solving the corresponding classical planning task $P_{\Lambda}$.
+Completeness. Any plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$ is computable solving the corresponding classical planning task $P'$.
 \end{mylemma}
 
-\begin{proof}[Proof sketch]
+\begin{proof}[Proof]
 \begin{small}
@@
 The compilation does not discard any possible set of action models $\mathcal{M}'$ definable within $\Psi_v$ that satisfies the observed state trajectory and action sequence of $\tau$.
-This means that for every $\mathcal{M}'$ that solves $\Lambda$, there exists a plan $\pi_\Lambda$ that can be built selecting the appropriate programming, apply and validate actions from the $P_{\Lambda}$ compilation.
-  %a solution plan $\pi_\Lambda$ can be built selecting the corresponding  $\mathsf{programPre_{p,\xi}}$ and $\mathsf{programEff_{p,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\tau$.
+This means that for every plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, we can build a plan $\pi'$ selecting the appropriate actions for inserting prenconcition and efects to the corresponding action model and then selecting the  $\mathsf{apply_{\xi,\omega}}$ actions that allows to transform the initial state $I$ into a state $G \subseteq s_n$.
 \end{small}
@@
 
-\subsubsection{Size}
-The size of the planning task $P_{\Lambda}$ output by the compilation approach depends on:
-\begin{itemize}
-\item The arity of the actions and the fluents in $\tau$ given as input in $\Lambda$.
-The larger the arity, the larger the size of the $\Psi_{\xi}$ sets.
-This is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents and the corresponding set of {\em programming} actions.
-\item The length of the observed action sequence and state trajectory of $\tau$.
-The larger the number of observed actions, $a_i\in\tau$ s.t.
-$1\leq i\leq n$, the more $\{at_i\}$ fluents.
-The larger the number of observed states, $s_j\in\tau$ s.t.
-$1\leq j\leq m$, the more $\{test_j\}$ fluents and $\{\mathsf{validate_{j}}\}$ actions in $P_{\Lambda}$.
-\end{itemize}
-
 \subsubsection{The bias of the initially empty action model}
 Since in the initial state of the classical planning compilation all the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ are false, our compilation introduces a bias to solve the $P=\tup{F,A[\cdot],I,G}$ classical planning task.
-This bias can be eliminated definin a cost landscape where any actions for determinign a precondition or an action effect has zero cost.
-Since classical planners are nor proficincy when optimizing the cost of solutions with this kind of cost landscapes we use a different aprpoach to disregard the cost of the actions that add a precondtion/effect to an action model.
+This bias can be eliminated defining a cost landscape where any actions for determinign a precondition or an action effect has zero cost.
+
+In practice, since classical planners are not proficiency when optimizing the cost of solutions with this kind of cost landscapes we use a different aprpoach to disregard the cost of the actions that add a precondtion/effect to an action model.
 Our approach is to use a SAT-based planner that applies the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
 Actions for programming action effects can also be applied in a single planning step so the plan horizon for programming any action model is always 2 which in adddition, significantly reduces the planning horizon.
+The SAT-based planning approach is also convenient because its ability to deal with planning instances populated with dead-ends and because achive good perdormance without defning symeti breaking strategies over the actions for inserting preconditions/effects into the action model.
+
+\subsubsection{Compilation size}
+The size of the classical planning task $P'$ output by the compilation approach depends on the arity of the {\em predicates} $\Psi$ that shape the propositional state variables $F$ and the arity of the action headers given by $A[\cdot]$.
+The larger the arity, the larger the size of the ${\mathcal I}_{\Psi,\xi}$ sets.
+This is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents and the corresponding set of {\em programming} actions.
+Note that for planning models that allow object typing, types can be used to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ significanlty reducing the size of the classcial planning task output by the compilation.
 
@@
 
+\section{Related Work}
+\label{sec:evaluation}
+The problem of {\em classical planning with unknown domain models} has been previously addressed for SAS+ action models~\cite{SternJ17}.
+In this work we formulate this task for \strips\ actoin models and evidence the relevance of this classcial planning setting since its allows to address the {\em goal recognition} task when the action model of the observed agent is not available (which typically is a too strong  assumption).
+
+Related work to {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}, where model edition is used to conform the PDDL models of two agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed optimal plan computed with one of the two models.
+{\em Model recognition}, however, conforms every input model $\mathcal{M}$ with another model $\mathcal{M}'$ that is not given as input but instead computed for every $\mathcal{M}$, and which is consistent with a partial observation of a plan execution.
+
 \section{Conclusions}

### 2019-01-29 doc/ijcai-19-goalRecognition/goalRecognition.tex
updates added to the ijcai19 paper on goal recognition

@@
 This setting of classical planning is very related to the learning of action models for planning~\cite{SternJ17}.
-In particular it can be consired an extreme learning scenario when a model has to be learned from a single learning example that contains only two state observations, an initial state and the goals.
+In particular it can be considered an extreme learning scenario when a model has to be learned from a single learning example that contains only two state observations, an initial state and the goals.
 
@@
 	}
-	\caption{\small Plan computed when solving the classical planning problem output by our compiltion for solving a classical planning with unknown domain models.}
+	\caption{\small Plan computed when solving the classical planning problem output by our compilation for solving a classical planning with unknown domain models.}
 	\label{fig:plan-lplan}
@@
 
-The plan of Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving the classical planning problem output by our compiltion for solving a classical planning with unknown domain models.
+The plan of Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving the classical planning problem output by our compilation for solving a classical planning with unknown domain models.
 In the initial state of that problem the robot is holding the blockA while the blockB is clear and the problem goal is having blockA on blockB.
-The plan shows the insert actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the action models to achive the goals.
+The plan shows the insert actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the action models to achieve the goals.
 Note that an equivalent solution could be found by inserting the same preconditions and effects into the {\em unstack} action model and then applying action (unstack blockA blockB) instead of (stack blockA blockB).
@@
 \subsection{Compilation properties}
-Now we present some theoretical properties of the compilation scheme.
+Now we present the properties of the compilation scheme.
 
@@
 Once a given precondition or effect is inserted into an action model it can never be removed back.
-In addition, $P'$ is only solvable if all the goals $G$ hold at the last state reached by $\pi'$.
-In the compiled problem the value of $F$, the set of fluents of the original problem, can exclusively be modified using $\mathsf{apply_{\xi,\omega}}$ actions.
-By the definition of these $\mathsf{apply_{\xi,\omega}}$ actions, the set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$.
-This means that the action model used by the $\mathsf{apply_{\xi,\omega}}$ actions has to be consistent with all the produced intermediate states and hence, that the subsequence of $\mathsf{apply_{\xi,\omega}}$ in $\pi'$ is a solution to $P=\tup{F,A[\cdot],I,G}$.
+In addition, $P'$ is only solvable iff goals $G$ hold at the last state reached by $\pi'$.
+In the compiled problem the value of $F$, the fluents of the original problem, can exclusively be modified via $\mathsf{apply_{\xi,\omega}}$ actions.
+The set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$.
+This means that the action model used by the $\mathsf{apply_{\xi,\omega}}$ actions has to be consistent with the traversed intermediate states.
+We know that this is true by the definition of the $\mathsf{apply_{\xi,\omega}}$ so hence, the sub-sequence of $\mathsf{apply_{\xi,\omega}}$ appearing in $\pi'$ to solve $P'$ is a solution plan to $P=\tup{F,A[\cdot],I,G}$.
 \end{small}
@@
 \begin{small}
-  By definition, $\Psi_{\xi}\subseteq \Psi_v$ fully captures the set of elements that can appear in an action model $\xi\in\mathcal{M}$.
-The compilation does not discard any possible set of action models $\mathcal{M}'$ definable within $\Psi_v$ that satisfies the observed state trajectory and action sequence of $\tau$.
-This means that for every plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, we can build a plan $\pi'$ selecting the appropriate actions for inserting prenconcition and efects to the corresponding action model and then selecting the  $\mathsf{apply_{\xi,\omega}}$ actions that allows to transform the initial state $I$ into a state $G \subseteq s_n$.
+ ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in an action model $\xi$ using the set of predicates $\Psi$.
+The compilation does not discard any possible action model definable within ${\mathcal I}_{\Psi,\xi}$.
+This means that for every plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, we can build a plan $\pi'$ selecting the appropriate actions for inserting precondition and effects to the corresponding action model and then selecting the  $\mathsf{apply_{\xi,\omega}}$ actions that allows to transform the initial state $I$ into a state $G \subseteq s_n$.
 \end{small}
@@
 \subsubsection{The bias of the initially empty action model}
-Since in the initial state of the classical planning compilation all the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ are false, our compilation introduces a bias to solve the $P=\tup{F,A[\cdot],I,G}$ classical planning task.
-This bias can be eliminated defining a cost landscape where any actions for determinign a precondition or an action effect has zero cost.
+Given that at the initial state of the classical planning compilation any $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluent is false, our compilation introduces a bias to solutions of the $P=\tup{F,A[\cdot],I,G}$ classical planning task.
 
-In practice, since classical planners are not proficiency when optimizing the cost of solutions with this kind of cost landscapes we use a different aprpoach to disregard the cost of the actions that add a precondtion/effect to an action model.
-Our approach is to use a SAT-based planner that applies the actions for programming preconditions in a single planning step (in parallel) because these actions do not interact.
-Actions for programming action effects can also be applied in a single planning step so the plan horizon for programming any action model is always 2 which in adddition, significantly reduces the planning horizon.
-The SAT-based planning approach is also convenient because its ability to deal with planning instances populated with dead-ends and because achive good perdormance without defning symeti breaking strategies over the actions for inserting preconditions/effects into the action model.
+This bias might be eliminated defining a cost landscape where any actions for {\em inserting} a precondition or an effect into a given action model has a zero cost while the remaining actions, $\mathsf{apply_{\xi,\omega}}$ actions has unit cost.
+In practice, since classical planners are not proficiency optimizing the cost of solutions with cost landscapes of this kind, we use a different approach that disregards the cost of the actions that insert a precondition/effect to an action model.
+
+Our approach is to use SAT-based planning because it can apply all the required actions for inserting preconditions in a single planning step (in parallel), because these actions do not interact.
+Actions for programming action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which in addition significantly reduces the planning horizon.
+The SAT-based planning approach is also convenient because its ability to deal with planning problems populated with dead-ends and because the symmetries in the insertion of preconditions/effects into the action model do not affect the performance of SAT-based planning.
 
 \subsubsection{Compilation size}
-The size of the classical planning task $P'$ output by the compilation approach depends on the arity of the {\em predicates} $\Psi$ that shape the propositional state variables $F$ and the arity of the action headers given by $A[\cdot]$.
+The size of the classical planning task $P'$ output by the compilation approach depends on the arity of the {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the arity of the action headers, given by $A[\cdot]$.
 The larger the arity, the larger the size of the ${\mathcal I}_{\Psi,\xi}$ sets.
-This is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents and the corresponding set of {\em programming} actions.
-Note that for planning models that allow object typing, types can be used to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ significanlty reducing the size of the classcial planning task output by the compilation.
+This is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
+
+Note that for planning models that allow object typing, types can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ significantly reducing the size of the classical planning task output by the compilation.
 
@@
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a planning problem where $G[\cdot]$ is the set of possible goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
+\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a classical planning problem where $G[\cdot]$ is the set of possible goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,I)$ produced by the execution of an unknown plan $\pi$ that reaches a goal $g\in G[\cdot]$ starting from the given initial state.
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that reaches a goal $g\in G[\cdot]$ starting from the given initial state.
 \end{itemize}
@@
 
-\subsection{Computing the $P(\mathcal{O}|g)$ with unknown domain models}
-Now we are ready to compute the target distribution $P(g|\mathcal{O})$ over the possible goals $g\in G[\cdot]$ given the observation $\mathcal{O}(\tau)$:
+\subsection{Estimating $P(\mathcal{O}|g)$ with classical planning}
+Now we are ready to compute an estimate of the target probability distribution $P(g|\mathcal{O})$ over the possible goals $g\in G[\cdot]$ given an observation $\mathcal{O}(\tau)$:
 \begin{enumerate}
-\item For each goal, we define the $P^{\top}$, that constrains the classical planning problem $P=\tup{F,A[\cdot],s_0,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
+\item For each goal $g\in G[\cdot]$,
+\begin{enumerate}
+\item Build the classical planning problem $P^{\top}_g$, that constrains the solutions to $P=\tup{F,A[\cdot],s_0,g}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
 Note that $s_0$ is the initial state in the given observation $\mathcal{O}(\tau)$.
-We use our adapted compilation to compute the classcial planning tasks $P^{\top}_\lambda$ and solve them using an off-the-shelf-classical planner.
-\item For each goal, we define $P^{\bot}$, that constrains $P=\tup{F,A,s_0,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ and that uses the action model $A$ used by the corresponding solution $\pi^\top$.
-\item We compute the cost difference $\Delta(cost(\pi_\top),cost(\pi_\bot))$ where these costs are defined as the length of the postfix of the $\pi^{\top}_\lambda$ and $\pi^{\bot}_\lambda$ plans and plug this cost difference into equation (2) to get the $P(g|\mathcal{O})$ likelihoods.
-\item Finally the previous likelihoods are plugged into the Bayes rule from which the goal posterior probabilities are obtained.
+\item Solve $P^{\top}_g$ using the proposed compilation and a classical planner.
+Extract from this solution $cost(\pi^\top)$ (by counting the number of $\mathsf{apply_{\xi,\omega}}$ actions in the solution) and the action model that was determined by the insert actions to achieve the goals.
+\item Build the classical planning problem $P^{\bot}_g$ that constrains $P=\tup{F,A,s_0,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ and that uses the action model $A$ determined by the solution to the proposed compilation.
+\item Compute the cost difference $\Delta(cost(\pi_\top),cost(\pi_\bot))$ and plug this cost difference into equation (2) to get the $P(g|\mathcal{O})$ likelihoods.
+\end{enumerate}
+\item Plug the likelihoods into the Bayes rule from which the goal posterior probabilities are obtained.
 In this case the $P\mathcal{O}(\tau)$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
@@
 \label{sec:evaluation}
-The problem of {\em classical planning with unknown domain models} has been previously addressed for SAS+ action models~\cite{SternJ17}.
-In this work we formulate this task for \strips\ actoin models and evidence the relevance of this classcial planning setting since its allows to address the {\em goal recognition} task when the action model of the observed agent is not available (which typically is a too strong  assumption).
+The problem of {\em classical planning with unknown domain models} has been previously addressed for $SAS^+$ action models~\cite{SternJ17}.
+In this work we formulate this task for \strips\ action models and evidence its relevance for addressing {\em goal recognition} tasks when the action model of the observed agent is not available (which typically is a too strong  assumption for many real-world applications).
 
-Related work to {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}, where model edition is used to conform the PDDL models of two agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed optimal plan computed with one of the two models.
-{\em Model recognition}, however, conforms every input model $\mathcal{M}$ with another model $\mathcal{M}'$ that is not given as input but instead computed for every $\mathcal{M}$, and which is consistent with a partial observation of a plan execution.
+The paper also evidenced that {\em goal recognition}, when the domain Model is unknown, is very related to the task of learning a classical planning action model.
+With this regard the classical planning compilation for learning \strips\ action model is very suitable because it allows to produce a \strips\ action model from minimal input knowledge (and initial state, goals pair), and to refine this model if more knowledge is available (e.g.
+observation constraints)~\cite{aineto2018learning}.
+Most of the existing approaches for learning action models aim maximizing the statistical consistency of the learned model with respect to observations so require large amounts of input knowledge and do not produce model that are guaranteed to be {\em logically consistent} with the given input knowledge.
 

### 2019-01-30 doc/ijcai-19-goalRecognition/goalRecognition.tex
first comments on goal recognition design added to the ijcai19 paper on goal recognition

@@
 
-\section{Classical planning with unknown domain models}
+\section{Planning with unknown domain models}
 \label{sec:planning}
@@
 
+Planning with an unkown model is also related to {\em goal recognition design}~\cite{KerenGK14} where they compute the modification of the initial state that allow an efficient online goal recognition.
+Our approach is related since we are encoding the space of propositional schemas planning state variables, we encode at the initial state the {\em empty} actoin model with no preconditions and effects and provide actions to modify this model to enable the {\em goal recognition} process.
+
 \section{Conclusions}

### 2019-01-30 doc/ijcai-19-goalRecognition/goalRecognition.tex
first comments on goal recognition design added to the ijcai19 paper on goal recognition

--- 2019-01-29 doc/ijcai-19-goalRecognition/goalRecognition.tex
@@
 
-\subsubsection{Soundness and completeness}
-
 \begin{mylemma}
@@
 
+Our approach for {\em planning with an unknown domain model} is also related to {\em goal recognition design}~\cite{KerenGK14}.
+The reason is that we are encoding the space of propositional schemes as state variables of the planning problem (we encode the {\em empty} action model with no preconditions and no effects at the initial state) and provide actions to modify the value of this state variables (the action model) to enable the {\em goal recognition} process.
+With this regard {\em goal recognition design} applied to our approach would compute the action model that allows to distinguish the best among the possible goals to be recognized.
+
 \section{Conclusions}

### 2019-01-30 doc/ijcai-19-goalRecognition/goalRecognition.tex
updates added to the ijcai19 paper on goal recognition

@@
 \begin{abstract}
-This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach for {\em goal recognition} that is a priori having an action model of the observed agents.
+This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em knowing the action model of the observed agents}.
 The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are known) and it shows that this formulation neatly fits with the popular {\em plan recognition as planning} approach for {\em goal recognition}.
-The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks without {\em a priori} knowing the action model of the observed agents and using an off-the-shelf classical planner.
+The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without {\em a priori} knowing the action model of the observed agents.
 \end{abstract}
@@
 \label{sec:introduction}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal and the classification examples are observations of agents acting to achieve one of that goals.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents acting to achieve one of that goals.
 Despite there exists a wide range of different approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most appealing and it is currently at the core of various activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
@@
 {\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to compute the most likely goal of that agents.
-In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which frequently becomes a too strong assumption when using {\em plan recognition as planning} at real-world applications.
-In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are given) and it shows that this formulation neatly fits with the {\em plan recognition as planning} approach.
-The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks without {\em a priori} knowing the action model of the observed agents and using an off-the-shelf classical planner.
+In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which frequently becomes a too strong assumption when applying {\em plan recognition as planning} at real-world problems.
+In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are known) and it shows that this formulation neatly fits with the successful {\em plan recognition as planning} approach.
+The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without {\em a priori} knowing the action model of the observed agents.
 
@@
 \label{sec:background}
-This section formalizes the {\em planning model} we follow, the kind of {\em observations} that are given as input to the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.
+This section formalizes the {\em planning model} we follow, the kind of {\em observations} given as input to the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.
 
@@
 
-A classical planning action $a\in A$ has: a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $cost(a)$.
+A {\em classical planning action} $a\in A$ has: a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $cost(a)$.
 The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
@@
 A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length} and $cost(\pi)=\sum_{a\in\pi} cost(a)$ its {\em plan cost}.
-The execution of $\pi$ on the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,s_0)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
-A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,s_0)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
+The execution of $\pi$ on the initial state of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,P)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
 A solution plan is {\em optimal} iff its cost is minimal.
 
-An {\em action with conditional effects} $a_c\in A$ is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
+We also define {\em actions with conditional effects} because they are useful to compactly formulate our approach for {\em goal recognition with unknown domain models}.
+An action $a_c\in A$ with conditional effects is a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
 Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
@@
 
-%which indicates that we observe $l$ actions, $1\leq l\leq |\pi|$, and $m$ states, $1\leq m \leq |\pi|+1$, from $\tau(\pi,P)$:
-
 \begin{itemize}
 \item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
-Specifically, the number of observed actions, $l$, can range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
-\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
+Specifically, the number of observed actions, $l$, can range from $0$ (fully unobserved action sequence) to $|\pi|$ (fully observed action sequence).
+\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observed.
 A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
@@
 Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
-    %Exceptionally, $s_m^o$ cannot be fully unobservable for the purpose of our task (we will elaborate on this issue later on).
+In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each observed states comprises $[1,|F|]$ fluents.
 \end{itemize}
@@
 Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
+In other words, having an input observation $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
 
 \subsection{Goal recognition with classical planning}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal $g\in G[\cdot]$ and there is a single classification example $\mathcal{O}(\tau)$ that represents the observation of agents acting to achieve one of the input goals in $g\in G[\cdot]$.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents the observation of agents acting to achieve one of the input goals in $G\in G[\cdot]$.
 Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
 \begin{align}
-argmax_{g\in G[\cdot]} P(\mathcal{O}|g) P(g).
+argmax_{G\in G[\cdot]} P(\mathcal{O}|G) P(G).
 \end{align}
 
-The {\em plan recognition as planning} approach shows how to compute estimates of the $P(\mathcal{O}|g)$ likelihood leveraging the action model of the observed agent and an off-the-shelf classical planner~\cite{ramirez2012plan}.
-Given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$, where $G[\cdot]$ represents the set of possible goals $P(\mathcal{O}|g)$ is estimated computing, for each goal $g\in G[\cdot]$, the cost difference of the solution plans to these two different classical planning problems:
+The {\em plan recognition as planning} approach shows that the $P(\mathcal{O}|G)$ likelihood can be estimated leveraging the action model of the observed agents and an off-the-shelf classical planner~\cite{ramirez2012plan}.
+Given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$ (where $G[\cdot]$ represents the set of possible goals) then $P(\mathcal{O}|G)$ is estimated computing, for each goal $G\in G[\cdot]$, the cost difference of the solution plans to these two different classical planning problems:
 \begin{itemize}
-\item $P^{\top}_g$, the classical planning problem built constraining $P=\tup{F,A,I,g}$ to achieve the particular goal $g\in G[\cdot]$ through a plan $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-\item $P^{\bot}_g$, the classical planning problem that constrains the solutions of $P=\tup{F,A,I,g}$ to plans $\pi^\bot$ that achieve $g\in G[\cdot]$ but are {\em inconsistent} with $\mathcal{O}(\tau)$.
+\item $P^{\top}_G$, the classical planning problem built constraining $P=\tup{F,A,I,G}$ to achieve the particular goal $G\in G[\cdot]$ through a plan $\pi^\top$ that is {\em consistent} with the input observation $\mathcal{O}(\tau)$.
+\item $P^{\bot}_G$, the classical planning problem that constrains the solutions of $P=\tup{F,A,I,G}$ to plans $\pi^\bot$ that achieve $G\in G[\cdot]$ but are {\em inconsistent} with $\mathcal{O}(\tau)$.
 \end{itemize}
 
-The higher the value of this cost difference $cost(\pi^\top)-cost(\pi^\bot)$, the higher probability of aiming to achieve the $g\in G[\cdot]$ goal.
+The higher the value of the $cost(\pi^\top)-cost(\pi^\bot)$ difference, the higher is the probability of the observed agents to aim goal $G\in G[\cdot]$.
 With this regard, {\em plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:
@@
 \begin{align}
-P(\mathcal{O}|g) = \frac{1}{1+e^{-\beta(cost(\pi^\top)-cost(\pi^\bot))}}
+P(\mathcal{O}|G) = \frac{1}{1+e^{-\beta(cost(\pi^\top)-cost(\pi^\bot))}}
 \end{align}
 
-This expression is derived from the assumption that while the observed agent is not perfectly rational, he is more likely to follow cheaper plans, according to a {\em Boltzmann} distribution.
-The larger the value of $\beta$, the more rational the agent, and the less likely that he will follow suboptimal plans.
-Recent works show that estimates of the $P(\mathcal{O}|g)$ likelihood can be faster computed using relaxations of the classical planning tasks~\cite{pereira2017landmark}.
+This expression is derived from the assumption that while the observed agents are not perfectly rational, they are more likely to follow cheaper plans, according to a {\em Boltzmann} distribution.
+The larger the value of $\beta$, the more rational the agents, and the less likely that they will follow suboptimal plans.
+Recent works show fast estimates of the $P(\mathcal{O}|G)$ likelihood computed using relaxations of the classical planning tasks~\cite{pereira2017landmark}.
 
-The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions~\cite{ramirez2009plan}.
-However, the approach applies to more expressive observation models that consider also state observations, like the one defined above, with a trivial three-fold extension:
+The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions.
+However, the approach applies also to more expressive observation models that consider state observations as well, like the observation model defined above, with a trivial three-fold extension:
 \begin{itemize}
-\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j\in\mathcal{O}(\tau)$ state observation.
-\item Adding $validated_m$ to every possible goal $g\in G[\cdot]$ to constraint the solution plans $\pi^\top$ to be consistent with all the state observations.
-\item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
+\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation.
+\item Adding $validated_m$ to every possible goal $G\in G[\cdot]$ to constrain solution plans $\pi^\top$ to be consistent with all the state observations.
+\item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
 \end{itemize}
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{validated_{j-1}\},\\
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{validated_{j-1}\},\\
 \cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg validated_{j-1}, validated_j\}.
@@
 This section introduces a novel formulation for classical planning in a setting where no action model is given.
-This setting of classical planning is very related to the learning of action models for planning~\cite{SternJ17}.
-In particular it can be considered an extreme learning scenario when a model has to be learned from a single learning example that contains only two state observations, an initial state and the goals.
-
-A {\em classical planning with unknown domain models} is a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
+This setting of classical planning has already shown related to the learning of action models for planning~\cite{SternJ17}.
+In particular it can be seen an extreme scenario when the action model is learned from a single example that contains only two state observations: an initial state and the goals.
+A {\em classical planning with unknown domain models} is then a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
@@
 A solution to this task is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
-one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for each {\small $1\leq i\leq n$}, and such that the reached final state met the goal conditions, $G \subseteq s_n$.
+one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state met the goal conditions, $G \subseteq s_n$.
 
@@
 
-For instance, in the {\em blocksworld} the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for a {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for a {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+For instance in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for the {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
@@
 In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\ so is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning classical planning action models.
 In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
@@
 \subsection{A classical planning compilation for planning with unknown domain models}
-Now we show how we adapt the classical planning compilation~\cite{aineto2018learning} to our propositional encoding.
-In more detail, given a classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$ we create another classical planning problem $P'=\tup{F',A',I,G}$ such that:
+Now we show how we adapt the classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} to our propositional encoding for addressing the task of planning with unknown domain models.
+
+Given a classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$ we create another classical planning problem $P'=\tup{F',A',I,G}$ such that:
 \begin{itemize}
 \item $F'$ extends $F$ with the fluents for the propositional encoding of the corresponding space of STRIPS action models.
-This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding parameters $pars(\xi)$.
+This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding action parameters $pars(\xi)$.
 
-\item $A'$ replaces the actions in $A$ with two new types of actions.
+\item $A'$ replaces the actions in $A$ with two types of actions.
 \begin{enumerate}
-\item Actions for {\em inserting} a component (precondition, positive effect or negative effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models. 
+\item Actions for {\em inserting} a component (precondition, positive effect or negative effect) in $\xi$ following the syntactic constraints of \strips\ models. 
 \begin{itemize}
-\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
-A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
+\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi$. A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
 
@@
 
-\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
+\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi$.
 
@@
 
-\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{ar(\xi)}$.
+\item Actions for {\em applying} an action model $\xi$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$, where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
 Since action headers are known, the variables $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position.
@@
 
-The dynamics of the actions for {\em applying} an action model $\xi\in\mathcal{M}$ is determined by the values of the model the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in the current state.
-For instance, executing the action {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$.
-This means that if the current state $s$ holds {\scriptsize$\{${\tt (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}$\} \subset s$}, then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
-Otherwise, a different set of precondition literals will be checked.
+The intuition of the compilation is that the dynamics of the actions for {\em applying} an action model $\xi$ is determined by the values of the model the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in the current state.
+For instance, executing {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in $s$.
+This means that if the current state $s$ holds $\{{\tt\scriptsize (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}\} \subset s$, then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
+Otherwise, a different set of precondition literals will be checked for the stack action.
 The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.
-Note that executing {\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if {\tt{\small stack}} has been correctly programmed by the insert actions.
+Executing {\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state only if {\tt{\small stack}} has been correctly programmed by the {\em insert} actions.
 
@@
 
-The plan of Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving the classical planning problem output by our compilation for solving a classical planning with unknown domain models.
-In the initial state of that problem the robot is holding the blockA while the blockB is clear and the problem goal is having blockA on blockB.
-The plan shows the insert actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the action models to achieve the goals.
-Note that an equivalent solution could be found by inserting the same preconditions and effects into the {\em unstack} action model and then applying action (unstack blockA blockB) instead of (stack blockA blockB).
+Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving the classical planning problem output by our compilation for solving a planning problem with unknown domain models.
+In the initial state of that problem the robot is holding {\tt blockA} while {\tt blockB} is clear and the problem goals are having {\tt blockA} on top of {\tt blockB}.
+The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the programmed action models to achieve the goals.
+Note that an equivalent solution could be found by inserting the same preconditions and effects into the {\em unstack} action model and then applying action {\tt\small (unstack blockA blockB)} instead of {\tt\small(stack blockA blockB)}.
 
@@
 The reason is that we are encoding the space of propositional schemes as state variables of the planning problem (we encode the {\em empty} action model with no preconditions and no effects at the initial state) and provide actions to modify the value of this state variables (the action model) to enable the {\em goal recognition} process.
-With this regard {\em goal recognition design} applied to our approach would compute the action model that allows to distinguish the best among the possible goals to be recognized.
+With this regard, {\em goal recognition design} applied to {\em goal recognition with unknown domain models} would compute the action model, in the space of possible models, that reveals any of the possible goals as early as possible.
 

### 2019-01-31 doc/ijcai-19-goalRecognition/goalRecognition.tex
updates added to the ijcai19 paper on goal recognition

@@
 
-The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions.
-However, the approach applies also to more expressive observation models that consider state observations as well, like the observation model defined above, with a trivial three-fold extension:
-\begin{itemize}
-\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation.
-\item Adding $validated_m$ to every possible goal $G\in G[\cdot]$ to constrain solution plans $\pi^\top$ to be consistent with all the state observations.
-\item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
-\end{itemize}
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{validated_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg validated_{j-1}, validated_j\}.
-\end{align*}
-\end{small}
-
 \section{Planning with unknown domain models}
@@
 \subsection{A classical planning compilation for planning with unknown domain models}
-Now we show how we adapt the classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} to our propositional encoding for addressing the task of planning with unknown domain models.
+Now we show how we adapt the classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} for addressing the task of planning with unknown domain models with our propositional encoding of STRIPS action models.
 
-Given a classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$ we create another classical planning problem $P'=\tup{F',A',I,G}$ such that:
+Given a classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$ we create a classical planning problem $P'=\tup{F',A',I,G}$ such that:
 \begin{itemize}
@@
 
-\item Actions for {\em applying} an action model $\xi$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$, where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-Since action headers are known, the variables $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position.
+\item Actions for {\em applying} an action model $\xi$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
+Since action headers are known, $pars(\xi)$ are bound to the objects in $\omega$ that appear in the same position.
 
@@
 
-The intuition of the compilation is that the dynamics of the actions for {\em applying} an action model $\xi$ is determined by the values of the model the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in the current state.
-For instance, executing {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in $s$.
+The intuition of the compilation is that the dynamics of the actions for {\em applying} an action model $\xi$ is determined by the values of the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ in the current state.
+For instance, executing {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of $\{pre\_e\_stack, eff\_e\_stack\}_{\forall e\in{\mathcal I}_{\Psi,stack}}$ fluents in $s$.
 This means that if the current state $s$ holds $\{{\tt\scriptsize (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}\} \subset s$, then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
 Otherwise, a different set of precondition literals will be checked for the stack action.
-The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.
+The same applies to the positive and negative effects.
 Executing {\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state only if {\tt{\small stack}} has been correctly programmed by the {\em insert} actions.
@@
 
-Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving the classical planning problem output by our compilation for solving a planning problem with unknown domain models.
-In the initial state of that problem the robot is holding {\tt blockA} while {\tt blockB} is clear and the problem goals are having {\tt blockA} on top of {\tt blockB}.
-The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the programmed action models to achieve the goals.
-Note that an equivalent solution could be found by inserting the same preconditions and effects into the {\em unstack} action model and then applying action {\tt\small (unstack blockA blockB)} instead of {\tt\small(stack blockA blockB)}.
+Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving a $P'=\tup{F',A',I,G}$ classical planning problem output by our compilation.
+In the initial state of that problem the robot is holding {\tt blockA} while {\tt blockB} is clear and the single problem goal is having {\tt blockA} on top of {\tt blockB}.
+The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the programmed action model to achieve the goals $G$.
+Note that another valid solution could be computed for instance by inserting the same preconditions and effects into the {\em unstack} action model and then applying action {\tt\small (unstack blockA blockB)} instead of {\tt\small(stack blockA blockB)}.
 
@@
 \begin{small}
-Once a given precondition or effect is inserted into an action model it can never be removed back.
-In addition, $P'$ is only solvable iff goals $G$ hold at the last state reached by $\pi'$.
+Once a given precondition or effect is inserted into an action model it can never be removed back and once an action model is applied it cannot be {\em reprogramed}.
 In the compiled problem the value of $F$, the fluents of the original problem, can exclusively be modified via $\mathsf{apply_{\xi,\omega}}$ actions.
@@
 \begin{small}
- ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in an action model $\xi$ using the set of predicates $\Psi$.
-The compilation does not discard any possible action model definable within ${\mathcal I}_{\Psi,\xi}$.
+By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
+Furthermore, the compilation does not discard any possible action model definable within ${\mathcal I}_{\Psi,\xi}$.
 This means that for every plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, we can build a plan $\pi'$ selecting the appropriate actions for inserting precondition and effects to the corresponding action model and then selecting the  $\mathsf{apply_{\xi,\omega}}$ actions that allows to transform the initial state $I$ into a state $G \subseteq s_n$.
@@
 
-\subsubsection{The bias of the initially empty action model}
-Given that at the initial state of the classical planning compilation any $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluent is false, our compilation introduces a bias to solutions of the $P=\tup{F,A[\cdot],I,G}$ classical planning task.
+\subsubsection{The bias of an initially {\em empty} action model}
+Given that at the initial state of the classical planning compilation any $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluent is false, our compilation can introduce a bias to the solutions of the $P=\tup{F,A[\cdot],I,G}$ classical planning task preferring solutions that imply action models with a smaller number of preconditions and effects (i.e., a les number of {\em insert} actions).
 
-This bias might be eliminated defining a cost landscape where any actions for {\em inserting} a precondition or an effect into a given action model has a zero cost while the remaining actions, $\mathsf{apply_{\xi,\omega}}$ actions has unit cost.
-In practice, since classical planners are not proficiency optimizing the cost of solutions with cost landscapes of this kind, we use a different approach that disregards the cost of the actions that insert a precondition/effect to an action model.
-
+This bias might be eliminated defining a cost landscape where {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has {\em positive constant cost}.
+In practice, since classical planners are not proficiency optimizing solutions with cost landscapes of this kind, we use a different approach to disregard the cost of actions that insert preconditions/effects.
 Our approach is to use SAT-based planning because it can apply all the required actions for inserting preconditions in a single planning step (in parallel), because these actions do not interact.
-Actions for programming action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which in addition significantly reduces the planning horizon.
-The SAT-based planning approach is also convenient because its ability to deal with planning problems populated with dead-ends and because the symmetries in the insertion of preconditions/effects into the action model do not affect the performance of SAT-based planning.
+Actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
+The SAT-based planning approach is also convenient because its ability to deal with populated with dead-ends and because symmetries in the insertion of preconditions/effects into the action models do not affect to performance.
 
 \subsubsection{Compilation size}
-The size of the classical planning task $P'$ output by the compilation approach depends on the arity of the {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the arity of the action headers, given by $A[\cdot]$.
-The larger the arity, the larger the size of the ${\mathcal I}_{\Psi,\xi}$ sets.
-This is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
+The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the arity of the action models, given by $A[\cdot]$.
+The larger these arities, the larger the size of the ${\mathcal I}_{\Psi,\xi}$ sets.
+This is the term that dominates the compilation size it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
 
-Note that for planning models that allow object typing, types can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ significantly reducing the size of the classical planning task output by the compilation.
+Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces the size of the classical planning task output by the compilation.
 
@@
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that reaches a goal $g\in G[\cdot]$ starting from the given initial state.
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that reaches a goal $G\in G[\cdot]$ starting from the given initial state.
 \end{itemize}
@@
 
-\subsection{Estimating $P(\mathcal{O}|g)$ with classical planning}
-Now we are ready to compute an estimate of the target probability distribution $P(g|\mathcal{O})$ over the possible goals $g\in G[\cdot]$ given an observation $\mathcal{O}(\tau)$:
+\subsection{Computing estimates of the $P(\mathcal{O}|G)$ likelyhood with classical planning}
+Now we are ready to build an estimate of the $P(\mathcal{O}|G)$ likelyhood.
+Our mechanism is the same as described in the {\em plan recognition as planning} approach~\cite{ramirez2009plan,ramirez2012plan} except that we compute $cost(\pi^\top)$ using now our compilation for {\em classical planning with unknown domain models}.
+
+In more detail, we build the estimate of the $P(\mathcal{O}|G)$ likelyhood following these four steps:
 \begin{enumerate}
-\item For each goal $g\in G[\cdot]$,
-\begin{enumerate}
-\item Build the classical planning problem $P^{\top}_g$, that constrains the solutions to $P=\tup{F,A[\cdot],s_0,g}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-Note that $s_0$ is the initial state in the given observation $\mathcal{O}(\tau)$.
-\item Solve $P^{\top}_g$ using the proposed compilation and a classical planner.
-Extract from this solution $cost(\pi^\top)$ (by counting the number of $\mathsf{apply_{\xi,\omega}}$ actions in the solution) and the action model that was determined by the insert actions to achieve the goals.
-\item Build the classical planning problem $P^{\bot}_g$ that constrains $P=\tup{F,A,s_0,g}$ to achieve $g\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ and that uses the action model $A$ determined by the solution to the proposed compilation.
-\item Compute the cost difference $\Delta(cost(\pi_\top),cost(\pi_\bot))$ and plug this cost difference into equation (2) to get the $P(g|\mathcal{O})$ likelihoods.
+\item {\em Build $P^{\top}_G$}, the classical planning problem that constrains solutions of the problem $P=\tup{F,A[\cdot],s_0^o,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
+Note that $s_0^o\in \mathcal{O}(\tau)$ is the initial state in the given observation.
+\item {\em Solve $P^{\top}_G$}, using the proposed compilation for classical planning with unknown domain models.
+Extract from this solution (1), $cost(\pi^\top)$ (by counting the number of $\mathsf{apply_{\xi,\omega}}$ actions in the solution) and (2), the action model $A$ that is determined by the insert actions used to achieve the goals.
+\item {\em Build $P^{\bot}_G$}, the classical planning problem that constrains $P=\tup{F,A,s_0^o,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ (where $A$ is the set of actions defined by the solution to the $P^{\top}_G$ problem).
+\item {\em Solve $P^{\bot}_G$} with a classical planner and extract $cost(\pi_\bot)$ as the length of the found solution plan.
+\item Compute the $cost(\pi^\top)-cost(\pi^\bot)$ difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
 \end{enumerate}
-\item Plug the likelihoods into the Bayes rule from which the goal posterior probabilities are obtained.
+
+To compute the target probability distribution $P(G|\mathcal{O})$ plug the $P(\mathcal{O}|G)$ likelihoods into the {\em Bayes rule} from which the goal posterior probabilities are obtained.
 In this case the $P\mathcal{O}(\tau)$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
-\end{enumerate}
+
+\subsection{Extending the observation model of {\em plan recognition as planning}}
+The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions.
+However, the approach applies also to more expressive observation models that consider state observations as well, like the observation model defined above, with a simple three-fold extension:
+\begin{itemize}
+\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation.
+\item Adding $validated_m$ to every possible goal $G\in G[\cdot]$ to constrain solution plans $\pi^\top$ to be consistent with all the state observations.
+\item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
+\end{itemize}
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{validated_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg validated_{j-1}, validated_j\}.
+\end{align*}
+\end{small}
 
@@
 \label{sec:evaluation}
-The problem of {\em classical planning with unknown domain models} has been previously addressed for $SAS^+$ action models~\cite{SternJ17}.
-In this work we formulate this task for \strips\ action models and evidence its relevance for addressing {\em goal recognition} tasks when the action model of the observed agent is not available (which typically is a too strong  assumption for many real-world applications).
+The problem of {\em classical planning with unknown domain models} has been previously addressed~\cite{SternJ17}.
+In this work we evidence the relevance of this task for addressing {\em goal recognition} tasks when the action model of the observed agent is not available (which typically is a too strong  assumption for many applications).
 
-The paper also evidenced that {\em goal recognition}, when the domain Model is unknown, is very related to the task of learning a classical planning action model.
-With this regard the classical planning compilation for learning \strips\ action model is very suitable because it allows to produce a \strips\ action model from minimal input knowledge (and initial state, goals pair), and to refine this model if more knowledge is available (e.g.
-observation constraints)~\cite{aineto2018learning}.
-Most of the existing approaches for learning action models aim maximizing the statistical consistency of the learned model with respect to observations so require large amounts of input knowledge and do not produce model that are guaranteed to be {\em logically consistent} with the given input knowledge.
+The paper also showed that {\em goal recognition}, when the domain model is unknown, is closely related to the learning a planning action model.
+With this regard, the classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} is very appealing because it allows to produce a \strips\ action model from minimal input knowledge (a singel initial state, goals pair), and to refine this model whenever more knowledge is available (e.g.
+observation constraints).
+Most of the existing approaches for learning action models aim maximizing an statistical consistency of the learned model with respect to the input observations so require large amounts of input knowledge and do not produce action models that are guaranteed to be {\em logically consistent} with the given input knowledge.
 
-Our approach for {\em planning with an unknown domain model} is also related to {\em goal recognition design}~\cite{KerenGK14}.
-The reason is that we are encoding the space of propositional schemes as state variables of the planning problem (we encode the {\em empty} action model with no preconditions and no effects at the initial state) and provide actions to modify the value of this state variables (the action model) to enable the {\em goal recognition} process.
-With this regard, {\em goal recognition design} applied to {\em goal recognition with unknown domain models} would compute the action model, in the space of possible models, that reveals any of the possible goals as early as possible.
+Our approach for {\em planning with an unknown domain model} is related to {\em goal recognition design}~\cite{KerenGK14}.
+The reason is that we are encoding the space of propositional schemes as state variables of the planning problem (the initial state encodes the {\em empty} action model with no preconditions and no effects) and provide actions to modify the value of this state variables an in a {\em goal recognition design} task.
+The aims of {\em goal recognition design} are however different.
+{\em Goal recognition design} applied to {\em goal recognition with unknown domain models} would compute the action model, in the space of possible models, that allows to reveal any of the possible goals as early as possible.
 

### 2019-01-31 doc/ijcai-19-goalRecognition/goalRecognition.tex
first version of the ijcai19 paper on goal recognition with unknown domain models (no experiments)

@@
 This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em knowing the action model of the observed agents}.
-The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are known) and it shows that this formulation neatly fits with the popular {\em plan recognition as planning} approach for {\em goal recognition}.
+The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the {\em plan recognition as planning} approach for {\em goal recognition}.
 The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without {\em a priori} knowing the action model of the observed agents.
@@
 \label{sec:introduction}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents acting to achieve one of that goals.
-Despite there exists a wide range of different approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most appealing and it is currently at the core of various activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents acting to achieve one of those goals.
+Despite there exists a wide range of different approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
 
@@
 In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which frequently becomes a too strong assumption when applying {\em plan recognition as planning} at real-world problems.
-In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action headers are known) and it shows that this formulation neatly fits with the successful {\em plan recognition as planning} approach.
+In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the successful {\em plan recognition as planning} approach.
 The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without {\em a priori} knowing the action model of the observed agents.
@@
 \label{sec:background}
-This section formalizes the {\em planning model} we follow, the kind of {\em observations} given as input to the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.
+This section formalizes the {\em planning model} we follow, the kind of {\em observations} that input the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.
 
@@
 \item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
-Specifically, the number of observed actions, $l$, can range from $0$ (fully unobserved action sequence) to $|\pi|$ (fully observed action sequence).
+The number of observed actions, $l$, ranges from $0$ (fully unobserved action sequence) to $|\pi|$ (fully observed action sequence).
 \item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observed.
@@
 Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each observed states comprises $[1,|F|]$ fluents.
+The number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each {\em observed} states comprises $[1,|F|]$ fluents (the observation can still miss intermediate states that are {\em unobserved}).
 \end{itemize}
@@
 \subsection{Goal recognition with classical planning}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents the observation of agents acting to achieve one of the input goals in $G\in G[\cdot]$.
+{\em Goal recognition} is a particular classification task in which each class represents a different possible goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents the observation of agents acting to achieve a goal $G\in G[\cdot]$.
+
 Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
@@
 The {\em plan recognition as planning} approach shows that the $P(\mathcal{O}|G)$ likelihood can be estimated leveraging the action model of the observed agents and an off-the-shelf classical planner~\cite{ramirez2012plan}.
-Given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$ (where $G[\cdot]$ represents the set of possible goals) then $P(\mathcal{O}|G)$ is estimated computing, for each goal $G\in G[\cdot]$, the cost difference of the solution plans to these two different classical planning problems:
+Given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$ (where $G[\cdot]$ represents the set of {\em recognizable} goals) then $P(\mathcal{O}|G)$ is estimated computing, for each goal $G\in G[\cdot]$, the cost difference of the solution plans to these two classical planning problems:
 \begin{itemize}
 \item $P^{\top}_G$, the classical planning problem built constraining $P=\tup{F,A,I,G}$ to achieve the particular goal $G\in G[\cdot]$ through a plan $\pi^\top$ that is {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-\item $P^{\bot}_G$, the classical planning problem that constrains the solutions of $P=\tup{F,A,I,G}$ to plans $\pi^\bot$ that achieve $G\in G[\cdot]$ but are {\em inconsistent} with $\mathcal{O}(\tau)$.
+\item $P^{\bot}_G$, the classical planning problem that constrains solutions of $P=\tup{F,A,I,G}$ to plans $\pi^\bot$, that achieve $G\in G[\cdot]$, but that are {\em inconsistent} with $\mathcal{O}(\tau)$.
 \end{itemize}
 
-The higher the value of the $cost(\pi^\top)-cost(\pi^\bot)$ difference, the higher is the probability of the observed agents to aim goal $G\in G[\cdot]$.
+The higher the value of the $cost(\pi^\top)-cost(\pi^\bot)$ difference, the higher the probability of the observed agents to aim goal $G\in G[\cdot]$.
 With this regard, {\em plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:
@@
 This section introduces a novel formulation for classical planning in a setting where no action model is given.
-This setting of classical planning has already shown related to the learning of action models for planning~\cite{SternJ17}.
-In particular it can be seen an extreme scenario when the action model is learned from a single example that contains only two state observations: an initial state and the goals.
+This setting has already shown related to the learning of action models for planning~\cite{SternJ17}.
+In particular it can be seen as an extreme scenario when the action model is learned from a single example that contains only two state observations: the initial state and the goals.
 A {\em classical planning with unknown domain models} is then a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
@@
 A solution to this task is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
-one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state met the goal conditions, $G \subseteq s_n$.
+one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state meets the goal conditions, $G \subseteq s_n$.
 
 Next we show that the space of possible STRIPS action models can be encoded as a set of propositional variables and a set of constraints over those variables.
-Then, we show how to exploit this encoding to solve $P=\tup{F,A[\cdot],I,G}$ problems with an off-the-shelf classical planner as well as the properties of this approach.
+Then, we show how to exploit this encoding to solve $P=\tup{F,A[\cdot],I,G}$ problems with an off-the-shelf classical planner and the properties of this approach.
 
@@
 \begin{enumerate}
-\item {\em Syntactic constraints}.
+\item {\bf Syntactic constraints}.
 \strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
@@
 {\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
-\item {\em Domain-specific constraints}.
+\item {\bf Domain-specific constraints}.
 One can introduce domain-specific knowledge to constrain further the space of possible schemata.
@@
 {\it State invariants} are also constraints of this kind~\cite{fox1998automatic}.
-\item {\em Observation constraints}. An observations $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
+\item {\bf Observation constraints}. An observation $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
 \end{enumerate}
@@
 This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning classical planning action models.
-In more detail, if {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
+In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
+If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
 Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
@@
 \subsection{A classical planning compilation for planning with unknown domain models}
-Now we show how we adapt the classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} for addressing the task of planning with unknown domain models with our propositional encoding of STRIPS action models.
+Now we show how we adapt the {\em classical planning compilation for learning \strips\ action models}~\cite{aineto2018learning} to address the task of {\em planning with unknown domain models}, using our propositional encoding of STRIPS action models.
 
@@
 \begin{itemize}
-\item $F'$ extends $F$ with the fluents for the propositional encoding of the corresponding space of STRIPS action models.
+\item $F'$ extends $F$ with a fluent $mode_{insert}$, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
 This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding action parameters $pars(\xi)$.
@@
 \begin{enumerate}
-\item Actions for {\em inserting} a component (precondition, positive effect or negative effect) in $\xi$ following the syntactic constraints of \strips\ models. 
+\item Actions for {\em inserting} a {\em precondition}, {\em positive} effect or {\em negative} effect in $\xi$ following the syntactic constraints of \strips\ models. 
 \begin{itemize}
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi)\},\\
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi), mode_{insert}\},\\
 \cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p}(\xi)\},\\
+\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p}(\xi), mode_{insert}\},\\
 \cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff_{p}(\xi)\}.
@@
 
-\item Actions for {\em applying} an action model $\xi$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
-Since action headers are known, $pars(\xi)$ are bound to the objects in $\omega$ that appear in the same position.
+\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
+The action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
+\end{enumerate}
+\end{itemize}
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
+\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
 \cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
-&\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\}.
+&\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\},\\
+&\{\emptyset\}\rhd\{mode_{insert}\}.
 \end{align*}
@@
 
-\end{enumerate}
-\end{itemize}
-
-The intuition of the compilation is that the dynamics of the actions for {\em applying} an action model $\xi$ is determined by the values of the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ in the current state.
-For instance, executing {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of $\{pre\_e\_stack, eff\_e\_stack\}_{\forall e\in{\mathcal I}_{\Psi,stack}}$ fluents in $s$.
+The intuition of the compilation is that the dynamics of the actions for {\em applying} an action model $\xi$ is determined by the values of the corresponding $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in the current state.
+For instance, executing {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small apply\_stack}} according to the values of $\{pre\_e\_stack, eff\_e\_stack\}_{\forall e\in{\mathcal I}_{\Psi,stack}}$ fluents in $s$.
 This means that if the current state $s$ holds $\{{\tt\scriptsize (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}\} \subset s$, then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
@@
 	}
-	\caption{\small Plan computed when solving the classical planning problem output by our compilation for solving a classical planning with unknown domain models.}
+	\caption{\small Plan computed when solving the classical planning problem output by our compilation corresponding to a classical planning with unknown domain models.}
 	\label{fig:plan-lplan}
@@
 The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the programmed action model to achieve the goals $G$.
-Note that another valid solution could be computed for instance by inserting the same preconditions and effects into the {\em unstack} action model and then applying action {\tt\small (unstack blockA blockB)} instead of {\tt\small(stack blockA blockB)}.
+Note that another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em unstack} action model and then applying the {\tt\small (unstack blockA blockB)} action instead of {\tt\small(stack blockA blockB)}.
 
 \subsection{Compilation properties}
-Now we present the properties of the compilation scheme.
+Now we present some properties of the compilation scheme.
 
@@
 This means that the action model used by the $\mathsf{apply_{\xi,\omega}}$ actions has to be consistent with the traversed intermediate states.
-We know that this is true by the definition of the $\mathsf{apply_{\xi,\omega}}$ so hence, the sub-sequence of $\mathsf{apply_{\xi,\omega}}$ appearing in $\pi'$ to solve $P'$ is a solution plan to $P=\tup{F,A[\cdot],I,G}$.
+We know that this must be true by the definition of the $\mathsf{apply_{\xi,\omega}}$ so hence, the sub-sequence of $\mathsf{apply_{\xi,\omega}}$ appearing in $\pi'$ to solve $P'$ is a solution plan to $P=\tup{F,A[\cdot],I,G}$.
 \end{small}
@@
 Furthermore, the compilation does not discard any possible action model definable within ${\mathcal I}_{\Psi,\xi}$.
-This means that for every plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, we can build a plan $\pi'$ selecting the appropriate actions for inserting precondition and effects to the corresponding action model and then selecting the  $\mathsf{apply_{\xi,\omega}}$ actions that allows to transform the initial state $I$ into a state $G \subseteq s_n$.
+This means that for every plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, we can build a plan $\pi'$ by selecting the appropriate actions for inserting precondition and effects to the corresponding action model and then selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state $I$ into a state $G \subseteq s_n$.
 \end{small}
@@
 \subsubsection{The bias of an initially {\em empty} action model}
-Given that at the initial state of the classical planning compilation any $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluent is false, our compilation can introduce a bias to the solutions of the $P=\tup{F,A[\cdot],I,G}$ classical planning task preferring solutions that imply action models with a smaller number of preconditions and effects (i.e., a les number of {\em insert} actions).
+Any $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluent is false at the initial state of our compilation.
+This fact can introduce a bias to the solutions of the $P=\tup{F,A[\cdot],I,G}$ classical planning task preferring solutions that imply action models with a smaller number of {\em preconditions}/{\em effects} (i.e., that imply a lower number of {\em insert} actions).
 
-This bias might be eliminated defining a cost landscape where {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has {\em positive constant cost}.
-In practice, since classical planners are not proficiency optimizing solutions with cost landscapes of this kind, we use a different approach to disregard the cost of actions that insert preconditions/effects.
+This bias might be eliminated defining a cost landscape where {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}.
+In practice, since classical planners are not proficiency optimizing cost landscapes of this kind, we use a different approach that disregard the cost of actions the {\em insert} actions.
 Our approach is to use SAT-based planning because it can apply all the required actions for inserting preconditions in a single planning step (in parallel), because these actions do not interact.
-Actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
-The SAT-based planning approach is also convenient because its ability to deal with populated with dead-ends and because symmetries in the insertion of preconditions/effects into the action models do not affect to performance.
+Further, actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
+The SAT-based planning approach is also convenient because its ability to deal with populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to performance.
 
 \subsubsection{Compilation size}
-The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the arity of the action models, given by $A[\cdot]$.
-The larger these arities, the larger the size of the ${\mathcal I}_{\Psi,\xi}$ sets.
-This is the term that dominates the compilation size it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
+The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
+The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$.
+The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
 
-Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces the size of the classical planning task output by the compilation.
+Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\Psi,\xi}|$ and hence, the size of the classical planning task output by the compilation.
 
@@
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a classical planning problem where $G[\cdot]$ is the set of possible goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
+\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a classical planning problem where $G[\cdot]$ is the set of {\em recognizable} goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that reaches a goal $G\in G[\cdot]$ starting from the given initial state.
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that reaches the goals $G\in G[\cdot]$ starting from the initial state $I$ in $P$.
 \end{itemize}
@@
 
-\subsection{Computing estimates of the $P(\mathcal{O}|G)$ likelyhood with classical planning}
+\subsection{Estimating the $P(\mathcal{O}|G)$ likelyhood with unknown domain models}
 Now we are ready to build an estimate of the $P(\mathcal{O}|G)$ likelyhood.
-Our mechanism is the same as described in the {\em plan recognition as planning} approach~\cite{ramirez2009plan,ramirez2012plan} except that we compute $cost(\pi^\top)$ using now our compilation for {\em classical planning with unknown domain models}.
+Our mechanism matches the {\em plan recognition as planning} approach~\cite{ramirez2012plan} except that we compute $cost(\pi^\top)$ using our compilation for {\em classical planning with unknown domain models}.
 
@@
 Note that $s_0^o\in \mathcal{O}(\tau)$ is the initial state in the given observation.
-\item {\em Solve $P^{\top}_G$}, using the proposed compilation for classical planning with unknown domain models.
-Extract from this solution (1), $cost(\pi^\top)$ (by counting the number of $\mathsf{apply_{\xi,\omega}}$ actions in the solution) and (2), the action model $A$ that is determined by the insert actions used to achieve the goals.
-\item {\em Build $P^{\bot}_G$}, the classical planning problem that constrains $P=\tup{F,A,s_0^o,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ (where $A$ is the set of actions defined by the solution to the $P^{\top}_G$ problem).
+\item {\em Solve $P^{\top}_G$}, using the proposed compilation for {\em classical planning with unknown domain models}.
+Extract from this solution (1), $cost(\pi^\top)$ (by counting the number of $\mathsf{apply_{\xi,\omega}}$ actions in the solution) but also (2), the action model $A$ that is determined by the {\em insert} actions used in $\pi^\top$ to achieve the goals $G$.
+\item {\em Build $P^{\bot}_G$}, the classical planning problem that constrains $P=\tup{F,A,s_0^o,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ (where $A$ is the set of actions extracted in step 2.).
 \item {\em Solve $P^{\bot}_G$} with a classical planner and extract $cost(\pi_\bot)$ as the length of the found solution plan.
@@
 To compute the target probability distribution $P(G|\mathcal{O})$ plug the $P(\mathcal{O}|G)$ likelihoods into the {\em Bayes rule} from which the goal posterior probabilities are obtained.
-In this case the $P\mathcal{O}(\tau)$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
+In this case the $P(\mathcal{O})$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
 
@@
 The problem of {\em classical planning with unknown domain models} has been previously addressed~\cite{SternJ17}.
-In this work we evidence the relevance of this task for addressing {\em goal recognition} tasks when the action model of the observed agent is not available (which typically is a too strong  assumption for many applications).
+In this work we evidence the relevance of this task for addressing {\em goal recognition} when the action model of the observed agent is not available (which it is typically a too strong assumption at many real-world applications).
 
-The paper also showed that {\em goal recognition}, when the domain model is unknown, is closely related to the learning a planning action model.
-With this regard, the classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} is very appealing because it allows to produce a \strips\ action model from minimal input knowledge (a singel initial state, goals pair), and to refine this model whenever more knowledge is available (e.g.
+The paper also showed that {\em goal recognition}, when the domain model is unknown, is closely related to the learning of planning action models.
+With this regard, the classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} is very appealing because it allows to produce a \strips\ action model from minimal input knowledge (a single initial state and goals pair), and to refine this model if more input knowledge is available (e.g.
 observation constraints).
@@
 Our approach for {\em planning with an unknown domain model} is related to {\em goal recognition design}~\cite{KerenGK14}.
-The reason is that we are encoding the space of propositional schemes as state variables of the planning problem (the initial state encodes the {\em empty} action model with no preconditions and no effects) and provide actions to modify the value of this state variables an in a {\em goal recognition design} task.
+The reason is that we are encoding the space of propositional schemes as state variables of the planning problem (the initial state encodes the {\em empty} action model with no preconditions and no effects) and provide actions to modify the value of this state variables as in {\em goal recognition design}.
 The aims of {\em goal recognition design} are however different.

### 2019-01-31 doc/ijcai-19-goalRecognition/goalRecognition.tex
first version of the ijcai19 paper on goal recognition with unknown domain models (no experiments)

@@
 &\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\},\\
-&\{\emptyset\}\rhd\{mode_{insert}\}.
+&\{\emptyset\}\rhd\{\neg mode_{insert}\}.
 \end{align*}

### 2019-02-04 doc/ijcai-19-deductivelearning/deductivelearning.tex
first commit ijcai19 paper on deductive learning

--- 2019-01-31 doc/ijcai-19-goalRecognition/goalRecognition.tex
@@
 \usepackage{amsthm}
+\usepackage{amssymb}
 \usepackage{listings}% http://ctan.org/pkg/listings
@@
 
-\title{Goal Recognition as Planning with Unknown Domain Models}
+\title{Learning \strips\ action models from {\em state-invaraints}}
 
@@
 Sergio Jim\'enez$^1$\and
-Eva Onaindia$^1$\And
-\and
-Miquel Ram\'irez$^2$
+Eva Onaindia$^1$
 \affiliations
-$^1${\small Departamento de Sistemas Inform\'aticos y Computaci\'on. Universitat Polit\`ecnica de Val\`encia. Valencia, Spain}\\
-$^2${\small School of Computing and Information Systems. The University of Melbourne. Melbourne, Victoria. Australia}
+$^1${\small Departamento de Sistemas Inform\'aticos y Computaci\'on. Universitat Polit\`ecnica de Val\`encia. Valencia, Spain}
 \emails
@@
 \begin{abstract}
-This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em knowing the action model of the observed agents}.
-The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the {\em plan recognition as planning} approach for {\em goal recognition}.
-The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without {\em a priori} knowing the action model of the observed agents.
 \end{abstract}
@@
 \label{sec:introduction}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents acting to achieve one of those goals.
-Despite there exists a wide range of different approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
-
-{\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to compute the most likely goal of that agents.
-In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which frequently becomes a too strong assumption when applying {\em plan recognition as planning} at real-world problems.
-In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the successful {\em plan recognition as planning} approach.
-The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without {\em a priori} knowing the action model of the observed agents.
+{\em Classical planing} is an interesting approach for learning \strips\ action models since it is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states)~\cite{aineto2018learning}.
+This paper shows that this flexibility of the compilation approach goes one step further since it also allows learning from {\em state-invariants}, logic formulae that constrain the set of possible states of a given domain.
+The experimental results show that {\em state-invariants} boost the performance of the {\em classical planing} compilation for learning \strips\ action models.
 
@@
 \label{sec:background}
-This section formalizes the {\em planning model} we follow, the kind of {\em observations} that input the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.
+This section formalizes the {\em classical planning model} we follow in this work and the kind of {\em knowledge} that can be given as input to the task of learning \strips\ action models.
 
@@
 
-\subsection{Goal recognition with classical planning}
-{\em Goal recognition} is a particular classification task in which each class represents a different possible goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents the observation of agents acting to achieve a goal $G\in G[\cdot]$.
+\subsection{State-invaraints}
+The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
+If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
+For instance, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
 
-Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
-\begin{align}
-argmax_{G\in G[\cdot]} P(\mathcal{O}|G) P(G).
-\end{align}
+{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
 
-The {\em plan recognition as planning} approach shows that the $P(\mathcal{O}|G)$ likelihood can be estimated leveraging the action model of the observed agents and an off-the-shelf classical planner~\cite{ramirez2012plan}.
-Given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$ (where $G[\cdot]$ represents the set of {\em recognizable} goals) then $P(\mathcal{O}|G)$ is estimated computing, for each goal $G\in G[\cdot]$, the cost difference of the solution plans to these two classical planning problems:
+The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
+For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}.
+There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.
+
+\begin{figure}[hbt!]
+  \begin{footnotesize}
+$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
+$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
+$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
+\end{footnotesize}
+ \caption{\small Example of the {\em strongest invariant} for the {\em blocksworld} domain.}
+\label{fig:strongest-invariant}
+\end{figure}
+
+A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a mutex because $block_A$ can only be on top of a single block.
+
+A {\em domain invariant} is an instance-independent invariant, i.e.
+holds for any possible initial state and set of objects.
+Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
+Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
+For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
+
+\section{Deductive learning of planning action models}
+\label{sec:evaluation}
+
+The task of learning action models is defined by $\Lambda=\tup{\mathcal{M_0},\Phi,\mathcal{O}(\tau)}$:
 \begin{itemize}
-\item $P^{\top}_G$, the classical planning problem built constraining $P=\tup{F,A,I,G}$ to achieve the particular goal $G\in G[\cdot]$ through a plan $\pi^\top$ that is {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-\item $P^{\bot}_G$, the classical planning problem that constrains solutions of $P=\tup{F,A,I,G}$ to plans $\pi^\bot$, that achieve $G\in G[\cdot]$, but that are {\em inconsistent} with $\mathcal{O}(\tau)$.
+\item $\mathcal{M_0}$ is the {\bf initial domain model} (set of action models).
+This set is {\em empty}, when learning from scratch, or {\em partially specified}, when some fragments of the action models are known a priori.
+\item $\Phi$, a set of {\em state-invariants} that constraint the set of possible states in the given domain.
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that reaches the goals $G\in G[\cdot]$ starting from the initial state $I$ in $P$.
 \end{itemize}
 
-The higher the value of the $cost(\pi^\top)-cost(\pi^\bot)$ difference, the higher the probability of the observed agents to aim goal $G\in G[\cdot]$.
-With this regard, {\em plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:
+A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O}(\tau)}$ is a domain model $\mathcal{M}'$ that is consistent with the information of $\mathcal{M}$ and with the observed plan trace $\tau$.
+This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan is consistent with the input observatio $\mathcal{O}(\tau)$ and only travers states that satisfy the states invariants.
 
-\begin{align}
-P(\mathcal{O}|G) = \frac{1}{1+e^{-\beta(cost(\pi^\top)-cost(\pi^\bot))}}
-\end{align}
-
-This expression is derived from the assumption that while the observed agents are not perfectly rational, they are more likely to follow cheaper plans, according to a {\em Boltzmann} distribution.
-The larger the value of $\beta$, the more rational the agents, and the less likely that they will follow suboptimal plans.
-Recent works show fast estimates of the $P(\mathcal{O}|G)$ likelihood computed using relaxations of the classical planning tasks~\cite{pereira2017landmark}.
-
-\section{Planning with unknown domain models}
-\label{sec:planning}
-This section introduces a novel formulation for classical planning in a setting where no action model is given.
-This setting has already shown related to the learning of action models for planning~\cite{SternJ17}.
-In particular it can be seen as an extreme scenario when the action model is learned from a single example that contains only two state observations: the initial state and the goals.
-A {\em classical planning with unknown domain models} is then a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-
-A solution to this task is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
-one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state meets the goal conditions, $G \subseteq s_n$.
-
-Next we show that the space of possible STRIPS action models can be encoded as a set of propositional variables and a set of constraints over those variables.
-Then, we show how to exploit this encoding to solve $P=\tup{F,A[\cdot],I,G}$ problems with an off-the-shelf classical planner and the properties of this approach.
-
-\subsection{A propositional encoding for the space of STRIPS action models}
-{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
-The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
-
-For instance in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for the {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the actual space of possible \strips\ schemata is bounded by constraints of three kinds:
-\begin{enumerate}
-\item {\bf Syntactic constraints}.
-\strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-{\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
-\item {\bf Domain-specific constraints}.
-One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\it State invariants} are also constraints of this kind~\cite{fox1998automatic}.
-\item {\bf Observation constraints}. An observation $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
-\end{enumerate}
-
-\begin{figure}
-  \begin{tiny}  
-  \begin{verbatim}
-(:action stack
-   :parameters (?v1 ?v2)
-   :precondition (and (holding ?v1) (clear ?v2))
-   :effect (and (not (holding ?v1)) (not (clear ?v2))
-                (clear ?v1) (handempty) (on ?v1 ?v2)))
-
-(pre_holding_v1_stack) (pre_clear_v2_stack)
-(eff_holding_v1_stack) (eff_clear_v2_stack)
-(eff_clear_v1_stack) (eff_handempty_stack) (eff_on_v1_v2_stack)
-  \end{verbatim}           
-  \end{tiny}  
- \caption{\small PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.}
-\label{fig:propositional}
-\end{figure}
-
-In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning classical planning action models.
-In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
-If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
-
-\subsection{A classical planning compilation for planning with unknown domain models}
-Now we show how we adapt the {\em classical planning compilation for learning \strips\ action models}~\cite{aineto2018learning} to address the task of {\em planning with unknown domain models}, using our propositional encoding of STRIPS action models.
-
-Given a classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$ we create a classical planning problem $P'=\tup{F',A',I,G}$ such that:
-\begin{itemize}
-\item $F'$ extends $F$ with a fluent $mode_{insert}$, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
-This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding action parameters $pars(\xi)$.
-
-\item $A'$ replaces the actions in $A$ with two types of actions.
-\begin{enumerate}
-\item Actions for {\em inserting} a {\em precondition}, {\em positive} effect or {\em negative} effect in $\xi$ following the syntactic constraints of \strips\ models. 
-\begin{itemize}
-\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi$. A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
+\subsection{Learning of planning action models with classical planning}
+Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
+Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space.
+With regard to our compilation, {\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema.
+For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi), mode_{insert}\},\\
-\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
-\end{align*}
-\end{small}
-
-\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi$.
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p}(\xi), mode_{insert}\},\\
-\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff_{p}(\xi)\}.
-\end{align*}
-\end{small}
-\end{itemize}
-
-\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
-The action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
-\end{enumerate}
-\end{itemize}
-
-\begin{small}
-\begin{align*}
-\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
-&\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\},\\
-&\{\emptyset\}\rhd\{mode_{insert}\}.
-\end{align*}
-\end{small}
-
-The intuition of the compilation is that the dynamics of the actions for {\em applying} an action model $\xi$ is determined by the values of the corresponding $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in the current state.
-For instance, executing {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small apply\_stack}} according to the values of $\{pre\_e\_stack, eff\_e\_stack\}_{\forall e\in{\mathcal I}_{\Psi,stack}}$ fluents in $s$.
-This means that if the current state $s$ holds $\{{\tt\scriptsize (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}\} \subset s$, then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
-Otherwise, a different set of precondition literals will be checked for the stack action.
-The same applies to the positive and negative effects.
-Executing {\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state only if {\tt{\small stack}} has been correctly programmed by the {\em insert} actions.
-
-\begin{figure}[hbt!]
-	{\scriptsize\tt
-		{\bf 00} : (insert\_pre\_stack\_holding\_v1) \\
-		01 : (insert\_pre\_stack\_clear\_v2)\\
-		{\bf 02} : (insert\_eff\_stack\_clear\_v1)\\
-		03 : (insert\_eff\_stack\_clear\_v2)\\
-		04 : (insert\_eff\_stack\_handempty)\\
-		05 : (insert\_eff\_stack\_holding\_v1)\\
-		06 : (insert\_eff\_stack\_on\_v1\_v2)\\
-		{\bf 07} : (apply\_stack blockA blockB)\\
-	}
-	\caption{\small Plan computed when solving the classical planning problem output by our compilation corresponding to a classical planning with unknown domain models.}
-	\label{fig:plan-lplan}
-\end{figure}
-
-Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving a $P'=\tup{F',A',I,G}$ classical planning problem output by our compilation.
-In the initial state of that problem the robot is holding {\tt blockA} while {\tt blockB} is clear and the single problem goal is having {\tt blockA} on top of {\tt blockB}.
-The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the programmed action model to achieve the goals $G$.
-Note that another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em unstack} action model and then applying the {\tt\small (unstack blockA blockB)} action instead of {\tt\small(stack blockA blockB)}.
-
-\subsection{Compilation properties}
-Now we present some properties of the compilation scheme.
-
-\begin{mylemma}
-Soundness. Any classical plan $\pi'$ that solves $P'$ produces a solution to the classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$.
-\end{mylemma}
-
-\begin{proof}[Proof]
-\begin{small}
-Once a given precondition or effect is inserted into an action model it can never be removed back and once an action model is applied it cannot be {\em reprogramed}.
-In the compiled problem the value of $F$, the fluents of the original problem, can exclusively be modified via $\mathsf{apply_{\xi,\omega}}$ actions.
-The set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$.
-This means that the action model used by the $\mathsf{apply_{\xi,\omega}}$ actions has to be consistent with the traversed intermediate states.
-We know that this must be true by the definition of the $\mathsf{apply_{\xi,\omega}}$ so hence, the sub-sequence of $\mathsf{apply_{\xi,\omega}}$ appearing in $\pi'$ to solve $P'$ is a solution plan to $P=\tup{F,A[\cdot],I,G}$.
-\end{small}
-\end{proof}
-
-\begin{mylemma}
-Completeness. Any plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$ is computable solving the corresponding classical planning task $P'$.
-\end{mylemma}
-
-\begin{proof}[Proof]
-\begin{small}
-By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
-Furthermore, the compilation does not discard any possible action model definable within ${\mathcal I}_{\Psi,\xi}$.
-This means that for every plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, we can build a plan $\pi'$ by selecting the appropriate actions for inserting precondition and effects to the corresponding action model and then selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state $I$ into a state $G \subseteq s_n$.
-\end{small}
-\end{proof}
-
-\subsubsection{The bias of an initially {\em empty} action model}
-Any $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluent is false at the initial state of our compilation.
-This fact can introduce a bias to the solutions of the $P=\tup{F,A[\cdot],I,G}$ classical planning task preferring solutions that imply action models with a smaller number of {\em preconditions}/{\em effects} (i.e., that imply a lower number of {\em insert} actions).
-
-This bias might be eliminated defining a cost landscape where {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}.
-In practice, since classical planners are not proficiency optimizing cost landscapes of this kind, we use a different approach that disregard the cost of actions the {\em insert} actions.
-Our approach is to use SAT-based planning because it can apply all the required actions for inserting preconditions in a single planning step (in parallel), because these actions do not interact.
-Further, actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
-The SAT-based planning approach is also convenient because its ability to deal with populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to performance.
-
-\subsubsection{Compilation size}
-The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
-The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$.
-The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
-
-Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\Psi,\xi}|$ and hence, the size of the classical planning task output by the compilation.
-
-\section{Goal recognition as planning with unknown domain models}
-\label{sec:recognition}
-
-We define the task of {\em goal recognition with unknown domain models} as a $\tup{P,\mathcal{O}(\tau)}$ pair, where:
-\begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a classical planning problem where $G[\cdot]$ is the set of {\em recognizable} goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that reaches the goals $G\in G[\cdot]$ starting from the initial state $I$ in $P$.
-\end{itemize}
-
-The {\em solution} to the {\em goal recognition with unknown domain models} task is again the subset of goals in $G[\cdot]$ that maximizes expression (1).
-
-\subsection{Estimating the $P(\mathcal{O}|G)$ likelyhood with unknown domain models}
-Now we are ready to build an estimate of the $P(\mathcal{O}|G)$ likelyhood.
-Our mechanism matches the {\em plan recognition as planning} approach~\cite{ramirez2012plan} except that we compute $cost(\pi^\top)$ using our compilation for {\em classical planning with unknown domain models}.
-
-In more detail, we build the estimate of the $P(\mathcal{O}|G)$ likelyhood following these four steps:
-\begin{enumerate}
-\item {\em Build $P^{\top}_G$}, the classical planning problem that constrains solutions of the problem $P=\tup{F,A[\cdot],s_0^o,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-Note that $s_0^o\in \mathcal{O}(\tau)$ is the initial state in the given observation.
-\item {\em Solve $P^{\top}_G$}, using the proposed compilation for {\em classical planning with unknown domain models}.
-Extract from this solution (1), $cost(\pi^\top)$ (by counting the number of $\mathsf{apply_{\xi,\omega}}$ actions in the solution) but also (2), the action model $A$ that is determined by the {\em insert} actions used in $\pi^\top$ to achieve the goals $G$.
-\item {\em Build $P^{\bot}_G$}, the classical planning problem that constrains $P=\tup{F,A,s_0^o,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ (where $A$ is the set of actions extracted in step 2.).
-\item {\em Solve $P^{\bot}_G$} with a classical planner and extract $cost(\pi_\bot)$ as the length of the found solution plan.
-\item Compute the $cost(\pi^\top)-cost(\pi^\bot)$ difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
-\end{enumerate}
-
-To compute the target probability distribution $P(G|\mathcal{O})$ plug the $P(\mathcal{O}|G)$ likelihoods into the {\em Bayes rule} from which the goal posterior probabilities are obtained.
-In this case the $P(\mathcal{O})$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
-
-\subsection{Extending the observation model of {\em plan recognition as planning}}
-The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions.
-However, the approach applies also to more expressive observation models that consider state observations as well, like the observation model defined above, with a simple three-fold extension:
-\begin{itemize}
-\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation.
-\item Adding $validated_m$ to every possible goal $G\in G[\cdot]$ to constrain solution plans $\pi^\top$ to be consistent with all the state observations.
-\item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
-\end{itemize}
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{validated_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg validated_{j-1}, validated_j\}.
+\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
+\cond(\mathsf{programPre_{f_1,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
 \end{align*}
@@
 \label{sec:evaluation}
-
-\section{Related Work}
-\label{sec:evaluation}
-The problem of {\em classical planning with unknown domain models} has been previously addressed~\cite{SternJ17}.
-In this work we evidence the relevance of this task for addressing {\em goal recognition} when the action model of the observed agent is not available (which it is typically a too strong assumption at many real-world applications).
-
-The paper also showed that {\em goal recognition}, when the domain model is unknown, is closely related to the learning of planning action models.
-With this regard, the classical planning compilation for learning \strips\ action models~\cite{aineto2018learning} is very appealing because it allows to produce a \strips\ action model from minimal input knowledge (a single initial state and goals pair), and to refine this model if more input knowledge is available (e.g.
-observation constraints).
-Most of the existing approaches for learning action models aim maximizing an statistical consistency of the learned model with respect to the input observations so require large amounts of input knowledge and do not produce action models that are guaranteed to be {\em logically consistent} with the given input knowledge.
-
-Our approach for {\em planning with an unknown domain model} is related to {\em goal recognition design}~\cite{KerenGK14}.
-The reason is that we are encoding the space of propositional schemes as state variables of the planning problem (the initial state encodes the {\em empty} action model with no preconditions and no effects) and provide actions to modify the value of this state variables as in {\em goal recognition design}.
-The aims of {\em goal recognition design} are however different.
-{\em Goal recognition design} applied to {\em goal recognition with unknown domain models} would compute the action model, in the space of possible models, that allows to reveal any of the possible goals as early as possible.
 

### 2019-02-04 doc/ijcai-19-deductivelearning/deductivelearning.tex
first commit ijcai19 paper on deductive learning

@@
 \label{sec:introduction}
-{\em Classical planing} is an interesting approach for learning \strips\ action models since it is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states)~\cite{aineto2018learning}.
-This paper shows that this flexibility of the compilation approach goes one step further since it also allows learning from {\em state-invariants}, logic formulae that constrain the set of possible states of a given domain.
+
+The specification of planning action models is a complex task, even for planning experts, which frequently causes a knowledge acquisition bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
+The {\em machine learning} of action models aims to relieve this knowledge acquisition bottleneck.
+Nowadays there is a wide range of different approaches for learning planning action models~\cite{arora:amodels:ker2018}.
+However, most of these approaches aim maximizing some measure of {\em statistical consistency} over large sets of input examples that typically are sequential plans.
+
+{\em Classical planning} is an interesting approach for learning \strips\ action models since it is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and guaranttees to output an action model that is {\em consistent} with the input knowledge~\cite{aineto2018learning}.
+In this paper we show that this flexibility of the compilation approach goes one step further since it also allows learning from {\em state-invariants}, logic formulae that constrain the set of possible states of a given domain.
 The experimental results show that {\em state-invariants} boost the performance of the {\em classical planing} compilation for learning \strips\ action models.
@@
 The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
-For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}.
-There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.
+For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for the {\em blocksworld} planning domain~\cite{slaney2001blocks}.
+There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard (as hard as testing plan existence~\cite{bylander:complexity:AIJ1994}).
 
@@
 \end{footnotesize}
- \caption{\small Example of the {\em strongest invariant} for the {\em blocksworld} domain.}
+ \caption{\small {\em Strongest invariant} for the {\em blocksworld} domain.}
 \label{fig:strongest-invariant}
@@
 
-\section{Deductive learning of planning action models}
+\section{Learning action models from {\em state-invaraints}}
 \label{sec:evaluation}
 
-The task of learning action models is defined by $\Lambda=\tup{\mathcal{M_0},\Phi,\mathcal{O}(\tau)}$:
+We define the task of learning action models as a tuple $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$, where:
 \begin{itemize}
-\item $\mathcal{M_0}$ is the {\bf initial domain model} (set of action models).
-This set is {\em empty}, when learning from scratch, or {\em partially specified}, when some fragments of the action models are known a priori.
-\item $\Phi$, a set of {\em state-invariants} that constraint the set of possible states in the given domain.
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that reaches the goals $G\in G[\cdot]$ starting from the initial state $I$ in $P$.
+\item $M$ is the {\em space of possible action model}.
+This set is the {\em full} space of action models, when learning from scratch, or {\em partially specified}, when some fragments of the action models are known a priori.
+A set of action models can be defined {\em explicitly}, enumerating all the models that belong to the set or {\em implicitly}, enumerating all the constraints that must satisfy any model that belongs to the set.
+A {\em partially specified} \strips\ action model is then a formalism for the {\em implicit} representation of a set of \strips\ schemes~\cite{sreedharan2018handling}.
+\item $\Phi$, a set of {\em state-invariants} that constrain the set of possible states in the given domain.
+\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves $P_\mathcal{O}$ (i.e.
+the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$ where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown since functions $\rho$ and/or $\theta$ of $a$ are undefined).
 \end{itemize}
 
-A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O}(\tau)}$ is a domain model $\mathcal{M}'$ that is consistent with the information of $\mathcal{M}$ and with the observed plan trace $\tau$.
-This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan is consistent with the input observatio $\mathcal{O}(\tau)$ and only travers states that satisfy the states invariants.
+A {\em solution} to a $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$ learning task is a model $\mathcal{M}'\in M$ that is {\em consistent} with the given input knowledge $\Phi$ and $\mathcal{O}(\tau)$.
+This means that let there exists a plan $\pi$ such that the observation $\mathcal{O}(\tau)$ is consistent with $\pi$ when the dynamics of $A[\cdot]$ is given by $\mathcal{M}'\in M$ and such that any state produced in the trajectory $\tau(\pi,P_\mathcal{O})$ satisfies the {\em state-invariants} $\Phi$.
 
-\subsection{Learning of planning action models with classical planning}
-Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
-Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space.
-With regard to our compilation, {\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema.
-For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:
+Next we show that the set $M$ of possible action models can be encoded as a set of propositional variables and a set of constraints over those variables.
+Then, we show how to exploit this encoding to solve the $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$ learning task with an off-the-shelf classical planner.
+
+\subsection{A propositional encoding for the space of STRIPS action models}
+{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
+The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
+
+For instance in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for the {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the actual space of possible \strips\ schemata is bounded by constraints of three kinds:
+\begin{enumerate}
+\item {\bf Syntactic constraints}.
+\strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
+{\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
+\item {\bf Domain-specific constraints}.
+One can introduce domain-specific knowledge to constrain further the space of possible schemata.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
+{\it State invariants} are also constraints of this kind.
+\item {\bf Observation constraints}. An observation $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
+\end{enumerate}
+
+\begin{figure}
+  \begin{tiny}  
+  \begin{verbatim}
+(:action stack
+   :parameters (?v1 ?v2)
+   :precondition (and (holding ?v1) (clear ?v2))
+   :effect (and (not (holding ?v1)) (not (clear ?v2))
+                (clear ?v1) (handempty) (on ?v1 ?v2)))
+
+(pre_holding_v1_stack) (pre_clear_v2_stack)
+(eff_holding_v1_stack) (eff_clear_v2_stack)
+(eff_clear_v1_stack) (eff_handempty_stack) (eff_on_v1_v2_stack)
+  \end{verbatim}           
+  \end{tiny}  
+ \caption{\small PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.}
+\label{fig:propositional}
+\end{figure}
+
+In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning classical planning action models.
+In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
+If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
+
+\subsection{Learning \strips\ action models with classical planning}
+Given a $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$ where $\Phi$ is a set of {\em domain mutex} $\phi\in\Phi$, we create a classical planning problem $P'=\tup{F',A',I,G}$ such that:
+\begin{itemize}
+\item $F'$ extends $F$ with a fluent $mode_{insert}$, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
+This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding action parameters $pars(\xi)$.
+
+\item $A'$ replaces the actions in $A$ with two types of actions.
+\begin{enumerate}
+\item Actions for {\em inserting} a {\em precondition}, {\em positive} effect or {\em negative} effect in $\xi$ following the syntactic constraints of \strips\ models.
+In the particular case that $M$ is a {\em partially specified model} then only the actions for inserting a possible {\em precondition} or {\em effect} are necessary.
+\begin{itemize}
+\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi$. A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
-\cond(\mathsf{programPre_{f_1,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi), mode_{insert}\},\\
+\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
 \end{align*}
 \end{small}
+
+\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p}(\xi), mode_{insert}\},\\
+\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff_{p}(\xi)\}.
+\end{align*}
+\end{small}
+\end{itemize}
+
+\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
+The action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
+\end{enumerate}
+\end{itemize}
+
+\begin{small}
+\begin{align*}
+\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
+&\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\},\\
+&\{\emptyset\}\rhd\{\neg mode_{insert}\}.
+\end{align*}
+\end{small}
+
+{\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema.
+For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:
 

### 2019-02-05 doc/ijcai-19-deductivelearning/deductivelearning.tex
first commit ijcai19 paper on deductive learning

@@
 
-\title{Learning \strips\ action models from {\em state-invaraints}}
+\title{Learning \strips\ action models from {\em state-invariants}}
 
@@
 \emails
-{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es, miquel.ramirez@unimelb.edu.au}}
+{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
 
@@
 
-The specification of planning action models is a complex task, even for planning experts, which frequently causes a knowledge acquisition bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
-The {\em machine learning} of action models aims to relieve this knowledge acquisition bottleneck.
-Nowadays there is a wide range of different approaches for learning planning action models~\cite{arora:amodels:ker2018}.
-However, most of these approaches aim maximizing some measure of {\em statistical consistency} over large sets of input examples that typically are sequential plans.
+The specification of planning action models is a complex process that limits, too often, the potential of model-based planning systems~\cite{kambhampati:modellite:AAAI2007}.
+The {\em machine learning} of action models can relieve this knowledge acquisition bottleneck of AI planning and nowadays there exists a wide range of effective approaches for learning planning action models~\cite{arora:amodels:ker2018}.
+Most of current approaches for learning planning action models are {\em inductive}, meaning that they compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so their success depends on the amount and quality of the input observations.
 
-{\em Classical planning} is an interesting approach for learning \strips\ action models since it is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and guaranttees to output an action model that is {\em consistent} with the input knowledge~\cite{aineto2018learning}.
-In this paper we show that this flexibility of the compilation approach goes one step further since it also allows learning from {\em state-invariants}, logic formulae that constrain the set of possible states of a given domain.
-The experimental results show that {\em state-invariants} boost the performance of the {\em classical planing} compilation for learning \strips\ action models.
+This paper proposes the exploitation of {\em deductive} knowledge to learn planning action models when input observations are minimal (just an {\em initial state} and the {\em goals}).
+We show that in such unfavourable scenario {\em state-invariant} (i.e logic formulae that specify constraints about the possible states of a given domain) help to learn better action models.
+Our approach is built on top of the {\em classical planning} compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
+This compilation approach is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and guarantees to output an action model that is {\em consistent} with the input knowledge.
+In this paper we show how to push the flexibility of the compilation approach for learning of \strips\ action models and boost its performance with {\em state-invariant} when few observations are given as input.
 

### 2019-02-05 doc/ijcai-19-deductivelearning/deductivelearning.tex
first version sections 1 and 2 ijcai19 paper on deductive learning

@@
 
-The specification of planning action models is a complex process that limits, too often, the potential of model-based planning systems~\cite{kambhampati:modellite:AAAI2007}.
-The {\em machine learning} of action models can relieve this knowledge acquisition bottleneck of AI planning and nowadays there exists a wide range of effective approaches for learning planning action models~\cite{arora:amodels:ker2018}.
-Most of current approaches for learning planning action models are {\em inductive}, meaning that they compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so their success depends on the amount and quality of the input observations.
+The specification of a planning action model is a complex process that limits, too often, the potential of model-based planning systems~\cite{kambhampati:modellite:AAAI2007}.
+The {\em machine learning} of action models can relieve this {\em knowledge acquisition bottleneck} of {\em AI planning} and nowadays, there exists a wide range of effective approaches for learning planning action models~\cite{arora:amodels:ker2018}.
+Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their success depends on the {\em amount} and {\em quality} of the input observations of plan executions.
 
-This paper proposes the exploitation of {\em deductive} knowledge to learn planning action models when input observations are minimal (just an {\em initial state} and the {\em goals}).
-We show that in such unfavourable scenario {\em state-invariant} (i.e logic formulae that specify constraints about the possible states of a given domain) help to learn better action models.
+This paper proposes the exploitation of {\em deductive} knowledge in the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations when learning action models.
 Our approach is built on top of the {\em classical planning} compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
 This compilation approach is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and guarantees to output an action model that is {\em consistent} with the input knowledge.
-In this paper we show how to push the flexibility of the compilation approach for learning of \strips\ action models and boost its performance with {\em state-invariant} when few observations are given as input.
+In this paper we show that, in unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models with the {\em classical planning} compilation.
 
@@
 This means that let there exists a plan $\pi$ such that the observation $\mathcal{O}(\tau)$ is consistent with $\pi$ when the dynamics of $A[\cdot]$ is given by $\mathcal{M}'\in M$ and such that any state produced in the trajectory $\tau(\pi,P_\mathcal{O})$ satisfies the {\em state-invariants} $\Phi$.
+
+%They compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model that is {\em consistent} with the input knowledge.
 

### 2019-02-05 doc/ijcai-19-deductivelearning/deductivelearning.tex
first version sections 1 and 2 ijcai19 paper on deductive learning

@@
 
-The specification of a planning action model is a complex process that limits, too often, the potential of model-based planning systems~\cite{kambhampati:modellite:AAAI2007}.
+The specification of a planning action model is a complex process that limits, too often, the application of model-based planning to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
 The {\em machine learning} of action models can relieve this {\em knowledge acquisition bottleneck} of {\em AI planning} and nowadays, there exists a wide range of effective approaches for learning planning action models~\cite{arora:amodels:ker2018}.
-Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their success depends on the {\em amount} and {\em quality} of the input observations of plan executions.
+Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance depends on the {\em amount} and {\em quality} of the input examples (namely observations of plan executions generated with the aimed action model).
 
-This paper proposes the exploitation of {\em deductive} knowledge in the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations when learning action models.
+This paper proposes the exploitation of {\em deductive} knowledge in the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations when learning \strips\ action models.
+Given an action model state-of-the-art planners use to compute {\em state-invariants} to reduce their search space and make the planning process more efficient~\cite{helmert2009concise}.
+In this paper we follow the oposite direction and exploit {\em state-invariants} to learn a planning action model.
+
 Our approach is built on top of the {\em classical planning} compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
-This compilation approach is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and guarantees to output an action model that is {\em consistent} with the input knowledge.
+This compilation is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
 In this paper we show that, in unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models with the {\em classical planning} compilation.

### 2019-02-05 doc/ijcai-19-deductivelearning/deductivelearning.tex
first updates to section 3 ijcai19 paper on deductive learning

@@
 
-\subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
-Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
-
-\begin{itemize}
-\item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
-The number of observed actions, $l$, ranges from $0$ (fully unobserved action sequence) to $|\pi|$ (fully observed action sequence).
-\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observed.
-A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
-Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
-Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-The number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each {\em observed} states comprises $[1,|F|]$ fluents (the observation can still miss intermediate states that are {\em unobserved}).
-\end{itemize}
-
-We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
-Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-In other words, having an input observation $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
-
 \subsection{State-invaraints}
@@
 
+\subsection{The observation model}
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
+Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
+
+\begin{itemize}
+\item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
+The number of observed actions, $l$, ranges from $0$ (fully unobserved action sequence) to $|\pi|$ (fully observed action sequence).
+\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observed.
+A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
+Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+The number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each {\em observed} states comprises $[1,|F|]$ fluents (the observation can still miss intermediate states that are {\em unobserved}).
+\end{itemize}
+
+We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
+In other words, having an input observation $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
+
 \section{Learning action models from {\em state-invaraints}}
@@
 
-We define the task of learning action models as a tuple $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$, where:
+We define the task of learning an action model from {\em state-invaraints} as a tuple $\Lambda=\tup{F,I,G,\Phi,M}$, where:
 \begin{itemize}
+\item $F$ is a state of fluents, $I$ is an initial state (a full assigment of the fluents), and $G$ is a set of goal conditions over the fluents.
+\item $\Phi$, a set of {\em state-invariants} that constrain the set of possible states in the given domain.
 \item $M$ is the {\em space of possible action model}.
@@
 A {\em partially specified} \strips\ action model is then a formalism for the {\em implicit} representation of a set of \strips\ schemes~\cite{sreedharan2018handling}.
-\item $\Phi$, a set of {\em state-invariants} that constrain the set of possible states in the given domain.
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves $P_\mathcal{O}$ (i.e.
-the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$ where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown since functions $\rho$ and/or $\theta$ of $a$ are undefined).
 \end{itemize}
 
-A {\em solution} to a $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$ learning task is a model $\mathcal{M}'\in M$ that is {\em consistent} with the given input knowledge $\Phi$ and $\mathcal{O}(\tau)$.
-This means that let there exists a plan $\pi$ such that the observation $\mathcal{O}(\tau)$ is consistent with $\pi$ when the dynamics of $A[\cdot]$ is given by $\mathcal{M}'\in M$ and such that any state produced in the trajectory $\tau(\pi,P_\mathcal{O})$ satisfies the {\em state-invariants} $\Phi$.
-
-%They compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model that is {\em consistent} with the input knowledge.
+A {\em solution} to a $\Lambda=\tup{F,I,G,\Phi,M}$ learning task is a model $\mathcal{M}'\in M$ that is {\em consistent} with the given input knowledge.
+This means that, let $P=\tup{F,A_{\mathcal{M}'},I,G}$ be a {\em classical planning problem} where the dynamics of the actions is given by the action model $\mathcal{M}'$, there exists a plan $\pi$ such that any state traversed in the trajectory $\tau(\pi,P_\mathcal{O})$ satisfies all the {\em state-invariants} in $\Phi$.
 
@@
 \subsection{A propositional encoding for the space of STRIPS action models}
+%They compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model that is {\em consistent} with the input knowledge.
+
 {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.

### 2019-02-05 doc/ijcai-19-deductivelearning/deductivelearning.tex
first updates to section 3 ijcai19 paper on deductive learning

--- 2019-02-05 doc/ijcai-19-deductivelearning/deductivelearning.tex
@@
 \begin{abstract}
+This paper addresses the learning of action models from {\em deductive} knowledge with the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations.
+Given an action model, state-of-the-art planners use it to infer {\em state-invariants} that reduce the planner search space and make the planning process more efficient~\cite{helmert2009concise}.
+In this paper we follow the opposite direction and exploit {\em state-invariants} to learn a planning action model.
 \end{abstract}
@@
 
-The specification of a planning action model is a complex process that limits, too often, the application of model-based planning to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
-The {\em machine learning} of action models can relieve this {\em knowledge acquisition bottleneck} of {\em AI planning} and nowadays, there exists a wide range of effective approaches for learning planning action models~\cite{arora:amodels:ker2018}.
-Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance depends on the {\em amount} and {\em quality} of the input examples (namely observations of plan executions generated with the aimed action model).
+The specification of planning action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
+The {\em machine learning} of action models can relieve the {\em knowledge acquisition bottleneck} of planning and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}.
+Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance depends on the {\em amount} and {\em quality} of the input examples (namely observations of plan executions generated by the aimed action model).
 
-This paper proposes the exploitation of {\em deductive} knowledge in the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations when learning \strips\ action models.
-Given an action model state-of-the-art planners use to compute {\em state-invariants} to reduce their search space and make the planning process more efficient~\cite{helmert2009concise}.
-In this paper we follow the oposite direction and exploit {\em state-invariants} to learn a planning action model.
+This paper addresses the learning of action models from {\em deductive} knowledge with the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations.
+Given an action model, state-of-the-art planners use it to infer {\em state-invariants} that reduce the planner search space and make the planning process more efficient~\cite{helmert2009concise}.
+In this paper we follow the opposite direction and exploit {\em state-invariants} to learn a planning action model.
 
@@
 
-\section{Learning action models from {\em state-invaraints}}
-\label{sec:evaluation}
+\section{Learning \strips\ action models from {\em state-invaraints}}
+\label{sec:learning}
 
-We define the task of learning action models as a tuple $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$, where:
+We define the task of learning an action model from {\em state-invaraints} as a tuple $\Lambda=\tup{P,\Phi,M}$, where:
 \begin{itemize}
-\item $M$ is the {\em space of possible action model}.
-This set is the {\em full} space of action models, when learning from scratch, or {\em partially specified}, when some fragments of the action models are known a priori.
+\item $P=\tup{F,A[\cdot],I,G}$, is a {\em classical planning problem} where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
+the functions $\rho$ and/or $\theta$ of $a$ are undefined).
+\item $\Phi$, a set of {\em state-invariants} that constrain the set of possible states in the given domain.
+\item $M$ is the {\em space of possible action model} for the $A[\cdot]$ actions.
+This set is the {\em full} space of action models, when learning from scratch, or a {\em partially specified action model}, when some fragments of the aimed action model are a priori known.
 A set of action models can be defined {\em explicitly}, enumerating all the models that belong to the set or {\em implicitly}, enumerating all the constraints that must satisfy any model that belongs to the set.
 A {\em partially specified} \strips\ action model is then a formalism for the {\em implicit} representation of a set of \strips\ schemes~\cite{sreedharan2018handling}.
-\item $\Phi$, a set of {\em state-invariants} that constrain the set of possible states in the given domain.
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that solves $P_\mathcal{O}$ (i.e.
-the classical planning problem $P_\mathcal{O}=\tup{F,A[\cdot],s_0^o,s_m^o}$ where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown since functions $\rho$ and/or $\theta$ of $a$ are undefined).
 \end{itemize}
 
-A {\em solution} to a $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$ learning task is a model $\mathcal{M}'\in M$ that is {\em consistent} with the given input knowledge $\Phi$ and $\mathcal{O}(\tau)$.
-This means that let there exists a plan $\pi$ such that the observation $\mathcal{O}(\tau)$ is consistent with $\pi$ when the dynamics of $A[\cdot]$ is given by $\mathcal{M}'\in M$ and such that any state produced in the trajectory $\tau(\pi,P_\mathcal{O})$ satisfies the {\em state-invariants} $\Phi$.
-
-%They compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model that is {\em consistent} with the input knowledge.
+A {\em solution} to a $\Lambda=\tup{P,\Phi,M}$ learning task is a model $\mathcal{M}'\in M$ such that there exists a plan $\pi$ satisfying that any state traversed in the trajectory $\tau(\pi,P)$ satisfies all the {\em state-invariants} in $\Phi$.
+Inductive approaches for the learning planning action models compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model in our case a solution to the addressed learning task is an action model that is {\em consistent} with the input knowledge.
 
 Next we show that the set $M$ of possible action models can be encoded as a set of propositional variables and a set of constraints over those variables.
-Then, we show how to exploit this encoding to solve the $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$ learning task with an off-the-shelf classical planner.
+Then, we show how to exploit this encoding to solve a $\Lambda=\tup{P,\Phi,M}$ learning task with an off-the-shelf classical planner.
 
@@
 \subsection{Learning \strips\ action models with classical planning}
-Given a $\Lambda=\tup{M,\Phi,\mathcal{O}(\tau)}$ where $\Phi$ is a set of {\em domain mutex} $\phi\in\Phi$, we create a classical planning problem $P'=\tup{F',A',I,G}$ such that:
+Given a $\Lambda=\tup{P,\Phi,M}$ where $\Phi$ is a set of {\em domain mutex} $\phi\in\Phi$, we create a classical planning problem $P'=\tup{F',A',I,G}$ such that:
 \begin{itemize}
-\item $F'$ extends $F$ with a fluent $mode_{insert}$, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
+\item $F'$ extends $F$ with a fluent ${\tt inconsistent}$, to indicate whether an action model is {\em inconsistent} with the input {\em state-invariants}, a fluent $mode_{insert}$, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
 This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding action parameters $pars(\xi)$.
@@
 \begin{align*}
-\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
+\pre(\mathsf{apply_{\xi,\omega}})=&\{\emptyset\},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge \neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
+&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
 &\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\},\\
 &\{\emptyset\}\rhd\{\neg mode_{insert}\}.
+\end{align*}
+\end{small}
+
+\section{Learning from {\em state-invariants} and observations of plan executions}
+\label{sec:observations}
+Most of the approachs for leanning planning action mdoels compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model that is {\em consistent} with the input knowledge.
+
+\subsection{The observation model}
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
+Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
+
+\begin{itemize}
+\item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
+The number of observed actions, $l$, ranges from $0$ (fully unobserved action sequence) to $|\pi|$ (fully observed action sequence).
+\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observed.
+A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
+Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+The number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each {\em observed} states comprises $[1,|F|]$ fluents (the observation can still miss intermediate states that are {\em unobserved}).
+\end{itemize}
+
+We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
+In other words, having an input observation $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
+
+\subsection{Extending the compilation to deal with {\em plan recognition as planning}}
+The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions.
+However, the approach applies also to more expressive observation models that consider state observations as well, like the observation model defined above, with a simple three-fold extension:
+\begin{itemize}
+\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation.
+Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions of $\tau$.
+The former is used to ensure that actions are executed in the same order as they are observed in $\tau$.
+The latter is used to iterate to the next planning step when solving $P_{\Lambda}$.
+\item Adding $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i \leq n$} to the initial state and $validated_m$ to every possible goal $G\in G[\cdot]$ to constrain solution plans $\pi^\top$ to be consistent with all the state observations.
+
+\item When the input plan trace contains observed actions, the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.\\
+
+\item Actions for {\em validating} the partially observed state $s_j\in\tau$, {\tt\small $1\leq j< m$}.
+These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observable data of the input plan trace $\tau$ follows after the execution of the apply actions.
+
+\item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
+\end{itemize}
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{validated_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg validated_{j-1}, validated_j\}.
 \end{align*}

### 2019-02-06 doc/ijcai-19-deductivelearning/deductivelearning.tex
complete version for sections 1,2 and 3 ijcai19 paper on deductive learning

@@
 \begin{abstract}
-This paper addresses the learning of action models from {\em deductive} knowledge with the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations.
-Given an action model, state-of-the-art planners use it to infer {\em state-invariants} that reduce the planner search space and make the planning process more efficient~\cite{helmert2009concise}.
-In this paper we follow the opposite direction and exploit {\em state-invariants} to learn a planning action model.
+This paper addresses the learning of action models from {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient learning examples.
+Our approach is built on top of a {\em classical planning} compilation that is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
+The experimental results show that, at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} are helpful to learn better \strips\ action models.
 \end{abstract}
@@
 
-The specification of planning action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
+The specification of planning action models is a complex process that limits, too often, the application of {\em model-based planning} systems to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
 The {\em machine learning} of action models can relieve the {\em knowledge acquisition bottleneck} of planning and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}.
-Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance depends on the {\em amount} and {\em quality} of the input examples (namely observations of plan executions generated by the aimed action model).
+Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance is linked to the {\em amount} and {\em quality} of the input examples (namely observations of plan executions generated by the aimed action model).
 
-This paper addresses the learning of action models from {\em deductive} knowledge with the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations.
-Given an action model, state-of-the-art planners use it to infer {\em state-invariants} that reduce the planner search space and make the planning process more efficient~\cite{helmert2009concise}.
-In this paper we follow the opposite direction and exploit {\em state-invariants} to learn a planning action model.
+This paper addresses the learning of action models from a different source of knowledge, {\em deductive} knowledge, with the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations.
+Given an action model, state-of-the-art planners infer {\em state-invariants} from that model to reduce the planner search space and make the planning process more efficient~\cite{helmert2009concise}.
+In this paper we follow the opposite direction and leverage {\em state-invariants} to learn a planning action model.
 
-Our approach is built on top of the {\em classical planning} compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
-This compilation is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
-In this paper we show that, in unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models with the {\em classical planning} compilation.
+Our approach extends the {\em classical planning} compilation for learning \strips\ action models~\cite{aineto2018learning}.
+Our extension is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
+The experimental results show that, at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models with the {\em classical planning} compilation.
 
@@
 
-\subsection{State-invaraints}
+\subsection{State-invariants}
 The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
@@
 
-\section{Learning \strips\ action models from {\em state-invaraints}}
+\section{Learning \strips\ action models from {\em state-invariants}}
 \label{sec:learning}
-
-We define the task of learning an action model from {\em state-invaraints} as a tuple $\Lambda=\tup{P,\Phi,M}$, where:
+We define the task of learning an action model from {\em state-invariants} as a tuple $\Lambda=\tup{P,\Phi,M}$, where:
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$, is a {\em classical planning problem} where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-\item $\Phi$, a set of {\em state-invariants} that constrain the set of possible states in the given domain.
-\item $M$ is the {\em space of possible action model} for the $A[\cdot]$ actions.
-This set is the {\em full} space of action models, when learning from scratch, or a {\em partially specified action model}, when some fragments of the aimed action model are a priori known.
-A set of action models can be defined {\em explicitly}, enumerating all the models that belong to the set or {\em implicitly}, enumerating all the constraints that must satisfy any model that belongs to the set.
-A {\em partially specified} \strips\ action model is then a formalism for the {\em implicit} representation of a set of \strips\ schemes~\cite{sreedharan2018handling}.
+\item $P=\tup{F,A[\cdot],I,G}$, is a {\em classical planning problem} where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is {\em unknown} (i.e.
+the functions $\rho$ and/or $\theta$ are undefined for $a\in A[\cdot]$).
+\item $\Phi$, a set of {\em state-invariants} that define constraints about the set of possible states.
+\item $M$ is the {\em space of possible action models} for the $A[\cdot]$ actions (i.e., the set of possible specifications of the $\rho$ and/or $\theta$ functions for each action $a\in A[\cdot]$).
 \end{itemize}
 
-A {\em solution} to a $\Lambda=\tup{P,\Phi,M}$ learning task is a model $\mathcal{M}'\in M$ such that there exists a plan $\pi$ satisfying that any state traversed in the trajectory $\tau(\pi,P)$ satisfies all the {\em state-invariants} in $\Phi$.
-Inductive approaches for the learning planning action models compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model in our case a solution to the addressed learning task is an action model that is {\em consistent} with the input knowledge.
+A {\em solution} to a $\Lambda=\tup{P,\Phi,M}$ learning task is a model $\mathcal{M}\in M$ such that there exists a plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, when the semantics of each action $a\in A[\cdot]$ is given by $\mathcal{M}$, and such that any state traversed by the trajectory $\tau(\pi,P)$ is {\em consistent} with the {\em state-invariants} in $\Phi$.
 
@@
 For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\it State invariants} are also constraints of this kind.
+{\it State invariants} are constraints of this kind.
 \item {\bf Observation constraints}. An observation $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
@@
 \subsection{Learning \strips\ action models with classical planning}
-Given a $\Lambda=\tup{P,\Phi,M}$ where $\Phi$ is a set of {\em domain mutex} $\phi\in\Phi$, we create a classical planning problem $P'=\tup{F',A',I,G}$ such that:
+Given a $\Lambda=\tup{P,\Phi,M}$ learning task we create a classical planning problem $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I,G_{\Lambda}}$ such that:
 \begin{itemize}
-\item $F'$ extends $F$ with a fluent ${\tt inconsistent}$, to indicate whether an action model is {\em inconsistent} with the input {\em state-invariants}, a fluent $mode_{insert}$, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
+\item $F_{\Lambda}$ extends $F$ with a fluent {\small$mode_{inval}$}, to indicate whether an action model is inconsistent with the input {\em state-invariants}, a fluent {\small$mode_{insert}$}, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
 This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding action parameters $pars(\xi)$.
 
-\item $A'$ replaces the actions in $A$ with two types of actions.
+\item $G_{\Lambda}= G\cup \{\neg mode_{inval}\}$ extends the original goals $G$ with the $\neg mode_{inval}$ literal to validate that only states consistent with the state constraints in $\Phi$ are traversed.
+
+\item $A_{\Lambda}$ replaces the actions in $A$ with two types of actions.
 \begin{enumerate}
@@
 The action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
+These actions validate also that any state traversed by a solution plan is {\em consistent} with the {\em state-invariants} in $\Phi$.
 \end{enumerate}
@@
 \begin{align*}
-\pre(\mathsf{apply_{\xi,\omega}})=&\{\emptyset\},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge \neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
-&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
+\pre(\mathsf{apply_{\xi,\omega}})=&\{\neg mode_{inval}\},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
 &\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\},\\
-&\{\emptyset\}\rhd\{\neg mode_{insert}\}.
+&\{pre_{p}(\xi)\wedge \neg p(\omega)\}\rhd\{mode_{inval}\}_{\forall p\in\Psi_\xi},\\
+&\{\neg\phi\}\rhd\{mode_{inval}\}_{\forall \phi\in\Phi},\\
+&\{\emptyset\}\rhd\{\neg mode_{insert}\},
 \end{align*}
@@
 
-\section{Learning from {\em state-invariants} and observations of plan executions}
+\subsection{Learning from partiallly specified models}
+The compilation is defined assuming $M$ represents the {\em full} space of \strips\ action models, that is when learning action models from scratch.
+However, in some contexts it is reasonable to assume that some parts of the action model are known so there is no need to learn the entire model from scratch~\cite{ZhuoNK13}.
+Here we show that the compilations is also flexible to this particular learning scenario where $M$ is given by a {\em partially specified action model}, when some fragments of the aimed action model are a priori known, so $M$ represents a subspace of the of the possible \strips\ action models\cite{sreedharan2018handling}.
+
+When the action model for a \strips\ schema $\xi$ is partially specified, the known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to true in the initial state.
+Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, thereby making the classical planning task $P_{\Lambda}$ easier to be solved.
+
+For example, supose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state $I$ is extended with litereals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the assocaited actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
+
+\subsection{Compilation properties}
+\begin{mylemma}
+Soundness. Any classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ produces a \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi,M}$ learning task.
+\end{mylemma}
+
+\begin{proof}[Proof]
+\begin{small}
+According to the $P_{\Lambda}$ compilation, once a given precondition or effect is inserted into the action model $\mathcal{M}$ it can never be removed back.
+In addition, once the action model $\mathcal{M}$ is applied it cannot be {\em reprogrammed}.
+In the compiled planning problem $P_{\Lambda}$, the value of the original fluents $F$ can exclusively be modified via $\mathsf{apply_{\xi,\omega}}$ actions.
+Therefore, the set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$ validating that every generated intermediate state is consistent with the input {\em state-invariants}.
+This is exactly the definition of a solution to the $\Lambda=\tup{P,\Phi,M}$ learning task.
+\end{small}
+\end{proof}
+
+\begin{mylemma}
+Completeness. Any \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi,M}$ learning task can be computed with a classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$.
+\end{mylemma}
+
+\begin{proof}[Proof]
+\begin{small}
+By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in a \strips\ action schema $\xi$ using predicates $\Psi$.
+In addition the $P_{\Lambda}$ compilation does not discard any possible action model $\mathcal{M}$ definable within ${\mathcal I}_{\Psi,\xi}$.
+This means that for every \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi,M}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate actions for inserting precondition and effects to the corresponding action model $\mathcal{M}$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state $I$ into a state that satisfies the goals $G$.
+\end{small}
+\end{proof}
+
+The size of the classical planning task $P_{\Lambda}$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
+The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$.
+The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
+
+Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\Psi,\xi}|$ and hence, the size of the classical planning task output by the compilation.
+
+\section{Learning from observations of plan executions}
 \label{sec:observations}
-Most of the approachs for leanning planning action mdoels compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model that is {\em consistent} with the input knowledge.
+So far we have explained the compilation for learning from a single input trace.
+However, the compilation is extensible to the more general case $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ is a set of plan traces.
+Taking this into account, a small modification is required in our compilation approach.
+In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_m^t\in \tau_t$, {\tt\small $1\leq t\leq k$} of a plan trace $\tau_t$ reset the current state and the current plan.
+These actions are now redefined as:
+
+Inductive approaches for the learning planning action models compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model in our case a solution to the addressed learning task is an action model that is {\em consistent} with the input knowledge.
 
@@
 
-\subsection{Extending the compilation to deal with {\em plan recognition as planning}}
+\subsection{Learning from observations with {\em classical planning}}
 The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions.

### 2019-02-06 doc/ijcai-19-deductivelearning/deductivelearning.tex
complete version for sections 1,2 and 3 ijcai19 paper on deductive learning

@@
 
+\subsection{Improving the compilation for {\em domain mutex}}
+We define a {\em domain mutex} as $p\leftrightarrow \neg q$ universally quantified formula where $p\in\Psi$ and $q\in\Psi$ are predicates that shape the fluents of a given planning problem.
+For instance, $\forall x_1\ holding(x_1)\leftrightarrow\neg clear(x_1)$  or $\forall x_1\ holding(x_1)\leftrightarrow\neg ontable(x_1)$ are {\em domain mutex} of the {\em blocksworld} domain.
+
+We pay special attention to this particular class of {\em state-invariants} because they are associated to the different {\em state-properties} of a given object and because they can be effectively exploited to improve our compilation for the learning /strips/ action models.
+Our approach is to extend the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions that determine the learned model with extra conditional effects indicating that the programmed model is {\em invalid} if it is inconsistent with a given {\em domain mutex}.
+Note that this {\em consistency} checking is more effective than the one implemented at the $\mathsf{apply_{\xi,\omega}}$ since it happens at an earlier stage of the planning process.
+
+Formally, given a $p\leftrightarrow \neg q$ we extend the actions for setting a precondition $p$ in a given action schema $\xi$ as follows:
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi),\\
+&mode_{insert},\neg mode_{inval}\},\\
+\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\},\\
+&\{pre_{q}(\xi)\}\rhd\{mode_{inval}\}.
+\end{align*}
+\end{small}
+The same procedure is applied for action $insertPre_{q,\xi}$ to ban this programming action iff $pre_{q}(\xi)$ precondition is already set.
+A similar procedure is also applied to $\mathsf{insertEff_{p,\xi}}$ and $\mathsf{insertEff_{q,\xi}}$ actions for banning two {\em negative effects} or (two {\em positive effects}) that are considered {\em domain mutex}.
+
 \subsection{Learning from partiallly specified models}

### 2019-02-06 doc/ijcai-19-deductivelearning/deductivelearning.tex
complete version for sections 1,2 and 3 ijcai19 paper on deductive learning

@@
 The same procedure is applied for action $insertPre_{q,\xi}$ to ban this programming action iff $pre_{q}(\xi)$ precondition is already set.
-A similar procedure is also applied to $\mathsf{insertEff_{p,\xi}}$ and $\mathsf{insertEff_{q,\xi}}$ actions for banning two {\em negative effects} or (two {\em positive effects}) that are considered {\em domain mutex}.
+A similar procedure is also applied to $\mathsf{insertEff_{p,\xi}}$ and $\mathsf{insertEff_{q,\xi}}$ actions for banning two {\em negative effects} or (two {\em positive effects}) that are considered {\em domain mutex}:
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p}(\xi), mode_{insert},\neg mode_{inval}\},\\
+\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff_{p}(\xi),\\
+&\{pre_{q}(\xi),eff_{q}(\xi),pre_{p}(\xi)\}\rhd\{mode_{inval}\},\\
+&\{\neg pre_{q}(\xi),eff_{q}(\xi),\neg pre_{p}(\xi)\}\rhd\{mode_{inval}\}.
+\end{align*}
+\end{small}
 
@@
 
+\section{Related work}
+\label{sec:related}
+
 \section{Conclusions}

### 2019-02-07 doc/ijcai-19-goalRecognition/goalRecognition.tex
example extended for ijai19 paper on goal recognition

@@
 \begin{figure}[hbt!]
-	{\scriptsize\tt
-		{\bf 00} : (insert\_pre\_stack\_holding\_v1) \\
-		01 : (insert\_pre\_stack\_clear\_v2)\\
-		{\bf 02} : (insert\_eff\_stack\_clear\_v1)\\
-		03 : (insert\_eff\_stack\_clear\_v2)\\
-		04 : (insert\_eff\_stack\_handempty)\\
-		05 : (insert\_eff\_stack\_holding\_v1)\\
-		06 : (insert\_eff\_stack\_on\_v1\_v2)\\
-		{\bf 07} : (apply\_stack blockA blockB)\\
+	{\tiny\tt
+
+\begin{tabular}{ll}
+		{\bf 00}:(insert\_pre\_stack\_holding\_v1) & 09:(insert\_pre\_pickup\_handempty)\\
+		01:(insert\_pre\_stack\_clear\_v2) & {\bf 10}:(insert\_eff\_pickup\_clear\_v1)\\
+		{\bf 02}:(insert\_eff\_stack\_clear\_v1) & 11:(insert\_eff\_pickup\_ontable\_v1)\\
+		03:(insert\_eff\_stack\_clear\_v2) & 12:(insert\_eff\_pickup\_handempty)\\
+		04:(insert\_eff\_stack\_handempty) & 13:(insert\_eff\_pickup\_holding\_v1)\\
+		05:(insert\_eff\_stack\_holding\_v1) & {\bf 14}:(apply\_pickup blockB)\\
+		06:(insert\_eff\_stack\_on\_v1\_v2) & 15:(apply\_stack blockB blockC)\\
+		{\bf 07}:(insert\_pre\_pickup\_clear\_v1) & 16:(apply\_pickup blockA)\\
+		08:(insert\_pre\_pickup\_ontable\_v1) & 17:(apply\_stack blockA blockB)\\             
+\end{tabular}
 }
@@
 Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving a $P'=\tup{F',A',I,G}$ classical planning problem output by our compilation.
-In the initial state of that problem the robot is holding {\tt blockA} while {\tt blockB} is clear and the single problem goal is having {\tt blockA} on top of {\tt blockB}.
-The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), and step $07$ is the plan postfix that applies the programmed action model to achieve the goals $G$.
-Note that another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em unstack} action model and then applying the {\tt\small (unstack blockA blockB)} action instead of {\tt\small(stack blockA blockB)}.
+In the initial state of that problem three  blocks (namely {\small\tt blockA},{\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table and the robot hand is empty.
+The problem goals is having the 3-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
+The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), steps $07-13$ insert the preconditions and effects of the {\tt\small pickup} action model and finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$.
+Note that another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em putdown} and {\em unstack} action models and then applying instead the four step postfix {\tt\small (putdown blockB), (unstack blockB blockC), (putdown blockA), (unstack blockA blockB)}.
 

### 2019-02-07 doc/ijcai-19-deductivelearning/deductivelearning.tex
bib updates to the ijcai19 paper on deductive learning

@@
 \begin{abstract}
-This paper addresses the learning of action models from {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient learning examples.
-Our approach is built on top of a {\em classical planning} compilation that is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
-The experimental results show that, at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} are helpful to learn better \strips\ action models.
+This paper addresses the learning of action models from {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to cushion the negative impact of insufficient learning examples.
+Our approach is a {\em classical planning} compilation that is flexible to different kinds of input knowledge (e.g., partially observations of plan executions including partially observed intermediate states and/or actions) and outputs an action model that is {\em consistent} with the given input knowledge.
+The experimental results show that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} are helpful to learn better \strips\ action models.
 \end{abstract}
@@
 The {\em machine learning} of action models can relieve the {\em knowledge acquisition bottleneck} of planning and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}.
-Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance is linked to the {\em amount} and {\em quality} of the input examples (namely observations of plan executions generated by the aimed action model).
+Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance is linked to the {\em amount} and {\em quality} of the input examples (which normally are observations of plan executions generated by the aimed action model).
 
-This paper addresses the learning of action models from a different source of knowledge, {\em deductive} knowledge, with the form of {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to reduce the negative effects of insufficient input observations.
-Given an action model, state-of-the-art planners infer {\em state-invariants} from that model to reduce the planner search space and make the planning process more efficient~\cite{helmert2009concise}.
-In this paper we follow the opposite direction and leverage {\em state-invariants} to learn a planning action model.
+This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge, with the form of {\em state-invariants} (i.e.
+logic formulae that specify constraints about the possible states of a given domain) to cushion the negative impact of insufficient learning examples.
+Given an action model, state-of-the-art planners infer {\em state-invariants} from that model to reduce search spaces and make the planning process more efficient~\cite{helmert2009concise}.
+In this paper we follow the opposite direction and leverage {\em state-invariants} to learn the planning action model.
 
-Our approach extends the {\em classical planning} compilation for learning \strips\ action models~\cite{aineto2018learning}.
-Our extension is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
-The experimental results show that, at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models with the {\em classical planning} compilation.
+Our approach for learning \strips\ action models from {\em state-invariants} is building a classical planning compilation that is inspired by recent work by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+Our compilation is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
+The experimental results show that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models with the {\em classical planning} compilation.
 
@@
 
+{\em State-invariants} have been previously used to infer a HTN planning model~\cite{lotinac2016constructing}.
+
 \section{Conclusions}

### 2019-02-07 doc/ijcai-19-deductivelearning/deductivelearning.tex
completed sections 1,2 and 3 of the ijcai19 paper on deductive learning

@@
 \label{sec:learning}
-We define the task of learning an action model from {\em state-invariants} as a tuple $\Lambda=\tup{P,\Phi,M}$, where:
+We define the task of learning a planning action model from {\em state-invariants} as a tuple $\Lambda=\tup{P,\Phi,M}$, where:
 \begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$, is a {\em classical planning problem} where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is {\em unknown} (i.e.
-the functions $\rho$ and/or $\theta$ are undefined for $a\in A[\cdot]$).
-\item $\Phi$, a set of {\em state-invariants} that define constraints about the set of possible states.
-\item $M$ is the {\em space of possible action models} for the $A[\cdot]$ actions (i.e., the set of possible specifications of the $\rho$ and/or $\theta$ functions for each action $a\in A[\cdot]$).
+\item $P=\tup{F,A[\cdot],I,G}$, is a {\em classical planning problem} where $A[\cdot]$ is a set of actions s.t., the {\em dynamics} of each action $a\in A[\cdot]$ is {\em unknown} (i.e.
+functions $\rho$ and/or $\theta$ are undefined for $a\in A[\cdot]$).
+\item $\Phi$ is a set of {\em state-invariants} that define constraints about the set of possible states.
+\item $M$ is the {\em space of possible action models} for the $A[\cdot]$ actions (i.e., the set of possible specifications of the $\rho$ and/or $\theta$ functions for each $a\in A[\cdot]$ action).
 \end{itemize}
 
-A {\em solution} to a $\Lambda=\tup{P,\Phi,M}$ learning task is a model $\mathcal{M}\in M$ such that there exists a plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, when the semantics of each action $a\in A[\cdot]$ is given by $\mathcal{M}$, and such that any state traversed by the trajectory $\tau(\pi,P)$ is {\em consistent} with the {\em state-invariants} in $\Phi$.
+A model $\mathcal{M}\in M$ is a {\em solution} to the $\Lambda=\tup{P,\Phi,M}$ learning task iff there exists a plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, when the semantics of each action $a\in A[\cdot]$ is given by $\mathcal{M}$, and such that any state traversed by the trajectory $\tau(\pi,P)$ is {\em consistent} with the {\em state-invariants} $\Phi$.
 
-Next we show that the set $M$ of possible action models can be encoded as a set of propositional variables and a set of constraints over those variables.
-Then, we show how to exploit this encoding to solve a $\Lambda=\tup{P,\Phi,M}$ learning task with an off-the-shelf classical planner.
+Next, we show that the set $M$ of possible action models can be compactly encoded as a set of propositional variables and a set of constraints over those variables.
+Then, we show how to exploit this compact encoding to solve a $\Lambda=\tup{P,\Phi,M}$ learning task with an off-the-shelf classical planner.
 
@@
 {\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
-The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
+Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters}, $pars(\xi)$.
+The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is the set of FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
 
@@
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the actual space of possible \strips\ schemata is bounded by constraints of three kinds:
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints of three kinds:
 \begin{enumerate}
@@
 {\it State invariants} are constraints of this kind.
-\item {\bf Observation constraints}. An observation $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
+\item {\bf Observation constraints}.
+The observation of the actions and states resulting from the execution of a plan depicts {\em semantic knowledge} that constraints further the space of possible action schemata.
 \end{enumerate}
@@
 In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning classical planning action models.
+This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ action models with classical planning.
 In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
 If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema using the {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
 
 \subsection{Learning \strips\ action models with classical planning}
-Given a $\Lambda=\tup{P,\Phi,M}$ learning task we create a classical planning problem $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I,G_{\Lambda}}$ such that:
+Our approach for computing an action model $\mathcal{M}\in M$ that solves the $\Lambda=\tup{P,\Phi,M}$ learning task is to build and solve a classical planning problem $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I,G_{\Lambda}}$ such that:
 \begin{itemize}
-\item $F_{\Lambda}$ extends $F$ with a fluent {\small$mode_{inval}$}, to indicate whether an action model is inconsistent with the input {\em state-invariants}, a fluent {\small$mode_{insert}$}, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
-This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding action parameters $pars(\xi)$.
+\item $F_{\Lambda}$ extends $F$ with a fluent {\small$mode_{inval}$}, to indicate whether an action model is {\em inconsistent} with the input {\em state-invariants} $\Phi$, a fluent {\small$mode_{insert}$}, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
+As explained, this is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$.
 
-\item $G_{\Lambda}= G\cup \{\neg mode_{inval}\}$ extends the original goals $G$ with the $\neg mode_{inval}$ literal to validate that only states consistent with the state constraints in $\Phi$ are traversed.
+\item $G_{\Lambda}= G\cup \{\neg mode_{inval}\}$ extends the original goals $G$ with the $\neg mode_{inval}$ literal to validate that only states {\em consistent} with the state constraints $\Phi$ are traversed by $P_{\Lambda}$ solutions.
 
@@
 \item Actions for {\em inserting} a {\em precondition}, {\em positive} effect or {\em negative} effect in $\xi$ following the syntactic constraints of \strips\ models. 
-In the particular case that $M$ is a {\em partially specified model} then only the actions for inserting a possible {\em precondition} or {\em effect} are necessary.
 \begin{itemize}
-\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi$. A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
+\item Actions which support the addition of a {\em precondition} $p\in {\mathcal I}_{\Psi,\xi}$ to the action model $\xi$.
+A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi), mode_{insert}\},\\
-\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre\_p\_\xi, \neg eff\_p\_\xi, mode_{insert}\},\\
+\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre\_p\_\xi\}.
 \end{align*}
@@
 
-\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi$.
+\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in {\mathcal I}_{\Psi,\xi}$ to the action model $\xi$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p}(\xi), mode_{insert}\},\\
-\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff_{p}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff\_p\_\xi, mode_{insert}\},\\
+\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff\_p\_\xi\}.
 \end{align*}
@@
 
-\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
+\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates, and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
 The action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
-These actions validate also that any state traversed by a solution plan is {\em consistent} with the {\em state-invariants} in $\Phi$.
+These actions validate also that any state traversed by $P_{\Lambda}$ solutions is {\em consistent} with the {\em state-invariants} $\Phi$.
 \end{enumerate}
@@
 \pre(\mathsf{apply_{\xi,\omega}})=&\{\neg mode_{inval}\},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
-&\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\},\\
-&\{pre_{p}(\xi)\wedge \neg p(\omega)\}\rhd\{mode_{inval}\}_{\forall p\in\Psi_\xi},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{pre\_p\_\xi\wedge eff\_p\_\xi\}\rhd\{\neg p(\omega)\}_{\forall p\in{\mathcal I}_{\Psi,\xi}},\\
+&\{\neg pre\_p\_\xi \wedge eff\_p\_\xi\}\rhd\{p(\omega)\}_{\forall p\in{\mathcal I}_{\Psi,\xi}}\},\\
+&\{pre\_p\_\xi \wedge \neg p(\omega)\}\rhd\{mode_{inval}\}_{\forall p\in{\mathcal I}_{\Psi,\xi}},\\
 &\{\neg\phi\}\rhd\{mode_{inval}\}_{\forall \phi\in\Phi},\\
@@
 \subsection{Improving the compilation for {\em domain mutex}}
-We define a {\em domain mutex} as $p\leftrightarrow \neg q$ universally quantified formula where $p\in\Psi$ and $q\in\Psi$ are predicates that shape the fluents of a given planning problem.
-For instance, $\forall x_1\ holding(x_1)\leftrightarrow\neg clear(x_1)$  or $\forall x_1\ holding(x_1)\leftrightarrow\neg ontable(x_1)$ are {\em domain mutex} of the {\em blocksworld} domain.
+We define a {\em domain mutex} as a $(p,q)$ predicates pair where $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of fluents $F$ of a given planning problem and such that they satisfy the following universally quantified formulae $p\leftrightarrow \neg q$.
+For instance, predicates $holding(x)$ and $clear(x)$ from the {\em blocksworld} are {\em domain mutex} since they satisfy $\forall\ x\ holding(x)\leftrightarrow\neg clear(x)$, in the same way predicates $clear(x)$ and $ontable(x)$ are not {\em domain mutex} because it is not always true that $\forall x\ clear(x)\leftrightarrow\neg ontable(x)$ .
 
-We pay special attention to this particular class of {\em state-invariants} because they are associated to the different {\em state-properties} of a given object and because they can be effectively exploited to improve our compilation for the learning /strips/ action models.
-Our approach is to extend the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions that determine the learned model with extra conditional effects indicating that the programmed model is {\em invalid} if it is inconsistent with a given {\em domain mutex}.
+We pay attention to this particular class of {\em state-invariants} because they are associated to the different {\em state-properties} of a given object and because they effectively prune the space of possible /strips/ action models.
+Our approach to implement this prunning is extending the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions (i.e., the actions that determine a solution model $\mathcal{M}$) with extra conditional effects indicating that the programmed model is {\em invalid} (is inconsistent with a {\em domain mutex} in $\Phi$).
 Note that this {\em consistency} checking is more effective than the one implemented at the $\mathsf{apply_{\xi,\omega}}$ since it happens at an earlier stage of the planning process.
 
-Formally, given a $p\leftrightarrow \neg q$ we extend the actions for setting a precondition $p$ in a given action schema $\xi$ as follows:
+Formally, given a {\em domain mutex} $(p,q)$ we extend the actions for setting a precondition $p$ in a given action schema $\xi$ as follows:
 
@@
 The same procedure is applied for action $insertPre_{q,\xi}$ to ban this programming action iff $pre_{q}(\xi)$ precondition is already set.
-A similar procedure is also applied to $\mathsf{insertEff_{p,\xi}}$ and $\mathsf{insertEff_{q,\xi}}$ actions for banning two {\em negative effects} or (two {\em positive effects}) that are considered {\em domain mutex}:
+A similar procedure is also applied to $\mathsf{insertEff_{p,\xi}}$ and $\mathsf{insertEff_{q,\xi}}$ actions for banning in this case, two {\em negative effects} (or two {\em positive effects}) that are {\em domain mutex}:
 

### 2019-02-07 doc/ijcai-19-deductivelearning/deductivelearning.tex
completed sections 1,2 and 3 of the ijcai19 paper on deductive learning

@@
 
-\subsection{Improving the compilation for {\em domain mutex}}
-We define a {\em domain mutex} as a $(p,q)$ predicates pair where $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of fluents $F$ of a given planning problem and such that they satisfy the following universally quantified formulae $p\leftrightarrow \neg q$.
-For instance, predicates $holding(x)$ and $clear(x)$ from the {\em blocksworld} are {\em domain mutex} since they satisfy $\forall\ x\ holding(x)\leftrightarrow\neg clear(x)$, in the same way predicates $clear(x)$ and $ontable(x)$ are not {\em domain mutex} because it is not always true that $\forall x\ clear(x)\leftrightarrow\neg ontable(x)$ .
+\subsection{Effective prunning of inconsistent action models with {\em domain mutex}}
+We define a {\em domain mutex} as a $(p,q)$ predicates pair where both $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of fluents $F$ of a given planning problem and such that they satisfy the following formulae $p\leftrightarrow \neg q$ where are the predicate variables are universally quantified.
+For instance, predicates $holding(x)$ and $clear(x)$ from the {\em blocksworld} are {\em domain mutex} since they satisfy $\forall x\ holding(x)\leftrightarrow\neg clear(x)$ while predicates $clear(x)$ and $ontable(x)$ (also from the {\em blocksworld}) are not {\em domain mutex} because they do not always satisfy $\forall x\ clear(x)\leftrightarrow\neg ontable(x)$.
 
-We pay attention to this particular class of {\em state-invariants} because they are associated to the different {\em state-properties} of a given object and because they effectively prune the space of possible /strips/ action models.
-Our approach to implement this prunning is extending the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions (i.e., the actions that determine a solution model $\mathcal{M}$) with extra conditional effects indicating that the programmed model is {\em invalid} (is inconsistent with a {\em domain mutex} in $\Phi$).
-Note that this {\em consistency} checking is more effective than the one implemented at the $\mathsf{apply_{\xi,\omega}}$ since it happens at an earlier stage of the planning process.
+We pay attention to this particular class of {\em state-invariants} because they define the {\em state-properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable an effectively pruning of inconsistent /strips/ action models.
+Our approach to implement this pruning is extending the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions (i.e., the actions that determine a solution model $\mathcal{M}$) with extra conditional effects indicating that the programmed model is {\em invalid} (i.e., inconsistent with a {\em domain mutex} in $\Phi$).
+Note that this {\em consistency} checking is more effective than the one implemented at the $\mathsf{apply_{\xi,\omega}}$ actions since $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions appear at an earlier stage of the planning process.
 
-Formally, given a {\em domain mutex} $(p,q)$ we extend the actions for setting a precondition $p$ in a given action schema $\xi$ as follows:
+Formally, given a {\em domain mutex} $(p,q)$, s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$, we extend the actions for setting a precondition $p$ in a given action schema $\xi$ as follows:
 
@@
 \end{small}
-The same procedure is applied for action $insertPre_{q,\xi}$ to ban this programming action iff $pre_{q}(\xi)$ precondition is already set.
-A similar procedure is also applied to $\mathsf{insertEff_{p,\xi}}$ and $\mathsf{insertEff_{q,\xi}}$ actions for banning in this case, two {\em negative effects} (or two {\em positive effects}) that are {\em domain mutex}:
+The same procedure is applied for action $insertPre_{q,\xi}$ to ban programming precondition $q$ iff $pre_{p}(\xi)$ precondition is already set.
+A similar procedure is also applied to $\mathsf{insertEff_{p,\xi}}$ and $\mathsf{insertEff_{q,\xi}}$ actions for banning in this case, two {\em negative effects} (or two {\em positive effects}) that are {\em domain mutex}.
+Now we show the actions that ban programming a positive (or negative) $p$ effect if its corresponding $q$ effect is already programmed:
 
@@
 \subsection{Learning from partiallly specified models}
-The compilation is defined assuming $M$ represents the {\em full} space of \strips\ action models, that is when learning action models from scratch.
-However, in some contexts it is reasonable to assume that some parts of the action model are known so there is no need to learn the entire model from scratch~\cite{ZhuoNK13}.
-Here we show that the compilations is also flexible to this particular learning scenario where $M$ is given by a {\em partially specified action model}, when some fragments of the aimed action model are a priori known, so $M$ represents a subspace of the of the possible \strips\ action models\cite{sreedharan2018handling}.
+The compilation is defined assuming $M$ represents the {\em full} space of \strips\ action models.
+In some contexts it is however reasonable to assume that the action model is not learned from scratch, e.g.
+because some parts of the action model are known~\cite{ZhuoNK13,sreedharan2018handling}.
+Here we show that the compilation approach is also flexible to this particular learning scenario.
 
 When the action model for a \strips\ schema $\xi$ is partially specified, the known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to true in the initial state.
-Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, thereby making the classical planning task $P_{\Lambda}$ easier to be solved.
+Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, making the classical planning task $P_{\Lambda}$ easier to be solved.
 
-For example, supose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state $I$ is extended with litereals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the assocaited actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
+For example, suppose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state $I$ is extended with literals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the associated actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
 
@@
 \begin{small}
-According to the $P_{\Lambda}$ compilation, once a given precondition or effect is inserted into the action model $\mathcal{M}$ it can never be removed back.
+According to the $P_{\Lambda}$ compilation, once a given precondition or effect is inserted into the action model $\mathcal{M}$ it cannot be removed back.
 In addition, once the action model $\mathcal{M}$ is applied it cannot be {\em reprogrammed}.
 In the compiled planning problem $P_{\Lambda}$, the value of the original fluents $F$ can exclusively be modified via $\mathsf{apply_{\xi,\omega}}$ actions.
-Therefore, the set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$ validating that every generated intermediate state is consistent with the input {\em state-invariants}.
-This is exactly the definition of a solution to the $\Lambda=\tup{P,\Phi,M}$ learning task.
+Therefore, the goals of the original $P$ classical planning task can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state $I=s_0$ reach a state $G \subseteq s_n$ validating that every generated intermediate state $s_i$, s.t.
+$0\leq i\leq n$, is consistent with the input {\em state-invariants}.
+This is exactly the definition of the solution condition for an action model $\mathcal{M}$ to solve the $\Lambda=\tup{P,\Phi,M}$ learning task.
 \end{small}
@@
 By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in a \strips\ action schema $\xi$ using predicates $\Psi$.
-In addition the $P_{\Lambda}$ compilation does not discard any possible action model $\mathcal{M}$ definable within ${\mathcal I}_{\Psi,\xi}$.
-This means that for every \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi,M}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate actions for inserting precondition and effects to the corresponding action model $\mathcal{M}$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state $I$ into a state that satisfies the goals $G$.
+In addition the $P_{\Lambda}$ compilation does not discard any possible action model $\mathcal{M}$ definable within ${\mathcal I}_{\Psi,\xi}$ while it can satisfy the domain mutex in $\Phi$.
+This means that for every \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi,M}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action model $\mathcal{M}$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state $I$ into a state that satisfies the goals $G$.
 \end{small}
@@
 The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$.
-The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
-
+The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $pre\_e\_\xi/eff\_e\_\xi$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
 Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\Psi,\xi}|$ and hence, the size of the classical planning task output by the compilation.
@@
 \label{sec:related}
+{\em State-invariants} have been previously used to infer a HTN lanning model~\cite{lotinac2016constructing}.
 
-{\em State-invariants} have been previously used to infer a HTN planning model~\cite{lotinac2016constructing}.
+In {\em Inductive Logic Programming} it is very common to make the hypothesis be consistent with some form deductive knowledge apart from the examples, what is ussually called {\em background knowledge}~\cite{muggleton1994inductive}.
 

### 2019-02-07 doc/ijcai-19-deductivelearning/deductivelearning.tex
first full version (no experiments though) of the ijcai19 paper on deductive learning

@@
 \label{sec:observations}
-So far we have explained the compilation for learning from a single input trace.
-However, the compilation is extensible to the more general case $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ is a set of plan traces.
-Taking this into account, a small modification is required in our compilation approach.
-In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_m^t\in \tau_t$, {\tt\small $1\leq t\leq k$} of a plan trace $\tau_t$ reset the current state and the current plan.
-These actions are now redefined as:
-
-Inductive approaches for the learning planning action models compute an action model that maximizes some notion of {\em statistical consistency} over a set of observations of plan executions so output an action model in our case a solution to the addressed learning task is an action model that is {\em consistent} with the input knowledge.
+Inductive approaches for the learning of planning action models compute an action model starting from an input set of observations of plan executions.
+Next we formalize the observation model for that kind of knowledge and the we show that the compilation can easily be extended to consider also inductive knowledge in the form of observations of plan executions.
 
@@
 \subsection{Learning from observations with {\em classical planning}}
-The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions.
-However, the approach applies also to more expressive observation models that consider state observations as well, like the observation model defined above, with a simple three-fold extension:
+To consider also inductive knowledge in the form of an $\mathcal{O}(\tau)$ observation of a plan execution, the compilation is extended in the following way:
 \begin{itemize}
 \item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation.
-Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions of $\tau$.
-The former is used to ensure that actions are executed in the same order as they are observed in $\tau$.
+Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions of $\mathcal{O}(\tau)$.
+The former is used to ensure that actions are executed in the same order as they are observed in $\mathcal{O}(\tau)$.
 The latter is used to iterate to the next planning step when solving $P_{\Lambda}$.
-\item Adding $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i \leq n$} to the initial state and $validated_m$ to every possible goal $G\in G[\cdot]$ to constrain solution plans $\pi^\top$ to be consistent with all the state observations.
-
-\item When the input plan trace contains observed actions, the extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.\\
-
-\item Actions for {\em validating} the partially observed state $s_j\in\tau$, {\tt\small $1\leq j< m$}.
+\item Adding $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i \leq n$} to the initial state and $validated_m$ to the goals $G$ of the classical planning problem to constrain solution plans to be consistent with all the state observations.
+\item Extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{pars(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\mathcal{O}(\tau)$.
+\item Actions for {\em validating} the partially observed state $s_j^o\in\mathcal{O}(\tau)$, {\tt\small $1\leq j< m$}.
 These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observable data of the input plan trace $\tau$ follows after the execution of the apply actions.
-
-\item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
+\item One $\mathsf{validate_{j}}$ action to constraint the solution plans to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
 \end{itemize}
@@
 
-{\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema.
-For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:
+So far we have explained the extension of the compilation for learning from a single observation $\mathcal{O}(\tau)$.
+The extension to the more general case of a set of observation $\{\mathcal{O}(\tau_1),\ldots,\mathcal{O}(\tau_k)\}$ is implemented with a small modification.
+In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_m^o\in \mathcal{O}(\tau_t)$, {\tt\small $1\leq t\leq k$} reset also the current state and the current plan step.
 

### 2019-02-08 doc/icaps-19/submitted/model-recognition.tex
new folders for preparing the icaps19 camera ready

--- 2018-12-04 doc/icaps-19/model-recognition.tex
@@
 
-\author{\#92}
+%\author{\#92}
 % Commented for blind submission
-%\author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
-%{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-%{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
-%{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
-%{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
-%\And Miquel Ram\'irez\\
-%{\scriptsize School of Computing and Information Systems}\\
-%{\scriptsize The University of Melbourne}\\
-%{\scriptsize Melbourne, Victoria. Australia}\\
-%{\scriptsize miquel.ramirez@nimelb.edu.au}}
+\author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
+{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
+{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
+{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
+{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
+\And Miquel Ram\'irez\\
+{\scriptsize School of Computing and Information Systems}\\
+{\scriptsize The University of Melbourne}\\
+{\scriptsize Melbourne, Victoria. Australia}\\
+{\scriptsize miquel.ramirez@nimelb.edu.au}}
 

### 2019-02-08 doc/ijcai-19-deductivelearning/deductivelearning.tex
first full version (no experiments though) of the ijcai19 paper on deductive learning

@@
 \begin{abstract}
-This paper addresses the learning of action models from {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain) to cushion the negative impact of insufficient learning examples.
-Our approach is a {\em classical planning} compilation that is flexible to different kinds of input knowledge (e.g., partially observations of plan executions including partially observed intermediate states and/or actions) and outputs an action model that is {\em consistent} with the given input knowledge.
-The experimental results show that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} are helpful to learn better \strips\ action models.
+This paper addresses the learning of action models from {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain).
+The benefit of exploiting {\em state-invariants} is two-fold, they allow to reduce the space of possible action models and to complete learning examples that are only partially observed.
+Our approach for the learning of \strips\ action models from {\em state-invariants} is a {\em classical planning} compilation.
+The compilation is flexible to different kinds of input knowledge (e.g., partially observations of plan executions including partially observed intermediate states and/or actions) and outputs an action model that is {\em consistent} with the given input knowledge.
+The experimental results show that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} are helpful to learn good quality \strips\ action models.
 \end{abstract}
@@
 The {\em machine learning} of action models can relieve the {\em knowledge acquisition bottleneck} of planning and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}.
-Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance is linked to the {\em amount} and {\em quality} of the input examples (which normally are observations of plan executions generated by the aimed action model).
+Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance is linked to the {\em amount} and {\em quality} of the input learning examples.
 
-This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge, with the form of {\em state-invariants} (i.e.
-logic formulae that specify constraints about the possible states of a given domain) to cushion the negative impact of insufficient learning examples.
-Given an action model, state-of-the-art planners infer {\em state-invariants} from that model to reduce search spaces and make the planning process more efficient~\cite{helmert2009concise}.
+This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge, to cushion the negative impact of insufficient learning examples.
+In more detail our approach leverages {\em state-invariants}, i.e.
+logic formulae that specify constraints about the possible states of a given domain.
+Given an action model, state-of-the-art planners are used to infer {\em state-invariants} from that model to reduce search spaces and make the planning process more efficient~\cite{helmert2009concise}.
 In this paper we follow the opposite direction and leverage {\em state-invariants} to learn the planning action model.
+The benefit of learning \strips\ action models from {\em state-invariants} is two-fold, {\em state-invariants} allow us to reduce the space of possible action models and to complete learning examples that are only partially observed.
 
-Our approach for learning \strips\ action models from {\em state-invariants} is building a classical planning compilation that is inspired by recent work by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+Our approach for learning \strips\ action models from {\em state-invariants} is compile the learning task into a classical planning task.
 Our compilation is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
-The experimental results show that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models with the {\em classical planning} compilation.
+The experimental results show that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models than with the existing {\em classical planning} compilation by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
 
@@
 Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
-
-The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
-For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for the {\em blocksworld} planning domain~\cite{slaney2001blocks}.
+For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em state invariants} for the {\em blocksworld} planning domain~\cite{slaney2001blocks}.
 There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard (as hard as testing plan existence~\cite{bylander:complexity:AIJ1994}).
@@
 \end{footnotesize}
- \caption{\small {\em Strongest invariant} for the {\em blocksworld} domain.}
+ \caption{\small Example of {\em state-invariants} for the {\em blocksworld} domain.}
 \label{fig:strongest-invariant}
@@
 A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a mutex because $block_A$ can only be on top of a single block.
+For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
 
@@
 Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
-For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
+For instance, $\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
 
@@
 functions $\rho$ and/or $\theta$ are undefined for $a\in A[\cdot]$).
-\item $\Phi$ is a set of {\em state-invariants} that define constraints about the set of possible states.
+\item $\Phi$ is a set of {\em state-invariants} that define constraints about the set of possible states in the previous planning problem $P$.
 \item $M$ is the {\em space of possible action models} for the $A[\cdot]$ actions (i.e., the set of possible specifications of the $\rho$ and/or $\theta$ functions for each $a\in A[\cdot]$ action).
@@
 
-A model $\mathcal{M}\in M$ is a {\em solution} to the $\Lambda=\tup{P,\Phi,M}$ learning task iff there exists a plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, when the semantics of each action $a\in A[\cdot]$ is given by $\mathcal{M}$, and such that any state traversed by the trajectory $\tau(\pi,P)$ is {\em consistent} with the {\em state-invariants} $\Phi$.
+We say that a given model $\mathcal{M}\in M$ is a {\em solution} to the $\Lambda=\tup{P,\Phi,M}$ learning task iff there exists a plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, when the semantics of each action $a\in A[\cdot]$ is given by $\mathcal{M}$, and such that any state traversed by a trajectory $\tau(\pi,P)$ is {\em consistent} with the input set of {\em state-invariants} $\Phi$.
 
-Next, we show that the set $M$ of possible action models can be compactly encoded as a set of propositional variables and a set of constraints over those variables.
+Next, we show that the set $M$, of possible action models, can be compactly encoded as a set of propositional variables and a set of constraints over those variables.
 Then, we show how to exploit this compact encoding to solve a $\Lambda=\tup{P,\Phi,M}$ learning task with an off-the-shelf classical planner.
@@
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints of three kinds:
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints of two kinds:
 \begin{enumerate}
@@
 {\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
-\item {\bf Domain-specific constraints}.
-One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\it State invariants} are constraints of this kind.
 \item {\bf Observation constraints}.
@@
 \end{enumerate}
+
+In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ action models with classical planning.
+In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
+If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema using the {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
 
@@
 
-In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ action models with classical planning.
-In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
-If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema using the {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
+In addition, one can introduce {\em domain-specific knowledge} to constrain further the space of possible schemata.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
+{\it State invariants} are {\em domain-specific knowledge} and they can be seen as either {\em syntactic} or {\em semantic} constraints because on the one hand, they constrain the space of possible action models but on the other hand, they can be used to complete partial observations of the states traversed by a plan.
 
@@
 These actions validate also that any state traversed by $P_{\Lambda}$ solutions is {\em consistent} with the {\em state-invariants} $\Phi$.
+The definition $\mathsf{apply_{\xi,\omega}}$ actions is also more compact in our compilation that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} since are not using disjunctions to code the possible preconditions of an action schema.
 \end{enumerate}
@@
 
-\subsection{Effective prunning of inconsistent action models with {\em domain mutex}}
+\subsection{Pruning inconsistent action models with {\em domain mutex}}
 We define a {\em domain mutex} as a $(p,q)$ predicates pair where both $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of fluents $F$ of a given planning problem and such that they satisfy the following formulae $p\leftrightarrow \neg q$ where are the predicate variables are universally quantified.
@@
 
-We pay attention to this particular class of {\em state-invariants} because they define the {\em state-properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable an effectively pruning of inconsistent /strips/ action models.
+We pay attention to this particular class of {\em state-invariants} because they define the {\em state-properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable an effectively pruning of inconsistent \strips\ action models.
 Our approach to implement this pruning is extending the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions (i.e., the actions that determine a solution model $\mathcal{M}$) with extra conditional effects indicating that the programmed model is {\em invalid} (i.e., inconsistent with a {\em domain mutex} in $\Phi$).
@@
 \end{small}
-
-\subsection{Learning from partiallly specified models}
-The compilation is defined assuming $M$ represents the {\em full} space of \strips\ action models.
-In some contexts it is however reasonable to assume that the action model is not learned from scratch, e.g.
-because some parts of the action model are known~\cite{ZhuoNK13,sreedharan2018handling}.
-Here we show that the compilation approach is also flexible to this particular learning scenario.
-
-When the action model for a \strips\ schema $\xi$ is partially specified, the known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to true in the initial state.
-Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, making the classical planning task $P_{\Lambda}$ easier to be solved.
-
-For example, suppose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state $I$ is extended with literals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the associated actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
 
@@
 Inductive approaches for the learning of planning action models compute an action model starting from an input set of observations of plan executions.
-Next we formalize the observation model for that kind of knowledge and the we show that the compilation can easily be extended to consider also inductive knowledge in the form of observations of plan executions.
+This section provides a formal model for such input observations and shows how to leverage {\em state-invariants} to automatically {\em complete} those input observations.
+The section ends with the extension of our compilation to exploit the {\em completed} observations for the learning of \strips\ action models.
 
@@
 
-\subsection{Learning from observations with {\em classical planning}}
-To consider also inductive knowledge in the form of an $\mathcal{O}(\tau)$ observation of a plan execution, the compilation is extended in the following way:
+\subsection{Completing observations with {\em domain mutex}}
+Our observation model follows the {\em open world} assumption, in other words, what is not observed is considered unknown.
+Here, we show that {\em state-invariants} are helpful to infer new knowledge that was unobserved.
+
+Given a {\em domain mutex} $(p,q)$ and a state observation $s_j^o\in \mathcal{O}(\tau)$, {\small $(1\leq j\leq m)$}, such that the literal $p(\omega)\in s_j^o$ is an instantiation of predicate $p$ over some subset of objects $\omega\subseteq\Omega^{|pars(p)|}$ then, the state observation can be safely completed adding the new literal $\neg q(\omega)$ (despite $\neg q(\omega)$ was actually unobserved).
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we have the {\em domain mutex} $\forall x\ holding(x)\leftrightarrow\neg clear(x)$ in the input set $\Phi$ of {\em state-invariants} we can safely add to the observation the literal {\tt\small $\neg$clear(blockA)} (despite this literal was actually unobserved).
+The process is repeated for all the observed states in $\mathcal{O}(\tau)$ and all the {\em domain mutex} in $\Phi$ to produce a new completed observation $\mathcal{O}(\tau)'$.
+
+\subsection{Learning from completed observations with {\em classical planning}}
+Let be $\mathcal{O}(\tau)'$ an observation completed as explained above, we extend here our compilation to constraint the possible \strips\ models with $\mathcal{O}(\tau)$':
 \begin{itemize}
-\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation.
-Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions of $\mathcal{O}(\tau)$.
-The former is used to ensure that actions are executed in the same order as they are observed in $\mathcal{O}(\tau)$.
+\item One fluent $\{observed_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)'$ state observation.
+Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions in $\mathcal{O}(\tau)'$.
+The former is used to ensure that actions are executed in the same order as they are observed.
 The latter is used to iterate to the next planning step when solving $P_{\Lambda}$.
-\item Adding $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i \leq n$} to the initial state and $validated_m$ to the goals $G$ of the classical planning problem to constrain solution plans to be consistent with all the state observations.
-\item Extra conditional effects $\{at_{i},plan(name(a_i),\Omega^{pars(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\mathcal{O}(\tau)$.
-\item Actions for {\em validating} the partially observed state $s_j^o\in\mathcal{O}(\tau)$, {\tt\small $1\leq j< m$}.
-These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observable data of the input plan trace $\tau$ follows after the execution of the apply actions.
-\item One $\mathsf{validate_{j}}$ action to constraint the solution plans to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
+\item Adding $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i \leq n$} to the initial state and $observed_m$ to the goals $G$ of the classical planning problem and hence, constrain the solution plans to be consistent with all the state observations.
+\item Adding the extra conditional effects {\small $\{at_{i},plan(name(a_i),\Omega^{pars(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$} to the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\mathcal{O}(\tau)'$.
+\item Actions for {\em validating} the partially observed state $s_j^o\in\mathcal{O}(\tau)'$, {\tt\small $1\leq j< m$}.
+These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observable data of the input observation $\mathcal{O}(\tau)$' follows after the execution of the apply actions.
+\item One $\mathsf{validate_{j}}$ action to constraint the solution plans to be consistent with the $s_j^o\in\mathcal{O}(\tau)'$ input state observation, {\small $(1\leq j\leq m)$}.  
 \end{itemize}
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{validated_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg validated_{j-1}, validated_j\}.
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{observed_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg observed_{j-1}, observed_j\}.
 \end{align*}
@@
 
-So far we have explained the extension of the compilation for learning from a single observation $\mathcal{O}(\tau)$.
+So far we explained the extension of the compilation for learning from a single observation $\mathcal{O}(\tau)'$.
 The extension to the more general case of a set of observation $\{\mathcal{O}(\tau_1),\ldots,\mathcal{O}(\tau_k)\}$ is implemented with a small modification.
@@
 
-In {\em Inductive Logic Programming} it is very common to make the hypothesis be consistent with some form deductive knowledge apart from the examples, what is ussually called {\em background knowledge}~\cite{muggleton1994inductive}.
+In {\em Inductive Logic Programming} it is very common to make the hypothesis be consistent with some form deductive knowledge apart from the examples, what is usually called {\em background knowledge}~\cite{muggleton1994inductive}.
 

### 2019-02-11 doc/ijcai-19-goalRecognition/goalRecognition.tex
updates to the ijcai19 paper on goal recognition

@@
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the actual space of possible \strips\ schemata is bounded by constraints of three kinds:
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints of two kinds:
 \begin{enumerate}
@@
 {\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
-\item {\bf Domain-specific constraints}.
-One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\it State invariants} are also constraints of this kind~\cite{fox1998automatic}.
-\item {\bf Observation constraints}. An observation $\mathcal{O}(\tau)$ depicts {\em semantic knowledge} that constraints further the space of possible action schemata.   
+\item {\bf Observation constraints}.
+The observation of the actions and states resulting from the execution of a plan depicts {\em semantic knowledge} that constraints further the space of possible action schemata.
 \end{enumerate}
+
+In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ action models with classical planning.
+In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
+If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
+Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema using the {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
 
@@
 
-In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning classical planning action models.
-In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
-If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
+In addition, one can introduce {\em domain-specific knowledge} to constrain further the space of possible schemata.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
+{\it State invariants} are {\em domain-specific knowledge}that can be exploited either as {\em syntactic} constrains to reduce the space of possible action models but also as {\em semantic} constraints to complete partial observations of the states traversed by a plan.
 
@@
 \begin{enumerate}
-\item Actions for {\em inserting} a {\em precondition}, {\em positive} effect or {\em negative} effect in $\xi$ following the syntactic constraints of \strips\ models. 
+\item Actions for {\em inserting} a {\em precondition}, {\em positive} effect or {\em negative} effect into the action model $\xi$ following the syntactic constraints of \strips\ models. 
 \begin{itemize}
-\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi$. A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
+\item Actions to insert an $e\in{\mathcal I}_{\Psi,\xi}$ {\em precondition} into $\xi$. The precondition is only inserted when neither $pre\_e\_\xi$ nor $eff\_e\_\xi$ exist in $\xi$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi), mode_{insert}\},\\
-\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre\_e\_\xi, \neg eff\_e\_\xi, mode_{insert}\},\\
+\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre\_e\_\xi\}.
 \end{align*}
@@
 
-\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi$.
+\item Actions to insert an $e\in{\mathcal I}_{\Psi,\xi}$ {\em effect} to the action model $\xi$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p}(\xi), mode_{insert}\},\\
-\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff_{p}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff\_e\_\xi, mode_{insert}\},\\
+\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff\_e\_\xi\}.
 \end{align*}
@@
 \begin{align*}
-\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
-&\{\neg pre_{p}(\xi)\wedge eff_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi}\},\\
+\pre(\mathsf{apply_{\xi,\omega}})=&\{pre\_e\_\xi\implies p(\omega)\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{pre\_e\_\xi\wedge eff\_e\_\xi\}\rhd\{\neg p(\omega)\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+&\{\neg pre\_e\_\xi \wedge eff\_e\_\xi\}\rhd\{p(\omega)\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\},\\
 &\{\emptyset\}\rhd\{\neg mode_{insert}\}.
@@
 \begin{tabular}{ll}
-		{\bf 00}:(insert\_pre\_stack\_holding\_v1) & 09:(insert\_pre\_pickup\_handempty)\\
-		01:(insert\_pre\_stack\_clear\_v2) & {\bf 10}:(insert\_eff\_pickup\_clear\_v1)\\
-		{\bf 02}:(insert\_eff\_stack\_clear\_v1) & 11:(insert\_eff\_pickup\_ontable\_v1)\\
-		03:(insert\_eff\_stack\_clear\_v2) & 12:(insert\_eff\_pickup\_handempty)\\
-		04:(insert\_eff\_stack\_handempty) & 13:(insert\_eff\_pickup\_holding\_v1)\\
-		05:(insert\_eff\_stack\_holding\_v1) & {\bf 14}:(apply\_pickup blockB)\\
-		06:(insert\_eff\_stack\_on\_v1\_v2) & 15:(apply\_stack blockB blockC)\\
-		{\bf 07}:(insert\_pre\_pickup\_clear\_v1) & 16:(apply\_pickup blockA)\\
-		08:(insert\_pre\_pickup\_ontable\_v1) & 17:(apply\_stack blockA blockB)\\             
+		{\bf 00}:(insert\_pre\_stack\_holding\_v1) & {\bf 10}:(insert\_eff\_pickup\_clear\_v1) \\
+		01:(insert\_pre\_stack\_clear\_v2) & 11:(insert\_eff\_pickup\_ontable\_v1)\\
+                {\bf 02}:(insert\_pre\_pickup\_handempty) & 12:(insert\_eff\_pickup\_handempty)\\
+                03:(insert\_pre\_pickup\_clear\_v1) & 13:(insert\_eff\_pickup\_holding\_v1)\\
+                04:(insert\_pre\_pickup\_ontable\_v1) & {\bf 14}:(apply\_pickup blockB)\\
+                {\bf 05}:(insert\_eff\_stack\_clear\_v1) & 15:(apply\_stack blockB blockC)\\
+                06:(insert\_eff\_stack\_clear\_v2) & 16:(apply\_pickup blockA)\\
+                07:(insert\_eff\_stack\_handempty) & 17:(apply\_stack blockA blockB) \\
+                08:(insert\_eff\_stack\_holding\_v1) & \\
+                09:(insert\_eff\_stack\_on\_v1\_v2) &             		 
 \end{tabular}
@@
 Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving a $P'=\tup{F',A',I,G}$ classical planning problem output by our compilation.
-In the initial state of that problem three  blocks (namely {\small\tt blockA},{\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table and the robot hand is empty.
-The problem goals is having the 3-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
+In the initial state of that problem three  blocks ({\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table and the robot hand is empty.
+The problem goal is having the 3-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
 The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), steps $07-13$ insert the preconditions and effects of the {\tt\small pickup} action model and finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$.
 Note that another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em putdown} and {\em unstack} action models and then applying instead the four step postfix {\tt\small (putdown blockB), (unstack blockB blockC), (putdown blockA), (unstack blockA blockB)}.
+
+\subsection{The bias of an initially {\em empty} action model}
+All $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our compilation.
+This fact might introduce a bias to the solutions of the $P=\tup{F,A[\cdot],I,G}$ classical planning task preferring solutions that imply action models with a smaller number of {\em preconditions}/{\em effects} (i.e., that imply a lower number of {\em insert} actions).
+
+This bias could be eliminated defining a cost landscape where {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}.
+In practice, since classical planners are not proficiency optimizing cost landscapes of this kind, we use a different approach that disregard the cost of actions the {\em insert} actions.
+Our approach is to use SAT-based planning because it can apply all the required actions for inserting preconditions in a single planning step (these actions do not interact).
+Further, actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
+The SAT-based planning approach is also convenient because its ability to deal with populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to performance.
+
+Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers, one for inserting the action preconditions, another for inserting the action effects followed by N layers of the SATPLAN encoding extended to apply the action models determined by the previous two initial layers.
+With regard to the example of Figure~\ref{fig:plan-lplan} this means that steps [00-04] are applied in paralel in a first SATPLAN layer, steps [05-13] are applied in paralel in a second SATPLAN layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer.
 
@@
 
-\subsubsection{The bias of an initially {\em empty} action model}
-Any $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluent is false at the initial state of our compilation.
-This fact can introduce a bias to the solutions of the $P=\tup{F,A[\cdot],I,G}$ classical planning task preferring solutions that imply action models with a smaller number of {\em preconditions}/{\em effects} (i.e., that imply a lower number of {\em insert} actions).
-
-This bias might be eliminated defining a cost landscape where {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}.
-In practice, since classical planners are not proficiency optimizing cost landscapes of this kind, we use a different approach that disregard the cost of actions the {\em insert} actions.
-Our approach is to use SAT-based planning because it can apply all the required actions for inserting preconditions in a single planning step (in parallel), because these actions do not interact.
-Further, actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
-The SAT-based planning approach is also convenient because its ability to deal with populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to performance.
-
-\subsubsection{Compilation size}
 The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
 The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$.
-The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
-
+The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $\{pre\_e\_\xi, eff\_e\_\xi\}$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
 Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\Psi,\xi}|$ and hence, the size of the classical planning task output by the compilation.
@@
 The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions.
-However, the approach applies also to more expressive observation models that consider state observations as well, like the observation model defined above, with a simple three-fold extension:
+However, the approach applies also to more expressive observation models that consider state observations as well, like the observation model defined above, with a simple two-fold extension:
 \begin{itemize}
-\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation.
-\item Adding $validated_m$ to every possible goal $G\in G[\cdot]$ to constrain solution plans $\pi^\top$ to be consistent with all the state observations.
+\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation and $validated_m$ is added to every possible goal $G\in G[\cdot]$ to constrain solution plans $\pi^\top$ to be consistent with all the state observations.
 \item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
@@
 \label{sec:conclusions}
+In some contexts it is however reasonable to assume that the action model is not learned from scratch, e.g.
+because some parts of the action model are known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
+Our compilation approach is also flexible to this particular learning scenario.
+The known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to true in the initial state.
+Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, making the classical planning task $P_{\Lambda}$ easier to be solved.
+For example, suppose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state $I$ is extended with literals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the associated actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
 

### 2019-02-11 doc/ijcai-19-goalRecognition/goalRecognition.tex
processing Miquels comments for pages 1 and 2 ijcai19 paper on goal recognition

@@
 \begin{abstract}
-This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em knowing the action model of the observed agents}.
+This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em that the observer knows the action model of the observed agents}.
 The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the {\em plan recognition as planning} approach for {\em goal recognition}.
-The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without {\em a priori} knowing the action model of the observed agents.
+We show that this novel formulation allow us to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without knowing beforehand the action model of the observed agents.
 \end{abstract}
@@
 \label{sec:introduction}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents acting to achieve one of those goals.
-Despite there exists a wide range of different approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents pursuing one of those goals.
+While there exist a wide range of different approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
 
-{\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to compute the most likely goal of that agents.
-In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which frequently becomes a too strong assumption when applying {\em plan recognition as planning} at real-world problems.
-In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the successful {\em plan recognition as planning} approach.
-The empirical evaluation evidences that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without {\em a priori} knowing the action model of the observed agents.
+{\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to estimate the most likely goal of the agents being observed.
+In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which can become a too strong requirement when applying {\em plan recognition as planning} on real-world problems.
+In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known beforehand) and it shows that this formulation neatly fits into the {\em plan recognition as planning} approach.
+The experimental results demonstrate that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without the requirement of having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents
 
@@
 The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
-Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$, i.e.~if its precondition holds in $s$.
+Then, $\rho(s,a)={\tt\small True}$ whenever $\pre(a)\subseteq s$ (i.e.~if its precondition holds in $s$) otherwise $\rho(s,a)={\tt\small False}$.
 The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
@@
 
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\in\mathcal{L}(F)$ is the set of goal conditions over the state variables.
-A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length} and $cost(\pi)=\sum_{a\in\pi} cost(a)$ its {\em plan cost}.
-The execution of $\pi$ on the initial state of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\subseteq\mathcal{L}(F)$ is the set of goal conditions over the state variables.
+A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ being the {\em length} of $\pi$ and $cost(\pi)=\sum_{a\in\pi} cost(a)$ being its {\em cost}.
+The execution of $\pi$ on the initial state of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $\rho(s_{i-1},a_i)={\tt\small True}$ and $s_i=\theta(s_{i-1},a_i)$.
 A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,P)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
@@
 \subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
+We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+Therefore, transiting between two consecutive observed states may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
+In other words, having an input observation does not imply knowing the actual length of the corresponding plan.
+
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define an interleaved combination of actions and states that represents the {\em observation from the execution of $\pi$ in $P$}.
 Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
-
 \begin{itemize}
@@
 The number of observed actions, $l$, ranges from $0$ (fully unobserved action sequence) to $|\pi|$ (fully observed action sequence).
-\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observed.
+\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observed states}, except for the initial state $s_0^o$, which is fully observed.
 A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
@@
 Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-The number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each {\em observed} states comprises $[1,|F|]$ fluents (the observation can still miss intermediate states that are {\em unobserved}).
+The number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each {\em observed} states comprises $[1,|F|]$ fluents ($\mathcal{O}(\tau)$ can still miss intermediate states that are {\em unobserved}).
 \end{itemize}
 
-We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
-Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-In other words, having an input observation $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
-
 \subsection{Goal recognition with classical planning}
-{\em Goal recognition} is a particular classification task in which each class represents a different possible goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents the observation of agents acting to achieve a goal $G\in G[\cdot]$.
+{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents the observation of agents acting to achieve a goal $G\in G[\cdot]$.
 
@@
 
-This expression is derived from the assumption that while the observed agents are not perfectly rational, they are more likely to follow cheaper plans, according to a {\em Boltzmann} distribution.
+This expression is derived from the assumption that while the observed agents are not perfectly rational, they are more likely to follow cheaper plans, according to a {\em Logistic} distribution.
 The larger the value of $\beta$, the more rational the agents, and the less likely that they will follow suboptimal plans.
-Recent works show fast estimates of the $P(\mathcal{O}|G)$ likelihood computed using relaxations of the classical planning tasks~\cite{pereira2017landmark}.
+Recent work exploits the structure of action {\em preconditions} and {\em effects} to compute fast estimates of the $P(\mathcal{O}|G)$ likelihood~\cite{pereira2017landmark}.
 
@@
 A solution to this task is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
-one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state meets the goal conditions, $G \subseteq s_n$.
+one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state $s_n$ meets the goal condition, $G \subseteq s_n$.
 

### 2019-02-11 doc/ijcai-19-goalRecognition/goalRecognition.tex
processing Miquels comments for pages 1 and 2 ijcai19 paper on goal recognition

@@
 \begin{abstract}
-This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em that the observer knows the action model of the observed agents}.
+This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em the observer knows the action model of the observed agents}.
 The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the {\em plan recognition as planning} approach for {\em goal recognition}.
-We show that this novel formulation allow us to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without knowing beforehand the action model of the observed agents.
+The experimental results demonstrate that this novel formulation for classical planning allow us to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without knowing beforehand the action model of the observed agents.
 \end{abstract}
@@
 {\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents pursuing one of those goals.
-While there exist a wide range of different approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2012plan} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
+While there exist diverse approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2010probabilistic} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
 
-{\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to estimate the most likely goal of the agents being observed.
+{\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to estimate the most likely goal of the agents being observed~\cite{ramirez2012plan}.
 In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which can become a too strong requirement when applying {\em plan recognition as planning} on real-world problems.
-In particular, the paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known beforehand) and it shows that this formulation neatly fits into the {\em plan recognition as planning} approach.
-The experimental results demonstrate that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without the requirement of having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents
+We introduce a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known beforehand) and we show that this formulation neatly fits into the {\em plan recognition as planning} approach.
+The experimental results demonstrate that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without requiring having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents
 
@@
 \label{sec:background}
-This section formalizes the {\em planning model} we follow, the kind of {\em observations} that input the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.
+This section formalizes the {\em classical planning} model that we follow in this work, the kind of {\em observations} that input the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.
 
@@
 A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ being the {\em length} of $\pi$ and $cost(\pi)=\sum_{a\in\pi} cost(a)$ being its {\em cost}.
-The execution of $\pi$ on the initial state of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, $\rho(s_{i-1},a_i)={\tt\small True}$ and $s_i=\theta(s_{i-1},a_i)$.
+The execution of $\pi$ on the initial state of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, then $\rho(s_{i-1},a_i)={\tt\small True}$ and $s_i=\theta(s_{i-1},a_i)$.
 A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,P)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
@@
 We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
-Therefore, transiting between two consecutive observed states may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
+This means that transiting between two consecutive observed states may require the execution of more than a single action.
 In other words, having an input observation does not imply knowing the actual length of the corresponding plan.
@@
 \subsection{Goal recognition with classical planning}
-{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents the observation of agents acting to achieve a goal $G\in G[\cdot]$.
+{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents an observation of agents acting to achieve a goal $G\in G[\cdot]$.
 
-Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
+Following the paradigm of the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
 \begin{align}
@@
 
-The higher the value of the $cost(\pi^\top)-cost(\pi^\bot)$ difference, the higher the probability of the observed agents to aim goal $G\in G[\cdot]$.
+The higher the value of the $\Delta(\pi^\top,\pi^\bot)$ cost difference, the higher the probability of the observed agents to aim goal $G\in G[\cdot]$.
 With this regard, {\em plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:
@@
 \begin{align}
-P(\mathcal{O}|G) = \frac{1}{1+e^{-\beta(cost(\pi^\top)-cost(\pi^\bot))}}
+P(\mathcal{O}|G) = \frac{1}{1+e^{-\beta \Delta(\pi^\top,\pi^\bot)}}
 \end{align}
@@
 This section introduces a novel formulation for classical planning in a setting where no action model is given.
-This setting has already shown related to the learning of action models for planning~\cite{SternJ17}.
-In particular it can be seen as an extreme scenario when the action model is learned from a single example that contains only two state observations: the initial state and the goals.
-A {\em classical planning with unknown domain models} is then a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
+This setting has already shown related to the learning of action models for planning~\cite{SternJ17} and it can be seen as an extreme scenario when the action model is {\em learned} from a single example that contains only two state observations: the initial state and the goals.
+
+A {\em classical planning with unknown domain models} is a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
+A {\em solution plan} is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
+one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)={\tt\small True}$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state $s_n$ meets the goal condition $G \subseteq s_n$.
 
-A solution to this task is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
-one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state $s_n$ meets the goal condition, $G \subseteq s_n$.
-
-Next we show that the space of possible STRIPS action models can be encoded as a set of propositional variables and a set of constraints over those variables.
-Then, we show how to exploit this encoding to solve $P=\tup{F,A[\cdot],I,G}$ problems with an off-the-shelf classical planner and the properties of this approach.
+Next we show that a propositional encoding for the space of STRIPS action models allows us to solve $P=\tup{F,A,I,G[\cdot]}$ problems with off-the-shelf classical planners.
 
 \subsection{A propositional encoding for the space of STRIPS action models}
-{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+{\em A \strips\ action schema} $\xi$ is defined by: A list of {\em parameters} $pars(\xi)$, and three sets of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters}, $pars(\xi)$.
-The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is the set of FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
+The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is the set of FOL interpretations of $\Psi$ over the parameters $pars(\xi)$, and is denoted as ${\mathcal I}_{\Psi,\xi}$.
 
-For instance in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for the {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+For instance in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for the case of the {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints of two kinds:
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints:
 \begin{enumerate}
@@
 
-In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ action models with classical planning.
+In this work we introduce a novel propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ that uses only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+This encoding exploits the syntactic constraints of \strips\, so it is more compact than previous encodings proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ actions with classical planning.
 In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
 If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema using the {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
+To illustrate this, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\em blocksworld} {\tt\small stack(?v1,?v2)} schema and our propositional representation for this schema using only {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
 
@@
 
-In addition, one can introduce {\em domain-specific knowledge} to constrain further the space of possible schemata.
+{\em Domain-specific knowledge} is also helpfull to constrain further the space of possible schemata.
 For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\it State invariants} are {\em domain-specific knowledge}that can be exploited either as {\em syntactic} constrains to reduce the space of possible action models but also as {\em semantic} constraints to complete partial observations of the states traversed by a plan.
+With this regard, {\it state invariants} can be exploited either as {\em syntactic} constraints (to reduce the space of possible action models) but also as {\em semantic} constraints (to complete partial observations of the states traversed by a plan)~\cite{fox:TIM:JAIR1998}.
 
 \subsection{A classical planning compilation for planning with unknown domain models}
-Now we show how we adapt the {\em classical planning compilation for learning \strips\ action models}~\cite{aineto2018learning} to address the task of {\em planning with unknown domain models}, using our propositional encoding of STRIPS action models.
+Inspired by the {\em classical planning compilation for learning \strips\ action models}~\cite{aineto2018learning}, we define here a compilation to address the task of {\em planning with unknown domain models} that uses our compact propositional encoding of STRIPS action models.
 
@@
 \begin{itemize}
-\item $F'$ extends $F$ with a fluent $mode_{insert}$, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
-This is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ such that $e\in{\mathcal I}_{\Psi,\xi}$ is a single element from the set of FOL interpretations of predicates $\Psi$ over the corresponding action parameters $pars(\xi)$.
+\item $F'$ extends $F$ with a fluent $mode_{insert}$, to indicate whether action models are being {\em programmed} or {\em applied}, and the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents for the propositional encoding of the corresponding space of STRIPS action models.
 
@@
 \begin{enumerate}
-\item Actions for {\em inserting} a {\em precondition}, {\em positive} effect or {\em negative} effect into the action model $\xi$ following the syntactic constraints of \strips\ models. 
+\item Actions for {\em inserting} a {\em precondition}, {\em positive}/{\em negative} effect into the action model $\xi$ following the syntactic constraints of \strips\ . 
 \begin{itemize}
@@
 \item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
-The action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
+Note that the action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
 \end{enumerate}
@@
 The problem goal is having the 3-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
-The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects), steps $07-13$ insert the preconditions and effects of the {\tt\small pickup} action model and finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$.
-Note that another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em putdown} and {\em unstack} action models and then applying instead the four step postfix {\tt\small (putdown blockB), (unstack blockB blockC), (putdown blockA), (unstack blockA blockB)}.
+The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $05-10$ insert the action model effects), steps $02-04$ insert the preconditions of the {\tt\small pickup} action model (while steps $10-13$ insert its effects) and finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$.
+Another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em unstack} and {\em putdown} action models and then applying instead the four step postfix {\tt\small (putdown blockB), (unstack blockB blockC), (putdown blockA), (unstack blockA blockB)}.
 
@@
 The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $\{pre\_e\_\xi, eff\_e\_\xi\}$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
-Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\Psi,\xi}|$ and hence, the size of the classical planning task output by the compilation.
 

### 2019-02-12 doc/ijcai-19-goalRecognition/goalRecognition.tex
processing Miquels comments for pages 1 and 2 ijcai19 paper on goal recognition

@@
 The intuition of the compilation is that the dynamics of the actions for {\em applying} an action model $\xi$ is determined by the values of the corresponding $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in the current state.
-For instance, executing {\tt{\small (apply\_stack blockB blockA)}} in a state $s$ implies activating the preconditions and effects of {\tt{\small apply\_stack}} according to the values of $\{pre\_e\_stack, eff\_e\_stack\}_{\forall e\in{\mathcal I}_{\Psi,stack}}$ fluents in $s$.
+For instance when executing {\tt{\small (apply\_stack blockB blockA)}} in a state $s$, its preconditions and effects are activated according to the values of the corresponding $\{pre\_e\_stack, eff\_e\_stack\}_{\forall e\in{\mathcal I}_{\Psi,stack}}$ fluents in $s$.
 This means that if the current state $s$ holds $\{{\tt\scriptsize (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}\} \subset s$, then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
 Otherwise, a different set of precondition literals will be checked for the stack action.
-The same applies to the positive and negative effects.
-Executing {\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state only if {\tt{\small stack}} has been correctly programmed by the {\em insert} actions.
+The same applies to the effects of {\tt{\small (apply\_stack blockB blockA)}}.
+Executing {\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state only if the $\{pre\_e\_stack, eff\_e\_stack\}_{\forall e\in{\mathcal I}_{\Psi,stack}}$ fluents has been correctly programmed by the corresponding {\em insert} actions.
 
@@
 Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving a $P'=\tup{F',A',I,G}$ classical planning problem output by our compilation.
-In the initial state of that problem three  blocks ({\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table and the robot hand is empty.
-The problem goal is having the 3-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
-The plan shows that the {\em insert} actions for the action model {\tt\small stack} (steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $05-10$ insert the action model effects), steps $02-04$ insert the preconditions of the {\tt\small pickup} action model (while steps $10-13$ insert its effects) and finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$.
-Another valid solution could be computed for instance, inserting the same preconditions and effects into the {\em unstack} and {\em putdown} action models and then applying instead the four step postfix {\tt\small (putdown blockB), (unstack blockB blockC), (putdown blockA), (unstack blockA blockB)}.
+In the initial state of that problem three  blocks ({\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table, the robot hand is empty.
+The problem goal is having the three-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
+The plan shows the {\em insert} actions for the {\tt\small stack} scheme (steps $00-01$ insert the preconditions, steps $05-10$ insert the effects), steps $02-04$ insert the preconditions of the {\tt\small pickup} scheme (while steps $10-13$ insert the effects of this scheme).
+Finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$ starting from $I$.
+Note that another valid solution could be computed for instance, by inserting the same preconditions and effects but into the {\em unstack} and {\em putdown} actions and then applying instead the four step postfix {\tt\small (putdown blockB), (unstack blockB blockC), (putdown blockA), (unstack blockA blockB)}.
 
-\subsection{The bias of an initially {\em empty} action model}
+\subsection{The bias of the initially {\em empty} action model}
 All $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our compilation.
-This fact might introduce a bias to the solutions of the $P=\tup{F,A[\cdot],I,G}$ classical planning task preferring solutions that imply action models with a smaller number of {\em preconditions}/{\em effects} (i.e., that imply a lower number of {\em insert} actions).
+Classical planners usually preffer shorter solution plans, so a bias may be introduced to the $P=\tup{F,A[\cdot],I,G}$ task preferring solutions that imply action models with a shorter number of {\em preconditions}/{\em effects} (i.e., classical plans that solve $P'=\tup{F',A',I,G}$ using a shorter number of {\em insert} actions).
 
-This bias could be eliminated defining a cost landscape where {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}.
-In practice, since classical planners are not proficiency optimizing cost landscapes of this kind, we use a different approach that disregard the cost of actions the {\em insert} actions.
-Our approach is to use SAT-based planning because it can apply all the required actions for inserting preconditions in a single planning step (these actions do not interact).
-Further, actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
-The SAT-based planning approach is also convenient because its ability to deal with populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to performance.
+This bias could be eliminated defining a cost landscape for the actions in $A'$ (e.g.
+{\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}).
+In practice, since classical planners are not proficiency optimizing cost landscapes of this kind, we use a different approach to disregard the cost of the {\em insert} actions.
+Our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
+Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
 
-Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers, one for inserting the action preconditions, another for inserting the action effects followed by N layers of the SATPLAN encoding extended to apply the action models determined by the previous two initial layers.
-With regard to the example of Figure~\ref{fig:plan-lplan} this means that steps [00-04] are applied in paralel in a first SATPLAN layer, steps [05-13] are applied in paralel in a second SATPLAN layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer.
+Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first one for inserting the action preconditions and a second one for inserting the action effects.
+These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers).
+Regarding again to the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so the example would be solved using just six layers).
+
+The SAT-based planning approach is also convenient for the task of {\em goal recognition as planning with unknown domain models} because its ability to deal with planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
 
 \subsection{Compilation properties}
-Now we present some properties of the compilation scheme.
-
 \begin{mylemma}
@@
 \item {\em Solve $P^{\bot}_G$} with a classical planner and extract $cost(\pi_\bot)$ as the length of the found solution plan.
-\item Compute the $cost(\pi^\top)-cost(\pi^\bot)$ difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
+\item Compute the $\Delta(\pi^\top,\pi^\bot)$ cost difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
 \end{enumerate}

### 2019-02-12 doc/ijcai-19-goalRecognition/goalRecognition.tex
processing Miquels comments for pages 1 and 2 ijcai19 paper on goal recognition

@@
 \subsection{The bias of the initially {\em empty} action model}
-All $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our compilation.
-Classical planners usually preffer shorter solution plans, so a bias may be introduced to the $P=\tup{F,A[\cdot],I,G}$ task preferring solutions that imply action models with a shorter number of {\em preconditions}/{\em effects} (i.e., classical plans that solve $P'=\tup{F',A',I,G}$ using a shorter number of {\em insert} actions).
+Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $P=\tup{F,A[\cdot],I,G}$ problems preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
+In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P'=\tup{F',A',I,G}$ compilation so classical planners tend to solve $P'$ with plans that require a shorter number of {\em insert} actions.
 
-This bias could be eliminated defining a cost landscape for the actions in $A'$ (e.g.
+This bias could be eliminated defining a cost function for the actions in $A'$ (e.g.
 {\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}).
-In practice, since classical planners are not proficiency optimizing cost landscapes of this kind, we use a different approach to disregard the cost of the {\em insert} actions.
-Our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
+In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} with zero-cost actions.
+Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
 Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
 
-Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first one for inserting the action preconditions and a second one for inserting the action effects.
-These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers).
-Regarding again to the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so the example would be solved using just six layers).
+Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects.
+These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers, the $\mathsf{apply_{\xi,\omega}}$ actions).
+Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
 
-The SAT-based planning approach is also convenient for the task of {\em goal recognition as planning with unknown domain models} because its ability to deal with planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
+The SAT-based planning approach is also convenient for the task of {\em goal recognition as planning with unknown domain models} because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
 
@@
 \begin{small}
-Once a given precondition or effect is inserted into an action model it can never be removed back and once an action model is applied it cannot be {\em reprogramed}.
-In the compiled problem the value of $F$, the fluents of the original problem, can exclusively be modified via $\mathsf{apply_{\xi,\omega}}$ actions.
-The set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$.
+Once a given precondition (or effect) is inserted into an action model it can never be removed back and once an action model is applied (via an $\mathsf{apply_{\xi,\omega}}$ action) it cannot longer be {\em programed}.
+The set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$ (the fluents of the original problem $F$ can exclusively be modified by the $\mathsf{apply_{\xi,\omega}}$ actions).
 This means that the action model used by the $\mathsf{apply_{\xi,\omega}}$ actions has to be consistent with the traversed intermediate states.
@@
 
-The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
+The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$ and the number of parameters of the action models, $|pars(\xi)|$.
 The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$.
-The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $\{pre\_e\_\xi, eff\_e\_\xi\}$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
+This term dominates the compilation size because it defines the $\{pre\_e\_\xi, eff\_e\_\xi\}$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects of the $\mathsf{apply_{\xi,\omega}}$ actions.
 

### 2019-02-13 doc/ijcai-19-deductivelearning/deductivelearning.tex
added comments after meeting with Diego and Eva to ijcai19 paper on deductive learning

@@
 
-\title{Learning \strips\ action models from {\em state-invariants}}
+\title{Learning action models from {\em state-invariants}}
 
@@
 \begin{abstract}
-This paper addresses the learning of action models from {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain).
-The benefit of exploiting {\em state-invariants} is two-fold, they allow to reduce the space of possible action models and to complete learning examples that are only partially observed.
-Our approach for the learning of \strips\ action models from {\em state-invariants} is a {\em classical planning} compilation.
-The compilation is flexible to different kinds of input knowledge (e.g., partially observations of plan executions including partially observed intermediate states and/or actions) and outputs an action model that is {\em consistent} with the given input knowledge.
-The experimental results show that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} are helpful to learn good quality \strips\ action models.
+This paper addresses the learning of \strips\ action models from {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain).
+The benefit of exploiting {\em state-invariants} is two-fold, they constrain the space of possible action models and they can complete learning examples that are only partially observed.
+Our approach for learning \strips\ action from {\em state-invariants} is a {\em classical planning} compilation that is flexible to different kinds of input knowledge (e.g., partially observations of plan executions including partially observed intermediate states and/or actions).
+The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} are helpful to learn good quality \strips\ action models.
 \end{abstract}
@@
 
-The specification of planning action models is a complex process that limits, too often, the application of {\em model-based planning} systems to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
-The {\em machine learning} of action models can relieve the {\em knowledge acquisition bottleneck} of planning and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}.
-Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, meaning that their performance is linked to the {\em amount} and {\em quality} of the input learning examples.
+The specification of planning action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
+The {\em machine learning} of action models relieves this {\em knowledge acquisition bottleneck} of {\em model-based planning} and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}.
+Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, linking learning performance exclusively to the {\em amount} and {\em quality} of the input learning examples.
 
-This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge, to cushion the negative impact of insufficient learning examples.
-In more detail our approach leverages {\em state-invariants}, i.e.
-logic formulae that specify constraints about the possible states of a given domain.
-Given an action model, state-of-the-art planners are used to infer {\em state-invariants} from that model to reduce search spaces and make the planning process more efficient~\cite{helmert2009concise}.
+This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge.
+Our approach leverages {\em state-invariants}, i.e.
+logic formulae that specify constraints about the possible states of a given domain, to cushion the negative impact of insufficient learning examples.
+Given an action model, state-of-the-art planners infer {\em state-invariants} from that model to reduce the search space and make the planning process more efficient~\cite{helmert2009concise}.
 In this paper we follow the opposite direction and leverage {\em state-invariants} to learn the planning action model.
-The benefit of learning \strips\ action models from {\em state-invariants} is two-fold, {\em state-invariants} allow us to reduce the space of possible action models and to complete learning examples that are only partially observed.
+The benefit of learning action models from {\em state-invariants} is two-fold, {\em state-invariants} constrain the space of possible action models and can complete learning examples that are only partially observed.
 
@@
 Our compilation is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
-The experimental results show that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models than with the existing {\em classical planning} compilation by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models.
 
@@
 \label{sec:background}
-This section formalizes the {\em classical planning model} we follow in this work and the kind of {\em knowledge} that can be given as input to the task of learning \strips\ action models.
+This section formalizes the {\em classical planning model} we follow in this work and introduces the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
+Finally, the section formalizes {\em state-invariants}.
 
@@
 
+\subsection{Learning action models with classical planning}
+This work is built on top of the {\em classical planning compilation} for the learning \strips\ action models~\cite{aineto2018learning}.
+This compilation receives as input an empty model $\mathcal{M}$ (which only contains the action headers), and an observation of a plan execution $\mathcal{O}(\tau)$ (extensible to a set of observations).
+The compilation outputs a model $\mathcal{M'}$ that specifies the preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of $\mathcal{O}(\tau)$ following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1}^o,a_i)$ for every observed action of $\mathcal{O}(\tau)$ and $s_i^o=\theta(s_{i-1}^o,a_i)$ for every observed state of $\mathcal{O}(\tau)$.
+
+Essentially, a solution plan to the classical planning problem that results from the compialtion is a sequence of: (a) \emph{insert actions} that insert preconditions and effects on the schemata of $\mathcal{M}$ to build $\mathcal{M'}$ and (b) \emph{apply actions} that validate the application of the $\mathcal{M'}$ model in the observation $\mathcal{O(\tau)}$.
+
+To illustrate this, Figure~\ref{fig:plan-lplan} shows a solution to a classical planning problem resulting from the~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} compilation.
+In the initial state of that problem three  blocks ({\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table, the robot hand is empty.
+The problem goal is having the three-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
+The plan shows the {\em insert} actions for the {\tt\small stack} scheme (steps $00-01$ insert the preconditions, steps $05-10$ insert the effects), steps $02-04$ insert the preconditions of the {\tt\small pickup} scheme (while steps $10-13$ insert the effects of this scheme).
+Finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$ starting from $I$.
+
+\begin{figure}[hbt!]
+	{\tiny\tt
+\begin{tabular}{ll}
+		{\bf 00}:(insert\_pre\_stack\_holding\_v1) & {\bf 10}:(insert\_eff\_pickup\_clear\_v1) \\
+		01:(insert\_pre\_stack\_clear\_v2) & 11:(insert\_eff\_pickup\_ontable\_v1)\\
+                {\bf 02}:(insert\_pre\_pickup\_handempty) & 12:(insert\_eff\_pickup\_handempty)\\
+                03:(insert\_pre\_pickup\_clear\_v1) & 13:(insert\_eff\_pickup\_holding\_v1)\\
+                04:(insert\_pre\_pickup\_ontable\_v1) & {\bf 14}:(apply\_pickup blockB)\\
+                {\bf 05}:(insert\_eff\_stack\_clear\_v1) & 15:(apply\_stack blockB blockC)\\
+                06:(insert\_eff\_stack\_clear\_v2) & 16:(apply\_pickup blockA)\\
+                07:(insert\_eff\_stack\_handempty) & 17:(apply\_stack blockA blockB) \\
+                08:(insert\_eff\_stack\_holding\_v1) &  {\bf 18}:(validate\_1)\\
+                09:(insert\_eff\_stack\_on\_v1\_v2) &             		 
+\end{tabular}
+}
+	\caption{\small Plan computed when solving a problem output by the classical planning compilation.}
+	\label{fig:plan-lplan}
+\end{figure}
+
 \subsection{State-invariants}
@@
 If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
-For instance, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
+For example, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
 
@@
 Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
-For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em state invariants} for the {\em blocksworld} planning domain~\cite{slaney2001blocks}.
-There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard (as hard as testing plan existence~\cite{bylander:complexity:AIJ1994}).
+For instance, Figure~\ref{fig:strongest-invariant} shows five clauses that define {\em state invariants} for the {\em blocksworld} planning domain~\cite{slaney2001blocks}.
 
@@
 
-A {\em domain invariant} is an instance-independent invariant, i.e.
-holds for any possible initial state and set of objects.
+A {\em domain invariant} is an instance-independent state-invariant, i.e.
+holds for any possible initial state and any possible set of objects.
 Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
 Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
-For instance, $\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
+
+We define a {\em domain mutex} as a $\tup{p,q}$ predicates pair where both $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of fluents $F$ of a given planning problem and such that they satisfy the following formulae $p\rightarrow \neg q$ where the predicate variables are universally quantified.
+For instance, predicates $holding(x)$ and $clear(x)$ from the {\em blocksworld} are {\em domain mutex} while predicates $clear(x)$ and $ontable(x)$ are not ($\forall x\ clear(x)\leftrightarrow\neg ontable(x)$ does not hold for every possible {\em blocksworld} state).
+We pay special attention to {\em domain mutex} because they identify the {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effectively pruning of inconsistent \strips\ action models and (2) effective completion of partially observed states.
 
@@
 \label{sec:learning}
-We define the task of learning a planning action model from {\em state-invariants} as a tuple $\Lambda=\tup{P,\Phi,M}$, where:
-\begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G}$, is a {\em classical planning problem} where $A[\cdot]$ is a set of actions s.t., the {\em dynamics} of each action $a\in A[\cdot]$ is {\em unknown} (i.e.
-functions $\rho$ and/or $\theta$ are undefined for $a\in A[\cdot]$).
-\item $\Phi$ is a set of {\em state-invariants} that define constraints about the set of possible states in the previous planning problem $P$.
-\item $M$ is the {\em space of possible action models} for the $A[\cdot]$ actions (i.e., the set of possible specifications of the $\rho$ and/or $\theta$ functions for each $a\in A[\cdot]$ action).
-\end{itemize}
+First, this section defines the sampling space and the space of possible action models. Then, the section formalizes the task of learning \strips\ action models from {\em state-invariants}.
 
-We say that a given model $\mathcal{M}\in M$ is a {\em solution} to the $\Lambda=\tup{P,\Phi,M}$ learning task iff there exists a plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, when the semantics of each action $a\in A[\cdot]$ is given by $\mathcal{M}$, and such that any state traversed by a trajectory $\tau(\pi,P)$ is {\em consistent} with the input set of {\em state-invariants} $\Phi$.
+\subsection{The sampling space}
+We define a {\em learning example} $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ as a sequence of partial states, except for the initial state $s_0^o$, which is fully observed.
+A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
+Each {\em observed} states comprises $[1,|F|]$ fluents.
+The observation can still miss intermediate states that are {\em unobserved} so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
 
-Next, we show that the set $M$, of possible action models, can be compactly encoded as a set of propositional variables and a set of constraints over those variables.
-Then, we show how to exploit this compact encoding to solve a $\Lambda=\tup{P,\Phi,M}$ learning task with an off-the-shelf classical planner.
+Our sampling space follows the {\em open world} assumption, i.e.
+what is not observed is considered unknown.
+{\em State-invariants} are helpful to infer new knowledge that was unobserved in the learning examples.
+Given a {\em domain mutex} $\tup{p,q}$ and a state observation $s_j^o\in \mathcal{O}(\tau)$, {\small $(1\leq j\leq m)$}, such that literal $p(\omega)\in s_j^o$ is an instantiation of predicate $p$ over some subset of objects $\omega\subseteq\Omega^{|pars(p)|}$ then, the state observation $s_j^o$ can be safely completed adding the new literal $\neg q(\omega)$ (despite $\neg q(\omega)$ was actually unobserved).
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we know the {\em domain mutex} $\forall x\ holding(x)\leftrightarrow\neg clear(x)$ we can safely extend the observation with literal {\tt\small $\neg$clear(blockA)} (despite this literal was actually unobserved).
 
-\subsection{A propositional encoding for the space of STRIPS action models}
-{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+\subsection{The space of \strips\ action models}
+{\em A \strips\ action schema} $\xi$ is defined by: A list of {\em parameters} $pars(\xi)$, and three sets of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters}, $pars(\xi)$.
@@
 
-In this work we introduce a propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ using only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
+In this work we introduce a novel propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ that uses only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
 This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ action models with classical planning.
@@
 
-In addition, one can introduce {\em domain-specific knowledge} to constrain further the space of possible schemata.
+One can also introduce {\em domain-specific knowledge} to constrain further the space of possible schemata.
 For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\it State invariants} are {\em domain-specific knowledge} and they can be seen as either {\em syntactic} or {\em semantic} constraints because on the one hand, they constrain the space of possible action models but on the other hand, they can be used to complete partial observations of the states traversed by a plan.
+{\it State invariants} are {\em domain-specific knowledge} that can be seen either as {\em syntactic} or {\em semantic} constraints.
+On the one hand, {\it state invariants} constrain the space of possible action models but on the other hand, they can complete partial observations of the states traversed by a plan.
 
-\subsection{Learning \strips\ action models with classical planning}
-Our approach for computing an action model $\mathcal{M}\in M$ that solves the $\Lambda=\tup{P,\Phi,M}$ learning task is to build and solve a classical planning problem $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I,G_{\Lambda}}$ such that:
+\subsection{The learning task}
+We define the task of learning a planning action model from {\em state-invariants} as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, where:
+\begin{itemize}
+\item $\mathcal{M}$ is the {\em initial empty model} that contains only the name, $name(\xi)$, and parameters, $pars(\xi)$, of each action model $\xi$ to be learned.
+\item $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ is a single learning example.
+This example can be reduced to its minimal expression $\mathcal{O}^*=\tup{s_0^o, s_m^o}$ meaning that it only contains two state observations, a full initial state $s_0^o$ and a partially observed state $s_m^o$.
+Note that the set of predicates $\Psi$ is deducible from $\mathcal{O}$ since $s_0^o$ is a fully observed state.
+\item $\Phi$ is a set of {\em state-invariants} that define constraints about the set of possible states.
+\end{itemize}
+
+$M$ is the {\em space of possible action models} for the $A[\cdot]$ actions (i.e., the set of possible specifications of the $\rho$ and/or $\theta$ functions for each $a\in A[\cdot]$ action).
+We say that a given model $\mathcal{M}\in M$ is a {\em solution} to the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task iff there exists a plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, when the semantics of each action $a\in A[\cdot]$ is given by $\mathcal{M}$, and such that any state traversed by a trajectory $\tau(\pi,P)$ is {\em consistent} with the input set of {\em state-invariants} $\Phi$.
+
+Now, we show how to exploit our compact encoding of \strips\ action models to solve a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task with an off-the-shelf classical planner.
+
+\section{Learning action models from {\em state-invariants} with classical planning}
+\label{sec:compilation}
+Given a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task we build and solve a classical planning problem $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I,G_{\Lambda}}$ such that:
 \begin{itemize}
@@
 \subsection{Pruning inconsistent action models with {\em domain mutex}}
-We define a {\em domain mutex} as a $(p,q)$ predicates pair where both $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of fluents $F$ of a given planning problem and such that they satisfy the following formulae $p\leftrightarrow \neg q$ where are the predicate variables are universally quantified.
-For instance, predicates $holding(x)$ and $clear(x)$ from the {\em blocksworld} are {\em domain mutex} since they satisfy $\forall x\ holding(x)\leftrightarrow\neg clear(x)$ while predicates $clear(x)$ and $ontable(x)$ (also from the {\em blocksworld}) are not {\em domain mutex} because they do not always satisfy $\forall x\ clear(x)\leftrightarrow\neg ontable(x)$.
-
-We pay attention to this particular class of {\em state-invariants} because they define the {\em state-properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable an effectively pruning of inconsistent \strips\ action models.
 Our approach to implement this pruning is extending the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions (i.e., the actions that determine a solution model $\mathcal{M}$) with extra conditional effects indicating that the programmed model is {\em invalid} (i.e., inconsistent with a {\em domain mutex} in $\Phi$).
@@
 
+\subsection{The bias of the initially {\em empty} action model}
+Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $P=\tup{F,A[\cdot],I,G}$ problems preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
+In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P'=\tup{F',A',I,G}$ compilation so classical planners tend to solve $P'$ with plans that require a shorter number of {\em insert} actions.
+
+This bias could be eliminated defining a cost function for the actions in $A'$ (e.g.
+{\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}).
+In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} with zero-cost actions.
+Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
+Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
+
+Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects.
+These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers, the $\mathsf{apply_{\xi,\omega}}$ actions).
+Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
+
+The SAT-based planning approach is also convenient because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
+
 \subsection{Compilation properties}
 \begin{mylemma}
-Soundness. Any classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ produces a \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi,M}$ learning task.
+Soundness. Any classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ produces a \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi}$ learning task.
 \end{mylemma}
@@
 
-\subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
-Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
-
-\begin{itemize}
-\item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
-The number of observed actions, $l$, ranges from $0$ (fully unobserved action sequence) to $|\pi|$ (fully observed action sequence).
-\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observed.
-A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
-Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
-Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-The number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each {\em observed} states comprises $[1,|F|]$ fluents (the observation can still miss intermediate states that are {\em unobserved}).
-\end{itemize}
-
-We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
-Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-In other words, having an input observation $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
-
-\subsection{Completing observations with {\em domain mutex}}
-Our observation model follows the {\em open world} assumption, in other words, what is not observed is considered unknown.
-Here, we show that {\em state-invariants} are helpful to infer new knowledge that was unobserved.
-
-Given a {\em domain mutex} $(p,q)$ and a state observation $s_j^o\in \mathcal{O}(\tau)$, {\small $(1\leq j\leq m)$}, such that the literal $p(\omega)\in s_j^o$ is an instantiation of predicate $p$ over some subset of objects $\omega\subseteq\Omega^{|pars(p)|}$ then, the state observation can be safely completed adding the new literal $\neg q(\omega)$ (despite $\neg q(\omega)$ was actually unobserved).
-For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we have the {\em domain mutex} $\forall x\ holding(x)\leftrightarrow\neg clear(x)$ in the input set $\Phi$ of {\em state-invariants} we can safely add to the observation the literal {\tt\small $\neg$clear(blockA)} (despite this literal was actually unobserved).
-The process is repeated for all the observed states in $\mathcal{O}(\tau)$ and all the {\em domain mutex} in $\Phi$ to produce a new completed observation $\mathcal{O}(\tau)'$.
-
-\subsection{Learning from completed observations with {\em classical planning}}
-Let be $\mathcal{O}(\tau)'$ an observation completed as explained above, we extend here our compilation to constraint the possible \strips\ models with $\mathcal{O}(\tau)$':
-\begin{itemize}
-\item One fluent $\{observed_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)'$ state observation.
-Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions in $\mathcal{O}(\tau)'$.
-The former is used to ensure that actions are executed in the same order as they are observed.
-The latter is used to iterate to the next planning step when solving $P_{\Lambda}$.
-\item Adding $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i \leq n$} to the initial state and $observed_m$ to the goals $G$ of the classical planning problem and hence, constrain the solution plans to be consistent with all the state observations.
-\item Adding the extra conditional effects {\small $\{at_{i},plan(name(a_i),\Omega^{pars(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$} to the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\mathcal{O}(\tau)'$.
-\item Actions for {\em validating} the partially observed state $s_j^o\in\mathcal{O}(\tau)'$, {\tt\small $1\leq j< m$}.
-These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observable data of the input observation $\mathcal{O}(\tau)$' follows after the execution of the apply actions.
-\item One $\mathsf{validate_{j}}$ action to constraint the solution plans to be consistent with the $s_j^o\in\mathcal{O}(\tau)'$ input state observation, {\small $(1\leq j\leq m)$}.  
-\end{itemize}
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{observed_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg observed_{j-1}, observed_j\}.
-\end{align*}
-\end{small}
-
-So far we explained the extension of the compilation for learning from a single observation $\mathcal{O}(\tau)'$.
-The extension to the more general case of a set of observation $\{\mathcal{O}(\tau_1),\ldots,\mathcal{O}(\tau_k)\}$ is implemented with a small modification.
-In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_m^o\in \mathcal{O}(\tau_t)$, {\tt\small $1\leq t\leq k$} reset also the current state and the current plan step.
-
 \section{Evaluation}

### 2019-02-14 doc/ijcai-19-deductivelearning/deductivelearning.tex
updates added to ijcai19 paper on deductive learning

@@
 The benefit of exploiting {\em state-invariants} is two-fold, they constrain the space of possible action models and they can complete learning examples that are only partially observed.
-Our approach for learning \strips\ action from {\em state-invariants} is a {\em classical planning} compilation that is flexible to different kinds of input knowledge (e.g., partially observations of plan executions including partially observed intermediate states and/or actions).
-The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} are helpful to learn good quality \strips\ action models.
+Our approach for learning \strips\ action from {\em state-invariants} is a {\em classical planning} compilation that is flexible to different sources of input knowledge including partial observations of plan executions and {\em state-invariants}.
+The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (a single learning example that comprises just a full initial state and a partially observed state), {\em state-invariant} are helpful to learn good quality \strips\ action models.
 \end{abstract}
@@
 \label{sec:introduction}
-
-The specification of planning action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
+The specification of action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
 The {\em machine learning} of action models relieves this {\em knowledge acquisition bottleneck} of {\em model-based planning} and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}.
-Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, linking learning performance exclusively to the {\em amount} and {\em quality} of the input learning examples.
+Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, linking learning performance exclusively to the {\em amount} and {\em quality} of the input learning examples (which typically are observation of plan executions).
 
 This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge.
-Our approach leverages {\em state-invariants}, i.e.
-logic formulae that specify constraints about the possible states of a given domain, to cushion the negative impact of insufficient learning examples.
+Our approach leverages {\em state-invariants} (i.e.
+logic formulae that specify constraints about the possible states of a given domain) to cushion the negative impact of insufficient learning examples.
 Given an action model, state-of-the-art planners infer {\em state-invariants} from that model to reduce the search space and make the planning process more efficient~\cite{helmert2009concise}.
@@
 
-Our approach for learning \strips\ action models from {\em state-invariants} is compile the learning task into a classical planning task.
-Our compilation is flexible to different kinds of input knowledge (e.g., partially/fully observations of actions of plan executions as well as partially/fully observed intermediate states) and outputs an action model that is {\em consistent} with the given input knowledge.
-The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (just an {\em initial state} and the {\em goals}), {\em state-invariant} help to learn better \strips\ models.
+Our approach for learning \strips\ action models from {\em state-invariants} is to compile the learning task into a classical planning task.
+Our compilation is built on top of the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} and it is flexible to different kinds of input knowledge including both partial observations of plan executions and {\em state-invariants}.
+The compilation outputs an \strips\ action model that is {\em consistent} with all the given input knowledge.
+The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (a single learning example that comprises just a full initial state and a partially observed state), {\em state-invariant} are helpful to learn good quality \strips\ action models.
 
@@
 \label{sec:background}
-This section formalizes the {\em classical planning model} we follow in this work and introduces the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
+This section formalizes the {\em planning model} we follow in this work and introduces the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
 Finally, the section formalizes {\em state-invariants}.
@@
 \subsection{Learning action models with classical planning}
-This work is built on top of the {\em classical planning compilation} for the learning \strips\ action models~\cite{aineto2018learning}.
-This compilation receives as input an empty model $\mathcal{M}$ (which only contains the action headers), and an observation of a plan execution $\mathcal{O}(\tau)$ (extensible to a set of observations).
-The compilation outputs a model $\mathcal{M'}$ that specifies the preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of $\mathcal{O}(\tau)$ following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1}^o,a_i)$ for every observed action of $\mathcal{O}(\tau)$ and $s_i^o=\theta(s_{i-1}^o,a_i)$ for every observed state of $\mathcal{O}(\tau)$.
+The {\em classical planning compilation} for the learning \strips\ action models~\cite{aineto2018learning} receives as input an empty model $\mathcal{M}$ (which contains just the action headers), and an observation of a set of observations of plan executions.
+The compilation outputs a model $\mathcal{M'}$ that specifies the preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of the observations following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1}^o,a_i)$ for every observed action and $s_i^o=\theta(s_{i-1}^o,a_i)$ for every observed state.
 
-Essentially, a solution plan to the classical planning problem that results from the compialtion is a sequence of: (a) \emph{insert actions} that insert preconditions and effects on the schemata of $\mathcal{M}$ to build $\mathcal{M'}$ and (b) \emph{apply actions} that validate the application of the $\mathcal{M'}$ model in the observation $\mathcal{O(\tau)}$.
-
-To illustrate this, Figure~\ref{fig:plan-lplan} shows a solution to a classical planning problem resulting from the~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} compilation.
-In the initial state of that problem three  blocks ({\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table, the robot hand is empty.
+A solution plan to the classical planning problem that results from the compialtion is a sequence of: (a) \emph{insert actions}, that insert preconditions and effects on a schemata of $\mathcal{M}$ to build $\mathcal{M'}$ and (b) \emph{apply actions} that validate the application of the $\mathcal{M'}$ model in the input observations.
+Figure~\ref{fig:plan-lplan} shows a solution to a classical planning problem resulting from the~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} compilation corresponding to the {\em blocksworld}~\cite{slaney2001blocks}.
+In the initial state of that problem the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table.
 The problem goal is having the three-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
-The plan shows the {\em insert} actions for the {\tt\small stack} scheme (steps $00-01$ insert the preconditions, steps $05-10$ insert the effects), steps $02-04$ insert the preconditions of the {\tt\small pickup} scheme (while steps $10-13$ insert the effects of this scheme).
-Finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$ starting from $I$.
+The plan shows the {\em insert} actions for the {\tt\small stack} scheme (steps $00-01$ insert the preconditions, steps $05-10$ insert the effects), the plan steps $02-04$ that  insert the preconditions of the {\tt\small pickup} scheme and steps $10-13$ that insert the effects of this scheme.
+Finally, steps $14-17$ is a plan postfix with actions that apply the programmed model to achieve the goals starting from the given initial state.
 
@@
 }
-	\caption{\small Plan computed when solving a problem output by the classical planning compilation.}
+	\caption{\small Example of a solution to a problem output by the classical planning compilation for the learning \strips\ action models.}
 	\label{fig:plan-lplan}
@@
 Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
-For instance, Figure~\ref{fig:strongest-invariant} shows five clauses that define {\em state invariants} for the {\em blocksworld} planning domain~\cite{slaney2001blocks}.
+For instance, Figure~\ref{fig:strongest-invariant} shows five clauses that define {\em state invariants} for the {\em blocksworld} planning domain.
 
@@
 
-We define a {\em domain mutex} as a $\tup{p,q}$ predicates pair where both $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of fluents $F$ of a given planning problem and such that they satisfy the following formulae $p\rightarrow \neg q$ where the predicate variables are universally quantified.
-For instance, predicates $holding(x)$ and $clear(x)$ from the {\em blocksworld} are {\em domain mutex} while predicates $clear(x)$ and $ontable(x)$ are not ($\forall x\ clear(x)\leftrightarrow\neg ontable(x)$ does not hold for every possible {\em blocksworld} state).
-We pay special attention to {\em domain mutex} because they identify the {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effectively pruning of inconsistent \strips\ action models and (2) effective completion of partially observed states.
-
 \section{Learning \strips\ action models from {\em state-invariants}}
@@
 \subsection{The sampling space}
-We define a {\em learning example} $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ as a sequence of partial states, except for the initial state $s_0^o$, which is fully observed.
-A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+We define a {\em learning example} as a sequence $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ of partial states, except for the initial state $s_0^o$, which is fully observed.
+Intermediate states can be missing meaning that are {\em unobserved}, so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
+A partially observed state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
 Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
-Each {\em observed} states comprises $[1,|F|]$ fluents.
-The observation can still miss intermediate states that are {\em unobserved} so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-
-Our sampling space follows the {\em open world} assumption, i.e.
-what is not observed is considered unknown.
-{\em State-invariants} are helpful to infer new knowledge that was unobserved in the learning examples.
-Given a {\em domain mutex} $\tup{p,q}$ and a state observation $s_j^o\in \mathcal{O}(\tau)$, {\small $(1\leq j\leq m)$}, such that literal $p(\omega)\in s_j^o$ is an instantiation of predicate $p$ over some subset of objects $\omega\subseteq\Omega^{|pars(p)|}$ then, the state observation $s_j^o$ can be safely completed adding the new literal $\neg q(\omega)$ (despite $\neg q(\omega)$ was actually unobserved).
-For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we know the {\em domain mutex} $\forall x\ holding(x)\leftrightarrow\neg clear(x)$ we can safely extend the observation with literal {\tt\small $\neg$clear(blockA)} (despite this literal was actually unobserved).
 
@@
 \item $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ is a single learning example.
-This example can be reduced to its minimal expression $\mathcal{O}^*=\tup{s_0^o, s_m^o}$ meaning that it only contains two state observations, a full initial state $s_0^o$ and a partially observed state $s_m^o$.
-Note that the set of predicates $\Psi$ is deducible from $\mathcal{O}$ since $s_0^o$ is a fully observed state.
+This example can be reduced to its minimal expression $\mathcal{O}^*=\tup{s_0^o, s_m^o}$ comprissing at least two state observations, a full initial state $s_0^o$ and a partially observed state $s_m^o$.
+The set of predicates $\Psi$ that shape the setf of fluents $F$ is deducible from $\mathcal{O}$ since $s_0^o$ is a fully observed state.
 \item $\Phi$ is a set of {\em state-invariants} that define constraints about the set of possible states.
@@
 
-$M$ is the {\em space of possible action models} for the $A[\cdot]$ actions (i.e., the set of possible specifications of the $\rho$ and/or $\theta$ functions for each $a\in A[\cdot]$ action).
-We say that a given model $\mathcal{M}\in M$ is a {\em solution} to the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task iff there exists a plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, when the semantics of each action $a\in A[\cdot]$ is given by $\mathcal{M}$, and such that any state traversed by a trajectory $\tau(\pi,P)$ is {\em consistent} with the input set of {\em state-invariants} $\Phi$.
-
-Now, we show how to exploit our compact encoding of \strips\ action models to solve a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task with an off-the-shelf classical planner.
+Next section shows how to exploit our compact encoding of \strips\ action models to solve a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task with an off-the-shelf classical planner.
 
@@
 \label{sec:compilation}
+
+Our sampling space follows the {\em open world} assumption, i.e.
+what is not observed is considered unknown.
+{\em State-invariants} are helpful to infer new knowledge that was unobserved in the learning examples.
+Given a {\em domain mutex} $\tup{p,q}$ and a state observation $s_j^o\in \mathcal{O}(\tau)$, {\small $(1\leq j\leq m)$}, such that literal $p(\omega)\in s_j^o$ is an instantiation of predicate $p$ over some subset of objects $\omega\subseteq\Omega^{|pars(p)|}$ then, the state observation $s_j^o$ can be safely completed adding the new literal $\neg q(\omega)$ (despite $\neg q(\omega)$ was actually unobserved).
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we know the {\em domain mutex} $\forall x\ holding(x)\leftrightarrow\neg clear(x)$ we can safely extend the observation with literal {\tt\small $\neg$clear(blockA)} (despite this literal was actually unobserved).
+
 Given a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task we build and solve a classical planning problem $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I,G_{\Lambda}}$ such that:
@@
 \subsection{Pruning inconsistent action models with {\em domain mutex}}
+We define a {\em domain mutex} as a $\tup{p,q}$ predicates pair where both $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of fluents $F$ of a given planning problem and such that they satisfy the following formulae $p\rightarrow \neg q$ where the predicate variables are universally quantified.
+For instance, predicates $holding(x)$ and $clear(x)$ from the {\em blocksworld} are {\em domain mutex} while predicates $clear(x)$ and $ontable(x)$ are not ($\forall x\ clear(x)\leftrightarrow\neg ontable(x)$ does not hold for every possible {\em blocksworld} state).
+We pay special attention to {\em domain mutex} because they identify the {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effectively pruning of inconsistent \strips\ action models and (2) effective completion of partially observed states.
+
 Our approach to implement this pruning is extending the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions (i.e., the actions that determine a solution model $\mathcal{M}$) with extra conditional effects indicating that the programmed model is {\em invalid} (i.e., inconsistent with a {\em domain mutex} in $\Phi$).

### 2019-02-14 doc/ijcai-19-deductivelearning/deductivelearning.tex
first full version (no experiments though) ijcai19 paper on deductive learning

@@
 %%%%%%%%%%%%%%%%%% Added for this paper
+\usepackage{enumitem}
 \usepackage{amsthm}
@@
 
+%The {\em classical planning compilation} for the learning \strips\ action models~\cite{aineto2018learning} can be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects.
+These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers).
+Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
+
 \subsection{State-invariants}
@@
 
-{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
-For instance, Figure~\ref{fig:strongest-invariant} shows five clauses that define {\em state invariants} for the {\em blocksworld} planning domain.
+{\em State-invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state-invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
+For instance, Figure~\ref{fig:strongest-invariant} shows five clauses that define {\em state-invariants} for the {\em blocksworld} planning domain.
 
@@
 
-A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+A {\em mutex} (mutually exclusive) is a {\em state-invariant} that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
 For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
@@
 Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
-Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
+Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants~\cite{rintanen:schematicInvariants:AAAI2017}).
 
@@
 \subsection{The sampling space}
-We define a {\em learning example} as a sequence $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ of partial states, except for the initial state $s_0^o$, which is fully observed.
+We define a {\em learning example} as a sequence $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ of partial states, except for the initial state $s_0^o$ which is fully observed.
 Intermediate states can be missing meaning that are {\em unobserved}, so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-A partially observed state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+A partially observed state $s_i^o$, ${\small 1\leq i\leq m}$, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
 Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
@@
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints of two kinds:
+Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by:
 \begin{enumerate}
@@
 \item {\bf Observation constraints}.
-The observation of the actions and states resulting from the execution of a plan depicts {\em semantic knowledge} that constraints further the space of possible action schemata.
+The {\em learning examples}, that are observations of plan executions, depict {\em semantic knowledge} that constraints further the space of possible action schemata.
 \end{enumerate}
@@
 For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\it State invariants} are {\em domain-specific knowledge} that can be seen either as {\em syntactic} or {\em semantic} constraints.
-On the one hand, {\it state invariants} constrain the space of possible action models but on the other hand, they can complete partial observations of the states traversed by a plan.
+{\it State-invariants} are {\em domain-specific knowledge} that can be used to define new {\em syntactic} and {\em semantic} constraints over the space of possible action schemata.
 
@@
 This example can be reduced to its minimal expression $\mathcal{O}^*=\tup{s_0^o, s_m^o}$ comprissing at least two state observations, a full initial state $s_0^o$ and a partially observed state $s_m^o$.
-The set of predicates $\Psi$ that shape the setf of fluents $F$ is deducible from $\mathcal{O}$ since $s_0^o$ is a fully observed state.
+The set of predicates $\Psi$ and objects $\Omega$ that shape the set of fluents $F$ is deducible from $\mathcal{O}$ since $s_0^o$ is a fully observed state.
 \item $\Phi$ is a set of {\em state-invariants} that define constraints about the set of possible states.
@@
 
-Next section shows how to exploit our compact encoding of \strips\ action models to solve a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task with an off-the-shelf classical planner.
+A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ is a model $\mathcal{M}'$ that is consistent with the the {\em initial empty model} $\mathcal{M}$, the learning example $\mathcal{O}$ and the set of {\em state invariants} in $\Phi$.
 
@@
 \label{sec:compilation}
+This section shows how to exploit our compact encoding of \strips\ action models to solve a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task with an off-the-shelf classical planner.
 
+\subsection{Completing partially observed states with {\em schematic mutex}}
 Our sampling space follows the {\em open world} assumption, i.e.
 what is not observed is considered unknown.
-{\em State-invariants} are helpful to infer new knowledge that was unobserved in the learning examples.
-Given a {\em domain mutex} $\tup{p,q}$ and a state observation $s_j^o\in \mathcal{O}(\tau)$, {\small $(1\leq j\leq m)$}, such that literal $p(\omega)\in s_j^o$ is an instantiation of predicate $p$ over some subset of objects $\omega\subseteq\Omega^{|pars(p)|}$ then, the state observation $s_j^o$ can be safely completed adding the new literal $\neg q(\omega)$ (despite $\neg q(\omega)$ was actually unobserved).
-For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we know the {\em domain mutex} $\forall x\ holding(x)\leftrightarrow\neg clear(x)$ we can safely extend the observation with literal {\tt\small $\neg$clear(blockA)} (despite this literal was actually unobserved).
+Here we describe a pre-processing mechanism to add new knowledge that completes states that are partially observed in a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task using a particular case of {\em state-invariants} that we call {\em schematic mutex}.
+We pay special attention to {\em schematic mutex} because they identify the {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effectively pruning of inconsistent \strips\ action models and (2) effective completion of partially observed states.
 
-Given a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task we build and solve a classical planning problem $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I,G_{\Lambda}}$ such that:
+We define a {\em schematic mutex} as a $\tup{p,q}$ predicates pair where both $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of state variables $F$ and such that they satisfy the following formulae $p\rightarrow \neg q$, where the predicate variables are universally quantified.
+For instance, predicates $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while predicates $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1\ clear(v_1)\leftrightarrow\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
+
+Given a {\em schematic mutex} $\tup{p,q}$ and a state observation $s_j^o\in \mathcal{O}$, {\small $(1\leq j\leq m)$}, such that $p(\omega)\in s_j^o$ is a literal built instantiating predicate $p$ with some subset of objects $\omega\subseteq\Omega^{|pars(p)|}$ then, the state observation $s_j^o$ can be safely completed adding the new literal $\neg q(\omega)$ (despite $\neg q(\omega)$ was actually unobserved).
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we know the {\em schematic mutex} $\forall x\ holding(x)\rightarrow\neg clear(x)$ we can safely extend that state observation with literal {\tt\small $\neg$clear(blockA)} (despite this literal was actually unobserved).
+
+\subsection{Pruning inconsistent action models with {\em schematic mutex}}
+Our approach to implement this pruning is adding new conditional effects to the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} that capture when the programmed model is inconsistent with a {\em schematic mutex} in $\Phi$.
+Given a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task we build and solve a classical planning problem $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I,G_{\Lambda}}$ that extends the original compilation as follows:
 \begin{itemize}
-\item $F_{\Lambda}$ extends $F$ with a fluent {\small$mode_{inval}$}, to indicate whether an action model is {\em inconsistent} with the input {\em state-invariants} $\Phi$, a fluent {\small$mode_{insert}$}, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
+\item $F_{\Lambda}$ extends $F$ with a fluent {\small$mode_{inval}$}, to indicate whether an action model is {\em inconsistent} with the input {\em state-invariants} $\Phi$.
+Like the original compilation we define a fluent {\small$mode_{insert}$}, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
 As explained, this is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$.
@@
 
-\item $A_{\Lambda}$ replaces the actions in $A$ with two types of actions.
-\begin{enumerate}
-\item Actions for {\em inserting} a {\em precondition}, {\em positive} effect or {\em negative} effect in $\xi$ following the syntactic constraints of \strips\ models. 
-\begin{itemize}
-\item Actions which support the addition of a {\em precondition} $p\in {\mathcal I}_{\Psi,\xi}$ to the action model $\xi$.
-A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
+\item $A_{\Lambda}$ contains actions of two kinds, like in the original compilation:
+\begin{enumerate}[leftmargin=*]
+\item Actions for {\em inserting} a {\em precondition}, {\em positive}/{\em negative} effect in $\xi$ following the syntactic constraints of \strips\ models. 
+\begin{itemize}[leftmargin=*]
+\item A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
+In addition, a conditional effect is added now for every {\em schematic mutex} $\tup{p,q}$ s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ to ban the insertion of two preconditions that are {\em schematic mutex}.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre\_p\_\xi, \neg eff\_p\_\xi, mode_{insert}\},\\
-\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre\_p\_\xi\}.
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre\_p\_\xi, \neg eff\_p\_\xi,\\
+&mode_{insert},\neg mode_{inval}\},\\
+\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre\_p\_\xi\},\\
+&\{pre\_q\_\xi\}\rhd\{mode_{inval}\}.
 \end{align*}
@@
 
-\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in {\mathcal I}_{\Psi,\xi}$ to the action model $\xi$.
+\item Actions which support the addition of a {\em negative}/{\em positive} effect $p\in {\mathcal I}_{\Psi,\xi}$ to the action model $\xi$ are extended with two conditional effects, for every {\em schematic mutex} $\tup{p,q}$ s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$, to ban the insertion of two positive/negative effects that are {\em schematic mutex}:
+\end{itemize}
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff\_p\_\xi, mode_{insert}\},\\
-\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff\_p\_\xi\}.
+\pre(\mathsf{insertEff_{p,\xi}})&=\{\neg eff\_p\_\xi, mode_{insert},\neg mode_{inval}\},\\
+\cond(\mathsf{insertEff_{p,\xi}})&=\{\emptyset\}\rhd\{eff\_p\_\xi\},\\
+&\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\},\\
+&\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\}.
 \end{align*}
 \end{small}
-\end{itemize}
 
-\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates, and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
+\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$.
 The action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
-These actions validate also that any state traversed by $P_{\Lambda}$ solutions is {\em consistent} with the {\em state-invariants} $\Phi$.
-The definition $\mathsf{apply_{\xi,\omega}}$ actions is also more compact in our compilation that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} since are not using disjunctions to code the possible preconditions of an action schema.
+The definition $\mathsf{apply_{\xi,\omega}}$ actions is more compact than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} since we are not using disjunctions to code the possible preconditions of an action schema.
 \end{enumerate}
@@
 &\{pre\_p\_\xi \wedge \neg p(\omega)\}\rhd\{mode_{inval}\}_{\forall p\in{\mathcal I}_{\Psi,\xi}},\\
-&\{\neg\phi\}\rhd\{mode_{inval}\}_{\forall \phi\in\Phi},\\
 &\{\emptyset\}\rhd\{\neg mode_{insert}\},
@@
 
-\subsection{Pruning inconsistent action models with {\em domain mutex}}
-We define a {\em domain mutex} as a $\tup{p,q}$ predicates pair where both $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of fluents $F$ of a given planning problem and such that they satisfy the following formulae $p\rightarrow \neg q$ where the predicate variables are universally quantified.
-For instance, predicates $holding(x)$ and $clear(x)$ from the {\em blocksworld} are {\em domain mutex} while predicates $clear(x)$ and $ontable(x)$ are not ($\forall x\ clear(x)\leftrightarrow\neg ontable(x)$ does not hold for every possible {\em blocksworld} state).
-We pay special attention to {\em domain mutex} because they identify the {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effectively pruning of inconsistent \strips\ action models and (2) effective completion of partially observed states.
-
-Our approach to implement this pruning is extending the conditional effects of the $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions (i.e., the actions that determine a solution model $\mathcal{M}$) with extra conditional effects indicating that the programmed model is {\em invalid} (i.e., inconsistent with a {\em domain mutex} in $\Phi$).
-Note that this {\em consistency} checking is more effective than the one implemented at the $\mathsf{apply_{\xi,\omega}}$ actions since $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertPre_{p,\xi}}$ actions appear at an earlier stage of the planning process.
-
-Formally, given a {\em domain mutex} $(p,q)$, s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$, we extend the actions for setting a precondition $p$ in a given action schema $\xi$ as follows:
+For every {\em state-invariant} $\phi\in\Phi$ we can extend $\mathsf{apply_{\xi,\omega}}$ actions with a conditional effect $\{\neg\phi\}\rhd\{mode_{inval}\}$ that checks the consistency of the {\em state-invariant} $\phi$ at every state traversed by a solution to the $P_\Lambda$ problem.
+Cheking arbitrary $\phi$ formulae can be too expensive for current classical planners.
+Instead, we prefer to add these conditional effects into $\mathsf{apply_{\xi,\omega}}$ actions for checking the consistency of a {\em schematic mutex} $\tup{p,q}$ s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi),\\
-&mode_{insert},\neg mode_{inval}\},\\
-\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\},\\
-&\{pre_{q}(\xi)\}\rhd\{mode_{inval}\}.
-\end{align*}
-\end{small}
-The same procedure is applied for action $insertPre_{q,\xi}$ to ban programming precondition $q$ iff $pre_{p}(\xi)$ precondition is already set.
-A similar procedure is also applied to $\mathsf{insertEff_{p,\xi}}$ and $\mathsf{insertEff_{q,\xi}}$ actions for banning in this case, two {\em negative effects} (or two {\em positive effects}) that are {\em domain mutex}.
-Now we show the actions that ban programming a positive (or negative) $p$ effect if its corresponding $q$ effect is already programmed:
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p}(\xi), mode_{insert},\neg mode_{inval}\},\\
-\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff_{p}(\xi),\\
-&\{pre_{q}(\xi),eff_{q}(\xi),pre_{p}(\xi)\}\rhd\{mode_{inval}\},\\
-&\{\neg pre_{q}(\xi),eff_{q}(\xi),\neg pre_{p}(\xi)\}\rhd\{mode_{inval}\}.
+&\{pre\_p\_\xi \wedge q(\omega)\}\rhd\{mode_{inval}\}_{\forall p\in{\mathcal I}_{\Psi,\xi}},\\
 \end{align*}
@@
 \subsection{The bias of the initially {\em empty} action model}
-Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $P=\tup{F,A[\cdot],I,G}$ problems preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
-In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P'=\tup{F',A',I,G}$ compilation so classical planners tend to solve $P'$ with plans that require a shorter number of {\em insert} actions.
+Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
+In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a shorter number of {\em insert} actions.
 
-This bias could be eliminated defining a cost function for the actions in $A'$ (e.g.
-{\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}).
-In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} with zero-cost actions.
-Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
-Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
-
-Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects.
-These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers, the $\mathsf{apply_{\xi,\omega}}$ actions).
-Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
-
+This bias could be eliminated defining a cost function for the actions in $P_\Lambda$ (e.g.
+{\em insert} actions have {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions have a {\em positive constant cost}).
+In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} when there are zero-cost actions.
+Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} that can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
+Further, the actions for inserting action effects are also applied in another single planning step.
+The plan horizon for programming any action model is then always bound to 2, which significantly reduces the planning horizon.
 The SAT-based planning approach is also convenient because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
@@
 \begin{mylemma}
-Soundness. Any classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ produces a \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi}$ learning task.
+Soundness. Any classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ produces a \strips\ model $\mathcal{M'}$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
 \end{mylemma}
@@
 According to the $P_{\Lambda}$ compilation, once a given precondition or effect is inserted into the action model $\mathcal{M}$ it cannot be removed back.
-In addition, once the action model $\mathcal{M}$ is applied it cannot be {\em reprogrammed}.
-In the compiled planning problem $P_{\Lambda}$, the value of the original fluents $F$ can exclusively be modified via $\mathsf{apply_{\xi,\omega}}$ actions.
-Therefore, the goals of the original $P$ classical planning task can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state $I=s_0$ reach a state $G \subseteq s_n$ validating that every generated intermediate state $s_i$, s.t.
-$0\leq i\leq n$, is consistent with the input {\em state-invariants}.
-This is exactly the definition of the solution condition for an action model $\mathcal{M}$ to solve the $\Lambda=\tup{P,\Phi,M}$ learning task.
+In addition, once the action model $\mathcal{M}$ is applied it cannot be {\em programmed}.
+In the compiled planning problem $P_{\Lambda}$, only $\mathsf{apply_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
+This means that a state consistent with an observation $s_n^o$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_i$, s.t.
+$0\leq i\leq n$, is consistent with the input state observations and {\em state-invariants}.
+This is exactly the definition of the solution condition for an action model $\mathcal{M}'$ to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
 \end{small}
@@
 \begin{mylemma}
-Completeness. Any \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi,M}$ learning task can be computed with a classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$.
+Completeness.
+Any \strips\ model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task can be computed with a classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$.
 \end{mylemma}
@@
 By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in a \strips\ action schema $\xi$ using predicates $\Psi$.
-In addition the $P_{\Lambda}$ compilation does not discard any possible action model $\mathcal{M}$ definable within ${\mathcal I}_{\Psi,\xi}$ while it can satisfy the domain mutex in $\Phi$.
-This means that for every \strips\ model $\mathcal{M}$ that solves the $\Lambda=\tup{P,\Phi,M}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action model $\mathcal{M}$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state $I$ into a state that satisfies the goals $G$.
+In addition the $P_{\Lambda}$ compilation does not discard any possible action model $\mathcal{M}'$ definable within ${\mathcal I}_{\Psi,\xi}$ that satisfies the domain mutex in $\Phi$.
+This means that, for every \strips\ model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action model $\mathcal{M}'$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
 \end{small}
@@
 
-\section{Learning from observations of plan executions}
-\label{sec:observations}
-Inductive approaches for the learning of planning action models compute an action model starting from an input set of observations of plan executions.
-This section provides a formal model for such input observations and shows how to leverage {\em state-invariants} to automatically {\em complete} those input observations.
-The section ends with the extension of our compilation to exploit the {\em completed} observations for the learning of \strips\ action models.
-
 \section{Evaluation}
 \label{sec:evaluation}
+This section evaluates the performance of our approach for learning \strips\ action models with different amounts of available input knowledge.
+
+\subsubsection{Reproducibility}
+The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+We only used 1 learning examples for each learning task and we fixed the examples for all the experiments so that we can evaluate the impact of the different amount and source of the input knowledge in the quality of the learned models.
+All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM.
+
+The classical planner we used to solve the instances that result from our compilations is the SAT-based planner{\sc Madagascar}~\cite{rintanen2014madagascar}.
+We used {\sc Madagascar} due to its ability to deal with planning instances populated with dead-ends~\cite{lopez2015deterministic}.
+
+For the sake of reproducibility, the compilation source code, evaluation scripts, used benchmarks and input {\em state-invariants} are fully available at the repository {\em https://github.com/anonsub/}.
 
@@
 \label{sec:related}
-{\em State-invariants} have been previously used to infer a HTN lanning model~\cite{lotinac2016constructing}.
+In {\em Inductive Logic Programming} it is common to make the hypothesis be consistent with the {\em background knowledge}, that is some form {\em deductive knowledge} apart from the examples~\cite{muggleton1994inductive}.
 
-In {\em Inductive Logic Programming} it is very common to make the hypothesis be consistent with some form deductive knowledge apart from the examples, what is usually called {\em background knowledge}~\cite{muggleton1994inductive}.
+{\em State-invariants} have also been previously used to improve the automatic construction of HTN planning model~\cite{lotinac2016constructing}.
 
@@
 because some parts of the action model are known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
-Our compilation approach is also flexible to this particular learning scenario.
+Our compilation is also flexible to this particular learning scenario.
 The known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to true in the initial state.
 Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, making the classical planning task $P_{\Lambda}$ easier to be solved.
-For example, suppose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state $I$ is extended with literals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the associated actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
+For example, suppose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state is extended with literals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the associated actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
 

### 2019-02-18 doc/ijcai-19-deductivelearning/deductivelearning.tex
added comments by Diego and Eva to the ijcai19 paper on deductive learning

@@
 
-\title{Learning action models from {\em state-invariants}}
+\title{One-shot learning: From domain knowledge to action models}
 
@@
 \begin{abstract}
-This paper addresses the learning of \strips\ action models from {\em state-invariants} (i.e logic formulae that specify constraints about the possible states of a given domain).
-The benefit of exploiting {\em state-invariants} is two-fold, they constrain the space of possible action models and they can complete learning examples that are only partially observed.
-Our approach for learning \strips\ action from {\em state-invariants} is a {\em classical planning} compilation that is flexible to different sources of input knowledge including partial observations of plan executions and {\em state-invariants}.
-The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (a single learning example that comprises just a full initial state and a partially observed state), {\em state-invariant} are helpful to learn good quality \strips\ action models.
+Most approaches to learning action planning models heavily rely on a significantly large volume of training samples or plan observations.
+In this paper, we adopt a different approach based on deductive learning from domain-specific knowledge, specifically from logic formulae that specify constraints about the possible states of a given domain.
+The minimal input observability required by our approach is a single example composed of a full initial state and a partial goal state.
+We will show that exploiting specific domain knowledge enable to constrain the space of possible action models as well as to complete partial observations, both of which turn out helpful to learn good-quality action models.
  \end{abstract}
@@
 \subsection{Learning action models with classical planning}
-The {\em classical planning compilation} for the learning \strips\ action models~\cite{aineto2018learning} receives as input an empty model $\mathcal{M}$ (which contains just the action headers), and an observation of a set of observations of plan executions.
-The compilation outputs a model $\mathcal{M'}$ that specifies the preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of the observations following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1}^o,a_i)$ for every observed action and $s_i^o=\theta(s_{i-1}^o,a_i)$ for every observed state.
+The {\em classical planning compilation} for the learning of \strips\ action models~\cite{aineto2018learning} receives as input an {\em empty} model (which contains just the {\em name} and {\em parameters} of each action schema), and a set of observations of plan executions.
+The compilation completes the {\em empty} model specifying the preconditions and effects of each action schema such that the validation of the completed model over the input observations is successful; i.e., there exists a plan computable with the completed model s.t.
+$\rho(s_{i-1}^o,a_i)$ and $s_i^o=\theta(s_{i-1}^o,a_i)$ holds for every observed state.
 
-A solution plan to the classical planning problem that results from the compialtion is a sequence of: (a) \emph{insert actions}, that insert preconditions and effects on a schemata of $\mathcal{M}$ to build $\mathcal{M'}$ and (b) \emph{apply actions} that validate the application of the $\mathcal{M'}$ model in the input observations.
+A solution plan to the classical planning problem that results from the compilation is then a sequence of:
+\begin{itemize}
+\item \emph{Insert actions}, that insert preconditions and effects on an action schema.
+\item \emph{Apply actions} that validate the application of the completed model in the input observations.
+\end{itemize}
 Figure~\ref{fig:plan-lplan} shows a solution to a classical planning problem resulting from the~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} compilation corresponding to the {\em blocksworld}~\cite{slaney2001blocks}.
-In the initial state of that problem the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table.
+In the initial state of that problem the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
 The problem goal is having the three-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
@@
 
-\subsection{State-invariants}
-The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
-If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
-For example, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
+\section{{\em One-shot} learning of planning action models from {\em domain-specific knowledge}}
+\label{sec:learning}
+We define the {\em one-shot} learning of planning action models from {\em domain-specific knowledge} as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, where:
+\begin{itemize}
+\item $\mathcal{M}$ is the {\em initial empty model} that contains only the name and parameters of each planning action to be learned.
+\item $\mathcal{O}$ is a single learning example that represents the observation of a sequence of states generated with the aimed planning action model. 
+\item $\Phi$ is a set of logic formulae defining {\em domain-specific knowledge} that constraint the set of possible states.
+\end{itemize}
 
-{\em State-invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state-invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
-For instance, Figure~\ref{fig:strongest-invariant} shows five clauses that define {\em state-invariants} for the {\em blocksworld} planning domain.
-
-\begin{figure}[hbt!]
-  \begin{footnotesize}
-$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
-$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
-$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
-\end{footnotesize}
- \caption{\small Example of {\em state-invariants} for the {\em blocksworld} domain.}
-\label{fig:strongest-invariant}
-\end{figure}
-
-A {\em mutex} (mutually exclusive) is a {\em state-invariant} that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
-
-A {\em domain invariant} is an instance-independent state-invariant, i.e.
-holds for any possible initial state and any possible set of objects.
-Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
-Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants~\cite{rintanen:schematicInvariants:AAAI2017}).
-
-\section{Learning \strips\ action models from {\em state-invariants}}
-\label{sec:learning}
-First, this section defines the sampling space and the space of possible action models. Then, the section formalizes the task of learning \strips\ action models from {\em state-invariants}.
-
-\subsection{The sampling space}
-We define a {\em learning example} as a sequence $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ of partial states, except for the initial state $s_0^o$ which is fully observed.
-Intermediate states can be missing meaning that are {\em unobserved}, so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-A partially observed state $s_i^o$, ${\small 1\leq i\leq m}$, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
-Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
+A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ is a model $\mathcal{M}'$ s.t.
+there exists a plan computable with $\mathcal{M}'$ that is consistent with the the {\em initial empty model} $\mathcal{M}$, the single learning example $\mathcal{O}$ and the given {\em domain-specific knowledge} in $\Phi$.
 
@@
 \item {\bf Observation constraints}.
-The {\em learning examples}, that are observations of plan executions, depict {\em semantic knowledge} that constraints further the space of possible action schemata.
+The {\em learning examples}, that in our case is the single observation of a sequence of states, depict {\em semantic knowledge} that constraints further the space of possible action schemata.
 \end{enumerate}
@@
 
-One can also introduce {\em domain-specific knowledge} to constrain further the space of possible schemata.
+\subsection{The sampling space}
+We define a {\em learning example} as a sequence $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ of {\em partially observed states}, except for the initial state $s_0^o$ which is a {\em full state}.
+The set of predicates $\Psi$ and the set of objects $\Omega$ that shape the fluents $F$ is then deducible from $\mathcal{O}$.
+A partially observed state $s_i^o$, ${\small 1\leq i\leq m}$, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ was not observed.
+Intermediate states can be {\em missing}, meaning that they are {\em unobserved}, so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$ (where ${\small k\geq 1}$ is unknown but finite).
+The minimal expression of a learning example must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed state $s_m^o$ so $m \geq 1$.
+
+To illustrate this Figure~\ref{fig:observation} shows a learning example that contains an initial state of the blocksworld where the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
+The second observation is a partially observed state in which {\tt blockA} is on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
+\begin{figure}[hbt!]
+  \begin{tiny}  
+  \begin{verbatim}
+(:predicates (on ?x ?y) (ontable ?x)
+	     (clear ?x) (handempty)
+	     (holding ?x))
+
+(:objects blockA blockB blockC)
+
+(:init (ontable blockA) (clear blockA)
+       (ontable blockB) (clear blockB)
+       (ontable blockC) (clear blockC)
+       (handempty))
+
+(:observations (on blockA blockB) (on blockB blockC))
+  \end{verbatim}           
+  \end{tiny}
+	\caption{\small Example of a two-state observationn for the learning \strips\ action models.}
+	\label{fig:observation}
+\end{figure}
+
+\subsection{Domain-specific knowledge}
+Our approach is to introduce {\em domain-specific knowledge} in the form of {\em state-constraints} to restrict further the space of possible schemata.
 For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-{\it State-invariants} are {\em domain-specific knowledge} that can be used to define new {\em syntactic} and {\em semantic} constraints over the space of possible action schemata.
+The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
+If we restrict ourselves to planning, {\em state-constraints} are abstractions for the compact specification of a set of states.
+For example, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em axiom} or {\em derived predicate} holds or the set of states that are considered goal states.
 
-\subsection{The learning task}
-We define the task of learning a planning action model from {\em state-invariants} as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, where:
-\begin{itemize}
-\item $\mathcal{M}$ is the {\em initial empty model} that contains only the name, $name(\xi)$, and parameters, $pars(\xi)$, of each action model $\xi$ to be learned.
-\item $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ is a single learning example.
-This example can be reduced to its minimal expression $\mathcal{O}^*=\tup{s_0^o, s_m^o}$ comprissing at least two state observations, a full initial state $s_0^o$ and a partially observed state $s_m^o$.
-The set of predicates $\Psi$ and objects $\Omega$ that shape the set of fluents $F$ is deducible from $\mathcal{O}$ since $s_0^o$ is a fully observed state.
-\item $\Phi$ is a set of {\em state-invariants} that define constraints about the set of possible states.
-\end{itemize}
+{\em State-invariants} is a kind of state-constraints useful for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state-invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
+For instance, Figure~\ref{fig:strongest-invariant} shows five clauses that define {\em state-invariants} for the {\em blocksworld} planning domain.
 
-A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ is a model $\mathcal{M}'$ that is consistent with the the {\em initial empty model} $\mathcal{M}$, the learning example $\mathcal{O}$ and the set of {\em state invariants} in $\Phi$.
+\begin{figure}[hbt!]
+  \begin{footnotesize}
+$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
+$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
+$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
+\end{footnotesize}
+ \caption{\small Example of {\em state-invariants} for the {\em blocksworld} domain.}
+\label{fig:strongest-invariant}
+\end{figure}
 
-\section{Learning action models from {\em state-invariants} with classical planning}
+A {\em mutex} (mutually exclusive) is a {\em state-invariant} that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
+
+A {\em domain invariant} is an instance-independent state-invariant, i.e.
+holds for any possible initial state and any possible set of objects.
+Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
+Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants~\cite{rintanen:schematicInvariants:AAAI2017}).
+
+In this work we exploit {\em domain-specific knowledge} that is given as  {\em schematic mutex}.
+We pay special attention to {\em schematic mutex} because they identify the {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effectively pruning of inconsistent \strips\ action models and (2) effective completion of partially observed states.
+We define a {\em schematic mutex} as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\Psi,\xi}$ represent predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, assuming that their corresponding variables are {\em universally} quantified.
+For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1\ \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
+
+\section{Learning \strips\ action models from {\em schematic mutex} with classical planning}
 \label{sec:compilation}
-This section shows how to exploit our compact encoding of \strips\ action models to solve a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task with an off-the-shelf classical planner.
+This section shows how to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task with an off-the-shelf classical planner.
 
@@
 what is not observed is considered unknown.
-Here we describe a pre-processing mechanism to add new knowledge that completes states that are partially observed in a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task using a particular case of {\em state-invariants} that we call {\em schematic mutex}.
-We pay special attention to {\em schematic mutex} because they identify the {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effectively pruning of inconsistent \strips\ action models and (2) effective completion of partially observed states.
+Here we describe a pre-processing mechanism to add new knowledge that completes states $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ that are partially observed in a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task using a given set $\Phi$ of {\em schematic mutex}.
 
-We define a {\em schematic mutex} as a $\tup{p,q}$ predicates pair where both $p\in\Psi$ and $q\in\Psi$ are predicates that shape the set of state variables $F$ and such that they satisfy the following formulae $p\rightarrow \neg q$, where the predicate variables are universally quantified.
-For instance, predicates $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while predicates $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1\ clear(v_1)\leftrightarrow\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
+Given a {\em schematic mutex} $\tup{p,q}$ and a state observation $s_j^o\in \mathcal{O}$, {\small $(1\leq j\leq m)$} then, the state observation $s_j^o$ can be safely completed adding the new literals $\neg q(\omega)$ that result from the unification of $p\implies \neg q$ with $s_j^o$ (assuming now that the corresponding variables of $p$ and $q$ are {\em existentially} quantified).
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we know the {\em schematic mutex} $\neg holding(v_1)\vee\neg clear(v_1)$, we can safely extend that state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal was actually unobserved).
 
-Given a {\em schematic mutex} $\tup{p,q}$ and a state observation $s_j^o\in \mathcal{O}$, {\small $(1\leq j\leq m)$}, such that $p(\omega)\in s_j^o$ is a literal built instantiating predicate $p$ with some subset of objects $\omega\subseteq\Omega^{|pars(p)|}$ then, the state observation $s_j^o$ can be safely completed adding the new literal $\neg q(\omega)$ (despite $\neg q(\omega)$ was actually unobserved).
-For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we know the {\em schematic mutex} $\forall x\ holding(x)\rightarrow\neg clear(x)$ we can safely extend that state observation with literal {\tt\small $\neg$clear(blockA)} (despite this literal was actually unobserved).
-
-\subsection{Pruning inconsistent action models with {\em schematic mutex}}
-Our approach to implement this pruning is adding new conditional effects to the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} that capture when the programmed model is inconsistent with a {\em schematic mutex} in $\Phi$.
-Given a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task we build and solve a classical planning problem $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I,G_{\Lambda}}$ that extends the original compilation as follows:
+\subsection{Pruning inconsistent action models with {\em domain-specific knowledge}}
+For every {\em state-constraint} $\phi\in\Phi$ we could extend $\mathsf{apply_{\xi,\omega}}$ actions with a conditional effect $\{\neg\phi\}\rhd\{mode_{inval}\}$ that checks the consistency of the {\em state-constraints} $\phi$ at every state traversed by a solution to the compiled problem.
+Cheking arbitrary $\phi$ formulae can however be too expensive for current classical planners.
+Instead, our approach is to check {\em state-constraints} in the form of {\em schematic mutex}.
+To implement this checking we add new conditional effects to {\em insert} and {\em apply} actions of the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
+These new conditional effects capture when the programmed model is inconsistent with a {\em schematic mutex} in $\Phi$.
 \begin{itemize}
-\item $F_{\Lambda}$ extends $F$ with a fluent {\small$mode_{inval}$}, to indicate whether an action model is {\em inconsistent} with the input {\em state-invariants} $\Phi$.
-Like the original compilation we define a fluent {\small$mode_{insert}$}, to indicate whether action models are being programmed, and the fluents for the propositional encoding of the corresponding space of STRIPS action models.
-As explained, this is a set of fluents of the type $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$.
-
-\item $G_{\Lambda}= G\cup \{\neg mode_{inval}\}$ extends the original goals $G$ with the $\neg mode_{inval}$ literal to validate that only states {\em consistent} with the state constraints $\Phi$ are traversed by $P_{\Lambda}$ solutions.
-
-\item $A_{\Lambda}$ contains actions of two kinds, like in the original compilation:
-\begin{enumerate}[leftmargin=*]
-\item Actions for {\em inserting} a {\em precondition}, {\em positive}/{\em negative} effect in $\xi$ following the syntactic constraints of \strips\ models. 
-\begin{itemize}[leftmargin=*]
-\item A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $eff_p$ exist in $\xi$.
-In addition, a conditional effect is added now for every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ to ban the insertion of two preconditions that are {\em schematic mutex}.
-
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre\_p\_\xi, \neg eff\_p\_\xi,\\
-&mode_{insert},\neg mode_{inval}\},\\
-\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre\_p\_\xi\},\\
-&\{pre\_q\_\xi\}\rhd\{mode_{inval}\}.
-\end{align*}
-\end{small}
-
-\item Actions which support the addition of a {\em negative}/{\em positive} effect $p\in {\mathcal I}_{\Psi,\xi}$ to the action model $\xi$ are extended with two conditional effects, for every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$, to ban the insertion of two positive/negative effects that are {\em schematic mutex}:
+\item For every {\em schematic mutex} $\tup{p,q}$ s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ a conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
+\item For every {\em schematic mutex} $\tup{p,q}$ s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions to ban the insertion of two positive/negative effects that are {\em schematic mutex}.
+\item The definition $\mathsf{apply_{\xi,\omega}}$ actions is more compact than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} since we are not using disjunctions to code the possible preconditions of an action schema.
 \end{itemize}
 
-\begin{small}
-\begin{align*}
-\pre(\mathsf{insertEff_{p,\xi}})&=\{\neg eff\_p\_\xi, mode_{insert},\neg mode_{inval}\},\\
-\cond(\mathsf{insertEff_{p,\xi}})&=\{\emptyset\}\rhd\{eff\_p\_\xi\},\\
-&\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\},\\
-&\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\}.
-\end{align*}
-\end{small}
+Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the classical planning compilation for the learning of \strips\ action models from {\em schematic mutex}.
+\begin{figure}
+\begin{tabular}{ll}
+{\bf Action} & {\bf New conditional effect}\\\hline
+$\mathsf{insertPre_{p,\xi}}$ & $\{pre\_q\_\xi\}\rhd\{mode_{inval}\},$\\
+$\mathsf{insertEff_{p,\xi}}$ & $\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\},$\\
+$\mathsf{insertEff_{p,\xi}}$ & $\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\},$\\
+$\mathsf{apply_{\xi,\omega}}$ & $\{pre\_p\_\xi \wedge \neg p(\omega)\}\rhd\{mode_{inval}\},$\\
+$\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\\
+& $q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\},$\\
+$\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
+&$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{mode_{inval}\},$
+\end{tabular}
+	\caption{\small Summary of the new conditional effects added to the classical planning compilation for the learning of \strips\ action models.}
+	\label{fig:ceffects}
+\end{figure}
 
-\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$.
-The action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
-The definition $\mathsf{apply_{\xi,\omega}}$ actions is more compact than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} since we are not using disjunctions to code the possible preconditions of an action schema.
-\end{enumerate}
-\end{itemize}
-
-\begin{small}
-\begin{align*}
-\pre(\mathsf{apply_{\xi,\omega}})=&\{\neg mode_{inval}\},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{pre\_p\_\xi\wedge eff\_p\_\xi\}\rhd\{\neg p(\omega)\}_{\forall p\in{\mathcal I}_{\Psi,\xi}},\\
-&\{\neg pre\_p\_\xi \wedge eff\_p\_\xi\}\rhd\{p(\omega)\}_{\forall p\in{\mathcal I}_{\Psi,\xi}}\},\\
-&\{pre\_p\_\xi \wedge \neg p(\omega)\}\rhd\{mode_{inval}\}_{\forall p\in{\mathcal I}_{\Psi,\xi}},\\
-&\{\emptyset\}\rhd\{\neg mode_{insert}\},
-\end{align*}
-\end{small}
-
-For every {\em state-invariant} $\phi\in\Phi$ we can extend $\mathsf{apply_{\xi,\omega}}$ actions with a conditional effect $\{\neg\phi\}\rhd\{mode_{inval}\}$ that checks the consistency of the {\em state-invariant} $\phi$ at every state traversed by a solution to the $P_\Lambda$ problem.
-Cheking arbitrary $\phi$ formulae can be too expensive for current classical planners.
-Instead, we prefer to add these conditional effects into $\mathsf{apply_{\xi,\omega}}$ actions for checking the consistency of a {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$.
-
-\begin{small}
-\begin{align*}
-&\{pre\_p\_\xi \wedge q(\omega)\}\rhd\{mode_{inval}\}_{\forall p\in{\mathcal I}_{\Psi,\xi}},\\
-\end{align*}
-\end{small}
+In addition, the goals of the classical planning problem output by the original compilation are extended with the $\neg mode_{inval}$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by the solution plans.
 

### 2019-02-19 experiments/aij18/comparison_FAMA.sh
Updated benchmarks and experiments of AIJ18

@@ -1,5 +1,5 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "satellite" "transport" "visitall" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "rovers" "satellite" "transport" "visitall" "zenotravel")
 
 ulimit -t 1000
 rm results*
@@ -16,9 +16,9 @@ touch sem-results.100.$j
 
 for d in "${arr[@]}"
 do
-   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 100 $j -t ten-observation -l 5 >> results.100.$j
+   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 100 $j -t ten-observation -l 0 5 >> results.100.$j
 
-   ../../src/compiler_new.py -v learned_domain.pddl ../../benchmarks/aij18/$d/ 100 $j -t ten-observation -l 5 >> sem-results.100.$j
+   ../../src/compiler_new.py -v learned_domain.pddl ../../benchmarks/aij18/$d/ 100 $j -t ten-observation -l 5 10 >> sem-results.100.$j
 
 done
 

### 2019-02-19 experiments/aij18/minimal_0_0.sh

@@ -1,11 +1,11 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "satellite" "transport" "visitall" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "rovers" "satellite" "transport" "visitall" "zenotravel")
 
 #declare -a arr=("transport" "visitall" "zenotravel")
 
 ulimit -t 1800
-rm results*
-rm sem-results*
+rm results.0.0
+rm sem-results.0.0
 
 LANG=en_US
 
@@ -15,8 +15,8 @@ touch sem-results.0.0
 
 for d in "${arr[@]}"
 do
-   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 0 0 -t ten-observation -l 2  >> results.0.0
+   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 0 0 -t ten-observation -l 0 2  >> results.0.0
 
-   ../../src/compiler_new.py -v learned_domain.pddl ../../benchmarks/aij18/$d/ 0 0 -t ten-observation -l 5 >> sem-results.0.0
+   ../../src/compiler_new.py -v learned_domain.pddl ../../benchmarks/aij18/$d/ 0 0 -t ten-observation -l 5 10 >> sem-results.0.0
 
 done

### 2019-02-19 experiments/aij18/minimal_100_10.sh

@@ -1,9 +1,9 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "satellite" "transport" "visitall" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "rovers" "satellite" "transport" "visitall" "zenotravel")
 
 ulimit -t 1000
-rm results*
-rm sem-results*
+rm results.100.10
+rm sem-results.100.10
 
 LANG=en_US
 
@@ -13,7 +13,7 @@ touch sem-results.100.10
 
 for d in "${arr[@]}"
 do
-   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 100 10 -t ten-observation -l 2 >> results.100.10
+   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 100 10 -t ten-observation -l 0 2 >> results.100.10
 
-   ../../src/compiler_new.py -v learned_domain.pddl ../../benchmarks/aij18/$d/ 100 10 -t ten-observation -l 5 >> sem-results.100.10 
+   ../../src/compiler_new.py -v learned_domain.pddl ../../benchmarks/aij18/$d/ 100 10 -t ten-observation -l 5 10 >> sem-results.100.10 
 done

### 2019-02-19 experiments/aij18/minimal_30_30.sh

@@ -1,8 +1,8 @@
 #!/bin/bash
-declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "satellite" "transport" "visitall" "zenotravel")
+declare -a arr=("blocks" "driverlog" "ferry" "floortile" "grid" "gripper" "hanoi" "miconic" "npuzzle" "parking" "rovers" "satellite" "transport" "visitall" "zenotravel")
 
 ulimit -t 1000
-rm results*
+rm results.30.30
 
 LANG=en_US
 
@@ -11,6 +11,6 @@ touch results.30.30
 
 for d in "${arr[@]}"
 do
-   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 30 30 -t ten-observation -l 2  >> results.30.30
+   ../../src/experimenter_new.py -s ../../benchmarks/aij18/$d/ 30 30 -t ten-observation -l 0 2  >> results.30.30
 
 done

### 2019-02-19 src/pddl_parser/parsing_functions.py
parsing_functions.py: improved trace parsing

@@ -84,7 +84,13 @@ def parse_condition_aux(alist, negated, type_dict, predicate_dict):
     """Parse a PDDL condition. The condition is translated into NNF on the fly."""
     tag = alist[0]
     if tag in ("and", "or", "not", "imply"):
-        args = alist[1:]
+        args = list()
+        for arg in alist[1:]:
+            if arg[0] == "=":
+                continue
+            if arg[0] == "not" and arg[1][0] == "=":
+                continue
+            args.append(arg)
         if tag == "imply":
             assert len(args) == 2
         if tag == "not":
@@ -105,8 +111,7 @@ def parse_condition_aux(alist, negated, type_dict, predicate_dict):
                 args[1], negated, type_dict, predicate_dict)]
         tag = "or"
     else:
-        parts = [parse_condition_aux(part, negated, type_dict, predicate_dict)
-                 for part in args]
+        parts = [parse_condition_aux(part, negated, type_dict, predicate_dict) for part in args]
 
     if tag == "and" and not negated or tag == "or" and negated:
         return pddl.Conjunction(parts)
@@ -508,7 +513,30 @@ def parse_task_pddl(task_pddl, type_dict, predicate_dict):
         assert False, entry
 
 
-def parse_trace_pddl(trace_pddl, predicates, action_observability=1, state_observability=1):
+def get_static_predicates(trajectory, predicates):
+
+    candidates = set([p.name for p in predicates])
+
+    trace_candidates = set()
+    for predicate in candidates:
+        static = True
+        init_literals = set([l for l in trajectory[0] if l.predicate == predicate])
+        for state in trajectory[1:]:
+            state_literals = set([l for l in state if l.predicate == predicate])
+
+            if init_literals != state_literals:
+                static = False
+                break
+
+        if static:
+            trace_candidates.add(predicate)
+
+    candidates = candidates.intersection(trace_candidates)
+
+    return candidates
+
+
+def parse_trace_pddl(trace_pddl, predicates, action_observability=1, state_observability=1, goal_observability=1, positive_goals=False, finite_steps=False):
     random.seed(123)
 
     iterator = iter(trace_pddl)
@@ -544,18 +572,53 @@ def parse_trace_pddl(trace_pddl, predicates, action_observability=1, state_obser
 
     for token in iterator:
         if token[0] == ':observations':
-            new_state = [literal for literal in parse_state(token[1:], all_literals) if random.random() <= state_observability]
-            states.append(new_state)
+            aux_state = parse_state(token[1:], all_literals)
[…]

### 2019-02-19 src/tim.py
tim.py: minor changes

@@ -725,7 +725,7 @@ def run_TIM(domain, problem):
     #                         newgen.append(new_s)
     #         if not p['attribute_space']:
     #             p['states'].extend(newgen)
-
+    #
     # print("=== Property sub-spaces ===")
     # for i in range(1, len(property_subspaces) + 1):
     #     print("PS {}:\n\t{}".format(i, property_space_to_string(property_subspaces[i - 1], inv_properties)))
@@ -738,7 +738,8 @@ def run_TIM(domain, problem):
     #         property_set = [inv_properties[i] for i in range(len(space['property_set'])) if
     #                         space['property_set'][i] == 1]
     #         for p in property_set:
-    #             construct_identity_invariant(p, predicate_arity_map, patterns)
+    #             num_invariants = construct_identity_invariant(p, predicate_arity_map, property_type_map, patterns,
+    #                                                           num_invariants)
     #         construct_state_membership_invariant(space, fd_task, patterns, inv_properties)
     #         construct_uniqueness_invariant(space, fd_task, patterns, inv_properties)
     #         # construct_binary_mutexes(space, predicate_arity_map, patterns, inv_properties)

### 2019-02-19 src/model_evaluator.py
model_evaluator.py: minor changes

@@ -71,9 +71,9 @@ def evaluate_matching(matchings, eva_actions, ref_actions):
                         eva_adds.add((action_evaluated, reform_literal(effect.literal, action_args, param_reform)))
                 break
 
-    pres_insertions = len(ref_pres) - len(ref_pres.intersection(eva_pres))
-    adds_insertions = len(ref_adds) - len(ref_adds.intersection(eva_adds))
-    dels_insertions = len(ref_dels) - len(ref_dels.intersection(eva_dels))
+    # pres_insertions = len(ref_pres) - len(ref_pres.intersection(eva_pres))
+    # adds_insertions = len(ref_adds) - len(ref_adds.intersection(eva_adds))
+    # dels_insertions = len(ref_dels) - len(ref_dels.intersection(eva_dels))
 
     pres_deletions = len(eva_pres) - len(ref_pres.intersection(eva_pres))
     adds_deletions = len(eva_adds) - len(ref_adds.intersection(eva_adds))
@@ -140,7 +140,7 @@ def evaluate(evaluation_domain_filename, reference_domain_filename, reformulatio
     if not reformulation:
         matches = list()
         for action_name, arity in actions_arity_list:
-            matches.append((action_name, action_name))
+            matches.append(tuple([action_name, action_name]+[i+1 for i in range(arity)]))
         matching_list = [matches]
     else:
         actions_by_arity = list()
@@ -173,24 +173,12 @@ def evaluate(evaluation_domain_filename, reference_domain_filename, reformulatio
                     if not valid_action_combination(comb, action_params_dict):
                         combinations.remove(comb)
             combinations_by_arity.append((arity, combinations))
-            # print(combinations)
 
 
         combinations_by_arity = [[[tuple([p[0][0]] + [e for e in p[1]]) for p in comb] for comb in combinations_by_arity[i][1]] for i in range(len(combinations_by_arity))]
-        # for e in combinations_by_arity:
-        #     print(e)
-
-        # combinations_by_arity2 = list()
-        # for ar, combs in combinations_by_arity:
-        #     params = [i for i in range(1,ar+1)]
-        #     aux = list()
-        #     for param_comb in itertools.permutations(params, ar):
-        #         for comb in combs:
-        #             print(tuple([comb[0][0]]+ [comb[0][1]] + [x for x in param_comb]))
 
         action_combinations = combinations_by_arity[0]
         for i in range(1, len(combinations_by_arity)):
[…]

### 2019-02-19 src/experimenter_new.py
experimenter_new.py: adapted to the new -l option

@@ -28,11 +28,13 @@ try:
 
     if "-l" in sys.argv:
         index = sys.argv.index("-l")
-        trace_limit = sys.argv[index+1]
-        sys.argv.remove("-l")
-        sys.argv.remove(sys.argv[index])
+        trace_min = int(sys.argv[index+1])
+        trace_max = int(sys.argv[index+2])
+        sys.argv.pop(index)
+        sys.argv.pop(index)
+        sys.argv.pop(index)
     else:
-        trace_limit = None
+        trace_min = None
 
     domain_folder_name  = sys.argv[1]
     action_observability = sys.argv[2]
@@ -50,8 +52,8 @@ except:
 # os.system(cmd)
 
 cmd = config.PROJECT_PATH + "src/compiler_new.py " + check_static_predicates + domain_folder_name + " " + action_observability + " " + state_observability + " -t " + trace_prefix
-if trace_limit:
-    cmd += " -l " + trace_limit
+if trace_min != None:
+    cmd += " -l {} {}".format(trace_min, trace_max)
 # print("\n\nExecuting... " + cmd)
 
 tic = time.time()

### 2019-02-19 src/FAMA_FSM.py
FAMA_FSM.py: extended with predicate filtering and counter examples

@@ -105,6 +105,14 @@ try:
     else:
         trace_prefix = "trace"
 
+    if "-nt" in sys.argv:
+        index = sys.argv.index("-nt")
+        nottrace_prefix = sys.argv[index+1]
+        sys.argv.remove("-nt")
+        sys.argv.remove(nottrace_prefix)
+    else:
+        nottrace_prefix = ""
+
     if "-l" in sys.argv:
         index = sys.argv.index("-l")
         trace_min = int(sys.argv[index+1])
@@ -128,15 +136,24 @@ except:
     sys.exit(-1)
 
 
-trace_filter = ["a", "b", "c", "d", "next", "head"]
-acceptor_state = "states4"
-# trace_filter = ["head"]
-pres_filter = []
-adds_filter = ['states0', 'states1', 'states2', 'states3', 'states4']
-# adds_filter = []
-dels_filter = ['states0', 'states1', 'states2', 'states3', 'states4']
+# trace_filter = ["a", "b", "c", "d", "next", "head"]
+# acceptor_state = "states4"
+# pres_filter = []-v ../benchmarks/icaps19/au-v ../benchmarks/icaps19/automata/class1/domain.pddl -v ../benchmarks/icaps19/automata/class1/domain.pddl tomata/class1/domain.pddl
+# adds_filter = ['states0', 'states1', 'states2', 'states3', 'states4']
+# dels_filter = ['states0', 'states1', 'states2', 'states3', 'states4']
 
+profile = open(domain_folder_name+'/profile', 'r')
+lines = profile.readlines()
 
+trace_filter = [s.strip().lower() for s in lines[0].strip().split(',')]
+acceptor_state = lines[1].strip()
+pres_filter = [s.strip().lower() for s in lines[2].strip().split(',')]
+adds_filter = [s.strip().lower() for s in lines[3].strip().split(',')]
+dels_filter = [s.strip().lower() for s in lines[4].strip().split(',')]
+
+profile.close()
+
+not_acceptor_states = ['states0', 'states1', 'states2', 'states3',]
 
 # Read the domain file
 if not validation_mode:
@@ -160,6 +177,25 @@ for trace in traces:
 
 if trace_min != None:
     traces = traces[trace_min:trace_max]
+
+
+# Negative examples
+if nottrace_prefix != "":
+    nottraces = list()
+    for filename in sorted(glob.glob(domain_folder_name + "/" + nottrace_prefix + "*")):
[…]

### 2019-02-19 doc/aij18/major-revision/evaluation.tex
Updated Dec 24 10.20

@@
 In this section we introduce the metrics used by \FAMA to evaluate the action models that result from solving a learning task $\Lambda$.
-\textcolor{red}{First, we will motivate the uso of two standard syntactic metrics ({\em precision} and {\em recall}).
+\textcolor{red}{First, we will motivate the use of two standard syntactic metrics ({\em precision} and {\em recall}).
 Then, section \ref{syntactic_precision_recall} will define these metrics for planning models, and section \ref{semantic_precision_recall} will introduce a semantic evaluation measure that builds upon {\em precision} and {\em recall}}.
@@
 The rationale behind the adaptation of precision and recall for planning models lies in counting the \emph{edit operations} that need to be applied in domain model $\mathcal{M}$ to transform it into the reference model $\mathcal{M'}$.
-Given a set of action models $\mathcal{M}$, the two allowed edit operations are:
+Given a domain model $\mathcal{M}$, the two allowed edit operations are:
 
@@
 \textcolor{red}{Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
+We interpret a sound learned model as one in which all preconditions and effects are correct with respect to the GTM, and so there is no need to remove anything.
+A complete model is one in which no precondition or effect is missing.
 %In more detail, precision is the fraction of relevant predicates among the predicates that make up the learned model, and recall is the fraction of relevant predicates in the learned model over all relevant predicates, understanding a relevant predicate as one that appears in the GTM.
@@
 A solution to this task is an \textbf{edited domain model} $\mathcal{M'}$ such that (1) $\mathcal{M'}$ is obtained by exclusively applying a finite sequence of \emph{deletion} and \emph{insertion} operations to $\mathcal{M}$ and (2) $\mathcal{M'}$ explains $\mathcal{T}$; i.e.
-$\mathcal{M'}$ is {\em compliant} with every plan trace $\tau\in\mathcal{T}$.
+$\mathcal{M'}$ is consistent with every plan trace $\tau\in\mathcal{T}$.
 It is always recommended for the test set to be different from the one used during learning and this is specially important for satisfying approaches such as \FAMA; otherwise $\mathcal{M'}$ = $\mathcal{M}$ since $\mathcal{M}$ would be able to explain $\mathcal{T}$ without any modification.
@@
   Given a domain model $\mathcal{M}$, and all the domain models $\mathcal{M'}$ able to explain the plan traces $\mathcal{T}$.
-The {\bf closest compliant domain model}, $\mathcal{M^*}$, is the comparable domain model closest to $\mathcal{M}$ (in terms of editions) that is able to explain $\mathcal{T}$;
+The {\bf closest consistent domain model}, $\mathcal{M^*}$, is the comparable domain model closest to $\mathcal{M}$ (in terms of editions) that is able to explain $\mathcal{T}$;
   \[\mathcal{M^*}=\underset{\forall \mathcal{M}' \rightarrow \mathcal{T}}{\arg\min} \ \left| INS(\mathcal{M},\mathcal{M'}) \cup DEL(\mathcal{M},\mathcal{M'}) \right|\]
@@
 
-The closest compliant domain model $\mathcal{M^*}$ allows us to define a semantic version of {\em precision} and {\em recall} following definitions \ref{syn-precision} and \ref{syn-recall}.
+The closest consistent domain model $\mathcal{M^*}$ allows us to define a semantic version of {\em precision} and {\em recall} following definitions \ref{syn-precision} and \ref{syn-recall}.
 
@@
 \begin{myproposition}
-When the closest compliant set of action models $\mathcal{M^*}$ of an evaluation task $\Lambda = \tup{\mathcal{M}, \mathcal{T}}$ is the GTM, the syntactic and semantic evaluation of $\mathcal{M}$ return the same values; that is, $Precision=sem\text{-}Precision$ and $Recall=sem\text{-}Recall$.
+When the closest consistent domain model $\mathcal{M^*}$ of an evaluation task $\Lambda = \tup{\mathcal{M}, \mathcal{T}}$ is the GTM, the syntactic and semantic evaluation of $\mathcal{M}$ return the same values; that is, $Precision=sem\text{-}Precision$ and $Recall=sem\text{-}Recall$.
 \end{myproposition}
 
-The intuition behind this evaluation is to {\em semantically} assess how well the learned domain model $\mathcal{M}$ explain a set of given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce the observations.
-Unlike the semantic metric defined by ARMS, our novel semantic definitions of precision and recall are not sensitive to flaws in the action model that manifest more than once in the plan traces since the flaws are corrected only once in the learned models instead of at every intermediate state of the plan traces.
+The intuition behind this evaluation is to {\em semantically} assess how well the learned domain model $\mathcal{M}$ explains a set of given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce the observations.
+\textcolor{red}{The interpretation of a sound and complete model in the semantic perspective is slightly different from the syntactic one.
+In this case, a sound model is one were there is no need to remove any precondition or effect in order to be consistent with some given plan traces.
+A complete model is one that is consistent with the given plan traces without adding any of its preconditions and effects.} Unlike the semantic metric defined by ARMS, our novel semantic definitions of precision and recall are not sensitive to flaws in the action model that manifest more than once in the plan traces since the flaws are corrected only once in the learned models instead of at every intermediate state of the plan traces.
 
@@
 
-Assuming we are using an optimal planner to solve $P_{\Lambda}'$, the solution plan of this problem will induce the \emph{closest compliant domain model} $\mathcal{M^*}$.
+Assuming we are using an optimal planner to solve $P_{\Lambda}'$, the solution plan of this problem will induce the \emph{closest consistent domain model} $\mathcal{M^*}$.
 Therefore, our compilation enables the straightforward computation of the semantic versions of \emph{precision} and \emph{recall}.
@@
 Considering this, $sem\text{-}Precision$ and $sem\text{-}Recall$ can be approximated if $P_{\Lambda}'$ is solved with a satisfying planner.
-In this case, no guarantees can be made that the edited model will be the closest compliant one, but a classical planner will always try to minimize the solution plan length and hence the number of edit operations applied to the input model.
+In this case, no guarantees can be made that the edited model will be the closest consistent one, but a classical planner will always try to minimize the solution plan length and hence the number of edit operations applied to the input model.

### 2019-02-19 doc/aij18/major-revision/experiments.tex

@@
   \item \textcolor[rgb]{1.00,0.00,0.00}{Other planners such as {\sc FastDownward} were also tested but provided worse experimental results}
+  \item \textcolor[rgb]{1.00,0.00,0.00}{The ability of {\sc Madagascar} to deal with instances populated with dead-ends, such as our compiled problem $P_\Lambda$, is very helpful~\cite{lopez2015deterministic}.}
   \item \textcolor[rgb]{1.00,0.00,0.00}{A SAT-based planner like {\sc Madagascar} is particularly suitable for tasks where the observed plan trace $\tau$ determines the horizon of the solution plan (case 1 presented in section \ref{task_definition})}.
 In this case, a SAT-based planner solves the prefix of the solution plan in two time steps because the actions for inserting preconditions can be applied in parallel in a single time step and the same for the actions inserting the effects.
-  \item \textcolor[rgb]{1.00,0.00,0.00}{When the length of the plan is unknown (case 2 presented in section \ref{task_definition}), the ability of {\sc Madagascar} to deal with instances populated with dead-ends is very helpful~\cite{lopez2015deterministic}.}
   \end{enumerate}
@@
 			Parking & 0.83 & 0.36 & 1.0 & 0.89 & 0.83 & 0.56 & 0.9 & 0.56& 160.64 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-car-to-car', u'move-car-to-car', 1, 2, 3)]
-			Rover & 0.43 & 0.73 & 0.24 & 0.47 & 0.56 & 0.38 & 0.39 & 0.61& 31.13 \\ % [(u'drop', u'drop', 1, 2), (u'navigate', u'navigate', 1, 3, 2), (u'sample-soil', u'sample-soil', 1, 2, 3), (u'sample-rock', u'sample-rock', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3, 4), (u'take-image', u'take-image', 1, 2, 3, 4, 5), (u'communicate-soil-data', u'communicate-rock-data', 1, 2, 4, 3, 5), (u'communicate-rock-data', u'communicate-rock-data', 1, 2, 3, 5, 4), (u'communicate-image-data', u'communicate-image-data', 1, 2, 3, 4, 6, 5)]
+			Rovers & 0.43 & 0.73 & 0.24 & 0.47 & 0.56 & 0.38 & 0.39 & 0.61& 31.13 \\ % [(u'drop', u'drop', 1, 2), (u'navigate', u'navigate', 1, 3, 2), (u'sample-soil', u'sample-soil', 1, 2, 3), (u'sample-rock', u'sample-rock', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3, 4), (u'take-image', u'take-image', 1, 2, 3, 4, 5), (u'communicate-soil-data', u'communicate-rock-data', 1, 2, 4, 3, 5), (u'communicate-rock-data', u'communicate-rock-data', 1, 2, 3, 5, 4), (u'communicate-image-data', u'communicate-image-data', 1, 2, 3, 4, 6, 5)]
 			Satellite & 0.6 & 0.21 & 0.56 & 1.0 & 0.5 & 0.5 & 0.56 & 0.43& 11.55 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
@@
 This broader space of solutions is also stressed in a diminished quality of the learned models.
-Thus, despite the learned models being compliant with the input data, they are further from the original GTM.
+Thus, despite the learned models being consistent with the input data, they are further from the original GTM.
 In Table~\ref{tab:results_minimum_0_0} we can observe the global values of precision and recall drop to 0.57 and 0.48, respectively.
@@
 Our last experiment is devoted to compare the scores provided by the syntactic and semantic versions of precision and recall.
-For that purpose, we will evaluate the models learned in Section \ref{minimal} both syntactically, using the GTM, and semantically, computing the set of action models closest to the learned models that is compliant with a testing set of five traces (see Definition \ref{compliant}).
-We must note that since we are using {\sc Madagascar}, a satisficing planner, the solution to the model evaluation may not be the closest compliant domain model, so the scores of sem-Precision and sem-Recall are approximate values.
+For that purpose, we will evaluate the models learned in Section \ref{minimal} both syntactically, using the GTM, and semantically, computing the set of action models closest to the learned models that is consistent with a testing set of five traces (see Definition \ref{compliant}).
+We must note that since we are using {\sc Madagascar}, a satisficing planner, the solution to the model evaluation may not be the closest consistent domain model, so the scores of sem-Precision and sem-Recall are approximate values.
 Our goal with this experiment is to gauge the suitability of the semantic metrics proposed in section \ref{sec:evaluation} with respect to their well-known counterparts.

### 2019-02-19 doc/aij18/major-revision/learning_task_def.tex

@@
 
-When the length of $\pi$ is given by $\tau$, the learning task is \textbf{SAT compilable}, and it is known that a boolean satisfiability problem is a NP-complete task~\cite{cook1971complexity}.
+    When the length of $\pi$ is given by $\tau$, the learning task is \textbf{SAT compilable}, and it is known that a Boolean satisfiability problem is a NP-complete task~\cite{cook1971complexity}.
 This is the reason why SAT solvers are commonly used in the approaches presented in section \ref{related_work}.
@@
 Likewise preconditions are derived lifting the minimal set of literals that appears in all the pre-states of the corresponding action.
+The challenge in this case comes from computing the least number of examples that are necessary to learn models within a given error rate~\cite{SternJ17}.
 

### 2019-02-19 doc/ijcai-19-conformant/conformant-learning.tex

@@
 \section{Batch learning of action models}
-Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a batch of state observations $\mathcal{O}_1,\ldots,\mathcal{O}_T$, this section shows how to implement a model learning algorithm that incrementally eliminates the action models that are not {\em consistent} with the input observations.
+Given a {\em classical planning frame} $\Phi=\tup{F,A[\cdot]}$ and a batch of state observations $\mathcal{O}_1,\ldots,\mathcal{O}_T$ within this frame, this section shows a model learning algorithm that incrementally eliminates action models that are not {\em consistent} with the input observations.
 The algorithm is defined as follows.
@@
 \item Initialize a {\em partially specified} model to the full space of possible action models, $\xi[\cdot]:=M$.
-\item For each observation $\mathcal{O}_t$, {\tt\small $1\leq t\leq T$}:
+\item For each input observation $\mathcal{O}_t$, {\tt\small $1\leq t\leq T$}:
 \begin{enumerate}
 \item Compute $M^*_t$, the {\em least-commitment} model that solves the corresponding $\tup{\Phi,\mathcal{O}_t}$ task.
-\item Update the {\em partially specified} model intersecting with the {\em least-commitment} model, $\xi[\cdot]:=\xi[\cdot]\cap M^*_t$.
-\item Continue until, the {\em partially specified} model is a singleton or no more observations left.
+\item Update the {\em partially specified} model intersecting it with the computed {\em least-commitment} model, $\xi[\cdot]:=\xi[\cdot]\cap M^*_t$.
+\item Continue until the {\em partially specified} model $\xi[\cdot]$ is a singleton, or no more observations left ($t=T$).
 \end{enumerate}
@@
 
+This model learning algorithm is {\em sound} given that, by definition, all the action models defined by $M^*_t$ are {\em consistent} with observation $\mathcal{O}_t$ and always holds that $\xi[\cdot]_t\subseteq \xi[\cdot]_{t-1}$, where $\xi[\cdot]_t$ is the value of the {\em partially specified} model after the $t^{th}$ iteration.
+The algorithm is also {\em complete} because it only discards from $\xi[\cdot]_0$ (the full space of possible action models) models that are not consistent with a given input observation.
+Further the complexity of our model learning algorithm depends only lineary on $T$, the numer of input observations, since it has never to re-process old observations.
+
 %Assuming that $\mathcal{O}_1,\ldots,\mathcal{O}_T$ are independent observations, and according to the PAC learning theory~\cite{kearns:PAC:1994}, we need to see at least $\frac{ln{(1/\delta)}+ln{|H|}}{\epsilon}$ observations to be sure that the prediction error is at most $\epsilon$ with probability $(1-\delta)$.
 In our case $|H|=\Pi_\xi2^{2\times|{\mathcal I}_{\Psi,\xi}|}$
+
+The challenge to implement this algorithm is then the computation of $M^*_t$, the {\em least-commitment} model that solves the corresponding $\tup{\Phi,\mathcal{O}_t}$ task.
 
@@
 The initial state $\Upsilon$ of the conformant planning task does not code an {\em initial} action model (typically {\em empty}) but the full space of action models.
-In more detail, the clauses in $\Upsilon$ comprises:
+In more detail, the clauses in $\Upsilon$ are:
       \begin{enumerate}
@@
       \item The clauses representing that the actual value of fluents {\tt\small pre\_e\_$\xi$}, {\tt\small eff\_e\_$\xi$} is unknown.
-In other words, that any model from the \strips\ space of models (following the previously mentioned {\em syntactic constraints}) can initially be part of the {\em least-commitment} action model.
-Formally, for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$, then $\Upsilon$ includes these two clauses:
+Any model from the \strips\ space of models (following the previously mentioned {\em syntactic constraints}) can initially be part of the {\em least-commitment} action model.
+Formally, for every $\xi$ and $e\in{\mathcal I}_{\Psi,\xi}$, $\Upsilon$ includes these two clauses:
             \begin{itemize}
-            \item {\tt\small pre\_e\_$\xi$} $\vee$ {\tt\small $\neg$pre\_e\_$\xi$}.
-            \item {\tt\small eff\_e\_$\xi$} $\vee$ {\tt\small $\neg$eff\_e\_$\xi$}.
+            \item {\tt\small pre\_e\_$\xi$} $\vee$ {\tt\small $\neg$pre\_e\_$\xi$}
+            \item {\tt\small eff\_e\_$\xi$} $\vee$ {\tt\small $\neg$eff\_e\_$\xi$}
             \end{itemize}
@@
       \end{enumerate}
-A conformant plan that solves this task induces a {\em partially specified model} that {\em explains} the input observation.
-To get the {\em least-commitment} action model, we have to compute the {\em optimal} plan for the resulting conformant planning task when: The actions that {\em program} a given precondition (or effect) of an action schema have an associated cost of 1.
-These actions represent now a reduction of one unit in the {\em uncertainty} of the initial {\em partially specified model}.
-In other words, these actions implement the different possible {\em immediate specification} of the current {\em partially specified model}.
-The remaining actions have a cost of 0.
+A conformant plan that solves this task only induces a {\em partially specified model} that {\em explains} the given input observation.
+To get a {\em partially specified model} that is {\em least-commitment}, we have to compute the {\em optimal} plan for this same conformant planning task using the following action costs: All actions have zero cost, except the actions that {\em program} a given precondition (or effect) of an action schema.
+These actions represent now a reduction of one unit in the {\em uncertainty} of the initial state and hence, have an associated cost of 1.
+In other words, each of these actions implement an {\em immediate specification} of the current {\em partially specified model}.
 
-\subsection{Computing the {\em least-commitment} model via classical planning}
-Inspired by the {\em classical planning compilation} $K_0$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose optimal solution induces the {\em least-commitment} action model for an observation $\mathcal{O}(\tau)$:
+\subsection{Computing a {\em superset} of the {\em least-commitment} action models via classical planning}
+Inspired by the {\em classical planning compilation} $K_0$ for conformant planning~\cite{palacios-conformant-JAIR09}, this section shows that we can build a {\em classical planning problem} $P=\tup{F',A',I',G'}$ whose optimal solution induces a {\em superset} of the {\em least-commitment} action model for an input observation $\mathcal{O}(\tau)$ (all actions in $A'$ have zero cost except {\em commit} actions that have a cost of one):
 \begin{itemize}
@@
 Similar actions are also defined for {\em committing} {\tt\small eff\_e\_$\xi$} to a positive/negative value but the value of {\tt\small eff\_e\_$\xi$} can only be committed once the value of the corresponding {\tt\small pre\_e\_$\xi$} is committed (i.e.
-once either $\mathsf{K\pre\_e\_\xi}$ or $\mathsf{K\neg\pre\_e\_\xi}$ holds in the current state).
+once either $\mathsf{K\pre\_e\_\xi}$ or $\mathsf{K\neg\pre\_e\_\xi}$ holds).
 \begin{small}
@@
 \hspace*{7pt}\pre(\mathsf{editable_{\xi}})=&\{\neg K\neg\pre\_e\_\xi\implies e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\\
-\cond(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi, \neg K\neg\eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
-&\{K\neg\pre\_e\_\xi, \neg K\neg\eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
+\cond(\mathsf{editable_{\xi}})=&\{K\pre\_e\_\xi, K\eff\_e\_\xi\}\rhd\{\neg e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
+&\{K\neg\pre\_e\_\xi, K\eff\_e\_\xi\}\rhd\{e\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}.
 \end{align*}
@@
  :effect (and
-   (when (and (Kpre_on_v1_v1_stack)(not(Knot_eff_on_v1_v1_stack))) (not (on ?o1 ?o1)))
-   (when (and (Kpre_on_v1_v2_stack)(not(Knot_eff_on_v1_v2_stack))) (not (on ?o1 ?o2)))
-   (when (and (Kpre_on_v2_v1_stack)(not(Knot_eff_on_v2_v1_stack))) (not (on ?o2 ?o1)))
-   (when (and (Kpre_on_v2_v2_stack)(not(Knot_eff_on_v2_v2_stack))) (not (on ?o2 ?o2)))
-   (when (and (Kpre_ontable_v1_stack)(not(Knot_eff_ontable_v1_stack))) (not (ontable ?o1)))
-   (when (and (Kpre_ontable_v2_stack)(not(Knot_eff_ontable_v2_stack))) (not (ontable ?o2)))
-   (when (and (Kpre_clear_v1_stack)(not(Knot_eff_clear_v1_stack))) (not (clear ?o1)))
-   (when (and (Kpre_clear_v2_stack)(not(Knot_eff_clear_v2_stack))) (not (clear ?o2)))
-   (when (and (Kpre_holding_v1_stack)(not(Knot_eff_holding_v1_stack))) (not (holding ?o1)))
-   (when (and (Kpre_holding_v2_stack)(not(Knot_eff_holding_v2_stack))) (not (holding ?o2)))
-   (when (and (Kpre_handempty_stack)(not(Knot_eff_handempty_stack))) (not (handempty)))
-   (when (and (Knot_pre_on_v1_v1_stack)(not(Knot_eff_on_v1_v1_stack))) (on ?o1 ?o1))
-   (when (and (Knot_pre_on_v1_v2_stack)(not(Knot_eff_on_v1_v2_stack))) (on ?o1 ?o2))
-   (when (and (Knot_pre_on_v2_v1_stack)(not(Knot_eff_on_v2_v1_stack))) (on ?o2 ?o1))
-   (when (and (Knot_pre_on_v2_v2_stack)(not(Knot_eff_on_v2_v2_stack))) (on ?o2 ?o2))
-   (when (and (Knot_pre_ontable_v1_stack)(not(Knot_eff_ontable_v1_stack))) (ontable ?o1))
-   (when (and (Knot_pre_ontable_v2_stack)(not(Knot_eff_ontable_v2_stack))) (ontable ?o2))
-   (when (and (Knot_pre_clear_v1_stack)(not(Knot_eff_clear_v1_stack))) (clear ?o1))
-   (when (and (Knot_pre_clear_v2_stack)(not(Knot_eff_clear_v2_stack))) (clear ?o2))
-   (when (and (Knot_pre_holding_v1_stack)(not(Knot_eff_holding_v1_stack))) (holding ?o1))
-   (when (and (Knot_pre_holding_v2_stack)(not(Knot_eff_holding_v2_stack))) (holding ?o2))
-   (when (and (Knot_pre_handempty_stack)(not(Knot_eff_handempty_stack))) (handempty))))
+   (when (and (Kpre_on_v1_v1_stack)(Keff_on_v1_v1_stack)) (not (on ?o1 ?o1)))
+   (when (and (Kpre_on_v1_v2_stack)(Keff_on_v1_v2_stack)) (not (on ?o1 ?o2)))
+   (when (and (Kpre_on_v2_v1_stack)(Keff_on_v2_v1_stack)) (not (on ?o2 ?o1)))
+   (when (and (Kpre_on_v2_v2_stack)(Keff_on_v2_v2_stack)) (not (on ?o2 ?o2)))
+   (when (and (Kpre_ontable_v1_stack)(Keff_ontable_v1_stack)) (not (ontable ?o1)))
+   (when (and (Kpre_ontable_v2_stack)(Keff_ontable_v2_stack)) (not (ontable ?o2)))
+   (when (and (Kpre_clear_v1_stack)(Keff_clear_v1_stack)) (not (clear ?o1)))
+   (when (and (Kpre_clear_v2_stack)(Keff_clear_v2_stack)) (not (clear ?o2)))
+   (when (and (Kpre_holding_v1_stack)(Keff_holding_v1_stack)) (not (holding ?o1)))
+   (when (and (Kpre_holding_v2_stack)(Keff_holding_v2_stack)) (not (holding ?o2)))
+   (when (and (Kpre_handempty_stack)(Keff_handempty_stack)) (not (handempty)))
+   (when (and (Knot_pre_on_v1_v1_stack)(Keff_on_v1_v1_stack)) (on ?o1 ?o1))
+   (when (and (Knot_pre_on_v1_v2_stack)(Keff_on_v1_v2_stack)) (on ?o1 ?o2))
+   (when (and (Knot_pre_on_v2_v1_stack)(Keff_on_v2_v1_stack)) (on ?o2 ?o1))
+   (when (and (Knot_pre_on_v2_v2_stack)(Keff_on_v2_v2_stack)) (on ?o2 ?o2))
+   (when (and (Knot_pre_ontable_v1_stack)(Keff_ontable_v1_stack)) (ontable ?o1))
+   (when (and (Knot_pre_ontable_v2_stack)(Keff_ontable_v2_stack)) (ontable ?o2))
+   (when (and (Knot_pre_clear_v1_stack)(Keff_clear_v1_stack)) (clear ?o1))
+   (when (and (Knot_pre_clear_v2_stack)(Keff_clear_v2_stack)) (clear ?o2))
+   (when (and (Knot_pre_holding_v1_stack)(Keff_holding_v1_stack)) (holding ?o1))
+   (when (and (Knot_pre_holding_v2_stack)(Keff_holding_v2_stack)) (holding ?o2))
+   (when (and (Knot_pre_handempty_stack)(Keff_handempty_stack)) (handempty))))
   \end{verbatim}           

### 2019-02-19 src/FAMA.py
FAMA.py: initial commit of FAMA with 2 vector compilation and support for invariants

@@ -1,0 +1,1181 @@
+#! /usr/bin/env python
+import glob, os, sys, copy, itertools
+import pddl, pddl_parser
+import config, fdtask_to_pddl
+import numpy as np
+
+
+def get_max_vars(actions):
+    max_vars = 0
+    for a in actions:
+        max_vars = max(max_vars, a.num_external_parameters)
+    return max_vars
+
+def get_max_steps(traces):
+    traces_steps = list()
+
+    for trace in traces:
+        not_empty_states = len([state for state in trace.states if state != []])
+        not_empty_actions = len([action for action in trace.actions if action != []])
+        traces_steps.append(max(not_empty_states, not_empty_actions))
+    return sum(traces_steps), max(traces_steps)
+
+def get_all_types(task, itype):
+   output=[itype]
+   for t in task.types:
+       if t.basetype_name == itype:
+           output.append(str(t.name))
+   return output
+
+def possible_pred_for_action(task, p, a, tup):
+    if (len(p.arguments) > len(a.parameters)):
+        return False
+
+    action_types = [set([a.parameters[int(tup[i])-1].type_name]) for i in range(len(tup))]
+    predicate_types = [set(get_all_types(task, x.type_name)) for x in p.arguments]
+
+    fits = [len(action_types[i].intersection(predicate_types[i])) >= 1 for i in range(len(action_types))]
+    return all(fits)
+
+def get_static_predicates(state_trajectories, predicates):
+
+    candidates = set([p.name for p in predicates])
+
+    for trajectory in state_trajectories:
+        trace_candidates = set()
+        for predicate in candidates:
+            static = True
+            init_literals = set([l for l in trajectory[0] if l.predicate == predicate])
+            for state in trajectory[1:]:
+                state_literals = set([l for l in state if l.predicate == predicate])
+
+                if init_literals != state_literals:
+                    static = False
+                    break
+
+            if static:
+                trace_candidates.add(predicate)
+
+        candidates = candidates.intersection(trace_candidates)
+
+    # reflexive_static_predicates = dict()
+    # for candidate in candidates:
+    #     reflexive_static_predicates[candidate] = True
+    #     for trace in traces:
+    #         init_literals = set([l for l in trace.init if l.predicate == candidate])
+    #         for literal in init_literals:
+    #             if len(literal.args) == 1 or len(set(literal.args)) != 1:
+    #                 reflexive_static_predicates[candidate] = False
+    #                 break
+
+    return candidates
[…]

### 2019-02-19 src/experimenter_FAMA.py
experimenter_FAMA.py: initial commit of the experimenter for FAMA.py

@@ -1,0 +1,140 @@
+#! /usr/bin/env python
+import glob, os, sys, copy, itertools, math, time
+import pddl, pddl_parser
+import fdtask_to_pddl
+import numpy as np
+import model_evaluator
+import config
+
+# **************************************#
+# MAIN
+# **************************************#
+
+try:
+    if "-s" in sys.argv:
+        index = sys.argv.index("-s")
+        check_static_predicates = "-s "
+        sys.argv.pop(index)
+    else:
+        check_static_predicates = ""
+
+
+    if "-i" in sys.argv:
+        index = sys.argv.index("-i")
+        use_invariants = "-i "
+        sys.argv.pop(index)
+    else:
+        use_invariants = ""
+
+
+    if "-g" in sys.argv:
+        index = sys.argv.index("-g")
+        positive_goal = "-g "
+        sys.argv.pop(index)
+    else:
+        positive_goal = ""
+
+
+    if "-f" in sys.argv:
+        index = sys.argv.index("-f")
+        finite_steps = "-f "
+        sys.argv.pop(index)
+    else:
+        finite_steps = ""
+
+
+    if "-t" in sys.argv:
+        index = sys.argv.index("-t")
+        trace_prefix = sys.argv[index+1]
+        sys.argv.pop(index)
+        sys.argv.pop(index)
+    else:
+        trace_prefix = "trace"
+
+    if "-l" in sys.argv:
+        index = sys.argv.index("-l")
+        trace_min = int(sys.argv[index+1])
+        trace_max = int(sys.argv[index+2])
+        sys.argv.pop(index)
+        sys.argv.pop(index)
+        sys.argv.pop(index)
+    else:
+        trace_min = None
+
+    if "-avg" in sys.argv:
+        index = sys.argv.index("-avg")
+        avg = int(sys.argv[index+1])
+        sys.argv.pop(index)
+        sys.argv.pop(index)
+    else:
+        avg = None
+
+    domain_folder_name  = sys.argv[1]
+    action_observability = sys.argv[2]
+    state_observability = sys.argv[3]
+    goal_observability = sys.argv[4]
+
+except:
+    print "Usage:"
+    print sys.argv[0] + "[-s] <domain folder> <action observability (0-100)> <state observability (0-100)> <goal observability (0-100)>"
+    sys.exit(-1)
+
+
+# outdir = "results/" + s + "-"+ d + "/"
+# cmd = "mkdir " + outdir
+# print("\n\nExecuting... " + cmd)
+# os.system(cmd)
[…]

### 2019-02-19 src/FAMA.py
FAMA.py: fixed mutex check in the effects

@@ -546,11 +546,14 @@ for a in actions:
                     if p.name == m[0] and m_tup == tup:
                         pass
                     else:
+
+                        m_vars = ["var" + str(t) for t in m_tup]
+
                         condition = pddl.conditions.Conjunction(
                             [pddl.conditions.NegatedAtom("pre_" + "_".join([p.name] + [a.name] + vars), []),
-                             pddl.conditions.Atom("pre_" + "_".join([p.name] + [a.name] + vars), []),
-                             pddl.conditions.NegatedAtom("pre_" + "_".join([m[0]] + [a.name] + vars), []),
-                             pddl.conditions.Atom("eff_" + "_".join([m[0]] + [a.name] + vars), []),
+                             pddl.conditions.Atom("eff_" + "_".join([p.name] + [a.name] + vars), []),
+                             pddl.conditions.NegatedAtom("pre_" + "_".join([m[0]] + [a.name] + m_vars), []),
+                             pddl.conditions.Atom("eff_" + "_".join([m[0]] + [a.name] + m_vars), []),
                              pddl.conditions.Atom(m[0], ["?o" + str(t) for t in m_tup])])
                         eff = eff + [
                             pddl.effects.Effect([], condition,
@@ -558,8 +561,8 @@ for a in actions:
 
                         condition = pddl.conditions.Conjunction(
                             [pddl.conditions.NegatedAtom("pre_" + "_".join([p.name] + [a.name] + vars), []),
-                             pddl.conditions.Atom("pre_" + "_".join([p.name] + [a.name] + vars), []),
-                             pddl.conditions.NegatedAtom("eff_" + "_".join([m[0]] + [a.name] + vars), []),
+                             pddl.conditions.Atom("eff_" + "_".join([p.name] + [a.name] + vars), []),
+                             pddl.conditions.NegatedAtom("eff_" + "_".join([m[0]] + [a.name] + m_vars), []),
                              pddl.conditions.Atom(m[0], ["?o" + str(t) for t in m_tup])])
                         eff = eff + [
                             pddl.effects.Effect([], condition,

### 2019-02-19 doc/ijcai-19-deductivelearning/deductivelearning.tex
added new conditional effects table to ijcai19 paper on deductive learning

@@
 These new conditional effects capture when the programmed model is inconsistent with a {\em schematic mutex} in $\Phi$.
-\begin{itemize}
-\item For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ a conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
-\item For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions to ban the insertion of two positive/negative effects that are {\em schematic mutex}.
-\item The definition $\mathsf{apply_{\xi,\omega}}$ actions is more compact than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} since we are not using disjunctions to code the possible preconditions of an action schema.
-\end{itemize}
 
 Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the classical planning compilation for the learning of \strips\ action models from {\em schematic mutex}.
+Next we describe each of them in detail:
+\begin{enumerate}
+\item[1-3.] For every {\em schematic mutex} $\tup{p,q}$ s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ a conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
+Likewise, two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions to ban the insertion of two positive/negative effects that are {\em schematic mutex}.
+\item[4-5.] For every {\em schematic mutex} $\tup{p,q}$ s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban a positive effect that is inconsistent with an input observation.
+\end{enumerate}
+
 \begin{figure}
-\begin{tabular}{ll}
-{\bf Action} & {\bf New conditional effect}\\\hline
-$\mathsf{insertPre_{p,\xi}}$ & $\{pre\_q\_\xi\}\rhd\{mode_{inval}\},$\\
-$\mathsf{insertEff_{p,\xi}}$ & $\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\},$\\
-$\mathsf{insertEff_{p,\xi}}$ & $\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\},$\\
-$\mathsf{apply_{\xi,\omega}}$ & $\{pre\_p\_\xi \wedge \neg p(\omega)\}\rhd\{mode_{inval}\},$\\
-$\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\\
-& $q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\},$\\
-$\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
-&$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{mode_{inval}\},$
+\begin{tabular}{lll}
+{\bf ID} & {\bf Action} & {\bf New conditional effect}\\\hline
+1.& $\mathsf{insertPre_{p,\xi}}$ & $\{pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
+2.& $\mathsf{insertEff_{p,\xi}}$ & $\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
+3.& $\mathsf{insertEff_{p,\xi}}$ & $\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
+5.& $\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
+& & $q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
+6.& $\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
+& &$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{mode_{inval}\}$
 \end{tabular}
@@
 
-In addition, the goals of the classical planning problem output by the original compilation are extended with the $\neg mode_{inval}$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by the solution plans.
+The goals of the classical planning problem output by the original compilation are extended with the $\neg mode_{inval}$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by the solution plans.
+This allows us to introduce a more compact definition $\mathsf{apply_{\xi,\omega}}$ actions than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} that do not require disjunctions to code the possible preconditions of an action schema.
 

### 2019-02-19 doc/ijcai-19-deductivelearning/deductivelearning.tex
added new conditional effects table to ijcai19 paper on deductive leaarning

@@
 \item[4-5.] For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban a positive effect that is inconsistent with an input observation.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation.
 \end{enumerate}
@@
 3.& $\mathsf{insertEff_{p,\xi}}$ & $\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
+4.& $\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
+& & $q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
 5.& $\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
-& & $q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
-6.& $\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
 & &$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{mode_{inval}\}$

### 2019-02-20 doc/ijcai-19-deductivelearning/deductivelearning.tex
Introduction updated

@@
 \usepackage{algorithmic}
+\usepackage{xspace}
 \urlstyle{same}
@@
 \newcommand{\strips}{\textsc{Strips}}
+
+\newcommand{\ARMS}{{\small {\sffamily ARMS}}\xspace}
+\newcommand{\CAMA}{{\small {\sffamily CAMA}}\xspace}
+\newcommand{\SLAF}{{\small {\sffamily SLAF}}\xspace}
+\newcommand{\LAMP}{{\small {\sffamily LAMP}}\xspace}
+\newcommand{\NOISTA}{{\small {\sffamily NOISTA}}\xspace}
+\newcommand{\LOCM}{{\small {\sffamily LOCM}}\xspace}
+\newcommand{\LOCMtwo}{{\small {\sffamily LOCM2}}\xspace}
+\newcommand{\LOP}{{\small {\sffamily LOP}}\xspace}
+\newcommand{\AMAN}{{\small {\sffamily AMAN}}\xspace}
+\newcommand{\LOUGA}{{\small {\sffamily LOUGA}}\xspace}
 
@@
 \label{sec:introduction}
-The specification of action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
+
+The learning of action models in planning has been typically addressed with inductive learning data-intensive approaches.
+From the pioneer learning system \ARMS~\cite{yang2007learning} to other more recent ones \cite{MouraoZPS12,zhuo2013action,kuvcera2018louga}, all of them require of the order of thousands of actions (input plan observations) to obtain and validate an action model.
+Generally speaking, the rationale behind these approaches is to obtain the statistically significant model $M$ that best explains the plan observations by minimizing some error and redundancy metrics.
+Specifically, given a plan observation $o=\langle s_0,a_1,s_1,\ldots,s_{n-1},a_n,s_n\rangle$, being $s_i, i>0$, a possibly partially observable state; if a solution plan $\pi$ to the planning task $\tup{s_0,s_n}$ can be produced with $M$ such that $\pi$ contains the observed actions $\tup{a_1,\ldots,a_n}$ of $o$, and the trace resulting from the execution of $\pi$ comprises the (partially) observed states $\tup{s_1,\ldots,s_{n-1}}$ of $o$, then $M$ explains $o$.
+Validating the explainability of a model as an optimization task over a test set of observations neither guarantees completeness (the model may not explain all the observations) nor correctness (a state that results from the execution of $\pi$ may contain contradictory information even though the state comprises the observed state).
+
+Differently, other approaches rely on a symbolic-via learning.
+The Simultaneous Learning and Filtering (\SLAF) approach~\cite{AmirC08} exploits logical inference and builds a complete explanation of observations through a CNF formula that represents the initial belief state, and an observation $o$ that contains partially observable states.
+The formula is updated with every action and state of $o$, thus representing all possible transition relations consistent with $o$.
+\SLAF extracts all satisfying models of the learned formula with a SAT solver although the algorithm cannot effectively learn the preconditions of actions.
+A more recent approach addresses the learning of action models from $o$ as a planning task that searches the space of the all possible action models~\cite{aineto2018learning}.
+A plan here is conceived as a series of steps that determine the preconditions and effects of the model actions plus other steps that validate the formed actions in $o$.
+The advantage of this approach is that only requires about a total of 50 actions in the set $o$.
+
+After this brief analysis, we raise the following question: can an action model be learnt from \emph{no observations} but using domain-specific knowledge?.
+The question is motivated by (a) the assumption that obtaining enough training observations is often difficult and costly, if not impossible in some domains~\cite{Zhuo15}; (b) we may not know the physics of the real-world domain being modeled but we may know certain pieces of knowledge about the domain; and (c) obtaining correct action models that are usable beyond their applicability to a set of testing observations.
+To this end, we opted for checking our hypothesis in the framework proposed in ~\cite{aineto2018learning} since this planning-based satisfiability approach allows us to configure additional constraints in the compilation scheme, it is able to work under a minimal set of observations and uses an off-the-shelf planner\footnote{We thank authors for providing us with the source files of their learning system.}.
+Ultimately, we aim to compare the informational power of domain observations against the representational power of domain-specific knowledge.
+
+The paper is organized as follows ...
+
+%The specification of action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
 The {\em machine learning} of action models relieves this {\em knowledge acquisition bottleneck} of {\em model-based planning} and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}.
@@
 
-This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge.
+%This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge.
 Our approach leverages {\em state-invariants} (i.e.
@@
 
-Our approach for learning \strips\ action models from {\em state-invariants} is to compile the learning task into a classical planning task.
+%Our approach for learning \strips\ action models from {\em state-invariants} is to compile the learning task into a classical planning task.
 Our compilation is built on top of the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} and it is flexible to different kinds of input knowledge including both partial observations of plan executions and {\em state-invariants}.

### 2019-02-20 doc/ijcai-19-deductivelearning/deductivelearning.tex
added pointer to roni's paper to ijcai19 paper on deductive learning

@@
 
-\subsection{The bias of the initially {\em empty} action model}
-Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
-In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a shorter number of {\em insert} actions.
-
-This bias could be eliminated defining a cost function for the actions in $P_\Lambda$ (e.g.
-{\em insert} actions have {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions have a {\em positive constant cost}).
-In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} when there are zero-cost actions.
-Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} that can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
-Further, the actions for inserting action effects are also applied in another single planning step.
-The plan horizon for programming any action model is then always bound to 2, which significantly reduces the planning horizon.
-The SAT-based planning approach is also convenient because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
-
 \subsection{Compilation properties}
@@
 
+Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
+In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a shorter number of {\em insert} actions.
+
+This bias could be eliminated defining a cost function for the actions in $P_\Lambda$ (e.g.
+{\em insert} actions have {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions have a {\em positive constant cost}).
+In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} when there are zero-cost actions.
+Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} that can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
+Further, the actions for inserting action effects are also applied in another single planning step.
+The plan horizon for programming any action model is then always bound to 2, which significantly reduces the planning horizon.
+The SAT-based planning approach is also convenient because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
+
 \section{Evaluation}
@@
 
+Our learning setting is related to the classical planning formulation where no action model is given~\cite{SternJ17}.
+This planning setting can can be seen as an scenario when the action model is {\em learned} from a single example that contains only two state observations: the initial state and the goals.
+
 \section{Conclusions}

### 2019-02-20 doc/ijcai-19-deductivelearning/deductivelearning.tex
added pointer to roni's paper to ijcai19 paper on deductive learning

@@
 
-(:observations (on blockA blockB) (on blockB blockC))
+(:observation (on blockA blockB) (on blockB blockC))
   \end{verbatim}
@@
 
-\subsection{Domain-specific knowledge}
+\subsection{The domain-specific knowledge}
 Our approach is to introduce {\em domain-specific knowledge} in the form of {\em state-constraints} to restrict further the space of possible schemata.
@@
 
-\subsection{Pruning inconsistent action models with {\em domain-specific knowledge}}
+\subsection{Pruning inconsistent action models with {\em schematic mutex}}
 For every {\em state-constraint} $\phi\in\Phi$ we could extend $\mathsf{apply_{\xi,\omega}}$ actions with a conditional effect $\{\neg\phi\}\rhd\{mode_{inval}\}$ that checks the consistency of the {\em state-constraints} $\phi$ at every state traversed by a solution to the compiled problem.
@@
 \begin{enumerate}
-\item[1-3.] For every {\em schematic mutex} $\tup{p,q}$ s.t.
+\item[1-3] For every {\em schematic mutex} $\tup{p,q}$ s.t.
 both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ a conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
 Likewise, two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions to ban the insertion of two positive/negative effects that are {\em schematic mutex}.
-\item[4-5.] For every {\em schematic mutex} $\tup{p,q}$ s.t.
+\item[4-5] For every {\em schematic mutex} $\tup{p,q}$ s.t.
 both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation.
@@
 \begin{figure}
+\begin{footnotesize}
 \begin{tabular}{lll}
 {\bf ID} & {\bf Action} & {\bf New conditional effect}\\\hline
-1.& $\mathsf{insertPre_{p,\xi}}$ & $\{pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
-2.& $\mathsf{insertEff_{p,\xi}}$ & $\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
-3.& $\mathsf{insertEff_{p,\xi}}$ & $\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
-4.& $\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
-& & $q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
-5.& $\mathsf{apply_{\xi,\omega}}$ & $\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
-& &$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{mode_{inval}\}$
+1&$\mathsf{insertPre_{p,\xi}}$&$\{pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
+2&$\mathsf{insertEff_{p,\xi}}$&$\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
+3&$\mathsf{insertEff_{p,\xi}}$&$\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
+4&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
+&&$q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
+5&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
+&&$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{mode_{inval}\}$
 \end{tabular}
+\end{footnotesize}
 	\caption{\small Summary of the new conditional effects added to the classical planning compilation for the learning of \strips\ action models.}
@@
 The goals of the classical planning problem output by the original compilation are extended with the $\neg mode_{inval}$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by the solution plans.
-This allows us to introduce a more compact definition $\mathsf{apply_{\xi,\omega}}$ actions than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} that do not require disjunctions to code the possible preconditions of an action schema.
+The $\neg mode_{inval}$ literal allows us also to define a more compact definition $\mathsf{apply_{\xi,\omega}}$ actions than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} that do not require a disjunction to code the possible preconditions of an action schema.
 

### 2019-02-21 doc/ijcai-19-deductivelearning/deductivelearning.tex
added more details to section 4 of the ijcai19 paper on deductive learning

@@
 Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
-Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants~\cite{rintanen:schematicInvariants:AAAI2017}).
+Domain invariants are often compactly defined as {\em lifted invariants} (also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}).
 
 In this work we exploit {\em domain-specific knowledge} that is given as  {\em schematic mutex}.
-We pay special attention to {\em schematic mutex} because they identify the {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effectively pruning of inconsistent \strips\ action models and (2) effective completion of partially observed states.
-We define a {\em schematic mutex} as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\Psi,\xi}$ represent predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, assuming that their corresponding variables are {\em universally} quantified.
-For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1\ \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
+We pay special attention to {\em schematic mutex} because they identify mutually exclusive {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effective completion of partially observed states and (2) effectively pruning of inconsistent \strips\ action models.
+We define a {\em schematic mutex} as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\Psi,\xi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are {\em universally quantified}.
+For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
 
@@
 what is not observed is considered unknown.
-Here we describe a pre-processing mechanism to add new knowledge that completes states $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ that are partially observed in a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task using a given set $\Phi$ of {\em schematic mutex}.
+Here we describe a pre-processing mechanism to add new knowledge that completes the states $\tup{s_1^o \ldots, s_m^o}$ that are partially observed in a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task using an input set of {\em schematic mutex} $\Phi$.
 
-Given a {\em schematic mutex} $\tup{p,q}$ and a state observation $s_j^o\in \mathcal{O}$, {\small $(1\leq j\leq m)$} then, the state observation $s_j^o$ can be safely completed adding the new literals $\neg q(\omega)$ that result from the unification of $p\implies \neg q$ with $s_j^o$ (assuming now that the corresponding variables of $p$ and $q$ are {\em existentially} quantified).
-For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and we know the {\em schematic mutex} $\neg holding(v_1)\vee\neg clear(v_1)$, we can safely extend that state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal was actually unobserved).
-
-\subsection{Pruning inconsistent action models with {\em schematic mutex}}
-For every {\em state-constraint} $\phi\in\Phi$ we could extend $\mathsf{apply_{\xi,\omega}}$ actions with a conditional effect $\{\neg\phi\}\rhd\{mode_{inval}\}$ that checks the consistency of the {\em state-constraints} $\phi$ at every state traversed by a solution to the compiled problem.
-Cheking arbitrary $\phi$ formulae can however be too expensive for current classical planners.
-Instead, our approach is to check {\em state-constraints} in the form of {\em schematic mutex}.
-To implement this checking we add new conditional effects to {\em insert} and {\em apply} actions of the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
-These new conditional effects capture when the programmed model is inconsistent with a {\em schematic mutex} in $\Phi$.
-
-Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the classical planning compilation for the learning of \strips\ action models from {\em schematic mutex}.
-Next we describe each of them in detail:
-\begin{enumerate}
-\item[1-3] For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ a conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
-Likewise, two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions to ban the insertion of two positive/negative effects that are {\em schematic mutex}.
-\item[4-5] For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation.
-\end{enumerate}
+Let us define the production rule $p\rightarrow\neg q$ such that $\tup{p,q}$ is a {\em schematic mutex}.
+Given a {\em schematic mutex} $\tup{p,q}\in \Phi$ and a state observation $s_j^o\in \mathcal{O}$, {\small $(1\leq j\leq m)$} then $s_j^o$ can be safely completed adding the new literals $\neg q(\omega)$ that result from the unification of the corresponding production rule with $s_j^o$.
+$\omega\subseteq\Omega^{pars(q)}$ represents the objects that unify the variables in $q$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and $\Phi$ contains the {\em schematic mutex} $\neg holding(v_1)\vee\neg clear(v_1)$, we can safely extend that state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal was actually unobserved).
 
@@
 {\bf ID} & {\bf Action} & {\bf New conditional effect}\\\hline
-1&$\mathsf{insertPre_{p,\xi}}$&$\{pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
-2&$\mathsf{insertEff_{p,\xi}}$&$\{pre\_q\_\xi,eff\_q\_\xi,pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
-3&$\mathsf{insertEff_{p,\xi}}$&$\{\neg pre\_q\_\xi,eff\_q\_\xi,\neg pre\_p\_\xi\}\rhd\{mode_{inval}\}$\\
+1&$\mathsf{insertPre_{p,\xi}}$&$\{pre\_q\_\xi\}\rhd\{invalid\}$\\
+2&$\mathsf{insertEff_{p,\xi}}$&$\{pre\_q\_\xi\wedge eff\_q\_\xi\wedge pre\_p\_\xi\}\rhd\{invalid\}$\\
+3&$\mathsf{insertEff_{p,\xi}}$&$\{\neg pre\_q\_\xi\wedge eff\_q\_\xi\wedge \neg pre\_p\_\xi\}\rhd\{invalid\}$\\
 4&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
-&&$q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{mode_{inval}\}$\\
+&&$q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{invalid\}$\\
 5&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
-&&$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{mode_{inval}\}$
+&&$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{invalid\}$
 \end{tabular}
@@
 
-The goals of the classical planning problem output by the original compilation are extended with the $\neg mode_{inval}$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by the solution plans.
-The $\neg mode_{inval}$ literal allows us also to define a more compact definition $\mathsf{apply_{\xi,\omega}}$ actions than the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} that do not require a disjunction to code the possible preconditions of an action schema.
+\subsection{Pruning inconsistent action models with {\em schematic mutex}}
+We could extend the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} to check the consistency of the {\em state-constraints} in $\Phi$ at every state traversed by a solution to the compiled problem.
+Unfortunately, checking arbitrary $\phi$ formulae is too expensive for current classical planners.
+
+Instead, our approach is to define a mechanism to check {\em state-constraints} in the form of {\em schematic mutex}.
+To implement this checking mechanism we add new conditional effects to the {\em insert} and {\em apply} actions of the classical planning compilation.
+Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the compilation and next, we describe them in detail:
+\begin{enumerate}
+\item[1-3] For every {\em schematic mutex} $\tup{p,q}$ s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ one conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
+Likewise, two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions, one to ban the insertion of two positive effects that are {\em schematic mutex} and another one to ban two mutex negative effects.
+\item[4-5] For every {\em schematic mutex} $\tup{p,q}$ s.t.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation (in $\mathsf{apply_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
+\end{enumerate}
+In theory, conditional effects of the kind $4$ and $5$ are enough to guarantee that all the states traversed by a plan produced by the compilation are {\em consistent} with the input set of {\em schematic mutex} $\Phi$ (of course, provided that the input initial state $s_0^o$ is a valid state).
+In practice we include also conditional effects of the kind $1$, $2$ and $3$ because they prune {\em invalid} action models at an earlier stage of the planning process (these effects extend the {\em insert} actions that always appear first in the solution plans).
+
+The goals of the classical planning problem output by the original compilation are extended with the $\neg invalid$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by solution plans.
+Remarkably, the $\neg invalid$ literal allows us also to define $\mathsf{apply_{\xi,\omega}}$ actions more compactly than in the original compilation by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+Disjunctions are no longer required to code the possible preconditions of an action schema since they can now be encoded with conditional effects of this kind $\{pre\_p\_\xi\wedge \neg p(\omega)\}\rhd\{invalid\}$.
 

### 2019-02-22 doc/ijcai-19-deductivelearning/deductivelearning.tex
Intro completed

@@
 The learning of action models in planning has been typically addressed with inductive learning data-intensive approaches.
-From the pioneer learning system \ARMS~\cite{yang2007learning} to other more recent ones \cite{MouraoZPS12,zhuo2013action,kuvcera2018louga}, all of them require of the order of thousands of actions (input plan observations) to obtain and validate an action model.
-Generally speaking, the rationale behind these approaches is to obtain the statistically significant model $M$ that best explains the plan observations by minimizing some error and redundancy metrics.
-Specifically, given a plan observation $o=\langle s_0,a_1,s_1,\ldots,s_{n-1},a_n,s_n\rangle$, being $s_i, i>0$, a possibly partially observable state; if a solution plan $\pi$ to the planning task $\tup{s_0,s_n}$ can be produced with $M$ such that $\pi$ contains the observed actions $\tup{a_1,\ldots,a_n}$ of $o$, and the trace resulting from the execution of $\pi$ comprises the (partially) observed states $\tup{s_1,\ldots,s_{n-1}}$ of $o$, then $M$ explains $o$.
-Validating the explainability of a model as an optimization task over a test set of observations neither guarantees completeness (the model may not explain all the observations) nor correctness (a state that results from the execution of $\pi$ may contain contradictory information even though the state comprises the observed state).
+From the pioneer learning system \ARMS~\cite{yang2007learning} to more recent ones \cite{MouraoZPS12,zhuo2013action,kuvcera2018louga}, all of them require thousands of plan observations or training samples, i.e., sequences of actions as evidence of the execution of an observed agent, to obtain and validate an action model.
+These approaches return the statistically significant model that best explains the plan observations by minimizing some error metric.
+A model explains an observation if a plan containing the observed actions can be generated with the model and the states induced by this plan also include the possibly partially observed states.
+The limitation of posing model validation as an optimization task over a testing set of observations is that it neither guarantees completeness (the model may not explain all the observations) nor correctness (the states induced by the execution of the plan generated with the model may contain contradictory information).
 
-Differently, other approaches rely on a symbolic-via learning.
-The Simultaneous Learning and Filtering (\SLAF) approach~\cite{AmirC08} exploits logical inference and builds a complete explanation of observations through a CNF formula that represents the initial belief state, and an observation $o$ that contains partially observable states.
-The formula is updated with every action and state of $o$, thus representing all possible transition relations consistent with $o$.
+Differently, other approaches rely on symbolic-via learning.
+The Simultaneous Learning and Filtering (\SLAF) approach~\cite{AmirC08} exploits logical inference and builds a complete explanation through a CNF formula that represents the initial belief state, and a plan observation.
+The formula is updated with every action and state of the observation, thus representing all possible transition relations consistent with it.
 \SLAF extracts all satisfying models of the learned formula with a SAT solver although the algorithm cannot effectively learn the preconditions of actions.
-A more recent approach addresses the learning of action models from $o$ as a planning task that searches the space of the all possible action models~\cite{aineto2018learning}.
-A plan here is conceived as a series of steps that determine the preconditions and effects of the model actions plus other steps that validate the formed actions in $o$.
-The advantage of this approach is that only requires about a total of 50 actions in the set $o$.
+A more recent approach addresses the learning of action models from plan observations as a planning task which searches the space of all possible action models~\cite{aineto2018learning}.
+A plan here is conceived as a series of steps that determine the preconditions and effects of the action models plus other steps that validate the formed actions in the observations.
+The advantage of this approach is that it only requires input samples of about a total of 50 actions.
 
-After this brief analysis, we raise the following question: can an action model be learnt from \emph{no observations} but using domain-specific knowledge?.
-The question is motivated by (a) the assumption that obtaining enough training observations is often difficult and costly, if not impossible in some domains~\cite{Zhuo15}; (b) we may not know the physics of the real-world domain being modeled but we may know certain pieces of knowledge about the domain; and (c) obtaining correct action models that are usable beyond their applicability to a set of testing observations.
-To this end, we opted for checking our hypothesis in the framework proposed in ~\cite{aineto2018learning} since this planning-based satisfiability approach allows us to configure additional constraints in the compilation scheme, it is able to work under a minimal set of observations and uses an off-the-shelf planner\footnote{We thank authors for providing us with the source files of their learning system.}.
-Ultimately, we aim to compare the informational power of domain observations against the representational power of domain-specific knowledge.
+This paper studies the impact of using mixed input data, i.e, automatically-collected plan observations and human-encoded domain-specific knowledge, in the learning of action models.
+Particularly, we aim to stress the extreme case of having a single observation sample and answer the question to whether the lack of training samples can be overcome with the supply of domain knowledge.
+The question is motivated by (a) the assumption that obtaining enough training observations is often difficult and costly, if not impossible in some domains~\cite{Zhuo15}; (b) the fact that although the physics of the real-world domain being modeled are unknown, the user may know certain pieces of knowledge about the domain; and (c) the desire for correct action models that are usable beyond their applicability to a set of testing observations.
+To this end, we opted for checking our hypothesis in the framework proposed in~\cite{aineto2018learning} since this planning-based satisfiability approach allows us to configure additional constraints in the compilation scheme, it is able to work under a minimal set of observations and uses an off-the-shelf planner\footnote{We thank authors for providing us with the source files of their learning system.}.
+Ultimately, we aim to compare the informational power of domain observations (information quantity) with the representational power of domain-specific knowledge (information quality).
+Complementarily, we restrict our attention to solely observations over fluents as in many applications the actual actions of an agent may not be observable~\cite{SohrabiRU16}.
 
-The paper is organized as follows ...
+Next section summarizes basic planning concepts and outlines the baseline learning approach~\cite{aineto2018learning}.
+Then we formalize our one-shot learning task with domain knowledge and subsequently we explain the task-solving process.
+Section 5 presents the experimental evaluation and last section concludes.
 

### 2019-02-22 doc/ijcai-19-deductivelearning/deductivelearning.tex
Background shorten

@@
 
-%The specification of action models is a complex process that limits, too often, the application of {\em model-based planning} to real-world tasks~\cite{kambhampati:modellite:AAAI2007}.
-The {\em machine learning} of action models relieves this {\em knowledge acquisition bottleneck} of {\em model-based planning} and nowadays, there exists a wide range of effective approaches for learning action models~\cite{arora:amodels:ker2018}.
-Many of the most successful approaches for learning planning action models are however purely {\em inductive}~\cite{yang2007learning,pasula2007learning,mourao2010learning,zhuo2013action}, linking learning performance exclusively to the {\em amount} and {\em quality} of the input learning examples (which typically are observation of plan executions).
-
-%This paper addresses the learning of action models exploiting a different source of knowledge, {\em deductive} knowledge.
-Our approach leverages {\em state-invariants} (i.e.
-logic formulae that specify constraints about the possible states of a given domain) to cushion the negative impact of insufficient learning examples.
-Given an action model, state-of-the-art planners infer {\em state-invariants} from that model to reduce the search space and make the planning process more efficient~\cite{helmert2009concise}.
-In this paper we follow the opposite direction and leverage {\em state-invariants} to learn the planning action model.
-The benefit of learning action models from {\em state-invariants} is two-fold, {\em state-invariants} constrain the space of possible action models and can complete learning examples that are only partially observed.
-
-%Our approach for learning \strips\ action models from {\em state-invariants} is to compile the learning task into a classical planning task.
-Our compilation is built on top of the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} and it is flexible to different kinds of input knowledge including both partial observations of plan executions and {\em state-invariants}.
-The compilation outputs an \strips\ action model that is {\em consistent} with all the given input knowledge.
-The experimental results demonstrate that, even at unfavorable scenarios where input observations are minimal (a single learning example that comprises just a full initial state and a partially observed state), {\em state-invariant} are helpful to learn good quality \strips\ action models.
-
 \section{Background}
 \label{sec:background}
-This section formalizes the {\em planning model} we follow in this work and introduces the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning}.
-Finally, the section formalizes {\em state-invariants}.
 
-\subsection{Classical planning with conditional effects}
-Let $F$ be the set of  propositional state variables ({\em fluents}) describing a state.
-A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e.
-either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values).
-Given $L$, let $\neg L=\{\neg l:l\in L\}$ be its complement.
+We denote as $F$ (fluents) the set of  propositional state variables.
+A partial assignment of values to fluents is represented by $L$ (literals).
+We adopt the \emph{open world assumption} (what is not known to be true in a state is unknown) to implicitly represent the unobserved literals of a state.
+Hence, a state $s$ includes positive literals ($f$) and negative literals ($\neg f$) and it is defined as a full assignment of values to fluents; $|s|=|F|$.
 We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.
-A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$.
 
-A {\em classical planning action} $a\in A$ has: a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $cost(a)$.
-The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
+A {\em planning action} $a$ has a precondition list $\pre(a)\in\mathcal{L}(F)$ and a effect list $\eff(a)\in\mathcal{L}(F)$.
+The semantics of an action $a$ is specified with two functions: $\rho(s,a)$ denotes whether $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying $a$ in a state $s$.
 Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$, i.e.~if its precondition holds in $s$.
-The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
-Subtracting the complement of $\eff(a)$ from $s$ ensures that $\theta(s,a)$ remains a well-defined state.
-The subset of action effects that assign a positive value to a state fluent is called {\em positive effects} and denoted by $\eff^+(a)\in \eff(a)$ while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects} of an action $a\in A$.
+The result of executing an applicable action $a$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$, where $\neg\eff(a)$ is the complement of $\eff(a)$, and it is subtracted from $s$ so as to ensure that $\theta(s,a)$ remains a well-defined state.
 
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\in\mathcal{L}(F)$ is the set of goal conditions over the state variables.
-A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length} and $cost(\pi)=\sum_{a\in\pi} cost(a)$ its {\em plan cost}.
-The execution of $\pi$ on the initial state of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
-A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,P)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
-A solution plan is {\em optimal} iff its cost is minimal.
+%The subset of action effects that assign a positive value to a state fluent is called {\em positive effects} and denoted by $\eff^+(a)\in \eff(a)$ while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects} of an action $a\in A$.
 
-We also define {\em actions with conditional effects} because they are useful to compactly formulate our approach for {\em goal recognition with unknown domain models}.
-An action $a_c\in A$ with conditional effects is a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
-Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\neg\eff_c(s,a)\cup\eff_c(s,a)\}$ where $\eff_c(s,a)$ are the {\em triggered effects} resulting from the action application (conditional effects whose conditions hold in $s$):
-\[
-\eff_c(s,a)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
-\]
+A {\em planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\in\mathcal{L}(F)$ is the set of goal conditions over the state variables.
+A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length}.
+The execution of $\pi$ in $I$ induces a {\em trajectory} $\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A plan $\pi$ solves $P$ iff the induced trajectory reaches a final state $s_n$ such that $G \subseteq s_n$.
+
+The baseline learning approach (hereafter \textbf{BLA} for short) our proposal draws upon uses {\em actions with conditional effects}~\cite{aineto2018learning}.
+The conditional effects of an action $a_c$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+The {\em triggered effects} resulting from the action application (conditional effects whose conditions hold in $s$) is defined as $\eff_c(s,a)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E$.
 
 \subsection{Learning action models with classical planning}
+\label{FAMA}
+
 The {\em classical planning compilation} for the learning of \strips\ action models~\cite{aineto2018learning} receives as input an {\em empty} model (which contains just the {\em name} and {\em parameters} of each action schema), and a set of observations of plan executions.

### 2019-02-23 doc/ijcai-19-deductivelearning/deductivelearning.tex
Updated evaluation section IJCAI19

@@
 The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-We only used 1 learning examples for each learning task and we fixed the examples for all the experiments so that we can evaluate the impact of the different amount and source of the input knowledge in the quality of the learned models.
-All experiments are run on an Intel Core i5 3.10 GHz x 4 with 8 GB of RAM.
+For each domain we generated 10 trajectories of length 10 via random walks to be used as training examples through all the experiments.
+We also introduce a new parameter, the {\em degree of observability} $\sigma$, which indicates de probability of observing a literal in an intermediate state.
+This parameter is used to build training examples with varying degrees of incompleteness from the generated trajectories.
+All experiments are run on an Intel Core i5 3.10 GHz x 4 with 16 GB of RAM.
 
-The classical planner we used to solve the instances that result from our compilations is the SAT-based planner{\sc Madagascar}~\cite{rintanen2014madagascar}.
+%Through all the experiments, we only used one learning example for each learning task and we fixed the examples for all the experiments so that we can evaluate the impact of the different amount and source of the input knowledge in the quality of the learned models.
+
+%The classical planner we used to solve the instances that result from our compilations is the SAT-based planner{\sc Madagascar}~\cite{rintanen2014madagascar}.
 We used {\sc Madagascar} due to its ability to deal with planning instances populated with dead-ends~\cite{lopez2015deterministic}.
@@
 For the sake of reproducibility, the compilation source code, evaluation scripts, used benchmarks and input {\em state-invariants} are fully available at the repository {\em https://github.com/anonsub/}.
+
+\subsubsection{Metrics}
+The learned models are evaluated using the {\em precision} and {\em recall} metrics for action models proposed in \cite{aineto2018learning}, which compare the learned models against the reference model.
+
+Precision measures the correctness of the learned models.
+Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that appear in both the learned and reference action models) and $fp$ is the number of false positives (predicates that appear in the learned action model but not in the reference model).
+Recall, on the other hand, measures the completeness of the model and is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
+
+%Precision measures the {\em correctness} while recall gives a notion of the {\em completeness} of the learned models.
+Formally, $Precision=\frac{tp}{tp+fp}$, where $tp$ is the number of true positives (predicates that correctly appear in the action model) and $fp$ is the number of false positives (predicates appear in the learned action model that should not appear).
+Recall is formally defined as $Recall=\frac{tp}{tp+fn}$ where $fn$ is the number of false negatives (predicates that should appear in the learned action model but are missing).
+
+\subsection{Observability versus Knowledge}
+In our first experiment we seek to answer the question of whether knowledge can substitute observations. To that end, we evaluate the following 4 settings:
+
+\begin{itemize}
+	\item \textbf{Neither observability nor knowledge:} This is the baseline setting where the input sample is reduced to the minimum and only the initial and final state are known ($\sigma = 0$).
+	\item \textbf{Only knowledge:} Here we add domain-specific knowledge encoded as schematic mutexes to the baseline scenario.
+	\item \textbf{Only observability:} In this one, instead of knowledge we use a more complete input example where part of the intermediate states is known ($\sigma = 0.2$).
+	\item \textbf{Both observability and knowledge:} In the last setting we use both more complete input examples and schematic mutexes.
+\end{itemize}
+
+\begin{table}[hbt!]
+	\begin{center}
+		\begin{footnotesize}
+			\resizebox{0.45\textwidth}{!}{	
+				\begin{tabular}{l|c|c|c|c|c|c|c|c|c|}
+					& & \multicolumn{2}{|c|}{$\sigma = 0$} & \multicolumn{2}{|c|}{$\sigma = 0$ with $\Phi$} & \multicolumn{2}{|c|}{$\sigma = 0.2$} & \multicolumn{2}{|c|}{$\sigma = 0.2$ with $\Phi$}\\ \cline{3-10}			
+					& \multicolumn{1}{|c|}{$|\Phi|$} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} &  \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} \\
+					\hline
+					blocks & 9 & 0.52 & 0.38 & 0.53 & 0.21 & 0.66 & 0.56 & 0.77 & 0.68 \\
+					driverlog & 8 & 0.49 & 0.33 & 0.33 & 0.31 & 0.54 & 0.38 & 0.70 & 0.53 \\
+					ferry & 2 & 0.50 & 0.40 & 0.57 & 0.41 & 0.59 & 0.64 & 0.59 & 0.70 \\
+					floor-tile & 7 & 0.30 & 0.40 & 0.58 & 0.46 & 0.68 & 0.46 & 0.75 & 0.48 \\
+					grid & 3 & 0.47 & 0.40 & 0.47 & 0.37 & 0.43 & 0.34 & 0.43 & 0.32 \\
+					gripper & 5 & 0.77 & 0.56 & 0.77 & 0.54 & 0.85 & 0.74 & 0.96 & 0.83 \\
+					hanoi & 3 & 0.84 & 0.76 & 0.76 & 0.75 & 0.96 & 0.75 & 0.97 & 0.79 \\
+					n-puzzle & 3 & 0.94 & 0.86 & 0.93 & 0.86 & 0.99 & 0.87 & 1.00 & 0.87 \\
+					parking & 8 & 0.48 & 0.37 & 0.60 & 0.40 & 0.58 & 0.45 & 0.67 & 0.49 \\
+					transport & 4 & 0.45 & 0.45 & 0.53 & 0.46 & 0.99 & 0.51 & 0.94 & 0.79 \\
+					zeno-travel & 4 & 0.72 & 0.39 & 0.75 & 0.40 & 0.80 & 0.42 & 0.93 & 0.55 \\
+					\hline
+					&  & 0.59 & 0.48 & 0.62 & 0.47 & 0.73 & 0.56 & 0.79 & 0.64 \\
+
+				\end{tabular}
+			}
+		\end{footnotesize}			
+	\end{center}
+	\caption{\small Observability versus knowledge}
+	\label{tab:observability_vs_knowledge}
+\end{table}
+
+Table \ref{tab:observability_vs_knowledge} compiles the average precision (P) and recall (R) for each domain in the different settings tested.
+The table also reports the number of schematic mutexes ($|\Phi|$) used for each domain.
+Comparing the settings wtih only knowledge and only observability, it is clear that having more complete training examples is preferable.
+In fact, the improvement of using domain knowledge is marginal with respect to the baseline case.
+This is not the case for the last setting, where the use of domain knowledge shows a significant improvement in the quality of the learned models when compared to the setting with only observability.
+
+\subsection{Experiment 2}
+In the previous experiment we have stablished that domain knowledge is no substitute to observations, and that, given a minimum of observability, domain knowledge is able to enrich both the observations and learning process thus procuring better learned models.
+In this next experiment we are going to measure the improvement provided by domain knowledge at increasing degrees of observability.
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=\linewidth]{figures/comparison_precision.eps}
+	\caption{Comparison of the precision of the learned models for increasing degrees of observability.}
+	\label{fig:comparison_precision}
+\end{figure}
+
+\begin{figure}[hbt!]
+	\centering
+	\includegraphics[width=\linewidth]{figures/comparison_recall.eps}
+	\caption{Comparison of the recall of the learned models for increasing degrees of observability.}
+	\label{fig:comparison_recall}
+\end{figure}
+
+Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare the precision and recall of the learned models in the settings with and without domain knowledge.
+The values plotted in these figures are averages across all the domains seen in the previous experiment.
+The results show that using domain knowledge significantly improves the learned models no matter how complete the training examples are.
+Another interesting aspect is that domain knowledge is able to enrich observations in the range of 30\% observability to the level of fully observable trajectories, which means that domain knowledge can make up for a lack of completeness in the training examples.
 

### 2019-02-23 doc/ijcai-19-deductivelearning/deductivelearning.tex
updated sections 4 and 5 of IJCAI19

@@
 \subsection{Completing partially observed states with {\em schematic mutex}}
-Our sampling space follows the {\em open world} assumption, i.e.
+%Our sampling space follows the {\em open world} assumption, i.e.
 what is not observed is considered unknown.
 Here we describe a pre-processing mechanism to add new knowledge that completes the states $\tup{s_1^o \ldots, s_m^o}$ that are partially observed in a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task using an input set of {\em schematic mutex} $\Phi$.
+
+Here we describe a pre-processing mechanism to add new literals to complete the partial states $\tup{s_1^o \ldots, s_m^o}$ of an observation using a set of {\em schematic mutex} $\Phi$.
 
@@
 $\omega\subseteq\Omega^{pars(q)}$ represents the objects that unify the variables in $q$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and $\Phi$ contains the {\em schematic mutex} $\neg holding(v_1)\vee\neg clear(v_1)$, we can safely extend that state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal was actually unobserved).
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and $\Phi$ contains the {\em schematic mutex} $\neg holding(v_1)\vee\neg clear(v_1)$, we can safely extend that state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal being initially unknown).
 
@@
 \subsection{Pruning inconsistent action models with {\em schematic mutex}}
-We could extend the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} to check the consistency of the {\em state-constraints} in $\Phi$ at every state traversed by a solution to the compiled problem.
+%We could extend the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} to check the consistency of the {\em state-constraints} in $\Phi$ at every state traversed by a solution to the compiled problem.
 Unfortunately, checking arbitrary $\phi$ formulae is too expensive for current classical planners.
 
-Instead, our approach is to define a mechanism to check {\em state-constraints} in the form of {\em schematic mutex}.
+%Instead, our approach is to define a mechanism to check {\em state-constraints} in the form of {\em schematic mutex}.
 To implement this checking mechanism we add new conditional effects to the {\em insert} and {\em apply} actions of the classical planning compilation.
 Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the compilation and next, we describe them in detail:
+
+Our approach to learning action models consistent with the set of {\em state-constraints} in $\Phi$ is to ensure that newly generated states produced by the learned actions cannot introduce any inconsistencies.
+This is implemented by adding new conditional effects to the {\em insert} and {\em apply} actions of the classical planning compilation.
+Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the compilation and next, we describe them in detail:
+
 \begin{enumerate}
@@
 \begin{small}
-According to the $P_{\Lambda}$ compilation, once a given precondition or effect is inserted into the action model $\mathcal{M}$ it cannot be removed back.
-In addition, once the action model $\mathcal{M}$ is applied it cannot be {\em programmed}.
+According to the $P_{\Lambda}$ compilation, once a given precondition or effect is inserted into the domain model $\mathcal{M}$ it cannot be removed back.
+In addition, once an action model is applied it cannot be modified.
 In the compiled planning problem $P_{\Lambda}$, only $\mathsf{apply_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
@@
 $0\leq i\leq n$, is consistent with the input state observations and {\em state-invariants}.
-This is exactly the definition of the solution condition for an action model $\mathcal{M}'$ to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
+This is exactly the definition of the solution condition for model $\mathcal{M}'$ to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
 \end{small}
@@
 By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in a \strips\ action schema $\xi$ using predicates $\Psi$.
-In addition the $P_{\Lambda}$ compilation does not discard any possible action model $\mathcal{M}'$ definable within ${\mathcal I}_{\Psi,\xi}$ that satisfies the domain mutex in $\Phi$.
-This means that, for every \strips\ model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action model $\mathcal{M}'$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
+In addition the $P_{\Lambda}$ compilation does not discard any possible domain model $\mathcal{M}'$ definable within ${\mathcal I}_{\Psi,\xi}$ that satisfies the domain mutex in $\Phi$.
+This means that, for every \strips\ model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
 \end{small}
@@
 Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
-In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a shorter number of {\em insert} actions.
+In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a smaller number of {\em insert} actions.
 
@@
 {\em insert} actions have {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions have a {\em positive constant cost}).
-In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} when there are zero-cost actions.
+In practice we use a different approach to disregard the cost of {\em insert} actions since classical planners are not proficient at optimizing {\em plan cost} when there are zero-cost actions.
 Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} that can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
@@
 The plan horizon for programming any action model is then always bound to 2, which significantly reduces the planning horizon.
-The SAT-based planning approach is also convenient because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
+The SAT-based planning approach is also convenient because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect the planning performance.
 
@@
 \begin{itemize}
-	\item \textbf{Neither observability nor knowledge:} This is the baseline setting where the input sample is reduced to the minimum and only the initial and final state are known ($\sigma = 0$).
+	\item \textbf{Neither observability nor knowledge:} This is the baseline setting where the input sample is reduced to the minimum and only the initial and final states are known ($\sigma = 0$).
 	\item \textbf{Only knowledge:} Here we add domain-specific knowledge encoded as schematic mutexes to the baseline scenario.
@@
 
-\subsection{Experiment 2}
+\subsection{Using knowledge to counter incompleteness}
 In the previous experiment we have stablished that domain knowledge is no substitute to observations, and that, given a minimum of observability, domain knowledge is able to enrich both the observations and learning process thus procuring better learned models.

### 2019-02-23 doc/ijcai-19-deductivelearning/deductivelearning.tex
plural of mutex is mutexes

@@
 
-\section{Learning \strips\ action models from {\em schematic mutex} with classical planning}
+\section{Learning \strips\ action models from {\em schematic mutexes} with classical planning}
 \label{sec:compilation}
@@
 
-\subsection{Completing partially observed states with {\em schematic mutex}}
+\subsection{Completing partially observed states with {\em schematic mutexes}}
 %Our sampling space follows the {\em open world} assumption, i.e.
@@
 
-Here we describe a pre-processing mechanism to add new literals to complete the partial states $\tup{s_1^o \ldots, s_m^o}$ of an observation using a set of {\em schematic mutex} $\Phi$.
+Here we describe a pre-processing mechanism to add new literals to complete the partial states $\tup{s_1^o \ldots, s_m^o}$ of an observation using a set of {\em schematic mutexes} $\Phi$.
 
@@
 
-\subsection{Pruning inconsistent action models with {\em schematic mutex}}
+\subsection{Pruning inconsistent action models with {\em schematic mutexes}}
 %We could extend the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} to check the consistency of the {\em state-constraints} in $\Phi$ at every state traversed by a solution to the compiled problem.
@@
 \end{enumerate}
-In theory, conditional effects of the kind $4$ and $5$ are enough to guarantee that all the states traversed by a plan produced by the compilation are {\em consistent} with the input set of {\em schematic mutex} $\Phi$ (of course, provided that the input initial state $s_0^o$ is a valid state).
+In theory, conditional effects of the kind $4$ and $5$ are enough to guarantee that all the states traversed by a plan produced by the compilation are {\em consistent} with the input set of {\em schematic mutexes} $\Phi$ (of course, provided that the input initial state $s_0^o$ is a valid state).
 In practice we include also conditional effects of the kind $1$, $2$ and $3$ because they prune {\em invalid} action models at an earlier stage of the planning process (these effects extend the {\em insert} actions that always appear first in the solution plans).
@@
 By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in a \strips\ action schema $\xi$ using predicates $\Psi$.
-In addition the $P_{\Lambda}$ compilation does not discard any possible domain model $\mathcal{M}'$ definable within ${\mathcal I}_{\Psi,\xi}$ that satisfies the domain mutex in $\Phi$.
+In addition the $P_{\Lambda}$ compilation does not discard any possible domain model $\mathcal{M}'$ definable within ${\mathcal I}_{\Psi,\xi}$ that satisfies the mutexes in $\Phi$.
 This means that, for every \strips\ model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.

### 2019-02-23 doc/ijcai-19-deductivelearning/deductivelearning.tex
Changed schema for model in proof

@@
 \begin{small}
-By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in a \strips\ action schema $\xi$ using predicates $\Psi$.
+By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
 In addition the $P_{\Lambda}$ compilation does not discard any possible domain model $\mathcal{M}'$ definable within ${\mathcal I}_{\Psi,\xi}$ that satisfies the mutexes in $\Phi$.
-This means that, for every \strips\ model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
+This means that, for every model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
 \end{small}

### 2019-02-23 doc/ijcai-19-deductivelearning/deductivelearning.tex
Revisado hasta 3.2

@@
 \usepackage{xspace}
+\usepackage{xcolor}
 \urlstyle{same}
@@
 % Check the ijcai19-multiauthor.tex file for detailed instructions
-\author{
-Diego Aineto$^1$\and
-Sergio Jim\'enez$^1$\and
-Eva Onaindia$^1$
-\affiliations
-$^1${\small Departamento de Sistemas Inform\'aticos y Computaci\'on. Universitat Polit\`ecnica de Val\`encia. Valencia, Spain}
-\emails
-{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
+%\author{
+%Diego Aineto$^1$\and
+%Sergio Jim\'enez$^1$\and
+%Eva Onaindia$^1$
+%\affiliations
+%$^1${\small Departamento de Sistemas Inform\'aticos y Computaci\'on. Universitat Polit\`ecnica de Val\`encia. Valencia, Spain}
+%\emails
+%{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
 
@@
 These approaches return the statistically significant model that best explains the plan observations by minimizing some error metric.
-A model explains an observation if a plan containing the observed actions can be generated with the model and the states induced by this plan also include the possibly partially observed states.
+A model explains an observation if a plan containing the observed actions is computable with the model and the states induced by this plan also include the possibly partially observed states.
 The limitation of posing model validation as an optimization task over a testing set of observations is that it neither guarantees completeness (the model may not explain all the observations) nor correctness (the states induced by the execution of the plan generated with the model may contain contradictory information).
@@
 A {\em planning action} $a$ has a precondition list $\pre(a)\in\mathcal{L}(F)$ and a effect list $\eff(a)\in\mathcal{L}(F)$.
-The semantics of an action $a$ is specified with two functions: $\rho(s,a)$ denotes whether $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying $a$ in a state $s$.
-Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$, i.e.~if its precondition holds in $s$.
-The result of executing an applicable action $a$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$, where $\neg\eff(a)$ is the complement of $\eff(a)$, and it is subtracted from $s$ so as to ensure that $\theta(s,a)$ remains a well-defined state.
+The semantics of an action $a$ is specified with two functions: $\rho(s,a)$ denotes whether $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results from applying $a$ in a state $s$.
+Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$, i.e.~if its preconditions hold in $s$.
+The result of executing an applicable action $a$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$, where $\neg\eff(a)$ is the complement of $\eff(a)$, which is subtracted from $s$ so as to ensure that $\theta(s,a)$ remains a well-defined state.
+The subset of effects of an action $a$ that assign a positive value to a fluent is called {\em positive effects} and denoted by $\eff^+(a)\in \eff(a)$ while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects}.
 
@@
 
-The baseline learning approach (hereafter \textbf{BLA} for short) our proposal draws upon uses {\em actions with conditional effects}~\cite{aineto2018learning}.
+The baseline learning approach our proposal draws upon uses {\em actions with conditional effects}~\cite{aineto2018learning}.
 The conditional effects of an action $a_c$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
@@
 
-\subsection{Learning action models with classical planning}
+\subsection{Learning action models as planning}
 \label{FAMA}
 
-The {\em classical planning compilation} for the learning of \strips\ action models~\cite{aineto2018learning} receives as input an {\em empty} model (which contains just the {\em name} and {\em parameters} of each action schema), and a set of observations of plan executions.
-The compilation completes the {\em empty} model specifying the preconditions and effects of each action schema such that the validation of the completed model over the input observations is successful; i.e., there exists a plan computable with the completed model s.t.
-$\rho(s_{i-1}^o,a_i)$ and $s_i^o=\theta(s_{i-1}^o,a_i)$ holds for every observed state.
+The approach for learning \strips\ action models presented in~\cite{aineto2018learning}, which we will use as our baseline learning system (hereafter BLS, for short), is a compilation scheme that transforms the problem of learning the preconditions and effects of action models into a planning task $P'$.
+A \strips\ \emph{action model} $\xi$ is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
 
-A solution plan to the classical planning problem that results from the compilation is then a sequence of:
-\begin{itemize}
-\item \emph{Insert actions}, that insert preconditions and effects on an action schema.
-\item \emph{Apply actions} that validate the application of the completed model in the input observations.
-\end{itemize}
-Figure~\ref{fig:plan-lplan} shows a solution to a classical planning problem resulting from the~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} compilation corresponding to the {\em blocksworld}~\cite{slaney2001blocks}.
+The BLS receives as input an empty domain model, which only contains the headers of the action models, and a set of observations of plan executions, and creates a propositional encoding of the planning task $P'$.
+Let $\Psi$ be the set of {\em predicates}\footnote{The initial state of an observation is a full assignment of values to fluents, $|s_0|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $|s_0|$.} that shape the variables $F$.
+The set of propositions of $P'$ that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\Psi,\xi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
+For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), \\
+on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\Psi,\xi}$ will shape the preconditions, positive and negative effects of the action model $\xi$.
+
+The decision as to whether or not an element of ${\mathcal I}_{\Psi,\xi}$ will be part of $pre(\xi)$, $del(\xi)$ or $add(\xi)$ is given by the plan that solves $P'$.
+Specifically, two different sets of actions are included in the definition of $P'$: \emph{insert actions}, which insert preconditions and effects on an action model; and \emph{apply actions}, which validate the application of the learned action models in the input observations.
+Roughly speaking, in the \emph{blocksworld} domain, the \emph{insert actions} of a plan that solves $P'$ will look like {\tt{\footnotesize (insert\_pre\_stack\_holding\_v1),\\
+(insert\_eff\_stack\_clear\_v1),\\
+(insert\_eff\_stack\_clear\_v2)}}, where the second action denotes a positive effect and the third one a negative effect both to be inserted in the model of {\tt{\small stack}}; and the second set of actions of the plan that solves $P'$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the {\tt {\small apply}} actions must be validated with the observations of plan executions.
+
+In a nutshell, the output of the BLS compilation is a plan that completes the empty input domain model by specifying the preconditions and effects of each action model such that the validation of the completed model over the input observations is successful.
+
+%; i.e., there exists a plan computable with $M$ such that $\rho(s_{i-1}^o,a_i)$ and $s_i^o=\theta(s_{i-1}^o,a_i)$ holds for every observed state.
+
+%Figure~\ref{fig:plan-lplan} shows a solution to a classical planning problem resulting from the~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} compilation corresponding to the {\em blocksworld}~\cite{slaney2001blocks}.
 In the initial state of that problem the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
@@
 
-\begin{figure}[hbt!]
-	{\tiny\tt
-\begin{tabular}{ll}
-		{\bf 00}:(insert\_pre\_stack\_holding\_v1) & {\bf 10}:(insert\_eff\_pickup\_clear\_v1) \\
-		01:(insert\_pre\_stack\_clear\_v2) & 11:(insert\_eff\_pickup\_ontable\_v1)\\
-                {\bf 02}:(insert\_pre\_pickup\_handempty) & 12:(insert\_eff\_pickup\_handempty)\\
-                03:(insert\_pre\_pickup\_clear\_v1) & 13:(insert\_eff\_pickup\_holding\_v1)\\
-                04:(insert\_pre\_pickup\_ontable\_v1) & {\bf 14}:(apply\_pickup blockB)\\
-                {\bf 05}:(insert\_eff\_stack\_clear\_v1) & 15:(apply\_stack blockB blockC)\\
-                06:(insert\_eff\_stack\_clear\_v2) & 16:(apply\_pickup blockA)\\
-                07:(insert\_eff\_stack\_handempty) & 17:(apply\_stack blockA blockB) \\
-                08:(insert\_eff\_stack\_holding\_v1) &  {\bf 18}:(validate\_1)\\
-                09:(insert\_eff\_stack\_on\_v1\_v2) &             		
-\end{tabular}
-}
-	\caption{\small Example of a solution to a problem output by the classical planning compilation for the learning \strips\ action models.}
-	\label{fig:plan-lplan}
-\end{figure}
+%\begin{figure}[hbt!]
+%	{\tiny\tt
+%\begin{tabular}{ll}
+%	{\bf 00}:(insert\_pre\_stack\_holding\_v1) & {\bf 10}:(insert\_eff\_pickup\_clear\_v1) \\
+%		01:(insert\_pre\_stack\_clear\_v2) & 11:(insert\_eff\_pickup\_ontable\_v1)\\
+%                {\bf 02}:(insert\_pre\_pickup\_handempty) & 12:(insert\_eff\_pickup\_handempty)\\
+%                03:(insert\_pre\_pickup\_clear\_v1) & 13:(insert\_eff\_pickup\_holding\_v1)\\
+%                04:(insert\_pre\_pickup\_ontable\_v1) & {\bf 14}:(apply\_pickup blockB)\\
+%                {\bf 05}:(insert\_eff\_stack\_clear\_v1) & 15:(apply\_stack blockB blockC)\\
+%                06:(insert\_eff\_stack\_clear\_v2) & 16:(apply\_pickup blockA)\\
+%                07:(insert\_eff\_stack\_handempty) & 17:(apply\_stack blockA blockB) \\
+%                08:(insert\_eff\_stack\_holding\_v1) &  {\bf 18}:(validate\_1)\\
+%                09:(insert\_eff\_stack\_on\_v1\_v2) &             		
+%\end{tabular}
+%}
+%	\caption{\small Example of a solution to a problem output by the classical planning compilation for the learning \strips\ action models.}
+%	\label{fig:plan-lplan}
+%\end{figure}
 
@@
 
-\section{{\em One-shot} learning of planning action models from {\em domain-specific knowledge}}
+\section{{\em One-shot} learning task}
 \label{sec:learning}
-We define the {\em one-shot} learning of planning action models from {\em domain-specific knowledge} as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, where:
+
+The {\em one-shot} learning task to learn action models from {\em domain-specific knowledge} is defined as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, where:
+
 \begin{itemize}
-\item $\mathcal{M}$ is the {\em initial empty model} that contains only the name and parameters of each planning action to be learned.
-\item $\mathcal{O}$ is a single learning example that represents the observation of a sequence of states generated with the aimed planning action model.
-\item $\Phi$ is a set of logic formulae defining {\em domain-specific knowledge} that constraint the set of possible states.
+\item $\mathcal{M}$ is the {\em initial empty model} that contains only the header of each action model to be learned.
+\item $\mathcal{O}$ is a single learning example or plan observation; i.e.
+a sequence of (partially) observable states representing the evidence of the execution of an observed agent \textcolor[rgb]{1.00,0.00,0.00}{the observation of a sequence of states generated with the aimed planning action model?????}.
+\item $\Phi$ is a set of logic formulae that define {\em domain-specific knowledge}.
 \end{itemize}
@@
 A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ is a model $\mathcal{M}'$ s.t.
-there exists a plan computable with $\mathcal{M}'$ that is consistent with the the {\em initial empty model} $\mathcal{M}$, the single learning example $\mathcal{O}$ and the given {\em domain-specific knowledge} in $\Phi$.
+there exists a plan computable with $\mathcal{M}'$ that is consistent with the headers of $\mathcal{M}$, the observed states of $\mathcal{O}$ and the given domain knowledge in $\Phi$.
 
 \subsection{The space of \strips\ action models}
-{\em A \strips\ action schema} $\xi$ is defined by: A list of {\em parameters} $pars(\xi)$, and three sets of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters}, $pars(\xi)$.
-The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is the set of FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ and is denoted as ${\mathcal I}_{\Psi,\xi}$.
 
-For instance in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for the {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+We analyze here the search space of a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$; i.e., the space of \strips\ action models.
+In principle, for a given action model $\xi$, any element of ${\mathcal I}_{\Psi,\xi}$ can potentially appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
+In practice, the actual space of possible \strips\ schemata is bounded by:
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by:
 \begin{enumerate}
 \item {\bf Syntactic constraints}.
-\strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-{\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
+The solution $\mathcal{M}'$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
 \item {\bf Observation constraints}.
-The {\em learning examples}, that in our case is the single observation of a sequence of states, depict {\em semantic knowledge} that constraints further the space of possible action schemata.
+The solution $\mathcal{M}'$ must be consistent with these \emph{semantic constraints} derived from  learning samples, which in our case is a single plan observation.
+Specifically, the states induced by the plan computable with $M'$ must comprise the observed states of the sample.
 \end{enumerate}
+
+Considering only the syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\Psi,\xi}$ can appear both in the preconditions and effects of $\xi$.
+{\em Typing constraints} would also be a type of syntactic constraint~\cite{mcdermott1998pddl} in non-\strips\ models.
+Additionally, observation samples further constrains the space of possible action models.
+
+==================================
+\textcolor[rgb]{1.00,0.00,0.00}{HABLAR DE ESTO}
 
@@
 
+=====================================================
+
 \subsection{The sampling space}
-We define a {\em learning example} as a sequence $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$ of {\em partially observed states}, except for the initial state $s_0^o$ which is a {\em full state}.
+
+The single plan observation is defined as $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$, a sequence of possibly {\em partially observed states} except for the initial state $s_0^o$ which is a {\em fully observable} state.
 The set of predicates $\Psi$ and the set of objects $\Omega$ that shape the fluents $F$ is then deducible from $\mathcal{O}$.
@@
 
-To illustrate this Figure~\ref{fig:observation} shows a learning example that contains an initial state of the blocksworld where the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
-The second observation is a partially observed state in which {\tt blockA} is on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
+Figure~\ref{fig:observation} shows a learning example that contains an initial state of the blocksworld where the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
+The observation represents a partially observed state in which {\tt\small{blockA}} is on top of {\tt\small{blockB}} and {\tt\small{blockB}} on top of {\tt\small{blockC}}.
+
 \begin{figure}[hbt!]
@@
 
-\section{Learning \strips\ action models from {\em schematic mutexes} with classical planning}
+\section{Action model learning from {\em schematic mutexes}}
 \label{sec:compilation}

### 2019-02-24 doc/ijcai-19-deductivelearning/deductivelearning.tex
Section 3 updated

@@
 The BLS receives as input an empty domain model, which only contains the headers of the action models, and a set of observations of plan executions, and creates a propositional encoding of the planning task $P'$.
-Let $\Psi$ be the set of {\em predicates}\footnote{The initial state of an observation is a full assignment of values to fluents, $|s_0|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $|s_0|$.} that shape the variables $F$.
+Let $\Psi$ be the set of {\em predicates}\footnote{The initial state of an observation is a full assignment of values to fluents, $|s_0|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0$.} that shape the variables $F$.
 The set of propositions of $P'$ that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\Psi,\xi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
@@
 
-%; i.e., there exists a plan computable with $M$ such that $\rho(s_{i-1}^o,a_i)$ and $s_i^o=\theta(s_{i-1}^o,a_i)$ holds for every observed state.
-
-%Figure~\ref{fig:plan-lplan} shows a solution to a classical planning problem resulting from the~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} compilation corresponding to the {\em blocksworld}~\cite{slaney2001blocks}.
-In the initial state of that problem the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
-The problem goal is having the three-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
-The plan shows the {\em insert} actions for the {\tt\small stack} scheme (steps $00-01$ insert the preconditions, steps $05-10$ insert the effects), the plan steps $02-04$ that  insert the preconditions of the {\tt\small pickup} scheme and steps $10-13$ that insert the effects of this scheme.
-Finally, steps $14-17$ is a plan postfix with actions that apply the programmed model to achieve the goals starting from the given initial state.
-
-%\begin{figure}[hbt!]
-%	{\tiny\tt
-%\begin{tabular}{ll}
-%	{\bf 00}:(insert\_pre\_stack\_holding\_v1) & {\bf 10}:(insert\_eff\_pickup\_clear\_v1) \\
-%		01:(insert\_pre\_stack\_clear\_v2) & 11:(insert\_eff\_pickup\_ontable\_v1)\\
-%                {\bf 02}:(insert\_pre\_pickup\_handempty) & 12:(insert\_eff\_pickup\_handempty)\\
-%                03:(insert\_pre\_pickup\_clear\_v1) & 13:(insert\_eff\_pickup\_holding\_v1)\\
-%                04:(insert\_pre\_pickup\_ontable\_v1) & {\bf 14}:(apply\_pickup blockB)\\
-%                {\bf 05}:(insert\_eff\_stack\_clear\_v1) & 15:(apply\_stack blockB blockC)\\
-%                06:(insert\_eff\_stack\_clear\_v2) & 16:(apply\_pickup blockA)\\
-%                07:(insert\_eff\_stack\_handempty) & 17:(apply\_stack blockA blockB) \\
-%                08:(insert\_eff\_stack\_holding\_v1) &  {\bf 18}:(validate\_1)\\
-%                09:(insert\_eff\_stack\_on\_v1\_v2) &             		
-%\end{tabular}
-%}
-%	\caption{\small Example of a solution to a problem output by the classical planning compilation for the learning \strips\ action models.}
-%	\label{fig:plan-lplan}
-%\end{figure}
-
-%The {\em classical planning compilation} for the learning \strips\ action models~\cite{aineto2018learning} can be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects.
-These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers).
-Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
-
 \section{{\em One-shot} learning task}
@@
 \item $\mathcal{M}$ is the {\em initial empty model} that contains only the header of each action model to be learned.
-\item $\mathcal{O}$ is a single learning example or plan observation; i.e.
-a sequence of (partially) observable states representing the evidence of the execution of an observed agent \textcolor[rgb]{1.00,0.00,0.00}{the observation of a sequence of states generated with the aimed planning action model?????}.
+\item $\mathcal{O}$ is a single learning example or plan observation; i.e. a sequence of (partially) observable states representing the evidence of the execution of an observed agent.
 \item $\Phi$ is a set of logic formulae that define {\em domain-specific knowledge}.
@@
 The solution $\mathcal{M}'$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+{\em Typing constraints} would also be a type of syntactic constraint~\cite{mcdermott1998pddl}.
 \item {\bf Observation constraints}.
-The solution $\mathcal{M}'$ must be consistent with these \emph{semantic constraints} derived from  learning samples, which in our case is a single plan observation.
+The solution $\mathcal{M}'$ must be consistent with these \emph{semantic constraints} derived from  the learning samples $\mathcal{O}$, which in our case is a single plan observation.
 Specifically, the states induced by the plan computable with $M'$ must comprise the observed states of the sample.
@@
 
-Considering only the syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\Psi,\xi}$ can appear both in the preconditions and effects of $\xi$.
-{\em Typing constraints} would also be a type of syntactic constraint~\cite{mcdermott1998pddl} in non-\strips\ models.
+Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\Psi,\xi}$ can appear both in the preconditions and effects of $\xi$.
+In the propositional encoding of this work, a refined version over the encoding of the BLS, the belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with two binary representations, one for the preconditions and one for the effects.
+This way, the four possible combinations stand for the four possibilities of belonging of $e$ to $\xi$: no belonging, it only appears in $pre(\xi)$, it appears in $pre(\xi)$ and $del(\xi)$ and it only appears in $add(\xi)$.
+This compact encoding allows for a more effective exploitation of the syntactic constraints, and also yields the solution space of $\Lambda$ to be the same as its search space.
+
 Additionally, observation samples further constrains the space of possible action models.
-
-==================================
-\textcolor[rgb]{1.00,0.00,0.00}{HABLAR DE ESTO}
-
-In this work we introduce a novel propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ that uses only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\, so it is more compact that the one previously proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ action models with classical planning.
-In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
-If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
-Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema using the {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
-
-\begin{figure}
-  \begin{tiny}
-  \begin{verbatim}
-(:action stack
-   :parameters (?v1 ?v2)
-   :precondition (and (holding ?v1) (clear ?v2))
-   :effect (and (not (holding ?v1)) (not (clear ?v2))
-                (clear ?v1) (handempty) (on ?v1 ?v2)))
-
-(pre_holding_v1_stack) (pre_clear_v2_stack)
-(eff_holding_v1_stack) (eff_clear_v2_stack)
-(eff_clear_v1_stack) (eff_handempty_stack) (eff_on_v1_v2_stack)
-  \end{verbatim}
-  \end{tiny}
- \caption{\small PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.}
-\label{fig:propositional}
-\end{figure}
-
-=====================================================
 
@@
 
-The single plan observation is defined as $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$, a sequence of possibly {\em partially observed states} except for the initial state $s_0^o$ which is a {\em fully observable} state.
-The set of predicates $\Psi$ and the set of objects $\Omega$ that shape the fluents $F$ is then deducible from $\mathcal{O}$.
+The single plan observation of $\mathcal{O}$ is defined as $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$, a sequence of possibly {\em partially observed states} except for the initial state $s_0^o$ which is a {\em fully observable} state.
+As commented before, the predicates $\Psi$ and the objects that shape the fluents $F$ are then deducible from $s_0^o$.
 A partially observed state $s_i^o$, ${\small 1\leq i\leq m}$, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ was not observed.
-Intermediate states can be {\em missing}, meaning that they are {\em unobserved}, so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$ (where ${\small k\geq 1}$ is unknown but finite).
-The minimal expression of a learning example must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed state $s_m^o$ so $m \geq 1$.
+Intermediate states can be {\em missing}, meaning that they are unobservable, so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than one action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$ (with ${\small k\geq 1}$ is unknown but finite).
+The minimal expression of a learning sample must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed final state $s_m^o$ so $m \geq 1$.
 
 Figure~\ref{fig:observation} shows a learning example that contains an initial state of the blocksworld where the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
-The observation represents a partially observed state in which {\tt\small{blockA}} is on top of {\tt\small{blockB}} and {\tt\small{blockB}} on top of {\tt\small{blockC}}.
+The observation represents a partially observable final state in which {\tt\small{blockA}} is on top of {\tt\small{blockB}} and {\tt\small{blockB}} on top of {\tt\small{blockC}}.
 
@@
 \subsection{The domain-specific knowledge}
-Our approach is to introduce {\em domain-specific knowledge} in the form of {\em state-constraints} to restrict further the space of possible schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
-If we restrict ourselves to planning, {\em state-constraints} are abstractions for the compact specification of a set of states.
-For example, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em axiom} or {\em derived predicate} holds or the set of states that are considered goal states.
 
-{\em State-invariants} is a kind of state-constraints useful for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state-invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
-For instance, Figure~\ref{fig:strongest-invariant} shows five clauses that define {\em state-invariants} for the {\em blocksworld} planning domain.
+Our approach is to introduce {\em domain-specific knowledge} in the form of {\em state constraints} to further restrict the space of the action models.
+Back to the {\em blocksworld} domain, one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of any action model $\xi$ because, in this specific domain, a block cannot be on top of itself.
+The notion of state constraint is very general and has been used in different areas of AI and for different purposes.
+In planning, state constraints are compact and abstract representations that relate the values of variables in each state traversed by a plan, and allow to specify the set of states where a given action is applicable, the set of states where a given {\em axiom} or {\em derived predicate} holds or the set of states that are considered goal states~\cite{HaslumIR0TSN18}.
+
+{\em State invariants} is a useful type of state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a planning problem $P=\tup{F,A,I,G}$, a state invariant is a formula $\phi$ that holds in $I$, $I\models \phi$, and in every state $s$ built out of $F$ that is reachable by applying actions of $A$ in $I$.
+
+A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
+
+A {\em domain invariant} is an instance-independent state invariant that holds for any possible initial state and any possible set of objects.
+Therefore, if $s\nvDash \phi$ such that $\phi$ is a domain invariant, it means that $s$ is not a valid state.
+Domain invariants are often compactly defined as {\em lifted invariants} (also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}).
+
+In this work we exploit domain-specific knowledge that is given as  {\em schematic mutex}.
+We pay special attention to {\em schematic mutex} because they identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) an effective completion of a partially observed state and (2) an effective pruning of inconsistent \strips\ action models.
+
+We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\Psi,\xi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are {\em universally quantified}.
+For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
 
@@
 \end{footnotesize}
- \caption{\small Example of {\em state-invariants} for the {\em blocksworld} domain.}
+ \caption{\small Example of {\em state invariants} for the {\em blocksworld} domain.}
 \label{fig:strongest-invariant}
@@
 
-A {\em mutex} (mutually exclusive) is a {\em state-invariant} that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
-
-A {\em domain invariant} is an instance-independent state-invariant, i.e.
-holds for any possible initial state and any possible set of objects.
-Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
-Domain invariants are often compactly defined as {\em lifted invariants} (also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}).
-
-In this work we exploit {\em domain-specific knowledge} that is given as  {\em schematic mutex}.
-We pay special attention to {\em schematic mutex} because they identify mutually exclusive {\em properties} of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) effective completion of partially observed states and (2) effectively pruning of inconsistent \strips\ action models.
-We define a {\em schematic mutex} as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\Psi,\xi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are {\em universally quantified}.
-For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
+Figure~\ref{fig:strongest-invariant} shows five clauses that define state invariants for the {\em blocksworld} domain.
+The first two clauses define domain invariants and the three subsequent clauses define schematic mutexes.
 
@@
 \label{sec:compilation}
-This section shows how to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task with an off-the-shelf classical planner.
+
+In this section, we show how to exploit \emph{schematic mutexes} for solving the learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$.
+
+DECIDIR NOTATION:
+
+{\tt{\small (insert\_pre\_stack\_holding\_v1)}}
+
+$\mathcal{I}_{\xi,\Psi}$
+
+${\tt{\small (insert\_pre)_{\xi,p}}}$
+
+For $p \in \mathcal{I}_{\xi,\Psi}$ if $p$ belongs to $pre(\xi)$ then  $pre\_\xi\_p$.
+
+For $p \in \mathcal{I}_{\xi,\Psi}$ if $p$ belongs to $add(\xi)$  or $del(\xi)$ then $eff\_\xi\_p$.
+
+============================
 
@@
 
-Here we describe a pre-processing mechanism to add new literals to complete the partial states $\tup{s_1^o \ldots, s_m^o}$ of an observation using a set of {\em schematic mutexes} $\Phi$.
+Adding new literals to complete the partial states $\tup{s_1^o \ldots, s_m^o}$ of an observation $\mathcal{O}$  using a set of schematic mutexes $\Phi$ is done in a pre-processing stage.
 
@@
 Our approach to learning action models consistent with the set of {\em state-constraints} in $\Phi$ is to ensure that newly generated states produced by the learned actions cannot introduce any inconsistencies.
-This is implemented by adding new conditional effects to the {\em insert} and {\em apply} actions of the classical planning compilation.
+This is implemented by adding new conditional effects to the {\texttt \small {insert}} and {\texttt\small {apply}} actions of the classical planning compilation.
 Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the compilation and next, we describe them in detail:

### 2019-02-24 deductivelearning.tex
Small updates

--- 2019-02-24 doc/ijcai-19-deductivelearning/deductivelearning.tex
@@
 Let $\Psi$ be the set of {\em predicates}\footnote{The initial state of an observation is a full assignment of values to fluents, $|s_0|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0$.} that shape the variables $F$.
-The set of propositions of $P'$ that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\Psi,\xi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
-For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), \\
+The set of propositions of $P'$ that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
+For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), \\
 on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\Psi,\xi}$ will shape the preconditions, positive and negative effects of the action model $\xi$.
+Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of the action model $\xi$.
 
-The decision as to whether or not an element of ${\mathcal I}_{\Psi,\xi}$ will be part of $pre(\xi)$, $del(\xi)$ or $add(\xi)$ is given by the plan that solves $P'$.
+The decision as to whether or not an element of ${\mathcal I}_{\xi,\Psi}$ will be part of $pre(\xi)$, $del(\xi)$ or $add(\xi)$ is given by the plan that solves $P'$.
 Specifically, two different sets of actions are included in the definition of $P'$: \emph{insert actions}, which insert preconditions and effects on an action model; and \emph{apply actions}, which validate the application of the learned action models in the input observations.
-Roughly speaking, in the \emph{blocksworld} domain, the \emph{insert actions} of a plan that solves $P'$ will look like {\tt{\footnotesize (insert\_pre\_stack\_holding\_v1),\\
-(insert\_eff\_stack\_clear\_v1),\\
+Roughly speaking, in the \emph{blocksworld} domain, the \emph{insert actions} of a plan that solves $P'$ will look like {\tt{\footnotesize(insert\_pre\_stack\_holding\_v1)}},\\
+{\tt{\footnotesize(insert\_eff\_stack\_clear\_v1),\\
 (insert\_eff\_stack\_clear\_v2)}}, where the second action denotes a positive effect and the third one a negative effect both to be inserted in the model of {\tt{\small stack}}; and the second set of actions of the plan that solves $P'$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the {\tt {\small apply}} actions must be validated with the observations of plan executions.
@@
 We analyze here the search space of a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$; i.e., the space of \strips\ action models.
-In principle, for a given action model $\xi$, any element of ${\mathcal I}_{\Psi,\xi}$ can potentially appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
+In principle, for a given action model $\xi$, any element of ${\mathcal I}_{\xi,\Psi}$ can potentially appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
 In practice, the actual space of possible \strips\ schemata is bounded by:
@@
 
-Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\Psi,\xi}$ can appear both in the preconditions and effects of $\xi$.
+Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\xi,\Psi}|}$ because one element in $\mathcal{I}_{\Psi,\xi}$ can appear both in the preconditions and effects of $\xi$.
 In the propositional encoding of this work, a refined version over the encoding of the BLS, the belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with two binary representations, one for the preconditions and one for the effects.
@@
 
-A {\em domain invariant} is an instance-independent state invariant that holds for any possible initial state and any possible set of objects.
+Recently, some works point at extracting \emph{lifted} invariants, also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}, that hold for any possible state and any possible set of objects.
+Invariant
+templates obtained by inspecting the lifted representation of the domain have also been exploited for deriving \emph{lifted mutex}~\cite{BernardiniFS18}.
+In this work we exploit domain-specific knowledge that is given as {\em schematic mutex}.
+We pay special attention to {\em schematic mutex} because they identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) an effective completion of a partially observed state and (2) an effective pruning of inconsistent \strips\ action models.
+
+%A {\em domain invariant} is an instance-independent state invariant that holds for any possible initial state and any possible set of objects.
 Therefore, if $s\nvDash \phi$ such that $\phi$ is a domain invariant, it means that $s$ is not a valid state.
@@
 
-In this work we exploit domain-specific knowledge that is given as  {\em schematic mutex}.
-We pay special attention to {\em schematic mutex} because they identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) an effective completion of a partially observed state and (2) an effective pruning of inconsistent \strips\ action models.
-
-We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\Psi,\xi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are {\em universally quantified}.
+We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\xi,\Psi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are {\em universally quantified}.
 For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
@@
 {\bf ID} & {\bf Action} & {\bf New conditional effect}\\\hline
-1&$\mathsf{insertPre_{p,\xi}}$&$\{pre\_q\_\xi\}\rhd\{invalid\}$\\
-2&$\mathsf{insertEff_{p,\xi}}$&$\{pre\_q\_\xi\wedge eff\_q\_\xi\wedge pre\_p\_\xi\}\rhd\{invalid\}$\\
-3&$\mathsf{insertEff_{p,\xi}}$&$\{\neg pre\_q\_\xi\wedge eff\_q\_\xi\wedge \neg pre\_p\_\xi\}\rhd\{invalid\}$\\
-4&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
+%1&$\mathsf{insertPre_{p,\xi}}$&$\{pre\_q\_\xi\}\rhd\{invalid\}$\\
+1&${\tt{\small (insert\_pre)_{\xi,p}}}$&$\{pre\_q\_\xi\}\rhd\{invalid\}$\\
+%2&$\mathsf{insertEff_{p,\xi}}$&$\{pre\_q\_\xi\wedge eff\_q\_\xi\wedge pre\_p\_\xi\}\rhd\{invalid\}$\\
+2&${\tt{\small (insert\_eff)_{\xi,p}}}$&$\{pre\_q\_\xi\wedge eff\_q\_\xi\wedge pre\_p\_\xi\}\rhd\{invalid\}$\\
+%3&$\mathsf{insertEff_{p,\xi}}$&$\{\neg pre\_q\_\xi\wedge eff\_q\_\xi\wedge \neg pre\_p\_\xi\}\rhd\{invalid\}$\\
+3&${\tt{\small (insert\_eff)_{\xi,p}}}$&$\{\neg pre\_q\_\xi\wedge eff\_q\_\xi\wedge \neg pre\_p\_\xi\}\rhd\{invalid\}$\\
+%4&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
+%&&$q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{invalid\}$\\
+4&${\tt{\small (apply)_{\xi,\omega}}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
 &&$q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{invalid\}$\\
-5&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
-&&$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{invalid\}$
+5&${\tt{\small (apply)_{\xi,\omega}}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
+%&&$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{invalid\}$
 \end{tabular}
@@
 \item[1-3] For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ one conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$ one conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
 Likewise, two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions, one to ban the insertion of two positive effects that are {\em schematic mutex} and another one to ban two mutex negative effects.
 \item[4-5] For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation (in $\mathsf{apply_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
+both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation (in $\mathsf{apply_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
 \end{enumerate}
@@
 \begin{small}
-By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
-In addition the $P_{\Lambda}$ compilation does not discard any possible domain model $\mathcal{M}'$ definable within ${\mathcal I}_{\Psi,\xi}$ that satisfies the mutexes in $\Phi$.
+By definition ${\mathcal I}_{\xi,\Psi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
+In addition the $P_{\Lambda}$ compilation does not discard any possible domain model $\mathcal{M}'$ definable within ${\mathcal I}_{\xi,\Psi}$ that satisfies the mutexes in $\Phi$.
 This means that, for every model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
@@
 The size of the classical planning task $P_{\Lambda}$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
-The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$.
-The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $pre\_e\_\xi/eff\_e\_\xi$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
-Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\Psi,\xi}|$ and hence, the size of the classical planning task output by the compilation.
+The larger these arities, the larger $|{\mathcal I}_{\xi,\Psi}|$.
+The size of the ${\mathcal I}_{\xi,\Psi}$ set is the term that dominates the compilation size because it defines the $pre\_e\_\xi/eff\_e\_\xi$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
+Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\xi,\Psi}|$ and hence, the size of the classical planning task output by the compilation.
 
 Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
-In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a smaller number of {\em insert} actions.
+In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a smaller number of {\em insert} actions.
 
@@
 Our compilation is also flexible to this particular learning scenario.
-The known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to true in the initial state.
+The known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ to true in the initial state.
 Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, making the classical planning task $P_{\Lambda}$ easier to be solved.

### 2019-02-24 doc/ijcai-19-deductivelearning/deductivelearning.tex
Unified notation for section 4

@@
 
-Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\Psi,\xi}$ can appear both in the preconditions and effects of $\xi$.
-In the propositional encoding of this work, a refined version over the encoding of the BLS, the belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with two binary representations, one for the preconditions and one for the effects.
-This way, the four possible combinations stand for the four possibilities of belonging of $e$ to $\xi$: no belonging, it only appears in $pre(\xi)$, it appears in $pre(\xi)$ and $del(\xi)$ and it only appears in $add(\xi)$.
+Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
+In this work, the belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre\_\xi\_e$ and $eff\_\xi\_e$, instead of the three used in the BLS.
+The four possible combinations of these fluents are sumarized in Figure \ref{fig:combinations}.
 This compact encoding allows for a more effective exploitation of the syntactic constraints, and also yields the solution space of $\Lambda$ to be the same as its search space.
+
+\begin{figure}
+	\begin{footnotesize}
+		\begin{tabular}{lll}
+			{\bf Combination} & {\bf Meaning}\\\hline
+			$\neg pre\_\xi\_e \wedge \neg eff\_\xi\_e $&Neither precondition nor effect\\
+			$pre\_\xi\_e \wedge \neg eff\_\xi\_e $&Precondition\\
+			$\neg pre\_\xi\_e \wedge eff\_\xi\_e $&Positive effect\\
+			$pre\_\xi\_e \wedge eff\_\xi\_e  $&Negative effect\\
+		\end{tabular}
+	\end{footnotesize}
+	\caption{\small Summary of possible combinations of the propositional encoding fluents and their meaning}
+	\label{fig:combinations}
+\end{figure}
 
@@
 
-In this work we exploit domain-specific knowledge that is given as  {\em schematic mutex}.
-We pay special attention to {\em schematic mutex} because they identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) an effective completion of a partially observed state and (2) an effective pruning of inconsistent \strips\ action models.
-
-We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\Psi,\xi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are {\em universally quantified}.
+We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\xi,\Psi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are {\em universally quantified}.
 For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
+Figure~\ref{fig:strongest-invariant} shows four clauses that define schematic mutexes for the {\em blocksworld} domain.
 
@@
   \begin{footnotesize}
-$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
-$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
+$\forall x_1,x_2\ \neg ontable(x_1)\vee\neg on(x_1,x_2)$.\\
+$\forall x_1,x_2\ \neg clear(x_1)\vee\neg on(x_2,x_1)$.\\
 $\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
 $\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
-$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
 \end{footnotesize}
- \caption{\small Example of {\em state invariants} for the {\em blocksworld} domain.}
+ \caption{\small Example of {\em schematic mutexes} for the {\em blocksworld} domain.}
 \label{fig:strongest-invariant}
 \end{figure}
-
-Figure~\ref{fig:strongest-invariant} shows five clauses that define state invariants for the {\em blocksworld} domain.
-The first two clauses define domain invariants and the three subsequent clauses define schematic mutexes.
 
@@
 In this section, we show how to exploit \emph{schematic mutexes} for solving the learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$.
-
-DECIDIR NOTATION:
-
-{\tt{\small (insert\_pre\_stack\_holding\_v1)}}
-
-$\mathcal{I}_{\xi,\Psi}$
-
-${\tt{\small (insert\_pre)_{\xi,p}}}$
-
-For $p \in \mathcal{I}_{\xi,\Psi}$ if $p$ belongs to $pre(\xi)$ then  $pre\_\xi\_p$.
-
-For $p \in \mathcal{I}_{\xi,\Psi}$ if $p$ belongs to $add(\xi)$  or $del(\xi)$ then $eff\_\xi\_p$.
-
-============================
 
@@
 
-Let us define the production rule $p\rightarrow\neg q$ such that $\tup{p,q}$ is a {\em schematic mutex}.
+Let $\Omega$ be the set of objects that appear in $F$ as the values of the arguments of the predicates $\Psi$, and $p\rightarrow\neg q$ a production rule such that $\tup{p,q}$ is a {\em schematic mutex}.
 Given a {\em schematic mutex} $\tup{p,q}\in \Phi$ and a state observation $s_j^o\in \mathcal{O}$, {\small $(1\leq j\leq m)$} then $s_j^o$ can be safely completed adding the new literals $\neg q(\omega)$ that result from the unification of the corresponding production rule with $s_j^o$.
-$\omega\subseteq\Omega^{pars(q)}$ represents the objects that unify the variables in $q$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
+$\omega\subseteq\Omega^{args(q)}$ represents the objects that unify the variables in $q$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
 For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and $\Phi$ contains the {\em schematic mutex} $\neg holding(v_1)\vee\neg clear(v_1)$, we can safely extend that state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal being initially unknown).
@@
 			{\bf ID} & {\bf Action} & {\bf New conditional effect}\\\hline
-1&$\mathsf{insertPre_{p,\xi}}$&$\{pre\_q\_\xi\}\rhd\{invalid\}$\\
-2&$\mathsf{insertEff_{p,\xi}}$&$\{pre\_q\_\xi\wedge eff\_q\_\xi\wedge pre\_p\_\xi\}\rhd\{invalid\}$\\
-3&$\mathsf{insertEff_{p,\xi}}$&$\{\neg pre\_q\_\xi\wedge eff\_q\_\xi\wedge \neg pre\_p\_\xi\}\rhd\{invalid\}$\\
-4&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
-&&$q(\omega)\wedge \neg pre\_q\_\xi\}\rhd\{invalid\}$\\
-5&$\mathsf{apply_{\xi,\omega}}$&$\{\neg pre\_p\_\xi \wedge eff\_p\_\xi \wedge $\\
-&&$q(\omega)\wedge pre\_q\_\xi \wedge \neg eff\_q\_\xi\}\rhd\{invalid\}$
+			1&${\tt (insert\_pre)_{\xi,p}}$&$\{pre\_\xi\_q\}\rhd\{invalid\}$\\
+			2&${\tt (insert\_eff)_{\xi,p}}$&$\{pre\_\xi\_q\wedge eff\_\xi\_q\wedge pre\_\xi\_p\}\rhd\{invalid\}$\\
+			3&${\tt (insert\_eff)_{\xi,p}}$&$\{\neg pre\_\xi\_q\wedge eff\_\xi\_q\wedge \neg pre\_\xi\_p\}\rhd\{invalid\}$\\
+			4&${\tt (apply)_{\xi,\omega}}$&$\{\neg pre\_\xi\_p \wedge eff\_\xi\_p \wedge $\\
+			&&$q(\omega)\wedge \neg pre\_\xi\_q\}\rhd\{invalid\}$\\
+			5&${\tt (apply)_{\xi,\omega}}$&$\{\neg pre\_\xi\_p \wedge eff\_\xi\_p \wedge $\\
+			&&$q(\omega)\wedge \neg eff\_\xi\_q\}\rhd\{invalid\}$
 		\end{tabular}
@@
 Our approach to learning action models consistent with the set of {\em state-constraints} in $\Phi$ is to ensure that newly generated states produced by the learned actions cannot introduce any inconsistencies.
-This is implemented by adding new conditional effects to the {\texttt \small {insert}} and {\texttt\small {apply}} actions of the classical planning compilation.
+This is implemented by adding new conditional effects to the ${\tt \small insert}$ and ${\tt\small apply}$ actions of the classical planning compilation.
 Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the compilation and next, we describe them in detail:
@@
 	\item[1-3] For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ one conditional effect is added to the $\mathsf{insertPre_{p,\xi}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
-Likewise, two conditional effects are added to the $\mathsf{insertEff_{p,\xi}}$ actions, one to ban the insertion of two positive effects that are {\em schematic mutex} and another one to ban two mutex negative effects.
+both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$ one conditional effect is added to the ${\tt \small (insert\_pre)_{\xi,p}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
+Likewise, two conditional effects are added to the ${\tt \small (insert\_eff)_{\xi,p}}$ actions, one to ban the insertion of two positive effects that are {\em schematic mutex} and another one to ban two mutex negative effects.
 	\item[4-5] For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\Psi,\xi}$ two conditional effects are added to the $\mathsf{apply_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation (in $\mathsf{apply_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
+both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$ two conditional effects are added to the ${\tt \small (apply)_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation (in ${\tt \small (apply)_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
 \end{enumerate}
 In theory, conditional effects of the kind $4$ and $5$ are enough to guarantee that all the states traversed by a plan produced by the compilation are {\em consistent} with the input set of {\em schematic mutexes} $\Phi$ (of course, provided that the input initial state $s_0^o$ is a valid state).
-In practice we include also conditional effects of the kind $1$, $2$ and $3$ because they prune {\em invalid} action models at an earlier stage of the planning process (these effects extend the {\em insert} actions that always appear first in the solution plans).
+In practice we include also conditional effects of the kind $1$, $2$ and $3$ because they prune {\em invalid} action models at an earlier stage of the planning process (these effects extend the ${\tt \small insert}$ actions that always appear first in the solution plans).
 
 The goals of the classical planning problem output by the original compilation are extended with the $\neg invalid$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by solution plans.
-Remarkably, the $\neg invalid$ literal allows us also to define $\mathsf{apply_{\xi,\omega}}$ actions more compactly than in the original compilation by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
-Disjunctions are no longer required to code the possible preconditions of an action schema since they can now be encoded with conditional effects of this kind $\{pre\_p\_\xi\wedge \neg p(\omega)\}\rhd\{invalid\}$.
+Remarkably, the $\neg invalid$ literal allows us also to define ${\tt \small (apply)_{\xi,\omega}}$ actions more compactly than in the original compilation by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
+Disjunctions are no longer required to code the possible preconditions of an action schema since they can now be encoded with conditional effects of this kind $\{pre\_\xi\_p\wedge \neg p(\omega)\}\rhd\{invalid\}$.
 
@@
 \begin{mylemma}
-Soundness. Any classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ produces a \strips\ model $\mathcal{M'}$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
+Soundness. Any classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ produces a model $\mathcal{M'}$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
 \end{mylemma}
@@
 In addition, once an action model is applied it cannot be modified.
-In the compiled planning problem $P_{\Lambda}$, only $\mathsf{apply_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
-This means that a state consistent with an observation $s_n^o$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_i$, s.t.
+In the compiled planning problem $P_{\Lambda}$, only ${\tt \small (apply)_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
+This means that a state consistent with an observation $s_n^o$ can only be achieved executing an applicable sequence of ${\tt \small (apply)_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_i$, s.t.
 $0\leq i\leq n$, is consistent with the input state observations and {\em state-invariants}.
@@
 \begin{mylemma}
-Completeness.
-Any \strips\ model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task can be computed with a classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$.
+Completeness. Any model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task can be computed with a classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$.
 \end{mylemma}
@@
 \begin{small}
-By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
-In addition the $P_{\Lambda}$ compilation does not discard any possible domain model $\mathcal{M}'$ definable within ${\mathcal I}_{\Psi,\xi}$ that satisfies the mutexes in $\Phi$.
-This means that, for every model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$ actions for {\em programming} the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
+By definition ${\mathcal I}_{\xi,\Psi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
+In addition the $P_{\Lambda}$ compilation does not discard any possible domain model $\mathcal{M}'$ definable within ${\mathcal I}_{\xi,\Psi}$ that satisfies the mutexes in $\Phi$.
+This means that, for every model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate ${\tt \small (insert\_pre)_{\xi,e}}$ and ${\tt \small (insert\_eff)_{\xi,e}}$ actions for {\em programming} the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding ${\tt \small (apply)_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
 \end{small}
@@
 The size of the classical planning task $P_{\Lambda}$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
-The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$.
-The size of the ${\mathcal I}_{\Psi,\xi}$ set is the term that dominates the compilation size because it defines the $pre\_e\_\xi/eff\_e\_\xi$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects in the $\mathsf{apply_{\xi,\omega}}$ actions.
-Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\Psi,\xi}|$ and hence, the size of the classical planning task output by the compilation.
+The larger these arities, the larger $|{\mathcal I}_{\xi,\Psi}|$.
+The size of the ${\mathcal I}_{\xi,\Psi}$ set is the term that dominates the compilation size because it defines the $pre\_\xi\_e/eff\_\xi\_e$ fluents, the corresponding set of ${\tt \small insert}$ actions, and the number of conditional effects in the ${\tt \small (apply)_{\xi,\omega}}$ actions.
+Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\xi,\Psi}|$ and hence, the size of the classical planning task output by the compilation.
 
 Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
-In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a smaller number of {\em insert} actions.
+In more detail, all $\{pre\_\xi\_e, eff\_\xi\_e\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a smaller number of {\em insert} actions.
 
 This bias could be eliminated defining a cost function for the actions in $P_\Lambda$ (e.g.
-{\em insert} actions have {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions have a {\em positive constant cost}).
+${\tt \small insert}$ actions have {\em zero cost} while ${\tt \small (apply)_{\xi,\omega}}$ actions have a {\em positive constant cost}).
 In practice we use a different approach to disregard the cost of {\em insert} actions since classical planners are not proficient at optimizing {\em plan cost} when there are zero-cost actions.

### 2019-02-24 doc/ijcai-19-deductivelearning/deductivelearning.tex
Until properties

@@
 Specifically, two different sets of actions are included in the definition of $P'$: \emph{insert actions}, which insert preconditions and effects on an action model; and \emph{apply actions}, which validate the application of the learned action models in the input observations.
-Roughly speaking, in the \emph{blocksworld} domain, the \emph{insert actions} of a plan that solves $P'$ will look like {\tt{\footnotesize(insert\_pre\_stack\_holding\_v1)}},\\
+Roughly speaking, in the \emph{blocksworld} domain, the insert actions of a plan that solves $P'$ will look like {\tt{\footnotesize(insert\_pre\_stack\_holding\_v1)}},\\
 {\tt{\footnotesize(insert\_eff\_stack\_clear\_v1),\\
@@
 Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
-In this work, the belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre\_\xi\_e$ and $eff\_\xi\_e$, instead of the three used in the BLS.
-The four possible combinations of these fluents are sumarized in Figure \ref{fig:combinations}.
+In this work, the belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre\_\xi\_e$ and $eff\_\xi\_e$, instead of the three fluents used in the BLS.
+The four possible combinations of these two fluents are sumarized in Figure \ref{fig:combinations}.
 This compact encoding allows for a more effective exploitation of the syntactic constraints, and also yields the solution space of $\Lambda$ to be the same as its search space.
@@
 			{\bf Combination} & {\bf Meaning}\\\hline
-			$\neg pre\_\xi\_e \wedge \neg eff\_\xi\_e $&Neither precondition nor effect\\
-			$pre\_\xi\_e \wedge \neg eff\_\xi\_e $&Precondition\\
-			$\neg pre\_\xi\_e \wedge eff\_\xi\_e $&Positive effect\\
-			$pre\_\xi\_e \wedge eff\_\xi\_e  $&Negative effect\\
+			$\neg pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ belongs neither to the preconditions \\
+             & nor effects of $\xi$ \\
+             & ($e \notin pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$)\\
+			$pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ is only a precondition of $\xi$\\
+               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$) \\
+			$\neg pre\_\xi\_e \wedge eff\_\xi\_e $& $e$ is a positive effect of $\xi$ \\
+               &  ($e \notin pre(\xi) \wedge e \in add(\xi) \wedge e \notin del(\xi)$) \\
+			$pre\_\xi\_e \wedge eff\_\xi\_e  $& $e$ is a negative effect of $\xi$ \\
+               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \in add(\xi)$) \\
 		\end{tabular}
 	\end{footnotesize}
-	\caption{\small Summary of possible combinations of the propositional encoding fluents and their meaning}
+	\caption{\small Combinations of the fluent propositional encoding and their meaning}
 	\label{fig:combinations}
@@
 
-We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\xi,\Psi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and such that they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are {\em universally quantified}.
-For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible {\em blocksworld} state.
+We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\xi,\Psi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are universally quantified.
+For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible state.
 Figure~\ref{fig:strongest-invariant} shows four clauses that define schematic mutexes for the {\em blocksworld} domain.
@@
 \end{footnotesize}
- \caption{\small Example of {\em schematic mutexes} for the {\em blocksworld} domain.}
+ \caption{\small {\em Schematic mutexes} for the {\em blocksworld} domain.}
 \label{fig:strongest-invariant}
@@
 
-Adding new literals to complete the partial states $\tup{s_1^o \ldots, s_m^o}$ of an observation $\mathcal{O}$  using a set of schematic mutexes $\Phi$ is done in a pre-processing stage.
+The addition of new literals to complete the partial states $\tup{s_1^o \ldots, s_m^o}$ of an observation $\mathcal{O}$  using a set of schematic mutexes $\Phi$ is done in a pre-processing stage.
 
-Let $\Omega$ be the set of objects that appear in $F$ as the values of the arguments of the predicates $\Psi$, and $p\rightarrow\neg q$ a production rule such that $\tup{p,q}$ is a {\em schematic mutex}.
-Given a {\em schematic mutex} $\tup{p,q}\in \Phi$ and a state observation $s_j^o\in \mathcal{O}$, {\small $(1\leq j\leq m)$} then $s_j^o$ can be safely completed adding the new literals $\neg q(\omega)$ that result from the unification of the corresponding production rule with $s_j^o$.
-$\omega\subseteq\Omega^{args(q)}$ represents the objects that unify the variables in $q$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-For instance, if the literal {\tt\small holding(blockA)} is observed in a particular blocksword state and $\Phi$ contains the {\em schematic mutex} $\neg holding(v_1)\vee\neg clear(v_1)$, we can safely extend that state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal being initially unknown).
+Let $\Omega$ be the set of objects that appear in $F$ as the values of the arguments of the predicates $\Psi$, and $\Phi=\tup{p,q}$ a schematic mutex.
+There exist many possible instantiations of $\Phi$ of the type $\tup{p(\omega),q(\omega')}$ with objects of $\Omega$, where $\omega\subseteq\Omega^{args(p)}$ and $\omega'\subseteq\Omega^{args(q)}$.
+Let us now assume that the instantiation $p(\omega) \in s_j^o$, {\small $(1\leq j\leq m)$}, being $s_j^o$ a partially observed state of $\mathcal{O}$.
+Then, two situations may occur: (a) $q(\omega') \in s_j^o$, in which case the expression $\neg p(\omega) \vee \neg q(\omega')$ holds in $s_j^o$; or (b) $q(\omega') \notin s_j^o$, in which case the literal $\neg q(\omega')$ has not been observed in $s_j^o$ and so we can safely complete the state with $\neg q(\omega')$ (the same applies inversely, when $q(\omega') \in s_j^o$ but $\neg p(\omega) \notin s_j^o$).
+In other words, if we find that one component of a schematic mutex is positively observed in a state and the other component is not observable in such state, we can complete the state with the missing negative literal.
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular state and $\Phi$ contains the schematic mutex $\neg holding(v_1)\vee\neg clear(v_1)$, we extend the state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal being initially unknown).
 
@@
 
-Our approach to learning action models consistent with the set of {\em state-constraints} in $\Phi$ is to ensure that newly generated states produced by the learned actions cannot introduce any inconsistencies.
-This is implemented by adding new conditional effects to the ${\tt \small insert}$ and ${\tt\small apply}$ actions of the classical planning compilation.
+Our approach to learning action models consistent with the schematic mutexes in $\Phi$ is to ensure that newly generated states induced by the learned actions do not introduce any inconsistency.
+This is implemented by adding new conditional effects to the ${\tt \small insert}$ and ${\tt\small apply}$ actions of the BLS compilation.
 Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the compilation and next, we describe them in detail:
@@
 \begin{enumerate}
-	\item[1-3] For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$ one conditional effect is added to the ${\tt \small (insert\_pre)_{\xi,p}}$ actions to ban the insertion of two preconditions that are {\em schematic mutex}.
-Likewise, two conditional effects are added to the ${\tt \small (insert\_eff)_{\xi,p}}$ actions, one to ban the insertion of two positive effects that are {\em schematic mutex} and another one to ban two mutex negative effects.
-	\item[4-5] For every {\em schematic mutex} $\tup{p,q}$ s.t.
-both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$ two conditional effects are added to the ${\tt \small (apply)_{\xi,\omega}}$ actions to ban positive effects that are inconsistent with an input observation (in ${\tt \small (apply)_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
+	\item[1-3] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$, one conditional effect is added to the ${\tt \small (insert\_pre)_{\xi,p}}$ actions to prevent the insertion of two preconditions that are {\em schematic mutex}.
+Likewise, two conditional effects are added to the ${\tt \small (insert\_eff)_{\xi,p}}$ actions, one to prevent the insertion of two positive effects that are {\em schematic mutex} and another one to prevent two mutex negative effects.
+	\item[4-5] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$, two conditional effects are added to the ${\tt \small (apply)_{\xi,\omega}}$ actions to prevent positive effects that are inconsistent with an input observation (in ${\tt \small (apply)_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
 \end{enumerate}
-In theory, conditional effects of the kind $4$ and $5$ are enough to guarantee that all the states traversed by a plan produced by the compilation are {\em consistent} with the input set of {\em schematic mutexes} $\Phi$ (of course, provided that the input initial state $s_0^o$ is a valid state).
-In practice we include also conditional effects of the kind $1$, $2$ and $3$ because they prune {\em invalid} action models at an earlier stage of the planning process (these effects extend the ${\tt \small insert}$ actions that always appear first in the solution plans).
 
-The goals of the classical planning problem output by the original compilation are extended with the $\neg invalid$ literal to validate that only states {\em consistent} with the state constraints defined in $\Phi$ are traversed by solution plans.
-Remarkably, the $\neg invalid$ literal allows us also to define ${\tt \small (apply)_{\xi,\omega}}$ actions more compactly than in the original compilation by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning}.
-Disjunctions are no longer required to code the possible preconditions of an action schema since they can now be encoded with conditional effects of this kind $\{pre\_\xi\_p\wedge \neg p(\omega)\}\rhd\{invalid\}$.
+In theory, conditional effects of the type $4$-$5$ are sufficient to guarantee that all the states traversed by a plan produced by the compilation are {\em consistent} with the input set of schematic mutexes $\Phi$ (obviously provided that the input initial state $s_0^o$ is a valid state).
+In practice we include also conditional effects of the type $1$-$3$ because they prune {\em invalid} action models at an earlier stage of the planning process (these effects extend the ${\tt \small insert}$ actions that always appear first in the solution plans).
+
+The goals of the planning task $P'$ generated by the original BLS compilation are extended with the $\neg invalid$ literal to validate that only states consistent with the state constraints defined in $\Phi$ are traversed by solution plans.
+Remarkably, the $\neg invalid$ literal allows us also to define ${\tt \small (apply)_{\xi,\omega}}$ actions more compactly than in the original compilation.
+Disjunctions are no longer required to code the possible preconditions of an action schema since they can now be encoded with conditional effects of the type $\{pre\_\xi\_p\wedge \neg p(\omega)\}\rhd\{invalid\}$.
 

### 2019-02-24 doc/ijcai-19-deductivelearning/deductivelearning.tex
New updates

@@
 \begin{enumerate}
-	\item[1-3] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$, one conditional effect is added to the ${\tt \small (insert\_pre)_{\xi,p}}$ actions to prevent the insertion of two preconditions that are {\em schematic mutex}.
-Likewise, two conditional effects are added to the ${\tt \small (insert\_eff)_{\xi,p}}$ actions, one to prevent the insertion of two positive effects that are {\em schematic mutex} and another one to prevent two mutex negative effects.
+	\item[1-3] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$, one conditional effect is added to the ${\tt \small (insert\_pre)_{\xi,p}}$ actions to prevent the insertion of two preconditions that are schematic mutex.
+Likewise, two conditional effects are added to the ${\tt \small (insert\_eff)_{\xi,p}}$ actions, one to prevent the insertion of two positive effects that are schematic mutex and another one to prevent two mutex negative effects.
 	\item[4-5] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$, two conditional effects are added to the ${\tt \small (apply)_{\xi,\omega}}$ actions to prevent positive effects that are inconsistent with an input observation (in ${\tt \small (apply)_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
@@
 
-The goals of the planning task $P'$ generated by the original BLS compilation are extended with the $\neg invalid$ literal to validate that only states consistent with the state constraints defined in $\Phi$ are traversed by solution plans.
+%The goals of the planning task $P'$ generated by the original BLS compilation are extended with the $\neg invalid$ literal to validate that only states consistent with the state constraints defined in $\Phi$ are traversed by solution plans.
 Remarkably, the $\neg invalid$ literal allows us also to define ${\tt \small (apply)_{\xi,\omega}}$ actions more compactly than in the original compilation.
@@
 \begin{mylemma}
-Soundness. Any classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ produces a model $\mathcal{M'}$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
+Soundness.
+Any classical plan $\pi$ that solves $P'$ (planning task that results from the compilation) produces a model $\mathcal{M'}$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
 \end{mylemma}
@@
 \begin{small}
-According to the $P_{\Lambda}$ compilation, once a given precondition or effect is inserted into the domain model $\mathcal{M}$ it cannot be removed back.
+According to the $P'$ compilation, once a given precondition or effect is inserted into the domain model $\mathcal{M}$ it cannot be undone.
 In addition, once an action model is applied it cannot be modified.
-In the compiled planning problem $P_{\Lambda}$, only ${\tt \small (apply)_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
+In the compiled planning task $P'$, only ${\tt \small (apply)_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
 This means that a state consistent with an observation $s_n^o$ can only be achieved executing an applicable sequence of ${\tt \small (apply)_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_i$, s.t.
@@
 \begin{mylemma}
-Completeness. Any model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task can be computed with a classical plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$.
+Completeness. Any model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task can be computed with a classical plan $\pi$ that solves $P'$.
 \end{mylemma}
@@
 By definition ${\mathcal I}_{\xi,\Psi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
-In addition the $P_{\Lambda}$ compilation does not discard any possible domain model $\mathcal{M}'$ definable within ${\mathcal I}_{\xi,\Psi}$ that satisfies the mutexes in $\Phi$.
-This means that, for every model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi_{\Lambda}$ that solves $P_{\Lambda}$ by selecting the appropriate ${\tt \small (insert\_pre)_{\xi,e}}$ and ${\tt \small (insert\_eff)_{\xi,e}}$ actions for {\em programming} the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding ${\tt \small (apply)_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
+In addition the $P'$ compilation does not discard any model $\mathcal{M}'$ definable within ${\mathcal I}_{\xi,\Psi}$ that satisfies the mutexes in $\Phi$.
+This means that, for every model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi$ that solves $P'$ by selecting the appropriate ${\tt \small (insert\_pre)_{\xi,e}}$ and ${\tt \small (insert\_eff)_{\xi,e}}$ actions for programming the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding ${\tt \small (apply)_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
 \end{small}
@@
 
-The size of the classical planning task $P_{\Lambda}$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$, that shape the propositional state variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
+The size of $P'$ depends on the arity of the predicates in $\Psi$, that shape variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
 The larger these arities, the larger $|{\mathcal I}_{\xi,\Psi}|$.
-The size of the ${\mathcal I}_{\xi,\Psi}$ set is the term that dominates the compilation size because it defines the $pre\_\xi\_e/eff\_\xi\_e$ fluents, the corresponding set of ${\tt \small insert}$ actions, and the number of conditional effects in the ${\tt \small (apply)_{\xi,\omega}}$ actions.
-Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$ which significantly reduces $|{\mathcal I}_{\xi,\Psi}|$ and hence, the size of the classical planning task output by the compilation.
+The size of ${\mathcal I}_{\xi,\Psi}$ is the most dominant factor of the compilation because it defines the $pre\_\xi\_e/eff\_\xi\_e$ fluents, the corresponding set of ${\tt \small insert}$ actions, and the number of conditional effects in the ${\tt \small (apply)_{\xi,\omega}}$ actions.
+Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$, which will significantly reduce $|{\mathcal I}_{\xi,\Psi}|$ and hence the size of $P'$ output by the compilation.
 
-Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
-In more detail, all $\{pre\_\xi\_e, eff\_\xi\_e\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ fluents are false at the initial state of our $P_\Lambda$ compilation so classical planners tend to solve $P_\Lambda$ with plans that require a smaller number of {\em insert} actions.
+Classical planners tend to prefer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of preconditions/effects.
+In more detail, all $\{pre\_\xi\_e, eff\_\xi\_e\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ fluents are false at the initial state of our $P'$ compilation so classical planners tend to solve $P'$ with plans that require a smaller number of ${\tt{\small insert}}$ actions.
 
-This bias could be eliminated defining a cost function for the actions in $P_\Lambda$ (e.g.
+This bias can be eliminated defining a cost function for the actions in $P'$ (e.g.
 ${\tt \small insert}$ actions have {\em zero cost} while ${\tt \small (apply)_{\xi,\omega}}$ actions have a {\em positive constant cost}).
-In practice we use a different approach to disregard the cost of {\em insert} actions since classical planners are not proficient at optimizing {\em plan cost} when there are zero-cost actions.
+In practice we use a different approach to disregard the cost of ${\tt \small insert}$ actions since classical planners are not proficient at optimizing plan cost with zero-cost actions.
 Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} that can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
 Further, the actions for inserting action effects are also applied in another single planning step.
-The plan horizon for programming any action model is then always bound to 2, which significantly reduces the planning horizon.
-The SAT-based planning approach is also convenient because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect the planning performance.
+The plan horizon for programming any action model is then always bounded to 2.
+The SAT-based planning approach is also convenient for its ability to deal with planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect the planning performance.
 
@@
 \subsection{Observability versus Knowledge}
-In our first experiment we seek to answer the question of whether knowledge can substitute observations. To that end, we evaluate the following 4 settings:
+In our first experiment we seek to answer the question of whether domain knowledge can substitute observations. To this end, we evaluate the following 4 settings:
 
@@
 	\item \textbf{Neither observability nor knowledge:} This is the baseline setting where the input sample is reduced to the minimum and only the initial and final states are known ($\sigma = 0$).
-	\item \textbf{Only knowledge:} Here we add domain-specific knowledge encoded as schematic mutexes to the baseline scenario.
-	\item \textbf{Only observability:} In this one, instead of knowledge we use a more complete input example where part of the intermediate states is known ($\sigma = 0.2$).
-	\item \textbf{Both observability and knowledge:} In the last setting we use both more complete input examples and schematic mutexes.
+	\item \textbf{Only knowledge:} Here we add domain knowledge encoded as schematic mutexes to the baseline scenario.
+	\item \textbf{Only observability:} In this one, we use a more complete observation where part of the intermediate states is known ($\sigma = 0.2$).
+	\item \textbf{Both observability and knowledge:} In the last setting we use we use both more complete input examples and schematic mutexes.
 \end{itemize}
@@
 The table also reports the number of schematic mutexes ($|\Phi|$) used for each domain.
-Comparing the settings wtih only knowledge and only observability, it is clear that having more complete training examples is preferable.
+Comparing the settings wtih only knowledge and only observability (second and third columns), it is clear that having more complete training examples is preferable.
 In fact, the improvement of using domain knowledge is marginal with respect to the baseline case.
@@
 
-\section{Related work}
-\label{sec:related}
-In {\em Inductive Logic Programming} it is common to make the hypothesis be consistent with the {\em background knowledge}, that is some form {\em deductive knowledge} apart from the examples~\cite{muggleton1994inductive}.
-
-{\em State-invariants} have also been previously used to improve the automatic construction of HTN planning model~\cite{lotinac2016constructing}.
+\section{Conclusions}
+\label{sec:conclusions}
 
@@
 
-\section{Conclusions}
-\label{sec:conclusions}
 In some contexts it is however reasonable to assume that the action model is not learned from scratch, e.g.

### 2019-02-24 doc/ijcai-19-deductivelearning/deductivelearning.tex
Unified nomenclature for section 5

@@
 \begin{enumerate}
-	\item[1-3] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$, one conditional effect is added to the ${\tt \small (insert\_pre)_{\xi,p}}$ actions to prevent the insertion of two preconditions that are schematic mutex.
+	\item[1-3] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to ${\mathcal I}_{\xi,\Psi}$, one conditional effect is added to the ${\tt \small (insert\_pre)_{\xi,p}}$ actions to prevent the insertion of two preconditions that are schematic mutex.
 Likewise, two conditional effects are added to the ${\tt \small (insert\_eff)_{\xi,p}}$ actions, one to prevent the insertion of two positive effects that are schematic mutex and another one to prevent two mutex negative effects.
-	\item[4-5] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to $\in{\mathcal I}_{\xi,\Psi}$, two conditional effects are added to the ${\tt \small (apply)_{\xi,\omega}}$ actions to prevent positive effects that are inconsistent with an input observation (in ${\tt \small (apply)_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
+	\item[4-5] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to ${\mathcal I}_{\xi,\Psi}$, two conditional effects are added to the ${\tt \small (apply)_{\xi,\omega}}$ actions to prevent positive effects that are inconsistent with an input observation (in ${\tt \small (apply)_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
 \end{enumerate}
@@
 In the compiled planning task $P'$, only ${\tt \small (apply)_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
-This means that a state consistent with an observation $s_n^o$ can only be achieved executing an applicable sequence of ${\tt \small (apply)_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_i$, s.t.
-$0\leq i\leq n$, is consistent with the input state observations and {\em state-invariants}.
+This means that a state consistent with an observation $s_m^o$ can only be achieved executing an applicable sequence of ${\tt \small (apply)_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_j$, s.t.
+$0 < j\leq m$, is consistent with the input state observations and {\em state-invariants}.
 This is exactly the definition of the solution condition for model $\mathcal{M}'$ to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
@@
 \label{sec:evaluation}
-This section evaluates the performance of our approach for learning \strips\ action models with different amounts of available input knowledge.
+This section evaluates the improvement of using domain-specific knowledge for learning action models.
 
@@
 The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-For each domain we generated 10 trajectories of length 10 via random walks to be used as training examples through all the experiments.
+For each domain we generated 10 learning examples of length 10 via random walks.
 We also introduce a new parameter, the {\em degree of observability} $\sigma$, which indicates de probability of observing a literal in an intermediate state.
-This parameter is used to build training examples with varying degrees of incompleteness from the generated trajectories.
+This parameter is used to build observations with varying degrees of incompleteness.
 All experiments are run on an Intel Core i5 3.10 GHz x 4 with 16 GB of RAM.
@@
 
-For the sake of reproducibility, the compilation source code, evaluation scripts, used benchmarks and input {\em state-invariants} are fully available at the repository {\em https://github.com/anonsub/}.
+For the sake of reproducibility, the compilation source code, evaluation scripts, used benchmarks and input {\em state-invariants} are fully available at the repository {\em https://github.com/anonsub/oneshot-learning}.
 
@@
 \begin{itemize}
-	\item \textbf{Neither observability nor knowledge:} This is the baseline setting where the input sample is reduced to the minimum and only the initial and final states are known ($\sigma = 0$).
-	\item \textbf{Only knowledge:} Here we add domain knowledge encoded as schematic mutexes to the baseline scenario.
-	\item \textbf{Only observability:} In this one, we use a more complete observation where part of the intermediate states is known ($\sigma = 0.2$).
-	\item \textbf{Both observability and knowledge:} In the last setting we use we use both more complete input examples and schematic mutexes.
+	\item \textbf{Minimal observation:} This is the baseline setting where the learning example is reduced to the minimum and only the initial state and part final state are known ($\sigma = 0$ w/o $\Phi$).
+	\item \textbf{Only $\Phi$:} Here we add domain knowledge encoded as schematic mutexes to the baseline scenario ($\sigma = 0$ with $\Phi$).
+	\item \textbf{Only observability:} In this one, we use a more complete observation where intermediate states are partially observed ($\sigma = 0.2$ w/o $\Phi$).
+	\item \textbf{Both observability and $\Phi$:} In the last setting we use both more complete observations and schematic mutexes ($\sigma = 0.2$ with $\Phi$).
 \end{itemize}
@@
 				\begin{tabular}{l|c|c|c|c|c|c|c|c|c|}
-					& & \multicolumn{2}{|c|}{$\sigma = 0$} & \multicolumn{2}{|c|}{$\sigma = 0$ with $\Phi$} & \multicolumn{2}{|c|}{$\sigma = 0.2$} & \multicolumn{2}{|c|}{$\sigma = 0.2$ with $\Phi$}\\ \cline{3-10}			
+					& & \multicolumn{2}{|c|}{$\sigma = 0$ w/o $\Phi$} & \multicolumn{2}{|c|}{$\sigma = 0$ with $\Phi$} & \multicolumn{2}{|c|}{$\sigma = 0.2$ w/o $\Phi$} & \multicolumn{2}{|c|}{$\sigma = 0.2$ with $\Phi$}\\ \cline{3-10}
 					& \multicolumn{1}{|c|}{$|\Phi|$} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} &  \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} \\
 					\hline
-					blocks & 9 & 0.52 & 0.38 & 0.53 & 0.21 & 0.66 & 0.56 & 0.77 & 0.68 \\
-					driverlog & 8 & 0.49 & 0.33 & 0.33 & 0.31 & 0.54 & 0.38 & 0.70 & 0.53 \\
-					ferry & 2 & 0.50 & 0.40 & 0.57 & 0.41 & 0.59 & 0.64 & 0.59 & 0.70 \\
-					floor-tile & 7 & 0.30 & 0.40 & 0.58 & 0.46 & 0.68 & 0.46 & 0.75 & 0.48 \\
-					grid & 3 & 0.47 & 0.40 & 0.47 & 0.37 & 0.43 & 0.34 & 0.43 & 0.32 \\
-					gripper & 5 & 0.77 & 0.56 & 0.77 & 0.54 & 0.85 & 0.74 & 0.96 & 0.83 \\
-					hanoi & 3 & 0.84 & 0.76 & 0.76 & 0.75 & 0.96 & 0.75 & 0.97 & 0.79 \\
-					n-puzzle & 3 & 0.94 & 0.86 & 0.93 & 0.86 & 0.99 & 0.87 & 1.00 & 0.87 \\
-					parking & 8 & 0.48 & 0.37 & 0.60 & 0.40 & 0.58 & 0.45 & 0.67 & 0.49 \\
-					transport & 4 & 0.45 & 0.45 & 0.53 & 0.46 & 0.99 & 0.51 & 0.94 & 0.79 \\
-					zeno-travel & 4 & 0.72 & 0.39 & 0.75 & 0.40 & 0.80 & 0.42 & 0.93 & 0.55 \\
+					blocks & 9 & 0.51 & 0.36 & 0.54 & 0.23 & 0.59 & 0.49 & 0.79 & 0.70 \\
+					driverlog & 8 & 0.48 & 0.36 & 0.34 & 0.34 & 0.41 & 0.31 & 0.69 & 0.49 \\
+					ferry & 2 & 0.47 & 0.39 & 0.58 & 0.43 & 0.50 & 0.51 & 0.61 & 0.72 \\
+					floor-tile & 7 & 0.39 & 0.39 & 0.48 & 0.45 & 0.64 & 0.48 & 0.74 & 0.52 \\
+					grid & 3 & 0.40 & 0.31 & 0.42 & 0.31 & 0.43 & 0.31 & 0.50 & 0.37 \\
+					gripper & 5 & 0.76 & 0.50 & 0.77 & 0.51 & 0.85 & 0.74 & 0.92 & 0.81 \\
+					hanoi & 3 & 0.88 & 0.71 & 0.73 & 0.81 & 0.94 & 0.78 & 1.00 & 0.81 \\
+					n-puzzle & 3 & 0.94 & 0.76 & 0.95 & 0.81 & 0.97 & 0.86 & 0.97 & 0.89 \\
+					parking & 8 & 0.54 & 0.41 & 0.60 & 0.40 & 0.51 & 0.40 & 0.51 & 0.41 \\
+					transport & 4 & 0.45 & 0.45 & 0.53 & 0.46 & 0.49 & 0.37 & 0.94 & 0.75 \\
+					zeno-travel & 4 & 0.73 & 0.36 & 0.80 & 0.36 & 0.79 & 0.36 & 0.89 & 0.50 \\
 					\hline
-					&  & 0.59 & 0.48 & 0.62 & 0.47 & 0.73 & 0.56 & 0.79 & 0.64 \\
+					&  & 0.60 & 0.45 & 0.61 & 0.46 & 0.65 & 0.51 & 0.78 & 0.63
 
@@
 The table also reports the number of schematic mutexes ($|\Phi|$) used for each domain.
-Comparing the settings wtih only knowledge and only observability (second and third columns), it is clear that having more complete training examples is preferable.
+Comparing the settings wtih only domain knowledge and only observability (second and third columns), it is clear that having more complete learning examples is preferable.
 In fact, the improvement of using domain knowledge is marginal with respect to the baseline case.
-This is not the case for the last setting, where the use of domain knowledge shows a significant improvement in the quality of the learned models when compared to the setting with only observability.
+This is not the same for the last setting, where the use of domain knowledge shows a significant improvement in the quality of the learned models when compared to the setting with only observability.
 
@@
 The values plotted in these figures are averages across all the domains seen in the previous experiment.
-The results show that using domain knowledge significantly improves the learned models no matter how complete the training examples are.
-Another interesting aspect is that domain knowledge is able to enrich observations in the range of 30\% observability to the level of fully observable trajectories, which means that domain knowledge can make up for a lack of completeness in the training examples.
+The results show that using domain knowledge significantly improves the learned models no matter how complete the learning examples are.
+Another interesting aspect is that domain knowledge is able to enrich observations in the range of 30\% observability to the level of fully observable examples, which means that domain knowledge can make up for a lack of completeness in the learning examples.
 

### 2019-02-25 doc/ijcai-19-deductivelearning/deductivelearning.tex
Final version #1

@@
 \label{sec:evaluation}
-This section evaluates the improvement of using domain-specific knowledge for learning action models.
+This section evaluates the improvement when using domain-specific knowledge for learning action models.
 
@@
 \subsection{Observability versus Knowledge}
-In our first experiment we seek to answer the question of whether domain knowledge can substitute observations. To this end, we evaluate the following 4 settings:
+In our first experiment, we seek to answer the question as to whether the plan observation (learning example) of $\mathcal{O}$ is replaceable by the domain knowledge encoded in $\Phi$.
+To this end, we evaluate the following 4 settings:
 
-\begin{itemize}
-	\item \textbf{Minimal observation:} This is the baseline setting where the learning example is reduced to the minimum and only the initial state and part final state are known ($\sigma = 0$ w/o $\Phi$).
-	\item \textbf{Only $\Phi$:} Here we add domain knowledge encoded as schematic mutexes to the baseline scenario ($\sigma = 0$ with $\Phi$).
-	\item \textbf{Only observability:} In this one, we use a more complete observation where intermediate states are partially observed ($\sigma = 0.2$ w/o $\Phi$).
-	\item \textbf{Both observability and $\Phi$:} In the last setting we use both more complete observations and schematic mutexes ($\sigma = 0.2$ with $\Phi$).
-\end{itemize}
+\begin{enumerate}
+	\item \textbf{Minimal observability:} This is the baseline setting where we use the minimal expression of the learning example; i.e., a fully observed initial state and a partially observed final state.
+This setting is labeled as $\sigma = 0$ with no $\Phi$.
+	\item \textbf{Only knowledge ($\Phi$):} In this setting we add domain knowledge encoded as schematic mutexes to the baseline scenario ($\sigma = 0$ with $\Phi$).
+	\item \textbf{Only observability:} We use a more complete observation where intermediate states are partially observed ($\sigma = 0.2$ with no $\Phi$).
+	\item \textbf{Both observability and $\Phi$:} We use both a more complete observation and schematic mutexes ($\sigma = 0.2$ with $\Phi$).
+\end{enumerate}
 
@@
 			\resizebox{0.45\textwidth}{!}{	
-				\begin{tabular}{l|c|c|c|c|c|c|c|c|c|}
-					& & \multicolumn{2}{|c|}{$\sigma = 0$ w/o $\Phi$} & \multicolumn{2}{|c|}{$\sigma = 0$ with $\Phi$} & \multicolumn{2}{|c|}{$\sigma = 0.2$ w/o $\Phi$} & \multicolumn{2}{|c|}{$\sigma = 0.2$ with $\Phi$}\\ \cline{3-10}
-					& \multicolumn{1}{|c|}{$|\Phi|$} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} &  \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} \\
+				\begin{tabular}{l|c|c|c||c|c||c|c||c|c|}
+					& & \multicolumn{2}{|c||}{setting 1} & \multicolumn{2}{|c||}{setting 2} & \multicolumn{2}{|c||}{setting 3} & \multicolumn{2}{|c|}{setting 4}\\ %\cline{3-10}
+					& & \multicolumn{2}{|c||}{$\sigma = 0$ w/o $\Phi$} & \multicolumn{2}{|c||}{$\sigma = 0$ with $\Phi$} & \multicolumn{2}{|c||}{$\sigma = 0.2$ w/o $\Phi$} & \multicolumn{2}{|c|}{$\sigma = 0.2$ with $\Phi$}\\ \cline{3-10}
+					& \multicolumn{1}{|c|}{$|\Phi|$} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c||}{R} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c||}{R} & \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c||}{R} &  \multicolumn{1}{|c|}{P} & \multicolumn{1}{|c|}{R} \\
 					\hline
@@
 
-Table \ref{tab:observability_vs_knowledge} compiles the average precision (P) and recall (R) for each domain in the different settings tested.
+Table \ref{tab:observability_vs_knowledge} shows the average precision (P) and recall (R) for each domain in the four tested settings.
 The table also reports the number of schematic mutexes ($|\Phi|$) used for each domain.
-Comparing the settings wtih only domain knowledge and only observability (second and third columns), it is clear that having more complete learning examples is preferable.
-In fact, the improvement of using domain knowledge is marginal with respect to the baseline case.
-This is not the same for the last setting, where the use of domain knowledge shows a significant improvement in the quality of the learned models when compared to the setting with only observability.
+Comparing the settings \emph{only domain knowledge} (setting 2) with \emph{only observability} (setting 3), we can see that slightly better results are obtained with the latter, meaning that observability is slightly more informative than domain knowledge.
+On the other hand, the gain of using $\Phi$ under minimal observability (setting 1 compared to setting 2) is rather marginal.
+While these results might indicate a general preference for observations over knowledge, when comparing setting 3 with setting 4, we can observe a significant improvement in the quality of the learned models.
+This indicates that the payoff of using $\Phi$ is noticeable when the learning example has a certain degree of observability.
 
 \subsection{Using knowledge to counter incompleteness}
-In the previous experiment we have stablished that domain knowledge is no substitute to observations, and that, given a minimum of observability, domain knowledge is able to enrich both the observations and learning process thus procuring better learned models.
-In this next experiment we are going to measure the improvement provided by domain knowledge at increasing degrees of observability.
+
+The previous experiment reveals that observations are not totally replaceable by domain knowledge; but also shows that given a minimum degree of observability, using $\Phi$ enriches both the observations and the learning process and better models are learnable.
+In this next experiment we measure the improvement provided by $\Phi$ at increasing degrees of observability of the learning example.
 
@@
 
-Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare the precision and recall of the learned models in the settings with and without domain knowledge.
-The values plotted in these figures are averages across all the domains seen in the previous experiment.
-The results show that using domain knowledge significantly improves the learned models no matter how complete the learning examples are.
-Another interesting aspect is that domain knowledge is able to enrich observations in the range of 30\% observability to the level of fully observable examples, which means that domain knowledge can make up for a lack of completeness in the learning examples.
+Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare the Precision and Recall of the learned models with and without domain knowledge.
+The values plotted in these figures are averages across all the domains presented in Table \ref{tab:observability_vs_knowledge}.
+The results show that using $\Phi$ significantly improves the learned models no matter how complete the learning examples are.
+An interesting and revealing aspect from the figures is that the quality of the action models learned with 30\%-observable learning examples and $\Phi$ is comparable to the quality obtained with a 100\%-observable example.
+Hence, domain knowledge can make up for the lack of completeness in the learning examples.
 
@@
 
-Our learning setting is related to the classical planning formulation where no action model is given~\cite{SternJ17}.
-This planning setting can can be seen as an scenario when the action model is {\em learned} from a single example that contains only two state observations: the initial state and the goals.
+We present an approach to learn action models that builds upon a former compilation-to-planning learning system~\cite{aineto2018learning}.
+Our proposal studies the gains of using domain-specific knowledge at the cost of limiting the learning examples to the minimal observability possible.
+Introducing domain knowledge encoded as schematic mutexes allows to narrow down the search space of the learning task and improve overall the performance of the learning system to the point that it offsets the lack of learning examples.
 
-In some contexts it is however reasonable to assume that the action model is not learned from scratch, e.g.
-because some parts of the action model are known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
-Our compilation is also flexible to this particular learning scenario.
-The known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ to true in the initial state.
-Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, making the classical planning task $P_{\Lambda}$ easier to be solved.
-For example, suppose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state is extended with literals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the associated actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
+In a theoretical work that analyzes the relation between the number of observed trajectory plans and the guarantee for a learned action model to achieve the goal~\cite{SternJ17}, authors conclude that the number of trajectories needed scales gracefully and the guarantee grows linearly with the number of predicates and quasi-linearly with the number of actions.
+This evidences that learning accurate models is heavily dependent on the number and quality (observability) of the learning examples.
+In this sense, our proposal comes to alleviate this dependency by relying on easily deducible domain knowledge.
+It is not only capable of learning from a single non-fully observable learning example but also proves that learning from a 30\%-observable example with domain-specific knowledge is comparable to learning from a complete plan observation.
 

### 2019-02-25 doc/ijcai-19-deductivelearning/deductivelearning.tex
Submission number

@@
 %{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
+
+\author{\#2922}
 
@@
 
-Table \ref{tab:observability_vs_knowledge} shows the average precision (P) and recall (R) for each domain in the four tested settings.
+Table \ref{tab:observability_vs_knowledge} shows the average precision (P) and recall (R) for each domain across the 10 learning examples in the four tested settings.
 The table also reports the number of schematic mutexes ($|\Phi|$) used for each domain.
@@
 We present an approach to learn action models that builds upon a former compilation-to-planning learning system~\cite{aineto2018learning}.
-Our proposal studies the gains of using domain-specific knowledge at the cost of limiting the learning examples to the minimal observability possible.
+Our proposal studies the gains of using domain-specific knowledge when the availability of learning examples is very limited.
 Introducing domain knowledge encoded as schematic mutexes allows to narrow down the search space of the learning task and improve overall the performance of the learning system to the point that it offsets the lack of learning examples.

### 2019-02-26 doc/ijcai-19-deductivelearning/deductivelearning.tex
averages
submitted version

@@
 The minimal input observability required by our approach is a single example composed of a full initial state and a partial goal state.
-We will show that exploiting specific domain knowledge enable to constrain the space of possible action models as well as to complete partial observations, both of which turn out helpful to learn good-quality action models.
+We will show that exploiting specific domain knowledge enables to constrain the space of possible action models as well as to complete partial observations, both of which turn out helpful to learn good-quality action models.
  \end{abstract}
@@
 A model explains an observation if a plan containing the observed actions is computable with the model and the states induced by this plan also include the possibly partially observed states.
-The limitation of posing model validation as an optimization task over a testing set of observations is that it neither guarantees completeness (the model may not explain all the observations) nor correctness (the states induced by the execution of the plan generated with the model may contain contradictory information).
+The limitation of posing model learning and validation as optimization tasks over a set of observations is that it neither guarantees completeness (the model may not explain all the observations) nor correctness (the states induced by the execution of the plan generated with the model may contain contradictory information).
 
@@
 Particularly, we aim to stress the extreme case of having a single observation sample and answer the question to whether the lack of training samples can be overcome with the supply of domain knowledge.
-The question is motivated by (a) the assumption that obtaining enough training observations is often difficult and costly, if not impossible in some domains~\cite{Zhuo15}; (b) the fact that although the physics of the real-world domain being modeled are unknown, the user may know certain pieces of knowledge about the domain; and (c) the desire for correct action models that are usable beyond their applicability to a set of testing observations.
+The question is motivated by (a) the assumption that obtaining enough training observations is often difficult and costly, if not impossible in some domains~\cite{Zhuo15}; (b) the fact that although the physics of the real-world domain being modeled are unknown, the user may know certain pieces of knowledge about the domain; and (c) the desire for correct action models that are usable beyond their fitness to a set of testing observations.
 To this end, we opted for checking our hypothesis in the framework proposed in~\cite{aineto2018learning} since this planning-based satisfiability approach allows us to configure additional constraints in the compilation scheme, it is able to work under a minimal set of observations and uses an off-the-shelf planner\footnote{We thank authors for providing us with the source files of their learning system.}.
@@
 on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of the action model $\xi$.
+Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi$.
 
@@
 {\tt{\footnotesize(insert\_eff\_stack\_clear\_v1),\\
-(insert\_eff\_stack\_clear\_v2)}}, where the second action denotes a positive effect and the third one a negative effect both to be inserted in the model of {\tt{\small stack}}; and the second set of actions of the plan that solves $P'$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the {\tt {\small apply}} actions must be validated with the observations of plan executions.
+(insert\_eff\_stack\_clear\_v2)}}, where the second action denotes a positive effect and the third one a negative effect both to be inserted in the model of {\tt{\small stack}}; and the second set of actions of the plan that solves $P'$ will be like {\tt{\small (apply\_unstack blockB blockA),(validate\_1),(apply\_putdown blockB),(validate\_2)}}, where the {\tt {\small validate}} actions denote the points at which the states generated through the {\tt {\small apply}} actions must be validated with the observations of plan executions.
 
@@
 
-We analyze here the search space of a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$; i.e., the space of \strips\ action models.
+We analyze here the solution space of a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$; i.e., the space of \strips\ action models.
 In principle, for a given action model $\xi$, any element of ${\mathcal I}_{\xi,\Psi}$ can potentially appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
@@
 The solution $\mathcal{M}'$ must be consistent with these \emph{semantic constraints} derived from  the learning samples $\mathcal{O}$, which in our case is a single plan observation.
-Specifically, the states induced by the plan computable with $M'$ must comprise the observed states of the sample.
+Specifically, the states induced by the plan computable with $\mathcal{M}'$ must comprise the observed states of the sample, which further constrains the space of possible action models.
 \end{enumerate}
@@
 \end{figure}
-
-Additionally, observation samples further constrains the space of possible action models.
 
@@
 For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible state.
-Figure~\ref{fig:strongest-invariant} shows four clauses that define schematic mutexes for the {\em blocksworld} domain.
+Figure~\ref{fig:strongest-invariant} shows an example of four clauses that define schematic mutexes for the {\em blocksworld} domain.
 
@@
 \subsection{Completing partially observed states with {\em schematic mutexes}}
-%Our sampling space follows the {\em open world} assumption, i.e.
-what is not observed is considered unknown.
-Here we describe a pre-processing mechanism to add new knowledge that completes the states $\tup{s_1^o \ldots, s_m^o}$ that are partially observed in a $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task using an input set of {\em schematic mutex} $\Phi$.
 
@@
 
-Let $\Omega$ be the set of objects that appear in $F$ as the values of the arguments of the predicates $\Psi$, and $\Phi=\tup{p,q}$ a schematic mutex.
-There exist many possible instantiations of $\Phi$ of the type $\tup{p(\omega),q(\omega')}$ with objects of $\Omega$, where $\omega\subseteq\Omega^{args(p)}$ and $\omega'\subseteq\Omega^{args(q)}$.
+Let $\Omega$ be the set of objects that appear in $F$ as the values of the arguments of the predicates $\Psi$, and $\phi=\tup{p,q}$ a schematic mutex.
+There exist many possible instantiations of $\phi$ of the type $\tup{p(\omega),q(\omega')}$ with objects of $\Omega$, where $\omega\subseteq\Omega^{|args(p)|}$ and $\omega'\subseteq\Omega^{|args(q)|}$.
 Let us now assume that the instantiation $p(\omega) \in s_j^o$, {\small $(1\leq j\leq m)$}, being $s_j^o$ a partially observed state of $\mathcal{O}$.
-Then, two situations may occur: (a) $q(\omega') \in s_j^o$, in which case the expression $\neg p(\omega) \vee \neg q(\omega')$ holds in $s_j^o$; or (b) $q(\omega') \notin s_j^o$, in which case the literal $\neg q(\omega')$ has not been observed in $s_j^o$ and so we can safely complete the state with $\neg q(\omega')$ (the same applies inversely, when $q(\omega') \in s_j^o$ but $\neg p(\omega) \notin s_j^o$).
+Then, two situations may occur: (a) $\neg q(\omega') \in s_j^o$, in which case the expression $\neg p(\omega) \vee \neg q(\omega')$ holds in $s_j^o$; or (b) $\neg q(\omega') \notin s_j^o$, in which case the literal has not been observed in $s_j^o$ and so we can safely complete the state with $\neg q(\omega')$ (the same applies inversely, when $q(\omega') \in s_j^o$ but $\neg p(\omega) \notin s_j^o$).
 In other words, if we find that one component of a schematic mutex is positively observed in a state and the other component is not observable in such state, we can complete the state with the missing negative literal.
@@
 In the compiled planning task $P'$, only ${\tt \small (apply)_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
-This means that a state consistent with an observation $s_m^o$ can only be achieved executing an applicable sequence of ${\tt \small (apply)_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_j$, s.t.
-$0 < j\leq m$, is consistent with the input state observations and {\em state-invariants}.
+This means that a state consistent with an observation $s_m^o$ can only be achieved executing an applicable sequence of ${\tt \small (apply)_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_j$ $(0 < j\leq m)$, is consistent with the input state observations and {\em state-invariants}.
 This is exactly the definition of the solution condition for model $\mathcal{M}'$ to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
@@
 
-Classical planners tend to prefer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of preconditions/effects.
+Classical planners tend to prefer shorter solution plans, so our compilation (as well as the BLS) may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of preconditions/effects.
 In more detail, all $\{pre\_\xi\_e, eff\_\xi\_e\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ fluents are false at the initial state of our $P'$ compilation so classical planners tend to solve $P'$ with plans that require a smaller number of ${\tt{\small insert}}$ actions.
@@
 The domains used in the evaluation are IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
-For each domain we generated 10 learning examples of length 10 via random walks.
+For each domain we generated 10 learning examples of length 10 via random walks and report average values (only a single example is considered at each learning episode).
 We also introduce a new parameter, the {\em degree of observability} $\sigma$, which indicates de probability of observing a literal in an intermediate state.
@@
 \subsection{Observability versus Knowledge}
-In our first experiment, we seek to answer the question as to whether the plan observation (learning example) of $\mathcal{O}$ is replaceable by the domain knowledge encoded in $\Phi$.
+In our first experiment, we seek to answer the question as to whether the plan observation (single learning example) of $\mathcal{O}$ is replaceable by the domain knowledge encoded in $\Phi$.
 To this end, we evaluate the following 4 settings:
@@
 \begin{enumerate}
-	\item \textbf{Minimal observability:} This is the baseline setting where we use the minimal expression of the learning example; i.e., a fully observed initial state and a partially observed final state.
+	\item \textbf{Minimal observability:} This is the baseline setting where we use the minimal expression of the single learning example; i.e., a fully observed initial state and a partially observed final state.
 This setting is labeled as $\sigma = 0$ with no $\Phi$.
 	\item \textbf{Only knowledge ($\Phi$):} In this setting we add domain knowledge encoded as schematic mutexes to the baseline scenario ($\sigma = 0$ with $\Phi$).
-	\item \textbf{Only observability:} We use a more complete observation where intermediate states are partially observed ($\sigma = 0.2$ with no $\Phi$).
+	\item \textbf{Only observability:} We use a more complete observation where intermediate states of the learning example are partially observed ($\sigma = 0.2$ with no $\Phi$).
 	\item \textbf{Both observability and $\Phi$:} We use both a more complete observation and schematic mutexes ($\sigma = 0.2$ with $\Phi$).
@@
 
-Table \ref{tab:observability_vs_knowledge} shows the average precision (P) and recall (R) for each domain across the 10 learning examples in the four tested settings.
+Table \ref{tab:observability_vs_knowledge} shows the average values of precision (P) and recall (R) for each domain in the four tested settings.
 The table also reports the number of schematic mutexes ($|\Phi|$) used for each domain.
-Comparing the settings \emph{only domain knowledge} (setting 2) with \emph{only observability} (setting 3), we can see that slightly better results are obtained with the latter, meaning that observability is slightly more informative than domain knowledge.
+Comparing the settings \emph{only domain knowledge} (setting 2) with \emph{only observability} (setting 3), we can see that slightly better results are obtained with the latter, meaning that observability is more informative than the used domain knowledge.
 On the other hand, the gain of using $\Phi$ under minimal observability (setting 1 compared to setting 2) is rather marginal.
@@
 Figures \ref{fig:comparison_precision} and \ref{fig:comparison_recall} compare the Precision and Recall of the learned models with and without domain knowledge.
-The values plotted in these figures are averages across all the domains presented in Table \ref{tab:observability_vs_knowledge}.
+The points plotted in these figures are average values across all the domains presented in Table \ref{tab:observability_vs_knowledge}.
 The results show that using $\Phi$ significantly improves the learned models no matter how complete the learning examples are.
@@
 We present an approach to learn action models that builds upon a former compilation-to-planning learning system~\cite{aineto2018learning}.
-Our proposal studies the gains of using domain-specific knowledge when the availability of learning examples is very limited.
+Our proposal studies the gains of using domain-specific knowledge when the availability (amount and observability) of learning examples is very limited.
 Introducing domain knowledge encoded as schematic mutexes allows to narrow down the search space of the learning task and improve overall the performance of the learning system to the point that it offsets the lack of learning examples.

### 2019-02-26 doc/icaps-19/camera-ready/model-recognition.tex
Fixed Miquel's email address
ICAPS19: fixed effects notation for open world assumption

--- 2019-02-08 doc/icaps-19/submitted/model-recognition.tex
@@
 {\scriptsize Melbourne, Victoria. Australia}\\
-{\scriptsize miquel.ramirez@nimelb.edu.au}}
+{\scriptsize miquel.ramirez@unimelb.edu.au}}
 
@@
 
-A {\em planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
+%A {\em planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
 An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\in\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\in\mathcal{L}(F)$.
@@
 And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
+
+A {\em planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
+An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$,  and {\em effects} $\eff(a)\in\mathcal{L}(F)$.
+The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
+Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$.
+And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\neg\eff(a))\cup\eff(a)\}$, with $\neg\eff(a) = \{\neg l : l \in \eff(a)\}$.
 
@@
 
+%An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
+Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+%\[
+%triggered(s,a_c)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
+%\]
+%
+%The result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$, where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
+
 An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
 Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+An action $a_c\in A$ is applicable in a state $s$ if and only if $\pre(a_c)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
 \[
-triggered(s,a_c)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
+triggered(s,a_c)=\bigcup\limits_{C\rhd E\in\cond(a_c),C\subseteq s} E. 
 \]
-
-The result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$, where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
+The result of applying $a_c$ in state $s$ follows the same definition of successor state, $\theta(s,a)$, but applied to the conditional effects in $triggered(s,a_c)$.
 

### 2019-02-27 doc/icaps-19/camera-ready/model-recognition.tex
ICAPS19: Minor corrections to section 4

@@
 
-Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters} $pars(\xi)$.
+Let $\Psi$ be the set of {\em predicates} that shape the propositional state variables $F$.
 The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is given by FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
@@
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is constrained by a set ${\mathcal C}$ that includes:
+Although any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, the space of possible \strips\ schemata is constrained by a set ${\mathcal C}$ that includes:
 
@@
 Given a model $\mathcal{M}$, the set $M^*$ of the {\bf closest consistent models} is the set of models $\mathcal{M}'$ that: (1) produce a trajectory $\tau(\pi,P)$ that reaches the goals in $P$ and is {\em consistent} with $\mathcal{O}(\tau)$ and (2) their {\em edit distance} to $\mathcal{M}$ is minimal;
-  \[\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
+  \[M^*=\underset{\forall \mathcal{M}' \rightarrow \mathcal{O}}{\arg\min}\ \delta(\mathcal{M},\mathcal{M}') \]
 \end{definition}
@@
 \end{align}
-Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because is independent of the input model $\mathcal{M}\in M$.
+Note that the term $P(\tau|\mathcal{M^*})$ is taken out of the maximization because it is independent of the input model $\mathcal{M}\in M$.
 
@@
 The largest term corresponds here to the shortest trajectory $\tau^*$ consistent with the observation (which now is assumed to be unique).
-Note that the more complete is the observation of the plan execution the more accurate is our estimate because it becomes more likely to assume that there is only one trajectory consistent with the input observation.
+Note that the more complete the observation of the plan execution is the more accurate our estimate is because it becomes more likely to assume that there is only one trajectory consistent with the input observation.
 
@@
 \end{align}
-where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than do not applying it.
+where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than not applying it.
 

### 2019-02-27 doc/icaps-19/camera-ready/model-recognition.tex
ICAPS19: added mention to removing actions in the compilation

@@
     \item actions for editing $\xi \in \mathcal{M}$ that follow the \emph{syntactic constraints} for well-defined \strips\ action schemata.
-Hence, $A'$ contains actions for inserting a new precondition $pre\_e\_\xi$ or a new positive effect $add\_e\_\xi$ in $\xi$ when neither $pre\_e\_\xi$, $del\_e\_\xi$ or $add\_e\_\xi$ exist in $\xi$, and actions for inserting a new negative effect when neither $del\_e\_\xi$ or $add\_e\_\xi$ appear in $\xi$ but $pre\_e\_\xi$ does.
+Hence, $A'$ contains actions for inserting and removing a precondition $pre\_e\_\xi$, a positive effect $add\_e\_\xi$ or a negative effect $del\_e\_\xi$ in $\xi$.
     \item actions for applying the new action schemata of the edited model $\mathcal{M'}$ and validating the observed states of $\mathcal{O(\tau)}$.

### 2019-02-27 doc/icaps-19/camera-ready/model-recognition.tex
ICAPS19: typos by reviewer#2

@@
 Even worse, the number of models $\mathcal{M}'$ that can be generated {\em transforming} a given classical planning model $\mathcal{M}$ explodes combinatorially with the number of state variables.
-Instead, our approach is to estimate of $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} referred to \strips\ planning models.
+Instead, our approach is to estimate $P(\mathcal{O}|\mathcal{M})$ using an {\em edit distance} defined for \strips\ planning models.
 {\em Edit distances} are similarity metrics, traditionally computed over {\em strings} or {\em graphs}, which have been proved successful for {\em pattern recognition}~\cite{MasekP80,Bunke97}.
@@
 
-$\delta^o(\mathcal{M},\mathcal{O})$ can also be defined through the edition that the observation $\mathcal{O}(\tau)$ requires to fit $\mathcal{M}$.
+$\delta^o(\mathcal{M},\mathcal{O})$ can also be defined through the editing that the observation $\mathcal{O}(\tau)$ requires to fit $\mathcal{M}$.
 This implies defining {\em edit operations} that modify the observation $\mathcal{O}(\tau)$ instead of the model $\mathcal{M}$~\cite{yang2007learning,SohrabiRU16}.
@@
 
-Now we we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ action schemata.
+Now we are ready to formulate an informative estimate of the $P(\mathcal{O}|\mathcal{M})$ likelihood for the particular case where models $\mathcal{M}$ are specified with \strips\ action schemata.
 
@@
 
-In other words, we are modeling the edition of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.
+In other words, we are modeling the editing of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.
 
@@
 Other planners, such as FastDownward were also tested but provided worse experimental results.
-We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned the maximum distance obtained.
+We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned the maximum distance.
 The compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em https://github.com/anonsub/model-recognition} so any experimental data reported in the paper can be reproduced.
@@
 
-An interesting research line related to {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}, in which model edition is used to conform the PDDL models of two agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed optimal plan computed with one of the two models.
+An interesting research line related to {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}, in which model editing is used to conform the PDDL models of two agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed optimal plan computed with one of the two models.
 {\em Model recognition}, however, conforms every input model $\mathcal{M}$ with another model $\mathcal{M}'$ that is not given as input but instead computed for every $\mathcal{M}$, and which is consistent with a partial observation of a plan execution.

### 2019-03-04 doc/lecture-notes-EBL/ebl.tex
first commit lecture notes on strips ebl

% This is samplepaper.tex, a sample chapter demonstrating the
% LLNCS macro package for Springer Computer Science proceedings;
% Version 2.20 of 2017/10/04
%
\documentclass[runningheads]{llncs}
%
\usepackage{graphicx}
% Used for displaying a sample figure. If possible, figure files should
% be included in EPS format.
%
% If you use the hyperref package, please uncomment the following line
% to display URLs in blue roman font according to Springer's eBook style:
% \renewcommand\UrlFont{\color{blue}\rmfamily}

\begin{document}
%
\title{Explanation-based learning of action models}
%
%\titlerunning{Abbreviated paper title}
% If the paper title is too long for the running head, you can set
% an abbreviated paper title here
%
\author{Diego Aineto\inst{1}\orcidID{} \and
Sergio Jiménez\inst{1}\orcidID{0000-0003-0561-4880} \and
Eva Onaindia\inst{1}\orcidID{0000-0001-6931-8293}}
%
\authorrunning{D. Aineto et al.}
% First names are abbreviated in the running head.
% If there are more than two authors, 'et al.' is used.
%
\institute{
{\small Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
{\small Universitat Polit\`ecnica de Val\`encia.}\\
{\small Camino de Vera s/n. 46022 Valencia, Spain}\\
{\small \{dieaigar,serjice,onaindia\}@dsic.upv.es}}
%
\maketitle              % typeset the header of the contribution
%
\begin{abstract}

\end{abstract}

\keywords{Learning action models \and Planning and Learning \and Classical Planning.}

\section{Introduction}

\section{Background}

\section{Explanation-based learning of action models}

\end{document}

### 2019-03-04 doc/lecture-notes-EBL/ebl.tex
creating paper sections

@@
 
+%%% Defintions For this paper
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{multirow}
+\usepackage{tikz}
+\usetikzlibrary{arrows,automata}
+\usepackage{comment}
+
+\newcommand{\tup}[1]{{\langle #1 \rangle}}
+\newcommand{\pre}{\mathsf{pre}}     % precondition
+\newcommand{\del}{\mathsf{del}}     % effect
+\newcommand{\add}{\mathsf{add}}     % effect
+\newcommand{\eff}{\mathsf{eff}}     % effect
+\newcommand{\cond}{\mathsf{cond}}   % conditional effect
+\newcommand{\true}{\mathsf{true}}   % true
+\newcommand{\false}{\mathsf{false}} % false
+\newcommand{\PE}{\mathrm{PE}}     % precondition
+\newcommand{\strips}{\textsc{Strips}}     % precondition
+%%%
+
 \begin{document}
 %
-\title{Explanation-based learning of action models}
+\title{Explanation-based learning of \strips\ action models}
 %
@@
 
+\subsection{Classical planning with conditional effects}
+$F$ is the set of {\em fluents} or {\em state variables} (propositional variables).
+A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e.
+either~$l=f$ or $l=\neg f$.
+$L$ is a set of literals that represents a partial assignment of values to fluents, and $\mathcal{L}(F)$ is the set of all literals sets on $F$, i.e.~all partial assignments of values to fluents.
+A {\em state} $s$ is a full assignment of values to fluents.
+We explicitly include negative literals $\neg f$ in states and so $|s|=|F|$ and the size of the state space is $2^{|F|}$.
+
+A {\em planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
+An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$,  and {\em effects} $\eff(a)\in\mathcal{L}(F)$.
+The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
+Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$.
+And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\neg\eff(a))\cup\eff(a)\}$, with $\neg\eff(a) = \{\neg l : l \in \eff(a)\}$.
+
+A {\em planning problem} is defined as a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state in which all the fluents of $F$ are assigned a value true/false and $G$ is the goal set.
+A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, and $|\pi|=n$ denotes its {\em plan length}.
+The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A trajectory $\tau(\pi,P)$ that solves $P$ is one in which $G \subseteq s_n$.
+
+An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
+Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+An action $a_c\in A$ is applicable in a state $s$ if and only if $\pre(a_c)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
+\[
+triggered(s,a_c)=\bigcup\limits_{C\rhd E\in\cond(a_c),C\subseteq s} E. 
+\]
+The result of applying $a_c$ in state $s$ follows the same definition of successor state, $\theta(s,a)$, but applied to the conditional effects in $triggered(s,a_c)$.
+
+\subsection{The observation model}
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
+Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
+
+%which indicates that we observe $l$ actions, $1\leq l\leq |\pi|$, and $m$ states, $1\leq m \leq |\pi|+1$, from $\tau(\pi,P)$:
+
+\begin{itemize}
+\item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
+Specifically, the number of observed actions, $l$, can range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
+\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
+A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
+Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
+    %Exceptionally, $s_m^o$ cannot be fully unobservable for the purpose of our task (we will elaborate on this issue later on).
+\end{itemize}
+
+We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
+In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
+
+Figure~\ref{fig:grid-example} illustrates a partial observation of a six-state trajectory \{{\tt\scriptsize<(xcoord 0)(ycoord 0)>, <(xcoord 1)(ycoord 0)>, <(xcoord 2)(ycoord 0)>, <(xcoord 3)(ycoord 0)>, <(xcoord 3)(ycoord 1)>, <(xcoord 2)(ycoord 1)>}\}.
+This observation only contains fluents of the predicates {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, corresponding to predicates {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unobservable in the six states.
+
 \section{Explanation-based learning of action models}
+The {\em one-shot} learning task to learn action models from {\em domain-specific knowledge} is defined as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, where:
+
+\begin{itemize}
+\item $\mathcal{M}$ is the {\em initial empty model} that contains only the header of each action model to be learned.
+\item $\mathcal{O}$ is a single learning example or plan observation; i.e. a sequence of (partially) observable states representing the evidence of the execution of an observed agent.
+\item $\Phi$ is a set of logic formulae that define {\em domain-specific knowledge}.
+\end{itemize}
+
+A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ is a model $\mathcal{M}'$ s.t.
+there exists a plan computable with $\mathcal{M}'$ that is consistent with the headers of $\mathcal{M}$, the observed states of $\mathcal{O}$ and the given domain knowledge in $\Phi$.
+
+\subsection{The space of \strips\ action models}
+
+We analyze here the solution space of a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$; i.e., the space of \strips\ action models.
+In principle, for a given action model $\xi$, any element of ${\mathcal I}_{\xi,\Psi}$ can potentially appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
+In practice, the actual space of possible \strips\ schemata is bounded by:
+
+\begin{enumerate}
+\item {\bf Syntactic constraints}.
+The solution $\mathcal{M}'$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+{\em Typing constraints} would also be a type of syntactic constraint~\cite{mcdermott1998pddl}.
+\item {\bf Observation constraints}.
+The solution $\mathcal{M}'$ must be consistent with these \emph{semantic constraints} derived from  the learning samples $\mathcal{O}$, which in our case is a single plan observation.
+Specifically, the states induced by the plan computable with $\mathcal{M}'$ must comprise the observed states of the sample, which further constrains the space of possible action models.
+\end{enumerate}
+
+Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
+In this work, the belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre\_\xi\_e$ and $eff\_\xi\_e$, instead of the three fluents used in the BLS.
+The four possible combinations of these two fluents are sumarized in Figure \ref{fig:combinations}.
+This compact encoding allows for a more effective exploitation of the syntactic constraints, and also yields the solution space of $\Lambda$ to be the same as its search space.
+
+\begin{figure}
+	\begin{footnotesize}
+		\begin{tabular}{lll}
+			{\bf Combination} & {\bf Meaning}\\\hline
+			$\neg pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ belongs neither to the preconditions \\
+             & nor effects of $\xi$ \\
+             & ($e \notin pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$)\\
+			$pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ is only a precondition of $\xi$\\
+               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$) \\
+			$\neg pre\_\xi\_e \wedge eff\_\xi\_e $& $e$ is a positive effect of $\xi$ \\
+               &  ($e \notin pre(\xi) \wedge e \in add(\xi) \wedge e \notin del(\xi)$) \\
+			$pre\_\xi\_e \wedge eff\_\xi\_e  $& $e$ is a negative effect of $\xi$ \\
+               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \in add(\xi)$) \\
+		\end{tabular}
+	\end{footnotesize}
+	\caption{\small Combinations of the fluent propositional encoding and their meaning}
+	\label{fig:combinations}
+\end{figure}
+
+\subsection{The sampling space}
+
+The single plan observation of $\mathcal{O}$ is defined as $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$, a sequence of possibly {\em partially observed states} except for the initial state $s_0^o$ which is a {\em fully observable} state.
+As commented before, the predicates $\Psi$ and the objects that shape the fluents $F$ are then deducible from $s_0^o$.
+A partially observed state $s_i^o$, ${\small 1\leq i\leq m}$, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ was not observed.
+Intermediate states can be {\em missing}, meaning that they are unobservable, so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than one action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$ (with ${\small k\geq 1}$ is unknown but finite).
+The minimal expression of a learning sample must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed final state $s_m^o$ so $m \geq 1$.
+
+Figure~\ref{fig:observation} shows a learning example that contains an initial state of the blocksworld where the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
+The observation represents a partially observable final state in which {\tt\small{blockA}} is on top of {\tt\small{blockB}} and {\tt\small{blockB}} on top of {\tt\small{blockC}}.
+
+\begin{figure}[hbt!]
+  \begin{small}
+  \begin{verbatim}
+(:predicates (on ?x ?y) (ontable ?x) (clear ?x) (handempty) (holding ?x))
+
+(:objects blockA blockB blockC)
+
+(:init (ontable blockA) (clear blockA) (ontable blockB) (clear blockB)
+       (ontable blockC) (clear blockC) (handempty))
+
+(:observation (on blockA blockB) (on blockB blockC))
+  \end{verbatim}
+  \end{small}
+	\caption{\small Example of a two-state observationn for the learning \strips\ action models.}
+	\label{fig:observation}
+\end{figure}
+
+\subsection{The domain-specific knowledge}
+One can introduce domain-specific knowledge to constrain further the space of possible schemata.
+For instance, back to the {\em blocksworld} domain, one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of any action model $\xi$ because, in this specific domain, a block cannot be on top of itself.
+
+\section{Learning \strips\ action models with classical planning}
+
+\section{Experimental results}
+
+\section{Conclussions}
 

### 2019-03-04 doc/lecture-notes-EBL/ebl.tex
full structure given to lecture notes on ebl

@@
 \newcommand{\strips}{\textsc{Strips}}     % precondition
+
+\newtheorem{mytheorem}{Theorem}
+\newtheorem{mylemma}[mytheorem]{Lemma}
+\newtheorem{mydefinition}[mytheorem]{Definition}
+\newtheorem{myproposition}[mytheorem]{Proposition}
+\newtheorem{myconstruction}{Construction}
 %%%
@@
 %
-\title{Explanation-based learning of \strips\ action models}
+\title{Explanation-based learning of action models}
 %
@@
 \section{Background}
+This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
 
@@
 \subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as an interleaved combination of actions and states that represents the observation from the execution of $\pi$ in $P$.
-Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as a sequence of partial states that represents the observation from the execution of $\pi$ in $P$.
 
-%which indicates that we observe $l$ actions, $1\leq l\leq |\pi|$, and $m$ states, $1\leq m \leq |\pi|+1$, from $\tau(\pi,P)$:
-
-\begin{itemize}
-\item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
-Specifically, the number of observed actions, $l$, can range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
-\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
+Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, $s_0^o=I$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
 A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
@@
 This observation only contains fluents of the predicates {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, corresponding to predicates {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unobservable in the six states.
+
+We introduce a particular class of $\mathcal{O}(\tau)$ observations.
+This new class allows us to distinguish between {\em observable} state variables, whose value may be read from sensors, and {\em hidden} or {\em latent} state variables that cannot be observed.
+
+\begin{definition}[$\Phi$-observation]
+Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
+\end{definition}
+
+We consider that the {\bf observed actions} are fluents consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$ s.t.
+$a_i^o\in s_i^o$ ({\small $0\leq i <m$}).
+Specifically, the number of observed actions, $l$, can range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
 
@@
 For instance, back to the {\em blocksworld} domain, one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of any action model $\xi$ because, in this specific domain, a block cannot be on top of itself.
+Further we can use domain-specific knowledge to complete partially observed states.
+
+{\em State invariants} is a useful type of state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a planning problem $P=\tup{F,A,I,G}$, a state invariant is a formula $\phi$ that holds in $I$, $I\models \phi$, and in every state $s$ built out of $F$ that is reachable by applying actions of $A$ in $I$.
+Recently, some works point at extracting \emph{lifted} invariants, also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}, that hold for any possible state and any possible set of objects.
+Invariant templates obtained by inspecting the lifted representation of the domain have also been exploited for deriving \emph{lifted mutex}~\cite{BernardiniFS18}.
+{\em Schematic invariants} can identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998}.
+
+We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\xi,\Psi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are universally quantified.
+For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible state.
+Figure~\ref{fig:strongest-invariant} shows an example of four clauses that define schematic mutexes for the {\em blocksworld} domain.
+
+\begin{figure}[hbt!]
+  \begin{footnotesize}
+$\forall x_1,x_2\ \neg ontable(x_1)\vee\neg on(x_1,x_2)$.\\
+$\forall x_1,x_2\ \neg clear(x_1)\vee\neg on(x_2,x_1)$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
+\end{footnotesize}
+ \caption{\small {\em Schematic mutexes} for the {\em blocksworld} domain.}
+\label{fig:strongest-invariant}
+\end{figure}
 
 \section{Learning \strips\ action models with classical planning}
+
+Our proposal to address a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to transform $\Lambda$ into a planning task $P_{\Lambda}$.
+The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$.
+Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition, a positive effect or a negative effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\tau$, including that the final reached state comprises $s_n$.
+Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states of $\tau$.
+
+\subsection{Compilation}
+Our compilation scheme builds upon the approach presented in \cite{aineto2018learning} but \FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.
+
+\vspace{0.1cm}
+
+A learning task $\Lambda=\tup{\mathcal{M},\tau}$ is compiled into a planning task $P_{\Lambda}$ with conditional effects in the context of a planning frame $\Phi=\tup{F,A}$.
+We use conditional effects because they allow us to compactly define actions whose effects depend on the current state.
+An action $a\in A$ with conditional effects is defined as a set of preconditions $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
+Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+An action $a\in A$ is applicable in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$; that is, $triggered(s,a)=\bigcup\limits_{C\rhd E\in\cond(a),C\subseteq s} E$.
+The result of applying $a$ in state $s$ follows the same definition of successor state, $\theta(s,a)$, introduced in section \ref{basic_planning} but applied to the conditional effects in $triggered(s,a)$.
+
+\vspace{0.25cm}
+
+A solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output domain model $\mathcal{M}'$ that solves the learning task $\Lambda$. Specifically, a solution plan $\pi_\Lambda$ serves two purposes:
+
+\begin{enumerate}
+\item {\bf To build the action models of $\mathcal{M}'$}.
+$\pi_\Lambda$ comprises a first block of actions (plan {\em prefix}) that set the predicates $p\in \Psi_{\xi}$ of $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\mathcal{M}$.
+\item {\bf To validate the action models of $\mathcal{M}'$}.
+$\pi_\Lambda$ also comprises a second block of actions (plan {\em postfix}) which is aimed at validating of the observed plan trace $\tau$ with the built action models $\mathcal{M}'$.
+\end{enumerate}
+
+%\subsection{Learning from observations of plan executions}
+Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$, with $\tau$ formed by an $n$-action sequence $\tup{a_1, \ldots, a_n}$ and a $m$-state trajectory $\tup{s_0, s_1, \ldots, s_m}$ ($\tau = \langle s_0, a_1, \ldots, a_n, s_m \rangle$), the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
+
+\begin{itemize}
+
+\item $F_{\Lambda}$ extends $F$ with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of $\tau$.
+Specifically, $F_{\Lambda}$ contains:
+\begin{itemize}
+\item The set of fluents obtained from $s_0$; i.e., $F$.
+\item The model fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section \ref{propositional_encoding}
+\item A set of fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to represent the $i^{th}$ observable action of $\tau$.
+In the example of Figure~\ref{fig:example-plans}, the two observed actions {\small \texttt{(putdown B)}} and {\small \texttt{(stack  A  B)}} would be encoded as fluents  {\small \texttt{(plan-putdown B i1)}} and {\small \texttt{(plan-stack A B i2)}} to indicate that {\small \texttt{(putdown B)}} is observed in the first place and {\small \texttt{(stack  A  B)}} is the second observed action.
+\item Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions of $\tau$.
+The former is used to ensure that actions are executed in the same order as they are observed in $\tau$.
+The latter is used to iterate to the next planning step when solving $P_{\Lambda}$.
+\item A set of fluents $\{test_j\}_{0\leq j\leq m}$, to point at the state observation $s_j\in\tau$ where the action model is
+validated. In the example of Figure~\ref{fig:example-plans} two tests are required to validate the programmed action model, one test at $s_0$ and another one at $s_4$.
+\item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated.
+\end{itemize}
+
+\item $I_{\Lambda}$ encodes $s_0$ and the following fluents set to true: $mode_{prog}$, $test_0$, $F_{\pi}$, $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i \leq n$}.
+Our compilation assumes that action models are initially programmed with no precondition, no negative effect and no positive effect.
+
+\item $G_{\Lambda}$ includes the positive literals $at_n$ and $test_m$.
+When these two goals are achieved by the solution plan $\pi_\Lambda$, we will be certain that the action models of $\mathcal{M'}$ are validated in all the actions and states observed in the input plan trace $\tau$.
+
+\item $A_{\Lambda}$ includes three types of actions that give rise to the actions of $\pi_\Lambda$.
+\begin{enumerate}
+\item Actions for {\em inserting} a component (precondition, positive effect or negative effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models.
+These actions will form the prefix of the solution plan $\pi_\Lambda$.
+Among the \emph{inserting} actions, we find:
+\begin{itemize}
+\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
+A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $del_p$ nor $add_p$ exist in $\xi$.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
+\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
+\end{align*}
+\end{small}
+
+\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
+A positive effect is inserted in $\xi$ under the same conditions of a precondition insertion, and a negative effect is inserted in $\xi$ when neither $del_p$ nor $add_p$ appear in $\xi$ but $pre_p$ does.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
+\cond(\mathsf{insertEff_{p,\xi}})=&\{pre_{p}(\xi)\}\rhd\{del_{p}(\xi)\},\\
+& \{\neg pre_{p}(\xi)\}\rhd\{add_{p}(\xi)\}.
+\end{align*}
+\end{small}
+\end{itemize}
+
+For instance, given $name(\xi)=${\tt{\small stack}} and $C_{pre-stack}=\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_holding\_v2), (pre\_stack\_on\_v1\_v2),(pre\_stack\_clear\_v1),(pre\_stack\_clear\_v1),}}$\ldots \}$, the insertion of each item $c \in C_{pre-stack}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$ as long as $c \notin pre(\xi)$,  $c \notin add(\xi)$ and $c \notin del(\xi)$.
+The same applies to effects with respect to sets $C_{add-stack}$ and $C_{del-stack}$ that would include all fluents starting with prefix  {\tt{\small add}} and {\tt{\small del}}, respectively.
+
+\vspace{0.1cm}
+
+Note that executing an insert action, e.g.{\tt{\small (insert\_pre\_stack\_holding\_v1)}}, will add the corresponding model fluent {\tt{\small (pre\_stack\_holding\_v1)}} to the successor state.
+Hence, the execution of the insert actions of $\pi_\Lambda$ yield a state containing the valuation of the model fluents that shape every $\xi \in \mathcal{M}$.
+For example, executing the insert actions that shape the action model $name(\xi)=${\tt{\small putdown}} leads to a state containing the positive literals {\tt{\small (pre\_putdown\_holding\_v1),(eff\_putdown\_holding\_v1),\\ (eff\_putdown\_clear\_v1),
+(eff\_putdown\_ontable\_v1),(eff\_putdown\_handempty)}}.
+
+\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{ar(\xi)}$.
+Since action headers are known, the variables $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
+\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
+&\{add_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi},\\
+&\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
+\end{align*}
+\end{small}
+
+    These actions will be part of the postfix of the plan $\pi_\Lambda$ and they determine the application of the learned action models according to the values of the model fluents in the current state configuration}.
+Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator.
+\textcolor[rgb]{1.00,0.00,0.00}{Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$.
+Executing this action in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$.
+For example, if  $\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}}$\} \subset s$ then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
+Otherwise, a different set of precondition literals will be checked.
+The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.
+
+Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if $name(\xi)=${\tt{\small stack}} has been correctly programmed by the insert actions.
+Hence, while \textbf{insert actions} add the values of the \textbf{model fluents} that shape $\xi$, the \textbf{apply actions} add the values of the \textbf{fluents of $F$} that result from the execution of $\xi$.
+
+\begin{figure}[hbt!]
+\begin{center}
+\begin{scriptsize}
+\begin{verbatim}
+(:action apply_stack
+  :parameters (?o1 - object ?o2 - object)
+  :precondition
+   (and (or (not (pre_stack_on_v1_v1)) (on ?o1 ?o1))
+        (or (not (pre_stack_on_v1_v2)) (on ?o1 ?o2))
+        (or (not (pre_stack_on_v2_v1)) (on ?o2 ?o1))
+        (or (not (pre_stack_on_v2_v2)) (on ?o2 ?o2))
+        (or (not (pre_stack_ontable_v1)) (ontable ?o1))
+        (or (not (pre_stack_ontable_v2)) (ontable ?o2))
+        (or (not (pre_stack_clear_v1)) (clear ?o1))
+        (or (not (pre_stack_clear_v2)) (clear ?o2))
+        (or (not (pre_stack_holding_v1)) (holding ?o1))
+        (or (not (pre_stack_holding_v2)) (holding ?o2))
+        (or (not (pre_stack_handempty)) (handempty)))
+  :effect
+   (and (when (del_stack_on_v1_v1) (not (on ?o1 ?o1)))
+        (when (del_stack_on_v1_v2) (not (on ?o1 ?o2)))
+        (when (del_stack_on_v2_v1) (not (on ?o2 ?o1)))
+        (when (del_stack_on_v2_v2) (not (on ?o2 ?o2)))
+        (when (del_stack_ontable_v1) (not (ontable ?o1)))
+        (when (del_stack_ontable_v2) (not (ontable ?o2)))
+        (when (del_stack_clear_v1) (not (clear ?o1)))
+        (when (del_stack_clear_v2) (not (clear ?o2)))
+        (when (del_stack_holding_v1) (not (holding ?o1)))
+        (when (del_stack_holding_v2) (not (holding ?o2)))
+        (when (del_stack_handempty) (not (handempty)))
+        (when (add_stack_on_v1_v1) (on ?o1 ?o1))
+        (when (add_stack_on_v1_v2) (on ?o1 ?o2))
+        (when (add_stack_on_v2_v1) (on ?o2 ?o1))
+        (when (add_stack_on_v2_v2) (on ?o2 ?o2))
+        (when (add_stack_ontable_v1) (ontable ?o1))
+        (when (add_stack_ontable_v2) (ontable ?o2))
+        (when (add_stack_clear_v1) (clear ?o1))
+        (when (add_stack_clear_v2) (clear ?o2))
+        (when (add_stack_holding_v1) (holding ?o1))
+        (when (add_stack_holding_v2) (holding ?o2))
+        (when (add_stack_handempty) (handempty))
+        (when (modeProg) (not (modeProg)))))
+\end{verbatim}
+\end{scriptsize}
+ \caption{\small PDDL action for applying an already programmed model for $stack$ (implications are coded as disjunctions).}
+\label{fig:compilation}
+\end{center}
+\end{figure}
+
+When the input plan trace contains observed actions, the extra conditional effects
+
+$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.\\
+
+\item Actions for {\em validating} the partially observed state $s_j\in\tau$, {\tt\small $1\leq j< m$}.
+These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observable data of the input plan trace $\tau$ follows after the execution of the apply actions.
+
+\begin{small}
+\begin{align*}
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j\}.
+\end{align*}
+\end{small}
+
+There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$.
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.
+
+\end{enumerate}
+\end{itemize}
+
+In some contexts, it is reasonable to assume that some parts of the action model are known and so there is no need to learn the entire model from scratch \cite{ZhuoNK13}.
+In \FAMA, when an action model $\xi$ is partially specified, the known preconditions and effects are encoded as fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$ set to true in the initial state $I_{\Lambda}$.
+In this case, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, thereby making the classical planning task $P_{\Lambda}$ easier to be solved.
+
+So far we have explained the compilation for learning from a single input trace.
+However, the compilation is extensible to the more general case $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ is a set of plan traces.
+Taking this into account, a small modification is required in our compilation approach.
+In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_m^t\in \tau_t$, {\tt\small $1\leq t\leq k$} of a plan trace $\tau_t$ reset the current state and the current plan.
+These actions are now redefined as:
+
+\begin{small}
+	\begin{align*}
+	\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_m^t\cup\{test_{j-1}\}\cup \{\neg mode_{prog}\},\\
+	\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_j\} \cup \\
+	&\{\neg f\}_{\forall f\in s_m^t, f \notin s_0^{t+1}}\cup \{f\}_{\forall f\in s_0^{t+1}, f \notin s_m^t},\\
+	&\{\neg f\}_{\forall f\in F_{\pi_t}}\cup \{f\}_{\forall f\in F_{\pi_{t+1}}}.\\
+	\end{align*}
+\end{small}
+
+Finally, we will detail the composition of a solution plan $\pi_\Lambda$ to a planning task $P_\Lambda$ and the mechanism to extract the action models of $\mathcal{M}'$ from $\pi_\Lambda$.
+The plan of Figure~\ref{fig:plan-lplan} shows a solution to the task $P_{\Lambda}$ that encodes a learning task $\Lambda=\tup{\mathcal{M},\tau}$ for obtaining the action models of the {\em blocksworld} domain, where the models for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already specified in $\mathcal{M}$.
+Therefore, the plan shows the insert actions and validate action for the action model {\tt\small stack} using the input plan trace of Figure~\ref{fig:example-plans}.
+Plan steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects, and steps $07-11$ form the plan postfix that applies the action models (only the {\tt\small stack} model is learned) and validates the result in the plan trace of Figure~\ref{fig:example-plans}.
+
+\begin{figure}[hbt!]
+	{\footnotesize\tt
+		{\bf 00} : (insert\_pre\_stack\_holding\_v1) \\
+		01 : (insert\_pre\_stack\_clear\_v2)\\
+		{\bf 02} : (insert\_eff\_stack\_clear\_v1)\\
+		03 : (insert\_eff\_stack\_clear\_v2)\\
+		04 : (insert\_eff\_stack\_handempty)\\
+		05 : (insert\_eff\_stack\_holding\_v1)\\
+		06 : (insert\_eff\_stack\_on\_v1\_v2)\\
+		{\bf 07} : (apply\_unstack blockB blockA i1 i2)\\
+		08 : (apply\_putdown blockB i2 i3)\\
+		09 : (apply\_pickup blockA i3 i4)\\
+		10 : (apply\_stack blockA blockB i4 i5)\\
+		{\bf 11} : (validate\_1)
+	}
+	\caption{\small Plan for programming and validating the $stack$ action model (using the plan trace $\tau$ of Figure~\ref{fig:example-plans}) as well as previously specified action models for $pickup$, $putdown$ and $unstack$.}
+	\label{fig:plan-lplan}
+\end{figure}
+
+Given a solution plan $\pi_\Lambda$ that solves $P_{\Lambda}$, the set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ are computed in linear time and space.
+In order to do so, $\pi_\Lambda$ is executed in the initial state $I_{\Lambda}$ and the action model $\mathcal{M}'$ will be given by the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that are set to true in the last state reached by $\pi_\Lambda$, $s_g=\theta(I_\Lambda,\pi_\Lambda)$.
+For each $\xi \in \mathcal{M'}$, we build the sets of preconditions, positive effects and negative effects as follows:
+
+\begin{small}
+	\begin{align*}
+	\hspace*{7pt}pre(\xi)=& \{p ~|~ pre_p(\xi) \in s_g\}_{\forall p \in \Psi_\xi},\\
+	\hspace*{7pt}add(\xi)=& \{p ~|~ add_p(\xi) \in s_g\}_{\forall p \in \Psi_\xi},\\
+	\hspace*{7pt}del(\xi)=& \{p ~|~ del_p(\xi) \in s_g\}_{\forall p \in \Psi_\xi}.
+	\end{align*}
+\end{small}
+
+The logical inference process our approach is based on has trouble learning preconditions that do not appear as negative effects since in this case no change is observed between the pre-state and post-state of an action.
+This is specially relevant for static predicates that never change and, hence, only appear as preconditions in the actions.
+In order to address this shortcoming and complete the list of learned preconditions}, we apply a post-process based on the one proposed in~\cite{kuvcera2018louga}.
+The idea lies in going through every action and counting the number of cases where a literal is present before the action is executed and the number of cases where it is not present.
+If a literal is present in all the cases before the action, the literal is considered to be a precondition.
+
+In order to obtain a complete trace, the proposal in ~\cite{kuvcera2018louga} applies the sequence of actions of the input trace and infers the preconditions from this \FO action sequence}.
+In our case, since the sequence of actions of the input trace might not be fully observable, we produce the traces by applying the actions found in the validation part of the solution plan.
+For instance, in the example of the figure \ref{fig:plan-lplan}, the sequence of actions used to produce the complete trace would be {\tt{\small(unstack blockB blockA)}}, {\tt{\small(put-down blockB)}}, {\tt{\small(pick-up blockA)}}, and {\tt{\small(stack blockA blockB)}.
+
+%This post-process allows \FAMA to learn more complete lists of preconditions and deal with the always problematic static predicates.
+
+\subsection{Properties of the compilation}
+\begin{mylemma}
+Soundness.
+Any classical plan $\pi$ that solves $P'$ (planning task that results from the compilation) produces a model $\mathcal{M'}$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
+\end{mylemma}
+
+\begin{proof}[Proof]
+\begin{small}
+According to the $P'$ compilation, once a given precondition or effect is inserted into the domain model $\mathcal{M}$ it cannot be undone.
+In addition, once an action model is applied it cannot be modified.
+In the compiled planning task $P'$, only ${\tt \small (apply)_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
+This means that a state consistent with an observation $s_m^o$ can only be achieved executing an applicable sequence of ${\tt \small (apply)_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_j$ $(0 < j\leq m)$, is consistent with the input state observations and {\em state-invariants}.
+This is exactly the definition of the solution condition for model $\mathcal{M}'$ to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
+\end{small}
+\end{proof}
+
+\begin{mylemma}
+Completeness. Any model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task can be computed with a classical plan $\pi$ that solves $P'$.
+\end{mylemma}
+
+\begin{proof}[Proof]
+\begin{small}
+By definition ${\mathcal I}_{\xi,\Psi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
+In addition the $P'$ compilation does not discard any model $\mathcal{M}'$ definable within ${\mathcal I}_{\xi,\Psi}$ that satisfies the mutexes in $\Phi$.
+This means that, for every model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi$ that solves $P'$ by selecting the appropriate ${\tt \small (insert\_pre)_{\xi,e}}$ and ${\tt \small (insert\_eff)_{\xi,e}}$ actions for programming the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding ${\tt \small (apply)_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
+\end{small}
+\end{proof}
+
+The size of $P'$ depends on the arity of the predicates in $\Psi$, that shape variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
+The larger these arities, the larger $|{\mathcal I}_{\xi,\Psi}|$.
+The size of ${\mathcal I}_{\xi,\Psi}$ is the most dominant factor of the compilation because it defines the $pre\_\xi\_e/eff\_\xi\_e$ fluents, the corresponding set of ${\tt \small insert}$ actions, and the number of conditional effects in the ${\tt \small (apply)_{\xi,\omega}}$ actions.
+Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$, which will significantly reduce $|{\mathcal I}_{\xi,\Psi}|$ and hence the size of $P'$ output by the compilation.
+
+Classical planners tend to prefer shorter solution plans, so our compilation (as well as the BLS) may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of preconditions/effects.
+In more detail, all $\{pre\_\xi\_e, eff\_\xi\_e\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ fluents are false at the initial state of our $P'$ compilation so classical planners tend to solve $P'$ with plans that require a smaller number of ${\tt{\small insert}}$ actions.
+
+This bias can be eliminated defining a cost function for the actions in $P'$ (e.g.
+${\tt \small insert}$ actions have {\em zero cost} while ${\tt \small (apply)_{\xi,\omega}}$ actions have a {\em positive constant cost}).
+In practice we use a different approach to disregard the cost of ${\tt \small insert}$ actions since classical planners are not proficient at optimizing plan cost with zero-cost actions.
+Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} that can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
+Further, the actions for inserting action effects are also applied in another single planning step.
+The plan horizon for programming any action model is then always bounded to 2.
+The SAT-based planning approach is also convenient for its ability to deal with planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect the planning performance.
+
+An interesting aspect of our approach is that when a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given in $\Lambda$, the $P_{\Lambda}$ compilation also serves to validate whether the observed $\tau$ follows the given model $\mathcal{M}$:
+
+\begin{itemize}
+	\item $\mathcal{M}$ is proved to be a {\em valid} action model for the given input data in $\tau$ iff a solution plan for $P_{\Lambda}$ can be found.
+	\item $\mathcal{M}$ is proved to be a {\em invalid} action model for the given input data $\tau$ iff $P_{\Lambda}$ is unsolvable.
+This means that $\mathcal{M}$ cannot be consistent with the given observation of the plan execution.
+\end{itemize}
+
+The validation capacity of our compilation is beyond the functionality of VAL (the plan validation tool~\cite{howey2004val}) because our $P_{\Lambda}$ compilation is able to address {\em model validation} of a partial (or even an empty) action model with a partially observed plan trace.
+VAL, however, requires a full plan and a full action model for plan validation.
 
@@
 
+\bibliographystyle{named}
+\bibliography{planlearnbibliography}
 \end{document}

### 2019-03-05 doc/lecture-notes-EBL/ebl.tex
introduction and background updated for lecture notes on ebl

@@
 %%% Defintions For this paper
+\usepackage[numbers]{natbib}
 \usepackage{amsmath}
@@
 
+\newcommand{\ARMS}{{\small {\sffamily ARMS}}\xspace}
+\newcommand{\CAMA}{{\small {\sffamily CAMA}}\xspace}
+\newcommand{\SLAF}{{\small {\sffamily SLAF}}\xspace}
+\newcommand{\LAMP}{{\small {\sffamily LAMP}}\xspace}
+\newcommand{\NOISTA}{{\small {\sffamily NOISTA}}\xspace}
+\newcommand{\LOCM}{{\small {\sffamily LOCM}}\xspace}
+\newcommand{\LOCMtwo}{{\small {\sffamily LOCM2}}\xspace}
+\newcommand{\LOP}{{\small {\sffamily LOP}}\xspace}
+\newcommand{\AMAN}{{\small {\sffamily AMAN}}\xspace}
+\newcommand{\LOUGA}{{\small {\sffamily LOUGA}}\xspace}
+\newcommand{\FAMA}{{\small {\sffamily FAMA}}\xspace}
+
 \newtheorem{mytheorem}{Theorem}
@@
 \author{Diego Aineto\inst{1}\orcidID{} \and
-Sergio Jiménez\inst{1}\orcidID{0000-0003-0561-4880} \and
+Sergio Jim\'enez\inst{1}\orcidID{0000-0003-0561-4880} \and
 Eva Onaindia\inst{1}\orcidID{0000-0001-6931-8293}}
@@
 \begin{abstract}
-
+This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this inductive learning task into a classical planning task.
+The compilation approach is flexible to different amounts of available input knowledge; learning examples can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just pairs of initial and final states where no intermediate action/state is observed.
+The compilation accepts also partially specified action models and it can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified or the observation is incomplete.
 \end{abstract}
@@
 \section{Introduction}
+Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
+In both tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
+Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
+
+{\em Machine Learning} (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
+The application of inductive ML to learning \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
+\begin{itemize}
+\item The {\em input} to ML algorithms (the learning/training data) is usually a finite set of vectors that represent the value of some fixed object features.
+The input for learning planning action models is, however, observations of plan executions (where each plan possibly has a different length).
+\item The {\em output} of ML algorithms is usually a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
+When learning action models the output is, for each action, the preconditions and effects that define the possible state transitions of a planning task.
+\end{itemize}
+
+Learning \strips\ action models is a well-studied problem with sophisticated algorithms such as {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}, which do not require full knowledge of the intermediate states traversed by the example plans.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper describes an innovative planning compilation approach for learning \strips\ action models.
+The compilation approach is appealing by itself because it opens up the door to the bootstrapping of planning action models, but also because:
+
+\begin{enumerate}
+\item It is flexible to various amounts of input knowledge.
+Learning examples range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just a pair of initial and final states where no intermediate state/action is observed.
+\item It accepts previous knowledge about the structure of the actions in the form of partially specified action models.
+In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
+\end{enumerate}
 
 \section{Background}
-This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
+This section formalizes the models for {\em classical planning}, for the {\em observation} of the execution of a classical plan and for the explanation of an observation.
 
@@
 The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
-A trajectory $\tau(\pi,P)$ that solves $P$ is one in which $G \subseteq s_n$.
+A plan $\pi$ solves $P$ if the induced trajectory $\tau(\pi,P)$ holds that $G \subseteq s_n$.
+A plan is optimal iff its length is minimal.
 
+Now we define actions with conditional effects because they allow to compactly define our compilation.
 An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
@@
 \subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define the \emph{observation of the trajectory} as a sequence of partial states that represents the observation from the execution of $\pi$ in $P$.
-
+Given a planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$, the \emph{observation of the trajectory} $\tau(\pi,P)$ is a sequence of partial states that captures what is observed from the execution of $\pi$ in $P$.
 Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, $s_0^o=I$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
 A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
-Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
-Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+
+This {\em observation model} comprises the case $|s_i^o| = 0$, when an intermediate state is fully unobservable.
+Further, this model consider also {\em observed actions} as fluents that indicate the action applied in a given state.
+This means that a sequence of {\em observed actions} $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$ s.t.
+$a_i^o\in s_i^o$ ({\small $0\leq i <m$}).
+The number of {\em observed actions}, $l$, can then range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
+
+Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be {\em consistent} with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
 In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
-
-We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
+This means that we assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
 Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
@@
 
-Figure~\ref{fig:grid-example} illustrates a partial observation of a six-state trajectory \{{\tt\scriptsize<(xcoord 0)(ycoord 0)>, <(xcoord 1)(ycoord 0)>, <(xcoord 2)(ycoord 0)>, <(xcoord 3)(ycoord 0)>, <(xcoord 3)(ycoord 1)>, <(xcoord 2)(ycoord 1)>}\}.
-This observation only contains fluents of the predicates {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, corresponding to predicates {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unobservable in the six states.
+Last but not least, our observation model can distinguish between {\em observable state variables}, whose value may be read from sensors, and {\em hidden} (or {\em latent}) {\em state variables}, that cannot be observed.
+Given a subset of fluents $\Gamma\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Gamma$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Gamma$.
 
-We introduce a particular class of $\mathcal{O}(\tau)$ observations.
-This new class allows us to distinguish between {\em observable} state variables, whose value may be read from sensors, and {\em hidden} or {\em latent} state variables that cannot be observed.
+\subsection{The explanation model}
+Let $\Phi_\mathcal{O}=\tup{F,A}$ be a {\em classical planning frame} where the fluents $F$ are deducible from $\mathcal{O}$, since $s_0^o$ is fully observed.
+Let also $P_\mathcal{O}$ be the classical planning problem $P_\mathcal{O}=\tup{F,A,s_0^o,s_m^o}$.
 
-\begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
+\begin{definition}[Explanation]
+We say that a plan $\pi$ {\em explains} $\mathcal{O}$ (denoted $\pi\mapsto\mathcal{O}$) iff $\pi$ is a solution for $P$ that is {\em consistent} with the state trajectory constraints imposed by the sequence of partial states $\mathcal{O}$.
 \end{definition}
 
-We consider that the {\bf observed actions} are fluents consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$ s.t.
-$a_i^o\in s_i^o$ ({\small $0\leq i <m$}).
-Specifically, the number of observed actions, $l$, can range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
+If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}(\tau)$.
 
 \section{Explanation-based learning of action models}
-The {\em one-shot} learning task to learn action models from {\em domain-specific knowledge} is defined as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, where:
+The task of learning action models from the observation of a plan execution is defined as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O},\kappa}$, where:
 
 \begin{itemize}
-\item $\mathcal{M}$ is the {\em initial empty model} that contains only the header of each action model to be learned.
-\item $\mathcal{O}$ is a single learning example or plan observation; i.e. a sequence of (partially) observable states representing the evidence of the execution of an observed agent.
-\item $\Phi$ is a set of logic formulae that define {\em domain-specific knowledge}.
+\item $\mathcal{M}$ is the {\em initial empty model} that contains only the {\em header} (i.e., the corresponding {\em name} and {\em parameters}) of each action model to be learned.
+\item $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$ is a sequence of partially observed states.
+\item $\kappa$ is a set of logic formulae that define {\em domain-specific knowledge}.
 \end{itemize}
 
-A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ is a model $\mathcal{M}'$ s.t.
-there exists a plan computable with $\mathcal{M}'$ that is consistent with the headers of $\mathcal{M}$, the observed states of $\mathcal{O}$ and the given domain knowledge in $\Phi$.
+Now we are ready to define a {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\kappa}$ as a model $\mathcal{M}'$ s.t.
+$\mathcal{M}'$ explains $\mathcal{O}$ and it is consistent with the headers of $\mathcal{M}$ and the given domain knowledge in $\kappa$.
+We say that a model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}$.
 
 \subsection{The space of \strips\ action models}
-
 We analyze here the solution space of a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$; i.e., the space of \strips\ action models.
@@
 Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
-In this work, the belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre\_\xi\_e$ and $eff\_\xi\_e$, instead of the three fluents used in the BLS.
+The belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre\_\xi\_e$ and $eff\_\xi\_e$, instead of the three fluents used in the original compilation~\cite{aineto2018learning}.
 The four possible combinations of these two fluents are sumarized in Figure \ref{fig:combinations}.

### 2019-03-05 doc/lecture-notes-EBL/ebl.tex
updated sections 1,2 and 3 for lecture notes on ebl

@@
 
+With regard to this definition, an {\em observation} can be considered a sequence of ordered {\em landmarks} for the $P_\mathcal{O}$ classical planning problem, because the literals in the observatoin must be achieved by all the plans that solve this problem in the same order~\cite{hoffmann2004ordered}.
+
 \section{Explanation-based learning of action models}
@@
 We analyze here the solution space of a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$; i.e., the space of \strips\ action models.
+
+A \strips\ \emph{action model} $\xi$ is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
+
+Let $\Psi$ be the set of {\em predicates} (the initial state of an observation is a full assignment of values to fluents, $|s_0|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0$) that shape the variables $F$.
+The set of propositions that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
+For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), \\
+on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi$.
+
 In principle, for a given action model $\xi$, any element of ${\mathcal I}_{\xi,\Psi}$ can potentially appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
 In practice, the actual space of possible \strips\ schemata is bounded by:
-
 \begin{enumerate}
@@
 	\begin{footnotesize}
-		\begin{tabular}{lll}
-			{\bf Combination} & {\bf Meaning}\\\hline
-			$\neg pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ belongs neither to the preconditions \\
-             & nor effects of $\xi$ \\
-             & ($e \notin pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$)\\
+		\begin{tabular}{c | c}
+			& {\bf Meaning}\\\hline
+$\neg pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ belongs neither to the preconditions nor effects of $\xi$ \\
+             & ($e \notin pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$)\\\\
 $pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ is only a precondition of $\xi$\\
-               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$) \\
+               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$) \\\\
 $\neg pre\_\xi\_e \wedge eff\_\xi\_e $& $e$ is a positive effect of $\xi$ \\
-               &  ($e \notin pre(\xi) \wedge e \in add(\xi) \wedge e \notin del(\xi)$) \\
+               &  ($e \notin pre(\xi) \wedge e \in add(\xi) \wedge e \notin del(\xi)$) \\\\
 $pre\_\xi\_e \wedge eff\_\xi\_e  $& $e$ is a negative effect of $\xi$ \\
-               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \in add(\xi)$) \\
+               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \in add(\xi)$)
 		\end{tabular}
@@
 \subsection{The sampling space}
-
-The single plan observation of $\mathcal{O}$ is defined as $\mathcal{O}=\tup{s_0^o,s_1^o \ldots, s_m^o}$, a sequence of possibly {\em partially observed states} except for the initial state $s_0^o$ which is a {\em fully observable} state.
-As commented before, the predicates $\Psi$ and the objects that shape the fluents $F$ are then deducible from $s_0^o$.
-A partially observed state $s_i^o$, ${\small 1\leq i\leq m}$, is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ was not observed.
-Intermediate states can be {\em missing}, meaning that they are unobservable, so transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than one action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$ (with ${\small k\geq 1}$ is unknown but finite).
-The minimal expression of a learning sample must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed final state $s_m^o$ so $m \geq 1$.
+According to our observation model the minimal expression of an observation must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed final state $s_m^o$ so $m \geq 1$.
+The maximal expression of an observation corresponds to a fully observed trajectory (where all traversed states and applied actions are fully observed).
+In between there is a grey scale of  different kind of observations including the observation of the initial state and a plan that is frequently used for previous system that learn planning action models such as {\sc ARMS}~\cite{yang2007learning} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
 
@@
 For instance, back to the {\em blocksworld} domain, one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of any action model $\xi$ because, in this specific domain, a block cannot be on top of itself.
-Further we can use domain-specific knowledge to complete partially observed states.
+In addition, we can use {\em domain-specific knowledge} to complete partially observed states.
 
@@
 {\em Schematic invariants} can identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998}.
-
-We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\xi,\Psi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are universally quantified.
 For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible state.

### 2019-03-05 doc/lecture-notes-EBL/ebl.tex
updated sections 1,2 and 3 for lecture notes on ebl

@@
 
-\newcommand{\ARMS}{{\small {\sffamily ARMS}}\xspace}
-\newcommand{\CAMA}{{\small {\sffamily CAMA}}\xspace}
-\newcommand{\SLAF}{{\small {\sffamily SLAF}}\xspace}
-\newcommand{\LAMP}{{\small {\sffamily LAMP}}\xspace}
-\newcommand{\NOISTA}{{\small {\sffamily NOISTA}}\xspace}
-\newcommand{\LOCM}{{\small {\sffamily LOCM}}\xspace}
-\newcommand{\LOCMtwo}{{\small {\sffamily LOCM2}}\xspace}
-\newcommand{\LOP}{{\small {\sffamily LOP}}\xspace}
-\newcommand{\AMAN}{{\small {\sffamily AMAN}}\xspace}
-\newcommand{\LOUGA}{{\small {\sffamily LOUGA}}\xspace}
-\newcommand{\FAMA}{{\small {\sffamily FAMA}}\xspace}
-
 \newtheorem{mytheorem}{Theorem}
@@
 %
-\author{Diego Aineto\inst{1}\orcidID{} \and
-Sergio Jim\'enez\inst{1}\orcidID{0000-0003-0561-4880} \and
-Eva Onaindia\inst{1}\orcidID{0000-0001-6931-8293}}
+\author{Diego Aineto\inst{1} \and
+Sergio Jim\'enez\inst{1} \and
+Eva Onaindia\inst{1}}
 %
@@
 \begin{abstract}
-This paper presents a novel approach for learning \strips\ action models from observations of plan executions that compiles this inductive learning task into a classical planning task.
-The compilation approach is flexible to different amounts of available input knowledge; learning examples can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just pairs of initial and final states where no intermediate action/state is observed.
-The compilation accepts also partially specified action models and it can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified or the observation is incomplete.
+The paper presents the classical planning compilation for learning \strips\ action models from observations of plan executions.
+The compilation is flexible to different amount and kind of available input knowledge; learning examples can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just pairs of initial and final states where no intermediate action/state is observed.
+The compilation accepts also partially specified action models and it can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified or the input observation is incomplete.
 \end{abstract}
 
-\keywords{Learning action models \and Planning and Learning \and Classical Planning.}
+\keywords{Learning action models \and Classical planning.}
 
@@
 In both tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
-Unfortunately, building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
+Unfortunately building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
 
@@
 \begin{itemize}
-\item The {\em input} to ML algorithms (the learning/training data) is usually a finite set of vectors that represent the value of some fixed object features.
+\item The {\em input} to ML algorithms (the learning/training data) is usually a finite vector that represents the value of some fixed object features.
 The input for learning planning action models is, however, observations of plan executions (where each plan possibly has a different length).
@@
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper describes an innovative planning compilation approach for learning \strips\ action models.
-The compilation approach is appealing by itself because it opens up the door to the bootstrapping of planning action models, but also because:
-
+The compilation approach is appealing by itself because it opens up the door to the bootstrapping of planning action models, but also because it is flexible to different amount and kind of available input knowledge:
 \begin{enumerate}
-\item It is flexible to various amounts of input knowledge.
-Learning examples range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just a pair of initial and final states where no intermediate state/action is observed.
-\item It accepts previous knowledge about the structure of the actions in the form of partially specified action models.
-In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if this model is not fully specified.
+\item Learning examples can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just a pair of initial and final states where no intermediate state/action is observed.
+\item Partially specified action models expressing {\em a priori} knowledge about the structure of actions can also be provided to the compilation.
+In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if the model is not fully specified or the observation is incomplete.
 \end{enumerate}
@@
 The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
-Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$.
-And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\neg\eff(a))\cup\eff(a)\}$, with $\neg\eff(a) = \{\neg l : l \in \eff(a)\}$.
+Therefore $\rho(s,a)$ holds iff $\pre(a)\subseteq s$ and the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\neg\eff(a))\cup\eff(a)\}$, with $\neg\eff(a) = \{\neg l : l \in \eff(a)\}$.
 
@@
 A plan $\pi$ solves $P$ if the induced trajectory $\tau(\pi,P)$ holds that $G \subseteq s_n$.
-A plan is optimal iff its length is minimal.
+A solution plan is {\em optimal} iff its length is minimal.
 
@@
 Given a planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$, the \emph{observation of the trajectory} $\tau(\pi,P)$ is a sequence of partial states that captures what is observed from the execution of $\pi$ in $P$.
-Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, $s_0^o=I$ is a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$, which is fully observable.
+Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, $s_0^o=I$ is a sequence of possibly {\em partially observable states} (except for the initial state $s_0^o$ which is fully observable).
 A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
@@
 $a_i^o\in s_i^o$ ({\small $0\leq i <m$}).
-The number of {\em observed actions}, $l$, can then range from $0$ (fully unobservable action sequence) to $|\pi|$ (fully observable action sequence).
+The number of {\em observed actions}, $l$, can then range from $0$ (in a fully unobservable action sequence) to $|\pi|$ (a fully observable action sequence).
 
-Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be {\em consistent} with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
+The sequence of observed states $\mathcal{O}(\tau)$ must be {\em consistent} with the sequence of states in $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+In practice, the number $m$ of observed states ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
 This means that we assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
 Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
+In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of the trajectory $\tau$.
 
-Last but not least, our observation model can distinguish between {\em observable state variables}, whose value may be read from sensors, and {\em hidden} (or {\em latent}) {\em state variables}, that cannot be observed.
+This observation model can also distinguish between {\em observable state variables}, whose value may be read from sensors, and {\em hidden} (or {\em latent}) {\em state variables}, that cannot be observed.
 Given a subset of fluents $\Gamma\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Gamma$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Gamma$.
@@
 \subsection{The explanation model}
-Let $\Phi_\mathcal{O}=\tup{F,A}$ be a {\em classical planning frame} where the fluents $F$ are deducible from $\mathcal{O}$, since $s_0^o$ is fully observed.
+Let $\Phi_\mathcal{O}=\tup{F,A}$ be a {\em classical planning frame} where the fluents $F$ are deduced from the fully observed $s_0^o\in \mathcal{O}$.
 Let also $P_\mathcal{O}$ be the classical planning problem $P_\mathcal{O}=\tup{F,A,s_0^o,s_m^o}$.
@@
 We say that a plan $\pi$ {\em explains} $\mathcal{O}$ (denoted $\pi\mapsto\mathcal{O}$) iff $\pi$ is a solution for $P$ that is {\em consistent} with the state trajectory constraints imposed by the sequence of partial states $\mathcal{O}$.
+If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}(\tau)$.
 \end{definition}
 
-If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}(\tau)$.
-
-With regard to this definition, an {\em observation} can be considered a sequence of ordered {\em landmarks} for the $P_\mathcal{O}$ classical planning problem, because the literals in the observatoin must be achieved by all the plans that solve this problem in the same order~\cite{hoffmann2004ordered}.
+An {\em observation} can then be considered a sequence of ordered {\em landmarks} for the $P_\mathcal{O}$ classical planning problem, because the literals in the observation must be achieved by all the plans that solve this problem and in the same order~\cite{hoffmann2004ordered}.
 
@@
 
-Now we are ready to define a {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\kappa}$ as a model $\mathcal{M}'$ s.t.
-$\mathcal{M}'$ explains $\mathcal{O}$ and it is consistent with the headers of $\mathcal{M}$ and the given domain knowledge in $\kappa$.
+A {\em solution} to a $\Lambda=\tup{\mathcal{M},{\mathcal O},\kappa}$ learning task is a model $\mathcal{M}'$ s.t.
+it is consistent with the headers of $\mathcal{M}$ and the given domain knowledge in $\kappa$ and it explains $\mathcal{O}$.
 We say that a model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}$.
@@
 
-Let $\Psi$ be the set of {\em predicates} (the initial state of an observation is a full assignment of values to fluents, $|s_0|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0$) that shape the variables $F$.
+Let $\Psi$ be the set of {\em predicates} that shape the fluents $F$ (the initial state of an observation is a full assignment of values to fluents, $|s_0^o|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0^o$).
 The set of propositions that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
-For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), \\
-on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi$.
+For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+Hence, solving a $\Lambda=\tup{\mathcal{M},{\mathcal O},\kappa}$ learning task is determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi\in \mathcal{M}$.
 
-In principle, for a given action model $\xi$, any element of ${\mathcal I}_{\xi,\Psi}$ can potentially appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
+In principle, for a given \strips\ action model $\xi$, any element of ${\mathcal I}_{\xi,\Psi}$ can potentially appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$.
 In practice, the actual space of possible \strips\ schemata is bounded by:
@@
 \subsection{The sampling space}
-According to our observation model the minimal expression of an observation must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed final state $s_m^o$ so $m \geq 1$.
-The maximal expression of an observation corresponds to a fully observed trajectory (where all traversed states and applied actions are fully observed).
-In between there is a grey scale of  different kind of observations including the observation of the initial state and a plan that is frequently used for previous system that learn planning action models such as {\sc ARMS}~\cite{yang2007learning} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
+According to our {\em observation model} the minimal expression of an observation must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed final state $s_m^o$ so $m \geq 1$.
+The maximal expression of an observation corresponds to a fully observed trajectory $\mathcal{O}(\tau)=\tau$ (meaning that all traversed states and applied actions are fully observed).
+In between there is a grey scale of different kind of possible observations, including the observation of the initial state and a plan that is frequently used for previous system that learn planning action models such as {\sc ARMS}~\cite{yang2007learning} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
 
@@
   \end{small}
-	\caption{\small Example of a two-state observationn for the learning \strips\ action models.}
+	\caption{\small Example of a two-state observation for the learning of \strips\ action models in the {\em blocksworld} domain.}
 	\label{fig:observation}
@@
 For instance, back to the {\em blocksworld} domain, one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of any action model $\xi$ because, in this specific domain, a block cannot be on top of itself.
-In addition, we can use {\em domain-specific knowledge} to complete partially observed states.
+In addition, {\em domain-specific knowledge} can be used to complete partially observed states.
 
-{\em State invariants} is a useful type of state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+A compact formalism for expressing domain-specific knowledge are {\em schematic invariants}.
+For instance {\em schematic invariants} can identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} (Figure~\ref{fig:strongest-invariant} shows an example of four clauses that define schematic invariants for the {\em blocksworld} domain).
+%{\em State invariants} is a useful type of state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
 Given a planning problem $P=\tup{F,A,I,G}$, a state invariant is a formula $\phi$ that holds in $I$, $I\models \phi$, and in every state $s$ built out of $F$ that is reachable by applying actions of $A$ in $I$.
@@
 Invariant templates obtained by inspecting the lifted representation of the domain have also been exploited for deriving \emph{lifted mutex}~\cite{BernardiniFS18}.
-{\em Schematic invariants} can identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998}.
-For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible state.
-Figure~\ref{fig:strongest-invariant} shows an example of four clauses that define schematic mutexes for the {\em blocksworld} domain.
 
@@
 \subsection{Compilation}
-Our compilation scheme builds upon the approach presented in \cite{aineto2018learning} but \FAMA comes up with a more general and flexible scheme able to capture any type of input plan trace.
+Our compilation scheme builds upon the approach presented in \cite{aineto2018learning} but {\sc FAMA} comes up with a more general and flexible scheme able to capture any type of input plan trace.
 
@@
 In some contexts, it is reasonable to assume that some parts of the action model are known and so there is no need to learn the entire model from scratch \cite{ZhuoNK13}.
-In \FAMA, when an action model $\xi$ is partially specified, the known preconditions and effects are encoded as fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$ set to true in the initial state $I_{\Lambda}$.
+In {\sc FAMA}, when an action model $\xi$ is partially specified, the known preconditions and effects are encoded as fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$ set to true in the initial state $I_{\Lambda}$.
 In this case, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, thereby making the classical planning task $P_{\Lambda}$ easier to be solved.

### 2019-03-05 doc/icaps-19/camera-ready/model-recognition.tex
ICAPS19: added Acks

@@
 
+The $P(M)$ probability expresses whether one model is known to be a priori more likely than the others.
+When this probability is not given as input, we can reasonable assume that, a priori, all models are equiprobable.
+This is precisely what distance-based model edits systems assume, that all the input models are equiprobable, because this kind of prior knowledge is not exploitable when using only distances as classification metric.
+Consequently, a probability-based formulation allows for a more general definition of a Model Recognition framework.
+
 \subsection{Formulating the $P(\mathcal{O}|\mathcal{M})$ likelihood}
 
-The $P(\mathcal{M})$ probability expresses whether one model is known to be a priori more likely than the others.
-If this probability is not given as input it is reasonable to assume that, {\em a priori}, all models are equiprobable.
 The challenge of our formulation to the model recognition task is the definition of $P(\mathcal{O}|\mathcal{M})$, the likelihood that expresses the probability of observing $\mathcal{O}(\tau)$ when $\mathcal{M}$ is the planning model.
@@
 
-We know from {\em pattern recognition}~\cite{devroye2013probabilistic} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}.
-Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
-\begin{align}
-CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d (1-p)^{N-d}
-\end{align}
-where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than not applying it.
+We are modeling the editing of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.
+Using a Bernoulli process implies that editions are uniformly random and independent \cite{devroye2013probabilistic}.
 
@@
 \end{align}
-where $d=\delta(\mathcal{M},\mathcal{M}')$.
+where $d=\delta(\mathcal{M},\mathcal{M}')$, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than not applying it.
+This means that the lower the value of $d$, the closer a model M' is to the original model M and so the more likely it is.
+Note also that all models at a same distance $d$ will be assigned the same $P(\mathcal{M'}|\mathcal{M})$ by this equation, which means that this probability only depends on the two given models and is "blind" to the observation.
+
+The $P(\mathcal{M^*}|\mathcal{M})$ probability is then given by equation (4), and computing the distance from $\mathcal{M}$ to the {\em closest consistents models} $\mathcal{M^*}$ is given by the {\em observation distance}, $d=\delta^o(\mathcal{M},\mathcal{O})$.
+
+%We know from {\em pattern recognition}~\cite{devroye2013probabilistic} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}.
+Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
+%\begin{align}
+%CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d (1-p)^{N-d}
+%\end{align}
+%where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than not applying it.
+%
+%With this regard, and considering that \strips\ models $\mathcal{M}\in M$ can be encoded with a propositional representation of fixed length $N$, we formulate the $P(\mathcal{M'}|\mathcal{M})$ probability distribution mapping the distance $\delta(\mathcal{M},\mathcal{M'})$ according to equation:
+%\begin{align}
+%P(\mathcal{M'}|\mathcal{M}) = p^d  (1-p)^{N-d}
+%\end{align}
+%where $d=\delta(\mathcal{M},\mathcal{M}')$.
 Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation (5) but instead, the parameter $d$ is given by the {\em observation distance}, $d=\delta^o(\mathcal{M},\mathcal{O})$.
-
-In other words, we are modeling the editing of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.
+%
+%In other words, we are modeling the editing of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.
 
@@
 
-This section shows that $\delta^o(\mathcal{M},\mathcal{O})$, and hence an approximation to $P(\mathcal{O}|\mathcal{M})$), can be computed with a compilation-to-planning approach as the one proposed in~\cite{aineto2018learning} (AJO approach hereafter) for learning \strips\ models.
+This section shows that $\delta^o(\mathcal{M},\mathcal{O})$, and hence an approximation to $P(\mathcal{O}|\mathcal{M})$, can be computed with a compilation-to-planning approach as the one proposed in~\cite{aineto2018learning} (AJO approach hereafter) for learning \strips\ models.
 The AJO approach receives as input an empty model $\mathcal{M}$, which only contains the headers of the action schemata formed of $\xi=\tup{name(\xi),pars(\xi)}$, and an observation of a plan execution $\mathcal{O}(\tau)$ (extensible to a set of observations), and it returns a model $\mathcal{M'}$ with specification of preconditions and effects of each action schema included in $\mathcal{M}$ such that the validation of $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$ following $\mathcal{M'}$ is successful; i.e., it holds $\rho(s_{i-1},a_i)$ for every observed action of $\mathcal{O}(\tau)$ and $s_i=\theta(s_{i-1},a_i)$ for every observed state of $\mathcal{O}(\tau)$.
 
-Essentially, the task $\tup{\mathcal{M},\mathcal{O}}$ of the AJO approach -- learning $\mathcal{M}$ that satisfies $\mathcal{O}$-- is interpreted as editing the empty action schemata of $\mathcal{M}$ introducing preconditions and effects until $\mathcal{O}$ is validated with the resulting model.
-We leverage the same idea to compute $\delta^o(\mathcal{M},\mathcal{O})$ with the exception that now our $\mathcal{M}$ is not empty but $\mathcal{M}=\{\xi_1, \ldots, \xi_m\}$, where $\xi_i=\tup{name(\xi_i),pars(\xi_i),pre(\xi_i),add(\xi_i),del(\xi_i)}$, $1 \leq i \leq m$.
+Essentially, the task $\tup{\mathcal{M},\mathcal{O}}$ of the AJO approach can be used for editing the empty action schemata of $\mathcal{M}$ introducing preconditions and effects until $\mathcal{O}$ is validated with the resulting model.
+We leverage the same idea to compute $\delta^o(\mathcal{M},\mathcal{O})$ with the exception that now our $\mathcal{M}$ is not empty but $\mathcal{M}=\{\xi_1, \ldots, \xi_n\}$, where $\xi_i=\tup{name(\xi_i),pars(\xi_i),pre(\xi_i),add(\xi_i),del(\xi_i)}$, $1 \leq i \leq m$.
 
-Editing/learning the action schemata of $\mathcal{M}$ in the AJO approach is addressed converting the task into a classical planning problem, which is later solved with a planner.
+Editing the action schemata of $\mathcal{M}$ in the AJO approach is addressed converting the task into a classical planning problem, which is later solved with a planner.
 The intuition behind this compilation is that a solution plan to the problem is a sequence of: (a) \emph{edit actions} on the schemata of $\mathcal{M}$ to build $\mathcal{M'}$ and (b) \emph{validate actions} that apply $\mathcal{M'}$ in $\mathcal{O(\tau)}$.
@@
 
-\newpage
+%\newpage
+
+\subsection*{Acknowledgement}
+This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.
+Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it RYC15/18009}, both programs funded by the Spanish government.
 

### 2019-03-06 doc/lecture-notes-EBL/ebl.tex
updated sections 1,2 and 3 for lecture notes on ebl

@@
 \begin{abstract}
-The paper presents the classical planning compilation for learning \strips\ action models from observations of plan executions.
+The paper presents the first classical planning compilation for learning \strips\ action models from partial observations of plan executions.
 The compilation is flexible to different amount and kind of available input knowledge; learning examples can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just pairs of initial and final states where no intermediate action/state is observed.
-The compilation accepts also partially specified action models and it can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if this model is not fully specified or the input observation is incomplete.
+The compilation accepts also partially specified action models and it can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if the given model or the given observation are incomplete.
 \end{abstract}
@@
 Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-In both tasks, an automated planner is required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
+In both tasks, off-the-shelf automated planners are required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
 Unfortunately building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
 
-{\em Machine Learning} (ML) has shown to be able to compute a wide range of different kinds of models from examples~\cite{michalski2013machine}.
+{\em Machine Learning} (ML) has shown to be able to induce a wide range of different kinds of models from examples~\cite{michalski2013machine}.
 The application of inductive ML to learning \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
@@
 \item The {\em input} to ML algorithms (the learning/training data) is usually a finite vector that represents the value of some fixed object features.
-The input for learning planning action models is, however, observations of plan executions (where each plan possibly has a different length).
+The input for learning planning action models is, however, observations of plan executions (where each plan possibly has a different length and plan length is not {\em a priori} bound).
 \item The {\em output} of ML algorithms is usually a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
-When learning action models the output is, for each action, the preconditions and effects that define the possible state transitions of a planning task.
+When learning action models the output is, for each action, the set of preconditions and effects that define the possible state transitions of a planning task.
 \end{itemize}
@@
 Learning \strips\ action models is a well-studied problem with sophisticated algorithms such as {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}, which do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2016hierarchical,segovia2017generating}, this paper describes an innovative planning compilation approach for learning \strips\ action models.
-The compilation approach is appealing by itself because it opens up the door to the bootstrapping of planning action models, but also because it is flexible to different amount and kind of available input knowledge:
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2017generating,segovia2018computing,segovia2019computing}, this paper describes an innovative planning compilation approach for learning \strips\ action models.
+The compilation approach is appealing by itself, because it opens up the door to the bootstrapping of planning action models, but also because it is flexible to different amount and kind of available input knowledge:
 \begin{enumerate}
-\item Learning examples can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just a pair of initial and final states where no intermediate state/action is observed.
-\item Partially specified action models expressing {\em a priori} knowledge about the structure of actions can also be provided to the compilation.
+\item {\em Learning examples} can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just a pair of initial and final states where no intermediate state/action is observed.
+\item {\em Partially specified action models} expressing {\em a priori} knowledge about the structure of actions can also be provided to the compilation.
 In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if the model is not fully specified or the observation is incomplete.
@@
 \section{Background}
-This section formalizes the models for {\em classical planning}, for the {\em observation} of the execution of a classical plan and for the explanation of an observation.
+This section formalizes the models we follow for {\em classical planning}, for the {\em observations} of executions of classical plans and for the explanation of a given observation.
 
@@
 
-Now we define actions with conditional effects because they allow to compactly define our compilation.
+Now we define actions with conditional effects because they allow us to compactly define our compilation.
 An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
@@
 Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, $s_0^o=I$ is a sequence of possibly {\em partially observable states} (except for the initial state $s_0^o$ which is fully observable).
-A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
+A partially observable state is one in which $|s_i^o| < |F|$, {\small $1\leq i\leq n$}; i.e., a state in which at least a fluent of $F$ is not observable.
 
-This {\em observation model} comprises the case $|s_i^o| = 0$, when an intermediate state is fully unobservable.
-Further, this model consider also {\em observed actions} as fluents that indicate the action applied in a given state.
+The {\em observation model} comprises the case $|s_i^o| = 0$, when an intermediate state is fully unobservable.
+This model consider also {\em observed actions} as fluents that indicate the action applied in a given state.
 This means that a sequence of {\em observed actions} $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$ s.t.
 $a_i^o\in s_i^o$ ({\small $0\leq i <m$}).
-The number of {\em observed actions}, $l$, can then range from $0$ (in a fully unobservable action sequence) to $|\pi|$ (a fully observable action sequence).
+The number of {\em observed actions}, $l$, can then range from $0$ (in a fully unobservable action sequence) to $|\pi|$ (in a fully observed action sequence).
 
-The sequence of observed states $\mathcal{O}(\tau)$ must be {\em consistent} with the sequence of states in $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+The sequence of observed states $\mathcal{O}(\tau)$ must be {\em consistent} with the sequence of states in $\tau(\pi,P)$.
 In practice, the number $m$ of observed states ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
@@
 
-\subsection{The explanation model}
-Let $\Phi_\mathcal{O}=\tup{F,A}$ be a {\em classical planning frame} where the fluents $F$ are deduced from the fully observed $s_0^o\in \mathcal{O}$.
-Let also $P_\mathcal{O}$ be the classical planning problem $P_\mathcal{O}=\tup{F,A,s_0^o,s_m^o}$.
+\subsection{Explaining observations with classical planning}
+Given a {\em classical planning frame} $\Phi=\tup{F,A}$ and a observation of the execution of a plan within the given planning frame $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$, then $P_\mathcal{O}$ is a classical planning problem that is built as follows $P_\mathcal{O}=\tup{F,A,s_0^o,s_m^o}$.
 
@@
 
-An {\em observation} can then be considered a sequence of ordered {\em landmarks} for the $P_\mathcal{O}$ classical planning problem, because the literals in the observation must be achieved by all the plans that solve this problem and in the same order~\cite{hoffmann2004ordered}.
+The {\em observation} $\mathcal{O}$ is then considered a sequence of ordered {\em landmarks} for the $P_\mathcal{O}$ classical planning problem, because all the literals in the observation must be achieved by any plan that solves $P_\mathcal{O}$ and in the same order as are defined in the observation~\cite{hoffmann2004ordered}.
 
-\section{Explanation-based learning of action models}
-The task of learning action models from the observation of a plan execution is defined as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O},\kappa}$, where:
+\section{Explanation-based learning of \strips\ action models}
+The task of learning action models by explaining the observation of a plan execution is defined as a tuple $\Lambda=\tup{\mathcal{M},{\mathcal O}}$, where:
 
@@
 \item $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$ is a sequence of partially observed states.
-\item $\kappa$ is a set of logic formulae that define {\em domain-specific knowledge}.
 \end{itemize}
 
-A {\em solution} to a $\Lambda=\tup{\mathcal{M},{\mathcal O},\kappa}$ learning task is a model $\mathcal{M}'$ s.t.
-it is consistent with the headers of $\mathcal{M}$ and the given domain knowledge in $\kappa$ and it explains $\mathcal{O}$.
-We say that a model $\mathcal{M}$ {\em explains} an observation $\mathcal{O}$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by $\mathcal{M}$, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}$.
+A {\em solution} to a $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task is a model $\mathcal{M}'$ s.t.
+it is consistent with the headers of $\mathcal{M}$ and that it explains $\mathcal{O}$.
+We say that a model {\em explains} an observation $\mathcal{O}$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by that model, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}$.
 
 \subsection{The space of \strips\ action models}
-We analyze here the solution space of a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$; i.e., the space of \strips\ action models.
+We analyze here the solution space of the addressed learning task; i.e., the space of \strips\ action models.
 
-A \strips\ \emph{action model} $\xi$ is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
+A \strips\ \emph{action model} is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
 
@@
 For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-Hence, solving a $\Lambda=\tup{\mathcal{M},{\mathcal O},\kappa}$ learning task is determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi\in \mathcal{M}$.
+Hence, solving a $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task is determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi\in \mathcal{M}$.
 
@@
 \item {\bf Observation constraints}.
-The solution $\mathcal{M}'$ must be consistent with these \emph{semantic constraints} derived from  the learning samples $\mathcal{O}$, which in our case is a single plan observation.
+The solution $\mathcal{M}'$ must be consistent with these \emph{semantic constraints} derived from the input observation $\mathcal{O}$.
 Specifically, the states induced by the plan computable with $\mathcal{M}'$ must comprise the observed states of the sample, which further constrains the space of possible action models.
@@
 	\begin{footnotesize}
-		\begin{tabular}{c | c}
-			& {\bf Meaning}\\\hline
+		\begin{tabular}{c@{\hskip .3in}  c}
+	{\bf Encoding} & {\bf Meaning}\\\hline
 $\neg pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ belongs neither to the preconditions nor effects of $\xi$ \\
@@
 According to our {\em observation model} the minimal expression of an observation must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed final state $s_m^o$ so $m \geq 1$.
-The maximal expression of an observation corresponds to a fully observed trajectory $\mathcal{O}(\tau)=\tau$ (meaning that all traversed states and applied actions are fully observed).
-In between there is a grey scale of different kind of possible observations, including the observation of the initial state and a plan that is frequently used for previous system that learn planning action models such as {\sc ARMS}~\cite{yang2007learning} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
+On the other hand, the maximal expression of an observation corresponds to a fully observed trajectory $\mathcal{O}(\tau)=\tau$ (meaning that all traversed states and applied actions are fully observed).
+In between there is a grey scale of different kinds of possible observations, including the observation of the initial state and the executed plan, that is frequently used for previous system that learn planning action models such as {\sc ARMS}~\cite{yang2007learning} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
 
-Figure~\ref{fig:observation} shows a learning example that contains an initial state of the blocksworld where the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
-The observation represents a partially observable final state in which {\tt\small{blockA}} is on top of {\tt\small{blockB}} and {\tt\small{blockB}} on top of {\tt\small{blockC}}.
+Figure~\ref{fig:observation} shows an example of an observation that contains only two states.
+An initial state of the blocksworld where the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
+The observation represents also a partially observable final state in which {\tt\small{blockA}} is on top of {\tt\small{blockB}} and {\tt\small{blockB}} on top of {\tt\small{blockC}}.
 
@@
 	\label{fig:observation}
-\end{figure}
-
-\subsection{The domain-specific knowledge}
-One can introduce domain-specific knowledge to constrain further the space of possible schemata.
-For instance, back to the {\em blocksworld} domain, one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of any action model $\xi$ because, in this specific domain, a block cannot be on top of itself.
-In addition, {\em domain-specific knowledge} can be used to complete partially observed states.
-
-A compact formalism for expressing domain-specific knowledge are {\em schematic invariants}.
-For instance {\em schematic invariants} can identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} (Figure~\ref{fig:strongest-invariant} shows an example of four clauses that define schematic invariants for the {\em blocksworld} domain).
-%{\em State invariants} is a useful type of state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a planning problem $P=\tup{F,A,I,G}$, a state invariant is a formula $\phi$ that holds in $I$, $I\models \phi$, and in every state $s$ built out of $F$ that is reachable by applying actions of $A$ in $I$.
-Recently, some works point at extracting \emph{lifted} invariants, also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}, that hold for any possible state and any possible set of objects.
-Invariant templates obtained by inspecting the lifted representation of the domain have also been exploited for deriving \emph{lifted mutex}~\cite{BernardiniFS18}.
-
-\begin{figure}[hbt!]
-  \begin{footnotesize}
-$\forall x_1,x_2\ \neg ontable(x_1)\vee\neg on(x_1,x_2)$.\\
-$\forall x_1,x_2\ \neg clear(x_1)\vee\neg on(x_2,x_1)$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
-\end{footnotesize}
- \caption{\small {\em Schematic mutexes} for the {\em blocksworld} domain.}
-\label{fig:strongest-invariant}
 \end{figure}

### 2019-03-06 doc/lecture-notes-EBL/ebl.tex
first updates to section 4 of lecture notes on ebl

@@
 \subsection{The space of \strips\ action models}
+\label{sec:strips-space}
 We analyze here the solution space of the addressed learning task; i.e., the space of \strips\ action models.
@@
 \section{Learning \strips\ action models with classical planning}
+Our approach to address a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is to compile $\Lambda$ into a planning task $P_{\Lambda}$.
+The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the observed actions and states in ${\mathcal O}$.
 
-Our proposal to address a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to transform $\Lambda$ into a planning task $P_{\Lambda}$.
-The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$.
-Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition, a positive effect or a negative effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\tau$, including that the final reached state comprises $s_n$.
-Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states of $\tau$.
+A solution plan $\pi_\Lambda$ will hence comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition or a effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned modes while successively \textbf{validating} the effects of the action application in every partial state in ${\mathcal O}$, including that the final reached state comprises $s_n$.
+Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1)}, {\tt\small (insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_holding\_v1)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1)}, {\tt\small (validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states in ${\mathcal O}$.
 
 \subsection{Compilation}
-Our compilation scheme builds upon the approach presented in \cite{aineto2018learning} but {\sc FAMA} comes up with a more general and flexible scheme able to capture any type of input plan trace.
-
-\vspace{0.1cm}
-
-A learning task $\Lambda=\tup{\mathcal{M},\tau}$ is compiled into a planning task $P_{\Lambda}$ with conditional effects in the context of a planning frame $\Phi=\tup{F,A}$.
-We use conditional effects because they allow us to compactly define actions whose effects depend on the current state.
-An action $a\in A$ with conditional effects is defined as a set of preconditions $\pre(a)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a)$.
-Each conditional effect $C\rhd E\in\cond(a)$ is composed of two sets of literals $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a\in A$ is applicable in a state $s$ if and only if $\pre(a)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$; that is, $triggered(s,a)=\bigcup\limits_{C\rhd E\in\cond(a),C\subseteq s} E$.
-The result of applying $a$ in state $s$ follows the same definition of successor state, $\theta(s,a)$, introduced in section \ref{basic_planning} but applied to the conditional effects in $triggered(s,a)$.
-
-\vspace{0.25cm}
-
-A solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output domain model $\mathcal{M}'$ that solves the learning task $\Lambda$. Specifically, a solution plan $\pi_\Lambda$ serves two purposes:
+A learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is compiled into a classical planning problem $P_{\Lambda}$ with conditional effects.
+A solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output domain model $\mathcal{M}'$ that solves the learning task $\Lambda$.
+Specifically, a solution plan $\pi_\Lambda$ serves two purposes:
 
@@
 \item {\bf To validate the action models of $\mathcal{M}'$}.
-$\pi_\Lambda$ also comprises a second block of actions (plan {\em postfix}) which is aimed at validating of the observed plan trace $\tau$ with the built action models $\mathcal{M}'$.
+$\pi_\Lambda$ also comprises a second block of actions (plan {\em postfix}) which is aimed at validating the observation ${\mathcal O}$ with the built action models $\mathcal{M}'$.
 \end{enumerate}
 
-%\subsection{Learning from observations of plan executions}
-Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$, with $\tau$ formed by an $n$-action sequence $\tup{a_1, \ldots, a_n}$ and a $m$-state trajectory $\tup{s_0, s_1, \ldots, s_m}$ ($\tau = \langle s_0, a_1, \ldots, a_n, s_m \rangle$), the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
-
+Given a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
 \begin{itemize}
-
-\item $F_{\Lambda}$ extends $F$ with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of $\tau$.
+\item $F_{\Lambda}$ extends $F$ with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of ${\mathcal O}$.
 Specifically, $F_{\Lambda}$ contains:
 \begin{itemize}
-\item The set of fluents obtained from $s_0$; i.e., $F$.
-\item The model fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section \ref{propositional_encoding}
-\item A set of fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to represent the $i^{th}$ observable action of $\tau$.
-In the example of Figure~\ref{fig:example-plans}, the two observed actions {\small \texttt{(putdown B)}} and {\small \texttt{(stack  A  B)}} would be encoded as fluents  {\small \texttt{(plan-putdown B i1)}} and {\small \texttt{(plan-stack A B i2)}} to indicate that {\small \texttt{(putdown B)}} is observed in the first place and {\small \texttt{(stack  A  B)}} is the second observed action.
-\item Two fluents, $at_i$ and $next_{i,i+1}$, {\small $1\leq i \leq n$}, to iterate through the $n$ observed actions of $\tau$.
-The former is used to ensure that actions are executed in the same order as they are observed in $\tau$.
-The latter is used to iterate to the next planning step when solving $P_{\Lambda}$.
-\item A set of fluents $\{test_j\}_{0\leq j\leq m}$, to point at the state observation $s_j\in\tau$ where the action model is
-validated. In the example of Figure~\ref{fig:example-plans} two tests are required to validate the programmed action model, one test at $s_0$ and another one at $s_4$.
+\item The set of fluents obtained from $s_0^o$; i.e., $F$.
+\item The model fluents $pre_p(\xi)$ and $eff_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section~\ref{sec:strips-space}.
+\item A set of fluents $\{test_j\}_{0\leq j\leq m}$, to point at the state observation $s_j^o\in {\mathcal O}$ where the action model is validated.
+In the example of Figure~\ref{fig:observation} two tests are required to validate the programmed action model, one corresponding to the initial state and the second one corresponding to the final state.
 \item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated.
@@
 
-\item $I_{\Lambda}$ encodes $s_0$ and the following fluents set to true: $mode_{prog}$, $test_0$, $F_{\pi}$, $at_1$ and $\{next_{i,i+1}\}$, {\small $1\leq i \leq n$}.
+\item $I_{\Lambda}$ encodes $s_0$ and the following fluents set to true: $mode_{prog}$, $test_0$.
 Our compilation assumes that action models are initially programmed with no precondition, no negative effect and no positive effect.
 
-\item $G_{\Lambda}$ includes the positive literals $at_n$ and $test_m$.
-When these two goals are achieved by the solution plan $\pi_\Lambda$, we will be certain that the action models of $\mathcal{M'}$ are validated in all the actions and states observed in the input plan trace $\tau$.
+\item $G_{\Lambda}$ includes the positive literal $test_m$.
+When this goal is achieved by the solution plan $\pi_\Lambda$, we will be certain that the action models of $\mathcal{M'}$ are validated in the input observation.
 

### 2019-03-06 doc/lecture-notes-EBL/ebl.tex
updates to section 4 of lecture notes on ebl

@@
 \section{Learning \strips\ action models with classical planning}
-Our approach to address a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is to compile $\Lambda$ into a planning task $P_{\Lambda}$.
-The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the observed actions and states in ${\mathcal O}$.
+Our approach to address a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is to compile it into a planning task $P_{\Lambda}$.
+The intuition behind the compilation is that when $P_{\Lambda}$ is solved, the solution plan $\pi_\Lambda$ is a sequence of actions that build the output model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the observation ${\mathcal O}$.
 
-A solution plan $\pi_\Lambda$ will hence comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition or a effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned modes while successively \textbf{validating} the effects of the action application in every partial state in ${\mathcal O}$, including that the final reached state comprises $s_n$.
-Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1)}, {\tt\small (insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_holding\_v1)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1)}, {\tt\small (validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states in ${\mathcal O}$.
+A solution plan $\pi_\Lambda$ comprise then two differentiated blocks of actions: a plan prefix with a set of actions each defining the \textbf{insertion} of a fluent as a precondition or a effect of an action model $\xi \in \mathcal{M'}$; and a plan postfix with a set of actions that determine the \textbf{application} of the learned modes while successively \textbf{validating} the effects of the action application in every partial state in ${\mathcal O}$.
+Roughly speaking, in the \emph{blocksworld}, the format of the first set of actions of $\pi_\Lambda$ looks like {\tt{\small (insert\_pre\_stack\_holding\_v1)}, {\tt\small (insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_holding\_v1)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ is like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1)}, {\tt\small (validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states in ${\mathcal O}$.
 
 \subsection{Compilation}
-A learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is compiled into a classical planning problem $P_{\Lambda}$ with conditional effects.
+A learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is compiled into a classical planning problem with conditional effects $P_{\Lambda}$.
 A solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output domain model $\mathcal{M}'$ that solves the learning task $\Lambda$.
@@
 
-\item $I_{\Lambda}$ encodes $s_0$ and the following fluents set to true: $mode_{prog}$, $test_0$.
+\item $I_{\Lambda}$ encodes $s_0^o$ and the following fluents set to true: $mode_{prog}$, $test_0$.
 Our compilation assumes that action models are initially programmed with no precondition, no negative effect and no positive effect.
@@
 \begin{enumerate}
-\item Actions for {\em inserting} a component (precondition, positive effect or negative effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models.
+\item Actions for {\em inserting} a component (precondition or effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models.
 These actions will form the prefix of the solution plan $\pi_\Lambda$.
@@
 \item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
-A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $del_p$ nor $add_p$ exist in $\xi$.
+A precondition $p$ is inserted in $\xi$ when neither $pre_p$ nor $eff_p$ exist in $\xi$.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi), mode_{prog}\},\\
 \cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.

### 2019-03-06 doc/lecture-notes-EBL/ebl.tex
updated sections 1,2,3,4 and 6 lecture notes on ebl

@@
 \usepackage[numbers]{natbib}
+\usepackage{amsthm}
 \usepackage{amsmath}
@@
 Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
-The belonging of an $e \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre\_\xi\_e$ and $eff\_\xi\_e$, instead of the three fluents used in the original compilation~\cite{aineto2018learning}.
+The belonging of a $p \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre_{p,\xi}$ and $eff_{p,\xi}$, instead of the three fluents used in the original compilation~\cite{aineto2018learning}.
 The four possible combinations of these two fluents are sumarized in Figure \ref{fig:combinations}.
@@
 	\begin{footnotesize}
-		\begin{tabular}{c@{\hskip .3in}  c}
+		\begin{tabular}{c@{\hskip .2in} |@{\hskip .1in} c}
 	{\bf Encoding} & {\bf Meaning}\\\hline
-$\neg pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ belongs neither to the preconditions nor effects of $\xi$ \\
-             & ($e \notin pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$)\\\\
-$pre\_\xi\_e \wedge \neg eff\_\xi\_e $& $e$ is only a precondition of $\xi$\\
-               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \notin del(\xi)$) \\\\
-$\neg pre\_\xi\_e \wedge eff\_\xi\_e $& $e$ is a positive effect of $\xi$ \\
-               &  ($e \notin pre(\xi) \wedge e \in add(\xi) \wedge e \notin del(\xi)$) \\\\
-$pre\_\xi\_e \wedge eff\_\xi\_e  $& $e$ is a negative effect of $\xi$ \\
-               &  ($e \in pre(\xi) \wedge e \notin add(\xi) \wedge e \in add(\xi)$)
+$\neg pre_{p,\xi} \wedge \neg eff_{p,\xi} $& $p$ belongs neither to the preconditions nor effects of $\xi$ \\
+             & ($p \notin pre(\xi) \wedge p \notin add(\xi) \wedge p \notin del(\xi)$)\\\\
+$pre_{p,\xi} \wedge \neg eff_{p,\xi} $& $p$ is only a precondition of $\xi$\\
+               &  ($p \in pre(\xi) \wedge p \notin add(\xi) \wedge p \notin del(\xi)$) \\\\
+$\neg pre_{p,\xi} \wedge eff_{p,\xi} $& $p$ is a positive effect of $\xi$ \\
+               &  ($p \notin pre(\xi) \wedge p \in add(\xi) \wedge p \notin del(\xi)$) \\\\
+$pre_{p,\xi} \wedge eff_{p,\xi} $& $p$ is a negative effect of $\xi$ \\
+               &  ($p \in pre(\xi) \wedge p \notin add(\xi) \wedge p \in add(\xi)$)
 		\end{tabular}
@@
 	\label{fig:combinations}
+\end{figure}
+
+To illustrate better this encoding, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
+
+\begin{figure}
+  \begin{small}  
+  \begin{verbatim}
+(:action stack
+   :parameters (?v1 ?v2)
+   :precondition (and (holding ?v1) (clear ?v2))
+   :effect (and (not (holding ?v1)) (not (clear ?v2)) (clear ?v1) (handempty) (on ?v1 ?v2)))
+
+(pre_holding_v1_stack) (pre_clear_v2_stack)
+(eff_holding_v1_stack) (eff_clear_v2_stack)
+(eff_clear_v1_stack) (eff_handempty_stack) (eff_on_v1_v2_stack)
+  \end{verbatim}           
+  \end{small}  
+ \caption{\small PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.}
+\label{fig:propositional}
 \end{figure}
@@
 \item {\bf To validate the action models of $\mathcal{M}'$}.
-$\pi_\Lambda$ also comprises a second block of actions (plan {\em postfix}) which is aimed at validating the observation ${\mathcal O}$ with the built action models $\mathcal{M}'$.
+$\pi_\Lambda$ also comprises a second block of actions (plan {\em postfix}) which is aimed at {\em explaining} the observation ${\mathcal O}$ with the built action models $\mathcal{M}'$.
 \end{enumerate}
@@
 \item The set of fluents obtained from $s_0^o$; i.e., $F$.
-\item The model fluents $pre_p(\xi)$ and $eff_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section~\ref{sec:strips-space}.
+\item The model fluents $pre_{p,\xi}$ and $eff_{p,\xi}$, for every $p\in \Psi_{\xi}$ and $\xi \in \mathcal{M}$, defined as explained in section~\ref{sec:strips-space}.
 \item A set of fluents $\{test_j\}_{0\leq j\leq m}$, to point at the state observation $s_j^o\in {\mathcal O}$ where the action model is validated.
 In the example of Figure~\ref{fig:observation} two tests are required to validate the programmed action model, one corresponding to the initial state and the second one corresponding to the final state.
-\item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated.
+\item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated and a fluent {\small\tt invalid} to indicate that the programmed action model is inconsistent with the input observation.
 \end{itemize}
@@
 
-\item $G_{\Lambda}$ includes the positive literal $test_m$.
-When this goal is achieved by the solution plan $\pi_\Lambda$, we will be certain that the action models of $\mathcal{M'}$ are validated in the input observation.
+\item $G_{\Lambda}$ includes the positive literal $test_m$ and the negative literal $\neg${\small\tt invalid}.
+When these goals are achieved by the solution plan $\pi_\Lambda$, we will be certain that the action models of $\mathcal{M'}$ are validated in the input observation.
 
@@
 A precondition $p$ is inserted in $\xi$ when neither $pre_p$ nor $eff_p$ exist in $\xi$.
-
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p}(\xi), \neg eff_{p}(\xi), mode_{prog}\},\\
-\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p,\xi}, \neg eff_{p,\xi}, mode_{prog}\},\\
+\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p,\xi}\}.
 \end{align*}
@@
 
-\item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
-A positive effect is inserted in $\xi$ under the same conditions of a precondition insertion, and a negative effect is inserted in $\xi$ when neither $del_p$ nor $add_p$ appear in $\xi$ but $pre_p$ does.
-
+\item Actions which support the addition of a effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$. 
 \begin{small}
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg del_{p}(\xi),\neg add_{p}(\xi), mode_{prog}\},\\
-\cond(\mathsf{insertEff_{p,\xi}})=&\{pre_{p}(\xi)\}\rhd\{del_{p}(\xi)\},\\
-& \{\neg pre_{p}(\xi)\}\rhd\{add_{p}(\xi)\}.
+\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff_{p,\xi}, mode_{prog}\},\\
+\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff_{p,\xi}\}
 \end{align*}
@@
 
-For instance, given $name(\xi)=${\tt{\small stack}} and $C_{pre-stack}=\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_holding\_v2), (pre\_stack\_on\_v1\_v2),(pre\_stack\_clear\_v1),(pre\_stack\_clear\_v1),}}$\ldots \}$, the insertion of each item $c \in C_{pre-stack}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$ as long as $c \notin pre(\xi)$,  $c \notin add(\xi)$ and $c \notin del(\xi)$.
-The same applies to effects with respect to sets $C_{add-stack}$ and $C_{del-stack}$ that would include all fluents starting with prefix  {\tt{\small add}} and {\tt{\small del}}, respectively.
+For instance, given $name(\xi)=${\tt{\small stack}} and $F_{pre,stack}=\{${\tt{\small (pre\_stack\_holding\_v1),} {\tt\small (pre\_stack\_holding\_v2), (pre\_stack\_on\_v1\_v2),(pre\_stack\_clear\_v1)}, {\tt\small(pre\_stack\_clear\_v1)}}, $\ldots \}$, the insertion of each item $p \in F_{pre,stack}\subseteq F_{\Lambda}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$.
+The same applies to effects $F_{eff,stack}=\{${\tt{\small (eff\_stack\_holding\_v1),(eff\_stack\_holding\_v2), (eff\_stack\_on\_v1\_v2),(eff\_stack\_clear\_v1),(eff\_stack\_clear\_v1),}}$\ldots \}$.
 
@@
 \item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{ar(\xi)}$.
+These actions will be part of the postfix of the plan $\pi_\Lambda$ and they determine the application of the learned action models according to the values of the model fluents in the current state configuration.
 Since action headers are known, the variables $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{pre_{p}(\xi)\implies p(\omega)\}_{\forall p\in\Psi_\xi},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{del_{p}(\xi)\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
-&\{add_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi},\\
+\hspace*{7pt}\pre(\mathsf{apply_{\xi,\omega}})=&\{\},\\
+\cond(\mathsf{apply_{\xi,\omega}})=& \{pre_{p,\xi}\wedge eff_{p,\xi}\}\rhd\{\neg p(\omega)\}_{\forall p\in\Psi_\xi},\\
+& \{\neg pre_{p,\xi}\wedge eff_{p,\xi}\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi},\\
+& \{pre_{p,\xi}\wedge \neg p(\omega)\}_{\forall p\in\Psi_\xi}\rhd\{invalid\},\\
 &\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
@@
 
-    These actions will be part of the postfix of the plan $\pi_\Lambda$ and they determine the application of the learned action models according to the values of the model fluents in the current state configuration}.
 Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator.
-\textcolor[rgb]{1.00,0.00,0.00}{Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$.
+Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$.
 Executing this action in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$.
@@
   :parameters (?o1 - object ?o2 - object)
-  :precondition
-   (and (or (not (pre_stack_on_v1_v1)) (on ?o1 ?o1))
-        (or (not (pre_stack_on_v1_v2)) (on ?o1 ?o2))
-        (or (not (pre_stack_on_v2_v1)) (on ?o2 ?o1))
-        (or (not (pre_stack_on_v2_v2)) (on ?o2 ?o2))
-        (or (not (pre_stack_ontable_v1)) (ontable ?o1))
-        (or (not (pre_stack_ontable_v2)) (ontable ?o2))
-        (or (not (pre_stack_clear_v1)) (clear ?o1))
-        (or (not (pre_stack_clear_v2)) (clear ?o2))
-        (or (not (pre_stack_holding_v1)) (holding ?o1))
-        (or (not (pre_stack_holding_v2)) (holding ?o2))
-        (or (not (pre_stack_handempty)) (handempty)))
-  :effect
-   (and (when (del_stack_on_v1_v1) (not (on ?o1 ?o1)))
-        (when (del_stack_on_v1_v2) (not (on ?o1 ?o2)))
-        (when (del_stack_on_v2_v1) (not (on ?o2 ?o1)))
-        (when (del_stack_on_v2_v2) (not (on ?o2 ?o2)))
-        (when (del_stack_ontable_v1) (not (ontable ?o1)))
-        (when (del_stack_ontable_v2) (not (ontable ?o2)))
-        (when (del_stack_clear_v1) (not (clear ?o1)))
-        (when (del_stack_clear_v2) (not (clear ?o2)))
-        (when (del_stack_holding_v1) (not (holding ?o1)))
-        (when (del_stack_holding_v2) (not (holding ?o2)))
-        (when (del_stack_handempty) (not (handempty)))
-        (when (add_stack_on_v1_v1) (on ?o1 ?o1))
-        (when (add_stack_on_v1_v2) (on ?o1 ?o2))
-        (when (add_stack_on_v2_v1) (on ?o2 ?o1))
-        (when (add_stack_on_v2_v2) (on ?o2 ?o2))
-        (when (add_stack_ontable_v1) (ontable ?o1))
-        (when (add_stack_ontable_v2) (ontable ?o2))
-        (when (add_stack_clear_v1) (clear ?o1))
-        (when (add_stack_clear_v2) (clear ?o2))
-        (when (add_stack_holding_v1) (holding ?o1))
-        (when (add_stack_holding_v2) (holding ?o2))
-        (when (add_stack_handempty) (handempty))
+  :precondition (and ) 
+  :effect (and (when (and (pre_stack_on_v1_v1) (eff_stack_on_v1_v1)) (not (on ?o1 ?o1)))
+               (when (and (pre_stack_on_v1_v2) (eff_stack_on_v1_v2)) (not (on ?o1 ?o2)))
+               (when (and (pre_stack_on_v2_v1) (eff_stack_on_v2_v1)) (not (on ?o2 ?o1)))
+               (when (and (pre_stack_on_v2_v2) (eff_stack_on_v2_v2)) (not (on ?o2 ?o2)))
+               (when (and (pre_stack_ontable_v1) (eff_stack_ontable_v1)) (not (ontable ?o1)))
+               (when (and (pre_stack_ontable_v2) (eff_stack_ontable_v2)) (not (ontable ?o2)))
+               (when (and (pre_stack_clear_v1) (eff_stack_clear_v1)) (not (clear ?o1)))
+               (when (and (pre_stack_clear_v2) (eff_stack_clear_v2)) (not (clear ?o2)))
+               (when (and (pre_stack_holding_v1) (eff_stack_holding_v1)) (not (holding ?o1)))
+               (when (and (pre_stack_holding_v2) (eff_stack_holding_v2)) (not (holding ?o2)))
+               (when (and (pre_stack_handempty) (eff_stack_handempty)) (not (handempty)))
+               (when (and (not (pre_stack_on_v1_v1)) (eff_stack_on_v1_v1)) (on ?o1 ?o1))
+               (when (and (not (pre_stack_on_v1_v2)) (eff_stack_on_v1_v2)) (on ?o1 ?o2))
+               (when (and (not (pre_stack_on_v2_v1)) (eff_stack_on_v2_v1)) (on ?o2 ?o1))
+               (when (and (not (pre_stack_on_v2_v2)) (eff_stack_on_v2_v2)) (on ?o2 ?o2))
+               (when (and (not (pre_stack_ontable_v1)) (eff_stack_ontable_v1)) (ontable ?o1))
+               (when (and (not (pre_stack_ontable_v2)) (eff_stack_ontable_v2)) (ontable ?o2))
+               (when (and (not (pre_stack_clear_v1)) (eff_stack_clear_v1)) (clear ?o1))
+               (when (and (not (pre_stack_clear_v2)) (eff_stack_clear_v2)) (clear ?o2))
+               (when (and (not (pre_stack_holding_v1)) (eff_stack_holding_v1)) (holding ?o1))
+               (when (and (not (pre_stack_holding_v2)) (eff_stack_holding_v2)) (holding ?o2))
+               (when (and (not (pre_stack_handempty)) (eff_stack_handempty)) (handempty))
+               (when (and (pre_stack_on_v1_v1) (not (on ?o1 ?o1))) (invalid))
+               (when (and (pre_stack_on_v1_v2) (not (on ?o1 ?o2))) (invalid))
+               (when (and (pre_stack_on_v2_v1) (not (on ?o2 ?o1))) (invalid))
+               (when (and (pre_stack_on_v2_v2) (not (on ?o2 ?o2))) (invalid))
+               (when (and (pre_stack_ontable_v1) (not (ontable ?o1))) (invalid))
+               (when (and (pre_stack_ontable_v2) (not (ontable ?o2))) (invalid))
+               (when (and (pre_stack_clear_v1) (not (clear ?o1))) (invalid))
+               (when (and (pre_stack_clear_v2) (not (clear ?o2))) (invalid))
+               (when (and (pre_stack_holding_v1) (not (holding ?o1))) (invalid))
+               (when (and (pre_stack_holding_v2) (not (holding ?o2))) (invalid))
+               (when (and (pre_stack_handempty) (not (handempty))) (invalid))
                (when (modeProg) (not (modeProg)))))
@@
 
-\item Actions for {\em validating} the partially observed state $s_j\in\tau$, {\tt\small $1\leq j< m$}.
-These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observable data of the input plan trace $\tau$ follows after the execution of the apply actions.
+\item Actions for {\em validating} the partially observed state $s_j\in\mathcal{O}$, {\tt\small $1\leq j< m$}.
+These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observation $\mathcal{O}$ follows after the execution of the apply actions.
 
@@
 
-There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$.
-The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.
+There will be a validate action in $\pi_\Lambda$ for every observed state in $\mathcal{O}$.
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\mathcal{O}$.
 
@@
 In some contexts, it is reasonable to assume that some parts of the action model are known and so there is no need to learn the entire model from scratch \cite{ZhuoNK13}.
-In {\sc FAMA}, when an action model $\xi$ is partially specified, the known preconditions and effects are encoded as fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$ set to true in the initial state $I_{\Lambda}$.
+In our compilation approach, when an action model $\xi$ is partially specified, the known preconditions and effects are encoded as fluents $pre_{p,\xi}$ and $eff_{p,\xi}$ set to true in the initial state $I_{\Lambda}$.
 In this case, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, thereby making the classical planning task $P_{\Lambda}$ easier to be solved.
 
-So far we have explained the compilation for learning from a single input trace.
-However, the compilation is extensible to the more general case $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\tau_1,\ldots,\tau_k\}$ is a set of plan traces.
+So far we explained the compilation for learning from a single input trace.
+However, the compilation is extensible to the more general case $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\mathcal{O}_1,\ldots,\mathcal{O}_k\}$ is a set of $k$ observations.
 Taking this into account, a small modification is required in our compilation approach.
-In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_m^t\in \tau_t$, {\tt\small $1\leq t\leq k$} of a plan trace $\tau_t$ reset the current state and the current plan.
+In particular, the actions in $P_{\Lambda}$ for {\em validating} the last state $s_{m,t}^o\in \mathcal{O}_t$, {\tt\small $1\leq t\leq k$} of an observation $\mathcal{O}_t$ reset the current state.
 These actions are now redefined as:
@@
 	\begin{align*}
-	\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_m^t\cup\{test_{j-1}\}\cup \{\neg mode_{prog}\},\\
+	\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_{m,t}^o\cup\{test_{j-1}\}\cup \{\neg mode_{prog}\},\\
 	\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_j\} \cup \\
-	&\{\neg f\}_{\forall f\in s_m^t, f \notin s_0^{t+1}}\cup \{f\}_{\forall f\in s_0^{t+1}, f \notin s_m^t},\\
-	&\{\neg f\}_{\forall f\in F_{\pi_t}}\cup \{f\}_{\forall f\in F_{\pi_{t+1}}}.\\
+	&\{\neg f\}_{\forall f\in s_{m,t}^o, f \notin s_{0,t+1}^o}\cup \{f\}_{\forall f\in s_{0,t+1}^o, f \notin s_{m,t}^o}.
 	\end{align*}
@@
 Finally, we will detail the composition of a solution plan $\pi_\Lambda$ to a planning task $P_\Lambda$ and the mechanism to extract the action models of $\mathcal{M}'$ from $\pi_\Lambda$.
-The plan of Figure~\ref{fig:plan-lplan} shows a solution to the task $P_{\Lambda}$ that encodes a learning task $\Lambda=\tup{\mathcal{M},\tau}$ for obtaining the action models of the {\em blocksworld} domain, where the models for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already specified in $\mathcal{M}$.
-Therefore, the plan shows the insert actions and validate action for the action model {\tt\small stack} using the input plan trace of Figure~\ref{fig:example-plans}.
-Plan steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects, and steps $07-11$ form the plan postfix that applies the action models (only the {\tt\small stack} model is learned) and validates the result in the plan trace of Figure~\ref{fig:example-plans}.
+The plan of Figure~\ref{fig:plan-lplan} shows a solution to the task $P_{\Lambda}$ that encodes a learning task $\Lambda=\tup{\mathcal{M},\mathcal{O}}$ for obtaining the action models of the {\em blocksworld} domain, where the models for {\tt\small pickup}, {\tt\small putdown} and {\tt\small unstack} are already specified in $\mathcal{M}$.
+Therefore, the plan shows the insert actions and validate action for the action model {\tt\small stack}.
+Plan steps $00-01$ insert the preconditions of the {\tt\small stack} model, steps $02-06$ insert the action model effects, and steps $07-11$ form the plan postfix that applies the action models (only the {\tt\small stack} model is learned) and validates the result in the input observation.
 
@@
 	}
-	\caption{\small Plan for programming and validating the $stack$ action model (using the plan trace $\tau$ of Figure~\ref{fig:example-plans}) as well as previously specified action models for $pickup$, $putdown$ and $unstack$.}
+	\caption{\small Plan for programming the $stack$ action model and for validating the programmed $stack$ action model with previously specified action models for $pickup$, $putdown$ and $unstack$.}
 	\label{fig:plan-lplan}
@@
 
-Given a solution plan $\pi_\Lambda$ that solves $P_{\Lambda}$, the set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ are computed in linear time and space.
-In order to do so, $\pi_\Lambda$ is executed in the initial state $I_{\Lambda}$ and the action model $\mathcal{M}'$ will be given by the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that are set to true in the last state reached by $\pi_\Lambda$, $s_g=\theta(I_\Lambda,\pi_\Lambda)$.
+Given a solution plan $\pi_\Lambda$ that solves $P_{\Lambda}$, the set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\mathcal{O}}$ learning task are computed in linear time and space.
+In order to do so, $\pi_\Lambda$ is executed in the initial state $I_{\Lambda}$ and the action model $\mathcal{M}'$ will be given by the fluents $pre_{p,\xi}$, and $eff_{p,\xi}$ that are set to true in the last state reached by $\pi_\Lambda$, $s_g=\theta(I_\Lambda,\pi_\Lambda)$.
 For each $\xi \in \mathcal{M'}$, we build the sets of preconditions, positive effects and negative effects as follows:
@@
 	\begin{align*}
-	\hspace*{7pt}pre(\xi)=& \{p ~|~ pre_p(\xi) \in s_g\}_{\forall p \in \Psi_\xi},\\
-	\hspace*{7pt}add(\xi)=& \{p ~|~ add_p(\xi) \in s_g\}_{\forall p \in \Psi_\xi},\\
-	\hspace*{7pt}del(\xi)=& \{p ~|~ del_p(\xi) \in s_g\}_{\forall p \in \Psi_\xi}.
+	  \hspace*{7pt}pre(\xi)=& \{p ~|~ pre_{p,\xi} \in s_g\}_{\forall p \in \Psi_\xi},\\
+	  \hspace*{7pt}del(\xi)=& \{p ~|~ pre_{p,\xi} \wedge eff_{p,\xi} \in s_g\}_{\forall p \in \Psi_\xi},\\
+	  \hspace*{7pt}add(\xi)=& \{p ~|~ \neg pre_{p,\xi} \wedge eff_{p,\xi} \in s_g\}_{\forall p \in \Psi_\xi}.          
 	\end{align*}
@@
 
-The logical inference process our approach is based on has trouble learning preconditions that do not appear as negative effects since in this case no change is observed between the pre-state and post-state of an action.
+The plain compilation has trouble learning preconditions that do not appear as negative effects since in this case no change is observed between the pre-state and post-state of an action.
 This is specially relevant for static predicates that never change and, hence, only appear as preconditions in the actions.
-In order to address this shortcoming and complete the list of learned preconditions}, we apply a post-process based on the one proposed in~\cite{kuvcera2018louga}.
-The idea lies in going through every action and counting the number of cases where a literal is present before the action is executed and the number of cases where it is not present.
+To address this shortcoming and complete the list of learned preconditions}, we apply a post-process based on the one proposed by the {\sc LOUGA} system~\cite{kuvcera2018louga}.
+The intuition is going through every action counting the number of cases where a literal is present before the action is executed.
 If a literal is present in all the cases before the action, the literal is considered to be a precondition.
-
-In order to obtain a complete trace, the proposal in ~\cite{kuvcera2018louga} applies the sequence of actions of the input trace and infers the preconditions from this \FO action sequence}.
-In our case, since the sequence of actions of the input trace might not be fully observable, we produce the traces by applying the actions found in the validation part of the solution plan.
-For instance, in the example of the figure \ref{fig:plan-lplan}, the sequence of actions used to produce the complete trace would be {\tt{\small(unstack blockB blockA)}}, {\tt{\small(put-down blockB)}}, {\tt{\small(pick-up blockA)}}, and {\tt{\small(stack blockA blockB)}.
-
-%This post-process allows \FAMA to learn more complete lists of preconditions and deal with the always problematic static predicates.
+Since intermmediate states/actions may not be fully observed in a observation $\mathcal{O}$, we consider the actions/states found in the validation part of the solution plan $\pi_\Lambda$.
+For instance, in the example of Figure \ref{fig:plan-lplan}, the used sequence of actions is {\tt\small(unstack blockB blockA)}, {\tt\small(put-down blockB)}, {\tt\small(pick-up blockA)}, and {\tt\small(stack blockA blockB)}.
 
@@
 \begin{mylemma}
-Soundness.
-Any classical plan $\pi$ that solves $P'$ (planning task that results from the compilation) produces a model $\mathcal{M'}$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
+Soundness. Any classical plan $\pi$ that solves $P_{\Lambda}$ produces a model $\mathcal{M'}$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task.
 \end{mylemma}
@@
 \begin{small}
-According to the $P'$ compilation, once a given precondition or effect is inserted into the domain model $\mathcal{M}$ it cannot be undone.
+According to the $P_{\Lambda}$ compilation, once a given precondition or effect is inserted into the domain model $\mathcal{M}$ it cannot be undone.
 In addition, once an action model is applied it cannot be modified.
-In the compiled planning task $P'$, only ${\tt \small (apply)_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
-This means that a state consistent with an observation $s_m^o$ can only be achieved executing an applicable sequence of ${\tt \small (apply)_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_j$ $(0 < j\leq m)$, is consistent with the input state observations and {\em state-invariants}.
-This is exactly the definition of the solution condition for model $\mathcal{M}'$ to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task.
+In the compiled planning problem $P_{\Lambda}$, only ${\tt \small (apply)_{\xi,\omega}}$ actions can update the value of the state fluents $F$.
+This means that a state consistent with an observation $s_m^o$ can only be achieved executing an applicable sequence of ${\tt \small (apply)_{\xi,\omega}}$ actions that, starting in the corresponding initial state $s_0^o$, validates that every generated intermediate state $s_j$ $(0 < j\leq m)$, is consistent with the input state observations.
+This is exactly the definition of the solution condition for model $\mathcal{M}'$ to solve the $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task.
 \end{small}
@@
 \begin{mylemma}
-Completeness. Any model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning task can be computed with a classical plan $\pi$ that solves $P'$.
+Completeness. Any model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task can be computed with a classical plan $\pi$ that solves $P_{\Lambda}$.
 \end{mylemma}
@@
 By definition ${\mathcal I}_{\xi,\Psi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
-In addition the $P'$ compilation does not discard any model $\mathcal{M}'$ definable within ${\mathcal I}_{\xi,\Psi}$ that satisfies the mutexes in $\Phi$.
-This means that, for every model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$, we can build a plan $\pi$ that solves $P'$ by selecting the appropriate ${\tt \small (insert\_pre)_{\xi,e}}$ and ${\tt \small (insert\_eff)_{\xi,e}}$ actions for programming the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding ${\tt \small (apply)_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
+In addition the $P_{\Lambda}$ compilation does not discard any model $\mathcal{M}'$ definable within ${\mathcal I}_{\xi,\Psi}$.
+This means that, for every model $\mathcal{M}'$ that solves the $\Lambda=\tup{\mathcal{M},{\mathcal O}}$, we can build a plan $\pi$ that solves $P_{\Lambda}$ by selecting the appropriate ${\tt \small (insert\_pre)_{p,,\xi}}$ and ${\tt \small (insert\_eff)_{p,\xi}}$ actions for programming the precondition and effects of the corresponding action models in $\mathcal{M}'$ and then, selecting the corresponding ${\tt\small (apply)_{\xi,\omega}}$ actions that transform the initial state observation $s_0^o$ into the final state observation $s_m^o$.
 \end{small}
@@
 
-The size of $P'$ depends on the arity of the predicates in $\Psi$, that shape variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
+The size of the classical planning proble $P_{\Lambda}$ depends on the arity of the predicates in $\Psi$, that shape variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
 The larger these arities, the larger $|{\mathcal I}_{\xi,\Psi}|$.
-The size of ${\mathcal I}_{\xi,\Psi}$ is the most dominant factor of the compilation because it defines the $pre\_\xi\_e/eff\_\xi\_e$ fluents, the corresponding set of ${\tt \small insert}$ actions, and the number of conditional effects in the ${\tt \small (apply)_{\xi,\omega}}$ actions.
-Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$, which will significantly reduce $|{\mathcal I}_{\xi,\Psi}|$ and hence the size of $P'$ output by the compilation.
+The size of ${\mathcal I}_{\xi,\Psi}$ is the most dominant factor of the compilation because it defines the $pre_{p,_\xi}/eff_{p,\xi}$ fluents, the corresponding set of ${\tt\small insert}$ actions, and the number of conditional effects in the ${\tt\small (apply)_{\xi,\omega}}$ actions.
+Note that {\em typing} can be used straightforward to constrain the FOL interpretations of $\Psi$ over the parameters $pars(\xi)$, which will significantly reduce $|{\mathcal I}_{\xi,\Psi}|$ and hence the size of $P_{\Lambda}$ output by the compilation.
 
-Classical planners tend to prefer shorter solution plans, so our compilation (as well as the BLS) may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O},\Phi}$ learning tasks preferring solutions that are referred to action models with a shorter number of preconditions/effects.
-In more detail, all $\{pre\_\xi\_e, eff\_\xi\_e\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ fluents are false at the initial state of our $P'$ compilation so classical planners tend to solve $P'$ with plans that require a smaller number of ${\tt{\small insert}}$ actions.
+Classical planners tend to prefer shorter solution plans, so our compilation may introduce a bias to $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning tasks preferring solutions that are referred to action models with a shorter number of preconditions/effects.
+In more detail, all $\{pre_{p,\xi}, eff_{p,\xi}\}_{\forall e\in{\mathcal I}_{\xi,\Psi}}$ fluents are false at the initial state of our $P_{\Lambda}$ compilation so classical planners tend to solve $P_{\Lambda}$ with plans that require a smaller number of ${\tt{\small insert}}$ actions.
 
-This bias can be eliminated defining a cost function for the actions in $P'$ (e.g.
+This bias can be eliminated defining a cost function for the actions in $P_{\Lambda}$ (e.g.
 ${\tt \small insert}$ actions have {\em zero cost} while ${\tt \small (apply)_{\xi,\omega}}$ actions have a {\em positive constant cost}).
@@
 
-An interesting aspect of our approach is that when a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given in $\Lambda$, the $P_{\Lambda}$ compilation also serves to validate whether the observed $\tau$ follows the given model $\mathcal{M}$:
-
+An interesting aspect of our approach is that when a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given in $\Lambda$, the $P_{\Lambda}$ compilation also serves to validate whether the observatoin $\mathcal{O}$ follows the given model $\mathcal{M}$:
 \begin{itemize}
-	\item $\mathcal{M}$ is proved to be a {\em valid} action model for the given input data in $\tau$ iff a solution plan for $P_{\Lambda}$ can be found.
-	\item $\mathcal{M}$ is proved to be a {\em invalid} action model for the given input data $\tau$ iff $P_{\Lambda}$ is unsolvable.
+	\item $\mathcal{M}$ is proved to be a {\em valid} action model for the given input data $\mathcal{O}$ iff a solution plan for $P_{\Lambda}$ can be found.
+	\item $\mathcal{M}$ is proved to be a {\em invalid} action model for the given input data $\mathcal{O}$ iff $P_{\Lambda}$ is unsolvable.
 This means that $\mathcal{M}$ cannot be consistent with the given observation of the plan execution.
 \end{itemize}
-
-The validation capacity of our compilation is beyond the functionality of VAL (the plan validation tool~\cite{howey2004val}) because our $P_{\Lambda}$ compilation is able to address {\em model validation} of a partial (or even an empty) action model with a partially observed plan trace.
+This validation capacity of our compilation is beyond the functionality of VAL (the plan validation tool~\cite{howey2004val}) because our $P_{\Lambda}$ compilation is able to address {\em model validation} of a partial (or even an empty) action model with a partially observed plan trace.
 VAL, however, requires a full plan and a full action model for plan validation.
@@
 \section{Conclussions}
+\section{Conclusions}
+We presented a novel approach for learning \strips\ action models from examples using classical planning.
+To the best of our knowledge, this is the first approach on learning action models that is exhaustively evaluated over a wide range of domains and uses exclusively an {\em off-the-shelf} classical planner.
+The work in~\cite{SternJ17} proposes a planning compilation for learning action models from plan traces following the {\em finite domain} representation for the state variables.
+This is a theoretical study on the boundaries of the learned models and no experimental results are reported.
+
+When example plans are available, we can compute accurate action models from small sets of learning examples (five examples per domain) in little computation time (less than a second).
+When action plans are not available, our approach still produces action models that are compliant with the input information.
+In this case, since learning is not constrained by actions, operators can be reformulated changing their semantics, in which case the comparison with a reference model turns out to be tricky.
+
+An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
+
+Generating {\em informative} examples for learning planning action models is still an open issue.
+Planning actions include preconditions that are only satisfied by specific sequences of actions which have low probability of being chosen by chance~\cite{fern2004learning}.
+The success of recent algorithms for exploring planning tasks~\cite{FrancesRLG17} motivates the development of novel techniques that enable to autonomously collect informative learning examples.
+The combination of such exploration techniques with our learning approach is an appealing research direction that opens up the door to the bootstrapping of planning action models.
+
+% Commented for blind submission
+%\begin{small}
+\subsection*{Acknowledgments}
+This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.
+Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it RYC15/18009}, both programs funded by the Spanish government.
+%\end{small}
 

### 2019-03-07 doc/lecture-notes-EBL/ebl.tex
first full version lecture notes on EBL for strips action models

@@
 \usepackage[numbers]{natbib}
-\usepackage{amsthm}
+
 \usepackage{amsmath}
 \usepackage{amssymb}
-\usepackage{amsthm}
 \usepackage{multirow}
@@
 \begin{abstract}
-The paper presents the first classical planning compilation for learning \strips\ action models from partial observations of plan executions.
+The paper presents a classical planning compilation for learning \strips\ action models from partial observations of plan executions.
 The compilation is flexible to different amount and kind of available input knowledge; learning examples can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just pairs of initial and final states where no intermediate action/state is observed.
-The compilation accepts also partially specified action models and it can be used to validate whether the observation of a plan execution follows a given \strips\ action model, even if the given model or the given observation are incomplete.
+The compilation accepts also partially specified action models and it can be used to validate whether an observation of a plan execution follows a given \strips\ action model, even if the given model or the given observation are incomplete.
 \end{abstract}
@@
 Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-In both tasks, off-the-shelf automated planners are required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
+In both tasks, automated planners are required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
 Unfortunately building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
@@
 \item The {\em input} to ML algorithms (the learning/training data) is usually a finite vector that represents the value of some fixed object features.
-The input for learning planning action models is, however, observations of plan executions (where each plan possibly has a different length and plan length is not {\em a priori} bound).
+The input for learning planning action models is, however, the observation of plan executions (where each plan possibly has a different length, since plan length is not {\em a priori} bound, and may also refer to a different number of objects).
 \item The {\em output} of ML algorithms is usually a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
@@
 Learning \strips\ action models is a well-studied problem with sophisticated algorithms such as {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}, which do not require full knowledge of the intermediate states traversed by the example plans.
-Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2017generating,segovia2018computing,segovia2019computing}, this paper describes an innovative planning compilation approach for learning \strips\ action models.
+Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2017generating,segovia2018computing,segovia2019computing}, this paper describes a classical planning compilation approach for learning \strips\ action models.
 The compilation approach is appealing by itself, because it opens up the door to the bootstrapping of planning action models, but also because it is flexible to different amount and kind of available input knowledge:
@@
 \item {\em Learning examples} can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just a pair of initial and final states where no intermediate state/action is observed.
-\item {\em Partially specified action models} expressing {\em a priori} knowledge about the structure of actions can also be provided to the compilation.
-In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if the model is not fully specified or the observation is incomplete.
+\item {\em Partially specified action models}, expressing {\em a priori} knowledge about the structure of actions, can also be provided to the compilation.
+In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if the model is not fully specified or the input observation is incomplete.
 \end{enumerate}
@@
 
-Now we define actions with conditional effects because they allow us to compactly define our compilation.
+Now we define actions with {\em conditional effects} because they allow us to compactly define our compilation.
 An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
@@
 Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, $s_0^o=I$ is a sequence of possibly {\em partially observable states} (except for the initial state $s_0^o$ which is fully observable).
-A partially observable state is one in which $|s_i^o| < |F|$, {\small $1\leq i\leq n$}; i.e., a state in which at least a fluent of $F$ is not observable.
+A partially observable state is one in which $|s_i^o| < |F|$, {\small $1\leq i\leq m\leq n$}; i.e., a state in which at least a fluent of $F$ is not observable.
 
 The {\em observation model} comprises the case $|s_i^o| = 0$, when an intermediate state is fully unobservable.
-This model consider also {\em observed actions} as fluents that indicate the action applied in a given state.
+This model considers also {\em observed actions} as fluents that indicate the action applied in a given state.
 This means that a sequence of {\em observed actions} $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$ s.t.
-$a_i^o\in s_i^o$ ({\small $0\leq i <m$}).
+$a_i^o\in s_i^o$ ({\small $0\leq i\leq l <m$}).
 The number of {\em observed actions}, $l$, can then range from $0$ (in a fully unobservable action sequence) to $|\pi|$ (in a fully observed action sequence).
@@
 The sequence of observed states $\mathcal{O}(\tau)$ must be {\em consistent} with the sequence of states in $\tau(\pi,P)$.
-In practice, the number $m$ of observed states ranges from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
+In practice, the number $m$ of observed states ranges from 1 (the initial state, at least), to $|\pi|+1$, and each of the observed intermediate states comprises a number of fluents between $[1,|F|]$.
 This means that we assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
 Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of the trajectory $\tau$.
+In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of the trajectory $\tau(\pi,P)$.
 
@@
 \subsection{Explaining observations with classical planning}
-Given a {\em classical planning frame} $\Phi=\tup{F,A}$ and a observation of the execution of a plan within the given planning frame $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$, then $P_\mathcal{O}$ is a classical planning problem that is built as follows $P_\mathcal{O}=\tup{F,A,s_0^o,s_m^o}$.
+Given a {\em classical planning frame} $\Phi=\tup{F,A}$ and a observation of the execution of a plan within the given planning frame $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$, then $P_\mathcal{O}$ is the classical planning problem that is built as follows $P_\mathcal{O}=\tup{F,A,s_0^o,s_m^o}$.
 
 \begin{definition}[Explanation]
-We say that a plan $\pi$ {\em explains} $\mathcal{O}$ (denoted $\pi\mapsto\mathcal{O}$) iff $\pi$ is a solution for $P$ that is {\em consistent} with the state trajectory constraints imposed by the sequence of partial states $\mathcal{O}$.
-If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}(\tau)$.
+We say that a plan $\pi$ {\em explains} $\mathcal{O}$ iff $\pi$ is a solution for $P_\mathcal{O}$ that is {\em consistent} with the state trajectory constraints imposed by the sequence of partial states $\mathcal{O}$.
+If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}$.
 \end{definition}
 
-The {\em observation} $\mathcal{O}$ is then considered a sequence of ordered {\em landmarks} for the $P_\mathcal{O}$ classical planning problem, because all the literals in the observation must be achieved by any plan that solves $P_\mathcal{O}$ and in the same order as are defined in the observation~\cite{hoffmann2004ordered}.
+The {\em observation} $\mathcal{O}$ is then a sequence of ordered {\em landmarks} for the $P_\mathcal{O}$ classical planning problem~\cite{hoffmann2004ordered}, because all the literals in the observation must be achieved by any plan that solves $P_\mathcal{O}$ and in the same order as are defined in the observation $\mathcal{O}$.
 
@@
 \begin{itemize}
-\item $\mathcal{M}$ is the {\em initial empty model} that contains only the {\em header} (i.e., the corresponding {\em name} and {\em parameters}) of each action model to be learned.
-\item $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$ is a sequence of partially observed states.
+\item $\mathcal{M}$ is the {\em initial empty model} that contains only the {\em header} (i.e., the {\em name} and {\em parameters}) of each action model to be learned.
+\item $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$ is a sequence of partially observed states, except for the initial state $s_0^o$ which is fully observable.
 \end{itemize}
@@
 it is consistent with the headers of $\mathcal{M}$ and that it explains $\mathcal{O}$.
-We say that a model {\em explains} an observation $\mathcal{O}$ iff, when the $\tup{\rho,\theta}$ functions of the actions in $P_\mathcal{O}$ are given by that model, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}$.
+We say that a model {\em explains} an observation $\mathcal{O}$ iff, when the $\tup{\rho,\theta}$ functions of the actions in the corresponding $P_\mathcal{O}$ problem are given by that model, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}$.
+The set of fluents $F\in P_\mathcal{O}$ is induced from $s_0^o\in \mathcal{O}$ since it represents a full state.
 
@@
 \label{sec:strips-space}
-We analyze here the solution space of the addressed learning task; i.e., the space of \strips\ action models.
+We analyze here the solution space of the addressed learning task; in this case the space of \strips\ action models.
 
@@
 For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-Hence, solving a $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task is determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi\in \mathcal{M}$.
+Hence, solving a $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task is determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of the corresponding action model.
 
@@
 The solution $\mathcal{M}'$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-{\em Typing constraints} would also be a type of syntactic constraint~\cite{mcdermott1998pddl}.
+{\em Typing constraints} are also a type of syntactic constraint that reduce the size of ${\mathcal I}_{\xi,\Psi}$~\cite{mcdermott1998pddl}.
 \item {\bf Observation constraints}.
 The solution $\mathcal{M}'$ must be consistent with these \emph{semantic constraints} derived from the input observation $\mathcal{O}$.
-Specifically, the states induced by the plan computable with $\mathcal{M}'$ must comprise the observed states of the sample, which further constrains the space of possible action models.
+Specifically, the states induced by plans computable with $\mathcal{M}'$ must comprise the observed states of the sample, which further constrains the space of possible action models.
 \end{enumerate}
@@
 The belonging of a $p \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre_{p,\xi}$ and $eff_{p,\xi}$, instead of the three fluents used in the original compilation~\cite{aineto2018learning}.
-The four possible combinations of these two fluents are sumarized in Figure \ref{fig:combinations}.
-This compact encoding allows for a more effective exploitation of the syntactic constraints, and also yields the solution space of $\Lambda$ to be the same as its search space.
+The four possible combinations of these two fluents are summarized in Figure \ref{fig:combinations}.
+This compact encoding allows for a more effective exploitation of the syntactic constraints, and also yields the solution space of $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ to be the same as its search space.
 
@@
 $pre_{p,\xi} \wedge eff_{p,\xi} $& $p$ is a negative effect of $\xi$ \\
-               &  ($p \in pre(\xi) \wedge p \notin add(\xi) \wedge p \in add(\xi)$)
+               &  ($p \in pre(\xi) \wedge p \notin add(\xi) \wedge p \in del(\xi)$)
 		\end{tabular}
 	\end{footnotesize}
-	\caption{\small Combinations of the fluent propositional encoding and their meaning}
+	\caption{\small Combinations of the propositional encoding and their meaning}
 	\label{fig:combinations}
@@
 
-To illustrate better this encoding, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
+To illustrate better this encoding, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with $pre_{p,stack}$ and $eff_{p,stack}$ fluents ($p\in{\mathcal I}_{\Psi,stack}$).
 
@@
    :precondition (and (holding ?v1) (clear ?v2))
-   :effect (and (not (holding ?v1)) (not (clear ?v2)) (clear ?v1) (handempty) (on ?v1 ?v2)))
+    :effect (and (not (holding ?v1)) (not (clear ?v2))
+                 (clear ?v1) (handempty) (on ?v1 ?v2)))
 
@@
 \subsection{The sampling space}
-According to our {\em observation model} the minimal expression of an observation must comprise at least two state observations, a full initial state $s_0^o$ and a partially observed final state $s_m^o$ so $m \geq 1$.
-On the other hand, the maximal expression of an observation corresponds to a fully observed trajectory $\mathcal{O}(\tau)=\tau$ (meaning that all traversed states and applied actions are fully observed).
-In between there is a grey scale of different kinds of possible observations, including the observation of the initial state and the executed plan, that is frequently used for previous system that learn planning action models such as {\sc ARMS}~\cite{yang2007learning} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
+According to our {\em observation model} the minimal expression of an observation must comprise at least two state observations $\mathcal{O}(\tau)=\tup{s_0^o,s_m^o}$, a full initial state $s_0^o$ and a partially observed final state $s_m^o$.
+Figure~\ref{fig:observation} shows an example of a $\mathcal{O}(\tau)=\tup{s_0^o,s_m^o}$ observation that contains only two states.
+An initial state of the blocksworld where the robot hand is empty and there are two blocks ({\tt\small{blockB}} on top of  {\tt\small{blockA}}).
+The observation represents also a partially observable final state in which {\tt\small{blockA}} is on top of {\tt\small{blockB}}.
 
-Figure~\ref{fig:observation} shows an example of an observation that contains only two states.
-An initial state of the blocksworld where the robot hand is empty and three blocks (namely {\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are on top of the table and clear.
-The observation represents also a partially observable final state in which {\tt\small{blockA}} is on top of {\tt\small{blockB}} and {\tt\small{blockB}} on top of {\tt\small{blockC}}.
+On the other hand, the maximal expression of an observation corresponds to a fully observed trajectory $\mathcal{O}(\tau)=\tau$ (meaning that all traversed states, and applied actions, are fully observed).
+In between there is a grey scale of different kinds of possible observations, including the observation of the initial state and the executed plan, that is frequently used for previous systems that learn planning action models such as {\sc ARMS}~\cite{yang2007learning} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
 
@@
 
-(:init (ontable blockA) (clear blockA) (ontable blockB) (clear blockB)
-       (ontable blockC) (clear blockC) (handempty))
+(:init (ontable blockA) (on blockB blockA) (clear blockB) (handempty))
 
-(:observation (on blockA blockB) (on blockB blockC))
+(:observation (on blockA blockB))
   \end{verbatim}
@@
 \section{Learning \strips\ action models with classical planning}
-Our approach to address a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is to compile it into a planning task $P_{\Lambda}$.
+Our approach to address a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is to compile it into a classical planning problem $P_{\Lambda}$.
 The intuition behind the compilation is that when $P_{\Lambda}$ is solved, the solution plan $\pi_\Lambda$ is a sequence of actions that build the output model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the observation ${\mathcal O}$.
 
-A solution plan $\pi_\Lambda$ comprise then two differentiated blocks of actions: a plan prefix with a set of actions each defining the \textbf{insertion} of a fluent as a precondition or a effect of an action model $\xi \in \mathcal{M'}$; and a plan postfix with a set of actions that determine the \textbf{application} of the learned modes while successively \textbf{validating} the effects of the action application in every partial state in ${\mathcal O}$.
-Roughly speaking, in the \emph{blocksworld}, the format of the first set of actions of $\pi_\Lambda$ looks like {\tt{\small (insert\_pre\_stack\_holding\_v1)}, {\tt\small (insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_holding\_v1)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ is like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1)}, {\tt\small (validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states in ${\mathcal O}$.
+A solution plan $\pi_\Lambda$ comprise then two differentiated blocks of actions: a plan prefix with a set of actions each defining the \textbf{insertion} of a fluent as a {\em precondition} or a {\em effect} of an action model; and a plan postfix with a set of actions that determine the \textbf{application} of the learned modes while successively \textbf{validating} the effects of the action application in every partial state in ${\mathcal O}$.
+Roughly speaking, in the \emph{blocksworld}, the format of the first set of actions of $\pi_\Lambda$ looks like {\tt{\small (insert\_pre\_stack\_holding\_v1)}, {\tt\small (insert\_eff\_stack\_clear\_v1)}, {\tt\small (insert\_eff\_stack\_holding\_v1)}}\ldots where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ is like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1)}, {\tt\small (validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states in ${\mathcal O}$.
 
 \subsection{Compilation}
-A learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is compiled into a classical planning problem with conditional effects $P_{\Lambda}$.
-A solution plan $\pi_\Lambda$ to $P_{\Lambda}$ induces the output domain model $\mathcal{M}'$ that solves the learning task $\Lambda$.
-Specifically, a solution plan $\pi_\Lambda$ serves two purposes:
-
-\begin{enumerate}
-\item {\bf To build the action models of $\mathcal{M}'$}.
-$\pi_\Lambda$ comprises a first block of actions (plan {\em prefix}) that set the predicates $p\in \Psi_{\xi}$ of $pre(\xi)$, $del(\xi)$ and $add(\xi)$ for each $\xi\in\mathcal{M}$.
-\item {\bf To validate the action models of $\mathcal{M}'$}.
-$\pi_\Lambda$ also comprises a second block of actions (plan {\em postfix}) which is aimed at {\em explaining} the observation ${\mathcal O}$ with the built action models $\mathcal{M}'$.
-\end{enumerate}
-
 Given a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
 \begin{itemize}
-\item $F_{\Lambda}$ extends $F$ with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of ${\mathcal O}$.
-Specifically, $F_{\Lambda}$ contains:
+\item $F_{\Lambda}$ extends the set of fluents $F$ (obtained from $s_0^o$) with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of ${\mathcal O}$.
+Specifically, $F_{\Lambda}$ contains also:
 \begin{itemize}
-\item The set of fluents obtained from $s_0^o$; i.e., $F$.
-\item The model fluents $pre_{p,\xi}$ and $eff_{p,\xi}$, for every $p\in \Psi_{\xi}$ and $\xi \in \mathcal{M}$, defined as explained in section~\ref{sec:strips-space}.
+\item Fluents $pre_{p,\xi}$ and $eff_{p,\xi}$, defined as explained in section~\ref{sec:strips-space}.
 \item A set of fluents $\{test_j\}_{0\leq j\leq m}$, to point at the state observation $s_j^o\in {\mathcal O}$ where the action model is validated.
 In the example of Figure~\ref{fig:observation} two tests are required to validate the programmed action model, one corresponding to the initial state and the second one corresponding to the final state.
-\item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated and a fluent {\small\tt invalid} to indicate that the programmed action model is inconsistent with the input observation.
+\item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated and a fluent {\em invalid} to indicate that the programmed action model is inconsistent with the input observation.
 \end{itemize}
@@
 
-\item $G_{\Lambda}$ includes the positive literal $test_m$ and the negative literal $\neg${\small\tt invalid}.
+\item $G_{\Lambda}$ includes the positive literal $test_m$ and the negative literal $\neg${\em invalid}.
 When these goals are achieved by the solution plan $\pi_\Lambda$, we will be certain that the action models of $\mathcal{M'}$ are validated in the input observation.
@@
 \begin{enumerate}
-\item Actions for {\em inserting} a component (precondition or effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models.
+\item Actions for {\em inserting} a precondition or effect into $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models.
 These actions will form the prefix of the solution plan $\pi_\Lambda$.
@@
 \begin{itemize}
-\item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
-A precondition $p$ is inserted in $\xi$ when neither $pre_p$ nor $eff_p$ exist in $\xi$.
+\item Actions for inserting a {\em precondition} $p\in{\mathcal I}_{\xi,\Psi}$ into $\xi$.
 \begin{small}
@@
 
-\item Actions which support the addition of a effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$. 
+\item Actions for inserting a {\em effect} $p\in{\mathcal I}_{\xi,\Psi}$ into $\xi$. 
 \begin{small}
@@
 
-For instance, given $name(\xi)=${\tt{\small stack}} and $F_{pre,stack}=\{${\tt{\small (pre\_stack\_holding\_v1),} {\tt\small (pre\_stack\_holding\_v2), (pre\_stack\_on\_v1\_v2),(pre\_stack\_clear\_v1)}, {\tt\small(pre\_stack\_clear\_v1)}}, $\ldots \}$, the insertion of each item $p \in F_{pre,stack}\subseteq F_{\Lambda}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$.
-The same applies to effects $F_{eff,stack}=\{${\tt{\small (eff\_stack\_holding\_v1),(eff\_stack\_holding\_v2), (eff\_stack\_on\_v1\_v2),(eff\_stack\_clear\_v1),(eff\_stack\_clear\_v1),}}$\ldots \}$.
+For instance, given $name(\xi)=${\tt{\small stack}} and $\{${\tt{\small (pre\_stack\_holding\_v1),} {\tt\small (pre\_stack\_holding\_v2)}, {\tt\small(pre\_stack\_on\_v1\_v2)}, {\tt\small(pre\_stack\_clear\_v1)}, {\tt\small(pre\_stack\_clear\_v1)}}, $\ldots \}$, the insertion of each item $p \in {\mathcal I}_{\xi,\Psi}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$.
+The same applies to effects $\{${\tt{\small (eff\_stack\_holding\_v1)}, {\tt\small(eff\_stack\_holding\_v2)}, {\tt\small (eff\_stack\_on\_v1\_v2)}, {\tt\small (eff\_stack\_clear\_v1)}, {\tt\small (eff\_stack\_clear\_v1),}}$\ldots \}$.
 
@@
 \end{scriptsize}
- \caption{\small PDDL action for applying an already programmed model for $stack$ (implications are coded as disjunctions).}
+ \caption{\small PDDL action for applying an already programmed model for $stack$.}
 \label{fig:compilation}
@@
 
-When the input plan trace contains observed actions, the extra conditional effects
+When the input plan trace contains observed actions extra preconditions have to be added to ensure that actions are applied in the same order as they appear in $\mathcal{O}$~\cite{aineto2018learning}.\\
 
-$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.\\
-
-\item Actions for {\em validating} the partially observed state $s_j\in\mathcal{O}$, {\tt\small $1\leq j< m$}.
+\item Actions for {\em validating} partially observed states $s_j^o\in\mathcal{O}$.
 These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observation $\mathcal{O}$ follows after the execution of the apply actions.
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j\cup\{test_{j-1}\},\\
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{test_{j-1}\},\\
 \cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j\}.
@@
 There will be a validate action in $\pi_\Lambda$ for every observed state in $\mathcal{O}$.
-The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\mathcal{O}$.
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j^o\in\mathcal{O}$.
 
@@
 In our compilation approach, when an action model $\xi$ is partially specified, the known preconditions and effects are encoded as fluents $pre_{p,\xi}$ and $eff_{p,\xi}$ set to true in the initial state $I_{\Lambda}$.
-In this case, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, thereby making the classical planning task $P_{\Lambda}$ easier to be solved.
+In this case, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary making the classical planning task $P_{\Lambda}$ easier to be solved.
 
 So far we explained the compilation for learning from a single input trace.
-However, the compilation is extensible to the more general case $\Lambda=\tup{\mathcal{M},\mathcal{T}}$, where $\mathcal{T}=\{\mathcal{O}_1,\ldots,\mathcal{O}_k\}$ is a set of $k$ observations.
+However, the compilation is extensible to the more general case $\Lambda=\tup{\mathcal{M},\mathcal{O}_1,\ldots,\mathcal{O}_k}$ where there is an input set of $k$ observations.
 Taking this into account, a small modification is required in our compilation approach.
@@
 
-The plain compilation has trouble learning preconditions that do not appear as negative effects since in this case no change is observed between the pre-state and post-state of an action.
+The plain compilation has trouble learning preconditions that do not appear as negative effects since in this case no change can be observed between the {\em pre-state} and {\em post-state} of an action application.
 This is specially relevant for static predicates that never change and, hence, only appear as preconditions in the actions.
-To address this shortcoming and complete the list of learned preconditions}, we apply a post-process based on the one proposed by the {\sc LOUGA} system~\cite{kuvcera2018louga}.
+To address this shortcoming and complete the list of learned preconditions, we apply a post-process based on the one proposed by the {\sc LOUGA} system~\cite{kuvcera2018louga}.
 The intuition is going through every action counting the number of cases where a literal is present before the action is executed.
 If a literal is present in all the cases before the action, the literal is considered to be a precondition.
-Since intermmediate states/actions may not be fully observed in a observation $\mathcal{O}$, we consider the actions/states found in the validation part of the solution plan $\pi_\Lambda$.
+Since intermediate states/actions may not be fully observed in a observation $\mathcal{O}$, we consider the actions/states found in the validation part of the solution plan $\pi_\Lambda$.
 For instance, in the example of Figure \ref{fig:plan-lplan}, the used sequence of actions is {\tt\small(unstack blockB blockA)}, {\tt\small(put-down blockB)}, {\tt\small(pick-up blockA)}, and {\tt\small(stack blockA blockB)}.
@@
 
-\begin{proof}[Proof]
+\begin{proof}[]
 \begin{small}
@@
 
-The size of the classical planning proble $P_{\Lambda}$ depends on the arity of the predicates in $\Psi$, that shape variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
+The size of the classical planning problem $P_{\Lambda}$ depends on the arity of the predicates in $\Psi$, that shape variables $F$, and the number of parameters of the action models, $|pars(\xi)|$.
 The larger these arities, the larger $|{\mathcal I}_{\xi,\Psi}|$.
@@
 
-An interesting aspect of our approach is that when a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given in $\Lambda$, the $P_{\Lambda}$ compilation also serves to validate whether the observatoin $\mathcal{O}$ follows the given model $\mathcal{M}$:
+An interesting aspect of our approach is that when a {\em fully} or {\em partially specified} \strips\ action model $\mathcal{M}$ is given in $\Lambda$, the $P_{\Lambda}$ compilation also serves to validate whether the observation $\mathcal{O}$ follows the given model $\mathcal{M}$:
 \begin{itemize}
@@
 
-\section{Conclussions}
 \section{Conclusions}
-We presented a novel approach for learning \strips\ action models from examples using classical planning.
+We presented a classical planning compilation for learning \strips\ action models from partial observations of plan executions.
 To the best of our knowledge, this is the first approach on learning action models that is exhaustively evaluated over a wide range of domains and uses exclusively an {\em off-the-shelf} classical planner.

### 2019-03-08 doc/icaps-19/camera-ready/model-recognition.tex
ICAPS19: assumptions revisited

@@
 The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+The {\em trajectory length} of $\tau(\pi,P)$ is given by the plan length of $\pi$.
 A trajectory $\tau(\pi,P)$ that solves $P$ is one in which $G \subseteq s_n$.
@@
 \begin{enumerate}
-\item The {\em principle of rationality}: The expectation that intentional agents will tend to choose actions that achieve their goals most efficiently, given their knowledge of the actual world state~\cite{Dennett83}.
-\item {\em The best plans for $P$ are unique or have different cost}.
+\item Similarly as the rationality principle \cite{Dennett83} applied to agents acting rationally, we assume that agents will perform the least needed amount of adjustments to its model needed to explain the observation.
+\item A model can only produce one trajectory consistent with the given observation $\mathcal{O}$ and any such trajectory is equally likely.
+This is a strong assumption that can be relaxed in future work by weighting the edit distance against the trajectory length.
 \end{enumerate}
 
-Similar assumptions were taken in previous work for {\em goal recognition}~\cite{ramirez2012plan}.
-They allow us to assume that the sum in equation (2) is dominated by its largest term, so other terms in the sum are not added up.
-The largest term corresponds here to the shortest trajectory $\tau^*$ consistent with the observation (which now is assumed to be unique).
-Note that the more complete the observation of the plan execution is the more accurate our estimate is because it becomes more likely to assume that there is only one trajectory consistent with the input observation.
+These assumptions mean that the sum in equation (2) is dominated by its largest term, so other terms in the sum are not added up.
+The largest term corresponds here to the closest consistent model $\mathcal{M^*}$.
+Note that the more complete the observation of the plan execution
+is the more accurate our estimate becomes, because the space of possible trajectories becomes more constrained by the observation.
 
-Again, we have that there is a single model $\mathcal{M}^*$ that can produce a given trajectory $\tau^*$.
+Under assumption (2), both $P(O|\tau)$ and $P(\tau|\mathcal{M}')$ are constant and can be taken out of the equation so we have that equation (1) simplifies once again to equation (3) with $\mathcal{M}^*$ being any model in $M^*$.
+
+%Again, we have that there is a single model $\mathcal{M}^*$ that can produce a given trajectory $\tau^*$.
 With this said, both $P(O|\tau^*)$ and $P(\tau^*|\mathcal{M}^*)$ are independent of $\mathcal{M} \in M$ so equation (1) simplifies once again to equation (3) under the previous two assumptions.
@@
 \subsection{Recognizing failures in a non-deterministic {\em blocksworld}}
-With this second experiment we aim to validate the two assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood: (1) that agents act rationally and so $P(\mathcal{O}|\mathcal{M})$ is dominated by the shortest trajectory that explains the observation $\mathcal{O}(\tau)$ and (2) that there exists a single such trajectory.
+With this second experiment we aim to validate the two assumptions underlying our approximation of the $P(\mathcal{O}|\mathcal{M})$ likelihood: (1) that agents adjust their model rationally so $P(\mathcal{O}|\mathcal{M})$ is dominated by the closest consistent model that explains the observation $\mathcal{O}(\tau)$ and (2) that this model can only produce a single trajectory consistent with $\mathcal{O}(\tau)$.
 Our assumptions are more correct with higher observability and become tautological when the trajectory is fully observed ($\mathcal{O(\tau)} = \tau$).
@@
 Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it RYC15/18009}, both programs funded by the Spanish government.
+M.
+Ramirez research is partially funded by DST Group Joint \& Operations Analysis Division.
 

### 2019-03-10 doc/icaps-19/camera-ready/model-recognition.tex
last updates

@@
 
-A growing interest in addressing the task of recognizing the mission (plans and goals) of a planning agent from a set of given observations has recently emerged.
-Approaches based on rule-based systems, parsing, graph-covering or Bayesian nets have been proposed for the plan/goal recognition tasks~\cite{sukthankar14}.
-{\em Plan recognition as planning} is the model-based approach for this task that leverages the action model of the observed agent to compute its most likely goal and predict its future actions~\cite{ramirez2009plan,ramirez2012plan}.
-While action models are learnable from observations of the agent~\cite{yang2007learning,cresswell2013acquiring,aineto2018learning}, in real-world applications developing in a particular context, approximate behavioral models (manually-introduced or
-automatically-learned) of the involved agents are usually already available, so there is rarely a need to learn a model from scratch.
-A more relevant and applicable task in these scenarios is thereby to predict the behavior of an agent from a set of given models.
+Addressing the task of recognizing the mission (plans and goals) of a planning agent from a set of given observations has recently captured much attention.
+Rule-based systems, parsing, graph-covering or Bayesian nets have been proposed for plan/goal recognition tasks~\cite{sukthankar14}.
+{\em Plan recognition as planning} is the model-based approach that leverages the action model of the observed agent to compute its most likely goal and predict its future actions~\cite{ramirez2009plan,RamirezG10}.
+This paper presents the novel task of \emph{model recognition}, where the target of the recognition is not a goal or a plan but the model that shapes the behaviour of the observed agent.
+Given a \emph{partially observed plan execution}, and a \emph{set of possible planning models}, model recognition is the task of identifying the model in the set with the highest probability of producing the given observation.
+Model recognition enables the identification of the agent'e behaviour without need of being aware of the agent's intention.
 
-This paper formalizes the novel task of {\em model recognition}, where the object to recognize is not a goal or a plan but the model that shapes the behavior of the observed agent.
-Given a partially observed plan execution, and a set of possible planning models (models that share the same state variables but different action schemata), {\em model recognition} is the task of identifying the model in the set with the highest probability of producing the given observation.
-We posit this problem as a \emph{classification task} where classes are given as generative planning models and samples are sequences of observations.
-We accomplish this task with off-the-shelf classical planning algorithms that compute for each model (class) $\mathcal{M}$ in the set a model $\mathcal{M'}$ that explains the observation, and subsequently distances between the two models are measured.
-Two key contributions of our {\em model recognition as planning} approach are: (1) the ability of handling partial observations of plan executions that contain no actions and incomplete -- even empty- intermediate states; and (2) a thorough formalization that induces a probability distribution of a model $\mathcal{M}$ to explain an observed plan execution.
+As an example of model recognition, imagine a robot in a $N\times N$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
+According to this model, when {\tt\small $q0$} ({\tt\small $q1$}) holds, the robot is at an even (odd) row and can move right (left) by incrementing (decrementing) its x-coordinate.
+Other models that shape different strategies of robot navigation can be defined within the same state variables altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated.
+Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, model recognition aims to identify the robot action schemata that explains this observation.
 
-\begin{figure}
-  \begin{scriptsize}
-  \begin{center}
-  \begin{tikzpicture}[scale=.4]
-          \begin{scope}
-            \draw (0, 0) grid (5, 5);
-            \node[anchor=center] at (2.5, 1.5) {R};
-            \node[anchor=center] at (0.5, 0.5) {};
-            \draw[thick,style=dotted] (0.5,0.5) -- (3.5,0.5);
-            \draw[thick,style=dotted] (3.5,0.5) -- (3.5,1.5);
-            \draw[thick,style=dotted,->] (3.5,1.5) -- (2.7,1.5);
-          \end{scope}
-        \end{tikzpicture}
-\hspace*{1cm}
-	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-	  \node[state] (A)              {$q_0$};
-	  \node[state] (B) [right of=A] {$q_1$};
-	  \path
-(A) edge [bend left, align=center] node {inc-y,\\dec-y} (B)
-(B) edge [bend left, align=center] node {inc-y,\\dec-y} (A);
-	\end{tikzpicture}
-  \end{center}
-  \end{scriptsize}
- \caption{\small ({\em Left)} Robot navigating a $5\times 5$ grid.
-{\em (Right)} Automata to control that the robot only increments its x-coordinate when $q_0$ holds (actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and switch the automata state).}
-\label{fig:grid-example}
-\end{figure}
-
-To better illustrate {\em model recognition}, imagine a robot in a $N\times N$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model, the robot can increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
+%the robot can increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
 at even rows if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at odd rows if {\tt\small $q0$} holds initially).
@@
 Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where state variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, {\em model recognition} aims to identify the robot action schemata that explains this observation.
-
-We provide a formalization of {\em model recognition} that is applicable to any planning model but we restrict our attention to \strips\texttt{-}compilable models.
-It is well-known that diverse automata representations like finite state controllers, {\sc GOLOG} programs or reactive policies can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,segovia2017generating}.
-As for the evaluation, the effectiveness of our approach is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a proof-of-concept classification task; (2) a simulation of a non-deterministic \emph{blocksworld} under different degrees of observability; and (3) a navigation task where observations contain an unbounded number of unobservable states.
-
-%: a classical string classification task; identification of the model that encodes a failure present in an observation; and recognition of a robot navigation policy.
-
-%The paper also introduces {\em model recognition as planning}; a novel method to estimate the probability of a given \strips\ model to produce an observed plan execution.
-Our method is built on top of off-the-shelf classical planning algorithms and is robust to missing intermediate states and actions in the observed plan execution.
 
@@
 
+Our proposal to tackle the task of model recognition is to compute, for each model $\mathcal{M}$ in the set, a model $\mathcal{M'}$ by editing $\mathcal{M}$ with the least needed amount of adjustments that are required for $\mathcal{M'}$ to explain the observation.
+We apply an \emph{edit distance} that determines the number of insertions and deletions of preconditions and effects that are required to transform $\mathcal{M}$ into $\mathcal{M'}$.
+With this in mind, model recognition resembles other tasks like \emph{model reconciliation}, which uses model editing to conform two PDDL models, where one of them can be an annotated model or an empty model, with respect to a fully observed optimal plan computed with one of the two models~\cite{ChakrabortiSZK17,sreedharan2018handling,Chakraborti18}.
+Also, the \emph{model drift} approach estimates the error between a computational model, and a ground-truth model which changes as the domain evolves, as the normalized symmetric difference of the propositions that appear in the preconditions and effects of both models.
+In this case, plan observations are used to hypothesize variations of the model.
+
+%, which are then checked via observing query answers, validity of plans or allowing manual edits to the planning model.
+This mixed-initiative approach estimates the error between the two models as the normalized symmetric difference of the propositions that appear in the preconditions and effects of both models.
+
+Unlike model reconciliation and model drift, our proposal computes the model $\mathcal{M'}$ from \emph{partial observations} of plan executions that contain no actions and incomplete -- even empty- intermediate states.
+Moreover, we leverage the two-model edit distance to elaborate a probabilistic distribution of a model $\mathcal{M}$, which is used to posit the model recognition task as a \emph{classification task} with as many classes as models in the set.
+This way, we account for the possibly different prior probability of each model, a feature that cannot be considered with purely distance-based metrics.
+
+One last consideration is related to the use of an action model learning approach with partial observability to induce the model $\mathcal{M'}$.
+While most approaches can learn from empty or partially specified models ~\cite{yang2007learning,AmirC08,cresswell2013acquiring}, model recognition requires the ability of editing a possibly inconsistent model $\mathcal{M}$ so that the resulting model $\mathcal{M'}$ is consistent with a partial observation in which all or many of the actions are probably missing.
+This poses the problem of \emph{filling the gaps} of the incomplete observation and so it justifies using an approach to \emph{learning action models as planning}~\cite{aineto2018learning}.
+Otherwise, in the case of full observability, our formalization might as well use any model editing approach such as model reconciliation or model drift.
+
+We provide a formalization of model recognition that is applicable to any planning model but we restrict our attention to \strips\texttt{-}compilable models.
+It is well-known that diverse automata representations like finite state controllers, {\sc GOLOG} programs or reactive policies can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,segovia2017generating}.
+As for the evaluation, the effectiveness of our approach is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a proof-of-concept classification task; (2) a simulation of a non-deterministic \emph{blocksworld} under different degrees of observability; and (3) a navigation task where observations contain an unbounded number of unobservable states.
+
+\begin{figure}
+  \begin{scriptsize}
+  \begin{center}
+  \begin{tikzpicture}[scale=.4]
+          \begin{scope}
+            \draw (0, 0) grid (5, 5);
+            \node[anchor=center] at (2.5, 1.5) {R};
+            \node[anchor=center] at (0.5, 0.5) {};
+            \draw[thick,style=dotted] (0.5,0.5) -- (3.5,0.5);
+            \draw[thick,style=dotted] (3.5,0.5) -- (3.5,1.5);
+            \draw[thick,style=dotted,->] (3.5,1.5) -- (2.7,1.5);
+          \end{scope}
+        \end{tikzpicture}
+\hspace*{1cm}
+	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
+	  \node[state] (A)              {$q_0$};
+	  \node[state] (B) [right of=A] {$q_1$};
+	  \path
+(A) edge [bend left, align=center] node {inc-y,\\dec-y} (B)
+(B) edge [bend left, align=center] node {inc-y,\\dec-y} (A);
+	\end{tikzpicture}
+  \end{center}
+  \end{scriptsize}
+ \caption{\small ({\em Left)} Robot navigating a $5\times 5$ grid.
+{\em (Right)} Automata to control that the robot only increments its x-coordinate when $q_0$ holds (actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and switch the automata state).}
+\label{fig:grid-example}
+\end{figure}
+
 \section{Background}
 \label{sec:background}
-This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
+%This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
 
@@
 Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a_c\in A$ is applicable in a state $s$ if and only if $\pre(a_c)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
-\[
-triggered(s,a_c)=\bigcup\limits_{C\rhd E\in\cond(a_c),C\subseteq s} E. 
-\]
-The result of applying $a_c$ in state $s$ follows the same definition of successor state, $\theta(s,a)$, but applied to the conditional effects in $triggered(s,a_c)$.
+An action $a_c\in A$ is applicable in a state $s$ if and only if $\pre(a_c)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$: $triggered(s,a_c)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E$.
 
@@
 In this experiment we included the 0\% case which corresponds to observations where only the initial and final states are observed.
-The figure shows that for 0\% observability we were unable to unmistakably identify the generating navigation policies, but from 10\% of observability onwards, we start to correctly classify half of the observations.
+The figure shows that for 0\% observability we were unable to unmistakably identify the navigation policies, but from 10\% of observability onwards, we start to correctly classify half of the observations.
 Accuracy stabilizes after 40\% observability in the range 0.875 to 1 which means that at most only one out of the 8 observations was not correctly classified.
@@
 Finally, the robustness of the approach is shown in the third experiment where actions and intermediates states are unobservable.
-A meaningful lesson from the experiments is that our approach allows us to predict the future actions of agents whose behavior is ruled by a model other than a goal-driven deliberative planning model; this includes, for instance, models of fixed programs or finite state machines (regular automata).
-
-An interesting research line related to {\em model recognition} is {\em model reconciliation}~\cite{ChakrabortiSZK17}, in which model editing is used to conform the PDDL models of two agents, or the model of an agent with a given {\em annotated model}~\cite{sreedharan2018handling}, with respect to a fully observed optimal plan computed with one of the two models.
-{\em Model recognition}, however, conforms every input model $\mathcal{M}$ with another model $\mathcal{M}'$ that is not given as input but instead computed for every $\mathcal{M}$, and which is consistent with a partial observation of a plan execution.
+A significant lesson from the experiments is that our approach allows us to predict behaviours that are not ruled by a goal-driven deliberative planning model (e.g., models of fixed programs or finite state machines --regular automata-).
 

### 2019-03-10 doc/icaps-19/camera-ready/model-recognition.tex
ICAPS19: intro example changed

@@
 
-As an example of model recognition, imagine a robot in a $N\times N$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
+%As an example, imagine you are given a set of navigation policies for a $N\times N$ grid.
+Figure~\ref{fig:model-example} shows an example of such policy, according to which, the agent can only move right when he is at an even row ({\tt\small $q0$} holds), and left when he is at an odd row ({\tt\small $q1$} holds).
+Other models that shape different strategies of robot navigation can be defined within the same state variables altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated.
+Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, model recognition aims to identify the navigation policy that better explains this observation.
+
+As an example, imagine you are given a set of models, each one encoding a different navigation policy for a $N\times N$ grid.
+Figure~\ref{fig:model-example} shows an example of one such model, according to which, the agent can only move right when he is at an even row ({\tt\small $q0$} holds), and left when he is at an odd row ({\tt\small $q1$} holds).
+Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, model recognition aims to identify the navigation policy (model) that better explains this observation.
+
+%As an example of model recognition, imagine a robot in a $N\times N$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
 According to this model, when {\tt\small $q0$} ({\tt\small $q1$}) holds, the robot is at an even (odd) row and can move right (left) by incrementing (decrementing) its x-coordinate.
@@
 The first experiment is a proof-of-concept \emph{classification task} that perfectly recognizes the regular automata of a given string.
-Additionally, the two assumptions underpinning our formal theory are experimentally proved in the second experiment by varying the degree of observability.
+The second experiment, where the number of trajectories explaining the observation is bounded, shows that a more accurate model prediction is achieved with more complete observations.
 Finally, the robustness of the approach is shown in the third experiment where actions and intermediates states are unobservable.

### 2019-03-10 doc/icaps-19/camera-ready/model-recognition.tex
last update in conclusions

@@
 Our {\em model recognition as planning} approach builds on top of off-the-shelf classical planning algorithms and it is robust to missing actions and incomplete or missing intermediate states in the observation.
-Once the planning model of the observed agent is recognized, the model-based machinery for automated planning becomes fully applicable for other recognition tasks like {\em goal recognition}~\cite{ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} or {\em counter-planning}~\cite{PozancoEFB18}, which require beforehand a model of the observed agent.
+Once the planning model of the observed agent is recognized, the model-based machinery for automated planning becomes fully applicable for other recognition tasks like {\em goal recognition}~\cite{RamirezG10,ramirez2012plan}, {\em goal recognition design}~\cite{KerenGK14} or {\em counter-planning}~\cite{PozancoEFB18}, which require beforehand a model of the observed agent.
 
@@
 Finally, the robustness of the approach is shown in the third experiment where actions and intermediates states are unobservable.
-A significant lesson from the experiments is that our approach allows us to predict behaviours that are not ruled by a goal-driven deliberative planning model (e.g., models of fixed programs or finite state machines --regular automata-).
+
+A significant lesson from the experiments is that our approach allows us to predict behaviours that are not ruled by a goal-driven deliberative planning model such as, for instance, models of fixed programs or finite state machines (regular automata).
 

### 2019-03-10 doc/icaps-19/camera-ready/model-recognition.tex
ICAPS19:minor changes

@@
 As an example, imagine you are given a set of models, each one encoding a different navigation policy for a $N\times N$ grid.
-Figure~\ref{fig:model-example} shows an example of one such model, according to which, the agent can only move right when he is at an even row ({\tt\small $q0$} holds), and left when he is at an odd row ({\tt\small $q1$} holds).
+Figure~\ref{fig:model-example} shows an example of one such model, according to which, the agent can only move right when it is at an even row ({\tt\small $q0$} holds), and left when it is at an odd row ({\tt\small $q1$} holds).
 Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, model recognition aims to identify the navigation policy (model) that better explains this observation.
@@
 
-\subsection*{Acknowledgement}
+\subsection*{Acknowledgements}
 This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.

### 2019-03-25 doc/icaps-19/camera-ready-aaai/MAIN-AinetoD.92.tex
Added icaps19 camera-ready aaai instructions (submitted version)

--- 2019-03-10 doc/icaps-19/camera-ready/model-recognition.tex
@@
-\def\year{2017}\relax
+\def\year{2019}\relax
 %File: formatting-instruction.tex
@@
 \usepackage{courier}  %Required
-\usepackage{url}  %Required
+\usepackage{url}\urlstyle{rm}  %Required
 \usepackage{graphicx}  %Required
@@
 \setlength{\pdfpageheight}{11in}  %Required
-%PDF Info Is Required:
-  \pdfinfo{}
 \setcounter{secnumdepth}{0}
 
+% Addiotional packages
 \usepackage{amsmath}
@@
 \usepackage{multirow}
+\usepackage{multicol}
 \usepackage{tikz}
 \usetikzlibrary{arrows,automata}
-\usepackage{comment}
-
-\usepackage{graphicx}
-\usepackage{caption}
-\usepackage{subcaption}
-\usepackage{listings}
-
-\lstset{
-  basicstyle=\ttfamily,
-  mathescape
-}
-
-\usepackage{multicol}
-\usepackage{arydshln}
 \usetikzlibrary{calc,backgrounds,positioning,fit}
@@
 
-\begin{document}
+%PDF Info Is Required:
+\pdfinfo{
+	/Title (Model Recognition as Planning)
+	/Author (Diego Aineto, Sergio Jimenez, Eva Onaindia, Miquel Ramirez)}
 
@@
 
-%\author{\#92}
-% Commented for blind submission
 \author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
-{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
-{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
-{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
-{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
+Departamento de Sistemas Inform\'aticos y Computaci\'on\\
+Universitat Polit\`ecnica de Val\`encia.\\
+Camino de Vera s/n. 46022 Valencia, Spain\\
+\{dieaigar,serjice,onaindia\}@dsic.upv.es
 \And Miquel Ram\'irez\\
-{\scriptsize School of Computing and Information Systems}\\
-{\scriptsize The University of Melbourne}\\
-{\scriptsize Melbourne, Victoria. Australia}\\
-{\scriptsize miquel.ramirez@unimelb.edu.au}}
+School of Computing and Information Systems\\
+The University of Melbourne\\
+Melbourne, Victoria. Australia\\
+miquel.ramirez@unimelb.edu.au}
 
+\begin{document}
 \maketitle
@@
 Given a \emph{partially observed plan execution}, and a \emph{set of possible planning models}, model recognition is the task of identifying the model in the set with the highest probability of producing the given observation.
-Model recognition enables the identification of the agent'e behaviour without need of being aware of the agent's intention.
-
-%As an example, imagine you are given a set of navigation policies for a $N\times N$ grid.
-Figure~\ref{fig:model-example} shows an example of such policy, according to which, the agent can only move right when he is at an even row ({\tt\small $q0$} holds), and left when he is at an odd row ({\tt\small $q1$} holds).
-Other models that shape different strategies of robot navigation can be defined within the same state variables altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated.
-Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, model recognition aims to identify the navigation policy that better explains this observation.
+Model recognition enables the identification of the agent's behaviour without need of being aware of the agent's intention.
 
@@
 Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, model recognition aims to identify the navigation policy (model) that better explains this observation.
-
-%As an example of model recognition, imagine a robot in a $N\times N$ grid whose navigation is determined by the \strips\ model of Figure~\ref{fig:model-example}.
-According to this model, when {\tt\small $q0$} ({\tt\small $q1$}) holds, the robot is at an even (odd) row and can move right (left) by incrementing (decrementing) its x-coordinate.
-Other models that shape different strategies of robot navigation can be defined within the same state variables altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated.
-Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, model recognition aims to identify the robot action schemata that explains this observation.
-
-%the robot can increment its {\em x-coordinate} when {\tt\small $q0$} holds (i.e.
-at even rows if {\tt\small $q0$} holds initially) and decrement it when {\tt\small $q1$} holds (at odd rows if {\tt\small $q0$} holds initially).
-Besides this model, other models that shape different kinds of robot navigation can be defined within the same state variables (e.g.
-altering the way {\tt\small $q0$} and {\tt\small $q1$} are required and updated).
-Given the partially observed plan execution of Figure~\ref{fig:grid-example}, where state variables {\small\tt $q0$} and {\small\tt $q1$} are unobserved, {\em model recognition} aims to identify the robot action schemata that explains this observation.
 
@@
   \end{tiny}
- \caption{\small Example of a \strips\ action schemata (given in PDDL) for robot navigation in a $N\times N$ grid.}
+ \caption{Example of an action schemata (given in PDDL) for robot navigation in a $N\times N$ grid.}
    \label{fig:model-example}
@@
 With this in mind, model recognition resembles other tasks like \emph{model reconciliation}, which uses model editing to conform two PDDL models, where one of them can be an annotated model or an empty model, with respect to a fully observed optimal plan computed with one of the two models~\cite{ChakrabortiSZK17,sreedharan2018handling,Chakraborti18}.
-Also, the \emph{model drift} approach estimates the error between a computational model, and a ground-truth model which changes as the domain evolves, as the normalized symmetric difference of the propositions that appear in the preconditions and effects of both models.
+Also, the \emph{model drift}~\cite{BryceBB16} approach estimates the error between a computational model, and a ground-truth model which changes as the domain evolves, as the normalized symmetric difference of the propositions that appear in the preconditions and effects of both models.
 In this case, plan observations are used to hypothesize variations of the model.
 
-%, which are then checked via observing query answers, validity of plans or allowing manual edits to the planning model.
-This mixed-initiative approach estimates the error between the two models as the normalized symmetric difference of the propositions that appear in the preconditions and effects of both models.
-
-Unlike model reconciliation and model drift, our proposal computes the model $\mathcal{M'}$ from \emph{partial observations} of plan executions that contain no actions and incomplete -- even empty- intermediate states.
-Moreover, we leverage the two-model edit distance to elaborate a probabilistic distribution of a model $\mathcal{M}$, which is used to posit the model recognition task as a \emph{classification task} with as many classes as models in the set.
+Unlike model reconciliation and model drift, our proposal computes the model $\mathcal{M'}$ from \emph{partial observations} of plan executions that contain no actions and incomplete -- even empty-- intermediate states.
+Moreover, we leverage the two-model edit distance to elaborate a probabilistic distribution of a model $\mathcal{M}$, which is used to posit the model recognition task as a \emph{classification task} with as many classes as input models.
 This way, we account for the possibly different prior probability of each model, a feature that cannot be considered with purely distance-based metrics.
@@
 We provide a formalization of model recognition that is applicable to any planning model but we restrict our attention to \strips\texttt{-}compilable models.
-It is well-known that diverse automata representations like finite state controllers, {\sc GOLOG} programs or reactive policies can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,segovia2017generating}.
+It is well-known that diverse automata representations like finite state controllers, {\sc GOLOG} programs or reactive policies can be encoded as classical planning models~\cite{BaierFM07,Geffner:FSM:AAAI10,segovia2018computing,segovia2019computing}.
 As for the evaluation, the effectiveness of our approach is tested in three experiments, each encoding a set of different \strips\ models and all using empty-action observations: (1) a proof-of-concept classification task; (2) a simulation of a non-deterministic \emph{blocksworld} under different degrees of observability; and (3) a navigation task where observations contain an unbounded number of unobservable states.
@@
   \end{scriptsize}
- \caption{\small ({\em Left)} Robot navigating a $5\times 5$ grid.
-{\em (Right)} Automata to control that the robot only increments its x-coordinate when $q_0$ holds (actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and switch the automata state).}
+ \caption{(Left) Robot navigating a $5\times 5$ grid.
+(Right) Automata to control that the robot only increments its x-coordinate when $q_0$ holds (actions {\tt inc-y} and {\tt dec-y} update the robot y-coordinate and switch the automata state).}
 \label{fig:grid-example}
@@
 \label{sec:background}
-%This section formalizes the models for {\em classical planning} and for the {\em observation} of the execution of a classical plan.
 
@@
 We explicitly include negative literals $\neg f$ in states and so $|s|=|F|$ and the size of the state space is $2^{|F|}$.
-
-%A {\em state} $s$ is a full assignment of values to fluents and we explicitly include negative literals $\neg f$ in states; i.e.
-$|s|=|F|$, so the size of the state space is $2^{|F|}$.
-Like in PDDL~\cite{fox2003pddl2}, we assume that fluents $F$ are instantiated from a set of {\em predicates} $\Psi$.
-Each predicate $p\in\Psi$ has an argument list of arity $ar(p)$.
-Given a set of {\em objects} $\Omega$, the set of fluents $F$ is induced by assigning objects in $\Omega$ to the arguments of predicates in $\Psi$; i.e.~$F=\{p(\omega):p\in\Psi,\omega\in\Omega^{ar(p)}\}$ such that $\Omega^k$ is the $k$-th Cartesian power of $\Omega$.
-
-%A {\em planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}.
-An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$, {\em positive effects}, $\eff^+(a)\in\mathcal{L}(F)$, and {\em negative effects} $\eff^-(a)\in\mathcal{L}(F)$.
-The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
-Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$.
-And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
 
@@
 
-%An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
-Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$:
-%\[
-%triggered(s,a_c)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
-%\]
-%
-%The result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\eff_c^-(s,a))\cup\eff_c^+(s,a)\}$, where $\eff_c^-(s,a)\subseteq triggered(s,a)$ and $\eff_c^+(s,a)\subseteq triggered(s,a)$ are, respectively, the triggered {\em negative} and {\em positive} effects.
-
 An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
@@
 
-%which indicates that we observe $l$ actions, $1\leq l\leq |\pi|$, and $m$ states, $1\leq m \leq |\pi|+1$, from $\tau(\pi,P)$:
-
 \begin{itemize}
@@
 Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
-Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
+Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$.
 In practice, the number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and the observed intermediate states will comprise a number of fluents between $[1,|F|]$.
-    %Exceptionally, $s_m^o$ cannot be fully unobservable for the purpose of our task (we will elaborate on this issue later on).
 \end{itemize}
@@
 We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
-Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
+Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${k\geq 1}$ is unknown but finite.
 In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of $\pi$.
 
-Figure~\ref{fig:grid-example} illustrates a partial observation of a six-state trajectory \{{\tt\scriptsize<(xcoord 0)(ycoord 0)>, <(xcoord 1)(ycoord 0)>, <(xcoord 2)(ycoord 0)>, <(xcoord 3)(ycoord 0)>, <(xcoord 3)(ycoord 1)>, <(xcoord 2)(ycoord 1)>}\}.
+Figure~\ref{fig:grid-example} illustrates a partial observation of a six-state trajectory $\langle${\tt\small\{(xcoord 0)(ycoord 0)\},\{(xcoord 1)(ycoord 0)\},\ldots,\{(xcoord 2)(ycoord 1)\}}$\rangle$.
 This observation only contains fluents of the predicates {\tt\small (xcoord ?v)} and {\tt\small (ycoord ?v)}, and the value of the remaining fluents, corresponding to predicates {\tt\small (next ?v1 ?v2)}, {\tt\small (q0)} and {\tt\small (q1)}, is unobservable in the six states.
@@
   \end{scriptsize}
- \caption{\small {\em Bayesian network} representing that model $\mathcal{M}$ is {\em transformable} into a model $\mathcal{M}'$ that produces a trajectory $\tau(\pi,P)$ that (1) reaches the goals in $P$ and (2) it is consistent with $\mathcal{O}(\tau)$.}
+ \caption{Bayesian network representing that model $\mathcal{M}$ is transformable into a model $\mathcal{M}'$ that produces a trajectory $\tau(\pi,P)$ that (1) reaches the goals in $P$ and (2) it is consistent with $\mathcal{O}(\tau)$.}
 \label{fig:net}
@@
 
-Our formalization of a model recognition task is valid for any planning model but our \emph{edit distance} measure is exclusively for \strips\ models and so we restrict our attention to \strips\texttt{-}compilable planning models\footnote{An edit distance for other planning models is also definable}.
+Our formalization of a model recognition task is valid for any planning model but our \emph{edit distance} measure is exclusively for \strips\ models and so we restrict our attention to \strips\texttt{-}compilable planning models\footnote{An edit distance for other planning models is also definable.}.
 Thus, we focus on the recognition of $\mathcal{M}\in M$, where $M$ is a set of \strips\ planning models and $\mathcal{M}$ defines the semantics of the actions through a set of \strips\ action schemata.
@@
 
-{\em A \strips\ action schema} $\xi$ is defined by four lists: A list of {\em parameters} $pars(\xi)$, and three list of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
+{\em A \strips\ action schema} $\xi$ is defined by a list of {\em parameters} $pars(\xi)$, and three lists of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
 
 \begin{definition}[Comparable \strips\ action schemata]
-Two \strips\ schemata $\xi$ and $\xi'$ are {\bf comparable} iff $pars(\xi)=pars(\xi')$, i.e, both share the same list of parameters.\footnote{In \strips\ models, $pars(\xi)=pars(\xi')$ implies the number of parameters must be the same.
-For other planning models that allow object typing, the equality implies that parameters share the same type}
+Two \strips\ schemata $\xi$ and $\xi'$ are {\bf comparable} iff $pars(\xi)=pars(\xi')$, i.e, both share the same list of parameters\footnote{In \strips\ models, $pars(\xi)=pars(\xi')$ implies the number of parameters must be the same.
+For other planning models that allow object typing, the equality implies that parameters share the same type.}.
 \end{definition}
@@
 
-\vspace{0.02cm}
-
 An observation of the execution of a plan generated with $\mathcal{M}$ further constraints the space of possible action schemata of $\mathcal{M}$.
@@
 \item A model can only produce one trajectory consistent with the given observation $\mathcal{O}$ and any such trajectory is equally likely.
-This is a strong assumption that can be relaxed in future work by weighting the edit distance against the trajectory length.
+As future work we plan to relax this assumption by weighting the edit distance against the trajectory length.
 \end{enumerate}
@@
 Under assumption (2), both $P(O|\tau)$ and $P(\tau|\mathcal{M}')$ are constant and can be taken out of the equation so we have that equation (1) simplifies once again to equation (3) with $\mathcal{M}^*$ being any model in $M^*$.
-
-%Again, we have that there is a single model $\mathcal{M}^*$ that can produce a given trajectory $\tau^*$.
-With this said, both $P(O|\tau^*)$ and $P(\tau^*|\mathcal{M}^*)$ are independent of $\mathcal{M} \in M$ so equation (1) simplifies once again to equation (3) under the previous two assumptions.
 
@@
 The $P(\mathcal{M^*}|\mathcal{M})$ probability is then given by equation (4), and computing the distance from $\mathcal{M}$ to the {\em closest consistents models} $\mathcal{M^*}$ is given by the {\em observation distance}, $d=\delta^o(\mathcal{M},\mathcal{O})$.
-
-%We know from {\em pattern recognition}~\cite{devroye2013probabilistic} that, if string symbols are uniformly random and independent, the distance of a given string to a fixed string follows a {\em Binomial distribution}.
-Moreover, the probability that a particular string will be within a distance $D$ is the Cumulative Distribution Function (CDF) for the {\em Binomial distribution}~\cite{wilcox1981review}:
-%\begin{align}
-%CDF(D)=\sum_{d=0}^D{{N}\choose{d}}  p^d (1-p)^{N-d}
-%\end{align}
-%where $N$ is the length of the string, and $p<0.5$ since we consider that the cost of applying an {\em edit operation} is higher than not applying it.
-%
-%With this regard, and considering that \strips\ models $\mathcal{M}\in M$ can be encoded with a propositional representation of fixed length $N$, we formulate the $P(\mathcal{M'}|\mathcal{M})$ probability distribution mapping the distance $\delta(\mathcal{M},\mathcal{M'})$ according to equation:
-%\begin{align}
-%P(\mathcal{M'}|\mathcal{M}) = p^d  (1-p)^{N-d}
-%\end{align}
-%where $d=\delta(\mathcal{M},\mathcal{M}')$.
-Likewise we can formulate $P(\mathcal{M^*}|\mathcal{M})$ with this same equation (5) but instead, the parameter $d$ is given by the {\em observation distance}, $d=\delta^o(\mathcal{M},\mathcal{O})$.
-%
-%In other words, we are modeling the editing of a \strips\ planning model as a {\em Bernoulli process} in which there is a sequence of $N$ independent events representing $N$ binary decisions (the $N$ possible applications of the edition operations) such that for every of these events $P(X=\top)=p$ and $P(X=\perp)=1-p$.
 
@@
 \item $G'$ contains the necessary fluents to check that all the observed actions and states of $\mathcal{O(\tau)}$ are generated correctly following the edited model $\mathcal{M'}$.
-    %Here we include the necessary fluents to check the final reached state is consistent with the final state $s_m^o$ of $\mathcal{O(\tau)}$.
 \end{itemize}
@@
 Consider that $\mathcal{M}$ is the model of Figure~\ref{fig:model-example} except that the schema {\tt\small inc-x} is defined without preconditions and its positive/negative effects are swapped with respect to Figure~\ref{fig:model-example}.
-And consider an observation that only contains the initial and final state of Figure~\ref{fig:grid-example}; i.e., $\mathcal{O(\tau)}=\tup{s_0^o,s_m^o}=\tup{\{{\texttt{\small{(xcoord 0)(ycoord 0)}}}\}, \{{\texttt{\small{(xcoord 2)(ycoord 1)}}}\}}$.
+And consider an observation that only contains the initial and final state of Figure~\ref{fig:grid-example}; i.e., $\mathcal{O(\tau)}=\tup{s_0^o,s_m^o}=\langle$\{{\texttt{\small{(xcoord 0)(ycoord 0)}}}\},\{{\texttt{\small{(xcoord 2)(ycoord 1)}}}\}$\rangle $.
 The plan found by a planner to $P'$ is shown in figure~\ref{fig:plan-pdistance}.
@@
 \begin{figure}
-{\tt\tiny
+\begin{tiny}
+{\tt
 \begin{tabular}{ll}
@@
 }
- \caption{\small A plan for $\tup{\mathcal{M},\mathcal{O}}$ where $\mathcal{M}$ is a modification of the model of Figure \ref{fig:model-example} and $\mathcal{O}$ a partial observation from Figure~\ref{fig:grid-example}.}
+\end{tiny}
+ \caption{A plan for $\tup{\mathcal{M},\mathcal{O}}$ where $\mathcal{M}$ is a modification of the model of Figure \ref{fig:model-example} and $\mathcal{O}$ a partial observation from Figure~\ref{fig:grid-example}.}
 \label{fig:plan-pdistance}
@@
 
-In this section, we evaluate the empirical performance of our approach in three different applications of {\em model recognition}.
+In this section, we evaluate the empirical performance of our approach\footnote{Available at https://github.com/anonsub/model-recognition.} in three different applications of {\em model recognition}.
 We will assume in the three experiments that $\mathcal{O}(\tau)$ contains an \emph{empty sequence of observed actions} and so we will only work with the available observed states.
@@
 \begin{definition}[$\Phi$-observation]
-Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
+Given a subset of fluents $\Phi\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Phi$-observation of the execution of $\pi$ on $P$ iff, for every ${1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Phi$.
 \end{definition}
@@
 We set up a timeout of 1000s for the computation of $\delta^o(\mathcal{M},\mathcal{O})$, at which point it is assigned the maximum distance.
-The compilation source code, evaluation scripts and benchmarks are fully available at this anonymous repository {\em https://github.com/anonsub/model-recognition} so any experimental data reported in the paper can be reproduced.
 
-\subsection{Recognition of {\em regular automata}.} The first experiment, which doubles as a proof of concept, exploits {\em model recognition as planning} for a classical {\em string classification} problem.
+\subsection{Recognition of {\em regular automata}} The first experiment, which doubles as a proof of concept, exploits {\em model recognition as planning} for a classical {\em string classification} problem.
 In this experiment, the system receives (1) the string to classify (the $\mathcal{O}(\tau)$ observation) and (2) a set $M$ of different planning models, where each $\mathcal{M}\in M$ represents a different class; i.e.
@@
 For instance, executing the planning model that encodes the {\em regular automata} of Figure~\ref{fig:regautomatae} with the input string $abcdabcd$ produces the following 8-action plan {\small $(\langle a,q_0\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow q_3)$, $(\langle d,q_3\rangle\rightarrow \underline{q_4})$, $(\langle a,\underline{q_4}\rangle\rightarrow q_1)$, $(\langle b,q_1\rangle\rightarrow q_2)$, $(\langle c,q_2\rangle\rightarrow \underline{q_3})$, $(\langle d,q_3\rangle\rightarrow \underline{q_4})$}.
-
-%\begin{figure}
-%  \begin{scriptsize}
-%  \begin{center}
-%	\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=2cm,semithick]
-%	  \node[state] (A)              {$q_0$};
-%	  \node[state] (B) [right of=A] {$q_1$};
-%	  \node[state] (C) [right of=B] {$q_2$};
-%	  \node[state] (D) [right of=C] {\underline{$q_3$}};
-%	  \path
-%(A) edge  node {a} (B)
-%(B) edge  node {b} (C)
-%(C) edge  node {c} (D)
-%(D) edge [bend left]  node {a} (B)
-%;
-%	\end{tikzpicture}
-%  \end{center}
-%  \end{scriptsize}
-% \caption{\small A 3-symbol 4-state {\em regular automata} for recognizing the $(abc)^+$ language (\underline{$q_3$} is the acceptor state).}
-%\label{fig:regautomatae}
-%\end{figure}
 
@@
   \end{scriptsize}
- \caption{\small A 4-symbol and 5-state {\em regular automata} for recognizing the $(abcd)^+$ language (\underline{$q_4$} is the acceptor state).}
+ \caption{A 4-symbol and 5-state regular automata for recognizing the $(abcd)^+$ language (\underline{$q_4$} is the acceptor state).}
 \label{fig:regautomatae}
 \end{figure}
-
-%Note that the the string to classify is a {\em $\Phi$-observation} since it does not contain any information about the structure of the possible regular automata.
-This means that $\mathcal{O}(\tau)$ is a $\Phi$-observation where fluents representing the internal machine state are unknown and the applied transitions (actions) are unobserved.
 
@@
 
-%More precisely observations $\mathcal{O}(\tau)$ are $\Phi$-observations where: the applied automata transitions and the fluents representing the internal automata state are unobserved.
-
 Table \ref{tab:conf_matrix} shows the {\em confusion matrix} resulting from classifying the 100 strings with our method.
@@
 The outstanding results reveal that our approach is very suitable for the classification of generative planning models that are more restrictive than \strips\ models, as it is the case of regular automata.
-
-%These results prove the suitability of {\em model recognition as planning} for classification tasks where classes are given as generative planning models and examples are sequence of observations.
 
@@
 	\end{tabular}
-	\caption{Confusion matrix for {\em regular automata recognition}.}
+	\caption{Confusion matrix for regular automata recognition.}
 	\label{tab:conf_matrix}
@@
 
-% Given that the two assumptions are correct when there is full observability of the executed plan, we can easily conclude that the more observability, the fewer trajectories that explain $\mathcal{O}(\tau)$ and hence the \emph{more correct} the two assumptions.
-In order to show the assumptions validation:
-
 Figure \ref{fig:blocks_acc} shows the {\em accuracy} of our approach when identifying failures over 30 different observations across different degrees of observability.
@@
 
-%The overall high accuracy validates our assumption that the trajectory that better explains the observation is the shortest one.
-
-%The increasing trend of the accuracy as observability increases and so the number of possible explanation lowers, validates our two assumptions.
-
-%We limited the failures to be of one kind per observation however, it would be possible to recognize different combinations of them.
-
 \begin{figure}
 	\centering
-	\includegraphics[width=0.9\linewidth]{figures/blocks.eps}
-	\caption{Accuracy for the recognition of failures in a non-deterministic {\em blocksworld}.}
+	\includegraphics[width=.95\columnwidth]{fig_blocks.pdf}
+	\caption{Accuracy for the recognition of failures in a non-deterministic blocksworld.}
 	\label{fig:blocks_acc}
@@
 \subsection{Recognition of navigation policies}
-%This experiment is also aimed at validating the accuracy of our approach to identify the model that generates an observation under different degrees of observability.
+
 The novel aspect of this experiment is that, unlike the previous experiment, we do not guarantee that the observed states contain at least one fluent.
@@
 	\centering
-	\includegraphics[width=0.9\linewidth]{figures/navigation.eps}
+	\includegraphics[width=.95\columnwidth]{fig_navigation.pdf}
 	\caption{Classification accuracy for the recognition of navigation models.}
@@
 
-We show the effectiveness of our approach in three experiments that encode a set of different \strips\ models.
-The first experiment is a proof-of-concept \emph{classification task} that perfectly recognizes the regular automata of a given string.
+We show the effectiveness of our approach in three experiments that encode a set of different models.
+The first experiment is a proof-of-concept \emph{classification task} that recognizes the regular automata of a given string.
 The second experiment, where the number of trajectories explaining the observation is bounded, shows that a more accurate model prediction is achieved with more complete observations.
-Finally, the robustness of the approach is shown in the third experiment where actions and intermediates states are unobservable.
+The robustness of the approach is shown in the third experiment where actions and intermediates states are unobservable.
+A significant lesson from the experiments is that our approach allows us to predict behaviours that are not ruled by a goal-driven deliberative planning model; e.g., models of fixed programs or finite state machines (regular automata).
 
-A significant lesson from the experiments is that our approach allows us to predict behaviours that are not ruled by a goal-driven deliberative planning model such as, for instance, models of fixed programs or finite state machines (regular automata).
+\subsection*{Acknowledgments}
+This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.
+D.
+Aineto is partially supported by the {\it FPU16/03184} and S.
+Jim\'enez by the {\it RYC15/18009}.
+M.
+Ram\'irez research is partially funded by DST Group Joint \& Operations Analysis Division.
 
-%Last but not least, previous work on the learning of \strips\ action models also defined semantic error metrics to quantify the errors of a model with respect to the observation of a plan execution~\cite{yang2007learning}.
-Our approach for quantifying this error is based on the definition of a edit distance for the model which allow us to not accumulate the repetition of errors coming from the the same model flaw.
-
-%Remarkably, the extension of this piece of work to the FOND planning setting~\cite{muise2012improved} is straightforward by simply considering the {\em all-outcomes} determiniztion of the actions with non-determinitic effects~\cite{yoon2007ff}.
-An interesting research direction is however to understand how to apply our approach to planning models where the planning models include actions with probabilistic effects~\cite{younes2005first}.
-
-%\newpage
-
-\subsection*{Acknowledgements}
-This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.
-Diego Aineto is partially supported by the {\it FPU16/03184} and Sergio Jim\'enez by the {\it RYC15/18009}, both programs funded by the Spanish government.
-M.
-Ramirez research is partially funded by DST Group Joint \& Operations Analysis Division.
-
+\bibliography{planlearnbibliography}
 \bibliographystyle{aaai}
-\bibliography{planlearnbibliography}
 

### 2019-03-27 doc/lecture-notes-EBL/ebl.tex
Modified up to section 3

@@
 The paper presents a classical planning compilation for learning \strips\ action models from partial observations of plan executions.
-The compilation is flexible to different amount and kind of available input knowledge; learning examples can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just pairs of initial and final states where no intermediate action/state is observed.
+The compilation is flexible to different amounts and types of input knowledge, from learning samples that comprise partially observed intermediate states of the plan execution to samples in which only the initial and final states are observed.
 The compilation accepts also partially specified action models and it can be used to validate whether an observation of a plan execution follows a given \strips\ action model, even if the given model or the given observation are incomplete.
@@
 \section{Introduction}
-Besides {\em plan synthesis}~\cite{ghallab2004automated}, planning action models are also useful for {\em plan/goal recognition}~\cite{ramirez2012plan}.
-In both tasks, automated planners are required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
+
+Action models in planning are not only required for plan synthesis~\cite{ghallab2004automated} but also for other tasks like plan/goal recognition~\cite{ramirez2009plan,ramirez2012plan}.
+In both cases, automated planners are required to reason about action models that correctly and completely capture the possible world transitions~\cite{geffner:book:2013}.
 Unfortunately building planning action models is complex, even for planning experts, and this knowledge acquisition task is a bottleneck that limits the potential of AI planning~\cite{kambhampati:modellite:AAAI2007}.
 
-{\em Machine Learning} (ML) has shown to be able to induce a wide range of different kinds of models from examples~\cite{michalski2013machine}.
+{\em Machine Learning} (ML) techniques have shown to be suitable to learn a wide range of different kinds of models from examples~\cite{michalski2013machine}.
 The application of inductive ML to learning \strips\ action models, the vanilla action model for planning~\cite{fikes1971strips}, is not straightforward though:
+
 \begin{itemize}
-\item The {\em input} to ML algorithms (the learning/training data) is usually a finite vector that represents the value of some fixed object features.
-The input for learning planning action models is, however, the observation of plan executions (where each plan possibly has a different length, since plan length is not {\em a priori} bound, and may also refer to a different number of objects).
-\item The {\em output} of ML algorithms is usually a scalar value (an integer, in the case of {\em classification} tasks, or a real value, in the case of {\em regression} tasks).
+\item The input to ML algorithms (the learning/training data) is usually a finite vector that represents the value of some fixed object features.
+The input for learning planning action models is, however, the observation of plan executions, where each plan has a possibly different length (plan length is not a priori bounded) and refer to a different number of objects.
+\item The output of ML algorithms is usually a scalar value (an integer, in the case of classification tasks, or a real value, in the case of regression tasks).
 When learning action models the output is, for each action, the set of preconditions and effects that define the possible state transitions of a planning task.
@@
 
-Learning \strips\ action models is a well-studied problem with sophisticated algorithms such as {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}, which do not require full knowledge of the intermediate states traversed by the example plans.
+Learning \strips\ action models is a well-studied problem with sophisticated algorithms such as {\sc ARMS}~\cite{yang2007learning}, {\sc SLAF}~\cite{amir:alearning:JAIR08} or {\sc LOCM}~\cite{cresswell2013acquiring}.
+All of these learning systems are capable of dealing with partial or null observability of the intermediates states traversed along the plan execution but they also require a full specification of the sequence of actions of the learning examples.
 Motivated by recent advances on the synthesis of different kinds of generative models with classical planning~\cite{bonet2009automatic,segovia2017generating,segovia2018computing,segovia2019computing}, this paper describes a classical planning compilation approach for learning \strips\ action models.
-The compilation approach is appealing by itself, because it opens up the door to the bootstrapping of planning action models, but also because it is flexible to different amount and kind of available input knowledge:
+The compilation approach is appealing by itself, because it opens up the door to the bootstrapping of planning action models, but also because it is flexible to different amounts and types of available input knowledge:
+
 \begin{enumerate}
-\item {\em Learning examples} can range from a set of plans (with their corresponding initial and final states, as well as partially observed intermediate states) to just a pair of initial and final states where no intermediate state/action is observed.
-\item {\em Partially specified action models}, expressing {\em a priori} knowledge about the structure of actions, can also be provided to the compilation.
-In the extreme, the compilation can validate whether an observed plan execution is valid for a given \strips\ action model, even if the model is not fully specified or the input observation is incomplete.
+\item {\em Learning examples} can range from plans that comprise partially observed intermediate states of the plan execution to samples in which no intermediate state/action is observed, that is, only the initial and final states are observed.
+
+\item {\em Partially specified action models}, expressing prior knowledge about the structure of actions, can also be provided to the compilation.
+In the extreme, the compilation can validate whether an observed plan execution is consistent with a given \strips\ action model, even if the model is not fully specified or the input observation is incomplete.
 \end{enumerate}
@@
 \section{Background}
-This section formalizes the models we follow for {\em classical planning}, for the {\em observations} of executions of classical plans and for the explanation of a given observation.
+
+In this section we formalize the {\em classical planning} model, for the {\em observation} model to represent the execution of a classical plan and the model for the \emph{explanation of a given observation}.
 
@@
 A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, and $|\pi|=n$ denotes its {\em plan length}.
-The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
-A plan $\pi$ solves $P$ if the induced trajectory $\tau(\pi,P)$ holds that $G \subseteq s_n$.
+The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A plan $\pi$ solves $P$ if $G$ holds in the last state of the induced trajectory $\tau$; i.e., $G \subseteq s_n$.
 A solution plan is {\em optimal} iff its length is minimal.
@@
 \subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$ and a plan $\pi$, the \emph{observation of the trajectory} $\tau(\pi,P)$ is a sequence of partial states that captures what is observed from the execution of $\pi$ in $P$.
-Formally, $\mathcal{O}(\tau)=\tup{s_0^o,s_1^o \ldots , s_m^o}$, $s_0^o=I$ is a sequence of possibly {\em partially observable states} (except for the initial state $s_0^o$ which is fully observable).
+
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ that solves $P$, and the corresponding trajectory $\tau$ induced by the execution of $\pi$ in $I$, $\tau=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$; there exist as many observations of $\tau$ as combinations of observable actions and observable fluents of the states of $\tau$.
+\emph{The observation model of the trajectory} $\tau$ comprises all possible combinations of observable elements of $\tau$.
+We will refer to the set of observations of $\tau$ as $Obs(\tau)$.
+
+Formally, one observation in $Obs(\tau)$ is defined as $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$, $s_0^o=I$,  a sequence of possibly {\em partially observable states}, except for the initial state $s_0^o$ which is fully observable.
 A partially observable state is one in which $|s_i^o| < |F|$, {\small $1\leq i\leq m\leq n$}; i.e., a state in which at least a fluent of $F$ is not observable.
+It may be also the case that $|s_i^o| = 0$ when an intermediate state is fully unobservable.
+The \emph{minimal observation} needed by our model is $\mathcal{O}=\tup{s_0^o,s_1^o}$, where $s_0^0$ is the fully observable initial state and $s_1^o$ is a partially observable final state.
 
-The {\em observation model} comprises the case $|s_i^o| = 0$, when an intermediate state is fully unobservable.
-This model considers also {\em observed actions} as fluents that indicate the action applied in a given state.
-This means that a sequence of {\em observed actions} $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$ s.t.
-$a_i^o\in s_i^o$ ({\small $0\leq i\leq l <m$}).
-The number of {\em observed actions}, $l$, can then range from $0$ (in a fully unobservable action sequence) to $|\pi|$ (in a fully observed action sequence).
+The observation model can also include {\em observed actions} as fluents indicating the applied action in a given state.
+This means that a sequence of observed actions $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi=\tup{a_1, \ldots, a_n}$ such that $a_i^o \in s_{i-1}^o$, {\small $0\leq i \leq l$}.
+Consequently, the number of fluents that represent observed actions, $l$, can range from $0$ (in a fully unobservable action sequence) to $|\pi|=n$ (in a fully observed action sequence).
 
-The sequence of observed states $\mathcal{O}(\tau)$ must be {\em consistent} with the sequence of states in $\tau(\pi,P)$.
-In practice, the number $m$ of observed states ranges from 1 (the initial state, at least), to $|\pi|+1$, and each of the observed intermediate states comprises a number of fluents between $[1,|F|]$.
-This means that we assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
-Therefore, transiting between two consecutive observed states in $\mathcal{O}(\tau)$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
-In other words, having $\mathcal{O}(\tau)$ does not imply knowing the actual length of the trajectory $\tau(\pi,P)$.
+Given $\mathcal{O} \in Obs(\tau)$, the number of observed states of $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$ ranges from 2 (at least the initial and final state, as explained above) to $|\pi|+1$.
+The number of fluents of the full observable state $s_0^o$ will be $|F|$, or $|F|+1$ in case the fluent of the applied action in $s_0$ is also observed.
+Every observable intermediate state will comprise a number of fluents between $[1,|F|+1]$, where a single fluent may represent a sensing fluent of the state or the observation of the applied action.
 
 This observation model can also distinguish between {\em observable state variables}, whose value may be read from sensors, and {\em hidden} (or {\em latent}) {\em state variables}, that cannot be observed.
-Given a subset of fluents $\Gamma\subseteq F$ we say that $\mathcal{O}(\tau)$ is a $\Gamma$-observation of the execution of $\pi$ on $P$ iff, for every ${\small 1\leq i\leq m}$, each observed state $s_i^o$ only contains fluents in $\Gamma$.
+Given a subset of fluents $\Gamma\subseteq F$ we say that $\mathcal{O}$ is a $\Gamma$-observation of the execution of $\pi$ on $P$ iff for every observed state $s_i^o$, ${\small 1\leq i\leq m}$, $s_i^o$ only contains fluents in $\Gamma$.
 
 \subsection{Explaining observations with classical planning}
-Given a {\em classical planning frame} $\Phi=\tup{F,A}$ and a observation of the execution of a plan within the given planning frame $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$, then $P_\mathcal{O}$ is the classical planning problem that is built as follows $P_\mathcal{O}=\tup{F,A,s_0^o,s_m^o}$.
+
+In this section we will explore the relationship between a trajectory $\tau$ and an observation $\mathcal{O}$.
+Particularly, we are interested in determining the necessary conditions for $\mathcal{O}$ to belong to $Obs(\tau)$.
+When the membership $\mathcal{O} \in Obs(\tau)$ is established, we say that $\mathcal{O}$ is \emph{consistent} with $\tau$ or that $\tau$ \emph{explains} $\mathcal{O}$.
+
+For the sake of simplicity, and given that our observation model encodes the observed applicable actions as fluents in the corresponding state, we will denote a trajectory as $\tau=\tup{s'_0, s'_1, \ldots, s'_n}$, where $s'_i$ comprises a fluent representing the applicable action $a_{i+1}$ in $s'_i$.
+
+Given an observation $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$ and a trajectory $\tau=\tup{s'_0, s'_1, \ldots, s'_n}$, where $m \leq n$, $s_0^o = s'_0$ and $s_m^o \subseteq s'_n$, it holds that $\mathcal{O} \in Obs(\tau)$ iff $\tau$ embeds $\mathcal{O}$; i.e., if there is a monotonic function $f$ mapping the observation indices $j=1,2, \ldots,m$ into the trajectory indices $i=1,2, \ldots,n$ such that $s_j^o \subseteq s'_{f(j)}$.
+This definition is a generalization of the one introduced in~\cite{ramirez2009plan}, which states the conditions under which an action sequence satisfies an observation sequence.
+Since all the elements (sets) of $\mathcal{O}$ are associated to an element (set) of $\tau$, but not viceversa, the fluents of a set of $\mathcal{O}$  are all included in the corresponding set of $\tau$, we can say that $\tau$ is a superset of $\mathcal{O}$.
+All this means that transiting between two consecutive observed states in $\mathcal{O}$ may require the execution of more than a single action ($\theta(s_i^o,\tup{a_1,\ldots,a_k})=s_{i+1}^o$, where ${\small k\geq 1}$ is unknown but finite.
+In other words, the information of $\mathcal{O}$ does not imply knowing the actual length of the trajectory $\tau$.
+
+Given a planning frame $\Phi=\tup{F,A}$ and an observation of a plan execution, $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$, we define $P_\mathcal{O}$, within the given planning frame, as the planning problem that is built as follows: $P_\mathcal{O}=\tup{F,A,s_0^o,s_m^o}$.
+
+%\begin{definition}[Explanation]
+%We say that a plan $\pi$ {\em explains} $\mathcal{O}$ iff $\pi$ is a solution for $P_\mathcal{O}$ that is {\em consistent} with the state trajectory constraints imposed by the sequence of partial states $\mathcal{O}$.
+If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}$.
+%\end{definition}
 
 \begin{definition}[Explanation]
-We say that a plan $\pi$ {\em explains} $\mathcal{O}$ iff $\pi$ is a solution for $P_\mathcal{O}$ that is {\em consistent} with the state trajectory constraints imposed by the sequence of partial states $\mathcal{O}$.
-If $\pi$ is also optimal, we say that $\pi$ is the {\em best explanation} for the input observation $\mathcal{O}$.
+A plan $\pi$ (or the trajectory $\tau$) \textbf{explains} $\mathcal{O}$ iff $\pi$ is a solution for $P_\mathcal{O}$ and $\mathcal{O} \in Obs(\tau)$.
 \end{definition}
 
-The {\em observation} $\mathcal{O}$ is then a sequence of ordered {\em landmarks} for the $P_\mathcal{O}$ classical planning problem~\cite{hoffmann2004ordered}, because all the literals in the observation must be achieved by any plan that solves $P_\mathcal{O}$ and in the same order as are defined in the observation $\mathcal{O}$.
+There may exist more than one solution plan for $P_\mathcal{O}$, one or more of which will be optimal solutions if their plan length is minimal.
+Additionally, other solutions longer than the optimal plan can also be found.
+
+\begin{definition}[Best explanation]
+A plan $\pi$ (or the trajectory $\tau$) that solves $P_\mathcal{O}$ is the \textbf{best explanation} for $\mathcal{O}$ iff $|\pi|=n$ and for every other $\tau_i$ s.t.
+$\mathcal{O} \in Obs(\tau_i)$, $|\pi_i|>n$.
+\end{definition}
+
+That is, in case that $\pi$ is optimal, we say that $\pi$ is the \textbf{best explanation} for the input observation $\mathcal{O}$.
+
+The observation $\mathcal{O}$ can also be regarded as a sequence of ordered {\em landmarks} for the planning problem $P_\mathcal{O}$~\cite{hoffmann2004ordered} since all the fluents of the sets in $\mathcal{O}$ must be achieved by any plan that solves $P_\mathcal{O}$ and in the same order as defined in the observation $\mathcal{O}$.
 
@@
 
-A {\em solution} to a $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task is a model $\mathcal{M}'$ s.t.
-it is consistent with the headers of $\mathcal{M}$ and that it explains $\mathcal{O}$.
-We say that a model {\em explains} an observation $\mathcal{O}$ iff, when the $\tup{\rho,\theta}$ functions of the actions in the corresponding $P_\mathcal{O}$ problem are given by that model, there exists a solution plan for $P_\mathcal{O}$ that {\em explains} $\mathcal{O}$.
+A {\em solution} to a $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task is a model $\mathcal{M}'$ that is consistent with the headers of $\mathcal{M}$ and that explains $\mathcal{O}$.
+We say that a model $\mathcal{M}'$ explains an observation $\mathcal{O}$ iff there exists a solution plan for $P_\mathcal{O}=\tup{F,A,s_0^o,s_m^o}$, where the semantics of the set of actions $A$ are given by $\mathcal{M}'$, such that $\pi$ explains $\mathcal{O}$.
 The set of fluents $F\in P_\mathcal{O}$ is induced from $s_0^o\in \mathcal{O}$ since it represents a full state.
@@
 Considering only the syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
-The belonging of a $p \in \mathcal{I}_{\Psi,\xi}$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a refined propositional encoding that uses fluents of two types, $pre_{p,\xi}$ and $eff_{p,\xi}$, instead of the three fluents used in the original compilation~\cite{aineto2018learning}.
+Given $p \in \mathcal{I}_{\Psi,\xi}$, the belonging of $p$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a propositional encoding that uses fluents of two types, $pre_{p,\xi}$ and $eff_{p,\xi}$.
 The four possible combinations of these two fluents are summarized in Figure \ref{fig:combinations}.
@@
 \subsection{The sampling space}
-According to our {\em observation model} the minimal expression of an observation must comprise at least two state observations $\mathcal{O}(\tau)=\tup{s_0^o,s_m^o}$, a full initial state $s_0^o$ and a partially observed final state $s_m^o$.
-Figure~\ref{fig:observation} shows an example of a $\mathcal{O}(\tau)=\tup{s_0^o,s_m^o}$ observation that contains only two states.
+According to our {\em observation model} the minimal expression of an observation must comprise at least two state observations $\mathcal{O}=\tup{s_0^o,s_m^o}$, a fully observable initial state $s_0^o$ and a partially observed final state $s_m^o$.
+Figure~\ref{fig:observation} shows an example of $\mathcal{O}=\tup{s_0^o,s_m^o}$ observation that contains only two states.
 An initial state of the blocksworld where the robot hand is empty and there are two blocks ({\tt\small{blockB}} on top of  {\tt\small{blockA}}).
@@
 
-On the other hand, the maximal expression of an observation corresponds to a fully observed trajectory $\mathcal{O}(\tau)=\tau$ (meaning that all traversed states, and applied actions, are fully observed).
-In between there is a grey scale of different kinds of possible observations, including the observation of the initial state and the executed plan, that is frequently used for previous systems that learn planning action models such as {\sc ARMS}~\cite{yang2007learning} or {\sc SLAF}~\cite{amir:alearning:JAIR08}.
+On the other hand, the maximal expression of an observation corresponds to a fully observed trajectory $\mathcal{O}=\tau$, meaning that all traversed states, and applied actions, are fully observed.
+Between our minimal and maximal expressions of observation, there exists a whole range of possible degrees of observability.
+For example, the majority of learning systems such as {\sc ARMS}~\cite{yang2007learning} or {\sc SLAF}~\cite{amir:alearning:JAIR08} use observations that comprise the initial state and all the actions of the executed plan.
 
@@
 
-An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{haslum:axiomsoptimal:ijcai15}.
+An interesting research direction related to this issue is {\em domain reformulation} to use actions in a more efficient way, reduce the set of actions identifying dispensable information or exploiting features that allow more compact solutions like the {\em reachable} or {\em movable} features in the {\em Sokoban} domain~\cite{IvankovicH15}.
 

### 2019-03-28 doc/lecture-notes-EBL/ebl.tex
EBL: Added experiments

@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p,\xi}, \neg eff_{p,\xi}, mode_{prog}\},\\
+\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre_{p,\xi}, mode_{prog}\},\\
 \cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre_{p,\xi}\}.
@@
 
-For instance, given $name(\xi)=${\tt{\small stack}} and $\{${\tt{\small (pre\_stack\_holding\_v1),} {\tt\small (pre\_stack\_holding\_v2)}, {\tt\small(pre\_stack\_on\_v1\_v2)}, {\tt\small(pre\_stack\_clear\_v1)}, {\tt\small(pre\_stack\_clear\_v1)}}, $\ldots \}$, the insertion of each item $p \in {\mathcal I}_{\xi,\Psi}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$.
+For instance, given $name(\xi)=${\tt{\small stack}} and $\{${\tt{\small (pre\_stack\_holding\_v1)}, {\tt\small (pre\_stack\_holding\_v2)}, {\tt\small(pre\_stack\_on\_v1\_v2)}, {\tt\small(pre\_stack\_clear\_v1)}, {\tt\small(pre\_stack\_clear\_v1)}}, $\ldots \}$, the insertion of each item $p \in {\mathcal I}_{\xi,\Psi}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$.
 The same applies to effects $\{${\tt{\small (eff\_stack\_holding\_v1)}, {\tt\small(eff\_stack\_holding\_v2)}, {\tt\small (eff\_stack\_on\_v1\_v2)}, {\tt\small (eff\_stack\_clear\_v1)}, {\tt\small (eff\_stack\_clear\_v1),}}$\ldots \}$.
@@
 Hence, the execution of the insert actions of $\pi_\Lambda$ yield a state containing the valuation of the model fluents that shape every $\xi \in \mathcal{M}$.
-For example, executing the insert actions that shape the action model $name(\xi)=${\tt{\small putdown}} leads to a state containing the positive literals {\tt{\small (pre\_putdown\_holding\_v1),(eff\_putdown\_holding\_v1),\\ (eff\_putdown\_clear\_v1),
+For example, executing the insert actions that shape the action model $name(\xi)=$ {\tt{\small putdown}} leads to a state containing the positive literals {\tt{\small (pre\_putdown\_holding\_v1),(eff\_putdown\_holding\_v1),\\ (eff\_putdown\_clear\_v1),
 (eff\_putdown\_ontable\_v1),(eff\_putdown\_handempty)}}.
 
-\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{ar(\xi)}$.
+\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ built by the insert actions and bound to objects $\omega\subseteq\Omega^{ar(\xi)}$.
 These actions will be part of the postfix of the plan $\pi_\Lambda$ and they determine the application of the learned action models according to the values of the model fluents in the current state configuration.
@@
 
-Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if $name(\xi)=${\tt{\small stack}} has been correctly programmed by the insert actions.
+Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if $name(\xi)=$ {\tt{\small stack}} has been correctly programmed by the insert actions.
 Hence, while \textbf{insert actions} add the values of the \textbf{model fluents} that shape $\xi$, the \textbf{apply actions} add the values of the \textbf{fluents of $F$} that result from the execution of $\xi$.
@@
 	\begin{align*}
-	\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_{m,t}^o\cup\{test_{j-1}\}\cup \{\neg mode_{prog}\},\\
-	\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_j\} \cup \\
-	&\{\neg f\}_{\forall f\in s_{m,t}^o, f \notin s_{0,t+1}^o}\cup \{f\}_{\forall f\in s_{0,t+1}^o, f \notin s_{m,t}^o}.
+	\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_{m,t}^o\cup\{test_{j-1}\}\cup  
+	\{\neg mode_{prog}\},\\
+	\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg  
+	test_{j-1},test_j\} \cup \\
+	&\{\neg f\}_{\forall f\in F, f \notin s_{0,t+1}^o}\cup \{f\}_{\forall  
+		f\in s_{0,t+1}^o}.
 	\end{align*}
@@
 	  \hspace*{7pt}pre(\xi)=& \{p ~|~ pre_{p,\xi} \in s_g\}_{\forall p \in \Psi_\xi},\\
-	  \hspace*{7pt}del(\xi)=& \{p ~|~ pre_{p,\xi} \wedge eff_{p,\xi} \in s_g\}_{\forall p \in \Psi_\xi},\\
-	  \hspace*{7pt}add(\xi)=& \{p ~|~ \neg pre_{p,\xi} \wedge eff_{p,\xi} \in s_g\}_{\forall p \in \Psi_\xi}.
+	  \hspace*{7pt}del(\xi)=& \{p ~|~ pre_{p,\xi} \in s_g \wedge eff_{p,\xi} \in s_g\}_{\forall p \in \Psi_\xi},\\
+	  \hspace*{7pt}add(\xi)=& \{p ~|~ \neg pre_{p,\xi} \in s_g \wedge eff_{p,\xi} \in s_g\}_{\forall p \in \Psi_\xi}.
 	\end{align*}
@@
 
-The plain compilation has trouble learning preconditions that do not appear as negative effects since in this case no change can be observed between the {\em pre-state} and {\em post-state} of an action application.
-This is specially relevant for static predicates that never change and, hence, only appear as preconditions in the actions.
-To address this shortcoming and complete the list of learned preconditions, we apply a post-process based on the one proposed by the {\sc LOUGA} system~\cite{kuvcera2018louga}.
+An optimally solved learning task will learn the minimum set of required preconditions; i.e, those that are at the same time negative effects.
+Optionally, it is possible to infer the maximum set of preconditions that is consistent with the observation and the learned model.
+This is done via a post-process based on the one proposed by the {\sc LOUGA} system~\cite{kuvcera2018louga}.
 The intuition is going through every action counting the number of cases where a literal is present before the action is executed.
 If a literal is present in all the cases before the action, the literal is considered to be a precondition.
-Since intermediate states/actions may not be fully observed in a observation $\mathcal{O}$, we consider the actions/states found in the validation part of the solution plan $\pi_\Lambda$.
+This is done by traversing the actions/states found in the validation part of the solution plan $\pi_\Lambda$.
 For instance, in the example of Figure \ref{fig:plan-lplan}, the used sequence of actions is {\tt\small(unstack blockB blockA)}, {\tt\small(put-down blockB)}, {\tt\small(pick-up blockA)}, and {\tt\small(stack blockA blockB)}.
@@
 
+We have tested the proposed explanation-based learning approach in 12 IPC domains that satisfy the \strips\ requirement~\cite{fox2003pddl2}, taken from the {\sc planning.domains} repository~\cite{muise2016planning}.
+In our experiments, we use a set of 5 observations of length 5 to 7 as learning examples.
+Each observation corresponds to plan executions generated via random walks.
+All experiments are run on an Intel Core i5 3.10 GHz x 4 with 16 GB of RAM.
+
+The learned models are evaluated using the {\em precision} and {\em recall} metrics for action models proposed in \cite{aineto2018learning}, which compare the learned models against the reference model.
+Precision measures the correctness of the learned models, while recall measures their completeness.
+Formally:
+
+\[
+Precision=\frac{tp}{tp+fp}
+\]
+\[
+Recall=\frac{tp}{tp+fn}
+\]
+
+where $tp$ ({\em true positives}) is the number of predicates that appear in both the learned and reference action models, $fp$ ({\em false positives}) is the number of predicates that appear in the learned action model but not in the reference model, and $fn$ ({\em false negatives}) is the number of predicates that should appear in the learned action model but are missing.
+
+\subsection{Learning from labeled plans}
+For our first experiment, we use the setting typically followed by most approaches, that is, learning from observations consisting of initial and final states, and the full sequence of actions between these two.
+In this setting, the number of trajectories that explain a given observation is bounded by the length of the observation and further constrained by the observed sequence of actions.
+
+The results of this experiment are compiled in Table~\ref{tab:results_plans}.
+Precision ({\bf P}) and recall ({\bf R}) are computed separately for the preconditions ({\bf Pre}), positive effects ({\bf Add}) and negative effects ({\bf Del}), while the last two columns and the last row report average scores.
+The table show high scores across all domains, with an average precision of 0.93 and average recall of 0.86.
+Recall is noticeably lower for preconditions at 0.74, which is to be expected given that any relaxation on the preconditions of the reference model will still be able to generate an explanation for the observation.
+
+\begin{table}[h]
+	\begin{center}
+		\begin{tabular}{l|*{6}{p{1cm}|}|*{2}{p{1cm}|}}
+			& \multicolumn{2}{c|}{\bf Pre} & \multicolumn{2}{c|}{\bf Add} & \multicolumn{2}{c||}{\bf Del} & \multicolumn{2}{c}{\bf Global}\\ \cline{2-9}			
+			& {\bf P} & {\bf R} &{\bf P} & {\bf R} & {\bf P} & {\bf R} &  {\bf P} & {\bf R} \\
+			\hline
+			Blocks & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Driverlog & 0.9 & 0.64 & 0.56 & 0.71 & 0.86 & 0.86 & 0.78 & 0.73\\
+			Ferry & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86\\
+			Floortile & 0.68 & 0.68 & 0.89 & 0.73 & 1.0 & 0.82 & 0.86 & 0.74\\
+			Grid & 0.79 & 0.65 & 1.0 & 0.86 & 0.88 & 1.0 & 0.89 & 0.83 \\
+			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.89\\
+			Hanoi & 0.75 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.92 & 0.92\\
+			Miconic & 0.89 & 0.89 & 1.0 & 0.75 & 0.75 & 1.0 & 0.88 & 0.88\\
+			Satellite & 0.82 & 0.64 & 1.0 & 1.0 & 1.0 & 0.75 & 0.94 & 0.80\\
+			Transport & 1.0 & 0.70 & 0.83 & 1.0 & 1.0 & 0.80 & 0.94 & 0.83\\
+			Visitall & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0\\
+			Zenotravel & 1.0 & 0.64 & 0.88 & 1.0 & 1.0 & 0.71 & 0.96 & 0.79\\
+			\hline
+			\bf  & 0.90 & 0.74 & 0.93 & 0.92 & 0.96 & 0.91 & 0.93 & 0.86\\
+		\end{tabular}	
+	\end{center}
+	\caption{Precision and recall scores for learning tasks from labeled plans.}
+	\label{tab:results_plans}
+\end{table}
+
+\subsection{Learning from initial/final state pairs}
+Now, we evaluate our approach when observations are reduced to their minimal expression $\mathcal{O}=\tup{s_0^o,s_m^o}$; i.e., only the initial and final state are observed.
+In contrast to the previous experiment, this setting presents an unbounded number of trajectories consistent with the observation.
+Moreover, the planner must determine how many "gaps" need to be filled between the two observed states.
+
+Table~\ref{tab:results_min} summarizes the results obtained for this experiment.
+Values for the {\em Zenotravel} and {\em Grid} domains are not reported because no solutions were found under the given timeout of 1000 seconds.
+Although the learned models are able to produce explanations for the input observations, we can see that the values of precision and recall are significantly lower than in Table ~\ref{tab:results_plans}.
+This is indicative that the learned models are now considerably different from the reference ones, which is caused by the larger solution space originated from the removal of some observation constraints.
+
+\begin{table}[h]
+	\begin{center}
+		\begin{tabular}{l|*{6}{p{1cm}|}|*{2}{p{1cm}|}}
+			& \multicolumn{2}{c|}{\bf Pre} & \multicolumn{2}{c|}{\bf Add} & \multicolumn{2}{c||}{\bf Del} & \multicolumn{2}{c}{\bf Global}\\ \cline{2-9}			
+			& {\bf P} & {\bf R} &{\bf P} & {\bf R} & {\bf P} & {\bf R} &  {\bf P} & {\bf R} \\
+			\hline
+			Blocks & 0.75 & 0.67 & 0.86 & 0.67 & 0.86 & 0.67 & 0.82 & 0.67 \\
+			Driverlog & 1 & 0.29 & 0.5 & 0.71 & 0.67 & 0.29 & 0.72 & 0.43 \\
+			Ferry & 1 & 0.57 & 1 & 1 & 1 & 1 & 1 & 0.86 \\
+			Floortile & 0.57 & 0.36 & 1 & 0.64 & 0.67 & 0.36 & 0.75 & 0.45 \\
+			Grid & - & - & - & - & - & - & - & - \\
+			Gripper & 1 & 0.67 & 1 & 1 & 1 & 1 & 1 & 0.89 \\
+			Hanoi & 1 & 0.5 & 1 & 1 & 1 & 1 & 1 & 0.83 \\
+			Miconic & 0.5 & 0.11 & 0.67 & 0.5 & 0.5 & 0.33 & 0.56 & 0.31 \\
+			Satellite & 0.5 & 0.21 & 0.57 & 0.8 & 0.75 & 0.75 & 0.61 & 0.59 \\
+			Transport & 1 & 0.3 & 0.71 & 1 & 1 & 0.6 & 0.9 & 0.63 \\
+			Visitall & - & - & - & - & - & - & - & - \\
+			Zenotravel & 1 & 0.29 & 0.57 & 0.57 & 1 & 0.57 & 0.86 & 0.48 \\
+			\hline
+			& 0.83 & 0.4 & 0.79 & 0.79 & 0.85 & 0.66 & 0.82 & 0.61 \\			
+		\end{tabular}
+	\end{center}
+	\caption{Precision and recall scores for learning tasks from initial and final states.}
+	\label{tab:results_min}
+\end{table}
+
 \section{Conclusions}

### 2019-03-28 doc/lecture-notes-EBL/ebl.tex
Final updates

@@
 Our approach to address a learning task $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ is to compile it into a classical planning problem $P_{\Lambda}$.
-The intuition behind the compilation is that when $P_{\Lambda}$ is solved, the solution plan $\pi_\Lambda$ is a sequence of actions that build the output model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the observation ${\mathcal O}$.
+The intuition behind the compilation is that when $P_{\Lambda}$ is solved, the solution plan $\pi_\Lambda$ is a sequence of actions that build the output model $\mathcal{M'}$ and verify that $\mathcal{M'}$ explains the observation ${\mathcal O}$.
 
-A solution plan $\pi_\Lambda$ comprise then two differentiated blocks of actions: a plan prefix with a set of actions each defining the \textbf{insertion} of a fluent as a {\em precondition} or a {\em effect} of an action model; and a plan postfix with a set of actions that determine the \textbf{application} of the learned modes while successively \textbf{validating} the effects of the action application in every partial state in ${\mathcal O}$.
-Roughly speaking, in the \emph{blocksworld}, the format of the first set of actions of $\pi_\Lambda$ looks like {\tt{\small (insert\_pre\_stack\_holding\_v1)}, {\tt\small (insert\_eff\_stack\_clear\_v1)}, {\tt\small (insert\_eff\_stack\_holding\_v1)}}\ldots where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ is like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1)}, {\tt\small (validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states in ${\mathcal O}$.
+A solution plan $\pi_\Lambda$ includes then two differentiated blocks of actions: a plan prefix with a set of actions, each defining the \textbf{insertion} of a fluent as a precondition or an effect of an action model; and a plan postfix with a set of actions that determine the \textbf{application} of the learned modes while successively \textbf{validating} the effects of the action application in every partial state of ${\mathcal O}$.
+Roughly speaking, in the \emph{blocksworld}, the format of the first block of actions of $\pi_\Lambda$ looks like {\tt{\small (insert\_pre\_stack\_holding\_v1)}, {\tt\small (insert\_eff\_stack\_clear\_v1)}, {\tt\small (insert\_eff\_stack\_holding\_v1)}}\ldots where the first effect denotes a positive effect and the second one a negative fluent to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second block of actions of $\pi_\Lambda$ is like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1)}, {\tt\small (validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states in ${\mathcal O}$.
 
@@
 \begin{itemize}
-\item $F_{\Lambda}$ extends the set of fluents $F$ (obtained from $s_0^o$) with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of ${\mathcal O}$.
+\item $F_{\Lambda}$ extends the set of fluents $F$ (obtained from $s_0^o$) with the model fluents that are used to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of ${\mathcal O}$.
 Specifically, $F_{\Lambda}$ contains also:
@@
 \begin{enumerate}
+
 \item Actions for {\em inserting} a precondition or effect into $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models.
@@
 Among the \emph{inserting} actions, we find:
+
 \begin{itemize}
@@
 
-\item Actions for inserting a {\em effect} $p\in{\mathcal I}_{\xi,\Psi}$ into $\xi$.
+\item Actions for inserting an {\em effect} $p\in{\mathcal I}_{\xi,\Psi}$ into $\xi$.
 \begin{small}
@@
 
-\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ built by the insert actions and bound to objects $\omega\subseteq\Omega^{ar(\xi)}$.
+\item Actions for {\em applying} the action models $\xi\in\mathcal{M}$ built by the insert actions and bounded to objects $\omega\subseteq\Omega^{ar(\xi)}$.
 These actions will be part of the postfix of the plan $\pi_\Lambda$ and they determine the application of the learned action models according to the values of the model fluents in the current state configuration.

### 2019-04-16 doc/aaai19-infiniteMacros/aaai19-infiniteMacros.tex
first notes on the learning of recursive state features

\def\year{2019}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
\usepackage{aaai19}  %Required
\usepackage{times}  %Required
\usepackage{helvet}  %Required
\usepackage{courier}  %Required
\usepackage{url}\urlstyle{rm}  %Required
\usepackage{graphicx}  %Required
\frenchspacing  %Required
\setlength{\pdfpagewidth}{8.5in}  %Required
\setlength{\pdfpageheight}{11in}  %Required
\setcounter{secnumdepth}{0}

% Addiotional packages
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{amsthm}
\usepackage{multirow}
\usepackage{multicol}
\usepackage{tikz}
\usetikzlibrary{arrows,automata}
\usetikzlibrary{calc,backgrounds,positioning,fit}

\newcommand{\tup}[1]{{\langle #1 \rangle}}

\newcommand{\pre}{\mathsf{pre}}     % precondition
\newcommand{\del}{\mathsf{del}}     % effect
\newcommand{\add}{\mathsf{add}}     % effect
\newcommand{\eff}{\mathsf{eff}}     % effect
\newcommand{\cond}{\mathsf{cond}}   % conditional effect
\newcommand{\true}{\mathsf{true}}   % true
\newcommand{\false}{\mathsf{false}} % false
\newcommand{\PE}{\mathrm{PE}}     % precondition
\newcommand{\strips}{\textsc{Strips}}     % precondition

\newtheorem{theorem}{Theorem}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{definition}[theorem]{Definition}

%PDF Info Is Required:
\pdfinfo{}

\title{Synthesis of infinite macro-actions}

\author{
Departamento de Sistemas Inform\'aticos y Computaci\'on\\
Universitat Polit\`ecnica de Val\`encia.\\
Camino de Vera s/n. 46022 Valencia, Spain\\
\{dieaigar,serjice,onaindia\}@dsic.upv.es}

\begin{document}
\maketitle
\begin{abstract}
\end{abstract}

\section{Introduction}
\label{sec:introduction}
A {\em macro-action} is a parameterized sequence of actions that has the form of a standard classical planning action so it can be reused straightforward to enrich a given planning domain theory. {\em Macro-planning} is a well-studied field and a wide number of macro planners exist, e.g., Marvin [Coles and Smith, 2007], MUM [Chrpa, 2010], MACRO FF [Boteaet al., 2005a], or DBMP/S [Hofmann et al., 2017] that leverage macro-actions to reduce the depth of the planning search space. All these existing macro-planners are restricted to {\em macro-actions} that represent a {\em finite and fixed} sequence of actions.

In this work we extend the notion of macros and define macro-actions that refer to a possibly {\em infinite} sequence of actions, then we show a representation of this kind of macro-actions using features with the form of {\em recursive derived predicates}. Last but not least, we present a classical planing compilation approach for the synthesis of {\em infinite macro-actions} from plans with an off-the-shelf classical planner.

To illustrate the notion of {\em infinite macros} Figure~\ref{fig:infinite-macro1} shows {\tt infinite-WALK}, an example of an {\em infinite macro-action} that is represented with the PDDL recursive derived predicate {\tt infinite-path} and that encodes a possibly infinite sequence of {\tt WALK} actions from the driverlog domain.

\begin{figure}
  \begin{tiny}
  \begin{verbatim}
;;;   
;;; Origininal action
;;; 
(:action WALK
  :parameters (?d - driver ?from ?to - location)
  :precondition (and (at ?d ?from) (path ?from ?to))
  :effect (and (not (at ?d ?from)) (at ?d ?to)))

;;; 
;;; Infinite macro-action
;;;

(:action infinite-WALK
  :parameters (?d - driver ?from ?to - location)
  :precondition (and (at ?d ?from) (infinite-path ?from ?to))
  :effect (and (not (at ?d ?from)) (at ?d ?to)))

(:derived infinite-path (?from ?to - location)
 (or (path ?from ?to)   ;;; base case
     (and (exists (?x - location) ;;; recursive case
                  (and (path ?from ?x) (infinite-path ?x ?to)))))) 
  \end{verbatim}
  \end{tiny}
 \caption{{\em Infinite macro-action} represented with the PDDL recursive derived predicate {\tt infinite-path} and that encodes a possibly inifinite sequence of {\tt WALK} actions from the driverlog domain.}
   \label{fig:infinite-macro1}
\end{figure}

\section{Background}
\label{sec:background}
This section introduces the classical plannning model and the classical planning compilation for the learning of \strips\ actions.

\subsection{Classical planning with conditional effects}
$F$ is the set of {\em fluents} or {\em state variables} (propositional variables). A {\em literal} $l$ is a valuation of a fluent $f\in F$, i.e. either~$l=f$ or $l=\neg f$. $L$ is a set of literals that represents a partial assignment of values to fluents, and $\mathcal{L}(F)$ is the set of all literals sets on $F$, i.e.~all partial assignments of values to fluents. A {\em state} $s$ is a full assignment of values to fluents. We explicitly include negative literals $\neg f$ in states and so $|s|=|F|$ and the size of the state space is $2^{|F|}$.

A {\em planning frame} is a tuple $\Phi=\tup{F,A}$, where $F$ is a set of fluents and $A$ is a set of \emph{actions}. An action $a\in A$ is defined with {\em preconditions}, $\pre(a)\in\mathcal{L}(F)$,  and {\em effects} $\eff(a)\in\mathcal{L}(F)$. The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$. Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$. And the result of applying $a$ in $s$ is $\theta(s,a)=\{s\setminus\neg\eff(a))\cup\eff(a)\}$, with $\neg\eff(a) = \{\neg l : l \in \eff(a)\}$.

A {\em planning problem} is defined as a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state in which all the fluents of $F$ are assigned a value true/false and $G$ is the goal set. A {\em plan} $\pi$ for $P$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, and $|\pi|=n$ denotes its {\em plan length}. The execution of $\pi$ in the initial state $I$ of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$. The {\em trajectory length} of $\tau(\pi,P)$ is given by the plan length of $\pi$. A trajectory $\tau(\pi,P)$ that solves $P$ is one in which $G \subseteq s_n$.

An action $a_c\in A$ with conditional effects is defined as a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$. Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}. An action $a_c\in A$ is applicable in a state $s$ if and only if $\pre(a_c)\subseteq s$, and the {\em triggered effects} resulting from the action application are the effects whose conditions hold in $s$: $triggered(s,a_c)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E$.

\subsection{Learning action models as planning}
\label{FAMA}

The approach for learning \strips\ action models presented in~\cite{aineto2018learning}, which we will use as our baseline learning system (hereafter BLS, for short), is a compilation scheme that transforms the problem of learning the preconditions and effects of action models into a planning task $P'$. A \strips\ \emph{action model} $\xi$ is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.

The BLS receives as input an empty domain model, which only contains the headers of the action models, and a set of observations of plan executions, and creates a propositional encoding of the planning task $P'$. Let $\Psi$ be the set of {\em predicates}\footnote{The initial state of an observation is a full assignment of values to fluents, $|s_0|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0$.} that shape the variables $F$. The set of propositions of $P'$ that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$. For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), \\
on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}. Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi$.

The decision as to whether or not an element of ${\mathcal I}_{\xi,\Psi}$ will be part of $pre(\xi)$, $del(\xi)$ or $add(\xi)$ is given by the plan that solves $P'$. Specifically, two different sets of actions are included in the definition of $P'$: \emph{insert actions}, which insert preconditions and effects on an action model; and \emph{apply actions}, which validate the application of the learned action models in the input observations. Roughly speaking, in the \emph{blocksworld} domain, the insert actions of a plan that solves $P'$ will look like {\tt{\footnotesize(insert\_pre\_stack\_holding\_v1)}},\\
{\tt{\footnotesize(insert\_eff\_stack\_clear\_v1),\\
(insert\_eff\_stack\_clear\_v2)}}, where the second action denotes a positive effect and the third one a negative effect both to be inserted in the model of {\tt{\small stack}}; and the second set of actions of the plan that solves $P'$ will be like {\tt{\small (apply\_unstack blockB blockA),(validate\_1),(apply\_putdown blockB),(validate\_2)}}, where the {\tt {\small validate}} actions denote the points at which the states generated through the {\tt {\small apply}} actions must be validated with the observations of plan executions.

In a nutshell, the output of the BLS compilation is a plan that completes the empty input domain model by specifying the preconditions and effects of each action model such that the validation of the completed model over the input observations is successful.

\section{Synthesis of infinite macro-actions}
\label{sec:infinite}

\section{Synthesis of infinite macro-actions with classical planning}
\label{sec:asPlanning}
Our approach is leveraging the classical planning compilation for the learning of strips actions to learn the preconditions and effects of two actions, one that represents the {\em case base} and another one that represents the {\em recursive case} of a recursively defined predicate~\cite{aineto2018learning}. Figure~\ref{fig:infinite-macro2} shows the two \strips\ actions that represent a macro-action encoding a possibly inifinite sequence of WALK ations from the driverlog domain.

\begin{figure}
  \begin{tiny}
  \begin{verbatim}
(:action WALK-baseCase
  :parameters (?d - driver ?from ?x ?to - location)
  :precondition (and (at ?d ?x) (evaluating-path ?from ?to) (path ?x ?to))
  :effect (and (not (at ?d ?x)) (at ?d ?to)
               (infinite-path ?from ?to)
               (not (evaluating-path ?from ?to))))

(:action WALK-recursiveCase
  :parameters (?d - driver ?from ?x ?to - location)
  :precondition (and (evaluating-path ?from ?x) (at ?d ?x) (path ?from ?x))
  :effect (and (not (at ?d ?from)) (at ?d ?x)
               (not (evaluating-path ?from ?x))
               (evaluating-path ?from ?to))) 
  \end{verbatim}
  \end{tiny}
 \caption{Two \strips\ actions representing a macro-action that encodes possibly inifinite sequence of WALK ations.}
   \label{fig:infinite-macro2}
\end{figure}

Figure~\ref{fig:plan} shows a sequential plan for sythesizing the {\em infinite macro-action} shown Figure~\ref{fig:infinite-macro2}. The plan is sythesized in a graph of locations that corresponds to a five-nodes linked list.
\begin{figure}
  \begin{tiny}
  \begin{verbatim}
  actions for programming the precs/effs of the strips action WALK-baseCase
  actions for programming the precs/effs of WALK-recursiveCase
  (apply-WALK-recursiveCase d1 l1 l1 l2)  
  (apply-WALK-recursiveCase d1 l1 l2 l3)  
  (apply-WALK-recursiveCase d1 l1 l3 l4)
  (apply-WALK-baseCase d1 l1 l4 l5)  
  \end{verbatim}
  \end{tiny}
 \caption{Sequential plan for learning the {\em macro-action} that encodes possibly inifinite sequence of WALK ations.}
   \label{fig:plan}
\end{figure}

\subsection*{Acknowledgments}
This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R. D. Aineto is partially supported by the {\it FPU16/03184} and S. Jim\'enez by the {\it RYC15/18009}. M. Ram\'irez research is partially funded by DST Group Joint \& Operations Analysis Division.

\bibliography{planlearnbibliography}
\bibliographystyle{aaai}

\end{document}

### 2019-05-01 doc/aij18/minor-revision/background.tex
AIJ18 minor revision folder created

--- 2018-12-26 doc/aij18/major-revision/background.tex
@@
 A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not assign conflicting values to any fluent).
-\textcolor[rgb]{1.00,0.00,0.00}{The complement of $L$ is defined as $\neg L=\{\neg l:l\in L\}$.} We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
+The complement of $L$ is defined as $\neg L=\{\neg l:l\in L\}$.
+We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$, i.e.~all partial assignments of values to fluents.
 
-\textcolor[rgb]{1.00,0.00,0.00}{We will adopt the \emph{open world assumption}, that is, what is not known to be true in a state is unknown, to implicitly represent the unobserved literals of states.
-Consequently, states will explicitly include positive literals ($f$) and negative literals ($\neg f$) such that literals that are not in a state are unknown or unobserved.} Hence, a {\em state} $s$ is a full assignment of values to fluents; i.e.
+We will adopt the \emph{open world assumption}, that is, what is not known to be true in a state is unknown, to implicitly represent the unobserved literals of states.
+Consequently, states will explicitly include positive literals ($f$) and negative literals ($\neg f$) such that literals that are not in a state are unknown or unobserved.
+Hence, a {\em state} $s$ is a full assignment of values to fluents; i.e.
 $|s|=|F|$, so the size of the state space is $2^{|F|}$.
@@
 An action $a\in A$ is applicable in a given state $s$ iff $pre(a)\subseteq s$, i.e.~if the literals $pre(a)$ hold in $s$.
-\textcolor[rgb]{1.00,0.00,0.00}{The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$.
+The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$.
 Note that subtracting the complement of $\eff(a)$ from $s$ ensures that $\theta(s,a)$ remains a well-defined state with positive and negative literals.
-Then:}
+Then:
 
 \begin{itemize}
-\item \textcolor[rgb]{1.00,0.00,0.00}{$\eff^+(a)\in\mathcal{L}(F)$ is the {\em positive effects} of $a$, the subset of action effects that assert a positive literal in the state resulting after the application of $a$}
-\item \textcolor[rgb]{1.00,0.00,0.00}{$\eff^-(a)\in\mathcal{L}(F)$ is the {\em negative effects} of $a$, the subset of action effects that assert a negative literal in the state resulting after the application of $a$}
+\item $\eff^+(a)\in\mathcal{L}(F)$ is the {\em positive effects} of $a$, the subset of action effects that assert a positive literal in the state resulting after the application of $a$
+\item $\eff^-(a)\in\mathcal{L}(F)$ is the {\em negative effects} of $a$, the subset of action effects that assert a negative literal in the state resulting after the application of $a$
 \end{itemize}
 
-\textcolor[rgb]{1.00,0.00,0.00}{Since we restrict our attention to \strips\ action models learning, we will assume the set of syntactic constraints imposed by \strips\ models, namely }that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$.
+Since we restrict our attention to \strips\ action models learning, we will assume the set of syntactic constraints imposed by \strips\ models, namely that $\eff^-(a)\subseteq \pre(a)$, $\eff^-(a)\cap \eff^+(a)=\emptyset$ and $\pre(a)\cap \eff^+(a)=\emptyset$.
 Additionally, actions $a\in A$ are instantiated from given action schemas, as in PDDL.
-
-%We say that an action $a\in A$ is {\em applicable} in a state $s$ iff $\pre(a)\subseteq s$.
-The result of applying $a$ in $s$ is the {\em successor state} denoted by $\theta(s,a)=\{s\setminus\eff^-(a))\cup\eff^+(a)\}$.
 
@@
 \begin{enumerate}
-\item we say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and \textcolor[rgb]{1.00,0.00,0.00}{there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$};  that is $\theta(s_i,\tup{a})=s_{i+1}$.
-\textcolor[rgb]{1.00,0.00,0.00}{This case clearly states that $\tau_s = s$, meaning that $\forall s_i \in \tau_s$, $s_i$ comprises all the literals of the corresponding state in the trajectory $s$ of the plan $\pi$.
-Formally, $\forall i, 1 \leq i  < n, |s_i| = |F|$.}
+\item we say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$;  that is $\theta(s_i,\tup{a})=s_{i+1}$.
+This case clearly states that $\tau_s = s$, meaning that $\forall s_i \in \tau_s$, $s_i$ comprises all the literals of the corresponding state in the trajectory $s$ of the plan $\pi$.
+Formally, $\forall i, 1 \leq i  < n, |s_i| = |F|$.
 \item we say that $\tau_s$ is a partially-observable (\PO) state trajectory if at least one intermediate state of $\tau_s$ is a partial assignment of values to fluents.
-\textcolor[rgb]{1.00,0.00,0.00}{Formally, $\exists i, 1 \leq i  < n, |s_i| < |F|$.} This means that one or more literals are missing in the intermediate $s_i$, all of which may be missing.
+Formally, $\exists i, 1 \leq i  < n, |s_i| < |F|$.
+This means that one or more literals are missing in the intermediate $s_i$, all of which may be missing.
 When all literals are missing, $s_i$ is a \emph{missing} or \emph{empty state} ($s_i = \emptyset$).
@@
 \item when \textbf{all} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, $\tau_s$ is a \emph{non-observable} (\NO) state trajectory.
-\textcolor[rgb]{1.00,0.00,0.00}{Formally, $\forall i, 1 \leq i < n,  s_i= \emptyset$; i.e., $|s_i| = 0$.}
+Formally, $\forall i, 1 \leq i < n,  s_i= \emptyset$; i.e., $|s_i| = 0$.
 \item when \textbf{none} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, we will refer to $\tau_s$ as a \POstar state trajectory.
-\textcolor[rgb]{1.00,0.00,0.00}{Formally, $\exists i, 1 \leq i < n, s_i \neq \emptyset$; i.e., $0 < |s_i| < |F|$.}
+Formally, $\exists i, 1 \leq i < n, s_i \neq \emptyset$; i.e., $0 < |s_i| < |F|$.
 \end{enumerate}
@@
 	     & {\bf \# intermediate states} & {\bf state type} \\ \hline
-    \FO & $n-1$  & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ is a full assignment \textcolor[rgb]{1.00,0.00,0.00}{$|s_i|=|F|$} \\ \hline
-    \multirow{1}{*}{\POstar} & $n-1$ & {\small $\exists i, 1 \leq i < n$}  \\ & & $s_i$ is a partial assignment \textcolor[rgb]{1.00,0.00,0.00}{$0 < |s_i|< |F|$}\\ \hline
-    \multirow{1}{*}{\PO} & $\leq n-1$ & {\small $\exists i, 1 \leq i < n$}   \\  & & $s_i$ is a partial assignment \textcolor[rgb]{1.00,0.00,0.00}{$|s_i|< |F|$}\\ \hline
-    \NO & 0 & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ \textcolor[rgb]{1.00,0.00,0.00}{is an empty state  $|s_i|=0$}
+    \FO & $n-1$  & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ is a full assignment $|s_i|=|F|$ \\ \hline
+    \multirow{1}{*}{\POstar} & $n-1$ & {\small $\exists i, 1 \leq i < n$}  \\ & & $s_i$ is a partial assignment $0 < |s_i|< |F|$\\ \hline
+    \multirow{1}{*}{\PO} & $\leq n-1$ & {\small $\exists i, 1 \leq i < n$}   \\  & & $s_i$ is a partial assignment $|s_i|< |F|$\\ \hline
+    \NO & 0 & {\small $\forall i, 1 \leq i < n$}  \\  & & $s_i$ is an empty state  $|s_i|=0$
 \end{tabular}
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{\FAMA can also deal with partial observability in the observed actions of $\tau$, that we will refer to as $\tau_a$.
-We identify three levels of observability, from the greatest to the lowest:}
+\FAMA can also deal with partial observability in the observed actions of $\tau$, that we will refer to as $\tau_a$. We identify three levels of observability, from the greatest to the lowest:
 
 \begin{enumerate}
-\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$.
+\item when \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$.
 In this case, $\tau_a$ contains all the necessary actions to transit every state $s_{i-1}$ to its corresponding successor state $s_{i}$, from $s_0$ to $s_n$.
-This is the type of input trace accepted by all the existing learning approaches (see section \ref{related_work} for details)}.
-\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence.
+This is the type of input trace accepted by all the existing learning approaches (see section \ref{related_work} for details).
+\item when \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence.
 In this case, at least one of the necessary actions of the plan $\pi$ is missing in $\tau_a$.
-Formally, $\exists i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.}
-\item \textcolor[rgb]{1.00,0.00,0.00}{when \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence.
+Formally, $\exists i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.
+\item when \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence.
 Formally, $\forall i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.
-That is, $\tau_a = \emptyset$.}
+That is, $\tau_a = \emptyset$.
 \end{enumerate}
@@
 Plan traces can be classified accordingly to the type of observed state trajectory (\FO, \POstar, \PO or \NO) and action sequence (\FO, \PO or \NO).
-\textcolor[rgb]{1.00,0.00,0.00}{In section \ref{task_definition}, we expose the impact of the combinations of observed state trajectories and observed action sequences when solving a learning task.}
-
-%Additionally, $\tau$ is a plan trace with PO states if $\tup{s_0, s_1, \ldots, s_n}$ is a PO state sequence; and a plan trace with PO actions if $\tup{a_1, \ldots, a_n}$ is a PO action sequence.
+In section \ref{task_definition}, we expose the impact of the combinations of observed state trajectories and observed action sequences when solving a learning task.
 
@@
 All the approaches analyzed in this work accept the more constrained definition of partial observations of intermediate states \POstar, two of them also allow the sequence of intermediate states to be empty (\PO) and the majority accept \NO state trajectories.
-\textcolor[rgb]{1.00,0.00,0.00}{Exceptionally, \LOCM is the only approach capable of learning from a fully-empty state trajectory, with neither initial nor final state.}
-
-%With the exception of \LOCM, a \NO state trajectory is the minimum state sequence required by the approaches the minimum state sequence of a plan trace required by the approaches that appear labeled as PO states* in the first column of Table \ref{table:models_comparison1} is $\tup{s_0, s_n}$.
-
-%Most approaches assume that a set of predicates and a set of action headers are provided alongside the input traces.
-Others do not explicitly say so but the fact is that predicates and actions headers are easily extractable from the state sequence and action sequence of the input plan traces, respectively.
-A requirement, however, is that the input plan traces comprise at least a grounded sample of every predicate and operator schema in the domain model.
+Exceptionally, \LOCM is the only approach capable of learning from a fully-empty state trajectory, with neither initial nor final state.
 
@@
 In the following, we present a comprehensive insight of the particularities of the seven systems presented in Table \ref{table:models_comparison1} and Table \ref{table:models_comparison2}.
-This exposition will also help us to highlight in section \label{task_definition} the value of our contribution \FAMA.
+This exposition will also help us to highlight in section \ref{task_definition} the value of our contribution \FAMA.
 
@@
 The \ARMS system became a benchmark in action-model learning, showing empirically that is is feasible lo learn a model in a reasonably efficient way using a weighted MAX-SAT even with \NO state trajectories.
-%Nevertheless, we argue that using the home-made error and redundancy metrics to evaluate the model correctness is questionable since a model that exhibits a wrong precondition or effect in any of its actions could still be eligible as the minimal-error output model.
 
@@
          \hline
-        %\multirow{2}{*}{\SLAF} & PO states and FO actions & existential quantifiers in $\pre$ \\ & & universal quantifiers in $\eff$ &  bayesian learning \\ & & & logical filtering \\
-         %\hline
 		\multirow{2}{*}{\LAMP} & \PO states &  quantifiers &  Markov logic networks \\  & \FO actions & logical implications &  \\
@@
          \hline
-         \textcolor[rgb]{1.00,0.00,0.00}{\LOUGA} & \textcolor[rgb]{1.00,0.00,0.00}{\NO states} & \textcolor[rgb]{1.00,0.00,0.00}{\strips} & \textcolor[rgb]{1.00,0.00,0.00}{Genetic algorithm} \\ & \textcolor[rgb]{1.00,0.00,0.00}{\FO actions} & \textcolor[rgb]{1.00,0.00,0.00}{negative preconditions} & \\
+         \LOUGA & \NO states & \strips & Genetic algorithm \\ & \FO actions & negative preconditions & \\
          \hline
-         \LOCMtwo & \textcolor[rgb]{1.00,0.00,0.00}{--- }&  predicates and types & Finite State Machines \\ & \FO actions & & \\
+         \LOCMtwo & --- &  predicates and types & Finite State Machines \\ & \FO actions & & \\
          \hline
@@
 
-%\LAMP allows observations of intermediate states to be empty up to a minimum percentage of 1/5 of observed states as well as different degrees of PO in the number of propositions in each state.
-
 \begin{table}
@@
          \hline
-       \textcolor{red}{\LOUGA} & \textcolor[rgb]{1.00,0.00,0.00}{cross-validation with} & \textcolor[rgb]{1.00,0.00,0.00}{redundant effects} & \textcolor[rgb]{1.00,0.00,0.00}{5} \\  & \textcolor[rgb]{1.00,0.00,0.00}{a test set of plan traces} & \textcolor[rgb]{1.00,0.00,0.00}{differences wrt the test set} & \textcolor[rgb]{1.00,0.00,0.00}{800 - 3200 actions (160 traces)}\\
+       \LOUGA & cross-validation with & redundant effects & 5 \\  & a test set of plan traces & differences wrt the test set & 800 - 3200 actions (160 traces)\\
          \hline
@@
 
-%On the other hand, a test comparing plan traces with and without noisy actions reveals that the accuracy in the case \emph{with noise} becomes more and more close to \emph{without noise} as the number of iterations of \AMAN algorithm increases because the distribution of noise in the plan becomes gradually closer to the real distribution.
-
 Another interesting approach that deals with noisy and incomplete observations of states is presented in \cite{MouraoZPS12}.
@@
 
-%To capture the knowledge from the crowd, \CAMA enumerates all possible preconditions and effects for each action, considering all predicates $p$  whose parameters are included by the parameters of the action $a$.
-\CAMA queries the crowed whether $p$ is a precondition/positive effect/negative effect of $a$, and human annotators reply \emph{yes}, \emph{no} or \emph{cannot tell}.
-The goal is to find an optimal estimator of the \texttt{true} labels given the observation, minimizing the average bit-wise error rate.
-
-\textcolor[rgb]{1.00,0.00,0.00}{One of the latest incorporations to the family of action model learning algorithms is \LOUGA.
+One of the latest incorporations to the family of action model learning algorithms is \LOUGA.
 This system uses a genetic algorithm to learn the effects of actions.
@@
 \LOUGA evaluates the learned models via cross-validation using the same metrics they use in their fitness function.
-In more detail, they measure (1) redundant positive and negative effects, (2) preconditions not met, and (3) literals observed in the input trace but not in the corresponding execution of the plan with the learned model.}
+In more detail, they measure (1) redundant positive and negative effects, (2) preconditions not met, and (3) literals observed in the input trace but not in the corresponding execution of the plan with the learned model.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{Finally, we present the Learning Object-Centred Models (\textbf{\LOCM}), possibly the most distinctive learning system due to its ability of learning with minimal input knowledge.} \textbf{\LOCM} only requires the \FO action sequence of the plan trace, without need for providing any information about the predicates or the state trajectory, not even the initial or final state~\cite{CresswellMW09,cresswell2013acquiring}.
+Finally, we present the Learning Object-Centred Models (\textbf{\LOCM}), possibly the most distinctive learning system due to its ability of learning with minimal input knowledge.
+\textbf{\LOCM} only requires the \FO action sequence of the plan trace, without need for providing any information about the predicates or the state trajectory, not even the initial or final state~\cite{CresswellMW09,cresswell2013acquiring}.
 The lack of available state information is overcome by exploiting assumptions about the structure of the actions.
-Particularly, \LOCM assumes that objects found in the same position in the header of actions are grouped as a collection of objects \textcolor[rgb]{1.00,0.00,0.00}{named \emph{sort}} whose defined set of states is captured by a parameterized Finite State Machine (FSM).
+Particularly, \LOCM assumes that objects found in the same position in the header of actions are grouped as a collection of objects named \emph{sort} whose defined set of states is captured by a parameterized Finite State Machine (FSM).
 The intuitive assumptions of \LOCM, like the continuity of object transitions or the association of parameters between consecutive actions in the training dataset, yield a learning model heavily reliant on the kind of domain structure.
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{The distinctive feature of the \LOCM family lies in the capacity to learn the state variables (fluents) because predicates are neither provided as input nor they are deducible from the plan trace as no state observability is allowed.
+The distinctive feature of the \LOCM family lies in the capacity to learn the state variables (fluents) because predicates are neither provided as input nor they are deducible from the plan trace as no state observability is allowed.
 In contrast, \FAMA and the rest of approaches either assume the set of predicates are provided alongside the input traces or assume they are extractable from the observed states of the plan trace, in which case the plan trace must comprise at least a grounded sample of every predicate.
-Similarly, the syntax of an action header (the action name and its parameters) is either extractable from the action sequence of the plan trace or it must be explicitly provided.}
-
- %Because {\sf LOCM} approaches induce similar models for domains with similar structures, they face problems at generating models for domains that are only distinguished by whether or not they contain static relations (e.g.
-{\em blocksworld} and {\em freecell}).
-In order to mitigate this drawback, {\sf LOP} applies a post-processing step after the {\sf LOCM} analysis which requires additional information about the plans, namely a set of optimal plans to be used in the learning phase.
-
-%Recently classical planning compilations have been defined to learn different kinds of generative models from examples.
-The existing compilations for computing FSCs for generalized planning follow a {\it top-down} approach that interleaves {\it programming} the FSC with validating it and hence, they tightly integrate planning and generalization.
-To keep the computation of FSCs tractable, they limit the space of possible solutions bounding the maximum size of the FSC.
-In addition, they impose that the instances to solve share, not only the domain theory (actions and predicates schemes) but the set of fluents~\cite{javi-Gplanning-ICAPS16} or a subset of {\it observable} fluents~\cite{Geffner:FSM:AAAI10}.
-Programs increase the readability of FSCs separating the control-flow structures from the primitive actions.
-Like FSCs, programs can also be computed following a {\it top-down} approach, e.g.~exploiting compilations that program and validate the program on instances with the same state and action space~\cite{javi-Gplanning-ICAPS16}.
-Since these {\it top-down} approaches search in the space of solutions, it is helpful to limit the set of different control-flow instructions.
-For instance using only {\it conditional gotos} that can both implement branching and loops~\cite{Jimenez15}.
+Similarly, the syntax of an action header (the action name and its parameters) is either extractable from the action sequence of the plan trace or it must be explicitly provided.

### 2019-05-01 doc/aij18/minor-revision/evaluation.tex
AIJ18 minor revision folder created

--- 2019-02-19 doc/aij18/major-revision/evaluation.tex
@@
 In this section we introduce the metrics used by \FAMA to evaluate the action models that result from solving a learning task $\Lambda$.
-\textcolor{red}{First, we will motivate the use of two standard syntactic metrics ({\em precision} and {\em recall}).
-Then, section \ref{syntactic_precision_recall} will define these metrics for planning models, and section \ref{semantic_precision_recall} will introduce a semantic evaluation measure that builds upon {\em precision} and {\em recall}}.
+First, we will motivate the use of two standard syntactic metrics ({\em precision} and {\em recall}).
+Then, section \ref{syntactic_precision_recall} will define these metrics for planning models, and section \ref{semantic_precision_recall} will introduce a semantic evaluation measure that builds upon {\em precision} and {\em recall}.
 Finally, section \ref{edit_distance} explains how \FAMA computes our novel semantic-based metrics.
 
-\textcolor{red}{We can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the the total number of all the possible preconditions and effects of an action model.
+We can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the the total number of all the possible preconditions and effects of an action model.
 This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM.
@@
 To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}.
-These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM:}
-
-%When the planning reference model of the input observations (i.e., the GTM) is available, the quality of the learned action models is measurable using two well-studied syntax-based metrics, {\em precision} and {\em recall}, commonly used in tasks such as information retrieval and recommender systems~\cite{davis2006relationship}.
-These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM.
-Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models:
+These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM:
 
@@
 	\begin{small}
-		\textcolor{red}{
+
 			\begin{align*}
@@
 			\end{align*}
-		}
+
 	\end{small}
@@
 Let $\mathcal{M}$ be a domain model and let $\mathcal{M'}$ be the GTM.
-We know that \textcolor{red}{$size(\mathcal{M}) = \sum_{\xi \in \mathcal{M}}\left|pre(\xi)\right| + \left|add(\xi)\right| + \left|del(\xi)\right|$} and by definition the number of preconditions and effects of the learned action models is equal to the sum of {\em true positives} and {\em false positives}; that is, $size(\mathcal{M}) = tp + fp$.
+We know that $size(\mathcal{M}) = \sum_{\xi \in \mathcal{M}}\left|pre(\xi)\right| + \left|add(\xi)\right| + \left|del(\xi)\right|$ and by definition the number of preconditions and effects of the learned action models is equal to the sum of {\em true positives} and {\em false positives}; that is, $size(\mathcal{M}) = tp + fp$.
 
@@
 \begin{mydefinition} \label{syn-precision} The precision of $\mathcal{M}$ relative to the GTM is defined as the fraction of the common preconditions and effects between $\mathcal{M}$ and the GTM among all prediconditions and effects of $\mathcal{M}$.
-%This gives an intuitive measure of the soundness of $\mathcal{M}$.
 	\begin{small}
@@
 \begin{mydefinition} \label{syn-recall} The recall of $\mathcal{M}$ relative to the GTM is defined as the fraction of the common preconditions and effects between $\mathcal{M}$ and the GTM among all preconditions and effects of the GTM.
-%This gives an intuitive measure of the completeness of $\mathcal{M}$.
 
@@
 
-\textcolor{red}{Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
+Intuitively, precision gives a notion of {\em soundness} while recall gives a notion of the {\em completeness} of the learned models.
 We interpret a sound learned model as one in which all preconditions and effects are correct with respect to the GTM, and so there is no need to remove anything.
 A complete model is one in which no precondition or effect is missing.
-%In more detail, precision is the fraction of relevant predicates among the predicates that make up the learned model, and recall is the fraction of relevant predicates in the learned model over all relevant predicates, understanding a relevant predicate as one that appears in the GTM.
-As an example, a precision of 0.5 means that only half of the predicates that make up the learned domain model are present in the GTM, while a recall of 0.5 means that only half of the predicates that make up the GTM are present in the learned domain model.}
+As an example, a precision of 0.5 means that only half of the predicates that make up the learned domain model are present in the GTM, while a recall of 0.5 means that only half of the predicates that make up the GTM are present in the learned domain model.
 
@@
 Semantic evaluation metrics add a distinctive value over the syntactic ones, which is that they evaluate the learned model with a set of observations of plan executions.
-\textcolor{red}{These metrics measure how well a model can reproduce a given plan execution, so the use of the word "semantic" here is in reference to the degree to which the learned model is able to capture the physics of the domain.} Semantic metrics are appropriate for scenarios where:
+These metrics measure how well a model can reproduce a given plan execution, so the use of the word "semantic" here is in reference to the degree to which the learned model is able to capture the physics of the domain.
+Semantic metrics are appropriate for scenarios where:
 
@@
 	\item The GTM is unknown. This is the most common scenario in ML, where models are both learned and evaluated with respect to datasets.
-%	\item The GTM is known but a test-based evaluation on a dataset is preferable (or needed as complementary to a syntactic evaluation).
+	\item We are interested in measuring the ability of a model to explain a given plan trace, which is a good indicator of how the model will perform in actual planning tasks.
 As a rule of thumb, it is preferable to evaluate the learned models wrt a dataset because a learned model can be semantically correct though syntactically incorrect (different from the GTM).
-We refer to this phenomenon as \emph{model reformulation}.
-	\item \textcolor{red}{We are interested in measuring the ability of a model to explain a given plan trace, which is a good indicator of how the model will perform in actual planning tasks.} As a rule of thumb, it is preferable to evaluate the learned models wrt a dataset because a learned model can be semantically correct though syntactically incorrect (different from the GTM).
 We refer to this phenomenon as \emph{model reformulation}.
@@
 
-%These semantic alterations typically appear in the learned models when the observed input data given in $\tau$ is scarce.
-Defining a proper semantic evaluation is key because the application of syntax-based metrics may report low scores for learned models that are actually {\em sound} and {\em complete} but correspond to {\em reformulations} of the GTM.
-In the following, we introduce a novel evaluation metric that is robust to different types of reformulation.
-
-\textcolor{red}{The \ARMS system was the first to show that a semantic evaluation can be done via validation of a set of plan traces with the learned model~\cite{yang2007learning}.
+The \ARMS system was the first to show that a semantic evaluation can be done via validation of a set of plan traces with the learned model~\cite{yang2007learning}.
 The underlying idea is that an error indication of the learned action models is obtained by counting the number of preconditions that are not satisfied during the execution of the plan trace with the learned models, similarly to the functionality provided by the automatic validation tool VAL~\cite{howey2004val} used in the IPCs.
@@
 In other words, modifying the plan trace to fit the model.
-Inspired by this approach, we present novel semantic-based error measure that builds upon the {\em precision} and {\em recall} metrics, and instead, determines the modifications required by a learned model to explain the given plan traces.}
-
-%Definitions \ref{syn-precision} and \ref{syn-recall} are syntax-based metrics to evaluate $\mathcal{M}$ with respect to the GTM.
-A semantic metric, on the other hand, evaluates $\mathcal{M}$ with respect to a given set of plan traces.
+Inspired by this approach, we present novel semantic-based error measure that builds upon the {\em precision} and {\em recall} metrics, and instead, determines the modifications required by a learned model to explain the given plan traces.
 
@@
 The intuition behind this evaluation is to {\em semantically} assess how well the learned domain model $\mathcal{M}$ explains a set of given observations of plan executions according to the amount of {\em edition} required by $\mathcal{M}$ to induce the observations.
-\textcolor{red}{The interpretation of a sound and complete model in the semantic perspective is slightly different from the syntactic one.
+The interpretation of a sound and complete model in the semantic perspective is slightly different from the syntactic one.
 In this case, a sound model is one were there is no need to remove any precondition or effect in order to be consistent with some given plan traces.
-A complete model is one that is consistent with the given plan traces without adding any of its preconditions and effects.} Unlike the semantic metric defined by ARMS, our novel semantic definitions of precision and recall are not sensitive to flaws in the action model that manifest more than once in the plan traces since the flaws are corrected only once in the learned models instead of at every intermediate state of the plan traces.
-
-%This semantic evaluation approach is again flexible to various amount and kind of available input knowledge.
+A complete model is one that is consistent with the given plan traces without adding any of its preconditions and effects.
+Unlike the semantic metric defined by ARMS, our novel semantic definitions of precision and recall are not sensitive to flaws in the action model that manifest more than once in the plan traces since the flaws are corrected only once in the learned models instead of at every intermediate state of the plan traces.
 
@@
 \label{edit_distance}
-
-%Our compilation is extensible to compute the {\em closest compliant set of action models} and hence, our semantic versions of the {\em precision} and {\em recall} metrics.
-This extension considers that the input models $\mathcal{M}$, is {\em non-empty} so instead of learning an action model from scratch we simply edit $\mathcal{M}$ until it satisfies the given input observations.
-In other words, now $\mathcal{M}$ is a set of given operator schemas, wherein each $\xi\in\mathcal{M}$ initially contains the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ sets.
-A solution to the classical planning task resulting from the extended compilation is a sequence of actions that:
 

### 2019-05-01 doc/aij18/minor-revision/experiments.tex
AIJ18 minor revision folder created

--- 2019-02-19 doc/aij18/major-revision/experiments.tex
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{After presenting the setup of the experiments in section \ref{setup}, we introduce three experiments that measure different aspects like the minimal number of traces necessary to obtain good quality models (section \ref{size_knowledge}), a comparison of the quality of models obtained by \FAMA compared to \ARMS (section \ref{compare_ARMS}) and an analysis of the quality of models when using a very small number of input traces in section \ref{minimal} In these three experiments, the quality of the models is measured using the syntactic-based precision and recall metrics presented in section \ref{syntactic_precision_recall}, that is, we measure the precision and recall of the learned models with respect to the GTM for each domain.
+After presenting the setup of the experiments in section \ref{setup}, we introduce three experiments that measure different aspects like the minimal number of traces necessary to obtain good quality models (section \ref{size_knowledge}), a comparison of the quality of models obtained by \FAMA compared to \ARMS (section \ref{compare_ARMS}) and an analysis of the quality of models when using a very small number of input traces in section \ref{minimal} In these three experiments, the quality of the models is measured using the syntactic-based precision and recall metrics presented in section \ref{syntactic_precision_recall}, that is, we measure the precision and recall of the learned models with respect to the GTM for each domain.
 Finally, the syntactic evaluation of section \ref{minimal} is compared against a semantic evaluation, i.e.
-how well the learned models are able to reproduce the input plan traces, in section \ref{semantic}.}
-
-%We also devote one experiment to show the suitability of the novel evaluation metrics introduced in section \ref{sec:evaluation}.
+how well the learned models are able to reproduce the input plan traces, in section \ref{semantic}.
 
@@
   \begin{enumerate}
-  \item \textcolor[rgb]{1.00,0.00,0.00}{Other planners such as {\sc FastDownward} were also tested but provided worse experimental results}
-  \item \textcolor[rgb]{1.00,0.00,0.00}{The ability of {\sc Madagascar} to deal with instances populated with dead-ends, such as our compiled problem $P_\Lambda$, is very helpful~\cite{lopez2015deterministic}.}
-  \item \textcolor[rgb]{1.00,0.00,0.00}{A SAT-based planner like {\sc Madagascar} is particularly suitable for tasks where the observed plan trace $\tau$ determines the horizon of the solution plan (case 1 presented in section \ref{task_definition})}.
+  \item Other planners such as {\sc FastDownward} were also tested but provided worse experimental results
+  \item The ability of {\sc Madagascar} to deal with instances populated with dead-ends, such as our compiled problem $P_\Lambda$, is very helpful~\cite{lopez2015deterministic}.
+  \item A SAT-based planner like {\sc Madagascar} is particularly suitable for tasks where the observed plan trace $\tau$ determines the horizon of the solution plan (case 1 presented in section \ref{task_definition}).
 In this case, a SAT-based planner solves the prefix of the solution plan in two time steps because the actions for inserting preconditions can be applied in parallel in a single time step and the same for the actions inserting the effects.
@@
 Figures \ref{fig:np_quality} and \ref{fig:np_time} show the quality of the models and computation time, respectively, for the case \FO/\PO.
-\textcolor{red}{The values plotted in these figures are averages over the 15 domains.} In Figure \ref{fig:np_quality} we see that, after three traces, precision stabilizes at 0.84 whereas recall stabilizes at 0.95.
+The values plotted in these figures are averages over the 15 domains.
+In Figure \ref{fig:np_quality} we see that, after three traces, precision stabilizes at 0.84 whereas recall stabilizes at 0.95.
 These results show that \FAMA does, in fact, not need big amounts of input knowledge to learn sound and complete models as opposite to other approaches in the literature where models are learned using around 100 traces (see Table \ref{table:models_comparison2}).
@@
 
-Figure~\ref{fig:pspace_quality} displays \textcolor{red}{the average precision and recall of the 15 learned models} in the scenario with unknown plan horizon.
+Figure~\ref{fig:pspace_quality} displays the average precision and recall of the 15 learned models in the scenario with unknown plan horizon.
 As expected, the quality of the learned models is lower than when the horizon of the plan is known.
-\textcolor{red}{The higher complexity of this setting is also reflected in the appearance of some timeouts when solving the learning task.
-The first timeout is found in the \emph{grid} domain at $|\mathcal{T}| = 3$; the \emph{floortile} times out with $|\mathcal{T}| = 4$; and by the time $|\mathcal{T}|$ reaches 10 the number of domains were no solution is found is 6, adding the \emph{npuzzle, parking, rovers and zenotravel} domains.} We can observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
+The higher complexity of this setting is also reflected in the appearance of some timeouts when solving the learning task.
+The first timeout is found in the \emph{grid} domain at $|\mathcal{T}| = 3$; the \emph{floortile} times out with $|\mathcal{T}| = 4$; and by the time $|\mathcal{T}|$ reaches 10 the number of domains were no solution is found is 6, adding the \emph{npuzzle, parking, rovers and zenotravel} domains.
+We can observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.
 The drop in the score is caused by the increasing number of timeouts, meaning that no solution is found in many tasks within the given time-bound, and consequently a value of $0$ for precision and recall is assigned to these experiments.
@@
 With unknown plan horizons, on the other hand, the learned models are generally more different from the GTM; while timeouts are the main cause of the drop in the score, we must point out that pure syntax-based metrics are not adequate to evaluate such under-constrained tasks since the phenomenon of \emph{reformulation} occurs and this largely impacts the results (we will provide experimental evidence of this in section \ref{minimal}).
-\textcolor{red}{These results emphasize a relevant feature our approach: the small size of the training set required by \FAMA in comparison with other approaches (see Table \ref{table:models_comparison2}).
+These results emphasize a relevant feature our approach: the small size of the training set required by \FAMA in comparison with other approaches (see Table \ref{table:models_comparison2}).
 Unlike extensive-data approaches, our work explores an alternative research direction to learn action models from small amounts of plan traces.
-This is an important advantage, particularly in domains in which it is costly or impossible to obtain a significant number of training samples.}
-
-%For the PSPACE-complete scenario, {\em precision} and {\em recall} scores have to be taken with a grain of salt because {\bf pure syntax-based metrics are not informative when reformulations appear} (we provide experimental evidence of this in Section \ref{minimal}).
+This is an important advantage, particularly in domains in which it is costly or impossible to obtain a significant number of training samples.
 
@@
 \end{figure}
-
-%The conclusion we can draw is that learning from few input samples is both our strong point and our limitation.
-Results on the NP-complete scenario show that the learned models are considerably sound and complete (to put these results into a richer perspective we will also compare ourselves with another approach in the following section).
-The PSPACE-complete scenario, on the other hand, shows that the learned models have less quality.
-Nevertheless, we argue that syntax-based metrics are not appropriate for the PSPACE-complete scenario due to the vast amount of reformulations that can appear in such under-constrained learning task and we will show proof of this in section \ref{minimal}.
 
@@
 
-%Remarkably, the overall \emph{precision} is now $0.98$, which means that the contents of the learned models is highly reliable.
-The value of \emph{recall}, 0.87, is an indication that the learned models still miss some information (preconditions are again the component more difficult to be fully learned).
-Overall, the results confirm the previous trend: the more input knowledge of the task, the better the models and the less planning time.
-Additionally, the solution plans required for this task are smaller because it is only necessary to program half of the actions (the other half are included in the input knowledge $\mathcal{M}$).
-{\em Visitall} and {\em Hanoi} are excluded from this evaluation because they only contain one action schema.
-
 \subsection{Learning with minimal input knowledge}
 \label{minimal}
-
-%In the previous experiments we have given a general view of the performance of \FAMA under different conditions.
-So far, experiments have shown that \FAMA is able to learn with very small amounts of input knowledge, be it due to low observability or few training samples.
 
@@
 				\hline
-				Blocks & 0.86 & 0.67 & 1.0 & 0.67 & 0.8 & 0.44 & 0.89 & 0.59& 0.24 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
-				Driverlog & 0.6 & 0.86 & 0.36 & 0.57 & 0.67 & 0.29 & 0.53 & 0.64& 0.58 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'unload-truck', 1, 2, 3), (u'board-truck', u'board-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 1, 2, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
-				Ferry & 0.7 & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0& 0.37 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
-				Floortile & 0.69 & 1.0 & 0.55 & 1.0 & 1.0 & 0.82 & 0.69 & 0.95& 1.38 \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'up', 1, 2, 3), (u'down', u'down', 1, 2, 3), (u'right', u'right', 1, 2, 3), (u'left', u'left', 1, 2, 3), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
-				Grid & 0.68 & 0.88 & 0.5 & 0.86 & 0.88 & 1.0 & 0.67 & 0.9& 0.65 \\ % [(u'move', u'move', 1, 2), (u'pickup', u'pickup', 1, 2), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
-				Gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 0.18 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
-				Hanoi & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.8 & 1.0& 0.36 \\ % [(u'move', u'move', 1, 2, 3)]
-				Miconic & 1.0 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 0.84 & 1.0& 0.3 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 1, 2), (u'down', u'down', 1, 2)]
-				Npuzzle & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.88 & 1.0& 0.26 \\ % [(u'move', u'move', 1, 2, 3)]
-				Parking & 0.78 & 1.0 & 0.69 & 1.0 & 1.0 & 1.0 & 0.8 & 1.0& 0.24 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-car-to-car', u'move-car-to-car', 1, 2, 3)]
-				Rovers & 0.54 & 1.0 & 0.3 & 0.76 & 1.0 & 0.46 & 0.48 & 0.85 & 2.14 \\ % [(u'drop', u'drop', 1, 2), (u'navigate', u'navigate', 1, 2, 3), (u'sample-soil', u'sample-soil', 1, 2, 3), (u'sample-rock', u'sample-rock', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3, 4), (u'take-image', u'take-image', 1, 2, 3, 4, 5), (u'communicate-soil-data', u'communicate-soil-data', 1, 2, 3, 4, 5), (u'communicate-rock-data', u'communicate-soil-data', 1, 2, 4, 3, 5), (u'communicate-image-data', u'communicate-image-data', 1, 2, 3, 4, 5, 6)]
-				Satellite & 0.93 & 1.0 & 0.56 & 1.0 & 1.0 & 0.75 & 0.81 & 0.96& 0.4 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
-				Transport & 0.83 & 1.0 & 0.5 & 1.0 & 0.6 & 0.6 & 0.67 & 0.9& 0.19 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
-				Visitall & 1.0 & 1.0 & 0.25 & 0.5 & 1.0 & 1.0 & 0.57 & 0.8& 1.31 \\ % [(u'move', u'move', 1, 2)]
-				Zenotravel & 0.9 & 0.64 & 0.5 & 0.71 & 0.83 & 0.71 & 0.73 & 0.68& 0.25 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 4, 5), (u'zoom', u'zoom', 1, 2, 3, 4, 5, 6)]
+				Blocks & 0.86 & 0.67 & 1.0 & 0.67 & 0.8 & 0.44 & 0.89 & 0.59& 0.24 \\ 
+				Driverlog & 0.6 & 0.86 & 0.36 & 0.57 & 0.67 & 0.29 & 0.53 & 0.64& 0.58 \\
+				Ferry & 0.7 & 1.0 & 0.36 & 1.0 & 1.0 & 1.0 & 0.6 & 1.0& 0.37 \\ 
+				Floortile & 0.69 & 1.0 & 0.55 & 1.0 & 1.0 & 0.82 & 0.69 & 0.95& 1.38 \\
+				Grid & 0.68 & 0.88 & 0.5 & 0.86 & 0.88 & 1.0 & 0.67 & 0.9& 0.65 \\ 
+				Gripper & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 0.18 \\ 
+				Hanoi & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.8 & 1.0& 0.36 \\
+				Miconic & 1.0 & 1.0 & 0.57 & 1.0 & 1.0 & 1.0 & 0.84 & 1.0& 0.3 \\ 
+				Npuzzle & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.88 & 1.0& 0.26 \\ 
+				Parking & 0.78 & 1.0 & 0.69 & 1.0 & 1.0 & 1.0 & 0.8 & 1.0& 0.24 \\
+				Rovers & 0.54 & 1.0 & 0.3 & 0.76 & 1.0 & 0.46 & 0.48 & 0.85 & 2.14 \\
+				Satellite & 0.93 & 1.0 & 0.56 & 1.0 & 1.0 & 0.75 & 0.81 & 0.96& 0.4 \\ 
+				Transport & 0.83 & 1.0 & 0.5 & 1.0 & 0.6 & 0.6 & 0.67 & 0.9& 0.19 \\
+				Visitall & 1.0 & 1.0 & 0.25 & 0.5 & 1.0 & 1.0 & 0.57 & 0.8& 1.31 \\
+				Zenotravel & 0.9 & 0.64 & 0.5 & 0.71 & 0.83 & 0.71 & 0.73 & 0.68& 0.25 \\ 
 				\hline
@@
 			\hline
-			Blocks & 0.89 & 0.89 & 0.8 & 0.89 & 0.83 & 0.56 & 0.84 & 0.78& 0.77 \\ % [(u'pick-up', u'pick-up', 1), (u'put-down', u'put-down', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
-			Driverlog & 0.57 & 0.29 & 0.31 & 0.57 & 0.4 & 0.29 & 0.4 & 0.36& 7.35 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'unload-truck', 1, 2, 3), (u'board-truck', u'walk', 1, 2, 3), (u'disembark-truck', u'board-truck', 1, 2, 3), (u'walk', u'walk', 2, 1, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
-			Ferry & 0.83 & 0.71 & 0.36 & 1.0 & 1.0 & 1.0 & 0.62 & 0.87& 3.38 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
-			Floortile & - & - & - & - & - & - & - & - & - \\ % [(u'change-color', u'change-color', 1, 2, 3), (u'up', u'change-color', 1, 3, 2), (u'down', u'up', 1, 2, 3), (u'right', u'up', 1, 3, 2), (u'left', u'down', 1, 3, 2), (u'paint-up', u'paint-up', 1, 2, 3, 4), (u'paint-down', u'paint-down', 1, 2, 3, 4)]
-			Grid & - & - & - & - & - & - & - & - & - \\ % [(u'move', u'move', 1, 2), (u'pickup', u'move', 2, 1), (u'putdown', u'putdown', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 1, 2, 3, 4)]
-			Gripper & 1.0 & 1.0 & 0.8 & 1.0 & 1.0 & 1.0 & 0.93 & 1.0& 0.17 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
-			Hanoi & 0.67 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 0.75& 132.69 \\ % [(u'move', u'move', 1, 2, 3)]
-			Miconic & 1.0 & 0.33 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.56& 0.71 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'up', 1, 2), (u'down', u'down', 1, 2)]
-			Npuzzle & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86& 13.48 \\ % [(u'move', u'move', 1, 2, 3)]
-			Parking & 0.83 & 0.36 & 1.0 & 0.89 & 0.83 & 0.56 & 0.9 & 0.56& 160.64 \\ % [(u'move-curb-to-curb', u'move-curb-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-car', 1, 2, 3), (u'move-car-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-car-to-car', u'move-car-to-car', 1, 2, 3)]
-			Rovers & 0.43 & 0.73 & 0.24 & 0.47 & 0.56 & 0.38 & 0.39 & 0.61& 31.13 \\ % [(u'drop', u'drop', 1, 2), (u'navigate', u'navigate', 1, 3, 2), (u'sample-soil', u'sample-soil', 1, 2, 3), (u'sample-rock', u'sample-rock', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3, 4), (u'take-image', u'take-image', 1, 2, 3, 4, 5), (u'communicate-soil-data', u'communicate-rock-data', 1, 2, 4, 3, 5), (u'communicate-rock-data', u'communicate-rock-data', 1, 2, 3, 5, 4), (u'communicate-image-data', u'communicate-image-data', 1, 2, 3, 4, 6, 5)]
-			Satellite & 0.6 & 0.21 & 0.56 & 1.0 & 0.5 & 0.5 & 0.56 & 0.43& 11.55 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
-			Transport & 1.0 & 0.2 & 0.57 & 0.8 & 1.0 & 0.4 & 0.73 & 0.4& 23.39 \\ % [(u'drive', u'drive', 1, 2, 3), (u'pick-up', u'drop', 1, 2, 3, 4, 5), (u'drop', u'pick-up', 1, 2, 3, 4, 5)]
-			Visitall & 0.67 & 1.0 & 0.5 & 0.5 & 1.0 & 1.0 & 0.67 & 0.8& 3.13 \\ % [(u'move', u'move', 1, 2)]
-			Zenotravel & 1.0 & 0.36 & 0.4 & 0.29 & 1.0 & 0.43 & 0.77 & 0.36& 226.27 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 5, 4), (u'zoom', u'zoom', 1, 2, 3, 5, 6, 4)]
+			Blocks & 0.89 & 0.89 & 0.8 & 0.89 & 0.83 & 0.56 & 0.84 & 0.78& 0.77 \\ 
+			Driverlog & 0.57 & 0.29 & 0.31 & 0.57 & 0.4 & 0.29 & 0.4 & 0.36& 7.35 \\ 
+			Ferry & 0.83 & 0.71 & 0.36 & 1.0 & 1.0 & 1.0 & 0.62 & 0.87& 3.38 \\ 
+			Floortile & - & - & - & - & - & - & - & - & - \\ 
+			Grid & - & - & - & - & - & - & - & - & - \\ 
+			Gripper & 1.0 & 1.0 & 0.8 & 1.0 & 1.0 & 1.0 & 0.93 & 1.0& 0.17 \\ 
+			Hanoi & 0.67 & 0.5 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86 & 0.75& 132.69 \\ 
+			Miconic & 1.0 & 0.33 & 1.0 & 1.0 & 1.0 & 0.67 & 1.0 & 0.56& 0.71 \\ 
+			Npuzzle & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86& 13.48 \\ 
+			Parking & 0.83 & 0.36 & 1.0 & 0.89 & 0.83 & 0.56 & 0.9 & 0.56& 160.64 \\
+			Rovers & 0.43 & 0.73 & 0.24 & 0.47 & 0.56 & 0.38 & 0.39 & 0.61& 31.13 \\ 
+			Satellite & 0.6 & 0.21 & 0.56 & 1.0 & 0.5 & 0.5 & 0.56 & 0.43& 11.55 \\ 
+			Transport & 1.0 & 0.2 & 0.57 & 0.8 & 1.0 & 0.4 & 0.73 & 0.4& 23.39 \\ 
+			Visitall & 0.67 & 1.0 & 0.5 & 0.5 & 1.0 & 1.0 & 0.67 & 0.8& 3.13 \\ 
+			Zenotravel & 1.0 & 0.36 & 0.4 & 0.29 & 1.0 & 0.43 & 0.77 & 0.36& 226.27 \\
 			\hline
@@
 			\hline
-			Blocks & 0.6 & 0.67 & 0.33 & 0.22 & 0.67 & 0.44 & 0.55 & 0.44& 0.26 \\ % [(u'pick-up', u'put-down', 1), (u'put-down', u'pick-up', 1), (u'stack', u'stack', 1, 2), (u'unstack', u'unstack', 1, 2)]
-			Driverlog & 0.5 & 0.29 & 0.33 & 0.57 & 0.0 & 0.0 & 0.38 & 0.29& 0.88 \\ % [(u'load-truck', u'load-truck', 1, 2, 3), (u'unload-truck', u'board-truck', 1, 2, 3), (u'board-truck', u'unload-truck', 1, 2, 3), (u'disembark-truck', u'disembark-truck', 1, 2, 3), (u'walk', u'walk', 2, 1, 3), (u'drive-truck', u'drive-truck', 1, 2, 3, 4)]
-			Ferry & 0.5 & 0.43 & 0.5 & 0.25 & 0.67 & 0.5 & 0.55 & 0.4& 0.45 \\ % [(u'sail', u'sail', 1, 2), (u'board', u'board', 1, 2), (u'debark', u'debark', 1, 2)]
-			Floortile & 0.48 & 0.45 & 0.27 & 0.36 & 0.46 & 0.55 & 0.41 & 0.45& 58.38 \\ % [(u'change-color', u'change-color', 1, 3, 2), (u'up', u'left', 1, 3, 2), (u'down', u'up', 1, 3, 2), (u'right', u'down', 1, 3, 2), (u'left', u'right', 1, 3, 2), (u'paint-up', u'paint-up', 1, 3, 2, 4), (u'paint-down', u'paint-down', 1, 3, 2, 4)]
-			Grid & 0.25 & 0.24 & 0.33 & 0.43 & 0.14 & 0.14 & 0.25 & 0.26& 234.63 \\ % [(u'move', u'move', 2, 1), (u'pickup', u'putdown', 1, 2), (u'putdown', u'pickup', 1, 2), (u'pickup-and-loose', u'pickup-and-loose', 1, 2, 3), (u'unlock', u'unlock', 2, 1, 3, 4)]
-			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86& 0.16 \\ % [(u'move', u'move', 1, 2), (u'pick', u'pick', 1, 2, 3), (u'drop', u'drop', 1, 2, 3)]
-			Hanoi & 0.6 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.78 & 0.88& 6.32 \\ % [(u'move', u'move', 3, 2, 1)]
-			Miconic & 0.63 & 0.56 & 0.6 & 0.75 & 0.25 & 0.33 & 0.53 & 0.56& 0.25 \\ % [(u'board', u'board', 1, 2), (u'depart', u'depart', 1, 2), (u'up', u'down', 1, 2), (u'down', u'up', 2, 1)]
-			Npuzzle & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 1.52 \\ % [(u'move', u'move', 1, 2, 3)]
-			Parking & 0.57 & 0.29 & 0.2 & 0.11 & 0.8 & 0.44 & 0.53 & 0.28& 17.43 \\ % [(u'move-curb-to-curb', u'move-car-to-curb', 1, 2, 3), (u'move-curb-to-car', u'move-curb-to-curb', 1, 3, 2), (u'move-car-to-curb', u'move-car-to-car', 3, 2, 1), (u'move-car-to-car', u'move-car-to-car', 1, 3, 2)]
-			Rovers & 0.38 & 0.64 & 0.07 & 0.24 & 0.13 & 0.54 & 0.22 & 0.53& 1.74 \\ % [(u'drop', u'drop', 1, 2), (u'navigate', u'navigate', 1, 2, 3), (u'sample-soil', u'sample-rock', 1, 2, 3), (u'sample-rock', u'sample-soil', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3, 4), (u'take-image', u'take-image', 1, 2, 3, 4, 5), (u'communicate-soil-data', u'communicate-soil-data', 1, 2, 5, 3, 4), (u'communicate-rock-data', u'communicate-soil-data', 1, 2, 3, 5, 4), (u'communicate-image-data', u'communicate-image-data', 1, 2, 3, 4, 5, 6)]
-			Satellite & 0.86 & 0.43 & 0.43 & 0.6 & 0.75 & 0.75 & 0.67 & 0.52& 3.15 \\ % [(u'switch-on', u'switch-on', 1, 2), (u'switch-off', u'switch-off', 1, 2), (u'turn-to', u'turn-to', 1, 2, 3), (u'calibrate', u'calibrate', 1, 2, 3), (u'take-image', u'take-image', 1, 2, 3, 4)]
-			Transport & 0.29 & 0.2 & 0.38 & 0.6 & 0.67 & 0.4 & 0.39 & 0.35& 1.06 \\ % [(u'drive', u'drive', 1, 3, 2), (u'pick-up', u'pick-up', 1, 2, 3, 4, 5), (u'drop', u'drop', 1, 2, 3, 4, 5)]
-			Visitall & 0.0 & 0.0 & 1.0 & 0.5 & 0.0 & 0.0 & 1.0 & 0.2& 1.21 \\ % [(u'move', u'move', 2, 1)]
-			Zenotravel & 0.4 & 0.29 & 0.25 & 0.29 & 0.2 & 0.14 & 0.3 & 0.25& 17.48 \\ % [(u'board', u'board', 1, 2, 3), (u'debark', u'debark', 1, 2, 3), (u'refuel', u'refuel', 1, 2, 3, 4), (u'fly', u'fly', 1, 2, 3, 5, 4), (u'zoom', u'zoom', 1, 2, 3, 4, 5, 6)]
-
+			Blocks & 0.6 & 0.67 & 0.33 & 0.22 & 0.67 & 0.44 & 0.55 & 0.44& 0.26 \\ 
+			Driverlog & 0.5 & 0.29 & 0.33 & 0.57 & 0.0 & 0.0 & 0.38 & 0.29& 0.88 \\
+			Ferry & 0.5 & 0.43 & 0.5 & 0.25 & 0.67 & 0.5 & 0.55 & 0.4& 0.45 \\ 
+			Floortile & 0.48 & 0.45 & 0.27 & 0.36 & 0.46 & 0.55 & 0.41 & 0.45& 58.38 \\
+			Grid & 0.25 & 0.24 & 0.33 & 0.43 & 0.14 & 0.14 & 0.25 & 0.26& 234.63 \\ 
+			Gripper & 1.0 & 0.67 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 0.86& 0.16 \\ 
+			Hanoi & 0.6 & 0.75 & 1.0 & 1.0 & 1.0 & 1.0 & 0.78 & 0.88& 6.32 \\ 
+			Miconic & 0.63 & 0.56 & 0.6 & 0.75 & 0.25 & 0.33 & 0.53 & 0.56& 0.25 \\ 
+			Npuzzle & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0 & 1.0& 1.52 \\ 
+			Parking & 0.57 & 0.29 & 0.2 & 0.11 & 0.8 & 0.44 & 0.53 & 0.28& 17.43 \\
+			Rovers & 0.38 & 0.64 & 0.07 & 0.24 & 0.13 & 0.54 & 0.22 & 0.53& 1.74 \\
+			Satellite & 0.86 & 0.43 & 0.43 & 0.6 & 0.75 & 0.75 & 0.67 & 0.52& 3.15 \\
+			Transport & 0.29 & 0.2 & 0.38 & 0.6 & 0.67 & 0.4 & 0.39 & 0.35& 1.06 \\
+			Visitall & 0.0 & 0.0 & 1.0 & 0.5 & 0.0 & 0.0 & 1.0 & 0.2& 1.21 \\ 
+			Zenotravel & 0.4 & 0.29 & 0.25 & 0.29 & 0.2 & 0.14 & 0.3 & 0.25& 17.48 \\ 			
 			\hline

### 2019-05-01 doc/aij18/minor-revision/introduction.tex
AIJ18 minor revision folder created

--- 2018-12-26 doc/aij18/major-revision/introduction.tex
@@
 
-A system for learning planning action models receives as an input observations of the agent's plan execution and \textcolor[rgb]{1.00,0.00,0.00}{outputs an approximation of the actions that embody the physics of the real-world domain being modeled.} The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards, \textcolor[rgb]{1.00,0.00,0.00}{but there exists as well a large variety of planning-related tasks that rely upon the existence of a planning model.
+A system for learning planning action models receives as an input observations of the agent's plan execution and outputs an approximation of the actions that embody the physics of the real-world domain being modeled.
+The primary underlying motivation for acquiring planning action models is to solve model-based planning tasks afterwards, but there exists as well a large variety of planning-related tasks that rely upon the existence of a planning model.
 Among these tasks, we might cite: \emph{goal and plan recognition} approaches based on a domain theory~\cite{ramirez2009plan,ramirez2012plan,SohrabiRU16}; \emph{transparent planning}, in which an agent implicitly communicates its true goal by making its intentions and its action selection transparent (recognizable) to observers~\cite{MacNallyLRP18}; or \emph{deceptive path-planning}, which draws on the definition of path-planning domains and aims at finding a path such that the probability of an observer identifying the final destination is minimised~\cite{MastersS17}.
 Planning models are also used in \emph{explainable AI planning} to form a common basis for communicating with users and facilitate the generation of transparent and explainable decisions~\cite{FoxLM17} as well as explanations in terms of the differences with a human mental model~\cite{ChakrabortiSK18}.
-\emph{Counterplanning} requires a model of the opponent agent in order to recognize its goals \cite{PozancoEFB18} and \emph{model reconciliation} aims to conform the models of two agents with respect to an observation of a plan computed with one of the two models~\cite{ChakrabortiSZK17}.}
+\emph{Counterplanning} requires a model of the opponent agent in order to recognize its goals \cite{PozancoEFB18} and \emph{model reconciliation} aims to conform the models of two agents with respect to an observation of a plan computed with one of the two models~\cite{ChakrabortiSZK17}.
 
-\textcolor[rgb]{1.00,0.00,0.00}{Motivated by the requirement for a planning model in many different tasks and the recent advances on the use of classical planning for the generation of different types of planning models (regular automata, context-free grammars, finite-state machines, \strips)} \cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behavior is not available.
+Motivated by the requirement for a planning model in many different tasks and the recent advances on the use of classical planning for the generation of different types of planning models (regular automata, context-free grammars, finite-state machines, \strips) \cite{bonet2009automatic,segovia2016generalized,segovia2016hierarchical,segovia2017generating}, in this paper we claim that a planning model is learnable even though an accurate representation of the agent's behavior is not available.
 Particularly, we present a novel learning algorithm, called \FAMA, capable of inferring the preconditions and effects of \strips\ action models, the vanilla action model for automated planning~\cite{fikes1971strips}, under minimal observability.
 
-%Up to date, all existing learning approaches assume that a given input plan trace contains a fully observed sequence of the executed actions.
-This heavily restricts the applicability of the learning approach to contexts in which the behaviour of the agent is fully observable and a human annotator correctly labels the executed actions.
-
-\textcolor[rgb]{1.00,0.00,0.00}{Current learning approaches assume that the observation of the agent's plan execution (plan trace) encompasses the fully observed sequence of the executed actions; i.e, they assume all the actions performed by the agent are observable.
+Current learning approaches assume that the observation of the agent's plan execution (plan trace) encompasses the fully observed sequence of the executed actions; i.e, they assume all the actions performed by the agent are observable.
 This heavily restricts the applicability of the learning approaches to contexts where the behaviour of the agent is fully observable, which also commonly entails a human annotator that correctly labels the executed actions.
-On the other hand, learning approaches accept a varying degree of observability in the states traversed in the plan trace, ranging from fully observable to fully unobservable states (see section \ref{related_work} for details).} In contrast, \FAMA allows for an incomplete or empty sequence of observable actions, and the minimum observability case acceptable by \FAMA is when the algorithm is only fed with the initial and final state of a plan trace.
+On the other hand, learning approaches accept a varying degree of observability in the states traversed in the plan trace, ranging from fully observable to fully unobservable states (see section \ref{related_work} for details).
+In contrast, \FAMA allows for an incomplete or empty sequence of observable actions, and the minimum observability case acceptable by \FAMA is when the algorithm is only fed with the initial and final state of a plan trace.
 Like many Machine Learning (ML) techniques, \FAMA is able to operate with only input/output pairs of states and an unknown or a partially known model of the agent.
@@
 
-\textcolor{red}{\FAMA is a new learning approach, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}.
+FAMA is a new learning approach, based on AI planning technology, that automatically compiles the task of learning \strips\ actions into a planning task which is then solved with a planner~\cite{aineto2018learning}.
 The construction of a \strips\ planning model starts out from a set of plan traces containing the observation of several plan executions.
@@
 Hence, a solution to this planning task is a plan that determines the preconditions and effects of the actions of a \strips\ planning model $M$ while ensuring consistency with the input traces.
-We say that a model $M$ is \emph{consistent} with a plan trace when $M$ can produce a solution $\pi$ to the planning problem $\tup{s_0,s_f}$ so that: (1) $\pi$ contains the observed actions of the plan trace, if any, and (2) the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.}
+We say that a model $M$ is \emph{consistent} with a plan trace when $M$ can produce a solution $\pi$ to the planning problem $\tup{s_0,s_f}$ so that: (1) $\pi$ contains the observed actions of the plan trace, if any, and (2) the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.
 
@@
 
-%Like other learning approaches, \FAMA also accepts as an input an initial action model of the agent's behaviour, either complete or partially specified~\cite{ZhuoNK13,ZhuoK17}.
-In this case, the output model of \FAMA is compliant with both the input model and with the observations.
-Following this behaviour, we can define a distance measure between the two models which yields an assessment of the accuracy with which the input model explains the observations.
-In other words, the distance measure provides a mechanism to explicitly correct or adjust a model to the observations.
-By using this distance-based concept, \FAMA contributes with two novel semantic evaluation metrics that are generally more informative than counting the number of errors between two models.
-Moreover, this new evaluation method opens up a way towards model reconciliation~\cite{KulkarniCZVZK16}.
-
-%All in all, \FAMA is a planning-based solving scheme that outputs a \strips\ action model using a planning model that is automatically built from minimal input knowledge.
-Unlike extensive-data ML approaches, \FAMA only requires a small amount of input plan traces.
-Unlike most relevant action-model learning algorithms, \FAMA does not require the traces to contain any observed action executed by the agent.
-
-\textcolor[rgb]{1.00,0.00,0.00}{In summary, \FAMA is a novel learning approach characterized by:
+In summary, \FAMA is a novel learning approach characterized by:
 \begin{itemize}
@@
 \item a semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model.
-\end{itemize} }
+\end{itemize}
 
@@
 Section~\ref{sec:background} introduces classical planning concepts and reviews related work on learning planning action models.
-\textcolor{red}{Section~\ref{task_definition} formally defines the learning task and motivates our compilation-to-planning approach for learning action models.} Section~\ref{sec:learning} presents the compilation scheme, the core of \FAMA.
+Section~\ref{task_definition} formally defines the learning task and motivates our compilation-to-planning approach for learning action models.
+Section~\ref{sec:learning} presents the compilation scheme, the core of \FAMA.
 Sections~\ref{sec:evaluation} explains the evaluation of a learned model with respect to a reference model (syntactic evaluation) and with respect to a set of plan traces (semantic evaluation).

### 2019-05-01 doc/aij18/minor-revision/learning_action_models.tex
AIJ18 minor revision folder created

--- 2018-12-26 doc/aij18/major-revision/learning_action_models.tex
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{Our proposal to address a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to transform $\Lambda$ into a planning task $P_{\Lambda}$.
+Our proposal to address a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is to transform $\Lambda$ into a planning task $P_{\Lambda}$.
 The intuition behind the compilation is that when $P_{\Lambda}$ is solved with a planner, the solution plan $\pi_\Lambda$ is a sequence of actions that build the action models of the output domain model $\mathcal{M'}$ and verify that $\mathcal{M'}$ is consistent with the actions and states of the observed plan trace $\tau = \tup {s_0, \ldots, s_n}$.
 Hence, $\pi_\Lambda$ will comprise two differentiated blocks of actions: a first set of actions each defining the \textbf{insertion} of a fluent as a precondition, a positive effect or a negative effect of an action model $\xi \in \mathcal{M'}$; and a second set of actions that determine the \textbf{application} of the learned $\xi$s while successively \textbf{validating} the effects of the action application in every observable point of $\tau$, including that the final reached state comprises $s_n$.
-Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states of $\tau$.}
+Roughly speaking, in the \emph{blocksworld} domain, the format of the first set of actions of $\pi_\Lambda$ will look like {\tt{\small (insert\_pre\_stack\_holding\_v1),(insert\_eff\_stack\_clear\_v1),(insert\_eff\_stack\_clear\_v2)}}, where the first effect denotes a positive effect and the second one a negative effect to be inserted in $name(\xi)=${\tt{\small stack}}; and the format of the second set of actions of $\pi_\Lambda$ will be like {\tt{\small (apply\_unstack blockB blockA),(apply\_putdown blockB)}} and {\tt{\small (validate\_1),(validate\_2)}}, where the last two actions denote the points at which the states generated through the action application must be validated with the observed states of $\tau$.
 
@@
 
-We are now ready to define the propositional encoding of the \textcolor[rgb]{1.00,0.00,0.00}{model fluents} of $pre(\xi)$, $del(\xi)$ and $add(\xi)$. For every $\xi$ and $p\in \Psi_{\xi}$, we create:
+We are now ready to define the propositional encoding of the model fluents of $pre(\xi)$, $del(\xi)$ and $add(\xi)$. For every $\xi$ and $p\in \Psi_{\xi}$, we create:
 
 \begin{itemize}
-\item $pre_p(\xi)$: \textcolor[rgb]{1.00,0.00,0.00}{model fluent} formed by the combination of the prefixes {\small \texttt{pre}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
+\item $pre_p(\xi)$: model fluent formed by the combination of the prefixes {\small \texttt{pre}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
 {\small \texttt{pre\_stack\_on\_v1\_v2}}, for $name(\xi)={\small\texttt{stack}}$ and $p={\small\texttt{on($v_1,v_2$)}}$)
-\item $del_p(\xi)$: \textcolor[rgb]{1.00,0.00,0.00}{model fluent} formed by the combination of the prefixes {\small \texttt{del}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
+\item $del_p(\xi)$: model fluent formed by the combination of the prefixes {\small \texttt{del}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
 {\small  \texttt{del\_stack\_on\_v1\_v2}})
-\item $add_p(\xi)$: \textcolor[rgb]{1.00,0.00,0.00}{model fluent} formed by the combination of the prefixes {\small \texttt{add}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
+\item $add_p(\xi)$: model fluent formed by the combination of the prefixes {\small \texttt{add}} and $name(\xi)$ plus a fluent of arity 0 that results from appending the elements of $p$ (e.g.
 {\small \texttt{add\_stack\_on\_v1\_v2}})
@@
 For a given action model $\xi$, if a fluent $pre_p(\xi)/del_p(\xi)/add_p(\xi)$ holds in a state, it means that $p$ is a precondition/negative/positive effect of $\xi$.
-For instance, Figure~\ref{fig:encodedstack} shows the conjunction of \textcolor[rgb]{1.00,0.00,0.00}{model fluents} that represents the propositional encoding of the preconditions, negative effects and positive effects of the action model corresponding to the {\em stack} operator shown in Figure~\ref{fig:stack}.
+For instance, Figure~\ref{fig:encodedstack} shows the conjunction of model fluents that represents the propositional encoding of the preconditions, negative effects and positive effects of the action model corresponding to the {\em stack} operator shown in Figure~\ref{fig:stack}.
 
@@
 
-%\subsection{Learning from observations of plan executions}
 Given a learning task $\Lambda=\tup{\mathcal{M},\tau}$, with $\tau$ formed by an $n$-action sequence $\tup{a_1, \ldots, a_n}$ and a $m$-state trajectory $\tup{s_0, s_1, \ldots, s_m}$ ($\tau = \langle s_0, a_1, \ldots, a_n, s_m \rangle$), the compilation outputs a classical planning task $P_{\Lambda}=\tup{F_{\Lambda},A_{\Lambda},I_{\Lambda},G_{\Lambda}}$ such that:
@@
 
-\item $F_{\Lambda}$ \textcolor[rgb]{1.00,0.00,0.00}{extends $F$ with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of $\tau$}.
+\item $F_{\Lambda}$ extends $F$ with the model fluents to represent the preconditions and effects of each $\xi\in\mathcal{M}$ as well as some other fluents to keep track of the validation of $\tau$.
 Specifically, $F_{\Lambda}$ contains:
@@
 \item The set of fluents obtained from $s_0$; i.e., $F$.
-\item The \textcolor[rgb]{1.00,0.00,0.00}{model fluents} $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section \ref{propositional_encoding}
+\item The model fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$, for every $\xi \in \mathcal{M}$ and $p\in \Psi_{\xi}$, defined as explained in section \ref{propositional_encoding}
 \item A set of fluents $F_{\pi}=\{plan(name(a_i),\Omega^{ar(a_i)},i)\}_{\small 1\leq i\leq n}$ to represent the $i^{th}$ observable action of $\tau$.
@@
 
-\item $A_{\Lambda}$ includes \textcolor[rgb]{1.00,0.00,0.00}{three types of actions that give rise to the actions of $\pi_\Lambda$}.
+\item $A_{\Lambda}$ includes three types of actions that give rise to the actions of $\pi_\Lambda$.
 \begin{enumerate}
-\item Actions for \textcolor[rgb]{1.00,0.00,0.00}{{\em inserting} a component (precondition, positive effect or negative effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models}.
+\item Actions for {\em inserting} a component (precondition, positive effect or negative effect) in $\xi \in \mathcal{M}$ following the syntactic constraints of \strips\ models.
 These actions will form the prefix of the solution plan $\pi_\Lambda$.
@@
 \item Actions which support the addition of a {\em precondition} $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
-\textcolor[rgb]{1.00,0.00,0.00}{A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $del_p$ nor $add_p$ exist in $\xi$.}
+A precondition $p$ is inserted in $\xi$ when neither $pre_p$, $del_p$ nor $add_p$ exist in $\xi$.
 
@@
 \item Actions which support the addition of a {\em negative} or {\em positive} effect $p\in \Psi_{\xi}$ to the action model $\xi\in\mathcal{M}$.
-\textcolor[rgb]{1.00,0.00,0.00}{A positive effect is inserted in $\xi$ under the same conditions of a precondition insertion, and a negative effect is inserted in $\xi$ when neither $del_p$ nor $add_p$ appear in $\xi$ but $pre_p$ does.}
+A positive effect is inserted in $\xi$ under the same conditions of a precondition insertion, and a negative effect is inserted in $\xi$ when neither $del_p$ nor $add_p$ appear in $\xi$ but $pre_p$ does.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{For instance, given $name(\xi)=${\tt{\small stack}} and $C_{pre-stack}=\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_holding\_v2), (pre\_stack\_on\_v1\_v2),(pre\_stack\_clear\_v1),(pre\_stack\_clear\_v1),}}$\ldots \}$, the insertion of each item $c \in C_{pre-stack}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$ as long as $c \notin pre(\xi)$,  $c \notin add(\xi)$ and $c \notin del(\xi)$.
-The same applies to effects with respect to sets $C_{add-stack}$ and $C_{del-stack}$ that would include all fluents starting with prefix  {\tt{\small add}} and {\tt{\small del}}, respectively.}
+For instance, given $name(\xi)=${\tt{\small stack}} and $C_{pre-stack}=\{${\tt{\small (pre\_stack\_holding\_v1),(pre\_stack\_holding\_v2), (pre\_stack\_on\_v1\_v2),(pre\_stack\_clear\_v1),(pre\_stack\_clear\_v1),}}$\ldots \}$, the insertion of each item $c \in C_{pre-stack}$ in $\xi$ will generate a different alternative in the search space when solving $P_{\Lambda}$ as long as $c \notin pre(\xi)$,  $c \notin add(\xi)$ and $c \notin del(\xi)$.
+The same applies to effects with respect to sets $C_{add-stack}$ and $C_{del-stack}$ that would include all fluents starting with prefix  {\tt{\small add}} and {\tt{\small del}}, respectively.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{Note that executing an insert action, e.g.{\tt{\small (insert\_pre\_stack\_holding\_v1)}}, will add the corresponding model fluent {\tt{\small (pre\_stack\_holding\_v1)}} to the successor state.
+Note that executing an insert action, e.g.{\tt{\small (insert\_pre\_stack\_holding\_v1)}}, will add the corresponding model fluent {\tt{\small (pre\_stack\_holding\_v1)}} to the successor state.
 Hence, the execution of the insert actions of $\pi_\Lambda$ yield a state containing the valuation of the model fluents that shape every $\xi \in \mathcal{M}$.
 For example, executing the insert actions that shape the action model $name(\xi)=${\tt{\small putdown}} leads to a state containing the positive literals {\tt{\small (pre\_putdown\_holding\_v1),(eff\_putdown\_holding\_v1),\\ (eff\_putdown\_clear\_v1),
-(eff\_putdown\_ontable\_v1),(eff\_putdown\_handempty)}}.}
+(eff\_putdown\_ontable\_v1),(eff\_putdown\_handempty)}}.
 
@@
 
-    These actions will be part of the postfix of the plan $\pi_\Lambda$ and \textcolor[rgb]{1.00,0.00,0.00}{they determine the application of the learned action models according to the values of the model fluents in the current state configuration}.
+    These actions will be part of the postfix of the plan $\pi_\Lambda$ and they determine the application of the learned action models according to the values of the model fluents in the current state configuration.
 Figure~\ref{fig:compilation} shows the PDDL encoding of {\tt{\small (apply\_stack)}} for applying the action model of the {\em stack} operator.
-\textcolor[rgb]{1.00,0.00,0.00}{Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$.
+Let's assume the action {\tt{\small (apply\_stack blockB blockA)}} is in $\pi_\Lambda$.
 Executing this action in a state $s$ implies activating the preconditions and effects of {\tt{\small (apply\_stack)}} according to the values of the model fluents in $s$.
@@
 Otherwise, a different set of precondition literals will be checked.
-The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.}
+The same applies to the conditional effects, generating the corresponding literals according to the values of the model fluents of $s$.
 
-\textcolor[rgb]{1.00,0.00,0.00}{    Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if $name(\xi)=${\tt{\small stack}} has been correctly programmed by the insert actions.
-Hence, while \textbf{insert actions} add the values of the \textbf{model fluents} that shape $\xi$, the \textbf{apply actions} add the values of the \textbf{fluents of $F$} that result from the execution of $\xi$.}
+Note that executing an apply action, e.g.{\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state if $name(\xi)=${\tt{\small stack}} has been correctly programmed by the insert actions.
+Hence, while \textbf{insert actions} add the values of the \textbf{model fluents} that shape $\xi$, the \textbf{apply actions} add the values of the \textbf{fluents of $F$} that result from the execution of $\xi$.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$.
-The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.}
+There will be a validate action in $\pi_\Lambda$ for every observed state in $\tau$.
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.
 
@@
 
-%As shown in Figure \ref{fig:plan-lplan}, the learned action models $\mathcal{M'}$ can be easily inferred from the solution plan $\pi_\Lambda$, just by looking at the programming actions.
-However, for the shake of formalizing the extraction of the learned models, we will infer them from the final state of the solution plan $s_g=\theta(I_\Lambda,\pi_\Lambda)$ by looking at the values for the fluents $pre_p(\xi)$, $del_p(\xi)$ and $add_p(\xi)$ which encode the preconditions and effects of the validated action models.
-For each $\xi \in \mathcal{M'}$, we build the sets of preconditions, positive effects and negative effects as follows:
-
 \begin{small}
@@
 This is specially relevant for static predicates that never change and, hence, only appear as preconditions in the actions.
-In order to address this shortcoming \textcolor[rgb]{1.00,0.00,0.00}{and complete the list of learned preconditions}, we apply a post-process based on the one proposed in~\cite{kuvcera2018louga}.
-\textcolor[rgb]{1.00,0.00,0.00}{The idea lies in going through every action and counting the number of cases where a literal is present before the action is executed and the number of cases where it is not present.
-If a literal is present in all the cases before the action, the literal is considered to be a precondition.}
+In order to address this shortcoming and complete the list of learned preconditions, we apply a post-process based on the one proposed in~\cite{kuvcera2018louga}.
+The idea lies in going through every action and counting the number of cases where a literal is present before the action is executed and the number of cases where it is not present.
+If a literal is present in all the cases before the action, the literal is considered to be a precondition.
 
-\textcolor[rgb]{1.00,0.00,0.00}{In order to obtain a complete trace, the proposal in ~\cite{kuvcera2018louga} applies the sequence of actions of the input trace and infers the preconditions from this \FO action sequence}.
+In order to obtain a complete trace, the proposal in ~\cite{kuvcera2018louga} applies the sequence of actions of the input trace and infers the preconditions from this \FO action sequence.
 In our case, since the sequence of actions of the input trace might not be fully observable, we produce the traces by applying the actions found in the validation part of the solution plan.
@@
 
-%This post-process allows \FAMA to learn more complete lists of preconditions and deal with the always problematic static predicates.
-
 \subsection{Properties of the compilation}
 \label{properties}
-
-%Given a plan $\pi_\Lambda$ that solves $P_\Lambda$, and the learned action models $\mathcal{M'}$ that solve $\Lambda=\tup{\mathcal{M}, \tau}$, \FAMA contributes with the following compilation properties:
 
@@
 This means that for every $\mathcal{M}'$ that solves $\Lambda$, there exists a plan $\pi_\Lambda$ that can be built selecting the appropriate programming, apply and validate actions from the $P_{\Lambda}$ compilation.
-  %a solution plan $\pi_\Lambda$ can be built selecting the corresponding  $\mathsf{programPre_{p,\xi}}$ and $\mathsf{programEff_{p,\xi}}$ actions according to $\mathcal{M}'$ and later, selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ and $\mathsf{validate_{i}}$ actions according to $\tau$.
 \end{small}
@@
 VAL, however, requires a full plan and a full action model for plan validation.
-
-%\subsection{Optimizing the compilation with background knowledge}
-%A distinctive feature of Inductive Logic Programming (ILP) is that ILP can leverage {\em background knowledge} to learn logic programs from data~\cite{muggleton1994inductive}.
-Inspired by ILP, we show that our approach for the learning of \strips\ action models can also leverage {\em background knowledge} in this case to optimize the performance of the $P_{\Lambda}$ compilation.
-
-%\subsubsection{Static predicates}
-%A {\em static predicate} $p \in \Psi_\Lambda$ is a predicate that does not appear in the effects of any action~\cite{fox:TIM:JAIR1998}.
-Therefore, one can get rid of the mechanism for programming these predicates in the effects of any action schema while keeping the compilation complete.
-Given a static predicate $p$:
-%\begin{itemize}
-%\item Fluents $del_p(\xi)$ and $add_p(\xi)$, such that $p\in \Psi_v$ is the static predicate, can be discarded for every $\xi\in\mathcal{M}$.
-%\item Actions $\mathsf{programEff_{p,\xi}}$ (s.t. $p\in \Psi_v$ is a parameterized static predicate) can also be discarded for every $\xi\in\mathcal(M)$.
-%\end{itemize}
-
-%Static predicates can also constrain the space of possible preconditions by looking at the given set of state observations in $\tau$.
-One can assume that if a precondition $p\in \Psi_v$ (s.t.
-$p\in \Psi_v$ is a static predicate parameterized with $\Omega_v$) is not compliant with the observations in $\tau$ then, fluents $pre_p(\xi)$ and actions $\mathsf{programPre_{p,\xi}}$ can be discarded for every $\xi\in\mathcal{M}$.
-For instance, in the {\em zenotravel}~\cite{long20033rd} domain $pre\_next\_board\_v1\_v1$, $pre\_next\_debark\_v1\_v1$, $pre\_next\_fly\_v1\_v1$, $pre\_next\_zoom\_v1\_v1$, $pre\_next\_refuel\_v1\_v1$ can be discarded (and their corresponding programming actions) because a precondition {\tt\small(next ?v1 ?v1 - flevel)} will never hold at any state in $\tau$.
-
-%Furthermore looking as well at the given plan traces, fluents $pre_p(\xi)$ and actions $\mathsf{programPre_{p,\xi}}$ are also discardable for every $\xi\in\mathcal(M)$ if a precondition $p\in \Psi_v$ (s.t.
-$p\in \Psi_v$ is a static predicate parameterized with $\Omega_v$) is not possible according to $\tau$.
-Back to the {\em zenotravel} domain, if a plan trace $\tau$ contains the action {\tt\small (fly plane1 city2 city0 fl3 fl2)} and the corresponding state observations contain the static literal {\tt\small (next fl2 fl3)} but does not contain {\tt\small (next fl2 fl2)}, {\tt\small (next fl3 fl3)} or {\tt\small (next fl3 fl2)} the only possible precondition including the static predicate is $pre\_next\_fly\_v5\_v4$.
-
-%\subsubsection{State constraints}
-%The notion of {\em state-constraint} is very general and has been used in different areas of AI and for different purposes.
-If we restrict ourselves to planning, {\em state-constraints} are abstractions for compactly specifying sets of states.
-For instance, {\em state-constraints} in planning allow to specify the set of states where a given action is applicable, the set of states where a given {\em derived predicate} holds or the set of states that are considered goal states.
-
-%{\em State invariants} is a kind of state-constraints useful for computing more compact state representations~\cite{helmert2009concise} or making {\em satisfiability planning} and {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a classical planning problem $P=\tup{F,A,I,G}$, a {\em state invariant} is a formula $\phi$ that holds at the initial state of a given classical planning problem, $I\models \phi$, and at every state $s$, built from $F$, that is reachable from $I$ by applying actions in $A$.
-
-%The formula $\phi_{I,A}^*$ represents the {\em strongest invariant} and exactly characterizes the set of all states reachable from $I$ with the actions in $A$.
-For instance Figure~\ref{fig:strongest-invariant} shows five clauses that define the {\em strongest invariant} for {\em blocksworld}.
-There are infinitely many strongest invariants, but they are all logically equivalent, and computing the strongest invariant is PSPACE-hard as hard as testing plan existence.
-
-%\begin{figure}[hbt!]
-%  \begin{footnotesize}
-%    \begin{center}
-%$\forall x_1,x_2\ ontable(x_1)\leftrightarrow\neg on(x_1,x_2)$.\\
-%$\forall x_1,x_2\ clear(x_1)\leftrightarrow\neg on(x_2,x_1)$.\\
-%$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
-%$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
-%$\forall x_1,\ldots,x_n\ \neg(on(x_1,x_2)\wedge on(x_2,x_3)\wedge\ldots\wedge on(x_{n-1},x_n)\wedge on(x_n,x_1)).$
-%    \end{center}
-%\end{footnotesize}
-% \caption{\small An example of the strongest invariant for the {\em blocksworld} domain.}
-%\label{fig:strongest-invariant}
-%\end{figure}
-
-%A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-block {\em blocksworld}, $\phi_1=\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a mutex because $block_A$ can only be on top of a single block.
-
-%A {\em domain invariant} is an instance-independent invariant, i.e.
-holds for any possible initial state and set of objects.
-Therefore, if a given state $s$ holds $s\nvDash \phi$ such that $\phi$ is a {\em domain invariant}, it means that $s$ is not a valid state.
-Domain invariants are often compactly defined as {\em lifted invariants} (also called schematic invariants)~\cite{rintanen:schematicInvariants:AAAI2017}.
-For instance, $\phi_2=\forall x:\ (\neg handempty\vee \neg holding(x))$, is a {\em domain mutex} for the {\em blocksworld} because the robot hand is never empty and holding a block at the same time.
-
-%An interesting contribution of our compilation is that the validation of an action model can also be done with {\em state constraints}.
-Given $\Phi$, a set of either state or trajectory constraints, our validate actions can be adapted to check also that the learned model satisfy a constraint $\phi\in\Phi$:
-%\begin{small}
-%\begin{align*}
-%\hspace*{7pt}\pre(\mathsf{validate_{i}})=&\phi\cup\{test_j\}_{j\in 1\leq j<i}\cup\{\neg test_j\}_{j\in i\leq j\leq |\Phi|}\cup \{mode_{val}\},\\
-%\cond(\mathsf{validate_{i}})=&\{\emptyset\}\rhd\{test_i,\neg mode_{val}\}.
-%\end{align*}
-%\end{small}
-
-%This redefinition of validate actions applies also to {\em trajectory constraints} because LTL formulae can be represented using classical action preconditions and goals by encoding the Non-deterministic B\"{u}chi Automaton (NBA), that is equivalent to the corresponding LTL formula, as part of the classical planning tasks~\cite{baier2006planning}.
-
-%Because of the combinatorial nature of the search for a solution plan, the sooner unpromising nodes are pruned from the search the more efficient the computation of a solution plan.
-Constraints can be used to confine earlier the set of possible \strips\ action models and reduce then the learning hypothesis space.
-With regard to our compilation, {\em domain mutex} are useful to reduce the amount of applicable actions for programming a precondition or an effect for a given action schema.
-For example given the {\em domain mutex} $\phi=(\neg f_1\vee \neg f_2)$ such that $f_1\in F_v(\xi)$ and $f_2\in F_v(\xi)$, we can redefine the corresponding programming actions for {\bf removing} the {\em precondition} $f_1\in F_v(\xi)$ from the action schema $\xi\in\mathcal{M}$ as:
-
-%\begin{small}
-%\begin{align*}
-%\hspace*{7pt}\pre(\mathsf{programPre_{f_1,\xi}})=&\{\neg del_{f_1}(\xi),\neg add_{f_1}(\xi), mode_{prog}, pre_{f_1}(\xi), pre_{f_2}(\xi)\},\\
-%\cond(\mathsf{programPre_{f_1,\xi}})=&\{\emptyset\}\rhd\{\neg pre_{f_1}(\xi)\}.
-%\end{align*}
-%\end{small}

### 2019-05-01 doc/aij18/minor-revision/learning_task_def.tex
AIJ18 minor revision folder created

--- 2019-02-19 doc/aij18/major-revision/learning_task_def.tex
@@
 	\item The name, $name(\xi)$, and parameters, $pars(\xi)$, of the action model define the {\em header} of the model.
-	\item $pre(\xi)$, $del(\xi)$ and $add(\xi)$ represent the preconditions, negative effects and positive effects of the action model, respectively, \textcolor[rgb]{1.00,0.00,0.00}{which follow the set of syntactic \strips\ constraints defined in section \ref{basic_planning}; specifically,} $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+	\item $pre(\xi)$, $del(\xi)$ and $add(\xi)$ represent the preconditions, negative effects and positive effects of the action model, respectively, which follow the set of syntactic \strips\ constraints defined in section \ref{basic_planning}; specifically, $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
 \end{itemize}
@@
 In this example $\tau=\langle s_0,${\small\tt (putdown\ B),(stack\ A\ B)}, $s_4\rangle$.
-\textcolor[rgb]{1.00,0.00,0.00}{Therefore, $\tau$ contains a \NO state trajectory because only the initial and final state are observed and the three intermediate states, $s_1$, $s_2$ and $s_3$, are missing; and a \PO action sequence where actions $a_2$ and $a_3$ are observed while $a_1$ and $a_4$ are unknown.} The initial domain model $\mathcal{M}$ only contains two of the four needed headers, but can be completed with the headers {\small\tt(putdown ?v1)} and {\small\tt(stack ?v1 ?v2)} inferred from $\tau$.
+Therefore, $\tau$ contains a \NO state trajectory because only the initial and final state are observed and the three intermediate states, $s_1$, $s_2$ and $s_3$, are missing; and a \PO action sequence where actions $a_2$ and $a_3$ are observed while $a_1$ and $a_4$ are unknown.
+The initial domain model $\mathcal{M}$ only contains two of the four needed headers, but can be completed with the headers {\small\tt(putdown ?v1)} and {\small\tt(stack ?v1 ?v2)} inferred from $\tau$.
 
 A {\em solution} to a learning task $\Lambda=\tup{\mathcal{M},\tau}$ is a domain model $\mathcal{M}'$ that is consistent with the information of $\mathcal{M}$ and with the observed plan trace $\tau$.
-\textcolor[rgb]{1.00,0.00,0.00}{This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan encompass the plan trace $\tau$.}
+This means that the action sequence (plan) that solves the planning problem $\tup {s_0,s_n}$ with $\mathcal{M}'$ along with the state trajectory induced by this plan encompass the plan trace $\tau$.
 
@@
 
-\textcolor[rgb]{1.00,0.00,0.00}{The key to understanding the intricacies of solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ lies in the type of plan trace $\tau = \tup {s_0, \ldots, s_n}$.
+The key to understanding the intricacies of solving a learning task $\Lambda=\tup{\mathcal{M},\tau}$ lies in the type of plan trace $\tau = \tup {s_0, \ldots, s_n}$.
 Let $\pi$ be the plan that solves the planning problem $\tup {s_0,s_n}$ with a learned domain model $\mathcal{M}'$ and $\tau$ be the observation of $\pi$.
 Since $\tau$ is a (partial) observation of the execution of $\pi$, actions, fluents or states traversed by $\pi$ may be missing in $\tau$ (we will use $\tau_s$ and $\tau_a$ to refer to the observed states and observed actions of $\tau$, respectively).
-We distinguish two well differentiated cases:}
+We distinguish two well differentiated cases:
 
 \begin{enumerate}
-\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ determines the length of $\pi$}.
-This happens in three scenarios: (1) when $\tau_a$ is a \FO action sequence (we know the actions of $\pi$); (2) when $\tau_s$ is a \FO state trajectory (the states induced by the actions of $\pi$ are fully known); or (3) when $\tau$ is a \POstar state trajectory (we have at least one fluent for every state induced by $\pi$, in which case we know there is a single action that transitions from every state of $\tau_s$).}
+\item \textbf{$\tau$ determines the length of $\pi$}.
+This happens in three scenarios: (1) when $\tau_a$ is a \FO action sequence (we know the actions of $\pi$); (2) when $\tau_s$ is a \FO state trajectory (the states induced by the actions of $\pi$ are fully known); or (3) when $\tau$ is a \POstar state trajectory (we have at least one fluent for every state induced by $\pi$, in which case we know there is a single action that transitions from every state of $\tau_s$).
 
@@
 Normally, the process of obtaining state feedback from sensors (or the processing of the sensor readings) is associated with a given sampling frequency that misses intermediate data between two subsequent sensor readings.
-    \item \textcolor[rgb]{1.00,0.00,0.00}{The assumption of having \POstar state trajectories seems more appropriate to reflect a real-world sensor reading but still requires that at least one fluent of every state traversed by $\pi$ is captured by the sensors.}
+    \item The assumption of having \POstar state trajectories seems more appropriate to reflect a real-world sensor reading but still requires that at least one fluent of every state traversed by $\pi$ is captured by the sensors.
     \end{itemize}
@@
 
-\item \textcolor[rgb]{1.00,0.00,0.00}{\textbf{$\tau$ does not identify the length of $\pi$}.
+\item \textbf{$\tau$ does not identify the length of $\pi$}.
 This happens when $\tau_a$ and $\tau_s$ are both partially observed (\PO) or non-observable (\NO) action/state trajectories.
@@
 This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$.
-As it is also well known, the worst case complexity of \strips\ planning is PSPACE-complete so we can conclude our learning task becomes PSPACE-complete instead of NP-complete when combining \PO and \NO state trajectories and action sequences.}
+As it is also well known, the worst case complexity of \strips\ planning is PSPACE-complete so we can conclude our learning task becomes PSPACE-complete instead of NP-complete when combining \PO and \NO state trajectories and action sequences.
 
-   \textcolor[rgb]{1.00,0.00,0.00}{ In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).
-Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.}
+In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).
+Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.
 
-%\FAMA shows that classical planning is a suitable approach for this particular scenario
 \end{enumerate}
-
-%\FAMA represents one step ahead towards learning action models without assuming observed actions.
-The main novelty of \FAMA with respect to other approaches lies in that our system is capable of handling \PO and \NO action sequences, which combined with \PO and \NO state trajectories, make the learning task more challenging.
-This essentially brings one key difference: the transition between two given observed states may now involve more than one action; i.e., $\theta(s_i,\tup{a_1,\ldots,a_k})=s_{i+1}$, with $k \geq 1$, $k$ unknown and unbound, and so the horizon of the input plan traces is no longer known now.
-Table \ref{tab:complex} shows that the worst case complexity of learning \strips\ action models becomes PSPACE-complete when combining \PO and \NO state trajectories and action sequences.
-
-%\begin{table}[ht]
-%\centering
-%\begin{tabular}{c|c|c|c|c|}
-%	& \multicolumn{4}{c|}{\emph{state observability}} \\ \cline{2-5}
-%	\multirow{1}{*}{\emph{action}} & \FO & \POstar & \PO & \NO\\ {\emph{observability}} & & & & \\ \hline
-%	\FO & - & NP-complete & NP-complete & NP-complete \\ \hline
-%	\PO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
-%	\NO & NP-complete & NP-complete & \textbf{PSPACE-complete} & \textbf{PSPACE-complete} \\ \hline
-%\end{tabular}
-%\caption{Complexity of learning tasks according to the type of input trace}
-%\label{tab:complex}
-%\end{table}

### 2019-05-01 doc/aij18/minor-revision/aij18.tex
AIJ18 minor revision folder created
Addressed question 1 and minor comments

--- 2018-12-26 doc/aij18/major-revision/aij18.tex
@@
 \input{background}
-%\input{motivation}
 \input{learning_task_def}
@@
 This great flexibility of \FAMA to different amounts of minimal input knowledge impacts the complexity of the planning tasks.
-When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, the planning task to be solved becomes a PSPACE-complete scenario.
+When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, \textcolor{red}{the learning task to be solved can no longer be addressed using current approaches}.
 This is one of the reasons that justifies our compilation-to-planning scheme in contrast to the most extended SAT-solving scheme used in many existing learning systems.
@@
 The application of the semantic-based precision and recall allows us to assess the validity of the learned models even in domains where a reference model is not available.
-%Remarkably, the action models of the {\em gripper-strips}, {\em n-puzzle} and {\em grid-visit-all} domains were perfectly learned from only observations of the initial and final states.
 
@@
 The combination of such exploration techniques with our learning approach is an appealing research direction towards the bootstrapping of planning action models.
-
-%When action plans are not available, our approach still produces action models that are compliant with the input information.
-In this case, since learning is not constrained by actions, operators can be reformulated changing their semantics, in which case the comparison with a reference model turns out to be tricky.
-
-%When example plans are available, our approach is also able to compute accurate action models from small sets of learning examples.
-Our experimental results report that \FAMA achieves {\em precision} and {\em recall} scores that are over 0.7 for ten out of the fourteen IPC domains, using only two plan traces (with ten actions each) and in little computation time (less than a second).
-Our experimental evaluation also shows that, in the particular setting of minimal input data, \FAMA outperforms \ARMS.
-
-%On the other hand, this work also introduced a semantic version of the {\em precision} and {\em recall} metrics.
-These two metrics, widely used in ML, separately measure the soundness and completeness of the learned models, which facilitate the identification of model flaws.
-Our brand new semantic version of {\em precision} and {\em recall} do not require a GTM and evaluate \strips\ action models with respect to a given set of observations.
-In addition, when no reformulation occurs, these new metrics behave similarly to their syntactic counterparts.
 

### 2019-05-01 doc/aij18/minor-revision/evaluation.tex
Addressed question 1 and minor comments

@@
 
-We can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the the total number of all the possible preconditions and effects of an action model.
+We can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the total number of all the possible preconditions and effects of an action model.
 This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM.

### 2019-05-01 doc/aij18/minor-revision/experiments.tex
Addressed question 1 and minor comments

@@
 The higher complexity of this setting is also reflected in the appearance of some timeouts when solving the learning task.
-The first timeout is found in the \emph{grid} domain at $|\mathcal{T}| = 3$; the \emph{floortile} times out with $|\mathcal{T}| = 4$; and by the time $|\mathcal{T}|$ reaches 10 the number of domains were no solution is found is 6, adding the \emph{npuzzle, parking, rovers and zenotravel} domains.
+The first timeout is found in the \emph{grid} domain at $|\mathcal{T}| = 3$; the \emph{floortile} times out with $|\mathcal{T}| = 4$; and by the time $|\mathcal{T}|$ reaches 10 the number of domains where no solution is found is 6, adding the \emph{npuzzle, parking, rovers and zenotravel} domains.
 We can observe in Figure~\ref{fig:pspace_quality} the opposite behaviour to Figure~\ref{fig:np_quality}; that is, we find a drop of the quality as the input knowledge increases.

### 2019-05-01 doc/aij18/minor-revision/introduction.tex
Addressed question 1 and minor comments

@@
 \begin{itemize}
-\item compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces.
-\item the plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states.
-\item the planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions.
+\item compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces,
+\item the plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states,
+\item the planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions,
 \item a semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model.

### 2019-05-01 doc/aij18/minor-revision/learning_action_models.tex
Addressed question 1 and minor comments

@@
 
-Given a solution plan $\pi_\Lambda$ that solves $P_{\Lambda}$, the set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ are computed in linear time and space.
+Given a solution plan $\pi_\Lambda$ that solves $P_{\Lambda}$, the set of action models $\mathcal{M}'$ that solves $\Lambda=\tup{\mathcal{M},\tau}$ is computed in linear time and space.
 In order to do so, $\pi_\Lambda$ is executed in the initial state $I_{\Lambda}$ and the action model $\mathcal{M}'$ will be given by the fluents $pre_f(\xi)$, $del_f(\xi)$ and $add_f(\xi)$ that are set to true in the last state reached by $\pi_\Lambda$, $s_g=\theta(I_\Lambda,\pi_\Lambda)$.

### 2019-05-01 doc/aij18/minor-revision/learning_task_def.tex
Addressed question 1 and minor comments

@@
 \FAMA addresses the learning and evaluation of PDDL action models that follow the \strips\ requirement~\cite{mcdermott1998pddl,fox2003pddl2}.
-An \strips\ action model is a tuple $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
+A \strips\ action model is a tuple $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$ where:
 
@@
 This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$.
-As it is also well known, the worst case complexity of \strips\ planning is PSPACE-complete so we can conclude our learning task becomes PSPACE-complete instead of NP-complete when combining \PO and \NO state trajectories and action sequences.
-
+\textcolor{red}{SAT approaches, on the other hand, are no longer obviously applicable given the lack of a length bound}.
 In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).

### 2019-05-01 doc/aij18/minor-revision/learning_action_models.tex
AIJ18-minor: Addressed questions 2 and 3

@@
 \item A fluent, $mode_{prog}$, to indicate whether action models are being programmed or validated.
+\item \textcolor{red}{A fluent, $action_{applied}$, to force the execution of at least one action between two observed states.}
 \end{itemize}
@@
 &\{add_{p}(\xi)\}\rhd\{p(\omega)\}_{\forall p\in\Psi\xi},\\
+& \textcolor{red}{\{\emptyset\}\rhd\{action_{applied}\},}\\
 &\{mode_{prog}\}\rhd\{\neg mode_{prog}\}.
@@
 When the input plan trace contains observed actions, the extra conditional effects
-
-$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.\\
+$\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.
+\textcolor{red}{As a side note, if we are sure that the input trace contains all actions (\FO sequence of actions) this conditional effect should be redefined as regular preconditions and effects.
+Doing so will make it impossible for the planner to execute apply actions outside of the observed ones.
+Additionally, for traces with \POstar state trajectories, we can force that only one apply action is executed between two observed states by adding the preconditions $\neg action_{applied}$.}\\
 
@@
 The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j\in\tau$.
+\textcolor{red}{If the input plan trace contains observed actions, we add an extra precondition, $at_{i+1}$, where $i$ is the index of the last observed action before the state we are validating.
+This additional precondition ensures that applied actions are also ordered with respect to the observed states}
 
@@
 	\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1},test_j\} \cup \\
-	&\{\neg f\}_{\forall f\in s_m^t, f \notin s_0^{t+1}}\cup \{f\}_{\forall f\in s_0^{t+1}, f \notin s_m^t},\\
+	&\textcolor{red}{ \{\neg action_{applied}\} \cup }
+	 \\
+	&\{\neg f\}_{\forall f\in s_m^t, f \notin s_0^{t+1}}\cup \{f\}_{\forall f\in s_0^{t+1}, f \notin s_m^t} \cup \\
 	&\{\neg f\}_{\forall f\in F_{\pi_t}}\cup \{f\}_{\forall f\in F_{\pi_{t+1}}}.\\

### 2019-05-01 doc/aij18/minor-revision/background.tex
minor modifications

@@
 
-Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2} -- almost all approaches rely on a comparison between the learned model and a {\em Ground-Truth Model}), the metrics used in the comparison (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
+Table \ref{table:models_comparison2} summarizes the main characteristics of the evaluation of the learned action models based on the type of evaluation method (first column of Table \ref{table:models_comparison2}), the metrics used in the evaluation (second column of Table \ref{table:models_comparison2}) and the number of tested domains alongside the size of the training dataset (third column of Table \ref{table:models_comparison2}).
+\textcolor[rgb]{1.00,0.00,0.00}{Regarding the evaluation method, almost all approaches rely on a comparison between the learned model and a {\em Ground-Truth Model} (GTM).
+By GTM we refer to an engineered model that knowledge modelers have adopted as being correct with respect to a domain, and that the planning community accepts as such.}
 

### 2019-05-01 doc/aij18/minor-revision/evaluation.tex
minor modifications

@@
 
-We can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the total number of all the possible preconditions and effects of an action model.
+\textcolor[rgb]{1.00,0.00,0.00}{As commented before, the large majority of approaches shown in Table \ref{table:models_comparison2} evaluate the learned action models wrt GTM, where the GTM is used to generate the test set of plan traces that the algorithms use to learn from.
+That is, learning approaches are evaluated in domains taken from the IPCs.} We can also observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the total number of all the possible preconditions and effects of an action model.
 This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM.

### 2019-05-01 doc/aij18/minor-revision/learning_action_models.tex
minor modifications

@@
 $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.
-\textcolor{red}{As a side note, if we are sure that the input trace contains all actions (\FO sequence of actions) this conditional effect should be redefined as regular preconditions and effects.
+\textcolor{red}{As a side note, when the input trace contains all actions (\FO sequence of actions), this conditional effect should be redefined as regular preconditions and effects.
 Doing so will make it impossible for the planner to execute apply actions outside of the observed ones.

### 2019-05-01 doc/aij18/minor-revision/learning_task_def.tex
minor modifications

@@
 \textcolor{red}{SAT approaches, on the other hand, are no longer obviously applicable given the lack of a length bound}.
-In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).
+\textcolor[rgb]{0.00,0.00,1.00}{In this case, the learning task is not SAT compilable and thereby we can no longer leverage the potential of SAT solvers given the lack of a length bound.} In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).
 Therefore, when we assume partial observability in both actions and states, a learning approach must consider that the length of the observed plan traces is not an indication of the actual length of the plan, which motivates and justifies the use of planning, as our proposal of compiling the learning task to a classical planning problem.

### 2019-05-02 doc/aij18/minor-revision/aij18.tex
AIJ18-minor: minor fixes

@@
 
+\textcolor{red}{The lack of observed actions and of a length bound in the input examples is addressed using a compilation-to-planning scheme, which allows a planner to {\em fill the gaps} in the input traces in a way that is consistent with what is observed.
+Additionally, our planning-based solving scheme allows us to leverage {\em off-the-shelf} classical planners and benefit from the multiple advances in classical planning research.}
+
+%This great flexibility of \FAMA to different amounts of minimal input knowledge impacts the complexity of the planning tasks.
+When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, \textcolor{red}{the learning task to be solved can no longer be addressed using current approaches}.
+This is one of the reasons that justifies our compilation-to-planning scheme in contrast to the most extended SAT-solving scheme used in many existing learning systems.
+
 Besides its capacity of working with minimal observability, \FAMA is also able to learn from very small amounts of input knowledge, a clear advantage in domains where obtaining enough training samples is difficult or costly.
@@
 Unlike extensive-data ML approaches, our work explores an alternative research direction that exploits logic reasoning to learn sound models from minimal input observability.
-
-This great flexibility of \FAMA to different amounts of minimal input knowledge impacts the complexity of the planning tasks.
-When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, \textcolor{red}{the learning task to be solved can no longer be addressed using current approaches}.
-This is one of the reasons that justifies our compilation-to-planning scheme in contrast to the most extended SAT-solving scheme used in many existing learning systems.
-Additionally, our planning-based solving scheme allows us to leverage {\em off-the-shelf} classical planners and benefit from the multiple advances in classical planning research.
 

### 2019-05-02 doc/aij18/minor-revision/background.tex
AIJ18-minor: minor fixes

@@
 \begin{enumerate}
-\item we say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$;  that is $\theta(s_i,\tup{a})=s_{i+1}$.
+\item We say that $\tau_s$ is a fully-observable (\FO) state trajectory if every observed intermediate state of $\tau_s$ is a full assignment of values to fluents, and there exists a single action that transitions from every state $s_i$ to state $s_{i+1}$ in $\tau_s$;  that is $\theta(s_i,\tup{a})=s_{i+1}$.
 This case clearly states that $\tau_s = s$, meaning that $\forall s_i \in \tau_s$, $s_i$ comprises all the literals of the corresponding state in the trajectory $s$ of the plan $\pi$.
 Formally, $\forall i, 1 \leq i  < n, |s_i| = |F|$.
-\item we say that $\tau_s$ is a partially-observable (\PO) state trajectory if at least one intermediate state of $\tau_s$ is a partial assignment of values to fluents.
+\item We say that $\tau_s$ is a partially-observable (\PO) state trajectory if at least one intermediate state of $\tau_s$ is a partial assignment of values to fluents.
 Formally, $\exists i, 1 \leq i  < n, |s_i| < |F|$.
@@
 \begin{enumerate}
-\item when \textbf{all} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, $\tau_s$ is a \emph{non-observable} (\NO) state trajectory.
+\item When \textbf{all} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, $\tau_s$ is a \emph{non-observable} (\NO) state trajectory.
 Formally, $\forall i, 1 \leq i < n,  s_i= \emptyset$; i.e., $|s_i| = 0$.
-\item when \textbf{none} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, we will refer to $\tau_s$ as a \POstar state trajectory.
+\item When \textbf{none} of the $n-1$ intermediate states of $s$ are \textbf{missing} in $\tau_s$, we will refer to $\tau_s$ as a \POstar state trajectory.
 Formally, $\exists i, 1 \leq i < n, s_i \neq \emptyset$; i.e., $0 < |s_i| < |F|$.
@@
 \begin{enumerate}
-\item when \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$.
+\item When \textbf{all} of the actions of $\pi$ appear in $\tau_a$, we say that $\tau_a$ is a fully-observable (\FO) action sequence; i.e.,  $\tau_a=\pi$.
 In this case, $\tau_a$ contains all the necessary actions to transit every state $s_{i-1}$ to its corresponding successor state $s_{i}$, from $s_0$ to $s_n$.
 This is the type of input trace accepted by all the existing learning approaches (see section \ref{related_work} for details).
-\item when \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence.
+\item When \textbf{some} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a partially observable (\PO) action sequence.
 In this case, at least one of the necessary actions of the plan $\pi$ is missing in $\tau_a$.
 Formally, $\exists i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.
-\item when \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence.
+\item When \textbf{none} of the actions of $\pi$  appear in $\tau_a$, we say that $\tau_a$ is a non-observable (\NO) action sequence.
 Formally, $\forall i, 1 \leq i \leq n, a_i \in \pi \wedge a_i \notin \tau_a$.

### 2019-05-02 doc/aij18/minor-revision/introduction.tex
AIJ18-minor: minor fixes

@@
 \begin{itemize}
-\item compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces,
-\item the plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states,
-\item the planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions,
-\item a semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model.
+\item Compiling the task of learning a \strips\ planning model into a planning task that is automatically built from a set of input plan traces.
+\item The plan traces are correct (no noise is considered in the observations) but may be incomplete in the number of observed actions as well as in the number and contents of the observed states.
+\item The planning task resulting from the compilation comprises actions for programming the preconditions and effects of the \strips\ actions and actions for validating the learned \strips\ actions.
+\item A semantic evaluation proposal that enables to assess a learned model beyond a merely syntactic comparison to a reference model.
 \end{itemize}

### 2019-05-02 doc/aij18/minor-revision/learning_action_models.tex
AIJ18-minor: minor fixes

@@
 $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.
-\textcolor{red}{As a side note, when the input trace contains all actions (\FO sequence of actions), this conditional effect should be redefined as regular preconditions and effects.
+\textcolor{red}{As a side note, when the input trace contains all actions (\FO sequence of actions), this conditional effect is redefined as regular preconditions and effects.
 Doing so will make it impossible for the planner to execute apply actions outside of the observed ones.
-Additionally, for traces with \POstar state trajectories, we can force that only one apply action is executed between two observed states by adding the preconditions $\neg action_{applied}$.}\\
+Additionally, for traces with \POstar or \FO state trajectories, we can force that only one apply action is executed between two observed states by adding the preconditions $\neg action_{applied}$.}\\
 

### 2019-05-02 doc/aij18/minor-revision/experiments.tex
AIJ18-minor: rephrasing some points

@@
 \item {\bf Reproducibility}.
-We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this repository {\em https://github.com/sjimenezgithub/strips-learning} so any experimental data reported in the paper is fully reproducible.
+We make fully available the compilation source code, the evaluation scripts and the used benchmarks at this repository {\em https://github.com/daineto/meta-planning} so any experimental data reported in the paper is fully reproducible.
 \end{itemize}

### 2019-05-02 doc/aij18/minor-revision/learning_action_models.tex
AIJ18-minor: rephrasing some points

@@
 
-When the input plan trace contains observed actions, the extra conditional effects
+When the input plan trace contains \PO observed actions, the extra conditional effects
 $\{at_{i},plan(name(a_i),\Omega^{ar(a_i)},i)\}\rhd\{\neg at_{i},at_{i+1}\}_{\forall i\in [1,n]}$ are included in the $\mathsf{apply_{\xi,\omega}}$ actions to ensure that actions are applied in the same order as they appear in $\tau$.
-\textcolor{red}{As a side note, when the input trace contains all actions (\FO sequence of actions), this conditional effect is redefined as regular preconditions and effects.
-Doing so will make it impossible for the planner to execute apply actions outside of the observed ones.
-Additionally, for traces with \POstar or \FO state trajectories, we can force that only one apply action is executed between two observed states by adding the preconditions $\neg action_{applied}$.}\\
+\textcolor{red}{If the sequence of actions is FO, then this conditional effect is redefined as regular preconditions and effects in the apply actions.
+Doing so will make it impossible for the planner to execute apply actions other than the observed actions in the plan trace.
+Furthermore, for traces with \POstar or \FO state trajectories, we force that only one apply action is executed between two observed states by adding the precondition $\neg action_{applied}$ in every apply action.}
 
@@
 \textcolor{red}{If the input plan trace contains observed actions, we add an extra precondition, $at_{i+1}$, where $i$ is the index of the last observed action before the state we are validating.
-This additional precondition ensures that applied actions are also ordered with respect to the observed states}
+This additional precondition ensures that applied actions are also ordered with respect to the observed states.}
 

### 2019-05-02 doc/aij18/minor-revision/letter.tex
AIJ18-minor: added response letter

\documentclass{article}

\usepackage[framemethod=tikz]{mdframed}
\usepackage{xspace}

\mathchardef\mh="2D
\newcommand{\pre}{\mathsf{pre}}       % precondition
\newcommand{\eff}{\mathsf{eff}}       % effect
\newcommand{\cond}{\mathsf{cond}}     % conditional effect
\newcommand{\add}{\mathsf{add}}       % add effect
\newcommand{\del}{\mathsf{del}}       % delete effect
\newcommand{\PE}{\mathrm{PE}}         % precondition
\newcommand{\PSPACE}{\mathrm{PSPACE}} % PSPACE
\newcommand{\strips}{\textsc{Strips}} % Strips
\newcommand{\FAMA}{{\small {\sffamily FAMA}}\xspace}

\usepackage{amsthm}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage{times}
\usepackage{helvet}
\usepackage{courier}
\usepackage{pstricks}
\usepackage{pst-node}
\usepackage{multirow}
\usepackage{listings}
\usepackage{xspace}

\title{\textbf{Response Letter}}
\begin{document}
\maketitle

We thank the reviewers and editors for their helpful comments to improve the paper.

\vspace{0.3cm}

Following, we explain how the four remarks pointed in the editors' comments have been addressed.

\vspace{0.1cm}

\begin{mdframed}[hidealllines=true,backgroundcolor=gray!20]
That your learning problem can be solved by planning does not mean it it PSPACE-complete; only that it is a member of PSPACE. To show PSPACE-hardness, you would have to show that arbitrary planning problems can arise in your setup. You either need to provide an argument for the latter, or to reformulate our text removing the claim of PSPACE-completeness and merely pointing out that SAT approaches are no longer obvously applicable given the lack of a length bound.
\end{mdframed}

We agree with the editors' observations as well as reviewer 1's remarks about the complexity of the learning task. Certainly, the fact that the learning task is solved via planning does not make it a PSPACE-complete task; in order to prove this, we should indeed show a planning problem is reducible to a learning problem. Actually, our intuition is that both tasks, learning and planning, are equivalent, and that reducibility is applicable in both directions. Nevertheless, this is an issue we must still study in depth and we expect to tackle it in the near future.

For the purpose of this paper, we have removed any reference to PSPACE-completeness. We simply highlight that when the length of the plan trace is unknown, we can no longer leverage the advantages of using a SAT solver as the large majority of the current learning approaches do. This is actually the key contribution of this paper: the minimal observation needed by \FAMA to learn an action model is a plan trace composed solely of a full observable initial state and a partially observable final state (goal state).

Clarifications regarding this remark are highlighted in red in page 11. We also rephrased some sentences of the section Conclusions.

\vspace{0.1cm}

\begin{mdframed} [hidealllines=true,backgroundcolor=gray!20]
It needs to be made clearer whether and how ordering constraints between actions and states are handled in your compilation scheme.
\end{mdframed}

This comment along with the next one are two insightful observations about the compilation scheme. Actually, we do have a mechanism in \FAMA to tackle both issues, although we did not explain it in the previous version of the document because we considered it was a rather technical issue. We have now fixed this.

The set $F_{\Lambda}$ includes a fluent $action_{applied}$ to force the execution of at least one action between two observed states (page 13). Then, when validating a plan trace that contains an interleaved sequence of actions and states, the \texttt{validate} actions include an extra precondition, $at_{i+1}$, where $i$ is the index of the last observed action before the state we are validating. This additional precondition ensures that applied actions are also ordered with respect to the observed states. This is highlighted in red in page 15.

\vspace{0.1cm}

\begin{mdframed} [hidealllines=true,backgroundcolor=gray!20]
It needs to be made clearer whether and how action/state insertion can be forbidden in your compilation scheme when part of every state is observed.
\end{mdframed}

When at least a fluent of every state is observed in the plan trace, we need to ensure that only one action is executed between two observed states. This is done by adding the precondition $\neg action_{applied}$ in every \texttt{apply} action.

In the case of a fully known plan trace (every action is observed and every state too even partially), the conditional effects formed by fluents $F_{\pi}$ and $at_i$ are transformed into a regular precondition+effect in the \texttt{apply} actions. This mechanism prevents the planner from executing  \texttt{apply} actions other than the ones that appear in the plan trace.

This is highlighted in red in page 14.

\begin{mdframed} [hidealllines=true,backgroundcolor=gray!20]
The term GTM must be explained and properly discussed.
\end{mdframed}

We agree with the remarks point out by reviewer 3. We borrow the definition of GTM provided by the reviewer (highlighted in red in page 5). Likewise, in section 5, we clarify that approaches that use the GTM as the reference point to compare the learned models, use plan traces generated with the GTM; i.e., they  use the domains of the IPC benchmarks.

\end{document}

### 2019-05-02 doc/aij18/minor-revision/aij18.tex
AIJ18-minor: rephrasing some points
AIJ18-minor: final version

@@
 \title{Learning action models with minimal observability}
-\author[label1]{Diego Aineto}
+\author[label1]{Diego Aineto\corref{cor1}}
+\ead{dieaigar@dsic.upv.es}
 \author[label1]{Sergio Jim\'{e}nez Celorrio}
+\ead{serjice@dsic.upv.es}
 \author[label1]{Eva Onaindia}
-\address[label1]{Department of Computer Systems and Computation, Universitat Polit\`ecnica de Val\`encia. Spain}
+\ead{onaindia@dsic.upv.es}
 
-%% use optional labels to link authors explicitly to addresses:
-%% \author[label1,label2]{<author name>}
-%% \address[label1]{<address>}
-%% \address[label2]{<address>}
+\cortext[cor]{Corresponding author}
+
+\address[label1]{Department of Computer Systems and Computation, Universitat Polit\`ecnica de Val\`encia. Camino de Vera s/n. 46022 Valencia, Spain}
 
@@
 In this paper we have presented \FAMA, an approach for learning \strips\ action models based on the compilation of the learning task into a planning task.
-The distinctive characteristic of \FAMA over other state-of-the-art approaches is its ability to learn from minimal input knowledge and, more particularly, from minimal observability plan traces with no observed actions and no observed intermediate states.
+The distinctive characteristic of \FAMA over other state-of-the-art approaches is its ability to learn from minimal observability, i.e., plan traces with no observed actions and no observed intermediate states.
 By relaxing the input constraints, \FAMA opens up the way for action model learning to operate on real-world problems, as opposite to current approaches where the heavy input restrictions have limited their applicability to synthetic benchmarks.
-Our approach is thus very well suited for learning action models in data-based applications where the only observable information is a possibly incomplete sequence of partially observed states.
+Our approach is thus suitable for learning action models in data-based applications where the only observable information is a possibly incomplete sequence of partially observed states.
 
@@
 Additionally, our planning-based solving scheme allows us to leverage {\em off-the-shelf} classical planners and benefit from the multiple advances in classical planning research.}
-
-%This great flexibility of \FAMA to different amounts of minimal input knowledge impacts the complexity of the planning tasks.
-When the length of the input plan traces is unknown, that is, we ignore the number of steps of the decision-making process of the agent, \textcolor{red}{the learning task to be solved can no longer be addressed using current approaches}.
-This is one of the reasons that justifies our compilation-to-planning scheme in contrast to the most extended SAT-solving scheme used in many existing learning systems.
 

### 2019-05-02 doc/aij18/minor-revision/background.tex
AIJ18-minor: final version

@@
 
-In the following, we present a comprehensive insight of the particularities of the seven systems presented in Table \ref{table:models_comparison1} and Table \ref{table:models_comparison2}.
+In the following, we present a comprehensive insight of the particularities of the eight systems presented in Table \ref{table:models_comparison1} and Table \ref{table:models_comparison2}.
 This exposition will also help us to highlight in section \ref{task_definition} the value of our contribution \FAMA.

### 2019-05-02 doc/aij18/minor-revision/evaluation.tex
AIJ18-minor: final version

--- 2019-05-01 doc/aij18/minor-revision/evaluation.tex
@@
 
-We can observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the total number of all the possible preconditions and effects of an action model.
+\textcolor[rgb]{1.00,0.00,0.00}{As commented before, the large majority of approaches shown in Table \ref{table:models_comparison2} evaluate the learned action models wrt a GTM.
+These GTM are also  used to generate the learning plan traces and they are normally domain definitions taken from the IPC.} We can also observe in Table \ref{table:models_comparison2} that most of the approaches use a similar syntax-based metric that consists in (1) counting the missing and extra fluents that appear in the learned model wrt the GTM and (2) normalizing this error by the total number of all the possible preconditions and effects of an action model.
 This is an \emph{optimistic} metric since error rates are not normalized by the size of the actual GTM.
 The set of preconditions and effects of the GTM is usually smaller than the set of all possible preconditions and effects and thereby it turns out that these syntax-based metrics may output error rates below 100\% for totally wrong learned models.
-To overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}.
+
+In order to overcome this limitation we propose to use two standard metrics from ML, {\em precision} and {\em recall}, that are frequently used in pattern recognition, information retrieval and binary classification~\cite{davis2006relationship}.
 These two syntactic metrics are generally more informative than counting the number of errors between the learned action models and the GTM:

### 2019-05-02 doc/aij18/minor-revision/introduction.tex
AIJ18-minor: final version

@@
 The construction of a \strips\ planning model starts out from a set of plan traces containing the observation of several plan executions.
-As mentioned above, a plan trace may comprise none of the actions executed by the agent but must include, at least, the initial state $s_0$ and final state $s_f$ of the execution.
+As mentioned above, a plan trace may comprise none of the actions executed by the agent but must include, at least, the initial state $s_0$ and final state $s_n$ of the execution.
 The compilation scheme lies in defining a planning task from the set of plan traces using a set of \emph{building actions} that insert the preconditions and effects of a learned action, and a set of \emph{validating actions} that validate the learned actions in the plan traces.
 Hence, a solution to this planning task is a plan that determines the preconditions and effects of the actions of a \strips\ planning model $M$ while ensuring consistency with the input traces.
-We say that a model $M$ is \emph{consistent} with a plan trace when $M$ can produce a solution $\pi$ to the planning problem $\tup{s_0,s_f}$ so that: (1) $\pi$ contains the observed actions of the plan trace, if any, and (2) the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.
+We say that a model $M$ is \emph{consistent} with a plan trace when $M$ can produce a solution $\pi$ to the planning problem $\tup{s_0,s_n}$ so that: (1) $\pi$ contains the observed actions of the plan trace, if any, and (2) the states generated by the application of $\pi$ to $s_0$ will encompass all the (possibly) partially observed states of the trace.
 

### 2019-05-02 doc/aij18/minor-revision/learning_task_def.tex
AIJ18-minor: minor fixes
AIJ18-minor: final version

--- 2019-05-01 doc/aij18/minor-revision/learning_task_def.tex
@@
 	\begin{verbatim}
-	;;; State observation
+	;;; Final state observation
 	(clear A) (on A B) (ontable B)
@@
 This justifies the use of \textbf{planning techniques} for solving the learning task, which can now be interpreted as \emph{filling the gap} between two observable points of $\tau$.
-\textcolor{red}{SAT approaches, on the other hand, are no longer obviously applicable given the lack of a length bound}.
+\textcolor{red}{SAT approaches, on the other hand, are no longer straightforwardly applicable given the lack of a length bound for the observed plan trace}.
 In this particular scenario, the actual number of plans consistent with the given observed plan trace is also unbounded and grows exponentially with the actual length of the plans (that is now unknown).

### 2019-05-15 doc/aaai20-goalRecognition/goalRecognition.tex
added first aaai20 paper on model-free goal recognition

--- 2019-02-12 doc/ijcai-19-goalRecognition/goalRecognition.tex
@@
-%%%% ijcai19.tex
+\def\year{2020}\relax
+%File: formatting-instruction.tex
+\documentclass[letterpaper]{article} %DO NOT CHANGE THIS
+\usepackage{aaai20}  %Required
+\usepackage{times}  %Required
+\usepackage{helvet}  %Required
+\usepackage{courier}  %Required
+\usepackage{url}\urlstyle{rm}  %Required
+\usepackage{graphicx}  %Required
+\frenchspacing  %Required
+\setlength{\pdfpagewidth}{8.5in}  %Required
+\setlength{\pdfpageheight}{11in}  %Required
+\setcounter{secnumdepth}{0}
 
-\typeout{IJCAI-19 Instructions for Authors}
+% Addiotional packages
+\usepackage{amsmath}
+\usepackage{amssymb}
+\usepackage{amsthm}
+\usepackage{multirow}
+\usepackage{multicol}
+\usepackage{tikz}
+\usetikzlibrary{arrows,automata}
+\usetikzlibrary{calc,backgrounds,positioning,fit}
 
-% These are the instructions for authors for IJCAI-19.
+\newcommand{\tup}[1]{{\langle #1 \rangle}}
 
-\documentclass{article}
-\pdfpagewidth=8.5in
-\pdfpageheight=11in
-% The file ijcai19.sty is NOT the same than previous years'
-\usepackage{ijcai19}
-
-% Use the postscript times font!
-\usepackage{times}
-\usepackage{soul}
-\usepackage{url}
-\usepackage[hidelinks]{hyperref}
-\usepackage[utf8]{inputenc}
-\usepackage[small]{caption}
-\usepackage{graphicx}
-\usepackage{amsmath}
-\usepackage{booktabs}
-\usepackage{algorithm}
-\usepackage{algorithmic}
-\urlstyle{same}
-
-%%%%%%%%%%%%%%%%%% Added for this paper
-\usepackage{amsthm}
-\usepackage{listings}% http://ctan.org/pkg/listings
-\lstset{
-  basicstyle=\ttfamily,
-  mathescape
-}
-\usepackage{ wasysym }
-\newcommand{\tup}[1]{{\langle #1 \rangle}}
 \newcommand{\pre}{\mathsf{pre}}     % precondition
@@
 \newcommand{\PE}{\mathrm{PE}}     % precondition
-\newcommand{\strips}{\textsc{Strips}}
+\newcommand{\strips}{\textsc{Strips}}     % precondition
 
@@
 
-%%%%%%%%%%%%%
-
-% the following package is optional:
-%\usepackage{latexsym}
-
-% Following comment is from ijcai97-submit.tex:
-% The preparation of these files was supported by Schlumberger Palo Alto
-% Research, AT\&T Bell Laboratories, and Morgan Kaufmann Publishers.
-% Shirley Jowell, of Morgan Kaufmann Publishers, and Peter F.
-% Patel-Schneider, of AT\&T Bell Laboratories collaborated on their
-% preparation.
-
-% These instructions can be modified and used in other conferences as long
-% as credit to the authors and supporting agencies is retained, this notice
-% is not changed, and further modification or reuse is not restricted.
-% Neither Shirley Jowell nor Peter F. Patel-Schneider can be listed as
-% contacts for providing assistance without their prior permission.
-
-% To use for other conferences, change references to files and the
-% conference appropriate and use other authors, contacts, publishers, and
-% organizations.
-% Also change the deadline and address for returning papers and the length and
-% page charge instructions.
-% Put where the files are available in the appropriate places.
+%PDF Info Is Required:
+\pdfinfo{}
 
@@
 
-% Single author syntax
-%\author{
-%    Sarit Kraus
-%    \affiliations
-%    Department of Computer Science, Bar-Ilan University, Israel \emails
-%    pcchair@ijcai19.org
-%}
-
-% Multiple author syntax (remove the single-author syntax above and the \iffalse ... \fi here)
-% Check the ijcai19-multiauthor.tex file for detailed instructions
-\author{
-Diego Aineto$^1$\and
-Sergio Jim\'enez$^1$\and
-Eva Onaindia$^1$\And
-\and
-Miquel Ram\'irez$^2$
-\affiliations
-$^1${\small Departamento de Sistemas Inform\'aticos y Computaci\'on. Universitat Polit\`ecnica de Val\`encia. Valencia, Spain}\\
-$^2${\small School of Computing and Information Systems. The University of Melbourne. Melbourne, Victoria. Australia}
-\emails
-{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es, miquel.ramirez@unimelb.edu.au}}
+%\author{\#}
+% Commented for blind submission
+\author{Diego Aineto \and Sergio Jim\'enez\and Eva Onaindia\\
+{\scriptsize Departamento de Sistemas Inform\'aticos y Computaci\'on}\\
+{\scriptsize Universitat Polit\`ecnica de Val\`encia.}\\
+{\scriptsize Camino de Vera s/n. 46022 Valencia, Spain}\\
+{\scriptsize \{dieaigar,serjice,onaindia\}@dsic.upv.es}
+\And Miquel Ram\'irez\\
+{\scriptsize School of Computing and Information Systems}\\
+{\scriptsize The University of Melbourne}\\
+{\scriptsize Melbourne, Victoria. Australia}\\
+{\scriptsize miquel.ramirez@nimelb.edu.au}}
 
@@
 \maketitle
-
 \begin{abstract}
@@
 
-%% The file named.bst is a bibliography style file for BibTeX 0.99c
-\bibliographystyle{named}
+\subsection*{Acknowledgments}
+This work is supported by the Spanish MINECO project TIN2017-88476-C2-1-R.
+D.
+Aineto is partially supported by the {\it FPU16/03184} and S.
+Jim\'enez by the {\it RYC15/18009}.
+M.
+Ram\'irez research is partially funded by DST Group Joint \& Operations Analysis Division.
+
 \bibliography{planlearnbibliography}
+\bibliographystyle{aaai}
 

### 2019-05-15 doc/aaai20-goalRecognition/goalRecognition.tex
updates added to the aaai20 paper on goal recognition

@@
 \begin{abstract}
-This paper shows how to relax a strong assumption of the {\em plan recognition as planning} approach that is {\em the observer knows the action model of the observed agents}.
-The paper introduces a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known) and it shows that this formulation neatly fits with the {\em plan recognition as planning} approach for {\em goal recognition}.
-The experimental results demonstrate that this novel formulation for classical planning allow us to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without knowing beforehand the action model of the observed agents.
+The {\em plan recognition as planning} approach assumes that {\em observers} must have both correct and complete knowledge of the action model of the observed agents.
+This paper shows that this assumption can be relaxed formulating a novel setup for classical planning where action models are unknown but the state variables, action parameters and a single plan observation are however known.
+The experimental results demonstrate that this novel classical planning setup allows us to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without knowing beforehand the precise action model of the observed agents.
 \end{abstract}
@@
 \label{sec:introduction}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal and classification examples are observations of agents pursuing one of those goals.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal for the observed agents and classification examples are observations of the agents pursuing one of those goals.
 While there exist diverse approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2010probabilistic} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
 
-{\em Plan recognition as planning} leverages the action model of the observed agents and an off-the-shelf classical planner to estimate the most likely goal of the agents being observed~\cite{ramirez2012plan}.
+{\em Plan recognition as planning} leverages the action model of the observed agents, a single plan observation, and an off-the-shelf classical planner to estimate the most likely goal of the observed agents~\cite{ramirez2012plan}.
 In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which can become a too strong requirement when applying {\em plan recognition as planning} on real-world problems.
-We introduce a novel formulation for classical planning in a setting where no action model is given (instead, only the state variables and the action parameters are known beforehand) and we show that this formulation neatly fits into the {\em plan recognition as planning} approach.
-The experimental results demonstrate that this novel formulation allows to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without requiring having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents
+We formulate a novel set up for classical planning where no action model is given (instead, only the state variables, a single plan observation and the action parameters are known beforehand) and we show that this formulation neatly fits into the {\em plan recognition as planning} approach.
+The experimental results demonstrate that we can solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without requiring having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents
 
@@
 \subsection{Classical planning with conditional effects}
-Let $F$ be the set of  propositional state variables ({\em fluents}) describing a state.
-A {\em literal} $l$ is a valuation of a fluent $f\in F$; i.e.
-either~$l=f$ or $l=\neg f$.
-A set of literals $L$ represents a partial assignment of values to fluents (without loss of generality, we will assume that $L$ does not contain conflicting values).
-Given $L$, let $\neg L=\{\neg l:l\in L\}$ be its complement.
+We denote as $F$ ({\em fluents}) the set of  propositional state variables.
+A partial assignment of values to fluents is represented by $L$ ({\em literals}).
+We adopt the \emph{open world assumption} (what is not known to be true in a state is unknown) to implicitly represent the unobserved literals of a state.
+Hence, a state $s$ includes positive literals ($f$) and negative literals ($\neg f$) and it is defined as a full assignment of values to fluents; $|s|=|F|$.
 We use $\mathcal{L}(F)$ to denote the set of all literal sets on $F$; i.e.~all partial assignments of values to fluents.
-A {\em state} $s$ is a full assignment of values to fluents; $|s|=|F|$.
 
-A {\em classical planning action} $a\in A$ has: a precondition $\pre(a)\in\mathcal{L}(F)$, a set of effects $\eff(a)\in\mathcal{L}(F)$, and a positive action cost $cost(a)$.
-The semantics of actions $a\in A$ is specified with two functions: $\rho(s,a)$ denotes whether action $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denotes the {\em successor state} that results of applying action $a$ in a state $s$.
-Then, $\rho(s,a)={\tt\small True}$ whenever $\pre(a)\subseteq s$ (i.e.~if its precondition holds in $s$) otherwise $\rho(s,a)={\tt\small False}$.
-The result of executing an applicable action $a\in A$ in a state $s$ is a new state $\theta(s,a)=(s\setminus \neg\eff(a))\cup\eff(a)$.
-Subtracting the complement of $\eff(a)$ from $s$ ensures that $\theta(s,a)$ remains a well-defined state.
-The subset of action effects that assign a positive value to a state fluent is called {\em positive effects} and denoted by $\eff^+(a)\in \eff(a)$ while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects} of an action $a\in A$.
+A {\em planning action} $a$ comprises a set of preconditions $\pre(a)\in\mathcal{L}(F)$ and a set of effects $\eff(a)\in\mathcal{L}(F)$.
+The semantics of an action $a$ is specified with two functions: $\rho(s,a)$ denoting whether $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denoting the {\em successor state} that results from applying $a$ in a state $s$.
+Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$, i.e.~ the action preconditions hold in the given state.
+The result of executing an applicable action $a$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$, where $\neg\eff(a)$ is the complement of $\eff(a)$, which is subtracted from $s$ so as to ensure that $\theta(s,a)$ remains a well-defined state.
+The subset of effects of an action $a$ that assign a positive value to a fluent is called {\em positive effects} and denoted by $\eff^+(a)\in \eff(a)$ while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects}.
 
-A {\em classical planning problem} is a tuple $P=\tup{F,A,I,G}$, where $I$ is the initial state and $G\subseteq\mathcal{L}(F)$ is the set of goal conditions over the state variables.
-A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ being the {\em length} of $\pi$ and $cost(\pi)=\sum_{a\in\pi} cost(a)$ being its {\em cost}.
-The execution of $\pi$ on the initial state of $P$ induces a {\em trajectory} $\tau(\pi,P)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, then $\rho(s_{i-1},a_i)={\tt\small True}$ and $s_i=\theta(s_{i-1},a_i)$.
-A plan $\pi$ solves $P$ iff the induced {\em trajectory} $\tau(\pi,P)$ reaches a final state $G \subseteq s_n$, where all goal conditions are met.
-A solution plan is {\em optimal} iff its cost is minimal.
+Planning actions can also define {\em conditional effects}.
+Formally the conditional effects of an action $a_c$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
+The {\em triggered effects} resulting from the action application (conditional effects whose conditions hold in $s$) is defined as $\eff_c(s,a)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E$.
 
-We also define {\em actions with conditional effects} because they are useful to compactly formulate our approach for {\em goal recognition with unknown domain models}.
-An action $a_c\in A$ with conditional effects is a set of preconditions $\pre(a_c)\in\mathcal{L}(F)$ and a set of {\em conditional effects} $\cond(a_c)$.
-Each conditional effect $C\rhd E\in\cond(a_c)$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-An action $a_c$ is applicable in a state $s$ if $\rho(s,a_c)$ is true, and the result of applying action $a_c$ in state $s$ is $\theta(s,a_c)=\{s\setminus\neg\eff_c(s,a)\cup\eff_c(s,a)\}$ where $\eff_c(s,a)$ are the {\em triggered effects} resulting from the action application (conditional effects whose conditions hold in $s$):
-\[
-\eff_c(s,a)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E,
-\]
+A {\em planning problem} is a tuple $P=\tup{F,A,I,G}$, where $A$ is a set of actions, $I$ is the initial state and $G\in\mathcal{L}(F)$ is the set of goal conditions over the state variables.
+A {\em plan} $\pi$ is an action sequence $\pi=\tup{a_1, \ldots, a_n}$, with $|\pi|=n$ denoting its {\em plan length}.
+The execution of $\pi$ in $I$ induces a {\em trajectory} $\tau=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and, for each {\small $1\leq i\leq n$}, it holds $\rho(s_{i-1},a_i)$ and $s_i=\theta(s_{i-1},a_i)$.
+A plan $\pi$ solves $P$ iff the induced trajectory reaches a final state $s_n$ such that $G \subseteq s_n$.
 
 \subsection{The observation model}
-We assume a bijective monotone mapping between actions/states of trajectories and observations~\cite{ramirez2009plan}, thus also granting the inverse consistency relationship (the trajectory is a superset of the observation).
-This means that transiting between two consecutive observed states may require the execution of more than a single action.
-In other words, having an input observation does not imply knowing the actual length of the corresponding plan.
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ that solves $P$, and the corresponding trajectory $\tau$ induced by the execution of $\pi$ in $I$, $\tau=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$; there exist as many observations of $\tau$ as combinations of observable actions and observable fluents of the states of $\tau$.
+We refer to the set of all possible combinations of observable elements of $\tau$ as $Obs(\tau)$.
 
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ and a trajectory $\tau(\pi,P)$, we define an interleaved combination of actions and states that represents the {\em observation from the execution of $\pi$ in $P$}.
-Formally, $\mathcal{O}(\tau)=\tup{s_0^o,a_1^o,s_1^o \ldots , a_l^o, s_m^o}$, $s_0^o=I$, and:
-\begin{itemize}
-\item The {\bf observed actions} are consistent with $\pi$, which means that $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi$.
-The number of observed actions, $l$, ranges from $0$ (fully unobserved action sequence) to $|\pi|$ (fully observed action sequence).
-\item The {\bf observed states} $\tup{s_0^o, s_1^o, \ldots, s_m^o}$ is a sequence of possibly {\em partially observed states}, except for the initial state $s_0^o$, which is fully observed.
-A partially observable state $s_i^o$ is one in which $|s_i^o| < |F|$; i.e., a state in which at least a fluent of $F$ is not observable.
-Note that this definition also comprises the case $|s_i^o| = 0$, when the state is fully unobservable.
-Whatever the sequence of observed states of $\mathcal{O}(\tau)$ is, it must be consistent with the sequence of states of $\tau(\pi,P)$, meaning that $\forall i, s_i^o \subseteq s_i$.
-The number of observed states, $m$, range from 1 (the initial state, at least), to $|\pi|+1$, and each {\em observed} states comprises $[1,|F|]$ fluents ($\mathcal{O}(\tau)$ can still miss intermediate states that are {\em unobserved}).
-\end{itemize}
+We define {\em an observation} $\mathcal{O}\in Obs(\tau)$ as a sequence of possibly {\em partially observable states}, $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$, except for the initial state $s_0^o=I$ which is fully observable.
+A {\em partially observable state} is one in which $|s_i^o| < |F|$, {\small $1\leq i\leq m\leq n$}; i.e., a state in which at least a fluent of $F$ is not observable.
+It may be also the case that $|s_i^o| = 0$ when an intermediate state is fully unobservable.
+
+The observation model can also include {\em observed actions} as fluents indicating the applied action in a given state.
+This means that a sequence of observed actions $\tup{a_1^o, \ldots, a_l^o}$ is a sub-sequence of $\pi=\tup{a_1, \ldots, a_n}$ such that $a_i^o \in s_{i-1}^o$, {\small $0\leq i \leq l$}.
+Consequently, the number of fluents that represent observed actions, $l$, can range from $0$ (in a fully unobservable action sequence) to $|\pi|=n$ (in a fully observed action sequence).
+
+Given $\mathcal{O} \in Obs(\tau)$, the number of observed states of $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$ ranges from 2 (at least the initial and final state, as explained above) to $|\pi|+1$.
+The number of fluents of the full observable state $s_0^o$ will be $|F|$, or $|F|+1$ in case the fluent of the applied action in $s_0$ is also observed.
+Every observable intermediate state will comprise a number of fluents between $[1,|F|+1]$, where a single fluent may represent a sensing fluent of the state or the observation of the applied action.
+
+This observation model can distinguish between {\em observable state variables}, whose value may be read from sensors, and {\em hidden} (or {\em latent}) {\em state variables}, that cannot be observed.
+Given a subset of fluents $\Gamma\subseteq F$ we say that $\mathcal{O}$ is a $\Gamma$-observation of the execution of $\pi$ on $P$ iff for every observed state $s_i^o$, ${\small 1\leq i\leq m}$, $s_i^o$ only contains fluents in $\Gamma$.
 
 \subsection{Goal recognition with classical planning}
-{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ and there is a single classification example, $\mathcal{O}(\tau)$, that represents an observation of agents acting to achieve a goal $G\in G[\cdot]$.
+{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ (where $G[\cdot]$ represents the full set of {\em recognizable} goals) and there is a single classification example $\mathcal{O}(\tau)$ (that represents the observation of a plan where agents act to achieve a goal $G\in G[\cdot]$).
 
-Following the paradigm of the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
+Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
 \begin{align}
@@
 The {\em plan recognition as planning} approach shows that the $P(\mathcal{O}|G)$ likelihood can be estimated leveraging the action model of the observed agents and an off-the-shelf classical planner~\cite{ramirez2012plan}.
-Given a {\em classical planning problem} $P=\tup{F,A,I,G[\cdot]}$ (where $G[\cdot]$ represents the set of {\em recognizable} goals) then $P(\mathcal{O}|G)$ is estimated computing, for each goal $G\in G[\cdot]$, the cost difference of the solution plans to these two classical planning problems:
+Given $P=\tup{F,A,I,G[\cdot]}$ then $P(\mathcal{O}|G)$ is estimated computing, for each goal $G\in G[\cdot]$, the cost difference of the solution plans to two classical planning problems:
 \begin{itemize}
-\item $P^{\top}_G$, the classical planning problem built constraining $P=\tup{F,A,I,G}$ to achieve the particular goal $G\in G[\cdot]$ through a plan $\pi^\top$ that is {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-\item $P^{\bot}_G$, the classical planning problem that constrains solutions of $P=\tup{F,A,I,G}$ to plans $\pi^\bot$, that achieve $G\in G[\cdot]$, but that are {\em inconsistent} with $\mathcal{O}(\tau)$.
+\item $P^{\top}_G$, the classical planning problem built constraining $P=\tup{F,A,I,G}$ to achieve the particular goal $G\in G[\cdot]$ through a plan $\pi^\top$ that is {\em consistent} with the input observation $\mathcal{O}$.
+\item $P^{\bot}_G$, the classical planning problem that constrains solutions of $P=\tup{F,A,I,G}$ to plans $\pi^\bot$, that achieve $G\in G[\cdot]$, but that are {\em inconsistent} with $\mathcal{O}$.
 \end{itemize}
@@
 The higher the value of the $\Delta(\pi^\top,\pi^\bot)$ cost difference, the higher the probability of the observed agents to aim goal $G\in G[\cdot]$.
-With this regard, {\em plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:
+{\em Plan recognition as planning} uses the {\em sigmoid function} to map the previous cost difference into a likelihood:
 
@@
 The larger the value of $\beta$, the more rational the agents, and the less likely that they will follow suboptimal plans.
-Recent work exploits the structure of action {\em preconditions} and {\em effects} to compute fast estimates of the $P(\mathcal{O}|G)$ likelihood~\cite{pereira2017landmark}.
+Recent work on {\em goal recognition} exploit the structure of action {\em preconditions} and {\em effects} to compute fast estimates of the $P(\mathcal{O}|G)$ likelihood~\cite{pereira2017landmark}.
+
+\section{Goal recognition as planning with unknown domain models}
+\label{sec:recognition}
+
+We define the task of {\em goal recognition with unknown domain models} as a $\tup{P,\mathcal{O}}$ pair, where:
+\begin{itemize}
+\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a classical planning problem where $G[\cdot]$ is the set of {\em recognizable} goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
+the functions $\rho$ and/or $\theta$ of $a$ are undefined).
+\item $\mathcal{O}$ is the observation of the execution of an unknown plan $\pi$ that reaches the goals $G\in G[\cdot]$ starting from the initial state $I$ in $P$.
+\end{itemize}
+
+The {\em solution} to the {\em goal recognition with unknown domain models} task is again the subset of goals in $G[\cdot]$ that maximizes expression (1).
+
+\subsection{Estimating the $P(\mathcal{O}|G)$ likelyhood with unknown domain models}
+To build an estimate of the $P(\mathcal{O}|G)$ likelyhood our mechanism matches the {\em plan recognition as planning} approach~\cite{ramirez2012plan} except that we compute $cost(\pi^\top)$ using our compilation for {\em classical planning with unknown domain models}.
+
+In more detail, we build the estimate of the $P(\mathcal{O}|G)$ likelyhood following these four steps:
+\begin{enumerate}
+\item {\em Build $P^{\top}_G$}, the classical planning problem that constrains solutions of the problem $P=\tup{F,A[\cdot],s_0^o,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
+Note that $s_0^o\in \mathcal{O}(\tau)$ is the initial state in the given observation.
+\item {\em Solve $P^{\top}_G$}, using the proposed compilation for {\em classical planning with unknown domain models}.
+Extract from this solution (1), $cost(\pi^\top)$ (by counting the number of $\mathsf{apply_{\xi,\omega}}$ actions in the solution) but also (2), the action model $A$ that is determined by the {\em insert} actions used in $\pi^\top$ to achieve the goals $G$.
+\item {\em Build $P^{\bot}_G$}, the classical planning problem that constrains $P=\tup{F,A,s_0^o,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ (where $A$ is the set of actions extracted in step 2.).
+\item {\em Solve $P^{\bot}_G$} with a classical planner and extract $cost(\pi_\bot)$ as the length of the found solution plan.
+\item Compute the $\Delta(\pi^\top,\pi^\bot)$ cost difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
+\end{enumerate}
+
+To compute the target probability distribution $P(G|\mathcal{O})$ plug the $P(\mathcal{O}|G)$ likelihoods into the {\em Bayes rule} from which the goal posterior probabilities are obtained.
+In this case the $P(\mathcal{O})$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
 
@@
 \label{sec:planning}
-This section introduces a novel formulation for classical planning in a setting where no action model is given.
-This setting has already shown related to the learning of action models for planning~\cite{SternJ17} and it can be seen as an extreme scenario when the action model is {\em learned} from a single example that contains only two state observations: the initial state and the goals.
+This section formulates {\em planning with unknown domain models}, a novel setup for classical planning where no action model is given.
+The state variables, action parameters and a single plan observation are known though.
+This setup is closely related to the learning of planning action models~\cite{SternJ17} and can be seen as an extreme learning task where action models must be {\em learned} from a single example that contain only two state observations: the initial state and the goals.
 
-A {\em classical planning with unknown domain models} is a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
+A {\em classical planning with unknown domain models} is defined as a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-A {\em solution plan} is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution induces a trajectory $\tau(\pi,I)=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
-one possible definition of the $\rho$ and $\theta$ functions within the given state variables) that satisfies $\rho(s_{i-1},a_i)={\tt\small True}$ and $s_i=\theta(s_{i-1},a_i)$, for every {\small $1\leq i\leq n$}, and such that the reached final state $s_n$ meets the goal condition $G \subseteq s_n$.
+A {\em solution plan} is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution on $I$ induces a trajectory $\tau=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
+one possible definition of the $\rho$ and $\theta$ functions within the given state variables) satisfying:
+\begin{itemize}
+\item For every {\small $1\leq i\leq n$} then $\rho(s_{i-1},a_i)={\tt\small True}$ .
+\item For every {\small $1\leq i\leq n$} then $s_i=\theta(s_{i-1},a_i)$.
+\item The reached final state $s_n$ meets the goals $G \subseteq s_n$. 
+\end{itemize}
 
@@
 
-\subsection{A propositional encoding for the space of STRIPS action models}
-{\em A \strips\ action schema} $\xi$ is defined by: A list of {\em parameters} $pars(\xi)$, and three sets of predicates (namely $pre(\xi)$, $del(\xi)$ and $add(\xi)$) that shape the kind of fluents that can appear in the {\em preconditions}, {\em negative effects} and {\em positive effects} of the actions induced from that schema.
-Let be $\Psi$ the set of {\em predicates} that shape the propositional state variables $F$, and a list of {\em parameters}, $pars(\xi)$.
-The set of elements that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of the \strips\ action schema $\xi$ is the set of FOL interpretations of $\Psi$ over the parameters $pars(\xi)$, and is denoted as ${\mathcal I}_{\Psi,\xi}$.
+\subsection{A propositional encoding for STRIPS action}
+A \strips\ \emph{action model} is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
 
-For instance in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\Psi,\xi}$ set contains only five elements for the case of the {\small \tt pickup($v_1$)} schemata, ${\mathcal I}_{\Psi,pickup}$={\small\tt\{handempty, holding($v_1$), clear($v_1$), ontable($v_1$), on($v_1,v_1$)\}} while it contains eleven elements for the {\small \tt stack($v_1$,$v_2$)} schemata, ${\mathcal I}_{\Psi,stack}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$), clear($v_2$), ontable($v_1$), ontable($v_2$), on($v_1,v_1$), on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+Let $\Psi$ be the set of {\em predicates} that shape the fluents $F$ (the initial state of an observation is a full assignment of values to fluents, $|s_0^o|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0^o$).
+The set of propositions that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
+For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+Hence, solving a $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task is determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of the corresponding action model.
 
-Despite any element of ${\mathcal I}_{\Psi,\xi}$ can {\em a priori} appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of schema $\xi$, in practice the actual space of possible \strips\ schemata is bounded by constraints:
-\begin{enumerate}
-\item {\bf Syntactic constraints}.
-\strips\ constraints require $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-Considering exclusively these syntactic constraints, the size of the space of possible \strips\ schemata is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$.
-{\em Typing constraints} are also of this kind~\cite{mcdermott1998pddl}.
-\item {\bf Observation constraints}.
-The observation of the actions and states resulting from the execution of a plan depicts {\em semantic knowledge} that constraints further the space of possible action schemata.
-\end{enumerate}
-
-In this work we introduce a novel propositional encoding of the {\em preconditions}, {\em negative}, and {\em positive} effects of a \strips\ action schema $\xi$ that uses only fluents of two kinds {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} (where $e\in{\mathcal I}_{\Psi,\xi}$).
-This encoding exploits the syntactic constraints of \strips\, so it is more compact than previous encodings proposed by~\citeauthor{aineto2018learning}~\citeyear{aineto2018learning} for learning \strips\ actions with classical planning.
-In more detail, if {\tt\small pre\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em precondition} in $\xi$.
-If {\tt\small pre\_e\_$\xi$} and {\tt\small eff\_e\_$\xi$} holds it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em negative effect} in $\xi$ while if $pre\_e\_\xi$ does not hold but {\tt\small eff\_e\_$\xi$} holds, it means that $e\in{\mathcal I}_{\Psi,\xi}$ is a {\em positive effect} in $\xi$.
-To illustrate this, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\em blocksworld} {\tt\small stack(?v1,?v2)} schema and our propositional representation for this schema using only {\tt\small pre\_e\_stack} and {\tt\small eff\_e\_stack} fluents ($e\in{\mathcal I}_{\Psi,stack}$).
+A valid actoin model $\xi$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+{\em Typing constraints} are also a type of syntactic constraint that reduce the size of ${\mathcal I}_{\xi,\Psi}$~\cite{mcdermott1998pddl}.
+Considering only these syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
+Given $p \in \mathcal{I}_{\Psi,\xi}$, the belonging of $p$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a propositional encoding that uses fluents of two types, $pre_{p,\xi}$ and $eff_{p,\xi}$.
+The four possible combinations of these two fluents are summarized in Figure \ref{fig:combinations}.
 
 \begin{figure}
-  \begin{tiny}  
+	\begin{scriptsize}
+		\begin{tabular}{c@{\hskip .2in} |@{\hskip .1in} c}
+	{\bf Encoding} & {\bf Meaning}\\\hline
+$\neg pre_{p,\xi} \wedge \neg eff_{p,\xi} $& $p$ belongs neither to the preconditions nor effects of $\xi$ \\
+             & ($p \notin pre(\xi) \wedge p \notin add(\xi) \wedge p \notin del(\xi)$)\\\\
+$pre_{p,\xi} \wedge \neg eff_{p,\xi} $& $p$ is only a precondition of $\xi$\\
+               &  ($p \in pre(\xi) \wedge p \notin add(\xi) \wedge p \notin del(\xi)$) \\\\
+$\neg pre_{p,\xi} \wedge eff_{p,\xi} $& $p$ is a positive effect of $\xi$ \\
+               &  ($p \notin pre(\xi) \wedge p \in add(\xi) \wedge p \notin del(\xi)$) \\\\
+$pre_{p,\xi} \wedge eff_{p,\xi} $& $p$ is a negative effect of $\xi$ \\
+               &  ($p \in pre(\xi) \wedge p \notin add(\xi) \wedge p \in del(\xi)$)
+		\end{tabular}
+	\end{scriptsize}
+	\caption{\small Combinations of the propositional encoding and their meaning}
+	\label{fig:combinations}
+\end{figure}
+
+To illustrate better this encoding, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with $pre_{p,stack}$ and $eff_{p,stack}$ fluents ($p\in{\mathcal I}_{\Psi,stack}$).
+
+\begin{figure}
+  \begin{scriptsize}
   \begin{verbatim}
@@
   \end{verbatim}
-  \end{tiny}  
+  \end{scriptsize}
  \caption{\small PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema.}
@@
 \end{figure}
-
-{\em Domain-specific knowledge} is also helpfull to constrain further the space of possible schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-With this regard, {\it state invariants} can be exploited either as {\em syntactic} constraints (to reduce the space of possible action models) but also as {\em semantic} constraints (to complete partial observations of the states traversed by a plan)~\cite{fox:TIM:JAIR1998}.
 
@@
 
-\subsection{The bias of the initially {\em empty} action model}
-Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $P=\tup{F,A[\cdot],I,G}$ problems preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
-In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P'=\tup{F',A',I,G}$ compilation so classical planners tend to solve $P'$ with plans that require a shorter number of {\em insert} actions.
-
-This bias could be eliminated defining a cost function for the actions in $A'$ (e.g.
-{\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}).
-In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} with zero-cost actions.
-Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
-Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
-
-Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects.
-These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers, the $\mathsf{apply_{\xi,\omega}}$ actions).
-Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
-
-The SAT-based planning approach is also convenient for the task of {\em goal recognition as planning with unknown domain models} because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
-
 \subsection{Compilation properties}
@@
 
-\section{Goal recognition as planning with unknown domain models}
-\label{sec:recognition}
+\subsection{The bias of the initially {\em empty} action model}
+Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $P=\tup{F,A[\cdot],I,G}$ problems preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
+In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P'=\tup{F',A',I,G}$ compilation so classical planners tend to solve $P'$ with plans that require a shorter number of {\em insert} actions.
 
-We define the task of {\em goal recognition with unknown domain models} as a $\tup{P,\mathcal{O}(\tau)}$ pair, where:
-\begin{itemize}
-\item $P=\tup{F,A[\cdot],I,G[\cdot]}$ is a classical planning problem where $G[\cdot]$ is the set of {\em recognizable} goals and $A[\cdot]$ is a set of actions s.t., for each $a\in A[\cdot]$, the semantics of $a$ is unknown (i.e.
-the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-\item $\mathcal{O}(\tau)$ is an observation of a trajectory $\tau(\pi,P)$ produced by the execution of an unknown plan $\pi$ that reaches the goals $G\in G[\cdot]$ starting from the initial state $I$ in $P$.
-\end{itemize}
+This bias could be eliminated defining a cost function for the actions in $A'$ (e.g.
+{\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}).
+In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} with zero-cost actions.
+Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
+Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
 
-The {\em solution} to the {\em goal recognition with unknown domain models} task is again the subset of goals in $G[\cdot]$ that maximizes expression (1).
+Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects.
+These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers, the $\mathsf{apply_{\xi,\omega}}$ actions).
+Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
 
-\subsection{Estimating the $P(\mathcal{O}|G)$ likelyhood with unknown domain models}
-Now we are ready to build an estimate of the $P(\mathcal{O}|G)$ likelyhood.
-Our mechanism matches the {\em plan recognition as planning} approach~\cite{ramirez2012plan} except that we compute $cost(\pi^\top)$ using our compilation for {\em classical planning with unknown domain models}.
+The SAT-based planning approach is also convenient for the task of {\em goal recognition as planning with unknown domain models} because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
 
-In more detail, we build the estimate of the $P(\mathcal{O}|G)$ likelyhood following these four steps:
-\begin{enumerate}
-\item {\em Build $P^{\top}_G$}, the classical planning problem that constrains solutions of the problem $P=\tup{F,A[\cdot],s_0^o,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-Note that $s_0^o\in \mathcal{O}(\tau)$ is the initial state in the given observation.
-\item {\em Solve $P^{\top}_G$}, using the proposed compilation for {\em classical planning with unknown domain models}.
-Extract from this solution (1), $cost(\pi^\top)$ (by counting the number of $\mathsf{apply_{\xi,\omega}}$ actions in the solution) but also (2), the action model $A$ that is determined by the {\em insert} actions used in $\pi^\top$ to achieve the goals $G$.
-\item {\em Build $P^{\bot}_G$}, the classical planning problem that constrains $P=\tup{F,A,s_0^o,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ (where $A$ is the set of actions extracted in step 2.).
-\item {\em Solve $P^{\bot}_G$} with a classical planner and extract $cost(\pi_\bot)$ as the length of the found solution plan.
-\item Compute the $\Delta(\pi^\top,\pi^\bot)$ cost difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
-\end{enumerate}
+\subsection{Constraining the action model space}
+In some contexts it is however reasonable to assume that some portions of the action model are known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
+Our compilation approach is also flexible to this particular scenario.
+The known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to true in the initial state.
+Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, making the classical planning task $P_{\Lambda}$ easier to be solved.
 
-To compute the target probability distribution $P(G|\mathcal{O})$ plug the $P(\mathcal{O}|G)$ likelihoods into the {\em Bayes rule} from which the goal posterior probabilities are obtained.
-In this case the $P(\mathcal{O})$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
+For example, suppose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state $I$ is extended with literals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the associated actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
 
-\subsection{Extending the observation model of {\em plan recognition as planning}}
-The work on {\em plan recognition as planning} usually assumes an observation model that is referred only to logs of executed actions.
-However, the approach applies also to more expressive observation models that consider state observations as well, like the observation model defined above, with a simple two-fold extension:
-\begin{itemize}
-\item One fluent $\{validated_j\}_{0\leq j\leq m}$ to point at every $s_j^o\in\mathcal{O}(\tau)$ state observation and $validated_m$ is added to every possible goal $G\in G[\cdot]$ to constrain solution plans $\pi^\top$ to be consistent with all the state observations.
-\item One $\mathsf{validate_{j}}$ action to constraint $\pi^\top$ to be consistent with the $s_j^o\in\mathcal{O}(\tau)$ input state observation, {\small $(1\leq j\leq m)$}.  
-\end{itemize}
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{validated_{j-1}\},\\
-\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg validated_{j-1}, validated_j\}.
-\end{align*}
-\end{small}
+{\em Domain-specific knowledge} is also helpfull to constrain further the space of possible action schemata.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
+With this regard, {\it state invariants} can be exploited either as {\em syntactic} constraints (to reduce the space of possible action models) but also as {\em semantic} constraints (to complete partial observations of the states traversed by a plan)~\cite{fox:TIM:JAIR1998}.
 
@@
 The problem of {\em classical planning with unknown domain models} has been previously addressed~\cite{SternJ17}.
-In this work we evidence the relevance of this task for addressing {\em goal recognition} when the action model of the observed agent is not available (which it is typically a too strong assumption at many real-world applications).
+In this work we evidence the relevance of this task for addressing {\em goal recognition} when the action model of the observed agent is not available.
 
@@
 \label{sec:conclusions}
-In some contexts it is however reasonable to assume that the action model is not learned from scratch, e.g.
-because some parts of the action model are known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
-Our compilation approach is also flexible to this particular learning scenario.
-The known preconditions and effects are encoded setting the corresponding fluents $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ to true in the initial state.
-Further, the corresponding insert actions, $\mathsf{insertPre_{p,\xi}}$ and $\mathsf{insertEff_{p,\xi}}$, become unnecessary and are removed from $A_{\Lambda}$, making the classical planning task $P_{\Lambda}$ easier to be solved.
-For example, suppose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state $I$ is extended with literals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the associated actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
 

### 2019-05-15 doc/aaai20-goalRecognition/goalRecognition.tex
updates added to the aaai20 paper on goal recognition

@@
 The {\em plan recognition as planning} approach assumes that {\em observers} must have both correct and complete knowledge of the action model of the observed agents.
-This paper shows that this assumption can be relaxed formulating a novel setup for classical planning where action models are unknown but the state variables, action parameters and a single plan observation are however known.
+This paper shows that this assumption can be relaxed formulating a novel setup for classical planning where action models are unknown but the state variables and the action parameters are however known.
 The experimental results demonstrate that this novel classical planning setup allows us to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without knowing beforehand the precise action model of the observed agents.
@@
 In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which can become a too strong requirement when applying {\em plan recognition as planning} on real-world problems.
-We formulate a novel set up for classical planning where no action model is given (instead, only the state variables, a single plan observation and the action parameters are known beforehand) and we show that this formulation neatly fits into the {\em plan recognition as planning} approach.
+We formulate a novel set up for classical planning where no action model is given (instead, only the state variables and the action parameters are known beforehand) and we show that this formulation neatly fits into the {\em plan recognition as planning} approach.
 The experimental results demonstrate that we can solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without requiring having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents
@@
 
-\subsection{Goal recognition with classical planning}
-{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ (where $G[\cdot]$ represents the full set of {\em recognizable} goals) and there is a single classification example $\mathcal{O}(\tau)$ (that represents the observation of a plan where agents act to achieve a goal $G\in G[\cdot]$).
+\subsection{Goal recognition as planning}
+{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ (where $G[\cdot]$ represents the set of {\em recognizable} goals) and there is a single classification example $\mathcal{O}(\tau)$ (that represents the observation of a plan where agents act to achieve a goal $G\in G[\cdot]$).
 
@@
 
-The {\em plan recognition as planning} approach shows that the $P(\mathcal{O}|G)$ likelihood can be estimated leveraging the action model of the observed agents and an off-the-shelf classical planner~\cite{ramirez2012plan}.
+The {\em plan recognition as planning} approach for goal recognition shows that the $P(\mathcal{O}|G)$ likelihood can be estimated leveraging the action model of the observed agents and an off-the-shelf classical planner~\cite{ramirez2012plan}.
 Given $P=\tup{F,A,I,G[\cdot]}$ then $P(\mathcal{O}|G)$ is estimated computing, for each goal $G\in G[\cdot]$, the cost difference of the solution plans to two classical planning problems:
@@
 
+To compute the target probability distribution $P(G|\mathcal{O})$ plug the $P(\mathcal{O}|G)$ likelihoods into the {\em Bayes rule} from which the goal posterior probabilities are obtained.
+In this case the $P(\mathcal{O})$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
+
 \section{Goal recognition as planning with unknown domain models}
@@
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
-\item $\mathcal{O}$ is the observation of the execution of an unknown plan $\pi$ that reaches the goals $G\in G[\cdot]$ starting from the initial state $I$ in $P$.
+\item $\mathcal{O}$ is the observation of the execution of an unknown plan $\pi$ to reach goal $G\in G[\cdot]$ starting from the given initial state $I=s_0^o$.
 \end{itemize}
@@
 The {\em solution} to the {\em goal recognition with unknown domain models} task is again the subset of goals in $G[\cdot]$ that maximizes expression (1).
+With this regard this new task is defined exactly as the original goal recognition task~\cite{ramirez2012plan} except that the preconditions and effects of the input set of actions are now unknown.
 
-\subsection{Estimating the $P(\mathcal{O}|G)$ likelyhood with unknown domain models}
-To build an estimate of the $P(\mathcal{O}|G)$ likelyhood our mechanism matches the {\em plan recognition as planning} approach~\cite{ramirez2012plan} except that we compute $cost(\pi^\top)$ using our compilation for {\em classical planning with unknown domain models}.
+\subsection{Estimating $P(\mathcal{O}|G)$ with unknown action models}
+Our approach to estimate the $P(\mathcal{O}|G)$ likelyhood is to leverage the available input knowledge to build a reasonable model for the actions in $A[\cdot]$, use such model to compute solution plans to the classical planning problems $P^{\top}_G$ and $P^{\bot}_G$ and, estimate $P(\mathcal{O}|G)$ as in the {\em plan recognition as planning} approach for goal recognition.
 
-In more detail, we build the estimate of the $P(\mathcal{O}|G)$ likelyhood following these four steps:
+In more detail, given $P$ and $\mathcal{O}$, we compute the estimate of the $P(\mathcal{O}|G)$ likelyhood following these steps:
 \begin{enumerate}
-\item {\em Build $P^{\top}_G$}, the classical planning problem that constrains solutions of the problem $P=\tup{F,A[\cdot],s_0^o,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}(\tau)$.
-Note that $s_0^o\in \mathcal{O}(\tau)$ is the initial state in the given observation.
-\item {\em Solve $P^{\top}_G$}, using the proposed compilation for {\em classical planning with unknown domain models}.
-Extract from this solution (1), $cost(\pi^\top)$ (by counting the number of $\mathsf{apply_{\xi,\omega}}$ actions in the solution) but also (2), the action model $A$ that is determined by the {\em insert} actions used in $\pi^\top$ to achieve the goals $G$.
-\item {\em Build $P^{\bot}_G$}, the classical planning problem that constrains $P=\tup{F,A,s_0^o,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}(\tau)$ (where $A$ is the set of actions extracted in step 2.).
-\item {\em Solve $P^{\bot}_G$} with a classical planner and extract $cost(\pi_\bot)$ as the length of the found solution plan.
+\item Compute an action model $A$ that is able to solve {\em $P^{\top}_G$}, the classical planning problem that constrains solutions of $\tup{F,A[\cdot],I,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}$.
+\item Replace $A[\cdot]$ with $A$, {\em solve $P^{\top}_G$} and take $cost(\pi^\top)$ from that solution.
+\item Replace $A[\cdot]$ with $A$, {\em solve $P^{\bot}_G$} (i.e., the classical planning problem that constrains $\tup{F,A,I,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}$) and take $cost(\pi^\top)$ from that solution.
 \item Compute the $\Delta(\pi^\top,\pi^\bot)$ cost difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
@@
 
-To compute the target probability distribution $P(G|\mathcal{O})$ plug the $P(\mathcal{O}|G)$ likelihoods into the {\em Bayes rule} from which the goal posterior probabilities are obtained.
-In this case the $P(\mathcal{O})$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
-
-\section{Planning with unknown domain models}
-\label{sec:planning}
+\subsection{Planning with unknown domain models}
 This section formulates {\em planning with unknown domain models}, a novel setup for classical planning where no action model is given.
@@
 This setup is closely related to the learning of planning action models~\cite{SternJ17} and can be seen as an extreme learning task where action models must be {\em learned} from a single example that contain only two state observations: the initial state and the goals.
+This planning setup is of interest because it allows to formulate steps 1.
+and 2.
+as a single classical planning problem.
 
@@
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
+
 A {\em solution plan} is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution on $I$ induces a trajectory $\tau=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
@@
 \item For every {\small $1\leq i\leq n$} then $s_i=\theta(s_{i-1},a_i)$.
-\item The reached final state $s_n$ meets the goals $G \subseteq s_n$. 
+\item Goals are met at the final state $G\subseteq s_n$. 
 \end{itemize}
 
-Next we show that a propositional encoding for the space of STRIPS action models allows us to solve $P=\tup{F,A,I,G[\cdot]}$ problems with off-the-shelf classical planners.
+We can define further constraints to reduce the hypothesis space for the functions $\rho(s,a)$ and $\theta(s,a)$ corresponding to the $A[\cdots]$ actions:
+\begin{itemize}
+\item {\em Observations}.
+The action model must be consistent with the input observation.
+Specifically, the states induced by plans computable with such model must comprise the observed states of the sample, which further constrains the space of possible action models.
+The observation $\mathcal{O}$ can also be regarded as a sequence of ordered {\em landmarks} for the planning problem $P_\mathcal{O}$~\cite{hoffmann2004ordered} since all the fluents of the sets in $\mathcal{O}$ must be achieved by any plan that solves $P_\mathcal{O}$ and in the same order as defined in the observation $\mathcal{O}$.
 
-\subsection{A propositional encoding for STRIPS action}
+\item {\em Domain-specific knowledge} is also helpfull to constrain further the space of possible action schemata.
+For instance, in the {\em blocksworld} one can argue that {\small\tt on(?x,?x)} will not appear in the preconditions/effects of an action because, in this specific domain, a block cannot be on top of itself.
+\item {\em Partially-specified action models}.
+In some contexts it is however reasonable to assume that some portions of the action model are known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
+\end{itemize}
+
+\section{Planning with unknown \strips\ models}
+\label{sec:planning}
+This section shows that, when the unknown precondition and effects of the actions in $A[\cdot]$ follow the \strips\ model, we can solve the task of {\em planning with unknown domain models} (and hence estimate the $P(\mathcal{O}|G)$ likelyhood) with an off-the-shelf classical planner.
+
+\subsection{The space of possible action models}
 A \strips\ \emph{action model} is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
 
-Let $\Psi$ be the set of {\em predicates} that shape the fluents $F$ (the initial state of an observation is a full assignment of values to fluents, $|s_0^o|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0^o$).
+Let $\Psi$ be the set of {\em predicates} that shape the fluents $F$ (the initial state is a full assignment of values to fluents so the predicates $\Psi$ are extractable from $I$).
 The set of propositions that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
 For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-Hence, solving a $\Lambda=\tup{\mathcal{M},{\mathcal O}}$ learning task is determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of the corresponding action model.
 
-A valid actoin model $\xi$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+\subsection{Syntactic constrains}
+An action model $\xi$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
 {\em Typing constraints} are also a type of syntactic constraint that reduce the size of ${\mathcal I}_{\xi,\Psi}$~\cite{mcdermott1998pddl}.
@@
 \subsection{A classical planning compilation for planning with unknown domain models}
-Inspired by the {\em classical planning compilation for learning \strips\ action models}~\cite{aineto2018learning}, we define here a compilation to address the task of {\em planning with unknown domain models} that uses our compact propositional encoding of STRIPS action models.
+The approach for learning \strips\ action models presented in~\cite{aineto2018learning}, which we will use as our baseline learning system (hereafter BLS, for short), is a compilation scheme that transforms the problem of learning the preconditions and effects of action models into a planning task $P'$.
+A \strips\ \emph{action model} $\xi$ is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
 
-Given a classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$ we create a classical planning problem $P'=\tup{F',A',I,G}$ such that:
-\begin{itemize}
-\item $F'$ extends $F$ with a fluent $mode_{insert}$, to indicate whether action models are being {\em programmed} or {\em applied}, and the $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents for the propositional encoding of the corresponding space of STRIPS action models.
+The BLS receives as input an empty domain model, which only contains the headers of the action models, and a set of observations of plan executions, and creates a propositional encoding of the planning task $P'$.
+Let $\Psi$ be the set of {\em predicates}\footnote{The initial state of an observation is a full assignment of values to fluents, $|s_0|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0$.} that shape the variables $F$.
+The set of propositions of $P'$ that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
+For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), \\
+on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi$.
 
-\item $A'$ replaces the actions in $A$ with two types of actions.
-\begin{enumerate}
-\item Actions for {\em inserting} a {\em precondition}, {\em positive}/{\em negative} effect into the action model $\xi$ following the syntactic constraints of \strips\ . 
-\begin{itemize}
-\item Actions to insert an $e\in{\mathcal I}_{\Psi,\xi}$ {\em precondition} into $\xi$. The precondition is only inserted when neither $pre\_e\_\xi$ nor $eff\_e\_\xi$ exist in $\xi$.
+The decision as to whether or not an element of ${\mathcal I}_{\xi,\Psi}$ will be part of $pre(\xi)$, $del(\xi)$ or $add(\xi)$ is given by the plan that solves $P'$.
+Specifically, two different sets of actions are included in the definition of $P'$: \emph{insert actions}, which insert preconditions and effects on an action model; and \emph{apply actions}, which validate the application of the learned action models in the input observations.
+Roughly speaking, in the \emph{blocksworld} domain, the insert actions of a plan that solves $P'$ will look like {\tt{\footnotesize(insert\_pre\_stack\_holding\_v1)}},\\
+{\tt{\footnotesize(insert\_eff\_stack\_clear\_v1),\\
+(insert\_eff\_stack\_clear\_v2)}}, where the second action denotes a positive effect and the third one a negative effect both to be inserted in the model of {\tt{\small stack}}; and the second set of actions of the plan that solves $P'$ will be like {\tt{\small (apply\_unstack blockB blockA),(validate\_1),(apply\_putdown blockB),(validate\_2)}}, where the {\tt {\small validate}} actions denote the points at which the states generated through the {\tt {\small apply}} actions must be validated with the observations of plan executions.
+
+In a nutshell, the output of the BLS compilation is a plan that completes the empty input domain model by specifying the preconditions and effects of each action model such that the validation of the completed model over the input observations is successful.
+
+\subsection{Observations}
+This require the compilation to include actions for {\em validating} partially observed states $s_j^o\in\mathcal{O}$.
+These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observation $\mathcal{O}$ follows after the execution of the apply actions.
 
@@
 \begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertPre_{p,\xi}})=&\{\neg pre\_e\_\xi, \neg eff\_e\_\xi, mode_{insert}\},\\
-\cond(\mathsf{insertPre_{p,\xi}})=&\{\emptyset\}\rhd\{pre\_e\_\xi\}.
+\hspace*{7pt}\pre(\mathsf{validate_{j}})=&s_j^o\cup\{test_{j-1}\},\\
+\cond(\mathsf{validate_{j}})=&\{\emptyset\}\rhd\{\neg test_{j-1}, test_j\}.
 \end{align*}
@@
 
-\item Actions to insert an $e\in{\mathcal I}_{\Psi,\xi}$ {\em effect} to the action model $\xi$.
+There will be a validate action in $\pi_\Lambda$ for every observed state in $\mathcal{O}$.
+The position of the validate actions in $\pi_\Lambda$ will be determined by the planner by checking that the state resulting after the execution of an apply action comprises the observed state $s_j^o\in\mathcal{O}$.
 
-\begin{small}
-\begin{align*}
-\hspace*{7pt}\pre(\mathsf{insertEff_{p,\xi}})=&\{\neg eff\_e\_\xi, mode_{insert}\},\\
-\cond(\mathsf{insertEff_{p,\xi}})=&\{\emptyset\}\rhd\{eff\_e\_\xi\}.
-\end{align*}
-\end{small}
-\end{itemize}
-
-\item Actions for {\em applying} an action model $\xi$ built by the {\em insert} actions and bounded to objects $\omega\subseteq\Omega^{|pars(\xi)|}$ (where $\Omega$ is the set of {\em objects} used to induce the fluents $F$ by assigning objects in $\Omega$ to the $\Psi$ predicates and $\Omega^k$ is the $k$-th Cartesian power of $\Omega$).
-Note that the action parameters, $pars(\xi)$, are bound to the objects in $\omega$ that appear in the same position.
 \end{enumerate}
@@
 
-\begin{small}
-\begin{align*}
-\pre(\mathsf{apply_{\xi,\omega}})=&\{pre\_e\_\xi\implies p(\omega)\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
-\cond(\mathsf{apply_{\xi,\omega}})=&\{pre\_e\_\xi\wedge eff\_e\_\xi\}\rhd\{\neg p(\omega)\}_{\forall e\in{\mathcal I}_{\Psi,\xi}},\\
-&\{\neg pre\_e\_\xi \wedge eff\_e\_\xi\}\rhd\{p(\omega)\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}\},\\
-&\{\emptyset\}\rhd\{\neg mode_{insert}\}.
-\end{align*}
-\end{small}
+\subsection{Domain-specific constraints}
 
-The intuition of the compilation is that the dynamics of the actions for {\em applying} an action model $\xi$ is determined by the values of the corresponding $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents in the current state.
-For instance when executing {\tt{\small (apply\_stack blockB blockA)}} in a state $s$, its preconditions and effects are activated according to the values of the corresponding $\{pre\_e\_stack, eff\_e\_stack\}_{\forall e\in{\mathcal I}_{\Psi,stack}}$ fluents in $s$.
-This means that if the current state $s$ holds $\{{\tt\scriptsize (pre\_stack\_holding\_v1),(pre\_stack\_clear\_v2)}\} \subset s$, then it must be checked that positive literals {\tt{\small (holding blockB)}} and {\tt{\small (clear blockA)}} hold in $s$.
-Otherwise, a different set of precondition literals will be checked for the stack action.
-The same applies to the effects of {\tt{\small (apply\_stack blockB blockA)}}.
-Executing {\tt{\small (apply\_stack blockB blockA)}}, will add the literals {\tt{\small (on blockB blockA),(clear blockB),(not(clear blockA)),(handempty)}} and {\tt{\small(not(clear blockB))}} to the successor state only if the $\{pre\_e\_stack, eff\_e\_stack\}_{\forall e\in{\mathcal I}_{\Psi,stack}}$ fluents has been correctly programmed by the corresponding {\em insert} actions.
+Our approach is to introduce {\em domain-specific knowledge} in the form of {\em state constraints} to further restrict the space of the action models.
+Back to the {\em blocksworld} domain, one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of any action model $\xi$ because, in this specific domain, a block cannot be on top of itself.
+The notion of state constraint is very general and has been used in different areas of AI and for different purposes.
+In planning, state constraints are compact and abstract representations that relate the values of variables in each state traversed by a plan, and allow to specify the set of states where a given action is applicable, the set of states where a given {\em axiom} or {\em derived predicate} holds or the set of states that are considered goal states~\cite{HaslumIR0TSN18}.
+
+{\em State invariants} is a useful type of state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+Given a planning problem $P=\tup{F,A,I,G}$, a state invariant is a formula $\phi$ that holds in $I$, $I\models \phi$, and in every state $s$ built out of $F$ that is reachable by applying actions of $A$ in $I$.
+
+A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
+
+Recently, some works point at extracting \emph{lifted} invariants, also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}, that hold for any possible state and any possible set of objects.
+Invariant
+templates obtained by inspecting the lifted representation of the domain have also been exploited for deriving \emph{lifted mutex}~\cite{BernardiniFS18}.
+In this work we exploit domain-specific knowledge that is given as {\em schematic mutex}.
+We pay special attention to {\em schematic mutex} because they identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) an effective completion of a partially observed state and (2) an effective pruning of inconsistent \strips\ action models.
+
+%A {\em domain invariant} is an instance-independent state invariant that holds for any possible initial state and any possible set of objects.
+Therefore, if $s\nvDash \phi$ such that $\phi$ is a domain invariant, it means that $s$ is not a valid state.
+Domain invariants are often compactly defined as {\em lifted invariants} (also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}).
+
+We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\xi,\Psi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are universally quantified.
+For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible state.
+Figure~\ref{fig:strongest-invariant} shows an example of four clauses that define schematic mutexes for the {\em blocksworld} domain.
 
 \begin{figure}[hbt!]
-	{\tiny\tt
-
-\begin{tabular}{ll}
-		{\bf 00}:(insert\_pre\_stack\_holding\_v1) & {\bf 10}:(insert\_eff\_pickup\_clear\_v1) \\
-		01:(insert\_pre\_stack\_clear\_v2) & 11:(insert\_eff\_pickup\_ontable\_v1)\\
-                {\bf 02}:(insert\_pre\_pickup\_handempty) & 12:(insert\_eff\_pickup\_handempty)\\
-                03:(insert\_pre\_pickup\_clear\_v1) & 13:(insert\_eff\_pickup\_holding\_v1)\\
-                04:(insert\_pre\_pickup\_ontable\_v1) & {\bf 14}:(apply\_pickup blockB)\\
-                {\bf 05}:(insert\_eff\_stack\_clear\_v1) & 15:(apply\_stack blockB blockC)\\
-                06:(insert\_eff\_stack\_clear\_v2) & 16:(apply\_pickup blockA)\\
-                07:(insert\_eff\_stack\_handempty) & 17:(apply\_stack blockA blockB) \\
-                08:(insert\_eff\_stack\_holding\_v1) & \\
-                09:(insert\_eff\_stack\_on\_v1\_v2) &             		 
-\end{tabular}
-}
-	\caption{\small Plan computed when solving the classical planning problem output by our compilation corresponding to a classical planning with unknown domain models.}
-	\label{fig:plan-lplan}
+  \begin{footnotesize}
+$\forall x_1,x_2\ \neg ontable(x_1)\vee\neg on(x_1,x_2)$.\\
+$\forall x_1,x_2\ \neg clear(x_1)\vee\neg on(x_2,x_1)$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
+\end{footnotesize}
+ \caption{\small {\em Schematic mutexes} for the {\em blocksworld} domain.}
+\label{fig:strongest-invariant}
 \end{figure}
 
-Figure~\ref{fig:plan-lplan} shows a solution plan computed when solving a $P'=\tup{F',A',I,G}$ classical planning problem output by our compilation.
-In the initial state of that problem three  blocks ({\small\tt blockA}, {\small\tt blockB} and {\small\tt blockC}) are clear and on top of the table, the robot hand is empty.
-The problem goal is having the three-block tower {\tt blockA} on top of {\tt blockB} and {\tt blockB} on top of {\tt blockC}.
-The plan shows the {\em insert} actions for the {\tt\small stack} scheme (steps $00-01$ insert the preconditions, steps $05-10$ insert the effects), steps $02-04$ insert the preconditions of the {\tt\small pickup} scheme (while steps $10-13$ insert the effects of this scheme).
-Finally, steps $14-17$ is the plan postfix that applies the programmed action model to achieve the goals $G$ starting from $I$.
-Note that another valid solution could be computed for instance, by inserting the same preconditions and effects but into the {\em unstack} and {\em putdown} actions and then applying instead the four step postfix {\tt\small (putdown blockB), (unstack blockB blockC), (putdown blockA), (unstack blockA blockB)}.
+{\it state invariants} can be exploited either as {\em syntactic} constraints (to reduce the space of possible action models) but also as {\em semantic} constraints (to complete partial observations of the states traversed by a plan)~\cite{fox:TIM:JAIR1998}.
 
-\subsection{Compilation properties}
-\begin{mylemma}
-Soundness. Any classical plan $\pi'$ that solves $P'$ produces a solution to the classical planning problem with unknown domain models $P=\tup{F,A[\cdot],I,G}$.
-\end{mylemma}
+\subsubsection{Completing partially observed states with {\em schematic mutexes}}
 
-\begin{proof}[Proof]
-\begin{small}
-Once a given precondition (or effect) is inserted into an action model it can never be removed back and once an action model is applied (via an $\mathsf{apply_{\xi,\omega}}$ action) it cannot longer be {\em programed}.
-The set of goals $G$ can only be achieved executing an applicable sequence of $\mathsf{apply_{\xi,\omega}}$ actions that, starting in the corresponding initial state reach a state $G \subseteq s_n$ (the fluents of the original problem $F$ can exclusively be modified by the $\mathsf{apply_{\xi,\omega}}$ actions).
-This means that the action model used by the $\mathsf{apply_{\xi,\omega}}$ actions has to be consistent with the traversed intermediate states.
-We know that this must be true by the definition of the $\mathsf{apply_{\xi,\omega}}$ so hence, the sub-sequence of $\mathsf{apply_{\xi,\omega}}$ appearing in $\pi'$ to solve $P'$ is a solution plan to $P=\tup{F,A[\cdot],I,G}$.
-\end{small}
-\end{proof}
+The addition of new literals to complete the partial states $\tup{s_1^o \ldots, s_m^o}$ of an observation $\mathcal{O}$  using a set of schematic mutexes $\Phi$ is done in a pre-processing stage.
 
-\begin{mylemma}
-Completeness. Any plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$ is computable solving the corresponding classical planning task $P'$.
-\end{mylemma}
+Let $\Omega$ be the set of objects that appear in $F$ as the values of the arguments of the predicates $\Psi$, and $\phi=\tup{p,q}$ a schematic mutex.
+There exist many possible instantiations of $\phi$ of the type $\tup{p(\omega),q(\omega')}$ with objects of $\Omega$, where $\omega\subseteq\Omega^{|args(p)|}$ and $\omega'\subseteq\Omega^{|args(q)|}$.
+Let us now assume that the instantiation $p(\omega) \in s_j^o$, {\small $(1\leq j\leq m)$}, being $s_j^o$ a partially observed state of $\mathcal{O}$.
+Then, two situations may occur: (a) $\neg q(\omega') \in s_j^o$, in which case the expression $\neg p(\omega) \vee \neg q(\omega')$ holds in $s_j^o$; or (b) $\neg q(\omega') \notin s_j^o$, in which case the literal has not been observed in $s_j^o$ and so we can safely complete the state with $\neg q(\omega')$ (the same applies inversely, when $q(\omega') \in s_j^o$ but $\neg p(\omega) \notin s_j^o$).
+In other words, if we find that one component of a schematic mutex is positively observed in a state and the other component is not observable in such state, we can complete the state with the missing negative literal.
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular state and $\Phi$ contains the schematic mutex $\neg holding(v_1)\vee\neg clear(v_1)$, we extend the state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal being initially unknown).
 
-\begin{proof}[Proof]
-\begin{small}
-By definition ${\mathcal I}_{\Psi,\xi}$ fully captures the set of elements that can appear in an action model $\xi$ using predicates $\Psi$.
-Furthermore, the compilation does not discard any possible action model definable within ${\mathcal I}_{\Psi,\xi}$.
-This means that for every plan $\pi$ that solves $P=\tup{F,A[\cdot],I,G}$, we can build a plan $\pi'$ by selecting the appropriate actions for inserting precondition and effects to the corresponding action model and then selecting the corresponding $\mathsf{apply_{\xi,\omega}}$ actions that transform the initial state $I$ into a state $G \subseteq s_n$.
-\end{small}
-\end{proof}
+\begin{figure}
+	\begin{footnotesize}
+		\begin{tabular}{lll}
+			{\bf ID} & {\bf Action} & {\bf New conditional effect}\\\hline
+			1&${\tt (insert\_pre)_{\xi,p}}$&$\{pre\_\xi\_q\}\rhd\{invalid\}$\\
+			2&${\tt (insert\_eff)_{\xi,p}}$&$\{pre\_\xi\_q\wedge eff\_\xi\_q\wedge pre\_\xi\_p\}\rhd\{invalid\}$\\
+			3&${\tt (insert\_eff)_{\xi,p}}$&$\{\neg pre\_\xi\_q\wedge eff\_\xi\_q\wedge \neg pre\_\xi\_p\}\rhd\{invalid\}$\\
+			4&${\tt (apply)_{\xi,\omega}}$&$\{\neg pre\_\xi\_p \wedge eff\_\xi\_p \wedge $\\
+			&&$q(\omega)\wedge \neg pre\_\xi\_q\}\rhd\{invalid\}$\\
+			5&${\tt (apply)_{\xi,\omega}}$&$\{\neg pre\_\xi\_p \wedge eff\_\xi\_p \wedge $\\
+			&&$q(\omega)\wedge \neg eff\_\xi\_q\}\rhd\{invalid\}$
+		\end{tabular}
+	\end{footnotesize}
+	\caption{\small Summary of the new conditional effects added to the classical planning compilation for the learning of \strips\ action models.}
+	\label{fig:ceffects}
+\end{figure}
 
-The size of the classical planning task $P'$ output by our compilation depends on the arity of the given {\em predicates} $\Psi$ and the number of parameters of the action models, $|pars(\xi)|$.
-The larger these arities, the larger $|{\mathcal I}_{\Psi,\xi}|$.
-This term dominates the compilation size because it defines the $\{pre\_e\_\xi, eff\_e\_\xi\}$ fluents, the corresponding set of {\em insert} actions, and the number of conditional effects of the $\mathsf{apply_{\xi,\omega}}$ actions.
+\subsubsection{Pruning inconsistent action models with {\em schematic mutexes}}
+We could extend the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} to check the consistency of the {\em state-constraints} in $\Phi$ at every state traversed by a solution to the compiled problem.
+Unfortunately, checking arbitrary $\phi$ formulae is too expensive for current classical planners.
 
-\subsection{The bias of the initially {\em empty} action model}
-Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $P=\tup{F,A[\cdot],I,G}$ problems preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
-In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P'=\tup{F',A',I,G}$ compilation so classical planners tend to solve $P'$ with plans that require a shorter number of {\em insert} actions.
+Instead, our approach is to define a mechanism to check {\em state-constraints} in the form of {\em schematic mutex}.
+To implement this checking mechanism we add new conditional effects to the {\em insert} and {\em apply} actions of the classical planning compilation.
+Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the compilation and next, we describe them in detail:
 
-This bias could be eliminated defining a cost function for the actions in $A'$ (e.g.
-{\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}).
-In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} with zero-cost actions.
-Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
-Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
+Our approach to learning action models consistent with the schematic mutexes in $\Phi$ is to ensure that newly generated states induced by the learned actions do not introduce any inconsistency.
+This is implemented by adding new conditional effects to the ${\tt \small insert}$ and ${\tt\small apply}$ actions of the BLS compilation.
+Figure~\ref{fig:ceffects} summarizes the new conditional effects added to the compilation and next, we describe them in detail:
 
-Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects.
-These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers, the $\mathsf{apply_{\xi,\omega}}$ actions).
-Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
+\begin{enumerate}
+	\item[1-3] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to ${\mathcal I}_{\xi,\Psi}$, one conditional effect is added to the ${\tt \small (insert\_pre)_{\xi,p}}$ actions to prevent the insertion of two preconditions that are schematic mutex.
+Likewise, two conditional effects are added to the ${\tt \small (insert\_eff)_{\xi,p}}$ actions, one to prevent the insertion of two positive effects that are schematic mutex and another one to prevent two mutex negative effects.
+	\item[4-5] For every schematic mutex $\tup{p,q}$, where both $p$ and $q$ belong to ${\mathcal I}_{\xi,\Psi}$, two conditional effects are added to the ${\tt \small (apply)_{\xi,\omega}}$ actions to prevent positive effects that are inconsistent with an input observation (in ${\tt \small (apply)_{\xi,\omega}}$ actions the variables in $pars(\xi)$ are bounded to the objects in $\omega$ that appear in the same position).
+\end{enumerate}
 
-The SAT-based planning approach is also convenient for the task of {\em goal recognition as planning with unknown domain models} because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
+In theory, conditional effects of the type $4$-$5$ are sufficient to guarantee that all the states traversed by a plan produced by the compilation are {\em consistent} with the input set of schematic mutexes $\Phi$ (obviously provided that the input initial state $s_0^o$ is a valid state).
+In practice we include also conditional effects of the type $1$-$3$ because they prune {\em invalid} action models at an earlier stage of the planning process (these effects extend the ${\tt \small insert}$ actions that always appear first in the solution plans).
 
-\subsection{Constraining the action model space}
-In some contexts it is however reasonable to assume that some portions of the action model are known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
+The goals of the planning task $P'$ generated by the original BLS compilation are extended with the $\neg invalid$ literal to validate that only states consistent with the state constraints defined in $\Phi$ are traversed by solution plans.
+Remarkably, the $\neg invalid$ literal allows us also to define ${\tt \small (apply)_{\xi,\omega}}$ actions more compactly than in the original compilation.
+Disjunctions are no longer required to code the possible preconditions of an action schema since they can now be encoded with conditional effects of the type $\{pre\_\xi\_p\wedge \neg p(\omega)\}\rhd\{invalid\}$.
+
+\subsection{Partially specified action models}
 Our compilation approach is also flexible to this particular scenario.
@@
 For example, suppose that the preconditions of the {\em blocksworld} action schema {\tt stack} are known, then the initial state $I$ is extended with literals, {\small\tt(pre\_holding\_v1\_stack)} and {\small\tt(pre\_clear\_v2\_stack)} and the associated actions $\mathsf{insertPre_{holding_v1,stack}}$ and $\mathsf{insertPre_{clear_v2,stack}}$ can be safely removed from the $A_{\Lambda}$ action set without altering the {\em soundness} and {\em completeness} of the $P_{\Lambda}$ compilation.
-
-{\em Domain-specific knowledge} is also helpfull to constrain further the space of possible action schemata.
-For instance, in the {\em blocksworld} one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ lists of an action schema $\xi$ because, in this specific domain, a block cannot be on top of itself.
-With this regard, {\it state invariants} can be exploited either as {\em syntactic} constraints (to reduce the space of possible action models) but also as {\em semantic} constraints (to complete partial observations of the states traversed by a plan)~\cite{fox:TIM:JAIR1998}.
 
@@
 \label{sec:conclusions}
+Classical planners tend to preffer shorter solution plans, so our compilation may introduce a bias to $P=\tup{F,A[\cdot],I,G}$ problems preferring solutions that are referred to action models with a shorter number of {\em preconditions}/{\em effects}.
+In more detail, all $\{pre\_e\_\xi, eff\_e\_\xi\}_{\forall e\in{\mathcal I}_{\Psi,\xi}}$ fluents are false at the initial state of our $P'=\tup{F',A',I,G}$ compilation so classical planners tend to solve $P'$ with plans that require a shorter number of {\em insert} actions.
+
+This bias could be eliminated defining a cost function for the actions in $A'$ (e.g.
+{\em insert} actions has {\em zero cost} while $\mathsf{apply_{\xi,\omega}}$ actions has a {\em positive constant cost}).
+In practice we use a different approach to disregard the cost of {\em insert} actions because classical planners are not proficiency optimizing {\em plan cost} with zero-cost actions.
+Instead, our approach is to use a SAT-based planner~\cite{rintanen2014madagascar} because it can apply all actions for inserting preconditions in a single planning step (these actions do not interact).
+Further, the actions for inserting action effects are also applied in a single planning step so the plan horizon for programming any action model is always bound to 2, which significantly reduces the planning horizon.
+
+Our compilation for {\em planning with unknown domain models} can then be understood as an extension of the SATPLAN approach for classical planning~\cite{kautz1992planning} with two additional initial layers: a first layer for inserting the action preconditions and a second one for inserting the action effects.
+These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers, the $\mathsf{apply_{\xi,\omega}}$ actions).
+Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
+
+The SAT-based planning approach is also convenient for the task of {\em goal recognition as planning with unknown domain models} because its ability to deal with classical planning problems populated with dead-ends and because symmetries in the insertion of preconditions/effects into an action model do not affect to the planning performance.
 

### 2019-05-16 doc/aaai20-goalRecognition/goalRecognition.tex
reshaping the aaai20 paper on goal recognition

@@
 
-\title{Goal Recognition as Planning with Unknown Domain Models}
+\title{Goal Recognition as Planning with Unknown Action Models}
 
@@
 \begin{abstract}
-The {\em plan recognition as planning} approach assumes that {\em observers} must have both correct and complete knowledge of the action model of the observed agents.
-This paper shows that this assumption can be relaxed formulating a novel setup for classical planning where action models are unknown but the state variables and the action parameters are however known.
+{\em Plan recognition as planning} assumes that {\em observers} must have correct and complete knowledge of the action model of the observed agents.
+We relax this assumption formulating a novel setup for classical planning where action models are unknown but the state variables and the action parameters are however known.
 The experimental results demonstrate that this novel classical planning setup allows us to solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without knowing beforehand the precise action model of the observed agents.
@@
 \label{sec:introduction}
-{\em Goal recognition} is a particular classification task in which each class represents a different goal for the observed agents and classification examples are observations of the agents pursuing one of those goals.
+{\em Goal recognition} is a particular classification task in which each class represents a different goal of the observed agents and classification examples are observations of the agents pursuing one of those goals.
 While there exist diverse approaches for {\em goal recognition}, {\em plan recognition as planning}~\cite{ramirez2009plan,ramirez2010probabilistic} is one of the most popular and it is currently at the core of various model-based activity recognition tasks such as, {\em goal recognition design}~\cite{KerenGK14}, {\em deceptive planning}~\cite{masters2017deceptive}, {\em planning for transparency}~\cite{macnally2018action} or {\em counter-planning}~\cite{PozancoEFB18}.
@@
 {\em Plan recognition as planning} leverages the action model of the observed agents, a single plan observation, and an off-the-shelf classical planner to estimate the most likely goal of the observed agents~\cite{ramirez2012plan}.
-In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which can become a too strong requirement when applying {\em plan recognition as planning} on real-world problems.
+In this paper we show how to relax the assumption of {\em knowing the action model of the observed agents}, which can become a too strong requirement.
 We formulate a novel set up for classical planning where no action model is given (instead, only the state variables and the action parameters are known beforehand) and we show that this formulation neatly fits into the {\em plan recognition as planning} approach.
-The experimental results demonstrate that we can solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without requiring having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents
+The experimental results demonstrate that we can solve standard goal recognition benchmarks, still using an off-the-shelf classical planner, but without requiring having at hand a model of the {\em preconditions} and {\em effects} of the actions of the observed agents.
 
@@
 \label{sec:background}
-This section formalizes the {\em classical planning} model that we follow in this work, the kind of {\em observations} that input the {\em goal recognition} task, and the {\em plan recognition as planning} approach for {\em goal recognition}.
+This section formalizes the {\em classical planning} model that we follow in this work, the kind of {\em observations} of plan executions that input the {\em goal recognition} task, and the {\em plan recognition as planning} approach for the {\em goal recognition} task.
 
@@
 A partial assignment of values to fluents is represented by $L$ ({\em literals}).
-We adopt the \emph{open world assumption} (what is not known to be true in a state is unknown) to implicitly represent the unobserved literals of a state.
+To implicitly represent the unobserved literals of a state we adopt the \emph{open world assumption} (i.e.
+what is not known to be true in a state is {\em unknown}).
 Hence, a state $s$ includes positive literals ($f$) and negative literals ($\neg f$) and it is defined as a full assignment of values to fluents; $|s|=|F|$.
@@
 A {\em planning action} $a$ comprises a set of preconditions $\pre(a)\in\mathcal{L}(F)$ and a set of effects $\eff(a)\in\mathcal{L}(F)$.
-The semantics of an action $a$ is specified with two functions: $\rho(s,a)$ denoting whether $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denoting the {\em successor state} that results from applying $a$ in a state $s$.
-Then, $\rho(s,a)$ holds iff $\pre(a)\subseteq s$, i.e.~ the action preconditions hold in the given state.
-The result of executing an applicable action $a$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$, where $\neg\eff(a)$ is the complement of $\eff(a)$, which is subtracted from $s$ so as to ensure that $\theta(s,a)$ remains a well-defined state.
-The subset of effects of an action $a$ that assign a positive value to a fluent is called {\em positive effects} and denoted by $\eff^+(a)\in \eff(a)$ while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects}.
+The semantics of an action $a$ is specified with two functions: $\rho(s,a)$ denoting whether $a$ is {\em applicable} in a state $s$ and $\theta(s,a)$ denoting the {\em successor state} that results from applying $a$ in $s$.
+Then $\rho(s,a)$ holds, i.e.~the action preconditions hold in the given state iff $\pre(a)\subseteq s$.
+The result of executing an applicable action $a$ in a state $s$ is a new state $\theta(s,a)=\{s\setminus \neg\eff(a)\cup\eff(a)\}$, where $\neg\eff(a)$ is the complement of $\eff(a)$, which is subtracted from $s$ to ensure that $\theta(s,a)$ remains a well-defined state.
+The subset of effects of an action $a$ that assign a positive value to a fluent is called {\em positive effects}, and denoted by $\eff^+(a)\in \eff(a)$, while $\eff^-(a)\in \eff(a)$ denotes the {\em negative effects}.
 
-Planning actions can also define {\em conditional effects}.
-Formally the conditional effects of an action $a_c$ is composed of two sets of literals: $C\in\mathcal{L}(F)$, the {\em condition}, and $E\in\mathcal{L}(F)$, the {\em effect}.
-The {\em triggered effects} resulting from the action application (conditional effects whose conditions hold in $s$) is defined as $\eff_c(s,a)=\bigcup_{C\rhd E\in\cond(a_c),C\subseteq s} E$.
+A planning action $a$ can also define a set of {\em conditional effects} $cond(a)$.
+Formally a conditional effect $C\rhd E\in cond(a)$ is composed of two sets of literals: the {\em condition}, $C\in\mathcal{L}(F)$, and the {\em effect}, $E\in\mathcal{L}(F)$.
+The {\em triggered effects} resulting from the action application (conditional effects whose conditions hold in $s$) is defined as $\eff_c(s,a)=\bigcup_{C\rhd E\in\cond(a),C\subseteq s} E$.
 
@@
 \subsection{The observation model}
-Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ that solves $P$, and the corresponding trajectory $\tau$ induced by the execution of $\pi$ in $I$, $\tau=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$; there exist as many observations of $\tau$ as combinations of observable actions and observable fluents of the states of $\tau$.
+Given a planning problem $P=\tup{F,A,I,G}$, a plan $\pi$ that solves $P$, and the corresponding trajectory $\tau$ induced by the execution of $\pi$ in $I$, $\tau=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$; there exist as many possible observations of $\tau$ as combinations of observable actions and observable fluents of the states in $\tau$.
 We refer to the set of all possible combinations of observable elements of $\tau$ as $Obs(\tau)$.
 
-We define {\em an observation} $\mathcal{O}\in Obs(\tau)$ as a sequence of possibly {\em partially observable states}, $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$, except for the initial state $s_0^o=I$ which is fully observable.
-A {\em partially observable state} is one in which $|s_i^o| < |F|$, {\small $1\leq i\leq m\leq n$}; i.e., a state in which at least a fluent of $F$ is not observable.
-It may be also the case that $|s_i^o| = 0$ when an intermediate state is fully unobservable.
+In this work we define {\em an observation} $\mathcal{O}\in Obs(\tau)$ as a sequence of {\em partially observed states}, $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$, except for the initial state $s_0^o=I$ which is fully observed.
+A {\em partially observed state} is one in which $|s_i^o| < |F|$, {\small $1\leq i\leq m\leq n$}; i.e., a state in which at least the value of a fluent in $F$ is {\em unknown}.
+It may be also the case that $|s_i^o| = 0$ when an intermediate state is fully unknown.
+This observation model allow us to distinguish between {\em observable state variables}, whose value may be read from sensors, and {\em hidden} (or {\em latent}) {\em state variables}, that cannot be observed.
 
@@
 
-Given $\mathcal{O} \in Obs(\tau)$, the number of observed states of $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$ ranges from 2 (at least the initial and final state, as explained above) to $|\pi|+1$.
+Given $\mathcal{O} \in Obs(\tau)$, the number of observed states of $\mathcal{O}=\tup{s_0^o,s_1^o \ldots , s_m^o}$ ranges from 2 (at least the initial and final state) to $|\pi|+1$.
 The number of fluents of the full observable state $s_0^o$ will be $|F|$, or $|F|+1$ in case the fluent of the applied action in $s_0$ is also observed.
@@
 
-This observation model can distinguish between {\em observable state variables}, whose value may be read from sensors, and {\em hidden} (or {\em latent}) {\em state variables}, that cannot be observed.
-Given a subset of fluents $\Gamma\subseteq F$ we say that $\mathcal{O}$ is a $\Gamma$-observation of the execution of $\pi$ on $P$ iff for every observed state $s_i^o$, ${\small 1\leq i\leq m}$, $s_i^o$ only contains fluents in $\Gamma$.
+\subsection{Goal recognition as planning}
+{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ (where $G[\cdot]$ is the set of {\em recognizable} goals) and there is a single classification example $\mathcal{O}$ (that represents the observation of a plan execution where agents act to achieve a goal $G\in G[\cdot]$).
 
-\subsection{Goal recognition as planning}
-{\em Goal recognition} is a specific classification task in which each class represents a different possible goal $G\in G[\cdot]$ (where $G[\cdot]$ represents the set of {\em recognizable} goals) and there is a single classification example $\mathcal{O}(\tau)$ (that represents the observation of a plan where agents act to achieve a goal $G\in G[\cdot]$).
-
-Following the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
+According to the {\em naive Bayes classifier}, the {\em solution} to the {\em goal recognition} task is the subset of goals in $G[\cdot]$ that maximizes this expression.
 \begin{align}
@@
 
-The {\em plan recognition as planning} approach for goal recognition shows that the $P(\mathcal{O}|G)$ likelihood can be estimated leveraging the action model of the observed agents and an off-the-shelf classical planner~\cite{ramirez2012plan}.
+The {\em plan recognition as planning} approach for goal recognition shows that the $P(\mathcal{O}|G)$ likelihood can be estimated leveraging the action model of the observed agents and off-the-shelf classical planners~\cite{ramirez2012plan}.
 Given $P=\tup{F,A,I,G[\cdot]}$ then $P(\mathcal{O}|G)$ is estimated computing, for each goal $G\in G[\cdot]$, the cost difference of the solution plans to two classical planning problems:
 \begin{itemize}
-\item $P^{\top}_G$, the classical planning problem built constraining $P=\tup{F,A,I,G}$ to achieve the particular goal $G\in G[\cdot]$ through a plan $\pi^\top$ that is {\em consistent} with the input observation $\mathcal{O}$.
+\item $P^{\top}_G$, the classical planning problem built constraining $P=\tup{F,A,I,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}$.
 \item $P^{\bot}_G$, the classical planning problem that constrains solutions of $P=\tup{F,A,I,G}$ to plans $\pi^\bot$, that achieve $G\in G[\cdot]$, but that are {\em inconsistent} with $\mathcal{O}$.
@@
 
-To compute the target probability distribution $P(G|\mathcal{O})$ plug the $P(\mathcal{O}|G)$ likelihoods into the {\em Bayes rule} from which the goal posterior probabilities are obtained.
+To compute the target probability distribution $P(G|\mathcal{O})$ the $P(\mathcal{O}|G)$ likelihoods can be plugged into the {\em Bayes rule}.
 In this case the $P(\mathcal{O})$ probabilities are obtained by normalization (goal probabilities must add up to 1 when summed over all possible goals).
 
-\section{Goal recognition as planning with unknown domain models}
+\section{Goal recognition as planning with unknown action models}
 \label{sec:recognition}
+This section formalizes the goal recognition task addressed in this paper, defines a setup for classical planning with unknown action models, and shows how to leverage this setup to solve goal recognition tasks when the action model of the observed agents is unknown.
 
-We define the task of {\em goal recognition with unknown domain models} as a $\tup{P,\mathcal{O}}$ pair, where:
+\subsection{Goal recognition with unknown action models}
+We define the task of {\em goal recognition with unknown domain models} as a $\tup{P,\mathcal{O}}$ pair:
 \begin{itemize}
@@
 \end{itemize}
+The parameters of the actions in $A[\cdot]$ are known. This means that they are either given in $P$ or deducible from $\mathcal{O}$.
 
 The {\em solution} to the {\em goal recognition with unknown domain models} task is again the subset of goals in $G[\cdot]$ that maximizes expression (1).
-With this regard this new task is defined exactly as the original goal recognition task~\cite{ramirez2012plan} except that the preconditions and effects of the input set of actions are now unknown.
-
-\subsection{Estimating $P(\mathcal{O}|G)$ with unknown action models}
-Our approach to estimate the $P(\mathcal{O}|G)$ likelyhood is to leverage the available input knowledge to build a reasonable model for the actions in $A[\cdot]$, use such model to compute solution plans to the classical planning problems $P^{\top}_G$ and $P^{\bot}_G$ and, estimate $P(\mathcal{O}|G)$ as in the {\em plan recognition as planning} approach for goal recognition.
-
-In more detail, given $P$ and $\mathcal{O}$, we compute the estimate of the $P(\mathcal{O}|G)$ likelyhood following these steps:
-\begin{enumerate}
-\item Compute an action model $A$ that is able to solve {\em $P^{\top}_G$}, the classical planning problem that constrains solutions of $\tup{F,A[\cdot],I,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}$.
-\item Replace $A[\cdot]$ with $A$, {\em solve $P^{\top}_G$} and take $cost(\pi^\top)$ from that solution.
-\item Replace $A[\cdot]$ with $A$, {\em solve $P^{\bot}_G$} (i.e., the classical planning problem that constrains $\tup{F,A,I,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}$) and take $cost(\pi^\top)$ from that solution.
-\item Compute the $\Delta(\pi^\top,\pi^\bot)$ cost difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
-\end{enumerate}
+With this regard this new task is defined exactly as the original goal recognition task~\cite{ramirez2012plan} except that the {\em preconditions} and {\em effects} of the input set of {\em actions} are now unknown.
 
 \subsection{Planning with unknown domain models}
-This section formulates {\em planning with unknown domain models}, a novel setup for classical planning where no action model is given.
+Our approach to estimate the $P(\mathcal{O}|G)$ likelihoods is to compute solution plans to the classical planning problems $P^{\top}_G$ and $P^{\bot}_G$, as in the {\em plan recognition as planning} approach for goal recognition.
+The only difference of our approach is that we solve $P^{\top}_G$ while we build a reasonable model for the actions in $A[\cdot]$ leveraging the available input knowledge $\tup{P,\mathcal{O}}$.
+To do so, we formulate {\em planning with unknown domain models}, a setup for classical planning where no action model is given.
 The state variables, action parameters and a single plan observation are known though.
-This setup is closely related to the learning of planning action models~\cite{SternJ17} and can be seen as an extreme learning task where action models must be {\em learned} from a single example that contain only two state observations: the initial state and the goals.
-This planning setup is of interest because it allows to formulate steps 1.
-and 2.
-as a single classical planning problem.
+This setup is closely related to the learning of planning action models~\cite{SternJ17} since it can be understood as {\em one-shot learning} (i.e.
+learning from a single example).
 
@@
 A {\em solution plan} is a sequence of actions $\pi=\tup{a_1, \ldots, a_n}$ whose execution on $I$ induces a trajectory $\tau=\tup{s_0, a_1, s_1, \ldots, a_n, s_n}$ such that $s_0=I$ and {\em there exists} at least one possible action model (e.g.
-one possible definition of the $\rho$ and $\theta$ functions within the given state variables) satisfying:
+one possible definition of the $\rho$ and $\theta$ functions within the given state variables) satisfying the following constraints:
 \begin{itemize}
@@
 
-We can define further constraints to reduce the hypothesis space for the functions $\rho(s,a)$ and $\theta(s,a)$ corresponding to the $A[\cdots]$ actions:
+Further constraints can be defined to build more accurate models for the $\rho(s,a)$ and $\theta(s,a)$ functions of the $A[\cdot]$ actions:
 \begin{itemize}
 \item {\em Observations}.
-The action model must be consistent with the input observation.
-Specifically, the states induced by plans computable with such model must comprise the observed states of the sample, which further constrains the space of possible action models.
-The observation $\mathcal{O}$ can also be regarded as a sequence of ordered {\em landmarks} for the planning problem $P_\mathcal{O}$~\cite{hoffmann2004ordered} since all the fluents of the sets in $\mathcal{O}$ must be achieved by any plan that solves $P_\mathcal{O}$ and in the same order as defined in the observation $\mathcal{O}$.
-
-\item {\em Domain-specific knowledge} is also helpfull to constrain further the space of possible action schemata.
+The states trajectories induced by plans computable with the actions in $A[\cdot]$ must be {\em consistent} with the observed states in $\mathcal{O}$.
+Observations $\mathcal{O}$ act as ordered {\em landmarks} for the planning problem $P$~\cite{hoffmann2004ordered} since all the fluents of the sets in $\mathcal{O}$ must be achieved by any plan that solves $P$ and in the same order as defined in the observation $\mathcal{O}$.
+\item {\em Domain-specific knowledge}.
+{\em State invariants} is a useful type of state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+{\em State invariants} reduce the space of the possible preconditions and effects of the actions in $A[\cdot]$.
 For instance, in the {\em blocksworld} one can argue that {\small\tt on(?x,?x)} will not appear in the preconditions/effects of an action because, in this specific domain, a block cannot be on top of itself.
 \item {\em Partially-specified action models}.
-In some contexts it is however reasonable to assume that some portions of the action model are known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
+In some contexts some portions of the preconditions and effects of the actions in $A[\cdot]$ can be known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
 \end{itemize}
 
-\section{Planning with unknown \strips\ models}
+\subsection{Estimating $P(\mathcal{O}|G)$ by planning with unknown action models}
+Our formalization of {\em classical planning with unknown domain models} can be used to estimate the $P(\mathcal{O}|G)$ likelihoods following the {\em plan recognition as planning} approach for goal recognition:
+\begin{enumerate}
+\item Compute an action model $A$ that can solve {\em $P^{\top}_G$}, now the classical planning problem {\em with unknown action model} that constrains solutions of $\tup{F,A[\cdot],I,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}$, and take $cost(\pi^\top)$ from that solution.
+\item Solve $P^{\bot}_G$, the classical planning problem that constrains $\tup{F,A,I,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}$, and take $cost(\pi^\bot)$ from that solution.
+\item Compute the $\Delta(\pi^\top,\pi^\bot)$ cost difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
+\end{enumerate}
+
+\begin{figure}
+  \begin{footnotesize}
+$\forall x_1,x_2\ \neg ontable(x_1)\vee\neg on(x_1,x_2)$.\\
+$\forall x_1,x_2\ \neg clear(x_1)\vee\neg on(x_2,x_1)$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
+$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
+\end{footnotesize}
+ \caption{\small {\em Schematic mutexes} for the {\em blocksworld} domain.}
+\label{fig:strongest-invariant}
+\end{figure}
+
+\subsection{Completing observations with {\em schematic mutexes}}
+The estimate of the $P(\mathcal{O}|G)$ likelihoods can be optimizing by adding a pre-processing step:
+\begin{itemize}
+\item[$0.$]  Complete the input observation $\mathcal{O}$ using the domain-specific knowledge if available.
+\end{itemize}
+
+Next we provide the details os this pre-process.
+A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
+For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
+
+Recently, some works point at extracting \emph{lifted} invariants, also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}, that hold for any possible state and any possible set of objects.
+Invariant
+templates obtained by inspecting the lifted representation of the domain have also been exploited for deriving \emph{lifted mutex}~\cite{BernardiniFS18}.
+In this work we exploit domain-specific knowledge that is given as {\em schematic mutex}.
+We pay special attention to {\em schematic mutex} because they identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) an effective completion of a partially observed state and (2) an effective pruning of inconsistent \strips\ action models.
+
+We define a schematic mutex as a $\tup{p,q}$ pair of predicates in $\Psi$ (the initial state is a full assignment of values to fluents so the predicates $\Psi$ are extractable from $I$) that satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding parameters are variables universally quantified.
+For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible state.
+Figure~\ref{fig:strongest-invariant} shows an example of four clauses that define schematic mutexes for the {\em blocksworld} domain.
+
+Let $\Omega$ be the set of objects that appear in $F$ as the values of the arguments of the predicates $\Psi$, and $\phi=\tup{p,q}$ a schematic mutex.
+There exist many possible instantiations of $\phi$ of the type $\tup{p(\omega),q(\omega')}$ with objects of $\Omega$, where $\omega\subseteq\Omega^{|args(p)|}$ and $\omega'\subseteq\Omega^{|args(q)|}$.
+Let us now assume that the instantiation $p(\omega) \in s_j^o$, {\small $(1\leq j\leq m)$}, being $s_j^o$ a partially observed state of $\mathcal{O}$.
+Then, two situations may occur: (a) $\neg q(\omega') \in s_j^o$, in which case the expression $\neg p(\omega) \vee \neg q(\omega')$ holds in $s_j^o$; or (b) $\neg q(\omega') \notin s_j^o$, in which case the literal has not been observed in $s_j^o$ and so we can safely complete the state with $\neg q(\omega')$ (the same applies inversely, when $q(\omega') \in s_j^o$ but $\neg p(\omega) \notin s_j^o$).
+In other words, if we find that one component of a schematic mutex is positively observed in a state and the other component is not observable in such state, we can complete the state with the missing negative literal.
+For instance, if the literal {\tt\small holding(blockA)} is observed in a particular state and $\Phi$ contains the schematic mutex $\neg holding(v_1)\vee\neg clear(v_1)$, we extend the state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal being initially unknown).
+
+\section{Classical planning with unknown \strips\ models}
 \label{sec:planning}
 This section shows that, when the unknown precondition and effects of the actions in $A[\cdot]$ follow the \strips\ model, we can solve the task of {\em planning with unknown domain models} (and hence estimate the $P(\mathcal{O}|G)$ likelyhood) with an off-the-shelf classical planner.
-
-\subsection{The space of possible action models}
-A \strips\ \emph{action model} is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
-
-Let $\Psi$ be the set of {\em predicates} that shape the fluents $F$ (the initial state is a full assignment of values to fluents so the predicates $\Psi$ are extractable from $I$).
-The set of propositions that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
-For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-
-\subsection{Syntactic constrains}
-An action model $\xi$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-{\em Typing constraints} are also a type of syntactic constraint that reduce the size of ${\mathcal I}_{\xi,\Psi}$~\cite{mcdermott1998pddl}.
-Considering only these syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
-Given $p \in \mathcal{I}_{\Psi,\xi}$, the belonging of $p$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a propositional encoding that uses fluents of two types, $pre_{p,\xi}$ and $eff_{p,\xi}$.
-The four possible combinations of these two fluents are summarized in Figure \ref{fig:combinations}.
 
@@
 
-To illustrate better this encoding, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with $pre_{p,stack}$ and $eff_{p,stack}$ fluents ($p\in{\mathcal I}_{\Psi,stack}$).
-
 \begin{figure}
@@
 \end{figure}
+
+\subsection{A propositional encoding for \strips\ actions models}
+A \strips\ \emph{action model} is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
+
+Let $\Psi$ be the set of {\em predicates} that shape the fluents $F$.
+The set of propositions that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
+
+For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
+
+An action model $\xi$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
+{\em Typing constraints} are also a type of syntactic constraint that reduce the size of ${\mathcal I}_{\xi,\Psi}$~\cite{mcdermott1998pddl}.
+Considering only these syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
+Given $p \in \mathcal{I}_{\Psi,\xi}$, the belonging of $p$ to the preconditions, positive effects or negative effects of $\xi$ is handled with a propositional encoding that uses fluents of two types, $pre_{p,\xi}$ and $eff_{p,\xi}$.
+The four possible combinations of these two fluents are summarized in Figure \ref{fig:combinations}.
+
+To illustrate better this encoding, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with $pre_{p,stack}$ and $eff_{p,stack}$ fluents ($p\in{\mathcal I}_{\Psi,stack}$).
 
@@
 
-\end{enumerate}
-\end{itemize}
-
-\subsection{Domain-specific constraints}
-
-Our approach is to introduce {\em domain-specific knowledge} in the form of {\em state constraints} to further restrict the space of the action models.
-Back to the {\em blocksworld} domain, one can argue that {\small\tt on($v_1$,$v_1$)} and {\small\tt on($v_2$,$v_2$)} will not appear in the $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of any action model $\xi$ because, in this specific domain, a block cannot be on top of itself.
-The notion of state constraint is very general and has been used in different areas of AI and for different purposes.
-In planning, state constraints are compact and abstract representations that relate the values of variables in each state traversed by a plan, and allow to specify the set of states where a given action is applicable, the set of states where a given {\em axiom} or {\em derived predicate} holds or the set of states that are considered goal states~\cite{HaslumIR0TSN18}.
-
-{\em State invariants} is a useful type of state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-Given a planning problem $P=\tup{F,A,I,G}$, a state invariant is a formula $\phi$ that holds in $I$, $I\models \phi$, and in every state $s$ built out of $F$ that is reachable by applying actions of $A$ in $I$.
-
-A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
-
-Recently, some works point at extracting \emph{lifted} invariants, also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}, that hold for any possible state and any possible set of objects.
-Invariant
-templates obtained by inspecting the lifted representation of the domain have also been exploited for deriving \emph{lifted mutex}~\cite{BernardiniFS18}.
-In this work we exploit domain-specific knowledge that is given as {\em schematic mutex}.
-We pay special attention to {\em schematic mutex} because they identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) an effective completion of a partially observed state and (2) an effective pruning of inconsistent \strips\ action models.
-
-%A {\em domain invariant} is an instance-independent state invariant that holds for any possible initial state and any possible set of objects.
-Therefore, if $s\nvDash \phi$ such that $\phi$ is a domain invariant, it means that $s$ is not a valid state.
-Domain invariants are often compactly defined as {\em lifted invariants} (also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}).
-
-We define a schematic mutex as a $\tup{p,q}$ pair where both $p,q\in{\mathcal I}_{\xi,\Psi}$ are predicates that shape the preconditions or effects of a given action scheme $\xi$ and they satisfy the formulae $\neg p\vee \neg q$, considering that their corresponding variables are universally quantified.
-For instance, $holding(v_1)$ and $clear(v_1)$ from the {\em blocksworld} are {\em schematic mutex} while $clear(v_1)$ and $ontable(v_1)$ are not because $\forall v_1, \neg clear(v_1)\vee\neg ontable(v_1)$ does not hold for every possible state.
-Figure~\ref{fig:strongest-invariant} shows an example of four clauses that define schematic mutexes for the {\em blocksworld} domain.
-
-\begin{figure}[hbt!]
-  \begin{footnotesize}
-$\forall x_1,x_2\ \neg ontable(x_1)\vee\neg on(x_1,x_2)$.\\
-$\forall x_1,x_2\ \neg clear(x_1)\vee\neg on(x_2,x_1)$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_1,x_2)\vee\neg on(x_1,x_3)\ such\ that\ x_2\neq x_3$.\\
-$\forall x_1,x_2,x_3\ \neg on(x_2,x_1)\vee\neg on(x_3,x_1)\ such\ that\ x_2\neq x_3$.\\
-\end{footnotesize}
- \caption{\small {\em Schematic mutexes} for the {\em blocksworld} domain.}
-\label{fig:strongest-invariant}
-\end{figure}
-
-{\it state invariants} can be exploited either as {\em syntactic} constraints (to reduce the space of possible action models) but also as {\em semantic} constraints (to complete partial observations of the states traversed by a plan)~\cite{fox:TIM:JAIR1998}.
-
-\subsubsection{Completing partially observed states with {\em schematic mutexes}}
-
-The addition of new literals to complete the partial states $\tup{s_1^o \ldots, s_m^o}$ of an observation $\mathcal{O}$  using a set of schematic mutexes $\Phi$ is done in a pre-processing stage.
-
-Let $\Omega$ be the set of objects that appear in $F$ as the values of the arguments of the predicates $\Psi$, and $\phi=\tup{p,q}$ a schematic mutex.
-There exist many possible instantiations of $\phi$ of the type $\tup{p(\omega),q(\omega')}$ with objects of $\Omega$, where $\omega\subseteq\Omega^{|args(p)|}$ and $\omega'\subseteq\Omega^{|args(q)|}$.
-Let us now assume that the instantiation $p(\omega) \in s_j^o$, {\small $(1\leq j\leq m)$}, being $s_j^o$ a partially observed state of $\mathcal{O}$.
-Then, two situations may occur: (a) $\neg q(\omega') \in s_j^o$, in which case the expression $\neg p(\omega) \vee \neg q(\omega')$ holds in $s_j^o$; or (b) $\neg q(\omega') \notin s_j^o$, in which case the literal has not been observed in $s_j^o$ and so we can safely complete the state with $\neg q(\omega')$ (the same applies inversely, when $q(\omega') \in s_j^o$ but $\neg p(\omega) \notin s_j^o$).
-In other words, if we find that one component of a schematic mutex is positively observed in a state and the other component is not observable in such state, we can complete the state with the missing negative literal.
-For instance, if the literal {\tt\small holding(blockA)} is observed in a particular state and $\Phi$ contains the schematic mutex $\neg holding(v_1)\vee\neg clear(v_1)$, we extend the state observation with literal {\tt\small $\neg$clear(blockA)} (despite this particular literal being initially unknown).
-
 \begin{figure}
@@
 
-\subsubsection{Pruning inconsistent action models with {\em schematic mutexes}}
+\subsection{Schematic mutexes}
 We could extend the classical planning compilation for the learning of \strips\ action models~\cite{aineto2018learning} to check the consistency of the {\em state-constraints} in $\Phi$ at every state traversed by a solution to the compiled problem.

### 2019-05-16 doc/aaai20-goalRecognition/goalRecognition.tex
adding updates to the aaai20 paper on goal recognition

@@
 With this regard this new task is defined exactly as the original goal recognition task~\cite{ramirez2012plan} except that the {\em preconditions} and {\em effects} of the input set of {\em actions} are now unknown.
+Note that in this new framework observing a plan prefix does no longer means knowing the current state because the action model is unknown.
 
@@
 Our approach to estimate the $P(\mathcal{O}|G)$ likelihoods is to compute solution plans to the classical planning problems $P^{\top}_G$ and $P^{\bot}_G$, as in the {\em plan recognition as planning} approach for goal recognition.
-The only difference of our approach is that we solve $P^{\top}_G$ while we build a reasonable model for the actions in $A[\cdot]$ leveraging the available input knowledge $\tup{P,\mathcal{O}}$.
+The main difference of our approach is that we solve $P^{\top}_G$ while we build a reasonable model for the actions in $A[\cdot]$ leveraging the available input knowledge $\tup{P,\mathcal{O}}$.
 To do so, we formulate {\em planning with unknown domain models}, a setup for classical planning where no action model is given.
 The state variables, action parameters and a single plan observation are known though.
-This setup is closely related to the learning of planning action models~\cite{SternJ17} since it can be understood as {\em one-shot learning} (i.e.
+This setup is closely related to the learning of planning action models~\cite{SternJ17} and it can be understood as {\em one-shot learning} (i.e.
 learning from a single example).
 
-A {\em classical planning with unknown domain models} is defined as a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
+A {\em planning problem with unknown domain models} is defined as a tuple $P=\tup{F,A[\cdot],I,G}$, where $A[\cdot]$ is a set of actions s.t., the semantics of each action $a\in A[\cdot]$ is unknown (i.e.
 the functions $\rho$ and/or $\theta$ of $a$ are undefined).
@@
 
-Further constraints can be defined to build more accurate models for the $\rho(s,a)$ and $\theta(s,a)$ functions of the $A[\cdot]$ actions:
+Additional constraints can be defined to build more accurate models for the $\rho(s,a)$ and $\theta(s,a)$ functions of the $A[\cdot]$ actions:
 \begin{itemize}
@@
 The states trajectories induced by plans computable with the actions in $A[\cdot]$ must be {\em consistent} with the observed states in $\mathcal{O}$.
-Observations $\mathcal{O}$ act as ordered {\em landmarks} for the planning problem $P$~\cite{hoffmann2004ordered} since all the fluents of the sets in $\mathcal{O}$ must be achieved by any plan that solves $P$ and in the same order as defined in the observation $\mathcal{O}$.
+Observations $\mathcal{O}$ act as ordered {\em landmarks}~\cite{hoffmann2004ordered} since all the fluents of the sets in $\mathcal{O}$ must be achieved by any solution plan and in the same order as defined in the observation $\mathcal{O}$.
 \item {\em Domain-specific knowledge}.
-{\em State invariants} is a useful type of state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
-{\em State invariants} reduce the space of the possible preconditions and effects of the actions in $A[\cdot]$.
+{\em State invariants} are known to be useful state constraints for computing more compact state representations of a given planning problem~\cite{helmert2009concise} and for making {\em satisfiability planning} or {\em backward search} more efficient~\cite{rintanen2014madagascar,alcazar2015reminder}.
+{\em State invariants} can also be used to reduce the space of the possible preconditions and effects of the actions in $A[\cdot]$.
 For instance, in the {\em blocksworld} one can argue that {\small\tt on(?x,?x)} will not appear in the preconditions/effects of an action because, in this specific domain, a block cannot be on top of itself.
 \item {\em Partially-specified action models}.
-In some contexts some portions of the preconditions and effects of the actions in $A[\cdot]$ can be known~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
+Some portions of the preconditions and effects of the actions in $A[\cdot]$ can be known in advance~\cite{ZhuoNK13,sreedharan2018handling,pereira2018heuristic}.
 \end{itemize}
@@
 \subsection{Estimating $P(\mathcal{O}|G)$ by planning with unknown action models}
-Our formalization of {\em classical planning with unknown domain models} can be used to estimate the $P(\mathcal{O}|G)$ likelihoods following the {\em plan recognition as planning} approach for goal recognition:
+Our formalization for {\em planning with unknown domain models} can be used straightforward to estimate the $P(\mathcal{O}|G)$ likelihoods following the {\em plan recognition as planning} approach:
 \begin{enumerate}
-\item Compute an action model $A$ that can solve {\em $P^{\top}_G$}, now the classical planning problem {\em with unknown action model} that constrains solutions of $\tup{F,A[\cdot],I,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}$, and take $cost(\pi^\top)$ from that solution.
+\item Compute an action model $A$ that solves {\em $P^{\top}_G$}, the classical planning problem {\em with unknown action model} that constrains solutions of $\tup{F,A[\cdot],I,G}$ to plans $\pi^\top$ {\em consistent} with the input observation $\mathcal{O}$, and take $cost(\pi^\top)$ from that solution.
 \item Solve $P^{\bot}_G$, the classical planning problem that constrains $\tup{F,A,I,G}$ to achieve $G\in G[\cdot]$ through a plan $\pi^\bot$ {\em inconsistent} with $\mathcal{O}$, and take $cost(\pi^\bot)$ from that solution.
-\item Compute the $\Delta(\pi^\top,\pi^\bot)$ cost difference and plug it into equation (2) to get the $P(\mathcal{O}|G)$ likelihoods.
+\item Compute the $\Delta(\pi^\top,\pi^\bot)$ cost difference and plug it into equation (2).
 \end{enumerate}
@@
 \subsection{Completing observations with {\em schematic mutexes}}
-The estimate of the $P(\mathcal{O}|G)$ likelihoods can be optimizing by adding a pre-processing step:
+The quality of the $P(\mathcal{O}|G)$ estimates can be improved pre-processing the input observation:
 \begin{itemize}
-\item[$0.$]  Complete the input observation $\mathcal{O}$ using the domain-specific knowledge if available.
+\item[$0.$]  Complete the partial states of the input observation $\mathcal{O}$ with domain-specific knowledge if available.
 \end{itemize}
@@
 Next we provide the details os this pre-process.
-A {\em mutex} (mutually exclusive) is a state invariant that takes the form of a binary clause and indicates a pair of different properties that cannot be simultaneously true~\cite{kautz:mutex:IJCAI1999}.
-For instance in a three-block {\em blocksworld}, $\neg on(block_A,block_B)\vee \neg on(block_A,block_C)$ is a {\em mutex} because $block_A$ can only be on top of a single block.
-
 Recently, some works point at extracting \emph{lifted} invariants, also called {\em schematic} invariants~\cite{rintanen:schematicInvariants:AAAI2017}, that hold for any possible state and any possible set of objects.
-Invariant
-templates obtained by inspecting the lifted representation of the domain have also been exploited for deriving \emph{lifted mutex}~\cite{BernardiniFS18}.
+Invariant templates obtained by inspecting the lifted representation of the domain have also been exploited for deriving \emph{lifted mutex}~\cite{BernardiniFS18}.
 In this work we exploit domain-specific knowledge that is given as {\em schematic mutex}.
-We pay special attention to {\em schematic mutex} because they identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998} and because they enable (1) an effective completion of a partially observed state and (2) an effective pruning of inconsistent \strips\ action models.
+We pay special attention to {\em schematic mutex} because they identify mutually exclusive properties of a given type of objects~\cite{fox:TIM:JAIR1998}.
 
@@
 An action model $\xi$ must be consistent with the \strips\ constraints: $del(\xi)\subseteq pre(\xi)$, $del(\xi)\cap add(\xi)=\emptyset$ and $pre(\xi)\cap add(\xi)=\emptyset$.
-{\em Typing constraints} are also a type of syntactic constraint that reduce the size of ${\mathcal I}_{\xi,\Psi}$~\cite{mcdermott1998pddl}.
 Considering only these syntactic constraints, the size of the space of possible \strips\ models is given by $2^{2\times|{\mathcal I}_{\Psi,\xi}|}$ because one element in $\mathcal{I}_{\xi,\Psi}$ can appear both in the preconditions and effects of $\xi$.
@@
 The four possible combinations of these two fluents are summarized in Figure \ref{fig:combinations}.
-
 To illustrate better this encoding, Figure~\ref{fig:propositional} shows the PDDL encoding of the {\tt\small stack(?v1,?v2)} schema and our propositional representation for this same schema with $pre_{p,stack}$ and $eff_{p,stack}$ fluents ($p\in{\mathcal I}_{\Psi,stack}$).
 
+{\em Typing constraints} can also be used here to reduce the size of ${\mathcal I}_{\xi,\Psi}$~\cite{mcdermott1998pddl}.
+
 \subsection{A classical planning compilation for planning with unknown domain models}
-The approach for learning \strips\ action models presented in~\cite{aineto2018learning}, which we will use as our baseline learning system (hereafter BLS, for short), is a compilation scheme that transforms the problem of learning the preconditions and effects of action models into a planning task $P'$.
-A \strips\ \emph{action model} $\xi$ is defined as $\xi=\tup{name(\xi),pars(\xi),pre(\xi),add(\xi),del(\xi)}$, where $name(\xi)$ and parameters, $pars(\xi)$, define the header of $\xi$; and $pre(\xi)$, $del(\xi)$ and $add(\xi)$) are sets of fluents that represent the {\em preconditions}, {\em negative effects} and {\em positive effects}, respectively, of the actions induced from the action model $\xi$.
+To plan with unknown \strips\ action models we adapt the classical planinng compilation for learning \strips\ action models~\cite{aineto2018learning}.
 
-The BLS receives as input an empty domain model, which only contains the headers of the action models, and a set of observations of plan executions, and creates a propositional encoding of the planning task $P'$.
-Let $\Psi$ be the set of {\em predicates}\footnote{The initial state of an observation is a full assignment of values to fluents, $|s_0|=|F|$, and so the predicates $\Psi$ are extractable from the observed state $s_0$.} that shape the variables $F$.
-The set of propositions of $P'$ that can appear in $pre(\xi)$, $del(\xi)$ and $add(\xi)$ of a given $\xi$, denoted as ${\mathcal I}_{\xi,\Psi}$, are FOL interpretations of $\Psi$ over the parameters $pars(\xi)$.
-For instance, in a four-operator {\em blocksworld}~\cite{slaney2001blocks}, the ${\mathcal I}_{\xi,\Psi}$ set contains five elements for the {\small \tt pickup($v_1$)} model, ${\mathcal I}_{pickup,\Psi}$={\small\tt\{handempty, holding($v_1$),clear($v_1$),ontable($v_1$), on($v_1,v_1$)\}} and eleven elements for the model of {\small \tt stack($v_1$,$v_2$)}, ${\mathcal I}_{stack,\Psi}$={\small\tt\{handempty, holding($v_1$), holding($v_2$), clear($v_1$),clear($v_2$),ontable($v_1$),ontable($v_2$), \\
-on($v_1,v_1$),on($v_1,v_2$), on($v_2,v_1$), on($v_2,v_2$)\}}.
-Hence, solving $P'$ consists in determining which elements of ${\mathcal I}_{\xi,\Psi}$ will shape the preconditions, positive and negative effects of each action model $\xi$.
+In a nutshell, the output of the compilation is a plan that completes the empty input domain model by specifying the preconditions and effects of each action model such that the validation of the completed model over the input observations is successful.
+The compilation receives as input an empty domain model, which only contains the headers of the action models, and determines whether or not an element of ${\mathcal I}_{\xi,\Psi}$ is part of $pre(\xi)$, $del(\xi)$ or $add(\xi)$.
 
-The decision as to whether or not an element of ${\mathcal I}_{\xi,\Psi}$ will be part of $pre(\xi)$, $del(\xi)$ or $add(\xi)$ is given by the plan that solves $P'$.
-Specifically, two different sets of actions are included in the definition of $P'$: \emph{insert actions}, which insert preconditions and effects on an action model; and \emph{apply actions}, which validate the application of the learned action models in the input observations.
-Roughly speaking, in the \emph{blocksworld} domain, the insert actions of a plan that solves $P'$ will look like {\tt{\footnotesize(insert\_pre\_stack\_holding\_v1)}},\\
+Briefly, the compilation defines two different sets of actions : \emph{insert actions}, which insert preconditions and effects on an action model; and \emph{apply actions}, which validate the application of the learned action models in the input observations.
+Roughly speaking, in the \emph{blocksworld} domain, the insert actions of a plan that solves a compiled problem will look like {\tt{\footnotesize(insert\_pre\_stack\_holding\_v1)}},\\
 {\tt{\footnotesize(insert\_eff\_stack\_clear\_v1),\\
-(insert\_eff\_stack\_clear\_v2)}}, where the second action denotes a positive effect and the third one a negative effect both to be inserted in the model of {\tt{\small stack}}; and the second set of actions of the plan that solves $P'$ will be like {\tt{\small (apply\_unstack blockB blockA),(validate\_1),(apply\_putdown blockB),(validate\_2)}}, where the {\tt {\small validate}} actions denote the points at which the states generated through the {\tt {\small apply}} actions must be validated with the observations of plan executions.
+(insert\_eff\_stack\_clear\_v2)}}, where the second action denotes a positive effect and the third one a negative effect both to be inserted in the model of {\tt{\small stack}}; and the second set of actions of the plan that solves a compiled problem will be like {\tt{\small (apply\_unstack blockB blockA),(validate\_1),(apply\_putdown blockB),(validate\_2)}}, where the {\tt {\small validate}} actions denote the points at which the states generated through the {\tt {\small apply}} actions must be validated with the observations of plan executions.
 
-In a nutshell, the output of the BLS compilation is a plan that completes the empty input domain model by specifying the preconditions and effects of each action model such that the validation of the completed model over the input observations is successful.
+Next we show how to extend the compilation to include constraints that  are useful for the {\em Goal Recognition as Planning with Unknown Action Models}.
 
 \subsection{Observations}
-This require the compilation to include actions for {\em validating} partially observed states $s_j^o\in\mathcal{O}$.
+The compilation includes actions for {\em validating} partially observed states $s_j^o\in\mathcal{O}$.
 These actions are also part of the postfix of the solution plan $\pi_\Lambda$ and they are aimed at checking that the observation $\mathcal{O}$ follows after the execution of the apply actions.
@@
 \begin{figure}
-	\begin{footnotesize}
+	\begin{scriptsize}
 		\begin{tabular}{lll}
@@
 		\end{tabular}
-	\end{footnotesize}
+	\end{scriptsize}
 	\caption{\small Summary of the new conditional effects added to the classical planning compilation for the learning of \strips\ action models.}
@@
 
-The goals of the planning task $P'$ generated by the original BLS compilation are extended with the $\neg invalid$ literal to validate that only states consistent with the state constraints defined in $\Phi$ are traversed by solution plans.
+The goals of the planning task generated by the original compilation are extended with the $\neg invalid$ literal to validate that only states consistent with the state constraints defined in $\Phi$ are traversed by solution plans.
 Remarkably, the $\neg invalid$ literal allows us also to define ${\tt \small (apply)_{\xi,\omega}}$ actions more compactly than in the original compilation.
@@
 These two extra layers are followed by the typical $N$ layers of the SATPLAN encoding (extended however to apply the action models that are determined by the previous two initial layers, the $\mathsf{apply_{\xi,\omega}}$ actions).
-Regarding again the example of Figure~\ref{fig:plan-lplan}, this means that steps [00-04] are applied in paralel in the first SATPLAN layer, steps [05-13] are applied in paralel in the second layer and each step [14-17] is applied sequentially and correponds to a differerent SATPLAN layer (so just six layers are necesary to compute the example plan of Figure~\ref{fig:plan-lplan}).
 
